### 2013-10-26 thesis/thesis.tex
Add figures of nordea 15m returns as well as garch estimate results.

\documentclass{book}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}

\input{../physics_common.tex}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

\title{Mathematical Aspects of the Capital Market}
\author{Xie Xiaolei}
\date{\today}
\begin{document}

\maketitle
\tableofcontents

\chapter{Introduction}
\section{Stylized facts}

\section{GARCH models}
The Auto-correlation function of $\epsilon_t^2$ is given by (c.f. \cite{Bollerslev87})
$$
\rho_n = \sum_{i=1}^{max(p,q)} (\alpha_i + \beta_i) \rho_{n-i}
\;n > p
$$
where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
zeros. From these equations, it is clear that the partial
auto-correlation function cuts off at $max(p, q)$.

\chapter{Case Study of Stocks}
\section{Nordea Bank}
\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=103 236 118
  130]{../pics/nordea_price_20120116-20120420.pdf}
  \caption{Nordea Bank paid prices 2012/01/16-2012/04/20}
  \label{fig:Nordea}
\end{figure}
Figure \ref{fig:Nordea} shows the paid prices of Nordea Bank from
2012/01/16 to 2012/04/20.

\subsection{15min Returns}
\subsubsection{GARCH model}
The auto-correlation function (ACF) of $\epsilon_t^2$ are shown in figure
\ref{fig:nordea_15min_acf}. At the absence of GARCH effects, the ACF
will have an asymptotic Gaussian distribution with mean 0 and variance
1/T. The first 5 auto-correlations have considerable sizes and do not
fall off as in the Gaussian case. This observation suggests a GARCH
model. Moreover, from \ref{fig:nordea_15min_vlt_acf}, one sees
apparent auto-correlations between volatilities of consecutive time
intervals. This again points to a GARCH model.
\begin{figure}[htb!]
  \centering
  \subfigure[ACF of squared returns]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    130]{../pics/nordea_15min_acf.pdf}
    \label{fig:nordea_15min_acf}
  }
  \subfigure[ACF of volatilities]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    130]{../pics/nordea_15min_vlt_acf.pdf}
    \label{fig:nordea_15min_vlt_acf}
  }
  \caption{Nordea Bank 15m autocorrelations}
\end{figure}

To obtain the functional form of the conditional distribution of
$\epsilon_t$, we look at the qq-plot of $\epsilon_t/\sigma_t$
\ref{fig:nordea_15min_epOversig_qq}, which is consistent with a
standard Gaussian distribution except for a few outliers at the two
ends. Hence the fat tails of the probability density function can be
entirely accounted for by GARCH effects, implying a Gaussian
conditional distribution is appropriate for $\epsilon_t$. 
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.6, clip=true, trim=100 236 118
    130]{../pics/nordea_15min_epOversig_qq.pdf}
  \caption{QQ-plot of $\epsilon_t / \sigma_t$. $\epsilon_t$ are
    derived from Nordea Bank 15m returns while $\sigma_t$ are realized
    volatilities calculated using 30s returns within each 15m
    interval.}
  \label{fig:nordea_15min_epOversig_qq}
\end{figure}

To determine the orders p and q of the GARCH model, we again check
the partial auto-correlation function of $\epsilon_t^2$. As just
mentioned, the first 5 partial auto-correlations are significant. Thus
we can tentatively estimate a GARCH(1, 5) model. Then by applying the
Lagrange multiplier test and accordingly removing insignificant
parameters, we obtain a GARCH(1, 3) model
\ref{tab:nordea_15min_garch}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{c|c|c|c}
    parameter & value & std. error & t statistic \\
    \hline
     $\alpha_0$ &   1.05668e-06 &   3.56871e-07 &    2.96096\\
     \hline
     $\beta_1$  &     0.687212  &   0.0299314   &    22.9596\\
     \hline
     $\alpha_1$ &      0.128614 &   0.0268144   &    4.79643\\
     \hline
     $\alpha_3$ &    0.0568022  &   0.0301503   &    1.88397
  \end{tabular}
  \caption{GARCH(1, 3) model of Nordea Bank 15m returns}
  \label{tab:nordea_15min_garch}
\end{table}

\subsection{15min Returns}

\bibliographystyle{plain}
\bibliography{econophysics}
\end{document}

### 2013-10-28 intra_ret_distr.m
Revise the thesis. Try fitting 45m returns' excess extra fat tail with Johnson Su distribution.

@@ -2,18 +2,16 @@ clear all
 close all
 timefmt = 'yyyy-mm-dd HH:MM:SS';
 % in units of minute
-interval = 5;
-start_day = '2012-03-16';
+dt = 45;
+start_day = '2012-01-16';
 end_day = '2012-04-20';
 company = 'nordea_bank';
 
-get_data = 0;
-
-[ret, sig] = get_intra_ret_simple(company, '2012-01-16', '2012-04-20', interval);
-r = (ret - mean(ret))./std(ret);
+[ret, sig] = get_intra_ret_simple(company, '2012-01-16', '2012-04-20', dt);
+r = (ret - mean(ret))./sig;
 
 q1 = [-2.1, -0.7, 0.7, 2.1];
-probs = normcdf(q1);
+probs = cdf('Normal', q1, 0, 1);
 q2 = quantile(r, probs);
 [R1, type1] = johnsrnd([q1; q2], 1e5, 1);
 [F1, X1] = ecdf(R1);
@@ -25,4 +23,24 @@ stairs(X1, F1, 'r');
 grid on
 hold off;
 
+sk = 0.1264;
+kts = 3.1940;
+
+f1 = @(x) - sqrt(0.5 * x(1) .* (x(1)-1)) .* (...
+     x(1) .* (x(1) + 2) .* sinh(3*x(2)) + 3*sinh(x(2))) ./ ...
+     (x(1) .* cosh(2*x(2)) + 1).^1.5;
+
+f2 = @(x) (x(1).^2 .* (x(1).^4 + 2*x(1).^3 + 3*x(1).^2 - 3) .* cosh(4*x(2))...
+      + 4 * x(1).^2 .* (x(1)+2) .* cosh(2*x(2)) + 3*(2*x(1) + 1)) ./...
+      (2 * (x(1) .* cosh(2*x(2)) + 1).^2);
+
+f = @(x) [f1(x) - sk; f2(x) - kts];
+
+g = -0.7373;
+d = 4.787;
+
+X = fsolve(f, [exp(1/d^2); g/d]);
+
+d = log(X(1)).^(-1/2);
+g = X(2) .* d;
 

### 2013-11-17 thesis/thesis.tex
Revise the thesis. Try fitting 45m returns' excess extra fat tail with Johnson Su distribution.
Revision to the thesis.

@@
 \usepackage{subfigure}
+\usepackage{amssymb}
 \usepackage{enumerate}
@@
 \chapter{Introduction}
+$\mathbb{E}(x)$
+
 \section{Stylized facts}
@@
 \section{GARCH models}
-The Auto-correlation function of $\epsilon_t^2$ is given by (c.f. \cite{Bollerslev87})
+The Auto-correlation function of $\epsilon_t^2$ is given by
+(c.f. \cite{Bollerslev87})
 $$
@@
   \includegraphics[scale=0.5, clip=true, trim=103 236 118
-  130]{../pics/nordea_price_20120116-20120420.pdf}
+  200]{../pics/nordea_price_20120116-20120420.pdf}
   \caption{Nordea Bank paid prices 2012/01/16-2012/04/20}
@@
     \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    130]{../pics/nordea_15min_acf.pdf}
+    200]{../pics/nordea_15min_acf.pdf}
     \label{fig:nordea_15min_acf}
@@
     \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    130]{../pics/nordea_15min_vlt_acf.pdf}
+    200]{../pics/nordea_15min_vlt_acf.pdf}
     \label{fig:nordea_15min_vlt_acf}
@@
 To obtain the functional form of the conditional distribution of
-$\epsilon_t$, we look at the qq-plot of $\epsilon_t/\sigma_t$
-\ref{fig:nordea_15min_epOversig_qq}, which is consistent with a
-standard Gaussian distribution except for a few outliers at the two
-ends. Hence the fat tails of the probability density function can be
-entirely accounted for by GARCH effects, implying a Gaussian
-conditional distribution is appropriate for $\epsilon_t$. 
+$\epsilon_t$ \footnote{We use the realized volatility as an estimate of the
+conditional volatility. See \cite{Andersen03}.}, we look at the
+qq-plot of $\epsilon_t/\sigma_t$ \ref{fig:nordea_15min_epOversig_qq},
+which is consistent with a standard Gaussian distribution except for a
+few outliers at the two ends. Hence the fat tails of the probability
+density function can be entirely accounted for by GARCH effects,
+implying a Gaussian conditional distribution is appropriate for $\epsilon_t$.
 \begin{figure}[htb!]
@@
     \includegraphics[scale=0.6, clip=true, trim=100 236 118
-    130]{../pics/nordea_15min_epOversig_qq.pdf}
+    200]{../pics/nordea_15min_epOversig_qq.pdf}
   \caption{QQ-plot of $\epsilon_t / \sigma_t$. $\epsilon_t$ are
@@
 
-\subsection{15min Returns}
+\subsubsection{ARFIMA(p,d,q) volatility model}
+With the availability of transactional data, estimating conditional
+volatility using returns sampled at a higher frequency (realized
+volatility) gives superior acuracy and reliability. However, at which
+frequency the time series should be sampled in order to give an
+unbiased and consistent estimate of the volatility is not a trivial
+question. Naively one would argue by the doctrine of statistics that
+the often the series is sampled, the better the estimate. This is also
+well supported in theory, since the quadratic variation which is the
+theoretial variance of the time series (see \cite{Andersen03}, theorem
+1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
+sec.6, theorem 22):
+\[
+\var(p_{t+h} - p_t) = [r, r]_{t+h} - [r, r]_t = \lim_{n \to \infty}
+\sum_{T^i \in \pi_n} (p^{T_{i+1}} - p^{T_i})^2
+\]
+where $p_t$ is the log price process and $r_t = p_t - p_0$. $\pi_n$
+is a refining partition of the interval $[t, t+h]$.
+
+However, due to the noise introduced by market micro-structure, there
+is actually an optimal sampling frequency relative to $h$. While the
+method to determine this optimal frequency is a subject of debate (see
+for example \cite{Sahalia05}), it is not difficult to find a fairly
+satisfactory frequency in practice.
+
+Andersen and Bollerslev et al proved in \cite{Andersen03} that, with
+the conditional volatility well approximated by the realized
+volatility, $r_{t+h, h} = p_{t+h} - p_t$ is governed by a Gaussian
+distribution (see theorem 2 of \cite{Andersen03}):
+\[
+r_{t+h, h} | \sigma(\mu_{t+s}, \sigma_{t+s})_{s \in [0, h]} \sim
+N(\int_{0}^h \mu_{t+s} ds, \int_{0}^h \sigma_{t+s}^2 ds)
+\]
+
+Thus one can try a few frequencies and compare the distribution of
+$r_{t+h, h}/\hat{\sigma_{t+h, h}}$ with the standard Gaussian. If the
+two match, $\hat{\sigma_{t+h, h}}$ is a good estimate of $\sigma_{t+h,
+  h}$, the volatility of $r_{t+h, h}$.
+
+For the present case with $h = 15\text{min}$, a reasonally good
+sampling frequency is 1/30sec, as confirmed by the QQ-plot
+\ref{fig:nordea_15min_epOversig_qq} as well as by the auto-correlation
+functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of $(r_{t+h, h} /
+\sigma_{t+h, h})^2$ which are shown in
+\ref{fig:nordea_15min_quotient_acf} and
+\ref{fig:nordea_15min_quotient_squared_acf} respectively.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[ACF of $r_{t+h, h} / \sigma_{t+h, h}$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_quotient_acf.pdf}
+    \label{fig:nordea_15min_quotient_acf}
+  }
+  \subfigure[ACF of $(r_{t+h, h} / \sigma_{t+h, h})^2$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
+    \label{fig:nordea_15min_quotient_squared_acf}
+  }
+\end{figure}
+
+We see in figure \ref{fig:nordea_15min_quotient_acf} and
+\ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
+auto-correlation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
+$(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
+series is i.i.d Gaussian distributed. As a result, the problem of
+modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
+  h}$. Andersen and Bollerslev et al reported (See \cite{Andersen03})
+that, for the exchange rates between Deutch mark, yen and dollar, $\ln
+\sigma_{t+h, h}$ is gaussian distributed. This is, however, not the
+case for our series in question. In fact, in our case, $\ln
+\sigma_{t+h, h}$ is right skewed (skewness 0.3342) and leptokurtic
+(kurtosis 6.1006). See figure \ref{fig:nordea_15min_logvol_pdf} and
+\ref{fig:nordea_15min_logvol_qq}
+\begin{figure}[htb!]
+  \centering
+  \subfigure[PDF of $\ln\sigma_{t+h,h}$]{
+    \includegraphics[scale=0.32, clip=true, trim=52 173 46
+    43]{../pics/nordea_15min_logvol_hist.pdf}
+    \label{fig:nordea_15min_logvol_pdf}
+  }
+  \subfigure[QQ-plot of $\ln\sigma_{t+h,h}$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_logvol_qq.pdf}
+    \label{fig:nordea_15min_logvol_qq}
+  }
+\end{figure}
+
+Moreover, the auto-correlation function of $\ln\sigma_{t+h, h}$
+shows a long memory of the time series (see figure
+\ref{fig:nordea_15min_logvol_acf} and
+\ref{fig:nordea_15min_logvol_pcf}.
+
+\subsection{45min Returns}
 

### 2013-11-19 thesis/thesis.tex
Modified scripts for investigating nordea 15min log volatility. Revised the thesis accordingly.

@@
 \section{GARCH models}
-The Auto-correlation function of $\epsilon_t^2$ is given by
+The autocorrelation function of $\epsilon_t^2$ is given by
 (c.f. \cite{Bollerslev87})
@@
 zeros. From these equations, it is clear that the partial
-auto-correlation function cuts off at $max(p, q)$.
+autocorrelation function cuts off at $max(p, q)$.
 
@@
 \subsubsection{GARCH model}
-The auto-correlation function (ACF) of $\epsilon_t^2$ are shown in figure
+The autocorrelation function (ACF) of $\epsilon_t^2$ are shown in figure
 \ref{fig:nordea_15min_acf}. At the absence of GARCH effects, the ACF
 will have an asymptotic Gaussian distribution with mean 0 and variance
-1/T. The first 5 auto-correlations have considerable sizes and do not
+1/T. The first 5 autocorrelations have considerable sizes and do not
 fall off as in the Gaussian case. This observation suggests a GARCH
 model. Moreover, from \ref{fig:nordea_15min_vlt_acf}, one sees
-apparent auto-correlations between volatilities of consecutive time
+apparent autocorrelations between volatilities of consecutive time
 intervals. This again points to a GARCH model.
@@
 To determine the orders p and q of the GARCH model, we again check
-the partial auto-correlation function of $\epsilon_t^2$. As just
-mentioned, the first 5 partial auto-correlations are significant. Thus
+the partial autocorrelation function of $\epsilon_t^2$. As just
+mentioned, the first 5 partial autocorrelations are significant. Thus
 we can tentatively estimate a GARCH(1, 5) model. Then by applying the
@@
 
-\subsubsection{ARFIMA(p,d,q) volatility model}
+\subsubsection{$(0,1,1)\times(0, 1, 1)$ seasonal volatility model}
 With the availability of transactional data, estimating conditional
@@
 sampling frequency is 1/30sec, as confirmed by the QQ-plot
-\ref{fig:nordea_15min_epOversig_qq} as well as by the auto-correlation
+\ref{fig:nordea_15min_epOversig_qq} as well as by the autocorrelation
 functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of $(r_{t+h, h} /
@@
   }
+  \caption{Nordea 15min $r_t/\sigma_t$ and $(r_t/\sigma_t)^2$ ACF.}
 \end{figure}
@@
 \ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
-auto-correlation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
+autocorrelation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
 $(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
@@
 modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
-  h}$. Andersen and Bollerslev et al reported (See \cite{Andersen03})
-that, for the exchange rates between Deutch mark, yen and dollar, $\ln
-\sigma_{t+h, h}$ is gaussian distributed. This is, however, not the
-case for our series in question. In fact, in our case, $\ln
-\sigma_{t+h, h}$ is right skewed (skewness 0.3342) and leptokurtic
-(kurtosis 6.1006). See figure \ref{fig:nordea_15min_logvol_pdf} and
+  h}$. Andersen and Bollerslev et al reported that, for the exchange
+rates between Deutch mark, yen and dollar, $\ln \sigma_{t+h, h}$ is
+gaussian distributed \cite{Andersen03}. This is, however, not the case
+for our series in question. In fact, in our case, $\ln \sigma_{t+h,
+  h}$ is right skewed (skewness 0.3342) and leptokurtic (kurtosis
+6.1006). See figure \ref{fig:nordea_15min_logvol_pdf} and
 \ref{fig:nordea_15min_logvol_qq}
@@
   }
+  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ unconditional
+    distribution}
 \end{figure}
 
-Moreover, the auto-correlation function of $\ln\sigma_{t+h, h}$
-shows a long memory of the time series (see figure
-\ref{fig:nordea_15min_logvol_acf} and
-\ref{fig:nordea_15min_logvol_pcf}.
+Moreover, the autocorrelation function of $\ln\sigma_{t+h, h}$
+shows long-lasting and periodic autocorrelations in the time series
+with an apparent period of 33 (see figure
+\ref{fig:nordea_15min_logvol_acf}). This suggests simplifying the
+series by differencing \cite{BoxJenkins94}:
+\[
+w_t = (1-B)(1-B^s)\ln\sigma_{t+h, h}
+\]
+where $B$ is the back-shift operator. The length of the time interval
+h, which is 15min in the present case, is left implicit.
 
+The autocorrelation function of the differenced process $w_t$, as shown in
+figure \ref{fig:nordea_15min_w_acf}, clearly points to a seasonal $(0,
+0, 1)\times(0, 0, 1)$ model: There are only 4 non-zero
+autocorrelations in the plot, located at lags 1, 32, 33, 34,
+respectively; furthermore, the two at 32 and 34 are approximately
+equal. Thus we can write down the model as
+\begin{eqnarray*}
+  w_t = (1 - \theta B)(1 - \Theta B^s) y_t
+\end{eqnarray*}
+where $\theta$ and $\Theta$ are parameters to be determined and $y_t$
+is a noise process with constant variance $\sigma_y^2$ and mean
+0. $y_t$ is often refered to as the residuals.
+
+The above $(0, 0, 1)\times(0, 0, 1)$ model has the following
+autocovariance structure \cite{BoxJenkins94}:
+\begin{eqnarray*}
+  \gamma_0 &=& \sigma_y^2 (1 + \theta^2)(1 + \Theta^2) \\
+  \gamma_1 &=& -\sigma_y^2\theta(1 + \Theta^2) \\
+  \gamma_s &=& -\sigma_y^2\Theta(1 + \theta^2) \\
+  \gamma_{s+1} &=& \gamma_{s-1}\;=\;\sigma_y^2\theta\Theta
+\end{eqnarray*}
+These equations together with the measured autocorrelations make
+possible an initial estimate of the parameters $\theta$ and $\Theta$:
+\begin{eqnarray*}
+  {\rho_{s+1}/\rho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
+    1 + \theta^2} \\
+  {\rho_{s+1}/\rho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
+    1 + \Theta^2} \\
+\end{eqnarray*}
+Substituting in the measured values shown in table
+\ref{tab:nordea_15min_w_acf},
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{c|c|c|c}
+    $\rho_1$ & $\rho_{s-1}$ & $\rho_s$ & $\rho_{s+1}$ \\
+    \hline
+    -0.4703 &  0.2053 & -0.4564 &  0.2212
+  \end{tabular}
+  \caption{Nordea 15min $w_t$ autocorrelations}
+  \label{tab:nordea_15min_w_acf}
+\end{table}
+we get
+\begin{eqnarray*}
+  \theta &=& 0.6890 \\
+  \Theta &=& 0.6378
+\end{eqnarray*}
+Among the two roots of each of the 2nd order equations in the above,
+we have chosen the one in the range $(-1, 1)$ so as to ensure
+invertibility of the model \cite{BoxJenkins94}.
+
+By inverting the seasonal moving average model of $w_t$ we can then
+infer the noise process $y_t$ \footnotemark, i.e. the residuals:
+\begin{eqnarray*}
+  y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
+  &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
+  w_t \\
+  &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
+  w_{t-k}
+\end{eqnarray*}
+where $\bot(\cdot)$ denotes the floor function, i.e. the largest
+integer smaller than its argument.
+
+In order to forecst the $w_t$ process, and hence the price process
+itself, we must also know the distribution of $y_t$. Moreover, to
+properly estimate the parameters of the model in the sense of maximum
+likelihood, we are also in need of the distribution of $y_t$.
+
+If $y_t$ is indeed i.i.d random variables as we assumed, we only need
+to work out its unconditional distribution.
+
+\footnotetext{
+  The autoregressive process resulting from the inversion is inifinite
+  in extent. Hence one must choose a cut-off value for k. For the
+  present case of Nordea 15min log-volatility, I choose the value
+  $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
+  $k_{\top}$, is around 550. To avoid data reduction of this amount,
+  we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
+  corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $E(w_t) = E(y_t
+  - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
+}
+\begin{figure}[htb!]
+  \centering
+  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 230 112
+    210]{../pics/nordea_15min_logvol_acf.pdf}
+    \label{fig:nordea_15min_logvol_acf}
+  }
+  \subfigure[ACF of $w_t$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 230 112
+    210]{../pics/nordea_15min_w_acf.pdf}
+    \label{fig:nordea_15min_w_acf}
+  }
+  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$ autocorrelations}
+\end{figure}
 \subsection{45min Returns}

### 2013-11-19 nonlcon.m

@@ -1,0 +1,15 @@
+function [C, Ceq] = nonlcon(X)
+minv = 8.0e-2;
+maxv = 1.0e-1;
+ep = 1.0e-3;
+
+mmt = johnson_su_moments12(X(1:4));
+
+% The variance of the innovations must be in [8.0e-2, 1.0e-1];
+% The mean of the innovations must be within [-ep, ep]
+C(1) = minv + ep - mmt(2);
+C(2) = mmt(2) - maxv + ep;
+C(3) = mmt(1) - ep;
+C(4) = -ep - mmt(1);
+
+Ceq = 0;

### 2013-11-19 test.m

@@ -21,25 +21,18 @@ lv = lv - mu;
 s = 33;
 w = ts_difference(lv, s, 1);
 w = ts_difference(w, 1, 1);
-acf = autocorr(w, 120) .* var(w);
+acf = autocorr(w, 120);
 acf = acf(2:end);
 
 % Nordea
 
 if 1
     theta = zeros(1, s+1);
-    % theta(1) = 0.6263;
-    % theta(s) = 0.5870;
-    theta(1) = 0.845552;
-    theta(s) = 0.804139;
-    % theta(1) = 0.878236;
-    % theta(s) = 0.912821;
+    m = mean([acf(s-1), acf(s+1)]);
+    theta(1) = fzero(@(x) (1+x^2)*m/acf(s) + x, 0);
+    theta(s) = fzero(@(x) (1+x^2)*m/acf(1) + x, 0);
     theta(s+1) = theta(1)*theta(s);
-    theta(s-1) = theta(s+1);
-    V = mean([acf(s-1), acf(s+1)])/theta(1)/theta(s);
-    
     y = ma_infer(w, theta(1), theta(s), s);
-
 else
 
     % Assume Gaussian distribution to obtain
@@ -53,12 +46,12 @@ else
 end
 
 param = NaN(1, 6);
-param(5) = model.MA{1};
-param(6) = model.SMA{33};
+param(5) = theta(1);
+param(6) = theta(s);
 mmt = [mean(y), var(y), skewness(y), kurtosis(y)];
 
 func = @(x) johnson_su_moments34(x) - mmt(3:4);
-param([1:2]) = fsolve(func, [-0.1922, 1.445]);
+param([1:2]) = fsolve(func, [-0.1461, 1.513]);
 m = johnson_su_moments12([param([1:2]), 1, 0]);
 param(3) = sqrt(mmt(2)/m(2));
 param(4) = mmt(1) - m(1)*param(3);
@@ -66,15 +59,15 @@ type = cmp_johnson_su(y);
 
 func = @(x) objective(x, w);
 
+% theta & Theta must be in (-1, 1)
 lb = -ones(1, 6) .* Inf;
 lb(5:6) = [-0.99, -0.99];
 ub = ones(1, 6) .* Inf;
 ub(5:6) = [0.99, 0.99];
 
-% options = optimoptions('fmincon','GradObj','on');
-% problem = struct('objective', func, 'x0', param, 'lb', lb, 'ub', ub, ...
-%                  'options', options);
[…]

### 2013-12-13 thesis/thesis.tex
New scripts.

@@
 $$
-\rho_n = \sum_{i=1}^{max(p,q)} (\alpha_i + \beta_i) \rho_{n-i}
+\rho_n = \sum_{i=1}^{p \vee q} (\alpha_i + \beta_i) \rho_{n-i}
 \;n > p
@@
 where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
-zeros. From these equations, it is clear that the partial
-autocorrelation function cuts off at $max(p, q)$.
+zeros. $p \vee q$ denotes the maximum of p and q. From these
+equations, it is clear that the partial autocorrelation function cuts
+off at $max(p, q)$.
 
@@
 will have an asymptotic Gaussian distribution with mean 0 and variance
-1/T. The first 5 autocorrelations have considerable sizes and do not
-fall off as in the Gaussian case. This observation suggests a GARCH
-model. Moreover, from \ref{fig:nordea_15min_vlt_acf}, one sees
-apparent autocorrelations between volatilities of consecutive time
-intervals. This again points to a GARCH model.
+1/T \cite{Bollerslev86, Bollerslev87}. The first 5 autocorrelations
+have considerable sizes and do not fall off as in the Gaussian
+case. This observation suggests a GARCH model. Moreover, from
+\ref{fig:nordea_15min_vlt_acf}, one sees apparent autocorrelations
+between volatilities of consecutive time intervals. This again points
+to a GARCH model.
 \begin{figure}[htb!]
@@
   }
-  \caption{Nordea Bank 15m autocorrelations}
+  \caption{Nordea Bank 15min autocorrelations}
 \end{figure}
@@
   \caption{QQ-plot of $\epsilon_t / \sigma_t$. $\epsilon_t$ are
-    derived from Nordea Bank 15m returns while $\sigma_t$ are realized
-    volatilities calculated using 30s returns within each 15m
+    derived from Nordea Bank 15min returns while $\sigma_t$ are realized
+    volatilities calculated using 30s returns within each 15min
     interval.}
@@
   \end{tabular}
-  \caption{GARCH(1, 3) model of Nordea Bank 15m returns}
+  \caption{GARCH(1, 3) model of Nordea Bank 15min returns}
   \label{tab:nordea_15min_garch}
@@
     \includegraphics[scale=0.32, clip=true, trim=52 173 46
-    43]{../pics/nordea_15min_logvol_hist.pdf}
+    113]{../pics/nordea_15min_logvol_hist.pdf}
     \label{fig:nordea_15min_logvol_pdf}
@@
     \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_logvol_qq.pdf}
+    220]{../pics/nordea_15min_logvol_qq.pdf}
     \label{fig:nordea_15min_logvol_qq}
@@
 0. $y_t$ is often refered to as the residuals.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 230 112
+    235]{../pics/nordea_15min_logvol_acf.pdf}
+    \label{fig:nordea_15min_logvol_acf}
+  }
+  \subfigure[ACF of $w_t$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 230 112
+    235]{../pics/nordea_15min_w_acf.pdf}
+    \label{fig:nordea_15min_w_acf}
+  }
+  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$ autocorrelations}
+\end{figure}
 
@@
 
-By inverting the seasonal moving average model of $w_t$ we can then
-infer the noise process $y_t$ \footnotemark, i.e. the residuals:
-\begin{eqnarray*}
-  y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
-  &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
-  w_t \\
-  &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
-  w_{t-k}
-\end{eqnarray*}
-where $\bot(\cdot)$ denotes the floor function, i.e. the largest
-integer smaller than its argument.
+With an estimate of $\theta$ and $\Theta$, one can then infer the
+noise process i.e. the residuals $y_t$:
+\begin{equation}
+  \label{eq:infer_y}
+  y_t = w_t + \theta y_{t-1} + \Theta y_{t-s} - \theta \Theta y_{t-s-1}  
+\end{equation}
+where we substitute $y_t\;(t \leq 0)$ with their unconditional
+expectation 0.
+% By inverting the seasonal moving average model of $w_t$ we can then
+% infer the noise process $y_t$ \footnotemark, i.e. the residuals:
+% \begin{eqnarray*}
+%   y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
+%   &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
+%   w_t \\
+%   &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
+%   w_{t-k}
+% \end{eqnarray*}
+% where $\bot(\cdot)$ denotes the floor function, i.e. the largest
+% integer smaller than its argument.
+% \footnotetext{
+%   The autoregressive process resulting from the inversion is inifinite
+%   in extent. Hence one must choose a cut-off value for k. For the
+%   present case of Nordea 15min log-volatility, I choose the value
+%   $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
+%   $k_{\top}$, is around 550. To avoid data reduction of this amount,
+%   we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
+%   corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $E(w_t) = E(y_t
+%   - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
+% }
 
@@
 If $y_t$ is indeed i.i.d random variables as we assumed, we only need
-to work out its unconditional distribution.
-
-\footnotetext{
-  The autoregressive process resulting from the inversion is inifinite
-  in extent. Hence one must choose a cut-off value for k. For the
-  present case of Nordea 15min log-volatility, I choose the value
-  $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
-  $k_{\top}$, is around 550. To avoid data reduction of this amount,
-  we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
-  corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $E(w_t) = E(y_t
-  - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
-}
+to work out its unconditional distribution. This is shown in figure
+\ref{fig:nordea_15min_y_qq} and \ref{fig:nordea_15min_y_acf}.
 \begin{figure}[htb!]
   \centering
-  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
+  \subfigure[QQ-plot of $y_t$]{
     \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_logvol_acf.pdf}
-    \label{fig:nordea_15min_logvol_acf}
+    210]{../pics/nordea_15min_y_qq.pdf}
+    \label{fig:nordea_15min_y_qq}
   }
-  \subfigure[ACF of $w_t$]{
+  \subfigure[ACF of $y_t$]{
     \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_w_acf.pdf}
-    \label{fig:nordea_15min_w_acf}
+    210]{../pics/nordea_15min_y_acf.pdf}
+    \label{fig:nordea_15min_y_acf}
   }
-  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$ autocorrelations}
+  \caption{Nordea 15min $y_t$ QQ-plot and autocorrelations}
 \end{figure}
+In addition, the measured moments of $y_t$ are shown in table
+\ref{tab:nordea_15min_y_moments}:
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{c|c|c|c}
+    mean & variance & skewness & kurtosis \\
+    \hline
+    0.0012 & 0.0935 & 0.2988 & 6.8691
+  \end{tabular}
+  \caption{Nordea 15min $y_t$ moments}
+  \label{tab:nordea_15min_y_moments}
+\end{table}
+
+If we ignore the weak autocorrelations in this crude estimate of
+$y_t$ and assume $y_t$ to be i.i.d random variables governed
+by a skewed and leptokertic distribution, the parameters of this
+distribution together with $\theta$ and $\Theta$ can be determined via
+maximum likelihood estimate (MLE).
+
+It turns out that $y_t$ can be well described by a Johnson Su
+distribution \cite{Shang2004}:
+\[
+  y_t = \xi + \lambda\sinh{z_t - \gamma \over \delta}
+\]
+where $\gamma, \delta, \lambda, \xi$ are parameters to be determined
+and $z_t \sim N(0, 1)$. The goodness of fitting is demonstrated in
+figure \ref{fig:nordea_15min_y_js_fit} by the
+empirical cummulative distribution function in comparison to the
+theoretical one.
+\begin{figure}[htb!]
+  \centering
+    \includegraphics[scale=0.5, clip=true, trim=112 272 107
+    140]{../pics/nordea_15min_y_js_fit.pdf}
+    \caption{Nordea 15min $y_t$ fitted to a Johnson Su distribution}
+    \label{fig:nordea_15min_y_js_fit}
+\end{figure}
+
+The first 4 moments of the Johnson Su distribution are expressible
+in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}:
+\begin{eqnarray*}
+  w &=& \exp{1 \over \delta^2} \\
+  \Omega &=& {\gamma \over \delta} \\
+  \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
+  \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
+    1)\right]^{1/2} \\
+  \text{skewness}(y) &=& {
+    \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
+    \over
+    (w\cosh 2\Omega + 1)^{3/2}} \\
+  \text{kurtosis}(y) &=& {
+    w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
+    + 3(2w+1) \over
+    2(w\cosh 2\Omega + 1)^2 }
+\end{eqnarray*}
+By matching the theoretical expressions of the moments with their
+measured values, and taking help from published tables
+\cite{Johnson1965}, one can solve for the parameters $\gamma, \delta,
+\lambda, \xi$.
+
+Under the assumption of i.i.d Johnson Su distributed residuals, the
+log-likelihood function of the parameters $\gamma, \delta$
+conditional on the sample $w_t$ can be written as
+\[
+L(\theta, \Theta) = -{1 \over 2}\sum_{t=1}^n z_t^2 + n \ln{\delta
+  \over \lambda \sqrt{2\pi}} - {1 \over 2}\sum_{t=1}^n \ln\left[
+  1 + \left({y_t - \xi \over \lambda}\right)^2
+\right]
+\]
+where $y_t$ are infered from $w_t$ using eq.\ref{eq:infer_y}
+and $z_t$ from $y_t$ using
+\[
+z_t = \delta \sinh^{-1}{y - \xi \over \lambda} + \gamma
+\]
+Note that $\gamma, \delta, \lambda, \xi$ are not really free
+parameters but rather are implied by $\theta$ and $\Theta$: Once the
+latter have been chosen and the corresponding $y_t$ infered, the
+former are determined by the moments of $y_t$.
+
+The eventual MLE is done in Matlab with the ``active set''
+algorithm. The initial as well as the final estimation results are
+listed in table \ref{tab:nordea_15min_js_param}:
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{c|c|c|c|c|c|c}
+    & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
+    \hline
+    initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
+    0.6378 \\
+    \hline
+    MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
+  \end{tabular}
+  \caption{Nordea 15min estimation results}
+  \label{tab:nordea_15min_js_param}
+\end{table}
+
+The probability density function of the residuals $y_t$ that result
+from the MLE estimate is shown in figure
+\ref{fig:nordea_15min_y1_hist} together with its counterpart from the
+initial estimate. 
 \subsection{45min Returns}

### 2013-12-13 LevyMarkov.m

@@ -1,0 +1,34 @@
+clear all
+close all
+
+N = 50;
+T = 1e3;
+fold = 2000;
+alpha = 3/2;
+gam = 1;
+dist = struct('name', 'Levy', 'prmt', [alpha, gam]);
+phi = [0:5] * 0.15;
+spec = ['b', 'c', 'g', 'm', 'r', 'k'];
+for i = 1:length(phi)
+    if (exist(sprintf('./MarkovEig-%d.mat', i-1), 'file') == 2)
+        continue;
+    end
+    ev = NaN(N, fold);
+    for j = 1:fold
+        R = gen_ret_mtx(N, T, dist, [phi(i)]);
+        s = diag(1./std(R'));
+        M = s * R;
+        C = T.^(-2/alpha) * M*M';
+        ev(:, j) = eig(C);
+    end
+    save(sprintf('MarkovEig-%d.mat', i-1), 'ev');
+end
+
+for i = 1:length(phi)
+    load(sprintf('MarkovEig-%d.mat', i-1), 'ev');
+    maxeig = max(ev);
+    [y, x] = ecdf(maxeig);
+    plot(log(x), log(y), spec(i));
+    hold on
+end
+hold off

### 2013-12-13 interday_ret_distr.m

@@ -2,33 +2,70 @@ clear all
 close all
 % Volvo B 2007-06-01 -- 2008-12-31 down
 % Volvo B 2009-01-01 -- 2010-12-31 up
-xi = 2.0512;
+% filename = 'data/volvo_b_20121203-20131202.csv';
 
-filename = 'data/ericsson_b.csv';
-price = dlmread(filename, ',', [1, 4, 3560, 4]);
+%% daily returns
+filename = 'data/nasdaq.csv';
+price = dlmread(filename, ',', [1, 4, 10807, 4]);
+% filename = 'data/sp.csv';
+% price = dlmread(filename, ',', [1, 4, 16084, 4]);
 price = flipud(price); 
-
-
 ret = price2ret(price);
 r = (ret - mean(ret))/std(ret);
 
-figure
-[f, x] = ecdf(r);
-plot(x, f, 'b');
-R = randn(1e4, 1);
-hold on
-[f, x] = ecdf(R);
-plot(x, f, 'r');
-hold off
-grid on
-title('CDF of volvo daily ret. dist.');
 
-figure
-qqplot(r);
-grid on
+%% intraday returns
+% company = 'volvo_b';
+% first_day = '2013-10-10';
+% last_day = '2013-12-04';
+% dt = 5;
+% % interval for calculating realized volatility. in seconds
+% delta = 10;
+% [ret, v] = get_intra_ret_simple(...
+%     company, first_day, last_day, dt, delta);
+% r = (ret - mean(ret))/std(ret);
 
-[p1, p2] = extract_xpnt(r);
-fprintf('xpnt1=%f, xpnt2=%f\n', p1, p2);
+[p1, p2] = extract_xpnt(r, 2.56e-2/std(ret), 5.0e-2/std(ret));
+% [p1, p2] = extract_xpnt(r, 0.045/std(ret));
+fprintf('Left tail exponent: %.4f\nRight tail exponent: %.4f\n', ...
+        p1, p2);
+
+% figure;
+
+% p = hist(r, x)/length(r)/(x(2) - x(1));
+% plot(x, p);
+
+% js = johnson_su_params([mean(r), var(r), skewness(r), ...
+%                     kurtosis(r)]);
+% pt = johnson_su_pdf(johnson_su_struct(js), x);
+% plot(x, pt, 'g');
+
+
+% param = mle(r, 'pdf',...
+%             @(x, xi1, xi2)...
+%             NormalPowerLaw_pdf([xi1, xi2], x),...
[…]

### 2013-12-13 johnson_su_param_array.m

@@ -1,0 +1,3 @@
+function param = johnson_su_param_array(js)
+    param = [js.Gamma, js.Delta, js.Lambda, js.Xi];
+    

### 2013-12-13 loglikelihood.m

@@ -6,18 +6,34 @@ function l = loglikelihood(param, obs)
 n = length(obs);
 s = 33;
 
-a = param(1);
-b = param(2);
-c = param(3);
-m = param(4);
-theta = param(5);
-Theta = param(6);
+% a: gamma
+% b: delta
+% c: lambda
+% m: xi
+% a = param(1);
+% b = param(2);
+% c = param(3);
+% m = param(4);
+theta = param(1);
+Theta = param(2);
 % mmt = johnson_su_moments12(param);
 
 y = ma_infer(obs, theta, Theta, s);
+moments = [mean(y), var(y), skewness(y), kurtosis(y)];
+if (moments(3) < 0 || moments(3) > 0.5 ||...
+    moments(4) < 3.6 || moments(4) > 8.0)
+    l = -Inf;
+    return;
+end
+
+jsp = johnson_su_params(moments);
+a = jsp(1);
+b = jsp(2);
+c = jsp(3);
+m = jsp(4);
 
 z = b*asinh((y - m)/c) + a;
-l = -z'*z/2 -sum(log(1 + ((y - m)/c).^2))/2;
+l = n*log(b/c) - z'*z/2  - sum(log(1 + ((y - m)/c).^2))/2 - n*log(2*pi)/2;
 
 % gradz = johnson_su_grad(param([1:4]), y, z);
 % grad = NaN(4, 1);

### 2013-12-13 tmp.m

@@ -1,270 +1,11 @@
-A = NaN(21, 11);
-A(:, 1) = [
-    1.0000
-    0.0976
-    0.0992
-    0.0665
-    0.0570
-    0.0598
-    0.0177
-    0.0502
-    0.0183
-    0.0508
-    0.0253
-    0.0014
-    0.0263
-    0.0408
-    0.0852
-    0.0273
-    0.0532
-    0.0204
-    0.0063
-    0.0059
-   -0.0325
-          ];
+clear all
+m = 902;
+l = 49;
 
+k = 1:m/2;
+s1 = prod(2*k ./ (2*k-1));
 
-A(:, 2) =[
-    1.0000
-    0.1041
-    0.0623
-    0.0732
-    0.0618
-    0.0620
-    0.0240
-    0.0286
-    0.0541
-    0.0178
-    0.0053
-    0.0538
-   -0.0092
-   -0.0187
-   -0.0192
-    0.0168
-   -0.0470
-   -0.0067
-    0.0182
-    0.0284
-   -0.0192
-         ];
+k=1:(l+1)/2;
+s2 = prod((m+2*k-2) ./ (m+l+2*k));
 
-
-A(:, 3) = [
-    1.0000
-    0.1626
-    0.0874
-    0.0538
-    0.0762
-    0.0302
-    0.0260
-    0.0089
-    0.0019
-    0.0527
-   -0.0111
-   -0.0060
-   -0.0421
-    0.0113
-    0.0159
-   -0.0253
-   -0.0017
-    0.0119
-    0.0086
-    0.0280
-    0.0508
-          ];
-
-A(:, 4) = [
-    1.0000
-    0.0752
-    0.0502
-    0.0659
-    0.0538
-    0.0268
-    0.0150
-    0.0264
-   -0.0122
-    0.0029
-   -0.0346
-    0.0373
-   -0.0433
-    0.0135
-    0.0247
-    0.0612
-    0.0332
-    0.0232
-    0.0180
-    0.0763
-    0.0371
-          ];
-
-
-A(:, 5) = [
-    1.0000
-    0.0830
-    0.0495
-    0.0853
-    0.0009
-    0.0784
-   -0.0051
-   -0.0084
-   -0.0218
-   -0.0252
-   -0.0004
-   -0.0056
-    0.0356
-    0.0898
-    0.0302
-    0.0208
-    0.1037
-    0.0579
-    0.1132
-    0.0424
-    0.0424
-          ];
-
[…]

### 2013-12-13 uc_wishart_pdf.m

@@ -1,0 +1,5 @@
+function p = uc_wishart_pdf(num, u)
+%% computes the probability density at u of the num-th largest eigen
+
+
+

### 2013-12-13 volvo_2hr.m

@@ -1,0 +1,38 @@
+clear all
+close all
+
+company = 'volvo_b';
+first_day = '2013-10-10';
+last_day = '2013-12-04';
+% dt in units of one minute
+% dt = 15min, delta=30sec
+% dt = 45min, delta=90sec,
+% dt = 30min, delta=60sec
+dt = 102;
+% interval for calculating realized volatility. in seconds
+delta = 204;
+
+s = 4;
+h = 0;
+
+[ret_a, v_a] = get_intra_ret_simple(...
+    company, first_day, last_day, dt, delta);
+ret = ret_a(1:end-h);
+lv_a = log(v_a);
+
+v = v_a(1:end-h);
+lv = lv_a(1:end-h);
+
+clear model0;
+model0 = arima('ARLags', [1:5], 'SARLags', [1:6]*4,...
+               'Distribution', 'Gaussian');
+load('volvo_2hr.mat', 'info');
+info.options.MaxFunEvals = 1e4;
+% [model0, VarCov, LogL, info] = ...
+%     estimate(model0, lv);
+[model0, VarCov, LogL, info] = ...
+    estimate(model0, lv, 'options', info.options);
+[Y, V] = infer(model0, lv);
+save('volvo_2hr.mat', 'info');
+
+save('/tmp/models.mat', 'model0');

### 2014-02-07 thesis/_region_.tex
A lot of things...

--- 2013-11-17 thesis/thesis.tex
@@
-\documentclass{book}
+\message{ !name(thesis.tex)}\documentclass{book}
 \usepackage{graphicx}
 \usepackage{subfigure}
+\usepackage{wrapfig}
 \usepackage{amssymb}
+\usepackage{amsfonts, amsmath}
+\usepackage{mathrsfs}
 \usepackage{enumerate}
@@
 
-\maketitle
-\tableofcontents
+\message{ !name(thesis.tex) !offset(1382) }
+\cite{NIST:DLMF}:
 
-\chapter{Introduction}
-$\mathbb{E}(x)$
+\message{ !name(thesis.tex) !offset(1837) }
 
-\section{Stylized facts}
-
-\section{GARCH models}
-The Auto-correlation function of $\epsilon_t^2$ is given by
-(c.f. \cite{Bollerslev87})
-$$
-\rho_n = \sum_{i=1}^{max(p,q)} (\alpha_i + \beta_i) \rho_{n-i}
-\;n > p
-$$
-where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
-zeros. From these equations, it is clear that the partial
-auto-correlation function cuts off at $max(p, q)$.
-
-\chapter{Case Study of Stocks}
-\section{Nordea Bank}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=103 236 118
-  200]{../pics/nordea_price_20120116-20120420.pdf}
-  \caption{Nordea Bank paid prices 2012/01/16-2012/04/20}
-  \label{fig:Nordea}
-\end{figure}
-Figure \ref{fig:Nordea} shows the paid prices of Nordea Bank from
-2012/01/16 to 2012/04/20.
-
-\subsection{15min Returns}
-\subsubsection{GARCH model}
-The auto-correlation function (ACF) of $\epsilon_t^2$ are shown in figure
-\ref{fig:nordea_15min_acf}. At the absence of GARCH effects, the ACF
-will have an asymptotic Gaussian distribution with mean 0 and variance
-1/T. The first 5 auto-correlations have considerable sizes and do not
-fall off as in the Gaussian case. This observation suggests a GARCH
-model. Moreover, from \ref{fig:nordea_15min_vlt_acf}, one sees
-apparent auto-correlations between volatilities of consecutive time
-intervals. This again points to a GARCH model.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of squared returns]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_acf.pdf}
-    \label{fig:nordea_15min_acf}
-  }
-  \subfigure[ACF of volatilities]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_vlt_acf.pdf}
-    \label{fig:nordea_15min_vlt_acf}
-  }
-  \caption{Nordea Bank 15m autocorrelations}
-\end{figure}
-
-To obtain the functional form of the conditional distribution of
-$\epsilon_t$ \footnote{We use the realized volatility as an estimate of the
-conditional volatility. See \cite{Andersen03}.}, we look at the
-qq-plot of $\epsilon_t/\sigma_t$ \ref{fig:nordea_15min_epOversig_qq},
-which is consistent with a standard Gaussian distribution except for a
-few outliers at the two ends. Hence the fat tails of the probability
-density function can be entirely accounted for by GARCH effects,
-implying a Gaussian conditional distribution is appropriate for $\epsilon_t$.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.6, clip=true, trim=100 236 118
-    200]{../pics/nordea_15min_epOversig_qq.pdf}
-  \caption{QQ-plot of $\epsilon_t / \sigma_t$. $\epsilon_t$ are
-    derived from Nordea Bank 15m returns while $\sigma_t$ are realized
-    volatilities calculated using 30s returns within each 15m
-    interval.}
-  \label{fig:nordea_15min_epOversig_qq}
-\end{figure}
-
-To determine the orders p and q of the GARCH model, we again check
-the partial auto-correlation function of $\epsilon_t^2$. As just
-mentioned, the first 5 partial auto-correlations are significant. Thus
-we can tentatively estimate a GARCH(1, 5) model. Then by applying the
-Lagrange multiplier test and accordingly removing insignificant
-parameters, we obtain a GARCH(1, 3) model
-\ref{tab:nordea_15min_garch}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{c|c|c|c}
-    parameter & value & std. error & t statistic \\
-    \hline
-     $\alpha_0$ &   1.05668e-06 &   3.56871e-07 &    2.96096\\
-     \hline
-     $\beta_1$  &     0.687212  &   0.0299314   &    22.9596\\
-     \hline
-     $\alpha_1$ &      0.128614 &   0.0268144   &    4.79643\\
-     \hline
-     $\alpha_3$ &    0.0568022  &   0.0301503   &    1.88397
-  \end{tabular}
-  \caption{GARCH(1, 3) model of Nordea Bank 15m returns}
-  \label{tab:nordea_15min_garch}
-\end{table}
-
-\subsubsection{ARFIMA(p,d,q) volatility model}
-With the availability of transactional data, estimating conditional
-volatility using returns sampled at a higher frequency (realized
-volatility) gives superior acuracy and reliability. However, at which
-frequency the time series should be sampled in order to give an
-unbiased and consistent estimate of the volatility is not a trivial
-question. Naively one would argue by the doctrine of statistics that
-the often the series is sampled, the better the estimate. This is also
-well supported in theory, since the quadratic variation which is the
-theoretial variance of the time series (see \cite{Andersen03}, theorem
-1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
-sec.6, theorem 22):
-\[
-\var(p_{t+h} - p_t) = [r, r]_{t+h} - [r, r]_t = \lim_{n \to \infty}
-\sum_{T^i \in \pi_n} (p^{T_{i+1}} - p^{T_i})^2
-\]
-where $p_t$ is the log price process and $r_t = p_t - p_0$. $\pi_n$
-is a refining partition of the interval $[t, t+h]$.
-
-However, due to the noise introduced by market micro-structure, there
-is actually an optimal sampling frequency relative to $h$. While the
-method to determine this optimal frequency is a subject of debate (see
-for example \cite{Sahalia05}), it is not difficult to find a fairly
-satisfactory frequency in practice.
-
-Andersen and Bollerslev et al proved in \cite{Andersen03} that, with
-the conditional volatility well approximated by the realized
-volatility, $r_{t+h, h} = p_{t+h} - p_t$ is governed by a Gaussian
-distribution (see theorem 2 of \cite{Andersen03}):
-\[
-r_{t+h, h} | \sigma(\mu_{t+s}, \sigma_{t+s})_{s \in [0, h]} \sim
-N(\int_{0}^h \mu_{t+s} ds, \int_{0}^h \sigma_{t+s}^2 ds)
-\]
-
-Thus one can try a few frequencies and compare the distribution of
-$r_{t+h, h}/\hat{\sigma_{t+h, h}}$ with the standard Gaussian. If the
-two match, $\hat{\sigma_{t+h, h}}$ is a good estimate of $\sigma_{t+h,
-  h}$, the volatility of $r_{t+h, h}$.
-
-For the present case with $h = 15\text{min}$, a reasonally good
-sampling frequency is 1/30sec, as confirmed by the QQ-plot
-\ref{fig:nordea_15min_epOversig_qq} as well as by the auto-correlation
-functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of $(r_{t+h, h} /
-\sigma_{t+h, h})^2$ which are shown in
-\ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} respectively.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of $r_{t+h, h} / \sigma_{t+h, h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_acf.pdf}
-    \label{fig:nordea_15min_quotient_acf}
-  }
-  \subfigure[ACF of $(r_{t+h, h} / \sigma_{t+h, h})^2$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
-    \label{fig:nordea_15min_quotient_squared_acf}
-  }
-\end{figure}
-
-We see in figure \ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
-auto-correlation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
-$(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
-series is i.i.d Gaussian distributed. As a result, the problem of
-modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
-  h}$. Andersen and Bollerslev et al reported (See \cite{Andersen03})
-that, for the exchange rates between Deutch mark, yen and dollar, $\ln
-\sigma_{t+h, h}$ is gaussian distributed. This is, however, not the
-case for our series in question. In fact, in our case, $\ln
-\sigma_{t+h, h}$ is right skewed (skewness 0.3342) and leptokurtic
-(kurtosis 6.1006). See figure \ref{fig:nordea_15min_logvol_pdf} and
-\ref{fig:nordea_15min_logvol_qq}
-\begin{figure}[htb!]
-  \centering
-  \subfigure[PDF of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.32, clip=true, trim=52 173 46
-    43]{../pics/nordea_15min_logvol_hist.pdf}
-    \label{fig:nordea_15min_logvol_pdf}
-  }
-  \subfigure[QQ-plot of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_logvol_qq.pdf}
-    \label{fig:nordea_15min_logvol_qq}
-  }
-\end{figure}
-
-Moreover, the auto-correlation function of $\ln\sigma_{t+h, h}$
-shows a long memory of the time series (see figure
-\ref{fig:nordea_15min_logvol_acf} and
-\ref{fig:nordea_15min_logvol_pcf}.
-
-\subsection{45min Returns}
-
-\bibliographystyle{plain}
-\bibliography{econophysics}
 \end{document}

### 2014-02-07 thesis/thesis.tex
A lot of things...

@@
 \usepackage{subfigure}
+\usepackage{wrapfig}
 \usepackage{amssymb}
+\usepackage{amsfonts, amsmath}
+\usepackage{mathrsfs}
 \usepackage{enumerate}
@@
 \chapter{Introduction}
-$\mathbb{E}(x)$
+The mathematics of the financial market has always been a topic that
+arouses interest and imagination, and with little doubt, has been
+studied from many aspects and in many different ways.
 
-\section{Stylized facts}
+Central to all these studies are the concepts of probability
+distributions and correlations. The values of stocks, futures,
+options, etc. are stochastic in nature and are governed by the laws of
+stochastic processes, which are expressed in terms of probability
+distributions and correlations.
+
+On the other hand, the observables of the market are prices, volumes,
+turn-over (the amount of money paid in a trade), names of the brokers,
+and time of the trades. These quantities don't make much sense
+by themselves but do reveal the probabilistic dynamics of the market
+when put together and turned into statistics.
+
+Here by dynamics we mean how the value of an asset is affected by its
+own history as well as by the histories and current values of other
+assets in the market. The influence from the asset's own history is
+termed autocorrelations, i.e. correlations in time, while
+the influence from other assets are termed cross correlations.
+
+Mathematical models that describe the aforementioned observables in
+terms of probability distributions, autocorrelations and cross
+correlations give predictive power and help determine a fair price of
+a given asset. Hence they are a central topic of the mathematics of
+the financial market.
+
+Once a model has been tentatively constructed, it must be
+verified. The verification again relies on the statistics of the
+observables. Over the years, a number of phenomena have been
+consistently reported and become generally accepted. These are referred
+to as stylized facts, which are detailed in \S
+\ref{sec:StylizedFacts} and constitute a standardized test suit for
+model verification.
+
+In \S \ref{sec:FundamentalConcepts} we present a few very
+important concepts and notations that we will often refer to in later
+chapters. Then the rest of this thesis is organized as follows:
+
+\section{Fundamental Concepts \& Notations}\label{sec:FundamentalConcepts}
+This section is a list of a few concepts that may be unfamiliar to the
+reader and that we will often refer to later in the thesis:
+\begin{itemize}
+\item Return. Given a fixed time interval $[t, t']$, for example, a
+  day, a week, a month, etc, and a particular asset, for example, a
+  share in company ABC, the return of this asset over the time
+  interval is defined as
+  \begin{eqnarray*}
+    r_{t', \Delta t} &=& \ln p_{t'} - \ln p_t \\
+    &=& \ln \left(1 + {p_{t'} - p_t \over p_t}\right) \\
+    &\approx& {p_{t'} - p_t \over p_t}
+  \end{eqnarray*}
+  where $p_t$ is the price of the asset at time $t$ and $\Delta t = t'
+  - t$ is sometimes called the time-lag of the return. Quite often,
+  where confusion is not possible, we will just write $r_t$ to mean
+  $r_{t, \Delta t}$, the time lag either does not matter or is clear
+  from the context.
+
+\item Autocorrelation. By autocorrelation, denoted $\rho_k$ here, we
+  mean the correlation between two temporally separated observations
+  of the same time series:
+  \begin{eqnarray*}
+    \rho_k &=& {
+      E\left[(a_t -E(a_t))(a_{t-k} - E(a_{t-k})\right]
+      \over
+      \sqrt{\text{var}(a_t)}\sqrt{\text{var}(a_{t-k})}
+    }
+  \end{eqnarray*}
+  Here $E(x)$ stands for the expectation value of $x$, and
+  $\text{var}(x)$ stands for the variance of $x$. $k$ is called the
+  time-lag and is the temporal seperation of the two observations
+  measured by the number of observations in between. For example, the
+  time-lag between the 1st and 3rd observation is 2.
+
+\item Cross-Correlation Matrix. Correlations between the returns of a
+  group of assets are described by the cross-correlation matrix. When
+  the asset returns are described by a stable distribution law with
+  L\'evy index $\alpha$, the empirical cross-correlation matrix is
+  constructed as
+  \begin{equation}
+    \label{eq:cross-correlation}
+    \begin{aligned}
+      C_{ij} &= {1 \over T^{2/\alpha}} \sum_{t=1}^T [r_{i,t}-E(r_i)]
+      [r_{j,t}-E(r_j)] \\
+      &= {1 \over T^{2/\alpha}} RR'
+    \end{aligned}
+  \end{equation}
+  where $r_{i,t}$ is the return of the i-th asset at time t and is
+  placed at the entry (i, t) of matrix R; $R'$ denotes the
+  transpose of R. In most practical situations, one has abundant data
+  for each and every asset. Thus $T \geq N$ is assumed throughout this
+  thesis.
+
+\item Auto-regressive processes. A time series $r_t$ is called an
+  auto-regressive process of order $p$ and denoted AR(p), if it can be
+  written in the following form:
+  \begin{eqnarray*}
+    r_t &=& \sum_{i=1}^p \phi_i r_{t-i} + a_t
+  \end{eqnarray*}
+  where, for all $i$, $\phi_i \in (-1, 1)$, and the $a_t$'s are
+  independent and identically distributed (iid.) random variables with
+  zero mean. Obviously this implies the mean of $r_t$ is zero
+  too. Apart from this, their distribution of $a_t$ is not restricted
+  to any particular form.
+
+  Of particular interest to this thesis is the AR(1) process:
+  \begin{eqnarray*}
+    r_t = \phi r_{t-1} + a_t
+  \end{eqnarray*}
+  Its autocorrelation function $\rho_k = \text{corr}(r_t, r_{t-k})$
+  ($k = 0, 1, 2, \cdots$) can be easily shown to fall off
+  exponentially:
+  \begin{eqnarray*}
+    \rho_k &=& {E(r_tr_{t-k}) - E(r_{t-k})E(r_{t}) \over
+      \sqrt{\var(r_t) \var(r_{t-k})}} \\
+    &=& {\phi E(r_{t-1}r_{t-k}) + E(a_t r_{t-k})
+      \over
+      \sqrt{\var(r_t) \var(r_{t-k})}
+    } \\
+    &=& \phi \rho_{k-1}
+  \end{eqnarray*}
+  where $E(a_t r_{t-k}) = 0$ follows from the fact that any return
+  $r_{t1}$ must not depend on disturbances $a_{t2}$ that occur later
+  in time. The last equation means $\rho_k$ is a geometric series. Since
+  $\rho_0 = 1$, we have
+  \begin{eqnarray*}
+    \rho_k &=& \phi^k \\
+    |\rho_k| &=& e^{k\ln|\phi|} \\
+  \end{eqnarray*}
+  Note that $|\phi| < 1$ and hence $\ln|\phi| < 0$.
+
+  Although $k$ can only take integer values, it is still useful to
+  define a correlation time $\tau$ such that $\phi^\tau = 1/2$. Such a
+  quantity is more intuitive and constitutes a measure of
+  autocorrelations that is universal and comparable among different
+  time series' models. From the definition of $\tau$ we get
+  \begin{equation}
+    \label{eq:tau_def}
+    \begin{aligned}
+      \phi^\tau &= 1/2 \\
+      \tau &= -{\ln 2 \over \ln\phi} \\
+      \phi &= 2^{-1/\tau}
+    \end{aligned}
+  \end{equation}
+  Figure \ref{fig:AR1-autocorrelation} shows the autocorrelation
+  function of the AR(1) model. As proven above, this function decays
+  exponentially.
+  \begin{figure}[htb!]
+    \vspace{-15mm}
+    \centering
+    \includegraphics[scale=0.5, clip=true, trim=113 229 115
+    139]{../pics/AR1-autocorrelation.pdf}
+    \caption{\footnotesize Autocorrelation function of the AR(1) model
+      with $\phi=1/\sqrt{2}$. Red circles: autocorrelations at $k=0, 1,
+      2, \cdots$. Blue line: $e^{t\ln\phi}$.}
+    \label{fig:AR1-autocorrelation}
+  \end{figure}
+  For more details about conventional time series' models, see
+  \cite{BoxJenkins94}.
+\end{itemize}
+
+For purposes of later reference, we also list some notations that may
+cause confusion to the reader:
+\begin{itemize}
+\item $E(x)$ or $\mean{x}$: The expectation value of $x$.
+\item $\var(x)$: the variance of $x$.
+\item $\text{cov}(x,y)$: the covariance of $x$ and $y$.
+\item $\text{corr}(x,y)$: the correlation between $x$ and $y$, i.e.
+  \begin{equation*}
+    \text{corr}(x,y) = {\text{cov}(x,y) \over \sqrt{\var(x)\var(y)}}
+  \end{equation*}
+\end{itemize}
+
+\section{Stylized facts}\label{sec:StylizedFacts}
+This section presents and explains a few ``stylized facts'',
+i.e. phenomena that are widely observed and accepted as true.
+\begin{itemize}
+\item Fat tails. It has been consistently reported by various studies
+  - for example \cite{Potters2003} and \cite{Mantegna2000} - that the
+  probability density function (PDF) of stock/index returns are not
+  Gaussian. Unlike the Gaussian PDF, which is symmetric and falls off
+  very quickly as its argument moves from the center to the outskirts
+  (tails), the PDF of stock/index returns are higher on the tails,
+  i.e. the probability of large fluctuations is higher than is dictated
+  by a Gaussian distribution - in fact, even higher than dictated by
+  an exponential function. Empirical studies suggest the PDF of the
+  returns follows a power-law on the tails; the exponent of the power
+  is around 4, depending on the specific stock/index.
+
+  Figure \ref{fig:FatTail} illustrates this feature.
+  \begin{figure}[htb!]
+    \centering
+    \includegraphics[scale=0.5, clip=true, trim=92 229 110
+    140]{../pics/FatTail.pdf}
+    \caption{\footnotesize Fat tails. Blue: empirical PDF of 15min returns of
+      {\it Nordea Bank}. The returns are computed using paid
+      prices on the Stockholm OMX market during the period between
+      2013-10-10 and 2014-01-29. Red: Gaussian distribution with equal
+      mean and variance.}
+    \label{fig:FatTail}
+  \end{figure}
+
+\item Non-zero skewness. Apart from fat tails, the PDF of stock/index
+  returns are often also skewed. If the skewness is positive
+  (negative), the probability of very large positive (negative)
+  returns is higher than that of very large negative (positive)
+  returns, even though the mean of the returns is 0 or extremely
+  close to 0.
+
+  Table \ref{tab:EmpiricalSkewness} lists the skewness of a few
+  Swedish stocks traded on the Stockholm OMX market.
+  \begin{table}[htb!]
+    \footnotesize
+    \centering
+    \begin{tabular}{c|c|c|c|c|c|c}
+      Nordea Bank & Volvo B & Boliden & ABB Ltd & H\&M & Scania
+      B & Ericsson B \\
+      \hline
+      0.1362 & 0.0471 & 0.0567 & -0.0364 & -0.0168 &
+      -1.1554 & 0.2086 \\
+    \end{tabular}
+    \caption{\footnotesize Skewness of Stock Returns. All the returns
+      have time-lag of 15 minutes and are computed using paid prices
+      between 2013-10-10 and 2014-01-29. }
+    \label{tab:EmpiricalSkewness}
+  \end{table}
+
+\item Higher-order autocorrelation. 
+\end{itemize}
 
@@
 equations, it is clear that the partial autocorrelation function cuts
-off at $max(p, q)$.
+off at $\max(p, q)$.
 
-\chapter{Case Study of Stocks}
+\chapter{Price Models}
 \section{Nordea Bank}
-\begin{figure}[ht]
+\begin{figure}[htb!]
   \centering
@@
   200]{../pics/nordea_price_20120116-20120420.pdf}
-  \caption{Nordea Bank paid prices 2012/01/16-2012/04/20}
+  \caption{\footnotesize Nordea Bank paid prices 2012/01/16-2012/04/20}
   \label{fig:Nordea}
@@
   }
-  \caption{Nordea Bank 15min autocorrelations}
+  \caption{\footnotesize Nordea Bank 15min autocorrelations}
 \end{figure}
@@
 conditional volatility. See \cite{Andersen03}.}, we look at the
-qq-plot of $\epsilon_t/\sigma_t$ \ref{fig:nordea_15min_epOversig_qq},
-which is consistent with a standard Gaussian distribution except for a
-few outliers at the two ends. Hence the fat tails of the probability
-density function can be entirely accounted for by GARCH effects,
-implying a Gaussian conditional distribution is appropriate for $\epsilon_t$.
-\begin{figure}[htb!]
+qq-plot of $\epsilon_t/\sigma_t$ (figure
+\ref{fig:nordea_15min_epOversig_qq}), which is consistent with a
+standard Gaussian distribution except for a few outliers at the two
+ends. Hence the fat tails of the probability density function can be
+entirely accounted for by GARCH effects, implying a Gaussian
+conditional distribution is appropriate for $\epsilon_t$.
+% \begin{figure}[htb!]
+\begin{wrapfigure}{r}{0.5\textwidth}
   \centering
-    \includegraphics[scale=0.6, clip=true, trim=100 236 118
+    \includegraphics[scale=0.35, clip=true, trim=100 236 118
     200]{../pics/nordea_15min_epOversig_qq.pdf}
-  \caption{QQ-plot of $\epsilon_t / \sigma_t$. $\epsilon_t$ are
+  \caption{\scriptsize{QQ-plot of $\epsilon_t / \sigma_t$. $\epsilon_t$ are
     derived from Nordea Bank 15min returns while $\sigma_t$ are realized
     volatilities calculated using 30s returns within each 15min
-    interval.}
+    interval.}}
   \label{fig:nordea_15min_epOversig_qq}
-\end{figure}
+\end{wrapfigure}
 
@@
   \end{tabular}
-  \caption{GARCH(1, 3) model of Nordea Bank 15min returns}
+  \caption{\footnotesize GARCH(1, 3) model of Nordea Bank 15min returns}
   \label{tab:nordea_15min_garch}
@@
   }
-  \caption{Nordea 15min $r_t/\sigma_t$ and $(r_t/\sigma_t)^2$ ACF.}
+  \caption{\footnotesize Nordea 15min $r_t/\sigma_t$ and $(r_t/\sigma_t)^2$ ACF.}
 \end{figure}
@@
   }
-  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ unconditional
+  \caption{\footnotesize Nordea 15min $\ln\sigma_{t+h, h}$ unconditional
     distribution}
@@
   }
-  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$ autocorrelations}
+  \caption{\footnotesize Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$ autocorrelations}
 \end{figure}
@@
   \end{tabular}
-  \caption{Nordea 15min $w_t$ autocorrelations}
+  \caption{\footnotesize Nordea 15min $w_t$ autocorrelations}
   \label{tab:nordea_15min_w_acf}
@@
   }
-  \caption{Nordea 15min $y_t$ QQ-plot and autocorrelations}
+  \caption{\footnotesize Nordea 15min $y_t$ QQ-plot and autocorrelations}
 \end{figure}
@@
   \end{tabular}
-  \caption{Nordea 15min $y_t$ moments}
+  \caption{\footnotesize Nordea 15min $y_t$ moments}
   \label{tab:nordea_15min_y_moments}
@@
 \begin{figure}[htb!]
+  \vspace{-18mm}
   \centering
-    \includegraphics[scale=0.5, clip=true, trim=112 272 107
+    \includegraphics[scale=0.35, clip=true, trim=112 272 107
     140]{../pics/nordea_15min_y_js_fit.pdf}
-    \caption{Nordea 15min $y_t$ fitted to a Johnson Su distribution}
+    \caption{\footnotesize Nordea 15min $y_t$ fitted to a Johnson Su distribution}
     \label{fig:nordea_15min_y_js_fit}
@@
   \end{tabular}
-  \caption{Nordea 15min estimation results}
+  \caption{\footnotesize Nordea 15min estimation results}
   \label{tab:nordea_15min_js_param}
@@
 
-The probability density function of the residuals $y_t$ that result
-from the MLE estimate is shown in figure
-\ref{fig:nordea_15min_y1_hist} together with its counterpart from the
-initial estimate. 
+% The probability density function of the residuals $y_t$ that result
+% from the MLE estimate is shown in figure
+% \ref{fig:nordea_15min_y1_hist} together with its counterpart from the
+% initial estimate.
 \subsection{45min Returns}
+\section{Fat tail and stochastic volatility}
+In this section we show that fat tails of the unconditional PDF are
+the results of stochasticity in volatility.
 
+Specifically we consider the model
+\begin{equation}  \label{eq:UnconditionalPdf}
+  \begin{aligned}
+    r_t &= e^{\bar{v} + v_t} b_t \\
+    &= e^{\bar{v}} e^{v_t} b_t \\
+    &= e^{\bar{v}} r'_t
+  \end{aligned}
+\end{equation}
+where $b_t \sim N(0, 1)$, $\bar{v}$ is a constant and $v_t \sim N(0,
+\sigma)$ if there is no auto-correlation in $\{v_t\}$. If there is
+auto-correlation, $v_t$ is described by an appropriate time series
+model. For example, in the case of Volvo B 102min returns, $v_t$ is
+represented by a seasonal auto-regressive model with 5 orders of
+auto-regression and 6 orders of seasonal auto-regression with a
+seasonality of 4.
+
+Here we note that any stationary ARMA model can be represented as a
+pure moving average model, which is infinite in extent if
+auto-regressive components are present. Therefore auto-correlations in
+$\{v_t\}$ influence the unconditional distribution of $r_t$ only
+through $\sigma$.
+
+In the following we derive the unconditional PDF of $r'_t$, denoted
+$f_{r'}(x)$. Then the PDF of $r_t$ is $e^{-\bar{v}}f_{r'}(e^{-\bar{v}}x)$.
+\begin{eqnarray*}
+  P(r'_t < x) &=& P(b_t < xe^{-v_t}) \\
+  f_{r'}(x) &=& f_b(xe^{-v_t}) e^{-v_t}
+\end{eqnarray*}
+Averaging over all $v_t$, we get
+\begin{equation}\label{eq:UncondPDFSymmetric}
+  \begin{aligned}
+    f_{r'}(x) =& \int_{-\infty}^{\infty} dv (2\pi\sigma^2)^{-1/2}
+    e^{-v^2/2\sigma^2}(2\pi)^{-1/2} \exp(-x^2e^{-2v}/2) e^{-v} \\
+    =& {1 \over 2\pi\sigma} \int_{-\infty}^{\infty} dv
+    \exp\left(
+      -{1 \over 2\sigma^2} v^2 - v -{1 \over 2} x^2 e^{-2v}
+    \right)
+  \end{aligned}
+  \end{equation}
+The last part of the integrand, $e^{-x^2 e^{-2v} / 2}$, is plotted in
+figure \ref{fig:DoubleExp}.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=85 252 100
+  231]{../pics/DoubleExp.pdf}
+  \caption{\footnotesize Plot of $\exp(-{1 \over 2} x^2 e^{-2v})$}
+  \label{fig:DoubleExp}
+\end{figure}
+Therefore we make the following approximation:
+\[
+\exp\left(-{1 \over 2} x^2 e^{-2v}\right) \approx \left\{
+  \begin{array}{lr}
+    0 & \text{if } v < \ln|x| -{1 \over 2} \ln(2\ln 2) \\
+    1 & \text{otherwise}
+  \end{array}
+\right.
+\]
+Here we note that $\exp\left(-{1 \over 2} x^2 e^{-2v}\right) = 1/2$ at
+$v = \ln|x| -{1 \over 2} \ln(2\ln 2)$.
+
+With this approximation we have
+\begin{eqnarray*}
+  f_{r'}(x) &=& {1\over C}{1 \over 2\pi\sigma} \int_{\ln(|x|/\sqrt{\ln
+      4})}^{\infty} dv
+  \exp\left(-{1 \over 2\sigma^2} v^2 - v\right) \\
+  &=& {1\over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}} \text{erfc} \left(
+    {1 \over \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma
+      \over \sqrt{2}}
+  \right)
+\end{eqnarray*}
+where $1/C$ has been added for the purpose of normalization.
+
+At large $x$, we may use the asymptotic expansion of $\mathrm{erfc}$
+to write
+\begin{eqnarray*}
+  Cf_{r'}(x) &=& {e^{\sigma^2 / 2} \over \sqrt{8\pi}} \mathrm{erfc}(\xi) \\
+  &=& {e^{\sigma^2 / 2} \over \sqrt{8 \pi}}
+  \frac{e^{-\xi^2}}{\xi\sqrt{\pi}}\left[
+    1 +
+    \sum_{n=1}^N (-1)^n \frac{(2n-1)!!}{(2\xi^2)^n} \right] +
+  O(\xi^{-2N-1} e^{-\xi^2})
+\end{eqnarray*}
+where $\xi = {1 \over \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} +
+{\sigma \over \sqrt{2}}$. The slowest-decaying term is
+\[
+f_{r,0}(x) = {e^{\sigma^2 / 2} \over \sqrt{8\pi}}
+\frac{e^{-\xi^2}}{\xi\sqrt{\pi}}
+\]
+Let $\zeta = {|x| \over \sqrt{\ln 4}}$. With a bit manipulation one
+obtains
+\begin{equation}
+  f_{r,0}(x) = {1 \over \pi \sqrt{8}}{1 \over
+    \left(\ln\zeta/\sigma\sqrt{2} +
+      \sigma/\sqrt{2}\right)\zeta\zeta^{\ln \zeta / 2\sigma^2}
+  }
+\end{equation}
+From the last equation one can see that, at any neighborhood of large
+$x$, $f_{r'}(x)$ may be approximated by $C/|x|^\alpha$, i.e. a power law.
+
+In the following we work out the normalization constant C.
+\begin{eqnarray*}
+  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {1 \over C}{e^{\sigma^2 / 2} \over
+    \sqrt{8\pi}} \int_{-\infty}^{\infty} \text{erfc} \left({1 \over
+      \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma \over
+      \sqrt{2}} \right) dx\\
+  &=& {2 \over C}{e^{\sigma^2 / 2} \over
+    \sqrt{8\pi}} \int_{0}^{\infty} \text{erfc} \left({1 \over
+      \sqrt{2}\sigma} \ln{x \over \sqrt{\ln 4}} + {\sigma \over
+      \sqrt{2}} \right) dx\\
+\end{eqnarray*}
+Let
+\begin{eqnarray*}
+  a &=& {1 \over \sigma \sqrt 2} \\
+  b &=& -{1 \over 2 \sigma \sqrt 2}\ln\ln 4 + {\sigma \over \sqrt 2} \\
+  y &=& a\ln|x|+b \\
+\end{eqnarray*}
+Then
+\begin{eqnarray*}
+  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {2 \over C}{e^{\sigma^2 / 2} \over
+    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {e^{(y-b)/a} \over a}
+  \text{erfc}(y) \\
+  &=& {2 \over C}{e^{\sigma^2 / 2} \over
+    \sqrt{8\pi}} \left.e^{(y-b)/a}
+    \text{erfc}(y)\right|_{y=-\infty}^{\infty}
+  + {2 \over C}{e^{\sigma^2 / 2} \over
+    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {2 \over \sqrt \pi}
+  e^{(y-b)/a} e^{-y^2} \\
+  &=& {2 \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}  2 e^{\ln\ln 4/2
+    - \sigma^2/2}\\
+  C &=& \sqrt{2 \ln 4\over \pi} \\
+  &\approx& 0.9394
+\end{eqnarray*}
+Thus we can write in summary:
+\begin{eqnarray*}
+  f_r(x) &=& e^{-\bar{v}} f_{r'}(e^{-\bar{v}} x) \\
+    &=& {e^{-\bar{v}} \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}
+    \text{erfc} \left({1 \over \sqrt{2}\sigma} \ln{|e^{-\bar{v}}x| \over \sqrt{\ln
+          4}} + {\sigma \over \sqrt{2}}
+    \right) \\
+\end{eqnarray*}
+Using the same technique of integration as for normalization, the
+cummulative distribution function of $r_t$ is found to be $F(x)$,
+which is the following:
+\begin{enumerate}
+\item if $x < 0$
+  \begin{eqnarray*}
+    F(x) &=& {\sqrt{\ln 4} \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}
+    \left[
+      {e^{-\bar{v}}x \over \sqrt{\ln 4}}\text{erfc}\left(
+        {1 \over \sigma \sqrt 2} \ln{-e^{-\bar{v}}x \over \sqrt{\ln
+            4}} + {\sigma \over \sqrt 2}
+      \right) \right.\\
+      && \left. + e^{-\sigma^2 / 2} \text{erfc}\left(
+        {1 \over \sigma \sqrt 2} \ln{-e^{-\bar{v}}x \over \sqrt{\ln 4}}
+      \right)
+    \right]
+  \end{eqnarray*}
+\item if $x \geq 0$
+  \begin{eqnarray*}
+    F(x) &=& \frac{1}{2} + {\sqrt{\ln 4} \over C}{e^{\sigma^2 / 2} \over
+      \sqrt{8\pi}} \left[
+      {e^{-\bar{v}}x \over \sqrt{\ln 4}} \text{erfc}\left(
+        {1 \over \sigma \sqrt 2} \ln{e^{-\bar{v}}x \over \sqrt{\ln
+            4}} + {\sigma \over \sqrt 2}
+      \right) \right. \\
+      && \left. + e^{-\sigma^2/2} \text{erfc}\left(
+        -{1 \over \sigma \sqrt 2} \ln{e^{-\bar{v}}x \over \sqrt{\ln 4}}
+      \right)
+    \right]
+  \end{eqnarray*}
+\end{enumerate}
+
+To verify the validity of the model, we fit the above probability
+density function to the de-meaned\footnote{By ``de-meaned returns'' we
+  mean the quantity $r_t - \mean{r_t}$, where $r_t$ are the measured
+  returns and $\mean{r_t}$ is their mean over time.} 30min returns of
+Volvo B by means of maximum likelihood estimate using the MATLAB
+function ``mle''. Then for the parameters $\sigma$ and $\bar{v}$ we
+get
+\begin{eqnarray*}
+  \sigma &=& 0.6355 \\
+  \bar{v} &=& -6.0308
+\end{eqnarray*}
+Then we plot $P(r'_t > x)$ of the model against its empirical
+counterpart on a log-log scale, as shown in figure
+\ref{fig:volvo_30min_ret}.
+\begin{wrapfigure}{r}{0.5\textwidth}
+  \vspace{-20mm}
+%\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.4, clip=true, trim=98 231 116
+    126]{../pics/volvo_30min_ret.pdf}
+  \end{center}
+  \vspace{-5mm}
+  \caption{\footnotesize{$P(r'_t > x)$ for $x > 0$. Blue: empirical
+      probabilities. Red: probabilities predicted by the model.}}
+  \label{fig:volvo_30min_ret}
+\end{wrapfigure}
+
+In general figure \ref{fig:volvo_30min_ret} shows a good fit, but a
+closer look reveals that deviations are significant in the regions $r
+\in (0, \sigma_r)$ and $r \in (2\sigma_r, 3\sigma_r)$, where
+$\sigma_r$ stands for the empirical standard deviation of the
+de-meaned returns. These observations are
+shown in greater details in figure \ref{fig:volvo_30min_ret2}.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[$\ln(P(r > x))$ with $x \in (0, \sigma_r)$]{
+    \includegraphics[scale=0.4, clip=true, trim=93 224 115
+    125]{../pics/volvo_30min_ret_0-sigma.pdf}
+  }
+  \subfigure[$\ln(P(r > x))$ with $x \in (2\sigma_r, 3\sigma_r)$]{
+    \includegraphics[scale=0.4, clip=true, trim=93 224 115
+    125]{../pics/volvo_30min_ret_2sigma-3sigma.pdf}
+  }
+  \caption{\footnotesize Surviving probabilities. Blue: empirical values of $\ln(P(r
+    > x))$. Red: Predicted values of $\ln(P(r > x))$.}
+  \label{fig:volvo_30min_ret2}
+\end{figure}
+
+Moreover, the inefficiency of the model is also manifest in the
+skewness of the data. For the Volvo 30min returns, the data has a
+skewness of 0.2419, but our model is strictly symmetric, since
+$x$ appears in equation \ref{eq:UncondPDFSymmetric} only as $x^2$.
+
+Hence the above model needs to be improved to accommodate the non-zero
+skewness as well as to account for the discrepancies shown in figure
+\ref{fig:volvo_30min_ret2}.
+
+It has long been hypothesized in the liturature that skewness is the
+result of price-volatility correlation (for example
+\cite{Potters2003}). Therefore, the most apparent modification is to
+allow $v_t$ and $b_t$ to be correlated. For convenience we decompose
+$v_t \sim N(0, \sigma)$ as $v_t = \sigma a_t$ and assume
+\begin{eqnarray*}
+  \begin{pmatrix}
+    a_t \\
+    b_t
+  \end{pmatrix} \sim N(0, \Sigma)
+\end{eqnarray*}
+with a covariance matrix
+\begin{eqnarray*}
+    \Sigma &=&
+  \begin{pmatrix}
+    1 & \psi \\
+    \psi & 1
+  \end{pmatrix}
+\end{eqnarray*}
+where $|\psi| < 1$. Then we rewrite equation \ref{eq:UnconditionalPdf}
+as
+\begin{eqnarray*}
+    r_t &=& e^{\bar{v} + \sigma a_t} b_t \\
+    &=& e^{\bar{v}} r'_t \\
+    r'_t &=& e^{\sigma a_t} b_t \\
+\end{eqnarray*}
+Then
+\begin{eqnarray*}
+  && P(r'_t < x)\\
+  &=& P(b_t < e^{-a_t\sigma}x) \\
+  &=& {1 \over 2\pi |\det(\Sigma)|^{1/2}}
+  \int_{-\infty}^{\infty} da \int_{-\infty}^{e^{-a\sigma}x} db
+  \exp\left[
+    -{1 \over 2} (a, b) \Sigma^{-1}
+    \begin{pmatrix}
+      a \\
+      b
+    \end{pmatrix}
+  \right] \\
+\end{eqnarray*}
+After some manipulations we get
+\begin{equation}\label{eq:UncondCDFAsymmetric}
+  \begin{aligned}
+    & P(r'_t < x) \\  
+    &= {1 \over 2\pi \sqrt{1 - \psi^2}} \int_{-\infty}^{\infty} da
+    e^{-a^2/2} \int_{-\infty}^{e^{-a\sigma}x} db
+    \exp\left[
+      - {(b - a\psi)^2 \over 2(1 - \psi^2)}
+    \right] \\
+    &= {1 \over 2\sqrt{2\pi}} \int_{-\infty}^{\infty} e^{-a^2/2}
+    \text{erfc}{a\psi - e^{-a\sigma} x \over \sqrt{2(1-\psi^2)}} da
+  \end{aligned}
+\end{equation}
+% To evaluate the above integral we taylor-expand $\erfc(u)$ around $u =
+% 0$:
+% \begin{eqnarray*}
+%   \erfc(u) &=& 1 - {2 \over \sqrt \pi} \sum_{k=0}^\infty (-1)^k
+%   {x^{2k+1} \over (1 + 2k) k!}
+% \end{eqnarray*}
+% Substitution in equation \ref{eq:UncondCDFAsymmetric} gives
+% \begin{eqnarray*}
+%   && P(r'_t < x) \\
+%   &=& {1 \over 2} - {1 \over \pi\sqrt{2}} \sum_{k=0}^\infty (-1)^k {1 \over
+%     (2k+1)k!} {1 \over \left[2(1 - \psi^2)\right]^{k+1/2}}
+%   \int_{-\infty}^\infty da e^{-a^2/2}(a\psi - e^{-a\sigma}x)^{2k+1} \\
+%   &=& 
+% \end{eqnarray*}
+% Binomial expansion of $(a\psi - e^{-a\sigma}x)^{2k+1}$ yields
+% \begin{eqnarray*}
+%   && P(r'_t < x) \\
+%   &=& {1 \over 2} - {1 \over \pi\sqrt{2}} \sum_{k=0}^\infty
+%   (-1)^k {1 \over (2k+1)k!} {1 \over \left[2(1 -
+%       \psi^2)\right]^{k+1/2}} \sum_{i=0}^{2k+1} {2k+1 \choose i}
+%   \psi^i x^{2k+1-i} \times \\
+%   && \exp{\sigma^2 (2k+1-i)^2 \over 2} \int_{-\infty}^\infty da a^i
+%   \exp\left\{
+%     -{1 \over 2} [a + \sigma(2k+1-i)]^2
+%     \right\} \\
+%   \end{eqnarray*}
+%   The trailing integral can be also evaluated by binomial expansion:
+%   \begin{eqnarray*}
+%     && \int_{-\infty}^\infty da a^i
+%     \exp\left\{
+%       -{1 \over 2} [a + \sigma(2k+1-i)]^2
+%     \right\}\\
+%     &=& \int_{-\infty}^\infty \sqrt{2} da \left[
+%     \sqrt{2} a - \sigma(2k+1-i)
+%   \right]^i e^{-a^2} \\
+%   &=& \sum_{j=0}^i {i \choose j} 2^{j/2} \sigma^{i-j} (2k+1-i)^{i-j}
+%   \sqrt{2} \int_{-\infty}^\infty a^j e^{-a^2} da \\
+%   &=& \sum_{j=0}^i {i \choose j} 2^{j/2} \sigma^{i-j} (2k+1-i)^{i-j}
+%   \sqrt{2} \Gamma\left(
+%     {j+1 \over 2}
+%   \right)
+% \end{eqnarray*}
+% Finally we get
+% \begin{eqnarray*}
+%     && P(r'_t < x) \\
+%   &=& {1 \over 2} - {1 \over \pi\sqrt{2}} \sum_{k=0}^\infty
+%   (-1)^k {1 \over (2k+1)k!} {1 \over \left[2(1 -
+%       \psi^2)\right]^{k+1/2}} \sum_{i=0}^{2k+1} {2k+1 \choose i}
+%   \psi^i x^{2k+1-i} \times \\
+%   && \exp{\sigma^2 (2k+1-i)^2 \over 2} \sum_{j=0}^i {i \choose j}
+%   2^{j/2} \sigma^{i-j} (2k+1-i)^{i-j} \sqrt{2} \Gamma\left(
+%     {j+1 \over 2}
+%   \right)
+% \end{eqnarray*}
+% The above expression is not in a closed form but nonetheless suffices
+% for numerical evaluation. For that purpose we re-arrange the i index
+% and rewrite the above expression as
+% \begin{eqnarray*}
+%     && P(r'_t < x) \\
+%   &=& {1 \over 2} - {1 \over \pi\sqrt{2}} \sum_{k=0}^\infty
+%   (-1)^k {1 \over (2k+1)k!} {1 \over \left[2(1 -
+%       \psi^2)\right]^{k+1/2}} \sum_{i=0}^{2k+1} {2k+1 \choose i}
+%   \psi^{2k+1-i} x^i \times \\
+%   && \exp{\sigma^2 i^2 \over 2} \sum_{j=0}^{2k+1-i} {2k+1-i \choose j}
+%   2^{j/2} (\sigma i)^{2k+1-i-j} \sqrt{2} \Gamma\left(
+%     {j+1 \over 2}
+%   \right)
+% \end{eqnarray*}
+% To compare the model with empirical return distributions we need $P(r_t
+% < x)$:
+% \begin{eqnarray*}
+%   P(r_t < x) &=& P(e^{\bar{v}} r'_t < x) \\
+%   &=& P(r'_t < e^{-\bar{v}} x) \\
+% \end{eqnarray*}
+To evaluate this integral we make an approximation of $\erfc(x)$:
+\begin{equation*}
+  \erfc(u) \approx \left\{
+  \begin{array}{ll}
+    2 & \text{if } u \leq u_1 \\
+    Au^3 + Bu + 1 & \text{if } u_1 < u \leq u_2 \\
+    0 & \text{if } u > u_2
+  \end{array}
+  \right.
+\end{equation*}
+where $u_1$ and $u_2$ are constants to be determined by OLE methods.
+
+The 3 intervals of $u$ translate to 3 corresponding intervals of $a$:
+\begin{enumerate}[I]
+\item
+  \begin{equation*}
+    (-\infty, u_1) \Leftrightarrow \left\{
+    \begin{array}{ll}
+      (-\infty, a_1) & \text{ if } \psi > 0 \\
+      (a_1, \infty) & \text{ if } \psi < 0 \\
+    \end{array}
+    \right.
+  \end{equation*}
+  where $a_1$ is defined as the root of the
+  transcendental equation $a_1\psi - e^{-a_1\sigma}x = u_1 \sqrt{2(1-\psi^2)}$.
+\item
+  \begin{equation}\label{eq:RegionI}
+    (u_1, u_2) \Leftrightarrow \left\{
+    \begin{array}{ll}
+      (a_1, a_2) & \text{ if } \psi > 0 \\
+      (a_2, a_1) & \text{ if } \psi < 0 \\
+    \end{array}
+    \right.
+  \end{equation}
+  where $a_2$ is defined as the root of the
+  transcendental equation $a_2\psi - e^{-a_2\sigma}x = u_1 \sqrt{2(1-\psi^2)}$.
+\item
+  \begin{equation}\label{eq:RegionII}
+    (u_2, \infty) \Leftrightarrow \left\{
+    \begin{array}{ll}
+      (a_2, \infty) & \text{ if } \psi > 0 \\
+      (-\infty, a_2) & \text{ if } \psi < 0 \\
+    \end{array}
+    \right.
+  \end{equation}
+\end{enumerate}
+We denote the parts of the integral \ref{eq:UncondCDFAsymmetric} in
+these intervals with I, II and III correspondingly. Then we can write
+\begin{equation}\label{eq:DeducedCDF}
+    P(r'_t < x) = {1 \over 2\sqrt{2\pi}} \left(
+      \text{I + II}
+    \right)
+\end{equation}
+In the following we first consider the case $\psi > 0$. In this case
+$a_1 < a_2$
+\begin{eqnarray*}
+  \text{I} &=& 2 \int_{-\infty}^{a_1} e^{-a^2/2} da \\
+  &=& \sqrt{2\pi} \erfc(-a_1/\sqrt 2)
+\end{eqnarray*}
+\begin{eqnarray*}
+  \text{II} &=& \int_{a_1}^{a_2} da e^{-a^2/2} \left[
+    A {(a\psi - e^{-a\sigma} x)^3\over 2^{3/2} (1 - \psi^2)^{3/2}} +
+    B{a\psi - e^{-a\sigma}x \over \sqrt{2(1-\psi^2)}} + 1
+  \right] \\
+  &=& \int_{a_1}^{a_2} da e^{-a^2/2} \left[
+    A{-e^{-3a\sigma}x^3 - 3a^2\psi^2 x e^{-a\sigma}
+      +3 a \psi x^2 e^{-2a\sigma} + a^3\psi^3
+      \over
+      2^{3/2} (1 - \psi^2)^{3/2}
+    } + \right.\\
+    && \left. B{a\psi - e^{-a\sigma}x
+    \over
+    \sqrt{2(1 - \psi^2)}
+  } + 1
+\right] \\
+&=& \text{II1} + \text{II2} + \text{II3} + \text{II4} + \text{II5} + \text{II6} + \text{II7}
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+  \text{II1} &=& -{Ax^3 \over 2^{3/2} (1 -
+    \psi^2)^{3/2}}\int_{a_1}^{a_2} da e^{-a^2/2} e^{-3a\sigma}x^3 \\
+  &=& -{Ax^3e^{9\sigma^2/2}\sqrt{2\pi} \over 2^{5/2} (1 -
+    \psi^2)^{3/2}} \left[
+    \erfc\left(a_1 + 3\sigma \over \sqrt 2\right)
+    - \erfc\left(a_2 + 3\sigma \over \sqrt 2\right)
+  \right]
+\end{eqnarray*}
+\begin{eqnarray*}
+  \text{II2} &=& -{3A\psi^2 x \over 2^{3/2} (1 - \psi^2)^{3/2}}
+  \int_{a_1}^{a_2} e^{-a^2/2 - a\sigma} a^2 \\
+  &=& -{3A\psi^2 x \over 2^{3/2} (1 - \psi^2)^{3/2}} e^{\sigma^2/2}
+  \sqrt 2 \int_{(a_1+\sigma)/\sqrt 2}^{(a_2+\sigma)/\sqrt 2} da \left[
+    2a^2 - 2\sqrt{2}\sigma a + \sigma^2
+  \right] e^{-a^2} \\
+  &=& -{3A\psi^2 x \over 2^{3/2} (1 - \psi^2)^{3/2}} e^{\sigma^2/2}
+  \sqrt 2 \left[
+    2\times\text{II2.1} -2\sqrt{2}\sigma\times\text{II2.2} + \sigma^2\text{II2.3}
+  \right]
+\end{eqnarray*}
+\begin{itemize}
+\item II2.1
+  \begin{itemize}
+  \item if $(a_1 + \sigma)(a_2 + \sigma) < 0$
+    \begin{equation*}
+      \text{II2.1} =  {1 \over 2} \left\{
+        \gamma\left[ 3/2, {(a_2 + \sigma)^2 \over 2} \right] +
+        \gamma\left[ 3/2, {(a_1 + \sigma)^2 \over 2} \right]
+      \right\}
+    \end{equation*}
+  \item if $(a_1 + \sigma)(a_2 + \sigma) > 0$
+    \begin{equation*}
+      \text{II2.1} =  {1 \over 2} \left|
+        \gamma\left[3/2, {(a_2 + \sigma)^2 \over 2}\right] -
+        \gamma\left[3/2, {(a_1 + \sigma)^2\over 2}\right]
+      \right|
+    \end{equation*}
+  \end{itemize}
+  where $\gamma(s, x)$ denotes the lower incomplete gamma function:
+  \begin{equation*}
+    \gamma(s, x) = \int_0^x x'^{s-1} e^{-x'} dx'
+  \end{equation*}
+\item II2.2
+  \begin{equation*}
+    \text{II2.2} = {1 \over 2} \left|
+      e^{-(a_1 + \sigma)^2/2} - e^{-(a_2 + \sigma)^2/2}
+    \right|
+  \end{equation*}
+\item II2.3
+  \begin{equation*}
+    \text{II2.3} = {\sqrt{\pi} \over 2} \left(
+      \erfc {a_1 + \sigma \over \sqrt 2} - \erfc {a_2 + \sigma \over \sqrt 2}
+    \right)
+  \end{equation*}
+\end{itemize}
+\begin{eqnarray*}
+  \text{II3} &=& {3Ax^2\psi \over 2^{3/2}(1 - \psi^2)^{3/2}}
+  \int_{a_1}^{a_2} da e^{-a^2/2 - 2a\sigma} a \\
+  &=& {3Ax^2\psi \over 2^{3/2}(1 - \psi^2)^{3/2}} e^{2\sigma^2}
+  \int_{(a_1 + 2\sigma)/\sqrt 2}^{(a_2 + 2\sigma)/\sqrt 2} e^{-a^2}
+  (\sqrt{2} a - 2\sigma) \sqrt{2} da \\
+  &=& {3Ax^2\psi \over 2^{3/2}(1 - \psi^2)^{3/2}} e^{2\sigma^2} \left[
+    e^{-(a_1 + 2\sigma)^2/2} - e^{-(a_2 + 2\sigma)^2/2} - \right.\\
+    &&
+    \left. \sqrt{2\pi}\sigma \left(
+      \erfc{a_1 + 2\sigma \over \sqrt 2} - \erfc{a_2 + 2\sigma \over \sqrt 2}
+    \right)
+  \right]
+\end{eqnarray*}
+\begin{eqnarray*}
+  \text{II4} &=& {A\psi^3 \over 2^{3/2}(1 - \psi^2)^{3/2}} \left[
+    \gamma\left(2, {a_2^2 \over 2}\right) -
+    \gamma\left(2, {a_1^2 \over 2}\right)
+  \right]
+\end{eqnarray*}
+\begin{eqnarray*}
+  \text{II5} &=& {B\psi \over \sqrt{2(1 - \psi^2)}} \left(
+    e^{-a_1^2/2} - e^{-a_2^2/2}
+  \right)
+\end{eqnarray*}
+\begin{eqnarray*}
+  \text{II6} = -{Bx e^{\sigma^2/2}\over \sqrt{1 - \psi^2}}{\sqrt \pi
+    \over 2} \left[
+    \erfc{a_1 + \sigma \over \sqrt{2}} -
+    \erfc{a_2 + \sigma \over \sqrt{2}}
+  \right]
+\end{eqnarray*}
+\begin{eqnarray*}
+  \text{II7} &=& \sqrt{\pi \over 2} \left[
+    \erfc{a_1^2 \over 2} - \erfc{a_2^2 \over 2}
+  \right]
+\end{eqnarray*}
+
+So far our results are only valid in the case $\psi > 0$, i.e. $a_t$
+and $b_t$ are positively correlated. If, however, $\psi < 0$, then
+$a_1$ and $a_2$ as defined in \ref{eq:RegionI} and \ref{eq:RegionII}
+will have inverse orders, $a_2 < a_1$. To make the above results valid
+even in this case, we only need to exchange $a_1$ and $a_2$ in the
+formulas.
+
+To obtain $P(r_t < x)$, we note
+\begin{eqnarray*}
+P(r_t < x) &=& P(e^{\bar{v}} r'_t < x) \\
+&=& P(r'_t < xe^{-\bar{v}}) \\
+\end{eqnarray*}
+That is, by substituting $xe^{-\bar{v}}$ for $x$ in the formula of
+$P(r'_t < x)$ we get $P(r_t < x)$.
+
+\chapter{Cross-Correlation Matrix with Gaussian Returns}
+In this chapter we present some analytical and numerical results about
+the cross-correlation matrix (cf. \S
+\ref{sec:FundamentalConcepts}), in the special case where the involved
+asset returns are assumed to be Gaussian distributed. This assumption
+is of course an over-simplification, but nevertheless lends some
+insight into the problem.
+
+Our primary interest is in the influence of autocorrelations on the
+cross-correlation matrix. \S \ref{sec:GCC-analytical} discusses the
+distribution of the matrix entries and \S \ref{sec:GCC-numerical} the
+distribution of the eigenvalues.
+
+\section{Distribution of the Matrix Entries}\label{sec:GCC-analytical}
+If the returns follow a zero-mean Gaussian distribution, 
+i.e. $\vec{r}(t) \sim N(0, \Sigma)$, and are not
+auto-correlated, the L\'evy index $\alpha$ as appears in equation
+\ref{eq:cross-correlation} is 2, and $RR'$ is a Wishart matrix with
+the following distribution law \cite{Anderson2003}:
+\begin{equation}
+  \label{eq:WishartPDF}
+  f_{RR'}(X) = { |\det X|^{(T-N-1)/2} \exp\left(-{1 \over 2}\tr
+      \Sigma^{-1}X \right)
+    \over
+    2^{NT/2}\pi^{N(N-1)/4}|\det \Sigma|^{T/2}
+    \prod_{i=1}^N \Gamma\left[(n+1-i)/2\right]
+  }
+\end{equation}
+where $\Sigma$ is the true covariance matrix of the returns.
+
+The Wishart matrix $RR'$ is the main subject of study in the
+literature (see for example \cite{Anderson2003}, \cite{Chiani2012}),
+and we shall make no exception. Since the leading constant $1/T$ is
+the only difference, the results about $RR'$ are easily transformed
+to those about $RR'/T$.
+
+When autocorrelations do exist among the columns of R, the
+distribution of C is no longer Wishart but is nonethelss closely
+related to it. In the following we consider the situation where
+$r_{i,t}$ can be represented as a vector auto-regressive
+process. Specifically, suppose
+\begin{eqnarray*}
+  \vec{r}_t &=& \sum_{k=1}^p \phi_k \vec{r}_{t-k} + \vec{a}_t \\
+  \vec{a}_t &=& \vec{r}_t - \sum_{k=1}^p \phi_k \vec{r}_{t-k}
+\end{eqnarray*}
+where $\vec{a}_t = (a_{1,t}, a_{2,t}, \cdots, a_{N,t})' \sim N(0,
+\Sigma)$. Here $N(0, \Sigma)$ denotes the multivarate normal
+distribution with covariance matrix $\Sigma$. The last equation can be
+written in matrix form
+\[
+A = R M
+\]
+For example, in the case of an AR(1) process
+\begin{equation*}
+  \begin{pmatrix}
+    a_{1,1} & a_{1,2} & \cdots & a_{1,T} \\
+    \vdots & \ddots & \vdots \\
+    a_{N,1} & a_{N,2} & \cdots & a_{N,T} \\
+  \end{pmatrix} =
+  \begin{pmatrix}
+    r_{1,1} & r_{1,2} & \cdots & r_{1,T} \\
+    \vdots & \ddots & \vdots \\
+    r_{N,1} & r_{N,2} & \cdots & r_{N,T} \\
+  \end{pmatrix}
+  \begin{pmatrix}
+    1 & -\phi_1 &   &   & \\
+      & 1 & -\phi_1 &   & \\
+      &   & \ddots  & \ddots &   \\
+      &   &   & 1 & -\phi_1 \\
+      &   &   &   & 1 \\
+  \end{pmatrix}
+\end{equation*}
+Let $QR = RM$, then $Q = RMR^{-1}$. Since the set of $\{r_{ij}\}$ for
+which $\text{det } R = 0$ has probability zero, a matrix $Q$
+satisfying the above equation almost surely exists. Thus $A = RM =
+QR$ and $AA' = QRR'Q'$ Since the columns of A are not auto-correlated,
+$AA' \sim W(\Sigma, T)$. Then $RR' \sim W(Q^{-1} \Sigma Q'^{-1}, T)$
+follows from \cite{Anderson2003}, sec.7.3.3.
+
+Now we observe that $\Sigma$ enters $f(RR')$ through $\tr (\Sigma^{-1}
+RR')$ and $\det \Sigma$. $RR'$ enters through $\tr (\Sigma^{-1}
+RR')$ and $\det (RR')$. Clearly
+\begin{eqnarray*}
+  \det (Q^{-1}\Sigma Q'^{-1}) &=& \det \Sigma \over (\det Q)^2 \\
+  &=& \det \Sigma \over (\det M)^2 \\
+  &=& \det \Sigma
+\end{eqnarray*}
+As to $\tr (\Sigma^{-1} C)$, we have
+\begin{eqnarray*}
+  && \tr\left[(Q^{-1}\Sigma Q'^{-1})^{-1} RR'\right] \\
+  &=& \tr\left[Q'\Sigma^{-1} Q RR'\right] \\
+  &=& \tr\left[ R'^{-1}M'R'\Sigma^{-1} RMR^{-1} RR'\right] \\
+  &=& \tr\left[ \Sigma^{-1} RM (RM)'\right] \\
+  &=& \tr\left[ \Sigma^{-1} AA'\right] \\
+\end{eqnarray*}
+
+Moreover, $\det RR' = \det AM^{-1} M'^{-1} A' = \det AA'$.
+
+Thus if we use $f_R(\cdot)$ to denote the joint probability density of
+the entries of RR' and $f_A(\cdot)$ to denote that of AA', we can
+write
+\begin{equation}\label{eq:cross-corr-matrix-PDF}
+  f_R(RR') = f_A(RMM'R')
+\end{equation}
+
+Substituting for $f_A$ the Wishart probability density function (PDF)
+\ref{eq:WishartPDF}, we get the PDF of $RR'$. The result is not
+particularly simple. However, a relatively simple approximate
+expression for the off-diagonal entries of $RR'/T$ may be
+obtained. In the following we assume that $\text{var}(a_{i,t}) =
+\sigma^2$ for any $i$ and $t$, and as always, that $a_{i,t}$ are not
+autocorrelated, i.e. $\text{corr}(a_{i,t}, a_{i, t'}) = 0$ for any
+$i$, $t$ and $t'$. We have
+\begin{eqnarray*}
+  r_{i,t} r_{j, t} &=& \left[
+    \phi_1 r_{i, t-1} + a_{i, t}
+  \right] \left[
+    \phi_1 r_{j, t-1} + a_{j, t}
+  \right] \\
+  \mean{r_{i,t} r_{j, t}} &=& {1 \over T}\sum_{t=1}^T r_{i,t} r_{j, t} \\
+  &=& \phi_1^2 \mean{r_{i,t-1} r_{j, t-1}} + \phi_1\mean{\left(
+    r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t}
+  \right)} + \mean{ a_{i,t} a_{j,t}}
+\end{eqnarray*}
+We note that the two sums $\sum_{t=1}^T r_{i,t} r_{j, t}$ and
+$\sum_{t=1}^T r_{i,t-1} r_{j, t-1}$ only differ by the first and the
+last addend, which is negligible for sufficiently large T. Thus we
+have
+\begin{eqnarray*}
+  (1 - \phi_1^2)\mean{r_{i,t} r_{j, t}} &\approx& \phi_1\mean{\left(
+    r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t}
+  \right)} + \mean{ a_{i,t} a_{j,t}}
+\end{eqnarray*}
+Now we use the fact that $i$ and $j$ are symmetric on the left-hand
+side, and hence must be so as well on the right-hand side. So we write
+\begin{equation}\label{eq:Offdiag1}
+  (1 - \phi_1^2)\mean{r_{i,t} r_{j, t}} = 2\phi_1\mean{r_{i, t-1}
+    a_{j, t}} + \mean{ a_{i,t} a_{j,t}}
+\end{equation}
+Now we write the AR(1) process $r_{i,t}$ as an infinite moving-average
+process:
+\begin{eqnarray*}
+  r_{i, t} &=& \phi_1 r_{i, t-1} + a_{i,t} \\
+  (1 - \phi_1 B) r_{i, t} &=& a_{i,t} \\
+\end{eqnarray*}
+where $B$ is the back-shift operator. Then it follows from the above
+equation
+\begin{eqnarray*}
+  r_{i,t} &=& {1 \over 1 - \phi_1 B} a_{i,t} \\
+  &=& \sum_{k=0}^\infty \phi_1^k B^k a_{i,t} \\
+  &=& \sum_{k=0}^\infty \phi_1^k a_{i,t-k} \\
+\end{eqnarray*}
+where it is remembered that $a_{i, t}$ with $t \leq 0$ is zero, (in
+words, this implies that the $r_{i,t}$ process is not affected at all
+by events before $t = 1$.)
+
+% we have
+% \begin{eqnarray*}
+%   r_{i,t} &=& \sum_{k=0}^{t-1} \phi_1^k a_{i,t-k}
+% \end{eqnarray*}
+Substituting this into eq.\ref{eq:Offdiag1} for $r_{i,t-1}$ yields
+\begin{equation}
+  \label{eq:Offdiag2}
+  \begin{aligned}
+  (1 - \phi_1^2)\mean{r_{i,t} r_{j, t}} &\approx
+  2 \sum_{k=0}^\infty\phi_1^{k+1} \mean{a_{i, t-k-1} a_{j, t}}
+  + \mean{a_{i,t} a_{j,t}} \\
+  &=
+  2 \sum_{k=1}^\infty\phi_1^k \mean{a_{i, t-k} a_{j, t}}
+  + \mean{a_{i,t} a_{j,t}} \\
+  &=
+  {2 \over T} \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi_1^k a_{i, t-k} a_{j, t}
+  + {1 \over T} \sum_{t=1}^{T} a_{i,t} a_{j,t}\\
+  \end{aligned}
+\end{equation}
+
+% To proceed any further, one has to assume a particular correlation
+% structure among the a's. A simple yet non-trivial assumption is the
+% following:
+% \begin{equation*}
+%   \text{corr}(a_{i, t-k}, a_{j, t}) = \left\{
+%     \begin{array}{l l}
+%       1 & \text{ if $i=j$ and $k = 0$ }\\
+%       \rho & \text{ if $i \neq j$ and $k = 0$ }\\
+%       0 & \text{otherwise}
+%     \end{array}
+%   \right.
+% \end{equation*}
+
+% To obtain the PDF of $\mean{r_{i,t} r_{j, t}}$ from the above
+% equation, we first find the characteristic function of
+% $a_{i,t}a_{j,t'}$ for arbitrary $t, t' \geq 1$. Then the
+% characteristic function of the above weighted sum of ${T+1 \choose 2}$
+% such random variables can be found as a product of the characteristic
+% functions of the addends. In the end we compute the PDF of $(1 -
+% \phi_1^2)\mean{r_{i,t} r_{j, t}}$ by an inverse Fourier transform.
+
+Given two Gaussian random variables $x$ and $y$ with zero mean and
+covariance matrix 
+\begin{equation*}
+  \Sigma =
+  \begin{pmatrix}
+    1 & \rho \\
+    \rho & 1 \\
+  \end{pmatrix}
+\end{equation*}
+
+The PDF of $xy$ can be found by considering $P(xy < z)$:
+\begin{eqnarray*}
+  P(xy < z) &=& \left(\int_0^\infty dx \int_0^{z/x} dy
+    +\int_{-\infty}^0 dx \int_{z/x}^\infty dy \right)
+    {1 \over 2\pi \sqrt{1 - \rho^2}} \\
+    &&
+    \exp\left[
+      -{x^2 -2\rho xy + y^2
+        \over
+        2\sigma^2 (1 - \rho^2)} 
+    \right] \\
+  f_\text{{INN}}(z; \sigma, \rho) &=& {d \over dz} P(xy < z)\\
+  &=& {1 \over \pi \sqrt{1 - \rho^2}} \exp\left[
+    {\rho z \over \sigma^2 (1 - \rho^2)}\right] K_0\left[
+    {|z| \over \sigma^2 (1 - \rho^2)}
+  \right]
+\end{eqnarray*}
+where $K_n(z')$ is the modified Bessel function of the second
+kind. It is worth taking note that, when $\rho \neq 0$,
+$f_\text{{INN}}(z; \sigma, \rho)$ is not symmetric with respect to
+$z$. As a result, if $\rho > 0$, the mean of $f_\text{{INN}}(z;
+\sigma, \rho)$ is positive, and vice versa.
+
+Secondly, because $|\rho| < 1$ and
+\begin{equation*}
+  K_0(z) \sim \sqrt{\pi \over 2z} e^{-z}
+\end{equation*}
+as $z \to \infty$ \cite{Olver:2010:NHMF}, all the moments of
+$f_\text{{INN}}(z; \sigma, \rho)$ are finite, implying the
+applicability of the Lyapunov central limit theorem provided that $T$
+is large, which is very often the case and what we assume here.
+
+With this mind, we observe that $\phi_1$ only affects the first sum in
+\ref{eq:Offdiag2}. If $a_{i,t-k}$ and $a_{j,t}$ are not correlated for
+non-zero $k$, $\phi_1$ will not affect the mean of $\mean{r_{i,t}
+  r_{j, t}}$, the estimated covariance between asset $i$ and
+$j$. Furthermore, it is also clear from equation \ref{eq:Offdiag2}
+that the variance of $\mean{r_{i,t}r_{j, t}}$ is always increased by a
+non-zero $\phi_1$, regardless of the sign of $\phi_1$. We compute this
+increment in the following.
+
+In light of the above expression for $f_\text{{INN}}(z; \sigma,
+\rho)$, we rewrite equation \ref{eq:Offdiag2} as
+\begin{eqnarray*}
+  (1 - \phi_1^2)\mean{r_{i,t} r_{j, t}} &\approx&
+  {2 \over T}
+  \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi_1^k a_{i, t-k} a_{j, t} \\
+  && + {1 \over T} \sum_{t=1}^{T} \sigma^2 (1 - \rho^2) {a_{i,t} \over
+    \sigma \sqrt{1 - \rho^2}} {a_{j,t} \over \sigma \sqrt{1 - \rho^2}} \\
+\end{eqnarray*}
+In addition, we assume
+\begin{equation*}
+  \text{corr}(a_{i, t-k}, a_{j, t}) = \left\{
+    \begin{array}{l l}
+      1 & \text{ if $i=j$ and $k = 0$ }\\
+      \rho & -1 < \rho < 1. \text{ if $i \neq j$ and $k = 0$ }\\
+      0 & \text{otherwise}
+    \end{array}
+  \right.
+\end{equation*}
+Then, because $a_{i, t-k}$ and $ a_{j, t}$ with $i \neq j$ and $k
+> 0$ are not correlated, the mean of $a_{i, t-k} a_{j, t}$ is 0
+($f_\text{{INN}}(z; \sigma, 0)$ is symmetric), and the variance of it
+can be found to be $\sigma^6$ using formula (10.43.19) of \cite{NIST:DLMF}:
+\begin{equation*}
+  \int_0^\infty dt K_\nu(t) t^{\mu-1} = 2^{\mu-2}
+  \Gamma\left(
+    {\mu + \nu \over 2}
+  \right) \Gamma\left(
+    {\mu - \nu \over 2}
+  \right)
+\end{equation*}
+On the other hand, $a_{i,t}/\sigma \sqrt{1 - \rho^2}$ and
+$a_{j,t}/\sigma \sqrt{1 - \rho^2}$ have variance $1/(1 - \rho^2)$
+and are correlated - $\text{corr}(a_{i,t}, a_{j,t}) = \rho$. The mean
+of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ can be found using formula
+(10.43.22) of \cite{NIST:DLMF}, given that $-1 < \rho < 1$:
+\begin{eqnarray*}
+  \int_0^\infty t^{\mu - 1} e^{-at} K_\nu(t) &=& (\pi/2)^{1/2}
+  \Gamma(\mu + \nu) \Gamma(\mu - \nu)(1 - a^2)^{-\mu/2 + 1/4} \times\\
+  && P^{-\mu+1/2}_{\nu-1/2} (a)
+\end{eqnarray*}
+where $P^\mu_\nu(\cdot)$ is Ferrers function of the first kind. The
+result is
+\begin{eqnarray*}
+  E\left[{a_{i,t}a_{j,t} \over \sigma^2 (1 - \rho^2)}\right]
+  &=&
+  {1 \over \sqrt{2\pi} (1 - \rho^2)^{5/4}} \left[
+    P^{-3/2}_{-1/2}(-\rho) - P^{-3/2}_{-1/2}(\rho)
+  \right]
+\end{eqnarray*}
+Similarly, the variance of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ is
+found to be
+\begin{eqnarray*}
+  v^2(\rho) &=&
+  {4 \over \sqrt{2\pi} (1 - \rho^2)^{7/4}} \left[
+    P^{-5/2}_{-1/2}(\rho) + P^{-5/2}_{-1/2}(-\rho)
+  \right] \\
+  && - E^2\left[{a_{i,t}a_{j,t} \over
+      \sigma^2 (1 - \rho^2)}\right]
+\end{eqnarray*}
+Now we can write down the asymptotic Gaussian distribution of
+$\mean{r_{i,t}r_{j,t}}$, the empirical estimate
+$\text{corr}(r_{i,t},r_{j,t})$:
+\begin{equation*}
+\mean{r_{i,t}r_{j,t}} \sim N(\mu'_X, \sigma'_X)
+\end{equation*}
+where
+\begin{eqnarray*}
+  \mu'_X &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi_1^2)(1 -
+    \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
+    P^{-3/2}_{-1/2}(\rho)
+  \right] \\
+  \sigma'^2_X &=& {1 \over (1 - \phi_1^2)^2}\left[
+    \sum_{t=1}^T \sum_{k=1}^{t-1} \left(
+      2\phi_1^k \over T
+    \right)^2 \sigma^6 + \sum_{t=1}^T
+    {\sigma^4 (1 - \rho^2)^2 \over T^2} v^2(\rho)
+  \right] \\
+  &=& {4 \sigma^6 \over T (1 - \phi_1^2)^2} \left[
+    {\phi_1^2 \over 1 - \phi_1^2} -
+    {\phi_1^2 (1 - \phi_1^{2T}) \over
+      T(1 - \phi_1^2)}
+  \right] + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
+    T (1 - \phi_1^2)^2} \\
+  &\approx& {4 \sigma^6 \phi_1^2 \over T (1 - \phi_1^2)^3}
+  + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
+    T (1 - \phi_1^2)^2}
+\end{eqnarray*}
+Considering that, for any $i$ and $t$,
+\begin{equation*}
+  r_{i,t} = \sum_{k=0}^{t-1} \phi_1^k a_{i, t-k}
+\end{equation*}
+we have
+\begin{equation*}
+  \text{var}(r_{i,t}) \approx {\sigma^2 \over 1 - \phi^2}
+\end{equation*}
+Then it follows that
+\begin{eqnarray*}
+  \text{corr}(r_{i,t}, r_{j,t}) &=&
+  {\mean{r_{i,t}, r_{j,t}} \over
+    \sqrt{\text{var}(r_{i,t})\text{var}(r_{j,t})}} \\
+  &\sim& N(\mu_X, \sigma_X)
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+  \mu_X &=& {1 \over \sqrt{2\pi} (1 -
+    \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
+    P^{-3/2}_{-1/2}(\rho)
+  \right] \\
+  \sigma'^2_X   &\approx& {4 \sigma^2 \phi_1^2 \over T (1 - \phi_1^2)}
+  + {(1 - \rho^2)^2 v^2(\rho) \over T} \\
+  &=& {4 \sigma^2 \over T (2^{2/\tau} - 1)}
+  + {(1 - \rho^2)^2 v^2(\rho) \over T} \\ 
+\end{eqnarray*}
+In the last line, $\phi$ is replaced by the correlation time
+$\tau$, whose definition is in equation \ref{eq:tau_def}.
+
+What we have obtained is a quantitative description of what has been
+said before: The mean of the off-diagonal entries of the
+cross-correlation matrix is not affected by autocorrelations in the
+returns, given that innovations $a_{i,t}$ of different assets and
+different times are un-correlated; the variance, however, is always
+increased by autocorrelations no matter the autocorrelation is
+positive or nagative. The increment is in the form of an additive term
+that scales as $1 / (2^{2/\tau} - 1)$.
+
+% However, to write $f_A(RMM'R')$ as a function of the eigen
+% values of $RMM'R'$ is very involved and no theoretical results are
+% known.
+
+% In the more general case
+% \begin{eqnarray*}
+%   a_{i,t} &=& \left(
+%     1 - \sum_{k=1}^p \phi_{i,k} B^k
+%   \right) r_{i,t} \\
+%   &=& \Phi_i(B)} r_{i,t
+% \end{eqnarray*}
+% where $B$ denotes the back shift operator acting on subscript $t$, we
+% can write
+% \begin{eqnarray*}
+%   A &=&
+%   \begin{pmatrix}
+%     \Phi_1(B) & &\\
+%     & \ddots & \\
+%     && \Phi_N(B)
+%   \end{pmatrix} R \\
+%   &=& \Phi(B) R
+% \end{eqnarray*}
+
+% Following the same reasoning, $RR' \sim W(\Phi^{-1}(B)
+% \Sigma \Phi'^{-1}(B), T)$. Substituting $\Phi^{-1}(B)
+% \Sigma \Phi'^{-1}(B)$ for $\Sigma$ in the Wishart
+% probability density function, we get
+% \begin{eqnarray*}
+%   && \tr\left\{\left[\Phi^{-1}(B)\Sigma \Phi'^{-1}(B)\right]^{-1}
+%     RR'\right\}\\
+%   &=& \tr \left\{\Phi'(B)\Sigma^{-1} \Phi(B) RR'\right\}\\
+%   &=& \tr\left\{\Sigma^{-1}AA'\right\}\\
+% \end{eqnarray*}
+% As to $\det \Phi^{-1}(B)\Sigma \Phi'^{-1}(B)$, we note that $\Sigma$
+% is constant over time, and thus a function of $B$ has no effect on
+% it. Therefore $\det \Phi^{-1}(B)\Sigma \Phi'^{-1}(B) = \det \Sigma$.
+
+\section{Distribution of the Eigenvalues}\label{sec:GCC-numerical}
+For the Wishart matrix, theoretical results are available for the
+eigenvalue distribution \cite{Chiani2012}. In the simplest case where
+$\Sigma = I$, i.e. no cross-correlations are present, we have
+\begin{eqnarray*}
+  f_{\lambda}(x_1, \cdots, x_N) = K \prod_{i=1}^N e^{-x_i/2}
+  x_i^{(T-N-1)/2} \prod_{i<j}^N (x_i - x_j)
+\end{eqnarray*}
+where the eigenvalues have been indexed in descending order: $x_1 \geq
+x_2 \geq \cdots \geq x_N$. The normalization constant K is given by
+\begin{equation*}
+  K = {\pi^{N^2/2} \over 2^{NT/2}\Gamma_N(T/2) \Gamma_N(N/2)}
+\end{equation*}
+where the function $\Gamma_m(a)$ is defined as
+\begin{equation*}
+  \Gamma_m(a) = \pi^{m(m-1)/4} \prod_{k=1}^m \Gamma\left(a - {k-1
+      \over 2}\right)
+\end{equation*}
+However, as detailed in the derivation leading to equation
+\ref{eq:cross-corr-matrix-PDF}, the distribution of $RR'$ is not
+Wishart when the columns of $R$ are correlated. Deriving the
+eigenvalue distribution analytically in this case is difficult and
+beyond the scope of this thesis. Instead we resort to numerical
+methods.
+
+As before we consider the AR(1) process (see \S
+\ref{sec:FundamentalConcepts}):
+\begin{equation*}
+  \vec{r}_t = \phi \vec{r}_{t-1} + \vec{a}_{t}
+\end{equation*}
+where $\vec{a}_t \sim N(0, I)$, i.e. the elements of $\vec{a}_t$ are
+independent Gaussian random variable with zero mean and unit
+variance.
+
+Now we investigate how the eigenvalue distribution depends on
+$\phi$. Figure \ref{fig:GaussianMarkovSpectrumPDF} shows the results
+of the simulation.
+% \begin{wrapfigure}{r}{0.5\textwidth}
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[scale=0.5, clip=true, trim=90 228 115
+    226]{../pics/GaussianMarkovSpectrumPDF.pdf}
+  \end{center}
+  \vspace{-10mm}
+  \label{fig:GaussianMarkovSpectrumPDF}
+  \caption{\footnotesize{
+    Eigenvalue distribution with $\tau$ ranging from 0 to
+    3. The 1st blue line, which is shown as stairs, is the
+    theoretical eigenvalue distribution according to the
+    Marcenko-Pastur law: $f(\lambda) = \sqrt{(\lambda_+/\lambda - 1)(1
+      - \lambda/\lambda_-)}/2\pi q \sigma^2$, where $\lambda_{\pm} =
+    \sigma^2(1 \pm \sqrt{q})^2$ In the simulation we have
+    chosen $q = N/T = 50/1000 = 0.05$ and $\sigma=1$. For each value
+    of $\tau$ we generate 2000 instances of $N \times T$ random matrix
+    $R$, and compute C as $C=RR'/T$. Hence each curve in the figure is
+    constructed from 2000 sets of eigenvalues.
+  }}
+\end{figure}
+It is clear from the figure that, while the minimum eigenvalue remains
+unchanged from its value given by the Marcenko-Pastur law, the maximum
+eigenvalue moves consistently to the right as the value of $\phi$
+increases. This prompts a study of the maximum eigenvalue distribution.
+
+K. Johnsson \cite{Johnsson2000} and I. Johnstone \cite{Johnstone2001}
+showed that, at the {\it absence} of autocorrelations and in the asymptotic
+limit $N, T \to \infty$, $N/T \to q < \infty$, the maximum eigenvalue
+$\lambda_1$ follows the Tracy-Widom distribution (denote
+$\mathscr{TW}_1$ here) when properly relocated and rescaled:
+\begin{equation*}
+  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} \sim \mathscr{TW}_1
+\end{equation*}
+where $\mu_{NT}$ and $\sigma_{NT}$ are given by
+\begin{eqnarray*}
+  \mu_{NT} &=& \left(
+    \sqrt{N-1/2} + \sqrt{T - 1/2}
+  \right)^2 \\
+  \sigma_{NT} &=& \sqrt{\mu_{NT}} \left(
+    {1 \over \sqrt{N-1/2}} + {1 \over \sqrt{T-1/2}}
+  \right)^{1/3}
+\end{eqnarray*}
+
+The $\mathscr{TW}_1$ distribution has the following cummulative
+distribution function (CDF) \cite{Chiani2012}
+\begin{equation*}
+  F_1(x) = \exp\left[
+    -{1 \over 2} \int_x^\infty dy \left(q(y) + (y-x)q^2(y)\right)
+  \right]
+\end{equation*}
+where $q(y)$ is defined as the solution to the Painlev\'e II differential
+equation
+\begin{equation*}
+  q''(y) = yq(y) + 2q^3(y)
+\end{equation*}
+which is unique when imposing the condition
+\begin{equation*}
+  q(y) \sim \text{Ai}(y) \text{ as } y \to \infty
+\end{equation*}
+
+Then Marco Chiani showed recently that the $\mathscr{TW}_1$ distribution
+can be well approximated by a gamma distribution based on his proof
+that the exact distribution of the maximum eigenvalue is a mixture of
+gamma distributions. Specifically,
+\begin{equation}\label{eq:TracyWidom-Gamma}
+  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} + \alpha \sim
+  \mathscr{G}(k ,\theta)
+\end{equation}
+where $\mathscr{G}(k, \theta)$ denotes the Gamma distribution with
+parameters $k$ and $\theta$ \cite{Chiani2012}. The probability
+density function (PDF) of the gamma distribution is given by
+\begin{equation*}
+  f_\gamma(x; k, \theta) = {1 \over \Gamma(k) \theta^k} x^{k-1}
+  e^{-x/\theta}
+\end{equation*}
+Moreover, the first 3 moments of the distribution are simple:
+\begin{eqnarray*}
+  E &=& k\theta \\
+  \text{variance} &=& k\theta^2 \\
+  \text{skewness} &=& 2/\sqrt{k}
+\end{eqnarray*}
+
+In the following we first verify these results and extend them to the
+case {\it with} autocorrelations and then study how the
+distribution changes as $\phi$ increases. Figure
+\ref{fig:GaussianMarkov005MaxEigCDF_loglog} compares the empirical
+maximum eigenvalue CDF with the CDF of a gamma distribution. Each pair
+of CDFs correspond to a fixed autocorrelation strength ($\phi$). The
+parameters $k$ and $\theta$ of the gamma distribution are fit to
+data by matching the 2nd and the 3rd moments of the gamma distribution
+to the corresponding moments of the empirical distribution. Then the
+parameter $\alpha$ in equation \ref{eq:TracyWidom-Gamma} is chosen to
+be
+\begin{equation*}
+  \alpha = k\theta - \mean{{\lambda_1 - \mu_{NT} \over \sigma_{NT}}}
+\end{equation*}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.5, clip=true, trim=30 198 32
+    166]{../pics/GaussianMarkov005MaxEigCDF_loglog.pdf}
+    \label{fig:GaussianMarkov005MaxEigCDF_loglog}
+    \caption{\footnotesize Cummulative distribution function (CDF) of the maximum
+      eigenvalue ($\lambda_1$). Blue: empirical CDF. Red: CDF of the
+      fitting gamma distribution. Both axis's are in log-scale.}
+  \end{center}
+  \vspace{-10mm}
+\end{figure}
+
+We can see in figure \ref{fig:GaussianMarkov005MaxEigCDF_loglog} that
+the gamma distribution fits fairly well to its corresponding maximum
+eigenvalue distribution. So we conclude that a gamma distribution not
+only approximates the maximum eigenvalue distribution at the absence of
+autocorrelations but does so even at the {\it presence} of
+autocorrelations. Autocorrelations influence the maximum eigenvalue
+distribution only via the parameters $k$, $\theta$ and $\alpha$. Next
+we study this influcence in detail.
+
+% Since the maximum eigenvalue distribution is well approximated by a
+% gamma distribution, which is characterized by the parmaters $k$ and
+% $\theta$, connecting the maximum eigenvalue distribution with the
+% autocorrelation means connecting $k$ and $\theta$ with the
+% autocorrelation strength $\phi$ or alternatively with the correlation
+% time $\tau$.
+
+Figure \ref{fig:GaussianMarkovMaxEig_k-phi} shows how the parameter $k$
+varies as the autocorrelation strengthens ($\phi$ increases). 
+\begin{figure}[htb!]
+  \vspace{-10mm}
+  \centering
+    \includegraphics[scale=0.5, clip=true, trim=100 223 112
+    141]{../pics/GaussianMarkov05MaxEig_k-phi.pdf}
+    \caption{\footnotesize The parameter $k$ against autocorrelation strength
+      $\phi$. Blue crosses: Empirical values of $k$. Red: best fitting
+      line in terms of {\it Least Square Errors}.}
+  \label{fig:GaussianMarkovMaxEig_k-phi}
+\end{figure}
+From the equation of the fitting line we can directly read out
+\begin{eqnarray*}
+  k &=& a\phi + b \\
+  &=& a\left(1 \over 2\right)^{1/\tau} + b
+\end{eqnarray*}
+where $a = -26$ and $b = 47$.
+
+For the parameter $\theta$, its behavior is more conveniently
+described in terms of the correlation time $\tau$. Figure
+\ref{fig:GaussianMarkovMaxEig_theta-tau} plots the values of $\theta$
+against those of $\tau$ together with a fitting quadratic
+function. Higher order polynomials provide a slightly better
+fit, but coefficients of the 3rd order and above are less than 1/1000
+times the coefficients of the 2nd and the 1st order. Therefore a 2nd
+order polynomial has been chosen.
+\begin{figure}[htb!]
+  \vspace{-15mm}
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=99 230 114
+  139]{../pics/GaussianMarkov05MaxEig_theta-tau.pdf} 
+  \caption{\footnotesize parameter $\theta$ against correlation time $\tau$. Blue:
+    empirical values of $\theta$. Cyan: Fitting quadratic function.}
+  \label{fig:GaussianMarkovMaxEig_theta-tau}
+\end{figure}
+The equation of this polynomial is
+\begin{eqnarray*}
+  \theta &=& A\tau^2 + B\tau + C
+\end{eqnarray*}
+where $A = 2.1\times 10^{-4}$, $B = -1.1\times 10^{-4}$, $C =
+3.2\times 10^{-4}$.
+
+The parameter $\alpha$ shifts the gamma distribution $\mathscr{G}(k,
+\theta)$ to match the mean of ${(\lambda_1 -
+  \mu_{NT})/\sigma_{NT}}$. The behavior of $\alpha$ with respect to
+changing autocorrelation is shown in figure
+\ref{fig:GaussianMarkovMaxEig_alpha-tau}.
+\begin{figure}[htb!]
+  \vspace{-15mm}
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=104 229 114
+  139]{../pics/GaussianMarkovMaxEig_alpha-tau.pdf}
+  \caption{\footnotesize The mean-shift parameter $\alpha$ against
+    correlation time $\tau$. Blue: empirical values of $\alpha$. Cyan:
+    quadratic fit.}
+  \label{fig:GaussianMarkovMaxEig_alpha-tau}
+\end{figure}
+The equation of $\alpha$ is then inferred from the fitting line:
+\begin{equation*}
+  \alpha = D\tau^2 + E\tau + F
+\end{equation*}
+where $D = -3.4\times 10^{-3}$, $E = -4.6\times 10^{-2}$ and $F = 69$.
+
+So combining the results of $\alpha$, $k$ and $\theta$ we can express
+the moments of the Tracy-Widom variate ${(\lambda_1 -
+  \mu_{NT})/\sigma_{NT}}$:
+\begin{equation}\label{eq:lambda1_MeanVariance}
+  \begin{aligned}
+    E({\lambda_1 - \mu_{NT} \over \sigma_{NT}}) &= k\theta -
+    \alpha \\
+    &= \left[-a\left(1 \over 2\right)^{1/\tau} + b\right](A\tau^2 +
+    B\tau - C) - (D\tau^2 + E\tau + F) \\
+    \text{variance}({\lambda_1 - \mu_{NT} \over \sigma_{NT}}) &=
+    k\theta^2 \\
+    &= \left[-a\left(1 \over 2\right)^{1/\tau} + b\right](A\tau^2 +
+    B\tau - C)^2 \\
+  \end{aligned}
+\end{equation}
+Figure \ref{fig:GaussianMarkovMaxEig_tw_moments} shows the empirical
+moments of ${(\lambda_1 - \mu_{NT})/\sigma_{NT}}$ together with the
+corresponding values computed using the above formulas.
+\begin{figure}[htb!]
+  \vspace{-15mm}
+  \centering
+  \includegraphics[scale=0.46, clip=true, trim=48 243 6
+  134]{../pics/GaussianMarkovMaxEig_tw_moments.pdf}
+  \caption{\footnotesize empirical moments of ${(\lambda_1 -
+      \mu_{NT})/\sigma_{NT}}$ against their theoretical
+    counterparts. Left: Empirical/theoretical mean against $\tau$;
+    Right: Empirical/theoretical variance against $\tau$. Blue
+    crosses: empirical values; Red line: fitting curves. Horizontal
+    axis: correlation time $\tau$.}
+\label{fig:GaussianMarkovMaxEig_tw_moments}
+\end{figure}
+The good fitness shown in figure
+\ref{fig:GaussianMarkovMaxEig_tw_moments} allows us to conclude 
+that, within the range of the correlation time that we have studied,
+namely $\tau \in [0, 13.51]$, the $k$ parameter is a linear function
+of $\phi = 2^{-1/\tau}$ while $\theta$ and $\alpha$ are quadratic
+functions of $\tau$. The moments of the transformed maximum
+eigenvalue, namely ${(\lambda_1 - \mu_{NT})/\sigma_{NT}}$, are thus
+expressed as functions of the correlation time $\tau$ via the
+parameters $k$, $\theta$ and $\alpha$.
+
+Figure \ref{fig:GaussianMarkovMaxEigPDF_original} shows the PDF of the
+maximum eigenvalue ($\lambda_1$) for a range of values of the
+correlation time $\tau$. One can clearly see that the mean of the
+distribution moves to the right and the width of the distribution
+increases as $\tau$ takes on larger and larger values. However, one
+must not forget that the behavior of the mean and the width is random
+in nature. Eq. \ref{eq:lambda1_MeanVariance} describes such behavior in
+the asymptotic limit, i.e. if an infinite number of random matrices
+are generated and their respective maximum eigenvalues computed for
+each and every value of correlation time $\tau$.
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[scale=0.5, clip=true, trim=98 228 112
+    221]{../pics/GaussianMarkovMaxEigPDF_original.pdf}
+    \label{fig:GaussianMarkovMaxEigPDF_original}
+    \caption{Probability density function of the maximum eigenvalue
+      ($\lambda_1$).}
+  \end{center}
+\end{figure}
+
+% The above observations lead to the conclusion that the
+% autocorrelations are fully characterized by the parameters $k$ and
+% $\theta$, which determine the gamma distribution via the 1st and the
+% 2nd moments:
+% \begin{eqnarray*}
+%   E(\lambda_1) &=& k\theta \\
+%   \text{var}(\lambda_1) &=& k\theta^2
+% \end{eqnarray*}
+
+% Table \ref{tab:GammaDistParam} shows the values of $k$ and $\theta$
+% corresponding to different values of $\phi$.
+% \begin{table}
+%   \begin{footnotesize}
+%   \centering
+%   \begin{tabular}{c|c|c || c|c|c }
+%     \hline
+%     $\phi$ & $k$ & $\theta$ &  $\phi$ & $k$ & $\theta$ \\
+%     \hline
+%     0  &  48.4730  &   0.0002  &   0.5000  &  66.9245  &   0.0003  \\
+%     0.0500  &  32.3891  &   0.0002  &   0.5500  &  21.1857  &   0.0007  \\
+%     0.1000  &  54.4910  &   0.0002  &   0.6000  &  49.2886  &   0.0005  \\
+%     0.1500  &  42.6755  &   0.0002  &   0.6500  &  39.0294  &   0.0007  \\
+%     0.2000  &  39.7918  &   0.0002  &   0.7000  &  37.8492  &   0.0009  \\
+%     0.2500  &  30.0915  &   0.0003  &   0.7500  &  27.5861  &   0.0016  \\
+%     0.3000  &  60.3928  &   0.0002  &   0.8000  &  29.6914  &   0.0021  \\
+%     0.3500  &  26.1387  &   0.0004  &   0.8500  &  19.5804  &   0.0047  \\
+%     0.4000  &  36.8579  &   0.0003  &   0.9000  &  28.8284  &   0.0079  \\
+%     0.4500  &  29.0491  &   0.0004  &   0.9500  &  18.2793  &   0.0375  \\
+%     \hline
+%   \end{tabular}
+%   \label{tab:GammaDistParam}
+%   \caption{Values of $k$ and $\theta$ for different $\phi$}
+% \end{footnotesize}
+% \end{table}
+% The values of $k$ are random while those of $\theta$ have a trend to
+% increase. It is hence reasonable to hypothesize that more probability
+% is distributed to large values of $\theta$ as $\phi$ gets larger. The
+% values of $\theta$ are plotted against those of $\phi$ in figure
+% \ref{fig:GaussianMarkovThetaPhi}. One can see that the $\theta$ values
+% grow faster than an exponential function after exceeding a critical value.
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[scale=0.35, clip=true, trim=58 220 87
+%     133]{../pics/GaussianMarkovThetaPhi.pdf}
+%   \end{center}
+%   \caption{$\theta$ values againt $\phi$ values in linear scale
+%     (left) and in semi-log scale (right)}
+%   \label{fig:GaussianMarkovThetaPhi}
+% \end{figure}
+% The smallness of the $\theta$ values and their behavior of increasing
+% with $\phi$ explain the pattern shown in figure
+% \ref{fig:GaussianMarkovMaxEigPDF_original}: The mean as well as the variance
+% of the $\lambda_1$ distribution increases with $\phi$; The effect of
+% right-shifting dominates before a critical $\phi$ value, while the
+% effect of widening dominates after the critical value.
+
+\chapter{Cross-Correlation Matrix with Fat-Tailed Returns}
 \bibliographystyle{plain}

### 2014-02-07 AsymmetricModel.m

@@ -1,0 +1,61 @@
+clear all
+% filename = './data/OMXS30.csv';
+% filename = 'data/VOLV-B-1990-01-01-2014-01-09.csv';
+% filename = 'data/ERIC-B-1990-01-02-2014-01-16.csv';
+% data = dlmread(filename, ';', [1, 1, 6857, 3]);
+
+% price = data(data(:, 1) > 0 & data(:, 3) > 0 &...
+%             data(:, 1) ~= data(:, 2), 3);
+% % price = data(data(:, 3) > 0, 3);
+% price = flipud(price);
+% N = size(price, 1);
+% T = 15;
+% % ret = price2ret(price(mod(N, T):T:end));
+% % ret = ret - mean(ret);
+% r = price2ret(price);
+% r = r - mean(r);
+% sigma = std(r);
+% r = (r - mean(r))/std(r);
+
+
+%% intraday returns
+company = 'ericsson_b';
+first_day = '2013-10-10';
+last_day = '2014-01-20';
+dt = 30;
+ret = get_intra_ret(company, first_day, last_day, dt);
+r = ret - mean(ret);
+sigma = std(r);
+
+C = sqrt(2/pi)*exp(log(log(4))/2);
+f = @(x, s, A) A*...
+    exp(s^2/2)/sqrt(8*pi)/C .* ...
+    erfc(1/(s*sqrt(2)) * log(abs(x*A)/sqrt(log(4))) + s/sqrt(2));
+
+p = mle(r, 'pdf', f, 'start', [1, 1/std(r)], 'lowerbound', [1e-4, 1e-4]);
+% [y, x] = ecdf(r);
+% y1 = F(x, sig);
+% plot(x, y, x, y1, 'r');
+
+% x = linspace(min(r), max(r), length(r)/100);
+% y = hist(r, x)./length(r)/(x(2)-x(1));
+% plot(x, y, x, f(x, p(1), p(2)), 'r');
+
+
+% plot(x(I), y(I), x(I), f(x(I), p(1), p(2)), 'r');
+% plot(log(x(I)), log(y(I)), log(x(I)), log(f(x(I), p(1), p(2))),
+% 'r');
+[y, x] = ecdf(r);
+y1 = retcdf_symmetric(x, p(1), p(2));
+I = x < 0;
+% I = x < -2*sigma & x > -3*sigma;
+% I = x > 2*sigma & x < 3*sigma;
+% I = logical(ones(1, length(x)));
+
[…]

### 2014-02-07 OMXS30.m

@@ -1,0 +1,32 @@
+clear all
+filename = './data/OMXS30.csv';
+data = dlmread(filename, ';', [1, 1, 6849, 6]);
+
+% data = data(data(:, 1) > 0 && data(:, 6) > 0 &&...
+%             data(:, 1) ~= data(:, 2), :);
+price = data(data(:, 3) > 0, 3);
+price = flipud(price);
+N = size(price, 1);
+
+T = 15;
+ret = price2ret(price(mod(N, T):T:end));
+mu = mean(ret);
+r = price2ret(price(mod(N, T):1:end));
+r = r - mean(r);
+r = reshape(r, T, length(ret));
+sig = sqrt(sum(r.^2))';
+
+logv = log(sig);
+L = length(logv);
+rho = reshape(autocorr(logv, L-1), 1, L);
+
+a = nan(1, L);
+omg = nan(1, L);
+for n = 0:L-1
+    a(n+1) = 2*(rho./sqrt(L))*(cos(2*pi*n*[0:L-1]/L)'./sqrt(L));
+end
+omg = 2*pi*[0:L-1]/L;
+I = omg < 1;
+plot(omg(I), log(a(I)));
+% plot(log(omg(I)), log(a(I)), 'x');
+

### 2014-02-07 get_intra_ret.m

@@ -1,34 +1,21 @@
-function returns = get_intra_ret(company, first_day, last_day, ...
-                                          interval, width, cmpt, ...
-                                          agrgt)
-if cmpt
-    cmpt_intra_ret(company, first_day, last_day, interval, width);
-end
+function ret = get_intra_ret(company, first_day, last_day, dt)
+    days = get_recorded_days(company, first_day, last_day);
 
-stmt = sprintf('ls data/%s_%dsec_ret_*.mat', company, interval);
-[status, output] = system(stmt);
-files = strsplit(output);
+    num = ceil(60*8.5 - dt);
 
-fmt = 'yyyy-mm-dd';
-ds = datenum(first_day, fmt);
-de = datenum(last_day, fmt);
-returns = [];
-l = 1;
-done = 0;
-while l <= length(files) && ~done
-    daystr = regexp(files(l), '[0-9]{4}-[0-9]{2}-[0-9]{2}', 'match');
-    daynum = datenum(daystr{1}, fmt);
-    if daynum >= ds && daynum <= de
-        if agrgt
-            load(files{l}, 'meta');
-            returns = [returns; meta.agrgt_ret];
-        else
-            load(files{l}, 'ret');
-            returns = [returns; ret];
+    ret = NaN(length(days) * num, 1);
+    N = NaN(length(days), 1);
+    n = 0;
+    for d = 1:length(days)
+        price = get_interpolated_prices(company, days{d}, 'minute', 1);
+        if length(price) <= dt
+            continue;
         end
-        if daynum == de
-            done = 1;
-        end
-        l = l + 1;
+        logprice = log(price);
+        r = logprice(dt+1:end) - logprice(1:end-dt);
+        N(d) = length(r);
+        ret(n+1: n + N(d)) = r;
+        n = n + N(d);
     end
-end
+    ret = ret(1:n);
+    

### 2014-02-07 interday_ret_distr.m

@@ -5,27 +5,25 @@ close all
 % filename = 'data/volvo_b_20121203-20131202.csv';
 
 %% daily returns
-filename = 'data/nasdaq.csv';
-price = dlmread(filename, ',', [1, 4, 10807, 4]);
+% filename = 'data/nasdaq.csv';
+filename = 'data/VOLVO-B-1990-01-01-2014-01-09';
+price = dlmread(filename, ',', [1, 6, 6029, 6]);
 % filename = 'data/sp.csv';
 % price = dlmread(filename, ',', [1, 4, 16084, 4]);
 price = flipud(price); 
 ret = price2ret(price);
-r = (ret - mean(ret))/std(ret);
+r = (ret - mean(ret));
 
 
 %% intraday returns
 % company = 'volvo_b';
 % first_day = '2013-10-10';
-% last_day = '2013-12-04';
-% dt = 5;
-% % interval for calculating realized volatility. in seconds
-% delta = 10;
-% [ret, v] = get_intra_ret_simple(...
-%     company, first_day, last_day, dt, delta);
+% last_day = '2014-01-10';
+% dt = 30;
+% [ret, v] = get_intra_ret_simple(company, first_day, last_day, dt, dt*2);
 % r = (ret - mean(ret))/std(ret);
 
-[p1, p2] = extract_xpnt(r, 2.56e-2/std(ret), 5.0e-2/std(ret));
+[p1, p2] = extract_xpnt(r, 2, 3);
 % [p1, p2] = extract_xpnt(r, 0.045/std(ret));
 fprintf('Left tail exponent: %.4f\nRight tail exponent: %.4f\n', ...
         p1, p2);

### 2014-02-07 retcdf.m

@@ -1,0 +1,17 @@
+function p = retcdf(x, s, A)
+C = sqrt(2/pi)*exp(log(log(4))/2);
+p = NaN(size(x));
+I = isfinite(x);
+
+p(I) = (x(I)<0).*sqrt(log(4))./C.*exp(s^2./2)./sqrt(8.*pi).*(...
+    A.*x(I)./sqrt(log(4)).*erfc(...
+        1./(s.*sqrt(2)) .* log(abs(x(I).*A)./sqrt(log(4))) + s./sqrt(2))...
+    + exp(-s^2./2).*erfc(1./(s.*sqrt(2)) .* log(abs(x(I).*A)./sqrt(log(4)))))...
+       +(x(I)>=0).*(1./2) + (x(I)>=0).*sqrt(log(4))./C.*exp(s^2./2)./sqrt(8.*pi).*(...
+           A.*x(I)./sqrt(log(4)).*erfc(...
+               1./(s.*sqrt(2)) .* log(abs(x(I).*A)./sqrt(log(4))) + s./sqrt(2))...
+           + exp(-s^2./2).*erfc(-1./(s.*sqrt(2)) .* log(abs(x(I).*A)./sqrt(log(4)))));
+
+
+I = ~isfinite(x);
+p(I) = (x(I) > 0);

### 2014-02-07 tmp.m

@@ -1,11 +1,27 @@
 clear all
-m = 902;
-l = 49;
+syms N A b;
 
-k = 1:m/2;
-s1 = prod(2*k ./ (2*k-1));
+assume(N, 'integer');
+assume(b > 0);
+assumeAlso(b < 1);
 
-k=1:(l+1)/2;
-s2 = prod((m+2*k-2) ./ (m+l+2*k));
+R = sym('r%d%d', [2, 2]);
+assume(R, 'real');
 
-s = s1 * s2 / m / pi;
+M = eye(3);
+% b = 0.5;
+M = sym(M);
+for i = 1:size(M, 1)-1
+    M(i, i+1) = -b;
+end
+
+% A = R*M;
+
+% [V1, E1] = eig(R*R');
+% [V2, E2] = eig(R'*R);
+[V, E] = eig(M*M');
+
+% t1 = trace(A*A');
+% t2 = trace(A*A'*R*R');
+
+% isequaln(R'*V1, V2)

### 2014-04-30 thesis/cross_correlation_gaussian.tex
1st version that has been handed to Sven. Gaussian Wishart Matrices are normized by 1/T.

\chapter{Correlation Matrix of Gaussian Returns}\label{chp:Gaussian}
In this chapter we present some analytical and numerical results about
the cross-correlation matrix (cf. \S
\ref{sec:FundamentalConcepts}), in the special case where the involved
asset returns are assumed to be Gaussian distributed. This assumption
is of course an over-simplification, but nevertheless lends some
insight into the problem.

Our primary interest is in the influence of autocorrelations on the
cross-correlation matrix. \S \ref{sec:GCC-analytical} discusses the
distribution of the matrix entries and \S \ref{sec:GCC-numerical} the
distribution of the eigenvalues.

\section{Distribution of the Matrix Elements}\label{sec:GCC-analytical}
If the returns follow a zero-mean Gaussian distribution, 
i.e. $\vec{r}_t \sim N(0, \Sigma)$, and are not
auto-correlated, the L\'evy index $\alpha$ as appears in equation
\ref{eq:cross-correlation} is 2, and $RR'$ is a Wishart matrix with
the following distribution law \cite{Anderson2003}:
\begin{equation}
  \label{eq:WishartPDF}
  f_{RR'}(X) = { |\det X|^{(T-N-1)/2} \exp\left(-{1 \over 2}\tr
      \Sigma^{-1}X \right)
    \over
    2^{NT/2}\pi^{N(N-1)/4}|\det \Sigma|^{T/2}
    \prod_{i=1}^N \Gamma\left[(n+1-i)/2\right]
  }
\end{equation}
where $\Sigma$ is the true covariance matrix of the returns.

When $\vec{r}_t$ are indeed auto-correlated, $RR'$ no longer has the
Wishart distribution. The joint distribution function of its elements
can be expressed in terms of the Wishart PDF and the
auto-correlations. We show this in appendix
\ref{app:pdf_gaussian1}. In the rest of this section, we derive an
approximate expression for the asymptotic distribution of these matrix
elements.

In the following we denote the diagonal elements as $C_{ii}$ and the
non-diagonal elements as $C_{ij}$ ($i \neq j$). We assume that
$\var(a_{i,t}) = \sigma^2$ for any $i$ and $t$, and $a_{i,t}$ are not
autocorrelated, i.e. $\text{corr}(a_{i,t}, a_{i, t'}) = 0$ for any
$i$, $t$ and $t'$. Then we can write
\begin{eqnarray*}
  r_{i,t} r_{j, t} &=& \left[
    \phi_1 r_{i, t-1} + a_{i, t}
  \right] \left[
    \phi_1 r_{j, t-1} + a_{j, t}
  \right] \\
  C_{ij} &=& {1 \over T}\sum_{t=1}^T r_{i,t} r_{j, t} \\
  &=& \phi_1^2 {1 \over T}\sum_{t=1}^Tr_{i,t-1} r_{j, t-1} +
  \phi_1{1 \over T}\sum_{t=1}^T\left(
    r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t}
  \right) + {1 \over T}\sum_{t=1}^T a_{i,t} a_{j,t}
\end{eqnarray*}
We note that the two sums $\sum_{t=1}^T r_{i,t} r_{j, t}$ and
$\sum_{t=1}^T r_{i,t-1} r_{j, t-1}$ only differ by the first and the
last addend, which is negligible for sufficiently large T. Thus we
have
\begin{equation}\label{eq:Offdiag1}
  \begin{aligned}
    (1 - \phi_1^2)C_{ij} &\approx& \phi_1 {1 \over T} \sum_{t=1}^T
    \left(r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t} \right)
    + {1 \over T}\sum_{t=1}^T a_{i,t} a_{j,t}
  \end{aligned}
\end{equation}
Now we write the AR(1) process $r_{i,t}$ as an infinite moving-average
process:
\begin{eqnarray*}
  r_{i, t} &=& \phi_1 r_{i, t-1} + a_{i,t} \\
  (1 - \phi_1 B) r_{i, t} &=& a_{i,t} \\
\end{eqnarray*}
where $B$ is the back-shift operator. Then it follows from the above
equation
\begin{eqnarray*}
  r_{i,t} &=& {1 \over 1 - \phi_1 B} a_{i,t} \\
  &=& \sum_{k=0}^\infty \phi_1^k B^k a_{i,t} \\
  &=& \sum_{k=0}^\infty \phi_1^k a_{i,t-k} \\
\end{eqnarray*}
where it is left implicit that $a_{i, t}$ with $t \leq 0$ is zero (in
words, this implies that the $r_{i,t}$ process is not affected at all
by events before $t = 1$).

Substituting this into eq.\ref{eq:Offdiag1} for $r_{i,t-1}$ yields
\begin{equation}
  \label{eq:Offdiag2}
  \begin{aligned}
  (1 - \phi_1^2)C_{ij} &\approx
  {1 \over T} \sum_{t=1}^T \sum_{k=0}^\infty\phi_1^{k+1} (a_{i, t-k-1}
  a_{j, t} + a_{j, t-k-1} a_{i, t})
  + {1 \over T} \sum_{t=1}^T a_{i,t} a_{j,t} \\
  &=
  {1 \over T} \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi_1^k (a_{i, t-k}
  a_{j, t} + a_{j, t-k} a_{i, t})
  + {1 \over T} \sum_{t=1}^{T} a_{i,t} a_{j,t}\\
  \end{aligned}
\end{equation}

Given two Gaussian random variables $x$ and $y$ with zero mean and
covariance matrix 
\begin{equation*}
  \Sigma =
  \begin{pmatrix}
    1 & \rho \\
    \rho & 1 \\
  \end{pmatrix}
\end{equation*}

The PDF of $xy$ can be found by considering $P(xy < z)$:
\begin{eqnarray*}
  P(xy < z) &=& \left(\int_0^\infty dx \int_0^{z/x} dy
    +\int_{-\infty}^0 dx \int_{z/x}^\infty dy \right)
    {1 \over 2\pi \sqrt{1 - \rho^2}} \\
    &&
    \exp\left[
      -{x^2 -2\rho xy + y^2
        \over
        2\sigma^2 (1 - \rho^2)} 
    \right] \\
  f(z; \sigma, \rho) &=& {d \over dz} P(xy < z)\\
  &=& {1 \over \pi \sqrt{1 - \rho^2}} \exp\left[
    {\rho z \over \sigma^2 (1 - \rho^2)}\right] K_0\left[
    {|z| \over \sigma^2 (1 - \rho^2)}
  \right]
\end{eqnarray*}
where $K_n(z')$ is the modified Bessel function of the second
kind. It is worth taking note that, when $\rho \neq 0$,
$f(z; \sigma, \rho)$ is not symmetric with respect to
$z$. As a result, if $\rho > 0$, the mean of $f(z;
\sigma, \rho)$ is positive, and vice versa.

Secondly, because $|\rho| < 1$ and
\begin{equation*}
  K_0(z) \sim \sqrt{\pi \over 2z} e^{-z}
\end{equation*}
as $z \to \infty$ \cite{Olver:2010:NHMF}, all the moments of
$f(z; \sigma, \rho)$ are finite, implying the
applicability of the Lyapunov central limit theorem provided that $T$
is large, which is very often the case and what we assume here.

With this in mind, we observe that $\phi_1$ only affects the first sum in
\ref{eq:Offdiag2}. If $a_{i,t-k}$ and $a_{j,t}$ are not correlated for
non-zero $k$, $\phi_1$ will not affect the mean of
$C_{ij}$. Furthermore, it is also clear from equation
\ref{eq:Offdiag2} that the variance of $C_{ij}$ is
always increased by a non-zero $\phi_1$, regardless of the sign of
$\phi_1$. We compute this increment in the following.

In light of the above expression for $f(z; \sigma,
\rho)$, we rewrite equation \ref{eq:Offdiag2} as
\begin{equation}\label{eq:Cij_dist}
  \begin{aligned}
    (1 - \phi_1^2)C_{ij} \approx &
    {1 \over T}
    \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi_1^k (a_{i, t-k} a_{j, t} + a_{j,
      t-k} a_{i, t}) \\
    & + {1 \over T} \sum_{t=1}^{T} \sigma^2 (1 - \rho^2) {a_{i,t} \over
      \sigma \sqrt{1 - \rho^2}} {a_{j,t} \over \sigma \sqrt{1 - \rho^2}} \\
  \end{aligned}
\end{equation}
In addition, we assume
\begin{equation*}
  \text{corr}(a_{i, t-k}, a_{j, t}) = \left\{
    \begin{array}{l l}
      1 & \text{ if $i=j$ and $k = 0$ }\\
      \rho & -1 < \rho < 1. \text{ if $i \neq j$ and $k = 0$ }\\
      0 & \text{otherwise}
    \end{array}
  \right.
\end{equation*}
Then, because $a_{i, t-k}$ and $ a_{j, t}$ with $i \neq j$ and $k
> 0$ are not correlated, the mean of $a_{i, t-k} a_{j, t}$ is 0
($f(z; \sigma, 0)$ is symmetric), and the variance of it
can be found to be $\sigma^6$ using formula (10.43.19) of \cite{NIST:DLMF}:
\begin{equation*}
  \int_0^\infty dt K_\nu(t) t^{\mu-1} = 2^{\mu-2}
  \Gamma\left(
    {\mu + \nu \over 2}
  \right) \Gamma\left(
    {\mu - \nu \over 2}
  \right)
\end{equation*}
On the other hand, $a_{i,t}/\sigma \sqrt{1 - \rho^2}$ and
$a_{j,t}/\sigma \sqrt{1 - \rho^2}$ have variance $1/(1 - \rho^2)$
and are correlated - $\text{corr}(a_{i,t}, a_{j,t}) = \rho$. The mean
of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ can be found using formula
(10.43.22) of \cite{NIST:DLMF}, given that $-1 < \rho < 1$:
\begin{eqnarray*}
  \int_0^\infty t^{\mu - 1} e^{-at} K_\nu(t) dt &=& (\pi/2)^{1/2}
  \Gamma(\mu + \nu) \Gamma(\mu - \nu)(1 - a^2)^{-\mu/2 + 1/4} \times\\
  && P^{-\mu+1/2}_{\nu-1/2} (a)
\end{eqnarray*}
where $P^\mu_\nu(\cdot)$ is Ferrers function of the first
kind\footnote{ Ferrers function of the first kind is defined through
  the hypergeometric functon $F(a, b; c; z)$ \cite{NIST:DLMF}:
  \begin{equation}
    \label{eq:Ferrers_1st}
    \mathop{\mathsf{P}^{\mu}_{\nu}\/}\nolimits\!\left(x\right)=\left(\frac{1+x}{1-
        x}\right)^{\mu/2}\mathop{\mathbf{F}\/}\nolimits\!\left(\nu+1,-\nu;1-\mu;\tfrac
      {1}{2}-\tfrac{1}{2}x\right).
  \end{equation}
}. The result is
\begin{eqnarray*}
  \E\left[{a_{i,t}a_{j,t} \over \sigma^2 (1 - \rho^2)}\right]
  &=&
  {1 \over \sqrt{2\pi} (1 - \rho^2)^{5/4}} \left[
    P^{-3/2}_{-1/2}(-\rho) - P^{-3/2}_{-1/2}(\rho)
  \right]
\end{eqnarray*}
Similarly, the variance of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ is
found to be
\begin{eqnarray*}
  v^2(\rho) &=&
  {4 \over \sqrt{2\pi} (1 - \rho^2)^{7/4}} \left[
    P^{-5/2}_{-1/2}(\rho) + P^{-5/2}_{-1/2}(-\rho)
  \right] \\
  && - \E^2\left[{a_{i,t}a_{j,t} \over
      \sigma^2 (1 - \rho^2)}\right]
\end{eqnarray*}
Now we can apply the Lyapunov central limit theorem
\cite{Billingsley1995} to the sum in equation \ref{eq:Cij_dist}
and write down the asymptotic Gaussian distribution of $C_{ij}$:
\begin{equation*}
C_{ij} \sim N(\mu'_X, \sigma'_X)
\end{equation*}
where
\begin{eqnarray}
  \mu'_X &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi_1^2)(1 -
    \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
    P^{-3/2}_{-1/2}(\rho)
  \right] \label{eq:gaussian_mean}\\
  \sigma'^2_X &=& {1 \over (1 - \phi_1^2)^2}\left[
    \sum_{t=1}^T \sum_{k=1}^{t-1} 2\left(
      \phi_1^k \over T
    \right)^2 \sigma^6 + \sum_{t=1}^T
    {\sigma^4 (1 - \rho^2)^2 \over T^2} v^2(\rho)
  \right] \nonumber \\
  &=& {2 \sigma^6 \over T (1 - \phi_1^2)^2} \left[
    {\phi_1^2 \over 1 - \phi_1^2} -
    {\phi_1^2 (1 - \phi_1^{2T}) \over
      T(1 - \phi_1^2)}
  \right] + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
    T (1 - \phi_1^2)^2} \nonumber \\
  &\approx& {2 \sigma^6 \phi_1^2 \over T (1 - \phi_1^2)^3}
  + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
    T (1 - \phi_1^2)^2} \label{eq:gaussian_variance}
\end{eqnarray}
Equation \ref{eq:gaussian_mean} tells that, if two return series $i$
and $j$ are not correlated, auto-correlation in the returns does not
introduce a bias into the estimation of the cross-correlation; if,
however, the return series are indeed correlated, auto-correlation
in the returns rescales the cross-correlation through a multiplicative
factor $1/(1 - \phi_1^2)$.

In addition, equation \ref{eq:gaussian_variance} tells that
auto-correlation in the returns always makes the cross-correlation
estimation more noisy. Auto-correlation not only rescales the variance
of the no-autocorrelation estimation by $1/(1 - \phi_1^2)$ but even
adds an extra term ${2 \sigma^6 \phi_1^2 \over T (1 - \phi_1^2)^3}$.

% What we have obtained is a quantitative description of what has been
% said before: The mean of the non-diagonal entries of the
% cross-correlation matrix is not affected by autocorrelations in the
% returns, given that innovations $a_{i,t}$ of different assets and
% different times are un-correlated; the variance, however, is always
% increased by autocorrelations no matter the autocorrelation is
% positive or nagative. The increment is in the form of an additive term
% that scales as $\phi_1^2 / (1 - \phi_1^2)^3$.

We may apply a similar treatment to the diagonal elements of the
covariance matrix, which we denote as $C_{ii}$ here:
\begin{eqnarray*}
  C_{ii} &=& {1 \over T} \sum_{t=1}^T r^2_{it} \\
  &=& {1 \over T} \sum_{t=1}^T \sum_{l=0}^{t-1}\phi_1^l a_{i, t-l}
  \sum_{k=0}^{t-1}\phi_1^k a_{i, t-k} \\
  &=& {1 \over T} \sum_{t=1}^T \left[
    \sum_{k=0}^{t-1} \phi_1^{2k} a^2_{i, t-k} +
    \sum_{k,l = 0}^{t-1} \phi_1^{k+l} a_{i, t-k} a_{i, t-l}
  \right]
\end{eqnarray*}
By the Lyapunov central limit theorem under the assumption of large T,
the asymptotic distribution of $C_{ii}$ is Gaussian, the mean and
variance being
\begin{eqnarray}
  \E(C_{ii}) &=& {1 \over T}\left[
    \sum_{k=0}^{t-1} \phi_1^{2k} \sigma^2
  \right] \nonumber \\
  &=& {\sigma^2 \over (1 - \phi_1^2) T} \left[
    T - {\phi_1^2 (1 - \phi_1^{2T}) \over 1 - \phi_1^2}
  \right] \nonumber \\
  & \approx & {\sigma^2 \over 1 - \phi_1^2} \left[
    1 - {\phi_1^2 \over T}
  \right] \label{eq:gaussian_cii_mean}
\end{eqnarray}
and
\begin{eqnarray}
  \var(C_{ii}) &=& \sum_{t=1}^T \left[
    \sum_{k=0}^{t-1} {\phi_1^{4k} \sigma^4 \over T^2} 2 +
    \sum_{k,l=0}^{t-1} {\phi_1^{2(k+l)} \over T^2} \sigma^6
  \right] \nonumber \\
  &=& \sum_{t=1}^T \left[
    {2 \sigma^4 \over T^2} {1 - \phi_1^{4t} \over 1 - \phi_1^4} +
    {\sigma^6 \over T^2} \left(
      {1 - \phi_1^{2t} \over 1- \phi_1^2}
    \right)^2 \right] \nonumber \\
  &=& {2 \sigma^4 \over T (1 - \phi_1^4)} -
  {2 \sigma^4 \phi_1^4 (1 - \phi_1^{4T}) \over T^2(1 - \phi_1^4)^2} +
  \nonumber \\
  && {\sigma^6 \over T (1 -\phi_1^2)^2} -
  {2 \sigma^6 \phi_1^2 (1 - \phi_1^{2T}) \over T^2 (1 - \phi_1^2)^3} +
  {\sigma^6 \phi_1^4 (1 - \phi_1^{4T}) \over T^2 (1 - \phi_1^2)^2 (1 -
    \phi_1^4)} \nonumber \\
  &\approx& {2 \sigma^4 \over T (1 - \phi_1^4)} + {\sigma^6 \over T (1
    -\phi_1^2)^2} \label{eq:gaussian_cii_variance}
\end{eqnarray}
From equation \ref{eq:gaussian_cii_mean} we see that auto-correlation
in the returns increases the variance of the return series; and from
equation \ref{eq:gaussian_cii_variance} we see that the variance of
that variance estimation is also increased by
auto-correlations. Moreover, we note that $\var(C_{ii})$ scales with T
approximately as $1/T$, similar to the behavior of
$\var(C_{ij})$. This is to be compared with the case of GARCH returns
discussed in chapter \ref{chp:CrossCorrelationFat}.

% As for the diagonal elements of the cross-correlation matrix,
% i.e. $\var(r_{it})$, from the representation of $r_{it}$ as an
% inifinite moving average process, one can deduce immediately
% \begin{equation*}
%   \text{var}(r_{i,t}) \approx {\sigma^2 \over 1 - \phi^2}
% \end{equation*}
% as is mentioned earlier in equation \ref{eq:gaussian_Cii_variance}.

% However, to write $f_A(RMM'R')$ as a function of the eigen
% values of $RMM'R'$ is very involved and no theoretical results are
% known.

% In the more general case
% \begin{eqnarray*}
%   a_{i,t} &=& \left(
%     1 - \sum_{k=1}^p \phi_{i,k} B^k
%   \right) r_{i,t} \\
%   &=& \Phi_i(B)} r_{i,t
% \end{eqnarray*}
% where $B$ denotes the back shift operator acting on subscript $t$, we
% can write
% \begin{eqnarray*}
%   A &=&
%   \begin{pmatrix}
%     \Phi_1(B) & &\\
%     & \ddots & \\
%     && \Phi_N(B)
%   \end{pmatrix} R \\
%   &=& \Phi(B) R
% \end{eqnarray*}

% Following the same reasoning, $RR' \sim W(\Phi^{-1}(B)
% \Sigma \Phi'^{-1}(B), T)$. Substituting $\Phi^{-1}(B)
% \Sigma \Phi'^{-1}(B)$ for $\Sigma$ in the Wishart
% probability density function, we get
% \begin{eqnarray*}
%   && \tr\left\{\left[\Phi^{-1}(B)\Sigma \Phi'^{-1}(B)\right]^{-1}
%     RR'\right\}\\
%   &=& \tr \left\{\Phi'(B)\Sigma^{-1} \Phi(B) RR'\right\}\\
%   &=& \tr\left\{\Sigma^{-1}AA'\right\}\\
% \end{eqnarray*}
% As to $\det \Phi^{-1}(B)\Sigma \Phi'^{-1}(B)$, we note that $\Sigma$
% is constant over time, and thus a function of $B$ has no effect on
% it. Therefore $\det \Phi^{-1}(B)\Sigma \Phi'^{-1}(B) = \det \Sigma$.

\section{Distribution of the Eigenvalues}\label{sec:GCC-numerical}
For the Wishart matrix, theoretical results are available for the
eigenvalue distribution \cite{Chiani2012}. In the simplest case where
$\Sigma = I$, i.e. no cross-correlations are present, we have
\begin{eqnarray*}
  f_{\lambda}(x_1, \cdots, x_N) = K \prod_{i=1}^N e^{-x_i/2}
  x_i^{(T-N-1)/2} \prod_{i<j}^N (x_i - x_j)
\end{eqnarray*}
where the eigenvalues have been indexed in descending order: $x_1 \geq
x_2 \geq \cdots \geq x_N$. The normalization constant K is given by
\begin{equation*}
  K = {\pi^{N^2/2} \over 2^{NT/2}\Gamma_N(T/2) \Gamma_N(N/2)}
\end{equation*}
where the function $\Gamma_m(a)$ is defined as
\begin{equation*}
  \Gamma_m(a) = \pi^{m(m-1)/4} \prod_{k=1}^m \Gamma\left(a - {k-1
      \over 2}\right)
\end{equation*}
However, as detailed in the derivation leading to equation
\ref{eq:cross-corr-matrix-PDF}, the distribution of $RR'$ is not
Wishart when the columns of $R$ are correlated. Deriving the
eigenvalue distribution analytically in this case is difficult and
beyond the scope of this thesis. Instead we resort to numerical
methods.

As before we consider the AR(1) process (see \S
\ref{sec:FundamentalConcepts}):
\begin{equation*}
  \vec{r}_t = \phi \vec{r}_{t-1} + \vec{a}_{t}
\end{equation*}
where $\vec{a}_t \sim N(0, I)$, i.e. the elements of $\vec{a}_t$ are
independent Gaussian random variable with zero mean and unit
variance.

Now we investigate how the eigenvalue distribution depends on
$\phi$. Figure \ref{fig:GaussianMarkovSpectrumPDF} shows the results
of the simulation.
% \begin{wrapfigure}{r}{0.5\textwidth}
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[scale=0.5, clip=true, trim=90 228 115
    226]{../pics/GaussianMarkovSpectrumPDF.pdf}
  \end{center}
  % \vspace{-10mm}
  \caption{\small \it
      Eigenvalue distribution with $\tau$ ranging from 0 to
      3. The 1st blue line, which is shown as stairs, is the
      theoretical eigenvalue distribution according to the
      Marcenko-Pastur law: $f(\lambda) = \sqrt{(\lambda_+/\lambda - 1)(1
        - \lambda/\lambda_-)}/2\pi q \sigma^2$, where $\lambda_{\pm} =
      \sigma^2(1 \pm \sqrt{q})^2$ In the simulation we have
      chosen $q = N/T = 50/1000 = 0.05$ and $\sigma=1$. For each value
      of $\tau$ we generate 2000 instances of $N \times T$ random matrix
      $R$, and compute C as $C=RR'/T$. Hence each curve in the figure is
      constructed from 2000 sets of eigenvalues.
    }
  \label{fig:GaussianMarkovSpectrumPDF}
\end{figure}
It is clear from the figure that the maximum eigenvalue moves
consistently to the right as the value of $\phi$ increases. The change
of the minimum eigenvalue remains unclear from the figure, so we plot
in \ref{fig:Gaussian_mineig} their mean for each value of
$\phi$. There one can see that the minimum eigenvalue actually
increases with $\phi$.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=100 226 116
  133]{../pics/Gaussian_mineig.pdf}
  \caption{\small \it The the minimum eigenvalue versus
    auto-correlation strength $\phi$. For each value of $\phi$ 2000
    random matrices are generated and their eigenvalues are
    calculated. The minimum eigenvalue of each random matrix is noted
    and the mean of the 2000 such minimum eigenvalues are plotted
    against the chosen value of $\phi$. 20 values of $\phi$ are
    included in the plot, ranging from 0 to 0.95 with step size 0.05.}
  \label{fig:Gaussian_mineig}
\end{figure}

K. Johnsson \cite{Johnsson2000} and I. Johnstone \cite{Johnstone2001}
showed that, at the {\it absence} of autocorrelations and in the
asymptotic limit $N, T \to \infty$, $N/T \to q < \infty$, the maximum
eigenvalue  $\lambda_1$ follows the Tracy-Widom distribution (denote
$\mathscr{TW}_1$ here) when properly relocated and rescaled:
\begin{equation*}
  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} \sim \mathscr{TW}_1
\end{equation*}
where $\mu_{NT}$ and $\sigma_{NT}$ are given by
\begin{eqnarray*}
  \mu_{NT} &=& \left(
    \sqrt{N-1/2} + \sqrt{T - 1/2}
  \right)^2 \\
  \sigma_{NT} &=& \sqrt{\mu_{NT}} \left(
    {1 \over \sqrt{N-1/2}} + {1 \over \sqrt{T-1/2}}
  \right)^{1/3}
\end{eqnarray*}
So in the following we focus on the maximum eigenvalue distribution
and leave the minimum eigenvalue for future studies. The
$\mathscr{TW}_1$ distribution has the following cummulative
distribution function (CDF) \cite{Chiani2012}
\begin{equation*}
  F_1(x) = \exp\left[
    -{1 \over 2} \int_x^\infty dy \left(q(y) + (y-x)q^2(y)\right)
  \right]
\end{equation*}
where $q(y)$ is defined as the solution to the Painlev\'e II differential
equation
\begin{equation*}
  q''(y) = yq(y) + 2q^3(y)
\end{equation*}
which is unique when imposing the condition
\begin{equation*}
  q(y) \sim \text{Ai}(y) \text{ as } y \to \infty
\end{equation*}

Then Marco Chiani showed recently that the $\mathscr{TW}_1$ distribution
can be well approximated by a gamma distribution based on his proof
that the exact distribution of the maximum eigenvalue is a mixture of
gamma distributions. Specifically,
\begin{equation}\label{eq:TracyWidom-Gamma}
  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} + \alpha \sim
  \mathscr{G}(k ,\theta)
\end{equation}
where $\mathscr{G}(k, \theta)$ denotes the Gamma distribution with
parameters $k$ and $\theta$ \cite{Chiani2012}. The probability
density function (PDF) of the gamma distribution is given by
\begin{equation*}
  f_\gamma(x; k, \theta) = {1 \over \Gamma(k) \theta^k} x^{k-1}
  e^{-x/\theta}
\end{equation*}
Moreover, the first 3 moments of the distribution are simple:
\begin{eqnarray*}
  \text{mean} &=& k\theta \\
  \text{variance} &=& k\theta^2 \\
  \text{skewness} &=& 2/\sqrt{k}
\end{eqnarray*}

In the following we first verify these results and extend them to the
case {\it with} autocorrelations and then study how the
distribution changes as $\phi$ increases. Figure
\ref{fig:GaussianMarkov005MaxEigCDF_loglog} compares the empirical
maximum eigenvalue CDF with the CDF of a gamma distribution. Each pair
of CDFs correspond to a fixed autocorrelation strength ($\phi$). The
parameters $k$ and $\theta$ of the gamma distribution are fit to
data by matching the 2nd and the 3rd moments of the gamma distribution
to the corresponding moments of the empirical distribution. Then the
parameter $\alpha$ in equation \ref{eq:TracyWidom-Gamma} is chosen to
be
\begin{equation*}
  \alpha = k\theta - \E\left({\lambda_1 - \mu_{NT} \over \sigma_{NT}}\right)
\end{equation*}

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.5, clip=true, trim=0 238 0
    197]{../pics/GaussianMarkov005MaxEigCDF_loglog.pdf}
  \end{center}
  \caption{\small \it Cummulative distribution function (CDF) of the maximum
    eigenvalue ($\lambda_1$). Blue: empirical CDF. Red: CDF of the
    fitting gamma distribution. The curves are plotted on log-log
    scale.}
  \label{fig:GaussianMarkov005MaxEigCDF_loglog}
  %\vspace{-10mm}
\end{figure}

We can see in figure \ref{fig:GaussianMarkov005MaxEigCDF_loglog} that
the gamma distribution fits fairly well to its corresponding maximum
eigenvalue distribution. So we conclude that a gamma distribution not
only approximates the maximum eigenvalue distribution at the absence of
autocorrelations but does so even at the {\it presence} of
autocorrelations. Since the maximum eigenvalue distribution is
characterized by the parmaters $k$, $\theta$, and  $\alpha$, the
influence of the autocorrelations can be characterized by the
dependence of $k$, $\theta$, and  $\alpha$ on $\phi$. In the following
we study these dependences in more details.

Figure \ref{fig:param1_phi_dep} shows how the mean, variance,
and skewness of the fitted gamma distribution, as well as the
relocation parameter $\alpha$ denpend on $\phi$.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=37 240 50
    202]{../pics/param1_phi_dep.pdf}
  \caption{\small \it  the  mean, variance, and skewness of the fitted
    gamma distribution as well as the relocation parameter $\alpha$
    are plotted against the autocorrelation strength $\phi$. 20 values
    of $\phi$ are included in the plot, ranging from 0 to 0.95 with
    step size 0.05.}
  \label{fig:param1_phi_dep}
\end{figure}
Despite the monotonic behavior shown by 3 of these curves, the
relation between each of the aforementioned quantities and $\phi$ is
more complicated than it apears. However, the mean, i.e. $k \theta$ is
an exception: Good support in the data can be found for the following
approximate relation:
\begin{eqnarray}
  k\theta &=& a \tan^2{\pi \phi \over 2} + b\tan{\pi \phi \over 2} +
  c \label{eq:k_theta-phi}
\end{eqnarray}
To verify this relation, we first fit a 2nd order polynomial and
obtain the coefficients $a$, $b$, $c$; then for each data point
$k_n\theta_n$ we solve the quadratic equation
\begin{eqnarray}
  a \tan^2{\pi \phi'_n \over 2} + b\tan{\pi \phi'_n \over 2} + c -
  k_n\theta_n &=& 0\label{eq:k_theta-phi_2}
\end{eqnarray}
for $\tan{\pi \phi'_n \over 2}$. If relation \ref{eq:k_theta-phi} is a
good approximation, a close match between $\tan{\pi \phi'_n \over 2}$
and $\tan{\pi \phi_n \over 2}$ is expected. Figure
\ref{fig:phi_and_roots} plots $\tan{\pi \phi'_n \over 2}$ against
$\tan{\pi \phi_n \over 2}$ and fits a straight line to the data
points. It is seen that the data points lie fairly close to the fitted
line and the line has a slope very close to 1 and an intercept close to
0. This strongly suggests $\tan{\pi \phi'_n \over 2} = \tan{\pi \phi_n
  \over 2}$ and supports the relation \ref{eq:k_theta-phi}.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=37 217 39
    170]{../pics/phi_and_roots.pdf}
  \caption{\small \it Upper plot: $\tan{\pi \phi'_n \over 2}$ against
    $\tan{\pi \phi_n \over 2}$. The fitted line has equation $y_n =
    0.995 \tan{\pi \phi_n \over 2} + 0.0146$. Lower plot: Residuals of
    the linear fit, i.e. $\tan{\pi \phi'_n \over 2} - y_n$. 20 values
    of $\phi$ are included in the plot, ranging from 0 to 0.95 with
    step size 0.05.}
  \label{fig:phi_and_roots}
\end{figure}

% Figure \ref{fig:GaussianMarkovMaxEig_k-phi} shows how the mean of the
% fitted gamma distribution, i.e. $k\theta$, varies as the
% autocorrelation strengthens ($\phi$ increases). 
% \begin{figure}[htb!]
%   %\vspace{-10mm}
%   \centering
%     \includegraphics[scale=0.5, clip=true, trim=100 223 112
%     141]{../pics/GaussianMarkov05MaxEig_k-phi.pdf}
%     \caption{\small \it The parameter $k$ against autocorrelation strength
%       $\phi$. Blue crosses: Empirical values of $k$. Red: best fitting
%       line in terms of {\it Least Square Errors}.}
%   \label{fig:GaussianMarkovMaxEig_k-phi}
% \end{figure}
% From the equation of the fitting line we can directly read out
% \begin{eqnarray*}
%   k &=& a\phi + b \\
%   &=& a\left(1 \over 2\right)^{1/\tau} + b
% \end{eqnarray*}
% where $a = -26$ and $b = 47$.

% For the parameter $\theta$, its behavior is more conveniently
% described in terms of the correlation time $\tau$. Figure
% \ref{fig:GaussianMarkovMaxEig_theta-tau} plots the values of $\theta$
% against those of $\tau$ together with a fitting quadratic
% function. Higher order polynomials provide a slightly better
% fit, but coefficients of the 3rd order and above are less than 1/1000
% times the coefficients of the 2nd and the 1st order. Therefore a 2nd
% order polynomial has been chosen.
% \begin{figure}[htb!]
%   \vspace{-15mm}
%   \centering
%   \includegraphics[scale=0.5, clip=true, trim=99 230 114
%   139]{../pics/GaussianMarkov05MaxEig_theta-tau.pdf} 
%   \caption{\small \it Vertical axis: $\theta$; Horizontal axis:
%     correlation time $\tau$. Blue: empirical values of $\theta$. Cyan:
%     Fitting quadratic function.}
%   \label{fig:GaussianMarkovMaxEig_theta-tau}
% \end{figure}
% The equation of this polynomial is
% \begin{eqnarray*}
%   \theta &=& A\tau^2 + B\tau + C
% \end{eqnarray*}
% where $A = 2.1\times 10^{-4}$, $B = -1.1\times 10^{-4}$, $C =
% 3.2\times 10^{-4}$.

% The parameter $\alpha$ shifts the gamma distribution $\mathscr{G}(k,
% \theta)$ to match the mean of ${(\lambda_1 -
%   \mu_{NT})/\sigma_{NT}}$. The behavior of $\alpha$ with respect to
% changing autocorrelation is shown in figure
% \ref{fig:GaussianMarkovMaxEig_alpha-tau}.
% \begin{figure}[htb!]
%   \vspace{-15mm}
%   \centering
%   \includegraphics[scale=0.5, clip=true, trim=104 229 114
%   139]{../pics/GaussianMarkovMaxEig_alpha-tau.pdf}
%   \caption{\small \it The mean-shift parameter $\alpha$ against
%     correlation time $\tau$. Blue: empirical values of $\alpha$. Cyan:
%     quadratic fit.}
%   \label{fig:GaussianMarkovMaxEig_alpha-tau}
% \end{figure}
% The equation of $\alpha$ is then inferred from the fitting line:
% \begin{equation*}
%   \alpha = D\tau^2 + E\tau + F
% \end{equation*}
% where $D = -3.4\times 10^{-3}$, $E = -4.6\times 10^{-2}$ and $F = 69$.

% So combining the results of $\alpha$, $k$ and $\theta$ we can express
% the moments of the Tracy-Widom variate ${(\lambda_1 -
%   \mu_{NT})/\sigma_{NT}}$:
% \begin{equation}\label{eq:lambda1_MeanVariance}
%   \begin{aligned}
%     \E({\lambda_1 - \mu_{NT} \over \sigma_{NT}}) &= k\theta -
%     \alpha \\
%     &= \left[-a\left(1 \over 2\right)^{1/\tau} + b\right](A\tau^2 +
%     B\tau - C) - (D\tau^2 + E\tau + F) \\
%     \var({\lambda_1 - \mu_{NT} \over \sigma_{NT}}) &=
%     k\theta^2 \\
%     &= \left[-a\left(1 \over 2\right)^{1/\tau} + b\right](A\tau^2 +
%     B\tau - C)^2 \\
%   \end{aligned}
% \end{equation}
% Figure \ref{fig:GaussianMarkovMaxEig_tw_moments} shows the empirical
% moments of ${(\lambda_1 - \mu_{NT})/\sigma_{NT}}$ together with the
% corresponding values computed using the above formulas.
% \begin{figure}[htb!]
%   \vspace{-15mm}
%   \centering
%   \includegraphics[scale=0.46, clip=true, trim=48 243 6
%   134]{../pics/GaussianMarkovMaxEig_tw_moments.pdf}
%   \caption{\small \it empirical moments of ${(\lambda_1 -
%       \mu_{NT})/\sigma_{NT}}$ against their theoretical
%     counterparts. Left: Empirical/theoretical mean against $\tau$;
%     Right: Empirical/theoretical variance against $\tau$. Blue
%     crosses: empirical values; Red line: fitting curves. Horizontal
%     axis: correlation time $\tau$.}
% \label{fig:GaussianMarkovMaxEig_tw_moments}
% \end{figure}
% The good fitness shown in figure
% \ref{fig:GaussianMarkovMaxEig_tw_moments} allows us to conclude 
% that, within the range of the correlation time that we have studied,
% namely $\tau \in [0, 13.51]$, the $k$ parameter is a linear function
% of $\phi = 2^{-1/\tau}$ while $\theta$ and $\alpha$ are quadratic
% functions of $\tau$. The moments of the transformed maximum
% eigenvalue, namely ${(\lambda_1 - \mu_{NT})/\sigma_{NT}}$, are thus
% expressed as functions of the correlation time $\tau$ via the
% parameters $k$, $\theta$ and $\alpha$.

% Figure \ref{fig:GaussianMarkovMaxEigPDF_original} shows the PDF of the
% maximum eigenvalue ($\lambda_1$) for a range of values of the
% correlation time $\tau$. One can clearly see that the mean of the
% distribution moves to the right and the width of the distribution
% increases as $\tau$ takes on larger and larger values. However, one
% must not forget that the behavior of the mean and the width is random
% in nature. Eq. \ref{eq:lambda1_MeanVariance} describes such behavior in
% the asymptotic limit, i.e. if an infinite number of random matrices
% are generated and their respective maximum eigenvalues computed for
% each and every value of correlation time $\tau$.
% \begin{figure}[htb!]
%   \begin{center}
%     \includegraphics[scale=0.5, clip=true, trim=98 228 112
%     221]{../pics/GaussianMarkovMaxEigPDF_original.pdf}
%     \caption{\small \it Probability density function of the maximum eigenvalue
%       ($\lambda_1$).}
%   \end{center}
%   \label{fig:GaussianMarkovMaxEigPDF_original}
% \end{figure}

### 2014-04-30 thesis/diary.tex
1st version that has been handed to Sven. Gaussian Wishart Matrices are normized by 1/T.

\documentclass{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[swedish]{babel}

\title{Project Diary}
\author{Xie Xiaolei}
\date{\today}
\begin{document}

\maketitle

\begin{itemize}
\item 2013/08/26 - 2013/09/01

Read Sven's article about Wishart-Levy matrices as well as two
others referenced by it.

\item 2013/09/02 - 2013/09/06

Read Free probability theory which is used to derive the eigenvalue
spectrum of the Wishart-Levy matrix.

\item 2013/09/07 - 2013/09/12

Read the book ``An application to probability theory and its
applications'', trying to understand the concepts of the concepts of
probability measures, measurable sets and sigma algebras.

\item 2013/09/13 - 2013/09/24

Work on the Matlab program that analyzes the intra-day transaction
data. The purpose to study the tail behavior. The main difficulty is
that, in order to obtain good statistics, a lot of data and hence
computation time are needed.

\item 2013/09/25 - 2013/09/29

Improve the program to make full use of data and to make it more
efficient to go through the data.

\item 2013/09/30 - 2013/10/04

Study the auto-correlations of intra-day returns. These are expected to
decay exponentially. It turns out that the auto-correlations between
the returns decay to 0 in approximately the same time as the time lag
of the returns.

\item 2013/10/05 - 2013/10/11

Study the GARCH models. Fit a GARCH model to 15-minute returns of
Nordea Bank.

\item 2013/10/12

Read a paper about using the Johnson SU distribution to model the
residuals in a GARCH model.

\item 2013/10/13 - 2013/10/15

Calculate the log-likelihood function of a GARCH model whose residuals
have Johnson SU distributions.

\item 2013/10/16 - 2013/10/21

Write a program to perform maximum-likelihood estimate for GARCH
models with Johnson SU residuals.

\item 2013/10/22 - 2013/10/30

Read articles about ``realized volatility'':
\begin{enumerate}
\item realized volatility is the square root of the sum of squared
  returns that are sampled at a higher frequency that those one
  intends to model.
\item The square of realized volatility is the sample analogue of the
  quadratic variation of a stochastic process.
\item At what frequency should the returns be sampled to give a good
  estimate of the quadratic variation: Ideally, the higher the
  better. But due to market micro-structure, there is an optimal
  frequency, the determination of which is, however, a subject of
  debate.
\end{enumerate}

\item 2013/11/01 - 2013/11/02

Read a paper about alternative volatility proxies to the realized
volatility. The trading range, i.e. (highest - lowest)/lowest, is
proposed among others. The use of realized volatility relies on the
availability of intra-day transaction data. I have such data but am
limited to a few stocks in disconnected time periods.

\item 2013/11/03 - 2013/11/05

Try out the proposed volatility proxies against real data.

\item 2013/11/06 - 2013/11/07

Study the log-volatility of Nordea Bank. They are not Gaussian
distributed but rather have fat tails.

\item 2013/11/08 - 2013/11/15

Study time series analysis.

\item 2013/11/16 - 2013/11/24

Study the book ``Stochastic integration and differential
equations''. This book is probably meant for mathematicians. Theorems,
corollaries, and lemma follow one another.

\item 2013/11/25 - 2013/11/30
Study the book ``Stochastic Differential equation - An introduction
with Applications''. It is easier than the other book and more
oriented to applications.

\item 2013/12/01 - 2013/12/03
Make a Matlab function that retrieve n-minute returns together with
their realized volatility from transaction data.
Estimate a stochastic log-volatility model for 15min returns of
Nordea Bank. The realized volatilities are calculated using 30sec
returns.

\item 2013/12/04 - 2013/12/07
Analyze the tail behavior of the simplest possible stochastic
log-volatility model: r = m + a exp(s b) where a and b are independent
standard Gaussian r.v. Eventually I can express p(r) as
p(r) = A erfc(a ln|r| + b)

\item 2013/12/08 - 2013/12/15
Read about Wishart matrices and multivariate Gaussian distribution.
It is understood that the elements' distribution of a Wishart
matrix, a GOE, a Wiger ensemble, are characterized by the
eigenvalues.

\item 2013/12/16 - 2013/12/20
Derive an expression for the elements' distribution of RR', where
R is an NxT matrix, each row of which represents a time series.
$R_{it} = \phi R_{i, t-1} + a_{it}$. It turns out The joint
probability density of RR' is simply p(RR') = p(AA') where R = AM and
hence $A = RM^{-1}$.

\item 2013/12/21 - 2013/12/23
Try to find an analytic expression for the eigenvalue distribution
of RR' at the presence of autocorrelations.

\item 2013/12/27 - 2013/12/29
Read an article about the maximum eigenvalue distribution of the
Wishart matrix when the covariance matrix is just the identity
matrix. The paper gives exact expressions for the PDF and the CDF of
the maximum eigenvalue. It also shows that the PDF is a mixture of
gamma distributions with different weights and hence can be
approximated by a gamma distribution, namely the dominant one in the
mixture.

\item 2014/01/03 - 2013/01/10
Try to make a Matlab program that computes the exact maximum
eigenvalue PDF. The difficulty is to deal with quotients of gamma
functions where both the denominator and the numerator tend to
infinity at large T. One also needs to handle a determinant that tends
to 0 at large T.

\item 2014/01/11 - 2013/01/15
Use the approximate gamma distribution instead. Write Matlab programs
to simulate the return matrices, to fit the simulated maximum
eigenvalue distribution to a gamma distribution. Study how the fitting
gamma distributions change in response to increasing auto-correlations
among the returns.

\item 2014/01/16 - 2013/01/21
Even more simulations. We need the autocorrelations to change more
smoothly in order to find out how the fitting gamma distributions
change.

\item 2014/01/22 - 2013/01/28
Conclude the chapter about cross-correlation matrix with Gaussian
returns. Plan the next chapter of correlation matrix with fat-tailed
returns.

\item 2014/01/29 - 2013/02/05
Further develop my simple stochastic log-volatility model
\begin{eqnarray*}
  r_t &=& A e^{\sigma a_t} b_t
\end{eqnarray*}
The current model does not account for skewness. Instinct suggests
that allowing $a_t$ and $b_t$ to be correlated will introduce
skewness. Calculate the analytic PDF in this case. An integral has to
be solved to express the PDF in closed form.

\item 2014/02/06 - 2013/02/10
Still on the integral of the PDF. Have tried different ways to
approximate it. The integrand is really complicated.

\item 2014/02/11 - 2013/02/14
Study the method of steepest descent in order to approximate the
integral. It turns out to be inapplicable. Its main use is in studying
asymptotic behavior of integrals when a large parameter is present in
the exponent and makes it valid to expand the exponent around a few
local maxima.

\item 2014/02/14 - 2013/02/20
Read a paper about regularly varying tails of GARCH(1,1) models, a
book about extreme events that systematically treats different types
of heavy tails, and an article about the heavy-tail central limit
theorem, i.e. convergence to the $\alpha$-stable laws.

\item 2014/02/21 - 2014/02/26
Study the correlation matrix built from GARCH returns, write up the
results.

\end{itemize}
\end{document}

### 2014-04-30 thesis/price_models.tex
1st version that has been handed to Sven. Gaussian Wishart Matrices are normized by 1/T.

\chapter{Return Models}\label{chp:PriceModels}
In this chapter we review some of the discrete-time return models and
fit them to intraday returns. The intention is to assess the
applicability of these models and to understand their statistical
properties. The implications of these properties on the returns'
correlation and covariance matrix are studied in chapters
\ref{chp:Gaussian} and \ref{chp:CrossCorrelationFat}.

\S\ref{sec:Garch_model} reviews the GARCH model using the Nordea
15-minute returns as an example. \S\ref{sec:SLV_model} examines the
stochastic log-volatility model and discusses the possibility of
describing the residuals of the log-volotility series using the
Johnson Su distribution. The same Nordea return series is fitted as an
example. \S\ref{sec:XieCalc} derives the unconditional distribution
function of the stochastic log-volotility model and compares the
analytic results to the 30-minute returns of a number of Swedish
stocks.

\section{Gaussian Distribution}
The justification of modeling return series as independent,
identically distributed Gaussian variates comes from imagining the
price process $S(t)$ as a Brownian motion whose increment $\sigma
dw_t$ at each time step is independent and scales as $\sigma
\sqrt{dt}$. Then, by adding a drift term $\mu dt$ that represents some
deterministic trend in the price process, one can express the
price $S(t)$ as a stochastic differential equation:
\begin{eqnarray*}
  dS &=& S\mu dt + S\sigma dw_t \\
  d(\ln S) &=& (\mu - \frac{1}{2} \sigma^2)dt + \sigma dw_t \\
\end{eqnarray*}
It follows from this equation that $\ln S(t) - \ln S_0$, where $S_0$
is the price at time 0, has Gaussian distribution with mean $(\mu -
\sigma^2/2)t$ and variance $\sigma^2 t$. Therefore, in a discrete-time
model, where the length of each time step $\Delta t$ is fixed, the
return $r_t = \ln S(t + \Delta t) - \ln S(t)$ is assumed to be
Gaussian distributed and have mean and variance that are functions of
$\Delta t$.

The picture depicted above is of course overly simplified, and as we discussed in
\S\ref{sec:StylizedFacts}, the distribution of returns is not really
Gaussian. Nevertheless, the assumption of Gaussian distributed returns
underlies such important theories as Black and Scholes theory of
option pricing. Thus, albeit hugely inaccurate, the Gaussian
distribution as a returns model shall not be forgotten. In the next
few sections, we discuss some more realistic returns models.

\section{GARCH models} \label{sec:Garch_model}
``GARCH'' is the acronym for ``Generalized Autoregressive Conditional
Heteroscedasticity''. A GARCH(p, q) model is defined by the following
equation system:
\begin{equation}
  \label{eq:garch_def}
  \begin{aligned}
    r_t &= \mu + \epsilon_t \\
    \sigma_t^2 &= \alpha_0 + \sum_{i=1}^q \alpha_i \epsilon_{t-i}^2 +
    \sum_{i=1}^p \beta_i \sigma_{t-i}^2
  \end{aligned}
\end{equation}
where $\epsilon_t$ is a random variable of a prescribed distribution,
the most common ones being Gaussian, Student's t, Asymmetric Student's
t, etc; $\sigma_t^2$ is the variance of $\epsilon_t$
\cite{Bollerslev86}. The autocorrelation function of $\epsilon_t^2$ is
given by (c.f. \cite{Bollerslev87})
\begin{eqnarray*}
\rho_n &=& \sum_{i=1} ^{p \vee q} (\alpha_i + \beta_i) \rho_{n-i}
\text{ for $n > p$}
\end{eqnarray*}
where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
zeros. $p \vee q$ denotes the maximum of p and q. From these
equations, it is clear that the partial autocorrelation function cuts
off at $\max(p, q)$.

To assess the applicability of GARCH models at the face of intra-day
returns, we study the {\it Nordea Bank 15min} returns sampled during
the period 2012/01/16 - 2012/04/20. Firstly we plot the autocorrelation
function (ACF) of $\epsilon_t^2$ from this sample, as shown
in figure \ref{fig:nordea_15min_acf}.

At the absence of GARCH effects, the ACF is expected to have an
asymptotic Gaussian distribution with mean 0 and variance 1/T
\cite{Bollerslev86, Bollerslev87}, which is clearly not the case in
figure \ref{fig:nordea_15min_acf}: The first 5 autocorrelations have
comparable sizes and do not fall off as in a Gaussian
scheme. Moreover, figure \ref{fig:nordea_15min_vlt_acf} shows 
even more clearly that the conditional variances of the series are
correlated. These observations suggest a GARCH(p, q) model can be
appropriate.

\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    200]{../pics/nordea_15min_acf.pdf}
    \label{fig:nordea_15min_acf}
  }
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    200]{../pics/nordea_15min_vlt_acf.pdf}
    \label{fig:nordea_15min_vlt_acf}
  }
  \caption{\small \it Auto-correlations among Nordea Bank 15min
    returns. \ref{fig:nordea_15min_acf}: Auto-correlations (ACF) of the
    squared returns; \ref{fig:nordea_15min_vlt_acf}: Auto-correlations
    (ACF) of the conditional variances.}
\end{figure}

According to Andersen and Bollerslev et al \cite{Andersen03},
$\epsilon_t$ can be decomposed as $\epsilon_t = \sigma_t z_t$, where
$z_t \sim N(0, 1)$ and $\sigma_t$ can be estimated using realized
volatility. To find out whether this is true for the aforementioned
series we look at the probability plot of $\epsilon_t/\sigma_t$
(figure \ref{fig:nordea_bank_15min_z_prob}), which is consistent with
a standard Gaussian distribution except for a few outliers at the two
ends. Hence the fat tails of the probability density function can be
entirely accounted for by GARCH effects, implying a Gaussian
conditional distribution is sufficient for $z_t$.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=80 258 104
    220]{../pics/nordea_bank_15min_z_prob.pdf}
  \caption{\small \it{Probability plot of $z_t = \epsilon_t /
      \sigma_t$. $\epsilon_t$ are derived from Nordea Bank 15min
      returns while $\sigma_t$ are realized volatilities calculated
      using 30s returns within each 15min interval. Horizontal axis:
      $z_t$}; Vertical axis: cummulative probability function (CDF) of
    $z_t$, arranged on such a scale that the CDF of the standard
    Gaussian is a straight line.}
  \label{fig:nordea_bank_15min_z_prob}
\end{figure}

To determine the orders p and q of the GARCH model, we again check
the partial autocorrelation function of $\epsilon_t^2$. As just
mentioned, the first 5 partial autocorrelations are significant. Thus
we can tentatively estimate a GARCH(1, 5) model. Then by applying the
Lagrange multiplier test and accordingly removing insignificant
parameters, we obtain a GARCH(1, 3) model
\ref{tab:nordea_15min_garch}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    parameter & value & std. error & t statistic \\
    \hline
     $\alpha_0$ &   1.05668e-06 &   3.56871e-07 &    2.96096\\
     \hline
     $\beta_1$  &     0.687212  &   0.0299314   &    22.9596\\
     \hline
     $\alpha_1$ &      0.128614 &   0.0268144   &    4.79643\\
     \hline
     $\alpha_3$ &    0.0568022  &   0.0301503   &    1.88397 \\
     \hline
  \end{tabular}
  \caption{\small \it GARCH(1, 3) model of Nordea Bank 15min returns}
  \label{tab:nordea_15min_garch}
\end{table}

\section{ARIMA Log-volotility Models}\label{sec:SLV_model}
With the availability of transaction data, estimating conditional
volatility using returns sampled at a higher frequency (realized
volatility) gives superior acuracy and reliability. However, at which
frequency the time series should be sampled in order to give an
unbiased and consistent estimate of the volatility is not a trivial
question. Naively one would argue by the doctrine of statistics that
the often the series is sampled, the better the estimate. This is also
well supported in theory, since the quadratic variation which is the
theoretial variance of the time series (see \cite{Andersen03}, theorem
1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
\S6, theorem 22):
\[
\var(p_{t+h} - p_t) = [r, r]_{t+h} - [r, r]_t = \lim_{n \to \infty}
\sum_{T^i \in \pi_n} (p^{T_{i+1}} - p^{T_i})^2
\]
where $p_t$ is the log price process and $r_t = p_t - p_0$. $\pi_n$
is a refining partition of the interval $[t, t+h]$.

However, due to the noise introduced by market micro-structure, there
is actually an optimal sampling frequency relative to $h$. While the
method to determine this optimal frequency is a subject of debate (see
for example \cite{Sahalia05}), it is not difficult to find a fairly
satisfactory frequency in practice.

Andersen and Bollerslev et al proved in \cite{Andersen03} that, with
the conditional volatility well approximated by the realized
volatility, $r_{t+h, h} = p_{t+h} - p_t$ is governed by a Gaussian
distribution (see theorem 2 of \cite{Andersen03}):
\[
r_{t+h, h} | \sigma(\mu_{t+s}, \sigma_{t+s})_{s \in [0, h]} \sim
N(\int_{0}^h \mu_{t+s} ds, \int_{0}^h \sigma_{t+s}^2 ds)
\]

Thus one can try a few frequencies and compare the distribution of
$r_{t+h, h}/\hat{\sigma}_{t+h, h}$ with the standard Gaussian. If the
two match, $\hat{\sigma}_{t+h, h}$ is a good estimate of $\sigma_{t+h,
  h}$, the volatility of $r_{t+h, h}$.

For the Nordea Bank 15-minute returns, $h = 15\text{min}$, and a
reasonally good sampling frequency is 1/30sec, as confirmed by the
probability plot \ref{fig:nordea_bank_15min_z_prob} as well as by the
autocorrelation functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of
$(r_{t+h, h} / \sigma_{t+h, h})^2$ which are shown in
\ref{fig:nordea_15min_quotient_acf} and
\ref{fig:nordea_15min_quotient_squared_acf} respectively.
\begin{figure}[htb!]
  \centering
  \subfigure[ACF of $r_{t+h, h} / \sigma_{t+h, h}$]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    200]{../pics/nordea_15min_quotient_acf.pdf}
    \label{fig:nordea_15min_quotient_acf}
  }
  \subfigure[ACF of $(r_{t+h, h} / \sigma_{t+h, h})^2$]{
    \includegraphics[scale=0.4, clip=true, trim=95 236 118
    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
    \label{fig:nordea_15min_quotient_squared_acf}
  }
  \caption{\small \it Nordea 15min $r_t/\sigma_t$ and
    $(r_t/\sigma_t)^2$ ACF.}
\end{figure}

We see in figure \ref{fig:nordea_15min_quotient_acf} and
\ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
autocorrelation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
$(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
series is i.i.d Gaussian distributed. As a result, the problem of
modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
  h}$. Andersen and Bollerslev et al reported that, for the exchange
rates between Deutch mark, yen and dollar, $\ln \sigma_{t+h, h}$ is
gaussian distributed \cite{Andersen03}. This is, however, not the case
for our series in question. In fact, in our case, $\ln \sigma_{t+h,
  h}$ is right skewed (skewness 0.3342) and leptokurtic (kurtosis
6.1006). See figure \ref{fig:nordea_15min_logvol_prob}
\begin{figure}[htb!]
  \centering
  %\vspace{-15mm}
  \includegraphics[scale=0.5, clip=true, trim=80 223 107
  4]{../pics/nordea_15min_logvol_prob.pdf}
  \caption{\small \it Probability plot of Nordea 15min
    $\ln\sigma_{t+h, h}$ unconditional distribution}
  \label{fig:nordea_15min_logvol_prob}
\end{figure}

Moreover, the series of $\ln\sigma_{t+h, h}$ shows long-lasting and
periodic autocorrelations with an apparent period of 33 (see figure
\ref{fig:nordea_15min_logvol_acf}). This suggests the series may be
described by an integrated seasonal ARMA\footnote{ARMA ---
  Autoregressive Moving Average} model. But before we set out
to specify and estimate this model of the logvolatility
$\ln\sigma_{t+h, h}$, it is useful to first write down the model of
the returns $r_t$:
\begin{eqnarray}
   r_{t+h, h} &=& \ln p_{t+h} - \ln p_t \nonumber \\
   r_{t+h, h} &=& \mu + \sigma_{t+h, h} b_t \label{eq:SLV_spec}
\end{eqnarray}
where $p_t$ is the price of the asset at time $t$ and $b_t \sim N(0,
1)$. For simplicity, in the rest of this chapter we shall just write
$t$ for the subscript ``t+h, h'', since the time interval $h$ is fixed
for each time series and is a known constant. In the case of the
Nordea Bank 15-minute returns, h = 15 minutes.

As has just been said, $\ln \sigma_t$ for the series in question has
long-lasting auto-correlations (often referred to as ``long
memory''). This suggests simplifying the series by differencing
\cite{BoxJenkins94}:
\[
w_t = (1-B)(1-B^s)\ln\sigma_t
\]
where $B$ is the back-shift operator and $s=33$ is the seasonality.

The autocorrelation function of the differenced process $w_t$, as shown in
figure \ref{fig:nordea_15min_w_acf}, clearly points to a seasonal
moving-average model: There are only 4 non-zero autocorrelations in
the plot, located at lags 1, 32, 33, 34, respectively; furthermore,
the two at 32 and 34 are approximately equal. Thus we can write down
the model as
\begin{eqnarray}
  w_t &=& (1 - \theta B)(1 - \Theta B^s) y_t \label{eq:nordea_w}
\end{eqnarray}
where $\theta$ and $\Theta$ are parameters to be determined and $y_t$
is a noise process with constant variance $\sigma_y^2$ and mean
0. $y_t$ is often refered to as the residuals.
\begin{figure}[htb!]
  \centering
  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
    \includegraphics[scale=0.4, clip=true, trim=95 230 112
    235]{../pics/nordea_15min_logvol_acf.pdf}
    \label{fig:nordea_15min_logvol_acf}
  }
  \subfigure[ACF of $w_t$]{
    \includegraphics[scale=0.4, clip=true, trim=95 230 112
    235]{../pics/nordea_15min_w_acf.pdf}
    \label{fig:nordea_15min_w_acf}
  }
  \caption{\small \it Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$
    autocorrelations}
\end{figure}

The above seasonal moving average model has the following
autocovariance structure \cite{BoxJenkins94}:
\begin{eqnarray*}
  \gamma_0 &=& \sigma_y^2 (1 + \theta^2)(1 + \Theta^2) \\
  \gamma_1 &=& -\sigma_y^2\theta(1 + \Theta^2) \\
  \gamma_s &=& -\sigma_y^2\Theta(1 + \theta^2) \\
  \gamma_{s+1} &=& \gamma_{s-1}\;=\;\sigma_y^2\theta\Theta
\end{eqnarray*}
These equations together with the measured autocorrelations make
possible an initial estimate of the parameters $\theta$ and $\Theta$:
\begin{eqnarray*}
  {\rho_{s+1}/\rho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
    1 + \theta^2} \\
  {\rho_{s+1}/\rho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
    1 + \Theta^2} \\
\end{eqnarray*}
Substituting in the measured values shown in table
\ref{tab:nordea_15min_w_acf},
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    $\rho_1$ & $\rho_{s-1}$ & $\rho_s$ & $\rho_{s+1}$ \\
    \hline
    -0.4703 &  0.2053 & -0.4564 &  0.2212 \\
    \hline
  \end{tabular}
  \caption{\small \it Nordea 15min $w_t$ autocorrelations}
  \label{tab:nordea_15min_w_acf}
\end{table}
we get
\begin{eqnarray*}
  \theta &=& 0.6890 \\
  \Theta &=& 0.6378
\end{eqnarray*}
Among the two roots of each of the 2nd order equations in the above,
we have chosen the one in the range $(-1, 1)$ so as to ensure
invertibility of the model \cite{BoxJenkins94}.

With an estimate of $\theta$ and $\Theta$, one can then infer the
noise process i.e. the residuals $y_t$:
\begin{equation}
  \label{eq:infer_y}
  y_t = w_t + \theta y_{t-1} + \Theta y_{t-s} - \theta \Theta y_{t-s-1}
\end{equation}
where we substitute $y_t\;(t \leq 0)$ with their unconditional
expectation 0.
% By inverting the seasonal moving average model of $w_t$ we can then
% infer the noise process $y_t$ \footnotemark, i.e. the residuals:
% \begin{eqnarray*}
%   y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
%   &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
%   w_t \\
%   &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
%   w_{t-k}
% \end{eqnarray*}
% where $\bot(\cdot)$ denotes the floor function, i.e. the largest
% integer smaller than its argument.
% \footnotetext{
%   The autoregressive process resulting from the inversion is inifinite
%   in extent. Hence one must choose a cut-off value for k. For the
%   present case of Nordea 15min log-volatility, I choose the value
%   $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
%   $k_{\top}$, is around 550. To avoid data reduction of this amount,
%   we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
%   corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $\E(w_t) = \E(y_t
%   - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
% }

In order to forecst the $w_t$ process, and hence the return process
itself, we must also know the distribution of $y_t$. Moreover, to
properly estimate the parameters of the model in the sense of maximum
likelihood, we are also in need of the distribution of $y_t$.

If $y_t$ is indeed i.i.d random variables as we assumed, we only need
to work out its unconditional distribution. This is shown in figure
\ref{fig:nordea_15min_y_qq} and \ref{fig:nordea_15min_y_acf}.
\begin{figure}[htb!]
  \centering
  \subfigure[QQ-plot of $y_t$]{
    \includegraphics[scale=0.4, clip=true, trim=95 230 112
    210]{../pics/nordea_15min_y_qq.pdf}
    \label{fig:nordea_15min_y_qq}
  }
  \subfigure[ACF of $y_t$]{
    \includegraphics[scale=0.4, clip=true, trim=95 230 112
    210]{../pics/nordea_15min_y_acf.pdf}
    \label{fig:nordea_15min_y_acf}
  }
  \caption{\small \it Nordea 15min $y_t$ QQ-plot and autocorrelations}
\end{figure}
In addition, the measured moments of $y_t$ are shown in table
\ref{tab:nordea_15min_y_moments}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    mean & variance & skewness & kurtosis \\
    \hline
    0.0012 & 0.0935 & 0.2988 & 6.8691 \\
    \hline
  \end{tabular}
  \caption{\small \it Nordea 15min $y_t$ moments}
  \label{tab:nordea_15min_y_moments}
\end{table}

If we ignore the weak autocorrelations in this crude estimate of
$y_t$ and assume $y_t$ to be i.i.d random variables governed
by a skewed and leptokertic distribution, the parameters of this
distribution together with $\theta$ and $\Theta$ can be determined via
maximum likelihood estimate (MLE).

It turns out that $y_t$ can be well described by a Johnson Su
distribution \cite{Shang2004}:
\[
  y_t = \xi + \lambda\sinh{z_t - \gamma \over \delta}
\]
where $\gamma, \delta, \lambda, \xi$ are parameters to be determined
and $z_t \sim N(0, 1)$. The goodness of fitting is demonstrated in
figure \ref{fig:nordea_15min_y_js_fit} by the
empirical cummulative distribution function in comparison to the
theoretical one.
\begin{figure}[htb!]
  %\vspace{-18mm}
  \centering
    \includegraphics[scale=0.5, clip=true, trim=92 229 116
    133]{../pics/nordea_15min_y_js_fit.pdf}
    \caption{\small \it Nordea 15min $y_t$ fitted to a Johnson Su
      distribution. Horizontal: values of $y_t$, denoted x; Vertical:
      $\ln\left(P(y_t < x)\right)$.}
    \label{fig:nordea_15min_y_js_fit}
\end{figure}

The first 4 moments of the Johnson Su distribution are expressible
in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}:
\begin{eqnarray*}
  w &=& \exp{1 \over \delta^2} \\
  \Omega &=& {\gamma \over \delta} \\
  \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
  \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
    1)\right]^{1/2} \\
  \text{skewness}(y) &=& {
    \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
    \over
    (w\cosh 2\Omega + 1)^{3/2}} \\
  \text{kurtosis}(y) &=& {
    w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
    + 3(2w+1) \over
    2(w\cosh 2\Omega + 1)^2 }
\end{eqnarray*}
By matching the theoretical expressions of the moments with their
measured values, and taking help from published tables
\cite{Johnson1965}, one can solve for the parameters $\gamma, \delta,
\lambda, \xi$.

Under the assumption of i.i.d Johnson Su distributed residuals, the
log-likelihood function of the parameters $\gamma, \delta$
conditional on the sample $w_t$ can be written as
\[
L(\theta, \Theta) = -{1 \over 2}\sum_{t=1}^n z_t^2 + n \ln{\delta
  \over \lambda \sqrt{2\pi}} - {1 \over 2}\sum_{t=1}^n \ln\left[
  1 + \left({y_t - \xi \over \lambda}\right)^2
\right]
\]
where $y_t$ are inferred from $w_t$ using eq.\ref{eq:infer_y}
and $z_t$ from $y_t$ using
\[
z_t = \delta \sinh^{-1}{y - \xi \over \lambda} + \gamma
\]
Note that $\gamma, \delta, \lambda, \xi$ are not really free
parameters but rather are implied by $\theta$ and $\Theta$: Once the
latter have been chosen and the corresponding $y_t$ inferred, the
former are determined by the moments of $y_t$.

The eventual MLE is done in Matlab with the ``active set''
algorithm. The initial as well as the final estimation results are
listed in table \ref{tab:nordea_15min_js_param}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
    \hline
    initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
    0.6378 \\
    \hline
    MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
    \\
    \hline
  \end{tabular}
  \caption{\small \it Nordea 15min estimation results}
  \label{tab:nordea_15min_js_param}
\end{table}

To verify that the log-volatility can in general be described by
ARIMA(p, d, q) models, we apply the above procedure to a number of
Swedish stocks whose data have been collected by the author with
home-developed software. In \S\ref{sec:nordea} and \S\ref{sec:volvo}
we model the log-volatilities of Nordea Bank 15-minute returns and
Volvo B 30-minute returns. Both return series span the period
2013/10/10 - 2014/04/04. They are computed using minute-by-minute
average prices, and their variances, i.e. volatilities squared, are
computed as the sum of squared higher-frequency returns, which are in
turn computed using second-by-second prices. Whenever a needed average
price cannot be computed due to the lack of price observations in the
relevent time interval, the value of the previous average price is
assumed.

After modeling, we perform 1-step-ahead forecast using the estimated
models, and compare the predicted log-volatilities with the actual
data.

\subsection{Nordea Bank Log-Volatility}\label{sec:nordea}
In this section we model the log-volatility of {\it Nordea Bank}
15-minute returns during the period 2013/10/10 - 2014/04/04, which
contains 3879 data points. The squared volatilities, i.e. the
variances, are computed as the sum of squared 30-second returns. Among
the 3879 data points, we use the first 80\% for model estimation and
the last 20\%, i.e. 776 data points for forecast verification.

This Nordea series has a auto-correlation structure similar to its
predecessor that has been studied in previous sections. We show it in
figure \ref{fig:nordea2_w_acf}. The reader may want to compare it with
figure \ref{fig:nordea_15min_w_acf}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=95 230 112
  235]{../pics/nordea2_w_acf.pdf}
  \caption{\small \it Auto-correlation function of $w_t$}
  \label{fig:nordea2_w_acf}
\end{figure}
The significant auto-correlations and their respective lags are listed
in table \ref{tab:nordea2_w_acf}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    lag & 1 & 32 & 33 & 34 \\
    \hline
    ACF &   -0.4805 & 0.2861 &  -0.5034 & 0.2093 \\
    \hline
  \end{tabular}
  \caption{\small \it Significant auto-correlations of $w_t$}
  \label{tab:nordea2_w_acf}
\end{table}

Apparently this series also has the seasonality $s = 33$, but judging
from the relatively big differences between auto-correlations at lags
32 and 34, we don't expect the moving average process to be in the
multiplicative form of equation \ref{eq:nordea_w}. Instead, we model
$\ln\sigma_t$ has
\begin{eqnarray}
  (1 - B)(1 - B^s) \ln\sigma_t &=& (1 - \theta_1 B - \theta_s B^s -
  \theta_{s+1} B^{s+1} ) y_t \label{eq:nordea2_v}
\end{eqnarray}
The parameters are etimated to be
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    parameter & $\theta_1$ & $\theta_{33}$ & $\theta_{34}$ & $\var(y_t)$ \\
    \hline
    value & 0.8339 & 0.8448 & -0.7067 & 0.10071 \\
    \hline
  \end{tabular}
  \caption{\small \it: Parameter values of $\ln \sigma_t$}
\end{table}

The residuals $y_t$ do not have a Gaussian distribution, as confirmed
by the probability plot shown on the left of figure
\ref{fig:nordea2_y}. The non-normality, however, cannot be accounted
by the sine hypobolic transformation of Johnson Su, because of the
negative skewness of -1.2669 found in table
\ref{tab:nordea2_y_mmt}. How the non-normality can be accomodated is
quite another discussion and we do not pursue it here.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=20 220 30
  115]{../pics/nordea2_y.pdf}
  \caption{\small \it The residuals' distribution and auto-correlations. Left:
    cummulative distribution function (CDF) of the residuals $y_t$; Right:
    auto-correlation function of $y_t$.}
  \label{fig:nordea2_y}
\end{figure}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Moment & mean & variance & skewness & kurtosis \\
    \hline
    value & -0.0022 & 0.1007 & -1.2669 & 15.5965 \\
    \hline
  \end{tabular}
  \caption{\small \it: Moments of the residuals $y_t$}
  \label{tab:nordea2_y_mmt}
\end{table}

The auto-correlations of the residuals $y_t$ are insignificant, as
shown on the right of figure \ref{fig:nordea2_y}. Though the Ljung-Box Q
test rejects the null hypothesis of no-autocorrelation in the
residuals, the weak auto-correlations, if any at all,  are not worth
further complicating the model.

Forecast of $\ln\sigma_t$ is done by reverting equation
\ref{eq:nordea2_v} and write
\begin{eqnarray*}
  \ln\sigma_t &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
  \ln\sigma_{t-s-1} + (1 - \theta_1 B - \theta_s B^s + \theta_{s+1}
  B^{s+1} ) y_t \\
  \E(\ln\sigma_t | \mathcal{F}_{t-1}) &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
  \ln\sigma_{t-s-1} - \theta_1 y_{t-1} - \theta_s y_{t-s} +
  \theta_{s+1} y_{t-s-1}
\end{eqnarray*}

Thus the forecast, namely $\E(\ln\sigma_t | \mathcal{F}_{t-1})$, is
found in terms of previous values in the same series and of previous
residuals.

To compare the performance, we also estimate a GARCH model with
Gaussian conditional distribution for the Nordea Bank 15-minute
return. The procedure is standard and we only list the results here
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_5$ & $\beta_1$ \\
    \hline
    Value & $7.3686 \times 10^{-7}$ & 0.154131 & 0.0360 & 0.491832 \\
    \hline
  \end{tabular}
  \caption{\small \it Parameters of GARCH(1, 5) model of Nordea
    Bank 15-minute returns}
  \label{tab:nordea2_garch}
\end{table}

Forecasting using both models, we get two series of one-step-ahead
forecast, $\ln \sigma^{\text{ARIMA}}_t$ and $\ln \sigma^G_t$,
respectively. We denote the empirical log-volatilities as $\ln
\hat{\sigma}_t$. Then the accuracy of the two models can be compared
by comparing $\ln \sigma^{\text{ARIMA}}_t - \ln \hat{\sigma}_t$ and
$\ln \sigma^G_t - \ln \hat{\sigma}_t$. As a reference, we also include
in the comparison the naive approach of taking the sample mean of
$\ln \sigma_t$ as the forecast.

Table \ref{tab:slv_garch_comparison} shows the standard deviation of
$\ln \sigma^F_t - \ln \hat{\sigma}_t$ for the 3 kinds of
forecast. Here we use $\ln \sigma^F_t$ to denote the forecast of $\ln
\sigma_t$. Clear the ARIMA model has the smallest standard deviation,
the sample mean has the largest, and the GARCH model sits in between.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & ARIMA & GARCH & Sample mean \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ &
    0.2876  &  0.3096 &   0.3349 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:slv_garch_comparison}
\end{table}
Even more informative about the quality of forecast is the
distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$, which we show
in figure \ref{fig:slv_garch_comparison}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=0 292 0
  255]{../pics/slv_garch_comparison.pdf}
  \caption{\small \it Comparison of the forecast from ARIMA, GARCH,
    and sample mean. Blue: ARIMA forecast; Green: GARCH forecast; Red:
    Sample mean forecast.}
  \label{fig:slv_garch_comparison}
\end{figure}
The left plot in the figure shows the situation of under-estimates,
i.e. $\ln \sigma^F_t - \ln \hat{\sigma}_t < 0$. We see that, to the
far left, the blue line is below the green, which is in turn below the
red, meaning the ARIMA forecast has the lowest probability of
under-estimate. The GARCH forecast comes second and the sample
mean forecast under-estimates most.

The same ranking of forecast quality is also seen on the right plot of
figure \ref{fig:slv_garch_comparison}, which shows the complementary
distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$ for postive
values, i.e. over-estimates: The ARIMA forecast has the lowest
probability of over-estimate; GARCH comes second and the sample mean
is most likely to over-estimate.

As another measure of the quality of forecast, the fraction of
``good'' forecasts that differ no more than a given percentage from
the measured value is worth a few words. Table
\ref{tab:nordea_good_percentage} compares this fraction of good
forecasts from the 3 alternatives with respect to different
requirements of goodness.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    ARIMA & GARCH & Sample Mean \\
    \hline
    1\% & 0.1920 & 0.2010 & 0.1843 \\
    \hline
    5\% & 0.7887 & 0.7436 & 0.7229 \\
    \hline
    10\% & 0.9665 & 0.9601 & 0.9369 \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:nordea_good_percentage}
\end{table}
Clearly, although ARIMA and GARCH rank differently with respect to
different requirements of goodness, they both out-perform the sample
mean forecast. So the efforts of building models and estimating them
have paid off.
\subsection{Volvo B Log-Volatility}\label{sec:volvo}
In this section we model the log-volatility of {\it Volvo B} 30-minute
returns during the period 2013/10/10 - 2014/04/04. This series
contains 1884 log-volatilities computed using 2-minute returns in each
30-minute interval. Among them we use the first 1507 for model
estimation and the last 377 for forecast and model verification.

The left plot of figure \ref{fig:volvo_inno_and_lv_acf} shows the
distribution of $(r_t - \E(r_t))/\sigma_t$. We observe in the figure a
nice Gaussian variate, so we can be sure that the sum of squared
2-minute returns makes a good approximation to the variance of
30-minute returns in this particular case.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=0 256 0
    151]{../pics/volvo_inno_and_lv_acf.pdf}
    \label{fig:volvo_inno_and_lv_acf}
  \caption{\small \it Left: probability plot of $(r_t -
    \E(r_t))/\sigma_t$; Right: auto-correlations of $\ln \sigma_t$}
\end{figure}

Guided by the auto-correlations of $\ln\sigma_t$ shown in the right
plot of figure \ref{fig:volvo_inno_and_lv_acf} we find the following
model:
\begin{eqnarray}
  && (1 - B)(1 - B^s) \ln \sigma_t \nonumber \\
  &=& (1 - \theta_1 B - \theta_2B^2 -
  \theta_3B^3 - \theta_4B^4) (1 - \Theta B^s)
  y_t \label{eq:volvo_lv_model}
\end{eqnarray}
where $s = 16$ is the seasonality and is apparent from the
auto-correlations of $\ln\sigma_t$. Fitting this model to the
actual log-volatilities yields the parameter values listed in table
\ref{tab:volvo_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ &
    $\Theta$ & residual variance \\
    \hline
    Value & 0.7305 & 0.0575 & 0.0574 & 0.0346 & 0.8324 & 0.1340\\
    \hline
  \end{tabular}
  \caption{\small \it Volvo B log-volatility parameters}
  \label{tab:volvo_params}
\end{table}
Forecasting using the estimated model parameters gives the forecast
series $\ln \sigma^{\text{ARIMA}}_t$. To access the quality of the
forecast, we also estimate a GARCH model using the returns. The result
is a GARCH(1, 1) model:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$  & $\beta_1$ \\
    \hline
    Value & $3.125 \times 10^{-7}$ & 0.05 & 0.90 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH(1, 1) model of Volvo B 30-minute returns}
  \label{tab:volvo_garch}
\end{table}

The forecasts from ARIMA, GARCH, and the sample mean are compared
using the difference $\ln \sigma^F_t - \ln\hat{\sigma}_t$, where $\ln
\sigma^F_t$ stands for the forecast. The distributions of this
difference is plotted in figure \ref{fig:volvo_slv_garch_cmp}, and
the standard deviation of the distributions are listed in table
\ref{tab:volvo_slv_garch_cmp}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & ARIMA & GARCH & Sample mean \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3261 &
    0.4250 & 0.3708 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:volvo_slv_garch_cmp}
\end{table}

\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=10 282 0
    243]{../pics/volvo_slv_garch_cmp.pdf}
    \label{fig:volvo_slv_garch_cmp}
  \caption{\small \it Comparison of the forecast from ARIMA, GARCH,
    and sample mean. Blue: ARIMA forecast; Green: GARCH forecast; Red:
    Sample mean forecast.}
\end{figure}
Figure \ref{fig:volvo_slv_garch_cmp} shows, as in the previous case of
Nordea Bank 15-minute returns, the ARIMA model performs the best, GARCH
the second, and the sample mean the worst. However, when it comes to
over-estimates, the sample mean appears to be the best estimator,
while ARIMA excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
the sample for estimation (1507 data points) and over the sample for
comparison (377 data points) reveals that the first sample has mean
-6.3127 while the second has -6.2072. This increment explains the low
probability of over-estimation when using the first sample mean as
forecast.

Table \ref{tab:volvo_good_percentage} compares the fraction of
``good'' estimates as measured by ${|\ln\sigma^F_t -
  \ln\hat{\sigma}_t| \over |\ln \hat{\sigma}_t|}$ being less than 1\%,
5\% and 10\%.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    ARIMA & GARCH & Sample Mean \\
    \hline
    1\% & 0.2175 & 0.1220 & 0.1406 \\
    \hline
    5\% & 0.7162 & 0.6154 & 0.6631 \\
    \hline
    10\% & 0.9231 & 0.8806 & 0.8992 \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:volvo_good_percentage}
\end{table}
We see from the table that the ARIMA model consistently excels over
the other two alternatives. Once again, it confirms the validity and
good performance of an ARIMA log-volatility model.

% \subsection{Ericsson B Log-Volatility}\label{sec:ericsson}
% The probability density function of the residuals $y_t$ that result
% from the MLE estimate is shown in figure
% \ref{fig:nordea_15min_y1_hist} together with its counterpart from the
% initial estimate.
% \subsection{45min Returns}
\section{Unconditional Distribution Functions of ARIMA Log-Volatility
  Models}\label{sec:XieCalc}
In this section we study the unconditional distribution function of
the Stochastic Log-Volatility model specified as equation
\ref{eq:SLV_spec}. As is discussed in \S\ref{sec:SLV_model},
$\ln\sigma_t$ can be described by an ARMA or ARIMA \footnote{ARIMA ---
  Integrated Autoregressive Moving Average} model, possibly with
seasonal components. Here we note that all these models can be
re-written as a moving average model, which is infinite in extent if
autoregressive components are present:
\begin{eqnarray*}
  \ln \sigma_t &=& y_t + \sum_{n=1}^\infty c_n y_{t-n} + \text{Const.}
\end{eqnarray*}
Since the $y_t$ are independent and identically distributed,
\begin{eqnarray*}
  y_t + \sum_{n=1}^\infty c_n y_{t-n}  
\end{eqnarray*}
has Gaussian distribution by the central limit theorem, on condition
that $y_t$ for all $t$ have finite second moment --- this is what we
assume in the rest of this section. It follows from the above equation
that the unconditional distribution of $\ln \sigma_t$ is the same as the
distribution of $\bar{v} + v$ where $v \sim N(0, \sigma)$ and
$\bar{v}$, $\sigma$ are constants. Now we can state that the
unconditional distribution of the returns $r_t$
\begin{eqnarray*}
  r_t &=& \mu + \sigma_t b_t\\
  &=& \mu + \exp\left(
    y_t + \sum_{n=1}^\infty c_n y_{t-n} + \text{Const.}
  \right)
\end{eqnarray*}
 is the same as
\begin{equation}  \label{eq:UnconditionalPdf}
  \begin{aligned}
    r &= \mu + e^{\bar{v} + v} b \\
  \end{aligned}
\end{equation}
where $b \sim N(0, 1)$. For convenience, let $r' = e^v b$

In \S\ref{sec:SLV_Symmetric} we first study the model in the
relatively simple case when $v$ and $b$ are uncorrelated and $\mu
= 0$. If this simplified version proves inadequate, one may resort to
the general model studied in \S\ref{sec:SLV_Asymmetric}.

\subsection{The Simplified model}\label{sec:SLV_Symmetric}
In the following we derive the unconditional PDF of $r'$, denoted
$f_{r'}(x)$. Then the PDF of $r$ is $e^{-\bar{v}}f_{r'}(e^{-\bar{v}}x)$.
\begin{eqnarray*}
  P(r' < x) &=& P(b < xe^{-v}) \\
  f_{r'}(x) &=& f_b(xe^{-v}) e^{-v}
\end{eqnarray*}
Averaging over all $v$, we get
\begin{equation}\label{eq:UncondPDFSymmetric}
  \begin{aligned}
    f_{r'}(x) =& \int_{-\infty}^{\infty} dv (2\pi\sigma^2)^{-1/2}
    e^{-v^2/2\sigma^2}(2\pi)^{-1/2} \exp(-x^2e^{-2v}/2) e^{-v} \\
    =& {1 \over 2\pi\sigma} \int_{-\infty}^{\infty} dv
    \exp\left(
      -{1 \over 2\sigma^2} v^2 - v -{1 \over 2} x^2 e^{-2v}
    \right)
  \end{aligned}
  \end{equation}
The last part of the integrand, $e^{-x^2 e^{-2v} / 2}$, is plotted in
figure \ref{fig:DoubleExp}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=85 252 100
  231]{../pics/DoubleExp.pdf}
  \caption{\small \it Plot of $\exp(-{1 \over 2} x^2 e^{-2v})$}
  \label{fig:DoubleExp}
\end{figure}
Therefore we make the following approximation:
\[
\exp\left(-{1 \over 2} x^2 e^{-2v}\right) \approx \left\{
  \begin{array}{lr}
    0 & \text{if } v < \ln|x| -{1 \over 2} \ln(2\ln 2) \\
    1 & \text{otherwise}
  \end{array}
\right.
\]
Here we note that $\exp\left(-{1 \over 2} x^2 e^{-2v}\right) = 1/2$ at
$v = \ln|x| -{1 \over 2} \ln(2\ln 2)$.

With this approximation we have
\begin{eqnarray*}
  f_{r'}(x) &=& {1\over C}{1 \over 2\pi\sigma} \int_{\ln(|x|/\sqrt{\ln
      4})}^{\infty} dv
  \exp\left(-{1 \over 2\sigma^2} v^2 - v\right) \\
  &=& {1\over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}} \text{erfc} \left(
    {1 \over \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma
      \over \sqrt{2}}
  \right)
\end{eqnarray*}
where $1/C$ has been added for the purpose of normalization.

At large $x$, we may use the asymptotic expansion of $\mathrm{erfc}$
to write
\begin{eqnarray*}
  Cf_{r'}(x) &=& {e^{\sigma^2 / 2} \over \sqrt{8\pi}} \mathrm{erfc}(\xi) \\
  &=& {e^{\sigma^2 / 2} \over \sqrt{8 \pi}}
  \frac{e^{-\xi^2}}{\xi\sqrt{\pi}}\left[
    1 +
    \sum_{n=1}^N (-1)^n \frac{(2n-1)!!}{(2\xi^2)^n} \right] +
  O(\xi^{-2N-1} e^{-\xi^2})
\end{eqnarray*}
where $\xi = {1 \over \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} +
{\sigma \over \sqrt{2}}$. The slowest-decaying term is
\[
f_{r,0}(x) = {e^{\sigma^2 / 2} \over \sqrt{8\pi}}
\frac{e^{-\xi^2}}{\xi\sqrt{\pi}}
\]
Let $\zeta = {|x| \over \sqrt{\ln 4}}$. With a bit manipulation one
obtains
\begin{equation*}
  f_{r,0}(x) = {1 \over \pi \sqrt{8}}{1 \over
    \left(\ln\zeta/\sigma\sqrt{2} +
      \sigma/\sqrt{2}\right)\zeta\zeta^{\ln \zeta / 2\sigma^2}
  }
\end{equation*}
From the last equation one can see that, at any neighborhood of large
$x$, $f_{r'}(x)$ may be approximated by $C/|x|^\alpha$, i.e. a power law.

In the following we work out the normalization constant C.
\begin{eqnarray*}
  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {1 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} \text{erfc} \left({1 \over
      \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma \over
      \sqrt{2}} \right) dx\\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{0}^{\infty} \text{erfc} \left({1 \over
      \sqrt{2}\sigma} \ln{x \over \sqrt{\ln 4}} + {\sigma \over
      \sqrt{2}} \right) dx\\
\end{eqnarray*}
Let
\begin{eqnarray*}
  a &=& {1 \over \sigma \sqrt 2} \\
  b &=& -{1 \over 2 \sigma \sqrt 2}\ln\ln 4 + {\sigma \over \sqrt 2} \\
  y &=& a\ln|x|+b \\
\end{eqnarray*}
Then
\begin{eqnarray*}
  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {e^{(y-b)/a} \over a}
  \text{erfc}(y) \\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \left.e^{(y-b)/a}
    \text{erfc}(y)\right|_{y=-\infty}^{\infty}
  + {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {2 \over \sqrt \pi}
  e^{(y-b)/a} e^{-y^2} \\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}  2 e^{\ln\ln 4/2
    - \sigma^2/2}\\
  C &=& \sqrt{2 \ln 4\over \pi} \\
  &\approx& 0.9394
\end{eqnarray*}
Thus we can write in summary:
\begin{eqnarray*}
  f_r(x) &=& e^{-\bar{v}} f_{r'}(e^{-\bar{v}} x) \\
    &=& {e^{-\bar{v}} \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}
    \text{erfc} \left({1 \over \sqrt{2}\sigma} \ln{|e^{-\bar{v}}x| \over \sqrt{\ln
          4}} + {\sigma \over \sqrt{2}}
    \right) \\
\end{eqnarray*}
Using the same technique for integration as for normalization, the
cummulative distribution function of $r$ is found to be $F(x)$,
which is the following:
\begin{enumerate}
\item if $x < 0$
  \begin{eqnarray*}
    F(x) &=& {\sqrt{\ln 4} \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}
    \left[
      {e^{-\bar{v}}x \over \sqrt{\ln 4}}\text{erfc}\left(
        {1 \over \sigma \sqrt 2} \ln{-e^{-\bar{v}}x \over \sqrt{\ln
            4}} + {\sigma \over \sqrt 2}
      \right) \right.\\
      && \left. + e^{-\sigma^2 / 2} \text{erfc}\left(
        {1 \over \sigma \sqrt 2} \ln{-e^{-\bar{v}}x \over \sqrt{\ln 4}}
      \right)
    \right]
  \end{eqnarray*}
\item if $x \geq 0$
  \begin{eqnarray*}
    F(x) &=& \frac{1}{2} + {\sqrt{\ln 4} \over C}{e^{\sigma^2 / 2} \over
      \sqrt{8\pi}} \left[
      {e^{-\bar{v}}x \over \sqrt{\ln 4}} \text{erfc}\left(
        {1 \over \sigma \sqrt 2} \ln{e^{-\bar{v}}x \over \sqrt{\ln
            4}} + {\sigma \over \sqrt 2}
      \right) \right. \\
      && \left. + e^{-\sigma^2/2} \text{erfc}\left(
        -{1 \over \sigma \sqrt 2} \ln{e^{-\bar{v}}x \over \sqrt{\ln 4}}
      \right)
    \right]
  \end{eqnarray*}
\end{enumerate}

To verify the validity of the model, we fit the above probability
density function to the de-meaned 30min returns of Volvo B
\footnote{By ``de-meaned returns'' we mean the quantity $r_t - 
  \mean{r_t}$, where $r_t$ are the measured returns and $\mean{r_t}$
  is the sample mean. The data set covers the transaction records of
  Volvo B on the OMX market (Stockholm) between 2013-10-10 and
  2014-03-12. The returns are computed using 1-minute mean prices.}
by means of maximum likelihood estimate using the MATLAB function
``mle''. Then for the parameters $\sigma$ and $\bar{v}$ we get
\begin{eqnarray*}
  \sigma &=& 0.6355 \\
  \bar{v} &=& -6.0308
\end{eqnarray*}
Then we plot $P(r' > x)$ of the model against its empirical
counterpart on a log-log scale, as shown in figure
\ref{fig:volvo_30min_ret}.
\begin{figure}[htb!]
  % \vspace{-15mm}
  \begin{center}
    \includegraphics[scale=0.4, clip=true, trim=98 231 116
    126]{../pics/volvo_30min_ret.pdf}
  \end{center}
  %\vspace{-5mm}
  \caption{\small \it{$P(r' > x)$ for $x > 0$. Blue: empirical
      probabilities. Red: probabilities predicted by the model.}}
  \label{fig:volvo_30min_ret}
\end{figure}

In general figure \ref{fig:volvo_30min_ret} shows a good fit, but a
closer look reveals that deviations are significant in the regions $r
\in (0, \sigma_r)$ and $r \in (2\sigma_r, 3\sigma_r)$, where
$\sigma_r$ stands for the empirical standard deviation of the
de-meaned returns. These observations are
shown in greater details in figure \ref{fig:volvo_30min_ret2}.
\begin{figure}[htb!]
  \centering
  \subfigure[$\ln(P(r > x))$ with $x \in (0, \sigma_r)$]{
    \includegraphics[scale=0.4, clip=true, trim=93 224 115
    125]{../pics/volvo_30min_ret_0-sigma.pdf}
  }
  \subfigure[$\ln(P(r > x))$ with $x \in (2\sigma_r, 3\sigma_r)$]{
    \includegraphics[scale=0.4, clip=true, trim=93 224 115
    125]{../pics/volvo_30min_ret_2sigma-3sigma.pdf}
  }
  \caption{\small \it Surviving probabilities. Blue: empirical values of $\ln(P(r
    > x))$. Red: Predicted values of $\ln(P(r > x))$.}
  \label{fig:volvo_30min_ret2}
\end{figure}

Moreover, the inefficiency of the model is also manifest in the
skewness of the data. For the Volvo 30min returns, the data has a
skewness of 0.2419, but our model is strictly symmetric, since
$x$ appears in equation \ref{eq:UncondPDFSymmetric} only as $x^2$.
Hence the above model needs to be improved to accommodate the non-zero
skewness as well as to account for the discrepancies shown in figure
\ref{fig:volvo_30min_ret2}. This is the subject of the next section.

\subsection{The General model}\label{sec:SLV_Asymmetric}
It has long been hypothesized in the liturature that skewness is the
result of price-volatility correlation (for example
\cite{Potters2003}). Therefore, the most apparent modification is to
allow $v$ and $b$ to be correlated. For convenience we decompose
$v \sim N(0, \sigma)$ as $v = \sigma a$ and assume
\begin{eqnarray*}
  \begin{pmatrix}
    a \\
    b
  \end{pmatrix} \sim N(0, \Sigma)
\end{eqnarray*}
with a covariance matrix
\begin{eqnarray*}
    \Sigma &=&
  \begin{pmatrix}
    1 & \psi \\
    \psi & 1
  \end{pmatrix}
\end{eqnarray*}
where $|\psi| < 1$. Then we rewrite equation \ref{eq:UnconditionalPdf}
as
\begin{equation}
  \label{eq:r_t}
  \begin{aligned}
    r &= \mu + e^{\bar{v}} r' \\
    r' &= e^{\sigma a} b \\
  \end{aligned}
\end{equation}
Then
\begin{eqnarray*}
  && P(r' < x)\\
  &=& P(b < e^{-a\sigma}x) \\
  &=& {1 \over 2\pi |\det(\Sigma)|^{1/2}}
  \int_{-\infty}^{\infty} da \int_{-\infty}^{e^{-a\sigma}x} db
  \exp\left[
    -{1 \over 2} (a, b) \Sigma^{-1}
    \begin{pmatrix}
      a \\
      b
    \end{pmatrix}
  \right] \\
\end{eqnarray*}
After some manipulations we get
\begin{equation}\label{eq:UncondCDFAsymmetric}
  \begin{aligned}
    & P(r' < x) \\  
    &= {1 \over 2\pi \sqrt{1 - \psi^2}} \int_{-\infty}^{\infty} da
    e^{-a^2/2} \int_{-\infty}^{e^{-a\sigma}x} db
    \exp\left[
      - {(b - a\psi)^2 \over 2(1 - \psi^2)}
    \right] \\
    &= {1 \over 2\sqrt{2\pi}} \int_{-\infty}^{\infty} e^{-a^2/2}
    \text{erfc}{a\psi - e^{-a\sigma} x \over \sqrt{2(1-\psi^2)}} da
  \end{aligned}
\end{equation}
Differentiating with respect to $x$ yields
\begin{eqnarray}
  f_{r'}(x) &=& {1 \over 2\pi \sqrt{1 - \psi^2}} \times \nonumber \\
  && \int_{-\infty}^\infty da \exp
  \left[- {
      a^2 + 2\sigma(1 - \psi^2)a - 2a\psi e^{-a\sigma} x + e^{-2a\sigma} x^2
      \over
      2(1 - \psi^2)
  }
  \right] \label{eq:UncondPDFAsymmetric} \\
  f_r(x) &=&  \td{r'}{r} f_{r'}[e^{-\bar{v}} (x - \mu)] \nonumber \\
  &=& e^{-\bar{v}} f_{r'}[e^{-\bar{v}}
  (x-\mu)] \label{eq:UncondPDFAsymmetric1}
\end{eqnarray}
Unlike \ref{eq:UncondPDFSymmetric} where a relatively simple
approximation can be found and thus leads to an analytic result of the
integral, no such approximation has been found by the
author for the integral \ref{eq:UncondPDFAsymmetric}. However, the
moment generating functions (MGF) of PDF \ref{eq:UncondPDFAsymmetric}
and \ref{eq:UncondPDFAsymmetric1} are easy enough to find and give the
moments about the origin in closed form. Then by matching the analytic
expressions of the moments with their statistical values from the
sample, the parameters $\sigma$, $\psi$ and $\bar{v}$ corresponding to the
sample can be obtained.

The MGF of \ref{eq:UncondPDFAsymmetric1} can be found as follows:
\begin{eqnarray*}
  && M_{r}(t) \\
  &=& \E(e^{tr}) \\
  &=& {1 \over 2\pi \sqrt{1 - \psi^2}} \int_{-\infty}^{\infty} da \exp \left(
    -{a^2 \over 2} - \sigma a
  \right) \int_{-\infty}^{\infty} dx \exp\left[
    t(e^{\bar{v}}x + \mu) - {
      (a\psi - e^{-a \sigma} x)^2 \over 2 (1 - \psi^2)
    }
  \right] \\
  &=& {1 \over \sqrt{2 \pi}} \int_{-\infty}^{\infty} da \exp\left[
    -{a^2 \over 2} + {1 \over 2} e^{2 a \sigma + 2\bar{v}} (1 - \psi^2) t^2 +
    a \psi e^{a \sigma + \bar{v}} t + \mu t
  \right]
\end{eqnarray*}
With the MGF, the first 4 moments of $r$ can be computed:
\begin{equation}
  \label{eq:AsymmetricMoments}
  \begin{aligned}
    \E(r) &= \mu +e^{\bar{v}+\frac{\sigma ^2}{2}} \sigma  \psi\\
    \E(r^2) &= \mu ^2+2 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu  \sigma  \psi
    +e^{2 \left(\bar{v}+\sigma ^2\right)} \left(1+4 \sigma ^2 \psi^2\right)\\
    \E(r^3) &= \mu ^3+3 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu ^2 \sigma \psi
    +9 e^{3 \bar{v}+\frac{9 \sigma ^2}{2}} \sigma \psi \left(1+3 \sigma ^2
      \psi ^2\right)
    +3e^{2 \left(\bar{v}+\sigma ^2\right)} \mu  \left(1+4 \sigma ^2 \psi
      ^2\right) \\
    \E(r^4) &= \mu ^4+4 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu ^3 \sigma  \psi
    +36 e^{3 \bar{v}+\frac{9 \sigma ^2}{2}} \mu  \sigma  \psi  \left(1+3
      \sigma ^2 \psi ^2\right)\\ 
    & +6 e^{2 \left(\bar{v}+\sigma ^2\right)} \mu ^2 \left(1+4 \sigma ^2 \psi
      ^2\right)
    +e^{4 \bar{v}+8 \sigma ^2} \left(3+96 \sigma ^2 \psi ^2+256 \sigma ^4
      \psi ^4\right)
  \end{aligned}
\end{equation}
These equations are rather complicated and directly solving them to
obtain the parameters $\mu$, $\sigma$, $\psi$ and $\bar{v}$ is
infeasible. However, in practice, $\E(r)$ is often very small --- so, to
get a rough estimate of the parameters, we may set $\mu = 0$ in the
above equations and, with a bit of manipulation, find the following
equation for $\psi\sigma$:
\begin{equation}
  \label{eq:moment1}
  {\E(r^3) \E(r^4)^{-3/8} \over \E(r^2)^{3/4}} =
  {9\sigma \psi (1 + 3\sigma^2\psi^2) \over (1 +
    4\sigma^2\psi^2)^{3/4}}
  {1
    \over
    (3 + 96\sigma^2\psi^2 + 256\sigma^4 \psi^4)^{3/8}
  }
\end{equation}
This equation can be solved numerically for a given sample to yield an
estimate for $\sigma\psi$, which in turn can be substituted in
equations \ref{eq:AsymmetricMoments}, where $\mu$ has been set to 0,
to give estimates for all the 4 parameters. These estimates then serve
as initial values in a numerical solution to
\ref{eq:AsymmetricMoments} where $\mu$ is kept as a free
variable. This full solution can now be used as the initial estimate
in an MLE procedure.

It is particularly important in our situation to have a good initial
estimate, because, as has been shown earlier, the PDF and CDF cannot
be obtained in closed forms and consequently have to be evaluated by
numerical integration. This is a rather costly procedure especially in
the context of MLE. So the computation of the moments under the
assumption of $\mu = 0$ is worthwhile.

Following the aforementioned procedure, i.e. computing the initial
estimate by matching the moments and then refining the estimate by
MLE, we obtain the parameter values for a number of return series,
including the Volvo B 30-minute returns described in
\S\ref{sec:SLV_Symmetric}. In the following, we list the parameter
values and show the comparison between the empirical distribution
functions and their counterparts from the model.

Clearly these figures show a fairly good match of the empirical
distribution functions to their model counterparts. So we are once
again assured of the validity of ARIMA log-volatility models. However,
we also see that the skewness of the returns of Volvo B and Ericsson B
have rather different values when computed from the sample and from
the model. This could be the consequence of the limited sample size or
deficiencies in the estimation procedure described above. We leave
these issues to future studies.

%  This indicates the parameter $\psi$, which is the
% correlation between the log-volatility and the return, may be more
% complicated than the simple constant that has been assumed so far. We
% leave this to later studies.

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    & $\psi$ & $\sigma$ & $\bar{v}$ & $\mu$ \\
    \hline
    Volvo B & -1.5747e-02 & 4.6929e-01 & -6.2304e+00 &
    -5.3040e-05 \\
    Nordea Bank & 1.7234e-02 & 4.2780e-01 & -6.3758e+00 & 7.5230e-05
    \\
    Ericsson B & 1.8486e-02 & 4.4905e-01 & -6.2834e+00 & -9.2817e-05 \\
    \hline
  \end{tabular}
  \caption{\small \it Parameter Values of Selected Assets'
    returns. The time series are 30-minute returns and run from
    2013/10/10 to 2014/03/12.}
  \label{tab:assets_params}
\end{table}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    & mean & std & skewness & kurtosis \\
    \hline
    \multirow{2}{*}{Volvo B} & -8.6766e-05 & 2.4500e-03 &
    -2.7217e-02 & 7.3394e+00 \\
    & -6.9281e-05 & 2.4538e-03 & -7.2677e-02 & 7.2472e+00 \\
    \hline
    \multirow{2}{*}{Nordea Bank} & 7.4821e-05 & 2.0500e-03 &
    6.9433e-02 & 6.7427e+00 \\
    & 8.8983e-05 & 2.0443e-03 & 6.7125e-02 & 6.2446e+00 \\
    \hline
    \multirow{2}{*}{Ericsson B} & -8.2606e-05 & 2.2939e-03 &
    2.6586e-01 & 7.6716e+00 \\
    & -7.5675e-05 & 2.2844e-03 & 7.8573e-02 & 6.7299e+00 \\
    \hline
  \end{tabular}
  \caption{\small \it Moments of Selected Assets' returns. For each return
    series, the 1st row contains the sample moments, while the 2nd
    constains those computed with MLE parameters. The time series are
    30-minute returns and run from 2013/10/10 to 2014/03/12.}
  \label{tab:assets_moments}
\end{table}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=50 233 63
  136]{../pics/Volvo_B_30m_returns.pdf}
  \caption{\small \it Volvo B 30min returns' unconditional distribution fit to
    the Stochastic Volatility model. The time series runs from
    2013/10/10 to 2014/03/12. Left: $P(r < x)$ with $x < 0$; Right:
    $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
  \label{fig:Volvo_B_30m_returns}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=65 241 72
  143]{../pics/Nordea_30m_returns.pdf}
  \caption{\small \it Nordea Bank 30min returns' unconditional
    distribution fit to the Stochastic Volatility model. The time
    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
  \label{fig:Nordea_30m_returns}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=21 241 28
  146]{../pics/Ericsson_30m_returns.pdf}
  \caption{\small \it Ericsson B 30min returns' unconditional
    distribution fit to the Stochastic Volatility model. The time
    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
  \label{fig:Ericsson_30m_returns}
\end{figure}

\subsection{Relation to Conditional Distribution Functions}
In the last section we have shown that the unconditional distribution of
the returns are skewed, which implies the variate in the
log-volatility $a$ is correlated to the variate in the return
$b$.

In the context of conditional distributions and forecast, this
correlation translates to the correlation between the residual of the
log-volatility, denoted $y_t$ in \S\ref{sec:nordea} and
\S\ref{sec:volvo}, and $b_t$. Now let us consider the forecast
function of an ARIMA model:
\begin{eqnarray*}
  \ln \sigma_t &=& y_t + \sum_{i=1}^P \phi_i \ln \sigma_{t-i} -
  \sum_{i=1}^Q \theta_i y_{t-i}
\end{eqnarray*}
where $\phi_i \leq 1$ if the model involves integration, and $\phi_i <
1$ otherwise. Comparing this equation with equation \ref{eq:r_t}, one
immediately realizes the PDF of $r_t$
\begin{eqnarray*}
  r_t &=& \mu + \sigma_t b_t
\end{eqnarray*}
is given by equation \ref{eq:UncondPDFAsymmetric1} and its moments
given by the equations \ref{eq:AsymmetricMoments} if one makes the
following substitutions:
\begin{eqnarray*}
  \sigma &\to& \text{std}(y_t) \\
  \bar{v} &\to& \sum_{i=1}^P \phi_i \ln \sigma_{t-i} - \sum_{i=1}^Q
  \theta_i y_{t-i} \\
  \psi &\to& \text{corr}(y_t, b_t)
\end{eqnarray*}

For the Nordea series considered in \S\ref{sec:nordea},
$\text{corr}(y_t, b_t)$ is found to be $3.56 \times 10^{-2}$; while
for the Volvo series considered in \S\ref{sec:volvo},
$\text{corr}(y_t, b_t)$ is found to be $-7.3 \times 10^{-3}$. These
values are comparable to those in table \ref{tab:assets_params}, where
$\psi$ for the Nordea series is $1.7234 \times 10^{-2}$ and $-1.5747
\times 10^{-2}$ for the Volvo series. The similarity in these values
provides some evidence about the validity of the model.

It is certainly impossible to compare a predicted conditional
distribution with observations, but conditional distributions are of
great interest in the context of risk management and derivative
pricing, so we point out how they may be calculated for ARIMA
log-volatility models.

### 2014-04-30 thesis/thesis.tex
1st version that has been handed to Sven. Gaussian Wishart Matrices are normized by 1/T.

--- 2013-11-19 thesis/thesis.tex
@@
-\documentclass{book}
+\documentclass{report}
 \usepackage{graphicx}
 \usepackage{subfigure}
+\usepackage{multirow}
+\usepackage{wrapfig}
 \usepackage{amssymb}
+\usepackage{amsfonts, amsmath}
+\usepackage{mathrsfs}
 \usepackage{enumerate}
@@
 
-\title{Mathematical Aspects of the Capital Market}
+% \title{Stocks' Cross-Correlations}
+% \subtitle{}
+
 \author{Xie Xiaolei}
@@
 
-\maketitle
+\begin{titlepage}
+\begin{center}
+
+% Upper part of the page. The '~' is needed because \\
+% only works if a paragraph has started.
+\includegraphics[width=0.5\textwidth]{../pics/lund_uni-logo_s}~\\[1cm]
+
+% \textsc{\LARGE Lund University}\\[1.5cm]
+
+\textsc{\Large Master's Thesis Project}\\[0.5cm]
+
+% Title
+%\HRule \\[0.4cm]
+{ \huge \bfseries Returns Models and Their Implications
+  on Cross-Correlation Matrix \\[0.4cm] }
+
+%\HRule \\[1.5cm]
+
+% Author and supervisor
+\begin{minipage}{0.4\textwidth}
+\begin{flushleft} \large
+\emph{Author:}\\
+\textsc{Xie} Xiaolei
+\end{flushleft}
+\end{minipage}
+\begin{minipage}{0.4\textwidth}
+\begin{flushright} \large
+\emph{Supervisor:} \\
+Prof. Sven \textsc{\AA berg}
+\end{flushright}
+\end{minipage}
+
+\vfill
+
+% Bottom of the page
+{\large \today}
+
+\end{center}
+\end{titlepage}
+% \maketitle
 \tableofcontents
@@
 \chapter{Introduction}
-$\mathbb{E}(x)$
+The mathematics of the financial market has always been a topic that
+arouses interest and imagination, and with no doubt, has been
+studied from many aspects and in many different ways.
 
-\section{Stylized facts}
+Central to all these studies are the concepts of probability
+distributions and correlations. The values of stocks, futures,
+options, etc. are stochastic in nature and are governed by the laws of
+stochastic processes, which are expressed in terms of probability
+distributions and correlations.
 
-\section{GARCH models}
-The autocorrelation function of $\epsilon_t^2$ is given by
-(c.f. \cite{Bollerslev87})
-$$
-\rho_n = \sum_{i=1}^{max(p,q)} (\alpha_i + \beta_i) \rho_{n-i}
-\;n > p
-$$
-where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
-zeros. From these equations, it is clear that the partial
-autocorrelation function cuts off at $max(p, q)$.
+Meanwhile, the observables of the market are prices, volumes,
+turn-over (the amount of money paid in a trade), names of the brokers,
+and time of the trades. These quantities don't make much sense
+by themselves but do reveal the probabilistic dynamics of the market
+when put together and turned into statistics.
 
-\chapter{Case Study of Stocks}
-\section{Nordea Bank}
-\begin{figure}[ht]
+Here by dynamics we mean how the value of an asset is affected by its
+own history as well as by the histories and current values of other
+assets in the market. The influence from the asset's own history is
+termed autocorrelations, i.e. correlations in time, while
+the influence from other assets are termed cross correlations.
+
+Mathematical models that describe the aforementioned observables in
+terms of probability distributions, autocorrelations and cross
+correlations give predictive power and help determine a fair price of
+a given asset. Hence they are a central topic of the mathematics of
+the financial market.
+
+Once a model has been tentatively constructed, it must be
+verified. The verification again relies on the statistics of the
+observables. Over the years, a number of phenomena have been
+consistently reported and become generally accepted. These are referred
+to as stylized facts, which are detailed in \S
+\ref{sec:StylizedFacts} and constitute a standardized test suit for
+model verification.
+
+% Judged by the stylized facts, each of the models proposed in the
+% literature has its own advantages and disadvantages. The most
+% prominent and influential among them are arguably the GARCH
+% models. Their first and simplest variant was proposed by Bollerslev in
+% 1990 \cite{Bollerslev86} and has been developed tremendously since
+% then.
+
+% An apparent advantage of the GARCH models is their simplicity. They
+% are linear models, where the conditional variance is modeled as a
+% linear combination of past variances and past innovations squared. As
+% a result, it is relatively easy to fit a model to a given set of data,
+% i.e. to estimate the model parameters with respect to the
+% data. Because of the linearity, the estimation of model parameters
+% often give consistent results, regardless of the choice of initial
+% parameter values or the choice of pre-sample data.
+
+% In contrast, many of the other types of models are often specified in
+% a complicated, although more advanced way, for example, by stochastic
+% differential equations. This kind of specifications certainly have the
+% advantage of directly linking to the theory of stochastic processes,
+% but nonetheless are difficult to fit to data, one of the most tricky
+% problems being the high sensitivity to the choice of initial parameter
+% values.
+
+% Another advantage of GARCH models is that their tail behavior is
+% better understood than that of their competitors. The simplest variant
+% of GARCH, namely the GARCH(1,1) models are proven to have regularly
+% varying tails, or in common terms, power-law tails
+% \cite{mikosch2000}. In particular, if the $\alpha_1$ and $\beta_1$
+% parameters of the model sum up to 1, the return series generated by
+% the model has a power-law tail with power exponent 2, and consequently
+% the squared returns have a tail with power exponent 1 -- by the
+% central limit theorem the sum of the squared returns will converge to
+% the Cauchy distribution, a stable distribution with L\'evy index 1. We
+% explore this fact in more details in chapter
+% \ref{chp:CrossCorrelationFat} where the cross-correlation matrix is
+% studied.
+
+% The GARCH model with $\alpha_1 + \beta_1 = 1$ is more often referred
+% to as an IGARCH model, aka. an integrated GARCH model. It is of
+% particular interest because, very often, when fitting a GARCH(1,1)
+% model to real financial data, one finds $\alpha_1 + \beta_1$ extremely
+% close to 1, i.e. at values like 0.99, 0.995, etc.
+
+% A third advantage of GARCH models is their flexibility. Namely one can
+% choose any appropriate distribution to model the innovations. While
+% the Gaussian and the student's t distributions are the most common,
+% the use of other distributions such as asymmetric student's t and
+% Johnson Su has also been studied in the literature (see
+% \cite{Simonato2012}).
+
+% However, GARCH models have their disadvantages too. The original
+% GARCH model does not account for skewness in the distribution of
+% the returns. As a result, the skewness, often attributed to
+% price-volatility correlations, have to be added ``by hand''. This kind
+% of treatment fixes the model on the one hand, but introduces some
+% arbitrariness on the other. Put in another way, skewness does not
+% follow naturally from the structure of the model but rather is added
+% explicitly.
+
+% Moreover, fitting a GARCH model often results in a large number of
+% parameters; 16 to 30 are not uncommon when intra-day data are being
+% dealt with. This raises the question of true fitness: the model
+% specification may indeed fit the selected data set but very well fall
+% short of an accurate description of the time series in question. In
+% other words, such a model is unstable and hence unreliable: if one
+% removes a small fraction of data from the beginning of the set and
+% appends some to the end -- a necessary operation for the evolution of
+% the model -- the model parameters change drastically and so do the
+% predictions.
+
+% As intra-day transaction data become more accessible, the volatility,
+% which has been considered an unobservable quantity by models like
+% GARCH, can be estimated by the so called realized volatility, which is
+% the square root of the sum of squared returns that have been sampled
+% at a higher frequency than are those that one intends to model. For
+% example, it has been shown that the volatility of daily returns is
+% best estimated using 17-minute intra-day returns \cite{Sahalia05}.
+
+% Now that the volatility can be estimated, models that take advantage
+% of this are expected to achieve significantly higher accuracy than
+% those that do not. The model presented in this thesis is built indeed
+% in this spirit.
+
+% Like the GARCH models, our proposed model has the simple structure
+% $r_t = \sigma_t z_t$, where $z_t$ has unit variance and is
+% drawn from any appropriate distribution; $\sigma_t$ is the
+% volatility. Unlike GARCH, which treats $\sigma_t$ as deterministic at
+% time $t$, we treat $\sigma_t$ as random and model $\ln\sigma_t$ using
+% time series analysis. This is possible because past values of
+% $\ln\sigma_t$ can be estimated using past values of the returns at a
+% higher frequency, as has just been discussed.
+
+% Our model has a number of advantages too. Like the original
+% GARCH model, it is simple. It has two random variables $z_t$ and
+% $\ln\sigma_t$, the latter being the subject of time series
+% modeling. The use of conventional time series analysis means the
+% model can be not only more accurate but also more economic -- long
+% memory of the volatility, which is the cause of a large number of
+% parameters in a GARCH model, can be modeled by differencing or
+% fractional differencing. As a result, much fewer parameters are needed
+% for the model.
+
+% Secondly, unlike GARCH models, skewness follows directly from the
+% correlation between the two random variables $z_t$ and
+% $\ln\sigma_t$. As shown in chapter \ref{chp:PriceModels},
+% positive/negative correlation leads to positive/negative skewness,
+% and skewness increases with the strength of correlation.
+
+% Thirdly, our model has sub-exponential tails, as shown in chapter
+% \ref{chp:PriceModels}. This type of tails are not as fat as regularly
+% varying tails, but in our specific case, the tails do behave
+% approximately as a power-law in the regions relevant to stock and
+% index returns. After all, power-law behavior is an empirical result
+% obtained in the returns data, and such data are not available to a
+% statistically significant amount at extremely large values. In plain
+% words, situations where one makes a 20\%+ profit on a single stock
+% trade in a single day happen too rarely to allow any statistical
+% statement.
+
+Another important part of the thesis is about the cross-correlation
+matrix, for which we study the distribution of its elements as well
+as its eigenvalues. When the cross-correlation matrix is constructed
+from returns with simple Gaussian distribution, the matrix is called a
+Wishart Matrix and has been studied extensively in the literature. If
+the returns have L\'evy distributions, the matrix is termed
+Wishart-L\'evy and has been studied to some extent, particularly
+regarding its eigenvalue distribution \cite{politi2010}.
+
+However, it is understood that real stock/index returns are much more
+complicated than a straight-forward Gaussian or L\'evy distribution can
+describe - instead, one needs structured models for their description.
+
+Therefore in this thesis we study how the elements and the eigenvalues
+of the cross-correlation matrix are distributed when the matrix is
+built from returns described by different returns' models, specifically
+GARCH(1,1) and ARIMA-log-volatility models, whose the tail behavior is
+understood. Moreover, we also study how auto-correlations in the
+returns influence the aforementioned distributions -- such
+auto-correlations, known as second-order auto-correlations decay
+exponentially but may still leave footprints in the cross-correlation
+matrix.
+
+The rest of this report is organized as follows: in
+\S\ref{sec:FundamentalConcepts} we present a few very important
+concepts and notations that we will often refer to in later
+chapters. Then chapter \ref{chp:PriceModels} reviews some of the most
+influential returns' models, calibrates them to a few intra-day return
+series and compares their performance. The unconditional distribution
+functions of ARIMA-log-volatility models, which is not seen to be
+given a full discussion in the literature, are also calculated and
+discussed. Chapter \ref{chp:Gaussian} investigates the elements' and
+eigenvalues' distributions of the Wishart matrix, where the influence
+of auto-correlations in the constituent returns are studied in
+detail. Chapter \ref{chp:CrossCorrelationFat} looks at the elements'
+and eigenvalues' distributions
+
+\section{Fundamental Concepts \&
+  Notations}\label{sec:FundamentalConcepts}
+This section is a list of a few concepts that may be unfamiliar to the
+reader and that we will often refer to later in the thesis:
+\begin{itemize}
+\item Return. Given a fixed time interval $[t - \Delta t, t]$, for example, a
+  day, a week, a month, etc, and a particular asset, for example, a
+  share in company ABC, the return of this asset over the time
+  interval is defined as
+  \begin{eqnarray*}
+    r_{t, \Delta t} &=& \ln p_{t} - \ln p_{t - \Delta t} \\
+    &=& \ln \left(1 + {p_{t} - p_{t - \Delta t} \over p_{t - \Delta t}}\right) \\
+    &\approx& {p_{t} - p_{t - \Delta t} \over p_{t - \Delta t}}
+  \end{eqnarray*}
+  where $p_t$ is the price of the asset at time $t$. $\Delta t$ is
+  sometimes called the time-lag of the return. Quite often, where
+  confusion is not possible, we will just write $r_t$ to mean $r_{t,
+    \Delta t}$, the time lag either does not matter or is clear from
+  the context.
+
+\item Autocorrelation. By autocorrelation, denoted $\rho_k$ here, we
+  mean the correlation between two temporally separated observations
+  of the same time series:
+  \begin{eqnarray*}
+    \rho_k &=& {
+      \E\left[(a_t -\E(a_t))(a_{t-k} - \E(a_{t-k})\right]
+      \over
+      \sqrt{\text{var}(a_t)}\sqrt{\text{var}(a_{t-k})}
+    }
+  \end{eqnarray*}
+  Here $\E(x)$ stands for the expectation value of $x$, and
+  $\text{var}(x)$ stands for the variance of $x$. $k$ is called the
+  time-lag and is the temporal seperation of the two observations
+  measured by the number of observations in between. For example, the
+  time-lag between the 1st and 3rd observation is 2.
+
+\item Cross-Correlation Matrix. Correlations between the returns of a
+  group of assets are described by the cross-correlation matrix. When
+  the asset returns are described by a stable distribution law with
+  L\'evy index $\alpha$, the empirical cross-correlation matrix is
+  constructed as
+  \begin{equation}
+    \label{eq:cross-correlation}
+    \begin{aligned}
+      C_{ij} &= {1 \over T^{2/\alpha}} \sum_{t=1}^T [r_{i,t}-\E(r_i)]
+      [r_{j,t}-\E(r_j)] \\
+      &= {1 \over T^{2/\alpha}} RR'
+    \end{aligned}
+  \end{equation}
+  where $r_{i,t}$ is the return of the i-th asset at time t and is
+  placed at the entry (i, t) of matrix R; $R'$ denotes the
+  transpose of R. In most practical situations, one has abundant data
+  for each and every asset. Thus $T \geq N$ is assumed throughout this
+  thesis.
+
+\item Auto-regressive processes. A time series $r_t$ is called an
+  auto-regressive process of order $p$ and denoted AR(p), if it can be
+  written in the following form:
+  \begin{eqnarray*}
+    r_t &=& \sum_{i=1}^p \phi_i r_{t-i} + a_t
+  \end{eqnarray*}
+  where, for all $i$, $\phi_i \in (-1, 1)$, and the $a_t$'s are
+  independent and identically distributed (iid.) random variables with
+  zero mean. Obviously this implies the mean of $r_t$ is zero
+  too. Apart from this, their distribution of $a_t$ is not restricted
+  to any particular form.
+
+  Of particular interest to this thesis is the AR(1) process:
+  \begin{eqnarray*}
+    r_t = \phi r_{t-1} + a_t
+  \end{eqnarray*}
+  Its autocorrelation function $\rho_k = \text{corr}(r_t, r_{t-k})$
+  ($k = 0, 1, 2, \cdots$) can be easily shown to fall off
+  exponentially:
+  \begin{eqnarray*}
+    \rho_k &=& {\E(r_tr_{t-k}) - \E(r_{t-k})\E(r_{t}) \over
+      \sqrt{\var(r_t) \var(r_{t-k})}} \\
+    &=& {\phi \E(r_{t-1}r_{t-k}) + \E(a_t r_{t-k})
+      \over
+      \sqrt{\var(r_t) \var(r_{t-k})}
+    } \\
+    &=& \phi \rho_{k-1}
+  \end{eqnarray*}
+  where $\E(a_t r_{t-k}) = 0$ follows from the fact that any return
+  $r_{t1}$ must not depend on disturbances $a_{t2}$ that occur later
+  in time. The last equation means $\rho_k$ is a geometric series. Since
+  $\rho_0 = 1$, we have
+  \begin{eqnarray*}
+    \rho_k &=& \phi^k \\
+    |\rho_k| &=& e^{k\ln|\phi|} \\
+  \end{eqnarray*}
+  Note that $|\phi| < 1$ and hence $\ln|\phi| < 0$.
+
+  Although $k$ can only take integer values, it is still useful to
+  define a correlation time $\tau$ such that $\phi^\tau = 1/2$. Such a
+  quantity is more intuitive and constitutes a measure of
+  autocorrelations that is universal and comparable among different
+  time series' models. From the definition of $\tau$ we get
+  \begin{equation}
+    \label{eq:tau_def}
+    \begin{aligned}
+      \phi^\tau &= 1/2 \\
+      \tau &= -{\ln 2 \over \ln\phi} \\
+      \phi &= 2^{-1/\tau}
+    \end{aligned}
+  \end{equation}
+  Figure \ref{fig:AR1-autocorrelation} shows the autocorrelation
+  function of the AR(1) model. As proven above, this function decays
+  exponentially.
+  \begin{figure}[htb!]
+    %\vspace{-15mm}
     \centering
-  \includegraphics[scale=0.5, clip=true, trim=103 236 118
-  200]{../pics/nordea_price_20120116-20120420.pdf}
-  \caption{Nordea Bank paid prices 2012/01/16-2012/04/20}
-  \label{fig:Nordea}
+    \includegraphics[scale=0.5, clip=true, trim=113 229 115
+    139]{../pics/AR1-autocorrelation.pdf}
+    \caption{\small \it Autocorrelation function of the AR(1) model
+      with $\phi=1/\sqrt{2}$. Red circles: autocorrelations at $k=0, 1,
+      2, \cdots$. Blue line: $e^{t\ln\phi}$.}
+    \label{fig:AR1-autocorrelation}
   \end{figure}
-Figure \ref{fig:Nordea} shows the paid prices of Nordea Bank from
-2012/01/16 to 2012/04/20.
+  For more details about conventional time series' models, see
+  \cite{BoxJenkins94}.
+\end{itemize}
 
-\subsection{15min Returns}
-\subsubsection{GARCH model}
-The autocorrelation function (ACF) of $\epsilon_t^2$ are shown in figure
-\ref{fig:nordea_15min_acf}. At the absence of GARCH effects, the ACF
-will have an asymptotic Gaussian distribution with mean 0 and variance
-1/T. The first 5 autocorrelations have considerable sizes and do not
-fall off as in the Gaussian case. This observation suggests a GARCH
-model. Moreover, from \ref{fig:nordea_15min_vlt_acf}, one sees
-apparent autocorrelations between volatilities of consecutive time
-intervals. This again points to a GARCH model.
+For purposes of later reference, we also list some notations that may
+cause confusion to the reader:
+\begin{itemize}
+\item $\E(x)$ or $\mean{x}$: The expectation value of $x$.
+\item $\var(x)$: the variance of $x$.
+\item $\text{cov}(x,y)$: the covariance of $x$ and $y$.
+\item $\text{corr}(x,y)$: the correlation between $x$ and $y$, i.e.
+  \begin{equation*}
+    \text{corr}(x,y) = {\text{cov}(x,y) \over \sqrt{\var(x)\var(y)}}
+  \end{equation*}
+\end{itemize}
+
+\section{Stylized facts}\label{sec:StylizedFacts}
+This section presents and explains a few ``stylized facts'',
+i.e. phenomena that are widely observed and accepted as true.
+\begin{itemize}
+\item Fat tails. It has been consistently reported by various studies
+  - for example \cite{Potters2003} and \cite{Mantegna2000} - that the
+  probability density function (PDF) of stock/index returns are not
+  Gaussian. Unlike the Gaussian PDF, which is symmetric and falls off
+  very quickly as its argument moves from the center to the outskirts
+  (tails), the PDF of stock/index returns are higher on the tails,
+  i.e. the probability of large fluctuations is higher than is dictated
+  by a Gaussian distribution - in fact, even higher than dictated by
+  an exponential function. Empirical studies suggest the distribution
+  function of the returns follows a power-law on the tails --- the
+  exponent of the power depends on the specific stock/index.
+
+  Figure \ref{fig:FatTail} illustrates this feature.
   \begin{figure}[htb!]
     \centering
-  \subfigure[ACF of squared returns]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_acf.pdf}
-    \label{fig:nordea_15min_acf}
-  }
-  \subfigure[ACF of volatilities]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_vlt_acf.pdf}
-    \label{fig:nordea_15min_vlt_acf}
-  }
-  \caption{Nordea Bank 15m autocorrelations}
+    \includegraphics[scale=0.6, clip=true, trim=92 229 110
+    140]{../pics/FatTail.pdf}
+    \caption{\small \it Fat tails of {\it Nordea Bank} 15-minute
+      returns during the period 2013-10-10 and 2014-01-29. The returns
+      are computed using minute-by-minute average prices.}
+    \label{fig:FatTail}
   \end{figure}
 
-To obtain the functional form of the conditional distribution of
-$\epsilon_t$ \footnote{We use the realized volatility as an estimate of the
-conditional volatility. See \cite{Andersen03}.}, we look at the
-qq-plot of $\epsilon_t/\sigma_t$ \ref{fig:nordea_15min_epOversig_qq},
-which is consistent with a standard Gaussian distribution except for a
-few outliers at the two ends. Hence the fat tails of the probability
-density function can be entirely accounted for by GARCH effects,
-implying a Gaussian conditional distribution is appropriate for $\epsilon_t$.
+\item Non-zero skewness. Apart from fat tails, the PDF of stock/index
+  returns are often also skewed. If the skewness is positive
+  (negative), the probability of very large positive (negative)
+  returns is higher than that of very large negative (positive)
+  returns, even though the mean of the returns is 0 or extremely
+  close to 0.
+
+  Table \ref{tab:EmpiricalSkewness} lists the skewness of a few
+  Swedish stocks traded on the Stockholm OMX market.
+  \begin{table}[htb!]
+    \footnotesize
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|c|}
+      \hline
+      Nordea Bank & Volvo B & Boliden & ABB Ltd & H\&M & Scania
+      B & Ericsson B \\
+      \hline
+      0.1362 & 0.0471 & 0.0567 & -0.0364 & -0.0168 &
+      -1.1554 & 0.2086 \\
+      \hline
+    \end{tabular}
+    \caption{\small \it Skewness of Stock Returns. All the returns
+      have time-lag of 15 minutes and are computed using paid prices
+      between 2013-10-10 and 2014-01-29. }
+    \label{tab:EmpiricalSkewness}
+  \end{table}
+
+\item Higher-order autocorrelation. To the lowest order, return series
+  are not auto-correlated --- if they are, the auto-correlations would
+  present an obvious opportunity of making easy profit and be
+  exploited and vanish as soon as they appear. However, the squared
+  returns do have significant auto-correlations, as figure
+  \ref{fig:nordea_15min_acf} illustrates. These auto-correlations
+  suggest the variances of the returns at subsequent time steps are
+  correlated. These are referred to as higher-order auto-correlations
+  and make a subjet of returns' models.
+
+\end{itemize}
+
+\input{price_models}
+\input{cross_correlation_gaussian}
+
+\chapter{Correlation Matrix of GARCH(1,1) Returns}
+\label{chp:CrossCorrelationFat}
+
+% In this chapter we study the cross-correlation matrix $C$ constructed from
+% returns that have a fat-tailed distribution, i.e.
+% \begin{eqnarray*}
+%   C &=& {1 \over T^{2/\alpha}} RR'
+% \end{eqnarray*}
+% where $R$ is an $N\times T$ returns' matrix whose entries are
+% \begin{eqnarray}\label{eq:CrossCorrelationMatrix}
+%   R_{it} &=& \left\{
+%   \begin{array}{ll}
+%     a_{it} & \text{if $t=1$} \\
+%     \phi R_{i, t-1} + a_{it} & \text{if $t>1$} \\
+%   \end{array} \right.
+% \end{eqnarray}
+% Here $a_{it}$ are random variables drawn from a stable distribution
+% with l\'evy index $\alpha$. In \S\ref{sec:LevyWishart} we look at
+% the case where the stable distribution is chosen to be Cauchy; in
+% \S\ref{sec:XieWishart} we look at the case where $a_{it}$ is
+% described by an ARIMA log-volatility model, which is detailed in
+% \S\ref{sec:SLV_model}; finally in \S\ref{sec:GarchWishart} we
+% investigate the case where $a_{it}$ is a GARCH(1, 1) process, which is
+% proven to have regularly varying tails by Mikosch et al
+% \cite{mikosch2000}.
+
+% \section{Wishart-L\'evy Matrices}\label{sec:LevyWishart}
+% In this section we study the case where the distribution of $a_{it}$
+% as appear in equation \ref{eq:CrossCorrelationMatrix} is chosen to be
+% Cauchy, a stable distribution with L\'evy index $\alpha =
+% 1$. Politi et al derived the eigenvalue distribution of Wishart-L\'evy
+% matrices in \cite{politi2010} using free probability theory. Here we
+% show, qualitatively, how the distribution changes when
+% auto-correlations are present among the return series. To be precise,
+% the PDF of $a_{it}$ is
+% \begin{eqnarray*}
+%   f_a(x) &=& {\gamma \over \pi} {1 \over \gamma^2 + x^2}
+% \end{eqnarray*}
+
+% By numerically simulating a large number of cross-correlation matrices
+% and solving for the eigenvalues of each simulated matrix, one can
+% obtain the eigenvalue distribution corresponding to each value of
+% $\phi$. Figure \ref{fig:WishartLevySpectraPositive} shows the
+% distribution for a range of positive $\phi$ values, and figure
+% \ref{fig:WishartLevySpectraNegative} shows those for a range of
+% negative $\phi$ values.
+% \begin{figure}[htb!]
+%   \centering
+%   \subfigure[]{
+%     \includegraphics[scale=0.35, clip=true, trim=94 223 109
+%     134]{../pics/WishartLevyEigDist.pdf}
+%     \label{fig:WishartLevySpectraPositive}
+%   }
+%   \subfigure[]{
+%     \includegraphics[scale=0.35, clip=true, trim=94 223 109
+%     134]{../pics/WishartLevyNegEigDist.pdf}
+%     \label{fig:WishartLevySpectraNegative}
+%   }
+%   \caption{\small \it Eigenvalue distribution (PDF) of Wishart-L\'evy
+%     Matrices. Horizontal axis: eigenvalues ($\lambda$) on log-scale;
+%     Vertical axis: PDF ($f_\phi(\lambda)$) of the eigenvalues on
+%     log-scale. \ref{fig:WishartLevySpectraPositive}: $\phi > 0$;
+%     \ref{fig:WishartLevySpectraNegative}: $\phi < 0$;  The $|\phi|$
+%     values 0, 0.5, 0.8, and 0.99 correspond to $\tau$ (correlation
+%     time) values 0, 1.0, 3.10, and 68.97.  The correlation time is
+%     independent of the sign of $\phi$.}
+% \end{figure}
+
+% It is seen in these figures
+% \begin{itemize}
+% \item the eigenvalue distribution is identical for $\phi$ and
+%   $-\phi$;
+% \item the distribution is not bounded within a finite interval but rather
+%   extends to infinity, in constrast to the Marcenko-Pastur law that
+%   bounds the maximum eigenvalue at $\sigma^2(1 + \sqrt{N/T})^2$;
+% \item as auto-correlation increases, the distribution flatens,
+%   implying probability being transferred to large eigenvalues.
+% \end{itemize}
+
+% \section{Cross-Correlation Matrix of ARIMA Log-volatility
+%   returns}\label{sec:XieWishart}
+% In this section we study the cross-correlation matrix of return series
+% specified as follows:
+% \begin{equation}
+%   \label{eq:CrossCorrSLV1}
+%   \begin{aligned}
+%     R_{it} &= \phi R_{i, t-1} + a_{it} \\
+%     a_{it} &= e^{v_{it}} b_{it}
+%   \end{aligned}
+% \end{equation}
+% where $v_{it}$ is typically described by an ARIMA model as discussed
+% in \S\ref{sec:SLV_model}. The unconditional distribution of
+% such returns is calculated in \S\ref{sec:XieCalc}, where it is shown
+% that all the moments of the distribution are finite. Based on this
+% observation, we expect the Marcenko-Pastur law to emerge for the
+% eigenvalue distribution of the cross-correlation matrix C, when
+% $\phi=0$, i.e. when no auto-correlations are absent. In this case C is
+% built as
+% \begin{eqnarray*}
+%   C &=& {1 \over T} \sum_{t=1}^T (DR)(DR)' \\
+% \end{eqnarray*}
+% where $R'$ denotes the transpose of $R$ and
+% \begin{eqnarray*}
+%   D &=&
+%   \begin{pmatrix}
+%     1/\sigma_1 & 0 & \cdots & 0 \\
+%     0 & 1/\sigma_2 & \cdots & 0 \\ 
+%     \vdots & \vdots & \ddots & 0 \\
+%     0 & 0 & \cdots & 1/\sigma_N
+%   \end{pmatrix}
+% \end{eqnarray*}
+% $\sigma_i$ is the standard deviation of return series $i$. By
+% numerical simulation as described above, the eigenvalue distribution
+% of C is obtained. We plot it in figure \ref{fig:FatWishartSpectra}.
+% \begin{figure}[htb!]
+%   \centering
+%   \vspace{-20mm}
+%   \includegraphics[scale=0.5, clip=true, trim=91 229 116
+%   137]{../pics/FatWishartEigDist.pdf}
+%   \caption{\small \it Eigenvalue distribution (PDF) of a
+%     cross-correlation matrix with returns described by
+%     \ref{eq:CrossCorrSLV1}. $\phi$ = 0, 0.5, 0.8, 0.95 correspond to
+%     correlation time $\tau$ = 0.00, 1.00, 3.11, 13.51}
+%   \label{fig:FatWishartSpectra}
+% \end{figure}
+% As expected, the case with zero auto-correlation has an eigenvalue
+% distribution governed by the Marcenko-Pastur law. As auto-correlations
+% become stronger, the distribution deforms, but in a rather different
+% way if compared to the case of Gaussian returns. Figure
+% \ref{fig:FatWishartSpectra} shows that the minimum eigenvalue
+% decreases with increasing $\phi$ in the present case of ARIMA
+% log-volatility returns, while figure \ref{fig:Gaussian_mineig} shows
+% that, in the case of Gaussian returns, the minimum eigenvalue actually
+% increases with $\phi$. We leave this interesting comparison to future
+% investigations.
+
+% \section{Cross-Correlation Matrix of GARCH(1,1) returns}
+% \label{sec:GarchWishart}
+In this chapter we consider the cross-correlation matrix of N
+identically specified, possibly auto-correlated GARCH(1, 1)
+processes:
+\begin{eqnarray}
+  r_{it} &=& \phi r_{i, t-1} + \epsilon_{it} \nonumber \\
+  \epsilon_{it} &=& \sigma_{it} a_{it} \label{eq:garch_spec}
+\end{eqnarray}
+where $i=1,2,...,N$; $t=1,2,...,T$; $a_{it}$ is independent,
+identically distributed, and
+\begin{eqnarray*}
+  \sigma_{it}^2 &=& \alpha_0 + \alpha_1 a_{i, t-1}^2 + \beta_1
+  \sigma_{i,t-1}^2
+\end{eqnarray*}
+Mikosch et al showed in \cite{mikosch2000} that a GARCH(1,1) process
+satisfying
+\begin{eqnarray*}
+  \alpha_0 &>& 0 \\
+  \E(\alpha_1 Z^2 + \beta_1) &<& 0 \\
+\end{eqnarray*}
+and
+\begin{eqnarray*}
+  \E[(\alpha_1 z^2 + \beta_1)^{p/2}] &\geq& 1 \\
+  \E|Z|^p \ln|Z| &\leq& \infty
+\end{eqnarray*}
+for some $p > 0$, is stationary and has regularly varying tails. The
+tail exponent $\alpha$ is determined by:
+\begin{equation}\label{eq:garch_alpha}
+  \E[(\alpha_1 Z^2 + \beta_1)^{\alpha/2}] = 1
+\end{equation}
+Here $Z$ is a random variable that has the same distribution as
+$a_{it}$. In our simulations described hereafter $a_{it}, Z \sim N(0,
+1)$. In this section and the next, we first study situations where no
+auto-correlations are present among the returns, i.e. $\phi = 0$; then
+in \S\ref{sec:garch_nonzero_phi} we look at how auto-correlations
+change the picture.
+
+With regularly varying tails, the eigenvalue distribution of a
+cross-correlation matrix built from GARCH(1,1) returns is expected to
+differ from the Marcenko-Pastur law. In figure
+\ref{fig:garch_ev_cii} we simulate N=50 independent GARCH(1,1)
+returns series, each with identical parameters, namely $\alpha_0 =
+2.3\times 10^{-6}$, $\alpha_1 = 0.15$, $\beta_1 = 0.84$, $\phi = 0$
+and T=$8\times10^4$ time steps, then we build the cross-correlation
+matrix as
+\begin{eqnarray*}
+  C &=& {1 \over T^{2/\alpha}} RR'
+\end{eqnarray*}
+where R is an $N\times T$ matrix, with elements $r_{it}$ specified by
+\ref{eq:garch_spec}. The PDF of the eigenvalue distribution of C is
+plotted in figure \ref{fig:garch_ev_cii_linear} and CDF of the
+distribution is plotted on log-log scale in
+\ref{fig:garch_ev_cii_loglog}.
 \begin{figure}[htb!]
   \centering
-    \includegraphics[scale=0.6, clip=true, trim=100 236 118
-    200]{../pics/nordea_15min_epOversig_qq.pdf}
-  \caption{QQ-plot of $\epsilon_t / \sigma_t$. $\epsilon_t$ are
-    derived from Nordea Bank 15m returns while $\sigma_t$ are realized
-    volatilities calculated using 30s returns within each 15m
-    interval.}
-  \label{fig:nordea_15min_epOversig_qq}
+  \subfigure[]{
+    \includegraphics[scale=0.40, clip=true, trim=104 271 111
+    216]{../pics/garch_ev_cii.pdf}
+    \label{fig:garch_ev_cii_linear}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.40, clip=true, trim=111 241 110
+    134]{../pics/garch_ev_cii_loglog.pdf}
+    \label{fig:garch_ev_cii_loglog}
+  }
+  \caption{\small \it \ref{fig:garch_ev_cii_linear}: Eigenvalues \&
+    Diagonal elements' distribution of a cross-correlation matrix
+    built from independent GARCH return series. Blue: PDF of
+    eigenvalues; Green: PDF of diagonal elements; Red: PDF of a
+    $\alpha$-stable distribution fitted to the diagonal
+    elements. \ref{fig:garch_ev_cii_loglog}: CDF of the same
+    quantities in 10-based log-log scale.}
+  \label{fig:garch_ev_cii}
+\end{figure}
+Also plotted in the same figure is the distribution of the diagonal
+elements of C. It is clear from the figure that the two PDFs coincide,
+implying C is diagonal. This is further confirmed by figure
+\ref{fig:garch_cij} which shows the distribution of the non-diagonal
+elements of C. One can see the non-diagonal elements are distributed
+symmetrically around 0 with a very small width in comparison to the
+distribution of the diagonal elements - in fact, 1 order of magnitude
+smaller ($2.14\times10^{-5}$ v.s. $7.31\times 10^{-4}$). Hence $C$ is
+very close to a diagonal matrix.
+\begin{figure}[htb!]
+  \centering
+    \includegraphics[scale=0.5, clip=true, trim=95 253 91
+    223]{../pics/garch_cij.pdf}
+  \caption{\small \it Distribution of the non-diagonal elements of
+    the cross-correlation matrix. Blue: PDF of the non-diagonal
+    elements; Green: $\alpha$-stable distribution fitted to the
+    non-diagonal elements' PDF.}
+  \label{fig:garch_cij}
 \end{figure}
 
-To determine the orders p and q of the GARCH model, we again check
-the partial autocorrelation function of $\epsilon_t^2$. As just
-mentioned, the first 5 partial autocorrelations are significant. Thus
-we can tentatively estimate a GARCH(1, 5) model. Then by applying the
-Lagrange multiplier test and accordingly removing insignificant
-parameters, we obtain a GARCH(1, 3) model
-\ref{tab:nordea_15min_garch}:
+Figure \ref{fig:garch_ev_cii} also shows the two curves are well fitted
+by an $\alpha$-stable distribution, with $\alpha \approx 1.38$. This
+is really an expected result for the diagonal elements. Using
+$\alpha_1 = 0.15$, $\beta_1 = 0.84$, which are the values used for
+simulating the GARCH returns, one can obtain, by solving equation
+\ref{eq:garch_alpha}, $\alpha=2.96$. Then according to Mikosch et al
+\cite{mikosch2000}
+\begin{eqnarray*}
+  P(|r_t| > x) &\sim& {\E(z^\alpha) c_0 \over x^\alpha} \\
+  P(|r_t|^2 > x) &\sim& {\E(z^\alpha) c_0 \over x^{\alpha/2}} \\
+\end{eqnarray*}
+for some constant $c_0$. Now that $P(|r_t|^2 > x)$ has power-law tail
+behavior with power $\alpha/2 < 2$, one can deduce
+\begin{equation}
+  \label{eq:stable_CLT}
+  \begin{aligned}
+    \sum_{t=1}^T r_t^2 &\xrightarrow{d} S(\alpha/2,
+    1, \gamma, \mu) \text{ as $T \to \infty$}
+  \end{aligned}
+\end{equation}
+where $\xrightarrow{d}$ denotes convergence in distribution, and
+$S(\alpha/2, 1, \gamma, \mu)$ denotes an $\alpha$-stable distribution
+with parameters $(\alpha/2, 1, \gamma, \mu)$. Here $\alpha/2$ is the
+L\'evy index, 1 is the asymmetry, $\gamma$ is the scale parameter
+and $\mu$ is the mean value of the distribution. Asymmetry being 1
+means a random variable so distributed only takes positive values
+\cite{Bilik2008, Embrechts1997}.
+
+The mean $\mu$ in equation \ref{eq:stable_CLT} is given by
+\begin{eqnarray*}
+  \mu &=& T \E(|r_t|^2) \\
+  &=& T{\alpha_0 \over 1 - \alpha_1 - \beta_1}
+\end{eqnarray*}
+and the scale parameter $\gamma$ in \ref{eq:stable_CLT} is determined
+by the limit \cite{Bilik2008}
+\begin{eqnarray*}
+  \lim_{T\to\infty} {T \E(z^{\alpha/2}) c_0 \over \gamma^{\alpha/2}}
+  &=& C_{\alpha/2}
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+  C_{\alpha/2} &=& \left( \int_0^\infty {\sin x \over x^{\alpha/2}} dx
+  \right)^{-1} \\
+  &\approx& {1 \over \sqrt{2 \pi}}
+\end{eqnarray*}
+Therefore
+\begin{eqnarray*}
+  \gamma^{\alpha/2} &=& \sqrt{2\pi} T \E\left(
+    |z|^{\alpha/2}
+  \right) c_0 \\
+  \gamma &=& (2\pi)^{1/\alpha} T^{2/\alpha} \E\left(
+    |z|^{\alpha/2}
+  \right)^{2/\alpha} c_0^{2/\alpha}
+\end{eqnarray*}
+Here we note that an $\alpha$-stable distribution $S(\alpha, \beta,
+\gamma, \mu)$ has characteristic function \cite{Guhr2007}
+\begin{eqnarray*}
+  \phi(k; \alpha, \beta, \gamma, \mu) &=& \exp\left[
+    i\mu k - \gamma^\alpha |k|^\alpha \left(
+      1 - i \beta {k \over |k|} \tan{\pi \alpha \over 2}
+    \right) \right] \text{ for $\alpha \neq 1$}
+\end{eqnarray*}
+from which we see $\phi(ak; \alpha, \beta, \gamma, \mu) = \phi(k;
+\alpha, \beta, a\gamma, a\mu)$, implying that, if $x \sim S(\alpha,
+\beta, \gamma, \mu)$, then $ax \sim S(\alpha, \beta, a\gamma, a\mu)$.
+
+Now that
+\begin{eqnarray*}
+  \sum_{t=1}^T r_t^2 &\xrightarrow{d} S(\alpha/2,
+  1, \gamma, \mu) \text{ as $T \to \infty$}  
+\end{eqnarray*}
+we have
+\begin{eqnarray*}
+  C_{ii} &=& {1 \over T^{2/\alpha}}\sum_{t=1}^T r_{it}^2 \\
+  &\xrightarrow{d}&
+  S(\alpha/2, 1, \gamma_D, \mu_D) \text{ as $T \to \infty$}
+\end{eqnarray*}
+where
+\begin{eqnarray}\label{eq:garch_wishart_cij_params}
+  \gamma_D &=& (2\pi)^{1/\alpha} \E\left(|z|^{\alpha/2}
+  \right)^{2/\alpha} c_0^{2/\alpha} \\
+  \mu_D &=& {\alpha_0 \over 1 - \alpha_1 - \beta_1} T^{1 - 2/\alpha}
+  \nonumber
+\end{eqnarray}
+So the diagonal elements of the cross-correlation matrix converge to
+an $\alpha$-stable distribution with L\'evy index $\alpha/2 \approx
+1.48$. This is consistent with the measured index 1.38 within
+measurement errors.
+
+Now we look at the distribution of the non-diagonal elements. At first
+glance, one might be tempted to think that figure \ref{fig:garch_cij}
+shows a Gaussian distribution, but in fact, it is not. This is
+illustrated in the probability plot in figure
+\ref{fig:garch_nondiag_probplot}. In comparison to the PDF of a
+Gaussian, the PDF of the nondiagonal elements is significantly fatter
+on the tails.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=81 229 114
+    121]{../pics/garch_cij_prob.pdf}
+    \caption{\small \it Probability plot of the non-diagonal
+      elements. Blue: The accumulative probability function
+      (CDF) of the non-diagonal elements. Black, dashed: CDF of the
+      Gaussian distribution that has the same mean and variance as the
+      sample. The graph is arranged on such a scale that the Gaussian
+      CDF is a straight line.
+    }
+  \label{fig:garch_nondiag_probplot}
+\end{figure}
+
+The parameters of the non-diagonal elements' distribution $S(\alpha',
+\beta', \gamma', \mu')$ can be obtained via fitting. The results have
+been shown in the legend of figure \ref{fig:garch_cij}. In table
+\ref{tab:garch_wishart_cij_params} we list them with a higher
+precision.
 \begin{table}[htb!]
   \centering
-  \begin{tabular}{c|c|c|c}
-    parameter & value & std. error & t statistic \\
+  \begin{tabular}{|c|c|c|c|}
     \hline
-     $\alpha_0$ &   1.05668e-06 &   3.56871e-07 &    2.96096\\
+    $\alpha'$ & $\beta'$ & $\gamma'$ & $\mu'$ \\
     \hline
-     $\beta_1$  &     0.687212  &   0.0299314   &    22.9596\\
+    1.9453 & 0.0018 & $2.1381 \times 10^{-5}$ & $2.0637\times 10^{-9}$ \\
     \hline
-     $\alpha_1$ &      0.128614 &   0.0268144   &    4.79643\\
-     \hline
-     $\alpha_3$ &    0.0568022  &   0.0301503   &    1.88397
   \end{tabular}
-  \caption{GARCH(1, 3) model of Nordea Bank 15m returns}
-  \label{tab:nordea_15min_garch}
+  \caption{\small \it Parameters of the non-diagonal elements' distribution}
+  \label{tab:garch_wishart_cij_params}
 \end{table}
+Since $r_{it}$ and $r_{jt}$ are independent of each other, $\beta'$
+and $\mu'$ are expected to be 0 --- but with a finite T, some deviation
+from 0 is not surprising.
 
-\subsubsection{$(0,1,1)\times(0, 1, 1)$ seasonal volatility model}
-With the availability of transactional data, estimating conditional
-volatility using returns sampled at a higher frequency (realized
-volatility) gives superior acuracy and reliability. However, at which
-frequency the time series should be sampled in order to give an
-unbiased and consistent estimate of the volatility is not a trivial
-question. Naively one would argue by the doctrine of statistics that
-the often the series is sampled, the better the estimate. This is also
-well supported in theory, since the quadratic variation which is the
-theoretial variance of the time series (see \cite{Andersen03}, theorem
-1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
-sec.6, theorem 22):
-\[
-\var(p_{t+h} - p_t) = [r, r]_{t+h} - [r, r]_t = \lim_{n \to \infty}
-\sum_{T^i \in \pi_n} (p^{T_{i+1}} - p^{T_i})^2
-\]
-where $p_t$ is the log price process and $r_t = p_t - p_0$. $\pi_n$
-is a refining partition of the interval $[t, t+h]$.
+Now that the parameters' values have been obtained, the way they scale
+with T, i.e. the length of the return series, can be deduced. Consider
+\begin{equation}\label{eq:garch_cij1}
+  T^{2/\alpha} C_{ij} = \sum_{t=1}^T r_{it} r_{jt} \xrightarrow{d} S(\alpha',
+  0, \gamma', 0)
+\end{equation}
+where the width $\gamma'$ is determined by \cite{Bilik2008},
+\begin{eqnarray*}
+  \lim_{T\to\infty} {T C \over \gamma'^{\alpha'}}
+  &=& C_{\alpha'} \\
+  C_{\alpha'} &=& \left( \int_0^\infty {\sin x \over x^{\alpha'}} dx
+  \right)^{-1} \\
+\end{eqnarray*}
+Here the constant C is such that
+\begin{equation*}
+  P(|r_{it}r_{jt}| > x) \sim {C \over x^{\alpha'}} \text{ as $x \to \infty$}
+\end{equation*}
+So we have
+\begin{eqnarray*}
+  \gamma' &=& {\left( CT \over C_{\alpha'}\right)^{1/\alpha'}}
+\end{eqnarray*}
+Divide throughout equation \ref{eq:garch_cij1} by $T^{2/\alpha}$ then
+gives
+\begin{equation}\label{eq:garch_cij2}
+  C_{ij} = {1 \over T^{2/\alpha}}\sum_{t=1}^T r_{it} r_{jt}
+  \xrightarrow{d}
+  S(\alpha', 0, \gamma_N, 0)
+\end{equation}
+where
+\begin{eqnarray*}
+  \gamma_N &=& {\left(C \over C_{\alpha'}\right)^{1/\alpha'}}
+  T^{1/\alpha'-2/\alpha} \\
+\end{eqnarray*}
+So we see that distribution of the non-diagonal elements has a width
+that scales with T as $T^{1/\alpha' - 2/\alpha} \approx T^{-1/6}$,
+while the width of the diagonal elements' distribution, as shown in
+equation \ref{eq:garch_wishart_cij_params}, does not scale with
+T. This is to be compared with the Wishart case where the return
+series have Gaussian distribution, and hence the asymptotic
+distributions of both the diagonal and the non-diagonal elements of the
+covariance matrix are Gaussian with a variance that scales as $1/T$
+(c.f. \ref{sec:GCC-analytical}).
 
-However, due to the noise introduced by market micro-structure, there
-is actually an optimal sampling frequency relative to $h$. While the
-method to determine this optimal frequency is a subject of debate (see
-for example \cite{Sahalia05}), it is not difficult to find a fairly
-satisfactory frequency in practice.
+% We also notice that, given our particular choice of $\alpha_1$ and
+% $\beta_1$, the covariance matrix's convergence to a diagonal matrix
+% is slow, since the non-diagonal elements decay to zero only
 
-Andersen and Bollerslev et al proved in \cite{Andersen03} that, with
-the conditional volatility well approximated by the realized
-volatility, $r_{t+h, h} = p_{t+h} - p_t$ is governed by a Gaussian
-distribution (see theorem 2 of \cite{Andersen03}):
-\[
-r_{t+h, h} | \sigma(\mu_{t+s}, \sigma_{t+s})_{s \in [0, h]} \sim
-N(\int_{0}^h \mu_{t+s} ds, \int_{0}^h \sigma_{t+s}^2 ds)
-\]
+% In the Wishart case, the diagonal and the non-diagonal elements
+% scale in the same way with T and thus have comparable sizes at large
+% T --- so the matrix is not diagonal and the eigenvalue distribution has
+% the Marcenko-Pastur law. In the GARCH case, the diagonal elements do
+% not scale with T while the non-diagonal elements scale as $T^{-1/6}$,
+% thus at large T, the non-diagonal elements are significantly smaller
+% than the diagonal ones --- so a diagonal matrix arises.
 
-Thus one can try a few frequencies and compare the distribution of
-$r_{t+h, h}/\hat{\sigma_{t+h, h}}$ with the standard Gaussian. If the
-two match, $\hat{\sigma_{t+h, h}}$ is a good estimate of $\sigma_{t+h,
-  h}$, the volatility of $r_{t+h, h}$.
+\section{Implications of Finite Number of Observations}
+In the last section we have discussed the limiting situation where the
+return series of the covariance matrix have an infinite number of
+observations ($T \to \infty$), and in the simulation studies we have
+generated a large number of observations for each return series,
+namely $T = 8 \times 10^4$. However, in practice, one often does not
+have such a large number of observations available. Therefore, it is
+useful to investigate situations where $T$ only has a modest size.
 
-For the present case with $h = 15\text{min}$, a reasonally good
-sampling frequency is 1/30sec, as confirmed by the QQ-plot
-\ref{fig:nordea_15min_epOversig_qq} as well as by the autocorrelation
-functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of $(r_{t+h, h} /
-\sigma_{t+h, h})^2$ which are shown in
-\ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} respectively.
+Figure \ref{fig:GarchCiiEig1} shows the diagonal elements' as well as
+the eigenvalues' distributions when the number of return series (N) is
+fixed at 250 and the number of observations in each series (T) is
+increased from 3000 to 6000.
 \begin{figure}[htb!]
   \centering
-  \subfigure[ACF of $r_{t+h, h} / \sigma_{t+h, h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_acf.pdf}
-    \label{fig:nordea_15min_quotient_acf}
+    \includegraphics[scale=0.5, clip=true, trim=81 248 75
+    195]{../pics/GarchCiiEig1.pdf}
+  \caption{\small \it The Diagonal elements' and the eigenvalues'
+    distributions with modest T. Blue: eigenvalues' CDF;
+    Green: Diagonal elements' CDF; Red: CDF of the $\alpha$-stable
+    distribution fitted to the diagonal elements. All the curves are
+    drawn on 10-based log-log scale.
   }
-  \subfigure[ACF of $(r_{t+h, h} / \sigma_{t+h, h})^2$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
-    \label{fig:nordea_15min_quotient_squared_acf}
-  }
-  \caption{Nordea 15min $r_t/\sigma_t$ and $(r_t/\sigma_t)^2$ ACF.}
+  \label{fig:GarchCiiEig1}
+\end{figure}
+It is seen from the figure that, compared to the earlier case where $T
+= 8 \times 10^4$, the eigenvalue distribution and the diagonal
+elements' distribution do not coincide as well but differ rather
+siginificantly for small values. For large values, however, the two do
+coincide and comply with the fitted $\alpha$-stable
+distribution. Moreover, the difference between the eigenvalues'
+distribution and the diagonal elements' distribution, as well as that
+between the diagonal elements' distribution and the fitting
+$\alpha$-stable distribution, is also seen to diminish as T
+increases.
+
+The convergence of the eigenvalues ($\lambda$) towards the diagonal
+elements $C_{ii}$ as $\lambda \to \infty$ is really an anticipated
+result. For convenience, we order the diagonal elements so that
+\begin{eqnarray*}
+  C_{11} < C_{22} < \cdots < C_{NN}
+\end{eqnarray*}
+where, as before, $N$ is the dimension of the covariance matrix
+$C$. Then, as $x \to \infty$,
+\begin{eqnarray*}
+  P(C_{ii} > x) &\sim& {c \over x^{\alpha/2}} \\
+  f(C_{ii}) &\sim& {c \over x^{\alpha/2 + 1}} \\
+\end{eqnarray*}
+where $f(\cdot)$ denotes the PDF of the diagonal elements and $c$ is
+some constant. Thus, at the limit $C_{ii} \to \infty$, the distance
+between two adjacent diagonal elements $C_{ii}$ and $C_{i+1, i+1}$ can
+be expressed as
+\begin{eqnarray*}
+{1 \over N (C_{i+1, i+1} - C_{ii})} &=& f(C_{ii})\\
+C_{i+1, i+1} - C_{ii} &\sim& {C_{ii}^{\alpha/2 + 1} \over c N}
+\end{eqnarray*}
+
+Thus as $C_{ii} \to \infty$, $C_{i+1, i+1} - C_{ii} \to \infty$ while
+$C_{ij} \to 0$ ($i \neq j$). Now that the spacing between adjacent
+diagonal elements become wider and the non-diagonal elements become
+smaller, the eigenstates considered as a mixture of the basis states,
+become more and more localized to a prominent basis state. This
+localization can be measured by the size of the component in each
+eigenvector that has the largest absolute value($|c|_\M$), provided
+that the eigenvectors have been normalized.
+
+Figure \ref{fig:garch_eigenvec_Cmax} shows how $|c|_\M$ changes
+in response to increasing $\lambda$. It is seen that as T increases,
+localization of the eigenstates proceeds from those with very
+large eigenvalues towards those with relatively smaller
+eigenvalues. At the same time, the minimum of $|c|_\M$ increases
+and so does the minimum of the eigenvalues. The last point here is
+further illustrated in figure \ref{fig:garch_eigenvec_Cmax_dist} where
+the PDF of $|c|_\M$ is plotted. We see in this figure that an
+increased value of T leads to advancement of $\min(|c|_\M)$ to larger
+values as well as to an increased proportion of large $|c|_\M$. The
+mean of $|c|_\M$ has apparently been increased too.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=14 199 31
+  155]{../pics/garch_eigenvec_Cmax.pdf}
+  \caption{\small \it largest component ($|c|_\M$) corresponding to
+    the eigenvalue. N is fixed to 50.}
+  \label{fig:garch_eigenvec_Cmax}
 \end{figure}
 
-We see in figure \ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
-autocorrelation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
-$(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
-series is i.i.d Gaussian distributed. As a result, the problem of
-modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
-  h}$. Andersen and Bollerslev et al reported that, for the exchange
-rates between Deutch mark, yen and dollar, $\ln \sigma_{t+h, h}$ is
-gaussian distributed \cite{Andersen03}. This is, however, not the case
-for our series in question. In fact, in our case, $\ln \sigma_{t+h,
-  h}$ is right skewed (skewness 0.3342) and leptokurtic (kurtosis
-6.1006). See figure \ref{fig:nordea_15min_logvol_pdf} and
-\ref{fig:nordea_15min_logvol_qq}
 \begin{figure}[htb!]
   \centering
-  \subfigure[PDF of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.32, clip=true, trim=52 173 46
-    43]{../pics/nordea_15min_logvol_hist.pdf}
-    \label{fig:nordea_15min_logvol_pdf}
-  }
-  \subfigure[QQ-plot of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_logvol_qq.pdf}
-    \label{fig:nordea_15min_logvol_qq}
-  }
-  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ unconditional
-    distribution}
+  \includegraphics[scale=0.5, clip=true, trim=86 257 112
+  136]{../pics/garch_eigenvec_Cmax_dist.pdf}
+  \caption{\small \it PDF of the largest component ($|c|_\M$) of
+    each eigenvector. N is fixed to 50.}
+  \label{fig:garch_eigenvec_Cmax_dist}
 \end{figure}
 
-Moreover, the autocorrelation function of $\ln\sigma_{t+h, h}$
-shows long-lasting and periodic autocorrelations in the time series
-with an apparent period of 33 (see figure
-\ref{fig:nordea_15min_logvol_acf}). This suggests simplifying the
-series by differencing \cite{BoxJenkins94}:
-\[
-w_t = (1-B)(1-B^s)\ln\sigma_{t+h, h}
-\]
-where $B$ is the back-shift operator. The length of the time interval
-h, which is 15min in the present case, is left implicit.
+Another informative quantity that measures the localization is the
+``Inverse Participation Ratio'' (IPR). For a given normalized eigenvector
+$\vec{c}_i = (c_{1, i}, c_{2, i}, ..., c_{N, i})$, the IPR is defined
+as \cite{Aberg2013}
+\begin{eqnarray*}
+  \text{IPR}(\vec{c}_i) &=& \sum_{k=1}^N c_{k,i}^4 \\
+\end{eqnarray*}
+Figure \ref{fig:garch_eigenvec_PR} shows ${1 \over
+  N \cdot \text{IPR}(\vec{c}_i)}$ in correspondence to the
+eigenvalues. This quantity is sometimes termed the
+normalized participation ratio (PR) and measures the proportion of
+basis vectors that contribute considerably to the eigenvector in
+question. From this figure we see that, for all values of T, if an
+eigenvalue is larger than $10^{-1.5} \approx 0.03$, its corresponding
+participation ratio is less than $10^{-1.6} = 2.5\%$, meaning less
+than $50 \times 0.025 = 1.26$ basis vectors contribute --- each of the
+corresponding eigenvectors is localized to a single basis vector and
+hence the distribution of such large eigenvalues is the same as the
+diagonal elements' distribution.
 
-The autocorrelation function of the differenced process $w_t$, as shown in
-figure \ref{fig:nordea_15min_w_acf}, clearly points to a seasonal $(0,
-0, 1)\times(0, 0, 1)$ model: There are only 4 non-zero
-autocorrelations in the plot, located at lags 1, 32, 33, 34,
-respectively; furthermore, the two at 32 and 34 are approximately
-equal. Thus we can write down the model as
-\begin{eqnarray*}
-  w_t = (1 - \theta B)(1 - \Theta B^s) y_t
-\end{eqnarray*}
-where $\theta$ and $\Theta$ are parameters to be determined and $y_t$
-is a noise process with constant variance $\sigma_y^2$ and mean
-0. $y_t$ is often refered to as the residuals.
-
-The above $(0, 0, 1)\times(0, 0, 1)$ model has the following
-autocovariance structure \cite{BoxJenkins94}:
-\begin{eqnarray*}
-  \gamma_0 &=& \sigma_y^2 (1 + \theta^2)(1 + \Theta^2) \\
-  \gamma_1 &=& -\sigma_y^2\theta(1 + \Theta^2) \\
-  \gamma_s &=& -\sigma_y^2\Theta(1 + \theta^2) \\
-  \gamma_{s+1} &=& \gamma_{s-1}\;=\;\sigma_y^2\theta\Theta
-\end{eqnarray*}
-These equations together with the measured autocorrelations make
-possible an initial estimate of the parameters $\theta$ and $\Theta$:
-\begin{eqnarray*}
-  {\rho_{s+1}/\rho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
-    1 + \theta^2} \\
-  {\rho_{s+1}/\rho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
-    1 + \Theta^2} \\
-\end{eqnarray*}
-Substituting in the measured values shown in table
-\ref{tab:nordea_15min_w_acf},
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{c|c|c|c}
-    $\rho_1$ & $\rho_{s-1}$ & $\rho_s$ & $\rho_{s+1}$ \\
-    \hline
-    -0.4703 &  0.2053 & -0.4564 &  0.2212
-  \end{tabular}
-  \caption{Nordea 15min $w_t$ autocorrelations}
-  \label{tab:nordea_15min_w_acf}
-\end{table}
-we get
-\begin{eqnarray*}
-  \theta &=& 0.6890 \\
-  \Theta &=& 0.6378
-\end{eqnarray*}
-Among the two roots of each of the 2nd order equations in the above,
-we have chosen the one in the range $(-1, 1)$ so as to ensure
-invertibility of the model \cite{BoxJenkins94}.
-
-By inverting the seasonal moving average model of $w_t$ we can then
-infer the noise process $y_t$ \footnotemark, i.e. the residuals:
-\begin{eqnarray*}
-  y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
-  &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
-  w_t \\
-  &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
-  w_{t-k}
-\end{eqnarray*}
-where $\bot(\cdot)$ denotes the floor function, i.e. the largest
-integer smaller than its argument.
-
-In order to forecst the $w_t$ process, and hence the price process
-itself, we must also know the distribution of $y_t$. Moreover, to
-properly estimate the parameters of the model in the sense of maximum
-likelihood, we are also in need of the distribution of $y_t$.
-
-If $y_t$ is indeed i.i.d random variables as we assumed, we only need
-to work out its unconditional distribution.
-
-\footnotetext{
-  The autoregressive process resulting from the inversion is inifinite
-  in extent. Hence one must choose a cut-off value for k. For the
-  present case of Nordea 15min log-volatility, I choose the value
-  $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
-  $k_{\top}$, is around 550. To avoid data reduction of this amount,
-  we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
-  corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $E(w_t) = E(y_t
-  - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
-}
+Figure \ref{fig:garch_eigenvec_PR_dist} shows the PDF of the
+normalized PR. We see that as T increases, the distribution of PR is
+compressed towards 0, suggesting increased localization of the
+eigenvectors.
 \begin{figure}[htb!]
   \centering
-  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_logvol_acf.pdf}
-    \label{fig:nordea_15min_logvol_acf}
+  \includegraphics[scale=0.5, clip=true, trim=49 208 70
+  162]{../pics/garch_eigenvec_PR.pdf}
+  \caption{\small \it Normalized participation ratio (PR)
+    versus eigenvalue.}.
+  \label{fig:garch_eigenvec_PR}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=97 259 113
+  226]{../pics/garch_eigenvec_PR_dist.pdf}
+  \caption{\small \it PDF of the normalized PR. N is fixed to
+    50.}
+  \label{fig:garch_eigenvec_PR_dist}
+\end{figure}
+In conclusion, localization of the eigenvectors, which implies 
+coincident eigenvalue and diagonal elements' distributions, begins with
+those associated to large eigenvalues. Increased observation points
+lead to increased localization and hence increased coincident sections
+of the eigenvalue and diagonal elements' distributions. However, this
+increment with T is slow, because the diagonal elements mean $\mu_D$
+increases only as a fractional power of $T$, namely $T^{1 -
+  2/\alpha}$, and the non-diagonal elements' variance decreases only
+as a fractional power too, namely $T^{1/\alpha' - 2/\alpha}$. These
+have been detailed in equations \ref{eq:garch_wishart_cij_params} and
+\ref{eq:garch_cij2}.
+
+\section{Influence of auto-correlations}
+\label{sec:garch_nonzero_phi}
+In the previous two sections we have studied situations where
+$\phi=0$ in the specification \ref{eq:garch_spec}, i.e. no
+auto-correlation is in the returns. In this section we investigate how
+auto-correlations change the picture.
+
+% When auto-correlations are present among the returns, the
+% cross-correlation matrix is expected to change. How exactly it changes
+% is the subject of this section. Here we consider a model specified
+% as follows:
+% \begin{eqnarray*}
+%   r_{it} &=& \phi r_{i, t-1} + \epsilon_{it} \\
+%   \epsilon_{it} &=& \sigma_{it} z_{it} \\
+%   \sigma_{it}^2 &=& \alpha_0 + \alpha_1 \epsilon_{i, t-1}^2 + \beta_1
+%   \sigma_{i, t-1}^2 \\
+%   z_{it} &\sim& N(0, 1)
+% \end{eqnarray*}
+% where $r_{it}$ is the element of the R matrix at the $i$-th row and the
+% $t$-th column. In the simulations described below R has N=50 rows and
+% $T = 8 \times 10^4$ columns. The parameters $\alpha_1$ and $\beta_1$
+% are the same as in the previous case of zero auto-correlations, namely
+% $\alpha_1 = 0.15$ and $\beta_1 = 0.84$. The cross-correlation matrix C
+% is built from R using
+% \begin{eqnarray*}
+%   C &=& {1 \over T^{2/\alpha}} RR'
+% \end{eqnarray*}
+% where $\alpha$ is 2.96 as before.
+
+Figure \ref{fig:GarchEigDiag1} shows the eigenvalue as well as the
+diagonal elements' distribution when $\phi = 0.95$, i.e. $\tau =
+13.51$. The values of N and T are 50 and $8\times 10^4$ as before. The
+GARCH(1,1) parameters are also unchanged, namely $\alpha_0 = 2.3\times
+10^{-6}$, $\alpha_1 = 0.15$, and $\beta_1 = 0.84$. Figure
+\ref{fig:GarchNondiag1} shows the non-diagonal elements' distribution
+in the same setup. Included in these plots are $50 \times 2000 =
+1\times 10^5$ eigenvalues and diagonal elements, as well as ${50
+  \choose 2} \times 2000 = 2,450,000$ non-diagonal elements. These
+data come from 2000 simulated matrices.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.42, clip=true, trim=112 272 103
+    217]{../pics/GarchEigDiag1.pdf}
+    \label{fig:GarchEigDiag1}
   }
-  \subfigure[ACF of $w_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_w_acf.pdf}
-    \label{fig:nordea_15min_w_acf}
+  \subfigure[]{
+    \includegraphics[scale=0.42, clip=true, trim=105 272 101
+    213]{../pics/GarchNondiag1.pdf}
+    \label{fig:GarchNondiag1}
   }
-  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$ autocorrelations}
+  \caption{\small \it \ref{fig:GarchEigDiag1}: Eigenvalues' and
+    diagonal elements' distribution when $\phi = 0.95$, i.e. $\tau$ =
+    13.51; \ref{fig:GarchNondiag1}: Non-diagonal elements'
+    distribution in the same situation.}
 \end{figure}
-\subsection{45min Returns}
+
+From figure \ref{fig:GarchEigDiag1} we see that, as auto-correlations
+become significant, the distribution of the eigenvalues no longer
+coincides with the diagonal elements' distribution --- instead it
+becomes wider and fatter on the tails. We also notice that the widths
+of both the diagonal and the non-diagonal elements' PDF's have
+increased.  In figure \ref{fig:garch_ev_cii} we see that, when no
+auto-correlation is present, the PDF of the diagonal elements has
+width ($\gamma$) $7.31 \times 10^{-4}$, while in figure
+\ref{fig:GarchEigDiag1} we see that the width has become $7.91 \times
+10^{-3}$ as $\tau$ becomes 13.51. Similarly the non-diagonal elements'
+PDF has width $2.14 \times 10^{-5}$ when $\tau=0$, as shown in figure
+\ref{fig:garch_cij}, and this width becomes $9.60 \times 10^{-4}$ when
+$\tau = 13.51$, as shown in figure \ref{fig:GarchNondiag1}.
+
+Figure \ref{fig:GarchSpectrumAutocorrelated} shows the eigenvalues'
+distribution corresponding to a range of $\phi$ values. The number of
+eigenvalues in each curve is the same as in figure
+\ref{fig:GarchEigDiag1}.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=85 258 103
+  229]{../pics/GarchSpectrumAutocorrelated.pdf}
+  \caption{\small \it Eigenvalue distribution of cross-correlation
+    matrix built from auto-correlated GARCH processes. $\phi$ = 0,
+    0.5, 0.8, 0.955, 0.97 correspond to correlation time $\tau$ =
+    0, 1.00, 3.11, 15.05, 22.76.}
+  \label{fig:GarchSpectrumAutocorrelated}
+\end{figure}
+From this figure one can see that, as auto-correlation strengthens,
+\begin{itemize}
+\item the PDF of the eigenvalue distribution flattens and widens;
+\item the minimum as well as the maximum eigenvalues increase.
+\end{itemize}
+
+To find out more about this series of deformation, we first look at
+how the largest component and the normalized participation ratio of
+the eigenvectors change as $\phi$ takes on larger values. Figure
+\ref{fig:garch_eigenvec_Cmax_corr} shows the largest eigenvector
+component $|c|_\M$ in correspondence to the eigenvalue. Apparently, as
+auto-correlation strengthens, the eigenvectors' composition
+fractures, leading to a reduced degree of localization and even
+reduced certainty of localization --- for a fixed eigenvalue, $|c|_\M$
+now varies in a larger range than it does with smaller $\phi$.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=0 197 0
+  154]{../pics/garch_eigenvec_Cmax_corr.pdf}
+  \caption{\small \it Eigenvectors' largest component in
+    correspondence to the eigenvalues. N is fixed to 50.}
+  \label{fig:garch_eigenvec_Cmax_corr}
+\end{figure}
+
+The same story of reduced localization is also evident from the plot
+of the normalized participation ratio (PR), shown in figure
+\ref{fig:garch_eigenvec_PR_dist_corr}, and from the PDF of $|c|_\m$
+shown in figure \ref{fig:garch_eigenvec_Cmax_dist_corr}. It is seen
+in \ref{fig:garch_eigenvec_PR_dist_corr} that the peak at the left 
+of the plot, representing the group of localized eigenvectors, falls
+with increased auto-correlation, and essentially disappears when
+$\phi$ reaches the extreme value 0.99. Figure
+\ref{fig:garch_eigenvec_Cmax_dist_corr} shows the proportion of large
+$|c|_\M$ values is severely reduced and the mean of $|c|_\M$ is pushed
+to smaller values by increased auto-correlation.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=95 260 111
+    232]{../pics/garch_eigenvec_PR_dist_corr.pdf}
+    \label{fig:garch_eigenvec_PR_dist_corr}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=95 260 111
+    232]{../pics/garch_eigenvec_Cmax_dist_corr.pdf}
+    \label{fig:garch_eigenvec_Cmax_dist_corr}
+  }
+  \caption{\small \it \ref{fig:garch_eigenvec_PR_dist_corr}: PDF of
+    the normalized participation ratio
+    (PR). \ref{fig:garch_eigenvec_Cmax_dist_corr}: PDF of the largest
+    component of the eigenvectors. N is fixed to 50. $\phi$ values of
+    0, 0.6000, 0.9000, 0.9900 correspond to correlation time $\tau$ =
+    0, 1.3569, 6.5788, 68.9676}
+\end{figure}
+
+It is also useful to look at how the fraction of localized eigenvectors
+changes with the auto-correlation. For definiteness, we classify
+an eigenvector as being localized when (1) its number of participating
+basis vectors is less than 2, or (2) the largest of its components'
+absolute values is larger than 0.9.
+
+Figure \ref{fig:localization_ratio} shows how the ratio of localized
+eigenvectors depends on the auto-correlation strength $\phi$. In
+either way of classification, the ratio falls with $\phi$ in
+accordance with a power law, the power exponent lying a bit below 2.
+This is further confirmed in plot \ref{fig:localization_ratio2}, where
+the ratios are plotted versus $\phi$ on log-log scale.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.38, clip=true, trim=93 229 115
+    134]{../pics/localization_ratio.pdf}
+    \label{fig:localization_ratio}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.38, clip=true, trim=87 227 115
+    133]{../pics/localization_ratio2.pdf}
+    \label{fig:localization_ratio2}
+  }
+  \caption{\small \it \ref{fig:localization_ratio}: Ratio of localized
+    eigenvectors versus the auto-correlation strength $\phi$. Black
+    ``+'': ratio of localized eigenvectors as measured by the number
+    of participating basis vectors being lower than 2; Black ``x'':
+    ratio of localized eigenvectors as measured by the largest
+    component being larger than 0.9. Blue curve: quadradic function
+    fitted to ``+''. Green curve: quadratic function fitted to
+    ``x''. There are 27 data points in the plot, corresponding to 27
+    $\phi$ values: 0 to 0.8 with step size 0.05, and 0.9 to 0.99 with
+    step size 0.01. The corresponding values of the correlation time
+    $\tau$ range from 0 to 69. \ref{fig:localization_ratio2}: $\ln
+    (f_\M - f)$ is plotted versus $\ln \phi$, where $f$ stands for the
+    ratio of localized eigenvectors. The fitted curves are linear.}
+\end{figure}
+
+% \subsection{Correlation Matrix of a Factor Model}
+%% Result of normalizing by 1/T
+% In this section we consider the estimation of correlation matrix in
+% the text of a factor model: the observed return series
+% $\{r_{it}\}$ ($1 \leq i \leq N$, $1 \leq t \leq T$) are driven by a
+% number of unobserved and uncorrelated random variables $z_{it}$
+% (factors), each of which is an auto-correlated GARCH(1,1) process as
+% specified by equation \ref{eq:garch_spec}. Factor GARCH models have
+% been studied extensively in the literature. A good review is found in
+% \cite{Mikosch2009}. Here we are interested in the impact of
+% auto-correlations in $z_{it}$ on the cross-correlation matrix of
+% $r_{it}$.
+
+% To be specific we adopt the generalized orthogonal GARCH model by
+% Weide \cite{Weide2002}:
+% \begin{eqnarray*}
+%   \vec{r_t} &=& \mtx{W}\vec{z_t}
+% \end{eqnarray*}
+% where $\mtx{W}$ is an $N \times N$ invertible matrix; $\vec{z_t}$ and
+% $\vec{r_t}$ are $N \times 1$ column vectors. Moreover, to meet our
+% interest, $\vec{z_t}$ are assumed to be auto-correlated:
+% \begin{eqnarray*}
+%   \vec{z}_t &=& \phi \vec{z}_{t-1} + \vec{y}_t \\
+%   \vec{y}_t &=& \mtx{H}^y_t \vec{x}_t \\
+% \end{eqnarray*}
+% where $\vec{x_t} \sim N(\vec{0}, \mtx{I})$, and $\mtx{H}^y_t$,
+% the conditional covariance matrix of $\vec{y}_t$, is given by
+% \begin{eqnarray*}
+%   \mtx{H}^y_t &=& (\mtx{I} - \mtx A - \mtx B) + \mtx A \odot
+%   (\vec{r}_{t-1} \vec{r}'_{t-1}) + \mtx B
+%   \mtx{H}^y_{t-1}
+% \end{eqnarray*}
+% Here $\mtx A$ and $\mtx B$ are diagonal $N \times N$ matrix and
+% $\odot$ denotes element-wise multiplication. The form of the constant
+% term, $\mtx{I} - \mtx A - \mtx B$, ensures that the unconditional
+% variance of $\vec{y}_t$ is $\mtx I$. In the simulations described in
+% the rest of this section, we use the parametrization $\mtx A =
+% \alpha_1 \mtx I$, $\mtx B = \beta_1 \mtx I$, where $\alpha_1 = 0.15$,
+% $\beta_1 = 0.84$. For tractability, we fix $N=4$ and consider one
+% particular covariance matrix of $\vec{r}_{it}$:
+% \begin{eqnarray*}
+%   \mtx C &=& \text{cov}(\vec{r}_{it}) \\
+%   &=& \begin{pmatrix}
+%     1 & \rho & \rho & \rho \\
+%     \rho & 1 & \rho & \rho \\
+%     \rho & \rho & 1 & \rho \\
+%     \rho & \rho & \rho & 1 \\
+%   \end{pmatrix}
+% \end{eqnarray*}
+% where $\rho = 0.5$. The eigenvalue decomposition of $\mtx C$ is
+% \begin{eqnarray*}
+%   \mtx C &=& \mtx{U} \mtx{\Lambda} \mtx{U}' \\
+%   \mtx U &=&
+%   \begin{pmatrix}
+%     -1/\sqrt{2} & -1/\sqrt{6} & -1/2\sqrt{3} & 1/2 \\
+%     1/\sqrt{2} & -1/\sqrt{6} & -1/2\sqrt{3} & 1/2 \\
+%     0 & \sqrt{2}/\sqrt{3} & -1/2\sqrt{3} & 1/2 \\
+%     0 & 0 & \sqrt{3}/2 & 1/2 \\
+%   \end{pmatrix} \\
+%   \mtx \Lambda &=&
+%   \begin{pmatrix}
+%     1-\rho & 0 & 0 & 0 \\
+%     0 & 1-\rho & 0 & 0 \\
+%     0 & 0 & 1-\rho & 0 \\
+%     0 & 0 & 0 & 1+ 3\rho \\
+%   \end{pmatrix}
+% \end{eqnarray*}
+
+%% Abandoned.
+% In the last section we have investigated what the measured
+% covariance matrix looks like when the return series have
+% auto-correlated GARCH(1,1) returns but are actually independent of
+% each other. In this section we investigate the situation when the
+% return series are truly correlated. In particular, we study the
+% following case:
+% \begin{equation}\label{eq:garch_correlated_returns}
+%   \begin{aligned}
+%     x_{1t} &= r_{1t} & \\
+%     x_{it} &= \sqrt{1 - \rho^2} r_{it} + \rho r_{1t} & \text{i = 2, 3,
+%       ..., N}\\
+%   \end{aligned}
+% \end{equation}
+% where $x_{it}$ are the return series whose covariance matrix is the
+% subject of the current study and $r_{it}$ are independent
+% auto-correlated GARCH(1,1) processes as specified by equation
+% \ref{eq:garch_spec}.
+
+% The covariance matrix C of the returns $\{x_{it}\}_{t=1}^T$ can be trivially
+% computed 
+% \begin{eqnarray}
+%   C_{ij} &=& {1 \over T^{2/\alpha}} \sum_{t=1}^T x_{it} x_{jt}
+%   \nonumber \\
+%   C &\xrightarrow{T \to \infty}&
+%   \begin{pmatrix}
+%     \sigma_1^2 & \rho \sigma_1^2 & \cdots & \rho \sigma_1^2 \\
+%     \rho \sigma_1^2 & (1 - \rho^2) \sigma_2^2 + \rho^2 \sigma_1^2 & \cdots &
+%     \rho^2 \sigma_1^2 \\
+%     \vdots & \vdots & \ddots & \vdots \\
+%     \rho \sigma_1^2 & \rho^2 \sigma_1^2 & \cdots & (1 - \rho^2)
+%     \sigma_N^2 + \rho^2 \sigma_1^2 
+%   \end{pmatrix} \label{eq:garch_cov}
+% \end{eqnarray}
+% where $\sigma_i$ is the standard deviation of the GARCH(1,1) series
+% $\{r_{it}\}_{t=1}^T$:
+% \begin{eqnarray*}
+%   \sigma_i = {1 \over T^{2/\alpha}} \sum_{t=1}^T r_{it}^2
+% \end{eqnarray*}
+
+% We would like to find out, when the covariance matrix of the $x$
+% series is estimated, how the auto-correlation in the $r$ series affects
+% the result --- specifically, how the auto-correlation strength $\phi$
+% (c.f. equation \ref{eq:garch_spec}) affects the distribution of
+% $C_{ij}$ when T is not very large.
+
+% Figure \ref{fig:garch_correlated_Cij} shows the distribution of
+% $C_{1j} / C_{11}$ ($j \geq 2$) and $C_{i1} / C_{11}$ ($i \geq
+% 2$). These rescaled elements, as seen from equation
+% \ref{eq:garch_cov}, have mean $\rho$. This is exactly what one sees in
+% figure \ref{fig:garch_correlated_Cij}. It is also clear from the
+% figure that $\phi$, the strength of auto-correlation, does not
+% introduce a bias but does increase the variance of the
+% estimation. This is the same conclusion as drawn in
+% \S\ref{sec:garch_nonzero_phi}.
+% \begin{figure}[htb!]
+%   \centering
+%     \includegraphics[scale=0.5, clip=true, trim=96 260 112
+%     227]{../pics/garch_correlated_Cij.pdf}
+%   \caption{\small \it Probability density function (PDF) of $C_{1j} /
+%     C_{11}$ ($j \geq 2$) and $C_{i1} / C_{11}$ ($i \geq 2$). The
+%     covariance matrix is constructed from 50 return series
+%     $\{x_{it}\}$, $i = 1, 2, ..., 50$, and $t=1, 2, ..., 600$, as
+%     specified by equation \ref{eq:garch_correlated_returns}. $\rho =
+%     0.5$. 4000 instances of the covariance matrix are generated for
+%     each value of $\phi$.}
+%   \label{fig:garch_correlated_Cij}
+% \end{figure}
+% One can also find support in the data that the aforementioned
+% distribution has power-law tails. This is shown in figure
+% \ref{fig:garch_mix_Cij_tail} where the power of the left
+% and the right tail is found to be 3.74 and 3.40, respectively.
+% \begin{figure}[htb!]
+%   \centering
+%     \includegraphics[scale=0.5, clip=true, trim=105 229 107
+%     120]{../pics/garch_mix_Cij_tail.pdf}
+%   \caption{\small \it Horizontal axis (x) : the elements $C_{1j} /
+%     C_{11}$ ($j \geq 2$) and $C_{i1} / C_{11}$ ($i \geq 2$) shifted to
+%     center around 0. The section with $0.15 < x < 0.35$ is
+%     plotted. Vertical axis: See the legend. }
+%   \label{fig:garch_mix_Cij_tail}
+% \end{figure}
+
+% It is also interesting to see how the eigenvalues and eigenvectors
+% behave given the non-trivial cross-correlation and the
+% auto-correlation in the returns. However, the cross-correlation matrix
+% of equation \ref{eq:garch_cov} has rather complicated eigenvalues and
+% eigenvectors. It has been useful in studying the non-diagonal matrix
+% elements but becomes clumsy with regard to eigenvalues and
+% eigenvectors. Therefore we look at the following return series
+% instead:
+% \begin{eqnarray}
+%   \begin{pmatrix}
+%     x_{1t} \\
+%     x_{2t} \\
+%   \end{pmatrix} &=&
+%   \begin{pmatrix}
+%     \cos\theta/\sigma_1 & -\sqrt{2} \sin\theta/\sigma_2  \\
+%      \sin\theta/\sigma_1 & \sqrt{2} \cos\theta/\sigma_2 \\
+%   \end{pmatrix}
+%   \begin{pmatrix}
+%     r_{1t} \\
+%     r_{2t} \\
+%   \end{pmatrix} \label{eq:x-r_transformation}
+% \end{eqnarray}
+% where $r_{it}$ are independent, auto-correlated GARCH(1,1) processes
+% as specified by equation \ref{eq:garch_spec}. Each of them has
+% standard deviation $\sigma_i$ ($i = 1, 2$). It is easy to
+% find the covariance matrix C of the series $\{x_{it}\}_{t=1}^T$ ($i =
+% 1,2$):
+% \begin{eqnarray*}
+%   C &=& 
+%   \begin{pmatrix}
+%     1 + \sin\theta^2 & -\sin\theta\cos\theta \\
+%     -\sin\theta\cos\theta & 1 + \cos\theta^2 \\
+%   \end{pmatrix}
+% \end{eqnarray*}
+% The eigenvalues and eigenvectors of $C$ are
+% \begin{eqnarray*}
+%   CX &=& X
+%   \begin{pmatrix}
+%     1 & \\
+%     & 2 \\
+%   \end{pmatrix} \\
+%   X &=& 
+%   \begin{pmatrix}
+%     \cos\theta & -\sin\theta \\
+%     \sin\theta & \cos\theta \\
+%   \end{pmatrix}
+% \end{eqnarray*}
+
+% Figure \ref{fig:garch_rotated_Cii} shows the probability plot of the
+% diagonal elements of $C$. Evidently, their distribution is very close
+% to a Gaussian, although the Anderson-Darling test rejects Gaussianity
+% at the 5\% significance level. Through equation
+% \ref{eq:x-r_transformation} the differences in variance of the
+% realizations of $\{r_{1t}\}_{t=1}^T$ and $\{r_{2t}\}_{t=1}^T$ are not
+% carried over to $\{x_{1t}\}_{t=1}^T$ and $\{x_{2t}\}_{t=1}^T$, and
+% therefore the differences in variance of the latter series are simply
+% the results of numerical computation and hence expected to be small
+% and Gaussian distributed. One can also see from figure
+% \ref{fig:garch_rotated_Cii} that the different auto-correlations
+% (different $\phi$ values) in the $\{r_{1t}\}_{t=1}^T$ and
+% $\{r_{2t}\}_{t=1}^T$ series do not change the Gaussianity. The
+% variances of these Gaussian distributions, however, are indeed
+% increased by increased auto-correlation.
+% \begin{figure}[htb!]
+%   \centering
+%     \includegraphics[scale=0.4, clip=true, trim=24 152 44
+%     18]{../pics/garch_rotated_Cii.pdf}
+%   \caption{\small \it Probability plot of $C_{11}$ corresponding to
+%     different auto-correlation strength ($\phi$) in the
+%     $r$-series. When $\phi = 0, 0.2, 0.4, 0.6$, the variance is $3.380
+%     \times 10^{-4}$, $3.639 \times 10^{-4}$, $4.512 \times 10^{-4}$,
+%     $7.506 \times 10^{-4}$, respectively.} 
+%   \label{fig:garch_rotated_Cii}
+% \end{figure}
+
+% On the other hand, the non-diagonal elements of $C$ have a very
+% different distribution from the diagonal ones, as shown in figure
+% \ref{fig:garch_rotated_Cij}.
+% \begin{figure}[htb!]
+%   \centering
+%     \includegraphics[scale=0.5, clip=true, trim=24 303 30
+%     274]{../pics/garch_rotated_Cij.pdf}
+%   \caption{\small \it Probability density function (PDF) of the
+%     non-diagonal elements of C. Corresponding to $\phi = 0, 0.2, 0.4,
+%     0.6, 0.8$, the variance of the distribution is
+%     $6.77\times10^{-8}$, $1.529\times10^{-7}$, $3.896\times10^{-7}$,
+%     $1.124\times10^{-6}$, $5.445\times10^{-6}$.}
+%   \label{fig:garch_rotated_Cij}
+% \end{figure}
+% One can see that the distribution of the non-diagonal elements is
+% left-skewed, regardless of the auto-correlation $\phi$. This is to be
+% compared with the distributions shown in figure
+% \ref{fig:garch_correlated_Cij}, where the skewness is small and
+% changes sign for different $\phi$ values, whereas in figure
+% \ref{fig:garch_rotated_Cij}, the skewness is consistently negative and
+% takes comparable values as $\phi$ changes from 0 to 0.9. We list these
+% skewness values in table \ref{tab:garch_rotated_var_skw}.
+% \begin{table}[htb!]
+%   \centering
+%   \begin{tabular}{|c|c|c|c|c|c|}
+%     \hline
+%     $\phi$ & 0 & 0.1 & 0.2 & 0.3 & 0.4 \\
+%     \hline 
+%     variance& 6.770e-08 & 1.008e-07 & 1.529e-07 & 2.339e-07 & 3.896e-07 \\
+%     \hline
+%     skewness & -1.62 & -1.65 & -1.96 & -1.86 & -1.67 \\
+%     \hline
+%   \end{tabular}
+%   \begin{tabular}{|c|c|c|c|c|c|}
+%     \hline
+%     $\phi$ & 0.5 & 0.6 & 0.7 & 0.8 & 0.9 \\
+%     \hline
+%     variance& 6.022e-07 & 1.124e-06 & 2.156e-06 & 5.445e-06 & 2.389e-05 \\
+%     \hline
+%     skewness & -1.62 & -1.65 & -1.96 & -1.86 & -1.67 \\
+%     \hline
+%   \end{tabular}
+%   \caption{\small \it Variance and skewness of $C_{12}$ and $C_{21}$}
+%   \label{tab:garch_rotated_var_skw}
+% \end{table}
+% From table \ref{tab:garch_rotated_var_skw} and figure
+% \ref{fig:garch_rotated_Cij} one can also see that the
+% variance steadily increases as $\phi$ increases --- a similar
+% observation that has been seen in figure
+% \ref{fig:garch_correlated_Cij}.
+
+% \begin{figure}[htb!]
+%   \centering
+%   \subfigure[]{
+%     \includegraphics[scale=0.35, clip=true, trim=107 278 104
+%     240]{../pics/garch_rotated_eig1_dist.pdf}
+%     \label{fig:garch_rotated_eig1_dist}
+%   }
+%   \subfigure[]{
+%     \includegraphics[scale=0.35, clip=true, trim=103 270 99 
+%     227]{../pics/garch_rotated_eig2_dist.pdf}
+%     \label{fig:garch_rotated_eig2_dist}
+%   }
+%   \caption{\small \it Eigenvalue distribution}
+%   \label{fig:garch_rotated_eig_dist}
+% \end{figure}
+
+% \begin{table}[htb!]
+%   \centering
+%   \begin{tabular}{|c|c|c|c|c|c|}
+%     \hline
+%     $\phi$ & 0 & 0.2 & 0.4 & 0.6 & 0.8 \\
+%     \hline
+%     variance& 3.338e-07 & 4.113e-07 & 8.415e-07 & 2.351e-06 & 9.580e-06 \\
+%     \hline
+%     skewness & -3.92 & -1.65 & -1.22 & -1.38 & -0.90 \\
+%     \hline
+%   \end{tabular}
+%   \caption{variance and skewness of the distribution of $\lambda_1$}
+%   \label{tab:garch_rotated_eig1_var_skw}
+% \end{table}
+
+% \begin{table}[htb!]
+%   \centering
+%   \begin{tabular}{|c|c|c|c|c|c|}
+%     \hline
+%     $\phi$ & 0 & 0.2 & 0.4 & 0.6 & 0.8 \\
+%     \hline
+%     variance& 4.947e-07 & 7.852e-07 & 1.767e-06 & 4.813e-06 & 2.258e-05 \\
+%     \hline
+%     skewness & 2.93 & 2.14 & 2.41 & 2.42 & 2.49 \\
+%     \hline
+%   \end{tabular}
+%   \caption{variance and skewness of the distribution of $\lambda_2$}
+%   \label{tab:garch_rotated_eig2_var_skw}
+% \end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% To find out more about this series of deformation, we investigate 
+% how the characteristic function changes in response to increasing
+% auto-correlation (increasing $\phi$). When $\phi = 0$ and T is very
+% large, like $8 \times 10^4$ in our simulation, we have found the
+% eigenvalue distribution converges to an $\alpha$-stable distribution
+% whose characteristic function is known in analytic form.
+
+% The empirical characteristic function $\varphi(k)$ is computed from
+% the simulated eigenvalues according to:
+% \begin{eqnarray*}
+%   \varphi(k) &=& \E(e^{ik\lambda}) \\
+%   &=& {1 \over M}\sum_{n=1}^M e^{ik\lambda_n}
+% \end{eqnarray*}
+% where $M$, the total number of eigenvalues in the sample, is $1\times
+% 10^5$ in our simulation, as mentioned earlier.
+
+% \begin{eqnarray}\label{eq:PDF_fourier_expansion}
+%   p(\lambda) &=& \sum_{n=-\infty}^\infty c_n \exp\left(
+%     -i {2\pi n \lambda \over L}
+%   \right) \\ \nonumber
+%   &\approx& \sum_{n=-M}^M c_n \exp\left(
+%     -i {2\pi n \lambda \over L}
+%   \right) \\
+%   L &=& \lambda_\M - \lambda_\m \nonumber
+% \end{eqnarray}
+% where $M$ is a large integer at which the series expansion is
+% truncated. For the simulations described in this section, we choose M
+% = 500.
+
+% The Fourier coefficients $c_n$ used in the above formula are computed
+% as
+% \begin{eqnarray*}
+%   c_n &=& {1 \over L} \int_{\lambda_\m}^{\lambda_\M} \exp\left(
+%     i {2\pi n \lambda \over L}
+%   \right) p(\lambda) d\lambda \\
+%   &=& {1 \over L} \E \left[
+%     \exp\left(
+%       i {2\pi n \lambda \over L}
+%     \right)
+%   \right] \\
+%   &=& {1 \over LS} \sum_{k=1}^S \exp\left(
+%     i {2\pi n \lambda_k \over L}
+%   \right)
+% \end{eqnarray*}
+% where $S = 1 \times 10^5$ is the aforementioned sample size of the
+% simulated eigenvalue. Figure \ref{fig:GarchEigPDFFourierCoef}
+% shows the Fourier coefficients computed according to the above
+% formula.
+% \begin{figure}[htb!]
+%   \centering
+%   \includegraphics[scale=0.56, clip=true, trim=10 203 6
+%   89]{../pics/GarchEigPDFFourierCoef.pdf}
+%   \caption{\small \it Fourier Coefficients of the empirical PDF of
+%     the Eigenvalue distribution. Blue/Red: Real/imaginary parts of the
+%     Fourier coefficients. The $\phi$ values 0, 0.2, 0.5, 0.8, 0.95, 0.96,
+%     0.97, 0.98 correspond to correlation time $\tau$ = 0, 0.43, 1.00,
+%     3.11, 13.51, 16.98, 22.76, 34.31.}
+%   \label{fig:GarchEigPDFFourierCoef}
+% \end{figure}
+
+% Before we proceed to investigate the properties of the empirical PDF
+% using the Fourier series expansion, we need to ensure the expansion
+% does provide a sufficiently accurate approximation to the empirical
+% PDF. However, to directly compare the Fourier series expansion
+% with the empirical PDF, the empirical PDF will have to be evaluated
+% --- the accuracy of the result will then largely depend on the number
+% of bins to which one chooses to sort the eigenvalues so as to produce
+% the empirical PDF --- a choice that is rather arbitrary and
+% subjective. To avoid such a procedure, we can compare the empirical
+% CDF with its Fourier expansion instead. Formally we integrate equation
+% \ref{eq:PDF_fourier_expansion} to obtain
+% \begin{equation}
+%   \label{eq:CDF_fourier_expansion}
+%   \begin{aligned}
+%     F(\lambda) &= \int_{\lambda_\m}^{\lambda} p(x) dx \\
+%     &= \sum_{n \in
+%       \mathbb{Z}\backslash\{0\}} c_n i {L \over 2\pi n} \left[
+%       \exp\left(
+%         -i {2\pi n \lambda \over L}
+%       \right) - 
+%       \exp\left(
+%         -i {2\pi n \lambda_\m \over L}
+%       \right) \right] \\
+%     & + (\lambda - \lambda_\m) c_0
+%   \end{aligned}
+% \end{equation}
+% The CDF on the left-hand-side of equation
+% \ref{eq:CDF_fourier_expansion} can be evaluated directly using the
+% simulated eigenvalues; meanwhile, the right-hand-side can be evaluated
+% for any $\lambda$ using $c_n$.
+
+% Figure \ref{fig:CDF_fourier_expansion} shows how the two sides of
+% equation \ref{eq:CDF_fourier_expansion} compare to each other.
+% \begin{figure}[htb!]
+%   \centering
+%   \includegraphics[scale=0.6, clip=true, trim=20 240 15
+%   200]{../pics/garch_cdf_fourier_expansion.pdf}
+%   \caption{\small \it The empirical cummulative distribution
+%     function (CDF) and its Fourier series expansion. Blue: empirical
+%     CDF. Red: Fourier series expansion. The plots are on log-log
+%     scale. The $\phi$ values 0, 0.2, 0.5, 0.8, 0.95, 0.96,
+%     0.97, 0.98 correspond to correlation time $\tau$ = 0, 0.43, 1.00,
+%     3.11, 13.51, 16.98, 22.76, 34.31.}
+%   \label{fig:CDF_fourier_expansion}
+% \end{figure}
+% We see that the two curves match fairly well --- they differ from each
+% other only at their lower ends, which are statistically insignificant
+% due to the very small portion of data points in that region.
+
+% Now that one is assured that the Fourier series expansion is
+% sufficiently accurate, it makes sense to study the properties of the
+% Fourier coefficients $c_n$. A few features immediately present
+% themselves in figure \ref{fig:GarchEigPDFFourierCoef}:
+% \begin{itemize}
+% \item $\re c_n = \re c_{-n}$, where $\re$ denotes the real part of a
+%   complex number;
+% \item $\re c_n$ resembles a dumped cosine wave;
+% \item $\re c_n$ and $\im c_n$ have the same frequency with a constant
+%   phase difference. Here $\im$ denotes the imaginary part of $c_n$.
+% \end{itemize}
+% The last conjecture about the frequency and the phase difference can
+% be verified by looking at the phase plot of $c_n$, which is shown in
+% figure \ref{fig:garch_Cn_phase}.
+% \begin{figure}[htb!]
+%   \centering
+%   \includegraphics[scale=0.6, clip=true, trim=21 248 14
+%   198]{../pics/garch_Cn_phase.pdf}
+%   \caption{\small \it Phases of the Fourier coefficients ($c_n$) of
+%     the empirical eigenvalue PDF. Horizonal axis: $n$; Vertical Axis:
+%     phases of $c_n$. The $\phi$ values 0, 0.2, 0.5, 0.8, 0.95, 0.96,
+%     0.97, 0.98 correspond to correlation time $\tau$ = 0, 0.43, 1.00,
+%     3.11, 13.51, 16.98, 22.76, 34.31.}
+%   \label{fig:garch_Cn_phase}
+% \end{figure}
+% From figure \ref{fig:garch_Cn_phase} one can see that the phase of
+% $c_n$ ($\ph c_n$) is approximately linear when $\phi$ is
+% small --- however, as $\phi$ becomes larger and larger, the
+% non-linearity of $\ph c_n$ becomes stronger and stronger.
+
+% In the relatively simple situation of small $\phi$, i.e. weak
+% auto-correlation, it lends some insight to consider the following
+% approximate formula:
+% \begin{eqnarray}
+%   \re c_n &=& e^{
+%     -u_r(\phi) - v_r(\phi) |n|
+%   } \cos[\omega(\phi) n] \label{eq:garch_fourier_coef_real}\\
+%   \im c_n &=& e^{
+%     -u_i(\phi) - v_i(\phi) |n|
+%   } \sin[\omega(\phi) n] \label{eq:garch_fourier_coef_imag}
+% \end{eqnarray}
+% Apparently, $\omega(\phi)$ can be directly read out from figure
+% \ref{fig:garch_Cn_phase}. The results are listed in table
+% \ref{tab:garch_phase_velocity}.
+% \begin{table}[htb!]
+%   \footnotesize
+%   \centering
+%   \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+%     \hline
+%     $\phi$ & 0 & 0.2 & 0.5 & 0.8 & 0.95 & 0.96 & 0.97 & 0.98 \\
+%     \hline
+%     $\tau$ & 0 & 0.4307 & 1.0000 & 3.1063 &  13.5134 & 16.9797 &
+%     22.7566 & 34.3096 \\
+%     \hline
+%     $\omega(\phi)$ & 0.0442 & 0.0538 & 0.0348 & 0.0414 & 0.0180 &
+%     0.0320 & 0.0154 & 0.0242 \\
+%     \hline
+%   \end{tabular}
+%   \caption{\small \it Phase velocity $\omega$ of the Fourier
+%     coefficients of the eigenvalue PDF.}
+%   \label{tab:garch_phase_velocity}
+% \end{table}
+% Using these values as initial estimate, the parameters in equation
+% \ref{eq:garch_fourier_coef_real} and \ref{eq:garch_fourier_coef_imag},
+% namely $u_r(\phi)$, $v_r(\phi)$ and $v_i(\phi)$ can be estimated by
+% least-square-error methods. We use the Matlab function ``lsqcurvefit''
+% for this purpose and list the estimated parameter values in table
+% \ref{tab:garch_fourier_coef_real_param}.
+% \begin{table}[htb!]
+%   \footnotesize
+%   \centering
+%   \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+%     \hline
+%     $\phi$ & 0 & 0.1 & 0.2 & 0.3 & 0.4 & 0.5 & 0.6 & 0.7 & 0.8\\
+%     \hline
+%     $\tau$ & 0 & 0.3010 & 0.4307 & 0.5757 & 0.7565 & 1.0000 & 1.3569 &
+%     1.9434 & 3.1063 \\
+%     \hline
+%     $u_r(\phi)$ & 0.0938 & 0.0936 & -0.0812 & 0.1649 & 0.2627 & 0.6285 &
+%     0.5443 & 0.5720 & 1.1610 \\
+%     \hline
+%     $v_r(\phi)$ & 0.0044 & 0.0045 & 0.0057 & 0.0046 & 0.0045 & 0.0034 &
+%     0.0046 & 0.0059 & 0.0048 \\
+%     \hline
+%     $\omega(\phi)$ & 0.0420 & 0.0424 & 0.0511 & 0.0428 & 0.0421 &
+%     0.0333 & 0.0416 & 0.0497 & 0.0394 \\
+%     \hline
+%   \end{tabular}
+%   % \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
+%   %   \hline
+%   %   $\phi$ & 0 & 0.1 & 0.2 & 0.3 & 0.4 & 0.5 & 0.6 & 0.7 & 0.8 & 0.9 \\
+%   %   \hline
+%   %   $\tau$ & 0 & 0.3010 & 0.4307 & 0.5757 & 0.7565 & 1.0000 & 1.3569 &
+%   %   1.9434 & 3.1063 & 6.5788 \\
+%   %   \hline
+%   %   $u_r(\phi)$ & 0.0938 & 0.0936 & -0.0812 & 0.1649 & 0.2627 & 0.6285 &
+%   %   0.5443 & 0.5720 & 1.1610 & 1.2347 \\
+%   %   \hline
+%   %   $v_r(\phi)$ & 0.0044 & 0.0045 & 0.0057 & 0.0046 & 0.0045 & 0.0034 &
+%   %   0.0046 & 0.0059 & 0.0048 & 0.0098 \\
+%   %   \hline
+%   %   $\omega(\phi)$ & 0.0420 & 0.0424 & 0.0511 & 0.0428 & 0.0421 &
+%   %   0.0333 & 0.0416 & 0.0497 & 0.0394 & 0.0667 \\
+%   %   \hline
+%   % \end{tabular}
+%   \caption{\small \it Least-Square-Error estimate of parameter values of $\re
+%     c_n$}
+%   \label{tab:garch_fourier_coef_real_param}
+% \end{table}
+% As is said earlier, the equation \ref{eq:garch_fourier_coef_real} and
+% \ref{eq:garch_fourier_coef_imag} only apply in the cases of small
+% $\phi$, so the estimation has only been done for $\phi$ values up to
+% 0.8. To obtain better statistics about the relation between $\phi$ and
+% the parameters, a few more values of $\phi$ have been added. The
+% corresponding Fourier coefficients have been computed and validated
+% against the empirical CDF. Figure \ref{fig:garch_fourier_coef_real}
+% shows how the equation \ref{eq:garch_fourier_coef_real} fits to $\re
+% c_n$ with the estimated parameter values.
+% \begin{figure}[htb!]
+%   \centering
+%   \includegraphics[scale=0.56, clip=true, trim=10 197 11
+%   147]{../pics/garch_fourier_coef_real.pdf}
+%   \caption{\small \it A dumped cosine function fit to $\re
+%     c_n$. Blue: empirical Fourier coefficients; Green: Dumped cosine
+%     function.}
+%   \label{fig:garch_fourier_coef_real}
+% \end{figure}
+% We see that the two curves in figure \ref{fig:garch_fourier_coef_real}
+% match fairly well. Obvious deviations are found only in regions of
+% large n, where slight inaccuracy in $\omega(\phi)$ is magnified. This
+% confirms the validity of formula \ref{eq:garch_fourier_coef_real}.
+
+% % To find out the relation between $\phi$ and the parameters $u_r(\phi)$,
+% % $v_r(\phi)$ and $\omega(\phi)$, we plot these parameters' values
+% % against $\phi$ in figure \ref{fig:Cn_real_param}.
+% % \begin{figure}[htb!]
+% %   \centering
+% %   \includegraphics[scale=0.56, clip=true, trim=6 311 0
+% %   264]{../pics/Cn_real_param.pdf}
+% %   \caption{\small \it Linear fit of the parameters in $\re
+% %     c_n$. Blue circles: parameter values normalized by the sample
+% %     mean; Green line: straight line fitted to the data points.}
+% %   \label{fig:Cn_real_param}
+% % \end{figure}
+% % Instead of directly plotting the parameters' values, figure
+% % \ref{fig:Cn_real_param} plots the ratio of these values over their
+% % respective sample mean, so that one can see how much each
+% % individual data point (blue circles) deviates from the sample mean as
+% % well as from the fitted line. The title of each plot in figure
+% % \ref{fig:Cn_real_param} is the equation of the fitted line.
+
+% % From figure \ref{fig:Cn_real_param} we see that $u_r(\phi)$ is
+% % approximately linear in $\phi$ --- the coefficient of the linear term
+% % in its equation is an order of magnitude larger than the constant
+% % term. In contrast, $v_r(\phi)$ and $\omega(\phi)$ are approximately
+% % constant --- the coefficient of the linear term in their equation is
+% % an order of magnitude smaller than the constant term. So, also
+% % considering $0 < \phi < 1$ and that we have a rather small sample
+% % size, we estimate $v_r(\phi)$ and $\omega(\phi)$ to be constant. In
+% % summary
+% % \begin{eqnarray*}
+% %   u_r(\phi) &=& 1.2364 \phi - 0.1124 \\
+% %   v_r &=& 4.7148 \times 10^{-3} \\
+% %   \omega &=& 3.3476 \times 10^{-2} \\
+% % \end{eqnarray*}
+% % Now that $v_r(\phi)$ and $\omega(\phi)$ are constant, we shall write
+% % them only as $v_r$ and $\omega$ to avoid confusion.
+
+% Turning our attention to the imaginary part of $c_n$($\im c_n$) and
+% applying the same methods, we find the corresponding parameter values
+% for $\im c_n$ (table \ref{tab:garch_fourier_coef_imag_param}). Figure
+% \ref{fig:garch_fourier_coef_imag} shows how these values together with
+% equation \ref{eq:garch_fourier_coef_imag} fit to $\im c_n$. Comparing
+% table \ref{tab:garch_fourier_coef_real_param} and
+% \ref{tab:garch_fourier_coef_imag_param}, we see that $v_r$ and $v_i$
+% are essentially equal, while $u_r(\phi)$ is consistently larger than
+% $u_i(\phi)$.
+
+% \begin{table}[htb!]
+%   \footnotesize
+%   \centering
+%   \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
+%     \hline
+%     $\phi$ & 0 & 0.1 & 0.2 & 0.3 & 0.4 & 0.5 & 0.6 & 0.7 & 0.8\\
+%     \hline
+%     $\tau$ & 0 & 0.3010 & 0.4307 & 0.5757 & 0.7565 & 1.0000 & 1.3569 &
+%     1.9434 & 3.1063 \\
+%     \hline
+%     $u_i(\phi)$ & 0.0555 & 0.0565 & -0.1105 & 0.1291 & 0.2236 &
+%     0.6007 & 0.5031 & 0.5341 & 1.1273 \\
+%     \hline
+%     $v_i(\phi)$ & 0.0047 & 0.0047 & 0.0059 & 0.0048 & 0.0048 &
+%     0.0035 & 0.0049 & 0.0062 & 0.0049 \\
+%     \hline
+%     $\omega(\phi)$ & 0.0420 & 0.0424 & 0.0511 & 0.0428 & 0.0421 &
+%     0.0333 & 0.0416 & 0.0497 & 0.0394 \\
+%     \hline
+%   \end{tabular}
+%   \caption{\small \it Least-Square-Error estimate of parameter values of $\im c_n$}
+%   \label{tab:garch_fourier_coef_imag_param}
+% \end{table}
+
+% \begin{figure}[htb!]
+%   \centering
+%   \includegraphics[scale=0.56, clip=true, trim=18 221 19
+%   171]{../pics/garch_fourier_coef_imag.pdf}
+%   \caption{\small \it A dumped sine function fit to the imaginary
+%     part of $c_n$ ($\im c_n$). Red: empirical Fourier coefficients;
+%     Green: Dumped sine function.} 
+%   \label{fig:garch_fourier_coef_imag}
+% \end{figure}
+
+%  As is done for $\re c_n$, we can also find an approximate
+% linear expression for $u_i(\phi)$ by fitting a line to the data. The
+% result is
+% \begin{eqnarray*}
+%   u_i(\phi) &=& 1.2356 \phi - 0.1480
+% \end{eqnarray*}
+% So, combining the results for $\re c_n$ and $\im c_n$, we may write
+% \begin{eqnarray*}
+%   c_n &=& e^{-v|n|}\left[
+%     e^{-a_r\phi - b_r}\cos(\omega n) + ie^{-a_i\phi - b_i}\sin(\omega n)
+%   \right] \\
+% \end{eqnarray*}
+% where
+% \begin{equation}
+%   \label{eq:garch_Cn_parameters}
+%   \begin{aligned}
+%     v &= 4.7148 \times 10^{-3} \\
+%     a_r &= 1.2364 \\
+%     b_r &= -0.1124 \\
+%     a_i &= 1.2356 \\
+%     b_i &= -0.1480 \\
+%     \omega &= 3.3476 \times 10^{-2}
+%   \end{aligned}
+% \end{equation}
+% It should be emphasized that the above parameters are only constants
+% with respect to the GARCH parameters $\alpha_1$, $\beta_1$ and hence
+% the resulting tail exponent $\alpha$. If the GARCH parameters are
+% changed, the parameters in \ref{eq:garch_Cn_parameters} are expected
+% to follow.
+
+%% TODO:
+% 1. find the power-exponent of the tail of the SLV (stochastic
+% log-volatility) model. Normalize the returns using the sample
+% standard deviation when computing the cross-correlation matrix.
+% 
+% 2. Hows does the power-law return distribution change the eigenvalue
+% distribution in comparison to the Marcenko-Pastur law?
+%
+% 3. How do the maximum and the minimum eigenvalue change
+% corresponding to different values of autocorrelation?
+%
+% 4. Use different combinations of q = N/T and investigate how the
+% eigenvalue distribution changes in the case of the SLV model.
+
+\chapter{Summary of the Results}
+The results of this thesis project are the following:
+\begin{enumerate}
+\item ARIMA log-volatility model
+  \begin{enumerate}
+  \item ARIMA log-volatility returns' models are justified and tested
+    against Nordea Bank 15-minute returns and Volvo B 30-minute returns
+    during the period 2013/10/10 - 2014/04/04. Historical volatility's
+    are estimated using ``realized volatilities'', denoted
+    $\hat{\sigma}_t$ here:
+    \begin{eqnarray*}
+      \hat{\sigma}_t &=& \sqrt{
+        \sum_{i = 1}^n \left[\ln {p_{t - \Delta t + i \Delta t / n} \over
+            p_{t - \Delta t + (i-1) \Delta t / n}} \right]^2
+      }
+    \end{eqnarray*}
+    where $p_{t'}$ is the price of the asset at time $t'$, and $n$ is
+    the number of sub intervals into which the time interval $[t -
+    \Delta t, t]$ is divided. A discussion of the choice of $n$ is found
+    in \S\ref{sec:SLV_model}.
+
+    Using the first 80\% of the data for estimating the model and the
+    last 20\% for comparing with the model forecast, we find the ARIMA
+    models out-perform their GARCH competitors in most cases.
+
+  \item The unconditional distribution functions of ARIMA
+    log-volatility models are calculated (equation
+    \ref{eq:UncondPDFAsymmetric} and \ref{eq:UncondPDFAsymmetric1}):
+    The unconditional distribution of  the returns $r_t$ is the same as
+    $\mu + e^{\bar{v} + \sigma a} b$, where $(a, b) \sim N(0, \Sigma)$
+    \begin{eqnarray*}
+      \Sigma &=&
+      \begin{pmatrix}
+        1 & \psi \\
+        \psi & 1
+      \end{pmatrix}
+    \end{eqnarray*}
+    Then the PDF of the unconditional distribution $f_r(x)$ can be expressed as
+    \begin{eqnarray*}
+      f_{r'}(x) &=& {1 \over 2\pi \sqrt{1 - \psi^2}} \times \\
+      && \int_{-\infty}^\infty da \exp
+      \left[- {
+          a^2 + 2\sigma(1 - \psi^2)a - 2a\psi e^{-a\sigma} x + e^{-2a\sigma} x^2
+          \over
+          2(1 - \psi^2)
+        }
+      \right] \\
+      f_r(x) &=&  \td{r'}{r} f_{r'}[e^{-\bar{v}} (x - \mu)] \\
+      &=& e^{-\bar{v}} f_{r'}[e^{-\bar{v}} (x-\mu)] \\
+    \end{eqnarray*}
+    where $r' = e^{\sigma a}b$. The moment generating function of the
+    unconditional distribution is found as an integral, and all the
+    moments can be evaluated in closed form. See equations
+    \ref{eq:AsymmetricMoments}.
+
+  \end{enumerate}
+\item Correlation matrix of Gaussian returns. $N$ series of
+  auto-correlated, standard Gaussian distributed returns are simulated
+  to form $N\times N$ correlation matrix $C$:
+  \begin{eqnarray*}
+    r_{it} &=& \phi r_{i,t-1} + a_t \\
+    C_{ij} &=& {1 \over T} \sum_{t=1}^T r_{it} r_{jt}
+  \end{eqnarray*}
+  where $i = 1, 2, ..., N$ and $t = 1,2, ..., T$. $a_t \sim N(0,
+  1)$.
+  \begin{enumerate}
+  \item Distribution of the elements of C. The asymptotic distribution
+    of the matrix elements $C_{ii}$ and $C_{ij}$ ($i \neq j$) are 
+    Gaussian whose mean and variance are functions of $\phi$:
+    \begin{eqnarray*}
+      \E(C_{ij}) &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi^2)(1 -
+        \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
+        P^{-3/2}_{-1/2}(\rho)
+      \right] \\
+      \var(C_{ij}) &=& {2 \sigma^6 \over T (1 - \phi^2)^2} \left[
+        {\phi^2 \over 1 - \phi^2} -
+        {\phi^2 (1 - \phi^{2T}) \over
+          T(1 - \phi^2)}
+      \right] + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
+        T (1 - \phi^2)^2} \\
+      &\approx& {2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}
+      + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
+        T (1 - \phi^2)^2}
+    \end{eqnarray*}
+    where $\rho = \text{corr}(a_{it}, a_{jt})$, $i \neq
+    j$. $P^\mu_\nu(\cdot)$ is Ferrers function of the first kind (See
+    equation \ref{eq:Ferrers_1st}) and
+    \begin{eqnarray*}
+      v^2(\rho) &=&
+      {4 \over \sqrt{2\pi} (1 - \rho^2)^{7/4}} \left[
+        P^{-5/2}_{-1/2}(\rho) + P^{-5/2}_{-1/2}(-\rho)
+      \right] \\
+      && - {1 \over 2\pi (1 - \rho^2)^{5/2}} \left[
+        P^{-3/2}_{-1/2}(-\rho) - P^{-3/2}_{-1/2}(\rho)
+      \right]^2
+    \end{eqnarray*}
+    In comparison,
+    \begin{eqnarray*}
+      \E(C_{ii}) & \approx & {\sigma^2 \over 1 - \phi_1^2} \left[
+        1 - {\phi_1^2 \over T}
+      \right] \\
+      \var(C_{ii}) &\approx& {2 \sigma^4 \over T (1 - \phi_1^4)} +
+      {\sigma^6 \over T (1 -\phi_1^2)^2}
+    \end{eqnarray*}
+
+  \item Distribution of the maximum eigenvalue of C. It is shown that
+    the maximum eigenvalue has approximately a gamma distribution even
+    when auto-correlations are present among the returns. The mean of
+    the gamma distribution is approximately quadratic in $\tan{\pi\phi
+      \over 2}$.
+  \end{enumerate}
+
+\item Covariance matrix C of GARCH(1,1) processes. In the limit of
+  very long return series, the covariance matrix of properly
+  specified GARCH(1,1) processes (see chapter
+  \ref{chp:CrossCorrelationFat}) is 
+  shown to be diagonal. The distribution of its eigenvalues thus
+  coincides with that of its diagonal elements. Their distribution
+  converges to an $\alpha$-stable distribution. If the returns $r_{it}$
+  have tail behavior $P(|r_{it}| > x) \sim {c \over x^\alpha}$, the
+  asymptotic stable distribution has L\'evy index
+  $\alpha/2$. Moreover, the mean of this stable distribution scales
+  with the length of the return series (T) as $T^{1 - 2/\alpha}$,
+  while the width is independent of T.
+
+  In contrast, the distribution of the non-diagonal elements of the
+  covariance matrix converges to a symmetric, zero-centered, stable
+  distribution with a different L\'evy index than $\alpha$. Denote it
+  $\alpha'$, then the variance of this stable distribution scales with
+  T as $T^{1/\alpha' - 2/\alpha}$.
+
+  When the return series only have a modest length, the situation is
+  studied in terms of the localization of the eigenvectors. For a
+  given T, the largest eigenvalues correspond to the most localized
+  eigenvectors, and the extent of localization decreases as the
+  eigenvalue decreases. As T increases, the extent of localization
+  increases for all eigenvectors.
+
+  Auto-correlations among the returns, on the other hand, reduces the
+  extent of localization, and also the certainty of localization with
+  respect to the eigenvalues. In other words, if the return series
+  have considerable auto-correlations, the eigenvectors corresponding
+  to the largest eigenvalues will have a rather fractured composition.
+\end{enumerate}
+
+\chapter{Appendix}
+\section{PDF of the covariance matrix of auto-correlated Gaussian
+  Returns} \label{app:pdf_gaussian1}
+When autocorrelations exist among the columns of the Gaussian returns
+matrix $\mtx R$, the distribution of the covariance matrix $\mtx C =
+\mtx{R} \mtx{R'}$ is no longer Wishart but is nonethelss closely
+related to it. In the following we consider the situation where
+$r_{i,t}$ can be represented as a vector auto-regressive
+process. Specifically, suppose
+\begin{eqnarray*}
+  \vec{r}_t &=& \sum_{k=1}^p \phi_k \vec{r}_{t-k} + \vec{a}_t \\
+  \vec{a}_t &=& \vec{r}_t - \sum_{k=1}^p \phi_k \vec{r}_{t-k}
+\end{eqnarray*}
+where $\vec{a}_t = (a_{1,t}, a_{2,t}, \cdots, a_{N,t})' \sim N(0,
+\Sigma)$ and comprise the columes of $\mtx A$; $\vec{r}_t$ are the
+columes of $\mtx R$. Here $N(0, \Sigma)$ denotes the multivarate
+normal distribution with covariance matrix $\Sigma$. The last equation
+can be written in matrix form
+\[
+A = R M
+\]
+For example, in the case of an AR(1) process
+\begin{equation*}
+  \begin{pmatrix}
+    a_{1,1} & a_{1,2} & \cdots & a_{1,T} \\
+    \vdots & \ddots & \vdots \\
+    a_{N,1} & a_{N,2} & \cdots & a_{N,T} \\
+  \end{pmatrix} =
+  \begin{pmatrix}
+    r_{1,1} & r_{1,2} & \cdots & r_{1,T} \\
+    \vdots & \ddots & \vdots \\
+    r_{N,1} & r_{N,2} & \cdots & r_{N,T} \\
+  \end{pmatrix}
+  \begin{pmatrix}
+    1 & -\phi_1 &   &   & \\
+      & 1 & -\phi_1 &   & \\
+      &   & \ddots  & \ddots &   \\
+      &   &   & 1 & -\phi_1 \\
+      &   &   &   & 1 \\
+  \end{pmatrix}
+\end{equation*}
+Let $QR = RM$, then $Q = RMR^{-1}$. Since the set of $\{r_{ij}\}$ for
+which $\text{det } R = 0$ has probability zero, a matrix $Q$
+satisfying the above equation almost surely exists. Thus $A = RM =
+QR$ and $AA' = QRR'Q'$ Since the columns of A are not auto-correlated,
+$AA' \sim W(\Sigma, T)$. Then $RR' \sim W(Q^{-1} \Sigma Q'^{-1}, T)$
+follows from \cite{Anderson2003}, \S7.3.3.
+
+Now we observe that $\Sigma$ enters $f(RR')$ through $\tr (\Sigma^{-1}
+RR')$ and $\det \Sigma$. $RR'$ enters through $\tr (\Sigma^{-1}
+RR')$ and $\det (RR')$. Clearly
+\begin{eqnarray*}
+  \det (Q^{-1}\Sigma Q'^{-1}) &=& \det \Sigma \over (\det Q)^2 \\
+  &=& \det \Sigma \over (\det M)^2 \\
+  &=& \det \Sigma
+\end{eqnarray*}
+As to $\tr (\Sigma^{-1} C)$, we have
+\begin{eqnarray*}
+  && \tr\left[(Q^{-1}\Sigma Q'^{-1})^{-1} RR'\right] \\
+  &=& \tr\left[Q'\Sigma^{-1} Q RR'\right] \\
+  &=& \tr\left[ R'^{-1}M'R'\Sigma^{-1} RMR^{-1} RR'\right] \\
+  &=& \tr\left[ \Sigma^{-1} RM (RM)'\right] \\
+  &=& \tr\left[ \Sigma^{-1} AA'\right] \\
+\end{eqnarray*}
+
+Moreover, $\det RR' = \det AM^{-1} M'^{-1} A' = \det AA'$.
+
+Thus if we use $f_R(\cdot)$ to denote the joint probability density of
+the entries of RR' and $f_A(\cdot)$ to denote that of AA', we can
+write
+\begin{equation}\label{eq:cross-corr-matrix-PDF}
+  f_R(RR') = f_A(RMM'R')
+\end{equation}
+
+Substituting for $f_A$ the Wishart probability density function (PDF)
+\ref{eq:WishartPDF}, we get the PDF of $RR'$.
 

### 2014-04-30 LevyFT.m

@@ -1,0 +1,11 @@
+function y = LevyFT(k, a, b, g, m)
+if a >= 2
+    y = NaN;
+elseif a == 1
+    ly = -g .* abs(k) -i * b * g .* k .* (2/pi) .* log(abs(k))...
+         + i * m .* k;
+    y = exp(ly);
+else
+    y = exp(i*m*k - (g^a .* abs(k).^a) .* ...
+            + (1 - i .* b .* sign(k) .* tan(pi * a / 2)));
+end

### 2014-04-30 eigen_dist_levy.m

@@ -1,0 +1,376 @@
+clear all
+close all
+
+T = 8e+4;
+N = 50;
+fold = 2e3;
+
+%% eigen values generated from the volvo stochastic log-volatility
+%% model.
+
+spec = cellstr(['b  '; 'c  '; 'g  '; 'm  '; 'r  '; 'k  ';...
+                'b--'; 'c--'; 'g--'; 'm--'; 'r--'; 'k--';...
+                'b: '; 'c: '; 'g: '; 'm: '; 'r: '; 'k: ';...
+                'b-.'; 'c-.'; 'g-.'; 'm-.'; 'r-.'; 'k-.';...
+                ]);
+% for i = [1:5]
+%     load(sprintf('./eigen-%d.mat', i));
+%     ev = reshape(ev, N, fold);
+%     maxeig = max(ev);
+%     [y, x] = ecdf(maxeig);
+%     y1 = diff(y) ./ diff(x);
+%     x1 = (x(1:end-1) + x(2:end))/2;
+%     plot(log(x1), log(y1), spec(i));
+%     hold on
+% end
+% hold off
+% grid on
+
+
+%% White Wishart matrix eigen values
+% AR = [0:0.05:0.8, 0.9, 0.95:0.005:0.99];
+AR = 0:0.1:0.9;
+tau = -log(2)./log(AR);
+% dist = struct('name', 'Cauchy', 'prmt', 1);
+dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.15, 0.84], 'distr', ...
+              'Gaussian', 'TailExponent', 2.9664);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.1, 0.80], 'distr', ...
+%               struct('Name', 't', 'DoF', 3), 'TailExponent', 0.7744);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.1, 0.725], 'distr', ...
+%               struct('Name', 't', 'DoF', 3), 'TailExponent', 1.8187);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.15, 0.84], 'distr', ...
+%               struct('Name', 't', 'DoF', 60), 'TailExponent', 2.4477);
+% for j = AR
+%     if (exist(sprintf('GarchWishartT8e4Eig-%.3f.mat', j), 'file') == 2)
+%         continue;
[…]

### 2014-04-30 eigen_dist_levy2.m

@@ -1,0 +1,144 @@
+clear all
+close all
+
+fold = 1e+4;
+AR = [0:0.1:0.9];
+tau = -log(2)./log(AR);
+
+% N = 50;
+% rho = 0.5;
+% q = 12;
+
+N = 4;
+rho = 0;
+q = 1500;
+
+% N = 2;
+% theta = pi/4;
+% q = 3000;
+
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.15, 0.84], 'distr', ...
+%               'Gaussian', 'TailExponent', 2.9664);
+dist = struct('name', 'Garch1_1', 'prmt', [0.01, 0.15, 0.84], 'distr', ...
+              'Gaussian', 'TailExponent', 2.9664);
+
+spec = cellstr(['b  '; 'c  '; 'g  '; 'm  '; 'r  '; 'k  ';...
+                'b--'; 'c--'; 'g--'; 'm--'; 'r--'; 'k--';...
+                'b: '; 'c: '; 'g: '; 'm: '; 'r: '; 'k: ';...
+                'b-.'; 'c-.'; 'g-.'; 'm-.'; 'r-.'; 'k-.';...
+                ]);
+hold on
+c = 1;
+% for phi = 0
+% for phi = [0, 0.1, 0.3, 0.5, 0.6, 0.7, 0.9]
+for phi = [0]
+    %% The matrix elements 
+    % load(sprintf('GarchWishartProperN%dQ%dRho%.3fC-%.3f.mat', ...
+    %              N, q, rho, phi), 'C');
+    % x = NaN(1, 2*(N-1)*fold);
+    % for k = 1:fold
+    %     x((k-1)*2*(N-1)+1 : k*2*(N-1)) = ...
+    %         [C(1, 2:end, k), C(2:end, 1, k)'] ./ C(1, 1, k);
+    % end
+    % [u, v] = epdf(x, 4, 0, 1, 320, spec{c});
+    % param = stblfit(x(1:1e+5), 'ecf');
+    % v1 = stblpdf(u, param(1), param(2), param(3), param(4));
+    % plot(u, log10(v), u, log10(v1));
+
+    % load(sprintf('GarchWishartProperN%dQ%dRho%.3fC-%.3f.mat', ...
+    %              N, q, rho, phi), 'C');
+    % % C11 = reshape(C(1, 1, :), 1, fold);
[…]

### 2014-04-30 eigen_dist_levy_tmp.m

@@ -1,0 +1,262 @@
+clear all
+close all
+
+T = 8e+4;
+N = 50;
+fold = 2e3;
+
+q = N/T;
+
+%% eigen values generated from the volvo stochastic log-volatility
+%% model.
+
+spec = cellstr(['b  '; 'c  '; 'g  '; 'm  '; 'r  '; 'k  ';...
+                'b--'; 'c--'; 'g--'; 'm--'; 'r--'; 'k--';...
+                'b-.'; 'c-.'; 'g-.'; 'm-.'; 'r-.'; 'k-.';...
+                ]);
+% for i = [1:5]
+%     load(sprintf('./eigen-%d.mat', i));
+%     ev = reshape(ev, N, fold);
+%     maxeig = max(ev);
+%     [y, x] = ecdf(maxeig);
+%     y1 = diff(y) ./ diff(x);
+%     x1 = (x(1:end-1) + x(2:end))/2;
+%     plot(log(x1), log(y1), spec(i));
+%     hold on
+% end
+% hold off
+% grid on
+
+
+%% White Wishart matrix eigen values
+
+AR = 0:0.05:0.8;
+% AR = [0];
+tau = -log(2)./log(AR);
+% dist = struct('name', 'Cauchy', 'prmt', 1);
+dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.15, 0.84], 'distr', ...
+              'Gaussian', 'TailExponent', 2.9664);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.15, 0.84], 'distr', ...
+%               struct('Name', 't', 'DoF', 60), 'TailExponent', 2.4477);
+for j = AR
+    if (exist(sprintf('GarchWishartT8e4Eig-%.3f.mat', j), 'file') == 2)
+        continue;
+    end
+    ev = NaN(N, fold);
+    Cij = NaN(N*(N-1)/2, fold);
+    Cii = NaN(N, fold);
+    for i = 1:fold
+        R = gen_ret_mtx(N, T, dist, j);
+        if strcmp(dist.name, 'Cauchy') == 1
+            R = R ./ dist.prmt ./ T;
+        elseif strcmp(dist.name, 'Garch1_1') == 1
+            % R = diag(1./std(R')) * R;
+            R = R ./ T^(1/dist.TailExponent);
+        end
+        C = R*R';
[…]

### 2014-04-30 slv_cdf.m

@@ -1,0 +1,10 @@
+function y = slv_cdf(x, psi, sig)
+x = reshape(x , length(x), 1);
+da = 1.0e-2;
+a = -10:da:10;
+Ki = (ones(length(x), 1) * a) .* psi - x * exp(-sig .* a);
+Ki = Ki ./ sqrt(2 * (1 - psi^2));
+Ki = erfc(Ki);
+Kr = exp(-a.^2 ./ 2) .* da;
+
+y = Ki * Kr' ./ 2 ./ sqrt(2 * pi);

### 2014-04-30 test12.m

@@ -1,0 +1,86 @@
+clear all
+close all
+
+fold = 4e+3;
+% fold = 1;
+
+%AR = 0:0.1:0.9;
+AR = [0:0.2:0.8];
+% AR = 0;
+tau = -log(2)./log(AR);
+% dist = struct('name', 'Gaussian', 'TailExponent', 2);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.15, 0.84], 'distr', ...
+%               struct('Name', 'Gaussian'), 'TailExponent', 2.9664);
+dist = struct('name', 'Garch1_1', 'prmt', [0.01, 0.15, 0.84], 'distr', ...
+              struct('Name', 'Gaussian'), 'TailExponent', 2.9664);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.1, 0.725], 'distr', ...
+%               struct('Name', 't', 'DoF', 3), 'TailExponent',
+%               1.8187);
+%for q = [12, 16, 20, 24]
+N = 4;
+% theta = pi/4;
+rho = 0;
+C = eye(N);
+C(~logical(eye(N))) = rho;
+[U, D] = eig(C);
+W = U*sqrt(D);
+
+for q = 1500
+    T = N * q;
+    for j = 0
+        % if (exist(sprintf('GarchWishartN%dQ%dDiag-%.3f.mat', N, ...
+        %                   T, j), 'file') == 2)
+        %     continue;
+        % end
+        ev1 = NaN(N, fold);
+        X1 = NaN(N, N, fold);
+        C1 = NaN(N, N, fold);
+        for i = 1:fold
+            R = gen_ret_mtx(N, T, dist, j, eye(N));
+            % if strcmp(dist.name, 'Cauchy') == 1
+            %     R = R ./ dist.prmt ./ T;
+            % elseif strcmp(dist.name, 'Garch1_1') == 1
+            %     % R = R ./ T^(1/dist.TailExponent);
+            %     ;
+            % end
+            if rho == 0
+                % C1(:, :, i) = R*R' ./ T^(2/dist.TailExponent);
+                C1(:, :, i) = R*R' ./ T;
+            else
+                Z = W * R;
+                % C1(:, :, i) = Z*Z' ./ T^(2/dist.TailExponent);
+                C1(:, :, i) = Z*Z' ./ T;
+            end
+            [V, D] = eig(C1(:, :, i));
+            ev1(:, i) = D(logical(eye(N)));
+            X1(:, :, i) = V;
[…]

### 2014-05-05 thesis/thesis.tex
This is the version to be used for job searching.

--- 2013-12-13 thesis/thesis.tex
@@
 
-\maketitle
+\begin{titlepage}
+\begin{center}
+
+% Upper part of the page. The '~' is needed because \\
+% only works if a paragraph has started.
+\includegraphics[width=0.5\textwidth]{../pics/lund_uni-logo_s}~\\[1cm]
+
+% \textsc{\LARGE Lund University}\\[1.5cm]
+
+\textsc{\Large Master's Thesis Project}\\[0.5cm]
+
+% Title
+%\HRule \\[0.4cm]
+{ \huge \bfseries A Comparison of GARCH and Stochastic Volatility Models in
+  the Study of Intraday Returns \\[0.4cm] }
+
+%\HRule \\[1.5cm]
+
+% Author and supervisor
+\begin{minipage}{0.4\textwidth}
+\begin{flushleft} \large
+\emph{Author:}\\
+\textsc{Xie} Xiaolei
+\end{flushleft}
+\end{minipage}
+\begin{minipage}{0.4\textwidth}
+\begin{flushright} \large
+\emph{Supervisor:} \\
+Prof. Sven \textsc{\AA berg}
+\end{flushright}
+\end{minipage}
+
+\vfill
+
+% Bottom of the page
+{\large \today}
+
+\end{center}
+\end{titlepage}
+% \maketitle
 \tableofcontents
@@
 \chapter{Introduction}
-$\mathbb{E}(x)$
+The mathematics of the financial market has always been a topic that
+arouses interest and imagination, and with no doubt, has been
+studied from many aspects and in many different ways.
 
-\section{Stylized facts}
+Central to all these studies are the concepts of probability
+distributions and correlations. The values of stocks, futures,
+options, etc. are stochastic in nature and are governed by the laws of
+stochastic processes, which are expressed in terms of probability
+distributions and correlations.
+
+Meanwhile, the observables of the market are prices, volumes,
+turn-over (the amount of money paid in a trade), names of the brokers,
+and time of the trades. These quantities don't make much sense
+by themselves but do reveal the probabilistic dynamics of the market
+when put together and turned into statistics.
+
+Here by dynamics we mean how the value of an asset is affected by its
+own history as well as by the histories and current values of other
+assets in the market. The influence from the asset's own history is
+termed autocorrelations, i.e. correlations in time, while
+the influence from other assets are termed cross correlations.
+
+Mathematical models that describe the aforementioned observables in
+terms of probability distributions, autocorrelations and cross
+correlations give predictive power and help determine a fair price of
+a given asset. Hence they are a central topic of the mathematics of
+the financial market.
+
+Once a model has been tentatively constructed, it must be
+verified. The verification again relies on the statistics of the
+observables. Over the years, a number of phenomena have been
+consistently reported and become generally accepted. These are referred
+to as stylized facts, which are detailed in \S
+\ref{sec:StylizedFacts} and constitute a standardized test suit for
+model verification.
+
+The rest of this report is organized as follows: in
+\S\ref{sec:FundamentalConcepts} we present a few very important
+concepts and notations that we will often refer to in later
+chapters; in chapter \ref{chp:nordea_15min} we study 15-minute returns
+of Nordea Bank during the period 2012/01/16 - 2012/04/20, using GARCH
+and stochastic volatility models; in chapter \ref{chp:Nordea2} we
+study 15-minute and 30-minute returns of Nordea Bank in the period
+2013/10/10 - 2014/04/04; and in chapter \ref{sec:volvo} we look at
+15-minute and 30-minute returns of Volvo B in the same period.
+
+In chapter \ref{chp:SLV_unconditional} a calculation of the
+unconditional distribution functions of a stochastic volatility model
+is presented. The results are checked against 30-minute returns of
+Nordea Bank, Volvo B and Ericsson B. Finally chapter \ref{chp:summary2}
+concludes the report.
+
+In these studies we approximate the conditional volatility with the
+``realized volatility'', which is computed from higher-frequency
+returns \cite{Mikosch2009}.
+The intraday returns involved in the studies are calculated from
+minute-by-minute average prices recorded in the Stockholm Stock
+Exchange (OMX). If no transaction is record in a interval $[t, t +
+\Delta t]$, the average price of the previous interval $[t - \Delta t,
+t]$ is assumed.
+
+\section{Fundamental Concepts \& Notations}
+\label{sec:FundamentalConcepts}
+This section is a list of a few concepts that may be unfamiliar to the
+reader and that we will often refer to later in the thesis:
+\begin{itemize}
+\item Return. Given a fixed time interval $[t - \Delta t, t]$, for example, a
+  day, a week, a month, etc, and a particular asset, for example, a
+  share in company ABC, the return of this asset over the time
+  interval is defined as
+  \begin{eqnarray*}
+    r_{t, \Delta t} &=& \ln p_{t} - \ln p_{t - \Delta t} \\
+    &=& \ln \left(1 + {p_{t} - p_{t - \Delta t} \over p_{t - \Delta t}}\right) \\
+    &\approx& {p_{t} - p_{t - \Delta t} \over p_{t - \Delta t}}
+  \end{eqnarray*}
+  where $p_t$ is the price of the asset at time $t$. $\Delta t$ is
+  sometimes called the time-lag of the return. Quite often, where
+q  confusion is not possible, we will just write $r_t$ to mean $r_{t,
+    \Delta t}$, the time lag either does not matter or is clear from
+  the context.
+
+\item Autocorrelation. By autocorrelation, denoted $\rho_k$ here, we
+  mean the correlation between two temporally separated observations
+  of the same time series:
+  \begin{eqnarray*}
+    \rho_k &=& {
+      \E\left[(a_t -\E(a_t))(a_{t-k} - \E(a_{t-k})\right]
+      \over
+      \sqrt{\text{var}(a_t)}\sqrt{\text{var}(a_{t-k})}
+    }
+  \end{eqnarray*}
+  Here $\E(x)$ stands for the expectation value of $x$, and
+  $\text{var}(x)$ stands for the variance of $x$. $k$ is called the
+  time-lag and is the temporal seperation of the two observations
+  measured by the number of observations in between. For example, the
+  time-lag between the 1st and 3rd observation is 2.
+
+\item Cross-Correlation Matrix. Correlations between the returns of a
+  group of assets are described by the cross-correlation matrix. When
+  the asset returns are described by a stable distribution law with
+  L\'evy index $\alpha$, the empirical cross-correlation matrix is
+  constructed as
+  \begin{equation}
+    \label{eq:cross-correlation}
+    \begin{aligned}
+      C_{ij} &= {1 \over T^{2/\alpha}} \sum_{t=1}^T [r_{i,t}-\E(r_i)]
+      [r_{j,t}-\E(r_j)] \\
+      &= {1 \over T^{2/\alpha}} RR'
+    \end{aligned}
+  \end{equation}
+  where $r_{i,t}$ is the return of the i-th asset at time t and is
+  placed at the entry (i, t) of matrix R; $R'$ denotes the
+  transpose of R. In most practical situations, one has abundant data
+  for each and every asset. Thus $T \geq N$ is assumed throughout this
+  thesis.
+
+\item Auto-regressive processes. A time series $r_t$ is called an
+  auto-regressive process of order $p$ and denoted AR(p), if it can be
+  written in the following form:
+  \begin{eqnarray*}
+    r_t &=& \sum_{i=1}^p \phi_i r_{t-i} + a_t
+  \end{eqnarray*}
+  where, for all $i$, $\phi_i \in (-1, 1)$, and the $a_t$'s are
+  independent and identically distributed (iid.) random variables with
+  zero mean. Obviously this implies the mean of $r_t$ is zero
+  too. Apart from this, their distribution of $a_t$ is not restricted
+  to any particular form.
+
+  Of particular interest to this thesis is the AR(1) process:
+  \begin{eqnarray*}
+    r_t = \phi r_{t-1} + a_t
+  \end{eqnarray*}
+  Its autocorrelation function $\rho_k = \text{corr}(r_t, r_{t-k})$
+  ($k = 0, 1, 2, \cdots$) can be easily shown to fall off
+  exponentially:
+  \begin{eqnarray*}
+    \rho_k &=& {\E(r_tr_{t-k}) - \E(r_{t-k})\E(r_{t}) \over
+      \sqrt{\var(r_t) \var(r_{t-k})}} \\
+    &=& {\phi \E(r_{t-1}r_{t-k}) + \E(a_t r_{t-k})
+      \over
+      \sqrt{\var(r_t) \var(r_{t-k})}
+    } \\
+    &=& \phi \rho_{k-1}
+  \end{eqnarray*}
+  where $\E(a_t r_{t-k}) = 0$ follows from the fact that any return
+  $r_{t1}$ must not depend on disturbances $a_{t2}$ that occur later
+  in time. The last equation means $\rho_k$ is a geometric series. Since
+  $\rho_0 = 1$, we have
+  \begin{eqnarray*}
+    \rho_k &=& \phi^k \\
+    |\rho_k| &=& e^{k\ln|\phi|} \\
+  \end{eqnarray*}
+  Note that $|\phi| < 1$ and hence $\ln|\phi| < 0$.
+
+  Although $k$ can only take integer values, it is still useful to
+  define a correlation time $\tau$ such that $\phi^\tau = 1/2$. Such a
+  quantity is more intuitive and constitutes a measure of
+  autocorrelations that is universal and comparable among different
+  time series' models. From the definition of $\tau$ we get
+  \begin{equation}
+    \label{eq:tau_def}
+    \begin{aligned}
+      \phi^\tau &= 1/2 \\
+      \tau &= -{\ln 2 \over \ln\phi} \\
+      \phi &= 2^{-1/\tau}
+    \end{aligned}
+  \end{equation}
+  Figure \ref{fig:AR1-autocorrelation} shows the autocorrelation
+  function of the AR(1) model. As proven above, this function decays
+  exponentially.
+  \begin{figure}[htb!]
+    %\vspace{-15mm}
+    \centering
+    \includegraphics[scale=0.5, clip=true, trim=113 229 115
+    139]{../pics/AR1-autocorrelation.pdf}
+    \caption{\small \it Autocorrelation function of the AR(1) model
+      with $\phi=1/\sqrt{2}$. Red circles: autocorrelations at $k=0, 1,
+      2, \cdots$. Blue line: $e^{t\ln\phi}$.}
+    \label{fig:AR1-autocorrelation}
+  \end{figure}
+  For more details about conventional time series' models, see
+  \cite{BoxJenkins94}.
+\end{itemize}
+
+For purposes of later reference, we also list some notations that may
+cause confusion to the reader:
+\begin{itemize}
+\item $\E(x)$ or $\mean{x}$: The expectation value of $x$.
+\item $\var(x)$: the variance of $x$.
+\item $\text{cov}(x,y)$: the covariance of $x$ and $y$.
+\item $\text{corr}(x,y)$: the correlation between $x$ and $y$, i.e.
+  \begin{equation*}
+    \text{corr}(x,y) = {\text{cov}(x,y) \over \sqrt{\var(x)\var(y)}}
+  \end{equation*}
+\end{itemize}
+
+\section{Stylized facts}\label{sec:StylizedFacts}
+This section presents and explains a few ``stylized facts'',
+i.e. phenomena that are widely observed and accepted as true.
+\begin{itemize}
+\item Fat tails. It has been consistently reported by various studies
+  - for example \cite{Potters2003} and \cite{Mantegna2000} - that the
+  probability density function (PDF) of stock/index returns are not
+  Gaussian. Unlike the Gaussian PDF, which is symmetric and falls off
+  very quickly as its argument moves from the center to the outskirts
+  (tails), the PDF of stock/index returns are higher on the tails,
+  i.e. the probability of large fluctuations is higher than is dictated
+  by a Gaussian distribution - in fact, even higher than dictated by
+  an exponential function. Empirical studies suggest the distribution
+  function of the returns follows a power-law on the tails --- the
+  exponent of the power depends on the specific stock/index.
+
+  Figure \ref{fig:FatTail} illustrates this feature.
+  \begin{figure}[htb!]
+    \centering
+    \includegraphics[scale=0.6, clip=true, trim=92 229 110
+    140]{../pics/FatTail.pdf}
+    \caption{\small \it Fat tails of {\it Nordea Bank} 15-minute
+      returns during the period 2013-10-10 and 2014-01-29. The returns
+      are computed using minute-by-minute average prices.}
+    \label{fig:FatTail}
+  \end{figure}
+
+\item Non-zero skewness. Apart from fat tails, the PDF of stock/index
+  returns are often also skewed. If the skewness is positive
+  (negative), the probability of very large positive (negative)
+  returns is higher than that of very large negative (positive)
+  returns, even though the mean of the returns is 0 or extremely
+  close to 0.
+
+  Table \ref{tab:EmpiricalSkewness} lists the skewness of a few
+  Swedish stocks traded on the Stockholm OMX market.
+  \begin{table}[htb!]
+    \footnotesize
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|c|}
+      \hline
+      Nordea Bank & Volvo B & Boliden & ABB Ltd & H\&M & Scania
+      B & Ericsson B \\
+      \hline
+      0.1362 & 0.0471 & 0.0567 & -0.0364 & -0.0168 &
+      -1.1554 & 0.2086 \\
+      \hline
+    \end{tabular}
+    \caption{\small \it Skewness of Stock Returns. All the returns
+      have time-lag of 15 minutes and are computed using paid prices
+      between 2013-10-10 and 2014-01-29. }
+    \label{tab:EmpiricalSkewness}
+  \end{table}
+
+\item Higher-order autocorrelation. To the lowest order, return series
+  are not auto-correlated --- if they are, the auto-correlations would
+  present an obvious opportunity of making easy profit and be
+  exploited and vanish as soon as they appear. However, the squared
+  returns do have significant auto-correlations, as figure
+  \ref{fig:nordea_15min_acf} illustrates. These auto-correlations
+  suggest the variances of the returns at subsequent time steps are
+  correlated. These are referred to as higher-order
+  auto-correlations.
+\end{itemize}
 
 \section{GARCH models}
-The autocorrelation function of $\epsilon_t^2$ is given by
+``GARCH'' is the acronym for ``Generalized Autoregressive Conditional
+Heteroscedasticity''. A GARCH(p, q) model is defined by the following
+equation system:
+\begin{equation}
+  \label{eq:garch_def}
+  \begin{aligned}
+    r_t &= \mu + \epsilon_t \\
+    \sigma_t^2 &= \alpha_0 + \sum_{i=1}^q \alpha_i \epsilon_{t-i}^2 +
+    \sum_{i=1}^p \beta_i \sigma_{t-i}^2
+  \end{aligned}
+\end{equation}
+where $\epsilon_t$ is termed the innovation of the return $r_t$ and is
+a random variable of a prescribed distribution, the most common ones
+being Gaussian, Student's t, Asymmetric Student's t, etc; $\sigma_t^2$
+is the variance of $\epsilon_t$ \cite{Bollerslev86}. The
+autocorrelation function of $\epsilon_t^2$ is given by
 (c.f. \cite{Bollerslev87})
-$$
-\rho_n = \sum_{i=1}^{p \vee q} (\alpha_i + \beta_i) \rho_{n-i}
-\;n > p
-$$
+\begin{eqnarray*}
+\rho_n &=& \sum_{i=1} ^{p \vee q} (\alpha_i + \beta_i) \rho_{n-i}
+\text{ for $n > p$}
+\end{eqnarray*}
 where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
@@
 equations, it is clear that the partial autocorrelation function cuts
-off at $max(p, q)$.
+off at $\max(p, q)$.
 
-\chapter{Case Study of Stocks}
-\section{Nordea Bank}
-\begin{figure}[ht]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=103 236 118
-  200]{../pics/nordea_price_20120116-20120420.pdf}
-  \caption{Nordea Bank paid prices 2012/01/16-2012/04/20}
-  \label{fig:Nordea}
-\end{figure}
-Figure \ref{fig:Nordea} shows the paid prices of Nordea Bank from
-2012/01/16 to 2012/04/20.
-
-\subsection{15min Returns}
-\subsubsection{GARCH model}
-The autocorrelation function (ACF) of $\epsilon_t^2$ are shown in figure
-\ref{fig:nordea_15min_acf}. At the absence of GARCH effects, the ACF
-will have an asymptotic Gaussian distribution with mean 0 and variance
-1/T \cite{Bollerslev86, Bollerslev87}. The first 5 autocorrelations
-have considerable sizes and do not fall off as in the Gaussian
-case. This observation suggests a GARCH model. Moreover, from
-\ref{fig:nordea_15min_vlt_acf}, one sees apparent autocorrelations
-between volatilities of consecutive time intervals. This again points
-to a GARCH model.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of squared returns]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_acf.pdf}
-    \label{fig:nordea_15min_acf}
-  }
-  \subfigure[ACF of volatilities]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_vlt_acf.pdf}
-    \label{fig:nordea_15min_vlt_acf}
-  }
-  \caption{Nordea Bank 15min autocorrelations}
-\end{figure}
-
-To obtain the functional form of the conditional distribution of
-$\epsilon_t$ \footnote{We use the realized volatility as an estimate of the
-conditional volatility. See \cite{Andersen03}.}, we look at the
-qq-plot of $\epsilon_t/\sigma_t$ \ref{fig:nordea_15min_epOversig_qq},
-which is consistent with a standard Gaussian distribution except for a
-few outliers at the two ends. Hence the fat tails of the probability
-density function can be entirely accounted for by GARCH effects,
-implying a Gaussian conditional distribution is appropriate for $\epsilon_t$.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.6, clip=true, trim=100 236 118
-    200]{../pics/nordea_15min_epOversig_qq.pdf}
-  \caption{QQ-plot of $\epsilon_t / \sigma_t$. $\epsilon_t$ are
-    derived from Nordea Bank 15min returns while $\sigma_t$ are realized
-    volatilities calculated using 30s returns within each 15min
-    interval.}
-  \label{fig:nordea_15min_epOversig_qq}
-\end{figure}
-
-To determine the orders p and q of the GARCH model, we again check
-the partial autocorrelation function of $\epsilon_t^2$. As just
-mentioned, the first 5 partial autocorrelations are significant. Thus
-we can tentatively estimate a GARCH(1, 5) model. Then by applying the
-Lagrange multiplier test and accordingly removing insignificant
-parameters, we obtain a GARCH(1, 3) model
-\ref{tab:nordea_15min_garch}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{c|c|c|c}
-    parameter & value & std. error & t statistic \\
-    \hline
-     $\alpha_0$ &   1.05668e-06 &   3.56871e-07 &    2.96096\\
-     \hline
-     $\beta_1$  &     0.687212  &   0.0299314   &    22.9596\\
-     \hline
-     $\alpha_1$ &      0.128614 &   0.0268144   &    4.79643\\
-     \hline
-     $\alpha_3$ &    0.0568022  &   0.0301503   &    1.88397
-  \end{tabular}
-  \caption{GARCH(1, 3) model of Nordea Bank 15min returns}
-  \label{tab:nordea_15min_garch}
-\end{table}
-
-\subsubsection{$(0,1,1)\times(0, 1, 1)$ seasonal volatility model}
-With the availability of transactional data, estimating conditional
+\section{Stochastic Volatility Models}\label{sec:SLV_model}
+With the availability of transaction data, estimating conditional
 volatility using returns sampled at a higher frequency (realized
@@
 1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
-sec.6, theorem 22):
+\S6, theorem 22):
 \[
@@
 
-Thus one can try a few frequencies and compare the distribution of
-$r_{t+h, h}/\hat{\sigma_{t+h, h}}$ with the standard Gaussian. If the
-two match, $\hat{\sigma_{t+h, h}}$ is a good estimate of $\sigma_{t+h,
-  h}$, the volatility of $r_{t+h, h}$.
+Thus in practice one can try a few frequencies and compare the
+distribution of $r_{t+h, h}/\hat{\sigma}_{t+h, h}$ with the standard
+Gaussian. If the two match, $\hat{\sigma}_{t+h, h}$ is a good estimate
+of $\sigma_{t+h, h}$, the volatility of $r_{t+h, h}$.
 
-For the present case with $h = 15\text{min}$, a reasonally good
-sampling frequency is 1/30sec, as confirmed by the QQ-plot
-\ref{fig:nordea_15min_epOversig_qq} as well as by the autocorrelation
-functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of $(r_{t+h, h} /
-\sigma_{t+h, h})^2$ which are shown in
-\ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} respectively.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of $r_{t+h, h} / \sigma_{t+h, h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_acf.pdf}
-    \label{fig:nordea_15min_quotient_acf}
-  }
-  \subfigure[ACF of $(r_{t+h, h} / \sigma_{t+h, h})^2$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
-    \label{fig:nordea_15min_quotient_squared_acf}
-  }
-  \caption{Nordea 15min $r_t/\sigma_t$ and $(r_t/\sigma_t)^2$ ACF.}
-\end{figure}
+Now that the computation of realized volatility is clear, we specify
+the model of the return and the log-volatility:
+\begin{eqnarray}
+   r_{t+h, h} &=& \ln p_{t+h} - \ln p_t \nonumber \\
+   r_{t+h, h} &=& \mu + \sigma_{t+h, h} b_t \label{eq:SLV_spec}
+\end{eqnarray}
+where $p_t$ is the price of the asset at time $t$ and $b_t \sim N(0,
+1)$. For simplicity, in the rest of this chapter we shall just write
+$t$ for the subscript ``t+h, h'', since the time interval $h$ is fixed
+for each time series and is a known constant. In the later chapters,
+we show that $\ln \sigma_t$ can be well described by ARIMA models.
 
-We see in figure \ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
-autocorrelation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
-$(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
-series is i.i.d Gaussian distributed. As a result, the problem of
-modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
-  h}$. Andersen and Bollerslev et al reported that, for the exchange
-rates between Deutch mark, yen and dollar, $\ln \sigma_{t+h, h}$ is
-gaussian distributed \cite{Andersen03}. This is, however, not the case
-for our series in question. In fact, in our case, $\ln \sigma_{t+h,
-  h}$ is right skewed (skewness 0.3342) and leptokurtic (kurtosis
-6.1006). See figure \ref{fig:nordea_15min_logvol_pdf} and
-\ref{fig:nordea_15min_logvol_qq}
-\begin{figure}[htb!]
-  \centering
-  \subfigure[PDF of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.32, clip=true, trim=52 173 46
-    113]{../pics/nordea_15min_logvol_hist.pdf}
-    \label{fig:nordea_15min_logvol_pdf}
-  }
-  \subfigure[QQ-plot of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    220]{../pics/nordea_15min_logvol_qq.pdf}
-    \label{fig:nordea_15min_logvol_qq}
-  }
-  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ unconditional
-    distribution}
-\end{figure}
+% \input{price_models}
+% \input{cross_correlation_gaussian}
+% \input{GARCH_correlation}
+% \input{summary.tex}
 
-Moreover, the autocorrelation function of $\ln\sigma_{t+h, h}$
-shows long-lasting and periodic autocorrelations in the time series
-with an apparent period of 33 (see figure
-\ref{fig:nordea_15min_logvol_acf}). This suggests simplifying the
-series by differencing \cite{BoxJenkins94}:
-\[
-w_t = (1-B)(1-B^s)\ln\sigma_{t+h, h}
-\]
-where $B$ is the back-shift operator. The length of the time interval
-h, which is 15min in the present case, is left implicit.
-
-The autocorrelation function of the differenced process $w_t$, as shown in
-figure \ref{fig:nordea_15min_w_acf}, clearly points to a seasonal $(0,
-0, 1)\times(0, 0, 1)$ model: There are only 4 non-zero
-autocorrelations in the plot, located at lags 1, 32, 33, 34,
-respectively; furthermore, the two at 32 and 34 are approximately
-equal. Thus we can write down the model as
-\begin{eqnarray*}
-  w_t = (1 - \theta B)(1 - \Theta B^s) y_t
-\end{eqnarray*}
-where $\theta$ and $\Theta$ are parameters to be determined and $y_t$
-is a noise process with constant variance $\sigma_y^2$ and mean
-0. $y_t$ is often refered to as the residuals.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    235]{../pics/nordea_15min_logvol_acf.pdf}
-    \label{fig:nordea_15min_logvol_acf}
-  }
-  \subfigure[ACF of $w_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    235]{../pics/nordea_15min_w_acf.pdf}
-    \label{fig:nordea_15min_w_acf}
-  }
-  \caption{Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$ autocorrelations}
-\end{figure}
-
-The above $(0, 0, 1)\times(0, 0, 1)$ model has the following
-autocovariance structure \cite{BoxJenkins94}:
-\begin{eqnarray*}
-  \gamma_0 &=& \sigma_y^2 (1 + \theta^2)(1 + \Theta^2) \\
-  \gamma_1 &=& -\sigma_y^2\theta(1 + \Theta^2) \\
-  \gamma_s &=& -\sigma_y^2\Theta(1 + \theta^2) \\
-  \gamma_{s+1} &=& \gamma_{s-1}\;=\;\sigma_y^2\theta\Theta
-\end{eqnarray*}
-These equations together with the measured autocorrelations make
-possible an initial estimate of the parameters $\theta$ and $\Theta$:
-\begin{eqnarray*}
-  {\rho_{s+1}/\rho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
-    1 + \theta^2} \\
-  {\rho_{s+1}/\rho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
-    1 + \Theta^2} \\
-\end{eqnarray*}
-Substituting in the measured values shown in table
-\ref{tab:nordea_15min_w_acf},
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{c|c|c|c}
-    $\rho_1$ & $\rho_{s-1}$ & $\rho_s$ & $\rho_{s+1}$ \\
-    \hline
-    -0.4703 &  0.2053 & -0.4564 &  0.2212
-  \end{tabular}
-  \caption{Nordea 15min $w_t$ autocorrelations}
-  \label{tab:nordea_15min_w_acf}
-\end{table}
-we get
-\begin{eqnarray*}
-  \theta &=& 0.6890 \\
-  \Theta &=& 0.6378
-\end{eqnarray*}
-Among the two roots of each of the 2nd order equations in the above,
-we have chosen the one in the range $(-1, 1)$ so as to ensure
-invertibility of the model \cite{BoxJenkins94}.
-
-With an estimate of $\theta$ and $\Theta$, one can then infer the
-noise process i.e. the residuals $y_t$:
-\begin{equation}
-  \label{eq:infer_y}
-  y_t = w_t + \theta y_{t-1} + \Theta y_{t-s} - \theta \Theta y_{t-s-1}  
-\end{equation}
-where we substitute $y_t\;(t \leq 0)$ with their unconditional
-expectation 0.
-% By inverting the seasonal moving average model of $w_t$ we can then
-% infer the noise process $y_t$ \footnotemark, i.e. the residuals:
-% \begin{eqnarray*}
-%   y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
-%   &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
-%   w_t \\
-%   &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
-%   w_{t-k}
-% \end{eqnarray*}
-% where $\bot(\cdot)$ denotes the floor function, i.e. the largest
-% integer smaller than its argument.
-% \footnotetext{
-%   The autoregressive process resulting from the inversion is inifinite
-%   in extent. Hence one must choose a cut-off value for k. For the
-%   present case of Nordea 15min log-volatility, I choose the value
-%   $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
-%   $k_{\top}$, is around 550. To avoid data reduction of this amount,
-%   we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
-%   corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $E(w_t) = E(y_t
-%   - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
-% }
-
-In order to forecst the $w_t$ process, and hence the price process
-itself, we must also know the distribution of $y_t$. Moreover, to
-properly estimate the parameters of the model in the sense of maximum
-likelihood, we are also in need of the distribution of $y_t$.
-
-If $y_t$ is indeed i.i.d random variables as we assumed, we only need
-to work out its unconditional distribution. This is shown in figure
-\ref{fig:nordea_15min_y_qq} and \ref{fig:nordea_15min_y_acf}.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[QQ-plot of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_y_qq.pdf}
-    \label{fig:nordea_15min_y_qq}
-  }
-  \subfigure[ACF of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_y_acf.pdf}
-    \label{fig:nordea_15min_y_acf}
-  }
-  \caption{Nordea 15min $y_t$ QQ-plot and autocorrelations}
-\end{figure}
-In addition, the measured moments of $y_t$ are shown in table
-\ref{tab:nordea_15min_y_moments}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{c|c|c|c}
-    mean & variance & skewness & kurtosis \\
-    \hline
-    0.0012 & 0.0935 & 0.2988 & 6.8691
-  \end{tabular}
-  \caption{Nordea 15min $y_t$ moments}
-  \label{tab:nordea_15min_y_moments}
-\end{table}
-
-If we ignore the weak autocorrelations in this crude estimate of
-$y_t$ and assume $y_t$ to be i.i.d random variables governed
-by a skewed and leptokertic distribution, the parameters of this
-distribution together with $\theta$ and $\Theta$ can be determined via
-maximum likelihood estimate (MLE).
-
-It turns out that $y_t$ can be well described by a Johnson Su
-distribution \cite{Shang2004}:
-\[
-  y_t = \xi + \lambda\sinh{z_t - \gamma \over \delta}
-\]
-where $\gamma, \delta, \lambda, \xi$ are parameters to be determined
-and $z_t \sim N(0, 1)$. The goodness of fitting is demonstrated in
-figure \ref{fig:nordea_15min_y_js_fit} by the
-empirical cummulative distribution function in comparison to the
-theoretical one.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=112 272 107
-    140]{../pics/nordea_15min_y_js_fit.pdf}
-    \caption{Nordea 15min $y_t$ fitted to a Johnson Su distribution}
-    \label{fig:nordea_15min_y_js_fit}
-\end{figure}
-
-The first 4 moments of the Johnson Su distribution are expressible
-in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}:
-\begin{eqnarray*}
-  w &=& \exp{1 \over \delta^2} \\
-  \Omega &=& {\gamma \over \delta} \\
-  \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
-  \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
-    1)\right]^{1/2} \\
-  \text{skewness}(y) &=& {
-    \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
-    \over
-    (w\cosh 2\Omega + 1)^{3/2}} \\
-  \text{kurtosis}(y) &=& {
-    w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
-    + 3(2w+1) \over
-    2(w\cosh 2\Omega + 1)^2 }
-\end{eqnarray*}
-By matching the theoretical expressions of the moments with their
-measured values, and taking help from published tables
-\cite{Johnson1965}, one can solve for the parameters $\gamma, \delta,
-\lambda, \xi$.
-
-Under the assumption of i.i.d Johnson Su distributed residuals, the
-log-likelihood function of the parameters $\gamma, \delta$
-conditional on the sample $w_t$ can be written as
-\[
-L(\theta, \Theta) = -{1 \over 2}\sum_{t=1}^n z_t^2 + n \ln{\delta
-  \over \lambda \sqrt{2\pi}} - {1 \over 2}\sum_{t=1}^n \ln\left[
-  1 + \left({y_t - \xi \over \lambda}\right)^2
-\right]
-\]
-where $y_t$ are infered from $w_t$ using eq.\ref{eq:infer_y}
-and $z_t$ from $y_t$ using
-\[
-z_t = \delta \sinh^{-1}{y - \xi \over \lambda} + \gamma
-\]
-Note that $\gamma, \delta, \lambda, \xi$ are not really free
-parameters but rather are implied by $\theta$ and $\Theta$: Once the
-latter have been chosen and the corresponding $y_t$ infered, the
-former are determined by the moments of $y_t$.
-
-The eventual MLE is done in Matlab with the ``active set''
-algorithm. The initial as well as the final estimation results are
-listed in table \ref{tab:nordea_15min_js_param}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{c|c|c|c|c|c|c}
-    & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
-    \hline
-    initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
-    0.6378 \\
-    \hline
-    MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
-  \end{tabular}
-  \caption{Nordea 15min estimation results}
-  \label{tab:nordea_15min_js_param}
-\end{table}
-
-The probability density function of the residuals $y_t$ that result
-from the MLE estimate is shown in figure
-\ref{fig:nordea_15min_y1_hist} together with its counterpart from the
-initial estimate. 
-\subsection{45min Returns}
+\input{Nordea_15min}
+\input{NordeaVolvo}
+\input{SLV_Unconditional}
+\input{summary2}
 

### 2014-05-05 epdf.m

@@ -13,5 +13,5 @@ y = tsmovavg(y, 's', lag);
 y = y(lag:end);
 
 if ~isempty(spec)
-    plot(log10(x), log10(y), spec, 'LineWidth', 2);
+    plot((x), (y), spec, 'LineWidth', 2);
 end

### 2014-05-06 thesis/cross_correlation_gaussian.tex
Downsized to 49 pages in the main text. Many images are moved to the appendix.

@@
 \section{Distribution of the Eigenvalues}\label{sec:GCC-numerical}
-For the Wishart matrix, theoretical results are available for the
-eigenvalue distribution \cite{Chiani2012}. In the simplest case where
-$\Sigma = I$, i.e. no cross-correlations are present, we have
-\begin{eqnarray*}
-  f_{\lambda}(x_1, \cdots, x_N) = K \prod_{i=1}^N e^{-x_i/2}
-  x_i^{(T-N-1)/2} \prod_{i<j}^N (x_i - x_j)
-\end{eqnarray*}
-where the eigenvalues have been indexed in descending order: $x_1 \geq
-x_2 \geq \cdots \geq x_N$. The normalization constant K is given by
-\begin{equation*}
-  K = {\pi^{N^2/2} \over 2^{NT/2}\Gamma_N(T/2) \Gamma_N(N/2)}
-\end{equation*}
-where the function $\Gamma_m(a)$ is defined as
-\begin{equation*}
-  \Gamma_m(a) = \pi^{m(m-1)/4} \prod_{k=1}^m \Gamma\left(a - {k-1
-      \over 2}\right)
-\end{equation*}
+For the Wishart matrix $\mtx{RR'}$, theoretical results are available
+for the eigenvalue distribution, and we summarize them in appendix
+\ref{sec:wishart_eigen_dist}. In short, the joint probability density
+function of the eigenvalues is given by equation
+\ref{eq:wishart_eigen_pdf} when neither auto-correlation nor
+cross-correlation is present in $\mtx R$. Moreover, the mariginal
+distribution of the largest eigenvalue is approximated by a gamma
+distribution \cite{Chiani2012}.
+
 However, as detailed in the derivation leading to equation
 \ref{eq:cross-corr-matrix-PDF}, the distribution of $RR'$ is not
-Wishart when the columns of $R$ are correlated. Deriving the
-eigenvalue distribution analytically in this case is difficult and
-beyond the scope of this thesis. Instead we resort to numerical
-methods.
+Wishart when the columns of $R$ are correlated
+(auto-correlation). Deriving the eigenvalue distribution analytically
+in this case is beyond the scope of this thesis. Instead, we resort to
+numerical methods.
 
-As before we consider the AR(1) process (see \S
-\ref{sec:FundamentalConcepts}):
+As before we consider the AR(1) process (see
+\S\ref{sec:FundamentalConcepts}):
 \begin{equation*}
@@
 of the simulation. 
-% \begin{wrapfigure}{r}{0.5\textwidth}
-\begin{figure}[htb!]
-  \begin{center}
-    \includegraphics[scale=0.5, clip=true, trim=90 228 115
-    226]{../pics/GaussianMarkovSpectrumPDF.pdf}
-  \end{center}
-  % \vspace{-10mm}
-  \caption{\small \it
-      Eigenvalue distribution with $\tau$ ranging from 0 to
-      3. The 1st blue line, which is shown as stairs, is the
-      theoretical eigenvalue distribution according to the
-      Marcenko-Pastur law: $f(\lambda) = \sqrt{(\lambda_+/\lambda - 1)(1
-        - \lambda/\lambda_-)}/2\pi q \sigma^2$, where $\lambda_{\pm} =
-      \sigma^2(1 \pm \sqrt{q})^2$ In the simulation we have
-      chosen $q = N/T = 50/1000 = 0.05$ and $\sigma=1$. For each value
-      of $\tau$ we generate 2000 instances of $N \times T$ random matrix
-      $R$, and compute C as $C=RR'/T$. Hence each curve in the figure is
-      constructed from 2000 sets of eigenvalues.
-    }
-  \label{fig:GaussianMarkovSpectrumPDF}
-\end{figure}
 It is clear from the figure that the maximum eigenvalue moves
-consistently to the right as the value of $\phi$ increases. The change
-of the minimum eigenvalue remains unclear from the figure, so we plot
-in \ref{fig:Gaussian_mineig} their mean for each value of
-$\phi$. There one can see that the minimum eigenvalue actually
-increases with $\phi$.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=100 226 116
-  133]{../pics/Gaussian_mineig.pdf}
-  \caption{\small \it The the minimum eigenvalue versus
-    auto-correlation strength $\phi$. For each value of $\phi$ 2000
-    random matrices are generated and their eigenvalues are
-    calculated. The minimum eigenvalue of each random matrix is noted
-    and the mean of the 2000 such minimum eigenvalues are plotted
-    against the chosen value of $\phi$. 20 values of $\phi$ are
-    included in the plot, ranging from 0 to 0.95 with step size 0.05.}
-  \label{fig:Gaussian_mineig}
-\end{figure}
+consistently to the right as the value of $\phi$ increases, and, as
+shown in figure \ref{fig:Gaussian_mineig}, the minimum eigenvalue
+also increases with $\phi$.
 
-K. Johnsson \cite{Johnsson2000} and I. Johnstone \cite{Johnstone2001}
-showed that, at the {\it absence} of autocorrelations and in the
-asymptotic limit $N, T \to \infty$, $N/T \to q < \infty$, the maximum
-eigenvalue  $\lambda_1$ follows the Tracy-Widom distribution (denote
-$\mathscr{TW}_1$ here) when properly relocated and rescaled:
-\begin{equation*}
-  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} \sim \mathscr{TW}_1
-\end{equation*}
-where $\mu_{NT}$ and $\sigma_{NT}$ are given by
-\begin{eqnarray*}
-  \mu_{NT} &=& \left(
-    \sqrt{N-1/2} + \sqrt{T - 1/2}
-  \right)^2 \\
-  \sigma_{NT} &=& \sqrt{\mu_{NT}} \left(
-    {1 \over \sqrt{N-1/2}} + {1 \over \sqrt{T-1/2}}
-  \right)^{1/3}
-\end{eqnarray*}
-So in the following we focus on the maximum eigenvalue distribution
-and leave the minimum eigenvalue for future studies. The
-$\mathscr{TW}_1$ distribution has the following cummulative
-distribution function (CDF) \cite{Chiani2012}
-\begin{equation*}
-  F_1(x) = \exp\left[
-    -{1 \over 2} \int_x^\infty dy \left(q(y) + (y-x)q^2(y)\right)
-  \right]
-\end{equation*}
-where $q(y)$ is defined as the solution to the Painlev\'e II differential
-equation
-\begin{equation*}
-  q''(y) = yq(y) + 2q^3(y)
-\end{equation*}
-which is unique when imposing the condition
-\begin{equation*}
-  q(y) \sim \text{Ai}(y) \text{ as } y \to \infty
-\end{equation*}
+% In the following we focus on the maximum eigenvalue, which is shown by
+% Chiani to have approximately a gamma distribution \cite{Chiani2012}.
+% So we first verify Chiani's result, and then extend it to the
+% case {\it with} autocorrelations and study how the
+% distribution changes as $\phi$ increases.
+Chiani showed that the maximum eigenvalue ($\lambda_1$) had
+approximately a gamma distribution \cite{Chiani2012}. So we compare in
+figure \ref{fig:GaussianMarkov005MaxEigCDF_loglog} the empirical
+cumulative distribution function (CDF) of the maximum eigenvalue 
+with the CDF of a gamma distribution. The cases where
+auto-correlations are present ($\phi > 0$) have also been included.
 
-Then Marco Chiani showed recently that the $\mathscr{TW}_1$ distribution
-can be well approximated by a gamma distribution based on his proof
-that the exact distribution of the maximum eigenvalue is a mixture of
-gamma distributions. Specifically,
-\begin{equation}\label{eq:TracyWidom-Gamma}
-  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} + \alpha \sim
-  \mathscr{G}(k ,\theta)
-\end{equation}
-where $\mathscr{G}(k, \theta)$ denotes the Gamma distribution with
-parameters $k$ and $\theta$ \cite{Chiani2012}. The probability
-density function (PDF) of the gamma distribution is given by
-\begin{equation*}
-  f_\gamma(x; k, \theta) = {1 \over \Gamma(k) \theta^k} x^{k-1}
-  e^{-x/\theta}
-\end{equation*}
-Moreover, the first 3 moments of the distribution are simple:
-\begin{eqnarray*}
-  \text{mean} &=& k\theta \\
-  \text{variance} &=& k\theta^2 \\
-  \text{skewness} &=& 2/\sqrt{k}
-\end{eqnarray*}
+It is seen in the figure that gamma distributions with different
+parameters fit fairly well in all cases. So we conclude that a
+gamma distribution not only approximates the maximum eigenvalue
+distribution at the absence of autocorrelations but does so even at
+the {\it presence} of autocorrelations.
 
-In the following we first verify these results and extend them to the
-case {\it with} autocorrelations and then study how the
-distribution changes as $\phi$ increases. Figure
-\ref{fig:GaussianMarkov005MaxEigCDF_loglog} compares the empirical
-maximum eigenvalue CDF with the CDF of a gamma distribution. Each pair
-of CDFs correspond to a fixed autocorrelation strength ($\phi$). The
-parameters $k$ and $\theta$ of the gamma distribution are fit to
-data by matching the 2nd and the 3rd moments of the gamma distribution
-to the corresponding moments of the empirical distribution. Then the
-parameter $\alpha$ in equation \ref{eq:TracyWidom-Gamma} is chosen to
-be
-\begin{equation*}
-  \alpha = k\theta - \E\left({\lambda_1 - \mu_{NT} \over \sigma_{NT}}\right)
-\end{equation*}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[scale=0.5, clip=true, trim=0 238 0
-    197]{../pics/GaussianMarkov005MaxEigCDF_loglog.pdf}
-  \end{center}
-  \caption{\small \it Cummulative distribution function (CDF) of the maximum
-    eigenvalue ($\lambda_1$). Blue: empirical CDF. Red: CDF of the
-    fitting gamma distribution. The curves are plotted on log-log
-    scale.}
-  \label{fig:GaussianMarkov005MaxEigCDF_loglog}
-  %\vspace{-10mm}
-\end{figure}
-
-We can see in figure \ref{fig:GaussianMarkov005MaxEigCDF_loglog} that
-the gamma distribution fits fairly well to its corresponding maximum
-eigenvalue distribution. So we conclude that a gamma distribution not
-only approximates the maximum eigenvalue distribution at the absence of
-autocorrelations but does so even at the {\it presence} of
-autocorrelations. Since the maximum eigenvalue distribution is
-characterized by the parmaters $k$, $\theta$, and  $\alpha$, the
-influence of the autocorrelations can be characterized by the
-dependence of $k$, $\theta$, and  $\alpha$ on $\phi$. In the following
-we study these dependences in more details.
-
-Figure \ref{fig:param1_phi_dep} shows how the mean, variance,
-and skewness of the fitted gamma distribution, as well as the
-relocation parameter $\alpha$ denpend on $\phi$.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=37 240 50
-    202]{../pics/param1_phi_dep.pdf}
-  \caption{\small \it  the  mean, variance, and skewness of the fitted
-    gamma distribution as well as the relocation parameter $\alpha$
-    are plotted against the autocorrelation strength $\phi$. 20 values
-    of $\phi$ are included in the plot, ranging from 0 to 0.95 with
-    step size 0.05.}
-  \label{fig:param1_phi_dep}
-\end{figure}
-Despite the monotonic behavior shown by 3 of these curves, the
-relation between each of the aforementioned quantities and $\phi$ is
-more complicated than it apears. However, the mean, i.e. $k \theta$ is
-an exception: Good support in the data can be found for the following
+Since the maximum eigenvalue distribution is characterized by the
+parameters $k$, $\theta$, and $\alpha$, the influence of the
+autocorrelations can be characterized by the dependence of $k$,
+$\theta$, and  $\alpha$ on $\phi$. While these dependences are rather
+intricate,
+% Figure \ref{fig:param1_phi_dep} shows how the mean, variance,
+% and skewness of the fitted gamma distribution, as well as the
+% relocation parameter $\alpha$ denpend on $\phi$.
+% \begin{figure}[htb!]
+%   \centering
+%     \includegraphics[scale=0.5, clip=true, trim=37 240 50
+%     202]{../pics/param1_phi_dep.pdf}
+%   \caption{\small \it  the  mean, variance, and skewness of the fitted
+%     gamma distribution as well as the relocation parameter $\alpha$
+%     are plotted against the autocorrelation strength $\phi$. 20 values
+%     of $\phi$ are included in the plot, ranging from 0 to 0.95 with
+%     step size 0.05.}
+%   \label{fig:param1_phi_dep}
+% \end{figure}
+% Despite the monotonic behavior shown by 3 of these curves, the
+% relation between each of the aforementioned quantities and $\phi$ is
+% more complicated than it apears. However, the mean, i.e. $k \theta$ is
+% an exception: 
+good support can be found in the data for the following
 approximate relation:

### 2014-05-06 thesis/price_models2.tex
Downsized to 49 pages in the main text. Many images are moved to the appendix.

--- 2014-04-30 thesis/price_models.tex
@@
 \begin{eqnarray*}
-\rho_n &=& \sum_{i=1} ^{p \vee q} (\alpha_i + \beta_i) \rho_{n-i}
+\rho_n &=& \sum_{i=1} ^{\max(p, q)} (\alpha_i + \beta_i) \rho_{n-i}
 \text{ for $n > p$}
@@
 where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
-zeros. $p \vee q$ denotes the maximum of p and q. From these
-equations, it is clear that the partial autocorrelation function cuts
-off at $\max(p, q)$.
+zeros. It is clear from this equation that the partial autocorrelation
+function cuts off at $\max(p, q)$.
 
@@
 including the Volvo B 30-minute returns described in
-\S\ref{sec:SLV_Symmetric}. In the following, we list the parameter
-values and show the comparison between the empirical distribution
-functions and their counterparts from the model.
-
-Clearly these figures show a fairly good match of the empirical
-distribution functions to their model counterparts. So we are once
-again assured of the validity of ARIMA log-volatility models. However,
-we also see that the skewness of the returns of Volvo B and Ericsson B
-have rather different values when computed from the sample and from
-the model. This could be the consequence of the limited sample size or
+\S\ref{sec:SLV_Symmetric}. Table \ref{tab:assets_params} and
+\ref{tab:assets_moments} present the obtained parameter values and the
+resulting moments, respectively. Figure \ref{fig:Volvo_B_30m_returns},
+\ref{fig:Nordea_30m_returns}, and \ref{fig:Ericsson_30m_returns}
+compare the empirical distribution functions with the analytic
+distribution functions evaluated at these parameters.
+
+Clearly these figures show a fairly good match of the empirical and
+the analytic distribution functions. However, we also see that the
+skewness of the returns of Volvo B and Ericsson B have rather
+different values when computed from the sample and from the
+model. This could be the consequence of the limited sample size or
 deficiencies in the estimation procedure described above. We leave
@@
 % leave this to later studies.
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    & $\psi$ & $\sigma$ & $\bar{v}$ & $\mu$ \\
-    \hline
-    Volvo B & -1.5747e-02 & 4.6929e-01 & -6.2304e+00 &
-    -5.3040e-05 \\
-    Nordea Bank & 1.7234e-02 & 4.2780e-01 & -6.3758e+00 & 7.5230e-05
-    \\
-    Ericsson B & 1.8486e-02 & 4.4905e-01 & -6.2834e+00 & -9.2817e-05 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Parameter Values of Selected Assets'
-    returns. The time series are 30-minute returns and run from
-    2013/10/10 to 2014/03/12.}
-  \label{tab:assets_params}
-\end{table}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    & mean & std & skewness & kurtosis \\
-    \hline
-    \multirow{2}{*}{Volvo B} & -8.6766e-05 & 2.4500e-03 &
-    -2.7217e-02 & 7.3394e+00 \\
-    & -6.9281e-05 & 2.4538e-03 & -7.2677e-02 & 7.2472e+00 \\
-    \hline
-    \multirow{2}{*}{Nordea Bank} & 7.4821e-05 & 2.0500e-03 &
-    6.9433e-02 & 6.7427e+00 \\
-    & 8.8983e-05 & 2.0443e-03 & 6.7125e-02 & 6.2446e+00 \\
-    \hline
-    \multirow{2}{*}{Ericsson B} & -8.2606e-05 & 2.2939e-03 &
-    2.6586e-01 & 7.6716e+00 \\
-    & -7.5675e-05 & 2.2844e-03 & 7.8573e-02 & 6.7299e+00 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Moments of Selected Assets' returns. For each return
-    series, the 1st row contains the sample moments, while the 2nd
-    constains those computed with MLE parameters. The time series are
-    30-minute returns and run from 2013/10/10 to 2014/03/12.}
-  \label{tab:assets_moments}
-\end{table}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=50 233 63
-  136]{../pics/Volvo_B_30m_returns.pdf}
-  \caption{\small \it Volvo B 30min returns' unconditional distribution fit to
-    the Stochastic Volatility model. The time series runs from
-    2013/10/10 to 2014/03/12. Left: $P(r < x)$ with $x < 0$; Right:
-    $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
-  \label{fig:Volvo_B_30m_returns}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=65 241 72
-  143]{../pics/Nordea_30m_returns.pdf}
-  \caption{\small \it Nordea Bank 30min returns' unconditional
-    distribution fit to the Stochastic Volatility model. The time
-    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
-    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
-  \label{fig:Nordea_30m_returns}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=21 241 28
-  146]{../pics/Ericsson_30m_returns.pdf}
-  \caption{\small \it Ericsson B 30min returns' unconditional
-    distribution fit to the Stochastic Volatility model. The time
-    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
-    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
-  \label{fig:Ericsson_30m_returns}
-\end{figure}
 

### 2014-05-06 thesis/thesis.tex
Downsized to 49 pages in the main text. Many images are moved to the appendix.

--- 2014-04-30 thesis/thesis.tex
@@
 
-\chapter{Correlation Matrix of GARCH(1,1) Returns}
+\chapter{Covariance Matrix of GARCH(1,1) Returns}
 \label{chp:CrossCorrelationFat}
@@
 % \label{sec:GarchWishart}
-In this chapter we consider the cross-correlation matrix of N
+GARCH models and particularly GARCH(1,1) models are widely used to
+model financial return series of various time scales, ranging from
+daily and monthly returns that have been studied extensively in the
+literature to intraday returns that we have selectively investigated
+in chapter \ref{chp:PriceModels}. One particularly nice feature of
+GARCH(1,1) models is that they have regularly varying tails (power-law
+tails) \cite{Mikosch2009, mikosch2000} even when the innovations
+(denoted $z_t$ in the following text) are normally distributed ---
+something not shared by other classes of models, e.g. not by ARIMA
+log-volatility models, as shown in \S\ref{sec:XieCalc}. However, what
+this tail behavior implies for the covariance matrix is much less
+understood, especially when the constituent return series are
+auto-correlated.
+
+So in this chapter we consider the covariance matrix of N
 identically specified, possibly auto-correlated GARCH(1, 1)
@@
   r_{it} &=& \phi r_{i, t-1} + \epsilon_{it} \nonumber \\
-  \epsilon_{it} &=& \sigma_{it} a_{it} \label{eq:garch_spec}
+  \epsilon_{it} &=& \sigma_{it} z_{it} \label{eq:garch_spec}
 \end{eqnarray}
-where $i=1,2,...,N$; $t=1,2,...,T$; $a_{it}$ is independent,
+where $i=1,2,...,N$; $t=1,2,...,T$; $z_{it}$ is independent,
 identically distributed, and
 \begin{eqnarray*}
-  \sigma_{it}^2 &=& \alpha_0 + \alpha_1 a_{i, t-1}^2 + \beta_1
+  \sigma_{it}^2 &=& \alpha_0 + \alpha_1 z_{i, t-1}^2 + \beta_1
   \sigma_{i,t-1}^2
@@
 Here $Z$ is a random variable that has the same distribution as
-$a_{it}$. In our simulations described hereafter $a_{it}, Z \sim N(0,
+$z_{it}$. In our simulations described hereafter $z_{it}, Z \sim N(0,
 1)$. In this section and the next, we first study situations where no
@@
 \ref{fig:GarchEigDiag1}.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=85 258 103
-  229]{../pics/GarchSpectrumAutocorrelated.pdf}
-  \caption{\small \it Eigenvalue distribution of cross-correlation
-    matrix built from auto-correlated GARCH processes. $\phi$ = 0,
-    0.5, 0.8, 0.955, 0.97 correspond to correlation time $\tau$ =
-    0, 1.00, 3.11, 15.05, 22.76.}
-  \label{fig:GarchSpectrumAutocorrelated}
-\end{figure}
 From this figure one can see that, as auto-correlation strengthens,
@@
 now varies in a larger range than it does with smaller $\phi$.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=0 197 0
-  154]{../pics/garch_eigenvec_Cmax_corr.pdf}
-  \caption{\small \it Eigenvectors' largest component in
-    correspondence to the eigenvalues. N is fixed to 50.}
-  \label{fig:garch_eigenvec_Cmax_corr}
-\end{figure}
 
@@
 to smaller values by increased auto-correlation.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=95 260 111
-    232]{../pics/garch_eigenvec_PR_dist_corr.pdf}
-    \label{fig:garch_eigenvec_PR_dist_corr}
-  }
-  \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=95 260 111
-    232]{../pics/garch_eigenvec_Cmax_dist_corr.pdf}
-    \label{fig:garch_eigenvec_Cmax_dist_corr}
-  }
-  \caption{\small \it \ref{fig:garch_eigenvec_PR_dist_corr}: PDF of
-    the normalized participation ratio
-    (PR). \ref{fig:garch_eigenvec_Cmax_dist_corr}: PDF of the largest
-    component of the eigenvectors. N is fixed to 50. $\phi$ values of
-    0, 0.6000, 0.9000, 0.9900 correspond to correlation time $\tau$ =
-    0, 1.3569, 6.5788, 68.9676}
-\end{figure}
 
@@
 the ratios are plotted versus $\phi$ on log-log scale.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[]{
-    \includegraphics[scale=0.38, clip=true, trim=93 229 115
-    134]{../pics/localization_ratio.pdf}
-    \label{fig:localization_ratio}
-  }
-  \subfigure[]{
-    \includegraphics[scale=0.38, clip=true, trim=87 227 115
-    133]{../pics/localization_ratio2.pdf}
-    \label{fig:localization_ratio2}
-  }
-  \caption{\small \it \ref{fig:localization_ratio}: Ratio of localized
-    eigenvectors versus the auto-correlation strength $\phi$. Black
-    ``+'': ratio of localized eigenvectors as measured by the number
-    of participating basis vectors being lower than 2; Black ``x'':
-    ratio of localized eigenvectors as measured by the largest
-    component being larger than 0.9. Blue curve: quadradic function
-    fitted to ``+''. Green curve: quadratic function fitted to
-    ``x''. There are 27 data points in the plot, corresponding to 27
-    $\phi$ values: 0 to 0.8 with step size 0.05, and 0.9 to 0.99 with
-    step size 0.01. The corresponding values of the correlation time
-    $\tau$ range from 0 to 69. \ref{fig:localization_ratio2}: $\ln
-    (f_\M - f)$ is plotted versus $\ln \phi$, where $f$ stands for the
-    ratio of localized eigenvectors. The fitted curves are linear.}
-\end{figure}
 
@@
 
+\section{Eigenvalue distribution of Wishart Matrix}
+\label{sec:wishart_eigen_dist}
+In this section we summarize the analytic results regarding the
+eigenvalue distribution of a Wishart matrix $\mtx{RR'}$. In the
+simplest case where the elements of $\mtx R$ are all independent of
+each other, i.e. neither auto-correlation nor cross-correlation
+exists, we have \cite{Chiani2012}
+\begin{eqnarray}
+  f_{\lambda}(x_1, \cdots, x_N) &=& K \prod_{i=1}^N e^{-x_i/2}
+  x_i^{(T-N-1)/2} \prod_{i<j}^N (x_i -
+  x_j) \label{eq:wishart_eigen_pdf}\\
+\end{eqnarray}
+where the eigenvalues have been indexed in descending order: $x_1 \geq
+x_2 \geq \cdots \geq x_N$. The normalization constant K is given by
+\begin{equation*}
+  K = {\pi^{N^2/2} \over 2^{NT/2}\Gamma_N(T/2) \Gamma_N(N/2)}
+\end{equation*}
+where the function $\Gamma_m(a)$ is defined as
+\begin{equation*}
+  \Gamma_m(a) = \pi^{m(m-1)/4} \prod_{k=1}^m \Gamma\left(a - {k-1
+      \over 2}\right)
+\end{equation*}
+Moreover, K. Johnsson \cite{Johnsson2000} and I. Johnstone
+\cite{Johnstone2001} showed that, at the {\it absence} of
+autocorrelations and in the asymptotic limit $N, T \to \infty$, $N/T
+\to q < \infty$, the maximum eigenvalue  $\lambda_1$ follows the
+Tracy-Widom distribution (denote $\mathscr{TW}_1$ here) when properly
+relocated and rescaled:
+\begin{equation*}
+  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} \sim \mathscr{TW}_1
+\end{equation*}
+where $\mu_{NT}$ and $\sigma_{NT}$ are given by
+\begin{eqnarray*}
+  \mu_{NT} &=& \left(
+    \sqrt{N-1/2} + \sqrt{T - 1/2}
+  \right)^2 \\
+  \sigma_{NT} &=& \sqrt{\mu_{NT}} \left(
+    {1 \over \sqrt{N-1/2}} + {1 \over \sqrt{T-1/2}}
+  \right)^{1/3}
+\end{eqnarray*}
+The $\mathscr{TW}_1$ distribution has the following cummulative
+distribution function (CDF) \cite{Chiani2012}
+\begin{equation*}
+  F_1(x) = \exp\left[
+    -{1 \over 2} \int_x^\infty dy \left(q(y) + (y-x)q^2(y)\right)
+  \right]
+\end{equation*}
+where $q(y)$ is defined as the solution to the Painlev\'e II differential
+equation
+\begin{equation*}
+  q''(y) = yq(y) + 2q^3(y)
+\end{equation*}
+which is unique when imposing the condition
+\begin{equation*}
+  q(y) \sim \text{Ai}(y) \text{ as } y \to \infty
+\end{equation*}
+
+Then Marco Chiani showed recently that the $\mathscr{TW}_1$ distribution
+can be well approximated by a gamma distribution based on his proof
+that the exact distribution of the maximum eigenvalue is a mixture of
+gamma distributions. Specifically,
+\begin{equation}\label{eq:TracyWidom-Gamma}
+  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} + \alpha \sim
+  \mathscr{G}(k ,\theta)
+\end{equation}
+where $\mathscr{G}(k, \theta)$ denotes the Gamma distribution with
+parameters $k$ and $\theta$ \cite{Chiani2012}. The probability
+density function (PDF) of the gamma distribution is given by
+\begin{equation*}
+  f_\gamma(x; k, \theta) = {1 \over \Gamma(k) \theta^k} x^{k-1}
+  e^{-x/\theta}
+\end{equation*}
+Moreover, the first 3 moments of the distribution are simple:
+\begin{eqnarray*}
+  \text{mean} &=& k\theta \\
+  \text{variance} &=& k\theta^2 \\
+  \text{skewness} &=& 2/\sqrt{k}
+\end{eqnarray*}
+
+\section{Collection of Figures and Tables}
+\begin{figure}[htb!]
+  \centering
+  \subfigure[ACF of $z_t$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_quotient_acf.pdf}
+    \label{fig:nordea_15min_quotient_acf}
+  }
+  \subfigure[ACF of $z_t^2$]{
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
+    \label{fig:nordea_15min_quotient_squared_acf}
+  }
+  \caption{\small \it Nordea 15min $z_t$ and $z_t^2$ auto-correlation
+    function (ACF).}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Normal Probability Plot of $y_t$]{
+    \includegraphics[scale=0.4, clip=true, trim=76 256 109
+    214]{../pics/nordea2_y_normplot.pdf}
+    \label{fig:nordea_15min_y_qq}
+  }
+  \subfigure[ACF of $y_t$]{
+    \includegraphics[scale=0.4, clip=true, trim=92 258 104
+    218]{../pics/nordea_15min_y_acf.pdf}
+    \label{fig:nordea_15min_y_acf}
+  }
+  \caption{\small \it Nordea 15min $y_t$ normal probability plot and
+    autocorrelations}
+\end{figure}
+
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[scale=0.4, clip=true, trim=90 228 115
+    226]{../pics/GaussianMarkovSpectrumPDF.pdf}
+  \end{center}
+  \vspace{-10mm}
+  \caption{\small \it
+      Eigenvalue distribution with $\tau$ ranging from 0 to
+      3. The 1st blue line, which is shown as stairs, is the
+      theoretical eigenvalue distribution according to the
+      Marcenko-Pastur law: $f(\lambda) = \sqrt{(\lambda_+/\lambda - 1)(1
+        - \lambda/\lambda_-)}/2\pi q \sigma^2$, where $\lambda_{\pm} =
+      \sigma^2(1 \pm \sqrt{q})^2$ In the simulation we have
+      chosen $q = N/T = 50/1000 = 0.05$ and $\sigma=1$. For each value
+      of $\tau$ we generate 2000 instances of $N \times T$ random matrix
+      $R$, and compute C as $C=RR'/T$. Hence each curve in the figure is
+      constructed from 2000 sets of eigenvalues.
+    }
+  \label{fig:GaussianMarkovSpectrumPDF}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4, clip=true, trim=100 226 116
+  133]{../pics/Gaussian_mineig.pdf}
+  \caption{\small \it The minimum eigenvalue versus
+    auto-correlation strength $\phi$. For each value of $\phi$ 2000
+    random matrices are generated and their eigenvalues are
+    calculated. The minimum eigenvalue of each random matrix is noted
+    and the mean of the 2000 such minimum eigenvalues are plotted
+    against the chosen value of $\phi$. 20 values of $\phi$ are
+    included in the plot, ranging from 0 to 0.95 with step size 0.05.}
+  \label{fig:Gaussian_mineig}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.5, clip=true, trim=0 238 0
+    197]{../pics/GaussianMarkov005MaxEigCDF_loglog.pdf}
+  \end{center}
+  \caption{\small \it Cummulative distribution function (CDF) of the maximum
+    eigenvalue ($\lambda_1$). Blue: empirical CDF. Red: CDF of the
+    fitting gamma distribution. Each pair of CDFs correspond to a
+    fixed autocorrelation strength ($\phi$). The parameters $k$ and
+    $\theta$ of the gamma distribution are fit to data by matching the
+    2nd and the 3rd moments of the gamma distribution to the
+    corresponding moments of the empirical distribution. Then the
+    parameter $\alpha$ in equation \ref{eq:TracyWidom-Gamma} is chosen
+    to be $\alpha = k\theta - \E\left({\lambda_1 - \mu_{NT} \over
+        \sigma_{NT}}\right)$. The curves are plotted on log-log
+    scale.}
+  \label{fig:GaussianMarkov005MaxEigCDF_loglog}
+  %\vspace{-10mm}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=50 233 63
+  136]{../pics/Volvo_B_30m_returns.pdf}
+  \caption{\small \it Volvo B 30min returns' unconditional distribution fit to
+    the Stochastic Volatility model. The time series runs from
+    2013/10/10 to 2014/03/12. Left: $P(r < x)$ with $x < 0$; Right:
+    $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
+  \label{fig:Volvo_B_30m_returns}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=65 241 72
+  143]{../pics/Nordea_30m_returns.pdf}
+  \caption{\small \it Nordea Bank 30min returns' unconditional
+    distribution fit to the Stochastic Volatility model. The time
+    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
+    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
+  \label{fig:Nordea_30m_returns}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=21 241 28
+  146]{../pics/Ericsson_30m_returns.pdf}
+  \caption{\small \it Ericsson B 30min returns' unconditional
+    distribution fit to the Stochastic Volatility model. The time
+    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
+    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
+  \label{fig:Ericsson_30m_returns}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=0 197 0
+  154]{../pics/garch_eigenvec_Cmax_corr.pdf}
+  \caption{\small \it Eigenvectors' largest component in
+    correspondence to the eigenvalues. N is fixed to 50.}
+  \label{fig:garch_eigenvec_Cmax_corr}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=85 258 103
+  229]{../pics/GarchSpectrumAutocorrelated.pdf}
+  \caption{\small \it Eigenvalue distribution of cross-correlation
+    matrix built from auto-correlated GARCH processes. $\phi$ = 0,
+    0.5, 0.8, 0.955, 0.97 correspond to correlation time $\tau$ =
+    0, 1.00, 3.11, 15.05, 22.76.}
+  \label{fig:GarchSpectrumAutocorrelated}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=95 260 111
+    232]{../pics/garch_eigenvec_PR_dist_corr.pdf}
+    \label{fig:garch_eigenvec_PR_dist_corr}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=95 260 111
+    232]{../pics/garch_eigenvec_Cmax_dist_corr.pdf}
+    \label{fig:garch_eigenvec_Cmax_dist_corr}
+  }
+  \caption{\small \it \ref{fig:garch_eigenvec_PR_dist_corr}: PDF of
+    the normalized participation ratio
+    (PR). \ref{fig:garch_eigenvec_Cmax_dist_corr}: PDF of the largest
+    component of the eigenvectors. N is fixed to 50. $\phi$ values of
+    0, 0.6000, 0.9000, 0.9900 correspond to correlation time $\tau$ =
+    0, 1.3569, 6.5788, 68.9676}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.38, clip=true, trim=93 229 115
+    134]{../pics/localization_ratio.pdf}
+    \label{fig:localization_ratio}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.38, clip=true, trim=87 227 115
+    133]{../pics/localization_ratio2.pdf}
+    \label{fig:localization_ratio2}
+  }
+  \caption{\small \it \ref{fig:localization_ratio}: Ratio of localized
+    eigenvectors versus the auto-correlation strength $\phi$. Black
+    ``+'': ratio of localized eigenvectors as measured by the number
+    of participating basis vectors being lower than 2; Black ``x'':
+    ratio of localized eigenvectors as measured by the largest
+    component being larger than 0.9. Blue curve: quadradic function
+    fitted to ``+''. Green curve: quadratic function fitted to
+    ``x''. There are 27 data points in the plot, corresponding to 27
+    $\phi$ values: 0 to 0.8 with step size 0.05, and 0.9 to 0.99 with
+    step size 0.01. The corresponding values of the correlation time
+    $\tau$ range from 0 to 69. \ref{fig:localization_ratio2}: $\ln
+    (f_\M - f)$ is plotted versus $\ln \phi$, where $f$ stands for the
+    ratio of localized eigenvectors. The fitted curves are linear.}
+\end{figure}
+
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    & $\psi$ & $\sigma$ & $\bar{v}$ & $\mu$ \\
+    \hline
+    Volvo B & -1.5747e-02 & 4.6929e-01 & -6.2304e+00 &
+    -5.3040e-05 \\
+    Nordea Bank & 1.7234e-02 & 4.2780e-01 & -6.3758e+00 & 7.5230e-05
+    \\
+    Ericsson B & 1.8486e-02 & 4.4905e-01 & -6.2834e+00 & -9.2817e-05 \\
+    \hline
+  \end{tabular}
+  \caption{\small \it Parameter Values of Selected Assets'
+    returns. The time series are 30-minute returns and run from
+    2013/10/10 to 2014/03/12.}
+  \label{tab:assets_params}
+\end{table}
+
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    & mean & std & skewness & kurtosis \\
+    \hline
+    \multirow{2}{*}{Volvo B} & -8.6766e-05 & 2.4500e-03 &
+    -2.7217e-02 & 7.3394e+00 \\
+    & -6.9281e-05 & 2.4538e-03 & -7.2677e-02 & 7.2472e+00 \\
+    \hline
+    \multirow{2}{*}{Nordea Bank} & 7.4821e-05 & 2.0500e-03 &
+    6.9433e-02 & 6.7427e+00 \\
+    & 8.8983e-05 & 2.0443e-03 & 6.7125e-02 & 6.2446e+00 \\
+    \hline
+    \multirow{2}{*}{Ericsson B} & -8.2606e-05 & 2.2939e-03 &
+    2.6586e-01 & 7.6716e+00 \\
+    & -7.5675e-05 & 2.2844e-03 & 7.8573e-02 & 6.7299e+00 \\
+    \hline
+  \end{tabular}
+  \caption{\small \it Moments of Selected Assets' returns. For each return
+    series, the 1st row contains the sample moments, while the 2nd
+    constains those computed with MLE parameters. The time series are
+    30-minute returns and run from 2013/10/10 to 2014/03/12.}
+  \label{tab:assets_moments}
+\end{table}
+
 \bibliographystyle{plain}

### 2014-05-08 thesis/price_models.tex
This is the version to be used for job searching.
Change the section about Nordea 15min in 1st quarter 2012.

@@
 In this chapter we review some of the discrete-time return models and
-fit them to intraday returns. The intention is to assess the
-applicability of these models and to understand their statistical
-properties. The implications of these properties on the returns'
-correlation and covariance matrix are studied in chapters
-\ref{chp:Gaussian} and \ref{chp:CrossCorrelationFat}.
-
-\S\ref{sec:Garch_model} reviews the GARCH model using the Nordea
-15-minute returns as an example. \S\ref{sec:SLV_model} examines the
-stochastic log-volatility model and discusses the possibility of
-describing the residuals of the log-volotility series using the
-Johnson Su distribution. The same Nordea return series is fitted as an
-example. \S\ref{sec:XieCalc} derives the unconditional distribution
-function of the stochastic log-volotility model and compares the
-analytic results to the 30-minute returns of a number of Swedish
-stocks.
-
-\section{Gaussian Distribution}
+fit them to intraday returns. The intention is to compare these models
+in terms of forecast accuracy and to understand their statistical
+properties. In the following we first describe these models briefly,
+then in
+
+% \S\ref{sec:Garch_model} reviews the GARCH model using the Nordea
+% 15-minute returns as an example. \S\ref{sec:SLV_model} examines the
+% stochastic log-volatility model and discusses the possibility of
+% describing the residuals of the log-volotility series using the
+% Johnson Su distribution. The same Nordea return series is fitted as an
+% example. \S\ref{sec:XieCalc} derives the unconditional distribution
+% function of the stochastic log-volotility model and compares the
+% analytic results to the 30-minute returns of a number of Swedish
+% stocks.
+
+\begin{enumerate}
+\item{\bf Gaussian Distribution}
+
 The justification of modeling return series as independent,
@@
 
-The picture depicted above is of course overly simplified, and as we discussed in
-\S\ref{sec:StylizedFacts}, the distribution of returns is not really
-Gaussian. Nevertheless, the assumption of Gaussian distributed returns
-underlies such important theories as Black and Scholes theory of
-option pricing. Thus, albeit hugely inaccurate, the Gaussian
-distribution as a returns model shall not be forgotten. In the next
-few sections, we discuss some more realistic returns models.
-
-\section{GARCH models} \label{sec:Garch_model}
+The picture depicted above is of course overly simplified, and as we
+discussed in \S\ref{sec:StylizedFacts}, the distribution of returns is
+not really Gaussian. Nevertheless, the assumption of Gaussian
+distributed returns underlies such important theories as Black and
+Scholes theory of option pricing. Thus, albeit hugely inaccurate, the
+Gaussian distribution as a returns model shall not be forgotten. In
+the next few sections, we discuss some more realistic returns models.
+
+\item{\bf GARCH models}
 ``GARCH'' is the acronym for ``Generalized Autoregressive Conditional
@@
 \end{equation}
-where $\epsilon_t$ is a random variable of a prescribed distribution,
-the most common ones being Gaussian, Student's t, Asymmetric Student's
-t, etc; $\sigma_t^2$ is the variance of $\epsilon_t$
-\cite{Bollerslev86}. The autocorrelation function of $\epsilon_t^2$ is
-given by (c.f. \cite{Bollerslev87})
+where $\epsilon_t$ is termed the innovation of the return $r_t$ and is
+a random variable of a prescribed distribution, the most common ones
+being Gaussian, Student's t, Asymmetric Student's t, etc; $\sigma_t^2$
+is the variance of $\epsilon_t$ \cite{Bollerslev86}. The
+autocorrelation function of $\epsilon_t^2$ is given by
+(c.f. \cite{Bollerslev87})
 \begin{eqnarray*}
@@
 
-To assess the applicability of GARCH models at the face of intra-day
-returns, we study the {\it Nordea Bank 15min} returns sampled during
-the period 2012/01/16 - 2012/04/20. Firstly we plot the autocorrelation
-function (ACF) of $\epsilon_t^2$ from this sample, as shown
-in figure \ref{fig:nordea_15min_acf}.
-
-At the absence of GARCH effects, the ACF is expected to have an
-asymptotic Gaussian distribution with mean 0 and variance 1/T
-\cite{Bollerslev86, Bollerslev87}, which is clearly not the case in
-figure \ref{fig:nordea_15min_acf}: The first 5 autocorrelations have
-comparable sizes and do not fall off as in a Gaussian
-scheme. Moreover, figure \ref{fig:nordea_15min_vlt_acf} shows 
+\item {\bf Stochastic Volatility Models}\label{sec:SLV_model}
+For the purpose of intraday returns, we specify the stochastic
+volatility model as
+\begin{eqnarray}
+   r_{t, t-h} &=& \ln p_{t} - \ln p_{t-h} \nonumber \\
+   r_{t, t-h} &=& \mu + \sigma_{t, t-h} b_t \label{eq:SLV_spec}
+\end{eqnarray}
+where $p_t$ is the price of the asset at time $t$; $b_t \sim N(0,
+1)$; $r_{t, t-h}$ is the return in the time interval $[t-h, t]$.
+For simplicity, in the rest of this chapter we shall just write
+$t$ for the subscript ``t, t-h'', since the time interval $h$ is fixed
+for each time series and is a known constant.
+
+Andersen and Bollerslev et al proved the following in
+\cite{Andersen03} (theorem 2):
+\begin{equation}
+  \label{eq:normal_r}
+  r_t|\mathcal{F}_{t-h} \sim N(\int_{0}^h \mu_{t-h+s} ds, \int_{0}^h
+  \sigma_{t-h+s}^2 ds)
+\end{equation}
+In plain words, given all the information up to time $t-h$, the
+distribution of $r_t$ is Gaussian with integrated mean and variance.
+The variance of this conditional distribution, $\int_{0}^h
+\sigma_{t-h+s}^2 ds$, can be approximated by \cite{Protter05,
+  Andersen03}:
+\begin{equation}
+  \label{eq:rv_def}
+  \int_{0}^h \sigma_{t-h+s}^2 ds = \sum_{k=1}^n \left(
+    p_{t-h+kh/n} - p_{t-h+(k-1)h/n} \right)^2
+\end{equation}
+where $n$ is a chosen constant. The square root of the right hand side
+of equation \ref{eq:rv_def} is the realized volatility, call it
+$\hat{\sigma}_t$.
+
+With the availability of transaction data, estimating conditional
+volatility using returns sampled at a higher frequency (realized
+volatility) gives superior accuracy and reliability. However, at which
+frequency the time series should be sampled (the choice of $n$) in
+order to give an unbiased and consistent estimate of the volatility is
+not a trivial question. Naively one would believe that the often the
+series is sampled, the better the estimate, but in fact, due to noise
+introduced by market micro-structure, there is an optimal sampling
+frequency, depending on $h$. While the method to determine this
+optimal frequency is a subject of debate (see for example
+\cite{Sahalia05}), it is not difficult to find a fairly satisfactory
+frequency in practice:
+
+Keeping equation \ref{eq:normal_r} in mind, one can simply try a few
+frequencies and compare the distribution of $(r_t -
+\E(r_t))/\hat{\sigma}_t$ with the standard Gaussian. If the two match,
+$\hat{\sigma}_t$ is a good approximation of
+\[
+\sigma_t = \left(\int_{0}^h \sigma_{t-h+s}^2 ds \right)^{1/2}
+\]
+
+\end{enumerate}
+
+To compare GARCH and stochastic log-volatility models at the face of
+intraday returns, we study the {\it Nordea Bank} 15-minute returns
+during the period 2012/01/16 - 2012/04/20 in \S\ref{chp:nordea_15min},
+and {\it Volvo B} 30-minute returns during 2013/10/10 - 2014/04/04.
+
+\section{Case Study: Nordea in 1st quarter 2012}
+\label{chp:nordea_15min}
+In this section we investigate the Nordea 15-minute returns sampled
+during the period 2012/01/16 - 2012/04/20. In total, these amount to
+2022 returns. We use the first 80\% (1617) for model estimation and  the
+remaining 20\% (405) for comparing with model forcasts. In section
+\ref{sec:nordea_15min_garch} we study the series with a GARCH model 
+and in section \ref{sec:nordea_15min_arima} we study it with a
+stochastic volatility model.
+
+\subsection{GARCH Model}\label{sec:nordea_15min_garch}
+First of all, to check for GARCH effects, we plot the auto-correlation
+function (ACF) of the squared returns. This is shown in figure
+\ref{fig:nordea_15min_acf}. At the absence of GARCH effects, the ACF
+is expected to have an asymptotic Gaussian distribution with mean 0
+and variance 1/T \cite{Bollerslev86, Bollerslev87}, which is clearly
+not the case in figure \ref{fig:nordea_15min_acf} --- the first 5
+autocorrelations have comparable sizes and do not fall off as in a
+Gaussian scheme. Moreover, figure \ref{fig:nordea_15min_vlt_acf} shows
 even more clearly that the conditional variances of the series are
@@
 
+Starting with a GARCH(1,1) model and taking advantage of the knowledge
+that the log-volatility $\ln \sigma_t$ has seasonality $s=33$, we fit
+to the return series a GARCH(33, 33) model, limiting to lags 1 and 33
+for both ARCH and GARCH parameters.
+\begin{eqnarray*}
+  r_t &=& \mu + €_t \\
+  €_t &=& \sigma_t z_t \\
+  \sigma^2_t &=& \alpha_0 + \alpha_1 €^2_{t-1} + \alpha_s €^2_{t-s} +
+  \beta_1 \sigma^2_{t-1} + \beta_s \sigma^2_{t-s}
+\end{eqnarray*}
+Via maximum likelihood estimation, parameter values listed in table
+\ref{tab:nordea_15min_garch} are obtained.
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|c|}
+    \hline
+    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_s$ & $\beta_1$ &
+    $\beta_s$ \\
+    \hline
+    Value & $4.7833 \times 10^{-7}$ & 0.1600 & 0.0667 & 0.6846 &
+    0.0342 \\
+    \hline
+  \end{tabular}
+  \caption{\small \it GARCH model parameters}
+  \label{tab:nordea_15min_garch}
+\end{table}
+
+\subsection{Stochastic Volatility Model}\label{sec:nordea_15min_arima}
+For the Nordea Bank 15-minute returns under consideration, it can be
+verified that the square root of the sum of squared 30-second returns
+makes a good proxy for the volatility. This can be seen from the
+probability plot of $z_t = (r_t - \E(r_t))/\hat{\sigma}_t$ (figure
+\ref{fig:nordea_bank_15min_z_prob}), i.e. the quotient of the 
+15-minute returns over the volatility proxy.
 \begin{figure}[htb!]
   \centering
-  \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_acf.pdf}
-    \label{fig:nordea_15min_acf}
-  }
-  \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_vlt_acf.pdf}
-    \label{fig:nordea_15min_vlt_acf}
-  }
-  \caption{\small \it Auto-correlations among Nordea Bank 15min
-    returns. \ref{fig:nordea_15min_acf}: Auto-correlations (ACF) of the
-    squared returns; \ref{fig:nordea_15min_vlt_acf}: Auto-correlations
-    (ACF) of the conditional variances.}
-\end{figure}
-
-According to Andersen and Bollerslev et al \cite{Andersen03},
-$\epsilon_t$ can be decomposed as $\epsilon_t = \sigma_t z_t$, where
-$z_t \sim N(0, 1)$ and $\sigma_t$ can be estimated using realized
-volatility. To find out whether this is true for the aforementioned
-series we look at the probability plot of $\epsilon_t/\sigma_t$
-(figure \ref{fig:nordea_bank_15min_z_prob}), which is consistent with
-a standard Gaussian distribution except for a few outliers at the two
-ends. Hence the fat tails of the probability density function can be
-entirely accounted for by GARCH effects, implying a Gaussian
-conditional distribution is sufficient for $z_t$.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=80 258 104
+    \includegraphics[scale=0.4, clip=true, trim=80 258 104
     220]{../pics/nordea_bank_15min_z_prob.pdf}
-  \caption{\small \it{Probability plot of $z_t = \epsilon_t /
-      \sigma_t$. $\epsilon_t$ are derived from Nordea Bank 15min
-      returns while $\sigma_t$ are realized volatilities calculated
+  \caption{\small \it{Probability plot of $z_t =
+      (\epsilon_t-\E(\epsilon_t))/\sigma_t$. $\epsilon_t$ are derived
+      from Nordea Bank 15min returns while $\sigma_t$ are realized
+      volatilities calculated
       using 30s returns within each 15min interval. Horizontal axis:
@@
 \end{figure}
-
-To determine the orders p and q of the GARCH model, we again check
-the partial autocorrelation function of $\epsilon_t^2$. As just
-mentioned, the first 5 partial autocorrelations are significant. Thus
-we can tentatively estimate a GARCH(1, 5) model. Then by applying the
-Lagrange multiplier test and accordingly removing insignificant
-parameters, we obtain a GARCH(1, 3) model
-\ref{tab:nordea_15min_garch}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    parameter & value & std. error & t statistic \\
-    \hline
-     $\alpha_0$ &   1.05668e-06 &   3.56871e-07 &    2.96096\\
-     \hline
-     $\beta_1$  &     0.687212  &   0.0299314   &    22.9596\\
-     \hline
-     $\alpha_1$ &      0.128614 &   0.0268144   &    4.79643\\
-     \hline
-     $\alpha_3$ &    0.0568022  &   0.0301503   &    1.88397 \\
-     \hline
-  \end{tabular}
-  \caption{\small \it GARCH(1, 3) model of Nordea Bank 15min returns}
-  \label{tab:nordea_15min_garch}
-\end{table}
-
-\section{ARIMA Log-volotility Models}\label{sec:SLV_model}
-With the availability of transaction data, estimating conditional
-volatility using returns sampled at a higher frequency (realized
-volatility) gives superior acuracy and reliability. However, at which
-frequency the time series should be sampled in order to give an
-unbiased and consistent estimate of the volatility is not a trivial
-question. Naively one would argue by the doctrine of statistics that
-the often the series is sampled, the better the estimate. This is also
-well supported in theory, since the quadratic variation which is the
-theoretial variance of the time series (see \cite{Andersen03}, theorem
-1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
-\S6, theorem 22):
-\[
-\var(p_{t+h} - p_t) = [r, r]_{t+h} - [r, r]_t = \lim_{n \to \infty}
-\sum_{T^i \in \pi_n} (p^{T_{i+1}} - p^{T_i})^2
-\]
-where $p_t$ is the log price process and $r_t = p_t - p_0$. $\pi_n$
-is a refining partition of the interval $[t, t+h]$.
-
-However, due to the noise introduced by market micro-structure, there
-is actually an optimal sampling frequency relative to $h$. While the
-method to determine this optimal frequency is a subject of debate (see
-for example \cite{Sahalia05}), it is not difficult to find a fairly
-satisfactory frequency in practice.
-
-Andersen and Bollerslev et al proved in \cite{Andersen03} that, with
-the conditional volatility well approximated by the realized
-volatility, $r_{t+h, h} = p_{t+h} - p_t$ is governed by a Gaussian
-distribution (see theorem 2 of \cite{Andersen03}):
-\[
-r_{t+h, h} | \sigma(\mu_{t+s}, \sigma_{t+s})_{s \in [0, h]} \sim
-N(\int_{0}^h \mu_{t+s} ds, \int_{0}^h \sigma_{t+s}^2 ds)
-\]
-
-Thus one can try a few frequencies and compare the distribution of
-$r_{t+h, h}/\hat{\sigma}_{t+h, h}$ with the standard Gaussian. If the
-two match, $\hat{\sigma}_{t+h, h}$ is a good estimate of $\sigma_{t+h,
-  h}$, the volatility of $r_{t+h, h}$.
-
-For the Nordea Bank 15-minute returns, $h = 15\text{min}$, and a
-reasonally good sampling frequency is 1/30sec, as confirmed by the
-probability plot \ref{fig:nordea_bank_15min_z_prob} as well as by the
-autocorrelation functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of
-$(r_{t+h, h} / \sigma_{t+h, h})^2$ which are shown in
-\ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} respectively.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of $r_{t+h, h} / \sigma_{t+h, h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_acf.pdf}
-    \label{fig:nordea_15min_quotient_acf}
-  }
-  \subfigure[ACF of $(r_{t+h, h} / \sigma_{t+h, h})^2$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
-    \label{fig:nordea_15min_quotient_squared_acf}
-  }
-  \caption{\small \it Nordea 15min $r_t/\sigma_t$ and
-    $(r_t/\sigma_t)^2$ ACF.}
-\end{figure}
-
-We see in figure \ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
-autocorrelation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
-$(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
-series is i.i.d Gaussian distributed. As a result, the problem of
-modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
-  h}$. Andersen and Bollerslev et al reported that, for the exchange
-rates between Deutch mark, yen and dollar, $\ln \sigma_{t+h, h}$ is
+% In addition, one can see
+% from figure \ref{fig:nordea_15min_quotient_acf} and
+% \ref{fig:nordea_15min_quotient_squared_acf} that there is essentially
+% no auto-correlation in the $z_t$ or the $z_t^2$ series.
+% \begin{figure}[htb!]
+%   \centering
+%   \subfigure[ACF of $z_t$]{
+%     \includegraphics[scale=0.4, clip=true, trim=95 236 118
+%     200]{../pics/nordea_15min_quotient_acf.pdf}
+%     \label{fig:nordea_15min_quotient_acf}
+%   }
+%   \subfigure[ACF of $z_t^2$]{
+%     \includegraphics[scale=0.4, clip=true, trim=95 236 118
+%     200]{../pics/nordea_15min_quotient_squared_acf.pdf}
+%     \label{fig:nordea_15min_quotient_squared_acf}
+%   }
+%   \caption{\small \it Nordea 15min $z_t$ and $z_t^2$ ACF.}
+% \end{figure}
+
+Andersen and Bollerslev et al reported that, for the exchange
+rates between Deutch mark, yen and dollar, $\ln \sigma_t$ is
 gaussian distributed \cite{Andersen03}. This is, however, not the case
-for our series in question. In fact, in our case, $\ln \sigma_{t+h,
-  h}$ is right skewed (skewness 0.3342) and leptokurtic (kurtosis
+for our series in question. In fact, in our case, $\ln \sigma_t$ is
+right skewed (skewness 0.3342) and leptokurtic (kurtosis 
 6.1006). See figure \ref{fig:nordea_15min_logvol_prob}
@@
   %\vspace{-15mm}
-  \includegraphics[scale=0.5, clip=true, trim=80 223 107
+  \includegraphics[scale=0.4, clip=true, trim=80 223 107
   4]{../pics/nordea_15min_logvol_prob.pdf}
   \caption{\small \it Probability plot of Nordea 15min
-    $\ln\sigma_{t+h, h}$ unconditional distribution}
+    $\ln\sigma_t$ unconditional distribution}
   \label{fig:nordea_15min_logvol_prob}
 \end{figure}
-
-Moreover, the series of $\ln\sigma_{t+h, h}$ shows long-lasting and
+Moreover, the series of $\ln\sigma_t$ shows long-lasting and
 periodic autocorrelations with an apparent period of 33 (see figure
 \ref{fig:nordea_15min_logvol_acf}). This suggests the series may be
-described by an integrated seasonal ARMA\footnote{ARMA ---
-  Autoregressive Moving Average} model. But before we set out
-to specify and estimate this model of the logvolatility
-$\ln\sigma_{t+h, h}$, it is useful to first write down the model of
-the returns $r_t$:
-\begin{eqnarray}
-   r_{t+h, h} &=& \ln p_{t+h} - \ln p_t \nonumber \\
-   r_{t+h, h} &=& \mu + \sigma_{t+h, h} b_t \label{eq:SLV_spec}
-\end{eqnarray}
-where $p_t$ is the price of the asset at time $t$ and $b_t \sim N(0,
-1)$. For simplicity, in the rest of this chapter we shall just write
-$t$ for the subscript ``t+h, h'', since the time interval $h$ is fixed
-for each time series and is a known constant. In the case of the
-Nordea Bank 15-minute returns, h = 15 minutes.
-
-As has just been said, $\ln \sigma_t$ for the series in question has
-long-lasting auto-correlations (often referred to as ``long
-memory''). This suggests simplifying the series by differencing
-\cite{BoxJenkins94}:
+described by a seasonal ARIMA model\footnote{ARIMA ---
+  Integrated Autoregressive Moving Average} model. This we first
+simplify the series by differencing \cite{BoxJenkins94}:
 \[
@@
 
-The autocorrelation function of the differenced process $w_t$, as shown in
-figure \ref{fig:nordea_15min_w_acf}, clearly points to a seasonal
+The autocorrelation function of the differenced process $w_t$, as
+shown in figure \ref{fig:nordea_15min_w_acf}, clearly points to a seasonal
 moving-average model: There are only 4 non-zero autocorrelations in
@@
   \centering
-  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
+  \subfigure[ACF of $\ln\sigma_t$]{
     \includegraphics[scale=0.4, clip=true, trim=95 230 112
@@
   }
-  \caption{\small \it Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$
+  \caption{\small \it Nordea 15min $\ln\sigma_t$ and $w_t$
     autocorrelations}
+  \label{fig:nordea1_15min_acf}
 \end{figure}
@@
 expectation 0.
-% By inverting the seasonal moving average model of $w_t$ we can then
-% infer the noise process $y_t$ \footnotemark, i.e. the residuals:
-% \begin{eqnarray*}
-%   y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
-%   &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
-%   w_t \\
-%   &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
-%   w_{t-k}
-% \end{eqnarray*}
-% where $\bot(\cdot)$ denotes the floor function, i.e. the largest
-% integer smaller than its argument.
-% \footnotetext{
-%   The autoregressive process resulting from the inversion is inifinite
-%   in extent. Hence one must choose a cut-off value for k. For the
-%   present case of Nordea 15min log-volatility, I choose the value
-%   $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
-%   $k_{\top}$, is around 550. To avoid data reduction of this amount,
-%   we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
-%   corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $\E(w_t) = \E(y_t
-%   - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
-% }
-
-In order to forecst the $w_t$ process, and hence the return process
+
+In order to forecast the $w_t$ process, and hence the return process
 itself, we must also know the distribution of $y_t$. Moreover, to
@@
 
-If $y_t$ is indeed i.i.d random variables as we assumed, we only need
-to work out its unconditional distribution. This is shown in figure
-\ref{fig:nordea_15min_y_qq} and \ref{fig:nordea_15min_y_acf}.
+Figure \ref{fig:nordea_15min_y_qq} shows the normal probability plot of
+$y_t$. It is evident from this figure that $y_t$ has fat tails.
 \begin{figure}[htb!]
   \centering
-  \subfigure[QQ-plot of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_y_qq.pdf}
+  \includegraphics[scale=0.4, clip=true, trim=78 255 109
+  123]{../pics/nordea2_y_normplot.pdf}
+  \caption{\small \it Nordea 15min $y_t$ normal probability plot.}
   \label{fig:nordea_15min_y_qq}
-  }
-  \subfigure[ACF of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_y_acf.pdf}
-    \label{fig:nordea_15min_y_acf}
-  }
-  \caption{\small \it Nordea 15min $y_t$ QQ-plot and autocorrelations}
 \end{figure}
-In addition, the measured moments of $y_t$ are shown in table
-\ref{tab:nordea_15min_y_moments}:
+In addition, a simple calculation reveals that the distribution of
+$y_t$ has skewness 0.2988 (shown in table
+\ref{tab:nordea_15min_y_moments}).
 \begin{table}[htb!]
@@
 \end{table}
-
-If we ignore the weak autocorrelations in this crude estimate of
-$y_t$ and assume $y_t$ to be i.i.d random variables governed
-by a skewed and leptokertic distribution, the parameters of this
-distribution together with $\theta$ and $\Theta$ can be determined via
-maximum likelihood estimate (MLE).
-
-It turns out that $y_t$ can be well described by a Johnson Su
-distribution \cite{Shang2004}:
+Based on this information, we find that $y_t$ can be well
+described by a Johnson Su distribution \cite{Shang2004}:
 \[
@@
   \centering
-    \includegraphics[scale=0.5, clip=true, trim=92 229 116
+    \includegraphics[scale=0.4, clip=true, trim=92 229 116
     133]{../pics/nordea_15min_y_js_fit.pdf}
-    \caption{\small \it Nordea 15min $y_t$ fitted to a Johnson Su
-      distribution. Horizontal: values of $y_t$, denoted x; Vertical:
-      $\ln\left(P(y_t < x)\right)$.}
+    \caption{\small \it Nordea 15min residuals $y_t$ fitted to a
+      Johnson Su distribution. Horizontal: values of $y_t$, denoted x;
+      Vertical: $\ln\left(P(y_t < x)\right)$.}
     \label{fig:nordea_15min_y_js_fit}
@@
 
-To verify that the log-volatility can in general be described by
-ARIMA(p, d, q) models, we apply the above procedure to a number of
-Swedish stocks whose data have been collected by the author with
-home-developed software. In \S\ref{sec:nordea} and \S\ref{sec:volvo}
-we model the log-volatilities of Nordea Bank 15-minute returns and
-Volvo B 30-minute returns. Both return series span the period
-2013/10/10 - 2014/04/04. They are computed using minute-by-minute
-average prices, and their variances, i.e. volatilities squared, are
-computed as the sum of squared higher-frequency returns, which are in
-turn computed using second-by-second prices. Whenever a needed average
-price cannot be computed due to the lack of price observations in the
-relevent time interval, the value of the previous average price is
-assumed.
-
-After modeling, we perform 1-step-ahead forecast using the estimated
-models, and compare the predicted log-volatilities with the actual
-data.
-
-\subsection{Nordea Bank Log-Volatility}\label{sec:nordea}
-In this section we model the log-volatility of {\it Nordea Bank}
-15-minute returns during the period 2013/10/10 - 2014/04/04, which
-contains 3879 data points. The squared volatilities, i.e. the
-variances, are computed as the sum of squared 30-second returns. Among
-the 3879 data points, we use the first 80\% for model estimation and
-the last 20\%, i.e. 776 data points for forecast verification.
-
-This Nordea series has a auto-correlation structure similar to its
-predecessor that has been studied in previous sections. We show it in
-figure \ref{fig:nordea2_w_acf}. The reader may want to compare it with
-figure \ref{fig:nordea_15min_w_acf}.
+\subsection{Comparison of the Forecasts}
+In this section we compare the one-step-ahead forecasts from the GARCH
+model and from the stochastic volatility model. For this purpose, we
+compute the difference between a forecast $\ln \sigma^F_t$ and its measured
+counterpart, i.e. the realized volatility of the same period $\ln
+\hat{\sigma}_t$.
+
+First of all, we look at the means and standard deviations of $\ln \sigma^F_t -
+\ln \hat{\sigma}_t$, which are listed in table \ref{tab:nordea_2012}.
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    & SV & GARCH & Sample mean \\
+    \hline
+    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0040 & -0.0008 &
+    -0.2210 \\
+    \hline
+    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.2659 & 0.3011 &
+    0.2893 \\
+    \hline
+  \end{tabular}
+  \caption{\small \it Mean and standard Deviation of the forecasts'
+    distribution. ``Sample mean`` refers to the mean of the realized
+    volatilities in the first 80\% of the data set. ``SV'': Stochastic
+    volatility.}
+  \label{tab:nordea_2012}
+\end{table}
+It is seen from table \ref{tab:nordea_2012} that, on average, the SV
+model over-estimates while GARCH under-estimates. In terms of the
+standard deviation of $\ln \sigma^F_t - \ln \hat{\sigma}_t$, the SV
+model wins with a small margin. In contrast, the sample mean forecast
+clearly under-estimates the log-volatilities to a large extent --- the
+efforts of building models has not been wasted.
+
+Figure \ref{fig:nordea_2012} compares the 3 kinds of forecasts by
+plotting the distribution function and the complementary distribution
+function of $\ln \sigma^F_t - \ln \hat{\sigma}_t$. Here one can see
+that the SV model yields a better quality of forecasts than does GARCH 
+with regard to the under-estimates, and essentially the same quality
+with regard to the over-estimates.
+
 \begin{figure}[htb!]
   \centering
-  \includegraphics[scale=0.4, clip=true, trim=95 230 112
-  235]{../pics/nordea2_w_acf.pdf}
-  \caption{\small \it Auto-correlation function of $w_t$}
-  \label{fig:nordea2_w_acf}
+    \includegraphics[scale=0.55, clip=true, trim=27 287 0
+    154]{../pics/nordea_2012.pdf}
+  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
+    sample mean forecasts. Left: cumulative distribution function;
+    Right: complementary distribution function.}
+  \label{fig:nordea_2012}
 \end{figure}
-The significant auto-correlations and their respective lags are listed
-in table \ref{tab:nordea2_w_acf}.
+
+Another measure of the forecasts' quality can be the percentage of good
+forecasts, where the criterion of ``good'' is defined, respectively,
+as deviating no more than 1\%, 5\%, or 10\% from the measured value of
+the realized volatility. Table \ref{tab:nordea_2012_good} shows the
+respective percentage of the 3 kinds of forecasts. Again in this table
+it is seen that the SV model gives more accurate forecasts than does
+GARCH.
 \begin{table}[htb!]
   \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    lag & 1 & 32 & 33 & 34 \\
-    \hline
-    ACF &   -0.4805 & 0.2861 &  -0.5034 & 0.2093 \\
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    ${\ln \sigma^F_t - \ln \hat{\sigma}_t \over |\ln \hat{\sigma}_t|}$
+    & SV & GARCH & sample mean \\
+    \hline
+    1\% & 0.2173 & 0.1383 & 0.1136 \\
+    \hline
+    5\% & 0.7457 & 0.7062 & 0.5333 \\
+    \hline
+    10\% & 0.9728 & 0.9383 & 0.8914 \\
     \hline
   \end{tabular}
-  \caption{\small \it Significant auto-correlations of $w_t$}
-  \label{tab:nordea2_w_acf}
+  \caption{\small \it The percentage of ``good'' forecasts when the
+    criterion of being good is deviating no more than 1\%, 5\% or
+    10\%.}
+  \label{tab:nordea_2012_good}
 \end{table}
 
-Apparently this series also has the seasonality $s = 33$, but judging
-from the relatively big differences between auto-correlations at lags
-32 and 34, we don't expect the moving average process to be in the
-multiplicative form of equation \ref{eq:nordea_w}. Instead, we model
-$\ln\sigma_t$ has
-\begin{eqnarray}
-  (1 - B)(1 - B^s) \ln\sigma_t &=& (1 - \theta_1 B - \theta_s B^s -
-  \theta_{s+1} B^{s+1} ) y_t \label{eq:nordea2_v}
-\end{eqnarray}
-The parameters are etimated to be
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    parameter & $\theta_1$ & $\theta_{33}$ & $\theta_{34}$ & $\var(y_t)$ \\
-    \hline
-    value & 0.8339 & 0.8448 & -0.7067 & 0.10071 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it: Parameter values of $\ln \sigma_t$}
-\end{table}
-
-The residuals $y_t$ do not have a Gaussian distribution, as confirmed
-by the probability plot shown on the left of figure
-\ref{fig:nordea2_y}. The non-normality, however, cannot be accounted
-by the sine hypobolic transformation of Johnson Su, because of the
-negative skewness of -1.2669 found in table
-\ref{tab:nordea2_y_mmt}. How the non-normality can be accomodated is
-quite another discussion and we do not pursue it here.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.4, clip=true, trim=20 220 30
-  115]{../pics/nordea2_y.pdf}
-  \caption{\small \it The residuals' distribution and auto-correlations. Left:
-    cummulative distribution function (CDF) of the residuals $y_t$; Right:
-    auto-correlation function of $y_t$.}
-  \label{fig:nordea2_y}
-\end{figure}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    Moment & mean & variance & skewness & kurtosis \\
-    \hline
-    value & -0.0022 & 0.1007 & -1.2669 & 15.5965 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it: Moments of the residuals $y_t$}
-  \label{tab:nordea2_y_mmt}
-\end{table}
-
-The auto-correlations of the residuals $y_t$ are insignificant, as
-shown on the right of figure \ref{fig:nordea2_y}. Though the Ljung-Box Q
-test rejects the null hypothesis of no-autocorrelation in the
-residuals, the weak auto-correlations, if any at all,  are not worth
-further complicating the model.
-
-Forecast of $\ln\sigma_t$ is done by reverting equation
-\ref{eq:nordea2_v} and write
-\begin{eqnarray*}
-  \ln\sigma_t &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
-  \ln\sigma_{t-s-1} + (1 - \theta_1 B - \theta_s B^s + \theta_{s+1}
-  B^{s+1} ) y_t \\
-  \E(\ln\sigma_t | \mathcal{F}_{t-1}) &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
-  \ln\sigma_{t-s-1} - \theta_1 y_{t-1} - \theta_s y_{t-s} +
-  \theta_{s+1} y_{t-s-1}
-\end{eqnarray*}
-
-Thus the forecast, namely $\E(\ln\sigma_t | \mathcal{F}_{t-1})$, is
-found in terms of previous values in the same series and of previous
-residuals.
-
-To compare the performance, we also estimate a GARCH model with
-Gaussian conditional distribution for the Nordea Bank 15-minute
-return. The procedure is standard and we only list the results here
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_5$ & $\beta_1$ \\
-    \hline
-    Value & $7.3686 \times 10^{-7}$ & 0.154131 & 0.0360 & 0.491832 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Parameters of GARCH(1, 5) model of Nordea
-    Bank 15-minute returns}
-  \label{tab:nordea2_garch}
-\end{table}
-
-Forecasting using both models, we get two series of one-step-ahead
-forecast, $\ln \sigma^{\text{ARIMA}}_t$ and $\ln \sigma^G_t$,
-respectively. We denote the empirical log-volatilities as $\ln
-\hat{\sigma}_t$. Then the accuracy of the two models can be compared
-by comparing $\ln \sigma^{\text{ARIMA}}_t - \ln \hat{\sigma}_t$ and
-$\ln \sigma^G_t - \ln \hat{\sigma}_t$. As a reference, we also include
-in the comparison the naive approach of taking the sample mean of
-$\ln \sigma_t$ as the forecast.
-
-Table \ref{tab:slv_garch_comparison} shows the standard deviation of
-$\ln \sigma^F_t - \ln \hat{\sigma}_t$ for the 3 kinds of
-forecast. Here we use $\ln \sigma^F_t$ to denote the forecast of $\ln
-\sigma_t$. Clear the ARIMA model has the smallest standard deviation,
-the sample mean has the largest, and the GARCH model sits in between.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    & ARIMA & GARCH & Sample mean \\
-    \hline
-    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ &
-    0.2876  &  0.3096 &   0.3349 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
-    \ln\hat{\sigma}_t$}
-  \label{tab:slv_garch_comparison}
-\end{table}
-Even more informative about the quality of forecast is the
-distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$, which we show
-in figure \ref{fig:slv_garch_comparison}.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=0 292 0
-  255]{../pics/slv_garch_comparison.pdf}
-  \caption{\small \it Comparison of the forecast from ARIMA, GARCH,
-    and sample mean. Blue: ARIMA forecast; Green: GARCH forecast; Red:
-    Sample mean forecast.}
-  \label{fig:slv_garch_comparison}
-\end{figure}
-The left plot in the figure shows the situation of under-estimates,
-i.e. $\ln \sigma^F_t - \ln \hat{\sigma}_t < 0$. We see that, to the
-far left, the blue line is below the green, which is in turn below the
-red, meaning the ARIMA forecast has the lowest probability of
-under-estimate. The GARCH forecast comes second and the sample
-mean forecast under-estimates most.
-
-The same ranking of forecast quality is also seen on the right plot of
-figure \ref{fig:slv_garch_comparison}, which shows the complementary
-distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$ for postive
-values, i.e. over-estimates: The ARIMA forecast has the lowest
-probability of over-estimate; GARCH comes second and the sample mean
-is most likely to over-estimate.
-
-As another measure of the quality of forecast, the fraction of
-``good'' forecasts that differ no more than a given percentage from
-the measured value is worth a few words. Table
-\ref{tab:nordea_good_percentage} compares this fraction of good
-forecasts from the 3 alternatives with respect to different
-requirements of goodness.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ &
-    ARIMA & GARCH & Sample Mean \\
-    \hline
-    1\% & 0.1920 & 0.2010 & 0.1843 \\
-    \hline
-    5\% & 0.7887 & 0.7436 & 0.7229 \\
-    \hline
-    10\% & 0.9665 & 0.9601 & 0.9369 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Fraction of ``good'' forecasts as defined by
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
-  \label{tab:nordea_good_percentage}
-\end{table}
-Clearly, although ARIMA and GARCH rank differently with respect to
-different requirements of goodness, they both out-perform the sample
-mean forecast. So the efforts of building models and estimating them
-have paid off.
-\subsection{Volvo B Log-Volatility}\label{sec:volvo}
+\section{Case Study: Volvo in 2013/14}
+\label{sec:volvo}
 In this section we model the log-volatility of {\it Volvo B} 30-minute
@@
 auto-correlations of $\ln\sigma_t$. Fitting this model to the
-actual log-volatilities yields the parameter values listed in table
+measured realized volatilities yields the parameter values listed in
+table
 \ref{tab:volvo_params}.
@@
 Forecasting using the estimated model parameters gives the forecast
-series $\ln \sigma^{\text{ARIMA}}_t$. To access the quality of the
+series $\ln \sigma^{\text{SV}}_t$. To access the quality of the
 forecast, we also estimate a GARCH model using the returns. The result
-is a GARCH(1, 1) model:
+is a GARCH(1, 1) model, whose parameter values are listed in table
+\ref{tab:volvo_garch}.
 \begin{table}[htb!]
@@
 
-The forecasts from ARIMA, GARCH, and the sample mean are compared
+The forecasts from SV, GARCH, and the sample mean are compared
 using the difference $\ln \sigma^F_t - \ln\hat{\sigma}_t$, where $\ln
 \sigma^F_t$ stands for the forecast. The distributions of this
-difference is plotted in figure \ref{fig:volvo_slv_garch_cmp}, and
-the standard deviation of the distributions are listed in table
-\ref{tab:volvo_slv_garch_cmp}.
+difference is plotted in figure \ref{fig:volvo_slv_garch_cmp}; the
+mean and the standard deviation of the distributions are listed in
+table \ref{tab:volvo_slv_garch_cmp}.
 \begin{table}[htb!]
@@
     \hline
-    & ARIMA & GARCH & Sample mean \\
+    & SV & GARCH & Sample mean \\
+    \hline
+    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & -0.0123 &
+    0.0242 & -0.1055 \\
     \hline
@@
     \label{fig:volvo_slv_garch_cmp}
-  \caption{\small \it Comparison of the forecast from ARIMA, GARCH,
-    and sample mean. Blue: ARIMA forecast; Green: GARCH forecast; Red:
+  \caption{\small \it Comparison of the forecast from SV, GARCH,
+    and sample mean. Blue: SV forecast; Green: GARCH forecast; Red:
     Sample mean forecast.}
@@
 Figure \ref{fig:volvo_slv_garch_cmp} shows, as in the previous case of
-Nordea Bank 15-minute returns, the ARIMA model performs the best, GARCH
+Nordea Bank 15-minute returns, the SV model performs the best, GARCH
 the second, and the sample mean the worst. However, when it comes to
 over-estimates, the sample mean appears to be the best estimator,
-while ARIMA excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
+while SV excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
 the sample for estimation (1507 data points) and over the sample for
@@
       \hat{\sigma}_t|}$ &
-    ARIMA & GARCH & Sample Mean \\
+    SV & GARCH & Sample Mean \\
     \hline
@@
 \end{table}
-We see from the table that the ARIMA model consistently excels over
-the other two alternatives. Once again, it confirms the validity and
-good performance of an ARIMA log-volatility model.
-
-% \subsection{Ericsson B Log-Volatility}\label{sec:ericsson}
-% The probability density function of the residuals $y_t$ that result
-% from the MLE estimate is shown in figure
-% \ref{fig:nordea_15min_y1_hist} together with its counterpart from the
-% initial estimate.
-% \subsection{45min Returns}
-\section{Unconditional Distribution Functions of ARIMA Log-Volatility
-  Models}\label{sec:XieCalc}
+We see from the table that the SV model consistently excels over
+the other two alternatives.
+
+\section{Unconditional Distribution Functions of SV Models}
+\label{sec:XieCalc}
 In this section we study the unconditional distribution function of
@@
 correlation translates to the correlation between the residual of the
-log-volatility, denoted $y_t$ in \S\ref{sec:nordea} and
+log-volatility, denoted $y_t$ in \S\ref{chp:nordea_15min} and
 \S\ref{sec:volvo}, and $b_t$. Now let us consider the forecast
@@
 
-For the Nordea series considered in \S\ref{sec:nordea},
+For the Nordea series considered in \S\ref{chp:nordea_15min},
 $\text{corr}(y_t, b_t)$ is found to be $3.56 \times 10^{-2}$; while

### 2014-05-08 thesis/thesis.tex
Change the section about Nordea 15min in 1st quarter 2012.

@@
 \section{Collection of Figures and Tables}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_acf.pdf}
+    \label{fig:nordea_15min_acf}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_vlt_acf.pdf}
+    \label{fig:nordea_15min_vlt_acf}
+  }
+  \caption{\small \it Auto-correlations among Nordea Bank 15min
+    returns. \ref{fig:nordea_15min_acf}: Auto-correlations (ACF) of the
+    squared returns; \ref{fig:nordea_15min_vlt_acf}: Auto-correlations
+    (ACF) of squared realized volatilities ($\hat{\sigma}^2_t$).}
+\end{figure}
+
 \begin{figure}[htb!]
@@
 
-\begin{figure}[htb!]
-  \centering
-  \subfigure[Normal Probability Plot of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=76 256 109
-    214]{../pics/nordea2_y_normplot.pdf}
-    \label{fig:nordea_15min_y_qq}
-  }
-  \subfigure[ACF of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=92 258 104
-    218]{../pics/nordea_15min_y_acf.pdf}
-    \label{fig:nordea_15min_y_acf}
-  }
-  \caption{\small \it Nordea 15min $y_t$ normal probability plot and
-    autocorrelations}
-\end{figure}
+% \begin{figure}[htb!]
+%   \centering
+%   \subfigure[Normal Probability Plot of $y_t$]{
+%     \includegraphics[scale=0.4, clip=true, trim=76 256 109
+%     214]{../pics/nordea2_y_normplot.pdf}
+%     \label{fig:nordea_15min_y_qq}
+%   }
+%   \subfigure[ACF of $y_t$]{
+%     \includegraphics[scale=0.4, clip=true, trim=92 258 104
+%     218]{../pics/nordea_15min_y_acf.pdf}
+%     \label{fig:nordea_15min_y_acf}
+%   }
+%   \caption{\small \it Nordea 15min $y_t$ normal probability plot and
+%     autocorrelations}
+% \end{figure}
 

### 2014-05-08 thesis/price_models.tex
Add the additions of stock case studies.

@@
 \label{sec:volvo}
-In this section we model the log-volatility of {\it Volvo B} 30-minute
-returns during the period 2013/10/10 - 2014/04/04. This series
-contains 1884 log-volatilities computed using 2-minute returns in each
-30-minute interval. Among them we use the first 1507 for model
-estimation and the last 377 for forecast and model verification.
-
-The left plot of figure \ref{fig:volvo_inno_and_lv_acf} shows the
-distribution of $(r_t - \E(r_t))/\sigma_t$. We observe in the figure a
-nice Gaussian variate, so we can be sure that the sum of squared
-2-minute returns makes a good approximation to the variance of
-30-minute returns in this particular case.
+In this section we study the volatility of Volvo B 15-minute returns
+during the period 2013/10/10 - 2014/04/04.
+The subject of modeling is the realized volatility
+computed as the square root of the sum of squared 1-minute
+returns. The normal probability plot of the quotient $(r_t -
+\E(r_t))/\sigma_t$ is shown in figure \ref{fig:volvo_15_quotient},
+from which one can see it is normally distributed, confirming the
+choice of 1-minute returns for computing the realized volatility.
 \begin{figure}[htb!]
   \centering
-    \includegraphics[scale=0.5, clip=true, trim=0 256 0
-    151]{../pics/volvo_inno_and_lv_acf.pdf}
-    \label{fig:volvo_inno_and_lv_acf}
-  \caption{\small \it Left: probability plot of $(r_t -
-    \E(r_t))/\sigma_t$; Right: auto-correlations of $\ln \sigma_t$}
+  \includegraphics[scale=0.4, clip=true, trim=83 259 110 
+  220]{../pics/volvo_15_quotient.pdf}
+  \caption{\small \it Normal probability plot of $(r_t -
+    \E(r_t))/\sigma_t$}
+  \label{fig:volvo_15_quotient}
 \end{figure}
-
-Guided by the auto-correlations of $\ln\sigma_t$ shown in the right
-plot of figure \ref{fig:volvo_inno_and_lv_acf} we find the following
-model:
-\begin{eqnarray}
-  && (1 - B)(1 - B^s) \ln \sigma_t \nonumber \\
-  &=& (1 - \theta_1 B - \theta_2B^2 -
-  \theta_3B^3 - \theta_4B^4) (1 - \Theta B^s)
-  y_t \label{eq:volvo_lv_model}
-\end{eqnarray}
-where $s = 16$ is the seasonality and is apparent from the
-auto-correlations of $\ln\sigma_t$. Fitting this model to the
-measured realized volatilities yields the parameter values listed in
-table
-\ref{tab:volvo_params}.
+The auto-correlations of $\ln \sigma_t$ and $w_t = (1-B)(1-B^s) \ln
+\sigma_t$ are plotted in figure \ref{fig:volvo_15_lv_acf} and
+\ref{fig:volvo_15_w_acf}. The former clearly shows the seasonality
+$s = 33$ in the auto-correlations of $\ln \sigma_t$.
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+  \includegraphics[scale=0.4, clip=true, trim=90 259 103
+  220]{../pics/volvo_15_lv_acf.pdf}
+  \label{fig:volvo_15_lv_acf}
+  }
+  \subfigure[]{
+  \includegraphics[scale=0.4, clip=true, trim=90 259 103
+  220]{../pics/volvo_15_w_acf.pdf}
+  \label{fig:volvo_15_w_acf}
+  }
+  \caption{\small \it Left: auto-correlations of $\ln \sigma_t$;
+    Right: auto-correlations of $w_t = (1-B)(1-B^s) \ln \sigma_t$.}
+\end{figure}
+
+Based on the auto-correlations of $w_t$, a seasonal moving average
+model is estimated:
+\begin{eqnarray*}
+  (1-B)(1-B^s) \ln\sigma_t &=& (1-\theta_1B - \theta_2B^2 - \theta_3
+  B^3)(1 - \Theta B^s)y_t
+\end{eqnarray*}
+By maximum likelihood estimation, the parameter values listed in table
+\ref{tab:volvo_15_sv_param} are obtained.
 \begin{table}[htb!]
   \centering
+  \begin{tabular}{|c|c|c|c|c|c|}
+  \hline
+  Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\Theta$ & 
+ $\var(y_t)$ \\
+ \hline
+ Value & 0.7491 & 0.1170 & 0.0571 & 0.8646 & 0.1245 \\
+  \hline
+  \end{tabular}
+  \caption{\small \it Parameter values of the stochastoc volatility
+    (SV) model.}
+  \label{tab:volvo_15_sv_param}
+\end{table}
+
+A GARCH model (c.f. equation \ref{eq:garch_def}) assuming Gaussian
+innovations is also fitted to the same series. The parameter values
+are listed in table \ref{tab:volvo_15_garch_param}:
+\begin{table}[htb!]
+  \centering
   \begin{tabular}{|c|c|c|c|c|c|c|}
   \hline
-    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ &
-    $\Theta$ & residual variance \\
-    \hline
-    Value & 0.7305 & 0.0575 & 0.0574 & 0.0346 & 0.8324 & 0.1340\\
+  Parameter & $\alpha_0$ & $\alpha_1$ & $\beta_{1}$ \\
+  \hline
+  Value & $6.54 \times 10^{-7}$ & 0.1565 & 0.6445\\
   \hline
   \end{tabular}
-  \caption{\small \it Volvo B log-volatility parameters}
-  \label{tab:volvo_params}
+  \caption{\small \it Parameter values of the GARCH model.}
+  \label{tab:volvo_15_garch_param}
 \end{table}
-Forecasting using the estimated model parameters gives the forecast
-series $\ln \sigma^{\text{SV}}_t$. To access the quality of the
-forecast, we also estimate a GARCH model using the returns. The result
-is a GARCH(1, 1) model, whose parameter values are listed in table
-\ref{tab:volvo_garch}.
+
+Performing one-step-ahead forecasts using both models gives the series
+$\ln\sigma^F_t - \ln\hat{\sigma}_t$, where $\ln\hat{\sigma}_t$ are the
+measured realized volatilities. The mean and standard deviation of
+$\ln\sigma^F_t - \ln\hat{\sigma}_t$ are shown in table
+\ref{tab:volvo_15_stat}.
 \begin{table}[htb!]
@@
     \hline
-    Parameter & $\alpha_0$ & $\alpha_1$  & $\beta_1$ \\
-    \hline
-    Value & $3.125 \times 10^{-7}$ & 0.05 & 0.90 \\
+     & SV & GARCH & Sample mean \\
+     \hline
+    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0035 & 0.0313 & -0.1448 \\
+    \hline
+     $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3356 &
+     0.3894 & 0.3826 \\
     \hline
  \end{tabular}
-  \caption{\small \it GARCH(1, 1) model of Volvo B 30-minute returns}
-  \label{tab:volvo_garch}
+  \caption{\small \it Mean and standard deviation of the 3 kinds of
+    forecasts}
+  \label{tab:volvo_15_stat}
 \end{table}
-
-The forecasts from SV, GARCH, and the sample mean are compared
-using the difference $\ln \sigma^F_t - \ln\hat{\sigma}_t$, where $\ln
-\sigma^F_t$ stands for the forecast. The distributions of this
-difference is plotted in figure \ref{fig:volvo_slv_garch_cmp}; the
-mean and the standard deviation of the distributions are listed in
-table \ref{tab:volvo_slv_garch_cmp}.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    & SV & GARCH & Sample mean \\
-    \hline
-    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & -0.0123 &
-    0.0242 & -0.1055 \\
-    \hline
-    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3261 &
-    0.4250 & 0.3708 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
-    \ln\hat{\sigma}_t$}
-  \label{tab:volvo_slv_garch_cmp}
-\end{table}
-
+We see that both the SV and the GARCH model give biased forecasts, but
+the SV forecasts are only 1/10 as biased as those of GARCH (0.0035
+vs. 0.0313). In addition, the standard deviation of the SV forecasts
+is smaller too. These results are confirmed by the distribution of
+$\ln \sigma^F_t - \ln \hat{\sigma}_t$ and the fraction of ``good''
+forecasts, which are shown in figure \ref{fig:volvo_15_diff} and table
+\ref{tab:volvo_15_diff}, respectively.
 \begin{figure}[htb!]
   \centering
-    \includegraphics[scale=0.5, clip=true, trim=10 282 0
-    243]{../pics/volvo_slv_garch_cmp.pdf}
-    \label{fig:volvo_slv_garch_cmp}
-  \caption{\small \it Comparison of the forecast from SV, GARCH,
-    and sample mean. Blue: SV forecast; Green: GARCH forecast; Red:
-    Sample mean forecast.}
+    \includegraphics[scale=0.54, clip=true, trim=27 276 0
+    236]{../pics/volvo_15_diff.pdf}
+  \caption{\small \it Comparison of the forecasts from SV, GARCH,
+    and sample mean. Blue: SV forecasts; Green: GARCH forecasts; Red:
+    Sample mean forecasts.}
+  \label{fig:volvo_15_diff}
 \end{figure}
-Figure \ref{fig:volvo_slv_garch_cmp} shows, as in the previous case of
-Nordea Bank 15-minute returns, the SV model performs the best, GARCH
-the second, and the sample mean the worst. However, when it comes to
-over-estimates, the sample mean appears to be the best estimator,
-while SV excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
-the sample for estimation (1507 data points) and over the sample for
-comparison (377 data points) reveals that the first sample has mean
--6.3127 while the second has -6.2072. This increment explains the low
-probability of over-estimation when using the first sample mean as
-forecast.
-
-Table \ref{tab:volvo_good_percentage} compares the fraction of
-``good'' estimates as measured by ${|\ln\sigma^F_t -
-  \ln\hat{\sigma}_t| \over |\ln \hat{\sigma}_t|}$ being less than 1\%,
-5\% and 10\%.
 \begin{table}[htb!]
@@
     \hline
-    1\% & 0.2175 & 0.1220 & 0.1406 \\
-    \hline
-    5\% & 0.7162 & 0.6154 & 0.6631 \\
-    \hline
-    10\% & 0.9231 & 0.8806 & 0.8992 \\
+    1\% & 0.5760 & 0.5979 & 0.3930 \\
+    \hline
+    5\% & 0.8351 & 0.8363 & 0.6611 \\
+    \hline
+    10\% & 0.9716 & 0.9536 & 0.9085 \\
     \hline
   \end{tabular}
-  \caption{\small \it Fraction of ``good'' forecasts as defined by
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
-  \label{tab:volvo_good_percentage}
+  \caption{\small \it Percentage of ``good'' forecasts as defined by
+    deviating nore more than 1\%, 5\% and 10\% from the measured
+    realized volatility.}
+  \label{tab:volvo_15_diff}
 \end{table}
-We see from the table that the SV model consistently excels over
-the other two alternatives.
 

### 2014-05-08 thesis/thesis.tex
Add the additions of stock case studies.

--- 2014-05-06 thesis/thesis.tex
@@
 
-\chapter{Summary of the Results}
-The results of this thesis project are the following:
-\begin{enumerate}
-\item ARIMA log-volatility model
-  \begin{enumerate}
-  \item ARIMA log-volatility returns' models are justified and tested
-    against Nordea Bank 15-minute returns and Volvo B 30-minute returns
-    during the period 2013/10/10 - 2014/04/04. Historical volatility's
-    are estimated using ``realized volatilities'', denoted
-    $\hat{\sigma}_t$ here:
-    \begin{eqnarray*}
-      \hat{\sigma}_t &=& \sqrt{
-        \sum_{i = 1}^n \left[\ln {p_{t - \Delta t + i \Delta t / n} \over
-            p_{t - \Delta t + (i-1) \Delta t / n}} \right]^2
-      }
-    \end{eqnarray*}
-    where $p_{t'}$ is the price of the asset at time $t'$, and $n$ is
-    the number of sub intervals into which the time interval $[t -
-    \Delta t, t]$ is divided. A discussion of the choice of $n$ is found
-    in \S\ref{sec:SLV_model}.
-
-    Using the first 80\% of the data for estimating the model and the
-    last 20\% for comparing with the model forecast, we find the ARIMA
-    models out-perform their GARCH competitors in most cases.
-
-  \item The unconditional distribution functions of ARIMA
-    log-volatility models are calculated (equation
-    \ref{eq:UncondPDFAsymmetric} and \ref{eq:UncondPDFAsymmetric1}):
-    The unconditional distribution of  the returns $r_t$ is the same as
-    $\mu + e^{\bar{v} + \sigma a} b$, where $(a, b) \sim N(0, \Sigma)$
-    \begin{eqnarray*}
-      \Sigma &=&
-      \begin{pmatrix}
-        1 & \psi \\
-        \psi & 1
-      \end{pmatrix}
-    \end{eqnarray*}
-    Then the PDF of the unconditional distribution $f_r(x)$ can be expressed as
-    \begin{eqnarray*}
-      f_{r'}(x) &=& {1 \over 2\pi \sqrt{1 - \psi^2}} \times \\
-      && \int_{-\infty}^\infty da \exp
-      \left[- {
-          a^2 + 2\sigma(1 - \psi^2)a - 2a\psi e^{-a\sigma} x + e^{-2a\sigma} x^2
-          \over
-          2(1 - \psi^2)
-        }
-      \right] \\
-      f_r(x) &=&  \td{r'}{r} f_{r'}[e^{-\bar{v}} (x - \mu)] \\
-      &=& e^{-\bar{v}} f_{r'}[e^{-\bar{v}} (x-\mu)] \\
-    \end{eqnarray*}
-    where $r' = e^{\sigma a}b$. The moment generating function of the
-    unconditional distribution is found as an integral, and all the
-    moments can be evaluated in closed form. See equations
-    \ref{eq:AsymmetricMoments}.
-
-  \end{enumerate}
-\item Correlation matrix of Gaussian returns. $N$ series of
-  auto-correlated, standard Gaussian distributed returns are simulated
-  to form $N\times N$ correlation matrix $C$:
-  \begin{eqnarray*}
-    r_{it} &=& \phi r_{i,t-1} + a_t \\
-    C_{ij} &=& {1 \over T} \sum_{t=1}^T r_{it} r_{jt}
-  \end{eqnarray*}
-  where $i = 1, 2, ..., N$ and $t = 1,2, ..., T$. $a_t \sim N(0,
-  1)$.
-  \begin{enumerate}
-  \item Distribution of the elements of C. The asymptotic distribution
-    of the matrix elements $C_{ii}$ and $C_{ij}$ ($i \neq j$) are 
-    Gaussian whose mean and variance are functions of $\phi$:
-    \begin{eqnarray*}
-      \E(C_{ij}) &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi^2)(1 -
-        \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
-        P^{-3/2}_{-1/2}(\rho)
-      \right] \\
-      \var(C_{ij}) &=& {2 \sigma^6 \over T (1 - \phi^2)^2} \left[
-        {\phi^2 \over 1 - \phi^2} -
-        {\phi^2 (1 - \phi^{2T}) \over
-          T(1 - \phi^2)}
-      \right] + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
-        T (1 - \phi^2)^2} \\
-      &\approx& {2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}
-      + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
-        T (1 - \phi^2)^2}
-    \end{eqnarray*}
-    where $\rho = \text{corr}(a_{it}, a_{jt})$, $i \neq
-    j$. $P^\mu_\nu(\cdot)$ is Ferrers function of the first kind (See
-    equation \ref{eq:Ferrers_1st}) and
-    \begin{eqnarray*}
-      v^2(\rho) &=&
-      {4 \over \sqrt{2\pi} (1 - \rho^2)^{7/4}} \left[
-        P^{-5/2}_{-1/2}(\rho) + P^{-5/2}_{-1/2}(-\rho)
-      \right] \\
-      && - {1 \over 2\pi (1 - \rho^2)^{5/2}} \left[
-        P^{-3/2}_{-1/2}(-\rho) - P^{-3/2}_{-1/2}(\rho)
-      \right]^2
-    \end{eqnarray*}
-    In comparison,
-    \begin{eqnarray*}
-      \E(C_{ii}) & \approx & {\sigma^2 \over 1 - \phi_1^2} \left[
-        1 - {\phi_1^2 \over T}
-      \right] \\
-      \var(C_{ii}) &\approx& {2 \sigma^4 \over T (1 - \phi_1^4)} +
-      {\sigma^6 \over T (1 -\phi_1^2)^2}
-    \end{eqnarray*}
-
-  \item Distribution of the maximum eigenvalue of C. It is shown that
-    the maximum eigenvalue has approximately a gamma distribution even
-    when auto-correlations are present among the returns. The mean of
-    the gamma distribution is approximately quadratic in $\tan{\pi\phi
-      \over 2}$.
-  \end{enumerate}
-
-\item Covariance matrix C of GARCH(1,1) processes. In the limit of
-  very long return series, the covariance matrix of properly
-  specified GARCH(1,1) processes (see chapter
-  \ref{chp:CrossCorrelationFat}) is 
-  shown to be diagonal. The distribution of its eigenvalues thus
-  coincides with that of its diagonal elements. Their distribution
-  converges to an $\alpha$-stable distribution. If the returns $r_{it}$
-  have tail behavior $P(|r_{it}| > x) \sim {c \over x^\alpha}$, the
-  asymptotic stable distribution has L\'evy index
-  $\alpha/2$. Moreover, the mean of this stable distribution scales
-  with the length of the return series (T) as $T^{1 - 2/\alpha}$,
-  while the width is independent of T.
-
-  In contrast, the distribution of the non-diagonal elements of the
-  covariance matrix converges to a symmetric, zero-centered, stable
-  distribution with a different L\'evy index than $\alpha$. Denote it
-  $\alpha'$, then the variance of this stable distribution scales with
-  T as $T^{1/\alpha' - 2/\alpha}$.
-
-  When the return series only have a modest length, the situation is
-  studied in terms of the localization of the eigenvectors. For a
-  given T, the largest eigenvalues correspond to the most localized
-  eigenvectors, and the extent of localization decreases as the
-  eigenvalue decreases. As T increases, the extent of localization
-  increases for all eigenvectors.
-
-  Auto-correlations among the returns, on the other hand, reduces the
-  extent of localization, and also the certainty of localization with
-  respect to the eigenvalues. In other words, if the return series
-  have considerable auto-correlations, the eigenvectors corresponding
-  to the largest eigenvalues will have a rather fractured composition.
-\end{enumerate}
-
-\chapter{Appendix}
-\section{PDF of the covariance matrix of auto-correlated Gaussian
-  Returns} \label{app:pdf_gaussian1}
-When autocorrelations exist among the columns of the Gaussian returns
-matrix $\mtx R$, the distribution of the covariance matrix $\mtx C =
-\mtx{R} \mtx{R'}$ is no longer Wishart but is nonethelss closely
-related to it. In the following we consider the situation where
-$r_{i,t}$ can be represented as a vector auto-regressive
-process. Specifically, suppose
-\begin{eqnarray*}
-  \vec{r}_t &=& \sum_{k=1}^p \phi_k \vec{r}_{t-k} + \vec{a}_t \\
-  \vec{a}_t &=& \vec{r}_t - \sum_{k=1}^p \phi_k \vec{r}_{t-k}
-\end{eqnarray*}
-where $\vec{a}_t = (a_{1,t}, a_{2,t}, \cdots, a_{N,t})' \sim N(0,
-\Sigma)$ and comprise the columes of $\mtx A$; $\vec{r}_t$ are the
-columes of $\mtx R$. Here $N(0, \Sigma)$ denotes the multivarate
-normal distribution with covariance matrix $\Sigma$. The last equation
-can be written in matrix form
-\[
-A = R M
-\]
-For example, in the case of an AR(1) process
-\begin{equation*}
-  \begin{pmatrix}
-    a_{1,1} & a_{1,2} & \cdots & a_{1,T} \\
-    \vdots & \ddots & \vdots \\
-    a_{N,1} & a_{N,2} & \cdots & a_{N,T} \\
-  \end{pmatrix} =
-  \begin{pmatrix}
-    r_{1,1} & r_{1,2} & \cdots & r_{1,T} \\
-    \vdots & \ddots & \vdots \\
-    r_{N,1} & r_{N,2} & \cdots & r_{N,T} \\
-  \end{pmatrix}
-  \begin{pmatrix}
-    1 & -\phi_1 &   &   & \\
-      & 1 & -\phi_1 &   & \\
-      &   & \ddots  & \ddots &   \\
-      &   &   & 1 & -\phi_1 \\
-      &   &   &   & 1 \\
-  \end{pmatrix}
-\end{equation*}
-Let $QR = RM$, then $Q = RMR^{-1}$. Since the set of $\{r_{ij}\}$ for
-which $\text{det } R = 0$ has probability zero, a matrix $Q$
-satisfying the above equation almost surely exists. Thus $A = RM =
-QR$ and $AA' = QRR'Q'$ Since the columns of A are not auto-correlated,
-$AA' \sim W(\Sigma, T)$. Then $RR' \sim W(Q^{-1} \Sigma Q'^{-1}, T)$
-follows from \cite{Anderson2003}, \S7.3.3.
-
-Now we observe that $\Sigma$ enters $f(RR')$ through $\tr (\Sigma^{-1}
-RR')$ and $\det \Sigma$. $RR'$ enters through $\tr (\Sigma^{-1}
-RR')$ and $\det (RR')$. Clearly
-\begin{eqnarray*}
-  \det (Q^{-1}\Sigma Q'^{-1}) &=& \det \Sigma \over (\det Q)^2 \\
-  &=& \det \Sigma \over (\det M)^2 \\
-  &=& \det \Sigma
-\end{eqnarray*}
-As to $\tr (\Sigma^{-1} C)$, we have
-\begin{eqnarray*}
-  && \tr\left[(Q^{-1}\Sigma Q'^{-1})^{-1} RR'\right] \\
-  &=& \tr\left[Q'\Sigma^{-1} Q RR'\right] \\
-  &=& \tr\left[ R'^{-1}M'R'\Sigma^{-1} RMR^{-1} RR'\right] \\
-  &=& \tr\left[ \Sigma^{-1} RM (RM)'\right] \\
-  &=& \tr\left[ \Sigma^{-1} AA'\right] \\
-\end{eqnarray*}
-
-Moreover, $\det RR' = \det AM^{-1} M'^{-1} A' = \det AA'$.
-
-Thus if we use $f_R(\cdot)$ to denote the joint probability density of
-the entries of RR' and $f_A(\cdot)$ to denote that of AA', we can
-write
-\begin{equation}\label{eq:cross-corr-matrix-PDF}
-  f_R(RR') = f_A(RMM'R')
-\end{equation}
-
-Substituting for $f_A$ the Wishart probability density function (PDF)
-\ref{eq:WishartPDF}, we get the PDF of $RR'$.
-
-\section{Eigenvalue distribution of Wishart Matrix}
-\label{sec:wishart_eigen_dist}
-In this section we summarize the analytic results regarding the
-eigenvalue distribution of a Wishart matrix $\mtx{RR'}$. In the
-simplest case where the elements of $\mtx R$ are all independent of
-each other, i.e. neither auto-correlation nor cross-correlation
-exists, we have \cite{Chiani2012}
-\begin{eqnarray}
-  f_{\lambda}(x_1, \cdots, x_N) &=& K \prod_{i=1}^N e^{-x_i/2}
-  x_i^{(T-N-1)/2} \prod_{i<j}^N (x_i -
-  x_j) \label{eq:wishart_eigen_pdf}\\
-\end{eqnarray}
-where the eigenvalues have been indexed in descending order: $x_1 \geq
-x_2 \geq \cdots \geq x_N$. The normalization constant K is given by
-\begin{equation*}
-  K = {\pi^{N^2/2} \over 2^{NT/2}\Gamma_N(T/2) \Gamma_N(N/2)}
-\end{equation*}
-where the function $\Gamma_m(a)$ is defined as
-\begin{equation*}
-  \Gamma_m(a) = \pi^{m(m-1)/4} \prod_{k=1}^m \Gamma\left(a - {k-1
-      \over 2}\right)
-\end{equation*}
-Moreover, K. Johnsson \cite{Johnsson2000} and I. Johnstone
-\cite{Johnstone2001} showed that, at the {\it absence} of
-autocorrelations and in the asymptotic limit $N, T \to \infty$, $N/T
-\to q < \infty$, the maximum eigenvalue  $\lambda_1$ follows the
-Tracy-Widom distribution (denote $\mathscr{TW}_1$ here) when properly
-relocated and rescaled:
-\begin{equation*}
-  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} \sim \mathscr{TW}_1
-\end{equation*}
-where $\mu_{NT}$ and $\sigma_{NT}$ are given by
-\begin{eqnarray*}
-  \mu_{NT} &=& \left(
-    \sqrt{N-1/2} + \sqrt{T - 1/2}
-  \right)^2 \\
-  \sigma_{NT} &=& \sqrt{\mu_{NT}} \left(
-    {1 \over \sqrt{N-1/2}} + {1 \over \sqrt{T-1/2}}
-  \right)^{1/3}
-\end{eqnarray*}
-The $\mathscr{TW}_1$ distribution has the following cummulative
-distribution function (CDF) \cite{Chiani2012}
-\begin{equation*}
-  F_1(x) = \exp\left[
-    -{1 \over 2} \int_x^\infty dy \left(q(y) + (y-x)q^2(y)\right)
-  \right]
-\end{equation*}
-where $q(y)$ is defined as the solution to the Painlev\'e II differential
-equation
-\begin{equation*}
-  q''(y) = yq(y) + 2q^3(y)
-\end{equation*}
-which is unique when imposing the condition
-\begin{equation*}
-  q(y) \sim \text{Ai}(y) \text{ as } y \to \infty
-\end{equation*}
-
-Then Marco Chiani showed recently that the $\mathscr{TW}_1$ distribution
-can be well approximated by a gamma distribution based on his proof
-that the exact distribution of the maximum eigenvalue is a mixture of
-gamma distributions. Specifically,
-\begin{equation}\label{eq:TracyWidom-Gamma}
-  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} + \alpha \sim
-  \mathscr{G}(k ,\theta)
-\end{equation}
-where $\mathscr{G}(k, \theta)$ denotes the Gamma distribution with
-parameters $k$ and $\theta$ \cite{Chiani2012}. The probability
-density function (PDF) of the gamma distribution is given by
-\begin{equation*}
-  f_\gamma(x; k, \theta) = {1 \over \Gamma(k) \theta^k} x^{k-1}
-  e^{-x/\theta}
-\end{equation*}
-Moreover, the first 3 moments of the distribution are simple:
-\begin{eqnarray*}
-  \text{mean} &=& k\theta \\
-  \text{variance} &=& k\theta^2 \\
-  \text{skewness} &=& 2/\sqrt{k}
-\end{eqnarray*}
-
-\section{Collection of Figures and Tables}
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of $z_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_acf.pdf}
-    \label{fig:nordea_15min_quotient_acf}
-  }
-  \subfigure[ACF of $z_t^2$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
-    \label{fig:nordea_15min_quotient_squared_acf}
-  }
-  \caption{\small \it Nordea 15min $z_t$ and $z_t^2$ auto-correlation
-    function (ACF).}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \subfigure[Normal Probability Plot of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=76 256 109
-    214]{../pics/nordea2_y_normplot.pdf}
-    \label{fig:nordea_15min_y_qq}
-  }
-  \subfigure[ACF of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=92 258 104
-    218]{../pics/nordea_15min_y_acf.pdf}
-    \label{fig:nordea_15min_y_acf}
-  }
-  \caption{\small \it Nordea 15min $y_t$ normal probability plot and
-    autocorrelations}
-\end{figure}
-
-\begin{figure}[htb!]
-  \begin{center}
-    \includegraphics[scale=0.4, clip=true, trim=90 228 115
-    226]{../pics/GaussianMarkovSpectrumPDF.pdf}
-  \end{center}
-  \vspace{-10mm}
-  \caption{\small \it
-      Eigenvalue distribution with $\tau$ ranging from 0 to
-      3. The 1st blue line, which is shown as stairs, is the
-      theoretical eigenvalue distribution according to the
-      Marcenko-Pastur law: $f(\lambda) = \sqrt{(\lambda_+/\lambda - 1)(1
-        - \lambda/\lambda_-)}/2\pi q \sigma^2$, where $\lambda_{\pm} =
-      \sigma^2(1 \pm \sqrt{q})^2$ In the simulation we have
-      chosen $q = N/T = 50/1000 = 0.05$ and $\sigma=1$. For each value
-      of $\tau$ we generate 2000 instances of $N \times T$ random matrix
-      $R$, and compute C as $C=RR'/T$. Hence each curve in the figure is
-      constructed from 2000 sets of eigenvalues.
-    }
-  \label{fig:GaussianMarkovSpectrumPDF}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.4, clip=true, trim=100 226 116
-  133]{../pics/Gaussian_mineig.pdf}
-  \caption{\small \it The minimum eigenvalue versus
-    auto-correlation strength $\phi$. For each value of $\phi$ 2000
-    random matrices are generated and their eigenvalues are
-    calculated. The minimum eigenvalue of each random matrix is noted
-    and the mean of the 2000 such minimum eigenvalues are plotted
-    against the chosen value of $\phi$. 20 values of $\phi$ are
-    included in the plot, ranging from 0 to 0.95 with step size 0.05.}
-  \label{fig:Gaussian_mineig}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[scale=0.5, clip=true, trim=0 238 0
-    197]{../pics/GaussianMarkov005MaxEigCDF_loglog.pdf}
-  \end{center}
-  \caption{\small \it Cummulative distribution function (CDF) of the maximum
-    eigenvalue ($\lambda_1$). Blue: empirical CDF. Red: CDF of the
-    fitting gamma distribution. Each pair of CDFs correspond to a
-    fixed autocorrelation strength ($\phi$). The parameters $k$ and
-    $\theta$ of the gamma distribution are fit to data by matching the
-    2nd and the 3rd moments of the gamma distribution to the
-    corresponding moments of the empirical distribution. Then the
-    parameter $\alpha$ in equation \ref{eq:TracyWidom-Gamma} is chosen
-    to be $\alpha = k\theta - \E\left({\lambda_1 - \mu_{NT} \over
-        \sigma_{NT}}\right)$. The curves are plotted on log-log
-    scale.}
-  \label{fig:GaussianMarkov005MaxEigCDF_loglog}
-  %\vspace{-10mm}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=50 233 63
-  136]{../pics/Volvo_B_30m_returns.pdf}
-  \caption{\small \it Volvo B 30min returns' unconditional distribution fit to
-    the Stochastic Volatility model. The time series runs from
-    2013/10/10 to 2014/03/12. Left: $P(r < x)$ with $x < 0$; Right:
-    $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
-  \label{fig:Volvo_B_30m_returns}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=65 241 72
-  143]{../pics/Nordea_30m_returns.pdf}
-  \caption{\small \it Nordea Bank 30min returns' unconditional
-    distribution fit to the Stochastic Volatility model. The time
-    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
-    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
-  \label{fig:Nordea_30m_returns}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=21 241 28
-  146]{../pics/Ericsson_30m_returns.pdf}
-  \caption{\small \it Ericsson B 30min returns' unconditional
-    distribution fit to the Stochastic Volatility model. The time
-    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
-    $x < 0$; Right: $P(r > x)$ with $x > 0$. Both plots are on log-log scale.}
-  \label{fig:Ericsson_30m_returns}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=0 197 0
-  154]{../pics/garch_eigenvec_Cmax_corr.pdf}
-  \caption{\small \it Eigenvectors' largest component in
-    correspondence to the eigenvalues. N is fixed to 50.}
-  \label{fig:garch_eigenvec_Cmax_corr}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=85 258 103
-  229]{../pics/GarchSpectrumAutocorrelated.pdf}
-  \caption{\small \it Eigenvalue distribution of cross-correlation
-    matrix built from auto-correlated GARCH processes. $\phi$ = 0,
-    0.5, 0.8, 0.955, 0.97 correspond to correlation time $\tau$ =
-    0, 1.00, 3.11, 15.05, 22.76.}
-  \label{fig:GarchSpectrumAutocorrelated}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=95 260 111
-    232]{../pics/garch_eigenvec_PR_dist_corr.pdf}
-    \label{fig:garch_eigenvec_PR_dist_corr}
-  }
-  \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=95 260 111
-    232]{../pics/garch_eigenvec_Cmax_dist_corr.pdf}
-    \label{fig:garch_eigenvec_Cmax_dist_corr}
-  }
-  \caption{\small \it \ref{fig:garch_eigenvec_PR_dist_corr}: PDF of
-    the normalized participation ratio
-    (PR). \ref{fig:garch_eigenvec_Cmax_dist_corr}: PDF of the largest
-    component of the eigenvectors. N is fixed to 50. $\phi$ values of
-    0, 0.6000, 0.9000, 0.9900 correspond to correlation time $\tau$ =
-    0, 1.3569, 6.5788, 68.9676}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \subfigure[]{
-    \includegraphics[scale=0.38, clip=true, trim=93 229 115
-    134]{../pics/localization_ratio.pdf}
-    \label{fig:localization_ratio}
-  }
-  \subfigure[]{
-    \includegraphics[scale=0.38, clip=true, trim=87 227 115
-    133]{../pics/localization_ratio2.pdf}
-    \label{fig:localization_ratio2}
-  }
-  \caption{\small \it \ref{fig:localization_ratio}: Ratio of localized
-    eigenvectors versus the auto-correlation strength $\phi$. Black
-    ``+'': ratio of localized eigenvectors as measured by the number
-    of participating basis vectors being lower than 2; Black ``x'':
-    ratio of localized eigenvectors as measured by the largest
-    component being larger than 0.9. Blue curve: quadradic function
-    fitted to ``+''. Green curve: quadratic function fitted to
-    ``x''. There are 27 data points in the plot, corresponding to 27
-    $\phi$ values: 0 to 0.8 with step size 0.05, and 0.9 to 0.99 with
-    step size 0.01. The corresponding values of the correlation time
-    $\tau$ range from 0 to 69. \ref{fig:localization_ratio2}: $\ln
-    (f_\M - f)$ is plotted versus $\ln \phi$, where $f$ stands for the
-    ratio of localized eigenvectors. The fitted curves are linear.}
-\end{figure}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    & $\psi$ & $\sigma$ & $\bar{v}$ & $\mu$ \\
-    \hline
-    Volvo B & -1.5747e-02 & 4.6929e-01 & -6.2304e+00 &
-    -5.3040e-05 \\
-    Nordea Bank & 1.7234e-02 & 4.2780e-01 & -6.3758e+00 & 7.5230e-05
-    \\
-    Ericsson B & 1.8486e-02 & 4.4905e-01 & -6.2834e+00 & -9.2817e-05 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Parameter Values of Selected Assets'
-    returns. The time series are 30-minute returns and run from
-    2013/10/10 to 2014/03/12.}
-  \label{tab:assets_params}
-\end{table}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    & mean & std & skewness & kurtosis \\
-    \hline
-    \multirow{2}{*}{Volvo B} & -8.6766e-05 & 2.4500e-03 &
-    -2.7217e-02 & 7.3394e+00 \\
-    & -6.9281e-05 & 2.4538e-03 & -7.2677e-02 & 7.2472e+00 \\
-    \hline
-    \multirow{2}{*}{Nordea Bank} & 7.4821e-05 & 2.0500e-03 &
-    6.9433e-02 & 6.7427e+00 \\
-    & 8.8983e-05 & 2.0443e-03 & 6.7125e-02 & 6.2446e+00 \\
-    \hline
-    \multirow{2}{*}{Ericsson B} & -8.2606e-05 & 2.2939e-03 &
-    2.6586e-01 & 7.6716e+00 \\
-    & -7.5675e-05 & 2.2844e-03 & 7.8573e-02 & 6.7299e+00 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Moments of Selected Assets' returns. For each return
-    series, the 1st row contains the sample moments, while the 2nd
-    constains those computed with MLE parameters. The time series are
-    30-minute returns and run from 2013/10/10 to 2014/03/12.}
-  \label{tab:assets_moments}
-\end{table}
+\input{summary}
+\input{appendix}
 

### 2014-05-10 thesis/cross_correlation_gaussian.tex
Revision following Sven's advice.

@@
 \chapter{Correlation Matrix of Gaussian Returns}\label{chp:Gaussian}
-In this chapter we present some analytical and numerical results about
-the cross-correlation matrix (cf. \S
-\ref{sec:FundamentalConcepts}), in the special case where the involved
-asset returns are assumed to be Gaussian distributed. This assumption
-is of course an over-simplification, but nevertheless lends some
-insight into the problem.
+In this chapter we present some results about the cross-correlation
+matrix of Gaussian returns. This assumption is of course an
+over-simplification, but nevertheless lends some insight into the
+problem.
 
 Our primary interest is in the influence of autocorrelations on the
-cross-correlation matrix. \S \ref{sec:GCC-analytical} discusses the
-distribution of the matrix entries and \S \ref{sec:GCC-numerical} the
-distribution of the eigenvalues.
+cross-correlation matrix. \S\ref{sec:GCC-analytical} and
+\S\ref{sec:GCC-numerical} discuss the distribution of the matrix
+elements and the eigenvalues, respectively.
 
@@
 i.e. $\vec{r}_t \sim N(0, \Sigma)$, and are not
-auto-correlated, the L\'evy index $\alpha$ as appears in equation
-\ref{eq:cross-correlation} is 2, and $RR'$ is a Wishart matrix with
+auto-correlated, then $RR'$ is a Wishart matrix with
 the following distribution law \cite{Anderson2003}:
@@
 
-As before we consider the AR(1) process (see
-\S\ref{sec:FundamentalConcepts}):
+As before we consider the AR(1) process:
 \begin{equation*}

### 2014-05-10 thesis/price_models.tex
Revision following Sven's advice.

--- 2014-05-08 thesis/price_models.tex
@@
 identically distributed Gaussian variates comes from imagining the
-price process $S(t)$ as a Brownian motion whose increment $\sigma
-dw_t$ at each time step is independent and scales as $\sigma
+price process $S(t)$ as a geometric Brownian motion whose increment
+$\sigma dw_t$ at each time step is independent and scales as $\sigma
 \sqrt{dt}$. Then, by adding a drift term $\mu dt$ that represents some
@@
 
-The picture depicted above is of course overly simplified, and as we
-discussed in \S\ref{sec:StylizedFacts}, the distribution of returns is
-not really Gaussian. Nevertheless, the assumption of Gaussian
-distributed returns underlies such important theories as Black and
-Scholes theory of option pricing. Thus, albeit hugely inaccurate, the
-Gaussian distribution as a returns model shall not be forgotten. In
-the next few sections, we discuss some more realistic returns models.
+The picture depicted above is of course overly simplified, and the
+distribution of returns is not really Gaussian. Nevertheless, the
+assumption of Gaussian distributed returns underlies such important
+theories as Black and Scholes theory of option pricing. Thus, albeit
+inaccurate in describing very large returns, the Gaussian distribution
+as a returns model shall not be forgotten. In the next few sections,
+we discuss some more realistic returns models.
 
 \item{\bf GARCH models}
+
 ``GARCH'' is the acronym for ``Generalized Autoregressive Conditional
 Heteroscedasticity''. A GARCH(p, q) model is defined by the following
-equation system:
+equation system \cite{Bollerslev86}:
 \begin{equation}
@@
   \begin{aligned}
-    r_t &= \mu + \epsilon_t \\
-    \sigma_t^2 &= \alpha_0 + \sum_{i=1}^q \alpha_i \epsilon_{t-i}^2 +
-    \sum_{i=1}^p \beta_i \sigma_{t-i}^2
+    r_t &= \mu_t + \epsilon_t \\
   \end{aligned}
 \end{equation}
-where $\epsilon_t$ is termed the innovation of the return $r_t$ and is
-a random variable of a prescribed distribution, the most common ones
-being Gaussian, Student's t, Asymmetric Student's t, etc; $\sigma_t^2$
-is the variance of $\epsilon_t$ \cite{Bollerslev86}. The
-autocorrelation function of $\epsilon_t^2$ is given by
-(c.f. \cite{Bollerslev87})
-\begin{eqnarray*}
-\rho_n &=& \sum_{i=1} ^{p \vee q} (\alpha_i + \beta_i) \rho_{n-i}
-\text{ for $n > p$}
+where $p$ and $q$ are constant integers; $\mu(t)$ is the mean process
+and considered a small constant for intraday returns. $\epsilon_t$ is
+called the innovation of the return and is a variate whose conditional
+distribution \footnotemark is Gaussian and has variance $\sigma^2_t$:
+\begin{eqnarray*}
+  \sigma_t^2 &=& \alpha_0 + \sum_{i=1}^q \alpha_i \epsilon_{t-i}^2 +
+  \sum_{i=1}^p \beta_i \sigma_{t-i}^2
+\end{eqnarray*}
+\footnotetext{
+By conditional distribution we mean the distribution conditional on
+the variate's history: The conditional distribution at time $t$ means
+the distribution conditional on the history up to time $t-1$
+in the discrete-time setting.
+}
+
+It can be shown that the autocorrelation function of $\epsilon_t^2$ is
+given by \cite{Bollerslev87}:
+\begin{eqnarray*}
+\text{corr}(\epsilon_t^2, \epsilon^2_{t-n}) &=& \sum_{i=1} ^{p \vee q}
+(\alpha_i + \beta_i) \rho_{n-i} \text{ for $n > p$}
 \end{eqnarray*}
@@
 \item {\bf Stochastic Volatility Models}\label{sec:SLV_model}
+
 For the purpose of intraday returns, we specify the stochastic
@@
 \begin{eqnarray}
-   r_{t, t-h} &=& \ln p_{t} - \ln p_{t-h} \nonumber \\
+   r_{t, t-h} &=& \ln S_{t} - \ln S_{t-h} \nonumber \\
    r_{t, t-h} &=& \mu + \sigma_{t, t-h} b_t \label{eq:SLV_spec}
 \end{eqnarray}
-where $p_t$ is the price of the asset at time $t$; $b_t \sim N(0,
-1)$; $r_{t, t-h}$ is the return in the time interval $[t-h, t]$.
+where $S_t$ is the price of the asset at time $t$; $b_t \sim N(0,
+1)$; $r_{t, t-h}$ is the return over the time interval $[t-h, t]$.
 For simplicity, in the rest of this chapter we shall just write
@@
   \int_{0}^h \sigma_{t-h+s}^2 ds = \sum_{k=1}^n \left(
-    p_{t-h+kh/n} - p_{t-h+(k-1)h/n} \right)^2
+    S_{t-h+kh/n} - S_{t-h+(k-1)h/n} \right)^2
 \end{equation}
@@
 during the period 2012/01/16 - 2012/04/20 in \S\ref{chp:nordea_15min},
-and {\it Volvo B} 30-minute returns during 2013/10/10 - 2014/04/04.
-
-\section{Case Study: Nordea in 1st quarter 2012}
+and {\it Volvo B} 30-minute returns during 2013/10/10 - 2014/04/04 in
+\S\ref{sec:volvo}. Another 5 intraday series are also studied and
+collected in appendix \ref{chp:appendix2}.
+
+\section{Case Study: Nordea 15-minute Returns}
 \label{chp:nordea_15min}
@@
 \subsection{GARCH Model}\label{sec:nordea_15min_garch}
-First of all, to check for GARCH effects, we plot the auto-correlation
-function (ACF) of the squared returns. This is shown in figure
-\ref{fig:nordea_15min_acf}. At the absence of GARCH effects, the ACF
-is expected to have an asymptotic Gaussian distribution with mean 0
-and variance 1/T \cite{Bollerslev86, Bollerslev87}, which is clearly
-not the case in figure \ref{fig:nordea_15min_acf} --- the first 5
-autocorrelations have comparable sizes and do not fall off as in a
-Gaussian scheme. Moreover, figure \ref{fig:nordea_15min_vlt_acf} shows
-even more clearly that the conditional variances of the series are
-correlated. These observations suggest a GARCH(p, q) model can be
-appropriate.
+First of all, to determine that the volatilities are auto-correlated
+and/or that the squared returns and the volatilities are correlated, and
+hence a GARCH(p, q) model is appropriate for the return series under
+investigation, we plot the auto-correlation function (ACF) of the
+squared returns. This is shown in figure
+\ref{fig:nordea_15min_acf}. If the aforementioned features are absent
+from the series, the auto-correlations are expected to be Gaussian
+distributed with mean 0 and variance 1/T and hence mostly reside
+within the confidence bounds set by $\pm 2/\sqrt{T}$
+\cite{Bollerslev86, Bollerslev87}.
+
+Clearly this is not the case in figure \ref{fig:nordea_15min_acf} ---
+the first 5 auto-correlations are rather significant. Moreover, figure
+\ref{fig:nordea_15min_vlt_acf} shows even more clearly that the
+conditional variances of the series are correlated. These observations
+suggest a GARCH(p, q) model can be appropriate.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=90 257 105
+    220]{../pics/nordea_15min_acf.pdf}
+    \label{fig:nordea_15min_acf}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=90 259 104
+    218]{../pics/nordea_15min_vlt_acf.pdf}
+    \label{fig:nordea_15min_vlt_acf}
+  }
+  \caption{\small \it \ref{fig:nordea_15min_acf}: Auto-correlations
+    (ACF) of the squared returns; \ref{fig:nordea_15min_vlt_acf}:
+    Auto-correlations (ACF) of squared realized volatilities
+    ($\hat{\sigma}^2_t$). The blue lines are confident bounds set at
+    $\pm 2/\sqrt{T}$.}
+\end{figure}
 
@@
 \begin{eqnarray*}
-  r_t &=& \mu + €_t \\
-  €_t &=& \sigma_t z_t \\
-  \sigma^2_t &=& \alpha_0 + \alpha_1 €^2_{t-1} + \alpha_s €^2_{t-s} +
-  \beta_1 \sigma^2_{t-1} + \beta_s \sigma^2_{t-s}
+  r_t &=& \mu + \epsilon_t \\
+  \epsilon_t &=& \sigma_t z_t \\
+  \sigma^2_t &=& \alpha_0 + \alpha_1 \epsilon^2_{t-1} + \alpha_s
+  \epsilon^2_{t-s} + \beta_1 \sigma^2_{t-1} + \beta_s \sigma^2_{t-s}
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-  {\rho_{s+1}/\rho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
+  {\varrho_{s+1}/\varrho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
     1 + \theta^2} \\
-  {\rho_{s+1}/\rho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
+  {\varrho_{s+1}/\varrho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
     1 + \Theta^2} \\
@@
     \hline
-    $\rho_1$ & $\rho_{s-1}$ & $\rho_s$ & $\rho_{s+1}$ \\
+    $\varrho_1$ & $\varrho_{s-1}$ & $\varrho_s$ & $\varrho_{s+1}$ \\
     \hline
@@
 The first 4 moments of the Johnson Su distribution are expressible
-in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}:
-\begin{eqnarray*}
-  w &=& \exp{1 \over \delta^2} \\
-  \Omega &=& {\gamma \over \delta} \\
-  \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
-  \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
-    1)\right]^{1/2} \\
-  \text{skewness}(y) &=& {
-    \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
-    \over
-    (w\cosh 2\Omega + 1)^{3/2}} \\
-  \text{kurtosis}(y) &=& {
-    w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
-    + 3(2w+1) \over
-    2(w\cosh 2\Omega + 1)^2 }
-\end{eqnarray*}
+in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}.
+% \begin{eqnarray*}
+%   w &=& \exp{1 \over \delta^2} \\
+%   \Omega &=& {\gamma \over \delta} \\
+%   \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
+%   \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
+%     1)\right]^{1/2} \\
+%   \text{skewness}(y) &=& {
+%     \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
+%     \over
+%     (w\cosh 2\Omega + 1)^{3/2}} \\
+%   \text{kurtosis}(y) &=& {
+%     w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
+%     + 3(2w+1) \over
+%     2(w\cosh 2\Omega + 1)^2 }
+% \end{eqnarray*}
 By matching the theoretical expressions of the moments with their
@@
 
-The eventual MLE is done in Matlab with the ``active set''
-algorithm. The initial as well as the final estimation results are
-listed in table \ref{tab:nordea_15min_js_param}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|c|c|}
-    \hline
-    & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
-    \hline
-    initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
-    0.6378 \\
-    \hline
-    MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
-    \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Nordea 15min estimation results}
-  \label{tab:nordea_15min_js_param}
-\end{table}
+% The eventual MLE is done in Matlab with the ``active set''
+% algorithm. The initial as well as the final estimation results are
+% listed in table \ref{tab:nordea_15min_js_param}:
+% \begin{table}[htb!]
+%   \centering
+%   \begin{tabular}{|c|c|c|c|c|c|c|}
+%     \hline
+%     & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
+%     \hline
+%     initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
+%     0.6378 \\
+%     \hline
+%     MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
+%     \\
+%     \hline
+%   \end{tabular}
+%   \caption{\small \it Nordea 15min estimation results}
+%   \label{tab:nordea_15min_js_param}
+% \end{table}
 
@@
   \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
-    sample mean forecasts. Left: cumulative distribution function;
-    Right: complementary distribution function.}
+    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
+    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
   \label{fig:nordea_2012}
@@
     \hline
-    1\% & 0.2173 & 0.1383 & 0.1136 \\
-    \hline
-    5\% & 0.7457 & 0.7062 & 0.5333 \\
-    \hline
-    10\% & 0.9728 & 0.9383 & 0.8914 \\
+    1\% & 21.73\% & 13.83\% & 11.36\% \\
+    \hline
+    5\% & 74.57\% & 70.62\% & 53.33\% \\
+    \hline
+    10\% & 97.28\% & 93.83\% & 89.14\% \\
     \hline
@@
 
-\section{Case Study: Volvo in 2013/14}
-\label{sec:volvo}
+\section{Case study: Volvo 30-minute Returns}
+\label{sec:volvo_30}
 In this section we model the log-volatility of {\it Volvo B} 30-minute
@@
   \centering
-    \includegraphics[scale=0.5, clip=true, trim=0 256 0
+    \includegraphics[scale=0.4, clip=true, trim=0 256 0
     151]{../pics/volvo_inno_and_lv_acf.pdf}
@@
     \label{fig:volvo_slv_garch_cmp}
-  \caption{\small \it Comparison of the forecast from SV, GARCH,
-    and sample mean. Blue: SV forecast; Green: GARCH forecast; Red:
-    Sample mean forecast.}
+  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
+    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln
+    \hat{\sigma}_t < x)$; Right: $P(\ln \sigma^F_t - \ln
+    \hat{\sigma}_t > x)$. Horizontal: $x$.}
 \end{figure}
@@
     \hline
-    1\% & 0.2175 & 0.1220 & 0.1406 \\
-    \hline
-    5\% & 0.7162 & 0.6154 & 0.6631 \\
-    \hline
-    10\% & 0.9231 & 0.8806 & 0.8992 \\
+    1\% & 21.75\% & 12.20\% & 14.06\% \\
+    \hline
+    5\% & 71.62\% & 61.54\% & 66.31\% \\
+    \hline
+    10\% & 92.31\% & 88.06\% & 89.92\% \\
     \hline
@@
 We see from the table that the SV model consistently excels over
-the other two alternatives.
+the other two alternatives. In addition, it is also noted that the
+GARCH forecast is even worse than the sample mean. This is surprising
+but not totally unexpected --- with only 3 parameters, the GARCH(1,1)
+model can only describe the most prominent auto-correlations in the
+volatility. When the volatility is influenced by relatively weak
+auto-correlations at several different time lags, the GARCH forecast
+cannot be expected to have good accuracy.
+
+In this particular case, we see that the ARIMA model has 3 relatively
+small moving average coefficients, located at lags 2, 3, and 4 and
+evaluated to 0.06, 0.06, 0.03, suggesting a scattered auto-correlation
+structure, so the GARCH model cannot be expected to perform very
+well. In contrast, the Volvo 15-minute returns studied in
+\S\ref{sec:volvo} has more concentrated auto-correlations --- 0.12 and
+0.06 at lags 2 and 3 --- thus the GARCH(1,1) model is also found to
+perform better and even marginally better than the SV model for the
+same series.
 
@@
 $x$, $f_{r'}(x)$ may be approximated by $C/|x|^\alpha$, i.e. a power law.
-
-In the following we work out the normalization constant C.
-\begin{eqnarray*}
-  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {1 \over C}{e^{\sigma^2 / 2} \over
-    \sqrt{8\pi}} \int_{-\infty}^{\infty} \text{erfc} \left({1 \over
-      \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma \over
-      \sqrt{2}} \right) dx\\
-  &=& {2 \over C}{e^{\sigma^2 / 2} \over
-    \sqrt{8\pi}} \int_{0}^{\infty} \text{erfc} \left({1 \over
-      \sqrt{2}\sigma} \ln{x \over \sqrt{\ln 4}} + {\sigma \over
-      \sqrt{2}} \right) dx\\
-\end{eqnarray*}
-Let
-\begin{eqnarray*}
-  a &=& {1 \over \sigma \sqrt 2} \\
-  b &=& -{1 \over 2 \sigma \sqrt 2}\ln\ln 4 + {\sigma \over \sqrt 2} \\
-  y &=& a\ln|x|+b \\
-\end{eqnarray*}
-Then
-\begin{eqnarray*}
-  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {2 \over C}{e^{\sigma^2 / 2} \over
-    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {e^{(y-b)/a} \over a}
-  \text{erfc}(y) \\
-  &=& {2 \over C}{e^{\sigma^2 / 2} \over
-    \sqrt{8\pi}} \left.e^{(y-b)/a}
-    \text{erfc}(y)\right|_{y=-\infty}^{\infty}
-  + {2 \over C}{e^{\sigma^2 / 2} \over
-    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {2 \over \sqrt \pi}
-  e^{(y-b)/a} e^{-y^2} \\
-  &=& {2 \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}  2 e^{\ln\ln 4/2
-    - \sigma^2/2}\\
-  C &=& \sqrt{2 \ln 4\over \pi} \\
-  &\approx& 0.9394
-\end{eqnarray*}
-Thus we can write in summary:
+The normalization constant $C$ is calculated in appendix
+\ref{chp:symmetric_SV_norm_const} to be $C = \sqrt{2 \ln 4\over
+  \pi}$. So, in summary, we can write:
 \begin{eqnarray*}

### 2014-05-10 thesis/thesis.tex
Revision following Sven's advice.

@@
 %\HRule \\[0.4cm]
-{ \huge \bfseries Returns Models and Their Implications
-  on Cross-Correlation Matrix \\[0.4cm] }
+{ \huge \bfseries Returns Models and Their Covariance Matrix
+  \\[0.4cm] }
 
@@
 a given asset. Hence they are a central topic of the mathematics of
-the financial market.
-
-Once a model has been tentatively constructed, it must be
-verified. The verification again relies on the statistics of the
-observables. Over the years, a number of phenomena have been
-consistently reported and become generally accepted. These are referred
-to as stylized facts, which are detailed in \S
-\ref{sec:StylizedFacts} and constitute a standardized test suit for
-model verification.
+the financial market. In the domain of discrete-time returns' models,
+GARCH and stochastic volatility models (SV) are the most popular, so in
+this thesis project we study a total of 7 intraday return series 
+using GARCH and SV models and compare their performance in volatility
+forecast.
+
+% Once a model has been tentatively constructed, it must be
+% verified. The verification again relies on the statistics of the
+% observables. Over the years, a number of phenomena have been
+% consistently reported and become generally accepted. These are referred
+% to as stylized facts, which are detailed in \S\ref{sec:StylizedFacts}
+% and constitute a standardized test suit for model verification.
 
@@
 complicated than a straight-forward Gaussian or L\'evy distribution can
-describe - instead, one needs structured models for their description.
+describe --- instead, one needs structured models for their description.
 
@@
 built from returns described by different returns' models, specifically
-GARCH(1,1) and ARIMA-log-volatility models, whose the tail behavior is
+GARCH(1,1) and stochastic volatility models, whose the tail behavior is
 understood. Moreover, we also study how auto-correlations in the
-returns influence the aforementioned distributions -- such
+returns influence the aforementioned distributions --- such
 auto-correlations, known as second-order auto-correlations decay
@@
 
-The rest of this report is organized as follows: in
-\S\ref{sec:FundamentalConcepts} we present a few very important
-concepts and notations that we will often refer to in later
-chapters. Then chapter \ref{chp:PriceModels} reviews some of the most
-influential returns' models, calibrates them to a few intra-day return
-series and compares their performance. The unconditional distribution
-functions of ARIMA-log-volatility models, which is not seen to be
-given a full discussion in the literature, are also calculated and
-discussed. Chapter \ref{chp:Gaussian} investigates the elements' and
-eigenvalues' distributions of the Wishart matrix, where the influence
-of auto-correlations in the constituent returns are studied in
-detail. Chapter \ref{chp:CrossCorrelationFat} looks at the elements'
-and eigenvalues' distributions
-
-\section{Fundamental Concepts \&
-  Notations}\label{sec:FundamentalConcepts}
-This section is a list of a few concepts that may be unfamiliar to the
-reader and that we will often refer to later in the thesis:
-\begin{itemize}
-\item Return. Given a fixed time interval $[t - \Delta t, t]$, for example, a
-  day, a week, a month, etc, and a particular asset, for example, a
-  share in company ABC, the return of this asset over the time
-  interval is defined as
-  \begin{eqnarray*}
-    r_{t, \Delta t} &=& \ln p_{t} - \ln p_{t - \Delta t} \\
-    &=& \ln \left(1 + {p_{t} - p_{t - \Delta t} \over p_{t - \Delta t}}\right) \\
-    &\approx& {p_{t} - p_{t - \Delta t} \over p_{t - \Delta t}}
-  \end{eqnarray*}
-  where $p_t$ is the price of the asset at time $t$. $\Delta t$ is
-  sometimes called the time-lag of the return. Quite often, where
-  confusion is not possible, we will just write $r_t$ to mean $r_{t,
-    \Delta t}$, the time lag either does not matter or is clear from
-  the context.
-
-\item Autocorrelation. By autocorrelation, denoted $\rho_k$ here, we
-  mean the correlation between two temporally separated observations
-  of the same time series:
-  \begin{eqnarray*}
-    \rho_k &=& {
-      \E\left[(a_t -\E(a_t))(a_{t-k} - \E(a_{t-k})\right]
-      \over
-      \sqrt{\text{var}(a_t)}\sqrt{\text{var}(a_{t-k})}
-    }
-  \end{eqnarray*}
-  Here $\E(x)$ stands for the expectation value of $x$, and
-  $\text{var}(x)$ stands for the variance of $x$. $k$ is called the
-  time-lag and is the temporal seperation of the two observations
-  measured by the number of observations in between. For example, the
-  time-lag between the 1st and 3rd observation is 2.
-
-\item Cross-Correlation Matrix. Correlations between the returns of a
-  group of assets are described by the cross-correlation matrix. When
-  the asset returns are described by a stable distribution law with
-  L\'evy index $\alpha$, the empirical cross-correlation matrix is
-  constructed as
-  \begin{equation}
-    \label{eq:cross-correlation}
-    \begin{aligned}
-      C_{ij} &= {1 \over T^{2/\alpha}} \sum_{t=1}^T [r_{i,t}-\E(r_i)]
-      [r_{j,t}-\E(r_j)] \\
-      &= {1 \over T^{2/\alpha}} RR'
-    \end{aligned}
-  \end{equation}
-  where $r_{i,t}$ is the return of the i-th asset at time t and is
-  placed at the entry (i, t) of matrix R; $R'$ denotes the
-  transpose of R. In most practical situations, one has abundant data
-  for each and every asset. Thus $T \geq N$ is assumed throughout this
-  thesis.
-
-\item Auto-regressive processes. A time series $r_t$ is called an
-  auto-regressive process of order $p$ and denoted AR(p), if it can be
-  written in the following form:
-  \begin{eqnarray*}
-    r_t &=& \sum_{i=1}^p \phi_i r_{t-i} + a_t
-  \end{eqnarray*}
-  where, for all $i$, $\phi_i \in (-1, 1)$, and the $a_t$'s are
-  independent and identically distributed (iid.) random variables with
-  zero mean. Obviously this implies the mean of $r_t$ is zero
-  too. Apart from this, their distribution of $a_t$ is not restricted
-  to any particular form.
-
-  Of particular interest to this thesis is the AR(1) process:
-  \begin{eqnarray*}
-    r_t = \phi r_{t-1} + a_t
-  \end{eqnarray*}
-  Its autocorrelation function $\rho_k = \text{corr}(r_t, r_{t-k})$
-  ($k = 0, 1, 2, \cdots$) can be easily shown to fall off
-  exponentially:
-  \begin{eqnarray*}
-    \rho_k &=& {\E(r_tr_{t-k}) - \E(r_{t-k})\E(r_{t}) \over
-      \sqrt{\var(r_t) \var(r_{t-k})}} \\
-    &=& {\phi \E(r_{t-1}r_{t-k}) + \E(a_t r_{t-k})
-      \over
-      \sqrt{\var(r_t) \var(r_{t-k})}
-    } \\
-    &=& \phi \rho_{k-1}
-  \end{eqnarray*}
-  where $\E(a_t r_{t-k}) = 0$ follows from the fact that any return
-  $r_{t1}$ must not depend on disturbances $a_{t2}$ that occur later
-  in time. The last equation means $\rho_k$ is a geometric series. Since
-  $\rho_0 = 1$, we have
-  \begin{eqnarray*}
-    \rho_k &=& \phi^k \\
-    |\rho_k| &=& e^{k\ln|\phi|} \\
-  \end{eqnarray*}
-  Note that $|\phi| < 1$ and hence $\ln|\phi| < 0$.
-
-  Although $k$ can only take integer values, it is still useful to
-  define a correlation time $\tau$ such that $\phi^\tau = 1/2$. Such a
-  quantity is more intuitive and constitutes a measure of
-  autocorrelations that is universal and comparable among different
-  time series' models. From the definition of $\tau$ we get
-  \begin{equation}
-    \label{eq:tau_def}
-    \begin{aligned}
-      \phi^\tau &= 1/2 \\
-      \tau &= -{\ln 2 \over \ln\phi} \\
-      \phi &= 2^{-1/\tau}
-    \end{aligned}
-  \end{equation}
-  Figure \ref{fig:AR1-autocorrelation} shows the autocorrelation
-  function of the AR(1) model. As proven above, this function decays
-  exponentially.
-  \begin{figure}[htb!]
-    %\vspace{-15mm}
-    \centering
-    \includegraphics[scale=0.5, clip=true, trim=113 229 115
-    139]{../pics/AR1-autocorrelation.pdf}
-    \caption{\small \it Autocorrelation function of the AR(1) model
-      with $\phi=1/\sqrt{2}$. Red circles: autocorrelations at $k=0, 1,
-      2, \cdots$. Blue line: $e^{t\ln\phi}$.}
-    \label{fig:AR1-autocorrelation}
-  \end{figure}
-  For more details about conventional time series' models, see
-  \cite{BoxJenkins94}.
-\end{itemize}
-
-For purposes of later reference, we also list some notations that may
-cause confusion to the reader:
-\begin{itemize}
-\item $\E(x)$ or $\mean{x}$: The expectation value of $x$.
-\item $\var(x)$: the variance of $x$.
-\item $\text{cov}(x,y)$: the covariance of $x$ and $y$.
-\item $\text{corr}(x,y)$: the correlation between $x$ and $y$, i.e.
-  \begin{equation*}
-    \text{corr}(x,y) = {\text{cov}(x,y) \over \sqrt{\var(x)\var(y)}}
-  \end{equation*}
-\end{itemize}
-
-\section{Stylized facts}\label{sec:StylizedFacts}
-This section presents and explains a few ``stylized facts'',
-i.e. phenomena that are widely observed and accepted as true.
-\begin{itemize}
-\item Fat tails. It has been consistently reported by various studies
-  - for example \cite{Potters2003} and \cite{Mantegna2000} - that the
-  probability density function (PDF) of stock/index returns are not
-  Gaussian. Unlike the Gaussian PDF, which is symmetric and falls off
-  very quickly as its argument moves from the center to the outskirts
-  (tails), the PDF of stock/index returns are higher on the tails,
-  i.e. the probability of large fluctuations is higher than is dictated
-  by a Gaussian distribution - in fact, even higher than dictated by
-  an exponential function. Empirical studies suggest the distribution
-  function of the returns follows a power-law on the tails --- the
-  exponent of the power depends on the specific stock/index.
-
-  Figure \ref{fig:FatTail} illustrates this feature.
-  \begin{figure}[htb!]
-    \centering
-    \includegraphics[scale=0.6, clip=true, trim=92 229 110
-    140]{../pics/FatTail.pdf}
-    \caption{\small \it Fat tails of {\it Nordea Bank} 15-minute
-      returns during the period 2013-10-10 and 2014-01-29. The returns
-      are computed using minute-by-minute average prices.}
-    \label{fig:FatTail}
-  \end{figure}
-
-\item Non-zero skewness. Apart from fat tails, the PDF of stock/index
-  returns are often also skewed. If the skewness is positive
-  (negative), the probability of very large positive (negative)
-  returns is higher than that of very large negative (positive)
-  returns, even though the mean of the returns is 0 or extremely
-  close to 0.
-
-  Table \ref{tab:EmpiricalSkewness} lists the skewness of a few
-  Swedish stocks traded on the Stockholm OMX market.
-  \begin{table}[htb!]
-    \footnotesize
-    \centering
-    \begin{tabular}{|c|c|c|c|c|c|c|}
-      \hline
-      Nordea Bank & Volvo B & Boliden & ABB Ltd & H\&M & Scania
-      B & Ericsson B \\
-      \hline
-      0.1362 & 0.0471 & 0.0567 & -0.0364 & -0.0168 &
-      -1.1554 & 0.2086 \\
-      \hline
-    \end{tabular}
-    \caption{\small \it Skewness of Stock Returns. All the returns
-      have time-lag of 15 minutes and are computed using paid prices
-      between 2013-10-10 and 2014-01-29. }
-    \label{tab:EmpiricalSkewness}
-  \end{table}
-
-\item Higher-order autocorrelation. To the lowest order, return series
-  are not auto-correlated --- if they are, the auto-correlations would
-  present an obvious opportunity of making easy profit and be
-  exploited and vanish as soon as they appear. However, the squared
-  returns do have significant auto-correlations, as figure
-  \ref{fig:nordea_15min_acf} illustrates. These auto-correlations
-  suggest the variances of the returns at subsequent time steps are
-  correlated. These are referred to as higher-order auto-correlations
-  and make a subjet of returns' models.
-
-\end{itemize}
+The rest of this report is organized as follows:
+
+% \S\ref{sec:FundamentalConcepts} presents a few very important
+% concepts and notations that will often be referred to in later
+% chapters; \S\ref{sec:StylizedFacts} gives a review of the ``stylized
+% facts'', i.e. a number of consistently reported and generally accepted
+% phenomena that must be accounted for by a returns' model.
+
+Chapter \ref{chp:PriceModels} reviews some of the most
+influential returns' models, fits them to a few intraday return
+series and compares their performance. A calculation of the
+unconditional distribution functions of stochastic volatility models,
+especially in the case where the residual of the log-volatility
+and the innovation of the return are correlated normal variates, is
+also presented. Chapter \ref{chp:Gaussian} investigates the elements' and
+eigenvalues' distributions of the Wishart matrix, and studies the influence
+of auto-correlated returns. Chapter \ref{chp:CrossCorrelationFat}
+looks at the elements' and eigenvalues' distributions of a covariance
+matrix that is built from identically specified GARCH(1,1) series.
 
@@
 \input{summary}
+\appendix
 \input{appendix}

### 2014-06-11 thesis/cross_correlation_gaussian.tex
Move matlab files to a separate folder.

@@
-\chapter{Correlation Matrix of Gaussian Returns}\label{chp:Gaussian}
-In this chapter we present some results about the cross-correlation
-matrix of Gaussian returns. This assumption is of course an
-over-simplification, but nevertheless lends some insight into the
-problem.
+\chapter{Covariance Matrix of Gaussian Returns}
+\label{chp:Gaussian}
+In this chapter we present some results about the covariance
+matrix of Gaussian returns. The assumption of Gaussian returns is of
+course an over-simplification of reallity, but nevertheless lends some
+insight into the problem of elements' and eigenvalues' distributions of
+a covariance matrix.
+Our primary interest is in the influence that auto-correlations in
+returns exert on the covariance matrix. Section \ref{sec:GCC-analytical} and
+section \ref{sec:GCC-numerical} discuss how the distributions of the matrix
+elements and the eigenvalues are affected by such autocorrelations,
+respectively.
 
-Our primary interest is in the influence of autocorrelations on the
-cross-correlation matrix. \S\ref{sec:GCC-analytical} and
-\S\ref{sec:GCC-numerical} discuss the distribution of the matrix
-elements and the eigenvalues, respectively.
+\section{Distribution of the Matrix Elements}
+\label{sec:GCC-analytical}
+In this section we study the distribution of the elements of a
+covariance matrix $C = \mtx{RR'}/T$, where
+\begin{eqnarray*}
+  \mtx R &=&
+  \begin{pmatrix}
+    r_{11} & r_{12} & \cdots & r_{1T} \\
+    r_{21} & r_{22} & \cdots & r_{2T} \\
+    \vdots & \vdots & \ddots & \vdots \\
+    r_{N1} & r_{N22} & \cdots & r_{NT} \\
+  \end{pmatrix}
+\end{eqnarray*}
+In words, an element $r_{it}$ of $\mtx R$ is the return of asset $i$
+at time $t$, with $i=1,\cdots, N$ and $t = 1, \cdots, T$. If each
+column of $\mtx R$ follows a zero-mean Gaussian distribution, i.e.
+\begin{eqnarray*}
+  \begin{pmatrix}
+    r_{1t} \\
+    \vdots \\
+    r_{Nt}
+  \end{pmatrix} \sim N(0, \mtx \Sigma)
+\end{eqnarray*}
+for all $t = 1, \cdots, T$, and none of the return series is
+auto-correlated, i.e. $\text{corr}(r_{it}, r_{i,t'}) = 0$ for all $i =
+1, \cdots, N$ and $t \neq t'$, then $\mtx{RR'}$ is a Wishart matrix whose
+probability density function is well known \cite{Anderson2003}.
 
-\section{Distribution of the Matrix Elements}\label{sec:GCC-analytical}
-If the returns follow a zero-mean Gaussian distribution, 
-i.e. $\vec{r}_t \sim N(0, \Sigma)$, and are not
-auto-correlated, then $RR'$ is a Wishart matrix with
-the following distribution law \cite{Anderson2003}:
-\begin{equation}
-  \label{eq:WishartPDF}
-  f_{RR'}(X) = { |\det X|^{(T-N-1)/2} \exp\left(-{1 \over 2}\tr
-      \Sigma^{-1}X \right)
-    \over
-    2^{NT/2}\pi^{N(N-1)/4}|\det \Sigma|^{T/2}
-    \prod_{i=1}^N \Gamma\left[(n+1-i)/2\right]
+When auto-correlations are indeed present in the returns, $\mtx{RR'}$
+no longer follows the Wishart distribution. However, the joint distribution
+function of its elements can be expressed in terms of the Wishart \gls{pdf}
+and the auto-correlations. We show this in appendix
+\ref{app:pdf_gaussian1}. Also, in appendix
+\ref{chp:gaussian_elements_dist} we derive an approximate expression
+for the asymptotic distribution of these matrix elements, assuming
+$r_{it}$ is an \gls{ar}(1) process \footnotemark, i.e. $r_{it} =
+\phi r_{i, t-1} + a_{it}$, and
+\footnotetext{
+  It is straight forward to derive the auto-correlation function
+  $\varrho_k$ of an \gls{ar}(1) process with autoregressive
+  coefficient $\phi$:
+  \begin{eqnarray*}
+    \varrho_k &=& \text{corr}(r_{it}, r_{i, t-k}) \\
+    &=& \phi \varrho_{k-1} \\
+    &=& \phi^k
+  \end{eqnarray*}
+  Let $\tau$ denote the time lag at which $\varrho_{\tau} = 1/2$, then
+  it follows $\tau = -\ln 2/\ln \phi$.
 }
-\end{equation}
-where $\Sigma$ is the true covariance matrix of the returns.
+\begin{eqnarray*}
+  \begin{pmatrix}
+    a_{1t} \\
+    \vdots \\
+    a_{Nt}
+  \end{pmatrix} \sim N(0, \mtx \Sigma)
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+  \mtx \Sigma &=& \sigma^2
+  \begin{pmatrix}
+    1 & \rho & \cdots & \rho \\
+    \rho & 1 & \cdots & \rho \\
+    \vdots & \vdots & \ddots & \vdots \\
+    \rho & \rho & \cdots & 1
+  \end{pmatrix}
+\end{eqnarray*}
+where $-1 < \rho < 1$ is a constant parameter describing the correlation
+between $a_{it}$ and $a_{jt}$ --- so, by construction, we assume such
+correlations are constant across all asset pairs and over all
+time.
 
-When $\vec{r}_t$ are indeed auto-correlated, $RR'$ no longer has the
-Wishart distribution. The joint distribution function of its elements
-can be expressed in terms of the Wishart PDF and the
-auto-correlations. We show this in appendix
-\ref{app:pdf_gaussian1}. In the rest of this section, we derive an
-approximate expression for the asymptotic distribution of these matrix
-elements.
-
-In the following we denote the diagonal elements as $C_{ii}$ and the
-non-diagonal elements as $C_{ij}$ ($i \neq j$). We assume that
-$\var(a_{i,t}) = \sigma^2$ for any $i$ and $t$, and $a_{i,t}$ are not
-autocorrelated, i.e. $\text{corr}(a_{i,t}, a_{i, t'}) = 0$ for any
-$i$, $t$ and $t'$. Then we can write
-\begin{eqnarray*}
-  r_{i,t} r_{j, t} &=& \left[
-    \phi_1 r_{i, t-1} + a_{i, t}
-  \right] \left[
-    \phi_1 r_{j, t-1} + a_{j, t}
-  \right] \\
-  C_{ij} &=& {1 \over T}\sum_{t=1}^T r_{i,t} r_{j, t} \\
-  &=& \phi_1^2 {1 \over T}\sum_{t=1}^Tr_{i,t-1} r_{j, t-1} +
-  \phi_1{1 \over T}\sum_{t=1}^T\left(
-    r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t}
-  \right) + {1 \over T}\sum_{t=1}^T a_{i,t} a_{j,t}
-\end{eqnarray*}
-We note that the two sums $\sum_{t=1}^T r_{i,t} r_{j, t}$ and
-$\sum_{t=1}^T r_{i,t-1} r_{j, t-1}$ only differ by the first and the
-last addend, which is negligible for sufficiently large T. Thus we
-have
-\begin{equation}\label{eq:Offdiag1}
-  \begin{aligned}
-    (1 - \phi_1^2)C_{ij} &\approx& \phi_1 {1 \over T} \sum_{t=1}^T
-    \left(r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t} \right)
-    + {1 \over T}\sum_{t=1}^T a_{i,t} a_{j,t}
-  \end{aligned}
-\end{equation}
-Now we write the AR(1) process $r_{i,t}$ as an infinite moving-average
-process:
-\begin{eqnarray*}
-  r_{i, t} &=& \phi_1 r_{i, t-1} + a_{i,t} \\
-  (1 - \phi_1 B) r_{i, t} &=& a_{i,t} \\
-\end{eqnarray*}
-where $B$ is the back-shift operator. Then it follows from the above
-equation
-\begin{eqnarray*}
-  r_{i,t} &=& {1 \over 1 - \phi_1 B} a_{i,t} \\
-  &=& \sum_{k=0}^\infty \phi_1^k B^k a_{i,t} \\
-  &=& \sum_{k=0}^\infty \phi_1^k a_{i,t-k} \\
-\end{eqnarray*}
-where it is left implicit that $a_{i, t}$ with $t \leq 0$ is zero (in
-words, this implies that the $r_{i,t}$ process is not affected at all
-by events before $t = 1$).
-
-Substituting this into eq.\ref{eq:Offdiag1} for $r_{i,t-1}$ yields
-\begin{equation}
-  \label{eq:Offdiag2}
-  \begin{aligned}
-  (1 - \phi_1^2)C_{ij} &\approx
-  {1 \over T} \sum_{t=1}^T \sum_{k=0}^\infty\phi_1^{k+1} (a_{i, t-k-1}
-  a_{j, t} + a_{j, t-k-1} a_{i, t})
-  + {1 \over T} \sum_{t=1}^T a_{i,t} a_{j,t} \\
-  &=
-  {1 \over T} \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi_1^k (a_{i, t-k}
-  a_{j, t} + a_{j, t-k} a_{i, t})
-  + {1 \over T} \sum_{t=1}^{T} a_{i,t} a_{j,t}\\
-  \end{aligned}
-\end{equation}
-
-Given two Gaussian random variables $x$ and $y$ with zero mean and
-covariance matrix 
-\begin{equation*}
-  \Sigma =
-  \begin{pmatrix}
-    1 & \rho \\
-    \rho & 1 \\
-  \end{pmatrix}
-\end{equation*}
-
-The PDF of $xy$ can be found by considering $P(xy < z)$:
-\begin{eqnarray*}
-  P(xy < z) &=& \left(\int_0^\infty dx \int_0^{z/x} dy
-    +\int_{-\infty}^0 dx \int_{z/x}^\infty dy \right)
-    {1 \over 2\pi \sqrt{1 - \rho^2}} \\
-    &&
-    \exp\left[
-      -{x^2 -2\rho xy + y^2
-        \over
-        2\sigma^2 (1 - \rho^2)} 
-    \right] \\
-  f(z; \sigma, \rho) &=& {d \over dz} P(xy < z)\\
-  &=& {1 \over \pi \sqrt{1 - \rho^2}} \exp\left[
-    {\rho z \over \sigma^2 (1 - \rho^2)}\right] K_0\left[
-    {|z| \over \sigma^2 (1 - \rho^2)}
-  \right]
-\end{eqnarray*}
-where $K_n(z')$ is the modified Bessel function of the second
-kind. It is worth taking note that, when $\rho \neq 0$,
-$f(z; \sigma, \rho)$ is not symmetric with respect to
-$z$. As a result, if $\rho > 0$, the mean of $f(z;
-\sigma, \rho)$ is positive, and vice versa.
-
-Secondly, because $|\rho| < 1$ and
-\begin{equation*}
-  K_0(z) \sim \sqrt{\pi \over 2z} e^{-z}
-\end{equation*}
-as $z \to \infty$ \cite{Olver:2010:NHMF}, all the moments of
-$f(z; \sigma, \rho)$ are finite, implying the
-applicability of the Lyapunov central limit theorem provided that $T$
-is large, which is very often the case and what we assume here.
-
-With this in mind, we observe that $\phi_1$ only affects the first sum in
-\ref{eq:Offdiag2}. If $a_{i,t-k}$ and $a_{j,t}$ are not correlated for
-non-zero $k$, $\phi_1$ will not affect the mean of
-$C_{ij}$. Furthermore, it is also clear from equation
-\ref{eq:Offdiag2} that the variance of $C_{ij}$ is
-always increased by a non-zero $\phi_1$, regardless of the sign of
-$\phi_1$. We compute this increment in the following.
-
-In light of the above expression for $f(z; \sigma,
-\rho)$, we rewrite equation \ref{eq:Offdiag2} as
-\begin{equation}\label{eq:Cij_dist}
-  \begin{aligned}
-    (1 - \phi_1^2)C_{ij} \approx &
-    {1 \over T}
-    \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi_1^k (a_{i, t-k} a_{j, t} + a_{j,
-      t-k} a_{i, t}) \\
-    & + {1 \over T} \sum_{t=1}^{T} \sigma^2 (1 - \rho^2) {a_{i,t} \over
-      \sigma \sqrt{1 - \rho^2}} {a_{j,t} \over \sigma \sqrt{1 - \rho^2}} \\
-  \end{aligned}
-\end{equation}
-In addition, we assume
-\begin{equation*}
-  \text{corr}(a_{i, t-k}, a_{j, t}) = \left\{
-    \begin{array}{l l}
-      1 & \text{ if $i=j$ and $k = 0$ }\\
-      \rho & -1 < \rho < 1. \text{ if $i \neq j$ and $k = 0$ }\\
-      0 & \text{otherwise}
-    \end{array}
-  \right.
-\end{equation*}
-Then, because $a_{i, t-k}$ and $ a_{j, t}$ with $i \neq j$ and $k
-> 0$ are not correlated, the mean of $a_{i, t-k} a_{j, t}$ is 0
-($f(z; \sigma, 0)$ is symmetric), and the variance of it
-can be found to be $\sigma^6$ using formula (10.43.19) of \cite{NIST:DLMF}:
-\begin{equation*}
-  \int_0^\infty dt K_\nu(t) t^{\mu-1} = 2^{\mu-2}
-  \Gamma\left(
-    {\mu + \nu \over 2}
-  \right) \Gamma\left(
-    {\mu - \nu \over 2}
-  \right)
-\end{equation*}
-On the other hand, $a_{i,t}/\sigma \sqrt{1 - \rho^2}$ and
-$a_{j,t}/\sigma \sqrt{1 - \rho^2}$ have variance $1/(1 - \rho^2)$
-and are correlated - $\text{corr}(a_{i,t}, a_{j,t}) = \rho$. The mean
-of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ can be found using formula
-(10.43.22) of \cite{NIST:DLMF}, given that $-1 < \rho < 1$:
-\begin{eqnarray*}
-  \int_0^\infty t^{\mu - 1} e^{-at} K_\nu(t) dt &=& (\pi/2)^{1/2}
-  \Gamma(\mu + \nu) \Gamma(\mu - \nu)(1 - a^2)^{-\mu/2 + 1/4} \times\\
-  && P^{-\mu+1/2}_{\nu-1/2} (a)
-\end{eqnarray*}
-where $P^\mu_\nu(\cdot)$ is Ferrers function of the first
-kind\footnote{ Ferrers function of the first kind is defined through
-  the hypergeometric functon $F(a, b; c; z)$ \cite{NIST:DLMF}:
-  \begin{equation}
-    \label{eq:Ferrers_1st}
-    \mathop{\mathsf{P}^{\mu}_{\nu}\/}\nolimits\!\left(x\right)=\left(\frac{1+x}{1-
-        x}\right)^{\mu/2}\mathop{\mathbf{F}\/}\nolimits\!\left(\nu+1,-\nu;1-\mu;\tfrac
-      {1}{2}-\tfrac{1}{2}x\right).
-  \end{equation}
-}. The result is
-\begin{eqnarray*}
-  \E\left[{a_{i,t}a_{j,t} \over \sigma^2 (1 - \rho^2)}\right]
-  &=&
-  {1 \over \sqrt{2\pi} (1 - \rho^2)^{5/4}} \left[
-    P^{-3/2}_{-1/2}(-\rho) - P^{-3/2}_{-1/2}(\rho)
-  \right]
-\end{eqnarray*}
-Similarly, the variance of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ is
-found to be
-\begin{eqnarray*}
-  v^2(\rho) &=&
-  {4 \over \sqrt{2\pi} (1 - \rho^2)^{7/4}} \left[
-    P^{-5/2}_{-1/2}(\rho) + P^{-5/2}_{-1/2}(-\rho)
-  \right] \\
-  && - \E^2\left[{a_{i,t}a_{j,t} \over
-      \sigma^2 (1 - \rho^2)}\right]
-\end{eqnarray*}
-Now we can apply the Lyapunov central limit theorem
-\cite{Billingsley1995} to the sum in equation \ref{eq:Cij_dist}
-and write down the asymptotic Gaussian distribution of $C_{ij}$:
-\begin{equation*}
-C_{ij} \sim N(\mu'_X, \sigma'_X)
-\end{equation*}
-where
+The elements $C_{ij}$ ($i \neq j$) of the covariance matrix, $C
+= \mtx{RR'}/T$, are found to be normally distributed with mean
 \begin{eqnarray}
-  \mu'_X &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi_1^2)(1 -
+  \mu'_X &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi^2)(1 -
     \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
     P^{-3/2}_{-1/2}(\rho)
-  \right] \label{eq:gaussian_mean}\\
-  \sigma'^2_X &=& {1 \over (1 - \phi_1^2)^2}\left[
+  \right] \label{eq:gaussian_mean2}
+\end{eqnarray}
+and variance
+\begin{eqnarray}
+  \sigma'^2_X &=& {1 \over (1 - \phi^2)^2}\left[
     \sum_{t=1}^T \sum_{k=1}^{t-1} 2\left(
-      \phi_1^k \over T
+      \phi^k \over T
     \right)^2 \sigma^6 + \sum_{t=1}^T
@@
   \right] \nonumber \\
-  &=& {2 \sigma^6 \over T (1 - \phi_1^2)^2} \left[
-    {\phi_1^2 \over 1 - \phi_1^2} -
-    {\phi_1^2 (1 - \phi_1^{2T}) \over
-      T(1 - \phi_1^2)}
+  &=& {2 \sigma^6 \over T (1 - \phi^2)^2} \left[
+    {\phi^2 \over 1 - \phi^2} -
+    {\phi^2 (1 - \phi^{2T}) \over
+      T(1 - \phi^2)}
   \right] + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
-    T (1 - \phi_1^2)^2} \nonumber \\
-  &\approx& {2 \sigma^6 \phi_1^2 \over T (1 - \phi_1^2)^3}
+    T (1 - \phi^2)^2} \nonumber \\
+  &\approx& {2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}
   + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
-    T (1 - \phi_1^2)^2} \label{eq:gaussian_variance}
+    T (1 - \phi^2)^2} \label{eq:gaussian_variance2}
 \end{eqnarray}
-Equation \ref{eq:gaussian_mean} tells that, if two return series $i$
-and $j$ are not correlated, auto-correlation in the returns does not
-introduce a bias into the estimation of the cross-correlation; if,
-however, the return series are indeed correlated, auto-correlation
-in the returns rescales the cross-correlation through a multiplicative
-factor $1/(1 - \phi_1^2)$.
+where $P^\mu_\nu(\cdot)$ is Ferrer's function of the first
+kind. See appendix \ref{chp:gaussian_elements_dist}. For the
+definition of Ferrer's function, see equation
+\ref{eq:Ferrers_1st}.
 
-In addition, equation \ref{eq:gaussian_variance} tells that
-auto-correlation in the returns always makes the cross-correlation
-estimation more noisy. Auto-correlation not only rescales the variance
-of the no-autocorrelation estimation by $1/(1 - \phi_1^2)$ but even
-adds an extra term ${2 \sigma^6 \phi_1^2 \over T (1 - \phi_1^2)^3}$.
+Equation \ref{eq:gaussian_mean2} tells that, if
+two return series $i$ and $j$ are not correlated, i.e. $\rho = 0$,
+auto-correlation in the returns does not introduce a bias into the
+estimation of their covariance, i.e. $\mu'_X$, since the difference
+between the two Ferrer's functions evaluate to 0 in equation
+\ref{eq:gaussian_mean2}; if, however, the return series are indeed
+correlated, auto-correlation in the returns rescales the
+covariance through a multiplicative factor $1/(1 - \phi^2)$.
 
-% What we have obtained is a quantitative description of what has been
-% said before: The mean of the non-diagonal entries of the
-% cross-correlation matrix is not affected by autocorrelations in the
-% returns, given that innovations $a_{i,t}$ of different assets and
-% different times are un-correlated; the variance, however, is always
-% increased by autocorrelations no matter the autocorrelation is
-% positive or nagative. The increment is in the form of an additive term
-% that scales as $\phi_1^2 / (1 - \phi_1^2)^3$.
+In addition, equation \ref{eq:gaussian_variance2} tells that
+auto-correlation in the returns always makes the covariance
+estimation more noisy --- auto-correlation not only rescales the
+variance of the no-autocorrelation estimation by $1/(1 - \phi^2)^2$
+but even adds an extra term ${2 \sigma^6 \phi^2 \over T (1 -
+  \phi^2)^3}$.
 
-We may apply a similar treatment to the diagonal elements of the
-covariance matrix, which we denote as $C_{ii}$ here:
-\begin{eqnarray*}
-  C_{ii} &=& {1 \over T} \sum_{t=1}^T r^2_{it} \\
-  &=& {1 \over T} \sum_{t=1}^T \sum_{l=0}^{t-1}\phi_1^l a_{i, t-l}
-  \sum_{k=0}^{t-1}\phi_1^k a_{i, t-k} \\
-  &=& {1 \over T} \sum_{t=1}^T \left[
-    \sum_{k=0}^{t-1} \phi_1^{2k} a^2_{i, t-k} +
-    \sum_{k,l = 0}^{t-1} \phi_1^{k+l} a_{i, t-k} a_{i, t-l}
-  \right]
-\end{eqnarray*}
-By the Lyapunov central limit theorem under the assumption of large T,
-the asymptotic distribution of $C_{ii}$ is Gaussian, the mean and
-variance being
+For the diagonal elements of the covariance matrix $C$ we have
 \begin{eqnarray}
   \E(C_{ii}) &=& {1 \over T}\left[
-    \sum_{k=0}^{t-1} \phi_1^{2k} \sigma^2
+    \sum_{k=0}^{t-1} \phi^{2k} \sigma^2
   \right] \nonumber \\
-  &=& {\sigma^2 \over (1 - \phi_1^2) T} \left[
-    T - {\phi_1^2 (1 - \phi_1^{2T}) \over 1 - \phi_1^2}
+  &=& {\sigma^2 \over (1 - \phi^2) T} \left[
+    T - {\phi^2 (1 - \phi^{2T}) \over 1 - \phi^2}
   \right] \nonumber \\
-  & \approx & {\sigma^2 \over 1 - \phi_1^2} \left[
-    1 - {\phi_1^2 \over T}
-  \right] \label{eq:gaussian_cii_mean}
+  & \approx & {\sigma^2 \over 1 - \phi^2} \left[
+    1 - {\phi^2 \over T (1 - \phi^2)}
+  \right] \label{eq:gaussian_cii_mean2}
 \end{eqnarray}
@@
   \var(C_{ii}) &=& \sum_{t=1}^T \left[
-    \sum_{k=0}^{t-1} {\phi_1^{4k} \sigma^4 \over T^2} 2 +
-    \sum_{k,l=0}^{t-1} {\phi_1^{2(k+l)} \over T^2} \sigma^6
+    \sum_{k=0}^{t-1} {\phi^{4k} \sigma^4 \over T^2} 2 +
+    \sum_{k,l=0}^{t-1} {\phi^{2(k+l)} \over T^2} \sigma^6
   \right] \nonumber \\
   &=& \sum_{t=1}^T \left[
-    {2 \sigma^4 \over T^2} {1 - \phi_1^{4t} \over 1 - \phi_1^4} +
+    {2 \sigma^4 \over T^2} {1 - \phi^{4t} \over 1 - \phi^4} +
     {\sigma^6 \over T^2} \left(
-      {1 - \phi_1^{2t} \over 1- \phi_1^2}
+      {1 - \phi^{2t} \over 1- \phi^2}
     \right)^2 \right] \nonumber \\
-  &=& {2 \sigma^4 \over T (1 - \phi_1^4)} -
-  {2 \sigma^4 \phi_1^4 (1 - \phi_1^{4T}) \over T^2(1 - \phi_1^4)^2} +
+  &=& {2 \sigma^4 \over T (1 - \phi^4)} -
+  {2 \sigma^4 \phi^4 (1 - \phi^{4T}) \over T^2(1 - \phi^4)^2} +
   \nonumber \\
-  && {\sigma^6 \over T (1 -\phi_1^2)^2} -
-  {2 \sigma^6 \phi_1^2 (1 - \phi_1^{2T}) \over T^2 (1 - \phi_1^2)^3} +
-  {\sigma^6 \phi_1^4 (1 - \phi_1^{4T}) \over T^2 (1 - \phi_1^2)^2 (1 -
-    \phi_1^4)} \nonumber \\
-  &\approx& {2 \sigma^4 \over T (1 - \phi_1^4)} + {\sigma^6 \over T (1
-    -\phi_1^2)^2} \label{eq:gaussian_cii_variance}
+  && {\sigma^6 \over T (1 -\phi^2)^2} -
+  {2 \sigma^6 \phi^2 (1 - \phi^{2T}) \over T^2 (1 - \phi^2)^3} +
+  {\sigma^6 \phi^4 (1 - \phi^{4T}) \over T^2 (1 - \phi^2)^2 (1 -
+    \phi^4)} \nonumber \\
+  &\approx& {2 \sigma^4 \over T (1 - \phi^4)} + {\sigma^6 \over T (1
+    -\phi^2)^2} \label{eq:gaussian_cii_variance2}
 \end{eqnarray}
-From equation \ref{eq:gaussian_cii_mean} we see that auto-correlation
+
+From equation \ref{eq:gaussian_cii_mean2} we see that auto-correlation
 in the returns increases the variance of the return series; and from
-equation \ref{eq:gaussian_cii_variance} we see that the variance of
+equation \ref{eq:gaussian_cii_variance2} we see that the variance of
 that variance estimation is also increased by
@@
 approximately as $1/T$, similar to the behavior of
-$\var(C_{ij})$. This is to be compared with the case of GARCH returns
-discussed in chapter \ref{chp:CrossCorrelationFat}.
+$\var(C_{ij})$. This is to be compared with the case of \gls{garch}
+returns discussed in chapter \ref{chp:CrossCorrelationFat}.
 
-% As for the diagonal elements of the cross-correlation matrix,
-% i.e. $\var(r_{it})$, from the representation of $r_{it}$ as an
-% inifinite moving average process, one can deduce immediately
-% \begin{equation*}
-%   \text{var}(r_{i,t}) \approx {\sigma^2 \over 1 - \phi^2}
-% \end{equation*}
-% as is mentioned earlier in equation \ref{eq:gaussian_Cii_variance}.
-
-% However, to write $f_A(RMM'R')$ as a function of the eigen
-% values of $RMM'R'$ is very involved and no theoretical results are
-% known.
-
-% In the more general case
-% \begin{eqnarray*}
-%   a_{i,t} &=& \left(
-%     1 - \sum_{k=1}^p \phi_{i,k} B^k
-%   \right) r_{i,t} \\
-%   &=& \Phi_i(B)} r_{i,t
-% \end{eqnarray*}
-% where $B$ denotes the back shift operator acting on subscript $t$, we
-% can write
-% \begin{eqnarray*}
-%   A &=&
-%   \begin{pmatrix}
-%     \Phi_1(B) & &\\
-%     & \ddots & \\
-%     && \Phi_N(B)
-%   \end{pmatrix} R \\
-%   &=& \Phi(B) R
-% \end{eqnarray*}
-
-% Following the same reasoning, $RR' \sim W(\Phi^{-1}(B)
-% \Sigma \Phi'^{-1}(B), T)$. Substituting $\Phi^{-1}(B)
-% \Sigma \Phi'^{-1}(B)$ for $\Sigma$ in the Wishart
-% probability density function, we get
-% \begin{eqnarray*}
-%   && \tr\left\{\left[\Phi^{-1}(B)\Sigma \Phi'^{-1}(B)\right]^{-1}
-%     RR'\right\}\\
-%   &=& \tr \left\{\Phi'(B)\Sigma^{-1} \Phi(B) RR'\right\}\\
-%   &=& \tr\left\{\Sigma^{-1}AA'\right\}\\
-% \end{eqnarray*}
-% As to $\det \Phi^{-1}(B)\Sigma \Phi'^{-1}(B)$, we note that $\Sigma$
-% is constant over time, and thus a function of $B$ has no effect on
-% it. Therefore $\det \Phi^{-1}(B)\Sigma \Phi'^{-1}(B) = \det \Sigma$.
-
-\section{Distribution of the Eigenvalues}\label{sec:GCC-numerical}
-For the Wishart matrix $\mtx{RR'}$, theoretical results are available
+\section{Distribution of the Eigenvalues}
+\label{sec:GCC-numerical}
+For a Wishart matrix $\mtx{RR'}$, theoretical results are available
 for the eigenvalue distribution, and we summarize them in appendix
@@
 However, as detailed in the derivation leading to equation
-\ref{eq:cross-corr-matrix-PDF}, the distribution of $RR'$ is not
-Wishart when the columns of $R$ are correlated
+\ref{eq:cross-corr-matrix-PDF}, the distribution of $\mtx{RR'}$ is not
+Wishart when the columns of $\mtx{R}$ are correlated
 (auto-correlation). Deriving the eigenvalue distribution analytically
@@
 
-As before we consider the AR(1) process:
+As before we consider the \gls{ar}(1) process:
 \begin{equation*}
@@
 \end{equation*}
-where $\vec{a}_t \sim N(0, I)$, i.e. the elements of $\vec{a}_t$ are
+where $\vec{a}_t \sim N(0, \mtx{I})$, i.e. the elements of $\vec{a}_t$ are
 independent Gaussian random variable with zero mean and unit
-variance.
-
-Now we investigate how the eigenvalue distribution depends on
-$\phi$. Figure \ref{fig:GaussianMarkovSpectrumPDF} shows the results
-of the simulation. 
+variance. Now we investigate how the eigenvalue distribution depends on
+the auto-correlation strength parameter $\phi$. Figure
+\ref{fig:GaussianMarkovSpectrumPDF} shows the results of the
+simulation.
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[scale=0.4, clip=true, trim=90 228 115
+    226]{../pics/GaussianMarkovSpectrumPDF.pdf}
+  \end{center}
+  \vspace{-10mm}
+  \caption{\small \it
+      Eigenvalue distribution with correlation time $\tau$ ranging
+      from 0 to 3. The 1st blue line, which is shown as stairs, is the
+      theoretical eigenvalue distribution according to the
+      Marcenko-Pastur law (see eq.\ref{eq:MP_pdf}). In the simulation
+      we have chosen $q = N/T = 50/1000 = 0.05$ and the standard
+      deviation of the returns $\sigma=1$. For each value of $\tau$ we
+      generate 2000 instances of $N \times T$ random matrix $R$, and
+      compute C as $C=RR'/T$. Hence each curve in the figure is
+      constructed from 2000 sets of eigenvalues. The correspondance
+      between the correlation time $\tau$ and the auto-regressive
+      coefficient $\phi$ is $\tau = -\ln 2 / \ln \phi$.
+    }
+  \label{fig:GaussianMarkovSpectrumPDF}
+\end{figure}
 It is clear from the figure that the maximum eigenvalue moves
@@
 also increases with $\phi$.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4, clip=true, trim=100 226 116
+  133]{../pics/Gaussian_mineig.pdf}
+  \caption{\small \it The minimum eigenvalue versus
+    auto-correlation strength $\phi$. For each value of $\phi$ 2000
+    random matrices are generated and their eigenvalues are
+    calculated. The minimum eigenvalue of each random matrix is noted
+    and the mean of the 2000 such minimum eigenvalues are plotted
+    against the chosen value of $\phi$. 20 values of $\phi$ are
+    included in the plot, ranging from 0 to 0.95 with step size 0.05.}
+  \label{fig:Gaussian_mineig}
+\end{figure}
 
-% In the following we focus on the maximum eigenvalue, which is shown by
-% Chiani to have approximately a gamma distribution \cite{Chiani2012}.
-% So we first verify Chiani's result, and then extend it to the
-% case {\it with} autocorrelations and study how the
-% distribution changes as $\phi$ increases.
-Chiani showed that the maximum eigenvalue ($\lambda_1$) had
-approximately a gamma distribution \cite{Chiani2012}. So we compare in
-figure \ref{fig:GaussianMarkov005MaxEigCDF_loglog} the empirical
+Chiani showed that the marginal distribution of the maximum eigenvalue
+($\lambda_1$) is approximately gamma when neither cross-correlation
+nor auto-correlation is present\cite{Chiani2012}. So we compare
+in figure \ref{fig:GaussianMarkov005MaxEigCDF_loglog} the empirical
 cumulative distribution function (CDF) of the maximum eigenvalue 
@@
 auto-correlations are present ($\phi > 0$) have also been included.
-
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.5, clip=true, trim=0 238 0
+    197]{../pics/GaussianMarkov005MaxEigCDF_loglog.pdf}
+  \end{center}
+  \caption{\small \it Cummulative distribution function (CDF) of the
+    maximum eigenvalue ($\lambda_1$) from numerical simulations (blue
+    lines) are compared to the fitted gamma distribution (red
+    lines). Each pair of CDFs correspond to a fixed autocorrelation
+    strength ($\phi$). The parameters $k$ and $\theta$ of the gamma
+    distribution are fit to data by matching the 2nd and the 3rd
+    moments of the gamma distribution to the corresponding moments of
+    the empirical distribution. Then the parameter $\alpha$ in
+    equation \ref{eq:TracyWidom-Gamma} is chosen to be $\alpha =
+    k\theta - \E\left({\lambda_1 - \mu_{NT} \over
+        \sigma_{NT}}\right)$. The curves are plotted on log-log
+    scale.}
+  \label{fig:GaussianMarkov005MaxEigCDF_loglog}
+  %\vspace{-10mm}
+\end{figure}
 It is seen in the figure that gamma distributions with different
@@
 
-Since the maximum eigenvalue distribution is characterized by the
-parameters $k$, $\theta$, and $\alpha$, the influence of the
-autocorrelations can be characterized by the dependence of $k$,
-$\theta$, and  $\alpha$ on $\phi$. While these dependences are rather
-intricate,
-% Figure \ref{fig:param1_phi_dep} shows how the mean, variance,
-% and skewness of the fitted gamma distribution, as well as the
-% relocation parameter $\alpha$ denpend on $\phi$.
-% \begin{figure}[htb!]
-%   \centering
-%     \includegraphics[scale=0.5, clip=true, trim=37 240 50
-%     202]{../pics/param1_phi_dep.pdf}
-%   \caption{\small \it  the  mean, variance, and skewness of the fitted
-%     gamma distribution as well as the relocation parameter $\alpha$
-%     are plotted against the autocorrelation strength $\phi$. 20 values
-%     of $\phi$ are included in the plot, ranging from 0 to 0.95 with
-%     step size 0.05.}
-%   \label{fig:param1_phi_dep}
-% \end{figure}
-% Despite the monotonic behavior shown by 3 of these curves, the
-% relation between each of the aforementioned quantities and $\phi$ is
-% more complicated than it apears. However, the mean, i.e. $k \theta$ is
-% an exception: 
-good support can be found in the data for the following
-approximate relation:
+Since the maximum eigenvalue distribution is approximated by a gamma
+distribution characterized by parameters $k$, $\theta$, and $\alpha$
+\footnotemark, the influence of the autocorrelations can be
+characterized by the dependence of $k$, $\theta$, and  $\alpha$ on
+$\phi$. While these dependences are rather intricate, good support can
+be found in the data for the following approximate relation:
+\footnotetext{
+  The mean, variance and skewness of the gamma distribution are given
+  by
+  \begin{eqnarray*}
+    \text{mean} &=& k\theta \\
+    \text{variance} &=& k\theta^2 \\
+    \text{skewness} &=& 2/\sqrt{k}
+  \end{eqnarray*}
+}
 \begin{eqnarray}
@@
 \end{eqnarray}
+Here we note that $k\theta$ is the mean of the gamma distribution.
 To verify this relation, we first fit a 2nd order polynomial and
@@
 good approximation, a close match between $\tan{\pi \phi'_n \over 2}$
-and $\tan{\pi \phi_n \over 2}$ is expected. Figure
-\ref{fig:phi_and_roots} plots $\tan{\pi \phi'_n \over 2}$ against
-$\tan{\pi \phi_n \over 2}$ and fits a straight line to the data
-points. It is seen that the data points lie fairly close to the fitted
-line and the line has a slope very close to 1 and an intercept close to
-0. This strongly suggests $\tan{\pi \phi'_n \over 2} = \tan{\pi \phi_n
-  \over 2}$ and supports the relation \ref{eq:k_theta-phi}.
+and $\tan{\pi \phi_n \over 2}$ is expected. From figure 
+\ref{fig:phi_and_roots} one can see this is indeed the case.
+
 \begin{figure}[htb!]

### 2014-06-11 thesis/price_models.tex
Move matlab files to a separate folder.

@@
 properties. In the following we first describe these models briefly,
-then in
-
-% \S\ref{sec:Garch_model} reviews the GARCH model using the Nordea
-% 15-minute returns as an example. \S\ref{sec:SLV_model} examines the
+then in section \ref{chp:nordea_15min}, section \ref{sec:volvo_30} and appendix
+\ref{chp:appendix2} we describe how the \gls{garch} and the \gls{sv} models
+are fitted to intraday returns and compare their forecasts. In section
+\ref{sec:XieCalc} we calculate the unconditional distribution
+functions of the \gls{sv} model.
+
+% \gls{frt} is good.
+% section \ref{sec:Garch_model} reviews the \gls{garch} model using the Nordea
+% 15-minute returns as an example. Section \ref{sec:SLV_model} examines the
 % stochastic log-volatility model and discusses the possibility of
@@
 % Johnson Su distribution. The same Nordea return series is fitted as an
-% example. \S\ref{sec:XieCalc} derives the unconditional distribution
+% example. Section \ref{sec:XieCalc} derives the unconditional distribution
 % function of the stochastic log-volotility model and compares the
@@
   dS &=& S\mu dt + S\sigma dw_t \\
+\end{eqnarray*}
+Then by It\^o's lemma, a stochastic differential equation for the
+logarithmic price $\ln S$ can be obtained
+\begin{eqnarray*}
   d(\ln S) &=& (\mu - \frac{1}{2} \sigma^2)dt + \sigma dw_t \\
@@
 is the price at time 0, has Gaussian distribution with mean $(\mu -
-\sigma^2/2)t$ and variance $\sigma^2 t$. Therefore, in a discrete-time
-model, where the length of each time step $\Delta t$ is fixed, the
-return $r_t = \ln S(t + \Delta t) - \ln S(t)$ is assumed to be
-Gaussian distributed and have mean and variance that are functions of
-$\Delta t$.
+\sigma^2/2)t$ and variance $\sigma^2 t$ \cite{Bernt2000}. Therefore,
+in a discrete-time model, where the length of each time step $\Delta
+t$ is fixed, the return $r_t = \ln S(t) - \ln S(t - \Delta t)$ is
+assumed to be Gaussian distributed and have mean and variance that are
+functions of $\Delta t$.
 
@@
 inaccurate in describing very large returns, the Gaussian distribution
-as a returns model shall not be forgotten. In the next few sections,
-we discuss some more realistic returns models.
-
-\item{\bf GARCH models}
+as a return model shall not be forgotten. In the next few sections,
+we discuss some more realistic return models.
+
+\item{\bf \gls{garch} models}
 
 ``GARCH'' is the acronym for ``Generalized Autoregressive Conditional
-Heteroscedasticity''. A GARCH(p, q) model is defined by the following
+Heteroscedasticity''. A \gls{garch}(p, q) model is defined by the following
 equation system \cite{Bollerslev86}:
@@
 \end{equation}
-where $p$ and $q$ are constant integers; $\mu(t)$ is the mean process
-and considered a small constant for intraday returns. $\epsilon_t$ is
-called the innovation of the return and is a variate whose conditional
-distribution \footnotemark is Gaussian and has variance $\sigma^2_t$:
+where $r_t$ is the return as mentioned earlier; $p$ and $q$ are
+constant integers; $\mu_t$ is the mean process and considered a small
+constant for intraday returns. $\epsilon_t$ is called the innovation
+of the return and is a variate whose conditional distribution
+\footnotemark is Gaussian and has variance $\sigma^2_t$:
 \begin{eqnarray*}
@@
 
-It can be shown that the autocorrelation function of $\epsilon_t^2$ is
-given by \cite{Bollerslev87}:
-\begin{eqnarray*}
-\text{corr}(\epsilon_t^2, \epsilon^2_{t-n}) &=& \sum_{i=1} ^{p \vee q}
-(\alpha_i + \beta_i) \rho_{n-i} \text{ for $n > p$}
+It can be shown that the autocorrelation function $\varrho_n =
+\text{corr}(\epsilon_t, \epsilon_{t-n})$ for $n > \text{max}(p,q)$ of
+$\epsilon_t^2$ is given by \cite{Bollerslev87}:
+\begin{eqnarray*}
+\varrho_n &=& \sum_{i=1} ^{p \vee q}
+(\alpha_i + \beta_i) \varrho_{n-i} \text{ for $n > p$}
 \end{eqnarray*}
@@
 
+Here we note that, in the \gls{garch} model, the volatility $\sigma_t$ is
+$\mathcal{F}_{t-1}$ measurable: given the history up to time $t-1$,
+$\sigma_t$ is a deterministic quantity. In contrast, \gls{sv} models,
+which we discuss shortly, treat $\sigma_t$ as a random variable even
+given the same history.
+
 \item {\bf Stochastic Volatility Models}\label{sec:SLV_model}
 
-For the purpose of intraday returns, we specify the stochastic
-volatility model as
+For the purpose of intraday returns, we specify the \gls{sv} model as
 \begin{eqnarray}
@@
 
-Andersen and Bollerslev et al proved the following in
-\cite{Andersen03} (theorem 2):
+Andersen et al proved the following in \cite{Andersen03} (theorem 2):
 \begin{equation}
@@
 \end{equation}
-In plain words, given all the information up to time $t-h$, the
+In plain words, conditional on the information up to time $t-h$, the
 distribution of $r_t$ is Gaussian with integrated mean and variance.
@@
   \int_{0}^h \sigma_{t-h+s}^2 ds = \sum_{k=1}^n \left(
-    S_{t-h+kh/n} - S_{t-h+(k-1)h/n} \right)^2
+    \ln S_{t-h+kh/n} - \ln S_{t-h+(k-1)h/n} \right)^2
 \end{equation}
@@
 
+Once a good approximation of $\sigma_t$ has been found, it is
+convenient to model $\ln\sigma_t$ so that the positivity of $\sigma_t$
+is implied by construction \cite{Mikosch2009}. As is seen in later
+sections, \gls{arima} models often serve well for this purpose. An
+\gls{arima}(p, d, q) model, where $p,d,q$ are integers, is defined as
+\cite{BoxJenkins94}
+\begin{eqnarray*}
+  (1 - \sum_{i=1}^p \phi_i B^i) (1 - B)^d \ln \sigma_t &=& (1 -
+  \sum_{i=1}^q \theta_i B^i) y_t
+\end{eqnarray*}
+where $B$ is the back-shift operator such that $B x_t = x_{t-1}$ for
+any time series $x_t$. $y_t$ are termed the residuals of the
+model, $d$ is the order of integration, and $p, q$  are orders of
+autoregression and moving average, respectively. $\phi_i$ and
+$\theta_i$ are constant parameters.
+
+Quite often, the auto-correlation function of the time series in
+question manifests periodic patterns, thus seasonal components are
+added to the above model to account for the seasonality:
+\begin{eqnarray*}
+&& (1 - \sum_{i=1}^P \Phi_i B^{is}) (1 - \sum_{i=1}^p \phi_i B^i) (1 - B^s)^D (1 -
+B)^d\ln \sigma_t \\
+&& = (1 - \sum_{i=1}^Q \Theta_i B^{is}) (1 -
+\sum_{i=1}^q \theta_i B^i)y_t
+\end{eqnarray*}
+where $D$, analogous to $d$ in the non-seasonal model, is the order of
+seasonal integration; $P, Q$ are orders of seasonal autoregression and
+moving average, respectively. $\Phi_i$ and $\Theta_i$ are constant
+parameters. In the most general situations, the seasonal and the
+non-seasonal components do not necessarily combine in the above
+multiplicative fashion, so the following model is also of interest:
+\begin{eqnarray*}
+(1 - \sum_{i=1}^{p+P} \phi_i B^i) (1 - B^s)^D(1 - B)^d \ln
+\sigma_t &=& (1 - \sum_{i=1}^{q+Q} \theta_i B^i) y_t
+\end{eqnarray*}
+In all the above cases, if $d = 0$ and $D = 0$, the model does not
+involve integration and hence is called an \gls{arma}(p,q) model. An
+even simpler case is where $q = 0$ and $Q = 0$ in addition to $d = 0$
+and $D = 0$. These conditions make the model a pure \gls{ar}
+process, denoted AR(p).
 \end{enumerate}
 
-To compare GARCH and stochastic log-volatility models at the face of
-intraday returns, we study the {\it Nordea Bank} 15-minute returns
-during the period 2012/01/16 - 2012/04/20 in \S\ref{chp:nordea_15min},
+To compare \gls{garch} and \gls{sv} models at the face of intraday returns,
+we study the {\it Nordea Bank} 15-minute returns during the period
+2012/01/16 - 2012/04/20 in section \ref{chp:nordea_15min}, 
 and {\it Volvo B} 30-minute returns during 2013/10/10 - 2014/04/04 in
-\S\ref{sec:volvo}. Another 5 intraday series are also studied and
-collected in appendix \ref{chp:appendix2}.
+section \ref{sec:volvo_30}. Another 4 intraday series are also studied and
+collected in appendix \ref{chp:appendix2}. We have chosen these
+particular stocks because they have the largest trading volumes in the
+Swedish market and hence provide the largest amount of data for
+analysis. The time intervals of 15 and 30 minutes are chosen because
+they are relatively short and hence give a large amount of data and
+yet they are not too short for noise of market
+micro-structure\footnotemark to become a concern.
+\footnotetext{
+Noise of market micro-structure is due to, most importantly,
+the difference between the bid and asked prices (bid-ask spread), and
+the discreteness of price changes.
+}
 
@@
 remaining 20\% (405) for comparing with model forcasts. In section
-\ref{sec:nordea_15min_garch} we study the series with a GARCH model 
+\ref{sec:nordea_15min_garch} we study the series with a \gls{garch} model 
 and in section \ref{sec:nordea_15min_arima} we study it with a
-stochastic volatility model.
+\gls{sv} model.
 
 \subsection{GARCH Model}\label{sec:nordea_15min_garch}
-First of all, to determine that the volatilities are auto-correlated
-and/or that the squared returns and the volatilities are correlated, and
-hence a GARCH(p, q) model is appropriate for the return series under
-investigation, we plot the auto-correlation function (ACF) of the
-squared returns. This is shown in figure
-\ref{fig:nordea_15min_acf}. If the aforementioned features are absent
-from the series, the auto-correlations are expected to be Gaussian
-distributed with mean 0 and variance 1/T and hence mostly reside
-within the confidence bounds set by $\pm 2/\sqrt{T}$
-\cite{Bollerslev86, Bollerslev87}.
+When volatilities are auto-correlated or squared returns and
+volatilities are correlated, a \gls{garch}(p, q) model may be
+appropriate for the return series under investigation. To find out
+whehter this is true in our case, we plot the \gls{acf} of the squared
+returns. This is shown in figure \ref{fig:nordea_15min_acf}. If the
+aforementioned features are absent from the series, the
+auto-correlations are expected to be Gaussian distributed with mean 0
+and variance 1/T, and hence mostly reside within the confidence bounds
+set by $\pm 2/\sqrt{T}$ \cite{Bollerslev86, Bollerslev87}.
 
@@
 conditional variances of the series are correlated. These observations
-suggest a GARCH(p, q) model can be appropriate.
+suggest a \gls{garch}(p, q) model can be appropriate.
 \begin{figure}[htb!]
@@
     ($\hat{\sigma}^2_t$). The blue lines are confident bounds set at
-    $\pm 2/\sqrt{T}$.}
-\end{figure}
-
-Starting with a GARCH(1,1) model and taking advantage of the knowledge
-that the log-volatility $\ln \sigma_t$ has seasonality $s=33$, we fit
-to the return series a GARCH(33, 33) model, limiting to lags 1 and 33
-for both ARCH and GARCH parameters.
+    $\pm 2/\sqrt{T}$. $T$ is the length of the time series.}
+\end{figure}
+
+Starting with a \gls{garch}(1,1) model and taking advantage of the knowledge
+that the log-volatility $\ln \sigma_t$ has seasonality $s=33$ (see
+figure \ref{fig:nordea_15min_logvol_acf}), we fit to the return series
+a \gls{garch}(33, 33) model, limiting to lags 1 and 33 for both ARCH and
+\gls{garch} parameters.
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-Via maximum likelihood estimation, parameter values listed in table
-\ref{tab:nordea_15min_garch} are obtained.
+where the mean process of $r_t$, denoted $\mu_t$ earlier, is simplified
+to a mere constant $\mu$ owing to its smallness. by \gls{mle},
+parameter values listed in table \ref{tab:nordea_15min_garch} are obtained.
 \begin{table}[htb!]
@@
     220]{../pics/nordea_bank_15min_z_prob.pdf}
-  \caption{\small \it{Probability plot of $z_t =
-      (\epsilon_t-\E(\epsilon_t))/\sigma_t$. $\epsilon_t$ are derived
+  \caption{\small \it Probability plot of $z_t =
+    (r_t-\E(r_t))/\sigma_t$. $\epsilon_t$ are derived 
       from Nordea Bank 15min returns while $\sigma_t$ are realized
-      volatilities calculated
-      using 30s returns within each 15min interval. Horizontal axis:
-      $z_t$}; Vertical axis: cummulative probability function (CDF) of
+      volatilities calculated using 30s returns within each 15min
+      interval. Horizontal axis: $z_t$; Vertical axis: CDF of
     $z_t$, arranged on such a scale that the CDF of the standard
@@
 rates between Deutch mark, yen and dollar, $\ln \sigma_t$ is
-gaussian distributed \cite{Andersen03}. This is, however, not the case
-for our series in question. In fact, in our case, $\ln \sigma_t$ is
-right skewed (skewness 0.3342) and leptokurtic (kurtosis 
-6.1006). See figure \ref{fig:nordea_15min_logvol_prob}
+Gaussian distributed \cite{Andersen03}. This is, however, not the case
+for our modestly sized series. In fact, in our case, $\ln \sigma_t$ is
+right skewed (skewness 0.3342) and leptokurtic (kurtosis 6.1006). See
+figure \ref{fig:nordea_15min_logvol_prob}
 \begin{figure}[htb!]
@@
   4]{../pics/nordea_15min_logvol_prob.pdf}
-  \caption{\small \it Probability plot of Nordea 15min
+  \caption{\small \it Probability plot of Nordea 15min log-volatility
     $\ln\sigma_t$ unconditional distribution}
@@
 \ref{fig:nordea_15min_logvol_acf}). This suggests the series may be
-described by a seasonal ARIMA model\footnote{ARIMA ---
-  Integrated Autoregressive Moving Average} model. This we first
-simplify the series by differencing \cite{BoxJenkins94}:
-\[
+described by a seasonal \gls{arima} model. Thus we first simplify the
+series by differencing \cite{BoxJenkins94}:
+\begin{equation}
+  \label{eq:differenced_lv}
   w_t = (1-B)(1-B^s)\ln\sigma_t  
-\]
-where $B$ is the back-shift operator and $s=33$ is the seasonality.
+\end{equation}
+where $B$ is the back-shift operator\footnote{For example, $B\,x_t =
+  x_{t-1}$} and $s=33$ is the seasonality.
 
@@
   \centering
-  \subfigure[ACF of $\ln\sigma_t$]{
+  \subfigure[ACF of log-volatility ($\ln\sigma_t$)]{
     \includegraphics[scale=0.4, clip=true, trim=95 230 112
@@
   }
-  \subfigure[ACF of $w_t$]{
+  \subfigure[ACF of differenced log-volatility ($w_t$)]{
     \includegraphics[scale=0.4, clip=true, trim=95 230 112
@@
   }
-  \caption{\small \it Nordea 15min $\ln\sigma_t$ and $w_t$
-    autocorrelations}
+  \caption{\small \it Auto-correlations of Nordea 15min log-volatility
+    ($\ln\sigma_t$) and differenced log-volatility ($w_t$).}
   \label{fig:nordea1_15min_acf}
@@
   \end{tabular}
-  \caption{\small \it Nordea 15min $w_t$ autocorrelations}
+  \caption{\small \it autocorrelations of differenced log-volatility
+    ($w_t$)}
   \label{tab:nordea_15min_w_acf}
@@
   123]{../pics/nordea2_y_normplot.pdf}
-  \caption{\small \it Nordea 15min $y_t$ normal probability plot.}
+  \caption{\small \it Normal probability plot of residuals of
+    log-volatility ($y_t$)}
   \label{fig:nordea_15min_y_qq}
@@
   \end{tabular}
-  \caption{\small \it Nordea 15min $y_t$ moments}
+  \caption{\small \it Moments of log-volatility residuals ($y_t$)}
   \label{tab:nordea_15min_y_moments}
@@
     133]{../pics/nordea_15min_y_js_fit.pdf}
-    \caption{\small \it Nordea 15min residuals $y_t$ fitted to a
+    \caption{\small \it Log-volatility residuals $y_t$ fitted to a
       Johnson Su distribution. Horizontal: values of $y_t$, denoted x;
@@
 Under the assumption of i.i.d Johnson Su distributed residuals, the
-log-likelihood function of the parameters $\gamma, \delta$
+log-likelihood function of the parameters $\theta, \Theta$
 conditional on the sample $w_t$ can be written as
@@
 \subsection{Comparison of the Forecasts}
-In this section we compare the one-step-ahead forecasts from the GARCH
-model and from the stochastic volatility model. For this purpose, we
+In this section we compare the one-step-ahead forecasts from the \gls{garch}
+model and from the \gls{sv} model. For this purpose, we
 compute the difference between a forecast $\ln \sigma^F_t$ and its measured
 counterpart, i.e. the realized volatility of the same period $\ln
-\hat{\sigma}_t$.
+\hat{\sigma}_t$. As a reference, we also consider the results obtained
+by taking the mean of the realized volatilities of the first 80\% of
+the data set as forecast for the volatilities of the remaining 20\%. We
+call this naive forecast the ``sample mean''.
 
@@
     \hline
-    & SV & GARCH & Sample mean \\
+    & \gls{sv} & \gls{garch} & Sample mean \\
     \hline
@@
   \caption{\small \it Mean and standard Deviation of the forecasts'
-    distribution. ``Sample mean`` refers to the mean of the realized
-    volatilities in the first 80\% of the data set. ``SV'': Stochastic
-    volatility.}
+    distribution}
   \label{tab:nordea_2012}
 \end{table}
-It is seen from table \ref{tab:nordea_2012} that, on average, the SV
-model over-estimates while GARCH under-estimates. In terms of the
-standard deviation of $\ln \sigma^F_t - \ln \hat{\sigma}_t$, the SV
+It is seen from table \ref{tab:nordea_2012} that, on average, the \gls{sv}
+model over-estimates while \gls{garch} under-estimates. In terms of the
+standard deviation of $\ln \sigma^F_t - \ln \hat{\sigma}_t$, the \gls{sv}
 model wins with a small margin. In contrast, the sample mean forecast
@@
 function of $\ln \sigma^F_t - \ln \hat{\sigma}_t$. Here one can see
-that the SV model yields a better quality of forecasts than does GARCH 
-with regard to the under-estimates, and essentially the same quality
-with regard to the over-estimates.
-
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.55, clip=true, trim=27 287 0
-    154]{../pics/nordea_2012.pdf}
+that the \gls{sv} model yields a better quality of forecasts than does \gls{garch} 
+with respect to both under-estimates and over-estimates.
+
+\begin{figure}[htb!]
+  \centering
+    \includegraphics[scale=0.55, clip=true, trim=40 296 16
+    260]{../pics/nordea_2012.pdf}
   \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
@@
 forecasts, where the criterion of ``good'' is defined, respectively,
-as deviating no more than 1\%, 5\%, or 10\% from the measured value of
-the realized volatility. Table \ref{tab:nordea_2012_good} shows the
+as the forecast lying within 1\%, 5\%, or 10\% of the corresponding
+realized volatility. Table \ref{tab:nordea_2012_good} shows the
 respective percentage of the 3 kinds of forecasts. Again in this table
-it is seen that the SV model gives more accurate forecasts than does
-GARCH.
+it is seen that the \gls{sv} model gives more accurate forecasts than does
+\gls{garch}. For example, defining a ``good'' forecast as one that lies
+within 5\% of the realized volatility, the probability of obtaining
+such a good forecast is 75\% using the \gls{sv} model, 71\% using the
+\gls{garch} model, and only 53\% using the sample mean.
 \begin{table}[htb!]
@@
     \hline
-    ${\ln \sigma^F_t - \ln \hat{\sigma}_t \over |\ln \hat{\sigma}_t|}$
-    & SV & GARCH & sample mean \\
-    \hline
-    1\% & 21.73\% & 13.83\% & 11.36\% \\
-    \hline
-    5\% & 74.57\% & 70.62\% & 53.33\% \\
-    \hline
-    10\% & 97.28\% & 93.83\% & 89.14\% \\
+    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln \hat{\sigma}_t|}$
+    & \gls{sv} & \gls{garch} & sample mean \\
+    \hline
+    1\% & 22\% & 14\% & 11\% \\
+    \hline
+    5\% & 75\% & 71\% & 53\% \\
+    \hline
+    10\% & 97\% & 94\% & 89\% \\
     \hline
   \end{tabular}
+  % \begin{tabular}{|c|c|c|c|}
+  %   \hline
+  %   ${\ln \sigma^F_t - \ln \hat{\sigma}_t \over |\ln \hat{\sigma}_t|}$
+  %   & \gls{sv} & GARCH & sample mean \\
+  %   \hline
+  %   1\% & 21.73\% & 13.83\% & 11.36\% \\
+  %   \hline
+  %   5\% & 74.57\% & 70.62\% & 53.33\% \\
+  %   \hline
+  %   10\% & 97.28\% & 93.83\% & 89.14\% \\
+  %   \hline
+  % \end{tabular}
   \caption{\small \it The percentage of ``good'' forecasts when the
@@
   151]{../pics/volvo_inno_and_lv_acf.pdf}
-    \label{fig:volvo_inno_and_lv_acf}
   \caption{\small \it Left: probability plot of $(r_t -
     \E(r_t))/\sigma_t$; Right: auto-correlations of $\ln \sigma_t$}
+  \label{fig:volvo_inno_and_lv_acf}
 \end{figure}
@@
 series $\ln \sigma^{\text{SV}}_t$. To access the quality of the
-forecast, we also estimate a GARCH model using the returns. The result
-is a GARCH(1, 1) model, whose parameter values are listed in table
+forecast, we also estimate a \gls{garch} model using the returns. The result
+is a \gls{garch}(1, 1) model, whose parameter values are listed in table
 \ref{tab:volvo_garch}.
@@
 
-The forecasts from SV, GARCH, and the sample mean are compared
+The forecasts from \gls{sv}, \gls{garch}, and the sample mean are compared
 using the difference $\ln \sigma^F_t - \ln\hat{\sigma}_t$, where $\ln
@@
   243]{../pics/volvo_slv_garch_cmp.pdf}
-    \label{fig:volvo_slv_garch_cmp}
   \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
@@
     \hat{\sigma}_t > x)$. Horizontal: $x$.}
+  \label{fig:volvo_slv_garch_cmp}
 \end{figure}
 Figure \ref{fig:volvo_slv_garch_cmp} shows, as in the previous case of
-Nordea Bank 15-minute returns, the SV model performs the best, GARCH
+Nordea Bank 15-minute returns, the \gls{sv} model performs the best, \gls{garch}
 the second, and the sample mean the worst. However, when it comes to
 over-estimates, the sample mean appears to be the best estimator,
-while SV excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
+while \gls{sv} excels over \gls{garch}. But a check of $\E(\ln\sigma_t)$ over
 the sample for estimation (1507 data points) and over the sample for
@@
       \hat{\sigma}_t|}$ &
-    SV & GARCH & Sample Mean \\
-    \hline
-    1\% & 21.75\% & 12.20\% & 14.06\% \\
-    \hline
-    5\% & 71.62\% & 61.54\% & 66.31\% \\
-    \hline
-    10\% & 92.31\% & 88.06\% & 89.92\% \\
+    \gls{sv} & \gls{garch} & Sample Mean \\
+    \hline
+    1\% & 22\% & 12\% & 14\% \\
+    \hline
+    5\% & 72\% & 62\% & 66\% \\
+    \hline
+    10\% & 92\% & 88\% & 90\% \\
     \hline
   \end{tabular}
+  % \begin{tabular}{|c|c|c|c|}
+  %   \hline
+  %   ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
+  %     \hat{\sigma}_t|}$ &
+  %   \gls{sv} & \gls{garch} & Sample Mean \\
+  %   \hline
+  %   1\% & 21.75\% & 12.20\% & 14.06\% \\
+  %   \hline
+  %   5\% & 71.62\% & 61.54\% & 66.31\% \\
+  %   \hline
+  %   10\% & 92.31\% & 88.06\% & 89.92\% \\
+  %   \hline
+  % \end{tabular}
   \caption{\small \it Fraction of ``good'' forecasts as defined by
@@
 \end{table}
-We see from the table that the SV model consistently excels over
+We see from the table that the \gls{sv} model consistently excels over
 the other two alternatives. In addition, it is also noted that the
-GARCH forecast is even worse than the sample mean. This is surprising
-but not totally unexpected --- with only 3 parameters, the GARCH(1,1)
+\gls{garch} forecast is even worse than the sample mean. This is surprising
+but not totally unexpected --- with only 3 parameters, the \gls{garch}(1,1)
 model can only describe the most prominent auto-correlations in the
 volatility. When the volatility is influenced by relatively weak
-auto-correlations at several different time lags, the GARCH forecast
+auto-correlations at several different time lags, the \gls{garch} forecast
 cannot be expected to have good accuracy.
 
-In this particular case, we see that the ARIMA model has 3 relatively
+In this particular case, we see that the \gls{arima} model has 3 relatively
 small moving average coefficients, located at lags 2, 3, and 4 and
 evaluated to 0.06, 0.06, 0.03, suggesting a scattered auto-correlation
-structure, so the GARCH model cannot be expected to perform very
+structure, so the \gls{garch} model cannot be expected to perform very
 well. In contrast, the Volvo 15-minute returns studied in
-\S\ref{sec:volvo} has more concentrated auto-correlations --- 0.12 and
-0.06 at lags 2 and 3 --- thus the GARCH(1,1) model is also found to
-perform better and even marginally better than the SV model for the
+section \ref{sec:volvo} has more concentrated auto-correlations --- 0.12 and
+0.06 at lags 2 and 3 --- thus the \gls{garch}(1,1) model is also found to
+perform better and even marginally better than the \gls{sv} model for the
 same series.
@@
 In this section we study the unconditional distribution function of
-the Stochastic Log-Volatility model specified as equation
-\ref{eq:SLV_spec}. As is discussed in \S\ref{sec:SLV_model},
-$\ln\sigma_t$ can be described by an ARMA or ARIMA \footnote{ARIMA ---
-  Integrated Autoregressive Moving Average} model, possibly with
-seasonal components. Here we note that all these models can be
-re-written as a moving average model, which is infinite in extent if
+the \gls{sv} model specified as equation
+\ref{eq:SLV_spec}. As is discussed at the beginning of this chapter,
+$\ln\sigma_t$ can be described by an \gls{arma} or \gls{arima} model,
+possibly with seasonal components. Here we note that all these models
+can be re-written as a moving average model, which is infinite in extent if
 autoregressive components are present:
@@
 that $y_t$ for all $t$ have finite second moment --- this is what we
-assume in the rest of this section. It follows from the above equation
-that the unconditional distribution of $\ln \sigma_t$ is the same as the
-distribution of $\bar{v} + v$ where $v \sim N(0, \sigma)$ and
-$\bar{v}$, $\sigma$ are constants. Now we can state that the
-unconditional distribution of the returns $r_t$
+assume in the rest of this section. It follows that the unconditional
+distribution of $\ln \sigma_t$ is the same as the distribution of
+$\bar{v} + v$ where $v \sim N(0, \sigma)$ and $\bar{v}$, $\sigma$ are
+constants. Now we can state that the unconditional distribution of the
+returns $r_t$
 \begin{eqnarray*}
   r_t &=& \mu + \sigma_t b_t\\
-  &=& \mu + \exp\left(
+  &=& \mu + b_t \exp\left(
     y_t + \sum_{n=1}^\infty c_n y_{t-n} + \text{Const.}
@@
 
-In \S\ref{sec:SLV_Symmetric} we first study the model in the
+In section \ref{sec:SLV_Symmetric} we first study the model in the
 relatively simple case when $v$ and $b$ are uncorrelated and $\mu
 = 0$. If this simplified version proves inadequate, one may resort to
-the general model studied in \S\ref{sec:SLV_Asymmetric}.
+the general model studied in section \ref{sec:SLV_Asymmetric}.
 
 \subsection{The Simplified model}\label{sec:SLV_Symmetric}
-In the following we derive the unconditional PDF of $r'$, denoted
-$f_{r'}(x)$. Then the PDF of $r$ is $e^{-\bar{v}}f_{r'}(e^{-\bar{v}}x)$.
+In the following we derive the unconditional \gls{pdf} of $r'$, the
+de-meaned and rescaled return. Denote this \gls{pdf} $f_{r'}(x)$. Then
+the \gls{pdf} of $r$ is $e^{-\bar{v}}f_{r'}(e^{-\bar{v}}x)$. First we
+consider
 \begin{eqnarray*}
@@
   2014-03-12. The returns are computed using 1-minute mean prices.}
-by means of maximum likelihood estimate using the MATLAB function
-``mle''. Then for the parameters $\sigma$ and $\bar{v}$ we get
+by means of \gls{mle} using the MATLAB function ``mle''. Then for the
+parameters $\sigma$ and $\bar{v}$ we get
 \begin{eqnarray*}
@@
 skewness of the data. For the Volvo 30min returns, the data has a
-skewness of 0.2419, but our model is strictly symmetric, since
-$x$ appears in equation \ref{eq:UncondPDFSymmetric} only as $x^2$.
+skewness of 0.2419, but our model is strictly symmetric, since the
+return $x$ appears in equation \ref{eq:UncondPDFSymmetric} only as $x^2$.
 Hence the above model needs to be improved to accommodate the non-zero
@@
 \cite{Potters2003}). Therefore, the most apparent modification is to
-allow $v$ and $b$ to be correlated. For convenience we decompose
-$v \sim N(0, \sigma)$ as $v = \sigma a$ and assume
+allow $v$, the zero-mean Gaussian variate in the log-volatility, and
+$b$, the zero-mean Gaussian variate in the innovation of the return,
+to be correlated. For convenience we decompose $v \sim N(0, \sigma)$
+as $v = \sigma a$ and assume
 \begin{eqnarray*}
@@
 \end{equation}
-Then
+We then get
 \begin{eqnarray*}
@@
 \end{equation}
-Differentiating with respect to $x$ yields
+Differentiating with respect to $x$ yields the \gls{pdf}
 \begin{eqnarray}
@@
 author for the integral \ref{eq:UncondPDFAsymmetric}. However, the
-moment generating functions (MGF) of PDF \ref{eq:UncondPDFAsymmetric}
+\gls{mgf} of \gls{pdf} \ref{eq:UncondPDFAsymmetric}
 and \ref{eq:UncondPDFAsymmetric1} are easy enough to find and give the
@@
 
-The MGF of \ref{eq:UncondPDFAsymmetric1} can be found as follows:
+The \gls{mgf} of \ref{eq:UncondPDFAsymmetric1} can be found as follows:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-With the MGF, the first 4 moments of $r$ can be computed:
+With the \gls{mgf}, the first 4 moments of $r$ can be computed:
 \begin{equation}
@@
 variable. This full solution can now be used as the initial estimate
-in an MLE procedure.
+in an \gls{mle} procedure.
 
 It is particularly important in our situation to have a good initial
-estimate, because, as has been shown earlier, the PDF and CDF cannot
-be obtained in closed forms and consequently have to be evaluated by
-numerical integration. This is a rather costly procedure especially in
-the context of MLE. So the computation of the moments under the
-assumption of $\mu = 0$ is worthwhile.
+estimate, because, as has been shown earlier, the \gls{pdf} and the
+\gls{cdf} cannot be obtained in closed forms and consequently have to
+be evaluated by numerical integration. This is a rather costly
+procedure especially in the context of \gls{mle}. So the computation of the
+moments under the assumption of $\mu = 0$ is worthwhile.
 
@@
 estimate by matching the moments and then refining the estimate by
-MLE, we obtain the parameter values for a number of return series,
+\gls{mle}, we obtain the parameter values for a number of return series,
 including the Volvo B 30-minute returns described in
-\S\ref{sec:SLV_Symmetric}. Table \ref{tab:assets_params} and
+section \ref{sec:SLV_Symmetric}. Table \ref{tab:assets_params} and
 \ref{tab:assets_moments} present the obtained parameter values and the
@@
 distribution functions evaluated at these parameters.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=50 233 63
+  136]{../pics/Volvo_B_30m_returns.pdf}
+  \caption{\small \it Volvo B 30min returns' unconditional distribution fit to
+    the SV model. The time series runs from
+    2013/10/10 to 2014/03/12. Left: $P(r < x)$ with $x < 0$; Right:
+    $P(r > x)$ with $x > 0$. Blue: empirical CDF obtained from data;
+    Green: model CDF computed using equation
+    \ref{eq:UncondCDFAsymmetric} with $x$ replaced by
+    $(x-\mu)e^{-\bar{v}}$. Both plots are on log-log scale.}
+  \label{fig:Volvo_B_30m_returns}
+\end{figure}
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=65 241 72
+  143]{../pics/Nordea_30m_returns.pdf}
+  \caption{\small \it Nordea Bank 30min returns' unconditional
+    distribution fit to the SV model. The time
+    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
+    $x < 0$; Right: $P(r > x)$ with $x > 0$. Blue: empirical CDF
+    obtained from data; Green: model CDF computed using equation
+    \ref{eq:UncondCDFAsymmetric} with $x$ replaced by
+    $(x-\mu)e^{-\bar{v}}$. Both plots are on log-log scale.}
+  \label{fig:Nordea_30m_returns}
+\end{figure}
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5, clip=true, trim=21 241 28
+  146]{../pics/Ericsson_30m_returns.pdf}
+  \caption{\small \it Ericsson B 30min returns' unconditional
+    distribution fit to the SV model. The time
+    series runs from 2013/10/10 to 2014/03/12. Left: $P(r < x)$ with
+    $x < 0$; Right: $P(r > x)$ with $x > 0$. Blue: empirical CDF
+    obtained from data; Green: model CDF computed using equation
+    \ref{eq:UncondCDFAsymmetric} with $x$ replaced by
+    $(x-\mu)e^{-\bar{v}}$. Both plots are on log-log scale.}
+  \label{fig:Ericsson_30m_returns}
+\end{figure}
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    & $\psi$ & $\sigma$ & $\bar{v}$ & $\mu$ \\
+    \hline
+    Volvo B & -1.6$\times 10^{-2}$ & 4.7$\times 10^{-1}$ & -6.2&
+    -5.3$\times 10^{-5}$ \\
+    Nordea Bank & 1.7$\times 10^{-2}$ & 4.3$\times 10^{-1}$ & -6.4 &
+    7.5$\times 10^{-5}$ \\
+    Ericsson B & 1.8$\times 10^{-2}$ & 4.5$\times 10^{-1}$ & -6.3 &
+    -9.3$\times 10^{-5}$ \\
+    \hline
+  \end{tabular}
+  % \begin{tabular}{|c|c|c|c|c|}
+  %   \hline
+  %   & $\psi$ & $\sigma$ & $\bar{v}$ & $\mu$ \\
+  %   \hline
+  %   Volvo B & -1.5747e-02 & 4.6929e-01 & -6.2304e+00 &
+  %   -5.3040e-05 \\
+  %   Nordea Bank & 1.7234e-02 & 4.2780e-01 & -6.3758e+00 & 7.5230e-05
+  %   \\
+  %   Ericsson B & 1.8486e-02 & 4.4905e-01 & -6.2834e+00 & -9.2817e-05 \\
+  %   \hline
+  % \end{tabular}
+  \caption{\small \it Parameter Values of Selected Assets'
+    returns. The time series are 30-minute returns and run from
+    2013/10/10 to 2014/03/12.}
+  \label{tab:assets_params}
+\end{table}
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    & mean & std & skewness & kurtosis \\
+    \hline
+    \multirow{2}{*}{Volvo B} & -8.7$\times 10^{-5}$ & 2.5$\times 10^{-3}$ &
+    -2.7$\times 10^{-2}$ & 7.3 \\
+    & -6.9$\times 10^{-5}$ & 2.5$\times 10^{-3}$ & -7.3$\times
+    10^{-2}$ & 7.2 \\
+    \hline
+    \multirow{2}{*}{Nordea Bank} & 7.5$\times 10^{-5}$ & 2.0$\times 10^{-3}$ &
+    6.9$\times 10^{-2}$ & 6.7 \\
+    & 8.9$\times 10^{-5}$ & 2.0$\times 10^{-3}$ &
+    6.7$\times 10^{-2}$ & 6.2 \\
+    \hline
+    \multirow{2}{*}{Ericsson B} & -8.3$\times 10^{-5}$ & 2.3$\times 10^{-3}$ &
+    2.7$\times 10^{-1}$ & 7.7 \\
+    & -7.6$\times 10^{-5}$ & 2.3$\times 10^{-3}$ &
+    7.9$\times 10^{-2}$ & 6.7 \\
+    \hline
+  \end{tabular}
+  \caption{\small \it Moments of Selected Assets' returns. For each return
+    series, the 1st row contains the sample moments, while the 2nd
+    constains those computed with MLE parameters. The time series are
+    30-minute returns and run from 2013/10/10 to 2014/03/12.}
+  \label{tab:assets_moments}
+\end{table}
 
@@
 \subsection{Relation to Conditional Distribution Functions}
+\label{sec:SV_Conditional}
 In the last section we have shown that the unconditional distribution of
-the returns are skewed, which implies the variate in the
-log-volatility $a$ is correlated to the variate in the return
-$b$.
+the returns are skewed, which implies the zero-mean Gaussian variate $a$ in the
+log-volatility is correlated to the zero-mean Gaussian variate $b$ in
+the return (c.f. equation \ref{eq:UnconditionalPdf}).
 
@@
 correlation translates to the correlation between the residual of the
-log-volatility, denoted $y_t$ in \S\ref{chp:nordea_15min} and
-\S\ref{sec:volvo}, and $b_t$. Now let us consider the forecast
-function of an ARIMA model:
+log-volatility, denoted $y_t$ in section \ref{chp:nordea_15min} and
+section \ref{sec:volvo}, and $b_t$. Now let us consider the forecast
+function of an \gls{arima} model:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\phi_i \leq 1$ if the model involves integration, and $\phi_i <
+where $\phi_i < 2$ if the model involves integration, and $\phi_i <
 1$ otherwise. Comparing this equation with equation \ref{eq:r_t}, one
-immediately realizes the PDF of $r_t$
-\begin{eqnarray*}
-  r_t &=& \mu + \sigma_t b_t
+immediately realizes that the conditional \gls{pdf} of $r_t$
+\begin{eqnarray*}
+  r_t &=& \mu + \sigma_t b_t \\
+  &=& \mu + \exp\left(
+    y_t + \sum_{i=1}^P \phi_i \ln \sigma_{t-i} -
+    \sum_{i=1}^Q \theta_i y_{t-i}
+  \right) b_t
 \end{eqnarray*}
@@
 
-For the Nordea series considered in \S\ref{chp:nordea_15min},
+For the Nordea series considered in section \ref{chp:nordea_15min},
 $\text{corr}(y_t, b_t)$ is found to be $3.56 \times 10^{-2}$; while
-for the Volvo series considered in \S\ref{sec:volvo},
+for the Volvo series considered in section \ref{sec:volvo},
 $\text{corr}(y_t, b_t)$ is found to be $-7.3 \times 10^{-3}$. These
 values are comparable to those in table \ref{tab:assets_params}, where
-$\psi$ for the Nordea series is $1.7234 \times 10^{-2}$ and $-1.5747
+$\psi$ is $1.7 \times 10^{-2}$ for the Nordea series and $-1.6
 \times 10^{-2}$ for the Volvo series. The similarity in these values
@@
 
-It is certainly impossible to compare a predicted conditional
-distribution with observations, but conditional distributions are of
-great interest in the context of risk management and derivative
-pricing, so we point out how they may be calculated for ARIMA
-log-volatility models.
+% It is certainly impossible to compare a predicted conditional
+% distribution with observations, but conditional distributions are of
+% great interest in the context of risk management and derivative
+% pricing, so we point out how they may be calculated for ARIMA
+% log-volatility models.

### 2014-06-11 thesis/thesis.tex
Move matlab files to a separate folder.

@@
 \usepackage{amssymb}
-\usepackage{amsfonts, amsmath}
+% \usepackage{amsfonts, amsmath}
+\usepackage{amsmath}
 \usepackage{mathrsfs}
@@
 \usepackage[bookmarks=true]{hyperref}
+\usepackage[acronym]{glossaries}
 \usepackage{bookmark}
@@
 \input{../physics_common.tex}
+\input{./acronyms.tex}
 \DeclareGraphicsExtensions{.pdf,.png,.jpg}
@@
 % \subtitle{}
-
+\makeglossaries
 \author{Xie Xiaolei}
@@
 %\HRule \\[0.4cm]
-{ \huge \bfseries Returns Models and Their Covariance Matrix
+{ \huge \bfseries Return Models and Covariance Matrices
   \\[0.4cm] }
@@
 % \maketitle
+\begin{abstract}
+Return models and covariance matrices of return series have been
+studied. In particular, \gls{garch} and \gls{sv} models
+are compared with respect to their forecasting accuracy when applied
+to intraday return series. \gls{sv} models are found to be considerably more
+accurate and more consistent in accuracy in forecasting.
+
+Covariance matrices formed from Gaussian and \gls{garch} return series, and
+in particular, return series auto-correlated as an AR(1) process, have
+been studied. In the case of Gaussian returns, the largest eigenvalue
+is found to approximately follow a gamma distribution also when the
+returns are auto-correlated. Expressions relating the mean and the
+variance of the asymptotic Gaussian distribution of the matrix
+elements are derived. In the case of \gls{garch} returns, both the largest
+and the smallest eigenvalues of the covariance matrix are seen to
+increase with increasing auto-correlation. The matrix elements are
+found to follow L\'evy distributions with different L\'evy indices for
+the diagonal and the non-diagonal elements.
+
+Localization of eigenvectors of covariance matrices of
+returns from \gls{garch} processes has been investigated. It is
+found that the localization is reduced as the auto-correlation is
+increased. Quantitatively, the number of localized eigenvectors
+decreases approximately as a quadratic function with the
+auto-correlation strength, i.e. the autoregressive coefficient of the
+AR(1) process.
+\end{abstract}
+
+\section*{Acknowledgements}
+First of all, I thank my supervisor, Prof. Sven \AA berg, for the many
+hours of mentoring, discussion, encouragement and guidance. Without
+them, this report will not exist. Then I wish to give my gratitude to
+the people in the division of mathematical physics. The casual
+discussions that I had with them truely inspired me, and the friendly
+and lively atmosphere that they create has helped bring peace and clarity
+to my mind.
+
+Last but not least, I want to thank my wife, Wang Xiaodan, for her
+unconditional support and understanding.
+
 \tableofcontents
+
+% \chapter*{List of Acronyms}
+
+\printglossaries
 
@@
 
-Here by dynamics we mean how the value of an asset is affected by its
-own history as well as by the histories and current values of other
-assets in the market. The influence from the asset's own history is
-termed autocorrelations, i.e. correlations in time, while
-the influence from other assets are termed cross correlations.
-
-Mathematical models that describe the aforementioned observables in
-terms of probability distributions, autocorrelations and cross
-correlations give predictive power and help determine a fair price of
-a given asset. Hence they are a central topic of the mathematics of
-the financial market. In the domain of discrete-time returns' models,
-GARCH and stochastic volatility models (SV) are the most popular, so in
-this thesis project we study a total of 7 intraday return series 
-using GARCH and SV models and compare their performance in volatility
-forecast.
-
-% Once a model has been tentatively constructed, it must be
-% verified. The verification again relies on the statistics of the
-% observables. Over the years, a number of phenomena have been
-% consistently reported and become generally accepted. These are referred
-% to as stylized facts, which are detailed in \S\ref{sec:StylizedFacts}
-% and constitute a standardized test suit for model verification.
-
-% Judged by the stylized facts, each of the models proposed in the
-% literature has its own advantages and disadvantages. The most
-% prominent and influential among them are arguably the GARCH
-% models. Their first and simplest variant was proposed by Bollerslev in
-% 1990 \cite{Bollerslev86} and has been developed tremendously since
-% then.
-
-% An apparent advantage of the GARCH models is their simplicity. They
-% are linear models, where the conditional variance is modeled as a
-% linear combination of past variances and past innovations squared. As
-% a result, it is relatively easy to fit a model to a given set of data,
-% i.e. to estimate the model parameters with respect to the
-% data. Because of the linearity, the estimation of model parameters
-% often give consistent results, regardless of the choice of initial
-% parameter values or the choice of pre-sample data.
-
-% In contrast, many of the other types of models are often specified in
-% a complicated, although more advanced way, for example, by stochastic
-% differential equations. This kind of specifications certainly have the
-% advantage of directly linking to the theory of stochastic processes,
-% but nonetheless are difficult to fit to data, one of the most tricky
-% problems being the high sensitivity to the choice of initial parameter
-% values.
-
-% Another advantage of GARCH models is that their tail behavior is
-% better understood than that of their competitors. The simplest variant
-% of GARCH, namely the GARCH(1,1) models are proven to have regularly
-% varying tails, or in common terms, power-law tails
-% \cite{mikosch2000}. In particular, if the $\alpha_1$ and $\beta_1$
-% parameters of the model sum up to 1, the return series generated by
-% the model has a power-law tail with power exponent 2, and consequently
-% the squared returns have a tail with power exponent 1 -- by the
-% central limit theorem the sum of the squared returns will converge to
-% the Cauchy distribution, a stable distribution with L\'evy index 1. We
-% explore this fact in more details in chapter
-% \ref{chp:CrossCorrelationFat} where the cross-correlation matrix is
-% studied.
-
-% The GARCH model with $\alpha_1 + \beta_1 = 1$ is more often referred
-% to as an IGARCH model, aka. an integrated GARCH model. It is of
-% particular interest because, very often, when fitting a GARCH(1,1)
-% model to real financial data, one finds $\alpha_1 + \beta_1$ extremely
-% close to 1, i.e. at values like 0.99, 0.995, etc.
-
-% A third advantage of GARCH models is their flexibility. Namely one can
-% choose any appropriate distribution to model the innovations. While
-% the Gaussian and the student's t distributions are the most common,
-% the use of other distributions such as asymmetric student's t and
-% Johnson Su has also been studied in the literature (see
-% \cite{Simonato2012}).
-
-% However, GARCH models have their disadvantages too. The original
-% GARCH model does not account for skewness in the distribution of
-% the returns. As a result, the skewness, often attributed to
-% price-volatility correlations, have to be added ``by hand''. This kind
-% of treatment fixes the model on the one hand, but introduces some
-% arbitrariness on the other. Put in another way, skewness does not
-% follow naturally from the structure of the model but rather is added
-% explicitly.
-
-% Moreover, fitting a GARCH model often results in a large number of
-% parameters; 16 to 30 are not uncommon when intra-day data are being
-% dealt with. This raises the question of true fitness: the model
-% specification may indeed fit the selected data set but very well fall
-% short of an accurate description of the time series in question. In
-% other words, such a model is unstable and hence unreliable: if one
-% removes a small fraction of data from the beginning of the set and
-% appends some to the end -- a necessary operation for the evolution of
-% the model -- the model parameters change drastically and so do the
-% predictions.
-
-% As intra-day transaction data become more accessible, the volatility,
-% which has been considered an unobservable quantity by models like
-% GARCH, can be estimated by the so called realized volatility, which is
-% the square root of the sum of squared returns that have been sampled
-% at a higher frequency than are those that one intends to model. For
-% example, it has been shown that the volatility of daily returns is
-% best estimated using 17-minute intra-day returns \cite{Sahalia05}.
-
-% Now that the volatility can be estimated, models that take advantage
-% of this are expected to achieve significantly higher accuracy than
-% those that do not. The model presented in this thesis is built indeed
-% in this spirit.
-
-% Like the GARCH models, our proposed model has the simple structure
-% $r_t = \sigma_t z_t$, where $z_t$ has unit variance and is
-% drawn from any appropriate distribution; $\sigma_t$ is the
-% volatility. Unlike GARCH, which treats $\sigma_t$ as deterministic at
-% time $t$, we treat $\sigma_t$ as random and model $\ln\sigma_t$ using
-% time series analysis. This is possible because past values of
-% $\ln\sigma_t$ can be estimated using past values of the returns at a
-% higher frequency, as has just been discussed.
-
-% Our model has a number of advantages too. Like the original
-% GARCH model, it is simple. It has two random variables $z_t$ and
-% $\ln\sigma_t$, the latter being the subject of time series
-% modeling. The use of conventional time series analysis means the
-% model can be not only more accurate but also more economic -- long
-% memory of the volatility, which is the cause of a large number of
-% parameters in a GARCH model, can be modeled by differencing or
-% fractional differencing. As a result, much fewer parameters are needed
-% for the model.
-
-% Secondly, unlike GARCH models, skewness follows directly from the
-% correlation between the two random variables $z_t$ and
-% $\ln\sigma_t$. As shown in chapter \ref{chp:PriceModels},
-% positive/negative correlation leads to positive/negative skewness,
-% and skewness increases with the strength of correlation.
-
-% Thirdly, our model has sub-exponential tails, as shown in chapter
-% \ref{chp:PriceModels}. This type of tails are not as fat as regularly
-% varying tails, but in our specific case, the tails do behave
-% approximately as a power-law in the regions relevant to stock and
-% index returns. After all, power-law behavior is an empirical result
-% obtained in the returns data, and such data are not available to a
-% statistically significant amount at extremely large values. In plain
-% words, situations where one makes a 20\%+ profit on a single stock
-% trade in a single day happen too rarely to allow any statistical
-% statement.
-
-Another important part of the thesis is about the cross-correlation
-matrix, for which we study the distribution of its elements as well
-as its eigenvalues. When the cross-correlation matrix is constructed
-from returns with simple Gaussian distribution, the matrix is called a
-Wishart Matrix and has been studied extensively in the literature. If
-the returns have L\'evy distributions, the matrix is termed
-Wishart-L\'evy and has been studied to some extent, particularly
-regarding its eigenvalue distribution \cite{politi2010}.
+The dynamics of an asset is affected by its own history as well as by
+the histories and current values of other assets in the market. The
+influence from the asset's own history is termed autocorrelations,
+i.e. correlations in time, while the influence from other assets are
+termed cross correlations.
+
+\Gls{sv} models built on historical data describe the time evolution
+of the aforementioned observables in terms of probability
+distributions, autocorrelations and cross correlations. They may have
+some predictive power and help determine a fair price of a given
+asset.
+
+Quite often, instead of an asset's price, the relative price change,
+i.e. the return, is studied. Analytically solvable models often assume
+Gaussian return distribution, although data suggest fatter tails. In
+numerical return models, realistic reproduction of historical data may
+be achieved.
+
+% In the domain of discrete-time returns' models,
+% \gls{garch} and stochastic volatility models (SV) are the most popular, so in
+% this thesis project we study a total of 7 intraday return series 
+% using \gls{garch} and SV models and compare their performance in volatility
+% forecast.
+In addition to the model of a single asset, the correlation between
+a group of assets is also of great interest. For example, in principle
+component analysis, one wishes to identify a number of factors that
+``drive'' the price evolution of a group of assets in the sense that
+each of the assets' returns can be expressed as a linear combination
+of the factors' returns. In this scenario, the eigenvalues of the
+covariance matrix are the variances of the factors' returns and their
+corresponding orthogonalized eigenvectors give the composition of the
+factors, i.e. the coefficients with which the factors are contructed
+as a linear combination of the assets.
+
+Therefore, in this thesis we also study the elements and the
+eigenvalues distribution, as well as the eigenvectors composition of
+the covariance matrix. When the matrix is constructed from returns
+with simple Gaussian distribution, the matrix is termed a Wishart matrix
+and has been studied extensively in the literature. If the returns
+have L\'evy distributions, the matrix is termed Wishart-L\'evy and has
+been studied to some extent, particularly regarding its eigenvalue
+distribution \cite{politi2010}.
 
@@
 complicated than a straight-forward Gaussian or L\'evy distribution can
-describe --- instead, one needs structured models for their description.
-
-Therefore in this thesis we study how the elements and the eigenvalues
-of the cross-correlation matrix are distributed when the matrix is
-built from returns described by different returns' models, specifically
-GARCH(1,1) and stochastic volatility models, whose the tail behavior is
-understood. Moreover, we also study how auto-correlations in the
-returns influence the aforementioned distributions --- such
-auto-correlations, known as second-order auto-correlations decay
-exponentially but may still leave footprints in the cross-correlation
-matrix.
-
-The rest of this report is organized as follows:
-
-% \S\ref{sec:FundamentalConcepts} presents a few very important
-% concepts and notations that will often be referred to in later
-% chapters; \S\ref{sec:StylizedFacts} gives a review of the ``stylized
-% facts'', i.e. a number of consistently reported and generally accepted
-% phenomena that must be accounted for by a returns' model.
-
-Chapter \ref{chp:PriceModels} reviews some of the most
-influential returns' models, fits them to a few intraday return
-series and compares their performance. A calculation of the
-unconditional distribution functions of stochastic volatility models,
+describe --- instead, one needs structured models. For this reason, we
+are particularly interested in a covariance matrix obtained for
+realistic return models. The so called \gls{garch}(1,1) model is a realistic
+return model proven to have regularly varying tails
+\cite{mikosch2000}. So we study properties of eigenvalues and
+eigenvectors of such covariance matrices. Moreover, we also study how
+auto-correlations in the returns influence the aforementioned
+properties. Such auto-correlations, known as second-order
+auto-correlations decay exponentially but may still leave footprints
+in the covariance matrix.
+
+This report is organized as follows: Chapter \ref{chp:PriceModels}
+reviews some of the most influential return model. Parameters of the
+models are fitted to a few intraday return series and the predictive
+power of the models is compared. A calculation of the
+unconditional distribution functions of \gls{sv} models,
 especially in the case where the residual of the log-volatility
 and the innovation of the return are correlated normal variates, is
-also presented. Chapter \ref{chp:Gaussian} investigates the elements' and
-eigenvalues' distributions of the Wishart matrix, and studies the influence
-of auto-correlated returns. Chapter \ref{chp:CrossCorrelationFat}
-looks at the elements' and eigenvalues' distributions of a covariance
-matrix that is built from identically specified GARCH(1,1) series.
+also presented. In chapter \ref{chp:Gaussian} we investigate
+distributions of eigenvalues of the Wishart matrix, and study the
+influence of auto-correlated returns. In chapter
+\ref{chp:CrossCorrelationFat} distributions of elements and
+eigenvalues of the covariance matrix of identically specified
+\gls{garch}(1,1) series are studied. Finally, chapter \ref{chp:summary}
+summarizes the results. Supplementary materials are provided in the
+appendices.
 
@@
 \label{chp:CrossCorrelationFat}
-
-% In this chapter we study the cross-correlation matrix $C$ constructed from
-% returns that have a fat-tailed distribution, i.e.
-% \begin{eqnarray*}
-%   C &=& {1 \over T^{2/\alpha}} RR'
-% \end{eqnarray*}
-% where $R$ is an $N\times T$ returns' matrix whose entries are
-% \begin{eqnarray}\label{eq:CrossCorrelationMatrix}
-%   R_{it} &=& \left\{
-%   \begin{array}{ll}
-%     a_{it} & \text{if $t=1$} \\
-%     \phi R_{i, t-1} + a_{it} & \text{if $t>1$} \\
-%   \end{array} \right.
-% \end{eqnarray}
-% Here $a_{it}$ are random variables drawn from a stable distribution
-% with l\'evy index $\alpha$. In \S\ref{sec:LevyWishart} we look at
-% the case where the stable distribution is chosen to be Cauchy; in
-% \S\ref{sec:XieWishart} we look at the case where $a_{it}$ is
-% described by an ARIMA log-volatility model, which is detailed in
-% \S\ref{sec:SLV_model}; finally in \S\ref{sec:GarchWishart} we
-% investigate the case where $a_{it}$ is a GARCH(1, 1) process, which is
-% proven to have regularly varying tails by Mikosch et al
-% \cite{mikosch2000}.
-
-% \section{Wishart-L\'evy Matrices}\label{sec:LevyWishart}
-% In this section we study the case where the distribution of $a_{it}$
-% as appear in equation \ref{eq:CrossCorrelationMatrix} is chosen to be
-% Cauchy, a stable distribution with L\'evy index $\alpha =
-% 1$. Politi et al derived the eigenvalue distribution of Wishart-L\'evy
-% matrices in \cite{politi2010} using free probability theory. Here we
-% show, qualitatively, how the distribution changes when
-% auto-correlations are present among the return series. To be precise,
-% the PDF of $a_{it}$ is
-% \begin{eqnarray*}
-%   f_a(x) &=& {\gamma \over \pi} {1 \over \gamma^2 + x^2}
-% \end{eqnarray*}
-
-% By numerically simulating a large number of cross-correlation matrices
-% and solving for the eigenvalues of each simulated matrix, one can
-% obtain the eigenvalue distribution corresponding to each value of
-% $\phi$. Figure \ref{fig:WishartLevySpectraPositive} shows the
-% distribution for a range of positive $\phi$ values, and figure
-% \ref{fig:WishartLevySpectraNegative} shows those for a range of
-% negative $\phi$ values.
-% \begin{figure}[htb!]
-%   \centering
-%   \subfigure[]{
-%     \includegraphics[scale=0.35, clip=true, trim=94 223 109
-%     134]{../pics/WishartLevyEigDist.pdf}
-%     \label{fig:WishartLevySpectraPositive}
-%   }
-%   \subfigure[]{
-%     \includegraphics[scale=0.35, clip=true, trim=94 223 109
-%     134]{../pics/WishartLevyNegEigDist.pdf}
-%     \label{fig:WishartLevySpectraNegative}
-%   }
-%   \caption{\small \it Eigenvalue distribution (PDF) of Wishart-L\'evy
-%     Matrices. Horizontal axis: eigenvalues ($\lambda$) on log-scale;
-%     Vertical axis: PDF ($f_\phi(\lambda)$) of the eigenvalues on
-%     log-scale. \ref{fig:WishartLevySpectraPositive}: $\phi > 0$;
-%     \ref{fig:WishartLevySpectraNegative}: $\phi < 0$;  The $|\phi|$
-%     values 0, 0.5, 0.8, and 0.99 correspond to $\tau$ (correlation
-%     time) values 0, 1.0, 3.10, and 68.97.  The correlation time is
-%     independent of the sign of $\phi$.}
-% \end{figure}
-
-% It is seen in these figures
-% \begin{itemize}
-% \item the eigenvalue distribution is identical for $\phi$ and
-%   $-\phi$;
-% \item the distribution is not bounded within a finite interval but rather
-%   extends to infinity, in constrast to the Marcenko-Pastur law that
-%   bounds the maximum eigenvalue at $\sigma^2(1 + \sqrt{N/T})^2$;
-% \item as auto-correlation increases, the distribution flatens,
-%   implying probability being transferred to large eigenvalues.
-% \end{itemize}
-
-% \section{Cross-Correlation Matrix of ARIMA Log-volatility
-%   returns}\label{sec:XieWishart}
-% In this section we study the cross-correlation matrix of return series
-% specified as follows:
-% \begin{equation}
-%   \label{eq:CrossCorrSLV1}
-%   \begin{aligned}
-%     R_{it} &= \phi R_{i, t-1} + a_{it} \\
-%     a_{it} &= e^{v_{it}} b_{it}
-%   \end{aligned}
-% \end{equation}
-% where $v_{it}$ is typically described by an ARIMA model as discussed
-% in \S\ref{sec:SLV_model}. The unconditional distribution of
-% such returns is calculated in \S\ref{sec:XieCalc}, where it is shown
-% that all the moments of the distribution are finite. Based on this
-% observation, we expect the Marcenko-Pastur law to emerge for the
-% eigenvalue distribution of the cross-correlation matrix C, when
-% $\phi=0$, i.e. when no auto-correlations are absent. In this case C is
-% built as
-% \begin{eqnarray*}
-%   C &=& {1 \over T} \sum_{t=1}^T (DR)(DR)' \\
-% \end{eqnarray*}
-% where $R'$ denotes the transpose of $R$ and
-% \begin{eqnarray*}
-%   D &=&
-%   \begin{pmatrix}
-%     1/\sigma_1 & 0 & \cdots & 0 \\
-%     0 & 1/\sigma_2 & \cdots & 0 \\ 
-%     \vdots & \vdots & \ddots & 0 \\
-%     0 & 0 & \cdots & 1/\sigma_N
-%   \end{pmatrix}
-% \end{eqnarray*}
-% $\sigma_i$ is the standard deviation of return series $i$. By
-% numerical simulation as described above, the eigenvalue distribution
-% of C is obtained. We plot it in figure \ref{fig:FatWishartSpectra}.
-% \begin{figure}[htb!]
-%   \centering
-%   \vspace{-20mm}
-%   \includegraphics[scale=0.5, clip=true, trim=91 229 116
-%   137]{../pics/FatWishartEigDist.pdf}
-%   \caption{\small \it Eigenvalue distribution (PDF) of a
-%     cross-correlation matrix with returns described by
-%     \ref{eq:CrossCorrSLV1}. $\phi$ = 0, 0.5, 0.8, 0.95 correspond to
-%     correlation time $\tau$ = 0.00, 1.00, 3.11, 13.51}
-%   \label{fig:FatWishartSpectra}
-% \end{figure}
-% As expected, the case with zero auto-correlation has an eigenvalue
-% distribution governed by the Marcenko-Pastur law. As auto-correlations
-% become stronger, the distribution deforms, but in a rather different
-% way if compared to the case of Gaussian returns. Figure
-% \ref{fig:FatWishartSpectra} shows that the minimum eigenvalue
-% decreases with increasing $\phi$ in the present case of ARIMA
-% log-volatility returns, while figure \ref{fig:Gaussian_mineig} shows
-% that, in the case of Gaussian returns, the minimum eigenvalue actually
-% increases with $\phi$. We leave this interesting comparison to future
-% investigations.
-
-% \section{Cross-Correlation Matrix of GARCH(1,1) returns}
-% \label{sec:GarchWishart}
-GARCH models and particularly GARCH(1,1) models are widely used to
+\gls{garch} models and particularly \gls{garch}(1,1) models are widely used to
 model financial return series of various time scales, ranging from
@@
 in chapter \ref{chp:PriceModels}. One particularly nice feature of
-GARCH(1,1) models is that they have regularly varying tails (power-law
+\gls{garch}(1,1) models is that they have regularly varying tails (power-law
 tails) \cite{Mikosch2009, mikosch2000} even when the innovations
 (denoted $z_t$ in the following text) are normally distributed ---
-something not shared by other classes of models, e.g. not by ARIMA
-log-volatility models, as shown in \S\ref{sec:XieCalc}. However, what
-this tail behavior implies for the covariance matrix is much less
-understood, especially when the constituent return series are
+something not shared by other classes of models (e.g. not by
+\gls{sv} models, as shown in section \ref{sec:XieCalc}) but well
+documented for realistic returns data by empirical studies
+\cite{Mantegna2000, Potters2003, Guhr2007}. However, what this tail
+behavior implies for the covariance matrix is much less understood,
+especially when the return series of the covariance matrix are
 auto-correlated.
@@
 So in this chapter we consider the covariance matrix of N
-identically specified, possibly auto-correlated GARCH(1, 1)
+identically specified, possibly auto-correlated \gls{garch}(1, 1)
 processes:
@@
 \end{eqnarray*}
-Mikosch et al showed in \cite{mikosch2000} that a GARCH(1,1) process
-satisfying
+Mikosch and Starica showed in \cite{mikosch2000} that a
+\gls{garch}(1,1) process satisfying
 \begin{eqnarray*}
   \alpha_0 &>& 0 \\
-  \E(\alpha_1 Z^2 + \beta_1) &<& 0 \\
+  \E\ln(\alpha_1 Z^2 + \beta_1) &<& 0 \\
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-  \E[(\alpha_1 z^2 + \beta_1)^{p/2}] &\geq& 1 \\
+  \E[(\alpha_1 Z^2 + \beta_1)^{p/2}] &\geq& 1 \\
   \E|Z|^p \ln|Z| &\leq& \infty
@@
 Here $Z$ is a random variable that has the same distribution as
-$z_{it}$. In our simulations described hereafter $z_{it}, Z \sim N(0,
-1)$. In this section and the next, we first study situations where no
-auto-correlations are present among the returns, i.e. $\phi = 0$; then
-in \S\ref{sec:garch_nonzero_phi} we look at how auto-correlations
-change the picture.
+$z_{it}$. In our simulations described hereafter, $z_{it}$ and $Z$
+have standard Gaussian distribution. In this section and the next, we
+first study situations where no auto-correlations are present among
+the returns, i.e. $\phi = 0$; then in section \ref{sec:garch_nonzero_phi} we
+look at how auto-correlations change the picture.
 
 With regularly varying tails, the eigenvalue distribution of a
-cross-correlation matrix built from GARCH(1,1) returns is expected to
-differ from the Marcenko-Pastur law. In figure
-\ref{fig:garch_ev_cii} we simulate N=50 independent GARCH(1,1)
+covariance matrix built from \gls{garch}(1,1) returns is expected to
+differ from the Marcenko-Pastur law discussed in section
+\ref{sec:wishart_eigen_dist}. In figure \ref{fig:garch_ev_cii} we
+simulate N=50 independent \gls{garch}(1,1) 
 returns series, each with identical parameters, namely $\alpha_0 =
 2.3\times 10^{-6}$, $\alpha_1 = 0.15$, $\beta_1 = 0.84$, $\phi = 0$
-and T=$8\times10^4$ time steps, then we build the cross-correlation
+and T=$8\times10^4$ time steps, then we build the covariance
 matrix as
 \begin{eqnarray*}
-  C &=& {1 \over T^{2/\alpha}} RR'
-\end{eqnarray*}
-where R is an $N\times T$ matrix, with elements $r_{it}$ specified by
-\ref{eq:garch_spec}. The PDF of the eigenvalue distribution of C is
-plotted in figure \ref{fig:garch_ev_cii_linear} and CDF of the
-distribution is plotted on log-log scale in
-\ref{fig:garch_ev_cii_loglog}.
+  \mtx C &=& {1 \over T^{2/\alpha}} \mtx{RR'}
+\end{eqnarray*}
+where $\mtx R$ is an $N\times T$ matrix, whose elements $r_{it}$ are
+specified by equation \ref{eq:garch_spec} with $\phi = 0$. The
+normalization factor $1 \over T^{2/\alpha}$ has been chosen such that the
+eigenvalue distribution is independent of $T$ in the limit $T \to
+\infty$ \cite{politi2010, Cizeau1994}.
+
+The \gls{pdf} of the eigenvalue distribution of C is plotted in figure
+\ref{fig:garch_ev_cii_linear} and \gls{cdf} of the distribution is
+plotted on log-log scale in \ref{fig:garch_ev_cii_loglog}.
 \begin{figure}[htb!]
@@
   \subfigure[]{
-    \includegraphics[scale=0.40, clip=true, trim=111 241 110
-    134]{../pics/garch_ev_cii_loglog.pdf}
+    \includegraphics[scale=0.40, clip=true, trim=104 274 104
+    226]{../pics/garch_ev_cii_loglog.pdf}
     \label{fig:garch_ev_cii_loglog}
   }
-  \caption{\small \it \ref{fig:garch_ev_cii_linear}: Eigenvalues \&
-    Diagonal elements' distribution of a cross-correlation matrix
+  \caption{\small \it \ref{fig:garch_ev_cii_linear}: Eigenvalues and
+    Diagonal elements' distribution of a covariance matrix
     built from independent GARCH return series. Blue: PDF of
@@
 Also plotted in the same figure is the distribution of the diagonal
-elements of C. It is clear from the figure that the two PDFs coincide,
-implying C is diagonal. This is further confirmed by figure
-\ref{fig:garch_cij} which shows the distribution of the non-diagonal
-elements of C. One can see the non-diagonal elements are distributed
-symmetrically around 0 with a very small width in comparison to the
-distribution of the diagonal elements - in fact, 1 order of magnitude
-smaller ($2.14\times10^{-5}$ v.s. $7.31\times 10^{-4}$). Hence $C$ is
-very close to a diagonal matrix.
+elements of $\mtx C$. It is clear from the figure that the two PDFs
+coincide, implying $\mtx C$ is diagonal. This is further confirmed by
+figure \ref{fig:garch_cij} which shows the distribution of the
+non-diagonal elements of C. One can see the non-diagonal elements are
+distributed symmetrically around 0 with a very small width in
+comparison to the distribution of the diagonal elements --- in fact, 1
+order of magnitude smaller ($2.14\times10^{-5}$ v.s. $7.31\times
+10^{-4}$). Hence $\mtx C$ is very close to a diagonal matrix.
 \begin{figure}[htb!]
@@
   \caption{\small \it Distribution of the non-diagonal elements of
-    the cross-correlation matrix. Blue: PDF of the non-diagonal
+    the covariance matrix. Blue: PDF of the non-diagonal
     elements; Green: $\alpha$-stable distribution fitted to the
@@
 Figure \ref{fig:garch_ev_cii} also shows the two curves are well fitted
-by an $\alpha$-stable distribution, with $\alpha \approx 1.38$. This
+by an $\alpha$-stable distribution. An estimate of the L\'evy index
+$\alpha$ of the stable distribution is also obtained via fitting,
+$\alpha \approx 1.38$, as shown in figure \ref{fig:garch_ev_cii}. This
 is really an expected result for the diagonal elements. Using
 $\alpha_1 = 0.15$, $\beta_1 = 0.84$, which are the values used for
-simulating the GARCH returns, one can obtain, by solving equation
-\ref{eq:garch_alpha}, $\alpha=2.96$. Then according to Mikosch et al
+simulating the \gls{garch} returns, one can obtain, by solving equation
+\ref{eq:garch_alpha}, $\alpha=2.96$. Then according to Mikosch and Starica
 \cite{mikosch2000}
@@
 with parameters $(\alpha/2, 1, \gamma, \mu)$. Here $\alpha/2$ is the
-L\'evy index, 1 is the asymmetry, $\gamma$ is the scale parameter
+L\'evy index, 1 is the asymmetry, $\gamma$ is the scaling parameter
 and $\mu$ is the mean value of the distribution. Asymmetry being 1
@@
 \end{eqnarray*}
-and the scale parameter $\gamma$ in \ref{eq:stable_CLT} is determined
-by the limit \cite{Bilik2008}
+where we have used the result $\E(|r_t|^2) = \alpha_0 / (1 - \alpha_1
+- \beta_1)$ \cite{Bollerslev86}. The scaling parameter $\gamma$ in
+\ref{eq:stable_CLT} is determined by the limit \cite{Bilik2008}
 \begin{eqnarray*}
@@
   \right) c_0 \\
-  \gamma &=& (2\pi)^{1/\alpha} T^{2/\alpha} \E\left(
+  \gamma &=& (2\pi)^{1/\alpha} T^{2/\alpha} \left(\E
     |z|^{\alpha/2}
@@
 \begin{eqnarray*}
-  \phi(k; \alpha, \beta, \gamma, \mu) &=& \exp\left[
+  \varphi(k; \alpha, \beta, \gamma, \mu) &=& \exp\left[
     i\mu k - \gamma^\alpha |k|^\alpha \left(
@@
 \end{eqnarray*}
-from which we see $\phi(ak; \alpha, \beta, \gamma, \mu) = \phi(k;
+from which we see $\varphi(ak; \alpha, \beta, \gamma, \mu) = \varphi(k;
 \alpha, \beta, a\gamma, a\mu)$, implying that, if $x \sim S(\alpha,
@@
 we have
-\begin{eqnarray*}
+\begin{eqnarray}
   C_{ii} &=& {1 \over T^{2/\alpha}}\sum_{t=1}^T r_{it}^2 \\
@@
   S(\alpha/2, 1, \gamma_D, \mu_D) \text{ as $T \to \infty$}
-\end{eqnarray*}
+  \label{eq:garch_cii_dist}
+\end{eqnarray}
 where
@@
 \end{eqnarray}
-So the diagonal elements of the cross-correlation matrix converge to
+So the diagonal elements of the covariance matrix converge to
 an $\alpha$-stable distribution with L\'evy index $\alpha/2 \approx
-1.48$. This is consistent with the measured index 1.38 within
-measurement errors.
-
-Now we look at the distribution of the non-diagonal elements. At first
-glance, one might be tempted to think that figure \ref{fig:garch_cij}
-shows a Gaussian distribution, but in fact, it is not. This is
-illustrated in the probability plot in figure
-\ref{fig:garch_nondiag_probplot}. In comparison to the PDF of a
-Gaussian, the PDF of the nondiagonal elements is significantly fatter
-on the tails.
+1.48$. This is comparable to the index value 1.38 obtained by
+fitting. Considering the slow convergene of regularly varying tails,
+this is a reasonably good match.
+
+Now we look at the distribution of the non-diagonal elements. Figure
+\ref{fig:garch_cij} shows that an $\alpha$-stable distribution fits
+rather well, and additionally, figure \ref{fig:garch_nondiag_probplot}
+shows that the distribution of these non-diagonal elements has fat
+tails, supporting an $\alpha$-stable distribution.
 \begin{figure}[htb!]
@@
 
-The parameters of the non-diagonal elements' distribution $S(\alpha',
-\beta', \gamma', \mu')$ can be obtained via fitting. The results have
-been shown in the legend of figure \ref{fig:garch_cij}. In table
-\ref{tab:garch_wishart_cij_params} we list them with a higher
-precision.
+The parameters of the fitted $\alpha$-stable distribution $S(\alpha',
+\beta', \gamma', \mu')$ are obtained in the procedure of fitting. The
+results have been shown in the legend of figure
+\ref{fig:garch_cij}. In table \ref{tab:garch_wishart_cij_params} we
+list them with a higher precision.
 \begin{table}[htb!]
@@
 
-Figure \ref{fig:GarchCiiEig1} shows the diagonal elements' as well as
-the eigenvalues' distributions when the number of return series (N) is
-fixed at 250 and the number of observations in each series (T) is
-increased from 3000 to 6000.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=81 248 75
+Figure \ref{fig:GarchCiiEig1} shows the distributions of the diagonal
+elements as well as the eigenvalues when the number of return series
+(N) is fixed at 250 and the number of observations in each series (T)
+is gradually increased from 3000 to 6000.
+\begin{figure}[htb!]
+  \centering
+    \includegraphics[scale=0.7, clip=true, trim=81 248 75
     195]{../pics/GarchCiiEig1.pdf}
@@
 elements' distribution do not coincide as well but differ rather
-siginificantly for small values. For large values, however, the two do
-coincide and comply with the fitted $\alpha$-stable
-distribution. Moreover, the difference between the eigenvalues'
-distribution and the diagonal elements' distribution, as well as that
-between the diagonal elements' distribution and the fitting
-$\alpha$-stable distribution, is also seen to diminish as T
+siginificantly for small values. For large values of the diagonal
+elements and the eigenvalues, the two do coincide and comply with the
+fitted $\alpha$-stable distribution. Moreover, the difference between
+the eigenvalues' distribution and the diagonal elements' distribution,
+as well as that between the diagonal elements' distribution and the
+fitting $\alpha$-stable distribution, is also seen to diminish as T
 increases.
@@
 \end{eqnarray*}
-where $f(\cdot)$ denotes the PDF of the diagonal elements and $c$ is
+where $f(\cdot)$ denotes the \gls{pdf} of the diagonal elements and $c$ is
 some constant. Thus, at the limit $C_{ii} \to \infty$, the distance
@@
 further illustrated in figure \ref{fig:garch_eigenvec_Cmax_dist} where
-the PDF of $|c|_\M$ is plotted. We see in this figure that an
+the \gls{pdf} of $|c|_\M$ is plotted. We see in this figure that an
 increased value of T leads to advancement of $\min(|c|_\M)$ to larger
@@
   \centering
-  \includegraphics[scale=0.5, clip=true, trim=14 199 31
+  \subfigure[]{
+  \includegraphics[scale=0.29, clip=true, trim=14 199 31
   155]{../pics/garch_eigenvec_Cmax.pdf}
-  \caption{\small \it largest component ($|c|_\M$) corresponding to
-    the eigenvalue. N is fixed to 50.}
     \label{fig:garch_eigenvec_Cmax}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=86 257 112
+  }
+  \subfigure[]{
+  \includegraphics[scale=0.41, clip=true, trim=86 257 112
   136]{../pics/garch_eigenvec_Cmax_dist.pdf}
-  \caption{\small \it PDF of the largest component ($|c|_\M$) of
-    each eigenvector. N is fixed to 50.}
   \label{fig:garch_eigenvec_Cmax_dist}
+  }
+  \caption{\small \it \ref{fig:garch_eigenvec_Cmax}: Horizontal:
+    eigenvalues ($\log \lambda$); Vertical: $\log |c|_\M$ of the
+    corresponding eigenvector. From left to right and 1st to 2nd row,
+    T has values 800, 1000, 1200,
+    80000. \ref{fig:garch_eigenvec_Cmax_dist}: PDF of the largest
+    component ($|c|_\M$) of each eigenvector. In both plots the number
+    of returns (N) is 50.}
+%  \label{fig:garch_eigenvec_Cmax_both}
+%   \includegraphics[scale=0.4, clip=true, trim=14 199 31
+%   155]{../pics/garch_eigenvec_Cmax.pdf}
+%   \caption{\small \it largest component ($|c|_\M$) corresponding to
+%     the eigenvalue. N is fixed to 50.}
+%   \label{fig:garch_eigenvec_Cmax}
+% \end{figure}
+
+% \begin{figure}[htb!]
+%   \centering
+%   \includegraphics[scale=0.4, clip=true, trim=86 257 112
+%   136]{../pics/garch_eigenvec_Cmax_dist.pdf}
+%   \caption{\small \it PDF of the largest component ($|c|_\M$) of
+%     each eigenvector. N is fixed to 50.}
+%   \label{fig:garch_eigenvec_Cmax_dist}
 \end{figure}
@@
 ``Inverse Participation Ratio'' (IPR). For a given normalized eigenvector
-$\vec{c}_i = (c_{1, i}, c_{2, i}, ..., c_{N, i})$, the IPR is defined
+$\vec{c}_i = (c_{1, i}, c_{2, i}, ..., c_{N, i})$, the \gls{ipr} is defined
 as \cite{Aberg2013}
-\begin{eqnarray*}
-  \text{IPR}(\vec{c}_i) &=& \sum_{k=1}^N c_{k,i}^4 \\
-\end{eqnarray*}
+\begin{eqnarray}
+  \text{IPR}(\vec{c}_i) &=& \sum_{k=1}^N c_{k,i}^4 \label{eq:IPR_def}
+\end{eqnarray}
 Figure \ref{fig:garch_eigenvec_PR} shows ${1 \over
@@
 eigenvalues. This quantity is sometimes termed the
-normalized participation ratio (PR) and measures the proportion of
+normalized \gls{pr} and measures the proportion of
 basis vectors that contribute considerably to the eigenvector in
@@
 
-Figure \ref{fig:garch_eigenvec_PR_dist} shows the PDF of the
-normalized PR. We see that as T increases, the distribution of PR is
-compressed towards 0, suggesting increased localization of the
-eigenvectors.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=49 208 70
+Figure \ref{fig:garch_eigenvec_PR_dist} shows the \gls{pdf} of the
+normalized \gls{pr}. We see that as T increases, the distribution of
+\gls{pr} is compressed towards 0, suggesting increased localization of
+the eigenvectors.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \label{fig:garch_eigenvec_PR}
+    \includegraphics[scale=0.34, clip=true, trim=49 208 70
     162]{../pics/garch_eigenvec_PR.pdf}
-  \caption{\small \it Normalized participation ratio (PR)
-    versus eigenvalue.}.
-  \label{fig:garch_eigenvec_PR}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=97 259 113
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=97 259 113
     226]{../pics/garch_eigenvec_PR_dist.pdf}
-  \caption{\small \it PDF of the normalized PR. N is fixed to
+  \label{fig:garch_eigenvec_PR_dist}
+  }
+  \caption{\small \it \ref{fig:garch_eigenvec_PR}: Normalized
+    participation ratio (PR) versus eigenvalue
+    ($\lambda$). Horizontal: $\log \lambda$; Vertical: $\log
+    \text{PR}$ of the corresponding
+    eigenvector. From left to right and 1st to 2nd row, T has values
+    800, 1000, 1200, 80000. \ref{fig:garch_eigenvec_PR_dist}: PDF of
+    the normalized PR. In both plots, the number of returns (N) is
     50.}
-  \label{fig:garch_eigenvec_PR_dist}
 \end{figure}
@@
 % When auto-correlations are present among the returns, the
-% cross-correlation matrix is expected to change. How exactly it changes
+% covariance matrix is expected to change. How exactly it changes
 % is the subject of this section. Here we consider a model specified
@@
 % are the same as in the previous case of zero auto-correlations, namely
-% $\alpha_1 = 0.15$ and $\beta_1 = 0.84$. The cross-correlation matrix C
+% $\alpha_1 = 0.15$ and $\beta_1 = 0.84$. The covariance matrix C
 % is built from R using
@@
 13.51$. The values of N and T are 50 and $8\times 10^4$ as before. The
-GARCH(1,1) parameters are also unchanged, namely $\alpha_0 = 2.3\times
+\gls{garch}(1,1) parameters are also unchanged, namely $\alpha_0 = 2.3\times
 10^{-6}$, $\alpha_1 = 0.15$, and $\beta_1 = 0.84$. Figure
@@
 increased.  In figure \ref{fig:garch_ev_cii} we see that, when no
-auto-correlation is present, the PDF of the diagonal elements has
+auto-correlation is present, the \gls{pdf} of the diagonal elements has
 width ($\gamma$) $7.31 \times 10^{-4}$, while in figure
@@
 10^{-3}$ as $\tau$ becomes 13.51. Similarly the non-diagonal elements'
-PDF has width $2.14 \times 10^{-5}$ when $\tau=0$, as shown in figure
+\gls{pdf} has width $2.14 \times 10^{-5}$ when $\tau=0$, as shown in figure
 \ref{fig:garch_cij}, and this width becomes $9.60 \times 10^{-4}$ when
@@
 \ref{fig:GarchEigDiag1}.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4, clip=true, trim=85 258 103
+  229]{../pics/GarchSpectrumAutocorrelated.pdf}
+  \caption{\small \it Eigenvalue distribution of covariance
+    matrix built from auto-correlated GARCH processes. $\phi$ = 0,
+    0.5, 0.8, 0.955, 0.97 correspond to correlation time $\tau$ =
+    0, 1.00, 3.11, 15.05, 22.76.}
+  \label{fig:GarchSpectrumAutocorrelated}
+\end{figure}
 From this figure one can see that, as auto-correlation strengthens,
 \begin{itemize}
-\item the PDF of the eigenvalue distribution flattens and widens;
+\item the \gls{pdf} of the eigenvalue distribution flattens and widens;
 \item the minimum as well as the maximum eigenvalues increase.
@@
 now varies in a larger range than it does with smaller $\phi$.
-
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4, clip=true, trim=0 197 0
+  154]{../pics/garch_eigenvec_Cmax_corr.pdf}
+  \caption{\small \it Eigenvectors' largest component versus
+    eigenvalue, for 4 auto-correlation strengths $\phi$ = 0, 0.6, 0.9,
+    and 0.99. The number of returns (N) is 50.}
+  \label{fig:garch_eigenvec_Cmax_corr}
+\end{figure}
 The same story of reduced localization is also evident from the plot
-of the normalized participation ratio (PR), shown in figure
-\ref{fig:garch_eigenvec_PR_dist_corr}, and from the PDF of $|c|_\m$
+of the normalized \gls{pr}, shown in figure
+\ref{fig:garch_eigenvec_PR_dist_corr}, and from the \gls{pdf} of $|c|_\M$
 shown in figure \ref{fig:garch_eigenvec_Cmax_dist_corr}. It is seen
@@
 to smaller values by increased auto-correlation.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=95 260 111
+    232]{../pics/garch_eigenvec_PR_dist_corr.pdf}
+    \label{fig:garch_eigenvec_PR_dist_corr}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.4, clip=true, trim=95 260 111
+    232]{../pics/garch_eigenvec_Cmax_dist_corr.pdf}
+    \label{fig:garch_eigenvec_Cmax_dist_corr}
+  }
+  \caption{\small \it \ref{fig:garch_eigenvec_PR_dist_corr}: PDF of
+    the normalized participation ratio
+    (PR). \ref{fig:garch_eigenvec_Cmax_dist_corr}: PDF of the largest
+    component of the eigenvectors. The number of returns (N) is
+    50. $\phi$ values of 0, 0.6000, 0.9000, 0.9900 correspond to
+    correlation time $\tau$ = 0, 1.3569, 6.5788, 68.9676}
+\end{figure}
 
@@
 the ratios are plotted versus $\phi$ on log-log scale.
-
-% \subsection{Correlation Matrix of a Factor Model}
-%% Result of normalizing by 1/T
-% In this section we consider the estimation of correlation matrix in
-% the text of a factor model: the observed return series
-% $\{r_{it}\}$ ($1 \leq i \leq N$, $1 \leq t \leq T$) are driven by a
-% number of unobserved and uncorrelated random variables $z_{it}$
-% (factors), each of which is an auto-correlated GARCH(1,1) process as
-% specified by equation \ref{eq:garch_spec}. Factor GARCH models have
-% been studied extensively in the literature. A good review is found in
-% \cite{Mikosch2009}. Here we are interested in the impact of
-% auto-correlations in $z_{it}$ on the cross-correlation matrix of
-% $r_{it}$.
-
-% To be specific we adopt the generalized orthogonal GARCH model by
-% Weide \cite{Weide2002}:
-% \begin{eqnarray*}
-%   \vec{r_t} &=& \mtx{W}\vec{z_t}
-% \end{eqnarray*}
-% where $\mtx{W}$ is an $N \times N$ invertible matrix; $\vec{z_t}$ and
-% $\vec{r_t}$ are $N \times 1$ column vectors. Moreover, to meet our
-% interest, $\vec{z_t}$ are assumed to be auto-correlated:
-% \begin{eqnarray*}
-%   \vec{z}_t &=& \phi \vec{z}_{t-1} + \vec{y}_t \\
-%   \vec{y}_t &=& \mtx{H}^y_t \vec{x}_t \\
-% \end{eqnarray*}
-% where $\vec{x_t} \sim N(\vec{0}, \mtx{I})$, and $\mtx{H}^y_t$,
-% the conditional covariance matrix of $\vec{y}_t$, is given by
-% \begin{eqnarray*}
-%   \mtx{H}^y_t &=& (\mtx{I} - \mtx A - \mtx B) + \mtx A \odot
-%   (\vec{r}_{t-1} \vec{r}'_{t-1}) + \mtx B
-%   \mtx{H}^y_{t-1}
-% \end{eqnarray*}
-% Here $\mtx A$ and $\mtx B$ are diagonal $N \times N$ matrix and
-% $\odot$ denotes element-wise multiplication. The form of the constant
-% term, $\mtx{I} - \mtx A - \mtx B$, ensures that the unconditional
-% variance of $\vec{y}_t$ is $\mtx I$. In the simulations described in
-% the rest of this section, we use the parametrization $\mtx A =
-% \alpha_1 \mtx I$, $\mtx B = \beta_1 \mtx I$, where $\alpha_1 = 0.15$,
-% $\beta_1 = 0.84$. For tractability, we fix $N=4$ and consider one
-% particular covariance matrix of $\vec{r}_{it}$:
-% \begin{eqnarray*}
-%   \mtx C &=& \text{cov}(\vec{r}_{it}) \\
-%   &=& \begin{pmatrix}
-%     1 & \rho & \rho & \rho \\
-%     \rho & 1 & \rho & \rho \\
-%     \rho & \rho & 1 & \rho \\
-%     \rho & \rho & \rho & 1 \\
-%   \end{pmatrix}
-% \end{eqnarray*}
-% where $\rho = 0.5$. The eigenvalue decomposition of $\mtx C$ is
-% \begin{eqnarray*}
-%   \mtx C &=& \mtx{U} \mtx{\Lambda} \mtx{U}' \\
-%   \mtx U &=&
-%   \begin{pmatrix}
-%     -1/\sqrt{2} & -1/\sqrt{6} & -1/2\sqrt{3} & 1/2 \\
-%     1/\sqrt{2} & -1/\sqrt{6} & -1/2\sqrt{3} & 1/2 \\
-%     0 & \sqrt{2}/\sqrt{3} & -1/2\sqrt{3} & 1/2 \\
-%     0 & 0 & \sqrt{3}/2 & 1/2 \\
-%   \end{pmatrix} \\
-%   \mtx \Lambda &=&
-%   \begin{pmatrix}
-%     1-\rho & 0 & 0 & 0 \\
-%     0 & 1-\rho & 0 & 0 \\
-%     0 & 0 & 1-\rho & 0 \\
-%     0 & 0 & 0 & 1+ 3\rho \\
-%   \end{pmatrix}
-% \end{eqnarray*}
-
-%% Abandoned.
-% In the last section we have investigated what the measured
-% covariance matrix looks like when the return series have
-% auto-correlated GARCH(1,1) returns but are actually independent of
-% each other. In this section we investigate the situation when the
-% return series are truly correlated. In particular, we study the
-% following case:
-% \begin{equation}\label{eq:garch_correlated_returns}
-%   \begin{aligned}
-%     x_{1t} &= r_{1t} & \\
-%     x_{it} &= \sqrt{1 - \rho^2} r_{it} + \rho r_{1t} & \text{i = 2, 3,
-%       ..., N}\\
-%   \end{aligned}
-% \end{equation}
-% where $x_{it}$ are the return series whose covariance matrix is the
-% subject of the current study and $r_{it}$ are independent
-% auto-correlated GARCH(1,1) processes as specified by equation
-% \ref{eq:garch_spec}.
-
-% The covariance matrix C of the returns $\{x_{it}\}_{t=1}^T$ can be trivially
-% computed 
-% \begin{eqnarray}
-%   C_{ij} &=& {1 \over T^{2/\alpha}} \sum_{t=1}^T x_{it} x_{jt}
-%   \nonumber \\
-%   C &\xrightarrow{T \to \infty}&
-%   \begin{pmatrix}
-%     \sigma_1^2 & \rho \sigma_1^2 & \cdots & \rho \sigma_1^2 \\
-%     \rho \sigma_1^2 & (1 - \rho^2) \sigma_2^2 + \rho^2 \sigma_1^2 & \cdots &
-%     \rho^2 \sigma_1^2 \\
-%     \vdots & \vdots & \ddots & \vdots \\
-%     \rho \sigma_1^2 & \rho^2 \sigma_1^2 & \cdots & (1 - \rho^2)
-%     \sigma_N^2 + \rho^2 \sigma_1^2 
-%   \end{pmatrix} \label{eq:garch_cov}
-% \end{eqnarray}
-% where $\sigma_i$ is the standard deviation of the GARCH(1,1) series
-% $\{r_{it}\}_{t=1}^T$:
-% \begin{eqnarray*}
-%   \sigma_i = {1 \over T^{2/\alpha}} \sum_{t=1}^T r_{it}^2
-% \end{eqnarray*}
-
-% We would like to find out, when the covariance matrix of the $x$
-% series is estimated, how the auto-correlation in the $r$ series affects
-% the result --- specifically, how the auto-correlation strength $\phi$
-% (c.f. equation \ref{eq:garch_spec}) affects the distribution of
-% $C_{ij}$ when T is not very large.
-
-% Figure \ref{fig:garch_correlated_Cij} shows the distribution of
-% $C_{1j} / C_{11}$ ($j \geq 2$) and $C_{i1} / C_{11}$ ($i \geq
-% 2$). These rescaled elements, as seen from equation
-% \ref{eq:garch_cov}, have mean $\rho$. This is exactly what one sees in
-% figure \ref{fig:garch_correlated_Cij}. It is also clear from the
-% figure that $\phi$, the strength of auto-correlation, does not
-% introduce a bias but does increase the variance of the
-% estimation. This is the same conclusion as drawn in
-% \S\ref{sec:garch_nonzero_phi}.
-% \begin{figure}[htb!]
-%   \centering
-%     \includegraphics[scale=0.5, clip=true, trim=96 260 112
-%     227]{../pics/garch_correlated_Cij.pdf}
-%   \caption{\small \it Probability density function (PDF) of $C_{1j} /
-%     C_{11}$ ($j \geq 2$) and $C_{i1} / C_{11}$ ($i \geq 2$). The
-%     covariance matrix is constructed from 50 return series
-%     $\{x_{it}\}$, $i = 1, 2, ..., 50$, and $t=1, 2, ..., 600$, as
-%     specified by equation \ref{eq:garch_correlated_returns}. $\rho =
-%     0.5$. 4000 instances of the covariance matrix are generated for
-%     each value of $\phi$.}
-%   \label{fig:garch_correlated_Cij}
-% \end{figure}
-% One can also find support in the data that the aforementioned
-% distribution has power-law tails. This is shown in figure
-% \ref{fig:garch_mix_Cij_tail} where the power of the left
-% and the right tail is found to be 3.74 and 3.40, respectively.
-% \begin{figure}[htb!]
-%   \centering
-%     \includegraphics[scale=0.5, clip=true, trim=105 229 107
-%     120]{../pics/garch_mix_Cij_tail.pdf}
-%   \caption{\small \it Horizontal axis (x) : the elements $C_{1j} /
-%     C_{11}$ ($j \geq 2$) and $C_{i1} / C_{11}$ ($i \geq 2$) shifted to
-%     center around 0. The section with $0.15 < x < 0.35$ is
-%     plotted. Vertical axis: See the legend. }
-%   \label{fig:garch_mix_Cij_tail}
-% \end{figure}
-
-% It is also interesting to see how the eigenvalues and eigenvectors
-% behave given the non-trivial cross-correlation and the
-% auto-correlation in the returns. However, the cross-correlation matrix
-% of equation \ref{eq:garch_cov} has rather complicated eigenvalues and
-% eigenvectors. It has been useful in studying the non-diagonal matrix
-% elements but becomes clumsy with regard to eigenvalues and
-% eigenvectors. Therefore we look at the following return series
-% instead:
-% \begin{eqnarray}
-%   \begin{pmatrix}
-%     x_{1t} \\
-%     x_{2t} \\
-%   \end{pmatrix} &=&
-%   \begin{pmatrix}
-%     \cos\theta/\sigma_1 & -\sqrt{2} \sin\theta/\sigma_2  \\
-%      \sin\theta/\sigma_1 & \sqrt{2} \cos\theta/\sigma_2 \\
-%   \end{pmatrix}
-%   \begin{pmatrix}
-%     r_{1t} \\
-%     r_{2t} \\
-%   \end{pmatrix} \label{eq:x-r_transformation}
-% \end{eqnarray}
-% where $r_{it}$ are independent, auto-correlated GARCH(1,1) processes
-% as specified by equation \ref{eq:garch_spec}. Each of them has
-% standard deviation $\sigma_i$ ($i = 1, 2$). It is easy to
-% find the covariance matrix C of the series $\{x_{it}\}_{t=1}^T$ ($i =
-% 1,2$):
-% \begin{eqnarray*}
-%   C &=& 
-%   \begin{pmatrix}
-%     1 + \sin\theta^2 & -\sin\theta\cos\theta \\
-%     -\sin\theta\cos\theta & 1 + \cos\theta^2 \\
-%   \end{pmatrix}
-% \end{eqnarray*}
-% The eigenvalues and eigenvectors of $C$ are
-% \begin{eqnarray*}
-%   CX &=& X
-%   \begin{pmatrix}
-%     1 & \\
-%     & 2 \\
-%   \end{pmatrix} \\
-%   X &=& 
-%   \begin{pmatrix}
-%     \cos\theta & -\sin\theta \\
-%     \sin\theta & \cos\theta \\
-%   \end{pmatrix}
-% \end{eqnarray*}
-
-% Figure \ref{fig:garch_rotated_Cii} shows the probability plot of the
-% diagonal elements of $C$. Evidently, their distribution is very close
-% to a Gaussian, although the Anderson-Darling test rejects Gaussianity
-% at the 5\% significance level. Through equation
-% \ref{eq:x-r_transformation} the differences in variance of the
-% realizations of $\{r_{1t}\}_{t=1}^T$ and $\{r_{2t}\}_{t=1}^T$ are not
-% carried over to $\{x_{1t}\}_{t=1}^T$ and $\{x_{2t}\}_{t=1}^T$, and
-% therefore the differences in variance of the latter series are simply
-% the results of numerical computation and hence expected to be small
-% and Gaussian distributed. One can also see from figure
-% \ref{fig:garch_rotated_Cii} that the different auto-correlations
-% (different $\phi$ values) in the $\{r_{1t}\}_{t=1}^T$ and
-% $\{r_{2t}\}_{t=1}^T$ series do not change the Gaussianity. The
-% variances of these Gaussian distributions, however, are indeed
-% increased by increased auto-correlation.
-% \begin{figure}[htb!]
-%   \centering
-%     \includegraphics[scale=0.4, clip=true, trim=24 152 44
-%     18]{../pics/garch_rotated_Cii.pdf}
-%   \caption{\small \it Probability plot of $C_{11}$ corresponding to
-%     different auto-correlation strength ($\phi$) in the
-%     $r$-series. When $\phi = 0, 0.2, 0.4, 0.6$, the variance is $3.380
-%     \times 10^{-4}$, $3.639 \times 10^{-4}$, $4.512 \times 10^{-4}$,
-%     $7.506 \times 10^{-4}$, respectively.} 
-%   \label{fig:garch_rotated_Cii}
-% \end{figure}
-
-% On the other hand, the non-diagonal elements of $C$ have a very
-% different distribution from the diagonal ones, as shown in figure
-% \ref{fig:garch_rotated_Cij}.
-% \begin{figure}[htb!]
-%   \centering
-%     \includegraphics[scale=0.5, clip=true, trim=24 303 30
-%     274]{../pics/garch_rotated_Cij.pdf}
-%   \caption{\small \it Probability density function (PDF) of the
-%     non-diagonal elements of C. Corresponding to $\phi = 0, 0.2, 0.4,
-%     0.6, 0.8$, the variance of the distribution is
-%     $6.77\times10^{-8}$, $1.529\times10^{-7}$, $3.896\times10^{-7}$,
-%     $1.124\times10^{-6}$, $5.445\times10^{-6}$.}
-%   \label{fig:garch_rotated_Cij}
-% \end{figure}
-% One can see that the distribution of the non-diagonal elements is
-% left-skewed, regardless of the auto-correlation $\phi$. This is to be
-% compared with the distributions shown in figure
-% \ref{fig:garch_correlated_Cij}, where the skewness is small and
-% changes sign for different $\phi$ values, whereas in figure
-% \ref{fig:garch_rotated_Cij}, the skewness is consistently negative and
-% takes comparable values as $\phi$ changes from 0 to 0.9. We list these
-% skewness values in table \ref{tab:garch_rotated_var_skw}.
-% \begin{table}[htb!]
-%   \centering
-%   \begin{tabular}{|c|c|c|c|c|c|}
-%     \hline
-%     $\phi$ & 0 & 0.1 & 0.2 & 0.3 & 0.4 \\
-%     \hline 
-%     variance& 6.770e-08 & 1.008e-07 & 1.529e-07 & 2.339e-07 & 3.896e-07 \\
-%     \hline
-%     skewness & -1.62 & -1.65 & -1.96 & -1.86 & -1.67 \\
-%     \hline
-%   \end{tabular}
-%   \begin{tabular}{|c|c|c|c|c|c|}
-%     \hline
-%     $\phi$ & 0.5 & 0.6 & 0.7 & 0.8 & 0.9 \\
-%     \hline
-%     variance& 6.022e-07 & 1.124e-06 & 2.156e-06 & 5.445e-06 & 2.389e-05 \\
-%     \hline
-%     skewness & -1.62 & -1.65 & -1.96 & -1.86 & -1.67 \\
-%     \hline
-%   \end{tabular}
-%   \caption{\small \it Variance and skewness of $C_{12}$ and $C_{21}$}
-%   \label{tab:garch_rotated_var_skw}
-% \end{table}
-% From table \ref{tab:garch_rotated_var_skw} and figure
-% \ref{fig:garch_rotated_Cij} one can also see that the
-% variance steadily increases as $\phi$ increases --- a similar
-% observation that has been seen in figure
-% \ref{fig:garch_correlated_Cij}.
-
-% \begin{figure}[htb!]
-%   \centering
-%   \subfigure[]{
-%     \includegraphics[scale=0.35, clip=true, trim=107 278 104
-%     240]{../pics/garch_rotated_eig1_dist.pdf}
-%     \label{fig:garch_rotated_eig1_dist}
-%   }
-%   \subfigure[]{
-%     \includegraphics[scale=0.35, clip=true, trim=103 270 99 
-%     227]{../pics/garch_rotated_eig2_dist.pdf}
-%     \label{fig:garch_rotated_eig2_dist}
-%   }
-%   \caption{\small \it Eigenvalue distribution}
-%   \label{fig:garch_rotated_eig_dist}
-% \end{figure}
-
-% \begin{table}[htb!]
-%   \centering
-%   \begin{tabular}{|c|c|c|c|c|c|}
-%     \hline
-%     $\phi$ & 0 & 0.2 & 0.4 & 0.6 & 0.8 \\
-%     \hline
-%     variance& 3.338e-07 & 4.113e-07 & 8.415e-07 & 2.351e-06 & 9.580e-06 \\
-%     \hline
-%     skewness & -3.92 & -1.65 & -1.22 & -1.38 & -0.90 \\
-%     \hline
-%   \end{tabular}
-%   \caption{variance and skewness of the distribution of $\lambda_1$}
-%   \label{tab:garch_rotated_eig1_var_skw}
-% \end{table}
-
-% \begin{table}[htb!]
-%   \centering
-%   \begin{tabular}{|c|c|c|c|c|c|}
-%     \hline
-%     $\phi$ & 0 & 0.2 & 0.4 & 0.6 & 0.8 \\
-%     \hline
-%     variance& 4.947e-07 & 7.852e-07 & 1.767e-06 & 4.813e-06 & 2.258e-05 \\
-%     \hline
-%     skewness & 2.93 & 2.14 & 2.41 & 2.42 & 2.49 \\
-%     \hline
-%   \end{tabular}
-%   \caption{variance and skewness of the distribution of $\lambda_2$}
-%   \label{tab:garch_rotated_eig2_var_skw}
-% \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% To find out more about this series of deformation, we investigate 
-% how the characteristic function changes in response to increasing
-% auto-correlation (increasing $\phi$). When $\phi = 0$ and T is very
-% large, like $8 \times 10^4$ in our simulation, we have found the
-% eigenvalue distribution converges to an $\alpha$-stable distribution
-% whose characteristic function is known in analytic form.
-
-% The empirical characteristic function $\varphi(k)$ is computed from
-% the simulated eigenvalues according to:
-% \begin{eqnarray*}
-%   \varphi(k) &=& \E(e^{ik\lambda}) \\
-%   &=& {1 \over M}\sum_{n=1}^M e^{ik\lambda_n}
-% \end{eqnarray*}
-% where $M$, the total number of eigenvalues in the sample, is $1\times
-% 10^5$ in our simulation, as mentioned earlier.
-
-% \begin{eqnarray}\label{eq:PDF_fourier_expansion}
-%   p(\lambda) &=& \sum_{n=-\infty}^\infty c_n \exp\left(
-%     -i {2\pi n \lambda \over L}
-%   \right) \\ \nonumber
-%   &\approx& \sum_{n=-M}^M c_n \exp\left(
-%     -i {2\pi n \lambda \over L}
-%   \right) \\
-%   L &=& \lambda_\M - \lambda_\m \nonumber
-% \end{eqnarray}
-% where $M$ is a large integer at which the series expansion is
-% truncated. For the simulations described in this section, we choose M
-% = 500.
-
-% The Fourier coefficients $c_n$ used in the above formula are computed
-% as
-% \begin{eqnarray*}
-%   c_n &=& {1 \over L} \int_{\lambda_\m}^{\lambda_\M} \exp\left(
-%     i {2\pi n \lambda \over L}
-%   \right) p(\lambda) d\lambda \\
-%   &=& {1 \over L} \E \left[
-%     \exp\left(
-%       i {2\pi n \lambda \over L}
-%     \right)
-%   \right] \\
-%   &=& {1 \over LS} \sum_{k=1}^S \exp\left(
-%     i {2\pi n \lambda_k \over L}
-%   \right)
-% \end{eqnarray*}
-% where $S = 1 \times 10^5$ is the aforementioned sample size of the
-% simulated eigenvalue. Figure \ref{fig:GarchEigPDFFourierCoef}
-% shows the Fourier coefficients computed according to the above
-% formula.
-% \begin{figure}[htb!]
-%   \centering
-%   \includegraphics[scale=0.56, clip=true, trim=10 203 6
-%   89]{../pics/GarchEigPDFFourierCoef.pdf}
-%   \caption{\small \it Fourier Coefficients of the empirical PDF of
-%     the Eigenvalue distribution. Blue/Red: Real/imaginary parts of the
-%     Fourier coefficients. The $\phi$ values 0, 0.2, 0.5, 0.8, 0.95, 0.96,
-%     0.97, 0.98 correspond to correlation time $\tau$ = 0, 0.43, 1.00,
-%     3.11, 13.51, 16.98, 22.76, 34.31.}
-%   \label{fig:GarchEigPDFFourierCoef}
-% \end{figure}
-
-% Before we proceed to investigate the properties of the empirical PDF
-% using the Fourier series expansion, we need to ensure the expansion
-% does provide a sufficiently accurate approximation to the empirical
-% PDF. However, to directly compare the Fourier series expansion
-% with the empirical PDF, the empirical PDF will have to be evaluated
-% --- the accuracy of the result will then largely depend on the number
-% of bins to which one chooses to sort the eigenvalues so as to produce
-% the empirical PDF --- a choice that is rather arbitrary and
-% subjective. To avoid such a procedure, we can compare the empirical
-% CDF with its Fourier expansion instead. Formally we integrate equation
-% \ref{eq:PDF_fourier_expansion} to obtain
-% \begin{equation}
-%   \label{eq:CDF_fourier_expansion}
-%   \begin{aligned}
-%     F(\lambda) &= \int_{\lambda_\m}^{\lambda} p(x) dx \\
-%     &= \sum_{n \in
-%       \mathbb{Z}\backslash\{0\}} c_n i {L \over 2\pi n} \left[
-%       \exp\left(
-%         -i {2\pi n \lambda \over L}
-%       \right) - 
-%       \exp\left(
-%         -i {2\pi n \lambda_\m \over L}
-%       \right) \right] \\
-%     & + (\lambda - \lambda_\m) c_0
-%   \end{aligned}
-% \end{equation}
-% The CDF on the left-hand-side of equation
-% \ref{eq:CDF_fourier_expansion} can be evaluated directly using the
-% simulated eigenvalues; meanwhile, the right-hand-side can be evaluated
-% for any $\lambda$ using $c_n$.
-
-% Figure \ref{fig:CDF_fourier_expansion} shows how the two sides of
-% equation \ref{eq:CDF_fourier_expansion} compare to each other.
-% \begin{figure}[htb!]
-%   \centering
-%   \includegraphics[scale=0.6, clip=true, trim=20 240 15
-%   200]{../pics/garch_cdf_fourier_expansion.pdf}
-%   \caption{\small \it The empirical cummulative distribution
-%     function (CDF) and its Fourier series expansion. Blue: empirical
-%     CDF. Red: Fourier series expansion. The plots are on log-log
-%     scale. The $\phi$ values 0, 0.2, 0.5, 0.8, 0.95, 0.96,
-%     0.97, 0.98 correspond to correlation time $\tau$ = 0, 0.43, 1.00,
-%     3.11, 13.51, 16.98, 22.76, 34.31.}
-%   \label{fig:CDF_fourier_expansion}
-% \end{figure}
-% We see that the two curves match fairly well --- they differ from each
-% other only at their lower ends, which are statistically insignificant
-% due to the very small portion of data points in that region.
-
-% Now that one is assured that the Fourier series expansion is
-% sufficiently accurate, it makes sense to study the properties of the
-% Fourier coefficients $c_n$. A few features immediately present
-% themselves in figure \ref{fig:GarchEigPDFFourierCoef}:
-% \begin{itemize}
-% \item $\re c_n = \re c_{-n}$, where $\re$ denotes the real part of a
-%   complex number;
-% \item $\re c_n$ resembles a dumped cosine wave;
-% \item $\re c_n$ and $\im c_n$ have the same frequency with a constant
-%   phase difference. Here $\im$ denotes the imaginary part of $c_n$.
-% \end{itemize}
-% The last conjecture about the frequency and the phase difference can
-% be verified by looking at the phase plot of $c_n$, which is shown in
-% figure \ref{fig:garch_Cn_phase}.
-% \begin{figure}[htb!]
-%   \centering
-%   \includegraphics[scale=0.6, clip=true, trim=21 248 14
-%   198]{../pics/garch_Cn_phase.pdf}
-%   \caption{\small \it Phases of the Fourier coefficients ($c_n$) of
-%     the empirical eigenvalue PDF. Horizonal axis: $n$; Vertical Axis:
-%     phases of $c_n$. The $\phi$ values 0, 0.2, 0.5, 0.8, 0.95, 0.96,
-%     0.97, 0.98 correspond to correlation time $\tau$ = 0, 0.43, 1.00,
-%     3.11, 13.51, 16.98, 22.76, 34.31.}
-%   \label{fig:garch_Cn_phase}
-% \end{figure}
-% From figure \ref{fig:garch_Cn_phase} one can see that the phase of
-% $c_n$ ($\ph c_n$) is approximately linear when $\phi$ is
-% small --- however, as $\phi$ becomes larger and larger, the
-% non-linearity of $\ph c_n$ becomes stronger and stronger.
-
-% In the relatively simple situation of small $\phi$, i.e. weak
-% auto-correlation, it lends some insight to consider the following
-% approximate formula:
-% \begin{eqnarray}
-%   \re c_n &=& e^{
-%     -u_r(\phi) - v_r(\phi) |n|
-%   } \cos[\omega(\phi) n] \label{eq:garch_fourier_coef_real}\\
-%   \im c_n &=& e^{
-%     -u_i(\phi) - v_i(\phi) |n|
-%   } \sin[\omega(\phi) n] \label{eq:garch_fourier_coef_imag}
-% \end{eqnarray}
-% Apparently, $\omega(\phi)$ can be directly read out from figure
-% \ref{fig:garch_Cn_phase}. The results are listed in table
-% \ref{tab:garch_phase_velocity}.
-% \begin{table}[htb!]
-%   \footnotesize
-%   \centering
-%   \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-%     \hline
-%     $\phi$ & 0 & 0.2 & 0.5 & 0.8 & 0.95 & 0.96 & 0.97 & 0.98 \\
-%     \hline
-%     $\tau$ & 0 & 0.4307 & 1.0000 & 3.1063 &  13.5134 & 16.9797 &
-%     22.7566 & 34.3096 \\
-%     \hline
-%     $\omega(\phi)$ & 0.0442 & 0.0538 & 0.0348 & 0.0414 & 0.0180 &
-%     0.0320 & 0.0154 & 0.0242 \\
-%     \hline
-%   \end{tabular}
-%   \caption{\small \it Phase velocity $\omega$ of the Fourier
-%     coefficients of the eigenvalue PDF.}
-%   \label{tab:garch_phase_velocity}
-% \end{table}
-% Using these values as initial estimate, the parameters in equation
-% \ref{eq:garch_fourier_coef_real} and \ref{eq:garch_fourier_coef_imag},
-% namely $u_r(\phi)$, $v_r(\phi)$ and $v_i(\phi)$ can be estimated by
-% least-square-error methods. We use the Matlab function ``lsqcurvefit''
-% for this purpose and list the estimated parameter values in table
-% \ref{tab:garch_fourier_coef_real_param}.
-% \begin{table}[htb!]
-%   \footnotesize
-%   \centering
-%   \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-%     \hline
-%     $\phi$ & 0 & 0.1 & 0.2 & 0.3 & 0.4 & 0.5 & 0.6 & 0.7 & 0.8\\
-%     \hline
-%     $\tau$ & 0 & 0.3010 & 0.4307 & 0.5757 & 0.7565 & 1.0000 & 1.3569 &
-%     1.9434 & 3.1063 \\
-%     \hline
-%     $u_r(\phi)$ & 0.0938 & 0.0936 & -0.0812 & 0.1649 & 0.2627 & 0.6285 &
-%     0.5443 & 0.5720 & 1.1610 \\
-%     \hline
-%     $v_r(\phi)$ & 0.0044 & 0.0045 & 0.0057 & 0.0046 & 0.0045 & 0.0034 &
-%     0.0046 & 0.0059 & 0.0048 \\
-%     \hline
-%     $\omega(\phi)$ & 0.0420 & 0.0424 & 0.0511 & 0.0428 & 0.0421 &
-%     0.0333 & 0.0416 & 0.0497 & 0.0394 \\
-%     \hline
-%   \end{tabular}
-%   % \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
-%   %   \hline
-%   %   $\phi$ & 0 & 0.1 & 0.2 & 0.3 & 0.4 & 0.5 & 0.6 & 0.7 & 0.8 & 0.9 \\
-%   %   \hline
-%   %   $\tau$ & 0 & 0.3010 & 0.4307 & 0.5757 & 0.7565 & 1.0000 & 1.3569 &
-%   %   1.9434 & 3.1063 & 6.5788 \\
-%   %   \hline
-%   %   $u_r(\phi)$ & 0.0938 & 0.0936 & -0.0812 & 0.1649 & 0.2627 & 0.6285 &
-%   %   0.5443 & 0.5720 & 1.1610 & 1.2347 \\
-%   %   \hline
-%   %   $v_r(\phi)$ & 0.0044 & 0.0045 & 0.0057 & 0.0046 & 0.0045 & 0.0034 &
-%   %   0.0046 & 0.0059 & 0.0048 & 0.0098 \\
-%   %   \hline
-%   %   $\omega(\phi)$ & 0.0420 & 0.0424 & 0.0511 & 0.0428 & 0.0421 &
-%   %   0.0333 & 0.0416 & 0.0497 & 0.0394 & 0.0667 \\
-%   %   \hline
-%   % \end{tabular}
-%   \caption{\small \it Least-Square-Error estimate of parameter values of $\re
-%     c_n$}
-%   \label{tab:garch_fourier_coef_real_param}
-% \end{table}
-% As is said earlier, the equation \ref{eq:garch_fourier_coef_real} and
-% \ref{eq:garch_fourier_coef_imag} only apply in the cases of small
-% $\phi$, so the estimation has only been done for $\phi$ values up to
-% 0.8. To obtain better statistics about the relation between $\phi$ and
-% the parameters, a few more values of $\phi$ have been added. The
-% corresponding Fourier coefficients have been computed and validated
-% against the empirical CDF. Figure \ref{fig:garch_fourier_coef_real}
-% shows how the equation \ref{eq:garch_fourier_coef_real} fits to $\re
-% c_n$ with the estimated parameter values.
-% \begin{figure}[htb!]
-%   \centering
-%   \includegraphics[scale=0.56, clip=true, trim=10 197 11
-%   147]{../pics/garch_fourier_coef_real.pdf}
-%   \caption{\small \it A dumped cosine function fit to $\re
-%     c_n$. Blue: empirical Fourier coefficients; Green: Dumped cosine
-%     function.}
-%   \label{fig:garch_fourier_coef_real}
-% \end{figure}
-% We see that the two curves in figure \ref{fig:garch_fourier_coef_real}
-% match fairly well. Obvious deviations are found only in regions of
-% large n, where slight inaccuracy in $\omega(\phi)$ is magnified. This
-% confirms the validity of formula \ref{eq:garch_fourier_coef_real}.
-
-% % To find out the relation between $\phi$ and the parameters $u_r(\phi)$,
-% % $v_r(\phi)$ and $\omega(\phi)$, we plot these parameters' values
-% % against $\phi$ in figure \ref{fig:Cn_real_param}.
-% % \begin{figure}[htb!]
-% %   \centering
-% %   \includegraphics[scale=0.56, clip=true, trim=6 311 0
-% %   264]{../pics/Cn_real_param.pdf}
-% %   \caption{\small \it Linear fit of the parameters in $\re
-% %     c_n$. Blue circles: parameter values normalized by the sample
-% %     mean; Green line: straight line fitted to the data points.}
-% %   \label{fig:Cn_real_param}
-% % \end{figure}
-% % Instead of directly plotting the parameters' values, figure
-% % \ref{fig:Cn_real_param} plots the ratio of these values over their
-% % respective sample mean, so that one can see how much each
-% % individual data point (blue circles) deviates from the sample mean as
-% % well as from the fitted line. The title of each plot in figure
-% % \ref{fig:Cn_real_param} is the equation of the fitted line.
-
-% % From figure \ref{fig:Cn_real_param} we see that $u_r(\phi)$ is
-% % approximately linear in $\phi$ --- the coefficient of the linear term
-% % in its equation is an order of magnitude larger than the constant
-% % term. In contrast, $v_r(\phi)$ and $\omega(\phi)$ are approximately
-% % constant --- the coefficient of the linear term in their equation is
-% % an order of magnitude smaller than the constant term. So, also
-% % considering $0 < \phi < 1$ and that we have a rather small sample
-% % size, we estimate $v_r(\phi)$ and $\omega(\phi)$ to be constant. In
-% % summary
-% % \begin{eqnarray*}
-% %   u_r(\phi) &=& 1.2364 \phi - 0.1124 \\
-% %   v_r &=& 4.7148 \times 10^{-3} \\
-% %   \omega &=& 3.3476 \times 10^{-2} \\
-% % \end{eqnarray*}
-% % Now that $v_r(\phi)$ and $\omega(\phi)$ are constant, we shall write
-% % them only as $v_r$ and $\omega$ to avoid confusion.
-
-% Turning our attention to the imaginary part of $c_n$($\im c_n$) and
-% applying the same methods, we find the corresponding parameter values
-% for $\im c_n$ (table \ref{tab:garch_fourier_coef_imag_param}). Figure
-% \ref{fig:garch_fourier_coef_imag} shows how these values together with
-% equation \ref{eq:garch_fourier_coef_imag} fit to $\im c_n$. Comparing
-% table \ref{tab:garch_fourier_coef_real_param} and
-% \ref{tab:garch_fourier_coef_imag_param}, we see that $v_r$ and $v_i$
-% are essentially equal, while $u_r(\phi)$ is consistently larger than
-% $u_i(\phi)$.
-
-% \begin{table}[htb!]
-%   \footnotesize
-%   \centering
-%   \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-%     \hline
-%     $\phi$ & 0 & 0.1 & 0.2 & 0.3 & 0.4 & 0.5 & 0.6 & 0.7 & 0.8\\
-%     \hline
-%     $\tau$ & 0 & 0.3010 & 0.4307 & 0.5757 & 0.7565 & 1.0000 & 1.3569 &
-%     1.9434 & 3.1063 \\
-%     \hline
-%     $u_i(\phi)$ & 0.0555 & 0.0565 & -0.1105 & 0.1291 & 0.2236 &
-%     0.6007 & 0.5031 & 0.5341 & 1.1273 \\
-%     \hline
-%     $v_i(\phi)$ & 0.0047 & 0.0047 & 0.0059 & 0.0048 & 0.0048 &
-%     0.0035 & 0.0049 & 0.0062 & 0.0049 \\
-%     \hline
-%     $\omega(\phi)$ & 0.0420 & 0.0424 & 0.0511 & 0.0428 & 0.0421 &
-%     0.0333 & 0.0416 & 0.0497 & 0.0394 \\
-%     \hline
-%   \end{tabular}
-%   \caption{\small \it Least-Square-Error estimate of parameter values of $\im c_n$}
-%   \label{tab:garch_fourier_coef_imag_param}
-% \end{table}
-
-% \begin{figure}[htb!]
-%   \centering
-%   \includegraphics[scale=0.56, clip=true, trim=18 221 19
-%   171]{../pics/garch_fourier_coef_imag.pdf}
-%   \caption{\small \it A dumped sine function fit to the imaginary
-%     part of $c_n$ ($\im c_n$). Red: empirical Fourier coefficients;
-%     Green: Dumped sine function.} 
-%   \label{fig:garch_fourier_coef_imag}
-% \end{figure}
-
-%  As is done for $\re c_n$, we can also find an approximate
-% linear expression for $u_i(\phi)$ by fitting a line to the data. The
-% result is
-% \begin{eqnarray*}
-%   u_i(\phi) &=& 1.2356 \phi - 0.1480
-% \end{eqnarray*}
-% So, combining the results for $\re c_n$ and $\im c_n$, we may write
-% \begin{eqnarray*}
-%   c_n &=& e^{-v|n|}\left[
-%     e^{-a_r\phi - b_r}\cos(\omega n) + ie^{-a_i\phi - b_i}\sin(\omega n)
-%   \right] \\
-% \end{eqnarray*}
-% where
-% \begin{equation}
-%   \label{eq:garch_Cn_parameters}
-%   \begin{aligned}
-%     v &= 4.7148 \times 10^{-3} \\
-%     a_r &= 1.2364 \\
-%     b_r &= -0.1124 \\
-%     a_i &= 1.2356 \\
-%     b_i &= -0.1480 \\
-%     \omega &= 3.3476 \times 10^{-2}
-%   \end{aligned}
-% \end{equation}
-% It should be emphasized that the above parameters are only constants
-% with respect to the GARCH parameters $\alpha_1$, $\beta_1$ and hence
-% the resulting tail exponent $\alpha$. If the GARCH parameters are
-% changed, the parameters in \ref{eq:garch_Cn_parameters} are expected
-% to follow.
-
-%% TODO:
-% 1. find the power-exponent of the tail of the SLV (stochastic
-% log-volatility) model. Normalize the returns using the sample
-% standard deviation when computing the cross-correlation matrix.
-% 
-% 2. Hows does the power-law return distribution change the eigenvalue
-% distribution in comparison to the Marcenko-Pastur law?
-%
-% 3. How do the maximum and the minimum eigenvalue change
-% corresponding to different values of autocorrelation?
-%
-% 4. Use different combinations of q = N/T and investigate how the
-% eigenvalue distribution changes in the case of the SLV model.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.38, clip=true, trim=93 229 115
+    134]{../pics/localization_ratio.pdf}
+    \label{fig:localization_ratio}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.38, clip=true, trim=87 227 115
+    133]{../pics/localization_ratio2.pdf}
+    \label{fig:localization_ratio2}
+  }
+  \caption{\small \it \ref{fig:localization_ratio}: Ratio of localized
+    eigenvectors versus the auto-correlation strength $\phi$. Black
+    ``+'': ratio of localized eigenvectors as measured by the number
+    of participating basis vectors being lower than 2; Black ``x'':
+    ratio of localized eigenvectors as measured by the largest
+    component being larger than 0.9. Blue curve: quadradic function
+    fitted to ``+''. Green curve: quadratic function fitted to
+    ``x''. There are 27 data points in the plot, corresponding to 27
+    $\phi$ values: 0 to 0.8 with step size 0.05, and 0.9 to 0.99 with
+    step size 0.01. The corresponding values of the correlation time
+    $\tau$ range from 0 to 69. \ref{fig:localization_ratio2}: $\ln
+    (f_\M - f)$ is plotted versus $\ln \phi$, where $f$ stands for the
+    ratio of localized eigenvectors. The fitted curves are linear.}
+\end{figure}
 
 \input{summary}
+
+\chapter{Self-reflection}
+First of all, during the thesis project I learned a lot about models
+of time series in finance, such as \gls{garch}, and \gls{sv} models,
+as well as traditional time series analysis, for example, \gls{arima}
+models. In addition, a study of continuous-time models gave me an
+opportunity to increase my knowledge of stochastic differential
+equations, probability theory, and statistics. I am a beginner, but
+nonetheless find it comfortable to read books and articles on these
+subjects after the thesis project.
+
+Moreover, I have introduced myself to the random matrix theory,
+multivariate analysis, and extreme value theory. These became relevant
+when I worked on the covariance matrix of returns described by
+Gaussian or \gls{garch} models. I am still a long way from being able
+to do serious proofs or derivations in these areas, but I know what to
+read and can understand the basics.
+
+In addition, I have become better at Matlab programming and dealing
+with databases, and have grown more confident in numerical analysis.
+
 \appendix
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 \bibliography{econophysics}

### 2014-06-11 m/HestonRetDistr.m

@@ -1,0 +1,20 @@
+% Stationary distribution of the returns in the Hestion model
+function prob_den = HestonRetDistr(model, t, x)
+rho = model.Correlation(2, 1);
+
+prob_den = zeros(length(x), 1);
+for n = 1:length(x)
+    func = @(p) exp(i*p*x(n) + HestonRetDistrFT(model, p, t));
+    lim = 50;
+    Q = integral(func, -lim, lim);
+    diff = NaN;
+    errtol = 1.0e-4;
+    increment = 10;
+    while isnan(diff) || abs(diff) > errtol 
+        diff = integral(func, -lim - increment, -lim) +...
+               integral(func, lim, lim + increment);
+        lim = lim + increment;
+        Q = Q + diff;
+    end
+    prob_den(n) = Q;
+end

### 2014-06-11 m/NormalPowerLaw_pdf.m

@@ -1,0 +1,27 @@
+% function loglf = NormalPowerLaw_pdf(param, x)
+function p = NormalPowerLaw_pdf(param, x)
+p = NaN(size(x));
+
+xi1 = param(1);
+alpha1 = xi1 ./ normcdf(-xi1) .* exp(-xi1.^2 / 2) / sqrt(2*pi) + 1;
+a1 = xi1^alpha1 * exp(-xi1^2 / 2) / sqrt(2*pi);
+
+xi2 = param(2);
+alpha2 = xi2 ./ (1 - normcdf(xi2)) .* exp(-xi2.^2 / 2) / sqrt(2*pi) + 1;
+a2 = xi2^alpha2 * exp(-xi2^2 / 2) / sqrt(2*pi);
+
+I1 = x < -xi1;
+I2 = x > xi2;
+I3 = -xi1 <= x & x <= xi2;
+
+p(I3) = normpdf(x(I3));
+
+%% The left tail
+
+p(I1) = a1 ./ abs(x(I1)).^alpha1;    
+
+%% The right tail
+
+p(I2) = a2 ./ x(I2).^alpha2;    
+
+%loglf = sum(log(p));

### 2014-06-11 m/SymmetricModel.m

@@ -1,0 +1,61 @@
+clear all
+% filename = './data/OMXS30.csv';
+% filename = 'data/VOLV-B-1990-01-01-2014-01-09.csv';
+% filename = 'data/ERIC-B-1990-01-02-2014-01-16.csv';
+% data = dlmread(filename, ';', [1, 1, 6857, 3]);
+
+% price = data(data(:, 1) > 0 & data(:, 3) > 0 &...
+%             data(:, 1) ~= data(:, 2), 3);
+% % price = data(data(:, 3) > 0, 3);
+% price = flipud(price);
+% N = size(price, 1);
+% T = 15;
+% % ret = price2ret(price(mod(N, T):T:end));
+% % ret = ret - mean(ret);
+% r = price2ret(price);
+% r = r - mean(r);
+% sigma = std(r);
+% r = (r - mean(r))/std(r);
+
+
+%% intraday returns
+company = 'ericsson_b';
+first_day = '2013-10-10';
+last_day = '2014-01-20';
+dt = 30;
+ret = get_intra_ret(company, first_day, last_day, dt);
+r = ret - mean(ret);
+sigma = std(r);
+
+C = sqrt(2/pi)*exp(log(log(4))/2);
+f = @(x, s, evbar) 1/evbar*...
+    exp(s^2/2)/sqrt(8*pi)/C .* ...
+    erfc(1/(s*sqrt(2)) * log(abs(x./evbar)/sqrt(log(4))) + s/sqrt(2));
+
+p = mle(r, 'pdf', f, 'start', [1, std(r)], 'lowerbound', [1e-4, 1e-4]);
+% [y, x] = ecdf(r);
+% y1 = F(x, sig);
+% plot(x, y, x, y1, 'r');
+
+% x = linspace(min(r), max(r), length(r)/100);
+% y = hist(r, x)./length(r)/(x(2)-x(1));
+% plot(x, y, x, f(x, p(1), p(2)), 'r');
+
+
+% plot(x(I), y(I), x(I), f(x(I), p(1), p(2)), 'r');
+% plot(log(x(I)), log(y(I)), log(x(I)), log(f(x(I), p(1), p(2))),
+% 'r');
+[y, x] = ecdf(r);
+y1 = retcdf_symmetric(x, p(1), p(2));
+I = x < 0;
+% I = x < -2*sigma & x > -3*sigma;
+% I = x > 2*sigma & x < 3*sigma;
+% I = logical(ones(1, length(x)));
+
[…]

### 2014-06-11 m/correlation.m

@@ -1,0 +1,58 @@
+clear all
+clc
+
+genesis = 0;
+filename1 = '../data/corr_eig_0.4.txt';
+filename2 = '../data/corr_eig_0.6.txt';
+filename = '../data/corr_eig_0.2.txt';
+
+sigma = [
+    1    0    0    0    0    0    0    0    0    0    0    0
+    0.2  1    0    0    0    0    0    0    0    0    0    0
+    0.2  0.2  1    0    0    0    0    0    0    0    0    0
+    0.2  0.2  0.2  1    0    0    0    0    0    0    0    0
+    0.2  0.2  0.2  0.2  1    0    0    0    0    0    0    0
+    0.2  0.2  0.2  0.2  0.2  1    0    0    0    0    0    0
+    0.2  0.2  0.2  0.2  0.2  0.2  1    0    0    0    0    0
+    0.2  0.2  0.2  0.2  0.2  0.2  0.2  1    0    0    0    0
+    0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  1    0    0    0
+    0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  1    0    0
+    0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  1    0
+    0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  0.2  1
+        ];
+sigma = sigma + sigma' - eye(size(sigma, 1));
+N = size(sigma, 1);
+
+if genesis
+    filename = '../data/corr_eig_0.2.txt';
+    %N = 800;
+    T = 800;
+    %Q = T/N;
+
+    for k = 1:400
+        random_matrix_eig(T, sigma, filename);
+    end
+else
+    lambda = [load(filename), load(filename1), load(filename2)];
+
+    a = min(lambda);
+    b = max(lambda);
+
+    cen = linspace(min(a), max(b), 300);
+    dx = cen(2) - cen(1);
+    pdf = double(hist(lambda, cen)) ./ (N*400*dx);
+    plot(cen, pdf);
+end
+
+% a1 = 1+1/Q-2*sqrt(1/Q);
+% b1 = 1+1/Q+2*sqrt(1/Q);
+
+% x = linspace(a1, b1, 200);
[…]

### 2014-06-11 m/eigen_dist_levy.m

@@ -1,0 +1,399 @@
+clear all
+close all
+
+T = 8e+4;
+N = 50;
+fold = 2e3;
+
+%% eigen values generated from the volvo stochastic log-volatility
+%% model.
+
+spec = cellstr(['b  '; 'c  '; 'g  '; 'm  '; 'r  '; 'k  ';...
+                'b--'; 'c--'; 'g--'; 'm--'; 'r--'; 'k--';...
+                'b: '; 'c: '; 'g: '; 'm: '; 'r: '; 'k: ';...
+                'b-.'; 'c-.'; 'g-.'; 'm-.'; 'r-.'; 'k-.';...
+                ]);
+% for i = [1:5]
+%     load(sprintf('./eigen-%d.mat', i));
+%     ev = reshape(ev, N, fold);
+%     maxeig = max(ev);
+%     [y, x] = ecdf(maxeig);
+%     y1 = diff(y) ./ diff(x);
+%     x1 = (x(1:end-1) + x(2:end))/2;
+%     plot(log(x1), log(y1), spec(i));
+%     hold on
+% end
+% hold off
+% grid on
+
+
+%% White Wishart matrix eigen values
+% AR = [0:0.05:0.8, 0.9, 0.95:0.005:0.99];
+AR = 0:0.1:0.9;
+tau = -log(2)./log(AR);
+% dist = struct('name', 'Cauchy', 'prmt', 1);
+dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.15, 0.84], 'distr', ...
+              'Gaussian', 'TailExponent', 2.9664);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.1, 0.80], 'distr', ...
+%               struct('Name', 't', 'DoF', 3), 'TailExponent', 0.7744);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.1, 0.725], 'distr', ...
+%               struct('Name', 't', 'DoF', 3), 'TailExponent', 1.8187);
+% dist = struct('name', 'Garch1_1', 'prmt', [2.3e-6, 0.15, 0.84], 'distr', ...
+%               struct('Name', 't', 'DoF', 60), 'TailExponent', 2.4477);
+% for j = AR
+%     if (exist(sprintf('GarchWishartT8e4Eig-%.3f.mat', j), 'file') == 2)
+%         continue;
+%     end
[…]

### 2014-06-11 m/gen_ret_mtx.m

@@ -1,0 +1,88 @@
+%% Generate a matrix of returns series with specfied AR. coefficients
+function R = gen_ret_mtx (n, T, distr, coef, varargin)
+
+% if strcmp(distr.name, 'garch')
+%     model = garch('Constant', 2.3e-6, 'ARCH', 0.15, 'GARCH', ...
+%                   0.84, 'Distribution', distr.prmt);
+%     [V, R] = simulate(model, T, 'numPaths', n);
+%     R = R';
+%     return;
+% end
+
+% First find the auto-correlation coefficients
+m = length(coef);
+% R = NaN(n, T+m);
+
+% x = NaN(n, T+m);
+% x(:, 1:m) = zeros(n, m);
+switch (distr.name)
+  case {'Normal', 'Gaussian'}
+    x = randn(n, T);
+  case {'Lorentzian', 'Cauchy'}
+    % A uniform distr. in (-1/2, 1/2) %
+    x = rand(n, T) - 1/2;
+    x = distr.prmt(1) .* tan(pi .* x);
+    % y = randn(n, 2);
+    % x = y(:, 1) ./ abs(y(:, 2));
+  case 'Levy'
+    alpha = distr.prmt(1);
+    gam = distr.prmt(2);
+    u = rand(n, T);
+    v = rand(n, T);
+    phi = pi * (v - 0.5);
+    x = gam * (...
+        -log(u) .* cos(phi) ./...
+        cos((1 - alpha) * phi) ...
+        ).^(1 - 1/alpha) .*...
+        sin(alpha * phi) ./ cos(phi);
+  case 'student t'
+    dof = distr.prmt(1);
+    y = randn(n, dof + 1);
+    x = y(:, 1) .* sqrt(dof) ./ sqrt(sum(y(:, 2:end).^2, 2));
+  case 'Garch1_1'
+    a0 = distr.prmt(1);
+    a1 = distr.prmt(2);
+    b1 = distr.prmt(3);
+
+    if strcmp(distr.distr.Name, 'Gaussian') == 1
+        h = [ones(n, 1)*(a0/(1 - a1 - b1)), NaN(n, T-1)];
+        if isempty(varargin)
+            x = [randn(n, 1), NaN(n, T-1)];
+            for t = 2:T
+                h(:, t) = a0 + a1.*x(:, t-1).^2 + b1.*h(:, t-1);
+                x(:, t) = randn(n, 1).*sqrt(h(:, t));
[…]

### 2014-06-11 m/get_mysql.m

@@ -1,0 +1,15 @@
+function mysql = get_mysql()
+timefmt = 'yyyy-mm-dd HH:MM:SS';
+if exist('/nfs/users3/xiexiaol', 'dir') == 7
+    hostname = '83.176.196.41';
+    jdbc = ['/nfs/users3/xiexiaol/lib/mysql-connector-java-5.1.26/'...
+            'mysql-connector-java-5.1.26-bin.jar'];
+    if isempty(strfind(javaclasspath, jdbc))
+        javaaddpath(jdbc);
+    end
+else
+    hostname = 'localhost';
+end
+mysql = database('avanza', 'sinbaski', 'q1w2e3r4',...
+                 'com.mysql.jdbc.Driver', ...
+                 sprintf('jdbc:mysql://%s:3306/avanza', hostname));

### 2014-06-11 m/random_matrix_eig.m

@@ -1,0 +1,12 @@
+%% Generate random matrices
+%function lambda = random_matrix_eig(N, T, Sigma, filename)
+function lambda = random_matrix_eig(T, Sigma, filename)
+% M = random('Normal', 0, 1, [N, T]);
+% C = M * M'/T;
+C = wishrnd(Sigma, T);
+lambda = eig(C);
+if exist(filename, 'file')
+    save(filename, 'lambda', '-append', '-ascii', '-double', '-tabs');
+else
+    save(filename, 'lambda', '-ascii', '-double', '-tabs');
+end

### 2014-06-11 m/test.m

@@ -1,0 +1,20 @@
+clear all
+close all
+
+company = 'nordea_bank';
+first_day = '2012-01-16';
+last_day = '2012-04-20';
+% dt in units of one minute
+% dt = 15min, delta=30sec
+% dt = 45min, delta=90sec,
+% dt = 30min, delta=60sec
+dt = 15;
+% interval for calculating realized volatility. in seconds
+delta = 30;
+s = 33;
+h = 0;
+
+[r, v] = get_intra_ret_simple(company, first_day, last_day, dt, delta);
+ep = r - mean(r);
+z = ep ./ v;
+probplot(log(v));

### 2014-08-13 thesis/cross_correlation_gaussian.tex
Finalized thesis.

@@
 \label{chp:Gaussian}
-In this chapter we present some results about the covariance
-matrix of Gaussian returns. The assumption of Gaussian returns is of
-course an over-simplification of reallity, but nevertheless lends some
-insight into the problem of elements' and eigenvalues' distributions of
-a covariance matrix.
-Our primary interest is in the influence that auto-correlations in
-returns exert on the covariance matrix. Section \ref{sec:GCC-analytical} and
-section \ref{sec:GCC-numerical} discuss how the distributions of the matrix
-elements and the eigenvalues are affected by such autocorrelations,
-respectively.
+In chapter \ref{chp:PriceModels} we have studied \gls{garch} and
+\gls{sv} models and seen their power of forecasting future
+volatilities. However, we have not considered the important fact that
+a financial market comprises many assets and the volatilities of these
+assets are correlated to each other in a complicated
+manner. Practically useful volatility forecasts require good
+understanding of these correlations.
+
+In the literature, covariance matrices of Gaussian and L\'evy
+distributed returns have been studied (see e.g. \cite{politi2010,
+  Chiani2012, Lalley2013}). However, as is seen in chapter
+\ref{chp:PriceModels}, real-world returns are not described by any
+particular distribution but rather by stochastic processes that
+account for auto-correlations in the returns and the volatilities.
+
+Therefore, the focus of this and the next chapter is on covariance
+matrices of realistic return series, and especially covariance matrices
+in the case when the return series have considerable
+auto-correlations. In particular, we study covariance matrices of
+\gls{garch} (1,1) return series in chapter
+\ref{chp:CrossCorrelationFat} and show the influence of
+auto-correlations on these matrices. But before that, it is useful to
+first understand the influence of auto-correlations on covariance
+matrices of Gaussian return series. When auto-correlations are absent,
+these matrices are called Wishart matrices and have been studied
+extensively. The results we obtain in this chapter will provide a
+reference to the studies in chapter \ref{chp:CrossCorrelationFat}.
+
+% In this chapter we present some results about the covariance
+% matrix of Gaussian returns. The assumption of Gaussian returns is of
+% course an over-simplification of reallity, but nevertheless lends some
+% insight into the problem of elements' and eigenvalues' distributions of
+% a covariance matrix.
+% Our primary interest is in the influence that auto-correlations in
+% returns exert on the covariance matrix.
+
+In section \ref{sec:GCC-analytical} we discuss how the distributions
+of the matrix elements are affected by autocorrelations, and in
+section \ref{sec:GCC-numerical} we investigate the distribution of the
+eigenvalues.
 
@@
 \end{eqnarray*}
+and $\mtx R'$ denotes the transpose of $\mtx R$.
 In words, an element $r_{it}$ of $\mtx R$ is the return of asset $i$

### 2014-08-13 thesis/price_models.tex
Finalized thesis.

@@
 The justification of modeling return series as independent,
-identically distributed Gaussian variates comes from imagining the
-price process $S(t)$ as a geometric Brownian motion whose increment
-$\sigma dw_t$ at each time step is independent and scales as $\sigma
-\sqrt{dt}$. Then, by adding a drift term $\mu dt$ that represents some
-deterministic trend in the price process, one can express the
-price $S(t)$ as a stochastic differential equation:
-\begin{eqnarray*}
-  dS &=& S\mu dt + S\sigma dw_t \\
-\end{eqnarray*}
-Then by It\^o's lemma, a stochastic differential equation for the
-logarithmic price $\ln S$ can be obtained
-\begin{eqnarray*}
-  d(\ln S) &=& (\mu - \frac{1}{2} \sigma^2)dt + \sigma dw_t \\
+identically distributed normal variates comes from imagining the
+price process $S_t$ as a stochastic process with
+  independent and normally distributed relative increments $dS_t / S_t
+  = \sigma dw_t$, where $w_t$ represents the path of a geometric
+  Brownian motion and $dw_t$ is normally distributed with mean 0 and
+  variance $dt$.
+
+By adding a drift term $\mu dt$ to allow a deterministic trend in the
+relative price change, one can express the price $S_t$ as a stochastic
+differential equation:
+\begin{eqnarray*}
+  {dS_t \over S_t} &=& \mu dt + \sigma dw_t \\
+\end{eqnarray*}
+By It\^o's lemma, a stochastic differential
+equation for the logarithmic price $\ln S_t$ can be obtained
+\begin{eqnarray*}
+  d(\ln S_t) &=& (\mu - \frac{1}{2} \sigma^2)dt + \sigma dw_t \\
+\end{eqnarray*}
+Solving this equation gives
+\begin{eqnarray*}
+  \ln S_t - \ln S_0 &=& (\mu - \frac{1}{2} \sigma^2)dt + \sigma w_t \\
 \end{eqnarray*}
@@
 }
+The quantity $\sigma_t$ introduced here is termed the volatility, and
+its meaning and significance has been discussed in the introduction
+\ref{chp:introduction}. In the model of geometric Brownian motion it
+is assumed constant and hence simply denoted $\sigma$.
+
+% The quantity $\sigma_t$ introduced here is of great importance in
+% finance and termed the volatility. Statistically, it is the
+% standard deviation of the distribution of $r_t$ and measures how
+% volatile $r_t$ is. The volatility of an asset constitutes a major
+% risk of investments on the asset and hence is very important.
 
@@
 Based on this information, we find that $y_t$ can be well
-described by a Johnson Su distribution \cite{Shang2004}:
+described by a Johnson Su distribution. This is a flexible
+distribution that features fat tails and positive skewness. It is
+defined by the following transformation \cite{Shang2004}:
 \[
@@
 model wins with a small margin. In contrast, the sample mean forecast
-clearly under-estimates the log-volatilities to a large extent --- the
-efforts of building models has not been wasted.
+clearly under-estimates the log-volatilities to a large
+extent. So the efforts of building models have indeed led
+  to more accurate forecasts.
 
@@
 forecasts, where the criterion of ``good'' is defined, respectively,
-as the forecast lying within 1\%, 5\%, or 10\% of the corresponding
-realized volatility. Table \ref{tab:nordea_2012_good} shows the
-respective percentage of the 3 kinds of forecasts. Again in this table
-it is seen that the \gls{sv} model gives more accurate forecasts than does
-\gls{garch}. For example, defining a ``good'' forecast as one that lies
-within 5\% of the realized volatility, the probability of obtaining
-such a good forecast is 75\% using the \gls{sv} model, 71\% using the
+as the forecast deviating less than 1\%, 5\%, or 10\% from the
+corresponding realized log-volatility. Table
+\ref{tab:nordea_2012_good} shows the respective percentage of the 3
+kinds of forecasts. Again in this table it is seen that the \gls{sv}
+model gives more accurate forecasts than does \gls{garch}. For
+example, defining a ``good'' forecast as one that deviates less than
+5\% from the realized volatility, the probability of obtaining such a
+good forecast is 75\% using the \gls{sv} model, 71\% using the
 \gls{garch} model, and only 53\% using the sample mean.
@@
 In the following we derive the unconditional \gls{pdf} of $r'$, the
-de-meaned and rescaled return. Denote this \gls{pdf} $f_{r'}(x)$. Then
-the \gls{pdf} of $r$ is $e^{-\bar{v}}f_{r'}(e^{-\bar{v}}x)$. First we
-consider
+de-meaned\footnote{By ``de-mean'' we mean subtracting from
+  a random variable its expectation value.} and rescaled
+return. Denote this \gls{pdf} $f_{r'}(x)$. Then the \gls{pdf} of $r$ is
+$e^{-\bar{v}}f_{r'}(e^{-\bar{v}}x)$. First we consider
 \begin{eqnarray*}
@@
 The last part of the integrand, $e^{-x^2 e^{-2v} / 2}$, is plotted in
-figure \ref{fig:DoubleExp}.
+figure \ref{fig:DoubleExp}. We see from the figure that the landscape
+defined by this function, roughly speaking, consists of two levels,
+the lower one lying at height 0 and the higher one at height 1. The
+slope connecting the two levels is rather steep. It changes as a
+double exponential along lines of constant $x$ and as $e^{-x^2}$ along
+lines of constant $v$. Therefore we approximate this function as a
+2-dimensional step function:
 \begin{figure}[htb!]
@@
 \end{figure}
-Therefore we make the following approximation:
 \[

### 2014-08-13 thesis/thesis.tex
Finalized thesis.

@@
 \usepackage{amssymb}
-% \usepackage{amsfonts, amsmath}
+\usepackage[usenames,dvipsnames]{color}
 \usepackage{amsmath}
@@
 
-% \title{Stocks' Cross-Correlations}
-% \subtitle{}
 \makeglossaries
@@
 to intraday return series. \gls{sv} models are found to be considerably more
-accurate and more consistent in accuracy in forecasting.
+accurate and more consistent in accuracy of forecasting.
 
@@
 is found to approximately follow a gamma distribution also when the
-returns are auto-correlated. Expressions relating the mean and the
-variance of the asymptotic Gaussian distribution of the matrix
-elements are derived. In the case of \gls{garch} returns, both the largest
-and the smallest eigenvalues of the covariance matrix are seen to
-increase with increasing auto-correlation. The matrix elements are
-found to follow L\'evy distributions with different L\'evy indices for
-the diagonal and the non-diagonal elements.
+returns are auto-correlated. Expressions relating auto-correlation
+strength with the mean and the variance of the asymptotic Gaussian
+distribution of the matrix elements are derived. In the case of
+\gls{garch} returns, both the largest and the smallest eigenvalues of
+the covariance matrix are seen to increase with increasing
+auto-correlation. The matrix elements are found to follow L\'evy
+distributions with different L\'evy indices for the diagonal and the
+non-diagonal elements.
 
@@
 AR(1) process.
+
+%% Popular Abstract
+% Can we predict the movements of the financial market?
+The direct answer is "no", but there is more to be said.
+One cannot predict the price movement in an efficient market, but one can indeed predict how "volatile" the prices are (volatility).
+This does not lead to profits, but is useful for risk management --- if one pursues a profit, one must take a risk.
+Knowing and quantitatively managing the risks associated with investments, one often makes wiser decisions.
+
+% This report entails, first of all, a comparison of return models in the context of intraday returns.
+6 return series are studied: Nordea, Volvo and Ericsson in the time scales of 15 and 30 minutes.
+Stochastic Volatility (SV) models and GARCH models are fitted to these series and the resulting volatility forecasts are compared.
+SV models are found to be more accurate in most cases.
+
+% Covariance matrices, which give the correlation between different companies, are also studied.
+In particular, the eigenvectors and eigenvalues of a covariance matrix, which respectively represent a set of driving factors of the companies' shares and the variances of these factors, are studied.
+It is found that when the return series are auto-correlated, i.e.
+when observations at different times are correlated, the eigenvectors are delocalized --- their expansion as a linear combination of the basis vectors involve more appreciable coefficients.
 \end{abstract}
@@
 the people in the division of mathematical physics. The casual
-discussions that I had with them truely inspired me, and the friendly
-and lively atmosphere that they create has helped bring peace and clarity
-to my mind.
+discussions that I had with them truly inspired me, and
+the friendly and lively atmosphere that they create has helped bring
+peace and clarity to my mind.
 
@@
 
-\chapter{Introduction}
+\chapter{Introduction} \label{chp:introduction}
 The mathematics of the financial market has always been a topic that
@@
 
-\Gls{sv} models built on historical data describe the time evolution
-of the aforementioned observables in terms of probability
-distributions, autocorrelations and cross correlations. They may have
-some predictive power and help determine a fair price of a given
-asset.
-
 Quite often, instead of an asset's price, the relative price change,
 i.e. the return, is studied. Analytically solvable models often assume
-Gaussian return distribution, although data suggest fatter tails. In
-numerical return models, realistic reproduction of historical data may
-be achieved.
+Gaussian return distribution, although data suggest fatter tails
+\footnotemark. By numerical methods, statistical features of
+historical returns data, such as fat tails, can be reproduced.
+\footnotetext{
+  While the exact nature of the tails of the distribution of
+  historical returns $r$ is debatable, it is agreed $P(|r| > x) > a
+  e^{-bx^2}$, for any positive constants $a$ and $b$. Here $P(\cdot)$
+  denotes probability. We adopt this as the definition of ``fat''
+  tails.
+}
+
+Statistical models built on historical data describe the time
+evolution of the aforementioned observables in terms of probability
+distributions, autocorrelations and cross correlations. They have been
+used to forecast future volatilities and help determine a pair price
+of a contingent claim, e.g. an option. Here the term ``volatility''
+refers to the standard deviation of the distribution of returns, and
+measures how volatile the returns are. The volatility of an asset
+constitutes a major risk of investments on the asset and hence is very
+important.
 
@@
 \gls{garch}(1,1) series are studied. Finally, chapter \ref{chp:summary}
-summarizes the results. Supplementary materials are provided in the
+summarizes the results. Supplementary material is provided in the
 appendices.
@@
 models. In addition, a study of continuous-time models gave me an
-opportunity to increase my knowledge of stochastic differential
-equations, probability theory, and statistics. I am a beginner, but
-nonetheless find it comfortable to read books and articles on these
-subjects after the thesis project.
+opportunity to increase my knowledge of stochastic processes,
+probability theory and statistics.
 
@@
 when I worked on the covariance matrix of returns described by
-Gaussian or \gls{garch} models. I am still a long way from being able
-to do serious proofs or derivations in these areas, but I know what to
-read and can understand the basics.
+Gaussian or \gls{garch} models.
 

### 2014-08-13 m/garch_tail_exponent.m

@@ -1,15 +1,18 @@
 function k = garch_tail_exponent(dist, param, T);
+% function k = garch_tail_exponent(param);
 
-if param(1) <= 0 || (param(2) + param(3) >= 1)
-    k = NaN;
-    return;
-end
+a1 = param(1);
+b1 = param(2);
+
 if strcmp(dist.name, 'Gaussian') == 1
     R = randn(1, T);
 elseif strcmp(dist.name, 'Student t') == 1
     R = trnd(dist.DoF, 1, T);
 end
-
+if param(1) <= 0
+    k = NaN;
+    return;
+end
 A = param(2) .* R.^2 + param(3);
 s = mean(log(A));
 
@@ -24,5 +27,4 @@ while mean(A.^a) < 1
 end
 
 f = @(k) mean(A.^(k/2)) - 1;
-options = optimoptions('fsolve','Display','off');
-k = fsolve(f, 2*a, options);
+k = fsolve(f, 2*a);

### 2014-08-13 m/gen_ret_mtx.m

@@ -43,40 +43,29 @@ switch (distr.name)
     a0 = distr.prmt(1);
     a1 = distr.prmt(2);
     b1 = distr.prmt(3);
-
-    if strcmp(distr.distr.Name, 'Gaussian') == 1
-        h = [ones(n, 1)*(a0/(1 - a1 - b1)), NaN(n, T-1)];
-        if isempty(varargin)
-            x = [randn(n, 1), NaN(n, T-1)];
-            for t = 2:T
-                h(:, t) = a0 + a1.*x(:, t-1).^2 + b1.*h(:, t-1);
-                x(:, t) = randn(n, 1).*sqrt(h(:, t));
-            end
-        else
-            Sigma = varargin{1};
-            x = [mvnrnd(zeros(n, 1), Sigma, 1)', NaN(n, T-1)];
-            for t = 2:T
-                h(:, t) = a0 + a1.*x(:, t-1).^2 + b1.*h(:, t-1);
-                x(:, t) =mvnrnd(zeros(n, 1), Sigma, 1)'.*sqrt(h(:, t));
-            end
-        end
-
-        % elseif strcmp(distr.distr.Name, 't') == 1
-    %     d = distr.distr.DoF;
-    %     s = sqrt(d / (d - 2));
-
-    %     h = [ones(n, 1), NaN(n, T-1)];
-    %     x = [trnd(d, n, T) ./ s, NaN(n, T-1)];
-    %     for t = 2:T
-    %         h(:, t) = a0 + a1.*x(:, t-1).^2 + b1.*h(:, t-1);
-    %         x(:, t) = trnd(n, 1) ./ s .* sqrt(h(:, t));
-    %     end
+    if a0 > 0 && (a1 + b1 < 1)
+        h0 = a0/(1 - a1 - b1);
+    else
+        h0 = 1;
     end
-
-    % model = garch('Constant', a0, 'ARCH', a1, 'GARCH', ...
-    %               b1, 'Distribution', distr.distr);
-    % x = simulate(model, T, 'numPaths', n)';
+    h = [ones(n, 1)*h0, NaN(n, T-1)];
+    x = [randn(n, 1).*sqrt(h0), NaN(n, T-1)];
+    for t = 2:T
+        h(:, t) = a0 + a1.*x(:, t-1).^2 + b1.*h(:, t-1);
[…]

### 2014-08-13 m/get_intra_ret_simple.m

@@ -23,6 +23,8 @@ function [ret, sig] = get_intra_ret_simple(...
     end
     ret = ret(1: n);
 
+    return;
+    
     if isnan(delta) || isinf(abs(delta))
         return;
     end

### 2014-08-13 m/test1.m

@@ -1,17 +1,14 @@
 clear all
-dist = struct('name', 'Gaussian');
-k = NaN(18*19, 1);
-c = 1;
-a0 = 0.01;
-for a1 = 0.05:0.05:0.9
-    n = 10;
-    db = (1 - a1)/n;
-    for b1 = 0:0.05:0.9
-        param = [0.01, a1, b1];
-        k(c) = garch_tail_exponent(dist, param, 1e+6);
-        c = c + 1;
-    end
-end
-save('garch_tail_exponent.txt', 'k', '-ascii');
+close all
 
+N = 250;
+q = 24;
+T = N * q;
+a = 0.94;
 
+load(sprintf(['../data/IGarchWishartN%dQ%dA%.4fEig.mat'], ...
+             N, q, a), 'ev');
+
+ev = reshape(ev, 1, prod(size(ev)));
+[x1, y1] = epdf(ev, 4, min(ev), max(ev), 200, '');
+plot(x1, y1);

### 2014-09-29 thesis/NordeaVolvo.tex
Add the draft of the paper.

\chapter{Nordea in 2013/14}
\label{chp:Nordea2}
In this chapter we model the volatility of {\it Nordea Bank}
15-minute and 30-minute returns during the period 2013/10/10 -
2014/04/04.

\section{15-minute Returns}\label{sec:nordea2_15min}
This return series contains 3879 data points. The squared
volatilities, i.e. the variances, are computed as the sum of squared
50-second returns. The normal probability plot of the quotient $(r_t - \E(r_t))/
\sigma_t$, is shown in figure \ref{fig:nordea2_b}. We see in this figure that
$(r_t - \E(r_t)) / \sigma_t$ is approximately Gaussian, confirming the choice of
50-second returns for the computation of the realized
volatility. Among the 3879 data points, we use the first 80\% for
model estimation and the last 20\%, i.e. 776 data points for forecast
verification.

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=79 259 108
  121]{../pics/nordea2_b.pdf}
  \caption{\small \it Normal probability plot of $(r_t - \E(r_t))/
    \sigma_t$. $\sigma_t^2$ is computed as the sum of squared
    50-second returns.}
  \label{fig:nordea2_b}
\end{figure}

The log-volatility series $\ln \sigma_t$ in this case has an
auto-correlation structure similar to its predecessor studied in
chapter \ref{chp:nordea_15min}. The seasonality is 33 as before. We
show the auto-correlations of $\ln \sigma_t$ in figure
\ref{fig:nordea2_15_lv_acf}, and those of $w_t = (1 - B)(1-B^s)\ln
\sigma_t$ in figure \ref{fig:nordea2_15_w_acf}. The reader may want to
compare them with figure \ref{fig:nordea1_15min_acf}.

\begin{figure}[htb!]
  \centering
  \subfigure[]{
  \includegraphics[scale=0.4, clip=true, trim=86 254 98
  123]{../pics/nordea2_15min_lv_acf.pdf}
  \label{fig:nordea2_15_lv_acf}
  }
  \subfigure[]{
  \includegraphics[scale=0.4, clip=true, trim=95 230 112
  235]{../pics/nordea2_w_acf.pdf}
  \label{fig:nordea2_15_w_acf}
  }
  \caption{\small \it Auto-correlation function of $\ln \sigma_t$ and
    $w_t = (1 - B)(1-B^s)\ln \sigma_t$} 
\end{figure}
The significant auto-correlations of $w_t$ and their respective lags are listed
in table \ref{tab:nordea2_w_acf}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    lag & 1 & 32 & 33 & 34 \\
    \hline
    ACF &   -0.4805 & 0.2861 &  -0.5034 & 0.2093 \\
    \hline
  \end{tabular}
  \caption{\small \it Significant auto-correlations of $w_t$}
  \label{tab:nordea2_w_acf}
\end{table}
Although the log-volatilities have the seasonality $s = 33$, judging
from the relatively big differences between the auto-correlations at
lags 32 and 34, we don't expect the moving average process to be in the
multiplicative form of equation \ref{eq:nordea_w}. Instead, we model
$\ln\sigma_t$ as
\begin{eqnarray}
  (1 - B)(1 - B^s) \ln\sigma_t &=& (1 - \theta_1 B - \theta_s B^s -
  \theta_{s+1} B^{s+1} ) y_t \label{eq:nordea2_v}
\end{eqnarray}
where $y_t$ is assumed to be normally distributed, neglecting slight
non-normality. By maximum likelihood estimation, we fit the above
model to the realized volatilities. The resulting parameter values are
listed in table \ref{tab:nordea2_15_SV}.

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    parameter & $\theta_1$ & $\theta_{33}$ & $\theta_{34}$ & $\var(y_t)$ \\
    \hline
    value & 0.8339 & 0.8448 & -0.7067 & 0.10071 \\
    \hline
  \end{tabular}
  \caption{\small \it: Parameter values of $\ln \sigma_t$}
  \label{tab:nordea2_15_SV}
\end{table}

The auto-correlations of the residuals $y_t$ are insignificant, as
shown in figure \ref{fig:nordea2_y}. Though the Ljung-Box Q
test rejects the null hypothesis of no autocorrelation in the
residuals, we rather neglect these weak auto-correlations in favor of
a relatively simple model.

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=86 256 101
  121]{../pics/nordea2_y_acf.pdf}
  \caption{\small \it Auto-correlations of the residuals ($y_t$).}
  \label{fig:nordea2_y}
\end{figure}

Forecast of $\ln\sigma_t$ is done by reverting equation
\ref{eq:nordea2_v} and write
\begin{eqnarray*}
  \ln\sigma_t &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
  \ln\sigma_{t-s-1} + (1 - \theta_1 B - \theta_s B^s + \theta_{s+1}
  B^{s+1} ) y_t \\
  \E(\ln\sigma_t | \mathcal{F}_{t-1}) &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
  \ln\sigma_{t-s-1} - \theta_1 y_{t-1} - \theta_s y_{t-s} +
  \theta_{s+1} y_{t-s-1}
\end{eqnarray*}

Thus the forecast, namely $\E(\ln\sigma_t | \mathcal{F}_{t-1})$, is
found in terms of previous values in the same series and of previous
residuals. To compare the performance, we also estimate a GARCH model with
Gaussian conditional distribution. The procedure is standard and we
only list the results here
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_5$ & $\beta_1$ \\
    \hline
    Value & $7.3686 \times 10^{-7}$ & 0.154131 & 0.0360 & 0.491832 \\
    \hline
  \end{tabular}
  \caption{\small \it Parameters of GARCH(1, 5) model of Nordea
    Bank 15-minute returns}
  \label{tab:nordea2_garch}
\end{table}

Forecasting using both models, we get two series of one-step-ahead
forecast, $\ln \sigma^{\text{SV}}_t$ and $\ln \sigma^G_t$,
respectively. We denote the empirical log-volatilities as $\ln
\hat{\sigma}_t$. Then the accuracy of the two models can be compared
by comparing $\ln \sigma^{\text{SV}}_t - \ln \hat{\sigma}_t$ and
$\ln \sigma^G_t - \ln \hat{\sigma}_t$. As a reference, we also include
in the comparison the naive approach of taking the sample mean of
$\ln \sigma_t$ as the forecast.

Table \ref{tab:slv_garch_comparison} shows the mean and the standard
deviation of $\ln \sigma^F_t - \ln \hat{\sigma}_t$ for the 3 kinds of
forecast. Here we use $\ln \sigma^F_t$ to denote the forecast of $\ln
\sigma_t$. Clearly the SV model has the smallest mean and standard
deviation for $\ln \sigma^F_t - \ln \hat{\sigma}_t$, the sample mean
has the largest, and the GARCH model sits in between.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & SV & GARCH & Sample mean \\
    \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & -0.0033 & 0.0356 &
    0.0544 \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ &
    0.2876  &  0.3096 &   0.3349 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:slv_garch_comparison}
\end{table}
Even more informative about the quality of forecast is the
distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$, which we show
in figure \ref{fig:slv_garch_comparison}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=0 292 0
  255]{../pics/slv_garch_comparison.pdf}
  \caption{\small \it Comparison of the forecasts from SV, GARCH,
    and sample mean. Blue: SV forecasts; Green: GARCH forecasts; Red:
    Sample mean forecasts.}
  \label{fig:slv_garch_comparison}
\end{figure}
The left plot in the figure shows the situation of under-estimates,
i.e. $\ln \sigma^F_t - \ln \hat{\sigma}_t < 0$. We see that, to the
far left, the blue line is below the green, which is in turn below the
red, meaning the SV forecast has the lowest probability of
under-estimate. The GARCH forecast comes second and the sample
mean forecast under-estimates most.

The same ranking of forecast quality is also seen on the right plot of
figure \ref{fig:slv_garch_comparison}, which shows the complementary
distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$ for postive
values, i.e. over-estimates: The SV forecast has the lowest
probability of over-estimate; GARCH comes second and the sample mean
is most likely to over-estimate.

As another measure of the quality of forecast, the fraction of
``good'' forecasts that differ no more than a given percentage from
the measured realized volatilities is also computed. Table
\ref{tab:nordea_good_percentage} compares this fraction of good
forecasts from the 3 alternatives with respect to different
criteria of goodness.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    SV & GARCH & Sample Mean \\
    \hline
    1\% & 0.1920 & 0.2010 & 0.1843 \\
    \hline
    5\% & 0.7887 & 0.7436 & 0.7229 \\
    \hline
    10\% & 0.9665 & 0.9601 & 0.9369 \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:nordea_good_percentage}
\end{table}
Clearly, although SV and GARCH rank differently with respect to
different criteria of goodness, they both out-perform the sample
mean forecast.

\section{30-minute Returns}\label{sec:nordea2_30min}
In this section we study the volatility of Nordea Bank 30-minute
returns in the period 2013/10/10 - 2014/04/04. The realized
volatilities that approximate the volatilities of these 30-minute
returns are computed using 1-minute returns. As in the previous cases,
this choice of 1-minute returns is confirmed by the normality of the
quotient $(r_t - \E(r_t))/\sigma_t$, which is shown in figure
\ref{fig:nordea3_quotient}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=78 255 108
  120]{../pics/nordea3_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -
    \E(r_t))/\sigma_t$}
  \label{fig:nordea3_quotient}
\end{figure}
The auto-correlations of $\ln \sigma_t$ is shown in figure
\ref{fig:nordea3_lv_acf}, where we see an apparent seasonality of
16. By differencing we get $w_t = (1-B)(1-B^s)\ln\sigma_t$ where
$s=16$. Its auto-correlations are shown in figure
\ref{fig:nordea3_w_acf}.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=86 256 101
    121]{../pics/nordea3_lv_acf.pdf}
    \label{fig:nordea3_lv_acf}
  }
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=86 256 103
    122]{../pics/nordea3_w_acf.pdf}
    \label{fig:nordea3_w_acf}
  }
  \caption{\small \it \ref{fig:nordea3_lv_acf}: Auto-correlations
    (ACF) of $\ln \sigma_t$; \ref{fig:nordea3_w_acf}: Auto-correlations
    (ACF) of $w_t = (1-B)(1-B^s)\ln\sigma_t$.}
\end{figure}
Once again, this auto-correlation structure
points to a seasonal moving average model:
\begin{eqnarray*}
  w_t &=& (1-\theta B)(1-\Theta B^s) y_t
\end{eqnarray*}
where $y_t$ is assumed to have Gaussian distribution, neglecting
slight non-normality as before. Maximum likelihood estimation gives
the parameter values listed in table \ref{tab:nordea3_sv_param}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\theta$ & $\Theta$ & $\var(y_t)$\\
    \hline
    Value & 0.7612 & 0.8036 & 0.0736\\
    \hline
  \end{tabular}
  \caption{\small \it Parameter values of the Seasonal Moving Average
    model}
  \label{tab:nordea3_sv_param}
\end{table}
To fit a GARCH model to the same series, we assume Gaussian
innovations (c.f. equation \ref{eq:garch_def}). The result is a
GARCH(1,1) model. Its parameters are listed in table
\ref{tab:nordea3_garch_param}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\beta_1$ \\
    \hline
    Value & $2.3 \times 10^{-7}$ & 0.05 & 0.9\\
    \hline
  \end{tabular}
  \caption{\small \it Parameter values of GARCH(1,1) model fitted to
    Nordea 30-minute returns.}
  \label{tab:nordea3_garch_param}
\end{table}

As before, we compare the accuracies of the GARCH and the stochastic
volatility model by comparing their one-ste-ahead forecasts. We
compute the difference between their forecasts and the realized
volatilities and then look at the statistics of these difference
values. Firstly we show the mean and the standard deviation of the
differences in table \ref{tab:nordea3_diff_1}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
     & SV & GARCH & Sample mean \\
     \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & -0.0047 & 0.0130 & 0.0584 \\
    \hline
     $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.2602 &
     0.3194 & 0.3093 \\
    \hline
 \end{tabular}
  \caption{\small \it Mean and standard deviation of the 3 kinds of
    forecasts}
  \label{tab:nordea3_diff_1}
\end{table}
It is seen in this table that the SV forecasts have a mean whose
absolute value is just above 1/3 of that of the GARCH forecasts. The
standard deviation of the SV forecasts is also smaller than that of
GARCH. In addition, we check the distribution of
$\ln \sigma^F_t - \ln \hat{\sigma}_t$, and the percentage of ``good''
forecasts with respect to different criteria of goodness. These are
shown in figure \ref{fig:nordea3_diff} and table
\ref{tab:nordea3_diff_2}, respectively.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.54, clip=true, trim=21 312 0
    179]{../pics/nordea3_diff.pdf}
  \caption{\small \it Comparison of the forecasts from SV, GARCH,
    and sample mean. Blue: SV forecasts; Green: GARCH forecasts; Red:
    Sample mean forecasts.}
  \label{fig:nordea3_diff}
\end{figure}
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    SV & GARCH & Sample Mean \\
    \hline
    1\% & 0.1777 & 0.1777 & 0.1538 \\
    \hline
    5\% & 0.8090 & 0.7109 & 0.6923 \\
    \hline
    10\% & 0.9814 & 0.9310 & 0.9523 \\
    \hline
  \end{tabular}
  \caption{\small \it Percentage of ``good'' forecasts as defined by
    deviating nore more than 1\%, 5\% and 10\% from the measured
    realized volatility.}
  \label{tab:nordea3_diff_2}
\end{table}
It is clear that the SV forecasts are considerably more accurate.

\chapter{Volvo in 2013/14}\label{sec:volvo}
In this chapter we study the volatility of Volvo B during the period
2013/10/10 - 2014/04/04. Section \ref{sec:volvo_15} and section \ref{sec:volvo_30}
investigate the 15-minute and 30-minute returns respectively.

\section{15-minute Returns}\label{sec:volvo_15}
In this section we look at the volatility of Volvo 15-minute
returns. The subject of modeling is the realized volatility, which is
computed as the square root of the sum of squared 1-minute
returns. The normal probability plot of the quotient $(r_t -
\E(r_t))/\sigma_t$ is shown in figure \ref{fig:volvo_15_quotient},
from which one can see it is normally distributed, confirming the
choice of 1-minute returns for computing the realized volatility.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=83 259 110 
  220]{../pics/volvo_15_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -
    \E(r_t))/\sigma_t$}
  \label{fig:volvo_15_quotient}
\end{figure}
The auto-correlations of $\ln \sigma_t$ and $w_t = (1-B)(1-B^s) \ln
\sigma_t$ are plotted in figure \ref{fig:volvo_15_lv_acf} and
\ref{fig:volvo_15_w_acf}. The former clearly shows the seasonality
$s = 33$ in the auto-correlations of $\ln \sigma_t$.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
  \includegraphics[scale=0.4, clip=true, trim=90 259 103
  220]{../pics/volvo_15_lv_acf.pdf}
  \label{fig:volvo_15_lv_acf}
  }
  \subfigure[]{
  \includegraphics[scale=0.4, clip=true, trim=90 259 103
  220]{../pics/volvo_15_w_acf.pdf}
  \label{fig:volvo_15_w_acf}
  }
  \caption{\small \it Left: auto-correlations of $\ln \sigma_t$;
    Right: auto-correlations of $w_t = (1-B)(1-B^s) \ln \sigma_t$.}
\end{figure}
Based on the auto-correlations of $w_t$, a seasonal moving average
model is estimated:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1-\theta_1B - \theta_2B^2 - \theta_3
  B^3)(1 - \Theta B^s)y_t
\end{eqnarray*}
By maximum likelihood estimation, the parameter values listed in table
\ref{tab:volvo_15_sv_param} are obtained.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|}
  \hline
  Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\Theta$ & 
 $\var(y_t)$ \\
 \hline
 Value & 0.7491 & 0.1170 & 0.0571 & 0.8646 & 0.1245 \\
  \hline
  \end{tabular}
  \caption{\small \it Parameter values of the stochastoc volatility
    (SV) model.}
  \label{tab:volvo_15_sv_param}
\end{table}

A GARCH model (c.f. equation \ref{eq:garch_def}) assuming Gaussian
innovations is also fitted to the same series. The parameter values
are listed in table \ref{tab:volvo_15_garch_param}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
  Parameter & $\alpha_0$ & $\alpha_1$ & $\beta_{1}$ \\
  \hline
  Value & $6.54 \times 10^{-7}$ & 0.1565 & 0.6445\\
  \hline
  \end{tabular}
  \caption{\small \it Parameter values of the GARCH model.}
  \label{tab:volvo_15_garch_param}
\end{table}

Performing one-step-ahead forecasts using both models gives the series
$\ln\sigma^F_t - \ln\hat{\sigma}_t$, where $\ln\hat{\sigma}_t$ are the
measured realized volatilities. The mean and standard deviation of
$\ln\sigma^F_t - \ln\hat{\sigma}_t$ are shown in table
\ref{tab:volvo_15_stat}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
     & SV & GARCH & Sample mean \\
     \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0035 & 0.0313 & -0.1448 \\
    \hline
     $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3356 &
     0.3894 & 0.3826 \\
    \hline
 \end{tabular}
  \caption{\small \it Mean and standard deviation of the 3 kinds of
    forecasts}
  \label{tab:volvo_15_stat}
\end{table}
We see that both the SV and the GARCH model give biased forecasts, but
the SV forecasts are only 1/10 as biased as those of GARCH (0.0035
vs. 0.0313). In addition, the standard deviation of the SV forecasts
is smaller too. These results are confirmed by the distribution of
$\ln \sigma^F_t - \ln \hat{\sigma}_t$ and the fraction of ``good''
forecasts, which are shown in figure \ref{fig:volvo_15_diff} and table
\ref{tab:volvo_15_diff}, respectively.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.54, clip=true, trim=27 276 0
    236]{../pics/volvo_15_diff.pdf}
  \caption{\small \it Comparison of the forecasts from SV, GARCH,
    and sample mean. Blue: SV forecasts; Green: GARCH forecasts; Red:
    Sample mean forecasts.}
  \label{fig:volvo_15_diff}
\end{figure}
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    SV & GARCH & Sample Mean \\
    \hline
    1\% & 0.5760 & 0.5979 & 0.3930 \\
    \hline
    5\% & 0.8351 & 0.8363 & 0.6611 \\
    \hline
    10\% & 0.9716 & 0.9536 & 0.9085 \\
    \hline
  \end{tabular}
  \caption{\small \it Percentage of ``good'' forecasts as defined by
    deviating nore more than 1\%, 5\% and 10\% from the measured
    realized volatility.}
  \label{tab:volvo_15_diff}
\end{table}

\section{30-minute Returns}\label{sec:volvo_30}
In this section we model the log-volatility of {\it Volvo B} 30-minute
returns during the period 2013/10/10 - 2014/04/04. This series
contains 1884 log-volatilities computed using 2-minute returns in each
30-minute interval. Among them we use the first 1507 for model
estimation and the last 377 for forecast and model verification.

The left plot of figure \ref{fig:volvo_inno_and_lv_acf} shows the
distribution of $(r_t - \E(r_t))/\sigma_t$. We observe in the figure a
nice Gaussian variate, so we can be sure that the sum of squared
2-minute returns makes a good approximation to the variance of
30-minute returns in this particular case.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.4, clip=true, trim=0 256 0
    151]{../pics/volvo_inno_and_lv_acf.pdf}
    \label{fig:volvo_inno_and_lv_acf}
  \caption{\small \it Left: probability plot of $(r_t -
    \E(r_t))/\sigma_t$; Right: auto-correlations of $\ln \sigma_t$}
\end{figure}

Guided by the auto-correlations of $\ln\sigma_t$ shown in the right
plot of figure \ref{fig:volvo_inno_and_lv_acf} we find the following
model:
\begin{eqnarray}
  && (1 - B)(1 - B^s) \ln \sigma_t \nonumber \\
  &=& (1 - \theta_1 B - \theta_2B^2 -
  \theta_3B^3 - \theta_4B^4) (1 - \Theta B^s)
  y_t \label{eq:volvo_lv_model}
\end{eqnarray}
where $s = 16$ is the seasonality and is apparent from the
auto-correlations of $\ln\sigma_t$. Fitting this model to the
measured realized volatilities yields the parameter values listed in
table
\ref{tab:volvo_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ &
    $\Theta$ & residual variance \\
    \hline
    Value & 0.7305 & 0.0575 & 0.0574 & 0.0346 & 0.8324 & 0.1340\\
    \hline
  \end{tabular}
  \caption{\small \it Volvo B log-volatility parameters}
  \label{tab:volvo_params}
\end{table}
Forecasting using the estimated model parameters gives the forecast
series $\ln \sigma^{\text{SV}}_t$. To access the quality of the
forecast, we also estimate a GARCH model using the returns. The result
is a GARCH(1, 1) model, whose parameter values are listed in table
\ref{tab:volvo_garch}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$  & $\beta_1$ \\
    \hline
    Value & $3.125 \times 10^{-7}$ & 0.05 & 0.90 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH(1, 1) model of Volvo B 30-minute returns}
  \label{tab:volvo_garch}
\end{table}

The forecasts from SV, GARCH, and the sample mean are compared
using the difference $\ln \sigma^F_t - \ln\hat{\sigma}_t$, where $\ln
\sigma^F_t$ stands for the forecast. The distributions of this
difference is plotted in figure \ref{fig:volvo_slv_garch_cmp}; the
mean and the standard deviation of the distributions are listed in
table \ref{tab:volvo_slv_garch_cmp}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & SV & GARCH & Sample mean \\
    \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & -0.0123 &
    0.0242 & -0.1055 \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3261 &
    0.4250 & 0.3708 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:volvo_slv_garch_cmp}
\end{table}

\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=10 282 0
    243]{../pics/volvo_slv_garch_cmp.pdf}
    \label{fig:volvo_slv_garch_cmp}
  \caption{\small \it Comparison of the forecast from SV, GARCH,
    and sample mean. Blue: SV forecast; Green: GARCH forecast; Red:
    Sample mean forecast.}
\end{figure}
Figure \ref{fig:volvo_slv_garch_cmp} shows, as in the previous case of
Nordea Bank 15-minute returns, the SV model performs the best, GARCH
the second, and the sample mean the worst. However, when it comes to
over-estimates, the sample mean appears to be the best estimator,
while SV excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
the sample for estimation (1507 data points) and over the sample for
comparison (377 data points) reveals that the first sample has mean
-6.3127 while the second has -6.2072. This increment explains the low
probability of over-estimation when using the first sample mean as
forecast.

Table \ref{tab:volvo_good_percentage} compares the fraction of
``good'' estimates as measured by ${|\ln\sigma^F_t -
  \ln\hat{\sigma}_t| \over |\ln \hat{\sigma}_t|}$ being less than 1\%,
5\% and 10\%.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    SV & GARCH & Sample Mean \\
    \hline
    1\% & 0.2175 & 0.1220 & 0.1406 \\
    \hline
    5\% & 0.7162 & 0.6154 & 0.6631 \\
    \hline
    10\% & 0.9231 & 0.8806 & 0.8992 \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:volvo_good_percentage}
\end{table}
We see from the table that the SV model consistently excels over
the other two alternatives.

\chapter{Ericsson in 2013/14}
\label{chp:ericsson}
In this chapter we model the volatility of {\it Ericsson B}
15-minute and 30-minute returns during the period 2013/10/10 -
2014/04/04.

\section{15-minute Returns}\label{sec:ericsson_15min}
Using the same method as before, we first find the right
higher-frequency returns that best approximate the volatility of the
15-minute returns. These turn out to be 50-second returns, as figure
\ref{fig:ericsson_15_quotient} shows.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=79 259 108
  121]{../pics/ericsson_15_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -\E(r_t))/
    \sigma_t$. $\sigma_t^2$ is computed as the sum of squared
    50-second returns.}
  \label{fig:ericsson_15_quotient}
\end{figure}

Following the procedure described in previous sections, an ARIMA model is
found for this series:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1- \theta_1B - \theta_2B^2 -
  \theta_3B^3)(1 - \Theta B^s) y_t
\end{eqnarray*}
where the seasonality $s$ is 33 and the other parameter values are
estimated to be those listed in table \ref{tab:ericsson_15_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\Theta$ &
    $\var(y_t)$ \\
    \hline
    Value & 0.8078 & 0.0454 & 0.0943 & 0.8798 & 0.1242 \\
    \hline
  \end{tabular}
  \caption{\small \it Ericsson B log-volatility parameters}
  \label{tab:ericsson_15_params}
\end{table}

Then a GARCH model is also found with parameter values listed in
table \ref{tab:ericsson_15_garch_params}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_{s}$ & $\beta_{s}$ \\
    \hline
    Value & $2.4181 \times 10^{-7}$ & 0.1513 & 0.1409 & 0.6027 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH model of Volvo B 30-minute returns}
  \label{tab:ericsson_15_garch_params}
\end{table}

The forecasts from both models are compared as follows: Table
\ref{tab:ericsson_15_diff1} shows the mean and standard deviation of
the 3 kinds of forecasts:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & SV & GARCH & Sample mean \\
    \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0035 &
    0.0600 & 0.1116 \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3278 &
    0.3667 & 0.3590 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:ericsson_15_diff1}
\end{table}
Consistent with previous results, the bias introduced by the SV model
is significantly smaller than that introduced by GARCH. Figure
\ref{fig:ericsson_15_diff2} shows the distribution of the difference
between a forecast and its corresponding measured realized volatility,
i.e. $\ln\sigma^F_t - \ln\hat{\sigma}_t$; table
\ref{tab:ericsson_15_diff3} compares the percentage of ``good''
forecasts using the 3 alternatives.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=45 298 21
    165]{../pics/ericsson_15_diff2.pdf}
  \caption{\small \it Comparison of the forecast from SV, GARCH,
    and sample mean. Blue: SV forecast; Green: GARCH forecast; Red:
    Sample mean forecast.}
  \label{fig:ericsson_15_diff2}
\end{figure}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    SV & GARCH & Sample Mean \\
    \hline
    1\% & 0.1673 & 0.1673 & 0.1634 \\
    \hline
    5\% & 0.7268 & 0.6850 & 0.6941 \\
    \hline
    10\% & 0.9621 & 0.9307 & 0.9399 \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:ericsson_15_diff3}
\end{table}
It is clear from figure \ref{fig:ericsson_15_diff2} and table
\ref{tab:ericsson_15_diff3} that the SV model out-performs GARCH. We
see that the SV model yields considerably higher fractions of good
forecasts by all criteria, and in particular, gives much few
under-estimates.

\section{30-minute Returns}\label{sec:ericsson_30min}
In this section we look at the 30-minute returns of Ericsson B during
the period 2013/10/10 - 2014/04/04. Since the methods are the same as
with other series, we shall only present the results here.

First of all, the volatility of this series is found to be well
approximated by realized volatilities computed from 75-second
returns. The normal probability plot is shown in figure
\ref{fig:ericsson_30_quotient}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=79 259 108
  121]{../pics/ericsson_30_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -\E(r_t))/
    \sigma_t$. $\sigma_t^2$ is computed as the sum of squared
    75-second returns.}
  \label{fig:ericsson_30_quotient}
\end{figure}
The following ARIMA model is found for the log-volatility $\ln
\sigma_t$:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1- \theta_1B - \theta_2B^2) (1 -
  \Theta B^s) y_t 
\end{eqnarray*}
where the seasonality $s = 16$. The parameter values are estimated to
be those listed in table
\ref{tab:ericsson_30_params}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\Theta$ &
    $\var(y_t)$ \\
    \hline
    Value & 0.6842 & 0.2470 & 0.8391 & 0.0918 \\
    \hline
  \end{tabular}
  \caption{\small \it Ericsson B log-volatility parameters}
  \label{tab:ericsson_30_params}
\end{table}
A GARCH model is also found with parameters listed in table
\ref{tab:ericsson_30_garch_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_{s}$ & $\beta_{s}$ \\
    \hline
    Value & $7.4856 \times 10^{-7}$ & 0.0532 & 0.1506 & 0.6778 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH(1, 1) model of Ericsson B 30-minute
    returns}
  \label{tab:ericsson_30_garch_params}
\end{table}

Table \ref{tab:ericsson_30_diff1} shows the mean and standard
deviation of the difference series $\ln\sigma^F_t -
\ln\hat{\sigma}_t$, where $\ln\sigma^F_t$ is the forecast
log-volatility and $\ln\hat{\sigma}_t$ is the measured realized
volatility.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & SV & GARCH & Sample mean \\
    \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0004 &
    0.0526 & 0.1039 \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.2519 &
    0.3019 & 0.2963 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:ericsson_30_diff1}
\end{table}
Apparently, the GARCH forecasts have a fairly large bias compared with
those of the SV model. Turning to figure \ref{fig:ericsson_30_diff2}
and table \ref{tab:ericsson_30_diff3}, we also see the SV model
performs consistently better --- a higher percentage of ``good''
forecasts are delivered and, while it under-estimates to around the
same extent as does GARCH, it certainly over-estimates a lot
less. These are confirmative to what we have observed for the other
series.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=33 281 11
    150]{../pics/ericsson_30_diff2.pdf}
  \caption{\small \it Comparison of the forecast from SV, GARCH,
    and sample mean. Blue: SV forecast; Green: GARCH forecast; Red:
    Sample mean forecast.}
  \label{fig:ericsson_30_diff2}
\end{figure}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    SV & GARCH & Sample Mean \\
    \hline
    1\% & 0.2210 & 0.1617 & 0.1563 \\
    \hline
    5\% & 0.7844 & 0.6900 & 0.6873 \\
    \hline
    10\% & 0.9838 & 0.9650 & 0.9596 \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:ericsson_30_diff3}
\end{table}

### 2014-09-29 thesis/SLV_Unconditional.tex
Add the draft of the paper.

--- 2014-04-30 thesis/price_models.tex
@@
-\chapter{Return Models}\label{chp:PriceModels}
-In this chapter we review some of the discrete-time return models and
-fit them to intraday returns. The intention is to assess the
-applicability of these models and to understand their statistical
-properties. The implications of these properties on the returns'
-correlation and covariance matrix are studied in chapters
-\ref{chp:Gaussian} and \ref{chp:CrossCorrelationFat}.
-
-\S\ref{sec:Garch_model} reviews the GARCH model using the Nordea
-15-minute returns as an example. \S\ref{sec:SLV_model} examines the
-stochastic log-volatility model and discusses the possibility of
-describing the residuals of the log-volotility series using the
-Johnson Su distribution. The same Nordea return series is fitted as an
-example. \S\ref{sec:XieCalc} derives the unconditional distribution
-function of the stochastic log-volotility model and compares the
-analytic results to the 30-minute returns of a number of Swedish
-stocks.
-
-\section{Gaussian Distribution}
-The justification of modeling return series as independent,
-identically distributed Gaussian variates comes from imagining the
-price process $S(t)$ as a Brownian motion whose increment $\sigma
-dw_t$ at each time step is independent and scales as $\sigma
-\sqrt{dt}$. Then, by adding a drift term $\mu dt$ that represents some
-deterministic trend in the price process, one can express the
-price $S(t)$ as a stochastic differential equation:
-\begin{eqnarray*}
-  dS &=& S\mu dt + S\sigma dw_t \\
-  d(\ln S) &=& (\mu - \frac{1}{2} \sigma^2)dt + \sigma dw_t \\
-\end{eqnarray*}
-It follows from this equation that $\ln S(t) - \ln S_0$, where $S_0$
-is the price at time 0, has Gaussian distribution with mean $(\mu -
-\sigma^2/2)t$ and variance $\sigma^2 t$. Therefore, in a discrete-time
-model, where the length of each time step $\Delta t$ is fixed, the
-return $r_t = \ln S(t + \Delta t) - \ln S(t)$ is assumed to be
-Gaussian distributed and have mean and variance that are functions of
-$\Delta t$.
-
-The picture depicted above is of course overly simplified, and as we discussed in
-\S\ref{sec:StylizedFacts}, the distribution of returns is not really
-Gaussian. Nevertheless, the assumption of Gaussian distributed returns
-underlies such important theories as Black and Scholes theory of
-option pricing. Thus, albeit hugely inaccurate, the Gaussian
-distribution as a returns model shall not be forgotten. In the next
-few sections, we discuss some more realistic returns models.
-
-\section{GARCH models} \label{sec:Garch_model}
-``GARCH'' is the acronym for ``Generalized Autoregressive Conditional
-Heteroscedasticity''. A GARCH(p, q) model is defined by the following
-equation system:
-\begin{equation}
-  \label{eq:garch_def}
-  \begin{aligned}
-    r_t &= \mu + \epsilon_t \\
-    \sigma_t^2 &= \alpha_0 + \sum_{i=1}^q \alpha_i \epsilon_{t-i}^2 +
-    \sum_{i=1}^p \beta_i \sigma_{t-i}^2
-  \end{aligned}
-\end{equation}
-where $\epsilon_t$ is a random variable of a prescribed distribution,
-the most common ones being Gaussian, Student's t, Asymmetric Student's
-t, etc; $\sigma_t^2$ is the variance of $\epsilon_t$
-\cite{Bollerslev86}. The autocorrelation function of $\epsilon_t^2$ is
-given by (c.f. \cite{Bollerslev87})
-\begin{eqnarray*}
-\rho_n &=& \sum_{i=1} ^{p \vee q} (\alpha_i + \beta_i) \rho_{n-i}
-\text{ for $n > p$}
-\end{eqnarray*}
-where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
-zeros. $p \vee q$ denotes the maximum of p and q. From these
-equations, it is clear that the partial autocorrelation function cuts
-off at $\max(p, q)$.
-
-To assess the applicability of GARCH models at the face of intra-day
-returns, we study the {\it Nordea Bank 15min} returns sampled during
-the period 2012/01/16 - 2012/04/20. Firstly we plot the autocorrelation
-function (ACF) of $\epsilon_t^2$ from this sample, as shown
-in figure \ref{fig:nordea_15min_acf}.
-
-At the absence of GARCH effects, the ACF is expected to have an
-asymptotic Gaussian distribution with mean 0 and variance 1/T
-\cite{Bollerslev86, Bollerslev87}, which is clearly not the case in
-figure \ref{fig:nordea_15min_acf}: The first 5 autocorrelations have
-comparable sizes and do not fall off as in a Gaussian
-scheme. Moreover, figure \ref{fig:nordea_15min_vlt_acf} shows 
-even more clearly that the conditional variances of the series are
-correlated. These observations suggest a GARCH(p, q) model can be
-appropriate.
-
-\begin{figure}[htb!]
-  \centering
-  \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_acf.pdf}
-    \label{fig:nordea_15min_acf}
-  }
-  \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_vlt_acf.pdf}
-    \label{fig:nordea_15min_vlt_acf}
-  }
-  \caption{\small \it Auto-correlations among Nordea Bank 15min
-    returns. \ref{fig:nordea_15min_acf}: Auto-correlations (ACF) of the
-    squared returns; \ref{fig:nordea_15min_vlt_acf}: Auto-correlations
-    (ACF) of the conditional variances.}
-\end{figure}
-
-According to Andersen and Bollerslev et al \cite{Andersen03},
-$\epsilon_t$ can be decomposed as $\epsilon_t = \sigma_t z_t$, where
-$z_t \sim N(0, 1)$ and $\sigma_t$ can be estimated using realized
-volatility. To find out whether this is true for the aforementioned
-series we look at the probability plot of $\epsilon_t/\sigma_t$
-(figure \ref{fig:nordea_bank_15min_z_prob}), which is consistent with
-a standard Gaussian distribution except for a few outliers at the two
-ends. Hence the fat tails of the probability density function can be
-entirely accounted for by GARCH effects, implying a Gaussian
-conditional distribution is sufficient for $z_t$.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=80 258 104
-    220]{../pics/nordea_bank_15min_z_prob.pdf}
-  \caption{\small \it{Probability plot of $z_t = \epsilon_t /
-      \sigma_t$. $\epsilon_t$ are derived from Nordea Bank 15min
-      returns while $\sigma_t$ are realized volatilities calculated
-      using 30s returns within each 15min interval. Horizontal axis:
-      $z_t$}; Vertical axis: cummulative probability function (CDF) of
-    $z_t$, arranged on such a scale that the CDF of the standard
-    Gaussian is a straight line.}
-  \label{fig:nordea_bank_15min_z_prob}
-\end{figure}
-
-To determine the orders p and q of the GARCH model, we again check
-the partial autocorrelation function of $\epsilon_t^2$. As just
-mentioned, the first 5 partial autocorrelations are significant. Thus
-we can tentatively estimate a GARCH(1, 5) model. Then by applying the
-Lagrange multiplier test and accordingly removing insignificant
-parameters, we obtain a GARCH(1, 3) model
-\ref{tab:nordea_15min_garch}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    parameter & value & std. error & t statistic \\
-    \hline
-     $\alpha_0$ &   1.05668e-06 &   3.56871e-07 &    2.96096\\
-     \hline
-     $\beta_1$  &     0.687212  &   0.0299314   &    22.9596\\
-     \hline
-     $\alpha_1$ &      0.128614 &   0.0268144   &    4.79643\\
-     \hline
-     $\alpha_3$ &    0.0568022  &   0.0301503   &    1.88397 \\
-     \hline
-  \end{tabular}
-  \caption{\small \it GARCH(1, 3) model of Nordea Bank 15min returns}
-  \label{tab:nordea_15min_garch}
-\end{table}
-
-\section{ARIMA Log-volotility Models}\label{sec:SLV_model}
-With the availability of transaction data, estimating conditional
-volatility using returns sampled at a higher frequency (realized
-volatility) gives superior acuracy and reliability. However, at which
-frequency the time series should be sampled in order to give an
-unbiased and consistent estimate of the volatility is not a trivial
-question. Naively one would argue by the doctrine of statistics that
-the often the series is sampled, the better the estimate. This is also
-well supported in theory, since the quadratic variation which is the
-theoretial variance of the time series (see \cite{Andersen03}, theorem
-1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
-\S6, theorem 22):
-\[
-\var(p_{t+h} - p_t) = [r, r]_{t+h} - [r, r]_t = \lim_{n \to \infty}
-\sum_{T^i \in \pi_n} (p^{T_{i+1}} - p^{T_i})^2
-\]
-where $p_t$ is the log price process and $r_t = p_t - p_0$. $\pi_n$
-is a refining partition of the interval $[t, t+h]$.
-
-However, due to the noise introduced by market micro-structure, there
-is actually an optimal sampling frequency relative to $h$. While the
-method to determine this optimal frequency is a subject of debate (see
-for example \cite{Sahalia05}), it is not difficult to find a fairly
-satisfactory frequency in practice.
-
-Andersen and Bollerslev et al proved in \cite{Andersen03} that, with
-the conditional volatility well approximated by the realized
-volatility, $r_{t+h, h} = p_{t+h} - p_t$ is governed by a Gaussian
-distribution (see theorem 2 of \cite{Andersen03}):
-\[
-r_{t+h, h} | \sigma(\mu_{t+s}, \sigma_{t+s})_{s \in [0, h]} \sim
-N(\int_{0}^h \mu_{t+s} ds, \int_{0}^h \sigma_{t+s}^2 ds)
-\]
-
-Thus one can try a few frequencies and compare the distribution of
-$r_{t+h, h}/\hat{\sigma}_{t+h, h}$ with the standard Gaussian. If the
-two match, $\hat{\sigma}_{t+h, h}$ is a good estimate of $\sigma_{t+h,
-  h}$, the volatility of $r_{t+h, h}$.
-
-For the Nordea Bank 15-minute returns, $h = 15\text{min}$, and a
-reasonally good sampling frequency is 1/30sec, as confirmed by the
-probability plot \ref{fig:nordea_bank_15min_z_prob} as well as by the
-autocorrelation functions of $r_{t+h, h} / \sigma_{t+h, h}$ and of
-$(r_{t+h, h} / \sigma_{t+h, h})^2$ which are shown in
-\ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} respectively.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of $r_{t+h, h} / \sigma_{t+h, h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_acf.pdf}
-    \label{fig:nordea_15min_quotient_acf}
-  }
-  \subfigure[ACF of $(r_{t+h, h} / \sigma_{t+h, h})^2$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 236 118
-    200]{../pics/nordea_15min_quotient_squared_acf.pdf}
-    \label{fig:nordea_15min_quotient_squared_acf}
-  }
-  \caption{\small \it Nordea 15min $r_t/\sigma_t$ and
-    $(r_t/\sigma_t)^2$ ACF.}
-\end{figure}
-
-We see in figure \ref{fig:nordea_15min_quotient_acf} and
-\ref{fig:nordea_15min_quotient_squared_acf} there is essentially no
-autocorrelation in the series $r_{t+h, h} / \sigma_{t+h, h}$ and
-$(r_{t+h, h} / \sigma_{t+h, h})^2$, supporting the assumption that the
-series is i.i.d Gaussian distributed. As a result, the problem of
-modeling $r_{t+h, h}$ boils down to modeling $\sigma_{t+h,
-  h}$. Andersen and Bollerslev et al reported that, for the exchange
-rates between Deutch mark, yen and dollar, $\ln \sigma_{t+h, h}$ is
-gaussian distributed \cite{Andersen03}. This is, however, not the case
-for our series in question. In fact, in our case, $\ln \sigma_{t+h,
-  h}$ is right skewed (skewness 0.3342) and leptokurtic (kurtosis
-6.1006). See figure \ref{fig:nordea_15min_logvol_prob}
-\begin{figure}[htb!]
-  \centering
-  %\vspace{-15mm}
-  \includegraphics[scale=0.5, clip=true, trim=80 223 107
-  4]{../pics/nordea_15min_logvol_prob.pdf}
-  \caption{\small \it Probability plot of Nordea 15min
-    $\ln\sigma_{t+h, h}$ unconditional distribution}
-  \label{fig:nordea_15min_logvol_prob}
-\end{figure}
-
-Moreover, the series of $\ln\sigma_{t+h, h}$ shows long-lasting and
-periodic autocorrelations with an apparent period of 33 (see figure
-\ref{fig:nordea_15min_logvol_acf}). This suggests the series may be
-described by an integrated seasonal ARMA\footnote{ARMA ---
-  Autoregressive Moving Average} model. But before we set out
-to specify and estimate this model of the logvolatility
-$\ln\sigma_{t+h, h}$, it is useful to first write down the model of
-the returns $r_t$:
-\begin{eqnarray}
-   r_{t+h, h} &=& \ln p_{t+h} - \ln p_t \nonumber \\
-   r_{t+h, h} &=& \mu + \sigma_{t+h, h} b_t \label{eq:SLV_spec}
-\end{eqnarray}
-where $p_t$ is the price of the asset at time $t$ and $b_t \sim N(0,
-1)$. For simplicity, in the rest of this chapter we shall just write
-$t$ for the subscript ``t+h, h'', since the time interval $h$ is fixed
-for each time series and is a known constant. In the case of the
-Nordea Bank 15-minute returns, h = 15 minutes.
-
-As has just been said, $\ln \sigma_t$ for the series in question has
-long-lasting auto-correlations (often referred to as ``long
-memory''). This suggests simplifying the series by differencing
-\cite{BoxJenkins94}:
-\[
-w_t = (1-B)(1-B^s)\ln\sigma_t
-\]
-where $B$ is the back-shift operator and $s=33$ is the seasonality.
-
-The autocorrelation function of the differenced process $w_t$, as shown in
-figure \ref{fig:nordea_15min_w_acf}, clearly points to a seasonal
-moving-average model: There are only 4 non-zero autocorrelations in
-the plot, located at lags 1, 32, 33, 34, respectively; furthermore,
-the two at 32 and 34 are approximately equal. Thus we can write down
-the model as
-\begin{eqnarray}
-  w_t &=& (1 - \theta B)(1 - \Theta B^s) y_t \label{eq:nordea_w}
-\end{eqnarray}
-where $\theta$ and $\Theta$ are parameters to be determined and $y_t$
-is a noise process with constant variance $\sigma_y^2$ and mean
-0. $y_t$ is often refered to as the residuals.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[ACF of $\ln\sigma_{t+h,h}$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    235]{../pics/nordea_15min_logvol_acf.pdf}
-    \label{fig:nordea_15min_logvol_acf}
-  }
-  \subfigure[ACF of $w_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    235]{../pics/nordea_15min_w_acf.pdf}
-    \label{fig:nordea_15min_w_acf}
-  }
-  \caption{\small \it Nordea 15min $\ln\sigma_{t+h, h}$ and $w_t$
-    autocorrelations}
-\end{figure}
-
-The above seasonal moving average model has the following
-autocovariance structure \cite{BoxJenkins94}:
-\begin{eqnarray*}
-  \gamma_0 &=& \sigma_y^2 (1 + \theta^2)(1 + \Theta^2) \\
-  \gamma_1 &=& -\sigma_y^2\theta(1 + \Theta^2) \\
-  \gamma_s &=& -\sigma_y^2\Theta(1 + \theta^2) \\
-  \gamma_{s+1} &=& \gamma_{s-1}\;=\;\sigma_y^2\theta\Theta
-\end{eqnarray*}
-These equations together with the measured autocorrelations make
-possible an initial estimate of the parameters $\theta$ and $\Theta$:
-\begin{eqnarray*}
-  {\rho_{s+1}/\rho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
-    1 + \theta^2} \\
-  {\rho_{s+1}/\rho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
-    1 + \Theta^2} \\
-\end{eqnarray*}
-Substituting in the measured values shown in table
-\ref{tab:nordea_15min_w_acf},
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    $\rho_1$ & $\rho_{s-1}$ & $\rho_s$ & $\rho_{s+1}$ \\
-    \hline
-    -0.4703 &  0.2053 & -0.4564 &  0.2212 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Nordea 15min $w_t$ autocorrelations}
-  \label{tab:nordea_15min_w_acf}
-\end{table}
-we get
-\begin{eqnarray*}
-  \theta &=& 0.6890 \\
-  \Theta &=& 0.6378
-\end{eqnarray*}
-Among the two roots of each of the 2nd order equations in the above,
-we have chosen the one in the range $(-1, 1)$ so as to ensure
-invertibility of the model \cite{BoxJenkins94}.
-
-With an estimate of $\theta$ and $\Theta$, one can then infer the
-noise process i.e. the residuals $y_t$:
-\begin{equation}
-  \label{eq:infer_y}
-  y_t = w_t + \theta y_{t-1} + \Theta y_{t-s} - \theta \Theta y_{t-s-1}
-\end{equation}
-where we substitute $y_t\;(t \leq 0)$ with their unconditional
-expectation 0.
-% By inverting the seasonal moving average model of $w_t$ we can then
-% infer the noise process $y_t$ \footnotemark, i.e. the residuals:
-% \begin{eqnarray*}
-%   y_t &=& {1 \over 1 - \theta B}{1 \over 1 - \Theta B^s} w_t \\
-%   &=& \sum_{i=0}^{\infty}\sum_{j=0}^{\infty} \theta^i \Theta^j B^{i+j}
-%   w_t \\
-%   &=& w_t + \sum_{k=1}^{\infty} \sum_{j=0}^{\bot(k/s)} \theta^{k-sj} \Theta^j
-%   w_{t-k}
-% \end{eqnarray*}
-% where $\bot(\cdot)$ denotes the floor function, i.e. the largest
-% integer smaller than its argument.
-% \footnotetext{
-%   The autoregressive process resulting from the inversion is inifinite
-%   in extent. Hence one must choose a cut-off value for k. For the
-%   present case of Nordea 15min log-volatility, I choose the value
-%   $k=sj$ for which $|\Theta|^j < 0.02$. This value of k, call it
-%   $k_{\top}$, is around 550. To avoid data reduction of this amount,
-%   we add to the beginning of $\{w_t\}_{t=1}^n$ $k_{\top}$ zeros
-%   corresponding to t=0, -1, ..., -$k_{\top}$+1. Recall $\E(w_t) = \E(y_t
-%   - \theta y_{t-1} - \Theta y_{t-s} + \theta \Theta y_{t-s-1}) = 0$.
-% }
-
-In order to forecst the $w_t$ process, and hence the return process
-itself, we must also know the distribution of $y_t$. Moreover, to
-properly estimate the parameters of the model in the sense of maximum
-likelihood, we are also in need of the distribution of $y_t$.
-
-If $y_t$ is indeed i.i.d random variables as we assumed, we only need
-to work out its unconditional distribution. This is shown in figure
-\ref{fig:nordea_15min_y_qq} and \ref{fig:nordea_15min_y_acf}.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[QQ-plot of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_y_qq.pdf}
-    \label{fig:nordea_15min_y_qq}
-  }
-  \subfigure[ACF of $y_t$]{
-    \includegraphics[scale=0.4, clip=true, trim=95 230 112
-    210]{../pics/nordea_15min_y_acf.pdf}
-    \label{fig:nordea_15min_y_acf}
-  }
-  \caption{\small \it Nordea 15min $y_t$ QQ-plot and autocorrelations}
-\end{figure}
-In addition, the measured moments of $y_t$ are shown in table
-\ref{tab:nordea_15min_y_moments}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    mean & variance & skewness & kurtosis \\
-    \hline
-    0.0012 & 0.0935 & 0.2988 & 6.8691 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Nordea 15min $y_t$ moments}
-  \label{tab:nordea_15min_y_moments}
-\end{table}
-
-If we ignore the weak autocorrelations in this crude estimate of
-$y_t$ and assume $y_t$ to be i.i.d random variables governed
-by a skewed and leptokertic distribution, the parameters of this
-distribution together with $\theta$ and $\Theta$ can be determined via
-maximum likelihood estimate (MLE).
-
-It turns out that $y_t$ can be well described by a Johnson Su
-distribution \cite{Shang2004}:
-\[
-  y_t = \xi + \lambda\sinh{z_t - \gamma \over \delta}
-\]
-where $\gamma, \delta, \lambda, \xi$ are parameters to be determined
-and $z_t \sim N(0, 1)$. The goodness of fitting is demonstrated in
-figure \ref{fig:nordea_15min_y_js_fit} by the
-empirical cummulative distribution function in comparison to the
-theoretical one.
-\begin{figure}[htb!]
-  %\vspace{-18mm}
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=92 229 116
-    133]{../pics/nordea_15min_y_js_fit.pdf}
-    \caption{\small \it Nordea 15min $y_t$ fitted to a Johnson Su
-      distribution. Horizontal: values of $y_t$, denoted x; Vertical:
-      $\ln\left(P(y_t < x)\right)$.}
-    \label{fig:nordea_15min_y_js_fit}
-\end{figure}
-
-The first 4 moments of the Johnson Su distribution are expressible
-in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}:
-\begin{eqnarray*}
-  w &=& \exp{1 \over \delta^2} \\
-  \Omega &=& {\gamma \over \delta} \\
-  \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
-  \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
-    1)\right]^{1/2} \\
-  \text{skewness}(y) &=& {
-    \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
-    \over
-    (w\cosh 2\Omega + 1)^{3/2}} \\
-  \text{kurtosis}(y) &=& {
-    w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
-    + 3(2w+1) \over
-    2(w\cosh 2\Omega + 1)^2 }
-\end{eqnarray*}
-By matching the theoretical expressions of the moments with their
-measured values, and taking help from published tables
-\cite{Johnson1965}, one can solve for the parameters $\gamma, \delta,
-\lambda, \xi$.
-
-Under the assumption of i.i.d Johnson Su distributed residuals, the
-log-likelihood function of the parameters $\gamma, \delta$
-conditional on the sample $w_t$ can be written as
-\[
-L(\theta, \Theta) = -{1 \over 2}\sum_{t=1}^n z_t^2 + n \ln{\delta
-  \over \lambda \sqrt{2\pi}} - {1 \over 2}\sum_{t=1}^n \ln\left[
-  1 + \left({y_t - \xi \over \lambda}\right)^2
-\right]
-\]
-where $y_t$ are inferred from $w_t$ using eq.\ref{eq:infer_y}
-and $z_t$ from $y_t$ using
-\[
-z_t = \delta \sinh^{-1}{y - \xi \over \lambda} + \gamma
-\]
-Note that $\gamma, \delta, \lambda, \xi$ are not really free
-parameters but rather are implied by $\theta$ and $\Theta$: Once the
-latter have been chosen and the corresponding $y_t$ inferred, the
-former are determined by the moments of $y_t$.
-
-The eventual MLE is done in Matlab with the ``active set''
-algorithm. The initial as well as the final estimation results are
-listed in table \ref{tab:nordea_15min_js_param}:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|c|c|}
-    \hline
-    & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
-    \hline
-    initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
-    0.6378 \\
-    \hline
-    MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
-    \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Nordea 15min estimation results}
-  \label{tab:nordea_15min_js_param}
-\end{table}
-
-To verify that the log-volatility can in general be described by
-ARIMA(p, d, q) models, we apply the above procedure to a number of
-Swedish stocks whose data have been collected by the author with
-home-developed software. In \S\ref{sec:nordea} and \S\ref{sec:volvo}
-we model the log-volatilities of Nordea Bank 15-minute returns and
-Volvo B 30-minute returns. Both return series span the period
-2013/10/10 - 2014/04/04. They are computed using minute-by-minute
-average prices, and their variances, i.e. volatilities squared, are
-computed as the sum of squared higher-frequency returns, which are in
-turn computed using second-by-second prices. Whenever a needed average
-price cannot be computed due to the lack of price observations in the
-relevent time interval, the value of the previous average price is
-assumed.
-
-After modeling, we perform 1-step-ahead forecast using the estimated
-models, and compare the predicted log-volatilities with the actual
-data.
-
-\subsection{Nordea Bank Log-Volatility}\label{sec:nordea}
-In this section we model the log-volatility of {\it Nordea Bank}
-15-minute returns during the period 2013/10/10 - 2014/04/04, which
-contains 3879 data points. The squared volatilities, i.e. the
-variances, are computed as the sum of squared 30-second returns. Among
-the 3879 data points, we use the first 80\% for model estimation and
-the last 20\%, i.e. 776 data points for forecast verification.
-
-This Nordea series has a auto-correlation structure similar to its
-predecessor that has been studied in previous sections. We show it in
-figure \ref{fig:nordea2_w_acf}. The reader may want to compare it with
-figure \ref{fig:nordea_15min_w_acf}.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.4, clip=true, trim=95 230 112
-  235]{../pics/nordea2_w_acf.pdf}
-  \caption{\small \it Auto-correlation function of $w_t$}
-  \label{fig:nordea2_w_acf}
-\end{figure}
-The significant auto-correlations and their respective lags are listed
-in table \ref{tab:nordea2_w_acf}.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    lag & 1 & 32 & 33 & 34 \\
-    \hline
-    ACF &   -0.4805 & 0.2861 &  -0.5034 & 0.2093 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Significant auto-correlations of $w_t$}
-  \label{tab:nordea2_w_acf}
-\end{table}
-
-Apparently this series also has the seasonality $s = 33$, but judging
-from the relatively big differences between auto-correlations at lags
-32 and 34, we don't expect the moving average process to be in the
-multiplicative form of equation \ref{eq:nordea_w}. Instead, we model
-$\ln\sigma_t$ has
-\begin{eqnarray}
-  (1 - B)(1 - B^s) \ln\sigma_t &=& (1 - \theta_1 B - \theta_s B^s -
-  \theta_{s+1} B^{s+1} ) y_t \label{eq:nordea2_v}
-\end{eqnarray}
-The parameters are etimated to be
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    parameter & $\theta_1$ & $\theta_{33}$ & $\theta_{34}$ & $\var(y_t)$ \\
-    \hline
-    value & 0.8339 & 0.8448 & -0.7067 & 0.10071 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it: Parameter values of $\ln \sigma_t$}
-\end{table}
-
-The residuals $y_t$ do not have a Gaussian distribution, as confirmed
-by the probability plot shown on the left of figure
-\ref{fig:nordea2_y}. The non-normality, however, cannot be accounted
-by the sine hypobolic transformation of Johnson Su, because of the
-negative skewness of -1.2669 found in table
-\ref{tab:nordea2_y_mmt}. How the non-normality can be accomodated is
-quite another discussion and we do not pursue it here.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.4, clip=true, trim=20 220 30
-  115]{../pics/nordea2_y.pdf}
-  \caption{\small \it The residuals' distribution and auto-correlations. Left:
-    cummulative distribution function (CDF) of the residuals $y_t$; Right:
-    auto-correlation function of $y_t$.}
-  \label{fig:nordea2_y}
-\end{figure}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    Moment & mean & variance & skewness & kurtosis \\
-    \hline
-    value & -0.0022 & 0.1007 & -1.2669 & 15.5965 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it: Moments of the residuals $y_t$}
-  \label{tab:nordea2_y_mmt}
-\end{table}
-
-The auto-correlations of the residuals $y_t$ are insignificant, as
-shown on the right of figure \ref{fig:nordea2_y}. Though the Ljung-Box Q
-test rejects the null hypothesis of no-autocorrelation in the
-residuals, the weak auto-correlations, if any at all,  are not worth
-further complicating the model.
-
-Forecast of $\ln\sigma_t$ is done by reverting equation
-\ref{eq:nordea2_v} and write
-\begin{eqnarray*}
-  \ln\sigma_t &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
-  \ln\sigma_{t-s-1} + (1 - \theta_1 B - \theta_s B^s + \theta_{s+1}
-  B^{s+1} ) y_t \\
-  \E(\ln\sigma_t | \mathcal{F}_{t-1}) &=& \ln\sigma_{t-1} + \ln\sigma_{t-s} -
-  \ln\sigma_{t-s-1} - \theta_1 y_{t-1} - \theta_s y_{t-s} +
-  \theta_{s+1} y_{t-s-1}
-\end{eqnarray*}
-
-Thus the forecast, namely $\E(\ln\sigma_t | \mathcal{F}_{t-1})$, is
-found in terms of previous values in the same series and of previous
-residuals.
-
-To compare the performance, we also estimate a GARCH model with
-Gaussian conditional distribution for the Nordea Bank 15-minute
-return. The procedure is standard and we only list the results here
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_5$ & $\beta_1$ \\
-    \hline
-    Value & $7.3686 \times 10^{-7}$ & 0.154131 & 0.0360 & 0.491832 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Parameters of GARCH(1, 5) model of Nordea
-    Bank 15-minute returns}
-  \label{tab:nordea2_garch}
-\end{table}
-
-Forecasting using both models, we get two series of one-step-ahead
-forecast, $\ln \sigma^{\text{ARIMA}}_t$ and $\ln \sigma^G_t$,
-respectively. We denote the empirical log-volatilities as $\ln
-\hat{\sigma}_t$. Then the accuracy of the two models can be compared
-by comparing $\ln \sigma^{\text{ARIMA}}_t - \ln \hat{\sigma}_t$ and
-$\ln \sigma^G_t - \ln \hat{\sigma}_t$. As a reference, we also include
-in the comparison the naive approach of taking the sample mean of
-$\ln \sigma_t$ as the forecast.
-
-Table \ref{tab:slv_garch_comparison} shows the standard deviation of
-$\ln \sigma^F_t - \ln \hat{\sigma}_t$ for the 3 kinds of
-forecast. Here we use $\ln \sigma^F_t$ to denote the forecast of $\ln
-\sigma_t$. Clear the ARIMA model has the smallest standard deviation,
-the sample mean has the largest, and the GARCH model sits in between.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    & ARIMA & GARCH & Sample mean \\
-    \hline
-    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ &
-    0.2876  &  0.3096 &   0.3349 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
-    \ln\hat{\sigma}_t$}
-  \label{tab:slv_garch_comparison}
-\end{table}
-Even more informative about the quality of forecast is the
-distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$, which we show
-in figure \ref{fig:slv_garch_comparison}.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=0 292 0
-  255]{../pics/slv_garch_comparison.pdf}
-  \caption{\small \it Comparison of the forecast from ARIMA, GARCH,
-    and sample mean. Blue: ARIMA forecast; Green: GARCH forecast; Red:
-    Sample mean forecast.}
-  \label{fig:slv_garch_comparison}
-\end{figure}
-The left plot in the figure shows the situation of under-estimates,
-i.e. $\ln \sigma^F_t - \ln \hat{\sigma}_t < 0$. We see that, to the
-far left, the blue line is below the green, which is in turn below the
-red, meaning the ARIMA forecast has the lowest probability of
-under-estimate. The GARCH forecast comes second and the sample
-mean forecast under-estimates most.
-
-The same ranking of forecast quality is also seen on the right plot of
-figure \ref{fig:slv_garch_comparison}, which shows the complementary
-distribution of $\ln \sigma^F_t - \ln \hat{\sigma}_t$ for postive
-values, i.e. over-estimates: The ARIMA forecast has the lowest
-probability of over-estimate; GARCH comes second and the sample mean
-is most likely to over-estimate.
-
-As another measure of the quality of forecast, the fraction of
-``good'' forecasts that differ no more than a given percentage from
-the measured value is worth a few words. Table
-\ref{tab:nordea_good_percentage} compares this fraction of good
-forecasts from the 3 alternatives with respect to different
-requirements of goodness.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ &
-    ARIMA & GARCH & Sample Mean \\
-    \hline
-    1\% & 0.1920 & 0.2010 & 0.1843 \\
-    \hline
-    5\% & 0.7887 & 0.7436 & 0.7229 \\
-    \hline
-    10\% & 0.9665 & 0.9601 & 0.9369 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Fraction of ``good'' forecasts as defined by
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
-  \label{tab:nordea_good_percentage}
-\end{table}
-Clearly, although ARIMA and GARCH rank differently with respect to
-different requirements of goodness, they both out-perform the sample
-mean forecast. So the efforts of building models and estimating them
-have paid off.
-\subsection{Volvo B Log-Volatility}\label{sec:volvo}
-In this section we model the log-volatility of {\it Volvo B} 30-minute
-returns during the period 2013/10/10 - 2014/04/04. This series
-contains 1884 log-volatilities computed using 2-minute returns in each
-30-minute interval. Among them we use the first 1507 for model
-estimation and the last 377 for forecast and model verification.
-
-The left plot of figure \ref{fig:volvo_inno_and_lv_acf} shows the
-distribution of $(r_t - \E(r_t))/\sigma_t$. We observe in the figure a
-nice Gaussian variate, so we can be sure that the sum of squared
-2-minute returns makes a good approximation to the variance of
-30-minute returns in this particular case.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=0 256 0
-    151]{../pics/volvo_inno_and_lv_acf.pdf}
-    \label{fig:volvo_inno_and_lv_acf}
-  \caption{\small \it Left: probability plot of $(r_t -
-    \E(r_t))/\sigma_t$; Right: auto-correlations of $\ln \sigma_t$}
-\end{figure}
-
-Guided by the auto-correlations of $\ln\sigma_t$ shown in the right
-plot of figure \ref{fig:volvo_inno_and_lv_acf} we find the following
-model:
-\begin{eqnarray}
-  && (1 - B)(1 - B^s) \ln \sigma_t \nonumber \\
-  &=& (1 - \theta_1 B - \theta_2B^2 -
-  \theta_3B^3 - \theta_4B^4) (1 - \Theta B^s)
-  y_t \label{eq:volvo_lv_model}
-\end{eqnarray}
-where $s = 16$ is the seasonality and is apparent from the
-auto-correlations of $\ln\sigma_t$. Fitting this model to the
-actual log-volatilities yields the parameter values listed in table
-\ref{tab:volvo_params}.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|c|c|}
-    \hline
-    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ &
-    $\Theta$ & residual variance \\
-    \hline
-    Value & 0.7305 & 0.0575 & 0.0574 & 0.0346 & 0.8324 & 0.1340\\
-    \hline
-  \end{tabular}
-  \caption{\small \it Volvo B log-volatility parameters}
-  \label{tab:volvo_params}
-\end{table}
-Forecasting using the estimated model parameters gives the forecast
-series $\ln \sigma^{\text{ARIMA}}_t$. To access the quality of the
-forecast, we also estimate a GARCH model using the returns. The result
-is a GARCH(1, 1) model:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    Parameter & $\alpha_0$ & $\alpha_1$  & $\beta_1$ \\
-    \hline
-    Value & $3.125 \times 10^{-7}$ & 0.05 & 0.90 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it GARCH(1, 1) model of Volvo B 30-minute returns}
-  \label{tab:volvo_garch}
-\end{table}
-
-The forecasts from ARIMA, GARCH, and the sample mean are compared
-using the difference $\ln \sigma^F_t - \ln\hat{\sigma}_t$, where $\ln
-\sigma^F_t$ stands for the forecast. The distributions of this
-difference is plotted in figure \ref{fig:volvo_slv_garch_cmp}, and
-the standard deviation of the distributions are listed in table
-\ref{tab:volvo_slv_garch_cmp}.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    & ARIMA & GARCH & Sample mean \\
-    \hline
-    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3261 &
-    0.4250 & 0.3708 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
-    \ln\hat{\sigma}_t$}
-  \label{tab:volvo_slv_garch_cmp}
-\end{table}
-
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=10 282 0
-    243]{../pics/volvo_slv_garch_cmp.pdf}
-    \label{fig:volvo_slv_garch_cmp}
-  \caption{\small \it Comparison of the forecast from ARIMA, GARCH,
-    and sample mean. Blue: ARIMA forecast; Green: GARCH forecast; Red:
-    Sample mean forecast.}
-\end{figure}
-Figure \ref{fig:volvo_slv_garch_cmp} shows, as in the previous case of
-Nordea Bank 15-minute returns, the ARIMA model performs the best, GARCH
-the second, and the sample mean the worst. However, when it comes to
-over-estimates, the sample mean appears to be the best estimator,
-while ARIMA excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
-the sample for estimation (1507 data points) and over the sample for
-comparison (377 data points) reveals that the first sample has mean
--6.3127 while the second has -6.2072. This increment explains the low
-probability of over-estimation when using the first sample mean as
-forecast.
-
-Table \ref{tab:volvo_good_percentage} compares the fraction of
-``good'' estimates as measured by ${|\ln\sigma^F_t -
-  \ln\hat{\sigma}_t| \over |\ln \hat{\sigma}_t|}$ being less than 1\%,
-5\% and 10\%.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ &
-    ARIMA & GARCH & Sample Mean \\
-    \hline
-    1\% & 0.2175 & 0.1220 & 0.1406 \\
-    \hline
-    5\% & 0.7162 & 0.6154 & 0.6631 \\
-    \hline
-    10\% & 0.9231 & 0.8806 & 0.8992 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Fraction of ``good'' forecasts as defined by
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
-  \label{tab:volvo_good_percentage}
-\end{table}
-We see from the table that the ARIMA model consistently excels over
-the other two alternatives. Once again, it confirms the validity and
-good performance of an ARIMA log-volatility model.
-
-% \subsection{Ericsson B Log-Volatility}\label{sec:ericsson}
-% The probability density function of the residuals $y_t$ that result
-% from the MLE estimate is shown in figure
-% \ref{fig:nordea_15min_y1_hist} together with its counterpart from the
-% initial estimate.
-% \subsection{45min Returns}
-\section{Unconditional Distribution Functions of ARIMA Log-Volatility
-  Models}\label{sec:XieCalc}
-In this section we study the unconditional distribution function of
-the Stochastic Log-Volatility model specified as equation
-\ref{eq:SLV_spec}. As is discussed in \S\ref{sec:SLV_model},
+\chapter{Unconditional Distribution Functions of Stochastic
+  Log-Volatility Models}
+\label{chp:SLV_unconditional}
+In this chapter we study the unconditional distribution functions of
+Stochastic Log-Volatility models specified as equation
+\ref{eq:SLV_spec}. As is discussed in section \ref{sec:SLV_model},
 $\ln\sigma_t$ can be described by an ARMA or ARIMA \footnote{ARIMA ---
@@
 
-In \S\ref{sec:SLV_Symmetric} we first study the model in the
+In section \ref{sec:SLV_Symmetric} we first study the model in the
 relatively simple case when $v$ and $b$ are uncorrelated and $\mu
 = 0$. If this simplified version proves inadequate, one may resort to
-the general model studied in \S\ref{sec:SLV_Asymmetric}.
+the general model studied in section \ref{sec:SLV_Asymmetric}.
 
-\subsection{The Simplified model}\label{sec:SLV_Symmetric}
+\section{The Simplified model}\label{sec:SLV_Symmetric}
 In the following we derive the unconditional PDF of $r'$, denoted
@@
 
-\subsection{The General model}\label{sec:SLV_Asymmetric}
+\section{The General model}\label{sec:SLV_Asymmetric}
 It has long been hypothesized in the liturature that skewness is the
@@
 including the Volvo B 30-minute returns described in
-\S\ref{sec:SLV_Symmetric}. In the following, we list the parameter
+section \ref{sec:SLV_Symmetric}. In the following, we list the parameter
 values and show the comparison between the empirical distribution
@@
 
-\subsection{Relation to Conditional Distribution Functions}
+\section{Relation to Conditional Distribution Functions}
 In the last section we have shown that the unconditional distribution of
@@
 In the context of conditional distributions and forecast, this
-correlation translates to the correlation between the residual of the
-log-volatility, denoted $y_t$ in \S\ref{sec:nordea} and
-\S\ref{sec:volvo}, and $b_t$. Now let us consider the forecast
-function of an ARIMA model:
+correlation translates to the correlation between $b_t$ and the
+residual of the log-volatility, denoted $y_t$ in the previous
+chapters. Now let us consider the forecast function of an ARIMA model:
 \begin{eqnarray*}
@@
 
-For the Nordea series considered in \S\ref{sec:nordea},
+For the Nordea 15-minute series considered in section \ref{sec:nordea2_15min},
 $\text{corr}(y_t, b_t)$ is found to be $3.56 \times 10^{-2}$; while
-for the Volvo series considered in \S\ref{sec:volvo},
+for the Volvo series considered in section \ref{sec:volvo},
 $\text{corr}(y_t, b_t)$ is found to be $-7.3 \times 10^{-3}$. These

### 2014-09-29 thesis/appendix.tex
Add the draft of the paper.

% \section{Fundamental Concepts \& Notations}
% \label{sec:FundamentalConcepts}
% This section is a list of a few concepts that may be unfamiliar to the
% reader and that we will often refer to later in the thesis:
% \begin{itemize}
% \item Return. Given a fixed time interval $[t - \Delta t, t]$, for example, a
%   day, a week, a month, etc, and a particular asset, for example, a
%   share in company ABC, the return of this asset over the time
%   interval is defined as
%   \begin{eqnarray*}
%     r_{t, \Delta t} &=& \ln S_{t} - \ln S_{t - \Delta t} \\
%     &=& \ln \left(1 + {S_{t} - S_{t - \Delta t} \over S_{t - \Delta t}}\right) \\
%     &\approx& {S_{t} - S_{t - \Delta t} \over S_{t - \Delta t}}
%   \end{eqnarray*}
%   where $S_t$ is the price of the asset at time $t$. $\Delta t$ is
%   sometimes called the time-lag of the return. Quite often, where
%   confusion is not possible, we will just write $r_t$ to mean $r_{t,
%     \Delta t}$, the time lag either does not matter or is clear from
%   the context.

% \item Autocorrelation. By autocorrelation, denoted $\rho_k$ here, we
%   mean the correlation between two temporally separated observations
%   of the same time series:
%   \begin{eqnarray*}
%     \rho_k &=& {
%       \E\left[(a_t -\E(a_t))(a_{t-k} - \E(a_{t-k})\right]
%       \over
%       \sqrt{\text{var}(a_t)}\sqrt{\text{var}(a_{t-k})}
%     }
%   \end{eqnarray*}
%   Here $\E(x)$ stands for the expectation value of $x$, and
%   $\text{var}(x)$ stands for the variance of $x$. $k$ is called the
%   time-lag and is the temporal seperation of the two observations
%   measured by the number of observations in between. For example, the
%   time-lag between the 1st and 3rd observation is 2.

% \item Covariance Matrix. Correlations between the returns of a
%   group of assets are described by the covariance matrix. When
%   the asset returns are described by a stable distribution law with
%   L\'evy index $\alpha$, the empirical covariance matrix is
%   constructed as
%   \begin{equation}
%     \label{eq:covariance}
%     \begin{aligned}
%       C_{ij} &= {1 \over T^{2/\alpha}} \sum_{t=1}^T [r_{i,t}-\E(r_i)]
%       [r_{j,t}-\E(r_j)] \\
%       &= {1 \over T^{2/\alpha}} RR'
%     \end{aligned}
%   \end{equation}
%   where $r_{i,t}$ is the return of the i-th asset at time t and is
%   placed at the entry (i, t) of matrix R; $R'$ denotes the
%   transpose of R. In most practical situations, one has abundant data
%   for each and every asset. Thus $T \geq N$ is assumed throughout this
%   thesis.

% \item Auto-regressive processes. A time series $r_t$ is called an
%   auto-regressive process of order $p$ and denoted AR(p), if it can be
%   written in the following form:
%   \begin{eqnarray*}
%     r_t &=& \sum_{i=1}^p \phi_i r_{t-i} + a_t
%   \end{eqnarray*}
%   where, for all $i$, $\phi_i \in (-1, 1)$, and the $a_t$'s are
%   independent and identically distributed (iid.) random variables with
%   zero mean. Obviously this implies the mean of $r_t$ is zero
%   too. Apart from this, their distribution of $a_t$ is not restricted
%   to any particular form.

%   Of particular interest to this thesis is the AR(1) process:
%   \begin{eqnarray*}
%     r_t = \phi r_{t-1} + a_t
%   \end{eqnarray*}
%   Its autocorrelation function $\rho_k = \text{corr}(r_t, r_{t-k})$
%   ($k = 0, 1, 2, \cdots$) can be easily shown to fall off
%   exponentially:
%   \begin{eqnarray*}
%     \rho_k &=& {\E(r_tr_{t-k}) - \E(r_{t-k})\E(r_{t}) \over
%       \sqrt{\var(r_t) \var(r_{t-k})}} \\
%     &=& {\phi \E(r_{t-1}r_{t-k}) + \E(a_t r_{t-k})
%       \over
%       \sqrt{\var(r_t) \var(r_{t-k})}
%     } \\
%     &=& \phi \rho_{k-1}
%   \end{eqnarray*}
%   where $\E(a_t r_{t-k}) = 0$ follows from the fact that any return
%   $r_{t1}$ must not depend on disturbances $a_{t2}$ that occur later
%   in time. The last equation means $\rho_k$ is a geometric series. Since
%   $\rho_0 = 1$, we have
%   \begin{eqnarray*}
%     \rho_k &=& \phi^k \\
%     |\rho_k| &=& e^{k\ln|\phi|} \\
%   \end{eqnarray*}
%   Note that $|\phi| < 1$ and hence $\ln|\phi| < 0$.

%   Although $k$ can only take integer values, it is still useful to
%   define a correlation time $\tau$ such that $\phi^\tau = 1/2$. Such a
%   quantity is more intuitive and constitutes a measure of
%   autocorrelations that is universal and comparable among different
%   time series' models. From the definition of $\tau$ we get
%   \begin{equation}
%     \label{eq:tau_def}
%     \begin{aligned}
%       \phi^\tau &= 1/2 \\
%       \tau &= -{\ln 2 \over \ln\phi} \\
%       \phi &= 2^{-1/\tau}
%     \end{aligned}
%   \end{equation}
%   Figure \ref{fig:AR1-autocorrelation} shows the autocorrelation
%   function of the AR(1) model. As proven above, this function decays
%   exponentially.
%   \begin{figure}[htb!]
%     %\vspace{-15mm}
%     \centering
%     \includegraphics[scale=0.5, clip=true, trim=113 229 115
%     139]{../pics/AR1-autocorrelation.pdf}
%     \caption{\small \it Autocorrelation function of the AR(1) model
%       with $\phi=1/\sqrt{2}$. Red circles: autocorrelations at $k=0, 1,
%       2, \cdots$. Blue line: $e^{t\ln\phi}$.}
%     \label{fig:AR1-autocorrelation}
%   \end{figure}
%   For more details about conventional time series' models, see
%   \cite{BoxJenkins94}.
% \end{itemize}

% % For purposes of later reference, we also list some notations that may
% % cause confusion to the reader:
% % \begin{itemize}
% % \item $\E(x)$ or $\mean{x}$: The expectation value of $x$.
% % \item $\var(x)$: the variance of $x$.
% % \item $\text{cov}(x,y)$: the covariance of $x$ and $y$.
% % \item $\text{corr}(x,y)$: the correlation between $x$ and $y$, i.e.
% %   \begin{equation*}
% %     \text{corr}(x,y) = {\text{cov}(x,y) \over \sqrt{\var(x)\var(y)}}
% %   \end{equation*}
% % \end{itemize}

% \section{Stylized facts}\label{sec:StylizedFacts}
% This section presents and explains a few ``stylized facts'',
% i.e. phenomena that are widely observed and accepted as true.
% \begin{itemize}
% \item Fat tails. It has been consistently reported by various studies
%   - for example \cite{Potters2003} and \cite{Mantegna2000} - that the
%   probability density function (PDF) of stock/index returns are not
%   Gaussian. Unlike the Gaussian PDF, which is symmetric and falls off
%   very quickly as its argument moves from the center to the outskirts
%   (tails), the PDF of stock/index returns are higher on the tails,
%   i.e. the probability of large fluctuations is higher than is dictated
%   by a Gaussian distribution - in fact, even higher than dictated by
%   an exponential function. Empirical studies suggest the distribution
%   function of the returns follows a power-law on the tails --- the
%   exponent of the power depends on the specific stock/index.

%   Figure \ref{fig:FatTail} illustrates this feature.
%   \begin{figure}[htb!]
%     \centering
%     \includegraphics[scale=0.6, clip=true, trim=92 229 110
%     140]{../pics/FatTail.pdf}
%     \caption{\small \it Fat tails of {\it Nordea Bank} 15-minute
%       returns during the period 2013-10-10 and 2014-01-29. The returns
%       are computed using minute-by-minute average prices.}
%     \label{fig:FatTail}
%   \end{figure}

% \item Non-zero skewness. Apart from fat tails, the PDF of stock/index
%   returns are often also skewed. If the skewness is positive
%   (negative), the probability of very large positive (negative)
%   returns is higher than that of very large negative (positive)
%   returns, even though the mean of the returns is 0 or extremely
%   close to 0.

%   Table \ref{tab:EmpiricalSkewness} lists the skewness of a few
%   Swedish stocks traded on the Stockholm OMX market.
%   \begin{table}[htb!]
%     \footnotesize
%     \centering
%     \begin{tabular}{|c|c|c|c|c|c|c|}
%       \hline
%       Nordea Bank & Volvo B & Boliden & ABB Ltd & H\&M & Scania
%       B & Ericsson B \\
%       \hline
%       0.1362 & 0.0471 & 0.0567 & -0.0364 & -0.0168 &
%       -1.1554 & 0.2086 \\
%       \hline
%     \end{tabular}
%     \caption{\small \it Skewness of Stock Returns. All the returns
%       have time-lag of 15 minutes and are computed using paid prices
%       between 2013-10-10 and 2014-01-29. }
%     \label{tab:EmpiricalSkewness}
%   \end{table}

% \item Higher-order autocorrelation. To the lowest order, return series
%   are not auto-correlated --- if they are, the auto-correlations would
%   present an obvious opportunity of making easy profit and be
%   exploited and vanish as soon as they appear. However, the squared
%   returns do have significant auto-correlations, as figure
%   \ref{fig:nordea_15min_acf} illustrates. These auto-correlations
%   suggest the variances of the returns at subsequent time steps are
%   correlated. These are referred to as higher-order auto-correlations
%   and make a subjet of returns' models.

% \end{itemize}
\chapter{Case Study of Some Intraday Series}
\label{chp:appendix2}
\section{Nordea 30-minute Returns}
\label{sec:nordea2_30min}
In this section we study the volatility of Nordea Bank 30-minute
returns in the period 2013/10/10 - 2014/04/04. The realized
volatilities that approximate the volatilities of these 30-minute
returns are computed using 1-minute returns. As in the previous cases,
this choice of 1-minute returns is confirmed by the normality of the
quotient $(r_t - \E(r_t))/\sigma_t$, which is shown in figure
\ref{fig:nordea3_quotient}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=78 255 108
  120]{../pics/nordea3_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -
    \E(r_t))/\sigma_t$}
  \label{fig:nordea3_quotient}
\end{figure}
The auto-correlations of $\ln \sigma_t$ is shown in figure
\ref{fig:nordea3_lv_acf}, where we see an apparent seasonality of
16. By differencing we get $w_t = (1-B)(1-B^s)\ln\sigma_t$ where
$s=16$. Its auto-correlations are shown in figure
\ref{fig:nordea3_w_acf}.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=86 256 101
    121]{../pics/nordea3_lv_acf.pdf}
    \label{fig:nordea3_lv_acf}
  }
  \subfigure[]{
    \includegraphics[scale=0.4, clip=true, trim=86 256 103
    122]{../pics/nordea3_w_acf.pdf}
    \label{fig:nordea3_w_acf}
  }
  \caption{\small \it \ref{fig:nordea3_lv_acf}: Auto-correlations
    (ACF) of $\ln \sigma_t$; \ref{fig:nordea3_w_acf}: Auto-correlations
    (ACF) of $w_t = (1-B)(1-B^s)\ln\sigma_t$.}
\end{figure}
Once again, this auto-correlation structure
points to a seasonal moving average model:
\begin{eqnarray*}
  w_t &=& (1-\theta B)(1-\Theta B^s) y_t
\end{eqnarray*}
where $y_t$ is assumed to have Gaussian distribution, neglecting
slight non-normality as before. Maximum likelihood estimation gives
the parameter values listed in table \ref{tab:nordea3_sv_param}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\theta$ & $\Theta$ & $\var(y_t)$\\
    \hline
    Value & 0.7612 & 0.8036 & 0.0736\\
    \hline
  \end{tabular}
  \caption{\small \it Parameter values of the Seasonal Moving Average
    model}
  \label{tab:nordea3_sv_param}
\end{table}
To fit a \gls{garch} model to the same series, we assume Gaussian
innovations (c.f. equation \ref{eq:garch_def}). The result is a
\gls{garch}(1,1) model. Its parameters are listed in table
\ref{tab:nordea3_garch_param}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\beta_1$ \\
    \hline
    Value & $2.3 \times 10^{-7}$ & 0.05 & 0.9\\
    \hline
  \end{tabular}
  \caption{\small \it Parameter values of GARCH(1,1) model fitted to
    Nordea 30-minute returns.}
  \label{tab:nordea3_garch_param}
\end{table}

As before, we compare the accuracies of the \gls{garch} and the stochastic
volatility model by comparing their one-ste-ahead forecasts. We
compute the difference between their forecasts and the realized
volatilities and then look at the statistics of these difference
values. Firstly we show the mean and the standard deviation of the
differences in table \ref{tab:nordea3_diff_1}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
     & \gls{sv} & \gls{garch} & Sample mean \\
     \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & -0.0047 & 0.0130 & 0.0584 \\
    \hline
     $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.2602 &
     0.3194 & 0.3093 \\
    \hline
 \end{tabular}
  \caption{\small \it Mean and standard deviation of the 3 kinds of
    forecasts}
  \label{tab:nordea3_diff_1}
\end{table}
It is seen in this table that the \gls{sv} forecasts have a mean whose
absolute value is just above 1/3 of that of the \gls{garch} forecasts. The
standard deviation of the \gls{sv} forecasts is also smaller than that of
\gls{garch}. In addition, we check the distribution of
$\ln \sigma^F_t - \ln \hat{\sigma}_t$, and the percentage of ``good''
forecasts with respect to different criteria of goodness. These are
shown in figure \ref{fig:nordea3_diff} and table
\ref{tab:nordea3_diff_2}, respectively.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.54, clip=true, trim=21 312 0
    179]{../pics/nordea3_diff.pdf}
  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
  \label{fig:nordea3_diff}
\end{figure}
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    \gls{sv} & \gls{garch} & Sample Mean \\
    \hline
    1\% & 18\% & 18\% & 15\% \\
    \hline
    5\% & 81\% & 71\% & 69\% \\
    \hline
    10\% & 98\% & 93\% & 95\% \\
    \hline
  \end{tabular}
  % \begin{tabular}{|c|c|c|c|}
  %   \hline
  %   ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
  %     \hat{\sigma}_t|}$ &
  %   \gls{sv} & \gls{garch} & Sample Mean \\
  %   \hline
  %   1\% & 17.77\% & 17.77\% & 15.38\% \\
  %   \hline
  %   5\% & 80.90\% & 71.09\% & 69.23\% \\
  %   \hline
  %   10\% & 98.14\% & 93.10\% & 95.23\% \\
  %   \hline
  % \end{tabular}
  \caption{\small \it Percentage of ``good'' forecasts as defined by
    deviating nore more than 1\%, 5\% and 10\% from the measured
    realized volatility.}
  \label{tab:nordea3_diff_2}
\end{table}
It is clear that the \gls{sv} forecasts are considerably more accurate.

\section{Volvo 15-minute returns in 2013/14}
\label{sec:volvo}
In this section we study the volatility of Volvo B 15-minute returns
during the period 2013/10/10 - 2014/04/04.
The subject of modeling is the realized volatility
computed as the square root of the sum of squared 1-minute
returns. The normal probability plot of the quotient $(r_t -
\E(r_t))/\sigma_t$ is shown in figure \ref{fig:volvo_15_quotient},
from which one can see it is normally distributed, confirming the
choice of 1-minute returns for computing the realized volatility.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=83 259 110 
  220]{../pics/volvo_15_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -
    \E(r_t))/\sigma_t$}
  \label{fig:volvo_15_quotient}
\end{figure}
The auto-correlations of $\ln \sigma_t$ and $w_t = (1-B)(1-B^s) \ln
\sigma_t$ are plotted in figure \ref{fig:volvo_15_lv_acf} and
\ref{fig:volvo_15_w_acf}. The former clearly shows the seasonality
$s = 33$ in the auto-correlations of $\ln \sigma_t$.

\begin{figure}[htb!]
  \centering
  \subfigure[]{
  \includegraphics[scale=0.4, clip=true, trim=90 259 103
  220]{../pics/volvo_15_lv_acf.pdf}
  \label{fig:volvo_15_lv_acf}
  }
  \subfigure[]{
  \includegraphics[scale=0.4, clip=true, trim=90 259 103
  220]{../pics/volvo_15_w_acf.pdf}
  \label{fig:volvo_15_w_acf}
  }
  \caption{\small \it Left: auto-correlations of $\ln \sigma_t$;
    Right: auto-correlations of $w_t = (1-B)(1-B^s) \ln \sigma_t$.}
\end{figure}

Based on the auto-correlations of $w_t$, a seasonal moving average
model is estimated:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1-\theta_1B - \theta_2B^2 - \theta_3
  B^3)(1 - \Theta B^s)y_t
\end{eqnarray*}
By maximum likelihood estimation, the parameter values listed in table
\ref{tab:volvo_15_sv_param} are obtained.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|}
  \hline
  Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\Theta$ & 
 $\var(y_t)$ \\
 \hline
 Value & 0.7491 & 0.1170 & 0.0571 & 0.8646 & 0.1245 \\
  \hline
  \end{tabular}
  \caption{\small \it Parameter values of the stochastoc volatility
    (SV) model.}
  \label{tab:volvo_15_sv_param}
\end{table}

A \gls{garch} model (c.f. equation \ref{eq:garch_def}) assuming Gaussian
innovations is also fitted to the same series. The parameter values
are listed in table \ref{tab:volvo_15_garch_param}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
  \hline
  Parameter & $\alpha_0$ & $\alpha_1$ & $\beta_{1}$ \\
  \hline
  Value & $6.54 \times 10^{-7}$ & 0.1565 & 0.6445\\
  \hline
  \end{tabular}
  \caption{\small \it Parameter values of the GARCH model.}
  \label{tab:volvo_15_garch_param}
\end{table}

Performing one-step-ahead forecasts using both models gives the series
$\ln\sigma^F_t - \ln\hat{\sigma}_t$, where $\ln\hat{\sigma}_t$ are the
measured realized volatilities. The mean and standard deviation of
$\ln\sigma^F_t - \ln\hat{\sigma}_t$ are shown in table
\ref{tab:volvo_15_stat}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
     & \gls{sv} & \gls{garch} & Sample mean \\
     \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0035 & 0.0313 & -0.1448 \\
    \hline
     $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3356 &
     0.3894 & 0.3826 \\
    \hline
 \end{tabular}
  \caption{\small \it Mean and standard deviation of the 3 kinds of
    forecasts}
  \label{tab:volvo_15_stat}
\end{table}
We see that both the \gls{sv} and the \gls{garch} model give biased forecasts, but
the \gls{sv} forecasts are only 1/10 as biased as those of \gls{garch} (0.0035
vs. 0.0313). In addition, the standard deviation of the \gls{sv} forecasts
is smaller too. These results are confirmed by the distribution of
$\ln \sigma^F_t - \ln \hat{\sigma}_t$ and the fraction of ``good''
forecasts, which are shown in figure \ref{fig:volvo_15_diff} and table
\ref{tab:volvo_15_diff}, respectively.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.54, clip=true, trim=27 276 0
    236]{../pics/volvo_15_diff.pdf}
  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
  \label{fig:volvo_15_diff}
\end{figure}
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    \gls{sv} & \gls{garch} & Sample Mean \\
    \hline
    1\% & 58\% & 60\% & 39\% \\
    \hline
    5\% & 84\% & 84\% & 66\% \\
    \hline
    10\% & 97\% & 95\% & 91\% \\
    \hline
  \end{tabular}
  \caption{\small \it Percentage of ``good'' forecasts as defined by
    deviating nore more than 1\%, 5\% and 10\% from the measured
    realized volatility.}
  \label{tab:volvo_15_diff}
\end{table}

\section{Ericsson 15-minute Returns}
\label{sec:ericsson_15min}
In this section we model the volatility of {\it Ericsson B}
15-minute returns during the period 2013/10/10 - 2014/04/04.
Using the same method as with other series, we first find the right
higher-frequency returns that best approximate the volatility of the
15-minute returns. These turn out to be 50-second returns, as figure
\ref{fig:ericsson_15_quotient} shows.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=79 259 108
  121]{../pics/ericsson_15_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -\E(r_t))/
    \sigma_t$. $\sigma_t^2$ is computed as the sum of squared
    50-second returns.}
  \label{fig:ericsson_15_quotient}
\end{figure}

Following the procedure described in previous sections, an ARIMA model is
found for this series:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1- \theta_1B - \theta_2B^2 -
  \theta_3B^3)(1 - \Theta B^s) y_t
\end{eqnarray*}
where the seasonality $s$ is 33 and the other parameter values are
estimated to be those listed in table \ref{tab:ericsson_15_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\Theta$ &
    $\var(y_t)$ \\
    \hline
    Value & 0.8078 & 0.0454 & 0.0943 & 0.8798 & 0.1242 \\
    \hline
  \end{tabular}
  \caption{\small \it Ericsson B log-volatility parameters}
  \label{tab:ericsson_15_params}
\end{table}

Then a \gls{garch} model is also found with parameter values listed in
table \ref{tab:ericsson_15_garch_params}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_{s}$ & $\beta_{s}$ \\
    \hline
    Value & $2.4181 \times 10^{-7}$ & 0.1513 & 0.1409 & 0.6027 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH model of Volvo B 30-minute returns}
  \label{tab:ericsson_15_garch_params}
\end{table}

The forecasts from both models are compared as follows: Table
\ref{tab:ericsson_15_diff1} shows the mean and standard deviation of
the 3 kinds of forecasts:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & \gls{sv} & \gls{garch} & Sample mean \\
    \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0035 &
    0.0600 & 0.1116 \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3278 &
    0.3667 & 0.3590 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:ericsson_15_diff1}
\end{table}
Consistent with previous results, the bias introduced by the \gls{sv} model
is significantly smaller than that introduced by \gls{garch}. Figure
\ref{fig:ericsson_15_diff2} shows the distribution of the difference
between a forecast and its corresponding measured realized volatility,
i.e. $\ln\sigma^F_t - \ln\hat{\sigma}_t$; table
\ref{tab:ericsson_15_diff3} compares the percentage of ``good''
forecasts using the 3 alternatives.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=45 298 21
    165]{../pics/ericsson_15_diff2.pdf}
  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
  \label{fig:ericsson_15_diff2}
\end{figure}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    \gls{sv} & \gls{garch} & Sample Mean \\
    \hline
    1\% & 17\% & 17\% & 16\% \\
    \hline
    5\% & 73\% & 69\% & 69\% \\
    \hline
    10\% & 96\% & 93\% & 94\% \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:ericsson_15_diff3}
\end{table}
It is clear from figure \ref{fig:ericsson_15_diff2} and table
\ref{tab:ericsson_15_diff3} that the \gls{sv} model out-performs \gls{garch}. We
see that the \gls{sv} model yields considerably higher fractions of good
forecasts by all criteria, and in particular, gives much few
under-estimates.

\section{Ericsson 30-minute Returns}\label{sec:ericsson_30min}
In this section we look at the 30-minute returns of Ericsson B during
the period 2013/10/10 - 2014/04/04. Since the methods are the same as
with other series, we shall only present the results here.

First of all, the volatility of this series is found to be well
approximated by realized volatilities computed from 75-second
returns. The normal probability plot is shown in figure
\ref{fig:ericsson_30_quotient}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4, clip=true, trim=79 259 108
  121]{../pics/ericsson_30_quotient.pdf}
  \caption{\small \it Normal probability plot of $(r_t -\E(r_t))/
    \sigma_t$. $\sigma_t^2$ is computed as the sum of squared
    75-second returns.}
  \label{fig:ericsson_30_quotient}
\end{figure}
The following ARIMA model is found for the log-volatility $\ln
\sigma_t$:
\begin{eqnarray*}
  (1-B)(1-B^s) \ln\sigma_t &=& (1- \theta_1B - \theta_2B^2) (1 -
  \Theta B^s) y_t 
\end{eqnarray*}
where the seasonality $s = 16$. The parameter values are estimated to
be those listed in table
\ref{tab:ericsson_30_params}:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\theta_1$ & $\theta_2$ & $\Theta$ &
    $\var(y_t)$ \\
    \hline
    Value & 0.6842 & 0.2470 & 0.8391 & 0.0918 \\
    \hline
  \end{tabular}
  \caption{\small \it Ericsson B log-volatility parameters}
  \label{tab:ericsson_30_params}
\end{table}
A \gls{garch} model is also found with parameters listed in table
\ref{tab:ericsson_30_garch_params}.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Parameter & $\alpha_0$ & $\alpha_1$ & $\alpha_{s}$ & $\beta_{s}$ \\
    \hline
    Value & $7.4856 \times 10^{-7}$ & 0.0532 & 0.1506 & 0.6778 \\
    \hline
  \end{tabular}
  \caption{\small \it GARCH(1, 1) model of Ericsson B 30-minute
    returns}
  \label{tab:ericsson_30_garch_params}
\end{table}

Table \ref{tab:ericsson_30_diff1} shows the mean and standard
deviation of the difference series $\ln\sigma^F_t -
\ln\hat{\sigma}_t$, where $\ln\sigma^F_t$ is the forecast
log-volatility and $\ln\hat{\sigma}_t$ is the measured realized
volatility.
\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    & \gls{sv} & \gls{garch} & Sample mean \\
    \hline
    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.0004 &
    0.0526 & 0.1039 \\
    \hline
    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.2519 &
    0.3019 & 0.2963 \\
    \hline
  \end{tabular}
  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
    \ln\hat{\sigma}_t$}
  \label{tab:ericsson_30_diff1}
\end{table}
Apparently, the \gls{garch} forecasts have a fairly large bias compared with
those of the \gls{sv} model. Turning to figure \ref{fig:ericsson_30_diff2}
and table \ref{tab:ericsson_30_diff3}, we also see the \gls{sv} model
performs consistently better --- a higher percentage of ``good''
forecasts are delivered and, while it under-estimates to around the
same extent as does \gls{garch}, it certainly over-estimates a lot
less. These are confirmative to what we have observed for the other
series.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.5, clip=true, trim=33 281 11
    150]{../pics/ericsson_30_diff2.pdf}
  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
  \label{fig:ericsson_30_diff2}
\end{figure}

\begin{table}[htb!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ &
    \gls{sv} & \gls{garch} & Sample Mean \\
    \hline
    1\% & 22\% & 16\% & 16\% \\
    \hline
    5\% & 78\% & 69\% & 69\% \\
    \hline
    10\% & 98\% & 97\% & 96\% \\
    \hline
  \end{tabular}
  \caption{\small \it Fraction of ``good'' forecasts as defined by
    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
  \label{tab:ericsson_30_diff3}
\end{table}

\section{Volatility of Forecast accuracy}
\label{sec:forecast_volatility}
In this section we compare the volatility of the accuracy of
\gls{garch} and \gls{sv} models' forecasts. Table
\ref{tab:5percent_values} lists the the percentage of forecasts
that deviate less than 5\% from the corresponding realized
volatilities, and also gives the standard deviation of the 6 numbers.
It is seen that the forecasts of \gls{sv} models have standard
deviation 0.0432 while those of \gls{garch} have 0.0722. Clearly the
accuracy of \gls{sv} forecasts varies considerably less than does that
of \gls{garch} forecasts. In other words, a \gls{sv} model performs
more consistently around a level of accuracy.
  \begin{table}[htb!]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      & Volvo 15m & Volvo 30m & Ericsson 15m & Ericsson 30m \\
      \hline
      SV & 84\% & 72\% & 73\% & 78\% \\
      \hline
      GARCH & 84\% & 62\% & 69\% & 69\% \\
      \hline
      \hline
      & Nordea 1 15m & Nordea 2 30m & \multicolumn{2}{c|}{standard deviation}\\
      \hline
      SV & 75\% & 81\% & \multicolumn{2}{c|}{0.0432} \\
      \hline
      GARCH & 71\% & 71\% & \multicolumn{2}{c|}{0.0722} \\
      \hline
    \end{tabular}
    % \begin{tabular}{|c|c|c|c|c|}
    %   \hline
    %   & Volvo 15m & Volvo 30m & Ericsson 15m & Ericsson 30m \\
    %   \hline
    %   SV & 0.8351 & 0.7162 & 0.7268 & 0.7844 \\
    %   \hline
    %   GARCH & 0.8363 & 0.6154 & 0.6850 & 0.6900 \\
    %   \hline
    %   \hline
    %   & Nordea 1 15m & Nordea 2 30m & std.\\
    %   \hline
    %   SV & 0.7457 & 0.8090 & 0.0432 \\
    %   \hline
    %   GARCH & 0.7062 & 0.7109 & 0.0722 \\
    %   \hline
    % \end{tabular}
    \caption{\small \it Comparison of the percentage of forecasts that
      lie within 5\% of the corresponding realized
      volatilities. Nordea 1 refers to the Nordea data set of
      2012/01/16 - 2012/04/20 (see section \ref{chp:nordea_15min})
      and Nordea 2 refers to the data set of 2013/10/10 - 2014/04/04
      (see section \ref{sec:nordea2_30min}).}
    \label{tab:5percent_values}
  \end{table}

\chapter{Normalization Constant in the Unconditional PDF of the
  Symmetric SV model}
\label{chp:symmetric_SV_norm_const}
In the following we work out the normalization constant C used in
section \ref{sec:SLV_Symmetric}:
\begin{eqnarray*}
  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {1 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} \text{erfc} \left({1 \over
      \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma \over
      \sqrt{2}} \right) dx\\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{0}^{\infty} \text{erfc} \left({1 \over
      \sqrt{2}\sigma} \ln{x \over \sqrt{\ln 4}} + {\sigma \over
      \sqrt{2}} \right) dx\\
\end{eqnarray*}
Let
\begin{eqnarray*}
  a &=& {1 \over \sigma \sqrt 2} \\
  b &=& -{1 \over 2 \sigma \sqrt 2}\ln\ln 4 + {\sigma \over \sqrt 2} \\
  y &=& a\ln|x|+b \\
\end{eqnarray*}
Then
\begin{eqnarray*}
  \int_{-\infty}^{\infty} f_{r'}(x) dx &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {e^{(y-b)/a} \over a}
  \text{erfc}(y) \\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \left.e^{(y-b)/a}
    \text{erfc}(y)\right|_{y=-\infty}^{\infty}
  + {2 \over C}{e^{\sigma^2 / 2} \over
    \sqrt{8\pi}} \int_{-\infty}^{\infty} dy {2 \over \sqrt \pi}
  e^{(y-b)/a} e^{-y^2} \\
  &=& {2 \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}  2 e^{\ln\ln 4/2
    - \sigma^2/2}\\
  C &=& \sqrt{2 \ln 4\over \pi} \\
  &\approx& 0.9394
\end{eqnarray*}

\chapter{PDF of the covariance matrix of
auto-correlated Gaussian Returns}
\label{app:pdf_gaussian1}
When autocorrelations exist among the columns of the Gaussian returns
matrix $\mtx R$, the distribution of the covariance matrix $\mtx C =
\mtx{R} \mtx{R'}$ is no longer Wishart but is nonethelss closely
related to it. In the following we consider the situation where
$r_{i,t}$ can be represented as a vector auto-regressive
process. Specifically, suppose
\begin{eqnarray*}
  \vec{r}_t &=& \sum_{k=1}^p \phi_k \vec{r}_{t-k} + \vec{a}_t \\
  \vec{a}_t &=& \vec{r}_t - \sum_{k=1}^p \phi_k \vec{r}_{t-k}
\end{eqnarray*}
where $\vec{a}_t = (a_{1,t}, a_{2,t}, \cdots, a_{N,t})' \sim N(0,
\mtx \Sigma)$ and comprise the columes of $\mtx A$; $\vec{r}_t$ are the
columes of $\mtx R$. Here $N(0, \mtx \Sigma)$ denotes the multivarate
normal distribution with covariance matrix $\mtx \Sigma$. The last equation
can be written in matrix form
\[
\mtx{A} = \mtx{R M}
\]
For example, in the case of an AR(1) process
\begin{equation*}
  \begin{pmatrix}
    a_{1,1} & a_{1,2} & \cdots & a_{1,T} \\
    \vdots & \ddots & \vdots \\
    a_{N,1} & a_{N,2} & \cdots & a_{N,T} \\
  \end{pmatrix} =
  \begin{pmatrix}
    r_{1,1} & r_{1,2} & \cdots & r_{1,T} \\
    \vdots & \ddots & \vdots \\
    r_{N,1} & r_{N,2} & \cdots & r_{N,T} \\
  \end{pmatrix}
  \begin{pmatrix}
    1 & -\phi &   &   & \\
      & 1 & -\phi &   & \\
      &   & \ddots  & \ddots &   \\
      &   &   & 1 & -\phi \\
      &   &   &   & 1 \\
  \end{pmatrix}
\end{equation*}
Let $\mtx{QR} = \mtx{RM}$, then $\mtx{Q} = \mtx{RMR^{-1}}$. Since the
set of $\{r_{ij}\}$ for which $\text{det} \mtx{R} = 0$ has probability
zero, a matrix $\mtx Q$ satisfying the above equation almost surely
exists. Thus $\mtx{A} = \mtx{RM} = \mtx{QR}$ and $\mtx{AA'} =
\mtx{QRR'Q'}$ Since the columns of A are not auto-correlated, $\mtx{AA'}
\sim W(\mtx{\Sigma}, T)$. Then $\mtx{RR'} \sim W(\mtx{Q^{-1} \Sigma
  Q'^{-1}}, T)$ follows from \cite{Anderson2003}, section 7.3.3. Now we
observe the Wishart \gls{pdf}
\begin{equation}
  \label{eq:WishartPDF}
  f_W(\mtx{X}) = { |\det X|^{(T-N-1)/2} \exp\left(-{1 \over 2}\tr
      \mtx{\Sigma^{-1}X} \right)
    \over
    2^{NT/2}\pi^{N(N-1)/4}|\det \mtx{\Sigma}|^{T/2}
    \prod_{i=1}^N \Gamma\left[(n+1-i)/2\right]
  }
\end{equation}
where, assuming $\mtx X = \mtx{UU'}$, $\mtx \Sigma$ is the covariance
matrix of the columns of $\mtx U$. we notice in \ref{eq:WishartPDF}
that $\mtx \Sigma$ enters $f_W(\cdot)$ through $\tr (\mtx{\Sigma^{-1}
  RR'})$ and $\det \mtx{\Sigma}$. $\mtx{RR'}$ enters through $\tr
\mtx{\Sigma^{-1} RR'}$ and $\det \mtx{RR'}$. Clearly
\begin{eqnarray*}
  \det \mtx{Q^{-1}\Sigma Q'^{-1}} &=& \det \mtx{\Sigma \over (\det Q)^2} \\
  &=& \det \mtx{\Sigma} \over (\det \mtx{M})^2 \\
  &=& \det \mtx{\Sigma}
\end{eqnarray*}
As to $\tr \mtx{\Sigma^{-1} C}$, we have
\begin{eqnarray*}
  && \tr\left[\mtx{(Q^{-1}\Sigma Q'^{-1})^{-1} RR'}\right] \\
  &=& \tr\left[\mtx{Q'\Sigma^{-1} Q RR'}\right] \\
  &=& \tr\left[\mtx{R'^{-1}M'R'\Sigma^{-1} RMR^{-1} RR'}\right] \\
  &=& \tr\left[\mtx{\Sigma^{-1} RM (RM)'}\right] \\
  &=& \tr\left[\mtx{\Sigma^{-1} AA'}\right] \\
\end{eqnarray*}

Moreover, $\det \mtx{RR'} = \det \mtx{AM^{-1} M'^{-1} A'} = \det
\mtx{AA'}$. Thus if we use $f_R(\cdot)$ to denote the joint
probability density of the entries of $\mtx{RR'}$ and $f_A(\cdot)$ to
denote that of $\mtx{AA'}$, we can write
\begin{equation}\label{eq:cross-corr-matrix-PDF}
  f_R(\mtx{RR'}) = f_A(\mtx{RMM'R'})
\end{equation}
Substituting for $f_A$ the Wishart \gls{pdf} \ref{eq:WishartPDF}, we
get the \gls{pdf} of $\mtx{RR'}$.

\chapter{Asymptotic Distributions of the Elements of a Covariance
  Matrix of Autocorrelated Gaussian Returns}
\label{chp:gaussian_elements_dist}
In this appendix we derive the approximate distributions of the
elements of a Covariance Matrix of autocorrelated Gaussian returns.

In the following we denote the diagonal elements as $C_{ii}$ and the
non-diagonal elements as $C_{ij}$ ($i \neq j$). We assume that
$\var(a_{i,t}) = \sigma^2$ for any $i$ and $t$, and $a_{i,t}$ are not
autocorrelated, i.e. $\text{corr}(a_{i,t}, a_{i, t'}) = 0$ for any
$i$, $t$ and $t'$. Then we can write
\begin{eqnarray*}
  r_{i,t} r_{j, t} &=& \left[
    \phi r_{i, t-1} + a_{i, t}
  \right] \left[
    \phi r_{j, t-1} + a_{j, t}
  \right] \\
  C_{ij} &=& {1 \over T}\sum_{t=1}^T r_{i,t} r_{j, t} \\
  &=& \phi^2 {1 \over T}\sum_{t=1}^Tr_{i,t-1} r_{j, t-1} +
  \phi{1 \over T}\sum_{t=1}^T\left(
    r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t}
  \right) + {1 \over T}\sum_{t=1}^T a_{i,t} a_{j,t}
\end{eqnarray*}
We note that the two sums $\sum_{t=1}^T r_{i,t} r_{j, t}$ and
$\sum_{t=1}^T r_{i,t-1} r_{j, t-1}$ only differ by the first and the
last addend, which is negligible for sufficiently large T. Thus we
have
\begin{equation}\label{eq:Offdiag1}
  \begin{aligned}
    (1 - \phi^2)C_{ij} &\approx& \phi {1 \over T} \sum_{t=1}^T
    \left(r_{i, t-1} a_{j, t} + r_{j, t-1} a_{i, t} \right)
    + {1 \over T}\sum_{t=1}^T a_{i,t} a_{j,t}
  \end{aligned}
\end{equation}
Now we write the AR(1) process $r_{i,t}$ as an infinite moving-average
process:
\begin{eqnarray*}
  r_{i, t} &=& \phi r_{i, t-1} + a_{i,t} \\
  (1 - \phi B) r_{i, t} &=& a_{i,t} \\
\end{eqnarray*}
where $B$ is the back-shift operator. Then it follows from the above
equation
\begin{eqnarray*}
  r_{i,t} &=& {1 \over 1 - \phi B} a_{i,t} \\
  &=& \sum_{k=0}^\infty \phi^k B^k a_{i,t} \\
  &=& \sum_{k=0}^\infty \phi^k a_{i,t-k} \\
\end{eqnarray*}
where it is left implicit that $a_{i, t}$ with $t \leq 0$ is zero (in
words, this implies that the $r_{i,t}$ process is not affected at all
by events before $t = 1$).

Substituting this into eq.\ref{eq:Offdiag1} for $r_{i,t-1}$ yields
\begin{equation}
  \label{eq:Offdiag2}
  \begin{aligned}
  (1 - \phi^2)C_{ij} &\approx
  {1 \over T} \sum_{t=1}^T \sum_{k=0}^\infty\phi^{k+1} (a_{i, t-k-1}
  a_{j, t} + a_{j, t-k-1} a_{i, t})
  + {1 \over T} \sum_{t=1}^T a_{i,t} a_{j,t} \\
  &=
  {1 \over T} \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi^k (a_{i, t-k}
  a_{j, t} + a_{j, t-k} a_{i, t})
  + {1 \over T} \sum_{t=1}^{T} a_{i,t} a_{j,t}\\
  \end{aligned}
\end{equation}

Given two Gaussian random variables $x$ and $y$ with zero mean and
covariance matrix 
\begin{equation*}
  \Sigma = \sigma^2
  \begin{pmatrix}
    1 & \rho \\
    \rho & 1 \\
  \end{pmatrix}
\end{equation*}

The \gls{pdf} of $xy$ can be found by considering $P(xy < z)$:
\begin{eqnarray*}
  P(xy < z) &=& \left(\int_0^\infty dx \int_{-\infty}^{z/x} dy
    +\int_{-\infty}^0 dx \int_{z/x}^\infty dy \right)
    {1 \over 2\pi\sigma\sqrt{1 - \rho^2}} \\
    &&
    \exp\left[
      -{x^2 -2\rho xy + y^2
        \over
        2\sigma^2 (1 - \rho^2)} 
    \right] \\
  f(z; \sigma, \rho) &=& {d \over dz} P(xy < z)\\
  &=& {1 \over \pi \sigma \sqrt{1 - \rho^2}} \exp\left[
    {\rho z \over \sigma^2 (1 - \rho^2)}\right] K_0\left[
    {|z| \over \sigma^2 (1 - \rho^2)}
  \right]
\end{eqnarray*}
where $K_n(z')$ is the modified Bessel function of the second
kind. It is worth taking note that, when $\rho \neq 0$,
$f(z; \sigma, \rho)$ is not symmetric with respect to
$z$. As a result, if $\rho > 0$, the mean of $f(z;
\sigma, \rho)$ is positive, and vice versa.

Secondly, because $|\rho| < 1$ and
\begin{equation*}
  K_0(z) \sim \sqrt{\pi \over 2z} e^{-z}
\end{equation*}
as $z \to \infty$ \cite{Olver:2010:NHMF}, all the moments of
$f(z; \sigma, \rho)$ are finite, implying the
applicability of the Lyapunov central limit theorem provided that $T$
is large, which is very often the case and what we assume here.

With this in mind, we observe that $\phi$ only affects the first sum in
\ref{eq:Offdiag2}. If $a_{i,t-k}$ and $a_{j,t}$ are not correlated for
non-zero $k$, $\phi$ will not affect the mean of
$C_{ij}$. Furthermore, it is also clear from equation
\ref{eq:Offdiag2} that the variance of $C_{ij}$ is
always increased by a non-zero $\phi$, regardless of the sign of
$\phi$. We compute this increment in the following.

In light of the above expression for $f(z; \sigma,
\rho)$, we rewrite equation \ref{eq:Offdiag2} as
\begin{equation}\label{eq:Cij_dist}
  \begin{aligned}
    (1 - \phi^2)C_{ij} \approx &
    {1 \over T}
    \sum_{t=1}^{T} \sum_{k=1}^{t-1}\phi^k (a_{i, t-k} a_{j, t} + a_{j,
      t-k} a_{i, t}) \\
    & + {1 \over T} \sum_{t=1}^{T} \sigma^2 (1 - \rho^2) {a_{i,t} \over
      \sigma \sqrt{1 - \rho^2}} {a_{j,t} \over \sigma \sqrt{1 - \rho^2}} \\
  \end{aligned}
\end{equation}
In addition, we assume
\begin{equation*}
  \text{corr}(a_{i, t-k}, a_{j, t}) = \left\{
    \begin{array}{l l}
      1 & \text{ if $i=j$ and $k = 0$ }\\
      \rho & -1 < \rho < 1. \text{ if $i \neq j$ and $k = 0$ }\\
      0 & \text{otherwise}
    \end{array}
  \right.
\end{equation*}
Then, because $a_{i, t-k}$ and $ a_{j, t}$ with $i \neq j$ and $k
> 0$ are not correlated, the mean of $a_{i, t-k} a_{j, t}$ is 0
($f(z; \sigma, 0)$ is symmetric), and the variance of it
can be found to be $\sigma^6$ using formula (10.43.19) of \cite{NIST:DLMF}:
\begin{equation*}
  \int_0^\infty dt K_\nu(t) t^{\mu-1} = 2^{\mu-2}
  \Gamma\left(
    {\mu + \nu \over 2}
  \right) \Gamma\left(
    {\mu - \nu \over 2}
  \right)
\end{equation*}
On the other hand, $a_{i,t}/\sigma \sqrt{1 - \rho^2}$ and
$a_{j,t}/\sigma \sqrt{1 - \rho^2}$ have variance $1/(1 - \rho^2)$
and are correlated - $\text{corr}(a_{i,t}, a_{j,t}) = \rho$. The mean
of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ can be found using formula
(10.43.22) of \cite{NIST:DLMF}, given that $-1 < \rho < 1$:
\begin{eqnarray*}
  \int_0^\infty t^{\mu - 1} e^{-at} K_\nu(t) dt &=& (\pi/2)^{1/2}
  \Gamma(\mu + \nu) \Gamma(\mu - \nu)(1 - a^2)^{-\mu/2 + 1/4} \times\\
  && P^{-\mu+1/2}_{\nu-1/2} (a)
\end{eqnarray*}
where $P^\mu_\nu(\cdot)$ is Ferrers function of the first
kind\footnote{ Ferrers function of the first kind is defined through
  the hypergeometric functon $F(a, b; c; z)$ \cite{NIST:DLMF}:
  \begin{equation}
    \label{eq:Ferrers_1st}
    \mathop{\mathsf{P}^{\mu}_{\nu}\/}\nolimits\!\left(x\right)=\left(\frac{1+x}{1-
        x}\right)^{\mu/2}\mathop{\mathbf{F}\/}\nolimits\!\left(\nu+1,-\nu;1-\mu;\tfrac
      {1}{2}-\tfrac{1}{2}x\right).
  \end{equation}
}. The result is
\begin{eqnarray*}
  \E\left[{a_{i,t}a_{j,t} \over \sigma^2 (1 - \rho^2)}\right]
  &=&
  {1 \over \sqrt{2\pi} (1 - \rho^2)^{5/4}} \left[
    P^{-3/2}_{-1/2}(-\rho) - P^{-3/2}_{-1/2}(\rho)
  \right]
\end{eqnarray*}
Similarly, the variance of $a_{i,t}a_{j,t}/\sigma^2 (1 - \rho^2)$ is
found to be
\begin{eqnarray*}
  v^2(\rho) &=&
  {4 \over \sqrt{2\pi} (1 - \rho^2)^{7/4}} \left[
    P^{-5/2}_{-1/2}(\rho) + P^{-5/2}_{-1/2}(-\rho)
  \right] \\
  && - \E^2\left[{a_{i,t}a_{j,t} \over
      \sigma^2 (1 - \rho^2)}\right]
\end{eqnarray*}
Now we can apply the Lyapunov central limit theorem
\cite{Billingsley1995} to the sum in equation \ref{eq:Cij_dist}
and write down the asymptotic Gaussian distribution of $C_{ij}$:
\begin{equation*}
C_{ij} \sim N(\mu'_X, \sigma'_X)
\end{equation*}
where
\begin{eqnarray}
  \mu'_X &=& {\sigma^2 \over \sqrt{2\pi} (1 - \phi^2)(1 -
    \rho^2)^{1/4}} \left[ P^{-3/2}_{-1/2}(-\rho) -
    P^{-3/2}_{-1/2}(\rho)
  \right] \label{eq:gaussian_mean}\\
  \sigma'^2_X &=& {1 \over (1 - \phi^2)^2}\left[
    \sum_{t=1}^T \sum_{k=1}^{t-1} 2\left(
      \phi^k \over T
    \right)^2 \sigma^6 + \sum_{t=1}^T
    {\sigma^4 (1 - \rho^2)^2 \over T^2} v^2(\rho)
  \right] \nonumber \\
  &=& {2 \sigma^6 \over T (1 - \phi^2)^2} \left[
    {\phi^2 \over 1 - \phi^2} -
    {\phi^2 (1 - \phi^{2T}) \over
      T(1 - \phi^2)}
  \right] + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
    T (1 - \phi^2)^2} \nonumber \\
  &\approx& {2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}
  + {\sigma^4 (1 - \rho^2)^2 v^2(\rho) \over
    T (1 - \phi^2)^2} \label{eq:gaussian_variance}
\end{eqnarray}
% Equation \ref{eq:gaussian_mean} tells that, if two return series $i$
% and $j$ are not correlated, auto-correlation in the returns does not
% introduce a bias into the estimation of the covariance; if,
% however, the return series are indeed correlated, auto-correlation
% in the returns rescales the covariance through a multiplicative
% factor $1/(1 - \phi^2)$.

% In addition, equation \ref{eq:gaussian_variance} tells that
% auto-correlation in the returns always makes the covariance
% estimation more noisy. Auto-correlation not only rescales the variance
% of the no-autocorrelation estimation by $1/(1 - \phi^2)$ but even
% adds an extra term ${2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}$.

We may apply a similar treatment to the diagonal elements of the
covariance matrix, which we denote as $C_{ii}$ here:
\begin{eqnarray*}
  C_{ii} &=& {1 \over T} \sum_{t=1}^T r^2_{it} \\
  &=& {1 \over T} \sum_{t=1}^T \sum_{l=0}^{t-1}\phi^l a_{i, t-l}
  \sum_{k=0}^{t-1}\phi^k a_{i, t-k} \\
  &=& {1 \over T} \sum_{t=1}^T \left[
    \sum_{k=0}^{t-1} \phi^{2k} a^2_{i, t-k} +
    \sum_{k,l = 0}^{t-1} \phi^{k+l} a_{i, t-k} a_{i, t-l}
  \right]
\end{eqnarray*}
By the Lyapunov central limit theorem under the assumption of large T,
the asymptotic distribution of $C_{ii}$ is Gaussian, the mean and
variance being
\begin{eqnarray}
  \E(C_{ii}) &=& {1 \over T}\left[
    \sum_{k=0}^{t-1} \phi^{2k} \sigma^2
  \right] \nonumber \\
  &=& {\sigma^2 \over (1 - \phi^2) T} \left[
    T - {\phi^2 (1 - \phi^{2T}) \over 1 - \phi^2}
  \right] \nonumber \\
  & \approx & {\sigma^2 \over 1 - \phi^2} \left[
    1 - {\phi^2 \over T}
  \right] \label{eq:gaussian_cii_mean}
\end{eqnarray}
and
\begin{eqnarray}
  \var(C_{ii}) &=& \sum_{t=1}^T \left[
    \sum_{k=0}^{t-1} {\phi^{4k} \sigma^4 \over T^2} 2 +
    \sum_{k,l=0}^{t-1} {\phi^{2(k+l)} \over T^2} \sigma^6
  \right] \nonumber \\
  &=& \sum_{t=1}^T \left[
    {2 \sigma^4 \over T^2} {1 - \phi^{4t} \over 1 - \phi^4} +
    {\sigma^6 \over T^2} \left(
      {1 - \phi^{2t} \over 1- \phi^2}
    \right)^2 \right] \nonumber \\
  &=& {2 \sigma^4 \over T (1 - \phi^4)} -
  {2 \sigma^4 \phi^4 (1 - \phi^{4T}) \over T^2(1 - \phi^4)^2} +
  \nonumber \\
  && {\sigma^6 \over T (1 -\phi^2)^2} -
  {2 \sigma^6 \phi^2 (1 - \phi^{2T}) \over T^2 (1 - \phi^2)^3} +
  {\sigma^6 \phi^4 (1 - \phi^{4T}) \over T^2 (1 - \phi^2)^2 (1 -
    \phi^4)} \nonumber \\
  &\approx& {2 \sigma^4 \over T (1 - \phi^4)} + {\sigma^6 \over T (1
    -\phi^2)^2} \label{eq:gaussian_cii_variance}
\end{eqnarray}
% Equation \ref{eq:gaussian_mean} tells that, if two return series $i$
% and $j$ are not correlated, auto-correlation in the returns does not
% introduce a bias into the estimation of the covariance; if,
% however, the return series are indeed correlated, auto-correlation
% in the returns rescales the covariance through a multiplicative
% factor $1/(1 - \phi^2)$.

% In addition, equation \ref{eq:gaussian_variance} tells that
% auto-correlation in the returns always makes the covariance
% estimation more noisy. Auto-correlation not only rescales the variance
% of the no-autocorrelation estimation by $1/(1 - \phi^2)$ but even
% adds an extra term ${2 \sigma^6 \phi^2 \over T (1 - \phi^2)^3}$.

\chapter{Eigenvalues Distribution of Wishart Matrix}
\label{sec:wishart_eigen_dist}
In this section we summarize the analytic results regarding the
eigenvalue distribution of a Wishart matrix $\mtx{RR'}$. In the
simplest case where the elements of $\mtx R$ are all independent of
each other, i.e. neither auto-correlation nor cross-correlation
exists, we have \cite{Chiani2012}
\begin{eqnarray}
  f(x_1, \cdots, x_N) &=& K \prod_{i=1}^N e^{-x_i/2}
  x_i^{(T-N-1)/2} \prod_{i<j}^N (x_i -
  x_j) \label{eq:wishart_eigen_pdf}
\end{eqnarray}
where the eigenvalues have been indexed in descending order: $x_1 \geq
x_2 \geq \cdots \geq x_N$. The normalization constant K is given by
\begin{equation*}
  K = {\pi^{N^2/2} \over 2^{NT/2}\Gamma_N(T/2) \Gamma_N(N/2)}
\end{equation*}
where the function $\Gamma_m(a)$ is defined as
\begin{equation*}
  \Gamma_m(a) = \pi^{m(m-1)/4} \prod_{k=1}^m \Gamma\left(a - {k-1
      \over 2}\right)
\end{equation*}
If the order of the eigenvalues is ignored, the \gls{pdf} of their
distribution is given by the Marcenko-Pastur law \cite{Guhr2007}
\begin{eqnarray}
  f(x) &=& {1 \over 2\pi \sigma^2 q} {
    \sqrt{(x_2 - x)(x - x_1)} \over x
  } \label{eq:MP_pdf}
\end{eqnarray}
where it is assumed $r_{it} \sim N(0, \sigma^2)$ and
\begin{eqnarray*}
  q &=& \lim_{N,T \to \infty} {N \over T} \\
  x_1 &=& \sigma^2 (1 - \sqrt q)^2 \\
  x_2 &=& \sigma^2 (1 + \sqrt q)^2 \\
\end{eqnarray*}
It is imposed that $q = N/T < 1$.

Moreover, K. Johnsson \cite{Johnsson2000} and I. Johnstone
\cite{Johnstone2001} showed that, at the {\it absence} of
autocorrelations and in the asymptotic limit $N, T \to \infty$, $N/T
\to q < \infty$, the maximum eigenvalue  $\lambda_1$ follows the
Tracy-Widom distribution (denote $\mathscr{TW}_1$ here) when properly
relocated and rescaled:
\begin{equation*}
  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} \sim \mathscr{TW}_1
\end{equation*}
where $\mu_{NT}$ and $\sigma_{NT}$ are given by
\begin{eqnarray*}
  \mu_{NT} &=& \left(
    \sqrt{N-1/2} + \sqrt{T - 1/2}
  \right)^2 \\
  \sigma_{NT} &=& \sqrt{\mu_{NT}} \left(
    {1 \over \sqrt{N-1/2}} + {1 \over \sqrt{T-1/2}}
  \right)^{1/3}
\end{eqnarray*}
The $\mathscr{TW}_1$ distribution has the following cummulative
distribution function (CDF) \cite{Chiani2012}
\begin{equation*}
  F_1(x) = \exp\left[
    -{1 \over 2} \int_x^\infty dy \left(q(y) + (y-x)q^2(y)\right)
  \right]
\end{equation*}
where $q(y)$ is defined as the solution to the Painlev\'e II differential
equation
\begin{equation*}
  q''(y) = yq(y) + 2q^3(y)
\end{equation*}
which is unique when imposing the condition
\begin{equation*}
  q(y) \sim \text{Ai}(y) \text{ as } y \to \infty
\end{equation*}

Then Marco Chiani showed recently that the $\mathscr{TW}_1$ distribution
can be well approximated by a gamma distribution based on his proof
that the exact distribution of the maximum eigenvalue is a mixture of
gamma distributions. Specifically,
\begin{equation}\label{eq:TracyWidom-Gamma}
  {\lambda_1 - \mu_{NT} \over \sigma_{NT}} + \alpha \sim
  \mathscr{G}(k ,\theta)
\end{equation}
where $\mathscr{G}(k, \theta)$ denotes the Gamma distribution with
parameters $k$ and $\theta$ \cite{Chiani2012}. The \gls{pdf} of the
gamma distribution is given by
\begin{equation*}
  f_\gamma(x; k, \theta) = {1 \over \Gamma(k) \theta^k} x^{k-1}
  e^{-x/\theta}
\end{equation*}
Moreover, the first 3 moments of the distribution are simple:
\begin{eqnarray*}
  \text{mean} &=& k\theta \\
  \text{variance} &=& k\theta^2 \\
  \text{skewness} &=& 2/\sqrt{k}
\end{eqnarray*}

% \chapter{Collection of Figures and Tables}

% \begin{figure}[htb!]
%   \centering
%   \subfigure[ACF of $z_t$]{
%     \includegraphics[scale=0.4, clip=true, trim=95 236 118
%     200]{../pics/nordea_15min_quotient_acf.pdf}
%     \label{fig:nordea_15min_quotient_acf}
%   }
%   \subfigure[ACF of $z_t^2$]{
%     \includegraphics[scale=0.4, clip=true, trim=95 236 118
%     200]{../pics/nordea_15min_quotient_squared_acf.pdf}
%     \label{fig:nordea_15min_quotient_squared_acf}
%   }
%   \caption{\small \it Nordea 15min $z_t$ and $z_t^2$ auto-correlation
%     function (ACF).}
% \end{figure}

% \begin{figure}[htb!]
%   \centering
%   \subfigure[Normal Probability Plot of $y_t$]{
%     \includegraphics[scale=0.4, clip=true, trim=76 256 109
%     214]{../pics/nordea2_y_normplot.pdf}
%     \label{fig:nordea_15min_y_qq}
%   }
%   \subfigure[ACF of $y_t$]{
%     \includegraphics[scale=0.4, clip=true, trim=92 258 104
%     218]{../pics/nordea_15min_y_acf.pdf}
%     \label{fig:nordea_15min_y_acf}
%   }
%   \caption{\small \it Nordea 15min $y_t$ normal probability plot and
%     autocorrelations}
% \end{figure}

### 2014-09-29 thesis/draft4.tex
Add the draft of the paper.

\documentclass{article}
\usepackage{amssymb,amsmath,amsthm,bm}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}

\input{../physics_common.tex}
\DeclareGraphicsExtensions{.eps, .pdf}

\author{
  Prof. Sven \AA berg \\
  Xie Xiaolei}
\date{\today}
\title{Effect of Heteroscedasticity in Covariance Matrix}

\begin{document}
\maketitle

\section{Separation of the Heteroscedasticity Effect}
Consider a covariance matrix constructed as
$$
C_{ij} = {1 \over T}\sum_{t=1}^T r_{it} r_{jt}
$$
where $1 \leq i \leq N$. Adopt a stochastic volatility model for
$r_{it}$, i.e.
$$
r_{it} = \sigma_{it} \eta_{it}
$$
We can write
\begin{eqnarray}
C_{ij} &=& {1 \over T}\sum_{t=1}^T \sigma_{it} \eta_{it} \sigma_{jt}
\eta_{jt} \nonumber \\
\bm{C} &=& {1 \over T} (\bm{\sigma * \eta}) (\bm{\sigma *
  \eta})' \label{eq:C}
\end{eqnarray}
where the matrices $\bm{\sigma}$ and $\bm{\eta}$ have elements
$\sigma_{it}$ and $\eta_{it}$ respectively, and * denotes element-wise
multiplication. Due to the cyclic property of matrix trace, the
moment generating function of $\bm{C}$, $M_C(z)$, relates to $M_D(z)$,
the moment generating function of $\bm D$
\begin{eqnarray*}
  \bm{D} &=& {1 \over T} (\bm{\sigma * \eta})' (\bm{\sigma * \eta})
\end{eqnarray*}
through the equation
\begin{eqnarray*}
  M_C(z) &=& {1 \over q} M_D(z)
\end{eqnarray*}
where $q = N/T$. For later convenience, we make use of $T \times T$
matrices $\bm{\tilde{\sigma}}$ and $\bm{\tilde{\eta}}$, as well as a
projector
$$
\bm{P} = \text{diag}(\underbrace{1, \cdots, 1}_{\text{N 1's}}, 
\underbrace{0, \cdots, 0}_{\text{T-N 0's}})
$$
The first N rows of $\tilde{\bm{\sigma}}$ and $\bm{\tilde{\eta}}$ are
precisely those of $\bm{\sigma}$ and $\bm{\eta}$. This way, we have
\begin{eqnarray*}
\bm D &=& {1 \over T} (\bm{\sigma * \eta})' (\bm{\sigma * \eta}) \\
&=& {1 \over T} (\bm{\tilde{\sigma} * \tilde{\eta}})' \bm P'
\bm P (\bm{\tilde{\sigma} * \tilde{\eta}}) \\
&=& {1 \over T} (\bm{\tilde{\sigma} * \tilde{\eta}})'
\bm P (\bm{\tilde{\sigma} * \tilde{\eta}}) \\
\end{eqnarray*}
Again, by the cyclic property of matrix trace, the spectral properties
of the RHS of the last equation is equivalent to those of
\begin{eqnarray*}
  \bm E &=& {1 \over T} (\bm{\tilde{\sigma} * \tilde{\eta}}) (\bm{\tilde{\sigma}
    * \tilde{\eta}})' \bm P \\
\end{eqnarray*}
In the simple situation where $\sigma_{it} =
\sigma_{jt} = \sigma_t$ for some $\sigma_t$ and for all $i, j, t$,
\begin{eqnarray*}
  \bm{\tilde \sigma * \tilde \eta} &=& \bm{\tilde \eta}
  \begin{pmatrix}
    \sigma_1 &        & \\
        & \ddots & \\
        &        & \sigma_T
  \end{pmatrix} \\
  &=& \bm{\tilde \eta \bar \sigma}
\end{eqnarray*}
Thus
\begin{eqnarray}\label{eq:E_def}
  \bm E &=& {1 \over T}\bm{\tilde \eta \bar \sigma^2 \tilde \eta' P}
\end{eqnarray}
In the large T limit $\bm{\tilde \eta \bar \sigma^2 \tilde \eta'}$ and $P$ are
freely independent, therefore their S-transforms are multiplicative,
i.e.
\begin{eqnarray}
  S_E(z) &=& S_{\tilde \eta \bar \sigma^2 \tilde \eta'/T}(z) S_P(z)
  \nonumber \\
  &=& S_{\tilde \eta' \tilde \eta /T}(z) S_{\bar \sigma^2}(z) S_P(z) \label{eq:S_E}
\end{eqnarray}
For $S_{\tilde \eta' \tilde \eta /T}(z)$ there have been results when
the elements of $\bm{\tilde \eta}$ are identically distributed with finite
second moment \cite{burda2011} or alternatively, follow $\alpha$-stable
distribution \cite{politi2010}.

So our focus hereafter is on the spectral properties of $\bar{
\bm \sigma^2}$. The Green's function of $\bar{\bm \sigma^2}$
can be immediately written down:
\begin{eqnarray*}
G_{\bar \sigma^2} &=& \frac{1}{T}\sum_{t=1}^T \frac{1}{z - \sigma_t^2}
\end{eqnarray*}
Suppose the $\sigma_t^2$ process has correlation time $\tau$,
i.e. $\text{corr}(\sigma_t^2, \sigma^2_{t+\tau}) \approx 0$. Then we
have
\begin{eqnarray*}
G_{\bar \sigma^2} &=& {1 \over T}\sum_{n=0}^{\tau-1}\sum_{m=1}^{T/\tau}
\frac{1}{z - \sigma_{m\tau + n}^2} \\
&=& {1 \over \tau}\sum_{n=0}^{\tau-1} {1 \over T/\tau}
\sum_{m=1}^{T/\tau} \frac{1}{z - \sigma_{m\tau + n}^2} \\
&=& {1 \over \tau}\sum_{n=0}^{\tau-1}
\E_m\left(\frac{1}{z - \sigma_{m\tau + n}^2}\right)
\end{eqnarray*}
where $\E_m$ denotes averaging over the index $m$. On condition that
$\sigma_t^2$ is a stationary process, $\E_m[1 / (z -
    \sigma_{m\tau + n}^2)]$ is independent of $n$. Thus
\begin{eqnarray}\label{eq:greens}
G_{\bar \sigma^2} &=& \E_m\left(
  \frac{1}{z - \sigma_{m\tau + n}^2}
\right)
\end{eqnarray}
So autocorrelations of $\sigma_t^2$'s affect spectral properties
only through the stationary distribution of $\sigma_t^2$.
In the following subsections we discuss situations where $\sigma_t$'s
are chosen from different conditional distributions and have different
autocorrelation structures.
% Then the moment generating function $M_{\bar \sigma^2}$ is
% \begin{eqnarray*}
%   M_{\bar \sigma^2} &=& z G_{\bar \sigma^2} - 1 \\
%   &=& \frac{1}{T} \sum_{t=1}^T \frac{1}{1 - \sigma_t^2/z} - 1\\
%   &=& \frac{1}{T} \sum_{n=0}^{\infty} \frac{1}{z^n}\sum_{t=1}^T
%   \sigma_t^{2n} - 1\\
%   &=& \frac{1}{T} \sum_{n=1}^{\infty} \frac{1}{z^n}\sum_{t=1}^T
%   \sigma_t^{2n}
% \end{eqnarray*}

\section{Normal innovations \& log-normal volatilities with AR(1)
  autocorrelation}
In this subsection we consider the case
\begin{eqnarray}
  \log \sigma_t &=& \phi\log \sigma_{t-1} + x_t \nonumber \\
  &=& \sum_{l = 0}^{\infty} \phi^l x_{t-l} \label{eq:AR}
\end{eqnarray}
where $x_t \sim N(0, 1)$ and we have assumed that the $\log \sigma_t$
process was started in the infinite past with $\log \sigma_{-\infty} =
x_{-\infty}$. In addition, we assume $\log \sigma_t$ is a stationary
process with $|\phi| < 1$. Clearly $\log \sigma_t$ is normally
distributed with mean 0 and variance ${1 / (1 -
  \phi^2)}$. According to \eqref{eq:greens}, the spectral density of
$\bm E$ as appears in \eqref{eq:E_def} is the same as if
$\sigma_t^2$'s were uncorrelated but $\log \sigma_t \sim N(0,
1/(1 - \phi^2))$.

To find the spectral density of $\bm C$ with normally distributed $\bm
\eta$ and uncorrelated, log-normally distributed $\sigma_t$, we use
the result of Biroli et al in \cite{biroli2007student}, which gives
the Blue function of $\bm C$ as
\begin{eqnarray*}
  B(z) &=& \frac{1}{z} + \int \frac{\sigma ^2 P(s) ds}{1-q \sigma ^2 z}
\end{eqnarray*}
where $q = N/T$ and $\sigma$ has the stationary distribution of
$\sigma_t$, whose probability density function is $P(s)$. Now we are
specifically considering lognormally distributed $\sigma_t$, so we
rewrite the above equation as
\begin{eqnarray}
B(z) &=& \int_{-\infty }^{\infty } \frac{e^{2 s} \exp \left(-\frac{s^2}{2
      v}\right)}{\sqrt{2 \pi  v} \left(1-q e^{2 s} z\right)} \,
ds+\frac{1}{z} \label{eq:LognormalBlue}
\end{eqnarray}
where $v = 1/(1 - \phi^2)$, reflecting the effect
of autocorrelation in $\log \sigma_t$'s. However, no closed form of
this integral is known to the author. To obtain the spectral density,
which is given by the Sokhotski-Weierstrass formula as
\begin{eqnarray}
  \rho(\lambda) &=& -{1 \over \pi} \lim_{\epsilon \to 0} \im G(\lambda +
  i\epsilon) \label{eq:Sokhotski-Weierstrass}
\end{eqnarray}
we equate the real part of RHS of \eqref{eq:LognormalBlue} to $\lambda \in
R$ and the imaginery part to 0, then we can solve \eqref{eq:LognormalBlue}
numerically and obtain $\rho(\lambda)$ via
\eqref{eq:Sokhotski-Weierstrass}. Specifically we solve the following
equation system:
\begin{eqnarray}
\lambda &=& \int_{-\infty}^{\infty} \frac{ds}{\sqrt{2\pi v}}
\frac{
  e^{-s^2/2v}(e^{-2s} - qx)
}{
  (e^{-2s} - qx)^2 + q^2 y^2  
} +
\frac{x}{x^2 + y^2} \label{eq:LognormalBlueReal}\\
0 &=& \int_{-\infty}^{\infty} \frac{ds}{\sqrt{2\pi v}}
\frac{
  e^{-s^2/2v} q
}{
  (e^{-2s} - qx)^2 + q^2 y^2  
} -
\frac{1}{x^2 + y^2} \label{eq:LognormalBlueImag}
\end{eqnarray}
This is the method used by Biroli et al in
\cite{biroli2007student}. Figure \ref{fig:LognormalBlue}
shows the plot of the Blue function \eqref{eq:LognormalBlue}. $-\pi
\rho(\lambda)$ is the projection of the contour $\im B(z) = 0$ onto the $\im
G$ axis.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.45, clip=true, trim=66 0 84
    0]{../pics/LognormalBlue.eps}
  \caption{\small \it The Blue function $B$. q = 0.45, v = 3}
  \label{fig:LognormalBlue}
\end{figure}
The spectral density function $\rho(\lambda)$ obtained as numerical
solutions is shown in figure \ref{fig:LognormalSpectra} and compared
to the Marcenko-Pastur law with several values of $q = N/T$ and $v$.
\begin{figure}[htb!]
  \centering
  \subfigure[$q = 0.1$]{
    \includegraphics[scale=0.35]{../pics/Lognormal_q0.1.eps}
    \label{fig:Lognormal_q0.1}
  }
  \subfigure[$q = 0.2$]{
    \includegraphics[scale=0.35]{../pics/Lognormal_q0.2.eps}
    \label{fig:Lognormal_q0.2}
  }
  \subfigure[$q = 0.5$]{
    \includegraphics[scale=0.35]{../pics/Lognormal_q0.5.eps}
    \label{fig:Lognormal_q0.5}
  }
  \subfigure[$q = 1.0$]{
    \includegraphics[scale=0.35]{../pics/Lognormal_q1.0.eps}
    \label{fig:Lognormal_q1.0}
  }
  % \subfigure[$v=1$, $q = 1$]{
  %   \includegraphics[scale=0.35]{../pics/Lognormal_v1_q1.eps}
  %   \label{fig:Lognormal_v1_q1}
  % }
  % \subfigure[$v=1$, $q = 0.1$]{
  %   \includegraphics[scale=0.35]{../pics/Lognormal_v1_q0.1.eps}
  %   \label{fig:Lognormal_v1_q0.1}
  % }
  % \subfigure[$v=0.0001$, $q = 1$]{
  %   \includegraphics[scale=0.35]{../pics/Lognormal_v0.0001_q1.eps}
  %   \label{fig:Lognormal_v0.0001_q1}
  % }
  % \subfigure[$v=0.0001$, $q = 0.1$]{
  %   \includegraphics[scale=0.35]{../pics/Lognormal_v0.0001_q0.1.eps}
  %   \label{fig:Lognormal_v0.0001_q0.1}
  % }
  \caption{\small \it Spectra with lognormal volatilities. The
    empirical probability density functions are plotted as stairs. The
  density function given by the Marcenko-Pastur law is plotted as a
  thick red curve.}
  \label{fig:LognormalSpectra}
\end{figure}
From figures \ref{fig:LognormalSpectra} one can see that the
Marcenko-Pastur law is recovered when $v \to 0$, as expected.

Now we consider equation \eqref{eq:LognormalBlueReal} and
\eqref{eq:LognormalBlueImag}. As $\lambda$ tends to $\lambda_\M$ with
$\lambda_\M \leq \infty$, $y \to 0$. Meanwhile, $x$ can neither tend
to 0 nor to $\infty$: If $x \to 0$, then $\frac{1}{x^2 + y^2} \to
\infty$ while the integral in \eqref{eq:LognormalBlueImag} tends to
$e^{8v}$ -- thus \eqref{eq:LognormalBlueImag} cannot hold. If instead
$x \to \infty$, RHS of \eqref{eq:LognormalBlueReal} tends to 0, failing
\eqref{eq:LognormalBlueReal}. So $x$ must tend to a finite value as $\lambda
\to \lambda_\M$. This limiting value, call it $a$, can be found from
\eqref{eq:LognormalBlueImag} by setting $y=0$:
\begin{eqnarray}
{1 \over a^2} &=& \int_{-\infty}^{\infty} \frac{ds}{\sqrt{2\pi v}}
\frac{
  e^{-s^2/2v} q
}{
  (e^{-2s} - qa)^2
} \label{eq:LognormalBlueImag2}
\end{eqnarray}
Clearly $a$ is a function of $v$ and $q$. Denote this function $f$
such that $a = f(v, q)$. As has been argued, $f(v,q)$ is finite for
all $v > 0$ and $0 < q \leq 1$. Using $f$, the integral in
\eqref{eq:LognormalBlueReal}, in the limit $\lambda \to \lambda_\M$,
can be expressed
\begin{eqnarray*}
  {1 \over \sqrt{2\pi v}} \int_{-\infty}^{\infty} \frac{
    e^{-2s - s^2/2v}
    }{
      (e^{-2s} - qa)^2
    } ds &=& {1 \over q e^{2v} f^2(v, qe^{-4v})}
\end{eqnarray*}
It immediately follows from \eqref{eq:LognormalBlueReal}
\begin{eqnarray*}
  \lambda_\M &=& {1 \over q e^{2v} f^2(v, qe^{-4v})}
\end{eqnarray*}
Since $q \leq 1$ and $e^{-2v} < 1$, $f(v, qe^{-2v})$ and hence
$\lambda_\M$ are finite. The spectrum is bounded.

The function $f(v,q)$ as defined by the largest root of
\eqref{eq:LognormalBlueImag2} is plotted in figure
\ref{fig:FFunction1}, seen as the intersection between the two curves,
which represent the two sides of \eqref{eq:LognormalBlueImag2},
respectively.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.35]{../pics/FFunctionOverview.eps}
    \label{fig:FFunctionOverview}
  }
  \subfigure[]{
    \includegraphics[scale=0.35]{../pics/FFunction1.eps}
    \label{fig:FFunction1}
  }
  %% \subfigure[$q = 0.5$]{
  %%   \includegraphics[scale=0.35]{../pics/FFunction3.eps}
  %%   \label{fig:FFunction2}
  %% }
  \caption{\small \it The implicit function $f(v, q)$ plotted in
    log-log scale. Blue: the integral $\int_{-\infty}^{\infty}
    \frac{ds}{\sqrt{2\pi v}}
    \frac{
      e^{-s^2/2v} q
    }{
      (e^{-2s} - qa)^2
    }$as a function of $a$. Green: the function $1/a^2$. v=0.1, q=0.1.
}
  \label{fig:FFunction}
\end{figure}

% To see the effect of heteroscedasticity we turn to simulations. Figure
% \ref{fig:LognormalSpectraAutocorrelated} shows the empirical spectral
% density when the conditional log-volatilities are given by \eqref{eq:AR}.
% \begin{figure}[htb!]
%   \centering
%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.01$,
%   $\phi = 0$, $\text{var}(\ln\sigma_t) = 0.01$]{
%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.1_phi0.eps}
%     \label{fig:Lognormal_sig0.1_phi_0}
%   }
%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.01$,
%   $\phi = 0.25$, $\text{var}(\ln\sigma_t) = 0.0107$]{
%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.1_phi0.25.eps}
%     \label{fig:Lognormal_sig0.1_phi_0.25}
%   }
%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.01$,
%   $\phi = 0.5$, $\text{var}(\ln\sigma_t) = 0.0133$]{
%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.1_phi0.5.eps}
%     \label{fig:Lognormal_sig0.1_phi_0.5}
%   }
%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.25$,
%   $\phi = 0$, $\text{var}(\ln\sigma_t) = 0.25$]{
%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.5_phi0.eps}
%     \label{fig:Lognormal_sig0.5_phi_0}
%   }
%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.25$,
%   $\phi = 0.25$, $\text{var}(\ln\sigma_t) = 0.2667$]{
%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.5_phi0.25.eps}
%     \label{fig:Lognormal_sig0.5_phi_0.25}
%   }
%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.25$,
%   $\phi = 0.5$, $\text{var}(\ln\sigma_t) = 0.3333$]{
%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.5_phi0.5.eps}
%     \label{fig:Lognormal_sig0.5_phi_0.5}
%   }
%   \caption{\small \it Spectra with autocorrelated normal log-volatilities.}
%   \label{fig:LognormalSpectraAutocorrelated}
% \end{figure}
% It is seen that the empirical spectral density agrees very well with
% the density function obtained as numerical solutions to $B(\re G -
% i\pi\rho) = \lambda$.

\bibliographystyle{unsrt}
\bibliography{econophysics}
\end{document}

### 2014-09-29 thesis/popular_abstract.tex
Add the draft of the paper.

\documentclass{article}
\author{Xie Xiaolei \\
Supervised by Prof. Sven \AA berg}

\title{Volatility and Correlations --- Eyes that the wise have in their head}
\begin{document}
\maketitle

Can we predict the movements of the financial market? The direct answer is "no", but there is more to be said. One cannot predict the price movement in an efficient market, but one can indeed predict how "volatile" the prices are (volatility). This does not lead to profits, but is useful for risk management --- if one pursues a profit, one must take a risk. Knowing and quantitatively managing the risks associated with investments, one often makes wiser decisions.

This report entails, first of all, a comparison of return models in the context of intraday returns. 6 return series are studied: Nordea, Volvo and Ericsson in the time scales of 15 and 30 minutes. Stochastic Volatility (SV) models and GARCH models are fitted to these series and the resulting volatility forecasts are compared. SV models are found to be more accurate in most cases.

Covariance matrices, which give the correlation between different companies, are also studied. In particular, the eigenvectors and eigenvalues of a covariance matrix, which respectively represent a set of driving factors of the companies' shares and the variances of these factors, are studied. It is found that when the return series are auto-correlated, i.e. when observations at different times are correlated, the eigenvectors are delocalized --- their expansion as a linear combination of the basis vectors involve more appreciable coefficients.
\end{document}

### 2014-09-29 thesis/summary2.tex
Add the draft of the paper.

\chapter{Conclusions}
\label{chp:summary2}
In the thesis project we have studied the volatilities of a total of 7
intraday return series: 15-minute returns of Nordea Bank during the
period 2012/01/16 - 2012/04/20 and 15-minute and 30-minute returns of
Nordea Bank, Volvo B and Ericsson B in the period 2013/10/10 -
2014/04/04. Every series is investigated with a stochastic volatility
(SV) model and a GARCH model. In fact, when we fit the GARCH model, 3
GARCH variants have been considered: the basic GARCH, EGARCH, and
GJR. But in all cases, the basic GARCH stands out to be the most
accurate in forecasting the realized volatilities.

For comparison, each model makes one-step-ahead forecasts and
then the two series of forecasts are compared. The quality of a
forecast series $\ln\sigma^F_t$ is judged by $\ln\sigma^F_t - \ln
\hat{\sigma}_t$, where $\hat{\sigma}_t$ is the measured realized
volatility. For each series of $\ln\sigma^F_t - \ln \hat{\sigma}_t$, we
have compared the mean, standard deviation, cumulative distribution
function, complementary distribution function, as well as the
percentage of ``good'' forecasts, where ``good'' is defined by
deviating less than 1\%, 5\% or 10\% from the realized
volatility. What we find through the studies is the following:
\begin{itemize}
\item In all the 7 studied series, the log-volatility $\ln \sigma_t$
  is well described by a seasonally integrated moving average
  model. Long memory in these series is accounted for by the compounded
  difference operator $(1-B)(1-B^s)$ where $B$ denotes the back-shift
  operator. $s$ at here stands for seasonality, which is 33 in the
  cases of 15-minute returns and 16 in the cases of 30-minute
  returns.

\item An SV model generally yields more accurate forecasts than does
  the GARCH model for the same series. This is certainly well expected,
  considering that the SV model incorporates much more data than does
  GARCH. However, it must be noted that the validity of the
  aforementioned comparison is underlain by the accuracy of realized
  volatility as a proxy to the true conditional volatility.

\item The accuracy of forecasts varies more for GARCH than for SV
  models. For example, the percentage of forecasts
  that lie within 5\% of the corresponding realized volatilities has
  the values listed in table \ref{tab:5percent_values}.
  \begin{table}[htb!]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      & Volvo 15m & Volvo 30m & Ericsson 15m & Ericsson 30m \\
      \hline
      SV & 0.8351 & 0.7162 & 0.7268 & 0.7844 \\
      \hline
      GARCH & 0.8363 & 0.6154 & 0.6850 & 0.6900 \\
      \hline
      \hline
      & Nordea 1 15m & Nordea 2 15m & Nordea 2 30m & std.\\
      \hline
      SV & 0.7457 & 0.7887 & 0.8090 & 0.0432 \\
      \hline
      GARCH & 0.7062 & 0.7436 & 0.7109 & 0.0722 \\
      \hline
    \end{tabular}
    \caption{\small \it Comparison of the percentage of forecasts that
      lie within 5\% of the corresponding realized
      volatilities. Nordea 1 refers to the Nordea 15-minute returns
      studied in chapter \ref{chp:nordea_15min} and Nordea 2 refers to
      the 15-minute and 30-minute returns studied in chapter
      \ref{chp:Nordea2}. 15m and 30m are acronyms for 15-minute and
      30-minute, respectively.}
    \label{tab:5percent_values}
  \end{table}
  The standard deviation of this percentage is 0.0432 for SV models
  but 0.0722 for GARCH.

\end{itemize}

In chapter \ref{chp:SLV_unconditional} we have presented our
calculation of the unconditional distribution functions of the
stochastic volatility models used throughout the project, i.e. the
distribution of $\mu + e^{\bar{v} + \sigma a}b$, where $a, b \sim N(0,
\mtx \Sigma)$. In the simplest case where $\mtx \Sigma = \mtx I$, this
is simply the product of a lognormal and a normal variate, but when
$\mtx \Sigma \neq \mtx I$, the distribution becomes more
complicated. In the case $\mtx \Sigma \neq \mtx I$, we obtained the
moment generating function and calculated the first four moments of
the distribution. The parameters of the unconditional distribution are
estimated with respect to the 30-minute returns of Nordea Bank, Volvo
B and Ericsson B. The resulting distributions are seen to fit the data
very nicely.

### 2014-09-29 m/LognormalEndsFun1.m

@@ -2,7 +2,9 @@
 % q: scalar
 % a: scalar or column vector
 % return: scalar or column vector
-function [f, fder] = LognormalEndsFun1(v, q, a)
+% function [f, fder] = LognormalEndsFun1(v, q, a)
+function f = LognormalEndsFun1(v, q, a)
+a = reshape(a, length(a), 1);
 L = 4000;
 s = linspace(-6*sqrt(v), 6*sqrt(v), L);
 M1 = exp(-s.^2./2./v)*(s(2) - s(1))/sqrt(2*pi*v);
@@ -11,5 +13,5 @@ M2 = repmat(exp(-2.*s), length(a), 1) - q.*repmat(a, 1, L);
 M3 = 1./M2.^2;
 f = M3 * M1' * q; 
 
-M4 = 1./M2.^3;
-fder = M4 * M1' .* (2*q^2);
+% M4 = 1./M2.^3;
+% fder = M4 * M1' .* (2*q^2);

### 2014-09-29 m/LognormalEndsFun3.m

@@ -1,6 +1,6 @@
 % function [f, fder] = LognormalEndsFun3(v, q, a)
 function lam = LognormalEndsFun3(v, q, a)
-L = 4000;
+L = 6000;
 s = linspace(-6*sqrt(v), 6*sqrt(v), L);
 M1 = exp(-s.^2./2./v - 2*s)*(s(2) - s(1))/sqrt(2*pi*v);
 M2 = repmat(exp(-2.*s), length(a), 1) - q.*repmat(a, 1, L);

### 2014-09-29 m/test3.m

@@ -15,27 +15,28 @@ texts = {};
 c = 1;
 hold on
 q = 0.1;
-%for sig = [0.1, 0.5, 1]
+% for sig = [0.1, 0.5, 1]
 for sig = 0.5
+    v = sig^2;
     load(sprintf(['../matfys/data/sv/normal_ret/lognormal_vol/q%.1f/Eig-' ...
                   'sig%.4f-phi%.4f.mat'], q, sig, 0));
 
     % ev = reshape(ev, 1, prod(size(ev)));
     eigmax = max(ev)';
-    [lam, den] = epdf(eigmax, 1, min(eigmax), max(eigmax), 400, spec{c});
-    
+    [lam, den] = epdf(eigmax, 4, min(eigmax), max(eigmax), 400, spec{c});
+    % g = LognormalGreen(mean(eigmax), v, q);
     % [Z2, G2] = epdf(ev, 1, min(ev), 10, 100, '');
 
     % lammax1= max(ev);
-    [u, lammax2] = LognormalEnds(sig^2, q);
-    a = LognormalEndsConstA(sig^2, q);
-    lammax3 = LognormalEndsFun3(sig^2, q, a);
-    % a = LognormalEndsConstA(sig^2, q);
-    texts{c} = sprintf('\\lambda_{max} = %.2f', lammax2);
-    % Z1 = linspace(max(min(ev),0.05), max(ev), 1000);
-
-    
+    % [u, lammax2] = LognormalEnds(v, q);
+    A = LognormalEndsConstA(v, q*exp(-4*v), 0.1);
+    lammax2 = 1/exp(2*v)/q/A^2;
+    a = LognormalEndsConstA(v, q, 10^(-1));
+    lammax3 = LognormalEndsFun3(v, q, a);
+    % texts{c} = sprintf('\\lambda_{max} = %.2f', lammax2);
     % Z1 = linspace(lammax1, lammax2, 1000);
+
+    % Z1 = linspace(max(ev)*0.9, max(ev), 1000);
     % G1 = LognormalGreen(Z1, sig.^2, q);
     % plot(Z1, real(G1), spec{c});
     % texts{c} = sprintf('a = %e', a);
@@ -45,8 +46,8 @@ for sig = 0.5
     % plot(Z1, -imag(G1)./pi, spec{c}, 'LineWidth', 1);
 
     % legend('Theoretical', 'Simulation');
-    % texts{2*c-1} = sprintf('v = %.2f, em', sig^2);
[…]

### 2014-11-04 thesis/draft4.tex
New scripts for verifying the theory of Mikosch.

@@
 \frac{
-  e^{-s^2/2v} q
+  e^{-s^2/2v} q y
 }{
@@
 } -
-\frac{1}{x^2 + y^2} \label{eq:LognormalBlueImag}
+\frac{y}{x^2 + y^2} \label{eq:LognormalBlueImag}
 \end{eqnarray}
@@
 Now we consider equation \eqref{eq:LognormalBlueReal} and
-\eqref{eq:LognormalBlueImag}. As $\lambda$ tends to $\lambda_\M$ with
-$\lambda_\M \leq \infty$, $y \to 0$. Meanwhile, $x$ can neither tend
-to 0 nor to $\infty$: If $x \to 0$, then $\frac{1}{x^2 + y^2} \to
-\infty$ while the integral in \eqref{eq:LognormalBlueImag} tends to
-$e^{8v}$ -- thus \eqref{eq:LognormalBlueImag} cannot hold. If instead
-$x \to \infty$, RHS of \eqref{eq:LognormalBlueReal} tends to 0, failing
-\eqref{eq:LognormalBlueReal}. So $x$ must tend to a finite value as $\lambda
-\to \lambda_\M$. This limiting value, call it $a$, can be found from
-\eqref{eq:LognormalBlueImag} by setting $y=0$:
+\eqref{eq:LognormalBlueImag}. When $q y$ and $v$ are both small, RHS
+of \eqref{eq:LognormalBlueReal} and \eqref{eq:LognormalBlueImag} have
+two sharp peaks, one at $s = 0$ and the other at $s = -{1 \over
+  2}\ln(qx)$. We expand the integrands in
+\eqref{eq:LognormalBlueReal} and \eqref{eq:LognormalBlueImag} around
+these two peaks. For \eqref{eq:LognormalBlueImag} We write
 \begin{eqnarray}
-{1 \over a^2} &=& \int_{-\infty}^{\infty} \frac{ds}{\sqrt{2\pi v}}
-\frac{
-  e^{-s^2/2v} q
-}{
-  (e^{-2s} - qa)^2
-} \label{eq:LognormalBlueImag2}
+  {y \over x^2 + y^2} &=& \int_{-a}^{a} {e^{-s^2/2v}ds \over
+    \sqrt{2\pi v}} \left[\frac{qy}{(1-qx)^2 + q^2y^2}\right] \label{eq:lambda}\\
+  && - \int_{-b}^{b} {dt \over \sqrt{2\pi v}} {1 \over 2(t + q x)}
+  \exp \left[-\frac{\log ^2(q x+t)}{8v}\right] {qy \over t^2 +
+    q^2y^2} \nonumber \\
+  &=& \erf\left({a \over \sqrt{2\pi}}\right)\left[\frac{qy}{(1-qx)^2 +
+      q^2 y^2}\right] \nonumber \\
+  && -{1 \over \sqrt{2\pi v}}\left\{
+  \exp\left(-{\ln^2qx \over 8v}\right) \frac{x \cot ^{-1}\left(\frac{q
+      y}{b}\right)+y \coth ^{-1}\left(\frac{q x}{b}\right)}{q x^2+q
+    y^2} \right. \nonumber \\
+  && \left.-\exp\left(-{\ln^2qx \over 8v}\right)
+  \frac{y \left[x \log \left(1-\frac{2 b}{b+q x}\right)+2 y
+    \cot^{-1}\left(\frac{q y}{b}\right)\right]}{2
+    \left(x^2+y^2\right)} {\ln qx \over 4qvx}
+  \right\} \nonumber
 \end{eqnarray}
-Clearly $a$ is a function of $v$ and $q$. Denote this function $f$
-such that $a = f(v, q)$. As has been argued, $f(v,q)$ is finite for
-all $v > 0$ and $0 < q \leq 1$. Using $f$, the integral in
-\eqref{eq:LognormalBlueReal}, in the limit $\lambda \to \lambda_\M$,
-can be expressed
-\begin{eqnarray*}
-  {1 \over \sqrt{2\pi v}} \int_{-\infty}^{\infty} \frac{
-    e^{-2s - s^2/2v}
-    }{
-      (e^{-2s} - qa)^2
-    } ds &=& {1 \over q e^{2v} f^2(v, qe^{-4v})}
-\end{eqnarray*}
-It immediately follows from \eqref{eq:LognormalBlueReal}
-\begin{eqnarray*}
-  \lambda_\M &=& {1 \over q e^{2v} f^2(v, qe^{-4v})}
-\end{eqnarray*}
-Since $q \leq 1$ and $e^{-2v} < 1$, $f(v, qe^{-2v})$ and hence
-$\lambda_\M$ are finite. The spectrum is bounded.
+where $a$ and $b$ are small positive numbers. In particular, $b < q x$
+such that $t + qx > 0$. Similarly, for \eqref{eq:LognormalBlueReal} we have
+\begin{eqnarray}
+  \lambda &=& {x \over x^2 + y^2} + \erf\left({a \over \sqrt{2v}}\right){1 - q x
+    \over (1 - q x)^2 + q^2 y^2 } \nonumber \\
+  && - {1 \over \sqrt{2\pi v}}
+  \left\{\exp\left(-{\ln^2qx \over 8v}\right) \frac{x \log
+    \left(1-\frac{2 b}{b+q x}\right)+2 y \cot ^{-1}\left(\frac{q
+      y}{b}\right)}{2 q \left(x^2+y^2\right)} \right. \nonumber \\
+  &&  
+  \left. -\frac{\log (q x)}{4 q v x} \exp\left(-{\ln^2qx \over
+    8v}\right) \frac{x \left[x \coth ^{-1}\left(\frac{q x}{b}\right)-y
+    \cot ^{-1}\left(\frac{q y}{b}\right)\right]}{x^2+y^2}
+  \right\} \label{eq:constraint}
+\end{eqnarray}
+%% Now we consider equation \eqref{eq:LognormalBlueReal} and
+%% \eqref{eq:LognormalBlueImag}. As $\lambda$ tends to $\lambda_\M$ with
+%% $\lambda_\M \leq \infty$, $y \to 0$. Meanwhile, $x$ can neither tend
+%% to 0 nor to $\infty$: If $x \to 0$, then $\frac{1}{x^2 + y^2} \to
+%% \infty$ while the integral in \eqref{eq:LognormalBlueImag} tends to
+%% $e^{8v}$ -- thus \eqref{eq:LognormalBlueImag} cannot hold. If instead
+%% $x \to \infty$, RHS of \eqref{eq:LognormalBlueReal} tends to 0, failing
+%% \eqref{eq:LognormalBlueReal}. So $x$ must tend to a finite value as $\lambda
+%% \to \lambda_\M$. This limiting value, call it $a$, can be found from
+%% \eqref{eq:LognormalBlueImag} by setting $y=0$:
+%% \begin{eqnarray}
+%% {1 \over a^2} &=& \int_{-\infty}^{\infty} \frac{ds}{\sqrt{2\pi v}}
+%% \frac{
+%%   e^{-s^2/2v} q
+%% }{
+%%   (e^{-2s} - qa)^2
+%% } \label{eq:LognormalBlueImag2}
+%% \end{eqnarray}
+%% Clearly $a$ is a function of $v$ and $q$. Denote this function $f$
+%% such that $a = f(v, q)$. As has been argued, $f(v,q)$ is finite for
+%% all $v > 0$ and $0 < q \leq 1$. Using $f$, the integral in
+%% \eqref{eq:LognormalBlueReal}, in the limit $\lambda \to \lambda_\M$,
+%% can be expressed
+%% \begin{eqnarray*}
+%%   {1 \over \sqrt{2\pi v}} \int_{-\infty}^{\infty} \frac{
+%%     e^{-2s - s^2/2v}
+%%     }{
+%%       (e^{-2s} - qa)^2
+%%     } ds &=& {1 \over q e^{2v} f^2(v, qe^{-4v})}
+%% \end{eqnarray*}
+%% It immediately follows from \eqref{eq:LognormalBlueReal}
+%% \begin{eqnarray*}
+%%   \lambda_\M &=& {1 \over q e^{2v} f^2(v, qe^{-4v})}
+%% \end{eqnarray*}
+%% Since $q \leq 1$ and $e^{-2v} < 1$, $f(v, qe^{-2v})$ and hence
+%% $\lambda_\M$ are finite. The spectrum is bounded.
 
-The function $f(v,q)$ as defined by the largest root of
-\eqref{eq:LognormalBlueImag2} is plotted in figure
-\ref{fig:FFunction1}, seen as the intersection between the two curves,
-which represent the two sides of \eqref{eq:LognormalBlueImag2},
-respectively.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[]{
-    \includegraphics[scale=0.35]{../pics/FFunctionOverview.eps}
-    \label{fig:FFunctionOverview}
-  }
-  \subfigure[]{
-    \includegraphics[scale=0.35]{../pics/FFunction1.eps}
-    \label{fig:FFunction1}
-  }
-  %% \subfigure[$q = 0.5$]{
-  %%   \includegraphics[scale=0.35]{../pics/FFunction3.eps}
-  %%   \label{fig:FFunction2}
+%% The function $f(v,q)$ as defined by the largest root of
+%% \eqref{eq:LognormalBlueImag2} is plotted in figure
+%% \ref{fig:FFunction1}, seen as the intersection between the two curves,
+%% which represent the two sides of \eqref{eq:LognormalBlueImag2},
+%% respectively.
+%% \begin{figure}[htb!]
+%%   \centering
+%%   \subfigure[]{
+%%     \includegraphics[scale=0.35]{../pics/FFunctionOverview.eps}
+%%     \label{fig:FFunctionOverview}
 %%   }
-  \caption{\small \it The implicit function $f(v, q)$ plotted in
-    log-log scale. Blue: the integral $\int_{-\infty}^{\infty}
-    \frac{ds}{\sqrt{2\pi v}}
-    \frac{
-      e^{-s^2/2v} q
-    }{
-      (e^{-2s} - qa)^2
-    }$as a function of $a$. Green: the function $1/a^2$. v=0.1, q=0.1.
-}
-  \label{fig:FFunction}
-\end{figure}
+%%   \subfigure[]{
+%%     \includegraphics[scale=0.35]{../pics/FFunction1.eps}
+%%     \label{fig:FFunction1}
+%%   }
+%%   %% \subfigure[$q = 0.5$]{
+%%   %%   \includegraphics[scale=0.35]{../pics/FFunction3.eps}
+%%   %%   \label{fig:FFunction2}
+%%   %% }
+%%   \caption{\small \it The implicit function $f(v, q)$ plotted in
+%%     log-log scale. Blue: the integral $\int_{-\infty}^{\infty}
+%%     \frac{ds}{\sqrt{2\pi v}}
+%%     \frac{
+%%       e^{-s^2/2v} q
+%%     }{
+%%       (e^{-2s} - qa)^2
+%%     }$as a function of $a$. Green: the function $1/a^2$. v=0.1, q=0.1.
+%% }
+%%   \label{fig:FFunction}
+%% \end{figure}
 

### 2014-11-04 m/get_mysql.m

@@ -7,6 +7,12 @@ if exist('/nfs/users3/xiexiaol', 'dir') == 7
     if isempty(strfind(javaclasspath, jdbc))
         javaaddpath(jdbc);
     end
+elseif exist('/home/lxb353', 'dir') == 7
+    hostname = '83.176.196.41';
+    jdbc = ['/home/lxb353/myspace/lib/mysql-connector-java-5.1.33-bin.jar'];
+    if isempty(strfind(javaclasspath, jdbc))
+        javaaddpath(jdbc);
+    end
 else
     hostname = 'localhost';
 end

### 2014-12-01 thesis/draft4.tex
Modified matlab scripts for Lund project.

--- 2014-09-29 thesis/draft4.tex
@@
 $$
-where $1 \leq i \leq N$. Adopt a stochastic volatility model for
-$r_{it}$, i.e.
+where $r_{it}$ is the log-return of the $i$-th stock at time $t$. $i =
+1, 2, \dots, N$, $t = 1,2,\dots,T$. Here and in the rest of the
+text we adopt a stochastic volatility model for $r_{it}$, i.e.
 $$
@@
 $$
-We can write
+where $\sigma_{it}$ symbolizes the volatility of the return while
+$\eta_{it}$ is assumed a standard normal random variable.
+So we can write
 \begin{eqnarray}
@@
 moment generating function of $\bm{C}$, $M_C(z)$, relates to $M_D(z)$,
-the moment generating function of $\bm D$
+the moment generating function of
 \begin{eqnarray*}
@@
 the elements of $\bm{\tilde \eta}$ are identically distributed with finite
-second moment \cite{burda2011} or alternatively, follow $\alpha$-stable
+second moment \cite{burda2011} or alternatively, follow a stable
 distribution \cite{politi2010}.
@@
   }
-  % \subfigure[$v=1$, $q = 1$]{
-  %   \includegraphics[scale=0.35]{../pics/Lognormal_v1_q1.eps}
-  %   \label{fig:Lognormal_v1_q1}
-  % }
-  % \subfigure[$v=1$, $q = 0.1$]{
-  %   \includegraphics[scale=0.35]{../pics/Lognormal_v1_q0.1.eps}
-  %   \label{fig:Lognormal_v1_q0.1}
-  % }
-  % \subfigure[$v=0.0001$, $q = 1$]{
-  %   \includegraphics[scale=0.35]{../pics/Lognormal_v0.0001_q1.eps}
-  %   \label{fig:Lognormal_v0.0001_q1}
-  % }
-  % \subfigure[$v=0.0001$, $q = 0.1$]{
-  %   \includegraphics[scale=0.35]{../pics/Lognormal_v0.0001_q0.1.eps}
-  %   \label{fig:Lognormal_v0.0001_q0.1}
-  % }
   \caption{\small \it Spectra with lognormal volatilities. The
@@
 
-Now we consider equation \eqref{eq:LognormalBlueReal} and
-\eqref{eq:LognormalBlueImag}. As $\lambda$ tends to $\lambda_\M$ with
-$\lambda_\M \leq \infty$, $y \to 0$. Meanwhile, $x$ can neither tend
-to 0 nor to $\infty$: If $x \to 0$, then $\frac{1}{x^2 + y^2} \to
-\infty$ while the integral in \eqref{eq:LognormalBlueImag} tends to
-$e^{8v}$ -- thus \eqref{eq:LognormalBlueImag} cannot hold. If instead
-$x \to \infty$, RHS of \eqref{eq:LognormalBlueReal} tends to 0, failing
-\eqref{eq:LognormalBlueReal}. So $x$ must tend to a finite value as $\lambda
-\to \lambda_\M$. This limiting value, call it $a$, can be found from
-\eqref{eq:LognormalBlueImag} by setting $y=0$:
+When $v$ is small, the integrands in \eqref{eq:LognormalBlueReal} and
+\eqref{eq:LognormalBlueImag} have sharp maxima at $v = 0$. Hence it is
+useful to approximate them by Taylor expansion. Let's write
+\begin{eqnarray*}
+  \lambda &=& \int _{-\infty }^{\infty }\!g \left( s,v \right) F \left( s
+  \right) {ds}+{\frac {x}{{x}^{2}+{y}^{2}}} \\
+  0 &=& \int _{-\infty }^{\infty }\!g \left( s,v \right) G \left( s
+  \right) {ds}-{\frac {y}{{x}^{2}+{y}^{2}}}
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+g &=& \frac{1}{\sqrt{2\mathop{\rm  }\pi \mathop{\rm  }v }}e^{-\frac{s
+    ^{2}}{2\mathop{\rm  }v }}\\
+F \left(s \right) &=& \frac{\left(e^{-2\mathop{\rm  }s }-q \mathop{\rm
+    }x \right)}{\left(e^{-2\mathop{\rm  }s }-q \mathop{\rm  }x
+  \right)^{2}+q ^{2}\mathop{\rm  }y ^{2}}\mathop{\rm  } \\
+G(s) &=& \frac{q \mathop{\rm  }y }{\left(e^{-2\mathop{\rm  }s }-q
+    \mathop{\rm  }x \right)^{2}+q ^{2}\mathop{\rm  }y ^{2}}\\
+\end{eqnarray*}
+
+Taylor-expanding $F(s)$ and $G(s)$ to 2nd order in $s$ and plugging
+them into \eqref{eq:LognormalBlueReal} and
+\eqref{eq:LognormalBlueImag} gives
 \begin{eqnarray}
-{1 \over a^2} &=& \int_{-\infty}^{\infty} \frac{ds}{\sqrt{2\pi v}}
-\frac{
-  e^{-s^2/2v} q
-}{
-  (e^{-2s} - qa)^2
-} \label{eq:LognormalBlueImag2}
+\lambda &=& {-q x+1 \over (-q x+1)^2+q^2 y^2} + {v \over 2}\left\{
+{4 \over (-q x+1)^2+q^2 y^2} -{24 (-q x+1) \over
+  [(-q x+1)^2+q^2 y^2]^2} \right. \nonumber \\
+&& \left. + {32 (-q x+1)^3 \over [(-q x+1)^2+q^2 y^2]^3} -8 {(-q
+    x+1)^2 \over [(-q x+1)^2+q^2 y^2]^2}
+\right\} \label{eq:lambda_linear}\\
+0 &=& {q y \over (-q x+1)^2+q^2 y^2} + {v \over 2} \left\{
+  {32 q y (-q x+1)^2 \over [(-q x+1)^2+q^2 y^2]^3}
+  -{8 q y \over [(-q x+1)^2+q^2 y^2]^2} \right. \nonumber \\
+  && \left.
+    -{8 q y (-q x+1) \over [(-q x+1)^2+q^2 y^2]^2}
+\right\} \label{eq:constraint_linear}
 \end{eqnarray}
-Clearly $a$ is a function of $v$ and $q$. Denote this function $f$
-such that $a = f(v, q)$. As has been argued, $f(v,q)$ is finite for
-all $v > 0$ and $0 < q \leq 1$. Using $f$, the integral in
-\eqref{eq:LognormalBlueReal}, in the limit $\lambda \to \lambda_\M$,
-can be expressed
-\begin{eqnarray*}
-  {1 \over \sqrt{2\pi v}} \int_{-\infty}^{\infty} \frac{
-    e^{-2s - s^2/2v}
-    }{
-      (e^{-2s} - qa)^2
-    } ds &=& {1 \over q e^{2v} f^2(v, qe^{-4v})}
-\end{eqnarray*}
-It immediately follows from \eqref{eq:LognormalBlueReal}
-\begin{eqnarray*}
-  \lambda_\M &=& {1 \over q e^{2v} f^2(v, qe^{-4v})}
-\end{eqnarray*}
-Since $q \leq 1$ and $e^{-2v} < 1$, $f(v, qe^{-2v})$ and hence
-$\lambda_\M$ are finite. The spectrum is bounded.
-
-The function $f(v,q)$ as defined by the largest root of
-\eqref{eq:LognormalBlueImag2} is plotted in figure
-\ref{fig:FFunction1}, seen as the intersection between the two curves,
-which represent the two sides of \eqref{eq:LognormalBlueImag2},
-respectively.
+Now that we look for a solution in the limit $y \to 0$, we set $y = 0$
+in \eqref{eq:lambda_linear}; in \eqref{eq:constraint_linear} we first
+cancel the factor $y$ and then set $y = 0$. This yields
+\begin{eqnarray}
+  \lambda &=& {1 \over (-q x+1)} + {2 v (q x+1) \over (-q x+1)^3} + {1
+    \over x} \label{eq:lambda_1} \\
+  0 &=& {q \over (-q x+1)^2} + {4 v q(q x+2) \over (-q x+1)^4} - {1
+    \over x^2} \label{eq:constraint_1}
+\end{eqnarray}
+Evaluating \eqref{eq:lambda_1} at $x = \lambda^{MP}_{+} = (1 +
+\sqrt{q})^2$ gives a rough estimate about the expected maximum
+eigenvalue, respectively. That is
+\begin{eqnarray}
+  \lambda_{+} &=& \lambda_{+}^{MP} [1+2v(2\sqrt q +
+  1)] \label{eq:lambda_3}
+\end{eqnarray}
+This result is checked against simulations, as shown in figure
+\ref{fig:lambda_verified}. Clearly one can see that the linear
+approximation \eqref{eq:lambda_3} fits fairly well to the estimated
+$\E \lambda_\M$.
 \begin{figure}[htb!]
   \centering
-  \subfigure[]{
-    \includegraphics[scale=0.35]{../pics/FFunctionOverview.eps}
-    \label{fig:FFunctionOverview}
+  \subfigure[$q = 0.1$]{
+    \includegraphics[scale=0.35]{../pics/lambda_max_q0.1.eps}
+    \label{fig:lambda_max_q0.1}
   }
-  \subfigure[]{
-    \includegraphics[scale=0.35]{../pics/FFunction1.eps}
-    \label{fig:FFunction1}
+  \subfigure[$q = 0.2$]{
+    \includegraphics[scale=0.35]{../pics/lambda_max.eps}
+    \label{fig:lambda_max_q0.2}
   }
-  %% \subfigure[$q = 0.5$]{
-  %%   \includegraphics[scale=0.35]{../pics/FFunction3.eps}
-  %%   \label{fig:FFunction2}
-  %% }
-  \caption{\small \it The implicit function $f(v, q)$ plotted in
-    log-log scale. Blue: the integral $\int_{-\infty}^{\infty}
-    \frac{ds}{\sqrt{2\pi v}}
-    \frac{
-      e^{-s^2/2v} q
-    }{
-      (e^{-2s} - qa)^2
-    }$as a function of $a$. Green: the function $1/a^2$. v=0.1, q=0.1.
-}
-  \label{fig:FFunction}
+  \caption{\small \it Estimated $\E \lambda_\M$ against its linear
+    approximation.}
+  \label{fig:lambda_verified}
 \end{figure}
-
-% To see the effect of heteroscedasticity we turn to simulations. Figure
-% \ref{fig:LognormalSpectraAutocorrelated} shows the empirical spectral
-% density when the conditional log-volatilities are given by \eqref{eq:AR}.
-% \begin{figure}[htb!]
-%   \centering
-%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.01$,
-%   $\phi = 0$, $\text{var}(\ln\sigma_t) = 0.01$]{
-%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.1_phi0.eps}
-%     \label{fig:Lognormal_sig0.1_phi_0}
-%   }
-%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.01$,
-%   $\phi = 0.25$, $\text{var}(\ln\sigma_t) = 0.0107$]{
-%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.1_phi0.25.eps}
-%     \label{fig:Lognormal_sig0.1_phi_0.25}
-%   }
-%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.01$,
-%   $\phi = 0.5$, $\text{var}(\ln\sigma_t) = 0.0133$]{
-%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.1_phi0.5.eps}
-%     \label{fig:Lognormal_sig0.1_phi_0.5}
-%   }
-%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.25$,
-%   $\phi = 0$, $\text{var}(\ln\sigma_t) = 0.25$]{
-%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.5_phi0.eps}
-%     \label{fig:Lognormal_sig0.5_phi_0}
-%   }
-%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.25$,
-%   $\phi = 0.25$, $\text{var}(\ln\sigma_t) = 0.2667$]{
-%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.5_phi0.25.eps}
-%     \label{fig:Lognormal_sig0.5_phi_0.25}
-%   }
-%   \subfigure[$q = 0.1$, $\text{var}(x_t) = 0.25$,
-%   $\phi = 0.5$, $\text{var}(\ln\sigma_t) = 0.3333$]{
-%     \includegraphics[scale=0.23]{../pics/Lognormal_sig0.5_phi0.5.eps}
-%     \label{fig:Lognormal_sig0.5_phi_0.5}
-%   }
-%   \caption{\small \it Spectra with autocorrelated normal log-volatilities.}
-%   \label{fig:LognormalSpectraAutocorrelated}
-% \end{figure}
-% It is seen that the empirical spectral density agrees very well with
-% the density function obtained as numerical solutions to $B(\re G -
-% i\pi\rho) = \lambda$.
 

### 2014-12-01 summary.tex

\documentclass{article}
\input{physics_common}
\title{Mathematical Aspects of the Capital Market}
\author{Xie Xiaolei}
\date{\today}
\begin{document}
\begin{enumerate}
\item The Nordea 15min returns are well described by the model:
\begin{eqnarray*}
  r_t &=& \sigma_t a_t \\
  (1-B)(1-B^s)(\ln\sigma_t - \mean{\ln\sigma}) &=& (1-\theta B)(1-\Theta
  B^s) \alpha_t
\end{eqnarray*}
where $a_t \sim N(0, 1)$ and $\alpha_t$ is distributed as Johnson Su:
\[
\alpha_t = m + c\sinh{z - a \over b}
\]
Here $z \sim N(0, 1)$. $s$ is the number of seasons in the above
seasonal model. $s=33$ for this particular series. $34 \times
15\text{min} = 8.5\text{hr}$, a full trading day in Stockholm.

\begin{enumerate}
\item 
Andersen and Bollerslev et al reported that $\ln\sigma_t$ is Gaussian
distributed for the exchange rate they studied. However, the
unconditional distribution of $\ln\sigma_t$ in our case is evidently
non-Gaussian.

\item
The residuals $\alpha_t$ have no significant auto-correlation. Their
distribution is, however, not Gaussian but rather fat-tailed and
right-skewed.

Nevertheless, the distribution of $\sigma_t$ is indeed leptokurtic and
right-skewed as Andersen and Bollerslev et al reported for the
exchange rates.

\item
$\sigma_t$ can be approximated by the realized volatility,
which is the sum of squared returns that are sampled at a higher
frequency. The optimal frequency for this is debatable. I choose the
frequency with which $a_t$ is best fitted to a standard Gaussian, and
it turns out to be 30sec for 15min returns.

\item
Right now I am developing programs to estimate the above
seasonal model. Functions for ARIMA model estimation exist in Matlab,
but only for residuals distributed as Gaussian or Student's t. Both
fail to describe the residuals in our case.

\end{enumerate}
\item The Nordea 30min and 45min returns haven't been studied in
  detail but are expected to follow ARIMA seasonal models. Preliminary
  results suggest $\ln\sigma_t$ for 45min returns are Gaussian
  distributed. 
\end{enumerate}
\end{document}

### 2014-12-01 thesis/Nordea_15min.tex

--- 2014-05-10 thesis/price_models.tex
@@
-\chapter{Return Models}\label{chp:PriceModels}
-In this chapter we review some of the discrete-time return models and
-fit them to intraday returns. The intention is to compare these models
-in terms of forecast accuracy and to understand their statistical
-properties. In the following we first describe these models briefly,
-then in
-
-% \S\ref{sec:Garch_model} reviews the GARCH model using the Nordea
-% 15-minute returns as an example. \S\ref{sec:SLV_model} examines the
-% stochastic log-volatility model and discusses the possibility of
-% describing the residuals of the log-volotility series using the
-% Johnson Su distribution. The same Nordea return series is fitted as an
-% example. \S\ref{sec:XieCalc} derives the unconditional distribution
-% function of the stochastic log-volotility model and compares the
-% analytic results to the 30-minute returns of a number of Swedish
-% stocks.
-
-\begin{enumerate}
-\item{\bf Gaussian Distribution}
-
-The justification of modeling return series as independent,
-identically distributed Gaussian variates comes from imagining the
-price process $S(t)$ as a geometric Brownian motion whose increment
-$\sigma dw_t$ at each time step is independent and scales as $\sigma
-\sqrt{dt}$. Then, by adding a drift term $\mu dt$ that represents some
-deterministic trend in the price process, one can express the
-price $S(t)$ as a stochastic differential equation:
-\begin{eqnarray*}
-  dS &=& S\mu dt + S\sigma dw_t \\
-  d(\ln S) &=& (\mu - \frac{1}{2} \sigma^2)dt + \sigma dw_t \\
-\end{eqnarray*}
-It follows from this equation that $\ln S(t) - \ln S_0$, where $S_0$
-is the price at time 0, has Gaussian distribution with mean $(\mu -
-\sigma^2/2)t$ and variance $\sigma^2 t$. Therefore, in a discrete-time
-model, where the length of each time step $\Delta t$ is fixed, the
-return $r_t = \ln S(t + \Delta t) - \ln S(t)$ is assumed to be
-Gaussian distributed and have mean and variance that are functions of
-$\Delta t$.
-
-The picture depicted above is of course overly simplified, and the
-distribution of returns is not really Gaussian. Nevertheless, the
-assumption of Gaussian distributed returns underlies such important
-theories as Black and Scholes theory of option pricing. Thus, albeit
-inaccurate in describing very large returns, the Gaussian distribution
-as a returns model shall not be forgotten. In the next few sections,
-we discuss some more realistic returns models.
-
-\item{\bf GARCH models}
-
-``GARCH'' is the acronym for ``Generalized Autoregressive Conditional
-Heteroscedasticity''. A GARCH(p, q) model is defined by the following
-equation system \cite{Bollerslev86}:
-\begin{equation}
-  \label{eq:garch_def}
-  \begin{aligned}
-    r_t &= \mu_t + \epsilon_t \\
-  \end{aligned}
-\end{equation}
-where $p$ and $q$ are constant integers; $\mu(t)$ is the mean process
-and considered a small constant for intraday returns. $\epsilon_t$ is
-called the innovation of the return and is a variate whose conditional
-distribution \footnotemark is Gaussian and has variance $\sigma^2_t$:
-\begin{eqnarray*}
-  \sigma_t^2 &=& \alpha_0 + \sum_{i=1}^q \alpha_i \epsilon_{t-i}^2 +
-  \sum_{i=1}^p \beta_i \sigma_{t-i}^2
-\end{eqnarray*}
-\footnotetext{
-By conditional distribution we mean the distribution conditional on
-the variate's history: The conditional distribution at time $t$ means
-the distribution conditional on the history up to time $t-1$
-in the discrete-time setting.
-}
-
-It can be shown that the autocorrelation function of $\epsilon_t^2$ is
-given by \cite{Bollerslev87}:
-\begin{eqnarray*}
-\text{corr}(\epsilon_t^2, \epsilon^2_{t-n}) &=& \sum_{i=1} ^{p \vee q}
-(\alpha_i + \beta_i) \rho_{n-i} \text{ for $n > p$}
-\end{eqnarray*}
-where $\alpha_i$ with $i > q$ and $\beta_i$ with $i > p$ are taken as
-zeros. $p \vee q$ denotes the maximum of p and q. From these
-equations, it is clear that the partial autocorrelation function cuts
-off at $\max(p, q)$.
-
-\item {\bf Stochastic Volatility Models}\label{sec:SLV_model}
-
-For the purpose of intraday returns, we specify the stochastic
-volatility model as
-\begin{eqnarray}
-   r_{t, t-h} &=& \ln S_{t} - \ln S_{t-h} \nonumber \\
-   r_{t, t-h} &=& \mu + \sigma_{t, t-h} b_t \label{eq:SLV_spec}
-\end{eqnarray}
-where $S_t$ is the price of the asset at time $t$; $b_t \sim N(0,
-1)$; $r_{t, t-h}$ is the return over the time interval $[t-h, t]$.
-For simplicity, in the rest of this chapter we shall just write
-$t$ for the subscript ``t, t-h'', since the time interval $h$ is fixed
-for each time series and is a known constant.
-
-Andersen and Bollerslev et al proved the following in
-\cite{Andersen03} (theorem 2):
-\begin{equation}
-  \label{eq:normal_r}
-  r_t|\mathcal{F}_{t-h} \sim N(\int_{0}^h \mu_{t-h+s} ds, \int_{0}^h
-  \sigma_{t-h+s}^2 ds)
-\end{equation}
-In plain words, given all the information up to time $t-h$, the
-distribution of $r_t$ is Gaussian with integrated mean and variance.
-The variance of this conditional distribution, $\int_{0}^h
-\sigma_{t-h+s}^2 ds$, can be approximated by \cite{Protter05,
-  Andersen03}:
-\begin{equation}
-  \label{eq:rv_def}
-  \int_{0}^h \sigma_{t-h+s}^2 ds = \sum_{k=1}^n \left(
-    S_{t-h+kh/n} - S_{t-h+(k-1)h/n} \right)^2
-\end{equation}
-where $n$ is a chosen constant. The square root of the right hand side
-of equation \ref{eq:rv_def} is the realized volatility, call it
-$\hat{\sigma}_t$.
-
-With the availability of transaction data, estimating conditional
-volatility using returns sampled at a higher frequency (realized
-volatility) gives superior accuracy and reliability. However, at which
-frequency the time series should be sampled (the choice of $n$) in
-order to give an unbiased and consistent estimate of the volatility is
-not a trivial question. Naively one would believe that the often the
-series is sampled, the better the estimate, but in fact, due to noise
-introduced by market micro-structure, there is an optimal sampling
-frequency, depending on $h$. While the method to determine this
-optimal frequency is a subject of debate (see for example
-\cite{Sahalia05}), it is not difficult to find a fairly satisfactory
-frequency in practice:
-
-Keeping equation \ref{eq:normal_r} in mind, one can simply try a few
-frequencies and compare the distribution of $(r_t -
-\E(r_t))/\hat{\sigma}_t$ with the standard Gaussian. If the two match,
-$\hat{\sigma}_t$ is a good approximation of
-\[
-\sigma_t = \left(\int_{0}^h \sigma_{t-h+s}^2 ds \right)^{1/2}
-\]
-
-\end{enumerate}
-
-To compare GARCH and stochastic log-volatility models at the face of
-intraday returns, we study the {\it Nordea Bank} 15-minute returns
-during the period 2012/01/16 - 2012/04/20 in \S\ref{chp:nordea_15min},
-and {\it Volvo B} 30-minute returns during 2013/10/10 - 2014/04/04 in
-\S\ref{sec:volvo}. Another 5 intraday series are also studied and
-collected in appendix \ref{chp:appendix2}.
-
-\section{Case Study: Nordea 15-minute Returns}
+\chapter{Nordea in 1st quarter 2012}
 \label{chp:nordea_15min}
-In this section we investigate the Nordea 15-minute returns sampled
+In this chapter we investigate the Nordea 15-minute returns sampled
 during the period 2012/01/16 - 2012/04/20. In total, these amount to
@@
 
-\subsection{GARCH Model}\label{sec:nordea_15min_garch}
-First of all, to determine that the volatilities are auto-correlated
-and/or that the squared returns and the volatilities are correlated, and
-hence a GARCH(p, q) model is appropriate for the return series under
-investigation, we plot the auto-correlation function (ACF) of the
-squared returns. This is shown in figure
-\ref{fig:nordea_15min_acf}. If the aforementioned features are absent
-from the series, the auto-correlations are expected to be Gaussian
-distributed with mean 0 and variance 1/T and hence mostly reside
-within the confidence bounds set by $\pm 2/\sqrt{T}$
-\cite{Bollerslev86, Bollerslev87}.
-
-Clearly this is not the case in figure \ref{fig:nordea_15min_acf} ---
-the first 5 auto-correlations are rather significant. Moreover, figure
-\ref{fig:nordea_15min_vlt_acf} shows even more clearly that the
-conditional variances of the series are correlated. These observations
-suggest a GARCH(p, q) model can be appropriate.
+\section{GARCH Model}\label{sec:nordea_15min_garch}
+First of all, to check for GARCH effects, we plot the auto-correlation
+function (ACF) of the squared returns. This is shown in figure
+\ref{fig:nordea_15min_acf}. At the absence of GARCH effects, the ACF
+is expected to have an asymptotic Gaussian distribution with mean 0
+and variance 1/T \cite{Bollerslev86, Bollerslev87}, which is clearly
+not the case in figure \ref{fig:nordea_15min_acf}: The first 5
+autocorrelations have comparable sizes and do not fall off as in a
+Gaussian scheme. Moreover, figure \ref{fig:nordea_15min_vlt_acf} shows
+even more clearly that the conditional variances of the series are
+correlated. These observations suggest a GARCH(p, q) model can be
+appropriate.
 \begin{figure}[htb!]
@@
   \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=90 257 105
-    220]{../pics/nordea_15min_acf.pdf}
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_acf.pdf}
     \label{fig:nordea_15min_acf}
@@
   \subfigure[]{
-    \includegraphics[scale=0.4, clip=true, trim=90 259 104
-    218]{../pics/nordea_15min_vlt_acf.pdf}
+    \includegraphics[scale=0.4, clip=true, trim=95 236 118
+    200]{../pics/nordea_15min_vlt_acf.pdf}
     \label{fig:nordea_15min_vlt_acf}
   }
-  \caption{\small \it \ref{fig:nordea_15min_acf}: Auto-correlations
-    (ACF) of the squared returns; \ref{fig:nordea_15min_vlt_acf}:
-    Auto-correlations (ACF) of squared realized volatilities
-    ($\hat{\sigma}^2_t$). The blue lines are confident bounds set at
-    $\pm 2/\sqrt{T}$.}
+  \caption{\small \it Auto-correlations among Nordea Bank 15min
+    returns. \ref{fig:nordea_15min_acf}: Auto-correlations (ACF) of the
+    squared returns; \ref{fig:nordea_15min_vlt_acf}: Auto-correlations
+    (ACF) of squared realized volatilities ($\hat{\sigma}^2_t$).}
 \end{figure}
@@
 \begin{eqnarray*}
-  r_t &=& \mu + \epsilon_t \\
-  \epsilon_t &=& \sigma_t z_t \\
-  \sigma^2_t &=& \alpha_0 + \alpha_1 \epsilon^2_{t-1} + \alpha_s
-  \epsilon^2_{t-s} + \beta_1 \sigma^2_{t-1} + \beta_s \sigma^2_{t-s}
+  r_t &=& \mu + €_t \\
+  €_t &=& \sigma_t z_t \\
+  \sigma^2_t &=& \alpha_0 + \alpha_1 €^2_{t-1} + \alpha_s €^2_{t-s} +
+  \beta_1 \sigma^2_{t-1} + \beta_s \sigma^2_{t-s}
 \end{eqnarray*}
@@
 
-\subsection{Stochastic Volatility Model}\label{sec:nordea_15min_arima}
+\section{Stochastic Volatility Model}\label{sec:nordea_15min_arima}
 For the Nordea Bank 15-minute returns under consideration, it can be
@@
 \begin{eqnarray*}
-  {\varrho_{s+1}/\varrho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
+  {\rho_{s+1}/\rho_s} &=& {\gamma_{s+1}/\gamma_s} \;=\; -{\theta \over
     1 + \theta^2} \\
-  {\varrho_{s+1}/\varrho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
+  {\rho_{s+1}/\rho_1} &=& {\gamma_{s+1}/\gamma_1} \;=\; -{\Theta \over
     1 + \Theta^2} \\
@@
     \hline
-    $\varrho_1$ & $\varrho_{s-1}$ & $\varrho_s$ & $\varrho_{s+1}$ \\
+    $\rho_1$ & $\rho_{s-1}$ & $\rho_s$ & $\rho_{s+1}$ \\
     \hline
@@
 The first 4 moments of the Johnson Su distribution are expressible
-in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}.
-% \begin{eqnarray*}
-%   w &=& \exp{1 \over \delta^2} \\
-%   \Omega &=& {\gamma \over \delta} \\
-%   \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
-%   \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
-%     1)\right]^{1/2} \\
-%   \text{skewness}(y) &=& {
-%     \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
-%     \over
-%     (w\cosh 2\Omega + 1)^{3/2}} \\
-%   \text{kurtosis}(y) &=& {
-%     w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
-%     + 3(2w+1) \over
-%     2(w\cosh 2\Omega + 1)^2 }
-% \end{eqnarray*}
+in closed form in $\gamma, \delta, \lambda, \xi$ \cite{Shang2004}:
+\begin{eqnarray*}
+  w &=& \exp{1 \over \delta^2} \\
+  \Omega &=& {\gamma \over \delta} \\
+  \text{E}(y) &=& -w^{1/2} \lambda \sinh\Omega + \xi\\
+  \text{std}(y) &=& \lambda \left[{1 \over 2}(w-1)(w\cosh 2\Omega +
+    1)\right]^{1/2} \\
+  \text{skewness}(y) &=& {
+    \sqrt{(1/2)w(w-1)} [w(w+2)\sinh 3\Omega + 3\sinh\Omega]
+    \over
+    (w\cosh 2\Omega + 1)^{3/2}} \\
+  \text{kurtosis}(y) &=& {
+    w^2(w^4 + 2w^3 + 3w^2 - 3)\cosh 4\Omega + 4w^2 (w+2) \cosh 2\Omega
+    + 3(2w+1) \over
+    2(w\cosh 2\Omega + 1)^2 }
+\end{eqnarray*}
 By matching the theoretical expressions of the moments with their
@@
 
-% The eventual MLE is done in Matlab with the ``active set''
-% algorithm. The initial as well as the final estimation results are
-% listed in table \ref{tab:nordea_15min_js_param}:
-% \begin{table}[htb!]
-%   \centering
-%   \begin{tabular}{|c|c|c|c|c|c|c|}
-%     \hline
-%     & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
-%     \hline
-%     initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
-%     0.6378 \\
-%     \hline
-%     MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
-%     \\
-%     \hline
-%   \end{tabular}
-%   \caption{\small \it Nordea 15min estimation results}
-%   \label{tab:nordea_15min_js_param}
-% \end{table}
+The eventual MLE is done in Matlab with the ``active set''
+algorithm. The initial as well as the final estimation results are
+listed in table \ref{tab:nordea_15min_js_param}:
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|c|c|}
+    \hline
+    & $\gamma$ & $\delta$ & $\lambda$ & $\xi$ & $\theta$ & $\Theta$ \\
+    \hline
+    initial estimate & 0.1476 & 1.5121 & 0.3633 & 0.0454 & 0.6890 &
+    0.6378 \\
+    \hline
+    MLE estimate & 0.1319 & 1.5266 & 0.3735 & 0.0410 & 0.6639 & 0.6025
+    \\
+    \hline
+  \end{tabular}
+  \caption{\small \it Nordea 15min estimation results}
+  \label{tab:nordea_15min_js_param}
+\end{table}
 
-\subsection{Comparison of the Forecasts}
+\section{Comparison of the Forecasts}
 In this section we compare the one-step-ahead forecasts from the GARCH
@@
   \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
-    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t < x)$;
-    Right: $P(\ln \sigma^F_t - \ln \hat{\sigma}_t > x)$. Horizontal: $x$.}
+    sample mean forecasts. Left: cumulative distribution function;
+    Right: complementary distribution function.}
   \label{fig:nordea_2012}
@@
     \hline
-    1\% & 21.73\% & 13.83\% & 11.36\% \\
+    1\% & 0.2173 & 0.1383 & 0.1136 \\
     \hline
-    5\% & 74.57\% & 70.62\% & 53.33\% \\
+    5\% & 0.7457 & 0.7062 & 0.5333 \\
     \hline
-    10\% & 97.28\% & 93.83\% & 89.14\% \\
+    10\% & 0.9728 & 0.9383 & 0.8914 \\
     \hline
@@
 \end{table}
-
-\section{Case study: Volvo 30-minute Returns}
-\label{sec:volvo_30}
-In this section we model the log-volatility of {\it Volvo B} 30-minute
-returns during the period 2013/10/10 - 2014/04/04. This series
-contains 1884 log-volatilities computed using 2-minute returns in each
-30-minute interval. Among them we use the first 1507 for model
-estimation and the last 377 for forecast and model verification.
-
-The left plot of figure \ref{fig:volvo_inno_and_lv_acf} shows the
-distribution of $(r_t - \E(r_t))/\sigma_t$. We observe in the figure a
-nice Gaussian variate, so we can be sure that the sum of squared
-2-minute returns makes a good approximation to the variance of
-30-minute returns in this particular case.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.4, clip=true, trim=0 256 0
-    151]{../pics/volvo_inno_and_lv_acf.pdf}
-    \label{fig:volvo_inno_and_lv_acf}
-  \caption{\small \it Left: probability plot of $(r_t -
-    \E(r_t))/\sigma_t$; Right: auto-correlations of $\ln \sigma_t$}
-\end{figure}
-
-Guided by the auto-correlations of $\ln\sigma_t$ shown in the right
-plot of figure \ref{fig:volvo_inno_and_lv_acf} we find the following
-model:
-\begin{eqnarray}
-  && (1 - B)(1 - B^s) \ln \sigma_t \nonumber \\
-  &=& (1 - \theta_1 B - \theta_2B^2 -
-  \theta_3B^3 - \theta_4B^4) (1 - \Theta B^s)
-  y_t \label{eq:volvo_lv_model}
-\end{eqnarray}
-where $s = 16$ is the seasonality and is apparent from the
-auto-correlations of $\ln\sigma_t$. Fitting this model to the
-measured realized volatilities yields the parameter values listed in
-table
-\ref{tab:volvo_params}.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|c|c|}
-    \hline
-    Parameter & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ &
-    $\Theta$ & residual variance \\
-    \hline
-    Value & 0.7305 & 0.0575 & 0.0574 & 0.0346 & 0.8324 & 0.1340\\
-    \hline
-  \end{tabular}
-  \caption{\small \it Volvo B log-volatility parameters}
-  \label{tab:volvo_params}
-\end{table}
-Forecasting using the estimated model parameters gives the forecast
-series $\ln \sigma^{\text{SV}}_t$. To access the quality of the
-forecast, we also estimate a GARCH model using the returns. The result
-is a GARCH(1, 1) model, whose parameter values are listed in table
-\ref{tab:volvo_garch}.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    Parameter & $\alpha_0$ & $\alpha_1$  & $\beta_1$ \\
-    \hline
-    Value & $3.125 \times 10^{-7}$ & 0.05 & 0.90 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it GARCH(1, 1) model of Volvo B 30-minute returns}
-  \label{tab:volvo_garch}
-\end{table}
-
-The forecasts from SV, GARCH, and the sample mean are compared
-using the difference $\ln \sigma^F_t - \ln\hat{\sigma}_t$, where $\ln
-\sigma^F_t$ stands for the forecast. The distributions of this
-difference is plotted in figure \ref{fig:volvo_slv_garch_cmp}; the
-mean and the standard deviation of the distributions are listed in
-table \ref{tab:volvo_slv_garch_cmp}.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    & SV & GARCH & Sample mean \\
-    \hline
-    $\E(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & -0.0123 &
-    0.0242 & -0.1055 \\
-    \hline
-    $\text{std}(\ln \sigma^F_t - \ln \hat{\sigma}_t)$ & 0.3261 &
-    0.4250 & 0.3708 \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Standard deviation of $\ln\sigma^F_t -
-    \ln\hat{\sigma}_t$}
-  \label{tab:volvo_slv_garch_cmp}
-\end{table}
-
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.5, clip=true, trim=10 282 0
-    243]{../pics/volvo_slv_garch_cmp.pdf}
-    \label{fig:volvo_slv_garch_cmp}
-  \caption{\small \it Blue: SV forecasts; Green: GARCH forecasts; Red:
-    sample mean forecasts. Left: $P(\ln \sigma^F_t - \ln
-    \hat{\sigma}_t < x)$; Right: $P(\ln \sigma^F_t - \ln
-    \hat{\sigma}_t > x)$. Horizontal: $x$.}
-\end{figure}
-Figure \ref{fig:volvo_slv_garch_cmp} shows, as in the previous case of
-Nordea Bank 15-minute returns, the SV model performs the best, GARCH
-the second, and the sample mean the worst. However, when it comes to
-over-estimates, the sample mean appears to be the best estimator,
-while SV excels over GARCH. But a check of $\E(\ln\sigma_t)$ over
-the sample for estimation (1507 data points) and over the sample for
-comparison (377 data points) reveals that the first sample has mean
--6.3127 while the second has -6.2072. This increment explains the low
-probability of over-estimation when using the first sample mean as
-forecast.
-
-Table \ref{tab:volvo_good_percentage} compares the fraction of
-``good'' estimates as measured by ${|\ln\sigma^F_t -
-  \ln\hat{\sigma}_t| \over |\ln \hat{\sigma}_t|}$ being less than 1\%,
-5\% and 10\%.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ &
-    SV & GARCH & Sample Mean \\
-    \hline
-    1\% & 21.75\% & 12.20\% & 14.06\% \\
-    \hline
-    5\% & 71.62\% & 61.54\% & 66.31\% \\
-    \hline
-    10\% & 92.31\% & 88.06\% & 89.92\% \\
-    \hline
-  \end{tabular}
-  \caption{\small \it Fraction of ``good'' forecasts as defined by
-    ${|\ln \sigma^F_t - \ln \hat{\sigma}_t| \over |\ln
-      \hat{\sigma}_t|}$ being less than 1\%, 5\% and 10\%.}
-  \label{tab:volvo_good_percentage}
-\end{table}
-We see from the table that the SV model consistently excels over
-the other two alternatives. In addition, it is also noted that the
-GARCH forecast is even worse than the sample mean. This is surprising
-but not totally unexpected --- with only 3 parameters, the GARCH(1,1)
-model can only describe the most prominent auto-correlations in the
-volatility. When the volatility is influenced by relatively weak
-auto-correlations at several different time lags, the GARCH forecast
-cannot be expected to have good accuracy.
-
-In this particular case, we see that the ARIMA model has 3 relatively
-small moving average coefficients, located at lags 2, 3, and 4 and
-evaluated to 0.06, 0.06, 0.03, suggesting a scattered auto-correlation
-structure, so the GARCH model cannot be expected to perform very
-well. In contrast, the Volvo 15-minute returns studied in
-\S\ref{sec:volvo} has more concentrated auto-correlations --- 0.12 and
-0.06 at lags 2 and 3 --- thus the GARCH(1,1) model is also found to
-perform better and even marginally better than the SV model for the
-same series.
-
-\section{Unconditional Distribution Functions of SV Models}
-\label{sec:XieCalc}
-In this section we study the unconditional distribution function of
-the Stochastic Log-Volatility model specified as equation
-\ref{eq:SLV_spec}. As is discussed in \S\ref{sec:SLV_model},
-$\ln\sigma_t$ can be described by an ARMA or ARIMA \footnote{ARIMA ---
-  Integrated Autoregressive Moving Average} model, possibly with
-seasonal components. Here we note that all these models can be
-re-written as a moving average model, which is infinite in extent if
-autoregressive components are present:
-\begin{eqnarray*}
-  \ln \sigma_t &=& y_t + \sum_{n=1}^\infty c_n y_{t-n} + \text{Const.}
-\end{eqnarray*}
-Since the $y_t$ are independent and identically distributed,
-\begin{eqnarray*}
-  y_t + \sum_{n=1}^\infty c_n y_{t-n}  
-\end{eqnarray*}
-has Gaussian distribution by the central limit theorem, on condition
-that $y_t$ for all $t$ have finite second moment --- this is what we
-assume in the rest of this section. It follows from the above equation
-that the unconditional distribution of $\ln \sigma_t$ is the same as the
-distribution of $\bar{v} + v$ where $v \sim N(0, \sigma)$ and
-$\bar{v}$, $\sigma$ are constants. Now we can state that the
-unconditional distribution of the returns $r_t$
-\begin{eqnarray*}
-  r_t &=& \mu + \sigma_t b_t\\
-  &=& \mu + \exp\left(
-    y_t + \sum_{n=1}^\infty c_n y_{t-n} + \text{Const.}
-  \right)
-\end{eqnarray*}
- is the same as
-\begin{equation}  \label{eq:UnconditionalPdf}
-  \begin{aligned}
-    r &= \mu + e^{\bar{v} + v} b \\
-  \end{aligned}
-\end{equation}
-where $b \sim N(0, 1)$. For convenience, let $r' = e^v b$
-
-In \S\ref{sec:SLV_Symmetric} we first study the model in the
-relatively simple case when $v$ and $b$ are uncorrelated and $\mu
-= 0$. If this simplified version proves inadequate, one may resort to
-the general model studied in \S\ref{sec:SLV_Asymmetric}.
-
-\subsection{The Simplified model}\label{sec:SLV_Symmetric}
-In the following we derive the unconditional PDF of $r'$, denoted
-$f_{r'}(x)$. Then the PDF of $r$ is $e^{-\bar{v}}f_{r'}(e^{-\bar{v}}x)$.
-\begin{eqnarray*}
-  P(r' < x) &=& P(b < xe^{-v}) \\
-  f_{r'}(x) &=& f_b(xe^{-v}) e^{-v}
-\end{eqnarray*}
-Averaging over all $v$, we get
-\begin{equation}\label{eq:UncondPDFSymmetric}
-  \begin{aligned}
-    f_{r'}(x) =& \int_{-\infty}^{\infty} dv (2\pi\sigma^2)^{-1/2}
-    e^{-v^2/2\sigma^2}(2\pi)^{-1/2} \exp(-x^2e^{-2v}/2) e^{-v} \\
-    =& {1 \over 2\pi\sigma} \int_{-\infty}^{\infty} dv
-    \exp\left(
-      -{1 \over 2\sigma^2} v^2 - v -{1 \over 2} x^2 e^{-2v}
-    \right)
-  \end{aligned}
-  \end{equation}
-The last part of the integrand, $e^{-x^2 e^{-2v} / 2}$, is plotted in
-figure \ref{fig:DoubleExp}.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5, clip=true, trim=85 252 100
-  231]{../pics/DoubleExp.pdf}
-  \caption{\small \it Plot of $\exp(-{1 \over 2} x^2 e^{-2v})$}
-  \label{fig:DoubleExp}
-\end{figure}
-Therefore we make the following approximation:
-\[
-\exp\left(-{1 \over 2} x^2 e^{-2v}\right) \approx \left\{
-  \begin{array}{lr}
-    0 & \text{if } v < \ln|x| -{1 \over 2} \ln(2\ln 2) \\
-    1 & \text{otherwise}
-  \end{array}
-\right.
-\]
-Here we note that $\exp\left(-{1 \over 2} x^2 e^{-2v}\right) = 1/2$ at
-$v = \ln|x| -{1 \over 2} \ln(2\ln 2)$.
-
-With this approximation we have
-\begin{eqnarray*}
-  f_{r'}(x) &=& {1\over C}{1 \over 2\pi\sigma} \int_{\ln(|x|/\sqrt{\ln
-      4})}^{\infty} dv
-  \exp\left(-{1 \over 2\sigma^2} v^2 - v\right) \\
-  &=& {1\over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}} \text{erfc} \left(
-    {1 \over \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} + {\sigma
-      \over \sqrt{2}}
-  \right)
-\end{eqnarray*}
-where $1/C$ has been added for the purpose of normalization.
-
-At large $x$, we may use the asymptotic expansion of $\mathrm{erfc}$
-to write
-\begin{eqnarray*}
-  Cf_{r'}(x) &=& {e^{\sigma^2 / 2} \over \sqrt{8\pi}} \mathrm{erfc}(\xi) \\
-  &=& {e^{\sigma^2 / 2} \over \sqrt{8 \pi}}
-  \frac{e^{-\xi^2}}{\xi\sqrt{\pi}}\left[
-    1 +
-    \sum_{n=1}^N (-1)^n \frac{(2n-1)!!}{(2\xi^2)^n} \right] +
-  O(\xi^{-2N-1} e^{-\xi^2})
-\end{eqnarray*}
-where $\xi = {1 \over \sqrt{2}\sigma} \ln{|x| \over \sqrt{\ln 4}} +
-{\sigma \over \sqrt{2}}$. The slowest-decaying term is
-\[
-f_{r,0}(x) = {e^{\sigma^2 / 2} \over \sqrt{8\pi}}
-\frac{e^{-\xi^2}}{\xi\sqrt{\pi}}
-\]
-Let $\zeta = {|x| \over \sqrt{\ln 4}}$. With a bit manipulation one
-obtains
-\begin{equation*}
-  f_{r,0}(x) = {1 \over \pi \sqrt{8}}{1 \over
-    \left(\ln\zeta/\sigma\sqrt{2} +
-      \sigma/\sqrt{2}\right)\zeta\zeta^{\ln \zeta / 2\sigma^2}
-  }
-\end{equation*}
-From the last equation one can see that, at any neighborhood of large
-$x$, $f_{r'}(x)$ may be approximated by $C/|x|^\alpha$, i.e. a power law.
-The normalization constant $C$ is calculated in appendix
-\ref{chp:symmetric_SV_norm_const} to be $C = \sqrt{2 \ln 4\over
-  \pi}$. So, in summary, we can write:
-\begin{eqnarray*}
-  f_r(x) &=& e^{-\bar{v}} f_{r'}(e^{-\bar{v}} x) \\
-    &=& {e^{-\bar{v}} \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}
-    \text{erfc} \left({1 \over \sqrt{2}\sigma} \ln{|e^{-\bar{v}}x| \over \sqrt{\ln
-          4}} + {\sigma \over \sqrt{2}}
-    \right) \\
-\end{eqnarray*}
-Using the same technique for integration as for normalization, the
-cummulative distribution function of $r$ is found to be $F(x)$,
-which is the following:
-\begin{enumerate}
-\item if $x < 0$
-  \begin{eqnarray*}
-    F(x) &=& {\sqrt{\ln 4} \over C}{e^{\sigma^2 / 2} \over \sqrt{8\pi}}
-    \left[
-      {e^{-\bar{v}}x \over \sqrt{\ln 4}}\text{erfc}\left(
-        {1 \over \sigma \sqrt 2} \ln{-e^{-\bar{v}}x \over \sqrt{\ln
-            4}} + {\sigma \over \sqrt 2}
-      \right) \right.\\
-      && \left. + e^{-\sigma^2 / 2} \text{erfc}\left(
-        {1 \over \sigma \sqrt 2} \ln{-e^{-\bar{v}}x \over \sqrt{\ln 4}}
-      \right)
-    \right]
-  \end{eqnarray*}
-\item if $x \geq 0$
-  \begin{eqnarray*}
-    F(x) &=& \frac{1}{2} + {\sqrt{\ln 4} \over C}{e^{\sigma^2 / 2} \over
-      \sqrt{8\pi}} \left[
-      {e^{-\bar{v}}x \over \sqrt{\ln 4}} \text{erfc}\left(
-        {1 \over \sigma \sqrt 2} \ln{e^{-\bar{v}}x \over \sqrt{\ln
-            4}} + {\sigma \over \sqrt 2}
-      \right) \right. \\
-      && \left. + e^{-\sigma^2/2} \text{erfc}\left(
-        -{1 \over \sigma \sqrt 2} \ln{e^{-\bar{v}}x \over \sqrt{\ln 4}}
-      \right)
-    \right]
-  \end{eqnarray*}
-\end{enumerate}
-
-To verify the validity of the model, we fit the above probability
-density function to the de-meaned 30min returns of Volvo B
-\footnote{By ``de-meaned returns'' we mean the quantity $r_t - 
-  \mean{r_t}$, where $r_t$ are the measured returns and $\mean{r_t}$
-  is the sample mean. The data set covers the transaction records of
-  Volvo B on the OMX market (Stockholm) between 2013-10-10 and
-  2014-03-12. The returns are computed using 1-minute mean prices.}
-by means of maximum likelihood estimate using the MATLAB function
-``mle''. Then for the parameters $\sigma$ and $\bar{v}$ we get
-\begin{eqnarray*}
-  \sigma &=& 0.6355 \\
-  \bar{v} &=& -6.0308
-\end{eqnarray*}
-Then we plot $P(r' > x)$ of the model against its empirical
-counterpart on a log-log scale, as shown in figure
-\ref{fig:volvo_30min_ret}.
-\begin{figure}[htb!]
-  % \vspace{-15mm}
-  \begin{center}
-    \includegraphics[scale=0.4, clip=true, trim=98 231 116
-    126]{../pics/volvo_30min_ret.pdf}
-  \end{center}
-  %\vspace{-5mm}
-  \caption{\small \it{$P(r' > x)$ for $x > 0$. Blue: empirical
-      probabilities. Red: probabilities predicted by the model.}}
-  \label{fig:volvo_30min_ret}
-\end{figure}
-
-In general figure \ref{fig:volvo_30min_ret} shows a good fit, but a
-closer look reveals that deviations are significant in the regions $r
-\in (0, \sigma_r)$ and $r \in (2\sigma_r, 3\sigma_r)$, where
-$\sigma_r$ stands for the empirical standard deviation of the
-de-meaned returns. These observations are
-shown in greater details in figure \ref{fig:volvo_30min_ret2}.
-\begin{figure}[htb!]
-  \centering
-  \subfigure[$\ln(P(r > x))$ with $x \in (0, \sigma_r)$]{
-    \includegraphics[scale=0.4, clip=true, trim=93 224 115
-    125]{../pics/volvo_30min_ret_0-sigma.pdf}
-  }
-  \subfigure[$\ln(P(r > x))$ with $x \in (2\sigma_r, 3\sigma_r)$]{
-    \includegraphics[scale=0.4, clip=true, trim=93 224 115
-    125]{../pics/volvo_30min_ret_2sigma-3sigma.pdf}
-  }
-  \caption{\small \it Surviving probabilities. Blue: empirical values of $\ln(P(r
-    > x))$. Red: Predicted values of $\ln(P(r > x))$.}
-  \label{fig:volvo_30min_ret2}
-\end{figure}
-
-Moreover, the inefficiency of the model is also manifest in the
-skewness of the data. For the Volvo 30min returns, the data has a
-skewness of 0.2419, but our model is strictly symmetric, since
-$x$ appears in equation \ref{eq:UncondPDFSymmetric} only as $x^2$.
-Hence the above model needs to be improved to accommodate the non-zero
-skewness as well as to account for the discrepancies shown in figure
-\ref{fig:volvo_30min_ret2}. This is the subject of the next section.
-
-\subsection{The General model}\label{sec:SLV_Asymmetric}
-It has long been hypothesized in the liturature that skewness is the
-result of price-volatility correlation (for example
-\cite{Potters2003}). Therefore, the most apparent modification is to
-allow $v$ and $b$ to be correlated. For convenience we decompose
-$v \sim N(0, \sigma)$ as $v = \sigma a$ and assume
-\begin{eqnarray*}
-  \begin{pmatrix}
-    a \\
-    b
-  \end{pmatrix} \sim N(0, \Sigma)
-\end{eqnarray*}
-with a covariance matrix
-\begin{eqnarray*}
-    \Sigma &=&
-  \begin{pmatrix}
-    1 & \psi \\
-    \psi & 1
-  \end{pmatrix}
-\end{eqnarray*}
-where $|\psi| < 1$. Then we rewrite equation \ref{eq:UnconditionalPdf}
-as
-\begin{equation}
-  \label{eq:r_t}
-  \begin{aligned}
-    r &= \mu + e^{\bar{v}} r' \\
-    r' &= e^{\sigma a} b \\
-  \end{aligned}
-\end{equation}
-Then
-\begin{eqnarray*}
-  && P(r' < x)\\
-  &=& P(b < e^{-a\sigma}x) \\
-  &=& {1 \over 2\pi |\det(\Sigma)|^{1/2}}
-  \int_{-\infty}^{\infty} da \int_{-\infty}^{e^{-a\sigma}x} db
-  \exp\left[
-    -{1 \over 2} (a, b) \Sigma^{-1}
-    \begin{pmatrix}
-      a \\
-      b
-    \end{pmatrix}
-  \right] \\
-\end{eqnarray*}
-After some manipulations we get
-\begin{equation}\label{eq:UncondCDFAsymmetric}
-  \begin{aligned}
-    & P(r' < x) \\  
-    &= {1 \over 2\pi \sqrt{1 - \psi^2}} \int_{-\infty}^{\infty} da
-    e^{-a^2/2} \int_{-\infty}^{e^{-a\sigma}x} db
-    \exp\left[
-      - {(b - a\psi)^2 \over 2(1 - \psi^2)}
-    \right] \\
-    &= {1 \over 2\sqrt{2\pi}} \int_{-\infty}^{\infty} e^{-a^2/2}
-    \text{erfc}{a\psi - e^{-a\sigma} x \over \sqrt{2(1-\psi^2)}} da
-  \end{aligned}
-\end{equation}
-Differentiating with respect to $x$ yields
-\begin{eqnarray}
-  f_{r'}(x) &=& {1 \over 2\pi \sqrt{1 - \psi^2}} \times \nonumber \\
-  && \int_{-\infty}^\infty da \exp
-  \left[- {
-      a^2 + 2\sigma(1 - \psi^2)a - 2a\psi e^{-a\sigma} x + e^{-2a\sigma} x^2
-      \over
-      2(1 - \psi^2)
-  }
-  \right] \label{eq:UncondPDFAsymmetric} \\
-  f_r(x) &=&  \td{r'}{r} f_{r'}[e^{-\bar{v}} (x - \mu)] \nonumber \\
-  &=& e^{-\bar{v}} f_{r'}[e^{-\bar{v}}
-  (x-\mu)] \label{eq:UncondPDFAsymmetric1}
-\end{eqnarray}
-Unlike \ref{eq:UncondPDFSymmetric} where a relatively simple
-approximation can be found and thus leads to an analytic result of the
-integral, no such approximation has been found by the
-author for the integral \ref{eq:UncondPDFAsymmetric}. However, the
-moment generating functions (MGF) of PDF \ref{eq:UncondPDFAsymmetric}
-and \ref{eq:UncondPDFAsymmetric1} are easy enough to find and give the
-moments about the origin in closed form. Then by matching the analytic
-expressions of the moments with their statistical values from the
-sample, the parameters $\sigma$, $\psi$ and $\bar{v}$ corresponding to the
-sample can be obtained.
-
-The MGF of \ref{eq:UncondPDFAsymmetric1} can be found as follows:
-\begin{eqnarray*}
-  && M_{r}(t) \\
-  &=& \E(e^{tr}) \\
-  &=& {1 \over 2\pi \sqrt{1 - \psi^2}} \int_{-\infty}^{\infty} da \exp \left(
-    -{a^2 \over 2} - \sigma a
-  \right) \int_{-\infty}^{\infty} dx \exp\left[
-    t(e^{\bar{v}}x + \mu) - {
-      (a\psi - e^{-a \sigma} x)^2 \over 2 (1 - \psi^2)
-    }
-  \right] \\
-  &=& {1 \over \sqrt{2 \pi}} \int_{-\infty}^{\infty} da \exp\left[
-    -{a^2 \over 2} + {1 \over 2} e^{2 a \sigma + 2\bar{v}} (1 - \psi^2) t^2 +
-    a \psi e^{a \sigma + \bar{v}} t + \mu t
-  \right]
-\end{eqnarray*}
-With the MGF, the first 4 moments of $r$ can be computed:
-\begin{equation}
-  \label{eq:AsymmetricMoments}
-  \begin{aligned}
-    \E(r) &= \mu +e^{\bar{v}+\frac{\sigma ^2}{2}} \sigma  \psi\\
-    \E(r^2) &= \mu ^2+2 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu  \sigma  \psi
-    +e^{2 \left(\bar{v}+\sigma ^2\right)} \left(1+4 \sigma ^2 \psi^2\right)\\
-    \E(r^3) &= \mu ^3+3 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu ^2 \sigma \psi
-    +9 e^{3 \bar{v}+\frac{9 \sigma ^2}{2}} \sigma \psi \left(1+3 \sigma ^2
-      \psi ^2\right)
-    +3e^{2 \left(\bar{v}+\sigma ^2\right)} \mu  \left(1+4 \sigma ^2 \psi
-      ^2\right) \\
-    \E(r^4) &= \mu ^4+4 e^{\bar{v}+\frac{\sigma ^2}{2}} \mu ^3 \sigma  \psi
-    +36 e^{3 \bar{v}+\frac{9 \sigma ^2}{2}} \mu  \sigma  \psi  \left(1+3
-      \sigma ^2 \psi ^2\right)\\ 
-    & +6 e^{2 \left(\bar{v}+\sigma ^2\right)} \mu ^2 \left(1+4 \sigma ^2 \psi
-      ^2\right)
-    +e^{4 \bar{v}+8 \sigma ^2} \left(3+96 \sigma ^2 \psi ^2+256 \sigma ^4
-      \psi ^4\right)
-  \end{aligned}
-\end{equation}
-These equations are rather complicated and directly solving them to
-obtain the parameters $\mu$, $\sigma$, $\psi$ and $\bar{v}$ is
-infeasible. However, in practice, $\E(r)$ is often very small --- so, to
-get a rough estimate of the parameters, we may set $\mu = 0$ in the
-above equations and, with a bit of manipulation, find the following
-equation for $\psi\sigma$:
-\begin{equation}
-  \label{eq:moment1}
-  {\E(r^3) \E(r^4)^{-3/8} \over \E(r^2)^{3/4}} =
-  {9\sigma \psi (1 + 3\sigma^2\psi^2) \over (1 +
-    4\sigma^2\psi^2)^{3/4}}
-  {1
-    \over
-    (3 + 96\sigma^2\psi^2 + 256\sigma^4 \psi^4)^{3/8}
-  }
-\end{equation}
-This equation can be solved numerically for a given sample to yield an
-estimate for $\sigma\psi$, which in turn can be substituted in
-equations \ref{eq:AsymmetricMoments}, where $\mu$ has been set to 0,
-to give estimates for all the 4 parameters. These estimates then serve
-as initial values in a numerical solution to
-\ref{eq:AsymmetricMoments} where $\mu$ is kept as a free
-variable. This full solution can now be used as the initial estimate
-in an MLE procedure.
-
-It is particularly important in our situation to have a good initial
-estimate, because, as has been shown earlier, the PDF and CDF cannot
-be obtained in closed forms and consequently have to be evaluated by
-numerical integration. This is a rather costly procedure especially in
-the context of MLE. So the computation of the moments under the
-assumption of $\mu = 0$ is worthwhile.
-
-Following the aforementioned procedure, i.e. computing the initial
-estimate by matching the moments and then refining the estimate by
-MLE, we obtain the parameter values for a number of return series,
-including the Volvo B 30-minute returns described in
-\S\ref{sec:SLV_Symmetric}. Table \ref{tab:assets_params} and
-\ref{tab:assets_moments} present the obtained parameter values and the
-resulting moments, respectively. Figure \ref{fig:Volvo_B_30m_returns},
-\ref{fig:Nordea_30m_returns}, and \ref{fig:Ericsson_30m_returns}
-compare the empirical distribution functions with the analytic
-distribution functions evaluated at these parameters.
-
-Clearly these figures show a fairly good match of the empirical and
-the analytic distribution functions. However, we also see that the
-skewness of the returns of Volvo B and Ericsson B have rather
-different values when computed from the sample and from the
-model. This could be the consequence of the limited sample size or
-deficiencies in the estimation procedure described above. We leave
-these issues to future studies.
-
-%  This indicates the parameter $\psi$, which is the
-% correlation between the log-volatility and the return, may be more
-% complicated than the simple constant that has been assumed so far. We
-% leave this to later studies.
-
-\subsection{Relation to Conditional Distribution Functions}
-In the last section we have shown that the unconditional distribution of
-the returns are skewed, which implies the variate in the
-log-volatility $a$ is correlated to the variate in the return
-$b$.
-
-In the context of conditional distributions and forecast, this
-correlation translates to the correlation between the residual of the
-log-volatility, denoted $y_t$ in \S\ref{chp:nordea_15min} and
-\S\ref{sec:volvo}, and $b_t$. Now let us consider the forecast
-function of an ARIMA model:
-\begin{eqnarray*}
-  \ln \sigma_t &=& y_t + \sum_{i=1}^P \phi_i \ln \sigma_{t-i} -
-  \sum_{i=1}^Q \theta_i y_{t-i}
-\end{eqnarray*}
-where $\phi_i \leq 1$ if the model involves integration, and $\phi_i <
-1$ otherwise. Comparing this equation with equation \ref{eq:r_t}, one
-immediately realizes the PDF of $r_t$
-\begin{eqnarray*}
-  r_t &=& \mu + \sigma_t b_t
-\end{eqnarray*}
-is given by equation \ref{eq:UncondPDFAsymmetric1} and its moments
-given by the equations \ref{eq:AsymmetricMoments} if one makes the
-following substitutions:
-\begin{eqnarray*}
-  \sigma &\to& \text{std}(y_t) \\
-  \bar{v} &\to& \sum_{i=1}^P \phi_i \ln \sigma_{t-i} - \sum_{i=1}^Q
-  \theta_i y_{t-i} \\
-  \psi &\to& \text{corr}(y_t, b_t)
-\end{eqnarray*}
-
-For the Nordea series considered in \S\ref{chp:nordea_15min},
-$\text{corr}(y_t, b_t)$ is found to be $3.56 \times 10^{-2}$; while
-for the Volvo series considered in \S\ref{sec:volvo},
-$\text{corr}(y_t, b_t)$ is found to be $-7.3 \times 10^{-3}$. These
-values are comparable to those in table \ref{tab:assets_params}, where
-$\psi$ for the Nordea series is $1.7234 \times 10^{-2}$ and $-1.5747
-\times 10^{-2}$ for the Volvo series. The similarity in these values
-provides some evidence about the validity of the model.
-
-It is certainly impossible to compare a predicted conditional
-distribution with observations, but conditional distributions are of
-great interest in the context of risk management and derivative
-pricing, so we point out how they may be calculated for ARIMA
-log-volatility models.

### 2014-12-01 thesis/_region_.tex
1st version that has been handed to Sven. Gaussian Wishart Matrices are normized by 1/T.
This is the version to be used for job searching.

@@
-\message{ !name(thesis.tex)}\documentclass{book}
+\message{ !name(draft4.tex)}\documentclass{article}
+\usepackage{amssymb,amsmath,amsthm,bm}
 \usepackage{graphicx}
 \usepackage{subfigure}
+\usepackage{multirow}
 \usepackage{wrapfig}
-\usepackage{amssymb}
-\usepackage{amsfonts, amsmath}
+\usepackage[usenames,dvipsnames]{color}
 \usepackage{mathrsfs}
@@
 \input{../physics_common.tex}
-\DeclareGraphicsExtensions{.pdf,.png,.jpg}
+\DeclareGraphicsExtensions{.eps, .pdf}
 
-\title{Mathematical Aspects of the Capital Market}
-\author{Xie Xiaolei}
-\date{\today}
 \begin{document}
 
-\message{ !name(thesis.tex) !offset(1382) }
-\cite{NIST:DLMF}:
+\message{ !name(draft4.tex) !offset(-3) }
 
-\message{ !name(thesis.tex) !offset(1837) }
+\section{effect of heteroscedasticity in covariance matrix}
+Consider a covariance matrix constructed as
+$$
+C_{ij} = {1 \over T}\sum_{t=1}^T r_{it} r_{jt}
+$$
+where $1 \leq i \leq N$. Adopt a stochastic volatility model for
+$r_{it}$, i.e.
+$$
+r_{it} = \sigma_{it} \eta_{it}
+$$
+We can write
+\begin{eqnarray*}
+C_{ij} &=& {1 \over T}\sum_{t=1}^T \sigma_{it} \eta_{it} \sigma_{jt}
+\eta_{jt} \\
+\bm{C} &=& {1 \over T} (\bm{\sigma * \eta}) (\bm{\sigma * \eta})'
+\end{eqnarray*}
+where the matrices $\bm{\sigma}$ and $\bm{\eta}$ have elements
+$\sigma_{it}$ and $\eta_{it}$ respectively, and * denotes element-wise
+multiplication. Due to the cyclic property of matrix trace, the
+moment generating function of $\bm{C}$, $M_C(z)$, relates to $M_D(z)$,
+the moment generating function of $\bm D$
+\begin{eqnarray*}
+  \bm{D} &=& {1 \over T} (\bm{\sigma * \eta})' (\bm{\sigma * \eta})
+\end{eqnarray*}
+through the equation
+\begin{eqnarray*}
+  M_C(z) &=& {1 \over q} M_D(z)
+\end{eqnarray*}
+where $q = N/T$. For later convenience, we make use of $T \times T$
+matrices $\bm{\tilde{\sigma}}$ and $\bm{\tilde{\eta}}$, as well as a
+projector
+$$
+\bm{P} = \text{diag}(\underbrace{1, \cdots, 1}_{\text{N 1's}}, 
+\underbrace{0, \cdots, 0}_{\text{T-N 0's}})
+$$
+The first N rows of $\bm{\tilde{\sigma}}$ and $\bm{\tilde{\eta}}$ are
+precisely those of $\bm{\sigma}$ and $\bm{\eta}$. This way, we have
+\begin{eqnarray*}
+\bm D &=& {1 \over T} (\bm{\sigma * \eta})' (\bm{\sigma * \eta}) \\
+&=& {1 \over T} (\bm{\tilde{\sigma} * \tilde{\eta}})' \bm P'
+\bm P (\bm{\tilde{\sigma} * \tilde{\eta}}) \\
+&=& {1 \over T} (\bm{\tilde{\sigma} * \tilde{\eta}})'
+\bm P (\bm{\tilde{\sigma} * \tilde{\eta}}) \\
+\end{eqnarray*}
+Again, by the cyclic property of matrix trace, the spectral properties
+of the RHS of the last equation is equivalent to those of
+\begin{eqnarray*}
+  \bm E &=& {1 \over T} (\bm{\tilde{\sigma} * \tilde{\eta}}) (\bm{\tilde{\sigma}
+    * \tilde{\eta}})' \bm P \\
+\end{eqnarray*}
+In the simple situation where $\sigma_{it} =
+\sigma_{jt} = \sigma_t$ for some $\sigma_t$ and for all $i, j, t$,
+\begin{eqnarray*}
+  \bm{\tilde \sigma * \tilde \eta} &=& \bm{\tilde \eta}
+  \begin{pmatrix}
+    \sigma_1 &        & \\
+        & \ddots & \\
+        &        & \sigma_T
+  \end{pmatrix} \\
+  &=& \bm{\tilde \eta \bar \sigma}
+\end{eqnarray*}
+Thus
+\begin{eqnarray*}
+  \bm E &=& {1 \over T}\bm{\tilde \eta \bar \sigma^2 \tilde \eta' P}
+\end{eqnarray*}
+In the large T limit $\bm{\tilde \eta \bar \sigma^2 \tilde \eta'}$ and $P$ are
+freely independent, therefore their S-transforms are multiplicative,
+i.e.
+\begin{eqnarray*}
+  S_E(z) &=& S_{\tilde \eta \bar \sigma^2 \tilde \eta'/T}(z) S_P(z) \\
+  &=& S_{\tilde \eta' \tilde \eta /T}(z) S_{\bar \sigma^2}(z) S_P(z)
+\end{eqnarray*}
+For $S_{\tilde \eta' \tilde \eta /T}(z)$ there have been results when
+the elements of $\bm{\tilde \eta}$ are identically distributed with finite
+second moment \cite{burda2011} or alternatively, follow $\alpha$-stable
+distribution \cite{politi2010}.
 
+So our focus hereafter is on the spectral properties of
+$\bar \sigma^2$. The Green's function of $\bar \sigma^2$ can be immediately
+written down:
+\begin{eqnarray*}
+  G_{\bar \sigma^2}(z) &=& {1 \over T} \sum_{t=1}^T \frac{1}{z - \sigma_t^2}
+\end{eqnarray*}
+If the $\sigma_t$'s are independent, $G_{\bar \sigma^2}(z)$ is deterministic and
+equal to $\mean{\frac{1}{z - \sigma_t^2}}$; If, however, the $\sigma_t$'s are
+correlated, $G_{\bar \sigma^2}(z)$ is random. In the extreme case where the
+$\sigma_t$'s have correlation 1 between any pair, $G_{\bar \sigma^2}(z)$ is a
+random variable $\frac{1}{z - \sigma_1^2}$. In the latter case, although
+the spectrum of $\bm \bar \sigma^2$ is random, it may still make sense to
+know the expectation of the spectrum, which one obtains from the
+expectation of the random Green's function. In the rest of this
+section we consider a few cases in regard to the distribution of
+$\sigma_t$'s as well as their correlations.
+
+\subsection{log-normal volatilities with AR(1) autocorrelation}
+In this subsection we consider the case
+\begin{eqnarray*}
+  \ln \sigma_t &=& \phi\ln \sigma_{t-1} + x_t
+\end{eqnarray*}
+where $x_t \sim N(0, 1)$. Then, assuming the process was started in
+the infinite past, i.e. $t = -\infty$, $\mean{\ln \sigma_t \ln \sigma_s} =
+\frac{\phi^{|t-s|}}{1 - \phi^2}$.
+
+\bibliographystyle{unsrt}
+\bibliography{econophysics}
 \end{document}
+
+\message{ !name(draft4.tex) !offset(-127) }

### 2014-12-01 thesis/draft3.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
% \usepackage[bookmarks=true]{hyperref}
% \usepackage[acronym]{glossaries}
%\usepackage{bookmark}
\input{../physics_common.tex}

\DeclareGraphicsExtensions{.eps, .pdf}

\begin{document}
\section{Spectra with same q and different $T$}
The following figure shows the spectra of correlation matrices with
fixed ratio $q = T/N$, and varying $T$ and $N$.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4]{../pics/spectra_q10.eps}
  \caption{\small \it Spectra with q = 10}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.3]{../pics/spectra_q24.eps}
  \caption{\small \it Spectra with q = 24}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.3]{../pics/spectra_q50.eps}
  \caption{\small \it Spectra with q = 50}
\end{figure}

Apparently the spectral density is changed even when the ratio $q$ is left
unchanged by varying $N$ and $T$. For this matter, I notice that for
Burda et al's weighted covariance matrix,
\begin{eqnarray*}
  C &=& {1 \over T} \mtx{RWR'}
\end{eqnarray*}
the spectral density also has $T$ dependence via the parameter $\beta$:
\begin{eqnarray*}
  \beta &=& T(1 - \alpha)
\end{eqnarray*}
where $\alpha$ in their IGARCH(1) model is analogous to $\beta_1$ in a
GARCH(1, 1) model.

\section{Why large deviation from MP at small $\alpha_1$}
Figure \ref{fig:Distance_to_MP} shows that when $\alpha_1$ is small,
deviation of the spectral density from the MP law is large with all
values of the tail exponent $\alpha$. Explanation to this can be found
in the autocorrelations of a GARCH(1,1) process.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.6]{../pics/Distance_to_MP.eps}
    \caption{\small \it Distance of the empirical spectral density to
      the Marchenko-Pastur law measured in terms of the mean square
      error between the two density functions}
    \label{fig:Distance_to_MP}
\end{figure}

A detailed calculation reveals the following autocorrelation
structure for a GARCH(1,1) process:
\begin{eqnarray*}
  \rho_n = \text{corr}(\epsilon_t^2, \epsilon_{t-n}^2) &=& (\alpha_1
  + \beta_1)^{n-1} \rho_1 \\
  \rho_1 = \text{corr}(\epsilon_t^2, \epsilon_{t-1}^2) &=& \alpha_1 +
  \beta_1 - \frac{\beta_1 [1 - (\alpha_1 + \beta_1)^2]}{
    1 - (\alpha_1 + \beta_1)^2 + \alpha_1^2
  }
\end{eqnarray*}
where $n \geq 2$. Figure \ref{fig:autocorr_1} and
\ref{fig:autocorr_1_full} shows $\rho_1$ as a function of $\alpha_1$
and $\beta_1$.
\begin{figure}[htb!]
  \centering
  \subfigure[$\rho_1 = \text{corr}(\epsilon_t^2, \epsilon_{t-1}^2)$ as
  a function of $\alpha_1$ and $\beta_1$. The plot is restricted to a
  region of small $\alpha_1$ and large $\beta_1$ such that $\alpha_1 +
  \beta_1 \approx 1$.]{
    \includegraphics[scale=0.8]{../pics/autocorr_1.pdf}
    \label{fig:autocorr_1}
  }
  \subfigure[$\rho_1 = \text{corr}(\epsilon_t^2, \epsilon_{t-1}^2)$ as
  a function of $\alpha_1$ and $\beta_1$. The plot covers the entire
  region of $\alpha_1 > 0$, $\beta_1 > 0$ and $\alpha_1 + \beta_1 <
  1$. Please ignore the spike, which is due to numerics.]{ 
    \includegraphics[scale=0.5]{../pics/autocorr_1_full.pdf}
    \label{fig:autocorr_1_full}
  }
\end{figure}
It is seen that, as $\alpha + \beta \to 1$, $\rho_1 \to 1$ even if
$\alpha_1$ is rather small. On a curve of constant-$\alpha$, as
$\alpha_1 \to 0$, $\alpha + \beta \to 1$ and hence $\rho_1 \to
1$. Thus for a fixed $\alpha$, very small $\alpha_1$ implies very
strong auto-correlation and hence large deviation of the spectral
density from the MP law.
\end{document}

### 2014-12-01 thesis/paper.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
% \usepackage[acronym]{glossaries}
%\usepackage{bookmark}
\input{../physics_common.tex}

\DeclareGraphicsExtensions{.eps}
\title{Spectral properties of a Wishart-GARCH Matrix}
\author{Prof. Sven \AA berg\\
  Xie Xiaolei}
\begin{document}
\maketitle
\section{The Spectral Distribution}
\subsection{Dependence on auto-correlation}
Figure \ref{fig:spectral_dist} shows the spectral density of the
correlation matrix of N identically distributed GARCH(1,1)
processes. These processes have such parameters that 
the $\alpha_1 + \beta_1 \approx 1$ and hence the auto-correlation
function $\text{corr}(r^2_t, r^2_{t-k})$ decays with a similar slow
rate for all the processes but from different initial values. The
parameter settings also make these processes have very similar tail
behavior with a tail exponent $\alpha$ in the range $2.17 \pm 0.01$.

Figure \ref{fig:spectral_dist1} shows the central part of the spectral
distribution and figure \ref{fig:spectral_dist2} shows how the
Hellinger distance between the empirical spectral density and the
Marcenko Pastur law changes with the parameter $\alpha_1$.
It is seen that, when measured by the Hellinger metric, the empirical
spectral density is most similar to the Marcenko-Pastur law when
$\alpha_1 = 1/2$.
\begin{figure}[htb!]
  \centering
  \subfigure[Spectral density of the correlation matrix plotted on
  log-log scale. The mean is 1 for all the empirical distributions and
  also for the Marcenko-Pastur law.]{
    \includegraphics[scale=0.32]{../pics/spectral_dist2.eps}
    \label{fig:spectral_dist2}
  }
  \hspace{5mm}
  \subfigure[Hellinger distance between the empirical spectral density
  and the Marcenko-Pastur law.]{
    \includegraphics[scale=0.32]{../pics/spectral_dist1.eps}
    \label{fig:spectral_dist1}
  }
  \caption{\small \it Spectral density of the
    correlation matrix of N identically distributed GARCH(1,1)
    processes.}
  \label{fig:spectral_dist}
\end{figure}
On the other hand, it is also seen that the variance of the spectral
distributions decreases as auto-correlation becomes stronger. This is
further shown in figure \ref{fig:spectral_dist3}.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../pics/spectral_dist3.eps}
  \caption{\small \it Variance of the spectral distributions.}
  \label{fig:spectral_dist3}
\end{figure}

\subsection{Dependence on the tail exponent}
Figure \ref{fig:spectral_density1} shows the dependence of the
spectral density of Wishart-GARCH matrices on the tail exponent
$\alpha$.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.4]{../pics/spectral_density1.eps}
    \caption{\small \it Dependence of spectral density (plotted on
      log-log scale) on the tail exponent. The means of the 3
      distributions are all equal to 1 and the variances are all equal
      to 0.04}
    \label{fig:spectral_density1}
\end{figure}
The mean and variance of the three density functions as well as their
Hellinger distances to the Marcenko-Pastur law are as follows:
\begin{table}[htb!]
  \centering
  \begin{tabular}{|l|c|c|c|}
    \hline
    Tail Exponent & 2.17 & 2.99 & 3.99 \\
    \hline
    mean & 1.0000 & 1.0000 & 1.0000 \\
    \hline
    variance & 0.0418 & 0.0409 & 0.0408 \\
    \hline
    Hellinger distance & 0.1012 & 0.0639 & 0.0586 \\
    \hline
  \end{tabular}
  \caption{mean, variance and Hellinger distance to the
    Marcenko-Pastur law of the three density functions in figure
    \ref{fig:spectral_density1}}
  \label{tab:spectral_density_hellinger}
\end{table}

\section{Distribution of the largest eigenvalue}
\subsection{Dependence on auto-correlation}
Chiani showed that the transformed largest eigenvalue of a Wishart
matrix obeyed approximately a gamma distribution \cite{Chiani2012}. Figure
\ref{fig:eig1_dist1} shows how this distribution is modified as the
auto-correlation between the squared returns becomes stronger; figure
\ref{fig:eig1_dist2} shows the growth of the difference between the
empirical distribution and a gamma distribution that has the same
mean and variance. It is clear that the logarithm of the Kolmogorov
distance is approximately linear in the GARCH parameter $\alpha_1$.

\begin{figure}[htb!]
  \centering
  \subfigure[Cumulative distribution of the largest eigenvalue]{
    \includegraphics[scale=0.33]{../pics/maxeigen_dist_garch.eps}
    \label{fig:eig1_dist1}
  }
  \subfigure[Kolmogorov distance between the empirical cumulative
  distribution and a gamma distribution with the same mean and variance]{
    \includegraphics[scale=0.45]{../pics/maxeigen_dist_garch_ks.eps}
    \label{fig:eig1_dist2}
  }
  \caption{\small \it Distribution of the largest eigenvalue and
    comparison to a gamma distribution.}
  \label{fig:eig1_dist}
\end{figure}

Figure \ref{fig:eig1_dist3} shows the density function of the largest
eigenvalue and figure \ref{fig:eig1_dist4} shows the tail of the
complementary distribution function of the largest eigenvalue. Here by
``normalized'' we mean the following transformation:
\begin{eqnarray*}
  \lambda'_1 &=& {\lambda_1 - \E(\lambda_1) \over \text{std}(\lambda_1)}  
\end{eqnarray*}

\begin{figure}[htb!]
  \centering
  \subfigure[Probability density function of the normalized largest
  eigenvalue]{
    \includegraphics[scale=0.32]{../pics/maxeigen_dist_pdf.eps}
    \label{fig:eig1_dist3}
  }
  \hspace{5mm}
  \subfigure[Complementary distribution function of the
  normalized largest eigenvalue on log-log scale.]{
    \includegraphics[scale=0.32]{../pics/maxeigen_dist_tail.eps}
    \label{fig:eig1_dist4}
  }
  \caption{\small \it Tail behavior of the largest eigenvalue.}
  \label{fig:eig1_dist_tail}
\end{figure}

\subsection{Dependence on the tail exponent}
Figure \ref{fig:eig1_dist_diff_alpha} shows how the distribution of
the largest eigenvalue depends on the tail exponent.
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.4]{../pics/eig1_dist_diff_alpha.eps}
    \caption{\small \it Dependence of the largest eigenvalue's
      probability density function on the tail exponent.}
    \label{fig:eig1_dist_diff_alpha}
\end{figure}

The mean and variance of the largest eigenvalue's distributions are
listed in table \ref{tab:eig1_dist_diff_alpha}
\begin{table}[htb!]
  \centering
  \begin{tabular}{|l|c|c|c|}
    \hline
    Tail Exponent & 2.17 & 2.99 & 3.99 \\
    \hline
    $\mean{\lambda_1}$ & 1.4306 & 1.4157 & 1.4137 \\
    \hline
    $\text{var}(\lambda_1)$ & 0.0012 & 0.0006 & 0.0005 \\
    \hline
  \end{tabular}
  \caption{mean and variance of the largest eigenvalue's distribution}
  \label{tab:eig1_dist_diff_alpha}
\end{table}

\bibliographystyle{unsrt}
\bibliography{econophysics}

\end{document}

### 2014-12-01 thesis/price_models2.tex

@@
 
-\S\ref{sec:Garch_model} reviews the GARCH model using the Nordea
-15-minute returns as an example. \S\ref{sec:SLV_model} examines the
+section \ref{sec:Garch_model} reviews the GARCH model using the Nordea
+15-minute returns as an example. Section \ref{sec:SLV_model} examines the
 stochastic log-volatility model and discusses the possibility of
@@
 Johnson Su distribution. The same Nordea return series is fitted as an
-example. \S\ref{sec:XieCalc} derives the unconditional distribution
+example. Section \ref{sec:XieCalc} derives the unconditional distribution
 function of the stochastic log-volotility model and compares the
@@
 The picture depicted above is of course overly simplified, and as we discussed in
-\S\ref{sec:StylizedFacts}, the distribution of returns is not really
+section \ref{sec:StylizedFacts}, the distribution of returns is not really
 Gaussian. Nevertheless, the assumption of Gaussian distributed returns
@@
 1 and corollary 1), is approximated as (see \cite{Protter05}, ch.II,
-\S6, theorem 22):
+section 6, theorem 22):
 \[
@@
 Swedish stocks whose data have been collected by the author with
-home-developed software. In \S\ref{sec:nordea} and \S\ref{sec:volvo}
+home-developed software. In section \ref{sec:nordea} and section \ref{sec:volvo}
 we model the log-volatilities of Nordea Bank 15-minute returns and
@@
 the Stochastic Log-Volatility model specified as equation
-\ref{eq:SLV_spec}. As is discussed in \S\ref{sec:SLV_model},
+\ref{eq:SLV_spec}. As is discussed in section \ref{sec:SLV_model},
 $\ln\sigma_t$ can be described by an ARMA or ARIMA \footnote{ARIMA ---
@@
 
-In \S\ref{sec:SLV_Symmetric} we first study the model in the
+In section \ref{sec:SLV_Symmetric} we first study the model in the
 relatively simple case when $v$ and $b$ are uncorrelated and $\mu
 = 0$. If this simplified version proves inadequate, one may resort to
-the general model studied in \S\ref{sec:SLV_Asymmetric}.
+the general model studied in section \ref{sec:SLV_Asymmetric}.
 
@@
 including the Volvo B 30-minute returns described in
-\S\ref{sec:SLV_Symmetric}. Table \ref{tab:assets_params} and
+section \ref{sec:SLV_Symmetric}. Table \ref{tab:assets_params} and
 \ref{tab:assets_moments} present the obtained parameter values and the
@@
 correlation translates to the correlation between the residual of the
-log-volatility, denoted $y_t$ in \S\ref{sec:nordea} and
-\S\ref{sec:volvo}, and $b_t$. Now let us consider the forecast
+log-volatility, denoted $y_t$ in section \ref{sec:nordea} and
+section \ref{sec:volvo}, and $b_t$. Now let us consider the forecast
 function of an ARIMA model:
@@
 
-For the Nordea series considered in \S\ref{sec:nordea},
+For the Nordea series considered in section \ref{sec:nordea},
 $\text{corr}(y_t, b_t)$ is found to be $3.56 \times 10^{-2}$; while
-for the Volvo series considered in \S\ref{sec:volvo},
+for the Volvo series considered in section \ref{sec:volvo},
 $\text{corr}(y_t, b_t)$ is found to be $-7.3 \times 10^{-3}$. These

### 2014-12-01 thesis/summary.tex

\chapter{Results}
\label{chp:summary}
Realistic dynamic models for relative price changes (returns) of
financial assets have been studied. In particular, the \gls{garch} and
\gls{sv} models have been utilized and compared in their forecast
accuracies by case studying of 15- and 30-minute returns of Nordea,
Volvo and Ericsson. The following are the conclusions from this
investigation:
\begin{itemize}
\item In all the 7 studied series, the log-volatility $\ln \sigma_t$
  is well described by a seasonally integrated moving average
  model. Long memory in these series is accounted for by the compounded
  difference operator $(1-B)(1-B^s)$ where $B$ denotes the back-shift
  operator. $s$ stands for seasonality, which is 33 in the cases of
  15-minute returns and 16 in the cases of 30-minute returns.

\item An \gls{sv} model generally yields more accurate forecasts than does
  the \gls{garch} model for the same series. This is certainly well expected,
  considering that the \gls{sv} model incorporates much more data than does
  \gls{garch}. However, it must be noted that the validity of the
  aforementioned comparison is underlain by the accuracy of realized
  volatility as a proxy to the true conditional volatility.

\item \gls{sv} models perform more consistently in terms of forecast
  accuracy than does \gls{garch}, as shown in appendix
  \ref{sec:forecast_volatility}.
\end{itemize}

Regarding the covariance matrix of Gaussian return series, we have
arrived at the following results:
\begin{itemize}
\item Auto-correlations in the returns rescale the covariance of two
  series by a factor of $1/(1 - \phi^2)$, where $\phi =
  \text{corr}(r_t, r_{t-1})$. See
  eq. \ref{eq:gaussian_mean2}. Moreover, they rescale the variance of
  each series by ${1 \over (1 - \phi^2)}\left[
    1- {\phi^2 \over T(1 - \phi^2)}
  \right]$. See eq. \ref{eq:gaussian_cii_mean2}.

\item Auto-correlations in the returns increase the variance of the
  covariance and also the variance of the variance. See
  eq. \ref{eq:gaussian_variance2} and
  \ref{eq:gaussian_cii_variance2}.

\item The largest eigenvalue obeys approximately a gamma distribution
  even when the returns are auto-correlated. It was shown by Chiani in
  2012 that the largest eigenvalue obeyed gamma distribution when the
  returns were {\it not} auto-correlated \cite{Chiani2012}. In
  addition, we find that the mean of the gamma distribution is
  approximately quadratic in $\tan{\pi\phi \over 2}$.
\end{itemize}

For the covariance matrix of \gls{garch}(1,1) series, our
contributions are the following:
\begin{itemize}
\item The diagonal and non-diagonal matrix elements both have L\'evy
  distributions but with different L\'evy indices.
\item The power-law tails of \gls{garch}(1,1) returns lead to a group
  of localized eigenvectors that correspond to large eigenvalues.
\item Auto-correlations in the returns reduce the localization of the
  eigenvectors. The fraction of localized eigenvectors decreases
  approximately as a quadratic function of $\phi = \text{corr}(r_t,
  r_{t-1})$. See figure \ref{fig:localization_ratio} and
  \ref{fig:localization_ratio2}.
\end{itemize}

In summary, auto-correlations in the returns create illusory
cross-correlations. To assess the true cross-correlations among the
assets in a market, one has to adopt a model for each of the return
series, infer the residuals and then assess the cross-correlations
among the residuals instead (c.f. chapter \ref{chp:PriceModels}).

However, in an efficient market, auto-correlations in the returns are
necessarily as weak as indistinguishable from measurement errors 
so that exploitable arbitrage opportunities do not exist. Hence
auto-correlations cannot be completely eliminated by taking residuals
of the returns. When estimating the covariance matrix and its
eigenvalues and eigenvectors, the illusory effects caused by
auto-correlations must be considered as an inherent source of
uncertainty.

% Correlations between sequences of stock returns from different assets
% have been studied by forming a covariance matrix. The stock returns
% have been simulated using different dynamical return models. Since a
% \gls{sv} model with Gaussian innovations have an unconditional
% distribution with all moments finite (see section \ref{sec:XieCalc}),
% we have focused our attention on \gls{garch}(1,1) models that were
% proven to have regularly varying (power-law) tails. We have
% investigated how a covariance matrix of \gls{garch}(1,1) returns differs from a
% covariance matrix of Gaussian returns, in terms of matrix elements and
% eigenvalues distributions. In particular, we have considered how the
% two matrices contrast with each other when the return series that they
% derive from are auto-correlated.

% In chapter \ref{chp:Gaussian} we studied how the distributions of
% elements and eigenvalues of a covariance matrix of Gaussian returns
% were altered by auto-correlations in the returns. For the matrix elements
% distribution, asymptotic Gaussian distributions were obtained. It was
% seen that the mean and the variance of the matrix elements
% distributions were changed by auto-correlations
% (Eq. \ref{eq:gaussian_mean2} to \ref{eq:gaussian_cii_variance2}). As 
% for the distribution of the maximum eigenvalue of C, it was shown that
% the maximum eigenvalue had approximately a gamma distribution even
% when the returns were auto-correlated. The mean of the gamma
% distribution was approximately quadratic in $\tan{\pi\phi \over 2}$,
% where $\phi$ was the coefficient of an \gls{ar}(1) process and
% characterized the strength of auto-correlation. On the other hand, the
% minimum eigenvalue was also increased by increased
% auto-correlations. As a whole, the eigenvalue spectrum widened as
% auto-correlation increased. See figure
% \ref{fig:GaussianMarkovSpectrumPDF}.

% In chapter \ref{chp:CrossCorrelationFat} we studied a covariance
% matrix of asset returns described by the \gls{garch}(1,1) 
% model. We found that both the diagonal and the non-diagonal matrix
% elements followed $\alpha$-stable (L\'evy) distributions, but had
% different L\'evy indices and scaling parameters (equations
% \ref{eq:garch_cii_dist} and \ref{eq:garch_cij2}). The eigenvalue
% distribution of the covariance matrix, as in the case of 
% Gaussian returns, was found to widen as auto-correlations
% increased. Meanwhile, both the largest and the smallest eigenvalues
% were seen to increase. See figure
% \ref{fig:GarchSpectrumAutocorrelated}.

% We have also investigated how localization of eigenvectors,
% i.e. dominance of one or a few basis vectors in the composition of an
% eigenvector, is affected by auto-correlations in the returns. Firstly,
% it is seen that localized eigenvectors correspond to large
% eigenvalues; secondly, it is found that strengthened auto-correlations
% reduce the extent of localization. These are consistent with what \AA
% berg found for Wishart-L\'evy matrices \cite{Aberg2013}.

% In particular, we find that the fraction of localized eigenvectors, as
% defined by an eigenvector's participation radio (see equation
% \ref{eq:IPR_def}) being lower than a given threshold or by its largest
% component (in the sense of absolute value) being higher than a
% threshold, falls approximately as a quadratic function of the
% auto-correlation strength (autoregressive coefficient of an
% \gls{ar}(1) process). See figure \ref{fig:localization_ratio} and
% \ref{fig:localization_ratio2}.

\chapter{Outlook}
Continuing from the obtained results, areas of future research may
include, first of all, a covariance matrix formed from \gls{sv} models
with power-law tails, which may be obtained by assuming, for example,
Student's t distribution or Normal Inverse Gaussian distribution for
the return's innovation.

Another interesting area could be models that treat the conditional
covariance matrix as an inherent part of the model specification
rather than treating it as an inferred quantity, which is the approach
taken by the current work. Predecessors on this path are the
multivariate \gls{garch} and \gls{sv} models
\cite{Mikosch2009}. To have sufficient flexibility, many of these
models involve a large number of parameters, which make it hard to fit
them to data and increase the chances of model mis-specification. A
model that strikes a good balance between flexibility and complexity
will be of great interest.
These approaches may well lead to more accurate volatility forecasts
and improved estimation of the uncertainties introduced by
auto-correlations in the returns.

Analytically deriving the eigenvalue and eigenvector distributions of
a covariance matrix formed from return series described by a realistic
model will be very challenging but interesting. It is also interesting to
see whether the methods developed for Wishart matrices, for example
the holonomic gradient method \cite{Hashiguchi2012}, can be applied to
the aforementioned covariance matrices.
Advancement in these areas will undoubtedly find applications in
e.g. principle component analysis and portfolio management.

### 2014-12-08 thesis/draft4.tex
Various changes.
Modify scripts to help with the Lund paper.

@@
 \usepackage{multirow}
-\usepackage{wrapfig}
+% \usepackage{wrapfig}
 \usepackage[usenames,dvipsnames]{color}
@@
 \end{eqnarray}
-where $x_t \sim N(0, 1)$ and we have assumed that the $\log \sigma_t$
-process was started in the infinite past with $\log \sigma_{-\infty} =
-x_{-\infty}$. In addition, we assume $\log \sigma_t$ is a stationary
-process with $|\phi| < 1$. Clearly $\log \sigma_t$ is normally
-distributed with mean 0 and variance ${1 / (1 -
-  \phi^2)}$. According to \eqref{eq:greens}, the spectral density of
-$\bm E$ as appears in \eqref{eq:E_def} is the same as if
+where $x_t \sim N(0, \theta^2)$ and we have assumed that the $\log
+\sigma_t$ process was started in the infinite past with $\log
+\sigma_{-\infty} = x_{-\infty}$. In addition, we assume $\log
+\sigma_t$ is a stationary process with $|\phi| < 1$. Clearly $\log
+\sigma_t$ is normally distributed with mean 0 and variance ${\theta^2
+  / (1 - \phi^2)}$. According to \eqref{eq:greens}, the spectral
+density of $\bm E$ as appears in \eqref{eq:E_def} is the same as if
 $\sigma_t^2$'s were uncorrelated but $\log \sigma_t \sim N(0,
-1/(1 - \phi^2))$.
+\theta^2/(1 - \phi^2))$.
 
@@
   }
+  \subfigure[$q = 0.5$]{
+    \includegraphics[scale=0.35]{../pics/lambda_max_q0.5.eps}
+    \label{fig:lambda_max_q0.5}
+  }
+  \subfigure[$q = 1.0$]{
+    \includegraphics[scale=0.35]{../pics/lambda_max_q1.0.eps}
+    \label{fig:lambda_max_q1.0}
+  }
   \caption{\small \it Estimated $\E \lambda_\M$ against its linear

### 2014-12-08 m/paper1/estimate_v_daily.m

@@ -1,0 +1,35 @@
+clear all
+close all
+mysql = get_mysql();
+name='SP500';
+stmt = sprintf('select symbol from %s_components;', name);
+symbols = fetch(mysql, stmt);
+
+p = size(symbols, 1);
+T = 3000;
+numRec = NaN(p, 1);
+
+for k = 1 : p
+    numRec(k) = cell2mat(fetch(mysql, sprintf(['select count(*) from %s_US'], ...
+                                              strrep(symbols{k}, '.', '_'))));
+end
+
+to_include = find(numRec >= T+1);
+p = length(to_include);
+
+R = NaN(T, p);
+logvol = NaN(T, p);
+c = 1;
+for k = to_include'
+    data = fetch(mysql, sprintf(...
+        ['select closing, (high - low)/low from ' ...
+         '%s_US order by day desc limit %d;'], ...
+        strrep(symbols{k}, '.', '_'), T+1));
+    R(:, c) = price2ret(flipud(cell2mat(data(:, 1))));
+    A = flipud(cell2mat(data(2:end, 2)));
+    A(A == 0) = min(A(A ~= 0));
+    logvol(:, c) = log(A);
+    logvol(:, c) = logvol(:, c) - mean(logvol(:, c));
+    c = c + 1;
+end
+close(mysql);

### 2014-12-15 thesis/draft4.tex
New additions.

@@
 \maketitle
+
+\section{Introduction}
 
@@
 \sigma_t$ is normally distributed with mean 0 and variance ${\theta^2
-  / (1 - \phi^2)}$. According to \eqref{eq:greens}, the spectral
+  / (1 - \phi^2)}$. The auto-correlation function $\rho(n)$ decays
+exponentially: $\rho_n = \phi^{n-1} = exp[-(n-1)\ln(1/\phi)]$.
+According to \eqref{eq:greens}, the spectral
 density of $\bm E$ as appears in \eqref{eq:E_def} is the same as if
@@
     0]{../pics/LognormalBlue.eps}
-  \caption{\small \it The Blue function $B$. q = 0.45, v = 3}
+  \caption{\small \it The Blue function $B$. q = 0.5, v = 0.5. The
+    thick black curve in the left figure is the $\im B = 0$ curve. The
+    smallest (largest) eigenvalue is achieved at the left (right) end
+    of the curve.}
   \label{fig:LognormalBlue}
@@
 When $v$ is small, the integrands in \eqref{eq:LognormalBlueReal} and
-\eqref{eq:LognormalBlueImag} have sharp maxima at $v = 0$. Hence it is
+\eqref{eq:LognormalBlueImag} have sharp maxima at $s = 0$. Hence it is
 useful to approximate them by Taylor expansion. Let's write
@@
 
-Taylor-expanding $F(s)$ and $G(s)$ to 2nd order in $s$ and plugging
-them into \eqref{eq:LognormalBlueReal} and
+Taylor-expand $F(s)$ and $G(s)$ to 2nd order in $s$
+\begin{eqnarray*}
+\lambda &=& F(0) + {v \over 2}F^{''}(0) + O(v^2) \\
+0 &=& G(0) + {v \over 2}G^{''}(0) + O(v^2) \\
+\end{eqnarray*}
+Plugging them into \eqref{eq:LognormalBlueReal} and
 \eqref{eq:LognormalBlueImag} gives

### 2014-12-16 thesis/draft4.tex
Modiciations related to the paper with Sven.

@@
 \section{Introduction}
+Applying free probability thoery to the study of spectra of large
+random matrices has acheived encouraging results. M. Politi, E. Scalas
+and G. Germano \cite{politi2010} studied the empirical covariance
+matrix of iid sequences having an $\alpha$-stable distribution, and
+derived formulas for the spectral density. Burda et al
+\cite{burda2011} investigated empirical covariance matrices of not
+only iid sequences but also those with auto- and
+cross-correlations. In the same paper, they also studied the case of
+I-GARCH sequences ($\sigma^2_t = \alpha \sigma^2_{t-1} + (1 - \alpha)
+r^2_{t-1}$) and derived formulas for the spectral density in the
+limiting situation $T(1-\alpha) \to \text{const}$.
 
-\section{Separation of the Heteroscedasticity Effect}
+However, covariance matrices of sequences of a stochastic volatility
+(SV) model, which often gives more accurate volatility forecast than
+do GARCH and its variates, seem to have eluded researchers'
+attention. In this paper, we present some results about the spectral
+density and minimum/maximum eigenvalues of a covariance matrix of SV,
+and in particular, lognormal-volatility return sequences where the
+volatilities have autocorrelations.
+
+Section \ref{sec:sep} summerizes the techniques of separating the
+heteroscedasticity effect, which are also found in
+\cite{biroli2007student} and \cite{burda2011}, and we argue that, in
+the special case where the sequences always have the same
+conditional volatilities, heteroscedasticity affects the spectral
+properties of the covariance matrix only through the unconditional
+distribution of the volatility.
+
+Section \ref{sec:lognormal} presents some results in the case where
+volatility is autocorrelated and conditionally lognormally
+distributed, while the innovations are iid normal.
+
+\section{Separation of the Heteroscedasticity Effect}\label{sec:sep}
 Consider a covariance matrix constructed as
@@
 only through the stationary distribution of $\sigma_t^2$.
-In the following subsections we discuss situations where $\sigma_t$'s
-are chosen from different conditional distributions and have different
-autocorrelation structures.
 % Then the moment generating function $M_{\bar \sigma^2}$ is
@@
 
-\section{Normal innovations \& log-normal volatilities with AR(1)
-  autocorrelation}
+\section{Normal innovations \& lognormal volatilities with AR(1)
+  autocorrelation}\label{sec:lognormal}
 In this subsection we consider the case
@@
   / (1 - \phi^2)}$. The auto-correlation function $\rho(n)$ decays
-exponentially: $\rho_n = \phi^{n-1} = exp[-(n-1)\ln(1/\phi)]$.
+exponentially: $\rho_n = \phi^{n-1} = \exp[-(n-1)\ln(1/\phi)]$.
 According to \eqref{eq:greens}, the spectral
@@
 \end{eqnarray}
-where $v = 1/(1 - \phi^2)$, reflecting the effect
+where $v = \theta^2/(1 - \phi^2)$, reflecting the effect
 of autocorrelation in $\log \sigma_t$'s. However, no closed form of
@@
   \centering
-    \includegraphics[scale=0.45, clip=true, trim=66 0 84
+    \includegraphics[scale=0.45, clip=true, trim=66 0 64
     0]{../pics/LognormalBlue.eps}
@@
 \end{figure}
-From figures \ref{fig:LognormalSpectra} one can see that the
-Marcenko-Pastur law is recovered when $v \to 0$, as expected.
 
@@
 \end{eqnarray}
-Now that we look for a solution in the limit $y \to 0$, we set $y = 0$
-in \eqref{eq:lambda_linear}; in \eqref{eq:constraint_linear} we first
-cancel the factor $y$ and then set $y = 0$. This yields
+When $v = 0$ and $\lambda = \lambda^{MP}_{\pm} = (1 \pm
+\sqrt{q})^2$ the Marcenko-Pastur law is recovered, as seen in figure
+\ref{fig:LognormalSpectra}. When $v > 0$, since we look for a solution
+in the limit $y \to 0$, we set $y = 0$ in \eqref{eq:lambda_linear}; in
+\eqref{eq:constraint_linear} we first cancel the factor $y$ and then
+set $y = 0$. This yields
 \begin{eqnarray}
@@
 \end{eqnarray}
+
 Evaluating \eqref{eq:lambda_1} at $x = \lambda^{MP}_{+} = (1 +
@@
   \lambda_{+} &=& \lambda_{+}^{MP} [1+2v(2\sqrt q +
-  1)] \label{eq:lambda_3}
+  1)] \label{eq:lambda_3} \\
+  \lambda_{-} &=& \lambda_{-}^{MP} [1+2v(2\sqrt q -
+  1)] \label{eq:lambda_4}
 \end{eqnarray}
@@
 \ref{fig:lambda_verified}. Clearly one can see that the linear
-approximation \eqref{eq:lambda_3} fits fairly well to the estimated
+approximation \eqref{eq:lambda_3} fits fairly well to the simulated
 $\E \lambda_\M$.
@@
   }
-  \caption{\small \it Estimated $\E \lambda_\M$ against its linear
+  \caption{\small \it Simulated $\E \lambda_\M$ against its linear
     approximation.}
   \label{fig:lambda_verified}
+\end{figure}
+
+The same comparison for the smallest eigenvalue is shown in figure
+\ref{fig:lambda_min_verified}.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[$q = 0.1$]{
+    \includegraphics[scale=0.35]{../pics/lambda_min_q0.1.eps}
+    \label{fig:lambda_min_q0.1}
+  }
+  \subfigure[$q = 0.2$]{
+    \includegraphics[scale=0.35]{../pics/lambda_min_q0.2.eps}
+    \label{fig:lambda_min_q0.2}
+  }
+  \subfigure[$q = 0.5$]{
+    \includegraphics[scale=0.35]{../pics/lambda_min_q0.5.eps}
+    \label{fig:lambda_min_q0.5}
+  }
+  \subfigure[$q = 1.0$]{
+    \includegraphics[scale=0.35]{../pics/lambda_min_q1.0.eps}
+    \label{fig:lambda_min_q1.0}
+  }
+  \caption{\small \it Simulated $\E \lambda_\m$ against its linear
+    approximation.}
+  \label{fig:lambda_min_verified}
 \end{figure}

### 2014-12-18 thesis/draft4.tex

@@
 \section{Introduction}
-Applying free probability thoery to the study of spectra of large
-random matrices has acheived encouraging results. M. Politi, E. Scalas
-and G. Germano \cite{politi2010} studied the empirical covariance
-matrix of iid sequences having an $\alpha$-stable distribution, and
-derived formulas for the spectral density. Burda et al
-\cite{burda2011} investigated empirical covariance matrices of not
-only iid sequences but also those with auto- and
-cross-correlations. In the same paper, they also studied the case of
-I-GARCH sequences ($\sigma^2_t = \alpha \sigma^2_{t-1} + (1 - \alpha)
-r^2_{t-1}$) and derived formulas for the spectral density in the
-limiting situation $T(1-\alpha) \to \text{const}$.
+In financial applications such as principle component analysis, it is
+of major interest to identify the most influential factors in a
+portfolio, whose variances are the largest eigenvalues of the
+covariance matrix of the portfolio \cite{Potters2003}. For this
+purpose, it is important to know the eigenvalue distribution of a
+large random covariance matrix. This distribution constitutes a
+background noise that must be filtered out before the true eigenvalues
+can be reliablly identified.
+
+Applying free probability thoery to the study of eigenvalue
+distributions of large random matrices has acheived encouraging
+results. M. Politi, E. Scalas and G. Germano \cite{politi2010} studied
+the empirical covariance matrix of iid sequences having an
+$\alpha$-stable distribution, and derived formulas for the spectral
+density. Burda et al \cite{burda2011} investigated empirical
+covariance matrices of not only iid sequences but also those with
+auto- and cross-correlations. In the same paper, they also studied the
+case of I-GARCH sequences ($\sigma^2_t = \alpha \sigma^2_{t-1} + (1 -
+\alpha) r^2_{t-1}$) and derived formulas for the spectral density in
+the limiting situation $T(1-\alpha) \to \text{const}$.
 
@@
 \end{eqnarray*}
-where $q = N/T$ and $\sigma$ has the stationary distribution of
-$\sigma_t$, whose probability density function is $P(s)$. Now we are
-specifically considering lognormally distributed $\sigma_t$, so we
-rewrite the above equation as
+where $z$ is a complex argument $z = x + i y$, $q = N/T$ and $\sigma$
+has the stationary distribution of $\sigma_t$, whose probability
+density function is $P(s)$. Now we are specifically considering
+lognormally distributed $\sigma_t$, so we rewrite the above equation as
 \begin{eqnarray}
@@
   1)] \label{eq:lambda_3} \\
-  \lambda_{-} &=& \lambda_{-}^{MP} [1+2v(2\sqrt q -
+  \lambda_{-} &=& \lambda_{-}^{MP} [1-2v(2\sqrt q -
   1)] \label{eq:lambda_4}
@@
   \caption{\small \it Simulated $\E \lambda_\m$ against its linear
-    approximation.}
+    approximation. Green: $\E \lambda_\m$ of simulated
+    eigenvalues. Red: linear approximation.}
   \label{fig:lambda_min_verified}

### 2014-12-22 thesis/draft4.tex
Work on the paper with Sven.

@@
 
+QQ-plots of the smallest/largest eigenvalues are shown in
+figure \ref{fig:eigmin_qq_plot} and \ref{fig:eigmax_qq_plot}. It is
+clearly seen that normal distributions are fairly good approximation
+to the distribution of the smallest eigenvalue in all cases, and the
+largest eigenvalue in all but the last two values of $v$, namely $v =
+0.05$ and $v = 0.08$.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.6]{../pics/eigmin_dist.eps}
+  \caption{\small \it QQ-plot of the smallest eigenvalues}
+  \label{fig:eigmin_qq_plot}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.6]{../pics/eigmax_dist.eps}
+  \caption{\small \it QQ-plot of the largest eigenvalues}
+  \label{fig:eigmax_qq_plot}
+\end{figure}
+
 \bibliographystyle{unsrt}

### 2014-12-22 m/test3.m

@@ -16,58 +16,60 @@ spec = cellstr(['b  '; 'c  '; 'g  '; 'm  '; 'r  '; 'k  ';...
 texts = {};
 c = 1;
 % hold on
-q = 1;
 lam1 = NaN(length(sig), length(phi), 2);
 lam2 = NaN(length(sig), length(phi), 2);
 variance = NaN(length(sig), length(phi));
 
-for m = 1 : length(sig)
-    for n = 1 : length(phi)
-        % for sig = 0.1
-        v = sig(m)^2/(1 - phi(n)^2);
-        variance(m, n) = v;
-        load(sprintf(['../matfys/data/sv/normal_ret/lognormal_vol/q%.1f/Eig-' ...
-                      'sig%.4f-phi%.4f.mat'], q, sig(m), phi(n)), 'ev');
+for q = [0.1, 0.2, 0.5, 1]
+    for m = 1 : length(sig)
+        for n = 1 : length(phi)
+            % for sig = 0.1
+            v = sig(m)^2/(1 - phi(n)^2);
+            variance(m, n) = v;
+            load(sprintf(['../matfys/data/sv/normal_ret/lognormal_vol/q%.1f/Eig-' ...
+                          'sig%.4f-phi%.4f.mat'], q, sig(m), phi(n)), 'ev');
 
-        % ev = reshape(ev, 1, prod(size(ev)));
-        eigmax = max(ev)';
-        eigmin = min(ev)';
-        
-        lam2(m, n, :) = [mean(eigmax), (1+sqrt(q))^2 * ...
-                            (1 + 2*v*(2*sqrt(q)+1))];
-        lam1(m, n, :) = [mean(eigmin), (1-sqrt(q))^2 * ...
-                            (1 - 2*v*(2*sqrt(q)-1))];
-        
-        c = c + 1;
+            % ev = reshape(ev, 1, prod(size(ev)));
+            eigmax = max(ev)';
+            eigmin = min(ev)';
+            
+            lam2(m, n, :) = [max(eigmax), (1+sqrt(q))^2 * ...
+                             (1 + 2*v*(2*sqrt(q)+1))];
+            lam1(m, n, :) = [min(eigmin), (1-sqrt(q))^2 * ...
+                             (1 - 2*v*(2*sqrt(q)-1))];
+            
+        end
     end
[…]

### 2015-01-04 cv/References.tex
Add CV and cover letters.

\title{References}
% \date{\today}
\author{
Xie Xiaolei
}

\documentclass[12pt]{article}

\begin{document}
\maketitle

My recommenders are:
\begin{itemize}
\item Sven \AA berg, professor and Head of Mathematical Physics at
  Lund University
\item Ferdi Aryasetiawan, professor of Mathematical Physics at
  Lund University
\end{itemize}

\end{document}

### 2015-01-04 cv/ResearchCoverLetter.tex
Add CV and cover letters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Long Lined Cover Letter
% LaTeX Template
% Version 1.0 (1/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Matthew J. Miller
% http://www.matthewjmiller.net/howtos/customized-cover-letter-scripts/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt,stdletter,dateno,sigleft]{newlfm} % Extra options: 'sigleft' for a left-aligned signature, 'stdletternofrom' to remove the from address, 'letterpaper' for US letter paper - consult the newlfm class manual for more options

\usepackage{charter} % Use the Charter font for the document text

% \newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
%\includegraphics[width=1.2\linewidth]{logo.png}}} % Company/institution logo at the top left of the page
% \makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

\newlfmP{sigsize=10pt} % Slightly decrease the height of the signature field
%\newlfmP{addrfromphone} % Print a phone number under the sender's address
%\newlfmP{addrfromemail} % Print an email address under the sender's address
% \PhrPhone{Phone} % Customize the "Telephone" text
% \PhrEmail{Email} % Customize the "E-mail" text

% \lthUiuc % Print the company/institution logo

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\namefrom{Xie Xiaolei} % Name

\addrfrom{
%\today\\[12pt] % Date
% Kiviksgatan 7C \\ % Address
% 21440 Malm\"o
% Sweden
}

% \phonefrom{+46 70 489 0472} % Phone number

% \emailfrom{xie.xiaolei@gmail.com} % Email address

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------

\greetto{Dear Sir or Madam,} % Greeting text
\closeline{Sincerely Yours,} % Closing text

\nameto{
% Evaluation Committee
} % Addressee of the letter above the to address

\addrto{
% Recruitment Officer \\ % To address
% The Corporation \\
% 123 Pleasant Lane \\
% City, State 12345
}

%----------------------------------------------------------------------------------------

\begin{document}
\begin{newlfm}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

I understand from your department's website that you are offering Ph.D
stipends, and found it a truly exciting opportunity. I have hence
contacted Prof. Thomas Valentin Mikosch and asked for his
supervision. After a kind discussion with me about my master's 
thesis, his research area, and a very interesting Ph.D project in
random matrix theory (RMT) and extreme value theory (EVT), he
encouraged me to apply. The project that he outlined perfectly
continued my master's studies and I really look forward to it. So I
would like to present to you what I did as my master's thesis project
and what my research interests are.

I wrote my thesis under the supervision of Prof. Sven \AA berg at Lund
university in the department of mathematical physics. I first compared
GARCH and stochastic volatility (SV) models by case studying of
intraday returns of a few Swedish stocks, and then I continued to
study the covariance matrix of Gaussian and GARCH(1,1) returns. There
I encountered the problem of eigenvalues distribution of a Wishart
matrix and found its large and charming literature.

By numerical simulation, I studied in the cases of Gaussian and
GARCH(1,1) returns, how the matrix elements and eigenvalues were
distributed, how the eigenvectors were composed, and in particular,
how the situation was changed by autocorrelations in the return
series. For example, I found in the case of GARCH(1,1) returns that
prolonged return series increased the extent of the eigenvectors'
localization while strengthened autocorrelations decreased the extent
of localization.

Among the many fascinating things that I was led to by the thesis
project were the extreme value theory and the random matrix
theory. In my studies of the intraday stock returns, heavy-tailed
distributions, and in particular L\'evy distributions, are very
important, so I read books such as ``Modeling Extremal Events'',
``Handbook of Financial Time Series'', and papers about the tail
behavior of GARCH and SV models, as well as those on the eigenvalues
distribution of random matrices. Quite naturally, these books and
papers led me to Prof. Mikosch. So when I met him and he outlined the
Ph.D project on RMT and EVT, which followed directly from my master's
project, I knew it was truly a great opportunity for me.

If I am so honored to be accepted, I will, first of all, learn more
about EVT in connection with multivariate time 
series, and also about its application in the research of RMT. Then I
would like to explore, for example, if and how the methods developed in the
research of the largest eigenvalue distribution and the spectral
density of random matrices, particularly Wishart matrices, can
be applied, presumably with EVT, to find the eigenvalues distribution
of a covariance matrix of heavy-tailed returns, such as those generated
by GARCH or SV models.

In particular, I understand that the problem of the exact
distribution of the largest eigenvalue of a Wishart matrix has been
attacked by the holonomic gradient method as an alternative to the
more conventional approach of series expansion around the origin. It
would be interesting to see if the same methods are applicable to the
aforementioned problem.

I look forward to an interview with you. If you would like to know more
about me, you are most welcome to call me at +46 70 489 0472 or find
me on LinkedIn as ``Xie Xiaolei''. If you want to contact my thesis
supervisor, Prof. Sven \AA berg, he is reachable via telephone +46 46
222 96 33 or mobile +46 70 557 35 69.

\end{newlfm}
\end{document}

### 2015-01-04 cv/TwoColumnCV.tex
Add CV and cover letters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Two Column One Page Curriculum Vitae
% LaTeX Template
% Version 1.1 (24/1/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Alessandro (The CV Inn)
%
% IMPORTANT: THIS TEMPLATE NEEDS TO BE COMPILED WITH XeLaTeX
%
% This template uses several fonts not included with Windows/Linux by
% default. If you get compilation errors saying a font is missing, find the line
% on which the font is used and either change it to a font included with your
% operating system or comment the line out to use the default font.
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
\setromanfont[Mapping=tex-text]{Gentium} % Main document font
\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Century Schoolbook
  L} % Font for your name at the top
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\color{text1} % Sets the default text color for the whole document to the color defined as 'text1'

%----------------------------------------------------------------------------------------
%	TITLE
%----------------------------------------------------------------------------------------

\par{\centering{\sffamily\Huge John Smith}\\ % Your name
{\color{headings}\fontspec[Variant = 2]{Century Schoolbook L} Curriculum {Vit\fontspec[Variant = 3]{Century Schoolbook L}\ae}\\[15pt]\par} % Curriculum vitae text in the Century Schoolbook L font

%----------------------------------------------------------------------------------------

\begin{minipage}[t]{0.5\textwidth} % Start the left-hand side of the page
\vspace{0pt} % Trick for alignment

%----------------------------------------------------------------------------------------
%	WORK EXPERIENCE
%----------------------------------------------------------------------------------------

\section{Work Experience}

%------------------------------------------------
% WORK EXPERIENCE 1
%------------------------------------------------

{\raggedleft\textsc{Current, from Jan 2009}\par}

{\raggedright\large Trader at Enoi s.p.a., Milan\\
\textit{Physical Natural Gas Trading}\\[5pt]}

\normalsize{Provided analysis of spreads, in particular time and location spreads, with emphasis on trading opportunities on the curve, also taking into account the availability of storage facilities. Became familiar with logistic implications of physical gas trading and developed knowledge of continental virtual hubs for gas exchange.}\\

%------------------------------------------------
% WORK EXPERIENCE 2
%------------------------------------------------

{\raggedleft\textsc{July 2008 -- Oct 2008}\par}

{\raggedright\large 1\textsuperscript{st} year Analyst at \textsc{Lehman Brothers}, London\\
\textit{Commodities Structured Trading}\\[5pt]}

\normalsize{Developed spreadsheets for risk analysis on exotic derivatives on a wide array of commodities (ags, oils, precious and base metals), managed blotter and secondary trades on structured notes, liaised with Middle Office, Sales and Structuring for bookkeeping.}\\

%------------------------------------------------
% WORK EXPERIENCE 3
%------------------------------------------------

{\raggedleft\textsc{Summer 2007}\par}

{\raggedright\large Summer Intern at \textsc{Lehman Brothers}\\
\textit{Capital Markets}\\[5pt]}

\normalsize{Received pre-placed offer from the Exotics Trading Desk as a result of very positive review. Rated ``truly distinctive" for Analytical Skills and Teamwork.}\\

%------------------------------------------------
% WORK EXPERIENCE 4
%------------------------------------------------

{\raggedleft\textsc{Current, from Jan 2009}\par}

{\raggedright\large Trader at Enoi s.p.a., Milan\\
\textit{Physical Natural Gas Trading}\\[5pt]}

\normalsize{Provided analysis of spreads, in particular time and location spreads, with emphasis on trading opportunities on the curve, also taking into account the availability of storage facilities. Became familiar with logistic implications of physical gas trading and developed knowledge of continental virtual hubs for gas exchange.}\\

%------------------------------------------------
% WORK EXPERIENCE 5
%------------------------------------------------

{\raggedleft\textsc{July 2008 -- Oct 2008}\par}

{\raggedright\large 1\textsuperscript{st} year Analyst at \textsc{Lehman Brothers}, London\\
\textit{Commodities Structured Trading}\\[5pt]}

\normalsize{Developed spreadsheets for risk analysis on exotic derivatives on a wide array of commodities (ags, oils, precious and base metals), managed blotter and secondary trades on structured notes, liaised with Middle Office, Sales and Structuring for bookkeeping.}\\

%----------------------------------------------------------------------------------------

\end{minipage} % End the left-hand side of the page
\hfill
\begin{minipage}[t]{0.44\textwidth} % Start the right-hand side of the page
\vspace{0pt} % Trick for alignment

%----------------------------------------------------------------------------------------
%	COLORED BOX
%----------------------------------------------------------------------------------------

\colorbox{shade}{\textcolor{text1}{
\begin{tabular}{c|p{7cm}}
\raisebox{-4pt}{\textifsymbol{18}} & 123 Broadway, City, State 12345 \\ % Address
\raisebox{-3pt}{\Mobilefone} & +1 (1)23 4567890 \\ % Phone number
\raisebox{-1pt}{\Letter} & \href{mailto:john@smith.com}{john@smith.com} \\ % Email address
\Keyboard & \href{http://www.johnsmith.com}{http://www.johnsmith.com} \\ % Website
\end{tabular}
}
}\\[10pt]

%----------------------------------------------------------------------------------------
%	EDUCATION
%----------------------------------------------------------------------------------------

\section{Education}

\begin{tabular}{rl} % Start a table with two columns, one for dates and one for qualifications

%------------------------------------------------
% EDUCATION 1
%------------------------------------------------

2011 -- \textsc{Present} & \textbf{Doctor of Philosophy} \\ 
& \textsc{Computer Science} \\ 
& \textit{The University of California}\\
&\\

%------------------------------------------------
% EDUCATION 2
%------------------------------------------------

2010 & \textbf{Master of Science} \\ 
& \small First Class Honours \\
& \textsc{Computer Science} \\ 
& \textit{The University of California}\\
&\\

%------------------------------------------------
% EDUCATION 3
%------------------------------------------------

2009 & \textbf{Postgraduate Diploma}\\
& \small Distinction \\
& \textsc{Computer Science} \\
& \textit{The University of California}\\
&\\

%------------------------------------------------
% EDUCATION 4
%------------------------------------------------

2006 -- 2008 & \textbf{Bachelor of Business Studies}\\
& \textsc{Commerce} \\
& \textit{The University of California}

%----------------------------------------------------------------------------------------

\end{tabular}\\[10pt]

%----------------------------------------------------------------------------------------
%	AWARDS
%----------------------------------------------------------------------------------------

\section{Awards}

\begin{tabular}{rl}
2010	 & \textbf{Faculty of Science Masters Scholarship}\\
& \textit{The University of California}\\ \\

%------------------------------------------------

2008	 & \textbf{Top Achiever Award -- Commerce}\\
& \textit{Business Association of California}
\end{tabular}\\[10pt]

%----------------------------------------------------------------------------------------
%	COMPUTER SKILLS
%----------------------------------------------------------------------------------------

\section{Computer Skills}

\begin{tabular}{rl}
Basic Knowledge
& \textsc{java}, Adobe Illustrator\\ 
& \\
Intermediate Knowledge
& \textsc{python}, \textsc{html}, \LaTeX,\\
& OpenOffice, Linux,\\
& Microsoft Windows \\
& Computer Hardware \& \\
& Support\\ \\
\end{tabular}

%----------------------------------------------------------------------------------------
%	COMMUNICATION SKILLS
%----------------------------------------------------------------------------------------

\section{Communication Skills}

\begin{tabular}{rl}
\textsc{Conferences}
& Oral Presentation at the California \\
& Computer Science Conference -- 2010\\
& \\
\textsc{Posters}
& Poster at the Annual Programming \\
& Conference in Oregon -- 2009 \\
\end{tabular}\\[10pt]

%----------------------------------------------------------------------------------------

\end{minipage} % End right-hand side of the page

\end{document}

### 2015-01-04 cv/XieXiaolei-Analyst.tex
Add CV and cover letters.

% Jason R. Blevins - Curriculum Vitae
%
% Copyright (C) 2004-2010 Jason R. Blevins
% http://jblevins.org/projects/cv-template/
%
% You may use use this document as a template to create your own CV
% and you may redistribute the source code freely. No attribution is
% required in any resulting documents. I do ask that you please leave
% this notice and the above URL in the source code if you choose to
% redistribute this file.

\documentclass[10pt,letterpaper]{article}

\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsfonts, amsmath}
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
% Comment the following line to use the default Computer Modern font
% instead of the Palatino font provided by the mathpazo package.
% Remove the 'osf' bit if you don't like the old style figures.
%\usepackage[sc,osf]{mathpazo}

% In practice, I use the following font packages instead of mathpazo.
% beramono provides a nice fixed-width font. xagaramon uses the
% (commercial) Adobe Garamond font.
%\usepackage[scaled=0.75]{beramono}
%\usepackage[osf]{xagaramon}

% Set your name here
\def\name{Xie Xiaolei}

% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {computer, mobile, information security},
  pdftitle = {\name: Curriculum Vitae},
  pdfsubject = {Curriculum Vitae},
  pdfpagemode = UseNone
}

\geometry{
  body={6.5in, 9.0in},
  left=1.0in,
  top=1.0in
}

% Customize page headers
\pagestyle{myheadings}
\markright{\name}
\thispagestyle{empty}

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\ttfamily\mdseries\Large}
\subsectionfont{\ttfamily\mdseries\itshape\large}

% Other possible font commands include:
% \ttfamily for teletype,
% \sffamily for sans serif,
% \bfseries for bold,
% \scshape for small caps,
% \normalsize, \large, \Large, \LARGE sizes.

% Don't indent paragraphs.
\setlength\parindent{0em}

% Make lists without bullets and compact spacing
%\renewenvironment{itemize}{
\newenvironment{aperiod}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
    \setlength{\itemsep}{1em}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0.25em}
  }
}{
  \end{list}
}

\begin{document}

% Place name at left
{\huge Xie Xiaolei}

% Alternatively, print name centered and bold:
%\centerline{\huge \bf \name}

\vspace{0.25in}

\begin{minipage}[t]{0.5\textwidth}
  Citizenship: Sweden \\
  Phone: +46 70 489 0472 \\
  Email: \href{mailto:xie.xiaolei@gmail.com}{\tt xie.xiaolei@gmail.com} \\
  Address: Kiviksgatan 7C, 21440 Malm\"{o}, Sweden
\end{minipage}

\section*{Education}
\begin{itemize}
\item Ph.D of Financial Mathematics, October 2014 - present
  \begin{itemize}
    \item {\it Reseach Areas:} I am mainly interested in the
      determination of the largest eigenvalues of a covariance matrix
      built from samples with heavy-tails. The purpose is dimension
      reduction, by which the behaviour of a diverse portfolio is
      replicated by one with much few components and hence analysis of
      the portfolio becomes more simplified and focused.

      I am also interested in efficient simulation of multivariate
      GARCH structures, which, among others, helps the determination of
      model parameters when fitting such a model to a large
      multi-dimensional data set.
  \end{itemize}

\item Master of Physics, Lund University, January 2011 - June 2014
  \begin{itemize}
  \item {\it Major Subject:} Econophysics. In my thesis I studied the
    average eigenvalue distribution of sample covariance matrices,
    with a focus on the impact of autocorrelations and heavy tails.

    % The returns' models are fitted to such Swedish stocks as Volvo B,
    % Ericsson B, and Nordea Bank, using both publicly available daily
    % statistics and intra-day transaction data that I have been
    % collecting using home-developed software.

    % The studies are carried out analytically wherever feasible, and
    % otherwise by approximation and simulation. The application of
    % probability theories, random matrices, and stochastic processes
    % are the most involved subjects.
  \end{itemize}

\item M.Sc.(Tech) Computer Science and Engineering, Helsinki
  University of Technology, 2005 - 2008.
  \begin{itemize}
  \item {\it Major Subject:} Mobile Computing - Services and Security
  \item {\it Average grade:} 4.17/5.
  \end{itemize}

\item B.Sc. Computer Science \& Technology, Beijing Institute of
  Technology, 2001- 2005.
  \begin{itemize}
  \item {\it Average grade:} 85.5/100.
  \item {\it Honors:} \textit{Distinguished Graduates of 2005}.
  \end{itemize}
\end{itemize}

\section*{Employment}
\begin{aperiod}
\item {\it \href{http://www.stericsson.com}{\tt
      ST-Ericsson}, Audio, Multimedia, Development Engineer, Sep 2008
    - Jan 2011}

  My main responsibility is software development in C, and sometimes
  C++ and Java, in the Linux operating system.

\item {\it \href{http://www.hiit.fi}{\tt Helsinki
      Institute of Information Technology}, Thesis Worker, Oct 2007 -
    May 2008}

  Funded by HIIT, I wrote my thesis ``A Mutual authentication and Key
  Agreement Protocol for the UMTS Network''. The algorithm of my
  protocol proposal was based on the Diffie-Hellman Exchange, but had
  significant modifications and extensions so that the mutual
  authentication between the mobile user and the network was delegated
  to the serving instead of the home network. The protocol design
  incorporated puzzle schemes that demanded unbalanced computing
  resources on the two sides to mitigate Denial-of-Service attacks.

\item {\it May 2007 - April 2004}
  I worked as software engineer in several companies. The programming
  languages used included C, C++, Java and python.
\end{aperiod}

\section*{Private Project}
{\it Trade data collector and analyzer}
With this home-developed software I have been collecting intra-day
transaction data and analyzing them in real-time with Matlab. I make
heavy use of the Matlab economitrics toolbox with its built-in
routines for estimating ARIMA and GARCH models and then forecasting
using the obtained model. The idea is to pre-estimate a model and thus
fix its structure, such as the distribution of the residuals, the
order of the auto-regressive and the moving-average components, the
order of differencing/fractional differencing, etc., and then to 
evolve the model parameters in real-time as new data are fed into
the system.
\end{document}

### 2015-01-04 cv/XieXiaolei-Computing.tex
Add CV and cover letters.

--- 2015-01-04 cv/XieXiaolei-Analyst.tex
@@
 \usepackage{geometry}
-\usepackage{amsfonts, amsmath}
 \usepackage[T1]{fontenc}
@@
 % Place name at left
-{\huge Xie Xiaolei}
+{\huge \name}
 
@@
 \begin{itemize}
-\item Ph.D of Financial Mathematics, October 2014 - present
+\item Master of Physics, Lund University, January 2011 - June 2014.
   \begin{itemize}
-    \item {\it Reseach Areas:} I am mainly interested in the
-      determination of the largest eigenvalues of a covariance matrix
-      built from samples with heavy-tails. The purpose is dimension
-      reduction, by which the behaviour of a diverse portfolio is
-      replicated by one with much few components and hence analysis of
-      the portfolio becomes more simplified and focused.
-
-      I am also interested in efficient simulation of multivariate
-      GARCH structures, which, among others, helps the determination of
-      model parameters when fitting such a model to a large
-      multi-dimensional data set.
+  \item {\it Major Subject:} Econophysics. My thesis project was
+    case studying of intraday returns of Nordea, Volvo and
+    Ericsson. Covariance matrices of GARCH and stochastic volatility
+    returns were also studied. Matlab programming was a crucial part.
   \end{itemize}
-
-\item Master of Physics, Lund University, January 2011 - June 2014
-  \begin{itemize}
-  \item {\it Major Subject:} Econophysics. In my thesis I studied the
-    average eigenvalue distribution of sample covariance matrices,
-    with a focus on the impact of autocorrelations and heavy tails.
-
-    % The returns' models are fitted to such Swedish stocks as Volvo B,
-    % Ericsson B, and Nordea Bank, using both publicly available daily
-    % statistics and intra-day transaction data that I have been
-    % collecting using home-developed software.
-
-    % The studies are carried out analytically wherever feasible, and
-    % otherwise by approximation and simulation. The application of
-    % probability theories, random matrices, and stochastic processes
-    % are the most involved subjects.
-  \end{itemize}
-
 \item M.Sc.(Tech) Computer Science and Engineering, Helsinki
@@
   \end{itemize}
-
 \item B.Sc. Computer Science \& Technology, Beijing Institute of
@@
 
-  My main responsibility is software development in C, and sometimes
-  C++ and Java, in the Linux operating system.
+  My responsibilities included developing audio low-level software,
+  typically device drivers, in the Linux OS. This involved designing and
+  implementing data structures and procedures so that the hardware
+  devices were used in the optimal way, in terms of parallelism,
+  power consumption, robustness, and ease of maintenance.
 
@@
   incorporated puzzle schemes that demanded unbalanced computing
-  resources on the two sides to mitigate Denial-of-Service attacks.
+  resources on the two sides to mitigate Denial-of-Service attacks. I
+  made C programs to test the puzzle schemes.
 
-\item {\it May 2007 - April 2004}
-  I worked as software engineer in several companies. The programming
-  languages used included C, C++, Java and python.
+\item {\it
+    \href{http://www.goodmillsystems.com}{\tt Goodmillsystems},
+    Software Designer, May 2007 - Sep 2007}
+
+  I developed a daemon program that ran on the company's smart router
+  product to collect statistical data about the router's network
+  traffic. The emphasis of the program was high performance in the
+  context of competing read and write accesses.
+
+\item {\it \href{http://www.cardinal.fi/}{\tt
+      Cardinal Information Systems}, Software Developer, Sep 2006 -
+    April 2007}
+
+  I implemented the FLUTE protocol for the company's DVB-H
+  mobile-end middleware. The implementation was multi-threaded and was
+  built on the {\it pthread} library. Good performance and robustness
+  were crucial requirements on the implementation.
+
+\item {\it
+    \href{http://www.digitalchocolate.com/}{\tt Digital Chocolate},
+    Game Engineer, Feb 2006 - Sep 2006}
+
+  At this job, my responsibilities were porting and customizing mobile
+  phone games for various handsets, and making game demos.
+
+\item {\it {\tt Beijing Holystone Ltd}, Software
+    Engineer, April 2004 - June 2005}
+
+  At this job I mainly developed JSP pages and database routines for
+  the company's information system.
 \end{aperiod}
 
-\section*{Private Project}
-{\it Trade data collector and analyzer}
-With this home-developed software I have been collecting intra-day
-transaction data and analyzing them in real-time with Matlab. I make
-heavy use of the Matlab economitrics toolbox with its built-in
-routines for estimating ARIMA and GARCH models and then forecasting
-using the obtained model. The idea is to pre-estimate a model and thus
-fix its structure, such as the distribution of the residuals, the
-order of the auto-regressive and the moving-average components, the
-order of differencing/fractional differencing, etc., and then to 
-evolve the model parameters in real-time as new data are fed into
-the system.
+\section*{Personal Project}
+{\it Trades data collector \& Analyzer}
+
+I have been developing my own stock-trading robot since the summer of
+2011. It is a software built on cURL, GLib, MySQL, and Matlab. I find
+it a fairly interesting, challenging, and rewarding combination of
+programming skills, data analysis techniques, and knowledge of
+financial time series. My knowledge in econophysics guides the
+development of this software.
+
+\section*{Non-English Language Skills}
+\begin{itemize}
+\item Swedish: informal communications. ``Swedish B'' passed.
+\item Chinese: Native speaker
+\end{itemize}
+
 \end{document}

### 2015-01-04 cv/XieXiaolei.tex
Add CV and cover letters.

--- 2015-01-04 cv/TwoColumnCV.tex
@@
 
+\usepackage{wasysym}
 \usepackage{marvosym} % Required for symbols in the colored box
@@
 
-\par{\centering{\sffamily\Huge John Smith}\\ % Your name
-{\color{headings}\fontspec[Variant = 2]{Century Schoolbook L} Curriculum {Vit\fontspec[Variant = 3]{Century Schoolbook L}\ae}\\[15pt]\par} % Curriculum vitae text in the Century Schoolbook L font
+\par{\centering{\sffamily\Huge Xie Xiaolei}\\ % Your name
+{\color{headings}\fontspec{Century Schoolbook L} Curriculum {Vit\fontspec{Century Schoolbook L}\ae}\\[15pt]\par} % Curriculum vitae text in the Century Schoolbook L font
 
@@
 \begin{minipage}[t]{0.5\textwidth} % Start the left-hand side of the page
-\vspace{0pt} % Trick for alignment
-
+%\vspace{0pt} % Trick for alignment
 %----------------------------------------------------------------------------------------
-%	WORK EXPERIENCE
+%	EDUCATION
 %----------------------------------------------------------------------------------------
 
-\section{Work Experience}
+\section{Education} 
+\begin{itemize} 
+\item \textit{Lund University, Sweden}
 
-%------------------------------------------------
-% WORK EXPERIENCE 1
-%------------------------------------------------
+\begin{tabular}{l p{0.7\textwidth}}
+  2011 -- \textsc{Present} & {M.Sc. Econophysics: Mathematical
+  Physics in Finance} \\
+  Thesis title & Returns Models and Their Covariance Matrix \\
+  Description & Compared accuracy of volatility forecasts of {\bf
+    GARCH} and {\bf stochastic volatility} (SV) models by case
+  studying Nordea, Volvo and Ericsson {\bf intraday} returns using the
+  two classes of models. Realized volatilities were computed and
+  compared to the forecasts. Also studied eigenvalues' distribution of
+  a covariance matrix of GARCH returns. {\bf MATLAB} programming was
+  an essential part.
+\end{tabular}
 
-{\raggedleft\textsc{Current, from Jan 2009}\par}
+\item {\it Helsinki University of Technology, Finland}
 
-{\raggedright\large Trader at Enoi s.p.a., Milan\\
-\textit{Physical Natural Gas Trading}\\[5pt]}
+\begin{tabular}{l p{0.7\textwidth}}
+  2005 -- 2008 & {\bf M.Sc. Computer Science} \\
+  Thesis Title & A Mutual Authentication and Key Agreement Protocol
+  for the UMTS Network \\
+  Description & The algorithm of the proposed protocol was based on
+  the Diffie-Hellman Exchange, but had significant modifications and
+  extensions so that the mutual authentication between the mobile user
+  and the network was delegated to the serving instead of the home
+  network. The protocol design incorporated puzzle schemes that
+  demanded unbalanced computing resources on the two sides to mitigate
+  Denial-of-Service attacks. \\
+  Average Grade & 4.17 / 5
+\end{tabular}
 
-\normalsize{Provided analysis of spreads, in particular time and location spreads, with emphasis on trading opportunities on the curve, also taking into account the availability of storage facilities.
-Became familiar with logistic implications of physical gas trading and developed knowledge of continental virtual hubs for gas exchange.}\\
+\item {\it Beijing Institute of Technology, China}
 
-%------------------------------------------------
-% WORK EXPERIENCE 2
-%------------------------------------------------
+\begin{tabular}{l p{0.7\textwidth}}
+  2001 -- 2005 & B.Sc. Computer Science \& Technology \\
+  Average Grade & 85.5 / 100 \\
+  Honor & Distinguished Graduates of 2005
+\end{tabular}
+\end{itemize}
 
-{\raggedleft\textsc{July 2008 -- Oct 2008}\par}
+\section{Personal Project}
+{\it Trades data collector and analyzer, June 2008 -- present}
 
-{\raggedright\large 1\textsuperscript{st} year Analyst at \textsc{Lehman Brothers}, London\\
-\textit{Commodities Structured Trading}\\[5pt]}
+Developed a Linux daemon program (programmed in C \& Matlab) to
+collect and analyze realtime transaction records on the Stockholm
+stock exchange. The program maintained and evolved stochastic
+volatility models and made volatility forecast in realtime. Collected
+data were managed with a MySQL database.
 
-\normalsize{Developed spreadsheets for risk analysis on exotic derivatives on a wide array of commodities (ags, oils, precious and base metals), managed blotter and secondary trades on structured notes, liaised with Middle Office, Sales and Structuring for bookkeeping.}\\
+\section{Non-English Language Skills}
+\begin{itemize}
+\item Swedish: Informal communications
+\item Chinese: Native speaker
+\end{itemize}
 
-%------------------------------------------------
-% WORK EXPERIENCE 3
-%------------------------------------------------
-
-{\raggedleft\textsc{Summer 2007}\par}
-
-{\raggedright\large Summer Intern at \textsc{Lehman Brothers}\\
-\textit{Capital Markets}\\[5pt]}
-
-\normalsize{Received pre-placed offer from the Exotics Trading Desk as a result of very positive review. Rated ``truly distinctive" for Analytical Skills and Teamwork.}\\
-
-%------------------------------------------------
-% WORK EXPERIENCE 4
-%------------------------------------------------
-
-{\raggedleft\textsc{Current, from Jan 2009}\par}
-
-{\raggedright\large Trader at Enoi s.p.a., Milan\\
-\textit{Physical Natural Gas Trading}\\[5pt]}
-
-\normalsize{Provided analysis of spreads, in particular time and location spreads, with emphasis on trading opportunities on the curve, also taking into account the availability of storage facilities.
-Became familiar with logistic implications of physical gas trading and developed knowledge of continental virtual hubs for gas exchange.}\\
-
-%------------------------------------------------
-% WORK EXPERIENCE 5
-%------------------------------------------------
-
-{\raggedleft\textsc{July 2008 -- Oct 2008}\par}
-
-{\raggedright\large 1\textsuperscript{st} year Analyst at \textsc{Lehman Brothers}, London\\
-\textit{Commodities Structured Trading}\\[5pt]}
-
-\normalsize{Developed spreadsheets for risk analysis on exotic derivatives on a wide array of commodities (ags, oils, precious and base metals), managed blotter and secondary trades on structured notes, liaised with Middle Office, Sales and Structuring for bookkeeping.}\\
+% \section{Non-English Language Skills}
+% \begin{tabular}{l p{0.7\textwidth}}
+%   Swedish & No problem with informal communications of talking,
+%   reading and writing. ``Swedish B'' passed.\\
+%   Chinese (Mandarin) & Native speaker \\
+% \end{tabular}
 
@@
 \begin{tabular}{c|p{7cm}}
-\raisebox{-4pt}{\textifsymbol{18}} & 123 Broadway, City, State 12345 \\ % Address
-\raisebox{-3pt}{\Mobilefone} & +1 (1)23 4567890 \\ % Phone number
-\raisebox{-1pt}{\Letter} & \href{mailto:john@smith.com}{john@smith.com} \\ % Email address
-\Keyboard & \href{http://www.johnsmith.com}{http://www.johnsmith.com} \\ % Website
+\raisebox{-1pt}{\Bowtie} & Swedish citizen \\
+\raisebox{-2pt}{\textifsymbol{18}} & Kiviksgatan 7C, 21440 Malm\"o, Sweden \\ % Address
+\raisebox{-2pt}{\Mobilefone} & +46 70 489 0472 \\ % Phone number
+\raisebox{-1pt}{\Letter} & \href{mailto:xie.xiaolei@gmail.com}{xie.xiaolei@gmail.com} \\ % Email address
+\Keyboard & \href{http://www.facebook.com/xie.xiaolei}{http://www.facebook.com/xie.xiaolei} \\ % Website
 \end{tabular}
-}
-}\\[10pt]
+}}\\[10pt]
 
 %----------------------------------------------------------------------------------------
-%	EDUCATION
+%	WORK EXPERIENCE
 %----------------------------------------------------------------------------------------
 
-\section{Education}
-
-\begin{tabular}{rl} % Start a table with two columns, one for dates and one for qualifications
+\section{Work Experience}
 
 %------------------------------------------------
-% EDUCATION 1
+% WORK EXPERIENCE 1
 %------------------------------------------------
 
-2011 -- \textsc{Present} & \textbf{Doctor of Philosophy} \\ 
-& \textsc{Computer Science} \\ 
-& \textit{The University of California}\\
-&\\
+{\raggedleft\textsc{Sep 2008 -- Jan 2011}\par}
+
+{\raggedright\large Development Engineer at \textsc{ST-Ericsson}, Lund\\
+  \textit{Software Development}\\[5pt]}
+
+\normalsize{Developed audio low-level software, typically device
+  drivers, in the Linux OS.}\\
 
 %------------------------------------------------
-% EDUCATION 2
+% WORK EXPERIENCE 2
 %------------------------------------------------
 
-2010 & \textbf{Master of Science} \\ 
-& \small First Class Honours \\
-& \textsc{Computer Science} \\ 
-& \textit{The University of California}\\
-&\\
+{\raggedleft\textsc{May 2007 -- Sep 2007}\par}
+
+{\raggedright\large Software Designer at \textsc{Goodmillsystems}, Helsinki\\
+  \textit{Software Development}\\[5pt]}
+
+\normalsize{Developed a daemon program that ran on the company's
+  smart router product for network traffic monitoring.}\\
 
 %------------------------------------------------
-% EDUCATION 3
+% WORK EXPERIENCE 3
 %------------------------------------------------
 
-2009 & \textbf{Postgraduate Diploma}\\
-& \small Distinction \\
-& \textsc{Computer Science} \\
-& \textit{The University of California}\\
-&\\
+{\raggedleft\textsc{Sep 2006 -- April 2007}\par}
+
+{\raggedright\large Software Developer at \textsc{Cardinal Information
+    Systems}, Helsinki\\
+  \textit{Software development}\\[5pt]}
+
+\normalsize{Implemented the FLUTE protocol for the company's DVB-H mobile-end
+  middleware.}\\
 
 %------------------------------------------------
-% EDUCATION 4
+% WORK EXPERIENCE 4
 %------------------------------------------------
+{\raggedleft\textsc{Feb 2006 - Sep 2006}\par}
 
-2006 -- 2008 & \textbf{Bachelor of Business Studies}\\
-& \textsc{Commerce} \\
-& \textit{The University of California}
+{\raggedright\large Game Engineer at \textsc{Digital Chocolate}, Helsinki\\
+  \textit{Java  mobile game development}\\[5pt]}
 
-%----------------------------------------------------------------------------------------
-
-\end{tabular}\\[10pt]
-
-%----------------------------------------------------------------------------------------
-%	AWARDS
-%----------------------------------------------------------------------------------------
-
-\section{Awards}
-
-\begin{tabular}{rl}
-2010	 & \textbf{Faculty of Science Masters Scholarship}\\
-& \textit{The University of California}\\ \\
+\normalsize{Ported and customized mobile phone games for various
+  handsets. Also developed some general tools for game customization.}\\
 
 %------------------------------------------------
+% WORK EXPERIENCE 5
+%------------------------------------------------
+{\raggedleft\textsc{April 2004 - June 2005}\par}
 
-2008	 & \textbf{Top Achiever Award -- Commerce}\\
-& \textit{Business Association of California}
-\end{tabular}\\[10pt]
+{\raggedright\large Software Engineer at \textsc{Holystone Ltd}, Beijing\\
+  \textit{Java Software development}\\[5pt]}
 
-%----------------------------------------------------------------------------------------
-%	COMPUTER SKILLS
-%----------------------------------------------------------------------------------------
+\normalsize{Developed JSP pages and database routines for the
+  company's information system.}\\
 
-\section{Computer Skills}
-
-\begin{tabular}{rl}
-Basic Knowledge
-& \textsc{java}, Adobe Illustrator\\ 
-& \\
-Intermediate Knowledge
-& \textsc{python}, \textsc{html}, \LaTeX,\\
-& OpenOffice, Linux,\\
-& Microsoft Windows \\
-& Computer Hardware \& \\
-& Support\\ \\
-\end{tabular}
-
-%----------------------------------------------------------------------------------------
-%	COMMUNICATION SKILLS
-%----------------------------------------------------------------------------------------
-
-\section{Communication Skills}
-
-\begin{tabular}{rl}
-\textsc{Conferences}
-& Oral Presentation at the California \\
-& Computer Science Conference -- 2010\\
-& \\
-\textsc{Posters}
-& Poster at the Annual Programming \\
-& Conference in Oregon -- 2009 \\
-\end{tabular}\\[10pt]
-
-%----------------------------------------------------------------------------------------
+\section{Programming Skills}
+\begin{itemize}
+\item Very experienced in C, Matlab
+\item Experienced in C++, Python, Java, SQL
+\item Very familiar with Linux
+\end{itemize}
 

### 2015-01-05 cv/SecondApplicationToKU.tex
Added CV and cover letters.

--- 2015-01-04 cv/ResearchCoverLetter.tex
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 
@@
 \begin{newlfm}
+I am currently a self-financing Ph.D student under the supervision of
+Prof. Thomas Mikosch. I understand that the department is providing
+free stipends to Ph.D students, so I would like to apply for a grant.
 
-%----------------------------------------------------------------------------------------
-%	LETTER CONTENT
-%----------------------------------------------------------------------------------------
+I started the Ph.D program here on 1st of October and have been taking
+the courses {\it Extreme Value Theory} given by Prof. Mikosch and {\it
+  Quantitative Risk Management} given by Prof. Collamore. At the same
+time I have done analysis on the eigenvalues of empirical covariance
+matrices of the S\&P500 and the DAX indices. A detailed plan of my
+Ph.D program is submitted along with this application.
 
-I understand from your department's website that you are offering Ph.D
-stipends, and found it a truly exciting opportunity. I have hence
-contacted Prof. Thomas Valentin Mikosch and asked for his
-supervision. After a kind discussion with me about my master's 
-thesis, his research area, and a very interesting Ph.D project in
-random matrix theory (RMT) and extreme value theory (EVT), he
-encouraged me to apply. The project that he outlined perfectly
-continued my master's studies and I really look forward to it. So I
-would like to present to you what I did as my master's thesis project
-and what my research interests are.
+I obtained my master's degree in mathematical physics from Lund
+University, with a thesis on econophysics supervised by Prof. Sven \AA
+berg. Based on the results of the thesis and with some continued work,
+Prof. \AA berg and I have written a paper together under the title
+``Effect of Heteroscedasticity on the spectrum of an Empirical
+Covariance Matrix'' and plan to submit it to {\it Quantitative
+  Finance} for publication. In the paper we argued that, on certain
+conditions, autocorrelations in the conditional volatilities could
+affect the empirical covariance matrixes' spectrum only through the
+unconditional distribution of the volatilities. On these
+conditions, we derived in the case of small autocorrelations the
+spectrum of an empirical covariance matrix. The derivations were based
+on results from free probability theory.
 
-I wrote my thesis under the supervision of Prof. Sven \AA berg at Lund
-university in the department of mathematical physics. I first compared
-GARCH and stochastic volatility (SV) models by case studying of
-intraday returns of a few Swedish stocks, and then I continued to
-study the covariance matrix of Gaussian and GARCH(1,1) returns. There
-I encountered the problem of eigenvalues distribution of a Wishart
-matrix and found its large and charming literature.
-
-By numerical simulation, I studied in the cases of Gaussian and
-GARCH(1,1) returns, how the matrix elements and eigenvalues were
-distributed, how the eigenvectors were composed, and in particular,
-how the situation was changed by autocorrelations in the return
-series. For example, I found in the case of GARCH(1,1) returns that
-prolonged return series increased the extent of the eigenvectors'
-localization while strengthened autocorrelations decreased the extent
-of localization.
-
-Among the many fascinating things that I was led to by the thesis
-project were the extreme value theory and the random matrix
-theory. In my studies of the intraday stock returns, heavy-tailed
-distributions, and in particular L\'evy distributions, are very
-important, so I read books such as ``Modeling Extremal Events'',
-``Handbook of Financial Time Series'', and papers about the tail
-behavior of GARCH and SV models, as well as those on the eigenvalues
-distribution of random matrices. Quite naturally, these books and
-papers led me to Prof. Mikosch. So when I met him and he outlined the
-Ph.D project on RMT and EVT, which followed directly from my master's
-project, I knew it was truly a great opportunity for me.
-
-If I am so honored to be accepted, I will, first of all, learn more
-about EVT in connection with multivariate time 
-series, and also about its application in the research of RMT. Then I
-would like to explore, for example, if and how the methods developed in the
-research of the largest eigenvalue distribution and the spectral
-density of random matrices, particularly Wishart matrices, can
-be applied, presumably with EVT, to find the eigenvalues distribution
-of a covariance matrix of heavy-tailed returns, such as those generated
-by GARCH or SV models.
-
-In particular, I understand that the problem of the exact
-distribution of the largest eigenvalue of a Wishart matrix has been
-attacked by the holonomic gradient method as an alternative to the
-more conventional approach of series expansion around the origin. It
-would be interesting to see if the same methods are applicable to the
-aforementioned problem.
-
-I look forward to an interview with you. If you would like to know more
-about me, you are most welcome to call me at +46 70 489 0472 or find
-me on LinkedIn as ``Xie Xiaolei''. If you want to contact my thesis
-supervisor, Prof. Sven \AA berg, he is reachable via telephone +46 46
-222 96 33 or mobile +46 70 557 35 69.
+Without a stipend, I am living on my own savings and borrowed money
+from my relatives. This admittedly causes hardships and
+distractions. So I really hope a stipend can be granted to me. It
+would be a great ease not only to me but also to my family. Thank you
+for your consideration.
 

### 2015-01-05 cv/XieXiaolei-Academic.tex
Added CV and cover letters.

--- 2015-01-04 cv/XieXiaolei-Computing.tex
@@
 \begin{minipage}[t]{0.5\textwidth}
-  Citizenship: Sweden \\
-  Phone: +46 70 489 0472 \\
+  Phone: +46 76 060 6638 \\
   Email: \href{mailto:xie.xiaolei@gmail.com}{\tt xie.xiaolei@gmail.com} \\
   Address: Kiviksgatan 7C, 21440 Malm\"{o}, Sweden
+  Citizenship: Sweden \\
 \end{minipage}
+
+% \section*{Objective}
+% A master of mathematical physics and Computer Science is looking for a
+% Ph.D position in the areas of stochastic processes, random matrix
+% theory and extreme value theory.
 
@@
 \begin{itemize}
-\item Master of Physics, Lund University, January 2011 - June 2014.
+\item Master of Physics, Lund University, Sweden, January 2011 - June
+  2014.
   \begin{itemize}
-  \item {\it Major Subject:} Econophysics. My thesis project was
-    case studying of intraday returns of Nordea, Volvo and
-    Ericsson. Covariance matrices of GARCH and stochastic volatility
-    returns were also studied. Matlab programming was a crucial part.
+  \item {\it Subject:} Mathematical Physics -- econophysics
+  \item My thesis project was a specialization in the area of economical
+    physics and focused on the influence that auto-correlations in 
+    return series had on the spectral distribution of an empirical
+    covariance matrix. It was titled ``Return Models and Covariance
+    Matrices'' and was supervised by Prof. Sven \AA berg.
   \end{itemize}
-\item M.Sc.(Tech) Computer Science and Engineering, Helsinki
-  University of Technology, 2005 - 2008.
+
+\item M.Sc.(Tech) Computer Science and Engineering, Aalto University,
+  Finland, 2005 - 2008.
   \begin{itemize}
@@
   \item {\it Average grade:} 4.17/5.
+  \item   Funded by HIIT, I wrote my thesis ``A Mutual authentication
+    and Key Agreement Protocol for the UMTS Network''. The algorithm of my
+    protocol proposal was based on the Diffie-Hellman Exchange, but had
+    significant modifications and extensions so that the mutual
+    authentication between the mobile user and the network was delegated
+    to the serving instead of the home network. The protocol design
+    incorporated puzzle schemes that demanded unbalanced computing
+    resources on the two sides to mitigate Denial-of-Service attacks. I
+    made C programs to test the puzzle schemes.
   \end{itemize}
+
 \item B.Sc. Computer Science \& Technology, Beijing Institute of
@@
 \section*{Employment}
-\begin{aperiod}
-\item {\it \href{http://www.stericsson.com}{\tt
-      ST-Ericsson}, Audio, Multimedia, Development Engineer, Sep 2008
-    - Jan 2011}
+\begin{itemize}
+\item Research Assistant, Lund University, June 10, 2014 - September
+  30, 2014
 
-  My responsibilities included developing audio low-level software,
-  typically device drivers, in the Linux OS. This involved designing and
-  implementing data structures and procedures so that the hardware
-  devices were used in the optimal way, in terms of parallelism,
-  power consumption, robustness, and ease of maintenance.
+  Based on the results from my thesis work, Prof. \AA berg and I wrote
+  a paper together. Titled ``Effect of Heteroscedasticity on the
+  spectrum of an Empirical Covariance Matrix'', we argued that, in
+  certain circumstances, autocorrelations in the conditional
+  volatilities were manifest in the empirical covariance matrixes'
+  spectrum only through the unconditional distribution of the
+  volatilities. In these circumstances, we derived in the case of small
+  autocorrelations the spectrum of an empirical covariance matrix. The
+  derivations were based on results from free probability theory.
 
-\item {\it \href{http://www.hiit.fi}{\tt Helsinki
-      Institute of Information Technology}, Thesis Worker, Oct 2007 -
-    May 2008}
+\item Software Engineer, April 2004 - January 2011, Helsinki
 
-  Funded by HIIT, I wrote my thesis ``A Mutual authentication and Key
-  Agreement Protocol for the UMTS Network''. The algorithm of my
-  protocol proposal was based on the Diffie-Hellman Exchange, but had
-  significant modifications and extensions so that the mutual
-  authentication between the mobile user and the network was delegated
-  to the serving instead of the home network. The protocol design
-  incorporated puzzle schemes that demanded unbalanced computing
-  resources on the two sides to mitigate Denial-of-Service attacks. I
-  made C programs to test the puzzle schemes.
+I worked in the IT industry as a software engineer. I designed and
+implemented software solutions in C, C++ and Java.
 
-\item {\it
-    \href{http://www.goodmillsystems.com}{\tt Goodmillsystems},
-    Software Designer, May 2007 - Sep 2007}
+\end{itemize}
 
-  I developed a daemon program that ran on the company's smart router
-  product to collect statistical data about the router's network
-  traffic. The emphasis of the program was high performance in the
-  context of competing read and write accesses.
+\section*{Skills}
+\begin{itemize}
+\item {\bf Programming:} Proficient in C, C++, Java, Matlab
+\item {\bf Scientific writing:} Latex
+\end{itemize}
 
-\item {\it \href{http://www.cardinal.fi/}{\tt
-      Cardinal Information Systems}, Software Developer, Sep 2006 -
-    April 2007}
+\section*{Association}
+I am a professional member of the {\it Association of Computing Machinery}.
 
-  I implemented the FLUTE protocol for the company's DVB-H
-  mobile-end middleware. The implementation was multi-threaded and was
-  built on the {\it pthread} library. Good performance and robustness
-  were crucial requirements on the implementation.
+\section*{Referees}
+\begin{itemize}
+\item Prof. Sven \AA berg. Email:
+  \href{mailto:sven.aberg@matfys.lth.se}{\tt
+    sven.aberg@matfys.lth.se}. Telephone: +46 46 222 96 33. Mobile:
+  +46 70 557 35 69.
 
-\item {\it
-    \href{http://www.digitalchocolate.com/}{\tt Digital Chocolate},
-    Game Engineer, Feb 2006 - Sep 2006}
-
-  At this job, my responsibilities were porting and customizing mobile
-  phone games for various handsets, and making game demos.
-
-\item {\it {\tt Beijing Holystone Ltd}, Software
-    Engineer, April 2004 - June 2005}
-
-  At this job I mainly developed JSP pages and database routines for
-  the company's information system.
-\end{aperiod}
-
-\section*{Personal Project}
-{\it Trades data collector \& Analyzer}
-
-I have been developing my own stock-trading robot since the summer of
-2011. It is a software built on cURL, GLib, MySQL, and Matlab. I find
-it a fairly interesting, challenging, and rewarding combination of
-programming skills, data analysis techniques, and knowledge of
-financial time series. My knowledge in econophysics guides the
-development of this software.
-
-\section*{Non-English Language Skills}
-\begin{itemize}
-\item Swedish: informal communications. ``Swedish B'' passed.
-\item Chinese: Native speaker
+\item Prof. Ferdi
+  Aryasetiawan. Email: \href{ferdi.aryasetiawan@teorfys.lu.se}{\tt
+    ferdi.aryasetiawan@teorfys.lu.se}. Telephone: +46 46 222 90 89.
 \end{itemize}

### 2015-01-06 cv/XieXiaolei-Academic.tex
Worked on EVT final test.

@@
 \begin{minipage}[t]{0.5\textwidth}
-  Phone: +46 76 060 6638 \\
+  Phone: +46 70 489 0472 \\
   Email: \href{mailto:xie.xiaolei@gmail.com}{\tt xie.xiaolei@gmail.com} \\
@@
 \section*{Employment}
-\begin{itemize}
-\item Research Assistant, Lund University, June 10, 2014 - September
-  30, 2014
-
-  Based on the results from my thesis work, Prof. \AA berg and I wrote
-  a paper together. Titled ``Effect of Heteroscedasticity on the
-  spectrum of an Empirical Covariance Matrix'', we argued that, in
-  certain circumstances, autocorrelations in the conditional
-  volatilities were manifest in the empirical covariance matrixes'
-  spectrum only through the unconditional distribution of the
-  volatilities. In these circumstances, we derived in the case of small
-  autocorrelations the spectrum of an empirical covariance matrix. The
-  derivations were based on results from free probability theory.
-
-\item Software Engineer, April 2004 - January 2011, Helsinki
-
-I worked in the IT industry as a software engineer. I designed and
-implemented software solutions in C, C++ and Java.
-
-\end{itemize}
+Between April 2004 and January 2011 I worked in the software industry
+as a software engineer. I designed and implemented software solutions
+in C, C++ and Java.
 
@@
 \begin{itemize}
-\item {\bf Programming:} Proficient in C, C++, Java, Matlab
+\item {\bf Programming:} Proficient in C/C++, Java, Matlab
 \item {\bf Scientific writing:} Latex

### 2015-01-06 EVT/MidTerm.tex

%% Created by Maple 17.00, Linux
%% Source Worksheet: MidTerm.mw
%% Generated: Mon Jan 05 22:30:46 CET 2015
\documentclass{article}
\usepackage{maplestd2e}
\def\emptyline{\vspace{12pt}}
\begin{document}
\pagestyle{empty}
\DefineParaStyle{Maple Heading 1}
\DefineParaStyle{Maple Text Output}
\DefineParaStyle{Maple Dash Item}
\DefineParaStyle{Maple Bullet Item}
\DefineParaStyle{Maple Normal}
\DefineParaStyle{Maple Heading 4}
\DefineParaStyle{Maple Heading 3}
\DefineParaStyle{Maple Heading 2}
\DefineParaStyle{Maple Warning}
\DefineParaStyle{Maple Title}
\DefineParaStyle{Maple Error}
\DefineCharStyle{Maple Hyperlink}
\DefineCharStyle{Maple 2D Math}
\DefineCharStyle{Maple Maple Input}
\DefineCharStyle{Maple 2D Output}
\DefineCharStyle{Maple 2D Input}
\begin{center}
\begin{Maple Normal}{
\begin{center}
\begin{Maple Normal}{
\textbf{Mid Term Test}}\end{Maple Normal}
\end{center}
}\end{Maple Normal}
\end{center}
\begin{center}
\begin{Maple Normal}{
\begin{center}
\begin{Maple Normal}{
\textbf{Xie Xiaolei}}\end{Maple Normal}
\end{center}
}\end{Maple Normal}
\end{center}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(a) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
For sufficiency, assume\mapleinline{inert}{2d}{t <= F(x)}{$\displaystyle t\leq F \left( x \right) $}
. Because}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("inf", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("s", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&in;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("R", italic = "true", mathvariant = "italic"), Typesetting:-mo(":", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("s", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&GreaterEqual;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{\[\displaystyle F ^{\eslarr }\left(x \right)=\inf \left(s \mathop{\rm  }\in R :\mathop{\rm  }F \left(s \right)\ge x \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
by definition, it must be so that\mapleinline{inert}{2d}{x >= `#msup(mi("F"),mo("&larr;"))`(t)}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( t \right) \leq x$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
For necessity, assume\mapleinline{inert}{2d}{`#msup(mi("F"),mo("&larr;"))`(t) <= x}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( t \right) \leq x$}
. If\mapleinline{inert}{2d}{F(x) < t}{$\displaystyle F \left( x \right) <t$}
, then by the definition of\mapleinline{inert}{2d}{`#msup(mi("F"),mo("&larr;"))`(t)}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( t \right) $}
,\mapleinline{inert}{2d}{x < `#msup(mi("F"),mo("&larr;"))`(t)}{$\displaystyle x<\mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( t \right) $}
. This contradicts out assumption. Therefore\mapleinline{inert}{2d}{F(x) >= t}{$\displaystyle t\leq F \left( x \right) $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(b)}Let\mapleinline{inert}{2d}{F(x)}{$\displaystyle F \left( x \right) $}
consists of point mass\mapleinline{inert}{2d}{p}{$\displaystyle p$}
at\mapleinline{inert}{2d}{X = 0}{$\displaystyle X=0$}
and\mapleinline{inert}{2d}{1-p}{$\displaystyle 1-p$}
at\mapleinline{inert}{2d}{X = 1}{$\displaystyle X=1$}
. Then\mapleinline{inert}{2d}{F(x) = p*`I__0&le; x<1__`+`I__x&ge;1`}{$\displaystyle F \left( x \right) =p{\it I\_\_0}\le \mbox {{\tt ` x$<$1\_\_`}}\\
\mbox{}+{\it I\_\_x}\ge \mbox {{\tt `1`}}$}
. Thus\mapleinline{inert}{2d}{`in`(F(X), {0, 1, p})}{$\displaystyle {\it in} \left( F \left( X \right) , \left\{ 0,1,p \right\}  \right) $}
and is not uniformly\mapleinline{inert}{2d}{U(0, 1)}{$\displaystyle U \left( 0,1 \right) $}
distributed.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(c) (i) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Provided that the distribution function F is continuous,\mapleinline{inert}{2d}{P(X[s] = X[t]) = 0}{$\displaystyle P \left( X_{{s}}=X_{{t}} \right) =0$}
for\mapleinline{inert}{2d}{s <> t}{$\displaystyle s\neq t$}
. Then for some\mapleinline{inert}{2d}{k}{$\displaystyle k$}
,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(t = tau[k]) = P(X[t] = M[t]) and P(X[t] = M[t]) = 1/t}{$\displaystyle P \left( t=\tau_{{k}} \right) =P \left( X_{{t}}=M_{{t}} \right) {\rm \: \wedge \:}P \left( X_{{t}}=M_{{t}} \right) ={t}^{-1}$}
\mapleinline{inert}{2d}{}{$\displaystyle $}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore the distribution of the record times are independent of F.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(c).(ii)}Because\mapleinline{inert}{2d}{F(x)}{$\displaystyle F \left( x \right) $}
is continuous,\mapleinline{inert}{2d}{`#msup(mi("F"),mo("&larr;"))`(x)}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( x \right) $}
is strictly increasing. Thus one can use\mapleinline{inert}{2d}{`#msup(mi("F"),mo("&larr;"))`(x)}{$\displaystyle \mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( x \right) $}
as a transformation to convert the record sequences.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(d)}Because F(x) is continuous on\mapleinline{inert}{2d}{RealRange(1, Open(infinity))}{$\displaystyle {\it RealRange} \left( 1,{\it Open} \left( \infty  \right)  \right) $}
, X can be simulated by a quantile transform from a uniformly U(0,1) distributed variable U: Let\mapleinline{inert}{2d}{U = F(X)}{$\displaystyle U=F \left( X \right) $}
. Then}\end{Maple Normal}

\begin{Maple Normal}{
\mapleinline{inert}{2d}{X = (1-U)^(-1/alpha)}{$\displaystyle X= \left( 1-U \right) ^{-{\alpha}^{-1}}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{1.(d)}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(X[n+1] > max(X[1], () .. (), X[n]))}{\[\displaystyle P \left( \max \left( X_{{1}},{\ldots },X_{{n}} \right) <X_{{n+1}} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("max", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mo("..", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.0em"), Typesetting:-mo(".", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-msub(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal")), mathvariant = "normal"))}{\[\displaystyle =P \left(X _{n +1}=\max \left(X _{1},...,X _{n +1}\right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{X[i]}{$\displaystyle X_{{i}}$}
are iid,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(X[n+1] = max(X[1], () .. (), X[n+1]))}{\[\displaystyle P \left( X_{{n+1}}=\max \left( X_{{1}},{\ldots },X_{{n+1}} \right)  \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))}{$\displaystyle =\mathop{\rm  }\frac{1}{n +1}$}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(a) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Consider}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(L(a*x)/L(x), x = infinity)}{\[\displaystyle \lim _{x\rightarrow \infty }{\frac {L \left( ax \right) }{L \left( x \right) }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accentunder = "false"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("`c__0`"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("`c__0`"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle =\mathop{\rm  }\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\frac{{\it `c__0` }\left(a \mathop{\rm  }x \right)}{{\it `c__0` }\left(x \right)}\exp \left(\Mapleoverset{a \mathop{\rm  }x }{\Mapleunderset{x }{\int }}\frac{\epsilon \left(t \right)}{t }\mathop{\rm d}t \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \frac{\epsilon \left(x \right)}{x }\essrarr 0$}
as\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&approx;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{\[\displaystyle \Mapleoverset{a \mathop{\rm  }x }{\Mapleunderset{x }{\int }}\frac{\epsilon \left(t \right)}{t }\mathop{\rm d}t \mathop{\rm  }\esapprox \frac{\epsilon \left(x \right)}{x }\left(a -1\right)x =\left(a -1\right)\epsilon \left(x \right)\essrarr 0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(L(a*x)/L(x), x = infinity)}{$\displaystyle \lim _{x\rightarrow \infty }{\frac {L \left( ax \right) }{L \left( x \right) }}$}
=\mapleinline{inert}{2d}{limit(c__0(a*x)*exp*{int(`&epsilon;`(t)/t, t = x .. a*x)}/c__0(x), x = infinity) = 1}{$\displaystyle \lim _{x\rightarrow \infty } \left( {\it c\_\_0} \left( ax \right)  \left( \exp \right) \, \left\{ \int _{x}^{ax}\!{\frac {\epsilon  \left( t \right) }{t}}{dt} \right\}  \left( {\it c\_\_0} \left( x \right) \\
\mbox{} \right) ^{-1} \right) =1$}
}\end{Maple Normal}

}\end{Maple Normal}

Thus, by definition,\mapleinline{inert}{2d}{L(x)}{$\displaystyle L \left( x \right) $}
is a slowing varying function.

\textbf{}\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(b) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
For\mapleinline{inert}{2d}{x > a}{$\displaystyle a<x$}
where\mapleinline{inert}{2d}{(log*log)*log*a > 0}{$\displaystyle 0<{{\it log}}^{3}a$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{[(log*log)*log(x)]^2 = exp[((2*log*log)*log*log)*x]}{$\displaystyle {[{{\it log}}^{2}{\it log} \left( x \right) ]}^{2}=\exp_{{2\,{{\it log}}^{4}x}}$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{diff([((log*log)*log*log)*x], x) = 1/(x*((ln(10))((log*log)*log*x))((log*log)*x)*log*x)}{\[\displaystyle {\frac {\partial }{\partial x}} \left( [{{\it log}}^{4}x] \right) ={\frac {1}{{x}^{2}\ln  \left( 10 \right)  \left( {{\it log}}^{3}x \right)  \left( {{\it log}}^{2}x \right)  \left( {\it log} \right) \\
\mbox{}}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore we can write}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{L(x) = c*exp[int(1/(x*((ln(10))((log*log)*log*x))((log*log)*x)*log*x), x = 10^4 .. x)]}{$\displaystyle L \left( x \right) =c\exp_{{\int _{10000}^{x}\!{\frac {1}{{x}^{2}\ln  \left( 10 \right)  \left( {{\it log}}^{3}x \right)  \left( {{\it log}}^{2}x \right) \\
\mbox{} \left( {\it log} \right) }}{dx}}}$}
for\mapleinline{inert}{2d}{x >= 10^4}{$\displaystyle 10000\leq x$}
and some constant c. As\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("10", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{\[\displaystyle \frac{1}{\mathop{\rm ln}\left(10\right)\left(\mathop{\rm log}\mathop{\rm  }\mathop{\rm log}\mathop{\rm  }\mathop{\rm log}\mathop{\rm  }x \right)\left(\mathop{\rm log}\mathop{\rm  }\mathop{\rm log}\mathop{\rm  }x \right)\mathop{\rm  }\mathop{\rm log}\mathop{\rm  }x }\essrarr 0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Thus\mapleinline{inert}{2d}{L(x)}{$\displaystyle L \left( x \right) $}
is a slowing varying function.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
For\mapleinline{inert}{2d}{x >= a}{$\displaystyle a\leq x$}
, where\mapleinline{inert}{2d}{a > 1}{$\displaystyle 1<a$}
is a constant,\mapleinline{inert}{2d}{L__0(x)}{$\displaystyle {\it L\_\_0} \left( x \right) $}
can be re-written as}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{L__0(x) = exp(sqrt(log*a))*exp*{int(1/(2*t*sqrt(log*t)), t = a .. x)}}{\[\displaystyle {\it L\_\_0} \left( x \right) ={{\rm e}^{ \sqrt{ \left( {\it log} \right) \,a}}} \left( \exp \right) \, \left\{ \int _{a}^{x}\!1/2\,{\frac {1}{t \sqrt{ \left( {\it log} \right) \,t}}}{dt} \right\} \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Clearly}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(1/(2*sqrt(x)), x = infinity) = 0}{\[\displaystyle \lim _{x\rightarrow \infty }1/2\, \left(  \sqrt{x} \right) ^{-1}=0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore\mapleinline{inert}{2d}{L__0(x)}{$\displaystyle {\it L\_\_0} \left( x \right) $}
is a slowing varying function.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2(c). Solution:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{log(x) = exp(ln*log*x)}{\[\displaystyle {\it log} \left( x \right) ={{\rm e}^{ \left( {\it ln} \right) \, \left( {\it log} \right) \,x}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("`x__0`")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("10", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), mathvariant = "normal"))}{\[\displaystyle =\mathop{\rm  }\exp \left(\Mapleoverset{x }{\Mapleunderset{{\it `x__0` }}{\int }}\frac{1}{\mathop{\rm ln}\left(10\right)t \mathop{\rm  }\mathop{\rm  }\mathop{\rm log}\left(t \right)}\mathop{\rm d}t \mathop{\rm  }+\mathop{\rm ln}\mathop{\rm  }\mathop{\rm log}\left(x _{0}\right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("`x__0`")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("10", mathvariant = "normal")), mathvariant = "normal"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("log", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), mathvariant = "normal")), Typesetting:-mi(""))}{\[\displaystyle =\mathop{\rm  }\mathop{\rm log}\left(x _{0}\right)\mathop{\rm  }\exp \left(\Mapleoverset{x }{\Mapleunderset{{\it `x__0` }}{\int }}\frac{1}{\mathop{\rm ln}\left(10\right)t \mathop{\rm  }\mathop{\rm  }\mathop{\rm log}\left(t \right)}\mathop{\rm d}t \mathop{\rm  }\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{1 < x[0] and x[0] <= x}{$\displaystyle 1<x_{{0}}{\rm \: \wedge \:}x_{{0}}\leq x$}
. We identify\mapleinline{inert}{2d}{`&epsilon;`(x) = 1/[ln(10)*log(x)]}{$\displaystyle \epsilon  \left( x \right) ={[\ln  \left( 10 \right) {\it log} \left( x \right) ]}^{-1}$}
. Clearly\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("&epsilon;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \epsilon \left(x \right)\mathop{\rm  }\essrarr 0$}
as\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
.\mapleinline{inert}{2d}{}{$\displaystyle $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(d) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("E", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle E \left(X ^{\delta }\right)=-\Mapleoverset{\infty }{\Mapleunderset{0}{\int }}x ^{\delta }\mathop{\rm d}\left(\Mapleoverset{{\it \_} }{F }\left(x \right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accentunder = "false"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic"))}{\[\displaystyle =\mathop{\rm  }\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}x ^{\delta }\Mapleoverset{{\it \_} }{F }\left(x \right)+\delta \mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{0}{\int }}x ^{\delta -\alpha -1}L \left(x \right)\mathop{\rm d}x \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{$\displaystyle \Mapleoverset{{\it \_} }{F }\left(x \right)$}
is regularly varying with index\mapleinline{inert}{2d}{-alpha}{$\displaystyle -\alpha$}
, it has representation}\end{Maple Normal}

}\end{Maple Normal}

\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle \Mapleoverset{{\it \_} }{F }\left(x \right)=c \left(x \right)\mathop{\rm  }\exp \left(\Mapleoverset{x }{\Mapleunderset{x _{0}}{\int }}\frac{\xi \left(t \right)}{t }\mathop{\rm d}t \right)\]}
for some\mapleinline{inert}{2d}{x[0] > 0}{$\displaystyle 0<x_{{0}}$}
,\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"))}{$\displaystyle c \left(x \right)\essrarr c _{0}$}
,\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"))}{$\displaystyle \xi \left(x \right)\essrarr -\alpha $}
as\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
. Therefore

\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle x ^{\delta }\mathop{\rm  }\Mapleoverset{{\it \_} }{F }\left(x \right)=c \left(x \right)\mathop{\rm  }\exp \left(\Mapleoverset{x }{\Mapleunderset{x _{0}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t +\delta \mathop{\rm  }\mathop{\rm ln}\left(x _{0}\right)\right)\]}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&int;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mo(" ", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo(" ", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("ln", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]")), Typesetting:-mi(""))}{\[\displaystyle =\mathop{\rm  }c \left(x \right)\mathop{\rm  }\exp \left(\Mapleoverset{x _{1}}{\Mapleunderset{x _{0}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t +\Mapleoverset{x }{\Mapleunderset{x _{1}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t \mathop{\rm  }+\delta \mathop{\rm  }\mathop{\rm ln}\left(x _{0}\right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \frac{\xi \left(t \right)+\delta }{t }\essrarr 0$}
as\mapleinline{inert}{2d}{proc (t) options operator, arrow; infinity end proc}{$\displaystyle t\mapsto \infty $}
, we can choose\mapleinline{inert}{2d}{x[1]}{$\displaystyle x_{{1}}$}
so large that}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`&approx;`(int((xi(t)+delta)/t, t = x[1] .. x), (delta-alpha)*(x-x[1])/x[1])}{\[\displaystyle \mbox {{\tt `\&approx;`}} \left( \int _{x_{{1}}}^{x}\!{\frac {\xi \left( t \right) +\delta}{t}}{dt},{\frac { \left( \delta-\alpha \right) \\
\mbox{} \left( x-x_{{1}} \right) }{x_{{1}}}} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{If}\mapleinline{inert}{2d}{delta < alpha}{$\displaystyle \delta<\alpha$}
,\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mo(" ", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&srarr;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&uminus0;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), Typesetting:-mi(""))}{$\displaystyle \Mapleoverset{x }{\Mapleunderset{x _{1}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t \mathop{\rm  }\essrarr -\infty $}
and\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle x ^{\delta }\mathop{\rm  }\Mapleoverset{{\it \_} }{F }\left(x \right)\essrarr 0$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Moreover, if\mapleinline{inert}{2d}{delta < alpha}{$\displaystyle \delta<\alpha$}
, i.e.\mapleinline{inert}{2d}{delta-alpha-1 < -1}{$\displaystyle \delta-\alpha<0$}
, by Karamata's theorem,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta -\alpha -1}L \left(x \right)\mathop{\rm  }\mathop{\rm d}x \estilde \frac{1}{\alpha -\delta }z ^{\delta -\alpha }L \left(z \right)\estilde \frac{1}{\alpha -\delta }z ^{\delta }\mathop{\rm  }\Mapleoverset{{\it \_} }{F }\left(z \right)\essrarr 0$}
as\mapleinline{inert}{2d}{proc (z) options operator, arrow; infinity end proc}{$\displaystyle z\mapsto \infty $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore\mapleinline{inert}{2d}{int(x^(delta-alpha-1)*L(x), x = 0 .. infinity) = 0 and 0 < infinity}{$\displaystyle \int _{0}^{\infty }\!{x}^{\delta-\alpha-1}L \left( x \right) {dx}=0{\rm \: \wedge \:}0<\infty $}
. So\mapleinline{inert}{2d}{E*X^delta < infinity}{$\displaystyle E{X}^{\delta}<\infty $}
when\mapleinline{inert}{2d}{delta < alpha}{$\displaystyle \delta<\alpha$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{if}\mapleinline{inert}{2d}{delta > alpha}{$\displaystyle \alpha<\delta$}
,\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mn("1", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), subscriptshift = "0")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), executable = "true", font_style_name = "2D Input", mathvariant = "normal"), Typesetting:-mo("+", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", executable = "true", font_style_name = "2D Input", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("d", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"), Typesetting:-mo(" ", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&srarr;", executable = "true", font_style_name = "2D Input", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("&infin;", italic = "true", executable = "true", font_style_name = "2D Input", mathvariant = "italic"))}{$\displaystyle \Mapleoverset{x }{\Mapleunderset{x _{1}}{\int }}\frac{\xi \left(t \right)+\delta }{t }\mathop{\rm d}t \mathop{\rm  }\essrarr \infty $}
and\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("_", italic = "true", mathvariant = "italic")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&infin;", italic = "true", mathvariant = "italic", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo(".", italic = "true", mathvariant = "italic", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"))}{\[\displaystyle x ^{\delta }\mathop{\rm  }\Mapleoverset{{\it \_} }{F }\left(x \right)\essrarr \infty .\]}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
and}\end{Maple Normal}

}\end{Maple Normal}

\textbf{}\mapleinline{inert}{2d}{int(x^(delta-alpha-1)*L(x), x = 0 .. infinity) = limit(int(t^(delta-alpha-1)*L(t), t = 0 .. x), x = infinity)}{\[\displaystyle \int _{0}^{\infty }\!{x}^{\delta-\alpha-1}L \left( x \right) {dx}=\lim _{x\rightarrow \infty } \left( \int _{0}^{x}\!{t}^{\delta-\alpha-1}\\
\mbox{}L \left( t \right) {dt} \right) \]}
\begin{Maple Normal}{
\begin{Maple Normal}{
By Karamata's theorem, as\mapleinline{inert}{2d}{proc (x) options operator, arrow; infinity end proc}{$\displaystyle x\mapsto \infty $}
,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("d", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic"))}{$\displaystyle \Mapleoverset{x }{\Mapleunderset{0}{\int }}t ^{\delta -\alpha -1}L \left(t \right)\mathop{\rm d}t \estilde \frac{1}{\delta -\alpha }x ^{\delta -\alpha }L \left(x \right)\essrarr \infty $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Thus\mapleinline{inert}{2d}{E*X^delta = infinity}{$\displaystyle E{X}^{\delta}=\infty $}
when\mapleinline{inert}{2d}{delta > alpha}{$\displaystyle \alpha<\delta$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{For an example of}\mapleinline{inert}{2d}{E*X^alpha = infinity}{$\displaystyle E{X}^{\alpha}=\infty $}
, consider\mapleinline{inert}{2d}{F(x) = 1-1/x}{$\displaystyle F \left( x \right) =1-{x}^{-1}$}
for\mapleinline{inert}{2d}{`in`(x, RealRange(1, Open(infinity)))}{$\displaystyle {\it in} \left( x,{\it RealRange} \left( 1,{\it Open} \left( \infty \\
\mbox{} \right)  \right)  \right) $}
, where\mapleinline{inert}{2d}{x[0] > 0}{$\displaystyle 0<x_{{0}}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\mapleinline{inert}{2d}{E*X = int(1/x, x = 1 .. infinity) and int(1/x, x = 1 .. infinity) = infinity}{\[\displaystyle EX=\int _{1}^{\infty }\!{x}^{-1}{dx}{\rm \: \wedge \:}\int _{1}^{\infty }\!{x}^{-1}{dx}=\infty \]}
\textbf{For an example of}\mapleinline{inert}{2d}{E*X^alpha < infinity}{$\displaystyle E{X}^{\alpha}<\infty $}
, consider\mapleinline{inert}{2d}{F(x) = 1-(a*ln^2*a)/(x*ln^2*x)}{$\displaystyle F \left( x \right) =1-{\frac {{a}^{2}}{{x}^{2}}}$}
for\mapleinline{inert}{2d}{`in`(x, RealRange(a, Open(infinity)))}{$\displaystyle {\it in} \left( x,{\it RealRange} \left( a,{\it Open} \left( \infty \\
\mbox{} \right)  \right)  \right) $}
,\mapleinline{inert}{2d}{a > 0}{$\displaystyle 0<a$}
, which is regularly varying with index -1:

\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(t*x)/`#mover(mi("F"),mo("&uminus0;"))`(x) = [x*ln^2*x]/[t*x*(ln*t+ln*x)^2] and [x*ln^2*x]/[t*x*(ln*t+ln*x)^2] = 1/t}{\[\displaystyle {\frac {\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( tx \right) }{\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) }}={\frac {[{x}^{2}{{\it ln}}^{2}]}{[tx \left(  \left( {\it ln} \right) \,t+ \left( {\it ln} \right) \,x \right) ^{2}]}}{\rm \: \wedge \:}{\frac {[{x}^{2}{{\it ln}}^{2}]}{[tx \left(  \left( {\it ln} \right) \,t+ \left( {\it ln} \right) \,x \right) ^{2}]}}={t}^{-1\\
\mbox{}}\]}
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("E", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("a", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dx", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("2", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dx", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi(""))}{\[\displaystyle E \mathop{\rm  }X =\Mapleoverset{\infty }{\Mapleunderset{a }{\int }}\frac{a \mathop{\rm  }\mathop{\rm ln}^{2}a }{x \mathop{\rm  }\mathop{\rm ln}^{2}x \mathop{\rm  }}{\it dx} =a \mathop{\rm  }\mathop{\rm ln}^{2}a \mathop{\rm  }\left(-\frac{1}{\mathop{\rm ln}\mathop{\rm  }a }+2\mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{a }{\int }}\frac{1}{x \mathop{\rm  }\mathop{\rm ln}^{2}x }{\it dx} \right)=a \mathop{\rm  }\mathop{\rm ln}\left(a \right)<\infty \]}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(e) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(`#mover(mi("\`c__n\`"),mo("&tilde;"))`) = `#mover(mi("F"),mo("&uminus0;"))`(`#mover(mi("\`c__n\`"),mo("&tilde;"))`/c__n*c__n)}{\[\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( \mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \right) =\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( \mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}$}
is regularly varying, by theorem A3.2,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(`#mover(mi("\`c__n\`"),mo("&tilde;"))`/c__n*c__n) = (`#mover(mi("\`c__n\`"),mo("&tilde;"))`/c__n)^(-alpha)*`#mover(mi("F"),mo("&uminus0;"))`(c__n)}{\[\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( \mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \right) = \left( {\frac {\mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{}}{{\it c\_\_n}}} \right) ^{-\alpha}\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it c\_\_n} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
But\mapleinline{inert}{2d}{`~`[`#mover(mi("F"),mo("&uminus0;"))`(c__n)]/n}{$\displaystyle {\frac {\mbox {{\tt `\~{}`}}_{{\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it c\_\_n} \right) }}}{n}}$}
and\mapleinline{inert}{2d}{`~`[`#mover(mi("F"),mo("&uminus0;"))`(`#mover(mi("\`c__n\`"),mo("&tilde;"))`)]/n}{$\displaystyle {\frac {\mbox {{\tt `\~{}`}}_{{\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( \mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \right) }}}{n}}$}
, therefore\mapleinline{inert}{2d}{`~`[c__n]*`#mover(mi("\`c__n\`"),mo("&tilde;"))`}{$\displaystyle \mbox {{\tt `\~{}`}}_{{{\it c\_\_n}}}\mbox {{\tt `\#mover(mi("`c\_\_n`"),mo("`}}\mbox {{\tt `\~{}`}}\\
\mbox{}\mbox {{\tt `"))`}}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(f) (i) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(x__F-1/x) = x^(-alpha)*L(x)}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it x\_\_F}-{x}^{-1} \right) ={x}^{-\alpha}L \left( x \right) $}
implies}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(X > x__F-1/x) = x^(-alpha)*L(x)}{\[\displaystyle P \left( {\it x\_\_F}-{x}^{-1}<X \right) ={x}^{-\alpha}L \left( x \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\mapleinline{inert}{2d}{P(1/x > x__F-X) = x^(-alpha)*L(x)}{\[\displaystyle P \left( {\it x\_\_F}-X<{x}^{-1} \right) ={x}^{-\alpha}L \left( x \right) \]}
\mapleinline{inert}{2d}{P(x < 1/(x__F-X)) = x^(-alpha)*L(x)}{\[\displaystyle P \left( x< \left( {\it x\_\_F}-X \right) ^{-1} \right) ={x}^{-\alpha}L \left( x \right) \]}
\mapleinline{inert}{2d}{`#mover(mi("G"),mo("&uminus0;"))`(x) = x^(-alpha)*L(x)}{$\displaystyle \mbox {{\tt `\#mover(mi("G"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) ={x}^{-\alpha}L \left( x \right) $}
\mapleinline{inert}{2d}{}{$\displaystyle $}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{2.(f) (ii)}\textbf{Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{1/(x__F-X)}{$\displaystyle  \left( {\it x\_\_F}-X \right) ^{-1}$}
has distribution function\mapleinline{inert}{2d}{G}{$\displaystyle G$}
with regularly varying tail as proven in (i),\mapleinline{inert}{2d}{`in`(1/(x__F-X), MDA(`&Phi;__&alpha;`))}{$\displaystyle {\it in} \left(  \left( {\it x\_\_F}-X \right) ^{-1},{\it MDA} \left( \Phi {\it \_\_}\\
\mbox{}\alpha  \right)  \right) $}
, which implies}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("`M__n`")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-msup(Typesetting:-mi("G", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal"), Typesetting:-mi("y", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("y", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), superscriptshift = "0"))}{\[\displaystyle P \left(\frac{1}{{\it `x__F` }-{\it `M__n` }}<G ^{\eslarr }\left(1-\frac{1}{n }\right)y \right)\essrarr e ^{-y ^{-\alpha }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{`#msup(mi("G"),mo("&larr;"))`(1-1/n) = 1/(x__F-`#msup(mi("F"),mo("&larr;"))`(1-1/n))}{$\displaystyle \mbox {{\tt `\#msup(mi("G"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) = \left( {\it x\_\_F}-\mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right)  \right) ^{-1}$}
, we write}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal")), Typesetting:-mi("")), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("`M__n`")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("y", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("y", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), superscriptshift = "0")), Typesetting:-mi(""))}{\[\displaystyle P \left(\frac{{\it `x__F` }-F ^{\eslarr }\left(1-\frac{1}{n }\right)}{{\it `x__F` }-{\it `M__n` }}<y \right)\essrarr e ^{-y ^{-\alpha }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("`M__n`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("`x__F`")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal")), Typesetting:-mi("")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("y", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("y", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), superscriptshift = "0"))}{$\displaystyle P \left(\frac{{\it `M__n` }-{\it `x__F` }}{{\it `x__F` }-F ^{\eslarr }\left(1-\frac{1}{n }\right)}<-\frac{1}{y }\right)\essrarr e ^{-y ^{-\alpha }}$}
\mapleinline{inert}{2d}{}{$\displaystyle $}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Let\mapleinline{inert}{2d}{-1/y = x}{$\displaystyle -{y}^{-1}=x$}
. Then}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("`M__n`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("`x__F`")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("`x__F`"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&larr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em")), superscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal")), Typesetting:-mi("")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msup(Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), superscriptshift = "0"))}{\[\displaystyle P \left(\frac{{\it `M__n` }-{\it `x__F` }}{{\it `x__F` }-F ^{\eslarr }\left(1-\frac{1}{n }\right)}<x \right)\essrarr e ^{-\left(-x \right)^{\alpha }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
This means, with constant\mapleinline{inert}{2d}{c__n = x__F-`#msup(mi("F"),mo("&larr;"))`(1-1/n)}{$\displaystyle {\it c\_\_n}={\it x\_\_F}-\mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) $}
,\mapleinline{inert}{2d}{`#mover(mo("&rarr;"),mi("d"))`*`~`[Y]*`&Psi;__&alpha;`/c__n(M__n-x__F)}{$\displaystyle {\frac {\mbox {{\tt `\#mover(mo("`}}\rightarrow \mbox {{\tt `"),mi("d"))`}}\\
\mbox{}\\
\mbox{}\,\mbox {{\tt `\~{}`}}_{{Y}}\Psi {\it \_\_}\alpha \\
\mbox{}}{{\it c\_\_n} \left( {\it M\_\_n}-{\it x\_\_F} \right) }}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{3. (a)}We show the statement is true for von. Vise's function. Since all function in\mapleinline{inert}{2d}{MDA(Lambda)}{$\displaystyle {\it MDA} \left( \Lambda \right) $}
have tails equivalent to that of the von Mise's function, that the statement is true for von Mise's function implies it is also true for the rest of the functions in}\end{Maple Normal}

\begin{Maple Normal}{
have tails equivalent to that of the von Mise's function, that the statement is true for von Mise's function implies it is also true for the rest of the functions in\mapleinline{inert}{2d}{MDA(Lambda)}{$\displaystyle {\it MDA} \left( \Lambda \right) $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accentunder = "false"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&infin;", italic = "true", mathvariant = "italic", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("y", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dy", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle \Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\frac{\Mapleoverset{-}{F }\left(t \mathop{\rm  }x \right)}{\Mapleoverset{-}{F }\left(x \right)}=\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\mathop{\rm  }\exp \left(-\Mapleoverset{t \mathop{\rm  }x }{\Mapleunderset{x }{\int }}\frac{1}{a \left(y \right)}{\it dy} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&infin;", italic = "true", mathvariant = "italic", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle =\mathop{\rm  }\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\mathop{\rm  }\exp \left(-\frac{x \left(t -1\right)}{a \left(x \right)}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
if\textbf{}\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi(""))}{$\displaystyle a \left(x \right)\essrarr \infty $}
,\mapleinline{inert}{2d}{limit(x(t-1)/a(x), x = infinity) = limit((t-1)/(diff(a(x), x)), x = infinity)}{$\displaystyle \lim _{x\rightarrow \infty }{\frac {x \left( t-1 \right) }{a \left( x \right) }}=\lim _{x\rightarrow \infty }{\frac {t-1}{{\frac {d}{dx}}a \left( x \right) }}$}
. Since\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("'", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle a \esapos \left(x \right)\essrarr 0$}
,\mapleinline{inert}{2d}{limit((t-1)/(diff(a(x), x)), x = infinity) = piecewise(0 < t and t < 1, -infinity, t > 1, infinity)}{$\displaystyle \lim _{x\rightarrow \infty }{\frac {t-1}{{\frac {d}{dx}}a \left( x \right) }}=\cases{-\infty &$0<t$\  and \ $t<1$\cr \infty &$1<t$\cr}$}
.and\mapleinline{inert}{2d}{limit(exp[-x(t-1)/a(x)], x = infinity) = piecewise(0 < t and t < 1, infinity, t > 1, 0)}{$\displaystyle \lim _{x\rightarrow \infty }\exp_{{-{\frac {x \left( t-1 \right) }{a \left( x \right) }}}}=\cases{\infty &$0<t$\  and \ $t<1$\cr 0&$1<t$\cr}$}

\begin{Maple Normal}{
\begin{Maple Normal}{
if\mapleinline{inert}{2d}{not (a(x) or 0)}{$\displaystyle {\rm \neg \:} a \left( x \right) {\rm \: \vee \:}0$}
or\mapleinline{inert}{2d}{not (a(x) or C)}{$\displaystyle {\rm \neg \:} a \left( x \right) {\rm \: \vee \:}C$}
, where\mapleinline{inert}{2d}{C}{$\displaystyle C$}
is a positive constant, it is clear\mapleinline{inert}{2d}{limit(x(t-1)/a(x), x = infinity) = piecewise(0 < t and t < 1, -infinity, t > 1, infinity)}{$\displaystyle \lim _{x\rightarrow \infty }{\frac {x \left( t-1 \right) }{a \left( x \right) }}=\cases{-\infty &$0<t$\  and \ $t<1$\cr \infty &$1<t$\cr}$}
.and\mapleinline{inert}{2d}{limit(exp[-x(t-1)/a(x)], x = infinity) = piecewise(0 < t and t < 1, infinity, t > 1, 0)}{$\displaystyle \lim _{x\rightarrow \infty }\exp_{{-{\frac {x \left( t-1 \right) }{a \left( x \right) }}}}=\cases{\infty &$0<t$\  and \ $t<1$\cr 0&$1<t$\cr}$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{3. (b) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("G", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]")), Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))}{\[\displaystyle \frac{\Mapleoverset{-}{F }\left(x \right)}{\Mapleoverset{-}{G }\left(x \right)}\estilde \frac{\exp \left(-\Mapleoverset{x }{\Mapleunderset{z }{\int }}\frac{1}{a \left(t \right)}{\it dt} \right)}{x ^{-\alpha }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle \estilde \mathop{\rm  }\exp \left(-\Mapleoverset{x }{\Mapleunderset{z }{\int }}\frac{1}{a \left(t \right)}{\it dt} +\alpha \mathop{\rm  }\mathop{\rm ln}\left(x \right)\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \estilde \exp \left(\Mapleoverset{x }{\Mapleunderset{\xi }{\int }}\left(\frac{\alpha }{t }-\frac{1}{a \left(t \right)}\right){\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{xi}{$\displaystyle \xi$}
is a very large number of our choice.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \exp \left(\Mapleoverset{x }{\Mapleunderset{\xi }{\int }}\left(\frac{\alpha }{t }-\frac{1}{a \left(t \right)}\right){\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \estilde \exp \left(\Mapleoverset{x }{\Mapleunderset{\xi }{\int }}\left(\frac{\alpha \mathop{\rm  }a \left(t \right)-t }{t \mathop{\rm  }a \left(t \right)}\right){\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Here we note that the integrand tends to a constant\mapleinline{inert}{2d}{-1/a__0}{$\displaystyle -{{\it a\_\_0}}^{-1}$}
as\mapleinline{inert}{2d}{proc (xi) options operator, arrow; infinity end proc}{$\displaystyle \xi\mapsto \infty $}
. Therefore}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&xi;", italic = "false", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \exp \left(\Mapleoverset{x }{\Mapleunderset{\xi }{\int }}\left(\frac{\alpha \mathop{\rm  }a \left(t \right)-t }{t \mathop{\rm  }a \left(t \right)}\right){\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`~exp`*{((x-xi)/(xi*a__0))[a__0*alpha-xi]}}{\[\displaystyle \mbox {{\tt `\~{}exp`}}\, \left\{ {\frac {x-\xi}{\xi\,{\it a\_\_0}}}_{{\alpha\,{\it a\_\_0}-\xi}} \right\} \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{xi}{$\displaystyle \xi$}
can be chosen arbitrarily large, the above expression can be arbitrarily close to 0. Hence}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(`#mover(mi("F"),mo("&uminus0;"))`(x)/`#mover(mi("G"),mo("&uminus0;"))`(x), x = infinity) = 0}{\[\displaystyle \lim _{x\rightarrow \infty }{\frac {\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) }{\mbox {{\tt `\#mover(mi("G"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) }}=0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{Now consider}\mapleinline{inert}{2d}{E*`X__+`^delta}{$\displaystyle E{\mbox {{\tt `X\_\_+`}}}^{\delta}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("E", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("`X__+`")), mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("`x__0`"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("")), accent = "true"), Typesetting:-mi(""))}{\[\displaystyle E \mathop{\rm  }\left({\it `X__+` }\right)^{\delta }=-\Mapleoverset{\infty }{\Mapleunderset{{\it `x__0` }\mathop{\rm  }}{\int }}x ^{\delta }\Mapleoverset{}{d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{x__0 = inf*supp(`X__+`)}{$\displaystyle {\it x\_\_0}=\inf\,{\it supp} \left( \mbox {{\tt `X\_\_+`}} \right) $}
. Then}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("E", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("`X__+`")), mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("`x__0`")), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mi(""))}{\[\displaystyle E \mathop{\rm  }\left({\it `X__+` }\right)^{\delta }=\mathop{\rm  }-\Mapleoverset{z }{\Mapleunderset{{\it `x__0` }}{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)-\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
for some large constant z. The 1st integral is clearly finite. As for the 2nd integral,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("F", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{\[\displaystyle -\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)<-\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{F }\left(x \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(x) = x^(-alpha)*L(x)}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) ={x}^{-\alpha}L \left( x \right) $}
is any regularly varying function with\mapleinline{inert}{2d}{alpha > 0}{$\displaystyle 0<\alpha$}
. Then}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("F", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munder(Typesetting:-mi("lim", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("C", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("dx", italic = "true", mathvariant = "italic"))}{\[\displaystyle -\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{F }\left(x \right)=-\Mapleunderset{x \essrarr \infty }{\mathop{\rm lim}}\mathop{\rm  }x ^{\delta -\alpha }L \left(x \right)+C \mathop{\rm  }+\mathop{\rm  }\delta \mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta -\alpha -1}L \left(x \right)\mathop{\rm  }{\it dx} \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where C is a constant. For\mapleinline{inert}{2d}{alpha > delta}{$\displaystyle \delta<\alpha$}
,}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(x^(delta-alpha)*L(x), x = infinity) = 0}{\[\displaystyle \lim _{x\rightarrow \infty }{x}^{\delta-\alpha}L \left( x \right) =0\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
and by Karamata's theorem}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&int;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("dx", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("L", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle \Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta -\alpha -1}L \left(x \right)\mathop{\rm  }{\it dx} \mathop{\rm  }\estilde \frac{1}{\alpha -\delta }z ^{\delta -\alpha }L \left(z \right)\essrarr 0$}
as\mapleinline{inert}{2d}{proc (z) options operator, arrow; infinity end proc}{$\displaystyle z\mapsto \infty $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Therefore\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("G", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("&delta;", italic = "false", mathvariant = "normal")), superscriptshift = "0"), Typesetting:-mi("d", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mover(Typesetting:-mrow(Typesetting:-mi("F", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("<", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mrow(Typesetting:-mi("&infin;", italic = "true", mathvariant = "italic")), Typesetting:-mi(""))}{$\displaystyle -\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{G }\left(x \right)<-\Mapleoverset{\infty }{\Mapleunderset{z }{\int }}x ^{\delta }d \mathop{\rm  }\Mapleoverset{-}{F }\left(x \right)<\infty $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{3. (c) Proof:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
For n large,\mapleinline{inert}{2d}{1/n = `#mover(mi("F"),mo("&uminus0;"))`(d__n)}{$\displaystyle {n}^{-1}=\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it d\_\_n} \right) $}
implies}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("c", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("`d__n`")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]"))}{\[\displaystyle \frac{1}{n }=c \mathop{\rm  }\exp \left(-\Mapleoverset{{\it `d__n` }}{\Mapleunderset{z }{\int }}\frac{1}{a \left(t \right)}{\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{z}{$\displaystyle z$}
is a large constant. We can write}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("z", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("`d__n`")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("ln", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("c", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{\[\displaystyle -\Mapleoverset{{\it `d__n` }}{\Mapleunderset{z }{\int }}\frac{1}{a \left(t \right)}{\it dt} =-\mathop{\rm ln}\mathop{\rm  }\left(n \mathop{\rm  }c \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Let\mapleinline{inert}{2d}{B(t) = t/a(t)}{$\displaystyle B \left( t \right) ={\frac {t}{a \left( t \right) }}$}
Then\mapleinline{inert}{2d}{(D(B))(t) = 1/a(t) and 1/a(t) > 0}{$\displaystyle \mbox {D} \left( B \right)  \left( t \right) = \left( a \left( t \right)  \right) ^{-1}{\rm \: \wedge \:}0< \left( a \left( t \right)  \right) ^{-1}$}
(notice that\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("'", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle a \esapos \left(t \right)\essrarr 0$}
as\mapleinline{inert}{2d}{proc (t) options operator, arrow; infinity end proc}{$\displaystyle t\mapsto \infty $}
). Hence\mapleinline{inert}{2d}{B(t)}{$\displaystyle B \left( t \right) $}
is strictly increasing and invertible. So we have}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{B(d__n)-B(z) = ln(n*c)}{\[\displaystyle B \left( {\it d\_\_n} \right) -B \left( z \right) =\ln  \left( nc \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{d__n = [ln(n*c)+B(z)]/B}{\[\displaystyle {\it d\_\_n}={\frac {[\ln  \left( nc \right) +B \left( z \right) ]}{B}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
This is a unique solution.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{3. (d)}We first show that}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(`#mover(mi("F"),mo("&uminus0;"))`(x+a(x)*t)/`#mover(mi("F"),mo("&uminus0;"))`(x), x = x__F) = e^(-t)}{\[\displaystyle \lim _{x\rightarrow {\it x\_\_F}}{\frac {\mbox {{\tt `\#mover(mi("F"),mo("`}}\\
\mbox{}-\mbox {{\tt `"))`}}\\
\mbox{} \left( x+a \left( x \right) t \right) }{\mbox {{\tt `\#mover(mi("F"),mo("`}}\\
\mbox{}-\mbox {{\tt `"))`}}\\
\mbox{} \left( x \right) }}={e}^{-t}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
The LHS is}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("u", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("du", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \frac{\Mapleoverset{-}{F }\left(x \mathop{\rm  }+\mathop{\rm  }a \left(x \right)\mathop{\rm  }t \right)}{\Mapleoverset{-}{F }\left(x \right)}=\exp \left(-\Mapleoverset{x +t \mathop{\rm  }a \left(x \right)}{\Mapleunderset{x }{\int }}\frac{1}{a \left(u \right)}\mathop{\rm  }{\it du} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Let\mapleinline{inert}{2d}{v = (u-x)/a(x)}{$\displaystyle v={\frac {u-x}{a \left( x \right) }}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mover(Typesetting:-mi("F", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em")), accent = "true"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("v", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dv", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle \frac{\Mapleoverset{-}{F }\left(x \mathop{\rm  }+\mathop{\rm  }a \left(x \right)\mathop{\rm  }t \right)}{\Mapleoverset{-}{F }\left(x \right)}=\exp \left(-\Mapleoverset{t }{\Mapleunderset{0}{\int }}\frac{a \left(x \right)}{a \left(v \mathop{\rm  }a \left(x \right)+x \right)}{\it dv} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Because\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mo("'", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("0", mathvariant = "normal"))}{$\displaystyle a \esapos \left(x \right)\essrarr 0$}
as\mapleinline{inert}{2d}{proc (x) options operator, arrow; x__F end proc}{$\displaystyle x\mapsto {\it x\_\_F}$}
we conclude\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("v", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"))}{$\displaystyle a \left(v \mathop{\rm  }a \left(x \right)+x \right)\essrarr a \left(x \right)$}
and thus}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("v", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "[", close = "]")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&srarr;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mn("1", mathvariant = "normal"))}{\[\displaystyle \frac{a \left(x \right)}{a \left(v \mathop{\rm  }a \left(x \right)+x \right)}\essrarr 1\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
It immediately follows}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(`#mover(mi("F"),mo("&uminus0;"))`(x+a(x)*t)/`#mover(mi("F"),mo("&uminus0;"))`(x), x = x__F) = e^(-t)}{\[\displaystyle \lim _{x\rightarrow {\it x\_\_F}}{\frac {\mbox {{\tt `\#mover(mi("F"),mo("`}}\\
\mbox{}-\mbox {{\tt `"))`}}\\
\mbox{} \left( x+a \left( x \right) t \right) }{\mbox {{\tt `\#mover(mi("F"),mo("`}}\\
\mbox{}-\mbox {{\tt `"))`}}\\
\mbox{} \left( x \right) }}={e}^{-t}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Now we choose\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(d__n) = 1/n}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it d\_\_n} \right) ={n}^{-1}$}
, i.e.\mapleinline{inert}{2d}{d__n = `#msup(mi("F"),mo("&larr;"))`(1-1/n)}{$\displaystyle {\it d\_\_n}=\mbox {{\tt `\#msup(mi("F"),mo("`}}\mbox {{\tt `\&larr;`}}\\
\mbox{}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) $}
we have}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{limit(n*`#mover(mi("F"),mo("&uminus0;"))`(d__n+a(d__n)*t), x = x__F) = e^(-t)}{\[\displaystyle \lim _{x\rightarrow {\it x\_\_F}}n\mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( {\it d\_\_n}+a \left( {\it d\_\_n} \right) t \right) ={e}^{-t}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
By Poisson's approximation}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(M__n <= a(d__n)*x+d__n)*`#mover(mo("&rarr;"),mi("d"))`*exp(-e^(-t))}{\[\displaystyle P \left( {\it M\_\_n}\leq a \left( {\it d\_\_n} \right) x+{\it d\_\_n} \right) \mbox {{\tt `\#mover(mo("`}}\rightarrow \mbox {{\tt `"),mi("d"))`}}\\
\mbox{}\\
\mbox{}\,{{\rm e}^{-{e}^{-t}}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
And thus\mapleinline{inert}{2d}{`in`(F, MDA(Lambda))}{$\displaystyle {\it in} \left( F,{\it MDA} \left( \Lambda \right)  \right) $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{4. (i) Solution:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(z) = C*(int((1+x^2/k)^(-(k+1)*(1/2)), x = z .. infinity))}{\[\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( z \right) =C\int _{z}^{\infty }\! \left( 1+{\frac {{x}^{2}}{k}} \right) ^{-1/2\,k-1/2}{dx}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("C", italic = "true", mathvariant = "italic"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"))), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), superscriptshift = "0"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), superscriptshift = "0"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("hypergeom", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"))), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"))), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mn("2", mathvariant = "normal"), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"))))), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))), mathvariant = "normal"))}{\[\displaystyle =\mathop{\rm  }C \mathop{\rm  }k ^{\frac{1}{2}~k -\frac{1}{2}}~z ^{-k }~\mathop{\rm hypergeom}\left(\left(\frac{1}{2}~k ,\frac{1}{2}~k +\frac{1}{2}\right),\left(1+\frac{1}{2}~k \right),-\frac{k }{z ^{2}}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
So\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(z)}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( z \right) $}
is regularly varying with index\mapleinline{inert}{2d}{-k}{$\displaystyle -k$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{4. (ii)}Since\mapleinline{inert}{2d}{`#mover(mi("F"),mo("&uminus0;"))`(x)}{$\displaystyle \mbox {{\tt `\#mover(mi("F"),mo("`}}-\mbox {{\tt `"))`}}\\
\mbox{}\\
\mbox{} \left( x \right) $}
is regularly varying,\mapleinline{inert}{2d}{`in`(F, MDA(`&Phi;__k`))}{$\displaystyle {\it in} \left( F,{\it MDA} \left( \Phi {\it \_\_k}\\
\mbox{} \right)  \right) $}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{4. (iii)}we choose\mapleinline{inert}{2d}{c__n = `#msup(mi("F"),mo("&larr;"))`(1-1/n) and `#msup(mi("F"),mo("&larr;"))`(1-1/n) = {C*hypergeom([(1/2)*k, (1/2)*k+1/2], [1+(1/2)*k], -k/z^2)*k^((1/2)*k-1/2)*n}^(1/k)}{$\displaystyle {\it c\_\_n}=\mbox {{\tt `\#msup(mi("F"),mo("`}}\\
\mbox{}\mbox {{\tt `\&larr;`}}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) {\rm \: \wedge \:}\mbox {{\tt `\#msup(mi("F"),mo("`}}\\
\mbox{}\mbox {{\tt `\&larr;`}}\mbox {{\tt `"))`}}\\
\mbox{} \left( 1-{n}^{-1} \right) ={ \left\{ C{\mbox{$_2$F$_1$}(1/2\,k,1/2\,k+1/2;\,1+1/2\,k;\,-{\frac {k}{{z}^{2}}})}\\
\mbox{}{k}^{1/2\,k-1/2}n \right\} }^{{k}^{-1}}$}
and\mapleinline{inert}{2d}{d__n = 0}{$\displaystyle {\it d\_\_n}=0$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{5. (i) Solution:}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(X^p > x)}{\[\displaystyle P \left( x<{X}^{p} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mo(">", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), superscriptshift = "0")), mathvariant = "normal"))}{\[\displaystyle =\mathop{\rm  }P \left(X >x ^{\frac{1}{p }}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), superscriptshift = "0"), Typesetting:-mi(""))}{\[\displaystyle =\mathop{\rm  }x ^{-\frac{\alpha }{p }}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
So\mapleinline{inert}{2d}{X^p}{$\displaystyle {X}^{p}$}
has regularly varying tail with index\mapleinline{inert}{2d}{-alpha/p}{$\displaystyle -{\frac {\alpha}{p}}$}
, and thus belong to\mapleinline{inert}{2d}{MDA(`&Phi;__&beta;`)}{$\displaystyle {\it MDA} \left( \Phi {\it \_\_}\beta  \right) $}
where\mapleinline{inert}{2d}{beta = alpha/p}{$\displaystyle \beta={\frac {\alpha}{p}}$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{P(ln*X > x)}{\[\displaystyle P \left( x< \left( {\it ln} \right) \,X \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi("X", italic = "true", mathvariant = "italic"), Typesetting:-mo(">", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), superscriptshift = "0")), mathvariant = "normal"))}{\[\displaystyle =\mathop{\rm  }P \left(X >e ^{x }\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msup(Typesetting:-mi("e", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("x", italic = "true", mathvariant = "italic")), superscriptshift = "0"))}{\[\displaystyle =\mathop{\rm  }e ^{-\alpha \mathop{\rm  }x }\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle =\mathop{\rm  }\exp \left(-\Mapleoverset{x }{\Mapleunderset{0}{\int }}\alpha \mathop{\rm  }{\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mi("exp", italic = "false", mathvariant = "normal"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&Integral;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mn("0", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("x", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-msup(Typesetting:-mi("&alpha;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mo("&uminus0;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi("dt", italic = "true", mathvariant = "italic")), mathvariant = "normal", open = "{", close = "}"))}{\[\displaystyle =\mathop{\rm  }\exp \left(-\Mapleoverset{x }{\Mapleunderset{0}{\int }}\frac{1}{\alpha ^{-1}}{\it dt} \right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
This is a von. Mise's function with auxilliary function\mapleinline{inert}{2d}{a(x) = 1/alpha}{$\displaystyle a \left( x \right) ={\alpha}^{-1}$}
. Obviously it satisfies\mapleinline{inert}{2d}{diff(a(x), x) = 0}{$\displaystyle {\frac {d}{dx}}a \left( x \right) =0$}
. Therefore it is in the MDA of Gumbel distribution.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{5. (ii)}}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
Following the formula for functions in MDA of Fréchet distribution, we choose, for\mapleinline{inert}{2d}{X^p}{$\displaystyle {X}^{p}$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{c__n = n^(p/alpha)}{\[\displaystyle {\it c\_\_n}={n}^{{\frac {p}{\alpha}}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
and\mapleinline{inert}{2d}{d__n = 0}{$\displaystyle {\it d\_\_n}=0$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\textbf{5. (iii)}Using the formula for von Mise's functions, we choose for\mapleinline{inert}{2d}{ln(X)}{$\displaystyle \ln  \left( X \right) $}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{d__n = ln(n)/alpha}{\[\displaystyle {\it d\_\_n}={\frac {\ln  \left( n \right) }{\alpha}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{c__n = 1/alpha}{\[\displaystyle {\it c\_\_n}={\alpha}^{-1}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{}{\[\displaystyle \]}
}\end{Maple Normal}
}\end{Maple Normal}
\end{document}

### 2015-01-26 thesis/Comparison_to_TW.tex
Add the file that compares the largest eigenvalue's distribution of the Tracy-Widom.

--- 2014-12-01 thesis/paper.tex
@@
 \usepackage[bookmarks=true]{hyperref}
-% \usepackage[acronym]{glossaries}
-%\usepackage{bookmark}
 \input{../physics_common.tex}
 
-\DeclareGraphicsExtensions{.eps}
-\title{Spectral properties of a Wishart-GARCH Matrix}
-\author{Prof. Sven \AA berg\\
-  Xie Xiaolei}
+\DeclareGraphicsExtensions{.pdf}
 \begin{document}
-\maketitle
-\section{The Spectral Distribution}
-\subsection{Dependence on auto-correlation}
-Figure \ref{fig:spectral_dist} shows the spectral density of the
-correlation matrix of N identically distributed GARCH(1,1)
-processes. These processes have such parameters that 
-the $\alpha_1 + \beta_1 \approx 1$ and hence the auto-correlation
-function $\text{corr}(r^2_t, r^2_{t-k})$ decays with a similar slow
-rate for all the processes but from different initial values. The
-parameter settings also make these processes have very similar tail
-behavior with a tail exponent $\alpha$ in the range $2.17 \pm 0.01$.
-
-Figure \ref{fig:spectral_dist1} shows the central part of the spectral
-distribution and figure \ref{fig:spectral_dist2} shows how the
-Hellinger distance between the empirical spectral density and the
-Marcenko Pastur law changes with the parameter $\alpha_1$.
-It is seen that, when measured by the Hellinger metric, the empirical
-spectral density is most similar to the Marcenko-Pastur law when
-$\alpha_1 = 1/2$.
 \begin{figure}[htb!]
   \centering
-  \subfigure[Spectral density of the correlation matrix plotted on
-  log-log scale. The mean is 1 for all the empirical distributions and
-  also for the Marcenko-Pastur law.]{
-    \includegraphics[scale=0.32]{../pics/spectral_dist2.eps}
-    \label{fig:spectral_dist2}
+  \subfigure[q = 0.1]{
+    \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot1.pdf}
   }
-  \hspace{5mm}
-  \subfigure[Hellinger distance between the empirical spectral density
-  and the Marcenko-Pastur law.]{
-    \includegraphics[scale=0.32]{../pics/spectral_dist1.eps}
-    \label{fig:spectral_dist1}
+  \subfigure[q = 0.2]{
+    \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot2.pdf}
   }
-  \caption{\small \it Spectral density of the
-    correlation matrix of N identically distributed GARCH(1,1)
-    processes.}
-  \label{fig:spectral_dist}
-\end{figure}
-On the other hand, it is also seen that the variance of the spectral
-distributions decreases as auto-correlation becomes stronger. This is
-further shown in figure \ref{fig:spectral_dist3}.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5]{../pics/spectral_dist3.eps}
-  \caption{\small \it Variance of the spectral distributions.}
-  \label{fig:spectral_dist3}
+  \caption{\small \it Comparison of the largest eigenvalue's
+    distribution with Tracy-Widom.}
+  \label{fig:eig1_TW}
 \end{figure}
 
-\subsection{Dependence on the tail exponent}
-Figure \ref{fig:spectral_density1} shows the dependence of the
-spectral density of Wishart-GARCH matrices on the tail exponent
-$\alpha$.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.4]{../pics/spectral_density1.eps}
-    \caption{\small \it Dependence of spectral density (plotted on
-      log-log scale) on the tail exponent. The means of the 3
-      distributions are all equal to 1 and the variances are all equal
-      to 0.04}
-    \label{fig:spectral_density1}
-\end{figure}
-The mean and variance of the three density functions as well as their
-Hellinger distances to the Marcenko-Pastur law are as follows:
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|l|c|c|c|}
-    \hline
-    Tail Exponent & 2.17 & 2.99 & 3.99 \\
-    \hline
-    mean & 1.0000 & 1.0000 & 1.0000 \\
-    \hline
-    variance & 0.0418 & 0.0409 & 0.0408 \\
-    \hline
-    Hellinger distance & 0.1012 & 0.0639 & 0.0586 \\
-    \hline
-  \end{tabular}
-  \caption{mean, variance and Hellinger distance to the
-    Marcenko-Pastur law of the three density functions in figure
-    \ref{fig:spectral_density1}}
-  \label{tab:spectral_density_hellinger}
-\end{table}
-
-\section{Distribution of the largest eigenvalue}
-\subsection{Dependence on auto-correlation}
-Chiani showed that the transformed largest eigenvalue of a Wishart
-matrix obeyed approximately a gamma distribution \cite{Chiani2012}. Figure
-\ref{fig:eig1_dist1} shows how this distribution is modified as the
-auto-correlation between the squared returns becomes stronger; figure
-\ref{fig:eig1_dist2} shows the growth of the difference between the
-empirical distribution and a gamma distribution that has the same
-mean and variance. It is clear that the logarithm of the Kolmogorov
-distance is approximately linear in the GARCH parameter $\alpha_1$.
-
-\begin{figure}[htb!]
-  \centering
-  \subfigure[Cumulative distribution of the largest eigenvalue]{
-    \includegraphics[scale=0.33]{../pics/maxeigen_dist_garch.eps}
-    \label{fig:eig1_dist1}
-  }
-  \subfigure[Kolmogorov distance between the empirical cumulative
-  distribution and a gamma distribution with the same mean and variance]{
-    \includegraphics[scale=0.45]{../pics/maxeigen_dist_garch_ks.eps}
-    \label{fig:eig1_dist2}
-  }
-  \caption{\small \it Distribution of the largest eigenvalue and
-    comparison to a gamma distribution.}
-  \label{fig:eig1_dist}
-\end{figure}
-
-Figure \ref{fig:eig1_dist3} shows the density function of the largest
-eigenvalue and figure \ref{fig:eig1_dist4} shows the tail of the
-complementary distribution function of the largest eigenvalue. Here by
-``normalized'' we mean the following transformation:
-\begin{eqnarray*}
-  \lambda'_1 &=& {\lambda_1 - \E(\lambda_1) \over \text{std}(\lambda_1)}  
-\end{eqnarray*}
-
-\begin{figure}[htb!]
-  \centering
-  \subfigure[Probability density function of the normalized largest
-  eigenvalue]{
-    \includegraphics[scale=0.32]{../pics/maxeigen_dist_pdf.eps}
-    \label{fig:eig1_dist3}
-  }
-  \hspace{5mm}
-  \subfigure[Complementary distribution function of the
-  normalized largest eigenvalue on log-log scale.]{
-    \includegraphics[scale=0.32]{../pics/maxeigen_dist_tail.eps}
-    \label{fig:eig1_dist4}
-  }
-  \caption{\small \it Tail behavior of the largest eigenvalue.}
-  \label{fig:eig1_dist_tail}
-\end{figure}
-
-\subsection{Dependence on the tail exponent}
-Figure \ref{fig:eig1_dist_diff_alpha} shows how the distribution of
-the largest eigenvalue depends on the tail exponent.
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.4]{../pics/eig1_dist_diff_alpha.eps}
-    \caption{\small \it Dependence of the largest eigenvalue's
-      probability density function on the tail exponent.}
-    \label{fig:eig1_dist_diff_alpha}
-\end{figure}
-
-The mean and variance of the largest eigenvalue's distributions are
-listed in table \ref{tab:eig1_dist_diff_alpha}
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{|l|c|c|c|}
-    \hline
-    Tail Exponent & 2.17 & 2.99 & 3.99 \\
-    \hline
-    $\mean{\lambda_1}$ & 1.4306 & 1.4157 & 1.4137 \\
-    \hline
-    $\text{var}(\lambda_1)$ & 0.0012 & 0.0006 & 0.0005 \\
-    \hline
-  \end{tabular}
-  \caption{mean and variance of the largest eigenvalue's distribution}
-  \label{tab:eig1_dist_diff_alpha}
-\end{table}
-
-\bibliographystyle{unsrt}
-\bibliography{econophysics}
-
 \end{document}

### 2015-01-28 m/LognormalSpectralDensity.m
Work on the paper with Sven.

@@ -1,0 +1,88 @@
+clear all
+close all
+mysql = get_mysql();
+name='SP500';
+stmt = sprintf('select symbol from %s_components;', name);
+symbols = fetch(mysql, stmt);
+
+N = size(symbols, 1);
+numRec = NaN(N, 1);
+day1 = '2000-01-02';
+day2 = '2015-01-23';
+for k = 1 : N
+    data = fetch(mysql, sprintf(['select count(*) ' ...
+                        'from %s_US where day between "%s" and "%s";'], ...
+                                strrep(symbols{k}, '.', '_'), day1, ...
+                                day2));
+    numRec(k) = cell2mat(data(1, 1));
+end
+T = 3788;
+to_include = find(numRec == T);
+% d1 = datenum(day1);
+% d2 = datenum(day2);
+% days = ones(1, d2 - d1 + 1);
+% for k = to_include'
+%     stmt = sprintf('select distinct day from %s_US;', strrep(symbols{k}, '.', '_'));
+%     data = datenum(fetch(mysql, stmt));
+%     data >= d1 & data <= d2
+% end
+
+N = length(to_include);
+R = NaN(T-1, N);
+j = 1;
+counter = NaN(1, N);
+for k = to_include'
+    stmt = sprintf('select closing from %s_US where day between "%s" and "%s" order by day;', ...
+                                   strrep(symbols{k}, '.', '_'), day1, day2);
+    closing = fetch(mysql, stmt);
+    R(:, j) = price2ret(cell2mat(closing));
+    counter(j) = length(closing);
+    j = j + 1;
+end
+close(mysql);
+
+c = 1;
+q = 0.8;
+
+n = floor(N/q);
+sections = floor(T/n);
+ev = NaN(N * sections, 1);
+for k = 1 : sections
+    A = R(n*(k-1)+1 : n*k, :);
+    C = A' * A;
+    ev(N*(k-1)+1 : N*k) = eig(C);
+end
+ev = sort(ev, 'descend');
+% Remove the largest from each section
+ev = ev(sections + 1 : end);
+sig = ev(1);
+ev = ev ./ sig;
+v = 0.8;
+
+ops = statset('mlecustom');
+ops.MaxFunEvals = 600;
+ops.MaxIter = 300;
[…]

### 2015-01-28 m/paper1/estimate_v_daily.m

@@ -6,27 +6,31 @@ stmt = sprintf('select symbol from %s_components;', name);
 symbols = fetch(mysql, stmt);
 
 p = size(symbols, 1);
-T = 3000;
 numRec = NaN(p, 1);
-
+day1 = '2000-01-02';
+day2 = '2015-01-23';
 for k = 1 : p
-    numRec(k) = cell2mat(fetch(mysql, sprintf(['select count(*) from %s_US'], ...
-                                              strrep(symbols{k}, '.', '_'))));
+    data = fetch(mysql, sprintf(['select count(*) ' ...
+                        'from %s_US where day between "%s" and "%s";'], ...
+                                strrep(symbols{k}, '.', '_'), day1, ...
+                                day2));
+    numRec(k) = cell2mat(data(1, 1));
 end
+T = 3788;
+to_include = find(numRec == T);
 
-to_include = find(numRec >= T+1);
 p = length(to_include);
 
-R = NaN(T, p);
-logvol = NaN(T, p);
+R = NaN(T-1, p);
+logvol = NaN(T - 1, p);
 c = 1;
 for k = to_include'
     data = fetch(mysql, sprintf(...
         ['select closing, (high - low)/low from ' ...
-         '%s_US order by day desc limit %d;'], ...
-        strrep(symbols{k}, '.', '_'), T+1));
-    R(:, c) = price2ret(flipud(cell2mat(data(:, 1))));
-    A = flipud(cell2mat(data(2:end, 2)));
+         '%s_US where day between "%s" and "%s" order by day;'], ...
+        strrep(symbols{k}, '.', '_'), day1, day2));
+    R(:, c) = price2ret(cell2mat(data(:, 1)));
+    A = cell2mat(data(2:end, 2));
     A(A == 0) = min(A(A ~= 0));
     logvol(:, c) = log(A);
     logvol(:, c) = logvol(:, c) - mean(logvol(:, c));

### 2015-01-28 thesis/Comparison_to_TW.tex
Estimated the spectral density of S&P-500 stocks.

@@
 \begin{document}
+How the maximum eigenvalue's distribution compares to Tracy-Widom:
+
 \begin{figure}[htb!]
   \centering
-  \subfigure[q = 0.1]{
   \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot1.pdf}
+  \caption{\small \it Comparison of the largest eigenvalue's
+    distribution with Tracy-Widom. $q=N/T=0.1$}
+  \label{fig:eig1_TW_q0.1}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot2.pdf}
+  \caption{\small \it Comparison of the largest eigenvalue's
+    distribution with Tracy-Widom. $q=N/T=0.2$}
+  \label{fig:eig1_TW_q0.2}
+\end{figure}
+
+Comparison between the empirical spectral density and the theoretical
+density. The value of $v$ used to obtain the theoretical density is
+the result of {\it maximum likelihood estimation}.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[q = 0.5, KL = 0.19]{
+    \includegraphics[scale=0.4, clip=true, trim=115 271 109
+    204]{../pics/spectral_density_q0dot5.pdf}
   }
-  \subfigure[q = 0.2]{
-    \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot2.pdf}
+  \subfigure[q = 0.65, KL = 0.15]{
+    \includegraphics[scale=0.4, clip=true, trim=115 271 109
+    204]{../pics/spectral_density_q0dot65.pdf}
   }
-  \caption{\small \it Comparison of the largest eigenvalue's
-    distribution with Tracy-Widom.}
-  \label{fig:eig1_TW}
+  \subfigure[q = 0.8, KL=0.14]{
+    \includegraphics[scale=0.4, clip=true, trim=115 271 109
+    204]{../pics/spectral_density_q0dot8.pdf}
+  }
+  \subfigure[q = 0.95, KL=0.15]{
+    \includegraphics[scale=0.4, clip=true, trim=115 271 109
+    204]{../pics/spectral_density_q0dot95.pdf}
+  }
+  \caption{\small \it Comparison of the empirical and theoretical
+    spectral densities. The largest empirical eigenvalue ($\lambda_1$)
+    is excluded. The rest is rescaled by $1/\lambda_1$. The
+    theoretical density functions are ajusted accordingly. KL is the
+    Kullback-Leibler divergence of the theoretical density function from
+    the empirical density function. Only intervals on which the
+    empirical density is non-zero are included in the calculation of
+    KL.}
+  \label{fig:spectral_density}
 \end{figure}

### 2015-01-29 thesis/Results.tex
Files belonging to the paper with Sven.

\documentclass{article}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{amsfonts, amsmath}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\input{../physics_common.tex}

\DeclareGraphicsExtensions{.pdf}
\begin{document}
Figure \ref{fig:eig1_TW_q0.5} and \ref{fig:eig1_TW_q1.0} are
quantile-to-quantile plots of the maximum eigenvalue's distribution
compared to Tracy-Widom. A straight line is expected if the two
distributions differ only by translation and/or scaling.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../r/eigmax_TW_q0dot5.pdf}
  \caption{\small \it Comparison of the largest eigenvalue's
    distribution (quantiles along the horizontal axis) with
    Tracy-Widom (quantiles along the vertical axis). $q=N/T=0.5$}
  \label{fig:eig1_TW_q0.5}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../r/eigmax_TW_q1dot0.pdf}
  \caption{\small \it Comparison of the largest eigenvalue's
    distribution (quantiles along the horizontal axis) with
    Tracy-Widom (quantiles along the vertical axis). $q=N/T=1.0$}
  \label{fig:eig1_TW_q1.0}
\end{figure}

\pagebreak
Figure \ref{fig:eigmax_HillPlot} shows the estimated tail-index of the
maximum eigenvalue distribution against the number of upper-order
statistics used in the estimation (Hill-plot). It is seen that the
region approximately at $(210, 250)$ is stable, suggesting a tail
index of 2.34.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{../r/TailIndex_eignmax_q1dot0_v0dot748.pdf}
  \caption{\small \it Hill plot of the largest eigenvalue's
    distribution. $q = 1$, $v = 0.748$.}
  \label{fig:eigmax_HillPlot}
\end{figure}

\pagebreak
Figure \ref{fig:spectral_density} compares the empirical spectral
density with the theoretical density. The value of $v$ used to obtain
the theoretical density is the result of {\it maximum likelihood
  estimation}.
\begin{figure}[htb!]
  \centering
  \subfigure[q = 0.5, v = 0.9415, KL = 0.19]{
    \includegraphics[scale=0.33, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot5.pdf}
  }
  \subfigure[q = 0.65, v=0.8409, KL = 0.15]{
    \includegraphics[scale=0.33, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot65.pdf}
  }
  \subfigure[q = 0.8, v=0.8808, KL=0.14]{
    \includegraphics[scale=0.33, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot8.pdf}
  }
  \subfigure[q = 0.95, v=0.8859, KL=0.15]{
    \includegraphics[scale=0.33, clip=true, trim=115 271 109
    204]{../pics/spectral_density_q0dot95.pdf}
  }
  \caption{\small \it Comparison of the empirical and theoretical
    spectral densities. The largest empirical eigenvalue ($\lambda_1$)
    is excluded. The rest is rescaled by $1/\lambda_1$. The
    theoretical density functions are rescaled accordingly. KL
    is the Kullback-Leibler divergence of the theoretical density
    function from the empirical density function. Only intervals on
    which the empirical density is non-zero are included in the
    calculation of KL.}
  \label{fig:spectral_density}
\end{figure}

\pagebreak
Figure \ref{fig:SP500-TailIndices} shows the upper- and lower-tail
indices of 441 stocks in the S\&P-500 index.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5, clip=true, trim=94 229 115
  121]{../pics/SP500-TailIndices.pdf}
  \caption{\small \it Upper- and lower-tail indices of 441 stocks in
    the S\&P-500 index. The horizontal coordinate of a point in the
    figure is the upper-tail index and the vertical coordinate is the
    lower-tail index.}
  \label{fig:SP500-TailIndices}
\end{figure}

\end{document}

### 2015-01-29 r/test.r

@@ -7,13 +7,13 @@ matlab <- Matlab();
 open(matlab);
 phi <- c(0, 0.5, 0.8160);
 sig <- c(0.1, 0.2, 0.5);
-q <- 1.0;
+q <- 0.5;
 m <- 3;
 n <- 3;
-## pdf(sprintf("eigmax_TW_q%.1f.pdf", q));
-## par(mfrow=c(3, 3));
-## for (m in 1:length(sig)) {
-##     for (n in 1:length(phi)) {
+pdf(sprintf("eigmax_TW_q%.2f.pdf", q));
+par(mfrow=c(3, 3));
+for (m in 1:length(sig)) {
+    for (n in 1:length(phi)) {
 
         v <- sig[m]^2/(1 - phi[n]^2);
         data <- readMat(sprintf(paste("../matfys/data/sv/normal_ret/lognormal_vol/",
@@ -24,11 +24,12 @@ n <- 3;
             eigmax[k] <- max(data$ev[, k])
         }
 
-        eigmax1 <- rWishartMax(length(eigmax), ndf=data$T, pdim=dim(data$ev)[1], var=exp(v/2), beta=1);
+        eigmax1 <- rWishartMax(length(eigmax), ndf=data$T, pdim=dim(data$ev)[1],
+                               var=exp(v/2), beta=1);
         qqplot(eigmax, eigmax1, xlab=sprintf("v=%.4f", v), ylab="", col="blue", type="l");
-##     }
-## }
-# dev.off();
+    }
+}
+dev.off();
 close(matlab);
 # rm(list = ls())
 

### 2015-01-29 cv/XieXiaolei-Analyst.tex
Looked at the mean excess function of a sample of largest eigenvalues.

@@
   Citizenship: Sweden \\
-  Phone: +46 70 489 0472 \\
+  Phone: +46 76 060 6638 \\
   Email: \href{mailto:xie.xiaolei@gmail.com}{\tt xie.xiaolei@gmail.com} \\

### 2015-02-03 cv/CoverLetter/SecondApplicationToKU.tex
Re-arrange documents of career.

--- 2015-01-05 cv/SecondApplicationToKU.tex
@@
 Covariance Matrix'' and plan to submit it to {\it Quantitative
-  Finance} for publication. In the paper we argued that, on certain
-conditions, autocorrelations in the conditional volatilities could
-affect the empirical covariance matrixes' spectrum only through the
+  Finance} for publication. In the paper we argue that, on certain
+conditions, autocorrelations in the conditional volatilities only
+affect the empirical covariance matrixes' spectrum through the
 unconditional distribution of the volatilities. On these
 conditions, we derived in the case of small autocorrelations the
-spectrum of an empirical covariance matrix. The derivations were based
-on results from free probability theory.
+spectrum of an empirical covariance matrix of identical sequences. The
+derivations are based on results from free probability theory.
 

### 2015-02-03 m/estimate_v.m

@@ -1,0 +1,42 @@
+clear all
+close all
+
+company = 'nordea_bank';
+dt = 30;
+first_day = '2013-10-10';
+last_day = '2014-08-21';
+[ret, sig] = get_intra_ret_simple(company, '2013-10-10', '2014-08-21', ...
+                                           dt, 60);
+X = log(sig);
+Z = ret ./ sig;
+
+Z = Z .* exp(mean(X));
+X = X - mean(X);
+
+
+%% Try (high - low)/low as volatility proxy.
+days = get_recorded_days(company, first_day, last_day);
+mysql = get_mysql();
+num = 8.5 * 60 / dt;
+range_proxy = NaN(size(sig));
+for d = 1:length(days)
+    stmt = sprintf(['select min(price) as low , max(price) as high, ' ...
+                    'floor(timestampdiff(MINUTE, "%s 09:00:00", tid)/%d) as ' ...
+                    't from %s where date(tid) = "%s" group by t order by ' ...
+                    't asc;'], days{d}, dt, company, days{d});
+    data = cell2mat(fetch(mysql, stmt));
+    A = NaN(num, 1);
+    A(data(:, 3)+1) = (data(:, 2) - data(:, 1))./data(:, 1);
+    c = NaN;
+    for k = 2 : num
+        if (isnan(A(k)))
+            A(k) = A(k-1);
+        elseif isnan(c)
+            c = k;
+        end
+    end
+    A(1:c-1) = A(c);
+    A(A == 0) = mean(A ~= 0);
+    range_proxy((d-1)*num+1 : d*num) = A;
+end
+close(mysql);

### 2015-02-09 cv/XieXiaolei-Analyst.tex
Added two cover letters.

@@
 
+\section*{My Goal}
+A self-funded Ph.D student of Finance \& Insurance mathematics
+at KU, who is very experienced in programming and statistical analysis, is
+looking for a part-time job as a quantitative analyst.
+
 \section*{Education}
 \begin{itemize}
-\item Ph.D of Financial Mathematics, October 2014 - present
+\item Ph.D of Finance \& Insurance Mathematics, October 2014 - present
   \begin{itemize}
@@
       determination of the largest eigenvalues of a covariance matrix
-      built from samples with heavy-tails. The purpose is dimension
+      of samples with heavy-tails. The purpose is dimension
       reduction, by which the behaviour of a diverse portfolio is
@@
 
-\item Master of Physics, Lund University, January 2011 - June 2014
+\item Master of Mathematical Physics, Lund University, January 2011 -
+  June 2014
   \begin{itemize}
@@
 
-    % The returns' models are fitted to such Swedish stocks as Volvo B,
-    % Ericsson B, and Nordea Bank, using both publicly available daily
-    % statistics and intra-day transaction data that I have been
-    % collecting using home-developed software.
+    The volatility models were fitted to such Swedish stocks as Volvo B,
+    Ericsson B, and Nordea Bank, using both publicly available daily
+    statistics and intra-day transaction data that I had collected
+    using home-developed software.
 
-    % The studies are carried out analytically wherever feasible, and
-    % otherwise by approximation and simulation. The application of
-    % probability theories, random matrices, and stochastic processes
-    % are the most involved subjects.
+    The studies were carried out analytically wherever feasible, and
+    otherwise by approximation and simulation. Software such as Maple,
+    Mathematica and Matlab were used extensively.
   \end{itemize}
@@
 
-  My main responsibility is software development in C, and sometimes
+  My main responsibility was software development in C, and sometimes
   C++ and Java, in the Linux operating system.
@@
 
-\section*{Private Project}
-{\it Trade data collector and analyzer}
-With this home-developed software I have been collecting intra-day
-transaction data and analyzing them in real-time with Matlab. I make
-heavy use of the Matlab economitrics toolbox with its built-in
-routines for estimating ARIMA and GARCH models and then forecasting
-using the obtained model. The idea is to pre-estimate a model and thus
-fix its structure, such as the distribution of the residuals, the
-order of the auto-regressive and the moving-average components, the
-order of differencing/fractional differencing, etc., and then to 
-evolve the model parameters in real-time as new data are fed into
-the system.
+\section*{Language Skills}
+\begin{itemize}
+\item English: Fluent
+\item Swedish: Fairly good
+\item Danish: Beginner
+\end{itemize}
+
+% \section*{Private Project}
+% {\it Trade data collector and analyzer}
+% With this home-developed software I have been collecting intra-day
+% transaction data and analyzing them in real-time with Matlab. I make
+% heavy use of the Matlab economitrics toolbox with its built-in
+% routines for estimating ARIMA and GARCH models and then forecasting
+% using the obtained model. The idea is to pre-estimate a model and thus
+% fix its structure, such as the distribution of the residuals, the
+% order of the auto-regressive and the moving-average components, the
+% order of differencing/fractional differencing, etc., and then to 
+% evolve the model parameters in real-time as new data are fed into
+% the system.
 \end{document}

### 2015-02-12 thesis/Figures.tex
Worked on the paper with Sven.

\documentclass{article}
\usepackage{amssymb,amsmath,amsthm,bm}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathrsfs}
\graphicspath{{/home/xxie/econophysics/pics/InCaseNeeded/}}

\input{/home/xxie/econophysics/physics_common.tex}

\begin{document}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6]{eigmax_TW.pdf}
  \caption{Distribution of simulated $\lambda_\M$ compared to
    Tracy-Widom. From the 1st to the 4th row, q = 0.1, 0.2, 0.5, 1.}
  \label{fig:eigmax_TW}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6, clip=true, trim=98 233 120
  127]{AppleLogvol-QQ.pdf}
  \caption{QQ-plot of $\ln{ S_\M - S_\m \over S_\m}$. The data are daily trading statistics of Apple from
    2005-01-02 to 2015-01-23. The lower tail of sample log-vol is a
    bit lighter and the upper tail a bit heavier than their
    counterparts of standard normal. }
  \label{fig:qqplot_logvol}
\end{figure}

\end{document}

### 2015-02-12 r/MeanExcessPlot.r

@@ -1,0 +1,22 @@
+library(R.matlab);
+
+data <- readMat(sprintf(paste("../matfys/data/sv/normal_ret/lognormal_vol/",
+                              "q%.1f/Eig-sig%.4f-phi%.4f.mat", sep=""),
+                        0.1, 0.1, 0));
+eigmax = seq(length.out=dim(data$ev)[2]) * NA;
+for (k in 1:length(eigmax)) {
+    eigmax[k] <- max(data$ev[, k])
+}
+Par = WishartMaxPar(ndf=data$T, pdim=dim(data$ev)[1],
+    var=exp(2*v), beta=1);
+eigmax = (eigmax - Par$centering) / Par$scaling;
+E = sort(eigmax, decreasing=FALSE);
+N = floor(length(E) / 5);
+thresholds = tail(E, N);
+meanExcess = rep(NA, N);
+
+for (i  in 1:N-1) {
+    meanExcess[i] = mean(tail(E, N-i )  - thresholds[i]);
+}
+plot(thresholds, meanExcess, type="p");
+

### 2015-02-12 thesis/Figures.tex
Worked on the paper with Sven.

@@
 \usepackage{mathrsfs}
-\graphicspath{{/home/xxie/econophysics/pics/InCaseNeeded/}}
+\usepackage{multirow}
 
-\input{/home/xxie/econophysics/physics_common.tex}
+\graphicspath{{../pics/InCaseNeeded/}}
+
+\input{../physics_common.tex}
 
@@
   \includegraphics[scale=0.6]{eigmax_TW.pdf}
-  \caption{Distribution of simulated $\lambda_\M$ compared to
-    Tracy-Widom. From the 1st to the 4th row, q = 0.1, 0.2, 0.5, 1.}
+  \caption{Distribution of simulated $\lambda_\M$ (blue) compared to
+    Tracy-Widom (red). From the 1st to the 4th row, q = 0.1, 0.2, 0.5,
+    1. $\lambda_\M$ has been centered and rescaled as is required in
+    the case of Wishart matrices. The Tracy-Widom density is computed
+    with variance $e^{2v}$, for $v \le 0.25$. $e^{2v}$ is $\E \sigma^2$,
+    where $\sigma \sim \text{Lognormal}(0, v)$.}
   \label{fig:eigmax_TW}
 \end{figure}
+
+When $v \ge 0.25$, a power-law is observed. Table
+\ref{tab:tail_indices} list the tail indices of the $\lambda_\M$
+corresponding to different values of $q$ and $v$.
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|c|c|}
+    \hline
+    \multicolumn{2}{|c|}{} & \multicolumn{5}{|c|}{$v$} \\
+    \hline
+    \multicolumn{2}{|c|}{} & 0.25 & 0.27 & 0.33 & 0.50 & 0.75 \\
+    \hline
+    \multirow{4}{*}{$q$} & 0.1 & 3.75 & 3.86 & 3.46 & 2.99 & 2.42 \\
+    & 0.2 & 3.23 & 3.98 & 3.24 & 2.90 & 2.73 \\
+    & 0.5 & 3.88 & 4.11 & 3.60 & 2.71 & 2.54 \\
+    & 1.0 & 3.41 & 3.66 & 4.02 & 2.90 & 2.55 \\
+    \hline
+  \end{tabular}
+  \caption{Tail indices of the distribution function of
+    $\lambda_\M$. By tail index $\alpha$, I mean $\lim_{x \to \infty}
+    x^\alpha P(\lambda_\M > x) = \text{const}$}
+  \label{tab:tail_indices}
+\end{table}
 
@@
   127]{AppleLogvol-QQ.pdf}
-  \caption{QQ-plot of $\ln{ S_\M - S_\m \over S_\m}$. The data are daily trading statistics of Apple from
-    2005-01-02 to 2015-01-23. The lower tail of sample log-vol is a
-    bit lighter and the upper tail a bit heavier than their
-    counterparts of standard normal. }
+  \caption{QQ-plot of $\ln{ S_\M - S_\m \over S_\m}$ against standard
+    normal. Here $S_\M$ and $S_\m$ stand for the highest and the
+    lowest price of each trading day, respectively. The data are daily
+    trading statistics of Apple Inc. from 2005-01-02 to 2015-01-23. The
+    lower tail of sample log-vol is a bit lighter and the upper tail a
+    bit heavier than their counterparts of standard normal.}
   \label{fig:qqplot_logvol}

### 2015-03-20 cv/XieXiaolei-Analyst.tex
Add the revised CV.

@@
 \section*{My Goal}
-A self-funded Ph.D student of Finance \& Insurance mathematics
-at KU, who is very experienced in programming and statistical analysis, is
-looking for a part-time job as a quantitative analyst.
+A Ph.D student of Economic \& Insurance mathematics at Copenhagen
+University, who is very experienced in programming and statistical
+analysis, is looking for a part-time job as a quantitative analyst.
 
@@
       reduction, by which the behaviour of a diverse portfolio is
-      replicated by one with much few components and hence analysis of
-      the portfolio becomes more simplified and focused.
+      replicated by one with much fewer components and hence analysis
+      of the portfolio becomes more simplified and focused.
 

### 2015-03-28 cv/XieXiaolei-Analyst.tex
Revise my CV as an analyst.

--- 2015-02-09 cv/XieXiaolei-Analyst.tex
@@
 
-\section*{My Goal}
-A self-funded Ph.D student of Finance \& Insurance mathematics
-at KU, who is very experienced in programming and statistical analysis, is
-looking for a part-time job as a quantitative analyst.
+% \section*{My Goal}
+% A PhD student of Finance \& Insurance Mathematics at Copenhagen
+% University, who is very experienced in programming and statistical
+% analysis,
+% % is looking for a part-time job as a quantitative analyst.
+% is looking for a summer internship as a quantitative analyst.
 
@@
 \begin{itemize}
-\item Ph.D of Finance \& Insurance Mathematics, October 2014 - present
+\item PhD of Finance \& Insurance Mathematics, Copenhagen University,
+  October 2014 - present
   \begin{itemize}
-    \item {\it Reseach Areas:} I am mainly interested in the
+    \item {\it Research Areas:} I am mainly interested in the
       determination of the largest eigenvalues of a covariance matrix
       of samples with heavy-tails. The purpose is dimension
-      reduction, by which the behaviour of a diverse portfolio is
-      replicated by one with much few components and hence analysis of
-      the portfolio becomes more simplified and focused.
+      reduction, by which the behavior of a large portfolio is
+      replicated by one with much fewer components and hence analysis of
+      the portfolio can be more simplified and focused.
 
@@
 
-\section*{Language Skills}
-\begin{itemize}
-\item English: Fluent
-\item Swedish: Fairly good
-\item Danish: Beginner
-\end{itemize}
+\section*{Programming Skills}
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{|l|c|| l |c|}
+    \hline
+    Skill & Experience (years) & Skill & Experience (years) \\
+    \hline
+    Matlab & 4.5 & R & 0.5 \\
+    C & 10 & C++ & 3 \\
+    SQL & 3 & Java & 3 \\
+    Python & 1 &&\\
+    \hline
+  \end{tabular}
+\end{table}
+
+% \section*{Language Skills}
+% \begin{itemize}
+% \item English: Fluent
+% \item Swedish: Fairly good
+% \item Danish: Beginner
+% \end{itemize}
 

### 2015-03-31 cv/XieXiaolei-Analyst.tex

@@
 
-% \section*{My Goal}
+\section*{My Goal}
 % A PhD student of Finance \& Insurance Mathematics at Copenhagen
-% University, who is very experienced in programming and statistical
-% analysis,
-% % is looking for a part-time job as a quantitative analyst.
+A PhD of Finance \& Insurance Mathematics at Copenhagen
+University, who is very experienced in programming and statistical
+analysis,
+% is looking for a part-time job as a quantitative analyst.
 % is looking for a summer internship as a quantitative analyst.
+is looking for a role of quantitative analysis.
 
@@
       multi-dimensional data set.
+
+      \item{\it Teaching:} I teach exercise classes of the
+        master-level course ``Statistical Methods of Non-life
+        insurance'', which includes Maximum-likelihood estimators,
+        hypothesis testing and linear regression, among others. I
+        first solve the problems myself and then present and explain the
+        solutions to the class upon the students' request.
+
   \end{itemize}

### 2015-05-04 papers/InequalityProof.tex
Separate the two documents.

\documentclass{article}
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}

\input{../physics_common}
\begin{document}
\section{Proof of the Lemma}
Here we prove that, for a symmetric, positive semidefinite $p \times
p$ matrix $\mtx M = (m_{ij})$, whose $k$th principle sub-matrix is denoted
$M_k$, satisfies
\[
\lambda_1(\mtx M) \leq \lambda_1(M_{k}) + \sum_{i=k+1}^p m_{ii}
\]
\begin{proof}
  Because $M_k$ is positive semidefinite for each $k$ , we only need
  to prove the claim for $k=p-1$. For notational convenience, write
  $A$ for $M_{p-1}$ and $d$ for $m_{pp}$. Next we prove the claim when
  $A$ is positive definite.
  \begin{eqnarray*}
    \mtx M &=&
    \begin{pmatrix}
      A & v \\
      v' & d
    \end{pmatrix} \\
    &=&
    \begin{pmatrix}
      A & 0 \\
      v' & d - v' A^{-1} v
    \end{pmatrix}\left(
      \begin{array}{cc}
        I & -A^{-1} v \\
        0 & 1
      \end{array}\right)^{-1} \\
    &=& U V
  \end{eqnarray*}
  From $\det(M) \geq 0$ it follows $\det(A) (d - v' A^{-1} v) \geq
  0$. Since $\det(A) > 0$, $d - v' A^{-1} v \geq 0$. Moreover, $A$ is
  assumed positive definite, so $v' A^{-1} v \geq 0$. According
  to Bai and Silverstein\cite{BaiSilverstein2010} we have
  $\lambda_1(M) \leq s_1(U)$, where $s_1(U)$ denotes the largest
  singular value of $U$.
  \begin{eqnarray*}
    \lambda_1(M) &\leq& s_1(U) \\
    &=& \max\left\{\lambda_1(A), d - v' A^{-1} v \right\} \\
    &\leq& \lambda_1(A) + d
  \end{eqnarray*}
  When $A$ is not positive definite, i.e. $\det(A)=0$,
  semi-positiveness implies $\det(A) d - v' v \geq 0$, hence $v =
  0$. It immediately follows
  \begin{eqnarray*}
    M &=&
    \begin{pmatrix}
      A & 0 \\
      0 & d
    \end{pmatrix} \\
    \lambda_1(M) &=& \max\{\lambda_1(A), d\} \\
    &\leq& \lambda_1(A) + d
  \end{eqnarray*}
\end{proof}
\bibliographystyle{unsrt}
\bibliography{../thesis/econophysics}
\end{document}

### 2015-05-11 papers/InequalityProof.tex

@@
   \end{eqnarray*}
-  When $A$ is not positive definite, i.e. $\det(A)=0$,
-  semi-positiveness implies $\det(A) d - v' v \geq 0$, hence $v =
-  0$. It immediately follows
+  Now we consider the case where $ d > 0$ and $A$ is not positive
+  definite, i.e. it contains zero eigenvalues. When $A$ has only zero
+  eigenvalues, our claim is clearly true; when $A$ has at least 1
+  non-zero eigenvalue, we may apply elementary transforms to $M$,
+  or in other words, multiply $M$ by an upper triangular matrix $T$
+  and its transpose $T'$
   \begin{eqnarray*}
-    M &=&
+    T &=& 
     \begin{pmatrix}
-      A & 0 \\
-      0 & d
+      1 & T_{12} & \cdots & T_{1,p} \\
+      0 & 1 & \cdots & T_{2, p} \\
+      \vdots & \vdots & \ddots & \vdots \\
+      0 & 0 & \cdots & 1 \\
+    \end{pmatrix}
+  \end{eqnarray*}
+  so that
+  \begin{eqnarray*}
+    T M T'&=&
+    \begin{pmatrix}
+      0 & 0 & u_1 \\
+      0 & A_2 & u_2 \\
+      u'_1 & u'_2 & d
     \end{pmatrix} \\
-    \lambda_1(M) &=& \max\{\lambda_1(A), d\} \\
-    &\leq& \lambda_1(A) + d
+    M  &=& T^{-1} M_1 T'^{-1}
   \end{eqnarray*}
+  where $A_2$ is positive definite and has the same rank as
+  A. Obviously $A_2$ has the same non-zero eigenvalues as $A$.
+  Because $M$ is positive semi-definite, so is $M_1$. Then it follows
+  $|M_{1;ij}| \leq \sqrt{M_{1;i,i} M_{1;j,j}}$ for all $1 \leq i,j
+  \leq p$, implying $u_1 = 0$. Now we have
+  \begin{eqnarray*}
+    M &=& T^{-1} M_1 T'^{-1} \\
+    &=& T^{-1}
+    \begin{pmatrix}
+      0 & 0 & 0 \\
+      0 & A_2 & u_2 \\
+      0 & u'_2 & d
+    \end{pmatrix} T'^{-1} \\
+  \end{eqnarray*}
+  By the same inequality found in \cite{BaiSilverstein2010},
+  \begin{eqnarray*}
+    \lambda_1(M) &\leq& s_1(T^{-1}) \lambda_1(M_1) s_1(T'^{-1}) \\
+    &=& \lambda_1(M_1) \\
+    &=& \lambda_1 \left[\left(
+      \begin{array}{cc}
+        A_2 & u_2 \\
+        u'_2 & d
+      \end{array} \right)\right] \\
+  &\leq& \lambda_1(A_2) + d
+  \end{eqnarray*}
+  where in the last inequality we apply the claim proven in the case
+  of positive definite principle sub-matrix.
 \end{proof}

### 2015-05-12 papers/InequalityProof.tex
Scripts of StatArb and revised papers.

@@
 \begin{document}
-\section{Proof of the Lemma}
+\section{Proof of the 1st Lemma}
 Here we prove that, for a symmetric, positive semidefinite $p \times
@@
   \end{eqnarray*}
-  so that
+  and in addition by an orthogonal matrix $Q$, and its transpose $Q'$,
+  where $Q$ is constructed by permuting the rows of the identity
+  matrix that has the same rank as $A$. Clearly $Q=Q'$ and $Q^2 = I$
+  and hence is orthogonal. Note the eigenvalues of $Q$ are either 1 or
+  -1.
+
+  The matrix $T$ uses a set of linearly independent row vectors of $A$
+  to eliminate the rest of the row vectors; $Q$ collects this set of
+  linearly independent vectors into a positive definite matrix that we
+  denote as $A_2$. By formulae
   \begin{eqnarray*}
-    T M T'&=&
+    QTMT'Q &=&
     \begin{pmatrix}
@@
       u'_1 & u'_2 & d
-    \end{pmatrix} \\
-    M  &=& T^{-1} M_1 T'^{-1}
+    \end{pmatrix} = M_1 \\
+    M  &=& T^{-1} Q M_1 Q T'^{-1}
   \end{eqnarray*}
@@
   \begin{eqnarray*}
-    M &=& T^{-1} M_1 T'^{-1} \\
-    &=& T^{-1}
+    M &=& T^{-1} Q M_1 Q T'^{-1} \\
+    &=& T^{-1} Q
     \begin{pmatrix}
@@
       0 & u'_2 & d
-    \end{pmatrix} T'^{-1} \\
+    \end{pmatrix} Q T'^{-1} \\
   \end{eqnarray*}
@@
   \begin{eqnarray*}
-    \lambda_1(M) &\leq& s_1(T^{-1}) \lambda_1(M_1) s_1(T'^{-1}) \\
+    \lambda_1(M) &\leq& s_1(Q)^2 s_1(T^{-1}) \lambda_1(M_1)
+    s_1(T'^{-1}) \\
     &=& \lambda_1(M_1) \\
@@
 \end{proof}
+
+\section{Proof of the 2nd Lemma}
+Here we prove that, for a symmetric, positive semi-definite $p \times
+p$ matrix
+\[
+M =
+\begin{pmatrix}
+  A & V \\
+  V' & B
+\end{pmatrix}
+\]
+we have
+\[
+\lambda_1(M) \leq \lambda_1(A) + \lambda_1(B) + 2
+\]
+\begin{proof}
+  Write $M$ as
+  \[
+  M =
+  \begin{pmatrix}
+    A & 0 \\
+    0 & B
+  \end{pmatrix} +
+  \begin{pmatrix}
+    0 & V \\
+    V' & 0
+  \end{pmatrix}
+  \]
+  Then by Weyle's inequality we have
+  \begin{eqnarray*}
+    \lambda_1(M) \leq \max\{\lambda_1(A), \lambda_1(B)\} +
+    s_1\left[\left(
+      \begin{array}{cc}
+        0 & V \\
+        V' & 0
+      \end{array}
+      \right)\right]
+  \end{eqnarray*}
+  But
+  \[
+  \begin{pmatrix}
+    0 & V \\
+    V' & 0
+  \end{pmatrix} =
+  \begin{pmatrix}
+    I & 0 \\
+    V' & I
+  \end{pmatrix} +
+  \begin{pmatrix}
+    I & V \\
+    0 & I
+  \end{pmatrix} - 2I
+  \]
+\end{proof}
 \bibliographystyle{unsrt}

### 2015-05-12 m/test3.m

@@ -15,40 +15,51 @@ spec = cellstr(['b  '; 'c  '; 'g  '; 'm  '; 'r  '; 'k  ';...
                 ]);
 texts = {};
 c = 1;
-% hold on
-lam1 = NaN(length(sig), length(phi), 2);
-lam2 = NaN(length(sig), length(phi), 2);
-variance = NaN(length(sig), length(phi));
-
-phi = 0.8160;
+parameters = NaN(4, 4);
+f = @(param, X) param(1) .* exp(param(2) .* X);
 for q = [0.1, 0.2, 0.5, 1]
+    variance = NaN(length(sig), length(phi));
+    B = NaN(length(sig), length(phi), length(q));
     for m = 1 : length(sig)
-        % for sig = 0.1
-        v = sig(m)^2/(1 - phi^2);
-        load(sprintf(['../matfys/data/sv/normal_ret/lognormal_vol/q%.1f/Eig-' ...
-                      'sig%.4f-phi%.4f.mat'], q, sig(m), phi), 'ev');
-        % fprintf('N=%d, T=%d, ');
-        % ev = reshape(ev, 1, prod(size(ev)));
-        eigmax = max(ev)';
-        [y, x] = ecdf(eigmax);
-        I = x > quantile(eigmax, 0.3) & x < quantile(eigmax, 0.85);
-        subplot(4, 3, c);
-        loglog(x(I), 1 - y(I), 'LineWidth', 2);
-        title(sprintf('q = %.1f, v=%.4f', q, v));
-        grid on
-        c = c + 1;
-        % eigmin = min(ev)';
-        
-        % lam1(m, n, :) = [mean(eigmin), (1-sqrt(q))^2 * ...
-        %                  (1 - 2*v*(2*sqrt(q)-1))];
-        % lam2(m, n, :) = [mean(eigmax), (1+sqrt(q))^2 * ...
-        %                  (1 + 2*v*(2*sqrt(q)+1))];
+        for n = 1 : length(phi)
+            v = sig(m)^2/(1 - phi(n)^2);
+            variance(m, n) = v;
+            load(sprintf(['../matfys/data/sv/normal_ret/lognormal_vol/q%.1f/Eig-' ...
+                          'sig%.4f-phi%.4f.mat'], q, sig(m), phi(n)), 'ev');
+            % fprintf('N=%d, T=%d, ');
[…]

### 2015-05-22 m/StatArb.m
Revised the paper and the matlab scripts.

@@ -7,8 +7,8 @@ p = size(R, 2);
 
 C = R' * R / T;
 [portfolios, D] = eig(C);
-% X = R * portfolios(:, end-1:end);
-X = R * portfolios(:, 20);
+X = R * portfolios(:, end-1);
+% X = R * portfolios(:, 20);
 % Y = ts_difference(X(:, 2), [10, 1; 1, 1]);
 Y = ts_difference(X, [10, 1; 1, 1]);
 % X = R * ones(p, 1);

### 2015-07-06 r/FixedDimEig.r
The first version of the paper.

@@ -2,7 +2,7 @@ rm(list=ls());
 library(FMStable)
 
 alpha <- 0.8;
-ns <- seq(2200, 5400, by=800);
+ns <- seq(1000, 2000, by=200);
 supremum <- rep(NA, length(ns));
 
 c <- 1;
@@ -10,6 +10,8 @@ for (n in ns) {
     p <- 5*floor(log(n));
     load(sprintf("dat/P-5xlogN_Dist-t/lambda.1.alpha%.1f_n%d.dat", alpha, n));
     X <- sort(lambda.1);
+    pars <- setMomentsFMstable(mean=0, sd=1, alpha=alpha/2);
+    F <- pEstable(X, pars)^p;
     supremum[c] <- max(c(abs(F[-p] - seq(1, p-1) / p), abs(F[-1] - seq(1, p-1) / p)));
     c <- c + 1;
 }

### 2015-07-06 papers/Number1/Applications.tex
Revised paper #1 and the R script.

\subsection{Rank-transformed S\&P 500 Data}

For real financial data, it is unreasonable to expect all
the time series under consideration to have the same tail index. So a
preocedure of standardization has to be adopted. In this section we
use the rank transform.

Suppose we have $p$ time series each of length $n$ and have arranged
them in a $p\times n$ matrix $X_t^{(i)}$, where $i = 1, 2, \dots, p$ and
$t=1,2, \dots, n$. For each row of $X$, we transform it as
\begin{eqnarray*}
  R_t^{(i)} &=& -\left[
    \log \left({1 \over n+1} \sum_{\tau=1}^n \1_{X_\tau^{(i)} \leq X_{t}^{(i)}} \right)
  \right]^{-1}
\end{eqnarray*}
When $X_t^{(i)}$ are iid, the argument to the $\log$ function is
uniformly distributed, and hence $R_t^{(i)}$ is a standard Fr\'echet r.v.

Figure \ref{fig:EigenRatio} shows the ratio of successive eigenvalues
of the sample covariance matrix of a number of selected time series
included in the S\&P 500 index.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6]{EigenRatio.pdf}
  \caption{$\lambda_{(i+1)} / \lambda_{(i)}$ versus $i$}
  \label{fig:EigenRatio}
\end{figure}

### 2015-07-07 papers/Number1/Applications.tex
Revised the paper with Richard.

@@
 Suppose we have $p$ time series each of length $n$ and have arranged
-them in a $p\times n$ matrix $X_t^{(i)}$, where $i = 1, 2, \dots, p$ and
+them in a $p\times n$ matrix $X_{it}$, where $i = 1, 2, \dots, p$ and
 $t=1,2, \dots, n$. For each row of $X$, we transform it as
 \begin{eqnarray*}
-  R_t^{(i)} &=& -\left[
-    \log \left({1 \over n+1} \sum_{\tau=1}^n \1_{X_\tau^{(i)} \leq X_{t}^{(i)}} \right)
+  R_{it} &=& -\left[
+    \log \left({1 \over n+1} \sum_{\tau=1}^n \1_{X_{i\tau} \leq X_{it}} \right)
   \right]^{-1}
 \end{eqnarray*}
-When $X_t^{(i)}$ are iid, the argument to the $\log$ function is
-uniformly distributed, and hence $R_t^{(i)}$ is a standard Fr\'echet r.v.
+When $X_{it}$ are iid for each $i$, the argument to the $\log$
+function is asymptotically uniformly distributed as $n \to \infty$,
+and hence $R_{it}$ is a standard Fr\'echet r.v.
 
 Figure \ref{fig:EigenRatio} shows the ratio of successive eigenvalues
-of the sample covariance matrix of a number of selected time series
-included in the S\&P 500 index.
+of the matrix $a_{np}^{-2}RR'$, where the rows of matrix $R$ comprise
+the rank-transformed sequences of a number of selected time series
+in the S\&P 500 index.
 \begin{figure}[htb!]
   \centering
-  \includegraphics[scale=0.6]{EigenRatio.pdf}
+  \subfigure[]{
+    \includegraphics[scale=0.40]{EigenRatioSP500_100_shown.pdf}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.40]{EigenRatioSP500_200_shown.pdf}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.40]{EigenRatioSP500_300_shown.pdf}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.40]{EigenRatioSP500_400_shown.pdf}
+  }
   \caption{$\lambda_{(i+1)} / \lambda_{(i)}$ versus $i$}
@@
 \end{figure}
+It is seen that, as $i \to \infty$, $\lambda_{(i+1)} / \lambda_{(i)}
+\to 1$. Indeed, as argued in previous sections, $ \lambda_{(i)} \to
+\Gamma_k^{-2/\alpha} v_{(l)}$.  Now consider $\lambda_{(i+1)}$. One
+possibility is $\lambda_{(i+1)} \to \Gamma_{k+1}^{-2/\alpha}
+v_{(l)}$. In this case
+\begin{eqnarray*}
+  {\lambda_{i+1} \over \lambda_{i}} &\to& \left(
+    \Gamma_{i} \over \Gamma_{i+1} \right)^{2/\alpha}
+  \overset{d}{=} U_{1,i} \overset{a.s.}{\longrightarrow} 1\text{, as }i
+  \to \infty
+\end{eqnarray*}
+where $U_{a,b}$ denotes the $a$-th upper order statistic in a uniform
+sample of size $b$.
+So $ \lambda_{i+1} \to \Gamma_{k+1}^{-2/\alpha} v_{(l)}$ can be untrue
+only if $\exists m\exists l'(m,l') \neq (1, l)$ such that $ \lambda_{i+1} \to
+\Gamma_{k+m}^{-2/\alpha} v_{(l')}$ and
+\begin{equation}
+  \label{eq:X_A}
+  \Gamma_{k+m}^{-2/\alpha} v_{(l')} \overset{a.s.}{\longrightarrow}
+  \Gamma_{k}^{-2/\alpha} v_{(l)}
+\end{equation}
+Suppose $m > 1$.
+\[
+\left(
+  \Gamma_{k} \over \Gamma_{k+m}
+\right)^{2/\alpha} \overset{d}{=} (U_{m,k+m-1})^{2/\alpha}
+\]
+The variance of $(U_{m,k+m-1})^{2/\alpha}$ remains finite regardless
+of $k$ and $m$ (\textcolor{red}{to be proven}), so the almost sure
+convergence in \eqref{eq:X_A} cannot happen.

### 2015-07-15 papers/Number1/Applications.tex
Correct the justification of the ratio of successive eigenvalues converging to 1.

@@
 It is seen that, as $i \to \infty$, $\lambda_{(i+1)} / \lambda_{(i)}
-\to 1$. Indeed, as argued in previous sections, $ \lambda_{(i)} \to
-\Gamma_k^{-2/\alpha} v_{(l)}$.  Now consider $\lambda_{(i+1)}$. One
-possibility is $\lambda_{(i+1)} \to \Gamma_{k+1}^{-2/\alpha}
-v_{(l)}$. In this case
-\begin{eqnarray*}
-  {\lambda_{i+1} \over \lambda_{i}} &\to& \left(
-    \Gamma_{i} \over \Gamma_{i+1} \right)^{2/\alpha}
-  \overset{d}{=} U_{1,i} \overset{a.s.}{\longrightarrow} 1\text{, as }i
-  \to \infty
-\end{eqnarray*}
-where $U_{a,b}$ denotes the $a$-th upper order statistic in a uniform
-sample of size $b$.
-So $ \lambda_{i+1} \to \Gamma_{k+1}^{-2/\alpha} v_{(l)}$ can be untrue
-only if $\exists m\exists l'(m,l') \neq (1, l)$ such that $ \lambda_{i+1} \to
-\Gamma_{k+m}^{-2/\alpha} v_{(l')}$ and
-\begin{equation}
-  \label{eq:X_A}
-  \Gamma_{k+m}^{-2/\alpha} v_{(l')} \overset{a.s.}{\longrightarrow}
-  \Gamma_{k}^{-2/\alpha} v_{(l)}
-\end{equation}
-Suppose $m > 1$.
-\[
-\left(
-  \Gamma_{k} \over \Gamma_{k+m}
-\right)^{2/\alpha} \overset{d}{=} (U_{m,k+m-1})^{2/\alpha}
-\]
-The variance of $(U_{m,k+m-1})^{2/\alpha}$ remains finite regardless
-of $k$ and $m$ (\textcolor{red}{to be proven}), so the almost sure
-convergence in \eqref{eq:X_A} cannot happen.
+\to 1$.
+This is in fact expected based on the theory explained in the previous sections.
+Suppose the sequences are iid and serial correlations are absent, i.e.
+the matrix $M=H H'$ has only one eigen value.
+Then it is clear ${\lambda_{i+1} \over \lambda_i}=\left({\Gamma_i \over \Gamma_{i+1}}\right)^{2/\alpha}=\left(1 - {E_{i+1} \over E_1 + E_2 + \dots E_{i+1}}\right)^{2/\alpha} \to 1$, where $E_k, k=1,2,\dots,i+1$ are iid.
+exponential rv.
+% Indeed, as argued in previous sections, $ \lambda_{(i)} \to
+% \Gamma_k^{-2/\alpha} v_{(l)}$.  Now consider $\lambda_{(i+1)}$. One
+% possibility is $\lambda_{(i+1)} \to \Gamma_{k+1}^{-2/\alpha}
+% v_{(l)}$. In this case
+% \begin{eqnarray*}
+%   {\lambda_{i+1} \over \lambda_{i}} &\to& \left(
+%     \Gamma_{i} \over \Gamma_{i+1} \right)^{2/\alpha}
+%   \overset{d}{=} U_{1,i} \overset{a.s.}{\longrightarrow} 1\text{, as }i
+%   \to \infty
+% \end{eqnarray*}
+% where $U_{a,b}$ denotes the $a$-th upper order statistic in a uniform
+% sample of size $b$.
+% So $ \lambda_{i+1} \to \Gamma_{k+1}^{-2/\alpha} v_{(l)}$ can be untrue
+% only if $\exists m\exists l'(m,l') \neq (1, l)$ such that $ \lambda_{i+1} \to
+% \Gamma_{k+m}^{-2/\alpha} v_{(l')}$ and
+% \begin{equation}
+%   \label{eq:X_A}
+%   \Gamma_{k+m}^{-2/\alpha} v_{(l')} \overset{a.s.}{\longrightarrow}
+%   \Gamma_{k}^{-2/\alpha} v_{(l)}
+% \end{equation}
+% Suppose $m > 1$.
+% \[
+% \left(
+%   \Gamma_{k} \over \Gamma_{k+m}
+% \right)^{2/\alpha} \overset{d}{=} (U_{m,k+m-1})^{2/\alpha}
+% \]
+% The variance of $(U_{m,k+m-1})^{2/\alpha}$ remains finite regardless
+% of $k$ and $m$ (\textcolor{red}{to be proven}), so the almost sure
+% convergence in \eqref{eq:X_A} cannot happen.

### 2015-08-06 papers/Number1/Applications.tex
Revised the paper with Johannes.

--- 2015-07-07 papers/Number1/Applications.tex
@@
 It is seen that, as $i \to \infty$, $\lambda_{(i+1)} / \lambda_{(i)}
-\to 1$. Indeed, as argued in previous sections, $ \lambda_{(i)} \to
+\to 1$.
+% This is in fact expected based on the theory explained in the
+% previous sections. Suppose the sequences are iid and serial
+% correlations are absent, i.e. the matrix $M=H H'$ has only one eigen
+% value. Then it is clear ${\lambda_{(i+1)} \over
+%   \lambda_i}=\left({\Gamma_i \over
+%     \Gamma_{i+1}}\right)^{2/\alpha}=\left(1 - {E_{i+1} \over E_1 + E_2
+%     + \dots E_{i+1}}\right)^{2/\alpha} \to 1$, where $E_k,
+% k=1,2,\dots,i+1$ are iid. exponential rv.
+Indeed, as argued in previous sections, $ \lambda_{(i)} \overset{d}{\to}
 \Gamma_k^{-2/\alpha} v_{(l)}$.  Now consider $\lambda_{(i+1)}$. One
-possibility is $\lambda_{(i+1)} \to \Gamma_{k+1}^{-2/\alpha}
+possibility is $\lambda_{(i+1)} \overset{d}{\to} \Gamma_{k+1}^{-2/\alpha}
 v_{(l)}$. In this case
 \begin{eqnarray*}
-  {\lambda_{i+1} \over \lambda_{i}} &\to& \left(
+  {\lambda_{(i+1)} \over \lambda_{i}} &\to& \left(
     \Gamma_{i} \over \Gamma_{i+1} \right)^{2/\alpha}
-  \overset{d}{=} U_{1,i} \overset{a.s.}{\longrightarrow} 1\text{, as }i
-  \to \infty
+  \overset{d}{=} (U_{1,i})^{2/\alpha} \overset{a.s.}{\longrightarrow}
+  1\text{, as }i \to \infty
 \end{eqnarray*}
 where $U_{a,b}$ denotes the $a$-th upper order statistic in a uniform
-sample of size $b$.
-So $ \lambda_{i+1} \to \Gamma_{k+1}^{-2/\alpha} v_{(l)}$ can be untrue
-only if $\exists m\exists l'(m,l') \neq (1, l)$ such that $ \lambda_{i+1} \to
+sample of size $b$. In words, $\lambda_{(i+1)}$ gets infinitely close to
+$\lambda_{(i)}$ in this case. For this reason, any other candidate for
+$\lambda_{(i+1)}$ must also be infinitely close to $\lambda_{(i)}$ and
+thereby infinitely close to $\Gamma_{k+1}^{-2/\alpha} v_{(l)}$.
+
+So $ \lambda_{(i+1)} \overset{d}{\to} \Gamma_{k+1}^{-2/\alpha}
+v_{(l)}$ can be untrue only if $\exists m\exists l'\,(m,l') \neq (1,
+l)$ such that $ \lambda_{(i+1)} \overset{d}{\to}
 \Gamma_{k+m}^{-2/\alpha} v_{(l')}$ and
@@
   \label{eq:X_A}
-  \Gamma_{k+m}^{-2/\alpha} v_{(l')} \overset{a.s.}{\longrightarrow}
-  \Gamma_{k}^{-2/\alpha} v_{(l)}
+  \P\left(\left|
+    {\Gamma_{k+m}^{-2/\alpha} v_{(l')} \over \Gamma_{k}^{-2/\alpha}
+      v_{(l)}} - 1\right| > \epsilon
+  \right) \overset{d}{\to} 0 \text{ as } i \to \infty
 \end{equation}
@@
 The variance of $(U_{m,k+m-1})^{2/\alpha}$ remains finite regardless
-of $k$ and $m$ (\textcolor{red}{to be proven}), so the almost sure
-convergence in \eqref{eq:X_A} cannot happen.
+of $k$ and $m$ (\textcolor{red}{to be proven}), so the convergence in
+\eqref{eq:X_A} cannot happen.

### 2015-08-07 papers/Number1/paper.tex
Add database scripts and revised the paper draft.

%last edited 21.07.2015
\documentclass[11pt,reqno]{amsart}
\usepackage{amsmath,epsfig,graphicx,color}
%\usepackage[english,ngerman]{babel}
%\usepackage[latin1]{inputenc}
\usepackage{ifthen}
\usepackage{dsfont}
\usepackage{shadethm}
\usepackage{framed}
%\usepackage{pstricks}
%\usepackage{graphics}
%\usepackage{tocbibind}
%\usepackage{showkeys}
%\usepackage{units}
\usepackage{amssymb,latexsym}
%\usepackage{amsfonts}
%\usepackage{makeidx,showidx}
%\usepackage[sc]{mathpazo}
%\linespread{1.05}

%\usepackage[backref=page]{hyperref}
\usepackage{hyperref}
\hypersetup{
%linktoc=page,
%bookmarks=true, % show bookmarks bar?
%unicode=false, % non-Latin characters in Acrobat? bookmarks
%pdftoolbar=true, % show Acrobat? toolbar?
%pdfmenubar=true, % show Acrobat? menu?
%pdffitwindow=true, % window fit to page when opened
%pdfstartview={FitH}, % fits the width of the page to the window
pdftitle={Paper}, % title
pdfauthor={Johannes Heiny}, % author
pdfsubject={Title}, % subject of the document
%pdfcreator={Creator}, % creator of the document
%pvdfproducer={Producer}, % producer of the document
pdfkeywords={multivariate regular variation} , % list of keywords
%pdfnewwindow=true, % links in new window
%colorlinks=true, % false: boxed links; true: colored links
%linkcolor=blue, % color of internal links
%citecolor=blue, % color of links to bibliography
%filecolor=blue, % color of file links
%urlcolor=blue, % color of external links
urlcolor=black, 
  menucolor=black, 
  citecolor=black, 
  anchorcolor=black, 
  filecolor=black, 
  linkcolor=black, 
  colorlinks=true,
}

\textwidth 6.50in
\topmargin -0.50in
\oddsidemargin 0in
\evensidemargin 0in
\textheight 9.00in
%\pagestyle{plain}

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\1}{\mathds{1}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\Fo}{\bar{F}}
\renewcommand{\b}[1]{\boldsymbol{#1}}
\newcommand{\Rq}{\mkern 1.5mu\overline{\mkern-1.5mu\R\mkern-3.0mu}\mkern 1.5mu}
\newcommand{\Frechet}{Fr\'{e}chet }
\renewcommand{\Finv}{F^{\gets}}
\DeclareMathOperator{\e}{e}
\newcommand{\inv}[1]{#1^{\gets}}
\newcommand{\x}{\boldsymbol{x}}
\newcommand{\y}{\boldsymbol{y}}
\newcommand{\X}{\boldsymbol{X}}
\newcommand{\Y}{\boldsymbol{Y}}
\newcommand{\bfZ}{\boldsymbol{Z}}
\newcommand{\0}{\boldsymbol{0}}
\newcommand{\dint}{\,\mathrm{d}}
\newcommand{\mB}{\mathcal{B}}
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\twonorm}[1]{\|#1\|_2}
\newcommand{\inftynorm}[1]{\|#1\|_\infty}
\newcommand{\vep}{\varepsilon}
\newcommand{\nto}{n \to \infty}
\newcommand{\xto}{x \to \infty}
\newcommand{\lhs}{left-hand side}
\newcommand{\rhs}{right-hand side}
\newcommand{\fidi}{finite-dimensional distribution}
\newcommand{\rv}{random variable}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\anp}{a_{np}^{-2}}
\newcommand{\Zbar}{\overline{Z}}
%\renewcommand*{\backref}[1]{}
%\renewcommand*{\backrefalt}[4]{%
%    \ifcase #1 (Not cited.)%
%    \or        (Cited on page~#2.)%
%    \else      (Cited on pages~#2.)%
%    \fi}

%von Schmock
\newcommand{\4}{\mathchoice{\mskip1.5mu}{\mskip1.5mu}{}{}}
\newcommand{\5}{\mathchoice{\mskip-1.5mu}{\mskip-1.5mu}{}{}}
\newcommand{\2}{\penalty250\mskip\thickmuskip\mskip-\thinmuskip} % after comma

\newcommand{\levy}{L\'evy}
\newcommand{\slln}{strong law of large numbers}
\newcommand{\clt}{central limit theorem}
\newcommand{\sde}{stochastic differential equation}
\newcommand{\It}{It\^o}
\newcommand{\sta}{St\u aric\u a}
\newcommand{\ex}{{\rm e}\,}

\def\theequation{\thesection.\arabic{equation}}
\def\tag{\refstepcounter{equation}\leqno }
\def\neqno{\refstepcounter{equation}\leqno(\thesection.\arabic{equation})}

\newtheorem{lemma}{Lemma}[section]
\newtheorem{figur}[lemma]{Figure}
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{proposition}[lemma]{Proposition}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{example}[lemma]{Example}
\newtheorem{exercise}[lemma]{Exercise}
\newtheorem{remark}[lemma]{Remark}
\newtheorem{fig}[lemma]{Figure}
\newtheorem{tab}[lemma]{Table}
\newtheorem{conjecture}[lemma]{Conjecture}

\newcommand{\cid}{\stackrel{d}{\rightarrow}}
\newcommand{\cip}{\stackrel{\P}{\rightarrow}}

\newcommand{\var}{{\rm var}}
\newcommand{\med}{{\rm med}}
\newcommand{\cov}{{\rm cov}}
\newcommand{\corr}{{\rm corr}}
\newcommand{\as}{{\rm a.s.}}
\newcommand{\io}{{\rm i.o.}}
\newcommand{\Holder}{H\"older}

\newcommand{\cadlag}{c\`adl\`ag}

%--------------------------------------------------------------------------------------------
\begin{document}
\today
\bibliographystyle{plain}
\title[The sample covariance matrix of a heavy-tailed multivariate time series]{Recent advances in random matrix theory applied to heavy-tailed multivariate time series}

\begin{abstract}

\end{abstract}

\maketitle
\tableofcontents

%------------------------------------------------------------------------
\section{Introduction}\setcounter{equation}{0}

\subsection{Motivation: Lam + Yao, PCA, Hill plot of upper/lower tails of S\&P data}
\subsection{Overview of light-tailed case}
\subsection{Finite $p$: Anderson}
\subsection{Growing $p$: Johnstone, Tao + Vu}

%------------------------------------------------------------------------
\section{The iid heavy-tailed case}

\subsection{$\alpha \in (0,2)$: Overview of existing results: Davis et al., Soshnikov, Auffiner et al.}
\subsection{$\alpha \in (0,2)$: Our extension of these results; only order statistics of $Z_{it}$'s necessary and no sums (proof by the technical result)}
\subsection{$\alpha \in [2,4)$: Overview of existing results: Davis et al., Soshnikov, Auffiner et al.}
\subsection{$\alpha \in [2,4)$: The problem of centering and restrictions on $p_n$}
\subsection{$\alpha \in [2,10/3)$: Our new result for the UNCENTERED sample covariance matrix; only order statistics of $Z_{it}$'s necessary and no sums (proof by the technical result)}
\subsection{$\alpha \in [10/3,4)$: A conjecture with partial proof for the UNCENTERED sample covariance matrix)}

%------------------------------------------------------------------------
\section{Extension to linear dependence, two-sided filters}

\subsection{$\alpha \in (0,10/3)$: Eigenvalues of the sample covariance matrix}
\subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample covariance matrix}
\subsection{$\alpha \in [2,4)$ and $p/n \to \gamma \in (0,1]$: Eigenvalues of the sample covariance matrix using Auffinger's result}
\subsection{Point process convergence}

%------------------------------------------------------------------------
\section{Applications of the results in the previous section}

\subsection{Ratio of second to largest eigenvalue (to power become uniform), self-normalizing sums, ratio of largest eigenvalue to trace, simulations, figures,...}
\subsection{Functions acting on the eigenvalues}
\subsection{Figures: Ratios of consecutive eigenvalues, illustrations for financial data, ...}
\subsection{S\&P data using ranks}

%------------------------------------------------------------------------
\section{A further extension: The sample autocovariance function}

\subsection{$\alpha \in (0,10/3)$: Eigenvalues of the sample autocovariance matrix}
\subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample autocovariance matrix}

%------------------------------------------------------------------------
\section{Applications of the results for the autocovariance function}

\subsection{Sums of largest eigenvalues vs.~ largest eigenvalues of sums, application to real data + plots}
\subsection{Explanation of these plots in the separable case, discussion of the two methods to ensure real eigenvalues, link to Lam + Yao}

%------------------------------------------------------------------------
\section{Non-linear theory, independent rows + large deviations, SV and GARCH, SV with multivariate Gaussian and t-distributed noise}
%------------------------------------------------------------------------
\section{Future directions, discussion of the conjecture}

%------------------------------------------------------------------------
\appendix 
\section{Regular variation, large deviations and point processes}
\section{The key technical result to identify the largest eigenvalues with the order statistics of the squares of the matrix entries in the iid case}
\section{A lemma which shows that the sample covariance matrix is approximated by its diagonal in the iid case (more involved than simple matrix norms)}
\section{The Proofs}

%-----------------------------------------------------------------------
%Bibliography
%Bibliography
\begin{thebibliography}{99}
\baselineskip12pt
%AAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAA
\bibitem{anderson:1963}
{\sc Anderson, T.W.,}\ (1963)
{\em Asymptotic theory for principal component analysis.}  {\em Ann.~Math.~Statist.} {\bf 34}, 122--148.
%\bibitem{anderson:guionnet:zeitouni:2008}
%{\sc Anderson, G.W., Guionnet, A. and Zeitouni, O.}\ (2008)
%{\em An Introduction to Random Matrices.}  Cambridge
%University Press, Cambridge (UK).

\bibitem{auffinger:arous:peche:2009}
{\sc Auffinger, A., Ben Arous, G. and P\'ech\'e, S.}\ (2009)
Poisson convergence for the largest eigenvalues of heavy tailed random
matrices.
{\em Ann. Inst. H. Poincar\'e, Probab. Statist.} {\bf 45}, 589--610.
%BBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBB
%\bibitem{bai:silverstein:2010}
%{\sc Bai, Z. and Silverstein, J.W.}\ (2010)
%{\em Spectral Analysis of Large Dimensional Random Matrices.} 2nd
%Edition. Springer, New York.
%\bibitem{belinschi:dembo:guionnet:2009}
%{\sc Belinschi, S.,  Dembo, A. and Guionnet, A.}\ (2009)
%Spectral measure of heavy tailed band and covariance random matrices.
%{\em Comm. Math. Phys.}  {\bf 289}, 1023--1055.
%\bibitem{arous:guionnet:2008}
%{\sc  Ben Arous, G. and  Guionnet, A.} \ (2008)
%The spectrum of heavy tailed random matrices.  {\em Comm. Math. Phys.}
%{\bf 278}, 715--751.
\bibitem{bhatia:1997}
{\sc Bhatia, R.}\ (1997)
{\em Matrix Analysis.} Springer, New York.
\bibitem{bingham:goldie:teugels:1987}
{\sc Bingham, N.H., Goldie, C.M.\ and Teugels, J.L.}\ (1987)
{\em Regular Variation.} Cambridge University Press, Cambridge.
%\bibitem{bose:hazra:saha:2009}
%{\sc Bose, A., Subhra Hazra, R. and  Saha, K.}\ (2009)
%Limiting spectral distribution of circulant type matrices
%with dependent inputs.  {\em Electron. J. Probab.}  {\bf 14},
%2463--2491.
%\bibitem{bose:hazra:saha:2010}
%{\sc Bose, A., Subhra Hazra, R. and  Saha, K.}\ (2010)
%Spectral norm of circulant type matrices with heavy tailed entries.
%{\em Electron. Commun. Probab.}  {\bf 15}, 299--313.
%CCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCCC
\bibitem{cline:hsing:1998}
{\sc Cline, D.B.H. and Hsing, T.}\ (1998)
Large deviation probabilities for sums of random variables
with heavy or subexponential tails. Technical report.
Statistics Dept., Texas A\&M University. %Available as \begin{verbatim}http://www.stat.tamu.edu/~dcline/Papers/large5.pdf \end{verbatim}
%DDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDDD
\bibitem{davis:resnick:1985}
{\sc
Davis, R.A. and Resnick, S.I.}\ (1985)
Limit theory for moving averages of random variables
with regularly varying tail probabilities.
{\em Ann. Probab.} {\bf 13}, 179--195.
\bibitem{davis:pfaffel:stelzer:2014}
{\sc Davis, R.A., Pfaffel, O. and Stelzer, R.}\ (2014)
Limit theory for the largest eigenvalues of
sample covariance matrices with heavy tails.
{\em Stoch. Proc. Appl.} {\bf 124}, 18--50.

\bibitem{denisov:dieker:shneer:2008}
{\sc Denisov, D., Dieker, A.B. and Shneer, V.}\ (2008)
Large deviations for random walks under subexponentiality:
the big-jump domain.
{\em Ann. Probab.} {\bf 36}, 1946--1991.

%EEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEE

\bibitem{embrechts:goldie:1980}
{\sc Embrechts, P. and Goldie,C.}\ (1980)
On closure and factorization theorems for subexponential and related
distributions. {\em J. Austral. Math. Soc.,} Ser. A {\bf 29}, 243--256.

%\bibitem{embrechts:kluppelberg:mikosch:1997}
%{\sc Embrechts, P., Kl\"uppelberg, C. and Mikosch, T.}\ (1997)
%{\em Modelling Extremal Events for Insurance and Finance.}
%Springer,  Berlin.

%GGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGGG

\bibitem{gohberg:goldberg:krupnik:2000}
{\sc Gohberg, I., Goldberg, S. and N. Krupnik}\ (2000)
{\em Traces and Determinants of Linear Operators.} Birkh\"auser, Basel.

\bibitem{johnstone:2001}
{\sc Johnstone, I.M.}\ (2001)
{\em on the distribution of the largest eigenvalue in principal component analysis.} {\em Ann.~Statist.} {\bf 29} (2), 295--327.

%MMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMMM

\bibitem{mikosch:samorodnitsky:2000}
{\sc Mikosch, T. and Samorodnitsky, G.} \ (2000)
The supremum of a negative drift random walk with
dependent heavy-tailed steps. {\em Ann. Appl. Probab.} {\bf 10}, 1025--1064.

%NNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNN

\bibitem{nagaev:1979}
{\sc Nagaev, S.V.}\ (1979)
Large deviations of sums of independent random
variables.
{\em Ann. Probab.} {\bf 7}, 745--789.

%PPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPPP

\bibitem{petrov:1995}
{\sc Petrov, V.V.} (1995)
{\em Limit Theorems of Probability Theory.} Oxford University Press,
Oxford.

%RRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRRR

\bibitem{resnick:1987}
{\sc Resnick, S.I.}\ (1987)
{\em Extreme Values, Regular Variation, and Point Processes.}
Sprin\-ger, New York.

\bibitem{resnick:2007}
{\sc Resnick, S.I.}\ (2007)
{\em Heavy-Tail Phenomena: Probabilistic and Statistical Modeling.}
Springer, New York.

%SSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSSS

\bibitem{samorodnitsky:taqqu:1994}
{\sc Samorodnitsky, G. and Taqqu, M.S.}\ (1994)
{\em Stable Non-Gaussian Random Processes.
Stochastic Models with Infinite Variance.} Chapman and Hall, London.

%\bibitem{soshnikov:2004}
%{\sc Soshnikov, A.}\ (2004)
%Poisson statistics for the largest eigenvalues of Wigner matrices with
%heavy tails. {\em Electr. Comm. Probab.} {\bf 9}, 82--91.

\bibitem{soshnikov:2006}
{\sc Soshnikov, A.}\ (2006)
Poisson statistics for the largest eigenvalues in random matrix
ensembles.
In: {\em Mathematical Physics of Quantum Mechanics.} Lecture Notes in
Physics {\bf 690}. Springer, Berlin, pp. 351--364.

\bibitem{tao:vu:2012}
{\sc Tao, T. and Vu, V.}\ (2012)
Random covariance matrices; universality of local statistics of eigenvalues.
{\em Ann.~Probab.} {\bf 40} (3), 1285--1315.

\end{thebibliography}
\end{document}

### 2015-08-07 db_scripts/load_from_yahoo.py

@@ -1,0 +1,68 @@
+#!/usr/bin/python
+
+import os
+import sys
+import _mysql
+import csv
+
+# conn = _mysql.connect(host=os.environ.get("PB"),
+#                      user="sinbaski",
+#                      passwd="q1w2e3r4",
+#                      db="avanza");
+conn = _mysql.connect(host="localhost",
+                     user="sinbaski",
+                     passwd="q1w2e3r4",
+                     db="avanza");
+index = "SP500";
+suffix = "_US";
+conn.query("select symbol from {0}_components where available=1;".format(index));
+
+
+Symbols = conn.store_result();
+while True:
+    symbol = Symbols.fetch_row(maxrows=1, how=0);
+    if not symbol:
+        break;
+
+    tbl = symbol[0][0].replace(".", "_");
+    tbl = tbl.replace("-", "_series_");
+    tbl += suffix;
+    stmt = 'insert into {0} values'.format(tbl)
+    csv_file_name = '/tmp/{0}.csv'.format(symbol[0][0]);
+    
+    conn.query('select max(day) as t from {0};'.format(tbl));
+    X = conn.use_result().fetch_row(maxrows = 1);
+    # last_date = X[0][0];
+    # conn.query('delete from {0};'.format(tbl));
+
+    num = 0;
+    with open(csv_file_name, 'rb') as csvfile:
+        my_reader = csv.reader(csvfile)
+        c = 0;
+        for row in my_reader:
+            c += 1;
+            # Skip the header
+            if c == 1 or row[2] == row[3]:
+                continue
+            elif X != None and row[0] <= X[0][0]:
+                break;
+            num += 1;
+            if num == 1:
+                stmt += ' ("{0}", {1}, {2}, {3}, {4})'.format(
+                    row[0], row[2], row[3], row[4], row[5]);
+            else:
+                stmt += ', ("{0}", {1}, {2}, {3}, {4})'.format(
+                    row[0], row[2], row[3], row[4], row[5]);
+    print '%d rows to insert into %s.' % (num, tbl);
+    if (num):
+        # stmt += ' on duplicate key update ';
+        # stmt += 'high=values(high), low=values(low),';
+        # stmt += 'closing=values(closing), volume=values(volume)';
+        stmt += ';';
+        logfile = open('/tmp/sql/%s.sql' % symbol[0][0], 'wb');
+        logfile.write('%s\n' % stmt);
+        logfile.close();
+        conn.query(stmt);
+    
+conn.close();
+

### 2015-08-07 db_scripts/load_from_yahoo.sh

@@ -1,0 +1,47 @@
+#!/bin/bash
+
+symbols=`echo "select symbol from SP500_components where available=1;" | mysql -N -u"sinbaski" -p"q1w2e3r4" -h localhost avanza`;
+# symbols=`echo "select symbol from OMXS30_components;" | mysql -N -u"sinbaski" -p"q1w2e3r4" -h localhost avanza`;
+# symbols=`echo "select symbol from indices;" | mysql -N -u"sinbaski" -p"q1w2e3r4" -h localhost avanza`;
+
+for symbol in `echo $symbols`; do
+    # if test -f ~/Downloads/$symbol.csv; then
+    # 	rm ~/Downloads/$symbol.csv
+    # fi
+    if [ -z "$str" ]; then str=$symbol; else str="$str,$symbol"; fi
+done
+curl http://real-chart.finance.yahoo.com/table.csv?s={"$str"} -o /tmp/#1.csv
+if [ ! -d /tmp/sql ]; then
+    mkdir /tmp/sql
+fi
+# python ./load_from_yahoo.py
+    # tbl=`echo $symbol | sed -e 's:\.:_:g'`;
+    # tbl=`echo $symbol | sed -e 's:\.:_:g' | sed -r -e 's:$:_US:g'`;
+#     echo "drop table if exists $tbl;" > /tmp/$symbol.txt;
+#     echo "create table if not exists $tbl (
+# day date primary key,
+# high decimal(7,2) not null,
+# low decimal(7,2) not null,
+# closing decimal(7,2) not null,
+# volume int
+# );" >> /tmp/$symbol.txt;
+
+#     if [ -f /tmp/$symbol.txt ]; then rm /tmp/$symbol.txt; fi
+#     echo "insert into $tbl values " >> /tmp/$symbol.txt;
+#     tail -n +2 ~/Downloads/$symbol.csv |
+#     head -n 7 |
+#     cut -f 1,3,4,5,6 -d',' |
+#     awk -F, '
+# {
+# if (FNR == 1) printf "(\"%s\", %.2f, %.2f, %.2f, %d)", $1,$2,$3,$4,$5
+# else printf ",\n(\"%s\", %.2f, %.2f, %.2f, %d)", $1,$2,$3,$4,$5
+# }
+# END {print ";"}' >> /tmp/$symbol.txt;
+
+
+    # printf "drop table if exists %s;" $symbol > /tmp/$symbol.txt
[…]

### 2015-08-08 papers/Number1/paper.tex
Added plot of estimated tail indices. Revised the R scripts.

@@
 %\usepackage{units}
+\usepackage{subfigure}
 \usepackage{amssymb,latexsym}
@@
 
-\subsection{Motivation: Lam + Yao, PCA, Hill plot of upper/lower tails of S\&P data}
+\subsection{Motivation: Lam + Yao, PCA, indices of upper/lower tails of S\&P
+  components}
+Figure \ref{fig:SP500_tail_indices} shows the estimated lower- and upper-tail indices of stocks
+included in the S\&P 500 index.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5]{SP500_tail_indices.pdf}
+  \caption{Tail indices of stocks included in the S\&P 500 index. The tail
+    indices are obtained as their Hill estimators. The red line is the line with equal lower- and upper
+    tail indices.}
+  \label{fig:SP500_tail_indices}
+\end{figure}
+
 \subsection{Overview of light-tailed case}
@@
 \subsection{S\&P data using ranks}
+For real financial data, it is unreasonable to expect all
+the time series under consideration to have the same tail index. So a
+preocedure of standardization has to be adopted. In this section we
+use the rank transform.
+
+Suppose we have $p$ time series each of length $n$ and have arranged
+them in a $p\times n$ matrix $X_{it}$, where $i = 1, 2, \dots, p$ and
+$t=1,2, \dots, n$. For each row of $X$, we transform it as
+\begin{eqnarray*}
+  R_{it} &=& -\left[
+    \log \left({1 \over n+1} \sum_{\tau=1}^n \1_{X_{i\tau} \leq X_{it}} \right)
+  \right]^{-1}
+\end{eqnarray*}
+When $X_{it}$ are iid for each $i$, the argument to the $\log$
+function is asymptotically uniformly distributed as $n \to \infty$,
+and hence $R_{it}$ is a standard Fr\'echet r.v.
+
+Figure \ref{fig:EigenRatio} shows the ratio of successive eigenvalues
+of the matrix $a_{np}^{-2}RR'$, where the rows of matrix $R$ comprise
+the rank-transformed sequences of a number of selected time series
+in the S\&P 500 index.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[]{
+    \includegraphics[scale=0.40]{EigenRatioSP500_100_shown.pdf}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.40]{EigenRatioSP500_200_shown.pdf}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.40]{EigenRatioSP500_300_shown.pdf}
+  }
+  \subfigure[]{
+    \includegraphics[scale=0.40]{EigenRatioSP500_400_shown.pdf}
+  }
+  \caption{$\lambda_{(i+1)} / \lambda_{(i)}$ versus $i$}
+  \label{fig:EigenRatio}
+\end{figure}
+It is seen that, as $i \to \infty$, $\lambda_{(i+1)} / \lambda_{(i)}
+\to 1$. This is in fact expected based on the theory explained in the
+previous sections. Suppose the sequences are iid and serial
+correlations are absent, i.e. the matrix $M=H H'$ has only one eigen
+value. Then it is clear ${\lambda_{(i+1)} \over
+  \lambda_i}=\left({\Gamma_i \over
+    \Gamma_{i+1}}\right)^{2/\alpha}=\left(1 - {E_{i+1} \over E_1 + E_2
+    + \dots E_{i+1}}\right)^{2/\alpha} \to 1$, where $E_k,
+k=1,2,\dots,i+1$ are iid. exponential rv.
 

### 2015-08-09 papers/Number1/paper.tex
Revised the paper.

@@
 \newcommand{\Zbar}{\overline{Z}}
+\newcommand{\M}{\text{max}}
 %\renewcommand*{\backref}[1]{}
@@
 \subsection{Overview of light-tailed case}
+Yin, Bai and Krishnaiah proved in their 1988 paper
+\cite{yin:bai:Krishnaiah:1988} the followng result for a $p \times n$
+matrix $\bf W$ with iid entries $w_{ij}$ under the assumption $\E
+w_{11} = 0$ and $\E w_{11}^4 < \infty$.
+\[
+\lim_{n \to \infty} \lambda_{\M}\left({1 \over n}\bf{WW'}\right) = \left(
+  1 + \sqrt{p \over n}
+\right)^2 \E w_{11}
+\]
+
 \subsection{Finite $p$: Anderson}
@@
 
+\bibitem{yin:bai:Krishnaiah:1988}
+{\sc Y.Q. Yin and Z.D. Bai and P.R. Krishnaiah}\ (1988)
+On the Limit of the Largest Eigenvalue of the Large Dimensional Sample
+Covariance Matrix.
+{\em Probability Theory and Related Fields} {\bf 78}, 509--521.
+
 \end{thebibliography}

### 2015-08-10 papers/Number1/paper.tex
Revised the review part.

@@
 \cite{yin:bai:Krishnaiah:1988} the followng result for a $p \times n$
-matrix $\bf W$ with iid entries $w_{ij}$ under the assumption $\E
-w_{11} = 0$ and $\E w_{11}^4 < \infty$.
+matrix $\bf X$ with iid entries $X_{ij}$ under the assumption $\E
+X_{11} = 0$ and $\E x_{11}^4 < \infty$.
 \[
-\lim_{n \to \infty} \lambda_{\M}\left({1 \over n}\bf{WW'}\right) = \left(
+\lim_{n \to \infty} \lambda_{\M}\left({1 \over n}\bf{XX'}\right) = \left(
   1 + \sqrt{p \over n}
-\right)^2 \E w_{11}
+\right)^2 \E x_{11}
+\]
+Later on, several authors \cite{Pillai:Yin:2011, tao:vu:2012,
+  Wang:2012} looked into empirical covariance matrix $S_{n, p} = {1
+  \over n} \bf{XX}'$, where the entries of $\bf X$, denoted $x_{ij}$
+are independent and have an exponential decay: $\exists B_1, B_2 > 0$
+such that
+\begin{equation}
+  \label{eq:C0}
+  \forall i \in \{1, 2, \dots, p\}\text{, }
+  \forall j \in \{1, 2, \dots, n\}\text{, }
+  \P(|x_{ij}| > t^{B_1}) \leq e^{-t}
+\end{equation}
+for all $t > B_2$. Pillai and Yin \cite{Pillai:Yin:2011} proved that,
+given condition \eqref{eq:C0} and in additon $1 < n/p < \infty$,
+the empirical eigenvalues are localized to their theoretical
+values: $\exists c, C > 0$ such that $\forall 1 \leq j \leq p$
+\[
+\P(\left|
+  \lambda_{(j)} - \gamma_{j}^{n,p}
+\right|) \geq  (\log p)
 \]
@@
 \subsection{$\alpha \in (0,2)$: Overview of existing results: Davis et al., Soshnikov, Auffiner et al.}
-\subsection{$\alpha \in (0,2)$: Our extension of these results; only order statistics of $Z_{it}$'s necessary and no sums (proof by the technical result)}
+\subsection{$\alpha \in (0,2)$: Our extension of these results; only
+  order statistics of $Z_{it}$'s necessary and no sums (proof by the
+  technical result)}
 \subsection{$\alpha \in [2,4)$: Overview of existing results: Davis et al., Soshnikov, Auffiner et al.}
 \subsection{$\alpha \in [2,4)$: The problem of centering and restrictions on $p_n$}
-\subsection{$\alpha \in [2,10/3)$: Our new result for the UNCENTERED sample covariance matrix; only order statistics of $Z_{it}$'s necessary and no sums (proof by the technical result)}
-\subsection{$\alpha \in [10/3,4)$: A conjecture with partial proof for the UNCENTERED sample covariance matrix)}
+\subsection{$\alpha \in [2,10/3)$: Our new result for the UNCENTERED
+  sample covariance matrix; only order statistics of $Z_{it}$'s
+  necessary and no sums (proof by the technical result)}
+\subsection{$\alpha \in [10/3,4)$: A conjecture with partial proof for
+  the UNCENTERED sample covariance matrix)}
 
@@
 \subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample covariance matrix}
-\subsection{$\alpha \in [2,4)$ and $p/n \to \gamma \in (0,1]$: Eigenvalues of the sample covariance matrix using Auffinger's result}
+\subsection{$\alpha \in [2,4)$ and $p/n \to \gamma \in (0,1]$:
+  Eigenvalues of the sample covariance matrix using Auffinger's
+  result}
 \subsection{Point process convergence}
@@
 
-\subsection{Sums of largest eigenvalues vs.~ largest eigenvalues of sums, application to real data + plots}
-\subsection{Explanation of these plots in the separable case, discussion of the two methods to ensure real eigenvalues, link to Lam + Yao}
+\subsection{Sums of largest eigenvalues vs.~ largest eigenvalues of
+  sums, application to real data + plots}
+\subsection{Explanation of these plots in the separable case,
+  discussion of the two methods to ensure real eigenvalues, link to
+  Lam + Yao}
 
 %------------------------------------------------------------------------
-\section{Non-linear theory, independent rows + large deviations, SV and GARCH, SV with multivariate Gaussian and t-distributed noise}
+\section{Non-linear theory, independent rows + large deviations, SV
+  and GARCH, SV with multivariate Gaussian and t-distributed noise}
 %------------------------------------------------------------------------
@@
 \section{Regular variation, large deviations and point processes}
-\section{The key technical result to identify the largest eigenvalues with the order statistics of the squares of the matrix entries in the iid case}
-\section{A lemma which shows that the sample covariance matrix is approximated by its diagonal in the iid case (more involved than simple matrix norms)}
+\section{The key technical result to identify the largest eigenvalues
+  with the order statistics of the squares of the matrix entries in
+  the iid case}
+\section{A lemma which shows that the sample covariance matrix is
+  approximated by its diagonal in the iid case (more involved than
+  simple matrix norms)}
 \section{The Proofs}
@@
 
+\bibitem{Pillai:Yin:2011}
+{\sc N. Pillai and J. Yin}\ (2011)
+Universality of covariance matrices
+{\em arXiv:1110.2501,2011}
+
+%K. Wang. Random covariance matrices: universality of local statistics of
+%eigenvalues up to the edge. RMTA, 1(1), 2012.
+\bibitem{Wang:2012}
+{\sc K. Wang}\ (2012)
+Random covariance matrices: universality of local statistics of
+eigenvalues up to the edge
+{\em RMTA} {\bf 1} (1)
+
 \end{thebibliography}

### 2015-08-11 papers/Number1/paper.tex
Revised the review part.

--- 2015-08-09 papers/Number1/paper.tex
@@
 %\usepackage{units}
+\usepackage{mathrsfs}
+\usepackage{enumerate}
 \usepackage{subfigure}
@@
 \newcommand{\M}{\text{max}}
+\newcommand{\m}{\text{min}}
 %\renewcommand*{\backref}[1]{}
@@
   \caption{Tail indices of stocks included in the S\&P 500 index. The tail
-    indices are obtained as their Hill estimators. The red line is the line with equal lower- and upper
-    tail indices.}
+    indices are obtained as their Hill estimators. The red line is the
+    line with equal lower- and upper tail indices.}
   \label{fig:SP500_tail_indices}
 \end{figure}
+Clearly the number of
 
@@
 \cite{yin:bai:Krishnaiah:1988} the followng result for a $p \times n$
-matrix $\bf W$ with iid entries $w_{ij}$ under the assumption $\E
-w_{11} = 0$ and $\E w_{11}^4 < \infty$.
+matrix $\bf X$ with iid entries $X_{ij}$ under the assumption $\E
+X_{11} = 0$, $\E X_{11}^4 < \infty$, and $q = \lim_{n \to \infty} p/n
+> 0$
 \[
-\lim_{n \to \infty} \lambda_{\M}\left({1 \over n}\bf{WW'}\right) = \left(
-  1 + \sqrt{p \over n}
-\right)^2 \E w_{11}
+\lambda_{\M}\left({1 \over n}\bf{XX'}\right) \overset{\as}{\to} \left(
+  1 + \sqrt{q}
+\right)^2 \E X_{11}^2 \text{ as }n \to \infty
 \]
+When $\E X_{11} = \infty$, Yin and Silverstein proved in 1988
+\cite{yin:silverstein:1988}
+\[
+\lim_{n \to \infty} \sup \lambda_{\M}\left({1 \over
+    n}\bf{XX'}\right) = \infty
+\]
+Later on, several authors \cite{Pillai:Yin:2011, tao:vu:2012,
+  Wang:2012} looked into empirical covariance matrix $S_{n, p} = {1
+  \over n} \bf{XX}'$, where the entries of $\bf X$, denoted $X_{ij}$
+are independent and satisfy the following condition which is refered
+to as condition {\bf C0}.
+\begin{enumerate}
+\item $\E X_{ij} = 0$, $\E X_{ij}^2 = 1$ and
+\item $\forall u > 1, \exists \theta > 0$ such that
+  \[
+  \P(X_{ij} > u) = \theta^{-1} e^{-u^{\theta}}
+  \]
+\end{enumerate}
+Under condition {\bf C0}, Pillai and Yin \cite{Pillai:Yin:2011} proved
+that, for $q = \lim_{n \to \infty} {p \over n} \in (0, \infty)
+\setminus \{1\}$, $\forall \zeta > 0, \exists C > 0, \exists \delta >
+0$,
+\[
+\P\left[
+  \left|
+    \lambda_{(j)}(S_{n,p}) - \gamma_{j}^{n, p}
+  \right| > \varphi^{C} p^{-2/3} \tilde{j}^{-1/3}
+\right] \leq p^{\delta} \exp(-\varphi^\zeta)
+\]
+where $\lambda_{(j)}(S_{n,p})$ is the $j$-th largest eigenvalue of
+$S_{n,p}$ and
+\begin{eqnarray*}
+  \varphi &=& (\log p)^{\log \log p} \\
+  \tilde j &=& \min\{(p \wedge n) + 1 - j, j\} \\
+  \gamma_{j}^{n, p} &=& \text{inf}\left\{\gamma: \int_\gamma^{\lambda_\M}
+  \mu_{n,p}(x) dx \leq {j \over p}\right\} \\
+  \lambda_\M &=& (1 + \sqrt q)^2 \\
+  \lambda_\m &=& (1 - \sqrt q)^2 \\
+  \mu_{n,p}(x) &=& {1 \over 2\pi q x} \sqrt{(x - \lambda_\m)(\lambda_\M - x)}
+\end{eqnarray*}
+
+Tao and Vu \cite{tao:vu:2012} relaxed the expoential decay condition
+of {\bf C0} and replaced it with a high-moment condition, and proved the
+four-moment theorem, which was later extended by Wang
+\cite{Wang:2012}. In their setting, a $p \times n$ matrix $\bf X$ is
+said to obey condition {\bf C1} with constant $C_0$ if $\forall 1
+\leq i \leq p, \forall 1 \leq j \leq n$, $\E X_{ij} = 0$, $\E
+X_{ij}^2 = 1 $ and $\sup_{i,j} \E|X_{ij}|^{C_0} < C$ for some constant
+$C$ independent of $p$ and $n$.
+
+Assume matrices $\bf X$ and $\bf Y$ both satisfy condition {\bf C1}
+with constant $C_0$ and in addition they match up to order $4$,
+i.e. for all $m, l \geq 0$ such that $m + l \leq 4$, $\E[\mathscr{
+R}(X_{ij})^m \mathscr{I}(X_{ij})^l] = \E[\mathscr{R}(Y_{ij})^m
+\mathscr{I}(Y_{ij})^l]$. Assume also $\lim_{n \to \infty} p/n = q \in
+(0, 1]$. Let $G: \mathbb R^k \to \mathbb R$ be a smooth function
+obeying the derivative bounds
+\[
+\left| \nabla^j G(x) \right| < n^{c_0}
+\]
+for some $c_0 > 0$. Then the following holds for all $1 \leq i_1 \leq
+i_2 \leq \cdots \leq i_{k} \leq p$, $1 \leq k \leq p$:
+\[
+\left|
+\E G(n\lambda_{i_1}, \dots, n\lambda_{i_k}) - 
+\E G(n\omega_{i_1}, \dots, n\omega_{i_k})
+\right| < n^{- c_0}
+\]
+where $\lambda_1 \leq \lambda_2, \leq \cdots \leq \lambda_p$ are the
+eigenvalues of ${1 \over n} XX'$ and $\omega_1 \leq \omega_2, \leq
+\cdots \leq \omega_p$ are those of ${1 \over n} YY'$.
 
@@
 
-\subsection{$\alpha \in (0,2)$: Overview of existing results: Davis et al., Soshnikov, Auffiner et al.}
-\subsection{$\alpha \in (0,2)$: Our extension of these results; only order statistics of $Z_{it}$'s necessary and no sums (proof by the technical result)}
-\subsection{$\alpha \in [2,4)$: Overview of existing results: Davis et al., Soshnikov, Auffiner et al.}
+\subsection{$\alpha \in (0,2)$: Overview of existing results: Davis et
+  al., Soshnikov, Auffiner et al.}
+\subsection{$\alpha \in (0,2)$: Our extension of these results; only
+  order statistics of $Z_{it}$'s necessary and no sums (proof by the
+  technical result)}
+\subsection{$\alpha \in [2,4)$: Overview of existing results: Davis et
+  al., Soshnikov, Auffiner et al.} 
 \subsection{$\alpha \in [2,4)$: The problem of centering and restrictions on $p_n$}
-\subsection{$\alpha \in [2,10/3)$: Our new result for the UNCENTERED sample covariance matrix; only order statistics of $Z_{it}$'s necessary and no sums (proof by the technical result)}
-\subsection{$\alpha \in [10/3,4)$: A conjecture with partial proof for the UNCENTERED sample covariance matrix)}
+\subsection{$\alpha \in [2,10/3)$: Our new result for the UNCENTERED
+  sample covariance matrix; only order statistics of $Z_{it}$'s
+  necessary and no sums (proof by the technical result)}
+\subsection{$\alpha \in [10/3,4)$: A conjecture with partial proof for
+  the UNCENTERED sample covariance matrix)}
 
@@
 
+\bibitem{yin:bai:Krishnaiah:1988}
+{\sc Y.Q. Yin and Z.D. Bai and P.R. Krishnaiah}\ (1988)
+On the Limit of the Largest Eigenvalue of the Large Dimensional Sample
+Covariance Matrix.
+{\em Probability Theory and Related Fields} {\bf 78}, 509--521.
+
+\bibitem{yin:silverstein:1988}
+{\sc Y.Q. Yin and Jack W. Silverstein}\ (1988)
+A Note on the Largest Eigenvalue of a Large Dimensional Sample
+Covariance Matrix
+{\em Journal of Multivariate Analysis} {\bf 26} (2), 166--168.
+
 \bibitem{resnick:2007}
@@
 
+\bibitem{Pillai:Yin:2011}
+{\sc N. Pillai and J. Yin}\ (2011)
+Universality of covariance matrices
+{\em arXiv:1110.2501,2011}
+
 \bibitem{tao:vu:2012}
@@
 
-\bibitem{yin:bai:Krishnaiah:1988}
-{\sc Y.Q. Yin and Z.D. Bai and P.R. Krishnaiah}\ (1988)
-On the Limit of the Largest Eigenvalue of the Large Dimensional Sample
-Covariance Matrix.
-{\em Probability Theory and Related Fields} {\bf 78}, 509--521.
+%K. Wang. Random covariance matrices: universality of local statistics of
+%eigenvalues up to the edge. RMTA, 1(1), 2012.
+\bibitem{Wang:2012}
+{\sc K. Wang}\ (2012)
+Random covariance matrices: universality of local statistics of
+eigenvalues up to the edge
+{\em RMTA} {\bf 1} (1)
 

### 2015-08-12 papers/Number1/paper.tex
Revised the paper.

@@
   components}
-Figure \ref{fig:SP500_tail_indices} shows the estimated lower- and upper-tail indices of stocks
-included in the S\&P 500 index.
+As reviewed in \S\ref{sec:light_tail_review}, much of the literature
+so far is focused on sample covariance matrices with a light-tailed atomic
+distribution, which in particular implies finite 4th moment. However,
+this is untrue when real financial data such as stock returns are
+concerned. Figure \ref{fig:SP500_tail_indices} shows the estimated
+lower- and upper-tail indices of stocks included in the S\&P 500 index.
 \begin{figure}[htb!]
@@
 \end{figure}
-Clearly the number of
+It is clear that a large number of stocks have a tail index smaller
+than 4, implying a possibly divergent 4th moment. Thus it is of great
+importance to study situations where the atomic distribution is
+heavy-tailed. This is the topic of the current paper. In particular,
+we look into sample covariance matrices with an atomic distribution
+that has regularly varying tails.
 
-\subsection{Overview of light-tailed case}
+\subsection{Overview of light-tailed case}\label{sec:light_tail_review}
 Yin, Bai and Krishnaiah proved in their 1988 paper

### 2015-08-12 papers/Number1/paper.tex
Corrected some typos.

@@
 Yin, Bai and Krishnaiah proved in their 1988 paper
-\cite{yin:bai:Krishnaiah:1988} the followng result for a $p \times n$
+\cite{yin:bai:Krishnaiah:1988} the following result for a $p \times n$
 matrix $\bf X$ with iid entries $X_{ij}$ under the assumption $\E
@@
   \over n} \bf{XX}'$, where the entries of $\bf X$, denoted $X_{ij}$
-are independent and satisfy the following condition which is refered
+are independent and satisfy the following condition which is referred
 to as condition {\bf C0}.
@@
 
-Tao and Vu \cite{tao:vu:2012} relaxed the expoential decay condition
+Tao and Vu \cite{tao:vu:2012} relaxed the exponential decay condition
 of {\bf C0} and replaced it with a high-moment condition, and proved the
@@
 \subsection{$\alpha \in [2,4)$: The problem of centering and restrictions on $p_n$}
-\subsection{$\alpha \in [2,10/3)$: Our new result for the UNCENTERED
+\subsection{$\alpha \in [2,10/3)$: Our new result for the UN-CENTERED
   sample covariance matrix; only order statistics of $Z_{it}$'s
@@
 \subsection{$\alpha \in [10/3,4)$: A conjecture with partial proof for
-  the UNCENTERED sample covariance matrix)}
+  the UN-CENTERED sample covariance matrix)}
 
@@
 the time series under consideration to have the same tail index. So a
-preocedure of standardization has to be adopted. In this section we
+procedure of standardization has to be adopted. In this section we
 use the rank transform.
@@
 previous sections. Suppose the sequences are iid and serial
-correlations are absent, i.e. the matrix $M=H H'$ has only one eigen
-value. Then it is clear ${\lambda_{(i+1)} \over
+correlations are absent, i.e. the matrix $M=H H'$ has only one
+eigenvalue. Then it is clear ${\lambda_{(i+1)} \over
   \lambda_i}=\left({\Gamma_i \over

### 2015-08-13 papers/Number1/paper.tex
Added confidence bands to the plot of eigenvalue ratios.

@@
 \newcommand{\inftynorm}[1]{\|#1\|_\infty}
+\newcommand{\pp}[1]{\mathscr E_{#1}}
 \newcommand{\vep}{\varepsilon}
@@
 \newcommand{\m}{\text{min}}
+
 %\renewcommand*{\backref}[1]{}
@@
 \end{eqnarray*}
-When $X_{it}$ are iid for each $i$, the argument to the $\log$
-function is asymptotically uniformly distributed as $n \to \infty$,
-and hence $R_{it}$ is a standard Fr\'echet r.v.
+When $X_{i1}, X_{i2}, \dots, X_{in}$ are iid for each $i$, the
+argument to the $\log$ function is asymptotically uniformly
+distributed as $n \to \infty$, and hence $R_{it}$ is a standard
+Fr\'echet r.v.
 
@@
   }
-  \caption{$\lambda_{(i+1)} / \lambda_{(i)}$ versus $i$}
+  \caption{$\lambda_{(i+1)} / \lambda_{(i)}$ versus $i$, The blue line
+  is the expected value of the ratio, the red line is the 5\% quantile
+  and the green line is the 95\% quantile.}
   \label{fig:EigenRatio}
@@
 \to 1$. This is in fact expected based on the theory explained in the
-previous sections. Suppose the sequences are iid and serial
-correlations are absent, i.e. the matrix $M=H H'$ has only one
-eigenvalue. Then it is clear ${\lambda_{(i+1)} \over
-  \lambda_i}=\left({\Gamma_i \over
-    \Gamma_{i+1}}\right)^{2/\alpha}=\left(1 - {E_{i+1} \over E_1 + E_2
-    + \dots E_{i+1}}\right)^{2/\alpha} \to 1$, where $E_k,
-k=1,2,\dots,i+1$ are iid. exponential rv.
+previous sections. Suppose the coefficients $h_{kl}$ is separable,
+i.e. $h_{kl} = \theta_k c_l$, then, as shown in previous sections,
+\[
+\sum_{i=1}^p \pp{a_{np}^{-2}\lambda_{(i)}} \overset{d}{\to}
+\sum_{i=1}^\infty \pp{\Gamma_i^{-2} \lambda_H}
+\]
+where $\lambda_H$ is the solo non-zero eigenvalue of $\bf H$. Note
+that the tail index $\alpha=1$ for the ranks. So it is clear
+\[
+{\lambda_{(i+1)} \over 
+  \lambda_{(i)}} \overset{d}{=} \left({\Gamma_i \over
+    \Gamma_{i+1}}\right)^{2}=\left(1 - {E_{i+1} \over E_1 + E_2
+    + \dots E_{i+1}}\right)^{2} \to 1 \text{ as } i \to \infty
+\]
+where $E_k, k=1,2,\dots,i+1$ are iid. exponential rv. In fact the
+distribution function of $\lambda_{(i+1)} / \lambda_{(i)}$ is rather
+simple. Recall
+\[
+{\Gamma_{i} \over \Gamma_{i+1}} \overset{d}{=} U_{1, i}
+\]
+where $U_{1, i}$ is the largest upper order statistic of a
+$\text{Uniform}(0, 1)$ sample of size $i$. So
+\begin{eqnarray*}
+\P\left[{\lambda_{(i+1)} \over \lambda_{(i)}} \leq x \right] &=& x^{i/2} \\
+\E \left[{\lambda_{(i+1)} \over \lambda_{(i)}} \right] &=& {i \over i + 2}
+\end{eqnarray*}
+The expected values and confident intervals marked in figure
+\ref{fig:EigenRatio} are calculated according to the above formulas.
 
@@
 
-\subsection{$\alpha \in (0,10/3)$: Eigenvalues of the sample autocovariance matrix}
-\subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample autocovariance matrix}
+\subsection{$\alpha \in (0,10/3)$: Eigenvalues of the sample
+  autocovariance matrix} 
+\subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample
+  autocovariance matrix}
 

### 2015-08-14 papers/Number1/paper.tex
Revised the script of RichardTransformed.r

@@
 \newcommand{\inftynorm}[1]{\|#1\|_\infty}
-\newcommand{\pp}[1]{\mathscr E_{#1}}
+\newcommand{\pp}[1]{\varepsilon_{#1}}
 \newcommand{\vep}{\varepsilon}
@@
 Suppose we have $p$ time series each of length $n$ and have arranged
-them in a $p\times n$ matrix $X_{it}$, where $i = 1, 2, \dots, p$ and
+them in a $p\times n$ matrix $R_{it}$, where $i = 1, 2, \dots, p$ and
 $t=1,2, \dots, n$. For each row of $X$, we transform it as
 \begin{eqnarray*}
-  R_{it} &=& -\left[
-    \log \left({1 \over n+1} \sum_{\tau=1}^n \1_{X_{i\tau} \leq X_{it}} \right)
+  X_{it} &=& -\left[
+    \log \left({1 \over n+1} \sum_{\tau=1}^n \1_{R_{i\tau} \leq R_{it}} \right)
   \right]^{-1}
 \end{eqnarray*}
-When $X_{i1}, X_{i2}, \dots, X_{in}$ are iid for each $i$, the
+When $R_{i1}, R_{i2}, \dots, R_{in}$ are iid for each $i$, the
 argument to the $\log$ function is asymptotically uniformly
-distributed as $n \to \infty$, and hence $R_{it}$ is a standard
+distributed as $n \to \infty$, and hence $X_{it}$ is a standard
 Fr\'echet r.v.
@@
 Figure \ref{fig:EigenRatio} shows the ratio of successive eigenvalues
-of the matrix $a_{np}^{-2}RR'$, where the rows of matrix $R$ comprise
+of the matrix $a_{np}^{-2}XX'$, where the rows of matrix $X$ comprise
 the rank-transformed sequences of a number of selected time series
@@
   \subfigure[]{
-    \includegraphics[scale=0.40]{EigenRatioSP500_100_shown.pdf}
+    \includegraphics[scale=0.40]{EigenRatioSP500_50_shown.pdf}
   }
   \subfigure[]{
-    \includegraphics[scale=0.40]{EigenRatioSP500_200_shown.pdf}
+    \includegraphics[scale=0.40]{EigenRatioSP500_150_shown.pdf}
   }
   \subfigure[]{
-    \includegraphics[scale=0.40]{EigenRatioSP500_300_shown.pdf}
+    \includegraphics[scale=0.40]{EigenRatioSP500_250_shown.pdf}
   }
   \subfigure[]{
-    \includegraphics[scale=0.40]{EigenRatioSP500_400_shown.pdf}
+    \includegraphics[scale=0.40]{EigenRatioSP500_350_shown.pdf}
   }
   \caption{$\lambda_{(i+1)} / \lambda_{(i)}$ versus $i$, The blue line
-  is the expected value of the ratio, the red line is the 5\% quantile
-  and the green line is the 95\% quantile.}
+  is the median of the ratio, the red line is the 1\% quantile
+  and the green line is the 99\% quantile.}
   \label{fig:EigenRatio}
@@
 \begin{eqnarray*}
-\P\left[{\lambda_{(i+1)} \over \lambda_{(i)}} \leq x \right] &=& x^{i/2} \\
-\E \left[{\lambda_{(i+1)} \over \lambda_{(i)}} \right] &=& {i \over i + 2}
+\P\left[{\lambda_{(i+1)} \over \lambda_{(i)}} \leq x \right] &=& x^{i/2}
+%\E \left[{\lambda_{(i+1)} \over \lambda_{(i)}} \right] &=& {i \over i + 2}
 \end{eqnarray*}
-The expected values and confident intervals marked in figure
+The median and confident intervals marked in figure
 \ref{fig:EigenRatio} are calculated according to the above formulas.

### 2015-08-14 r/RichardTransformed.r

@@ -122,19 +122,22 @@ E <- eigen((n.records * p)^(-2) * t(R.trfm) %*% R.trfm);
 
 ## plot(lambda[2:p]/lambda[1:p-1], type="b", xlim=c(1, 20), ylim=c(0,1));
 
-N = 400;
+N = 350;
 pdf(sprintf("EigenRatioSP500_%d_shown.pdf", N));
-plot(E$values[2:p]/E$values[1:p-1], type="b", pch=20, xlim=c(1, N), ylim=c(0, 1),
-     xlab=expression(i), ylab=expression(lambda[(i+1)]/lambda[(i)]));
+## plot(E$values[2:p]/E$values[1:p-1], type="b", pch=20, xlim=c(1, N), ylim=c(0, 1),
+##      xlab=expression(i), ylab=expression(lambda[(i+1)]/lambda[(i)]), log="y");
+plot(log(E$values[2:p]/E$values[1:p-1]), type="b", pch=20, xlim=c(1, N), ylim=c(-2.1, 0),
+     xlab=expression(i), ylab=expression(log(lambda[(i+1)]/lambda[(i)])));
 I <- 1:(N-1);
-q = 0.95;
-Q <- matrix(NA, ncol=2, nrow=N-1);
+q = 0.99;
+Q <- matrix(NA, ncol=3, nrow=N-1);
 Q[, 1] = q^(2/I);
-Q[, 2] = (1-q)^(2/I);
-expected <- I / (I+2);
-lines(I, expected, col="#0000FF", type="l");
-lines(I, Q[, 2], col="#FF0000");
-lines(I, Q[, 1], col="#00FF00");
+Q[, 2] = (0.5)^(2/I);
+Q[, 3] = (1-q)^(2/I);
+## expected <- I / (I+2);
+lines(I, log(Q[, 2]), col="#0000FF", type="l");
+lines(I, log(Q[, 3]), col="#FF0000");
+lines(I, log(Q[, 1]), col="#00FF00");
 dev.off();
 
 

### 2015-08-26 papers/Number1/paper.tex
illustrated the approximation of eigenvalues.

@@
   technical result)}
+To illustrate the validity of the results, we pick the following
+distribution function for $Z$ and compare the eigenvalues with
+the order statistics of $D_i = \sum_{t=1}^n Z_{it}^2$ and $Z^2$ using
+simulated data.
+\[
+f(x) = {d \over dx}\P(Z \leq x) = \left\{
+  \begin{array}{ll}
+    {\alpha \over (4x)^{\alpha + 1}} & |x| > 1/4 \\
+    \alpha & \text{ otherwise }
+  \end{array}
+\right.
+\]
+This function is shown in figure \ref{fig:densitySim}.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.40]{densitySim.pdf}
+  \caption{Density function of simulated data. $\alpha=1.6$ is chosen
+    for simulation.}
+  \label{fig:densitySim}
+\end{figure}
+Using the simulated data, empirical density functions are obtained for
+$a_{np}^{-2}\lambda_{(i)} - a_{np}^{-2}D_{(i)}$ and
+$a_{np}^{-2}\lambda_{(i)} - a_{np}^{-2}Z^2_{(i)}$. These are shown in
+figure \ref{fig:lambda_comparison}.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.40]{lambda_comparison.pdf}
+  \caption{A total of 20000 pairs are compared in each case.}
+  \label{fig:lambda_comparison}
+\end{figure}
+
 \subsection{$\alpha \in [2,4)$: Overview of existing results: Davis et

### 2015-09-07 papers/Number1/paper.tex

@@
 
-\subsection{Overview of light-tailed case}\label{sec:light_tail_review}
-Yin, Bai and Krishnaiah proved in their 1988 paper
+\subsection{Overview of light-tailed
+  case}\label{sec:light_tail_review}
+As early as in 1962, T. W. Anderson \cite{anderson:1963} showed that,
+when the entries of the data matrix has normal distribution ($\E
+X_{ij} = 0$ for $i,j = 1,2,\dots, p$; $\E X_{ii}^2 = 2\sigma^2, \E
+X_{ij}^2 = \sigma^2$ for $i \neq j$), the joint density function
+of the eigenvalues of the sample covariance matrix is given by
+\[
+f(\lambda_1, \lambda_2, \dots, \lambda_p; \sigma, p) \propto \prod_{i=1}^p \exp\left(
+  -{\lambda_i^2 \over 4\sigma^2}
+\right) \prod_{i < j} (\lambda_i - \lambda_j)
+\]
+In 1988, Yin, Bai and Krishnaiah proved in
 \cite{yin:bai:Krishnaiah:1988} the following result for a $p \times n$
@@
 \]
-When $\E X_{11} = \infty$, Yin and Silverstein proved in 1988
+When $\E X_{11}^4 = \infty$, Yin and Silverstein proved in 1988
 \cite{yin:silverstein:1988}
@@
 \]
+Later in 2001, I. M. Johnstone showed that, when the data matrix has
+standard normal distribution, i.e. $X_{ij} \sim N(0, 1)$, the largest
+eigenvalue of the sample covariance matrix follows the Tracy-Widom
+distribution:
+\[
+{ł_1 - \mu_{np} \over \sigma_{np}} \overset{d}{\to} W_1 \sim F_1
+\text{ as }{n, p \to \infty}
+\]
+where $\mu_{np}$ and $\sigma_{np}$ are constants dependent on $p$ and
+$n$ and the distribution function $F_1$ is given by
+\[
+F_1(s) = \exp\left\{
+  -{1 \over 2} \int_{s}^\infty \left[
+    q(x) + (x - s) q^2(x)
+  \right] dx
+\right\}
+\]
+where $q(x)$ is the unique solution to the Painlev\'e II differential
+equation
+\begin{eqnarray*}
+  q''(x) &=& xq(x) + 2 q^3(x) \\
+  q(x) &\sim& \text{Ai}(x) \text{ as } x \to \infty
+\end{eqnarray*}
+Figure \ref{fig:normal-TW} illustrates the density function of the largest
+eigenvalue from a simulated sample and that of the Tracy-Widom.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5]{normal-TW.pdf}
+  \caption{Sample density function of the largest eigenvalue compared
+    with the Tracy-Widom density function. The data matrix has
+    dimension $200 \times 1000$. An ensamble of 2000 matrices are
+    simulated.}
+  \label{fig:normal-TW}
+\end{figure}
+
 Later on, several authors \cite{Pillai:Yin:2011, tao:vu:2012,
-  Wang:2012} looked into empirical covariance matrix $S_{n, p} = {1
+  Wang:2012} looked into sample covariance matrix $S_{n, p} = {1
   \over n} \bf{XX}'$, where the entries of $\bf X$, denoted $X_{ij}$
@@
 that, for $q = \lim_{n \to \infty} {p \over n} \in (0, \infty)
-\setminus \{1\}$, $\forall \zeta > 0, \exists C > 0, \exists \delta >
-0$,
+\setminus \{1\}$, the probability of the $j$-th largest eigenvalue of
+the sample covariance matrix deviating significantly from its
+theoretical value computed as the $1 - j/p$ quantile of the
+Marcenko-Pastur distribution decreases exponentially\footnotemark.
+
+\footnotetext{
+In precise terms, this means $\forall \zeta > 0, \exists C > 0,
+\exists \delta > 0$,
 \[
@@
 \end{eqnarray*}
+}
 
@@
 \cite{Wang:2012}. In their setting, a $p \times n$ matrix $\bf X$ is
-said to obey condition {\bf C1} with constant $C_0$ if $\forall 1
+said to obey condition {\bf C1} with constant $c_0$ if $\forall 1
 \leq i \leq p, \forall 1 \leq j \leq n$, $\E X_{ij} = 0$, $\E
-X_{ij}^2 = 1 $ and $\sup_{i,j} \E|X_{ij}|^{C_0} < C$ for some constant
+X_{ij}^2 = 1 $ and $\sup_{i,j} \E|X_{ij}|^{c_0} < C$ for some constant
 $C$ independent of $p$ and $n$.
@@
 Assume matrices $\bf X$ and $\bf Y$ both satisfy condition {\bf C1}
-with constant $C_0$ and in addition they match up to order $4$,
+with constant $c_0$ and in addition they match up to order $4$,
 i.e. for all $m, l \geq 0$ such that $m + l \leq 4$, $\E[\mathscr{
 R}(X_{ij})^m \mathscr{I}(X_{ij})^l] = \E[\mathscr{R}(Y_{ij})^m
-\mathscr{I}(Y_{ij})^l]$. Assume also $\lim_{n \to \infty} p/n = q \in
-(0, 1]$. Let $G: \mathbb R^k \to \mathbb R$ be a smooth function
-obeying the derivative bounds
+\mathscr{I}(Y_{ij})^l]$ for all $1 \leq i,j \leq p$. Assume also
+$\lim_{n \to \infty} p/n = q \in (0, 1]$. Let $G: \mathbb R^k \to
+\mathbb R$ be a smooth function obeying the derivative bounds
 \[
-\left| \nabla^j G(x) \right| < n^{c_0}
+\left| \nabla^j G(\vec x) \right| < n^{c_0}
 \]
-for some $c_0 > 0$. Then the following holds for all $1 \leq i_1 \leq
-i_2 \leq \cdots \leq i_{k} \leq p$, $1 \leq k \leq p$:
+for some $c_0 > 0$ and for all $\vec x \in \mathbb R^k$ and all $0
+\leq j \leq 5$. Then the following holds for all $1 \leq i_1 \leq i_2
+\leq \cdots \leq i_{k} \leq p$, $1 \leq k \leq p$:
 \[
 \left|
-\E G(n\lambda_{i_1}, \dots, n\lambda_{i_k}) - 
-\E G(n\omega_{i_1}, \dots, n\omega_{i_k})
+\E G(\lambda_{i_1}, \dots, \lambda_{i_k}) - 
+\E G(\omega_{i_1}, \dots, \omega_{i_k})
 \right| < n^{- c_0}
@@
 where $\lambda_1 \leq \lambda_2, \leq \cdots \leq \lambda_p$ are the
-eigenvalues of ${1 \over n} XX'$ and $\omega_1 \leq \omega_2, \leq
-\cdots \leq \omega_p$ are those of ${1 \over n} YY'$.
+eigenvalues of $\bf XX'$ and $\omega_1 \leq \omega_2, \leq
+\cdots \leq \omega_p$ are those of $\bf YY'$.
 
-\subsection{Finite $p$: Anderson}
-\subsection{Growing $p$: Johnstone, Tao + Vu}
+Moreover, even when ${\bf XX'}$ and ${\bf YY'}$ only match to order
+3, the above inequality still holds provided that the bounds on the
+derivatives are strengthened to
+\[
+|\nabla^j G(\vec x)| < n^{-j c_1}
+\]
+for all $0 \leq j \leq 5$, all $\vec x \in \mathbb R^k$ and some $c_1$
+that is sufficiently large compared to $c_0$.
 
@@
   \begin{array}{ll}
-    {\alpha \over (4x)^{\alpha + 1}} & |x| > 1/4 \\
+    {\alpha \over (4|x|)^{\alpha + 1}} & |x| > 1/4 \\
     \alpha & \text{ otherwise }

### 2015-09-07 papers/Number1/paper.tex
Revise the procedure for simulating eigenvalues.

@@
 \end{eqnarray*}
-Figure \ref{fig:normal-TW} illustrates the density function of the largest
-eigenvalue from a simulated sample and that of the Tracy-Widom.
+Figure \ref{fig:3point-TW} illustrates the density function of the largest
+eigenvalue from a simulated 3-point sample compared with that of the
+Tracy-Widom.
 \begin{figure}[htb!]
   \centering
-  \includegraphics[scale=0.5]{normal-TW.pdf}
+  % \subfigure[Normal]{
+  %   \includegraphics[scale=0.5]{normal-TW.pdf}
+  % }
+  \subfigure[3-point: $\P(X=sqrt(3)) = \P(X=-sqrt(3)) = 1/6$, $\P(X=0)=2/3$]{
+    \includegraphics[scale=0.3]{3point-TW.pdf}
+  }
   \caption{Sample density function of the largest eigenvalue compared
@@
     simulated.}
-  \label{fig:normal-TW}
+  \label{fig:3point-TW}
 \end{figure}

### 2015-09-08 papers/Number1/paper.tex
Revised the paper wth Thomas, Richard and Johannes.

--- 2015-09-07 papers/Number1/paper.tex
@@
 \end{eqnarray*}
-Figure \ref{fig:normal-TW} illustrates the density function of the largest
-eigenvalue from a simulated sample and that of the Tracy-Widom.
+Figure \ref{fig:normal-3point-TW} compares the density function of the largest
+eigenvalue from simulated $N(0,1)$ and 3-point samples with
+that of the Tracy-Widom.
 \begin{figure}[htb!]
   \centering
-  \includegraphics[scale=0.5]{normal-TW.pdf}
+  \subfigure[$N(0, 1)$ entries]{
+    \includegraphics[scale=0.4]{normal-TW.pdf}
+  }
+  \subfigure[Entry distribution: $\P(X=\sqrt 3) = \P(X=-\sqrt 3 ) =
+  1/6$, $\P(X=0)=2/3$]{
+    \includegraphics[scale=0.4]{3point-TW.pdf}
+  }
   \caption{Sample density function of the largest eigenvalue compared
     with the Tracy-Widom density function. The data matrix has
-    dimension $200 \times 1000$. An ensamble of 2000 matrices are
+    dimension $200 \times 1000$. An ensemble of 2000 matrices are
     simulated.}
-  \label{fig:normal-TW}
+  \label{fig:normal-3point-TW}
 \end{figure}
-
+In the case of entry distributions with infinite 4th moment, the
+Tracy-Widom law no longer applies and gives way to the Frechet
+distribution. Figure \ref{fig:MyDist-Frechet} illustrates this with a
+simulated ensemble whose entries are distributed according to
+\eqref{eq:mydist}.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.5]{MyDist-Frechet.pdf}  
+  \caption{Sample distribution function of the largest eigenvalue compared with that of Frechet}
+  \label{fig:MyDist-Frechet}
+\end{figure}
 Later on, several authors \cite{Pillai:Yin:2011, tao:vu:2012,
@@
 simulated data.
-\[
+\begin{equation}
+  \label{eq:mydist}
   f(x) = {d \over dx}\P(Z \leq x) = \left\{
@@
   \right.
-\]
+\end{equation}
 This function is shown in figure \ref{fig:densitySim}.

### 2015-09-08 r/simulateEig.r

@@ -45,15 +45,51 @@ n <- 1000;
 mu <- (sqrt(n-1) + sqrt(p))^2;
 sigma <- (sqrt(n-1) + sqrt(p)) * (1/sqrt(n-1) + 1/sqrt(p))^(1/3);
 
-## Simulate using the 3-point distribution
+## Simulate using myDist distribution
 N <- 2000;
+a <- 1.6;
+a.np <- qMyDist(1 - 1/(n*p), a) * 1.6;
 lambda <- rep(NA, N);
-for (i in (2:1000)) {
-    X <- matrix(r3Point(p*n, 1/6, sqrt(3)), nrow=p, ncol=n);
+for (i in (1001:2000)) {
+    X <- matrix(rMyDist(p*n, a), nrow=p, ncol=n);
     C <- X %*% t(X);
     E <- eigen(C, only.values=TRUE);
     lambda[i] <- E$values[1];
 }
+W <- lambda[1:N]/a.np^2;
+W1 <- (lambda - mu)/sigma;
+
+Fn <- ecdf(W);
+X <- exp(seq(min(log(W)), max(log(W)), length.out=200));
+Y <- Fn(X);
+
+pdf("MyDist-Frechet.pdf")
+plot(log(X), (Y), type="p", xlab=expression(log(x)),
+     ylab=expression(P(lambda[(1)] < x)),
+     main="Sample Distribution Function and Frechet");
+Y.f <- pfrechet(X, shape=a/2);
+lines(log(X), Y.f, col="#FF0000");
+grid(nx=20);
+explanations <- c(
+    "sample",
+    "Frechet"
+    );
+legend("topleft", legend=explanations,
+       lty=c(1, 1), lwd=c(1, 1),
+       col=c("#000000", "#FF0000"));
+dev.off();
+
+
+
+## Simulate using the 3-point distribution
+## N <- 2000;
+## lambda <- rep(NA, N);
+## for (i in (2:1000)) {
+##     X <- matrix(r3Point(p*n, 1/6, sqrt(3)), nrow=p, ncol=n);
+##     C <- X %*% t(X);
+##     E <- eigen(C, only.values=TRUE);
+##     lambda[i] <- E$values[1];
+## }
 
 
 ## Simulate using N(0, 1)
@@ -66,23 +102,24 @@ for (i in (2:1000)) {
 ##     lambda[i] <- E$values[1];
 ## }
 
[…]

### 2015-09-09 papers/Number1/paper.tex
Revised the paper with Richard David.

@@
 
+Figure \ref{fig:LamYao} compares the largest eigenvalue of the sum of
+time-lagged covariance matrices with the sum of the largest
+eigenvalues of individual time-lagged covariance matrices. The data
+are rank-transformed return series of 478 stocks selected from the S\&P
+500 index.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[] {
+    \includegraphics[scale=0.4]{eigen_sum.pdf}
+    \label{fig:LamYao:a}
+  }
+  \subfigure[] {
+    \includegraphics[scale=0.4]{eigen_sum_plus.pdf}
+    \label{fig:LamYao:a}
+  }
+  \label{fig:LamYao}
+  \caption{The largest eigenvalue of summed time-lagged covariance
+    matrices and the sum of the largest eigenvalues of time-lagged
+    covariance matrices. In both cases $(C_i)_{lm} =
+    (np)^{-2}\sum_{t=1}^{n-i} X_{lt} X_{m,t+i}$}
+\end{figure}
+
 \subsection{Overview of light-tailed
@@
 
-\subsection{$\alpha \in (0,2)$: Overview of existing results: Davis et
-  al., Soshnikov, Auffiner et al.}
-\subsection{$\alpha \in (0,2)$: Our extension of these results; only
-  order statistics of $Z_{it}$'s necessary and no sums (proof by the
-  technical result)}
+\subsection{\texorpdfstring{$\alpha \in (0,2)$}{alpha in (0,2)}:
+  Overview of existing results: Davis et al., Soshnikov, Auffiner et
+  al.}
+\subsection{\texorpdfstring{$\alpha \in (0,2)$}{alpha in (0,2)}: Our
+  extension of these results; only order statistics of \texorpdfstring{$Z_{it}$}{Z[it]}'s
+  necessary and no sums (proof by the technical result)}
 To illustrate the validity of the results, we pick the following
@@
 
-\subsection{$\alpha \in [2,4)$: Overview of existing results: Davis et
+\subsection{\texorpdfstring{$\alpha \in [2,4)$}{alpha in [2,4)}: Overview of existing results: Davis et
   al., Soshnikov, Auffiner et al.}
-\subsection{$\alpha \in [2,4)$: The problem of centering and restrictions on $p_n$}
-\subsection{$\alpha \in [2,10/3)$: Our new result for the UN-CENTERED
-  sample covariance matrix; only order statistics of $Z_{it}$'s
+\subsection{\texorpdfstring{$\alpha \in [2,4)$}{alpha in [2,4)}: The problem of centering and restrictions on \texorpdfstring{$p_n$}{p[n]}}
+\subsection{\texorpdfstring{$\alpha \in [2,10/3)$}{alpha in [2,10/3)}: Our new result for the UN-CENTERED
+  sample covariance matrix; only order statistics of \texorpdfstring{\texorpdfstring{$Z_{it}$}{Z[it]}}{Z[it]}'s
   necessary and no sums (proof by the technical result)}
-\subsection{$\alpha \in [10/3,4)$: A conjecture with partial proof for
+\subsection{\texorpdfstring{$\alpha \in [10/3,4)$}{alpha in [10/3,4)}: A conjecture with partial proof for
   the UN-CENTERED sample covariance matrix)}
@@
 
-\subsection{$\alpha \in (0,10/3)$: Eigenvalues of the sample covariance matrix}
-\subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample covariance matrix}
-\subsection{$\alpha \in [2,4)$ and $p/n \to \gamma \in (0,1]$:
+\subsection{\texorpdfstring{$\alpha \in (0,10/3)$}{alpha in (0,10/3)}: Eigenvalues of the sample covariance matrix}
+\subsection{\texorpdfstring{$\alpha \in [10/3,4)$}{alpha in [10/3,4)}: Eigenvalues of the CENTERED sample covariance matrix}
+\subsection{\texorpdfstring{$\alpha \in [2,4)$}{alpha in [2,4)} and
+  \texorpdfstring{$p/n \to \gamma \in (0,1]$}{p/n tends to gamma in (0,1]}:
   Eigenvalues of the sample covariance matrix using Auffinger's
@@
 
+\subsection{Probability Mass Due To Correlations}
+In figure \ref{fig:ProbMass} the distribution function of
+${\lambda_{(1)} - \lambda_{(2)} \over \lambda_{(1)}}$ in the cases of
+iid data and correlated data, which are described in {\color{red} fill
+in here}, is illuatrated. The predicted probability mass at 3/4 is
+clearly seen.
+\begin{figure}[htb!]
+  \centering
+  \subfigure[iid data]{
+    \includegraphics[scale=0.4]{lambda_1_2_Gaps_iid.pdf}
+  }
+  \subfigure[Correlated data]{
+    \includegraphics[scale=0.4]{lambda_1_2_Gaps.pdf}
+  }
+  \caption{Distribution function of ${\lambda_{(1)} - \lambda_{(2)} \over
+      \lambda_{(1)}}$ in the cases of iid data and correlated
+    data. The data matrices have dimension $200 \times 1000$. 1000
+    data matrices are simulated in each case. The data have the
+    distribution described in \eqref{eq:mydist} with $\alpha=0.6$.}
+  \label{fig:ProbMass}
+\end{figure}
+
 %------------------------------------------------------------------------
@@
 
-\subsection{$\alpha \in (0,10/3)$: Eigenvalues of the sample
+\subsection{\texorpdfstring{$\alpha \in (0,10/3)$}{alpha in (0,10/3)}: Eigenvalues of the sample
   autocovariance matrix}
-\subsection{$\alpha \in [10/3,4)$: Eigenvalues of the CENTERED sample
+\subsection{\texorpdfstring{$\alpha \in [10/3,4)$}{alpha in [10/3,4)}: Eigenvalues of the CENTERED sample
   autocovariance matrix}

### 2015-09-09 papers/Number1/paper.tex
Revised the paper.

@@
 
+\newpage
 Figure \ref{fig:LamYao} compares the largest eigenvalue of the sum of
@@
 eigenvalues of individual time-lagged covariance matrices. The data
-are rank-transformed return series of 478 stocks selected from the S\&P
-500 index.
+are return series of 478 stocks selected from the S\&P 500 index.
 \begin{figure}[htb!]
@@
     \includegraphics[scale=0.4]{eigen_sum_plus.pdf}
-    \label{fig:LamYao:a}
+    \label{fig:LamYao:b}
   }
-  \label{fig:LamYao}
   \caption{The largest eigenvalue of summed time-lagged covariance
@@
     covariance matrices. In both cases $(C_i)_{lm} =
-    (np)^{-2}\sum_{t=1}^{n-i} X_{lt} X_{m,t+i}$}
+    \sum_{t=1}^{n-i} X_{lt} X_{m,t+i}$.}
+  \label{fig:LamYao}
 \end{figure}
+
+\newpage
 
@@
   \subfigure[Entry distribution: $\P(X=\sqrt 3) = \P(X=-\sqrt 3 ) =
-  1/6$, $\P(X=0)=2/3$]{
+  1/6$, $\P(X=0)=2/3$. Note $\E X = 0$, $\E X^2 = 1$, $\E X^3 = 0$ and
+  $\E X^4 = 3$, i.e. the first 4 moments match those of $N(0, 1).$]{
     \includegraphics[scale=0.4]{3point-TW.pdf}
@@
 \end{figure}
+
+\newpage
 In the case of entry distributions with infinite 4th moment, the
@@
 simulated ensemble whose entries are distributed according to
-\eqref{eq:mydist}.
+\eqref{eq:mydist} with $\alpha = 1.6$.
 \begin{figure}[htb!]
@@
   \includegraphics[scale=0.5]{MyDist-Frechet.pdf}  
-  \caption{Sample distribution function of the largest eigenvalue compared with that of Frechet}
+  \caption{Sample distribution function of the largest eigenvalue
+    compared with that of Frechet. The data matrices have dimensions
+    $200 \times 1000$. An ensemble of 2000 matrices are simulated.}
   \label{fig:MyDist-Frechet}
 \end{figure}
+
+\newpage
 Later on, several authors \cite{Pillai:Yin:2011, tao:vu:2012,

### 2015-10-05 r/RichardTransformed.r
Revised some pictures.

@@ -118,7 +118,7 @@ for (i in 1 : length(stocks.included)) {
     ## U <- U[U < 1];
     R.trfm[, i] <- -1/log(rank(R[,i])/(n.records+1));
 }
-# E <- eigen((n.records * p)^(-2) * t(R.trfm) %*% R.trfm);
+## E <- eigen((n.records * p)^(-2) * t(R.trfm) %*% R.trfm);
 E <- eigen((n.records * p)^(-2) * t(R) %*% R);
 
 ## plot(lambda[2:p]/lambda[1:p-1], type="b", xlim=c(1, 20), ylim=c(0,1));
@@ -128,8 +128,9 @@ pdf(sprintf("EigenRatioSP500_lin2_%d_shown.pdf", N));
 ## plot(E$values[2:p]/E$values[1:p-1], type="b", pch=20, xlim=c(1, N), ylim=c(0, 1),
 ##      xlab=expression(i), ylab=expression(lambda[(i+1)]/lambda[(i)]), log="y");
 
-plot((1:(N-1)), (E$values[2:N]/E$values[1:(N-1)]), type="b",
-     ylim=c(0.9, 1),
+plot((1:(N-1)), log(E$values[2:N]/E$values[1:(N-1)]), type="b",
+     xlim=c(0,50),
+     ylim=log(c(0.9, 1)),
      xlab=expression(i),
      ylab=expression(lambda[(i+1)]/lambda[(i)]));
 I <- 1:(N-1);

### 2015-10-05 r/simulateEig.r

@@ -7,42 +7,42 @@ source("liblxb353.r");
 p <- 200;
 n <- 1000;
 
-mu <- (sqrt(n-1) + sqrt(p))^2;
-sigma <- (sqrt(n-1) + sqrt(p)) * (1/sqrt(n-1) + 1/sqrt(p))^(1/3);
+## mu <- (sqrt(n-1) + sqrt(p))^2;
+## sigma <- (sqrt(n-1) + sqrt(p)) * (1/sqrt(n-1) + 1/sqrt(p))^(1/3);
 
 ## Simulate using myDist distribution
-N <- 2000;
-a <- 1.6;
-a.np <- qMyDist(1 - 1/(n*p), a) * 1.6;
-lambda <- rep(NA, N);
-for (i in (1001:2000)) {
-    X <- matrix(rMyDist(p*n, a), nrow=p, ncol=n);
-    C <- X %*% t(X);
-    E <- eigen(C, only.values=TRUE);
-    lambda[i] <- E$values[1];
-}
-W <- lambda[1:N]/a.np^2;
-W1 <- (lambda - mu)/sigma;
+## N <- 2000;
+## a <- 1.6;
+## a.np <- qMyDist(1 - 1/(n*p), a) * 1.6;
+## lambda <- rep(NA, N);
+## for (i in (1001:2000)) {
+##     X <- matrix(rMyDist(p*n, a), nrow=p, ncol=n);
+##     C <- X %*% t(X);
+##     E <- eigen(C, only.values=TRUE);
+##     lambda[i] <- E$values[1];
+## }
+## W <- lambda[1:N]/a.np^2;
+## W1 <- (lambda - mu)/sigma;
 
-Fn <- ecdf(W);
-X <- exp(seq(min(log(W)), max(log(W)), length.out=200));
-Y <- Fn(X);
+## Fn <- ecdf(W);
+## X <- exp(seq(min(log(W)), max(log(W)), length.out=200));
+## Y <- Fn(X);
 
-pdf("MyDist-Frechet.pdf")
-plot(log(X), (Y), type="p", xlab=expression(log(x)),
-     ylab=expression(P(lambda[(1)] < x)),
-     main="Sample Distribution Function and Frechet");
-Y.f <- pfrechet(X, shape=a/2);
-lines(log(X), Y.f, col="#FF0000");
-grid(nx=20);
-explanations <- c(
-    "sample",
-    "Frechet"
-    );
-legend("topleft", legend=explanations,
-       lty=c(1, 1), lwd=c(1, 1),
[…]

### 2015-11-05 r/HillEstimate_sectors.r
Revised the paper about fixed-dimensional sample covariance matrix.

@@ -144,10 +144,10 @@ dbClearResult(results);
 colors <- rainbow(length(sector.names));
 
 ## Record the eigenvalues
-lambda <- matrix(NA, nrow=1+length(sector.names), ncol=p);
+## lambda <- matrix(NA, nrow=1+length(sector.names), ncol=p);
 company.num <- rep(NA, length(sector.names));
 for (i in 1:length(sector.names)) {
-    I <- which(sectors[to.include] == sector.names[i]);
+    I <- which(sectors[to.include] != sector.names[i]);
     ## if (i == 1) {
     ##     plot(tailIndices[I, 1], tailIndices[I, 2],
     ##          col=colors[i], pch=i,
@@ -157,7 +157,7 @@ for (i in 1:length(sector.names)) {
     ##     points(tailIndices[I, 1], tailIndices[I, 2], col=colors[i], pch=i);
     ## }
 
-    ## pdf(sprintf("/tmp/%s_tail_indices.pdf", sector.names[i]));
+    ## pdf(sprintf("../papers/%s_tail_indices.pdf", sector.names[i]));
     ## plot(tailIndices[I, 1], tailIndices[I, 2],
     ##      col="#0000FF", pch=i,
     ##      xlim=c(a, b), ylim=c(a, b),
@@ -167,15 +167,76 @@ for (i in 1:length(sector.names)) {
     ## lines(X, X, col="#FF0000");
     ## grid(nx=20);
     ## dev.off();
+
+    pdf(sprintf("../papers/%s_eigenvalue_ratios.pdf", sector.names[i]));
     company.num[i] <- length(I);
-    R.trfm <- rank.transform(R, I);
-    E <- eigen((n.records * length(I))^(-2) * t(R.trfm) %*% R.trfm, only.values=TRUE);
-    lambda[i,1:length(I)] <- E$values;
-    ## dev.off();
+    # R.trfm <- rank.transform(R, I);
+    E <- eigen(t(R[, I]) %*% R[, I], only.values=TRUE);
+    l <- length(E$values);
+    plot(1:(l-1), (E$values[2:l] / E$values[1:(l-1)]), type="b",
+         xlab="", ylab="", xlim=c(1, 100), ylim=c(0.1, 1),
+         main=sprintf("\"%s\" removed", sector.names[i]));
+    dev.off();
 }
-R.trfm <- rank.transform(R, 1:(dim(R)[2]));
[…]

### 2015-11-22 papers/PDR.tex
PDR to be discussed with Mogens.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}
\usepackage{graphicx}

\newif\ifALL

\ALLtrue

\input{../physics_common}
\title{Performance and Development Review}
\author{Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

\ifALL
\section{Research Activities Oct. 2014 - Oct. 2015}
\begin{itemize}
\item Co-authored ``Extreme value analysis for the sample
  autocovariance matrices of heavy-tailed multivariate time
  series''. Paper submitted to the ``Extremes''.
\item Went to the {\it Extreme Value Analysis} Conference in Ann Arbor
  in June, 2015.
\end{itemize}

\section{Goals for Research in the comming year}
\begin{itemize}
\item Finish the ongoing paper with Prof. Jeffrey Collamore on
  importance sampling of probabilities of large exceedances by Matrix
  Markov Additive process.
\item Empirical study of the eigenvalues of covariance and correlation
  matrices, using data of Foreigen Exchange Rates and stock prices
  data.
\item I am most excited by the the possibility of explaining patterns
  found in real-world data using our mathematical model.
\end{itemize}

\section{Teaching Activities Oct. 2014 - Oct. 2015}
\begin{itemize}
\item Helped Prof. Jostein Paulsen in teaching the course
  ``Statistical Methods of Non-life Insurance''. Taught the excercise
  sessions.
\item I hope to provide the students with more insight into the course
  material via the exercise sessions.
\item I consult Prof. Paulsen whenever in doubt about solutions to the
  problems or about how to interact with the students.
\item I feel most satisfied when I myself obtain better understanding
  of the mathematics when discussing the problems with Prof. Paulsen
  and with the students.
\end{itemize}

\section{Career}
\begin{itemize}
\item My aspiration is to work in the industry and apply the methods
  and knowledge learned during my Ph.D studies.
\item I hope to have opportunities of working as a summer internship
  in a major investment bank or hedge fund.
\end{itemize}

\section{Development Plan}
Specific Development goals for the next year:
\begin{itemize}
\item ``Introduction to University Pedagogy'', January 25.-29. 2016.
\item ``Responsible Conduct of Research'', February 2. 2016.
\end{itemize}

\bibliographystyle{unsrt}
\bibliography{../thesis/econophysics}
\end{document}

### 2015-11-24 papers/PDR.tex
Empirical studies of the eigenvalues and eigenvectors of a correlation matrix.

@@
 \item My aspiration is to work in the industry and apply the methods
-  and knowledge learned during my Ph.D studies.
+  and knowledge developed during my Ph.D studies.
 \item I hope to have opportunities of working as a summer internship
-  in a major investment bank or hedge fund.
+  at a role of quantitative analysis.
 \end{itemize}

### 2015-11-24 papers/segments/CorrelationMatrix.tex
Empirical studies of the eigenvalues and eigenvectors of a correlation matrix.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{subfigure}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}
\usepackage{graphicx}

\input{../../physics_common}
\title{Eigenvalues and Eigenvectors of Correlation/Covariance Matrices}
\author{Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle
\section{Eigenvalues}
% \begin{figure}[htb!]
%   \centering
%   \subfigure[]{
%     \includegraphics[scale=0.3]{Eigenvalues_of_CovarianceMatrix_SP500.pdf}
%   }
%   \subfigure[]{
%     \includegraphics[scale=0.3]{Eigenvalues_of_CorrelationMatrix_SP500.pdf}
%   }
%   \caption{$\log_{10}(\lambda_{(i+1)} / \lambda_{(i)})$ versus $i$}
%   \label{fig:EigenRatio}
% \end{figure}
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Eigenvalues_Cor_n_Cov_Matrix.pdf}  
  \caption{$\log_{10}(\lambda_{(i)})$}
  \label{fig:EigenRatio}
\end{figure}

\section{Eigenvectors}
% \begin{figure}[htb!]
%   \centering
%   \subfigure[]{
%     \includegraphics[scale=0.3]{LargestComponent_of_Eigenvectors_of_CovarianceMatrix_SP500.pdf}
%   }
%   \subfigure[]{
%     \includegraphics[scale=0.3]{LargestComponent_of_Eigenvectors_of_CorrelationMatrix_SP500.pdf}
%   }
%   \caption{$\max_{1 \leq k \leq p} |V_{ki}|$, where $V_{ki}$ is the
%     $k$-th component of the eigenvector associated with the $i$-th
%     largest eigenvalue.}
%   \label{fig:Eigenvector_largest_comp}
% \end{figure}
\begin{figure}[htb!]
  \centering
  \subfigure[Covariance Matrix]{
    \includegraphics[scale=0.3]{V1_CovarianceMatrix.pdf}
  }
  \subfigure[Correlation Matrix]{
    \includegraphics[scale=0.3]{V1_CorrelationMatrix.pdf}
  }
  \caption{Eigenvector corresponding to the largest eigenvalue}
  \label{fig:Eigenvector_largest_comp}
\end{figure}
\bibliographystyle{unsrt}
\bibliography{../thesis/econophysics}
\end{document}

### 2015-11-26 papers/segments/CorrelationMatrix.tex
Empirical studies of the correlation matrix.

@@
 % \end{figure}
+
 \begin{figure}[htb!]
   \centering
-  \includegraphics[scale=0.5]{Eigenvalues_Cor_n_Cov_Matrix.pdf}  
-  \caption{$\log_{10}(\lambda_{(i)})$}
-  \label{fig:EigenRatio}
+  \subfigure[$\lambda_{(i)}$]{
+    \includegraphics[scale=0.3]{Eigenvalues_Cor_n_Cov_Matrix.pdf}
+  }
+  \subfigure[$\log_{10}\lambda^{\text{corr}}_{(i)}$
+  vs. $\log_{10}\lambda^{\text{cov}}_{(i)}$]{
+    \includegraphics[scale=0.3]{Eigenvalues_Cor_vs_Cov_Matrix.pdf}
+  }
 \end{figure}
+
+% \begin{figure}[htb!]
+%   \centering
+%   \includegraphics[scale=0.5]{Eigenvalues_Cor_n_Cov_Matrix.pdf}  
+%   \caption{$\lambda_{(i)}$}
+% \end{figure}
+
+% \begin{figure}[htb!]
+%   \centering
+%   \includegraphics[scale=0.5]{Eigenvalues_Cor_vs_Cov_Matrix.pdf}  
+%   \caption{$\lambda^{\text{corr}}_{(i)}$ vs. $\lambda^{\text{cov}}_{(i)}$}
+% \end{figure}
 
@@
 \end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{V2_CovarianceMatrix.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{V2_CorrelationMatrix.pdf}
+  }
+  \caption{Eigenvector corresponding to the 2nd largest eigenvalue}
+  \label{fig:Eigenvector_largest_comp}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{V3_CovarianceMatrix.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{V3_CorrelationMatrix.pdf}
+  }
+  \caption{Eigenvector corresponding to the 3rd largest eigenvalue}
+  \label{fig:Eigenvector_largest_comp}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{V4_CovarianceMatrix.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{V4_CorrelationMatrix.pdf}
+  }
+  \caption{Eigenvector corresponding to the 4th largest eigenvalue}
+  \label{fig:Eigenvector_largest_comp}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{V5_CovarianceMatrix.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{V5_CorrelationMatrix.pdf}
+  }
+  \caption{Eigenvector corresponding to the 5th largest eigenvalue}
+  \label{fig:Eigenvector_largest_comp}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{V6_CovarianceMatrix.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{V6_CorrelationMatrix.pdf}
+  }
+  \caption{Eigenvector corresponding to the 6th largest eigenvalue}
+  \label{fig:Eigenvector_largest_comp}
+\end{figure}
+
 \bibliographystyle{unsrt}

### 2015-12-09 papers/segments/CorrelationMatrix.tex
Revised the empirical studies' documents.

@@
 % \end{figure}
+\subsection{Covariance \& Correlation Matrices of Returns}
+
 \begin{figure}[htb!]
@@
   }
-  \caption{Eigenvector corresponding to the largest eigenvalue}
-  \label{fig:Eigenvector_largest_comp}
+  \caption{Eigenvector corresponding to the largest eigenvalue of the returns' covariance/correlation matrices}
 \end{figure}
@@
   }
-  \caption{Eigenvector corresponding to the 2nd largest eigenvalue}
-  \label{fig:Eigenvector_largest_comp}
+  \caption{Eigenvector corresponding to the 2nd largest eigenvalue of the returns' covariance/correlation matrices}
 \end{figure}
@@
   }
-  \caption{Eigenvector corresponding to the 3rd largest eigenvalue}
-  \label{fig:Eigenvector_largest_comp}
+  \caption{Eigenvector corresponding to the 3rd largest eigenvalue of the returns' covariance/correlation matrices}
 \end{figure}
@@
   }
-  \caption{Eigenvector corresponding to the 4th largest eigenvalue}
-  \label{fig:Eigenvector_largest_comp}
+  \caption{Eigenvector corresponding to the 4th largest eigenvalue of the returns' covariance/correlation matrices}
 \end{figure}
@@
   }
-  \caption{Eigenvector corresponding to the 5th largest eigenvalue}
-  \label{fig:Eigenvector_largest_comp}
+  \caption{Eigenvector corresponding to the 5th largest eigenvalue of the returns' covariance/correlation matrices}
 \end{figure}
@@
   }
-  \caption{Eigenvector corresponding to the 6th largest eigenvalue}
-  \label{fig:Eigenvector_largest_comp}
+  \caption{Eigenvector corresponding to the 6th largest eigenvalue of the returns' covariance/correlation matrices}
+\end{figure}
+
+\pagebreak
+\subsection{Covariance \& Correlation Matrices of Squared Returns}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{R2_cov_eigenvector1.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{R2_cor_eigenvector1.pdf}
+  }
+  \caption{Eigenvector corresponding to the largest eigenvalue of the squared returns' covariance/correlation matrices}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{R2_cov_eigenvector2.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{R2_cor_eigenvector2.pdf}
+  }
+  \caption{Eigenvector corresponding to the 2nd largest eigenvalue of the squared returns' covariance/correlation matrices}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{R2_cov_eigenvector3.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{R2_cor_eigenvector3.pdf}
+  }
+  \caption{Eigenvector corresponding to the 3rd largest eigenvalue of the squared returns' covariance/correlation matrices}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{R2_cov_eigenvector4.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{R2_cor_eigenvector4.pdf}
+  }
+  \caption{Eigenvector corresponding to the 4th largest eigenvalue of the squared returns' covariance/correlation matrices}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{R2_cov_eigenvector5.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{R2_cor_eigenvector5.pdf}
+  }
+  \caption{Eigenvector corresponding to the 5th largest eigenvalue of the squared returns' covariance/correlation matrices}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \subfigure[Covariance Matrix]{
+    \includegraphics[scale=0.3]{R2_cov_eigenvector6.pdf}
+  }
+  \subfigure[Correlation Matrix]{
+    \includegraphics[scale=0.3]{R2_cor_eigenvector6.pdf}
+  }
+  \caption{Eigenvector corresponding to the 6th largest eigenvalue of
+    the squared returns' covariance/correlation matrices}
 \end{figure}

### 2016-02-15 papers/Jeffrey1/RCR.tex
Wrote part of the essay for RCR.

\documentclass{article}

\title{Authorship Issues in Applied Probability Theory and Statistics}
\author{Xie Xiaolei}
\date{\today}

\begin{document}
\maketitle

\section{Introduction}
It is not unusual in the field of applied mathematics that several
researchers collaborate and produce a paper. However, their roles in
producing the paper vary. Quite often one or two people have an idea,
or in mathematical terminology, a conjecture; then, if feasible, they
try to verify the conjecture by computational simulations, i.e. they
develop computer software to create a large number of
realizations of the stochastic process of their interest and then
investigate the statistics of these realizations to find out whether
their initial conjecture is true. If the conjecture is not entirely
consistent with the result of simulation, they may adjust the
conjecture and run the simulation again. Not unexpectedly, these steps
are often iterated.

If the simulations eventually confirm the conjecture, the researchers
try to prove the conjecture with mathematical rigor. Again, this is a
complicated task that involves a considerable number of steps each
producing an intermediate result that is typically termed a lemma.

If the theoretical proof of the conjecture is successfully completed
with a number of supporting lemmas, the researchers are often
interested in real-world examples of their conjecture, which, now with
its rigorous proof, can be called a theorem. Once again, this often
imples programming: Computer software needs to be developed in order
to examine statistical data from sources of their interest, such as
stock trading, meteorological measurements, designed questionare,
etc. and look for signatures of the theorem, often in terms of
distributions of some random variables.

In summary, almost all significant research projects in applied
probability theory and statistics involve computer programming to some
extent. The holders of the initial idea, albeit clever and skilled,
may be not experts in programming and wish to have someone else do the work.
Moreover, even the theoretical development is often so
complicated that it has to be divided and ``outsourced'' to people who
may be unfamiliar with the central ideas of the initial conjecture.

Thus issues arise as to who should be listed as co-authors and in what
order. The rest of this essay discusses such potential issues in
details.

\end{document}

### 2016-02-16 papers/Jeffrey1/RCR.tex
Revised the paper and the RCR essay.

@@
 \author{Xie Xiaolei}
-\date{\today}
 
@@
 its rigorous proof, can be called a theorem. Once again, this often
-imples programming: Computer software needs to be developed in order
+implies programming: Computer software needs to be developed in order
 to examine statistical data from sources of their interest, such as
-stock trading, meteorological measurements, designed questionare,
+stock trading, meteorological measurements, designed question are,
 etc. and look for signatures of the theorem, often in terms of
@@
 
+\section{Software Developers' Roles \& Authorships}
+Development of software in this context often involves two aspects,
+firstly acquisition and preparation of data and secondly algorithmic
+processing of data. Obviously, only the second part of the development
+is related to the mathematical ideas of the work, and hence
+intuitively, only the undertakers of the second part should be counted
+as co-authors.
+
+However, the first part of the software development, i.e. acquisition
+and preparation of data can be creative by its own merit and according
+to the Vancouver Recommendations \cite{RCR} (p.36) ``may'' qualify its
+undertakers as co-authors. For example, while acquiring data, one may
+invent unconventional methods to collect data from sources that are
+otherwise inaccessible, or develop database routines that hugely
+improve efficiency. More often than not, these software solutions are
+not only laborious but also creative. That being said, these solutions
+most likely will not be mentioned in the text of the paper since they
+do not relate to the central mathematical ideas.
+
+On the other hand, the software developers in question are often
+master's students and won't be paid. Therefore, the only plausible way
+to acknowledge their efforts is authorship. This is also seen as a
+common and acceptable practice.
+
+Nevertheless, it is my opinion that this type of software development
+should not lead to authorship -- or more precisely, should not
+lead to authorship of a paper on mathematics. It is more appropriate
+to describe the innovations in a paper pertaining to computer
+science. A possible approach is to have a joint project between
+mathematics and computer science, in which researchers from both areas
+work together but produce separate papers each describing their own
+parts of the project.
+
+\section{Conclusion}
+Research in applied probability theory and statistics is often coupled
+with software development that may be innovative by its own merit. It
+is recommendable to have researchers from both areas work together but
+write separate papers so that different creative work can be described
+and published with the right authorships and in the right journals.
+
+\begin{thebibliography}{9}
+
+\bibitem{RCR}
+  Karsten Klint Jensen, Louise Whiteley and Peter Sand\o e.
+  University of Copenhagen,
+  1st edition,
+  2015.
+
+\end{thebibliography}
 \end{document}

### 2016-03-16 r/libxxie.r
Revised the paper with Jeff and also updated libxxie.r

@@ -82,21 +82,20 @@ and day between '%s' and '%s';", day1, day2));
 
         results <- dbSendQuery(
             database,
-            sprintf("select count(*) from %s where
-day between '%s' and '%s';", tables[i], day1, day2)
+            sprintf("select min(day) from %s;",
+                    tables[i])
             );
-        n.traded.days <- fetch(results)[1,1];
+        d1 <- fetch(results)[1, 1];
         dbClearResult(results);
 
         results <- dbSendQuery(
             database,
-            sprintf("select min(day) from %s;",
-                    tables[i])
+            sprintf("select max(day) from %s;", tables[i])
             );
-        d <- fetch(results)[1, 1];
+        d2 <- fetch(results)[1,1];
         dbClearResult(results);
 
-        if (n.traded.days < 1000 || d > day1) {
+        if (d1 > day1 || d2 < day2) {
             to.include[i] <- FALSE;
             next;
         }

### 2016-06-10 mw/garch_covariance/FX.tex
Worked on the minorization.

%% Created by Maple 2015.1, Linux
%% Source Worksheet: FX.mw
%% Generated: Thu Jun 09 15:57:19 CEST 2016
\documentclass{article}
\usepackage{maplestd2e}
\def\emptyline{\vspace{12pt}}
\begin{document}
\pagestyle{empty}
\DefineParaStyle{Maple Heading 1}
\DefineParaStyle{Maple Text Output}
\DefineParaStyle{Maple Dash Item}
\DefineParaStyle{Maple Bullet Item}
\DefineParaStyle{Maple Normal}
\DefineParaStyle{Maple Heading 4}
\DefineParaStyle{Maple Heading 3}
\DefineParaStyle{Maple Heading 2}
\DefineParaStyle{Maple Warning}
\DefineParaStyle{Maple Title}
\DefineParaStyle{Maple Error}
\DefineCharStyle{Maple Hyperlink}
\DefineCharStyle{Maple 2D Math}
\DefineCharStyle{Maple Maple Input}
\DefineCharStyle{Maple 2D Output}
\DefineCharStyle{Maple 2D Input}
\begin{Maple Normal}{
\begin{Maple Normal}{
From\mapleinline{inert}{2d}{x[i, t] := sigma[i, t]*z[i, t]}{$\displaystyle x_{{i,t}}\, := \,\sigma_{{i,t}}z_{{i,t}}$}
and\mapleinline{inert}{2d}{x[j, t] := sigma[j, t]*z[j, t]}{$\displaystyle x_{{j,t}}\, := \,\sigma_{{j,t}}z_{{j,t}}$}
we have}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{x[i, t]*x[j, t] = sigma[i, t]*sigma[j, t]*z[i, t]*z[j, t]}{\[\displaystyle x_{{i,t}}x_{{j,t}}=\sigma_{{i,t}}\sigma_{{j,t}}z_{{i,t}}z_{{j,t}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
Since\mapleinline{inert}{2d}{sigma[i, t] in R[-2*alpha[i]]}{$\displaystyle \sigma_{{i,t}} \in R_{{-2\,\alpha_{{i}}}}$}
,\mapleinline{inert}{2d}{sigma[j, t] in R[-2*alpha[j]]}{$\displaystyle \sigma_{{j,t}} \in R_{{-2\,\alpha_{{j}}}}$}
,\mapleinline{inert}{2d}{x[i, t]*x[j, t] in R[-2*min*{alpha[i], alpha[j]}]}{$\displaystyle x_{{i,t}}x_{{j,t}} \in R_{{-2\, \left( \min \right) \, \left\{ \alpha_{{i}},\alpha_{{j}} \right\} }}$}
}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{(sum(x[i, t]*x[j, t]*`#mover(mo("&rarr;"),mi("v"))`*xi[i, j], t = 1 .. n))/a[n]}{\[\displaystyle {\frac {\sum _{t=1}^{n}x_{{i,t}}x_{{j,t}}\mbox {{\tt `\#mover(mo("`}}\rightarrow \mbox {{\tt `"),mi("v"))`}}\\
\mbox{}\\
\mbox{}\,\xi_{{i,j}}}{a_{{n}}}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
where\mapleinline{inert}{2d}{a[n]}{$\displaystyle a_{{n}}$}
satisfies\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mi("P", italic = "true", mathvariant = "italic"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), subscriptshift = "0"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), subscriptshift = "0"), Typesetting:-mo(">", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), subscriptshift = "0")), mathvariant = "normal"), Typesetting:-mo("~", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.1111111em", rspace = "0.1111111em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"))}{$\displaystyle P \left(x _{i ,t }\cdot x _{j ,t }>a _{n }\right)\estilde \frac{1}{n }$}
.}\end{Maple Normal}

}\end{Maple Normal}

\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{(sum(x[i, t]^2, t = 1 .. n))/n}{\[\displaystyle {\frac {\sum _{t=1}^{n}{x_{{i,t}}}^{2}}{n}}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false")), mathvariant = "normal")), Typesetting:-mi(""))}{\[\displaystyle =\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\mathop{\rm  }\left(\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,1}b _{i ,1}^{l -1}x _{i ,t -l }^{2}+\frac{a _{i ,0}\left(2-b _{i ,1}\right)}{1-b _{i ,1}}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"))}{\[\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,1}b _{i ,1}^{l -1}\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\cdot x _{i ,t -l }^{2}+\frac{a _{i ,0}\left(2-b _{i ,1}\right)}{1-b _{i ,1}}\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\mathop{\rm  }\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi(""), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), mathvariant = "normal")), mathvariant = "normal", open = "[", close = "]"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mi(""), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("x", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mfrac(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-mfrac(Typesetting:-mn("1", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), linethickness = "1", denomalign = "center", numalign = "center", bevelled = "false"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("n", italic = "true", mathvariant = "italic")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"))}{$\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,1}b _{i ,1}^{l -1}\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\cdot x _{i ,t -l }^{2}-\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\right)\mathop{\rm  }\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}x _{i ,t -l }^{2}\right)\right)+\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,1}b _{i ,1}^{l -1}\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}\right)\mathop{\rm  }\left(\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}x _{i ,t -l }^{2}\right)+\frac{a _{i ,0}\left(2-b _{i ,1}\right)}{1-b _{i ,1}}\frac{1}{n }\Mapleoverset{n }{\Mapleunderset{t =1}{\sum }}z _{i ,t }^{2}$}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{sigma[i, t]^2 = sum((b[i, 1]*sigma[i, t-l]^2+a[i, 0])*a[i, 1]^(l-1)*(product(z[i, t-k]^2, k = 1 .. l-1)), l = 1 .. r)+a[i, 1]^r*x[i, t-r]^2*(product(z[i, t-k]^2, k = 1 .. l-1))}{$\displaystyle {\sigma_{{i,t}}}^{2}=\sum _{l=1}^{r} \left(  \left( b_{{i,1}}{\sigma_{{i,t-l}}}^{2}+a_{{i,0}} \right) {a_{{i,1}}}^{l-1}\\
\mbox{}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2} \right) +{a_{{i,1}}}^{r}{x_{{i,t-r}}}^{2}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2}$}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi(""))}{\[\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}\left(a _{i ,0}+b _{i ,1}\sigma _{i ,t -l }^{2}\right)a _{i ,1}^{l -1}\mathop{\rm  }\cdot \Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi(""))}{$\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}a _{i ,0}a _{i ,1}^{l -1}\mathop{\rm  }\cdot \Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}+\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}b _{i ,1}\cdot a _{i ,1}^{l -1}\cdot \sigma _{i ,t -l }^{2}\cdot \Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}$}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{sigma[i, t]^2*sigma[j, t]^2 = (sum(a[i, 0]*a[i, 1]^(l-1)*(product(z[i, t-k]^2, k = 1 .. l-1)), l = 1 .. infinity)+sum(b[i, 1]*a[i, 1]^(l-1)*sigma[i, t-l]^2*(product(z[i, t-k]^2, k = 1 .. l-1)), l = 1 .. infinity))*(sum(a[j, 0]*a[j, 1]^(p-1)*(product(z[j, t-q]^2, q = 1 .. p-1)), p = 1 .. infinity)+sum(b[j, 1]*a[j, 1]^(p-1)*sigma[j, t-p]^2*(product(z[j, t-q]^2, q = 1 .. p-1)), p = 1 .. infinity))}{\[\displaystyle {\sigma_{{i,t}}}^{2}{\sigma_{{j,t}}}^{2}= \left( \sum _{l=1}^{\infty } \left( a_{{i,0}}{a_{{i,1}}}^{l-1}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2} \right) +\sum _{l=1}^{\infty } \left( b_{{i,1}}{a_{{i,1}}}^{l-1}{\sigma_{{i,t-l}}}^{2}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2} \right) \\
\mbox{} \right)  \left( \sum _{p=1}^{\infty } \left( a_{{j,0}}{a_{{j,1}}}^{p-1}\prod _{q=1}^{p-1}{z_{{j,t-q}}}^{2} \right) +\sum _{p=1}^{\infty } \left( b_{{j,1}}{a_{{j,1}}}^{p-1}{\sigma_{{j,t-p}}}^{2}\prod _{q=1}^{p-1}{z_{{j,t-q}}}^{2} \right) \\
\mbox{} \right) \]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&Sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mi("&infin;", italic = "false", mathvariant = "normal"), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Product;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("q", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("q", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&Sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), accent = "false", accentunder = "false"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&sdot;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("p", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Product;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("q", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("q", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mi(""))}{$\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}\mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{p =1}{\sum }}\mathop{\rm  }a _{i ,0}\mathop{\rm  }a _{j ,0}\mathop{\rm  }a _{i ,1}^{l -1}~a _{j ,1}^{p -1}\mathop{\rm  }\left(\Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}\right)\mathop{\rm  }\left(\Mapleoverset{p -1}{\Mapleunderset{q =1}{\prod }}\mathop{\rm  }z _{j ,t -q }^{2}\right)+\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}\mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{p =1}{\sum }}b _{i ,1}b _{j ,1}\cdot a _{i ,1}^{l -1}a _{j ,1}^{p -1}\cdot \sigma _{i ,t -l }^{2}\mathop{\rm  }\sigma _{j ,t -p }^{2}\mathop{\rm  }\left(\Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}\right)\mathop{\rm  }\left(\Mapleoverset{p -1}{\Mapleunderset{q =1}{\prod }}\mathop{\rm  }z _{j ,t -q }^{2}\right)$}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{sum(sum(a[i, 0]*a[i, 1]^(l-1)*b[j, 1]*a[j, 1]^(p-1)*sigma[j, t-p]^2*(product(z[i, t-k]^2, k = 1 .. l-1))*(product(z[j, t-q]^2, q = 1 .. p-1)), p = 1 .. infinity), l = 1 .. infinity)+sum(sum(b[i, 1]*a[i, 1]^(l-1)*sigma[i, t-l]^2*a[j, 0]*a[j, 1]^(p-1)*(product(z[i, t-k]^2, k = 1 .. l-1))*(product(z[j, t-q]^2, q = 1 .. p-1)), p = 1 .. infinity), l = 1 .. infinity)}{$\displaystyle \sum _{l=1}^{\infty } \left( \sum _{p=1}^{\infty } \left( a_{{i,0}}{a_{{i,1}}}^{l-1}b_{{j,1}}{a_{{j,1}}}^{p-1}\\
\mbox{}{\sigma_{{j,t-p}}}^{2}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2}\prod _{q=1}^{p-1}{z_{{j,t-q}}}^{2} \right)  \right) +\sum _{l=1}^{\infty } \left( \sum _{p=1}^{\infty } \left( b_{{i,1}}{a_{{i,1}}}^{l-1}{\sigma_{{i,t-l}}}^{2}a_{{j,0}}\\
\mbox{}{a_{{j,1}}}^{p-1}\prod _{k=1}^{l-1}{z_{{i,t-k}}}^{2}\prod _{q=1}^{p-1}{z_{{j,t-q}}}^{2} \right)  \right) $}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
\mapleinline{inert}{2d}{Typesetting:-mrow(Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-munderover(Typesetting:-mo("&sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", foreground = "[128,0,128]", placeholder = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-munderover(Typesetting:-mo("&Sum;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mo("&infin;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), accent = "false", accentunder = "false"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo("&InvisibleTimes;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&prod;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("k", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("l", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("k", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-munderover(Typesetting:-mo("&Product;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "true", symmetric = "false", largeop = "true", movablelimits = "true", accent = "false", lspace = "0.0em", rspace = "0.1666667em"), Typesetting:-mrow(Typesetting:-mi("q", italic = "true", mathvariant = "italic"), Typesetting:-mo("=", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2777778em", rspace = "0.2777778em"), Typesetting:-mn("1", mathvariant = "normal")), Typesetting:-mrow(Typesetting:-mi("p", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mn("1", mathvariant = "normal")), accent = "false", accentunder = "false"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("z", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("q", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-mfenced(Typesetting:-mrow(Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-mi(""), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("p", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0"), Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("a", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("0", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mi(""), Typesetting:-mrow(Typesetting:-mo("+", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msub(Typesetting:-mi("b", italic = "true", mathvariant = "italic"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mn("1", mathvariant = "normal")), subscriptshift = "0"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("i", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("l", italic = "true", mathvariant = "italic")), Typesetting:-mrow(Typesetting:-mn("2", mathvariant = "normal")), superscriptshift = "0", subscriptshift = "0", Typesetting:-msemantics = "[none,none]"), Typesetting:-mo(" ", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.0em"), Typesetting:-msubsup(Typesetting:-mi("&sigma;", italic = "false", mathvariant = "normal"), Typesetting:-mrow(Typesetting:-mi("j", italic = "true", mathvariant = "italic"), Typesetting:-mo(",", mathvariant = "normal", fence = "false", separator = "true", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.0em", rspace = "0.3333333em"), Typesetting:-mrow(Typesetting:-mi("t", italic = "true", mathvariant = "italic"), Typesetting:-mo("&minus;", mathvariant = "normal", fence = "false", separator = "false", stretchy = "false", symmetric = "false", largeop = "false", movablelimits = "false", accent = "false", lspace = "0.2222222em", rspace = "0.2222222em"), Typesetting:-mi("p", italic = "true", mathvariant = "italic")), Typesetting:-mi("")), Typesetting:-mn("2", mathvariant = "normal"), superscriptshift = "0", subscriptshift = "0")), Typesetting:-mi("")), mathvariant = "normal"))}{\[\displaystyle =\Mapleoverset{\infty }{\Mapleunderset{l =1}{\sum }}\mathop{\rm  }\Mapleoverset{\infty }{\Mapleunderset{p =1}{\sum }}a _{i ,1}^{l -1}~a _{j ,1}^{p -1}\left(\Mapleoverset{l -1}{\Mapleunderset{k =1}{\prod }}z _{i ,t -k }^{2}\right)\mathop{\rm  }\left(\Mapleoverset{p -1}{\Mapleunderset{q =1}{\prod }}\mathop{\rm  }z _{j ,t -q }^{2}\right)\mathop{\rm  }\left(a _{i ,0}\mathop{\rm  }a _{j ,0}+a _{i ,0}b _{j ,1}\sigma _{j ,t -p }^{2}+a _{j ,0}b _{i ,1}\sigma _{i ,t -l }^{2}+b _{i ,1}b _{j ,1}\sigma _{i ,t -l }^{2}\mathop{\rm  }\sigma _{j ,t -p }^{2}\right)\]}
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\begin{Maple Normal}{
\begin{Maple Normal}{
}\end{Maple Normal}
}\end{Maple Normal}
\end{document}

### 2016-09-15 papers/FX/fx.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}
\usepackage{graphicx}

\newif\ifALL

\ALLtrue

\input{../../physics_common}
\title{On FX Rates and a Multivariate GARCH model}
\author{Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

\section{Notation}
We study the covariance matrix of a $p$-dimensional sequence with each
component modeled by a GARCH(1, 1) process:

\begin{eqnarray*}
  X_{i,t} &=& \sigma_{i,t} Z_{i,t} \\
  (Z_{1, t}, Z_{2,t}, \dots, Z_{p,t}) &\sim& N(0, \Sigma) \\
  \sigma_{i, t}^2 &=& \alpha_{i,0} + \alpha_{i,1} X_{i, t-1}^2 + \beta_{i,1} \sigma_{i, t-1}^2
\end{eqnarray*}

\section{Spectra of S\&P sectors}
\subsection{Utilities}
Fitting the daily returns' data of 25 companies that have been
included in the S\&P 500 index since 2010-01-01 and categorized in the
sector ``Utilities'', we obtained 23 GARCH(1, 1) models that satisfy
the condition
\begin{equation}
  \label{eq:garch_cond1}
  \alpha_1 + \beta_1 < 1  
\end{equation}
Then we simulate the GARCH(1, 1) processes with innovations having the
same covariance structure as those of the real data. Figure
\ref{fig:Utilities_eigenvalues} and \ref{fig:Utilities_eigenvectors1},
\ref{fig:Utilities_eigenvectors2}
compare the eigenvalues and eigenvectors, respectively, of the 
covariance matrices of the real and the simulated data.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6]{Utilities_eigenvalues.pdf}
  \caption{Eigenvalues of Utilities: real \& simulated}
  \label{fig:Utilities_eigenvalues}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Utilities_eigenvectors1.pdf}
  \caption{Eigenvectors of Utilities: real \& simulated}
  \label{fig:Utilities_eigenvectors1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Utilities_eigenvectors2.pdf}
  \caption{Eigenvectors of Utilities: real \& simulated continued}
  \label{fig:Utilities_eigenvectors2}
\end{figure}

\subsection{Materials}
A similar analysis on the ``Materials'' sector produces the following
spectra. In this sector, 21 out of 25 return sequencies are
successfully fitted to GARCH(1, 1) models that satisfy condition
\eqref{eq:garch_cond1}.

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.6]{Materials_eigenvalues.pdf}
  \caption{Eigenvalues of Materials: real \& simulated}
  \label{fig:Materials_eigenvalues}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Materials_eigenvectors1.pdf}
  \caption{Eigenvectors of Materials: real \& simulated}
  \label{fig:Materials_eigenvectors1}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{Materials_eigenvectors2.pdf}
  \caption{Eigenvectors of Materials: real \& simulated continued}
  \label{fig:Materials_eigenvectors2}
\end{figure}

\bibliographystyle{unsrt}
\bibliography{../thesis/econophysics}
\end{document}

### 2016-09-19 papers/FX/fx.tex
Summarized the results about the multivariate GARCH model with Thomas and Olivier.

@@
 
-\section{Notation}
+\section{Introduction}
 We study the covariance matrix of a $p$-dimensional sequence with each
@@
 
+\section{IGARCH(1,1) Processes}
+We first consider the cases where $\alpha_{i, 1} + \beta_{i, 1} = 1$, i.e. IGARCH models.
+\subsection{IARCH(1)}
+In this case $\alpha_{i, 1} = 1$, $\beta_{i, 1} = 0$. Thus, from the equation
+\begin{equation*}
+  \E (Z_i^2)^{\xi_i} = 1
+\end{equation*}
+we obtain $\xi_i = 1$. Suppose
+\begin{eqnarray*}
+  \E (Z_i^2 Z_j^2)^\xi &=& 1 \\
+  (Z_i, Z_j) &\sim& N(0, \Sigma) \\
+  \Sigma &=&
+  \begin{pmatrix}
+    1 & \rho \\
+    \rho & 1
+  \end{pmatrix}
+\end{eqnarray*}
+Re-writing the above equation gives
+\begin{equation}
+  \label{eq:iarch1_1st}
+  {1 \over 2\pi}
+  \int \int
+  \left(
+    \rho^2 x^4 + (1 - \rho^2) x^2 y^2 + 2 \rho \sqrt{1 - \rho^2} x^3 y
+  \right)^\xi
+  \exp\left(
+    -{x^2 + y^2 \over 2}
+  \right)
+  dx dy = 1
+\end{equation}
+When $\rho = \pm 1$, it is clear $\xi = 1/2$. When $-1 < \rho < 1$,
+we may re-write \eqref{eq:iarch1_1st} in polar coordinates to obtain
+\begin{equation*}
+  {1 \over 2\pi} \int_0^{\infty} r^{4\xi + 1} e^{-r^2 / 2} dr
+  \int_{0}^{2\pi} \cos(\theta)^{2 \xi}
+  \left(
+    2 \rho^2 \cos(\theta)^2
+    + 2 \rho \sqrt{1 - \rho^2} \cos(\theta) \sin(\theta) - \cos(\theta)^2 + 1 - \rho^2
+  \right)^\xi d\theta = 1
+\end{equation*}
+Integrating the radial part of the integral results in
+\begin{equation*}
+  \int_{0}^{2\pi}
+  \cos(\theta)^{2\xi}
+  \left(
+    2 \rho^2 \cos(\theta)^2 + \rho \sqrt{1 - \rho^2} \sin(2\theta) - \cos(\theta)^2 - \rho^2 + 1
+  \right)^\xi d\theta
+  = {2 \pi
+    \over
+    2^{2 \xi} \Gamma(2\xi + 1)
+  }
+\end{equation*}
+After some manipulations we arrive at
+\begin{equation}
+  \label{eq:iarch_2nd}
+  \int_{-1}^1 (\sin(\pi z) + \rho)^{2\xi} dz = {2 \over \Gamma(2\xi + 1)}
+\end{equation}
+Using this result, we immediately obtain for ARCH(1) models
+\begin{equation*}
+  \label{eq:arch_1st}
+  \int_{-1}^1 (\sin(\pi z) + \rho)^{2\xi} dz =
+  {2 \over \Gamma(2\xi + 1)}
+  {1 \over (\alpha_i \alpha_j)^\xi}
+\end{equation*}
+
+\section{General IARCH(1)}
+For general IARCH(1) models, the equation
+\begin{equation*}
+  \E \left[
+    (
+    \alpha_{11} Z_1^2 + 1 - \alpha_{11}
+    )^\xi
+    (
+    \alpha_{11} Z_1^2 + 1 - \alpha_{11}
+    )^\xi\right] = 1
+\end{equation*}
+has to be solved numerically.
 \section{Spectra of S\&P sectors}

### 2016-09-20 papers/FX/fx.tex
Investigated numericaly integration and analytical integration for IGARCH(1, 1) models Importance sampling for Jeff.

@@
 
-\section{IGARCH(1,1) Processes}
+\section{IGARCH(1) Processes}
 We first consider the cases where $\alpha_{i, 1} + \beta_{i, 1} = 1$, i.e. IGARCH models.
+
 \subsection{IARCH(1)}
@@
   {2 \over \Gamma(2\xi + 1)}
-  {1 \over (\alpha_i \alpha_j)^\xi}
+  {1 \over (\alpha_{i, 1} \alpha_{j, 1})^\xi}
 \end{equation*}
 
-\section{General IARCH(1)}
-For general IARCH(1) models, the equation
+\section{General IGARCH(1)}
+For general IGARCH(1) models, the equation
 \begin{equation*}
@@
     (
-    \alpha_{11} Z_1^2 + 1 - \alpha_{11}
+    \alpha_{i,1} Z_1^2 + 1 - \alpha_{i,1}
     )^\xi
     (
-    \alpha_{11} Z_1^2 + 1 - \alpha_{11}
+    \alpha_{j,1} Z_2^2 + 1 - \alpha_{j,1}
     )^\xi\right] = 1

### 2016-09-20 cpp/ImportanceSampling/main.cpp

@@ -175,9 +175,12 @@ int main(int argc, char* argv[])
 	1.0e-7, 0.11, 0.88
     };
     Garch1D<double> garch11(coef[0], coef[1], coef[2], stol(argv[1]));
+    printf("Lambda(%.4f) = %.4f\n", garch11.xi, log(garch11.moment_func(garch11.xi, 0, 0)));
+    for (double x = stod(argv[2]); x <= stod(argv[3]); x += 0.1) {
+	printf("%.4f    %.4f\n", x, log(garch11.moment_func(x)));
+    }
 
-    printf("lambda(-%.4f) = %.4f\n", garch11.xi, garch11.moment_func(-garch11.xi, 0, 0));
-
+#if 0
     /**
      * Find M. The set C = [-M, M]
      */
@@ -302,6 +305,6 @@ int main(int argc, char* argv[])
     // for (int i = 0; i < stoi(argv[1]); i++) {
     // 	cout << test() << endl;
     // }
-    
+#endif    
     return 0;
 }

### 2016-10-03 papers/FX/fx.tex
some progress on the multivariate GARCH model.

@@
 \usepackage{graphicx}
+\usepackage{subfigure}
 
@@
 \end{equation*}
+Fitting ARCH(1) models to the FX return series and then simulating the
+models produces a fairly good match between the spectra of the
+simulated and the real series, as shown in figure
+\ref{fig:FX_ARCH_eigenvalues} and \ref{fig:FX_ARCH_eigenvectors}.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4]{FX_eigenvalues.pdf}  
+  \caption{FX \& simulated ARCH eigenvalues}
+  \label{fig:FX_ARCH_eigenvalues}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.3]{FX_eigenvectors}  
+  \caption{FX \& ARCH eigenvectors}
+  \label{fig:FX_ARCH_eigenvectors}
+\end{figure}
 
@@
 \end{equation*}
-has to be solved numerically.
+has to be solved numerically. Figure \ref{fig:xi_rho0.5} shows $\xi$
+as a function of $(\alpha_{i,1}, \alpha_{j,1})$ as defined by the
+above equation.
+\begin{figure}[htb!]
+  \centering
+  % \includegraphics[scale=0.4]{igarch_rho0dot5.pdf}
+  \caption{$\xi(\alpha_1, \alpha_2)$ for $\rho = 1/2$}
+  \label{fig:xi_rho0.5}
+\end{figure}
+
 \section{Spectra of S\&P sectors}

### 2016-10-05 papers/FX/fx.tex
Some revision to the FX project and some to the GARCH(p,q) project.

@@
   \centering
-  \includegraphics[scale=0.3]{FX_eigenvectors}  
+  \includegraphics[scale=0.3]{FX_eigenvectors.pdf}  
   \caption{FX \& ARCH eigenvectors}
@@
   \centering
-  % \includegraphics[scale=0.4]{igarch_rho0dot5.pdf}
+  \includegraphics[scale=0.4]{igarch_rho0dot5.pdf}
   \caption{$\xi(\alpha_1, \alpha_2)$ for $\rho = 1/2$}
   \label{fig:xi_rho0.5}
+\end{figure}
+
+\section{Heavy-Tailed Innovations}
+Figure \ref{fig:ARCH_t_FX_eigenvalues} and Figure
+\ref{fig:ARCH_t_FX_eigenvectors} show the eigenvalues and
+eigenvectors of the simulated GARCH(1, 1) process with innovations
+following t(4) distribution. Clearly, the spectrum of the simulated
+series does not match that of the observed FX series.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4]{FX_eigenvalues_ARCH_t_inno.pdf}  
+  \caption{FX \& simulated ARCH eigenvalues. t-distributed innovations}
+  \label{fig:ARCH_t_FX_eigenvalues}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.3]{FX_eigenvectors_ARCH_t_inno.pdf}  
+  \caption{FX \& ARCH eigenvectors. t-distributed innovations}
+  \label{fig:ARCH_t_FX_eigenvectors}
 \end{figure}
@@
   \centering
-  \includegraphics[scale=0.6]{Utilities_eigenvalues.pdf}
+  \includegraphics[scale=0.4]{Utilities_eigenvalues.pdf}
   \caption{Eigenvalues of Utilities: real \& simulated}

### 2016-10-05 cpp/ImportanceSampling/ExtremeNumber2.cpp

@@ -180,6 +180,11 @@ double ExtremeNumber::operator () (long n) const
     return sign == 1 ? x : -x;
 }
 
+ExtremeNumber::operator double() const
+{
+    return pow(10, mylog);
+}
+
 long ExtremeNumber::power(void) const
 {
     return (long)floor(mylog);
@@ -271,6 +276,14 @@ double log(const ExtremeNumber &x)
 {
     return log10(x) * log(10);
 }
+
+// ExtremeNumber log10(const ExtremeNumber &x)
+// {
+//     assert(x.sign == 1);
+//     ExtremeNumber y = x;
+//     y.mylog = log10(x.mylog);
+//     return y;
+// }
 
 ExtremeNumber exp(const ExtremeNumber &x)
 {

### 2016-10-05 cpp/ImportanceSampling/ExtremeNumber2.hpp

@@ -32,6 +32,7 @@ public:
     
     double operator() (void) const;
     double operator() (long n) const;
+    operator double() const;
     long power(void) const;
 };
 
@@ -46,10 +47,13 @@ ExtremeNumber operator - (const ExtremeNumber &x, const ExtremeNumber &y);
 ExtremeNumber operator - (const ExtremeNumber &x, double y);
 ExtremeNumber operator ^ (const ExtremeNumber &x, double u);
 
+
 ExtremeNumber abs(const ExtremeNumber &x);
+ExtremeNumber exp(const ExtremeNumber &x);
+//ExtremeNumber log10(const ExtremeNumber &x);
 double log10(const ExtremeNumber &x);
 double log(const ExtremeNumber &x);
-double exp(const ExtremeNumber &x);
+
 
 
 #endif

### 2016-10-23 r/garch11_processes.r
Presentation for Rotterdam.

@@ -69,9 +69,10 @@ X <- getAssetReturns("2010-01-04", "2016-04-01", currencies, 1,
                      "rate", "localhost");
 n <- dim(X)[1];
 p <- dim(X)[2];
+X <- X - matrix(rep(apply(X, FUN=mean, MARGIN=2), n), nrow=n, ncol=p, byrow=TRUE);
 
 inno <- matrix(NA, nrow=n, ncol=p);
-coef <- matrix(0, nrow=p, ncol=3);
+params <- matrix(0, nrow=p, ncol=3);
 ics <- matrix(NA, nrow=p, ncol=4);
 for (i in 1:p) {
     ## spec <- ugarchspec(
@@ -96,7 +97,7 @@ for (i in 1:p) {
     ##     }
     ##     inno[, i] <- residuals(M1) / sigma(M1);
     ##     inno[, i] <- inno[, i] - mean(inno[, i]);
-    ##     coef[i, ] <- coef(M1);
+    ##     params[i, ] <- coef(M1);
     ## }
     ## if (is.null(M1) || convergence(M1)) {
     ##     stop(sprintf("Estimation failed for %d", i));
@@ -105,15 +106,14 @@ for (i in 1:p) {
     M <- garchFit(~garch(1,1),
                   data=X[, i],
                   trace=FALSE,
-                  cond.dist="std",
-                  shape=4,
+                  ## cond.dist="std",
+                  ## shape=4,
                   include.shape=FALSE,
-                  include.mean=TRUE,
+                  include.mean=FALSE,
                   include.delta=FALSE,
                   include.skew=FALSE
                   );
-    ## coef[i, ] <- M@fit$params$params[c(2,3,5)];
-    coef[i, 1:(length(M@fit$coef) - 1)] <- M@fit$coef[-1];
+    params[i, ] <- coef(M);
     inno[, i] <- M@residuals / M@sigma.t;
     inno[, i] <- inno[, i] - mean(inno[, i]);
     inno[, i] <- inno[, i] / sd(inno[, i]);
@@ -148,15 +148,15 @@ W <- matrix(NA, nrow=100*n, ncol=p);
 sig2 <- matrix(NA, nrow=dim(W)[1], ncol=p);
 # set the initial values
 for (i in 1:p) {
[…]

### 2016-10-28 r/FX.r
Revised the slides.

@@ -2,7 +2,6 @@ rm(list=ls());
 graphics.off();
 ## require("tseries");
 require("fGarch");
-require("rugarch");
 require("mvtnorm");
 source("libxxie.r");
 
@@ -75,9 +74,9 @@ N <- 8;
 
 E <- cov(X);
 Eig <- eigen(E);
-Y <- X %*% Eig$vectors[, 1:N];
-C <- t(Eig$vectors);
-C <- head(C, N);
+C <-Eig$vectors[, 1:N];
+
+Y <- X %*% C;
 
 params <- matrix(0, nrow=N, ncol=3);
 for (i in 1:N) {
@@ -121,16 +120,17 @@ D <- cov(W);
 ##     Res[, i] <- residuals(model);
 ## }
 ## Dep <- cov(Res);
-F <- t(C) %*% D %*% C;
+F <- C %*% D %*% t(C);
 Fig <- eigen(F);
 
-pdf("/tmp/FX_eigenvalues_comparison.pdf", width=10, height=10);
+pdf("/tmp/FX_OGARCH_eigenvalues.pdf", width=10, height=10);
 plot(1:p, Eig$values, type="p", pch=0,
      xlab=expression(i),
      ylab=expression(lambda[(i)]),
+     cex=2,
      main="FX and sim. spectrum", ylim=c(0, max(Eig$values[1], Fig$values[1]))
 );
-points(1:p, (Fig$values), pch=16, col="#FF0000");
+points(1:p, (Fig$values), pch=16, col="#FF0000", cex=2);
 
 ## ## points(1:p, (E1$values)/sum(E1$values), col="#FF0000", cex=2, pch=15);
 ## ## points(1:p, (F1$values)/sum(F1$values), col="#00FF00", cex=2, pch=16);
@@ -144,10 +144,10 @@ legend("topright",
 grid();
 dev.off();
 
-pdf("/tmp/FX_eigenvectors_comparison.pdf", width=20, height=10);
-par(mfrow=c(3,6));
+pdf("/tmp/FX_OGARCH_eigenvectors.pdf", width=20, height=10);
+par(mfrow=c(2,4));
 ## mse <- c(0, 0);
-for (i in 1:p) {
+for (i in 1:8) {
     V <- Eig$vectors[, i];
     Q <- Fig$vectors[, i];
     
@@ -161,13 +161,13 @@ for (i in 1:p) {
     
     plot(1:p, V * s, main=sprintf("FX & sim. V[%d]", i),
[…]

### 2017-02-01 papers/FX/draft.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

\input{../../physics_common}
\title{Tail Indices and Scale Parameters in Financial Time Series}
\author{Casper de Vries, Thomas Valentin Mikosch, Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

\begin{abstract}
We consider an investor with preferences that accord with Generalized
Disappointment Aversion. Such an investor cares about downside risk and
we assume he recognizes the heavy tail feature of asset return
distributions. We argue that when a market is dominated by rational
investors of this kind, the return distributions of equities that are
actively traded in this market must have very similar tail indices.

On the other hand, whether or not all equities in a multivariate model
have the same tail index is a dividing issue for multivariate GARCH
models proposed in the literature. Therefore, it is important to analyze
data of real equity returns and see how close to each other the
tail indices actually are.

In this work empirical results are also presented and they appear
to support the conclusion that the tail indices are very similar,
with respect to the confidence bands of estimation.
\end{abstract}

\section{Introduction}
It is generally believed that the utility function of an
investor is increasing and concave -- as John H. Cochrane
put it, ``the last bite is never as satisfying as the first''
\cite{cochrane2009asset}. This implies investors are considerably
more concerned about large losses than are cheered by large gains.
Furthermore, it is a stylized fact that equity return distributions
are heavy-tailed (see e.g. Jansen and de Vries \cite{jansen1991frequency}).
In particular, the lower-tail of an equity distribution is often heavier
than the upper-tail of the same distribution.

We show in this paper that concerns of the down-side risk and the
heavy-tail property of equity returns  imply that the lower-tail indices
of actively traded equities in a rational market must be fairly close to
one another.

%% A cornerstone of this conclusion is the empirical fact that equity
%% return distributions are heavy-tailed (see e.g. Jansen and de Vries
%% (1991) \cite{jansen1991frequency}). If investors and speculators alike
%% display concern for downside risk, as formulated via a utility function,
%% it is important to model this risk adequately.

In section \ref{sec:GDA}, we explicitly recognize the behavioral
concern for downside risk in the investor's evaluation of portfolios
using the framework of Generalized Disappointment Aversion (GDA)
introduced by Routledge and Zin \cite{routledge2010generalized}. GDA
is an extension of the concept of Disappointment Aversion (DA) of Gul
(1991) \cite{gul1991theory}. The Gul paper derives the DA from first
principles (axiomatic).

%% Secondly, the empirical fact that the loss return distribution is
%% non-normally fat tailed distributed is taken into account.
%% If investors and speculators alike display concern for downside risk,
%% as formulated by a utility function, it becomes important to model this 
%% risk adequately. see e.g. Jansen and de Vries (1991), exhibiting power
%% like tails rather than exponential tails. The risk is determined both
%% by the coefficient that regulates the hyperbolic power and the
%% scale. We reason how both determine the risk of investment apart from
%% the mean return.

%% \section{Notations}
%% Here is a summary of the notations that we use in the rest of the paper.
%% In this summary $x_F$ denotes a quantity whose value that should be
%% obvious from the context where the notation in question is used.
%% \begin{itemize}
%% \item $\sim$ stands for ``asymptotically equivalent to'', i.e.
%%   $F(x) \sim G(x)$ means
%%   \[
%%   \lim_{x \to x_F} {F(x) \over G(x)} = 1
%%   \]
%% \item $O(F(x))$ stands for terms of the same order as $F(x)$, i.e.
%%   \[
%%   \lim_{x \to x_F } {O(F(x)) \over F(x)} = \text{constant}
%%   \]
%% \item $o(F(x))$ stands for terms of higher orders than $F(x)$, i.e.
%%   \[
%%   \lim_{x \to x_F } {o(F(x)) \over F(x)} = 0
%%   \]
%% \item $\hot$ stands for ``higher order terms'', i.e.
%%   \[
%%   \lim_{x \to x_F} {F(x) + \hot \over F(x)} = 1
%%   \]
%% \end{itemize}

%% \section{Implications of heavy-tailed equity return}
%% \label{sec:1}
%% In this section we consider a rather general utility functional
%% $U(F)$ associated with the distribution function $F(\cdot)$ of
%% an equity return $X$. We define $U(F)$ as
%% \[
%% U(F) = \E u(X)
%% \]
%% for some function $u(\cdot)$ satisfying the following conditions:
%% \begin{enumerate}
%% \item $u(\cdot)$ is defined on $\mathbb R$, is increasing, concave
%%   and differentiable.
%% \item $u(\cdot)$ is independent of the distribution of $X$
%% \item $\lim_{x \to \infty} u(x) = \bar u$, where $\bar u$ is a
%%   constant
%% \end{enumerate}
%% %% An exampple of such utility functions is
%% %% \[
%% %% u(x) = \bar u - e^{-\lambda x}
%% %% \]
%% %% for some $\lambda > 0$.
%% In addition we assume $F(\cdot)$ has the following properties:
%% \begin{enumerate}
%% \item $\lim_{x \to -\infty} F(x) u(x) = 0$
%% \item $\lim_{x \to -\infty} {|x|^\alpha F(x) \over L(|x|)} = 1$
%% \item $\lim_{x \to \infty} {x^\alpha (1 - F(x)) \over L(x)} = 1$
%% \end{enumerate}
%% for some $\alpha > 1$ and some slowly varying function $L(x) > 0$
%% defined on $\mathbb R_+$.
%% $\E u(X)$ can be computed as follows
%% \begin{eqnarray*}
%%   && \E u(X) \\
%%   &=& \int_{-\infty}^{\infty} u(x) d F(x) \\
%%   &=& \left. u(x) F(x) \right|_{-\infty}^{\infty}
%%   - \int_{-\infty}^\infty F(x) u'(x) dx \\
%%   &=& \bar u - \left(
%%   \int_{-\infty}^{-M} + \int_{-M}^{M} + \int_{M}^{\infty}
%%   \right) F(x) u'(x) dx
%% \end{eqnarray*}
%% As $M \to \infty$ we have for the lower-tail integral
%% \begin{eqnarray*}
%%   && \int_{-\infty}^{-M} F(x) u'(x) dx \\
%%   &\sim& \int_{-\infty}^{-M} L(|x|) |x|^{-\alpha} u'(x) dx \\
%%   &=& \int_{M}^{\infty} L(x) x^{-\alpha} u'(-x) dx \\
%% \end{eqnarray*}
%% Applying the same calculation to the upper-tail integral and
%% using $1 - F(x) \sim L(x) x^{-\alpha}$ yields
%% \begin{eqnarray}
%%   && \E u(X) \nonumber \\
%%   &=& \bar u - \int_{M}^\infty L(x) x^{-\alpha} u'(-x) dx
%%   -\int_{-M}^M F(x) u'(x) dx \nonumber \\
%%   && -\int_{M}^\infty [1 - L(x) x^{-\alpha}] u'(x) dx
%%   + \hot \nonumber \\
%%   &=& u(M) - \int_{M}^\infty L(x) x^{-\alpha} [u'(-x) - u'(x)] dx
%%   + \hot \nonumber \\
%%   && - \int_{-M}^M F(x) u'(x) dx \label{eq:xxie0}
%% \end{eqnarray}
%% In the simple case where $L(x) = A$, i.e. a constant, we have
%% \begin{eqnarray}
%%   \pd{\E u(X)}{\alpha} &=& \int_M^{\infty} A x^{-\alpha} \ln(x)
%%      [u'(-x) - u'(x)] dx +\hot \nonumber \\
%%      \pd{\E u(X)}{A} &=& -\int_M^{\infty} x^{-\alpha} [u'(-x) - u'(x)] dx
%%      + \hot
%%   \label{eq:xxie01}   
%% \end{eqnarray}
%% By concavity $u'(-x) > u'(x)$ for $x \in (M, \infty)$. So for large
%% $M$,
%% \[
%% \pd{\E u(X)}{\alpha} > 0, \quad \pd{\E u(X)}{A} < 0
%% \]
%% %% \begin{eqnarray*}
%% %%   \pd{\E u(X)}{\alpha} &>& 0 \\
%% %%   \pd{\E u(X)}{A} &<& 0
%% %% \end{eqnarray*}
%% Now that each and every equity is assigned a score by the utility
%% functional, all equities actively traded in a rational market that
%% is dominated by agents exhibiting a utility function must have almost
%% equal scores -- otherwise trading will be concentrated on those
%% with top scores.

%% Suppose the equity with the highest score in the market has parameters
%% $(\alpha, A)$. We would like to find out how these parameters vary across
%% equities that score almost as high. Consider a parameter pair
%% $(\alpha + \Delta \alpha, A + \Delta A)$. If it scores the same as
%% $(\alpha, A)$, we have
%% \[
%% 0 = \Delta \E u(X) = \Delta \alpha \pd{\E u(X)}{\alpha}
%% + \Delta A \pd{\E u(X)}{A} + \hot
%% \]
%% Discarding higher order terms and plugging \eqref{eq:xxie01} into
%% this equation gives
%% \[
%%   {\Delta \alpha \over \Delta A}
%%   =
%%   {
%%     \int_M^{\infty} x^{-\alpha} [u'(-x) - u'(x)] dx    
%%     \over
%%     \int_M^{\infty} A x^{-\alpha} \ln(x) [u'(-x) - u'(x)] dx    
%%   }
%% \]
%% Taking the limit as $M \to \infty$, we get
%% \[
%% \lim_{M \to \infty} {\Delta \alpha \over \Delta A}
%% =
%% \lim_{M \to \infty} {1 \over A \ln(M)} = 0
%% \]
%% We see from the last equation that in the configuration space around
%% $(\alpha, A)$, a small variation of $\alpha$ induces a tremendous
%% variation of $A$. Furthermore, this conclusion does not depend on the
%% choice of the function $u(\cdot)$. Thus in a market dominated by rational
%% agents guided by their utility functionals, we can expect to see very
%% similar tail indices but hugely different scale parameters in equity
%% return distributions.

%% %% Due to concavity, $u'(-x) > u'(x)$ on the interval $(M, \infty)$.
%% %% The central region integral on $(-M, M)$ does not depend on $\alpha$.
%% %% So $\E u(X)$ is monotonically increasing with $\alpha$. Trying to
%% %% maximize his utility function, an agent will seek the largest $\alpha$
%% %% possible in the market. As a result, all equities traded with significant
%% %% volumes must eventually become so close that they are indistinguishable
%% %% by available statistical methods.

%% In section \ref{sec:GDA}, we compute the utility function within the
%% elaborate framework of {\it Generalized Disappointment Aversion} and
%% show that the same conclusion stands -- tail indices are similar while
%% scale parameters can be drastically different.

\section{Generalized disappointment aversion}
\label{sec:GDA}
Routledge and Zin (2010) \cite{routledge2010generalized} define the GDA
preferences over discrete states in their formula (2). A slight extension
is to consider a continuum of possible states. Suppose the state variable
$C$ is continuously supported on $(0, \infty)$ and has distribution
function $G(\cdot)$. The utility function of an agent exhibiting GDA
preferences reads
\begin{equation}
  u(v)= \E u(C) - b \int_{0}^{\delta v}
  \left[ u(\delta v) - u(x) \right] dG(x)\label{11}%
\end{equation}
where $\delta > 0$, and $v>0$. Equivalently
\begin{equation}
  \label{eq:xxie0}
  u(v) = \E u(C) - b \E[u(\delta v) - u(C); C < \delta v]
\end{equation}
Here $v$ can be thought of as the certainty payoff equivalent to the
risky payoff $C$. Note that when $b=0$, 
preferences are expected utility; when $\delta=1$ and $b>0$,
preferences follow Gul's (1991) original disappointment aversion.

\subsection{power-law utility function}
In this section we follow Routledge and Zin \cite{routledge2010generalized} and assume
that the utility function of consumption $C$ is a power function%
\[
u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%
\]
where $\gamma > 0$. For simplicity we assume the investor
initially has one unit of wealth. Let $r$ be the return on the
riskless state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation*}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation*}

\subsubsection{mildly risk-avert utility function}
We first consider the class of utility functions for which $\gamma < 1$.
In this case $\E u(C) < \infty$ only if the distribution of $X$, call
it $F(\cdot)$, has a light upper tail. We assume this for the rest of
this section.

We note that utility functions with $\gamma < 1$ represent investors
who are less risk-avert than are those represented by utility
functions with $\gamma > 1$.

Formally, we assume
\begin{equation*}
  F(x) = \left\{
  \begin{array}{lr}
    p M^\alpha/|x|^\alpha & \text{ if } x < -M \\
    F_u(x) & \text{ if } x \geq -M \\
  \end{array}
  \right.
\end{equation*}
where $0 < p < 1$ is a constant; $\alpha > 1$, $M > 1$ and $F_u(\cdot)$
is some distribution function satisfying
\begin{itemize}
\item $F_u(-M) = p$
\item
  \[
  \int_{-M}^\infty [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) < \infty
  \]
\end{itemize}
%% In addition we assume
%% \begin{equation}
%%   \label{eq:xxie0.1}
%% \phi < {4 e^{M + r} \over 4 e^{M + r} + 1}
%% \end{equation}

Rewriting \eqref{eq:xxie0} gives
\begin{eqnarray*}
  u(v) + b u(\delta v) &=& \E u(C) + b \E [u(C); C < \delta v] \\
  &=& \alpha p M^\alpha{1 + b \over 1 - \gamma}\int_{-\infty}^{-M}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
  && + {1 + b \over 1 - \gamma}\int_{-M}^{\delta v}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  && + \int_{\delta v}^{\infty}
  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
  &=& I_1 + I_2 + I_3
\end{eqnarray*}
Only the first term $I_1$ involves the tail parameters $\alpha$
and $M$ in which we are interested, so we will only compute this
term. The integral of $I_1$ can be written as
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=& \int_{M}^{\infty}
[(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} x^{-\alpha-1} dx
\end{eqnarray*}
Taylor expansion at $\phi e^{-x} = 0$ gives
\begin{eqnarray*}
  && [(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} \\
  &=& ((1 - \phi) e^r)^{1 - \gamma}
  + \sum_{n=1}^{\infty} {1 \over n!}
  {\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
  ((1 - \phi) e^r)^{1 - n - \gamma} \phi^n e^{-n x}
\end{eqnarray*}
Therefore
\begin{eqnarray*}
&& \int_{-\infty}^{-M}
[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
&=&
{1 \over \alpha} M^{-\alpha} ((1 - \phi) e^r)^{1 - \gamma}
+ \sum_{n=1}^{\infty} {1 \over n!}
{\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
((1 - \phi) e^r)^{1 - n - \gamma} n^\alpha \Gamma(-\alpha, n M)
\end{eqnarray*}
where $\Gamma(\alpha, x)$ is the upper incomplete gamma
function. For $x > 1$, this function may be expressed as (equation
8.11.2 of {\it NIST Digital Library of Mathematical Functions}
\cite{NIST.DLMF})
\[
\Gamma(a, z) = z^{a - 1} e^{-z}
\sum_{k=0}^\infty (-1)^k {\Gamma(a) \over \Gamma(a - k)} z^{-k}
\]
Thus after re-arranging terms we get
\begin{eqnarray}
  {1 \over 1 + b} I_3
  &=&
  {p \over 1 - \gamma} ((1 - \phi) e^r)^{1 - \gamma} \nonumber \\
  &&
  + \sum_{k=0}^\infty
  {\alpha p \over M^{k+1} ((1 - \phi) e^r)^{\gamma - 1}}
  \prod_{i=0}^{k-1} (\alpha + i)
  \sum_{n=1}^\infty {(-1)^{n-1} \over n! n^{k+1}}
  \left(
  {\phi \over e^{M + r} (1 - \phi)}
  \right)^n \times \nonumber \\
  && \prod_{i=1}^{n-1} (\gamma + i - 1)
  \label{eq:xxie1}
\end{eqnarray}
Observe that the sum over $n$ does not involve $\alpha$. In the
following we show this sum is positive for every $k \geq 0$, then it
is clear from \eqref{eq:xxie1} that $I_3$ and hence
$u(v) + b u(\delta v)$ is monotonically increasing with $\alpha$. To
see this, denote
\[
a_n(k) =
{(-1)^{n-1} \over n! n^{k+1}}
\left(
     {\phi \over e^{M + r} (1 - \phi)}
     \right)^n
     \prod_{i=1}^{n-1} (\gamma + i - 1)
\]
For each given $k$, $a_n(k)$ is an alternating series. To show the
positivity of $\sum_n a_n(k)$, it suffices to show
\[
  {a_n(k) \over |a_{n+1}(k)|} > 1
  \quad
  \text{ for } n=1,3,5,...
\]
Direct computation gives
\begin{eqnarray*}
  && {a_n(k) \over |a_{n+1}(k)|} \\
  &=& {
    e^{M + r} (1 - \phi) (n+1)^{k+2}
    \over
    \phi n^{k+1} (n + \kappa - 1)
  } \\
  &=&
  \left(1 + {2 - \kappa \over n + \kappa - 1}\right)
  \left( 1 + {1 \over n} \right)^{k+1}
  {e^{M + r} (1 - \phi) \over \phi}
\end{eqnarray*}

Thus we see that
$u(v) + u(\delta v)$ is monotonically increasing with $\alpha$. If
all investors are guided by and hence seek to maximize the same
utility function, they would all have constructed their portfolios by
choosing from a set of equities that have tail indices close to the
largest in the market. At market equilibrium, all actively traded
equities would have tail indices that are indistinguishable from each
other.

Notice that the summand of \eqref{eq:xxie1} decreases

\section{Review of multivariate GARCH models: same tail index or not}
Many multivariate GARCH models imply a common tail index for all
series under consideration, e.g. the Orthogonal GARCH model of
Alexander and Chibumba (1997) \cite{alexander1997multivariate}, its
generalization GO-GARCH of van der Weide (2002) \cite{van2002go}, as
well as the Full Factor GARCH model of Vrontos et al (2003)
\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
model of Lanne and Saikkonen (2007) \cite{lanne2007modeling}. These
models are characterized by their treatment of each series as a linear
combination of factors, and each of the factors is modeled as a GARCH
process. As a result of this treatment, each and every series under
consideration necessarily has the lowest of the tail indices of its
contributing factors (cf. Mikosch (2013) \cite{Mikosch2013} as well
as Andersen, Davis, Kreiss and Mikosch (2009)
\cite{andersen2009handbook}).

In contrast, several other multivariate volatility models imply, in
general, different tail indices for different series. For example, the
Constant Conditional Correlation (CCC-) GARCH model of Bollerslev
(1990) \cite{bollerslev1990modelling} specifies, for a $d$-dimensional
series $\vec X_t$
\begin{eqnarray}
  \cov(X_{i,t}, X_{j,t}) &=& \sqrt{h_{i,t} h_{j,t}} \rho_{i,j}
  \nonumber \\
  h_{i,t} &=& \omega_i + \sum_{l=1}^q A_{i,l} X_{i, t-l}^2 +
  \sum_{l=1}^p B_{i,l} h_{i, t-l} \label{eq:CCC-GARCH}
\end{eqnarray}
Consider the simple case when $p=q=1$. Assume
$X_{i, t} = \sqrt h_{i,t} Z_{i,t}$, $Z_{i, t}$ are iid for
all $t \in \mathbb Z$, and $\P(|Z_{i,t}| > u)/\P(\sqrt h_{i,t} > u) \to 0$ as
$u \to \infty$. Then the tail index of the stationary distribution of
$X_{i,t}$, call it $\kappa_i$, is given by
\[
\E (A_{i, 1} Z^2 + B_{i, 1})^{\kappa_i / 2} = 1
\]
where $Z \overset{d}{=} Z_{i,t}$.
Provided $(A_{i,1}, B_{i,1}) \neq (A_{j,1}, B_{j,1})$, it is clear
$\kappa_i \neq \kappa_j$.

As an extension to the CCC-GARCH model, Jeantheau (1998)
\cite{jeantheau1998strong} proposed his Extended CCC-GARCH by
modifying \eqref{eq:CCC-GARCH} to
\begin{equation}
  \label{eq:ECCC-GARCH}
  \vec h_t =
  \sum_{l=1}^q \mtx A_{l} (\vec X_{t-l} \odot \vec X_{t-l})
  +
  \sum_{l=1}^p \mtx B_{l} \vec h_{t-l}
  +
  \vec \omega
\end{equation}
where $\mtx A_l$ and $\mtx B_l$ are non-diagonal $d \times d$ matrices
and $\odot$ denotes element-wise multiplication. In the simple case $p=q=1$,
equation \eqref{eq:ECCC-GARCH} reduces to
\begin{eqnarray*}
  h_{i,t} &=&
  \sum_{j=1}^d A_{i,j} X_{j, t-1}^2
  + \sum_{j=1}^d B_{i,j} h_{j, t-1}
  + \omega_i
\end{eqnarray*}
or in vector form
\begin{eqnarray}
  \vec h_t &=& \left[
    \mtx A \diag(Z_{1, t-1}^2, \dots, Z_{d, t-1}^2) + \mtx B
    \right] \vec h_{t-1} + \vec \omega
  \label{eq:ECCC-GARCH2}
\end{eqnarray}
On appropriate conditions, \eqref{eq:ECCC-GARCH2} implies the
components of $\vec h_t$ are jointly regularly varying with some index
$\kappa$ as $t \to \infty$. This is a direct result of the Kesten-Goldie
theorem (cf. Kesten (1973)\cite{Kesten1973}). Thus all series under
consideration have the same tail index $\kappa$.

In summary, whether or not a common tail index is shared by all
series is a dividing issue for multivariate GARCH
models. Nevertheless, economic theoretical arguments are supportive of
a common tail index for all series. Therefore, it is important to
gather statistical evidence from financial time series if the issue is
to be settled definitely.

\section{Hill Estimator}
%% TODO: Show Hill estimates of real and simulated data with confidence
%% band. Argue that all estimated tail indices are indistinguishable
%% given the confidence band.
There have been several estimators for the tail index proposed in the
literature, e.g. Hill (1975) \cite{hill1975simple}, Pikands (1975)
\cite{pickands1975statistical}, Dekkers and Einmahl and De Haan (1989)
\cite{dekkers1989moment}. See also Embrechts and Kl\"uppelberg and
Mikosch (1997)\cite{Embrechts1997}. The most popular of these is
arguably Hill (1975) \cite{hill1975simple}:
\[
\hat \alpha = \left[
  {1 \over k} \sum_{i=1}^k \ln {x_{(i)} \over x_{(k+1)}}
  \right]^{-1}
\]
where $x_{(i)}$ is the $i$-th upper order statistic of a sample
$x_1, x_2, ..., x_n$. It has been shown that the Hill estimator is
applicable to 
\begin{itemize}
\item iid random variables (Mason (1982) \cite{mason1982laws})
\item a weakly dependent sequence (Rootz{\'e}n, Leadbetter,
  and de Haan \cite{rootzen1992tail})
\item a linear process (Resnick and St{\u{a}}ric{\u{a}} (1995)
  \cite{resnick1995consistency}).
\end{itemize}
In these cases it is shown that, if $k \to \infty$ and $k/n \to 0$
as $n \to \infty$, the Hill estimator is weakly consistent, i.e.
\[
\sqrt k (\hat \alpha - \alpha) \overset{d}{\to} N(0, \alpha^2)
\]

%% with continuous and strictly increasing
%% distribution function $F(\cdot)$. Cheng and Peng (2001)
%% \cite{cheng2001confidence} showed asymptotic normal distribution of
%% the estimator $\alpha_n$:
%% \begin{proposition}\label{prop:Hill_asymp_dist}
%%   Suppose $X_1, ..., X_n$ are independent identically distributed and
%%   their distribution function $F(\cdot)$ satisfies, as $x \to \infty$,
%%   \[
%%   1 - F(x) = A x^{-\alpha} [1 + B x^{-\beta} + o(x^{-\beta})]
%%   \]
%%   where $A, \alpha, \beta > 0$ and $B \neq 0$. Then
%%   \[
%%   \sqrt k (1/\alpha_n - 1/\alpha) \overset{d}{\to}
%%   N(0, {1 \over \alpha^2})
%%   \]
%%   if and only if $k(n) \to \infty$ and $k(n) = o(n^{2\beta/(\alpha +
%%     2\beta)})$.
%% \end{proposition}
%% \begin{proof}
%%   This is proposition 1 of Cheng and Peng (2001)
%%   \cite{cheng2001confidence}. See also Embrechts et al (1997)
%%   \cite{Embrechts1997}.
%% \end{proof}

\section{Peaks over Threshold}
An alternative to the Hill estimator is the method of ``Peaks over
Threshold'' (cf. Embrechts et al (1997)
\cite{Embrechts1997}). Consider random variable $X$ whose distribution
satisfies 
\[
\P(X > x) = A (x + a)^{-\alpha}
\]
for all $x > M$ for some $M > 0$. Then it immediately follows
\begin{eqnarray*}
  \P(X > u + x)
  &=& A (u + a)^{-\alpha} \left(
  {x \over u + a} + 1
  \right)^{-\alpha} \\
  \P(X > u + x | X > u) &=& \left(
    {x \over u + a} + 1
    \right)^{-\alpha}
\end{eqnarray*}
The density function of the above conditional probability is
\[
f_u(x) = {
  \alpha \over x + u + a
} \left(
{x \over u + a} + 1
\right)^{-\alpha}
\]

Now suppose we have a sample $X_1, ..., X_N$ where $X_1, ..., X_n > u$
and $X_i - u = x_i > 0$, for $i=1,2,...,n$. Using the above
conditional density function, the log-likelihood function of the
parameters $\alpha$ and $a$ can be found:
\[
L(u, a) = -(1 + \alpha) \sum_{i=1}^n \ln(x_i + u + a) +
u \alpha \ln(u + a) + n \ln(\alpha)
\]
It follows
\begin{eqnarray*}
  \pd{L(\alpha, a)}{\alpha} &=& \sum_{i=1}^n \ln(x_i + u + a)
  + n \ln(u + a) + n/\alpha \\
  \pd{L(\alpha, a)}{a} &=&
  -(1 + \alpha) \sum_{i=1}^n {1 \over x_i + u + a}
  + {n \alpha \over u + a}
\end{eqnarray*}
Setting the two derivatives to 0 gives
\begin{eqnarray*}
  \left[
    {1 \over n} \sum_{i=1}^n \ln(x_i + u + a)
    - \ln(u + a)
    \right]^{-1} &=& \alpha\\
  \left[
    {1 \over n} \sum_{i=1}^n (y_i + 1)
    \right] \left[
    {1 \over n} \sum_{i=1}^n e^{-y_i}
    \right] &=& 1
\end{eqnarray*}
where
\begin{eqnarray*}
  y_i &=& \ln\left(
  {x_i \over u + a} + 1
  \right)
\end{eqnarray*}
Solving the last equation numerically and substituting the result into
the first one gives an estimate for $\alpha$ and $a$, call them $\hat
\alpha$ and $\hat a$. Once $\hat \alpha$ and $\hat a$ have been
obtained, an estimate of $A$, call it $\hat A$, can be obtained by
the method of minimum mean square error.
\begin{eqnarray*}
  \hat P(X > x) = {1 \over N} \sum_{i=1}^N \1{X_i > x}
  &=& \hat A (x + \hat a)^{-\hat \alpha} \\
  \ln[\hat P(X > x)] &=& \ln(\hat A) - \hat \alpha \ln(x + \hat a)
\end{eqnarray*}
for all $x > M$. Using the upper order statistics of the
sample $X_1, ..., X_N$, we have
\[
\ln\left({k-1 \over N}\right)
=
\ln(\hat A) - \hat \alpha \ln(X_{(k)} + \hat a)
\]
where $X_{(k)}$ denotes the $k$-th upper order statistic of the
sample. Thus, by fitting $\ln\left({k-1 \over N}\right)$ as a linear
function of $\ln(X_{(k)} + \hat a)$, one can find $\ln(\hat A)$ as the
intercept.

Before computing the estimates, one should certainly check whether the
model makes sense for a given data series. One way to do this is to
plot the {\it mean excess function}. For the tail function in the
above, it follows
\begin{eqnarray}
  e(u) &=& \E (X - u | X > u) \nonumber \\
  &=& -\int_0^\infty x  d\P(X > u + x | X > u) \nonumber \\
  &=& \int_0^\infty \left(
  {x \over u + a} + 1
  \right)^{-\alpha} dx \nonumber \\
  &=& {u + a \over \alpha - 1} \label{eq:MEF}
\end{eqnarray}
where we assume $\alpha > 1$ so that the conditional expectation is
finite. It is seen from this formula that mean excesses are linear in
the thresholds $u$ -- this may prompt discarding the farthest tail
part of the data when carrying out estimation.
%% The appearance of these data points are so rare that
%% including them in the statistics significantly increases the variance
%% of the estimate. This is analogous to the large variance of the Hill
%% estimator when the number of upper order statistics used in the
%% estimation is too small.
\begin{wrapfigure}{r}{0.5\textwidth}
  \centering
  \includegraphics[width=0.45\textwidth]{ADM_MEF.pdf}
  \caption{\scriptsize The mean excess function of ADM. Thresholds are
    taken from upper order statistics.}
  \label{fig:ADM_MEF}
\end{wrapfigure}

For example, figure (\ref{fig:ADM_MEF}) shows the mean
excess function of the return series of the stock of {\it
  Archer-Daniels-Midland Co}, a company in the {\it Consumer Staples}
sector of S\&P 500 index.
One can see that the 8 largest observations of the sample are
inconsistent with the linear pattern with slope $1/(\alpha - 1)$.
%% So we would rather not include them in the maximum likelihood
%% estimation.

It is also possible to directly use equation \eqref{eq:MEF} as a means
of estimation. After picking out a portion of the mean excess function
that is consistent with \eqref{eq:MEF}, one may fit a linear function
to the empirical $\hat e(u)$ by minimizing the mean square error, and
obtain parameter estimates from the coefficients of the fitted line.
  %% \adjustbox{valign=t} {
  %%   \begin{minipage}[t]{0.45\textwidth}
  %%   \end{minipage}
  %% }\hfill
  %% \adjustbox{valign=t} {
  %%   \begin{minipage}[t]{0.5\textwidth}
  %%   \end{minipage}
  %% }
\section{Pickand's Estimator}
Pickands (1975) \cite{pickands1975statistical} proposed a different
approach to estimate the tail index than that of Hill. Assume
$\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
denote the quantile of $1 - 1/t$ of this distribution, i.e.
\[
P(X > U(t)) = {1 \over t}
\]
The key idea of Pickands' estimator is to observe
\[
\lim_{t \to \infty} {
  U(c(t) t) - U(t)
  \over
  U(t) - U(t/c(t))
} = 2^{1/\alpha}
\]
for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
Pickands formulated the estimation problem in terms of the extreme
value index $\xi$, which is the reciprocal of the tail index $\alpha$
when the distribution of $X$ is regularly varying (cf. Embrechtsm,
Kl\"uppelberg and Mikosch \cite{Embrechts1997}).

Building upon the aforementioned observation and taking $c(t)$ to
be simply constant 2, Pickands proposed the following estimator for
$\xi$:
\[
\hat \xi = {1 \over \ln 2} \ln {
  x_{(k)} - x_{(2k)}
  \over
  x_{(2k)} - x_{(4k)}
}
\]
Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
that Pickands' estimator was consistent in both the weak and the strong
sense, when the sequence $k(n)$ was chosen appropriately:
\begin{itemize}
\item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
  \[
  \hat \xi \overset{P}{\to} \xi
  \]
\item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
  as $n \to \infty$,
  \[
  \hat \xi \overset{a.s.}{\to} \xi
  \]
\end{itemize}
Dekkers and de Haan \cite{dekkers1989estimation} also showed the
asymptotic normality of the estimator, when some
extra technical conditions were satisfied by
$U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
\[
\sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
\]
where
\[
v(\xi) = {
  \xi^2 (2^{2 \xi + 1} + 1)
  \over
  [2 \ln(2) (2^\xi - 1)]^2
}
\]

\section{Hoga Test}
%% TODO: Show the test statistics of Hoga test applied to a sequence of
%% two concatenated return series. A changes in the scale is characterized
%% by a smooth bump in the test statistics while a change in the tail
%% index is characterized by a jump. Derive a theoretical result to
%% support this?

%% What is the probability of rejecting the null hypothesis when the test
%% is applied to $(X_1, X_2)$ where $X_1$ and $X_2$ have tail indices
%% $\kappa_1 \neq \kappa_2$? The probability of wrongly rejecting the
%% null hypothesis must be lower than $\alpha$.
\textcolor{red}{Review the test of Hoga}

\section{Data Analysis}
\subsection{Tail index and scale parameter}
In this section we apply the statistical tools reviewed in the
previous sections to real stock return series, denoted $X_t$
hereafter. We are interested in finding out the tail indices and the
scale parameters of the stationary distribution of $|X_t|$.

%% and show that, within a
%% comfortable confidence interval, one cannot reject the null hypothesis
%% that these return series have the same tail index.

\subsubsection{The Energy and the Consumer Staples sectors of S\&P 500}
Figure \ref{fig:Energy_Hill_upper} shows the Hill estimates of the
tail indices of 32 stocks in the {\it Energy} sector of S\&P 500
index; figure \ref{fig:Consumer_Staples_Hill_upper} shows the same of
the {\it Consumer Staples} sector.
\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth]{Energy_Hill_upper.pdf}
  \caption{Hill estimates of the upper tail indices of 32 stocks'
    daily return series in the {\it Energy} sector of Standard \& Poor
    500 index. Each circle coresponds to a Hill estimate; the grey
    triangles above and below it mark the 97.5\% and 2.5\% quantiles
    of its approximate normal distribution.
    The low and high red lines mark the median of the 2.5\% quantiles
    and median of the 97.5\% quantiles, respectively.
    The data are from {\it Yahoo Finance} and the labels on
    the horizontal axis are Yahoo symbols of the stocks. The data span
    from 2010-01-01 to 2015-01-01 and comprise $n=1304$
    observations. $k=(\ln n)^2 = 49$ upper order statistics are used
    for computing $\alpha_n$. 
  }
  \label{fig:Energy_Hill_upper}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth]{Consumer_Staples_Hill_upper.pdf}
  \caption{Hill estimates of the upper tail indices of 30 stocks'
    daily return series in the {\it Consumer Staples} sector of
    Standard \& Poor 500 index.
  }
  \label{fig:Consumer_Staples_Hill_upper}
\end{figure}
For our purpose, we have excluded a few stocks, though:
\begin{figure}[htb!]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{WBA_price.pdf}
    \caption{{\it Walgreens Boots Alliance} prices}
    \label{fig:EBA_prices}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{WFM_price.pdf}
    \caption{{\it Whole Foods Market} prices}
    \label{fig:WFM_prices}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{EL_price.pdf}
    \caption{{\it Estee Lauder Cos.} prices}    
    \label{fig:EL_prices}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{CL_price.pdf}
    \caption{{\it Colgate-Palmolive} prices}    
    \label{fig:CL_prices}
  \end{subfigure}
  \caption{Plot of prices of {\it Consumer Staples} stocks
    excluded from Hill estimates}
\end{figure}

\begin{itemize}
\item {\it Walgreens Boots Alliance}:

  As shown in figure (\ref{fig:EBA_prices}),
  this stock was traded only in a small fraction of the bank
  days and it is hence no surprise that the price varied
  tremendously.
\item {\it Whole Foods Market}, {\it Estee Lauder Cos.} and {\it
  Colgate-Palmolive}

  As seen in figures (\ref{fig:WFM_prices}), (\ref{fig:EL_prices}) and
  (\ref{fig:CL_prices}),
  the prices of these stocks were reduced by half during the
  period under consideration. But the price changes are the
  result of emission of new shares rather than random fluctuations.
\end{itemize}

Figure \ref{fig:Energy_Hill_upper} and
\ref{fig:Consumer_Staples_Hill_upper}
are suggestive of the conclusion that excluding price
movements due to deterministic events, tail indices of stock return
series tend to have the same value, within a reasonable confidence
interval.

If we look instead at the scales, differences among the stock return
series are however rather large. Figure (\ref{fig:Energy_Hill_Scales})
and (\ref{fig:Consumer_Staples_Hill_Scales}) show the Hill estimates of
the scale parameters of the stocks in the S\&P 500 {\it Energy} and
{\it Consumer Staples} sectors, respectively.
\begin{figure}[htb!]
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{Energy_Hill_scales.pdf}
    \caption{{\it Energy} scale estimates}    
    \label{fig:Energy_Hill_Scales}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{Consumer_Staples_Hill_scales.pdf}
    \caption{{\it Consumer Staples} scale estimates}    
    \label{fig:Consumer_Staples_Hill_Scales}
  \end{subfigure}
  \caption{Black boxes are Hill estimates of the scale parameters of
    the stock return series, while red circles are the scale estimates
    of simulated iid. Student's $t$ series with a degree of freedom
    equal to the mean value of the estimated tail indices of the stock
    return series. The estimates of the Student's $t$ series are
    included for comparison with those of the return series.}
\end{figure}
Clearly the estimated scale parameters can differ by several orders of
magnitude.

Using the ``Peaks over Threshold'' method and the mean excess
function, we also obtain estimates for the tail index ($\alpha$), the
scale parameter ($A$) and the shift parameter ($a$) for stocks in the
{\it Energy} and the {\it Consumer Staples} sectors of the S\&P 500
index. The results are plotted in figure
\ref{fig:Energy_OLS_estimates} and
\ref{fig:Consumer_Staples_OLS_estimates},
respectively.
\begin{figure}[htb!]
  \includegraphics[width=\textwidth]{Energy_OLS_estimates.pdf}
  \caption{Tail parameter estimates of {\it Energy} stocks. The top
    plot shows the tail index estimates, the middle plot shows the
    10-based logarithm of the scale estimates and bottom plot shows
    the estimates of the shift parameter.
  }
  \label{fig:Energy_OLS_estimates}
\end{figure}

\begin{figure}[htb!]
  \includegraphics[width=\textwidth]{Consumer_Staples_OLS_estimates.pdf}
  \caption{Tail parameter estimates of {\it Consumer Staples}
    stocks. The top 
    plot shows the tail index estimates, the middle plot shows the
    10-based logarithm of the scale estimates and bottom plot shows
    the estimates of the shift parameter.
  }
  \label{fig:Consumer_Staples_OLS_estimates}
\end{figure}

Using Pickands' estimator, the same result of very similar tail indices
is again obtained for the {\it Energy} and the {\it Consumer Staples}
sectors. Figure \ref{fig:Energy_Pickands} and
\ref{fig:Consumer_Staples_Pickands} illustrate this claim.
\begin{figure}[htb!]
  \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
  \caption{Pickands' estimates of {\it Energy} stocks.}
  \label{fig:Energy_Pickands}
\end{figure}

\begin{figure}[htb!]
  \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
  \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
  \label{fig:Consumer_Staples_Pickands}
\end{figure}

\subsubsection{The Information Technology sector of S\&P 500}
Compared with the sectors of energy and consumer staples, which enjoy
a relatively stable demand on their products, companies of information
technology are more subject to volatile demand and competition. As a
result, one can expect their stock prices to be more volatile, implying
lower tail indices and larger differences between the tail indices of
different companies. In this section we investigate how these companies
differ in terms of their tail indices and scale parameters.

Figure \ref{fig:Information_Technology_Hill} shows the Hill
estimates of 42 stocks in the {\it Information Technology} sector. Daily
closing prices of these stocks have been available since 2000-01-01 and
earlier.
\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth]{Information_Technology_Hill.pdf}
  \caption{Hill estimates of the tail indices of 42 stocks' daily
    return series in the {\it Information Technology} sector of Standard
    \& Poor 500 index.
  }
  \label{fig:Information_Technology_Hill}
\end{figure}
It is seen in the figure that most of the estimates fall within the
interval spanned by the confidence bands. Thus, given the estimation
precision, these are consistent with the similarity conclusion of 
\S\ref{sec:GDA}.

%% From the figure it appears a few pairs of stocks do have rather different
%% tail indices, e.g. ``ADI'', ``PAYX'' and ``XLNX'' versus ``CTSH'' and
%% ``NVDA''.

%% To gain more insight into these pairs, we concatenate the
%% squared return series of each pair and apply Hoga's test to the concatenated
%% sequence. Table \ref{tab:Information_Technology_Hoga} shows the test
%% statistic of each concatenated pair.
%% \begin{table}[htb!]
%%   \centering
%%   \begin{tabular}{c|c|c}
%%     \multicolumn{2}{c}{Stock(s)} & Hoga Test Statistic \\
%%     \hline
%%     ADI & CTSH & 15.52 \\
%%     ADI & NVDA & 4.726 \\
%%     PAYX & CTSH & 231.5 \\
%%     PAYX & NVDA & 110.6 \\
%%     XLNX & CTSH & 503.5 \\
%%     XLNX & NVDA & 146.5
%%   \end{tabular}
%%   \caption{Hoga Test statistic of pairs of stocks from the
%%     {\it Information Technology} sector of S\&P 500 index.
%%     The critical value at 95\% confidence level is 166.4}
%%   \label{tab:Information_Technology_Hoga}
%% \end{table}

\subsection{Temporal changes of the parameters}

%% Thomas: Try fitting data to a Burr distribution.
%% TODO: use GPD and POT to estimate the scale and tail index.
%% TODO: plot the Hill estimator instead of its inverse. The asymptotic
%% normal distribution of Hill estimator is found in ``Modeling Extremal
%% events''.

%% TODO: Does the scale estimator also have a convergence result from
%% which confidence bands can be derived?

%% TODO: Why does the Quintos Fan test almost always reject the null?

%% Hill (1975) \cite{hill1975simple} also proposed an estimator for the
%% scale parameter $A$:
%% \[
%% A_n = {k \over n}X_{(k+1)}^{\alpha_n}
%% \]

%% Review of CCC by Starica and others.
%% explain that the model implies that all series have the same tail index
%% Explain that economic arguments also lead to the same tail index

%% Explain that the tail functions of different series are
%% differentiated more by the scale parameter.

\section{Conclusion}
TODO: Summarize and conclude that all correlated series share the same
tail index; their tail functions differ from each other via their
scale parameters.

\bibliographystyle{unsrt}
\bibliography{../../thesis/econophysics}
\end{document}

### 2017-02-01 papers/FX/fx.tex
Revisions following discussion with Casper.

@@
 \usepackage{graphicx}
-\usepackage{subfigure}
+\usepackage{multirow}
+\usepackage{color}
+\usepackage[dvipsnames]{xcolor}
 
@@
 \input{../../physics_common}
-\title{On FX Rates and a Multivariate GARCH model}
+\title{To be Determined}
 \author{Xie Xiaolei}
@@
   (Z_{1, t}, Z_{2,t}, \dots, Z_{p,t}) &\sim& N(0, \Sigma) \\
-  \sigma_{i, t}^2 &=& \alpha_{i,0} + \alpha_{i,1} X_{i, t-1}^2 + \beta_{i,1} \sigma_{i, t-1}^2
+  \sigma_{i, t}^2 &=& \omega_i + \alpha_i X_{i, t-1}^2 + \beta_i
+  \sigma_{i, t-1}^2
 \end{eqnarray*}
@@
 \section{IGARCH(1) Processes}
-We first consider the cases where $\alpha_{i, 1} + \beta_{i, 1} = 1$, i.e. IGARCH models.
+We first consider the cases where $\alpha_i + \beta_i = 1$,
+i.e. IGARCH models.
 
 \subsection{IARCH(1)}
-In this case $\alpha_{i, 1} = 1$, $\beta_{i, 1} = 0$. Thus, from the equation
+In this case $\alpha_i = 1$, $\beta_i = 0$. Thus, from the equation
 \begin{equation*}
@@
 \begin{eqnarray*}
-  \E (Z_i^2 Z_j^2)^\xi &=& 1 \\
+  \E (Z_i^2 Z_j^2)^{\xi/2} &=& 1 \\
   (Z_i, Z_j) &\sim& N(0, \Sigma) \\
@@
     \rho^2 x^4 + (1 - \rho^2) x^2 y^2 + 2 \rho \sqrt{1 - \rho^2} x^3 y
-  \right)^\xi
+  \right)^{\xi/2}
   \exp\left(
@@
 \end{equation}
-When $\rho = \pm 1$, it is clear $\xi = 1/2$. When $-1 < \rho < 1$,
+When $\rho = \pm 1$, it is clear $\xi = 1$. When $-1 < \rho < 1$,
 we may re-write \eqref{eq:iarch1_1st} in polar coordinates to obtain
 \begin{equation*}
-  {1 \over 2\pi} \int_0^{\infty} r^{4\xi + 1} e^{-r^2 / 2} dr
-  \int_{0}^{2\pi} \cos(\theta)^{2 \xi}
+  {1 \over 2\pi} \int_0^{\infty} r^{2\xi + 1} e^{-r^2 / 2} dr
+  \int_{0}^{2\pi} \cos(\theta)^{\xi}
   \left(
     2 \rho^2 \cos(\theta)^2
-    + 2 \rho \sqrt{1 - \rho^2} \cos(\theta) \sin(\theta) - \cos(\theta)^2 + 1 - \rho^2
-  \right)^\xi d\theta = 1
+    + 2 \rho \sqrt{1 - \rho^2} \cos(\theta) \sin(\theta) -
+    \cos(\theta)^2 + 1 - \rho^2
+  \right)^{\xi/2} d\theta = 1
 \end{equation*}
@@
   \int_{0}^{2\pi}
-  \cos(\theta)^{2\xi}
+  \cos(\theta)^{\xi}
   \left(
-    2 \rho^2 \cos(\theta)^2 + \rho \sqrt{1 - \rho^2} \sin(2\theta) - \cos(\theta)^2 - \rho^2 + 1
-  \right)^\xi d\theta
+    2 \rho^2 \cos(\theta)^2 + \rho \sqrt{1 - \rho^2} \sin(2\theta) -
+    \cos(\theta)^2 - \rho^2 + 1
+  \right)^{\xi/2} d\theta
   = {2 \pi
     \over
-    2^{2 \xi} \Gamma(2\xi + 1)
+    2^{\xi} \Gamma(\xi + 1)
   }
@@
   \label{eq:iarch_2nd}
-  \int_{-1}^1 (\sin(\pi z) + \rho)^{2\xi} dz = {2 \over \Gamma(2\xi + 1)}
+  \int_{-1}^1 (\sin(\pi z) + \rho)^{\xi} dz = {2 \over \Gamma(\xi + 1)}
 \end{equation}
 Using this result, we immediately obtain for ARCH(1) models
-\begin{equation*}
+\begin{equation}
   \label{eq:arch_1st}
-  \int_{-1}^1 (\sin(\pi z) + \rho)^{2\xi} dz =
-  {2 \over \Gamma(2\xi + 1)}
-  {1 \over (\alpha_{i, 1} \alpha_{j, 1})^\xi}
-\end{equation*}
+  \int_{-1}^1 (\sin(\pi z) + \rho)^{\xi} dz =
+  {2 \over \Gamma(\xi + 1)}
+  {1 \over (\alpha_i \alpha_j)^{\xi/2}}
+\end{equation}
 Fitting ARCH(1) models to the FX return series and then simulating the
@@
 
+To compare the solution to \eqref{eq:arch_1st}, call it $\xi_{i,j}$
+with the tail indices of $\text{ARCH}(1)$ processes with parameters
+$\alpha_i$ and $\alpha_j$, we fix $\alpha_i \alpha_j$ and let
+$\alpha_i = \alpha_j = \sqrt{\alpha_i \alpha_j}$. Since $\xi_i$ as a
+function of $\alpha_i$ is monotonically decreasing, $\min\{\xi_i,
+\xi_j\}$ is at its maximum when $\alpha_i = \alpha_j$ for fixed
+$\alpha_i \alpha_j$. This is the reason for the above setup. Table
+\ref{tab:tail_indices_normal} lists the values of $\xi_i$ and
+$\xi_{i,j}$. Here $\text{corr}(Z_i, Z_j)$ varies from 0.09 to 0.99.
+\begin{table}[htb!]
+  \centering
+  \begin{scriptsize}
+  \begin{tabular}{c|c|cccccccccc}
+    \multicolumn{2}{c|}{$\alpha_i \alpha_j$} & 0.090 & 0.190 & 0.290 &
+    0.390 & 0.490 & 0.590 & 0.690 & 0.790 & 0.890 & 0.990 \\
+    \hline
+    \multicolumn{2}{c|}{$\alpha_i, \alpha_j$} & 0.300 & 0.436 & 0.539 &
+    0.624 & 0.700 & 0.768 & 0.831 & 0.889 & 0.943 & 0.995 \\
+    \hline
+    \multicolumn{2}{c|}{$\xi_i, \xi_j$} & 4.180 & 2.766 & 2.170 &
+    1.822 & 1.586 & 1.413 & 1.279 & 1.171 & 1.082 & 1.007 \\
+    \hline
+    \multirow{10}{*}{$\text{corr}(Z_i Z_j)$} & 0.1 & 8.041 & 5.375 &
+    4.238 & 3.568 &
+    3.113 & 2.778 & 2.518 & 2.308 & 2.134 & 1.987 \\
+    & 0.2 & 7.439 & 5.031 & 3.997 & 3.384 & 2.964 &
+    2.653 & 2.411 & 2.215 & 2.052 & 1.914 \\
+    & 0.3 & 6.846 & 4.652 & 3.713 & 3.156 &
+    2.774 & 2.491 & 2.269 & 2.089 & 1.939 & 1.812 \\
+    & 0.4 & 6.318 & 4.294 & 3.434 & 2.924 &
+    2.575 & 2.317 & 2.114 & 1.950 & 1.813 & 1.696 \\
+    & 0.5 & 5.852 & 3.970 & 3.174 & 2.704 &
+    2.383 & 2.146 & 1.960 & 1.810 & 1.684 & 1.577 \\
+    & 0.6 & 5.441 & 3.679 & 2.936 & 2.500 &
+    2.203 & 1.983 & 1.812 & 1.673 & 1.558 & 1.460 \\
+    & 0.7 & 5.074 & 3.417 & 2.720 & 2.312 &
+    2.035 & 1.830 & 1.671 & 1.543 & 1.436 & 1.345 \\
+    & 0.8 & 4.746 & 3.180 & 2.522 & 2.138 &
+    1.877 & 1.686 & 1.537 & 1.418 & 1.318 & 1.234 \\
+    & 0.9 & 4.449 & 2.964 & 2.340 & 1.975 &
+    1.729 & 1.548 & 1.408 & 1.296 & 1.203 & 1.124 \\
+    & 1.0 & 4.180 & 2.766 & 2.170 & 1.822 &
+    1.586 & 1.413 & 1.279 & 1.171 & 1.082 & 1.007
+  \end{tabular}
+  \end{scriptsize}
+  \caption{$\xi_i, \xi_j$ and $\xi_{i,j}$}
+  \label{tab:tail_indices_normal}
+\end{table}
+
 \section{General IGARCH(1)}
@@
     (
-    \alpha_{i,1} Z_1^2 + 1 - \alpha_{i,1}
-    )^\xi
+    \alpha_i Z_1^2 + 1 - \alpha_i
+    )^{\xi/2}
     (
-    \alpha_{j,1} Z_2^2 + 1 - \alpha_{j,1}
+    \alpha_j Z_2^2 + 1 - \alpha_j
     )^\xi\right] = 1
@@
 has to be solved numerically. Figure \ref{fig:xi_rho0.5} shows $\xi$
-as a function of $(\alpha_{i,1}, \alpha_{j,1})$ as defined by the
+as a function of $(\alpha_i, \alpha_j)$ as defined by the
 above equation.
@@
   \label{fig:ARCH_t_FX_eigenvectors}
+\end{figure}
+
+On the other hand, if each series $X_{i,t}$ is assumed to have
+conditional distribution $t(\nu_i)$ with $\nu_i, \nu_j$ potentially
+different from each other, the following parameters as listed in table
+\ref{tab:ARCH_t_params} are obtained:
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{c|cccc}
+    currency & $\omega$ & $\alpha$ & $\beta$ & $\nu$ \\
+    \hline
+    AUD & 1.1e-06 & 0.026 & 0.949 & 9.8 \\
+    CAD & 2.1e-07 & 0.015 & 0.980 & 8.3 \\
+    CNY & 2.6e-07 & 0.038 & 0.957 & 7.0 \\
+    CZK & 2.4e-07 & 0.035 & 0.955 & 8.3 \\
+    DKK & 6.5e-07 & 0.060 & 0.905 & 9.5 \\
+    EUR & 6.3e-07 & 0.061 & 0.905 & 9.6 \\
+    GBP & 7.9e-07 & 0.049 & 0.931 & 10.0 \\
+    HKD & 2.6e-07 & 0.041 & 0.955 & 8.8 \\
+    HUF & 5.8e-07 & 0.045 & 0.942 & 7.5 \\
+    JPY & 5.4e-07 & 0.052 & 0.942 & 8.5 \\
+    KRW & 1.4e-06 & 0.053 & 0.916 & 6.8 \\
+    MAD & 5.8e-07 & 0.045 & 0.938 & 3.7 \\
+    MXN & 3.8e-06 & 0.083 & 0.875 & 3.9 \\
+    NOK & 1.6e-07 & 0.030 & 0.962 & 6.7 \\
+    NZD & 4.3e-07 & 0.020 & 0.972 & 10.0 \\
+    SGD & 4.2e-07 & 0.040 & 0.947 & 10.0 \\
+    USD & 2.5e-07 & 0.040 & 0.956 & 8.5
+  \end{tabular}
+  \label{tab:ARCH_t_params}
+  \caption{parameters for univariate GARCH(1,1)-t models}
+\end{table}
+
+\section{Spectra of Squared returns}
+Figure \ref{fig:FX_squares_eigenvalues} and
+\ref{fig:FX_squares_eigenvectors} show the eigenvalues and
+eigenvectors of the sample covariance matrix of squared return series
+of the selected foreign exchange rates. Table
+\ref{tab:FX_squared_Hill} shows the Hill estimators of $X_{i,t}^2$ and
+$X_{i,t} X_{i-1,t}$.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4]{FX_squares_eigenvalues.pdf}  
+  \caption{Eigenvalues of sample covariance matrix of returns and
+    squared returns of FX}
+  \label{fig:FX_squares_eigenvalues}
+\end{figure}
+
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{c|c|c}
+    currency & Hill est. & Hill est. of prod.\\
+    \hline
+    AUD & 2.129957 \\
+    CAD & 2.388715 & 2.417197 \\
+    CNY & 2.483576 & 2.990937 \\
+    CZK & 2.341256 & 2.677970 \\
+    DKK & 2.303255 & 2.873808 \\
+    EUR & 2.496141 & 2.528287 \\
+    GBP & 2.355945 & \textcolor{BrickRed}{1.979052} \\
+    HKD & 1.814324 & 2.222141 \\
+    HUF & 1.829534 & 2.308966 \\
+    JPY & 1.923963 & 2.109159 \\
+    KRW & 1.983916 & 2.296491 \\
+    MAD & 5.000601 & 2.438606 \\
+    MXN & 5.288364 & \textcolor{BrickRed}{2.063652} \\
+    NOK & 1.814017 & 2.381056 \\
+    NZD & 2.224353 & 2.270946 \\
+    SGD & 2.003079 & 2.009848 \\
+    USD & 2.268252 & 2.452259 \\
+  \end{tabular}
+  \caption{Hill Estimator of $X_{i,t}^2$ and $X_{i,t} X_{j,t}$}
+  \label{tab:FX_squared_Hill}
+\end{table}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.3]{FX_squares_eigenvectors.pdf}  
+  \caption{Eigenvectors of sample covariance matrix of returns and
+    squared returns of FX}
+  \label{fig:FX_squares_eigenvectors}
 \end{figure}
@@
 
+\section{Factor Models}
+\begin{eqnarray*}
+  X_{i,t} &=& \sum_{k=1}^N C_{i,k} Y_{k,t} \\
+  Y_{k,t} &=& Z_{k,t} \sigma_{k,t} \\
+  \sigma_{k,t}^2 &=& \omega_k + \alpha_k Y_{k, t-1}^2 +
+  \beta_{k, t-1} \sigma_{k, t-1}^2
+\end{eqnarray*}
+
+Figures \ref{fig:FX_eigenvalues_factor} and
+\ref{fig:FX_eigenvectors_factor} show how the spectrum produced by
+simulation according to the factor model compares to the spectrum
+of real FX data.
+
+\subsection{Orthogonal GARCH}
+Kariya (1988) and Alexander and Chibumba (1997) proposed the
+orthogonal GARCH model, which treats each sequence as a linear
+combination of a number of common, indepdent factors, each of which
+is modeled by a variant of GARCH, e.g. standard GARCH,
+EGARCH, GJR-GARCH, etc. Figure \ref{fig:FX_eigenvalues_factor} and
+\ref{fig:FX_eigenvectors_factor} show the spectrum of real FX data
+together with a spectrum of data simulated from the orthogonal GARCH
+model. Here the innovations of the GARCH model are sampled from the
+$N(0, 1)$ distribution.
+
+GARCH parameters of the factors are estimated from the data, i.e. the
+sample covariance matrix of the observed series $X_{i,t}$ are first
+computed. The eigenvalues $\lambda_i$ and the eigenvectors $\vec{C_j}$
+of the sample covariane matrix are then computed. Let $C_{i,j}$
+denote the $i$-th element of the $j$-th eigenvector. Then the $j$-th
+factor $Y_{j,t}$ is constructed as
+\[
+Y_{j,t} = \sum_{i=1} C_{i, j} X_{i, t}
+\]
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4]{FX_eigenvalues_factor.pdf}  
+  \caption{FX \& simulated GARCH eigenvalues. normal innovations}
+  \label{fig:FX_eigenvalues_factor}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.3]{FX_eigenvectors_factor.pdf}  
+  \caption{FX \& simulated GARCH eigenvectors. normal innovations}
+  \label{fig:FX_eigenvectors_factor}
+\end{figure}
+
+If, however, the innovations are sampled from the Students' t
+distribution with degree of freedom estimated from the data, the
+simulated spectrum demonstrates a dominance pattern, as shown in
+figure \ref{fig:GARCH-t_eigenvalues} and
+\ref{fig:GARCH-t_eigenvectors}
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.4]{GARCH-t_eigenvalues.pdf}  
+  \caption{FX \& simulated GARCH eigenvalues. t-innovations}
+  \label{fig:GARCH-t_eigenvalues}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[scale=0.3]{GARCH-t_eigenvectors.pdf}  
+  \caption{FX \& simulated GARCH eigenvectors. t-innovations}
+  \label{fig:GARCH-t_eigenvectors}
+\end{figure}
+
+\subsection{Non-linear Factor Model}
+A survey of Hill estimated tail indices of 16 currencies are listed in
+table \ref{tab:currencies_hill_estimate}. It is visible that Asian
+currencies are associated with lower tail indices than are their
+counterparts in Europe, North America and Oceania. This observation is
+in contradition with classical factor models, by which all the
+currencies should have the same tail index that is equal to the lowest
+tail index of the factors, since the return series of each currency is
+a linear combination of the return series of the factors. The
+conclusion follows immediately from Mikosch 2013 \cite{Mikosch2013},
+lemma 3.1.
+\begin{table}[htb!]
+  \begin{tabular}{lr|lr|lr|lr}
+    \multicolumn{2}{c|}{Asia} & \multicolumn{2}{|c}{Europe} &
+    \multicolumn{2}{|c}{N. America} & \multicolumn{2}{|c}{Oceania}\\
+    \hline
+    HKD & 1.88 & EUR & 2.09 & CAD & 2.14 & AUD & 2.13 \\
+    JPY & 1.92 & NOK & 2.02 & USD & 2.23 & NZD & 2.17 \\
+    KRW & 1.98 & CZK & 2.17 & & & &\\
+    SGD & 1.98 & DKK & 2.13 & & & & \\
+    & &          GBP & 2.21 & & & & \\
+    & &          HUF & 2.11 & & & &
+  \end{tabular}
+  \caption{Hill estimates of currency tail indices by continents}
+  \label{tab:currencies_hill_estimate}
+\end{table}
+
+Therefore we propose
+\begin{eqnarray*}
+  X_{i,t} &=& \sum_{k=1}^N C_{i,k} \text{sgn}(Y_{k,t}) |Y_{k,t}|^{\delta_{i,k}}
+\end{eqnarray*}
+where $0 < \delta_{i,k} \leq 1$ for all $i, k$. In particular, if
+currencies $i_1, i_2, ..., i_m$ belong to the same group, e.g. the
+group of Asian currencies, $\delta_{i_1, k} = \delta_{i_2, k} = \cdots
+\delta_{i_m, k}$.
+
+\section{different tail indices or same tail index but different scales}
+The Hill estimator is rather inaccurate in estimating the tail index
+of a heavy-tailed series, particularly when the series is scaled
+differently. Figure \ref{fig:simulated.indices} shows the Hill
+estimates of two groups of simulated series of the form
+$X_{i,t} = C_i Y_{i, t} + G_t$, where the series 
+$Y_{i}$ and the common factor $G$ are independent and identical
+GARCH(1,1) processes. The first, shown in black squares, has $C_i$
+ranging from 0.1 to 1.5; the second, shown in red circles, has
+$C_i = 1$ for all $i$.
+\begin{figure}[htb!]
+  \includegraphics[scale=0.5]{simulated_indices.pdf}
+  \caption{Tail indices of linearly combined GARCH(1, 1) processes}
+  \label{fig:simulated.indices}
+\end{figure}
+
+%% Show simulation result of the tail index and scale using the script
+%% tail_indices_compare.r
+%% Use Hill estimators for both the tail index and the scale
+%% Use the optimal number of order statistics as given in Casper's
+%% paper.
+%% Maybe also show simulation result of the tail index and scale using
+%% iid Student's t series.
+
+%% CAUTION: Chen Zhou believes it is hopeless to answer definitely
+%% whether two series have the same tail index.
+
 \bibliographystyle{unsrt}
-\bibliography{../thesis/econophysics}
+\bibliography{../../thesis/econophysics}
 \end{document}

### 2017-02-01 papers/FX/draft.tex
Some more work on the power utility function.

@@
 that the utility function of consumption $C$ is a power function%
-\[
+\begin{equation}
+  \label{eq:power_utility}
   u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
-\]
+\end{equation}
 where $\gamma > 0$. For simplicity we assume the investor
@@
 
-\subsubsection{mildly risk-avert utility function}
+\subsubsection{mildly risk-avert utility functions}
 We first consider the class of utility functions for which $\gamma < 1$.
@@
   + \sum_{k=0}^\infty
-  {\alpha p \over M^{k+1} ((1 - \phi) e^r)^{\gamma - 1}}
+  {\alpha p ((1 - \phi) e^r)^{1 - \gamma}\over M^{k+1}}
   \prod_{i=0}^{k-1} (\alpha + i)
@@
 \end{eqnarray}
-Observe that the sum over $n$ does not involve $\alpha$. In the
-following we show this sum is positive for every $k \geq 0$, then it
-is clear from \eqref{eq:xxie1} that $I_3$ and hence
-$u(v) + b u(\delta v)$ is monotonically increasing with $\alpha$. To
-see this, denote
+Observe that the sum over $n$ does not involve $\alpha$. Denote
 \[
@@
 \]
-For each given $k$, $a_n(k)$ is an alternating series. To show the
-positivity of $\sum_n a_n(k)$, it suffices to show
-\[
-  {a_n(k) \over |a_{n+1}(k)|} > 1
-  \quad
-  \text{ for } n=1,3,5,...
-\]
-Direct computation gives
+For each given $k$, $a_n(k)$ is an alternating series.
 \begin{eqnarray*}
@@
 
-Thus we see that
-$u(v) + u(\delta v)$ is monotonically increasing with $\alpha$. If
-all investors are guided by and hence seek to maximize the same
-utility function, they would all have constructed their portfolios by
-choosing from a set of equities that have tail indices close to the
-largest in the market. At market equilibrium, all actively traded
-equities would have tail indices that are indistinguishable from each
-other.
-
-Notice that the summand of \eqref{eq:xxie1} decreases
+If ${e^{M + r} (1 - \phi) \over \phi} > 1$, it is clear
+$a_n(k)/|a_{n+1}(k)| > 1$. It follows that $I_3$ and hence $u(v) +
+u(\delta v)$ is monotonically increasing with $\alpha$. If all
+investors are guided by and hence seek to maximize the same utility
+function, they would all 
+have constructed their portfolios by choosing from a set of equities
+that have tail indices close to the largest in the market. At market
+equilibrium, all actively traded equities would have tail indices that
+are indistinguishable from each other.
+
+\textcolor{red}{
+  What happens when ${e^{M + r} (1 - \phi) \over \phi} \leq 1$?
+}
+
+\subsubsection{very risk-avert utility functions}
+In this section we consider power utility functions of
+\eqref{eq:power_utility} with $\gamma > 1$. For convenience, let $ĸ =
+\gamma - 1$.
 

### 2017-02-02 papers/FX/draft.tex
Some advancement.
1. Power utility function with gamma < 1 2. some progress on finding the right eigenfunction in the paper with Jeff.

@@
 \title{Tail Indices and Scale Parameters in Financial Time Series}
-\author{Casper de Vries, Thomas Valentin Mikosch, Xie Xiaolei}
+\author{Thomas Valentin Mikosch, Casper de Vries, Xie Xiaolei}
 \date{\today}
@@
 \begin{eqnarray}
-  {1 \over 1 + b} I_3
+  {1 \over 1 + b} I_1
   &=&
@@
   {\alpha p ((1 - \phi) e^r)^{1 - \gamma}\over M^{k+1}}
-  \prod_{i=0}^{k-1} (\alpha + i)
+  \left[ \prod_{i=0}^{k-1} (\alpha + i) \right]
   \sum_{n=1}^\infty {(-1)^{n-1} \over n! n^{k+1}}
@@
 For each given $k$, $a_n(k)$ is an alternating series.
+For $n = 1,3,5,\cdots$
 \begin{eqnarray*}
@@
 If ${e^{M + r} (1 - \phi) \over \phi} > 1$, it is clear
-$a_n(k)/|a_{n+1}(k)| > 1$. It follows that $I_3$ and hence $u(v) +
+$a_n(k)/|a_{n+1}(k)| > 1$. It follows that $I_1$ and hence $u(v) +
 u(\delta v)$ is monotonically increasing with $\alpha$. If all
@@
 equilibrium, all actively traded equities would have tail indices that
-are indistinguishable from each other.
+are indistinguishable from each other by standard estimation procedures.
+
+Furthermore, differentiating \eqref{eq:xxie1} gives
+\begin{eqnarray*}
+  && \opd{\phi} [u(v) + b u(\delta v)] \\
+  &=&
+  \pd{I_1}{\phi} + \opd{\phi}(I_2 + I_3) \\
+  &=&
+  -(1 + b) {
+    p ((1 - \phi) e^r)^{1- \kappa}
+    \over
+    1 - \phi
+  }
+  - (1 + b) \sum_{k=0}^\infty \sum_{n=1}^\infty
+  {\alpha p \over M^{k + 1}}
+  {\Gamma(\alpha + k) \over \Gamma(\alpha)}
+  {(-1)^{n-1} \over n! n^{k+1}} \times \\
+  &&
+  {\Gamma(\gamma + n - 1) \over \Gamma(\gamma)}
+  {1 \over \phi}
+  \left[
+    {\phi \over e^{M+r} (1 - \phi)}
+  \right]^n
+  e^r [(1 - \gamma) \phi + n] [(1 - \phi) e^r]^{-\gamma} \\
+  && + \partial_\phi I_2 + \partial_\phi I_3
+\end{eqnarray*}
+where we have used the short-hand notation $\partial_\phi I_i =
+\pd{I_i}{\phi}$. Truncating the infinite sum above after $k=0$ and
+$n=1$, and then setting the result to 0 provides a condition that the
+optimal choice of $\phi$ must satisfy:
+\begin{eqnarray*}
+  {
+    p [(1 - \phi) e^r]^{-\gamma}
+    [e^{-M} \alpha (1 - \phi (1 - \gamma)) - e^r M (1 - \phi)]
+    (1 + b)
+    \over
+    M (1 - \phi)
+  } + \partial_\phi I_2 + \partial_\phi I_3 = 0
+\end{eqnarray*}
+Solving for $\alpha$, we get
+\begin{eqnarray*}
+  \alpha &=& {
+    M e^M (1 - \bar \phi)
+    \left[
+      e^r p (1 + b) - ((1 - \bar \phi) e^r)^\gamma
+      (\partial_\phi I_2 + \partial_\phi I_3)_{\phi = \bar \phi}
+    \right]
+    \over
+    p [1 - (1 - \gamma) \bar \phi] (1 + b)
+  }
+\end{eqnarray*}
+where $\bar \phi$ denotes the optimal choice of $\phi$, the portion of
+one's wealth allocated to the equity.
 
@@
 In this section we consider power utility functions of
-\eqref{eq:power_utility} with $\gamma > 1$. For convenience, let $ĸ =
+\eqref{eq:power_utility} with $\gamma > 1$. For convenience, let $\kappa =
 \gamma - 1$.

### 2017-02-04 papers/FX/implied_tail_index.cpp
1. explored the idea of computing the right eigenfunction using    the infinite series representation from Nummelin et al 1987. 2. R scripts developed in Rotterdam.

@@ -1,0 +1,382 @@
+#include <stdio.h>
+#include <assert.h>
+#include <cmath>
+#include <algorithm>
+#include <iostream>
+#include <string>
+#include <armadillo>
+#include <gsl/gsl_math.h>
+#include <gsl/gsl_roots.h>
+#include <gsl/gsl_errno.h>
+#include <gsl/gsl_integration.h>
+#include <gsl/gsl_randist.h>
+#include <gsl/gsl_multimin.h>
+#include <gsl/gsl_linalg.h>
+#include <gsl/gsl_blas.h>
+#include <gsl/gsl_cblas.h>
+#include <gsl/gsl_eigen.h>
+#include <nlopt.hpp>
+
+#define ARRAY_SIZE(X) sizeof(X)/sizeof((X)[0])
+
+using namespace std;
+using namespace arma;
+
+double t_density_function(double x, void *params)
+{
+    double nu = *(double *)params;
+    return gsl_ran_tdist_pdf(x, nu);
+}
+
+double norm_density_function(double x, void *params)
+{
+    return gsl_ran_gaussian_pdf(x, 1);
+}
+
+/**
+   params:
+   0th: tail index
+   1st: prob. density func.
+   2nd: a vector of ARCH and GARCH parameters
+ */
+double uni_moment_integrand_function (double t, void *params)
+{
+    vector<void *> *par = (vector<void *> *)params;
+    double exponent = *(double *)(par->at(0));
+    gsl_function *pdf = (gsl_function *)(par->at(1));
+    vector<double> *mdl_par = (vector<double> *)(par->at(2));
+    double y;
+    if (mdl_par->size() == 1) {
+	y = pow(t * t, exponent) *
+	    pdf->function(t, pdf->params);
+    } else {
+	y = pow(t * t + mdl_par->at(1) / mdl_par->at(0), exponent) *
+	    pdf->function(t, pdf->params);
+    }
+    return y;
+}
+
+/**
+   params:
+   0th: a vector of ARCH and GARCH parameters
+   1st: prob. density func.
+ */
+double uni_moment_function(double index, void *params)
+{
+    vector<void *> *par = (vector<void *> *)params;
+    // double a = ((vector<double> *)par->at(0))->at(0);
+    gsl_integration_workspace *w 
+	= gsl_integration_workspace_alloc (1000);
+  
+    double result, error;
+    gsl_function G;
[…]

### 2017-02-21 papers/FX/draft.tex
Started the section of power-utility functions with $\gamma > 1$.
progress on the monotonicity proof.

--- 2017-02-01 papers/FX/draft.tex
@@
 preferences follow Gul's (1991) original disappointment aversion.
+An agent guided by this utility function will seek to maximize $u(v)$.
 
 \subsection{power-law utility function}
-In this section we follow Routledge and Zin \cite{routledge2010generalized} and assume
+In this section we follow Routledge and Zin
+\cite{routledge2010generalized} and assume 
 that the utility function of consumption $C$ is a power function%
@@
 are $\left(1-\phi\right)$ and $\phi$ respectively. Then
-\begin{equation*}
+\begin{equation}
+  \label{eq:xxie1}
   %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
@@
   C = (1 - \phi) e^r + \phi e^X
-\end{equation*}
-
-\subsubsection{mildly risk-avert utility functions}
-We first consider the class of utility functions for which $\gamma < 1$.
-In this case $\E u(C) < \infty$ only if the distribution of $X$, call
-it $F(\cdot)$, has a light upper tail. We assume this for the rest of
-this section.
-
-We note that utility functions with $\gamma < 1$ represent investors
-who are less risk-avert than are those represented by utility
-functions with $\gamma > 1$.
-
-Formally, we assume
+\end{equation}
+
+%% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
+
+\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
+In this section we consider power utility functions of
+\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
+let $\xi = \gamma - 1$. In this case, we assume the following
+distribution function $F(\cdot)$ of the equity return:
 \begin{equation*}
   F(x) = \left\{
-  \begin{array}{lr}
-    p M^\alpha/|x|^\alpha & \text{ if } x < -M \\
-    F_u(x) & \text{ if } x \geq -M \\
+  \begin{array}{ll}
+    p \left(
+    {K \over K - x}
+    \right)^\alpha & x < 0 \\
+    1 - (1 - p) \left(
+    {K' \over K' + x}
+    \right)^\beta & x > 0
   \end{array}
@@
 \end{equation*}
-where $0 < p < 1$ is a constant; $\alpha > 1$, $M > 1$ and $F_u(\cdot)$
-is some distribution function satisfying
-\begin{itemize}
-\item $F_u(-M) = p$
-\item
-  \[
-  \int_{-M}^\infty [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) < \infty
-  \]
-\end{itemize}
-%% In addition we assume
-%% \begin{equation}
-%%   \label{eq:xxie0.1}
-%% \phi < {4 e^{M + r} \over 4 e^{M + r} + 1}
-%% \end{equation}
-
-Rewriting \eqref{eq:xxie0} gives
+Assume $\alpha, \beta > 1$, $0 < p < 1$, $K > 0$. Equation
+\eqref{eq:xxie0} can be re-arranged as
+\[
+u(v) + b u(\delta v))
+=
+\E u(C) + (1 + b) \E [u(\delta v) - u(C); C < \delta v]
+\]
+The right-hand-side of the above equation is a functional of the
+distribution function of $X$ via eq.\eqref{eq:xxie1}. For convenience
+we denote it $\mathcal U(\cdot)$, i.e.
+\[
+\mathcal U F (\alpha, \beta, K, K', \phi)
+= 
+\E u(C) + (1 + b) \E [u(\delta v) - u(C); C < \delta v]
+\]
+\begin{theorem}
+  Let
+  \[
+  G(\alpha, \beta, K, K') =
+  \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
+  \]
+  $G(\alpha, \beta, K, K')$ is monotonically increasing with
+  $\alpha$.
+\end{theorem}
+
+\begin{proof}
+  We show
+  \[
+  \td{G}{\alpha} > 0
+  \]
+  Let
+  \begin{equation}
+    \label{eq:xxie2}
+    \hat \phi := \argmax_{0 < \phi \leq 1}
+    \mathcal U F(\alpha, \beta, K, K', \phi)
+  \end{equation}
+  We have
+  \[
+  G(\alpha, \beta, K, K') = \mathcal U F(\alpha, \beta, K, K', \hat \phi)
+  \]
+  It follows
+  \begin{eqnarray}
+    \td{G}{\alpha}
+    &=&
+    \pd{\mathcal U F}{\alpha}(\alpha, \beta, K, K', \hat \phi)
+    + \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi)
+    \pd{\hat \phi}{\alpha}(\alpha, \beta, K, K')
+    \label{eq:xxie3}
+  \end{eqnarray}
+  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
+  \begin{equation}
+    \label{eq:xxie4}
+    \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
+  \end{equation}
+  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
+  \[
+  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
+  \]
+  \begin{enumerate}[i.]
+  \item When $C < \delta v$ implies $X < -q$ for some $q > 0$, we have
     \begin{eqnarray*}
-  u(v) + b u(\delta v) &=& \E u(C) + b \E [u(C); C < \delta v] \\
-  &=& \alpha p M^\alpha{1 + b \over 1 - \gamma}\int_{-\infty}^{-M}
-  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
-  && + {1 + b \over 1 - \gamma}\int_{-M}^{\delta v}
-  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
-  && + \int_{\delta v}^{\infty}
-  [(1 - \phi) e^r + \phi e^x]^{1 - \gamma} d F_u(x) \\
-  &=& I_1 + I_2 + I_3
+      \pd{\mathcal U(F)}{\alpha}
+      &=&
+      - \frac{\partial}{\partial \alpha} \left\{
+      {\alpha p K^\alpha \over \xi}
+      \left[
+      (1 + b)\int_{-\infty}^{-q} +
+      \int_{-q}^0
+      \right] \right. \\
+      && \left.
+      \left\{
+      \left[
+        (1 - \phi) e^r + \phi e^x
+      \right]^{-\xi}
+      (K - x)^{-\alpha - 1}
+      \right\}
+      dx \right\} \\
+      &=&
+      \underbrace{
+      {p \over \xi}
+      \left[
+      (1 + b)\int_{q}^{\infty} +
+      \int_{0}^q
+      \right]
+      \left[
+        (1 - \phi) e^r + \phi e^{-x}
+      \right]^{-\xi}
+      \ln \left(
+      1 + {x \over K}
+      \right)
+            {\alpha K^{\alpha} dx
+              \over
+              (K + x)^{\alpha + 1}
+            }
+      }_{L_1}\\
+      && - \underbrace{
+        {p \over \xi}
+        \left[
+          (1 + b)\int_{q}^{\infty} +
+          \int_{0}^q
+          \right]
+        \left[
+          (1 - \phi) e^r + \phi e^{-x}
+          \right]^{-\xi}
+        {K^{\alpha} dx \over (K + x)^{\alpha + 1} }
+      }_{L_2} \\
     \end{eqnarray*}
-Only the first term $I_1$ involves the tail parameters $\alpha$
-and $M$ in which we are interested, so we will only compute this
-term. The integral of $I_1$ can be written as
+    Clearly $L_2 < \infty$. For $L_1$, observe
     \begin{eqnarray*}
-&& \int_{-\infty}^{-M}
-[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
-&=& \int_{M}^{\infty}
-[(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} x^{-\alpha-1} dx
+      L_1 &<& {
+        \alpha K^\alpha p (1 + b)
+        \over
+        \xi (1 - \phi)^\xi e^{r \xi}
+      } \int_0^\infty
+      {1 \over (x + K)^{\alpha + 1}}
+      \left(1 + {x \over K}\right)
+      dx \\
+      &=&
+      {
+        p \alpha (1 + b)
+        \over
+        \xi (\alpha - 1) (1 - \phi)^\xi e^{r \xi}
+      } < \infty
     \end{eqnarray*}
-Taylor expansion at $\phi e^{-x} = 0$ gives
+    Let
     \begin{eqnarray*}
-  && [(1 - \phi) e^r + \phi e^{-x}]^{1 - \gamma} \\
-  &=& ((1 - \phi) e^r)^{1 - \gamma}
-  + \sum_{n=1}^{\infty} {1 \over n!}
-  {\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
-  ((1 - \phi) e^r)^{1 - n - \gamma} \phi^n e^{-n x}
+      c &:=& {p \over \xi}
+      \left[
+      (1 + b)\int_{q}^{\infty} +
+      \int_{0}^q
+      \right]
+      \left[
+        (1 - \phi) e^r + \phi e^{-x}
+      \right]^{-\xi}
+      {
+        \alpha K^{\alpha} dx \over (K + x)^{\alpha + 2}
+      } < \infty
     \end{eqnarray*}
+    Define distribution function $H(x): (0, \infty) \to (0, 1)$ by its
+    density with respect to the Lebesgue measure
+    \[
+      {d H(x) \over d x}
+      =
+      {p \over c \xi}
+      \left[
+        (1 - \phi) e^r + \phi e^{-x}
+        \right]^{-\xi}
+      {
+        \alpha K^{\alpha}
+        \over
+        (K + x)^{\alpha + 2}
+      }
+      \times
+      \left\{
+      \begin{array}{ll}
+        1 & \text{ if } 0 < x < q \\
+        (1 + b) & \text{ if } x > q \\
+      \end{array}
+      \right.
+    \]
+    Let $Y \sim H$, then
+    \begin{eqnarray*}
+      \E Y &:=& {p \over c \xi}
+      \left[
+        (1 + b)\int_{q}^{\infty} +
+        \int_{0}^q
+        \right]
+      \left[
+        (1 - \phi) e^r + \phi e^{-x}
+        \right]^{-\xi}
+      \alpha K^{\alpha} (K + x)^{-\alpha - 1} dx \\
+      &=& {\alpha L_2 \over c} \\
+      L_1 &=& c \E[Y \ln(Y/K)]
+    \end{eqnarray*}
+    Notice
+    \[
+      {
+        d^2
+        \over
+        dx^2
+      }
+      \left[
+        x \ln\left({x \over K}\right)
+      \right] = {1 \over x} > 0
+      \quad \text{ for } x > 0
+    \]
+    That is, $x \ln(x/K)$ defined on $(0, \infty)$ is a convex
+    function. Now that $L_1, L_2 < \infty$, Jensen's inequality gives
+    \[
+      {L_1 \over c} = \E [Y \ln(Y/K)]
+      >
+      \E Y \times \ln(\E Y/K)
+      =
+      {\alpha L_2  \over c}
+      \ln\left(
+      {\alpha L_2  \over c K}
+      \right)
+    \]
+  \end{enumerate}
   Therefore
+  \[
+  L_1 - L_2 > L_2 \left[
+    \alpha \ln \left(
+    {\alpha L_2 \over c K}
+    \right) - 1
+  \right]
+  \]
+  Clearly $L_2 > 0$. So it remains to show
+  \[
+    \alpha \ln \left(
+    {\alpha L_2 \over c K}
+    \right) - 1 > 0
+  \]
+  Observe
   \begin{eqnarray*}
-&& \int_{-\infty}^{-M}
-[(1 - \phi) e^r + \phi e^x]^{1 - \gamma} |x|^{-\alpha-1} dx \\
-&=&
-{1 \over \alpha} M^{-\alpha} ((1 - \phi) e^r)^{1 - \gamma}
-+ \sum_{n=1}^{\infty} {1 \over n!}
-{\Gamma(2 - \gamma) \over \Gamma(2 - n -\gamma)}
-((1 - \phi) e^r)^{1 - n - \gamma} n^\alpha \Gamma(-\alpha, n M)
+    L_2 &>&
+      {p \over \xi}
+      \left[
+      (1 + b)\int_{q}^{\infty} +
+      \int_{0}^q
+      \right]
+      \left[
+        (1 - \phi) e^r + \phi
+      \right]^{-\xi}
+      {K^{\alpha} dx
+        \over
+        (K + x)^{\alpha + 1}
+      } \\
+      &=& {
+        p [(K + q)^{-\alpha} b + K^{-\alpha}]
+        \over
+        \alpha \xi [(1 - \phi) e^r + \phi]^\xi
+      } 
   \end{eqnarray*}
-where $\Gamma(\alpha, x)$ is the upper incomplete gamma
-function. For $x > 1$, this function may be expressed as (equation
-8.11.2 of {\it NIST Digital Library of Mathematical Functions}
-\cite{NIST.DLMF})
-\[
-\Gamma(a, z) = z^{a - 1} e^{-z}
-\sum_{k=0}^\infty (-1)^k {\Gamma(a) \over \Gamma(a - k)} z^{-k}
-\]
-Thus after re-arranging terms we get
-\begin{eqnarray}
-  {1 \over 1 + b} I_3
-  &=&
-  {p \over 1 - \gamma} ((1 - \phi) e^r)^{1 - \gamma} \nonumber \\
-  &&
-  + \sum_{k=0}^\infty
-  {\alpha p ((1 - \phi) e^r)^{1 - \gamma}\over M^{k+1}}
-  \prod_{i=0}^{k-1} (\alpha + i)
-  \sum_{n=1}^\infty {(-1)^{n-1} \over n! n^{k+1}}
-  \left(
-  {\phi \over e^{M + r} (1 - \phi)}
-  \right)^n \times \nonumber \\
-  && \prod_{i=1}^{n-1} (\gamma + i - 1)
-  \label{eq:xxie1}
-\end{eqnarray}
-Observe that the sum over $n$ does not involve $\alpha$. Denote
-\[
-a_n(k) =
-{(-1)^{n-1} \over n! n^{k+1}}
-\left(
-     {\phi \over e^{M + r} (1 - \phi)}
-     \right)^n
-     \prod_{i=1}^{n-1} (\gamma + i - 1)
-\]
-For each given $k$, $a_n(k)$ is an alternating series.
-\begin{eqnarray*}
-  && {a_n(k) \over |a_{n+1}(k)|} \\
-  &=& {
-    e^{M + r} (1 - \phi) (n+1)^{k+2}
-    \over
-    \phi n^{k+1} (n + \kappa - 1)
-  } \\
-  &=&
-  \left(1 + {2 - \kappa \over n + \kappa - 1}\right)
-  \left( 1 + {1 \over n} \right)^{k+1}
-  {e^{M + r} (1 - \phi) \over \phi}
-\end{eqnarray*}
-
-If ${e^{M + r} (1 - \phi) \over \phi} > 1$, it is clear
-$a_n(k)/|a_{n+1}(k)| > 1$. It follows that $I_3$ and hence $u(v) +
-u(\delta v)$ is monotonically increasing with $\alpha$. If all
-investors are guided by and hence seek to maximize the same utility
-function, they would all 
-have constructed their portfolios by choosing from a set of equities
-that have tail indices close to the largest in the market. At market
-equilibrium, all actively traded equities would have tail indices that
-are indistinguishable from each other.
-
-\textcolor{red}{
-  What happens when ${e^{M + r} (1 - \phi) \over \phi} \leq 1$?
-}
-
-\subsubsection{very risk-avert utility functions}
-In this section we consider power utility functions of
-\eqref{eq:power_utility} with $\gamma > 1$. For convenience, let $ĸ =
-\gamma - 1$.
+\end{proof}
+
+%% \[
+%% F(x) = \left\{
+%% \begin{array}{ll}
+%%     p M^\alpha / |x|^\alpha & \text{ if } x \leq -M \\
+%%     F_u(x) & \text{ if } -M < x < M\\
+%%     1 - p M^\alpha x^\alpha & \text{ if } x \geq M
+%% \end{array}
+%% \right.
+%% \]
+%% where the distribution function $F_u(\cdot)$ satisfies
+%% \begin{eqnarray*}
+%%   F_u(-M) &=& p \\
+%%   F_u(M) &=& 1 - p
+%% \end{eqnarray*}
+%% In this case we have
+%% \begin{eqnarray*}
+%%   && u(v) + b u(\delta v) \\
+%%   &=&
+%%   {
+%%     (1 + b) p \alpha M^\alpha
+%%     \over
+%%     \kappa
+%%   }
+%%   \int_{-\infty}^{-M}
+%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} |x|^{-\kappa - 1}
+%%   dx + \\
+%%   &&
+%%   (1 + b) \int_{-M}^{\delta v}
+%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} d F_u(x) + \\
+%%   &&
+%%   {
+%%     (1 + b) p \alpha M^\alpha
+%%     \over
+%%     \kappa
+%%   }
+%%   \int_{M}^{\infty}
+%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} x^{-\kappa - 1}
+%%   dx \\
+%%   &=&
+%%   I_1 + I_2 + I_3
+%% \end{eqnarray*}
 
@@
 
-\section{Hoga Test}
+%% \section{Hoga Test}
 %% TODO: Show the test statistics of Hoga test applied to a sequence of

### 2017-02-21 papers/FX/draft.tex
Finished the proof with the 1st power-utility function.

@@
 
-\input{../../physics_common}
+\usepackage{amssymb,amsmath,amsthm,amsfonts}
+\usepackage{mathrsfs}
+\usepackage{dsfont}
+\usepackage{enumerate}
+
+%% \newtheorem{mdef}{Definition}
+
+\newcommand{\eqsplit}[2]{
+  \begin{equation}\label{#2}
+    \begin{split}
+      #1
+    \end{split}
+  \end{equation}}
+\newcommand{\eqnsplit}[1]{
+  \begin{eqnarray*}
+    #1
+  \end{eqnarray*}}
+\newcommand{\tran}[1]{
+  \tilde{#1}
+}
+\newcommand{\td}[2]{
+  \frac{d #1}{d #2}
+}
+\newcommand{\pd}[2]{
+  \frac{\partial #1}{\partial #2}
+}
+\newcommand{\ppd}[2]{
+  \frac{\partial^2 #1}{\partial #2^2}
+}
+\newcommand{\pdd}[3]{
+  \frac{\partial^2 #1}{\partial #2 \partial #3}
+}
+\newcommand{\otd}[1]{
+  \frac{d}{d #1}
+}
+\newcommand{\opd}[1]{
+  \frac{\partial}{\partial #1}
+}
+\newcommand{\oppd}[1]{
+  \frac{\partial^2}{\partial #1^2}
+}
+\newcommand{\opdd}[2]{
+  \frac{\partial^2}{\partial #1 \partial #2}
+}
+\newcommand{\ket}[1]{
+  |#1\rangle
+}
+\newcommand{\bra}[1]{
+  \langle#1|
+}
+\newcommand{\inn}[1]{
+  \langle#1\rangle
+}
+\newcommand{\mean}[1]{
+  \langle#1\rangle
+}
+\newcommand{\tr}{
+  \text{tr}\,
+}
+\newcommand{\re}{
+  \text{Re}\,
+}
+\newcommand\im{
+  \text{Im}\,
+}
+\newcommand{\arcsinh}{
+  \sinh^{-1}
+}
+\newcommand{\arccosh}{
+  \cosh^{-1}
+}
+\newcommand{\erfc}{
+  \text{erfc}
+}
+\newcommand{\E}{
+  \mathbb{E}
+}
+\renewcommand{\P}{
+  \mathbb{P}
+}
+\newcommand{\I}{
+  \mathbf{1}
+}
+\newcommand{\1}[1]{
+  \mathbf{1}_{\{#1\}}
+}
+\newcommand{\diag}{
+  \text{diag\,}
+}
+\newcommand{\pop}[1]{
+\varepsilon_{#1}
+}
+
+\newcommand{\M}{
+  {\text{max}}
+}
+\newcommand{\m}{
+  {\text{min}}
+}
+% \newcommand{\ph}{
+%   {\text{arg}\,}
+% }
+\newcommand\erf{
+  \text{erf}
+}
+\renewcommand\vec[1]{
+  \mathbf{#1}
+}
+\newcommand\cov{
+  \text{cov}
+}
+\newcommand\var{
+  \text{var}
+}
+\newcommand\cor{
+  \text{cor}
+}
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
+\newcommand\mtx[1]{
+  \mathbf{#1}
+}
+\renewcommand \L {
+  \mathfrak{leb}
+}
+\newcommand{\hot}{
+  \text{h.o.t.}
+}
+
+%% \newtheorem{proposition}{Proposition}
+\newtheorem{theorem}{Theorem}
+\newtheorem{acknowledgement}{Acknowledgement}
+\newtheorem{algorithm}{Algorithm}
+\newtheorem{axiom}{Axiom}
+\newtheorem{case}{Case}
+\newtheorem{claim}{Claim}
+\newtheorem{conclusion}{Conclusion}
+\newtheorem{condition}{Condition}
+\newtheorem{conjecture}{Conjecture}
+\newtheorem{corollary}{Corollary}
+\newtheorem{criterion}{Criterion}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+\newtheorem{exercise}{Exercise}
+\newtheorem{lemma}{Lemma}
+\newtheorem{notation}{Notation}
+\newtheorem{problem}{Problem}
+\newtheorem{proposition}{Proposition}
+\newtheorem{remark}{Remark}
+\newtheorem{solution}{Solution}
+\newtheorem{summary}{Summary}
+
 \title{Tail Indices and Scale Parameters in Financial Time Series}
@@
 where $\gamma > 0$. For simplicity we assume the investor
-initially has one unit of wealth. Let $r$ be the return on the
-riskless state bond and $X$ be the return on equity. Portfolio weights
+initially has one unit of wealth. Let $r > 0$ be the return on the
+risk free state bond and $X$ be the return on equity. Portfolio weights
 are $\left(1-\phi\right)$ and $\phi$ respectively. Then
@@
 \end{equation*}
-Assume $\alpha, \beta > 1$, $0 < p < 1$, $K > 0$. Equation
+Assume $\alpha, \beta > 1$, $0 < p < 1$ and
+\begin{equation}
+  \label{eq:xxie1.1}
+  0 < K <
+  {\alpha \over \alpha + 1}
+  (e^r - 1)^\xi
+  \exp\left(
+    - r \xi
+    + {1 \over \alpha}
+  \right)
+\end{equation}
+Equation
 \eqref{eq:xxie0} can be re-arranged as
@@
 =
-\E u(C) + (1 + b) \E [u(\delta v) - u(C); C < \delta v]
+\E u(C) + b \E [u(C); C < \delta v]
 \]
@@
 = 
-\E u(C) + (1 + b) \E [u(\delta v) - u(C); C < \delta v]
+\E u(C) + b \E [u(C); C < \delta v]
 \]
@@
   \begin{enumerate}[i.]
-  \item When $C < \delta v$ implies $X < -q$ for some $q > 0$, we have
+  \item When $C < \delta v$ implies $X < -q$ for some $q \geq 0$, we have
     \begin{eqnarray*}
-      \pd{\mathcal U(F)}{\alpha}
+      && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\
       &=&
@@
           \left[
-        (1 - \phi) e^r + \phi e^x
+            (1 - \hat\phi) e^r + \hat\phi e^x
           \right]^{-\xi}
@@
         dx \right\} \\
-      &=&
-      \underbrace{
+      &=& \underbrace{
         {p \over \xi}
@@
         \left[
-        (1 - \phi) e^r + \phi e^{-x}
+          (1 - \hat\phi) e^r + \hat\phi e^{-x}
         \right]^{-\xi}
@@
         }
-      }_{L_1}\\
+      }_{L_1} \\
       && - \underbrace{
@@
         \left[
-          (1 - \phi) e^r + \phi e^{-x}
+          (1 - \hat\phi) e^r + \hat\phi e^{-x}
         \right]^{-\xi}
         {K^{\alpha} dx \over (K + x)^{\alpha + 1} }
-      }_{L_2} \\
+      }_{L_2}
     \end{eqnarray*}
@@
         \over
-        \xi (1 - \phi)^\xi e^{r \xi}
+        \xi (1 - \hat\phi)^\xi e^{r \xi}
       } \int_0^\infty
@@
         \over
-        \xi (\alpha - 1) (1 - \phi)^\xi e^{r \xi}
+        \xi (\alpha - 1) (1 - \hat\phi)^\xi e^{r \xi}
       } < \infty
@@
       \left[
-        (1 - \phi) e^r + \phi e^{-x}
+        (1 - \hat\phi) e^r + \hat\phi e^{-x}
       \right]^{-\xi}
@@
     \left[
-        (1 - \phi) e^r + \phi e^{-x}
+      (1 - \hat\phi) e^r + \hat\phi e^{-x}
     \right]^{-\xi}
@@
       \left[
-        (1 - \phi) e^r + \phi e^{-x}
-        \right]^{-\xi}
-      \alpha K^{\alpha} (K + x)^{-\alpha - 1} dx \\
-      &=& {\alpha L_2 \over c} \\
-      L_1 &=& c \E[Y \ln(Y/K)]
+        (1 - \hat\phi) e^r + \hat\phi e^{-x}
+      \right]^{-\xi} {
+        \alpha K^{\alpha} x dx
+        \over
+        (K + x)^{\alpha + 2}
+      } \\
+      &=& {\alpha L_2 \over c} - K \\
+      L_1 &=& c \E \left[
+        (Y + K) \ln \left(1 + {Y \over K} \right)
+      \right]
     \end{eqnarray*}
@@
     \left[
-        x \ln\left({x \over K}\right)
-      \right] = {1 \over x} > 0
+      (x + K) \ln\left(1 + {x \over K}\right)
+    \right] = {1 \over x + K} > 0
     \quad \text{ for } x > 0
     \]
-    That is, $x \ln(x/K)$ defined on $(0, \infty)$ is a convex
+    That is, $(x + K) \ln(1 + x/K)$ defined on $(0, \infty)$ is a convex
     function. Now that $L_1, L_2 < \infty$, Jensen's inequality gives
     \[
-      {L_1 \over c} = \E [Y \ln(Y/K)]
+    {L_1 \over c} = \E \left[
+      (Y + K) \ln \left(1 + {Y \over K} \right)
+    \right]
     >
-      \E Y \times \ln(\E Y/K)
+    (\E Y + K) \times \ln(1 + \E Y/K)
     =
@@
     \ln\left(
-      {\alpha L_2  \over c K}
+      {\alpha L_2  \over c}
     \right)
     \]
-  \end{enumerate}
     Therefore
@@
       \alpha \ln \left(
-    {\alpha L_2 \over c K}
+        {\alpha L_2 \over c}
       \right) - 1
@@
     \alpha \ln \left(
-    {\alpha L_2 \over c K}
+      {\alpha L_2 \over c}
     \right) - 1 > 0
     \]
+    Or equivalently
+    \begin{equation}
+      \label{eq:xxie5}
+      {L_2 \over c} > {1 \over \alpha} \exp\left( {1 \over \alpha} \right)
+    \end{equation}
     Observe
@@
       \left[
-        (1 - \phi) e^r + \phi
+        (1 - \hat\phi) e^r + \hat\phi
       \right]^{-\xi}
@@
       &=& {
-        p [(K + q)^{-\alpha} b + K^{-\alpha}]
-        \over
-        \alpha \xi [(1 - \phi) e^r + \phi]^\xi
+        p [(K + q)^{-\alpha} b + K^{-\alpha}] K^\alpha
+        \over
+        \alpha \xi [(1 - \hat\phi) e^r + \hat\phi]^\xi
       } 
     \end{eqnarray*}
+    Meanwhile
+    \begin{eqnarray*}
+      c &<& {p \over \xi}
+      \left[
+        (1 - \hat\phi) e^r
+      \right]^{-\xi}
+      \left[
+        (1 + b)\int_{q}^{\infty} +
+        \int_{0}^q
+      \right]
+      {
+        \alpha K^{\alpha} dx \over (K + x)^{\alpha + 2}
+      } \\
+      &=&
+      {
+        \alpha K^\alpha p \left[ (K + q)^{-\alpha - 1} b + K^{-\alpha - 1} \right]
+        \over
+        \xi (\alpha + 1) (1 - \hat\phi)^\xi e^{r \xi}
+      }
+    \end{eqnarray*}
+    Thus, to show inequality \eqref{eq:xxie5}, it is sufficient to show
+    \begin{equation}
+      \label{eq:xxie6}
+      {
+        \left[
+          b (K + q)^{-\alpha} + K^{-\alpha}
+        \right]
+        \left[
+          (1 - \hat\phi) e^r + \hat\phi
+        \right]^{-\xi}
+        (\alpha + 1)
+        \over
+        \alpha^2
+        \left[
+          b (K + q)^{-\alpha - 1} + K^{-\alpha - 1}
+        \right]
+        \left[
+          (1 - \hat\phi) e^r
+        \right]^{-\xi}
+      } > {1 \over \alpha} \exp\left( {1 \over \alpha} \right)
+    \end{equation}
+    This simplies to
+    \begin{equation}
+      \label{eq:xxie7}
+      \hat\phi < {
+        g - 1
+        \over
+        (1 - e^{-r}) g - 1
+      }
+    \end{equation}
+    where
+    \[
+    g = 
+    \left\{
+      {
+        \left[
+          (K + q)^{\alpha + 1} + b K^{\alpha + 1}
+        \right] \alpha e^{1/\alpha}
+        \over
+        (\alpha + 1) K (K + q) 
+        \left[
+          (K + q)^{\alpha} + b K^{\alpha}
+        \right]
+      }
+    \right\}^{1/\xi}
+    \]
+    Assumption \eqref{eq:xxie1.1} ensures $g > e^r/(e^r - 1)$. Hence the
+    right-hand-side of inequality \eqref{eq:xxie7} is larger than 1.
+    So \eqref{eq:xxie7} is satisfied by all $\hat\phi$. Thus
+    \eqref{eq:xxie6} holds and we conclude
+    \[
+    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
+    \]
+
+  \item When $C < \delta v$ implies $X < q$ for some $q > 0$, we have
+    \begin{eqnarray*}
+      && \mathcal U(F)(\alpha, \beta, K, K', \hat\phi) \\
+      &=& \alpha p K^\alpha (1 + b) \int_{-\infty}^0
+      [(1 - \hat\phi) e^r + \hat\phi e^x]^{-\xi}
+      (K - x)^{-\alpha-1} dx \\
+      &&
+      + (1 - p) \beta K'^\beta \left[
+        (1 + b) \int_0^q + \int_q^\infty
+      \right]
+      [(1 - \hat\phi) e^r + \hat\phi e^x]^{-\xi} (K' + x)^{-\beta-1} dx    
+    \end{eqnarray*}
+  \end{enumerate}
+  So the same calculation as in the previous case carries through
+  with $q = 0$. The same conclusion of
+  \[
+  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
+  \]
+  stands.
 \end{proof}
 
-%% \[
-%% F(x) = \left\{
-%% \begin{array}{ll}
-%%     p M^\alpha / |x|^\alpha & \text{ if } x \leq -M \\
-%%     F_u(x) & \text{ if } -M < x < M\\
-%%     1 - p M^\alpha x^\alpha & \text{ if } x \geq M
-%% \end{array}
-%% \right.
-%% \]
-%% where the distribution function $F_u(\cdot)$ satisfies
-%% \begin{eqnarray*}
-%%   F_u(-M) &=& p \\
-%%   F_u(M) &=& 1 - p
-%% \end{eqnarray*}
-%% In this case we have
-%% \begin{eqnarray*}
-%%   && u(v) + b u(\delta v) \\
-%%   &=&
-%%   {
-%%     (1 + b) p \alpha M^\alpha
-%%     \over
-%%     \kappa
-%%   }
-%%   \int_{-\infty}^{-M}
-%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} |x|^{-\kappa - 1}
-%%   dx + \\
-%%   &&
-%%   (1 + b) \int_{-M}^{\delta v}
-%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} d F_u(x) + \\
-%%   &&
-%%   {
-%%     (1 + b) p \alpha M^\alpha
-%%     \over
-%%     \kappa
-%%   }
-%%   \int_{M}^{\infty}
-%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} x^{-\kappa - 1}
-%%   dx \\
-%%   &=&
-%%   I_1 + I_2 + I_3
-%% \end{eqnarray*}
-
-\section{Review of multivariate GARCH models: same tail index or not}
+\section{Review of multivariate GARCH models:
+  same tail index or not}
 Many multivariate GARCH models imply a common tail index for all
@@
 \bibliographystyle{unsrt}
-\bibliography{../../thesis/econophysics}
+\bibliography{my_citations}
 \end{document}

### 2017-02-22 notes/PotentialProjects.tex
Added a number of scripts and files made in Rotterdam.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

\input{../physics_common}
\title{Potential Projects}
\author{Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

\begin{enumerate}
\item Eigenvalues and eigenvectors of tail dependence matrix
\item Extend the theory with Casper and Thomas to situations when the
  return distribution of the stock has a Poisson component.
\item With Cheng Dan: The old project with Thomas and Olivier.
\item Cellular automata. Find out whether the theories of random
  matrices and extreme values are useful.
\item {\bf Maybe we can model market crash as a phase transition when
  the probabilistic rule involves a heavy-tail distribution?}
  \begin{itemize}
  \item Critical Behavior of a Probabilistic Local and Nonlocal
    Site-Exchange Cellular Automaton
  \item Boccara, N.; Nasser, J.; Roger, M.
  \item International Journal of Modern Physics C, Volume 5, Issue 03,
    pp. 537-545 (1994). (IJMPC Homepage)
  \item Publication Date:	00/1994
  \item Keywords: Cellular Automata, Critical Behavior
  \item Abstract:
    We study the critical behavior of a probabilistic automata network
    whose local rule consists of two subrules. The first one, applied
    synchronously, is a probabilistic one-dimensional range-one cellular
    automaton rule. The second, applied sequentially, exchanges the values
    of a pair of sites. According to whether the two sites are
    first-neighbors or not, the exchange is said to be local or
    nonlocal. The evolution of the system depends upon two parameters, the
    probability p characterizing the probabilistic cellular automaton, and
    the degree of mixing m resulting from the exchange process. Depending
    upon the values of these parameters, the system exhibits a bifurcation
    similar to a second order phase transition characterized by a
    nonnegative order parameter, whose role is played by the stationary
    density of occupied sites. When m is very large, the correlations
    created by the application of the probabilistic cellular automaton
    rule are destroyed, and, as expected, the behavior of the system is
    then correctly predicted by a mean-field-type approximation. According
    to whether the exchange of the site values is local or nonlocal, the
    critical behavior is qualitatively different as m varies.
  \end{itemize}

\end{enumerate}
\bibliographystyle{unsrt}
\bibliography{../thesis/econophysics}
\end{document}

### 2017-02-22 papers/FX/check_fx_data.r

@@ -1,0 +1,17 @@
+graphics.off();
+database = dbConnect(MySQL(), user='root', password='q1w2e3r4',
+                     dbname='avanza', host="localhost");
+
+result <- dbSendQuery(database,
+                      "select distinct day from USD_SEK_Rates where day between \"2010-01-04\" and \"2016-04-01\";")
+D = fetch(result, n=-1)[[1]];
+X <- getAssetPrices("2010-01-04", "2016-04-01", currencies, 1,
+                     "rate", "localhost");
+
+par(mfrow=c(3,6));
+## mse <- c(0, 0);
+for (i in 1:p) {
+    plot(as.xts(X[, i], order.by=as.Date(D)),
+         main=names[i]);
+}
+

### 2017-02-22 papers/FX/draft.tex
Corrected the condition on K.

@@
 \end{equation*}
-Assume $\alpha, \beta > 1$, $0 < p < 1$ and
-\begin{equation}
-  \label{eq:xxie1.1}
-  0 < K <
-  {\alpha \over \alpha + 1}
-  (e^r - 1)^\xi
-  \exp\left(
-    - r \xi
-    + {1 \over \alpha}
-  \right)
-\end{equation}
+Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
 Equation
@@
   \]
-  $G(\alpha, \beta, K, K')$ is monotonically increasing with
-  $\alpha$.
+  Let the disappointment threshold $C < \delta v$ be equivalent to $X < -q$. Assume
+  \begin{equation}
+    \label{eq:xxie1.1}
+    r > -\ln \left\{
+      1 - \left[
+          {\alpha + 1 \over \alpha}
+          {
+            \left[
+              b K^{\alpha + 1} + (K + q)^{\alpha + 1}
+            \right]
+            \over
+            \left[
+              b K^{\alpha+1} \left(1 + {q \over K} \right) + (K + q)^{\alpha + 1}
+            \right]
+          }
+          {
+            1 \over K e^{1/\alpha}
+          }
+      \right]^{1/\xi}
+    \right\}
+  \end{equation}
+  Then $G(\alpha, \beta, K, K')$ is monotonically increasing with $\alpha$.
 \end{theorem}
@@
   \begin{enumerate}[i.]
-  \item When $C < \delta v$ implies $X < -q$ for some $q \geq 0$, we have
+  \item When $q \geq 0$, we have
     \begin{eqnarray*}
@@
     \end{eqnarray*}
-    Define distribution function $H(x): (0, \infty) \to (0, 1)$ by its
+    Define distribution function $H(x): [0, \infty] \to [0, 1]$ by its
     density with respect to the Lebesgue measure
@@
     \]
-    That is, $(x + K) \ln(1 + x/K)$ defined on $(0, \infty)$ is a convex
+    That is, $(x + K) \ln(1 + x/K)$ defined on $[0, \infty]$ is a convex
     function. Now that $L_1, L_2 < \infty$, Jensen's inequality gives
@@
     \end{equation}
-    This simplies to
-    \begin{equation}
+    This expression can be simplified to
+    \begin{eqnarray}
+      \left(
+        e^r + {\hat\phi \over 1 - \hat\phi}
+      \right)^{-\xi} &>&
+      {
+        \alpha e^{1/\alpha} \left[
+          (K + q)^{-\alpha - 1} b + K^{-\alpha - 1}
+        \right]
+        \over
+        (\alpha + 1) e^{r \xi} \left[
+          (K + q)^{-\alpha} b + K^{-\alpha}
+        \right]
+      } \nonumber \\
+      e^r + {\hat\phi \over 1 - \hat\phi} &<&
+      \left[
+        {
+          (K + q)^{\alpha + 1} K + b K^{\alpha + 1} (K + q)
+          \over
+          b K^{\alpha + 1} + (K + q)^{\alpha + 1}
+        } {
+          \alpha
+          \over
+          \alpha + 1
+        }
+      \right]^{-1/\xi}
+      \exp \left(
+        r - {1 \over \alpha \xi}
+      \right) \nonumber \\
+      &=&
+      \left\{
+        {\alpha + 1 \over \alpha}
+        {
+          \left[
+            b K^{\alpha + 1} + (K + q)^{\alpha + 1}
+          \right]
+          \over
+          \left[
+            b K^{\alpha+1} \left(1 + {q \over K} \right) + (K + q)^{\alpha + 1}
+          \right]
+        }
+        {
+          \exp \left(
+            r \xi - {1 \over \alpha}
+          \right) \over K
+        }
+      \right\}^{1/\xi} \nonumber \\
+      \hat\phi &<& {e^r - g \over e^r - g - 1}
       \label{eq:xxie7}
-      \hat\phi < {
-        g - 1
-        \over
-        (1 - e^{-r}) g - 1
-      }
-    \end{equation}
+    \end{eqnarray}
     where
@@
       \left\{
+        {\alpha + 1 \over \alpha}
         {
           \left[
-          (K + q)^{\alpha + 1} + b K^{\alpha + 1}
-        \right] \alpha e^{1/\alpha}
-        \over
-        (\alpha + 1) K (K + q) 
-        \left[
-          (K + q)^{\alpha} + b K^{\alpha}
-        \right]
+            b K^{\alpha + 1} + (K + q)^{\alpha + 1}
+          \right]
+          \over
+          \left[
+            b K^{\alpha+1} \left(1 + {q \over K} \right) + (K + q)^{\alpha + 1}
+          \right]
+        }
+        {
+          \exp \left(
+            r \xi - {1 \over \alpha}
+          \right) \over K
         }
@@
     \]
-    Assumption \eqref{eq:xxie1.1} ensures $g > e^r/(e^r - 1)$. Hence the
+    Assumption \eqref{eq:xxie1.1} ensures $g < e^r - 1$. Hence the
     right-hand-side of inequality \eqref{eq:xxie7} is larger than 1.
@@
 
-  \item When $C < \delta v$ implies $X < q$ for some $q > 0$, we have
+  \item When $q < 0$, we have
     \begin{eqnarray*}
@@
       + (1 - p) \beta K'^\beta \left[
-        (1 + b) \int_0^q + \int_q^\infty
+        (1 + b) \int_0^{-q} + \int_{-q}^\infty
       \right]
@@
 \end{proof}
+\begin{remark}
+  When $q = 0$, the lower-bound \eqref{eq:xxie1.1} on the interest
+  rate simplifies to
+  \begin{equation}
+    \label{eq:xxie8}
+    K < {
+      (\alpha + 1) e^{-1/\alpha}
+      \over
+      \alpha (1 - e^{-r})^\xi
+    }
+  \end{equation}
+  The right-hand-side of \eqref{eq:xxie8} is a very high upper bound
+  on the scale parameter $K$, as is revealed later in this paper.
+\end{remark}
 

### 2017-02-24 papers/FX/draft.tex
Adding a paragraph about the variability of the scale parameter.

@@
 \title{Tail Indices and Scale Parameters in Financial Time Series}
-\author{Thomas Valentin Mikosch, Casper de Vries, Xie Xiaolei}
+\author{Thomas Mikosch, Casper de Vries, Xie Xiaolei}
 \date{\today}
@@
 \]
+$C < \delta v$ implies
+\[
+X < -q(r, \phi, \delta v) := \ln \left( {
+  \delta v - (1 - \phi) e^r
+  \over
+  \phi
+} \right)
+\]
+In the rest of this section we write $q$ for $q(r, \phi, \delta v)$
+for convenience. We have
+\begin{eqnarray}
+  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
+  &=&
+  -{\alpha K^\alpha \over \xi} p
+  \int_{-\infty}^0
+      {
+       1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
+      }
+  {1 + b \1{x < -q} \over (K + x)^{\alpha + 1}} dx \nonumber \\
+  && -{\beta K'^\beta \over \xi} (1 - p)
+  \int_{0}^\infty
+  {
+   1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
+  }
+  {1 + b \1{x < -q} \over (K' + x)^{\beta + 1}} dx
+  \label{eq:xxie1.0}
+\end{eqnarray}
+
 \begin{theorem}
@@
     \label{eq:xxie1.1}
-    r > -\ln \left\{
-      1 - \left[
-          {\alpha + 1 \over \alpha}
-          {
-            \left[
-              b K^{\alpha + 1} + (K + q)^{\alpha + 1}
-            \right]
-            \over
-            \left[
-              b K^{\alpha+1} \left(1 + {q \over K} \right) + (K + q)^{\alpha + 1}
-            \right]
-          }
-          {
-            1 \over K e^{1/\alpha}
-          }
-      \right]^{1/\xi}
-    \right\}
+    K < {
+      \alpha e^{1/\alpha} (1 - e^{-r})^\xi
+      \over
+      (\alpha + 1) A
+    }
   \end{equation}
+  where
+  \[
+  A = \max_{x \geq 1} {b x + x^{\alpha + 1} \over b + x^{\alpha + 1}}
+  \]
   Then $G(\alpha, \beta, K, K')$ is monotonically increasing with $\alpha$.
@@
       {
-        \alpha K^\alpha p \left[ (K + q)^{-\alpha - 1} b + K^{-\alpha - 1} \right]
+        \alpha K^\alpha p \left[
+          (K + q)^{-\alpha - 1} b + K^{-\alpha - 1}
+        \right]
         \over
@@
         \left[
-          (1 - \hat\phi) e^r + \hat\phi
-        \right]^{-\xi}
+          (1 - \hat\phi) e^r
+        \right]^{\xi}
         (\alpha + 1)
@@
         \left[
-          (1 - \hat\phi) e^r
-        \right]^{-\xi}
+          (1 - \hat\phi) e^r + \hat\phi
+        \right]^{\xi}
       } > {1 \over \alpha} \exp\left( {1 \over \alpha} \right)
@@
       \left(
-        e^r + {\hat\phi \over 1 - \hat\phi}
+           {\hat\phi \over 1 - \hat\phi} e^{-r} + 1
       \right)^{-\xi} &>&
@@
         \over
-        (\alpha + 1) e^{r \xi} \left[
+        (\alpha + 1) \left[
           (K + q)^{-\alpha} b + K^{-\alpha}
@@
       } \nonumber \\
-      e^r + {\hat\phi \over 1 - \hat\phi} &<&
-      \left[
+      {\hat\phi \over 1 - \hat\phi}e^{-r} + 1
+      &<&
+      \left[
+        {\alpha + 1 \over \alpha}
         {
-          (K + q)^{\alpha + 1} K + b K^{\alpha + 1} (K + q)
+          b K^{\alpha + 1} (K + q)
+          + (K + q)^{\alpha + 1} K 
           \over
@@
         } {
-          \alpha
-          \over
-          \alpha + 1
-        }
-      \right]^{-1/\xi}
-      \exp \left(
-        r - {1 \over \alpha \xi}
-      \right) \nonumber \\
+          1
+          \over
+          e^{1/\alpha}
+        }
+     \right]^{1/\xi} \nonumber \\
       &=&
-      \left\{
+      \left[
         {\alpha + 1 \over \alpha}
         {
-          \left[
-            b K^{\alpha + 1} + (K + q)^{\alpha + 1}
-          \right]
-          \over
-          \left[
-            b K^{\alpha+1} \left(1 + {q \over K} \right) + (K + q)^{\alpha + 1}
-          \right]
-        }
-        {
-          \exp \left(
-            r \xi - {1 \over \alpha}
-          \right) \over K
-        }
-      \right\}^{1/\xi} \nonumber \\
-      \hat\phi &<& {e^r - g \over e^r - g - 1}
+          b \left(1 + {q \over K}\right)
+          + \left(1 + {q \over K}\right)^{\alpha + 1}
+          \over
+          b + \left(1 + {q \over K}\right)^{\alpha + 1}
+        } {
+          K
+          \over
+          e^{1/\alpha}
+        }
+      \right]^{1/\xi} \nonumber \\
+      \hat\phi &<& {1 - g \over 1 - g - e^{-r}}
       \label{eq:xxie7}
@@
     g = 
-      \left\{
+    \left[
       {\alpha + 1 \over \alpha}
       {
-          \left[
-            b K^{\alpha + 1} + (K + q)^{\alpha + 1}
+        b \left(1 + {q \over K}\right)
+        + \left(1 + {q \over K}\right)^{\alpha + 1}
+        \over
+        b + \left(1 + {q \over K}\right)^{\alpha + 1}
+      } {
+        K
+        \over
+        e^{1/\alpha}
+      }
+      \right]^{1/\xi}
+    \]
+    Inequality \eqref{eq:xxie7} is satisfied by all possible
+    $\hat\phi$ if
+    \begin{eqnarray*}
+      g &<& 1 - e^{-r}
+    \end{eqnarray*}
+    Observe that the expression
+    \[
+    {
+      b \left(1 + {q \over K}\right)
+      + \left(1 + {q \over K}\right)^{\alpha + 1}
+      \over
+      b + \left(1 + {q \over K}\right)^{\alpha + 1}
+    }
+    \]
+    is 1 and increasing at $q = 0$; as $q \to \infty$, it tends to
+    1. So it must have a maximum value for
+    $q \in (0, \infty)$. Let
+    \[
+    A = \max_{x \geq 1} {b x + x^{\alpha + 1} \over b + x^{\alpha + 1}}
+    \]
+    Then $g < 1 - e^{-r}$ is implied by
+    \begin{eqnarray*}
+      {\alpha + 1 \over \alpha}
+      {K \over e^{1/\alpha}}
+      A &<& (1 - e^{-r})^\xi \\
+      K &<& {
+        \alpha e^{1/\alpha} (1 - e^{-r})^\xi
+        \over
+        (\alpha + 1) A
+      }
+    \end{eqnarray*}
+    The last inequality is indeed assumption \eqref{eq:xxie1.1}.
+    So, under this assumption, \eqref{eq:xxie7} is satisfied by all
+    $\hat\phi$ and \eqref{eq:xxie6} holds. We conclude
+    \[
+    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
+    \]
+
+  \item When $q < 0$, it follows from \eqref{eq:xxie1.0}:
+    \begin{eqnarray*}
+      && \opd{\alpha}\mathcal U(F)(\alpha, \beta, K, K', \hat\phi) \\
+      &=& \opd{\alpha} \left[
+        -{\alpha K^\alpha \over \xi} p
+        \int_{-\infty}^0
+            {
+              1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
+            }
+            {1 + b \over (K + x)^{\alpha + 1}} dx
         \right]
-          \over
-          \left[
-            b K^{\alpha+1} \left(1 + {q \over K} \right) + (K + q)^{\alpha + 1}
-          \right]
-        }
-        {
-          \exp \left(
-            r \xi - {1 \over \alpha}
-          \right) \over K
-        }
-      \right\}^{1/\xi}
-    \]
-    Assumption \eqref{eq:xxie1.1} ensures $g < e^r - 1$. Hence the
-    right-hand-side of inequality \eqref{eq:xxie7} is larger than 1.
-    So \eqref{eq:xxie7} is satisfied by all $\hat\phi$. Thus
-    \eqref{eq:xxie6} holds and we conclude
-    \[
-    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
-    \]
-
-  \item When $q < 0$, we have
-    \begin{eqnarray*}
-      && \mathcal U(F)(\alpha, \beta, K, K', \hat\phi) \\
-      &=& \alpha p K^\alpha (1 + b) \int_{-\infty}^0
-      [(1 - \hat\phi) e^r + \hat\phi e^x]^{-\xi}
-      (K - x)^{-\alpha-1} dx \\
-      &&
-      + (1 - p) \beta K'^\beta \left[
-        (1 + b) \int_0^{-q} + \int_{-q}^\infty
-      \right]
-      [(1 - \hat\phi) e^r + \hat\phi e^x]^{-\xi} (K' + x)^{-\beta-1} dx    
     \end{eqnarray*}
@@
   So the same calculation as in the previous case carries through
-  with $q = 0$. The same conclusion of
+  with 0 substituted for $q$. The same conclusion of
   \[
@@
 \end{proof}
-\begin{remark}
-  When $q = 0$, the lower-bound \eqref{eq:xxie1.1} on the interest
-  rate simplifies to
-  \begin{equation}
-    \label{eq:xxie8}
-    K < {
-      (\alpha + 1) e^{-1/\alpha}
-      \over
-      \alpha (1 - e^{-r})^\xi
-    }
-  \end{equation}
-  The right-hand-side of \eqref{eq:xxie8} is a very high upper bound
-  on the scale parameter $K$, as is revealed later in this paper.
-\end{remark}
+Since $u(v) + b u(\delta v)$ given in \eqref{eq:xxie1.0} as a
+functional of the distribution function of $X$ is increasing with the
+lower tail index $\alpha$, all agents guided by this class of utility
+functions will seek the largest $\alpha$ found in a stock group e.g.
+a sector. As a result, one can expect all the actively traded equities
+in the group to feature the same $\alpha$. In section
+\ref{sec:data_analysis} we analyse real-world data of stock returns to
+verify this conclusion.
+
+In comparison to the theoretical constancy of $\alpha$, the scale and
+shift parameter $K$ in the lower-tail distribution of $X$ may vary
+hugely from one equity to another. This could be understood also by
+computing $d\mathcal U F / dK$:
 
@@
 
-\section{Data Analysis}
+\section{Data Analysis} \label{sec:data_analysis}
 \subsection{Tail index and scale parameter}

### 2017-02-27 papers/FX/draft.tex
Looked again at the variation of K.
We need to do some estimation about the assumed model.

@@
 distribution function $F(\cdot)$ of the equity return:
-\begin{equation*}
+\begin{equation}
+  \label{eq:X_distr}
   F(x) = \left\{
@@
   \right.
-\end{equation*}
+\end{equation}
 Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
@@
       }
-  {1 + b \1{x < -q} \over (K + x)^{\alpha + 1}} dx \nonumber \\
+  {1 + b \1{x < -q} \over (K - x)^{\alpha + 1}} dx \nonumber \\
   && -{\beta K'^\beta \over \xi} (1 - p)
@@
 
-%% with continuous and strictly increasing
-%% distribution function $F(\cdot)$. Cheng and Peng (2001)
-%% \cite{cheng2001confidence} showed asymptotic normal distribution of
-%% the estimator $\alpha_n$:
-%% \begin{proposition}\label{prop:Hill_asymp_dist}
-%%   Suppose $X_1, ..., X_n$ are independent identically distributed and
-%%   their distribution function $F(\cdot)$ satisfies, as $x \to \infty$,
-%%   \[
-%%   1 - F(x) = A x^{-\alpha} [1 + B x^{-\beta} + o(x^{-\beta})]
-%%   \]
-%%   where $A, \alpha, \beta > 0$ and $B \neq 0$. Then
-%%   \[
-%%   \sqrt k (1/\alpha_n - 1/\alpha) \overset{d}{\to}
-%%   N(0, {1 \over \alpha^2})
-%%   \]
-%%   if and only if $k(n) \to \infty$ and $k(n) = o(n^{2\beta/(\alpha +
-%%     2\beta)})$.
-%% \end{proposition}
-%% \begin{proof}
-%%   This is proposition 1 of Cheng and Peng (2001)
-%%   \cite{cheng2001confidence}. See also Embrechts et al (1997)
-%%   \cite{Embrechts1997}.
-%% \end{proof}
-
-\section{Peaks over Threshold}
-An alternative to the Hill estimator is the method of ``Peaks over
-Threshold'' (cf. Embrechts et al (1997)
-\cite{Embrechts1997}). Consider random variable $X$ whose distribution
-satisfies 
-\[
-\P(X > x) = A (x + a)^{-\alpha}
-\]
-for all $x > M$ for some $M > 0$. Then it immediately follows
-\begin{eqnarray*}
-  \P(X > u + x)
-  &=& A (u + a)^{-\alpha} \left(
-  {x \over u + a} + 1
-  \right)^{-\alpha} \\
-  \P(X > u + x | X > u) &=& \left(
-    {x \over u + a} + 1
-    \right)^{-\alpha}
-\end{eqnarray*}
-The density function of the above conditional probability is
-\[
-f_u(x) = {
-  \alpha \over x + u + a
-} \left(
-{x \over u + a} + 1
-\right)^{-\alpha}
-\]
-
-Now suppose we have a sample $X_1, ..., X_N$ where $X_1, ..., X_n > u$
-and $X_i - u = x_i > 0$, for $i=1,2,...,n$. Using the above
-conditional density function, the log-likelihood function of the
-parameters $\alpha$ and $a$ can be found:
-\[
-L(u, a) = -(1 + \alpha) \sum_{i=1}^n \ln(x_i + u + a) +
-u \alpha \ln(u + a) + n \ln(\alpha)
-\]
-It follows
-\begin{eqnarray*}
-  \pd{L(\alpha, a)}{\alpha} &=& \sum_{i=1}^n \ln(x_i + u + a)
-  + n \ln(u + a) + n/\alpha \\
-  \pd{L(\alpha, a)}{a} &=&
-  -(1 + \alpha) \sum_{i=1}^n {1 \over x_i + u + a}
-  + {n \alpha \over u + a}
-\end{eqnarray*}
-Setting the two derivatives to 0 gives
-\begin{eqnarray*}
-  \left[
-    {1 \over n} \sum_{i=1}^n \ln(x_i + u + a)
-    - \ln(u + a)
-    \right]^{-1} &=& \alpha\\
-  \left[
-    {1 \over n} \sum_{i=1}^n (y_i + 1)
-    \right] \left[
-    {1 \over n} \sum_{i=1}^n e^{-y_i}
-    \right] &=& 1
-\end{eqnarray*}
-where
-\begin{eqnarray*}
-  y_i &=& \ln\left(
-  {x_i \over u + a} + 1
-  \right)
-\end{eqnarray*}
-Solving the last equation numerically and substituting the result into
-the first one gives an estimate for $\alpha$ and $a$, call them $\hat
-\alpha$ and $\hat a$. Once $\hat \alpha$ and $\hat a$ have been
-obtained, an estimate of $A$, call it $\hat A$, can be obtained by
-the method of minimum mean square error.
-\begin{eqnarray*}
-  \hat P(X > x) = {1 \over N} \sum_{i=1}^N \1{X_i > x}
-  &=& \hat A (x + \hat a)^{-\hat \alpha} \\
-  \ln[\hat P(X > x)] &=& \ln(\hat A) - \hat \alpha \ln(x + \hat a)
-\end{eqnarray*}
-for all $x > M$. Using the upper order statistics of the
-sample $X_1, ..., X_N$, we have
-\[
-\ln\left({k-1 \over N}\right)
-=
-\ln(\hat A) - \hat \alpha \ln(X_{(k)} + \hat a)
-\]
-where $X_{(k)}$ denotes the $k$-th upper order statistic of the
-sample. Thus, by fitting $\ln\left({k-1 \over N}\right)$ as a linear
-function of $\ln(X_{(k)} + \hat a)$, one can find $\ln(\hat A)$ as the
-intercept.
-
-Before computing the estimates, one should certainly check whether the
-model makes sense for a given data series. One way to do this is to
-plot the {\it mean excess function}. For the tail function in the
-above, it follows
-\begin{eqnarray}
-  e(u) &=& \E (X - u | X > u) \nonumber \\
-  &=& -\int_0^\infty x  d\P(X > u + x | X > u) \nonumber \\
-  &=& \int_0^\infty \left(
-  {x \over u + a} + 1
-  \right)^{-\alpha} dx \nonumber \\
-  &=& {u + a \over \alpha - 1} \label{eq:MEF}
-\end{eqnarray}
-where we assume $\alpha > 1$ so that the conditional expectation is
-finite. It is seen from this formula that mean excesses are linear in
-the thresholds $u$ -- this may prompt discarding the farthest tail
-part of the data when carrying out estimation.
-%% The appearance of these data points are so rare that
-%% including them in the statistics significantly increases the variance
-%% of the estimate. This is analogous to the large variance of the Hill
-%% estimator when the number of upper order statistics used in the
-%% estimation is too small.
-\begin{wrapfigure}{r}{0.5\textwidth}
-  \centering
-  \includegraphics[width=0.45\textwidth]{ADM_MEF.pdf}
-  \caption{\scriptsize The mean excess function of ADM. Thresholds are
-    taken from upper order statistics.}
-  \label{fig:ADM_MEF}
-\end{wrapfigure}
-
-For example, figure (\ref{fig:ADM_MEF}) shows the mean
-excess function of the return series of the stock of {\it
-  Archer-Daniels-Midland Co}, a company in the {\it Consumer Staples}
-sector of S\&P 500 index.
-One can see that the 8 largest observations of the sample are
-inconsistent with the linear pattern with slope $1/(\alpha - 1)$.
-%% So we would rather not include them in the maximum likelihood
-%% estimation.
-
-It is also possible to directly use equation \eqref{eq:MEF} as a means
-of estimation. After picking out a portion of the mean excess function
-that is consistent with \eqref{eq:MEF}, one may fit a linear function
-to the empirical $\hat e(u)$ by minimizing the mean square error, and
-obtain parameter estimates from the coefficients of the fitted line.
-  %% \adjustbox{valign=t} {
-  %%   \begin{minipage}[t]{0.45\textwidth}
-  %%   \end{minipage}
-  %% }\hfill
-  %% \adjustbox{valign=t} {
-  %%   \begin{minipage}[t]{0.5\textwidth}
-  %%   \end{minipage}
-  %% }
-\section{Pickand's Estimator}
-Pickands (1975) \cite{pickands1975statistical} proposed a different
-approach to estimate the tail index than that of Hill. Assume
-$\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
-denote the quantile of $1 - 1/t$ of this distribution, i.e.
-\[
-P(X > U(t)) = {1 \over t}
-\]
-The key idea of Pickands' estimator is to observe
-\[
-\lim_{t \to \infty} {
-  U(c(t) t) - U(t)
-  \over
-  U(t) - U(t/c(t))
-} = 2^{1/\alpha}
-\]
-for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
-Pickands formulated the estimation problem in terms of the extreme
-value index $\xi$, which is the reciprocal of the tail index $\alpha$
-when the distribution of $X$ is regularly varying (cf. Embrechtsm,
-Kl\"uppelberg and Mikosch \cite{Embrechts1997}).
-
-Building upon the aforementioned observation and taking $c(t)$ to
-be simply constant 2, Pickands proposed the following estimator for
-$\xi$:
-\[
-\hat \xi = {1 \over \ln 2} \ln {
-  x_{(k)} - x_{(2k)}
-  \over
-  x_{(2k)} - x_{(4k)}
-}
-\]
-Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
-that Pickands' estimator was consistent in both the weak and the strong
-sense, when the sequence $k(n)$ was chosen appropriately:
-\begin{itemize}
-\item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
-  \[
-  \hat \xi \overset{P}{\to} \xi
-  \]
-\item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
-  as $n \to \infty$,
-  \[
-  \hat \xi \overset{a.s.}{\to} \xi
-  \]
-\end{itemize}
-Dekkers and de Haan \cite{dekkers1989estimation} also showed the
-asymptotic normality of the estimator, when some
-extra technical conditions were satisfied by
-$U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
-\[
-\sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
-\]
-where
-\[
-v(\xi) = {
-  \xi^2 (2^{2 \xi + 1} + 1)
-  \over
-  [2 \ln(2) (2^\xi - 1)]^2
-}
-\]
-
-%% \section{Hoga Test}
-%% TODO: Show the test statistics of Hoga test applied to a sequence of
-%% two concatenated return series. A changes in the scale is characterized
-%% by a smooth bump in the test statistics while a change in the tail
-%% index is characterized by a jump. Derive a theoretical result to
-%% support this?
-
-%% What is the probability of rejecting the null hypothesis when the test
-%% is applied to $(X_1, X_2)$ where $X_1$ and $X_2$ have tail indices
-%% $\kappa_1 \neq \kappa_2$? The probability of wrongly rejecting the
-%% null hypothesis must be lower than $\alpha$.
-\textcolor{red}{Review the test of Hoga}
-
-\section{Data Analysis} \label{sec:data_analysis}
-\subsection{Tail index and scale parameter}
-In this section we apply the statistical tools reviewed in the
-previous sections to real stock return series, denoted $X_t$
-hereafter. We are interested in finding out the tail indices and the
-scale parameters of the stationary distribution of $|X_t|$.
-
-%% and show that, within a
-%% comfortable confidence interval, one cannot reject the null hypothesis
-%% that these return series have the same tail index.
-
 \subsubsection{The Energy and the Consumer Staples sectors of S\&P 500}
-Figure \ref{fig:Energy_Hill_upper} shows the Hill estimates of the
-tail indices of 32 stocks in the {\it Energy} sector of S\&P 500
-index; figure \ref{fig:Consumer_Staples_Hill_upper} shows the same of
-the {\it Consumer Staples} sector.
+Figure \ref{fig:Energy_Hill_upper},
+\ref{fig:Consumer_Staples_Hill_upper},
+and \ref{fig:Information_Technology_lower}
+show the Hill estimates of the
+lower tail indices of the ``energy'', ``consumer staples'' and ``information
+technology'' sectors of the S\&P 500 index, respectively.
 \begin{figure}[htb!]
@@
 \end{figure}
-For our purpose, we have excluded a few stocks, though:
+
 \begin{figure}[htb!]
   \centering
-  \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\textwidth]{WBA_price.pdf}
-    \caption{{\it Walgreens Boots Alliance} prices}
-    \label{fig:EBA_prices}
-  \end{subfigure}
-  \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\textwidth]{WFM_price.pdf}
-    \caption{{\it Whole Foods Market} prices}
-    \label{fig:WFM_prices}
-  \end{subfigure}
-  \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\textwidth]{EL_price.pdf}
-    \caption{{\it Estee Lauder Cos.} prices}    
-    \label{fig:EL_prices}
-  \end{subfigure}
-  \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\textwidth]{CL_price.pdf}
-    \caption{{\it Colgate-Palmolive} prices}    
-    \label{fig:CL_prices}
-  \end{subfigure}
-  \caption{Plot of prices of {\it Consumer Staples} stocks
-    excluded from Hill estimates}
+  \includegraphics[width=\textwidth]{Information_Technology_lower.pdf}
+  \caption{Hill estimates of the lower-tail indices of 42 stocks' daily
+    return series in the {\it Information Technology} sector of S\&P
+    500 index.
+  }
+  \label{fig:Information_Technology_lower}
 \end{figure}
 
+\section{Hybrid estimation of the assumed distribution function}
+We adopt a hybrid procedure to estimate the parameters of the
+distribution function assumed in eq.\eqref{eq:X_distr}: First we
+obtain a Hill estimate of the lower- and upper-tail indices of an
+equity return and fix $\alpha$ and $\beta$ in \eqref{eq:X_distr}
+to their Hill estimates. Then we estimate $K$ and $K'$ by maximizing
+the likelihood function with fixed $\alpha$ and $\beta$.
+
+\section{Pickand's Estimator}
+Pickands (1975) \cite{pickands1975statistical} proposed a different
+approach to estimate the tail index than that of Hill. Assume
+$\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
+denote the quantile of $1 - 1/t$ of this distribution, i.e.
+\[
+P(X > U(t)) = {1 \over t}
+\]
+The key idea of Pickands' estimator is to observe
+\[
+\lim_{t \to \infty} {
+  U(c(t) t) - U(t)
+  \over
+  U(t) - U(t/c(t))
+} = 2^{1/\alpha}
+\]
+for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
+Pickands formulated the estimation problem in terms of the extreme
+value index $\xi$, which is the reciprocal of the tail index $\alpha$
+when the distribution of $X$ is regularly varying (cf. Embrechtsm,
+Kl\"uppelberg and Mikosch \cite{Embrechts1997}).
+
+Building upon the aforementioned observation and taking $c(t)$ to
+be simply constant 2, Pickands proposed the following estimator for
+$\xi$:
+\[
+\hat \xi = {1 \over \ln 2} \ln {
+  x_{(k)} - x_{(2k)}
+  \over
+  x_{(2k)} - x_{(4k)}
+}
+\]
+Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
+that Pickands' estimator was consistent in both the weak and the strong
+sense, when the sequence $k(n)$ was chosen appropriately:
 \begin{itemize}
-\item {\it Walgreens Boots Alliance}:
-
-  As shown in figure (\ref{fig:EBA_prices}),
-  this stock was traded only in a small fraction of the bank
-  days and it is hence no surprise that the price varied
-  tremendously.
-\item {\it Whole Foods Market}, {\it Estee Lauder Cos.} and {\it
-  Colgate-Palmolive}
-
-  As seen in figures (\ref{fig:WFM_prices}), (\ref{fig:EL_prices}) and
-  (\ref{fig:CL_prices}),
-  the prices of these stocks were reduced by half during the
-  period under consideration. But the price changes are the
-  result of emission of new shares rather than random fluctuations.
+\item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
+  \[
+  \hat \xi \overset{P}{\to} \xi
+  \]
+\item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
+  as $n \to \infty$,
+  \[
+  \hat \xi \overset{a.s.}{\to} \xi
+  \]
 \end{itemize}
-
-Figure \ref{fig:Energy_Hill_upper} and
-\ref{fig:Consumer_Staples_Hill_upper}
-are suggestive of the conclusion that excluding price
-movements due to deterministic events, tail indices of stock return
-series tend to have the same value, within a reasonable confidence
-interval.
-
-If we look instead at the scales, differences among the stock return
-series are however rather large. Figure (\ref{fig:Energy_Hill_Scales})
-and (\ref{fig:Consumer_Staples_Hill_Scales}) show the Hill estimates of
-the scale parameters of the stocks in the S\&P 500 {\it Energy} and
-{\it Consumer Staples} sectors, respectively.
-\begin{figure}[htb!]
-  \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\textwidth]{Energy_Hill_scales.pdf}
-    \caption{{\it Energy} scale estimates}    
-    \label{fig:Energy_Hill_Scales}
-  \end{subfigure}
-  \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\textwidth]{Consumer_Staples_Hill_scales.pdf}
-    \caption{{\it Consumer Staples} scale estimates}    
-    \label{fig:Consumer_Staples_Hill_Scales}
-  \end{subfigure}
-  \caption{Black boxes are Hill estimates of the scale parameters of
-    the stock return series, while red circles are the scale estimates
-    of simulated iid. Student's $t$ series with a degree of freedom
-    equal to the mean value of the estimated tail indices of the stock
-    return series. The estimates of the Student's $t$ series are
-    included for comparison with those of the return series.}
-\end{figure}
-Clearly the estimated scale parameters can differ by several orders of
-magnitude.
-
-Using the ``Peaks over Threshold'' method and the mean excess
-function, we also obtain estimates for the tail index ($\alpha$), the
-scale parameter ($A$) and the shift parameter ($a$) for stocks in the
-{\it Energy} and the {\it Consumer Staples} sectors of the S\&P 500
-index. The results are plotted in figure
-\ref{fig:Energy_OLS_estimates} and
-\ref{fig:Consumer_Staples_OLS_estimates},
-respectively.
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{Energy_OLS_estimates.pdf}
-  \caption{Tail parameter estimates of {\it Energy} stocks. The top
-    plot shows the tail index estimates, the middle plot shows the
-    10-based logarithm of the scale estimates and bottom plot shows
-    the estimates of the shift parameter.
-  }
-  \label{fig:Energy_OLS_estimates}
-\end{figure}
-
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{Consumer_Staples_OLS_estimates.pdf}
-  \caption{Tail parameter estimates of {\it Consumer Staples}
-    stocks. The top 
-    plot shows the tail index estimates, the middle plot shows the
-    10-based logarithm of the scale estimates and bottom plot shows
-    the estimates of the shift parameter.
-  }
-  \label{fig:Consumer_Staples_OLS_estimates}
-\end{figure}
+Dekkers and de Haan \cite{dekkers1989estimation} also showed the
+asymptotic normality of the estimator, when some
+extra technical conditions were satisfied by
+$U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
+\[
+\sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
+\]
+where
+\[
+v(\xi) = {
+  \xi^2 (2^{2 \xi + 1} + 1)
+  \over
+  [2 \ln(2) (2^\xi - 1)]^2
+}
+\]
+
+%% \section{Hoga Test}
+%% TODO: Show the test statistics of Hoga test applied to a sequence of
+%% two concatenated return series. A changes in the scale is characterized
+%% by a smooth bump in the test statistics while a change in the tail
+%% index is characterized by a jump. Derive a theoretical result to
+%% support this?
+
+%% What is the probability of rejecting the null hypothesis when the test
+%% is applied to $(X_1, X_2)$ where $X_1$ and $X_2$ have tail indices
+%% $\kappa_1 \neq \kappa_2$? The probability of wrongly rejecting the
+%% null hypothesis must be lower than $\alpha$.
+\textcolor{red}{Review the test of Hoga}
+
+\section{Data Analysis} \label{sec:data_analysis}
+\subsection{Tail index and scale parameter}
+In this section we apply the statistical tools reviewed in the
+previous sections to real stock return series, denoted $X_t$
+hereafter. We are interested in finding out the tail indices and the
+scale parameters of the stationary distribution of $|X_t|$.
+
+%% \subsubsection{The Energy and the Consumer Staples sectors of S\&P 500}
+%% Figure \ref{fig:Energy_Hill_upper} shows the Hill estimates of the
+%% tail indices of 32 stocks in the {\it Energy} sector of S\&P 500
+%% index; figure \ref{fig:Consumer_Staples_Hill_upper} shows the same of
+%% the {\it Consumer Staples} sector.
+%% \begin{figure}[htb!]
+%%   \centering
+%%   \includegraphics[width=\textwidth]{Energy_Hill_upper.pdf}
+%%   \caption{Hill estimates of the upper tail indices of 32 stocks'
+%%     daily return series in the {\it Energy} sector of Standard \& Poor
+%%     500 index. Each circle coresponds to a Hill estimate; the grey
+%%     triangles above and below it mark the 97.5\% and 2.5\% quantiles
+%%     of its approximate normal distribution.
+%%     The low and high red lines mark the median of the 2.5\% quantiles
+%%     and median of the 97.5\% quantiles, respectively.
+%%     The data are from {\it Yahoo Finance} and the labels on
+%%     the horizontal axis are Yahoo symbols of the stocks. The data span
+%%     from 2010-01-01 to 2015-01-01 and comprise $n=1304$
+%%     observations. $k=(\ln n)^2 = 49$ upper order statistics are used
+%%     for computing $\alpha_n$. 
+%%   }
+%%   \label{fig:Energy_Hill_upper}
+%% \end{figure}
+
+%% \begin{figure}[htb!]
+%%   \centering
+%%   \includegraphics[width=\textwidth]{Consumer_Staples_Hill_upper.pdf}
+%%   \caption{Hill estimates of the upper tail indices of 30 stocks'
+%%     daily return series in the {\it Consumer Staples} sector of
+%%     Standard \& Poor 500 index.
+%%   }
+%%   \label{fig:Consumer_Staples_Hill_upper}
+%% \end{figure}
+
+%% For our purpose, we have excluded a few stocks, though:
+%% \begin{figure}[htb!]
+%%   \centering
+%%   \begin{subfigure}{0.45\textwidth}
+%%     \includegraphics[width=\textwidth]{WBA_price.pdf}
+%%     \caption{{\it Walgreens Boots Alliance} prices}
+%%     \label{fig:EBA_prices}
+%%   \end{subfigure}
+%%   \begin{subfigure}{0.45\textwidth}
+%%     \includegraphics[width=\textwidth]{WFM_price.pdf}
+%%     \caption{{\it Whole Foods Market} prices}
+%%     \label{fig:WFM_prices}
+%%   \end{subfigure}
+%%   \begin{subfigure}{0.45\textwidth}
+%%     \includegraphics[width=\textwidth]{EL_price.pdf}
+%%     \caption{{\it Estee Lauder Cos.} prices}    
+%%     \label{fig:EL_prices}
+%%   \end{subfigure}
+%%   \begin{subfigure}{0.45\textwidth}
+%%     \includegraphics[width=\textwidth]{CL_price.pdf}
+%%     \caption{{\it Colgate-Palmolive} prices}    
+%%     \label{fig:CL_prices}
+%%   \end{subfigure}
+%%   \caption{Plot of prices of {\it Consumer Staples} stocks
+%%     excluded from Hill estimates}
+%% \end{figure}
+
+%% \begin{itemize}
+%% \item {\it Walgreens Boots Alliance}:
+
+%%   As shown in figure (\ref{fig:EBA_prices}),
+%%   this stock was traded only in a small fraction of the bank
+%%   days and it is hence no surprise that the price varied
+%%   tremendously.
+%% \item {\it Whole Foods Market}, {\it Estee Lauder Cos.} and {\it
+%%   Colgate-Palmolive}
+
+%%   As seen in figures (\ref{fig:WFM_prices}), (\ref{fig:EL_prices}) and
+%%   (\ref{fig:CL_prices}),
+%%   the prices of these stocks were reduced by half during the
+%%   period under consideration. But the price changes are the
+%%   result of emission of new shares rather than random fluctuations.
+%% \end{itemize}
+
+%% Figure \ref{fig:Energy_Hill_upper} and
+%% \ref{fig:Consumer_Staples_Hill_upper}
+%% are suggestive of the conclusion that excluding price
+%% movements due to deterministic events, tail indices of stock return
+%% series tend to have the same value, within a reasonable confidence
+%% interval.
+
+%% If we look instead at the scales, differences among the stock return
+%% series are however rather large. Figure (\ref{fig:Energy_Hill_Scales})
+%% and (\ref{fig:Consumer_Staples_Hill_Scales}) show the Hill estimates of
+%% the scale parameters of the stocks in the S\&P 500 {\it Energy} and
+%% {\it Consumer Staples} sectors, respectively.
+%% \begin{figure}[htb!]
+%%   \begin{subfigure}{0.45\textwidth}
+%%     \includegraphics[width=\textwidth]{Energy_Hill_scales.pdf}
+%%     \caption{{\it Energy} scale estimates}    
+%%     \label{fig:Energy_Hill_Scales}
+%%   \end{subfigure}
+%%   \begin{subfigure}{0.45\textwidth}
+%%     \includegraphics[width=\textwidth]{Consumer_Staples_Hill_scales.pdf}
+%%     \caption{{\it Consumer Staples} scale estimates}    
+%%     \label{fig:Consumer_Staples_Hill_Scales}
+%%   \end{subfigure}
+%%   \caption{Black boxes are Hill estimates of the scale parameters of
+%%     the stock return series, while red circles are the scale estimates
+%%     of simulated iid. Student's $t$ series with a degree of freedom
+%%     equal to the mean value of the estimated tail indices of the stock
+%%     return series. The estimates of the Student's $t$ series are
+%%     included for comparison with those of the return series.}
+%% \end{figure}
+%% Clearly the estimated scale parameters can differ by several orders of
+%% magnitude.
+
+%% Using the ``Peaks over Threshold'' method and the mean excess
+%% function, we also obtain estimates for the tail index ($\alpha$), the
+%% scale parameter ($A$) and the shift parameter ($a$) for stocks in the
+%% {\it Energy} and the {\it Consumer Staples} sectors of the S\&P 500
+%% index. The results are plotted in figure
+%% \ref{fig:Energy_OLS_estimates} and
+%% \ref{fig:Consumer_Staples_OLS_estimates},
+%% respectively.
+%% \begin{figure}[htb!]
+%%   \includegraphics[width=\textwidth]{Energy_OLS_estimates.pdf}
+%%   \caption{Tail parameter estimates of {\it Energy} stocks. The top
+%%     plot shows the tail index estimates, the middle plot shows the
+%%     10-based logarithm of the scale estimates and bottom plot shows
+%%     the estimates of the shift parameter.
+%%   }
+%%   \label{fig:Energy_OLS_estimates}
+%% \end{figure}
+
+%% \begin{figure}[htb!]
+%%   \includegraphics[width=\textwidth]{Consumer_Staples_OLS_estimates.pdf}
+%%   \caption{Tail parameter estimates of {\it Consumer Staples}
+%%     stocks. The top 
+%%     plot shows the tail index estimates, the middle plot shows the
+%%     10-based logarithm of the scale estimates and bottom plot shows
+%%     the estimates of the shift parameter.
+%%   }
+%%   \label{fig:Consumer_Staples_OLS_estimates}
+%% \end{figure}
 
@@
 
-\subsubsection{The Information Technology sector of S\&P 500}
-Compared with the sectors of energy and consumer staples, which enjoy
-a relatively stable demand on their products, companies of information
-technology are more subject to volatile demand and competition. As a
-result, one can expect their stock prices to be more volatile, implying
-lower tail indices and larger differences between the tail indices of
-different companies. In this section we investigate how these companies
-differ in terms of their tail indices and scale parameters.
-
-Figure \ref{fig:Information_Technology_Hill} shows the Hill
-estimates of 42 stocks in the {\it Information Technology} sector. Daily
-closing prices of these stocks have been available since 2000-01-01 and
-earlier.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\textwidth]{Information_Technology_Hill.pdf}
-  \caption{Hill estimates of the tail indices of 42 stocks' daily
-    return series in the {\it Information Technology} sector of Standard
-    \& Poor 500 index.
-  }
-  \label{fig:Information_Technology_Hill}
-\end{figure}
-It is seen in the figure that most of the estimates fall within the
-interval spanned by the confidence bands. Thus, given the estimation
-precision, these are consistent with the similarity conclusion of 
-\S\ref{sec:GDA}.
+%% \subsubsection{The Information Technology sector of S\&P 500}
+%% Compared with the sectors of energy and consumer staples, which enjoy
+%% a relatively stable demand on their products, companies of information
+%% technology are more subject to volatile demand and competition. As a
+%% result, one can expect their stock prices to be more volatile, implying
+%% lower tail indices and larger differences between the tail indices of
+%% different companies. In this section we investigate how these companies
+%% differ in terms of their tail indices and scale parameters.
+
+%% Figure \ref{fig:Information_Technology_Hill} shows the Hill
+%% estimates of 42 stocks in the {\it Information Technology} sector. Daily
+%% closing prices of these stocks have been available since 2000-01-01 and
+%% earlier.
+%% \begin{figure}[htb!]
+%%   \centering
+%%   \includegraphics[width=\textwidth]{Information_Technology_Hill.pdf}
+%%   \caption{Hill estimates of the tail indices of 42 stocks' daily
+%%     return series in the {\it Information Technology} sector of Standard
+%%     \& Poor 500 index.
+%%   }
+%%   \label{fig:Information_Technology_Hill}
+%% \end{figure}
+%% It is seen in the figure that most of the estimates fall within the
+%% interval spanned by the confidence bands. Thus, given the estimation
+%% precision, these are consistent with the similarity conclusion of 
+%% \S\ref{sec:GDA}.
 
@@
 
-\subsection{Temporal changes of the parameters}
+%% \subsection{Temporal changes of the parameters}
 
@@
 \bibliographystyle{unsrt}
-\bibliography{my_citations}
+\bibliography{../../thesis/econophysics}
 \end{document}

### 2017-02-28 papers/FX/draft.tex
Still struggling about K.
Started to correct the mistake related to the probability.

@@
 investors of this kind, the return distributions of equities that are
-actively traded in this market must have very similar tail indices.
+actively traded in this market must have very similar tail indices. We
+also show, in contrast, the scale parameters of the return
+distributions may differ hugely from one another.
 
@@
     {K \over K - x}
-    \right)^\alpha & x < 0 \\
+    \right)^\alpha & x \leq 0 \\
     1 - (1 - p) \left(
@@
 \eqref{eq:xxie0} can be re-arranged as
-\[
-u(v) + b u(\delta v))
-=
-\E u(C) + b \E [u(C); C < \delta v]
-\]
-The right-hand-side of the above equation is a functional of the
-distribution function of $X$ via eq.\eqref{eq:xxie1}. For convenience
+\begin{eqnarray}
+u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
+&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(-q) \label{eq:functional}
+\end{eqnarray}
+where
+\begin{eqnarray*}
+  q(r, \phi) &:=& -\ln \left( {
+    \delta v - (1 - \phi) e^r
+    \over
+    \phi
+} \right)
+\end{eqnarray*}
+Note $C < \delta v$ implies
+\[
+X < -q(r, \phi)
+\]
+The right-hand-side of \eqref{eq:functional} is a functional of the
+distribution function of $X$ via \eqref{eq:xxie1}. For convenience
 we denote it $\mathcal U(\cdot)$, i.e.
@@
 = 
-\E u(C) + b \E [u(C); C < \delta v]
-\]
-$C < \delta v$ implies
-\[
-X < -q(r, \phi, \delta v) := \ln \left( {
-  \delta v - (1 - \phi) e^r
-  \over
-  \phi
-} \right)
-\]
-In the rest of this section we write $q$ for $q(r, \phi, \delta v)$
+\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(-q)
+\]
+In the rest of this section we write $q$ for $q(r, \phi)$
 for convenience. We have
@@
   && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
-  &=&
+  &=& \underbrace{
   -{\alpha K^\alpha \over \xi} p
@@
       }
-  {1 + b \1{x < -q} \over (K - x)^{\alpha + 1}} dx \nonumber \\
+  {1 + b \1{x < -q} \over (K - x)^{\alpha + 1}} dx
+  }_{J_1} \nonumber \\
   && -{\beta K'^\beta \over \xi} (1 - p)
@@
   }
-  {1 + b \1{x < -q} \over (K' + x)^{\beta + 1}} dx
+  {1 + b \1{x < -q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
+  &&
+  \underbrace{
+    -b u(\delta v) p {K^\alpha \over (K + q)^\alpha} \1{q \geq 0}
+  }_{J_2}
+  - b u(\delta v) \left[
+    1 - (1 - p) {
+      K'^\beta
+      \over
+      (K' - q)^\beta
+    } \1{q < 0}
+  \right]
   \label{eq:xxie1.0}
@@
   So the second term of \eqref{eq:xxie3} vanishes. It remains to show
-  \[
-  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
-  \]
+  \begin{eqnarray*}
+    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
+    &=& \pd{J_1}{\alpha} + \pd{J_2}{\alpha} > 0
+  \end{eqnarray*}
+  The 2nd term is easily computed:
+  \begin{eqnarray*}
+    \pd{J_2}{\alpha} &=& b u(\delta v) p \1{q \geq 0}
+    \left(
+      {K \over K + q}
+    \right)^\alpha
+    \ln\left(
+      1 + {q \over K}
+    \right) > 0
+  \end{eqnarray*}
+  As for the 1st term we have
   \begin{enumerate}[i.]
-  \item When $q \geq 0$, we have
+  \item When $q \geq 0$,
     \begin{eqnarray*}
-      && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\
+      && \pd{J_1}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\
       &=&
@@
     \begin{eqnarray*}
-      && \opd{\alpha}\mathcal U(F)(\alpha, \beta, K, K', \hat\phi) \\
+      && \opd{\alpha}J_1(\alpha, \beta, K, K', \hat\phi) \\
       &=& \opd{\alpha} \left[
@@
 shift parameter $K$ in the lower-tail distribution of $X$ may vary
-hugely from one equity to another. This could be understood also by
-computing $d\mathcal U F / dK$:
+hugely from one equity to another. This could be understood by observing
+\[
+\lim_{K \to 0} F(x) = \left\{
+\begin{array}{ll}
+  p \delta_0((-\infty, x])) & \text{ if } x \leq 0 \\
+  1 - (1 - p) \left(
+    K' \over K' + x
+  \right)^\beta & \text{ if } x > 0
+\end{array}
+\right.
+\]
 

### 2017-03-01 papers/FX/draft.tex
Accomodating the mistake pointed out by Thomas.

@@
 u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
-&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(-q) \label{eq:functional}
+&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
 \end{eqnarray}
@@
 \begin{eqnarray*}
-  q(r, \phi) &:=& -\ln \left( {
+  q(r, \phi) &:=& \ln \left( {
     \delta v - (1 - \phi) e^r
@@
     \phi
-} \right)
+} \right) = \ln\left(
+e^r + {\delta v - e^r \over \phi}
+\right)
 \end{eqnarray*}
@@
 \[
-X < -q(r, \phi)
-\]
+X < q(r, \phi)
+\]
+Naturally, if an agent invests in a risky asset instead of a riskless
+bond, he expects to obtain a higher return from the risky asset than
+he is guarranteed from the riskless bond. In our notation,
+this means $\delta v > e^r$, which implies $q(r, \phi) > 0$.
 The right-hand-side of \eqref{eq:functional} is a functional of the
@@
 = 
-\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(-q)
-\]
-In the rest of this section we write $q$ for $q(r, \phi)$
-for convenience. We have
+\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
+\]
+Hereafter we abbreviate $q(r, \phi)$ to just $q$. We have
 \begin{eqnarray}
@@
       }
-  {1 + b \1{x < -q} \over (K - x)^{\alpha + 1}} dx
+  {1 + b \over (K - x)^{\alpha + 1}} dx
   }_{J_1} \nonumber \\
@@
   }
-  {1 + b \1{x < -q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
+  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
   &&
-  \underbrace{
-    -b u(\delta v) p {K^\alpha \over (K + q)^\alpha} \1{q \geq 0}
-  }_{J_2}
   - b u(\delta v) \left[
@@
       (K' - q)^\beta
-    } \1{q < 0}
+    }
   \right]
@@
     \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
-    &=& \pd{J_1}{\alpha} + \pd{J_2}{\alpha} > 0
+    &>& 0
   \end{eqnarray*}
-  The 2nd term is easily computed:
+  We have
     \begin{eqnarray*}
-    \pd{J_2}{\alpha} &=& b u(\delta v) p \1{q \geq 0}
-    \left(
-      {K \over K + q}
-    \right)^\alpha
-    \ln\left(
-      1 + {q \over K}
-    \right) > 0
-  \end{eqnarray*}
-  As for the 1st term we have
-  \begin{enumerate}[i.]
-  \item When $q \geq 0$,
-    \begin{eqnarray*}
-      && \pd{J_1}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\
+      && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\
       &=&
       - \frac{\partial}{\partial \alpha} \left\{
-        {\alpha p K^\alpha \over \xi}
-        \left[
-          (1 + b)\int_{-\infty}^{-q} +
-          \int_{-q}^0
-        \right] \right. \\
-      && \left.
-        \left\{
+        {\alpha p K^\alpha \over \xi} (1 + b)
+        \int_{-\infty}^{0}
           \left[
@@
           (K - x)^{-\alpha - 1}
-        \right\}
         dx \right\} \\
@@
         {p \over \xi}
-        \left[
-          (1 + b)\int_{q}^{\infty} +
-          \int_{0}^q
-        \right]
+          (1 + b)\int_{0}^{\infty}
         \left[
@@
         {p \over \xi}
-        \left[
-          (1 + b)\int_{q}^{\infty} +
-          \int_{0}^q
-        \right]
+        (1 + b)\int_{0}^{\infty}
         \left[
@@
     \begin{eqnarray*}
-      L_1 &<& {
+      L_1 &<&
+      {
         \alpha K^\alpha p (1 + b)
@@
         \xi (1 - \hat\phi)^\xi e^{r \xi}
-      } \int_0^\infty
+      }
+      \int_0^\infty
       {1 \over (x + K)^{\alpha + 1}}
@@
     Let
-    \begin{eqnarray*}
+    \begin{eqnarray}
       c &:=& {p \over \xi}
-      \left[
-        (1 + b)\int_{q}^{\infty} +
-        \int_{0}^q
-      \right]
+        (1 + b)\int_{0}^{\infty}
       \left[
@@
         \alpha K^{\alpha} dx \over (K + x)^{\alpha + 2}
+      } \nonumber \\
+      &<&
+      {
+        \alpha K^\alpha p (1 + b)
+        \over
+        \xi (1 - \hat\phi)^\xi e^{r \xi}
+      }
+      \int_{0}^{\infty}
+      {
+        dx \over (K + x)^{\alpha + 2}
+      } \nonumber \\
+      &=& {
+        \alpha p (1 + b)
+        \over
+        K \xi (\alpha + 1) (1 - \phi)^\xi e^{r \xi}
       } < \infty
-    \end{eqnarray*}
+      \label{eq:c_lower_bound}
+    \end{eqnarray}
     Define distribution function $H(x): [0, \infty] \to [0, 1]$ by its
@@
     =
-    {p \over c \xi}
+    {p (1 + b)\over c \xi}
     \left[
@@
     }
-    \times
-    \left\{
-      \begin{array}{ll}
-        1 & \text{ if } 0 < x < q \\
-        (1 + b) & \text{ if } x > q \\
-      \end{array}
-    \right.
     \]
@@
     \begin{eqnarray*}
-      \E Y &:=& {p \over c \xi}
-      \left[
-        (1 + b)\int_{q}^{\infty} +
-        \int_{0}^q
-      \right]
+      \E Y &:=& {p (1 + b)\over c \xi}
+      \int_0^\infty
       \left[
@@
       L_2 &>&
-      {p \over \xi}
-      \left[
-        (1 + b)\int_{q}^{\infty} +
-        \int_{0}^q
-      \right]
+      {p (1 + b) \over \xi}
+      \int_{0}^{\infty}
       \left[
@@
       &=& {
-        p [(K + q)^{-\alpha} b + K^{-\alpha}] K^\alpha
-        \over
-        \alpha \xi [(1 - \hat\phi) e^r + \hat\phi]^\xi
+        p (1 + b)
+        \over
+        \alpha \xi ((1 - \phi) e^r + \phi)^\xi
       } 
     \end{eqnarray*}
-    Meanwhile
-    \begin{eqnarray*}
-      c &<& {p \over \xi}
-      \left[
-        (1 - \hat\phi) e^r
-      \right]^{-\xi}
-      \left[
-        (1 + b)\int_{q}^{\infty} +
-        \int_{0}^q
-      \right]
-      {
-        \alpha K^{\alpha} dx \over (K + x)^{\alpha + 2}
-      } \\
-      &=&
-      {
-        \alpha K^\alpha p \left[
-          (K + q)^{-\alpha - 1} b + K^{-\alpha - 1}
-        \right]
-        \over
-        \xi (\alpha + 1) (1 - \hat\phi)^\xi e^{r \xi}
-      }
-    \end{eqnarray*}
+    Meanwhile, as computed earlier in \eqref{eq:c_lower_bound}
+    \[
+      c < {
+        \alpha p (1 + b)
+        \over
+        K \xi (\alpha + 1) (1 - \phi)^\xi e^{r \xi}
+      }
+    \]
     Thus, to show inequality \eqref{eq:xxie5}, it is sufficient to show
@@
     \]
-
-  \item When $q < 0$, it follows from \eqref{eq:xxie1.0}:
-    \begin{eqnarray*}
-      && \opd{\alpha}J_1(\alpha, \beta, K, K', \hat\phi) \\
-      &=& \opd{\alpha} \left[
-        -{\alpha K^\alpha \over \xi} p
-        \int_{-\infty}^0
-            {
-              1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
-            }
-            {1 + b \over (K + x)^{\alpha + 1}} dx
-        \right]
-    \end{eqnarray*}
-  \end{enumerate}
-  So the same calculation as in the previous case carries through
-  with 0 substituted for $q$. The same conclusion of
-  \[
-  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
-  \]
-  stands.
 \end{proof}

### 2017-03-02 papers/FX/draft.tex
It is only the proof of convergence of mu((-inf, x]) = P(X <= x | X <= 0) -> Dirac[0] that is left to be proven.
K is a bigger problem!

@@
   && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
-  &=& \underbrace{
+  &=&
   -{\alpha K^\alpha \over \xi} p
@@
   {1 + b \over (K - x)^{\alpha + 1}} dx
-  }_{J_1} \nonumber \\
+  \nonumber \\
   && -{\beta K'^\beta \over \xi} (1 - p)
@@
       \over
-      (K' - q)^\beta
+      (K' + q)^\beta
     }
@@
 \end{eqnarray}
-
-\begin{theorem}
 Let
@@
 \]
-  Let the disappointment threshold $C < \delta v$ be equivalent to $X < -q$. Assume
-  \begin{equation}
-    \label{eq:xxie1.1}
-    K < {
-      \alpha e^{1/\alpha} (1 - e^{-r})^\xi
-      \over
-      (\alpha + 1) A
-    }
-  \end{equation}
-  where
-  \[
-  A = \max_{x \geq 1} {b x + x^{\alpha + 1} \over b + x^{\alpha + 1}}
-  \]
-  Then $G(\alpha, \beta, K, K')$ is monotonically increasing with $\alpha$.
+
+\begin{theorem} \label{thrm:I}
+  Let the disappointment threshold $C < \delta v$ be equivalent to $X
+  < q$. Assume $q > 0$. Then $G(\alpha, \beta, K, K')$ is
+  monotonically increasing with $\alpha$.
 \end{theorem}
@@
         dx \right\} \\
-      &=& \underbrace{
+      &=&
       {p \over \xi}
-          (1 + b)\int_{0}^{\infty}
+      (1 + b)
+      \int_{0}^{\infty}
       \left[
@@
       \right]^{-\xi}
-        \ln \left(
+      \left[
+        \alpha \ln \left(
+          1 + {x \over K}
+        \right) - 1
+      \right]
+      {K^{\alpha} dx
+        \over
+        (K + x)^{\alpha + 1}
+      } \\
+      &=&
+      - {p \over \xi}
+      (1 + b)
+      \int_{0}^{h}
+      \left[
+        (1 - \hat\phi) e^r + \hat\phi e^{-x}
+      \right]^{-\xi}
+      \left[
+        1 - \alpha \ln \left(
           1 + {x \over K}
         \right)
-        {\alpha K^{\alpha} dx
+      \right]
+      {K^{\alpha} dx
         \over
         (K + x)^{\alpha + 1}
-        }
-      }_{L_1} \\
-      && - \underbrace{
-        {p \over \xi}
-        (1 + b)\int_{0}^{\infty}
+      } \\
+      &&
+      + {p \over \xi}
+      (1 + b)
+      \int_{h}^{\infty}
       \left[
@@
       \right]^{-\xi}
-        {K^{\alpha} dx \over (K + x)^{\alpha + 1} }
-      }_{L_2}
-    \end{eqnarray*}
-    Clearly $L_2 < \infty$. For $L_1$, observe
-    \begin{eqnarray*}
-      L_1 &<&
-      {
-        \alpha K^\alpha p (1 + b)
-        \over
-        \xi (1 - \hat\phi)^\xi e^{r \xi}
-      }
-      \int_0^\infty
-      {1 \over (x + K)^{\alpha + 1}}
-      \left(1 + {x \over K}\right)
-      dx \\
-      &=&
-      {
-        p \alpha (1 + b)
-        \over
-        \xi (\alpha - 1) (1 - \hat\phi)^\xi e^{r \xi}
-      } < \infty
-    \end{eqnarray*}
-    Let
-    \begin{eqnarray}
-      c &:=& {p \over \xi}
-        (1 + b)\int_{0}^{\infty}
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-x}
-      \right]^{-\xi}
-      {
-        \alpha K^{\alpha} dx \over (K + x)^{\alpha + 2}
-      } \nonumber \\
-      &<&
-      {
-        \alpha K^\alpha p (1 + b)
-        \over
-        \xi (1 - \hat\phi)^\xi e^{r \xi}
-      }
-      \int_{0}^{\infty}
-      {
-        dx \over (K + x)^{\alpha + 2}
-      } \nonumber \\
-      &=& {
-        \alpha p (1 + b)
-        \over
-        K \xi (\alpha + 1) (1 - \phi)^\xi e^{r \xi}
-      } < \infty
-      \label{eq:c_lower_bound}
-    \end{eqnarray}
-    Define distribution function $H(x): [0, \infty] \to [0, 1]$ by its
-    density with respect to the Lebesgue measure
-    \[
-    {d H(x) \over d x}
-    =
-    {p (1 + b)\over c \xi}
-    \left[
-      (1 - \hat\phi) e^r + \hat\phi e^{-x}
-    \right]^{-\xi}
-    {
-      \alpha K^{\alpha}
-      \over
-      (K + x)^{\alpha + 2}
-    }
-    \]
-    Let $Y \sim H$, then
-    \begin{eqnarray*}
-      \E Y &:=& {p (1 + b)\over c \xi}
-      \int_0^\infty
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-x}
-      \right]^{-\xi} {
-        \alpha K^{\alpha} x dx
-        \over
-        (K + x)^{\alpha + 2}
-      } \\
-      &=& {\alpha L_2 \over c} - K \\
-      L_1 &=& c \E \left[
-        (Y + K) \ln \left(1 + {Y \over K} \right)
-      \right]
-    \end{eqnarray*}
-    Notice
-    \[
-    {
-      d^2
-      \over
-      dx^2
-    }
-    \left[
-      (x + K) \ln\left(1 + {x \over K}\right)
-    \right] = {1 \over x + K} > 0
-    \quad \text{ for } x > 0
-    \]
-    That is, $(x + K) \ln(1 + x/K)$ defined on $[0, \infty]$ is a convex
-    function. Now that $L_1, L_2 < \infty$, Jensen's inequality gives
-    \[
-    {L_1 \over c} = \E \left[
-      (Y + K) \ln \left(1 + {Y \over K} \right)
-    \right]
-    >
-    (\E Y + K) \times \ln(1 + \E Y/K)
-    =
-    {\alpha L_2  \over c}
-    \ln\left(
-      {\alpha L_2  \over c}
-    \right)
-    \]
-    Therefore
-    \[
-    L_1 - L_2 > L_2 \left[
+      \left[
         \alpha \ln \left(
-        {\alpha L_2 \over c}
+          1 + {x \over K}
         \right) - 1
       \right]
-    \]
-    Clearly $L_2 > 0$. So it remains to show
-    \[
-    \alpha \ln \left(
-      {\alpha L_2 \over c}
-    \right) - 1 > 0
-    \]
-    Or equivalently
-    \begin{equation}
-      \label{eq:xxie5}
-      {L_2 \over c} > {1 \over \alpha} \exp\left( {1 \over \alpha} \right)
-    \end{equation}
-    Observe
-    \begin{eqnarray*}
-      L_2 &>&
-      {p (1 + b) \over \xi}
-      \int_{0}^{\infty}
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi
-      \right]^{-\xi}
       {K^{\alpha} dx
@@
       } \\
-      &=& {
-        p (1 + b)
-        \over
-        \alpha \xi ((1 - \phi) e^r + \phi)^\xi
-      } 
+      &>&
+      - {p \over \xi}
+      (1 + b)
+      \left[
+        (1 - \hat\phi) e^r + \hat\phi e^{-h}
+      \right]^{-\xi}
+      \int_{0}^{h}
+      \left[
+        1 - \alpha \ln \left(
+          1 + {x \over K}
+        \right)
+      \right]
+      {K^{\alpha} dx
+        \over
+        (K + x)^{\alpha + 1}
+      } \\
+      &&
+      + {p \over \xi}
+      (1 + b)
+      \left[
+        (1 - \hat\phi) e^r + \hat\phi e^{-h}
+      \right]^{-\xi}
+      \int_{h}^{\infty}
+      \left[
+        \alpha \ln \left(
+          1 + {x \over K}
+        \right) - 1
+      \right]
+      {K^{\alpha} dx
+        \over
+        (K + x)^{\alpha + 1}
+      } \\
+      &=&
+      {p \over \xi}
+      (1 + b)
+      \left[
+        (1 - \hat\phi) e^r + \hat\phi e^{-h}
+      \right]^{-\xi}
+      \int_{0}^{\infty}
+      \left[
+        \alpha \ln \left(
+          1 + {x \over K}
+        \right) - 1
+      \right]
+      {K^{\alpha} dx
+        \over
+        (K + x)^{\alpha + 1}
+      } \\
+      &=& 0
     \end{eqnarray*}
-    Meanwhile, as computed earlier in \eqref{eq:c_lower_bound}
-    \[
-      c < {
-        \alpha p (1 + b)
-        \over
-        K \xi (\alpha + 1) (1 - \phi)^\xi e^{r \xi}
-      }
-    \]
-    Thus, to show inequality \eqref{eq:xxie5}, it is sufficient to show
-    \begin{equation}
-      \label{eq:xxie6}
-      {
-        \left[
-          b (K + q)^{-\alpha} + K^{-\alpha}
-        \right]
-        \left[
-          (1 - \hat\phi) e^r
-        \right]^{\xi}
-        (\alpha + 1)
-        \over
-        \alpha^2
-        \left[
-          b (K + q)^{-\alpha - 1} + K^{-\alpha - 1}
-        \right]
-        \left[
-          (1 - \hat\phi) e^r + \hat\phi
-        \right]^{\xi}
-      } > {1 \over \alpha} \exp\left( {1 \over \alpha} \right)
-    \end{equation}
-    This expression can be simplified to
-    \begin{eqnarray}
-      \left(
-           {\hat\phi \over 1 - \hat\phi} e^{-r} + 1
-      \right)^{-\xi} &>&
-      {
-        \alpha e^{1/\alpha} \left[
-          (K + q)^{-\alpha - 1} b + K^{-\alpha - 1}
-        \right]
-        \over
-        (\alpha + 1) \left[
-          (K + q)^{-\alpha} b + K^{-\alpha}
-        \right]
-      } \nonumber \\
-      {\hat\phi \over 1 - \hat\phi}e^{-r} + 1
-      &<&
-      \left[
-        {\alpha + 1 \over \alpha}
-        {
-          b K^{\alpha + 1} (K + q)
-          + (K + q)^{\alpha + 1} K 
-          \over
-          b K^{\alpha + 1} + (K + q)^{\alpha + 1}
-        } {
-          1
-          \over
-          e^{1/\alpha}
-        }
-     \right]^{1/\xi} \nonumber \\
-      &=&
-      \left[
-        {\alpha + 1 \over \alpha}
-        {
-          b \left(1 + {q \over K}\right)
-          + \left(1 + {q \over K}\right)^{\alpha + 1}
-          \over
-          b + \left(1 + {q \over K}\right)^{\alpha + 1}
-        } {
-          K
-          \over
-          e^{1/\alpha}
-        }
-      \right]^{1/\xi} \nonumber \\
-      \hat\phi &<& {1 - g \over 1 - g - e^{-r}}
-      \label{eq:xxie7}
-    \end{eqnarray}
     where
     \[
-    g = 
-    \left[
-      {\alpha + 1 \over \alpha}
-      {
-        b \left(1 + {q \over K}\right)
-        + \left(1 + {q \over K}\right)^{\alpha + 1}
-        \over
-        b + \left(1 + {q \over K}\right)^{\alpha + 1}
-      } {
-        K
-        \over
-        e^{1/\alpha}
-      }
-      \right]^{1/\xi}
-    \]
-    Inequality \eqref{eq:xxie7} is satisfied by all possible
-    $\hat\phi$ if
-    \begin{eqnarray*}
-      g &<& 1 - e^{-r}
-    \end{eqnarray*}
-    Observe that the expression
-    \[
-    {
-      b \left(1 + {q \over K}\right)
-      + \left(1 + {q \over K}\right)^{\alpha + 1}
-      \over
-      b + \left(1 + {q \over K}\right)^{\alpha + 1}
-    }
-    \]
-    is 1 and increasing at $q = 0$; as $q \to \infty$, it tends to
-    1. So it must have a maximum value for
-    $q \in (0, \infty)$. Let
-    \[
-    A = \max_{x \geq 1} {b x + x^{\alpha + 1} \over b + x^{\alpha + 1}}
-    \]
-    Then $g < 1 - e^{-r}$ is implied by
-    \begin{eqnarray*}
-      {\alpha + 1 \over \alpha}
-      {K \over e^{1/\alpha}}
-      A &<& (1 - e^{-r})^\xi \\
-      K &<& {
-        \alpha e^{1/\alpha} (1 - e^{-r})^\xi
-        \over
-        (\alpha + 1) A
-      }
-    \end{eqnarray*}
-    The last inequality is indeed assumption \eqref{eq:xxie1.1}.
-    So, under this assumption, \eqref{eq:xxie7} is satisfied by all
-    $\hat\phi$ and \eqref{eq:xxie6} holds. We conclude
-    \[
-    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
+    h = (e^{1/\alpha} - 1)K
     \]
   \end{proof}
-Since $u(v) + b u(\delta v)$ given in \eqref{eq:xxie1.0} as a
+Since $u(v)$ given in \eqref{eq:xxie1.0} as a
 functional of the distribution function of $X$ is increasing with the
@@
 shift parameter $K$ in the lower-tail distribution of $X$ may vary
-hugely from one equity to another. This could be understood by observing
-\[
-\lim_{K \to 0} F(x) = \left\{
-\begin{array}{ll}
-  p \delta_0((-\infty, x])) & \text{ if } x \leq 0 \\
-  1 - (1 - p) \left(
-    K' \over K' + x
-  \right)^\beta & \text{ if } x > 0
-\end{array}
-\right.
-\]
-
+hugely from one equity to another. This could be explained by
+\[
+\lim_{K \to 0} \td{G}{K}(\alpha, \beta, K, K', \phi) = 0
+\]
+When $K$ is very small, as is the case in real-world stock return
+distributions, a change of $K$ makes little difference to the utility
+functional, so its value can vary drastically.
+
+\begin{theorem}
+  Assume $K \leq 1$.
+  \[
+  \lim_{K \to 0} \td{G}{K}(\alpha, \beta, K, K') = 0
+  \]
+\end{theorem}
+\begin{proof}
+  By the same argument as in the proof of theorem \ref{thrm:I}, it
+  suffices to show
+  \begin{eqnarray*}
+    \lim_{K \to 0}
+    \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat \phi) &=& 0
+  \end{eqnarray*}
+  The left-hand-side is
+  \begin{eqnarray*}
+    && \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat \phi) \\
+    &=&
+    -{\alpha p (1 + b) \over \xi}
+    \int_{-\infty}^0
+    {
+      1 \over
+      \left[
+        (1 - \phi) e^r + \phi e^x
+      \right]^\xi
+    }
+    {
+      K^{\alpha - 1}
+      (\alpha x + K)
+      \over
+      (K - x)^{\alpha + 2}
+    } dx  \\
+    &=&
+    {\alpha p (1 + b) \over \xi}
+    \int_0^{\infty}
+    {
+      1 \over
+      \left[
+        (1 - \phi) e^r + \phi e^x
+      \right]^\xi
+    }
+    {
+      K^{\alpha - 1}
+      (\alpha x - K)
+      \over
+      (K + x)^{\alpha + 2}
+    } dx
+  \end{eqnarray*}
+  By reverse Fatou's lemma, it follows
+  \begin{eqnarray*}
+    &&
+    \lim_{K \to 0} \sup_{0 \leq M \leq K}
+    \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat \phi) \\
+    &\leq&
+    {\alpha p (1 + b) \over \xi}
+    \int_0^{\infty}
+    {
+      1 \over
+      \left[
+        (1 - \phi) e^r + \phi e^x
+      \right]^\xi
+    }
+    \lim_{K \to 0} \sup_{0 \leq M \leq K}
+    {
+      M^{\alpha - 1}
+      (\alpha x - M)
+      \over
+      (M + x)^{\alpha + 2}
+    } dx
+  \end{eqnarray*}
+\end{proof}
 \section{Review of multivariate GARCH models:

### 2017-03-03 papers/FX/draft.tex
First success of generalization.

@@
 
-\subsection{power-law utility function}
-In this section we follow Routledge and Zin
-\cite{routledge2010generalized} and assume 
-that the utility function of consumption $C$ is a power function%
+Routledge and Zin \cite{routledge2010generalized} assumed
+a power-law utility function of $C$:
 \begin{equation}
@@
 \end{equation}
-where $\gamma > 0$. For simplicity we assume the investor
+where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
+an increasing function. For simplicity we assume the investor
 initially has one unit of wealth. Let $r > 0$ be the return on the
@@
 \end{equation}
-
+For the distribution of $X$, call it $F(\cdot)$, we assume
 %% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
-
-\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
-In this section we consider power utility functions of
-\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
-let $\xi = \gamma - 1$. In this case, we assume the following
-distribution function $F(\cdot)$ of the equity return:
+% \subsection{power-law utility function}
+% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
+% In this section we consider power utility functions of
+% \eqref{eq:power_utility} with $\gamma > 1$. For convenience,
+% let $\xi = \gamma - 1$. In this case, we assume the following
+% distribution function $F(\cdot)$ of the equity return:
 \begin{equation}
@@
 Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
-Equation
-\eqref{eq:xxie0} can be re-arranged as
+Equation \eqref{eq:xxie0} can be re-arranged as
 \begin{eqnarray}
@@
   &=&
-  -{\alpha K^\alpha \over \xi} p
+  \alpha K^\alpha  p
   \int_{-\infty}^0
-      {
-       1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
-      }
+  u\left[ (1 - \phi) e^r + \phi e^x \right]
   {1 + b \over (K - x)^{\alpha + 1}} dx
   \nonumber \\
-  && -{\beta K'^\beta \over \xi} (1 - p)
+  &&
+  \beta K'^\beta (1 - p)
   \int_{0}^\infty
-  {
-   1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
-  }
+  u\left[ (1 - \phi) e^r + \phi e^x \right]
   {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
@@
   Let the disappointment threshold $C < \delta v$ be equivalent to $X
-  < q$. Assume $q > 0$. Then $G(\alpha, \beta, K, K')$ is
-  monotonically increasing with $\alpha$.
+  < q$. Assume $q > 0$. Also assume $u(\cdot)$ is an increasing function.
+  Then
+  \[
+  \td{G}{K} < 0 < \td{G}{\alpha}
+  \]
 \end{theorem}
@@
 \begin{proof}
-  We show
-  \[
-  \td{G}{\alpha} > 0
-  \]
   Let
@@
     \]
+    As for $\td{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
+    $\td{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
+    \[
+    \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
+    \]
+    We have
+    \begin{eqnarray*}
+      && \pd{\mathcal UF}{K}(\alpha, \beta, K, K', \hat\phi) \\
+      &=&
+      -{\alpha p (1+ b) \over \xi}
+      \int_0^\infty
+      \left[
+        (1 - \hat\phi) e^r + \hat\phi e^{-x}
+      \right]^{-\xi}
+      {
+        K^{\alpha - 1} (\alpha x - K)
+        \over
+        (K + x)^{\alpha + 2}
+      }
+      dx \\
+      &=&
+      -{\alpha p (1+ b) \over \xi}
+      \left\{
+        - \int_{0}^{K/\alpha}
+        \left[
+          (1 - \hat\phi) e^r + \hat\phi e^{-x}
+        \right]^{-\xi}
+        {
+          K^{\alpha - 1} (K - \alpha x)
+          \over
+          (K + x)^{\alpha + 2}
+        }
+        dx \right. \\
+      &&
+      \left.
+        +
+        \int_{K/\alpha}^{\infty}
+        \left[
+          (1 - \hat\phi) e^r + \hat\phi e^{-x}
+        \right]^{-\xi}
+        {
+          K^{\alpha - 1} (\alpha x - K)
+          \over
+          (K + x)^{\alpha + 2}
+        }
+        dx
+      \right\} \\
+      &<&
+      -{\alpha p (1+ b) \over \xi}      
+      \left[
+        (1 - \hat\phi) e^r + \hat\phi e^{-M/\alpha}
+      \right]^{-\xi} \times \\
+      &&
+      \left\{
+        \int_{K/\alpha}^{\infty}
+        {
+          K^{\alpha - 1} (\alpha x - K)
+          \over
+          (K + x)^{\alpha + 2}
+        }
+        dx
+        - \int_{0}^{K/\alpha}
+        {
+          K^{\alpha - 1} (K - \alpha x)
+          \over
+          (K + x)^{\alpha + 2}
+        }
+        dx
+      \right\} \\
+      &=& 0
+    \end{eqnarray*}
   \end{proof}
@@
 functional of the distribution function of $X$ is increasing with the
-lower tail index $\alpha$, all agents guided by this class of utility
-functions will seek the largest $\alpha$ found in a stock group e.g.
-a sector. As a result, one can expect all the actively traded equities
-in the group to feature the same $\alpha$. In section
-\ref{sec:data_analysis} we analyse real-world data of stock returns to
-verify this conclusion.
-
-In comparison to the theoretical constancy of $\alpha$, the scale and
-shift parameter $K$ in the lower-tail distribution of $X$ may vary
-hugely from one equity to another. This could be explained by
-\[
-\lim_{K \to 0} \td{G}{K}(\alpha, \beta, K, K', \phi) = 0
-\]
-When $K$ is very small, as is the case in real-world stock return
-distributions, a change of $K$ makes little difference to the utility
-functional, so its value can vary drastically.
-
-\begin{theorem}
-  Assume $K \leq 1$.
-  \[
-  \lim_{K \to 0} \td{G}{K}(\alpha, \beta, K, K') = 0
-  \]
-\end{theorem}
-\begin{proof}
-  By the same argument as in the proof of theorem \ref{thrm:I}, it
-  suffices to show
-  \begin{eqnarray*}
-    \lim_{K \to 0}
-    \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat \phi) &=& 0
-  \end{eqnarray*}
-  The left-hand-side is
-  \begin{eqnarray*}
-    && \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat \phi) \\
-    &=&
-    -{\alpha p (1 + b) \over \xi}
-    \int_{-\infty}^0
-    {
-      1 \over
-      \left[
-        (1 - \phi) e^r + \phi e^x
-      \right]^\xi
-    }
-    {
-      K^{\alpha - 1}
-      (\alpha x + K)
-      \over
-      (K - x)^{\alpha + 2}
-    } dx  \\
-    &=&
-    {\alpha p (1 + b) \over \xi}
-    \int_0^{\infty}
-    {
-      1 \over
-      \left[
-        (1 - \phi) e^r + \phi e^x
-      \right]^\xi
-    }
-    {
-      K^{\alpha - 1}
-      (\alpha x - K)
-      \over
-      (K + x)^{\alpha + 2}
-    } dx
-  \end{eqnarray*}
-  By reverse Fatou's lemma, it follows
-  \begin{eqnarray*}
-    &&
-    \lim_{K \to 0} \sup_{0 \leq M \leq K}
-    \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat \phi) \\
-    &\leq&
-    {\alpha p (1 + b) \over \xi}
-    \int_0^{\infty}
-    {
-      1 \over
-      \left[
-        (1 - \phi) e^r + \phi e^x
-      \right]^\xi
-    }
-    \lim_{K \to 0} \sup_{0 \leq M \leq K}
-    {
-      M^{\alpha - 1}
-      (\alpha x - M)
-      \over
-      (M + x)^{\alpha + 2}
-    } dx
-  \end{eqnarray*}
-\end{proof}
+lower tail index $\alpha$ and decreasing with $K$, there is a curve of
+equal preference on the plane of $(\alpha, K)$. Moving along this
+curve in the direction of increasing $\alpha$, one expects the values
+of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
+should appear positively correlated. This is indeed the case in some
+real stock return data, e.g. the ``Energy'' and ``Information Technology''
+sectors of S\&P 500 companies. This is illustrated in figures
+\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
+
+% all agents guided by this class of utility
+% functions will seek the largest $\alpha$ and the smallest $K$ found in
+% a stock group e.g. a sector.
+
+% As a result, one can expect all the
+% actively traded equities in the group to feature the same $\alpha$ and
+% $K$. However, the estimated value of $K^\alpha$ may vary
+% significantly. This can be understood as a result of the smallness of
+% $K$, which as revealed in section \ref{sec:hybrid_estimation} is of the
+% order of $10^{-2}$, and the inaccuracy of the estimated value of
+% $\alpha$, which typically ranges from 2.5 to 5. Exponentiation hugely
+% magnifies the inaccuracy in $\alpha$.
+
 \section{Review of multivariate GARCH models:
@@
 \section{Hill Estimator}
-%% TODO: Show Hill estimates of real and simulated data with confidence
-%% band. Argue that all estimated tail indices are indistinguishable
-%% given the confidence band.
 There have been several estimators for the tail index proposed in the
@@
 
-\subsubsection{The Energy and the Consumer Staples sectors of S\&P 500}
-Figure \ref{fig:Energy_Hill_upper},
-\ref{fig:Consumer_Staples_Hill_upper},
+%% \subsubsection{The Energy and the Consumer Staples sectors of S\&P 500}
+Figure \ref{fig:Energy_lower},
+\ref{fig:Consumer_Staples_lower},
 and \ref{fig:Information_Technology_lower}
@@
   \centering
-  \includegraphics[width=\textwidth]{Energy_Hill_upper.pdf}
-  \caption{Hill estimates of the upper tail indices of 32 stocks'
+  \includegraphics[width=\textwidth]{Energy_lower.pdf}
+  \caption{Hill estimates of the lower tail indices of 32 stocks'
     daily return series in the {\it Energy} sector of Standard \& Poor
@@
     from 2010-01-01 to 2015-01-01 and comprise $n=1304$
-    observations. $k=(\ln n)^2 = 49$ upper order statistics are used
+    observations. $k=(\ln n)^2 = 49$ lower order statistics are used
     for computing $\alpha_n$. 
   }
-  \label{fig:Energy_Hill_upper}
+  \label{fig:Energy_lower}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{Consumer_Staples_Hill_upper.pdf}
-  \caption{Hill estimates of the upper tail indices of 30 stocks'
+  \includegraphics[width=\textwidth]{Consumer_Staples_lower.pdf}
+  \caption{Hill estimates of the lower tail indices of 30 stocks'
     daily return series in the {\it Consumer Staples} sector of
@@
   }
-  \label{fig:Consumer_Staples_Hill_upper}
+  \label{fig:Consumer_Staples_lower}
 \end{figure}
@@
 \section{Hybrid estimation of the assumed distribution function}
+\label{sec:hybrid_estimation}
 We adopt a hybrid procedure to estimate the parameters of the
-distribution function assumed in eq.\eqref{eq:X_distr}: First we
-obtain a Hill estimate of the lower- and upper-tail indices of an
-equity return and fix $\alpha$ and $\beta$ in \eqref{eq:X_distr}
-to their Hill estimates. Then we estimate $K$ and $K'$ by maximizing
-the likelihood function with fixed $\alpha$ and $\beta$.
+distribution function assumed in \eqref{eq:X_distr}: First we
+obtain a Hill estimate of the lower-tail indices of an
+equity return and fix $\alpha$ in \eqref{eq:X_distr}
+to its Hill estimate. Then we estimate $K$ by maximizing
+the likelihood function with fixed $\alpha$. Obviously we take only
+negative observations and hence condition on $X < 0$.
+
+Figure \ref{fig:Energy_alpha_K},
+\ref{fig:Consumer_Staples_alpha_K},
+and \ref{fig:Information_Technology_alpha_K}
+show the estimates of $K$ alongside the Hill estimates of $\alpha$ of
+the ``energy'', ``consumer staples'' and ``information
+technology'' sectors of the S\&P 500 index, respectively.
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\textwidth]{Energy_alpha_K.pdf}
+  \caption{Hybrid estimates of the lower-tail parameters of daily
+    return series in the {\it Energy} sector of S\&P
+    500 index.
+  }
+  \label{fig:Energy_alpha_K}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\textwidth]{Consumer_Staples_alpha_K.pdf}
+  \caption{Hybrid estimates of the lower-tail parameters of daily
+    return series in the {\it Consumer Staples} sector of S\&P
+    500 index.
+  }
+  \label{fig:Consumer_Staples_alpha_K}
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\textwidth]{Information_Technology_alpha_K.pdf}
+  \caption{Hybrid estimates of the lower-tail parameters of daily
+    return series in the {\it Information Technology} sector of S\&P
+    500 index.
+  }
+  \label{fig:Information_Technology_alpha_K}
+\end{figure}
 
@@
 \]
+Using Pickands' estimator, the same result of very similar tail indices
+is again obtained for the {\it Energy} and the {\it Consumer Staples}
+sectors. Figure \ref{fig:Energy_Pickands} and
+\ref{fig:Consumer_Staples_Pickands} illustrate this claim.
+\begin{figure}[htb!]
+  \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
+  \caption{Pickands' estimates of {\it Energy} stocks.}
+  \label{fig:Energy_Pickands}
+\end{figure}
+
+\begin{figure}[htb!]
+  \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
+  \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
+  \label{fig:Consumer_Staples_Pickands}
+\end{figure}
 
@@
 %% null hypothesis must be lower than $\alpha$.
-\textcolor{red}{Review the test of Hoga}
-
-\section{Data Analysis} \label{sec:data_analysis}
-\subsection{Tail index and scale parameter}
-In this section we apply the statistical tools reviewed in the
-previous sections to real stock return series, denoted $X_t$
-hereafter. We are interested in finding out the tail indices and the
-scale parameters of the stationary distribution of $|X_t|$.
+%% \textcolor{red}{Review the test of Hoga}
+
+% \section{Data Analysis} \label{sec:data_analysis}
+% \subsection{Tail index and scale parameter}
+% In this section we apply the statistical tools reviewed in the
+% previous sections to real stock return series, denoted $X_t$
+% hereafter. We are interested in finding out the tail indices and the
+% scale parameters of the stationary distribution of $|X_t|$.
 
@@
 
-Using Pickands' estimator, the same result of very similar tail indices
-is again obtained for the {\it Energy} and the {\it Consumer Staples}
-sectors. Figure \ref{fig:Energy_Pickands} and
-\ref{fig:Consumer_Staples_Pickands} illustrate this claim.
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
-  \caption{Pickands' estimates of {\it Energy} stocks.}
-  \label{fig:Energy_Pickands}
-\end{figure}
-
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
-  \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
-  \label{fig:Consumer_Staples_Pickands}
-\end{figure}
-
 %% \subsubsection{The Information Technology sector of S\&P 500}

### 2017-03-05 papers/FX/draft.tex
All that can be said in the univariate case has probably been said.

@@
 \end{eqnarray}
+In theorem \ref{thrm:I} we show
+\[
+\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
+\]
+is monotonically increasing with $\alpha$ and monotonically decreasing
+with $K$. But before that it is worthwhile to state the following
+simple observation:
+\begin{lemma} \label{lemma:I}
+  Assume distribution function $F(x; \theta)$ parameterized by
+  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
+  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
+  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
+  $(a, b)$ and $X \sim F$.
+  \begin{enumerate}
+  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, 
+    \[
+    \pd{\E h(X)}{\theta} > 0
+    \]
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$,
+    \[
+    \pd{\E h(X)}{\theta} > 0
+    \]
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$
+    \[
+    \pd{\E h(X)}{\theta} < 0
+    \]
+  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
+    \[
+    \pd{\E h(X)}{\theta} < 0
+    \]
+  \end{enumerate}
+\end{lemma}
+\begin{proof}
+  Clearly
+  \begin{eqnarray*}
+    && \pd{\E h(X)}{\theta} \\
+    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
+    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
+    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
+  \end{eqnarray*}
+  \begin{enumerate}
+  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
+    \theta) > 0$ on $(a, x_0)$
+    \[
+    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
+    \]
+    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
+    $h(x)$ is decreasing, we have
+    \begin{eqnarray*}
+      I_2 &=& \int_{x_0}^b -h(x_0)
+      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
+      &>& -h(x_0)
+      \int_{x_0}^b \left| 
+        \pd{f}{\theta}(x, \theta)
+      \right| dx
+    \end{eqnarray*}
+    Finally we have
+    \begin{eqnarray*}
+      && \pd{\E h(X)}{\theta} \\
+      &=& I_1 + I_2 \\
+      &>& h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx \\
+      &=& h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx \\
+      &=& h(x_0) \opd{\theta} 1 = 0
+    \end{eqnarray*}
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
+    \begin{eqnarray*}
+      I_1 &=&
+      \int_a^{x_0} -h(x)
+      \left| \pd{f}{\theta}(x, \theta) \right| dx \\
+      &>&
+      -h(x_0) \int_a^{x_0}
+      \left| \pd{f}{\theta}(x, \theta) \right| dx
+    \end{eqnarray*}
+    and obviously
+    \[
+    I_2 > h(x_0) \int_{x_0}^b
+    \pd{f}{\theta}(x, \theta) dx
+    \]
+    So
+    \[
+    \pd{\E h(X)}{\theta}
+    = I_1 + I_2
+    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
+    \]
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$
+
+    Obviously
+    \[
+    I_1 = \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx
+    < h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
+    \]
+    and
+    \begin{eqnarray*}
+      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
+      &<& -h(x_0) \int_{x_0}^b \left| \pd{f}{\theta}(x, \theta) \right| dx \\
+    \end{eqnarray*}
+    Thus
+    \[
+    \pd{\E h(X)}{\theta}
+    = I_1 + I_2
+    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
+    \]
+  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
+
+    In this case
+    \begin{eqnarray*}
+      I_1 &=& \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx \\
+      &<& -h(x_0) \int_a^{x_0} \left| \pd{f}{\theta}(x, \theta) \right| dx
+    \end{eqnarray*}
+    And
+    \begin{eqnarray*}
+      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
+      &<& h(x_0) \int_{x_0}^b \pd{f}{\theta}(x, \theta) dx
+    \end{eqnarray*}
+    So
+    \[
+    \pd{\E h(X)}{\theta}
+    = I_1 + I_2
+    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
+    \]
+  \end{enumerate}
+\end{proof}
+With lemma \ref{lemma:I} it is straightforward to show the
+monotonicity claimed earlier.
 Let
@@
 \]
-
-\begin{theorem} \label{thrm:I}
-  Let the disappointment threshold $C < \delta v$ be equivalent to $X
-  < q$. Assume $q > 0$. Also assume $u(\cdot)$ is an increasing function.
-  Then
-  \[
-  \td{G}{K} < 0 < \td{G}{\alpha}
+\begin{theorem}
+  \label{thrm:I}
+  \[
+  \pd{G}{\alpha} (\alpha, \beta, K, K') > 0,
+  \quad
+  \pd{G}{K} (\alpha, \beta, K, K') < 0
   \]
@@
   \begin{eqnarray}
-    \td{G}{\alpha}
+    \pd{G}{\alpha}
     &=&
@@
   \end{eqnarray*}
-  We have
-    \begin{eqnarray*}
-      && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\
+  From \eqref{eq:xxie1.0}, it follows
+  \begin{eqnarray*}
+    && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
+    \\
+    &=& p \opd{\alpha}\E[u((1 - \phi) e^r + \phi e^X) | X < 0]
+  \end{eqnarray*}
+  The function $u((1 - \phi) e^r + \phi e^X)$ is obviously increasing
+  with $X$. Furthermore
+  \begin{eqnarray*}
+    && \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
     &=&
-      - \frac{\partial}{\partial \alpha} \left\{
-        {\alpha p K^\alpha \over \xi} (1 + b)
-        \int_{-\infty}^{0}
+    - {K^\alpha \over (K - x)^{\alpha + 1}}
     \left[
-            (1 - \hat\phi) e^r + \hat\phi e^x
-          \right]^{-\xi}
-          (K - x)^{-\alpha - 1}
-        dx \right\} \\
-      &=&
-      {p \over \xi}
-      (1 + b)
-      \int_{0}^{\infty}
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-x}
-      \right]^{-\xi}
-      \left[
-        \alpha \ln \left(
-          1 + {x \over K}
+      \alpha
+      \ln\left(
+        1 - {x \over K}
       \right) - 1
     \right]
-      {K^{\alpha} dx
+  \end{eqnarray*}
+  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
+  \[
+  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  \]
+  and when $K(1 - e^{1/\alpha}) < x < 0$,
+  \[
+  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  \]
+  This is the second case of lemma \ref{lemma:I}. So we have
+  \begin{eqnarray*}
+    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
+  \end{eqnarray*}
+  As for $\pd{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
+  $\pd{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
+  \[
+  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
+  \]
+  We have
+  \[
+  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi)
+  = p \opd{K} \E \left[
+    u((1 - \phi) e^r + \phi e^X)
+    | X < 0
+  \right] 
+  \]
+  and
+  \begin{eqnarray*}
+    && \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+    &=&
+    -\alpha K^{\alpha - 1} {
+      \alpha x + K
       \over
-        (K + x)^{\alpha + 1}
-      } \\
-      &=&
-      - {p \over \xi}
-      (1 + b)
-      \int_{0}^{h}
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-x}
-      \right]^{-\xi}
-      \left[
-        1 - \alpha \ln \left(
-          1 + {x \over K}
-        \right)
-      \right]
-      {K^{\alpha} dx
-        \over
-        (K + x)^{\alpha + 1}
-      } \\
-      &&
-      + {p \over \xi}
-      (1 + b)
-      \int_{h}^{\infty}
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-x}
-      \right]^{-\xi}
-      \left[
-        \alpha \ln \left(
-          1 + {x \over K}
-        \right) - 1
-      \right]
-      {K^{\alpha} dx
-        \over
-        (K + x)^{\alpha + 1}
-      } \\
-      &>&
-      - {p \over \xi}
-      (1 + b)
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-h}
-      \right]^{-\xi}
-      \int_{0}^{h}
-      \left[
-        1 - \alpha \ln \left(
-          1 + {x \over K}
-        \right)
-      \right]
-      {K^{\alpha} dx
-        \over
-        (K + x)^{\alpha + 1}
-      } \\
-      &&
-      + {p \over \xi}
-      (1 + b)
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-h}
-      \right]^{-\xi}
-      \int_{h}^{\infty}
-      \left[
-        \alpha \ln \left(
-          1 + {x \over K}
-        \right) - 1
-      \right]
-      {K^{\alpha} dx
-        \over
-        (K + x)^{\alpha + 1}
-      } \\
-      &=&
-      {p \over \xi}
-      (1 + b)
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-h}
-      \right]^{-\xi}
-      \int_{0}^{\infty}
-      \left[
-        \alpha \ln \left(
-          1 + {x \over K}
-        \right) - 1
-      \right]
-      {K^{\alpha} dx
-        \over
-        (K + x)^{\alpha + 1}
-      } \\
-      &=& 0
+      (K - x)^{\alpha + 2}
+    }
   \end{eqnarray*}
-    where
-    \[
-    h = (e^{1/\alpha} - 1)K
-    \]
-    As for $\td{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
-    $\td{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
+  Clearly, when $x < -\alpha/K$
+  \[
+  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  \]
+  and when $-\alpha/K < x < 0$
+  \[
+  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  \]
+  So by the third case of lemma \ref{lemma:I} we conclude
   \[
@@
   \]
-    We have
-    \begin{eqnarray*}
-      && \pd{\mathcal UF}{K}(\alpha, \beta, K, K', \hat\phi) \\
-      &=&
-      -{\alpha p (1+ b) \over \xi}
-      \int_0^\infty
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-x}
-      \right]^{-\xi}
-      {
-        K^{\alpha - 1} (\alpha x - K)
-        \over
-        (K + x)^{\alpha + 2}
-      }
-      dx \\
-      &=&
-      -{\alpha p (1+ b) \over \xi}
-      \left\{
-        - \int_{0}^{K/\alpha}
-        \left[
-          (1 - \hat\phi) e^r + \hat\phi e^{-x}
-        \right]^{-\xi}
-        {
-          K^{\alpha - 1} (K - \alpha x)
-          \over
-          (K + x)^{\alpha + 2}
-        }
-        dx \right. \\
-      &&
-      \left.
-        +
-        \int_{K/\alpha}^{\infty}
-        \left[
-          (1 - \hat\phi) e^r + \hat\phi e^{-x}
-        \right]^{-\xi}
-        {
-          K^{\alpha - 1} (\alpha x - K)
-          \over
-          (K + x)^{\alpha + 2}
-        }
-        dx
-      \right\} \\
-      &<&
-      -{\alpha p (1+ b) \over \xi}      
-      \left[
-        (1 - \hat\phi) e^r + \hat\phi e^{-M/\alpha}
-      \right]^{-\xi} \times \\
-      &&
-      \left\{
-        \int_{K/\alpha}^{\infty}
-        {
-          K^{\alpha - 1} (\alpha x - K)
-          \over
-          (K + x)^{\alpha + 2}
-        }
-        dx
-        - \int_{0}^{K/\alpha}
-        {
-          K^{\alpha - 1} (K - \alpha x)
-          \over
-          (K + x)^{\alpha + 2}
-        }
-        dx
-      \right\} \\
-      &=& 0
-    \end{eqnarray*}
 \end{proof}
@@
 \ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
-
+The confidence bands of the Hill estimates of $\alpha$ of these
+sectors are shown in figures \ref{fig:Energy_lower},
+\ref{fig:Consumer_Staples_lower} and \ref{fig:Information_Technology_lower}.
 % all agents guided by this class of utility
@@
 \section{Hill Estimator}
+\label{sec:Hill}
 There have been several estimators for the tail index proposed in the

### 2017-03-06 papers/FX/draft.tex
A revision to the lemma. Applied the lemma to t-distribution.
Added the section about t-distributed equity return.

@@
 where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
-an increasing function. For simplicity we assume the investor
-initially has one unit of wealth. Let $r > 0$ be the return on the
-risk free state bond and $X$ be the return on equity. Portfolio weights
-are $\left(1-\phi\right)$ and $\phi$ respectively. Then
+increasing and concave. Before proceeding any further, it is worth
+stating the following simple observation:
+\begin{lemma} \label{lemma:I}
+  Assume distribution function $F(x; \theta)$ parameterized by
+  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
+  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
+  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
+  $(a, b)$, $X \sim F$, $\E h(X) < \infty$ and $\forall x \in (a, b)$
+  \[
+  \left| \pd{f(x, \theta)}{\theta} \right| < \infty
+  \]
+  \begin{enumerate}
+  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
+    \[
+    \pd{\E h(X)}{\theta} > 0
+    \]
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
+    \[
+    \pd{\E h(X)}{\theta} > 0
+    \]
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
+    \[
+    \pd{\E h(X)}{\theta} < 0
+    \]
+  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
+    \[
+    \pd{\E h(X)}{\theta} < 0
+    \]
+  \end{enumerate}
+\end{lemma}
+\begin{proof}
+  Clearly
+  \begin{eqnarray*}
+    && \pd{\E h(X)}{\theta} \\
+    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
+    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
+    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
+  \end{eqnarray*}
+  \begin{enumerate}
+  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
+    \theta) > 0$ on $(a, x_0)$
+    \[
+    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
+    \]
+    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
+    $h(x)$ is decreasing, we have
+    \begin{eqnarray*}
+      I_2 &=& \int_{x_0}^b -h(x_0)
+      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
+      &>& -h(x_0)
+      \int_{x_0}^b \left| 
+        \pd{f}{\theta}(x, \theta)
+      \right| dx
+    \end{eqnarray*}
+    Finally we have
+    \begin{eqnarray*}
+      && \pd{\E h(X)}{\theta} \\
+      &=& I_1 + I_2 \\
+      &>& h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx \\
+      &=& h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx \\
+      &=& h(x_0) \opd{\theta} 1 = 0
+    \end{eqnarray*}
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
+    \begin{eqnarray*}
+      I_1 &=&
+      \int_a^{x_0} -h(x)
+      \left| \pd{f}{\theta}(x, \theta) \right| dx \\
+      &>&
+      -h(x_0) \int_a^{x_0}
+      \left| \pd{f}{\theta}(x, \theta) \right| dx
+    \end{eqnarray*}
+    and obviously
+    \[
+    I_2 > h(x_0) \int_{x_0}^b
+    \pd{f}{\theta}(x, \theta) dx
+    \]
+    So
+    \[
+    \pd{\E h(X)}{\theta}
+    = I_1 + I_2
+    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
+    \]
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$
+
+    Obviously
+    \[
+    I_1 = \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx
+    < h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
+    \]
+    and
+    \begin{eqnarray*}
+      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
+      &<& -h(x_0) \int_{x_0}^b \left| \pd{f}{\theta}(x, \theta) \right| dx \\
+    \end{eqnarray*}
+    Thus
+    \[
+    \pd{\E h(X)}{\theta}
+    = I_1 + I_2
+    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
+    \]
+  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
+
+    In this case
+    \begin{eqnarray*}
+      I_1 &=& \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx \\
+      &<& -h(x_0) \int_a^{x_0} \left| \pd{f}{\theta}(x, \theta) \right| dx
+    \end{eqnarray*}
+    And
+    \begin{eqnarray*}
+      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
+      &<& h(x_0) \int_{x_0}^b \pd{f}{\theta}(x, \theta) dx
+    \end{eqnarray*}
+    So
+    \[
+    \pd{\E h(X)}{\theta}
+    = I_1 + I_2
+    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
+    \]
+  \end{enumerate}
+\end{proof}
+
+Now for simplicity we assume the investor initially has one unit of
+wealth. Let $r > 0$ be the return on the risk free state bond and $X$
+be the return on equity. Portfolio weights are $\left(1-\phi\right)$
+and $\phi$ respectively. Then
 \begin{equation}
   \label{eq:xxie1}
-  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
-  %% 0, 1 + (1-\phi) r  +\phi X
-  %% \right\}^{1-\gamma} \label{3}%
   C = (1 - \phi) e^r + \phi e^X
 \end{equation}
-For the distribution of $X$, call it $F(\cdot)$, we assume
-%% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
-% \subsection{power-law utility function}
-% \subsubsection[in case gamma > 1]{in case $\gamma > 1$}
-% In this section we consider power utility functions of
-% \eqref{eq:power_utility} with $\gamma > 1$. For convenience,
-% let $\xi = \gamma - 1$. In this case, we assume the following
-% distribution function $F(\cdot)$ of the equity return:
+Equation \eqref{eq:xxie0} can be re-arranged as
+\begin{eqnarray}
+u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
+&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
+\end{eqnarray}
+where
+\begin{eqnarray*}
+  q(r, \phi) &:=& \ln \left( {
+    \delta v - (1 - \phi) e^r
+    \over
+    \phi
+} \right) = \ln\left(
+e^r + {\delta v - e^r \over \phi}
+\right)
+\end{eqnarray*}
+Note $C < \delta v$ implies
+\[
+X < q(r, \phi)
+\]
+Hereafter we abbreviate $q(r, \phi)$ to just $q$. 
+Naturally, if an agent invests in a risky asset instead of a riskless
+bond, he expects to obtain a higher return from the risky asset than
+he is guarranteed from the riskless bond. In our notation,
+this means $\delta v > e^r$, which implies $q(r, \phi) > r > 0$.
+The right-hand-side of \eqref{eq:functional} is a functional of the
+distribution function of $X$ via \eqref{eq:xxie1}. For convenience
+we denote it $\mathcal U(\cdot)$, i.e.
+\[
+\mathcal U F (\alpha, \beta, K, K', \phi)
+= 
+\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
+\]
+In the rest of this section, we use $F(\cdot)$ to denote the
+distribution function of $X$ and use $f(\cdot)$ to denote its density
+function.
+
+\subsection{Pareto-distributed equity return}
+In this section, we look into the situation when $X$ has a shifted
+Pareto distribution. Assume
 \begin{equation}
@@
 Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
-Equation \eqref{eq:xxie0} can be re-arranged as
-\begin{eqnarray}
-u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
-&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
-\end{eqnarray}
-where
-\begin{eqnarray*}
-  q(r, \phi) &:=& \ln \left( {
-    \delta v - (1 - \phi) e^r
-    \over
-    \phi
-} \right) = \ln\left(
-e^r + {\delta v - e^r \over \phi}
-\right)
-\end{eqnarray*}
-Note $C < \delta v$ implies
-\[
-X < q(r, \phi)
-\]
-Naturally, if an agent invests in a risky asset instead of a riskless
-bond, he expects to obtain a higher return from the risky asset than
-he is guarranteed from the riskless bond. In our notation,
-this means $\delta v > e^r$, which implies $q(r, \phi) > 0$.
-The right-hand-side of \eqref{eq:functional} is a functional of the
-distribution function of $X$ via \eqref{eq:xxie1}. For convenience
-we denote it $\mathcal U(\cdot)$, i.e.
-\[
-\mathcal U F (\alpha, \beta, K, K', \phi)
-= 
-\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
-\]
-Hereafter we abbreviate $q(r, \phi)$ to just $q$. We have
+We have
 \begin{eqnarray}
@@
 \end{eqnarray}
-In theorem \ref{thrm:I} we show
-\[
-\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
-\]
-is monotonically increasing with $\alpha$ and monotonically decreasing
-with $K$. But before that it is worthwhile to state the following
-simple observation:
-\begin{lemma} \label{lemma:I}
-  Assume distribution function $F(x; \theta)$ parameterized by
-  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
-  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
-  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
-  $(a, b)$ and $X \sim F$.
-  \begin{enumerate}
-  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, 
-    \[
-    \pd{\E h(X)}{\theta} > 0
-    \]
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$,
-    \[
-    \pd{\E h(X)}{\theta} > 0
-    \]
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$
-    \[
-    \pd{\E h(X)}{\theta} < 0
-    \]
-  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
-    \[
-    \pd{\E h(X)}{\theta} < 0
-    \]
-  \end{enumerate}
-\end{lemma}
-\begin{proof}
-  Clearly
-  \begin{eqnarray*}
-    && \pd{\E h(X)}{\theta} \\
-    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
-    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
-    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
-  \end{eqnarray*}
-  \begin{enumerate}
-  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
-    \theta) > 0$ on $(a, x_0)$
-    \[
-    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
-    \]
-    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
-    $h(x)$ is decreasing, we have
-    \begin{eqnarray*}
-      I_2 &=& \int_{x_0}^b -h(x_0)
-      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
-      &>& -h(x_0)
-      \int_{x_0}^b \left| 
-        \pd{f}{\theta}(x, \theta)
-      \right| dx
-    \end{eqnarray*}
-    Finally we have
-    \begin{eqnarray*}
-      && \pd{\E h(X)}{\theta} \\
-      &=& I_1 + I_2 \\
-      &>& h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx \\
-      &=& h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx \\
-      &=& h(x_0) \opd{\theta} 1 = 0
-    \end{eqnarray*}
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
-    \begin{eqnarray*}
-      I_1 &=&
-      \int_a^{x_0} -h(x)
-      \left| \pd{f}{\theta}(x, \theta) \right| dx \\
-      &>&
-      -h(x_0) \int_a^{x_0}
-      \left| \pd{f}{\theta}(x, \theta) \right| dx
-    \end{eqnarray*}
-    and obviously
-    \[
-    I_2 > h(x_0) \int_{x_0}^b
-    \pd{f}{\theta}(x, \theta) dx
-    \]
-    So
-    \[
-    \pd{\E h(X)}{\theta}
-    = I_1 + I_2
-    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
-    \]
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$
-
-    Obviously
-    \[
-    I_1 = \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx
-    < h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
-    \]
-    and
-    \begin{eqnarray*}
-      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
-      &<& -h(x_0) \int_{x_0}^b \left| \pd{f}{\theta}(x, \theta) \right| dx \\
-    \end{eqnarray*}
-    Thus
-    \[
-    \pd{\E h(X)}{\theta}
-    = I_1 + I_2
-    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
-    \]
-  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
-
-    In this case
-    \begin{eqnarray*}
-      I_1 &=& \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx \\
-      &<& -h(x_0) \int_a^{x_0} \left| \pd{f}{\theta}(x, \theta) \right| dx
-    \end{eqnarray*}
-    And
-    \begin{eqnarray*}
-      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
-      &<& h(x_0) \int_{x_0}^b \pd{f}{\theta}(x, \theta) dx
-    \end{eqnarray*}
-    So
-    \[
-    \pd{\E h(X)}{\theta}
-    = I_1 + I_2
-    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
-    \]
-  \end{enumerate}
-\end{proof}
-With lemma \ref{lemma:I} it is straightforward to show the
-monotonicity claimed earlier.
+% In theorem \ref{thrm:I} we show
+% \[
+% \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
+% \]
+% is monotonically increasing with $\alpha$ and monotonically decreasing
+% with $K$. But before that it is worthwhile to state the following
+% simple observation:
+% With lemma \ref{lemma:I} it is straightforward to show the
+% monotonicity claimed earlier.
 Let
@@
   The function $u((1 - \phi) e^r + \phi e^X)$ is obviously increasing
-  with $X$. Furthermore
-  \begin{eqnarray*}
-    && \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+  with $X$. Furthermore, the density function of $X | X < 0$ is ${1
+    \over p} f(x; \alpha, K)$. It follows
+  \begin{eqnarray*}
+    && \pd{f(x; \alpha, K)}{\alpha} \\
+    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
     &=&
@@
   \begin{eqnarray*}
-    && \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+    && \pd{f(x; \alpha, K)}{K} \\
+    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
     &=&
@@
 sectors are shown in figures \ref{fig:Energy_lower},
-\ref{fig:Consumer_Staples_lower} and \ref{fig:Information_Technology_lower}.
-% all agents guided by this class of utility
-% functions will seek the largest $\alpha$ and the smallest $K$ found in
-% a stock group e.g. a sector.
-
-% As a result, one can expect all the
-% actively traded equities in the group to feature the same $\alpha$ and
-% $K$. However, the estimated value of $K^\alpha$ may vary
-% significantly. This can be understood as a result of the smallness of
-% $K$, which as revealed in section \ref{sec:hybrid_estimation} is of the
-% order of $10^{-2}$, and the inaccuracy of the estimated value of
-% $\alpha$, which typically ranges from 2.5 to 5. Exponentiation hugely
-% magnifies the inaccuracy in $\alpha$.
+\ref{fig:Consumer_Staples_lower} and
+\ref{fig:Information_Technology_lower}.
+
+\subsection{Student's t-distributed equity return}
+It is a common practice to use Student's t-distribution to model the
+stationary distribution of equity returns. So it is of interest to
+find out what implications this distribution has when it is combined
+with the PDA preference. Formally we assume
+\[
+f(x; \nu) = c(\nu) \left(
+  1 + {x^2 \over \nu}
+\right)^{-(\nu + 1)/2}
+\]
+where $\nu > 1$ and
+\[
+c(\nu) = {
+  \Gamma({\nu + 1 \over 2})
+  \over
+  \Gamma(\nu/2) \sqrt{\nu \pi}
+}
+\]
+Then it follows
+\begin{eqnarray*}
+\pd{f(x, \nu)}{\nu} &=& {
+  c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
+  -
+  \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
+  \over
+  2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
+}
+\end{eqnarray*}
+The nummerator of the right side of $\pd{f(x, \nu)}{\nu}$, if seen as
+a function of $x^2$, has a single root:
+\begin{equation}
+  \label{eq:xxie5}
+  x^2 = \nu\exp\left\{
+    W\left[
+      -\left(1 + {1 \over \nu}\right)
+      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+    \right]
+    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
+  \right\} - \nu
+\end{equation}
+where $W(\cdot)$ is the principle branch of the Lambert $W$
+function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
+the right side of \eqref{eq:xxie5} for positivity, we first note
+$c'(\nu) > 0$:
+\[
+c'(\nu) = {
+  \pi \Gamma(\nu/2 + 1/2) \left[
+    \nu \Psi(\nu/2 + 1/2) - \nu \Psi(\nu/2) - 1
+  \right]
+  \over
+  2 \Gamma(\nu/2) (\pi \nu)^{3/2}
+}
+\]
+where $\Psi(\cdot)$ is the digamma function:
+\[
+\Psi(x) = \td{\ln[\Gamma(x)]}{x}
+\]
+Since for all $x > 0$, $\Psi(x) > 0$ and $\nu > 1$, it is clear
+$c'(\nu) > 0$. Furthermore, we recall
+$W(\cdot)$ is increasing on its principle branch. So
+\begin{eqnarray*}
+  &&
+  W\left[
+    -\left( 1 + {1 \over \nu} \right)
+    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+  \right]
+  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+  &>& 
+  W\left[
+    -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
+    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+  \right]
+  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+  &=& W(-y e^{-y}) + y
+\end{eqnarray*}
+where
+\[
+y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
+\]
+Now notice
+\[
+\ln(y e^{-y}) = \ln(y) - y
+\]
+is a decreasing function. Thus $-y e^{-y}$ is an increasing
+function. Hence we have
+\begin{eqnarray*}
+  W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
+\end{eqnarray*}
+Now it is clear
+\[
+\nu\exp\left\{
+    W\left[
+      -\left(1 + {1 \over \nu}\right)
+      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+    \right]
+    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
+  \right\} - \nu > 0
+\]
 

### 2017-03-09 papers/FX/draft.tex
some progress on the student's t case.
close to finishing the proof in the Student's t case.

@@
 Then it follows
-\begin{eqnarray*}
+\begin{eqnarray}
 \pd{f(x, \nu)}{\nu} &=& {
@@
   2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
-}
-\end{eqnarray*}
-The nummerator of the right side of $\pd{f(x, \nu)}{\nu}$, if seen as
-a function of $x^2$, has a single root:
+  } \nonumber \\
+  &:=& {
+    P(x^2, \nu)
+    \over
+    2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
+  }
+  \label{eq:xxie4.1}
+\end{eqnarray}
+
+While the denominator of the right side of $\pd{f(x, \nu)}{\nu}$ is
+always positive, its nummerator $P(x^2, \nu)$ has a single root:
 \begin{equation}
   \label{eq:xxie5}
-  x^2 = \nu\exp\left\{
+  x_0^2 = \nu\exp\left\{
     W\left[
@@
 c'(\nu) = {
-  \pi \Gamma(\nu/2 + 1/2) \left[
-    \nu \Psi(\nu/2 + 1/2) - \nu \Psi(\nu/2) - 1
-  \right]
+  \pi \Gamma(\nu/2 + 1/2) \left\{
+    \nu \left[ \Psi(\nu/2 + 1/2) - \Psi(\nu/2)\right] - 1
+  \right\}
   \over
@@
 \]
-Since for all $x > 0$, $\Psi(x) > 0$ and $\nu > 1$, it is clear
-$c'(\nu) > 0$. Furthermore, we recall
+\begin{minipage}{0.48\textwidth}
+  $\Psi(x)$ \linebreak
+  \includegraphics[width=\textwidth]{digamma.png}
+\end{minipage}\hfill
+\begin{minipage}{0.5\textwidth}
+  As shown in the figure to the left, $\Psi(x)$ is increasing
+  for $x > 0$. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
+  So we have
+  \begin{eqnarray*}
+    && \nu \left[
+    \Psi(\nu/2 + 1/2) - \Psi(\nu/2)
+    \right] - 1 \\
+    &\geq& 1 \times \left[
+    \Psi(1/2 + 1/2) - \Psi(1/2)
+    \right] - 1 \\
+    &=& \ln(4) - \ln(e) \\
+    &>& 0
+  \end{eqnarray*}
+  Thus $c'(\nu) > 0$. Furthermore, we recall
   $W(\cdot)$ is increasing on its principle branch. So
+\end{minipage}
 \begin{eqnarray*}
@@
 \]
+Now that we have established that $\pd{f}{\nu}(x, \nu) = 0$ has a
+single positive root, it remains to determine the sign of
+$\pd{f}{\nu}(x, \nu)$ on the two sides of the root in order to apply
+lemma \ref{lemma:I}. For this purpose we observe
+\begin{equation}
+  \label{eq:xxie5.1}
+  P(0, \nu) = 2 \nu^2 c'(\nu) > 0
+\end{equation}
+So we want to investigate $\pd{P}{x^2}(x^2, \nu)$:
+\begin{eqnarray}
+  && \pd{P}{x^2}(x^2, \nu) \nonumber \\
+  &=&
+  2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \ln\left(
+  1 + {x^2 \over \nu}
+  \right)
+  \label{eq:xxie6}
+\end{eqnarray}
+From \eqref{eq:xxie5.1} and \eqref{eq:xxie6} it is clear
+$P(x^2, \nu) > 0$ for $0 \leq x^2 < x_0^2$ and
+$P(x^2, \nu) < 0$ for $x^2 > x_0^2$.
 

### 2017-03-09 papers/FX/draft.tex
The first part of the proof is more complicated than I imagined!.

@@
 this means $\delta v > e^r$, which implies $q(r, \phi) > r > 0$.
-The right-hand-side of \eqref{eq:functional} is a functional of the
-distribution function of $X$ via \eqref{eq:xxie1}. For convenience
-we denote it $\mathcal U(\cdot)$, i.e.
-\[
-\mathcal U F (\alpha, \beta, K, K', \phi)
+The right-hand-side of \eqref{eq:functional} is a functional of $\phi$
+and the distribution function of $X$ via \eqref{eq:xxie1}. For convenience
+we denote it $\mathcal U(\cdot, \cdot)$, i.e.
+\[
+\mathcal U(F, \phi)
 = 
@@
 distribution function of $X$ and use $f(\cdot)$ to denote its density
-function.
+function. Let
+\[
+G(F) = \max_{0 < \phi \leq 1} \mathcal U(F, \phi)
+\]
 
@@
 \end{eqnarray}
-% In theorem \ref{thrm:I} we show
-% \[
-% \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
-% \]
-% is monotonically increasing with $\alpha$ and monotonically decreasing
-% with $K$. But before that it is worthwhile to state the following
-% simple observation:
-% With lemma \ref{lemma:I} it is straightforward to show the
-% monotonicity claimed earlier.
-Let
-\[
-G(\alpha, \beta, K, K') =
-\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
-\]
 \begin{theorem}
@@
   \[
-  \pd{G}{\alpha} (\alpha, \beta, K, K') > 0,
+  \pd{G(F)}{\alpha} (\alpha, \beta, K, K') > 0,
   \quad
-  \pd{G}{K} (\alpha, \beta, K, K') < 0
+  \pd{G(F)}{K} (\alpha, \beta, K, K') < 0
   \]
@@
     \label{eq:xxie2}
-    \hat \phi := \argmax_{0 < \phi \leq 1}
-    \mathcal U F(\alpha, \beta, K, K', \phi)
+    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
   \end{equation}
@@
   \[
-  G(\alpha, \beta, K, K') = \mathcal U F(\alpha, \beta, K, K', \hat \phi)
+  G(F)(\alpha, \beta, K, K') = \mathcal U(F, \hat\phi)(\alpha, \beta, K, K')
   \]
@@
   \begin{eqnarray}
-    \pd{G}{\alpha}
+    \pd{G(F)}{\alpha}
     &=&
-    \pd{\mathcal U F}{\alpha}(\alpha, \beta, K, K', \hat \phi)
-    + \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi)
-    \pd{\hat \phi}{\alpha}(\alpha, \beta, K, K')
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
+    \nonumber \\
+    && + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, \beta, K, K', \hat\phi)
+    \pd{\hat\phi}{\alpha}(\alpha, \beta, K, K')
     \label{eq:xxie3}
@@
     \label{eq:xxie4}
-    \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
+    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
   \end{equation}
@@
   \begin{eqnarray*}
-    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
     &>& 0
@@
   \begin{eqnarray*}
-    && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
+    && \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
     \\
@@
   \begin{eqnarray*}
-    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
   \end{eqnarray*}
@@
   \[
-  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
   \]
@@
   \[
-  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi)
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi)
   = p \opd{K} \E \left[
@@
   \[
-  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0  
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi) < 0  
   \]
@@
 \]
+\begin{theorem}
+  \[
+  \pd{G(F)}{\nu} > 0
+  \]
+  where the distribution function $F(x, \nu)$ corresponds to the
+  density function $f(x, \nu)$ defined above.
+\end{theorem}
+\begin{proof}
+  By the same argument as in the proof of theorem \ref{thrm:I}, it
+  suffices to show
+  \begin{eqnarray*}
+    \pd{\mathcal U(F, \phi)}{\nu}(\nu, \hat\phi) &>& 0
+  \end{eqnarray*}
+  where $\hat\phi$ is defined in \eqref{eq:xxie2}. We have
+  \begin{eqnarray*}
+    && \mathcal U(F, \phi) \\
+    &=&
+    \int_{-\infty}^\infty
+    u[(1 - \phi) e^r + \phi e^x] (1 + b \1{x < q}) f(x, \nu) dx
+  \end{eqnarray*}
+
+\end{proof}
+
 Then it follows

### 2017-03-10 papers/FX/draft.tex
Revised the draft.
Student's t with log-utility is solved.

@@
   function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
-  $(a, b)$, $X \sim F$, $\E h(X) < \infty$ and $\forall x \in (a, b)$
-  \[
+  $(a, b)$, $X \sim F$, $\text{supp}(X) = (a,b)$,
+  $\E h(X) < \infty$ and
+  \begin{eqnarray*}
+    \opd{\theta}\left[
+      F(b) - F(a)
+    \right] = 0 \\
     \left| \pd{f(x, \theta)}{\theta} \right| < \infty
-  \]
+    \quad \forall x \in (a, b)
+  \end{eqnarray*}
   \begin{enumerate}
@@
     \]
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
-    \[
-    \pd{\E h(X)}{\theta} < 0
-    \]
-  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
-    \[
-    \pd{\E h(X)}{\theta} < 0
-    \]
   \end{enumerate}
 \end{lemma}
-\begin{proof}
-  Clearly
-  \begin{eqnarray*}
-    && \pd{\E h(X)}{\theta} \\
-    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
-    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
-    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
-  \end{eqnarray*}
+\begin{remark}
+  \label{remark:I}
+  Two other cases follow trivially from lemma \ref{lemma:I}:
   \begin{enumerate}
-  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
-    \theta) > 0$ on $(a, x_0)$
-    \[
-    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
-    \]
-    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
-    $h(x)$ is decreasing, we have
-    \begin{eqnarray*}
-      I_2 &=& \int_{x_0}^b -h(x_0)
-      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
-      &>& -h(x_0)
-      \int_{x_0}^b \left| 
-        \pd{f}{\theta}(x, \theta)
-      \right| dx
-    \end{eqnarray*}
-    Finally we have
-    \begin{eqnarray*}
-      && \pd{\E h(X)}{\theta} \\
-      &=& I_1 + I_2 \\
-      &>& h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx \\
-      &=& h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx \\
-      &=& h(x_0) \opd{\theta} 1 = 0
-    \end{eqnarray*}
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
-    \begin{eqnarray*}
-      I_1 &=&
-      \int_a^{x_0} -h(x)
-      \left| \pd{f}{\theta}(x, \theta) \right| dx \\
-      &>&
-      -h(x_0) \int_a^{x_0}
-      \left| \pd{f}{\theta}(x, \theta) \right| dx
-    \end{eqnarray*}
-    and obviously
-    \[
-    I_2 > h(x_0) \int_{x_0}^b
-    \pd{f}{\theta}(x, \theta) dx
-    \]
-    So
-    \[
-    \pd{\E h(X)}{\theta}
-    = I_1 + I_2
-    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
-    \]
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$
-
-    Obviously
-    \[
-    I_1 = \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx
-    < h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
-    \]
-    and
-    \begin{eqnarray*}
-      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
-      &<& -h(x_0) \int_{x_0}^b \left| \pd{f}{\theta}(x, \theta) \right| dx \\
-    \end{eqnarray*}
-    Thus
-    \[
-    \pd{\E h(X)}{\theta}
-    = I_1 + I_2
-    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
-    \]
-  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
-
-    In this case
-    \begin{eqnarray*}
-      I_1 &=& \int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx \\
-      &<& -h(x_0) \int_a^{x_0} \left| \pd{f}{\theta}(x, \theta) \right| dx
-    \end{eqnarray*}
-    And
-    \begin{eqnarray*}
-      I_2 &=& \int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx \\
-      &<& h(x_0) \int_{x_0}^b \pd{f}{\theta}(x, \theta) dx
-    \end{eqnarray*}
-    So
-    \[
-    \pd{\E h(X)}{\theta}
-    = I_1 + I_2
-    < h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
-    \]
+  \item If $h(\cdot)$ is increasing and $\pd{f}{\theta}$ satisfies the
+    same conditions of the 1st case of lemma \ref{lemma:I},
+    $\pd{\E h(X)}{\theta} < 0$. This immediately follows from applying
+    1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
+  \item By the same argument, if $h(\cdot)$ is decreasing and
+    $\pd{f}{\theta}$ satisfies the same conditions of the 2nd case of
+    lemma \ref{lemma:I}, $\pd{\E h(X)}{\theta} < 0$.
   \end{enumerate}
-\end{proof}
-
+\end{remark}
 Now for simplicity we assume the investor initially has one unit of
@@
 \begin{eqnarray*}
-  q(r, \phi) &:=& \ln \left( {
+  q = q(r, \phi) &:=& \ln \left( {
       \delta v - (1 - \phi) e^r
@@
 \end{eqnarray*}
-Note $C < \delta v$ implies
-\[
-X < q(r, \phi)
-\]
-Hereafter we abbreviate $q(r, \phi)$ to just $q$. 
-Naturally, if an agent invests in a risky asset instead of a riskless
-bond, he expects to obtain a higher return from the risky asset than
-he is guarranteed from the riskless bond. In our notation,
-this means $\delta v > e^r$, which implies $q(r, \phi) > r > 0$.
-The right-hand-side of \eqref{eq:functional} is a functional of $\phi$
-and the distribution function of $X$ via \eqref{eq:xxie1}. For convenience
-we denote it $\mathcal U(\cdot, \cdot)$, i.e.
-\[
+Note $C < \delta v$ implies $X < q$. Naturally, if an agent invests in
+a risky asset instead of a riskless bond, he expects to obtain a
+higher return from the risky asset than he is guarranteed from the
+riskless bond. In our notation, this means $\delta v > e^r$, which
+implies $q(r, \phi) > r > 0$. The right-hand-side of
+\eqref{eq:functional} is a functional of $\phi$ and the distribution
+function of $X$ via \eqref{eq:xxie1}. For convenience we denote it
+$\mathcal U(\cdot, \cdot)$, i.e.
+\begin{equation}
+  \label{eq:U_functional}
   \mathcal U(F, \phi)
@@
   \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
-\]
+\end{equation}
 In the rest of this section, we use $F(\cdot)$ to denote the
@@
 function. Let
-\[
+\begin{equation}
+  \label{eq:G_functional}
   G(F) = \max_{0 < \phi \leq 1} \mathcal U(F, \phi)  
-\]
+\end{equation}
 
@@
   u\left[ (1 - \phi) e^r + \phi e^x \right]
-  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
-  &&
-  - b u(\delta v) \left[
-    1 - (1 - p) {
-      K'^\beta
-      \over
-      (K' + q)^\beta
-    }
-  \right]
+  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx 
+  + \text{const.}
+  % &&
+  % - b u(\delta v) \left[
+  %   1 - (1 - p) {
+  %     K'^\beta
+  %     \over
+  %     (K' + q)^\beta
+  %   }
+  % \right]
   \label{eq:xxie1.0}
 \end{eqnarray}
+where the last term is a constant independent of $\phi$, $\alpha$ and $K$.
 \begin{theorem}
   \label{thrm:I}
-  \[
-  \pd{G(F)}{\alpha} (\alpha, \beta, K, K') > 0,
+  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
+  function $u(\cdot)$ is increasing and differentiable, then
+  \[
+  \pd{G(F)}{\alpha} > 0,
   \quad
-  \pd{G(F)}{K} (\alpha, \beta, K, K') < 0
+  \pd{G(F)}{K} < 0
   \]
 \end{theorem}
-
-\begin{proof}
-  Let
-  \begin{equation}
-    \label{eq:xxie2}
-    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
-  \end{equation}
-  We have
-  \[
-  G(F)(\alpha, \beta, K, K') = \mathcal U(F, \hat\phi)(\alpha, \beta, K, K')
-  \]
-  It follows
-  \begin{eqnarray}
-    \pd{G(F)}{\alpha}
-    &=&
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
-    \nonumber \\
-    && + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, \beta, K, K', \hat\phi)
-    \pd{\hat\phi}{\alpha}(\alpha, \beta, K, K')
-    \label{eq:xxie3}
-  \end{eqnarray}
-  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
-  \begin{equation}
-    \label{eq:xxie4}
-    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
-  \end{equation}
-  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
-  \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
-    &>& 0
-  \end{eqnarray*}
-  From \eqref{eq:xxie1.0}, it follows
-  \begin{eqnarray*}
-    && \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
-    \\
-    &=& p \opd{\alpha}\E[u((1 - \phi) e^r + \phi e^X) | X < 0]
-  \end{eqnarray*}
-  The function $u((1 - \phi) e^r + \phi e^X)$ is obviously increasing
-  with $X$. Furthermore, the density function of $X | X < 0$ is ${1
-    \over p} f(x; \alpha, K)$. It follows
-  \begin{eqnarray*}
-    && \pd{f(x; \alpha, K)}{\alpha} \\
-    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
-    &=&
-    - {K^\alpha \over (K - x)^{\alpha + 1}}
-    \left[
-      \alpha
-      \ln\left(
-        1 - {x \over K}
-      \right) - 1
-    \right]
-  \end{eqnarray*}
-  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
-  \[
-  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
-  \]
-  and when $K(1 - e^{1/\alpha}) < x < 0$,
-  \[
-  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
-  \]
-  This is the second case of lemma \ref{lemma:I}. So we have
-  \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
-  \end{eqnarray*}
-  As for $\pd{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
-  $\pd{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
-  \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
-  \]
-  We have
-  \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi)
-  = p \opd{K} \E \left[
-    u((1 - \phi) e^r + \phi e^X)
-    | X < 0
-  \right] 
-  \]
-  and
-  \begin{eqnarray*}
-    && \pd{f(x; \alpha, K)}{K} \\
-    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
-    &=&
-    -\alpha K^{\alpha - 1} {
-      \alpha x + K
-      \over
-      (K - x)^{\alpha + 2}
-    }
-  \end{eqnarray*}
-  Clearly, when $x < -\alpha/K$
-  \[
-  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
-  \]
-  and when $-\alpha/K < x < 0$
-  \[
-  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
-  \]
-  So by the third case of lemma \ref{lemma:I} we conclude
-  \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi) < 0  
-  \]
-\end{proof}
 Since $u(v)$ given in \eqref{eq:xxie1.0} as a
@@
 
+\begin{proof}
+  Let
+  \begin{equation}
+    \label{eq:xxie2}
+    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
+  \end{equation}
+  We have
+  \[
+  G(F) = \mathcal U(F, \hat\phi)
+  \]
+  It follows
+  \begin{eqnarray}
+    \pd{G(F)}{\alpha}
+    &=&
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
+    + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat\phi)
+    \pd{\hat\phi}{\alpha}(\alpha, K)
+    \label{eq:xxie3}
+  \end{eqnarray}
+  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
+  \begin{equation}
+    \label{eq:xxie4}
+    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat \phi) = 0
+  \end{equation}
+  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
+  \begin{eqnarray*}
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat \phi)
+    &>& 0
+  \end{eqnarray*}
+  From \eqref{eq:xxie1.0}, it follows
+  \begin{eqnarray*}
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
+    = \opd{\alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}]
+  \end{eqnarray*}
+  The function $u((1 - \hat\phi) e^r + \hat\phi e^X)$ is obviously increasing
+  with $X$. It follows
+  \begin{eqnarray*}
+    && \pd{f(x; \alpha, K)}{\alpha} \\
+    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+    &=&
+    - {K^\alpha \over (K - x)^{\alpha + 1}}
+    \left[
+      \alpha
+      \ln\left(
+        1 - {x \over K}
+      \right) - 1
+    \right]
+  \end{eqnarray*}
+  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
+  \[
+  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  \]
+  and when $K(1 - e^{1/\alpha}) < x < 0$,
+  \[
+  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  \]
+  This is the second case of lemma \ref{lemma:I}. So we have
+  \begin{eqnarray*}
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi) &>& 0
+  \end{eqnarray*}
+  As for $\pd{G(F)}{K}$, by the same argument as for
+  $\pd{G(F)}{\alpha}$, it suffices to show
+  \[
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0
+  \]
+  We have
+  \[
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi)
+  = \opd{K} \E \left[
+    u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}
+  \right] 
+  \]
+  and
+  \begin{eqnarray*}
+    && \pd{f(x; \alpha, K)}{K} \\
+    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+    &=&
+    -\alpha K^{\alpha - 1} {
+      \alpha x + K
+      \over
+      (K - x)^{\alpha + 2}
+    }
+  \end{eqnarray*}
+  Clearly, when $x < -\alpha/K$
+  \[
+  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  \]
+  and when $-\alpha/K < x < 0$
+  \[
+  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  \]
+  So by the 1st case of remark \ref{remark:I} we conclude
+  \[
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0  
+  \]
+\end{proof}
+
 \subsection{Student's t-distributed equity return}
@@
 \]
+\begin{lemma}
+  \label{lemma:II}
+  There exists $x_0 > 0$ such that $\pd{f}{\nu}(x_0, \nu) = 0$ and
+  $\forall x \in (0, x_0), \pd{f}{\nu}(x, \nu) > 0$ and
+  $\forall x \in (x_0, \infty), \pd{f}{\nu}(x, \nu) < 0$.
+\end{lemma}
+
+\subsubsection{log-utility function}
+We first consider the case where the utility function is simply
+$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
+\eqref{eq:power_utility} as $\gamma \to 0$.
 \begin{theorem}
-  \[
-  \pd{G(F)}{\nu} > 0
-  \]
-  where the distribution function $F(x, \nu)$ corresponds to the
-  density function $f(x, \nu)$ defined above.
+  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
+  function $u(\cdot) = \ln(\cdot)$, then
+  \[
+  \pd{G(F)}{\nu} < 0
+  \]
+  where $F(x, \nu)$ is the Student's t-distribution function.
 \end{theorem}
@@
   \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\nu}(\nu, \hat\phi) &>& 0
+    \pd{\mathcal U(F, \phi)}{\nu}(\nu, \hat\phi) &<& 0
   \end{eqnarray*}
@@
 \]
-is a decreasing function. Thus $-y e^{-y}$ is an increasing
-function. Hence we have
+is a decreasing function for $y > 1$. Thus $-y e^{-y}$ is an
+increasing function. Hence we have
 \begin{eqnarray*}
@@
 
+\appendix
+\section{Proof of lemma \ref{lemma:I}}
+\begin{proof}
+  Clearly
+  \begin{eqnarray*}
+    && \pd{\E h(X)}{\theta} \\
+    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
+    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
+    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
+  \end{eqnarray*}
+  \begin{enumerate}
+  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
+    \theta) > 0$ on $(a, x_0)$
+    \[
+    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
+    \]
+    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
+    $h(x)$ is decreasing, we have
+    \begin{eqnarray*}
+      I_2 &=& \int_{x_0}^b -h(x_0)
+      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
+      &>& -h(x_0)
+      \int_{x_0}^b \left| 
+        \pd{f}{\theta}(x, \theta)
+      \right| dx
+    \end{eqnarray*}
+    Finally we have
+    \begin{eqnarray*}
+      \pd{\E h(X)}{\theta}
+      > h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx
+      = h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx
+      = 0
+    \end{eqnarray*}
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
+    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
+    \begin{eqnarray*}
+      I_1 &=&
+      \int_a^{x_0} -h(x)
+      \left| \pd{f}{\theta}(x, \theta) \right| dx \\
+      &>&
+      -h(x_0) \int_a^{x_0}
+      \left| \pd{f}{\theta}(x, \theta) \right| dx
+    \end{eqnarray*}
+    and obviously
+    \[
+    I_2 > h(x_0) \int_{x_0}^b
+    \pd{f}{\theta}(x, \theta) dx
+    \]
+    So
+    \[
+    \pd{\E h(X)}{\theta}
+    = I_1 + I_2
+    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
+    \]
+  \end{enumerate}
+\end{proof}
+
+\section{Proof of lemma \ref{lemma:II}}
+
 \bibliographystyle{unsrt}

### 2017-03-13 papers/FX/draft.tex
There is still some hope in the Student's t and log-utility situation.

@@
   \label{eq:xxie1}
-  C = (1 - \phi) e^r + \phi e^X
+  C(X) = (1 - \phi) e^r + \phi e^X
 \end{equation}
@@
 implies $q(r, \phi) > r > 0$. The right-hand-side of
-\eqref{eq:functional} is a functional of $\phi$ and the distribution
+\eqref{eq:functional} is a functional of $\phi$ and of the distribution
 function of $X$ via \eqref{eq:xxie1}. For convenience we denote it
@@
     \int_{-\infty}^\infty
-    u[(1 - \phi) e^r + \phi e^x] (1 + b \1{x < q}) f(x, \nu) dx
-  \end{eqnarray*}
-
-\end{proof}
-
-Then it follows
+    \ln[C(x)] (1 + b \1{x < q})f(x, \nu) dx
+    - b \ln(\delta v) F(q, \nu) \\
+    &=& 
+    \int_{0}^\infty
+    \ln\left[C(x) C(-x)\right]
+    f(x, \nu) dx
+    +
+    b \int_{-\infty}^\infty
+    \ln
+    \left[
+      C(x)
+      \over
+      C(q)
+    \right]
+    \1{x < q} f(x, \nu) dx \\
+    &=&
+    \underbrace{
+      \int_{0}^\infty
+      \ln\left[C(x) C(-x)\right]
+      f(x, \nu) dx
+    }_{I_1}
+    +
+    b \underbrace{
+      \int_{0}^\infty
+      \left\{
+      \ln
+      \left[
+        C(x)
+        \over
+        C(q)
+        \right]
+      \1{x < q}
+      +
+      \ln\left[
+        C(-x)
+        \over
+        C(q)
+        \right]
+      \right\}
+      f(x, \nu) dx
+    }_{I_2}
+  \end{eqnarray*}
+  where $C(\cdot)$ is defined in \eqref{eq:xxie1}.
+  Observe that $\ln(C(x)C(-x))$ is an increasing function:
+  \begin{eqnarray*}
+    \opd{x}
+    \left[
+      \ln(C(x) C(-x))
+    \right]
+    &=&
+    {
+      \phi (1- \phi) e^r (e^x - e^{-x})
+      \over
+      ((1 - \phi) e^r + \phi e^x)
+      ((1 - \phi) e^r + \phi e^{-x})
+    } > 0
+  \end{eqnarray*}
+  Thus by 1st case of remark \ref{remark:I} and lemma \ref{lemma:II},
+  $\pd{I_1}{\nu} < 0$. If $b = 0$, i.e.
+  $\mathcal U(F, \phi) = \E u(C(X))$, then we have concluded
+  $\pd{G(F)}{\nu} < 0$. Otherwise we need to investigate $I_2$:
+  \begin{enumerate}
+  \item In case $q \leq x_0$, by lemma \ref{lemma:II},
+    $\pd{f}{\nu}(x, \nu) > 0$ for $x \leq q < x_0$. Moreover
+    \[
+    \ln\left[
+      C(-x)
+      \over
+      C(q)
+      \right] < 0
+    \quad
+    \forall x > 0
+    \]
+    Therefore
+    \begin{eqnarray*}
+      \pd{I_2}{\nu}
+      &<&
+      \int_{0}^\infty
+      \ln
+      \left[
+        C(x) C(-x)
+        \over
+        C(q)^2
+        \right]
+      \1{x < q}
+      \pd{f}{\nu}(x, \nu) dx \\
+      &&
+      + \int_{0}^\infty
+      \ln
+      \left[
+        C(-x)
+        \over
+        C(q)
+        \right]
+      \1{x > x_0}
+      \pd{f}{\nu}(x, \nu) dx \\
+    \end{eqnarray*}
+
+  \item If $q > x_0$
+  \end{enumerate}
+
   \begin{eqnarray}
@@
   \end{eqnarray*}
+%   where
+%   \[
+%   y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
+%   \]
+%   Now notice
+%   \[
+%   \ln(y e^{-y}) = \ln(y) - y
+%   \]
+%   is a decreasing function. Thus $-y e^{-y}$ is an increasing
+%   function. Hence we have
+%   \begin{eqnarray*}
+%     W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
+%   \end{eqnarray*}
+%   Now it is clear
+%   \[
+%   \nu\exp\left\{
+%   W\left[
+%     -\left(1 + {1 \over \nu}\right)
+%     e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+%   \right]
+%   + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+%   &=& W(-y e^{-y}) + y
+% \end{eqnarray*}
 where
@@
 
+\end{proof}
 \section{Review of multivariate GARCH models:

### 2017-03-13 papers/FX/draft.tex
I think we have said all that there is to be said in the case of t-distributed X.

--- 2017-03-10 papers/FX/draft.tex
@@
 $u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
-\eqref{eq:power_utility} as $\gamma \to 0$.
-\begin{theorem}
-  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
-  function $u(\cdot) = \ln(\cdot)$, then
-  \[
-  \pd{G(F)}{\nu} < 0
-  \]
-  where $F(x, \nu)$ is the Student's t-distribution function.
-\end{theorem}
-\begin{proof}
-  By the same argument as in the proof of theorem \ref{thrm:I}, it
-  suffices to show
-  \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\nu}(\nu, \hat\phi) &<& 0
-  \end{eqnarray*}
-  where $\hat\phi$ is defined in \eqref{eq:xxie2}. We have
+\eqref{eq:power_utility} as $\gamma \to 0$. Straightforward
+computation gives
 \begin{eqnarray*}
@@
   &=&
-    \int_{-\infty}^\infty
-    u[(1 - \phi) e^r + \phi e^x] (1 + b \1{x < q}) f(x, \nu) dx
-  \end{eqnarray*}
-
-\end{proof}
-
-Then it follows
-\begin{eqnarray}
-\pd{f(x, \nu)}{\nu} &=& {
-  c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
-  -
-  \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
-  \over
-  2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
-  } \nonumber \\
-  &:=& {
-    P(x^2, \nu)
-    \over
-    2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
-  }
-  \label{eq:xxie4.1}
-\end{eqnarray}
-
-While the denominator of the right side of $\pd{f(x, \nu)}{\nu}$ is
-always positive, its nummerator $P(x^2, \nu)$ has a single root:
-\begin{equation}
-  \label{eq:xxie5}
-  x_0^2 = \nu\exp\left\{
-    W\left[
-      -\left(1 + {1 \over \nu}\right)
-      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+  \int_{0}^{\infty}
+  \left[
+    \ln(C(x)) + (1 + b)\ln(C(-x))
   \right]
-    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
-  \right\} - \nu
-\end{equation}
-where $W(\cdot)$ is the principle branch of the Lambert $W$
-function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
-the right side of \eqref{eq:xxie5} for positivity, we first note
-$c'(\nu) > 0$:
-\[
-c'(\nu) = {
-  \pi \Gamma(\nu/2 + 1/2) \left\{
-    \nu \left[ \Psi(\nu/2 + 1/2) - \Psi(\nu/2)\right] - 1
-  \right\}
-  \over
-  2 \Gamma(\nu/2) (\pi \nu)^{3/2}
-}
-\]
-where $\Psi(\cdot)$ is the digamma function:
-\[
-\Psi(x) = \td{\ln[\Gamma(x)]}{x}
-\]
-\begin{minipage}{0.48\textwidth}
-  $\Psi(x)$ \linebreak
-  \includegraphics[width=\textwidth]{digamma.png}
-\end{minipage}\hfill
-\begin{minipage}{0.5\textwidth}
-  As shown in the figure to the left, $\Psi(x)$ is increasing
-  for $x > 0$. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
-  So we have
-  \begin{eqnarray*}
-    && \nu \left[
-    \Psi(\nu/2 + 1/2) - \Psi(\nu/2)
-    \right] - 1 \\
-    &\geq& 1 \times \left[
-    \Psi(1/2 + 1/2) - \Psi(1/2)
-    \right] - 1 \\
-    &=& \ln(4) - \ln(e) \\
-    &>& 0
-  \end{eqnarray*}
-  Thus $c'(\nu) > 0$. Furthermore, we recall
-  $W(\cdot)$ is increasing on its principle branch. So
-\end{minipage}
-\begin{eqnarray*}
+  f(x, \nu) dx
+  + b \int_{0}^{q}
+  \ln(C(x))
+  f(x, \nu) dx \\
   &&
-  W\left[
-    -\left( 1 + {1 \over \nu} \right)
-    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
-  \right]
-  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
-  &>& 
-  W\left[
-    -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
-    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
-  \right]
-  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
-  &=& W(-y e^{-y}) + y
-\end{eqnarray*}
-where
-\[
-y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
-\]
-Now notice
-\[
-\ln(y e^{-y}) = \ln(y) - y
-\]
-is a decreasing function for $y > 1$. Thus $-y e^{-y}$ is an
-increasing function. Hence we have
-\begin{eqnarray*}
-  W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
-\end{eqnarray*}
-Now it is clear
-\[
-\nu\exp\left\{
-    W\left[
-      -\left(1 + {1 \over \nu}\right)
-      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
-    \right]
-    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
-  \right\} - \nu > 0
-\]
-Now that we have established that $\pd{f}{\nu}(x, \nu) = 0$ has a
-single positive root, it remains to determine the sign of
-$\pd{f}{\nu}(x, \nu)$ on the two sides of the root in order to apply
-lemma \ref{lemma:I}. For this purpose we observe
-\begin{equation}
-  \label{eq:xxie5.1}
-  P(0, \nu) = 2 \nu^2 c'(\nu) > 0
-\end{equation}
-So we want to investigate $\pd{P}{x^2}(x^2, \nu)$:
-\begin{eqnarray}
-  && \pd{P}{x^2}(x^2, \nu) \nonumber \\
+  - b \ln(C(q)) F(q, \nu)
+\end{eqnarray*}
+where $C(\cdot)$ is defined in \eqref{eq:xxie1}. Notice that
+when $b = 0$, $\mathcal U(F, \phi)$  reduces to $\E u(C(X))$.
+Also observe $\ln(C(x)C(-x))$ is an increasing function:
+\begin{eqnarray*}
+  \pd{C(x) C(-x)}{x}
   &=&
-  2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \ln\left(
-  1 + {x^2 \over \nu}
-  \right)
-  \label{eq:xxie6}
-\end{eqnarray}
-From \eqref{eq:xxie5.1} and \eqref{eq:xxie6} it is clear
-$P(x^2, \nu) > 0$ for $0 \leq x^2 < x_0^2$ and
-$P(x^2, \nu) < 0$ for $x^2 > x_0^2$.
-
+  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
+\end{eqnarray*}
+Thus by 1st case of remark \ref{remark:I} and lemma \ref{lemma:II},
+$\pd{\mathcal U(F, \phi))}{\nu} < 0$ for all $\phi$. Hence all
+investors guided by this utility function would be seeking the
+smallest $\nu$, i.e. the heaviest tail in the market.
+
+However, this is not necessarily the case when $b > 0$. For example,
+consider the case when $q = 0$ and $\hat\phi = 1$. This may be
+viewed an approximation to the situation where the interest rate $r$ is
+vanishingly low while the investor is more conservative than
+greedy. In this case we have
+\begin{eqnarray*}
+  \mathcal U(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
+\end{eqnarray*}
+Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
+$\opd{\nu}\mathcal U(F, \phi) > 0$. In this extreme case, the investor
+would rather seek the largest $\nu$ or in other words, the lightest
+tail in the market.
+
+%   where
+%   \[
+%   y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
+%   \]
+%   Now notice
+%   \[
+%   \ln(y e^{-y}) = \ln(y) - y
+%   \]
+%   is a decreasing function. Thus $-y e^{-y}$ is an increasing
+%   function. Hence we have
+%   \begin{eqnarray*}
+%     W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
+%   \end{eqnarray*}
+%   Now it is clear
+%   \[
+%   \nu\exp\left\{
+%   W\left[
+%     -\left(1 + {1 \over \nu}\right)
+%     e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+%   \right]
+%   + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+%   &=& W(-y e^{-y}) + y
+% \end{eqnarray*}
 \section{Review of multivariate GARCH models:
@@
 \section{Proof of lemma \ref{lemma:II}}
+\begin{proof}
+  Straightforward computation gives
+  \begin{eqnarray}
+    \pd{f(x, \nu)}{\nu} &=& {
+      c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
+      -
+      \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
+      \over
+      2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
+    } \nonumber \\
+    &:=& {
+      P(x^2, \nu)
+      \over
+      2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
+    }
+    \label{eq:xxie4.1}
+  \end{eqnarray}
+
+  While the denominator of the right side of $\pd{f(x, \nu)}{\nu}$ is
+  always positive, its nummerator $P(x^2, \nu)$ has a single root:
+  \begin{equation}
+    \label{eq:xxie5}
+    x_0^2 = \nu\exp\left\{
+      W\left[
+        -\left(1 + {1 \over \nu}\right)
+        e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+      \right]
+      + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
+    \right\} - \nu
+  \end{equation}
+  where $W(\cdot)$ is the principle branch of the Lambert $W$
+  function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
+  the right side of \eqref{eq:xxie5} for positivity, we first note
+  $c'(\nu) > 0$:
+  \[
+  c'(\nu) = {
+    \pi \Gamma(\nu/2 + 1/2) \left\{
+      \nu \left[ \Psi(\nu/2 + 1/2) - \Psi(\nu/2)\right] - 1
+    \right\}
+    \over
+    2 \Gamma(\nu/2) (\pi \nu)^{3/2}
+  }
+  \]
+  where $\Psi(\cdot)$ is the digamma function:
+  \[
+  \Psi(x) = \td{\ln[\Gamma(x)]}{x}
+  \]
+  \begin{minipage}{0.48\textwidth}
+    $\Psi(x)$ \linebreak
+    \includegraphics[width=\textwidth]{digamma.png}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\textwidth}
+    As shown in the figure to the left, $\Psi(x)$ is increasing
+    for $x > 0$. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
+    So we have
+    \begin{eqnarray*}
+      && \nu \left[
+        \Psi(\nu/2 + 1/2) - \Psi(\nu/2)
+      \right] - 1 \\
+      &\geq& 1 \times \left[
+        \Psi(1/2 + 1/2) - \Psi(1/2)
+      \right] - 1 \\
+      &=& \ln(4) - \ln(e) \\
+      &>& 0
+    \end{eqnarray*}
+    Thus $c'(\nu) > 0$. Furthermore, we recall
+    $W(\cdot)$ is increasing on its principle branch. So
+  \end{minipage}
+  \begin{eqnarray*}
+    &&
+    W\left[
+      -\left( 1 + {1 \over \nu} \right)
+      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+    \right]
+    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+    &>& 
+    W\left[
+      -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
+      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+    \right]
+    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+    &=& W(-y e^{-y}) + y
+  \end{eqnarray*}
+  where
+  \[
+  y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
+  \]
+  Now notice
+  \[
+  \ln(y e^{-y}) = \ln(y) - y
+  \]
+  is a decreasing function for $y > 1$. Thus $-y e^{-y}$ is an
+  increasing function. Hence we have
+  \begin{eqnarray*}
+    W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
+  \end{eqnarray*}
+  Now it is clear
+  \[
+  \nu\exp\left\{
+    W\left[
+      -\left(1 + {1 \over \nu}\right)
+      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+    \right]
+    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
+  \right\} - \nu > 0
+  \]
+  Now that we have established that $\pd{f}{\nu}(x, \nu) = 0$ has a
+  single positive root, it remains to determine the sign of
+  $\pd{f}{\nu}(x, \nu)$ on the two sides of the root. For this purpose
+  we observe
+  \begin{equation}
+    \label{eq:xxie5.1}
+    P(0, \nu) = 2 \nu^2 c'(\nu) > 0
+  \end{equation}
+  So we want to investigate $\pd{P}{x}(x, \nu)$:
+  \begin{eqnarray}
+    \pd{P}{x}(x, \nu) &=&
+    2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \ln\left(
+      1 + {x \over \nu}
+    \right)
+    \label{eq:xxie6}
+  \end{eqnarray}
+  Clearly, $\pd{P}{x}(0, \nu) > 0$. Hence from \eqref{eq:xxie5.1} and
+  \eqref{eq:xxie6} it is clear
+  \begin{equation*}
+    \text{sign}\left[
+      \pd{f}{\nu}(x, \nu)
+    \right]
+    = \left\{
+    \begin{array}{rl}
+      1 & 0 < x < x_0 \\
+      -1 & x > x_0
+    \end{array}
+    \right.
+  \end{equation*}
+  where $x_0$ is the positive root of \eqref{eq:xxie5}.
+\end{proof}
 

### 2017-03-16 cpp/ImportanceSampling/GarchPQ.cpp
Wrote psuedo code.

@@ -60,9 +60,12 @@ double estimateLambda(const vector<double>& a,
 		 {
 		     x = unif(gen);
 		 });
-//	E_prev[i] = normalise(E_prev[i], "inf");
+	auto n = norm(E_prev[i], "Inf");
+	E_prev[i] /= n;
     }
     double Lambda = 0;
+    // According to me
+    // vector<double> Lambda_samples(K, 1);
     sd = 0;
     for (unsigned j = 0; j < N; j++) {
 	uniform_real_distribution<double> unif;
@@ -87,22 +90,41 @@ double estimateLambda(const vector<double>& a,
 #pragma omp critical
 	    U = unif(gen) * Q.back();
 	    unsigned l = upper_bound(Q.begin(), Q.end(), U) - Q.begin();
-//	    E[k] = normalise(A[k] * E_prev[l], "inf");
 	    E[k] = A[k] * E_prev[l];
 	    alpha[k] = norm(E[k], "inf");
-	    E[k] = E[k] / alpha[k];
+	    E[k] /= alpha[k];
 	    alpha[k] = pow(alpha[k], theta);
+
+	    // according to me
+	    // Lambda_samples[k] *= log(alpha[k]);
 	}
-	copy(E.begin(), E.end(), E_prev.begin());	
-	double lbt = log(Q.back()/K);
-	Lambda += lbt / N;
+	copy(E.begin(), E.end(), E_prev.begin());
+	// According to Anand
+	// double lbt = log(Q.back()/K);
+	Lambda += log(Q.back()/K) / N;
 	sd += pow(lbt, 2) / N;
     }
+    // According to Anand
     double lbt = log(accumulate(alpha.begin(), alpha.end(), 0.0)/K);
     Lambda += lbt/N;
     sd += pow(lbt, 2)/N;
     sd = sqrt(sd - pow(Lambda, 2));
     return Lambda;
+
+    // according to me
+    // Lambda = 
+    // accumulate(Lambda_samples.begin(), Lambda_samples.end(), 0.0,
+    // 	       [=](double s, double x) {
+    // 		   return s + x/K;
+    // 	       });
+    // sd =
+    // accumulate(Lambda_samples.begin(), Lambda_samples.end(), 0.0,
+    // 	       [=](double s, double x) {
+    // 		   double t = exp(x) - Lambda;
+    // 		   return s + t * t/K;
+    // 	       });
+    // sd = sqrt(sd);
[…]

### 2017-03-16 cpp/ImportanceSampling/plot_data.r

@@ -12,18 +12,20 @@ graphics.off();
 
 ## old <- read.table(file="./lambda0.txt");
 ## new <- read.table(file="./lambda0_new.txt");
-data <- read.table(file="Lambda.txt");
-pdf("Lambda.pdf")
-plot(data$V1, data$V2,
+data <- read.table("SP500_garch21_tailindex.dat");
+## data <- read.table("./test.dat");
+pdf("../../papers/Jeffrey1/SP500_xi.pdf");
+plot(data[, 1], data[, 2],
      ##     main=expression(sigma[t]^2 == 0.11 * R[t-1]^2 + 0.88 * sigma[t-1]^2 + 10^{-7}
      type="l", xlab=expression(alpha),
      ylab=expression(Lambda(alpha))
      );
+grid();
 ## lines(new$V1, new$V2, col="#00FF00");
 ## legend("topleft",
 ##        legend=c("old", "new"),
 ##        col=c("#000000", "#00FF00"),
 ##        lwd=c(1, 1),
 ##        );
-grid();
+
 dev.off();

### 2017-03-17 cpp/ImportanceSampling/GarchPQ.cpp
Some revisions.

@@ -11,7 +11,7 @@
 using namespace std;
 using namespace arma;
 
-// random_device gen;
+random_device gen;
 unsigned long SEED;
 
 mat& gen_rand_matrix(const vector<double> &alpha,
@@ -53,7 +53,7 @@ double estimateLambda(const vector<double>& a,
 #pragma omp parallel for
     for (unsigned i = 0; i < K; i++) {
 	uniform_real_distribution<double> unif;
-	mt19937 gen(i + K + SEED);
+	// mt19937 gen(i + K + SEED);
 	E_prev[i].set_size(a.size() + b.size() - 2);
 	for_each(E_prev[i].begin(), E_prev[i].end(),
 		 [&](double &x)
@@ -70,7 +70,7 @@ double estimateLambda(const vector<double>& a,
     for (unsigned j = 0; j < N; j++) {
 	uniform_real_distribution<double> unif;
 	chi_squared_distribution<double> chi2;
-	mt19937 gen(j + SEED);
+	// mt19937 gen(j + SEED);
 
 	vector<mat> A(K);
 	vector<double> Q(K);
@@ -79,7 +79,7 @@ double estimateLambda(const vector<double>& a,
 #pragma omp parallel for schedule(dynamic) shared(gen, chi2)
 	for (unsigned k = 0; k < K; k++) {
 	    double z2;
-#pragma omp critical
+// #pragma omp critical
 	    z2 = chi2(gen);
 	    gen_rand_matrix(a, b, z2, A[k]);
 	}
@@ -87,7 +87,7 @@ double estimateLambda(const vector<double>& a,
 #pragma omp parallel for shared(gen, unif)
 	for (unsigned k = 0; k < K; k++) {
 	    double U;
-#pragma omp critical
+// #pragma omp critical
 	    U = unif(gen) * Q.back();
 	    unsigned l = upper_bound(Q.begin(), Q.end(), U) - Q.begin();
 	    E[k] = A[k] * E_prev[l];
@@ -100,8 +100,8 @@ double estimateLambda(const vector<double>& a,
 	}
 	copy(E.begin(), E.end(), E_prev.begin());
 	// According to Anand
-	// double lbt = log(Q.back()/K);
-	Lambda += log(Q.back()/K) / N;
+	double lbt = log(Q.back()/K);
+	Lambda += lbt / N;
 	sd += pow(lbt, 2) / N;
     }
     // According to Anand
@@ -187,8 +187,8 @@ int main(int argc, char*argv[])
[…]

### 2017-03-17 papers/FX/ParetoDistr_Preference.r

@@ -1,0 +1,22 @@
+rm(list=ls());
+
+r <- 0.05;
+b <- 1;
+p <- 0.5;
+phi <- 0.5;
+
+utility <- function(x) {
+    return(log(x));
+}
+
+preference <- function(alpha, K, phi) {
+    f <- function(x) {
+        u <- log((1 - phi) * exp(r) + phi * exp(x));
+        y <- u * (1 + b)/(K - x)^(alpha + 1);
+        return(y);
+    }
+    y <- integrate(f, -Inf, 0);
+    y <- alpha * K^alpha * p * y;
+    return(y);
+}
+

### 2017-03-21 papers/FX/draft.tex
Some revisions.
There seems to be a jumping behaviour in the optimal phi when X is t-distributed...

@@
   {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx 
-  + \text{const.}
+  - b u(\delta v) F(q)
   % &&
@@
 \end{eqnarray}
-where the last term is a constant independent of $\phi$, $\alpha$ and $K$.
 \begin{theorem}
@@
 \end{proof}
+By numerically integration and optimization with respect to $\phi$,
+the value of $\hat\phi$ can be obtained for given $u(\cdot)$ and given
+values of $K, K', \alpha, \beta$. Figures \ref{fig:phi_hat_pareto5e-1}
+and \ref{fig:phi_hat_pareto4} show the values of $\hat\phi$ for
+$u(C) = -\frac{C^{-\xi}}{\xi}$ with $\xi = 1/2, 4$ respectively.
+$\beta=\alpha$ and $K'=K$ in both figures.
+\begin{figure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}
+    \caption{$\hat\phi$}
+    \label{fig:phi_hat_pareto5e-1}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
+    \caption{$\hat\phi$}
+    \label{fig:phi_hat_pareto4}
+  \end{subfigure}
+  \caption{$\hat\phi$, the optimal equity allocation when the utility
+    function is $u(C) = -{C^{-\xi} \over \xi}$ for $\xi = 1/2, 4$.
+  }
+\end{figure}
+
+\begin{minipage}{0.5\linewidth}
+  \includegraphics[width=\textwidth]{power_utilities.png}
+\end{minipage}\hfill
+\begin{minipage}{0.42\textwidth}
+  As shown to the left, the power utility function with a small
+  $\xi$ grows slower and saturates later than does one with a large
+  $\xi$. It represents an agent that is more tolerant of low
+  consumption and that seeks wealth more aggressively. In other
+  words, he is less risk-avert than is one with a higher $\xi$.
+  Such an agent will therefore invest more heavily on the equity as
+  is shown in figures \ref{fig:phi_hat_pareto5e-1} and
+  \ref{fig:phi_hat_pareto4}.
+\end{minipage}
+Certainly, it rarely happens in practice that the lower- and
+upper-tail indices of an equity are equal. Thus, to verify our result
+that $G(F)$ is increasing with $\alpha$ and decreasing with $K$, we
+need to verify that the 1st integral of \eqref{eq:xxie1.0} with $\phi$
+replaced by $\hat\phi$, i.e.
+\[
+I(\alpha, K) = 
+  \alpha K^\alpha  p
+  \int_{-\infty}^0
+  u\left[ (1 - \hat\phi(\alpha, K)) e^r + \hat\phi(\alpha, K) e^x \right]
+  {1 + b \over (K - x)^{\alpha + 1}} dx
+\]
+is increasing with $\alpha$ and decreasing with $K$. Figures
+\ref{fig:preference_pareto5e-1} and \ref{fig:preference_pareto4}
+show the values of the above integral when $\alpha$ and $K$ take
+a range of values.
+\begin{figure}[htb!]
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
+    \caption{$I(\alpha, K)$}
+    \label{fig:preference_pareto5e-1}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{preference_pareto4.pdf}
+    \caption{$I(\alpha, K)$}
+    \label{fig:preference_pareto4}
+  \end{subfigure}
+  \caption{$I(\alpha, K)$ when the utility function is
+    $u(C) = -{C^{-\xi} \over \xi}$ for $\xi = 1/2, 4$.
+  }
+\end{figure}
 
@@
 
-\subsubsection{log-utility function}
-We first consider the case where the utility function is simply
+We consider the case where the utility function is simply
 $u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of

### 2017-03-22 papers/FX/draft.tex
Rearranged the sections.

@@
     \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}
-    \caption{$\hat\phi$}
+    \caption{$\hat\phi$ when $\xi = 1/2$}
     \label{fig:phi_hat_pareto5e-1}
@@
     \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
-    \caption{$\hat\phi$}
+    \caption{$\hat\phi$ when $\xi = 4$}
     \label{fig:phi_hat_pareto4}
@@
 \]
+We can write $\mathcal U(F, \phi)$ as
+\begin{eqnarray}
+  && \mathcal U(F, \phi) \nonumber \\
+  &=&
+  \int_{0}^{\infty}
+  \left[
+    u(C(x, \phi)) + (1 + b)u(C(-x, \phi))
+  \right]
+  f(x, \nu) dx \nonumber \\
+  &&
+  + b \int_{0}^{q}
+  u(C(x, \phi))
+  f(x, \nu) dx - b u(C(q, \phi)) F(q, \nu)
+  \label{eq:t1}
+\end{eqnarray}
+where $C(\cdot)$ is defined in \eqref{eq:xxie1}.
+
 \begin{lemma}
@@
 
-We consider the case where the utility function is simply
+\subsubsection{log-utility function}
+We first consider the case where the utility function is simply
 $u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
-\eqref{eq:power_utility} as $\gamma \to 0$. Straightforward
-computation gives
-\begin{eqnarray*}
-  && \mathcal U(F, \phi) \\
-  &=&
-  \int_{0}^{\infty}
-  \left[
-    \ln(C(x)) + (1 + b)\ln(C(-x))
-  \right]
-  f(x, \nu) dx
-  + b \int_{0}^{q}
-  \ln(C(x))
-  f(x, \nu) dx \\
-  &&
-  - b \ln(C(q)) F(q, \nu)
-\end{eqnarray*}
-where $C(\cdot)$ is defined in \eqref{eq:xxie1}. Notice that
-when $b = 0$, $\mathcal U(F, \phi)$  reduces to $\E u(C(X))$.
-Also observe $\ln(C(x)C(-x))$ is an increasing function:
-\begin{eqnarray*}
-  \pd{C(x) C(-x)}{x}
+\eqref{eq:power_utility} as $\gamma \to 0$.
+
+First notice that when $b = 0$, $\mathcal U(F, \phi)$ reduces to
+$\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
+Its integrand becomes $\ln[C(x, \phi) C(-x, \phi)]$.
+Now observe $\ln(C(x, \phi)C(-x, \phi))$ is an increasing function of
+$x$:
+\begin{eqnarray*}
+  \pd{C(x, \phi) C(-x, \phi)}{x}
   &=&
@@
 \end{eqnarray*}
-Thus by 1st case of remark \ref{remark:I} and lemma \ref{lemma:II},
+Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
+\ref{lemma:II},
 $\pd{\mathcal U(F, \phi))}{\nu} < 0$ for all $\phi$. Hence all
@@
 tail in the market.
+
+\subsubsection{power-utility function}
+In this section we consider the case when the utility function
+$u(\cdot)$ takes the form
+\[
+u(C) = -{1 \over \xi} C^{-\xi}
+\]
+Another major difference caused by a non-zero value of $b$ is seen in
+the value of $\hat\phi$. When $b = 0$, $\hat\phi$ increases
+continuously with $\nu$ as shown in figure \ref{fig:phi_hat_t5e-1};
+when $b > 0$, $\hat\phi$ instead increases by jumping up at a sequence
+of values of $\nu$. This is shown in figure \ref{fig:phi_hat_b_t5e-1}.
+\begin{figure}[htb!]
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_t5e-1.pdf}
+    \caption{$\hat\phi$ when $b = 0$}
+    \label{fig:phi_hat_t5e-1}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t5e-1.pdf}
+    \caption{$\hat\phi$ when $b = 0.01$}
+    \label{fig:phi_hat_b_t5e-1}
+  \end{subfigure}
+  \caption{$\hat\phi$, the optimal equity allocation when the utility
+    function is $u(C) = -{2 \over \sqrt{C}}$
+  }
+\end{figure}
 
@@
 
-\section{Hill Estimator}
+\section{Data analysis}
+\subsection{Hill Estimator}
 \label{sec:Hill}
@@
 
-%% \subsubsection{The Energy and the Consumer Staples sectors of S\&P 500}
 Figure \ref{fig:Energy_lower},
@@
   \centering
-  \includegraphics[width=\textwidth]{Energy_lower.pdf}
-  \caption{Hill estimates of the lower tail indices of 32 stocks'
-    daily return series in the {\it Energy} sector of Standard \& Poor
+  \begin{subfigure}[b]{\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+                    {Energy_lower.pdf}
+    \caption{``Energy'' sector of S\&P 500}
+    \label{fig:Energy_lower}
+  \end{subfigure}
+  \begin{subfigure}[b]{\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+                    {Consumer_Staples_lower.pdf}
+    \caption{``Consumer Staples'' sector of S\&P 500}
+    \label{fig:Consumer_Staples_lower}
+  \end{subfigure}
+  \begin{subfigure}[b]{\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+                    {Information_Technology_lower.pdf}
+    \caption{``Information Technology'' sector of S\&P 500}
+    \label{fig:Information_Technology_lower}
+  \end{subfigure}
+  \caption{\small Hill estimates of lower tail indices of
+    daily return series in sectors of Standard \& Poor
     500 index. Each circle coresponds to a Hill estimate; the grey
@@
     from 2010-01-01 to 2015-01-01 and comprise $n=1304$
-    observations. $k=(\ln n)^2 = 49$ lower order statistics are used
+    observations. $k=50$ lower order statistics are used
     for computing $\alpha_n$. 
   }
-  \label{fig:Energy_lower}
 \end{figure}
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\textwidth]{Consumer_Staples_lower.pdf}
-  \caption{Hill estimates of the lower tail indices of 30 stocks'
-    daily return series in the {\it Consumer Staples} sector of
-    Standard \& Poor 500 index.
-  }
-  \label{fig:Consumer_Staples_lower}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\textwidth]{Information_Technology_lower.pdf}
-  \caption{Hill estimates of the lower-tail indices of 42 stocks' daily
-    return series in the {\it Information Technology} sector of S\&P
-    500 index.
-  }
-  \label{fig:Information_Technology_lower}
-\end{figure}
-
-\section{Hybrid estimation of the assumed distribution function}
+\subsection{Hybrid fitting of the shifted Pareto distribution}
 \label{sec:hybrid_estimation}
@@
 
-\section{Pickand's Estimator}
+\subsection{Pickand's Estimator}
 Pickands (1975) \cite{pickands1975statistical} proposed a different

### 2017-03-22 papers/FX/draft.tex
Revised the scripts to add error bars.

@@
 the likelihood function with fixed $\alpha$. Obviously we take only
-negative observations and hence condition on $X < 0$.
+negative observations and hence condition on $X < 0$. Regularity
+conditions on the maximum likelihood estimator are easily checked for
+the distribution \eqref{eq:X_distr}.
+The density function of $X$ conditional on $X < 0$ is
+\[
+f_L(x) = {\alpha K^{\alpha} \over (K - x)^{\alpha + 1}}
+\]
+Clearly, the log-likelihood function of the $K$ parameter is
+\[
+l(K | X < 0) = \ln(\alpha) + \alpha \ln(K) - (\alpha + 1)\ln(K - X)
+\]
+where, as is just mentioned, we pre-estimate $\alpha$ with Hill
+estimator and regard it as known at this step. With $n$ observations
+$X_1, \dots, X_n$, the log-likelihood function of $K$ is
+\[
+L(K | \forall i=1,2,...,n, X_i < 0) = n \ln(\alpha) + n \alpha \ln(K)
+- (\alpha + 1) \sum_{i=1}^n \ln(K - X_i)
+\]
+To obtain the maximum likelihood estimation, we need to solve the
+equation $\pd{L}{K} = 0$, which expands to
+\[
+{n \alpha \over K} - (\alpha + 1) \sum_{i=1}^n {1 \over K - X_i} = 0
+\]
+This equation has to be solved numerically.
+The Fisher information of $K$ follows as
+\begin{eqnarray*}
+  I(K) &=& \E\left[ \left( \opd{K} l(K | X < 0) \right)^2 \right] \\
+  &=& \alpha K^{\alpha - 2}
+  \int_{-\infty}^0 {(\alpha x + K)^2 \over (K - x)^{\alpha + 3}} dx \\
+  &=& {\alpha \over (2 + \alpha) K^2}
+\end{eqnarray*}
+So the maximum likelihood estimator, call it $\hat K$, is
+asymptotically normal:
+\[
+\sqrt n (\hat K - K) \to N \left(
+  0, \left( 1 + {2 \over \alpha} \right) K^2
+\right)
+\]
 
@@
 
-\subsection{Pickand's Estimator}
-Pickands (1975) \cite{pickands1975statistical} proposed a different
-approach to estimate the tail index than that of Hill. Assume
-$\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
-denote the quantile of $1 - 1/t$ of this distribution, i.e.
-\[
-P(X > U(t)) = {1 \over t}
-\]
-The key idea of Pickands' estimator is to observe
-\[
-\lim_{t \to \infty} {
-  U(c(t) t) - U(t)
-  \over
-  U(t) - U(t/c(t))
-} = 2^{1/\alpha}
-\]
-for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
-Pickands formulated the estimation problem in terms of the extreme
-value index $\xi$, which is the reciprocal of the tail index $\alpha$
-when the distribution of $X$ is regularly varying (cf. Embrechtsm,
-Kl\"uppelberg and Mikosch \cite{Embrechts1997}).
-
-Building upon the aforementioned observation and taking $c(t)$ to
-be simply constant 2, Pickands proposed the following estimator for
-$\xi$:
-\[
-\hat \xi = {1 \over \ln 2} \ln {
-  x_{(k)} - x_{(2k)}
-  \over
-  x_{(2k)} - x_{(4k)}
-}
-\]
-Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
-that Pickands' estimator was consistent in both the weak and the strong
-sense, when the sequence $k(n)$ was chosen appropriately:
-\begin{itemize}
-\item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
-  \[
-  \hat \xi \overset{P}{\to} \xi
-  \]
-\item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
-  as $n \to \infty$,
-  \[
-  \hat \xi \overset{a.s.}{\to} \xi
-  \]
-\end{itemize}
-Dekkers and de Haan \cite{dekkers1989estimation} also showed the
-asymptotic normality of the estimator, when some
-extra technical conditions were satisfied by
-$U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
-\[
-\sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
-\]
-where
-\[
-v(\xi) = {
-  \xi^2 (2^{2 \xi + 1} + 1)
-  \over
-  [2 \ln(2) (2^\xi - 1)]^2
-}
-\]
-Using Pickands' estimator, the same result of very similar tail indices
-is again obtained for the {\it Energy} and the {\it Consumer Staples}
-sectors. Figure \ref{fig:Energy_Pickands} and
-\ref{fig:Consumer_Staples_Pickands} illustrate this claim.
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
-  \caption{Pickands' estimates of {\it Energy} stocks.}
-  \label{fig:Energy_Pickands}
-\end{figure}
-
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
-  \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
-  \label{fig:Consumer_Staples_Pickands}
-\end{figure}
+% \subsection{Pickand's Estimator}
+% Pickands (1975) \cite{pickands1975statistical} proposed a different
+% approach to estimate the tail index than that of Hill. Assume
+% $\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
+% denote the quantile of $1 - 1/t$ of this distribution, i.e.
+% \[
+% P(X > U(t)) = {1 \over t}
+% \]
+% The key idea of Pickands' estimator is to observe
+% \[
+% \lim_{t \to \infty} {
+%   U(c(t) t) - U(t)
+%   \over
+%   U(t) - U(t/c(t))
+% } = 2^{1/\alpha}
+% \]
+% for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
+% Pickands formulated the estimation problem in terms of the extreme
+% value index $\xi$, which is the reciprocal of the tail index $\alpha$
+% when the distribution of $X$ is regularly varying (cf. Embrechtsm,
+% Kl\"uppelberg and Mikosch \cite{Embrechts1997}).
+
+% Building upon the aforementioned observation and taking $c(t)$ to
+% be simply constant 2, Pickands proposed the following estimator for
+% $\xi$:
+% \[
+% \hat \xi = {1 \over \ln 2} \ln {
+%   x_{(k)} - x_{(2k)}
+%   \over
+%   x_{(2k)} - x_{(4k)}
+% }
+% \]
+% Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
+% that Pickands' estimator was consistent in both the weak and the strong
+% sense, when the sequence $k(n)$ was chosen appropriately:
+% \begin{itemize}
+% \item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
+%   \[
+%   \hat \xi \overset{P}{\to} \xi
+%   \]
+% \item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
+%   as $n \to \infty$,
+%   \[
+%   \hat \xi \overset{a.s.}{\to} \xi
+%   \]
+% \end{itemize}
+% Dekkers and de Haan \cite{dekkers1989estimation} also showed the
+% asymptotic normality of the estimator, when some
+% extra technical conditions were satisfied by
+% $U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
+% \[
+% \sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
+% \]
+% where
+% \[
+% v(\xi) = {
+%   \xi^2 (2^{2 \xi + 1} + 1)
+%   \over
+%   [2 \ln(2) (2^\xi - 1)]^2
+% }
+% \]
+% Using Pickands' estimator, the same result of very similar tail indices
+% is again obtained for the {\it Energy} and the {\it Consumer Staples}
+% sectors. Figure \ref{fig:Energy_Pickands} and
+% \ref{fig:Consumer_Staples_Pickands} illustrate this claim.
+% \begin{figure}[htb!]
+%   \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
+%   \caption{Pickands' estimates of {\it Energy} stocks.}
+%   \label{fig:Energy_Pickands}
+% \end{figure}
+
+% \begin{figure}[htb!]
+%   \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
+%   \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
+%   \label{fig:Consumer_Staples_Pickands}
+% \end{figure}
 

### 2017-03-22 r/myLogLik.r

@@ -2,43 +2,43 @@ rm(list=ls());
 require(MASS);
 source("libxxie.r")
 #Energy
-## tables <- c (
-##     "APA",
-##     "APC",
-##     "BHI",
-##     "CHK",
-##     "COG",
-##     "COP",
-##     "DO",
-##     "DVN",
-##     "EOG",
-##     "EQT",
-##     "FTI",
-##     "HAL",
-##     "HES",
-##     "HP",
-##     "KMI",
-##     "MPC",
-##     "MRO",
-##     "MUR",
-##     "NBL",
-##     "NFX",
-##     "NOV",
-##     "OKE",
-##     "OXY",
-##     "PSX",
-##     "PXD",
-##     "RIG",
-##     "RRC",
-##     "SE",
-##     "SLB",
-##     "SWN",
-##     "TSO",
-##     "VLO",
-##     "WMB",
-##     "XEC",
-##     "XOM"
-## );
+tables <- c (
+    "APA",
+    "APC",
+    "BHI",
+    "CHK",
+    "COG",
+    "COP",
+    "DO",
+    "DVN",
+    "EOG",
+    "EQT",
+    "FTI",
+    "HAL",
+    "HES",
+    "HP",
+    "KMI",
+    "MPC",
+    "MRO",
+    "MUR",
+    "NBL",
+    "NFX",
+    "NOV",
+    "OKE",
+    "OXY",
+    "PSX",
+    "PXD",
+    "RIG",
+    "RRC",
+    "SE",
+    "SLB",
+    "SWN",
+    "TSO",
+    "VLO",
+    "WMB",
+    "XEC",
+    "XOM"
+);
 
 ## Consumer staples
 ## tables <- c(
@@ -79,50 +79,50 @@ source("libxxie.r")
 ## );
 
 ## Information Technology
-tables <- c(
-    "ADBE",
-    "ADI",
-    "ADP",
-    "ADSK",
-    "AKAM",
-    "AMAT",
-    "CA",
-    "CSCO",
-    "CTSH",
-    "CTXS",
-    "EA",
-    "EBAY",
-    "FFIV",
-    "FISV",
-    "HPQ",
-    "HRS",
-    "IBM",
-    "INTC",
-    "INTU",
-    "JNPR",
-    "KLAC",
-    "LLTC",
-    "LRCX",
-    "MCHP",
-    "MSFT",
-    "MSI",
-    "MU",
-    "NTAP",
[…]

### 2017-03-23 papers/FX/draft.tex
Reshuffled the sections in the FX paper.

@@
 
+\section{Empirical Results}
+\subsection{Hill estimates of lower tail indices}
+\label{sec:Hill}
+There have been several estimators for the tail index proposed in the
+literature, e.g. Hill (1975) \cite{hill1975simple}, Pikands (1975)
+\cite{pickands1975statistical}, Dekkers and Einmahl and De Haan (1989)
+\cite{dekkers1989moment}. See also Embrechts and Kl\"uppelberg and
+Mikosch (1997)\cite{Embrechts1997}. The most popular of these is
+arguably Hill (1975) \cite{hill1975simple}:
+\[
+\hat \alpha = \left[
+  {1 \over k} \sum_{i=1}^k \ln {x_{(i)} \over x_{(k+1)}}
+  \right]^{-1}
+\]
+where $x_{(i)}$ is the $i$-th upper order statistic of a sample
+$x_1, x_2, ..., x_n$. It has been shown that the Hill estimator is
+applicable to 
+\begin{itemize}
+\item iid random variables (Mason (1982) \cite{mason1982laws})
+\item a weakly dependent sequence (Rootz{\'e}n, Leadbetter,
+  and de Haan \cite{rootzen1992tail})
+\item a linear process (Resnick and St{\u{a}}ric{\u{a}} (1995)
+  \cite{resnick1995consistency}).
+\end{itemize}
+In these cases it is shown that, if $k \to \infty$ and $k/n \to 0$
+as $n \to \infty$, the Hill estimator is weakly consistent, i.e.
+\[
+\sqrt k (\hat \alpha - \alpha) \overset{d}{\to} N(0, \alpha^2)
+\]
+
+Figure \ref{fig:Energy_lower},
+\ref{fig:Consumer_Staples_lower},
+and \ref{fig:Information_Technology_lower}
+show the Hill estimates of the
+lower tail indices of the ``energy'', ``consumer staples'' and ``information
+technology'' sectors of the S\&P 500 index, respectively.
+\begin{figure}[htb!]
+  \centering
+  \begin{subfigure}[b]{\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+                    {Energy_lower.pdf}
+    \caption{``Energy'' sector of S\&P 500}
+    \label{fig:Energy_lower}
+  \end{subfigure}
+  \begin{subfigure}[b]{\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+                    {Consumer_Staples_lower.pdf}
+    \caption{``Consumer Staples'' sector of S\&P 500}
+    \label{fig:Consumer_Staples_lower}
+  \end{subfigure}
+  \begin{subfigure}[b]{\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+                    {Information_Technology_lower.pdf}
+    \caption{``Information Technology'' sector of S\&P 500}
+    \label{fig:Information_Technology_lower}
+  \end{subfigure}
+  \caption{\small Hill estimates of lower tail indices of
+    daily return series in sectors of Standard \& Poor
+    500 index. Each circle coresponds to a Hill estimate; the grey
+    triangles above and below it mark the 97.5\% and 2.5\% quantiles
+    of its approximate normal distribution.
+    The low and high red lines mark the median of the 2.5\% quantiles
+    and median of the 97.5\% quantiles, respectively.
+    The data are from {\it Yahoo Finance} and the labels on
+    the horizontal axis are Yahoo symbols of the stocks. The data span
+    from 2010-01-01 to 2015-01-01 and comprise $n=1304$
+    observations. $k=50$ lower order statistics are used
+    for computing $\alpha_n$. 
+  }
+\end{figure}
+
+\section{Review of multivariate GARCH models:
+  same tail index or not}
+Many multivariate GARCH models imply a common tail index for all
+series under consideration, e.g. the Orthogonal GARCH model of
+Alexander and Chibumba (1997) \cite{alexander1997multivariate}, its
+generalization GO-GARCH of van der Weide (2002) \cite{van2002go}, as
+well as the Full Factor GARCH model of Vrontos et al (2003)
+\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
+model of Lanne and Saikkonen (2007) \cite{lanne2007modeling}. These
+models are characterized by their treatment of each series as a linear
+combination of factors, and each of the factors is modeled as a GARCH
+process. As a result of this treatment, each and every series under
+consideration necessarily has the lowest of the tail indices of its
+contributing factors (cf. Mikosch (2013) \cite{Mikosch2013} as well
+as Andersen, Davis, Kreiss and Mikosch (2009)
+\cite{andersen2009handbook}).
+
+In contrast, several other multivariate volatility models imply, in
+general, different tail indices for different series. For example, the
+Constant Conditional Correlation (CCC-) GARCH model of Bollerslev
+(1990) \cite{bollerslev1990modelling} specifies, for a $d$-dimensional
+series $\vec X_t$
+\begin{eqnarray}
+  \cov(X_{i,t}, X_{j,t}) &=& \sqrt{h_{i,t} h_{j,t}} \rho_{i,j}
+  \nonumber \\
+  h_{i,t} &=& \omega_i + \sum_{l=1}^q A_{i,l} X_{i, t-l}^2 +
+  \sum_{l=1}^p B_{i,l} h_{i, t-l} \label{eq:CCC-GARCH}
+\end{eqnarray}
+Consider the simple case when $p=q=1$. Assume
+$X_{i, t} = \sqrt h_{i,t} Z_{i,t}$, $Z_{i, t}$ are iid for
+all $t \in \mathbb Z$, and $\P(|Z_{i,t}| > u)/\P(\sqrt h_{i,t} > u) \to 0$ as
+$u \to \infty$. Then the tail index of the stationary distribution of
+$X_{i,t}$, call it $\kappa_i$, is given by
+\[
+\E (A_{i, 1} Z^2 + B_{i, 1})^{\kappa_i / 2} = 1
+\]
+where $Z \overset{d}{=} Z_{i,t}$.
+Provided $(A_{i,1}, B_{i,1}) \neq (A_{j,1}, B_{j,1})$, it is clear
+$\kappa_i \neq \kappa_j$.
+
+As an extension to the CCC-GARCH model, Jeantheau (1998)
+\cite{jeantheau1998strong} proposed his Extended CCC-GARCH by
+modifying \eqref{eq:CCC-GARCH} to
+\begin{equation}
+  \label{eq:ECCC-GARCH}
+  \vec h_t =
+  \sum_{l=1}^q \mtx A_{l} (\vec X_{t-l} \odot \vec X_{t-l})
+  +
+  \sum_{l=1}^p \mtx B_{l} \vec h_{t-l}
+  +
+  \vec \omega
+\end{equation}
+where $\mtx A_l$ and $\mtx B_l$ are non-diagonal $d \times d$ matrices
+and $\odot$ denotes element-wise multiplication. In the simple case $p=q=1$,
+equation \eqref{eq:ECCC-GARCH} reduces to
+\begin{eqnarray*}
+  h_{i,t} &=&
+  \sum_{j=1}^d A_{i,j} X_{j, t-1}^2
+  + \sum_{j=1}^d B_{i,j} h_{j, t-1}
+  + \omega_i
+\end{eqnarray*}
+or in vector form
+\begin{eqnarray}
+  \vec h_t &=& \left[
+    \mtx A \diag(Z_{1, t-1}^2, \dots, Z_{d, t-1}^2) + \mtx B
+    \right] \vec h_{t-1} + \vec \omega
+  \label{eq:ECCC-GARCH2}
+\end{eqnarray}
+On appropriate conditions, \eqref{eq:ECCC-GARCH2} implies the
+components of $\vec h_t$ are jointly regularly varying with some index
+$\kappa$ as $t \to \infty$. This is a direct result of the Kesten-Goldie
+theorem (cf. Kesten (1973)\cite{Kesten1973}). Thus all series under
+consideration have the same tail index $\kappa$.
+
+In summary, whether or not a common tail index is shared by all
+series is a dividing issue for multivariate GARCH
+models. Nevertheless, economic theoretical arguments are supportive of
+a common tail index for all series. Therefore, it is important to
+gather statistical evidence from financial time series if the issue is
+to be settled definitely.
+
 \section{Generalized disappointment aversion}
@@
 
-\subsection{Student's t-distributed equity return}
-It is a common practice to use Student's t-distribution to model the
-stationary distribution of equity returns. So it is of interest to
-find out what implications this distribution has when it is combined
-with the PDA preference. Formally we assume
-\[
-f(x; \nu) = c(\nu) \left(
-  1 + {x^2 \over \nu}
-\right)^{-(\nu + 1)/2}
-\]
-where $\nu > 1$ and
-\[
-c(\nu) = {
-  \Gamma({\nu + 1 \over 2})
-  \over
-  \Gamma(\nu/2) \sqrt{\nu \pi}
-}
-\]
-We can write $\mathcal U(F, \phi)$ as
-\begin{eqnarray}
-  && \mathcal U(F, \phi) \nonumber \\
-  &=&
-  \int_{0}^{\infty}
-  \left[
-    u(C(x, \phi)) + (1 + b)u(C(-x, \phi))
-  \right]
-  f(x, \nu) dx \nonumber \\
-  &&
-  + b \int_{0}^{q}
-  u(C(x, \phi))
-  f(x, \nu) dx - b u(C(q, \phi)) F(q, \nu)
-  \label{eq:t1}
-\end{eqnarray}
-where $C(\cdot)$ is defined in \eqref{eq:xxie1}.
-
-\begin{lemma}
-  \label{lemma:II}
-  There exists $x_0 > 0$ such that $\pd{f}{\nu}(x_0, \nu) = 0$ and
-  $\forall x \in (0, x_0), \pd{f}{\nu}(x, \nu) > 0$ and
-  $\forall x \in (x_0, \infty), \pd{f}{\nu}(x, \nu) < 0$.
-\end{lemma}
-
-\subsubsection{log-utility function}
-We first consider the case where the utility function is simply
-$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
-\eqref{eq:power_utility} as $\gamma \to 0$.
-
-First notice that when $b = 0$, $\mathcal U(F, \phi)$ reduces to
-$\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
-Its integrand becomes $\ln[C(x, \phi) C(-x, \phi)]$.
-Now observe $\ln(C(x, \phi)C(-x, \phi))$ is an increasing function of
-$x$:
-\begin{eqnarray*}
-  \pd{C(x, \phi) C(-x, \phi)}{x}
-  &=&
-  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
-\end{eqnarray*}
-Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
-\ref{lemma:II},
-$\pd{\mathcal U(F, \phi))}{\nu} < 0$ for all $\phi$. Hence all
-investors guided by this utility function would be seeking the
-smallest $\nu$, i.e. the heaviest tail in the market.
-
-However, this is not necessarily the case when $b > 0$. For example,
-consider the case when $q = 0$ and $\hat\phi = 1$. This may be
-viewed an approximation to the situation where the interest rate $r$ is
-vanishingly low while the investor is more conservative than
-greedy. In this case we have
-\begin{eqnarray*}
-  \mathcal U(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
-\end{eqnarray*}
-Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
-$\opd{\nu}\mathcal U(F, \phi) > 0$. In this extreme case, the investor
-would rather seek the largest $\nu$ or in other words, the lightest
-tail in the market.
-
-\subsubsection{power-utility function}
-In this section we consider the case when the utility function
-$u(\cdot)$ takes the form
-\[
-u(C) = -{1 \over \xi} C^{-\xi}
-\]
-Another major difference caused by a non-zero value of $b$ is seen in
-the value of $\hat\phi$. When $b = 0$, $\hat\phi$ increases
-continuously with $\nu$ as shown in figure \ref{fig:phi_hat_t5e-1};
-when $b > 0$, $\hat\phi$ instead increases by jumping up at a sequence
-of values of $\nu$. This is shown in figure \ref{fig:phi_hat_b_t5e-1}.
-\begin{figure}[htb!]
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_t5e-1.pdf}
-    \caption{$\hat\phi$ when $b = 0$}
-    \label{fig:phi_hat_t5e-1}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t5e-1.pdf}
-    \caption{$\hat\phi$ when $b = 0.01$}
-    \label{fig:phi_hat_b_t5e-1}
-  \end{subfigure}
-  \caption{$\hat\phi$, the optimal equity allocation when the utility
-    function is $u(C) = -{2 \over \sqrt{C}}$
-  }
-\end{figure}
-
-%   where
-%   \[
-%   y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
-%   \]
-%   Now notice
-%   \[
-%   \ln(y e^{-y}) = \ln(y) - y
-%   \]
-%   is a decreasing function. Thus $-y e^{-y}$ is an increasing
-%   function. Hence we have
-%   \begin{eqnarray*}
-%     W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
-%   \end{eqnarray*}
-%   Now it is clear
-%   \[
-%   \nu\exp\left\{
-%   W\left[
-%     -\left(1 + {1 \over \nu}\right)
-%     e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
-%   \right]
-%   + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
-%   &=& W(-y e^{-y}) + y
-% \end{eqnarray*}
-\section{Review of multivariate GARCH models:
-  same tail index or not}
-Many multivariate GARCH models imply a common tail index for all
-series under consideration, e.g. the Orthogonal GARCH model of
-Alexander and Chibumba (1997) \cite{alexander1997multivariate}, its
-generalization GO-GARCH of van der Weide (2002) \cite{van2002go}, as
-well as the Full Factor GARCH model of Vrontos et al (2003)
-\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
-model of Lanne and Saikkonen (2007) \cite{lanne2007modeling}. These
-models are characterized by their treatment of each series as a linear
-combination of factors, and each of the factors is modeled as a GARCH
-process. As a result of this treatment, each and every series under
-consideration necessarily has the lowest of the tail indices of its
-contributing factors (cf. Mikosch (2013) \cite{Mikosch2013} as well
-as Andersen, Davis, Kreiss and Mikosch (2009)
-\cite{andersen2009handbook}).
-
-In contrast, several other multivariate volatility models imply, in
-general, different tail indices for different series. For example, the
-Constant Conditional Correlation (CCC-) GARCH model of Bollerslev
-(1990) \cite{bollerslev1990modelling} specifies, for a $d$-dimensional
-series $\vec X_t$
-\begin{eqnarray}
-  \cov(X_{i,t}, X_{j,t}) &=& \sqrt{h_{i,t} h_{j,t}} \rho_{i,j}
-  \nonumber \\
-  h_{i,t} &=& \omega_i + \sum_{l=1}^q A_{i,l} X_{i, t-l}^2 +
-  \sum_{l=1}^p B_{i,l} h_{i, t-l} \label{eq:CCC-GARCH}
-\end{eqnarray}
-Consider the simple case when $p=q=1$. Assume
-$X_{i, t} = \sqrt h_{i,t} Z_{i,t}$, $Z_{i, t}$ are iid for
-all $t \in \mathbb Z$, and $\P(|Z_{i,t}| > u)/\P(\sqrt h_{i,t} > u) \to 0$ as
-$u \to \infty$. Then the tail index of the stationary distribution of
-$X_{i,t}$, call it $\kappa_i$, is given by
-\[
-\E (A_{i, 1} Z^2 + B_{i, 1})^{\kappa_i / 2} = 1
-\]
-where $Z \overset{d}{=} Z_{i,t}$.
-Provided $(A_{i,1}, B_{i,1}) \neq (A_{j,1}, B_{j,1})$, it is clear
-$\kappa_i \neq \kappa_j$.
-
-As an extension to the CCC-GARCH model, Jeantheau (1998)
-\cite{jeantheau1998strong} proposed his Extended CCC-GARCH by
-modifying \eqref{eq:CCC-GARCH} to
-\begin{equation}
-  \label{eq:ECCC-GARCH}
-  \vec h_t =
-  \sum_{l=1}^q \mtx A_{l} (\vec X_{t-l} \odot \vec X_{t-l})
-  +
-  \sum_{l=1}^p \mtx B_{l} \vec h_{t-l}
-  +
-  \vec \omega
-\end{equation}
-where $\mtx A_l$ and $\mtx B_l$ are non-diagonal $d \times d$ matrices
-and $\odot$ denotes element-wise multiplication. In the simple case $p=q=1$,
-equation \eqref{eq:ECCC-GARCH} reduces to
-\begin{eqnarray*}
-  h_{i,t} &=&
-  \sum_{j=1}^d A_{i,j} X_{j, t-1}^2
-  + \sum_{j=1}^d B_{i,j} h_{j, t-1}
-  + \omega_i
-\end{eqnarray*}
-or in vector form
-\begin{eqnarray}
-  \vec h_t &=& \left[
-    \mtx A \diag(Z_{1, t-1}^2, \dots, Z_{d, t-1}^2) + \mtx B
-    \right] \vec h_{t-1} + \vec \omega
-  \label{eq:ECCC-GARCH2}
-\end{eqnarray}
-On appropriate conditions, \eqref{eq:ECCC-GARCH2} implies the
-components of $\vec h_t$ are jointly regularly varying with some index
-$\kappa$ as $t \to \infty$. This is a direct result of the Kesten-Goldie
-theorem (cf. Kesten (1973)\cite{Kesten1973}). Thus all series under
-consideration have the same tail index $\kappa$.
-
-In summary, whether or not a common tail index is shared by all
-series is a dividing issue for multivariate GARCH
-models. Nevertheless, economic theoretical arguments are supportive of
-a common tail index for all series. Therefore, it is important to
-gather statistical evidence from financial time series if the issue is
-to be settled definitely.
-
-\section{Data analysis}
-\subsection{Hill Estimator}
-\label{sec:Hill}
-There have been several estimators for the tail index proposed in the
-literature, e.g. Hill (1975) \cite{hill1975simple}, Pikands (1975)
-\cite{pickands1975statistical}, Dekkers and Einmahl and De Haan (1989)
-\cite{dekkers1989moment}. See also Embrechts and Kl\"uppelberg and
-Mikosch (1997)\cite{Embrechts1997}. The most popular of these is
-arguably Hill (1975) \cite{hill1975simple}:
-\[
-\hat \alpha = \left[
-  {1 \over k} \sum_{i=1}^k \ln {x_{(i)} \over x_{(k+1)}}
-  \right]^{-1}
-\]
-where $x_{(i)}$ is the $i$-th upper order statistic of a sample
-$x_1, x_2, ..., x_n$. It has been shown that the Hill estimator is
-applicable to 
-\begin{itemize}
-\item iid random variables (Mason (1982) \cite{mason1982laws})
-\item a weakly dependent sequence (Rootz{\'e}n, Leadbetter,
-  and de Haan \cite{rootzen1992tail})
-\item a linear process (Resnick and St{\u{a}}ric{\u{a}} (1995)
-  \cite{resnick1995consistency}).
-\end{itemize}
-In these cases it is shown that, if $k \to \infty$ and $k/n \to 0$
-as $n \to \infty$, the Hill estimator is weakly consistent, i.e.
-\[
-\sqrt k (\hat \alpha - \alpha) \overset{d}{\to} N(0, \alpha^2)
-\]
-
-Figure \ref{fig:Energy_lower},
-\ref{fig:Consumer_Staples_lower},
-and \ref{fig:Information_Technology_lower}
-show the Hill estimates of the
-lower tail indices of the ``energy'', ``consumer staples'' and ``information
-technology'' sectors of the S\&P 500 index, respectively.
-\begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{\linewidth}
-    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
-                    {Energy_lower.pdf}
-    \caption{``Energy'' sector of S\&P 500}
-    \label{fig:Energy_lower}
-  \end{subfigure}
-  \begin{subfigure}[b]{\linewidth}
-    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
-                    {Consumer_Staples_lower.pdf}
-    \caption{``Consumer Staples'' sector of S\&P 500}
-    \label{fig:Consumer_Staples_lower}
-  \end{subfigure}
-  \begin{subfigure}[b]{\linewidth}
-    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
-                    {Information_Technology_lower.pdf}
-    \caption{``Information Technology'' sector of S\&P 500}
-    \label{fig:Information_Technology_lower}
-  \end{subfigure}
-  \caption{\small Hill estimates of lower tail indices of
-    daily return series in sectors of Standard \& Poor
-    500 index. Each circle coresponds to a Hill estimate; the grey
-    triangles above and below it mark the 97.5\% and 2.5\% quantiles
-    of its approximate normal distribution.
-    The low and high red lines mark the median of the 2.5\% quantiles
-    and median of the 97.5\% quantiles, respectively.
-    The data are from {\it Yahoo Finance} and the labels on
-    the horizontal axis are Yahoo symbols of the stocks. The data span
-    from 2010-01-01 to 2015-01-01 and comprise $n=1304$
-    observations. $k=50$ lower order statistics are used
-    for computing $\alpha_n$. 
-  }
-\end{figure}
-
 \subsection{Hybrid fitting of the shifted Pareto distribution}
@@
 technology'' sectors of the S\&P 500 index, respectively.
-
 \begin{figure}[htb!]
   \centering
-  \includegraphics[width=\textwidth]{Energy_alpha_K.pdf}
-  \caption{Hybrid estimates of the lower-tail parameters of daily
-    return series in the {\it Energy} sector of S\&P
-    500 index.
-  }
+  \begin{subfigure}[b]{0.45\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Energy_alpha_K.pdf}
+                    \caption{``Energy''}
                     \label{fig:Energy_alpha_K}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Energy_alpha_K_ci.pdf}
+                    \caption{``Energy''}
+                    \label{fig:Energy_alpha_K_ci}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Consumer_Staples_alpha_K.pdf}
+                    \caption{``Consumer Staples''}
+                    \label{fig:Consumer_Staples_alpha_K}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Consumer_Staples_alpha_K_ci.pdf}
+                    \caption{``Information Technology''}
+                    \label{fig:Consumer_Staples_alpha_K_ci}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Information_Technology_alpha_K.pdf}
+                    \caption{``Information Technology''}
+                    \label{fig:Information_Technology_alpha_K}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Information_Technology_alpha_K_ci.pdf}
+                    \caption{``Information Technology''}
+                    \label{fig:Information_Technology_alpha_K_ci}
+  \end{subfigure}
+  \caption{\small Maximum likelihood estimates of $K$, the scale
+    parameter, of stocks in sectors of the S\&P 500 index. The points
+    are the
+    estimated values; the red bars are the upper bound of the 98\%
+    confidence interval. The lower bound is 0.
+  }
 \end{figure}
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\textwidth]{Consumer_Staples_alpha_K.pdf}
-  \caption{Hybrid estimates of the lower-tail parameters of daily
-    return series in the {\it Consumer Staples} sector of S\&P
-    500 index.
-  }
-  \label{fig:Consumer_Staples_alpha_K}
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\textwidth]{Information_Technology_alpha_K.pdf}
-  \caption{Hybrid estimates of the lower-tail parameters of daily
-    return series in the {\it Information Technology} sector of S\&P
-    500 index.
-  }
-  \label{fig:Information_Technology_alpha_K}
-\end{figure}
+%% \begin{figure}[htb!]
+%%   \centering
+%%   \includegraphics[width=\textwidth]{Energy_alpha_K.pdf}
+%%   \caption{Hybrid estimates of the lower-tail parameters of daily
+%%     return series in the {\it Energy} sector of S\&P
+%%     500 index.
+%%   }
+%%   \label{fig:Energy_alpha_K}
+%% \end{figure}
+
+%% \begin{figure}[htb!]
+%%   \centering
+%%   \includegraphics[width=\textwidth]{Consumer_Staples_alpha_K.pdf}
+%%   \caption{Hybrid estimates of the lower-tail parameters of daily
+%%     return series in the {\it Consumer Staples} sector of S\&P
+%%     500 index.
+%%   }
+%%   \label{fig:Consumer_Staples_alpha_K}
+%% \end{figure}
+
+%% \begin{figure}[htb!]
+%%   \centering
+%%   \includegraphics[width=\textwidth]{Information_Technology_alpha_K.pdf}
+%%   \caption{Hybrid estimates of the lower-tail parameters of daily
+%%     return series in the {\it Information Technology} sector of S\&P
+%%     500 index.
+%%   }
+%%   \label{fig:Information_Technology_alpha_K}
+%% \end{figure}
 
@@
 
+\subsection{Student's t-distributed equity return}
+It is a common practice to use Student's t-distribution to model the
+stationary distribution of equity returns. So it is of interest to
+find out what implications this distribution has when it is combined
+with the PDA preference. Formally we assume
+\[
+f(x; \nu) = c(\nu) \left(
+  1 + {x^2 \over \nu}
+\right)^{-(\nu + 1)/2}
+\]
+where $\nu > 1$ and
+\[
+c(\nu) = {
+  \Gamma({\nu + 1 \over 2})
+  \over
+  \Gamma(\nu/2) \sqrt{\nu \pi}
+}
+\]
+We can write $\mathcal U(F, \phi)$ as
+\begin{eqnarray}
+  && \mathcal U(F, \phi) \nonumber \\
+  &=&
+  \int_{0}^{\infty}
+  \left[
+    u(C(x, \phi)) + (1 + b)u(C(-x, \phi))
+  \right]
+  f(x, \nu) dx \nonumber \\
+  &&
+  + b \int_{0}^{q}
+  u(C(x, \phi))
+  f(x, \nu) dx - b u(C(q, \phi)) F(q, \nu)
+  \label{eq:t1}
+\end{eqnarray}
+where $C(\cdot)$ is defined in \eqref{eq:xxie1}.
+
+\begin{lemma}
+  \label{lemma:II}
+  There exists $x_0 > 0$ such that $\pd{f}{\nu}(x_0, \nu) = 0$ and
+  $\forall x \in (0, x_0), \pd{f}{\nu}(x, \nu) > 0$ and
+  $\forall x \in (x_0, \infty), \pd{f}{\nu}(x, \nu) < 0$.
+\end{lemma}
+
+\subsubsection{log-utility function}
+We first consider the case where the utility function is simply
+$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
+\eqref{eq:power_utility} as $\gamma \to 0$.
+
+First notice that when $b = 0$, $\mathcal U(F, \phi)$ reduces to
+$\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
+Its integrand becomes $\ln[C(x, \phi) C(-x, \phi)]$.
+Now observe $\ln(C(x, \phi)C(-x, \phi))$ is an increasing function of
+$x$:
+\begin{eqnarray*}
+  \pd{C(x, \phi) C(-x, \phi)}{x}
+  &=&
+  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
+\end{eqnarray*}
+Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
+\ref{lemma:II},
+$\pd{\mathcal U(F, \phi))}{\nu} < 0$ for all $\phi$. Hence all
+investors guided by this utility function would be seeking the
+smallest $\nu$, i.e. the heaviest tail in the market.
+
+However, this is not necessarily the case when $b > 0$. For example,
+consider the case when $q = 0$ and $\hat\phi = 1$. This may be
+viewed an approximation to the situation where the interest rate $r$ is
+vanishingly low while the investor is more conservative than
+greedy. In this case we have
+\begin{eqnarray*}
+  \mathcal U(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
+\end{eqnarray*}
+Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
+$\opd{\nu}\mathcal U(F, \phi) > 0$. In this extreme case, the investor
+would rather seek the largest $\nu$ or in other words, the lightest
+tail in the market.
+
+\subsubsection{power-utility function}
+In this section we consider the case when the utility function
+$u(\cdot)$ takes the form
+\[
+u(C) = -{1 \over \xi} C^{-\xi}
+\]
+Another major difference caused by a non-zero value of $b$ is seen in
+the value of $\hat\phi$. When $b = 0$, $\hat\phi$ increases
+continuously with $\nu$ as shown in figure \ref{fig:phi_hat_t5e-1};
+when $b > 0$, $\hat\phi$ instead increases by jumping up at a sequence
+of values of $\nu$. This is shown in figure \ref{fig:phi_hat_b_t5e-1}.
+\begin{figure}[htb!]
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_t5e-1.pdf}
+    \caption{$\hat\phi$ when $b = 0$}
+    \label{fig:phi_hat_t5e-1}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t5e-1.pdf}
+    \caption{$\hat\phi$ when $b = 0.01$}
+    \label{fig:phi_hat_b_t5e-1}
+  \end{subfigure}
+  \caption{$\hat\phi$, the optimal equity allocation when the utility
+    function is $u(C) = -{2 \over \sqrt{C}}$
+  }
+\end{figure}
+
+%   where
+%   \[
+%   y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
+%   \]
+%   Now notice
+%   \[
+%   \ln(y e^{-y}) = \ln(y) - y
+%   \]
+%   is a decreasing function. Thus $-y e^{-y}$ is an increasing
+%   function. Hence we have
+%   \begin{eqnarray*}
+%     W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
+%   \end{eqnarray*}
+%   Now it is clear
+%   \[
+%   \nu\exp\left\{
+%   W\left[
+%     -\left(1 + {1 \over \nu}\right)
+%     e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+%   \right]
+%   + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+%   &=& W(-y e^{-y}) + y
+% \end{eqnarray*}
 \section{Conclusion}

### 2017-03-23 papers/FX/draft.tex
Checked the Hoga test again. Use my own implementation.

@@
 \end{figure}
+One can see that all the confidence intervals of the ``Energy'' and
+``Consumer Staples'' sectors overlap. Hence it is informative to test
+the hypothesis that the tail indices of these return series are equal.
+Hoga \cite{Hoga2017} devised tests for this purpose. We adopt his
+approach but use our own implementation. The next section has more
+details on this.
+
+\subsection{Hoga's Test}
+Hoga \cite{Hoga2017} devised tests for the change of a specified
+extreme quantile.
+
+We first test whether there is a change of tail indices in each of two
+series, and if there is no change in either series, we concatenate
+them and test for a change. Figure \ref{fig:Hoga_Energy_Single} shows
+the test statistics of stocks in the ``Energy'' sector of the S\&P 500
+index. Clearly, the null hypothesis, i.e. there is a change in the
+extreme quantile can not be rejected at the specified significance
+level for any of the stocks.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\textwidth]{Hoga_Energy_Single.pdf}
+  \caption{Hoga Test Statistics}
+  \label{fig:Hoga_Energy_Single}
+\end{figure}
+
+% Figure \ref{fig:Hoga_Energy_pair} shows
+% the test statistics of pairs of stocks in the ``Energy'' sector of the
+% S\&P 500 index: We first order the stocks by their names and then
+% concatenate the 1st to the 2nd to form the 1st sequence, the 2nd and
+% the 3rd to form the 2nd sequence, and so forth.
+
+% Clearly, the null hypothesis, i.e. there is a change in the
+% extreme quantile  can not be rejected at the specified significance
+% level for any of the stocks.
 

### 2017-03-28 papers/FX/draft.tex
Added the result of Hoga test.

--- 2017-03-23 papers/FX/draft.tex
@@
 
-%% \newtheorem{mdef}{Definition}
-
-\newcommand{\eqsplit}[2]{
-  \begin{equation}\label{#2}
-    \begin{split}
-      #1
-    \end{split}
-  \end{equation}}
-\newcommand{\eqnsplit}[1]{
-  \begin{eqnarray*}
-    #1
-  \end{eqnarray*}}
-\newcommand{\tran}[1]{
-  \tilde{#1}
-}
-\newcommand{\td}[2]{
-  \frac{d #1}{d #2}
-}
-\newcommand{\pd}[2]{
-  \frac{\partial #1}{\partial #2}
-}
-\newcommand{\ppd}[2]{
-  \frac{\partial^2 #1}{\partial #2^2}
-}
-\newcommand{\pdd}[3]{
-  \frac{\partial^2 #1}{\partial #2 \partial #3}
-}
-\newcommand{\otd}[1]{
-  \frac{d}{d #1}
-}
-\newcommand{\opd}[1]{
-  \frac{\partial}{\partial #1}
-}
-\newcommand{\oppd}[1]{
-  \frac{\partial^2}{\partial #1^2}
-}
-\newcommand{\opdd}[2]{
-  \frac{\partial^2}{\partial #1 \partial #2}
-}
-\newcommand{\ket}[1]{
-  |#1\rangle
-}
-\newcommand{\bra}[1]{
-  \langle#1|
-}
-\newcommand{\inn}[1]{
-  \langle#1\rangle
-}
-\newcommand{\mean}[1]{
-  \langle#1\rangle
-}
-\newcommand{\tr}{
-  \text{tr}\,
-}
-\newcommand{\re}{
-  \text{Re}\,
-}
-\newcommand\im{
-  \text{Im}\,
-}
-\newcommand{\arcsinh}{
-  \sinh^{-1}
-}
-\newcommand{\arccosh}{
-  \cosh^{-1}
-}
-\newcommand{\erfc}{
-  \text{erfc}
-}
-\newcommand{\E}{
-  \mathbb{E}
-}
-\renewcommand{\P}{
-  \mathbb{P}
-}
-\newcommand{\I}{
-  \mathbf{1}
-}
-\newcommand{\1}[1]{
-  \mathbf{1}_{\{#1\}}
-}
-\newcommand{\diag}{
-  \text{diag\,}
-}
-\newcommand{\pop}[1]{
-\varepsilon_{#1}
-}
-
-\newcommand{\M}{
-  {\text{max}}
-}
-\newcommand{\m}{
-  {\text{min}}
-}
-% \newcommand{\ph}{
-%   {\text{arg}\,}
-% }
-\newcommand\erf{
-  \text{erf}
-}
-\renewcommand\vec[1]{
-  \mathbf{#1}
-}
-\newcommand\cov{
-  \text{cov}
-}
-\newcommand\var{
-  \text{var}
-}
-\newcommand\cor{
-  \text{cor}
-}
-\DeclareMathOperator*{\argmin}{argmin}
-\DeclareMathOperator*{\argmax}{argmax}
-\newcommand\mtx[1]{
-  \mathbf{#1}
-}
-\renewcommand \L {
-  \mathfrak{leb}
-}
-\newcommand{\hot}{
-  \text{h.o.t.}
-}
-
-%% \newtheorem{proposition}{Proposition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{acknowledgement}{Acknowledgement}
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{axiom}{Axiom}
-\newtheorem{case}{Case}
-\newtheorem{claim}{Claim}
-\newtheorem{conclusion}{Conclusion}
-\newtheorem{condition}{Condition}
-\newtheorem{conjecture}{Conjecture}
-\newtheorem{corollary}{Corollary}
-\newtheorem{criterion}{Criterion}
-\newtheorem{definition}{Definition}
-\newtheorem{example}{Example}
-\newtheorem{exercise}{Exercise}
-\newtheorem{lemma}{Lemma}
-\newtheorem{notation}{Notation}
-\newtheorem{problem}{Problem}
-\newtheorem{proposition}{Proposition}
-\newtheorem{remark}{Remark}
-\newtheorem{solution}{Solution}
-\newtheorem{summary}{Summary}
+\input{../../physics_common.tex}
 
@@
 \end{figure}
+One can see that all the confidence intervals of stocks in the
+``Energy'' and ``Consumer Staples'' sectors, as well as those of a
+large portion of stocks in the ``Information Technology'' sector,
+overlap. This naturally prompts an investigation into the extreme
+quantile estimates of the same stocks.
+
+%% Hence it is informative to test
+%% the hypothesis that the tail indices of these return series are equal.
+%% Hoga \cite{Hoga2017} devised tests suitable this purpose. We adopt his
+%% approach but use our own implementation. The next section has more
+%% details.
+
+%% \subsection{Hill estimates of extreme quantiles}
+
+%% By the design of Hoga's test, the null hypothesis can be
+%% rejected due to changed scale parameters instead of changed tail
+%% indices. Thus it is useful to apply the Hill scale estimator to the
+%% same stock return series.
+
+\subsection{Hoga's Test}
+Hoga \cite{Hoga2017} proposed his test for a time series
+$\{X_i\}_{i=1}^n$ with tail functions
+$\bar F_i(\cdot)$, i.e. $\P(X_i > x) = \bar F_i(x)$. The null
+hypothesis is
+\begin{hypothesis}
+  $\mathcal H_0$: For a given $p$, $\exists t^* \in [t_0, 1 - t_0]$
+  such that
+  \[
+  \bar F_1^{-1}(p) = \bar F_2^{-1}(p) = \cdots =
+  \bar F_{\floor{t^* n}}^{-1}(p) \neq
+  \bar F_{\ceil{t^* n}}^{-1}(p) = \cdots = F_n^{-1}(p)
+  \]
+\end{hypothesis}
+where $t_0 \in (0, 1/2)$. The alternative hypothesis is that no such
+$t^*$ exists. Hoga proposed the following test statistics:
+\[
+Q_1 = \sup_{t \in [t_0, 1 - t_0]}
+\left\{
+  {
+    \left[
+      t (1 - t) \ln\left(
+      \hat x_p(0, t)
+      \over
+      \hat x_p(t, 1)
+      \right)
+    \right]^2
+    \over
+    \int_{t_0}^t
+    \left[
+      s \ln \left(
+      \hat x_p(0, s)
+      \over
+      \hat x_p(0, t)
+      \right)
+    \right]^2 ds
+    +
+    \int_{t}^{1 - t_0}
+    \left[
+      (1 - s) \ln \left(
+      \hat x_p(s, 1)
+      \over
+      \hat x_p(t, 1)
+      \right)
+    \right]^2 ds
+  }
+\right\}
+\]
+Under the null hypothesis, Hoga showed that as $n \to \infty$,
+\begin{equation}
+  \label{eq:Q1_distr}
+  Q_1 \overset{d}{\to}
+  \sup_{t \in [t_0, 1 - t_0]} q_t
+\end{equation}
+where
+\[
+q_t = 
+  {
+    [W(t) - t W(1)]^2
+    \over
+    \int_{t_0}^t [W(s) - {s \over t}W(t)]^2 ds
+    +
+    \int_{t}^{1 - t_0}
+    \left[
+      W(1) - W(s) - {1 - s \over 1 - t}
+      (W(1) - W(t))
+      \right]^2 ds
+  }
+\]
+$W(t)$ with $t \in (0, 1)$ is a standard Brownian motion.
+
+When applied to our problem, We would like to test, first of all,
+whether there is a change of tail index in each of two 
+series and, if there is no change in either series, whether the two
+series have the same tail index. Obviously, for either question, if
+the tail index changes, the extreme quantile will change.
+For the first question, Hoga's test is directly applicable; for the
+second question, we concatenate the two series before applying the
+test.
+
+Figure \ref{fig:Hoga_Energy_Single} shows the test statistics of
+stocks in the ``Energy'' sector of the S\&P 500
+index. The null hypothesis here is that the tail index remains the
+same througout the selected period of time.
+\begin{figure}[htb!]
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{Hoga_Energy_Single.pdf}
+    \caption{``Energy''}
+    \label{fig:Hoga_Energy_Single}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{Hoga_Consumer_Staples_Single.pdf}
+    \caption{``Consumer Staples''}
+    \label{fig:Hoga_consumer_Staples_Single}
+  \end{subfigure}
+  \caption{Hoga Test Statistics of stocks in the ``Energy'' and
+    ``Consumer Staples'' sectors of
+    S\&P 500. The green, blue and red lines are, respectively the
+    85\%, 90\% and 95\% quantiles of the distribution of
+    \eqref{eq:Q1_distr}}.
+\end{figure}
 

### 2017-03-28 r/hoga.r

@@ -1,81 +1,81 @@
 rm(list=ls());
 source("libxxie.r")
 #Energy
-tables <- c (
-    "APA",
-    "APC",
-    "BHI",
-    "CHK",
-    "COG",
-    "COP",
-    "DO",
-    "DVN",
-    "EOG",
-    "EQT",
-    "FTI",
-    "HAL",
-    "HES",
-    "HP",
-    "KMI",
-    "MPC",
-    "MRO",
-    "MUR",
-    "NBL",
-    "NFX",
-    "NOV",
-    "OKE",
-    "OXY",
-    "PSX",
-    "PXD",
-    "RIG",
-    "RRC",
-    "SE",
-    "SLB",
-    "SWN",
-    "TSO",
-    "VLO",
-    "WMB",
-    "XEC",
-    "XOM"
-);
+## tables <- c (
+##     "APA",
+##     "APC",
+##     "BHI",
+##     "CHK",
+##     "COG",
+##     "COP",
+##     "DO",
+##     "DVN",
+##     "EOG",
+##     "EQT",
+##     "FTI",
+##     "HAL",
+##     "HES",
+##     "HP",
+##     "KMI",
+##     "MPC",
+##     "MRO",
+##     "MUR",
+##     "NBL",
+##     "NFX",
+##     "NOV",
+##     "OKE",
+##     "OXY",
+##     "PSX",
+##     "PXD",
+##     "RIG",
+##     "RRC",
+##     "SE",
+##     "SLB",
+##     "SWN",
+##     "TSO",
+##     "VLO",
+##     "WMB",
+##     "XEC",
+##     "XOM"
+## );
 
 ## Consumer staples
-## tables <- c(
-##     "ADM",
-##     "BF_series_B",
-##     "CAG",
-## ##    "CL",
-##     "CLX",
-##     "COST",
-##     "CPB",
-##     "CVS",
-##     "DPS",
-## ##    "EL",
-##     "GIS",
-##     "HRL",
-##     "HSY",
-##     "K",
-##     "KMB",
-##     "KO",
-##     "KR",
-##     "MDLZ",
-##     "MJN",
-##     "MKC",
-##     "MNST",
-##     "MO",
-##     "PEP",
-##     "PG",
-##     "PM",
-##     "RAI",
-##     "SJM",
-##     "STZ",
-##     "SYY",
-##     "TAP",
-##     "TSN",
-## ##    "WBA",
-## ##    "WFM",
[…]

### 2017-03-28 papers/FX/draft.tex
Revised the paper with Thomas.

@@
 
+To test whether two series share the same tail index, we concatenate
+the two series and apply the Hoga test. For the ``Energy'' and the
+``Consumer Staples'' sector, we summarise the results in figure
+\ref{fig:Hoga_Energy_pair} and \ref{fig:Hoga_consumer_Staples_pair}.
+\begin{figure}[htb!]
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{Hoga_Energy_pair.pdf}
+    \caption{``Energy''}
+    \label{fig:Hoga_Energy_pair}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{Hoga_Consumer_Staples_pair.pdf}
+    \caption{``Consumer Staples''}
+    \label{fig:Hoga_consumer_Staples_pair}
+  \end{subfigure}
+  \caption{Hoga Test Statistics of stocks in the ``Energy'' and
+    ``Consumer Staples'' sectors of
+    S\&P 500. The green, blue and red colors represent, respectively,
+    exceedances of the 85\%, 90\% and 95\% quantiles of the distribution of
+    \eqref{eq:Q1_distr}.
+    Grey points stand for pairs for which the test statistic is below
+    the 85\% quantile;
+    black points represent pairs for which
+    computation of the test statistics fails for given precision
+    requirements and time limits}.
+\end{figure}
+Clearly, these figures show that the null
+hypothesis of an equal tail index is rejected for more pairs in the
+``Energy'' sector than it is for those in the ``Consumer Staples''
+sector. This suggests that lower tail indices of stocks in the
+``energy'' sector are more spread out than are those of the ``Consumer
+Staples'' sector.
+
+Also observe that while 3 stocks, say A, B and C, test in favor of
+relations $\alpha_A = \alpha_B, \alpha_A \neq \alpha_C$, it often
+happens that another test on B and C is supportive of $\alpha_B =
+\alpha_C$ . This is due to the limited power of the test. Based
+on such results, one may guess $\alpha_B$ lies between $\alpha_A$ and
+$\alpha_C$. The test is unable to recognize the smaller differences
+between $\alpha_A, \alpha_B$ and between $\alpha_B, \alpha_C$.
+
+To have an idea about the power of Hoga's test, we run it on
+concatenated samples drawn from t-distributions with different degrees
+of freedom. Figure \ref{fig:t_sim_pair} shows the result.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\textwidth]{t_sim_pair.pdf}
+  \caption{Test Statistics of concatenated t-samples with different
+    degrees of freedom. Numbers on the axis are the degrees of freedom.}
+  \label{fig:t_sim_pair}
+\end{figure}
+From the figure one can see that the power of the test decreases as
+the minimum tail index of a concatenated pair increases.
+
 \section{Review of multivariate GARCH models:

### 2017-03-28 r/libxxie.r

@@ -1,5 +1,6 @@
 library(RMySQL);
 library(abind);
+library(parallel)
 ## library(rmgarch);
 
 ###
@@ -660,8 +661,10 @@ asymptoticDist <- function(n.paths, n.steps, t0)
 {
     W <- matrix(rnorm(n.steps * n.paths),
                 nrow=n.steps, ncol=n.paths);
+    ## W <- unlist(mclapply(1:n.steps, cumsum, mc.cores=detectCores()));
     W <- apply(W, MARGIN=2, FUN=cumsum);
-    f <- function(X) {
+    f <- function(index) {
+        X <- W[, index];
         n <- length(X);
         ds <- 1/n;
         a <- ceiling(t0 * n);
@@ -679,7 +682,9 @@ asymptoticDist <- function(n.paths, n.steps, t0)
         }
         return(max(results));
     }
-    samples <- apply(W, MARGIN=2, FUN=f);
+    ## samples <- apply(W, MARGIN=2, FUN=f);
+    samples <- unlist(mclapply(X=1:n.paths, FUN=f, mc.cores=detectCores()));
+    ## samples <- lapply(X=n.paths, FUN=f);
     return(ecdf(samples));
 }
 

### 2017-03-29 papers/FX/draft.tex
Revised the draft of FX.

@@
 investors guided by this utility function would be seeking the
-smallest $\nu$, i.e. the heaviest tail in the market.
-
-However, this is not necessarily the case when $b > 0$. For example,
-consider the case when $q = 0$ and $\hat\phi = 1$. This may be
-viewed an approximation to the situation where the interest rate $r$ is
-vanishingly low while the investor is more conservative than
-greedy. In this case we have
-\begin{eqnarray*}
-  \mathcal U(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
-\end{eqnarray*}
-Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
-$\opd{\nu}\mathcal U(F, \phi) > 0$. In this extreme case, the investor
-would rather seek the largest $\nu$ or in other words, the lightest
-tail in the market.
+smallest $\nu$, i.e. the heaviest tail in the market. Here note that
+Student's t is a symmetric distribution; heavier tails imply greater
+down-side risk as well as greater up-side potential.
+
+In fact, even when $b > 0$, the same monotonicity property holds. This
+is revealed by numerically maximizing $\mathcal U(F, \phi)$ with
+respect to $\phi$ for a sequence of selected values of $\nu$. The
+results are shown in figures \ref{fig:phi_hat_t_log} and
+\ref{fig:U_b_t_log}.
+\begin{figure}[htb!]
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
+    \caption{$\hat\phi(\nu)$ of log-utility function}
+    \label{fig:phi_hat_t_log}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_log.pdf}
+    \caption{$G(\nu)$ of log-utility function}
+    \label{fig:U_b_t_log}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
+    \caption{$\hat\phi(\nu)$ of power-utility function}
+    \label{fig:phi_hat_t_power}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
+    \caption{$G(\nu)$ of power-utility function}
+    \label{fig:U_b_t_power}
+  \end{subfigure}
+  \caption{$\hat\phi$ and $G(\nu)$
+  }
+\end{figure}
+
+%% However, this is not necessarily the case when $b > 0$. For example,
+%% consider the case when $q = 0$ and $\hat\phi = 1$. This may be
+%% viewed an approximation to the situation where the interest rate $r$ is
+%% vanishingly low while the investor is more conservative than
+%% greedy. In this case we have
+%% \begin{eqnarray*}
+%%   \mathcal U(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
+%% \end{eqnarray*}
+%% Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
+%% $\opd{\nu}\mathcal U(F, \phi) > 0$. In this extreme case, the investor
+%% would rather seek the largest $\nu$ or in other words, the lightest
+%% tail in the market.
 
@@
 \]
-Another major difference caused by a non-zero value of $b$ is seen in
-the value of $\hat\phi$. When $b = 0$, $\hat\phi$ increases
-continuously with $\nu$ as shown in figure \ref{fig:phi_hat_t5e-1};
-when $b > 0$, $\hat\phi$ instead increases by jumping up at a sequence
-of values of $\nu$. This is shown in figure \ref{fig:phi_hat_b_t5e-1}.
-\begin{figure}[htb!]
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_t5e-1.pdf}
-    \caption{$\hat\phi$ when $b = 0$}
-    \label{fig:phi_hat_t5e-1}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t5e-1.pdf}
-    \caption{$\hat\phi$ when $b = 0.01$}
-    \label{fig:phi_hat_b_t5e-1}
-  \end{subfigure}
-  \caption{$\hat\phi$, the optimal equity allocation when the utility
-    function is $u(C) = -{2 \over \sqrt{C}}$
-  }
-\end{figure}
+One can see from figure \ref{fig:phi_hat_t_power} and
+\ref{fig:U_b_t_power} that both $\hat \phi$ and $G(\nu)$ are monotonically 
+decreasing with $\nu$ as in the case of a log-utility function.
+%% \begin{figure}[htb!]
+%%   \begin{subfigure}[b]{0.5\linewidth}
+%%     \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
+%%     \caption{$\hat\phi(\nu)$}
+%%     \label{fig:phi_hat_t_power}
+%%   \end{subfigure}
+%%   \begin{subfigure}[b]{0.5\linewidth}
+%%     \includegraphics[width=\textwidth]{U_b_t_power.pdf}
+%%     \caption{$G(\nu)$}
+%%     \label{fig:U_b_t_power}
+%%   \end{subfigure}
+%%   \caption{$\hat\phi$ and $G(\nu)$ when the utility
+%%     function is $u(C) = -{2 \over \sqrt{C}}$
+%%   }
+%% \end{figure}
 

### 2017-03-29 papers/FX/draft.tex
Revised the scripts.
Added an introductory paragraph at the beginning of the section of empirical results.

@@
 put it, ``the last bite is never as satisfying as the first''
-\cite{cochrane2009asset}. This implies investors are considerably
-more concerned about large losses than are cheered by large gains.
+\cite{cochrane2009asset}. This implies investors are repelled to a
+larger extent by large losses than they are tempted by large gains. 
 Furthermore, it is a stylized fact that equity return distributions
@@
 
-We show in this paper that concerns of the down-side risk and the
-heavy-tail property of equity returns  imply that the lower-tail indices
-of actively traded equities in a rational market must be fairly close to
-one another.
-
-%% A cornerstone of this conclusion is the empirical fact that equity
-%% return distributions are heavy-tailed (see e.g. Jansen and de Vries
-%% (1991) \cite{jansen1991frequency}). If investors and speculators alike
-%% display concern for downside risk, as formulated via a utility function,
-%% it is important to model this risk adequately.
-
-In section \ref{sec:GDA}, we explicitly recognize the behavioral
+we explore in this paper the market implications of investor
+preferences and the heavy-tail property of equity return
+distributions. The rest of this paper is organized into the following
+sections:
+In section \ref{sec:empirical_results} we present a survey of
+statistical characteristics of equity returns. Then 
+in section \ref{sec:GDA}, we explicitly recognize the behavioral
 concern for downside risk in the investor's evaluation of portfolios
@@
 principles (axiomatic).
+In this section we infer characteristics of equity returns from the
+GDA theory and compare them with the empirical results.
 
@@
 \section{Empirical Results}
+\label{sec:empirical_results}
+As has been mentioned, we are interested in the heavy-tail properties
+of equity returns. Let $X_t$ be an equity return series. If the
+distribution of $X_t$ is heavy-tailed, we can write
+\begin{eqnarray*}
+  \lim_{x \to \infty} {x^{\beta}\P(X_t > x) \over L_+(x)} &=& 1 \\
+  \lim_{x \to \infty} {x^{\alpha}\P(X_t < -x) \over L_-(x)} &=& 1
+\end{eqnarray*}
+where $L_+(x)$ and $L_-(x)$ are slowly varying functions satisfying
+\[
+\lim_{x \to \infty} {L_*(a x) \over L_*(x)} = 1
+\]
+for all $a > 0$. Here ``*'' can be either ``+'' or ``-''. In the
+simplest case, $L_+(x)$ and $L_(x)$ are constants and the distribution
+has power-law tails. Let $J_*(1/x) = L_*(x)$. A Taylor expansion at
+$x=\infty$ gives
+\[
+J_*(1/x) = J_*(0) + {J'_*(0) \over x} + O(x^{-2})
+\]
+Thus
+\[
+x^{-\alpha} L_-(x) = J_-(0) x^{-\alpha} + J'_-(0) x^{-\alpha - 1} + O(x^{-\alpha-2})
+\]
+when $x \to \infty$. Clearly the tail indices $\alpha$ and $\beta$ as
+well as the scale parameters $J_-(0)$ and $J_+(0)$ play crutial roles
+in determining the tail distribution and hence in asset allocation and
+risk management. These parameters of the equities in a market are
+characteristics of the market and provide a useful description of the
+market. One can also use them to help build a model of the equities in
+the market. With these motives in mind, we present a survey of the
+tail indices and scale parameters of 3 sectors, namely ``Energy'',
+``Consumer Staples'' and ``Information Technology'', of the S\&P 500
+index.
+
 \subsection{Hill estimates of lower tail indices}
@@
 large portion of stocks in the ``Information Technology'' sector,
-overlap. This naturally prompts an investigation into the extreme
-quantile estimates of the same stocks.
+overlap. This suggests that the stocks in each sector may have very
+similar tail indices. Hoga's test on the change of extreme quantiles
+may provide some insight about how similar these tail indices are,
+because different tail indices are likely to result in different
+extreme quantiles. Nevertheless, changes in the extreme quantiles can
+be the result of changed tail indices but can also be the result of
+changed scale parameters. Thus we first investigate the scale
+parameters of stocks in the same sectors before applying Hoga's test.
+
+\subsection{Hill estimates of lower-tail scale parameters}
 
@@
 \subsection{Hoga's Test}
+\label{sec:Hoga}
+
 Hoga \cite{Hoga2017} proposed his test for a time series
@@
     computation of the test statistics fails for given precision
-    requirements and time limits}.
+    requirements and time limits.}
 \end{figure}
@@
                     {Consumer_Staples_alpha_K_ci.pdf}
-                    \caption{``Information Technology''}
+                    \caption{``Consumer Staples''}
                     \label{fig:Consumer_Staples_alpha_K_ci}
@@
     are the
-    estimated values; the red bars are the upper bound of the 98\%
+    estimated values; the bars are the upper bound of the 98\%
     confidence interval. The lower bound is 0.

### 2017-03-30 papers/FX/draft.tex
Revised the part about estimation of shifted pareto distributions.

@@
 \end{eqnarray*}
-where $L_+(x)$ and $L_-(x)$ are slowly varying functions satisfying
+where the constants $\alpha$ and $\beta$ are left and right tail
+indices, respectively, and $L_+(x)$ and $L_-(x)$ are slowly varying
+functions satisfying
 \[
@@
 $x=\infty$ gives
-\[
+\begin{equation}
+  \label{eq:scale_parameter}
   J_*(1/x) = J_*(0) + {J'_*(0) \over x} + O(x^{-2})
-\]
+\end{equation}
 Thus
-\[
-x^{-\alpha} L_-(x) = J_-(0) x^{-\alpha} + J'_-(0) x^{-\alpha - 1} + O(x^{-\alpha-2})
-\]
+\begin{equation}
+  \label{eq:lower_tail_expansion}
+  x^{-\alpha} L_-(x) =
+  J_-(0) x^{-\alpha} + J'_-(0) x^{-\alpha - 1} + O(x^{-\alpha-2})
+\end{equation}
 when $x \to \infty$. Clearly the tail indices $\alpha$ and $\beta$ as
@@
 arguably Hill (1975) \cite{hill1975simple}:
-\[
+\begin{equation}
+  \label{eq:Hill_index}
   \hat \alpha = \left[
@@
     \right]^{-1}
-\]
+\end{equation}
 where $x_{(i)}$ is the $i$-th upper order statistic of a sample
@@
 \subsection{Hill estimates of lower-tail scale parameters}
+\label{sec:HillScaleEstimates}
+
+For convenience, we denote the scale parameter by $A$. It is the
+constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
+Hill \cite{hill1975simple} put forward an estimator of $A$ using his
+tail index  estimator:
+\[
+\hat A = {k \over n} X_{(k+1)}^{\hat \alpha}
+\]
+where $X_{(k+1)}$ is the $(k+1)$-th upper order statistic of the
+sequence $X_t$ as defined in section\ref{sec:Hill}.
 
@@
 \end{theorem}
+The proof of theorem \ref{thrm:I} is given in secton
+\ref{sec:thrmI_proof}.
 Since $u(v)$ given in \eqref{eq:xxie1.0} as a
@@
 should appear positively correlated. This is indeed the case in some
-real stock return data, e.g. the ``Energy'' and ``Information Technology''
-sectors of S\&P 500 companies. This is illustrated in figures
-\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
+real stock return data, e.g. the ``Energy'', ``Consumer Staples'' and
+``Information Technology'' sectors of S\&P 500 companies. This is
+illustrated in figures 
+\ref{fig:Energy_alpha_K_ci}, \ref{fig:Consumer_Staples_alpha_K_ci}
+and \ref{fig:Information_Technology_alpha_K_ci}.
 The confidence bands of the Hill estimates of $\alpha$ of these
@@
 
-\begin{proof}
-  Let
-  \begin{equation}
-    \label{eq:xxie2}
-    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
-  \end{equation}
-  We have
-  \[
-  G(F) = \mathcal U(F, \hat\phi)
-  \]
-  It follows
-  \begin{eqnarray}
-    \pd{G(F)}{\alpha}
-    &=&
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
-    + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat\phi)
-    \pd{\hat\phi}{\alpha}(\alpha, K)
-    \label{eq:xxie3}
-  \end{eqnarray}
-  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
-  \begin{equation}
-    \label{eq:xxie4}
-    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat \phi) = 0
-  \end{equation}
-  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
-  \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat \phi)
-    &>& 0
-  \end{eqnarray*}
-  From \eqref{eq:xxie1.0}, it follows
-  \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
-    = \opd{\alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}]
-  \end{eqnarray*}
-  The function $u((1 - \hat\phi) e^r + \hat\phi e^X)$ is obviously increasing
-  with $X$. It follows
-  \begin{eqnarray*}
-    && \pd{f(x; \alpha, K)}{\alpha} \\
-    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
-    &=&
-    - {K^\alpha \over (K - x)^{\alpha + 1}}
-    \left[
-      \alpha
-      \ln\left(
-        1 - {x \over K}
-      \right) - 1
-    \right]
-  \end{eqnarray*}
-  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
-  \[
-  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
-  \]
-  and when $K(1 - e^{1/\alpha}) < x < 0$,
-  \[
-  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
-  \]
-  This is the second case of lemma \ref{lemma:I}. So we have
-  \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi) &>& 0
-  \end{eqnarray*}
-  As for $\pd{G(F)}{K}$, by the same argument as for
-  $\pd{G(F)}{\alpha}$, it suffices to show
-  \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0
-  \]
-  We have
-  \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi)
-  = \opd{K} \E \left[
-    u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}
-  \right] 
-  \]
-  and
-  \begin{eqnarray*}
-    && \pd{f(x; \alpha, K)}{K} \\
-    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
-    &=&
-    -\alpha K^{\alpha - 1} {
-      \alpha x + K
-      \over
-      (K - x)^{\alpha + 2}
-    }
-  \end{eqnarray*}
-  Clearly, when $x < -\alpha/K$
-  \[
-  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
-  \]
-  and when $-\alpha/K < x < 0$
-  \[
-  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
-  \]
-  So by the 1st case of remark \ref{remark:I} we conclude
-  \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0  
-  \]
-\end{proof}
 By numerically integration and optimization with respect to $\phi$,
@@
 
-Figure \ref{fig:Energy_alpha_K},
-\ref{fig:Consumer_Staples_alpha_K},
-and \ref{fig:Information_Technology_alpha_K}
+To obtain an estimate of $p$, we use the law of large number:
+\begin{eqnarray*}
+  p &=& \P(X < 0) = \lim_{n \to \infty}
+  {1 \over n} \sum_{t=1}^n \1{X_t < 0}
+\end{eqnarray*}
+By the central limit theorem,
+\begin{eqnarray*}
+  \sqrt n \left[
+    {1 \over n}\sum_{t=1}^n \1{X_t < 0} - p
+  \right] \to N(0, p - p^2))
+\end{eqnarray*}
+
+Figure \ref{fig:Energy_alpha_K_ci},
+\ref{fig:Consumer_Staples_alpha_K_ci},
+and \ref{fig:Information_Technology_alpha_K_ci}
 show the estimates of $K$ alongside the Hill estimates of $\alpha$ of
-the ``energy'', ``consumer staples'' and ``information
-technology'' sectors of the S\&P 500 index, respectively.
+the ``energy'', ``consumer staples'' and ``information technology''
+sectors of the S\&P 500 index, respectively.
+Figure \ref{fig:Energy_p}, \ref{fig:Consumer_Staples_p} and
+\ref{fig:Information_Technology_p} show the estimates of $p$ of the
+same sectors.
+
+From these figures, one can see that stocks of the ``Energy'' and the
+``Information Technology'' sectors generally have larger values of
+$p$, meaning their stock prices fall more often than do those in the
+``consumer Staples'' sector; Also, the values of $K$ in the ``Energy''
+and ``Information Technology'' sectors are larger than those in the
+``Consumer Staples'' sector. A larger value of $K$ means, for a given
+value of $p$, i.e. a given probability of losses, large losses are
+more probable. Thus one can conclude that these two sectors are
+considerablly riskier than the ``Consumer Staples'' sector. This is of
+course a confirmation of one's economic instinct.
+
+One can also see from figure \ref{fig:Energy_alpha_K_ci}
+and \ref{fig:Information_Technology_alpha_K_ci} that, while the
+``Energy'' and the ``Information Technology'' sectors are similar in
+riskiness, the correlation between $\alpha$ and $K$ is stronger in
+``Energy''. As discussed in the last section, moving along a curve of
+equal preference in the direction of increasing $\alpha$, $K$ also
+increases. So the strong positive correlation seen in the ``Energy''
+sector suggests these stocks might have very similar investor
+preferences. This in turn may be attributed to stronger business
+relations between the energy enterprises. While two IT companies
+may provide a variety of products and services and not depend on each
+other, two energy companies are more likely to depend on each other
+via relations of supplier and customer or otherwise to compete with each
+other if they are on the same link of the chain of energy production
+and distribution.
+
 \begin{figure}[htb!]
@@
     \includegraphics[width=\textwidth]
-                    {Energy_alpha_K.pdf}
+                    {Energy_p.pdf}
                     \caption{``Energy''}
-                    \label{fig:Energy_alpha_K}
+                    \label{fig:Energy_p}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]
-                    {Consumer_Staples_alpha_K.pdf}
+                    {Consumer_Staples_p.pdf}
                     \caption{``Consumer Staples''}
-                    \label{fig:Consumer_Staples_alpha_K}
+                    \label{fig:Consumer_Staples_p}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]
-                    {Information_Technology_alpha_K.pdf}
+                    {Information_Technology_p.pdf}
                     \caption{``Information Technology''}
-                    \label{fig:Information_Technology_alpha_K}
+                    \label{fig:Information_Technology_p}
   \end{subfigure}
@@
   \end{subfigure}
-  \caption{\small Maximum likelihood estimates of $K$, the scale
-    parameter, of stocks in sectors of the S\&P 500 index. The points
-    are the
-    estimated values; the bars are the upper bound of the 98\%
-    confidence interval. The lower bound is 0.
+  \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
+    S\&P 500 index. The points are the estimated values; the bars are
+    the $2\sigma$ confidence intervals.
   }
 \end{figure}
+To compare the maximum likelihood estimates with the results obtained
+from Hill's scale estimates, one can re-write the lower-tail
+distribution function: For all $x < 0$,
+\begin{eqnarray*}
+  \P(X < x)
+  &=&
+  p {K^\alpha \over (K - x)^\alpha} \\
+  &=&
+  p {K^\alpha |x|^{-\alpha} \over (1 - K/x)^\alpha} \\
+  &=&
+  p K^\alpha |x|^{-\alpha} \left(
+  1 + {\alpha K \over x} + O(|x|^{-2})
+  \right)
+\end{eqnarray*}
+Comparing this expression with \eqref{eq:lower_tail_expansion}, one
+can immediately identity $p K^\alpha$ as the scale parameter.
 
@@
 smallest $\nu$, i.e. the heaviest tail in the market. Here note that
-Student's t is a symmetric distribution; heavier tails imply greater
-down-side risk as well as greater up-side potential.
+Student's t is a symmetric distribution; a heavier tail implies not
+only greater down-side risk but also greater up-side potential.
 
@@
 
+\section{Proof of Theorem \ref{thrm:I}}
+\label{sec:thrmI_proof}
+\begin{proof}
+  Let
+  \begin{equation}
+    \label{eq:xxie2}
+    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
+  \end{equation}
+  We have
+  \[
+  G(F) = \mathcal U(F, \hat\phi)
+  \]
+  It follows
+  \begin{eqnarray}
+    \pd{G(F)}{\alpha}
+    &=&
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
+    + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat\phi)
+    \pd{\hat\phi}{\alpha}(\alpha, K)
+    \label{eq:xxie3}
+  \end{eqnarray}
+  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
+  \begin{equation}
+    \label{eq:xxie4}
+    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat \phi) = 0
+  \end{equation}
+  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
+  \begin{eqnarray*}
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat \phi)
+    &>& 0
+  \end{eqnarray*}
+  From \eqref{eq:xxie1.0}, it follows
+  \begin{eqnarray*}
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
+    = \opd{\alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}]
+  \end{eqnarray*}
+  The function $u((1 - \hat\phi) e^r + \hat\phi e^X)$ is obviously
+  increasing with $X$. It follows
+  \begin{eqnarray*}
+    && \pd{f(x; \alpha, K)}{\alpha} \\
+    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+    &=&
+    - {K^\alpha \over (K - x)^{\alpha + 1}}
+    \left[
+      \alpha
+      \ln\left(
+        1 - {x \over K}
+      \right) - 1
+    \right]
+  \end{eqnarray*}
+  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
+  \[
+  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  \]
+  and when $K(1 - e^{1/\alpha}) < x < 0$,
+  \[
+  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  \]
+  This is the second case of lemma \ref{lemma:I}. So we have
+  \begin{eqnarray*}
+    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi) &>& 0
+  \end{eqnarray*}
+  As for $\pd{G(F)}{K}$, by the same argument as for
+  $\pd{G(F)}{\alpha}$, it suffices to show
+  \[
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0
+  \]
+  We have
+  \[
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi)
+  = \opd{K} \E \left[
+    u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}
+  \right] 
+  \]
+  and
+  \begin{eqnarray*}
+    && \pd{f(x; \alpha, K)}{K} \\
+    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+    &=&
+    -\alpha K^{\alpha - 1} {
+      \alpha x + K
+      \over
+      (K - x)^{\alpha + 2}
+    }
+  \end{eqnarray*}
+  Clearly, when $x < -\alpha/K$
+  \[
+  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  \]
+  and when $-\alpha/K < x < 0$
+  \[
+  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  \]
+  So by the 1st case of remark \ref{remark:I} we conclude
+  \[
+  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0  
+  \]
+\end{proof}
+
 \bibliographystyle{unsrt}

### 2017-03-30 r/myLogLik.r

@@ -3,43 +3,43 @@ require(MASS);
 require(gplots);
 source("libxxie.r")
 #Energy
-## tables <- c (
-##     "APA",
-##     "APC",
-##     "BHI",
-##     "CHK",
-##     "COG",
-##     "COP",
-##     "DO",
-##     "DVN",
-##     "EOG",
-##     "EQT",
-##     "FTI",
-##     "HAL",
-##     "HES",
-##     "HP",
-##     "KMI",
-##     "MPC",
-##     "MRO",
-##     "MUR",
-##     "NBL",
-##     "NFX",
-##     "NOV",
-##     "OKE",
-##     "OXY",
-##     "PSX",
-##     "PXD",
-##     "RIG",
-##     "RRC",
-##     "SE",
-##     "SLB",
-##     "SWN",
-##     "TSO",
-##     "VLO",
-##     "WMB",
-##     "XEC",
-##     "XOM"
-## );
+tables <- c (
+    "APA",
+    "APC",
+    "BHI",
+    "CHK",
+    "COG",
+    "COP",
+    "DO",
+    "DVN",
+    "EOG",
+    "EQT",
+    "FTI",
+    "HAL",
+    "HES",
+    "HP",
+    "KMI",
+    "MPC",
+    "MRO",
+    "MUR",
+    "NBL",
+    "NFX",
+    "NOV",
+    "OKE",
+    "OXY",
+    "PSX",
+    "PXD",
+    "RIG",
+    "RRC",
+    "SE",
+    "SLB",
+    "SWN",
+    "TSO",
+    "VLO",
+    "WMB",
+    "XEC",
+    "XOM"
+);
 
 ## Consumer staples
 ## tables <- c(
@@ -80,50 +80,50 @@ source("libxxie.r")
 ## );
 
 ## Information Technology
-tables <- c(
-    "ADBE",
-    "ADI",
-    "ADP",
-    "ADSK",
-    "AKAM",
-    "AMAT",
-    "CA",
-    "CSCO",
-    "CTSH",
-    "CTXS",
-    "EA",
-    "EBAY",
-    "FFIV",
-    "FISV",
-    "HPQ",
-    "HRS",
-    "IBM",
-    "INTC",
-    "INTU",
-    "JNPR",
-    "KLAC",
-    "LLTC",
-    "LRCX",
-    "MCHP",
-    "MSFT",
-    "MSI",
-    "MU",
-    "NTAP",
[…]

### 2017-03-30 papers/FX/draft.tex
Added pictures of scale estimates.

@@
 where $X_{(k+1)}$ is the $(k+1)$-th upper order statistic of the
-sequence $X_t$ as defined in section\ref{sec:Hill}.
+sequence $X_t$ as defined in section\ref{sec:Hill}. $\hat \alpha$
+is the Hill tail index estimator.
+Estimates of $A$ of the ``Energy'', ``Consumer Staples'' and
+``Information Technology'' sectors of the S\&P 500 index are computed
+using this method. The results are shown in figure \ref{fig:Energy_scale},
+\ref{fig:Consumer_Staples_scale} and
+\ref{fig:Information_Technology_scale}, along with estimates obtained
+using the hybrid method described in section
+\ref{sec:hybrid_estimation}.
 
@@
 For the first question, Hoga's test is directly applicable; for the
-second question, we concatenate the two series before applying the
+second, we concatenate the two series before applying the
 test.
@@
   \centering
-  \includegraphics[width=\textwidth]{t_sim_pair.pdf}
+  \includegraphics[width=\textwidth, trim={1cm 1cm 1cm 2cm},
+  clip]{t_sim_pair.pdf}
   \caption{Test Statistics of concatenated t-samples with different
@@
 Comparing this expression with \eqref{eq:lower_tail_expansion}, one
-can immediately identity $p K^\alpha$ as the scale parameter.
+can immediately identity $p K^\alpha$ as the scale parameter. Figure
+\ref{fig:Energy_scale}, \ref{fig:Consumer_Staples_scale} and
+\ref{fig:Information_Technology_scale} compare the estimates obtained
+from Hill's scale estimator with those from the $p K^\alpha$ method.
+
+\begin{figure}[htb!]
+  \centering
+  \begin{subfigure}[b]{0.32\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Energy_scale.pdf}
+                    \caption{``Energy''}
+                    \label{fig:Energy_scale}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Consumer_Staples_scale.pdf}
+                    \caption{``Consumer Staples''}
+                    \label{fig:Consumer_Staples_scale}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\linewidth}
+    \includegraphics[width=\textwidth]
+                    {Information_Technology_scale.pdf}
+                    \caption{``IT''}
+                    \label{fig:Information_Technology_scale}
+  \end{subfigure}
+  \caption{\small Scale estimates obtained from Hill's scale estimator
+    and from $p K^\alpha$ . Red dots are Hill's scale estimates and
+    triangles are $p K^\alpha$ estimates.
+  }
+\end{figure}
 

### 2017-03-31 papers/FX/draft.tex
Revised the paper with Thomas and Casper.

@@
   \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{Hoga_Energy_pair.pdf}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 1cm, 1cm, 1cm}, clip
+    ]{Hoga_Energy_pair.pdf}
     \caption{``Energy''}
@@
   \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{Hoga_Consumer_Staples_pair.pdf}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 1cm, 1cm, 1cm}, clip
+    ]{Hoga_Consumer_Staples_pair.pdf}
     \caption{``Consumer Staples''}
@@
   \centering
-  \includegraphics[width=\textwidth, trim={1cm 1cm 1cm 2cm},
+  \includegraphics[width=\textwidth, trim={1.0cm 1.5cm 1cm 2cm},
   clip]{t_sim_pair.pdf}
@@
 the minimum tail index of a concatenated pair increases.
+
+Also of concern is the distribution of the test statistic under the
+null hypothesis. This distribution is only known asymptotically (given
+in equation \eqref{eq:Q1_distr}) when the sample size tends to
+infinity. The rate at which the finite-sample distribution tends to
+\eqref{eq:Q1_distr} is unclear. To find out more about this issue, we
+draw samples from the Student's t(3) distribution and compare the
+empirical distribution of the test statistic $Q_1$ with the
+empirical distribution of \eqref{eq:Q1_distr}, which we also obtain by
+simulating the Brownian motion $W(t)$.
 

### 2017-03-31 cv/20sec/template.tex
Added the 20seconds CV.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Twenty Seconds Resume/CV
% LaTeX Template
% Version 1.1 (8/1/17)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Carmine Spagnuolo (cspagnuolo@unisa.it) with major modifications by 
% Vel (vel@LaTeXTemplates.com)
%
% License:
% The MIT License (see included LICENSE file)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[letterpaper]{twentysecondcv} % a4paper for A4

%----------------------------------------------------------------------------------------
%	 PERSONAL INFORMATION
%----------------------------------------------------------------------------------------

% If you don't need one or more of the below, just remove the content leaving the command, e.g. \cvnumberphone{}

\profilepic{selfie1.jpg} % Profile picture

\cvname{Xie Xiaolei} % Your name
\cvjobtitle{Quantitative Analyst} % Job title/career

\cvdate{09 January 1983} % Date of birth
\cvaddress{Sweden} % Short address/location, use \newline if more than 1 line is required
\cvnumberphone{+46 760606638} % Phone number
\cvsite{https://www.linkedin.com/in/xie-xiaolei-bb0788b/} % Personal website
\cvmail{xie.xiaolei@gmail.com} % Email address

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	 ABOUT ME
%----------------------------------------------------------------------------------------

\aboutme{Alice is a sensible prepubescent girl from a wealthy English family who finds herself in a strange world ruled by imagination and fantasy. Alice feels comfortable with her identity and has a strong sense that her environment is comprised of clear, logical, and consistent rules and features. Alice's familiarity with the world has led one critic to describe her as a "disembodied intellect". Alice displays great curiosity and attempts to fit her diverse experiences into a clear understanding of the world.} % To have no About Me section, just remove all the text and leave \aboutme{}

%----------------------------------------------------------------------------------------
%	 SKILLS
%----------------------------------------------------------------------------------------

% Skill bar section, each skill must have a value between 0 an 6 (float)
\skills{{pursuer of rabbits/5.8},{good manners/4},{outgoing/4.3},{polite/4},{Java/0.01}}

%------------------------------------------------

% Skill text section, each skill must have a value between 0 an 6
\skillstext{{lovely/4},{narcissistic/3}}

%----------------------------------------------------------------------------------------

\makeprofile % Print the sidebar

%----------------------------------------------------------------------------------------
%	 INTERESTS
%----------------------------------------------------------------------------------------

\section{Interests}

The heroine and the dreamer of Wonderland; Alice is the principal character.

%----------------------------------------------------------------------------------------
%	 EDUCATION
%----------------------------------------------------------------------------------------

\section{Education}

\begin{twenty} % Environment for a list with descriptions
	\twentyitem{since 1865}{Ph.D. {\normalfont candidate in Computer Science}}{Wonderland}{\emph{A Quantified Theory of Social Cohesion.}}
	\twentyitem{1863-1865}{M.Sc. magna cum laude}{Wonderland}{Majoring in Computer Science}
	\twentyitem{1861-1863}{B.Sc. magna cum laude}{Wonderland}{Majoring in Computer Science}
	\twentyitem{1856-1861}{High school}{Wonderland}{Specializing in mathematics and physics.}
	%\twentyitem{<dates>}{<title>}{<location>}{<description>}
\end{twenty}

%----------------------------------------------------------------------------------------
%	 PUBLICATIONS
%----------------------------------------------------------------------------------------

\section{Publications}

\begin{twentyshort} % Environment for a short list with no descriptions
	\twentyitemshort{1865}{Chapter One, Down the Rabbit Hole.}
	\twentyitemshort{1865}{Chapter Two, The Pool of Tears.}
	\twentyitemshort{1865}{Chapter Three,  The Caucus Race and a Long Tale.}
	\twentyitemshort{1865}{Chapter Four,  The Rabbit Sends a Little Bill.}
	\twentyitemshort{1865}{Chapter Five,  Advice from a Caterpillar.}
	%\twentyitemshort{<dates>}{<title/description>}
\end{twentyshort}

%----------------------------------------------------------------------------------------
%	 AWARDS
%----------------------------------------------------------------------------------------

\section{Awards}

\begin{twentyshort} % Environment for a short list with no descriptions
	\twentyitemshort{1987}{All-Time Best Fantasy Novel.}
	\twentyitemshort{1998}{All-Time Best Fantasy Novel before 1990.}
	%\twentyitemshort{<dates>}{<title/description>}
\end{twentyshort}

%----------------------------------------------------------------------------------------
%	 EXPERIENCE
%----------------------------------------------------------------------------------------

\section{Experience}

\begin{twenty} % Environment for a list with descriptions
	\twentyitem{1900}{Alice in Wonderland-The Circra (1900's) Silent Film.}{Film}{The first Alice on film was over a hundred years ago.}
	\twentyitem{1933}{Alice in Wonderland 1933 version.}{Film}{This film stars Ethel griffies and Charlotte Henry. It was a box office flop when it was released.}
	\twentyitem{1951}{Disney Film.}{Film}{Walt Disney brings Lewis Carroll's fantasy story to life in this well done animated classic. Even though many elements from the book were dropped, such as the duchess with the baby pig and mock turtle, this version is without a doubt the most famous Alice adaption made.}
	%\twentyitem{<dates>}{<title>}{<location>}{<description>}
\end{twenty}

%----------------------------------------------------------------------------------------
%	 OTHER INFORMATION
%----------------------------------------------------------------------------------------

\section{Other information}

\subsection{Review}

Alice approaches Wonderland as an anthropologist, but maintains a strong sense of noblesse oblige that comes with her class status. She has confidence in her social position, education, and the Victorian virtue of good manners. Alice has a feeling of entitlement, particularly when comparing herself to Mabel, whom she declares has a ``poky little house," and no toys. Additionally, she flaunts her limited information base with anyone who will listen and becomes increasingly obsessed with the importance of good manners as she deals with the rude creatures of Wonderland. Alice maintains a superior attitude and behaves with solicitous indulgence toward those she believes are less privileged.

%----------------------------------------------------------------------------------------
%	 SECOND PAGE EXAMPLE
%----------------------------------------------------------------------------------------

%\newpage % Start a new page

%\makeprofile % Print the sidebar

%\section{Other information}

%\subsection{Review}

%Alice approaches Wonderland as an anthropologist, but maintains a strong sense of noblesse oblige that comes with her class status. She has confidence in her social position, education, and the Victorian virtue of good manners. Alice has a feeling of entitlement, particularly when comparing herself to Mabel, whom she declares has a ``poky little house," and no toys. Additionally, she flaunts her limited information base with anyone who will listen and becomes increasingly obsessed with the importance of good manners as she deals with the rude creatures of Wonderland. Alice maintains a superior attitude and behaves with solicitous indulgence toward those she believes are less privileged.

%\section{Other information}

%\subsection{Review}

%Alice approaches Wonderland as an anthropologist, but maintains a strong sense of noblesse oblige that comes with her class status. She has confidence in her social position, education, and the Victorian virtue of good manners. Alice has a feeling of entitlement, particularly when comparing herself to Mabel, whom she declares has a ``poky little house," and no toys. Additionally, she flaunts her limited information base with anyone who will listen and becomes increasingly obsessed with the importance of good manners as she deals with the rude creatures of Wonderland. Alice maintains a superior attitude and behaves with solicitous indulgence toward those she believes are less privileged.

%----------------------------------------------------------------------------------------

\end{document}

### 2017-03-31 papers/FX/draft.tex
Added the 20seconds CV.

@@
 
-\input{../../physics_common.tex}
+%\input{../../physics_common.tex}
+\input{physics_common.tex}
 
@@
 \eqref{eq:Q1_distr} is unclear. To find out more about this issue, we
-draw samples from the Student's t(3) distribution and compare the
-empirical distribution of the test statistic $Q_1$ with the
+draw samples from the Student's t(3) and t(4) distributions and
+compare the empirical distribution of the test statistic $Q_1$ with the
 empirical distribution of \eqref{eq:Q1_distr}, which we also obtain by
-simulating the Brownian motion $W(t)$.
+simulating the Brownian motion $W(t)$. The empirical density functions
+are shown in figure \ref{fig:Q1_distr}.
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
+  \caption{Comparison of the asymptotic distribution with empirical
+    distribution of $Q_1$ obtained from Student's t samples. The t
+    samples have the same size as the S\&P sectors' data, namely 1304
+    observations.}
+  \label{fig:Q1_distr}
+\end{figure}
+As seen in the figure, the asymptotic distribution \eqref{eq:Q1_distr}
+assigns significantly more mass to the tail than do the empirical
+distributions of size 1304. Table \ref{tab:HogaAsymptotic} lists a few
+quantiles of these distributions.
+\begin{table}[htb!]
+  \centering
+  \begin{tabular}{c|c|c|c|c}
+    & \multicolumn{4}{c}{Quantiles} \\
+    \hline
+    Distribution & 80\% & 85\% & 90\% & 95\% \\
+    \hline
+    Asymptotic & 47.48 & 59.61 & 78.90 & 113.12 \\
+    t(3) empirical & 23.30 & 28.27 & 35.04 & 46.75 \\
+    t(4) empirical & 25.76 & 31.13 & 38.81 & 56.32
+  \end{tabular}
+  \caption{Quantiles of the asymptotic and empirical test statistic
+    distributions}
+  \label{tab:HogaAsymptotic}
+\end{table}
 
@@
 \bibliographystyle{unsrt}
-\bibliography{../../thesis/econophysics}
+%% \bibliography{../../thesis/econophysics}
+\bibliography{econophysics}
 \end{document}

### 2017-04-02 cv/20sec/template.tex
Revised my CV.

@@
 
-\documentclass[letterpaper]{twentysecondcv} % a4paper for A4
+%% \documentclass[letterpaper]{twentysecondcv} % a4paper for A4
+\documentclass[a4paper]{twentysecondcv} % a4paper for A4
 
@@
 
-\aboutme{Alice is a sensible prepubescent girl from a wealthy English family who finds herself in a strange world ruled by imagination and fantasy.
-Alice feels comfortable with her identity and has a strong sense that her environment is comprised of clear, logical, and consistent rules and features.
-Alice's familiarity with the world has led one critic to describe her as a "disembodied intellect".
-Alice displays great curiosity and attempts to fit her diverse experiences into a clear understanding of the world.} % To have no About Me section, just remove all the text and leave \aboutme{}
+\aboutme{
+  \begin{itemize}
+  \item Ph.D of Mathematical Statistics at Copenhagen University
+  \item M.Sc. of Mathematical Physics
+  \item M.Sc. of Computer Science
+  \item 7 years work experience in programming 
+  \item strong in statistical analysis \& numerical methods
+  \end{itemize}
+} % To have no About Me section, just remove all the text and leave \aboutme{}
 
@@
 % Skill bar section, each skill must have a value between 0 an 6 (float)
-\skills{{pursuer of rabbits/5.8},{good manners/4},{outgoing/4.3},{polite/4},{Java/0.01}}
+\skills {{Python/3},{SQL/4},{GNU Scientific Library/5},{Matlab/5.2},{R/5.8},{C/5},{C++/5}}
 
@@
 % Skill text section, each skill must have a value between 0 an 6
-\skillstext{{lovely/4},{narcissistic/3}}
+%% \skillstext{{lovely/4},{narcissistic/3}}
 
@@
 
-\section{Interests}
-
-The heroine and the dreamer of Wonderland; Alice is the principal character.
+%% \section{Summary}
 
@@
 
-\begin{twenty} % Environment for a list with descriptions
-	\twentyitem{since 1865}{Ph.D. {\normalfont candidate in Computer Science}}{Wonderland}{\emph{A Quantified Theory of Social Cohesion.}}
-	\twentyitem{1863-1865}{M.Sc. magna cum laude}{Wonderland}{Majoring in Computer Science}
-	\twentyitem{1861-1863}{B.Sc. magna cum laude}{Wonderland}{Majoring in Computer Science}
-	\twentyitem{1856-1861}{High school}{Wonderland}{Specializing in mathematics and physics.}
-	%\twentyitem{<dates>}{<title>}{<location>}{<description>}
-\end{twenty}
+\begin{minipage}{0.15\linewidth}
+  Oct. 2014 - Sep. 2017
+\end{minipage} \hfill
+\begin{minipage}{0.85\linewidth}
+  \begin{itemize}
+  \item Ph.D. in Mathematical Statistics
+  \item Copenhagen University
+  \item Areas of research:
+    \begin{itemize}
+      \item {\bf Time series analysis}: ARIMA, GARCH, stochastic volatility
+        models
+      \item {\bf Asset allocation \& risk management}: Principle component
+        analysis; optimization of investor utility functions; extreme
+        value theory
+      \item {\bf Numerical methods}: Importance sampling, numerical
+        optimization, estimation of GARCH(p,q) tail indices
+    \end{itemize}
+  \end{itemize}
+\end{minipage}
 
-%----------------------------------------------------------------------------------------
-%	 PUBLICATIONS
-%----------------------------------------------------------------------------------------
+\vspace{3mm}
+
+\begin{minipage}{0.15\linewidth}
+  Sep. 2011 - June 2014
+\end{minipage} \hfill
+\begin{minipage}{0.85\linewidth}
+  \begin{itemize}
+  \item M.Sc. in Mathematical Physics
+  \item Lund University
+  \item Thesis project: Distribution of eigenvalues of large
+    sample covariance matrices in the infinite dimensional limit with
+    a focus on the influence of autocorrelations and heavy tails. The
+    theoretical derivation is based on the {\it free probability
+      theory} and is checked against {\it intraday tick-by-tick
+      prices} of high-volume Swedish stocks.
+  \item Course work:    
+    \begin{itemize}
+      \item Mathematical methods of Physics and Engineering
+      \item Statistical mechanics, Quantum mechanics, General
+        relativity
+    \end{itemize}
+  \end{itemize}
+\end{minipage}
+
+%% \begin{minipage}[b]{0.2\linewidth}
+%%   Oct. 2014 - Sep. 2017
+%% \end{minipage} \hfill
+%% \begin{minipage}{0.8\linewidth}
+%%   \begin{itemize}
+%%   \item Ph.D. in Mathematical Statistics
+%%   \item Copenhagen University
+%%   \item \emph{portfolio \& risk management} Balbla. Oh my god, I have
+%%     to write a lot more about my research
+%%   \end{itemize}
+%% \end{minipage}
 
@@
 
-\begin{twentyshort} % Environment for a short list with no descriptions
-	\twentyitemshort{1865}{Chapter One, Down the Rabbit Hole.}
-	\twentyitemshort{1865}{Chapter Two, The Pool of Tears.}
-	\twentyitemshort{1865}{Chapter Three,  The Caucus Race and a Long Tale.}
-	\twentyitemshort{1865}{Chapter Four,  The Rabbit Sends a Little Bill.}
-	\twentyitemshort{1865}{Chapter Five,  Advice from a Caterpillar.}
-	%\twentyitemshort{<dates>}{<title/description>}
-\end{twentyshort}
+\begin{itemize}
+\item Davis, R.A., Heiny, J., Mikosch, T. and {\bf Xie, X.},
+  2016. {\it Extreme value analysis for the sample autocovariance
+    matrices of heavy-tailed multivariate time series}.
+  \underline{Extremes}, 19(3), pp.517-547.
 
+  {\small
+  We provide, among others, a limit theory for the largest eigenvalues
+  of a sample covariance matrix of heavy-tailed, possibly linearly
+  dependent data with infinite dimensions. We prove weak convergence of
+  the point process of the largest eigenvalues to Poisson or cluster
+  Poisson processes and, based on this convergence result, derive limit
+  laws of functions of the ordered eigenvalues.}
+
+\item Jan\ss en, A., Mikosch, T., Rezapour, M., {\bf Xie, X},
+  2016. {\it The eigenvalues of the sample covariance matrix of a
+    multivariate heavy-tailed stochastic volatility model}.
+  To appear in the \underline{Bernoulli Journal}.
+
+  {\small
+  We derive multivariate $\alpha$-stable limit distributions of the
+  sample covariance matrix when the sample size was large. We
+  consider two possibilities of heavy-tails: 1. the innovation
+  sequence is heavy tailed; and 2. the volatility sequence is
+  heavy-tailed. While the eigen system of the first case is shown to
+  resemble that of i.i.d data, the eigen system in the second case is
+  found to allow dependencies in their limiting distributions, which
+  are determined by the structure of the volatility sequence.}
+\end{itemize}
+
+\section{Papers in Finalization}
+\begin{itemize}
+\item Mikosch, T., de Vries, C., {\bf Xie, X.}, 2017,
+  {\it Generalized Disappointment Aversion and heavy-tail property of
+    equity return distributions}
+
+  {\small
+    We study the heavy tail properties of real-world equity return
+    distributions and, based on the framework of Generalized
+    Disappointment Aversion, derive how an investor's preference varies
+    with the tail distribution characteristics, namely the tail index and
+    the scale parameter. We find, when the equity return is t-distributed,
+    the GDA preference actually decreases with the degree of freedom of
+    the distribution, i.e. the tail index; we interprete this as a
+    consequence of the symmetry of the distribution. When the equity
+    return is assumed to have a shifted Pareto distribution with
+    independent parameters of the left and the right tails, we find the
+    GDA preference increasing with both the lower tail index and the scale
+    parameter. We attribute this to the risk avert nature of rational
+    investors. By numerical methods, optimal asset allocation is also
+    solved in both cases.
+  }
+\end{itemize}
 %----------------------------------------------------------------------------------------
@@
 
-\section{Awards}
+%% \section{Awards}
 
-\begin{twentyshort} % Environment for a short list with no descriptions
-	\twentyitemshort{1987}{All-Time Best Fantasy Novel.}
-	\twentyitemshort{1998}{All-Time Best Fantasy Novel before 1990.}
-	%\twentyitemshort{<dates>}{<title/description>}
-\end{twentyshort}
+%% \begin{twentyshort} % Environment for a short list with no descriptions
+%% 	\twentyitemshort{1987}{All-Time Best Fantasy Novel.}
+%% 	\twentyitemshort{1998}{All-Time Best Fantasy Novel before 1990.}
+%% 	%\twentyitemshort{<dates>}{<title/description>}
+%% \end{twentyshort}
 
@@
 
-\section{Experience}
+%----------------------------------------------------------------------------------------
+%	 OTHER INFORMATION
+%----------------------------------------------------------------------------------------
+
+%----------------------------------------------------------------------------------------
+%	 SECOND PAGE EXAMPLE
+%----------------------------------------------------------------------------------------
+
+\newpage % Start a new page
+
+\makeprofile % Print the sidebar
+
+\section{Work Experience}
 
@@
 
-%----------------------------------------------------------------------------------------
-%	 OTHER INFORMATION
-%----------------------------------------------------------------------------------------
-
 \section{Other information}
@@
 
-Alice approaches Wonderland as an anthropologist, but maintains a strong sense of noblesse oblige that comes with her class status.
-She has confidence in her social position, education, and the Victorian virtue of good manners.
-Alice has a feeling of entitlement, particularly when comparing herself to Mabel, whom she declares has a ``poky little house," and no toys.
-Additionally, she flaunts her limited information base with anyone who will listen and becomes increasingly obsessed with the importance of good manners as she deals with the rude creatures of Wonderland.
-Alice maintains a superior attitude and behaves with solicitous indulgence toward those she believes are less privileged.
-
-%----------------------------------------------------------------------------------------
-%	 SECOND PAGE EXAMPLE
-%----------------------------------------------------------------------------------------
-
-%\newpage % Start a new page
-
-%\makeprofile % Print the sidebar
-
-%\section{Other information}
-
-%\subsection{Review}
-
-%Alice approaches Wonderland as an anthropologist, but maintains a strong sense of noblesse oblige that comes with her class status.
-She has confidence in her social position, education, and the Victorian virtue of good manners.
-Alice has a feeling of entitlement, particularly when comparing herself to Mabel, whom she declares has a ``poky little house," and no toys.
-Additionally, she flaunts her limited information base with anyone who will listen and becomes increasingly obsessed with the importance of good manners as she deals with the rude creatures of Wonderland.
-Alice maintains a superior attitude and behaves with solicitous indulgence toward those she believes are less privileged.
+Alice approaches Wonderland as an anthropologist, but maintains a
+strong sense of noblesse oblige that comes with her class status. She
+has confidence in her social position, education, and the Victorian
+virtue of good manners. Alice has a feeling of entitlement,
+particularly when comparing herself to Mabel, whom she declares has a
+``poky little house," and no toys. Additionally, she flaunts her
+limited information base with anyone who will listen and becomes
+increasingly obsessed with the importance of good manners as she deals
+with the rude creatures of Wonderland. Alice maintains a superior
+attitude and behaves with solicitous indulgence toward those she
+believes are less privileged.
 

### 2017-04-02 papers/FX/draft.tex
Revised my CV.

@@
 
-\title{Tail Indices and Scale Parameters in Financial Time Series}
+\title{Generalized Disappointment Aversion and heavy-tail property of
+  equity return distributions}
 \author{Thomas Mikosch, Casper de Vries, Xie Xiaolei}
@@
 \begin{abstract}
-We consider an investor with preferences that accord with Generalized
-Disappointment Aversion. Such an investor cares about downside risk and
-we assume he recognizes the heavy tail feature of asset return
-distributions. We argue that when a market is dominated by rational
-investors of this kind, the return distributions of equities that are
-actively traded in this market must have very similar tail indices. We
-also show, in contrast, the scale parameters of the return
-distributions may differ hugely from one another.
+We study the heavy tail properties of real-world equity return
+distributions and, based on the framework of Generalized
+Disappointment Aversion, derive how an investor's preference varies
+with the tail distribution characteristics, namely the tail index and
+the scale parameter. We find, when the equity return is t-distributed,
+the GDA preference actually decreases with the degree of freedom of
+the distribution, i.e. the tail index; we interprete this as a
+consequence of the symmetry of the distribution. When the equity
+return is assumed to have a shifted Pareto distribution with
+independent parameters of the left and the right tails, we find the
+GDA preference increasing with both the lower tail index and the scale
+parameter. We attribute this to the risk avert nature of rational
+investors. By numerical methods, optimal asset allocation is also
+solved in both cases.
+
+%% We consider investors with preferences that accord with Generalized
+%% Disappointment Aversion. Such an investor cares about downside risk and
+%% we assume he recognizes the heavy tail feature of asset return
+%% distributions. We argue that when a market is dominated by rational
+%% investors of this kind, the return distributions of equities that are
+%% actively traded in this market must have very similar tail indices. We
+%% also show, in contrast, the scale parameters of the return
+%% distributions may differ hugely from one another.
+
+%% It is generally believed that the utility function of an
+%% investor is increasing and concave -- as John H. Cochrane
+%% put it, ``the last bite is never as satisfying as the first''
+%% \cite{cochrane2009asset}. This implies investors are more repelled
+%% by large losses than they are tempted by large gains. 
+%% Furthermore, it is a stylized fact that equity return distributions
+%% are heavy-tailed (see e.g. Jansen and de Vries
+%% \cite{jansen1991frequency}).
+%% In particular, the lower-tail of an equity distribution is often heavier
+%% than the upper-tail of the same distribution.
+
+%% We explore in this paper the market implications of investor
+%% preferences and the heavy-tail property of equity return
+%% distributions.
 
@@
 
-In this work empirical results are also presented and they appear
-to support the conclusion that the tail indices are very similar,
-with respect to the confidence bands of estimation.
+%% In this work empirical results are also presented and they appear
+%% to support the conclusion that the tail indices are very similar,
+%% with respect to the confidence bands of estimation.
 \end{abstract}
@@
 put it, ``the last bite is never as satisfying as the first''
-\cite{cochrane2009asset}. This implies investors are repelled to a
-larger extent by large losses than they are tempted by large gains. 
+\cite{cochrane2009asset}. This implies investors are more repelled
+by large losses than they are tempted by large gains. 
 Furthermore, it is a stylized fact that equity return distributions
@@
 distributions. The rest of this paper is organized into the following 
-sections:
-In section \ref{sec:empirical_results} we present a survey of
-statistical characteristics of equity returns. Then 
+sections: In section \ref{sec:empirical_results} we present a survey
+of statistical characteristics of equity returns. Then 
 in section \ref{sec:GDA}, we explicitly recognize the behavioral
@@
 (1991) \cite{gul1991theory}. The Gul paper derives the DA from first
-principles (axiomatic).
-In this section we infer characteristics of equity returns from the
-GDA theory and compare them with the empirical results.
+principles (axiomatic). In this section we infer characteristics of
+equity returns from the GDA theory and compare them with the empirical
+results.
 

### 2017-04-02 cpp/ImportanceSampling/plot_data.r
Added some text about the estimation of xi.

@@ -13,7 +13,7 @@ graphics.off();
 ## old <- read.table(file="./lambda0.txt");
 ## new <- read.table(file="./lambda0_new.txt");
 data <- read.table("DAX_garch21_tailindex.dat");
-pdf("../../papers/Jeffrey1/DAX_xi.pdf");
+pdf("../../papers/Jeffrey1/SP500_xi.pdf");
 plot(data[, 1], data[, 2],
      ##     main=expression(sigma[t]^2 == 0.11 * R[t-1]^2 + 0.88 * sigma[t-1]^2 + 10^{-7}
      type="l", xlab=expression(alpha),

### 2017-04-04 papers/FX/draft.tex
In the process of merging Thomas' re-written introduction.

--- 2017-03-31 papers/FX/draft.tex
@@
-\documentclass{article}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{enumerate}
-\usepackage[bookmarks=true]{hyperref}
-\usepackage{bookmark}
+% \documentclass{article}
+% \usepackage{amsmath}
+% \usepackage{amsthm}
+% \usepackage{enumerate}
+% \usepackage{bookmark}
+% \usepackage{graphicx}
+% \usepackage{multirow}
+% \usepackage{adjustbox}
+% \usepackage{wrapfig}
+% \usepackage{subcaption}
+% \usepackage{color}
+% \usepackage[dvipsnames]{xcolor}
+
+% \usepackage{amssymb,amsmath,amsthm,amsfonts}
+% \usepackage{mathrsfs}
+% \usepackage{dsfont}
+% \usepackage{enumerate}
+
+% \input{physics_common.tex}
+
+\documentclass[11pt,a4]{amsart}
+%\usepackage{amsmath}
+%\usepackage{amsthm}
+%\usepackage{enumerate}
+%\usepackage[bookmarks=true]{hyperref}
+%\usepackage{bookmark}
 \usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{adjustbox}
-\usepackage{wrapfig}
-\usepackage{subcaption}
+\usepackage[bookmarks=true]{ hyperref}
+%\usepackage{multirow}
+%\usepackage{adjustbox}
+%\usepackage{wrapfig}
+%\usepackage{subcaption}
 \usepackage{color}
@@
 \usepackage{amssymb,amsmath,amsthm,amsfonts}
-\usepackage{mathrsfs}
-\usepackage{dsfont}
-\usepackage{enumerate}
+%\usepackage{mathrsfs}
+%\usepackage{dsfont}
+%\usepackage{enumerate}
 
 %\input{../../physics_common.tex}
-\input{physics_common.tex}
-
-\title{Tail Indices and Scale Parameters in Financial Time Series}
-\author{Thomas Mikosch, Casper de Vries, Xie Xiaolei}
+%\input{physics_common.tex}
+
+\textwidth 6.50in
+\topmargin -0.50in
+\oddsidemargin 0in
+\evensidemargin 0in
+\textheight 9.00in
+%\pagestyle{plain}
+\definecolor{darkblue}{rgb}{.2, 0.2,.8}
+\definecolor{darkgreen}{rgb}{0,0.5,0.3}
+\definecolor{darkred}{rgb}{.8, .1,.1}
+\newcommand{\pd}{\partial}
+\newcommand{\red}{\color{darkred}}
+\newcommand{\blue}{\color{darkblue}}
+\newcommand{\green}{\color{darkgreen}}
+\newcommand{\black}{\color{black}}
+\newcommand{\levy}{L\'evy}
+\newcommand{\slln}{strong law of large numbers}
+\newcommand{\clt}{central limit theorem}
+\newcommand{\sde}{stochastic \de}
+\newcommand{\de}{differential equation}
+\newcommand{\mregvar}{multivariate \regvar}
+\newcommand{\garch}{{\rm GARCH}$(1,1)$}
+\newcommand{\spr}{stochastic process}
+\newcommand{\It}{It\^o}
+\newcommand{\sta}{St\u aric\u a}
+\newcommand{\ex}{{\rm e}\,}
+\newcommand{\sas}{s$\alpha$s}
+\newcommand{\civ}{\stackrel{v}{\rightarrow}}
+\newcommand{\asy}{asymptotic}
+\newcommand {\hatt} {\hat{\theta}}
+\newcommand {\hattg} {\hat{\theta}_{MLE}}
+\newcommand {\hattz} {\hat{\theta}_0}
+\newcommand {\hatb} {\hat{\beta}}
+\newcommand {\hatbg} {\hat{\beta}_G}
+\newcommand {\spp} {^\prime}
+\newcommand{\ts}{time series}
+\newcommand{\tsa}{\ts\ analysis}
+\newcommand{\garchpq}{{\rm GARCH}$(p,q)$}
+\def\theequation{\thesection.\arabic{equation}}
+\def\tag{\refstepcounter{equation}\leqno }
+\def\neqno{\refstepcounter{equation}\leqno(\thesection.\arabic{equation})}
+\newtheorem{lemma}{Lemma}[section]
+\newtheorem{figur}[lemma]{Figure}
+\newtheorem{theorem}[lemma]{Theorem}
+\newcommand{\Leb}{{\mathbb L}{\mathbb E}{\mathbb B}}
+\newcommand{\bbc}{{\mathbb C}}
+\newtheorem{proposition}[lemma]{Proposition}
+\newtheorem{definition}[lemma]{Definition}
+\newtheorem{corollary}[lemma]{Corollary}
+\newtheorem{example}[lemma]{Example}
+\newtheorem{exercise}[lemma]{Exercise}
+\newtheorem{remark}[lemma]{Remark}
+\newtheorem{fig}[lemma]{Figure}
+\newtheorem{tab}[lemma]{Table}
+\newtheorem{hypothesis}{Hypothesis}
+\newcommand{\play}{\displaystyle}
+\newcommand{\rmk}{\noindent R{\sc emark}}
+\newcommand{\cid}{\stackrel{d}{\rightarrow}}
+\newcommand{\cip}{\stackrel{P}{\rightarrow}}
+\newcommand{\reference}{\hangindent=5pc\hangafter=1}
+\newcommand{\MC}{Markov chain}
+\newcommand{\bfQ}{{\bf Q}}
+\newcommand{\bfM}{{\bf M}}
+\newcommand{\bfu}{{\bf u}}
+\newcommand{\bfv}{{\bf v}}
+\newcommand{\bfV}{{\bf V}}
+\newcommand{\bfU}{{\bf U}}
+\newcommand{\bfR}{{\bf R}}
+\newcommand{\bfC}{{\bf C}}
+\newcommand{\bfF}{{\bf F}}
+\newcommand{\RV}{{\rm RV}}
+\newcommand{\LN}{{\rm LN}}
+\newcommand{\WE}{{\rm WE}}
+\newcommand{\bth}{\begin{theorem}}
+\newcommand{\ethe}{\end{theorem}}
+\newcommand{\sv}{stochastic volatility}
+\newcommand{\bre}{\begin{remark}\em }
+\newcommand{\ere}{\end{remark}}
+\newcommand{\fre}{frequenc}
+\newcommand{\ble}{\begin{lemma}}
+\newcommand{\ele}{\end{lemma}}
+\newcommand{\sre}{stochastic recurrence equation}
+\newcommand{\pp}{point process}
+\newcommand{\bde}{\begin{definition}}
+\newcommand{\ede}{\end{definition}}
+\newcommand{\bco}{\begin{corollary}}
+\newcommand{\eco}{\end{corollary}}
+
+\newcommand{\bpr}{\begin{proposition}}
+\newcommand{\epr}{\end{proposition}}
+
+\newcommand{\bexer}{\begin{exercise}}
+\newcommand{\eexer}{\end{exercise}}
+
+\newcommand{\bexam}{\begin{example}}
+\newcommand{\eexam}{\end{example}}
+
+\newcommand{\bfi}{\begin{fig}}
+\newcommand{\efi}{\end{fig}}
+
+\newcommand{\btab}{\begin{tab}}
+\newcommand{\etab}{\end{tab}}
+%\renewcommand{\thetab}{\arabic{section}.9.\arabic{tab}}
+
+\newcommand{\edf}{empirical distribution function}
+\newcommand{\epro}{empirical process}
+\newcommand{\per}{periodogram}
+\newcommand{\pf}{{\bf Proof.}}
+\newcommand{\lhs}{left-hand side}
+\newcommand{\fidi}{finite-dimensional distribution}
+\newcommand{\rv}{random variable}
+\newcommand{\DA}{{\rm DA}}
+\newcommand{\DNA}{{\rm DNA}}
+\newcommand{\MDA}{{\rm MDA}}
+\newcommand{\sign}{{\rm sign}}
+\newcommand{\PRM}{{\rm PRM}}
+\newcommand{\card}{{\rm card}}
+\newcommand{\var}{{\rm var}}
+\newcommand{\med}{{\rm med}}
+\newcommand{\cov}{{\rm cov}}
+\newcommand{\corr}{{\rm corr}}
+\newcommand{\as}{{\rm a.s.}}
+\newcommand{\io}{{\rm i.o.}}
+\newcommand{\Pois}{\rm Poisson}
+
+\newcommand{\bfth}{\mbox{\boldmath$\theta$}}
+\newcommand{\bfTh}{\mbox{\boldmath$\Theta$}}
+\newcommand{\bfmu}{\mbox{\boldmath$\mu$}}
+
+\newcommand{\bfvep}{\mbox{\boldmath$\vep$}}
+\newcommand{\rhs}{right-hand side}
+\newcommand{\df}{distribution function}
+\newcommand{\dsum}{\displaystyle\sum}
+%\newcommand{\dfrac}{\displaystyle\frac}
+\newcommand{\dint}{\displaystyle\int}
+\newcommand{\dlim}{\displaystyle\lim\limits}
+\newcommand{\dmax}{\displaystyle\max\limits}
+\newcommand{\dsup}{\displaystyle\sup\limits}
+\newcommand{\dmin}{\displaystyle\min\limits}
+\newcommand{\dprod}{\displaystyle\prod}
+
+\newcommand{\beao}{\begin{eqnarray*}}
+\newcommand{\eeao}{\end{eqnarray*}\noindent}
+
+\newcommand{\beam}{\begin{eqnarray}}
+\newcommand{\eeam}{\end{eqnarray}\noindent}
+
+\newcommand{\beqq}{\begin{equation}}
+\newcommand{\eeqq}{\end{equation}\noindent}
+
+\newcommand{\bce}{\begin{center}}
+\newcommand{\ece}{\end{center}}
+
+\newcommand{\barr}{\begin{array}}
+\newcommand{\earr}{\end{array}}
+\newcommand{\cadlag}{c\`adl\`ag}
+\newcommand{\tep}{\tilde\epsilon}
+\newcommand{\stp}{\stackrel{\P}{\rightarrow}}
+\newcommand{\std}{\stackrel{d}{\rightarrow}}
+\newcommand{\stas}{\stackrel{\rm a.s.}{\rightarrow}}
+\newcommand{\stj}{\stackrel{J_1}{\rightarrow}}
+\newcommand{\stv}{\stackrel{v}{\rightarrow}}
+\newcommand{\stw}{\stackrel{w}{\rightarrow}}
+\newcommand{\stww}{\stackrel{\wh w}{\rightarrow}}
+\newcommand{\eqd}{\stackrel{d}{=}}
+
+\newcommand{\refbib}{\cite}
+\newcommand{\vague}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
+                    \it{v} $}}{\rightarrow}}
+\newcommand{\weak}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
+                    \it{w} $}}{\rightarrow}}
+\newcommand{\what}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
+                    \it{\hat{w}} $}}{\rightarrow}}
+
+\newcommand{\bdis}{\begin{displaymath}}
+\newcommand{\edis}{\end{displaymath}\noindent}
+
+\newcommand{\Thetabold}{{\boldsymbol{\Theta}}}
+\newcommand{\Prob}{\operatorname{P}}
+\newcommand{\Dspace}{\mathbb{D}}
+\newcommand{\Espace}{\mathbb{E}}
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\Xbold}{{\mathbf{X}}}
+\newcommand{\xbold}{{\mathbf{x}}}
+\newcommand{\Ybold}{{\mathbf{Y}}}
+\newcommand{\ybold}{{\mathbf{y}}}
+\newcommand{\zerobold}{{\mathbf{0}}}
+\newcommand{\Sphere}{\mathbb{S}}
+\newcommand{\Dbar}{\overline{\mathbb{D}}_{0}}
+\newcommand{\ind}{independent}
+\newcommand{\w}{\omega}
+\newcommand{\W}{\Omega}
+\newcommand{\bw}{\bigwedge}
+\newcommand{\bv}{\bigvee}
+\newcommand{\nto}{n\to\infty}
+\newcommand{\kto}{k\to\infty}
+\newcommand{\xto}{x\to\infty}
+\newcommand{\tto}{T\to\infty}
+\newcommand{\uto}{u\to\infty}
+\newcommand{\ov}{\overline}
+\newcommand{\wt}{\widetilde}
+\newcommand{\wh}{\widehat}
+\newcommand{\vep}{\varepsilon}
+\newcommand{\ep}{\epsilon}
+\newcommand{\vap}{\varphi}
+\newcommand{\la}{\lambda}
+\newcommand{\La}{\Lambda}
+
+\newcommand{\halmos}{\hfill \qed}
+\newcommand{\regvary}{regularly varying}
+\newcommand{\slvary}{slowly varying}
+\newcommand{\regvar}{regular variation}
+%\newcommand{\bbr}{{I\!\!R}}
+%\newcommand{\bbn}{{I\!\!N}}
+%\newcommand{\bbz}{{I\!\!\!Z}}
+
+\newcommand{\ka}{\kappa}
+\newcommand{\bba}{{\mathcal A}}
+\newcommand{\bbe}{{\ov \bbr^{d}_0}}
+\newcommand{\bfT}{{\bf T}}
+\newcommand{\bbr}{{\mathbb R}}
+\newcommand{\bbm}{{\mathcal M}}
+\newcommand{\bbb}{{\mathcal B}}
+\newcommand{\bbz}{{\mathbb Z}}
+\newcommand{\bbn}{{\mathbb N}}
+\newcommand{\bbd}{{\mathbb D}}
+\newcommand{\bbf}{{\mathcal F}}
+\newcommand{\bbs}{{\mathbb S}}
+\newcommand{\bbE}{{\mathcal E}}
+
+\newcommand{\BM}{Brownian motion}
+\newcommand{\BB}{Brownian bridge}
+\newcommand{\con}{convergence}
+\newcommand{\ev}{extreme value}
+\newcommand{\evt}{extreme value theory}
+\newcommand{\evd}{extreme value distribution}
+\newcommand{\clp}{central limit problem}
+\newcommand{\nc}{norming constant}
+\newcommand{\st}{such that}
+\newcommand{\fif}{if and only if}
+\newcommand{\wrt}{with respect to}
+\newcommand{\chf}{characteristic function}
+\newcommand{\fct}{function}
+\newcommand{\mdoa}{maximum domain of attraction}
+\newcommand{\ds}{distribution}
+\newcommand{\gev}{generalized extreme value distribution}
+\newcommand{\gpd}{generalized Pareto distribution}
+\newcommand{\pwm}{probability-weighted moments}
+\newcommand{\mle}{maximum likelihood estimator}
+\newcommand{\mel}{mean excess function}
+\newcommand{\an}{asymptotically normal}
+\newcommand{\ac}{absolutely continuous}
+\newcommand{\rep}{representation}
+\newcommand{\cmt}{continuous mapping theorem}
+\newcommand{\seq}{sequence}
+\newcommand{\pr} {{\bf Proof.}}
+\newcommand{\ins}{insurance}
+\newcommand{\mat}{mathematic}
+\newcommand{\pro}{probabilit}
+\newcommand{\pros}{probabilities}
+\newcommand{\ms}{measure}
+\newcommand{\mgf}{moment generating function}
+\newcommand{\ld}{large deviation}
+\newcommand{\bfx}{{\bf x}}
+\newcommand{\bfX}{{\bf X}}
+\newcommand{\bfB}{{\bf B}}
+\newcommand{\bfY}{{\bf Y}}
+\newcommand{\bfy}{{\bf y}}
+\newcommand{\bfA}{{\bf A}}
+\newcommand{\bfG}{{\bf G}}
+\newcommand{\bfO}{{\bf 0}}
+\newcommand{\bfZ}{{\bf Z}}
+\newcommand{\bfz}{{\bf z}}
+\newcommand{\bfS}{{\bf S}}
+\newcommand{\bfa}{{\bf a}}
+\newcommand{\bfb}{{\bf b}}
+\newcommand{\bfe}{{\bf e}}
+\newcommand{\bft}{{\bf t}}
+\newcommand{\bfI}{{\bf I}}
+\newcommand{\bfs}{{\bf s}}
+\newcommand{\bfh}{{\bf h}}
+\newcommand{\bfunit}{{\bf h}}
+\newcommand{\ARCH}{{\rm ARCH}}
+\newcommand{\E }{{\mathbb E}}
+\renewcommand{\P }{{\mathbb P}}
+\newcommand{\Q }{{\mathbb Q}}
+\newcommand{\1}{{\mathbf 1}}
+
+\newcommand{\floor}[1]{
+  \lfloor #1\rfloor
+}
+\newcommand{\ceil}[1]{
+  \lceil #1\rceil
+}
+
+\allowdisplaybreaks
+\begin{document}
+\title{Tail-indices and scale parameters of returns series}
+\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei, }
 \date{\today}
-\begin{document}
+
 \maketitle
@@
 \begin{abstract}
-We consider an investor with preferences that accord with Generalized
-Disappointment Aversion. Such an investor cares about downside risk and
-we assume he recognizes the heavy tail feature of asset return
-distributions. We argue that when a market is dominated by rational
-investors of this kind, the return distributions of equities that are
-actively traded in this market must have very similar tail indices. We
-also show, in contrast, the scale parameters of the return
-distributions may differ hugely from one another.
-
-On the other hand, whether or not all equities in a multivariate model
-have the same tail index is a dividing issue for multivariate GARCH
-models proposed in the literature. Therefore, it is important to analyze
-data of real equity returns and see how close to each other the
-tail indices actually are.
-
-In this work empirical results are also presented and they appear
-to support the conclusion that the tail indices are very similar,
-with respect to the confidence bands of estimation.
+%We consider an investor with preferences that accord with Generalized
+%Disappointment Aversion. Such an investor cares about downside risk and
+%we assume he recognizes the heavy tail feature of asset return
+%distributions. We argue that when a market is dominated by rational
+%investors of this kind, the return distributions of equities that are
+%actively traded in this market must have very similar tail indices. We
+%also show, in contrast, the scale parameters of the return
+%distributions may differ hugely from one another.
+%On the other hand, whether or not all equities in a multivariate model
+%have the same tail index is a dividing issue for multivariate GARCH
+%models proposed in the literature. Therefore, it is important to analyze
+%data of real equity returns and see how close to each other the
+%tail indices actually are.
+%In this work empirical results are also presented and they appear
+%to support the conclusion that the tail indices are very similar,
+%with respect to the confidence bands of estimation.
 \end{abstract}
 
-\section{Introduction}
-It is generally believed that the utility function of an
-investor is increasing and concave -- as John H. Cochrane
-put it, ``the last bite is never as satisfying as the first''
-\cite{cochrane2009asset}. This implies investors are repelled to a
-larger extent by large losses than they are tempted by large gains. 
-Furthermore, it is a stylized fact that equity return distributions
-are heavy-tailed (see e.g. Jansen and de Vries \cite{jansen1991frequency}).
-In particular, the lower-tail of an equity distribution is often heavier
-than the upper-tail of the same distribution.
-
-we explore in this paper the market implications of investor
-preferences and the heavy-tail property of equity return
-distributions. The rest of this paper is organized into the following
-sections:
-In section \ref{sec:empirical_results} we present a survey of
-statistical characteristics of equity returns. Then 
-in section \ref{sec:GDA}, we explicitly recognize the behavioral
-concern for downside risk in the investor's evaluation of portfolios
+\section{Introduction}\setcounter{equation}{0}
+It is one of the stylized facts of financial econometrics that 
+returns of speculative prices are {\em heavy-tailed.} There is no agreement
+in the literature how heavy tails really are. For example, Barndorff-Nielsen and Shephard
+\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails, i.e.,
+tails which are comparable with those of a gamma \ds . On the other
+hand, tails of returns have been studied in great detail 
+in the extreme value community. Among extreme value specialists
+there is general agreement that returns $X_t$ have tails of power-law-type, i.e.,
+\beam\label{eq:1}
+\P(X_t>x)\sim c_+ \,x^{\alpha_{\rm up}}\quad\mbox{and}\quad \P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
+\eeam  
+where $c_{\pm}$ are positive constants.\footnote{Here and in what
+  follows, $f(x)\sim g(x)$ for positive \fct s $f$ and $g$ means 
+that $f(x)/g(x)\to 1$ as $\xto$.} See for example, Embrechts et
+al. \cite{embrechts:klueppelberg:mikosch:1997}, Jansen and de Vries
+\cite{jansen1991frequency}, Mikosch \cite{mikosch:2003}, Resnick
+\cite{resnick:2007}.  
+In the extreme value literature it is common to replace the constants $c_\pm$ by a suitable
+{\em slowly varying} \fct s; cf. Embrechts et
+al. \cite{embrechts:klueppelberg:mikosch:1997}, Chapter 3. In this
+paper, for the  
+sake of argument, we stick to the condition \eqref{eq:1}.
+\par
+There are some good theoretical reasons for the appearance of power-law tails in situations where certain moments 
+of data are believed to be infinite. Tails of the type \eqref{eq:1}
+describe the maximum domain of attraction of the  
+Fr\'echet \ds\ $\Phi_\alpha(x)=\exp(-x^{-\alpha})$ for $x>0$. Equivalently, power-law tails are prescribed by the 
+generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e.,
+\beao
+\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha)^{-\alpha}\,,\qquad u\to\infty\,.
+\eeao
+The aforementioned results 
+are considered very natural for iid and weakly dependent strictly stationary \seq s of random variables $(X_t)$; 
+in the world of extremes they are the analogs
+of the \clt\ from the world of sums.
+\par
+In the literature on extremes for return data one finds the 
+statement that {\em estimated} values $\hat \alpha_{\rm up}$ and $\hat \alpha_{\rm low}$ 
+of the tail-indices $\alpha_{\rm up}$ and $\alpha_{\rm low}$,
+respectively, typically have the tendency that $\hat \alpha_{\rm
+  up}>\hat \alpha_{\rm low}$. 
+This observation is often explained by the fact that investors are
+more prone to negative than to positive news in te market. 
+Moreover, in the literature the {\em estimated} tail-indices $\hat \alpha$ (both in the left and right tails) 
+are typically found in the range $(2.5,4)$. For an illustration, see
+Figure~\ref{fig:1} where estimates $\hat \alpha_{\rm low}$ 
+in three sectors of the Standard \& Poor 500 index  are shown. The
+estimates are based on 1304 observations of daily return 
+data from 1 January 2010 to 1 January 2015.
+\par
+When looking at Figure~\ref{fig:1} one might ask the following questions:
+\begin{itemize}
+\item
+In view of the wide \asy\ confidence bands of the estimators of tail-indices, 
+are the estimated tail-indices from different series really distinct?
+\item
+Are there some {\em theoretical} reasons in favor of the fact that the tail indices from different series are {\em not} 
+distinct?
+\end{itemize}
+In this paper, we try to find some answers to these questions. 
+\par
+The estimator of the tail-index $\alpha$  which favored in the literature  is the {\em Hill estimator}; the graphs in 
+Figure~\ref{fig:1} are based on this estimator. We introduce this estimator in Section~\ref{sec:1} and discuss 
+some of its virtues and vices. In addition to tail-index estimation we also discuss the related problem of
+estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). It turns out 
+that the simulateneous estimation of the tail-index and the scale parameter are strongly related.
+\par
+In Section~\ref{sec:2} we discuss the theoretical problem of appearance of power-law tails in models 
+for daily or, more generally, low \fre y 
+return data. In particular, we address the power-law tails of univariate and multivariate GARCH models as potential models
+for a set of return data from distinct assets. As a matter of fact, under mild conditions, the aforementioned models
+have power-tails due to their relation with so-called {\em \sre s}.
+\par
+In Section~\ref{sec:3} we discuss an economic argument for the fact that return data of similar assets
+(such as return series in a given sector of the S\&P 500 index) may have tail indices which are close to each other.
+We argue based on a  utility \fct\ approach. We explicitly recognize the behavioral
+concern for downside risk in an investor's evaluation of a portfolio
 using the framework of Generalized Disappointment Aversion (GDA)
@@
 is an extension of the concept of Disappointment Aversion (DA) of Gul
-(1991) \cite{gul1991theory}. The Gul paper derives the DA from first
-principles (axiomatic).
-In this section we infer characteristics of equity returns from the
-GDA theory and compare them with the empirical results.
-
-%% Secondly, the empirical fact that the loss return distribution is
-%% non-normally fat tailed distributed is taken into account.
-%% If investors and speculators alike display concern for downside risk,
-%% as formulated by a utility function, it becomes important to model this 
-%% risk adequately. see e.g. Jansen and de Vries (1991), exhibiting power
-%% like tails rather than exponential tails. The risk is determined both
-%% by the coefficient that regulates the hyperbolic power and the
-%% scale. We reason how both determine the risk of investment apart from
-%% the mean return.
-
-%% \section{Notations}
-%% Here is a summary of the notations that we use in the rest of the paper.
-%% In this summary $x_F$ denotes a quantity whose value that should be
-%% obvious from the context where the notation in question is used.
-%% \begin{itemize}
-%% \item $\sim$ stands for ``asymptotically equivalent to'', i.e.
-%%   $F(x) \sim G(x)$ means
-%%   \[
-%%   \lim_{x \to x_F} {F(x) \over G(x)} = 1
-%%   \]
-%% \item $O(F(x))$ stands for terms of the same order as $F(x)$, i.e.
-%%   \[
-%%   \lim_{x \to x_F } {O(F(x)) \over F(x)} = \text{constant}
-%%   \]
-%% \item $o(F(x))$ stands for terms of higher orders than $F(x)$, i.e.
-%%   \[
-%%   \lim_{x \to x_F } {o(F(x)) \over F(x)} = 0
-%%   \]
-%% \item $\hot$ stands for ``higher order terms'', i.e.
-%%   \[
-%%   \lim_{x \to x_F} {F(x) + \hot \over F(x)} = 1
-%%   \]
-%% \end{itemize}
-
-%% \section{Implications of heavy-tailed equity return}
-%% \label{sec:1}
-%% In this section we consider a rather general utility functional
-%% $U(F)$ associated with the distribution function $F(\cdot)$ of
-%% an equity return $X$. We define $U(F)$ as
-%% \[
-%% U(F) = \E u(X)
-%% \]
-%% for some function $u(\cdot)$ satisfying the following conditions:
-%% \begin{enumerate}
-%% \item $u(\cdot)$ is defined on $\mathbb R$, is increasing, concave
-%%   and differentiable.
-%% \item $u(\cdot)$ is independent of the distribution of $X$
-%% \item $\lim_{x \to \infty} u(x) = \bar u$, where $\bar u$ is a
-%%   constant
-%% \end{enumerate}
-%% %% An exampple of such utility functions is
-%% %% \[
-%% %% u(x) = \bar u - e^{-\lambda x}
-%% %% \]
-%% %% for some $\lambda > 0$.
-%% In addition we assume $F(\cdot)$ has the following properties:
-%% \begin{enumerate}
-%% \item $\lim_{x \to -\infty} F(x) u(x) = 0$
-%% \item $\lim_{x \to -\infty} {|x|^\alpha F(x) \over L(|x|)} = 1$
-%% \item $\lim_{x \to \infty} {x^\alpha (1 - F(x)) \over L(x)} = 1$
-%% \end{enumerate}
-%% for some $\alpha > 1$ and some slowly varying function $L(x) > 0$
-%% defined on $\mathbb R_+$.
-%% $\E u(X)$ can be computed as follows
-%% \begin{eqnarray*}
-%%   && \E u(X) \\
-%%   &=& \int_{-\infty}^{\infty} u(x) d F(x) \\
-%%   &=& \left. u(x) F(x) \right|_{-\infty}^{\infty}
-%%   - \int_{-\infty}^\infty F(x) u'(x) dx \\
-%%   &=& \bar u - \left(
-%%   \int_{-\infty}^{-M} + \int_{-M}^{M} + \int_{M}^{\infty}
-%%   \right) F(x) u'(x) dx
-%% \end{eqnarray*}
-%% As $M \to \infty$ we have for the lower-tail integral
-%% \begin{eqnarray*}
-%%   && \int_{-\infty}^{-M} F(x) u'(x) dx \\
-%%   &\sim& \int_{-\infty}^{-M} L(|x|) |x|^{-\alpha} u'(x) dx \\
-%%   &=& \int_{M}^{\infty} L(x) x^{-\alpha} u'(-x) dx \\
-%% \end{eqnarray*}
-%% Applying the same calculation to the upper-tail integral and
-%% using $1 - F(x) \sim L(x) x^{-\alpha}$ yields
-%% \begin{eqnarray}
-%%   && \E u(X) \nonumber \\
-%%   &=& \bar u - \int_{M}^\infty L(x) x^{-\alpha} u'(-x) dx
-%%   -\int_{-M}^M F(x) u'(x) dx \nonumber \\
-%%   && -\int_{M}^\infty [1 - L(x) x^{-\alpha}] u'(x) dx
-%%   + \hot \nonumber \\
-%%   &=& u(M) - \int_{M}^\infty L(x) x^{-\alpha} [u'(-x) - u'(x)] dx
-%%   + \hot \nonumber \\
-%%   && - \int_{-M}^M F(x) u'(x) dx \label{eq:xxie0}
-%% \end{eqnarray}
-%% In the simple case where $L(x) = A$, i.e. a constant, we have
-%% \begin{eqnarray}
-%%   \pd{\E u(X)}{\alpha} &=& \int_M^{\infty} A x^{-\alpha} \ln(x)
-%%      [u'(-x) - u'(x)] dx +\hot \nonumber \\
-%%      \pd{\E u(X)}{A} &=& -\int_M^{\infty} x^{-\alpha} [u'(-x) - u'(x)] dx
-%%      + \hot
-%%   \label{eq:xxie01}   
-%% \end{eqnarray}
-%% By concavity $u'(-x) > u'(x)$ for $x \in (M, \infty)$. So for large
-%% $M$,
-%% \[
-%% \pd{\E u(X)}{\alpha} > 0, \quad \pd{\E u(X)}{A} < 0
-%% \]
-%% %% \begin{eqnarray*}
-%% %%   \pd{\E u(X)}{\alpha} &>& 0 \\
-%% %%   \pd{\E u(X)}{A} &<& 0
-%% %% \end{eqnarray*}
-%% Now that each and every equity is assigned a score by the utility
-%% functional, all equities actively traded in a rational market that
-%% is dominated by agents exhibiting a utility function must have almost
-%% equal scores -- otherwise trading will be concentrated on those
-%% with top scores.
-
-%% Suppose the equity with the highest score in the market has parameters
-%% $(\alpha, A)$. We would like to find out how these parameters vary across
-%% equities that score almost as high. Consider a parameter pair
-%% $(\alpha + \Delta \alpha, A + \Delta A)$. If it scores the same as
-%% $(\alpha, A)$, we have
-%% \[
-%% 0 = \Delta \E u(X) = \Delta \alpha \pd{\E u(X)}{\alpha}
-%% + \Delta A \pd{\E u(X)}{A} + \hot
-%% \]
-%% Discarding higher order terms and plugging \eqref{eq:xxie01} into
-%% this equation gives
-%% \[
-%%   {\Delta \alpha \over \Delta A}
-%%   =
-%%   {
-%%     \int_M^{\infty} x^{-\alpha} [u'(-x) - u'(x)] dx    
-%%     \over
-%%     \int_M^{\infty} A x^{-\alpha} \ln(x) [u'(-x) - u'(x)] dx    
-%%   }
-%% \]
-%% Taking the limit as $M \to \infty$, we get
-%% \[
-%% \lim_{M \to \infty} {\Delta \alpha \over \Delta A}
-%% =
-%% \lim_{M \to \infty} {1 \over A \ln(M)} = 0
-%% \]
-%% We see from the last equation that in the configuration space around
-%% $(\alpha, A)$, a small variation of $\alpha$ induces a tremendous
-%% variation of $A$. Furthermore, this conclusion does not depend on the
-%% choice of the function $u(\cdot)$. Thus in a market dominated by rational
-%% agents guided by their utility functionals, we can expect to see very
-%% similar tail indices but hugely different scale parameters in equity
-%% return distributions.
-
-%% %% Due to concavity, $u'(-x) > u'(x)$ on the interval $(M, \infty)$.
-%% %% The central region integral on $(-M, M)$ does not depend on $\alpha$.
-%% %% So $\E u(X)$ is monotonically increasing with $\alpha$. Trying to
-%% %% maximize his utility function, an agent will seek the largest $\alpha$
-%% %% possible in the market. As a result, all equities traded with significant
-%% %% volumes must eventually become so close that they are indistinguishable
-%% %% by available statistical methods.
-
-%% In section \ref{sec:GDA}, we compute the utility function within the
-%% elaborate framework of {\it Generalized Disappointment Aversion} and
-%% show that the same conclusion stands -- tail indices are similar while
-%% scale parameters can be drastically different.
-
-\section{Empirical Results}
-\label{sec:empirical_results}
-As has been mentioned, we are interested in the heavy-tail properties
-of equity returns. Let $X_t$ be an equity return series. If the
-distribution of $X_t$ is heavy-tailed, we can write
-\begin{eqnarray*}
-  \lim_{x \to \infty} {x^{\beta}\P(X_t > x) \over L_+(x)} &=& 1 \\
-  \lim_{x \to \infty} {x^{\alpha}\P(X_t < -x) \over L_-(x)} &=& 1
-\end{eqnarray*}
-where the constants $\alpha$ and $\beta$ are left and right tail
-indices, respectively, and $L_+(x)$ and $L_-(x)$ are slowly varying
-functions satisfying
-\[
-\lim_{x \to \infty} {L_*(a x) \over L_*(x)} = 1
-\]
-for all $a > 0$. Here ``*'' can be either ``+'' or ``-''. In the
-simplest case, $L_+(x)$ and $L_(x)$ are constants and the distribution
-has power-law tails. Let $J_*(1/x) = L_*(x)$. A Taylor expansion at
-$x=\infty$ gives
-\begin{equation}
-  \label{eq:scale_parameter}
-  J_*(1/x) = J_*(0) + {J'_*(0) \over x} + O(x^{-2})
-\end{equation}
-Thus
-\begin{equation}
-  \label{eq:lower_tail_expansion}
-  x^{-\alpha} L_-(x) =
-  J_-(0) x^{-\alpha} + J'_-(0) x^{-\alpha - 1} + O(x^{-\alpha-2})
-\end{equation}
-when $x \to \infty$. Clearly the tail indices $\alpha$ and $\beta$ as
-well as the scale parameters $J_-(0)$ and $J_+(0)$ play crutial roles
-in determining the tail distribution and hence in asset allocation and
-risk management. These parameters of the equities in a market are
-characteristics of the market and provide a useful description of the
-market. One can also use them to help build a model of the equities in
-the market. With these motives in mind, we present a survey of the
-tail indices and scale parameters of 3 sectors, namely ``Energy'',
-``Consumer Staples'' and ``Information Technology'', of the S\&P 500
-index.
-
-\subsection{Hill estimates of lower tail indices}
-\label{sec:Hill}
-There have been several estimators for the tail index proposed in the
-literature, e.g. Hill (1975) \cite{hill1975simple}, Pikands (1975)
-\cite{pickands1975statistical}, Dekkers and Einmahl and De Haan (1989)
-\cite{dekkers1989moment}. See also Embrechts and Kl\"uppelberg and
-Mikosch (1997)\cite{Embrechts1997}. The most popular of these is
-arguably Hill (1975) \cite{hill1975simple}:
-\begin{equation}
-  \label{eq:Hill_index}
-  \hat \alpha = \left[
-    {1 \over k} \sum_{i=1}^k \ln {x_{(i)} \over x_{(k+1)}}
-    \right]^{-1}
-\end{equation}
-where $x_{(i)}$ is the $i$-th upper order statistic of a sample
-$x_1, x_2, ..., x_n$. It has been shown that the Hill estimator is
-applicable to 
-\begin{itemize}
-\item iid random variables (Mason (1982) \cite{mason1982laws})
-\item a weakly dependent sequence (Rootz{\'e}n, Leadbetter,
-  and de Haan \cite{rootzen1992tail})
-\item a linear process (Resnick and St{\u{a}}ric{\u{a}} (1995)
-  \cite{resnick1995consistency}).
-\end{itemize}
-In these cases it is shown that, if $k \to \infty$ and $k/n \to 0$
-as $n \to \infty$, the Hill estimator is weakly consistent, i.e.
-\[
-\sqrt k (\hat \alpha - \alpha) \overset{d}{\to} N(0, \alpha^2)
-\]
-
-Figure \ref{fig:Energy_lower},
-\ref{fig:Consumer_Staples_lower},
-and \ref{fig:Information_Technology_lower}
-show the Hill estimates of the
-lower tail indices of the ``energy'', ``consumer staples'' and ``information
-technology'' sectors of the S\&P 500 index, respectively.
+(1991) \cite{gul1991theory}. The Gul paper derives the DA from first principles (axiomatic).
+\par 
+In Section~\ref{sec:4} we summarize the discussion of the previous sections.
+
+\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
+In this section, we assume the model \eqref{eq:1} for the tails of the marginal
+\ds\ of a univariate return series $(X_t)$. For the sake of argument, we assume
+that this series constitutes a strictly stationary \seq . In what follows, we focus
+on the left tail of the \ds , i.e., on the losses. However, it is common to present the tail index estimators
+for positive data. Therefore we will multiply the losses by minus one, swapping the negative with the positive values.
+ The two parameters -- the tail-index  $\alpha_{\rm low}$ and the scale parameter
+$c_-$ -- are assumed positive. They play crucial roles for the understanding the risk hidden in the data, hence 
+for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
+useful description of the risk. Alternatively, these parameters can be
+used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
+%With these motives in mind, we present a survey of the
+%tail indices and scale parameters of 3 sectors, namely ``Energy'',
+%``Consumer Staples'' and ``Information Technology'', of the S\&P 500
+%index.
+
+\subsection{Hill estimates of lower tail indices}\label{sec:Hill}
+Various estimators of the tail index $\alpha_{\rm low}$  in the model
+\eqref{eq:1} have been proposed in the literature;
+see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, de
+Haan and Ferreira \cite{dehaan:ferreira:2006}, Resnick
+\cite{resnick:2007}. The most popular among them was introduced  by
+Hill \cite{hill1975simple}.
+Given a sample $X_1,\ldots,X_n$ from \eqref{eq:1}, one can calculate
+the order statistices $X_{(1)}\le \cdots\le X_{(n)}$ and construct the
+{\em Hill estimator}:
+\beao%\label{eq:Hill_index}
+  \hat \alpha_{\rm low} = \Big(
+    {1 \over k} \sum_{i=1}^k \ln {X_{(n-i+1)} \over X_{(n-k)}}
+    \Big)^{-1}\,.
+\eeao
+Here $k$  is the number of upper order statistics in the sample used
+for the estimation. The estimator $\hat \alpha_{\rm low}$ is 
+a maximum-likelihood estimator of $\alpha$ based on the $k$ upper
+order statistics in the pure Pareto model (recall that 
+we multiplied the data by minus one)
+\beao
+\P(X_t>x)= \dfrac{K^\alpha}{x^\alpha}\,,\quad x>K\,,
+\eeao
+under the hypothesis that we do not know the (high) threshold value
+$K$. The estimator has ``good'' properties such as 
+\asy\ consistency and \asy\ normality. These properties hold under
+strict stationarity assumptions on the data; Drees and Rootz\'en 
+\cite{drees:rootzen:2010} give perhaps most general conditions for
+dependent \seq s and de Haan and Ferreira provide
+a complete \asy\ theory in the iid case.
+\par
+A major problem for Hill estimation is the choice of the number $k$ of
+order statistics. As  matter of fact,
+if $k$ is too large the order statistics are too close to the center
+of the \ds\ of the $X_t$. This leads to bias
+of the estimator. On the other hand, by construction $\hat \alpha_{\rm
+  low}$ is an average of $k$ log-differences of the 
+data. Therefore, the variance of the estimator is the larger the
+smaller $k$. For these reasons, \asy\ theory requires 
+to choose $k=k_n$ \st\ $k_n\to\infty$ and $k_n/n\to 0$ as $\nto$. This
+fact does not make the estimation of $\alpha_{\rm low}$ an  
+easy matter: one has to choose a ``small'' value $k$ which is not
+``too large''. For practical purposes,
+a so-called {\em Hill plot} is recommended where $\hat \alpha$ is
+plotted for a variety of $k$-values, corresponding to
+some high quantile $X_{(n-k)}$ of the data. Then $k$ is chosen from a
+region in the plot where $\hat \alpha_{\rm low}$ is relatively
+stable. For example, in Figure~\ref{fig:1} we have chosen $k=50$ from
+a sample of size $n=1304$, corresponding to the 96\%-quantile of the
+data. In general, the estimation of the tail index is an art and
+requires some expertise; for some guidance
+see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997},
+Resnick \cite{resnick:2007} and Drees et al. 
+\cite{drees:resnick:how to make a Hill plot}. In Figure~\ref{fig:1},
+we exhibit 95\% \asy\ confidence bands which is based on the \clt
+\beam\label{eq:2}
+\sqrt k\, \big(\hat \alpha_{\rm low} - \alpha_{\rm low}\big) \std N(0, \alpha^2)\,,
+\eeam
+i.e., $\hat \alpha_{\rm low}$ is \asy ally unbiased and has variance
+$\alpha^2/k$. Since $k/n\to 0$ this means that the 
+confidence bands are significantly larger than the classical
+$\sqrt{n}$-rates. This fact is one explanation for the fact that it is
+difficult to say something meaningful about the true value of
+$\alpha_{\rm low}$. There exist various other reasons why one should
+not have 100\% trust in the confidence bands shown in
+Figure~\ref{fig:1}. Indeed, \eqref{eq:2}
+holds under rather subtle {\em second order conditions} on the tail
+$\P(X_t>x)$ which cannot be verified on data. However, 
+given a theoretical model such as the GARCH, these conditions can be
+verified based on the properties of the model. If they are not satisfied the 
+Hill estimator may exhibit significant bias; see Embrechts et
+al. \cite{embrechts:klueppelberg:mikosch:1997} for illustrations 
+of this fact leading to so-called ``Hill horror plots''. Moreover, the
+Hill estimator is rather sensitive to non-stationarity 
+of the data and to dependence. For example, results in Drees and
+Rootz\'en \cite{drees:rootzen:???} show that the \asy\ variance of the
+Hill estimator can be significantly larger than in the iid case. 
+Since return data are dependent, the \asy\ confidence bands should be
+even wider than exhibited in Figure~\ref{fig:1}. 
+Again, only under he assumption of a concrete model like GARCH these
+confidence bands could be evaluated.
 \begin{figure}[htb!]
-  \centering
-  \begin{subfigure}[b]{\linewidth}
+  \begin{minipage}{1.0\linewidth}
     \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
     {Energy_lower.pdf}
-    \caption{``Energy'' sector of S\&P 500}
-    \label{fig:Energy_lower}
-  \end{subfigure}
-  \begin{subfigure}[b]{\linewidth}
+  \end{minipage}
+  \begin{minipage}{1.0\linewidth}
     \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
     {Consumer_Staples_lower.pdf}
-    \caption{``Consumer Staples'' sector of S\&P 500}
-    \label{fig:Consumer_Staples_lower}
-  \end{subfigure}
-  \begin{subfigure}[b]{\linewidth}
+  \end{minipage}
+  \begin{minipage}{1.0\linewidth}
     \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
     {Information_Technology_lower.pdf}
-    \caption{``Information Technology'' sector of S\&P 500}
-    \label{fig:Information_Technology_lower}
-  \end{subfigure}
+  \end{minipage}
   \caption{\small Hill estimates of lower tail indices of
     daily return series in sectors of Standard \& Poor
-    500 index. Each circle coresponds to a Hill estimate; the grey
+    500 index. The figures from top to bottom are ``Energy'',
+    ``Consumer Staples'' and ``Information Technology'' sectors.
+    Each circle coresponds to a Hill estimate; the grey
     triangles above and below it mark the 97.5\% and 2.5\% quantiles
@@
     for computing $\alpha_n$. 
-  }
+  }\label{fig:1}
 \end{figure}
-One can see that all the confidence intervals of stocks in the
-``Energy'' and ``Consumer Staples'' sectors, as well as those of a
-large portion of stocks in the ``Information Technology'' sector,
-overlap. This suggests that the stocks in each sector may have very
-similar tail indices. Hoga's test on the change of extreme quantiles
+
+In Figure~\ref{fig:1} we see that confidence intervals of the Hill
+estimates of the losses in the ``Energy'' and ``Consumer Staples''
+sectors of the S\& P 500 index, as well as those of a 
+large portion of losses in the ``Information Technology'' sector,
+overlap. This fact indicates that the returns in each sector may 
+have comparable tail indices.
+
+Hoga's test on the change of extreme quantiles
 may provide some insight about how similar these tail indices are,
@@
 
-Figure \ref{fig:Hoga_Energy_Single} shows the test statistics of
-stocks in the ``Energy'' sector of the S\&P 500
-index. The null hypothesis here is that the tail index remains the
+Figure \ref{fig:Hoga_Single} shows the test statistics of
+stocks in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
+500 index. The null hypothesis here is that the tail index remains the
 same througout the selected period of time.
 \begin{figure}[htb!]
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{Hoga_Energy_Single.pdf}
-    \caption{``Energy''}
-    \label{fig:Hoga_Energy_Single}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{Hoga_Consumer_Staples_Single.pdf}
-    \caption{``Consumer Staples''}
-    \label{fig:Hoga_consumer_Staples_Single}
-  \end{subfigure}
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]
+    {Hoga_Energy_Single.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]
+    {Hoga_Consumer_Staples_Single.pdf}
+  \end{minipage}
   \caption{Hoga Test Statistics of stocks in the ``Energy'' and
-    ``Consumer Staples'' sectors of
-    S\&P 500. The green, blue and red lines are, respectively the
-    85\%, 90\% and 95\% quantiles of the distribution of
-    \eqref{eq:Q1_distr}}.
+    ``Consumer Staples'' sectors of S\&P 500. The green, blue and red
+    lines are, respectively the 85\%, 90\% and 95\% quantiles of the
+    distribution of \eqref{eq:Q1_distr}}.
+  \label{fig:Hoga_Single}
 \end{figure}
@@
 ``Consumer Staples'' sector, we summarise the results in figure
-\ref{fig:Hoga_Energy_pair} and \ref{fig:Hoga_consumer_Staples_pair}.
+\ref{fig:Hoga_pair}.
 \begin{figure}[htb!]
-  \begin{subfigure}[b]{0.5\linewidth}
+  \begin{minipage}{0.5\linewidth}
     \includegraphics[
@@
     ]{Hoga_Energy_pair.pdf}
-    \caption{``Energy''}
-    \label{fig:Hoga_Energy_pair}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
     \includegraphics[
@@
     ]{Hoga_Consumer_Staples_pair.pdf}
-    \caption{``Consumer Staples''}
-    \label{fig:Hoga_consumer_Staples_pair}
-  \end{subfigure}
+  \end{minipage}
   \caption{Hoga Test Statistics of stocks in the ``Energy'' and
@@
     requirements and time limits.}
+  \label{fig:Hoga_pair}
 \end{figure}
@@
   same tail index or not}
+\label{sec:2}
 Many multivariate GARCH models imply a common tail index for all
@@
   \vec h_t =
-  \sum_{l=1}^q \mtx A_{l} (\vec X_{t-l} \odot \vec X_{t-l})
+  \sum_{l=1}^q \mathbf A_{l} (\vec X_{t-l} \odot \vec X_{t-l})
   +
-  \sum_{l=1}^p \mtx B_{l} \vec h_{t-l}
+  \sum_{l=1}^p \mathbf B_{l} \vec h_{t-l}
   +
@@
 \end{equation}
-where $\mtx A_l$ and $\mtx B_l$ are non-diagonal $d \times d$ matrices
+where $\mathbf A_l$ and $\mathbf B_l$ are non-diagonal $d \times d$ matrices
 and $\odot$ denotes element-wise multiplication. In the simple case $p=q=1$,
@@
   \vec h_t &=& \left[
-    \mtx A \diag(Z_{1, t-1}^2, \dots, Z_{d, t-1}^2) + \mtx B
+    \mathbf A \text{diag}(Z_{1, t-1}^2, \dots, Z_{d, t-1}^2) + \mathbf B
     \right] \vec h_{t-1} + \vec \omega
@@
 \section{Generalized disappointment aversion}
-\label{sec:GDA}
+\label{sec:3}
 Routledge and Zin (2010) \cite{routledge2010generalized} define the GDA
@@
   \begin{eqnarray*}
-    \opd{\theta}\left[
+    {\partial \over \partial \theta}\left[
       F(b) - F(a)
@@
   \[
-  \pd{G(F)}{\alpha} > 0,
+  \frac{\pd G(F)}{\pd \alpha} > 0,
   \quad
-  \pd{G(F)}{K} < 0
+  \frac{\pd G(F)}{\pd K} < 0
   \]
@@
 ``Information Technology'' sectors of S\&P 500 companies. This is
-illustrated in figures 
-\ref{fig:Energy_alpha_K_ci}, \ref{fig:Consumer_Staples_alpha_K_ci}
-and \ref{fig:Information_Technology_alpha_K_ci}.
+illustrated in figure \ref{fig:sectors_parameters}.
 The confidence bands of the Hill estimates of $\alpha$ of these
-sectors are shown in figures \ref{fig:Energy_lower},
-\ref{fig:Consumer_Staples_lower} and
-\ref{fig:Information_Technology_lower}.
+sectors are shown in figure~\ref{fig:1}.
 
@@
 the value of $\hat\phi$ can be obtained for given $u(\cdot)$ and given
-values of $K, K', \alpha, \beta$. Figures \ref{fig:phi_hat_pareto5e-1}
-and \ref{fig:phi_hat_pareto4} show the values of $\hat\phi$ for
-$u(C) = -\frac{C^{-\xi}}{\xi}$ with $\xi = 1/2, 4$ respectively.
-$\beta=\alpha$ and $K'=K$ in both figures.
+values of $K, K', \alpha, \beta$. Figures \ref{fig:phi_hat_pareto}
+shows the values of $\hat\phi$ for $u(C) = -\frac{C^{-\xi}}{\xi}$ with
+$\xi = 1/2, 4$. $\beta=\alpha$ and $K'=K$ in both figures.
 \begin{figure}
-  \begin{subfigure}[b]{0.5\linewidth}
+  \begin{minipage}{0.5\linewidth}
     \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
-    \caption{$\hat\phi$ when $\xi = 1/2$}
-    \label{fig:phi_hat_pareto5e-1}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
     \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
-    \caption{$\hat\phi$ when $\xi = 4$}
-    \label{fig:phi_hat_pareto4}
-  \end{subfigure}
+  \end{minipage}
   \caption{$\hat\phi$, the optimal equity allocation when the utility
-    function is $u(C) = -{C^{-\xi} \over \xi}$ for $\xi = 1/2, 4$.
+    function is $u(C) = -{C^{-\xi} \over \xi}$ for $\xi = 1/2$
+    (left). and $\xi = 1/2$ (right).
   }
+  \label{fig:phi_hat_pareto}
 \end{figure}
@@
   consumption and that seeks wealth more aggressively. In other
-  words, he is less risk-avert than is one with a higher $\xi$.
+  words, he is less risk-averse than is one with a higher $\xi$.
   Such an agent will therefore invest more heavily on the equity as
-  is shown in figures \ref{fig:phi_hat_pareto5e-1} and
-  \ref{fig:phi_hat_pareto4}.
+  is shown in figures \ref{fig:phi_hat_pareto}.
 \end{minipage}
@@
 \]
-is increasing with $\alpha$ and decreasing with $K$. Figures
-\ref{fig:preference_pareto5e-1} and \ref{fig:preference_pareto4}
-show the values of the above integral when $\alpha$ and $K$ take
+is increasing with $\alpha$ and decreasing with
+$K$. Figure~\ref{fig:preference_pareto}
+shows the values of the above integral when $\alpha$ and $K$ take
 a range of values.
 \begin{figure}[htb!]
-  \begin{subfigure}[b]{0.5\linewidth}
+  \begin{minipage}{0.5\linewidth}
     \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
-    \caption{$I(\alpha, K)$}
-    \label{fig:preference_pareto5e-1}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
     \includegraphics[width=\textwidth]{preference_pareto4.pdf}
-    \caption{$I(\alpha, K)$}
-    \label{fig:preference_pareto4}
-  \end{subfigure}
+  \end{minipage}
   \caption{$I(\alpha, K)$ when the utility function is
@@
   }
+  \label{fig:preference_pareto}
 \end{figure}
@@
 \begin{eqnarray*}
-  I(K) &=& \E\left[ \left( \opd{K} l(K | X < 0) \right)^2 \right] \\
+  I(K) &=& \E\left[ \left( {\partial \over \partial K} l(K | X < 0) \right)^2 \right] \\
   &=& \alpha K^{\alpha - 2}
@@
 
-Figure \ref{fig:Energy_alpha_K_ci},
-\ref{fig:Consumer_Staples_alpha_K_ci},
-and \ref{fig:Information_Technology_alpha_K_ci}
+The plots in the 1st column of figure \ref{fig:sectors_parameters}
 show the estimates of $K$ alongside the Hill estimates of $\alpha$ of
 the ``energy'', ``consumer staples'' and ``information technology''
-sectors of the S\&P 500 index, respectively.
-Figure \ref{fig:Energy_p}, \ref{fig:Consumer_Staples_p} and
-\ref{fig:Information_Technology_p} show the estimates of $p$ of the
-same sectors.
+sectors of the S\&P 500 index;
+The plots in the 2nd column show the estimates of $p$ of the same
+sectors.
 
@@
 
-One can also see from figure \ref{fig:Energy_alpha_K_ci}
-and \ref{fig:Information_Technology_alpha_K_ci} that, while the
-``Energy'' and the ``Information Technology'' sectors are similar in
+One can also see from figure \ref{fig:sectors_parameters} that, while
+the ``Energy'' and the ``Information Technology'' sectors are similar in
 riskiness, the correlation between $\alpha$ and $K$ is stronger in
@@
   \centering
-  \begin{subfigure}[b]{0.45\linewidth}
+  % \begin{subfigure}[b]{0.45\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %                   {Energy_p.pdf}
+  %                   \caption{``Energy''}
+  %                   \label{fig:Energy_p}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.45\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %                   {Energy_alpha_K_ci.pdf}
+  %                   \caption{``Energy''}
+  %                   \label{fig:Energy_alpha_K_ci}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.45\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %                   {Consumer_Staples_p.pdf}
+  %                   \caption{``Consumer Staples''}
+  %                   \label{fig:Consumer_Staples_p}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.45\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %                   {Consumer_Staples_alpha_K_ci.pdf}
+  %                   \caption{``Consumer Staples''}
+  %                   \label{fig:Consumer_Staples_alpha_K_ci}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.45\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %                   {Information_Technology_p.pdf}
+  %                   \caption{``Information Technology''}
+  %                   \label{fig:Information_Technology_p}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.45\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %                   {Information_Technology_alpha_K_ci.pdf}
+  %                   \caption{``Information Technology''}
+  %                   \label{fig:Information_Technology_alpha_K_ci}
+  % \end{subfigure}
+  \begin{minipage}{0.45\linewidth}
     \includegraphics[width=\textwidth]
     {Energy_p.pdf}
-                    \caption{``Energy''}
-                    \label{fig:Energy_p}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\linewidth}
+  \end{minipage}\hfill
+  \begin{minipage}{0.45\linewidth}
     \includegraphics[width=\textwidth]
     {Energy_alpha_K_ci.pdf}
-                    \caption{``Energy''}
-                    \label{fig:Energy_alpha_K_ci}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\linewidth}
-    \includegraphics[width=\textwidth]
-                    {Consumer_Staples_p.pdf}
-                    \caption{``Consumer Staples''}
-                    \label{fig:Consumer_Staples_p}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\linewidth}
+  \end{minipage}
+  \begin{minipage}{0.45\linewidth}
     \includegraphics[width=\textwidth]
     {Consumer_Staples_alpha_K_ci.pdf}
-                    \caption{``Consumer Staples''}
-                    \label{fig:Consumer_Staples_alpha_K_ci}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\linewidth}
+  \end{minipage}\hfill
+  \begin{minipage}{0.45\linewidth}
     \includegraphics[width=\textwidth]
-                    {Information_Technology_p.pdf}
-                    \caption{``Information Technology''}
-                    \label{fig:Information_Technology_p}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\linewidth}
+    {Consumer_Staples_alpha_K_ci.pdf}
+  \end{minipage}
+  \begin{minipage}{0.45\linewidth}
     \includegraphics[width=\textwidth]
     {Information_Technology_alpha_K_ci.pdf}
-                    \caption{``Information Technology''}
-                    \label{fig:Information_Technology_alpha_K_ci}
-  \end{subfigure}
+  \end{minipage}\hfill
+  \begin{minipage}{0.45\linewidth}
+    \includegraphics[width=\textwidth]
+    {Information_Technology_alpha_K_ci.pdf}
+  \end{minipage}
   \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
@@
   }
+  \label{fig:sectors_parameters}
 \end{figure}
@@
   \centering
-  \begin{subfigure}[b]{0.32\linewidth}
+  \begin{minipage}{0.32\linewidth}
     \includegraphics[width=\textwidth]
     {Energy_scale.pdf}
-                    \caption{``Energy''}
-                    \label{fig:Energy_scale}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.32\linewidth}
+
+  \end{minipage}
+  \begin{minipage}{0.32\linewidth}
     \includegraphics[width=\textwidth]
     {Consumer_Staples_scale.pdf}
-                    \caption{``Consumer Staples''}
-                    \label{fig:Consumer_Staples_scale}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.32\linewidth}
+
+  \end{minipage}
+  \begin{minipage}{0.32\linewidth}
     \includegraphics[width=\textwidth]
     {Information_Technology_scale.pdf}
-                    \caption{``IT''}
-                    \label{fig:Information_Technology_scale}
-  \end{subfigure}
+  \end{minipage}
+  % \begin{subfigure}[b]{0.32\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %   {Energy_scale.pdf}
+  %   \caption{``Energy''}
+  %   \label{fig:Energy_scale}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.32\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %   {Consumer_Staples_scale.pdf}
+  %   \caption{``Consumer Staples''}
+  %   \label{fig:Consumer_Staples_scale}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.32\linewidth}
+  %   \includegraphics[width=\textwidth]
+  %   {Information_Technology_scale.pdf}
+  %   \caption{``IT''}
+  %   \label{fig:Information_Technology_scale}
+  % \end{subfigure}
   \caption{\small Scale estimates obtained from Hill's scale estimator
@@
 \begin{figure}[htb!]
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
-    \caption{$\hat\phi(\nu)$ of log-utility function}
-    \label{fig:phi_hat_t_log}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_log.pdf}
-    \caption{$G(\nu)$ of log-utility function}
-    \label{fig:U_b_t_log}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
-    \caption{$\hat\phi(\nu)$ of power-utility function}
-    \label{fig:phi_hat_t_power}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
-    \caption{$G(\nu)$ of power-utility function}
-    \label{fig:U_b_t_power}
-  \end{subfigure}
+  % \begin{subfigure}[b]{0.5\linewidth}
+  %   \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
+  %   \caption{$\hat\phi(\nu)$ of log-utility function}
+  %   \label{fig:phi_hat_t_log}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.5\linewidth}
+  %   \includegraphics[width=\textwidth]{U_b_t_log.pdf}
+  %   \caption{$G(\nu)$ of log-utility function}
+  %   \label{fig:U_b_t_log}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.5\linewidth}
+  %   \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
+  %   \caption{$\hat\phi(\nu)$ of power-utility function}
+  %   \label{fig:phi_hat_t_power}
+  % \end{subfigure}
+  % \begin{subfigure}[b]{0.5\linewidth}
+  %   \includegraphics[width=\textwidth]{U_b_t_power.pdf}
+  %   \caption{$G(\nu)$ of power-utility function}
+  %   \label{fig:U_b_t_power}
+  % \end{subfigure}
   \caption{$\hat\phi$ and $G(\nu)$
@@
 
-%   where
-%   \[
-%   y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
-%   \]
-%   Now notice
-%   \[
-%   \ln(y e^{-y}) = \ln(y) - y
-%   \]
-%   is a decreasing function. Thus $-y e^{-y}$ is an increasing
-%   function. Hence we have
-%   \begin{eqnarray*}
-%     W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
-%   \end{eqnarray*}
-%   Now it is clear
-%   \[
-%   \nu\exp\left\{
-%   W\left[
-%     -\left(1 + {1 \over \nu}\right)
-%     e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
-%   \right]
-%   + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
-%   &=& W(-y e^{-y}) + y
-% \end{eqnarray*}
 \section{Conclusion}
@@
       > h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx
-      = h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx
+      = h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx
       = 0
@@
     = I_1 + I_2
-    > h(x_0) \opd{\theta} \int_a^b f(x, \theta) dx = 0
+    > h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx = 0
     \]
@@
   \[
-  \Psi(x) = \td{\ln[\Gamma(x)]}{x}
+  \Psi(x) = {d \ln[\Gamma(x)] \over d x}
   \]
@@
     \label{eq:xxie2}
-    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
+    \hat \phi := \text{argmax}_{0 < \phi \leq 1} \mathcal U(F, \phi)
   \end{equation}
@@
     \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
-    = \opd{\alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}]
+    = {\partial \over \partial \alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}]
   \end{eqnarray*}
@@
     && \pd{f(x; \alpha, K)}{\alpha} \\
-    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+    &=& {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
     &=&
@@
   \[
-  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
   \]
@@
   \[
-  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
   \]
@@
   \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi)
-  = \opd{K} \E \left[
+  = {\partial \over \partial K} \E \left[
     u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}
@@
     && \pd{f(x; \alpha, K)}{K} \\
-    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
+    &=& {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
     &=&
@@
   \[
-  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
   \]
@@
   \[
-  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
   \]
@@
 \bibliographystyle{unsrt}
-%% \bibliography{../../thesis/econophysics}
-\bibliography{econophysics}
+\bibliography{../../thesis/econophysics}
 \end{document}

### 2017-04-05 papers/FX/draft.tex
Some more revision of the draft.

@@
 %\usepackage{enumerate}
-%\usepackage[bookmarks=true]{hyperref}
+\usepackage[bookmarks=true]{hyperref}
 %\usepackage{bookmark}
 \usepackage{graphicx}
-\usepackage[bookmarks=true]{ hyperref}
 %\usepackage{multirow}
@@
 \par 
-In Section~\ref{sec:4} we summarize the discussion of the previous sections.
+In Section~\ref{sec:4} we summarize the discussion of the previous
+sections.
 
@@
 ``Information Technology'' sectors of the S\&P 500 index are computed
-using this method. The results are shown in figure \ref{fig:Energy_scale},
-\ref{fig:Consumer_Staples_scale} and
-\ref{fig:Information_Technology_scale}, along with estimates obtained
+using this method. The results are shown in figure
+\ref{fig:sectors_parameters}, along with estimates obtained
 using the hybrid method described in section
@@
     \right] = 0 \\
-    \left| \pd{f(x, \theta)}{\theta} \right| < \infty
+    \left| \frac{\pd f(x, \theta)}{\pd \theta} \right| < \infty
     \quad \forall x \in (a, b)
@@
   \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
-    \[
-    \pd{\E h(X)}{\theta} > 0
+    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
+    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
+    \[
+    \frac{\pd \E h(X)}{\pd \theta} > 0
     \]
   \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
-    \[
-    \pd{\E h(X)}{\theta} > 0
+    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
+    \[
+    \frac{\pd \E h(X)}{\pd \theta} > 0
     \]
@@
   \begin{enumerate}
-  \item If $h(\cdot)$ is increasing and $\pd{f}{\theta}$ satisfies the
+  \item If $h(\cdot)$ is increasing and $\frac{\pd f}{\pd \theta}$
+    satisfies the
     same conditions of the 1st case of lemma \ref{lemma:I},
-    $\pd{\E h(X)}{\theta} < 0$. This immediately follows from applying
-    1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
+    $\frac{\pd \E h(X)}{\pd \theta} < 0$. This immediately follows
+    from applying 1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
   \item By the same argument, if $h(\cdot)$ is decreasing and
-    $\pd{f}{\theta}$ satisfies the same conditions of the 2nd case of
-    lemma \ref{lemma:I}, $\pd{\E h(X)}{\theta} < 0$.
+    ${\pd f \over \pd \theta}$ satisfies the same conditions of the 2nd case of
+    lemma \ref{lemma:I}, ${\pd \E h(X) \over \pd \theta} < 0$.
   \end{enumerate}
@@
 To obtain the maximum likelihood estimation, we need to solve the
-equation $\pd{L}{K} = 0$, which expands to
+equation ${\pd L \over \pd K} = 0$, which expands to
 \[
@@
   \centering
-  % \begin{subfigure}[b]{0.45\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %                   {Energy_p.pdf}
-  %                   \caption{``Energy''}
-  %                   \label{fig:Energy_p}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.45\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %                   {Energy_alpha_K_ci.pdf}
-  %                   \caption{``Energy''}
-  %                   \label{fig:Energy_alpha_K_ci}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.45\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %                   {Consumer_Staples_p.pdf}
-  %                   \caption{``Consumer Staples''}
-  %                   \label{fig:Consumer_Staples_p}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.45\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %                   {Consumer_Staples_alpha_K_ci.pdf}
-  %                   \caption{``Consumer Staples''}
-  %                   \label{fig:Consumer_Staples_alpha_K_ci}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.45\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %                   {Information_Technology_p.pdf}
-  %                   \caption{``Information Technology''}
-  %                   \label{fig:Information_Technology_p}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.45\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %                   {Information_Technology_alpha_K_ci.pdf}
-  %                   \caption{``Information Technology''}
-  %                   \label{fig:Information_Technology_alpha_K_ci}
-  % \end{subfigure}
   \begin{minipage}{0.45\linewidth}
@@
 can immediately identity $p K^\alpha$ as the scale parameter. Figure
-\ref{fig:Energy_scale}, \ref{fig:Consumer_Staples_scale} and
-\ref{fig:Information_Technology_scale} compare the estimates obtained
+\ref{fig:scale_estimates} compare the estimates obtained
 from Hill's scale estimator with those from the $p K^\alpha$ method.
@@
     {Energy_scale.pdf}
-
   \end{minipage}
@@
     {Consumer_Staples_scale.pdf}
-
   \end{minipage}
@@
   \end{minipage}
-  % \begin{subfigure}[b]{0.32\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %   {Energy_scale.pdf}
-  %   \caption{``Energy''}
-  %   \label{fig:Energy_scale}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.32\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %   {Consumer_Staples_scale.pdf}
-  %   \caption{``Consumer Staples''}
-  %   \label{fig:Consumer_Staples_scale}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.32\linewidth}
-  %   \includegraphics[width=\textwidth]
-  %   {Information_Technology_scale.pdf}
-  %   \caption{``IT''}
-  %   \label{fig:Information_Technology_scale}
-  % \end{subfigure}
   \caption{\small Scale estimates obtained from Hill's scale estimator
@@
   }
+  \label{fig:scale_estimates}
 \end{figure}
-
-%% \begin{figure}[htb!]
-%%   \centering
-%%   \includegraphics[width=\textwidth]{Energy_alpha_K.pdf}
-%%   \caption{Hybrid estimates of the lower-tail parameters of daily
-%%     return series in the {\it Energy} sector of S\&P
-%%     500 index.
-%%   }
-%%   \label{fig:Energy_alpha_K}
-%% \end{figure}
-
-%% \begin{figure}[htb!]
-%%   \centering
-%%   \includegraphics[width=\textwidth]{Consumer_Staples_alpha_K.pdf}
-%%   \caption{Hybrid estimates of the lower-tail parameters of daily
-%%     return series in the {\it Consumer Staples} sector of S\&P
-%%     500 index.
-%%   }
-%%   \label{fig:Consumer_Staples_alpha_K}
-%% \end{figure}
-
-%% \begin{figure}[htb!]
-%%   \centering
-%%   \includegraphics[width=\textwidth]{Information_Technology_alpha_K.pdf}
-%%   \caption{Hybrid estimates of the lower-tail parameters of daily
-%%     return series in the {\it Information Technology} sector of S\&P
-%%     500 index.
-%%   }
-%%   \label{fig:Information_Technology_alpha_K}
-%% \end{figure}
-
-% \subsection{Pickand's Estimator}
-% Pickands (1975) \cite{pickands1975statistical} proposed a different
-% approach to estimate the tail index than that of Hill. Assume
-% $\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
-% denote the quantile of $1 - 1/t$ of this distribution, i.e.
-% \[
-% P(X > U(t)) = {1 \over t}
-% \]
-% The key idea of Pickands' estimator is to observe
-% \[
-% \lim_{t \to \infty} {
-%   U(c(t) t) - U(t)
-%   \over
-%   U(t) - U(t/c(t))
-% } = 2^{1/\alpha}
-% \]
-% for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
-% Pickands formulated the estimation problem in terms of the extreme
-% value index $\xi$, which is the reciprocal of the tail index $\alpha$
-% when the distribution of $X$ is regularly varying (cf. Embrechtsm,
-% Kl\"uppelberg and Mikosch \cite{Embrechts1997}).
-
-% Building upon the aforementioned observation and taking $c(t)$ to
-% be simply constant 2, Pickands proposed the following estimator for
-% $\xi$:
-% \[
-% \hat \xi = {1 \over \ln 2} \ln {
-%   x_{(k)} - x_{(2k)}
-%   \over
-%   x_{(2k)} - x_{(4k)}
-% }
-% \]
-% Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
-% that Pickands' estimator was consistent in both the weak and the strong
-% sense, when the sequence $k(n)$ was chosen appropriately:
-% \begin{itemize}
-% \item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
-%   \[
-%   \hat \xi \overset{P}{\to} \xi
-%   \]
-% \item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
-%   as $n \to \infty$,
-%   \[
-%   \hat \xi \overset{a.s.}{\to} \xi
-%   \]
-% \end{itemize}
-% Dekkers and de Haan \cite{dekkers1989estimation} also showed the
-% asymptotic normality of the estimator, when some
-% extra technical conditions were satisfied by
-% $U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
-% \[
-% \sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
-% \]
-% where
-% \[
-% v(\xi) = {
-%   \xi^2 (2^{2 \xi + 1} + 1)
-%   \over
-%   [2 \ln(2) (2^\xi - 1)]^2
-% }
-% \]
-% Using Pickands' estimator, the same result of very similar tail indices
-% is again obtained for the {\it Energy} and the {\it Consumer Staples}
-% sectors. Figure \ref{fig:Energy_Pickands} and
-% \ref{fig:Consumer_Staples_Pickands} illustrate this claim.
-% \begin{figure}[htb!]
-%   \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
-%   \caption{Pickands' estimates of {\it Energy} stocks.}
-%   \label{fig:Energy_Pickands}
-% \end{figure}
-
-% \begin{figure}[htb!]
-%   \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
-%   \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
-%   \label{fig:Consumer_Staples_Pickands}
-% \end{figure}
-
-%% \section{Hoga Test}
-%% TODO: Show the test statistics of Hoga test applied to a sequence of
-%% two concatenated return series. A changes in the scale is characterized
-%% by a smooth bump in the test statistics while a change in the tail
-%% index is characterized by a jump. Derive a theoretical result to
-%% support this?
-
-%% What is the probability of rejecting the null hypothesis when the test
-%% is applied to $(X_1, X_2)$ where $X_1$ and $X_2$ have tail indices
-%% $\kappa_1 \neq \kappa_2$? The probability of wrongly rejecting the
-%% null hypothesis must be lower than $\alpha$.
-%% \textcolor{red}{Review the test of Hoga}
-
-% \section{Data Analysis} \label{sec:data_analysis}
-% \subsection{Tail index and scale parameter}
-% In this section we apply the statistical tools reviewed in the
-% previous sections to real stock return series, denoted $X_t$
-% hereafter. We are interested in finding out the tail indices and the
-% scale parameters of the stationary distribution of $|X_t|$.
-
-%% \subsubsection{The Energy and the Consumer Staples sectors of S\&P 500}
-%% Figure \ref{fig:Energy_Hill_upper} shows the Hill estimates of the
-%% tail indices of 32 stocks in the {\it Energy} sector of S\&P 500
-%% index; figure \ref{fig:Consumer_Staples_Hill_upper} shows the same of
-%% the {\it Consumer Staples} sector.
-%% \begin{figure}[htb!]
-%%   \centering
-%%   \includegraphics[width=\textwidth]{Energy_Hill_upper.pdf}
-%%   \caption{Hill estimates of the upper tail indices of 32 stocks'
-%%     daily return series in the {\it Energy} sector of Standard \& Poor
-%%     500 index. Each circle coresponds to a Hill estimate; the grey
-%%     triangles above and below it mark the 97.5\% and 2.5\% quantiles
-%%     of its approximate normal distribution.
-%%     The low and high red lines mark the median of the 2.5\% quantiles
-%%     and median of the 97.5\% quantiles, respectively.
-%%     The data are from {\it Yahoo Finance} and the labels on
-%%     the horizontal axis are Yahoo symbols of the stocks. The data span
-%%     from 2010-01-01 to 2015-01-01 and comprise $n=1304$
-%%     observations. $k=(\ln n)^2 = 49$ upper order statistics are used
-%%     for computing $\alpha_n$. 
-%%   }
-%%   \label{fig:Energy_Hill_upper}
-%% \end{figure}
-
-%% \begin{figure}[htb!]
-%%   \centering
-%%   \includegraphics[width=\textwidth]{Consumer_Staples_Hill_upper.pdf}
-%%   \caption{Hill estimates of the upper tail indices of 30 stocks'
-%%     daily return series in the {\it Consumer Staples} sector of
-%%     Standard \& Poor 500 index.
-%%   }
-%%   \label{fig:Consumer_Staples_Hill_upper}
-%% \end{figure}
-
-%% For our purpose, we have excluded a few stocks, though:
-%% \begin{figure}[htb!]
-%%   \centering
-%%   \begin{subfigure}{0.45\textwidth}
-%%     \includegraphics[width=\textwidth]{WBA_price.pdf}
-%%     \caption{{\it Walgreens Boots Alliance} prices}
-%%     \label{fig:EBA_prices}
-%%   \end{subfigure}
-%%   \begin{subfigure}{0.45\textwidth}
-%%     \includegraphics[width=\textwidth]{WFM_price.pdf}
-%%     \caption{{\it Whole Foods Market} prices}
-%%     \label{fig:WFM_prices}
-%%   \end{subfigure}
-%%   \begin{subfigure}{0.45\textwidth}
-%%     \includegraphics[width=\textwidth]{EL_price.pdf}
-%%     \caption{{\it Estee Lauder Cos.} prices}    
-%%     \label{fig:EL_prices}
-%%   \end{subfigure}
-%%   \begin{subfigure}{0.45\textwidth}
-%%     \includegraphics[width=\textwidth]{CL_price.pdf}
-%%     \caption{{\it Colgate-Palmolive} prices}    
-%%     \label{fig:CL_prices}
-%%   \end{subfigure}
-%%   \caption{Plot of prices of {\it Consumer Staples} stocks
-%%     excluded from Hill estimates}
-%% \end{figure}
-
-%% \begin{itemize}
-%% \item {\it Walgreens Boots Alliance}:
-
-%%   As shown in figure (\ref{fig:EBA_prices}),
-%%   this stock was traded only in a small fraction of the bank
-%%   days and it is hence no surprise that the price varied
-%%   tremendously.
-%% \item {\it Whole Foods Market}, {\it Estee Lauder Cos.} and {\it
-%%   Colgate-Palmolive}
-
-%%   As seen in figures (\ref{fig:WFM_prices}), (\ref{fig:EL_prices}) and
-%%   (\ref{fig:CL_prices}),
-%%   the prices of these stocks were reduced by half during the
-%%   period under consideration. But the price changes are the
-%%   result of emission of new shares rather than random fluctuations.
-%% \end{itemize}
-
-%% Figure \ref{fig:Energy_Hill_upper} and
-%% \ref{fig:Consumer_Staples_Hill_upper}
-%% are suggestive of the conclusion that excluding price
-%% movements due to deterministic events, tail indices of stock return
-%% series tend to have the same value, within a reasonable confidence
-%% interval.
-
-%% If we look instead at the scales, differences among the stock return
-%% series are however rather large. Figure (\ref{fig:Energy_Hill_Scales})
-%% and (\ref{fig:Consumer_Staples_Hill_Scales}) show the Hill estimates of
-%% the scale parameters of the stocks in the S\&P 500 {\it Energy} and
-%% {\it Consumer Staples} sectors, respectively.
-%% \begin{figure}[htb!]
-%%   \begin{subfigure}{0.45\textwidth}
-%%     \includegraphics[width=\textwidth]{Energy_Hill_scales.pdf}
-%%     \caption{{\it Energy} scale estimates}    
-%%     \label{fig:Energy_Hill_Scales}
-%%   \end{subfigure}
-%%   \begin{subfigure}{0.45\textwidth}
-%%     \includegraphics[width=\textwidth]{Consumer_Staples_Hill_scales.pdf}
-%%     \caption{{\it Consumer Staples} scale estimates}    
-%%     \label{fig:Consumer_Staples_Hill_Scales}
-%%   \end{subfigure}
-%%   \caption{Black boxes are Hill estimates of the scale parameters of
-%%     the stock return series, while red circles are the scale estimates
-%%     of simulated iid. Student's $t$ series with a degree of freedom
-%%     equal to the mean value of the estimated tail indices of the stock
-%%     return series. The estimates of the Student's $t$ series are
-%%     included for comparison with those of the return series.}
-%% \end{figure}
-%% Clearly the estimated scale parameters can differ by several orders of
-%% magnitude.
-
-%% Using the ``Peaks over Threshold'' method and the mean excess
-%% function, we also obtain estimates for the tail index ($\alpha$), the
-%% scale parameter ($A$) and the shift parameter ($a$) for stocks in the
-%% {\it Energy} and the {\it Consumer Staples} sectors of the S\&P 500
-%% index. The results are plotted in figure
-%% \ref{fig:Energy_OLS_estimates} and
-%% \ref{fig:Consumer_Staples_OLS_estimates},
-%% respectively.
-%% \begin{figure}[htb!]
-%%   \includegraphics[width=\textwidth]{Energy_OLS_estimates.pdf}
-%%   \caption{Tail parameter estimates of {\it Energy} stocks. The top
-%%     plot shows the tail index estimates, the middle plot shows the
-%%     10-based logarithm of the scale estimates and bottom plot shows
-%%     the estimates of the shift parameter.
-%%   }
-%%   \label{fig:Energy_OLS_estimates}
-%% \end{figure}
-
-%% \begin{figure}[htb!]
-%%   \includegraphics[width=\textwidth]{Consumer_Staples_OLS_estimates.pdf}
-%%   \caption{Tail parameter estimates of {\it Consumer Staples}
-%%     stocks. The top 
-%%     plot shows the tail index estimates, the middle plot shows the
-%%     10-based logarithm of the scale estimates and bottom plot shows
-%%     the estimates of the shift parameter.
-%%   }
-%%   \label{fig:Consumer_Staples_OLS_estimates}
-%% \end{figure}
-
-%% \subsubsection{The Information Technology sector of S\&P 500}
-%% Compared with the sectors of energy and consumer staples, which enjoy
-%% a relatively stable demand on their products, companies of information
-%% technology are more subject to volatile demand and competition. As a
-%% result, one can expect their stock prices to be more volatile, implying
-%% lower tail indices and larger differences between the tail indices of
-%% different companies. In this section we investigate how these companies
-%% differ in terms of their tail indices and scale parameters.
-
-%% Figure \ref{fig:Information_Technology_Hill} shows the Hill
-%% estimates of 42 stocks in the {\it Information Technology} sector. Daily
-%% closing prices of these stocks have been available since 2000-01-01 and
-%% earlier.
-%% \begin{figure}[htb!]
-%%   \centering
-%%   \includegraphics[width=\textwidth]{Information_Technology_Hill.pdf}
-%%   \caption{Hill estimates of the tail indices of 42 stocks' daily
-%%     return series in the {\it Information Technology} sector of Standard
-%%     \& Poor 500 index.
-%%   }
-%%   \label{fig:Information_Technology_Hill}
-%% \end{figure}
-%% It is seen in the figure that most of the estimates fall within the
-%% interval spanned by the confidence bands. Thus, given the estimation
-%% precision, these are consistent with the similarity conclusion of 
-%% \S\ref{sec:GDA}.
-
-%% From the figure it appears a few pairs of stocks do have rather different
-%% tail indices, e.g. ``ADI'', ``PAYX'' and ``XLNX'' versus ``CTSH'' and
-%% ``NVDA''.
-
-%% To gain more insight into these pairs, we concatenate the
-%% squared return series of each pair and apply Hoga's test to the concatenated
-%% sequence. Table \ref{tab:Information_Technology_Hoga} shows the test
-%% statistic of each concatenated pair.
-%% \begin{table}[htb!]
-%%   \centering
-%%   \begin{tabular}{c|c|c}
-%%     \multicolumn{2}{c}{Stock(s)} & Hoga Test Statistic \\
-%%     \hline
-%%     ADI & CTSH & 15.52 \\
-%%     ADI & NVDA & 4.726 \\
-%%     PAYX & CTSH & 231.5 \\
-%%     PAYX & NVDA & 110.6 \\
-%%     XLNX & CTSH & 503.5 \\
-%%     XLNX & NVDA & 146.5
-%%   \end{tabular}
-%%   \caption{Hoga Test statistic of pairs of stocks from the
-%%     {\it Information Technology} sector of S\&P 500 index.
-%%     The critical value at 95\% confidence level is 166.4}
-%%   \label{tab:Information_Technology_Hoga}
-%% \end{table}
-
-%% \subsection{Temporal changes of the parameters}
-
-%% Thomas: Try fitting data to a Burr distribution.
-%% TODO: use GPD and POT to estimate the scale and tail index.
-%% TODO: plot the Hill estimator instead of its inverse. The asymptotic
-%% normal distribution of Hill estimator is found in ``Modeling Extremal
-%% events''.
-
-%% TODO: Does the scale estimator also have a convergence result from
-%% which confidence bands can be derived?
-
-%% TODO: Why does the Quintos Fan test almost always reject the null?
-
-%% Hill (1975) \cite{hill1975simple} also proposed an estimator for the
-%% scale parameter $A$:
-%% \[
-%% A_n = {k \over n}X_{(k+1)}^{\alpha_n}
-%% \]
-
-%% Review of CCC by Starica and others.
-%% explain that the model implies that all series have the same tail index
-%% Explain that economic arguments also lead to the same tail index
-
-%% Explain that the tail functions of different series are
-%% differentiated more by the scale parameter.
 
@@
   \label{lemma:II}
-  There exists $x_0 > 0$ such that $\pd{f}{\nu}(x_0, \nu) = 0$ and
-  $\forall x \in (0, x_0), \pd{f}{\nu}(x, \nu) > 0$ and
-  $\forall x \in (x_0, \infty), \pd{f}{\nu}(x, \nu) < 0$.
+  There exists $x_0 > 0$ such that ${\pd f \over \pd \nu}(x_0, \nu) = 0$ and
+  $\forall x \in (0, x_0), {\pd f \over \pd \nu}(x, \nu) > 0$ and
+  $\forall x \in (x_0, \infty), {\pd f \over \pd \nu}(x, \nu) < 0$.
 \end{lemma}
@@
 \begin{eqnarray*}
-  \pd{C(x, \phi) C(-x, \phi)}{x}
+  {\pd C(x, \phi) C(-x, \phi) \over \pd x}
   &=&
@@
 \ref{lemma:II},
-$\pd{\mathcal U(F, \phi))}{\nu} < 0$ for all $\phi$. Hence all
+${\pd \mathcal U(F, \phi)) \over \pd \nu} < 0$ for all $\phi$. Hence all
 investors guided by this utility function would be seeking the
@@
 respect to $\phi$ for a sequence of selected values of $\nu$. The
-results are shown in figures \ref{fig:phi_hat_t_log} and
-\ref{fig:U_b_t_log}.
+results are shown in figure~\ref{fig:phi_hat_U}.
 \begin{figure}[htb!]
@@
   % \end{subfigure}
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_log.pdf}
+  \end{minipage}
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
+  \end{minipage}
   \caption{$\hat\phi$ and $G(\nu)$
+    \label{fig:phi_hat_U}
   }
@@
 \]
-One can see from figure \ref{fig:phi_hat_t_power} and
-\ref{fig:U_b_t_power} that both $\hat \phi$ and $G(\nu)$ are monotonically 
-decreasing with $\nu$ as in the case of a log-utility function.
-%% \begin{figure}[htb!]
-%%   \begin{subfigure}[b]{0.5\linewidth}
-%%     \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
-%%     \caption{$\hat\phi(\nu)$}
-%%     \label{fig:phi_hat_t_power}
-%%   \end{subfigure}
-%%   \begin{subfigure}[b]{0.5\linewidth}
-%%     \includegraphics[width=\textwidth]{U_b_t_power.pdf}
-%%     \caption{$G(\nu)$}
-%%     \label{fig:U_b_t_power}
-%%   \end{subfigure}
-%%   \caption{$\hat\phi$ and $G(\nu)$ when the utility
-%%     function is $u(C) = -{2 \over \sqrt{C}}$
-%%   }
-%% \end{figure}
+One can see from figure \ref{fig:phi_hat_U_power} that both $\hat
+\phi$ and $G(\nu)$ are monotonically decreasing with $\nu$ as in the
+case of a log-utility function.
+\begin{figure}[htb!]
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
+  \end{minipage}
+  \caption{$\hat\phi$ and $G(\nu)$ when the utility
+    function is $u(C) = -{2 \over \sqrt{C}}$
+  }
+  \label{fig:phi_hat_U_power}
+\end{figure}
 
 \section{Conclusion}
+\label{sec:4}
 TODO: Summarize and conclude that all correlated series share the same
@@
   \begin{eqnarray*}
-    && \pd{\E h(X)}{\theta} \\
-    &=& \int_a^b h(x) \pd{f}{\theta}(x, \theta) dx \\
-    &=& \underbrace{\int_a^{x_0} h(x) \pd{f}{\theta}(x, \theta) dx}_{I_1}
-    + \underbrace{\int_{x_0}^b h(x) \pd{f}{\theta}(x, \theta) dx}_{I_2}
+    && {\pd \E h(X) \over \pd \theta} \\
+    &=& \int_a^b h(x) {\pd f \over \pd \theta}(x, \theta) dx \\
+    &=& \underbrace{\int_a^{x_0} h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_1}
+    + \underbrace{\int_{x_0}^b h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_2}
   \end{eqnarray*}
   \begin{enumerate}
-  \item When $h(x)$ is decreasing on $(a, b)$ and $\pd{f}{\theta}(x,
+  \item When $h(x)$ is decreasing on $(a, b)$ and ${\pd f \over \pd \theta}(x,
     \theta) > 0$ on $(a, x_0)$
     \[
-    I_1 > h(x_0) \int_a^{x_0} \pd{f}{\theta}(x, \theta) dx
-    \]
-    Similarly, because $\pd{f}{\theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
+    I_1 > h(x_0) \int_a^{x_0} {\pd f \over \pd \theta}(x, \theta) dx
+    \]
+    Similarly, because ${\pd f \over \pd \theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
     $h(x)$ is decreasing, we have
@@
       I_2 &=& \int_{x_0}^b -h(x_0)
-      \left|\pd{f}{\theta}(x, \theta) \right| dx \\
+      \left|{\pd f \over \pd \theta}(x, \theta) \right| dx \\
       &>& -h(x_0)
       \int_{x_0}^b \left| 
-        \pd{f}{\theta}(x, \theta)
+        {\pd f \over \pd \theta}(x, \theta)
       \right| dx
@@
     \begin{eqnarray*}
-      \pd{\E h(X)}{\theta}
-      > h(x_0) \int_a^b \pd{f}{\theta}(x, \theta) dx
+      {\pd \E h(X) \over \pd \theta}
+      > h(x_0) \int_a^b {\pd f \over \pd \theta}(x, \theta) dx
       = h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx
@@
   \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\pd{f}{\theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\pd{f}{\theta}(x_0; \theta) > 0$ on $(x_0, b)$
+    ${\pd f \over \pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    ${\pd f \over \pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$
     \begin{eqnarray*}
@@
       \int_a^{x_0} -h(x)
-      \left| \pd{f}{\theta}(x, \theta) \right| dx \\
+      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx \\
       &>&
       -h(x_0) \int_a^{x_0}
-      \left| \pd{f}{\theta}(x, \theta) \right| dx
+      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx
     \end{eqnarray*}
@@
     I_2 > h(x_0) \int_{x_0}^b
-    \pd{f}{\theta}(x, \theta) dx
+    {\pd f \over \pd \theta}(x, \theta) dx
     \]
@@
     \[
-    \pd{\E h(X)}{\theta}
+    {\pd \E h(X) \over \pd \theta}
     = I_1 + I_2
@@
   \begin{eqnarray}
-    \pd{f(x, \nu)}{\nu} &=& {
+    {\pd f(x, \nu) \over \pd \nu} &=& {
       c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
@@
 
-  While the denominator of the right side of $\pd{f(x, \nu)}{\nu}$ is
+  While the denominator of the right side of ${\pd f(x, \nu) \over \pd \nu}$ is
   always positive, its nummerator $P(x^2, \nu)$ has a single root:
@@
   \]
-  Now that we have established that $\pd{f}{\nu}(x, \nu) = 0$ has a
+  Now that we have established that ${\pd f \over \pd \nu}(x, \nu) = 0$ has a
   single positive root, it remains to determine the sign of
-  $\pd{f}{\nu}(x, \nu)$ on the two sides of the root. For this purpose
+  ${\pd f \over \pd \nu}(x, \nu)$ on the two sides of the root. For this purpose
   we observe
@@
   \end{equation}
-  So we want to investigate $\pd{P}{x}(x, \nu)$:
+  So we want to investigate ${\pd P \over \pd x}(x, \nu)$:
   \begin{eqnarray}
-    \pd{P}{x}(x, \nu) &=&
+    {\pd P \over \pd x}(x, \nu) &=&
     2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \ln\left(
@@
   \end{eqnarray}
-  Clearly, $\pd{P}{x}(0, \nu) > 0$. Hence from \eqref{eq:xxie5.1} and
+  Clearly, $\frac{\pd P}{\pd x}(0, \nu) > 0$. Hence from
+  \eqref{eq:xxie5.1} and
   \eqref{eq:xxie6} it is clear

### 2017-04-05 papers/FX/draft.tex
Revised my bib file. Corrected the citations. Updated the proof.

@@
 in the literature how heavy tails really are. For example, Barndorff-Nielsen and Shephard
-\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails, i.e.,
-tails which are comparable with those of a gamma \ds . On the other
-hand, tails of returns have been studied in great detail 
-in the extreme value community. Among extreme value specialists
-there is general agreement that returns $X_t$ have tails of power-law-type, i.e.,
+\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor
+``semi-heavy'' tails, i.e., tails which are comparable with those of a
+gamma \ds . On the other hand, tails of returns have been studied in
+great detail in the extreme value community. Among extreme value
+specialists there is general agreement that returns $X_t$ have tails
+of power-law-type, i.e.,
 \beam\label{eq:1}
-\P(X_t>x)\sim c_+ \,x^{\alpha_{\rm up}}\quad\mbox{and}\quad \P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
+\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad
+\P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
 \eeam  
@@
 that $f(x)/g(x)\to 1$ as $\xto$.} See for example, Embrechts et
-al. \cite{embrechts:klueppelberg:mikosch:1997}, Jansen and de Vries
+al. \cite{Embrechts1997}, Jansen and de Vries
 \cite{jansen1991frequency}, Mikosch \cite{mikosch:2003}, Resnick
@@
 {\em slowly varying} \fct s; cf. Embrechts et
-al. \cite{embrechts:klueppelberg:mikosch:1997}, Chapter 3. In this
+al. \cite{Embrechts1997}, Chapter 3. In this
 paper, for the  
@@
 is an extension of the concept of Disappointment Aversion (DA) of Gul
-(1991) \cite{gul1991theory}. The Gul paper derives the DA from first principles (axiomatic).
+ \cite{gul1991theory}. The Gul paper derives the DA from first principles (axiomatic).
 \par 
@@
 \eqref{eq:1} have been proposed in the literature;
-see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, de
+see Embrechts et al. \cite{Embrechts1997}, de
 Haan and Ferreira \cite{dehaan:ferreira:2006}, Resnick
@@
 requires some expertise; for some guidance
-see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997},
+see Embrechts et al. \cite{Embrechts1997},
 Resnick \cite{resnick:2007} and Drees et al. 
-\cite{drees:resnick:how to make a Hill plot}. In Figure~\ref{fig:1},
+\cite{drees:resnick:2000}. In Figure~\ref{fig:1},
 we exhibit 95\% \asy\ confidence bands which is based on the \clt
@@
 Hill estimator may exhibit significant bias; see Embrechts et
-al. \cite{embrechts:klueppelberg:mikosch:1997} for illustrations 
+al. \cite{Embrechts1997} for illustrations 
 of this fact leading to so-called ``Hill horror plots''. Moreover, the
@@
 of the data and to dependence. For example, results in Drees and
-Rootz\'en \cite{drees:rootzen:???} show that the \asy\ variance of the
+Rootz\'en \cite{drees:rootzen:2010} show that the \asy\ variance of the
 Hill estimator can be significantly larger than in the iid case. 
@@
 
-For convenience, we denote the scale parameter by $A$. It is the
-constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
+For convenience, we denote the scale parameter by $A$.
+%% It is the constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
 Hill \cite{hill1975simple} put forward an estimator of $A$ using his
@@
 \begin{hypothesis}
-  $\mathcal H_0$: For a given $p$, $\exists t^* \in [t_0, 1 - t_0]$
-  such that
+  $\mathcal H_0$: For a given $p$ and all $1 \leq i < j \leq n$,
+  $F_i^{-1}(p) = F_j^{-1}(p)$
+\end{hypothesis}
+while the alternative is $\exists t^* \in [t_0, 1 - t_0]$ such that
 \[
@@
 \]
-\end{hypothesis}
-where $t_0 \in (0, 1/2)$. The alternative hypothesis is that no such
-$t^*$ exists. Hoga proposed the following test statistics:
+where $t_0 \in (0, 1/2)$ and $\hat x_p(s, t)$ is the Weissman
+estimator applied to the partial series
+$X_{\ceil{ns}}, \dots, X_{\floor{nt}}$ using $\ceil{k(t - s)}$
+upper-order statistics. See Embrechts, Kl\"uppelberg and T.Mikosch
+\cite{Embrechts1997}, p.347.
+
+Hoga proposed the following test statistics:
 \[
@@
 \]
-Under the null hypothesis, Hoga showed that as $n \to \infty$,
+Under the null hypothesis and suitable dependence conditions, Hoga
+showed that as $n \to \infty$,
 \begin{equation}
@@
 
-When applied to our problem, We would like to test, first of all,
+When applied to our problem, we would like to test, first of all,
 whether there is a change of tail index in each of two 
@@
 stocks in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
-500 index. The null hypothesis here is that the tail index remains the
-same througout the selected period of time.
+500 index. The null hypothesis here is that the extreme quantile
+remains the same througout the selected period of time.
 \begin{figure}[htb!]
@@
   \caption{Comparison of the asymptotic distribution with empirical
-    distribution of $Q_1$ obtained from Student's t samples. The t
-    samples have the same size as the S\&P sectors' data, namely 1304
+    distribution of $Q_1$ obtained from Student's t samples. Each t
+    sample has the same size as the S\&P sectors' data, namely 1304
     observations.}
@@
 series under consideration, e.g. the Orthogonal GARCH model of
-Alexander and Chibumba (1997) \cite{alexander1997multivariate}, its
-generalization GO-GARCH of van der Weide (2002) \cite{van2002go}, as
-well as the Full Factor GARCH model of Vrontos et al (2003)
+Alexander and Chibumba \cite{alexander1997multivariate}, its
+generalization GO-GARCH of van der Weide \cite{van2002go}, as
+well as the Full Factor GARCH model of Vrontos et al 
 \cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
-model of Lanne and Saikkonen (2007) \cite{lanne2007modeling}. These
+model of Lanne and Saikkonen \cite{lanne2007modeling}. These
 models are characterized by their treatment of each series as a linear
@@
 consideration necessarily has the lowest of the tail indices of its
-contributing factors (cf. Mikosch (2013) \cite{Mikosch2013} as well
-as Andersen, Davis, Kreiss and Mikosch (2009)
+contributing factors (cf. Mikosch \cite{Mikosch2013} as well
+as Andersen, Davis, Kreiss and Mikosch 
 \cite{andersen2009handbook}).
@@
 Constant Conditional Correlation (CCC-) GARCH model of Bollerslev
-(1990) \cite{bollerslev1990modelling} specifies, for a $d$-dimensional
+\cite{bollerslev1990modelling} specifies, for a $d$-dimensional
 series $\vec X_t$
@@
 \]
-where $Z \overset{d}{=} Z_{i,t}$.
-Provided $(A_{i,1}, B_{i,1}) \neq (A_{j,1}, B_{j,1})$, it is clear
-$\kappa_i \neq \kappa_j$.
-
-As an extension to the CCC-GARCH model, Jeantheau (1998)
+where $Z \overset{d}{=} Z_{i,t}$. It is clear $\kappa_i \neq \kappa_j$
+except for special values of $A_i, B_i$ and $A_j, B_j$.
+
+As an extension to the CCC-GARCH model, Jeantheau 
 \cite{jeantheau1998strong} proposed his Extended CCC-GARCH by
@@
 $\kappa$ as $t \to \infty$. This is a direct result of the Kesten-Goldie
-theorem (cf. Kesten (1973)\cite{Kesten1973}). Thus all series under
+theorem (cf. Kesten \cite{Kesten1973}). Thus all series under
 consideration have the same tail index $\kappa$.
@@
 \label{sec:3}
-Routledge and Zin (2010) \cite{routledge2010generalized} define the GDA
+Routledge and Zin \cite{routledge2010generalized} define the GDA
 preferences over discrete states in their formula (2). A slight extension
@@
 \end{equation}
-where $\delta > 0$, and $v>0$. Equivalently
+where $\delta > 0$, $v>0$ and $b > 0$. Equivalently
 \begin{equation}
@@
 preferences are expected utility; when $\delta=1$ and $b>0$,
-preferences follow Gul's (1991) original disappointment aversion.
+preferences follow Gul's original disappointment aversion.
 An agent guided by this utility function will seek to maximize $u(v)$.
@@
   \begin{eqnarray*}
-    {\partial \over \partial \theta}\left[
-      F(b) - F(a)
-    \right] = 0 \\
     \left| \frac{\pd f(x, \theta)}{\pd \theta} \right| < \infty
@@
   u\left[ (1 - \phi) e^r + \phi e^x \right]
-  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx 
+  {1 + b \1_{\{x < q\}} \over (K' + x)^{\beta + 1}} dx 
   - b u(\delta v) F(q)
@@
   p &=& \P(X < 0) = \lim_{n \to \infty}
-  {1 \over n} \sum_{t=1}^n \1{X_t < 0}
+  {1 \over n} \sum_{t=1}^n \1_{\{X_t < 0\}}
 \end{eqnarray*}
@@
   \sqrt n \left[
-    {1 \over n}\sum_{t=1}^n \1{X_t < 0} - p
+    {1 \over n}\sum_{t=1}^n \1_{\{X_t < 0\}} - p
   \right] \to N(0, p - p^2))
@@
 \end{eqnarray*}
-Comparing this expression with \eqref{eq:lower_tail_expansion}, one
-can immediately identity $p K^\alpha$ as the scale parameter. Figure
+% Comparing this expression with \eqref{eq:lower_tail_expansion},
+Then one can identity $p K^\alpha$ as the scale parameter. Figure
 \ref{fig:scale_estimates} compare the estimates obtained
@@
     As shown in the figure to the left, $\Psi(x)$ is increasing
-    for $x > 0$. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
+    for $x > 0$. This immediately follows from the series
+    representation
+    \[
+    \Psi(x + 1) = -\gamma + \sum_{n=1}^\infty {x \over n(n + x)}
+    \quad x \neq -1, -2, -3,\dots
+    \]
+    which gives
+    \[
+    {\pd \Psi(x + 1) \over \pd x}
+    = \sum_{n=1}^\infty {1 \over (n + x)^2} > 0
+    \]
+    See Abramowitz and Stegun \cite{abramowitz1972handbook}, p.259,
+    formula 6.3.16. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
     So we have
@@
     \text{sign}\left[
-      \pd{f}{\nu}(x, \nu)
+      {\pd f \over \pd\nu}(x, \nu)
     \right]
@@
   \begin{eqnarray}
-    \pd{G(F)}{\alpha}
+    {\pd G(F) \over \pd \alpha}
     &=&
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
-    + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat\phi)
-    \pd{\hat\phi}{\alpha}(\alpha, K)
+    {\pd \mathcal U(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi)
+    + {\pd \mathcal U(F, \phi) \over \pd \phi}(\alpha, K, \hat\phi)
+    {\pd \hat\phi \over \pd \alpha}(\alpha, K)
     \label{eq:xxie3}
@@
     \label{eq:xxie4}
-    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, K, \hat \phi) = 0
+    {\pd \mathcal U(F, \phi) \over \pd \phi}(\alpha, K, \hat \phi) = 0
   \end{equation}
@@
   \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat \phi)
+    {\pd \mathcal U(F, \phi) \over \pd \alpha}(\alpha, K, \hat \phi)
     &>& 0
@@
   \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi)
-    = {\partial \over \partial \alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}]
+    {\pd \mathcal U(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi)
+    = {\partial \over \partial \alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1_{\{X < 0\}}]
   \end{eqnarray*}
@@
   \begin{eqnarray*}
-    && \pd{f(x; \alpha, K)}{\alpha} \\
+    && {\pd f(x; \alpha, K) \over \pd \alpha} \\
     &=& {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
@@
   \begin{eqnarray*}
-    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, K, \hat\phi) &>& 0
-  \end{eqnarray*}
-  As for $\pd{G(F)}{K}$, by the same argument as for
-  $\pd{G(F)}{\alpha}$, it suffices to show
-  \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0
+    {\pd \mathcal U(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi) &>& 0
+  \end{eqnarray*}
+  As for ${\pd G(F) \over \pd K}$, by the same argument as for
+  ${\pd G(F) \over \pd \alpha}$, it suffices to show
+  \[
+  {\pd \mathcal U(F, \phi) \over \pd K}(\alpha, K, \hat\phi) < 0
   \]
@@
   \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi)
+  {\pd \mathcal U(F, \phi) \over \pd K}(\alpha, K, \hat\phi)
   = {\partial \over \partial K} \E \left[
-    u((1 - \hat\phi) e^r + \hat\phi e^X) \1{X < 0}
+    u((1 - \hat\phi) e^r + \hat\phi e^X) \1_{\{X < 0\}}
   \right] 
@@
   \begin{eqnarray*}
-    && \pd{f(x; \alpha, K)}{K} \\
+    && {\pd f(x; \alpha, K) \over \pd K} \\
     &=& {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
@@
   \[
-  \pd{\mathcal U(F, \phi)}{K}(\alpha, K, \hat\phi) < 0  
+  {\pd \mathcal U(F, \phi) \over \pd K}(\alpha, K, \hat\phi) < 0  
   \]

### 2017-04-05 cv/20sec/template.tex
Add parallel computing using OpenMP to my skill list.
Wrote an abstract of the FX paper and revised my CV accordingly.

@@
 % Skill bar section, each skill must have a value between 0 an 6 (float)
-\skills {{Python/3},{SQL/4},{GNU Scientific Library/5},{Matlab/5.2},{R/5.8},{C/5},{C++/5}}
+\skills {{Python/3},{SQL/4},{Matlab/5.2},{GNU Scientific Library/5},{R/5.8},{Parallel Computing OpenMP/4.5},{C/5},{C++/5}}
 
@@
 \item Mikosch, T., de Vries, C., {\bf Xie, X.}, 2017,
-  {\it Generalized Disappointment Aversion and heavy-tail property of
-    equity return distributions}
+  {\it Tail-indices and scale parameters of returns series}
 
   {\small
-    We study the heavy tail properties of real-world equity return
-    distributions and, based on the framework of Generalized
-    Disappointment Aversion, derive how an investor's preference varies
-    with the tail distribution characteristics, namely the tail index and
-    the scale parameter. We find, when the equity return is t-distributed,
-    the GDA preference actually decreases with the degree of freedom of
-    the distribution, i.e. the tail index; we interprete this as a
-    consequence of the symmetry of the distribution. When the equity
-    return is assumed to have a shifted Pareto distribution with
-    independent parameters of the left and the right tails, we find the
-    GDA preference increasing with both the lower tail index and the scale
-    parameter. We attribute this to the risk avert nature of rational
-    investors. By numerical methods, optimal asset allocation is also
-    solved in both cases.
+    %% We study the heavy tail properties of real-world equity return
+    %% distributions and, based on the framework of Generalized
+    %% Disappointment Aversion, derive how an investor's preference varies
+    %% with the tail distribution characteristics, namely the tail index and
+    %% the scale parameter. We find, when the equity return is t-distributed,
+    %% the GDA preference actually decreases with the degree of freedom of
+    %% the distribution, i.e. the tail index; we interprete this as a
+    %% consequence of the symmetry of the distribution. When the equity
+    %% return is assumed to have a shifted Pareto distribution with
+    %% independent parameters of the left and the right tails, we find the
+    %% GDA preference increasing with both the lower tail index and the scale
+    %% parameter. We attribute this to the risk avert nature of rational
+    %% investors. By numerical methods, optimal asset allocation is also
+    %% solved in both cases.
+    It is a stylized fact that equity return distributions are
+    heavy-tailed. We discuss how an investor's preference of an equity
+    varies with the parameters of the tail distribution of the return on
+    equity. These parameters can be seen as characteristics of the
+    market and provide a quantitative description of the status of the
+    market.
+    In addition, we also find optimal asset allocation schemes
+    corresponding to different investor preferences as captured by the
+    investors' utility functions. These results can be useful when
+    considering a problem of asset allocation. Real-world stock return
+    series, which have motivated our work, are analyzed and presented.
   }

### 2017-04-05 papers/FX/draft.tex
Wrote an abstract of the FX paper and revised my CV accordingly.

@@
 \title{Tail-indices and scale parameters of returns series}
-\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei, }
+\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
 \date{\today}
@@
 \begin{abstract}
+  It is a stylized fact that equity return distributions are
+  heavy-tailed. We discuss how an investor's preference of an equity
+  varies with the parameters of the tail distribution of the return on
+  equity. These parameters can be seen as characteristics of the
+  market and provide a quantitative description of the market's status.
+  In addition, we also find optimal asset allocation schemes
+  corresponding to different investor preferences as captured by the
+  investors' utility functions. These results can be useful when
+  considering a problem of asset allocation. Real-world stock return
+  series, which have motivated our work, are analyzed and presented.
+
 %We consider an investor with preferences that accord with Generalized
@@
   &&
+  +
   \beta K'^\beta (1 - p)
@@
     \includegraphics[width=\textwidth]
-    {Energy_p.pdf}
+    {Energy_alpha_K_ci.pdf}
   \end{minipage}\hfill
@@
     \includegraphics[width=\textwidth]
-    {Energy_alpha_K_ci.pdf}
+    {Energy_p.pdf}
   \end{minipage}
@@
     \includegraphics[width=\textwidth]
-    {Consumer_Staples_alpha_K_ci.pdf}
+    {Consumer_Staples_p.pdf}
   \end{minipage}
@@
     \includegraphics[width=\textwidth]
-    {Information_Technology_alpha_K_ci.pdf}
+    {Information_Technology_p.pdf}
   \end{minipage}

### 2017-04-07 cv/20sec/template.tex
Revised my CV.

@@
 %----------------------------------------------------------------------------------------
+\newenvironment{aperiod}{
+  \begin{list}{}{
+    \setlength{\leftmargin}{1.5em}
+    \setlength{\itemsep}{1em}
+    \setlength{\parskip}{0pt}
+    \setlength{\parsep}{0.25em}
+  }
+}{
+  \end{list}
+}
 
@@
   \item M.Sc. of Computer Science
-  \item 7 years work experience in programming 
+  \item 6.5 years work experience in programming 
   \item strong in statistical analysis \& numerical methods
@@
 % Skill bar section, each skill must have a value between 0 an 6 (float)
-\skills {{Python/3},{SQL/4},{Matlab/5.2},{GNU Scientific Library/5},{R/5.8},{Parallel Computing OpenMP/4.5},{C/5},{C++/5}}
+\skills {{Java/3},{Python/3},{SQL/4},{Matlab/5.2},{GNU Scientific Library/5},{R/5.8},{Parallel Computing OpenMP/4.5},{C/5},{C++/5}}
 
@@
   \begin{itemize}
-  \item Ph.D. in Mathematical Statistics
-  \item Copenhagen University
+  \item \underline{Ph.D. in Mathematical Statistics}
+  \item Copenhagen University, Denmark
   \item Areas of research:
@@
   \begin{itemize}
-  \item M.Sc. in Mathematical Physics
-  \item Lund University
+  \item \underline{M.Sc. in Mathematical Physics}
+  \item Lund University, Sweden
   \item Thesis project: Distribution of eigenvalues of large
@@
     \end{itemize}
+  \end{itemize}
+\end{minipage}
+\vspace{3mm}
+
+\begin{minipage}{0.15\linewidth}
+  Sep. 2005 - June 2008
+\end{minipage}\hfill
+\begin{minipage}{0.85\linewidth}
+  \begin{itemize}
+  \item \underline{M.Sc.Computer Science and Engineering}
+  \item Helsinki University of Technology
+  \item {\it Major Subject:} Mobile Computing - Services and Security
+  \item {\it Average grade:} 4.17/5.
+  \end{itemize}
+\end{minipage}
+\vspace{3mm}
+
+\begin{minipage}{0.15\linewidth}
+  Sep. 2001 - June 2005
+\end{minipage}\hfill
+\begin{minipage}{0.85\linewidth}
+  \begin{itemize}
+  \item \underline{B.Sc.Computer Science and Engineering}
+  \item Beijing Institute of Technology
   \end{itemize}
@@
 \begin{itemize}
-\item Davis, R.A., Heiny, J., Mikosch, T. and {\bf Xie, X.},
+\item {\bf Xie, X.}, Davis, R.A., Heiny, J. and Mikosch, T.
   2016. {\it Extreme value analysis for the sample autocovariance
@@
 
-\item Jan\ss en, A., Mikosch, T., Rezapour, M., {\bf Xie, X},
+\item {\bf Xie, X}, Jan\ss en, A., Mikosch, T. and Rezapour, M.
   2016. {\it The eigenvalues of the sample covariance matrix of a
@@
 \begin{itemize}
-\item Mikosch, T., de Vries, C., {\bf Xie, X.}, 2017,
+\item {\bf Xie, X.}, Mikosch, T. and de Vries, C., 2017,
   {\it Tail-indices and scale parameters of returns series}
@@
   {\small
-    %% We study the heavy tail properties of real-world equity return
-    %% distributions and, based on the framework of Generalized
-    %% Disappointment Aversion, derive how an investor's preference varies
-    %% with the tail distribution characteristics, namely the tail index and
-    %% the scale parameter. We find, when the equity return is t-distributed,
-    %% the GDA preference actually decreases with the degree of freedom of
-    %% the distribution, i.e. the tail index; we interprete this as a
-    %% consequence of the symmetry of the distribution. When the equity
-    %% return is assumed to have a shifted Pareto distribution with
-    %% independent parameters of the left and the right tails, we find the
-    %% GDA preference increasing with both the lower tail index and the scale
-    %% parameter. We attribute this to the risk avert nature of rational
-    %% investors. By numerical methods, optimal asset allocation is also
-    %% solved in both cases.
     It is a stylized fact that equity return distributions are
@@
     series, which have motivated our work, are analyzed and presented.
+  }
+\end{itemize}
+
+\pagebreak
+\makeprofile % Print the sidebar
+
+\section{Papers in Finalization}
+\begin{itemize}
+\item {\bf Xie, X.} Collamore, J, 2017
+  {\it An efficient estimator of the probability of very large losses posed by
+    GARCH(p,q) processes}
+
+  {\small
+    We propose an efficient importance sampling estimator of the
+    probability of excursion over high thresholds (causing large
+    losses) by Multivariate Markov chains e.g. a GARCH(p,q) process or
+    a multivariate GARCH process.
+    Although stationary distributions of these processes are known
+    to have power-law tails, precise evaluation of the tail
+    probabilities is difficult because the exact distribution is
+    unknown. Moreover, naive simulation methods lead to erractic
+    estimates. Our importance sampling estimator presents a solution.
   }
@@
 
-\newpage % Start a new page
+% \newpage % Start a new page
 
-\makeprofile % Print the sidebar
+% \makeprofile % Print the sidebar
 
 \section{Work Experience}
+% \begin{twenty} % Environment for a list with descriptions
+% 	\twentyitem{1900}{Alice in Wonderland-The Circra (1900's) Silent Film.}{Film}{The first Alice on film was over a hundred years ago.}
+% 	\twentyitem{1933}{Alice in Wonderland 1933 version.}{Film}{This film stars Ethel griffies and Charlotte Henry. It was a box office flop when it was released.}
+% 	\twentyitem{1951}{Disney Film.}{Film}{Walt Disney brings Lewis Carroll's fantasy story to life in this well done animated classic.
+Even though many elements from the book were dropped, such as the duchess with the baby pig and mock turtle, this version is without a doubt the most famous Alice adaption made.}
+% 	%\twentyitem{<dates>}{<title>}{<location>}{<description>}
+% \end{twenty}
+\begin{aperiod}
+%% \begin{small}
+\item {\it \href{http://www.stericsson.com}{\tt
+      ST-Ericsson, Lund, Sweden}, Audio, Multimedia, Development
+    Engineer, Sep 2008 - Jan 2011}
 
-\begin{twenty} % Environment for a list with descriptions
-	\twentyitem{1900}{Alice in Wonderland-The Circra (1900's) Silent Film.}{Film}{The first Alice on film was over a hundred years ago.}
-	\twentyitem{1933}{Alice in Wonderland 1933 version.}{Film}{This film stars Ethel griffies and Charlotte Henry. It was a box office flop when it was released.}
-	\twentyitem{1951}{Disney Film.}{Film}{Walt Disney brings Lewis Carroll's fantasy story to life in this well done animated classic.
-Even though many elements from the book were dropped, such as the duchess with the baby pig and mock turtle, this version is without a doubt the most famous Alice adaption made.}
-	%\twentyitem{<dates>}{<title>}{<location>}{<description>}
-\end{twenty}
+  My responsibilities included developing audio low-level software,
+  typically device drivers, in the Linux OS. This involved designing and
+  implementing data structures and procedures so that the hardware
+  devices were used in the optimal way, in terms of parallelism,
+  power consumption, robustness, and ease of maintenance.
 
-\section{Other information}
+  {\it C, C++, Java, Linux}
+\item {\it \href{http://www.hiit.fi}{\tt Helsinki
+      Institute of Information Technology, Helsinki}, Oct 2007 - May 2008}
 
-\subsection{Review}
+  Funded by HIIT, I wrote my thesis ``A Mutual authentication and Key
+  Agreement Protocol for the UMTS Network''. The algorithm of my
+  protocol proposal was based on the Diffie-Hellman Exchange, but had
+  significant modifications and extensions so that the mutual
+  authentication between the mobile user and the network was delegated
+  to the serving instead of the home network. The protocol design
+  incorporated puzzle schemes that demanded unbalanced computing
+  resources on the two sides to mitigate Denial-of-Service attacks.
 
-Alice approaches Wonderland as an anthropologist, but maintains a
-strong sense of noblesse oblige that comes with her class status. She
-has confidence in her social position, education, and the Victorian
-virtue of good manners. Alice has a feeling of entitlement,
-particularly when comparing herself to Mabel, whom she declares has a
-``poky little house," and no toys. Additionally, she flaunts her
-limited information base with anyone who will listen and becomes
-increasingly obsessed with the importance of good manners as she deals
-with the rude creatures of Wonderland. Alice maintains a superior
-attitude and behaves with solicitous indulgence toward those she
-believes are less privileged.
+  {\it C, C++, Linux}
+\item {\it
+    \href{http://www.goodmillsystems.com}{\tt Goodmillsystems,
+      Helsinki}, Software Designer, May 2007 - Sep 2007}
 
-%\section{Other information}
+  I developed a daemon program that ran on the company's smart router
+  product to collect statistical data about the router's network
+  traffic. The emphasis of the program was high performance in the
+  context of competing read and write accesses.
 
-%\subsection{Review}
+  {\it C, C++, Linux}
+\item {\it \href{http://www.cardinal.fi/}{\tt
+      Cardinal Information Systems, Helsinki}, Software
+    Developer, Sep 2006 - April 2007}
 
-%Alice approaches Wonderland as an anthropologist, but maintains a strong sense of noblesse oblige that comes with her class status.
-She has confidence in her social position, education, and the Victorian virtue of good manners.
-Alice has a feeling of entitlement, particularly when comparing herself to Mabel, whom she declares has a ``poky little house," and no toys.
-Additionally, she flaunts her limited information base with anyone who will listen and becomes increasingly obsessed with the importance of good manners as she deals with the rude creatures of Wonderland.
-Alice maintains a superior attitude and behaves with solicitous indulgence toward those she believes are less privileged.
+  I implemented the FLUTE protocol for the company's DVB-H
+  mobile-end middleware. The implementation was multi-threaded and was
+  built on the {\it pthread} library. Good performance and robustness
+  were crucial requirements on the implementation.
 
-%----------------------------------------------------------------------------------------
+  {\it C, C++, Linux}
+\item {\it
+    \href{http://www.digitalchocolate.com/}{\tt Digital Chocolate,
+      Helsinki, Finland}, Game Engineer, Feb 2006 - Sep 2006}
+
+  My responsibilities were porting and customizing mobile
+  phone games for various handsets.
+
+  {\it Java, J2ME, Bash}
+\item {\it {\tt Beijing Holystone Ltd, Beijing}, Software
+    Engineer, April 2004 - June 2005}
+
+  I developed JSP pages and database routines for
+  the company's information system.
+
+  {\it Java, J2EE}
+%% \end{small}
+\end{aperiod}
+
+\section{Interests}
+\begin{itemize}
+\item Algorithmic trading, alpha generation
+\item statistical inference, machine learning, neural network
+\end{itemize}
 

### 2017-04-09 cv/CoverLetter/CitiMQA_QuantitativeAnalyst.tex
Added a Cover letter for Citi.

%% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

\documentclass[10pt,a4paper, gentium]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

% moderncv themes
\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{red}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
\renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean

% adjust the page margins
\usepackage[scale=0.75]{geometry}
%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
%\setlength{\makecvtitlenamewidth}{10cm}           % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks. be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...

% personal data
\name{Xie}{Xiaolei}
%\title{Resumé title}                               % optional, remove / comment the line if not wanted
\address{Kiviksgatan 7C}{21440 Malm\"o}{Sweden}% optional, remove / comment the line if not wanted; the "postcode city" and and "country" arguments can be omitted or provided empty
\phone[mobile]{+46~76~060~6638}                   % optional, remove / comment the line if not wanted
%\phone[fixed]{+2~(345)~678~901}                    % optional, remove / comment the line if not wanted
%\phone[fax]{+3~(456)~789~012}                      % optional, remove / comment the line if not wanted
\email{xie.xiaolei@gmail.com}                               % optional, remove / comment the line if not wanted
%\homepage{www.johndoe.com}                         % optional, remove / comment the line if not wanted
%\extrainfo{additional information}                 % optional, remove / comment the line if not wanted
%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file
%\quote{Some quote}                                 % optional, remove / comment the line if not wanted

% to show numerical labels in the bibliography (default is to show no labels); only useful if you make citations in your resume
%\makeatletter
%\renewcommand*{\bibliographyitemlabel}{\@biblabel{\arabic{enumiv}}}
%\makeatother
%\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}% CONSIDER REPLACING THE ABOVE BY THIS

% bibliography with mutiple entries
%\usepackage{multibib}
%\newcites{book,misc}{{Books},{Others}}
%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%-----       letter       ---------------------------------------------------------
% recipient data
\recipient{Cross Asset Quantitative Analysis team}{Citi}{London}
\date{\today}
\opening{Dear Sir/Madam,}
\closing{Yours sincerely,}
%\enclosure[Attached]{curriculum vit\ae{}}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
% \enclosure[Attached]{curriculum vit\ae{}, Master's thesis in
%   Econophysics}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
\makelettertitle

I see on Citi's website that you are recruiting a quantitative
analyst to join your cross asset quantitative analysis team. In the
description of the job, you emphasized knowledge and skills in time
series analysis, C++ programming and understanding of the financial
market. These are indeed areas in which I find confidence. So I would
like to ask you for your consideration.

I am currently a Ph.D student at Copenhagen University, researching in
Mathematical statistics and particularly in time series analysis. I
have co-authored and published two papers on eigen systems of sample
covariance matrices with a focus on heavy-tailed time series. These
results can be helpful in principle component analysis.

I am also finalizing two other papers. One of them is about how an
investor's preference (as characterized by {\em Generalized
  Disappointment Aversion}) of an equity varies with the tail
parameters of the stationary distribution of the equity. I think this
result can be applied to problems of portfolio construction as well as
assessment of the overall market risk. As part of the paper, I also did
statistical analysis on the tail parameters of equities in the S\&P
500 index, specifically, the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors. R and C++ programming are
indispensible part of the work.

The other paper of mine is about estimation of the probability of very
large losses posed by portfolios whose volatility dynamics are
described by multivariate GARCH or GARCH(p,q) processes.

\makeletterclosing

\end{document}

### 2017-04-10 cv/CoverLetter/CitiMQA_QuantitativeAnalyst.tex
Revised my CV and the cover letter.

@@
 \opening{Dear Sir/Madam,}
-\closing{Yours sincerely,}
+\closing{Best Regards,}
 %\enclosure[Attached]{curriculum vit\ae{}}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
@@
 analyst to join your cross asset quantitative analysis team. In the
-description of the job, you emphasized knowledge and skills in time
+description of the job, you emphasised knowledge and skills in time
 series analysis, C++ programming and understanding of the financial
-market. These are indeed areas in which I find confidence. So I would
-like to ask you for your consideration.
+market, which are indeed areas that I know well and am passionate
+about. So I would like to ask you for your consideration.
 
-I am currently a Ph.D student at Copenhagen University, researching in
-Mathematical statistics and particularly in time series analysis. I
-have co-authored and published two papers on eigen systems of sample
-covariance matrices with a focus on heavy-tailed time series. These
-results can be helpful in principle component analysis.
+I am currently a PhD student at Copenhagen University, researching in
+Mathematical statistics and particularly in time series analysis,
+extreme value theory and random matrix theory. I am available for
+immediate full-time work and expect to receive my degree on 30th
+September, 2017. So far I have co-authored and published two papers on
+eigen systems of sample covariance matrices, with a focus on
+heavy-tailed (e.g. power-law tail) time series. These results are
+applicable to principle component analysis, for example.
 
-I am also finalizing two other papers. One of them is about how an
-investor's preference (as characterized by {\em Generalized
+I am also finalising two other papers. One of them is about how an
+investor's preference (as characterised by {\em Generalised
   Disappointment Aversion}) of an equity varies with the tail
-parameters of the stationary distribution of the equity. I think this
-result can be applied to problems of portfolio construction as well as
-assessment of the overall market risk. As part of the paper, I also did
-statistical analysis on the tail parameters of equities in the S\&P
-500 index, specifically, the ``Energy'', ``Consumer Staples'' and
-``Information Technology'' sectors. R and C++ programming are
-indispensible part of the work.
+parameters of the stationary distribution of the equity. This
+result can be applied to, for example, portfolio construction and
+assessment of the overall market risk. As part of the paper, I have
+also done statistical analysis on the tail parameters of equities in
+the S\&P 500 index, specifically, the ``Energy'', ``Consumer Staples''
+and ``Information Technology'' sectors. R and C++ programming are
+indispensable part of the work.
 
-The other paper of mine is about estimation of the probability of very
+The fourth paper of mine is about estimation of the probability of very
 large losses posed by portfolios whose volatility dynamics are
-described by multivariate GARCH or GARCH(p,q) processes.
+described by univariate GARCH(p,q) or multivariate GARCH
+processes. Because the stationary distribution of such a process is
+not known exactly, the aforementioned probability as well as the
+related risk measures {\em Value at Risk} and {\em Expected
+  Shortfall} cannot be computed precisely. In the paper I propose 
+an estimator based on importance sampling and Markov chain theory and
+naturally, would be more than happy to see it applied in practical
+risk management. I have implemented the estimator in C++, and in
+particular, have used the OpenMP directives to achieve high
+performance via parallel computing.
+
+Citi is a great investment bank that I have known since my
+childhood. I know taking a role in this esteemed organisation presents
+many challenges and opportunities. Through constant endeavour towards
+ever greater excellence, I hope to develop a productive career that I
+will be proud of by contributing my part to the success of Citi.
+
+I look forward to discussing the job with you in more
+details. You can always call me at +46 76 060 6638 or email me at
+xie.xiaolei@gmail.com. It would be my pleasure to hear from you.
 

### 2017-04-10 cv/CoverLetter/CitiMQA_QuantitativeAnalyst.tex
Revised the Cover letter again.

@@
 analyst to join your cross asset quantitative analysis team. In the
-description of the job, you emphasised knowledge and skills in time
+description of the job, you emphasise knowledge and skills in time
 series analysis, C++ programming and understanding of the financial
@@
 I am also finalising two other papers. One of them is about how an
-investor's preference (as characterised by {\em Generalised
+investor's preference (as quantified by {\em Generalised
   Disappointment Aversion}) of an equity varies with the tail
@@
 also done statistical analysis on the tail parameters of equities in
-the S\&P 500 index, specifically, the ``Energy'', ``Consumer Staples''
+the S\&P 500 index, specifically the ``Energy'', ``Consumer Staples''
 and ``Information Technology'' sectors. R and C++ programming are
@@
 large losses posed by portfolios whose volatility dynamics are
-described by univariate GARCH(p,q) or multivariate GARCH
-processes. Because the stationary distribution of such a process is
-not known exactly, the aforementioned probability as well as the
+time series modeled by univariate GARCH(p,q) or multivariate
+GARCH processes. Because the stationary distribution of such a process
+is not known exactly, the aforementioned probability as well as the
 related risk measures {\em Value at Risk} and {\em Expected

### 2017-04-11 cv/CoverLetter/CitiMQA_QuantitativeAnalyst.tex
1. Changed my cover letter template. 2. updated the title of the paper with Jeff.

@@
 Mathematical statistics and particularly in time series analysis,
-extreme value theory and random matrix theory. I am available for
-immediate full-time work and expect to receive my degree on 30th
+extreme value theory and random matrix theory. I will be available for
+full-time work on 1st July 2017. I expect to receive my degree on 30th
 September, 2017. So far I have co-authored and published two papers on
@@
 large losses posed by portfolios whose volatility dynamics are
-time series modeled by univariate GARCH(p,q) or multivariate
+time series modelled by univariate GARCH(p,q) or multivariate
 GARCH processes. Because the stationary distribution of such a process

### 2017-04-11 papers/FX/draft.tex
Merged Thomas' revision.

--- 2017-04-05 papers/FX/draft.tex
@@
 %\usepackage{enumerate}
-\usepackage[bookmarks=true]{hyperref}
+%\usepackage[bookmarks=true]{hyperref}
 %\usepackage{bookmark}
@@
 \newcommand{\bfX}{{\bf X}}
+\newcommand{\bfW}{{\bf W}}
 \newcommand{\bfB}{{\bf B}}
@@
 \begin{document}
-\title{Tail-indices and scale parameters of returns series}
-\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei, }
+\title{Do return series have power-law tails with the same index?}
+\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
 \date{\today}
@@
 returns of speculative prices are {\em heavy-tailed.} There is no agreement
-in the literature how heavy tails really are. For example, Barndorff-Nielsen and Shephard
-\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor
-``semi-heavy'' tails, i.e., tails which are comparable with those of a
-gamma \ds . On the other hand, tails of returns have been studied in
-great detail in the extreme value community. Among extreme value
-specialists there is general agreement that returns $X_t$ have tails
-of power-law-type, i.e.,
+in the literature about how heavy these tails really are. For example, Barndorff-Nielsen and Shephard
+\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails
+which are comparable with those of a gamma \ds . On the other
+hand, tails of returns have been studied in great detail 
+in the extreme value community. Among extreme value specialists
+there is general agreement that returns $X_t$ have tails of power-law-type, i.e.,
 \beam\label{eq:1}
-\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad
-\P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
+\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad \P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
 \eeam  
-where $c_{\pm}$ are positive constants.\footnote{Here and in what
+where $c_{\pm}$, $\alpha_{\rm up}$ and $\alpha_{\rm low}$ are positive constants.\footnote{Here and in what
   follows, $f(x)\sim g(x)$ for positive \fct s $f$ and $g$ means 
 that $f(x)/g(x)\to 1$ as $\xto$.} See for example, Embrechts et
-al. \cite{Embrechts1997}, Jansen and de Vries
+al. \cite{embrechts:klueppelberg:mikosch:1997}, Jansen and de Vries
 \cite{jansen1991frequency}, Mikosch \cite{mikosch:2003}, Resnick
 \cite{resnick:2007}.  
-In the extreme value literature it is common to replace the constants $c_\pm$ by a suitable
+In the extreme value literature it is common to replace the constants $c_\pm$ by suitable
 {\em slowly varying} \fct s; cf. Embrechts et
-al. \cite{Embrechts1997}, Chapter 3. In this
+al. \cite{embrechts:klueppelberg:mikosch:1997}, Chapter 3. In this
 paper, for the  
@@
 There are some good theoretical reasons for the appearance of power-law tails in situations where certain moments 
-of data are believed to be infinite. Tails of the type \eqref{eq:1}
+of data are believed to be infinite. Tails of type \eqref{eq:1}
 describe the maximum domain of attraction of the  
-Fr\'echet \ds\ $\Phi_\alpha(x)=\exp(-x^{-\alpha})$ for $x>0$. Equivalently, power-law tails are prescribed by the 
-generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e.,
+Fr\'echet \ds\ $\Phi_{\alpha_{\rm up}}(x)=\exp(-x^{-\alpha_{\rm up}})$ for $x>0$, i.e., scaled maxima of an iid \seq\ $(X_t)$ with
+upper tail described in \eqref{eq:1} converge in \ds\ to $\Phi_{\alpha_{\rm up}}$.
+Equivalently, power-law tails are prescribed by the 
+generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e., for a suitable positive scaling \fct\ $a(u)$,
 \beao
-\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha)^{-\alpha}\,,\qquad u\to\infty\,.
+\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha_{\rm up})^{-\alpha_{\rm up}}\,,\qquad u\to\infty\,.
 \eeao
@@
 Moreover, in the literature the {\em estimated} tail-indices $\hat \alpha$ (both in the left and right tails) 
-are typically found in the range $(2.5,4)$. For an illustration, see
+are typically found in the range $(2,4)$. For an illustration, see
 Figure~\ref{fig:1} where estimates $\hat \alpha_{\rm low}$ 
-in three sectors of the Standard \& Poor 500 index  are shown. The
+in three sectors of the Standard \& Poors 500 index  are shown. The
 estimates are based on 1304 observations of daily return 
-data from 1 January 2010 to 1 January 2015.
+data from 1 January 2010 to {\red why 1 January?} 1 January 2015.
 \par
@@
 \par
-The estimator of the tail-index $\alpha$  which favored in the literature  is the {\em Hill estimator}; the graphs in 
+The estimator of the tail-index $\alpha>0$  in the model
+\beao
+\P(X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
+\eeao
+favored in the literature  is the {\em Hill estimator}; the graphs in 
 Figure~\ref{fig:1} are based on this estimator. We introduce this estimator in Section~\ref{sec:1} and discuss 
@@
 estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). It turns out 
-that the simulateneous estimation of the tail-index and the scale parameter are strongly related.
+that the simultaneous estimation of the tail-index and the scale parameter are strongly related.
 \par
@@
 for daily or, more generally, low \fre y 
-return data. In particular, we address the power-law tails of univariate and multivariate GARCH models as potential models
+return data. In particular, in Section~\ref{subsec:garch} we address the power-law tails of univariate and multivariate GARCH models as potential models
 for a set of return data from distinct assets. As a matter of fact, under mild conditions, the aforementioned models
-have power-tails due to their relation with so-called {\em \sre s}.
+have power-law tails due to their relation with so-called {\em \sre s}. Moreover, some of the {\em standard} multivariate 
+GARCH models as the CCC or BEKK ensure that the marginal \ds s have power-law tails with the same index.
 \par
@@
 introduced by Routledge and Zin \cite{routledge2010generalized}. GDA
-is an extension of the concept of Disappointment Aversion (DA) of Gul
- \cite{gul1991theory}. The Gul paper derives the DA from first principles (axiomatic).
+is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991}. The Gul paper derives the DA from first principles (axiomatic).
 \par 
@@
 that this series constitutes a strictly stationary \seq . In what follows, we focus
-on the left tail of the \ds , i.e., on the losses. However, it is common to present the tail index estimators
-for positive data. Therefore we will multiply the losses by minus one, swapping the negative with the positive values.
- The two parameters -- the tail-index  $\alpha_{\rm low}$ and the scale parameter
-$c_-$ -- are assumed positive. They play crucial roles for the understanding the risk hidden in the data, hence 
+on the left tail of the \ds , i.e., on the losses. 
+However, it is common to present the tail index estimators
+for positive data. Therefore we will multiply the losses $X_t$ by minus one, swapping the negative with the positive values.
+For simplicity, we also suppress subscripts in the notation:
+\beam\label{eq:1a}
+\P(-X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
+\eeam
+where we assume that the two parameters -- the tail-index  $\alpha$ and the scale parameter
+$c$ -- are positive. They play crucial roles for the understanding of the risk hidden in the data, hence 
 for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
@@
 \subsection{Hill estimates of lower tail indices}\label{sec:Hill}
-Various estimators of the tail index $\alpha_{\rm low}$  in the model
-\eqref{eq:1} have been proposed in the literature;
-see Embrechts et al. \cite{Embrechts1997}, de
-Haan and Ferreira \cite{dehaan:ferreira:2006}, Resnick
+Various estimators of the tail index $\alpha$  in the model
+\eqref{eq:1a} have been proposed in the literature;
+see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, de
+Haan and Ferreira \cite{haan:ferreira:2006}, Resnick
 \cite{resnick:2007}. The most popular among them was introduced  by
 Hill \cite{hill1975simple}.
-Given a sample $X_1,\ldots,X_n$ from \eqref{eq:1}, one can calculate
+Given a sample $-X_1,\ldots,-X_n$ whose marginal \ds\ satisfies  \eqref{eq:1a}, calculate
 the order statistices $X_{(1)}\le \cdots\le X_{(n)}$  and construct the
@@
 \beao%\label{eq:Hill_index}
-  \hat \alpha_{\rm low} = \Big(
+  \hat \alpha_k = \Big(
     {1 \over k} \sum_{i=1}^k \ln {X_{(n-i+1)} \over X_{(n-k)}}
@@
 Here $k$  is the number of upper order statistics in the sample used
-for the estimation. The estimator $\hat \alpha_{\rm low}$ is 
+for the estimation. The estimator $\hat \alpha_k$ is 
 a maximum-likelihood estimator of $\alpha$ based on the $k$ upper
@@
 we multiplied the data by minus one)
-\beao
-\P(X_t>x)= \dfrac{K^\alpha}{x^\alpha}\,,\quad x>K\,,
-\eeao
+\beam\label{eq:3}
+\P(-X_t>x)= \dfrac{K^\alpha}{x^\alpha}\,,\qquad x>K\,,
+\eeam
 under the hypothesis that we do not know the (high) threshold value
@@
 \cite{drees:rootzen:2010} give perhaps most general conditions for
-dependent \seq s and de Haan and Ferreira provide
+dependent \seq s and de Haan and Ferreira \cite{haan:ferreira:2006} provide
 a complete \asy\ theory in the iid case.
@@
 if $k$ is too large the order statistics are too close to the center
-of the \ds\ of the $X_t$. This leads to bias
-of the estimator. On the other hand, by construction $\hat \alpha_{\rm
-  low}$ is an average of $k$ log-differences of the 
+of the \ds\ of the $-X_t$, leading to a bias
+of the estimator. On the other hand, by construction, $\hat \alpha_k$ is an average of $k$ log-differences of the 
 data. Therefore, the variance of the estimator is the larger the
@@
 to choose $k=k_n$ \st\ $k_n\to\infty$ and $k_n/n\to 0$ as $\nto$. This
-fact does not make the estimation of $\alpha_{\rm low}$ an  
+fact does not make the estimation of $\alpha$ an  
 easy matter: one has to choose a ``small'' value $k$ which is not
 ``too large''. For practical purposes,
-a so-called {\em Hill plot} is recommended where $\hat \alpha$ is
+a so-called {\em Hill plot} is recommended where $\hat \alpha_k$ is
 plotted for a variety of $k$-values, corresponding to
 some high  quantile $X_{(n-k)}$ of the data. Then $k$ is chosen from a
-region in the plot where $\hat \alpha_{\rm low}$ is relatively
+region in the plot where $\hat \alpha_k$ is relatively
 stable. For example, in Figure~\ref{fig:1} we have chosen $k=50$ from
@@
 requires some expertise; for some guidance
-see Embrechts et al. \cite{Embrechts1997},
+see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997},
 Resnick \cite{resnick:2007} and Drees et al. 
-\cite{drees:resnick:2000}. In Figure~\ref{fig:1},
-we exhibit 95\% \asy\ confidence bands which is based on the \clt
+\cite{drees:resnick:2000}. 
+\par
+In Figure~\ref{fig:1},
+we exhibit 95\% \asy\ confidence bands derived from  the \clt
 \beam\label{eq:2}
-\sqrt k\, \big(\hat \alpha_{\rm low} - \alpha_{\rm low}\big) \std N(0, \alpha^2)\,,
+\sqrt k\, \big(\hat \alpha_k - \alpha\big) \std N(0, \alpha^2)\,,
 \eeam
-i.e., $\hat \alpha_{\rm low}$ is \asy ally unbiased and has variance
+i.e., $\hat \alpha$ is \asy ally unbiased and has variance
 $\alpha^2/k$. Since $k/n\to 0$ this means that the 
 confidence bands are significantly larger than the classical
-$\sqrt{n}$-rates. This fact is one explanation for the fact that it is
+$1/\sqrt{n}$-rates. This fact is one explanation for the fact that it is
 difficult to say something meaningful about the true value of
-$\alpha_{\rm low}$. There exist various other reasons why one should
+$\alpha$. There exist various other reasons why one should
 not have 100\% trust in the confidence bands shown in
@@
 Hill estimator may exhibit significant bias; see Embrechts et
-al. \cite{Embrechts1997} for illustrations 
+al. \cite{embrechts:klueppelberg:mikosch:1997} and Resnick \cite{resnick:1987} for illustrations 
 of this fact leading to so-called ``Hill horror plots''. Moreover, the
 Hill estimator is rather sensitive to non-stationarity 
-of the data and to dependence. For example, results in Drees and
+of the data and to dependence. For example, results in Drees \cite{drees:2008}, and Drees and
 Rootz\'en \cite{drees:rootzen:2010} show that the \asy\ variance of the
@@
 Again, only under he assumption of a concrete model like GARCH these
-confidence bands could be evaluated.
+confidence bands can be evaluated and therefore the bands shown in Figure~\ref{fig:1}
+just show some benchmark which holds in the iid case and under additional conditions on the tail \asy s.
 \begin{figure}[htb!]
@@
   \end{minipage}
-  \caption{\small Hill estimates of lower tail indices of
-    daily return series in sectors of Standard \& Poor
-    500 index. The figures from top to bottom are ``Energy'',
+  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail indices $\alpha$ of
+    daily return series in sectors of the S\&P 500
+    index. The data span
+    from 2010-01-01 to 2015-01-01 and comprise $n=1304$ observations.
+The graphs from top to bottom correspond to the ``Energy'',
     ``Consumer Staples'' and ``Information Technology'' sectors.
-    Each circle coresponds to a Hill estimate; the grey
+    Each circle corresponds to a Hill estimate $\hat\alpha_k$; the grey
     triangles above and below it mark the 97.5\% and 2.5\% quantiles
-    of its approximate normal distribution.
-    The low and high red lines mark the median of the 2.5\% quantiles
-    and median of the 97.5\% quantiles, respectively.
-    The data are from {\it Yahoo Finance} and the labels on
-    the horizontal axis are Yahoo symbols of the stocks. The data span
-    from 2010-01-01 to 2015-01-01 and comprise $n=1304$
-    observations. $k=50$ lower order statistics are used
-    for computing $\alpha_n$. 
+    of its approximate normal distribution; see \eqref{eq:2} and the discussion following it for an 
+interpretation.
+    The lower and upper red lines mark the medians of the 2.5\% 
+    and 97.5\% quantiles, respectively, evaluated from all stocks in the sector.
+    The data are taken from {\it Yahoo Finance}; the labels on
+    the horizontal axes are Yahoo symbols of the stocks. 
   }\label{fig:1}
@@
 
-In Figure~\ref{fig:1} we see that confidence intervals of the Hill
+In Figure~\ref{fig:1} we see significant overlap of the confidence intervals of the Hill
 estimates of the losses in the ``Energy'' and ``Consumer Staples''
-sectors of the S\& P 500 index, as well as those of a 
-large portion of losses in the ``Information Technology'' sector,
-overlap. This fact indicates that the returns in each sector may 
+sectors of the S\&P 500 index, as well as those of a 
+large portion of losses in the ``Information Technology'' sector.
+This fact indicates that the returns in each sector may 
 have comparable tail indices.
-
-Hoga's test on the change of extreme quantiles
-may provide some insight about how similar these tail indices are,
-because different tail indices are likely to result in different
-extreme quantiles. Nevertheless, changes in the extreme quantiles can
-be the result of changed tail indices but can also be the result of
-changed scale parameters. Thus we first investigate the scale
-parameters of stocks in the same sectors before applying Hoga's test.
-
-\subsection{Hill estimates of lower-tail scale parameters}
-\label{sec:HillScaleEstimates}
-
-For convenience, we denote the scale parameter by $A$.
-%% It is the constant $J_*(0)$ in the Taylor expansion \eqref{eq:scale_parameter}.
-Hill \cite{hill1975simple} put forward an estimator of $A$ using his
-tail index  estimator:
-\[
-\hat A = {k \over n} X_{(k+1)}^{\hat \alpha}
-\]
-where $X_{(k+1)}$ is the $(k+1)$-th upper order statistic of the
-sequence $X_t$ as defined in section\ref{sec:Hill}. $\hat \alpha$
-is the Hill tail index estimator.
-Estimates of $A$ of the ``Energy'', ``Consumer Staples'' and
+\par
+Hoga's \cite{hoga:2016} test about the change of extreme quantiles
+may provide some further insight about how similar these tail indices are.
+Different tail indices are likely to result in different
+extreme quantiles. Nevertheless, changes in the extreme quantiles may also
+result from changing scale parameters in the tail. Therefore  we first investigate the scale
+parameters of daily stock returns in the same sectors of S\&P 500 before we apply the test.
+
+\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
+We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$. 
+Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
+derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
+Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
+It is given by
+\beao
+\hat K_k = {k \over n} X_{(n-k+1)}\,.
+\eeao
+Estimates of $K$ {\red or $K^\alpha$?} of the ``Energy'', ``Consumer Staples'' and
 ``Information Technology'' sectors of the S\&P 500 index are computed
-using this method. The results are shown in figure
-\ref{fig:sectors_parameters}, along with estimates obtained
+using this method. {\blue The results are shown in Figure~\ref{fig:sectors_parameters} {\red ???} along with estimates obtained
 using the hybrid method described in section
-\ref{sec:hybrid_estimation}.
+\ref{sec:hybrid_estimation}.} {\red Figures of the Hill estimates of $K$ or $K^\alpha$ with Hill confidence bands. 
+Discuss confidence bands and what you see in the data.}
 
@@
 %% the hypothesis that the tail indices of these return series are equal.
-%% Hoga \cite{Hoga2017} devised tests suitable this purpose. We adopt his
+%% Hoga \cite{hoga:2016} devised tests suitable this purpose. We adopt his
 %% approach but use our own implementation. The next section has more
@@
 
-\subsection{Hoga's Test}
-\label{sec:Hoga}
-
-Hoga \cite{Hoga2017} proposed his test for a time series
-$\{X_i\}_{i=1}^n$ with tail functions
-$\bar F_i(\cdot)$, i.e. $\P(X_i > x) = \bar F_i(x)$. The null
-hypothesis is
-\begin{hypothesis}
-  $\mathcal H_0$: For a given $p$ and all $1 \leq i < j \leq n$,
-  $F_i^{-1}(p) = F_j^{-1}(p)$
-\end{hypothesis}
-while the alternative is $\exists t^* \in [t_0, 1 - t_0]$ such that
-\[
-\bar F_1^{-1}(p) = \bar F_2^{-1}(p) = \cdots =
-\bar F_{\floor{t^* n}}^{-1}(p) \neq
-\bar F_{\ceil{t^* n}}^{-1}(p) = \cdots = F_n^{-1}(p)
-\]
-where $t_0 \in (0, 1/2)$ and $\hat x_p(s, t)$ is the Weissman
-estimator applied to the partial series
-$X_{\ceil{ns}}, \dots, X_{\floor{nt}}$ using $\ceil{k(t - s)}$
-upper-order statistics. See Embrechts, Kl\"uppelberg and T.Mikosch
-\cite{Embrechts1997}, p.347.
-
-Hoga proposed the following test statistics:
-\[
-Q_1 = \sup_{t \in [t_0, 1 - t_0]}
-\left\{
-  {
-    \left[
-      t (1 - t) \ln\left(
-      \hat x_p(0, t)
-      \over
-      \hat x_p(t, 1)
-      \right)
-    \right]^2
-    \over
-    \int_{t_0}^t
-    \left[
-      s \ln \left(
-      \hat x_p(0, s)
-      \over
-      \hat x_p(0, t)
-      \right)
-    \right]^2 ds
+\subsection{Tests for a change in the extreme tail}\label{sec:Hoga}
+To the best of our knowledge, there do not exist statistical tests for comparing the 
+tail indices or scale parameters of two or more dependent \ts\ in  a direct way. {\red construct a test for this}
+We borrow a test from a recent paper by Hoga \cite{hoga:2016}. 
+This test has been developed for a different kind of problem. Given a strictly stationary \ts\ $X_1,\ldots,X_n$
+with a marginal \ds\ $F$ with right power-law tail, the goal is to test whether there is a structural break 
+of the {\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to zero. If the tail index
+{\em or} the scale parameter in a \ds\ of type \eqref{eq:3} change inside a sample, then it is likely
+that the extreme quantiles change as well. We will test for a change of tail index or scale parameter 
+in this indirect way.
+\par
+The null hypothesis of the test in \cite{hoga:2016} is that there is no change of the extreme quantiles $F^{-1}(1-p)$ 
+for $p=p_n\to 0$ in any subsample with indices
+$t\in (n\,t_0,n(1-t_0))$ where $t_0$ is a fixed number in $(0,0.5)$. Writing $\hat x_p(a,b)$ for an estimator of the extreme $(1-p)$-quantile 
+based on the subsample with indices $t\in (na,nb)$, the test statistic is given by
+\beam\label{eq:4}
+T_n = \sup_{s \in [t_0, 1 - t_0]}
+  \dfrac{  \big[s (1 - s) \log \big(\hat x_p(0, s)/\hat x_p(s, 1)\big)
+    \big]^2}{
+    \int_{t_0}^s\big[r \log \big( \hat x_p(0, r)/\hat x_p(0, s)
+      \big)
+    \big]^2 dr
     +
-    \int_{t}^{1 - t_0}
-    \left[
-      (1 - s) \ln \left(
+    \int_{s}^{1 - t_0}
+    \big[
+      (1 - r) \log \big(
+      \hat x_p(r, 1)/
       \hat x_p(s, 1)
-      \over
-      \hat x_p(t, 1)
-      \right)
-    \right]^2 ds
-  }
-\right\}
-\]
-Under the null hypothesis and suitable dependence conditions, Hoga
-showed that as $n \to \infty$,
-\begin{equation}
-  \label{eq:Q1_distr}
-  Q_1 \overset{d}{\to}
-  \sup_{t \in [t_0, 1 - t_0]} q_t
-\end{equation}
-where
-\[
-q_t = 
-  {
-    [W(t) - t W(1)]^2
-    \over
-    \int_{t_0}^t [W(s) - {s \over t}W(t)]^2 ds
-    +
-    \int_{t}^{1 - t_0}
-    \left[
-      W(1) - W(s) - {1 - s \over 1 - t}
-      (W(1) - W(t))
-      \right]^2 ds
-  }
-\]
-$W(t)$ with $t \in (0, 1)$ is a standard Brownian motion.
-
-When applied to our problem, we would like to test, first of all,
-whether there is a change of tail index in each of two 
-series and, if there is no change in either series, whether the two
-series have the same tail index. Obviously, for either question, if
-the tail index changes, the extreme quantile will change.
-For the first question, Hoga's test is directly applicable; for the
-second, we concatenate the two series before applying the
-test.
-
-Figure \ref{fig:Hoga_Single} shows the test statistics of
-stocks in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
-500 index. The null hypothesis here is that the extreme quantile
-remains the same througout the selected period of time.
+      \big)
+    \big]^2 dr}\nonumber\\
+\eeam
+Under the null hypothesis, $(T_n)$ converges to a complicated \fct al of \BM\ on $[0,1]$; the \asy\ quantiles need to be 
+evaluated by simulation.  
+\par
+When applied to our problem we would like to test 
+whether there is a change of the tail index {\em or} scale parameter in \eqref{eq:3} in each of the S\&P 500 
+series in the distinct sectors. We also  want to get some indication about a possible change of tail index or
+scale parameter from one series to another within a given sector. For this reason, we choose any pair of series
+within a sector and concatenate each of the paired series. Then we run the test on the concatenated series.
+Of course, despite the fact that we test changes of tail index or scale parameter 
+{\em in a very indirect way} -- there may be many other reasons for the change of extreme quantiles in a sample -- 
+we also concatenate two rather distinct series. Even if we assume that the two series come from related models
+(such as GARCH), the parameters of these models will in general not be the same. Moreover, the concatenation
+of two strictly stationary \ts\ is in general not strictly stationary. Therefore we have to be careful
+with interpretations of the results of the tests.
+\par 
+In Figure~\ref{fig:Hoga_Single} we show the values of the test statistic $T_n$ (horizontal bars)
+for $t_0=0.1$ and daily return series of stock in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
+500 index. The ``null hypothesis'' is that the tail index and scale parameter remain the
+same througout the selected period of time. For most stocks, the hypothesis cannot be rejected even at the 85\% level.
+This fact may be an indication that the \ds\ inside a series is rather homogeneous.
+Alternatively, it may show that the power of the test is very low. A possible reason for this suspicion is that the \con\ rate
+of $(T_n)$ to its limit is very slow, i.e., the \asy\ \ds\ is not representative for the \ds\ of $T_n$ for the chosen $n$; for some
+simulation evidence, see below.
+\par
+To check whether any two series share the same tail index and scale parameter we concatenate
+any two series and apply the aforementioned test on the concatenated series. For the ``Energy'' and the
+``Consumer Staples'' sector we summarize the results in Figure~\ref{fig:Hoga_pair}.
+These graphs show that the ``null
+hypothesis'' of an equal tail index and scale parameter is rejected for more pairs in the
+``Energy'' sector than it is for those in the ``Consumer Staples''
+sector. This suggests that lower tail indices of stocks in the
+``Energy'' sector are more spread out than those of the ``Consumer
+Staples'' sector. Also observe that while 3 stocks, say A, B and C, test in favor of
+the relations $\alpha_A = \alpha_B, \alpha_A \neq \alpha_C$, it often
+happens that another test on B and C is supportive of $\alpha_B = \alpha_C$. Again,
+this is due to the limited power of the test. Based
+on such results, one may guess that $\alpha_B$ lies between $\alpha_A$ and
+$\alpha_C$. The test is unable to recognize the smaller differences
+between $\alpha_A, \alpha_B$ on one hand and between $\alpha_B, \alpha_C$ on the other hand.
+
 \begin{figure}[htb!]
@@
   \end{minipage}
-  \caption{Hoga Test Statistics of stocks in the ``Energy'' and
+  \caption{Test statistic $T_n$ from \eqref{eq:4} for the  stocks in the ``Energy'' and
     ``Consumer Staples'' sectors of S\&P 500. The green, blue and red
-    lines are, respectively the 85\%, 90\% and 95\% quantiles of the
-    distribution of \eqref{eq:Q1_distr}}.
+    lines correspond to the 85\%, 90\% and 95\% quantiles of the limit \ds\ of $T_n$.  They are derived
+by simulations from the limit \ds .}
   \label{fig:Hoga_Single}
@@
 
-To test whether two series share the same tail index, we concatenate
-the two series and apply the Hoga test. For the ``Energy'' and the
-``Consumer Staples'' sector, we summarise the results in figure
-\ref{fig:Hoga_pair}.
 \begin{figure}[htb!]
@@
   \end{minipage}
-  \caption{Hoga Test Statistics of stocks in the ``Energy'' and
+  \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail index or scale parameter based on concatenated series in the ``Energy'' and
     ``Consumer Staples'' sectors of
-    S\&P 500. The green, blue and red colors represent, respectively,
-    exceedances of the 85\%, 90\% and 95\% quantiles of the distribution of
-    \eqref{eq:Q1_distr}.
-    Grey points stand for pairs for which the test statistic is below
-    the 85\% quantile;
+    S\&P 500. The green, blue and red colors represent, respectively, the cases when  
+    the 85\%, 90\% and 95\% quantiles of the limit distribution of $T_n$ is exceeded.
+    Grey points stand for pairs for which $T_n$ is below
+    the 85\% \asy\ quantile;
     black points represent pairs for which
-    computation of the test statistics fails for given precision
+    the computation of $T_n$ fails for given precision
     requirements and time limits.}
@@
 \end{figure}
-Clearly, these figures show that the null
-hypothesis of an equal tail index is rejected for more pairs in the
-``Energy'' sector than it is for those in the ``Consumer Staples''
-sector. This suggests that lower tail indices of stocks in the
-``energy'' sector are more spread out than are those of the ``Consumer
-Staples'' sector.
-
-Also observe that while 3 stocks, say A, B and C, test in favor of
-relations $\alpha_A = \alpha_B, \alpha_A \neq \alpha_C$, it often
-happens that another test on B and C is supportive of $\alpha_B =
-\alpha_C$ . This is due to the limited power of the test. Based
-on such results, one may guess $\alpha_B$ lies between $\alpha_A$ and
-$\alpha_C$. The test is unable to recognize the smaller differences
-between $\alpha_A, \alpha_B$ and between $\alpha_B, \alpha_C$.
-
-To have an idea about the power of Hoga's test, we run it on
-concatenated samples drawn from t-distributions with different degrees
-of freedom. Figure \ref{fig:t_sim_pair} shows the result.
+
+\par
+To get an idea about the power of the test we run it on a
+sample concatenated from two independent iid samples of the same size $n=1304$ as the S\&P 500 series. 
+Both pieces are $t$-distributed with distinct degrees
+of freedom. The results are shown in Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the larger the minimum
+tail index in the concatenated pair.
 \begin{figure}[htb!]
@@
   clip]{t_sim_pair.pdf}
-  \caption{Test Statistics of concatenated t-samples with different
-    degrees of freedom. Numbers on the axis are the degrees of freedom.}
+  \caption{Test  of concatenated $t$-samples with different
+    degrees of freedom. Numbers on the axes are the degrees of freedom in the subsamples. For an interpretation of the colored
+bullets, see Figure~\ref{fig:Hoga_pair}. The graph shows the limited power of the test. In particular, if
+both degrees of freedom are relatively large it loses the capablity of  distinguishing the \ds .}
   \label{fig:t_sim_pair}
 \end{figure}
-From the figure one can see that the power of the test decreases as
-the minimum tail index of a concatenated pair increases.
-
-Also of concern is the distribution of the test statistic under the
-null hypothesis. This distribution is only known asymptotically (given
-in equation \eqref{eq:Q1_distr}) when the sample size tends to
-infinity. The rate at which the finite-sample distribution tends to
-\eqref{eq:Q1_distr} is unclear. To find out more about this issue, we
-draw samples from the Student's t(3) and t(4) distributions and
-compare the empirical distribution of the test statistic $Q_1$ with the
-empirical distribution of \eqref{eq:Q1_distr}, which we also obtain by
-simulating the Brownian motion $W(t)$. The empirical density functions
-are shown in figure \ref{fig:Q1_distr}.
+\par
+A major problem in this test is the \asy\ distribution of the test statistic under the
+null hypothesis.  The rate at which the finite-sample distribution tends to its limit is not known.
+To find out about this problem we compared the \ds s of $T_n$ 
+with $n=1304$ for $t$-distributed $X_t$ with  3 and 4 degrees of freedom and the limit \ds\ of $T_n$.
+The estimated density functions are shown in Figure~\ref{fig:Q1_distr}.
 \begin{figure}[htb!]
@@
   \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
-  \caption{Comparison of the asymptotic distribution with empirical
-    distribution of $Q_1$ obtained from Student's t samples. Each t
-    sample has the same size as the S\&P sectors' data, namely 1304
-    observations.}
+  \caption{{\red Change $Q_1$ on $x$-axis} Comparison of the \asy\ distribution of the test statistic $T_n$ in \eqref{eq:4} under the null hypothesis 
+and the \ds\ of $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees of freedom.}
   \label{fig:Q1_distr}
 \end{figure}
-As seen in the figure, the asymptotic distribution \eqref{eq:Q1_distr}
-assigns significantly more mass to the tail than do the empirical
-distributions of size 1304. Table \ref{tab:HogaAsymptotic} lists a few
-quantiles of these distributions.
+As seen in the graph, the asymptotic distribution 
+assigns significantly more mass to the tail than the \ds s of $T_n$ do. For comparison, we list a few quantiles of these \ds s in
+Table~\ref{tab:HogaAsymptotic}, showing major differences between the \asy\ and finite-sample \ds s.
 \begin{table}[htb!]
   \centering
-  \begin{tabular}{c|c|c|c|c}
-    & \multicolumn{4}{c}{Quantiles} \\
+  \begin{tabular}{l|c|c|c|r}
+    & \multicolumn{4}{c}{Quantiles} \\[2mm]
     \hline
-    Distribution & 80\% & 85\% & 90\% & 95\% \\
+    Distribution of $X_t$& 80\% & 85\% & 90\% & 95\% \\
     \hline
     Asymptotic & 47.48 & 59.61 & 78.90 & 113.12 \\
-    t(3) empirical & 23.30 & 28.27 & 35.04 & 46.75 \\
-    t(4) empirical & 25.76 & 31.13 & 38.81 & 56.32
+    t(3)  & 23.30 & 28.27 & 35.04 & 46.75 \\
+    t(4)  & 25.76 & 31.13 & 38.81 & 56.32\\[2mm]
   \end{tabular}
-  \caption{Quantiles of the asymptotic and empirical test statistic
-    distributions}
+  \caption{Quantiles of the test statistic $T_n$ for $n=1304$ $t$-distributed samples with 3 and 4 degrees of freedom as well
+as the corresponding quantiles for the limiting \ds\ of $T_n$. In particular, there are huge differences between the three \ds s  
+for the higher quantiles. 
+    }
   \label{tab:HogaAsymptotic}
@@
 
-\section{Review of multivariate GARCH models:
-  same tail index or not}
-\label{sec:2}
-Many multivariate GARCH models imply a common tail index for all
-series under consideration, e.g. the Orthogonal GARCH model of
+\newpage
+\section{Some theoretical arguments for equality of tail indices}\setcounter{equation}{0}\label{sec:2}
+\subsection{Multivariate GARCH models whose components have  equal tail indices}\label{subsec:garch}
+Among the models for returns the generalized autoregressive conditionally heteroscedastic (GARCH) model
+is certainly most popular because it is parsimonious, captures various of the stylized facts of real return data
+and can also be modified in various directions to capture specific behavior of \ts\ such a asymmetry, skewness, long memory; 
+see for example Andersen et al. \cite{andersen:davis:kreiss:mikosch:2009}, Part 1, for a collection of results
+on GARCH-type models.
+The original  {\em univariate} GARCH model of Bollerslev \cite{bollerslev:1986} is a \sv\ model of the type $X_t=\sigma_t\,Z_t$, where 
+$(Z_t)$ is an iid mean-zero unit-variance \seq . In the simple case of a \garch\ the {\em squared volatility} satisfies the {\em \sre} 
+\beam\label{eq:5}
+\sigma_t^2= \alpha_0+\alpha_1 X_{t-1}^2+\beta_1\sigma_{t-1}^2=\alpha_0+(\alpha_1Z_{t-1}^2+\beta_1)\sigma_{t-1}^2\,,\quad t\in\bbz\,.
+\eeam
+Here $\alpha_0>0$, $\alpha_1,\beta_1$ are non-negative constants. For suitable choices of $\alpha_1,\beta_1$ the equation \eqref{eq:5}
+can be solved and the solution $(\sigma_t^2)$ constitutes a strictly stationary \seq , implying that $(X_t)$ 
+is strictly stationary itself. A remarkable property of the process $(\sigma_t)$ is that it has a power-law tail 
+of the form
+\beam\label{eq:6}
+\P(\sigma_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
+\eeam
+for some positive $c>0$ and a positive tail index $\alpha$ being the unique solution of the equation 
+$\E [(\alpha_1 Z_1+\beta_1)^{\alpha/2}]=1$
+provided the latter soluton exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
+follows by an application of the Kesten-Goldie theorem; see Kesten \cite{kesten:1973}, Goldie \cite{goldie:1991}, cf. 
+Buraczewski et al. \cite{buraczewski:damek:mikosch:2016}. The latter result ensures power-law tails  for
+the strictly stationary  solution $(Y_t)$ to the \sre\ 
+\beam\label{eq:7}
+Y_t= A_t\,Y_{t-1}+B_t\,,\qquad t\in\bbz\,,
+\eeam
+for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^\alpha]=1$.  
+In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2$ and $B_t=\alpha_0$ to achieve
+\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016})
+it follows that
+\beao
+\P(\pm X_t>x)\sim \E[(Z_t)_\pm^\alpha]\,\P(\sigma_t>x)\,,\qquad \xto\,, 
+\eeao
+implying power-laws for the right and left tails of $X_t$ caused by the power-law tail of $\sigma_t$.
+A GARCH process of the order $(p,q)$ can be embedded in a multivariate equation of the type \eqref{eq:7},
+where $(\bfA_t)$ are iid random matrices and $\bfB=\bfB_t$ is a constant vector. Again, the Kesten theory \cite{kesten:1973}
+applies, implying that the marginal and \fidi s of the GARCH process are \regvary\ with a positive index $\alpha$.
+We refrain from explaining the notion of multivariate \regvar\ which is needed in this context. For further details,
+see Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} where the Kesten theorem and \regvar\ of GARCH processes
+are explained in detail.
+\par
+There exist various extensions of the univariate GARCH
+model to the multivariate case. For the sake of argument, we stick here to the 
+{\em constant conditional correlation} (CCC) model of Bollerslev
+\cite{bollerslev:1990} and Jeantheau \cite{jeantheau:1998}, and we only consider a special bivariate case.
+It is the model 
+\beao\bfX_t=
+\left(\barr{l}X_{1,t}\\
+X_{2,t}\earr\right)= \left(\barr{ll}\sigma_{1,t}& 0\\
+0&\sigma_{2,t}\earr\right)\,\left(\barr{l}Z_{1,t}\\Z_{2,t}\earr\right)=\Sigma_t\,\bfZ_t\,,\qquad t\in\bbz\,.
+\eeao
+Thus both return components $X_{i,t}$ have the form of a univariate \sv\ model $X_{i,t}=\sigma_{i,t}Z_{i,t}$ 
+with non-negative volatility $\sigma_{i,t}$ and an iid bivariate noise \seq\ $(\bfZ_t)$ with zero mean and unite variance components.
+We also have the specification
+\beam\label{eq:8}
+\bfY_t=\left(\barr{l}\sigma^2_{1,t}  \\  
+\sigma^2_{2,t}\earr
+\right)
+&=& \left(
+\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)
++\left(\barr{cc}\alpha_{11} & \alpha_{12}  \\
+      \alpha_{21} & \alpha_{22}\earr \right)\, 
+\left(\barr{l}X_{1,t-1}^2  \\X_{2,t-1}^2   \earr\right)
+ + \left(\barr{cc}\beta_{11} & \beta_{12}  \\\beta_{21} & \beta_{22} \earr
+ \right)\,\left(\barr{c}\sigma^2_{1,t-1}  \\\sigma^2_{2,t-1}\earr
+  \right)\nonumber\\
+&=& \left(
+\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)+\left(\barr{cc}\alpha_{11}Z_{1,t-1}^2+\beta_{11}&\alpha_{12}Z_{2,t-1}^2+
+\beta_{12}\\
+\alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
+\earr\right)\,\left(\barr{l}\sigma_{1,t-1}^2\\\sigma_{2,t-1}^2\earr
+\right)\,.
+\eeam
+Writing
+\beao
+\bfB_t= \left(
+\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)\quad\mbox{and}\quad
+\bfA_t=\left(\barr{cc}\alpha_{11}Z_{1,t-1}^2+\beta_{11}&\alpha_{12}Z_{2,t-1}^2+
+\beta_{12}\\
+\alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
+\earr\right)\,,
+\eeao
+we see that we are again in the framework of a \sre\ 
+but this time for vector-valued $\bfB_t$ and matrix-valued $\bfA_t$:
+\beam\label{eq:jan6b}
+\bfY_t=\bfA_t\,\bfY_{t-1}+\bfB_t\,,\qquad t\in\bbz\,.
+\eeam
+Kesten \cite{kesten:1973} also provided the corresponding theory  
+for stationarity and tails in this case. \sta\ \cite{starica:1999}
+dealt with the corresponding problems for CCC-GARCH processes,
+making use of the theory in Kesten \cite{kesten:1973},
+Bougerol and Picard \cite{bougerol:picard:1992}
+and its
+specification to the tails of GARCH models 
+in Basrak et al.~\cite{basrak:davis:mikosch:2002}. \sta\ \cite{starica:1999} assumed the 
+Kesten conditions for the matrices $\bfA_t$. These conditions ensure that the product matrices $\bfA_1\cdots\bfA_n$ 
+have positive entries for sufficiently large $n$. Then Kesten's theory implies that
+all components of the vector $\bfX_t$ have power-law tails with the same index $\alpha$ and also
+that the \fidi s of the process $(\bfX_t)$ are \regvary\ with index $\alpha$. 
+\par
+Various GARCH modifications are derived by considering linear combinations of CCC-GARCH models.
+The property of multivariate \regvar\ of multivariate GARCH ensures that, after linear transformations,  
+the new process in all components has again power-law tails with the same index as the original GARCH process; see 
+Basrak et al.~\cite{basrak:davis:mikosch:2002}. 
+Models which are constructed in this way are
+the Orthogonal GARCH model of
 Alexander and Chibumba \cite{alexander1997multivariate}, its
-generalization GO-GARCH of van der Weide \cite{van2002go}, as
-well as the Full Factor GARCH model of Vrontos et al 
+generalization GO-GARCH by van der Weide \cite{van2002go},  the Full Factor GARCH model of Vrontos et al.
 \cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
@@
 combination of factors, and each of the factors is modeled as a GARCH
-process. As a result of this treatment, each and every series under
-consideration necessarily has the lowest of the tail indices of its
-contributing factors (cf. Mikosch \cite{Mikosch2013} as well
-as Andersen, Davis, Kreiss and Mikosch 
-\cite{andersen2009handbook}).
-
-In contrast, several other multivariate volatility models imply, in
-general, different tail indices for different series. For example, the
-Constant Conditional Correlation (CCC-) GARCH model of Bollerslev
-\cite{bollerslev1990modelling} specifies, for a $d$-dimensional
-series $\vec X_t$
-\begin{eqnarray}
-  \cov(X_{i,t}, X_{j,t}) &=& \sqrt{h_{i,t} h_{j,t}} \rho_{i,j}
-  \nonumber \\
-  h_{i,t} &=& \omega_i + \sum_{l=1}^q A_{i,l} X_{i, t-l}^2 +
-  \sum_{l=1}^p B_{i,l} h_{i, t-l} \label{eq:CCC-GARCH}
-\end{eqnarray}
-Consider the simple case when $p=q=1$. Assume
-$X_{i, t} = \sqrt h_{i,t} Z_{i,t}$, $Z_{i, t}$ are iid for
-all $t \in \mathbb Z$, and $\P(|Z_{i,t}| > u)/\P(\sqrt h_{i,t} > u) \to 0$ as
-$u \to \infty$. Then the tail index of the stationary distribution of
-$X_{i,t}$, call it $\kappa_i$, is given by
-\[
-\E (A_{i, 1} Z^2 + B_{i, 1})^{\kappa_i / 2} = 1
-\]
-where $Z \overset{d}{=} Z_{i,t}$. It is clear $\kappa_i \neq \kappa_j$
-except for special values of $A_i, B_i$ and $A_j, B_j$.
-
-As an extension to the CCC-GARCH model, Jeantheau 
-\cite{jeantheau1998strong} proposed his Extended CCC-GARCH by
-modifying \eqref{eq:CCC-GARCH} to
-\begin{equation}
-  \label{eq:ECCC-GARCH}
-  \vec h_t =
-  \sum_{l=1}^q \mathbf A_{l} (\vec X_{t-l} \odot \vec X_{t-l})
-  +
-  \sum_{l=1}^p \mathbf B_{l} \vec h_{t-l}
-  +
-  \vec \omega
-\end{equation}
-where $\mathbf A_l$ and $\mathbf B_l$ are non-diagonal $d \times d$ matrices
-and $\odot$ denotes element-wise multiplication. In the simple case $p=q=1$,
-equation \eqref{eq:ECCC-GARCH} reduces to
-\begin{eqnarray*}
-  h_{i,t} &=&
-  \sum_{j=1}^d A_{i,j} X_{j, t-1}^2
-  + \sum_{j=1}^d B_{i,j} h_{j, t-1}
-  + \omega_i
-\end{eqnarray*}
-or in vector form
-\begin{eqnarray}
-  \vec h_t &=& \left[
-    \mathbf A \text{diag}(Z_{1, t-1}^2, \dots, Z_{d, t-1}^2) + \mathbf B
-    \right] \vec h_{t-1} + \vec \omega
-  \label{eq:ECCC-GARCH2}
-\end{eqnarray}
-On appropriate conditions, \eqref{eq:ECCC-GARCH2} implies the
-components of $\vec h_t$ are jointly regularly varying with some index
-$\kappa$ as $t \to \infty$. This is a direct result of the Kesten-Goldie
-theorem (cf. Kesten \cite{Kesten1973}). Thus all series under
-consideration have the same tail index $\kappa$.
-
-In summary, whether or not a common tail index is shared by all
-series is a dividing issue for multivariate GARCH
-models. Nevertheless, economic theoretical arguments are supportive of
-a common tail index for all series. Therefore, it is important to
-gather statistical evidence from financial time series if the issue is
-to be settled definitely.
-
-\section{Generalized disappointment aversion}
-\label{sec:3}
-Routledge and Zin \cite{routledge2010generalized} define the GDA
-preferences over discrete states in their formula (2). A slight extension
-is to consider a continuum of possible states. Suppose the state variable
-$C$ is continuously supported on $(0, \infty)$ and has distribution
-function $G(\cdot)$. The utility function of an agent exhibiting GDA
-preferences reads
-\begin{equation}
-  u(v)= \E u(C) - b \int_{0}^{\delta v}
-  \left[ u(\delta v) - u(x) \right] dG(x)\label{11}%
-\end{equation}
-where $\delta > 0$, $v>0$ and $b > 0$. Equivalently
-\begin{equation}
-  \label{eq:xxie0}
-  u(v) = \E u(C) - b \E[u(\delta v) - u(C); C < \delta v]
-\end{equation}
+process; see Silvennoinen and Ter\"asvirt\"a \cite{silventeras:2009}.
+\par
+Not all choices of $\alpha$- and $\beta$-parameters in the model \eqref{eq:8} allow for an
+application of the Kesten theory. For example, assume that only the diagonal elements $\alpha_{ii}$ and $\beta_{ii}$ are positive.
+Then $\bfA_t$ is diagonal and, hence, the condition that $\bfA_1\cdots\bfA_n$ have positive entries for sufficiently large $n$ 
+cannot be satisfied. In the latter situation, both $(X_{1,t})$ and $(X_{2,t})$ are univariate \garch\ processes. Assuming the
+conditions of the univariate Kesten-Goldie theorem for each component process, $(X_{1,t})$ and $(X_{2,t})$ have power-law tails
+with indices $\kappa_1$ and $\kappa_2$, respecticely,  given by the solutions to the equations 
+$\E [(\alpha_{ii} Z_{i,t}^2+\beta_{ii})^{\kappa_i}]=1$, $i=1,2$.
+In this model, one can introduce dependence between the two component series $(X_{1,t})$ and $(X_{2,t})$ by assuming dependence between the noise variables $Z_{1,t}$ and $Z_{2,t}$.
+Another situation when the Kesten theory fails
+appears when $\bfA_t$ is an upper or lower triangle matrix: then the products  $\bfA_1\cdots\bfA_n$ are always of the same triangular type.
+Similar remarks apply when one considers a CCC model in general dimension. Of course, one may argue that the latter models
+are not natural since they do not allow for a linear relationship between all squared volatilities
+on a given day.
+
+\subsection{An utility based argument for equal tail indices} \setcounter{equation}{0}\label{sec:3}
+In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
+We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
+Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
+it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
+Following Routledge and Zin \cite{routledge2010generalized},
+the utility of an agent with GDA preferences is given by
+\beao% \label{eq:xxie0}
+  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
+  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
+\eeao
+where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red meaning of these constants?} 
  Here $v$ can be thought of as the certainty payoff equivalent to the
-risky payoff $C$. Note that when $b=0$, 
-preferences are expected utility; when $\delta=1$ and $b>0$,
-preferences follow Gul's original disappointment aversion.
-An agent guided by this utility function will seek to maximize $u(v)$.
-
-Routledge and Zin \cite{routledge2010generalized} assumed
-a power-law utility function of $C$:
-\begin{equation}
-  \label{eq:power_utility}
-  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
-\end{equation}
-where $\gamma > 0$. For our discussion it suffices that $u(\cdot)$ is
-increasing and concave. Before proceeding any further, it is worth
-stating the following simple observation:
-\begin{lemma} \label{lemma:I}
-  Assume distribution function $F(x; \theta)$ parameterized by
-  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
-  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
-  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
-  $(a, b)$, $X \sim F$, $\text{supp}(X) = (a,b)$,
-  $\E h(X) < \infty$ and
-  \begin{eqnarray*}
-    \left| \frac{\pd f(x, \theta)}{\pd \theta} \right| < \infty
-    \quad \forall x \in (a, b)
-  \end{eqnarray*}
-  \begin{enumerate}
-  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
-    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
-    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
-    \[
-    \frac{\pd \E h(X)}{\pd \theta} > 0
-    \]
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
-    \[
-    \frac{\pd \E h(X)}{\pd \theta} > 0
-    \]
-  \end{enumerate}
-\end{lemma}
-\begin{remark}
-  \label{remark:I}
-  Two other cases follow trivially from lemma \ref{lemma:I}:
-  \begin{enumerate}
-  \item If $h(\cdot)$ is increasing and $\frac{\pd f}{\pd \theta}$
-    satisfies the
-    same conditions of the 1st case of lemma \ref{lemma:I},
-    $\frac{\pd \E h(X)}{\pd \theta} < 0$. This immediately follows
-    from applying 1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
-  \item By the same argument, if $h(\cdot)$ is decreasing and
-    ${\pd f \over \pd \theta}$ satisfies the same conditions of the 2nd case of
-    lemma \ref{lemma:I}, ${\pd \E h(X) \over \pd \theta} < 0$.
-  \end{enumerate}
-\end{remark}
-Now for simplicity we assume the investor initially has one unit of
-wealth. Let $r > 0$ be the return on the risk free state bond and $X$
-be the return on equity. Portfolio weights are $\left(1-\phi\right)$
-and $\phi$ respectively. Then
-\begin{equation}
-  \label{eq:xxie1}
-  C(X) = (1 - \phi) e^r + \phi e^X
-\end{equation}
-Equation \eqref{eq:xxie0} can be re-arranged as
-\begin{eqnarray}
-u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
-&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
-\end{eqnarray}
-where
-\begin{eqnarray*}
-  q = q(r, \phi) &:=& \ln \left( {
-      \delta v - (1 - \phi) e^r
-      \over
-      \phi
-    } \right) = \ln\left(
-    e^r + {\delta v - e^r \over \phi}
-  \right)
-\end{eqnarray*}
-Note $C < \delta v$ implies $X < q$. Naturally, if an agent invests in
+risky payoff $C$. If $b=0$, 
+preferences are the classical expected utility. If $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
+disappointment aversion which were generalized by Routledge and Zin \cite{routledge2010generalized}.
+An agent guided by the utility function $u$ will seek to maximize the \fct al $\wt u$.
+Routledge and Zin assumed
+a power-law utility function 
+\beam\label{eq:u}
+  u(x)=\frac{1}{-\xi}\,x^{-\xi}\,,\quad \xi>-1\,. 
+\eeam
+For the sake of argument, we assume that an investor initially has one unit of
+wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
+bond with interest rate $r>0$ and $\phi$ units in a risky asset with return $X$ over one time unit, i.e.,
+\beao%\label{eq:xxie1}
+  C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
+\eeao
+Then we have 
+\beao
+\wt u(\phi) &=& \E [u(C)] + b\, \E \big[u(C)\1(C < \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
+\eeao
+where $F_X$ is the \df\ of $X$ with density $f_X$, and
+\beao
+  q = %q(r, \phi) &:=& \ln \left( {
+      %\delta v - (1 - \phi) e^r
+      %\over
+      %\phi
+    % \right) 
+\log\Big(
+    \ex^r + {\delta \,v - \ex^r \over \phi}
+  \Big)\,.
+\eeao
+Note that $C < \delta v$ \fif\ $X < q$. 
+\par
+Naturally, if an agent invests in
 a risky asset instead of a riskless bond, he expects to obtain a
-higher return from the risky asset than he is guarranteed from the
-riskless bond. In our notation, this means $\delta v > e^r$, which
-implies $q(r, \phi) > r > 0$. The right-hand-side of
-\eqref{eq:functional} is a functional of $\phi$ and of the distribution
-function of $X$ via \eqref{eq:xxie1}. For convenience we denote it
-$\mathcal U(\cdot, \cdot)$, i.e.
-\begin{equation}
-  \label{eq:U_functional}
-  \mathcal U(F, \phi)
-  = 
-  \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
-\end{equation}
-In the rest of this section, we use $F(\cdot)$ to denote the
-distribution function of $X$ and use $f(\cdot)$ to denote its density
-function. Let
-\begin{equation}
-  \label{eq:G_functional}
-  G(F) = \max_{0 < \phi \leq 1} \mathcal U(F, \phi)  
-\end{equation}
-
-\subsection{Pareto-distributed equity return}
-In this section, we look into the situation when $X$ has a shifted
-Pareto distribution. Assume
-\begin{equation}
-  \label{eq:X_distr}
-  F(x) = \left\{
+higher (on average) return from the risky asset than he is guaranteed from the
+riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
+$q > r $. For given $b,\delta,v$, $\wt u$ is a functional of $\phi$ and $F_X$,
+$
+\wt u=\wt u (F_X, \phi)\,,
+$
+and we assume that
+\beao
+\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
+\eeao
+exists.
+
+\subsection{Pareto-distributed returns}
+We consider the case when $X$ has a two-sided Pareto \ds\ given by
+\beam\label{eq:pareto}
+  F_X(x) = \left\{
   \begin{array}{ll}
@@
     {K \over K - x}
-    \right)^\alpha & x \leq 0 \\
+    \right)^\alpha & x \leq 0 \,,\\
     1 - (1 - p) \left(
     {K' \over K' + x}
-    \right)^\beta & x > 0
+    \right)^\beta & x > 0\,,
   \end{array}
   \right.
-\end{equation}
-Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
-We have
-\begin{eqnarray}
-  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
+\eeam
+where 
+$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
+\beam
+ \wt u(F_X,\phi)\nonumber
   &=&
-  \alpha K^\alpha  p
+  \alpha\, K^\alpha\,  p\,
   \int_{-\infty}^0
-  u\left[ (1 - \phi) e^r + \phi e^x \right]
-  {1 + b \over (K - x)^{\alpha + 1}} dx
-  \nonumber \\
-  &&
-  \beta K'^\beta (1 - p)
+  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
+  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
+    &&+
+  \beta \,(K')^\beta\, (1 - p)
   \int_{0}^\infty
-  u\left[ (1 - \phi) e^r + \phi e^x \right]
-  {1 + b \1_{\{x < q\}} \over (K' + x)^{\beta + 1}} dx 
-  - b u(\delta v) F(q)
-  % &&
-  % - b u(\delta v) \left[
-  %   1 - (1 - p) {
-  %     K'^\beta
-  %     \over
-  %     (K' + q)^\beta
-  %   }
-  % \right]
-  \label{eq:xxie1.0}
-\end{eqnarray}
-\begin{theorem}
-  \label{thrm:I}
-  Suppose functional $G(\cdot)$ is given as \eqref{eq:G_functional},
-  function $u(\cdot)$ is increasing and differentiable, then
-  \[
-  \frac{\pd G(F)}{\pd \alpha} > 0,
-  \quad
-  \frac{\pd G(F)}{\pd K} < 0
-  \]
+  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
+  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
+  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
+\eeam
+{\red Is this worth a theorem?}
+\begin{theorem}\label{thrm:I}
+ Assume the two-sided Pareto model \eqref{eq:pp} and that the utility \fct\ $u$ is increasing and differentiable. Then
+$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
 \end{theorem}
-The proof of theorem \ref{thrm:I} is given in secton
-\ref{sec:thrmI_proof}.
-Since $u(v)$ given in \eqref{eq:xxie1.0} as a
-functional of the distribution function of $X$ is increasing with the
-lower tail index $\alpha$ and decreasing with $K$, there is a curve of
-equal preference on the plane of $(\alpha, K)$. Moving along this
+The proof is given in Appendix~\ref{sec:thrmI_proof}. 
+\par
+We illustrate the results of Theorem~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
+\fct\ \eqref{eq:u}.
+For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
+quantity is denoted by $I(\alpha, K)$.
+\par
+We conclude that
+$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
+equal preference on the $(\alpha, K)$-plane. Moving along this
 curve in the direction of increasing $\alpha$, one expects the values
-of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
-should appear positively correlated. This is indeed the case in some
-real stock return data, e.g. the ``Energy'', ``Consumer Staples'' and
-``Information Technology'' sectors of S\&P 500 companies. This is
-illustrated in figure \ref{fig:sectors_parameters}.
-The confidence bands of the Hill estimates of $\alpha$ of these
-sectors are shown in figure~\ref{fig:1}.
-
-By numerically integration and optimization with respect to $\phi$,
-the value of $\hat\phi$ can be obtained for given $u(\cdot)$ and given
-values of $K, K', \alpha, \beta$. Figures \ref{fig:phi_hat_pareto}
-shows the values of $\hat\phi$ for $u(C) = -\frac{C^{-\xi}}{\xi}$ with
-$\xi = 1/2, 4$. $\beta=\alpha$ and $K'=K$ in both figures.
+of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
+should appear positively dependent. This is indeed the case for some
+real return data, e.g. the ``Energy'', ``Consumer Staples'' and
+``Information Technology'' sectors of the S\&P 500 index; see 
+Figure \ref{fig:sectors_parameters}. {\red how can I see positive dependence?}
+\par
+The value of $\hat\phi$ can be calculated
+by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
+This is shown in 
+Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
 \begin{figure}
@@
   \end{minipage}
-  \caption{$\hat\phi$, the optimal equity allocation when the utility
-    function is $u(C) = -{C^{-\xi} \over \xi}$ for $\xi = 1/2$
-    (left). and $\xi = 1/2$ (right).
+  \caption{The graphs show $\hat\phi$, the optimal equity allocation
+as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
+the utility $u$ from \eqref{eq:u}
+    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$
+    (left) and $\xi = 4$ (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other figures, and use the same
+structure of the formulas.}
   }
@@
 \end{minipage}\hfill
-\begin{minipage}{0.42\textwidth}
-  As shown to the left, the power utility function with a small
-  $\xi$ grows slower and saturates later than does one with a large
-  $\xi$. It represents an agent that is more tolerant of low
-  consumption and that seeks wealth more aggressively. In other
-  words, he is less risk-averse than is one with a higher $\xi$.
-  Such an agent will therefore invest more heavily on the equity as
-  is shown in figures \ref{fig:phi_hat_pareto}.
+\begin{minipage}{0.42\textwidth}{\red change formulas in graph and $C$ to $x$.}
+  In this graph we show two power-utility functions $u$. We see that $u(x)$ grows slower
+  and saturates later for small $\xi$ than for large $\xi$. The latter case
+  represents an agent who is more tolerant about low
+  consumption and seeks wealth more aggressively. In other
+  words, he is less risk-averse than one with a larger $\xi$.
+  Such an agent will therefore invest more heavily in equity, as shown in Figure~\ref{fig:phi_hat_pareto}.
 \end{minipage}
-Certainly, it rarely happens in practice that the lower- and
-upper-tail indices of an equity are equal. Thus, to verify our result
-that $G(F)$ is increasing with $\alpha$ and decreasing with $K$, we
-need to verify that the 1st integral of \eqref{eq:xxie1.0} with $\phi$
-replaced by $\hat\phi$, i.e.
-\[
-I(\alpha, K) = 
-  \alpha K^\alpha  p
-  \int_{-\infty}^0
-  u\left[ (1 - \hat\phi(\alpha, K)) e^r + \hat\phi(\alpha, K) e^x \right]
-  {1 + b \over (K - x)^{\alpha + 1}} dx
-\]
-is increasing with $\alpha$ and decreasing with
-$K$. Figure~\ref{fig:preference_pareto}
-shows the values of the above integral when $\alpha$ and $K$ take
-a range of values.
+
 \begin{figure}[htb!]
@@
   \end{minipage}
-  \caption{$I(\alpha, K)$ when the utility function is
-    $u(C) = -{C^{-\xi} \over \xi}$ for $\xi = 1/2, 4$.
+  \caption{The quantities $I(\alpha, K)$ for the power-utility \fct\ with $\xi=1/2$ (left) and $\xi=4$ (right).
+The graphs clearly show that  $I(\alpha, K)$ increases with $\alpha$ and decreases with $K$.
   }
@@
   p &=& \P(X < 0) = \lim_{n \to \infty}
-  {1 \over n} \sum_{t=1}^n \1_{\{X_t < 0\}}
+  {1 \over n} \sum_{t=1}^n \1{X_t < 0}
 \end{eqnarray*}
@@
   \sqrt n \left[
-    {1 \over n}\sum_{t=1}^n \1_{\{X_t < 0\}} - p
+    {1 \over n}\sum_{t=1}^n \1{X_t < 0} - p
   \right] \to N(0, p - p^2))
@@
     S\&P 500 index. The points are the estimated values; the bars are
-    the $2\sigma$ confidence intervals.
+    the $2\sigma$ confidence intervals. The confidence bands of the
+    corresponding Hill estimates $\hat \alpha$ of these sectors are
+    shown in figure~\ref{fig:1}.
   }

### 2017-04-12 papers/FX/draft.tex
Revised the FX manuscript.

@@
 %\usepackage{enumerate}
-%\usepackage[bookmarks=true]{hyperref}
-%\usepackage{bookmark}
+\usepackage[bookmarks=true]{hyperref}
+%% \usepackage{bookmark}
 \usepackage{graphicx}
@@
 for some positive $c>0$ and a positive tail index $\alpha$ being the unique solution of the equation 
-$\E [(\alpha_1 Z_1+\beta_1)^{\alpha/2}]=1$
+$\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$
 provided the latter soluton exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
@@
 \eeao
-where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red meaning of these constants?} 
+where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red
+  meaning of these constants?}
+{\blue 
  Here $v$ can be thought of as the certainty payoff equivalent to the
-risky payoff $C$. If $b=0$, 
-preferences are the classical expected utility. If $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
-disappointment aversion which were generalized by Routledge and Zin \cite{routledge2010generalized}.
-An agent guided by the utility function $u$ will seek to maximize the \fct al $\wt u$.
-Routledge and Zin assumed
-a power-law utility function 
+ risky payoff $C$; $\delta$ tunes the threshold of {\em
+   disappointment} in proportion to $v$; $b$ determines the extra
+ weight given to the expected return of $C$ when $C$ is below the
+ disappoinment threshold $\delta v$.
+ If $b=0$, preferences are the classical expected utility. If
+ $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
+disappointment aversion which were generalized by Routledge and Zin
+\cite{routledge2010generalized}.
+}
+An agent guided by the utility function $u$ will seek to maximize the
+\fct al $\wt u$.
+Routledge and Zin assumed a power-law utility function 
 \beam\label{eq:u}
-  u(x)=\frac{1}{-\xi}\,x^{-\xi}\,,\quad \xi>-1\,. 
+u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
 \eeam
-For the sake of argument, we assume that an investor initially has one unit of
-wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
-bond with interest rate $r>0$ and $\phi$ units in a risky asset with return $X$ over one time unit, i.e.,
-\beao%\label{eq:xxie1}
+In practice it is more often to observe $\gamma > 1$ {\blue Let's
+consult Casper for references.} Therefore we adopt a more convenient
+notation
+\beao
+u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
+\eeao
+
+For the sake of argument, we assume that an investor initially has one
+unit of wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
+bond with interest rate $r>0$ and $\phi$ units in a risky asset with
+return $X$ over one time unit, i.e.,
+\beam\label{eq:xxie1}
   C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
-\eeao
+\eeam
 Then we have 
@@
 $\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
-\beam
+\beam\label{eq:xxie1.0}
  \wt u(F_X,\phi)\nonumber
@@
     {\red Is this worth a theorem?}
-\begin{theorem}\label{thrm:I}
+    {\blue No}
+\begin{remark}\label{thrm:I}
  Assume the two-sided Pareto model \eqref{eq:pp} and that the utility \fct\ $u$ is increasing and differentiable. Then
 $\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
-\end{theorem}
+\end{remark}
 The proof is given in Appendix~\ref{sec:thrmI_proof}. 
 \par
-We illustrate the results of Theorem~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
+We illustrate the results of remark~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
 \fct\ \eqref{eq:u}.
@@
 
-\subsection{Hybrid fitting of the shifted Pareto distribution}
+\subsection{Fitting the shifted Pareto distribution}
 \label{sec:hybrid_estimation}
 We adopt a hybrid procedure to estimate the parameters of the
-distribution function assumed in \eqref{eq:X_distr}: First we
+distribution function assumed in \eqref{eq:pareto}: First we
 obtain a Hill estimate of the lower-tail indices of an
-equity return and fix $\alpha$ in \eqref{eq:X_distr}
+equity return and fix $\alpha$ in \eqref{eq:pareto}
 to its Hill estimate. Then we estimate $K$ by maximizing
@@
 conditions on the maximum likelihood estimator are easily checked for
-the distribution \eqref{eq:X_distr}.
+the distribution \eqref{eq:pareto}.
 The density function of $X$ conditional on $X < 0$ is
@@
 $u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
-\eqref{eq:power_utility} as $\gamma \to 0$.
+\eqref{eq:u} as $\gamma \to 0$.
 
@@
 \appendix
-\section{Proof of lemma \ref{lemma:I}}
+\section{A monontonicity lemma}
+\begin{lemma} \label{lemma:I}
+  Assume distribution function $F(x; \theta)$ parameterized by
+  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
+  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
+  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
+  $(a, b)$, $X \sim F$, $\text{supp}(X) = (a,b)$,
+  $\E h(X) < \infty$ and
+  \begin{eqnarray*}
+    \left| \frac{\pd f(x, \theta)}{\pd \theta} \right| < \infty
+    \quad \forall x \in (a, b)
+  \end{eqnarray*}
+  \begin{enumerate}
+  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
+    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
+    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
+    \[
+    \frac{\pd \E h(X)}{\pd \theta} > 0
+    \]
+  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
+    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
+    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
+    \[
+    \frac{\pd \E h(X)}{\pd \theta} > 0
+    \]
+  \end{enumerate}
+\end{lemma}
+\begin{remark}
+  \label{remark:I}
+  Two other cases follow trivially from lemma \ref{lemma:I}:
+  \begin{enumerate}
+  \item If $h(\cdot)$ is increasing and $\frac{\pd f}{\pd \theta}$
+    satisfies the
+    same conditions of the 1st case of lemma \ref{lemma:I},
+    $\frac{\pd \E h(X)}{\pd \theta} < 0$. This immediately follows
+    from applying 1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
+  \item By the same argument, if $h(\cdot)$ is decreasing and
+    ${\pd f \over \pd \theta}$ satisfies the same conditions of the
+    2nd case of lemma \ref{lemma:I}, ${\pd \E h(X) \over \pd \theta} < 0$.
+  \end{enumerate}
+\end{remark}
+
 \begin{proof}
@@
     &=& \int_a^b h(x) {\pd f \over \pd \theta}(x, \theta) dx \\
-    &=& \underbrace{\int_a^{x_0} h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_1}
-    + \underbrace{\int_{x_0}^b h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_2}
+    &=& \underbrace{\int_a^{x_0} h(x) {\pd f \over \pd \theta}(x,
+      \theta) dx}_{I_1}
+    + \underbrace{\int_{x_0}^b h(x) {\pd f \over \pd \theta}(x,
+      \theta) dx}_{I_2}
   \end{eqnarray*}
   \begin{enumerate}
-  \item When $h(x)$ is decreasing on $(a, b)$ and ${\pd f \over \pd \theta}(x,
-    \theta) > 0$ on $(a, x_0)$
+  \item When $h(x)$ is decreasing on $(a, b)$ and ${\pd f \over \pd
+    \theta}(x, \theta) > 0$ on $(a, x_0)$
     \[
@@
     \]
-    Similarly, because ${\pd f \over \pd \theta}(x, \theta) < 0$ for $x \in (x_0, b)$ and
-    $h(x)$ is decreasing, we have
+    Similarly, because ${\pd f \over \pd \theta}(x, \theta) < 0$ for
+    $x \in (x_0, b)$ and $h(x)$ is decreasing, we have
     \begin{eqnarray*}

### 2017-04-12 papers/FX/draft.tex
Revised the FX manuscript.

@@
 However, it is common to present the tail index estimators
-for positive data. Therefore we will multiply the losses $X_t$ by minus one, swapping the negative with the positive values.
+for positive data. Therefore we will multiply the losses $X_t$ by
+minus one, swapping the negative with the positive values.
 For simplicity, we also suppress subscripts in the notation:
@@
 \beao
-\hat K_k = {k \over n} X_{(n-k+1)}\,.
+\hat K_k = \left({k \over n}\right)^{1/\hat\alpha} X_{(n-k)}\,.
 \eeao
-Estimates of $K$ {\red or $K^\alpha$?} of the ``Energy'', ``Consumer Staples'' and
+Using the asymptotic normality property of upper order statistics
+(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, theorem 2.2.1), one can show
+\begin{eqnarray*}
+  \sqrt k (\hat K_k - K) &\overset{d}{\to}& N\left(
+    0, \left( {K \over \alpha}  \right)^2
+  \right) \\
+  \sqrt k (\hat K_k^\alpha - K^\alpha) &\overset{d}{\to}& N\left(
+    0, \left( {K^{2\alpha}}  \right)^2
+  \right)
+\end{eqnarray*}
+where the tail index $\alpha$ is regarded as known. From the above
+asymptotic normality property, confidence bands of the estimate of $K$
+and $K^\alpha$ can be constructed. 
+Estimates of $K$ of the ``Energy'', ``Consumer Staples'' and
 ``Information Technology'' sectors of the S\&P 500 index are computed
-using this method. {\blue The results are shown in Figure~\ref{fig:sectors_parameters} {\red ???} along with estimates obtained
-using the hybrid method described in section
-\ref{sec:hybrid_estimation}.} {\red Figures of the Hill estimates of $K$ or $K^\alpha$ with Hill confidence bands. 
-Discuss confidence bands and what you see in the data.}
-
-%% Hence it is informative to test
-%% the hypothesis that the tail indices of these return series are equal.
-%% Hoga \cite{hoga:2016} devised tests suitable this purpose. We adopt his
-%% approach but use our own implementation. The next section has more
-%% details.
-
-%% \subsection{Hill estimates of extreme quantiles}
+using this method. The results are shown in
+Figure~\ref{fig:sectors_parameters}.
+One can see that $K$ generally takes a rather small value. For the
+more volatile sectors of ``Energy'' and ``Information Technology'',
+the average value of $K$ is around 0.01, while for the more stable
+sector of ``Consumer Staples'', the average value is around 0.005. Due
+to the smallness of $K$, mild variations of $\alpha$ would lead to
+huge variations of $K^\alpha$, as shown in the 2nd column of
+figure~\ref{fig:sectors_parameters}.
+
+Secondly, one can see that the values of $\alpha$ and $K$ are appear
+positively correlated. This prompts Pareto tails of the return
+distributions. We expain more about this in
+section~\ref{sec:pareto_tail}.
+
+\subsection{A test of equal tail indices based on Hill estimator}
+From \eqref{eq:2} one can deduce
+\begin{equation}
+  \label{eq:x1}
+  \sqrt k (\hat \alpha_A - \hat \alpha_B) \overset{d}{\to} N(0, \alpha_A^2 - \alpha_B^2)  
+\end{equation}
+where $\hat \alpha_A$ and $\hat \alpha_B$ are Hill estimates of
+$\alpha_A$ and $\alpha_B$, the tail indices of two series $A$ and
+$B$. The weak convergence of \eqref{eq:x1} allows one to construct a
+test of the equality of $\alpha_A$ and $\alpha_B$.
 
@@
 
-\subsection{Pareto-distributed returns}
+\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
 We consider the case when $X$ has a two-sided Pareto \ds\ given by
@@
 ``Information Technology'' sectors of the S\&P 500 index; see 
-Figure \ref{fig:sectors_parameters}. {\red how can I see positive dependence?}
+Figure \ref{fig:sectors_parameters}.
 \par
@@
   \caption{The graphs show $\hat\phi$, the optimal equity allocation
-as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
+    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
+    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
     the utility $u$ from \eqref{eq:u}
-    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$
-    (left) and $\xi = 4$ (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other figures, and use the same
-structure of the formulas.}
+    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$ (left) and $\xi = 4$
+    (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other
+      figures, and use the same structure of the formulas.}
   }
@@
     \includegraphics[width=\textwidth]
-    {Energy_alpha_K_ci.pdf}
+    {Energy_K.pdf}
   \end{minipage}\hfill
@@
     \includegraphics[width=\textwidth]
-    {Energy_p.pdf}
+    {Energy_scale.pdf}
   \end{minipage}
@@
     \includegraphics[width=\textwidth]
-    {Consumer_Staples_alpha_K_ci.pdf}
+    {Consumer_Staples_K.pdf}
   \end{minipage}\hfill
@@
     \includegraphics[width=\textwidth]
-    {Consumer_Staples_p.pdf}
+    {Consumer_Staples_scale.pdf}
   \end{minipage}
@@
     \includegraphics[width=\textwidth]
-    {Information_Technology_alpha_K_ci.pdf}
+    {Information_Technology_K.pdf}
   \end{minipage}\hfill
@@
     \includegraphics[width=\textwidth]
-    {Information_Technology_p.pdf}
+    {Information_Technology_scale.pdf}
   \end{minipage}
-  \caption{\small Estimates of $p$ and $K$ of stocks in sectors of the
-    S\&P 500 index. The points are the estimated values; the bars are
+  \caption{\small Estimates of $K$ and $K^\alpha$ of stocks in sectors
+    of the S\&P 500 index. The points are the estimated values; the bars are
     the $2\sigma$ confidence intervals. The confidence bands of the
     corresponding Hill estimates $\hat \alpha$ of these sectors are
-    shown in figure~\ref{fig:1}.
+    shown in figure~\ref{fig:1}. Note $\log_{10}(K^\alpha)$ is shown
+    instead of $K^\alpha$.
   }
@@
 
-\section{Proof of Theorem \ref{thrm:I}}
+\section{Proof of Remark \ref{thrm:I}}
 \label{sec:thrmI_proof}

### 2017-04-12 r/libxxie.r

@@ -91,7 +91,9 @@ inferInnovations <- function(X) {
 
 
 getInterpolatedPrices <- function(day1, day2, assetSet="",
-                                   tables="", suffix="") {
+                                   tables="", suffix="")
+{
+    require(RMySQL);
     database = dbConnect(MySQL(), user='sinbaski', password='q1w2e3r4',
         dbname='avanza', host="localhost");
     if (nchar(assetSet) > 0) {

### 2017-04-12 r/myLogLik.r

@@ -3,43 +3,43 @@ require(MASS);
 require(gplots);
 source("libxxie.r")
 #Energy
-## tables <- c (
-##     "APA",
-##     "APC",
-##     "BHI",
-##     "CHK",
-##     "COG",
-##     "COP",
-##     "DO",
-##     "DVN",
-##     "EOG",
-##     "EQT",
-##     "FTI",
-##     "HAL",
-##     "HES",
-##     "HP",
-##     "KMI",
-##     "MPC",
-##     "MRO",
-##     "MUR",
-##     "NBL",
-##     "NFX",
-##     "NOV",
-##     "OKE",
-##     "OXY",
-##     "PSX",
-##     "PXD",
-##     "RIG",
-##     "RRC",
-##     "SE",
-##     "SLB",
-##     "SWN",
-##     "TSO",
-##     "VLO",
-##     "WMB",
-##     "XEC",
-##     "XOM"
-## );
+tables <- c (
+    "APA",
+    "APC",
+    "BHI",
+    "CHK",
+    "COG",
+    "COP",
+    "DO",
+    "DVN",
+    "EOG",
+    "EQT",
+    "FTI",
+    "HAL",
+    "HES",
+    "HP",
+    "KMI",
+    "MPC",
+    "MRO",
+    "MUR",
+    "NBL",
+    "NFX",
+    "NOV",
+    "OKE",
+    "OXY",
+    "PSX",
+    "PXD",
+    "RIG",
+    "RRC",
+    "SE",
+    "SLB",
+    "SWN",
+    "TSO",
+    "VLO",
+    "WMB",
+    "XEC",
+    "XOM"
+);
 
 ## Consumer staples
 ## tables <- c(
@@ -80,50 +80,50 @@ source("libxxie.r")
 ## );
 
 ## Information Technology
-tables <- c(
-    "ADBE",
-    "ADI",
-    "ADP",
-    "ADSK",
-    "AKAM",
-    "AMAT",
-    "CA",
-    "CSCO",
-    "CTSH",
-    "CTXS",
-    "EA",
-    "EBAY",
-    "FFIV",
-    "FISV",
-    "HPQ",
-    "HRS",
-    "IBM",
-    "INTC",
-    "INTU",
-    "JNPR",
-    "KLAC",
-    "LLTC",
-    "LRCX",
-    "MCHP",
-    "MSFT",
-    "MSI",
-    "MU",
-    "NTAP",
-    "NVDA",
-    "ORCL",
[…]

### 2017-04-12 r/temp.r

@@ -1,6 +1,8 @@
 rm(list=ls());
 source("libxxie.r");
 library(parallel);
+library(doParallel);
+library(foreach);
 
 #Energy
 tables <- c (
@@ -137,12 +139,52 @@ assets <- data$assets;
 
 X <- diff(log(price));
 n <- dim(price)[1];
-A <- lapply(1:dim(X)[2],
-            FUN=function(i) {
-                scaleEstimate(-X[, i], max(floor(dim(X)[1]*0.03), 60))
-            });
-A <- unlist(A);
-plot(1:dim(X)[2], A)
+
+cl <- makeCluster(detectCores());
+registerDoParallel(cl);
+k <- max(floor(dim(X)[1]*0.03), 50);
+
+params <- foreach (i=1:dim(X)[2], .combine=rbind) %dopar% {
+    R <- sort(-X[, i], decreasing=TRUE);
+    alpha <- hillEstimate(-X[, i], k);
+    K <- R[k+1] * (k/n)^(1/alpha);
+    c(alpha, K)
+}
+
+pdf(file="../papers/FX/Consumer_Staples_K.pdf");
+sd <- params[, 2]/params[, 1]/sqrt(k);
+plotCI(params[, 1], params[, 2],
+       uiw=0, liw=0,
+       ui=params[, 2] + 2*sd,
+       li=params[, 2] - 2*sd,
+       barcol="#000000",
+       main="Consumer Staples",
+       col="red",
+##       pch=16,
+       lwd=1,
+       xlab=expression(alpha),
+       ylab=expression(K));
+dev.off();
+
+pdf(file="../papers/FX/Consumer_Staples_scale.pdf");
+values <- params[, 2]^params[, 1];
+sd <- values/sqrt(k);
+plotCI(1:dim(X)[2], log10(values),
+       uiw=0, liw=0,
+       ui=log10(values + 2*sd),
+       li=log10(values - 2*sd),
+       barcol="#000000",
+       main="Consumer Staples",
+       col="red",
+       lwd=1,
+       xlab="",
+       xaxt="n",
+       ylab=expression(log10(K^alpha)));
+axis(side=1, at=1:dim(X)[2],
+     labels=gsub("_series_", ".", gsub("_US", "", data$assets)),
+     las=2);
+dev.off();
+
 
 
 
[…]

### 2017-04-13 papers/FX/draft.tex
Revised the FX paper as well as the R scripts.

@@
 \subsection{A test of equal tail indices based on Hill estimator}
+Suppose we have two iid series $X_1, \dots, X_n \sim F_X$ and
+$Y_1, \dots, Y_n \sim F_Y$. Both $F_X$ and $F_Y$ have power-law tails
+with indices $\alpha_X$ and $\alpha_Y$, respectively.
 From \eqref{eq:2} one can deduce
+\[
+\sqrt k
+\begin{pmatrix}
+  {\hat \alpha_X - \alpha_X} \\
+  {\hat \alpha_X - \alpha_X} \\
+\end{pmatrix} \overset{d}{\to}
+\begin{pmatrix}
+  Z_X \\
+  Z_Y
+\end{pmatrix}
+\sim
+N\left(
+0, \text{diag}(\alpha_X^2, \alpha_Y^2)
+\right)
+\]
+where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimates of
+$\alpha_X$ and $\alpha_Y$. Then it follows from continuous mapping
+theorem
 \begin{equation}
   \label{eq:x1}
-  \sqrt k (\hat \alpha_A - \hat \alpha_B) \overset{d}{\to} N(0, \alpha_A^2 - \alpha_B^2)  
+  \sqrt k [(\hat \alpha_X -\alpha_X) - (\hat \alpha_Y - \alpha_Y)]
+  \overset{d}{\to}
+  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)
 \end{equation}
-where $\hat \alpha_A$ and $\hat \alpha_B$ are Hill estimates of
-$\alpha_A$ and $\alpha_B$, the tail indices of two series $A$ and
-$B$. The weak convergence of \eqref{eq:x1} allows one to construct a
-test of the equality of $\alpha_A$ and $\alpha_B$.
+The weak convergence of \eqref{eq:x1} allows one to construct a test
+of the equality $\alpha_X = \alpha_Y$. As the null hypothesis, assume
+$\alpha_X = \alpha_Y$. Under this hypothesis \eqref{eq:x1} means the
+test statistic
+$\hat \alpha_X - \hat \alpha_Y \overset{d}{\to} N(0, \alpha_X^2 +
+\alpha_Y^2)$
+We apply this test to the equities in the ``Energy'',
+``Consumder Staples'' and ``Information Technology''
+sectors of the S\&P 500 index. The results are shown in
+\ref{fig:PairTest}.
+\begin{figure}[htb!]
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[
+      width=\textwidth,
+    ]{HillTest_Energy.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 1cm, 1cm, 1cm}, clip
+    ]{Hoga_Energy_pair.pdf}
+  \end{minipage}
+  \caption{Test of equal tail indices}
+  \label{fig:PairTest} 
+\end{figure}
 

### 2017-04-14 papers/FX/draft.tex
Revised the script.
A major revision of the FX manuscript.

--- 2017-04-12 papers/FX/draft.tex
@@
 Figure~\ref{fig:sectors_parameters}.
+\begin{figure}[htb!]
+  \centering
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[width=\textwidth]
+    {Energy_K.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[width=\textwidth]
+    {Consumer_Staples_K.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[width=\textwidth]
+    {Information_Technology_K.pdf}
+  \end{minipage}
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[width=\textwidth]
+    {Energy_scale.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[width=\textwidth]
+    {Consumer_Staples_scale.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[width=\textwidth]
+    {Information_Technology_scale.pdf}
+  \end{minipage}
+  \caption{\small Estimates of $K$ and $K^\alpha$ of stocks in sectors
+    of the S\&P 500 index. The points are the estimated values; the bars are
+    the $2\sigma$ confidence intervals. The confidence bands of the
+    corresponding Hill estimates $\hat \alpha$ of these sectors are
+    shown in figure~\ref{fig:1}. Note $\log_{10}(K^\alpha)$ is shown
+    instead of $K^\alpha$.
+  }
+  \label{fig:sectors_parameters}
+\end{figure}
 One can see that $K$ generally takes a rather small value. For the
@@
 
-Secondly, one can see that the values of $\alpha$ and $K$ are appear
+Secondly, one can see that the values of $\alpha$ and $K$ appear
 positively correlated. This prompts Pareto tails of the return
@@
 
+Thirdly, figure~\ref{fig:sectors_parameters} shows the values of $K$
+in the ``Energy'' and ``Information Technology'' sectors are on
+average larger than those in the ``Consumer Staples'' sector. A larger
+value of $K$ means, for a given value of $p$, i.e. a given probability
+of losses, large losses are more probable. Thus one can conclude that
+these two sectors are considerablly riskier than the ``Consumer
+Staples'' sector. This is of course a confirmation of one's economic
+instinct.
+
+Yet another indication from figure \ref{fig:sectors_parameters} is
+that, while the ``Energy'' and the ``Information Technology'' sectors
+are similar in riskiness, the correlation between $\alpha$ and $K$ is
+stronger in ``Energy''. As discussed in the last section, moving along
+a curve of equal preference in the direction of increasing $\alpha$, $K$ also
+increases. So the strong positive correlation seen in the ``Energy''
+sector suggests these stocks might have very similar investor
+preferences. This in turn may be attributed to stronger business
+relations between the energy enterprises. While two IT companies
+may provide a variety of products and services and not depend on each
+other, two energy companies are more likely to depend on each other
+via relations of supplier and customer or otherwise to compete with each
+other if they are on the same link of the chain of energy production
+and distribution.
+
 \subsection{A test of equal tail indices based on Hill estimator}
+Suppose we have two weakly dependent series $X_1, \dots, X_n \sim F_X$ and
+$Y_1, \dots, Y_n \sim F_Y$. Both $F_X$ and $F_Y$ have power-law tails
+with indices $\alpha_X$ and $\alpha_Y$, respectively. Assume also
+$X_i$ is independent of $Y_j$ for all $1 \leq i \leq j \leq n$.
 From \eqref{eq:2} one can deduce
 \begin{equation}
+  \label{eq:x3}
+  \sqrt k
+  \begin{pmatrix}
+    {\hat \alpha_X - \alpha_X} \\
+    {\hat \alpha_X - \alpha_X} \\
+  \end{pmatrix} \overset{d}{\to}
+  \begin{pmatrix}
+    Z_X \\
+    Z_Y
+  \end{pmatrix}
+  \sim
+  N\left(
+    0, \text{diag}(\alpha_X^2, \alpha_Y^2)
+  \right)
+\end{equation}
+where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimates of
+$\alpha_X$ and $\alpha_Y$. Then it follows from continuous mapping
+theorem
+\begin{equation}
   \label{eq:x1}
-  \sqrt k (\hat \alpha_A - \hat \alpha_B) \overset{d}{\to} N(0, \alpha_A^2 - \alpha_B^2)  
+  \sqrt k [(\hat \alpha_X -\alpha_X) - (\hat \alpha_Y - \alpha_Y)]
+  \overset{d}{\to}
+  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)
 \end{equation}
-where $\hat \alpha_A$ and $\hat \alpha_B$ are Hill estimates of
-$\alpha_A$ and $\alpha_B$, the tail indices of two series $A$ and
-$B$. The weak convergence of \eqref{eq:x1} allows one to construct a
-test of the equality of $\alpha_A$ and $\alpha_B$.
+The weak convergence of \eqref{eq:x1} allows one to construct a test
+of the equality $\alpha_X = \alpha_Y$. As the null hypothesis, assume
+$\alpha_X = \alpha_Y$. Under this hypothesis, \eqref{eq:x1} means the
+test statistic $\hat \alpha_X - \hat \alpha_Y \overset{d}{\to} N(0,
+\alpha_X^2 + \alpha_Y^2)$. 
+We apply this test to the equities in the ``Energy'',
+``Consumder Staples'' and ``Information Technology''
+sectors of the S\&P 500 index. The results are shown in
+\ref{fig:PairTest}.
+\begin{figure}[htb!]
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{HillTest_Energy.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{HillTest_CS.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{HillTest_IT.pdf}
+  \end{minipage}
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{Hoga_Energy_pair.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{Hoga_CS_pair.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{Hoga_IT_pair.pdf}
+  \end{minipage}
+  \caption{
+    {\em 1st row}: Test for changing tail index using Hill estimator based on
+    concatenated series in the ``Energy'' and ``Consumer Staples''
+    sectors of S\&P 500. The green, blue and red colors represent,
+    respectively, the cases when the 85\%, 90\% and 95\% quantiles of
+    the limit distribution of the test statistic is exceeded. 
+    Grey points stand for pairs for which the test statistic is below
+    the 85\% \asy\ quantile;
+    {\em 2nd row:} Test for changing tail index or scale parameter
+    using Hoga's test based on the same concatenated series. black points represent
+    pairs for which the computation of $T_n$ fails for given precision
+    requirements and time limits.
+    The same number of upper-order statistics are used for both tests.
+  }
+  \label{fig:PairTest} 
+\end{figure}
+One can see from the figure that tail indices of equities in the
+``Energy'' or ``Information Technology'' sectors vary more than do
+their counterparts in the ``Consumer Staples'' sector, as the null
+hypothesis is rejected more often for members of these two sectors.
+Moreover, these figures suggest that the test based on Hill estimator
+is more powerful in distinguishing differences between tail indices,
+as it results in more rejections for the ``Energy'' and the
+``Information Technology'' sectors than does the other test. We
+argue this can be the result of slow convergence of the distribution
+of the test statistic to its asymptotic distribution. We elaborate on
+this in the section~\ref{sec:Hoga}.
+
+Nevertheless, one should to bear in mind that the weak convergence
+\eqref{eq:x3} is valid only on condition that both the $X_i$ and $Y_i$
+series are independent of each other, which is generally untrue for
+two return series in the same market.
 
@@
 any two series and apply the aforementioned test on the concatenated series. For the ``Energy'' and the
-``Consumer Staples'' sector we summarize the results in Figure~\ref{fig:Hoga_pair}.
+``Consumer Staples'' sector we summarize the results in Figure~\ref{fig:PairTest}.
 These graphs show that the ``null
@@
     \includegraphics[width=\textwidth]
-    {Hoga_Consumer_Staples_Single.pdf}
+    {Hoga_CS_Single.pdf}
   \end{minipage}
@@
 
+% \begin{figure}[htb!]
+%   \begin{minipage}{0.5\linewidth}
+%     \includegraphics[
+%       width=\textwidth,
+%       trim={0.3cm, 1cm, 1cm, 1cm}, clip
+%     ]{Hoga_Energy_pair.pdf}
+%   \end{minipage}\hfill
+%   \begin{minipage}{0.5\linewidth}
+%     \includegraphics[
+%       width=\textwidth,
+%       trim={0.3cm, 1cm, 1cm, 1cm}, clip
+%     ]{Hoga_CS_pair.pdf}
+%   \end{minipage}
+%   \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail index or scale parameter based on concatenated series in the ``Energy'' and
+%     ``Consumer Staples'' sectors of
+%     S\&P 500. The green, blue and red colors represent, respectively, the cases when  
+%     the 85\%, 90\% and 95\% quantiles of the limit distribution of $T_n$ is exceeded.
+%     Grey points stand for pairs for which $T_n$ is below
+%     the 85\% \asy\ quantile;
+%     black points represent pairs for which
+%     the computation of $T_n$ fails for given precision
+%     requirements and time limits.}
+%   \label{fig:Hoga_pair}
+% \end{figure}
+
+\par
+To get an idea about the power of the test we run it on a
+sample concatenated from two independent iid samples of the same size
+$n=1304$ as the S\&P 500 series. Both pieces are $t$-distributed with
+distinct degrees of freedom. The results are shown in
+Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
+larger the minimum tail index in the concatenated pair.
 \begin{figure}[htb!]
   \begin{minipage}{0.5\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 1cm, 1cm, 1cm}, clip
-    ]{Hoga_Energy_pair.pdf}
+    \includegraphics[width=\textwidth]{t_sim_pair.pdf}
   \end{minipage}\hfill
   \begin{minipage}{0.5\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 1cm, 1cm, 1cm}, clip
-    ]{Hoga_Consumer_Staples_pair.pdf}
+    \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
   \end{minipage}
-  \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail index or scale parameter based on concatenated series in the ``Energy'' and
-    ``Consumer Staples'' sectors of
-    S\&P 500. The green, blue and red colors represent, respectively, the cases when  
-    the 85\%, 90\% and 95\% quantiles of the limit distribution of $T_n$ is exceeded.
-    Grey points stand for pairs for which $T_n$ is below
-    the 85\% \asy\ quantile;
-    black points represent pairs for which
-    the computation of $T_n$ fails for given precision
-    requirements and time limits.}
-  \label{fig:Hoga_pair}
-\end{figure}
-
-\par
-To get an idea about the power of the test we run it on a
-sample concatenated from two independent iid samples of the same size $n=1304$ as the S\&P 500 series. 
-Both pieces are $t$-distributed with distinct degrees
-of freedom. The results are shown in Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the larger the minimum
-tail index in the concatenated pair.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\textwidth, trim={1.0cm 1.5cm 1cm 2cm},
-  clip]{t_sim_pair.pdf}
-  \caption{Test  of concatenated $t$-samples with different
-    degrees of freedom. Numbers on the axes are the degrees of freedom in the subsamples. For an interpretation of the colored
-bullets, see Figure~\ref{fig:Hoga_pair}. The graph shows the limited power of the test. In particular, if
-both degrees of freedom are relatively large it loses the capablity of  distinguishing the \ds .}
+  \caption{
+    {\em Left}: Test  of concatenated $t$-samples with different
+    degrees of freedom. Numbers on the axes are the degrees of freedom
+    in the subsamples. For an interpretation of the colored bullets,
+    see Figure~\ref{fig:PairTest}. The graph shows the limited power
+    of the test. In particular, if both degrees of freedom are
+    relatively large it loses the capablity of distinguishing the \ds\ .
+    {\em Right}: Comparison of the \asy\ distribution of the test statistic
+    $T_n$ in \eqref{eq:4} under the null hypothesis and the \ds\ of
+    $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees
+    of freedom.
+  }
   \label{fig:t_sim_pair}
 \end{figure}
+
 \par
@@
 with $n=1304$ for $t$-distributed $X_t$ with  3 and 4 degrees of freedom and the limit \ds\ of $T_n$.
-The estimated density functions are shown in Figure~\ref{fig:Q1_distr}.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
-  \caption{{\red Change $Q_1$ on $x$-axis} Comparison of the \asy\ distribution of the test statistic $T_n$ in \eqref{eq:4} under the null hypothesis 
-and the \ds\ of $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees of freedom.}
-  \label{fig:Q1_distr}
-\end{figure}
-As seen in the graph, the asymptotic distribution 
-assigns significantly more mass to the tail than the \ds s of $T_n$ do. For comparison, we list a few quantiles of these \ds s in
-Table~\ref{tab:HogaAsymptotic}, showing major differences between the \asy\ and finite-sample \ds s.
+The estimated density functions are shown in on the right of Figure~\ref{fig:t_sim_pair}.
+As seen in the graph, the asymptotic distribution assigns
+significantly more mass to the tail than the \ds s of $T_n$ do. For
+comparison, we list a few quantiles of these \ds s in
+Table~\ref{tab:HogaAsymptotic}, showing major differences between the
+\asy\ and finite-sample \ds s.
 \begin{table}[htb!]
@@
 
-\subsection{An utility based argument for equal tail indices} \setcounter{equation}{0}\label{sec:3}
+\subsection{An utility based argument for equal tail indices}\label{sec:3}
 In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
@@
 consult Casper for references.} Therefore we adopt a more convenient
-notation
+notation in this case
 \beao
@@
     {\red Is this worth a theorem?}
-    {\blue No}
+    {\blue No, let's change it to a remark.}
 \begin{remark}\label{thrm:I}
- Assume the two-sided Pareto model \eqref{eq:pp} and that the utility \fct\ $u$ is increasing and differentiable. Then
-$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
+ Assume the two-sided Pareto model \eqref{eq:pp} and that the utility
+ \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
+   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
+   K} < 0$.
 \end{remark}
@@
 
-\subsection{Fitting the shifted Pareto distribution}
-\label{sec:hybrid_estimation}
-We adopt a hybrid procedure to estimate the parameters of the
-distribution function assumed in \eqref{eq:pareto}: First we
-obtain a Hill estimate of the lower-tail indices of an
-equity return and fix $\alpha$ in \eqref{eq:pareto}
-to its Hill estimate. Then we estimate $K$ by maximizing
-the likelihood function with fixed $\alpha$. Obviously we take only
-negative observations and hence condition on $X < 0$. Regularity
-conditions on the maximum likelihood estimator are easily checked for
-the distribution \eqref{eq:pareto}.
-The density function of $X$ conditional on $X < 0$ is
-\[
-f_L(x) = {\alpha K^{\alpha} \over (K - x)^{\alpha + 1}}
-\]
-Clearly, the log-likelihood function of the $K$ parameter is
-\[
-l(K | X < 0) = \ln(\alpha) + \alpha \ln(K) - (\alpha + 1)\ln(K - X)
-\]
-where, as is just mentioned, we pre-estimate $\alpha$ with Hill
-estimator and regard it as known at this step. With $n$ observations
-$X_1, \dots, X_n$, the log-likelihood function of $K$ is
-\[
-L(K | \forall i=1,2,...,n, X_i < 0) = n \ln(\alpha) + n \alpha \ln(K)
-- (\alpha + 1) \sum_{i=1}^n \ln(K - X_i)
-\]
-To obtain the maximum likelihood estimation, we need to solve the
-equation ${\pd L \over \pd K} = 0$, which expands to
-\[
-{n \alpha \over K} - (\alpha + 1) \sum_{i=1}^n {1 \over K - X_i} = 0
-\]
-This equation has to be solved numerically.
-The Fisher information of $K$ follows as
-\begin{eqnarray*}
-  I(K) &=& \E\left[ \left( {\partial \over \partial K} l(K | X < 0) \right)^2 \right] \\
-  &=& \alpha K^{\alpha - 2}
-  \int_{-\infty}^0 {(\alpha x + K)^2 \over (K - x)^{\alpha + 3}} dx \\
-  &=& {\alpha \over (2 + \alpha) K^2}
-\end{eqnarray*}
-So the maximum likelihood estimator, call it $\hat K$, is
-asymptotically normal:
-\[
-\sqrt n (\hat K - K) \to N \left(
-  0, \left( 1 + {2 \over \alpha} \right) K^2
-\right)
-\]
-
-To obtain an estimate of $p$, we use the law of large number:
-\begin{eqnarray*}
-  p &=& \P(X < 0) = \lim_{n \to \infty}
-  {1 \over n} \sum_{t=1}^n \1{X_t < 0}
-\end{eqnarray*}
-By the central limit theorem,
-\begin{eqnarray*}
-  \sqrt n \left[
-    {1 \over n}\sum_{t=1}^n \1_{\{X_t < 0\}} - p
-  \right] \to N(0, p - p^2))
-\end{eqnarray*}
-
-The plots in the 1st column of figure \ref{fig:sectors_parameters}
-show the estimates of $K$ alongside the Hill estimates of $\alpha$ of
-the ``energy'', ``consumer staples'' and ``information technology''
-sectors of the S\&P 500 index;
-The plots in the 2nd column show the estimates of $p$ of the same
-sectors.
-
-From these figures, one can see that stocks of the ``Energy'' and the
-``Information Technology'' sectors generally have larger values of
-$p$, meaning their stock prices fall more often than do those in the
-``consumer Staples'' sector; Also, the values of $K$ in the ``Energy''
-and ``Information Technology'' sectors are larger than those in the
-``Consumer Staples'' sector. A larger value of $K$ means, for a given
-value of $p$, i.e. a given probability of losses, large losses are
-more probable. Thus one can conclude that these two sectors are
-considerablly riskier than the ``Consumer Staples'' sector. This is of
-course a confirmation of one's economic instinct.
-
-One can also see from figure \ref{fig:sectors_parameters} that, while
-the ``Energy'' and the ``Information Technology'' sectors are similar in
-riskiness, the correlation between $\alpha$ and $K$ is stronger in
-``Energy''. As discussed in the last section, moving along a curve of
-equal preference in the direction of increasing $\alpha$, $K$ also
-increases. So the strong positive correlation seen in the ``Energy''
-sector suggests these stocks might have very similar investor
-preferences. This in turn may be attributed to stronger business
-relations between the energy enterprises. While two IT companies
-may provide a variety of products and services and not depend on each
-other, two energy companies are more likely to depend on each other
-via relations of supplier and customer or otherwise to compete with each
-other if they are on the same link of the chain of energy production
-and distribution.
-
-\begin{figure}[htb!]
-  \centering
-  \begin{minipage}{0.45\linewidth}
-    \includegraphics[width=\textwidth]
-    {Energy_K.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.45\linewidth}
-    \includegraphics[width=\textwidth]
-    {Energy_scale.pdf}
-  \end{minipage}
-  \begin{minipage}{0.45\linewidth}
-    \includegraphics[width=\textwidth]
-    {Consumer_Staples_K.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.45\linewidth}
-    \includegraphics[width=\textwidth]
-    {Consumer_Staples_scale.pdf}
-  \end{minipage}
-  \begin{minipage}{0.45\linewidth}
-    \includegraphics[width=\textwidth]
-    {Information_Technology_K.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.45\linewidth}
-    \includegraphics[width=\textwidth]
-    {Information_Technology_scale.pdf}
-  \end{minipage}
-  \caption{\small Estimates of $K$ and $K^\alpha$ of stocks in sectors
-    of the S\&P 500 index. The points are the estimated values; the bars are
-    the $2\sigma$ confidence intervals. The confidence bands of the
-    corresponding Hill estimates $\hat \alpha$ of these sectors are
-    shown in figure~\ref{fig:1}. Note $\log_{10}(K^\alpha)$ is shown
-    instead of $K^\alpha$.
-  }
-  \label{fig:sectors_parameters}
-\end{figure}
-To compare the maximum likelihood estimates with the results obtained
-from Hill's scale estimates, one can re-write the lower-tail
-distribution function: For all $x < 0$,
-\begin{eqnarray*}
-  \P(X < x)
-  &=&
-  p {K^\alpha \over (K - x)^\alpha} \\
-  &=&
-  p {K^\alpha |x|^{-\alpha} \over (1 - K/x)^\alpha} \\
-  &=&
-  p K^\alpha |x|^{-\alpha} \left(
-  1 + {\alpha K \over x} + O(|x|^{-2})
-  \right)
-\end{eqnarray*}
-% Comparing this expression with \eqref{eq:lower_tail_expansion},
-Then one can identity $p K^\alpha$ as the scale parameter. Figure
-\ref{fig:scale_estimates} compare the estimates obtained
-from Hill's scale estimator with those from the $p K^\alpha$ method.
-
-\begin{figure}[htb!]
-  \centering
-  \begin{minipage}{0.32\linewidth}
-    \includegraphics[width=\textwidth]
-    {Energy_scale.pdf}
-  \end{minipage}
-  \begin{minipage}{0.32\linewidth}
-    \includegraphics[width=\textwidth]
-    {Consumer_Staples_scale.pdf}
-  \end{minipage}
-  \begin{minipage}{0.32\linewidth}
-    \includegraphics[width=\textwidth]
-    {Information_Technology_scale.pdf}
-  \end{minipage}
-  \caption{\small Scale estimates obtained from Hill's scale estimator
-    and from $p K^\alpha$ . Red dots are Hill's scale estimates and
-    triangles are $p K^\alpha$ estimates.
-  }
-  \label{fig:scale_estimates}
-\end{figure}
-
 \subsection{Student's t-distributed equity return}
@@
 where $C(\cdot)$ is defined in \eqref{eq:xxie1}.
-
-\begin{lemma}
-  \label{lemma:II}
-  There exists $x_0 > 0$ such that ${\pd f \over \pd \nu}(x_0, \nu) = 0$ and
-  $\forall x \in (0, x_0), {\pd f \over \pd \nu}(x, \nu) > 0$ and
-  $\forall x \in (x_0, \infty), {\pd f \over \pd \nu}(x, \nu) < 0$.
-\end{lemma}
 
@@
 
-\section{Proof of lemma \ref{lemma:II}}
+\section{A lemma on Student's t-distribution}
+\begin{lemma}
+  \label{lemma:II}
+  Let $f$ denotes the density function of the Student's
+  t-distribution, i.e.
+  \[
+  f(x; \nu) = c(\nu) \left(
+    1 + {x^2 \over \nu}
+  \right)^{-(\nu + 1)/2}
+  \]
+  where $\nu > 1$ and
+  \[
+  c(\nu) = {
+    \Gamma({\nu + 1 \over 2})
+    \over
+    \Gamma(\nu/2) \sqrt{\nu \pi}
+  }
+  \]
+  Then there exists $x_0 > 0$ such that ${\pd f \over
+    \pd \nu}(x_0, \nu) = 0$ and $\forall x \in (0, x_0), {\pd f \over
+    \pd \nu}(x, \nu) > 0$ and $\forall x \in (x_0, \infty), {\pd f
+    \over \pd \nu}(x, \nu) < 0$.
+\end{lemma}
 \begin{proof}

### 2017-04-17 papers/FX/draft.tex

@@
 %\usepackage{enumerate}
-\usepackage[bookmarks=true]{hyperref}
+% \usepackage[bookmarks=true]{hyperref}
 %% \usepackage{bookmark}
@@
 \eeao
+{\blue
 Using the asymptotic normality property of upper order statistics
@@
 ``Information Technology'' sectors than does the other test. We
-argue this can be the result of slow convergence of the distribution
-of the test statistic to its asymptotic distribution. We elaborate on
-this in the section~\ref{sec:Hoga}.
-
-Nevertheless, one should to bear in mind that the weak convergence
+argue this lack of power of the latter can be the result of slow
+convergence of the distribution of its test statistic to the
+asymptotic distribution. We elaborate on this in section~\ref{sec:Hoga}.
+
+Nevertheless, one should bear in mind that the weak convergence
 \eqref{eq:x3} is valid only on condition that both the $X_i$ and $Y_i$
@@
 two return series in the same market.
-
-%% By the design of Hoga's test, the null hypothesis can be
-%% rejected due to changed scale parameters instead of changed tail
-%% indices. Thus it is useful to apply the Hill scale estimator to the
-%% same stock return series.
+}
 
 \subsection{Tests for a change in the extreme tail}\label{sec:Hoga}
-To the best of our knowledge, there do not exist statistical tests for comparing the 
-tail indices or scale parameters of two or more dependent \ts\ in  a direct way. {\red construct a test for this}
-We borrow a test from a recent paper by Hoga \cite{hoga:2016}. 
-This test has been developed for a different kind of problem. Given a strictly stationary \ts\ $X_1,\ldots,X_n$
-with a marginal \ds\ $F$ with right power-law tail, the goal is to test whether there is a structural break 
-of the {\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to zero. If the tail index
-{\em or} the scale parameter in a \ds\ of type \eqref{eq:3} change inside a sample, then it is likely
-that the extreme quantiles change as well. We will test for a change of tail index or scale parameter 
-in this indirect way.
+To the best of our knowledge, there do not exist published statistical
+tests for comparing the tail indices or scale parameters of two or
+more dependent \ts\ in a direct way. In addition to the test described
+in the previous section, we borrow a test from
+a recent paper by Hoga \cite{hoga:2016}. This test has been developed
+for a different kind of problem. Given a strictly stationary
+\ts\ $X_1,\ldots,X_n$ with a marginal \ds\ $F$ with right power-law
+tail, the goal is to test whether there is a structural break of the
+{\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to
+zero. If the tail index {\em or} the scale parameter in a \ds\ of type
+\eqref{eq:3} change inside a sample, then it is likely that the
+extreme quantiles change as well. We will test for a change of tail
+index or scale parameter in this indirect way.
 \par
@@
 \eeao
-where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red
-  meaning of these constants?}
+where $\delta $ and $v$ are positive constants, and $b\ge 0$.
 {\blue 
@@
 \eeam
-    {\red Is this worth a theorem?}
-    {\blue No, let's change it to a remark.}
+{\blue
 \begin{remark}\label{thrm:I}
@@
 \end{remark}
+}
 The proof is given in Appendix~\ref{sec:thrmI_proof}. 
@@
     function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$ (left) and $\xi = 4$
-    (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other
-      figures, and use the same structure of the formulas.}
+    (right).
   }
@@
 \end{minipage}\hfill
-\begin{minipage}{0.42\textwidth}{\red change formulas in graph and $C$ to $x$.}
-  In this graph we show two power-utility functions $u$. We see that $u(x)$ grows slower
-  and saturates later for small $\xi$ than for large $\xi$. The latter case
-  represents an agent who is more tolerant about low
-  consumption and seeks wealth more aggressively. In other
-  words, he is less risk-averse than one with a larger $\xi$.
-  Such an agent will therefore invest more heavily in equity, as shown in Figure~\ref{fig:phi_hat_pareto}.
+\begin{minipage}{0.42\textwidth}
+  In this graph we show two power-utility functions $u$. We see that
+  $u(x)$ grows slower and saturates later for small $\xi$ than for
+  large $\xi$. The latter case represents an agent who is more
+  tolerant about low consumption and seeks wealth more
+  aggressively. In other words, he is less risk-averse than one with a
+  larger $\xi$. Such an agent will therefore invest more heavily in
+  equity, as shown in Figure~\ref{fig:phi_hat_pareto}.
 \end{minipage}
@@
 \]
-We can write $\mathcal U(F, \phi)$ as
+We can write $\wt u(F, \phi)$ as
 \begin{eqnarray}
-  && \mathcal U(F, \phi) \nonumber \\
+  && \wt u(F, \phi) \nonumber \\
   &=&
@@
 
-First notice that when $b = 0$, $\mathcal U(F, \phi)$ reduces to
+First notice that when $b = 0$, $\wt u(F, \phi)$ reduces to
 $\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
@@
 \ref{lemma:II},
-${\pd \mathcal U(F, \phi)) \over \pd \nu} < 0$ for all $\phi$. Hence all
+${\pd \wt u(F, \phi)) \over \pd \nu} < 0$ for all $\phi$. Hence all
 investors guided by this utility function would be seeking the
@@
 In fact, even when $b > 0$, the same monotonicity property holds. This
-is revealed by numerically maximizing $\mathcal U(F, \phi)$ with
+is revealed by numerically maximizing $\wt u(F, \phi)$ with
 respect to $\phi$ for a sequence of selected values of $\nu$. The
@@
 %% \begin{eqnarray*}
-%%   \mathcal U(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
+%%   \wt u(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
 %% \end{eqnarray*}
 %% Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
-%% $\opd{\nu}\mathcal U(F, \phi) > 0$. In this extreme case, the investor
+%% $\opd{\nu}\wt u(F, \phi) > 0$. In this extreme case, the investor
 %% would rather seek the largest $\nu$ or in other words, the lightest
@@
 \[
-u(C) = -{1 \over \xi} C^{-\xi}
+u(x) = -{1 \over \xi} x^{-\xi}
 \]
@@
     \label{eq:xxie2}
-    \hat \phi := \text{argmax}_{0 < \phi \leq 1} \mathcal U(F, \phi)
+    \hat \phi := \underset{0 < \phi \leq 1}{\text{argmax}}\,
+    \wt u(F, \phi)
   \end{equation}
@@
   \[
-  G(F) = \mathcal U(F, \hat\phi)
+  G(F) = \wt u(F, \hat\phi)
   \]
@@
     &=&
-    {\pd \mathcal U(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi)
-    + {\pd \mathcal U(F, \phi) \over \pd \phi}(\alpha, K, \hat\phi)
+    {\pd \wt u(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi)
+    + {\pd \wt u(F, \phi) \over \pd \phi}(\alpha, K, \hat\phi)
     {\pd \hat\phi \over \pd \alpha}(\alpha, K)
@@
     \label{eq:xxie4}
-    {\pd \mathcal U(F, \phi) \over \pd \phi}(\alpha, K, \hat \phi) = 0
+    {\pd \wt u(F, \phi) \over \pd \phi}(\alpha, K, \hat \phi) = 0
   \end{equation}
@@
   \begin{eqnarray*}
-    {\pd \mathcal U(F, \phi) \over \pd \alpha}(\alpha, K, \hat \phi)
+    {\pd \wt u(F, \phi) \over \pd \alpha}(\alpha, K, \hat \phi)
     &>& 0
@@
   \begin{eqnarray*}
-    {\pd \mathcal U(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi)
+    {\pd \wt u(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi)
     = {\partial \over \partial \alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1_{\{X < 0\}}]
@@
   \begin{eqnarray*}
-    {\pd \mathcal U(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi) &>& 0
+    {\pd \wt u(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi) &>& 0
   \end{eqnarray*}
@@
   \[
-  {\pd \mathcal U(F, \phi) \over \pd K}(\alpha, K, \hat\phi) < 0
+  {\pd \wt u(F, \phi) \over \pd K}(\alpha, K, \hat\phi) < 0
   \]
@@
   \[
-  {\pd \mathcal U(F, \phi) \over \pd K}(\alpha, K, \hat\phi)
+  {\pd \wt u(F, \phi) \over \pd K}(\alpha, K, \hat\phi)
   = {\partial \over \partial K} \E \left[
@@
   \[
-  {\pd \mathcal U(F, \phi) \over \pd K}(\alpha, K, \hat\phi) < 0  
+  {\pd \wt u(F, \phi) \over \pd K}(\alpha, K, \hat\phi) < 0  
   \]
@@
 \bibliographystyle{unsrt}
-\bibliography{../../thesis/econophysics}
+%% \bibliography{../../thesis/econophysics}
 \end{document}

### 2017-04-18 papers/FX/draft.tex
corrected the dates of the data used in the FX paper.

@@
 %\usepackage{enumerate}
-% \usepackage[bookmarks=true]{hyperref}
+\usepackage[bookmarks=true]{hyperref}
 %% \usepackage{bookmark}
@@
 estimates are based on 1304 observations of daily return 
-data from 1 January 2010 to {\red why 1 January?} 1 January 2015.
+data from 4th Jan. 2010 to 31st Dec. 2014.
 \par
@@
   \end{minipage}
-  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail indices $\alpha$ of
-    daily return series in sectors of the S\&P 500
-    index. The data span
-    from 2010-01-01 to 2015-01-01 and comprise $n=1304$ observations.
-The graphs from top to bottom correspond to the ``Energy'',
-    ``Consumer Staples'' and ``Information Technology'' sectors.
+  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail
+    indices $\alpha$ of daily return series in sectors of the S\&P 500
+    index. The data span from 4th Jan. 2010 to 31st Dec. 2014 and
+    comprise $n=1304$ observations. The graphs from top to bottom
+    correspond to the ``Energy'', ``Consumer Staples'' and
+    ``Information Technology'' sectors.
     Each circle corresponds to a Hill estimate $\hat\alpha_k$; the grey
@@
 are similar in riskiness, the correlation between $\alpha$ and $K$ is
-stronger in ``Energy''. As discussed in the last section, moving along
+stronger in ``Energy''. As discussed in section \ref{sec:pareto_tail},
+moving along
 a curve of equal preference in the direction of increasing $\alpha$, $K$ also
@@
 \bibliographystyle{unsrt}
-%% \bibliography{../../thesis/econophysics}
+\bibliography{../../thesis/econophysics}
 \end{document}

### 2017-04-19 papers/FX/draft.tex
1. revised the paper with Jeff 2. corrected figures in the FX paper.

@@
   \sqrt k (\hat K_k - K) &\overset{d}{\to}& N\left(
-    0, \left( {K \over \alpha}  \right)^2
+    0, {K^2 \over \alpha^2}
   \right) \\
   \sqrt k (\hat K_k^\alpha - K^\alpha) &\overset{d}{\to}& N\left(
-    0, \left( {K^{2\alpha}}  \right)^2
+    0, {K^{2\alpha}}
   \right)
@@
     \includegraphics[width=\textwidth]
-    {Consumer_Staples_K.pdf}
+    {CS_K.pdf}
   \end{minipage}\hfill
@@
     \includegraphics[width=\textwidth]
-    {Information_Technology_K.pdf}
+    {IT_K.pdf}
   \end{minipage}
@@
     \includegraphics[width=\textwidth]
-    {Consumer_Staples_scale.pdf}
+    {CS_scale.pdf}
   \end{minipage}\hfill
@@
     \includegraphics[width=\textwidth]
-    {Information_Technology_scale.pdf}
+    {IT_scale.pdf}
   \end{minipage}
@@
 to the smallness of $K$, mild variations of $\alpha$ would lead to
-huge variations of $K^\alpha$, as shown in the 2nd column of
+huge variations of $K^\alpha$, as shown in the 2nd row of
 figure~\ref{fig:sectors_parameters}.
@@
 $\alpha_X = \alpha_Y$. Under this hypothesis, \eqref{eq:x1} means the
-test statistic $\hat \alpha_X - \hat \alpha_Y \overset{d}{\to} N(0,
-\alpha_X^2 + \alpha_Y^2)$. 
+test statistic $\sqrt k (\hat \alpha_X - \hat \alpha_Y)
+\overset{d}{\to} N(0, \alpha_X^2 + \alpha_Y^2)$.
 We apply this test to the equities in the ``Energy'',

### 2017-04-24 cv/20sec/XieXiaolei.tex
Revised my CV.

--- 2017-04-07 cv/20sec/template.tex
@@
 
-\cvdate{09 January 1983} % Date of birth
-\cvaddress{Sweden} % Short address/location, use \newline if more than 1 line is required
+\cvdate{09 Jan. 1983, Swedish citizen} % Date of birth
+\cvaddress{Kivksgatan 7C, 21440 Malm\"o} % Short address/location, use \newline if more than 1 line is required
 \cvnumberphone{+46 760606638} % Phone number
@@
 \item {\it
-    \href{http://www.goodmillsystems.com}{\tt Goodmillsystems,
+    \href{http://www.goodmillsystems.com}{\tt Goodmill Systems,
       Helsinki}, Software Designer, May 2007 - Sep 2007}

### 2017-04-25 cv/CoverLetter/Eurex.tex

--- 2017-04-10 cv/CoverLetter/CitiMQA_QuantitativeAnalyst.tex
@@
 % recipient data
-\recipient{Cross Asset Quantitative Analysis team}{Citi}{London}
+% \recipient{Cross Asset Quantitative Analysis team}{Citi}{London}
+\recipient{Deutsche B\"orse Group}{Frankfurt am Main}{Germany}
 \date{\today}
@@
 
-I see on Citi's website that you are recruiting a quantitative
-analyst to join your cross asset quantitative analysis team. In the
-description of the job, you emphasised knowledge and skills in time
-series analysis, C++ programming and understanding of the financial
-market, which are indeed areas that I know well and am passionate
-about. So I would like to ask you for your consideration.
+I see on {\em Monster} that you are recruiting a quantitative risk
+analyst to join your team of {\em Risk methodology and New products}.
+From your description of the responsibilities and requirements, I
+understand you are looking for someone who is strong both in
+statistical analysis and in software development using Matlab, C++ and
+SQL.
 
-I am currently a PhD student at Copenhagen University, researching in
-Mathematical statistics and particularly in time series analysis,
-extreme value theory and random matrix theory. I am available for
-immediate full-time work and expect to receive my degree on 30th
-September, 2017. So far I have co-authored and published two papers on
-eigen systems of sample covariance matrices, with a focus on
-heavy-tailed (e.g. power-law tail) time series. These results are
-applicable to principle component analysis, for example.
+As is detailed in my CV, I am anticipating to get my PhD in {\em
+  Mathematical Statistics} from Copenhagen University this September
+and have co-authored 2 papers on the eigen systems of sample
+covariance matrices of heavy tailed time series. At the time of
+writing, I am also finalizing two other papers, one about tail
+parameters of equity returns and the other about estimation of
+the probabilities of large losses by importance sampling.
+These results are applicable to principle component analysis,
+portfolio optimization as well as risk management; and indeed, I am
+keen to apply them and see them making a real impact on business.
 
-I am also finalising two other papers. One of them is about how an
-investor's preference (as characterised by {\em Generalised
-  Disappointment Aversion}) of an equity varies with the tail
-parameters of the stationary distribution of the equity. This
-result can be applied to, for example, portfolio construction and
-assessment of the overall market risk. As part of the paper, I have
-also done statistical analysis on the tail parameters of equities in
-the S\&P 500 index, specifically, the ``Energy'', ``Consumer Staples''
-and ``Information Technology'' sectors. R and C++ programming are
-indispensable part of the work.
-
-The fourth paper of mine is about estimation of the probability of very
-large losses posed by portfolios whose volatility dynamics are
-described by univariate GARCH(p,q) or multivariate GARCH
-processes. Because the stationary distribution of such a process is
-not known exactly, the aforementioned probability as well as the
-related risk measures {\em Value at Risk} and {\em Expected
-  Shortfall} cannot be computed precisely. In the paper I propose 
-an estimator based on importance sampling and Markov chain theory and
-naturally, would be more than happy to see it applied in practical
-risk management. I have implemented the estimator in C++, and in
-particular, have used the OpenMP directives to achieve high
-performance via parallel computing.
-
-Citi is a great investment bank that I have known since my
-childhood. I know taking a role in this esteemed organisation presents
-many challenges and opportunities. Through constant endeavour towards
-ever greater excellence, I hope to develop a productive career that I
-will be proud of by contributing my part to the success of Citi.
+Speaking about the application of mathematical models, I think my past
+experience as a software developer is worth mentioning. As is detailed
+in my CV, I worked as a software developer for 6.5 years in several
+companies and accumulated abundant experience in C, C++, R, Matlab
+programming as well as in data analysis and algorithm implementation.
+Thus I am confident to take on the challenges of a quantitative risk
+analyst at {\em Eurex}.
 

### 2017-04-25 cv/References.tex
Added CV and cover letters.

@@
 
-My recommenders are:
+I would like to refer you to the following contacts for references of
+my education and employment.
 \begin{itemize}
-\item Sven \AA berg, professor and Head of Mathematical Physics at
-  Lund University
-\item Ferdi Aryasetiawan, professor of Mathematical Physics at
-  Lund University
+\item Thomas Mikosch, {\em professor of Mathematics at Copenhagen
+    University}, my PhD supervisor
+  \begin{itemize}
+  \item Email: mikosch@math.ku.dk
+  \item Office phone: +45 353 30414
+  \end{itemize}
+\item Sven \AA berg, {\em professor and Head of Mathematical Physics
+    at Lund University}, my Master thesis supervisor
+  \begin{itemize}
+  \item Email: sven.aberg@matfys.lth.se
+  \item Office phone: +46 46 222 96 33
+  \item Mobile phone: +46 70 557 35 69
+  \end{itemize}
+\item Ferdi Aryasetiawan, {\em professor of Mathematical Physics at
+    Lund University}, my teacher of multiple courses
+  \begin{itemize}
+  \item Email: ferdi.aryasetiawan@teorfys.lu.se
+  \item Office phone: +46 46 222 90 89
+  \end{itemize}
+\item Fredrik l\"of, {\em Technical Project Manager and Team Leader at
+    Bombardier Transportation}, my supervisor at ST-Ericsson
+  \begin{itemize}
+  \item Email: fredrik.n.lof@gmail.com
+  \item LinkedIn: https://www.linkedin.com/in/fredrik-lof/
+  \end{itemize}
 \end{itemize}

### 2017-04-26 papers/FX/draft.tex
Added Thomas' changes to the manuscript.

--- 2017-04-17 papers/FX/draft.tex
@@
 This observation is often explained by the fact that investors are
-more prone to negative than to positive news in te market. 
+more prone to negative than to positive news in the market. 
 Moreover, in the literature the {\em estimated} tail-indices $\hat \alpha$ (both in the left and right tails) 
@@
 estimates are based on 1304 observations of daily return 
-data from 1 January 2010 to {\red why 1 January?} 1 January 2015.
+data from 1 January 2010 to 31 December 2014.
 \par
@@
 some of its virtues and vices. In addition to tail-index estimation we also discuss the related problem of
-estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). It turns out 
+estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). {\red check} It turns out 
 that the simultaneous estimation of the tail-index and the scale parameter are strongly related.
@@
 In Section~\ref{sec:2} we discuss the theoretical problem of appearance of power-law tails in models 
-for daily or, more generally, low \fre y 
+for daily or, more generally, low-\fre y 
 return data. In particular, in Section~\ref{subsec:garch} we address the power-law tails of univariate and multivariate GARCH models as potential models
@@
 have power-law tails due to their relation with so-called {\em \sre s}. Moreover, some of the {\em standard} multivariate 
-GARCH models as the CCC or BEKK ensure that the marginal \ds s have power-law tails with the same index.
+GARCH models as the CCC ensure that the component-wise marginal \ds s have power-law tails with the same index.
 \par
 In Section~\ref{sec:3} we discuss an economic argument for the fact that return data of similar assets
-(such as return series in a given sector of the S\&P 500 index) may have tail indices which are close to each other.
+(such as return series in a given sector of the S\&P 500 index) may have tail-indices which are close to each other.
 We argue based on a  utility \fct\ approach. We explicitly recognize the behavioral
@@
 Given a sample $-X_1,\ldots,-X_n$ whose marginal \ds\ satisfies  \eqref{eq:1a}, calculate
-the order statistices $X_{(1)}\le \cdots\le X_{(n)}$  and construct the
+the order statistics $X_{(1)}\le \cdots\le X_{(n)}$  and construct the
 {\em Hill estimator}:
@@
   \hat \alpha_k = \Big(
-    {1 \over k} \sum_{i=1}^k \ln {X_{(n-i+1)} \over X_{(n-k)}}
+    {1 \over k} \sum_{i=1}^k \log {X_{(n-i+1)} \over X_{(n-k)}}
     \Big)^{-1}\,.
@@
 A major problem for Hill estimation is the choice of the number $k$ of
-order statistics. As  matter of fact,
+order statistics. As  a matter of fact,
 if $k$ is too large the order statistics are too close to the center
@@
     index. The data span
-    from 2010-01-01 to 2015-01-01 and comprise $n=1304$ observations.
+    from 1 January 2010 to 31 December 2011 and comprise $n=1304$ observations.
 The graphs from top to bottom correspond to the ``Energy'',
     ``Consumer Staples'' and ``Information Technology'' sectors.
-    Each circle corresponds to a Hill estimate $\hat\alpha_k$; the grey
+    Each circle corresponds to a Hill estimate $\hat\alpha_{50}$; the grey
     triangles above and below it mark the 97.5\% and 2.5\% quantiles
@@
 \beao
-\hat K_k = \left({k \over n}\right)^{1/\hat\alpha} X_{(n-k)}\,.
+\hat K_k = \left({k \over n}\right)^{1/\hat\alpha_k} X_{(n-k)}\,.
 \eeao
-{\blue
 Using the asymptotic normality property of upper order statistics
-(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, theorem 2.2.1), one can show
-\begin{eqnarray*}
-  \sqrt k (\hat K_k - K) &\overset{d}{\to}& N\left(
-    0, \left( {K \over \alpha}  \right)^2
-  \right) \\
-  \sqrt k (\hat K_k^\alpha - K^\alpha) &\overset{d}{\to}& N\left(
-    0, \left( {K^{2\alpha}}  \right)^2
-  \right)
+(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, Theorem 2.2.1), one can show
+\beao
+  \sqrt{k}\, (\hat K_k - K) &\std& N\big(
+    0, (K /\alpha)^2\big)\quad\mbox{and}\quad
+  \sqrt {k} \,(\hat K_k^\alpha - K^\alpha) \overset{d}{\to} N\big(
+    0, K^{2\alpha}  \big)\,,
 \end{eqnarray*}
-where the tail index $\alpha$ is regarded as known. From the above
-asymptotic normality property, confidence bands of the estimate of $K$
-and $K^\alpha$ can be constructed. 
-Estimates of $K$ of the ``Energy'', ``Consumer Staples'' and
+where the tail-index $\alpha$ is regarded as known. From the above
+asymptotic normality property, confidence bands of $\hat K_k$
+and $\hat K_k^{\hat\alpha}$ can be constructed.   
+Estimates $\hat K_k$ in the ``Energy'', ``Consumer Staples'' and
 ``Information Technology'' sectors of the S\&P 500 index are computed
@@
     \includegraphics[width=\textwidth]
-    {Information_Technology_scale.pdf}
+    {IT_scale.pdf}
   \end{minipage}
-  \caption{\small Estimates of $K$ and $K^\alpha$ of stocks in sectors
-    of the S\&P 500 index. The points are the estimated values; the bars are
-    the $2\sigma$ confidence intervals. The confidence bands of the
-    corresponding Hill estimates $\hat \alpha$ of these sectors are
-    shown in figure~\ref{fig:1}. Note $\log_{10}(K^\alpha)$ is shown
-    instead of $K^\alpha$.
+  \caption{\small Estimates $\hat K_k$ and $\hat K_k^{\hat \alpha}$ on $\log_{10}$-scale 
+of stocks in sectors
+    of the S\&P 500 index. The points are the estimated values, the bars 
+    the \asy\ 95\%-confidence intervals; the confidence bands of the
+    corresponding Hill estimates $\hat \alpha_k$ of these sectors are
+    shown in Figure~\ref{fig:1}. 
   }
@@
 to the smallness of $K$, mild variations of $\alpha$ would lead to
-huge variations of $K^\alpha$, as shown in the 2nd column of
-figure~\ref{fig:sectors_parameters}.
-
-Secondly, one can see that the values of $\alpha$ and $K$ appear
-positively correlated. This prompts Pareto tails of the return
-distributions. We expain more about this in
-section~\ref{sec:pareto_tail}.
-
-Thirdly, figure~\ref{fig:sectors_parameters} shows the values of $K$
-in the ``Energy'' and ``Information Technology'' sectors are on
-average larger than those in the ``Consumer Staples'' sector. A larger
-value of $K$ means, for a given value of $p$, i.e. a given probability
-of losses, large losses are more probable. Thus one can conclude that
-these two sectors are considerablly riskier than the ``Consumer
-Staples'' sector. This is of course a confirmation of one's economic
+huge variations of $K^\alpha$, as shown in the 2nd row of
+Figure~\ref{fig:sectors_parameters}.
+\par
+Secondly, it appears that there is positive dependence between the  values of $\alpha$ and $K$. 
+This prompts Pareto tails of the return
+distributions. For more explanations, see
+Section~\ref{sec:pareto_tail}.
+\par
+Thirdly, Figure~\ref{fig:sectors_parameters} shows that, on
+average, the values of $K$
+in the ``Energy'' and ``Information Technology'' sectors are  larger than those in the ``Consumer Staples'' sector. 
+For a given loss \pro y, 
+a larger
+value of $K$ implies that large losses are more probable. Thus one can conclude that
+these two sectors are considerably riskier than the ``Consumer
+Staples'' sector. {\red Let Casper decide about this sentence.} This is of course a confirmation of one's economic
 instinct.
-
-Yet another indication from figure \ref{fig:sectors_parameters} is
+\par
+Yet another indication from Figure \ref{fig:sectors_parameters} is
 that, while the ``Energy'' and the ``Information Technology'' sectors
-are similar in riskiness, the correlation between $\alpha$ and $K$ is
-stronger in ``Energy''. As discussed in the last section, moving along
+are similar in riskiness, the dependence  between $\alpha$ and $K$ is
+stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, moving along
 a curve of equal preference in the direction of increasing $\alpha$, $K$ also
@@
 
-\subsection{A test of equal tail indices based on Hill estimator}
-Suppose we have two weakly dependent series $X_1, \dots, X_n \sim F_X$ and
-$Y_1, \dots, Y_n \sim F_Y$. Both $F_X$ and $F_Y$ have power-law tails
-with indices $\alpha_X$ and $\alpha_Y$, respectively. Assume also
-$X_i$ is independent of $Y_j$ for all $1 \leq i \leq j \leq n$.
+\subsection{A test for equal tail-indices based on Hill estimation}
+Suppose we have two independent strictly stationary positive series $X_1, \dots, X_n $ and
+$Y_1, \dots, Y_n$ with corresponding \df s $F_X$ and $F_Y$ that have power-law tails
+with indices $\alpha_X$ and $\alpha_Y$, respectively. 
 From \eqref{eq:2} one can deduce
@@
 \end{equation}
-where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimates of
-$\alpha_X$ and $\alpha_Y$. Then it follows from continuous mapping
+where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimators of
+$\alpha_X$ and $\alpha_Y$; we suppress their dependence on $k$. Then it follows from the continuous mapping
 theorem
@@
   \overset{d}{\to}
-  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)
+  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)\,.
 \end{equation}
-The weak convergence of \eqref{eq:x1} allows one to construct a test
-of the equality $\alpha_X = \alpha_Y$. As the null hypothesis, assume
-$\alpha_X = \alpha_Y$. Under this hypothesis, \eqref{eq:x1} means the
-test statistic $\hat \alpha_X - \hat \alpha_Y \overset{d}{\to} N(0,
-\alpha_X^2 + \alpha_Y^2)$. 
+This relation allows one to construct an \asy\ test under 
+the null hypothesis $\alpha_X = \alpha_Y$ and with test statistic $\hat \alpha_X - \hat \alpha_Y$.
 We apply this test to the equities in the ``Energy'',
-``Consumder Staples'' and ``Information Technology''
-sectors of the S\&P 500 index. The results are shown in
-\ref{fig:PairTest}.
+``Consumer Staples'' and ``Information Technology''
+sectors of the S\&P 500 index. The results are shown in the top row of
+Figure~\ref{fig:PairTest}.
 \begin{figure}[htb!]
@@
   \end{minipage}
-  \caption{
-    {\em 1st row}: Test for changing tail index using Hill estimator based on
-    concatenated series in the ``Energy'' and ``Consumer Staples''
-    sectors of S\&P 500. The green, blue and red colors represent,
-    respectively, the cases when the 85\%, 90\% and 95\% quantiles of
-    the limit distribution of the test statistic is exceeded. 
-    Grey points stand for pairs for which the test statistic is below
-    the 85\% \asy\ quantile;
-    {\em 2nd row:} Test for changing tail index or scale parameter
-    using Hoga's test based on the same concatenated series. black points represent
+  \caption{{\em Top row}: Test for pairwise equality of tail-indices in the 
+    ``Energy'', ``Consumer Staples'' and ``Information Technology'' 
+    sectors of S\&P 500. The test statistic in \eqref{eq:x1} is based on Hill estimates 
+    of $\alpha_X$ and $\alpha_Y$. 
+    The green, blue and red points correspond to pairs of stock in a sector
+    when the test statistic is outside the intervals $[q_{0.075},q_{0.925}]$,
+ $[q_{0.05},q_{0.95}]$,  $[q_{0.025},q_{0.975}]$, repectively, where $q_p$ is the $p$-quantile of the limiting
+$N(0,\alpha_X^2+\alpha_Y^2)$-\ds\ of the test statistic in \eqref{eq:x1}. 
+    Grey points stand for pairs for which the test statistic is inside $[q_{0.075},q_{0.925}]$. 
+    {\em Bottom row:} Test for changing tail index or scale parameter     
+    using Hoga's test based on concatenated series of pairs of stocks. The green, blue and red points 
+correspond to pairs of stock in a sector 
+    when the test statistic $T_n$ is exceeds the $85\%$-, $90\%$-, $95\%$-quantile of the limit \ds .  
+  Grey points stand for pairs for which the test statistic is below the \asy\ $85\%$-quantile. Black points represent
     pairs for which the computation of $T_n$ fails for given precision  
     requirements and time limits.
-    The same number of upper-order statistics are used for both tests.
-  }
+    The same number of upper order statistics are used for both tests.}
   \label{fig:PairTest} 
 \end{figure}
-One can see from the figure that tail indices of equities in the
-``Energy'' or ``Information Technology'' sectors vary more than do
-their counterparts in the ``Consumer Staples'' sector, as the null
-hypothesis is rejected more often for members of these two sectors.
-Moreover, these figures suggest that the test based on Hill estimator
-is more powerful in distinguishing differences between tail indices,
-as it results in more rejections for the ``Energy'' and the
-``Information Technology'' sectors than does the other test. We
-argue this lack of power of the latter can be the result of slow
-convergence of the distribution of its test statistic to the
-asymptotic distribution. We elaborate on this in section~\ref{sec:Hoga}.
-
-Nevertheless, one should bear in mind that the weak convergence
-\eqref{eq:x3} is valid only on condition that both the $X_i$ and $Y_i$
-series are independent of each other, which is generally untrue for
+It indicates that tail indices of equities in the
+``Energy'' or ``Information Technology'' sectors are more variable 
+in the ``Consumer Staples'' sector, as the null
+hypothesis is rejected more often for members of these two former sectors.
+Moreover, these figures suggest that the test based on the Hill estimator
+is quite powerful in distinguishing between tail indices. In contrast to the test presented in Section~\ref{sec:Hoga}
+the present test results in  more rejections for the ``Energy'' and the
+``Information Technology'' sectors.
+\par
+As a caution, one should bear in mind that 
+\eqref{eq:x3} is valid on condition that the $X$- and $Y$-
+series are independent of each other (or weakly dependent on each other), which is generally untrue for
 two return series in the same market.
-}
-
-\subsection{Tests for a change in the extreme tail}\label{sec:Hoga}
-To the best of our knowledge, there do not exist published statistical
-tests for comparing the tail indices or scale parameters of two or
-more dependent \ts\ in a direct way. In addition to the test described
-in the previous section, we borrow a test from
+
+\subsection{A tests for a change in the extreme tail}\label{sec:Hoga}
+Here we apply a test from
 a recent paper by Hoga \cite{hoga:2016}. This test has been developed
@@
 \par
-To check whether any two series share the same tail index and scale parameter we concatenate
+To check whether any pair of stocks shares the same tail index and scale parameter we concatenate
 any two series and apply the aforementioned test on the concatenated series. For the ``Energy'' and the
-``Consumer Staples'' sector we summarize the results in Figure~\ref{fig:PairTest}.
+``Consumer Staples'' sector we summarize the results in the bottom row of Figure~\ref{fig:PairTest}.
 These graphs show that the ``null
@@
     in the subsamples. For an interpretation of the colored bullets,
-    see Figure~\ref{fig:PairTest}. The graph shows the limited power
+    see the bottom row of Figure~\ref{fig:PairTest}. The graph shows the limited power
     of the test. In particular, if both degrees of freedom are
-    relatively large it loses the capablity of distinguishing the \ds\ .
+    relatively large it loses the capablity of distinguishing between the \ds s.
     {\em Right}: Comparison of the \asy\ distribution of the test statistic
@@
 \eeam
-for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^\alpha]=1$.  
-In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2$ and $B_t=\alpha_0$ to achieve
-\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016})
+for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^{\alpha/2}]=1$.  
+In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2+\beta_1$ and $B_t=\alpha_0$ to achieve
+\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016}, p.~275)
 it follows that
@@
 Thus both return components $X_{i,t}$ have the form of a univariate \sv\ model $X_{i,t}=\sigma_{i,t}Z_{i,t}$ 
-with non-negative volatility $\sigma_{i,t}$ and an iid bivariate noise \seq\ $(\bfZ_t)$ with zero mean and unite variance components.
+with non-negative volatility $\sigma_{i,t}$ and an iid bivariate noise \seq\ $(\bfZ_t)$ with zero mean and unit variance components.
 We also have the specification
@@
 Similar remarks apply when one considers a CCC model in general dimension. Of course, one may argue that the latter models
-are not natural since they do not allow for a linear relationship between all squared volatilities
+are degenerate since they do not allow for a linear relationship between all squared volatilities
 on a given day.
@@
   \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
-  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
+  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
 \eeao
 where $\delta $ and $v$ are positive constants, and $b\ge 0$.
-{\blue 
- Here $v$ can be thought of as the certainty payoff equivalent to the
- risky payoff $C$; $\delta$ tunes the threshold of {\em
+  Here $v$ can be thought of as the {\em certainty payoff} equivalent to the
+ risky payoff $C$; $\delta$ tunes the {\em threshold of 
    disappointment} in proportion to $v$; $b$ determines the extra
@@
 \cite{routledge2010generalized}.
-}
+\par
 An agent guided by the utility function $u$ will seek to maximize the
@@
 \beam\label{eq:u}
-u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
+u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>-1\,. 
 \eeam
-In practice it is more often to observe $\gamma > 1$ {\blue Let's
-consult Casper for references.} Therefore we adopt a more convenient
-notation in this case
-\beao
-u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
-\eeao
-
+%In practice it is more often to observe $\gamma > 1$ {\blue Let's
+%consult Casper for references.} Therefore we adopt a more convenient
 For the sake of argument, we assume that an investor initially has one
@@
 \beao
-\wt u(\phi) &=& \E [u(C)] + b\, \E \big[u(C)\1(C < \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
+\wt u(\phi,F_X) &=& \E [u(C)] + b\, \E \big[u(C)\1(C \le  \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
 \eeao
@@
 \beao
-  q = %q(r, \phi) &:=& \ln \left( {
+  q = %q(r, \phi) &:=& \log \left( {
       %\delta v - (1 - \phi) e^r
@@
 \eeao
-Note that $C < \delta v$ \fif\ $X < q$. 
+Note that $C \le \delta v$ \fif\ $X \le  q$. 
 \par
@@
 riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
-$q > r $. For given $b,\delta,v$, $\wt u$ is a functional of $\phi$ and $F_X$,
-$
-\wt u=\wt u (F_X, \phi)\,,
-$
-and we assume that
+$q > r $. For given $b,\delta,v$, $\wt u$ is only a functional of $\phi$ and $F_X$,
+$\wt u=\wt u (F_X, \phi)$.
+We assume that
 \beao
@@
 \subsection{Pareto-distributed returns}\label{sec:pareto_tail}
+Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
+we assume the following toy model. 
 We consider the case when $X$ has a two-sided Pareto \ds\ given by
@@
 \eeam
-where 
+where {\red where needed?}
 $\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
 \beam\label{eq:xxie1.0}
- \wt u(F_X,\phi)\nonumber
+ \wt u(\phi,F_X)\nonumber
   &=&
@@
   u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
-  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
+  {1 + b \1(x \le  q) \over (K' + x)^{\beta + 1}} \,dx 
   - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
 \eeam
-{\blue
-\begin{remark}\label{thrm:I}
- Assume the two-sided Pareto model \eqref{eq:pp} and that the utility
+\par
+We observe the following property which also applies to \eqref{eq:xxie1.0}. 
+\ble\label{thrm:I}
+ Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
  \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
@@
    K} < 0$.
-\end{remark}
-}
+\ele
 The proof is given in Appendix~\ref{sec:thrmI_proof}. 
-\par
-We illustrate the results of remark~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
-\fct\ \eqref{eq:u}.
-For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
-quantity is denoted by $I(\alpha, K)$.
 \par
@@
 ``Information Technology'' sectors of the S\&P 500 index; see 
-Figure \ref{fig:sectors_parameters}.
+Figure~\ref{fig:sectors_parameters}.
+\par
+We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
+\fct\ \eqref{eq:u}.
+For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
+quantity is denoted by $I(\alpha, K)$.
 \par
@@
 This is shown in 
-Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
-\begin{figure}
+Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
+\begin{figure}[htb!]
   \begin{minipage}{0.5\linewidth}
@@
     \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
-    the utility $u$ from \eqref{eq:u}
-    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$ (left) and $\xi = 4$
+    the utility \fct\ $u$ from \eqref{eq:u}
+    for $\xi = 1/2$ (left) and $\xi = 4$
     (right).
@@
 \subsection{Student's t-distributed equity return}
+{\red This section is incomplete unless we can show results for $b>0$. Otherwise, we can do everything for $b=0$.} 
 It is a common practice to use Student's t-distribution to model the
@@
 find out what implications this distribution has when it is combined
-with the PDA preference. Formally we assume
+with the PDA preference. Formally we assume {\red instead of this function one can take a symmetric \ds\ with tails
+$\P(X>x)=0.5 (\kappa/(x+\kappa))^{\alpha}$. If the symmetry causes another behavior than in the previous section this would be interesting.}
+{\blue our tail index is always $\alpha$, not $\nu$.}
 \[
@@
 We first consider the case where the utility function is simply
-$u(\cdot) = \ln(\cdot)$. This corresponds to the limiting case of
+$u(\cdot) = \log(\cdot)$. This corresponds to the limiting case of
 \eqref{eq:u} as $\gamma \to 0$.
@@
 $\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
-Its integrand becomes $\ln[C(x, \phi) C(-x, \phi)]$.
-Now observe $\ln(C(x, \phi)C(-x, \phi))$ is an increasing function of
+Its integrand becomes $\log[C(x, \phi) C(-x, \phi)]$.
+Now observe $\log(C(x, \phi)C(-x, \phi))$ is an increasing function of
 $x$:
@@
 Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
-\ref{lemma:II},
+\ref{lemma:II}, {\red need to look at $\pd \wt u_{\rm max} \over \pd \alpha$?} 
 ${\pd \wt u(F, \phi)) \over \pd \nu} < 0$ for all $\phi$. Hence all
@@
       -
-      \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
+      \nu c(\nu) (x^2 + \nu) \log(1 + x^2/\nu)
       \over
@@
   \[
-  \Psi(x) = {d \ln[\Gamma(x)] \over d x}
+  \Psi(x) = {d \log[\Gamma(x)] \over d x}
   \]
@@
       \right] - 1 \\
-      &=& \ln(4) - \ln(e) \\
+      &=& \log(4) - \log(e) \\
       &>& 0
@@
   \[
-  \ln(y e^{-y}) = \ln(y) - y
+  \log(y e^{-y}) = \log(y) - y
   \]
@@
     {\pd P \over \pd x}(x, \nu) &=&
-    2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \ln\left(
+    2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \log\left(
       1 + {x \over \nu}
@@
 
-\section{Proof of Remark \ref{thrm:I}}
+\section{Proof of Lemma~\ref{thrm:I}}
 \label{sec:thrmI_proof}
@@
       \alpha
-      \ln\left(
+      \log\left(
         1 - {x \over K}

### 2017-04-29 papers/FX/draft.tex
Started writing the proof of the equal tail case.

@@
 \par
-We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
-\fct\ \eqref{eq:u}.
+We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto}
+for the power-utility \fct\ \eqref{eq:u}.
+A particularly interesting special case is when the left and the right
+tails have equal parameters, i.e. when $\alpha = \beta$ and $K =
+K'$. In this case \eqref{eq:pareto} becomes
+\begin{eqnarray*}
+  \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
+  \int_{0}^\infty {
+    u(C(x)) \left[
+      1 - {b \over 1 + b}\1_{\{x > q\}}
+      \right]
+    + u(C(-x))
+    \over
+    (K + x)^{\alpha + 1}
+  } dx
+  - b u(\delta v) F_X(q)
+\end{eqnarray*}
+A detailed analysis of the montonicity of
+\[
+u(C(x)) \left[
+  1 - {b \over 1 + b}\1_{\{x > q\}}
+  \right]
++ u(C(-x))
+\]
+which comprises lemma \ref{lemma:III}, allows us to apply lemma
+\ref{lemma:I} and conclude the following
+\begin{itemize}
+\item I am a good guy
+\item So are you
+\end{itemize}
+
 For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
@@
 
+\section{Lemma~\ref{lemma:III}}
+\begin{lemma}\label{lemma:III}
+
+\end{lemma}
+\begin{proof}
+  This is obvious.
+\end{proof}
+
 \bibliographystyle{unsrt}

### 2017-04-29 papers/FX/draft.tex
The proof is very tedious!

@@
 \begin{lemma}\label{lemma:III}
-
+Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:u} and \eqref{eq:xxie1} respectively. Define
+\[
+U(x) =
+u(C(x)) \left[
+  1 - {b \over 1 + b}\1_{\{x > q\}}
+  \right]
++ u(C(-x))
+\quad
+x \geq 0
+\]
+\begin{enumerate}
+\item  Hello
+\end{enumerate}
 \end{lemma}
 \begin{proof}
-  This is obvious.
+  It is convenient to re-write $U(x)$ as
+  \[
+  U(x) = 
+  -{\hat\phi^{-\xi} \over \xi} \left\{
+    \underbrace{
+    \left(1 - {b \over 1 + b}\1_{\{x > q\}}\right)
+    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^x\right]^{-\xi}
+    +
+    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^{-x}\right]^{-\xi}
+  }_{U'(x)}
+\right\}
+\]
+  Thus the monotonicity of $U(x)$ is the opposite of $U'(x)$. For
+  convenience of writing, let $a = (1 - \hat\phi)e^r/\hat\phi$.
+  First of all, we find the conditions on which $U'(x)$ is monotonic.
+  Direct differentiation yields
+  \begin{eqnarray*}
+    {\pd U'(x) \over \pd x}
+    &=&
+    -{
+      (a + e^x)^{-\xi - 1} \xi e^x
+      \over
+      1 + b \1_{\{x > q\}}
+    } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}
+    \quad
+    x \neq q
+  \end{eqnarray*}
+  ${\pd U'(x) \over \pd x} = 0$ is equivalent to
+  \begin{eqnarray*}
+    {a + y \over a y + 1} &=& 
+    \left( \1_{\{y \leq e^q\}} + c \1_{\{y > e^q\}} \right)
+    y^{1 -\xi \over 1+\xi} \\
+    \underbrace{\log(a + y) - \log(a y + 1)}_{f(y)} &=& \underbrace{{1
+        -\xi \over 1+\xi} \log(y)
+      + \log(c)\1_{\{y > e^q\}} }_{g(y)} 
+  \end{eqnarray*}
+  where we have defined $y = e^x$, $c = (1 + b)^{-1/(1 + \xi)}$ and
+  \begin{eqnarray*}
+    f(y) &=& \log(a + y) - \log(a y + 1) \\
+    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > e^q\}}
+  \end{eqnarray*}
+  Observe
+  \begin{eqnarray}
+    f(y) - g(y) &=& \log\left(
+      {a + y \over a y + 1}
+      y^{\xi - 1 \over \xi + 1}
+    \right) - \log(c)\1_{\{y > e^q\}} \label{eq:f-g}
+  \end{eqnarray}
+  and when $y \neq e^q$,
+  \begin{eqnarray}
+    {\pd f(y) \over \pd y} &=& {1 - a^2 \over (a + y) (a y + 1)}
+    \label{eq:fd}\\
+    {\pd g(y) \over \pd y} &=& {1 - \xi \over (1 + \xi) y}
+    \label{eq:gd}
+  \end{eqnarray}
+  Furthermore
+  \begin{equation}
+    \label{eq:f-g-der}
+    {\pd (f -g)(y) \over \pd y} = {
+      a (\xi - 1) y^2 + 2 (\xi - a^2) y + a(\xi - 1)
+      \over
+      (1 + \xi) (a + y) (a y + 1) y
+    }
+    \quad
+    y \neq e^q
+  \end{equation}
+  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots
+  \begin{eqnarray*}
+    y_\pm &=& {
+      a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
+      \over
+      a (\xi - 1)
+    }
+  \end{eqnarray*}
+  \begin{enumerate}
+  \item If $\min\{1, \xi\} < a < \max\{1, \xi\}$, $y_\pm$ are not
+    real, ${\pd (f -g)(y) \over \pd y} = 0$ has no solution on $(1,
+    \infty)\setminus \{e^q\}$; $(f -g)(y)$ is monotonic on both
+    $(1, e^q)$ and $(e^q, \infty)$.
+    \begin{enumerate}
+    \item If in addition $\xi > 1$, i.e. $1 < a < \xi$, $(f - g)(y)$
+      is monotonically increasing on both
+      $(1, e^q)$ and $(e^q, \infty)$, because the coefficient of the
+      $y^2$ term of the numerator of \eqref{eq:f-g-der}, i.e. $a (\xi
+      - 1)$ is positive.
+      \begin{itemize}
+      \item If $f(e^q) > g(e^q)$, then clearly $f(y) = g(y)$ has no
+        solution on $(e^q, \infty)$; we also note $f(1) = g(1) =
+        0$. Thus $f(y) = g(y)$ has no solution on $(1, e^q)$ either.
+        Now it can be concluded that $U'(\cdot)$ is monotonic on
+        $(1, q)$ and $(q, \infty)$. To determine the monotonicity of
+        $U'(\cdot)$, it remains to check the signs of
+        ${\pd U'(x) \over \pd x}$  on  $(0, q)$ and on $(q, \infty)$
+        as well as the values of $\lim_{x \uparrow q} U'(x)$ and
+        $\lim_{x \downarrow q} U'(x)$. Obviously
+        \begin{eqnarray*}
+          \left.{\pd U'(x) \over \pd x}\right|_{x=0} &=& 0 \\
+        \end{eqnarray*}
+        For a small $\epsilon > 0$, the sign of ${\pd U'(x) \over \pd
+          x}$ on $(0, \epsilon)$ is thus the same as
+        $\left.{\pd^2 U'(x) \over \pd x^2}\right|_{x=0}$:
+        \[
+        \left.{\pd^2 U'(x) \over \pd x^2}\right|_{x=0} =
+        2 (a + 1)^{-\xi - 2} (\xi - a) > 0
+        \]
+      \end{itemize}
+
+      % conversely, $f(e^q) < g(e^q)$ implies $f(y) = g(y)$ must have a
+      % solution on $(e^q, \infty)$, since \eqref{eq:f-g} tells us
+      % $(f-g)(y) \to \infty$ as $y \to \infty$.
+
+    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, $(f - g)(y)$
+      is monotonically decreasing.
+    \end{enumerate}
+
+  % \item If $a < \min\{1, \xi\}$ or $a > \max\{1, \xi\})$, ${\pd (f
+  %     -g)(y) \over \pd y} = 0$ has 2 real solutions.
+  % \end{enumerate}
+
+  % \begin{enumerate}
+  % \item If $a < 1$ and $\xi > 1$
+  %   \begin{enumerate}
+  %   \item when $x > q$, $U'(x)$ being monotonic is equivalent to ${\pd U'(x)
+  %       \over \pd x}$ having no zero on $(q, \infty)$.
+
+  %     \begin{enumerate}
+  %     \item If $f (e^q) > g(e^q)$ it is clear $f(y) = g(y)$
+  %       has no solution on $(e^q, \infty)$.
+  %     \item If $f (e^q) <  g(e^q)$ it is also clear $f(y) = g(y)$ must have
+  %       one solution on $(e^q, \infty)$.
+  %     \end{enumerate}
+  %   \end{enumerate}
+
+  %   \item If $a > 1$ and $\xi < 1$
+  %     \begin{enumerate}
+  %     \item when $x > q$, the same calculation that leads to
+  %       \eqref{eq:fd} and \eqref{eq:gd} gives ${\pd f(y) \over \pd y}
+  %       < 0$ and ${\pd g(y) \over \pd y} > 0$. So
+  %     \begin{enumerate}
+  %     \item If $f (e^q) >  g(e^q)$, $f(y) = g(y)$ must have
+  %       one solution on $(e^q, \infty)$.
+  %     \item If $f (e^q) < g(e^q)$, $f(y) = g(y)$ has no solution on $(e^q, \infty)$.
+  %     \end{enumerate}
+  %     \end{enumerate}
+
+  %     \item If $a < 1$ and $\xi < 1$
+  %       \begin{enumerate}
+  %       \item when $x > q$
+  %         \begin{enumerate}
+  %         \item If $f (e^q) -  g(e^q) > 0$, we observe $f(y) - g(y)
+  %           \to -\infty$ as $y \to \infty$. So by continuity $f(y) -
+  %           g(y) = 0$ must have at least one solution on $(e^q, \infty)$.
+  %         \end{enumerate}
+  %       \end{enumerate}
+  %     \item If $a > 1$ and $\xi > 1$
+  %       \begin{enumerate}
+  %       \item when $x > q$, both $f(y)$ and $g(y)$ decreases. Observe
+  %         In this case, if $f(e^q) > g(e^q)$, there is no solution 
+  %       \end{enumerate}
+  \end{enumerate}
 \end{proof}

### 2017-04-30 papers/FX/draft.tex
Added to the proof of double sided pareto tail same tail parameters.
It appears the combined utility function is never monotonic.

@@
   u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
-  {1 + b \1(x \le  q) \over (K' + x)^{\beta + 1}} \,dx 
+  {1 + b \1(x <  q) \over (K' + x)^{\beta + 1}} \,dx 
   - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
@@
     u(C(x)) \left[
-      1 - {b \over 1 + b}\1_{\{x > q\}}
+      1 - {b \over 1 + b}\1_{\{x \ge q\}}
       \right]
@@
 u(C(x)) \left[
-  1 - {b \over 1 + b}\1_{\{x > q\}}
+  1 - {b \over 1 + b}\1_{\{x \ge q\}}
   \right]
@@
     \underbrace{
-    \left(1 - {b \over 1 + b}\1_{\{x > q\}}\right)
+    \left(1 - {b \over 1 + b}\1_{\{x \ge q\}}\right)
     \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^x\right]^{-\xi}
@@
     {a + y \over a y + 1} &=& 
-    \left( \1_{\{y \leq e^q\}} + c \1_{\{y > e^q\}} \right)
+    \left( \1_{\{y < e^q\}} + c \1_{\{y > q\}} \right)
     y^{1 -\xi \over 1+\xi} \\
@@
         -\xi \over 1+\xi} \log(y)
-      + \log(c)\1_{\{y > e^q\}} }_{g(y)} 
+      + \log(c)\1_{\{y > q\}} }_{g(y)} 
   \end{eqnarray*}
@@
     f(y) &=& \log(a + y) - \log(a y + 1) \\
-    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > e^q\}}
+    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > q\}}
   \end{eqnarray*}
@@
   \end{equation}
-  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots
+  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots when $\xi \ne 1$
   \begin{eqnarray*}
@@
       $y^2$ term of the numerator of \eqref{eq:f-g-der}, i.e. $a (\xi
-      - 1)$ is positive.
-      \begin{itemize}
-      \item If $f(e^q) > g(e^q)$, then clearly $f(y) = g(y)$ has no
-        solution on $(e^q, \infty)$; we also note $f(1) = g(1) =
-        0$. Thus $f(y) = g(y)$ has no solution on $(1, e^q)$ either.
+      - 1)$ is positive. We note $f(1) = g(1) = 0$; thus on $(1, e^q)$,
+      $f(y) > g(y)$ and hence there is no solution to $f(y) = g(y)$ on
+      this interval.
+      \begin{enumerate}
+      \item If $ \lim_{y \to e^q}f(x) \ge \lim_{y \downarrow e^q} g(y)$,
+        then clearly $f(y) = g(y)$ has no
+        solution on $(e^q, \infty)$ either.
         Now it can be concluded that $U'(\cdot)$ is monotonic on
-        $(1, q)$ and $(q, \infty)$. To determine the monotonicity of
+        $(0, q)$ and $(q, \infty)$. To determine the monotonicity of
         $U'(\cdot)$, it remains to check the signs of
@@
         \]
+        Thus ${\pd U'(x) \over \pd x} > 0$ for $x \in (0, q)$.
+        To find out the sign of
+        ${\pd U'(x) \over \pd x}$ on the interval $(q, \infty)$,
+        one can look into the sign of ${\pd U'(x) \over \pd x}$
+        as $x \to \infty$. For
+        this purpose, substitute $x'$ for $e^{-x}$ and re-write
+        ${\pd U'(x) \over \pd x}$ as
+        \begin{eqnarray*}
+          {1 \over \xi}
+          {\pd U'(x) \over \pd x} &=& {x' \over (a + x')^{\xi + 1}}
+          - {x'^\xi \over (a x' + 1)^{\xi + 1} (1 + b)}
+        \end{eqnarray*}
+        Taylor-expanding the last expression at $x'=0$ gives
+        \[
+          {1 \over \xi}
+          {\pd U'(x) \over \pd x}
+         = {x' \over a^{\xi + 1}} + o(x')
+        \]
+        From this expression we can see that
+        ${\pd U'(x) \over \pd x} > 0$ for
+        sufficiently large $x$; then by the proven monotonicity,
+        ${\pd U'(x) \over \pd x} > 0$ for all $x \in (q, \infty)$.
+        It remains to compare $\lim_{x\uparrow q}U'(x)$ with
+        $\lim_{x\downarrow q}U'(x)$. Clearly
+        \[
+        \lim_{x\downarrow q}U'(x) - \lim_{x\uparrow q}U'(x) = - {
+          (a + e^q)^{-\xi} b
+          \over
+          1+ b
+        } < 0
+        \]
+        So $U'(x)$ and hence $U(x)$ as a whole is not monotonic on $(0, \infty)$.
+
+      \item If $ \lim_{y \to e^q}f(x) < \lim_{y \downarrow e^q} g(y)$,
+        $(f - g)(y)$ must have a solution on
+        $(e^q, \infty)$ since $(f-g)(y)$ is monotonically
+        increasing. So $U'(\cdot)$ is not monotonic in this case.
+      \end{enumerate}
+
+    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar argument as in the
+      previous case, $(f - g)(y)$ is monotonically decreasing on both
+      $(1, e^q)$ and $(e^q, \infty)$. From $f(1) = g(1) = 0$, it again
+      follows that $U'(x)$ is monotonic on $(0, q)$.
+      \begin{itemize}
+      \item
+        If $ \lim_{y \to e^q} f(y) > \lim_{y \downarrow e^q} g(y)$,
+        then there must be a solution to $f(y) = g(y)$ on $(e^q,
+        \infty)$. $U'(x)$ is not monotonic in this case.
+
+      \item
+        If $ \lim_{y \to e^q} f(y) \le \lim_{y \downarrow e^q} g(y)$,
+        there is no solution to $f(y) = g(y)$ on $(e^q,
+        \infty)$ either. $U'(x)$ is also monotonic on
+        $(e^q, \infty)$. By the calculation as before, it can be seen
+        ${\pd U'(x) \over \pd x} < 0$ for positive $x$ that is
+        sufficiently close to 0; and
+        ${\pd U'(x) \over \pd x} > 0$ for sufficiently large $x$.
+        So $U'(x)$ is not monotonic in this case.
       \end{itemize}
-
-      % conversely, $f(e^q) < g(e^q)$ implies $f(y) = g(y)$ must have a
-      % solution on $(e^q, \infty)$, since \eqref{eq:f-g} tells us
-      % $(f-g)(y) \to \infty$ as $y \to \infty$.
-
-    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, $(f - g)(y)$
-      is monotonically decreasing.
     \end{enumerate}
-
-  % \item If $a < \min\{1, \xi\}$ or $a > \max\{1, \xi\})$, ${\pd (f
-  %     -g)(y) \over \pd y} = 0$ has 2 real solutions.
-  % \end{enumerate}
-
-  % \begin{enumerate}
-  % \item If $a < 1$ and $\xi > 1$
-  %   \begin{enumerate}
-  %   \item when $x > q$, $U'(x)$ being monotonic is equivalent to ${\pd U'(x)
-  %       \over \pd x}$ having no zero on $(q, \infty)$.
-
-  %     \begin{enumerate}
-  %     \item If $f (e^q) > g(e^q)$ it is clear $f(y) = g(y)$
-  %       has no solution on $(e^q, \infty)$.
-  %     \item If $f (e^q) <  g(e^q)$ it is also clear $f(y) = g(y)$ must have
-  %       one solution on $(e^q, \infty)$.
-  %     \end{enumerate}
-  %   \end{enumerate}
-
-  %   \item If $a > 1$ and $\xi < 1$
-  %     \begin{enumerate}
-  %     \item when $x > q$, the same calculation that leads to
-  %       \eqref{eq:fd} and \eqref{eq:gd} gives ${\pd f(y) \over \pd y}
-  %       < 0$ and ${\pd g(y) \over \pd y} > 0$. So
-  %     \begin{enumerate}
-  %     \item If $f (e^q) >  g(e^q)$, $f(y) = g(y)$ must have
-  %       one solution on $(e^q, \infty)$.
-  %     \item If $f (e^q) < g(e^q)$, $f(y) = g(y)$ has no solution on $(e^q, \infty)$.
-  %     \end{enumerate}
-  %     \end{enumerate}
-
-  %     \item If $a < 1$ and $\xi < 1$
-  %       \begin{enumerate}
-  %       \item when $x > q$
-  %         \begin{enumerate}
-  %         \item If $f (e^q) -  g(e^q) > 0$, we observe $f(y) - g(y)
-  %           \to -\infty$ as $y \to \infty$. So by continuity $f(y) -
-  %           g(y) = 0$ must have at least one solution on $(e^q, \infty)$.
-  %         \end{enumerate}
-  %       \end{enumerate}
-  %     \item If $a > 1$ and $\xi > 1$
-  %       \begin{enumerate}
-  %       \item when $x > q$, both $f(y)$ and $g(y)$ decreases. Observe
-  %         In this case, if $f(e^q) > g(e^q)$, there is no solution 
-  %       \end{enumerate}
+  \item If $a < \min\{1, \xi\}$ and $\xi > 1$, i.e. $a < 1 < \xi$, it
+    is clear
+    \[
+    y_- = {
+      a^2 - \xi - \sqrt{(a^2 - 1) (a^2 - \xi^2)}
+      \over
+      a (\xi - 1)
+    } < 0
+    \]
+    It remains to compare $y_+$ with $1$ to determine whether $f(y) =
+    g(y)$ has a solution. Assume $y_+ > 1$. Then
+    \begin{eqnarray*}
+      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1) \\
+      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0
+    \end{eqnarray*}
+    This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
+    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotonically increasing
+    on both $(1, e^q)$ and $(e^q, \infty)$. Following the same
+    analysis as in the case (1.a.i), one can see $U'(\cdot)$ is not
+    monotonic on $(0, \infty)$.
+
+  \item If $a < \min\{1, \xi\}$ and $\xi < 1$, i.e. $a < \xi < 1$, it
+    is clear $y_- > 0$ and $y_+ < y_-$. Moreover, $y_- > 1$ is equivalent
+    to
+    \[
+    \sqrt{(a^2 - 1) (a^2 - \xi^2)} > a^2 - \xi a + a - \xi = (a + 1) (a - \xi)
+    \]
+    The last inequality is obviously true in this case. So $y_- >
+    1$. ${\pd (f-g)(y) \over \pd y}$ has a maximum at
+    $y_- \in (1, \infty)$. We know $(f - g)(y_-)$ is a maximum because
+    \eqref{eq:f-g-der} shows that, for $y > y_-$,
+    ${\pd (f-g)(y) \over \pd y} < 0$.
+
+    If $(f-g)(y_-) < 0$, $(f-g)(y) = 0$ has no solution on $(1,
+    \infty)$. Hence $U'(\cdot)$ is monotonic on both $(0, q)$ and $(q,
+    \infty)$. The same analysis as in the case (1.a.i) shows $U'(\cdot)$
+    is not monotonic on $(0, \infty)$.
+
+    If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
+    \infty)$. So $U'(\cdot)$ is not monotonic.
   \end{enumerate}

### 2017-05-02 papers/FX/draft.tex
Removed a paragraph that I think Thomas would want to remove.
The case when b is not zero is not monotonic.

@@
 \end{eqnarray*}
-A detailed analysis of the montonicity of
+First of all we note that the function
 \[
@@
 \]
+is not monotonic. To see this we may re-write the above as
+\begin{equation}
+  -{\phi^{-\xi} \over \xi} \left\{
+  \underbrace{
+    \left(1 - {b \over 1 + b}\1_{\{x \ge q\}}\right)
+    \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
+    +
+    \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
+  }_{U(x)} \right\}
+  \label{eq:frg}
+\end{equation}
+Direct computation gives
+\begin{eqnarray*}
+  {\pd U(x) \over \pd x}
+  &=&
+  -{
+    (a + e^x)^{-\xi - 1} \xi e^x
+    \over
+    1 + b \1_{\{x > q\}}
+  } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}
+  \quad x \neq q
+\end{eqnarray*}
+where $a = (1 - \phi) e^r/\phi$. For $x \ge q$, we have
+\begin{eqnarray*}
+  {1 \over \xi}
+  {\pd U(x) \over \pd x} &=& {x' \over (a + x')^{\xi + 1}}
+  - {x'^\xi \over (a x' + 1)^{\xi + 1} (1 + b)}
+\end{eqnarray*}
+where we have substituted $x'$ for $e^{-x}$. When $x$ is very large
+i.e. when $x'$ is very close to 0, it is useful to expand the last
+expression at $x' = 0$. This yields
+\[
+  {1 \over \xi}
+  {\pd U(x) \over \pd x}
+  = {x' \over a^{\xi + 1}} + o(x')
+\]
+So we see that when $x$ is sufficiently large
+${\pd U(x) \over \pd x} > 0$. On the other hand, at the discontinuity
+point $x = q$, we have
+\[
+\lim_{x\downarrow q}U(x) - \lim_{x\uparrow q}U(x) = - {
+  (a + e^q)^{-\xi} b
+  \over
+  1+ b
+} < 0
+\]
+So $U(x)$ and hence \eqref{eq:frg} is not monotonic.
+
 which comprises lemma \ref{lemma:III}, allows us to apply lemma
@@
   U(x) = 
-  -{\hat\phi^{-\xi} \over \xi} \left\{
+  -{\phi^{-\xi} \over \xi} \left\{
     \underbrace{
     \left(1 - {b \over 1 + b}\1_{\{x \ge q\}}\right)
-    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^x\right]^{-\xi}
+    \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
     +
-    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^{-x}\right]^{-\xi}
+    \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
   }_{U'(x)}
@@
   Thus the monotonicity of $U(x)$ is the opposite of $U'(x)$. For
-  convenience of writing, let $a = (1 - \hat\phi)e^r/\hat\phi$.
+  convenience of writing, let $a = (1 - \phi)e^r/\phi$.
   First of all, we find the conditions on which $U'(x)$ is monotonic.
@@
   where we have defined $y = e^x$, $c = (1 + b)^{-1/(1 + \xi)}$ and
-  \begin{eqnarray*}
-    f(y) &=& \log(a + y) - \log(a y + 1) \\
-    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > q\}}
-  \end{eqnarray*}
-  Observe
-  \begin{eqnarray}
-    f(y) - g(y) &=& \log\left(
-      {a + y \over a y + 1}
-      y^{\xi - 1 \over \xi + 1}
-    \right) - \log(c)\1_{\{y > e^q\}} \label{eq:f-g}
-  \end{eqnarray}
-  and when $y \neq e^q$,
-  \begin{eqnarray}
-    {\pd f(y) \over \pd y} &=& {1 - a^2 \over (a + y) (a y + 1)}
-    \label{eq:fd}\\
-    {\pd g(y) \over \pd y} &=& {1 - \xi \over (1 + \xi) y}
-    \label{eq:gd}
-  \end{eqnarray}
-  Furthermore
+  $f(y)$, $g(y)$ as above. Observe
+  % \begin{eqnarray}
+  %   f(y) - g(y) &=& \log\left(
+  %     {a + y \over a y + 1}
+  %     y^{\xi - 1 \over \xi + 1}
+  %   \right) - \log(c)\1_{\{y > e^q\}} \label{eq:f-g}
+  % \end{eqnarray}
+  % and when $y \neq e^q$,
+  % \begin{eqnarray}
+  %   {\pd f(y) \over \pd y} &=& {1 - a^2 \over (a + y) (a y + 1)}
+  %   \label{eq:fd}\\
+  %   {\pd g(y) \over \pd y} &=& {1 - \xi \over (1 + \xi) y}
+  %   \label{eq:gd}
+  % \end{eqnarray}
+  % Furthermore
   \begin{equation}
@@
         \]
-        So $U'(x)$ and hence $U(x)$ as a whole is not monotonic on $(0, \infty)$.
+        So $U'(x)$ and hence $U(x)$ as a whole is not monotonic on
+        $(0, \infty)$.
 
@@
 
-    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar argument as in the
+    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar
+      argument as in the
       previous case, $(f - g)(y)$ is monotonically decreasing on both
@@
     \]
-    It remains to compare $y_+$ with $1$ to determine whether $f(y) =
-    g(y)$ has a solution. Assume $y_+ > 1$. Then
-    \begin{eqnarray*}
-      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1) \\
-      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0
-    \end{eqnarray*}
+    It remains to compare $y_+$ with $1$ to determine whether
+    ${\pd (f -g)(y) \over \pd y} = 0$ has a solution on $(1, \infty)$.
+    Assume $y_+ > 1$. Then
+    \begin{eqnarray}
+      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1)
+      \nonumber \\
+      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0 \label{eq:mi6}
+    \end{eqnarray}
     This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
@@
     \[
-    \sqrt{(a^2 - 1) (a^2 - \xi^2)} > a^2 - \xi a + a - \xi = (a + 1) (a - \xi)
+    \sqrt{(a^2 - 1) (a^2 - \xi^2)} > a^2 - \xi a + a - \xi
+    = (a + 1) (a - \xi)
     \]
@@
     \infty)$. So $U'(\cdot)$ is not monotonic.
+
+    \item $a > \max\{1, \xi\}$ and $\xi > 1$, i.e. $1 < \xi < a$. By
+      the same argument that leads to \eqref{eq:mi6}, we see
+      $y_+ > 1$. From \eqref{eq:f-g-der} it is clear $(f-g)(y_+)$ is a
+      minimum. If $(f-g)(y_+) > 0$, there is no solution to
+      $(f-g)(y) = 0$ on $(1, \infty)$; if $(f-g)(y_+) < 0$,
+      there must be a solution to $(f-g)(y) = 0$ on $(y_+, \infty)$.
+
   \end{enumerate}

### 2017-05-02 papers/FX/draft.tex
Completed the proof about monotonicity.

--- 2017-04-29 papers/FX/draft.tex
@@
 estimates are based on 1304 observations of daily return 
-data from 1 January 2010 to 31 December 2014.
+data from 4th January 2010 to 31 December 2014.
 \par
@@
 \beam\label{eq:u}
-u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>-1\,. 
+u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
 \eeam
+In practice it is more often to observe $\gamma > 1$ {\blue Let's
+consult Casper for references.} Therefore we adopt a more convenient
+notation in this case
+\beao
+u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
+\eeao
+
 %In practice it is more often to observe $\gamma > 1$ {\blue Let's
@@
 \eeam
-where {\red where needed?}
-$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
+where
+$0 < p < 1$ and $K,K' > 0$. Then we have
 \beam\label{eq:xxie1.0}
@@
 K'$. In this case \eqref{eq:pareto} becomes
-\begin{eqnarray*}
+\begin{eqnarray}
   \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
@@
   - b u(\delta v) F_X(q)
-\end{eqnarray*}
-A detailed analysis of the montonicity of
-\[
+  \label{eq:ipf}
+\end{eqnarray}
+First of all we note that the function
+\[
+U_{\text{all}} =
 u(C(x)) \left[
-  1 - {b \over 1 + b}\1_{\{x > q\}}
+  1 - {b \over 1 + b}\1_{\{x \ge q\}}
   \right]
@@
 \]
-which comprises lemma \ref{lemma:III}, allows us to apply lemma
-\ref{lemma:I} and conclude the following
-\begin{itemize}
-\item I am a good guy
-\item So are you
-\end{itemize}
-
-For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
-quantity is denoted by $I(\alpha, K)$.
+is not monotonic. To see this we may re-write the above as
+\begin{equation}
+U_{\text{all}} =
+  -{\phi^{-\xi} \over \xi} \left\{
+  \underbrace{
+    \left(1 - {b \over 1 + b}\1_{\{x \ge q\}}\right)
+    \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
+    +
+    \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
+  }_{U(x)} \right\}
+  \label{eq:frg}
+\end{equation}
+Direct computation gives
+\begin{eqnarray*}
+  {\pd U(x) \over \pd x}
+  &=&
+  -{
+    (a + e^x)^{-\xi - 1} \xi e^x
+    \over
+    1 + b \1_{\{x > q\}}
+  } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}
+  \quad x \neq q
+\end{eqnarray*}
+where $a = (1 - \phi) e^r/\phi$. For $x \ge q$, we have
+\begin{eqnarray*}
+  {1 \over \xi}
+  {\pd U(x) \over \pd x} &=& {x' \over (a + x')^{\xi + 1}}
+  - {x'^\xi \over (a x' + 1)^{\xi + 1} (1 + b)}
+\end{eqnarray*}
+where we have substituted $x'$ for $e^{-x}$. When $x$ is very large
+i.e. when $x'$ is very close to 0, it is useful to expand the last
+expression at $x' = 0$. This yields
+\[
+  {1 \over \xi}
+  {\pd U(x) \over \pd x}
+  = {x' \over a^{\xi + 1}} + o(x')
+\]
+So we see that when $x$ is sufficiently large
+${\pd U(x) \over \pd x} > 0$. On the other hand, at the discontinuity
+point $x = q$, we have
+\[
+\lim_{x \downarrow 0} \left(
+  U(q + x) - U(q - x)
+\right)
+= - {
+  (a + e^q)^{-\xi} b
+  \over
+  1+ b
+} < 0
+\]
+So $U(x)$ and hence \eqref{eq:frg} is not monotonic. we resort to
+numerical methods to gain some understanding of how $\tilde u(F_X,
+\phi)$ changes with $\alpha$.
+For this reason, we replace $\phi$ in the first integral in
+\eqref{eq:pp} by the optimal $\hat\phi$. The resulting quantity is
+denoted by $I(\alpha, K)$.
 \par
 The value of $\hat\phi$ can be calculated
-by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
-This is shown in 
+by numerical integration and optimization with respect to $\phi$ for
+given values of $K, K', \alpha, \beta$.  This is shown in 
 Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
@@
 
+The following graph (figure \ref{fig:preference_pareto}) shows how
+$\tilde u(\alpha, K, \wt\phi)$ changes with $\alpha$ and $K$.
 \begin{figure}[htb!]
@@
   \end{minipage}
-  \caption{The quantities $I(\alpha, K)$ for the power-utility \fct\ with $\xi=1/2$ (left) and $\xi=4$ (right).
-The graphs clearly show that  $I(\alpha, K)$ increases with $\alpha$ and decreases with $K$.
+  \caption{The quantities $I(\alpha, K)$ for the power-utility \fct\
+    with $\xi=1/2$ (left) and $\xi=4$ (right). The graphs clearly show
+    that  $I(\alpha, K)$ increases with $\alpha$ and decreases with
+    $K$.
   }
@@
 \end{figure}
+
+When the parameter $b$ is very small, a GDA preference is closely
+approximated by a mean-utility preference, which GDA reduces to when
+$b = 0$. As shown in the proof of lemma \ref{lemma:III}, the function
+$U_{\text{all}}$ is monotonically increasing or decreasing when $\xi$
+and $(1 - \phi) e^r / \phi$ satisfy certain conditions. Moreover, it
+is also easily checked that, when $x \in (0, K(e^{1/\alpha} - 1)]$,
+the density function in the integral of \eqref{eq:ipf} is increasing
+with $\alpha$; when $x \in (K(e^{1/\alpha} - 1), \infty)$, it is
+decreasing with $\alpha$. So, following the same line of arguments
+using lemma \ref{lemma:I} as in the proof of lemma \ref{thrm:I},
+it can be seen $\tilde u_{\text{max}}(\alpha, K)$ is
+increasing/decreasing with $\alpha$ when $U_{\text{all}}$ is
+decreasing/increasing.
 
@@
 \begin{lemma}\label{lemma:III}
-Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:u} and \eqref{eq:xxie1} respectively. Define
-\[
-U(x) =
-u(C(x)) \left[
-  1 - {b \over 1 + b}\1_{\{x > q\}}
-  \right]
+Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:u} and
+\eqref{eq:xxie1} respectively. Define
+\begin{eqnarray*}
+U_{\text{all}} &=& u(C(x))
 + u(C(-x))
 \quad
-x \geq 0
-\]
+x \geq 0 \\
+a &=& {
+  (1 - \phi) e^r
+  \over
+  \phi
+} \\
+y_\pm &=& {
+  a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
+  \over
+  a (\xi - 1)
+}
+\end{eqnarray*}
+The following holds true:
 \begin{enumerate}
-\item  Hello
+\item If $\max\{a, 1\} < \xi$, $U_{\text{all}}(\cdot)$ is
+  monotonically decreasing.
+\item If $a < \xi < 1$ and $(a + y_-)/(a y_- + 1) <
+  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotonically
+  decreasing.
+\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotonically
+  increasing.
+\item If $1 < \xi < a$ and $(a + y_+)/(a y_+ + 1) >
+  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotonically
+  increasing.
+\item In other case, $U_{\text{all}}(\cdot)$ is not monotonic.
 \end{enumerate}
@@
 \begin{proof}
-  It is convenient to re-write $U(x)$ as
-  \[
-  U(x) = 
-  -{\hat\phi^{-\xi} \over \xi} \left\{
+  It is convenient to re-write $U_{\text{all}}$ as
+  \[
+  U_{\text{all}} =
+  -{\phi^{-\xi} \over \xi} \left\{
     \underbrace{
-    \left(1 - {b \over 1 + b}\1_{\{x > q\}}\right)
-    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^x\right]^{-\xi}
+      \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
       +
-    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^{-x}\right]^{-\xi}
-  }_{U'(x)}
+      \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
+    }_{U(x)}
   \right\}
   \]
-  Thus the monotonicity of $U(x)$ is the opposite of $U'(x)$. For
-  convenience of writing, let $a = (1 - \hat\phi)e^r/\hat\phi$.
-  First of all, we find the conditions on which $U'(x)$ is monotonic.
+  Thus the monotonicity of $U_{\text{all}}(x)$ is the opposite of $U(x)$. For
+  convenience of writing, let $a = (1 - \phi)e^r/\phi$.
+  First of all, we find the conditions on which $U(x)$ is monotonic.
   Direct differentiation yields
   \begin{eqnarray*}
-    {\pd U'(x) \over \pd x}
+    {\pd U(x) \over \pd x}
     &=&
@@
       (a + e^x)^{-\xi - 1} \xi e^x
-      \over
-      1 + b \1_{\{x > q\}}
     } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}
-    \quad
-    x \neq q
-  \end{eqnarray*}
-  ${\pd U'(x) \over \pd x} = 0$ is equivalent to
+  \end{eqnarray*}
+  ${\pd U(x) \over \pd x} = 0$ is equivalent to
   \begin{eqnarray*}
-    {a + y \over a y + 1} &=& 
-    \left( \1_{\{y \leq e^q\}} + c \1_{\{y > e^q\}} \right)
-    y^{1 -\xi \over 1+\xi} \\
-    \underbrace{\log(a + y) - \log(a y + 1)}_{f(y)} &=& \underbrace{{1
-        -\xi \over 1+\xi} \log(y)
-      + \log(c)\1_{\{y > e^q\}} }_{g(y)} 
-  \end{eqnarray*}
-  where we have defined $y = e^x$, $c = (1 + b)^{-1/(1 + \xi)}$ and
-  \begin{eqnarray*}
-    f(y) &=& \log(a + y) - \log(a y + 1) \\
-    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > e^q\}}
-  \end{eqnarray*}
-  Observe
-  \begin{eqnarray}
-    f(y) - g(y) &=& \log\left(
-      {a + y \over a y + 1}
-      y^{\xi - 1 \over \xi + 1}
-    \right) - \log(c)\1_{\{y > e^q\}} \label{eq:f-g}
-  \end{eqnarray}
-  and when $y \neq e^q$,
-  \begin{eqnarray}
-    {\pd f(y) \over \pd y} &=& {1 - a^2 \over (a + y) (a y + 1)}
-    \label{eq:fd}\\
-    {\pd g(y) \over \pd y} &=& {1 - \xi \over (1 + \xi) y}
-    \label{eq:gd}
-  \end{eqnarray}
-  Furthermore
+    {a + y \over a y + 1} &=& y^{1 -\xi \over 1+\xi} \\
+    \underbrace{\log(a + y) - \log(a y + 1)}_{f(y)} &=&
+    {1 -\xi \over 1+\xi} \log(y)
+  \end{eqnarray*}
+  where we have defined $y = e^x$, and $f(y)$, $g(y)$ as above. Observe
   \begin{equation}
@@
     }
-    \quad
-    y \neq e^q
   \end{equation}
-  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots
+  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots when $\xi \ne 1$
   \begin{eqnarray*}
@@
   \item If $\min\{1, \xi\} < a < \max\{1, \xi\}$, $y_\pm$ are not
-    real, ${\pd (f -g)(y) \over \pd y} = 0$ has no solution on $(1,
-    \infty)\setminus \{e^q\}$; $(f -g)(y)$ is monotonic on both
-    $(1, e^q)$ and $(e^q, \infty)$.
+    real, ${\pd (f -g)(y) \over \pd y} = 0$ has no solution on
+    $(1, \infty)$; $(f -g)(y)$ is monotonic on $(1, \infty)$.
     \begin{enumerate}
     \item If in addition $\xi > 1$, i.e. $1 < a < \xi$, $(f - g)(y)$
-      is monotonically increasing on both
-      $(1, e^q)$ and $(e^q, \infty)$, because the coefficient of the
-      $y^2$ term of the numerator of \eqref{eq:f-g-der}, i.e. $a (\xi
-      - 1)$ is positive.
-      \begin{itemize}
-      \item If $f(e^q) > g(e^q)$, then clearly $f(y) = g(y)$ has no
-        solution on $(e^q, \infty)$; we also note $f(1) = g(1) =
-        0$. Thus $f(y) = g(y)$ has no solution on $(1, e^q)$ either.
-        Now it can be concluded that $U'(\cdot)$ is monotonic on
-        $(1, q)$ and $(q, \infty)$. To determine the monotonicity of
-        $U'(\cdot)$, it remains to check the signs of
-        ${\pd U'(x) \over \pd x}$  on  $(0, q)$ and on $(q, \infty)$
-        as well as the values of $\lim_{x \uparrow q} U'(x)$ and
-        $\lim_{x \downarrow q} U'(x)$. Obviously
+      is monotonically increasing on $(1, \infty)$, because the
+      coefficient of the $y^2$ term of the numerator of
+      \eqref{eq:f-g-der}, i.e. $a (\xi - 1)$ is positive. We note
+      $f(1) = g(1) = 0$; thus on $(1, \infty)$, there is no solution
+      to $f(y) = g(y)$. It can be concluded that $U(\cdot)$ is
+      monotonic on $(0, \infty)$. Furthermore
         \begin{eqnarray*}
-          \left.{\pd U'(x) \over \pd x}\right|_{x=0} &=& 0 \\
-        \end{eqnarray*}
-        For a small $\epsilon > 0$, the sign of ${\pd U'(x) \over \pd
+          \left.{\pd U(x) \over \pd x}\right|_{x=0} &=& 0 \\
+        \end{eqnarray*}
+        For a small $\epsilon > 0$, the sign of ${\pd U(x) \over \pd
           x}$ on $(0, \epsilon)$ is thus the same as
-        $\left.{\pd^2 U'(x) \over \pd x^2}\right|_{x=0}$:
-        \[
-        \left.{\pd^2 U'(x) \over \pd x^2}\right|_{x=0} =
+        $\left.{\pd^2 U(x) \over \pd x^2}\right|_{x=0}$:
+        \[
+        \left.{\pd^2 U(x) \over \pd x^2}\right|_{x=0} =
         2 (a + 1)^{-\xi - 2} (\xi - a) > 0
         \]
-      \end{itemize}
-
-      % conversely, $f(e^q) < g(e^q)$ implies $f(y) = g(y)$ must have a
-      % solution on $(e^q, \infty)$, since \eqref{eq:f-g} tells us
-      % $(f-g)(y) \to \infty$ as $y \to \infty$.
-
-    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, $(f - g)(y)$
-      is monotonically decreasing.
+        Thus ${\pd U(x) \over \pd x} > 0$ for $x \in (0, \infty)$;
+        $U_{\text{all}}$ is monotonically decreasing.
+    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar
+      argument as in the previous case, $U_{\text{all}}$ is monotonically
+      increasing on $(0, \infty)$.
     \end{enumerate}
-
-  % \item If $a < \min\{1, \xi\}$ or $a > \max\{1, \xi\})$, ${\pd (f
-  %     -g)(y) \over \pd y} = 0$ has 2 real solutions.
-  % \end{enumerate}
-
-  % \begin{enumerate}
-  % \item If $a < 1$ and $\xi > 1$
-  %   \begin{enumerate}
-  %   \item when $x > q$, $U'(x)$ being monotonic is equivalent to ${\pd U'(x)
-  %       \over \pd x}$ having no zero on $(q, \infty)$.
-
-  %     \begin{enumerate}
-  %     \item If $f (e^q) > g(e^q)$ it is clear $f(y) = g(y)$
-  %       has no solution on $(e^q, \infty)$.
-  %     \item If $f (e^q) <  g(e^q)$ it is also clear $f(y) = g(y)$ must have
-  %       one solution on $(e^q, \infty)$.
-  %     \end{enumerate}
-  %   \end{enumerate}
-
-  %   \item If $a > 1$ and $\xi < 1$
-  %     \begin{enumerate}
-  %     \item when $x > q$, the same calculation that leads to
-  %       \eqref{eq:fd} and \eqref{eq:gd} gives ${\pd f(y) \over \pd y}
-  %       < 0$ and ${\pd g(y) \over \pd y} > 0$. So
-  %     \begin{enumerate}
-  %     \item If $f (e^q) >  g(e^q)$, $f(y) = g(y)$ must have
-  %       one solution on $(e^q, \infty)$.
-  %     \item If $f (e^q) < g(e^q)$, $f(y) = g(y)$ has no solution on $(e^q, \infty)$.
-  %     \end{enumerate}
-  %     \end{enumerate}
-
-  %     \item If $a < 1$ and $\xi < 1$
-  %       \begin{enumerate}
-  %       \item when $x > q$
-  %         \begin{enumerate}
-  %         \item If $f (e^q) -  g(e^q) > 0$, we observe $f(y) - g(y)
-  %           \to -\infty$ as $y \to \infty$. So by continuity $f(y) -
-  %           g(y) = 0$ must have at least one solution on $(e^q, \infty)$.
-  %         \end{enumerate}
-  %       \end{enumerate}
-  %     \item If $a > 1$ and $\xi > 1$
-  %       \begin{enumerate}
-  %       \item when $x > q$, both $f(y)$ and $g(y)$ decreases. Observe
-  %         In this case, if $f(e^q) > g(e^q)$, there is no solution 
-  %       \end{enumerate}
+  \item If $a < \min\{1, \xi\}$ and $\xi > 1$, i.e. $a < 1 < \xi$, it
+    is clear
+    \[
+    y_- = {
+      a^2 - \xi - \sqrt{(a^2 - 1) (a^2 - \xi^2)}
+      \over
+      a (\xi - 1)
+    } < 0
+    \]
+    It remains to compare $y_+$ with $1$ to determine whether
+    ${\pd (f -g)(y) \over \pd y} = 0$ has a solution on $(1, \infty)$.
+    Assume $y_+ > 1$. Then
+    \begin{eqnarray}
+      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1)
+      \nonumber \\
+      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0 \label{eq:mi6}
+    \end{eqnarray}
+    This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
+    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotonically increasing
+    on $(1, \infty)$. Following the same analysis as in the case
+    (1.a), one can see $U_{\text{all}}$ is monotonically decreasing.
+
+  \item If $a < \min\{1, \xi\}$ and $\xi < 1$, i.e. $a < \xi < 1$, it
+    is clear $y_- > 0$ and $y_+ < y_-$. Moreover, $y_- > 1$ is equivalent
+    to
+    \[
+    \sqrt{(a^2 - 1) (a^2 - \xi^2)} > a^2 - \xi a + a - \xi
+    = (a + 1) (a - \xi)
+    \]
+    The last inequality is obviously true in this case. So $y_- >
+    1$. ${\pd (f-g)(y) \over \pd y}$ has a maximum at
+    $y_- \in (1, \infty)$. We know $(f - g)(y_-)$ is a maximum because
+    \eqref{eq:f-g-der} shows that, for $y > y_-$,
+    ${\pd (f-g)(y) \over \pd y} < 0$.
+
+    If $(f-g)(y_-) < 0$, $(f-g)(y) = 0$ has no solution on $(1,
+    \infty)$. Hence $U(\cdot)$ is monotonic on both $(0, \infty)$.
+    The same analysis as in the case (1.a) shows $U_{\text{all}}$
+    is monotonically decreasing on $(0, \infty)$.
+
+    If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
+    \infty)$. So $U(\cdot)$ is not monotonic.
+
+    \item $a > \max\{1, \xi\}$ and $\xi > 1$, i.e. $1 < \xi < a$. By
+      the same argument that leads to \eqref{eq:mi6}, we see
+      $y_+ > 1$. From \eqref{eq:f-g-der} it is clear $(f-g)(y_+)$ is a
+      minimum. If $(f-g)(y_+) > 0$, there is no solution to
+      $(f-g)(y) = 0$ on $(1, \infty)$; $U(x)$ is monotonic. By the
+      same analysis as in the case (1.a), we know $U(x)$ is
+      monotonically decreasing and $U_{\text{all}}$ is monotonically increasing.
+
+      if $(f-g)(y_+) < 0$, there must be a solution to $(f-g)(y) = 0$
+      on $(y_+, \infty)$. $U_{\text{all}}$ is not monotonic.
   \end{enumerate}

### 2017-05-02 papers/FX/draft.tex
Corrected the proof of the monotonicity lemma.

@@
 \begin{lemma} \label{lemma:I}
-  Assume distribution function $F(x; \theta)$ parameterized by
-  $\theta$ is defined on $x \in (a, b)$, $a, b \in \mathbb R \cup
-  \{-\infty, \infty\}$, and in addition $F(x; \theta)$ has density
-  function $f(x; \theta)$. Assume function $h(\cdot)$ is defined on
-  $(a, b)$, $X \sim F$, $\text{supp}(X) = (a,b)$,
-  $\E h(X) < \infty$ and
-  \begin{eqnarray*}
-    \left| \frac{\pd f(x, \theta)}{\pd \theta} \right| < \infty
-    \quad \forall x \in (a, b)
-  \end{eqnarray*}
+  Assume distribution function $F(x, \theta)$ parameterized by 
+  $\theta \in \Theta$ is defined on $x \in (a, b)$,
+  $a, b \in \mathbb R \cup \{-\infty, \infty\}$,
+  and in addition $F(x, \theta)$ has density
+  function $f(x, \theta)$ that is differentiable with respect to
+  $\theta$ for all $\theta \in \Theta$.
+  Moreover, $f(x, \theta)$ satisfies
+  \begin{equation}\label{eq:frt}
+  \int_a^b {\pd f(x, \theta) \over \pd \theta} dx
+  < \infty
+  \end{equation}
+  Assume function $h(\cdot)$ is defined on
+  $(a, b)$ and is monotonic throughout this interval;
+  $X \sim F$ and $\E h(X) < \infty$. Then
   \begin{enumerate}
   \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
-    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(a, x_0)$ while
-    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(x_0, b)$, then
+    $\frac{\pd f}{\pd \theta}(x_0, \theta) > 0$ on $(a, x_0)$ while
+    $\frac{\pd f}{\pd \theta}(x_0, \theta) < 0$ on $(x_0, b)$, then
     \[
@@
   \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\frac{\pd f}{\pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    $\frac{\pd f}{\pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$, then
+    $\frac{\pd f}{\pd \theta}(x_0, \theta) < 0$ on $(a, x_0)$  while
+    $\frac{\pd f}{\pd \theta}(x_0, \theta) > 0$ on $(x_0, b)$, then
     \[
@@
 \begin{proof}
-  Clearly
+  Firstly we show
+  \[
+    {\pd \over \pd \theta} \int_a^b f(x, \theta) dx
+    =
+    \int_a^b {\pd \over \pd \theta} f(x, \theta) dx    
+  \]
+  Let
+  \[
+  f_{\Delta}(x, \theta) = {
+    f(x, \theta +  \Delta) - f(x, \theta)
+    \over
+    \Delta
+  }
+  \]
+  Since $f(x, \theta)$ is differentiable for all $\theta \in \Theta$,
+  \[
+  f_{\Delta}(x, \theta) \to {\pd f \over \pd \theta}(x, \theta)
+  \quad \Delta \to 0
+  \]
+  For any give $\epsilon > 0$ and sufficiently small
+  $\Delta$, we have
+  \[
+  |f_{\Delta}(x, \theta)| <
+  \left|
+      {\pd f \over \pd \theta}(x, \theta)
+  \right| (1 + \epsilon)
+  \]
+  Using \eqref{eq:frt} and dominated convergence theorem we obtain
+  \[
+  \lim_{\Delta \to 0} \int_a^b f_\Delta(x, \theta) dx
+  =
+  \int_a^b {\pd f \over \pd \theta}(x, \theta) dx
+  < \infty
+  \]
+  That is
+  \[
+    {\pd \over \pd \theta} \int_a^b f(x, \theta) dx
+    =
+    \lim_{\Delta \to 0} {
+      \int_a^b f(x, \theta + \Delta) dx
+      -
+      \int_a^b f(x, \theta) dx
+      \over
+      \Delta
+    }
+    =
+    \int_a^b {\pd f \over \pd \theta}(x, \theta) dx     
+  \]
+  From \eqref{eq:frt} it easily follows that the set on which
+  ${\pd f(x, \theta) \over \pd \theta} = \infty$ has Lesbegue
+  measure 0. Call this set $S_0$. We have
   \begin{eqnarray*}
     && {\pd \E h(X) \over \pd \theta} \\
-    &=& \int_a^b h(x) {\pd f \over \pd \theta}(x, \theta) dx \\
-    &=& \underbrace{\int_a^{x_0} h(x) {\pd f \over \pd \theta}(x,
-      \theta) dx}_{I_1}
-    + \underbrace{\int_{x_0}^b h(x) {\pd f \over \pd \theta}(x,
+    &=& \int_{(a,b) \setminus S_0} h(x)
+    {\pd f \over \pd \theta}(x, \theta) dx \\
+    &=& \underbrace{\int_{(a, x_0) \setminus S_0}
+      h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_1}
+    + \underbrace{\int_{(x_0, b) \setminus S_0}
+      h(x) {\pd f \over \pd \theta}(x,
       \theta) dx}_{I_2}
   \end{eqnarray*}
+  Moreover, $x_0$ being located in the interior of $(a, b)$ and
+  $h(\cdot)$ being monotonic implies $h(x_0) < \infty$.
   \begin{enumerate}
@@
     \[
-    I_1 > h(x_0) \int_a^{x_0} {\pd f \over \pd \theta}(x, \theta) dx
+    I_1 > h(x_0) \int_{(a, x_0) \setminus S_0}
+    {\pd f \over \pd \theta}(x, \theta) dx
     \]
@@
     \begin{eqnarray*}
-      I_2 &=& \int_{x_0}^b -h(x_0)
-      \left|{\pd f \over \pd \theta}(x, \theta) \right| dx \\
-      &>& -h(x_0)
-      \int_{x_0}^b \left| 
+      I_2 &=& \int_{(x_0, b) \setminus S_0} -h(x_0)
+      \left|{\pd f \over \pd \theta}(x, \theta) \right| dx
+      > -h(x_0)
+      \int_{(x_0, b) \setminus S_0} \left| 
         {\pd f \over \pd \theta}(x, \theta)
@@
       {\pd \E h(X) \over \pd \theta}
-      > h(x_0) \int_a^b {\pd f \over \pd \theta}(x, \theta) dx
-      = h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx
+      > h(x_0) \int_{(a, b) \setminus S_0}
+      {\pd f \over \pd \theta}(x, \theta) dx
+      = h(x_0) {\partial \over \partial \theta}
+      \int_{(a, b) \setminus S_0} f(x, \theta) dx
       = 0
@@
   \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    ${\pd f \over \pd \theta}(x_0; \theta) < 0$ on $(a, x_0)$  while
-    ${\pd f \over \pd \theta}(x_0; \theta) > 0$ on $(x_0, b)$
+    ${\pd f \over \pd \theta}(x_0, \theta) < 0$ on $(a, x_0)$  while
+    ${\pd f \over \pd \theta}(x_0, \theta) > 0$ on $(x_0, b)$
     \begin{eqnarray*}
       I_1 &=&
-      \int_a^{x_0} -h(x)
-      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx \\
-      &>&
-      -h(x_0) \int_a^{x_0}
+      \int_{(a, x_0) \setminus S_0} -h(x)
+      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx
+      >
+      -h(x_0) \int_{(a, x_0) \setminus S_0}
       \left| {\pd f \over \pd \theta}(x, \theta) \right| dx
@@
     \[
-    I_2 > h(x_0) \int_{x_0}^b
+    I_2 > h(x_0) \int_{(x_0, b) \setminus S_0}
     {\pd f \over \pd \theta}(x, \theta) dx
@@
     = I_1 + I_2
-    > h(x_0) {\partial \over \partial \theta} \int_a^b f(x, \theta) dx = 0
+    > h(x_0) {\partial \over \partial \theta}
+    \int_{(a, b) \setminus S_0} f(x, \theta) dx = 0
     \]

### 2017-05-03 papers/FX/draft.tex
Fixed equation numbers.
Revised the case of t-distributed returns.

@@
 notation in this case
-\beao
+\beam\label{eq:hjyr}
 u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
-\eeao
+\eeam
 
@@
 \[
-f(x; \nu) = c(\nu) \left(
-  1 + {x^2 \over \nu}
-\right)^{-(\nu + 1)/2}
-\]
-where $\nu > 1$ and
-\[
-c(\nu) = {
-  \Gamma({\nu + 1 \over 2})
+f(x; \alpha) = c(\alpha) \left(
+  1 + {x^2 \over \alpha}
+\right)^{-(\alpha + 1)/2}
+\]
+where $\alpha > 1$ and
+\[
+c(\alpha) = {
+  \Gamma({\alpha + 1 \over 2})
   \over
-  \Gamma(\nu/2) \sqrt{\nu \pi}
+  \Gamma(\alpha/2) \sqrt{\alpha \pi}
 }
 \]
-We can write $\wt u(F, \phi)$ as
+In the same way as for \eqref{eq:ipf}, we can write $\wt u(F, \phi)$
+as
 \begin{eqnarray}
@@
   &=&
+  (1 + b)
   \int_{0}^{\infty}
-  \left[
-    u(C(x, \phi)) + (1 + b)u(C(-x, \phi))
+  \underbrace{
+    {
+      u(C(x)) \left[
+        1 - {b \over 1 + b}\1_{\{x \ge q\}}
         \right]
-  f(x, \nu) dx \nonumber \\
-  &&
-  + b \int_{0}^{q}
-  u(C(x, \phi))
-  f(x, \nu) dx - b u(C(q, \phi)) F(q, \nu)
+      + u(C(-x))
+      \over
+      (K + x)^{\alpha + 1}
+    }
+  }_{U_{\text{all}}}
+  dx
+  - b u(\delta v) F_X(q)
   \label{eq:t1}
 \end{eqnarray}
-where $C(\cdot)$ is defined in \eqref{eq:xxie1}.
+where $C(\cdot)$ is defined in \eqref{eq:xxie1}. As shown in the last
+section, $U_{\text{all}}$ is not monotonic when $b > 0$ and $u(\cdot)$ 
+takes the power-form of \eqref{eq:hjyr}; when $b = 0$, however,
+$U_{\text{all}}$ is indeed monotonic depending on the values of $\xi$
+and $(1 - \phi) e^r / \phi$. In the first case, we have to resort to
+numerical methods; in the 2nd case, we can find some analytic results.
 
@@
 \ref{lemma:II}, {\red need to look at $\pd \wt u_{\rm max} \over \pd \alpha$?} 
-${\pd \wt u(F, \phi)) \over \pd \nu} < 0$ for all $\phi$. Hence all
+${\pd \wt u(F, \phi)) \over \pd \alpha} < 0$ for all $\phi$. Hence all
 investors guided by this utility function would be seeking the
-smallest $\nu$, i.e. the heaviest tail in the market. Here note that
+smallest $\alpha$, i.e. the heaviest tail in the market. Here note that
 Student's t is a symmetric distribution; a heavier tail implies not
@@
 is revealed by numerically maximizing $\wt u(F, \phi)$ with
-respect to $\phi$ for a sequence of selected values of $\nu$. The
+respect to $\phi$ for a sequence of selected values of $\alpha$. The
 results are shown in figure~\ref{fig:phi_hat_U}.
@@
   %   \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
-  %   \caption{$\hat\phi(\nu)$ of log-utility function}
+  %   \caption{$\hat\phi(\alpha)$ of log-utility function}
   %   \label{fig:phi_hat_t_log}
@@
   %   \includegraphics[width=\textwidth]{U_b_t_log.pdf}
-  %   \caption{$G(\nu)$ of log-utility function}
+  %   \caption{$G(\alpha)$ of log-utility function}
   %   \label{fig:U_b_t_log}
@@
   %   \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
-  %   \caption{$\hat\phi(\nu)$ of power-utility function}
+  %   \caption{$\hat\phi(\alpha)$ of power-utility function}
   %   \label{fig:phi_hat_t_power}
@@
   %   \includegraphics[width=\textwidth]{U_b_t_power.pdf}
-  %   \caption{$G(\nu)$ of power-utility function}
+  %   \caption{$G(\alpha)$ of power-utility function}
   %   \label{fig:U_b_t_power}
@@
   \end{minipage}
-  \caption{$\hat\phi$ and $G(\nu)$
+  \caption{$\hat\phi$ and $G(\alpha)$
     \label{fig:phi_hat_U}
@@
 %% \begin{eqnarray*}
-%%   \wt u(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
+%%   \wt u(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \alpha) dx
 %% \end{eqnarray*}
 %% Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
-%% $\opd{\nu}\wt u(F, \phi) > 0$. In this extreme case, the investor
-%% would rather seek the largest $\nu$ or in other words, the lightest
+%% $\opd{\alpha}\wt u(F, \phi) > 0$. In this extreme case, the investor
+%% would rather seek the largest $\alpha$ or in other words, the lightest
 %% tail in the market.
@@
 One can see from figure \ref{fig:phi_hat_U_power} that both $\hat
-\phi$ and $G(\nu)$ are monotonically decreasing with $\nu$ as in the
+\phi$ and $G(\alpha)$ are monotonically decreasing with $\alpha$ as in the
 case of a log-utility function.
@@
   \end{minipage}
-  \caption{$\hat\phi$ and $G(\nu)$ when the utility
+  \caption{$\hat\phi$ and $G(\alpha)$ when the utility
     function is $u(C) = -{2 \over \sqrt{C}}$
@@
 \section{A monontonicity lemma}
+\setcounter{equation}{0}
 \begin{lemma} \label{lemma:I}
@@
   \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
-    $\frac{\pd f}{\pd \theta}(x_0, \theta) > 0$ on $(a, x_0)$ while
-    $\frac{\pd f}{\pd \theta}(x_0, \theta) < 0$ on $(x_0, b)$, then
+    $\frac{\pd f}{\pd \theta}(x, \theta) > 0$ for $x \in (a, x_0)$ while
+    $\frac{\pd f}{\pd \theta}(x, \theta) < 0$ for $x \in (x_0, b)$, then
     \[
@@
   \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\frac{\pd f}{\pd \theta}(x_0, \theta) < 0$ on $(a, x_0)$  while
-    $\frac{\pd f}{\pd \theta}(x_0, \theta) > 0$ on $(x_0, b)$, then
+    $\frac{\pd f}{\pd \theta}(x, \theta) < 0$ for $x \in (a, x_0)$  while
+    $\frac{\pd f}{\pd \theta}(x, \theta) > 0$ for $x \in (x_0, b)$, then
     \[
@@
 \section{A lemma on Student's t-distribution}
+\setcounter{equation}{0}
 \begin{lemma}
@@
   \[
-  f(x; \nu) = c(\nu) \left(
-    1 + {x^2 \over \nu}
-  \right)^{-(\nu + 1)/2}
-  \]
-  where $\nu > 1$ and
-  \[
-  c(\nu) = {
-    \Gamma({\nu + 1 \over 2})
+  f(x; \alpha) = c(\alpha) \left(
+    1 + {x^2 \over \alpha}
+  \right)^{-(\alpha + 1)/2}
+  \]
+  where $\alpha > 1$ and
+  \[
+  c(\alpha) = {
+    \Gamma({\alpha + 1 \over 2})
     \over
-    \Gamma(\nu/2) \sqrt{\nu \pi}
+    \Gamma(\alpha/2) \sqrt{\alpha \pi}
   }
@@
   Then there exists $x_0 > 0$ such that ${\pd f \over
-    \pd \nu}(x_0, \nu) = 0$ and $\forall x \in (0, x_0), {\pd f \over
-    \pd \nu}(x, \nu) > 0$ and $\forall x \in (x_0, \infty), {\pd f
-    \over \pd \nu}(x, \nu) < 0$.
+    \pd \alpha}(x_0, \alpha) = 0$ and $\forall x \in (0, x_0), {\pd f \over
+    \pd \alpha}(x, \alpha) > 0$ and $\forall x \in (x_0, \infty), {\pd f
+    \over \pd \alpha}(x, \alpha) < 0$.
 \end{lemma}
@@
   \begin{eqnarray}
-    {\pd f(x, \nu) \over \pd \nu} &=& {
-      c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
+    {\pd f(x, \alpha) \over \pd \alpha} &=& {
+      c(\alpha) x^2 (\alpha + 1) + (2 \alpha x^2 + 2 \alpha^2) c'(\alpha)
       -
-      \nu c(\nu) (x^2 + \nu) \log(1 + x^2/\nu)
+      \alpha c(\alpha) (x^2 + \alpha) \log(1 + x^2/\alpha)
       \over
-      2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
+      2 \alpha (x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
     } \nonumber \\
     &:=& {
-      P(x^2, \nu)
+      P(x^2, \alpha)
       \over
-      2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
+      2 \alpha (x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
     }
@@
 
-  While the denominator of the right side of ${\pd f(x, \nu) \over \pd \nu}$ is
-  always positive, its nummerator $P(x^2, \nu)$ has a single root:
+  While the denominator of the right side of ${\pd f(x, \alpha) \over \pd \alpha}$ is
+  always positive, its nummerator $P(x^2, \alpha)$ has a single root:
   \begin{equation}
     \label{eq:xxie5}
-    x_0^2 = \nu\exp\left\{
+    x_0^2 = \alpha\exp\left\{
       W\left[
-        -\left(1 + {1 \over \nu}\right)
-        e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+        -\left(1 + {1 \over \alpha}\right)
+        e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
       \right]
-      + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
-    \right\} - \nu
+      + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)}
+    \right\} - \alpha
   \end{equation}
@@
   the right side of \eqref{eq:xxie5} for positivity, we first note
-  $c'(\nu) > 0$:
-  \[
-  c'(\nu) = {
-    \pi \Gamma(\nu/2 + 1/2) \left\{
-      \nu \left[ \Psi(\nu/2 + 1/2) - \Psi(\nu/2)\right] - 1
+  $c'(\alpha) > 0$:
+  \[
+  c'(\alpha) = {
+    \pi \Gamma(\alpha/2 + 1/2) \left\{
+      \alpha \left[ \Psi(\alpha/2 + 1/2) - \Psi(\alpha/2)\right] - 1
     \right\}
     \over
-    2 \Gamma(\nu/2) (\pi \nu)^{3/2}
+    2 \Gamma(\alpha/2) (\pi \alpha)^{3/2}
   }
@@
     See Abramowitz and Stegun \cite{abramowitz1972handbook}, p.259,
-    formula 6.3.16. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
+    formula 6.3.16. Therefore $\Psi(\alpha/2 + 1/2) - \Psi(\alpha/2) > 0$.
     So we have
     \begin{eqnarray*}
-      && \nu \left[
-        \Psi(\nu/2 + 1/2) - \Psi(\nu/2)
+      && \alpha \left[
+        \Psi(\alpha/2 + 1/2) - \Psi(\alpha/2)
       \right] - 1 \\
@@
     \end{eqnarray*}
-    Thus $c'(\nu) > 0$. Furthermore, we recall
+    Thus $c'(\alpha) > 0$. Furthermore, we recall
     $W(\cdot)$ is increasing on its principle branch. So
@@
     W\left[
-      -\left( 1 + {1 \over \nu} \right)
-      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+      -\left( 1 + {1 \over \alpha} \right)
+      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
     \right]
-    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \\
     &>& 
     W\left[
-      -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
-      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+      -\left( 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \right)
+      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
     \right]
-    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
+    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \\
     &=& W(-y e^{-y}) + y
@@
   \[
-  y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
+  y = 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} > 1
   \]
@@
   \[
-  \nu\exp\left\{
+  \alpha\exp\left\{
     W\left[
-      -\left(1 + {1 \over \nu}\right)
-      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
+      -\left(1 + {1 \over \alpha}\right)
+      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
     \right]
-    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
-  \right\} - \nu > 0
-  \]
-  Now that we have established that ${\pd f \over \pd \nu}(x, \nu) = 0$ has a
-  single positive root, it remains to determine the sign of
-  ${\pd f \over \pd \nu}(x, \nu)$ on the two sides of the root. For this purpose
-  we observe
+    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)}
+  \right\} - \alpha > 0
+  \]
+  Now that we have established that ${\pd f \over \pd \alpha}(x,
+  \alpha) = 0$ has a single positive root, it remains to determine the
+  sign of ${\pd f \over \pd \alpha}(x, \alpha)$ on the two sides of
+  the root. For this purpose we observe
   \begin{equation}
     \label{eq:xxie5.1}
-    P(0, \nu) = 2 \nu^2 c'(\nu) > 0
+    P(0, \alpha) = 2 \alpha^2 c'(\alpha) > 0
   \end{equation}
-  So we want to investigate ${\pd P \over \pd x}(x, \nu)$:
+  So we want to investigate ${\pd P \over \pd x}(x, \alpha)$:
   \begin{eqnarray}
-    {\pd P \over \pd x}(x, \nu) &=&
-    2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \log\left(
-      1 + {x \over \nu}
+    {\pd P \over \pd x}(x, \alpha) &=&
+    2 \alpha c'(\alpha) + c(\alpha) - \alpha c(\alpha) \log\left(
+      1 + {x \over \alpha}
     \right)
@@
   \end{eqnarray}
-  Clearly, $\frac{\pd P}{\pd x}(0, \nu) > 0$. Hence from
+  Clearly, $\frac{\pd P}{\pd x}(0, \alpha) > 0$. Hence from
   \eqref{eq:xxie5.1} and
@@
     \text{sign}\left[
-      {\pd f \over \pd\nu}(x, \nu)
+      {\pd f \over \pd\alpha}(x, \alpha)
     \right]
@@
 \section{Proof of Lemma~\ref{thrm:I}}
+\setcounter{equation}{0}
 \label{sec:thrmI_proof}
@@
 \section{Lemma~\ref{lemma:III}}
+\setcounter{equation}{0}
 \begin{lemma}\label{lemma:III}

### 2017-05-03 papers/FX/draft.tex
Corrected the proof of the monotonicity lemma.

@@
 \]
-is not monotonic. To see this we may re-write the above as
+is not monotonic. To see this we may plug in \eqref{eq:hjyr} and
+re-write the above as
 \begin{equation}
@@
   \underbrace{
-    {
+    \left\{
       u(C(x)) \left[
@@
       + u(C(-x))
-      \over
-      (K + x)^{\alpha + 1}
-    }
+    \right\}
   }_{U_{\text{all}}}
-  dx
+  f(x, \alpha) dx
   - b u(\delta v) F_X(q)
@@
 \end{eqnarray}
-where $C(\cdot)$ is defined in \eqref{eq:xxie1}. As shown in the last
-section, $U_{\text{all}}$ is not monotonic when $b > 0$ and $u(\cdot)$ 
-takes the power-form of \eqref{eq:hjyr}; when $b = 0$, however,
-$U_{\text{all}}$ is indeed monotonic depending on the values of $\xi$
-and $(1 - \phi) e^r / \phi$. In the first case, we have to resort to
-numerical methods; in the 2nd case, we can find some analytic results.
+where $C(\cdot)$ is defined in \eqref{eq:xxie1}. As shown in lemma
+\ref{lemma:III}, when $b = 0$ and $u(\cdot)$
+takes the power-form of \eqref{eq:hjyr}, $U_{\text{all}}$ is monotonic
+depending on the values of $\xi$ and $(1 - \phi) e^r / \phi$. As given
+in lemma \ref{lemma:II}, there is a point $x_0 > 0$ such that
+${\pd f \over \pd \alpha}(x_0, \alpha) = 0$ and
+$\forall x \in (0, x_0),
+{\pd f \over  \pd \alpha}(x, \alpha) > 0$ and
+$\forall x \in (x_0, \infty), {\pd f \over \pd \alpha}(x, \alpha) <
+0$. Thus it remains to verify
+$\int_0^\infty {\pd f \over \pd \alpha}(x, \alpha) dx < \infty$
+if we are to apply lemma \ref{lemma:I}.
+
+As computed in the proof of lemma \ref{lemma:II}, ${\pd f \over \pd
+  \alpha}(x, \alpha)$ is given by \eqref{eq:xxie4.1}. It is also shown
+there ${d \over d \alpha} c(\alpha) > 0$. Thus for
+$\int_0^\infty {\pd f \over \pd \alpha}(x, \alpha) dx < \infty$
+it suffices to show
+\[
+\int_ 0^\infty {
+x^2
+\over
+(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
+} dx < \infty
+\]
+We may write
+\begin{eqnarray*}
+&& \int_ 0^\infty {
+x^2
+\over
+(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
+} dx \\
+&=& \left(\int_0^1 + \int_1^\infty \right) {
+x^2
+\over
+(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
+} dx \\
+&=& I_1 + I_2
+\end{eqnarray*}
+Clearly
+\[
+I_1 < \int_0^1 {1 \over \alpha} dx < \infty
+\]
+while
+\begin{eqnarray*}
+  I_2 &=& \int_1^\infty {1 \over 1 + \alpha/x^2}
+  {1 \over (1/x^2 + 1/\alpha)^{(\alpha+1) / 2}} {dx \over x^{\alpha +
+      1}} \\
+  &<& \int_1^\infty \alpha^{(\alpha+1) / 2} {dx \over x^{\alpha + 1}}
+  < \infty
+\end{eqnarray*}
+So we conclude $\int_0^\infty {\pd f \over \pd \alpha}(x, \alpha) dx <
+\infty$, and by lemma \ref{lemma:I}, $\wt u_{\text{max}}$ is monotonically
+increasing/decreasing with $\alpha$ when $U_{\text{all}}(\cdot)$ is
+monotonically decreasing/increasing.
 
@@
 
-%% However, this is not necessarily the case when $b > 0$. For example,
-%% consider the case when $q = 0$ and $\hat\phi = 1$. This may be
-%% viewed an approximation to the situation where the interest rate $r$ is
-%% vanishingly low while the investor is more conservative than
-%% greedy. In this case we have
-%% \begin{eqnarray*}
-%%   \wt u(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \alpha) dx
-%% \end{eqnarray*}
-%% Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
-%% $\opd{\alpha}\wt u(F, \phi) > 0$. In this extreme case, the investor
-%% would rather seek the largest $\alpha$ or in other words, the lightest
-%% tail in the market.
-
-\subsubsection{power-utility function}
-In this section we consider the case when the utility function
-$u(\cdot)$ takes the form
-\[
-u(x) = -{1 \over \xi} x^{-\xi}
-\]
-One can see from figure \ref{fig:phi_hat_U_power} that both $\hat
-\phi$ and $G(\alpha)$ are monotonically decreasing with $\alpha$ as in the
-case of a log-utility function.
-\begin{figure}[htb!]
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
-  \end{minipage}
-  \caption{$\hat\phi$ and $G(\alpha)$ when the utility
-    function is $u(C) = -{2 \over \sqrt{C}}$
-  }
-  \label{fig:phi_hat_U_power}
-\end{figure}
-
 \section{Conclusion}
@@
   Assume distribution function $F(x, \theta)$ parameterized by 
-  $\theta \in \Theta$ is defined on $x \in (a, b)$,
-  $a, b \in \mathbb R \cup \{-\infty, \infty\}$,
+  $\theta \in \Theta \subseteq \mathbb R$ has support
+  $(a, b) \subseteq \mathbb R$,
   and in addition $F(x, \theta)$ has density
@@
   $\theta$ for all $\theta \in \Theta$.
-  Moreover, $f(x, \theta)$ satisfies
+  Let $X \sim F$ and assume function $h(\cdot)$ is defined on $(a, b)$
+  and is monotonic throughout this interval.
+  Moreover, we assume $h(x)$ and $f(x, \theta)$ satisfy
   \begin{equation}\label{eq:frt}
-  \int_a^b {\pd f(x, \theta) \over \pd \theta} dx
+%    \E |h(X)| < \infty, \quad
+    \int_a^b \left| {\pd f(x, \theta) \over \pd \theta} \right| dx
+    < \infty \text { and }
+    \int_a^b \left| h(x) {\pd f(x, \theta) \over \pd \theta} \right| dx
     < \infty
   \end{equation}
-  Assume function $h(\cdot)$ is defined on
-  $(a, b)$ and is monotonic throughout this interval;
-  $X \sim F$ and $\E h(X) < \infty$. Then
+  Then the following holds true:
   \begin{enumerate}
@@
 \begin{proof}
-  Firstly we show
-  \[
-    {\pd \over \pd \theta} \int_a^b f(x, \theta) dx
-    =
-    \int_a^b {\pd \over \pd \theta} f(x, \theta) dx    
-  \]
-  Let
-  \[
-  f_{\Delta}(x, \theta) = {
-    f(x, \theta +  \Delta) - f(x, \theta)
-    \over
-    \Delta
-  }
-  \]
-  Since $f(x, \theta)$ is differentiable for all $\theta \in \Theta$,
-  \[
-  f_{\Delta}(x, \theta) \to {\pd f \over \pd \theta}(x, \theta)
-  \quad \Delta \to 0
-  \]
-  For any give $\epsilon > 0$ and sufficiently small
-  $\Delta$, we have
-  \[
-  |f_{\Delta}(x, \theta)| <
-  \left|
-      {\pd f \over \pd \theta}(x, \theta)
-  \right| (1 + \epsilon)
-  \]
-  Using \eqref{eq:frt} and dominated convergence theorem we obtain
-  \[
-  \lim_{\Delta \to 0} \int_a^b f_\Delta(x, \theta) dx
-  =
-  \int_a^b {\pd f \over \pd \theta}(x, \theta) dx
-  < \infty
-  \]
-  That is
-  \[
-    {\pd \over \pd \theta} \int_a^b f(x, \theta) dx
-    =
-    \lim_{\Delta \to 0} {
-      \int_a^b f(x, \theta + \Delta) dx
-      -
-      \int_a^b f(x, \theta) dx
-      \over
-      \Delta
-    }
-    =
-    \int_a^b {\pd f \over \pd \theta}(x, \theta) dx     
-  \]
-  From \eqref{eq:frt} it easily follows that the set on which
-  ${\pd f(x, \theta) \over \pd \theta} = \infty$ has Lesbegue
-  measure 0. Call this set $S_0$. We have
+  % Firstly we show, for any $S \subseteq (a, b)$,
+  % \[
+  %   {\pd \over \pd \theta} \int_S f(x, \theta) dx
+  %   =
+  %   \int_S {\pd \over \pd \theta} f(x, \theta) dx    
+  % \]
+  % Let
+  % \[
+  % f_{\Delta}(x, \theta) = {
+  %   f(x, \theta +  \Delta) - f(x, \theta)
+  %   \over
+  %   \Delta
+  % }
+  % \]
+  % Since $f(x, \theta)$ is differentiable for all $\theta \in \Theta$,
+  % \[
+  % f_{\Delta}(x, \theta) \to {\pd f \over \pd \theta}(x, \theta)
+  % \quad \Delta \to 0
+  % \]
+  % For any give $\epsilon > 0$ and sufficiently small
+  % $\Delta$, we have
+  % \[
+  % |f_{\Delta}(x, \theta)| <
+  % \left|
+  %     {\pd f \over \pd \theta}(x, \theta)
+  % \right| (1 + \epsilon)
+  % \]
+  % Using \eqref{eq:frt} and dominated convergence theorem we obtain
+  % \[
+  % \lim_{\Delta \to 0} \int_S f_\Delta(x, \theta) dx
+  % =
+  % \int_S {\pd f \over \pd \theta}(x, \theta) dx
+  % < \infty
+  % \]
+  % That is
+  % \[
+  %   {\pd \over \pd \theta} \int_S f(x, \theta) dx
+  %   =
+  %   \lim_{\Delta \to 0} {
+  %     \int_S f(x, \theta + \Delta) dx
+  %     -
+  %     \int_S f(x, \theta) dx
+  %     \over
+  %     \Delta
+  %   }
+  %   =
+  %   \int_S {\pd f \over \pd \theta}(x, \theta) dx     
+  % \]
+  % From \eqref{eq:frt} it easily follows that the set on which
+  % ${\pd f(x, \theta) \over \pd \theta} = \infty$ has Lesbegue
+  % measure 0. Call this set $S_0$.
+  Firstly, by dominated convergence theorem, conditions \eqref{eq:frt}
+  imply, for all $S \subseteq (a, b)$,
+  \begin{eqnarray*}
+    {\pd \over \pd \theta}\int_S f(x, \theta) dx
+    &=&
+    \int_S {\pd \over \pd \theta} f(x, \theta) dx \\
+    {\pd \over \pd \theta}\int_S h(x) f(x, \theta) dx
+    &=&
+    \int_S h(x) {\pd \over \pd \theta} f(x, \theta) dx \\
+  \end{eqnarray*}
+  Thus we have
   \begin{eqnarray*}
     && {\pd \E h(X) \over \pd \theta} \\
-    &=& \int_{(a,b) \setminus S_0} h(x)
+    &=& \int_a^b h(x)
     {\pd f \over \pd \theta}(x, \theta) dx \\
-    &=& \underbrace{\int_{(a, x_0) \setminus S_0}
+    &=& \underbrace{\int_a^{x_0}
       h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_1}
-    + \underbrace{\int_{(x_0, b) \setminus S_0}
+    + \underbrace{\int_{x_0}^b
       h(x) {\pd f \over \pd \theta}(x,
@@
   \end{eqnarray*}
-  Moreover, $x_0$ being located in the interior of $(a, b)$ and
-  $h(\cdot)$ being monotonic implies $h(x_0) < \infty$.
+  $x_0$ being located in the interior of $(a, b)$ and
+  $h(\cdot)$ being monotonic imply $h(x_0) < \infty$.
   \begin{enumerate}
@@
     \[
-    I_1 > h(x_0) \int_{(a, x_0) \setminus S_0}
+    I_1 > h(x_0) \int_a^{x_0}
     {\pd f \over \pd \theta}(x, \theta) dx
@@
     \begin{eqnarray*}
-      I_2 &=& \int_{(x_0, b) \setminus S_0} -h(x_0)
+      I_2 &=& \int_{x_0}^b -h(x_0)
       \left|{\pd f \over \pd \theta}(x, \theta) \right| dx
       > -h(x_0)
-      \int_{(x_0, b) \setminus S_0} \left| 
+      \int_{x_0}^b \left| 
         {\pd f \over \pd \theta}(x, \theta)
@@
       {\pd \E h(X) \over \pd \theta}
-      > h(x_0) \int_{(a, b) \setminus S_0}
+      > h(x_0) \int_a^b
       {\pd f \over \pd \theta}(x, \theta) dx
       = h(x_0) {\partial \over \partial \theta}
-      \int_{(a, b) \setminus S_0} f(x, \theta) dx
+      \int_a^b f(x, \theta) dx
       = 0
@@
     ${\pd f \over \pd \theta}(x_0, \theta) < 0$ on $(a, x_0)$  while
-    ${\pd f \over \pd \theta}(x_0, \theta) > 0$ on $(x_0, b)$
-    \begin{eqnarray*}
-      I_1 &=&
-      \int_{(a, x_0) \setminus S_0} -h(x)
-      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx
-      >
-      -h(x_0) \int_{(a, x_0) \setminus S_0}
-      \left| {\pd f \over \pd \theta}(x, \theta) \right| dx
-    \end{eqnarray*}
-    and obviously
-    \[
-    I_2 > h(x_0) \int_{(x_0, b) \setminus S_0}
-    {\pd f \over \pd \theta}(x, \theta) dx
-    \]
-    So
-    \[
-    {\pd \E h(X) \over \pd \theta}
-    = I_1 + I_2
-    > h(x_0) {\partial \over \partial \theta}
-    \int_{(a, b) \setminus S_0} f(x, \theta) dx = 0
+    ${\pd f \over \pd \theta}(x_0, \theta) > 0$ on $(x_0, b)$, 
+    by similar arguments, one can show
+    \[
+    {\pd \E h(X) \over \pd \theta}> 0
     \]
@@
     \hat \phi := \underset{0 < \phi \leq 1}{\text{argmax}}\,
-    \wt u(F, \phi)
+    \wt u(F_X, \phi)
   \end{equation}
@@
   \[
-  G(F) = \wt u(F, \hat\phi)
+  \tilde u_{\text{max}}(F_X) = \wt u(F_X, \hat\phi)
   \]
@@
   \begin{eqnarray}
-    {\pd G(F) \over \pd \alpha}
+    {d \tilde u_{\text{max}}(F_X) \over d \alpha}
     &=&
-    {\pd \wt u(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi)
-    + {\pd \wt u(F, \phi) \over \pd \phi}(\alpha, K, \hat\phi)
-    {\pd \hat\phi \over \pd \alpha}(\alpha, K)
+    \left.{\pd \wt u(\alpha, \phi) \over \pd \alpha}\right|_{\phi = \hat\phi}
+    + \left.{\pd \wt u(\alpha, \phi) \over \pd \phi}\right|_{\phi = \hat\phi}
+    {\pd \hat\phi \over \pd \alpha}
     \label{eq:xxie3}
   \end{eqnarray}
-  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
+  The definition \eqref{eq:xxie2} implies for all $\alpha$
   \begin{equation}
     \label{eq:xxie4}
-    {\pd \wt u(F, \phi) \over \pd \phi}(\alpha, K, \hat \phi) = 0
+    \left.{\pd \wt u(\alpha, \phi) \over \pd \phi}\right|_{\phi = \hat\phi} = 0
   \end{equation}
@@
   \begin{eqnarray*}
-    {\pd \wt u(F, \phi) \over \pd \alpha}(\alpha, K, \hat \phi)
-    &>& 0
+    \left.{
+        \pd \wt u(\alpha, \phi) \over \pd \alpha}
+    \right|_{\phi = \hat\phi} &>& 0
   \end{eqnarray*}
@@
   \begin{eqnarray*}
-    {\pd \wt u(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi)
-    = {\partial \over \partial \alpha}\E[u((1 - \hat\phi) e^r + \hat\phi e^X) \1_{\{X < 0\}}]
-  \end{eqnarray*}
-  The function $u((1 - \hat\phi) e^r + \hat\phi e^X)$ is obviously
-  increasing with $X$. It follows
+    {\pd \wt u(\alpha, \phi) \over \pd \alpha}
+    = {\partial \over \partial \alpha}\E[u((1 - \phi) e^r + \phi e^x) \1_{\{X < 0\}}]
+  \end{eqnarray*}
+  The function $u((1 - \phi) e^r + \phi e^x)$ is obviously
+  increasing with $x$. It follows
   \begin{eqnarray*}
-    && {\pd f(x; \alpha, K) \over \pd \alpha} \\
+    && {\pd f_X(x; \alpha, K) \over \pd \alpha} \\
     &=& {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
@@
   \begin{eqnarray*}
-    {\pd \wt u(F, \phi) \over \pd \alpha}(\alpha, K, \hat\phi) &>& 0
-  \end{eqnarray*}
-  As for ${\pd G(F) \over \pd K}$, by the same argument as for
-  ${\pd G(F) \over \pd \alpha}$, it suffices to show
-  \[
-  {\pd \wt u(F, \phi) \over \pd K}(\alpha, K, \hat\phi) < 0
+    {\pd \wt u(\alpha, \phi) \over \pd \alpha} &>& 0
+  \end{eqnarray*}
+  As for ${\pd \tilde u_{\text{max}}(F_X) \over \pd K}$, by the same argument as for
+  ${\pd \tilde u_{\text{max}}(F_X) \over \pd \alpha}$, it suffices to show
+  \[
+  {\pd \wt u(K, \phi) \over \pd K} < 0
   \]
@@
   \[
-  {\pd \wt u(F, \phi) \over \pd K}(\alpha, K, \hat\phi)
+  {\pd \wt u(K, \phi) \over \pd K}
   = {\partial \over \partial K} \E \left[
-    u((1 - \hat\phi) e^r + \hat\phi e^X) \1_{\{X < 0\}}
+    u((1 - \phi) e^r + \phi e^x) \1_{\{X < 0\}}
   \right] 
@@
   \begin{eqnarray*}
-    && {\pd f(x; \alpha, K) \over \pd K} \\
+    && {\pd f_X(x; \alpha, K) \over \pd K} \\
     &=& {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
@@
   \end{eqnarray*}
-  Clearly, when $x < -\alpha/K$
+  Clearly, when $x < -K/\alpha$
   \[
@@
   \]
-  and when $-\alpha/K < x < 0$
+  and when $-K/\alpha < x < 0$
   \[
@@
   \[
-  {\pd \wt u(F, \phi) \over \pd K}(\alpha, K, \hat\phi) < 0  
+  {\pd \wt u(K, \phi) \over \pd K} < 0  
   \]

### 2017-05-05 papers/FX/draft.tex
Revised the FX paper.

@@
 %investors of this kind, the return distributions of equities that are
-%actively traded in this market must have very similar tail indices. We
+%actively traded in this market must have very similar tail-indices. We
 %also show, in contrast, the scale parameters of the return
@@
 %On the other hand, whether or not all equities in a multivariate model
-%have the same tail index is a dividing issue for multivariate GARCH
+%have the same tail-index is a dividing issue for multivariate GARCH
 %models proposed in the literature. Therefore, it is important to analyze
 %data of real equity returns and see how close to each other the
-%tail indices actually are.
+%tail-indices actually are.
 %In this work empirical results are also presented and they appear
-%to support the conclusion that the tail indices are very similar,
+%to support the conclusion that the tail-indices are very similar,
 %with respect to the confidence bands of estimation.
@@
 \item
-Are there some {\em theoretical} reasons in favor of the fact that the tail indices from different series are {\em not} 
+Are there some {\em theoretical} reasons in favor of the fact that the tail-indices from different series are {\em not} 
 distinct?
@@
 introduced by Routledge and Zin \cite{routledge2010generalized}. GDA
-is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991}. The Gul paper derives the DA from first principles (axiomatic).
-\par 
+is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991}. Gul derives the DA from first principles (axiomatic).
+
 In Section~\ref{sec:4} we summarize the discussion of the previous
@@
 on the left tail of the \ds , i.e., on the losses. 
-However, it is common to present the tail index estimators
+However, it is common to present the tail-index estimators
 for positive data. Therefore we will multiply the losses $X_t$ by
@@
 %With these motives in mind, we present a survey of the
-%tail indices and scale parameters of 3 sectors, namely ``Energy'',
+%tail-indices and scale parameters of 3 sectors, namely ``Energy'',
 %``Consumer Staples'' and ``Information Technology'', of the S\&P 500
@@
 
-\subsection{Hill estimates of lower tail indices}\label{sec:Hill}
-Various estimators of the tail index $\alpha$  in the model
+\subsection{Hill estimates of lower tail-indices}\label{sec:Hill}
+Various estimators of the tail-index $\alpha$  in the model
 \eqref{eq:1a} have been proposed in the literature;
@@
 a sample of size $n=1304$, corresponding to the 96\%-quantile of the
-data. In general, the estimation of the tail index is an art and
+data. In general, the estimation of the tail-index is an art and
 requires some expertise; for some guidance
@@
   \end{minipage}
-  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail indices $\alpha$ of
+  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail-indices $\alpha$ of
     daily return series in sectors of the S\&P 500
     index. The data span
-    from 1 January 2010 to 31 December 2011 and comprise $n=1304$ observations.
+    from 1 January 2010 to 31 December 2014 and comprise $n=1304$ observations.
 The graphs from top to bottom correspond to the ``Energy'',
@@
 This fact indicates that the returns in each sector may 
-have comparable tail indices.
+have comparable tail-indices.
 \par
 Hoga's \cite{hoga:2016} test about the change of extreme quantiles
-may provide some further insight about how similar these tail indices are.
-Different tail indices are likely to result in different
+may provide some further insight about how similar these tail-indices are.
+Different tail-indices are likely to result in different
 extreme quantiles. Nevertheless, changes in the extreme quantiles may also
@@
   \end{minipage}
-  \caption{\small Estimates $\hat K_k$ and $\hat K_k^{\hat \alpha}$ on $\log_{10}$-scale 
-of stocks in sectors
+  \caption{\small Estimates of $\hat K_k$ (top) and $\hat K_k^{\hat
+      \alpha}$ (bottom) on $\log_{10}$-scale of stocks in sectors
     of the S\&P 500 index. The points are the estimated values, the bars 
@@
 relations between the energy enterprises. While two IT companies
-may provide a variety of products and services and not depend on each
+may provide a variety of products and services and do not depend on each
 other, two energy companies are more likely to depend on each other
@@
   \end{minipage}
-  \caption{{\em Top row}: Test for pairwise equality of tail-indices in the 
-    ``Energy'', ``Consumer Staples'' and ``Information Technology'' 
-    sectors of S\&P 500. The test statistic in \eqref{eq:x1} is based on Hill estimates 
+  \caption{{\em Top row}: Test for pairwise equality of tail-indices
+    of losses in the  ``Energy'', ``Consumer Staples'' and
+    ``Information Technology'' 
+    sectors of S\&P 500. The test statistic in \eqref{eq:x1} is based
+    on Hill estimates 
     of $\alpha_X$ and $\alpha_Y$. 
@@
     when the test statistic is outside the intervals $[q_{0.075},q_{0.925}]$,
- $[q_{0.05},q_{0.95}]$,  $[q_{0.025},q_{0.975}]$, repectively, where $q_p$ is the $p$-quantile of the limiting
-$N(0,\alpha_X^2+\alpha_Y^2)$-\ds\ of the test statistic in \eqref{eq:x1}. 
-    Grey points stand for pairs for which the test statistic is inside $[q_{0.075},q_{0.925}]$. 
-    {\em Bottom row:} Test for changing tail index or scale parameter     
-    using Hoga's test based on concatenated series of pairs of stocks. The green, blue and red points 
+    $[q_{0.05},q_{0.95}]$,  $[q_{0.025},q_{0.975}]$, repectively, where
+    $q_p$ is the $p$-quantile of the limiting 
+    $N(0,\alpha_X^2+\alpha_Y^2)$-\ds\ of the test statistic in
+    \eqref{eq:x1}. Grey points stand for pairs for which the test
+    statistic is inside $[q_{0.075},q_{0.925}]$. 
+    {\em Bottom row:} Test for changing tail-index or scale parameter
+    of losses
+    using Hoga's test based on concatenated series of pairs of
+    stocks. The green, blue and red points
     correspond to pairs of stock in a sector 
-    when the test statistic $T_n$ is exceeds the $85\%$-, $90\%$-, $95\%$-quantile of the limit \ds .  
-  Grey points stand for pairs for which the test statistic is below the \asy\ $85\%$-quantile. Black points represent
+    when the test statistic $T_n$ is exceeds the $85\%$-, $90\%$-,
+    $95\%$-quantile of the limit \ds .
+  Grey points stand for pairs for which the test statistic is below
+  the \asy\ $85\%$-quantile. Black points represent 
     pairs for which the computation of $T_n$ fails for given precision  
@@
 \end{figure}
-It indicates that tail indices of equities in the
+It indicates that tail-indices of equities in the
 ``Energy'' or ``Information Technology'' sectors are more variable 
-in the ``Consumer Staples'' sector, as the null
+than in the ``Consumer Staples'' sector, as the null
 hypothesis is rejected more often for members of these two former sectors.
 Moreover, these figures suggest that the test based on the Hill estimator
-is quite powerful in distinguishing between tail indices. In contrast to the test presented in Section~\ref{sec:Hoga}
+is quite powerful in distinguishing between tail-indices. In contrast to the test presented in Section~\ref{sec:Hoga}
 the present test results in  more rejections for the ``Energy'' and the
@@
 
-\subsection{A tests for a change in the extreme tail}\label{sec:Hoga}
+\subsection{A test for a change in the extreme tail}\label{sec:Hoga}
 Here we apply a test from
@@
 {\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to
-zero. If the tail index {\em or} the scale parameter in a \ds\ of type
+zero. If the tail-index {\em or} the scale parameter in a \ds\ of type
 \eqref{eq:3} change inside a sample, then it is likely that the
@@
 When applied to our problem we would like to test 
-whether there is a change of the tail index {\em or} scale parameter in \eqref{eq:3} in each of the S\&P 500 
-series in the distinct sectors. We also  want to get some indication about a possible change of tail index or
+whether there is a change of the tail-index {\em or} scale parameter in \eqref{eq:3} in each of the S\&P 500 
+series in the distinct sectors. We also  want to get some indication about a possible change of tail-index or
 scale parameter from one series to another within a given sector. For this reason, we choose any pair of series
 within a sector and concatenate each of the paired series. Then we run the test on the concatenated series.
-Of course, despite the fact that we test changes of tail index or scale parameter 
+Of course, despite the fact that we test changes of tail-index or scale parameter 
 {\em in a very indirect way} -- there may be many other reasons for the change of extreme quantiles in a sample -- 
@@
 for $t_0=0.1$ and daily return series of stock in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
-500 index. The ``null hypothesis'' is that the tail index and scale parameter remain the
+500 index. The ``null hypothesis'' is that the tail-index and scale parameter remain the
 same througout the selected period of time. For most stocks, the hypothesis cannot be rejected even at the 85\% level.
@@
 \par
-To check whether any pair of stocks shares the same tail index and scale parameter we concatenate
+To check whether any pair of stocks shares the same tail-index and scale parameter we concatenate
 any two series and apply the aforementioned test on the concatenated series. For the ``Energy'' and the
@@
 These graphs show that the ``null
-hypothesis'' of an equal tail index and scale parameter is rejected for more pairs in the
+hypothesis'' of an equal tail-index and scale parameter is rejected for more pairs in the
 ``Energy'' sector than it is for those in the ``Consumer Staples''
-sector. This suggests that lower tail indices of stocks in the
+sector. This suggests that lower tail-indices of stocks in the
 ``Energy'' sector are more spread out than those of the ``Consumer
@@
 %   \end{minipage}
-%   \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail index or scale parameter based on concatenated series in the ``Energy'' and
+%   \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail-index or scale parameter based on concatenated series in the ``Energy'' and
 %     ``Consumer Staples'' sectors of
@@
 Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
-larger the minimum tail index in the concatenated pair.
+larger the minimum tail-index in the concatenated pair.
 \begin{figure}[htb!]
@@
 \newpage
-\section{Some theoretical arguments for equality of tail indices}\setcounter{equation}{0}\label{sec:2}
-\subsection{Multivariate GARCH models whose components have  equal tail indices}\label{subsec:garch}
+\section{Some theoretical arguments for equality of tail-indices}\setcounter{equation}{0}\label{sec:2}
+\subsection{Multivariate GARCH models whose components have  equal tail-indices}\label{subsec:garch}
 Among the models for returns the generalized autoregressive conditionally heteroscedastic (GARCH) model
@@
 \eeam
-for some positive $c>0$ and a positive tail index $\alpha$ being the unique solution of the equation 
+for some positive $c>0$ and a positive tail-index $\alpha$ being the unique solution of the equation 
 $\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$
@@
 with indices $\kappa_1$ and $\kappa_2$, respecticely,  given by the solutions to the equations 
-$\E [(\alpha_{ii} Z_{i,t}^2+\beta_{ii})^{\kappa_i}]=1$, $i=1,2$.
+$\E [(\alpha_{ii} Z_{i,t}^2+\beta_{ii})^{\kappa_i/2}]=1$, $i=1,2$.
 In this model, one can introduce dependence between the two component series $(X_{1,t})$ and $(X_{2,t})$ by assuming dependence between the noise variables $Z_{1,t}$ and $Z_{2,t}$.
@@
 
-\subsection{An utility based argument for equal tail indices}\label{sec:3}
-In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
+\subsection{An utility based argument for equal tail-indices}\label{sec:3}
+In this section we give an argument based on economic theory which  favors equality of  tail-indices for return series.
 We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
@@
 where
-$0 < p < 1$ and $K,K' > 0$. Then we have
+$\alpha, \beta > 0$, $K,K' > 0$, $0 < p < 1$ and
+$\alpha, \beta, K, K'$ are independent of each other. Then we have
 \beam\label{eq:xxie1.0}
@@
 $\P(X>x)=0.5 (\kappa/(x+\kappa))^{\alpha}$. If the symmetry causes another behavior than in the previous section this would be interesting.}
-{\blue our tail index is always $\alpha$, not $\nu$.}
+{\blue our tail-index is always $\alpha$, not $\nu$.}
 \[
@@
 {\pd f \over  \pd \alpha}(x, \alpha) > 0$ and
-$\forall x \in (x_0, \infty), {\pd f \over \pd \alpha}(x, \alpha) <
-0$. Thus it remains to verify
-$\int_0^\infty {\pd f \over \pd \alpha}(x, \alpha) dx < \infty$
+$\forall x \in (x_0, \infty)$,
+${\pd f \over \pd \alpha}(x, \alpha) < 0$. Thus it remains to verify
+$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$ and
+$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
 if we are to apply lemma \ref{lemma:I}.
@@
 there ${d \over d \alpha} c(\alpha) > 0$. Thus for
-$\int_0^\infty {\pd f \over \pd \alpha}(x, \alpha) dx < \infty$
+$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
 it suffices to show
@@
 \end{eqnarray*}
-So we conclude $\int_0^\infty {\pd f \over \pd \alpha}(x, \alpha) dx <
-\infty$, and by lemma \ref{lemma:I}, $\wt u_{\text{max}}$ is monotonically
+So we conclude
+$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
+To see
+$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$, we note
+\begin{eqnarray*}
+  \xi |U_{\text{all}}(x)|
+  &<& [(1 - \phi) e^r + e^x]^{-\xi} + [(1 - \phi) e^r + e^{-x}]^{-\xi} \\
+  &<& 1 + (1 - \phi)^{-\xi} e^{-r \xi}
+\end{eqnarray*}
+Since $\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$,
+it follows from the above inequality $\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
+Thus by lemma \ref{lemma:I}, $\wt u_{\text{max}}$ is monotonically
 increasing/decreasing with $\alpha$ when $U_{\text{all}}(\cdot)$ is
@@
 TODO: Summarize and conclude that all correlated series share the same
-tail index; their tail functions differ from each other via their
+tail-index; their tail functions differ from each other via their
 scale parameters.

### 2017-05-05 papers/FX/draft.tex
Fixed the bibliography.

@@
 Alexander and Chibumba \cite{alexander1997multivariate}, its
-generalization GO-GARCH by van der Weide \cite{van2002go},  the Full Factor GARCH model of Vrontos et al.
+generalization GO-GARCH by van der Weide \cite{Weide2002},  the Full Factor GARCH model of Vrontos et al.
 \cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
-model of Lanne and Saikkonen  \cite{lanne2007modeling}. These
+model of Lanne and Saikkonen  \cite{lanne2007modelling}. These
 models are characterized by their treatment of each series as a linear

### 2017-05-09 papers/FX/draft.tex
Revised the pictures of pareto.

@@
 \item
-In view of the wide \asy\ confidence bands of the estimators of tail-indices, 
+In view of the wide \asy\ confidence bands for the estimators of tail-indices, 
 are the estimated tail-indices from different series really distinct?
@@
 under the hypothesis that we do not know the (high) threshold value
-$K$. The estimator has ``good'' properties such as 
+$K$. The estimator has ``good'' theoretical properties such as 
 \asy\ consistency and \asy\ normality. These properties hold under
@@
 a curve of equal preference in the direction of increasing $\alpha$, $K$ also
-increases. So the strong positive correlation seen in the ``Energy''
+increases. So the strong positive dependence seen in the ``Energy''
 sector suggests these stocks might have very similar investor
@@
     correspond to pairs of stock in a sector 
-    when the test statistic $T_n$ is exceeds the $85\%$-, $90\%$-,
+    when the test statistic $T_n$ exceeds the $85\%$-, $90\%$-,
     $95\%$-quantile of the limit \ds .
@@
     {\em Left}: Test  of concatenated $t$-samples with different
-    degrees of freedom. Numbers on the axes are the degrees of freedom
+    degrees of freedom ($\alpha$). Numbers on the axes are the degrees of freedom
     in the subsamples. For an interpretation of the colored bullets,
@@
 To find out about this problem we compared the \ds s of $T_n$ 
-with $n=1304$ for $t$-distributed $X_t$ with  3 and 4 degrees of freedom and the limit \ds\ of $T_n$.
-The estimated density functions are shown in on the right of Figure~\ref{fig:t_sim_pair}.
+with $n=1304$ for $t$-distributed $X_t$ with  3 and 4 degrees of freedom ($\alpha$) and
+the limit \ds\ of $T_n$.
+The estimated density functions are shown on the right of Figure~\ref{fig:t_sim_pair}.
 As seen in the graph, the asymptotic distribution assigns
@@
   \end{tabular}
-  \caption{Quantiles of the test statistic $T_n$ for $n=1304$ $t$-distributed samples with 3 and 4 degrees of freedom as well
-as the corresponding quantiles for the limiting \ds\ of $T_n$. In particular, there are huge differences between the three \ds s  
-for the higher quantiles. 
+  \caption{Quantiles of the test statistic $T_n$ for $n=1304$
+    $t$-distributed samples with 3 and 4 degrees of freedom ($\alpha$)
+    as well as the corresponding quantiles for the limiting \ds\ of
+    $T_n$. In particular, there are huge differences between the three
+    \ds s  for the higher quantiles. 
     }
@@
 \eeam
-for some positive $c>0$ and a positive tail-index $\alpha$ being the unique solution of the equation 
+for some positive $c>0$ and a positive tail-index $\alpha$ which is the unique solution of the equation 
 $\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$
-provided the latter soluton exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
+provided that the soluton exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
 follows by an application of the Kesten-Goldie theorem; see Kesten \cite{kesten:1973}, Goldie \cite{goldie:1991}, cf. 
@@
 implying power-laws for the right and left tails of $X_t$ caused by the power-law tail of $\sigma_t$.
-A GARCH process of the order $(p,q)$ can be embedded in a multivariate equation of the type \eqref{eq:7},
-where $(\bfA_t)$ are iid random matrices and $\bfB=\bfB_t$ is a constant vector. Again, the Kesten theory \cite{kesten:1973}
-applies, implying that the marginal and \fidi s of the GARCH process are \regvary\ with a positive index $\alpha$.
-We refrain from explaining the notion of multivariate \regvar\ which is needed in this context. For further details,
-see Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} where the Kesten theorem and \regvar\ of GARCH processes
-are explained in detail.
+
+A GARCH process of the order $(p,q)$ can be embedded in a multivariate
+equation of the type \eqref{eq:7}, where $(\bfA_t)$ are iid random
+matrices and $\bfB=\bfB_t$ is a constant vector. Again, the Kesten
+theory \cite{kesten:1973} applies, implying that the marginal and
+\fidi s of the GARCH process are \regvary\ with a positive index
+$\alpha$. We refrain from explaining the notion of multivariate
+\regvar\ which is needed in this context. For further details, see
+Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} where the
+Kesten theorem and \regvar\ of GARCH processes are explained in detail.
 \par
@@
 
-\subsection{An utility based argument for equal tail-indices}\label{sec:3}
-In this section we give an argument based on economic theory which  favors equality of  tail-indices for return series.
-We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
-Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
-it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
-Following Routledge and Zin \cite{routledge2010generalized},
+\subsection{A utility based argument for equal tail-indices}\label{sec:3}
+In this section we give an argument based on economic theory that
+suggests equality of  tail-indices for equity return series.
+We follow an approach by Routledge and Zin
+\cite{routledge2010generalized} who introduced the notion of
+Generalized Disappointment Aversion (GDA). We consider the risky
+payoff $C$ of an investor and assume that it has a continuous \ds\ on
+$(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be
+increasing and concave on $(0,\infty)$.  Following Routledge and Zin
+\cite{routledge2010generalized},
 the utility of an agent with GDA preferences is given by
@@
   \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
-  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
+  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\
+  &=&
+  \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
 \eeao
@@
 \eeam
-In practice it is more often to observe $\gamma > 1$ {\blue Let's
-consult Casper for references.} Therefore we adopt a more convenient
-notation in this case
+% In practice it is more often to observe $\gamma > 1$ {\blue Let's
+% consult Casper for references.}
+We focus on the case $\gamma > 1$ and hence adopt a more convenient
+notation:
 \beam\label{eq:hjyr}
@@
 \beam\label{eq:xxie1}
-  C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
+  C(x) = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
 \eeam
@@
 \beao
-\wt u(\phi,F_X) &=& \E [u(C)] + b\, \E \big[u(C)\1(C \le  \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
+\wt u(F_X, \phi) &=& \E [u(C)] + b\, \E \big[u(C)\1(C \le  \delta v)\big] - b \,u(\delta\, v) F_X(q)\,,
 \eeao
@@
 \eeao
-exists.
+exists and that the maximum is achieved at a unique $\hat\phi \in (0, 1)$.
 
@@
 $\alpha, \beta > 0$, $K,K' > 0$, $0 < p < 1$ and
-$\alpha, \beta, K, K'$ are independent of each other. Then we have
+$\alpha, \beta, K, K'$ are independent of each other. We shall also
+use the density function of $X$, which we define as
+$f_X(x) = {\pd F_X(x) \over \pd x}$.
+Then we have
 \beam\label{eq:xxie1.0}
- \wt u(\phi,F_X)\nonumber
+ \wt u(F_X, \phi)\nonumber
   &=&
@@
 We conclude that
-$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
-equal preference on the $(\alpha, K)$-plane. Moving along this
-curve in the direction of increasing $\alpha$, one expects the values
-of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
-should appear positively dependent. This is indeed the case for some
+$\wt u_{\rm max}$ increases with $\alpha$ and decreases with
+$K$. Therefore there is a curve of  equal preference on the $(\alpha,
+K)$-plane. Moving along this curve in the direction of increasing
+$\alpha$, one expects the values of $K$ to increase too, i.e., the
+estimated values of $\alpha$ and $K$ 
+should appear positively dependent. Figure~\ref{fig:preference_pareto}
+illustrates this scenario for $\xi = 1/2, 4$ for the power-utility
+\fct\ \eqref{eq:u}.
+In fact, this positive dependence is indeed observed for some
 real return data, e.g. the ``Energy'', ``Consumer Staples'' and
@@
 \par
-We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto}
-for the power-utility \fct\ \eqref{eq:u}.
 A particularly interesting special case is when the left and the right
@@
 \begin{eqnarray}
-  \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
+  \wt u(F_X, \phi) &=& {\alpha \over 2} K^\alpha (1 + b)
   \int_{0}^\infty {
@@
   \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
+    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1_A.pdf}    
   \end{minipage}\hfill
   \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
+    \includegraphics[width=\textwidth]{phi_hat_pareto4_A.pdf}
   \end{minipage}
@@
     as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
-    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
+    \eqref{eq:pareto} with $K'=0.012$, $\beta = 1.4$. We choose
     the utility \fct\ $u$ from \eqref{eq:u}
@@
   \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
+    \includegraphics[width=\textwidth]{preference_pareto5e-1_A.pdf}
   \end{minipage}\hfill
   \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{preference_pareto4.pdf}
+    \includegraphics[width=\textwidth]{preference_pareto4_A.pdf}
   \end{minipage}
-  \caption{The quantities $I(\alpha, K)$ for the power-utility \fct\
+  \caption{$\tilde u_{\text{max}}(\alpha, K)$ for the power-utility \fct\
     with $\xi=1/2$ (left) and $\xi=4$ (right). The graphs clearly show
-    that  $I(\alpha, K)$ increases with $\alpha$ and decreases with
-    $K$.
+    that  $\tilde u_{\text{max}}(\alpha, K)$ increases with $\alpha$
+    and decreases with $K$. $K' = 0.012$ and $\beta = 1.4$ are fixed.
   }
@@
   \begin{eqnarray*}
-    && {\pd \E h(X) \over \pd \theta} \\
+    {\pd \E h(X) \over \pd \theta}
+    &=&
+    {\pd \over \pd \theta}
+    \int_a^b h(x)
+    f(x, \theta) dx \\
     &=& \int_a^b h(x)

### 2017-05-10 papers/FX/draft.tex
Revised the pictures in the case of equal tail indices.

@@
     ``Consumer Staples'' and ``Information Technology'' sectors.
-    Each circle corresponds to a Hill estimate $\hat\alpha_{50}$; the grey
+    Each circle corresponds to a Hill estimate $\hat\alpha_{50}$; the gray
     triangles above and below it mark the 97.5\% and 2.5\% quantiles
@@
     when the test statistic is outside the intervals $[q_{0.075},q_{0.925}]$,
-    $[q_{0.05},q_{0.95}]$,  $[q_{0.025},q_{0.975}]$, repectively, where
+    $[q_{0.05},q_{0.95}]$,  $[q_{0.025},q_{0.975}]$, respectively, where
     $q_p$ is the $p$-quantile of the limiting 
@@
 500 index. The ``null hypothesis'' is that the tail-index and scale parameter remain the
-same througout the selected period of time. For most stocks, the hypothesis cannot be rejected even at the 85\% level.
+same throughout the selected period of time. For most stocks, the hypothesis cannot be rejected even at the 85\% level.
 This fact may be an indication that the \ds\ inside a series is rather homogeneous.
@@
     of the test. In particular, if both degrees of freedom are
-    relatively large it loses the capablity of distinguishing between the \ds s.
+    relatively large it loses the capability of distinguishing between the \ds s.
     {\em Right}: Comparison of the \asy\ distribution of the test statistic
@@
 $\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$
-provided that the soluton exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
+provided that the solution exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
 follows by an application of the Kesten-Goldie theorem; see Kesten \cite{kesten:1973}, Goldie \cite{goldie:1991}, cf. 
@@
 Not all choices of $\alpha$- and $\beta$-parameters in the model \eqref{eq:8} allow for an
-application of the Kesten theory. For example, assume that only the diagonal elements $\alpha_{ii}$ and $\beta_{ii}$ are positive.
-Then $\bfA_t$ is diagonal and, hence, the condition that $\bfA_1\cdots\bfA_n$ have positive entries for sufficiently large $n$ 
-cannot be satisfied. In the latter situation, both $(X_{1,t})$ and $(X_{2,t})$ are univariate \garch\ processes. Assuming the
-conditions of the univariate Kesten-Goldie theorem for each component process, $(X_{1,t})$ and $(X_{2,t})$ have power-law tails
-with indices $\kappa_1$ and $\kappa_2$, respecticely,  given by the solutions to the equations 
-$\E [(\alpha_{ii} Z_{i,t}^2+\beta_{ii})^{\kappa_i/2}]=1$, $i=1,2$.
-In this model, one can introduce dependence between the two component series $(X_{1,t})$ and $(X_{2,t})$ by assuming dependence between the noise variables $Z_{1,t}$ and $Z_{2,t}$.
-Another situation when the Kesten theory fails
-appears when $\bfA_t$ is an upper or lower triangle matrix: then the products  $\bfA_1\cdots\bfA_n$ are always of the same triangular type.
-Similar remarks apply when one considers a CCC model in general dimension. Of course, one may argue that the latter models
-are degenerate since they do not allow for a linear relationship between all squared volatilities
-on a given day.
+application of the Kesten theory. For example, assume that only the
+diagonal elements $\alpha_{ii}$ and $\beta_{ii}$ are positive.
+Then $\bfA_t$ is diagonal and, hence, the condition that
+$\bfA_1\cdots\bfA_n$ have positive entries for sufficiently large $n$ 
+cannot be satisfied. In the latter situation, both $(X_{1,t})$ and
+$(X_{2,t})$ are univariate \garch\ processes. Assuming the 
+conditions of the univariate Kesten-Goldie theorem for each component
+process, $(X_{1,t})$ and $(X_{2,t})$ have power-law tails 
+with indices $\kappa_1$ and $\kappa_2$, respectively,  given by the solutions to the equations 
+$\E [(\alpha_{ii} Z_{i,t}^2+\beta_{ii})^{\kappa_i/2}]=1$, $i=1,2$. In
+this model, one can introduce dependence between the two component
+series $(X_{1,t})$ and $(X_{2,t})$ by assuming dependence between the
+noise variables $Z_{1,t}$ and $Z_{2,t}$. Another situation when the
+Kesten theory fails 
+appears when $\bfA_t$ is an upper or lower triangle matrix: then the
+products  $\bfA_1\cdots\bfA_n$ are always of the same triangular
+type. 
+Similar remarks apply when one considers a CCC model in general
+dimension. Of course, one may argue that the latter models 
+are degenerate since they do not allow for a linear relationship
+between all squared volatilities on a given day.
 
@@
  weight given to the expected return of $C$ when $C$ is below the
- disappoinment threshold $\delta v$.
+ disappointment threshold $\delta v$.
  If $b=0$, preferences are the classical expected utility. If
@@
 \]
-is not monotonic. To see this we may plug in \eqref{eq:hjyr} and
-re-write the above as
-\begin{equation}
+is not monotone.\footnote {
+To see this we may plug \eqref{eq:hjyr} in $U_{\text{all}}$ and
+re-write it as
+\[
 U_{\text{all}} =
@@
   }_{U(x)} \right\}
-  \label{eq:frg}
-\end{equation}
+\]
 Direct computation gives
@@
 expression at $x' = 0$. This yields
-\[
-  {1 \over \xi}
-  {\pd U(x) \over \pd x}
-  = {x' \over a^{\xi + 1}} + o(x')
-\]
+${1 \over \xi} {\pd U(x) \over \pd x} = {x' \over a^{\xi + 1}} + o(x')$
 So we see that when $x$ is sufficiently large
@@
 point $x = q$, we have
-\[
+$
 \lim_{x \downarrow 0} \left(
@@
 } < 0
-\]
-So $U(x)$ and hence \eqref{eq:frg} is not monotonic. we resort to
-numerical methods to gain some understanding of how $\tilde u(F_X,
-\phi)$ changes with $\alpha$.
-For this reason, we replace $\phi$ in the first integral in
-\eqref{eq:pp} by the optimal $\hat\phi$. The resulting quantity is
-denoted by $I(\alpha, K)$.
-\par
+$
+So $U(x)$ and hence $U_{\text{all}}$ is not monotone.}
+So, unlike the case when $\beta$ and $K'$
+are independent of $\alpha$ and $K$, we cannot use Lemma \ref{lemma:I}
+to determine the monotonicity of $\tilde u_{\rm max}(\alpha, K)$.
+Instead, we resort to numerical methods to gain some understanding of
+how $\tilde u(F_X, \phi)$ changes with $\alpha$.
+
 The value of $\hat\phi$ can be calculated
@@
 given values of $K, K', \alpha, \beta$.  This is shown in 
-Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
+Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\
+\eqref{eq:u}.
+One can see that, when $K'$ and $\beta$ are fixed, $\tilde
+u_{\text{max}}(\alpha, K)$ and hence $\hat\phi$ increases 
+with $\alpha$ and decreases with $K$. This is a verification of Lemma
+\ref{lemma:I}; however, when $\beta$ and $K'$ are equal to $\alpha$
+and $K$, respectively, $\tilde u_{\text{max}}(\alpha, K)$ is not
+monotone with respect to $\alpha$ or $K$. Rather, there is a most
+preferred value of $\alpha$ for each value of $K$. Furthermore, this
+most preferred value changes only slightly when the value of $K$
+changes. Also notice that the value of $\xi$ does make a qualitative
+difference in the behavior of $\tilde u_{\text{max}}(\alpha, K)$ and
+$\hat\phi$, as is seen in the 3rd and the 4th plots of
+\ref{fig:phi_hat_pareto}. So the power-form of the utility function
+and the equality of the lower- and upper-tail-indices are
+responsible for the observed behavior. If this form of utility
+function is common among investors in a market, one can expect the 
+equity return series in the market to have very similar tail indices.
 \begin{figure}[htb!]
-  \begin{minipage}{0.5\linewidth}
+  \begin{minipage}{0.25\linewidth}
     \includegraphics[width=\textwidth]{phi_hat_pareto5e-1_A.pdf}    
   \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
+  \begin{minipage}{0.25\linewidth}
     \includegraphics[width=\textwidth]{phi_hat_pareto4_A.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.25\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
+  \end{minipage}\hfill
+  \begin{minipage}{0.25\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
   \end{minipage}
- \caption{The graphs show $\hat\phi$, the optimal equity allocation
-    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
-    \eqref{eq:pareto} with $K'=0.012$, $\beta = 1.4$. We choose
-    the utility \fct\ $u$ from \eqref{eq:u}
-    for $\xi = 1/2$ (left) and $\xi = 4$
-    (right).
+ \caption{
+   {\em The 1st and 2nd} graphs show $\hat\phi$, the optimal equity
+   allocation as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
+   \eqref{eq:pareto} with $K'=0.012$, $\beta = 1.4$.
+   {\em The 3rd and 4th} graphs show $\hat\phi$ as a \fct\ of $\alpha$
+   and $K$ with $\beta = \alpha$ and $K' = K$.
+   We choose the utility \fct\ $u$ from \eqref{eq:u} for $\xi = 1/2$
+   and $\xi = 4$.
   }
@@
 \begin{figure}[htb!]
-  \begin{minipage}{0.5\linewidth}
+  \begin{minipage}{0.25\linewidth}
     \includegraphics[width=\textwidth]{preference_pareto5e-1_A.pdf}
   \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
+  \begin{minipage}{0.25\linewidth}
     \includegraphics[width=\textwidth]{preference_pareto4_A.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.25\linewidth}
+    \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.25\linewidth}
+    \includegraphics[width=\textwidth]{preference_pareto4.pdf}
   \end{minipage}
-  \caption{$\tilde u_{\text{max}}(\alpha, K)$ for the power-utility \fct\
-    with $\xi=1/2$ (left) and $\xi=4$ (right). The graphs clearly show
-    that  $\tilde u_{\text{max}}(\alpha, K)$ increases with $\alpha$
-    and decreases with $K$. $K' = 0.012$ and $\beta = 1.4$ are fixed.
+  \caption{
+   {\em The 1st and 2nd} graphs show
+   $\tilde u_{\text{max}}(\alpha, K)$, as a \fct\ of $\alpha$ and $K$
+   in the two-sided Pareto model \eqref{eq:pareto} with $K'=0.012$,
+   $\beta = 1.4$.
+   Clearly, $\tilde u_{\text{max}}(\alpha, K)$ increases with $\alpha$
+   and decreases with $K$ when $K'$ and $\beta$ are fixed.
+   {\em The 3rd and 4th} graphs show
+   $\tilde u_{\text{max}}(\alpha, K)$ as a \fct\ of $\alpha$
+   and $K$ with $\beta = \alpha$ and $K' = K$.
+   We choose the utility \fct\ $u$ from \eqref{eq:u} for $\xi = 1/2$
+   and $\xi = 4$.
+% $\tilde u_{\text{max}}(\alpha, K)$ for the power-utility \fct\
+%     with $\xi=1/2$ (left) and $\xi=4$ (right). The graphs clearly show
+%     that  $\tilde u_{\text{max}}(\alpha, K)$ increases with $\alpha$
+%     and decreases with $K$. $K' = 0.012$ and $\beta = 1.4$ are fixed.
   }
@@
 $b = 0$. As shown in the proof of lemma \ref{lemma:III}, the function
-$U_{\text{all}}$ is monotonically increasing or decreasing when $\xi$
+$U_{\text{all}}$ is monotone increasing or decreasing when $\xi$
 and $(1 - \phi) e^r / \phi$ satisfy certain conditions. Moreover, it
@@
 \ref{lemma:III}, when $b = 0$ and $u(\cdot)$
-takes the power-form of \eqref{eq:hjyr}, $U_{\text{all}}$ is monotonic
+takes the power-form of \eqref{eq:hjyr}, $U_{\text{all}}$ is monotone
 depending on the values of $\xi$ and $(1 - \phi) e^r / \phi$. As given
@@
 it follows from the above inequality $\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
-Thus by lemma \ref{lemma:I}, $\wt u_{\text{max}}$ is monotonically
+Thus by lemma \ref{lemma:I}, $\wt u_{\text{max}}$ is monotone
 increasing/decreasing with $\alpha$ when $U_{\text{all}}(\cdot)$ is
-monotonically decreasing/increasing.
+monotone decreasing/increasing.
 
@@
 \appendix
-\section{A monontonicity lemma}
+\section{A monotonicity lemma}
 \setcounter{equation}{0}
@@
   Let $X \sim F$ and assume function $h(\cdot)$ is defined on $(a, b)$
-  and is monotonic throughout this interval.
+  and is monotone throughout this interval.
   Moreover, we assume $h(x)$ and $f(x, \theta)$ satisfy
@@
   $x_0$ being located in the interior of $(a, b)$ and
-  $h(\cdot)$ being monotonic imply $h(x_0) < \infty$.
+  $h(\cdot)$ being monotone imply $h(x_0) < \infty$.
   \begin{enumerate}
@@
   While the denominator of the right side of ${\pd f(x, \alpha) \over \pd \alpha}$ is
-  always positive, its nummerator $P(x^2, \alpha)$ has a single root:
+  always positive, its numerator $P(x^2, \alpha)$ has a single root:
   \begin{equation}
@@
 \item If $\max\{a, 1\} < \xi$, $U_{\text{all}}(\cdot)$ is
-  monotonically decreasing.
+  monotone decreasing.
 \item If $a < \xi < 1$ and $(a + y_-)/(a y_- + 1) <
-  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotonically
+  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
   decreasing.
-\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotonically
+\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotone
   increasing.
 \item If $1 < \xi < a$ and $(a + y_+)/(a y_+ + 1) >
-  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotonically
+  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
   increasing.
-\item In other case, $U_{\text{all}}(\cdot)$ is not monotonic.
+\item In other case, $U_{\text{all}}(\cdot)$ is not monotone.
 \end{enumerate}
@@
   convenience of writing, let $a = (1 - \phi)e^r/\phi$.
-  First of all, we find the conditions on which $U(x)$ is monotonic.
+  First of all, we find the conditions on which $U(x)$ is monotone.
   Direct differentiation yields
@@
     real, ${\pd (f -g)(y) \over \pd y} = 0$ has no solution on
-    $(1, \infty)$; $(f -g)(y)$ is monotonic on $(1, \infty)$.
+    $(1, \infty)$; $(f -g)(y)$ is monotone on $(1, \infty)$.
     \begin{enumerate}
     \item If in addition $\xi > 1$, i.e. $1 < a < \xi$, $(f - g)(y)$
-      is monotonically increasing on $(1, \infty)$, because the
+      is monotone increasing on $(1, \infty)$, because the
       coefficient of the $y^2$ term of the numerator of
@@
       to $f(y) = g(y)$. It can be concluded that $U(\cdot)$ is
-      monotonic on $(0, \infty)$. Furthermore
+      monotone on $(0, \infty)$. Furthermore
         \begin{eqnarray*}
@@
         Thus ${\pd U(x) \over \pd x} > 0$ for $x \in (0, \infty)$;
-        $U_{\text{all}}$ is monotonically decreasing.
+        $U_{\text{all}}$ is monotone decreasing.
     \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar
-      argument as in the previous case, $U_{\text{all}}$ is monotonically
+      argument as in the previous case, $U_{\text{all}}$ is monotone
       increasing on $(0, \infty)$.
@@
     This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
-    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotonically increasing
+    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotone increasing
     on $(1, \infty)$. Following the same analysis as in the case
-    (1.a), one can see $U_{\text{all}}$ is monotonically decreasing.
+    (1.a), one can see $U_{\text{all}}$ is monotone decreasing.
 
@@
     If $(f-g)(y_-) < 0$, $(f-g)(y) = 0$ has no solution on $(1,
-    \infty)$. Hence $U(\cdot)$ is monotonic on both $(0, \infty)$.
+    \infty)$. Hence $U(\cdot)$ is monotone on both $(0, \infty)$.
     The same analysis as in the case (1.a) shows $U_{\text{all}}$
-    is monotonically decreasing on $(0, \infty)$.
+    is monotone decreasing on $(0, \infty)$.
 
     If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
-    \infty)$. So $U(\cdot)$ is not monotonic.
+    \infty)$. So $U(\cdot)$ is not monotone.
 
@@
       minimum. If $(f-g)(y_+) > 0$, there is no solution to
-      $(f-g)(y) = 0$ on $(1, \infty)$; $U(x)$ is monotonic. By the
+      $(f-g)(y) = 0$ on $(1, \infty)$; $U(x)$ is monotone. By the
       same analysis as in the case (1.a), we know $U(x)$ is
-      monotonically decreasing and $U_{\text{all}}$ is monotonically increasing.
+      monotone decreasing and $U_{\text{all}}$ is monotone increasing.
 
       if $(f-g)(y_+) < 0$, there must be a solution to $(f-g)(y) = 0$
-      on $(y_+, \infty)$. $U_{\text{all}}$ is not monotonic.
+      on $(y_+, \infty)$. $U_{\text{all}}$ is not monotone.
   \end{enumerate}

### 2017-05-11 papers/FX/draft.tex
Revised the FX paper on the Student's t case.

@@
     {\hat \alpha_X - \alpha_X} \\
-    {\hat \alpha_X - \alpha_X} \\
+    {\hat \alpha_Y - \alpha_Y} \\
   \end{pmatrix} \overset{d}{\to}
@@
 Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\
-\eqref{eq:u}.
+\eqref{eq:u}. The corresponding $\tilde u_{\rm max}(\alpha, K)$
+is shown in figure \ref{fig:preference_pareto}.
 One can see that, when $K'$ and $\beta$ are fixed, $\tilde
-u_{\text{max}}(\alpha, K)$ and hence $\hat\phi$ increases 
+u_{\rm max}(\alpha, K)$ and $\hat\phi$ increase 
 with $\alpha$ and decreases with $K$. This is a verification of Lemma
 \ref{lemma:I}; however, when $\beta$ and $K'$ are equal to $\alpha$
-and $K$, respectively, $\tilde u_{\text{max}}(\alpha, K)$ is not
-monotone with respect to $\alpha$ or $K$. Rather, there is a most
-preferred value of $\alpha$ for each value of $K$. Furthermore, this
-most preferred value changes only slightly when the value of $K$
-changes. Also notice that the value of $\xi$ does make a qualitative
-difference in the behavior of $\tilde u_{\text{max}}(\alpha, K)$ and
-$\hat\phi$, as is seen in the 3rd and the 4th plots of
-\ref{fig:phi_hat_pareto}. So the power-form of the utility function
-and the equality of the lower- and upper-tail-indices are
-responsible for the observed behavior. If this form of utility
-function is common among investors in a market, one can expect the 
-equity return series in the market to have very similar tail indices.
+and $K$, $\tilde u_{\rm max}(\alpha, K)$ is monotone
+decreasing with $\alpha$ but is rather insensitive to $K$. On the
+other hand, $\hat\phi$ is not monotone with respect to $\alpha$ or
+$K$. For each fixed $K$, it peaks at an $\alpha$ value somewhere below
+1. For typical realistic $\alpha$ values in $(2, 4)$, $\hat\phi$ is a
+small value below 5\%.
+Since $\tilde u_{\rm max}(\alpha, K)$ is monotone decreasing with
+$\alpha$, investors who seek to maximize it will prefer the smallest
+$\alpha$ in the market, resulting in similar values of $\alpha$ for
+different equities.
 \begin{figure}[htb!]
@@
    {\em The 1st and 2nd} graphs show
-   $\tilde u_{\text{max}}(\alpha, K)$, as a \fct\ of $\alpha$ and $K$
+   $\tilde u_{\rm max}(\alpha, K)$, as a \fct\ of $\alpha$ and $K$
    in the two-sided Pareto model \eqref{eq:pareto} with $K'=0.012$,
    $\beta = 1.4$.
-   Clearly, $\tilde u_{\text{max}}(\alpha, K)$ increases with $\alpha$
+   Clearly, $\tilde u_{\rm max}(\alpha, K)$ increases with $\alpha$
    and decreases with $K$ when $K'$ and $\beta$ are fixed.
    {\em The 3rd and 4th} graphs show
-   $\tilde u_{\text{max}}(\alpha, K)$ as a \fct\ of $\alpha$
+   $\tilde u_{\rm max}(\alpha, K)$ as a \fct\ of $\alpha$
    and $K$ with $\beta = \alpha$ and $K' = K$.
@@
    and $\xi = 4$.
-% $\tilde u_{\text{max}}(\alpha, K)$ for the power-utility \fct\
+% $\tilde u_{\rm max}(\alpha, K)$ for the power-utility \fct\
 %     with $\xi=1/2$ (left) and $\xi=4$ (right). The graphs clearly show
-%     that  $\tilde u_{\text{max}}(\alpha, K)$ increases with $\alpha$
+%     that  $\tilde u_{\rm max}(\alpha, K)$ increases with $\alpha$
 %     and decreases with $K$. $K' = 0.012$ and $\beta = 1.4$ are fixed.
@@
 using lemma \ref{lemma:I} as in the proof of lemma \ref{thrm:I},
-it can be seen $\tilde u_{\text{max}}(\alpha, K)$ is
+it can be seen $\tilde u_{\rm max}(\alpha, K)$ is
 increasing/decreasing with $\alpha$ when $U_{\text{all}}$ is
@@
 it follows from the above inequality $\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
-Thus by lemma \ref{lemma:I}, $\wt u_{\text{max}}$ is monotone
+Thus by lemma \ref{lemma:I}, $\wt u_{\rm max}$ is monotone
 increasing/decreasing with $\alpha$ when $U_{\text{all}}(\cdot)$ is
-monotone decreasing/increasing.
+monotone decreasing/increasing. Accordingly, an investor guided by the
+utility function will seek the largest/smallest $\alpha$ observed in
+the market.
+
+If however $b > 0$, Lemma \ref{lemma:I} is not applicable
+anymore. Nonetheless, numerical analysis lends some insight.
+As shown in figure \ref{fig:htfg}, $\hat\phi$ is monotone increasing
+for all 4 values of $b$, while $\tilde u_{\rm max}(\alpha)$ is
+increasing with $\alpha$ when $b$ is relatively large, but
+decreasing when $b$ is small. We note that a sizable value of $b$
+indicates a conservative, risk-averse investor.
+
+\begin{figure}[htb!]
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
+  \end{minipage}
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_power4.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_power4.pdf}
+  \end{minipage}
+  \caption{$\hat\phi$ (left) and ${\pd \tilde u_{\rm max} \over \pd
+      \alpha}$ (right). {\em top:} $\xi = 1/2$. {\em bottom:} $\xi =
+    4$.
+  }
+  \label{fig:htfg}
+\end{figure}
 
@@
   \end{minipage}
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
-  \end{minipage}
-  \caption{$\hat\phi$ and $G(\alpha)$
+  \caption{$\hat\phi$ and $\tilde u_{\rm max}(\alpha)$
     \label{fig:phi_hat_U}
@@
   \[
-  \tilde u_{\text{max}}(F_X) = \wt u(F_X, \hat\phi)
+  \tilde u_{\rm max}(F_X) = \wt u(F_X, \hat\phi)
   \]
@@
   \begin{eqnarray}
-    {d \tilde u_{\text{max}}(F_X) \over d \alpha}
+    {d \tilde u_{\rm max}(F_X) \over d \alpha}
     &=&
@@
   \end{eqnarray*}
-  As for ${\pd \tilde u_{\text{max}}(F_X) \over \pd K}$, by the same argument as for
-  ${\pd \tilde u_{\text{max}}(F_X) \over \pd \alpha}$, it suffices to show
+  As for ${\pd \tilde u_{\rm max}(F_X) \over \pd K}$, by the same argument as for
+  ${\pd \tilde u_{\rm max}(F_X) \over \pd \alpha}$, it suffices to show
   \[

### 2017-05-11 papers/FX/Preference.r

@@ -43,39 +43,12 @@ pareto.preference <- function(phi, alpha, alpha.r, K, K.r) {
     }, q, Inf)$value;
     y3 <- y3 * alpha.r * K.r^alpha.r * (1 - p);
 
-    y4 <- b * Pareto(q, alpha, alpha.r, K, K.r) *
-        utility(consumption(q, phi, r));
+    y4 <- b * utility(delta.v) * Pareto(q, alpha, alpha.r, K, K.r);
     
-    ## return(c(y1, y2, y3));
     return(y1 + y2 + y3 - y4);
 }
 
 t.preference <- function(phi, nu) {
-    ## q <- Inf;
-    ## if (phi > 0) {
-    ##     q <- log(exp(r) + (delta.v - exp(r))/phi);
-    ## }
-
-    ## y1 <- (1 + b) * integrate(function(x) {
-    ##     U <- utility(consumption(x, phi, r)) * dt(x, nu);
-    ##     return(U);
-    ## }, -Inf, 0)$value;
-
-    ## y2 <- (1 + b) * integrate(function(x) {
-    ##     U <- utility(consumption(x, phi, r)) * dt(x, nu);
-    ##     return(U);
-    ## }, 0, q)$value;
-    ## y2 <- y2 + integrate(function(x) {
-    ##     U <- utility(consumption(0, phi, r-x)) * dt(x, nu);
-    ##     return(U);
-    ## }, q, Inf)$value;
-    ## y2 <- y2 + integrate(function(x) {
-    ##     U <- (1/x^2 + 1/nu)^(-nu/2-1/2) * x^(-nu);
-    ##     return(U);
-    ## }, q, Inf)$value;
-    
-    ## y3 <- b * pt(q, nu) * utility(consumption(q, phi, r));
-    ## return(c(y1, y2, y3));
     q <- Inf;
     if (phi > 0) {
         q <- log(exp(r) + (delta.v - exp(r))/phi);
@@ -93,11 +66,9 @@ t.preference <- function(phi, nu) {
     }, 0, q)$value;
     y2 <- y2 * (1 + b) * (1 - p);
 
-    y4 <- b * Pareto(q, alpha, alpha.r, K, K.r) *
-        utility(consumption(q, phi, r));
+    y3 <- b * pt(q, nu) * utility(delta.v);
     
-    ## return(c(y1, y2, y3));
-    return(y1 + y2 + y3 - y4);
[…]

### 2017-05-12 papers/FX/draft.tex
Fixed the bibliography.

@@
 average, the values of $K$
-in the ``Energy'' and ``Information Technology'' sectors are  larger than those in the ``Consumer Staples'' sector. 
-For a given loss \pro y, 
-a larger
-value of $K$ implies that large losses are more probable. Thus one can conclude that
-these two sectors are considerably riskier than the ``Consumer
-Staples'' sector. {\red Let Casper decide about this sentence.} This is of course a confirmation of one's economic
-instinct.
-\par
+in the ``Energy'' and ``Information Technology'' sectors are  larger
+than those in the ``Consumer Staples'' sector.  For a given loss \pro y, 
+a larger value of $K$ implies that large losses are more
+probable. Thus one can conclude that these two sectors are
+considerably riskier than the ``Consumer Staples'' sector. {\red Let
+  Casper decide about this sentence.} This is of course a confirmation
+of one's economic instinct.
+
 Yet another indication from Figure \ref{fig:sectors_parameters} is
@@
 between $\alpha_A, \alpha_B$ on one hand and between $\alpha_B, \alpha_C$ on the other hand.
-
 \begin{figure}[htb!]
@@
 \end{figure}
-
 % \begin{figure}[htb!]
@@
 % \end{figure}
-
-\par
 To get an idea about the power of the test we run it on a
@@
 \end{figure}
-
-\par
 A major problem in this test is the \asy\ distribution of the test statistic under the
@@
 \end{table}
+We also acknowledge that there is yet another issue concerning the use
+of $T_n$ as a test for a change in the tail-index: It is designed as a
+test for a change in an extreme quantile and hence is not invariant
+under permutation of the observations. This is clear from
+\eqref{eq:4}. If the observations are iid, the null hypothesis is of
+course not rejected with or without permutation; If instead the
+observations are not iid but rather have small changes of the
+tail-index over time, the null hypothesis may be rejected with
+permutation but not rejected without. The latter may happen when the
+changes in the tail-index are small but just big enough for the
+accumulated change between two time points to exceed the given
+critical value. In other words, the test cannot detect slow changes.
+
+We illustrate this point in figure \ref{fig:PairTest:permuted} where
+we apply a random permutation of observations to each series before
+concatenating them and running the test.
+\begin{figure}[htb!]
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{Hoga_Energy_pair_permuted.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{Hoga_CS_pair_permuted.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.33\linewidth}
+    \includegraphics[
+      width=\textwidth,
+      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
+    ]{Hoga_IT_pair_permuted.pdf}
+  \end{minipage}
+  \caption{
+    Test for changing tail-index or scale parameter
+    of losses
+    using Hoga's test based on concatenated series of pairs of
+    stocks.
+    A random permutation is applied to the observations of each
+    series.
+    The green, blue and red points
+    correspond to pairs of stock in a sector 
+    when the test statistic $T_n$ exceeds the $85\%$-, $90\%$-,
+    $95\%$-quantile of the limit \ds .
+    Grey points stand for pairs for which the test statistic is below
+    the \asy\ $85\%$-quantile. Black points represent 
+    pairs for which the computation of $T_n$ fails for given precision  
+    requirements and time limits.
+  }
+  \label{fig:PairTest:permuted} 
+\end{figure}
 
@@
 the Orthogonal GARCH model of
-Alexander and Chibumba \cite{alexander1997multivariate}, its
+Alexander and Chibumba \cite{alexander:chibumba:1996}, its
 generalization GO-GARCH by van der Weide \cite{Weide2002},  the Full Factor GARCH model of Vrontos et al.
@@
    We choose the utility \fct\ $u$ from \eqref{eq:u} for $\xi = 1/2$
-   and $\xi = 4$.
+   and $\xi = 4$. $b = 0.01$ in all cases.
   }
@@
    We choose the utility \fct\ $u$ from \eqref{eq:u} for $\xi = 1/2$
-   and $\xi = 4$.
-% $\tilde u_{\rm max}(\alpha, K)$ for the power-utility \fct\
-%     with $\xi=1/2$ (left) and $\xi=4$ (right). The graphs clearly show
-%     that  $\tilde u_{\rm max}(\alpha, K)$ increases with $\alpha$
-%     and decreases with $K$. $K' = 0.012$ and $\beta = 1.4$ are fixed.
+   and $\xi = 4$. $b = 0.01$ in all cases.
   }

### 2017-05-12 papers/FX/Preference.r

@@ -104,20 +104,21 @@ b <- 0.01;
 ## b <- 0;
 p <- 0.5;
 delta.v <- exp(r)*1.05;
-xi <- 4;
+xi <- 0.5;
+
 
 ## alpha.r <- 3.5;
 ## alpha <- 3;
 ## K.r <- 0.5;
 ## K <- 0.1;
 
-## utility <- function(x) log(x);
-utility <- power.utility;
+utility <- function(x) log(x);
+## utility <- power.utility;
 
 scales <- seq(from=10^(-2.3), to=10^(-1.7), length.out=50);
 ## scales <- seq(from=0.1, to=1, length.out=50);
 ## indices <- seq(from=2, to=5, length.out=40);
-indices <- seq(from=0.1, to=4, length.out=40);
+indices <- seq(from=1.5, to=4, length.out=40);
 phi.hat <- matrix(NA, nrow=length(indices),
                   ncol=length(scales));
 U <- matrix(NA, nrow=length(indices),

### 2017-05-15 papers/FX/draft.tex
Revised the FX paper.

@@
 }
-
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
 \allowdisplaybreaks
@@
 % \end{figure}
+
 To get an idea about the power of the test we run it on a
@@
 \end{figure}
+
 A major problem in this test is the \asy\ distribution of the test statistic under the
@@
 \end{table}
-We also acknowledge that there is yet another issue concerning the use
-of $T_n$ as a test for a change in the tail-index: It is designed as a
-test for a change in an extreme quantile and hence is not invariant
-under permutation of the observations. This is clear from
-\eqref{eq:4}. If the observations are iid, the null hypothesis is of
+
+We also acknowledge that the test is designed to detect a sudden big
+change and cannot detect changes that are small but accumulate over
+time. This is due to the fact that $T_n$ is not invariant under
+permutation of the observations ($X_t$).
+If the observations are iid, the null hypothesis is of 
 course not rejected with or without permutation; If instead the
@@
 changes in the tail-index are small but just big enough for the
-accumulated change between two time points to exceed the given
+accumulated change between two time points to exceed the
 critical value. In other words, the test cannot detect slow changes.
@@
 \]
-is not monotone.\footnote {
+is not monotone. \footnote{
 To see this we may plug \eqref{eq:hjyr} in $U_{\text{all}}$ and
@@
 increasing with $\alpha$ when $b$ is relatively large, but
-decreasing when $b$ is small. We note that a sizable value of $b$
-indicates a conservative, risk-averse investor.
+decreasing with $\alpha$ when $b$ is small. We note that a sizable
+value of $b$ indicates a conservative, risk-averse investor.
 
@@
 
-\subsubsection{log-utility function}
-We first consider the case where the utility function is simply
-$u(\cdot) = \log(\cdot)$. This corresponds to the limiting case of
-\eqref{eq:u} as $\gamma \to 0$.
-
-First notice that when $b = 0$, $\wt u(F, \phi)$ reduces to
-$\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
-Its integrand becomes $\log[C(x, \phi) C(-x, \phi)]$.
-Now observe $\log(C(x, \phi)C(-x, \phi))$ is an increasing function of
-$x$:
-\begin{eqnarray*}
-  {\pd C(x, \phi) C(-x, \phi) \over \pd x}
-  &=&
-  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
-\end{eqnarray*}
-Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
-\ref{lemma:II}, {\red need to look at $\pd \wt u_{\rm max} \over \pd \alpha$?} 
-${\pd \wt u(F, \phi)) \over \pd \alpha} < 0$ for all $\phi$. Hence all
-investors guided by this utility function would be seeking the
-smallest $\alpha$, i.e. the heaviest tail in the market. Here note that
-Student's t is a symmetric distribution; a heavier tail implies not
-only greater down-side risk but also greater up-side potential.
-
-In fact, even when $b > 0$, the same monotonicity property holds. This
-is revealed by numerically maximizing $\wt u(F, \phi)$ with
-respect to $\phi$ for a sequence of selected values of $\alpha$. The
-results are shown in figure~\ref{fig:phi_hat_U}.
-\begin{figure}[htb!]
-  % \begin{subfigure}[b]{0.5\linewidth}
+% \subsubsection{log-utility function}
+% We first consider the case where the utility function is simply
+% $u(\cdot) = \log(\cdot)$. This corresponds to the limiting case of
+% \eqref{eq:u} as $\gamma \to 0$.
+
+% First notice that when $b = 0$, $\wt u(F, \phi)$ reduces to
+% $\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
+% Its integrand becomes $\log[C(x, \phi) C(-x, \phi)]$.
+% Now observe $\log(C(x, \phi)C(-x, \phi))$ is an increasing function of
+% $x$:
+% \begin{eqnarray*}
+%   {\pd C(x, \phi) C(-x, \phi) \over \pd x}
+%   &=&
+%   \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
+% \end{eqnarray*}
+% Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
+% \ref{lemma:II}, {\red need to look at $\pd \wt u_{\rm max} \over \pd \alpha$?} 
+% ${\pd \wt u(F, \phi)) \over \pd \alpha} < 0$ for all $\phi$. Hence all
+% investors guided by this utility function would be seeking the
+% smallest $\alpha$, i.e. the heaviest tail in the market. Here note that
+% Student's t is a symmetric distribution; a heavier tail implies not
+% only greater down-side risk but also greater up-side potential.
+
+% In fact, even when $b > 0$, the same monotonicity property holds. This
+% is revealed by numerically maximizing $\wt u(F, \phi)$ with
+% respect to $\phi$ for a sequence of selected values of $\alpha$. The
+% results are shown in figure~\ref{fig:phi_hat_U}.
+% \begin{figure}[htb!]
+%   \begin{minipage}{0.5\linewidth}
 %     \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
-  %   \caption{$\hat\phi(\alpha)$ of log-utility function}
-  %   \label{fig:phi_hat_t_log}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.5\linewidth}
+%   \end{minipage}\hfill
+%   \begin{minipage}{0.5\linewidth}
 %     \includegraphics[width=\textwidth]{U_b_t_log.pdf}
-  %   \caption{$G(\alpha)$ of log-utility function}
-  %   \label{fig:U_b_t_log}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.5\linewidth}
-  %   \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
-  %   \caption{$\hat\phi(\alpha)$ of power-utility function}
-  %   \label{fig:phi_hat_t_power}
-  % \end{subfigure}
-  % \begin{subfigure}[b]{0.5\linewidth}
-  %   \includegraphics[width=\textwidth]{U_b_t_power.pdf}
-  %   \caption{$G(\alpha)$ of power-utility function}
-  %   \label{fig:U_b_t_power}
-  % \end{subfigure}
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_log.pdf}
-  \end{minipage}
-  \caption{$\hat\phi$ and $\tilde u_{\rm max}(\alpha)$
-    \label{fig:phi_hat_U}
-  }
-\end{figure}
+%   \end{minipage}
+%   \caption{$\hat\phi$ and $\tilde u_{\rm max}(\alpha)$
+%     \label{fig:phi_hat_U}
+%   }
+% \end{figure}
 
@@
   So the second term of \eqref{eq:xxie3} vanishes. It remains to show
-  \begin{eqnarray*}
-    \left.{
-        \pd \wt u(\alpha, \phi) \over \pd \alpha}
-    \right|_{\phi = \hat\phi} &>& 0
-  \end{eqnarray*}
-  From \eqref{eq:xxie1.0}, it follows
+  the first term is positive. From \eqref{eq:xxie1.0}, it follows
   \begin{eqnarray*}
@@
   \begin{eqnarray*}
-    && {\pd f_X(x; \alpha, K) \over \pd \alpha} \\
-    &=& {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
-    &=&
-    - {K^\alpha \over (K - x)^{\alpha + 1}}
+    {\pd f_X(x; \alpha, K) \over \pd \alpha}
+    = {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} 
+    = - {K^\alpha \over (K - x)^{\alpha + 1}}
     \left[
@@
   \end{eqnarray*}
-  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
-  \[
-  {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
-  \]
-  and when $K(1 - e^{1/\alpha}) < x < 0$,
-  \[
-  {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
-  \]
+  It is easily checked
+  \[
+  {\partial \over \partial \alpha}
+  {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
+  \left\{
+    \begin{array}{ll}
+      > 0 & \text{ when } x < K(1 - e^{1/\alpha}) < 0 \\
+      < 0 & \text{ when } K(1 - e^{1/\alpha}) < x < 0
+    \end{array}
+  \right.
+  \]
+% that when $x < K(1 - e^{1/\alpha}) < 0$,
+%   \[
+%   {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+%   \]
+%   and when $K(1 - e^{1/\alpha}) < x < 0$,
+%   \[
+%   {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+%   \]
   This is the second case of lemma \ref{lemma:I}. So we have
-  \begin{eqnarray*}
-    {\pd \wt u(\alpha, \phi) \over \pd \alpha} &>& 0
-  \end{eqnarray*}
-  As for ${\pd \tilde u_{\rm max}(F_X) \over \pd K}$, by the same argument as for
-  ${\pd \tilde u_{\rm max}(F_X) \over \pd \alpha}$, it suffices to show
-  \[
-  {\pd \wt u(K, \phi) \over \pd K} < 0
-  \]
+  ${\pd \wt u(\alpha, \phi) \over \pd \alpha} > 0$.
+  As for ${\pd \tilde u_{\rm max}(F_X) \over \pd K}$, by the same
+  argument, it suffices to show ${\pd \wt u(K, \phi) \over \pd K} < 0$.
   We have
@@
   \begin{eqnarray*}
-    && {\pd f_X(x; \alpha, K) \over \pd K} \\
-    &=& {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
-    &=&
-    -\alpha K^{\alpha - 1} {
+    {\pd f_X(x; \alpha, K) \over \pd K}
+    = {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
+   = -\alpha K^{\alpha - 1} {
       \alpha x + K
@@
   \end{eqnarray*}
-  Clearly, when $x < -K/\alpha$
-  \[
-  {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
-  \]
-  and when $-K/\alpha < x < 0$
-  \[
-  {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
-  \]
-  So by the 1st case of remark \ref{remark:I} we conclude
-  \[
-  {\pd \wt u(K, \phi) \over \pd K} < 0  
-  \]
+  Clearly,
+  \[
+  {\partial \over \partial K}
+  {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
+  \left\{
+    \begin{array}{ll}
+      > 0 & \text{ when } x < -K/\alpha < 0\\
+      < 0 & \text{ when } -K/\alpha < x < 0
+    \end{array}
+  \right.
+  \]
+  % Clearly, when $x < -K/\alpha$
+  % \[
+  % {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
+  % \]
+  % and when $-K/\alpha < x < 0$
+  % \[
+  % {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
+  % \]
+  So by the 1st case of Remark \ref{remark:I} we conclude $  {\pd \wt u(K, \phi) \over \pd K} < 0$.
 \end{proof}

### 2017-06-02 papers/FX/draft.tex
Some revision on both papers.

@@
 \title{Do return series have power-law tails with the same index?}
-\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
+t\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
 \date{\today}
@@
 estimates are based on 1304 observations of daily return 
-data from 4th January 2010 to 31 December 2014.
+data from 4 January 2010 to 31 December 2014.
 \par
@@
 In view of the wide \asy\ confidence bands for the estimators of tail-indices, 
-are the estimated tail-indices from different series really distinct?
+are the tail-indices from different series really distinct?
 \item
-Are there some {\em theoretical} reasons in favor of the fact that the tail-indices from different series are {\em not} 
+Are there some {\em theoretical} reasons supporting the fact that the tail-indices from different series are {\em not} 
 distinct?
@@
 some of its virtues and vices. In addition to tail-index estimation we also discuss the related problem of
-estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). {\red check} It turns out 
-that the simultaneous estimation of the tail-index and the scale parameter are strongly related.
+estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). 
+%{\red check} It turns out 
+%that the simultaneous estimation of the tail-index and the scale parameter are strongly related.
 \par
@@
 introduced by Routledge and Zin \cite{routledge2010generalized}. GDA
-is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991}. Gul derives the DA from first principles (axiomatic).
+is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991} who derived DA from first principles (axiomatic).
 
@@
 for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
-useful description of the risk. Alternatively, these parameters can be
+useful description of the risk, for example in terms of high quantiles such as
+Value-at-Risk. Alternatively, these parameters can be
 used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
@@
 A major problem for Hill estimation is the choice of the number $k$ of
-order statistics. As  a matter of fact,
+upper order statistics. As  a matter of fact,
 if $k$ is too large the order statistics are too close to the center
@@
 \eeam
-i.e., $\hat \alpha$ is \asy ally unbiased and has variance
+i.e., $\hat \alpha_k$ is \asy ally unbiased and has variance
 $\alpha^2/k$. Since $k/n\to 0$ this means that the 
@@
 given a theoretical model such as the GARCH, these conditions can be
-verified based on the properties of the model. If they are not satisfied the 
+verified based on the theoretical properties of the model. If they are not satisfied the 
 Hill estimator may exhibit significant bias; see Embrechts et
@@
 Hoga's \cite{hoga:2016} test about the change of extreme quantiles
+in a sample
 may provide some further insight about how similar these tail-indices are.
@@
       \alpha}$ (bottom) on $\log_{10}$-scale of stocks in sectors
-    of the S\&P 500 index. The points are the estimated values, the bars 
+    of the S\&P 500 index. The estimates are ordered according to the 
+    corresponding estimated $\alpha$-values.
+    The points are the estimated values, the bars 
     the \asy\ 95\%-confidence intervals; the confidence bands of the
@@
 \par
-Secondly, it appears that there is positive dependence between the  values of $\alpha$ and $K$. 
-This prompts Pareto tails of the return
-distributions. For more explanations, see
-Section~\ref{sec:pareto_tail}.
+Secondly, it appears that there is positive dependence between the
+values of $\alpha$ and $K$.
+As argued in Section~\ref{sec:pareto_tail}, this is consistent with
+the assumption that the return series have Pareto tails on both sides
+with the tail parameters on each side independent of those on the
+other.
+
 \par
@@
 probable. Thus one can conclude that these two sectors are
-considerably riskier than the ``Consumer Staples'' sector. {\red Let
-  Casper decide about this sentence.} This is of course a confirmation
-of one's economic instinct.
+considerably riskier than the ``Consumer Staples'' sector. {\red Casper:
+we leave this sentence to your discretion.} This is of course a
+confirmation of one's economic instinct.
 
@@
 are similar in riskiness, the dependence  between $\alpha$ and $K$ is
-stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, moving along
-a curve of equal preference in the direction of increasing $\alpha$, $K$ also
+stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, when moving along
+a curve of equal preference in the direction of increasing $\alpha$, the parameter $K$ also
 increases. So the strong positive dependence seen in the ``Energy''
-sector suggests these stocks might have very similar investor
+sector suggests that these stocks might have very similar investor
 preferences. This in turn may be attributed to stronger business
@@
     ``Information Technology'' 
-    sectors of S\&P 500. The test statistic in \eqref{eq:x1} is based
+    sectors of S\&P 500. The test statistic in $\hat\alpha_X-\hat \alpha_Y$ is based
     on Hill estimates 
@@
     requirements and time limits.
-    The same number of upper order statistics are used for both tests.}
+    The same number (50) of upper order statistics is used for both tests.}
   \label{fig:PairTest} 
 \end{figure}
-It indicates that tail-indices of equities in the
+They indicate that tail-indices of equities in the
 ``Energy'' or ``Information Technology'' sectors are more variable 
@@
 \eqref{eq:3} change inside a sample, then it is likely that the
-extreme quantiles change as well. We will test for a change of tail
-index or scale parameter in this indirect way.
+extreme quantiles change as well. We will test for a change of tail-index or scale parameter in this indirect way.
 \par
@@
 \end{figure}
-% \begin{figure}[htb!]
-%   \begin{minipage}{0.5\linewidth}
-%     \includegraphics[
-%       width=\textwidth,
-%       trim={0.3cm, 1cm, 1cm, 1cm}, clip
-%     ]{Hoga_Energy_pair.pdf}
-%   \end{minipage}\hfill
-%   \begin{minipage}{0.5\linewidth}
-%     \includegraphics[
-%       width=\textwidth,
-%       trim={0.3cm, 1cm, 1cm, 1cm}, clip
-%     ]{Hoga_CS_pair.pdf}
-%   \end{minipage}
-%   \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail-index or scale parameter based on concatenated series in the ``Energy'' and
-%     ``Consumer Staples'' sectors of
-%     S\&P 500. The green, blue and red colors represent, respectively, the cases when  
-%     the 85\%, 90\% and 95\% quantiles of the limit distribution of $T_n$ is exceeded.
-%     Grey points stand for pairs for which $T_n$ is below
-%     the 85\% \asy\ quantile;
-%     black points represent pairs for which
-%     the computation of $T_n$ fails for given precision
-%     requirements and time limits.}
-%   \label{fig:Hoga_pair}
-% \end{figure}
 
@@
     {\em Left}: Test  of concatenated $t$-samples with different
-    degrees of freedom ($\alpha$). Numbers on the axes are the degrees of freedom
+    degrees of freedom $\alpha$. Numbers on the axes are the degrees of freedom
     in the subsamples. For an interpretation of the colored bullets,
-    see the bottom row of Figure~\ref{fig:PairTest}. The graph shows the limited power
+    see the caption for the bottom row of Figure~\ref{fig:PairTest}. The graph shows the limited power
     of the test. In particular, if both degrees of freedom are
@@
 
-A major problem in this test is the \asy\ distribution of the test statistic under the
+A major problem of this test is the \asy\ distribution of the test statistic under the
 null hypothesis.  The rate at which the finite-sample distribution tends to its limit is not known.
-To find out about this problem we compared the \ds s of $T_n$ 
-with $n=1304$ for $t$-distributed $X_t$ with  3 and 4 degrees of freedom ($\alpha$) and
+To find out about this problem we compared the \ds s of $T_{1304}$ 
+for $t$-distributed $X_t$ with  $\alpha=3$ and $\alpha=4$ degrees of freedom with
 the limit \ds\ of $T_n$.
@@
   \caption{Quantiles of the test statistic $T_n$ for $n=1304$
-    $t$-distributed samples with 3 and 4 degrees of freedom ($\alpha$)
+    $t$-distributed samples with $\alpha=3$ and $\alpha=4$ 
+degrees of freedom
     as well as the corresponding quantiles for the limiting \ds\ of
@@
 
-We also acknowledge that the test is designed to detect a sudden big
-change and cannot detect changes that are small but accumulate over
-time. This is due to the fact that $T_n$ is not invariant under
-permutation of the observations ($X_t$).
-If the observations are iid, the null hypothesis is of 
-course not rejected with or without permutation; If instead the
-observations are not iid but rather have small changes of the
-tail-index over time, the null hypothesis may be rejected with
-permutation but not rejected without. The latter may happen when the
-changes in the tail-index are small but just big enough for the
-accumulated change between two time points to exceed the
-critical value. In other words, the test cannot detect slow changes.
-
-We illustrate this point in figure \ref{fig:PairTest:permuted} where
-we apply a random permutation of observations to each series before
-concatenating them and running the test.
+Figure~\ref{fig:PairTest:permuted} points at another shortcoming
+of the test: we show $T_n$ for an 
+arbitrarily chosen random permutation of the concatenated data from two 
+different stocks. In this case the null hypothesis that two series
+have the same tail-index is rejected much more often, as a comparison
+with the bottom graphs of
+Figure~\ref{fig:PairTest} shows. If the data in the concatenated 
+series were iid a random permutation would not change the 
+\ds\ of $T_n$. Thus the value of the test statistic $T_n$  
+strongly depends on the dependence structure of the underlying data
+and therefore a test based on $T_n$ may be misleading.
+
 \begin{figure}[htb!]
@@
 follows by an application of the Kesten-Goldie theorem; see Kesten \cite{kesten:1973}, Goldie \cite{goldie:1991}, cf. 
-Buraczewski et al. \cite{buraczewski:damek:mikosch:2016}. The latter result ensures power-law tails  for
+Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} for a 
+recent textbook treatment. The latter result ensures power-law tails  for
 the strictly stationary  solution $(Y_t)$ to the \sre\ 
@@
 \earr\right)\,\left(\barr{l}\sigma_{1,t-1}^2\\\sigma_{2,t-1}^2\earr
-\right)\,.
+\right)\,,
 \eeam
+for positive $\alpha_{0i}$ and suitable non-negative 
+$\alpha_{ij},\beta_{ij}$, $i,j=1,2$.
 Writing
@@
 dimension. Of course, one may argue that the latter models 
-are degenerate since they do not allow for a linear relationship
+are not natural: they are degenerate since they do not allow 
+for a linear relationship
 between all squared volatilities on a given day.
@@
   \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
-  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\
-  &=&
-  \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
+  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\,,
+  %&=&
+  %\E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
 \eeao
@@
 Routledge and Zin assumed a power-law utility function 
-\beam\label{eq:u}
-u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
-\eeam
+%\beam\label{eq:u}
+%u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
+%\eeam
 % In practice it is more often to observe $\gamma > 1$ {\blue Let's
 % consult Casper for references.}
-We focus on the case $\gamma > 1$ and hence adopt a more convenient
-notation:
+%We focus on the case $\gamma > 1$ and hence adopt a more convenient
+%notation:
 \beam\label{eq:hjyr}
@@
 \beam\label{eq:xxie1}
-  C(x) = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
+  C(X) = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
 \eeam
@@
 \eeao
-where $F_X$ is the \df\ of $X$ with density $f_X$, and
+where $F_X$ is the \df\ of $X$  and
 \beao
@@
 riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
-$q > r $. For given $b,\delta,v$, $\wt u$ is only a functional of $\phi$ and $F_X$,
+$q > r $. For given $b,\delta,v$, the \fct al $\wt u$  depends only on $\phi$ and $F_X$,
 $\wt u=\wt u (F_X, \phi)$.
@@
 where
-$\alpha, \beta > 0$, $K,K' > 0$, $0 < p < 1$ and
-$\alpha, \beta, K, K'$ are independent of each other. We shall also
-use the density function of $X$, which we define as
-$f_X(x) = {\pd F_X(x) \over \pd x}$.
-Then we have
+$\alpha, \beta > 0$, $K,K' > 0$, $0 < p < 1$. 
+We also write $f_X$ for the density function of $F_X$.
+\par
+We have
 \beam\label{eq:xxie1.0}
@@
 \par
-We observe the following property which also applies to \eqref{eq:xxie1.0}. 
+We observe the following property 
+whose proof is given in Appendix~\ref{sec:thrmI_proof}. 
+% which also applies to the special case 
+%\eqref{eq:xxie1.0}. 
 \ble\label{thrm:I}
- Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
- \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
-   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
-   K} < 0$.
+Assume the two-sided Pareto model \eqref{eq:pareto},
+that there is no \fct al relationship between $\alpha,K$ and $\beta,K'$
+and the utility \fct\ $u$ is increasing and differentiable. Then
+$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and
+$\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
 \ele
-The proof is given in Appendix~\ref{sec:thrmI_proof}. 
-\par
 We conclude that
@@
 illustrates this scenario for $\xi = 1/2, 4$ for the power-utility
-\fct\ \eqref{eq:u}.
+\fct\ \eqref{eq:hjyr}.
 In fact, this positive dependence is indeed observed for some
@@
 \par
-A particularly interesting special case is when the left and the right
-tails have equal parameters, i.e. when $\alpha = \beta$ and $K =
-K'$. In this case \eqref{eq:pareto} becomes
+A particularly interesting case occurs  when 
+$F_X$ is symmetric, i.e., when $\alpha = \beta$, $K =K'$ and $p=0.5$.
+Then \eqref{eq:pareto} turns into 
 \begin{eqnarray}
@@
   } dx
-  - b u(\delta v) F_X(q)
+  - b u(\delta v) F_X(q)\,.\nonumber\\
   \label{eq:ipf}
 \end{eqnarray} 
-First of all we note that the function
-\[
-U_{\text{all}} =
+This situation is not covered by Lemma~\ref{thrm:I}:
+for the proof of the latter result we used Lemma \ref{lemma:I} whose
+assumptions are not satisfied in the present situation.
+Indeed, the integrand 
+\[
+U_{\text{all}}(x) =
 u(C(x)) \left[
-  1 - {b \over 1 + b}\1_{\{x \ge q\}}
+  1 - {b \over 1 + b}\1(x \ge q)
   \right]
@@
 \]
-is not monotone. \footnote{
-To see this we may plug \eqref{eq:hjyr} in $U_{\text{all}}$ and
-re-write it as
-\[
-U_{\text{all}} =
-  -{\phi^{-\xi} \over \xi} \left\{
+is not monotone.\footnote{To see this we may plug \eqref{eq:hjyr} 
+in $U_{\text{all}}$ and re-write it as
+\[U_{\text{all}}(x) =
+  -{\phi^{-\xi} \over \xi} \Big\{
   \underbrace{
-    \left(1 - {b \over 1 + b}\1_{\{x \ge q\}}\right)
-    \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
+    \Big(1 - {b \over 1 + b}\1(x \ge q)\Big)
+    \big[a + e^x\big]^{-\xi}
     +
-    \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
-  }_{U(x)} \right\}
-\]
+    \Big[a + e^{-x}\Big]^{-\xi}
+  }_{U(x)} \Big\}\,,
+\]
+where $a = (1 - \phi) e^r/\phi$.
 Direct computation gives
 \begin{eqnarray*}
-  {\pd U(x) \over \pd x}
+  U'(x)
   &=&
-  -{
-    (a + e^x)^{-\xi - 1} \xi e^x
+  -{(a + e^x)^{-\xi - 1} \xi e^x
     \over
-    1 + b \1_{\{x > q\}}
-  } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}
-  \quad x \neq q
+    1 + b \1(x \ge  q)
+  } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}\,,
+  \qquad x \neq q
 \end{eqnarray*} 
-where $a = (1 - \phi) e^r/\phi$. For $x \ge q$, we have
-\begin{eqnarray*}
-  {1 \over \xi}
-  {\pd U(x) \over \pd x} &=& {x' \over (a + x')^{\xi + 1}}
-  - {x'^\xi \over (a x' + 1)^{\xi + 1} (1 + b)}
-\end{eqnarray*}
-where we have substituted $x'$ for $e^{-x}$. When $x$ is very large
-i.e. when $x'$ is very close to 0, it is useful to expand the last
-expression at $x' = 0$. This yields
-${1 \over \xi} {\pd U(x) \over \pd x} = {x' \over a^{\xi + 1}} + o(x')$
-So we see that when $x$ is sufficiently large
-${\pd U(x) \over \pd x} > 0$. On the other hand, at the discontinuity
-point $x = q$, we have
-$
-\lim_{x \downarrow 0} \left(
-  U(q + x) - U(q - x)
-\right)
-= - {
-  (a + e^q)^{-\xi} b
-  \over
-  1+ b
-} < 0
-$
-So $U(x)$ and hence $U_{\text{all}}$ is not monotone.}
-So, unlike the case when $\beta$ and $K'$
-are independent of $\alpha$ and $K$, we cannot use Lemma \ref{lemma:I}
-to determine the monotonicity of $\tilde u_{\rm max}(\alpha, K)$.
-Instead, we resort to numerical methods to gain some understanding of
+The \fct\ $U(x)$, hence $U_{\text{all}}$, 
+is not monotone because $U'(x)>0$ for all large $x$ while $U(x)$ decreases
+in a small neighborhood of $q$.}
+\par
+We resort to numerical methods to gain some understanding of
 how $\tilde u(F_X, \phi)$ changes with $\alpha$.
-
 The value of $\hat\phi$ can be calculated
@@
 Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\
-\eqref{eq:u}. The corresponding $\tilde u_{\rm max}(\alpha, K)$
-is shown in figure \ref{fig:preference_pareto}.
-One can see that, when $K'$ and $\beta$ are fixed, $\tilde
-u_{\rm max}(\alpha, K)$ and $\hat\phi$ increase 
-with $\alpha$ and decreases with $K$. This is a verification of Lemma
-\ref{lemma:I}; however, when $\beta$ and $K'$ are equal to $\alpha$
-and $K$, $\tilde u_{\rm max}(\alpha, K)$ is monotone
-decreasing with $\alpha$ but is rather insensitive to $K$. On the
+\eqref{eq:hjyr} both for fixed $K', \beta$ and for $K=K'$, $\alpha=\beta$.  
+The corresponding values $\tilde u_{\rm max}(\alpha, K)$
+are shown in Figure \ref{fig:preference_pareto}.
+If $K'$ and $\beta$ are fixed, both $\wt u_{\rm max}(\alpha, K)$ 
+and $\hat\phi$ increase  with $\alpha$ and decrease with $K$. 
+This is in agreement with Lemma~\ref{thrm:I}. 
+In contrast, when $K=K'$ are equal to $\alpha=\beta$
+$\wt u_{\rm max}(\alpha, K)$ decreases with $\alpha$ 
+but is rather insensitive with respect to $K$. On the
 other hand, $\hat\phi$ is not monotone with respect to $\alpha$ or
-$K$. For each fixed $K$, it peaks at an $\alpha$ value somewhere below
-1. For typical realistic $\alpha$ values in $(2, 4)$, $\hat\phi$ is a
+$K$. For each fixed $K$, it peaks at an $\alpha$-value somewhere below
+1. For realistic values $\alpha\in (2, 4)$, $\hat\phi$ is a
 small value below 5\%.
-Since $\tilde u_{\rm max}(\alpha, K)$ is monotone decreasing with
-$\alpha$, investors who seek to maximize it will prefer the smallest
-$\alpha$ in the market, resulting in similar values of $\alpha$ for
-different equities.
+Since $\wt u_{\rm max}(\alpha, K)$ decreases with
+$\alpha$, investors who seek to maximize $\wt u_{\rm max}(\alpha, K)$
+will prefer the smallest $\alpha$ in the market, resulting in similar
+values of $\alpha$ for different equities.
 \begin{figure}[htb!]
@@
    allocation as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
-   \eqref{eq:pareto} with $K'=0.012$, $\beta = 1.4$.
+   \eqref{eq:pareto} for fixed $K'=0.012$, $\beta = 1.4$.
    {\em The 3rd and 4th} graphs show $\hat\phi$ as a \fct\ of $\alpha$
    and $K$ with $\beta = \alpha$ and $K' = K$.
-   We choose the utility \fct\ $u$ from \eqref{eq:u} for $\xi = 1/2$
-   and $\xi = 4$. $b = 0.01$ in all cases.
+   We choose the utility \fct\ $u$ from \eqref{eq:hjyr} for $\xi = 1/2$
+   and $\xi = 4$, $b = 0.01$ in all cases.
   }
@@
   larger $\xi$. Such an agent will therefore invest more heavily in
-  equity, as shown in Figure~\ref{fig:phi_hat_pareto}.
+  equity, as shown in Figure~\ref{fig:phi_hat_pareto}. Note that,
+  while $\hat \phi$ changes nearly in the same in both graphs, the
+  scales in the graphcs are very different.
+
 \end{minipage}
-
-The following graph (figure \ref{fig:preference_pareto}) shows how
-$\tilde u(\alpha, K, \wt\phi)$ changes with $\alpha$ and $K$.
+{\red Needed? The following graph (figure \ref{fig:preference_pareto}) shows how
+$\tilde u(\alpha, K, \wt\phi)$ changes with $\alpha$ and $K$.}
 \begin{figure}[htb!]
@@
    and $K$ with $\beta = \alpha$ and $K' = K$.
-   We choose the utility \fct\ $u$ from \eqref{eq:u} for $\xi = 1/2$
+   We choose the utility \fct\ $u$ from \eqref{eq:hjyr} for $\xi = 1/2$
    and $\xi = 4$. $b = 0.01$ in all cases.
@@
 
-When the parameter $b$ is very small, a GDA preference is closely
-approximated by a mean-utility preference, which GDA reduces to when
-$b = 0$. As shown in the proof of lemma \ref{lemma:III}, the function
-$U_{\text{all}}$ is monotone increasing or decreasing when $\xi$
-and $(1 - \phi) e^r / \phi$ satisfy certain conditions. Moreover, it
+If the parameter $b$ is very small the GDA preference is closely
+approximated by the mean-utility preference, corresponding to $b = 0$. 
+In this case, we show in the proof of Lemma \ref{lemma:III} that the function
+$U_{\text{all}}(x)$ may increase or decrease depending on particular conditions
+on the values of $\xi$ and $(1 - \phi) e^r / \phi$. {\red Conditions?} Moreover, it
 is also easily checked that, when $x \in (0, K(e^{1/\alpha} - 1)]$,
-the density function in the integral of \eqref{eq:ipf} is increasing
-with $\alpha$; when $x \in (K(e^{1/\alpha} - 1), \infty)$, it is
-decreasing with $\alpha$. So, following the same line of arguments
-using lemma \ref{lemma:I} as in the proof of lemma \ref{thrm:I},
-it can be seen $\tilde u_{\rm max}(\alpha, K)$ is
-increasing/decreasing with $\alpha$ when $U_{\text{all}}$ is
-decreasing/increasing.
+the {\red density?} function in the integral of \eqref{eq:ipf} increases
+with $\alpha$; when $x \in (K(e^{1/\alpha} - 1), \infty)$ it 
+decreases with $\alpha$. Following the arguments for Lemma~\ref{thrm:I},
+in particular  using Lemma \ref{lemma:I}, it can be seen that
+$\wt u_{\rm max}(\alpha, K)$ 
+increases/decreases with $\alpha$ when $U_{\text{all}}(x)$ 
+decreases/increases.
 
@@
 \begin{lemma}\label{lemma:III}
-Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:u} and
+Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:hjyr} and
 \eqref{eq:xxie1} respectively. Define

### 2017-06-05 papers/FX/draft.tex
Some revision to the importance sampling paper. Made C++ programs for Johannes' paper.

@@
 Routledge and Zin assumed a power-law utility function 
-%\beam\label{eq:u}
-%u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
-%\eeam
-% In practice it is more often to observe $\gamma > 1$ {\blue Let's
-% consult Casper for references.}
-%We focus on the case $\gamma > 1$ and hence adopt a more convenient
-%notation:
 \beam\label{eq:hjyr}
@@
 
-%In practice it is more often to observe $\gamma > 1$ {\blue Let's
-%consult Casper for references.} Therefore we adopt a more convenient
 For the sake of argument, we assume that an investor initially has one
@@
 \begin{minipage}{0.42\textwidth}
-  In this graph we show two power-utility functions $u$. We see that
+  The economitrical differences between the two utility functions
+  are illustrated in the graph to the left.
+  We see that
   $u(x)$ grows slower and saturates later for small $\xi$ than for
@@
   scales in the graphcs are very different.
-
 \end{minipage}
-{\red Needed? The following graph (figure \ref{fig:preference_pareto}) shows how
-$\tilde u(\alpha, K, \wt\phi)$ changes with $\alpha$ and $K$.}
+
 \begin{figure}[htb!]
@@
 $U_{\text{all}}(x)$ may increase or decrease depending on particular conditions
-on the values of $\xi$ and $(1 - \phi) e^r / \phi$. {\red Conditions?} Moreover, it
-is also easily checked that, when $x \in (0, K(e^{1/\alpha} - 1)]$,
-the {\red density?} function in the integral of \eqref{eq:ipf} increases
+on the values of $\xi$ and $(1 - \phi) e^r / \phi$:
+\begin{enumerate}
+\item If $\max\{a, 1\} < \xi$, $U_{\text{all}}(\cdot)$ is
+  monotone decreasing.
+\item If $a < \xi < 1$ and $(a + y_-)/(a y_- + 1) <
+  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
+  decreasing.
+\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotone
+  increasing.
+\item If $1 < \xi < a$ and $(a + y_+)/(a y_+ + 1) >
+  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
+  increasing.
+\item In other case, $U_{\text{all}}(\cdot)$ is not monotone.
+\end{enumerate}
+where
+\begin{eqnarray*}
+a &=& {
+  (1 - \phi) e^r
+  \over
+  \phi
+} \\
+y_\pm &=& {
+  a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
+  \over
+  a (\xi - 1)
+}
+\end{eqnarray*}
+Moreover, it is also easily checked that, when
+$x \in (0, K(e^{1/\alpha} - 1)]$,
+the density function in the integral of \eqref{eq:ipf} increases
 with $\alpha$; when $x \in (K(e^{1/\alpha} - 1), \infty)$ it 
 decreases with $\alpha$. Following the arguments for Lemma~\ref{thrm:I},
-in particular  using Lemma \ref{lemma:I}, it can be seen that
-$\wt u_{\rm max}(\alpha, K)$ 
-increases/decreases with $\alpha$ when $U_{\text{all}}(x)$ 
-decreases/increases.
-
-\subsection{Student's t-distributed equity return}
-{\red This section is incomplete unless we can show results for $b>0$. Otherwise, we can do everything for $b=0$.} 
-It is a common practice to use Student's t-distribution to model the
-stationary distribution of equity returns. So it is of interest to
-find out what implications this distribution has when it is combined
-with the PDA preference. Formally we assume {\red instead of this function one can take a symmetric \ds\ with tails
-$\P(X>x)=0.5 (\kappa/(x+\kappa))^{\alpha}$. If the symmetry causes another behavior than in the previous section this would be interesting.}
-{\blue our tail-index is always $\alpha$, not $\nu$.}
-\[
-f(x; \alpha) = c(\alpha) \left(
-  1 + {x^2 \over \alpha}
-\right)^{-(\alpha + 1)/2}
-\]
-where $\alpha > 1$ and
-\[
-c(\alpha) = {
-  \Gamma({\alpha + 1 \over 2})
-  \over
-  \Gamma(\alpha/2) \sqrt{\alpha \pi}
-}
-\]
-In the same way as for \eqref{eq:ipf}, we can write $\wt u(F, \phi)$
-as
-\begin{eqnarray}
-  && \wt u(F, \phi) \nonumber \\
-  &=&
-  (1 + b)
-  \int_{0}^{\infty}
-  \underbrace{
-    \left\{
-      u(C(x)) \left[
-        1 - {b \over 1 + b}\1_{\{x \ge q\}}
-        \right]
-      + u(C(-x))
-    \right\}
-  }_{U_{\text{all}}}
-  f(x, \alpha) dx
-  - b u(\delta v) F_X(q)
-  \label{eq:t1}
-\end{eqnarray}
-where $C(\cdot)$ is defined in \eqref{eq:xxie1}. As shown in lemma
-\ref{lemma:III}, when $b = 0$ and $u(\cdot)$
-takes the power-form of \eqref{eq:hjyr}, $U_{\text{all}}$ is monotone
-depending on the values of $\xi$ and $(1 - \phi) e^r / \phi$. As given
-in lemma \ref{lemma:II}, there is a point $x_0 > 0$ such that
-${\pd f \over \pd \alpha}(x_0, \alpha) = 0$ and
-$\forall x \in (0, x_0),
-{\pd f \over  \pd \alpha}(x, \alpha) > 0$ and
-$\forall x \in (x_0, \infty)$,
-${\pd f \over \pd \alpha}(x, \alpha) < 0$. Thus it remains to verify
-$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$ and
-$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
-if we are to apply lemma \ref{lemma:I}.
-
-As computed in the proof of lemma \ref{lemma:II}, ${\pd f \over \pd
-  \alpha}(x, \alpha)$ is given by \eqref{eq:xxie4.1}. It is also shown
-there ${d \over d \alpha} c(\alpha) > 0$. Thus for
-$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
-it suffices to show
-\[
-\int_ 0^\infty {
-x^2
-\over
-(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
-} dx < \infty
-\]
-We may write
-\begin{eqnarray*}
-&& \int_ 0^\infty {
-x^2
-\over
-(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
-} dx \\
-&=& \left(\int_0^1 + \int_1^\infty \right) {
-x^2
-\over
-(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
-} dx \\
-&=& I_1 + I_2
-\end{eqnarray*}
-Clearly
-\[
-I_1 < \int_0^1 {1 \over \alpha} dx < \infty
-\]
-while
-\begin{eqnarray*}
-  I_2 &=& \int_1^\infty {1 \over 1 + \alpha/x^2}
-  {1 \over (1/x^2 + 1/\alpha)^{(\alpha+1) / 2}} {dx \over x^{\alpha +
-      1}} \\
-  &<& \int_1^\infty \alpha^{(\alpha+1) / 2} {dx \over x^{\alpha + 1}}
-  < \infty
-\end{eqnarray*}
-So we conclude
-$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
-To see
-$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$, we note
-\begin{eqnarray*}
-  \xi |U_{\text{all}}(x)|
-  &<& [(1 - \phi) e^r + e^x]^{-\xi} + [(1 - \phi) e^r + e^{-x}]^{-\xi} \\
-  &<& 1 + (1 - \phi)^{-\xi} e^{-r \xi}
-\end{eqnarray*}
-Since $\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$,
-it follows from the above inequality $\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
-Thus by lemma \ref{lemma:I}, $\wt u_{\rm max}$ is monotone
-increasing/decreasing with $\alpha$ when $U_{\text{all}}(\cdot)$ is
-monotone decreasing/increasing. Accordingly, an investor guided by the
-utility function will seek the largest/smallest $\alpha$ observed in
-the market.
-
-If however $b > 0$, Lemma \ref{lemma:I} is not applicable
-anymore. Nonetheless, numerical analysis lends some insight.
-As shown in figure \ref{fig:htfg}, $\hat\phi$ is monotone increasing
-for all 4 values of $b$, while $\tilde u_{\rm max}(\alpha)$ is
-increasing with $\alpha$ when $b$ is relatively large, but
-decreasing with $\alpha$ when $b$ is small. We note that a sizable
-value of $b$ indicates a conservative, risk-averse investor.
-
-\begin{figure}[htb!]
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
-  \end{minipage}
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_power4.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_power4.pdf}
-  \end{minipage}
-  \caption{$\hat\phi$ (left) and ${\pd \tilde u_{\rm max} \over \pd
-      \alpha}$ (right). {\em top:} $\xi = 1/2$. {\em bottom:} $\xi =
-    4$.
-  }
-  \label{fig:htfg}
-\end{figure}
-
-% \subsubsection{log-utility function}
-% We first consider the case where the utility function is simply
-% $u(\cdot) = \log(\cdot)$. This corresponds to the limiting case of
-% \eqref{eq:u} as $\gamma \to 0$.
-
-% First notice that when $b = 0$, $\wt u(F, \phi)$ reduces to
-% $\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
-% Its integrand becomes $\log[C(x, \phi) C(-x, \phi)]$.
-% Now observe $\log(C(x, \phi)C(-x, \phi))$ is an increasing function of
-% $x$:
-% \begin{eqnarray*}
-%   {\pd C(x, \phi) C(-x, \phi) \over \pd x}
-%   &=&
-%   \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
-% \end{eqnarray*}
-% Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
-% \ref{lemma:II}, {\red need to look at $\pd \wt u_{\rm max} \over \pd \alpha$?} 
-% ${\pd \wt u(F, \phi)) \over \pd \alpha} < 0$ for all $\phi$. Hence all
-% investors guided by this utility function would be seeking the
-% smallest $\alpha$, i.e. the heaviest tail in the market. Here note that
-% Student's t is a symmetric distribution; a heavier tail implies not
-% only greater down-side risk but also greater up-side potential.
-
-% In fact, even when $b > 0$, the same monotonicity property holds. This
-% is revealed by numerically maximizing $\wt u(F, \phi)$ with
-% respect to $\phi$ for a sequence of selected values of $\alpha$. The
-% results are shown in figure~\ref{fig:phi_hat_U}.
-% \begin{figure}[htb!]
-%   \begin{minipage}{0.5\linewidth}
-%     \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
-%   \end{minipage}\hfill
-%   \begin{minipage}{0.5\linewidth}
-%     \includegraphics[width=\textwidth]{U_b_t_log.pdf}
-%   \end{minipage}
-%   \caption{$\hat\phi$ and $\tilde u_{\rm max}(\alpha)$
-%     \label{fig:phi_hat_U}
-%   }
-% \end{figure}
+and applying Lemma \ref{lemma:I}, it can be seen that
+$\wt u_{\rm max}(\alpha, K)$ increases/decreases with $\alpha$ when
+$U_{\text{all}}(x)$  decreases/increases.
 
@@
 \label{sec:4}
-TODO: Summarize and conclude that all correlated series share the same
-tail-index; their tail functions differ from each other via their
-scale parameters.
-
+{\red Casper, we believe you are the most knowledgeable to conclude.}
 \appendix
@@
 \begin{proof}
-  % Firstly we show, for any $S \subseteq (a, b)$,
-  % \[
-  %   {\pd \over \pd \theta} \int_S f(x, \theta) dx
-  %   =
-  %   \int_S {\pd \over \pd \theta} f(x, \theta) dx    
-  % \]
-  % Let
-  % \[
-  % f_{\Delta}(x, \theta) = {
-  %   f(x, \theta +  \Delta) - f(x, \theta)
-  %   \over
-  %   \Delta
-  % }
-  % \]
-  % Since $f(x, \theta)$ is differentiable for all $\theta \in \Theta$,
-  % \[
-  % f_{\Delta}(x, \theta) \to {\pd f \over \pd \theta}(x, \theta)
-  % \quad \Delta \to 0
-  % \]
-  % For any give $\epsilon > 0$ and sufficiently small
-  % $\Delta$, we have
-  % \[
-  % |f_{\Delta}(x, \theta)| <
-  % \left|
-  %     {\pd f \over \pd \theta}(x, \theta)
-  % \right| (1 + \epsilon)
-  % \]
-  % Using \eqref{eq:frt} and dominated convergence theorem we obtain
-  % \[
-  % \lim_{\Delta \to 0} \int_S f_\Delta(x, \theta) dx
-  % =
-  % \int_S {\pd f \over \pd \theta}(x, \theta) dx
-  % < \infty
-  % \]
-  % That is
-  % \[
-  %   {\pd \over \pd \theta} \int_S f(x, \theta) dx
-  %   =
-  %   \lim_{\Delta \to 0} {
-  %     \int_S f(x, \theta + \Delta) dx
-  %     -
-  %     \int_S f(x, \theta) dx
-  %     \over
-  %     \Delta
-  %   }
-  %   =
-  %   \int_S {\pd f \over \pd \theta}(x, \theta) dx     
-  % \]
-  % From \eqref{eq:frt} it easily follows that the set on which
-  % ${\pd f(x, \theta) \over \pd \theta} = \infty$ has Lesbegue
-  % measure 0. Call this set $S_0$.
   Firstly, by dominated convergence theorem, conditions \eqref{eq:frt}
@@
 
-\section{A lemma on Student's t-distribution}
+\section{When equity returns follow Student's t-distribution}
 \setcounter{equation}{0}
+It is a common practice to use Student's t-distribution to model the
+stationary distribution of equity returns. So it is of interest to
+find out what implications this distribution has when it is combined
+with the PDA preference. Formally we assume
+\[
+f(x; \alpha) = c(\alpha) \left(
+  1 + {x^2 \over \alpha}
+\right)^{-(\alpha + 1)/2}
+\]
+where $\alpha > 1$ and
+\[
+c(\alpha) = {
+  \Gamma({\alpha + 1 \over 2})
+  \over
+  \Gamma(\alpha/2) \sqrt{\alpha \pi}
+}
+\]
+In the same way as for \eqref{eq:ipf}, we can write $\wt u(F, \phi)$
+as
+\begin{small}
+  \begin{eqnarray}
+    \wt u(F, \phi)
+    &=&
+    (1 + b)
+    \int_{0}^{\infty}
+    \underbrace{
+      \left\{
+      u(C(x)) \left[
+        1 - {b \over 1 + b}\1_{\{x \ge q\}}
+        \right]
+      + u(C(-x))
+      \right\}
+    }_{U_{\text{all}}}
+    f(x, \alpha) dx - b u(\delta v) F_X(q)
+    \label{eq:t1}
+  \end{eqnarray}
+\end{small}
+where $C(\cdot)$ is defined in \eqref{eq:xxie1}. As shown in lemma
+\ref{lemma:III}, when $b = 0$ and $u(\cdot)$
+takes the power-form of \eqref{eq:hjyr}, $U_{\text{all}}$ is monotone
+depending on the values of $\xi$ and $(1 - \phi) e^r / \phi$. As given
+in lemma \ref{lemma:II}, there is a point $x_0 > 0$ such that
+${\pd f \over \pd \alpha}(x_0, \alpha) = 0$ and
+$\forall x \in (0, x_0),
+{\pd f \over  \pd \alpha}(x, \alpha) > 0$ and
+$\forall x \in (x_0, \infty)$,
+${\pd f \over \pd \alpha}(x, \alpha) < 0$. Thus it remains to verify
+$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$ and
+$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
+if we are to apply lemma \ref{lemma:I}.
+
+As computed in the proof of lemma \ref{lemma:II}, ${\pd f \over \pd
+  \alpha}(x, \alpha)$ is given by \eqref{eq:xxie4.1}. It is also shown
+there ${d \over d \alpha} c(\alpha) > 0$. Thus for
+$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
+it suffices to show
+\[
+\int_ 0^\infty {
+x^2
+\over
+(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
+} dx < \infty
+\]
+We may write
+\begin{eqnarray*}
+&& \int_ 0^\infty {
+x^2
+\over
+(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
+} dx \\
+&=& \left(\int_0^1 + \int_1^\infty \right) {
+x^2
+\over
+(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
+} dx \\
+&=& I_1 + I_2
+\end{eqnarray*}
+Clearly
+\[
+I_1 < \int_0^1 {1 \over \alpha} dx < \infty
+\]
+while
+\begin{eqnarray*}
+  I_2 &=& \int_1^\infty {1 \over 1 + \alpha/x^2}
+  {1 \over (1/x^2 + 1/\alpha)^{(\alpha+1) / 2}} {dx \over x^{\alpha +
+      1}} \\
+  &<& \int_1^\infty \alpha^{(\alpha+1) / 2} {dx \over x^{\alpha + 1}}
+  < \infty
+\end{eqnarray*}
+So we conclude
+$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
+To see
+$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$, we note
+\begin{eqnarray*}
+  \xi |U_{\text{all}}(x)|
+  &<& [(1 - \phi) e^r + e^x]^{-\xi} + [(1 - \phi) e^r + e^{-x}]^{-\xi} \\
+  &<& 1 + (1 - \phi)^{-\xi} e^{-r \xi}
+\end{eqnarray*}
+Since $\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$,
+it follows from the above inequality $\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
+Thus by lemma \ref{lemma:I}, $\wt u_{\rm max}$ is monotone
+increasing/decreasing with $\alpha$ when $U_{\text{all}}(\cdot)$ is
+monotone decreasing/increasing. Accordingly, an investor guided by the
+utility function will seek the largest/smallest $\alpha$ observed in
+the market.
+
+If however $b > 0$, Lemma \ref{lemma:I} is not applicable
+anymore. Nonetheless, numerical analysis lends some insight.
+As shown in figure \ref{fig:htfg}, $\hat\phi$ is monotone increasing
+for all 4 values of $b$, while $\tilde u_{\rm max}(\alpha)$ is
+increasing with $\alpha$ when $b$ is relatively large, but
+decreasing with $\alpha$ when $b$ is small. We note that a sizable
+value of $b$ indicates a conservative, risk-averse investor.
+
+\begin{figure}[htb!]
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
+  \end{minipage}
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{phi_hat_b_t_power4.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.5\linewidth}
+    \includegraphics[width=\textwidth]{U_b_t_power4.pdf}
+  \end{minipage}
+  \caption{$\hat\phi$ (left) and ${\pd \tilde u_{\rm max} \over \pd
+      \alpha}$ (right). {\em top:} $\xi = 1/2$. {\em bottom:} $\xi =
+    4$.
+  }
+  \label{fig:htfg}
+\end{figure}
 \begin{lemma}

### 2017-07-10 cv/20sec/XieXiaolei.tex
Added GPA to the CV.
Applied to BNY Mellon.
Modified the CV and the paper with jeff.

@@
 % Skill bar section, each skill must have a value between 0 an 6 (float)
-\skills {{Java/3},{Python/3},{SQL/4},{Matlab/5.2},{GNU Scientific Library/5},{R/5.8},{Parallel Computing OpenMP/4.5},{C/5},{C++/5}}
+\skills {{Java/3},{Python/3},{SQL/4.5},{Matlab/4},{GNU Scientific Library/4},{R/4},{Parallel Computing OpenMP/3.8},{C/4.3},{C++/4.3}}
 
@@
 \begin{minipage}{0.85\linewidth}
+  {\small
   \begin{itemize}
@@
   \end{itemize}
+  }
 \end{minipage}
 
-\vspace{3mm}
+\vspace{2mm}
 
@@
     \end{itemize}
+  \item Average grade: 87/100
   \end{itemize}
+  }
 \end{minipage}
-\vspace{3mm}
+\vspace{2mm}
 
@@
 \begin{minipage}{0.85\linewidth}
+  {\small
     \begin{itemize}
@@
     \item {\it Major Subject:} Mobile Computing - Services and Security
-  \item {\it Average grade:} 4.17/5.
+    \item Average grade: 4.17/5
     \end{itemize}
+  }
 \end{minipage}
-\vspace{3mm}
+\vspace{2mm}
 
@@
   \item Beijing Institute of Technology
+  \item Average grade: 89/100
   \end{itemize}
+  }
 \end{minipage}
@@
 \begin{itemize}
-\item {\bf Xie, X.}, Davis, R.A., Heiny, J. and Mikosch, T.
+\item Davis, R.A., Heiny, J., Mikosch, T., {\bf Xie, X.},
   2016. {\it Extreme value analysis for the sample autocovariance
@@
 
-\item {\bf Xie, X}, Jan\ss en, A., Mikosch, T. and Rezapour, M.
+\item Jan\ss en, A., Mikosch, T., Rezapour, M., {\bf Xie, X},
   2016. {\it The eigenvalues of the sample covariance matrix of a
@@
 \begin{itemize}
-\item {\bf Xie, X.}, Mikosch, T. and de Vries, C., 2017,
+\item Mikosch, T., de Vries, C., {\bf Xie, X.}, 2017,
   {\it Tail-indices and scale parameters of returns series}
@@
 \begin{itemize}
-\item {\bf Xie, X.} Collamore, J, 2017
-  {\it An efficient estimator of the probability of very large losses posed by
-    GARCH(p,q) processes}
+\item Collamore, J., {\bf Xie, X.}, 2017
+  {\it An efficient estimator of the probability of very large losses
+    posed by GARCH(p,q) processes}
 
@@
     probabilities is difficult because the exact distribution is
-    unknown. Moreover, naive simulation methods lead to erractic
+    unknown. Moreover, naive simulation methods lead to erratic
     estimates. Our importance sampling estimator presents a solution.

### 2017-07-14 cpp/ImportanceSampling/GarchPQ.cpp
Investigated the DJIA data set.

@@ -186,15 +186,21 @@ int main(int argc, char*argv[])
     // // madeup
     // vector<double> alpha({1.0e-7, 0.11, 0});
     // vector<double> beta({0.88});
+    // vector<double> alpha({1.0e-7, 0.2, 0});
+    // vector<double> beta({0.2});
     // DAX
-    vector<double> alpha({3.374294e-06, 2.074732e-02, 4.104949e-02});
-    vector<double> beta({9.102376e-01});
+    // vector<double> alpha({3.374294e-06, 2.074732e-02, 4.104949e-02});
+    // vector<double> beta({9.102376e-01});
     // FTSE100
     // vector<double> alpha({1.0e-7, 0.10623464, 0.02904907});
     // vector<double> beta({0.7829784});
     // // FTSE100 GARCH(1,1)
     // vector<double> alpha({1.0e-7, 1.268747e-01});
     // vector<double> beta({8.008847e-01});
+
+    // DJIA
+    vector<double> alpha({3.374294e-06, 0.061577928, 0.12795424});
+    vector<double> beta({0.6610499});
 
     // SP500
     // vector<double> alpha({9.376992e-06, 7.949678e-02, 8.765884e-02});

### 2017-07-14 papers/FX/draft.tex
Made slides for the Delft talk.

@@
 This is in agreement with Lemma~\ref{thrm:I}. 
-In contrast, when $K=K'$ are equal to $\alpha=\beta$
+In contrast, when $K=K'$ and $\alpha=\beta$
 $\wt u_{\rm max}(\alpha, K)$ decreases with $\alpha$ 
@@
 \label{sec:4}
-{\red Casper, we believe you are the most knowledgeable to conclude.}
+We have established that, in the case of an equity return series with
+two-sided, functionally independent Pareto tails, investor
+preference functionals are monotone increasing/decreasing with the
+tail index/scale parameters. Thus in a market dominated by such
+equities, the investors would pursue the largest tail index in the
+market, leading to a shared common tail index for all equities.
+
 \appendix

### 2017-07-18 papers/FX/draft.tex
Some revision to papers.

@@
 
+The empirical results presented in section \ref{sec:1} suggest this
+may well be the case for the ``Consumer Staples'' sector of S\&P 500,
+given the Hill estimates of tail indices shown in figure \ref{fig:1}
+and the largely positive results of tests for equal tail indices shown
+in figure \ref{fig:PairTest}.
+
+On the other hand, we have also seen that, when the left and the right
+tails have the same indices, investor preference over the equity has
+more sophisticated variations in the parameters' space including the
+tail parameters of the equity, the interest rate, the investor's risk
+apetite as captured by his utility function, and his threshold of
+disappointment.
+
+We also acknowledge that our model of the market and the investor is a
+simple one, not accoounting for the dependence between equities, nor
+the categorization of investors and their interactions. These are
+potential topics of future work.
+
 \appendix

### 2017-07-18 cpp/ImportanceSampling/GarchPQ_FX.r

@@ -69,21 +69,23 @@ R <- getAssetReturns("2012-01-01", "2015-01-01", assets, 1, "closing", "localhos
 coef <- matrix(NA, nrow=length(assets), ncol=3);
 for (i in 1:length(assets)) {
     X <- getAssetReturns("2012-01-01", "2015-01-01", assets[i], 1, "closing", "localhost");
-    M <- garchFit(~garch(2, 1), data=X, trace=FALSE,
+    garch21 <- garchFit(~garch(2, 1), data=X, trace=FALSE,
                   include.mean=FALSE);
-    coef[i, ] <- M@fit$coef[-1];
+    garch11 <- garchFit(~garch(1, 1), data=X, trace=FALSE,
+                  include.mean=FALSE);
+    coef[i, ] <- garch21@fit$coef[-1];
 }
 
 X <- getAssetReturns("2012-01-01", "2015-01-01", "SP500", 1, "closing", "localhost");
 M <- garchFit(~garch(2, 1), data=X, trace=FALSE);
 
 
-##            [,1]       [,2]      [,3]
-## [1,] 0.02749864 0.04228535 0.8968533
-## [2,] 0.10623464 0.02904907 0.7829784
-## [3,] 0.08835834 0.09685783 0.6543018
-## [4,] 0.07490443 0.12812118 0.6543123
-## [5,] 0.01677130 0.03622593 0.9232396
+##           alpha1       alpha2     beta1   sum
+## DAX     0.02074733 0.04104947 0.9102376   0.97
+## FTSE100 0.08806776 0.04548650 0.7808275   0.91
+## SP500   0.07949678 0.08765884 0.6683833   0.84
+## DJIA    0.06157793 0.12795424 0.6610499   0.85
+## OMXS30  0.00924104 0.04309249 0.9246996   0.98
 
 
 

### 2017-07-18 phd/Chaptercorrelation.tex
Added My PhD thesis.

%correlation Chapter
\chapter[Almost sure convergence of the largest and smallest eigenvalues of high-dimensional sample correlation matrices under infinite fourth moment]{{\huge Almost sure convergence of the largest and smallest eigenvalues of high-dimensional sample correlation matrices under infinite fourth moment}}\label{ch:corr}
\chaptermark{Extreme eigenvalues of sample correlation matrices}

\begin{center}
\textsc{Johannes Heiny \& Thomas Mikosch (2016)}
\end{center}

\begin{abstract}{In this paper, we show that the largest and smallest eigenvalues 
of a sample correlation matrix stemming from $n$ independent observations of a $p$-dimensional 
time series with iid components converge almost surely to $(1+\sqrt{\gamma})^2$ and $(1-\sqrt{\gamma})^2$, 
respectively, as $\nto$, if $p/n\to \gamma \in (0,1]$ and the truncated variance of the entry 
distribution is ``almost slowly varying'', a condition we describe via moment 
properties of self-normalized sums. Moreover, the empirical spectral distributions 
of these sample correlation matrices converge weakly, with probability $1$, to the \MP law, 
which extends a result in \cite{bai:zhou:2008}. We compare the behavior of the eigenvalues of the 
sample covariance and sample correlation matrices and argue that the latter seems more robust, 
in particular in the case of infinite fourth moment. We briefly address some practical issues 
for the estimation of extreme eigenvalues in a simulation study.   
\par
In our proofs we use the method of moments combined with a Path-Shortening Algorithm, 
which efficiently uses the structure of sample correlation matrices, 
to calculate precise bounds for matrix norms. We believe that this new approach could be of further use in random matrix theory. 
\medskip

\noindent {\bf Keywords:} Sample correlation matrix, infinite fourth moment, largest eigenvalue, smallest eigenvalue, 
spectral distribution, sample covariance matrix, self-normalization, regular variation, combinatorics.}
\end{abstract}

\newpage
\section{Introduction and notation}\label{sec:sdlgj}

%{\red Discuss $p/n \to 0$.} 
%{\blue Inconsistent use of referenes, sometimes with or without names. I cannot see a general rule.}
In modern statistical analyses one is often faced with large data sets where both the dimension of the observations and the sample size are large. The dramatic increase and improvement of computing power and data collection devices have triggered the necessity to study and interpret the sometimes overwhelming amounts of data in an efficient and tractable way. Huge data sets arise naturally in wireless communication, finance, natural sciences and genetic engineering.
For such data one commonly studies the dependence structure via covariances and correlations which can be estimated by their sample analogs. {\em Principal component analysis}, for example, uses an orthogonal transformation of the data such that only a few of the resulting vectors explain most of the variation in the data. The empirical variances of these so-called {\em principal component vectors} are the largest eigenvalues of the {\em sample covariance or correlation matrix}.

Throughout this paper we consider the $p\times n$ data matrix
\begin{equation*}
\X=\X_n=(X_{it})_{i=1,\ldots,p; t=1,\ldots,n}
\end{equation*}
of identically distributed entries $(X_{it})$ with generic element $X$, where we assume $\E[X]=0$ and $\E[X^2]=1$ if the first and second moments of $X$ are finite, respectively. A column of $\X$ represents an observation of a $p$-dimensional time series.

Random matrix theory provides a great variety of results on the ordered eigenvalues
\begin{equation}\label{eq:deflambda}
\la_{(1)} \ge \cdots \ge\la_{(p)}\,,
\end{equation}
of the (non-normalized) {\em sample covariance matrix} $\X\X'$. Here we will only discuss the case $p=p_n \to \infty$ and, unless stated otherwise, we assume the growth condition  
\begin{equation}\label{Cpn}
\lim_{\nto} \frac{p_n}{n}\to \gamma\in (0,1]\,. \tag{$G_{\gamma}$}
\end{equation} 
For the finite $p$ case, we refer to \cite{anderson:1963,muirhead,janssen:mikosch:rezapour:xie:2016}.
When studying the asymptotic properties of estimators under \eqref{Cpn}
one often obtains results that dramatically differ from the standard $p$ fixed, $\nto$ case, in which the spectrum of $(n^{-1}\X\X')$ converges to its population covariance spectrum. In 1967, Mar\v cenko and Pastur \cite{marchenko:pastur:1967} observed that even in the case of iid entries $(X_{it})$ with $\E[X^2]=1$ the eigenvalues $(\lambda_{(i)}/n)$ do not concentrate around $1$. For more examples, see \cite[Chapter 1]{bai:silverstein:2010} and \cite{elkaroui:2009}. Typical applications where \eqref{Cpn} seems reasonable are discussed in \cite{johnstone:2001,donoho:2000}. 
\par

In comparison with $(\lambda_{(i)})$, much less is known about the ordered eigenvalues
\begin{equation*}
\mu_{(1)} \ge \cdots \ge\mu_{(p)}\,
\end{equation*} 
of the {\em sample correlation matrix} $\bfR=\Y\Y'$ with entries
\begin{equation}\label{eq:corrR}
R_{ij}=\sum_{t=1}^n \frac{X_{it}X_{jt}}{\sqrt{D_i} \sqrt{D_j}} = \sum_{t=1}^n Y_{it}Y_{jt}\,, \quad i,j=1,\ldots,p\,.
\end{equation}
In this paper we will often make use of the notation $\Y=(Y_{it})=(X_{it}/\sqrt{D_i})$ and
\begin{equation}\label{eq:D}
D_i=D_i^{(n)}=\sum_{t=1}^n X_{it}^2\,, \qquad
i=1,\ldots,p;\;  n\ge 1\,.
\end{equation} 
Note that the dependence of $(\lambda_{(i)})$ and $(\mu_{(i)})$ on $n$ is suppressed in the notation.

%---------------------------------------------------------------------------------------
\subsection{The case $(X_{it})$ iid, $\E[X^4]<\infty$ and $\E[X^2]=1$}

In this case the behavior of the eigenvalues of the sample covariance matrix $\X\X'$ and the sample correlation matrix $\bfR$ are closely intertwined. 
\par

For any random $p\times p$ matrix $\bfA$ with real eigenvalues $\lambda_1(\bfA),\ldots,\lambda_p(\bfA)$ the {\em empirical spectral distribution} is defined by 
\begin{equation*}
F_{\bfA}(x)= \frac{1}{p}\; \sum_{i=1}^p \1_{\{ \lambda_i(\bfA)\le x \}}, \qquad x\in  \R\,.
\end{equation*}
Many functionals of the eigenvalues $\lambda_1(\bfA),\ldots,\lambda_p(\bfA)$ can be expressed in terms of $F_{\bfA}$ \cite{bai:fang:liang:2014}, for instance
\begin{equation*}
\det \bfA = \prod_{i=1}^p \lambda_i(\bfA) = \exp \Big( p \int_{0}^{\infty} \log x \;\dint F_{\bfA}(x) \Big)\,.
\end{equation*}
A major problem in random matrix theory is to find the weak limit of $(F_{\bfA_n})$ for suitable sequences $(\bfA_n)$; see for example \cite{bai:silverstein:2010,yao:zheng:bai:2015} for more details. By weak convergence of a sequence of probability distributions $(F_{\bfA_n})$ 
to a \pro y \ds\ $F$, we mean $\lim_{\nto} F_{\bfA_n}(x)=F(x) \,\as$ for all continuity points of $F$.
In this context a useful tool is the {\em Stieltjes transform}
of the empirical spectral distribution $F_{\bfA}$:
\begin{equation*}
s_{\bfA}(z)= \int_{\R} \frac{1}{x-z} \dint F_{\bfA}(x) = \frac{1}{p} \tr(\bfA -z \bfI)^{-1}\,, \quad z\in \C^+\,,
\end{equation*}
where $\C^+$ denotes the complex numbers with positive imaginary part. 
Weak convergence of $(F_{\bfA_n})$ to $F$ is equivalent to $s_{F_{\bfA_n}}(z) \to s_F(z)$ a.s. for all $z\in \C^+$.

Under the growth condition \eqref{Cpn}, the sequence of empirical spectral distributions of the normalized sample covariance matrix $n^{-1} \X\X'$ converges weakly to the \MP law with density
\begin{eqnarray}\label{eq:MPch4}
f_\gamma(x) = 
\left\{\begin{array}{cc}
\frac{1}{2\pi x\gamma} \sqrt{(b-x)(x-a)} \,, & \mbox{if } a\le x \le b, \\
 0 \,, & \mbox{otherwise,}
\end{array}\right.
\end{eqnarray}\noindent
where $\gamma\in (0,1]$, $a=(1 -\sqrt{\gamma})^2$ and $b=(1 +\sqrt{\gamma})^2$. This classical result is sometimes referred to as \MP theorem \cite{marchenko:pastur:1967}. Informally, the histogram of $(\lambda_{(i)}/n)$ is asymptotically non-random and the limiting shape depends only on the fraction $p/n$. For an illustration, see Figure~\ref{fig:alpha6}.
\par

The \MP law has $k$-th moment
\begin{equation}\label{eq:momentsmp}
\beta_k=\beta_k(\gamma)=\int_a^b x^k f_\gamma(x) \dint x=\sum_{r=1}^{k} \frac{1}{r} \binom{k}{r-1}\binom{k-1}{r-1}\gamma^{r-1}\,,\quad k\ge 1\,,
\end{equation}
and Stieltjes transform 
\beam\label{eq:stieltjestransform}
s(z)
=\int_{\R} \frac{1}{x-z}f_\gamma(x) \dint x 
= \frac{1-\gamma -z +\sqrt{(1+\gamma-z)^2-4\gamma}}{2 \gamma z} \,;
%+ \frac{1-\gamma}{\gamma z} \1_{\{\gamma>1\}}\,.
\eeam
see \cite[Chapter~3]{bai:silverstein:2010} or \cite{bai:fang:liang:2014,yao:zheng:bai:2015}.
\par

The \as~behavior of the extreme eigenvalues is more involved and therefore it has received significant attention in the literature. From the \MP theorem one can infer
\begin{equation}\label{eq:efsdf}
\limsup_{\nto} n^{-1}\lambda_{(p)} \le (1-\sqrt{\gamma})^2 \le (1+\sqrt{\gamma})^2 \le \liminf_{\nto} n^{-1}\lambda_{(1)}\, \quad \as
\end{equation}
The finiteness of the fourth moment of $X$ is necessary for the almost sure convergence of  $\lambda_{(1)}/n$; see \cite{baisilv}.
If $\E[X^4]<\infty$, one has (see \cite{bai:silverstein:2010}) 
\begin{equation}\label{eq:drtgdrghdr}
n^{-1} \la_{(1)} \to (1+\sqrt{\gamma})^2\, \quad \text{ and } \quad n^{-1} \la_{(p)} \to (1-\sqrt{\gamma})^2\, \quad \as
\end{equation}
\par

The minimal moment requirement for the convergence of the normalized smallest eigenvalue, however, was an open question for a long time.
Recently, it was proved in \cite{tikhomirov:2015} that $n^{-1} \la_{(p)} \to (1-\sqrt{\gamma})^2\, \as$ only requires a finite second moment.
Under suitable moment assumptions $\la_{(1)}$ and $\la_{(p)}$ possess {\em Tracy--Widom} fluctuations around their almost sure limits. For instance, 
the paper \cite{johnstone:2001} complemented \eqref{eq:drtgdrghdr} by the corresponding \clt\ in the special case of iid standard normal 
entries:
\beam\label{eq:tcch4}
 n^{2/3}\,\dfrac{(\sqrt{\gamma})^{1/3}}{\big(1+\sqrt{\gamma}\big)^{4/3}}\Big(\dfrac {\la_{(1)}}{n} -
\big(1+\sqrt{\tfrac pn }\big)^2\Big)
\std \xi\,,
\eeam
where the limiting \rv\ has a {\em Tracy--Widom \ds} of order 1. Notice that the centering
$\big(1+\sqrt{\tfrac pn }\big)^2$ can in general not be replaced by $(1+\sqrt{\gamma})^2$.
This \ds\ is ubiquitous in random matrix theory.
%It is defined via some ordinary differential equation; we refer to \cite{tracy:widom:2012} for a definition and properties.
Its distribution function $F_1$ is given by
\begin{equation*}
F_1(s) = \exp\Big\{
  -\frac{1}{2} \int_{s}^\infty [
    q(x) + (x - s) q^2(x)
 ] \dint x
\Big\}\,,
\end{equation*}
where $q(x)$ is the unique solution to the Painlev\'e II differential
equation
\begin{equation*}
  q''(x) = xq(x) + 2 q^3(x)\,,
\end{equation*}
where $ q(x)\sim {\rm Ai}(x)$ as $x \to \infty$ and Ai$(\cdot)$ is the Airy kernel; see Tracy and Widom~\cite{tracy:widom:2012} for details.
%We notice that the rate $n^{2/3}$ compares favorably to the $\sqrt{n}$-rate in the classical \clt\ for sums of iid finite variance \rv s.
\par

Sometimes practitioners would like to know ``to which extent the random matrix results would hold if one were concerned with sample correlation matrices and not sample covariance matrices \cite{elkaroui:2009}''.
A partial answer is that the aforementioned results also hold for the sample correlation matrix $\bfR$ and its eigenvalues $\mu_{(1)}\ge\cdots \ge \mu_{(p)}$. 
%The results on sample covariance matrices can be used to draw conclusions about the behavior of the eigenvalues of the sample correlation matrix.
With $\bfF= \diag (1/D_1,\ldots,1/D_p)$, we have $\bfR={\bfF}^{1/2} {\X}{\X}'{\bfF}^{1/2}$ which has the same eigenvalues as ${\X}{\X}'{\bfF}$. Weyl's inequality (see \cite{bhatia:1997}) yields
\begin{equation}\label{eq:lamu}
\begin{split}
\max_{i=1,\ldots,p} |\mu_{(i)}- n^{-1} \la_{(i)}| &\le \twonorm{{\X}{\X}'{\bfF}-n^{-1}{\X}{\X}' }\\
&\le n^{-1} \twonorm{{\X}{\X}'}  \twonorm{n \bfF- \bfI}\\
&= n^{-1} \la_{(1)} \max_{i=1,\ldots,p} \Big| \frac{n}{D_i} -1 \Big|\,,
\end{split}
\end{equation}
where for any matrix $\bfA$, $\twonorm{\bfA}$ denotes its spectral norm, i.e., its largest singular value.

Lemma 2 in \cite{bai:yin:1993} implies that $\E[X^4]<\infty$ is equivalent to
\begin{equation*}
\max_{i=1,\ldots,p} \Big| \frac{n}{D_i} -1 \Big| \cas 0\,,
\end{equation*}
while $n^{-1} \la_{(1)}\to (1+\sqrt{\gamma})^2~ \as$ Hence, $\max_{i=1,\ldots,p} |\mu_{(i)}- n^{-1} \la_{(i)}| \to 0~ \as$
This approach was used in \cite{jiang:2004,xiao:zhou:2010} to derive 
\begin{equation}\label{eq:drtgdrghdr1}
\mu_{(1)} \to (1+\sqrt{\gamma})^2\, \quad \text{ and } \quad  \mu_{(p)} \to (1-\sqrt{\gamma})^2\, \quad \as
\end{equation}
%Note that due to self-normalization of the sample correlation matrix we do not have to require $\E[X^2]=1$ which was needed for \eqref{eq:drtgdrghdr}. If $\E[X^2]\neq 1$, \eqref{eq:drtgdrghdr} would have to be adjusted to
%\begin{equation*}
%n^{-1} \la_{(1)} \to (1+\sqrt{\gamma})^2\E[X^2]\, \quad \text{ and } \quad n^{-1} \la_{(p)} \to (1-\sqrt{\gamma})^2\E[X^2]\, \quad \as
%\end{equation*}
If the assumption $\E[X^4]<\infty$ is weakened to $\lim_{\nto} n\,\P(X^4> n)=0$, the paper \cite{bai:yin:1993} proves that $n^{-1} \la_{(1)}\cip (1+\sqrt{\gamma})^2$ and $\max_{i=1,\ldots,p} \big| n/D_i -1 \big| \cip 0$. As a consequence, the limit results for $\mu_{(1)}$ and $\mu_{(p)}$ hold in probability instead of \as
\par
Distributional limit results have been derived for the appropriately centered and normalized eigenvalues of sample correlation matrices. The authors of \cite{bao:pan:zhou:2012} assumed iid, symmetric entries $X_{it}$ and that there exist positive constants $C,C'$ such that $\P(|X|\ge t^C)\le \e^{-t}, t\ge C'$. They showed \eqref{eq:tcch4} with $\lambda_{(1)}/n$ replaced by $\mu_{(1)}$.
%\beao
%n^{2/3}\,\dfrac{(\sqrt{\gamma})^{1/3}}{\big(1+\sqrt{\gamma}\big)^{4/3}}\Big(\mu_{(1)} -
%\big(1+\sqrt{\tfrac pn }\big)^2\Big)
%\std \xi\,,
%\eeao
A similar limit result holds for $\mu_{(p)}$.

\subsection{The case $(X_{it})$ iid and $\E[X^4]=\infty$}

Asymptotic theory for the eigenvalues of $\X\X'$ in the case of an entry distribution with infinite fourth moment was studied in \cite{soshnikov:2004,soshnikov:2006,auffinger:arous:peche:2009} in the cases when $p/n\to \gamma \in (0,\infty)$, while the authors of \cite{davis:mikosch:heiny:xie:2015,heiny:mikosch:2015:iid} allowed nearly arbitrary growth of the dimension $p$. 
In their model, the entries of $\X$ are regularly varying with index $\alpha>0$, implying that
\begin{equation}\label{eq:reg}
\P(|X|>x) = x^{-\alpha} L(x)\,,
\end{equation}
for a slowly varying function $L$. For $\alpha \in (0,4)$,  which implies an infinite fourth moment, they showed that $(a_{np}^{-2} \la_{(1)})$ converges to a \Frechet distributed random variable $\eta_{\alpha/2}$ with parameter $\alpha/2$ while $a_{np}^{-2} \la_{(p)}\cip 0$. Here the normalizing sequence $(a_n)$  is defined via $\P(|X|>a_n)\sim n^{-1}$, hence $n/a_{np}^{2}\to 0$.

%If $\gamma=1$, the sample covariance matrix is almost surely singular. 
To illustrate the stark contrast between the cases $\alpha>4$ and $\alpha<4$, assume \eqref{Cpn} and $\E[X]=0$ if $\E[|X|]<\infty$. Then it follows from \eqref{eq:drtgdrghdr} that
\begin{equation}\label{eq:forch1}
\begin{split}
\frac{\la_{(p)}}{\la_{(1)}} &\cas \frac{(1-\sqrt{\gamma})^2}{(1+\sqrt{\gamma})^2} \quad \text{ if } \alpha >4\,,\\
\frac{a_{np}^2}{n} \frac{\la_{(p)}}{\la_{(1)}} &\cid \frac{(1-\sqrt{\gamma})^2}{\eta_{\alpha/2}} \quad \text{ if } \alpha \in (2,4)\,,\\
\frac{a_{np}^2}{n} \frac{\la_{(p)}}{\la_{(1)}} &\cas 0 \quad \text{ if } \alpha \in (0,2)\,,
\end{split}
\end{equation}
where the rate $a_{np}^2/n\to \infty$ in the last line can even be increased. 
To the best of our knowledge, a suitable normalization $(b_n)$ such that $(b_n \lambda_{(p)})$ has a nontrivial limit is not available when $ \alpha \in (0,2)$.

Under \eqref{Cpn} the asymptotic behavior of the eigenvalues of sample correlation matrices can be very different from that of sample covariance matrices, especially for an entry distribution with infinite fourth moment. If $\alpha\in (2,4)$, the \MP theorem and Theorem 2.3 in \cite{bai:zhou:2008} assert that $(F_{n^{-1}\X\X'})$ and $(F_{\bfR})$  converge weakly to the \MP law. From \cite{baisilv} it is known that $\limsup_n \lambda_{(1)}/n =\infty \,\as$ 
\par

For $\E[X^4]=\infty$, the approach to sample correlation matrices from \eqref{eq:lamu} fails. No limit results for $\mu_{(1)}$ or $\mu_{(p)}$ seem to be available in the literature at this point, although Theorem 2.3 in \cite{bai:zhou:2008} ensures the weak convergence of the empirical spectral distribution $F_{\bfR}$ to the \MP law if $X$ is in the domain of attraction of the normal distribution. Analogously to \eqref{eq:efsdf}, the weak limit of $(F_{\bfR})$ provides a first idea what the limits of the extreme eigenvalues might be.

\subsection{$(X_{it})$ identically distributed, but dependent}
For practical purposes it is important to work with arbitrary population covariance matrices and not just $n^{-1} \E[\X\X']=\bfI$. Based on well understood results in the iid case, numerous generalizations and estimation techniques have been developed. For many models the limiting spectral distribution can only be characterized in terms of an integral equation (=\MP equation) for its Stieltjes transform. Explicit solutions are more involved; see the monographs \cite{bai:silverstein:2010,bai:fang:liang:2014,yao:zheng:bai:2015}. Over the last couple of years siginificant progress on limiting spectral distributions for dependent time series was achieved; see for example %by Banna, Merlev{\`e}de and Peligrad 
\cite{banna:merlevede:peligrad:2015,banna:merlevede:2015,banna:2016}.
Since the sample covariance matrix is a poor estimator for the population covariance matrix in high dimension, a different approach to the fundamental problem of estimating population eigenvalues is needed. In \cite{elkaroui:purdom:2016} the authors find that the bootstrap works for the top eigenvalues if they are sufficiently separated from the bulk. Among others, El Karoui \cite{elkaroui:2008} proposed to use the \MP equation, which basically requires more insight into the empirical spectral distribution and its support. This was achieved in \cite{dobriban:2015}, where an algorithm for calculating the spectral distribution based on certain approximate integral equations for its Stieltjes transform was presented.

In view of \cite{davis:pfaffel:stelzer:2014,davis:mikosch:pfaffel:2015,davis:mikosch:heiny:xie:2015} the behavior of the top eigenvalues is reasonably well understood in the case of linear dependence among the $X_{it}$ and $\E[X^4]=\infty$. 
If $\E[X^4]<\infty$, similar arguments to \eqref{eq:lamu} can be developed to show that methods for sample covariance matrices can be applied to sample correlation matrices; see for example \cite{elkaroui:2009}. Theorem 1 in \cite{elkaroui:2009} proves that if the spectral norm of the population correlation matrix is uniformly bounded and $\E[X^4 (\log X)^{2+\vep}]<\infty$, then the spectral properties of $\bfR$ and $n^{-1} \X\X'$ are asymptotically the same. In particular, if $\lambda_{(1)}/n\cas c$, then $\mu_{(1)}\cas c$.

For the sake of completeness we mention that the study of non-asymptotic high-dimensional sample covariance matrices was subject to an intense line of research in the last years. Good references are \cite{srivastava:vershynin:2013,adamczak:litvak:pajor:jaegermann:2010,adamczak:litvak:pajor:jaegermann:2011,yao:zheng:bai:2015}.

\subsection{About this paper}
In Section~\ref{sec:2ch4} we introduce the basic assumptions of this paper and discuss their meaning.
The main results are given in Section~\ref{sec:3}.
We show that the limiting spectral distribution of the 
sample correlation matrices is the \MP law (Theorem~\ref{thm:mpcorrelation}) 
and that the extreme eigenvalues converge  \as~to the endpoints of the limiting support (Theorem~\ref{thm:mu1convergence}) provided $\bfX$ has iid entries \sth\
their truncated variance is ``almost slowly varying''. 
In this sense, the limiting spectral distribution of sample correlation matrices is universal. 
A similar kind of universality holds for the limiting spectral distribution of sample covariance matrices given a finite variance, while the asymptotic behavior of their extreme eigenvalues is totally different if the fourth moment is infinite.
Thus the eigenvalues of sample correlation matrices exhibit a  ``more robust'' behavior than their sample covariance analogs. 
This is perhaps not surprising in view of the {\em self-normalizing property} of sample correlations.
Self-normalization also has the advantage that one does not have to worry about the correct normalization.
This is a crucial problem in the study of sample covariance matrices in the case of an infinite fourth moment 
where one needs a normalization stronger than the classical one.
We conclude Section~\ref{sec:3} with a small simulation study which shows that the \asy\ results work nicely.
\par
We continue with some technical results in Section~\ref{sec:5.3}. These are of independent interest
because they provide a {\em Path-Shortening Algorithm} for the calculation of bounds for the very high moments of 
$\mu_{(1)}$. We believe that this technique is novel and will be of further use for proving results
in random matrix theory. The proofs of our main results Theorems~\ref{thm:mu1convergence} and \ref{thm:mpcorrelation}
are given in Sections~ \ref{sec:5.2} and \ref{sec:6}, respectively. Both proofs heavily depend on the techniques 
developed in   Section~\ref{sec:5.3}. We conclude with an Appendix which contains some auxiliary analytical results.
\par
Condition \eqref{eq:assumptionq} is crucial for the proof of Theorem~\ref{thm:mpcorrelation}. In Section~\ref{sec:2ch4}
we discuss this condition and find out that it is very close to condition \eqref{eq:condX}
which in turn is very close (but not equivalent) to membership of the \ds\ of $X$ in the domain of attraction
of the Gaussian law. We conjecture that the statement of Theorem~\ref{thm:mpcorrelation} may be proved only under \eqref{eq:condX}.

%--------------------------------------------------------------------------------
\section{Assumptions}\label{sec:2ch4}
In this section we will present some distributional assumptions and discuss their meaning.
We assume that $(X_{it})$ is an iid field with generic element $X$. Recall the notation
\begin{equation}\label{eq:Yit}
Y_{it}=\frac{X_{it}}{\sqrt{D_i}}\,, \quad i=1,\ldots,p;\, t=1,\ldots,n\,.
\end{equation}
For ease of notation we will sometimes write $(Y_1, \ldots, Y_n)= (Y_{11}, \ldots, Y_{1n})$, $Y=Y_1$ and $D=D_1$.
\subsection{Domain of attraction type-condition for the Gaussian law}
One of the basic assumptions in this paper is
\beam\label{eq:condX}
\E \big[ Y_{1}Y_{2} \big] = o(n^{-2}) \quad  \text{ and } \quad  \E \big[ Y_{1}^4 \big] = o(n^{-1})\,,\qquad \nto\,.
\eeam
In \cite{gine:goetze:mason:1997} it was proved that condition \eqref{eq:condX} holds if the \ds\ of 
$X$ is in the domain of attraction of the normal law, which is equivalent to $\E[X^2 \1_{\{|X|\le x\}}]$ being slowly varying. 
\par
The converse implication is not valid. Indeed, let $h(\cdot)$ be a positive function such that $0<c_1 = \liminf_{x \to \infty}h(x) <\limsup_{x \to \infty}h(x) =c_2<\infty$ and consider a symmetric random variable $X$ with tail $\P(X>x)=\P(X<-x)=x^{-2} h(|x|)/2$ for $x$ sufficiently large. Then we have
\begin{equation*}
c_1 = \liminf_{x \to \infty} \frac{\P(|X|>x)}{x^2} <\limsup_{x \to \infty}\frac{\P(|X|>x)}{x^2} =c_2\,,
\end{equation*} 
and therefore $\E[X^2 \1_{\{|X|\le x\}}]$ is not slowly varying, or, equivalently, the \ds\ of $X$ is not in the domain of attraction of the normal law, but \eqref{eq:condX} is valid as a domination argument shows.

\subsection{Condition~\eqref{eq:assumptionq}}
This condition will be crucial for the proofs in this paper:\\[1mm]
{\em There exists a sequence $q=q_n\to \infty$ such that for some integer sequence $k=k_n$ 
with $k/\log n \to \infty$ we have $(k^3 q)/n \to 0$, and the moment inequality
\begin{equation}\label{eq:assumptionq}
\E[ Y_{1}^{2m_1}\cdots Y_{r}^{2m_r} ] \le 
 \frac{q_n}{n} \, \E[ Y_{1}^{2m_1}\cdots Y_{r-1}^{2m_{r-1}}Y_{r}^{2m_r-2} ]\, \tag{$C_q$}
\end{equation}
holds for $1\le r\le \ell-1$ and any positive integers $m_1,\ldots,m_r$ satisfying $m_1+\cdots +m_r=\ell$, where $\ell \le k$.}\\[1mm]
Next we shed some light on this condition. It turns out to be closely related to \eqref{eq:condX}. Indeed, assume \eqref{eq:assumptionq}. Iteration of \eqref{eq:assumptionq} for any fixed $\ell$ yields
\beao
\E[ Y_{1}^{2m_1}\cdots Y_{r}^{2m_r} ] \le  \big(\frac{q_n}{n}\big)^{\ell-r} \E[ Y_{1}^{2}\cdots Y_{r}^{2} ]
%\frac 1 {n(n-1)\cdots (n-r+1)} 
\sim \frac{q_n^{\ell-r}}{n^{\ell}} , \quad \nto\,.
\eeao
In particular, $n\,\E[Y_{1}^4]\le q_n/n\le (\log n)^{-3}$. Thus,  \eqref{eq:assumptionq} provides some precise rate at which 
$n\,\E[Y_{1}^4]$ converges to zero.
%Then we have
%\begin{equation*}
%\begin{split}
%\E[Y_{1}^4] &\sim n^{k-2} \E[Y_{1}^4 Y_{2}^2 \cdots Y_{k-1}^2]
%\le n^{k-2} \frac{q_n}{n}\E[Y_{1}^2 Y_{2}^2 \cdots Y_{k-1}^2]
%\sim q_n n^{-2}\,,
%\end{split}
%\end{equation*}
%which implies $\E[Y_{1}^4] = o(n^{-1})$. 
\par
Moreover, \eqref{eq:assumptionq} does not hold if $\vep=\liminf_{\nto} n\,\E [ Y_{1}^4 ] >0$.
If \eqref{eq:assumptionq} were valid we would have for large $n$,
\begin{equation*}
\vep/2 \le n\,\E[Y_1^4] \le \frac{q_n}{n-1} \, n(n-1) \E[Y_1^2Y_2^2]\le \frac{q_n}{n-1}\to 0\,.
\end{equation*}
For example, Proposition~1 in \cite{mason:zinn:2005} asserts that the \ds\ of $X^2$ is in the 
domain of attraction of an $\alpha/2$-stable distribution with $0< \alpha <2$ if and only if
\begin{equation}\label{eq:limita<2}
\lim_{\nto} n\,\E[Y_{1}^4] =1-\frac{\alpha}{2}\,,
\end{equation}
hence \eqref{eq:assumptionq} does not hold if $|X|$ has a regularly varying tail with index $0<\alpha<2$.
\par
%{\blue Where is this needed? It was said before.} 
%{\green If $X$ is symmetric and in the domain of attraction of the normal law, we have
%\begin{equation*}
%\begin{split}
%\E[ ( Y_{1}+\cdots+Y_{n})^4 ]&=
%\sum_{r=1}^4 \sum_{m_1+\cdots+m_r=4; m_i\ge 1} \binom{n}{r} \binom{4}{m_1,\ldots,m_r} \E[ Y_{1}^{m_1}\cdots Y_{r}^{m_r}] = 3-2n \E[Y_1^4]\,,
%\end{split}
%\end{equation*}
%where we used the identity $n(n-1) \E[ Y_{1}^{2} Y_{2}^{2}] = 1- n \E[ Y_{1}^{4}]$.
%By \cite{gine:goetze:mason:1997}, $\E[ ( Y_{1}+\cdots+Y_{n})^4 ]\to 3$, the fourth moment of the standard normal distribution, and hence $n\,\E[Y_{1}^4] = o(1)$.}

The expectations in \eqref{eq:assumptionq} can be calculated by using the following formula due to 
\cite{gine:goetze:mason:1997}:
\begin{equation}\label{eq:formulagine}
\E[ Y_{1}^{2m_1}\cdots Y_{r}^{2m_r} ] = \frac{1}{(k-1)!} \int_0^{\infty} 
\la^{k-1} (\E[\e^{-\la X^2}])^{n-r}  \prod_{j=1}^r \E[X^{2m_j}\e^{-\la X^2} ] \, \dint \la\,,
\end{equation}
where $1\le r\le k$, $m_1+\cdots+m_r=k$ and $m_i\ge 1$.
%This gives the explicit expression
%\begin{equation}\label{eq:fourthself}
%\E[Y_{1}^4]=  \int_0^{\infty} 
%\la (\E[\e^{-\la X^2}])^{n-1}  \E[X^{4}\e^{-\la X^2} ] \, \dint \la\,.
%\end{equation}
\par
We present some examples of distributions of $X$  which satisfy \eqref{eq:assumptionq}. %For any bounded random variable $|X|\le c$ we can choose $q_n=\log n$.

\begin{example}[Standard normal distribution]\label{ex:normal}{\em
Assume $X_i \sim N(0,1)$. We calculate $\E[Y_{1}^{2m_1}\cdots Y_{r}^{2m_r} ]$ for the standard normal distibution via \eqref{eq:formulagine}. Since $X_1^2$ has $\chi^2$-distribution we know for $\lambda\ge 0$ that $\E[\e^{-\la X^2}]=(1+2\la)^{-1/2}$.
We have 
\begin{equation*}
\frac{\dint^m}{\dint \la^m} \e^{-\la X^2} = (-X^2)^m \e^{-\la X^2}\,.
\end{equation*}
%Interchanging the order of differentiation and integration, which is justified for all $\lambda \ge 0$ since $\E[\e^{-\la X^2}]<\infty$, we obtain 
%\begin{equation}\label{eq:changelse}
%(-1)^{m} \E[X^{2m}\e^{-\la X^2} ]=  \E\Big[\frac{\dint^{m}}{\dint \la^{m}} \e^{-\la X^2}\Big]=\frac{\dint^{m}}{\dint \la^{m}} \E[\e^{-\la X^2}]\,.
%\end{equation}
Calculation yields 
\begin{equation}\label{eq:diffmgf}
(-1)^{m} \E[X^{2m}\e^{-\la X^2} ]
%\frac{\dint^m}{\dint \la^m} (1+2\la)^{-1/2} = \frac{2^m \Gamma(1/2)}{\Gamma(1/2-m)} (1+2\la)^{-1/2-m}%= \frac{(-2)^m (2m)!}{m!} (1+2\la)^{-1/2-m}
=(-1)^n (2m-1)!! (1+2\la)^{-1/2-m}\,.
\end{equation}
By \eqref{eq:formulagine} and \eqref{eq:diffmgf}, we have for $\ell\le k$
\begin{equation*}
\begin{split}
\E[Y_{1}^{2m_1}\cdots Y_{r}^{2m_r} ] &= \frac{1}{(\ell-1)!} \int_0^{\infty} 
\la^{\ell-1} (\E[\e^{-\la X^2}])^{n-r}  \prod_{j=1}^r \E[X^{2m_j}\e^{-\la X^2} ] \, \dint \la\\
%&= \frac{1}{(k-1)!} \int_0^{\infty} 
%\la^{k-1} (\E[\e^{-\la X^2}])^{n-r}  \prod_{j=1}^r (-1)^{m_j} \frac{\dint^{m_j}}{\dint \la^{m_j}} \E[\e^{-\la X^2}]  \, \dint \la\\
&= \frac{1}{(\ell-1)!} \int_0^{\infty} 
\la^{\ell-1} (1+2\la)^{-(n+2\ell)/2}  \, \dint \la \prod_{j=1}^r (2m_j-1)!!\,.
\end{split}
\end{equation*}
Since 
\begin{equation*}
\int_0^{\infty} 
\la^{\ell-1} (1+2\la)^{-(n+2\ell)/2}  \, \dint \la = \frac{\Gamma(n/2) \Gamma(\ell)}{2^{\ell} \Gamma(n/2+\ell)}\,,
\end{equation*}
one obtains
\begin{equation}\label{eq:momentsnormal}
\E[ Y_{1}^{2m_1}\cdots Y_{r}^{2m_r}] = \frac{\Gamma(n/2) }{2^\ell \Gamma(n/2+\ell)} \prod_{j=1}^r (2m_j-1)!!\,,
\end{equation}
which allows one to conclude that
\begin{equation*}
\begin{split}
\frac{\E[ Y_{1}^{2m_1}\cdots Y_{r}^{2m_r}]}{\E[ Y_{1}^{2m_1}\cdots Y_{r}^{2m_r-2}]}&=
\frac{2m_r-1}{n+2\ell-2}\le \frac{2k}{n}\,,
\end{split}
\end{equation*}
where we used $m_r\le \ell\le k$. Hence \eqref{eq:assumptionq} holds with $q_n=2k_n$.
}\end{example}

\begin{example}[Gamma distribution]\label{ex:gamma}{\em
 Assume $X^2\sim \text{Gamma}(\alpha,\beta)$, $\alpha,\beta>0$. In this case
\begin{equation*}
\frac{\dint^{m}}{\dint \la^{m}} \E[\e^{-\la X^2}]=\frac{\dint^{m}}{\dint \la^{m}} \Big(1+\frac{\la}{\beta}\Big)^{-\alpha} = \frac{\Gamma(1-\alpha )}{\Gamma(1-\alpha -m)} \beta^{-n} \Big(1+\frac{\la}{\beta}\Big)^{-\alpha-n}\,.
\end{equation*}
For $\ell\le k$ one can calculate
\begin{equation*}
\begin{split}
\E[ Y_{1}^{2m_1}\cdots Y_{r}^{2m_r}] 
&= \frac{1}{(\ell-1)!} \int_0^{\infty} 
\la^{\ell-1} (\E[\e^{-\la X^2}])^{n-r}  \prod_{j=1}^r (-1)^{m_j} \frac{\dint^{m_j}}{\dint \la^{m_j}} \E[\e^{-\la X^2}]  \, \dint \la\\
&= \frac{\Gamma(\alpha n)(-1)^{\ell}}{\Gamma(\alpha n +\ell)} \prod_{j=1}^r  \frac{\Gamma(1-\alpha )}{\Gamma(1-\alpha -m_j)}\,.
\end{split}
\end{equation*}
Similarly to the previous example, \eqref{eq:assumptionq} holds with $q_n=(k_n+\alpha)/\alpha$.
}\end{example}

%----------------------------------------------------------------------		
\section{Main results}\label{sec:3}
Our first result identifies the limit of the empirical spectral \ds\ 
$F_{\bfR}$ of the sample correlation matrix $\bfR$
for iid random fields $(X_{it})$ with generic element $X$.

\begin{theorem}[Limiting spectral distribution]\label{thm:mpcorrelation}
Assume the condition \eqref{Cpn}.
\begin{enumerate}
\item[(1)] If $X$ is centered and \eqref{eq:condX} holds
then the sequence $(F_{\bfR})$ converges weakly to the \MP law given in \eqref{eq:MPch4}.
\item[(2)]
If $X$ is symmetric and  \eqref{eq:condX} does not hold, i.e.,
$\liminf_{\nto} n\,\E [ Y^4 ] >0$,
then 
\beao
\liminf_{\nto} \E\Big[\int x^k F_{\bfR}(dx)\Big]> \beta_k(\gamma)\,,\qquad k\ge 1\,,
\eeao
where $\beta_k(\gamma)$ is the $k$-th moment of the \MP law given in \eqref{eq:momentsmp}.
\end{enumerate}
\end{theorem}
The proof of parts (1) and (2) will be given in Sections~\ref{sec:5.1} and \ref{sec:5.5}, repectively.
\begin{remark}{\em 
Part (1) with condition \eqref{eq:condX} replaced by $\E[X^2]<\infty$ was proved in \cite{jiang:2004}. Later,
in \cite{bai:zhou:2008} the finite variance assumption was replaced by the weaker condition that
the distribution of $X$ belongs to the domain of attraction of the normal law. We discussed in the previous
section that   \eqref{eq:condX} holds under the latter condition.
Part (2) shows that \eqref{eq:condX} is the minimal condition for part (1). By Lemma B.1 in \cite{bai:silverstein:2010},
$\lim_{\nto} \E\big[\int x^k F_{\bfR}(dx)\big]=\beta_k(\gamma)\,, k\ge 1\,,$
implies weak convergence of $F_{\bfR}$ to the \MP distribution as the latter is uniquely determined by its moments $(\beta_k(\gamma))_{k\ge 1}$.
}\end{remark}

If $X$ is symmetric, $n\,\E[Y^4]=o(1)$ and $p/n\to 0$, a slight modification of the proof of part (2) combined with the method of moments yields $F_{\bfR}\to \1_{[1,\infty)}$ weakly. Consequently, 
for any $\vep \in(0,1)$ the number of eigenvalues outside $(1-\vep,1+\vep)$ is $o(p)$ \as~ In particular, if $p$ is fixed, then $\mu_{(1)}$ and $\mu_{(p)}$ converge to $1$ \as~ In view of part (2), one concludes that $n\E[Y^4]=o(1)$ is a necessary and sufficient condition for the a.s. convergence of the eigenvalues $(\mu_{(i)})$ if $X$ is symmetric and $p$ fixed.

When $p\to \infty$ one has to deal with the potentially $o(p)$ eigenvalues outside the support of the limiting spectral distribution. We develop a method to overcome this problem at the expense of strengthening the assumption $n\,\E[Y^4]=o(1)$ to \eqref{eq:assumptionq}.

A Borel--Cantelli argument to obtain an upper bound for $\limsup_n \mu_{(1)}$ requires an adequate bound on $\E[\mu_{(1)}^{k_n}]$, where $k_n\to \infty$. To this end, we use the inequality
\begin{equation*}
\E[\mu_{(1)}^{k_n}]\le \E[\tr \bfR^{k_n}] %=p \E\big[\int x^{k_n} F_{\bfR}(dx)\big]
= \sum_{i_1,\ldots,i_{k_n}=1}^p \sum_{t_1,\ldots,t_{k_n}=1}^n \E[ Y_{i_1t_{k_n}} Y_{i_1t_1} Y_{i_2t_1}Y_{i_2t_2} \cdots Y_{i_{k_n}t_{{k_n}-1}} Y_{i_{k_n}t_{k_n}}  ]
\end{equation*}
and determine those summands on the \rhs~which are largest when weighted by their multiplicities. Using our {\em Path-Shortening Algorithm}, which is a novel technique that efficiently uses the inherent structure of sample correlation matrices, their contribution is calculated explicitly. The other summands can --with considerable technical effort-- be controlled by \eqref{eq:assumptionq}. 
Note that because of the identity $\E[\tr \bfR^{k_n}] =p\, \E\big[\int x^{k_n} F_{\bfR}(dx)\big]$ the behavior of moments of the empirical spectral distribution is closely linked to the above upper bound.

The following result provides general conditions for the a.s.~\con\ of 
the largest and smallest eigenvalues $\mu_{(1)}$ and $\mu_{(p)}$ of $\bfR$ to the endpoints of the 
\MP law. The proof of this result is given in Section~\ref{sec:5.2}.
\begin{theorem}[Limit of extreme eigenvalues]\label{thm:mu1convergence}
Assume \eqref{Cpn}. 
\begin{itemize}
\item[(1)] If $\E[X^4]<\infty$ and $\E[X]=0$ 
\item[(2)] or $X$ is symmetric and satisfies condition \eqref{eq:assumptionq},
\end{itemize}
then 
\begin{equation}\label{eq:mlmgd}
{\mu}_{(1)} \to (1+\sqrt{\gamma})^2 \, \quad \as\,,
\end{equation}
\begin{equation}\label{eq:limitmup}
{\mu}_{(p)} \to (1-\sqrt{\gamma})^2 \, \quad \as\phantom{\,,}
\end{equation}
\end{theorem}
\begin{remark}{\em Part (1) was proved in \cite{jiang:2004,xiao:zhou:2010}; see the discussion
in Section~\ref{sec:sdlgj}. 
Theorem~\ref{thm:mu1convergence} indicates that the a.s. convergence of the extreme eigenvalues of $\bfR$
does not depend on the finiteness of the fourth or even second moment. 
This is in stark contrast to the a.s. behavior of $n^{-1}\lambda_{(1)}$, the largest eigenvalue of the sample covariance matrix 
$n^{-1}\bfX\bfX'$. Note that there is a phase transition of the a.s. \asy\ behavior of the  
extreme eigenvalues at the border between finite and infinite fourth moment of $X$, 
while such a transition occurs for the empirical spectral distribution at the border between finite and infinite variance.}
\end{remark}
\subsection{Simulation study}
\begin{figure}[htb!]
  \centering
  \subfigure[Sample correlation: $\mu_{(p)}=0.086, \mu_{(1)}=2.898$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{corr_t_alpha6_n2000_p1000__mumax2.8988_b2.9142_mumin0.0860_a0.0858}.pdf}
  }
	\qquad
  \subfigure[\mbox{Sample covariance: $\lambda_{(p)}/(n \E{[X^2]})$} $=0.085,\lambda_{(1)}/(n \E{[X^2]})=2.908$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{cov_t_alpha6_n2000_p1000__mumax2.9086_b2.9142_mumin0.0855_a0.0858}.pdf}
  }
\caption{Histogram and \MP density for $X \sim t_6$, $n=2000, p=1000$.  $\gamma=0.5, (1-\sqrt{\gamma})^2=0.085, (1+\sqrt{\gamma})^2=2.914$.}
  \label{fig:alpha6}
\end{figure}

\begin{figure}[htb!]
  \centering
\subfigure[Sample correlation: $\mu_{(p)}=0.088, \mu_{(1)}=2.880$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{corr_t_alpha3_n2000_p1000__mumax2.8809_b2.9142_mumin0.0884_a0.0858}.pdf}
		}
		\qquad
	\subfigure[\mbox{Sample covariance: $\lambda_{(p)}/(n \E{[X^2]})$} $=0.083$, $\lambda_{(1)}/(n \E{[X^2]})=8.870$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{cov_t_alpha3_n2000_p1000__mumax8.8707_b2.9142_mumin0.0839_a0.0858}.pdf}
  }
\caption{Histogram and \MP density for $X \sim t_3$, $n=2000, p=1000$. Here $\gamma=p/n=0.5, (1-\sqrt{\gamma})^2=0.085, (1+\sqrt{\gamma})^2=2.914$.}
  \label{fig:alpha3}
\end{figure}

\begin{figure}[htb!]
  \centering
\subfigure[Sample correlation: $\mu_{(p)}=0.086, \mu_{(1)}=2.902$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{corr_symmetricpareto_alpha3.99_n2000_p1000__mumax2.9024_b2.9142_mumin0.0865_a0.0858}.pdf}
  }
	\qquad
  \subfigure[\mbox{Sample covariance: $\lambda_{(p)}/(n \E{[X^2]})$} $=0.083, \lambda_{(1)}/(n \E{[X^2]})=3.176$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{cov_symmetricpareto_alpha3.99_n2000_p1000__mumax3.1766_b2.9142_mumin0.0838_a0.0858}.pdf}
  }
	\caption{Histogram and \MP density: $X \eid Z_1-Z_2$ for $Z_i\sim \text{Pareto}(3.99)$, $n=2000, p=1000$. Here $\gamma=p/n=0.5, (1-\sqrt{\gamma})^2=0.085, (1+\sqrt{\gamma})^2=2.914$.}
	\label{fig:pareto399}
	\subfigure[Sample correlation: $\mu_{(p)}=0.469, \mu_{(1)}=1.731$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{corr_t_alpha2.1_n10000_p1000_mumax1.7318_b1.7325_mumin0.4697_a0.4675}.pdf}
  }
	\qquad
  \subfigure[\mbox{Sample covariance: $\lambda_{(p)}/(n \E{[X^2]})$} $=0.159, \lambda_{(1)}/(n \E{[X^2]})=35.319$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{cov_t_alpha2.1_n10000_p1000_mumax35.3196_b1.7325_mumin0.1596_a0.4675}.pdf}
  }
\caption{Histogram and \MP density for $X \sim t_{2.1}$, $n=10000, p=1000$. Here $\gamma=p/n=0.1, (1-\sqrt{\gamma})^2=0.467, (1+\sqrt{\gamma})^2=1.732$.}
  \label{fig:stability}
\end{figure}

\begin{figure}[htb!]
  \centering
\subfigure[$X \eid Z^2-\E Z^2$ for $Z\sim t_{1.5}$, $n=2000, p=1000, \gamma=0.5$.]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{corr_nonsymmt_alpha1.5_n2000_p1000__mumax3.4725_b2.9142_mumin0.0333_a0.0858}.pdf}
  }
	\qquad
\subfigure[$X \sim t_{1.8}$, $n=10000, p=1000, \gamma=0.1$]{
    \includegraphics[trim = 1.5in 3.4in 1.5in 3.6in, clip, scale=0.4]{{corr_t_alpha1.8_n10000_p1000_relativelystable}.pdf}
  }
\caption{Histogram of $(\mu_{(i)})$ and \MP density}
  \label{fig:smallalpha}
\end{figure}

In this subsection we simulate a large data matrix $\X$ 
of iid entries. We compare the spectra of $\X\X'/n$ and $\bfR$ to the limiting \MP spectral density with appropriate 
parameter $\gamma$; see Theorem \ref{thm:mpcorrelation}. We simulate from different distributions of $X$ and choose various values for $p$ and $n$ to cover \MP distributions of several shapes. 
In what follows, we assume $\E[X^2]=1$, whenever the second moment is finite. 
\par

In Figure~\ref{fig:alpha6} we simulated a $1000\times 2000$ data matrix $\X$ with iid entries drawn from a $t_6$-distribution which we renormalized to meet the requirement $\E[X^2]=1$. To illustrate the weak convergence of $(F_{\bfR})$ and $(F_{\X\X'/n})$ we plot the histograms of the eigenvalues $(\mu_{(i)})$ and $(\lambda_{(i)}/n)$ and compare them to the \MP distribution with $\gamma=1/2$. %Note that the histograms of $(\mu_{(i)})$ and $(\lambda_{(i)}/n)$ are practically identical.
As expected in the case $\E[X^4]<\infty$, the values $n^{-1} \la_{(1)}=2.9086$ and $n^{-1} \la_{(p)}=0.0855$ are very close to their theoretical almost sure limits $2.9142$ and $0.0858$, respectively. The same is valid for $\mu_{(1)}$ and $\mu_{(p)}$.
\par

In Figure~\ref{fig:alpha3} we simulate $X$ from a renormalized $t_3$-distribution with unit variance. The histograms of $(\mu_{(i)})$ and $(\lambda_{(i)}/n)$ resemble the corresponding \MP density $f_{1/2}$. Note that $\lambda_{(1)}/n$ can be larger than the right endpoint $(1+\sqrt{\gamma})^2$ since it has a different limit behavior than in the case $\E[X^4]<\infty$, while $\mu_{(p)}$ and $\mu_{(1)}$ are close to the endpoints $(1-\sqrt{\gamma})^2$ and $(1+\sqrt{\gamma})^2$, respectively, for which Theorem~\ref{thm:mu1convergence} provides a formal justification.  
\par

In Figures~\ref{fig:pareto399} and \ref{fig:stability} we simulated from distributions with infinite fourth moment. 
We drew from a symmetrized Pareto distribution with parameter $3.99$ to create the plots in Figure~\ref{fig:pareto399}. 
Note that in this case $\E[X^{3.99}]=\infty$, while $\E[X^{3.99-\vep}]<\infty$ for any $\vep >0$, i.e., 
we are at the ``border'' between finite and infinite fourth moment. 
The extreme eigenvalues in the sample correlation case are very close to their theoretical limits stated in Theorem~\ref{thm:mu1convergence}, whereas 
the largest eigenvalues of the sample covariance matrix cease to lie within the support of the \MP distribution. 
Note that the assumption $\E[X^2]=1$ is superfluous for the sample correlation plots due to self-normalization. For the histogram of $(\lambda_{(i)}/n)$ the knowledge of the correct value $\E[X^2]$ is crucial since, for instance, $\lambda_{(1)}/n \to (1+\sqrt{\gamma})^2 \E[X^2]\, \as$ 
In applications, $\E[X^2]$ needs to be estimated first and estimation errors might significantly alter the conclusion. One can easily imagine that Figure~\ref{fig:alpha6}(b) with a misspecified variance of the data could have resembled Figure~\ref{fig:pareto399}(b). In this respect sample correlations are more robust. 
\par

In Figure~\ref{fig:stability}, we choose $X$ from the standardized
$t_{2.1}$-\ds , moving closer to the infinite variance case. The histogram of $(\mu_{(i)})$ fits the \MP density very well and the extreme eigenvalues are located in a close proximity of the endpoints of the \MP support. The sample covariance case in (b) does not 
look particularly appealing due to the fact that there are a few relatively large eigenvalues. For example, $\lambda_{(1)}/n=35.3196$ while $(1+\sqrt{\gamma})^2$ is only $1.7325$. By \cite{auffinger:arous:peche:2009,heiny:mikosch:2015:iid,davis:mikosch:heiny:xie:2015}, the properly normalized $\lambda_{(1)}$ converges to a \Frechet distributed random variable. The correct normalization is roughly $n^{4/2.1}$ and hence it is expected that $\lambda_{(1)}/n$ is separated from the bulk, whose behavior ultimately determines the limiting spectral distribution, which is the \MP law with parameter $\gamma=0.1$. However, due to the separation between the top eigenvalues and the bulk, it is not obvious from a histogram with only $50$ classes that the \MP law provides a good fit to the spectral distribution in (b). This different behavior of sample correlations and covariances is an additional argument for the higher stability of results obtained from an analysis of the sample correlation matrix.  
\par

Finally, we present two histograms of $(\mu_{(i)})$ with $\E[X^2]=\infty$ in Figure~\ref{fig:smallalpha}. 
In (a), we choose the non-symmetric $X \eid Z^2-\E Z^2$ for $Z\sim t_{1.5}$. 
In (b), the simulated $X$ is standardized $t_{1.8}$. 
The plots look surprisingly stable, given that the empirical spectral distribution 
does not weakly converge to the \MP law; see Theorem~\ref{thm:mpcorrelation}(2). 
The extreme eigenvalues $\mu_{(1)}$ and $\mu_{(p)}$ are much further away from $(1-\sqrt{p/n})^2$ and $(1+\sqrt{p/n})^2$, respectively, than in all the other sample correlation histograms we have seen so far.

\subsection{A remark on the centered sample correlation matrix}

%{\blue ???} This remark is based on Section 2.3.4 in \cite{elkaroui:2009}. 
We presented results for the matrices $\bfR$ and $\X\X'$, assuming that $\E[X]=0$ when $\E[|X|]<\infty$. 
In practice, the expectation of $X$ typically has to be estimated. We discuss what has to be changed in the
aforementioned theory in this case. We consider the matrix $\widetilde{\X}\widetilde{\X}'$, where 
\begin{equation*}
\widetilde{X}_{it}= X_{it}- \overline{X}_i \quad \text{ and } \quad \overline{X}_i= \frac{1}{n} \sum_{t=1}^n X_{it}\,.
\end{equation*}
and the corresponding correlation matrix $\widetilde{\bfR} = \widetilde{\bfF}^{1/2} \widetilde{\X}\widetilde{\X}'\widetilde{\bfF}^{1/2}$, where $\widetilde{\bfF}$ is the $p\times p$ diagonal matrix with entries 
\begin{equation*}
\widetilde{F}_{ii}= \frac{1}{(\widetilde{\X}\widetilde{\X}')_{ii}}\,, \quad i=1,\ldots,p.
\end{equation*}
\par
In contrast to \eqref{eq:lamu} an application of Weyl's inequality \cite{bai:silverstein:2010} yields
\begin{equation}\label{eq:sgsgrs}
n^{-1} |\la_{(1)}(\X\X')- \la_{(1)}(\widetilde{\X}\widetilde{\X}')| \le n^{-1} \twonorm{\X\X'-\widetilde{\X}\widetilde{\X}'}\,,
\end{equation}
where, in general, the \rhs\ does not converge to zero. 
However, since $\X-\widetilde{\X}$ is a rank $1$ matrix, it is known from \cite{bai:silverstein:2010} that $n^{-1} \X\X'$ and $n^{-1}  \widetilde{\X}\widetilde{\X}'$ share the same limiting spectral distribution (if it exists) with right endpoint $b$ say. Therefore we have
\begin{equation*}
\liminf_{\nto} \frac{\lambda_{(1)}(\widetilde{\X}\widetilde{\X}')}{n} \ge b\,.
\end{equation*}
Following \cite{elkaroui:2009}, we let $\mathbf{H}= \bfI_n-n^{-1} \mathbf{1}\mathbf{1}'$, where $\mathbf{1}=(1,\ldots,1)'$. Then one can write $\widetilde{\X}=\X \mathbf{H}$ and since $\mathbf{H}$ is a symmetric matrix with $(n-1)$ eigenvalues equal to $1$ and one eigenvalue equal to $0$ we see that
\begin{equation*}
\la_{(1)}(\widetilde{\X}\widetilde{\X}') \le \la_{(1)}(\X\X')\,.
\end{equation*} 
We conclude
\begin{equation*}
\lim_{\nto} \frac{\lambda_{(1)}(\widetilde{\X}\widetilde{\X}')}{n} = b\quad \as
\end{equation*}
whenever $\la_{(1)}(\X\X')/n \to b$ a.s. Therefore the a.s. behavior of the largest eigenvalues of $\bfX\bfX'$ and $\wt \bfX\wt\bfX'$ are
closely related. %{\blue about the other eigenvalues? J: I am not aware of a nice result.} 
%{\red Find similar argument for $\widetilde{\bfR}$. that it is sufficient to study $\bfR$ as we do.}

Due to the shift and scale invariance of sample correlations, the aforementioned arguments remain valid for the ordered eigenvalues 
%{\blue is this obvious or does it follow from Jiang? J: They are the same.}
\begin{equation*}
\widetilde{\mu}_{(1)}\ge \cdots \ge \widetilde{\mu}_{(p)}
\end{equation*}
of $\widetilde{\bfR}$ if $\E[X^4]<\infty$ and $\E[X]=c$ (not necessarily zero), as shown in \cite{jiang:2004}. Then we have $\widetilde{\mu}_{(1)} \to (1+\sqrt{\gamma})^2$~ \as ~ and $\widetilde{\mu}_{(p)} \to (1-\sqrt{\gamma})^2$ ~ \as~
In Theorem 2 of \cite{jiang:2004} it was proven that if $\E[X^2]< \infty$ and $p/n\to \gamma\in (0,\infty)$, the empirical spectral distribution of $\widetilde{\bfR}$ converges weakly to the \MP law.

%----------------------------------------------------------------------

\section{Technical results}\label{sec:5.3}
In this section we provide most technical results required for the proofs of the main theorems. 
We develop a new approach which efficiently uses the structure of sample correlation matrices.
The goal of this section is to prove Proposition~\ref{prop:paths}.
\par
Throughout $(X_{it})$ are iid symmetric. We will study the moments
\begin{equation*}
\sum_{t_1,\ldots,t_k=1}^n \E[ Y_{i_1t_k} Y_{i_1t_1}  Y_{i_2t_1} Y_{i_2t_2} Y_{i_3t_2} Y_{i_3t_3} \cdots Y_{i_kt_{k-1}} Y_{i_kt_k}  ].
\end{equation*}
for $k\ge 1$  and various choices of {\em paths} $I=(i_1, i_2, \ldots, i_k)\in \{ 1,\ldots, p\}^k$. In this case, $\length(I)=k$ is the {\em length of the path}.
We say that a {\em path} $(i_1, i_2, \ldots, i_k)$ is an $r$-{\em path} if it contains exactly $r$ distinct components.
% i.e., $|\{i_1, i_2, \ldots, i_k\} |=r$. 
A path is {\em canonical} if $i_1=1$ and $i_l \le \max \{i_1, \ldots, i_{l-1}\} +1, l\ge 2$. 
A canonical $r$-path satisfies $\{i_1, i_2, \ldots, i_k\} = \{1, \ldots, r\}$.
Two paths are {\em isomorphic} if one becomes the other by a suitable permutation on $(1,\ldots,p)$. Each {\em isomorphism class} contains exactly one canonical path.
%Furthermore, we say that a path contains a duplicate pair (DP) if and only if there exist $1\le l_1<l_2<l_3<l_4\le k$ such that $i_{l_1}=i_{l_3}\neq i_{l_2}=i_{l_4}$.
For $k\ge 1$, define
\beao
f(I,T)=\E[ Y_{i_1t_k} Y_{i_1t_1}  Y_{i_2t_1} Y_{i_2t_2} Y_{i_3t_2} Y_{i_3t_3} \cdots Y_{i_kt_{k-1}} Y_{i_kt_k}  ]\,,\quad I,T \in \{1,\ldots,k \}^k\,.
\eeao
Finally, define $F(\emptyset)=n$ and
\beao
F(i_1,\ldots,i_k)=F_n(i_1,\ldots,i_k)=\sum_{t_1,\ldots,t_k=1}^n f((i_1,\ldots,i_k),(t_1,\ldots,t_k))\,.
\eeao 
Note that $F(I_1)=F(I_2)$ if $I_1,I_2$ lie in the same isomorphism class. Therefore, whenever we are interested in 
$F(I)$ we can assume without loss of generality that $I$ is canonical. 
\par
In what follows, 
we will consider transformations of the path $I$ leading to a new path $S(I)$. 
For ease of notation, we will also assume $S(I)$ canonical. 
If it is not canonical, we can always work with its {\em canonical representative}, 
the unique canonical path in its isomorphism class. 
\par
When calculating values of $F$, the path-shortening function $PS$ will be useful. 
Let $I=(i_1,\ldots,i_k)\in \{1,\ldots,k \}^k$. $PS(I)$ is the output of the following algorithm.

\subsection*{Path-Shortening Algorithm $PS(I)$.}
\begin{itemize}
\item[Input: ] Path $I=(i_1,\ldots,i_k)$. Set $J=I$ and $R=0, \runs =0$. 
\item[Step 0:] Set $l= \length(I)$. Go to Step 1.
\item[Step 1:] Erase runs. 
\begin{itemize}
\item If $i_j=i_{j+1}$ for some $1\le j\le l$, where we interpret $i_{l+1}$ as $i_1$, erase element $i_j$ from the path. Set $I=(i_1, \ldots, i_{j-1}, i_{j+1}, \ldots, i_l)$, $\runs = \runs +1$ and return to Step 0.
\item Otherwise proceed with Step 2.
\end{itemize}
\item[Step 2:] Let $R_1$ be the number of elements of the path $I$ which appear exactly once. Set $R:=R+R_1$. Then define $I$ to be the resulting (possibly shorter) path which is obtained by deleting those $R_1$ elements from the path $I$. Go to Step 3.
\item[Step 3:] 
\begin{itemize}
\item If $J=I$, then return $(I,R,\runs)$ as output. 
\item If $J\neq I$, set $J:=I$ and return to Step 0.
\end{itemize}
\end{itemize}

\begin{definition}\label{def:pathshortening}
The path-shortening function $PS$ is the output  $(S(I),R(I),\runs(I))$ of the Path-Shortening Algorithm (PSA) 
where $S(I)$ is the resulting shortened path and $R(I)$ is the total number of 
elements that were removed in Step 2 of the PSA. We write 
$PS(I)= (S(I),R(I),\runs(I))$.
\end{definition}

\subsection*{Properties of $PS(I)$.}
Clearly, ${\rm length}(S(I))\le {\rm length}(I)$. 
If $I=(1,\ldots,r)$ then $S(I)=\emptyset$, which shows that $S(I)$ can have length zero. 
Furthermore, all elements in $S(I)$ appear at least twice. If $I$ is an $r$-path then $R(I)\le r$.
%We introduce the notion of duplicate pairs, which are a part of every irreducible path under Algorithm~\ref{alg:psi}.
\begin{lemma}\label{prop:PSI}
For any $I \in \{1,\ldots,k \}^k$, we have $F(I)=F(S(I))\, n^{-R(I)}$.
\end{lemma}
\begin{proof}
%{\blue delete: Let $k\ge 1$ and $I \in \{1,\ldots,k \}^k$.} 
We shall look at the changes made to $I$ in Steps 1 and 2 of the PSA  separately.\\
{\em Assume we are in Step 1.}
\begin{itemize}
\item If $i_j=i_{j+1}$ for some $1\le j\le l$, where we interpret $i_{l+1}$ as $i_1$, erase element $i_j$ from the path. Set $S_1(I)=(i_1, \ldots, i_{j-1}, i_{j+1}, \ldots, i_l)$.
\item Otherwise, $S_1(I)=I$.
\end{itemize}
Since Step 1 does not influence the value of $R$ it suffices to show $F(I)=F(S_1(I))$. 
If $S_1(I)=I$ there is nothing to show. Therefore assume $i_j=i_{j+1}$ for some $j$. In this case,  we have
\beam\label{eq:proofstep1}
F(I)%&=\sum_{t_1,\ldots,t_k=1}^n \E[ Y_{i_1t_k} Y_{i_1t_1} \cdots Y_{i_{j-1}t_{j-2}} Y_{i_{j-1}t_{j-1}}
%Y_{i_{j}t_{j-1}} Y_{i_{j}t_{j}}^2 Y_{i_{j}t_{j+1}}
%Y_{i_{j+2}t_{j+1}} Y_{i_{j+2}t_{j+2}} \cdots Y_{i_kt_{k-1}} Y_{i_kt_k}  ]\\
&=& \sum_{t_1,\ldots,t_{j-1}, t_{j+1}, \ldots, t_k=1}^n \E[ Y_{i_1t_k} Y_{i_1t_1}\cdots Y_{i_{j-1}t_{j-2}} Y_{i_{j-1}t_{j-1}} 
Y_{i_{j}t_{j-1}} \sum_{t_j=1}^n Y_{i_{j}t_{j}}^2 \nonumber\\
&& \qquad \qquad Y_{i_{j}t_{j+1}} Y_{i_{j+2}t_{j+1}} Y_{i_{j+2}t_{j+2}} \cdots Y_{i_kt_{k-1}} Y_{i_kt_k}  ]\nonumber\\
&=& \sum_{t_1,\ldots,t_{j-1}, t_{j+1}, \ldots, t_k=1}^n \E[ Y_{i_1t_k} Y_{i_1t_1} \cdots Y_{i_{j-1}t_{j-2}} Y_{i_{j-1}t_{j-1}} 
Y_{i_{j}t_{j-1}}  Y_{i_{j}t_{j+1}}\nonumber\\
&& \qquad \qquad Y_{i_{j+2}t_{j+1}} Y_{i_{j+2}t_{j+2}} \cdots Y_{i_kt_{k-1}} Y_{i_kt_k} ]= F(S_1(I))\,,
\eeam
where we used $\sum_{t=1}^nY_{it}^2=1$. This proves that Step 1 poses no problem.\\[1mm]
{\em Next we turn to Step 2.} Without loss of generality we can assume that $I$ 
does not contain any runs. If all elements of $I$ appear at least twice there is nothing to prove. 
Therefore assume the $j$th element $i_j$ appears only once and $R_1=1$. 
Let $S_2(I)$ denote the path $I$ with the $j$th element removed. 
Thus we have to show $F(I)=F(S_2(I))n^{-1}$.  In this case,  we have
\beam\label{eq:proofstep2}
F(I)&=&\sum_{t_1,\ldots,t_{j-1}, t_{j+1}, \ldots, t_k=1}^n \sum_{t_j=1}^n \E[ Y_{i_{j}t_{j-1}} Y_{i_{j}t_{j}} ]\nonumber\\&&\qquad\times \E[ Y_{i_1t_k} Y_{i_1t_1} \cdots Y_{i_{j-1}t_{j-2}} Y_{i_{j-1}t_{j-1}}
Y_{i_{j+1}t_{j}} Y_{i_{j+1}t_{j+1}} \cdots Y_{i_kt_{k-1}} Y_{i_kt_k}  ]\nonumber\\
&=&\sum_{t_1,\ldots,t_{j-1}, t_{j+1}, \ldots, t_k=1}^n n^{-1} \E[ Y_{i_1t_k} Y_{i_1t_1} \cdots Y_{i_{j-1}t_{j-2}} Y_{i_{j-1}t_{j-1}}
Y_{i_{j+1}t_{j-1}}\nonumber\\&&\qquad\times
Y_{i_{j+1}t_{j+1}} \cdots Y_{i_kt_{k-1}} Y_{i_kt_k}  ]\nonumber\\
&=& F(S_2(I))n^{-1}\,.
\eeam
Here we used that $t_{j-1}=t_j$ is necessary for $\E[ Y_{i_{j}t_{j-1}} Y_{i_{j}t_{j}} ]$ to be non-zero. If $R_1>1$ we can apply the above argument iteratively to obtain $F(I)=F(S_2 \circ \cdots \circ S_2(I))n^{-R_1}$. The proof is complete.
\end{proof}
Define for $k\ge 1$, a function $g$ by $g(\emptyset)=1$ and
\begin{equation}\label{eq:defmaxg}
g(I)= \max_{T \in \{1,\ldots,k\}^k}\{ |T|\,:\, f(I,T)>0  \}\,,\qquad I \in \{1,\ldots,k \}^k\,.
\end{equation}
From now, on we assume the $I$-paths to be canonical.
\begin{lemma}\label{lem:9.5}
Let $I$ be a canonical $r$-path of length $k$.  
For any $T \in \{1,\ldots,k \}^k$ such that $f(I,T)>0$ we have $|T|\le k-r+1$. 
\end{lemma}
\begin{proof}
%Let $I=(i_1,\ldots,i_k)$ be a canonical $r$-path of length $k$, and $T= (t_1,\ldots,t_k)\in \{1,\ldots,k\}^k$. 
Without loss of generality we may assume that $T$ is canonical. 
We shall sometimes refer to the $t_i$'s as $t$-indices. 
In the beginning one should think of the $t$-indices as pairwise distinct whenever possible. 
Their actual values are not relevant for the value $f(I,T)$. In all cases, except $r=1$, there are certain $t$-indices that have to coincide such that $f(I,T)$ can be positive: $t_i=t_j$ for some $i,j$, $i\neq j$. This is due to the symmetry of $X$.
% and taken into account in Step 2 of the Path-Shortening Algorithm and the part of the proof of Proposition~\ref{prop:PSI} which is concerned with Step 2. 
We will see that in some cases these $i,j$ are not unique. More precisely, it may happen that there is a set $\{ t_{i_1}, t_{i_2}, \ldots,t_{i_{2a}}\}$ with $i_1, \ldots,i_{2a}$ distinct such that $|\{ t_{i_1}, t_{i_2}, \ldots,t_{i_{2a}}\}|\le a$ is necessary for $f(I,T)>0$. In these cases, the cardinality of
$T$ is less than $k$ provided $f(I,T)>0$.  
\par
We start with the two simplest cases. If $r=1$, we have $f(I,T)>0$ for any $T$, and 
hence $|T|\le k = k-r+1$. Moreover, if $r=k$, we have $f(I,T)>0$ if and only if $t_1=\cdots=t_k$, and hence $|T|=1 = k-r+1$.
\par
Now we assume $1<r<k$.
Our arguments will rely on the proof of Lemma~\ref{prop:PSI}.
Clearly, $1\le g(I)\le k$. From the definition of $\runs$ in the PSA we have $\runs(I)\le k-r$. From \eqref{eq:proofstep1} and \eqref{eq:proofstep2} one infers 
\begin{equation}\label{eq:sglmsoe}
\runs(I)+1\le g(I)\le k-R(I)\,.
\end{equation}
First, we analyze paths $I$ with $S(I)=\emptyset$, or equivalently $\length(S(I))=0$. This implies $\runs(I)=k-r$; otherwise the path-shortening function stops earlier and $S(I)\neq \emptyset$. Therefore, we get from the identity
\begin{equation}\label{eq:identityg}
g(I)=g(S(I)) + \runs(I)\,
\end{equation}
that $g(I)= k-r+1$, which finishes the proof in the case $S(I)=\emptyset$. 
Note that \eqref{eq:identityg} holds for all $I$. This follows from the proof of Lemma~\ref{prop:PSI}.
\par

Next, assume $S(I)\neq \emptyset$. In this case, we immediately see
\begin{equation}\label{eq:lengthsi}
\length(S(I))=k-R(I)-\runs(I)\,.
\end{equation}
Since each element in $S(I)$ has to appear at least twice and $r\ge 2$ we have $\length(S(I))\ge 4$. Moreover, $S(I)$ has $r-R(I)\ge 2$ distinct components.  
As a consequence, it must hold
\begin{equation}\label{eq:boundri}
\runs(I) \le k-r- (r-R(I)) =k-2r+R(I)\,.
\end{equation}
In view of \eqref{eq:identityg}, we have to bound $g(S(I))$. 
Without loss of generality $S(I)$ may be assumed canonical: there is exactly 
one canonical path in the isomorphism class of $S(I)$ and every path in an isomorphism class has the same $g$-value.
If, for example, $S(I)$ happens to be $(3,4,3,4)$, then we will work with the canonical representative $(1,2,1,2)$.
Write $S(I)= (s_1, \ldots, s_{\length(S(I))})$. Since $S(I)$ is canonical, we have
\begin{equation*}
\{s_1, \ldots, s_{\length(S(I))}\} = \{1,\ldots,r-R(I)\}\,.
\end{equation*}
For $ i=1,\ldots,r-R(I)$ define $N_i:= |\{1\le j\le \length(S(I)) : s_j=i\}|$. If we now let $L_i$ be the set of all $u$ such that $(i,t_u)$ appears as an index in $f(S(I),T)$, then $|L_i|=2 N_i$. Finally, define
\begin{equation*}
T_i:= \{t_j\,:\, j\in L_i \} \quad \text{ and } \quad \widetilde{T}_i:= (t_j\,:\, j\in L_i )\,, \quad i=1,\ldots,r-R(I)\,.
\end{equation*}
For example, if $I=(1,2,1,2,3,3)$, then $k=6,r=3$, $N_1=N_2=2$ and we have 
$(S(I),R(I),\runs(I))=((1,2,1,2),1,1)$ and $L_1=L_2=\{1,2,3,4\}$.
\par
By construction, $f(S(I),T)$ can only be positive if $|T_i| \le N_i$. More precisely every $t$-index in the vector $\widetilde{T}_i$ needs to coincide with at least $1$ other $t$-index of this vector. Otherwise,
$\E\big[\prod_{u\in L_i} Y_{i,t_u} \big]=0$
which would imply $f(S(I),T)=0$. The quantity $g(S(I))$ is the maximum number of distinct $t$-indices such that $f(S(I),T)>0$. Hence, there can be at most $\length(\widetilde{T}_i)/2$ distinct $t$-indices in $\widetilde{T}_i$. Since each $t_j$ appears exactly twice in $(\widetilde{T}_1, \ldots, \widetilde{T}_{r-R(I)})$, 
\begin{equation}\label{eq:gsings}
g(S(I))\le 0.5\,\length(S(I))\,.
\end{equation}
%A detailed argument to prove $\eqref{eq:gsings}$ is the following:  By setting two $t$-indices equal you can create at most two new pairs in all the vectors $\widetilde{T}_i, i=1,\ldots,r-R(I)$ combined. At the start $|T_i|+\cdots+|T_{r-R(I)}|=2 \length(S(I))$. After setting one pair of $t$-indices equal we have
%\begin{equation}\label{eq:gdrstgdrt}
%|T_i|+\cdots+|T_{r-R(I)}|\ge 2 \length(S(I)) -2\,.
%\end{equation}
%After setting $v$ pairs of $t$-indices equal we have $|T_i|+\cdots+|T_{r-R(I)}|\ge 2 \length(S(I)) -2 v$. Remember that we have to achieve $|T_i|+\cdots+|T_{r-R(I)}|\le  \length(S(I))$, for which $v\ge \length(S(I))/2$  is a necessary condition. Finally, $g(S(I))=\length(S(I))-v(I)$, where $v(I)\ge \length(S(I))/2$. This proofs $\eqref{eq:gsings}$.
\par
Now we are ready to finish the proof of the lemma.
By \eqref{eq:identityg}, \eqref{eq:gsings}, \eqref{eq:lengthsi}, \eqref{eq:boundri},  
in this order, one obtains
\begin{equation*}
\begin{split}
g(I)&= g(S(I))+ \runs(I) \le \frac{\length(S(I)}{2}+ \runs(I)\\
&= \frac{k-R(I)-\runs(I)}{2}+ \runs(I)\\
&\le \frac{k-R(I)+k-2r+R(I)}{2} =k-r\,.
\end{split}
\end{equation*}
\end{proof}
\begin{remark}\label{re:notunique}{\em 
The above proof reveals that $g(I)=k-r+1$ if and only if $S(I)= \emptyset$. For $r$-paths $I$ of length $k$ with $S(I)\neq \emptyset$, the bound $g(I)\le k-r$ is sharp. Consider for instance $I=(1,2,1,2)$, where 
\begin{equation*}
f(I,T)=(\E[ Y_{t_1}Y_{t_2}Y_{t_3}Y_{t_4} ])^2\,.
\end{equation*}
From this relation it is easily deduced that the only canonical representatives $T=(t_1,t_2,t_3,t_4)$ leading to $f(I,T)>0$ are $(1,1,2,2),(1,2,1,2),(1,2,2,1)$ and $(1,1,1,1)$. The first three of them have the highest number of distinct values. We conclude $g(I)=2$. In general, the canonical paths $T$ for which the maximum in \eqref{eq:defmaxg} is attained are not unique, whenever $g(I)\le k-r$. On the other hand, if $g(I)= k-r+1$ there exists exactly one canonical $(k-r+1)$-path $T$ of length $k$ for which the maximum is obtained. This is an immediate consequence of the above proofs. In \cite{bai:silverstein:2010}, Bai and Silverstein present a way to describe this $T$. 
}\end{remark}

For a canonical $r$-path $I$ of length $k$ let 
\begin{equation}\label{eq:defdr}
d(I)= k-r+1-g(I)\,.
\end{equation}
The function $d$ satisfies $0\le d(I) \le k-r$ and $d(S(I))=d(I)$.
The set of canonical $r$-paths of length $k$, denoted by $\mathcal{I}_{r,k}$, can be written as a disjoint union
\begin{equation*}
\mathcal{I}_{r,k}= \bigcup_{u=0}^{k-r} \mathcal{I}_{r,k}(u)\,,
\end{equation*} 
where $\mathcal{I}_{r,k}(u)$ contains those $I$ with $d(I)=u$.
\par
Lemma 3.4 in \cite{bai:silverstein:2010} determines the cardinality of $\mathcal{I}_{r,k}(0)$:
for $k\in \N$ and $r\le k$, 
\beam\label{lem:lemma3.4}
|\mathcal{I}_{r,k}(0)| = \frac{1}{r} \binom{k}{r-1} \binom{k-1}{r-1}\,.
\eeam
%\begin{remark}{\em 
%We already know $|\bigcup_{u=1}^{k-r} \mathcal{I}_{r,k}(u)|>0$, which contradicts the claim in \cite[p.~42]{bai:silverstein:2010} that ``in this category, $r+s=k$''.
%}\end{remark}
\begin{proposition}\label{prop:paths} Assume condition \eqref{eq:assumptionq}. Then the following statements hold for any 
$r$-path $I$ of length $k\ge 1$ and $1\le r\le k$: 
\begin{itemize}
\item[(1)] If $S(I)=\emptyset$, then $F(I)= n^{1-r}$.
\item[(2)] In general, we have 
\begin{equation}\label{eq:proppaths}
F(I)\le 2 \, n^{1-r-d(I)} (2k)^{d(I)} q^{d(I)}\,.
\end{equation}
\end{itemize}
\end{proposition}
\begin{proof}
$S(I)=\emptyset$ is equivalent to $R(I)=r$. By Lemma~\ref{prop:PSI}, 
\begin{equation*}
F(I)=F(S(I)) n^{-R(I)} =F(\emptyset)n^{-r}=n^{1-r}\,.
\end{equation*}
Therefore, we only have to prove \eqref{eq:proppaths} for paths $I$ with $d(I)\ge 1$. Without loss of generality we assume $S(I)$ is a canonical $(r-R(I))$-path. We use the notation for paths with $S(I)\neq \emptyset$ developed in the proof of Lemma~\ref{lem:9.5}. We know that
\begin{equation*}
S(I)= (\pi_1, \ldots, \pi_{\length(S(I))})\,,
\end{equation*}
where $\pi_1, \ldots, \pi_{\length(S(I))}$ is a permutation of the path
\begin{equation*}
I_0 = (\underbrace{1,\ldots,1}_{N_1},\underbrace{2,\ldots,2}_{N_2}, \ldots, \underbrace{r-R(I),\ldots,r-R(I)}_{N_{r-R(I)}})\,.
\end{equation*}
Clearly, $I_0 \in \mathcal{I}_{r-R(I),\length(S(I))}(0)= \mathcal{I}_{r-R(I),k-R(I)-\runs(I)}(0)$. By Lemma~\ref{lem:9.5}, 
\begin{equation*}
g(I_0)=(k-R(I)-\runs(I))-(r-R(I))+1 = k-r-\runs(I)+1
\end{equation*}
and by definition of the function $d(\cdot)$,
\begin{equation*}
g(S(I))=k-r-\runs(I)+1 -d(S(I)=k-r-\runs(I)+1-d(I)\,.
\end{equation*}
\par
The main idea will be to compare $F(S(I))$ to $F(I_0)$. Both of them are sums of expressions of the type
\begin{equation}\label{eq:type}
\prod_{i=1}^{r-R(I)} \E\Big[Y_{i1}^{2m_{i,1}}\cdots Y_{is_i}^{2m_{i,s_i}} \Big] = \prod_{i=1}^{r-R(I)} 
\E\Big[Y_{1}^{2m_{i,1}}\cdots Y_{s_i}^{2m_{i,s_i}} \Big]\,, 
\end{equation}
%\begin{equation}\label{eq:type}
%\prod_{i=1}^{r-R(I)} \E\Big[ \frac{X_{i1}^{2m_{i,1}}\cdots X_{is_i}^{2m_{i,s_i}}]}{D_i^{N_i}} \Big] = \prod_{i=1}^{r-R(I)} \E\Big[ \frac{X_{11}^{2m_{i,1}}\cdots X_{1s_i}^{2m_{i,s_i}}]}{D_1^{N_i}} \Big]\,, 
%\end{equation}
where for all $i=1,\ldots,r-R(I)$, $1\le s_i\le N_i$, $m_{i,j}\ge 1$ for all $j\ge 1$ and $m_{i,1}+\cdots +m_{i,s_i}=N_i$. We write 
\begin{equation}\label{eq:sdfesofj}
\mathbf{s}=(s_1,\ldots,s_{r-R(I)}) \quad \text{ and } \quad \mathbf{m}_i=(m_{i,1},\ldots,m_{i,s_i})\,,\quad i=1,\ldots,r-R(I)\,.
\end{equation}
Observe that in 
\begin{equation*}
\begin{split}
F(I_0) &= \sum_{t_1,\ldots,t_{N_1+\cdots+N_{r-R(I)}}=1}^n \E\Big[ Y_{t_{N_1+\cdots+N_{r-R(I)}}} Y_{t_1}^2 
\cdots Y_{t_{N_1-1}}^2Y_{t_{N_1}} \Big] \cdots \\
%\E\Big[ \frac{X_{2,t_{N_1}} X_{2,t_{N_1+1}}^2 \cdots X_{2,t_{N_1+N_2-1}}^2X_{2,t_{N_1+N_2}}}{D_2^{N_2}} \Big]\\ 
& \E\Big[Y_{t_{N_1+\cdots+N_{r-R(I)-1}}} Y_{t_{N_1+\cdots+N_{r-R(I)-1}}+1}^2 \cdots Y_{t_{N_1+\cdots+N_{r-R(I)}}-1}^2Y_{,t_{N_1+\cdots+N_{r-R(I)}}} \Big]
\end{split}
\end{equation*}
the non-zero summands have to satisfy $t_{N_1}=t_{N_2}= \cdots =t_{N_1+\cdots+N_{r-R(I)}}$.  Hence, the above sum is effectively a 
sum only over $g(I_0)$ $t$-indices. The point we want to stress is that there is never a choice, in the sense that even though there are $g(I_0)$ distinct $t$-indices, something like $t_{N_1}=t_1\neq t_2=t_{N_1+N_2}$ is never possible. The reason is that the associated canonical $g(I_0)$-path for the $t$-indices is unique.

For $S(I)\neq \emptyset$, however, the associated canonical $g(S(I))$-path for the $t$-indices is not unique, as mentioned in Remark~\ref{re:notunique}. Depending on the sets $L_i$ there are several possibilities. For instance, for $S(I)=(1,2,1,2)$ we have $L_1=L_2=\{1,2,3,4\}$, $d(I)=1$ and $\length(S(I))=4$. To produce a positive summand one needs $|\{ t_1,t_2,t_3,t_4\}|\le 2$ with every $t$-index appearing at least twice. In this case, $t_1$ has to take the same value as one of the other three $t$-indices. Then there are two $t$-indices left 
which all have to appear at least twice. In this specific example, there are three  canonical paths of $t$-indices 
which are listed in Remark~\ref{re:notunique} above.

We are interested in the general case. How many distinct canonical $g(S(I))$-paths $T$ of length $\length(S(I))$ with $f(S(I),T)>0$ can exist? With the reasoning which lead to \eqref{eq:gsings} one can show that this number is at most $(2d(I)+1)!!$. This bound is attained if $N_1=N_2=\length(S(I))/2$ which implies $L_1=L_2=\{1,\ldots,\length(S(I)\}$. %Then there is equality in \eqref{eq:gdrstgdrt}.

For our purpose we will use a much larger bound, namely
\begin{equation}\label{eq:nnnnnnnnn}
(2d(I)+1)!! = (2d(I)+1) (2d(I)-1)\cdots 3 \le (2d(I)+1)^{d(I)} \le (2k)^{d(I)}\,.
\end{equation}

Now let us compare $F(S(I))$ and $F(I_0)$, which look very similar at first sight. The main difference is the dimension of the index sets in the summation. While the sum for $F(I_0)$ contains $n^{g(I_0)}$ positive elements, the sum for $F(S(I))$ has at most $(2k)^{d(I)} n^{g(I_0)-d(I)}$ positive elements. Let $Q_{S(I)}$ denote the set of all canonical $T$ for which the maximum in \eqref{eq:defmaxg} is attained. By the above considerations, $| Q_{S(I)}| \le (2k)^{d(I)}$. Each element in $Q_{S(I)}$ corresponds to a different configuration of $t$-indices in $F(S(I))$, i.e., it tells us which $t$-indices have to be equal. Therefore, we have
\begin{equation}\label{eq:lkhdtoh}
F(S(I))\le \sum_{Q\in Q_{S(I)}} F_Q(S(I))\,,
\end{equation} 
where $F_Q(S(I))$ is defined as follows. Write $Q=(q_1,\ldots, q_{\length(S(I))})$. By construction,\\
 $\{ q_1,\ldots, q_{\length(S(I))}\} = \{1,\ldots, g(S(I))\}$. Set $K_j=\{ 1\le i\le \length(S(I)) :  q_i=j\}$. Then 
\begin{equation}\label{eq:FQ}
F_Q(S(I))= \mathop{\sum_{t_1,\ldots,t_{\length(S(I))}=1}}_{t_l=t_m \, \forall l,m \in K_j\,, 1\le j\le g(S(I))}^n f(S(I),(t_1, \ldots,t_{\length(S(I))}))\,.
\end{equation}
We will show later that 
\begin{equation}\label{eq:assmomentnow}
F_Q(S(I)) \le 2\, q^{d(I)} n^{-d(I)} F(I_0)\,, \quad Q\in Q_{S(I)}\,.
\end{equation}
Then it follows from \eqref{eq:lkhdtoh} and \eqref{eq:assmomentnow} that
\begin{equation*}
\begin{split}
F(S(I))&\le \sum_{Q\in Q_{S(I)}} F_Q(S(I))\\
&\le (2k)^{d(I)} 2\, q^{d(I)} n^{-d(I)} F(I_0) = 2\, (2k)^{d(I)} q^{d(I)} n^{-d(I)} n^{1-r+R(I)}\,.
\end{split}
\end{equation*}
Finally, an application of Lemma~\ref{prop:PSI} gives 
\begin{equation*}
F(I) =n^{-R(I)}F(S(I))\le 2\, (2k)^{d(I)} q^{d(I)} n^{-d(I)} n^{1-r}\,,
\end{equation*}
which completes the proof. 
\par
Next, we show \eqref{eq:assmomentnow} by matching each of the $n^{g(I_0)-d(I)}$ positive summands in \eqref{eq:FQ} with $n^{d(I)}$ of the $n^{g(I_0)}$ positive summands in $F(I_0)$, where we recall that
\begin{equation}\label{eq:FI0}
F(I_0)= \mathop{\sum_{t_1,\ldots,t_{\length(S(I))}=1}}_{t_{N_1}=t_{N_2}= \cdots =t_{N_1+\cdots+N_{r-R(I)}}}^n f(I_0,(t_1, \ldots,t_{\length(S(I))}))\,.
\end{equation}
\par
By {\em matching} we mean the following. Assume we want to prove 
\begin{equation}\label{eq:stggsrbjklfg}
\sum_{i=1}^n A_i \le  \sum_{j=1}^m B_j
\end{equation}
for nonnegative $A_i,B_i$ and $m\ge n$. If for every $i=1,\ldots,n$ there exists a $j_i\in \{1,\ldots, m\}$ such that $A_i\le B_{j_i}$ and the $j_i$'s are distinct, then \eqref{eq:stggsrbjklfg} holds. In this case, we say that each $A_i$ is matched by some $B_{j_i}$.
\par
We say that $f(S(I),(t_1, \ldots,t_{\length(S(I))}))$ and  $f(I_0,(t_1, \ldots,t_{\length(S(I))}))$ are in class $y=\sum_{i=1}^{r-R(I)}s_i$ if they can be written in the form 
\begin{equation}\label{eq:type11}
\prod_{i=1}^{r-R(I)} \E\big[ Y_{1}^{2m_{i,1}}\cdots Y_{s_i}^{2m_{i,s_i}} \big]\,.
\end{equation}
By construction, $y$ takes values in the set $\{ r-R(I), \ldots,\length(S(I)) \}$. A summand in class $y$ is fully determined by the vector $(\mathbf{s},\mathbf{m}_1,\ldots, \mathbf{m}_{r-R(I)})=:(\mathbf{s},\mathbf{m})$; see \eqref{eq:sdfesofj} for this notation.
Hence, we call this summand of type $(\mathbf{s},\mathbf{m})$ and denote it $f_{\mathbf{s},\mathbf{m}}$. Note that the class $y$ is comprised of all elements of type $(\mathbf{s},\mathbf{m})$ such that $\sum_{i=1}^{r-R(I)}s_i=y$ and $\mathbf{m}$ satisfies the restriction stated below equation \eqref{eq:type}.

Let $\mathcal{T}_0(y)$ and $\mathcal{T}_Q(y)$ be index sets which contain the exact type of all  summands (counted with multiplicity) of class $y$ in \eqref{eq:FI0} and \eqref{eq:FQ}, respectively. 
As mentioned before, we must have
\begin{equation*}
\sum_{y=r-R(I)}^{\length(S(I))} |\mathcal{T}_0(y)|=n^{g(I_0)} \quad \text{ and } \quad \sum_{y=r-R(I)}^{\length(S(I))} |\mathcal{T}_Q(y)|=n^{g(I_0)-d(I)}\,.
\end{equation*}
With this notation we can write
\beam\label{eq:rewrfi0}
2 \, F(I_0)&=&2\, \sum_{y=r-R(I)}^{\length(S(I))} \sum_{(\mathbf{s},\mathbf{m})\in \mathcal{T}_0(y)} f_{\mathbf{s},\mathbf{m}}\,,\\\label{eq:rewrfq}
n^{d(I)}\, F_Q(S(I))&=&n^{d(I)}\, \sum_{y=r-R(I)}^{\length(S(I))} \sum_{(\mathbf{s},\mathbf{m})\in \mathcal{T}_Q(y)} f_{\mathbf{s},\mathbf{m}}\,.
\eeam
We show \eqref{eq:assmomentnow} by a {\em matching argument}. 
We start by matching summands in class $\length(S(I))$. From \eqref{eq:type11} we see that elements of class $\length(S(I))$ are necessarily of the form 
\begin{equation*}
\prod_{i=1}^{r-R(I)} \E\big[ Y_{1}^{2}\cdots Y_{N_i}^{2} \big]\,, 
\end{equation*}
in other words they are all equal. Note that 
\begin{equation*}
\begin{split}
|\mathcal{T}_0(\length(S(I)))|&= n (n-1) \cdots (n-g(I_0)+1)\,,\\
|\mathcal{T}_Q(\length(S(I)))|&= n (n-1) \cdots (n-g(I_0)+d(I)+1)\,.
\end{split}
\end{equation*}
Therefore, we have 
\begin{equation*}
n^{d(I)}\, \sum_{(\mathbf{s},\mathbf{m})\in \mathcal{T}_Q(\length(S(I)))} f_{\mathbf{s},\mathbf{m}} \le 2\, \sum_{(\mathbf{s},\mathbf{m})\in \mathcal{T}_0(\length(S(I)))} f_{\mathbf{s},\mathbf{m}}\,.
\end{equation*}
This shows that for each summand in class $\length(S(I))$ on the \lhs~ of \eqref{eq:assmomentnow} we can find at least one summand of the same type on the  \rhs~ of \eqref{eq:assmomentnow}.

Since a large number of summands of the class $\length(S(I))$ have not been used for matching of summands from the same class, we can use \eqref{eq:assumptionq} to match them with summands of classes $\length(S(I))-1, \length(S(I))-2, \ldots, \max(\length(S(I))-d(I),r-R(I))$.

Applying \eqref{eq:assumptionq} to a summand of class $\length(S(I))-1$ we obtain that it is bounded by $q$ times a summand in class $\length(S(I))$. Hence, we can perform the matching in \eqref{eq:assmomentnow} also between different classes. Clearly, for $y\in\{ r-R(I)+1,\ldots,\length(S(I))\}$ the index set $\mathcal{T}_0(y)$ is much larger than $\mathcal{T}_0(y-1)$. In fact, we have $|\mathcal{T}_0(y)|=n\; c\;|\mathcal{T}_0(y-1)|$ for some constant $c>1$. Note that $y\mapsto |\mathcal{T}_Q(y)|$ is not a strictly increasing function since some $\mathcal{T}_Q(y)$ can be empty.

The matching is performed as follows: first match the class $\length(S(I))$ summands on the \lhs~ of \eqref{eq:assmomentnow}. Then match the class $\length(S(I))-1$ summands on the \lhs~ of \eqref{eq:assmomentnow} with the remaining class $\length(S(I))$ summands on the \rhs~which have not been used for the matching yet.

Let $r-R(I)\le u\le \length(S(I))$. The general strategy is to match class $u$ summands on the \lhs~ with class $u,\ldots,\min(u+d(I),\length(S(I)))$ summands on the \rhs. During the matching one tries to use the (still available) class $\min(u+d(I),\length(S(I)))$ summands on the \rhs~ first, then turns to class $\min(u+d(I),\length(S(I)))-1$, and so forth. Whenever a matching between different classes is performed an application of \eqref{eq:assumptionq} is necessary to ensure that the expression on the \lhs~ is bounded by whatever we have matched it with on the \rhs. This leads to powers of $q$ and since $q^{d(I)}$ is the highest possible power we have explained the factor $q^{d(I)}$ in  \eqref{eq:assmomentnow}.
\par
Note that the factor $2$ in  \eqref{eq:assmomentnow} is there to guarantee 
\begin{equation*}
|\mathcal{T}_Q(\length(S(I)))|< 2\, n^{-d(I)} \, |\mathcal{T}_0(\length(S(I)))|\,
\end{equation*}
for sufficiently large $n$, but it is of no central importance.

The last step in the procedure is the matching of the summands with the highest possible powers on the \lhs~of \eqref{eq:assmomentnow}, which appear when all $t$-indices are equal. They are elements of the class $r-R(I)$. We have
\begin{equation*}
|\mathcal{T}_0(r-R(I))|=|\mathcal{T}_Q(r-R(I))|=n\,,
\end{equation*}
which is a simple explanation why matching of \eqref{eq:rewrfi0} and \eqref{eq:rewrfq} with summands in the same class cannot work in general. Using \eqref{eq:assumptionq} $d(I)$ times, we can bound class $r-R(I)$ summands by class $r-R(I)+d(I)$ summands of which we originally have $|\mathcal{T}_0(r-R(I))|\approx n^{d(I)}|\mathcal{T}_0(r-R(I))|$, which explains the factor $n^{d(I)}$ in \eqref{eq:assmomentnow}. The general matching strategy applies and the proof of \eqref{eq:assmomentnow} is complete.
\end{proof}

\section{Proof of Theorem~\ref{thm:mu1convergence}}\label{sec:5.2}
The following proposition contains our main technical novelty. Its proof is given after the proof of Theorem~\ref{thm:mu1convergence}.
\begin{proposition}\label{thm:limsup}
Assume \eqref{Cpn} and that the  iid symmetric field $(X_{it})$ satisfies  \eqref{eq:assumptionq}.
Then the following limit results hold for the largest and smallest eigenvalues $\mu_{(1)}$ and $\mu_{(p)}$ of $\bfR$:
\beam\label{eq:dndfnseg}
\limsup_{\nto} \mu_{(1)} &\le& (1+\sqrt{\gamma})^2 \, \quad \as\\\label{eq:dndfnseg1}
\liminf_{\nto} \mu_{(p)} &\ge &(1-\sqrt{\gamma})^2 \, \quad \as
\eeam
\end{proposition}
\begin{proof}[Proof of Theorem~\ref{thm:mu1convergence}]
(1) If $\E[X^4]<\infty$, \eqref{eq:mlmgd} and \eqref{eq:limitmup} hold for any mean zero distribution as seen in \eqref{eq:drtgdrghdr1}.
\par
\noindent
(2) Now assume \eqref{eq:assumptionq}.
The convergence of $F_{\bfR}$ to a deterministic distribution 
supported on a compact interval implies that the number of the
eigenvalues outside this interval is $o(p)$. Since the right and left endpoints of the \MP law are 
$(1+\sqrt{\gamma})^2$ and $(1-\sqrt{\gamma})^2$, respectively, we conclude from Theorem~\ref{thm:mpcorrelation}(1) that 
\begin{equation*}
\liminf_{n\to \infty}\mu_{(1)} \ge (1+\sqrt{\gamma})^2\quad \as \quad\mbox{and}\quad
\limsup_{n\to \infty}\mu_{(p)} \le (1-\sqrt{\gamma})^2\quad \as \,;
\end{equation*}
see \cite{bai:silverstein:2010} for details. Together with Proposition~\ref{thm:limsup} this completes the proof.
\end{proof}

\subsection*{Proof of equation \eqref{eq:dndfnseg} in Proposition~\ref{thm:limsup}}\label{sec:5.4}

Following \cite{geman}, we prove \eqref{eq:dndfnseg} by showing 
\begin{equation}\label{eq:showsumexp1}
\sum_{n=1}^\infty \E\Big[\Big(\frac{\mu_{(1)}}{z}\Big)^k  \Big] <\infty\,,
\end{equation}
where $z>(1+\sqrt{\gamma})^2$ and $k=k_n$ satisfies $k/\log n \to \infty$ and $(k^3 q)/n \to 0$, which exists by condition \eqref{eq:assumptionq}. 
We use that $\E[\mu_{(1)}^k] \le \E[\tr(\bfR)^k]$ and 
\begin{equation*}
\begin{split}
\E[\tr(\bfR)^k]&= \sum_{i_1,\ldots,i_k=1}^p \sum_{t_1,\ldots,t_k=1}^n \E[ Y_{i_1t_k} Y_{i_1t_1} Y_{i_2t_1}Y_{i_2t_2}Y_{i_3t_2} Y_{i_3t_3} \cdots Y_{i_kt_{k-1}} Y_{i_kt_k}  ]\\
&= \sum_{i_1,\ldots,i_k=1}^p F(i_1,\ldots,i_k)\, .
\end{split}
\end{equation*}
We rewrite $\E[\tr(\bfR)^k]$ by sorting according to the number of distinct components in the path $(i_1, \ldots, i_k)$. Any $r$-path of length $k$ is an element in the disjoint union $\mathcal{J}_{r,k}(0) \cup \cdots \cup \mathcal{J}_{r,k}(k-r)$, where $\mathcal{J}_{r,k}(u)$ is the set of all $r$-paths $I$ of length $k$ with $d(I)=u$; see \eqref{eq:defdr} for the definition of $d(I)$. Hence we have
\begin{equation}\label{eq:disjointkds}
\{1,\ldots, p\}^k = \bigcup_{r=1}^k \bigcup_{u=0}^{k-r} \mathcal{J}_{r,k}(u)\,.
\end{equation}

Given a path $I\in \mathcal{J}_{r,k}(u)$ we can look at the positions where the $r$ distinct components appear for the first time. There are $r$ such positions. The first such position is always $1$, in general $i_1$ can take $p$ different values. For the second such position there are $(p-1)$ possibilities; the original $p$ minus the one from the first position. In total there are $p(p-1)\cdots (p-r+1)$ ways to assign values to these $r$ positions. For this reason
\begin{equation}\label{eq:slktej}
|\mathcal{J}_{r,k}(u)| = p(p-1)\cdots (p-r+1) |\mathcal{I}_{r,k}(u)|\,,
\end{equation}
where $\mathcal{I}_{r,k}(u)$ is the set of all canonical $r$-paths $I$ of length $k$ with $d(I)=u$. The only difference between the definitions of $\mathcal{J}_{r,k}(u)$ and $\mathcal{I}_{r,k}(u)$ is that the elements of the latter are canonical. 
Note that $\mathcal{I}_{k,k}(u)=\emptyset$ for all $u\ge 1$.
\par
In view of \eqref{eq:disjointkds} and \eqref{eq:slktej} we obtain
\beam\label{eq:oklj}
\E[\tr(\bfR)^k]&=& \sum_{r=1}^k \sum_{u=0}^{k-r} \sum_{I\in \mathcal{J}_{r,k}(u)} F(I)\nonumber\\
&=& \sum_{r=1}^k p(p-1)\cdots (p-r+1) \sum_{u=0}^{k-r} \sum_{I\in \mathcal{I}_{r,k}(u)} F(I)\label{eq:xx}\\
&\le& \sum_{r=1}^k p^r  \sum_{I\in \mathcal{I}_{r,k}(0)} F(I)
+\sum_{r=1}^{k-1} p^r \sum_{u=1}^{k-r} |\mathcal{I}_{r,k}(u)| \max_{I\in \mathcal{I}_{r,k}(u)}F(I)=:S_1+S_2\,.\nonumber
\eeam
By Proposition~\ref{prop:paths}, \eqref{lem:lemma3.4} and since $|\mathcal{I}_{r,k}(0)|\le \binom{k-1}{r-1}^2$, we have
\begin{equation}\label{eq:sumS1}
S_1 \le \sum_{r=1}^k p^r \binom{k-1}{r-1}^2 n^{1-r} = p \sum_{r=1}^k \binom{k-1}{r-1}^2 \Big(\frac{p}{n}\Big)^{r-1}\,.
\end{equation}

Next we bound $S_2$. Consider $1\le u\le k-r$. We will see how elements of $\mathcal{I}_{r,k}(u)$ can be constructed by modifying elements of $\mathcal{I}_{r,k}(0)$. Let $I\in \mathcal{I}_{r,k}(u)(N_1, \ldots, N_r)$ be the subset of $\mathcal{I}_{r,k}(u)$ for whose elements the integer $i$ appears exactly $N_i$ times as a component. Here $N_i, i=1,\ldots,r$ are positive integers satisfying $N_1+\cdots+N_r=k$. Obviously it is possible to obtain $I$ by permuting the components of any $I_0 \in \mathcal{I}_{r,k}(0)(N_1, \ldots, N_r)$. Consider the following permutation of $I_0$: two components of $I_0$ exchange places, all other remain untouched. We denote such a {\em switching permutation} by $SP$. The number of such permutations is bounded by $k^2/2$. Indeed, the first component can switch places with the remaining $k-1$ components, the second with $k-2$ components, etc. In total there are 
\begin{equation*}
(k-1)+(k-2)+ \cdots+ 1 = \sum_{j=1}^{k-1} j = \frac{(k-1)k}{2} \le \frac{k^2}{2}
\end{equation*}
ways how two components can switch positions.

Let $u=1$. For any $I\in \mathcal{I}_{r,k}(u)(N_1, \ldots, N_r)$ there exists at least one path $I_0 \in \mathcal{I}_{r,k}(0)(N_1, \ldots, N_r)$ and a switching permutation $SP$ such that $I= SP(I_0)$. Here $SP$ and $I_0$ are in general not unique. This is a consequence of the proof of Lemma~\ref{lem:9.5}. This implies
\beao
|\mathcal{I}_{r,k}(u)| \le |\mathcal{I}_{r,k}(0)| \frac{k^2}{2}\,.
\eeao
Similarly, for $1\le u\le k-r$ and $I\in \mathcal{I}_{r,k}(u)(N_1, \ldots, N_r)$ there exist $I_0 \in \mathcal{I}_{r,k}(0)(N_1, \ldots, N_r)$ and switching permutations $SP_1,\ldots, SP_u$ such that
$
I= SP_1 \circ \cdots \circ SP_u (I_0)\,,
$
which shows
\begin{equation}\label{eq:lsetseet}
|\mathcal{I}_{r,k}(u)| \le |\mathcal{I}_{r,k}(0)| \Big(\frac{k^2}{2}\Big)^u\,.
\end{equation}
\par
Now we are ready to bound $S_2$. From Proposition~\ref{prop:paths} we get 
\begin{equation*}
\max_{I\in \mathcal{I}_{r,k}(u)}F(I) \le 2 n^{1-r-u}(2k)^u q^u 
\end{equation*}
and therefore,
\begin{equation*}
\begin{split}
S_2&\le \sum_{r=1}^{k-1} p^r \sum_{u=1}^{k-r} \binom{k-1}{r-1}^2 \Big(\frac{k^2}{2}\Big)^u 2 n^{1-r-u}(2k)^u q^u\\
&= p \sum_{r=1}^{k-1} \binom{k-1}{r-1}^2 \Big(\frac{p}{n}\Big)^{r-1} 2 \sum_{u=1}^{k-r} \Big( \frac{k^3 q}{n} \Big)^u\\
&\le p \sum_{r=1}^{k-1} \binom{k-1}{r-1}^2 \Big(\frac{p}{n}\Big)^{r-1} 2 \Big[\Big(1- \frac{k^3 q}{n} \Big)^{-1}-1\Big]\,.
\end{split}
\end{equation*}
Finally, we have the bound
\begin{equation*}
\begin{split}
\E[\tr(\bfR)^k]&\le S_1+S_2\le p \sum_{r=1}^k \binom{k-1}{r-1}^2 \Big(\frac{p}{n}\Big)^{r-1} \Big( 1+ 2 \Big[\Big(1- \frac{k^3 q}{n} \Big)^{-1}-1\Big] \1_{\{r<k\}} \Big)\\
&\le p \sum_{r=1}^k \binom{2k-2}{2r-2} \Big(\frac{p}{n}\Big)^{r-1} \Big( 1+ 2 \Big[\Big(1- \frac{k^3 q}{n} \Big)^{-1}-1\Big]\1_{\{r<k\}}\Big)\\
&\le p\sum_{r=0}^{2k-2} \binom{2k-2}{r} \Big(\sqrt{\frac{p}{n}}\Big)^{r} \Big( 2 \Big(1- \frac{k^3 q}{n} \Big)^{-1}-1 \Big)^{2k-2-r}\\
&= \Big[p^{1/(k-1)} \Big( 2 \Big(1- \frac{k^3 q}{n} \Big)^{-1}-1 +\sqrt{\frac{p}{n}}\Big)^2\Big]^{k-1} \le \eta^k\,,
\end{split}
\end{equation*}
where $\eta$ is a constant satisfying $(1+\sqrt{\gamma})^2<\eta<z$. The last inequality follows from $p^{1/(k-1)} \to 1$ and
\begin{equation*}
\lim_{\nto} \Big( 2 \Big(1- \frac{k^3 q}{n} \Big)^{-1}-1 +\sqrt{\frac{p}{n}}\Big)^2 =(1+\sqrt{\gamma})^2\,.
\end{equation*}
This shows \eqref{eq:showsumexp1} which concludes the proof.

%\begin{remark}{\em 
%Assume $X$ is symmetric. We have seen that the contribution of paths $I$ with $d(I)\ge 1$ is smaller. If we use the same bound $F(I)\le n^{1-r}$ for all canonical $r$-paths $I$, then we obtain the following bound  
%\begin{equation*}
%\E[\tr(\Y\Y')^k]\le p \sum_{r=1}^k c_{k,r} \,(p-1) \cdots (p-r+1) n^{1-r} \le p \sum_{r=1}^k c_{k,r} \, \Big(\frac{p}{n}\Big)^{r-1}\;,
%\end{equation*}
%where the coefficients $c_{k,r}$ count all the different possibilities of choosing indices $i_1, \ldots, i_k \in \{1,\ldots,r\}$ such $(i_1,\ldots,i_k)$ is a canonical $r$-path. This includes paths with duplicate pairs. By distinguishing whether $i_l \in \{i_1,\ldots, i_{l-1}\}$ it can be shown that
%\begin{equation*}
%c_{k,r}=\mathop{\sum_{t_1,\ldots,t_{k-1}=0}^1}_{t_1+ \cdots + t_{k-1}=r-1} \prod_{i:t_i=0} %\Big(1+\sum_{m=1}^{i-1} t_m\Big)\,.
%\end{equation*}
%}\end{remark}
%------------------------------------------------------------------------------------

%--------------------------------------------------------------------------------
\subsection{Proof of equation \eqref{eq:dndfnseg1} in Proposition~\ref{thm:limsup}}\label{sec:5.6}
We start with the following result.
\begin{proposition}\label{prop:unified}
Assume \eqref{Cpn}. If the iid entries $(X_{it})$ are symmetric and satisfy condition \eqref{eq:assumptionq} then
\beam\label{eq:jo1}
\limsup_{n\to \infty} \twonorm{\bfR-(1+\gamma) \bfI} \le 2 \sqrt{\gamma}\,\quad \as
\eeam
\end{proposition}

\begin{proof}
The general idea is the same as in the proof of equation \eqref{eq:dndfnseg} in Proposition~\ref{thm:limsup}: we will 
bound the spectral norm of $\bfR-(1+\gamma) \bfI$ by the trace of high powers of this matrix and then take an appropriate root.
%We observe that $\twonorm{\bfR-(1+\gamma) \bfI}=\twonorm{(\bfR-(1+\gamma) \bfI)^2}^{1/2}$ where $(\bfR-(1+\gamma) \bfI)^2$ is positive semidefinite.
To this end we choose an integer \seq\ $k=k_n\to \infty$ such that $k/\log n \to \infty$ and $(k^3 q)/n \to 0$, which exists by condition \eqref{eq:assumptionq}. 
Since the matrices $\bfR$ and $(1+\gamma) \bfI$ commute we have
\begin{equation*}
(\bfR-(1+\gamma) \bfI)^{2k}= \sum_{i=0}^{2k} \binom{2k}{i} \bfR^i (-1)^i (1+\gamma)^{2k-i} \bfI\,.
\end{equation*}
By linearity of the trace,
\begin{equation}\label{eq:commutegh}
\E[ \tr(\bfR-(1+\gamma) \bfI)^{2k}]= p\,(1+\gamma)^{2k} \Big[1 + p^{-1} \sum_{i=1}^{2k} \binom{2k}{i}  \Big(\frac{-1}{1+\gamma}\Big)^i \E[\tr \bfR^i]\Big]\,.
\end{equation}
From \eqref{eq:oklj} combined with \eqref{lem:lemma3.4} we know that for $n$ sufficiently large
\begin{equation}\label{eq:gsegkasdf1}
\E[\tr \bfR^i]\ge p \sum_{r=1}^i \frac{(p-1)(p-2)\cdots (p-r+1)}{n^{r-1}} \frac{1}{r}\binom{i}{r-1}\binom{i-1}{r-1} =p\, \beta_i(\gamma)\,(1-\delta_n)\,,
\end{equation}
where $\delta_n=O(1/n)$. Additionally, we established
\begin{equation}\label{eq:gsegkasdf2}
\E[\tr \bfR^i]\le p\, \beta_i(\gamma)\,\Big(1+\frac{2k^3 q}{n}\Big) (1+\delta_n)\,.
\end{equation}
Hence, by \eqref{eq:commutegh}, Lemma~\ref{lem:fkgamma}, and noting that $f_k$ is continuous on $\R$, and $p/n\to \gamma\in(0,1]$, 
we have for $n$ sufficiently large,
\begin{equation*}
\begin{split}
\E[ \tr(\bfR-(1+\gamma) \bfI)^{2k}]%&= p(1+\gamma)^{2k} \Big[ 1 +p^{-1} \sum_{i=1}^{2k} \binom{2k}{i}  \Big(\frac{-1}{1+\gamma}\Big)^i \E[\tr \bfR^i] \Big]\\
& = p(1+\gamma)^{2k} \Big[ 1 +\sum_{i=1}^{2k} \binom{2k}{i}  \Big(\frac{-1}{1+\gamma}\Big)^i  \,\beta_i(\gamma)\Big]
%\sum_{r=1}^i \frac{1}{r} \binom{i}{r-1} \binom{i-1}{r-1} \gamma^{r-1}\Big]\\
\big(1+O(2k^3 q_n/n)\big)\\
&= p(1+\gamma)^{2k} f_k(\gamma) \big(1+O(2k^3 q_n/n)\big)\\
&\le p\,(1+\gamma) (4\gamma)^k \big(1+O(2k^3 q_n/n)\big) < z^{2k}\,,
\end{split}
\end{equation*}
for any $z>2\sqrt{\gamma}$. The last inequality follows from
\begin{equation*}
\lim_{\nto} p^{1/(2k)} \big(1+2k^3 q_n/n \big)^{1/(2k)}(1+\gamma)^{1/(2k)} 
= 1\,.
\end{equation*}
Using the same Borel-Cantelli argument as in the proof of \eqref{eq:dndfnseg}, one obtains the desired relation
\begin{equation*}
\limsup_{n\to \infty} \twonorm{\bfR-(1+\gamma) \bfI} \le  2 \sqrt{\gamma}\quad \as
\end{equation*}
\end{proof}
With Proposition~\ref{prop:unified} we can finish the proof of \eqref{eq:dndfnseg1}.
We have 
\begin{equation*}
\twonorm{\bfR-(1+\gamma) \bfI} = \max \{\mu_{(1)}- (1+\gamma), -\mu_{(p)}+(1+\gamma)\}\,.
\end{equation*}
From \eqref{eq:jo1} we conclude
%\begin{equation*}
%\limsup_{n\to \infty} \twonorm{\bfR-(1+\gamma) \bfI} \le 2 \sqrt{\gamma}\,,
%\end{equation*}
%then 
\begin{equation*}
\begin{split}
\limsup_{n\to \infty}\mu_{(1)} &\le \phantom{-} 2 \sqrt{\gamma}+1+\gamma=(1+\sqrt{\gamma})^2 \quad \as\,,\\
\liminf_{n\to \infty}\mu_{(p)} &\ge -2 \sqrt{\gamma}+1+\gamma=(1-\sqrt{\gamma})^2\quad \as
\end{split}
\end{equation*}

\section{Proof of Theorem~\ref{thm:mpcorrelation}}\label{sec:6}
\subsection{Proof of Theorem~\ref{thm:mpcorrelation}(1)}\label{sec:5.1}
We appeal to the proof of Theorem 2.3 in \cite{bai:zhou:2008}.
The following lemma is a version of Corollary 1.1 in \cite{bai:zhou:2008}.
\begin{lemma}\label{lem:baicorrected}
Let 
$\mathbf{B} =\mathbf{B}_n= (B_{jk)}$ be a non-random $n\times n$ matrix with bounded norm
and 
\begin{equation*}
\begin{split}
\mathcal{S}=\{ (i_1&,j_1,i_2,j_2): 1\le i_1,j_1,i_2,j_2\le n\}\\ &\backslash \{(i_1,j_1,i_2,j_2): i_1=i_2 , j_1=j_2 \text{ or } i_1=j_2\neq i_2=j_1  \}\,.
\end{split}
\end{equation*}
%Assume \eqref{Cpn}. 
If $\E[Y^4 ] = o(n^{-1})$,
\beam\label{eq:con1}
n\,\var(Y_{1}Y_{2}) &\to& 0\,,\\
\label{eq:con2}
V_n=n^2\, \sum_{\mathcal{S}} \big(\cov(Y_{i_1}Y_{j_1},Y_{i_2}Y_{j_2})\big)^2 &\to& 0\,, 
\eeam
then Condition 1 of Theorem 1.1 in \cite{bai:zhou:2008} holds, i.e.,
\begin{equation*}
\E[ |\Y_{0} \mathbf{B} \Y_{0}'- \tr(\mathbf{B}\,\E[\bfY_0\bfY_0'])|^2] = o(1),
\end{equation*}
where $\Y_{0}= (Y_{1}, \ldots, Y_{n})$.
\end{lemma}
\begin{proof}
We have for some constant $c>0$, 
\beao
\lefteqn{\E[ |\Y_{0} \mathbf{B} \Y_{0}'- \tr(\mathbf{B}\E[\bfY_0\bfY_0'])|^2]}\\ &=& 
\E\Big[ \Big| \sum_{i_1,j_1=1}^n B_{i_1j_1} (Y_{i_1} Y_{j_1} - \E[Y_{i_1} Y_{j_1}]) \Big|^2\Big]\\
&=& \sum_{i_1,j_1=1}^n \sum_{i_2,j_2=1}^n B_{i_1j_1}B_{i_2j_2}
\cov (Y_{i_1} Y_{j_1},Y_{i_2} Y_{j_2})\\
&\le &c\, \big[n  \,\var(Y_{1}^2)
 + n \, \var(Y_{11} Y_{12})\big] 
+ \sum_{\mathcal{S}} B_{i_1j_1}B_{i_2j_2} \cov(Y_{i_1} Y_{j_1}, Y_{i_2} Y_{j_2})\,.
\eeao
By assumption,
$n  \,\var(Y^2)  = n \,(\E[Y^4]-n^{-2})\to 0$. The second summand converges to zero by \eqref{eq:con1}. 
It is shown in 
\cite{bai:zhou:2008} that the last summand is bounded by
\begin{equation*}
c\, n \Big(\sum_{\mathcal{S}} \big( \cov(Y_{i_1} Y_{j_1},Y_{1i_2} Y_{1j_2})\big)^2 \Big)^{1/2}
\end{equation*}
which converges to zero by \eqref{eq:con2}.
\end{proof}
\begin{remark}{\em Lemma~\ref{lem:baicorrected} corrects the proof of Theorem 2.3 and Corollary 1.1 in \cite{bai:zhou:2008}. 
In the latter paper
it is claimed that
\beao
V_n'=n^2\sum_{\mathcal{S'}} \big(\cov(Y_{i_1}Y_{j_1},Y_{i_2}Y_{j_2}\big)^2 \to 0\,,
\eeao 
where 
\begin{equation*}
\begin{split}
\mathcal{S'}=\{ (i_1&,j_1,i_2,j_2): 1\le i_1,j_1,i_2,j_2\le n\}\\ & \backslash \{(i_1,j_1,i_2,j_2): i_1=i_2 \neq j_1=j_2 \text{ or } i_1=j_2\neq i_2=j_1  \}\,. 
\end{split}
\end{equation*}
However, $\mathcal{S}'$ contains the quadruples $(i,i,i,i)$. Hence
\beao
V_n'&\ge&  n p^2 \big( \var(Y^2)  \big)^2= n^{-1} \,p^2\, ( n\,\E[Y^4 ] )^2 - 2\, \frac{p^2}{n^2}\, (n\, \E[Y^4 ]) + \frac{p^2}{n^3}\,,
\eeao
which does not necessarily converge to zero since 
$n\,\E[Y^4 ]$ may converge to zero arbitrarily slowly. }
\end{remark}
\par
Now we are ready for the proof of Theorem \ref{thm:mpcorrelation}(1).
If the \ds\ of $X$ is in the domain  of attraction of the normal law
the claim follows from Theorem 2.3 in \cite{bai:zhou:2008}, using our Lemma~\ref{lem:baicorrected}.
\par
Now assume the alternative condition \eqref{eq:condX}.
%By construction, the columns of $\Y$ are dependent. 
We will apply Theorem 2.2 in \cite{bai:zhou:2008} and our Lemma~\ref{lem:baicorrected}. 
Our goal is to find the limiting spectral distribution of $\bfR=\Y\Y'$ via the limit of the 
Stieltjes transform of $\Y'\Y$, using the fact that
$\Y\Y'$ and $\Y'\Y$ have the same non-zero eigenvalues. 
%The empirical spectral \ds\ of $\bfY'\bfY$ for $x\in \R$ satisfies the relation
%\begin{equation*}
%\frac{1}{n} \sum_{i=1}^n \1_{\{ \lambda_{(i)}(\Y'\Y)\le x\}}=\Big(1-\frac{p}{n}\Big) \1_{[0,\infty)} + \frac{p}{n} \frac{1}{p} \sum_{i=1}^p \1_{\{ \lambda_{(i)}(\Y\Y')\le x\}}\,.
%\end{equation*}
Since $\lambda_{(i)}=0$ for any of these matrices whenever $i>n\vee p$ 
we obtain a connection between the two spectral distributions:
\beao
F_{\Y'\Y}=\Big(1-\frac{p}{n}\Big) \1_{[0,\infty)} + \frac{p}{n} F_{\Y\Y'}\,.
\eeao
Hence 
\beam\label{eq:sgdsglop}
s_{\bfR}(z)&=& \int \frac{1}{x-z} \dint F_{\bfR}(x)\nonumber\\
&=& \int \frac{1}{x-z} \dint \Big(\frac{n}{p} F_{\Y'\Y} - \Big( \frac{n}{p}-1 \Big) \1_{[0,\infty)} \Big)(x)\nonumber\\
&=& \frac{n}{p} s_{\Y'\Y}(z)- \Big( \frac{n}{p}-1 \Big) \frac{1}{-z}\,,\quad z\in\C^+\,,
\eeam
where we used that for a constant $c\neq 0$ we have $s_{c\bfA}(z)=c^{-1} s_{\bfA}(cz)$.
\par
We introduce the $n\times n$ matrix $\bfT=(T_{ij})=(p\,\E[Y_{i}Y_{j}])$ which is a 
a circulant matrix whose eigenvalues can be determined as $T_{11}+(n-1) T_{12}$ and $T_{11}-T_{12}$ where the latter appears with multiplicity $n-1$. By assumption \eqref{eq:condX}, we have $T_{ij}=o(n^{-1})$ for $i\neq j$ and hence $\twonorm{\bfT}$ is bounded. The empirical spectral distribution 
\begin{equation*}
F_{\bfT}(x) = \frac{1}{n} \sum_{j=1}^n \1_{\{\lambda_j(\bfT)\le x\}}= \frac{\1_{\{T_{11}+(n-1) T_{12}\le x\}}}{n} +\frac{n-1}{n} \1_{\{T_{11}-T_{12} \le x\}}
\end{equation*}
converges to the degenerate distribution $H_\gamma$ with all mass at $\lim_{\nto} (T_{11}-T_{12})= \lim_{\nto}p/n =\gamma$.
\par
Next we verify the assumptions of Lemma~\ref{lem:baicorrected}.
We have 
\beao
n\, \E[\var(Y_1Y_2)]&=&  n\,\big(\E[(Y_1 Y_2)^2] - (\E[Y_{1}Y_{2}])^2 \big)\\
&\le& n \Big(\frac1 {n(n-1)} - o(n^{-2})\Big) \to 0\,, \qquad \nto\,.
\eeao
This implies \eqref{eq:con1}.
\par Now we turn to $V_n$ in \eqref{eq:con2}.
If we distinguish  between the types of indices in $\mathcal{S}$ we find that either possible structure 
for the summands $(Y_{i_1}Y_{j_1}-\E[Y_{i_1}Y_{j_1}])( Y_{i_2}Y_{j_2}-\E[Y_{i_2}Y_{j_2}])$ in $V_n$
is of the type $Y_{1}^3Y_{2}, Y_{1}^2Y_{2}Y_{3}$ or $Y_{1}Y_{2}Y_{3}Y_{4}$. Keeping this in mind, we conclude that for some constant $c>0$,
\beao
V_n&\le& c\,\Big(n^4\,\big( \cov(Y_1^2,Y_1Y_2)\big)^2 + n^5  \,\big( \cov(Y_1^2,Y_2Y_3)\big)^2
+ n^5 \, \big( \cov(Y_1Y_2,Y_2Y_3)\big)^2\\
&&+n^6\,\big(\cov(Y_1Y_2,Y_3Y_4)\big)^2\Big)\\ 
&\le & c \,\Big( n^4\,\big(\E[Y_{1}^3 Y_{2}] -(1/n) \E[Y_{1} Y_{2}] \big)^2 
+ n^5 \, \big( \E[Y_{1}^2 Y_{2}Y_{3}] -(1/n)  \E[Y_{1} Y_{2}] \big)^2\\
&&+ n^5\, \big(  \E[Y_{1} Y_{2}^2 Y_{3}] -  (\E[Y_{1} Y_{2}])^2 \big)^2+  n^6\,\big( \E[Y_{1} Y_{2}Y_{3}Y_{4}] -  
(\E[Y_{1} Y_{2}])^2 \big)^2\Big)\,.
\eeao
The \rhs\ converges to zero in view of assumption \eqref{eq:condX} and because (see \cite{gine:goetze:mason:1997}) 
\begin{equation*}
\E[Y_{1}^3 Y_{2}] = O(n^{-2}), \quad \E[Y_{1}^2 Y_{2}Y_{3}] = O(n^{-3})\quad \text{ and } \quad 
\E[Y_{1} Y_{2}Y_{3}Y_{4}] = O(n^{-4})\,.
\end{equation*}
Applications of Theorem 2.2 in \cite{bai:zhou:2008} and our Lemma~\ref{lem:baicorrected} yield
for $s=\lim_{\nto}s_{\bfY'\bfY}$, 
\begin{equation*}
\begin{split}
s(z)&= \int \frac{1}{\omega (1-\gamma^{-1}-\gamma^{-1}zs(z))-z} \dint H_\gamma(\omega)\\
&= \frac{1}{\gamma (1-\gamma^{-1}-\gamma^{-1}zs(z))-z}\,,
\end{split}
\end{equation*}
Thus $s=s(z)$ is the solution of the quadratic equation
\begin{equation*}
s^2 z + s (1+z-\gamma) +1=0.
\end{equation*}
By convention of \cite{bai:silverstein:2010}, the square root of 
a complex number is the one with a positive imaginary part. Hence
\beao
s(z)= \frac{-(\gamma^{-1}z+\gamma^{-1}-1)+\sqrt{(\gamma^{-1}z-\gamma^{-1}-1)^2-4\gamma^{-1}}}{2\gamma^{-1}z}.
\eeao
Writing $m$ for the limiting Stieltjes transform of $F_{\Y\Y'}$, we conclude from \eqref{eq:sgdsglop} and since $n/p\to \gamma^{-1}$ that
\beao
m(z)&=&\gamma^{-1} s(z)+\frac{\gamma^{-1}-1}{z}\\
%\frac{-(\gamma^{-1}z-\gamma^{-1}+1)+\sqrt{(\gamma^{-1}z-\gamma^{-1}-1)^2-4\gamma^{-1}}}{2z}\\
&=&\frac{1-\gamma -z +\sqrt{(1+\gamma-z)^2-4\gamma}}{2 \gamma z}\,,
\eeao
which we recognize as the Stieltjes transform of the \MP law in \eqref{eq:MPch4}; see \eqref{eq:stieltjestransform}. The proof is complete.

\subsection{Proof of Theorem~\ref{thm:mpcorrelation}(2)}\label{sec:5.5}
%\begin{proof}[Proof of Theorem~\ref{thm:mpcorrelation}(2)]
Assume $\liminf_{\nto} n\,\E [ Y^4] =\delta >0$.
For $k\ge 1$, the expected moments of the empirical spectral distribution $F_\bfR$ are 
%\begin{equation*}
%\beta_k(\bfR)=\int x^k \dint F_{\bfR}(x)= \frac{1}{p} \tr \bfR^k=\frac{1}{p} \tr(\Y\Y')^k\,.
%\end{equation*}
%Therefore, the expected moments are 
\begin{equation}\label{eq:expmomentsr}
\wt \beta_k = \E\Big[\int x^k \dint F_{\bfR}(x)\Big]=p^{-1}\E[\tr\,\bfR^k]=p^{-1}\sum_{i_1,\ldots,i_k=1}^p F(i_1,\ldots,i_k)\,.
\end{equation}
From \eqref{eq:xx} we know that 
\begin{equation*}
\begin{split}
p^{-1}\E[\tr(\bfR)^k]%&=\sum_{r=1}^k (p-1)(p-2)\cdots (p-r+1) \sum_{u=0}^{k-r} \sum_{I\in \mathcal{I}_{r,k}(u)} F(I)\\
&\ge \sum_{r=1}^k (p-1)(p-2)\cdots (p-r+1) \Big(\sum_{I\in \mathcal{I}_{r,k}(0)} +\sum_{I\in \mathcal{I}_{r,k}(1)}\Big)
F(I)  =: S_3+S_4\,.
\end{split}
\end{equation*}
By Proposition~\ref{prop:paths} and \eqref{lem:lemma3.4}, we have
\begin{equation}\label{eq:sumS11}
\lim_{\nto} S_3 = \sum_{r=1}^{k} \frac{1}{r} \binom{k}{r-1}\binom{k-1}{r-1}\gamma^{r-1}=\beta_k(\gamma)\,,
\end{equation}
which we recognize from \eqref{eq:momentsmp} as the $k$-th moment of the \MP law. 
\par
Next, observe that for $k\ge 4$ and $2\le r\le k-2$, $\mathcal{I}_{r,k}(1)$ contains the element 
\begin{equation*}
I_r = (1,2,1,2,\underbrace{2,\ldots,2}_{k-r-2},3,\ldots,r)\,.
\end{equation*}
One checks that $R(I_r)=r-2$ and $S(I_r)=(1,2,1,2)$; consult the PSA and Definition~\ref{def:pathshortening} for the definitions of $R(\cdot)$ and $S(\cdot)$. Moreover, by symmetry of $Y_{it}$
we have	
\begin{equation*}
\begin{split}
F(1,2,1,2)&=
\sum_{t_1,\ldots,t_4=1}^n \E[ Y_{1t_1}Y_{1t_2}Y_{1t_3}Y_{1t_4} 
Y_{2t_1}Y_{2t_2}Y_{2t_3}Y_{2t_4}]\\ &= \sum_{t_1,\ldots,t_4=1}^n (\E[ Y_{t_1}Y_{t_2}Y_{t_3}Y_{t_4} ])^2\\
&= \sum_{t_1=1}^n (\E[ Y_{t_1}^4])^2 + 3 \sum_{t_1\neq t_2=1}^n (\E[ Y_{t_1}^2Y_{t_2}^2 ])^2\ge \frac{1}{n} (n\E[ Y^4])^2\,.
\end{split}
\end{equation*}
%Since $X$ is regularly varying with index $\alpha<2$ it follows that $n\E\Big[ \frac{X_{11}^4}{D_{1}^2}\Big]$ has a positive limit, say $\delta$.
By Lemma~\ref{prop:PSI} we have 
\begin{equation*}
F(I_r)= n^{2-r} F(1,2,1,2)\ge n^{1-r}(n\E[ Y^4])^2
\end{equation*}and consequently
\begin{equation*}
\begin{split}
\liminf_{\nto} S_4&\ge \liminf_{\nto} \sum_{r=2}^{k-2} (p-1)(p-2)\cdots (p-r+1)  F(I_r)\\
&\ge \liminf_{\nto} \sum_{r=2}^{k-2} (p-1)(p-2)\cdots (p-r+1)  n^{1-r}(n\E[ Y^4])^2=  \delta^2 \sum_{r=2}^{k-2} \gamma^{r-1}\,.
\end{split}
\end{equation*}
This together with \eqref{eq:sumS11} proves $\liminf_{\nto} \wt \beta_k>\beta_k(\gamma)$, as desired.

%--------------------------------------------------------------------------------
%\appendix
\section{Appendix}
In this section we provide some auxiliary tools for the proofs of the main results.
\begin{lemma}\label{lem:fkgammaneww}
Let $k\in \N$ and $1\le j\le k$. Then
\begin{equation}\label{eq:vdsvsg}
-\sum_{i=2j-1}^{2k} (-1)^i \binom{2k}{i} \binom{i-1}{2j-2}=1
\end{equation}
\end{lemma}
\begin{proof}
%If $j=k$, we have 
%\begin{equation*}
%-\sum_{i=2k-1}^{2k} (-1)^i \binom{2k}{i} \binom{i-1}{2k-2}=2k-(2k-1)=1\,.
%\end{equation*}
For $1\le j\le k$ we rewrite \eqref{eq:vdsvsg} as 
\begin{equation}\label{eq:vdsvsg2}
-\sum_{i=2j-1}^{2k} (-1)^i \binom{2k}{i} \frac{(i-1)!}{(i+1-2j)!}=(2j-2)!\,.
\end{equation}
We define the functions
\begin{equation*}
u(x)=(x-1)^{2k}-1\,,\quad v(x)=\sum_{i=1}^{2k}  \binom{2k}{i}(-1)^i x^{i-1}\,,\quad \text{and } \quad w(x)=\frac{1}{x}\,.
\end{equation*}
Then $v(x)=u(x)w(x)$ and since
\begin{equation*}
\frac{(i-1)!}{(i+1-2j)!}=(i-1)(i-2)\cdots (i-2j+2)\,,
\end{equation*}
equation \eqref{eq:vdsvsg2} is equivalent to an equation for the $(2j-2)$-th derivative of $v$ evaluated at $1$,
\begin{equation*}
v^{(2j-2)}(1)=-(2j-2)!\,.
\end{equation*}
By Leibniz's rule for differentiation, one gets
\begin{equation*}
v^{(2j-2)}(x)= (uw)^{(2j-2)}(x)= \sum_{\ell=0}^{2j-2} \binom{2j-2}{\ell} u^{(\ell)}(x) w^{(2j-2-\ell)}(x)\,.
\end{equation*}
Observe that $u^{(0)}(1)=-1$ and $u^{(\ell)}(1)=0$ for $1\le \ell \le 2j-2$. Furthermore we have $w^{(2j-2-\ell)}(1) = (2j-2-\ell)!$. Hence, we conclude
\begin{equation*}
v^{(2j-2)}(1)= \sum_{\ell=0}^{2j-2} \binom{2j-2}{\ell} u^{(\ell)}(1) w^{(2j-2-\ell)}(1)=- (2j-2)!\,,
\end{equation*}
completing the proof.
\end{proof}
\par
For $k\in \N$ and $x\in [0,1]$, define the function
\begin{equation}\label{eq:fkgamma}
f_k(x)=1+\sum_{i=1}^{2k} \binom{2k}{i} \Big( \frac{-1}{1+x}\Big)^i \sum_{r=1}^i \frac{1}{r} \binom{i}{r-1} \binom{i-1}{r-1} x^{r-1}\,.
\end{equation}
The following is our key lemma.

\begin{lemma}\label{lem:fkgamma}
We have for $k\in \N$ and $x\in [0,1]$
\begin{equation*}
f_k(x) = 1 - \sum_{j=1}^k \frac{x^{j-1}}{(1+x)^{2j-1}} \frac{(2j-2)!}{j! (j-1)!} \le \frac{ (4x)^k}{(1+x)^{2k-1}}\,.
\end{equation*}
\end{lemma}
\begin{proof}
%Let $x\in [0,1]$ and $k\in \N$. First, we find $f_1(x)=x/(1+x)$. 
%With long and tedious calculations one can show that 
%\begin{equation}\label{eq:sejfo}
%f_{j-1}(x)-f_j(x) =\frac{x^{j-1}}{(1+x)^{2j-1}} \frac{4^{k-1} \Gamma(j-1/2)}{\sqrt{\pi}\Gamma(j+1)} = \frac{x^{j-1}}{(1+x)^{2j-1}} \frac{(2j-2)!}{j! (j-1)!}\,, \quad j\ge 2\,,
%\end{equation}
%where we used the following identity for the Gamma function $\Gamma(\cdot)$,
%\begin{equation*}
%\Gamma(j-1/2)= \frac{(2j-2)! \sqrt{\pi}}{4^{j-1}(j-1)!}\,.
%\end{equation*}
%By a telescoping sum argument and \eqref{eq:sejfo} we get 
%\begin{equation*}
%\begin{split}
%f_k(x)&= f_1(x)- \sum_{j=2}^k (f_{j-1}(x)-f_j(x))\\
%&= \frac{x}{1+x} - \sum_{j=2}^k \frac{x^{j-1}}{(1+x)^{2j-1}} \frac{(2j-2)!}{j! (j-1)!}\,.
%\end{split}
%\end{equation*}
From \cite[page 41]{bai:silverstein:2010} we know that 
\begin{equation*}
\sum_{r=1}^i \frac{1}{r} \binom{i}{r-1} \binom{i-1}{r-1} x^{r-1}
= \sum_{r=0}^{\lfloor(i-1)/2 \rfloor} x^r (1+x)^{i-1-2r} \frac{(i-1)!}{(i-1-2r)! r! (r+1)!}\,.
\end{equation*}
Changing the order of summation one obtains
\begin{equation*}
\begin{split}
f_k(x)-1&= \sum_{r=0}^{k-1} \sum_{i=2r+1}^{2k} \binom{2k}{i} (-1)^i x^r (1+x)^{-1-2r}\frac{(i-1)!}{(i-1-2r)! r! (r+1)!}\\
&= \sum_{j=1}^k \frac{x^{j-1}}{(1+x)^{2j-1}} \frac{1}{j! (j-1)!} \sum_{i=2j-1}^{2k} \binom{2k}{i} (-1)^i \frac{(i-1)!}{(i+1-2j)!}\\
&= -\sum_{j=1}^k \frac{x^{j-1}}{(1+x)^{2j-1}} \frac{(2j-2)!}{j! (j-1)!}\,,
\end{split}
\end{equation*}
where the last equality followed from Lemma~\ref{lem:fkgammaneww} and its equivalent formulation \eqref{eq:vdsvsg2}.

For $j \in \N$ define
\begin{equation}\label{eq:gkgamma}
g_j(x)= \frac{(1+x)^{2j-1}}{x^j} f_j(x)\,.
\end{equation}
We have $g_1(x)=1$ and $g_2(x)=2+x$. A straightforward induction proves the recursion
\begin{equation}\label{eq:gkgammarec}
g_j(x)= \frac{(1+x)^2 g_{j-1}(x)-g_{j-1}(0)}{x}\,, \quad j\ge 2\,.
\end{equation}
From this recursive construction one deduces that $g_j(x)$ is a polynomial of degree $j-1$ with positive coefficients.

Next we show $g_k(x)\le 4^k$. Clearly we have $g_1(x)\le 4$ and $g_2(x)\le 4^2$. Therefore assume
\begin{equation*}
\| g_{k-1}\|_{[0,1]} := \sup_{y\in [0,1]} |g_{k-1}(y)| \le 4^{k-1}\,.
\end{equation*}
Then for $x\in [0,1]$,
\begin{equation*}
\begin{split}
g_k(x) &= \frac{x (2+x) g_{k-1}(x)+g_{k-1}(x)-g_{k-1}(0)}{x}\\
&\le (2+x) g_{k-1}(x)+g_{k-1}(1) \le (2+x+1) \| g_{k-1}\|_{[0,1]}\\
&\le (3+x) 4^{k-1} \le 4^k\,.
\end{split}
\end{equation*}
In view of \eqref{eq:gkgamma}, this finishes the proof.
\end{proof}

%--------------------------------------------------------------------------------

%\noindent{\bf Acknowledgments}
%A large part of this paper was written during a research visit of JH at the Statistics %Department of 
%Columbia University in the City of New York. JH thanks the Department for its great 5%hospitality.
%JH and TM  thank Richard A. Davis, Olivier Wintenberger and Gennady Samorodnitsky for %valuable discussions on the subject.

\begin{comment}
\section[An example]{Example: Weak limit of extreme eigenvalues under infinite fourth moment}\label{sec:anexample}
\sectionmark{An example}

Assume $X$ has a symmetric distribution with infinite fourth moment. In principle, \eqref{eq:assumptionq} can be verified by direct calculation using \eqref{eq:formulagine}. However, for some classical distributions, such as the $t$- and symmetrized Pareto distributions, \eqref{eq:formulagine} requires the calculation of a high-dimensional integral which might not always be as easy to compute as in Examples \ref{ex:normal} and \ref{ex:gamma}.

In this section, we provide an explicit example of a distribution with infinite fourth moment such that $\mu_{(1)}\to (1+\sqrt{\gamma})^2$ and $\mu_{(1)}\to (1-\sqrt{\gamma})^2$ in probability.

For a positive slowly varying sequence $\delta_n\to 0$ define $\widehat{X}_{it}^{(n)}=X_{it} \1_{\{|X_{it}|\le \delta_n \sqrt{n}\}}$ and set
\begin{equation}\label{eq:truncated}
\widehat{\X}_n= \big(\widehat{X}_{it}^{(n)}\big)_{i=1,\ldots,p;t=1,\ldots,n}\,.
\end{equation}
The sample correlation matrix of the truncated data $\widehat{\bfR}=\widehat{\Y}\widehat{\Y}'$ with $\widehat{Y_{it}}=\widehat{X}^{(n)}_{it}/\sqrt{\sum_{t=1}^n (\widehat{X}_{jt}^{(n)})^2}$ and 
\begin{equation}\label{eq:corrRhat}
\widehat{\bfR}= \big(\widehat{R}_{ij}\big) = \Big(\sum_{t=1}^n \frac{\widehat{X}_{it}^{(n)}\widehat{X}_{jt}^{(n)}}{\sqrt{\sum_{t=1}^n (\widehat{X}_{it}^{(n)})^2} \sqrt{\sum_{t=1}^n (\widehat{X}_{jt}^{(n)})^2}} \Big)\,, \quad i,j=1,\ldots,p\,,
\end{equation}
has largest and smallest eigenvalues $\widehat{\mu}_{(1)}$ and $\widehat{\mu}_{(p)}$, respectively. For ease of notation we will again write $(\widehat{Y}_1, \ldots, \widehat{Y}_n)= (\widehat{Y}_{11}, \ldots, \widehat{Y}_{1n})$ and $\widehat{D}=\widehat{D}_1=\sum_{t=1}^n (\widehat{X}_{1t}^{(n)})^2$.
\par

{\em {\bf{Condition}} \eqref{eq:assumptionqhat}: There exists a positive slowly varying sequence $\delta_n\to 0$ such that 
\begin{equation}\label{eq:gsdglsrk}
\lim_{\nto} n^2\, \P(|X|>\delta_n \sqrt{n}) =0\,
\end{equation}
and there exists a sequence $q=q_n\to \infty$ such that for some integer sequence $k=k_n$ with $k/\log n \to \infty$ we have $(k^3 q)/n \to 0$ and the moment inequality
\begin{equation}\label{eq:assumptionqhat}
\E[ \widehat{Y}_{1}^{2m_1}\cdots \widehat{Y}_{r}^{2m_r} ] \le 
 \frac{q_n}{n} \, \E[ \widehat{Y}_{1}^{2m_1}\cdots \widehat{Y}_{r-1}^{2m_{r-1}}\widehat{Y}_{r}^{2m_r-2} ]\, \tag{$\widehat{C}_q$}
\end{equation}
holds for $1\le r\le k-1$ and any $m_1,\ldots,m_r$ positive integers satisfying $m_1+\cdots +m_r=k$.}
\par

Loosely speaking, condition \eqref{eq:assumptionq} fails if the probability of one $Y_i^2$ being much larger than the sum of the other $Y_j^2$ is ``not small enough''. In some cases it might be easier to check condition \eqref{eq:assumptionqhat} instead of \eqref{eq:assumptionq} since one is allowed to include an extra truncation step which alleviates the aforementioned issue at the cost of the additional restriction \eqref{eq:gsdglsrk}.

If $\E[X^{4}] <\infty$, then by Lemma 2.2 in \cite{bai:yin:krishnaiah:1988}, $\P(\X_n  \neq \widehat{\X}_n, \text{ i.o.})=0$, which implies
\begin{equation*}
\max_{1\le i\le p} \{ |\mu_{(i)}-\widehat{\mu}_{(i)}|\} \cas 0\,.
\end{equation*}
Under \eqref{eq:assumptionqhat}, one only has $\P(\X_n  \neq \widehat{\X}_n)\to 0$, which yields
\begin{equation}\label{eq:gsejkbfsejkfn}
\max_{1\le i\le p} \{ |\mu_{(i)}-\widehat{\mu}_{(i)}| \} \cip 0\,.
\end{equation}

Next we present an extremely crude way to check \eqref{eq:assumptionqhat}.
\begin{corollary}\label{cor:trunc}
%Under the conditions of Theorem~\ref{thm:mu1convergence}, 
If $\E[X]=0$,$|X|\ge c>0$ and $\E[X^4]<\infty$, then condition \eqref{eq:assumptionqhat} is valid.
\end{corollary}
\begin{proof}
If $\E[X]=0$ and $\E[X^4]<\infty$, equation \eqref{eq:gsdglsrk} follows from Borel--Cantelli and the proof of Lemma 2.2 in \cite{bai:yin:krishnaiah:1988}.
Since $|X|\ge c>0$ and $\widehat{X}^2 \le \delta_n^2 n$ we get
\begin{equation*}
\begin{split}
\E[ \widehat{Y}_{1}^{2m_1}\cdots \widehat{Y}_{r}^{2m_r} ] &\le 
\delta_n^2 n \, \E[ \widehat{Y}_{1}^{2m_1}\cdots \widehat{Y}_{r-1}^{2m_{r-1}}\widehat{Y}_{r}^{2m_r-2} D^{-1}]\\
&\le \delta_n^2  \, \E[ \widehat{Y}_{1}^{2m_1}\cdots \widehat{Y}_{r-1}^{2m_{r-1}}\widehat{Y}_{r}^{2m_r-2}]/c\\
&= \frac{q_n}{n}  \, \E[ \widehat{Y}_{1}^{2m_1}\cdots \widehat{Y}_{r-1}^{2m_{r-1}}\widehat{Y}_{r}^{2m_r-2}]\,,
\end{split}
\end{equation*}
for $q_n=n \delta_n^2/c$.
Choose $\delta_n=(\log n)^{-2}$ and $k_n=(\log n)^{1+\vep/3}$ with $\vep \in (0,1)$. We calculate
\begin{equation*}
\frac{k_n^3 q_n}{n}=c^{-1}\,(\log n)^{3+\vep-4} \to 0\,,
\end{equation*}
which finishes the proof.
\end{proof}

\begin{proposition}\label{prop:exampleinfinite}
Let $u>4$ and $s\in (4/u,1)$. Consider a symmetric random variable $X$ satisfying $|X|\ge c>0$ and
\begin{equation*}
\P(|X|>x)=x^{-4} (\log x)^{-s}\,, \quad x\ge x_0\,.
\end{equation*}
Then $\E[X^4]=\infty$. % Assume the iid entries $X_{it}$ have the same distribution as $X$. 
If $p/n \to \gamma \in (0,1]$, we have for the largest eigenvalue $\mu_{(1)}$ and the smallest eigenvalue $\mu_{(p)}$ of the sample correlation matrix $\bfR$:
\begin{equation}\label{eq:convprob}
\mu_{(1)} \cip (1+\sqrt{\gamma})^2 \quad \text{ and } \quad \mu_{(p)} \cip (1-\sqrt{\gamma})^2\,.
\end{equation}
\end{proposition}
\begin{proof}
For the slowly varying sequence $\delta_n=(\log n)^{-u}$ and $\widehat{X}_{it}^{(n)}=X_{it} \1_{\{|X_{it}|\le \delta_n \sqrt{n}\}}$, we have due to $4u-s<0$,
\begin{equation*}
\begin{split}
\P(\X_n  \neq \widehat{\X}_n)&\le np \, \P(|X|>\delta_n \sqrt{n})\\
&\sim 2^{-s} (\log n)^{4u-s} \to 0\,.
\end{split}
\end{equation*}
%This proves
%\begin{equation}\label{eq:gsejkbfsejkfn}
%\max\{ |\mu_{(1)}-\widehat{\mu}_{(1)}|,  |\mu_{(p)}-\widehat{\mu}_{(p)}| \} \cip 0\,.
%\end{equation}
By Corollary~\ref{cor:trunc}, condition  \eqref{eq:assumptionqhat} holds. 
Thus, by construction, the truncated entries $\widehat{X}_{it}$ satisfy condition \eqref{eq:assumptionq} and are symmetric. Following the proof of Theorem~\ref{thm:mu1convergence} under condition \eqref{eq:assumptionq}, we get 
\begin{equation*}
\widehat{\mu}_{(1)} \cas (1+\sqrt{\gamma})^2 \quad \text{ and } \quad \widehat{\mu}_{(p)} \cas (1-\sqrt{\gamma})^2\,.
\end{equation*}
Together with \eqref{eq:gsejkbfsejkfn} this finishes the proof.
\end{proof}
Using the crude Corollary~\ref{cor:trunc}, the almost sure convergence of the extreme eigenvalues is replaced by weak convergence. 
\end{comment}

### 2017-07-18 phd/Chapterextremes.tex
Added My PhD thesis.

%extremes Chapter
\chapter[Extreme value analysis for the sample autocovariance matrices of heavy-tailed multivariate time series]{{\huge Extreme value analysis for the sample autocovariance matrices of heavy-tailed multivariate time series}}\label{ch:extremes}
\chaptermark{Sample autocovariance matrices}

\begin{center}
\textsc{Richard Davis, Johannes Heiny, \\Thomas Mikosch \& Xiaolei Xie\\
{\em Extremes 19}, 3 (2016), 517--547.}
\end{center}

\begin{abstract}{We  provide some  asymptotic theory for the largest eigenvalues of a sample covariance matrix
of a $p$-dimensional time series where the dimension $p=p_n$ converges to infinity when the sample size $n$ increases.
We give a short overview of the literature on the topic both in the light- and heavy-tailed cases when the data have
finite (infinite) fourth moment, respectively.
Our main focus is on the heavy-tailed case. In this case, one has a theory for the point process of the normalized eigenvalues
of the sample covariance matrix in the iid case but also when rows and columns of the data are linearly dependent.
We provide limit results for the weak convergence of these point processes to Poisson or cluster Poisson processes. Based o
this convergence we can also derive the limit laws of various functionals of the ordered eigenvalues such as the
joint convergence of a finite number of the largest order statistics, the joint limit law of the largest eigenvalue and the trace,
limit laws for successive ratios of ordered eigenvalues,
etc. We also develop some
limit theory for the singular values of the sample autocovariance matrices and their sums of squares. The theory is illustrated
for simulated data and for the components of the S\&P 500 stock index.
\medskip

\noindent {\bf Keywords:} Regular variation, sample covariance matrix, dependent entries,
largest  eigenvalues, trace, point process
  convergence, cluster Poisson limit,
infinite variance stable limit, Fr\'echet distribution.}
\end{abstract}

\newpage
\section[Estimation of the largest eigenvalues]{Estimation of the largest eigenvalues: an overview in the iid case}\label{sec:motivationch3}
\subsection{The light-tailed case}
%In recent years, \asy\ theory for large random matrices has a
%attracted a
%great deal of attention; see for example the monographs
%Bai and Silverstein \cite{bai:silverstein:2010} and Anderson et al.
%\cite{anderson:guionnet:zeitouni:2008}. The literature about {\em
%  heavy-tailed random matrices} is rather sparse. Here and in what
%follows, we call a random matrix {\em heavy-tailed} if its entries have
%\ds s with \regvary\ tails, typically with tail index below 4.
%The eigenvalues of heavy-tailed matrices
%with iid entries and dimension $p\times n$, where $p=p_n\to\infty$ and $p/n\to
%\gamma\in (0,\infty)$, were studied by Soshnikov \cite{soshnikov:2004,
%  soshnikov:2006} and  Auffinger et
%al. \cite{auffinger:arous:peche:2009}. In the latter reference and in
%Belinschi et al. \cite{belinschi:dembo:guionnet:2009},
%the sample covariance matrices of iid heavy-tailed \seq s were also studied.
%Bose et al. \cite{bose:hazra:saha:2009,bose:hazra:saha:2010}
%investigated the spectral norm of circulant-type heavy-tailed matrices
%and Ben Arous and Guionnet \cite{arous:guionnet:2008}
%studied the limits of heavy-tailed random Wigner matrices with
%infinite variance.
One of the exciting new areas of statistics is concerned with analyses of large data sets.
For such data one often studies the dependence structure via covariances and correlations.
In this paper we focus on one aspect: the estimation of the eigenvalues of the covariance matrix of a multivariate  \ts\
when the dimension $p$ of the series increases with the sample size $n$. In particular, we are interested in limit theory for the largest eigenvalues of the sample covariance matrix. This theory is closely related to topics from classical \evt\ such as
maximum domains of attraction with the corresponding normalizing and centering constants for maxima; cf. Embrechts et al.~\cite{embrechts:kluppelberg:mikosch:1997}, Resnick~\cite{resnick:2007,resnick:1987}.
Moreover, \pp\ \con\ with limiting Poisson and cluster Poisson processes enters in a natural way when one describes
the joint \con\ of the largest eigenvalues of the sample covariance matrix. Large deviation techniques find applications,
linking \evt\ with random walk theory and \pp\ \con . The objective of this paper is to illustrate some of the main
developments in random matrix theory for the particular case of the sample covariance matrix of multivariate \ts\ with
independent or dependent entries. We give special emphasis to the heavy-tailed case when \evt\ enters in a rather
straightforward way.
%We do not cover other parts of random matrix theory such as limit theory for the eigen-spectrum of random matrices with
%iid entries or other special types of matrices. Also in this case one can find many parallels with \evt.
\par
Classical multivariate \tsa\ deals with
observations which assume values in a $p$-dimensional space
where $p$ is ``relatively small'' compared to the sample size $n$.  With
the availability of large data sets $p$ can be ``large''
relative to $n$. One of the possible con\seq s is that standard asymptotics (such as the \clt )
break down and may even cause misleading results.
\par
The dependence
structure in multivariate data is often summarized by the covariance matrix which is typically
estimated by its sample analog.  For example, principal component analysis (PCA)
extracts principal component vectors corresponding to the largest
eigenvalues of the sample covariance matrix. The magnitudes of these eigenvalues provide an empirical \ms\ of the importance
of these components.
\par
If $p,n$ are fixed, a column of the $p\times n$ data matrix
\beao
\X =\X_n = \big(X_{it}\big)_{i=1,\ldots,p;t=1,\ldots,n}\,
\eeao
represents an observation of a $p$-dimensional \ts\ model with unknown parameters.
In this section we assume that the real-valued entries $X_{it}$ are iid, unless mentioned otherwise, and we write $X$ for a generic element.
One challenge is to infer information about the parameters
from the eigenvalues $\lambda_1,\ldots,\lambda_p$
of the {\em sample covariance matrix} $\X\X'$. In the notation we suppress the dependence of $(\la_i)$ on $n$ and $p$.
If $p$ and $n$ are finite and the columns of $\X$ are iid and multivariate normal,
Muirhead \cite{muirhead} derived a (rather complicated) formula for the joint distribution of the eigenvalues $(\lambda_i)$.
%Therefore a theoretical solution is already available in the finite $n$ case.
%The expression, however, contains an integral over the orthogonal group $\mathbb{O}(p)$.
%\begin{figure}[htb!]
%  \centering
 % \includegraphics[scale=0.5]{eigen_normal_qqplot1.pdf}
 % \caption{Largest eigenvalues of $500$ simulated sample covariance matrices against standard normal quantiles.}
 % \label{fig:QQ}
%\end{figure}

For $p$ fixed and $\nto$, assuming
$\X$ has centered normal entries and a diagonal covariance matrix $\Sigma$,
Anderson~\cite{anderson:1963}  derived the joint \asy\ density of $(\lambda_1, \ldots, \lambda_p)$.
%If the largest eigenvalue of the underlying covariance matrix appears with multiplicity one, he
%showed that the largest eigenvalue of the sample covariave matrix
%is asymptotically normal.
We quote from Johnstone~\cite{johnstone:2001}:
``The classic paper by Anderson~\cite{anderson:1963} gives the limiting joint distribution of the roots, but the
marginal distribution of the largest eigenvalue is hard to extract even in the null case'' (i.e., when the covariance matrix $\Sigma$ is
proportional to the identity matrix).
%If the largest entry of $\Sigma$ exceeds its second largest, one obtains a central limit type theorem for the largest eigenvalue of the sample covariance matrix.
%In Figure~\ref{fig:QQ} we chose $n=10000, p=50$ and $\Sigma =\diag(1024,1,\ldots,1)$, and plot the properly normalized and
%standardized largest eigenvalue against standard normal quantiles.

It turns out that limit theory for the largest eigenvalues becomes ``easier''
when the dimension $p$ increases with $n$.
Over the last 15 years there has been increasing interest in the case when $p=p_n\to\infty$ as $\nto$. In most of
the literature (exceptions are El Karoui \cite{elkaroui:2003}, Davis et al. \cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} and Heiny and Mikosch
\cite{heiny:mikosch:2015:iid})
one assumes that $p$ and $n$ grow at the same rate:
\beam\label{eq:gammach3}
\dfrac p n\to \gamma\qquad \mbox{for some $\gamma \in (0,\infty)$.}
\eeam
%One explanation for the commonly used condition \eqref{eq:gamma} is that much of the theory was developed for $n \times n$ {\em Wigner matrices}, which is a Hermitean matrix $H$ such that $\{ H_{ij}:i<j \}$ and $\{ H_{ii}:0\le i\le n \}$ are two sets of iid random variables. If $p =\gamma n$, it turns out that the limit theory for the eigenvalues of Wigner matrices and sample covariance matrices is closely connected by the convergence of the empirical spectral distributions to the semicircle and the Mar\v cenko--Pastur law, respectively; see Bai and Silverstein \cite{bai:silverstein:2010}, Chapters 1-3.

In random matrix theory,  the convergence of the {\em empirical spectral distributions} $(F_{n^{-1}\X \X'})$ of a sequence $(n^{-1}\X \X')$ of non-negative definite matrices is the principle object of study. The empirical spectral distribution $F_{n^{-1}\X \X'}$ is constructed from the eigenvalues via
\begin{equation*}
F_{n^{-1}\X \X'}(x)= \frac{1}{p}\; \# \{ 1\le j\le p : n^{-1} \lambda_j \le x \}, \quad x\in \R,\quad n\ge1.
\end{equation*}
In the literature convergence results for the sequence of empirical spectral distributions are established under the assumption that $p$ and $n$ grow at the same rate.
Suppose that the iid entries $Z_{it}$ have mean $0$ and variance $1$. If \eqref{eq:gammach3} holds, then, with probability one, $(F_{n^{-1}\X \X'})$ converges weakly to the celebrated Mar\v cenko--Pastur law $F_\gamma$. If $\gamma \in (0,1]$,  $F_\gamma$  has density,
\begin{eqnarray}\label{eq:MPch3}
f_\gamma(x) =
\left\{\begin{array}{ll}
\frac{1}{2\pi x\gamma} \sqrt{(b-x)(x-a)} \,, & \mbox{if } a\le x \le b, \\
 0 \,, & \mbox{otherwise,}
\end{array}\right.
\end{eqnarray}\noindent
where $a=(1-\sqrt{\gamma})^2$ and $b=(1+\sqrt{\gamma})^2$. If $\gamma>1$, the \MP law is a mixture of a point mass at $0$ and the density function $f_{1/\gamma}$ with weights $1-1/\gamma$ and $1/\gamma$, respectively.  The point mass at $0$ is intuitively explained by the fact that, with probability $1$, $\min(p,n)$ eigenvalues $\lambda_i$ are non-zero. When $n=(1/\gamma) \; p$ and $\gamma >1$ one sees that the proportion of non-zero eigenvalues of the sample covariance matrix is $1/\gamma$ while the proportion of zero eigenvalues is $1-1/\gamma$.

While the finite second moment is the central assumption to obtain the Mar\v cenko--Pastur law as the limiting spectral distribution, the finite fourth moment plays a crucial role when studying the largest eigenvalues
\beam\label{eq:orderch3}
\la_{(1)}\ge \cdots \ge \la_{(p)}
\eeam
of $\X\X'$, where we suppress the dependence on $n$ in the notation.

\begin{figure}[htb!]
  \centering
  \subfigure[Standard normal entries]{
    \includegraphics[scale=0.34]{normal-TW.pdf}
  }
  \subfigure[Entry distribution: $\P(X=\sqrt{3}) = \P(X=-\sqrt{3} ) =
  1/6$, $\P(X=0)=2/3$. Note $\E X = 0$, { $\E[ X^2] = 1$, $\E [X^3] = 0$ and
  $\E[ X^4] = 3$, i.e., the first 4 moments of $X$ match those of the standard normal \ds .}]{
    \includegraphics[scale=0.34]{3point-TW.pdf}
  }
  \caption{Sample density function of the largest eigenvalue compared
    with the Tracy--Widom density function. The data matrix $ \X$ has
    dimension $200 \times 1000$. An ensemble of 2000 matrices is
    simulated.}
  \label{fig:normal-3point-TW}
\end{figure}

Assuming \eqref{eq:gammach3} and iid entries $X_{it}$ with zero mean, unit variance and finite fourth moment,  Geman~\cite{geman}
showed that
\beam\label{eq:gemanch3}
\dfrac {\la_{(1)}}{n} \stas \big(1+\sqrt{\gamma}\big)^2\,,\qquad \nto\,.
\eeam
Johnstone \cite{johnstone:2001} complemented this \slln\ by the corresponding \clt\ in the special case of iid standard normal
entries:
\beam\label{eq:tcch3}
 n^{2/3}\,\dfrac{(\sqrt{\gamma})^{1/3}}{\big(1+\sqrt{\gamma}\big)^{4/3}}\Big(\dfrac {\la_{(1)}}{n} -
\big(1+\sqrt{\tfrac pn }\big)^2\Big)
\std {\rm TW}\,,
\eeam
where the limiting \rv\ has a {\em Tracy--Widom \ds} of order 1. Notice that the centering
$\big(1+\sqrt{\tfrac pn }\big)^2$ can in general not be replaced by $(1+\sqrt{\gamma})^2$.
This \ds\ is ubiquitous in random matrix theory.
%It is defined via some ordinary differential equation; we refer to \cite{tracy:widom:2012} for a definition and properties.
Its distribution function $F_1$ is given by
\begin{equation*}
F_1(s) = \exp\Big\{
  -\frac{1}{2} \int_{s}^\infty [
    q(x) + (x - s) q^2(x)
 ] \dint x
\Big\}\,,
\end{equation*}
where $q(x)$ is the unique solution to the Painlev\'e II differential
equation
\begin{equation*}
  q''(x) = xq(x) + 2 q^3(x)\,,
\end{equation*}
where $ q(x)\sim {\rm Ai}(x)$ as $x \to \infty$ and Ai$(\cdot)$ is the Airy kernel; see Tracy and Widom~\cite{tracy:widom:2012} for details.
We notice that the rate $n^{2/3}$ compares favorably to the $\sqrt{n}$-rate in the classical \clt\ for sums
of iid finite variance \rv s.
The calculation of the spectrum is facilitated by the fact that the distribution of
the classical Gaussian matrix ensembles is invariant under orthogonal transformations. The corresponding
computation for non-invariant matrices with non-Gaussian entries is more complicated and was a major challenge for several years; a first step was made by Johansson \cite{johansson}.
Johnstone's result was extended to matrices $\X$ with iid non-Gaussian entries
by Tao and Vu \cite[Theorem~1.16]{tao09b}. Assuming that the first four moments of the entry \ds\ match  those of
the standard normal \ds , they showed \eqref{eq:tcch3} by
employing {\em Lindeberg's replacement method}, i.e., the iid non-Gaussian entries are replaced
step-by-step by iid Gaussian ones.
This approach is well-known from summation theory for \seq s of iid \rv s. Tao and Vu's result is a consequence of the so-called {\em Four Moment Theorem}, which describes the insensitivity of the eigenvalues with respect to changes in the distribution of the entries. To some extent (modulo the strong moment matching conditions) it shows the universality of Johnstone's limit
result \eqref{eq:tcch3}. Later we will deal with entries with infinite fourth moment. In this case, the weak limit
for the normalized largest eigenvalue $\la_{(1)}$ is distinct from the Tracy--Widom \ds : the classical Fr\'echet extreme value
\ds\ appears.
In Figure~\ref{fig:normal-3point-TW} we illustrate how the Tracy--Widom approximation works for Gaussian and non-Gaussian entries of $\X$
and in Figure~\ref{fig:MyDist-Frechet} we also illustrate that this approach fails when $\E[X^4]=\infty$.

Figure \ref{fig:normal-3point-TW} compares the sample density function
of the properly normalized largest eigenvalue estimated from 2000 simulated sample covariance matrices $\X\X'$ ($n=1000, p=200$) with
the Tracy--Widom density. If  $X$ has infinite fourth moment and further regularity conditions on the tail hold then the
Tracy--Widom limiting law needs to be replaced by the \Frechet
distribution; see Section~\ref{subsec:1.2} for details. Figure \ref{fig:MyDist-Frechet} illustrates this fact with a
simulated ensemble whose entries are distributed according to
the heavy-tailed distribution from \eqref{eq:distrsimch3} below with $\alpha = 1.6$.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{MyDist-Frechet.pdf}
  \caption{Sample distribution function of the largest eigenvalue $\la_{(1)}$
    compared to the \Frechet distribution (solid line) with $\alpha=1.6$. The data matrices have dimension
    $200 \times 1000$ and iid entries with infinite fourth moment. The results are based on 2000 replicates.}
  \label{fig:MyDist-Frechet}
\end{figure}

\subsection{The heavy-tailed case}\label{subsec:1.2}
So far we focused on ``light-tailed'' $\X$
in the sense that its entries have finite fourth moment.  However, there is statistical evidence that the assumption
of finite fourth moment may be violated when dealing with data from
insurance, finance or telecommunications. We illustrate this fact
in Figure~\ref{fig:SP500_tail_indices} where we show the pairs $(\alpha_L,\alpha_U)$ of
lower and upper tail indices
of $p=478$  log-return series composing
the S\&P 500 index estimated from $n=1,345$ daily observations from 01/04/2010 to 02/28/2015.
This means we assume for every row  $(X_{it})_{t=1,\ldots,n}$ of $\X$ that the tails behave like
\beao
\P(X_{it}>x)\sim c_U\,x^{-\alpha_U}\qquad\mbox{and}\qquad \P(X_{it}<-x)\sim c_L\,x^{-\alpha_L}\,,\qquad \xto\,,
\eeao
for non-negative constants $c_L,c_U$. We apply the Hill estimator (see Embrechts et al.~\cite{embrechts:kluppelberg:mikosch:1997}, p.~330,
de Haan and Ferreira \cite{dehaan:ferreira:2006}, p.~69)
to the \ts\ of the gains and losses in a naive way,
neglecting the dependence and non-stationarity in the data; we also omit confidence bands.
From the figure it is evident that the majority of the return series have tail indices below
four, corresponding to an infinite fourth moment.
\begin{figure}[htb!]
\centering
\includegraphics[scale=0.5]{SP500_tail_indices.pdf}
\caption{Tail indices of log-returns of 478 \ts\ from the S\&P 500 index. The values  $(\hat \alpha_L,\hat \alpha_U)$ of the
lower and upper tail indices are provided by
Hill's estimator.  We also draw the line $\hat\alpha_U=\hat\alpha_L$.}
\label{fig:SP500_tail_indices}
\end{figure}
The behavior of the largest eigenvalue $\la_{(1)}$
changes dramatically when $\X$ has infinite fourth moment.
Bai and Silverstein~\cite{baisilv} proved for an $n\times n$ matrix $\X$ with iid centered entries
that
\beam\label{eq:wdfrch3}
\limsup_{\nto} \dfrac{\la_{(1)}}{n}=\infty \qquad {\rm a.s.}
\eeam
This is in stark contrast to Geman's result \eqref{eq:gemanch3}.
%It is also unlikely that the
%Tracy--Widom \ds\ yields a good approximation to the \ds\ of the largest eigenvalue of the sample covariance matrix;
%see Figure~\ref{fig:MyDist-Frechet} for some simulation evidence.
\par
In the heavy-tailed case it is common to assume a {\em \regvar\ condition:}
\beam\label{eq:regvarch3}
\P(X>x)\sim p_+\,\dfrac{L(x)}{x^{\alpha}}\qquad\mbox{and}\qquad \P(X<-x) \sim p_-\,\dfrac{L(x)}{x^\alpha}\,,\qquad \xto\,,
\eeam
where $ p_\pm$ are non-negative constants \sth\ $p_++p_-=1$ and $L$ is a \slvary\ \fct . In particular, if $\alpha<4$ we have $\E [X^4]= \infty$. The \regvar\ condition
on $X$ (we will also refer to $X$ as a \regvary\ \rv ) is needed for proving \asy\ theory for
the eigenvalues of $\X\X'$. This is similar to proving limit theory for sums
of iid \rv s with infinite variance stable limits; see for example Feller~\cite{feller}.
\par

In \eqref{eq:MPch3} we have seen that the sequence $(F_{n^{-1}\X \X'})$ of empirical spectral distributions converges to the Mar\v cenko--Pastur law if the centered iid entries possess a finite second moment. Now we will discuss the situation when the entries are still iid and centered, but have an infinite variance. Here we assume the entries to be regularly varying with index $\alpha \in (0,2)$.
Assuming \eqref{eq:gammach3} with $\gamma\in (0,1]$ in this infinite variance case, Belinschi et al.~\cite[Theorem~1.10]{belinschi:dembo:guionnet:2009} showed that the sequence $(F_{a_{n+p}^{-2}\X \X'})$ converges with probability one to a non-random probability measure with density $\rho_{\alpha}^\gamma$ satisfying
\begin{equation*}
\rho_{\alpha}^\gamma(x) x^{1+\alpha/2} \to \frac{\alpha \gamma}{2(1+\gamma)}, \quad x \to \infty,
\end{equation*}
see also Ben Arous and Guionnet \cite[Theorem~1.6]{arous:guionnet:2008}.
The normalization $(a_k)$ is chosen \sth\ $\P(|X|>a_k)\sim k^{-1}$ as $k\to\infty$. An application of the Potter bounds (see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}) shows that $a_{n+p}^2/n \to \infty$.

It is interesting to note that there is a phase change in the extreme eigenvalues in going from finite to infinite fourth moment, while the phase change occurs for the empirical spectral distribution going from finite to infinite variance.
%In the case of finite fourth moment, \eqref{eq:geman} and \eqref{eq:MP} imply that $\lambda_{(k)}/n$ has to converge to the right endpoint of the Mar\v cenko--Pastur law for any finite $k$. However, if we assume an infinite fourth moment but finite second moment the empirical spectral distributions  $(F_{{n}^{-1}\X \X'})$ still converge to the Mar\v cenko--Pastur law with the same support, while $\lambda_{(1)}/n$ will tend to infintity due to \eqref{eq:wdfr}. This means that the proportion of the normalized eigenvalues taking values larger than $(1+\sqrt{\gamma})^2$ has to be negligible; otherwise we would have a contradiction to the convergence of the empirical spectral distributions to the Mar\v cenko--Pastur law.

%Finally, if the second moment is infinity, i.e.~regularly varying entries with index $\alpha \in (0,2)$, then the much larger values of the eigenvalues are visible in the limiting spectral distribution even when using a much stronger normalization \cite[Theorem~1.10]{belinschi:dembo:guionnet:2009}.
\par

The theory for the largest eigenvalues of sample covariance matrices with heavy tails is less developed than in the light-tailed case.
Pioneering work for $\la_{(1)}$ in the case of iid \regvary\ entries $X_{it}$ with index $\alpha\in (0,2)$
is due to Soshnikov~\cite{soshnikov:2004,soshnikov:2006}. He showed the \pp\ \con\
\beam\label{eq:nn}
N_n=\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i} \std N=\sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeam
under the growth condition \eqref{eq:gammach3} on $(p_n)$.
%The normalization $(a_k)$ is chosen \st\ $\P(|X|>a_k)\sim k^{-1}$ as $k\to\infty$.
Here
\beam\label{eq:Gamma}
\Gamma_i=E_1+\cdots + E_i\,,\qquad i\ge 1\,,
\eeam
and $(E_i)$ is an iid standard exponential \seq . In other words, $N$ is a Poisson point process on $(0,\infty)$ with mean \ms\
$\mu(x,\infty)= x^{-\alpha/2}$, $x>0$. Convergence in \ds\ of \pp es is understood in the sense of weak \con\
in the space of point \ms s equipped with the vague topology; see Resnick \cite{resnick:2007,resnick:1987}.
We can easily derive the limiting distribution of $a_{np}^{-2} \lambda_{(k)}$ for fixed $k\ge 1$ from \eqref{eq:nn}:
\begin{equation*}
\begin{split}
\lim_{\nto}\P(a_{np}^{-2} \lambda_{(k)}\le x)&= \lim_{\nto}\P(N_n(x,\infty)<k)
=  \P(N(x,\infty)<k) =\P(\Gamma_k^{-2/\alpha}\le x)\\ &= \sum_{s=0}^{k-1} \frac{\big(\mu(x,\infty)\big)^s}{s!} \e^{-\mu(x,\infty)}, \quad  x>0.
\end{split}
\end{equation*}
In particular,
\beao
\dfrac{\la_{(1)}}{a_{np}^2}\std \Gamma_1^{-\alpha/2}\,,\qquad \nto\,,
\eeao
where the limit has {\em Fr\'echet \ds } with parameter $\alpha/2$ and distribution function
\beao
\Phi_{\alpha/2}(x) =\ex^{-x^{-\alpha/2}}\,,\qquad x>0\,.
\eeao

\noindent We mention that the tail balance condition \eqref{eq:regvarch3} may be replaced in this case by the weaker assumption
$\P(|X|>x)= L(x) x^{-\alpha}$ for a \slvary\ \fct\ $L$. Indeed, it follows from the proof %(see also the comments below)
that only the squares $X_{it}^2$ contribute to the \pp\ limits of the eigenvalues $(\la_i)$. A con\seq\ of the
\cmt\ and \eqref{eq:nn} is the joint \con\ of the upper order statistics: for any $k\ge 1$,
\beao
a_{np}^{-2} \big(\la_{(1)},\ldots,\la_{(k)}\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,,\qquad \nto\,.
\eeao

It follows from standard theory
for \pp es with iid points (e.g. Resnick \cite{resnick:2007,resnick:1987})
that \eqref{eq:nn} remains valid if
we replace  $N_n$  by the \pp\ $\sum_{i=1}^p\sum_{t=1}^n\vep_{X_{it}^2/a_{np}^2}$. Then we also have for any $k\ge 1$,
\beam\label{eq:nna}
a_{np}^{-2} \big(X_{(1),np}^2,\ldots,X_{(k),np}^2\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,,\qquad \nto\,,
\eeam
where
\beao
X_{(1),np}^2\ge \cdots \ge X_{(np),np}^2
\eeao
denote the order statistics of $(X_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$.
\par
Auffinger et al.~\cite{auffinger:arous:peche:2009} showed that \eqref{eq:nn}
remains valid under the \regvar\ condition \eqref{eq:regvarch3} for  $\alpha\in (2,4)$, the growth condition
\eqref{eq:gammach3} on $(p_n)$ and the
additional assumption $\E [X]=0$. Of course, \eqref{eq:nna} remains valid as well.
Davis et al.~\cite{davis:pfaffel:stelzer:2014} extended these results to the case when the rows of $\X$
are iid linear processes with iid \regvary\ noise. The Poisson \pp\ \con\ result of \eqref{eq:nn} remains valid in this case.
Different limit processes can only be expected if there is dependence across rows and columns.
\par
In what follows, we refer to the  {\em heavy-tailed case} when we assume the \regvar\ condition \eqref{eq:regvarch3}
for some $\alpha\in (0,4)$.%{\blue This hypothesis is confirmed in}

\subsection{Overview}

The primary objective of this overview  is to make a connection between extreme value theory and the behavior of the largest eigenvalues of sample covariance matrices from heavy-tailed multivariate time series.  For time series that are linearly dependent through time and across rows, it turns out that the extreme eigenvalues are essentially determined by the extreme order statistics from an array of iid random variables.  The asymptotic behavior of the extreme eigenvalues is then derived routinely from classical extreme value theory.  As such, explicit joint distributions of the extreme order statistics can be given which yield a plethora of ancillary results.  
%These include the limit behavior of ratios of consecutive extreme eigenvalues, ratios of extreme order statistics to the sample variance, etc..
%%The underlying multivariate \ts\ will assumed to be heavy-tailed and linearly dependent through time and across the rows.We consider asymptotic theory for the largest eigenvalues of a sample covariance matrix,
%and we will also address the largest singular values of the sample autocovariance matrix.
%To our knowledge, the present paper is the first to consider bonafide dependence among the components in the time series which renders a multivariate analysis, such as PCA, meaningful.
Convergence of the point process of extreme eigenvalues, properly normalized, plays a central role in establishing the main results.
\par
In Section~\ref{sec:2ch3} we continue the study of the case when the data matrix $\X$ consists of iid heavy-tailed entries.  We will consider power-law growth rates on the dimension $(p_n)$ that is more general than prescribed by \eqref{eq:gammach3}.
In Section~\ref{sec:model} we introduce a model for $X_{it}$ which allows for linear dependence across the rows and through time.
The point process convergence of normalized eigenvalues is presented in Section~\ref{sec:mainresult}.
This result lays the foundation for new insight into the spectral behavior of the sample covariance matrix, which is the content of Section~\ref{sec:samplecov}.

Sections~\ref{sec:samplecov} and \ref{sec:possemidef} are devoted to {\em sample autocovariance matrices}.  Motivated by \cite{lam:yao}, we study the eigenvalues of sums of transformed matrices and illustrate the results in two examples. These results are applied to the time series of S\&P 500 in Section \ref{sec:sp500}.
%The technical proofs are collected in Section~\ref{sec:proof}. 
Appendix~\ref{appendix:Ach3} contains useful facts about regular variation and point processes.

\section[General growth rates for $p_n$]{General growth rates for $p_n$ in the iid heavy-tailed case}\label{sec:2ch3}
This section is based on ideas in Heiny and Mikosch \cite{heiny:mikosch:2015:iid} where one can also find detailed proofs.
\subsection*{Growth conditions on $(p_n)$}\label{subsec:pn}
In many applications it is not realistic to assume
that the dimension $p$ of the data and the sample size $n$ grow at the same rate.
The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009} already indicate that the value $\gamma$ in the growth
rate~\eqref{eq:gammach3} does not appear in the \ds al limits.
This obervation is in contrast to the light-tailed case; see \eqref{eq:gemanch3}
and \eqref{eq:tcch3}.  Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}
and Heiny and Mikosch ~\cite{heiny:mikosch:2015:iid} allowed for more general rates for
$p_n\to\infty$ than linear growth in $n$.
Recall that $p=p_n\to\infty$ is the number of rows in the
matrix $\X_n$. We need to specify the growth rate of $(p_n)$ to ensure
a non-degenerate limit distribution of the normalized singular values of the sample autocovariance
matrices. To be precise, we assume
\begin{equation}\label{eq:pch3}
p=p_n=n^\beta \ell(n), \quad n\ge1,\tag{$C_p(\beta)$}
\end{equation}
where $\ell$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume $\ell(n) \to \infty$.
Condition~\ref{eq:pch3} is more general than the growth conditions in the literature; see
\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
\begin{theorem}\label{thm:intro}
Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvarch3} for some
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Let $(p_n)$ be an integer sequence
satisfying \ref{eq:pch3} with $\beta\ge 0$. In addition, we require
\begin{equation}\label{Cbeta}
\min(\beta,\beta^{-1})\in (\alpha/2-1,1] \qquad \mbox{ for } \alpha\in [2,4), \tag{$\widetilde{C}_\beta(\alpha)$}
\end{equation}
Then
\beam\label{eq:2}
\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeam
where the \con\ holds in the space of point \ms s with state space $(0,\infty)$ equipped with the vague toplogy; see Resnick \cite{resnick:2007}.
\end{theorem}

\subsubsection*{A discussion of the case $\beta\in [0,1]$}
We mentioned earlier that in the heavy-tailed case, limit theory for the
largest eigenvalues of the sample covariance matrix is rather insensitive to
the growth rate of $(p_n)$ and that the limits are essentially
determined by the diagonal of this matrix. This is confirmed by the following result.

\begin{proposition}\label{prop:offdiagonalch3}
Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvarch3} for some
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Then for any \seq\ $(p_n)$
satisfying \ref{eq:pch3} with $\beta\in [0,1]$ we have
\beao
a_{np}^{-2} \twonorm{\X \X' - \diag(\X \X')}\stp 0\,,\qquad\nto\,,
\eeao
where $\| \cdot\|_2$ denotes the spectral norm; see \eqref{specnorm} for its definition.
\end{proposition}

\par
Proposition~\ref{prop:offdiagonalch3} is not unexpected for two reasons:
\begin{itemize}
\item
It is well-known from classical theory (see Embrechts and Veraverbeke \cite{embrechts:veraverbeke:1982}) that for any iid \regvary\ non-negative \rv s
$Y,Y'$ with index  $\alpha'>0$, $Y\,Y'$ is \regvary\ with index $\alpha'$ while $Y^2$ is \regvary\ with index $\alpha'/2$.
Therefore $X^2$ and $X_{11}X_{12}$  are \regvary\ with indices $\alpha/2$ and $\alpha$, respectively.
\item
The aforementioned tail behavior is inherited by the entries of $\X\X'$ in the following sense.
By virtue of Nagaev-type \ld\ results for an iid \regvary\ \seq\ $(Y_i)$ with index $\alpha'\in (0,2)$ where we also assume that
$\E [Y_0]=0$ if $\E[|Y_0|]<\infty$
(see Theorem~\ref{thm:nagaevch3}) %S.V. Nagaev \cite{nagaev:1979} and
%Cline and Hsing \cite{cline:hsing:1998})
we have that $\P(Y_1+\cdots +Y_n >b_n)/(n \,\P(|Y_0|>b_n))$ converges to a non-negative constant
provided $b_n/a_n'\to\infty$, where $\P(|Y_0|>a_n')\sim n^{-1}$ as $\nto$. As a con\seq\ of the tail behaviors of $X_{it}^2$ and $X_{it}X_{jt}$ for $i\ne j$
and Nagaev's results we have for $(b_n)$ \sth\ $b_n/a_n^2\to\infty$,
\beam \label{eq:dfsdfjl}
\dfrac{\P\big((\X\X')_{ij}> b_n\big)}{\P\big((\X\X')_{ii}- c_n> b_n\big)}\sim \dfrac{n\,\P( X_{11}X_{12}>b_n)}{n\,\P(X^2>b_n)}\to 0\,,\qquad \nto\,,
\eeam
where $c_n=0$ or $n\,\E[X^2]$ according as $\alpha\in (0,2)$ or $\alpha\in (2,4)$.
This means that the diagonal and off-diagonal entries of
$\X\X'$ inherit the tails of $X_{it}^2$ and $X_{it}X_{jt}$, $i\ne j$, respectively, above the high threshold $b_n$.
\end{itemize}
\par
Proposition~\ref{prop:offdiagonalch3} has some immediate con\seq s for the approximation of the eigenvalues
of $\X\X'$ by those of ${\rm diag}(\X\X')$. Indeed, let $C$ be a symmetric $p\times p$ matrix with
eigenvalues $\la_1(C),\ldots,\la_p(C)$ and
ordered eigenvalues
\beam \label{eq:weyl}
\la_{(1)}(C)\ge \cdots \ge \la_{(p)}(C)\,.
\eeam
Then for any symmetric $p\times p$ matrices $A,B$, by {\em Weyl's inequality} (see Bhatia \cite{bhatia:1997}),
\beao
\max_{i=1,\ldots,p}\big|\la_{(i)}(A+B)-\la_{(i)}(A)\big|\le \|B\|_2\,.
\eeao
If we now choose
$A+B=\X\X'$ and $A= \diag (\X\X')$ we obtain the following result.
\begin{corollary}\label{cor:687ch3}
Under the conditions of Proposition~\ref{prop:offdiagonalch3},
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\diag(\X\X'))\big|\stp 0\,,\quad\nto \,.
\eeao
\end{corollary}
Thus the problem of deriving limit theory for the order statistics of $\X\X'$ has been reduced to limit theory for the order
statistics of the iid row-sums
\beao
D_i^{\rightarrow}= (\X\X')_{ii}=\sum_{t=1}^n X_{it}^2\,,\qquad i=1,\ldots,p\,,
\eeao
which are the eigenvalues of $\diag(\X\X')$. This theory is completely described by the \pp es constructed from the points $D_i^\rightarrow/a_{np}^2$ $i=1,\ldots,p$. Necessary
and sufficient conditions for the weak \con\ of these \pp es are provided by  Lemma~\ref{lem:pprch3}
which in combination with the Nagaev-type \ld\ results of  Theorem~\ref{thm:nagaevch3}
yield the following result; see also Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
\begin{lemma}\label{lem:ppch3}
Assume the conditions of Proposition~\ref{prop:offdiagonalch3} hold. Then
\beao
\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i^\rightarrow-c_n)}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeao
where $(\Gamma_i)$ is defined in \eqref{eq:Gamma} and $c_n=0$ if $\E[D^\rightarrow]=\infty$ and $c_n=\E[D^\rightarrow]$ otherwise.
\end{lemma}
In this result, centering is only needed for $\alpha\in [2,4)$ when $n/a_{np}^2\not\to 0$. Under the additional condition \ref{Cbeta},
%\begin{equation}\label{Cbetaor}
%\beta\in (\alpha/2-1,1], \quad \mbox{ for } \alpha\in [2,4),
%\end{equation}
$n/a_{np}^2\to 0$ in view of the Potter bounds; see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}.
Combining Lemma~\ref{lem:ppch3} and Corollary~\ref{cor:687ch3}, we conclude that Theorem~\ref{thm:intro} holds for $\beta\in [0,1]$.

%In  Davis et al. \cite{davis:mikosch:pfaffel;2015} relation \eqref{eq:2} was derived for $\alpha\in (0,1)$ even for \seq s
%$p_n\to\infty$ \st\ $p_n= O(\ex^{s_n/n})$, where $s_n=o(n)$ but with stronger limitations on the growth of $(p_n)$ for $[1,4]$.
%In the case $\alpha\in (2,4]$ they proved \eqref{eq:center} with $(\la_i)$ replaced by $(\la_i(\X\X'-n\E[ \X\X']))$. It is surprising that
%\eqref{eq:center} also holds since, in general, $\la_i(\X\X'-\E [\X\X'])\not = \la_i-n\,\E X$.

\subsubsection*{Extension to general $\beta$}%\label{subsec:generalb}
Next we explain that it suffices to consider only the case $\beta\in [0,1]$ and how to proceed when $\beta>1$.
The main reason is that
the $p \times p$ sample covariance matrix
$\X\X'$ and the $n \times n$ matrix $\X'\X$  have the same rank and their non-zero eigenvalues coincide; see Bhatia \cite[p.~64]{bhatia:1997}. When proving limit
theory for the eigenvalues  of the sample covariance matrix one may switch to $\X'\X$ and vice versa,
hereby interchanging the roles of $p$ and $n$. By switching to $\X'\X$, one basically replaces $\beta$ by $\beta^{-1}$. Since $\min(\beta,\beta^{-1})\in [0,1]$ for any $ \beta \ge 0$, one can assume without loss of generality that $\beta\in [0,1]$.  This trick allows one to extend results for
$(p_n)$ satisfying \ref{eq:pch3} with $\beta\in [0,1]$  to $\beta>1$. We illustrate this approach by providing the direct analogs of Proposition~\ref{prop:offdiagonalch3} and Corollary~\ref{cor:687ch3}.

\begin{proposition}\label{prop:offdiagonal1}
Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvarch3} for some
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Then for any \seq\ $(p_n)$
satisfying \ref{eq:pch3} with $\beta>1$ we have
\beao
a_{np}^{-2} \twonorm{\X' \X - \diag(\X' \X)}\stp 0\,,\qquad\nto\,,
\eeao
where $\| \cdot\|_2$ denotes the spectral norm.
\end{proposition}
Note that for $\beta>1$ we have $\lim_{\nto} p/n= \infty$. This means that
$\X'\X$ has  asymptotically a much smaller dimension than $\X\X'$ and therefore it is more convenient to work with $\X' \X$
when bounding the spectral norm.
\begin{corollary}\label{cor:6871}
Under the conditions of Proposition~\ref{prop:offdiagonal1},
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,n}\big|\la_{(i)}-\la_{(i)}(\diag(\X'\X))\big|\stp 0\,,\quad\nto \,.
\eeao
\end{corollary}
Now, Theorem~\ref{thm:intro} for $\beta > 1$ is a consequence of Corollary~\ref{cor:6871}.
\par

%-----------------------------------------------------------------------
\section{Introducing dependence between the rows and columns}\label{sec:model}
For details on the results of this section, we refer to Davis et al.~\cite{davis:mikosch:pfaffel:2015}, Heiny and Mikosch \cite{heiny:mikosch:2015:iid} and Heiny et al.~\cite{heiny:mikosch:2016:noniid}.

\subsection{The model}
When dealing with covariance matrices of a multivariate \ts\ $(\X_n)$ it is rather natural to assume dependence
between the entries $X_{it}$.
In this section we introduce a model which allows for {\em linear dependence}
between the rows and columns of $\X$:
\begin{equation}\label{eq:1}
X_{it}=\sum_{l\in \Z}\sum_{k\in \Z} h_{kl} Z_{i-k,t-l}\,,\qquad i,t\in\Z\,,
\end{equation}
where $(Z_{it})_{i,t\in \Z}$ is a field of iid \rv s and $(h_{kl})_{k,l\in\Z}$ is an array of real numbers.
Of course, linear dependence is restrictive in some sense. However, the particular dependence structure allows one to
determine those ingredients in the sample covariance matrix which contribute to its largest eigenvalues.
If the series in \eqref{eq:1} converges a.s. $(X_{it})$ constitutes a strictly stationary random field.
We denote generic elements of the $Z$- and $X$-fields by $Z$ and $X$, respectively. We assume that $Z$ is \regvary\ in the sense that
\begin{equation}\label{eq:27}
\P(Z>x)\sim p_+ \dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad  \P(Z\le -x)\sim p_-
\dfrac{L(x)}{x^{\alpha}}\,,\qquad \xto\,,
\end{equation}
for some tail index $\alpha>0$, constants $p_+,p_-\ge 0$ with $p_++p_-=1$ and a \slvary\ $L$. We will assume $\E[ Z]=0$ whenever $\E [Z^2]<\infty$.
Moreover, we require the summability condition
\begin{equation}\label{eq:2a}
\sum_{l \in \Z} \sum_{k\in \Z} |h_{kl}|^{\delta} <\infty
\end{equation}
for some $\delta\in (0,\min({\alpha/2},1))$ which ensures
the a.s.~absolute convergence of the series in \eqref{eq:1}. Under the conditions \eqref{eq:27} and \eqref{eq:2a}, the marginal and
\fidi s of the field $(X_{it})$ are \regvary\ with index $\alpha$; see
Embrechts et al. \cite{embrechts:kluppelberg:mikosch:1997}, Appendix A3.3. Therefore we also refer to $(X_{it})$ and $(Z_{it})$
as \regvary\ fields.
\par
The model \eqref{eq:1} was introduced by Davis et al. \cite{davis:pfaffel:stelzer:2014}, assuming the rows iid, and in the
present form by Davis et al. \cite{davis:mikosch:pfaffel:2015}.
\subsection{Sample covariance and autocovariance matrices}
From the field $(X_{it})$ we construct the $p\times n$ matrices
\begin{equation}\label{eq:26}
\X_n(s)= (X_{i,t+s})_{i=1,\ldots,p;t=1,\ldots,n}\,,\qquad s=0,1,2,\ldots\,,
\end{equation}
As before, we will write $\X=\X_n(0)$.
Now we can introduce the (non-normalized)
{\em sample autocovariance matrices}
\beam\label{eq:sample}
\X_n(0)\X_n(s)'\,,\qquad s=0,1,2,\ldots\,.
\eeam
We will refer to $s$ as the {\em lag}. For $s=0$, we obtain the {\em sample covariance matrix.}
In what follows, we will be interested in the asymptotic behavior (of \fct s) of the eigen- and singular values of the
sample covariance and autocovariance matrices in the heavy-tailed case. Recall that the {\em singular values} of a matrix $A$ are the square roots of the
eigenvalues of the non-negative definite matrix $AA'$ and its {\em spectral norm} $\twonorm{A}$ is its largest singular value. We notice that $\X_n(0)\X_n(s)'$ is not symmetric
and therefore its eigenvalues can be complex. To avoid this situation, we use
%\beao
%\X_n(0)\X_n(s)' +\X_n(s)\X_n(0)'
%\eeao
%(this matrix is symmetric and therefore it has real eigenvalues) or use
the squares
\beam\label{eq:squares}
\X_n(0)\X_n(s)' \X_n(s)\X_n(0)'
\eeam
%(this matrix is  non-negative definite  and therefore it has non-negative eigenvalues).
whose eigenvalues are the squares of the singular values of $\X_n(0)\X_n(s)'$.
The idea of using the sample autocovariance matrices and \fct s of their squares \eqref{eq:squares}
originates from a paper by Lam and Yao~\cite{lam:yao} who used a model different from \eqref{eq:1}.
This idea is quite natural in the context of \tsa .
\par
In Theorem~\ref{thm:mains} below, we provide a general approximation result for the ordered singular values of the sample
autocovariance matrices in the heavy-tailed case. This result is rather technical. To formulate it we introduce further notation.
As before, $p=p_n$ is any integer \seq\ converging to infinity.
\subsection{More notation}
Important roles are played by the quantities $(Z_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$ and their order statistics
 \begin{equation}\label{eq:zorderch3}
Z_{(1),np}^2 \ge Z_{(2),np}^2 \ge  \ldots \ge Z_{(np),np}^2, \qquad n,p\ge 1\,.
\end{equation}
As important are the row-sums
\begin{equation}\label{eq:llch3}
D_i^\rightarrow=D_i^{(n),\rightarrow}=\sum_{t=1}^n Z_{it}^2\,, \qquad
i=1,\ldots,p;\quad  n=1,2,\ldots\,,
\end{equation} with generic element $D^\rightarrow$ and their ordered values
\beam\label{eq:help6ch3}
D_{(1)}^\rightarrow=D_{L_1}^\rightarrow\ge \cdots \ge D_{(p)}^\rightarrow=D_{L_p}^\rightarrow\,,
\eeam
where we assume without loss of generality that $(L_1,\ldots,L_p)$ is
a permutation of $(1,\ldots,p)$ for fixed~$n$.
\par
Finally, we introduce the column-sums
\beao
D_t^\downarrow=D_t^{(n),\downarrow}= \sum_{i=1}^p Z_{it}^2\,,\qquad t=1,\ldots,n;\; \quad p=1,2,\ldots\,,
\eeao
with generic element $D^\downarrow$ and we also adapt the notation from \eqref{eq:help6ch3} to these quantities.
\subsubsection*{Matrix norms} For any $p\times n $ matrix $\A=(a_{ij})$, we will use the following norms:
\begin {itemize}
\item
{\em Spectral norm:}
\beam\label{specnorm}
\|\A\|_2=\sqrt{\la_{(1)}(\A\A')}\,,
\eeam
\item
{\em Frobenius norm:}
\beao
\frobnorm{\A}= \Big( \sum_{i=1}^p \sum_{j=1}^n |a_{ij}|^2\Big)^{1/2}\,.
\eeao
%\item
%{\em Max-row sum norm:}
%\beao
%\inftynorm{A}= \max_{i=1,\ldots,p} \sum_{j=1}^n |a_{ij}|\,.
%\eeao
%\item
%{\em Max-column sum norm:}
%\beao
%\|{A}\|_1 = \max_{j=1,\ldots,n} \sum_{i=1}^p |a_{ij}|\,.
%\eeao
\end{itemize}
We will frequently make use of the bound $\|\A\|_2\le \|\A\|_ F$. Standard references for matrix norms are \cite{belitski,bhatia:1997, horn, shores}.
\subsubsection*{Singular values of the sample autocovariance matrices} Fix integers $n\ge 1$ and $s\ge 0$. We recycle the $\la$-notation
for the singular values
$\la_1(s), \ldots, \la_p(s)$ of the sample autocovariance matrix $\X_n(0)\X_n(s)'$, suppressing the dependence on $n$.
Correspondingly, the  order statistics are denoted by
\begin{equation}\label{eq:sigma}
\la_{(1)}(s) \ge \cdots \ge \la_{(p)}(s)\,.
\end{equation}
When $s=0$ we typically write $\la_i$ instead of $\la_i(0)$.
\subsubsection*{The matrix $\M(s)$} We introduce some auxiliary matrices
derived from the coefficient matrix $\H=(h_{kl})_{k,l\in \Z}$:
\beao
\H(s)=(h_{k,l+s})_{k,l\in \Z}, \qquad \M(s)= \H(0)\H(s)'\qquad s\ge 0\,.
\eeao
Notice that
\begin{equation}\label{eq:m}
(\M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z .
\end{equation}
We denote the ordered singular values of $\M(s)$ by
\begin{equation}\label{eq:v1}
v_1(s) \ge v_2(s) \ge\cdots \,.
\end{equation}
Let $r(s)$ be the rank of $\M(s)$ so that $v_{r(s)}(s)>0$ while $v_{r(s)+1}(s)=0$ if $r(s)$ is
finite, otherwise $v_i(s)>0$ for all $i$. We also write $r=r(0)$.
\par
Under the summability condition \eqref{eq:2a} on $(h_{kl})$ for fixed $s\ge 0$,
\beam\label{eq:tracea}
\sum_{i=1}^\infty (v_i(s))^2 &=&  \frobnorm{\M(s)}^2= \sum_{i,j\in \Z} \sum_{l_1,l_2 \in \Z} h_{i,l_1} h_{j,l_1+s}h_{i,l_2}
h_{j,l_2+s}\nonumber\\
&\le& c \,\Big(\sum_{l_1,l_2 \in \Z} \sum_{i\in \Z} |h_{i,l_1} h_{i,l_2}|\Big)^2
\le c \,\sum_{l_1\in \Z} \sum_{i \in \Z} |h_{i,l_1}| <\infty\,.
\eeam
Therefore all singular values $v_i(s)$ are finite and the ordering
  \eqref{eq:v1}
is justified.
\par
{\em Here and in what follows, we write $c$ for any constant whose value is not of interest.}

\subsubsection*{Normalizing sequence}
We define $(a_k)$ by
\beao
\P(|Z|>a_k)\sim k^{-1}\,,\qquad k\to\infty\,,
\eeao
and choose the normalizing \seq\ for the singular values as $(a_{np}^2)$ for suitable \seq s $p=p_n\to\infty$.
\subsubsection*{Approximations to singular values}\label{subsec:defdelta}
We will give approximations to the singular values $\la_i(s)$  in
terms of the  $p$ largest ordered values for $s\ge 0$,
\beao
&&\delta_{(1)}(s)\ge \cdots \ge \delta_{(p)}(s)\,,\\
&&\gamma_{(1)}^\rightarrow(s)\ge \cdots \ge \gamma_{(p)}^\rightarrow(s)\,,\\
&&\gamma_{(1)}^\downarrow(s)\ge \cdots \ge \gamma_{(n)}^\downarrow(s)\,,
\eeao
from the sets
\beao
&&\big\{Z_{(i),np}^2 v_j(s)\,, i=1,\ldots,p\,;j=1,2,\ldots\big\}\,,\notag\\
&&\big\{D_{i}^\rightarrow v_j(s), i=1,\ldots,p\,;j=1,2,\ldots\big\}\,,\notag\\
&&\big\{D_{t}^\downarrow v_j(s), t=1,\ldots,n\,;j=1,2,\ldots\big\}\,,
\eeao
respectively.
%---------------------------------------------------------------------------
\subsection{Approximation of the singular values}\label{sec:mainresult}
%Furthermore, let $\tau_{(1)}(s)  \ge \ldots \ge \tau_{(p)}(s)$ be  the singular values of $\tfrac{1}{2}(\X_n(0)\X_n(s)'+\X_n(s)\X_n(0)' )$ and let $\tilde{\tau}_{(1)}(s) \ge  \ldots \ge\tilde{\tau}_{(p)}(s)$ be the singular values of
%\begin{equation*}
%\tfrac{1}{2}(\X_n(0)\X_n(s)'-\E[\X_n(0)\X_n(s)']+\X_n(s)\X_n(0)' -\E[\X_n(s)\X_n(0)'])
%\end{equation*}
In the following result we povide some useful approximations to the singular values of the sample autocovariance matrices of
the linear model \eqref{eq:1}.
\begin{theorem}\label{thm:mains}
Consider the linear process \eqref{eq:1} under
\begin{itemize}
\item
the \regvar\ condition \eqref{eq:27}
for some $\alpha\in (0,4)$,
\item the centering condition
$\E[Z]=0$ if $\E[|Z|]<\infty$,
\item
the summability condition
\eqref{eq:2a} on the coefficient matrix  $(h_{kl})$,
\item
the growth condition \ref{eq:pch3} on $(p_n)$ for some $\beta\ge 0$.
\end{itemize}
Then the following statements hold for $s\ge 0$:
\begin{enumerate}
\item We consider two disjoint cases:
$\alpha \in (0,2)$ and $\beta\in (0,\infty)$, or
$\alpha\in [2,4)$ and $\beta$ satisfying \ref{Cbeta}. Then
\begin{equation}\label{eq:mains1}
a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\delta_{(i)}(s)| \cip 0, \quad \nto.
\end{equation}
\item
Assume $\beta\in [0,1]$.
If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta\in (\alpha/2-1,1]$ then
\beao
a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\gamma_{(i)}^\rightarrow(s)| \cip 0, \quad \nto.
\eeao
Assume $\beta>1$. If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta^{-1}\in(\alpha/2-1,1]$. Then
\beao
a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\gamma_{(i)}^\downarrow(s)| \cip 0, \quad \nto.
\eeao
\end{enumerate}
\end{theorem}
\begin{remark}\em
The proof of Theorem~\ref{thm:mains} is given in Heiny et al.~\cite{heiny:mikosch:2016:noniid}. Part (2) of this  result
with more restrictive conditions on the growth rate of $(p_n)$ is contained in Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
These proofs are very technical and lengthy.
\end{remark}
\begin{remark}\em
If we consider a random array $(h_{kl})$ independent of  $(X_{it})$ and
assume that the summability
condition \eqref{eq:2a} holds a.s.
then Theorem~\ref{thm:mains} remains valid conditionally on
$(h_{kl})$, hence  unconditionally in $\P$-probability; see also \cite{davis:mikosch:pfaffel:2015}.\end{remark}
\subsection{Point process \con }\label{sec:lkasfj}
Theorem~\ref{thm:mains}  and arguments similar to the proofs in Davis et al. \cite{davis:mikosch:pfaffel:2015}
enable one to derive the weak \con\ of
the point processes of the normalized singular values. Recall
the representation of the points $(\Gamma_i)$ of a unit rate homogeneous Poisson process on $(0,\infty)$
given in \eqref{eq:Gamma}. For $s\ge 0$, we define the point processes of the normalized singular values:
\begin{equation}\label{eq:ppdef}
N_n^{\lambda,s}=\sum_{i=1}^p \vep_{a_{np}^{-2}(\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s))} \,.
\end{equation}

\begin{theorem}\label{cor:1ch3}
Assume the conditions of Theorem~\ref{thm:mains}.
Then $(N_n^{\lambda,s})$ converge weakly in the space of point measures
with state space $(0,\infty)^{s+1}$ equipped with the vague topology.
If either $\alpha \in (0,2]$, $\E[Z^2]=\infty$ and $\beta \ge 0$,
or $\alpha \in [2,4)$, $\E[Z^2]<\infty$ and \ref{Cbeta} hold then
\begin{equation}\label{eq:pp}
N_n^{\lambda,s} \cid N= \sum_{i=1}^\infty
\sum_{j=1}^{\infty} \vep_{\Gamma_i^{-2/\alpha} (v_j(0),\ldots,v_j(s))}, \qquad \nto.
\end{equation}
\end{theorem}

\begin{proof}
Regular variation of $Z^2$ is equivalent to
\begin{equation}\label{eq:wwa}
n\,p\, \P(a_{np}^{-2} Z^2  \in \cdot ) \civ \mu(\cdot),
\end{equation}
where $\civ$ denotes vague convergence of Radon measures on  $(0,\infty)$ and the \ms\ $\mu$ is given by $\mu(x,\infty)=x^{-\alpha/2}$, $x>0$.
In view of Resnick \cite{resnick:1987}, Proposition~3.21, \eqref{eq:wwa} is equivalent
to the weak \con\ of the following \pp es:
\begin{equation*}
\sum_{i=1}^p\sum_{t=1}^n \vep_{a_{np}^{-2} Z_{it}^2}=\sum_{i=1}^{np} \vep_{a_{np}^{-2} Z^2_{(i),np}}\cid
\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha} }=\widetilde{N}\,,\quad
\nto\,,
\end{equation*}
where the limit $\widetilde{N}$ is a Poisson random measure (PRM) with state space $(0,\infty)$ and
mean measure~$\mu$.
\par
Since $a_{np}^{-2} Z^2_{(p),np}\cip 0$ as $\nto$,
the point processes $\sum_{i=1}^{p} \vep_{a_{np}^{-2} Z^2_{(i),np}}$ converge weakly to the same PRM:
\begin{equation}\label{eq:ppc}
\sum_{i=1}^{p} \vep_{a_{np}^{-2} Z^2_{(i),np}}\cid
\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha} }\,,\quad
\nto\,.
\end{equation}
A continuous mapping argument together with the fact that $\sum_{i=1}^\infty (v_i(s))^2<\infty$ (see  \eqref{eq:tracea})
shows that
\begin{equation*}
\sum_{j=1}^{ \infty}\sum_{i=1}^p\vep_{a_{np}^{-2} Z^2_{(i),np}(v_j(0),\ldots,v_j(s))}\cid
\sum_{j=1}^{ \infty}\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha}  (v_j(0),\ldots,v_j(s))}\,.
\end{equation*}
If the assumptions of part (1) of Theorem~\ref{thm:mains} are satisfied
an application of \eqref{eq:mains1} (also recalling the definition of $(\delta_{(i)}(s))$) shows that \eqref{eq:ppc} remains
valid with the points  $(a_{np}^{-2} Z^2_{(i),np}(v_j(0),\ldots,v_j(s)))$ replaced by
$(a_{np}^{-2} (\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s))$.
\par
The only cases which are not covered by Theorem~\ref{thm:mains}(1) are $\alpha\in (0,2)$, $\beta=0$ and $\alpha=2$, $\E[Z^2]=\infty$, $\beta\ge0$.
In these cases we get from Theorem~\ref{thm:nagaevch3} that
\begin{equation*}
 p \, \P(a_{np}^{-2} D^\rightarrow>x ) \sim p\,n\, \P( Z^2>a_{np}^{2} x )\to \mu(x,\infty)\,,\quad x>0\,,
\end{equation*}
i.e., $
 p \, \P(a_{np}^{-2} D^\rightarrow\in \cdot ) \stackrel{v}{\rightarrow} \mu(\cdot)$.
It follows from Lemma~\ref{lem:pprch3} that
$\sum_{i=1}^p\vep_{a_{np}^{-2} D_i^\rightarrow}\std \widetilde{N}$. As before,
a continuous mapping argument in combination with
the approximation obtained in Theorem~\ref{thm:mains}(2)
justifies the replacement of the points $(a_{np}^{-2} D_{(i)}^\rightarrow(v_j(0),\ldots,v_j(s)))$ by
$(a_{np}^{-2} (\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s)))$ in the case $\beta\in [0,1]$. If $\beta>1$ one has to work with the
quantities $(D_i^\downarrow)_{i=1,\ldots,n}$ instead of  $(D_i^\rightarrow)_{i=1,\ldots,p}$ and one may follow the same argument as above.
This finishes the proof.
\end{proof}

\section{Some applications}
\subsection{Sample covariance matrices}\label{sec:samplecov}%\setcounter{equation}{0}
The sample covariance matrix $\X_n(0)\X_n(0)'=\X\X'$ is a non-negative definite matrix. Therefore
its eigenvalues and singular values coincide.  Moreover, $v_j=v_j(0)$, $j\ge 1$, are the eigenvalues of $\M=\M(0)$.
\par
Theorem~\ref{thm:mains}(1) yields an approximation of the ordered eigenvalues $(\la_{(i)})$ of $\X\X'$ by
the quantities $(\delta_{(i)})$ which are derived from the order statistics of $(Z_{it}^2)$.
Part (2) of this result provides an approximation of $(\la_{(i)})$ by the quantities
$(\gamma_{(i)}^{\rightarrow/\downarrow})$ which are derived from the order statistics
of the partial sums $(D_i^{\rightarrow/\downarrow})$.
\par
In the following example we illustrate the
quality of the two approximations.
\begin{example}\label{ex:xiao}\em
We choose a Pareto-type \ds\ for $Z$ with density
\beam\label{eq:distrsimch3}
f_Z(x) =
\left\{\begin{array}{cc}
 \frac{\alpha}{(4|x|)^{\alpha + 1}}\,, & \mbox{if } |x| > 1/4 \\
1\,, & \mbox{otherwise.}
\end{array}\right.
\eeam
We simulated $20,000$ matrices $\X_n$ for $n=1,000$ and $p=200$ whose
iid entries have this density. We assume $\beta=1$.
Note that $\M=\M(0)$ has rank one and $v_1=1$.
The estimated densities of the deviations $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ and
$a_{np}^{-2}(\lambda_{(1)}-Z^2_{(1),np})$  based on the simulations are shown in
Figure~\ref{fig:lambda_comparisonch3}. The approximation error is very small indeed.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.40]{lambda_comparison.pdf}
  \caption{Density of the approximation errors for the eigenvalues of $a_{np}^{-2}\X\X'$.
The entries of $\X$ are iid with density \eqref{eq:distrsimch3} and $\alpha=1.6$.}
  \label{fig:lambda_comparisonch3}
\end{figure}
According to the theory,
\beao
a_{np}^{-2}\sup_i|D_{(i)}^\rightarrow-\lambda_{(i)}| +a_{np}^{-2}\sup_i |Z^2_{(i),np}-\lambda_{(i)}|\stp 0\,,
\eeao
but  for finite $n$ the $(D_{(i)}^\rightarrow)$ sequence yields a better approximation to $(\la_{(i)})$. By construction,
the considered differences
have a tendency to be positive but Figure~\ref{fig:lambda_comparisonch3} also shows that the median of the
approximation error for $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ is almost zero.
\end{example}
\par
Theorem~\ref{cor:1ch3} and the continuous mapping theorem immediately yield results about
the joint convergence of the largest eigenvalues of the matrices $a_{np}^{-2}\X_n\X_n'$ for $\alpha\in (0,2)$ and
$\alpha\in (2,4)$ when $\beta$ satisfies \ref{Cbeta}.
For fixed $k\ge 1$ one gets
\begin{equation*}
a_{np}^{-2} \big(\la_{(1)},\ldots,\la_{(k)}\big)\cid
\big(d_{(1)},\ldots, d_{(k)}\big)\,,
\end{equation*}
where $d_{(1)}\ge \cdots\ge d_{(k)}$ are the $k$ largest ordered
values of the set $\{\Gamma_i^{-2/\alpha} v_j, i=1,2,\ldots,j=1,\ldots,r\}$.
The continuous mapping theorem yields for $k\ge 1$,
\begin{equation}\label{eq:u}
\dfrac{\la_{(1)}}{\la_{(1)}+\cdots+ \la_{(k)}}\cid
\dfrac{d_{(1)}}{d_{(1)}+\cdots+ d_{(k)}}\,,\quad \nto\,.
\end{equation}

An application of the continuous mapping theorem to  the distributional convergence of the point processes in Theorem~\ref{cor:1ch3} in the spirit of Resnick
\cite{resnick:2007}, Theorem 7.1, also yields the following result; see Davis et al. \cite{davis:mikosch:pfaffel:2015} for a proof and a similar result in the case $\alpha \in (2,4)$.
\begin{corollary}\label{cor:1q} Assume the conditions of Theorem~\ref{thm:mains}.
If $\alpha \in (0,2]$ and $\E[Z^2]=\infty$, then
\begin{equation*}
a_{np}^{-2}\Big(\la_{(1)},\sum_{i=1}^p \la_i\Big) \cid
\Big( v_1\,\Gamma_1^{-2/\alpha}\,,\sum_{j=1}^r v_j\,\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}\Big)\,,
\end{equation*}
where $\Gamma_1^{-2/\alpha}$ is Fr\'echet $\Phi_{\alpha/2}$-distributed.
and $\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}$ has the distribution
of a positive $\alpha/2$-stable random variable.
In particular,
\begin{equation}\label{eq:limit}
\dfrac{\la_{(1)}}{\la_{1}+\cdots+ \la_{p}}\cid
\dfrac{v_1}{\sum_{j=1}^r v_j}\;
\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty\Gamma_i^{-2/\alpha}}\,,\quad
\nto\,.
\end{equation}
\end{corollary}

\begin{remark}\label{rem:4.5}\em
The ratio
\begin{equation*}
\dfrac{\la_{(1)}+\cdots +\la_{(k)}}{\la_1+\cdots+\la_p}, \qquad k\ge 1\,,
\end{equation*}
plays an important role in PCA. It reflects the proportion of the total variance in the data that we can explain by the first $k$ principal components.
It follows from Corollary~\ref{cor:1q} that for fixed $k\ge 1$,
\begin{equation*}
\dfrac{\la_{(1)}+\cdots \la_{(k)}}{\la_1+\cdots+\la_p}\cid
\dfrac{d_{(1)}+\cdots +d_{(k)}} {d_{(1)}+d_{(2)}+\cdots}\,.
\end{equation*}
Unfortunately, the limiting variable does in general not have a clean
form.
An exception is the case when $r=1$; see Example~\ref{exam:separable}.
Also notice that the trace of $\X\X'$ coincides with $\la_1+\cdots+\la_p$.
\end{remark}

To illustrate the theory we consider a simple moving average example
taken from Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
\begin{example} \rm
\begin{figure}[htb!]
  \centering
  \subfigure[iid data]{
    \includegraphics[scale=0.34]{lambda_1_2_Gaps_iid.pdf}
  }
  \subfigure[data from model \eqref{ex:4.6}]{
    \includegraphics[scale=0.34]{lambda_1_2_Gaps.pdf}
  }
  \caption{Distribution function of $(\lambda_{(1)} - \lambda_{(2)})/\lambda_{(1)}$ for iid data (left) and
    data generated from the model \eqref{ex:4.6} (right). In each graph we compare the empirical \ds\ \fct\
(dotted line, based on
1000 simulations of $200 \times 1000$ matrices with $Z$-\ds\ \eqref{eq:distrsimch3}) with the theoretical curve (solid line).}
  \label{fig:ProbMass}
\end{figure}
Assume that $\alpha\in (0,2)$ and
\begin{equation}\label{ex:4.6}
X_{it}= Z_{it}+ Z_{i,t-1}-2 (Z_{i-1,t}- Z_{i-1,t-1})\,,\quad i,t\in\Z\,.
\end{equation}
In this case, the non-zero entries of $\H$ are
\begin{equation*}
h_{00}=1, h_{01}=1,h_{10}=-2 \quad \mbox{ and }\quad h_{11}=2.
\end{equation*}
Hence $\M=\H\H'$ has the positive eigenvalues
$v_1=8$ and $v_2=2$. The limit point process in \eqref{eq:pp} is
\begin{equation*}
N=\sum_{i=1}^\infty \vep_{8\Gamma_i^{-2/\alpha}}+ \sum_{i=1}^\infty \vep_{2\Gamma_i^{-2/\alpha}}\,,
\end{equation*}
so that
\begin{equation*}
a_{np}^{-2}\,\big(\la_{(1)},\la_{(2)}\big) \cid
\big(8\Gamma_{1}^{-2/\alpha}, 2\Gamma_1^{-2/\alpha}\vee 8
\Gamma_2^{-2/\alpha}\big)\,.
\end{equation*}
Using the fact that $U=\Gamma_1/\Gamma_2$ has a uniform \ds\ on $(0,1)$ we calculate
\begin{equation*}
\P(2\Gamma_1^{-2/\alpha}>8\Gamma_2^{-2/\alpha})= \P(\Gamma_1/\Gamma_2<2^{-\alpha})= 2^{-\alpha} \in (1/4,1).
\end{equation*}
In particular, we have for the normalized spectral gap
\begin{equation*}
a_{np}^{-2} \big(\la_{(1)}-\la_{(2)}\big)\cid
6 \,\Gamma_1^{-2/\alpha} \1_{\{\Gamma_1 4^{\alpha/2 }<\Gamma_2\}}+
8\,
\big(\Gamma_1^{-2/\alpha}-\Gamma_2^{-2/\alpha}\big)\1_{\{\Gamma_14^{\alpha/2
  }> \Gamma_2\}}
\end{equation*}
and for the self-normalized spectral gap (see also Example~\ref{exam:spectralgap} for a detailed analysis)
\begin{equation*}
\begin{split}
\dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}}&\cid
\dfrac{6}{8} \, \1_{\{\Gamma_1 2^{\alpha }<\Gamma_2\}}+
\big(1-(\Gamma_1/\Gamma_2)^{2/\alpha}\big)\1_{\{\Gamma_12^{\alpha
  }> \Gamma_2\}}\\
& =\dfrac{3}{4}\,  \1_{\{U 2^{\alpha }<1\}}+
\big(1-U^{2/\alpha}\big)\1_{\{U2^{\alpha
  }> 1\}}=Y\,.
\end{split}
\end{equation*}
The limit  distribution of the spectral gap
has an atom at $3/4$ with probability $2^{-\alpha}$, i.e.,~$\P(Y=3/4)=2^{-\alpha}$, and
\begin{equation*}
\P(Y\le x)= 1- (1-x)^{\alpha/2},\qquad x\in (0,3/4).
\end{equation*}
In the iid case the limit distribution of the self-normalized spectral gap has distribution function $F(x) = 1- (1-x)^{\alpha/2}$ for $x\in [0,1]$. This means that the atom disappears if the entries are iid. Figure~\ref{fig:ProbMass} compares the distribution function of $Y$
with $F$ for $\alpha=0.6$; the atom at $3/4$ is clearly visible.

Along the same lines, we also have
\begin{equation*}
(a_{np}^{-2}\lambda_{(1)},\lambda_{(2)}/\lambda_{(1)}) \cid(
8\,\Gamma_1^{-2/\alpha},\frac{1}{4}\,\1_{\{U<2^{-\alpha}\}}+U^{2/\alpha}\,\1_{\{U\ge
  2^{-\alpha}\}})\,
\end{equation*}
and hence the limit distribution of $\lambda_{(2)}/\lambda_{(1)}$ is supported on $[1/4,1)$ with  mass of $2^{-\alpha}$ at $1/4$.  The histogram of the ratio $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ based on 1000 replications from the model \eqref{ex:4.6} with noise given by a $t$-distribution with $\alpha=1.5$ degrees of freedom, $n=1000$ and $p=200$ is displayed in Figure \ref{fig:1}.
\begin{figure} [h]
\begin{center}
\includegraphics[scale=.35]{histogram.pdf}
\end{center}
\caption{Histogram based on 1000 replications of $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ from model \eqref{ex:4.6}.}
\label{fig:1}
\end{figure} Observing that $2^{-\alpha}=0.3536\ldots$,
the histogram is remarkably close to what one would expect from a sample from the truncated
uniform \ds , $2^{-\alpha}\, \1_{\{U<2^{-\alpha}\}}+U\, \1_{\{U\ge 2^{-\alpha} \}}$.
The mass of the limiting discrete component of the ratio can be much
larger if one conditions on $a_{np}^{-2}\lambda_{(1)}$ being large.
Specifically, for any $\epsilon\in (0,1/4)$ and $x>0$,
\begin{equation*}
\lim_{n\to\infty} \P(\epsilon<\lambda_{(2)}/\lambda_{(1)}\le 1/4|\lambda_{(1)}>a_{np}^2x)
=\P(\Gamma_1/\Gamma_2\le 2^{-\alpha}|\Gamma_1<(x/8)^{-\alpha/2})=G(x)\,.
\end{equation*}
The function $G$ approaches $1$ as $x\to\infty$ indicating the speed at which the
two largest eigenvalues get linearly related; see Figure \ref{fig:2} for a graph of $G$ in the case $\alpha=1.5$.
\begin{figure} [h]
\begin{center}
\includegraphics[scale=.35]{Hfunction.pdf}
\end{center}
\caption{Graph of $G(x)=\P(\Gamma_1/\Gamma_2\le2^{-\alpha}|\Gamma_1<(x/8)^{-\alpha/2})$ when $\alpha=1.5$.}
\label{fig:2}
\end{figure}
In addition, from Remark~\ref{rem:4.5}, we also have
\begin{equation*}
\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \cid \dfrac{4}{5}\,
\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}}\,.
\end{equation*}
Clearly, the limit \rv\ is stochastically smaller than what one would get in the iid case; see \eqref{eq:limit}.
\end{example}
\begin{example}\label{exam:spectralgap}\rm
The previous example also illustrates the behavior of the two largest eigenvalues in the general case
when the rank $r$ of the matrix $\M$ is larger than one.
We have in general
\begin{equation*}
\dfrac{\lambda_{(2)}}{\lambda_{(1)}}\cid
\frac{v_2}{v_1}\,\1_{\{U<(v_2/v_1)^{\alpha/2}\}}+U^{2/\alpha}\,\1_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
\end{equation*}
In particular, the limiting
{\em self-normalized spectral gap} has \rep
\beao
\dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}} \cid \dfrac{v_1-v_2}{v_1}\,\1_{\{U<(v_2/v_1)^{\alpha/2}\}}+(1-U^{2/\alpha})\,\1_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
\eeao
The limiting variable assumes values in $(0,1-v_2/v_1]$  and has an atom at the right end-point.
This is in contrast to the iid case
and to the case when $r=1$ (hence $v_2=0$) including the case of iid rows and the separable case; see Example~\ref{exam:separable}.
\end{example}
\begin{example}\label{exam:separable}
{\em We consider the separable case when  $h_{kl}=\theta_kc_l$, $k,l \in \Z$, where $(c_l)$, $(\theta_k)$ are
  real sequences such that the conditions on $(h_{kl})$ in
  Theorem~\ref{thm:mains} hold. In this case,
\begin{equation*}
\M= \sum_{l\in \Z} c_l^2\;\; (\theta_i\theta_j)_{i,j \in \Z}\,.
\end{equation*}
Note that $r=1$ with the only non-negative eigenvalue
\beao
v_1=\sum_{l\in \Z} c_l^2\;\;\sum_{k\in \Z}\theta_k^2\,.
\eeao
In this
case, the limiting point process in Theorem~\ref{cor:1ch3} is a PRM on
$(0,\infty)$ with mean measure of $(y,\infty)$ given by $(v_1/
y)^{\alpha/2}$, $y>0$. The normalized eigenvalues have similar asymptotic behavior as
in the case of iid entries. For example, the log-spacings have the same limit as in the iid case for fixed $k$,
\beao
\big(\log \la_{(1)}-\log \la_{(2)},\ldots,\log \la_{(k+1)}-\log \la_{(k)}\big)\std
-\dfrac{2}{\alpha}\,\big(\log(\Gamma_1/\Gamma_2),\ldots,\log (\Gamma_{k}/\Gamma_{k+1})\big)\,.
\eeao
The same observation applies to the ratio of the largest eigenvalue and the trace in the case $\alpha\in (0,2)$:
\beao
\dfrac{\la_{(1)}}{{\rm tr} (\X \X')}=\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \std
\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}}\,.
\eeao
We also mentioned in Example~\ref{exam:spectralgap} that the \ds al limit of the self-normalized spectral gap has no atom
as in the iid case.
}
\end{example}

\subsection{S\&P 500 data}\label{sec:sp500}
We conduct a short analysis of the largest eigenvalues of the univariate log-return time series which compose
the S\&P 500 stock index; see Section~\ref{subsec:1.2} for a description of the data.
Although there is strong empirical evidence that these univariate series have power-law tails
(see Figure~\ref{fig:SP500_tail_indices}) we  do not expect that they
have the same tail index. One way to proceed would be to ignore this fact because the tail indices are
in a close range and the differences are due to large sampling errors for estimating such quantities.  One could also collect \ts\ with similar
tail indices in the same group.
In this case, the dimension $p$ decreases. This grouping would be a rather arbitrary classification method.
We have chosen a third way: to use rank transforms.
This approach has its merits because it aims at standardizing the tails but it also has a major disadvantage: one destroys
the covariance structure underlying the data.
\par
Given a $p\times n$ matrix $(R_{it})_{i = 1, \cdots, p;t=1,\cdots, n}$,
we construct a matrix $\X$ via the rank transforms
\begin{equation*}
  X_{it} = -\Big[
    \log \Big(\frac{1}{n+1} \sum_{\tau=1}^n \1_{\{R_{i\tau} \leq R_{it}\}} \Big)
  \Big]^{-1} \,,\qquad i=1,\ldots,p;t=1,\ldots,n\,.
\end{equation*}
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.550]{EigenRatioSP500_ranks_log_50_shown.pdf}
  \caption{The logarithms of the ratios $\lambda_{(i+1)} / \lambda_{(i)}$ for the S\&P 500 series after rank transform.
We also show the 1, 50 and 99\% quantiles (bottom, middle, top lines, respectively) of the variables
$\log((\Gamma_i/ \Gamma_{i+1})^{2})$. }
  \label{fig:EigenRatio}
\end{figure}
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.550]{EigenRatioSP500_log_50_shown.pdf}
  \caption{The logarithms of the ratios $\lambda_{(i+1)} / \lambda_{(i)}$ for the original (non-rank transformed) S\&P 500 log-return data.
We also show the 1, 50 and 99\% quantiles (bottom, middle, top lines, respectively) of the variables
$\log((\Gamma_i/ \Gamma_{i+1})^{2/2.3})$; see also Figure~\ref{fig:EigenRatio} for comparison.}\label{fig:22}
\end{figure}

%\begin{figure}[htb!]
%  \centering
%    \includegraphics[scale=0.550]{EigenRatioSP500_lin2_250_shown.pdf}
%  \caption{{\green Same as Figure \ref{fig:EigenRatio} but on original scale, no logarithm. Choose one.}}
%\end{figure}

If the rows $R_{i1}, \ldots, R_{in}$ were iid (or, more generally, stationary ergodic)
with a continuous \ds\  then
the averages under the logarithm would be  asymptotically uniform
on $(0,1)$ as $n \to \infty$. Hence $X_{it}$ would  be \asy ally standard
Fr\'echet $\Phi_1$-distributed.
In what follows, we assume that the aforementioned univariate \ts\ of the S\&P 500 index have undergone the rank transform and that
their marginal \ds s are close to $\Phi_1$; we
always use the symbol $\X$ for the resulting multivariate series.
\par
In Figure~\ref{fig:EigenRatio} we show the ratios of the consecutive ordered eigenvalues $(\la_{(i+1)}/\la_{(i)})$
of the matrix $\X\X'$. This graph shows the rather surprising fact that the ratios are close to one even for small values $i$.
We also show the 1, 50 and 99 \% quantiles of the variables $((\Gamma_{i}/\Gamma_{i+1})^{2/\alpha})$ calculated from the formula
\beam\label{eq:iid}
\P\big((\Gamma_{i} /\Gamma_{i+1})^{2/\alpha}\le x \big) = x^{i \cdot \alpha/2}, \quad x\in (0,1)\,.
\eeam
For increasing $i$, the \ds\ is concentrated closely to 1, in agreement with the \slln\ which yields
$\Gamma_{i} /\Gamma_{i+1}\stas 1$ as $i\to\infty$.
The \asy\ \ds s \eqref{eq:iid} correspond to the case when the matrix
$\M$ has rank $r=1$. It includes the iid and separable cases;
see Example~\ref{exam:separable}. The shown \asy\ quantiles are in agreement with the rank $r=1$ hypothesis.
\par
For comparison, in Figure~\ref{fig:22} we also show the ratios  $(\la_{(i+1)}/\la_{(i)})$ for the 
non-rank transformed S\&P 500 data and the 1, 50 and 99\% quantiles of the variables
$\log((\Gamma_i/ \Gamma_{i+1})^{2/\alpha})$, where we choose $\alpha=2.3$ motivated by the estimated tail indices in 
Figure~\ref{fig:SP500_tail_indices}.
The two graphs in Figure~\ref{fig:EigenRatio} and Figure~\ref{fig:22} are quite similar but the smallest ratios for the original data are slightly larger than
for the rank-transformed data.

%--------------------------------------------------------------------------------------
\subsection{Sums of squares of sample autocovariance matrices}\label{sec:possemidef}
In this section we consider some additive \fct s
of the squares of $\A_n(s)=\X_n(0)\X_n(s)'$ given by $\A_n(s)\A_n(s)'$ for $s=0,1,\ldots$. By definition of the singular values of a matrix
(see  \eqref{eq:sigma}), the non-negative definite
matrix $\A_n(s)\A_n(s)'$ has eigenvalues $(\la_i^2(s))_{i=1,\ldots,p}$.
\par
The following result is a corollary of Theorem~\ref{thm:mains}.
\begin{proposition}\label{thm:mainstr} Consider the linear process \eqref{eq:1} under
the conditions of Theorem~\ref{thm:mains}. Then the following statements hold for $s\ge 0$:
\begin{enumerate}
\item[$(1)$]
We consider two disjoint cases:
$\alpha \in (0,2)$ and $\beta\in (0,\infty)$, or
$\alpha\in [2,4)$ and $\beta$ satisfying \ref{Cbeta}. Then
\beao
a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-\delta_{(i)}^2(s)| \cip 0, \quad \nto.
\eeao
\item[$(2)$]
Assume $\beta\in [0,1]$.
If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta \in (\alpha/2-1,1]$, then
\beao
a_{np}^{-4} \max_{i=1,\ldots,p} |\la_{(i)}^2(s)-(\gamma_{(i)}^\rightarrow(s))^2| \cip 0, \quad \nto.
\eeao
Assume $\beta>1$. If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta^{-1} \in (\alpha/2-1,1]$. Then
\beao
a_{np}^{-4} \max_{i=1,\ldots,p} |\la_{(i)}^2(s)-(\gamma_{(i)}^\downarrow(s))^2| \cip 0, \quad \nto.
\eeao
\end{enumerate}
\end{proposition}
To the best of our knowledge, sums of squares of sample autocovariance matrices were used first in the paper by Lam and Yao
\cite{lam:yao}; their \ts\ model is quite different from ours.
\begin{proof}
Part (1).
The proof follows from Theorem~\ref{thm:mains} if we can show that
\beao
a_{np}^{-2}\max_{i=1,\ldots,p} \big(\la_{(i)}(s)+\delta_{(i)}(s) \big)=O_\P(1)\,\quad \nto\,.
\eeao
We have by Theorem~\ref{cor:1ch3},
\beam \label{eq:gdfg}
a_{np}^{-2}\max_{i=1\ldots,p} \la_{(i)}(s)=a_{np}^{-2}\la_{(1)}(s) \cid c\, \xi_{\alpha/2}\,,
\eeam
where $\xi_{\alpha/2}$ has a $\Phi_{\alpha/2}$ \ds . In view of Theorem~\ref{thm:mains}(1) we also have
\beao
a_{np}^{-2}\max_{i=1\ldots,p} \delta_{(i)}(s)\cid c\, \xi_{\alpha/2}\,.
\eeao
Therefore, again using Theorem~\ref{thm:mains}(1), we have
\beao
\lefteqn{a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-\delta_{(i)}^2(s)|}\\
&\le &\big[a_{np}^{-2} \max_{i=1,\ldots,p} |\lambda_{(i)}(s)-\delta_{(i)}(s)|\big]\,
\big[a_{np}^{-2} \max_{i=1,\ldots,p}\big ( |\lambda_{(i)}(s)|+|\delta_{(i)}(s)|\big)\big]
\cip 0, \quad \nto.
\eeao
This proves part (1).\\[1mm]
Part (2). Now assume $\beta\in [0,1]$ and $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta\in (\alpha/2-1,1]$. Then \eqref{eq:gdfg} is still true and we have by Theorem~\ref{thm:mains}(2) and Theorem~\ref{cor:1ch3}
\beao
a_{np}^{-2}\max_{i=1\ldots,p} \gamma_{(i)}^{\rightarrow}(s)\cid c\, \xi_{\alpha/2}\,.
\eeao
We then have
\beao
\lefteqn{a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-(\gamma_{(i)}^\rightarrow(s))^2|}\\
&\le &\big[a_{np}^{-2} \max_{i=1,\ldots,p} |\lambda_{(i)}(s) -\gamma_{(i)}^\rightarrow(s)|\big]\,
\big[a_{np}^{-2} \max_{i=1,\ldots,p}\big ( \lambda_{(i)}(s)+\gamma_{(i)}^\rightarrow(s)
\big)\big]
\cip 0\,, \, \nto.
\eeao
The proof of the remaining part is similar and therefore omitted.
\end{proof}
Now, using Proposition~\ref{thm:mainstr} and a continuous mapping argument, we can show
limit theory for the eigenvalues
\beao
w_{(1)}(s_0,s_1)\ge \cdots \ge w_{(p)}(s_0,s_1)\,,\qquad 0\le s_0\le s_1\,,
\eeao
of the non-negative definite random matrices
\begin{equation}\label{eq:sumA}
\sum_{s=s_0}^{s_1} \A_n(s) \A_n(s)'\,.
\end{equation}

\begin{proposition}\label{prop:sumsmal}
Assume $0\le s_0\le s_1$ and the conditions of Theorem~\ref{thm:mains} hold.
If $\alpha \in (0,4)$ and $\beta\in (0,1] \cap  (\alpha/2-1,1]$ then
\begin{equation*}
a_{np}^{-4} \max_{i=1,\ldots,p} |w_{(i)}(s_0,s_1)-\omega_{(i)}(s_0,s_1)| \cip 0, \quad \nto,
\end{equation*}
where $\omega_{(i)}(s_0,s_1)$ are the ordered values of the set $\{Z_{(i),np}^4 v_j(s_0,s_1), i=1,\ldots,p;j=1,2,\ldots\}$
and $(v_j(s_0,s_1))$ are the ordered eigenvalues of $ \sum_{s=s_0}^{s_1} \M(s)\M(s)'$.
\end{proposition}

\begin{example}\label{exam:additive}{\em
\begin{figure}[htb!]
  \centering
  \subfigure[]
{
    \includegraphics[scale=0.34]{eigen_sum.pdf}
    \label{fig:LamYao:a}
  }
  \subfigure[] {
    \includegraphics[scale=0.31]{eigen_sum_ratio.pdf}
    \label{fig:LamYao:b}
  }
  \caption{The largest eigenvalues of the sums of the squared autocovariance
    matrices compared with the sums of the largest eigenvalues of these matrices
    for the S\&P 500 data for different values $s_1$. The two values are surprisingly close to each other; mind the scale
of the $y$-axis. We also show their ratios.}
  \label{fig:LamYao}
\end{figure}
Recall the separable case from Example~\ref{exam:separable}, i.e.,
$h_{kl}=\theta_kc_l$, $k,l \ge 0$, where $(c_l)$, $(\theta_k)$ are
real sequences such that the conditions on $(h_{kl})$ in
Theorem~\ref{thm:mains} hold.
%Let $\alpha \in (0,2)$ and set
%\begin{equation*}
%\theta = (\theta_0, \theta_1,\theta_2,\ldots)' \quad \mbox{and} \quad
%c= (c_0,c_1,c_2, \ldots)'.
%\end{equation*}
Write $\Theta_{ij}=\theta_i \theta_j$. It is symmetric and has rank one; the only non-zero
eigenvalue is $\gamma_\theta(0)=\sum_{k=0}^\infty \theta_k^2$. Hence $\Theta$  is non-negative definite.
We get from \eqref{eq:m} that
\begin{equation*}
\M(s)=\gamma_c(s)\, \Theta, \quad s\ge 0\,,
\end{equation*}
where
\beao
\gamma_c(s)=\sum_{l=0}^\infty c_lc_{l+s}\,,\qquad s\ge 0\,.
\eeao
The matrix $\M(s)$ has the only non-zero eigenvalue $\gamma_c(s)\gamma_\theta(0)$.
The factors $(\gamma_c(s))$ can be positive or negative; they constitute the autocovariance \fct\ of a stationary linear process
with coefficients $(c_l)$.
Accordingly, $\M(s)$ is either non-negative or non-positive definite. %This explains why a symmetrization of $\M(s)$ is irrelevant.
Now we consider the non-negative definite matrix
\beao
\sum_{s=s_0}^{s_1} \M(s)\,\M(s)'= \sum_{s=s_0}^{s_1}\gamma_c^2(s)\,\Theta\Theta'\,.
\eeao
This matrix has rank $1$ and its largest eigenvalue is given by
\begin{equation*}
C_{c,\theta}(s_0,s_1)=\sum_{s=s_0}^{s_1}\gamma_c^2(s)\,\gamma_\theta^2(0)\,.
\end{equation*}
An application of
Proposition~\ref{prop:sumsmal} yields that the ordered eigenvalues of the matrix $a_{np}^{-4}\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'$
are uniformly approximated by the quantities
\begin{equation}\label{eq:drtgdfg}
a_{np}^{-4} Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\,,\qquad  i=1,\ldots,p\,.
\end{equation}
Since
\beao
C_{c,\theta}(s_0,s_1)= \sum_{i=s_0}^{s_1} C_{c,\theta}(i,i)
\eeao
one gets the remarkable property that
\beao
&&a_{np}^{-4}\max_{i=1,\ldots,p} \Big| \la_{(i)}\Big(\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'\Big)- Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\Big|\\
&=&a_{np}^{-4}\max_{i=1,\ldots,p} \Big|\sum_{s=s_0}^{s_1}\la_{(i)}(\A_n(s)\A_n(s)')- Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\Big|+o_P(1)\,.
\eeao
In particular, for $s_1\ge s_0$ we get the weak \con\ of the \pp es towards a PRM:
\begin{equation*}
\begin{split}
\sum_{i=1}^p &\varepsilon_{a_{np}^{-4}\Big(\la_{i}\Big(\sum_{s=s_0}^{s_0}\A_n(s)\A_n(s)'\big)\,,\ldots,\la_{i}\big(\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'\big)\Big)}\\
&\std \sum_{i=1}^\infty \varepsilon_{\Gamma_i^{-4/\alpha} \Big(C_{c,\theta}(s_0,s_0),\ldots,C_{c,\theta}(s_0,s_1)\Big)}\,,\quad \nto\,.
\end{split}
\end{equation*}

}
\end{example}
\begin{example}\em
In Figure~\ref{fig:LamYao} we calculate the largest eigenvalues\\
$\la_{(1)}\big(\sum_{s=0}^{s_1}\A_n(s)\A_n(s)'\big)$ for $s_1=0,\ldots,5$ as well as the sums of the largest eigenvalues
$\sum_{s=0}^{s_1}\la_{(1)}\big(\A_n(s)\A_n(s)'\big)$
the log-return series from the S\&P 500 index described in Section~\ref{subsec:1.2}. The data are not rank-transformed.
We notice that
the two values are surprisingly close across the values $s_0=0,\ldots,5$. This phenomenon
could be explained by the structure of the eigenvalues in Example~\ref{exam:additive}.
Also note that the largest eigenvalue $\A_n(0)\A_n(0)'$ makes a major contribution to the values in Figure~\ref{fig:LamYao};
the contribution of the squares $\A_n(s)\A_n(s)'$, $s=1,\ldots,5,$ to the largest eigenvalue of the sum of squares is less substantial.
%Figure~\ref{fig:LamYao}(b) shows the same calculations for the matrices studied in Example~\ref{exam:4.6cont}. Here cancellation is possible (see \eqref{eq:Xrgsdf}) which is indicated by the fact that the red line does not follow the blue line as well as in (a).
\end{example}
%\section{\red Some conclusions?}\setcounter{equation}{0}
%{\em Say something about EVT and largest eigenvalues}

\section{Auxiliary results}\label{appendix:Ach3}

Let $(Z_i)$ be iid copies of $Z$ whose distribution satisfies
\begin{equation*}
\P(Z>x)\sim p_+ \dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad  \P(Z\le -x)\sim p_-
\dfrac{L(x)}{x^{\alpha}} \,,\quad \xto\,,
\end{equation*}
 for some tail index $\alpha>0$,
where $p_+,p_-\ge 0$ with $p_++p_-=1$ and $L$ is a slowly varying function. If $\E[|Z|]<\infty$ also assume $\E[Z]=0$. The product $Z_1Z_2$ is regular varying with the same index $\alpha$ and $\P(|Z_1Z_2|>x)= x^{-\alpha} L_1(x)$, where $L_1$ is slowly varying function different from $L$;
see Embrechts and Goldie \cite{embrechts:goldie:1980}.
Write
\begin{equation*}
S_n=Z_1+\cdots +Z_n\,,\quad n\ge 1,
\end{equation*} and consider a sequence $(a_n)$ such that $\P(|Z|>a_n)\sim n^{-1}$.

\subsection{Large deviation results}
The following theorem can be found in
Nagaev \cite{nagaev:1979} and Cline and Hsing
\cite{cline:hsing:1998} for $\alpha>2$ and $\alpha\le 2$,
respectively; see also  Denisov et al.~\cite{denisov:dieker:shneer:2008}.
\begin{theorem}\label{thm:nagaevch3}
Under the assumptions on the iid sequence $(Z_t)$
given above the following relation holds
\begin{equation*}
\sup_{x\ge c_n}\left|\dfrac{\P(S_n>x )}{n\P(|Z|>x)} -p_+ \right|\to 0\,,
\end{equation*}
where $(c_n)$ is any sequence satisfying $c_n/a_n\to  \infty$ for
$\alpha\le 2$ and $c_n\ge \sqrt{(\alpha-2)n\log n}$ for $\alpha>2$.
\end{theorem}

\subsection{A point process convergence result}
Assume that the conditions at the beginning of Appendix \ref{appendix:Ach3} hold.
Consider a sequence of iid copies $(S_{n}^{(t)})_{t=1,2,\ldots}$
of $S_n$ and the sequence of point processes
\begin{equation*}
N_n= \sum_{t=1}^{p} \vep_{a_{np}^{-1} S_{n}^{(t)}}, \quad n=1,2,\ldots\,,
\end{equation*}
for an integer sequence $p=p_n\to\infty$. We assume that the state space of the
point processes $N_n$ is $\overline{\R}_0=[\R\cup\{\pm \infty\}]\backslash \{0\}$.
\begin{lemma}\label{lem:pprch3}
Assume $\alpha \in (0,2)$ and the
conditions of  Appendix~\ref{appendix:Ach3}
on the iid sequence $(Z_t)$ and the normalizing sequence $(a_n)$. Then the limit relation
$N_n\cid N$ holds in the space of point measures on $\overline{\R}_0$
equipped with the vague topology (see
\cite{resnick:1987,resnick:2007})
for a  Poisson random measure $N$ with state space $\overline{\R}_0$ and intensity measure $\mu_\alpha(dx)=\alpha |x|^{-\alpha-1} (p_+ \1_{\{x>0\}}+ p_- \1_{\{x<0\}}) dx$.
\end{lemma}
\begin{proof}
According to Resnick \cite{resnick:1987}, Proposition 3.21, we need to
show that
$p\, \P(a_{np}^{-1}S_n\in \cdot)\civ \mu_\alpha
$,
where $\civ$ denotes vague convergence of Radon measures on  $\overline{\R}_0$.
Observe that we have $a_{np}/a_n\to\infty$ as $\nto$. This fact and
$\alpha\in (0,2)$ allow one to apply
Theorem~\ref{thm:nagaevch3}:
\begin{equation*}
\dfrac{\P( S_n >x a_{np})}{n\,\P(|Z|>  a_{np})}\to p_+ x^{-\alpha}
\quad \mbox{and}\quad \dfrac{\P( S_n \le -x a_{np})}{n\,\P(|Z|>
  a_{np})}\to p_-\, x^{-\alpha}\,,\quad x>0\,.
\end{equation*}
On the other hand, $n\,\P(|Z|>  a_{np})\sim p^{-1}$ as $\nto$.
This proves the lemma.
\end{proof}

%\section*{Acknowledgments}

%We thank Olivier Wintenberger for reading the manuscript and fruitful discussions.

### 2017-07-18 phd/Chapteriidpaper.tex
Added My PhD thesis.

%iid Chapter
\chapter[Eigenvalues and eigenvectors of heavy-tailed sample covariance matrices with general growth rates: the iid case]{{\huge Eigenvalues and eigenvectors of heavy-tailed sample covariance matrices with general growth rates: the iid case}}\label{ch:iid}
\chaptermark{Heavy-tailed sample covariance matrices}

\begin{center}
\textsc{Johannes Heiny \& Thomas Mikosch\\
{\em Stochastic Process. Appl.\/} (2016), 29.}
\end{center}

\begin{abstract}{In this paper we study the joint \ds al \con\ of the largest eigenvalues of the sample covariance matrix
of a $p$-dimensional \ts\ with iid entries when $p$ converges to infinity together with the sample size $n$. We consider
only heavy-tailed \ts\ in the sense that the entries satisfy some \regvar\ condition which ensures that
their fourth moment is infinite. In this case, Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and 
Auffinger et al.~\cite{auffinger:arous:peche:2009} proved the weak \con\ of the \pp es of the normalized eigenvalues 
of the sample covariance matrix towards an inhomogeneous Poisson process which implies in turn that the largest eigenvalue
converges in \ds\ to a Fr\'echet distributed \rv . They proved these results under the assumption that $p$ and $n$ are proportional
to each other. In this paper we show that the aforementioned results remain valid if $p$ grows at any polynomial rate. The proofs 
are different from those in \cite{auffinger:arous:peche:2009,soshnikov:2004,soshnikov:2006}; we employ large deviation 
techniques to achieve them. The proofs reveal that only the diagonal of the sample covariance matrix is relevant for the \asy\ behavior of the largest eigenvalues and the corresponding eigenvectors which are close to the canonical basis vectors.  
We also discuss extensions of the results to sample autocovariance matrices.
\medskip

\noindent {\bf Keywords:} Regular variation, sample covariance matrix, independent entries,
largest eigenvalues, eigenvectors, point process
  convergence, compound Poisson limit, Fr\'echet distribution.}
\end{abstract}
\newpage

%-------------------------------------------------------------------
\section{Introduction}\label{sec:motivation}
In recent years we have seen a vast increase in the number and sizes of data sets. 
Science  (meteorology, telecommunications, genomics, \ldots), society (social networks, finance, military and civil intelligence, \ldots) and industry 
need to extract valuable information from high-dimensional data sets
which are often too large or 
complex to be processed by traditional means. 
In order to explore the structure of data one often studies the dependence via 
(sample) covariances and correlations. %Here we focus on the {\em sample covariance matrix}.
Often dimension reduction techniques facilitate further analyzes of large data matrices. For example, 
{\em principal component analysis}  (PCA) transforms the data linearly such that only a few of the resulting vectors 
contain most of the variation in the data. These {\em principal component vectors} 
are the eigenvectors associated with the largest eigenvalues of the sample covariance matrix.
\par 
The aim of this paper is to investigate the \asy\ properties of the 
largest {\em eigenvalues} and their corresponding {\em eigenvectors} for
sample covariance matrices of high-dimensional heavy-tailed time series with iid entries.
Special emphasis is given to the case 
%we study the largest {\em eigenvalues} and the corresponding {\em eigenvectors} of the {\em sample covariance matrix} 
when the dimension $p$ and the sample size $n$ tend to infinity simultaneously, not necessarily at the same rate. 
%Classical limit theorems in probability theory and statistics are based on the assumption of finite dimension of data. 
%However, over the last decade data sets gathered by cheap and widely available devices have grown in 
%size exponentially. The question whether classical or large-dimensional limit results are closer to reality is addressed by Bai and Silverstein \cite[Chapter~1]{bai:silverstein:2010}.
\par
Throughout we consider the $p\times n$ data matrix 
\beao
\bfZ =\bfZ_n = \big(Z_{it}\big)_{i=1,\ldots,p;t=1,\ldots,n}\,
\eeao
A column of $\bfZ$ represents an observation of a $p$-dimensional time series.
We assume that the entries $Z_{it}$ are real-valued, independent and identically distributed (iid), 
unless stated otherwise. We write $Z$ for a generic element and assume $\E[Z]=0$ and $\E [Z^2]=1$ 
if the first and second moments of $Z$ are finite, respectively.
We are interested in limit theory for the eigenvalues $\lambda_1,\ldots,\lambda_p$ of the {\em sample covariance matrix} $\bfZ\bfZ'$
and their ordered values 
\beam\label{eq:order}
\la_{(1)}\ge \cdots \ge \la_{(p)}\,.
\eeam
In this notation we suppress 
the dependence of $(\la_i)$ on $n$. We will only discuss the case when $p \to \infty$; for the finite $p$ case 
we refer to \cite{anderson:1963,muirhead}.

\subsection{The light-tailed case}
In random matrix theory a lot of attention has been given to the {\em empirical spectral distribution \fct } of 
the sequence $(n^{-1}\z \z')$:
\begin{equation*}
F_{n^{-1}\z \z'}(x)= \frac{1}{p}\; \# \{ 1\le j\le p : n^{-1} \lambda_j \le x \}, \qquad x\ge 0\,,\quad n\ge1.
\end{equation*}
In the literature convergence results for $(F_{n^{-1}\z \z'})$  
are established under the assumption that $p$ and $n$ grow at the same rate:
\beam\label{eq:gamma}
\dfrac p n\to \gamma\qquad \mbox{for some $\gamma \in (0,\infty)$.}
\eeam 
Suppose that the iid entries $Z_{it}$ have mean $0$ and variance $1$. If \eqref{eq:gamma} holds then, with probability one, 
$(F_{n^{-1}\z \z'})$ converges weakly to the  Mar\v cenko--Pastur law $F_\gamma$. If $\gamma \in (0,1]$,  $F_\gamma$  has density,
\begin{eqnarray}\label{eq:MP}
f_\gamma(x) =
\left\{\begin{array}{ll}
\frac{1}{2\pi x\gamma} \sqrt{(b-x)(x-a)} \,, & \mbox{if } a\le x \le b, \\
 0 \,, & \mbox{otherwise,}
\end{array}\right.
\end{eqnarray}\noindent
where $a=(1-\sqrt{\gamma})^2$ and $b=(1+\sqrt{\gamma})^2$. If $\gamma>1$, the \MP law is a mixture of a point mass at $0$ and the density function $f_{1/\gamma}$ with weights $1-1/\gamma$ and $1/\gamma$, respectively. 
This  mass is intuitively explained by the fact that, with probability $1$, $\min(p,n)$ 
eigenvalues $\lambda_i$ are non-zero. When $n=(1/\gamma) \; p$ and $\gamma >1$ the fraction of non-zero eigenvalues 
is $1/\gamma$ while the fraction of zero eigenvalues is $1-1/\gamma$.
\par
The moment condition $\E[Z^2]<\infty$ is crucial for deriving the Mar\v cenko--Pastur limit law. 
When studying the largest eigenvalues of the sample covariance matrix $\z\z'$ the moment condition $\E[Z^4]<\infty$ plays a similarly 
important role; we assume it in the remainder of this subsection. 
%It will be convenient to write
%\beam\label{eq:order}
%\la_{(1)}\ge \cdots \ge \la_{(p)}
%\eeam
%for the ordered eigenvalues of $\z\z'$ where we suppress the dependence on $n$ in the notation. 
If \eqref{eq:gamma} holds and the iid entries $Z_{it}$ have zero mean and unit variance, %and finite fourth moment,  
Geman~\cite{geman} showed that
\beam\label{eq:geman}
\dfrac {\la_{(1)}}{n} \stas \big(1+\sqrt{\gamma}\big)^2\,,\qquad \nto\,.
\eeam
This means that $\la_{(1)}/n$ converges to the right endpoint of the Mar\v cenko--Pastur law in \eqref{eq:MP}.
Johnstone \cite{johnstone:2001} complemented this \slln\ by the corresponding \clt\ in the special case of iid standard normal 
entries:
\beam\label{eq:tc}
\frac{\lambda_{(1)}-\mu_{n,p}}{\sigma_{n,p}} \cid \xi,
\eeam
where the limiting \rv\ has a {\em Tracy--Widom \ds} of order 1 and the centering and scaling constants are
\begin{equation*}
\mu_{n,p}=(\sqrt{n-1}+ \sqrt{p})^2, \quad \sigma_{n,p}=(\sqrt{n-1}+ \sqrt{p}) \Big( \frac{1}{\sqrt{n-1}} +\frac{1}{\sqrt{p}}\Big)^{1/3}\,;
\end{equation*}
%This \ds\ appears as a generic limit \ds\ in random matrix theory. 
%Its distribution function $F_1$ is given by
%\begin{equation*}
%F_1(s) = \exp\Big\{
%  -\frac{1}{2} \int_{s}^\infty [
%    q(x) + (x - s) q^2(x)
% ] \dint x
%\Big\}
%\end{equation*}
%where $q(x)$ is the unique solution to the Painlev\'e II differential
%equation
%\begin{equation*}
%  q''(x) = xq(x) + 2 q^3(x)\,,
%\end{equation*}
%where $ q(x)\sim {\rm Ai}(x)$ as $x \to \infty$ and
%Ai$(\cdot)$ is the Airy kernel; 
see Tracy and Widom~\cite{tracy:widom:2012} for details. 
Ma~\cite{zongming:2012} showed Berry--Esseen-type bounds for  
\eqref{eq:tc}. %He show improves if one slightly modifies the centering and scaling of $\lambda_{(1)}$. 
%He proved that ``the difference between the distribution of $(\lambda_{(1)}-\widetilde{\mu}_{n,p})/\widetilde{\sigma}_{n,p}$ and $F_1$ reduces to `second order', being $O((n \wedge p)^{-2/3})$ rather than $O((n \wedge p)^{-1/3})$, that would apply by using \eqref{eq:tc}'', where the modified centering and scaling constants are
%\begin{equation*}
%\widetilde{\mu}_{n,p}=(\sqrt{n-1/2}+ \sqrt{p-1/2})^2, \quad \widetilde{\sigma}_{n,p}=(\sqrt{n-1/2}+ \sqrt{p-1/2}) \Big( \frac{1}{\sqrt{n-1/2}} +\frac{1}{\sqrt{p-1/2}}\Big)^{1/3}.
%\end{equation*}
\par
%{ The calculation of the spectrum} % is facilitated by the fact that the distribution of 
%the classical Gaussian matrix ensembles is invariant under orthogonal transformations. The corresponding  
%computation for non-invariant 
Asymptotic theory for the largest eigenvalues of sample covariance matrices with 
non-Gaussian entries is more complicated; %and was a major challenge for several years; 
pioneering work is due to Johansson \cite{johansson}. 
Johnstone's result was extended to matrices $\z$ with iid non-Gaussian entries 
by Tao and Vu \cite[Theorem~1.16]{tao09b}, assuming that the first four moments of $Z$ match  those of 
the normal \ds. % they showed \eqref{eq:tc} by
%employing {\em Lindeberg's replacement method}, i.e., the iid non-Gaussian entries are replaced 
%step-by-step by iid Gaussian ones.
%This approach is well-known from summation theory for \seq s of iid \rv s. 
Tao and Vu's result is a consequence of the so-called {\em Four Moment Theorem} 
which describes the insensitivity of the eigenvalues with respect to changes in the distribution of the entries. 
To some extent (modulo the strong moment matching conditions) it shows the universality of Johnstone's limit 
result \eqref{eq:tc}.% see Figure~\ref{fig:3point} for an illustration.
%\begin{figure}[htb!]
%  \centering
%    \includegraphics[scale=0.550]{3point-TW.pdf}
%  \caption{Entry distribution: $\P(Z=\sqrt{3}) = \P(Z=-\sqrt{3} ) =
%  1/6$, $\P(Z=0)=2/3$. Note $\E [Z] = 0$, { $\E[ Z^2] = 1$, $\E [Z^3] = 0$ and
%  $\E[ Z^4] = 3$, i.e., the first 4 moments of $Z$ match those of the standard normal \ds .}}
%  \label{fig:3point}
%\end{figure}
\par
In the light-tailed case little is known when $p$ and $n$ grow at different rates, i.e., $\lim p/n \in\{ 0,\infty\}$. 
Notable exceptions are El Karoui \cite{elkaroui:2003} who proved that Johnstone's result (assuming iid standard normal entries)
remains valid when $p/n\to 0$ or $n/p\to\infty$, and P{\'e}ch{\'e} \cite{peche:2009} who showed universality results for the largest eigenvalues of some sample covariance matrices with non-Gaussian entries.
%While the iid assumption may be somewhat weakened (see \cite{bai:silverstein:2010}) and one still obtains the same convergence results of the empirical spectral distribution.
%There seems to be no corresponding result if \eqref{eq:gamma} is violated.

\subsection{The heavy-tailed case}

Distributions of which certain moments cease to exist are often called heavy-tailed. 
So far we reviewed theoretical results where the data matrix $\z$ was ``light-tailed'' in the following sense:
for the \ds al convergence of the empirical 
spectral distribution  and the largest eigenvalue of the sample covariance matrix  towards the 
Mar\v cenko--Pastur 
and Tracy-Widom \ds s, respectively, we  
required finite second/fourth moments of the entries.
\par
The behavior of the largest eigenvalue $\la_{(1)}$ 
changes dramatically when $\E[Z^4]=\infty$.
Bai and Silverstein~\cite{baisilv} proved for an $n\times n$ matrix $\z$ with iid centered entries 
that
\beam\label{eq:wdfr}
\limsup_{\nto} \dfrac{\la_{(1)}}{n}=\infty \qquad {\rm a.s.}
\eeam
This is in stark contrast to Geman's result \eqref{eq:geman}. 
%It is also unlikely that the  
%Tracy--Widom \ds\ yields a good approximation to the \ds\ of the largest eigenvalue of the sample covariance matrix.
\par
Following classical limit theory for partial sum processes and maxima, we require more than an infinite fourth moment.
We assume a {\em \regvar\ condition} on the tail of $Z$:
\beam\label{eq:regvar}
\P(Z>x)\sim p_+\,\dfrac{L(x)}{x^{\alpha}}\qquad\mbox{and}\qquad \P(Z<-x) \sim p_-\,\dfrac{L(x)}{x^\alpha}\,,\qquad \xto\,,
\eeam
for some $\alpha\in (0,4)$, where $ p_\pm$ are non-negative constants \sth\ $p_++p_-=1$ and $L$ is a \slvary\ \fct .  
We will also refer to $Z$ as a \regvary\ \rv , $\z$ as a \regvary\ matrix, etc.
%In particular, we have $\E [|X|^\delta]= \infty$ for any $\delta>\alpha$. Moments higher than $\alpha$ cease to exist. 
%Condition \eqref{eq:regvar} on $Z$  is needed for proving \asy\ theory for the eigenvalues of $\z\z'$.
%\par
%We are interested in the case $\alpha \in (0,4)$. %In view of \eqref{eq:wdfr} one has to choose a stronger normalization than $n$, at least when $p=n$. It turns out that $a_{np}^2$ with is the appropriate choice.
{\em Here and in what follows, we normalize the eigenvalues $(\la_i)$ by $(a_{np}^2)$ where the \seq\
$(a_k)$ is chosen \sth } 
\beao
\P(|Z|>a_k)\sim k^{-1}\,,\quad k\to\infty.
\eeao
Standard theory for \regvary\ \fct s (e.g. Bingham et al. \cite{bingham:goldie:teugels:1987}, Feller \cite{feller})
%From standard references on extreme value theory (for example Embrechts et al.~\cite{embrechts:kluppelberg:mikosch:1997}) one knows 
yields that $a_n=n^{1/\alpha} \ell(n)$ where $\ell$ is a slowly varying function. Assuming \eqref{eq:gamma} for $p$,
the Potter bounds (see \cite[p.~25]{bingham:goldie:teugels:1987}) yield for $\alpha\in (0,4)$ that
\begin{equation}\label{eq:toinf}
 \frac{a_{np}^2}{n} \sim \frac{n^{4/\alpha} \gamma^{2/\alpha}\,\ell^2(n^2 \gamma)}{n} \to \infty, \qquad \nto\,,
\end{equation}
i.e., the normalization $a_{np}^2$ is stronger than $n$.
\par
The eigenvalues $(\lambda_i)$ of a heavy-tailed matrix $\z\z'$ were 
studied first by Soshnikov~\cite{soshnikov:2004,soshnikov:2006}. He showed under 
\eqref{eq:gamma} and \eqref{eq:regvar} for $\alpha \in (0,2)$ that  
\begin{equation}\label{eq:larg}
\dfrac{\lambda_{(1)}}{a_{np}^2} \cid \zeta, \quad \nto,
\end{equation}
where $\zeta$ follows a {\em \Frechet distribution} with parameter $\alpha/2$:
\beao
\Phi_{\alpha/2}(x) =\ex^{-x^{-\alpha/2}}\,,\qquad x>0\,.
\eeao
Later Auffinger et al.~\cite{auffinger:arous:peche:2009} established \eqref{eq:larg} also for $\alpha \in [2,4)$ 
under the additional assumption that the entries are centered. 
Both Soshnikov \cite{soshnikov:2004,soshnikov:2006}
and Auffinger et al.~\cite{auffinger:arous:peche:2009} proved convergence of the point processes of normalized eigenvalues, 
from which one can easily infer the joint limiting distribution of the $k$ largest eigenvalues.  
Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} 
extended these results allowing for more general growth of $p$ 
than dictated by \eqref{eq:gamma} and a linear dependence structure between the rows and columns of $\z$; see also Chakrabarty et al.~\cite{chakrabarty:hazra:roy:20013} and the overview
paper Davis et al.~\cite{davis:mikosch:heiny:xie:2015}. {\em The study of eigenvectors of heavy-tailed sample covariance matrices is a fresh topic, which has not been explored in the literature listed here.}
\par
For the sake of completeness we mention 
%In \eqref{eq:MP} we have seen that the sequence 
that, under \eqref{eq:gamma} with $\gamma\in (0,1]$,  
\eqref{eq:regvar} with $\alpha\in (0,2)$ and $\E[Z]=0$ if the latter expectation is defined,
the empirical spectral \ds\ $F_{a_{n+p}^{-2}\z \z'}$ converges weakly with \pro y one to a deterministic 
\pro y \ms\ whose density $\rho_{\alpha}^\gamma$ satisfies
\begin{equation*}
\rho_{\alpha}^\gamma(x) x^{1+\alpha/2} \to \frac{\alpha \gamma}{2(1+\gamma)}, \qquad x \to \infty\,,
\end{equation*}
%$(F_{n^{-1}\z\z'})$ of empirical spectral distribution converges to the Marchenko--Pastur if the centered iid entries possess a finite second moment. Now we will discuss the situation when the entries are still iid and centered, but have an infinite variane. For this aim we assume the entries to be regularly varying with index $\alpha \in (0,2)$. Then we obtain analogously to \eqref{eq:toinf} that $a_{n+p}^2/n \to \infty$. 
%Assuming \eqref{eq:gamma} with $\gamma\in (0,1]$ in this infinite variance case 
see Belinschi et al.~\cite[Theorem~1.10]{belinschi:dembo:guionnet:2009} 
%showed that the sequence $(F_{a_{n+p}^{-2}\z \z'})$ converges with probability one 
%to a non-random probability measure $\mu_{\alpha}^\gamma$, which has heavy tails. The corresponding density $\rho_{\alpha}^\gamma$ satisfies
and Ben Arous and Guionnet \cite[Theorem~1.6]{arous:guionnet:2008}.

%It is very interesting to note that the behavior of the eigenvalues itself changes dramatically if one moves from finite to infinite fourth moment, while the type of the limit of the empirical spectral distribution depends on the finiteness of the second moment. In the case of finite fourth moment, \eqref{eq:geman} and \eqref{eq:MP} imply that $\lambda_{(k)}/n$ has to converge to the right endpoint of the Marchenko--Pastur law for any finite $k$. However, if we assume infinite fourth moment but finite second moment the empirical spectral distributions  $(F_{{n}^{-1}\z \z'})$ still converge to the Marchenko--Pastur law with the same support, while $\lambda_{(1)}/n$ will tend to infintity due to \eqref{eq:wdfr}. This means that the proportion of the normalized eigenvalues taking values larger than $(1+\sqrt{\gamma})^2$ has to be negligible; otherwise we would have a contradiction to the convergence of the empirical spectral distributions to the Marchenko--Pastur law.
%Finally, if the second moment is infinity, i.e.~regularly varying entries with index $\alpha \in (0,2)$, then the much larger values of the eigenvalues are visible in the limiting spectral distribution even when using a much stronger normalization \cite[Theorem~1.10]{belinschi:dembo:guionnet:2009}.

\subsection{Structure of the paper}
The primary objective of this paper is to study the joint distribution of the largest eigenvalues of the sample covariance matrix $\z\z'$
in the case of iid regularly varying entries with infinite fourth moment.
We make a connection between extreme value theory, point process \con\ and the behavior of the largest eigenvalues. 
We study these eigenvalues under polynomial 
growth rates of the dimension $p$ relative to the sample size $n$. 
It turns out that they are essentially determined by the extreme diagonal elements of $\bfZ\bfZ'$ or, alternatively, 
by the extreme order statistics of the squared entries of $\bfZ$.  
\par
In Section~\ref{sec:2} we consider power-law growth rates of $(p_n)$, 
thereby generalizing proportional growth as prescribed by \eqref{eq:gamma}.
Our main results are presented in Section~\ref{sec:mainresults}. 
Theorem~\ref{thm:iidmain} provides approximations of the ordered eigenvalues of the sample covariance matrix either by
the ordered diagonal elements of $\bfZ\bfZ'$ or $\bfZ'\bfZ$, or by the order statistics of the squared entries of $\bfZ$.
These approximations provide a clear picture where the largest eigenvalues of the sample covariance matrix originate from. 
Our results generalize those in Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and 
Auffinger et al.~\cite{auffinger:arous:peche:2009} who assume proportionality of $p$ and $n$.
The employed techniques originate from extreme value analysis and large deviation theory; the proofs differ from those 
in the aforementioned literature. 
The same techniques can be applied when the entries of $\bfZ$ 
are heavy-tailed and allow for dependence through the rows and across the columns; see 
Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} for some recent attempts 
when the entries satisfy some linear dependence conditions. In the iid case, these results are covered by the present paper
and we also show that they remain valid under much more general growth conditions than in \cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}. In particular, we make clear that centering of the sample covariance matrix (as assumed in 
\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} when $Z$ has a finite second moment) is not needed. Thus, our techniques are applicable under rather general dependence structures. We refer to the recent work by Janssen et al.~\cite{janssen:mikosch:rezapour:xie:2016} on eigenvalues of stochastic volatility matrix models, where non-linear dependence was allowed.
% provided the rate of $p$ is sufficiently fast.

The convergence of the point processes of the properly normalized eigenvalues in Section~\ref{sec:ppconvergence} yields a multitude of useful findings connected to the joint distribution of the eigenvalues. As an application, the structure of the eigenvectors of $\z\z'$ is explored in Section~\ref{sec:eigenvectors}.
Technical proofs are collected in Section~\ref{sec:proofs}.
Section~\ref{sec:autocov} is devoted to an extension of the results to the singular values of the 
{\em sample autocovariance matrices}  which are a generalization of the traditional autocovariance function for time series to high-dimensional matrices. In applications, the analysis of sample autocovariance matrices for different lags might help to detect dependencies in the data; see Lam and Yao \cite{lam:yao} for related work.
 We conclude with Appendix~\ref{appendix:A} which contains useful facts about regular variation and point processes.

%------------------------------------------------------------------------
\section{Preliminaries}\label{sec:2}
In this section we will discuss growth rates for $p=p_n\to\infty$ and introduce some notation. 
\subsection{Growth rates for $p$}
In many applications it is not realistic to assume 
that the dimension $p$ of the data and the sample size $n$ grow at the same rate, i.e., condition \eqref{eq:gamma} is unlikely to be
satisfied.
The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al.~\cite{auffinger:arous:peche:2009} already show that the value $\gamma$ in the growth
rate~\eqref{eq:gamma} does not appear in the \ds al limits. 
This observation is in contrast to the light-tailed case; see \eqref{eq:MP} 
and \eqref{eq:geman}.  Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} allowed for more general rates for
$p_n\to\infty$ than linear growth in $n$. {\em However, they could not completely solve the technical difficulties arising with general growth rates of~$p$.}
%Recall that $p=p_n\to\infty$ is the number of rows in the
%matrix $\z_n$. 
In what follows, we 
specify the growth rate of $(p_n)$:
%a non-degenerate limit distribution of the normalized eigenvalues of the sample autocovariance matrices. 
%To be precise, we assume
\begin{equation}\label{eq:p}
p=p_n=n^\beta \ell(n), \qquad n\ge1,\tag{$C_p(\beta)$}
\end{equation}
where $\ell$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume $\ell(n) \to \infty$. 
Condition~\ref{eq:p} is more general than the growth conditions in the literature; see 
\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.

\subsection{Notation}\label{sec:section2.2}
Recall that $\z=\z_n= (Z_{it})_{i=1,\ldots,p;t=1,\ldots,n}$ is a $p\times n$ matrix with iid entries satisfying the regular variation 
condition \eqref{eq:regvar} for some $\alpha\in (0,4)$.
The sample covariance matrix $\z \z'$ has eigenvalues $\lambda_1, \ldots, \lambda_p$ whose order statistics were defined in \eqref{eq:order}.

Important roles are played by the quantities $(Z_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$ and their order statistics
\begin{equation}\label{eq:zorder}
Z_{(1),np}^2 \ge Z_{(2),np}^2 \ge  \ldots \ge Z_{(np),np}^2, \qquad n,p\ge 1\,.
\end{equation}
As important are the row-sums 
\begin{equation}\label{eq:ll}
D_i^\rightarrow=D_i^{(n),\rightarrow}=\sum_{t=1}^n Z_{it}^2\,, \qquad
i=1,\ldots,p\,;\quad  n=1,2,\ldots\,,
\end{equation} with generic element $D^\rightarrow$ and their ordered values
\beam\label{eq:help6}
D_{(1)}^\rightarrow=D_{L_1}^\rightarrow\ge \cdots \ge D_{(p)}^\rightarrow=D_{L_p}^\rightarrow\,,
\eeam
where we assume without loss of generality that $(L_1,\ldots,L_p)$ is
a permutation of $(1,\ldots,p)$ for fixed~$n$.
\par
Finally, we introduce the column-sums
\beam\label{eq:colsum}
D_t^\downarrow=D_t^{(n),\downarrow}= \sum_{i=1}^p Z_{it}^2\,,\qquad t=1,\ldots,n\,; \quad p=1,2,\ldots\,,
\eeam
with generic element $D^\downarrow$ and we also adapt the notation from \eqref{eq:help6} to these quantities.

\subsubsection*{Norms} 
%\subsubsection*{Vector norm} 
For any $p$-dimensional vector $\bfv$, $\ltwonorm{\bfv}$ denotes its Euclidean norm.
%In this paper we will exclusively use the {\em $\ell_2$-norm}:
%\beao
%\ltwonorm{v}= \Big( \sum_{i=1}^p |v_i|^2\Big)^{1/2}\,.
%\eeao
For any $p\times p$ matrix $\bfC$, we write $\la_i(\bfC)$ for its $p$ singular values and we denote their order statistics by
\beao
\la_{(1)}(\bfC) \ge\cdots \ge \la_{(p)}(\bfC)\,. 
\eeao 
For any $p\times n $ matrix $\A=(a_{ij})$, we will use the 
{\em spectral norm}
$\|\A\|_2=\sqrt{\la_{(1)}(\A\A')}$, the 
{\em Frobenius norm} 
$
\frobnorm{\A}= \Big( \sum_{i=1}^p \sum_{j=1}^n |a_{ij}|^2\Big)^{1/2}
$ and the 
{\em max-row sum norm}
$
\inftynorm{\bfA}= \max_{i=1,\ldots,p} \sum_{j=1}^n |a_{ij}|\,.
$
%\item
%{\em Max-column sum norm:}
%\beao
%\|{A}\|_1 = \max_{j=1,\ldots,n} \sum_{i=1}^p |a_{ij}|\,.
%\eeao
%\end{itemize}
%We will frequently make use of the bound $\|\A\|_2\le \|\A\|_ F$. Standard references for matrix norms are \cite{belitski,bhatia:1997, horn, shores}.

%-------------------------\subsubsection*{Eigenvectors} 
%We introduce the unit eigenvectors
%\begin{equation}\label{eq:eigenv}
%v_j^{\rightarrow}\,, \quad j=1,\ldots,p,
%\end{equation}
%of the sample covariance matrix $\z\z'$, i.e., $\ltwonorm{v_j}=1$, where $v_j^{\rightarrow}$ is associated with the $j$th largest eigenvalue $\lambda_{(j)}$. Furthermore, let 
%\begin{equation}\label{eq:eigenvd}
%v_j^{\downarrow}\,, \quad j=1,\ldots,n,
%\end{equation}
%denote the unit eigenvectors of the $n\times n$ matrix $\z'\z$.
%The canonical basis vectors are denoted by $\bfe_j$, where $(\bfe_j)_i=\delta_{ij}$ and $\delta_{ij}$ denotes the Kronecker-$\delta$.
%-----------------------------------------------
\section{Main results}\label{sec:mainresults}
\subsection{Basic approximations}
We commence with some basic approximation results for the eigenvalues and eigenvectors of $\z\z'$.
The approximating quantities have a simple structure and their \asy\ behavior is inherited by the eigenvalues
and has influence on the eigenvectors.
\begin{theorem}\label{thm:iidmain}
Consider a $p\times n$-dimensional matrix $\z$ with iid entries. We assume the following conditions:
\begin{itemize} \item
The \regvar\ condition \eqref{eq:regvar} for some  
$\alpha \in (0,4)$. 
\item $\E [Z]=0$ for $\alpha \ge 2$.
\item
The integer \seq\ $(p_n)$ 
has growth rate \ref{eq:p} for some  $\beta\ge 0$.
\end{itemize}
Then the following statements hold:
\begin{enumerate}
\item If $\beta\in [0,1]$, then
\beam\label{eq:rowa}
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\Dr_{(i)}\big|\stp 0\,.
\eeam
%and  for the eigenvectors we have
%\beao
%\ltwonorm{v_j^{\rightarrow} - \bfe_{L_j}} \cip 0\,, \qquad j=1,\ldots,p\,.
%\eeao
\item If $\beta > 1$, then
\beam\label{eq:cola}
a_{np}^{-2}\,\max_{i=1,\ldots,n}\big|\la_{(i)}-\Dc_{(i)}\big|\stp 0\,.
\eeam
\item If  $\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$, then
\beam\label{eq:eigena}
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-Z_{(i),np}^2\big|\stp 0\,.
\eeam
\end{enumerate}
\end{theorem}
\begin{remark}
\rm In \eqref{eq:cola} we have chosen to take maxima over the index set $\{1,\ldots,n\}$. We notice that
$\la_{(i)}=0$ for $i=p\wedge n+1,\ldots,p\vee n$. This is due to the fact that the $p\times p$ matrix $\bfZ\bfZ'$ and the $n\times n$ matrix
$\bfZ'\bfZ$ have the same positive eigenvalues. Moreover, for $n$ sufficiently large, $p\wedge n=p$ for $\beta\in (0,1)$ and $p\wedge n=n$ for $\beta>1$, i.e., only in the case 
$\beta=1$ both cases $n\le p$ or $p\le n$ are possible.
\end{remark} 
\begin{remark}\rm 
The condition $\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$ in part (3) is only a restriction when $\alpha\in (2,4)$. We notice 
that this condition implies $(n\vee p)/a_{np}^2\to 0$. In turn, this means that centering of the quantities $a_{np}^{-2}D_i^\rightarrow$ and $a_{np}^{-2}D_i^\downarrow$ in the limit theorems 
can be avoided.
This argument is relevant in various parts of the proofs.
\end{remark}
\begin{remark}\rm
In Figure~\ref{fig:lambda_comparison} we illustrate 
the different approximations of the eigenvalues $(\la_{(i)})$ by 
$(\Dr_{(i)})$ as suggested by \eqref{eq:rowa}
and $(Z_{(i),np}^2)$ as suggested by \eqref{eq:eigena}. For $Z$ we choose the density
\beam\label{eq:distrsim}
f_Z(x) =
\left\{\begin{array}{cc}
 \frac{\alpha}{(4|x|)^{\alpha + 1}}\,, & \mbox{if } |x| > 1/4 \\
1\,, & \mbox{otherwise.}
\end{array}\right.
\eeam
In the left graph, we focus on the largest eigenvalue $\la_{(1)}$. We show smoothed histograms of the approximation errors
$a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$, $a_{np}^{-2}(\lambda_{(1)}-Z_{(1),np}^2)$.
By Cauchy's interlacing theorem (see \cite[Lemma~22]{tao:vu:2012}), the considered differences
are non-negative.

In the right graph, we take the maxima as in \eqref{eq:rowa} and \eqref{eq:eigena} and show smoothed histograms of the approximation errors
$a_{np}^{-2}\max_{i\le p} |\lambda_{(i)}-D_{(i)}^\rightarrow|$, $a_{np}^{-2}\max_{i\le p}|\lambda_{(i)}-Z_{(i),np}^2|$. We take absolute values to deal with negative differences. 
Figure~\ref{fig:lambda_comparison} indicates 
that $(D_{(i)}^\rightarrow)$ yield a much better approximation to $(\la_{(i)})$ than $(Z_{(i),np}^2)$. Notice the different scaling on the $x$- and $y$-axes.
\begin{figure}[htb!]
  \centering
  \subfigure{
    \includegraphics[scale=0.34]{comparison_1.pdf}
  }
  \subfigure{
    \includegraphics[scale=0.34]{comparisonmax.pdf}
  }
  \caption{Smoothed histograms of the  approximation errors for the normalized eigenvalues $(a_{np}^{-2}\la_{(i)})$
for entries $Z_{it}$ with density \eqref{eq:distrsim}, $\alpha=1.6$, $\beta=1$, $n=1,000$ and $p=200$.
}
  \label{fig:lambda_comparison}
\end{figure}
\end{remark}
\par 
The proof of Theorem~\ref{thm:iidmain} will be given in Section~\ref{sec:proofs}. A main step in the proof is provided by the following
result whose proof will also be given in Section~\ref{sec:proofs};
a version of this theorem was proved in Davis et al.~\cite{davis:mikosch:pfaffel:2015} 
under more restrictive conditions on the growth rate of $(p_n)$.

\begin{theorem}\label{prop:offdiagonal}
Assume the conditions of Theorem~\ref{thm:iidmain} on  $\z$ and $(p_n)$.
\begin{enumerate}
\item
If $\beta\in [0,1]$ we have
%has iid entries satisfying the \regvar\ condition \eqref{eq:regvar} for some  
%$\alpha \in (0,4)$. If $\E[|Z|]<\infty$ we also suppose that $\E [Z]=0$. Then for any \seq\ $(p_n)$ 
%satisfying \ref{eq:p} with $\beta\in [0,1]$ 
\beao
a_{np}^{-2} \twonorm{\bfZ \bfZ' - \diag(\bfZ \bfZ')} \cip 0\,,\qquad\nto\,.
\eeao
\item
If $\beta\ge 1$ we have 
\beao
a_{np}^{-2} \twonorm{\bfZ' \bfZ - \diag(\bfZ' \bfZ)} \cip 0\,,\qquad\nto\,.
\eeao
\end{enumerate}
\end{theorem}
The second part of this theorem follows from the first one by an interchange of $n$ and $p$.
%observing that the $p \times p$ matrix $\z\z'$ and the $n \times n$ matrix $\z'\z$ have the same non-zero eigenvalues; see Bhatia \cite[p.~64]{bhatia:1997}. 
Indeed, if $\beta\ge 1$, we can write $n=p^{1/\beta}\ell(p)$ for some \slvary\
\fct\ $\ell$ and then part (2) follows from part (1).
\begin{remark}
Theorem \ref{prop:offdiagonal} shows that the largest eigenvalues of $\z\z'$ are determined by the largest diagonal entries. 
In the case of heavy-tailed Wigner matrices, however, the diagonal elements do not play any particular role.
\end{remark}

From this theorem one immediately obtains 
a result about the approximation of the eigenvalues
 of $\z\z'$ and $\z'\z$ by those of $\diag(\z\z')$ and $\diag(\z'\z)$, respectively. Indeed, 
for any symmetric $p\times p$ matrices $\bfA,\bfB$, by {\em Weyl's inequality} (see Bhatia \cite{bhatia:1997}),
\beam\label{eq:weylin}
\max_{i=1,\ldots,p}\big|\la_{(i)}(\bfA+\bfB)-\la_{(i)}(\bfA)\big|\le \|\bfB\|_2\,.
\eeam
If we now choose 
$\bfA+\bfB=\z\z'$ and $\bfA= \diag (\z\z')$ (or $\bfA+\bfB=\z'\z$ and $\bfA= \diag (\z'\z)$) we obtain the following result.

\begin{corollary}\label{cor:687}
Assume the conditions of Theorem~\ref{thm:iidmain} on  $\z$ and $(p_n)$.
\begin{enumerate}
\item
If $\beta\in [0,1]$ we have
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\diag(\z\z'))\big|\stp 0\,,\qquad \nto\,.
\eeao
\item
If $\beta> 1$  we have
\beao
a_{np}^{-2}\,
\max_{i=1,\ldots,n}\big|\la_{(i)}-\la_{(i)}(\diag(\z'\z))\big|
\stp 0\,,\qquad\nto \,.
\eeao
\end{enumerate}
\end{corollary}
Now \eqref{eq:rowa} and \eqref{eq:cola} are immediate con\seq s of this corollary. Indeed, we have
$\la_{(i)}(\diag(\z\z'))=D_{(i)}^\rightarrow$ and $\la_{(i)}(\z'\z)=D_{(i)}^\downarrow$, $i=1,\ldots,p\wedge n$.
%\begin{proposition}\label{prop:offdiagonal1}
%Assume that $\z=\z_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvar} for some  
%$\alpha \in (0,4)$. If $\E[|Z|]<\infty$ we also suppose that $\E [Z]=0$. Then for any \seq\ $(p_n)$ 
%satisfying \ref{eq:p} with $\beta>1$ we have
%\beao
%a_{np}^{-2} \twonorm{\z_n'\z_n - \diag(\z_n'\z_n)}\stp 0\,,\qquad\nto\,.
%\eeao
%\end{proposition}
%Note that for $\beta>1$ we have $\lim_{\nto} p/n= \infty$. This means that 
%$\z_n'\z_n$ has  asymptotically a much smaller dimension than $\z_n\z_n'$ and therefore it is more convenient to work with $\z_n'\z_n$ 
%when bounding the spectral norm. 
%\begin{corollary}\label{cor:6871}
%Under the conditions of Proposition~\ref{prop:offdiagonal1},
%\beao
%a_{np}^{-2}\,\max_{i=1,\ldots,n}\big|\la_{(i)}-\la_{(i)}(\diag(\z_n'\z_n))\big|\stp 0\,.
%\eeao
%\end{corollary}

\subsection{Point process \con }\label{sec:ppconvergence}
In this section we want to illustrate how the approximations from Theorem~\ref{thm:iidmain} can be used
to derive \asy\ theory for the largest eigenvalues of $\z\z'$ via the weak \con\ of suitable \pp es.
The limiting \pp\ involves the points of the Poisson process
\begin{equation}\label{eq:N}
N_\Gamma=\sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\end{equation}
where $\vep_y$ is the Dirac measure at $y$, 
\beao%\label{eq:Gamma}
\Gamma_i=E_1+\cdots + E_i\,,\qquad i\ge 1\,,
\eeao
and $(E_i)$ is a sequence of iid standard exponential random variables. In other words, $N_\Gamma$ is a Poisson point process on $(0,\infty)$ with mean \ms\
$\mu(x,\infty)= x^{-\alpha/2}$, $x>0$.
\par
\begin{lemma}\label{lem:pp}
Assume the conditions of Theorem~\ref{prop:offdiagonal} hold. 
\begin{enumerate}
\item
If $\beta\ge 0$, then
\beam\label{eq:immb}
\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i^\rightarrow-c_n)}\cid N_\Gamma\,,\qquad \nto\,,
\eeam  
where $c_n=0$ if $\E[D^\rightarrow]=\infty$ and $c_n=\E[D^\rightarrow]=n\,\E[Z^2]$ otherwise. 
\item
If $\beta\ge 0$, then
\beam\label{eq:imma}
\sum_{i=1}^p \vep_{a_{np}^{-2}Z_{(i),np}^2}\cid N_\Gamma\,,\qquad \nto\,,
\eeam  
\end{enumerate}
The weak \con\ of the \pp es holds in the space of 
point \ms s with state space $(0,\infty)$ equipped with the vague topology; see Resnick \cite{resnick:2007}.
\end{lemma}
\begin{remark} \rm
Similar results were used in the proofs of Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2015}.
We also mention that the centering $c_n$ in the finite variance case can be avoided if $n/a_{np}^2\to 0$. 
The latter condition is satisfied if $\beta>\alpha/2-1$.
\end{remark}
\begin{proof} 
Part (1) follows from Lemma~\ref{lem:ppr}. As regards part (2), we observe that
\beam\label{eq:gp}
\sum_{i=1}^p\sum_{t=1}^n \vep_{a_{np}^{-2} Z_{it}^2}\std N_\Gamma\,;
\eeam
see e.g. Resnick~\cite{resnick:1987}, Proposition 3.21. On the other hand, $a_{np}^{-2} Z_{(p),np}^2\stp 0$ which together with
\eqref{eq:gp} yields part (2).
\end{proof}

Theorem~\ref{thm:iidmain}  and arguments similar to the proofs in Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2015}
enable one to derive the weak \con\ of
the point processes of the normalized eigenvalues. 
%\begin{equation}\label{eq:ppdef}
%N_n=\sum_{i=1}^p \vep_{a_{np}^{-2} \lambda_i}, \quad \widetilde{N}_n^{\rightarrow}=\sum_{i=1}^p \vep_{a_{np}^{-2} (\lambda_i-n\,\E[Z^2])}, \qquad \mbox{and} \qquad
%\widetilde{N}_n^{\downarrow}=\sum_{i=1}^n \vep_{a_{np}^{-2} (\lambda_i-p\,\E[Z^2])}, \qquad n\ge 1\,.
%\end{equation}
\begin{theorem}\label{cor:1}
Assume the conditions of Theorem~\ref{thm:iidmain}. If
$\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$ 
then
\beam\label{eq:imm}
\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std N_\Gamma\,,
\eeam 
in the space of point measures 
with state space $(0,\infty)$ equipped with the vague topology. 
\end{theorem}
\begin{proof}
The limit relation \eqref{eq:imm} follows from \eqref{eq:imma} in combination with
\eqref{eq:eigena}. Alternatively, one can exploit \eqref{eq:immb} both for $(D_i^\rightarrow)$ and 
$(D_t^\downarrow)$ (notice that the \pp\ \con\ for the latter \seq\ follows by interchanging the roles of $n$ and $p$), the fact that
$(n\vee p)/a_{np}^2\to 0$ if $\min(\beta,\beta^{-1})\in ((\alpha/2-1)_+,1]$ (hence centering of the points $(D_i^\rightarrow)$ and  $(D_t^\downarrow)$ in \eqref{eq:immb}
can be avoided for $\E[Z^2]<\infty$) and finally using the approximations \eqref{eq:rowa} or \eqref{eq:cola}.
\end{proof}
The weak \con\ of the \pp es of the normalized eigenvalues of $\bfZ\bfZ'$ 
in Theorem~\ref{cor:1} allows one to use the conventional tools in this
field; see Resnick~\cite{resnick:2007,resnick:1987}. An immediate con\seq\ is
\beam \label{eq:helpgamma1}
a_{np}^{-2}\big(\la_{(1)},\ldots,\la_{(k)}\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)
\eeam
for any fixed $k\ge 1$. Using the methods of Davis et al.~\cite{davis:mikosch:heiny:xie:2015} shows for $\alpha \in (2,4)$
\beam \label{eq:helpgamma2}
a_{np}^{-2}\big(\la_{(1)}-(p\vee n) \E[Z^2],\ldots,\la_{(k)}-(p\vee n) \E[Z^2]\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,.
\eeam
Equations \eqref{eq:helpgamma1} and \eqref{eq:helpgamma2} yield that for $\alpha \in (0,4)$ and any fixed $k\ge 1$,
\beam \label{eq:helpgamma3}
a_{np}^{-2}\big(\la_{(1)}- \la_{(2)},\ldots, \la_{(k)}- \la_{(k+1)}\big)\std \big(\Gamma_1^{-2/\alpha} - \Gamma_{2}^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha} - \Gamma_{k+1}^{-2/\alpha} \big)\,.
\eeam
Related results can also be derived for an increasing number of order statistics, e.g.~the joint \con\ of the largest eigenvalue $a_{np}^{-2}\la_{(1)}$ and the trace $a_{np}^{-2}(\la_1+\cdots+\la_p)$. In particular, one obtains for
$\alpha\in (0,2)$ under the conditions of Theorem~\ref{cor:1} that
\beao
\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \std \dfrac{\Gamma_1^{-2/\alpha}}{\Gamma_1^{-2/\alpha}+\Gamma_2^{-2/\alpha}+\cdots}\,.
\eeao 
We refer to Davis et al.~\cite{davis:mikosch:pfaffel:2015} for details on the proofs and more examples.

In the next subsection we will show how the above results on the joint convergence of eigenvalues can be applied to approximate the eigenvectors of $\z\z'$.

%---------------------------------------------------------------------
\subsection{Eigenvectors}\label{sec:eigenvectors}
In this section we assume the conditions of Theorem~\ref{prop:offdiagonal} and $\beta \in [0,1]$.
From Theorem~\ref{prop:offdiagonal}(1) %and the proof of Corollary~\ref{cor:688} 
we know that $\z\z'$ is approximated in spectral norm by $\diag(\z\z')$.
% and $\Y_n$, both of which are diagonal matrices, are good approximations to the sample covariance matrix $\z_n\z_n'$ in the spectral norm. 
The unit eigenvectors %$\bfv_j$, $j=1,\dots,p$, 
of a $p\times p$ diagonal matrix are the canonical basis vectors $\bfe_j\in\R^p$, $j=1,\dots,p$. 
This raises the question as to whether $(\bfe_j)$ are good approximations of the eigenvectors $(\bfv_j)$ of $\z\z'$.
By $\bfv_{{j}}$ we denote the unit eigenvector associated with the $j$th largest eigenvalue $\la_{(j)}$. The unit eigenvector associated with the $j$th largest eigenvalue of $\diag(\z\z')$ is $\bfe_{L_j}$, where $L_j$ is defined in \eqref{eq:help6}. Our guess that $\bfv_j$ is approximated by $\bfe_{L_j}$ is confirmed by the following result.

\begin{theorem}\label{thm:eigenvec} Assume the conditions of Theorem~\ref{thm:iidmain} and let
$\beta\in [0,1]$. Then for any fixed $k\ge 1$,
\beao
\ltwonorm{\bfv_k - \bfe_{L_k}} \cip 0\,, \quad \nto \,.
\eeao
\end{theorem}

Indeed, $\bfv_j$ and $\bfe_{L_j}$ share another property: they are {\em localized} which means that they are concentrated only in a few components. Vectors which are not localized are called {\em delocalized}.
Figure~\ref{fig:eigenvectors} shows the outcome of a simulation example in which we visualize 
the components of the unit eigenvector associated with the largest eigenvalue of $\z\z'$ 
for a simulated data matrix $\z$ with iid Pareto(0.8) entries. In the right graph
we see that only one of the $p=200$ components is significant. 
Hence we can find a canonical basis vector $\bfe_{k}$ such that $\ltwonorm{\bfe_{k}-\bfv_{1}}$ is small. 
Therefore the eigenvector is localized. This is in stark contrast to the case of iid standard normal entries; see the left graph.
%When the entry distribution is standard normal, the picture looks completely different. 
Then many components are of similar magnitude, hence the eigenvector is delocalized. 
Typically, the eigenvectors tend to be localized when the entry distribution has an infinite fourth moment, 
while they tend to be delocalized  otherwise; see Benaych-Georges and P\'{e}ch\'{e} \cite{benaych:peche} for the case of Wigner matrices. 
\begin{figure}[htb!]
  \centering
  \subfigure%[Standard normal entries]
{
    \includegraphics[scale=0.34]{LargestEigenvector-NormalData.pdf}
  }
%  \subfigure[Pareto(0.8) entries]
{
    \includegraphics[scale=0.34]{LargestEigenvector-ParetoData.pdf}
  }
  \caption{The components of the eigenvector $\bfv_1$. Right: The case of iid Pareto(0.8) entries. Left: The case
of iid standard normal entries. 
%$ \max_{i=1,\dots, p} |V_{1, i}| = 1 - 10^{-5}???$ 
We choose $p=200$ and $n=1,000$.
}
  \label{fig:eigenvectors}
\end{figure}

%The observation from Figure~\ref{fig:eigenvectors} that the eigenvector appears to be localized 
%in a single simulation study does not mean that that the eigenvectors are ``close'' to the canonical basis 
%vectors in general. Also if many of the largest $k$ eigenvalues lie in a very small interval (as for 
%standard normal data) it is not easy to determine which basis vector should provide a good approximation to which eigenvector even if we assume the set of eigenvectors and the set of basis vectors to coincide. In applications small approximation errors might thus lead to grave mistakes. The distributional limit of the spectral gaps
%\begin{equation*}
%a_{np}^{-2} (\lambda_{(i)}- \lambda_{(i+1)})\,, \quad i=1,\ldots, k,
%\end{equation*}
%will provide a measure of the placing of $(\lambda_{(i)})$ on the positive real line. From \eqref{eq:jointconv} we conclude that
%\beam\label{eq:gap}
%a_{np}^{-2} (\lambda_{(i)}- \lambda_{(i+1)})\cid \Gamma_i^{-2/\alpha}-\Gamma_{i+1}^{-2/\alpha}\,, \quad i=1,\ldots, k.
%\eeam

\begin{proof}[Proof of Theorem~\ref{thm:eigenvec}]
Fix $k\ge 1$. Since $p\to \infty$ we can assume $k\le p$ for sufficiently large $n$.
We observe that for $j=1,\ldots,p$,
\begin{equation*}
\z\z' \,\bfe_j -\Dr_j \,\bfe_j = \Big( \sum_{t=1}^n Z_{1t}Z_{jt}, \ldots,\sum_{t=1}^n Z_{j-1,t}Z_{jt}, 0,  \sum_{t=1}^n Z_{j+1,t}Z_{jt}, \ldots,\sum_{t=1}^n Z_{pt}Z_{jt} \Big)'\,,
\end{equation*}
are the columns of $\z\z' -\diag(\z\z')$. By Theorem~\ref{prop:offdiagonal}(1),
\beam \label{eq:colas}
a_{np}^{-2} \max_{j=1,\ldots,p} \ltwonorm{\z\z' \bfe_j -\Dr_j \bfe_j}
 \le a_{np}^{-2} \twonorm{\bfZ \bfZ' - \diag(\bfZ \bfZ')} \cip 0\,,\,\nto.
\eeam

If we set $\bfH^{(n)}=a_{np}^{-2}\z\z'$, $\bfv^{(n)}=\bfe_{L_k} \in \R^p$ and $\lambda^{(n)}= a_{np}^{-2} \Dr_{L_k}$, we see that
\begin{equation*}
a_{np}^{-2}\z\z' \bfe_{L_k} = a_{np}^{-2} \Dr_{L_k} \bfe_{L_k} + \vep^{(n)} \bfw^{(n)},
\end{equation*}
 where $\bfw^{(n)}= \ltwonorm{\z\z' \bfe_{L_k} -\Dr_{L_k} \bfe_{L_k}}^{-1} (\z\z' \bfe_{L_k} -\Dr_{L_k} \bfe_{L_k})$ is a unit vector and $\vep^{(n)}= a_{np}^{-2} \ltwonorm{\z\z' \bfe_{L_k} -\Dr_{L_k} \bfe_{L_k}} \cip 0$ by \eqref{eq:colas}.

Before we can apply Proposition~\ref{prop:perturbation} we need to show that with probability converging to $1$, there are no other eigenvalues in a suitably small interval around $\lambda_{(k)}$. Let $s>1$. We define the set
\begin{equation*}
\Omega_n = \Omega_n(k,s)= \{a_{np}^{-2} |\la_{(k)}-\la_{(i)}|>s\, \vep^{(n)}\, :\, i\neq k =1,\ldots,p  \}\,.
\end{equation*} 
From \eqref{eq:colas} we get $s\, \vep^{(n)}\to 0$. Then using this and \eqref{eq:helpgamma3}, we obtain
\begin{equation*}
\begin{split}
\lim_{\nto} \P(\Omega_n^c) &= \lim_{\nto} \P( a_{np}^{-2} \min\{ \la_{(k-1)}-\la_{(k)}, \la_{(k)}-\la_{(k+1)}\} \le s\, \vep^{(n)}) =0
\end{split}
\end{equation*}

By Proposition~\ref{prop:perturbation} the unit eigenvector $\bfv_k$ associated with $\la_{(k)}$ and the projected vector $\bfP_{\bfe_{L_k}}(\bfv_{k})=(\bfv_{k})_{L_k}\bfe_{L_k}$ satisfy for fixed $\delta>0$:
\begin{equation*}
\begin{split}
\limsup_{\nto} \, &\P(\ltwonorm{\bfv_k-(\bfv_{k})_{L_k}\bfe_{L_k}}>\delta)\\
&\le \limsup_{\nto} \P(\{\ltwonorm{\bfv_k-(\bfv_{k})_{L_k}\bfe_{L_k}}>\delta\} \cap \Omega_n)+ \limsup_{\nto}\P(\Omega_n^c)\\
&\le \limsup_{\nto} \P(\{2\vep^{(n)}/(s\,\vep^{(n)}-\vep^{(n)})>\delta\} \cap \Omega_n)\\
&\le \limsup_{\nto} \P(\{2/(s-1)>\delta\})=  \1_{\{ 2/(s-1)>\delta\}}.
\end{split}
\end{equation*}
The right-hand side is zero for sufficiently large $s$. Since both $\bfv_k$ and $\bfe_{L_k}$ are unit vectors this means that
\beao
\ltwonorm{\bfv_k - \bfe_{L_k}} \cip 0\,, \quad \nto \,.
\eeao
This proves our result on eigenvectors.
\end{proof}

%-------------------------------------------------------------------------
%Under more restrictive conditions on the dimension $p$, Davis et al.~\cite{davis:pfaffel:stelzer:2014} showed that the eigenvalues of $\z\z'$ are basically determined by the diagonal entries of this matrix if 
%the norm of  $a_{np}^{-2} (\bfZ_n \bfZ_n' - \diag(\bfZ_n \bfZ_n'))$ is ``small''. Here
%\begin{equation*}
%(\bfZ_n \bfZ_n' - \diag(\bfZ_n \bfZ_n'))_{ij}= \sum_{t=1}^n Z_{it} Z_{jt}, \quad i\neq j =1, \ldots,p.
%\end{equation*}
%Under more restrictive conditions on the dimension $p$,  showed the fo
%that the eigenvalues of $\z\z'$ are basically determined by the diagonal entries of this matrix if 
%the norm of  $a_{np}^{-2} (\bfZ_n \bfZ_n' - \diag(\bfZ_n \bfZ_n'))$ is ``small''. Here
\section{Proof of Theorem~\ref{thm:iidmain} }\label{sec:proofs}
{\em In what follows, $c$ stands for any constant whose value is not of interest.}
We write $(Z_t)$ for an iid \seq\ with the same \ds\ as $Z$.
\par
The plan of the proof is as follows:
\begin{enumerate}
\item
We prove  Theorem~\ref{prop:offdiagonal} which implies \eqref{eq:rowa} and \eqref{eq:cola}; see Corollary~\ref{cor:687}.
In view of the arguments after Theorem~\ref{thm:iidmain} it suffices to consider only the case $\beta\in [0,1]$.
\item
We prove \eqref{eq:eigena}.
\end{enumerate} 
\subsection{Proof of Theorem~\ref{prop:offdiagonal}}
We proceed in several steps.
\begin{proof}[The case $\alpha \in (0,8/3)$]
If $\alpha\in [1,2)$ and $\E[|Z|]< \infty$, we have
\begin{equation*}
a_{np}^{-1} \twonorm{\bfZ - (\bfZ -\E[\bfZ])}= |\E[Z]| \frac{\sqrt{np}}{a_{np}}\to 0\,, \quad \nto.
\end{equation*}
Therefore, without loss of generality $\E[Z]$ can be assumed $0$ in this case.

From now on we assume $\E[Z]=0$ whenever $\E[|Z|]$ exists. Since the Frobenius norm $\frobnorm{\cdot}$ is an upper bound of the spectral norm we have
\beao
\twonorm{\bfZ \bfZ' - \diag(\bfZ \bfZ')}^2&\le& \frobnorm{\bfZ \bfZ' - \diag(\bfZ \bfZ')}^2\\
&=&\sum_{i, j =1; i\neq j}^p   \sum_{t=1}^n Z_{it}^2 Z_{jt}^2+\sum_{i, j =1; i\neq j}^p   \sum_{t_1,t_2=1; t_1\neq t_2}^n Z_{i,t_1} Z_{j,t_1}Z_{i,t_2} Z_{j,t_2}\\
&=&\sum_{i, j =1; i\neq j}^p   \sum_{t=1}^n Z_{it}^2 Z_{jt}^2 \big[\1_{\{ Z_{it}^2 Z_{jt}^2  > a_{np}^4 \}}+\1_{\{ Z_{it}^2 Z_{jt}^2  \le a_{np}^4 \}}\big]
%\sum_{i, j =1; i\neq j}^p   \sum_{t=1}^n Z_{it}^2 Z_{jt}^2 \1_{\{ Z_{it}^2 Z_{jt}^2\le   a_{np}^4 \}}
+I_2^{(n)}\\
&=&I_{11}^{(n)}+I_{12}^{(n)}+I_2^{(n)}\,.
\eeao
Thus it suffices to show that each of the expressions on the \rhs\ when normalized with $a_{np}^4$ converges to zero in \pro y.
%\begin{equation*}
%\begin{split}
%\P(a_{np}^{-2} \twonorm{\bfZ_n \bfZ_n' &- \diag(\bfZ_n \bfZ_n')}>c) \le \P(a_{np}^{-2} \frobnorm{\bfZ_n \bfZ_n' - \diag(\bfZ_n \bfZ_n')}>c)\\
%&\le \P \Big( \sum_{i, j =1; i\neq j}^p  \Big( \sum_{t=1}^n Z_{it} Z_{jt}\Big)^2 >c a_{np}^4 \Big)\\
%&\le \P \Big( \sum_{i, j =1; i\neq j}^p   \sum_{t=1}^n Z_{it}^2 Z_{jt}^2 >c a_{np}^4\Big) + \P \Big( \sum_{i, j =1; i\neq j}^p   \sum_{t_1,t_2=1; t_1\neq t_2}^n Z_{i,t_1} Z_{j,t_1}Z_{i,t_2} Z_{j,t_2} >c a_{np}^4 \Big)\\
%&= P_1^{(n)}+P_2^{(n)}.
%\end{split}
%\end{equation*}
%We proceed by showing that $P_1^{(n)}$ and $P_2^{(n)}$ converge to zero. One sees
We have for any $\epsilon>0$,
\begin{equation*}
\P \big( I_{11}^{(n)}>\epsilon\,a_{np}^4\big)\le   p^2\,n\, \P(Z_{1}^2 Z_{2}^2 > a_{np}^4) \to 0\,.
\end{equation*}
%We obtain $P_{11}^{(n)} \le p^2n \P(Z_{1}^2 Z_{2}^2 > a_{np}^4) \to 0$ as $\nto$. 
Here we also used the fact that $Z_1Z_2$ is regularly varying with index $\alpha$; see Embrechts and Goldie \cite{embrechts:goldie:1980}. 
An application of Markov's inequality and Lyapunov's moment inequality with $\gamma\in (\alpha/2, 4/3)$ if $\alpha \in [2,8/3)$ and $\gamma=1$ otherwise shows that 
\begin{equation*}
\P\big( I_{12}^{(n)}>\epsilon\,a_{np}^4\big)\le
c\, \frac{p^2n}{a_{np}^4} \Big(\E[|Z_{1} Z_{2}|^{2\gamma}\1_{\{ |Z_{1} Z_{2}| \le a_{np}^2 \}}]\Big)^{\frac{1}{\gamma}}\le 
c\, p^{2-\frac{2}{\gamma}}\,n^{1-\frac{2}{\gamma}+\delta}\to 0,
\end{equation*} 
where we used Karamata's theorem (see Bingham et al. \cite{bingham:goldie:teugels:1987}), and the constant $\delta>0$ can be chosen arbitrarily small due to the Potter bounds.
\par
In the case $\alpha \in (0,2)$ the probability $P_{2}^{(n)}=\P(I_2^{(n)}>\epsilon\,a_{np}^4)$ can be handled analogously. Next, we turn to $P_{2}^{(n)}$ in the case $\alpha \in (2,8/3)$. In particular, $\E [Z^2]<\infty$. With \v Cebychev's inequality, also using the fact that $\E [Z]=0$,  we find that
\begin{equation}\label{eq:sdvfg}
P_{2}^{(n)} \le c \frac{1}{a_{np}^8} \E\Big[ \Big( \sum_{i, j =1; i\neq j}^p   \sum_{t_1,t_2=1; t_1\neq t_2}^n Z_{i,t_1} Z_{j,t_1}Z_{i,t_2} Z_{j,t_2}\Big)^2 \Big]\le c\,\frac{(p\,n)^2}{a_{np}^8} \to 0.
\end{equation}
The case $\alpha=2$ is
most difficult because the second moment of $Z$ can be infinite. 
Without loss of generality we assume that $Z$ is continuous. Otherwise, we add independent centered normal random variables 
to each of the entries $Z_{it}$; due the normalization $a_{np}^2$ 
the \asy\ properties of the eigenvalues remain the same, i.e., the added normal components are \asy ally negligible.  
In view of Hult and Samorodnitsky \cite[Lemma~$4.2$]{hult:2008} there exist constants $C,K>0$ and  a 
function $h:[K,\infty) \to (0,\infty)$ such that 
\begin{equation}\label{eq:fcth}
\E[Z \1_{\{ -h(x) \le Z \le x \}}]=0 \quad \mbox{and} \quad C^{-1} \le \frac{h(x)}{x} \le C
\end{equation}
for all $x\ge K$.\footnote{Here we assume that $p_+\,p_->0$. If either $p_+=0$ or $p_-=0$ one can proceed 
in a similar way by modifying $h$ slightly; we omit details.} We have 
\begin{equation*}
I_2^{(n)}=
\sum_{i, j =1; i\neq j}^p   \sum_{t_1,t_2=1; t_1\neq t_2}^n Z_{i,t_1} Z_{j,t_1}Z_{i,t_2} Z_{j,t_2}\,
\big[\1_{A_{i,j,t_1,t_2}^c}+\1_{A_{i,j,t_1,t_2}}\big]=I_{21}^{(n)}+I_{22}^{(n)}\,,
\end{equation*}
where $A_{i,j,t_1,t_2}=\{ -h(a_{np}^4) \le Z_{i,t_1} , Z_{j,t_1},Z_{i,t_2} ,Z_{j,t_2}\le a_{np}^4 \}$. We see that
\beao
\P(I_{21}^{(n)}>\epsilon\,a_{np}^4)&\le& (p\,n)^2\, \P(A_{i,j,t_1,t_2}^c)\le c\, (p\,n)^2\, \P(|Z| >\min(h(a_{np}^4), a_{np}^4))\\
&\le& c\, (pn)^2\, \P(|Z| > \min(C,C^{-1}) \,a_{np}^4)\le c\, (np)^{-2+\delta}\to 0,
\eeao
where we used the second formula in \eqref{eq:fcth}. The small constant $\delta>0$  comes from a Potter bound argument.
Finally, using the first condition in \eqref{eq:fcth}, we may  conclude similarly to \eqref{eq:sdvfg} that 
\begin{equation*}
P_{22}^{(n)}=\P(I_{22}^{(n)}>\epsilon\,a_{np}^4) \le c\, \frac{(pn)^2}{a_{np}^8}\, \big(\E[Z^2 \1_{\{ -h(a_{np}^4) \le Z \le a_{np}^4 \}}]\big)^4.
\end{equation*}
Since 
\begin{equation*}
\E[Z^2 \1_{\{ |Z| \le \max(C,C^{-1}) x \}}] \ge \E[Z^2 \1_{\{ -h(x) \le Z \le x \}}]\,,
\end{equation*}
and the \lhs\
is slowly varying (see \cite{feller}), we have $P_{22}^{(n)}\to 0$. The proof is complete for $\alpha\in (0,8/3)$. 
\end{proof}

\subsubsection*{The case $\alpha \in [8/3,4)$}
Before we can proceed with the case $\alpha\in [8/3,4)$ we provide an auxiliary result.
Consider the following decomposition 
\beao 
[\bfZ \bfZ' - \diag(\bfZ \bfZ')]^2=\bfD+\bfF+\bfR\,,
\eeao
where 
\beao
\bfD=(D_{ij})_{i,j=1,\ldots,p}=\diag([\bfZ \bfZ' - \diag(\bfZ \bfZ')]^2)\,,
\eeao
The $p\times p$ matrix $\bfF$ has a zero-diagonal and 
\begin{equation*}
F_{ij}= \sum_{u=1;u\neq i,j}^p \sum_{t=1}^n  Z_{it}\, Z_{jt}\, Z_{ut}^2 ,\quad 1\le i\ne j\le p\,,
\end{equation*}
The $p\times p$ matrix $\bfR$ has a zero-diagonal and 
\begin{equation*}
R_{ij}= \sum_{u=1;u\neq i,j}^p \sum_{t_1=1}^n \sum_{t_2=1; t_2 \neq t_1}^n Z_{i,t_1}\, Z_{j,t_2}\, Z_{u,t_1}\,Z_{u,t_2} ,\quad 
1 \le i\neq j\le p\,.
\end{equation*}
\begin{lemma}\label{lem:diagonal} 
Assume the conditions of Theorem~\ref{prop:offdiagonal} and $\alpha\in(2,4)$. 
Then $a_{np}^{-4} \big(\twonorm{\bfD}+\twonorm{\bfF}+\|\bfR\|_2\big)\stp 0$.
\end{lemma}
In view of this lemma we have 
\beao
a_{np}^{-4}\twonorm{\bfZ \bfZ' - \diag(\bfZ \bfZ')}^2= 
a_{np}^{-4}\twonorm{[\bfZ \bfZ' - \diag(\bfZ \bfZ')]^2}
=a_{np}^{-4}\twonorm{\bfD+\bfF+\bfR}^2\stp 0\,.
\eeao
This finishes the  proof of Theorem~\ref{prop:offdiagonal}. It is left to prove Lemma~\ref{lem:diagonal}.
\begin{proof}[Proof of the $\bfD$-part]
%Let $\alpha\in (2,4)$. An entry $D_{ii}$, $i \in \{ 1, \ldots, p\}$ of the matrix $D$ has the following form
We have for $i=1,\ldots,p$,
\begin{equation*}
\begin{split}
D_{ii}&= \sum_{u=1}^p \sum_{t=1}^n  Z_{it}^2 Z_{ut}^2 \1_{\{ i\neq u\}}
+\sum_{u=1}^p \sum_{t_1=1}^n \sum_{t_2=1}^n Z_{i,t_1} Z_{u,t_1}Z_{u,t_2} Z_{i,t_2} \1_{\{ i\neq u\}}\1_{\{ t_1\neq t_2\}}\\ &= M_{ii}+N_{ii}\,.
\end{split}
\end{equation*}
We write $\bfM$ and $\bfN$ for diagonal matrices constructed from $(M_{ii})$ and $(N_{ii})$ such that $\bfD=\bfM+\bfN$.
%the uncentered and centered parts of the diagonal of $\diag(\bfD)$. 
First bounding $\|\bfN\|_2$ by the Frobenius norm and then applying Markov's inequality, one can  prove that $a_{np}^{-4} \twonorm{\bfN}\stp 0$.
We have 
\begin{equation*}
\frac{\E[M_{ii}]}{a_{np}^{4}} \le c\, \frac{np}{a_{np}^{4}} \to 0, \qquad \nto.
\end{equation*}
Therefore centering of $M_{ii}$ will not influence the limit of the spectral norm $a_{np}^{-4}\|\bfM\|_2$.
Writing $A_{i,u}=\{|\sum_{t=1}^n  (Z_{it}^2 Z_{ut}^2 -\E[Z_1^2Z_2^2])\1_{\{ i\neq u\}}| > a_{np}^2 \}$,
we have for $i=1, \ldots,p$,
\beao
M_{ii}-\E[M_{ii}]&=& \sum_{u=1}^p \sum_{t=1}^n  \big(Z_{it}^2 Z_{ut}^2 -\E[Z_1^2Z_2^2]\big)\,\1_{\{ i\neq u\}}\,\big[\1_{A_{i,u}}+
\1_{A_{i,u}^c}\big]\\
%\1_{\{|\sum_{t=1}^n  (Z_{it}^2 Z_{ut}^2 -\E[Z_1^2Z_2^2])\1_{\{ i\neq u\}}| > a_{np}^2 \}}+
%& \quad + \sum_{u=1}^p 
%\Big(\sum_{t=1}^n  (Z_{it}^2 Z_{ut}^2 -\E[Z_1^2Z_2^2])
%\1_{\{|\sum_{t=1}^n  (Z_{it}^2 Z_{ut}^2 -\E[Z_1^2Z_2^2])\1_{\{ i\neq u\}}| \le a_{np}^2 \}}\Big]\\
&=& M_{ii}^{(1)} + M_{ii}^{(2)}. 
\eeao
On the one hand, $\twonorm{M^{(2)}}\le p\, a_{np}^2$. Hence $a_{np}^{-4} \twonorm{M^{(2)}}\stp 0$. On the other hand, 
we obtain with Markov's inequality, Proposition~\ref{prop:karasumsbeta} and the Potter bounds for $\epsilon>0$ and small $\delta>0$,
\beao
\P(\twonorm{M^{(1)}} >\epsilon\, a_{np}^{4}) &=& \P(\max_{i=1,\ldots, p} |M_{ii}^{(1)}| >\epsilon\,a_{np}^{4})\\
%&\le &(p\,n)\,\P(|\sum_{t=1}^n  (Z_{1t}^2 Z_{2t}^2 -\E[Z_1^2Z_2^2])| > a_{np}^2)
&\le& \P\Big(\max_{i=1,\ldots, p}  \sum_{u=1}^p \Big|\sum_{t=1}^n  (Z_{it}^2 Z_{ut}^2 -\E[Z_1^2Z_2^2])
\1_{\{ i\neq u\}}\,\1_{A_{i,u}}\Big|
%{\{ i\neq u\}}| > a_{np}^2 \}}  
>\epsilon\, a_{np}^{4}\Big)\\
&\le& c \frac{p^2}{a_{np}^{4}} \E\Big[ \Big|\sum_{t=1}^n  (Z_{1t}^2 Z_{2t}^2 -\E[Z_1^2Z_2^2])\Big| 
\1_{A_{1,2}}%\{|\sum_{t=1}^n  (Z_{1t}^2 Z_{2t}^2 -\E[Z_1^2Z_2^2])| > a_{np}^2 \}} 
\Big]\\
&\sim& c\, \frac{p^2}{a_{np}^{4}} \,n\, a_{np}^{2}\, \P(Z_1^2Z_2^2>a_{np}^2) \le\dfrac{ p \,(np)^\delta}{a_{np}^{2}} \to 0,
\eeao
since $Z_1Z_2$ is \regvary\ with index $\alpha$. This finishes the proof of the $\bfD$-part.
\end{proof}
%\begin{lemma}\label{lem:Foffdiagonal}
%Assume $\alpha\in(2,4)$.
%Then $a_{np}^{-4} \twonorm{F}$ converges to $0$ in probability.
%\end{lemma}
\begin{proof}[Proof of the $\bfF$-part]
%The idea is to split $F$ into several parts.
Let $\delta>0$. We will use the following decomposition for $i\ne j$:
\begin{equation*}
\begin{split}
F_{ij}&= \sum_{u=1;u\neq i,j}^p \sum_{t=1}^n  Z_{it} Z_{jt} (Z_{ut}^2 - \E[Z^2 \1_{\{Z^2 \le a_{np}^{4-2\delta}\}}])+ \\ &\quad +\E [Z^2\1_{\{Z^2 \le a_{np}^{4-2\delta}\}}]\,(p-2) 
\sum_{t=1}^n  Z_{it} Z_{jt} =\wt F_{ij}+T_{ij}\,.
\end{split}
\end{equation*}
We observe that $\bfT=  \E [Z^2\1_{\{Z^2 \le a_{np}^{4-2\delta}\}}] \,(p-2) \,(\bfZ_n \bfZ_n' - \diag(\bfZ_n \bfZ_n'))$.
We have for some constant $c>0$, 
\beao
\|\bfT\|_2^2&=&\|\bfT^2\|_2\le
c\,p^2\,
\|(\bfZ_n \bfZ_n' - \diag(\bfZ_n \bfZ_n'))^2\|_2\\ %&\le& c\,p^2\,\big[\|\bfD+\bfF+\bfR\|_2\big]\\
&\le &c\,p^2\,\|\bfD+\wt \bfF+\bfR\|_2+ c\,p^2 \,\|\bfT\|_2\,.
\eeao
Therefore
\beam\label{eq:inequality}
\dfrac{\|\bfT\|_2}{a_{np}^4}\le c\,\dfrac{p}{a_{np}^2}\,
\Big(\dfrac{\|\bfD+\wt \bfF+\bfR\|_2}{a_{np}^4}\Big)^{1/2}+
c\,\dfrac{p}{a_{np}^2} \Big(\dfrac{\|\bfT\|_2}{a_{np}^4}\Big)^{1/2}\,. 
\eeam
In the course of the proof of this lemma we show that
\beao
\dfrac{\|\bfD+\wt \bfF+\bfR\|_2}{a_{np}^4}\stp 0\,.
\eeao
Moreover, there is a small $\vep>0$ \sth
\begin{equation*}
\delta_n=\frac{p}{a_{np}^{2}}\le n^{1-4/\alpha+\vep}, \quad 1-4/\alpha+\vep<0\,.
\end{equation*}
Therefore iteration of \eqref{eq:inequality} yields for $k\ge 1$
\beam\label{eq:ko}
\dfrac{\|\bfT\|_2}{a_{np}^4}&\le& o_\P(1)
+ c\,\delta_n\Big(\delta_n\,
\Big(\dfrac{\|\bfD+\wt \bfF+\bfR\|_2}{a_{np}^4}\Big)^{1/2}\Big)^{1/2}
+c\,\delta_n \Big(\delta_n \Big(\dfrac{\|\bfT\|_2}{a_{np}^4}\Big)^{1/2}\Big)^{1/2}\nonumber\\
&=& o_\P(1)+c\,\Big(\delta_n^{4+2}\dfrac{\|\bfT\|_2}{a_{np}^4}\Big)^{1/4}\nonumber\\
&\le & o_\P(1)+c\,\Big(\delta_n^{2^k+\cdots +2}\dfrac{\|\bfT\|_2}{a_{np}^4}\Big)^{1/{2^k}}\,.
\eeam
Using some elementary moment bounds for $\|\bfT\|_2$ (e.g. a bound by the Frobenius norm),
it is not difficult to show that $n^{-l} \|\bfT\|_2\stp 0$ for some sufficiently large $l$. Thus we achieve that the \rhs\
in \eqref{eq:ko} converges to zero in \pro y. 
%The problem of showing that $a_{np}^{-4} \twonorm{T} \cip 0$ is almost the same as in Proposition~\ref{prop:offdiagonal} itself, where one was faced with $a_{np}^{-2} \twonorm{\bfZ_n \bfZ_n' - \diag(\bfZ_n \bfZ_n')}$. Fortunately, it turns out that the factor 
\par
It remains to show that $a_{np}^{-4} \twonorm{\widetilde{\bfF}}\stp 0$.
With the notation 
$B_{u,t}=\{Z_{ut}^2 \le a_{np}^{4-2\delta}  \} $
for some small $\delta>0$,
we decompose $Z_{it} Z_{jt} (Z_{ut}^2 - \E[Z^2 \1_{\{Z^2 \le a_{np}^{4-2\delta}\}}])$ as follows:
\begin{equation*}
\begin{split}
&Z_{it} Z_{jt} (Z_{ut}^2 \1_{B_{u,t}} - \E[Z^2 \1_{\{Z^2 \le a_{np}^{4-2\delta}\}}]) + Z_{it} Z_{jt} Z_{ut}^2 \1_{B_{u,t}^c}\,.
\end{split}
\end{equation*}
We decompose the matrix $\widetilde{\bfF}$ accordingly: 
\begin{equation*}
\widetilde{\bfF}= \widetilde{\bfF}^{(1)} + \widetilde{\bfF}^{(2)}\,,
\end{equation*}
\sth, for example, 
\begin{equation*}
\widetilde{F}^{(1)}_{ij} = \sum_{u=1;u\neq i,j}^p \sum_{t=1}^n  Z_{it} Z_{jt} (Z_{ut}^2 \1_{B_{u,t}} - \E[Z^2 \1_{\{Z^2 \le a_{np}^{4-2\delta}\}}])\,, \quad i \neq j.
\end{equation*} 
%In what follows we are going to prove that $a_{np}^{-4} \twonorm{F^{(kl)}}, k,l \in \{1,2\}$ converge to $0$ in probability.
\noindent
$\widetilde{\bfF}^{(1)}$: 
Bounding the spectral norm by the Frobenius norm, applying Markov's inequality and using Karamata's theorem together with the Potter bounds one can check that for $\epsilon>0$ and small $\delta>0$,
\begin{equation*}
\begin{split}
\P&(\twonorm{\widetilde{\bfF}^{(1)}}> \epsilon\, a_{np}^4) \le c\, a_{np}^{-8}\, \E \Big[ \sum_{i,j=1}^p (\widetilde{F}^{(1)}_{ij})^2  \Big]\\
&\le  c \,\frac{p^3\, n}{a_{np}^{8}}\E[(Z_1Z_2)^2 ] \E[(Z^2 \1_{\{Z^2 \le a_{np}^{4-2\delta}\}} - \E[Z^2 \1_{\{Z^2 \le a_{np}^{4-2\delta}\}}])^2 ]\\
&\le c \,\frac{p^3\, n}{a_{np}^{8}} \E[Z^4 \1_{\{Z^2 \le a_{np}^{4-2\delta}\}}] \\
&\le c \frac{p^3 n}{a_{np}^{4 \delta}} \P(|Z|> a_{np}^{2-\delta}) \to 0\,,
%\le c \frac{p}{n} (np)^{-\delta(4/\alpha -1)+ \vep} \to 0, 
\qquad \nto,
\end{split}
\end{equation*}
%since $-\delta(4/\alpha -1)+ \vep<0$ for $\vep$ small enough.
$\widetilde{\bfF}^{(2)}$: %The set $B_{u,t}^c$ defined in \eqref{eq:ab}  is contained in 
%\beao
%\{ |Z_{ut}^2-\E[Z^2]|>\min(C,C^{-1})a_{np}^{4-2\delta}  \}.
%\eeao
%Since $\twonorm{\cdot} \le \inftynorm{\cdot}$ for symmetric matrices, we get
We have for small $\delta>0$,
\beao%\label{eq:boundf12}
\P(\twonorm{\widetilde{\bfF}^{(2)}}> \epsilon\, a_{np}^4) &\le& 
\P\Big(\bigcup_{1\le u\le p\,,1\le t\le n} B_{u,t}^c\Big)\\
&\le & p\,n\, \P(|Z|> a_{np}^{2-\delta})\to 0\,,\qquad\,\nto\,.
\eeao
The proof of the $\bfF$-part is complete.
\end{proof}
\begin{proof}[Proof of the $\bfR$-part]

We have
\beao
\E [\|\bfR\|_2^2]\le \E[\|\bfR\|_F^2]&\le& \sum_{i,j=1}^p \sum_{u=1}^p \sum_{t_1=1}^n \sum_{t_2=1}^n (\E[Z^2])^4\le c\, p^3\, n^2\,.
\eeao
Therefore and by Markov's inequality for $\epsilon>0$,
\begin{equation}\label{eq:ljo}
\P(\twonorm{\bfR}> \epsilon\, a_{np}^4) %\le \P(\|R\|_F^2> c a_{np}^8)
%\le c \,\frac{\E[\|R\|_F^2]}{a_{np}^8}
\le c\, \frac{p^3 \,n^2}{a_{np}^8}
 \to 0, \qquad \nto\,,
\end{equation}
as long as $\alpha \in (2,16/5)$.
For $\alpha \in [16/5,4)$  
we use a similar idea for the truncated entries. Write $\bfR=\overline{\bfR} + \widetilde{\bfR}$, where for $i\neq j$
\begin{equation*}
\begin{split}
\overline{R}_{ij} &= \sum_{u=1;u\neq i,j}^p \sum_{t_1=1}^n \sum_{t_1=1; t_1 \neq t_2}^n Z_{i,t_1} Z_{j,t_2} Z_{u,t_1}Z_{u,t_2}\, 
\1_{A_{i,j,t_1,t_2}},\\
\widetilde{R}_{ij} &= \sum_{u=1;u\neq i,j}^p \sum_{t_1=1}^n \sum_{t_1=1; t_1 \neq t_2}^n Z_{i,t_1} Z_{j,t_2} Z_{u,t_1}Z_{u,t_2}\, 
\1_{A_{i,j,t_1,t_2}^c},
\end{split}
\end{equation*}
with $A_{i,j,t_1,t_2}^c=\{ -h(a_{np}) \le Z_{i,t_1}, Z_{j,t_2}, Z_{u,t_1},Z_{u,t_2}\le a_{np} \}$ and $h$ as in \eqref{eq:fcth}. 
Analogously to \eqref{eq:ljo}, using the fact that  $Z_{i,t_1} Z_{j,t_2} Z_{u,t_1}Z_{u,t_2}\, 
\1_{A_{i,j,t_1,t_2}}$ are uncorrelated for the considered index set,
one obtains for $\epsilon>0$, 
\beao
\P(\twonorm{\overline{\bfR}}> \epsilon\, a_{np}^4) &\le& c \,\frac{p^3\, n^2}{a_{np}^8}
 \,\E[Z^2 \1_{\{ |Z| > \min(C,C^{-1})\,a_{np}\}}]\\ &\le& c \,\frac{p^3 \,n^2}{a_{np}^6} \,\P(|Z| > \min(C,C^{-1}) \,a_{np})\to 0\,,\quad \nto\,,
\eeao
where we used Karamata's theorem and $\P(A_{i,j,t_1,t_2}) \le c\, \P(|Z|>\min(C,C^{-1})\,a_{np})$.
\par
%In the following we apply the Frobenius norm to higher powers of $\widetilde{R}$. 
We introduce the truncated \rv s $\wt Z_{it}=Z_{it}\1_{\{ -h(a_{np}) \le Z_{it} \le a_{np} \}}$ with the generic element $\wt Z$. 
We will repeatedly use the following inequality which is valid
for a real symmetric matrix $\bfM$:
\beao
\twonorm{\bfM}^2 \le \frobnorm{\bfM}^2 = \tr(\bfM^2).
\eeao
Then we have for $k\ge 1$, $\twonorm{\widetilde{\bfR}^{2^{k-1}}}^2 = \twonorm{\widetilde{\bfR}^{2^k}}$ and 
\beao
 \twonorm{\widetilde{\bfR}}^{2^k} \le \tr(\widetilde{\bfR}^{2^k})= \sum_{i,j=1}^p (\widetilde{R}^{2^{k-1}})_{ij}^2\,.
\eeao
This together with the Markov inequality of order $2^k$ yields
\begin{equation}\label{eq:sdfgb}
\P(  \twonorm{\widetilde{\bfR}}> ca_{np}^{4}) \le c a_{np}^{-4 \cdot 2^k} \E \Big[\sum_{i,j=1}^p (\widetilde{R}^{2^{k-1}})_{ij}^2 \Big] .
\end{equation}
Next we study the structure of $\widetilde{\bfR}^{2^{k-1}}$. The $(i,j)$-entry of this matrix is
\begin{equation}\label{eq:powerexpl}
(\widetilde{R}^{2^{k-1}})_{ij} = \sum_{i_1=1}^p %\sum_{i_2=1}^p  
\cdots \sum_{i_{2^{k-1}-1}=1}^p \widetilde{R}_{i,i_1} \widetilde{R}_{i_1,i_2}\cdots  \widetilde{R}_{i_{2^{k-1}-2},i_{2^{k-1}-1}} \widetilde{R}_{i_{2^{k-1}-1},j}.
\end{equation}
In view of \eqref{eq:powerexpl} and by definition of $\widetilde{\bfR}$, $(\widetilde{R}^{2^{k-1}} )_{ij}$ contains
exactly $2^k-1$ sums running from 1 to $p$, and $2^k$ sums running from 1 to $n$. 
Now we consider the expectation on the right-hand side of \eqref{eq:sdfgb}. 
The highest and lowest powers of $\wt Z_{it}$ in this expectation are $2^k$ and $1$. Let $(I,T) = ((i_1,t_1), \ldots,(i_{2^k},t_{2^k}))$. We have 
\begin{equation*}
\E \Big[\sum_{i,j=1}^p (\widetilde{R}^{2^{k-1}})_{ij}^2 \Big] = \sum_{(I,T)\in S} \E[\widetilde{Z}_{i_1,t_1} \widetilde{Z}_{i_2,t_2} \cdots \widetilde{Z}_{i_{2^k},t_{2^k}} 
],
\end{equation*}
where $S\subset \{1, \ldots,p\}^{2^k} \times \{1, \ldots,n\}^{2^k}$ is the index set that 
covers all combinations of indices that arise on the left-hand side. 
Since $\E[\widetilde{Z}]=0$, each $\widetilde{Z}$ in $\widetilde{Z}_{i_1,t_1} \widetilde{Z}_{i_2,t_2} \cdots \widetilde{Z}_{i_{2^k},t_{2^k}} $ 
must appear at least twice for the expectation of this product to be non-zero. Let $S_1 \subset S$ be the set of all those indices 
that make  a non-zero contribution to the sum. From the specific structure of $\widetilde{\bfR}$,  \eqref{eq:powerexpl} and the considerations above it now follows that the cardinality of $S_1$ has the following bound
\begin{equation*}
|S_1| \le c(k)\, p^2\, p^{2^k-1}\, n^{2^k}=c\, p^{2^k+1}\,n^{2^k}\,.
\end{equation*}
For $l=2,3$ we can use $\E[|\widetilde{Z}^l|]\le c$. If $l\ge 4$, we infer with Karamata's theorem 
\begin{equation}\label{eq:23424}
\E[|\widetilde{Z}^l|] \le c\, a_{np}^{l}\, \P(|Z| > a_{np}).
\end{equation}
The subset of $S_1$ (say $S_l$) which generates a  $\widetilde{Z}^l$ for $l\ge 4$ is much smaller than $S_1$. Also its cardinality is divided by at least $n$ if we go from $l$ to $l+1$, i.e. $|S_l|\ge n |S_{l+1}|$. Observe that $n a_{np}^{-1}$ converges to infinity. This combined with \eqref{eq:23424} tells us that only the case of every $\widetilde{Z}$ appearing exactly twice is of interest since it has most 
influence on the expectation in \eqref{eq:sdfgb}. We conclude that
\begin{equation*}
\begin{split}
\dfrac{1}{a_{np}^{4 \cdot 2^k}}\, \E \Big[\sum_{i,j=1}^p (\widetilde{R}^{2^{k-1}})_{ij}^2 \Big]
&\le c\, \dfrac{|S_1|}{a_{np}^{4 \cdot 2^k}}\le c\,p\, \Big( \frac{np}{a_{np}^4} \Big)^{2^k}\le c\, (np)\,  \Big( \frac{np}{a_{np}^4} \Big)^{2^{k}}.
\end{split}
\end{equation*}
The  expression on the \rhs\ converges to $0$ if $1+2^{k}-2^{k+2}/\alpha <0$ or equivalently
\beao
k>\log\Big( \frac{\alpha}{4-\alpha}\Big) (\log 2)^{-1}.
\eeao
Since $k$ was arbitrary the proof of the $\bfR$-part is finished.
\end{proof}

\subsection{Proof of \eqref{eq:eigena}}
We define the $p\times p$ matrix $\Y_n^\rightarrow$ as the diagonal matrix with elements
\begin{equation*}
(\Y_n^\rightarrow)_{ii}= \max_{t=1, \ldots,n} Z_{it}^2\,, \qquad i=1,\ldots,p\,.
\end{equation*}
Correspondingly, we define the $n\times n$ matrix $\Y_n^\downarrow$ as the diagonal matrix with elements
\begin{equation*}
(\Y_n^\downarrow)_{tt}= \max_{i=1, \ldots,p} Z_{it}^2\,, \qquad t=1,\ldots,n\,.
\end{equation*}
\begin{lemma}\label{cor:688}
Assume the conditions of Theorem~\ref{thm:iidmain}. 
\begin{enumerate}
\item
If $\beta\in ((\alpha/2-1)_+,1]$ we have
%$\beta\in (0,1]\cap (\alpha/2-1,1]$ we have
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\Y_n^\rightarrow)\big|\stp 0\,,\qquad\nto \,.
\eeao
\item 
If $\beta^{-1}\in ((\alpha/2-1)_+,1)$ we have
%$\beta\in (0,1]\cap (\alpha/2-1,1]$ we have
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,n}\big|\la_{(i)}-\la_{(i)}(\Y_n^\downarrow)\big|\stp 0\,,\qquad\nto \,.
\eeao
\end{enumerate}
\end{lemma}
\begin{proof}
We restrict ourselves to the proof in the case $\beta\in (0,1]$; the case $\beta>1$ can again be handled by switching 
from $\bfZ\bfZ'$ to $\bfZ'\bfZ$.
An application of Weyl's inequality (see \eqref{eq:weylin}) and the triangle inequality yield
\begin{equation*}
\begin{split}
a_{np}^{-2}\!\max_{i=1,\ldots,p}\big|\la_{(i)}\!-\!\la_{(i)}(\Y_n^\rightarrow)\big|& \le 
a_{np}^{-2} \twonorm{\bfZ \bfZ' - \diag(\bfZ \bfZ')} + a_{np}^{-2} \twonorm{ \diag(\bfZ \bfZ')- \diag(\Y_n^\rightarrow)}.
\end{split}
\end{equation*}
The first term on the right-hand side converges to $0$ in probability by Theorem~\ref{prop:offdiagonal}(1).
As regards the second term we have 
\beao
a_{np}^{-2} \twonorm{ \diag(\bfZ \bfZ')- \diag(\Y_n^\rightarrow)}&=& a_{np}^{-2}\max_{i=1,\ldots,p} \Big|D_i^\rightarrow - \max_{t=1,\ldots,n} Z_{it}^2\Big|\,.
\eeao
The \rhs\ converges to zero in \pro y in view of Lemma~\ref{thm:keyresult} applied to $(Z_{it}^2)$.
\end{proof}
Now \eqref{eq:eigena} follows from the next result.
\begin{lemma}\label{lem:redorder}
Assume the conditions of Theorem~\ref{thm:iidmain}. 
\begin{enumerate}
\item
If $\beta\in ((\alpha/2-1)_+,1]$ we have
%$\beta\in (0,1]\cap (\alpha/2-1,1]$ we have
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}(\Y_n^\rightarrow)-Z_{(i),np}^2\big|\stp 0\,,\qquad\nto \,.
\eeao
\item 
If $\beta^{-1}\in ((\alpha/2-1)_+,1)$ we have
%$\beta\in (0,1]\cap (\alpha/2-1,1]$ we have
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,n}\big|\la_{(i)}(\Y_n^\downarrow)-Z_{(i),np}^2\big|\stp 0\,,\qquad\nto \,.
\eeao
\end{enumerate}
\end{lemma}
%\begin{theorem}\label{thm:main}
%Assume the conditions of Theorem~\ref{thm:iidmain}.
%$\alpha \in (0,4)$. { Check $\beta^{-1}$} If $\beta \in (0,1] \cap (\alpha/2-1,1]$ then 
%\begin{equation*}
%a_{np}^{-2} \max_{i=1,\ldots,p} |\lambda_{(i)}-Z_{(i),np}^2| \cip 0\,, \quad \nto.
%\end{equation*}
%\end{theorem}
\begin{proof}
We focus on part (1). We write 
$V_{(1)}\ge \cdots \ge V_{(p)}$ for the order statistics of $(\max_{t=1,\ldots,n}Z_{it}^2)$. By definition of the order statistics we have $Z_{(i),np}^2\ge V_{(i)}$ for $i=1,\ldots,p$. We choose $\delta$ such that $1>\delta >\frac{2+\beta}{2(1+\beta)}$ and define the event
\beao
B_{np}^{2\delta}= \{ \mbox{There is a row of $(Z_{it}^2)$ with at least two entries larger than $a_{np}^{2\delta}$.} \}\,.
\eeao
By Lemma~\ref{lem:onecolumn}, $\P(B_{np}^{2\delta})\to 0$.

Next, we choose $0<\vep < 1-\delta$. Then Lemma~\ref{prop:seqk} guarantees the existence of a sequence $k=k_n \to \infty$ such that the event 
\begin{equation*}
\Omega_n = \{ Z_{(k),np}^2 > a_{np}^{2(1-\vep)} \}
\end{equation*}
satisfies $\P(\Omega_n^c)\to 0$. On the event $(B_{np}^{2\delta})^c \cap \Omega_n$ we have 
\begin{equation*}
V_{(i)}-Z_{(i),np}^2=0\,, \quad i=1,\ldots,k.
\end{equation*}
This shows for $\gamma>0$,
\begin{equation*}
\begin{split}
\limsup_{\nto}& \,\P\big(a_{np}^{-2}\max_{i=1,\ldots,p} |V_{(i)}-Z_{(i),np}^2|>\gamma\big)\\
&\le \limsup_{\nto}\P\big(\{a_{np}^{-2}\max_{i=1,\ldots,p} |V_{(i)}-Z_{(i),np}^2|>\gamma\}\cap (B_{np}^{2\delta})^c \cap \Omega_n\big)\\
&\quad + \limsup_{\nto} \P(B_{np}^{2\delta})  +\limsup_{\nto} \P(\Omega_n^c)\\
&= \limsup_{\nto}\P\big(\{a_{np}^{-2}\max_{i=k+1,\ldots,p} |V_{(i)}-Z_{(i),np}^2|>\gamma\}\cap (B_{np}^{2\delta})^c \cap \Omega_n\big)\\
&\le \limsup_{\nto}\P\big( 2 \, a_{np}^{-2} \,Z_{(k+1),np}^2 >\gamma \big) =0.
\end{split}
\end{equation*}
\end{proof}

%--------------------------------------------------------------
\section{Generalization to autocovariance matrices}\label{sec:autocov}

An important topic in multivariate time series analysis is the study of the covariance structure. From the field $(Z_{it})$ we construct the $p\times n$ matrices
\beao
\z(s,k)=\z_n(s,k)=(Z_{i-s,t-k})_{i=1,\ldots,p;t=1,\ldots,n}\,, \quad s,k\in\Z\,.
\eeao
We introduce the (non-normalized) {\em generalized sample autocovariance matrices}  
\beao%\label{eq:Ygsrg}
(\z(0,0)\z(s,k)')\,, \quad s,k\in\Z\,,
\eeao
with entries
\beao%\label{eq:Ygsrg}
(\z(0,0)\z(s,k)')_{ij} = \sum_{t=1}^n Z_{i,t} \,Z_{j-s,t-k}\,,\qquad i,j=1,\ldots,p\,.
\eeao
If $\min(|s|,|k|) \neq 0$, the generalized sample autocovariance matrix $\z(0,0)\z(s,k)'$ is not symmetric and might thus have complex eigenvalues. In what follows, we will be interested in the {\em singular values} $\la_1(s,k), \ldots, \la_p(s,k)$ of $\z(0,0)\z(s,k)'$.  The singular values of a matrix $\bfA$ are the square roots of the eigenvalues of $\bfA\bfA'$. We reuse the notation $(\la_i(s,k))$ for the singular values 
%since in the case $s=k=0$ the singular values of the sample covariance matrix coincide with its eigenvalues.
and again write $\la_{(1)}(s,k)\ge \cdots  \ge\la_{(p)}(s,k)$ for their order statistics.
\par
%{ This notation has been introduced before.}
%We need some more notation. For $s\in \Z$ let $s_+, s_-\ge 0$ be the positive and negative parts of $s$, respectively. 
%We will use the order statistics 
%\begin{equation*}
%Z_{(1),(n(p-|s|)}^2 \ge Z_{(2),(n(p-|s|)}^2 \ge  \cdots \ge Z_{(n(p-|s|)),(n(p-|s|)}^2, \qquad n,p\ge 1\,,
%\end{equation*}
%of the random variables $(Z_{it}^2)$, $i=1+s_-,\ldots,p-s_+; t=1,\ldots,n$.
%Important objects are the row-sums 
%\begin{equation*}%\label{eq:ll}
%D_i^{\rightarrow}=D_i^{(s),(n),\rightarrow}=\sum_{t=1}^n Z_{it}^2\,, \qquad
%i=1+s_-,\ldots,p-s_+;\quad  s\in\Z, n=1,2,\ldots\,,
%\end{equation*} 
%and their ordered values
%$D_{(1)}^{(s),\rightarrow}\ge \cdots \ge D_{(p-|s|)}^{(s),\rightarrow}$. Similarly we also define the order statistics $D_{(1)}^{(s),\downarrow}\ge \cdots \ge D_{(n-|s|)}^{(s),\downarrow}$ of the column-sums $D_t^{\downarrow}$.
\begin{theorem}\label{thm:iidauto}
Assume $s,k\in \Z$. Consider the $p\times n$-dimensional matrices $\z(0,0)$ and $\z(s,k)$ with iid entries. We assume the following conditions:
\begin{itemize} \item
The \regvar\ condition \eqref{eq:regvar} for some  
$\alpha \in (0,4)$. 
\item $\E [Z]=0$ for $\alpha\ge 2$.
\item
The integer \seq\ $(p_n)$ 
has growth rate \ref{eq:p} for some  $\beta\ge 0$.
\end{itemize}
\begin{itemize}
\item[(1)] If $k\neq 0$, then 
\beao
a_{np}^{-2} \, \la_{(1)}(s,k) \cip 0\,.
\eeao
\end{itemize}
Now assume $k=0$ and recall the notation $D_{(i)}^{\rightarrow}$ and $D_{(i)}^{\downarrow}$ from Section \ref{sec:section2.2}. Then the following statements hold:
\begin{itemize}
\item[(2)] If $\beta\in [0,1]$, then
\beam\label{eq:rowauto}
a_{np}^{-2}\,\max_{i=1,\ldots,p-|s|}\big|\la_{(i)}(s, 0)-D_{(i)}^{\rightarrow}\big|\stp 0\,.
\eeam

\item[(3)] If $\beta > 1$, then
\beam\label{eq:colauto}
a_{np}^{-2}\,\max_{i=1,\ldots,n-|s|}\big|\la_{(i)}(s, 0)-D_{(i)}^{\downarrow}\big|\stp 0\,.
\eeam

\item[(4)] If  $\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$, then
\beam\label{eq:eigenauto}
a_{np}^{-2}\,\max_{i=1,\ldots,p-|s|}\big|\la_{(i)}(s, 0)-%Z_{(i),(n(p-|s|)}^2 
Z_{(i),np}^2\big|\stp 0\,.
\eeam
\end{itemize}
\end{theorem}
\begin{proof}
We focus on the case $\beta \in [0,1]$.  The proof is analogous to the proof of
Theorem~\ref{thm:iidmain} which was given in Section~\ref{sec:proofs}. This proof relied on the reduction of $\z\z'$ to its diagonal. If $k=0$, we will reduce $\z(0,0)\z(s,k)'$ to 
 a $p\times p$ matrix $\M^{(s,k)}$, which only takes values on its $s$th sub-diagonal.
The entries of the $s$th sub-diagonal of $\M^{(s,k)}$ are $\M^{(s,k)}_{i,i+s}$, $i=1+s_-,\ldots,p-s_+$. Here $s_+, s_-\ge 0$ are the positive and negative parts of $s$, respectively.

We sketch the steps of this reduction. Let $k\in\Z$. For simplicity of notation assume $s \ge 0$.
Define the $p\times p$ matrix $\M^{(s,k)}$,
\begin{equation*}
\M^{(s,k)}_{i,i+s}=\1_{\{ k=0\}}(\z(0,0)\z(s,0)')_{i,i+s}= \1_{\{ k=0\}}\sum_{t=1}^n Z_{it}^2\,, \quad i=1,\ldots,p-s\,,
\end{equation*}
and $\M^{(s,k)}_{ij}=0$ for all other $i,j$. We have
\begin{equation*}
\begin{split}
\big((\z(0,0)\z(s,k)'-&\M^{(s,k)}) (\z(0,0)\z(s,k)'-\M^{(s,k)})'\big)_{ij}\\&=
\sum_{u=1}^p \sum_{t_1=1}^n \sum_{t_2=1}^n Z_{i,t_1} Z_{j,t_2} Z_{u-s,t_1-k}Z_{u-s,t_2-k} \1_{\{ i\neq u-s, j\neq u-s \}}\\
&\quad \times (\1_{\{ i=j\}}+\1_{\{i\neq j, t_1=t_2\}}+\1_{\{i\neq j, t_1\neq t_2\}})\\
&= \bfD_{ij}+\bfF_{ij}+\bfR_{ij}\,.
\end{split}
\end{equation*}
Repeating the steps in the proof of Lemma~\ref{lem:diagonal}, one obtains 
\beao
a_{np}^{-4}\twonorm{\bfD+\bfF+\bfR}^2\stp 0\,.
\eeao
Therefore we also have
\begin{equation*}
\begin{split}
a_{np}^{-4}&\twonorm{\z(0,0)\z(s,k)'-\M^{(s,k)}}^2\\ &= 
a_{np}^{-4}\twonorm{(\z(0,0)\z(s,k)'-\M^{(s,k)}) (\z(0,0)\z(s,k)'-\M^{(s,k)})'}\stp 0\,.
\end{split}
\end{equation*}

This proves part (1). Since, with probability tending to $1$, the matrix $\M^{(s,k)}$ has the required singular values, part (2) follows by Weyl's inequality.

Finally, part (4) is a consequence of Lemma~\ref{lem:redorder}.
\end{proof}
We obtain the following result for the weak \con\ of the point processes of the points $\la_i(s,0)$, $s=0,\ldots,l$; the proof is similar
to the one of Theorem~\ref{thm:iidauto}. 
\begin{corollary}
Assume the conditions of Theorem~\ref{thm:iidauto}. 
%Let $\xi_{\alpha/2}$ be a \Frechet distributed random variable with parameter $\alpha/2$.
Then, with the notation of Theorem~\ref{cor:1}, the following \pp\ \con\ holds for $ l\ge 0$ and $(\beta,\beta^{-1})\in ((\alpha/2-1)_+,1]$,
%\begin{enumerate}
%\item If either $\alpha\in (0,2)$ and $\beta\ge 0$, or $\alpha\in [2,4)$ and $\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$, then
\beao
\sum_{i=1}^{ p} \vep_{a_{np}^{-2} \, \big(\la_{ (i)}({ 0},0),\ldots,\la_{ (i)}(l,0)\big)} \cid \sum_{i=1}^\infty 
\vep_{\Gamma_i^{-2/\alpha}\big(1,\ldots,1\big)} \,.
\eeao
%and,
%{\blue delete, not used in paper with Richard} for $\gamma<\alpha/2$,
%the \seq\ $\big((a_{np}^{-2} \la_{(1)}(s,0))^\gamma\big)_{n\ge 1}$ is uniformly integrable.
%\item If  $\alpha\in (2,4)$ and $\beta\ge 0$, then
%\begin{equation*}
%a_{np}^{-2} \, (\la_{(1)}(s,0) -(n\vee p) \E[Z^2] ) \cid \xi_{\alpha/2}\,,
%\end{equation*}
%{\blue delete this part, not used in the paper with Richard} and, for $\gamma<\alpha/2$,
%the \seq\ $\big((a_{np}^{-2} |\la_{(1)}(s,0) -(n\vee p) \E[Z^2] |)^\gamma\big)_{n\ge 1}$ is uniformly integrable.
%\end{enumerate}
\end{corollary}
The joint \con\ of a finite number of the random variables $\la_{(i)}(s,0)$, $i\ge 1$, $s\ge 0$, is an immediate con\seq\ of this result.
%{\red 
%\begin{proof}
%The distributional convergence in part (1) follows from Theorem~\ref{thm:iidauto}. Indeed, if $\alpha\in (0,2)$ and $\beta\ge 0$
%we have $a_{np}^{-2} D_{(1)}^{\rightarrow}\cid \xi_{\alpha/2}$ and $a_{np}^{-2} D_{(1)}^{\downarrow}\cid \xi_{\alpha/2}$ and if 
%$\alpha\in (2,4)$ and $\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$ we have the corresponding results $a_{np}^{-2} (D_{(1)}^{\rightarrow}-n\,\E[Z^2])\cid \xi_{\alpha/2}$
%and $a_{np}^{-2} (D_{(1)}^{\downarrow}-p\,\E[Z^2])\cid \xi_{\alpha/2}$; if $\alpha=1$ one needs the corresponding modification for the form of the centering constants. However, we also have $(p\vee n)/a_{np}^2\to 0$ \st\ centering is not needed.
%\begin{equation*}
%a_{np}^{-2} Z_{(1),(n(p-|s|)}^2 \cid \xi_{\alpha/2}\,,
%\end{equation*}
%if $\alpha\in [2,4)$ and $\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$.
%\par
%The same arguments combined with Lemma~\ref{lem:pp} conclude the proof of part (2).
%\end{proof}

%-----------------------------------------------------------------------

\section{Appendix}\label{appendix:A}

Let $(Z_i)$ be iid copies of $Z$ whose distribution satisfies 
\begin{equation*}
\P(Z>x)\sim p_+ \dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad  \P(Z\le -x)\sim p_-
\dfrac{L(x)}{x^{\alpha}} \,,\quad \xto\,,
\end{equation*}
 for some tail index $\alpha>0$,
where $p_+,p_-\ge 0$ with $p_++p_-=1$ and $L$ is a slowly varying function. We say that $Z$ is regularly varying with index $\alpha$. The monograph \cite{bingham:goldie:teugels:1987} contains many properties and useful tools for regularly varying functions. Theorem 1.5.6 therein, which is known as Potter bounds, asserts that a regularly varying function essentially lies between two power laws. In particular,  
for any $\delta>0$ and $C>1$ we have for $x$ sufficiently large,
\begin{equation*}
C^{-1} x^{-\delta} \le L(x) \le C x^{\delta}\,.
\end{equation*}

Theorem 1.6.1 in \cite{bingham:goldie:teugels:1987}, widely known as Karamata's theorem, describes the behavior of truncated moments of the regularly varying random variable $Z$. For $x\to \infty$,
\begin{equation*}
\begin{split}
\E[|Z|^\beta \1_{\{|Z|\leq x\}}] &\sim \frac{\alpha}{\beta -\alpha} x^\beta \P(|Z|>x), \quad \beta>\alpha,\\
\E[|Z|^\beta \1_{\{|Z| > x\}}] &\sim \frac{\alpha}{\alpha-\beta} x^\beta \P(|Z|>x), \quad \beta<\alpha.
\end{split}
\end{equation*}

If $\E[|Z|]<\infty$ also assume $\E[Z]=0$. The product $Z_1Z_2$ is regular varying with the same index $\alpha$ and $\P(|Z_1Z_2|>x)= x^{-\alpha} L_1(x)$, where $L_1$ is slowly varying function different from $L$;
see Embrechts and Goldie \cite{embrechts:goldie:1980}.
Write
\begin{equation*}
S_n=Z_1+\cdots +Z_n\,,\quad n\ge 1,
\end{equation*} and consider a sequence $(a_n)$ such that $\P(|Z|>a_n)\sim n^{-1}$.

\subsection{Large deviation results}
The following theorem can be found in
Nagaev \cite{nagaev:1979} and Cline and Hsing
\cite{cline:hsing:1998} for $\alpha>2$ and $\alpha\le 2$,
respectively; see also  Denisov et al.~\cite{denisov:dieker:shneer:2008}.
\begin{theorem}\label{thm:nagaev} 
Under the assumptions on the iid sequence $(Z_t)$
given above the following relation holds
\begin{equation*}
\sup_{x\ge c_n}\left|\dfrac{\P(S_n>x )}{n\P(|Z|>x)} -p_+ \right|\to 0\,,
\end{equation*}
where $(c_n)$ is any sequence satisfying $c_n/a_n\to  \infty$ for
$\alpha\le 2$ and $c_n\ge \sqrt{(\alpha-2)n\log n}$ for $\alpha>2$.
\end{theorem}

\subsection{Karamata theory for sums}
\begin{proposition}\label{prop:karasumsbeta}
Let $(c_n)$ be the threshold sequence in Theorem~\ref{thm:nagaev} for a
given $\alpha>0$,  and let $(d_n)$ be such that
$d_n/c_n\to\infty$ for $\alpha>2$ and $d_n=c_n$ for $\alpha\le 2$. Assume $0<\gamma< \alpha$.
Then we have for a sequence $x_n\ge d_n$
\begin{equation}
\E[|x_n^{-1} S_n|^\gamma \1_{\{|S_n|>x_n\}}] \sim \frac{\alpha}{\alpha -\gamma} n \P(|Z|>x_n), \qquad \nto.
\end{equation}
\end{proposition}
\begin{proof}
We use the notation $Y_n:=|x_n^{-1} S_n|$. Since $Y_n^\gamma \1_{\{Y_n>1\}}$ is a positive random variable one can write 
\begin{equation*}
\E[Y_n^\gamma \1_{\{Y_n>1\}}]=\int_0^\infty \P(Y_n^\gamma \1_{\{Y_n>1\}}>y) \dint y.
\end{equation*}
The probability inside the integral is 
\begin{equation*}
\begin{split}
\P(Y_n^\gamma\1_{\{Y_n>1\}}>y) &= \P(Y_n^\gamma \1_{\{Y_n>1\}}>y, Y_n>1)+\P(Y_n^\gamma \1_{\{Y_n>1\}}>y, Y_n<1)\\
&=\P(Y_n^\gamma >y, Y_n>1)= \P(Y_n > \max \{ y^{1/\gamma},1\})\\
&= \begin{cases}
\P(Y_n>1) & \text{if } y\le 1,\\ 
\P(Y_n>y^{1/\gamma}) & \text{if } y\ge 1.
\end{cases}
\end{split}
\end{equation*}
Therefore, using the uniform convergence result in Theorem~\ref{thm:nagaev}, we conclude that
\begin{equation*}
\begin{split}
\int_0^\infty \P(Y_n^\gamma \1_{\{Y_n>1\}}>y) \dint y &= \P(Y_n>1) + \int_1^\infty \P(Y_n>y^{1/\gamma}) \dint y\\
&\sim n \P(|Z|>x_n)+\int_1^\infty y^{-\frac{\alpha}{\gamma}} n \P(|Z|>x_n) \dint y\\
&= \frac{\alpha}{\alpha -\gamma} n \P(|Z|>x_n), \quad \nto.
\end{split}
\end{equation*}
\end{proof}

\subsection{A point process convergence result}
Assume that the conditions at the beginning of Appendix \ref{appendix:A} hold.
Consider a sequence of iid copies $(S_{n}^{(t)})_{t=1,2,\ldots}$
of $S_n$ and the sequence of point processes
\begin{equation*}
N_n= \sum_{t=1}^{p} \vep_{a_{np}^{-1} S_{n}^{(t)}}, \quad n=1,2,\ldots\,,
\end{equation*}
for an integer sequence $p=p_n\to\infty$. We assume that the state space of the
point processes $N_n$ is $\overline{\R}_0=[\R\cup\{\pm \infty\}]\backslash \{0\}$.
\begin{lemma}\label{lem:ppr} 
Assume $\alpha \in (0,2)$ and the
conditions of  Appendix~\ref{appendix:A}
on the iid sequence $(Z_t)$ and the normalizing sequence $(a_n)$. Then the limit relation
$N_n\cid N$ holds in the space of point measures on $\overline{\R}_0$
equipped with the vague topology (see
\cite{resnick:1987,resnick:2007})
for a  Poisson random measure $N$ with state space $\overline{\R}_0$ and intensity measure $\mu_\alpha(dx)=\alpha |x|^{-\alpha-1} (p_+ \1_{\{x>0\}}+ p_- \1_{\{x<0\}}) 
dx$.
\end{lemma}
\begin{proof}
According to Resnick \cite{resnick:1987}, Proposition 3.21, we need to
show that
$p\, \P(a_{np}^{-1}S_n\in \cdot)\civ \mu_\alpha
$,
where $\civ$ denotes vague convergence of Radon measures on  $\overline{\R}_0$.
Observe that we have $a_{np}/a_n\to\infty$ as $\nto$. This fact and
$\alpha\in (0,2)$ allow one to apply
Theorem~\ref{thm:nagaev}:
\begin{equation*}
\dfrac{\P( S_n >x a_{np})}{n\,\P(|Z|>  a_{np})}\to p_+ x^{-\alpha}
\quad \mbox{and}\quad \dfrac{\P( S_n \le -x a_{np})}{n\,\P(|Z|>
  a_{np})}\to p_-\, x^{-\alpha}\,,\quad x>0\,.
\end{equation*}
On the other hand, $n\,\P(|Z|>  a_{np})\sim p^{-1}$ as $\nto$.
This proves the lemma.
\end{proof}

\subsection{Auxiliary results}
%By construction, the entries of $\X_n$  are the order statistics $Z_{(1),np}^2, \ldots, Z_{(np),np}^2$. If we consider the first $k$ of them, we expect them to be larger than $a_{np}^{2\delta}$ with high probability, since $2\delta\le 2$ and we know that $a_{np}^{2}$ is the ``correct'' normalization according to the above point process results. This intuition will be confirmed by the next result.
Assume that the non-negative random variable $Z$ is regularly varying with index $\alpha \in (0,2)$  and $(a_n)$ is such that $n \,\P(Z>a_n)\sim 1$.
We also write 
\beao
Z_{(1)}\ge \cdots \ge Z_{(n)}\,,
\eeao
for the order statistics of the iid copies $Z_1,\ldots,Z_n$ of $Z$. 
\begin{lemma}\label{prop:seqk}
For every $\vep \in (0,0.5)$ there exists a sequence $k=k_n\to \infty$, $k<n$ such that 
\begin{equation*}
\lim_{\nto} \P(Z_{(k)} > a_{n}^{1-\vep}) = 1.
\end{equation*}
%{\blue ? possible choice for $(k_n)$ is $k_n= \vep \,\log (n)$.}
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{prop:seqk}]
From the theory of order statistics we know that
\beao
\P(Z_{(k)} \le a_{n}^{1-\vep}) &=& \sum_{r=0}^{k-1}\binom{n}{r} \P(Z>a_{n}^{1-\vep})^r \; \P(Z\le a_{n}^{1-\vep})^{n-r}\\
&\le& \big(\P(Z\le a_{n}^{1-\vep})\big)^{n} \sum_{r=0}^{k-1} \frac{1}{r!} \Big( \frac{n \, \P(Z>a_{n}^{1-\vep})}{\P(Z\le a_{n}^{1-\vep})} \Big)^r.
\eeao
We observe that
\beao
\big(\P(Z\le a_{n}^{1-\vep})\big)^n \sim \ex^{- n\,\big[\P(Z>a_n^{1-\vep})-0.5(\P(Z>a_n^{1-\vep}))^2(1+o(1))\big]\,} 
\eeao
Writing $\Gamma(k)$ and $\Gamma(k,y)$ for the gamma and incompete gamma \fct s, we have
\beao
\ex^{-y}\sum_{r=0}^{k-1} \frac{y^r}{r!} = \frac{\Gamma(k,y)}{\Gamma(k)}=\P(\Gamma_k>y), \qquad y\ge 0\,,
\eeao
where $\Gamma_k=E_1+\cdots +E_k$, $k\ge 1$, for an iid standard exponential \seq\ $(E_i)$.
Therefore
\beao\lefteqn{\P(Z_{(k)} \le a_{n}^{1-\vep})}\\ &\le& c\,\ex^{- n\,\big[\P(Z>a_n^{1-\vep})
-0.5(\P(Z>a_n^{1-\vep}))^2(1+o(1))\big]+  \big[n\, \P(Z>a_{n}^{1-\vep})/\P(Z\le a_{n}^{1-\vep})\big]}\\
&&\P\big(\Gamma_k> n\, \P(Z>a_{n}^{1-\vep})/\P(Z\le a_{n}^{1-\vep})\big) \\&=&
c\,\ex^{O\big(n\,(\P(Z>a_n^{1-\vep}))^2\big)}\,\P\big(k^{-1}\Gamma_k> k^{-1}n\, \P(Z>a_{n}^{1-\vep})/\P(Z\le a_{n}^{1-\vep})\big) \,.
\eeao
The \rhs\ converges to zero if $2\vep<1$ and $k\le n^{\vep'}$ for some $\vep'<\vep$. 
\end{proof}
Now consider a  $p\times n$ random matrix $\bfZ$ with iid non-negative entries $Z_{it}$ and generic element $Z$ as specified above. The number of rows $p$ satisfies the growth condition \ref{eq:p}.
\par
We write for $\delta>0$,
\beam
B_{np}^\delta&=& \{ \mbox{There is a row of $\bfZ$ with at least two entries larger than $a_{np}^\delta$.} \}\,.\label{eq:bnp}
%\widetilde{B}_{np}^\delta&=& \{ \mbox{There is a column of $\bfZ$ with two entries larger than $a_{np}^\delta$.}\}\,.\nonumber
\eeam
\begin{lemma}\label{lem:onecolumn}
Assume that $p=p_n$ satisfies the growth condition \ref{eq:p} with $\beta\in [0,1]$. Then we have 
\begin{equation*}
\lim_{\nto} \P(B_{np}^\delta) =0 \quad \mbox{for all} \quad \delta>\frac{2+\beta}{2(1+\beta)}.
\end{equation*} 
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:onecolumn}]
Assume  $\delta>\frac{2+\beta}{2(1+\beta)}$
and consider the counting variables
\begin{equation*}
N_i=\sum_{t=1}^n \1_{\{Z_{it}>a_{np}^\delta\}}, \qquad i=1,\ldots,p.
\end{equation*}
Clearly, $N_i$ are iid $\text{Bin}(n, q)$   
with $q=q_n= \P(Z>a_{np}^\delta)\to 0$ as $\nto$  and
\beao
\P(B_{np}^\delta)& =& \P(\max_{i=1,\ldots,p} N_i \ge 2)\\ &=& 1-\big(\P(N_1\le 1)\big)^p\\
&=& 1- \big((1-q)^{n-1} (1+(n-1)q)\big)^p\,.
\eeao
Thus it remains to show that the \rhs\ converges to 0. Taking  logarithms, we get
\beao%\label{eq:lem5a11}
p\,\log \big((1-q)^{n-1} \,(1+(n-1)q)\big)= p\, [(n-1) \log(1-q)+\log(1+(n-1)q)].
\eeao
%We observe that $nq\to 0$ if $\delta>(1+\beta)^{-1}$. %This conditions holds if $\delta>(2+\beta)/(2(1+\beta))$.
A second order Taylor expansion of the logarithm yields 
\beam\label{eq:lem5a21}
p\,(n-1)\, \log(1-q)+p\,\log(1+(n-1)q)= p\,q+p\,\frac{(nq)^2}{2} +O(p\,(nq^2+(nq)^3)\,)\,.
\eeam 
By the Potter bounds we conclude that
\eqref{eq:lem5a21} 
converges to zero if $\delta>\frac{2+\beta}{2(1+\beta)}$.
%\begin{equation}\label{eq:lem5a31}
%c\; p n^2q^2 \le c\; n^{\beta + \vep_1} n^2 n^{2(1+\beta-\vep_1)(-\delta+\vep_2)}= c\;n^{2+\beta+\vep_1+ 2(1+\beta-\vep_1)(-\delta+\vep_2)},
%\end{equation}
%where $\vep_1, \vep_2>0$ can be chosen arbitrarily small. The \rhs\ of \eqref{eq:lem5a31} converges to $0$ since the exponent of $n$ is negative, that is
%\begin{equation*}
%2+\beta+\vep_1+ 2(1+\beta-\vep_1)(-\delta+\vep_2)<0 \quad \Leftrightarrow \quad \delta>\frac{2+\beta+\vep_1}{2(1+\beta-\vep_1)}+\vep_2,
%\end{equation*}
%which is fulfilled for sufficiently small $\vep_1, \vep_2>0$. This finishes the proof of $\P(B_{np}^\delta) \to 0$.
The proof is complete. 
\end{proof}
For $\vep\in (0,1)$
define the events
\beao
A_i^{(n)}(\vep)&=& 
\Big\{ \sum_{t=1}^n Z_{it}- \max_{t=1, \ldots,n} Z_{it}> a_{np}^{1-\vep}  \Big\}, \qquad i=1,\ldots,p\,.
\eeao
The following result generalizes Lemma 5 in 
Auffinger et al.~\cite{auffinger:arous:peche:2009} (which in turn is a 
modified version of a result in Soshnikov~\cite{soshnikov:2004}) to the case of \regvary\ growth rates $(p_n)$.
The method of proof is different from the aforementioned literature.
\begin{lemma}\label{thm:keyresult}
Assume that $p=p_n=n^\beta \ell(n)$ where $\ell$ is a slowly varying function. 
Assume $ \beta\in (0,\infty)$ for $\alpha\in (0,1]$ and $\beta  \in (\alpha-1,\infty)$ for $\alpha\in [1,2)$.
There exists a constant $\vep\in (0,1)$ such that 
\begin{equation*}
\lim_{\nto} \P \Big( \bigcup_{i=1}^p A_i^{(n)}(\vep) \Big)=0\,.
\end{equation*}
\end{lemma}
\begin{proof}
Write $M_t=\max_{i=1,\ldots,t}Z_i$.
We observe that
\beao
\P \Big( \bigcup_{i=1}^p A_i^{(n)}(\vep) \Big)&\le & p\, \P(S_n-M_n>a_{np}^{1-\vep})\\
&=& n\,p\,\P(S_{n-1}>a_{np}^{1-\vep}\,,Z_n>M_{n-1})\\
&=&n\,p\,\int_0^\infty \P(S_{n-1}>a_{np}^{1-\vep}\,,z>M_{n-1})\, \dint \P(Z\le z)\,.
\eeao
We split the integration area into disjoint sets:
\beao
[0,\infty)= [0,a_n/h_n]\cup (a_n/h_n,a_{np}^{\gamma}]\cup (a_{np}^{\gamma},\infty) =\bigcup_{i=1}^3 B_i\,.
\eeao
We choose $h_n\to\infty$ \sth\ $n\,\P(Z>a_n/h_n)\sim 2\log(np)$. Then 
\beam\label{eq:hp}
\log (np)- n\,\P(Z>a_n/h_n)\to -\infty\,,\qquad n\,\big(\P(Z>a_n/h_n)\big)^2\to 0\,.
\eeam
Moreover, choose $\gamma$ and $\vep>0$ fixed \sth\ $\vep<1-(1\vee \alpha)/(1+\beta)$ and
\begin{itemize}
\item $\frac{1}{1+\beta}+\vep< \gamma <1-\frac{\vep}{1-\alpha}$ if $\alpha \in (0,1)$ and
\item $\frac{1}{1+\beta}+\vep< \gamma <1-\frac{2\vep}{2-\alpha}$ if $\alpha \in [1,2)$.
\end{itemize}
By virtue of \eqref{eq:hp} we have 
\begin{equation*}
\begin{split}
n\,p\,\int_{B_1} \P(S_{n-1}>a_{np}^{1-\vep}\,,z>M_{n-1})\,\dint \P(Z\le z)&\le n\,p\,\P(M_{n-1}\le a_n/h_n)\\
&= \ex^{\log (np)-n\,\P(Z>a_n/h_n)+o(1)}\to 0\,.
\end{split}
\end{equation*}
By definition of $\vep$, we have $(a_n+n)/a_{np}^{1-\vep}\to 0$ for $\alpha\in (0,2)$. Therefore an application of
Theorem~\ref{thm:nagaev} yields
\begin{equation*}
\begin{split}
n\,p\,\int_{B_3} \P(S_{n-1}>a_{np}^{1-\vep}\,,z>M_{n-1})\, \dint \P(Z\le z)&\le 
n\,p\,\P(S_{n-1}>a_{np}^{1-\vep})\,\P(Z>a_{np}^{\gamma})\\
&\sim \big(n\,p\,\P(Z>a_{np}^{1-\vep})\big)\,\big(n\,\P(Z>a_{np}^{\gamma})\big)\,.
\end{split}
\end{equation*}
The \rhs\ converges to zero due to the property $\gamma>1/(1+\beta)+\vep$. 
\par
Now assume $\alpha\in (0,1)$. Then we have by Markov's inequality and Karamata's theorem,
\beao
\lefteqn{n\,p\,\int_{B_2} \P(S_{n-1}>a_{np}^{1-\vep}\,,z>M_{n-1})\, \dint \P(Z\le z)}\nonumber\\&\le&
\dfrac{n^2\,p}{a_{np}^{1-\vep}}\,\int_{B_2} \E [Z\1_{\{Z\le z\}}]  \dint \P(Z\le z)\\
&\le &\dfrac{n^2\,p}{a_{np}^{1-\vep}}\,\E [Z\1_{\{Z\le a_{np}^{\gamma}\}}] \,\P(Z>a_n/h_n)\\
&\sim & c\,\dfrac{n\,p}{a_{np}^{1-\vep}}\, \big[a_{np}^{\gamma}\,\P(Z>a_{np}^{\gamma})\big] \,\log(np)\,.
\eeao
An application of the Potter bounds and using the fact that $\gamma <1-\vep/(1-\alpha)$ shows that the \rhs\ converges to zero for the chosen $\vep$.
\par
Now assume $\alpha\in [1,2)$ and $\beta >\alpha-1$. Due to the latter condition we have $n/a_{np}^{1-\vep}\to 0$. 
We obtain by \v Cebyshev's inequality and Karamata's theorem,
\beao
\lefteqn{n\,p\,\int_{B_2} \P(S_{n-1}>a_{np}^{1-\vep}\,,z>M_{n-1})\, \dint \P(Z\le z)}\nonumber\\&\!\!\le\!\!\!\!&
n\,p\int_{B_2} \P\Big(\sum_{t=1}^n Z_t\,\1_{\{Z_t\le a_{np}^{\gamma}\}} \! -n\E \big[Z\1_{\{Z\le a_{np}^{\gamma}\}}\big]>a_{np}^{1-\vep}-n\,
\E \big[Z\1_{\{Z\le a_{np}^{\gamma}\}}\big]) \dint \P(Z\le z)\\
&\!\!\le\!\! \!\!&n\,p \int_{B_2} 
\P\Big(\sum_{t=1}^n Z_t\,\1_{\{Z_t\le a_{np}^{\gamma}\}}\!-n\,\E \big[Z\,\1_{\{Z\le a_{np}^{\gamma}\}}\big]
>c\,a_{np}^{1-\vep}\Big)\, \dint \P(Z\le z)\\
&\!\!\le\!\! \!\!&n\,p \,\dfrac{\E [ Z^2\,\1_{\{Z\le a_{np}^{\gamma}\}}]}{a_{np}^{2(1-\vep)}}\,\big[n\,\P(Z>a_n/h_n)\big]\\
%n\,p\,\int_{B_2} \Big[\overline \Phi\Big(c\,a_{np}^{1-\vep}/\sqrt{p\,\E[Z^2\1_{\{Z\le a_p^{1+\delta}\}}]}\Big)\\
%&&+c\,\dfrac{1}{\sqrt{p}} \dfrac{\E[Z^3\1_{\{Z\le a_p^{1+\delta}\}}]}{(\E[Z^2\1_{\{Z\le a_p^{1+\delta}\}}])^{3/2}}\Big]
%\dfrac{1}{1+ c\,\Big(a_{np}^{1-\vep}/\sqrt{p \E[Z^2\1_{\{Z\le a_p^{1+\delta}\}}]}\Big)^3}
%\,d\P(Z\le z)\\
%&=&J_1+J_2\,,
&\!\!\sim\!\! \!\!&c\,n\,p \,\dfrac{a_{np}^{2\,\gamma} \P(Z>a_{np}^{\gamma})}{a_{np}^{2(1-\vep)}}\,\log(np)\,.
\eeao
The \rhs\ converges to zero since $\gamma <1-2\vep/(2-\alpha)$.
%We mentioned before that $p/a_{np}^{1-\vep}\to 0 $ for small $\vep$.
%If $\E [Z]<\infty$ the \rhs\ is bounded from below by  $c\,a_{np}^{1-\vep}$. Now we consider the remaining cases when $\E [Z]=\infty$.
%If $\alpha=1$, $\E \big[Z\,\1_{\{Z\le a_p^{1+\delta}\}}\big]$ is a \slvary\ \fct\ of $p$, hence the \rhs\ is again bounded from below by  
%$c\,a_{np}^{1-\vep}$. A calculation with Karamata's theorem shows that the same statement holds for $\alpha\in (0,1)$ and suitable choices
This finishes the proof.
\end{proof}

\subsection{Perturbation theory for eigenvectors}
We state Proposition~A.1 in Benaych-Georges and P\'{e}ch\'{e} \cite{benaych:peche}.
\begin{proposition}\label{prop:perturbation}
Let $\bfH$ be a Hermitean matrix and $\bfv$ a unit vector such that for some  $\la\in \R$, $\vep>0$,
\begin{equation*}
\bfH\,\bfv= \la\,\bfv + \vep\, \bfw\,,
\end{equation*}
where $\bfw$ is a unit vector such that $\bfw \perp \bfv$.
\begin{enumerate}
\item
Then $\bfH$ has an eigenvalue $\lambda_{\vep}$ \sth\ $|\la-\la_\vep|\le \vep$.
\item
If $\bfH$ has only one eigenvalue $\la_\vep$ (counted with multiplicity) \sth\ $|\la-\la_\vep|\le \vep$ and all other eigenvalues 
are at distance at least $d>\vep$ from $\la$. Then for a unit eigenvector $\bfv_\vep$ associated with $\lambda_{\vep}$ we have
\begin{equation*}
\ltwonorm{\bfv_{\vep} - \bfP_\bfv(\bfv_{\vep})} \le \frac{2\, \vep}{d-\vep}\,,
\end{equation*}
where $\bfP_{\bfv}$ denotes the orthogonal projection onto \rm{Span}$(\bfv)$.
\end{enumerate}
\end{proposition}

%\section*{Acknowledgments}

%We thank Richard A. Davis and Olivier Wintenberger for reading the manuscript and fruitful %discussions. Special thanks go to Xiaolei Xie for providing the graphs.

### 2017-07-18 phd/Chapterintroduction.tex
Added My PhD thesis.

\chapter{Introduction}\label{ch:intr}

In this thesis we study the largest and smallest eigenvalues of high-dimensional sample covariance and correlation matrices of heavy-tailed time series.

\section{Random Matrix Theory}\label{sec:rmt}

The field of Random Matrix Theory (RMT) is concerned with the spectral properties of high-dimensional random matrices. Its development was motivated by applications.
In quantum mechanics, for example, the energy levels of particles in a large system can be characterized by the eigenvalues of a random infinite-dimensional Hermitean operator $\bfW$ on a Hilbert space. It is common to work with discretizations in a finite-dimensional space. In this case $\bfW$ turns into a high-dimensional Hermitean random matrix. Such matrices are called Wigner matrices named after Eugene Paul Wigner. He proved in the $1950$s that 
if $\bfW$ has independent standard normal entries on and above the diagonal, then
the expected empirical distribution of the eigenvalues of $\bfW$ tends to the so-called semi-circle law defined in \eqref{eq:semi-circle}, as the dimension of $\bfW$ goes to infinity; see for example \cite{wigner:1955,wigner:1957}. Since then a great variety of asymptotic results has been proved for various classes of random matrices under different assumptions on the distribution of the entries and their dependence structure. The discovery of many results was triggered by the enormous improvement of computation power which led to numerous conjectures. In the second half of the $20$th century, the research on asymptotic spectral properties of large-dimensional random matrices attracted considerable interest among physicists, computer scientists and mathematicians.  
A breakthrough in the theory of spectral distributions of sample covariance matrices was achieved by Mar\v cenko and Pastur in $1967$. 
%In $2010$, the most famous conjecture about the circular law was solved in \cite{tao:vu:krishnapur:2010}.

For many years the main focus of research in RMT has been on limiting spectral distributions. More recently, the focus turned to linear spectral statistics, eigenvectors, limiting distributions of extreme eigenvalues and their spacings. This thesis addresses these four topics in the special setting of random matrices with  heavy-tailed entries.

RMT is a versatile and useful tool in many fields of modern sciences that are faced with high-dimensional data sets. It  employs techniques from probability theory, multivariate statistics, number theory and combinatorics. Moreover, RMT finds applications in quantum physics, signal processing, wireless communications and finance; see Bai et al.~\cite{bai:fang:liang:2014} for more detailed examples. 
%``Big data'' is a popular topic nowadays. To meet the demand for data analysis and dimension reduction techniques RMT brings together

\subsection{Limiting spectral distributions}

For any random $p\times p$ matrix $\bfA$ with real eigenvalues $\lambda_1(\bfA),\ldots,\lambda_p(\bfA)$ the {\em empirical spectral distribution} is defined by 
\begin{equation*}
F_{\bfA}(x)= \frac{1}{p}\; \sum_{i=1}^p \1_{\{ \lambda_i(\bfA)\le x \}}, \qquad x\in  \R\,.
\end{equation*}

A major problem in random matrix theory is to find the weak limit of $(F_{\bfA_n})$, the so-called {\em limiting spectral distribution}, for suitable sequences of Hermitean $p\times p$ matrices $(\bfA_n)$. By weak convergence of a sequence of probability distributions $(F_{\bfA_n})$ 
to a \pro y \ds\ $F$, we mean $\lim_{\nto} F_{\bfA_n}(x)=F(x) \,\as$ for all continuity points of $F$.
Although the eigenvalues of $\bfA_n$ are continuous functions of the entries of $\bfA_n$ there are no closed-form expressions if the dimension is larger than $4$. Therefore methods to identify and characterize the limiting spectral distribution are needed. We will briefly discuss the two most common ones: the method of moments and Stieltjes transforms.
\par

By Lemma B.3 in \cite{bai:silverstein:2010}, a distribution function $F$ is uniquely characterized by its sequence of moments 
\begin{equation*}
\beta_k=\int_0^{\infty} x^k \dint F (x)\,, \quad k=1,2,\ldots
\end{equation*}
if Carleman's condition $\sum_{k=1}^{\infty} \beta_{2k}^{-1/(2k)}=\infty$ is satisfied. In this case, weak convergence of $(F_{\bfA_n})$ to $F$ is equivalent to the convergence of moments, that is
\begin{equation}\label{eq:mmom}
\beta_k(\bfA_n)= \int_0^{\infty} x^k \dint F_{\bfA_n} (x)=\frac{1}{p} \tr(\bfA_n^k) \to \beta_k\,,\quad \nto, \,k=1,2,\ldots
\end{equation}
In many cases the calculation of $\tr(\bfA_n^k)$ is very demanding. Often its mean and variance are estimated by combinatorial techniques.
On the positive side, if $F$ has finite support, Carleman's condition holds automatically.
\par

Another useful tool is the {\em Stieltjes transform}
of the empirical spectral distribution $F_{\bfA}$:
\begin{equation*}
s_{F_{\bfA}}(z)= \int_{\R} \frac{1}{x-z} \dint F_{\bfA}(x) = \frac{1}{p} \tr\big((\bfA -z \bfI)^{-1}\big)\,, \quad z\in \C^+\,,
\end{equation*}
where $\C^+$ denotes the complex numbers with positive imaginary part. 
Weak convergence of $(F_{\bfA_n})$ to a distribution function $F$ is equivalent to $s_{F_{\bfA_n}}(z) \to s_F(z)$ a.s. for all $z\in \C^+$.
Notice that the Stieltjes transform $s_F$ determines a distribution function $F$ at all continuity points $a,b$ of $F$ via
\begin{equation*}
F(b)-F(a)= \lim_{v\to 0^+} \frac{1}{\pi} \int_a^b \Im s_F(x+\mathrm{i} v) \dint x\,.
\end{equation*}

\section{Sample covariance matrices}
For a sample of $n$ column vectors $\x_1,\ldots,\x_n$ of a $p$-dimensional time series the (non-normalized) {\em sample covariance matrix} is usually defined as
\begin{equation*}
\bfS=\sum_{i=1}^n (\x_i-\overline{\x})(\x_i-\overline{\x})'= \sum_{i=1}^n \x_i \x_i' - n\overline{\x}\,\overline{\x}'= \X\X' - n\overline{\x}\,\overline{\x}'\,,
\end{equation*}  
where $\overline{\x}=n^{-1}\sum_i \x_i$ is the sample mean and 
\begin{equation*}
\X=(\x_1, \ldots, \x_n) =(X_{it})_{i=1,\ldots,p;\,t=1,\ldots,n}
\end{equation*}
is the {\em data matrix}. The sample covariance matrix is of crucial importance in multivariate statistics, for instance in principal component analysis, canonical correlation analysis, multivariate regression, factor analysis, hypothesis testing and discriminant analysis. The case of multivariate normal observations has played a particular role in the development of statistical theory. Already in $1928$, Wishart \cite{wishart:1928} studied sample covariance matrices with normal entries. Through the $20$th century non-asymptotic procedures for Gaussian observations such as Fisher's test, Student's test, and the analysis of variance were developed. In practice, however, observations are often not normally distributed and asymptotic methods based on limit theorems for certain model parameters are employed instead of exact results which are difficult to obtain.

Most of the classical limit theorems are derived under the assumption that the dimension $p$ is fixed and the sample size $n$ goes to infinity. 
If the assumptions of the law of large numbers are satisfied, then $n^{-1} \bfS$ converges \as~to the covariance matrix $\Sigma$ of $\x_1$.

If $p$ is moderately large, it is known that $n^{-1} \bfS$ ceases to be a good estimate for $\Sigma$. This means that classical methods based on fixed dimension and large sample limits may lead to wrong conclusions when applied to high-dimensional data. One would need appropriate adjustments. RMT provides limit theory in the case of large $p$. New statistical methods can be built on these results.

In RMT one assumes that $p=p_n$ grows with $n$. The most common condition in the literature is
\begin{equation}\label{Cp}
\lim_{\nto} \frac{p_n}{n}\to \gamma\in (0,\infty)\,. %\tag{$G_{\gamma}$}
\end{equation}
The asymptotic spectral behaviors of two large matrices are the same if their difference is of finite rank. Therefore we will refer to $\X\X'$ as the {\em sample covariance matrix} from now on. Indeed, by the rank inequality (see Bai and Silverstein \cite[Theorem A.44]{bai:silverstein:2010}) we have for the supremum norm
\begin{equation*}
\| F_{n^{-1}\bfS}-F_{n^{-1}\X\X'} \| \le p^{-1} \mathrm{rank}((\overline{\x},\ldots,\overline{\x}))\,,
\end{equation*}
which means that the respective limiting spectral distributions coincide under \eqref{Cp}. For the same reason we can assume without loss of generality that the entries $(X_{it})$ are centered whenever the expectations exist. 
\par

The limiting spectral distribution of normalized sample covariance matrices was found by Mar\v cenko and Pastur.

\begin{theorem}[Debashis and Aue \cite{debashis:aue:2014}]\label{thm:mpch1}
Suppose that $\X$ has iid entries with common mean and variance $1$. If \eqref{Cp} holds, then, with probability one, $(F_{n^{-1}\X \X'})$ converges weakly to a non-random distribution, the so-called  Mar\v cenko--Pastur law $F_\gamma$. If $\gamma \in (0,1]$,  $F_\gamma$  has density,
\begin{eqnarray}\label{eq:MPch1}
f_\gamma(x) =
\left\{\begin{array}{ll}
\frac{1}{2\pi x\gamma} \sqrt{(b-x)(x-a)} \,, & \mbox{if } a\le x \le b, \\
 0 \,, & \mbox{otherwise,}
\end{array}\right.
\end{eqnarray}\noindent
where $a=(1-\sqrt{\gamma})^2$ and $b=(1+\sqrt{\gamma})^2$. If $\gamma>1$, the \MP law is a mixture of a point mass at $0$ and the density function $f_{1/\gamma}$ with weights $1-1/\gamma$ and $1/\gamma$, respectively.
\end{theorem}

The \MP law describes the global behavior of the eigenvalues of $\X \X'$. Theorem \ref{thm:mpch1} quantifies the spread of the eigenvalues around their mean $1$. Note that the range of the deviation increases when $\gamma$ increases from $0$ to $\infty$.

If $p/n\to 0$, the limiting spectral distribution in Theorem \ref{thm:mpch1} is the Dirac measure at $1$. After an appropriate transformation of the sample covariance matrix one can obtain the semi-circle law defined in \eqref{eq:semi-circle} as a non-degenerate limiting spectral distribution in this case.

The crucial assumptions in Theorem \ref{thm:mpch1} are the finiteness of the variance and that $p$ and $n$ tend to infinity at the same rate. By Theorem 2.8 in Bai \cite{Bai:1999}, the conclusion still holds if the entries are independent, have common mean and satisfy the Lindeberg-type condition
\begin{equation*}
\lim_{\nto} \frac{1}{\delta^2 n p}\sum_{i,t} \E[X_{it}^2 \1_{\{|X_{it}|>\delta \sqrt{n}  \}}] =0\,,\quad \delta>0\,.
\end{equation*}
\par

Many important test statistics in multivariate analysis are functions of the eigenvalues
$\la_{(1)}\ge \cdots \ge \la_{(p)}$ of the sample covariance matrix and can be expressed by means of the empirical spectral distribution of $n^{-1}\X\X'$.

\begin{example}{\em 
Let $(X_{it})$ be iid standard normal and consider
\begin{equation*}
V_n:=\log(\det (n^{-1}\X\X'))=
\log \prod_{i=1}^p \frac{\lambda_{(i)}}{n} =  p \int_{0}^{\infty} \log x \;\dint F_{n^{-1}\X\X'}(x)\,.
\end{equation*}
On the one hand, if $p$ is fixed we know from Example 1.1 in Yao et a.~\cite{yao:zheng:bai:2015} that 
\begin{equation}\label{eq:srgdrgd}
\sqrt{\frac{n}{p}}\, V_n \cid Y\sim N(0,2)\,,\quad \nto\,.
\end{equation}
On the other hand, if $p/n \to \gamma\in (0,1)$, one can use Theorem \ref{thm:mpch1} to obtain asymptotic values for $V_n$. On gets \as
\begin{equation*}
\frac{V_n}{p}\to \int_{0}^{\infty} \log x \;\dint F_{\gamma}(x)= \frac{\gamma-1}{\gamma} \log(1-\gamma)-1<0\,,
\end{equation*}
which implies $\sqrt{n/p}\, V_n \to -\infty$ \as~In view of \eqref{eq:srgdrgd} the asymptotic behavior of the test statistic $V_n$ crucially depends on the dimension $p$.}
\end{example}

A characterization of the limiting spectral distribution of sample covariance matrices with  general population covariance has been derived for many settings. As an example we state Theorem 1.1 in Bai and Zhou \cite{bai:zhou:2008}.
%a result that we will use in Chapter \ref{ch:corr} to find the limiting spectral distribution of sample correlation matrices.
\begin{theorem}\label{them:baizhou}
Assume \eqref{Cp} and the following conditions.
\begin{itemize}
\item For all $k$, $\E[X_{jk}X_{lk}]=T_{lj}$, and for any non-random $p\times p$ matrix $\bfB$ with bounded norm, $\E[(\x_k'\bfB \x_k -\tr(\bfB \bfT))^2]=o(n^{-2})$, where $\bfT=\bfT_n=(T_{jl})$ and $\x_k$ are the columns of $\X$.
\item The norm of $\bfT_n$ is uniformly bounded and $F_{\bfT_n}$ tends to a non-random probability distribution $H$.
\end{itemize}
Then, with probability $1$, $F_{n^{-1}\X\X'}$ tends to a probability distribution, whose Stieltjes transform $m(z)$ satisfies
\begin{equation}\label{eq:mpequation}
m(z)=\int_{\R} \frac{1}{t(1-\gamma-\gamma z m(z))-z} \dint H(t)\,.
\end{equation}
If $\underline{m}(z)=-(1-\gamma)/z+\gamma m(z)$, then \eqref{eq:mpequation} becomes
\begin{equation}\label{eq:silversteineq}
z=-\frac{1}{\underline{m}(z)}+\gamma \int_{\R} \frac{1}{1+\underline{m}(z) \,t} \dint H(t)\,.
\end{equation}
\end{theorem}
For historical reasons, \eqref{eq:mpequation} is often called {\em \MP equation}. In practice, the inversion of such integral equations can be very difficult. Therefore, a characterization of the limiting spectral distribution via \eqref{eq:mpequation} is of limited use. For numerical procedures, the {\em Silverstein equation} \eqref{eq:silversteineq} is sometimes preferred. 
Roughly speaking, the only known explicit examples of non-degenerate limiting spectral distributions are the \MP law, the circular law, the semi-circle law and the multivariate $F$-matrix. In contrast, the number of existence results is huge. 
We refer to the discussion in Yao et al.~\cite{yao:zheng:bai:2015} for further details.
\medskip

In Theorem \ref{thm:mpch1} we have seen that the sequence $(F_{n^{-1}\X \X'})$ converges to the Mar\v cenko--Pastur law if the iid entries possess a finite second moment. Now we will discuss the situation when the entries are still iid, but have an infinite variance. Here we assume the entries to be regularly varying with index $\alpha \in (0,2)$; see \eqref{eq:regvarch1} on page \pageref{eq:regvarch1} for the definition of regular variation.
Assuming \eqref{Cp} with $\gamma\in (0,1]$ in this infinite variance case, Belinschi et al.~\cite[Theorem~1.10]{belinschi:dembo:guionnet:2009} showed that the sequence $(F_{a_{n+p}^{-2}\X \X'})$ converges with probability one to a non-random probability measure with density $\rho_{\alpha}^\gamma$ satisfying
\begin{equation}\label{eq:esd}
\rho_{\alpha}^\gamma(x) x^{1+\alpha/2} \to \frac{\alpha \gamma}{2(1+\gamma)}, \quad x \to \infty\,;
\end{equation}
see also Ben Arous and Guionnet \cite[Theorem~1.6]{arous:guionnet:2008}.
Here the normalization $(a_k)$ is defined \sth\ 
\begin{equation}\label{eq:defanp}
\P(|X|>a_k)\sim k^{-1}\,,\quad k\to\infty\,.
\end{equation}
An application of the Potter bounds (see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}) shows that $a_{n+p}^2/n \to \infty$. To the best of our knowledge, explicit expressions or computational methods for the limiting spectral distribution in the infinite variance case are not available at this moment.

\section{Limits of extreme eigenvalues under finite fourth moment}

After the limiting spectral distribution for sample covariance matrices had been found, the focus shifted to the asymptotic behavior of the largest and smallest eigenvalues $\la_{(1)}$ and $\la_{(p)}$, respectively, of $\X\X'$. 
In this section, we assume that the entries of the data matrix $\X$ are iid with generic element $X$. Furthermore, suppose $p\le n$; otherwise $\la_{(p)}=0$ since $\X\X'$ has at most $\min(n,p)$ non-zero eigenvalues.
We will discuss the setting $\E[X^4]<\infty$, while the case $\E[X^4]=\infty$ is treated in Section \ref{sec:contr}.

Under condition \eqref{Cp} with $\gamma\le 1$, one can infer from Theorem \ref{thm:mpch1} that
\begin{equation}\label{eq:conlcu}
\liminf_{n\to \infty}\frac{\la_{(1)}}{n} \ge (1+\sqrt{\gamma})^2\quad \mbox{and}\quad
\limsup_{n\to \infty}\frac{\la_{(p)}}{n} \le (1-\sqrt{\gamma})^2\quad \as 
\end{equation}
We follow Bai and Silverstein \cite{bai:silverstein:2010} and derive necessary conditions for the \as~convergence of $n^{-1}\la_{(1)}$. Since the largest diagonal entry of a matrix is bounded by its largest eigenvalue, we have
\begin{equation}\label{eq:simplebo}
\frac{\la_{(1)}}{n}\ge \max_{i=1,\ldots,p} \frac{1}{n} \sum_{t=1}^n X_{it}^2\,.
\end{equation}
If $\E[X^4]=\infty$, then by Lemma B.25 in \cite{bai:silverstein:2010}
\begin{equation}\label{eq:sghdfgdfs}
\limsup_{n\to \infty}\max_{i=1,\ldots,p} \frac{1}{n} \sum_{t=1}^n X_{it}^2 =\infty \quad \as
\end{equation}
%Next note that Theorem \ref{thm:mpch1} holds if $\E[X]=c\neq 0$. 
If $\E[X^4]<\infty$ and $\E[X]=c\neq 0$, then
\begin{equation*}
\frac{1}{\sqrt{n}}\twonorm{\X}\ge \frac{1}{\sqrt{n}}\twonorm{\E[\X]}  -\frac{1}{\sqrt{n}}\twonorm{\X-\E[\X]}
\ge \frac{|c| p}{\sqrt{n}}  - \frac{1}{\sqrt{n}}\twonorm{\X-\E[\X]} \to\infty \quad \as,
\end{equation*}
where for any matrix $\bfA$, $\twonorm{\bfA}$ denotes its spectral norm, i.e., its largest singular value. This and \eqref{eq:sghdfgdfs} show that $\E[X]=0$ and $\E[X^4]<\infty$ are necessary conditions for the \as~convergence of $n^{-1}\la_{(1)}$.

{\em In what follows, we assume $\E[X]=0$ and $\E[X^2]=1$, whenever the respective moments exist.}

\subsection{Sample covariance matrices}
Extending work by Geman \cite{geman},
Bai et al.~\cite{bai:yin:krishnaiah:1988} 
showed under \eqref{Cp} that
\beam\label{eq:gemanch1}
\dfrac {\la_{(1)}}{n} \stas \big(1+\sqrt{\gamma}\big)^2\,,\qquad \nto\,,
\eeam
which is the optimal result in view of \eqref{eq:sghdfgdfs}. Later Bai and Yin \cite{bai:yin:1993} proved the following result under the additional assumption $\gamma\in (0,1)$:
\beam\label{eq:jo45}
\limsup_{n\to \infty} \twonorm{n^{-1} \X\X'-(1+\gamma) \bfI} \le 2 \sqrt{\gamma}\,\quad \as
\eeam
Because of 
\begin{equation*}
\twonorm{n^{-1} \X\X'-(1+\gamma) \bfI} = \max \{ \la_{(1)}/n- (1+\gamma), -\la_{(p)}/n+(1+\gamma)\}\,,
\end{equation*}
equations \eqref{eq:jo45} and \eqref{eq:conlcu} imply 
\begin{equation}\label{eq:sdfgsdfdf}
\lim_{n\to \infty}\frac{\la_{(1)}}{n} = (1+\sqrt{\gamma})^2\quad \mbox{and}\quad
\lim_{n\to \infty}\frac{\la_{(p)}}{n} = (1-\sqrt{\gamma})^2\quad \as \,
\end{equation}
The approach based on \eqref{eq:jo45} provides a lower bound on the smallest eigenvalue. Unfortunately, one cannot gain any information about the minimal conditions for the existence of a limit of $\la_{(p)}/n$ since the method treats $\la_{(1)}$ and $\la_{(p)}$ simultaneously. Therefore it can (at best) only be applied under $\E[X^4]<\infty$. Consequently, one loses sharpness of the conditions for the limit of $\la_{(p)}/n$. It was finally discovered by Tikhomirov \cite{tikhomirov:2015} that the \as~limit of $ \la_{(p)}/n$ is given by \eqref{eq:sdfgsdfdf} if $\E[X^2]=1$, whereas higher moments can be infinite.
\par

Under suitable moment assumptions, $\la_{(1)}$ and $\la_{(p)}$ possess {\em Tracy--Widom} fluctuations around their almost sure limits. For instance, 
Johnstone \cite{johnstone:2001} complemented \eqref{eq:sdfgsdfdf} by the corresponding \clt\ in the special case of iid standard normal 
entries:
\begin{equation*}%\label{eq:john}
 n^{2/3}\,\dfrac{(\sqrt{\gamma})^{1/3}}{\big(1+\sqrt{\gamma}\big)^{4/3}}\Big(\dfrac {\la_{(1)}}{n} -
\big(1+\sqrt{\tfrac pn }\big)^2\Big)
\std \xi\,,
\end{equation*}
where the limiting \rv\ has a {\em Tracy--Widom \ds} of order~$1$.
Its distribution function $G_1$ is given by
\begin{equation*}
G_1(s) = \exp\Big\{
  -\frac{1}{2} \int_{s}^\infty [
    q(x) + (x - s) q^2(x)
 ] \dint x
\Big\}\,,
\end{equation*}
where $q(x)$ is the unique solution to the Painlev\'e II differential
equation
\begin{equation*}
  q''(x) = xq(x) + 2 q^3(x)\,,
\end{equation*}
where $ q(x)\sim {\rm Ai}(x)$ as $x \to \infty$ and Ai$(\cdot)$ is the Airy kernel; see Tracy and Widom~\cite{tracy:widom:2012} for details.

\subsection{Sample correlation matrices}

In comparison with the eigenvalues of $\X\X'$, much less is known about the ordered eigenvalues
\begin{equation*}
\mu_{(1)} \ge \cdots \ge\mu_{(p)}\,
\end{equation*} 
of the {\em sample correlation matrix} $\bfR=\Y\Y'$ with entries
\begin{equation}\label{eq:corrRch1}
R_{ij}=\sum_{t=1}^n \frac{X_{it}X_{jt}}{\sqrt{D_i} \sqrt{D_j}} = \sum_{t=1}^n Y_{it}Y_{jt}\,, \quad i,j=1,\ldots,p\,.
\end{equation}
In this thesis we will often make use of the notation $\Y=(Y_{it})=(X_{it}/\sqrt{D_i})$ and
\begin{equation}\label{eq:Dch1}
D_i=D_i^{(n)}=\sum_{t=1}^n X_{it}^2\,, \qquad
i=1,\ldots,p;\;  n\ge 1\,.
\end{equation}

With $\bfF= \diag (1/D_1,\ldots,1/D_p)$, we have $\bfR={\bfF}^{1/2} {\X}{\X}'{\bfF}^{1/2}$ which has the same eigenvalues as ${\X}{\X}'{\bfF}$. Weyl's inequality (see \cite{bhatia:1997}) yields
\begin{equation}\label{eq:lamuch1}
\begin{split}
\max_{i=1,\ldots,p} |\mu_{(i)}- n^{-1} \la_{(i)}| &\le \twonorm{{\X}{\X}'{\bfF}-n^{-1}{\X}{\X}' }\\
&\le n^{-1} \twonorm{{\X}{\X}'}  \twonorm{n \bfF- \bfI}\\
&= n^{-1} \la_{(1)} \max_{i=1,\ldots,p} \Big| \frac{n}{D_i} -1 \Big|\,,
\end{split}
\end{equation}
which converges \as~to $0$ if $\E[X^4]<\infty$; see Chapter~\ref{ch:corr} for details.
Under \eqref{Cp} with $\gamma\in(0,1]$,  Jiang \cite{jiang:2004}, and Xiao and Zhou \cite{xiao:zhou:2010} used this approach to derive 
\begin{equation*}
\mu_{(1)} \to (1+\sqrt{\gamma})^2\, \quad \text{ and } \quad  \mu_{(p)} \to (1-\sqrt{\gamma})^2\, \quad \as
\end{equation*}

\section{Eigenvectors}

Eigenvectors of large random matrices and graphs play an essential role in statistical analysis, physics and computer science. Many properties of a matrix or a graph can be deduced from its eigenvectors. Popular algorithms for data analysis such as spectral clustering, principal component analysis, PageRank and community detection are based on the eigenvector-eigenvalue decomposition of a matrix.

If the data matrix $\X$ has iid standard normal entries, then $\X\X'$ is a {\em Wishart matrix}, whose eigenvectors are well studied; see for example Bai and Silverstein \cite[Ch.~10]{bai:silverstein:2010}. Due to the invariance of $\X\X'$ under orthogonal transformations, its matrix of properly normalized eigenvectors is Haar distributed, i.e., the distribution is uniform on the space of orthogonal $p\times p$ matrices. This result has been extended to different classes of matrices $\X$ by direct comparison with Wishart matrices. Such statements are called universality results and they often require that the new entry distribution is in some sense similar to the standard normal distribution. This is often achieved by moment conditions. Silverstin \cite{silverstein:1984} showed that the matrix of eigenvectors is asymptotically Haar distributed as $p/n\to \gamma\in(0,\infty)$ if the first four moments of the iid entries coincide with those of the standard normal distribution. On the one hand, this means that eigenvectors of $\X\X'$ are completely unstructured for light-tailed entry distributions.  On the other hand, the extreme eigenvalues converge to constants \as 
\par

Although eigenvectors play a minor role in this thesis, we summarize some results from the literature on the light-tailed case.  Consequently, our approximations of eigenvectors in the heavy-tailed case are put into context. The majority of studies on eigenvectors of large random matrices is conducted on Wigner matrices $\bfW=(W_{ij})$. They are symmetric, real-valued, $n\times n$ matrices with entries $W_{ij}, 1\le i \le j \le n$, that are iid, mean zero and unit variance random variables.

Roughly speaking, the spectral properties of the square of a Wigner matrix and a sample covariance matrix with $p=n$ are quite similar if they share the same sufficiently light-tailed entry distribution. For the purpose of exposition, we look at $\bfW^2=\bfW \bfW'$ instead of $\X\X'$. Indeed, apart from the additional symmetry restriction for $\bfW$ they have the same structure.

By Theorem 2.1 in \cite{bai:fang:liang:2014}, the limiting spectral distribution of $(\bfW /\sqrt{n})$ is the {\em semi-circle law} $G$ with density
\begin{equation}\label{eq:semi-circle}
g(x)=\frac{1}{2\pi} \sqrt{4-x^2}\, \1_{\{ |x|\le 2 \}}\,.
\end{equation}
The semi-circle law and the \MP law $F_{\gamma}$ with $\gamma=1$ are linked in the following way: if $Y\sim G$ then $Y^2\sim F_1$.

Since $\bfW^2$ and $\bfW$ have the same eigenvectors it will be sufficient to study the latter. In the remainder of this section we list some properties of $\bfv_1, \ldots, \bfv_n$, where $\bfv_i$ is an eigenvector associated to the $i$th largest eigenvalue of $\bfW$. In addition, we assume $\bfv_i$ are unit vectors, i.e. $\ltwonorm{\bfv_i}=1$, and that the first non-zero component of $\bfv_i$ is positive.

Often the entries are assumed to be {\em sub-exponential}. We call a random variable $W$ {\em sub-exponential} with exponent $\alpha>0$ if there exists a constant $\beta>0$ such that 
\begin{equation*}
\P(|W|>x)\le \beta \exp(-x^\alpha /\beta)\,,\quad x>0\,.
\end{equation*}

Theorem 1.4 in O'Rourke et al.~\cite{orourke:vu:wang:2016} focuses on Wigner matrices with sub-exponential entries with exponent $\alpha$. There exists a constant $C_{\alpha}>0$ such that the probability that the spectrum of $\bfW$ is simple and that every coordinate of every $\bfv_i$ is non-zero is at least $1-C_{\alpha} n^{-\alpha}$. By our sign convention, the eigenvectors are unique with high probability. In particular, if $W_{11}$ is standard normal, $\bfv_i$ is uniformly distributed on 
\begin{equation*}
\mathcal{S}_+^{n-1}:=\{ \x=(x_1,\ldots,x_n)': \ltwonorm{\x}=1 \text{ and } x_1>0 \}\,.
\end{equation*}
An eigenvector of a Wigner matrix with light-tailed entry distribution (for instance sub-exponential) behaves like a random vector uniformly distributed on $\mathcal{S}_+^{n-1}$. More precise quantitative statements are difficult to obtain. For details we refer to \cite{orourke:vu:wang:2016}.

If $\bfv=(v_1,\ldots,v_n)'$ is a random vector uniformly distributed on $\mathcal{S}_+^{n-1}$, probabilistic bounds on its coordinates are available. By Theorem 2.1 in \cite{orourke:vu:wang:2016}, we have 
\begin{equation}\label{eq:coordinates}
v_{\text{max}} := \max_{i=1,\ldots,n} |v_i| \le C\sqrt{\frac{\log n}{n}}\quad \text{ and } \quad v_{\text{min}} := \min_{i=1,\ldots,n} |v_i|\ge \frac{c}{n^{3/2}}
\end{equation}
with probability $1-o(1)$ for any $C>1$ and $c\in [0,1)$.

In RMT it is common to study the so-called {\em bulk} and {\em edge} spectra separately.
For $\vep\in (0,1)$ one distinguishes between the {\em bulk} eigenvectors $\bfv_i, i\in \{1\le t\le n: \vep n \le t \le (1-\vep) n\} :=B_\vep$ and the {\em edge} eigenvectors $\bfv_i, i \in \{1,\ldots,n\} \backslash B_\vep$. The associated eigenvalues are usually referred to as bulk and edge spectrum, respectively. Roughly speaking, the limiting spectral distribution of a sequence of random matrices depends on the bulk spectrum, while the edge spectrum influences the behavior of functionals of the eigenvalues.
\par

The behavior of the largest coordinates of $\bfv_1,\ldots, \bfv_n$ was studied in the case of certain sub-exponential entries with exponent $\alpha=2$ whose first four moments match those of the standard normal distribution. 
By Theorems 4.1 and 4.3 in \cite{orourke:vu:wang:2016}, one has for any bulk eigenvector $\bfv_i$ and appropriate constants $c_1,c_2>0$,
\begin{equation*}
c_1 \sqrt{\frac{\log n}{n}} \le \bfv_{i,\text{max}}\le c_2 \sqrt{\frac{\log n}{n}}
\end{equation*}
with probability $1-o(1)$. This result is astonishingly precise. Note that up to logarithmic corrections $\bfv_{i,\text{max}}$ is of the smallest possible magnitude $n^{-1/2}$. This property is called {\em complete delocalization}; see \cite{rudelson:vershynin:2015}.
For edge eigenvectors, however, it is proved that for a constant $c_3>0$,
\begin{equation*}
\bfv_{i,\text{max}}\le c_3 \frac{\log n}{\sqrt{n}}\,
\end{equation*}
with probability $1-o(1)$.
The optimal bound in the edge case remains an open problem. Additionally, by Corollary 5.4 in \cite{orourke:vu:wang:2016} the $\ell_p$-norms, $1\le p\le2$, of the $(\bfv_i)$ are of the same order of magnitude:
\begin{equation}\label{eq:lpefgs}
c_0 n^{1/p-1/2}\le \min_{i=1,\ldots,n} \| \bfv_i\|_{\ell_p} \le \max_{i=1,\ldots,n} \| \bfv_i\|_{\ell_p}  \le C_0 n^{1/p-1/2}
\end{equation}
with probability $1-o(1)$ for positive constants $c_0,C_0$.
\par

If the entry distribution of $\bfW$ has heavy tails, the behavior of its eigenvectors is completely different. In \cite{benaych:peche}, Benaych-Georges and P\'{e}ch\'{e} assumed $W_{11}$ to be regularly varying with index $\alpha\in (0,4)$; see \eqref{eq:regvarch1} on page \pageref{eq:regvarch1}, and provided an approximation of $\bfv_k$ for any fixed $k$. From their asymptotic result one can deduce that $\bfv_{k,\text{max}}$ converges to $1/\sqrt{2}$ in probability. Moreover, asymptotically there are only two coordinates of $\bfv_k$ with non-zero mass. Both are of magnitude $1/\sqrt{2}$. This is the opposite of complete delocalization: {\em complete localization}. The number of non-zero coordinates is bounded.

Note that in the presence of heavy tails the eigenvectors of $\bfW$ and $\X\X'$ are very different; see \eqref{eq:eigenvec} in Section \ref{sec:contr}.

\section{Contribution of this thesis}\label{sec:contr}

In this section we summarize our contribution to the spectral theory of high-dimensional sample covariance and correlation matrices. We focus on the case where the entries of the $(p\times n)$-dimensional data matrix $\X$ have an infinite fourth moment.

\subsection{Sample covariance matrices: the iid case}

While the finite second moment is the central assumption to obtain the Mar\v cenko--Pastur law as the limiting spectral distribution, the finite fourth moment plays a crucial role when studying the eigenvalues
\beam\label{eq:orderch1}
\la_{(1)}\ge \cdots \ge \la_{(p)}
\eeam
of the sample covariance matrix $\X\X'$. 
Unless stated otherwise, the entries $(X_{it})$ are iid regularly varying random variables with index $\alpha\in (0,4)$ (see \eqref{eq:regvarch1}) and $X$ is a generic random variable with the same distribution. This implies $\E[X^4]=\infty$. Here and in what follows, we will refer to this setting as the heavy-tailed case, in contrast to the light-tailed case in which $\E[X^4]$ is finite. Moreover, we assume $\E[X]=0$ and $\E[X^2]=1$, whenever $\E[X^2]<\infty$.

We normalize the eigenvalues $(\la_{(i)})$ by $(a_{np}^2)$ where the \seq\
$(a_k)$ is chosen \sth 
\beao
\P(|X|>a_k)\sim k^{-1}\,,\quad k\to\infty.
\eeao
Standard theory for \regvary\ \fct s (e.g. Bingham et al. \cite{bingham:goldie:teugels:1987}, Feller \cite{feller}) yields that $a_n=n^{1/\alpha} \ell(n)$ where $\ell$ is a slowly varying function.
Assuming the usual growth condition \eqref{Cp} for $p$,
the Potter bounds (see \cite[p.~25]{bingham:goldie:teugels:1987}) yield for $\alpha\in (0,4)$ that
\begin{equation}\label{eq:toinfch1}
 \frac{a_{np}^2}{n} \sim \frac{n^{4/\alpha} \gamma^{2/\alpha}\,\ell^2(n^2 \gamma)}{n} \to \infty, \qquad \nto\,,
\end{equation}
i.e., the normalization $a_{np}^2$ is stronger than $n$.

By \eqref{eq:simplebo}, we have $\la_{(1)}\ge X_{(1),np}^2$, where $X_{(1),np}^2\ge \cdots \ge X_{(np),np}^2$
denote the order statistics of $(X_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$. Classical extreme value theory yields that $a_{np}^{-2} X_{(1),np}^2$ converges weakly to a  
{\em \Frechet distribution} with parameter~$\alpha/2$:
\beam\label{eq:frechetch1}
\Phi_{\alpha/2}(x) =\ex^{-x^{-\alpha/2}}\,,\qquad x>0\,.
\eeam
\par

The theory for the largest eigenvalues of sample covariance matrices with heavy tails is less developed than in the light-tailed case.
Pioneering work for $\la_{(1)}$ under the growth condition \eqref{Cp} and $\alpha\in (0,2)$
is due to Soshnikov~\cite{soshnikov:2004,soshnikov:2006}. For $k\ge 1$ fixed, he showed that
\begin{equation}\label{eq:sosh}
\frac{\la_{(m)}}{X_{(m),np}^2}\cip 1\,,\quad \nto\,, \,1\le m\le k\,,
\end{equation}
which reveals that the limiting distribution of $a_{np}^{-2} X_{(1),np}^2$ is the \Frechet distribution \eqref{eq:frechetch1}. Furthermore he proved the \pp\ \con\
\beam\label{eq:nnch1}
N_n=\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i} \std N_{\Gamma}=\sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,.
\eeam
Here $\vep_y$ is the Dirac measure at $y$,
\beam\label{eq:Gammach1}
\Gamma_i=E_1+\cdots + E_i\,,\qquad i\ge 1\,,
\eeam
and $(E_i)$ is a sequence of iid standard exponential random variables. In other words, $N_{\Gamma}$ is a Poisson point process on $(0,\infty)$ with mean \ms\
$\mu(x,\infty)= x^{-\alpha/2}$, \mbox{$x>0$}. Convergence in \ds\ of \pp es is understood in the sense of weak \con\
in the space of point \ms s equipped with the vague topology; see Resnick \cite{resnick:2007,resnick:1987}.

Later Auffinger et al.~\cite{auffinger:arous:peche:2009} established \eqref{eq:nnch1} also for $\alpha \in [2,4)$. In their proofs they used truncation techniques and a combinatorial approach.

\subsubsection*{General growth rates for $p_n$}

In many applications it is not realistic to assume
that the dimension $p$ of the data and the sample size $n$ grow at the same rate.
In the light-tailed case little is known when $p$ and $n$ grow at different rates, i.e., $\lim p/n \in\{ 0,\infty\}$. 
Notable exceptions are El Karoui \cite{elkaroui:2003} who proved that Johnstone's result in \cite{johnstone:2001} (assuming iid standard normal entries)
remains valid when $p/n\to 0$ or $n/p\to\infty$, and P{\'e}ch{\'e} \cite{peche:2009} who showed universality results for the largest eigenvalues of some sample covariance matrices with non-Gaussian entries.

The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009} already indicate that the value $\gamma$ in the usual growth
rate~\eqref{Cp} does not appear in the \ds al limits.
In the heavy-tailed case, more general growth of $p$ than prescribed by \eqref{Cp} has been used in Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
In Chapters \ref{ch:iid} and \ref{ch:extremes} we will consider power-law growth rates on the dimension $(p_n)$. To be precise, we assume an integer sequence
\begin{equation}\label{eq:pch1}
p=p_n=n^\beta l(n), \quad n\ge1\,,%\tag{$C_p(\beta)$}
\end{equation}
where $l$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume \mbox{$l(n) \to \infty$}.
Our condition~\eqref{eq:pch1} is more general than the growth conditions in the literature; see
\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.

Note that the matrices $\X\X'$ and $\X'\X$ have the same non-zero eigenvalues. Therefore it is sufficient to consider $\beta\in [0,1]$. For details we refer to Chapters \ref{ch:iid} and \ref{ch:extremes}.

\subsubsection*{Our contribution}

%It is interesting to note that there is a phase change of the behavior of the extreme eigenvalues when going from finite to infinite fourth moment, while a similar phase change occurs for the empirical spectral distribution when going from finite to infinite variance.

In the heavy-tailed case and under the growth condition \eqref{eq:pch1} with $\beta \in [0,1]$ we prove with considerable technical effort that  
\begin{equation}\label{eq:offdiagonalch1}
a_{np}^{-2} \twonorm{\X \X' - \diag(\X \X')}\stp 0\,,\qquad\nto\,,
\end{equation}
where $\diag(\X \X')$ denotes the diagonal of $\X\X'$; see Theorem \ref{prop:offdiagonal}.

The employed techniques originate from extreme value analysis and large deviation theory. The $(i,j)$ entry of $\X\X'$ is 
\begin{equation*}
(\X\X')_{ij}=\sum_{t=1}^n X_{it}X_{jt}\,,\quad i,j=1\ldots,p\,.
\end{equation*}
From Embrechts and Veraverbeke \cite{embrechts:veraverbeke:1982} we know that  $X^2$ and $X_{11}X_{12}$  are \regvary\ with indices $\alpha/2$ and $\alpha$, respectively. By large deviation theory (see \eqref{eq:dfsdfjl}), the diagonal and off-diagonal entries of
$\X\X'$ inherit the tails of $X_{it}^2$ and $X_{it}X_{jt}$, $i\ne j$, respectively, above some high threshold. Therefore the random variables in the diagonal of $\X\X'$ have the heaviest tail. They dominate the spectral behavior of $\X\X'$ and thus \eqref{eq:offdiagonalch1} is not unexpected.
\par

Equation \eqref{eq:offdiagonalch1} has some immediate con\seq s for the approximation of the eigenvalues
of $\X\X'$ by those of ${\rm diag}(\X\X')$. Indeed, let $\bfC$ be a symmetric $p\times p$ matrix with
eigenvalues 
\beam \label{eq:weylch1}
\la_{(1)}(\bfC)\ge \cdots \ge \la_{(p)}(\bfC)\,.
\eeam
Then for any symmetric $p\times p$ matrices $\bfA,\bfB$, by {\em Weyl's inequality} (see Bhatia \cite{bhatia:1997}),
\beao
\max_{i=1,\ldots,p}\big|\la_{(i)}(\bfA+\bfB)-\la_{(i)}(\bfA)\big|\le \|\bfB\|_2\,.
\eeao
If we now choose
$\bfA+\bfB=\X\X'$ and $\bfA= \diag (\X\X')$ we obtain
\begin{equation}\label{eq:appch1}
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\diag(\X\X'))\big|\stp 0\,,\quad\nto \,.
\end{equation}

Thus the problem of deriving limit theory for the order statistics of $\X\X'$ has been reduced to limit theory for the order
statistics of the iid row-sums
\beao
D_i= (\X\X')_{ii}=\sum_{t=1}^n X_{it}^2\,,\qquad i=1,\ldots,p\,,
\eeao
which are the eigenvalues of $\diag(\X\X')$. 
This theory is completely described by the \pp es constructed from the points $D_i/a_{np}^2$, $i=1,\ldots,p$. Necessary
and sufficient conditions for the weak \con\ of these \pp es are provided by  Lemma~\ref{lem:pprch3}. In combination with the Nagaev-type \ld\ results of  Theorem~\ref{thm:nagaevch3} they yield the following result under \eqref{eq:pch1}:
\begin{equation}\label{eq:ppdiagch1}
\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i-c_n)}\std N_{\Gamma}\,,\qquad \nto\,,
\end{equation}
where $N_{\Gamma}$ was defined in \eqref{eq:nnch1} and  $c_n=0$ if $\E[D]=\infty$ and $c_n=\E[D]=n \E[Z^2]$ otherwise.
Note that the centering $c_n$ in the finite variance case can be avoided if $n/a_{np}^2\to 0$. The latter condition is satisfied if
\begin{equation}\label{eq:b}
\beta>\alpha/2-1\,.
\end{equation}

Combining \eqref{eq:offdiagonalch1}, \eqref{eq:appch1} and \eqref{eq:ppdiagch1}, we conclude that \eqref{eq:nnch1} holds under the general growth rate \eqref{eq:pch1} with $\beta\in [0,1]$, where for $\alpha\in [2,4)$ one needs to require \eqref{eq:b}.

The limiting point process \eqref{eq:nnch1} yields a plethora of ancillary results. For example, one can easily derive the limiting distribution of $a_{np}^{-2} \lambda_{(k)}$ for fixed $k\ge 1$:
\begin{equation*}
\begin{split}
\lim_{\nto}\P(a_{np}^{-2} \lambda_{(k)}\le x)&= \lim_{\nto}\P(N_n(x,\infty)<k)
=  \P(N(x,\infty)<k)\\ &=\P(\Gamma_k^{-2/\alpha}\le x)= \sum_{s=0}^{k-1} \frac{\big(x^{-\alpha/2}\big)^s}{s!} \e^{-x^{-\alpha/2}}, \quad  x>0\,.
\end{split}
\end{equation*}
Another immediate con\seq\ of \eqref{eq:nnch1} is
\beam \label{eq:helpgamma1ch1}
a_{np}^{-2}\big(\la_{(1)},\ldots,\la_{(k)}\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)
\eeam
for any fixed $k\ge 1$ and $\alpha\in (0,2]$. In Chapter \ref{ch:extremes} we show for $\alpha \in (2,4)$
\beam \label{eq:helpgamma2ch1}
a_{np}^{-2}\big(\la_{(1)}-n \,\E[Z^2],\ldots,\la_{(k)}-n\, \E[Z^2]\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,.
\eeam
Equations \eqref{eq:helpgamma1ch1} and \eqref{eq:helpgamma2ch1} yield that for $\alpha \in (0,4)$ and any fixed $k\ge 1$,
\beam \label{eq:helpgamma3ch1}
a_{np}^{-2}\big(\la_{(1)}\!-\! \la_{(2)},\ldots, \la_{(k)}\!- \!\la_{(k+1)}\big) \!\std \!\big(\Gamma_1^{-2/\alpha} \!- \! \Gamma_{2}^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha} \! - \! \Gamma_{k+1}^{-2/\alpha} \big).
\eeam
Related results were also derived for linear spectral statistics such as the trace $a_{np}^{-2}(\la_1+\cdots+\la_p)$. 
We refer to Chapter \ref{ch:extremes} and Davis et al.~\cite{davis:mikosch:pfaffel:2015} for details on the proofs and more examples.
\medskip

In the case of  fixed $p$, Janssen et al.~\cite{janssen:mikosch:rezapour:xie:2016} related the limiting distribution of the eigenvalues $(\la_{(i)})$ to stable distributions. They also used \eqref{eq:offdiagonalch1} and \eqref{eq:appch1}. In this case it is clear that for example
\begin{equation*}
\frac{\la_{(2)}}{X_{(2),np}^2}\cip 1\,,\quad \nto\,, 
\end{equation*}
cannot hold. If $X_{(1),np}^2$ and $X_{(2),np}^2$ lie in the same row of $\X$, then they appear on the same spot on the diagonal of $\X\X'$. Then $X_{(2),np}^2$ cannot be used for the approxiamtion of $\la_{(2)}$ in view of \eqref{eq:appch1}.  Indeed, the probability that this happens is approximately $1/p$ which does not tend to $0$ if $p$ is fixed.  This is in contrast to \eqref{eq:sosh}.

Recall that if $p$ is fixed, one has to use the approximation of the eigenvalues provided by \eqref{eq:appch1}, while if $p/n \to \gamma$ one can use either \eqref{eq:appch1} or \eqref{eq:sosh}. There is a phase change of the behavior of $(\la_{(i)})$ when going from finite $p$ to $p$ proportional to $n$. In our condition \eqref{eq:pch1} the growth of $p$ is essentially described by the parameter $\beta$.

Under \eqref{eq:pch1} our Theorem \ref{thm:iidmain} asserts that 
\beam\label{eq:eigch1}
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-X_{(i),np}^2\big|\stp 0\,,
\eeam
provided \eqref{eq:b} holds. Due to its simplicity, \eqref{eq:eigch1} is an elegant result. It reveals, for example, that the largest eigenvalue of a high-dimensional heavy-tailed matrix behaves like the maximum of its iid entries.

In view of Lemma \ref{thm:keyresult}, condition \eqref{eq:b} describes the precise $\beta$-region, up to the slowly varying function in the tail of $X$, where $\max_i a_{np}^{-2}|D_{L_i}-X_{(i),np}^2|$ is sufficiently small. 
Here $L_i$ encodes the location of the $i$th largest diagonal element of $\X\X'$; see \eqref{eq:help6} for the formal definition of $L_i$.
Therefore the critical value of $\beta$ at which the aforementioned phase change occurs is  $\max(0,\alpha/2-1)$.
\medskip

The study of eigenvectors of heavy-tailed sample covariance matrices is a fresh topic which has not been explored in the literature listed here.
Our approximation of $\X\X'$ in \eqref{eq:offdiagonalch1} and the limiting distributions of the spacings \eqref{eq:helpgamma3ch1} can be applied to approximate the unit eigenvectors $(\bfv_j)$ of $\X\X'$, where $\bfv_{{j}}$ is associated to $\la_{(j)}$. As for the eigenvectors of Wigner matrices we use the convention that their first non-zero coordinate is positive.
From \eqref{eq:offdiagonalch1} we know that $\X\X'$ is approximated in spectral norm by $\diag(\X\X')$. The unit eigenvectors of $\diag(\X\X')$ are the canonical basis vectors $\bfe_j\in\R^p$, $j=1,\dots,p$.

By Theorem \ref{thm:eigenvec}, $(\bfe_j)$ approximate the eigenvectors $(\bfv_j)$.
For $\beta\in [0,1]$ and any fixed $k\ge 1$, we have
\begin{equation}\label{eq:eigenvec}
\ltwonorm{\bfv_k - \bfe_{L_k}} \cip 0\,, \quad \nto \,.
\end{equation}

\subsection{Sample covariance matrices: the non-iid case}

Davis et al.~\cite{davis:pfaffel:stelzer:2014} extended the results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009}
to the case where the rows of $\X$
are iid linear processes with iid \regvary\ noise. After a multiplication of the mean measure $\mu$ by a constant the Poisson \pp\ \con\ result of \eqref{eq:nnch1} remains valid. Pfaffel and Schlemm \cite{pfaffel:schlemm:2011} described the Stieltjes transform of the limiting spectral distribution in this model.
Different limit processes can only be expected if there is dependence in both directions:
in Chapter~\ref{ch:extremes} we use a model for $(X_{it})$ which allows for linear dependence across the rows and through time (see also \cite{davis:mikosch:pfaffel:2015}): 
\begin{equation}\label{eq:1ch1}
X_{it}=\sum_{l\in \Z}\sum_{k\in \Z} h_{kl} Z_{i-k,t-l}\,,\qquad i,t\in\Z\,,
\end{equation}
where $(Z_{it})_{i,t\in \Z}$ is a field of iid regularly varying \rv s with index $\alpha\in (0,4)$ and $(h_{kl})_{k,l\in\Z}$ is an array of real numbers. Moreover, we require the summability condition
\begin{equation}\label{eq:2ach1}
\sum_{l \in \Z} \sum_{k\in \Z} |h_{kl}|^{\delta} <\infty
\end{equation}
for some $\delta\in (0,\min({\alpha/2},1))$ which ensures
the a.s.~absolute convergence of the series in \eqref{eq:1ch1}. Under the condition \eqref{eq:2ach1}, the marginal and
\fidi s of the field $(X_{it})$ are \regvary\ with index $\alpha$; see
Embrechts et al. \cite{embrechts:kluppelberg:mikosch:1997}, Appendix A3.3.

From the field $(X_{it})$ we construct the $p\times n$ matrices
\begin{equation*}
\X_n(s)= (X_{i,t+s})_{i=1,\ldots,p;t=1,\ldots,n}\,,\qquad s=0,1,2,\ldots\,,
\end{equation*}
As before, we will write $\X=\X_n(0)$.
Now we can introduce the (non-normalized)
{\em sample autocovariance matrices}
\beao%\label{eq:samplech1}
\X_n(0)\X_n(s)'\,,\qquad s=0,1,2,\ldots\,.
\eeao
We will refer to $s$ as the {\em lag}. For $s=0$, we obtain the {\em sample covariance matrix.}
In \cite{heiny:mikosch:2016:noniid1} (see page \pageref{eq:regvarch1}) and Chapter \ref{ch:extremes}, we study the asymptotic behavior of the eigen- and singular values of the
sample covariance and autocovariance matrices under the growth condition \eqref{eq:pch1}.

Theorem~\ref{thm:mains} provides a general approximation result for the ordered singular values of $\X_n(0)\X_n(s)'$. Their behavior is determined by the sums $(\sum_{t=1}^n Z_{it}^2)$ and the singular values of the matrix $\M(s)$ given by
\begin{equation*}
(\M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z \,.
\end{equation*}
In Section \ref{sec:lkasfj} the limiting point process of the singular values of $\X_n(0)\X_n(s)'$ is derived. 
Finally, we mention that our paper \cite{heiny:mikosch:2016:noniid1} deals with the eigenvectors of $\X_n(0)\X_n(s)'$; see also \eqref{eq:eigenvec}. One obtains more interesting structures than in the latter result. In fact, by choosing $(h_{kl})$ accordingly one can obtain arbitrary eigenvectors. This constitutes a valuable property in principal component analysis.

\subsection{Sample correlation matrices}

In Chapter \ref{ch:corr} we study the spectrum of the sample correlation matrix $\bfR$ defined in \eqref{eq:corrRch1}. We assume that the underlying data matrix $\X$ has iid centered entries and the usual growth condition \eqref{Cp} holds. Recall the notation $Y_{it}=X_{it}/\sqrt{D_i}$ from p.~\pageref{eq:corrRch1}. We will sometimes write $(Y_1, \ldots, Y_n)= (Y_{11}, \ldots, Y_{1n})$ and $Y=Y_1$.
\par

Consider the following domain of attraction type-condition for the Gaussian law:
\beam\label{eq:condXch1}
\E \big[ Y_{1}Y_{2} \big] = o(n^{-2}) \quad  \text{ and } \quad  \E \big[ Y^4 \big] = o(n^{-1})\,,\qquad \nto\,.
\eeam
By Gin{\'e} et al.~\cite{gine:goetze:mason:1997}, condition \eqref{eq:condXch1} holds if the \ds\ of 
$X$ is in the domain of attraction of the normal law. We use Theorem \ref{them:baizhou} to  show that under \eqref{eq:condXch1} the sequence $(F_{\bfR})$ converges weakly to the \MP law $F_{\gamma}$ defined in \eqref{eq:MPch1}; see Theorem \ref{thm:mpcorrelation}. We prove that the condition \eqref{eq:condXch1} is necessary. When \eqref{eq:condXch1} is not valid, the limiting spectral distribution of $(F_{\bfR})$ (if it exists) must have mean $1$, by virtue of the method of moments (see \eqref{eq:mmom}). This follows from the fact that the diagonal elements of $\bfR$ are $1$.
This together with our approximation of $\E [\beta_k(\bfR)]$ provides some information about this distribution; compare also with \eqref{eq:esd} for the sample covariance case.
\par

Our analysis of the almost sure convergence of the extreme eigenvalues $\mu_{(1)}$ and $\mu_{(p)}$ of $\bfR$ is carried out for symmetric $X$. Then condition \eqref{eq:condXch1} turns into 
\beam\label{eq:condXb}
n \, \E \big[ Y^4 \big] \to 0\,,\qquad \nto\,.
\eeam
Theorem \ref{thm:mu1convergence} asserts
\begin{equation}\label{eq:drtgdrghdr1ch1}
\mu_{(1)} \to (1+\sqrt{\gamma})^2\, \quad \text{ and } \quad  \mu_{(p)} \to (1-\sqrt{\gamma})^2\, \quad \as
\end{equation}
under some condition \eqref{eq:assumptionq} on p.~\pageref{eq:assumptionq} which is slightly more restrictive than \eqref{eq:condXb}. Condition \eqref{eq:assumptionq} essentially means that the convergence rate of $n \, \E \big[ Y^4 \big]$ is at least logarithmic. A detailed discussion is given in Section \ref{sec:2ch4}.
\par

Our proof requires an adequate bound on $\E[\mu_{(1)}^{k_n}]$, where $k_n\to \infty$. To this end, we use the inequality
\begin{equation*}
\E[\mu_{(1)}^{k_n}]\le \E[\tr \bfR^{k_n}] %=p \E\big[\int x^{k_n} F_{\bfR}(dx)\big]
= \sum_{i_1,\ldots,i_{k_n}=1}^p \sum_{t_1,\ldots,t_{k_n}=1}^n \E[ Y_{i_1t_{k_n}} Y_{i_1t_1}  \cdots Y_{i_{k_n}t_{{k_n}-1}} Y_{i_{k_n}t_{k_n}}  ]
\end{equation*}
and determine those summands on the \rhs~which are largest when weighted by their multiplicities. Employing our {\em Path-Shortening Algorithm}, which is a novel technique that efficiently uses the inherent structure of sample correlation matrices, their contribution is calculated explicitly. The other summands can --with considerable technical effort-- be controlled by \eqref{eq:assumptionq}. 
Note that because of the identity $\E[\tr \bfR^{k_n}] =p\, \E\big[\int x^{k_n} F_{\bfR}(dx)\big]$ the behavior of the moments of the empirical spectral distribution is closely linked to the above upper bound.

Equation \eqref{eq:drtgdrghdr1ch1} indicates that the a.s. convergence of the extreme eigenvalues of $\bfR$
does not depend on the finiteness of the fourth or even second moments. 
This is in stark contrast to the a.s. behavior of $n^{-1}\lambda_{(1)}$, the largest eigenvalue of the sample covariance matrix 
$n^{-1}\bfX\bfX'$. Note that there is a phase transition of the a.s. \asy\ behavior of the  
extreme eigenvalues at the border between finite and infinite fourth moment of $X$, 
while such a transition occurs for the empirical spectral distribution at the border between finite and infinite variance.
\par

The eigenvalues of sample correlation matrices exhibit a  ``more robust'' behavior than their sample covariance analogs. 
This is perhaps not surprising in view of the {\em self-normalizing property} of sample correlations.
Self-normalization also has the advantage that one does not have to worry about the correct normalization.
This is a crucial problem in the study of sample covariance matrices in the case of an infinite fourth moment 
where one needs a normalization stronger than the classical one; see \eqref{eq:toinfch1}.
A simulation study in Section~\ref{sec:3} shows that the \asy\ results for $\mu_{(1)}, \mu_{(p)}, \la_{(1)}$ and $\la_{(p)}$ work nicely.
They can be used to design new statistical tests; see for example~\eqref{eq:forch1}.

\section{Outlook}

The main objective of our work was to find explicit limiting distributions of the eigenvalues of large random matrices and functionals thereof. Thus our theory can be applied in a straightforward way. Sections \ref{sec:rmt}-\ref{sec:contr} listed fields where our results can be used. Another example is the analysis of high-frequency data which receives significant interest; see for example Podolskij and Heinrich \cite{heinrich:podolskij:2014}, and  Xia and Zheng \cite{xia:zheng:2014,xia:zheng:2016}.

For practical purposes it is important to work with arbitrary population covariance matrices. Numerous generalizations and estimation techniques have been developed. For many models the limiting spectral distribution can only be characterized in terms of an integral equation (=\MP equation) for its Stieltjes transform. Explicit solutions are more involved; see the discussion after Theorem \ref{them:baizhou}. 
In Dobriban \cite{dobriban:2015}, an algorithm for calculating the spectral distribution based on certain approximate integral equations for its Stieltjes transform was presented. Contributions like this one breathe life into abstract theoretical results. Research in this direction will attract major interest from the industry.

From a more theoretical point of view, it is interesting to study models with heavy tails in which the asymptotic behavior of the sample covariance matrix $\X\X'$ is not dominated by the squares of the entries of $\X$. Moreover, different tail indices of the rows of $\X$ could make the model more appealing to practitioners. In Chapter \ref{ch:extremes}, we will see that a simple transformation, such as the rank transform, does not entirely overcome this issue.

As regards the sample correlation matrix our methods can be applied to more general models such as spiked covariance/correlation structures. Our technical results in Section~\ref{sec:5.3} are of independent interest. They provide a {\em Path-Shortening Algorithm} for the calculation of bounds for the very high moments of 
$\mu_{(1)}$. This technique is novel and will be of further use for proving results
in Random Matrix Theory.
We also conjecture that \eqref{eq:drtgdrghdr1ch1} may be proved under \eqref{eq:condXch1} only.

Finally, in our working papers \cite{heiny:mikosch:2016:noniid1} and \cite{heiny:mikosch:2016:stochvol1} (see page \pageref{eq:regvarch1}), we analyze  sample autocovariance matrices and provide limit theory for high-dimensional stochastic
volatility matrices.
Again, we utilize the large deviations approach propagated in Chapters \ref{ch:iid} and \ref{ch:extremes}, now for dependent heavy-tailed time series. The corresponding large deviations theory is available in Mikosch and Wintenberger \cite{mikosch:wintenberger:2016}.

### 2017-07-18 phd/MyThesis.tex
Added My PhD thesis.

\input{Preamble}

\begin{document}

%\includepdf{delete}

\frontmatter
\include{Titlepage}

\newpage
\leavevmode\thispagestyle{empty}

%\vspace*{1cm}
\hspace*{0.5cm}  {\bf Xiaolei Xie}\\
\hspace*{1.5cm} Department of Mathematical Sciences\\
\hspace*{1.5cm} University of Copenhagen\\
\hspace*{1.5cm} Universitetsparken 5\\
\hspace*{1.5cm} 2100 København Ø\\
\hspace*{1.5cm} Denmark\\
\hspace*{1.5cm} xie@math.ku.dk\\
\hspace*{1.5cm} xie.xiaolei@gmail.com

\vspace*{3cm}

\begin{tabular}{ll}
Supervisor: & Prof.~Thomas Mikosch\\
& University of Copenhagen\\
& \\
Co-supervisor: & Associate Prof.~Jeffrey Collamore\\
& University of Copenhagen\\
& \\
Assessment committee: &  Prof.~Catalin Starica\\
& University of Neuchatel, CH\\
& \\

& Prof.~Olivier Wintenberger\\
& University of Marie \& Pierre Curie, Paris\\
& \\

& Lecturer.~Jesper Lund Pedersen (chairman)\\
& University of Copenhagen\\
\end{tabular}

\vspace*{5cm}
\noindent \hspace*{0.5cm} Copyright:\\
\hspace*{0.5cm} Chapters 1-4: \copyright Johannes Heiny (according to Danish legislation)\\
\hspace*{0.5cm} Chapter 2: \copyright Elsevier\\

\noindent \hspace*{0.5cm} ISBN: $978$-$87$-$7078$-$921$-$9$

\begin{comment}
{	\pagestyle{empty}
	\setlength{\topmargin}{0pt}
	\setlength{\headheight}{0pt}
	\setlength{\headsep}{0pt}
	\setlength{\footskip}{0pt}

	\setlength{\parskip}{12pt}

	%\begin{center}
		%\vspace*{2.5cm}
 \large \textbf{\textsc{Johannes Heiny}}

		\vspace*{1cm}

	%\end{center}

	\vspace*{\fill}}
\end{comment}

\chapter*{}
\vspace*{-2cm}
\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}

This thesis is concerned with asymptotic properties of the eigenvalues of high-dimensional sample covariance and correlation matrices under an infinite fourth moment of the entries.

In the first part, we study the joint distributional convergence of the largest eigenvalues of the sample covariance matrix of a $p$-dimensional heavy-tailed time series when $p$ converges to infinity together with the sample size $n$. We generalize the growth rates of $p$ existing in the literature. Assuming a regular variation condition with tail index $\alpha<4$, we employ a large deviations approach to show that the extreme eigenvalues are essentially determined by the extreme order statistics from an array of iid random variables.  The asymptotic behavior of the extreme eigenvalues is then derived routinely from classical extreme value theory. The resulting approximations are strikingly simple considering the high dimension of the problem at hand.

We develop a theory for the point process of the normalized eigenvalues
of the sample covariance matrix in the case where rows and columns of the data are linearly dependent. Based on the weak convergence of this point process we derive the limit laws of various functionals of the eigenvalues.

In the second part, we show that the largest and smallest eigenvalues 
of a high-dimensional sample correlation matrix possess almost sure non-random limits if the truncated variance of the entry distribution is ``almost slowly varying'', a condition we describe via moment properties of self-normalized sums. We compare the behavior of the eigenvalues of the sample covariance and sample correlation matrices and argue that the latter seems more robust, in particular in the case of infinite fourth moment.

\section*{Resum{\'e}}
%\addcontentsline{toc}{section}{Resum{\'e}}

%Fokus for afhandlingen ligger i / Afhandlingen fokuserer på 
Denne afhandling beskæftiger sig med de asymptotiske egenskaber af egenværdierne for højdimensionale empiriske korrelations- og kovariansmatricer, under antagelse af at matrixindgangene har uendeligt fjerde moment.

I første del undersøger vi konvergens i fordeling af de 
%principiale? 
største egenværdier for den observerede kovarians af en $p$-dimensional tunghalet tidsrække, når
%/hvor 
$p$ sammen med stikprøvestørrelsen går mod uendelig.  
Vi generaliserer de eksisterende vækstrater 
af $p$ 
fra litteraturen.  Under antagelse af regulær variation med haleindeks $\alpha < 4$, bruger vi en 
%"store afvigelser?"
large deviations-tilgang for at vise at de ekstremale egenværdier er essentielt bestemt ud fra de største værdier i et array af iid. stokastiske variable.
Herefter udleder vi rutinemæssigt de ekstremale egenværdiers asymptotiske egenskaber ved brug af klassisk 
%ekstremal
ekstremværditeori.

Vi fremfører en teori
%? 
for punktprocesser af de normalisererde egenværdier af den empiriske kovariansmatrix i tilfældet, hvor dens rækker og søjler er afhængige. Med udgangspunkt i, at denne punktproces konverger svagt, udleder vi grænsemomenter af 
%diverse?  
forskellige funktionaler af egenværdierne.

I den anden del af afhandlingen viser vi at de største og mindste egenværdier fra en højdimensional korrelationsmatrix har næsten sikre grænser, hvis den trunkerede varians af indgangsfordelingen er ,,næsten langsom varierende``, en betingelse som kan indentificeres ud fra momentegenskaber af selv-normaliserende summer. Vi sammenligner, hvordan egenværdierne for 
%hhv.
henholdsvis den empiriske korrelations- og den empiriske kovariansmatrix opfører sig, og argumenterer for at sidstnævnte tilfælde er mere robust, hvilket særligt gælder i tilfældet med uendeligt fjerde moment.

\newpage
\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}

First and foremost, I would like to thank my supervisor Thomas Mikosch for many fascinating discussions, his guidance and experience. I could not have wished for a better person to work with and learn from.  Furthermore I am grateful for the support I received from my co-supervisor Olivier Wintenberger.

I want to thank Richard Davis and the Statistics Department of Columbia University in the City of New York, where I spent a research visit in $2016$, for their great hospitality. The group meetings on Wednesdays led to a vivid exchange of ideas.
I would also like to thank Claudia Kl\"uppelberg. Her advice is highly appreciated.

Next, I want to thank Gennady Samorodnitsky and Mark Podolskij for inspiring discussions. I am also grateful to Jeffrey Collamore, Jesper Lund Pedersen, Mogens Steffensen and Jostein Paulsen.

It is my special pleasure to thank our lunch group at the University of Copenhagen which was an integral part of daily university life. I also want to thank my fellow PhD students at the University of Copenhagen and Columbia University for many fun hours. Special thanks go to all my office mates and in particular Xie Xiaolei.

I acknowledge the support by Danish Research Council Grant DFF-4002-00435 ``Large random matrices with heavy tails and dependence''.

Last but not least, I want to thank my family for the unconditional love and the invaluable support I have received over the years.

\bigskip

{\em Johannes Heiny \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}

\chapter*{Summary}
\addcontentsline{toc}{section}{Summary}

This PhD thesis provides asymptotic theory for the eigenvalues of the sample covariance matrix of a high-dimensional time series with infinite fourth moment. Its main part consists of the first three of the following research papers. They were written from March $2014$ until December $2016$.

\begin{enumerate}
\bibitem[P1]{heiny:mikosch:2015:iid1}
{\sc Heiny, J., and Mikosch, T.}
\newblock Eigenvalues and eigenvectors of heavy-tailed sample covariance
  matrices with general growth rates: the iid case.
\newblock {\em Stochastic Process. Appl.\/} (2016), 29. \href{http://www.sciencedirect.com/science/article/pii/S0304414916301934}{\blue{[pdf]}}

\bibitem[P2]{davis:mikosch:heiny:xie:20151}
{\sc Davis, R.~A., Heiny, J., Mikosch, T., and Xie, X.}
\newblock Extreme value analysis for the sample autocovariance matrices of
  heavy-tailed multivariate time series.
\newblock {\em Extremes 19}, 3 (2016), 517--547. \href{http://link.springer.com/article/10.1007/s10687-016-0251-7}{\blue{[pdf]}}

\bibitem[P3]{heiny:mikosch:2016:corr11}
{\sc Heiny, J., and Mikosch, T.}
\newblock Almost sure convergence of the largest and smallest eigenvalues of high-dimensional sample correlation matrices under infinite fourth moment.
\newblock {\em Submitted for publication}.

\bibitem[P4]{heiny:mikosch:2016:noniid1}
{\sc Heiny, J., and Mikosch, T.}
\newblock Limit theory for the singular values of the sample autocovariance
  matrix function of multivariate time series.
\newblock {\em In preparation}.

\bibitem[P5]{heiny:mikosch:2016:stochvol1}
{\sc Heiny, J., and Mikosch, T.}
\newblock Asymptotic theory for high-dimensional stochastic volatility matrices.
\newblock {\em In preparation}.
\end{enumerate}

The aforementioned infinite fourth moment is ensured by a {\em regular variation condition}. We say that a random variable $X$ and its distribution are {\em regularly varying with index} $\alpha>0$ if
\begin{equation}\label{eq:regvarch1}
\P(X>x)\sim p_+\,\dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad \P(X<-x) \sim p_-\,\dfrac{L(x)}{x^\alpha}\,,\qquad \xto\,,\tag{$RV_{\alpha}$}
\end{equation}
where $ p_\pm$ are non-negative constants \sth\ $p_++p_-=1$ and $L$ is a \slvary\ \fct . In particular, if $\alpha<4$ we have $\E [X^4]= \infty$. The \regvar\ condition is needed for proving \asy\ theory for the eigenvalues of the sample covariance matrix. Moreover, we will often use the concept of {\em heavy tails}.  A distribution is called {\em heavy-tailed} if certain moments are infinite. By construction, any regularly varying distribution is heavy-tailed.  
\par

Now we explain the structure of this thesis. In Chapter \ref{ch:intr} we provide an introduction to Random Matrix Theory and present the classical results in the light-tailed case. We give examples of high-dimensional statistical inference problems and indicate how the asymptotic theory applies. The empirical distribution of the eigenvalues of the widely used sample covariance matrix is studied. Furthermore the \as~limits of its largest and smallest eigenvalues are identified under finite fourth moment. Then we explain the typical behavior of its eigenvectors. 
{\em Section \ref{sec:contr} constitutes the main part of Chapter~\ref{ch:intr}.} There we present the contribution of this thesis and the novelties of \cite{heiny:mikosch:2015:iid1, davis:mikosch:heiny:xie:20151, heiny:mikosch:2016:corr11}.

Chapters \ref{ch:iid}-\ref{ch:corr} consist of these $3$ papers, respectively. Each chapter is self-contained with its own introduction and references. Chapter~\ref{ch:intr} can be used as a joint introduction to Chapters \ref{ch:iid}-\ref{ch:corr}. 
\par
In Chapter~\ref{ch:iid} we study the joint \ds al \con\ of the largest eigenvalues of the sample covariance matrix
of a $p$-dimensional \ts\ with iid entries when $p$ converges to infinity together with the sample size $n$. We generalize the growth rates of $p$ in the literature. Assuming the regular variation condition with $\alpha<4$, we employ a large deviations approach to show
that only the diagonal of the sample covariance matrix is relevant for the \asy\ behavior of the largest eigenvalues and the corresponding eigenvectors. The resulting approximations are strikingly simple considering the high dimension of the problem at hand.

In Chapter~\ref{ch:extremes} we generalize the results from the iid case that were presented in the previous chapter. 
We develop a theory for the \pp\ of the normalized eigenvalues
of the sample covariance matrix in the case when rows and columns of the data are linearly dependent.
We provide limit results for the weak \con\ of these \pp es to Poisson or cluster Poisson processes. Based on
this \con\ we can also derive the limit laws of various \fct als of the ordered eigenvalues such as the
joint \con\ of a finite number of the largest order statistics, the joint limit law of the largest eigenvalue and the trace,
limit laws for successive ratios of ordered eigenvalues,
etc. We also develop some
limit theory for the singular values of the sample autocovariance matrices and their sums of squares. The theory is illustrated
for simulated data and for the components of the S\&P 500 stock index. Further generalizations of the results of this chapter are made in \cite{heiny:mikosch:2016:noniid1} and \cite{heiny:mikosch:2016:stochvol1}, but are not part of this thesis.

In Chapter~\ref{ch:corr}, we show that the largest and smallest eigenvalues 
of a sample correlation matrix stemming from $n$ independent observations of a $p$-dimensional 
time series with iid components converge almost surely to $(1+\sqrt{\gamma})^2$ and $(1-\sqrt{\gamma})^2$, 
respectively, as $\nto$, if $p/n\to \gamma \in (0,1]$ and the truncated variance of the entry 
distribution is ``almost slowly varying'', a condition we describe via moment 
properties of self-normalized sums. We compare the behavior of the eigenvalues of the 
sample covariance and sample correlation matrices and argue that the latter seems more robust, 
in particular in the case of infinite fourth moment. We briefly address some practical issues 
for the estimation of extreme eigenvalues in a simulation study.

Chapter~\ref{ch:corr} is the most technical one of this thesis. %In addition to the submitted version we added Section~\ref{sec:anexample} which contains an attempt to adapt the widely employed truncation techniques to self-normalized sums with the goal of verifying our crucial distributional assumption. 
In our proofs we use the method of moments combined with a Path-Shortening Algorithm, 
which efficiently uses the structure of sample correlation matrices, 
to calculate precise bounds for matrix norms. We believe that this new approach could be of further use in Random Matrix Theory. 
\bigskip

{\em Johannes Heiny \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}

\newpage
\tableofcontents

\mainmatter

\include{Chapterintroduction}
%------------------------------------------------------------------------------------
\include{Chapteriidpaper}
\include{Chapterextremes}
\include{Chaptercorrelation}

\bibliography{libraryjohannes}

%for print version only
\newpage
\leavevmode\thispagestyle{empty}
\phantom{bla}
\end{document}

### 2017-07-21 papers/FX/draft.tex
Added my own thesis.

--- 2017-06-05 papers/FX/draft.tex
@@
-% \documentclass{article}
-% \usepackage{amsmath}
-% \usepackage{amsthm}
-% \usepackage{enumerate}
-% \usepackage{bookmark}
-% \usepackage{graphicx}
-% \usepackage{multirow}
-% \usepackage{adjustbox}
-% \usepackage{wrapfig}
-% \usepackage{subcaption}
-% \usepackage{color}
-% \usepackage[dvipsnames]{xcolor}
-
-% \usepackage{amssymb,amsmath,amsthm,amsfonts}
-% \usepackage{mathrsfs}
-% \usepackage{dsfont}
-% \usepackage{enumerate}
-
-% \input{physics_common.tex}
-
 \documentclass[11pt,a4]{amsart}
-%\usepackage{amsmath}
-%\usepackage{amsthm}
-%\usepackage{enumerate}
-\usepackage[bookmarks=true]{hyperref}
-%% \usepackage{bookmark}
-\usepackage{graphicx}
-%\usepackage{multirow}
-%\usepackage{adjustbox}
-%\usepackage{wrapfig}
-%\usepackage{subcaption}
-\usepackage{color}
-\usepackage[dvipsnames]{xcolor}
-
-\usepackage{amssymb,amsmath,amsthm,amsfonts}
-%\usepackage{mathrsfs}
-%\usepackage{dsfont}
-%\usepackage{enumerate}
-
-%\input{../../physics_common.tex}
-%\input{physics_common.tex}
-
-\textwidth 6.50in
-\topmargin -0.50in
-\oddsidemargin 0in
-\evensidemargin 0in
-\textheight 9.00in
-%\pagestyle{plain}
-\definecolor{darkblue}{rgb}{.2, 0.2,.8}
-\definecolor{darkgreen}{rgb}{0,0.5,0.3}
-\definecolor{darkred}{rgb}{.8, .1,.1}
-\newcommand{\pd}{\partial}
-\newcommand{\red}{\color{darkred}}
-\newcommand{\blue}{\color{darkblue}}
-\newcommand{\green}{\color{darkgreen}}
-\newcommand{\black}{\color{black}}
-\newcommand{\levy}{L\'evy}
-\newcommand{\slln}{strong law of large numbers}
-\newcommand{\clt}{central limit theorem}
-\newcommand{\sde}{stochastic \de}
-\newcommand{\de}{differential equation}
-\newcommand{\mregvar}{multivariate \regvar}
-\newcommand{\garch}{{\rm GARCH}$(1,1)$}
-\newcommand{\spr}{stochastic process}
-\newcommand{\It}{It\^o}
-\newcommand{\sta}{St\u aric\u a}
-\newcommand{\ex}{{\rm e}\,}
-\newcommand{\sas}{s$\alpha$s}
-\newcommand{\civ}{\stackrel{v}{\rightarrow}}
-\newcommand{\asy}{asymptotic}
-\newcommand {\hatt} {\hat{\theta}}
-\newcommand {\hattg} {\hat{\theta}_{MLE}}
-\newcommand {\hattz} {\hat{\theta}_0}
-\newcommand {\hatb} {\hat{\beta}}
-\newcommand {\hatbg} {\hat{\beta}_G}
-\newcommand {\spp} {^\prime}
-\newcommand{\ts}{time series}
-\newcommand{\tsa}{\ts\ analysis}
-\newcommand{\garchpq}{{\rm GARCH}$(p,q)$}
-\def\theequation{\thesection.\arabic{equation}}
-\def\tag{\refstepcounter{equation}\leqno }
-\def\neqno{\refstepcounter{equation}\leqno(\thesection.\arabic{equation})}
-\newtheorem{lemma}{Lemma}[section]
-\newtheorem{figur}[lemma]{Figure}
-\newtheorem{theorem}[lemma]{Theorem}
-\newcommand{\Leb}{{\mathbb L}{\mathbb E}{\mathbb B}}
-\newcommand{\bbc}{{\mathbb C}}
-\newtheorem{proposition}[lemma]{Proposition}
-\newtheorem{definition}[lemma]{Definition}
-\newtheorem{corollary}[lemma]{Corollary}
-\newtheorem{example}[lemma]{Example}
-\newtheorem{exercise}[lemma]{Exercise}
-\newtheorem{remark}[lemma]{Remark}
-\newtheorem{fig}[lemma]{Figure}
-\newtheorem{tab}[lemma]{Table}
-\newtheorem{hypothesis}{Hypothesis}
-\newcommand{\play}{\displaystyle}
-\newcommand{\rmk}{\noindent R{\sc emark}}
-\newcommand{\cid}{\stackrel{d}{\rightarrow}}
-\newcommand{\cip}{\stackrel{P}{\rightarrow}}
-\newcommand{\reference}{\hangindent=5pc\hangafter=1}
-\newcommand{\MC}{Markov chain}
-\newcommand{\bfQ}{{\bf Q}}
-\newcommand{\bfM}{{\bf M}}
-\newcommand{\bfu}{{\bf u}}
-\newcommand{\bfv}{{\bf v}}
-\newcommand{\bfV}{{\bf V}}
-\newcommand{\bfU}{{\bf U}}
-\newcommand{\bfR}{{\bf R}}
-\newcommand{\bfC}{{\bf C}}
-\newcommand{\bfF}{{\bf F}}
-\newcommand{\RV}{{\rm RV}}
-\newcommand{\LN}{{\rm LN}}
-\newcommand{\WE}{{\rm WE}}
-\newcommand{\bth}{\begin{theorem}}
-\newcommand{\ethe}{\end{theorem}}
-\newcommand{\sv}{stochastic volatility}
-\newcommand{\bre}{\begin{remark}\em }
-\newcommand{\ere}{\end{remark}}
-\newcommand{\fre}{frequenc}
-\newcommand{\ble}{\begin{lemma}}
-\newcommand{\ele}{\end{lemma}}
-\newcommand{\sre}{stochastic recurrence equation}
-\newcommand{\pp}{point process}
-\newcommand{\bde}{\begin{definition}}
-\newcommand{\ede}{\end{definition}}
-\newcommand{\bco}{\begin{corollary}}
-\newcommand{\eco}{\end{corollary}}
-
-\newcommand{\bpr}{\begin{proposition}}
-\newcommand{\epr}{\end{proposition}}
-
-\newcommand{\bexer}{\begin{exercise}}
-\newcommand{\eexer}{\end{exercise}}
-
-\newcommand{\bexam}{\begin{example}}
-\newcommand{\eexam}{\end{example}}
-
-\newcommand{\bfi}{\begin{fig}}
-\newcommand{\efi}{\end{fig}}
-
-\newcommand{\btab}{\begin{tab}}
-\newcommand{\etab}{\end{tab}}
-%\renewcommand{\thetab}{\arabic{section}.9.\arabic{tab}}
-
-\newcommand{\edf}{empirical distribution function}
-\newcommand{\epro}{empirical process}
-\newcommand{\per}{periodogram}
-\newcommand{\pf}{{\bf Proof.}}
-\newcommand{\lhs}{left-hand side}
-\newcommand{\fidi}{finite-dimensional distribution}
-\newcommand{\rv}{random variable}
-\newcommand{\DA}{{\rm DA}}
-\newcommand{\DNA}{{\rm DNA}}
-\newcommand{\MDA}{{\rm MDA}}
-\newcommand{\sign}{{\rm sign}}
-\newcommand{\PRM}{{\rm PRM}}
-\newcommand{\card}{{\rm card}}
-\newcommand{\var}{{\rm var}}
-\newcommand{\med}{{\rm med}}
-\newcommand{\cov}{{\rm cov}}
-\newcommand{\corr}{{\rm corr}}
-\newcommand{\as}{{\rm a.s.}}
-\newcommand{\io}{{\rm i.o.}}
-\newcommand{\Pois}{\rm Poisson}
-
-\newcommand{\bfth}{\mbox{\boldmath$\theta$}}
-\newcommand{\bfTh}{\mbox{\boldmath$\Theta$}}
-\newcommand{\bfmu}{\mbox{\boldmath$\mu$}}
-
-\newcommand{\bfvep}{\mbox{\boldmath$\vep$}}
-\newcommand{\rhs}{right-hand side}
-\newcommand{\df}{distribution function}
-\newcommand{\dsum}{\displaystyle\sum}
-%\newcommand{\dfrac}{\displaystyle\frac}
-\newcommand{\dint}{\displaystyle\int}
-\newcommand{\dlim}{\displaystyle\lim\limits}
-\newcommand{\dmax}{\displaystyle\max\limits}
-\newcommand{\dsup}{\displaystyle\sup\limits}
-\newcommand{\dmin}{\displaystyle\min\limits}
-\newcommand{\dprod}{\displaystyle\prod}
-
-\newcommand{\beao}{\begin{eqnarray*}}
-\newcommand{\eeao}{\end{eqnarray*}\noindent}
-
-\newcommand{\beam}{\begin{eqnarray}}
-\newcommand{\eeam}{\end{eqnarray}\noindent}
-
-\newcommand{\beqq}{\begin{equation}}
-\newcommand{\eeqq}{\end{equation}\noindent}
-
-\newcommand{\bce}{\begin{center}}
-\newcommand{\ece}{\end{center}}
-
-\newcommand{\barr}{\begin{array}}
-\newcommand{\earr}{\end{array}}
-\newcommand{\cadlag}{c\`adl\`ag}
-\newcommand{\tep}{\tilde\epsilon}
-\newcommand{\stp}{\stackrel{\P}{\rightarrow}}
-\newcommand{\std}{\stackrel{d}{\rightarrow}}
-\newcommand{\stas}{\stackrel{\rm a.s.}{\rightarrow}}
-\newcommand{\stj}{\stackrel{J_1}{\rightarrow}}
-\newcommand{\stv}{\stackrel{v}{\rightarrow}}
-\newcommand{\stw}{\stackrel{w}{\rightarrow}}
-\newcommand{\stww}{\stackrel{\wh w}{\rightarrow}}
-\newcommand{\eqd}{\stackrel{d}{=}}
-
-\newcommand{\refbib}{\cite}
-\newcommand{\vague}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
-                    \it{v} $}}{\rightarrow}}
-\newcommand{\weak}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
-                    \it{w} $}}{\rightarrow}}
-\newcommand{\what}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
-                    \it{\hat{w}} $}}{\rightarrow}}
-
-\newcommand{\bdis}{\begin{displaymath}}
-\newcommand{\edis}{\end{displaymath}\noindent}
-
-\newcommand{\Thetabold}{{\boldsymbol{\Theta}}}
-\newcommand{\Prob}{\operatorname{P}}
-\newcommand{\Dspace}{\mathbb{D}}
-\newcommand{\Espace}{\mathbb{E}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\Xbold}{{\mathbf{X}}}
-\newcommand{\xbold}{{\mathbf{x}}}
-\newcommand{\Ybold}{{\mathbf{Y}}}
-\newcommand{\ybold}{{\mathbf{y}}}
-\newcommand{\zerobold}{{\mathbf{0}}}
-\newcommand{\Sphere}{\mathbb{S}}
-\newcommand{\Dbar}{\overline{\mathbb{D}}_{0}}
-\newcommand{\ind}{independent}
-\newcommand{\w}{\omega}
-\newcommand{\W}{\Omega}
-\newcommand{\bw}{\bigwedge}
-\newcommand{\bv}{\bigvee}
-\newcommand{\nto}{n\to\infty}
-\newcommand{\kto}{k\to\infty}
-\newcommand{\xto}{x\to\infty}
-\newcommand{\tto}{T\to\infty}
-\newcommand{\uto}{u\to\infty}
-\newcommand{\ov}{\overline}
-\newcommand{\wt}{\widetilde}
-\newcommand{\wh}{\widehat}
-\newcommand{\vep}{\varepsilon}
-\newcommand{\ep}{\epsilon}
-\newcommand{\vap}{\varphi}
-\newcommand{\la}{\lambda}
-\newcommand{\La}{\Lambda}
-
-\newcommand{\halmos}{\hfill \qed}
-\newcommand{\regvary}{regularly varying}
-\newcommand{\slvary}{slowly varying}
-\newcommand{\regvar}{regular variation}
-%\newcommand{\bbr}{{I\!\!R}}
-%\newcommand{\bbn}{{I\!\!N}}
-%\newcommand{\bbz}{{I\!\!\!Z}}
-
-\newcommand{\ka}{\kappa}
-\newcommand{\bba}{{\mathcal A}}
-\newcommand{\bbe}{{\ov \bbr^{d}_0}}
-\newcommand{\bfT}{{\bf T}}
-\newcommand{\bbr}{{\mathbb R}}
-\newcommand{\bbm}{{\mathcal M}}
-\newcommand{\bbb}{{\mathcal B}}
-\newcommand{\bbz}{{\mathbb Z}}
-\newcommand{\bbn}{{\mathbb N}}
-\newcommand{\bbd}{{\mathbb D}}
-\newcommand{\bbf}{{\mathcal F}}
-\newcommand{\bbs}{{\mathbb S}}
-\newcommand{\bbE}{{\mathcal E}}
-
-\newcommand{\BM}{Brownian motion}
-\newcommand{\BB}{Brownian bridge}
-\newcommand{\con}{convergence}
-\newcommand{\ev}{extreme value}
-\newcommand{\evt}{extreme value theory}
-\newcommand{\evd}{extreme value distribution}
-\newcommand{\clp}{central limit problem}
-\newcommand{\nc}{norming constant}
-\newcommand{\st}{such that}
-\newcommand{\fif}{if and only if}
-\newcommand{\wrt}{with respect to}
-\newcommand{\chf}{characteristic function}
-\newcommand{\fct}{function}
-\newcommand{\mdoa}{maximum domain of attraction}
-\newcommand{\ds}{distribution}
-\newcommand{\gev}{generalized extreme value distribution}
-\newcommand{\gpd}{generalized Pareto distribution}
-\newcommand{\pwm}{probability-weighted moments}
-\newcommand{\mle}{maximum likelihood estimator}
-\newcommand{\mel}{mean excess function}
-\newcommand{\an}{asymptotically normal}
-\newcommand{\ac}{absolutely continuous}
-\newcommand{\rep}{representation}
-\newcommand{\cmt}{continuous mapping theorem}
-\newcommand{\seq}{sequence}
-\newcommand{\pr} {{\bf Proof.}}
-\newcommand{\ins}{insurance}
-\newcommand{\mat}{mathematic}
-\newcommand{\pro}{probabilit}
-\newcommand{\pros}{probabilities}
-\newcommand{\ms}{measure}
-\newcommand{\mgf}{moment generating function}
-\newcommand{\ld}{large deviation}
-\newcommand{\bfx}{{\bf x}}
-\newcommand{\bfX}{{\bf X}}
-\newcommand{\bfW}{{\bf W}}
-\newcommand{\bfB}{{\bf B}}
-\newcommand{\bfY}{{\bf Y}}
-\newcommand{\bfy}{{\bf y}}
-\newcommand{\bfA}{{\bf A}}
-\newcommand{\bfG}{{\bf G}}
-\newcommand{\bfO}{{\bf 0}}
-\newcommand{\bfZ}{{\bf Z}}
-\newcommand{\bfz}{{\bf z}}
-\newcommand{\bfS}{{\bf S}}
-\newcommand{\bfa}{{\bf a}}
-\newcommand{\bfb}{{\bf b}}
-\newcommand{\bfe}{{\bf e}}
-\newcommand{\bft}{{\bf t}}
-\newcommand{\bfI}{{\bf I}}
-\newcommand{\bfs}{{\bf s}}
-\newcommand{\bfh}{{\bf h}}
-\newcommand{\bfunit}{{\bf h}}
-\newcommand{\ARCH}{{\rm ARCH}}
-\newcommand{\E }{{\mathbb E}}
-\renewcommand{\P }{{\mathbb P}}
-\newcommand{\Q }{{\mathbb Q}}
-\newcommand{\1}{{\mathbf 1}}
-
-\newcommand{\floor}[1]{
-  \lfloor #1\rfloor
-}
-\newcommand{\ceil}[1]{
-  \lceil #1\rceil
-}
-\renewcommand{\thefootnote}{\fnsymbol{footnote}}
-\allowdisplaybreaks
+\input{../../ThomasDefs}
 \begin{document}
 \title{Do return series have power-law tails with the same index?}
-t\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
+\author{Thomas Mikosch, Casper de Vries, Xie Xiaolei}
 \date{\today}
@@
 
-\begin{abstract}
-%We consider an investor with preferences that accord with Generalized
-%Disappointment Aversion. Such an investor cares about downside risk and
-%we assume he recognizes the heavy tail feature of asset return
-%distributions. We argue that when a market is dominated by rational
-%investors of this kind, the return distributions of equities that are
-%actively traded in this market must have very similar tail-indices. We
-%also show, in contrast, the scale parameters of the return
-%distributions may differ hugely from one another.
-%On the other hand, whether or not all equities in a multivariate model
-%have the same tail-index is a dividing issue for multivariate GARCH
-%models proposed in the literature. Therefore, it is important to analyze
-%data of real equity returns and see how close to each other the
-%tail-indices actually are.
-%In this work empirical results are also presented and they appear
-%to support the conclusion that the tail-indices are very similar,
-%with respect to the confidence bands of estimation.
-\end{abstract}
-
-\section{Introduction}\setcounter{equation}{0}
-It is one of the stylized facts of financial econometrics that 
-returns of speculative prices are {\em heavy-tailed.} There is no agreement
-in the literature about how heavy these tails really are. For example, Barndorff-Nielsen and Shephard
-\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails
-which are comparable with those of a gamma \ds . On the other
-hand, tails of returns have been studied in great detail 
-in the extreme value community. Among extreme value specialists
-there is general agreement that returns $X_t$ have tails of power-law-type, i.e.,
-\beam\label{eq:1}
-\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad \P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
-\eeam  
-where $c_{\pm}$, $\alpha_{\rm up}$ and $\alpha_{\rm low}$ are positive constants.\footnote{Here and in what
-  follows, $f(x)\sim g(x)$ for positive \fct s $f$ and $g$ means 
-that $f(x)/g(x)\to 1$ as $\xto$.} See for example, Embrechts et
-al. \cite{embrechts:klueppelberg:mikosch:1997}, Jansen and de Vries
-\cite{jansen1991frequency}, Mikosch \cite{mikosch:2003}, Resnick
-\cite{resnick:2007}.  
-In the extreme value literature it is common to replace the constants $c_\pm$ by suitable
-{\em slowly varying} \fct s; cf. Embrechts et
-al. \cite{embrechts:klueppelberg:mikosch:1997}, Chapter 3. In this
-paper, for the  
-sake of argument, we stick to the condition \eqref{eq:1}.
-\par
-There are some good theoretical reasons for the appearance of power-law tails in situations where certain moments 
-of data are believed to be infinite. Tails of type \eqref{eq:1}
-describe the maximum domain of attraction of the  
-Fr\'echet \ds\ $\Phi_{\alpha_{\rm up}}(x)=\exp(-x^{-\alpha_{\rm up}})$ for $x>0$, i.e., scaled maxima of an iid \seq\ $(X_t)$ with
-upper tail described in \eqref{eq:1} converge in \ds\ to $\Phi_{\alpha_{\rm up}}$.
-Equivalently, power-law tails are prescribed by the 
-generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e., for a suitable positive scaling \fct\ $a(u)$,
-\beao
-\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha_{\rm up})^{-\alpha_{\rm up}}\,,\qquad u\to\infty\,.
-\eeao
-The aforementioned results 
-are considered very natural for iid and weakly dependent strictly stationary \seq s of random variables $(X_t)$; 
-in the world of extremes they are the analogs
-of the \clt\ from the world of sums.
-\par
-In the literature on extremes for return data one finds the 
-statement that {\em estimated} values $\hat \alpha_{\rm up}$ and $\hat \alpha_{\rm low}$ 
-of the tail-indices $\alpha_{\rm up}$ and $\alpha_{\rm low}$,
-respectively, typically have the tendency that $\hat \alpha_{\rm
-  up}>\hat \alpha_{\rm low}$. 
-This observation is often explained by the fact that investors are
-more prone to negative than to positive news in the market. 
-Moreover, in the literature the {\em estimated} tail-indices $\hat \alpha$ (both in the left and right tails) 
-are typically found in the range $(2,4)$. For an illustration, see
-Figure~\ref{fig:1} where estimates $\hat \alpha_{\rm low}$ 
-in three sectors of the Standard \& Poors 500 index  are shown. The
-estimates are based on 1304 observations of daily return 
-data from 4 January 2010 to 31 December 2014.
-\par
-When looking at Figure~\ref{fig:1} one might ask the following questions:
-\begin{itemize}
-\item
-In view of the wide \asy\ confidence bands for the estimators of tail-indices, 
-are the tail-indices from different series really distinct?
-\item
-Are there some {\em theoretical} reasons supporting the fact that the tail-indices from different series are {\em not} 
-distinct?
-\end{itemize}
-In this paper, we try to find some answers to these questions. 
-\par
-The estimator of the tail-index $\alpha>0$  in the model
-\beao
-\P(X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
-\eeao
-favored in the literature  is the {\em Hill estimator}; the graphs in 
-Figure~\ref{fig:1} are based on this estimator. We introduce this estimator in Section~\ref{sec:1} and discuss 
-some of its virtues and vices. In addition to tail-index estimation we also discuss the related problem of
-estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). 
-%{\red check} It turns out 
-%that the simultaneous estimation of the tail-index and the scale parameter are strongly related.
-\par
-In Section~\ref{sec:2} we discuss the theoretical problem of appearance of power-law tails in models 
-for daily or, more generally, low-\fre y 
-return data. In particular, in Section~\ref{subsec:garch} we address the power-law tails of univariate and multivariate GARCH models as potential models
-for a set of return data from distinct assets. As a matter of fact, under mild conditions, the aforementioned models
-have power-law tails due to their relation with so-called {\em \sre s}. Moreover, some of the {\em standard} multivariate 
-GARCH models as the CCC ensure that the component-wise marginal \ds s have power-law tails with the same index.
-\par
-In Section~\ref{sec:3} we discuss an economic argument for the fact that return data of similar assets
-(such as return series in a given sector of the S\&P 500 index) may have tail-indices which are close to each other.
-We argue based on a  utility \fct\ approach. We explicitly recognize the behavioral
-concern for downside risk in an investor's evaluation of a portfolio
-using the framework of Generalized Disappointment Aversion (GDA)
-introduced by Routledge and Zin \cite{routledge2010generalized}. GDA
-is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991} who derived DA from first principles (axiomatic).
-
-In Section~\ref{sec:4} we summarize the discussion of the previous
-sections.
-
-\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
-In this section, we assume the model \eqref{eq:1} for the tails of the marginal
-\ds\ of a univariate return series $(X_t)$. For the sake of argument, we assume
-that this series constitutes a strictly stationary \seq . In what follows, we focus
-on the left tail of the \ds , i.e., on the losses. 
-However, it is common to present the tail-index estimators
-for positive data. Therefore we will multiply the losses $X_t$ by
-minus one, swapping the negative with the positive values.
-For simplicity, we also suppress subscripts in the notation:
-\beam\label{eq:1a}
-\P(-X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
-\eeam
-where we assume that the two parameters -- the tail-index  $\alpha$ and the scale parameter
-$c$ -- are positive. They play crucial roles for the understanding of the risk hidden in the data, hence 
-for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
-useful description of the risk, for example in terms of high quantiles such as
-Value-at-Risk. Alternatively, these parameters can be
-used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
-%With these motives in mind, we present a survey of the
-%tail-indices and scale parameters of 3 sectors, namely ``Energy'',
-%``Consumer Staples'' and ``Information Technology'', of the S\&P 500
-%index.
-
-\subsection{Hill estimates of lower tail-indices}\label{sec:Hill}
-Various estimators of the tail-index $\alpha$  in the model
-\eqref{eq:1a} have been proposed in the literature;
-see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, de
-Haan and Ferreira \cite{haan:ferreira:2006}, Resnick
-\cite{resnick:2007}. The most popular among them was introduced  by
-Hill \cite{hill1975simple}.
-Given a sample $-X_1,\ldots,-X_n$ whose marginal \ds\ satisfies  \eqref{eq:1a}, calculate
-the order statistics $X_{(1)}\le \cdots\le X_{(n)}$  and construct the
-{\em Hill estimator}:
-\beao%\label{eq:Hill_index}
-  \hat \alpha_k = \Big(
-    {1 \over k} \sum_{i=1}^k \log {X_{(n-i+1)} \over X_{(n-k)}}
-    \Big)^{-1}\,.
-\eeao
-Here $k$  is the number of upper order statistics in the sample used
-for the estimation. The estimator $\hat \alpha_k$ is 
-a maximum-likelihood estimator of $\alpha$ based on the $k$ upper
-order statistics in the pure Pareto model (recall that 
-we multiplied the data by minus one)
-\beam\label{eq:3}
-\P(-X_t>x)= \dfrac{K^\alpha}{x^\alpha}\,,\qquad x>K\,,
-\eeam
-under the hypothesis that we do not know the (high) threshold value
-$K$. The estimator has ``good'' theoretical properties such as 
-\asy\ consistency and \asy\ normality. These properties hold under
-strict stationarity assumptions on the data; Drees and Rootz\'en 
-\cite{drees:rootzen:2010} give perhaps most general conditions for
-dependent \seq s and de Haan and Ferreira \cite{haan:ferreira:2006} provide
-a complete \asy\ theory in the iid case.
-\par
-A major problem for Hill estimation is the choice of the number $k$ of
-upper order statistics. As  a matter of fact,
-if $k$ is too large the order statistics are too close to the center
-of the \ds\ of the $-X_t$, leading to a bias
-of the estimator. On the other hand, by construction, $\hat \alpha_k$ is an average of $k$ log-differences of the 
-data. Therefore, the variance of the estimator is the larger the
-smaller $k$. For these reasons, \asy\ theory requires 
-to choose $k=k_n$ \st\ $k_n\to\infty$ and $k_n/n\to 0$ as $\nto$. This
-fact does not make the estimation of $\alpha$ an  
-easy matter: one has to choose a ``small'' value $k$ which is not
-``too large''. For practical purposes,
-a so-called {\em Hill plot} is recommended where $\hat \alpha_k$ is
-plotted for a variety of $k$-values, corresponding to
-some high  quantile $X_{(n-k)}$ of the data. Then $k$ is chosen from a
-region in the plot where $\hat \alpha_k$ is relatively
-stable. For example, in Figure~\ref{fig:1} we have chosen $k=50$ from
-a sample of size $n=1304$, corresponding to the 96\%-quantile of the
-data. In general, the estimation of the tail-index is an art and
-requires some expertise; for some guidance
-see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997},
-Resnick \cite{resnick:2007} and Drees et al. 
-\cite{drees:resnick:2000}. 
-\par
-In Figure~\ref{fig:1},
-we exhibit 95\% \asy\ confidence bands derived from  the \clt
-\beam\label{eq:2}
-\sqrt k\, \big(\hat \alpha_k - \alpha\big) \std N(0, \alpha^2)\,,
-\eeam
-i.e., $\hat \alpha_k$ is \asy ally unbiased and has variance
-$\alpha^2/k$. Since $k/n\to 0$ this means that the 
-confidence bands are significantly larger than the classical
-$1/\sqrt{n}$-rates. This fact is one explanation for the fact that it is
-difficult to say something meaningful about the true value of
-$\alpha$. There exist various other reasons why one should
-not have 100\% trust in the confidence bands shown in
-Figure~\ref{fig:1}. Indeed, \eqref{eq:2}
-holds under rather subtle {\em second order conditions} on the tail
-$\P(X_t>x)$ which cannot be verified on data. However, 
-given a theoretical model such as the GARCH, these conditions can be
-verified based on the theoretical properties of the model. If they are not satisfied the 
-Hill estimator may exhibit significant bias; see Embrechts et
-al. \cite{embrechts:klueppelberg:mikosch:1997} and Resnick \cite{resnick:1987} for illustrations 
-of this fact leading to so-called ``Hill horror plots''. Moreover, the
-Hill estimator is rather sensitive to non-stationarity 
-of the data and to dependence. For example, results in Drees \cite{drees:2008}, and Drees and
-Rootz\'en \cite{drees:rootzen:2010} show that the \asy\ variance of the
-Hill estimator can be significantly larger than in the iid case. 
-Since return data are dependent, the \asy\ confidence bands should be
-even wider than exhibited in Figure~\ref{fig:1}. 
-Again, only under he assumption of a concrete model like GARCH these
-confidence bands can be evaluated and therefore the bands shown in Figure~\ref{fig:1}
-just show some benchmark which holds in the iid case and under additional conditions on the tail \asy s.
-\begin{figure}[htb!]
-  \begin{minipage}{1.0\linewidth}
-    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
-    {Energy_lower.pdf}
-  \end{minipage}
-  \begin{minipage}{1.0\linewidth}
-    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
-    {Consumer_Staples_lower.pdf}
-  \end{minipage}
-  \begin{minipage}{1.0\linewidth}
-    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
-    {Information_Technology_lower.pdf}
-  \end{minipage}
-  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail-indices $\alpha$ of
-    daily return series in sectors of the S\&P 500
-    index. The data span
-    from 1 January 2010 to 31 December 2014 and comprise $n=1304$ observations.
-The graphs from top to bottom correspond to the ``Energy'',
-    ``Consumer Staples'' and ``Information Technology'' sectors.
-    Each circle corresponds to a Hill estimate $\hat\alpha_{50}$; the gray
-    triangles above and below it mark the 97.5\% and 2.5\% quantiles
-    of its approximate normal distribution; see \eqref{eq:2} and the discussion following it for an 
-interpretation.
-    The lower and upper red lines mark the medians of the 2.5\% 
-    and 97.5\% quantiles, respectively, evaluated from all stocks in the sector.
-    The data are taken from {\it Yahoo Finance}; the labels on
-    the horizontal axes are Yahoo symbols of the stocks. 
-  }\label{fig:1}
-\end{figure}
-
-In Figure~\ref{fig:1} we see significant overlap of the confidence intervals of the Hill
-estimates of the losses in the ``Energy'' and ``Consumer Staples''
-sectors of the S\&P 500 index, as well as those of a 
-large portion of losses in the ``Information Technology'' sector.
-This fact indicates that the returns in each sector may 
-have comparable tail-indices.
-\par
-Hoga's \cite{hoga:2016} test about the change of extreme quantiles
-in a sample
-may provide some further insight about how similar these tail-indices are.
-Different tail-indices are likely to result in different
-extreme quantiles. Nevertheless, changes in the extreme quantiles may also
-result from changing scale parameters in the tail. Therefore  we first investigate the scale
-parameters of daily stock returns in the same sectors of S\&P 500 before we apply the test.
-
-\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
-We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
-Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
-derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
-Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
-It is given by
-\beao
-\hat K_k = \left({k \over n}\right)^{1/\hat\alpha_k} X_{(n-k)}\,.
-\eeao
-Using the asymptotic normality property of upper order statistics
-(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, Theorem 2.2.1), one can show
-\beao
-  \sqrt{k}\, (\hat K_k - K) &\std& N\big(
-    0, (K /\alpha)^2\big)\quad\mbox{and}\quad
-  \sqrt {k} \,(\hat K_k^\alpha - K^\alpha) \overset{d}{\to} N\big(
-    0, K^{2\alpha}  \big)\,,
-\end{eqnarray*}
-where the tail-index $\alpha$ is regarded as known. From the above
-asymptotic normality property, confidence bands of $\hat K_k$
-and $\hat K_k^{\hat\alpha}$ can be constructed.   
-Estimates $\hat K_k$ in the ``Energy'', ``Consumer Staples'' and
-``Information Technology'' sectors of the S\&P 500 index are computed
-using this method. The results are shown in
-Figure~\ref{fig:sectors_parameters}.
-\begin{figure}[htb!]
-  \centering
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[width=\textwidth]
-    {Energy_K.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[width=\textwidth]
-    {CS_K.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[width=\textwidth]
-    {IT_K.pdf}
-  \end{minipage}
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[width=\textwidth]
-    {Energy_scale.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[width=\textwidth]
-    {CS_scale.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[width=\textwidth]
-    {IT_scale.pdf}
-  \end{minipage}
-  \caption{\small Estimates of $\hat K_k$ (top) and $\hat K_k^{\hat
-      \alpha}$ (bottom) on $\log_{10}$-scale of stocks in sectors
-    of the S\&P 500 index. The estimates are ordered according to the 
-    corresponding estimated $\alpha$-values.
-    The points are the estimated values, the bars 
-    the \asy\ 95\%-confidence intervals; the confidence bands of the
-    corresponding Hill estimates $\hat \alpha_k$ of these sectors are
-    shown in Figure~\ref{fig:1}. 
-  }
-  \label{fig:sectors_parameters}
-\end{figure}
-One can see that $K$ generally takes a rather small value. For the
-more volatile sectors of ``Energy'' and ``Information Technology'',
-the average value of $K$ is around 0.01, while for the more stable
-sector of ``Consumer Staples'', the average value is around 0.005. Due
-to the smallness of $K$, mild variations of $\alpha$ would lead to
-huge variations of $K^\alpha$, as shown in the 2nd row of
-Figure~\ref{fig:sectors_parameters}.
-\par
-Secondly, it appears that there is positive dependence between the
-values of $\alpha$ and $K$.
-As argued in Section~\ref{sec:pareto_tail}, this is consistent with
-the assumption that the return series have Pareto tails on both sides
-with the tail parameters on each side independent of those on the
-other.
-
-\par
-Thirdly, Figure~\ref{fig:sectors_parameters} shows that, on
-average, the values of $K$
-in the ``Energy'' and ``Information Technology'' sectors are  larger
-than those in the ``Consumer Staples'' sector.  For a given loss \pro y, 
-a larger value of $K$ implies that large losses are more
-probable. Thus one can conclude that these two sectors are
-considerably riskier than the ``Consumer Staples'' sector. {\red Casper:
-we leave this sentence to your discretion.} This is of course a
-confirmation of one's economic instinct.
-
-Yet another indication from Figure \ref{fig:sectors_parameters} is
-that, while the ``Energy'' and the ``Information Technology'' sectors
-are similar in riskiness, the dependence  between $\alpha$ and $K$ is
-stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, when moving along
-a curve of equal preference in the direction of increasing $\alpha$, the parameter $K$ also
-increases. So the strong positive dependence seen in the ``Energy''
-sector suggests that these stocks might have very similar investor
-preferences. This in turn may be attributed to stronger business
-relations between the energy enterprises. While two IT companies
-may provide a variety of products and services and do not depend on each
-other, two energy companies are more likely to depend on each other
-via relations of supplier and customer or otherwise to compete with each
-other if they are on the same link of the chain of energy production
-and distribution.
-
-\subsection{A test for equal tail-indices based on Hill estimation}
-Suppose we have two independent strictly stationary positive series $X_1, \dots, X_n $ and
-$Y_1, \dots, Y_n$ with corresponding \df s $F_X$ and $F_Y$ that have power-law tails
-with indices $\alpha_X$ and $\alpha_Y$, respectively. 
-From \eqref{eq:2} one can deduce
-\begin{equation}
-  \label{eq:x3}
-  \sqrt k
-  \begin{pmatrix}
-    {\hat \alpha_X - \alpha_X} \\
-    {\hat \alpha_Y - \alpha_Y} \\
-  \end{pmatrix} \overset{d}{\to}
-  \begin{pmatrix}
-    Z_X \\
-    Z_Y
-  \end{pmatrix}
-  \sim
-  N\left(
-    0, \text{diag}(\alpha_X^2, \alpha_Y^2)
-  \right)
-\end{equation}
-where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimators of
-$\alpha_X$ and $\alpha_Y$; we suppress their dependence on $k$. Then it follows from the continuous mapping
-theorem
-\begin{equation}
-  \label{eq:x1}
-  \sqrt k [(\hat \alpha_X -\alpha_X) - (\hat \alpha_Y - \alpha_Y)]
-  \overset{d}{\to}
-  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)\,.
-\end{equation}
-This relation allows one to construct an \asy\ test under 
-the null hypothesis $\alpha_X = \alpha_Y$ and with test statistic $\hat \alpha_X - \hat \alpha_Y$.
-We apply this test to the equities in the ``Energy'',
-``Consumer Staples'' and ``Information Technology''
-sectors of the S\&P 500 index. The results are shown in the top row of
-Figure~\ref{fig:PairTest}.
-\begin{figure}[htb!]
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{HillTest_Energy.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{HillTest_CS.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{HillTest_IT.pdf}
-  \end{minipage}
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{Hoga_Energy_pair.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{Hoga_CS_pair.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{Hoga_IT_pair.pdf}
-  \end{minipage}
-  \caption{{\em Top row}: Test for pairwise equality of tail-indices
-    of losses in the  ``Energy'', ``Consumer Staples'' and
-    ``Information Technology'' 
-    sectors of S\&P 500. The test statistic in $\hat\alpha_X-\hat \alpha_Y$ is based
-    on Hill estimates 
-    of $\alpha_X$ and $\alpha_Y$. 
-    The green, blue and red points correspond to pairs of stock in a sector
-    when the test statistic is outside the intervals $[q_{0.075},q_{0.925}]$,
-    $[q_{0.05},q_{0.95}]$,  $[q_{0.025},q_{0.975}]$, respectively, where
-    $q_p$ is the $p$-quantile of the limiting 
-    $N(0,\alpha_X^2+\alpha_Y^2)$-\ds\ of the test statistic in
-    \eqref{eq:x1}. Grey points stand for pairs for which the test
-    statistic is inside $[q_{0.075},q_{0.925}]$. 
-    {\em Bottom row:} Test for changing tail-index or scale parameter
-    of losses
-    using Hoga's test based on concatenated series of pairs of
-    stocks. The green, blue and red points
-    correspond to pairs of stock in a sector 
-    when the test statistic $T_n$ exceeds the $85\%$-, $90\%$-,
-    $95\%$-quantile of the limit \ds .
-  Grey points stand for pairs for which the test statistic is below
-  the \asy\ $85\%$-quantile. Black points represent 
-    pairs for which the computation of $T_n$ fails for given precision  
-    requirements and time limits.
-    The same number (50) of upper order statistics is used for both tests.}
-  \label{fig:PairTest} 
-\end{figure}
-They indicate that tail-indices of equities in the
-``Energy'' or ``Information Technology'' sectors are more variable 
-than in the ``Consumer Staples'' sector, as the null
-hypothesis is rejected more often for members of these two former sectors.
-Moreover, these figures suggest that the test based on the Hill estimator
-is quite powerful in distinguishing between tail-indices. In contrast to the test presented in Section~\ref{sec:Hoga}
-the present test results in  more rejections for the ``Energy'' and the
-``Information Technology'' sectors.
-\par
-As a caution, one should bear in mind that 
-\eqref{eq:x3} is valid on condition that the $X$- and $Y$-
-series are independent of each other (or weakly dependent on each other), which is generally untrue for
-two return series in the same market.
-
-\subsection{A test for a change in the extreme tail}\label{sec:Hoga}
-Here we apply a test from
-a recent paper by Hoga \cite{hoga:2016}. This test has been developed
-for a different kind of problem. Given a strictly stationary
-\ts\ $X_1,\ldots,X_n$ with a marginal \ds\ $F$ with right power-law
-tail, the goal is to test whether there is a structural break of the
-{\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to
-zero. If the tail-index {\em or} the scale parameter in a \ds\ of type
-\eqref{eq:3} change inside a sample, then it is likely that the
-extreme quantiles change as well. We will test for a change of tail-index or scale parameter in this indirect way.
-\par
-The null hypothesis of the test in \cite{hoga:2016} is that there is no change of the extreme quantiles $F^{-1}(1-p)$ 
-for $p=p_n\to 0$ in any subsample with indices
-$t\in (n\,t_0,n(1-t_0))$ where $t_0$ is a fixed number in $(0,0.5)$. Writing $\hat x_p(a,b)$ for an estimator of the extreme $(1-p)$-quantile 
-based on the subsample with indices $t\in (na,nb)$, the test statistic is given by
-\beam\label{eq:4}
-T_n = \sup_{s \in [t_0, 1 - t_0]}
-  \dfrac{  \big[s (1 - s) \log \big(\hat x_p(0, s)/\hat x_p(s, 1)\big)
-    \big]^2}{
-    \int_{t_0}^s\big[r \log \big( \hat x_p(0, r)/\hat x_p(0, s)
-      \big)
-    \big]^2 dr
-    +
-    \int_{s}^{1 - t_0}
-    \big[
-      (1 - r) \log \big(
-      \hat x_p(r, 1)/
-      \hat x_p(s, 1)
-      \big)
-    \big]^2 dr}\nonumber\\
-\eeam
-Under the null hypothesis, $(T_n)$ converges to a complicated \fct al of \BM\ on $[0,1]$; the \asy\ quantiles need to be 
-evaluated by simulation.  
-\par
-When applied to our problem we would like to test 
-whether there is a change of the tail-index {\em or} scale parameter in \eqref{eq:3} in each of the S\&P 500 
-series in the distinct sectors. We also  want to get some indication about a possible change of tail-index or
-scale parameter from one series to another within a given sector. For this reason, we choose any pair of series
-within a sector and concatenate each of the paired series. Then we run the test on the concatenated series.
-Of course, despite the fact that we test changes of tail-index or scale parameter 
-{\em in a very indirect way} -- there may be many other reasons for the change of extreme quantiles in a sample -- 
-we also concatenate two rather distinct series. Even if we assume that the two series come from related models
-(such as GARCH), the parameters of these models will in general not be the same. Moreover, the concatenation
-of two strictly stationary \ts\ is in general not strictly stationary. Therefore we have to be careful
-with interpretations of the results of the tests.
-\par 
-In Figure~\ref{fig:Hoga_Single} we show the values of the test statistic $T_n$ (horizontal bars)
-for $t_0=0.1$ and daily return series of stock in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
-500 index. The ``null hypothesis'' is that the tail-index and scale parameter remain the
-same throughout the selected period of time. For most stocks, the hypothesis cannot be rejected even at the 85\% level.
-This fact may be an indication that the \ds\ inside a series is rather homogeneous.
-Alternatively, it may show that the power of the test is very low. A possible reason for this suspicion is that the \con\ rate
-of $(T_n)$ to its limit is very slow, i.e., the \asy\ \ds\ is not representative for the \ds\ of $T_n$ for the chosen $n$; for some
-simulation evidence, see below.
-\par
-To check whether any pair of stocks shares the same tail-index and scale parameter we concatenate
-any two series and apply the aforementioned test on the concatenated series. For the ``Energy'' and the
-``Consumer Staples'' sector we summarize the results in the bottom row of Figure~\ref{fig:PairTest}.
-These graphs show that the ``null
-hypothesis'' of an equal tail-index and scale parameter is rejected for more pairs in the
-``Energy'' sector than it is for those in the ``Consumer Staples''
-sector. This suggests that lower tail-indices of stocks in the
-``Energy'' sector are more spread out than those of the ``Consumer
-Staples'' sector. Also observe that while 3 stocks, say A, B and C, test in favor of
-the relations $\alpha_A = \alpha_B, \alpha_A \neq \alpha_C$, it often
-happens that another test on B and C is supportive of $\alpha_B = \alpha_C$. Again,
-this is due to the limited power of the test. Based
-on such results, one may guess that $\alpha_B$ lies between $\alpha_A$ and
-$\alpha_C$. The test is unable to recognize the smaller differences
-between $\alpha_A, \alpha_B$ on one hand and between $\alpha_B, \alpha_C$ on the other hand.
-\begin{figure}[htb!]
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]
-    {Hoga_Energy_Single.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]
-    {Hoga_CS_Single.pdf}
-  \end{minipage}
-  \caption{Test statistic $T_n$ from \eqref{eq:4} for the  stocks in the ``Energy'' and
-    ``Consumer Staples'' sectors of S\&P 500. The green, blue and red
-    lines correspond to the 85\%, 90\% and 95\% quantiles of the limit \ds\ of $T_n$.  They are derived
-by simulations from the limit \ds .}
-  \label{fig:Hoga_Single}
-\end{figure}
-
-To get an idea about the power of the test we run it on a
-sample concatenated from two independent iid samples of the same size
-$n=1304$ as the S\&P 500 series. Both pieces are $t$-distributed with
-distinct degrees of freedom. The results are shown in
-Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
-larger the minimum tail-index in the concatenated pair.
-\begin{figure}[htb!]
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{t_sim_pair.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
-  \end{minipage}
-  \caption{
-    {\em Left}: Test  of concatenated $t$-samples with different
-    degrees of freedom $\alpha$. Numbers on the axes are the degrees of freedom
-    in the subsamples. For an interpretation of the colored bullets,
-    see the caption for the bottom row of Figure~\ref{fig:PairTest}. The graph shows the limited power
-    of the test. In particular, if both degrees of freedom are
-    relatively large it loses the capability of distinguishing between the \ds s.
-    {\em Right}: Comparison of the \asy\ distribution of the test statistic
-    $T_n$ in \eqref{eq:4} under the null hypothesis and the \ds\ of
-    $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees
-    of freedom.
-  }
-  \label{fig:t_sim_pair}
-\end{figure}
-
-A major problem of this test is the \asy\ distribution of the test statistic under the
-null hypothesis.  The rate at which the finite-sample distribution tends to its limit is not known.
-To find out about this problem we compared the \ds s of $T_{1304}$ 
-for $t$-distributed $X_t$ with  $\alpha=3$ and $\alpha=4$ degrees of freedom with
-the limit \ds\ of $T_n$.
-The estimated density functions are shown on the right of Figure~\ref{fig:t_sim_pair}.
-As seen in the graph, the asymptotic distribution assigns
-significantly more mass to the tail than the \ds s of $T_n$ do. For
-comparison, we list a few quantiles of these \ds s in
-Table~\ref{tab:HogaAsymptotic}, showing major differences between the
-\asy\ and finite-sample \ds s.
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l|c|c|c|r}
-    & \multicolumn{4}{c}{Quantiles} \\[2mm]
-    \hline
-    Distribution of $X_t$& 80\% & 85\% & 90\% & 95\% \\
-    \hline
-    Asymptotic & 47.48 & 59.61 & 78.90 & 113.12 \\
-    t(3)  & 23.30 & 28.27 & 35.04 & 46.75 \\
-    t(4)  & 25.76 & 31.13 & 38.81 & 56.32\\[2mm]
-  \end{tabular}
-  \caption{Quantiles of the test statistic $T_n$ for $n=1304$
-    $t$-distributed samples with $\alpha=3$ and $\alpha=4$ 
-degrees of freedom
-    as well as the corresponding quantiles for the limiting \ds\ of
-    $T_n$. In particular, there are huge differences between the three
-    \ds s  for the higher quantiles. 
-    }
-  \label{tab:HogaAsymptotic}
-\end{table}
-
-Figure~\ref{fig:PairTest:permuted} points at another shortcoming
-of the test: we show $T_n$ for an 
-arbitrarily chosen random permutation of the concatenated data from two 
-different stocks. In this case the null hypothesis that two series
-have the same tail-index is rejected much more often, as a comparison
-with the bottom graphs of
-Figure~\ref{fig:PairTest} shows. If the data in the concatenated 
-series were iid a random permutation would not change the 
-\ds\ of $T_n$. Thus the value of the test statistic $T_n$  
-strongly depends on the dependence structure of the underlying data
-and therefore a test based on $T_n$ may be misleading.
-
-\begin{figure}[htb!]
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{Hoga_Energy_pair_permuted.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{Hoga_CS_pair_permuted.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.33\linewidth}
-    \includegraphics[
-      width=\textwidth,
-      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
-    ]{Hoga_IT_pair_permuted.pdf}
-  \end{minipage}
-  \caption{
-    Test for changing tail-index or scale parameter
-    of losses
-    using Hoga's test based on concatenated series of pairs of
-    stocks.
-    A random permutation is applied to the observations of each
-    series.
-    The green, blue and red points
-    correspond to pairs of stock in a sector 
-    when the test statistic $T_n$ exceeds the $85\%$-, $90\%$-,
-    $95\%$-quantile of the limit \ds .
-    Grey points stand for pairs for which the test statistic is below
-    the \asy\ $85\%$-quantile. Black points represent 
-    pairs for which the computation of $T_n$ fails for given precision  
-    requirements and time limits.
-  }
-  \label{fig:PairTest:permuted} 
-\end{figure}
-
-\newpage
-\section{Some theoretical arguments for equality of tail-indices}\setcounter{equation}{0}\label{sec:2}
-\subsection{Multivariate GARCH models whose components have  equal tail-indices}\label{subsec:garch}
-Among the models for returns the generalized autoregressive conditionally heteroscedastic (GARCH) model
-is certainly most popular because it is parsimonious, captures various of the stylized facts of real return data
-and can also be modified in various directions to capture specific behavior of \ts\ such a asymmetry, skewness, long memory; 
-see for example Andersen et al. \cite{andersen:davis:kreiss:mikosch:2009}, Part 1, for a collection of results
-on GARCH-type models.
-The original  {\em univariate} GARCH model of Bollerslev \cite{bollerslev:1986} is a \sv\ model of the type $X_t=\sigma_t\,Z_t$, where 
-$(Z_t)$ is an iid mean-zero unit-variance \seq . In the simple case of a \garch\ the {\em squared volatility} satisfies the {\em \sre} 
-\beam\label{eq:5}
-\sigma_t^2= \alpha_0+\alpha_1 X_{t-1}^2+\beta_1\sigma_{t-1}^2=\alpha_0+(\alpha_1Z_{t-1}^2+\beta_1)\sigma_{t-1}^2\,,\quad t\in\bbz\,.
-\eeam
-Here $\alpha_0>0$, $\alpha_1,\beta_1$ are non-negative constants. For suitable choices of $\alpha_1,\beta_1$ the equation \eqref{eq:5}
-can be solved and the solution $(\sigma_t^2)$ constitutes a strictly stationary \seq , implying that $(X_t)$ 
-is strictly stationary itself. A remarkable property of the process $(\sigma_t)$ is that it has a power-law tail 
-of the form
-\beam\label{eq:6}
-\P(\sigma_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
-\eeam
-for some positive $c>0$ and a positive tail-index $\alpha$ which is the unique solution of the equation 
-$\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$
-provided that the solution exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
-follows by an application of the Kesten-Goldie theorem; see Kesten \cite{kesten:1973}, Goldie \cite{goldie:1991}, cf. 
-Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} for a 
-recent textbook treatment. The latter result ensures power-law tails  for
-the strictly stationary  solution $(Y_t)$ to the \sre\ 
-\beam\label{eq:7}
-Y_t= A_t\,Y_{t-1}+B_t\,,\qquad t\in\bbz\,,
-\eeam
-for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^{\alpha/2}]=1$.  
-In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2+\beta_1$ and $B_t=\alpha_0$ to achieve
-\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016}, p.~275)
-it follows that
-\beao
-\P(\pm X_t>x)\sim \E[(Z_t)_\pm^\alpha]\,\P(\sigma_t>x)\,,\qquad \xto\,, 
-\eeao
-implying power-laws for the right and left tails of $X_t$ caused by the power-law tail of $\sigma_t$.
-
-A GARCH process of the order $(p,q)$ can be embedded in a multivariate
-equation of the type \eqref{eq:7}, where $(\bfA_t)$ are iid random
-matrices and $\bfB=\bfB_t$ is a constant vector. Again, the Kesten
-theory \cite{kesten:1973} applies, implying that the marginal and
-\fidi s of the GARCH process are \regvary\ with a positive index
-$\alpha$. We refrain from explaining the notion of multivariate
-\regvar\ which is needed in this context. For further details, see
-Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} where the
-Kesten theorem and \regvar\ of GARCH processes are explained in detail.
-\par
-There exist various extensions of the univariate GARCH
-model to the multivariate case. For the sake of argument, we stick here to the 
-{\em constant conditional correlation} (CCC) model of Bollerslev
-\cite{bollerslev:1990} and Jeantheau \cite{jeantheau:1998}, and we only consider a special bivariate case.
-It is the model 
-\beao\bfX_t=
-\left(\barr{l}X_{1,t}\\
-X_{2,t}\earr\right)= \left(\barr{ll}\sigma_{1,t}& 0\\
-0&\sigma_{2,t}\earr\right)\,\left(\barr{l}Z_{1,t}\\Z_{2,t}\earr\right)=\Sigma_t\,\bfZ_t\,,\qquad t\in\bbz\,.
-\eeao
-Thus both return components $X_{i,t}$ have the form of a univariate \sv\ model $X_{i,t}=\sigma_{i,t}Z_{i,t}$ 
-with non-negative volatility $\sigma_{i,t}$ and an iid bivariate noise \seq\ $(\bfZ_t)$ with zero mean and unit variance components.
-We also have the specification
-\beam\label{eq:8}
-\bfY_t=\left(\barr{l}\sigma^2_{1,t}  \\  
-\sigma^2_{2,t}\earr
-\right)
-&=& \left(
-\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)
-+\left(\barr{cc}\alpha_{11} & \alpha_{12}  \\
-      \alpha_{21} & \alpha_{22}\earr \right)\, 
-\left(\barr{l}X_{1,t-1}^2  \\X_{2,t-1}^2   \earr\right)
- + \left(\barr{cc}\beta_{11} & \beta_{12}  \\\beta_{21} & \beta_{22} \earr
- \right)\,\left(\barr{c}\sigma^2_{1,t-1}  \\\sigma^2_{2,t-1}\earr
-  \right)\nonumber\\
-&=& \left(
-\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)+\left(\barr{cc}\alpha_{11}Z_{1,t-1}^2+\beta_{11}&\alpha_{12}Z_{2,t-1}^2+
-\beta_{12}\\
-\alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
-\earr\right)\,\left(\barr{l}\sigma_{1,t-1}^2\\\sigma_{2,t-1}^2\earr
-\right)\,,
-\eeam
-for positive $\alpha_{0i}$ and suitable non-negative 
-$\alpha_{ij},\beta_{ij}$, $i,j=1,2$.
-Writing
-\beao
-\bfB_t= \left(
-\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)\quad\mbox{and}\quad
-\bfA_t=\left(\barr{cc}\alpha_{11}Z_{1,t-1}^2+\beta_{11}&\alpha_{12}Z_{2,t-1}^2+
-\beta_{12}\\
-\alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
-\earr\right)\,,
-\eeao
-we see that we are again in the framework of a \sre\ 
-but this time for vector-valued $\bfB_t$ and matrix-valued $\bfA_t$:
-\beam\label{eq:jan6b}
-\bfY_t=\bfA_t\,\bfY_{t-1}+\bfB_t\,,\qquad t\in\bbz\,.
-\eeam
-Kesten \cite{kesten:1973} also provided the corresponding theory  
-for stationarity and tails in this case. \sta\ \cite{starica:1999}
-dealt with the corresponding problems for CCC-GARCH processes,
-making use of the theory in Kesten \cite{kesten:1973},
-Bougerol and Picard \cite{bougerol:picard:1992}
-and its
-specification to the tails of GARCH models 
-in Basrak et al.~\cite{basrak:davis:mikosch:2002}. \sta\ \cite{starica:1999} assumed the 
-Kesten conditions for the matrices $\bfA_t$. These conditions ensure that the product matrices $\bfA_1\cdots\bfA_n$ 
-have positive entries for sufficiently large $n$. Then Kesten's theory implies that
-all components of the vector $\bfX_t$ have power-law tails with the same index $\alpha$ and also
-that the \fidi s of the process $(\bfX_t)$ are \regvary\ with index $\alpha$. 
-\par
-Various GARCH modifications are derived by considering linear combinations of CCC-GARCH models.
-The property of multivariate \regvar\ of multivariate GARCH ensures that, after linear transformations,  
-the new process in all components has again power-law tails with the same index as the original GARCH process; see 
-Basrak et al.~\cite{basrak:davis:mikosch:2002}. 
-Models which are constructed in this way are
-the Orthogonal GARCH model of
-Alexander and Chibumba \cite{alexander:chibumba:1996}, its
-generalization GO-GARCH by van der Weide \cite{Weide2002},  the Full Factor GARCH model of Vrontos et al.
-\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
-model of Lanne and Saikkonen  \cite{lanne2007modelling}. These
-models are characterized by their treatment of each series as a linear
-combination of factors, and each of the factors is modeled as a GARCH
-process; see Silvennoinen and Ter\"asvirt\"a \cite{silventeras:2009}.
-\par
-Not all choices of $\alpha$- and $\beta$-parameters in the model \eqref{eq:8} allow for an
-application of the Kesten theory. For example, assume that only the
-diagonal elements $\alpha_{ii}$ and $\beta_{ii}$ are positive.
-Then $\bfA_t$ is diagonal and, hence, the condition that
-$\bfA_1\cdots\bfA_n$ have positive entries for sufficiently large $n$ 
-cannot be satisfied. In the latter situation, both $(X_{1,t})$ and
-$(X_{2,t})$ are univariate \garch\ processes. Assuming the 
-conditions of the univariate Kesten-Goldie theorem for each component
-process, $(X_{1,t})$ and $(X_{2,t})$ have power-law tails 
-with indices $\kappa_1$ and $\kappa_2$, respectively,  given by the solutions to the equations 
-$\E [(\alpha_{ii} Z_{i,t}^2+\beta_{ii})^{\kappa_i/2}]=1$, $i=1,2$. In
-this model, one can introduce dependence between the two component
-series $(X_{1,t})$ and $(X_{2,t})$ by assuming dependence between the
-noise variables $Z_{1,t}$ and $Z_{2,t}$. Another situation when the
-Kesten theory fails 
-appears when $\bfA_t$ is an upper or lower triangle matrix: then the
-products  $\bfA_1\cdots\bfA_n$ are always of the same triangular
-type. 
-Similar remarks apply when one considers a CCC model in general
-dimension. Of course, one may argue that the latter models 
-are not natural: they are degenerate since they do not allow 
-for a linear relationship
-between all squared volatilities on a given day.
-
-\subsection{A utility based argument for equal tail-indices}\label{sec:3}
-In this section we give an argument based on economic theory that
-suggests equality of  tail-indices for equity return series.
-We follow an approach by Routledge and Zin
-\cite{routledge2010generalized} who introduced the notion of
-Generalized Disappointment Aversion (GDA). We consider the risky
-payoff $C$ of an investor and assume that it has a continuous \ds\ on
-$(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be
-increasing and concave on $(0,\infty)$.  Following Routledge and Zin
-\cite{routledge2010generalized},
-the utility of an agent with GDA preferences is given by
-\beao% \label{eq:xxie0}
-  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
-  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\,,
-  %&=&
-  %\E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
-\eeao
-where $\delta $ and $v$ are positive constants, and $b\ge 0$.
-  Here $v$ can be thought of as the {\em certainty payoff} equivalent to the
- risky payoff $C$; $\delta$ tunes the {\em threshold of 
-   disappointment} in proportion to $v$; $b$ determines the extra
- weight given to the expected return of $C$ when $C$ is below the
- disappointment threshold $\delta v$.
- If $b=0$, preferences are the classical expected utility. If
- $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
-disappointment aversion which were generalized by Routledge and Zin
-\cite{routledge2010generalized}.
-\par
-An agent guided by the utility function $u$ will seek to maximize the
-\fct al $\wt u$.
-Routledge and Zin assumed a power-law utility function 
-\beam\label{eq:hjyr}
-u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
-\eeam
-
-For the sake of argument, we assume that an investor initially has one
-unit of wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
-bond with interest rate $r>0$ and $\phi$ units in a risky asset with
-return $X$ over one time unit, i.e.,
-\beam\label{eq:xxie1}
-  C(X) = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
-\eeam
-Then we have 
-\beao
-\wt u(F_X, \phi) &=& \E [u(C)] + b\, \E \big[u(C)\1(C \le  \delta v)\big] - b \,u(\delta\, v) F_X(q)\,,
-\eeao
-where $F_X$ is the \df\ of $X$  and
-\beao
-  q = %q(r, \phi) &:=& \log \left( {
-      %\delta v - (1 - \phi) e^r
-      %\over
-      %\phi
-    % \right) 
-\log\Big(
-    \ex^r + {\delta \,v - \ex^r \over \phi}
-  \Big)\,.
-\eeao
-Note that $C \le \delta v$ \fif\ $X \le  q$. 
-\par
-Naturally, if an agent invests in
-a risky asset instead of a riskless bond, he expects to obtain a
-higher (on average) return from the risky asset than he is guaranteed from the
-riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
-$q > r $. For given $b,\delta,v$, the \fct al $\wt u$  depends only on $\phi$ and $F_X$,
-$\wt u=\wt u (F_X, \phi)$.
-We assume that
-\beao
-\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
-\eeao
-exists and that the maximum is achieved at a unique $\hat\phi \in (0, 1)$.
-
-\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
-Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
-we assume the following toy model. 
-We consider the case when $X$ has a two-sided Pareto \ds\ given by
-\beam\label{eq:pareto}
-  F_X(x) = \left\{
-  \begin{array}{ll}
-    p \left(
-    {K \over K - x}
-    \right)^\alpha & x \leq 0 \,,\\
-    1 - (1 - p) \left(
-    {K' \over K' + x}
-    \right)^\beta & x > 0\,,
-  \end{array}
-  \right.
-\eeam
-where
-$\alpha, \beta > 0$, $K,K' > 0$, $0 < p < 1$. 
-We also write $f_X$ for the density function of $F_X$.
-\par
-We have
-\beam\label{eq:xxie1.0}
- \wt u(F_X, \phi)\nonumber
-  &=&
-  \alpha\, K^\alpha\,  p\,
-  \int_{-\infty}^0
-  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
-  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
-    &&+
-  \beta \,(K')^\beta\, (1 - p)
-  \int_{0}^\infty
-  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
-  {1 + b \1(x <  q) \over (K' + x)^{\beta + 1}} \,dx 
-  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
-\eeam
-\par
-We observe the following property 
-whose proof is given in Appendix~\ref{sec:thrmI_proof}. 
-% which also applies to the special case 
-%\eqref{eq:xxie1.0}. 
-\ble\label{thrm:I}
-Assume the two-sided Pareto model \eqref{eq:pareto},
-that there is no \fct al relationship between $\alpha,K$ and $\beta,K'$
-and the utility \fct\ $u$ is increasing and differentiable. Then
-$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and
-$\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
-\ele
-We conclude that
-$\wt u_{\rm max}$ increases with $\alpha$ and decreases with
-$K$. Therefore there is a curve of  equal preference on the $(\alpha,
-K)$-plane. Moving along this curve in the direction of increasing
-$\alpha$, one expects the values of $K$ to increase too, i.e., the
-estimated values of $\alpha$ and $K$ 
-should appear positively dependent. Figure~\ref{fig:preference_pareto}
-illustrates this scenario for $\xi = 1/2, 4$ for the power-utility
-\fct\ \eqref{eq:hjyr}.
-In fact, this positive dependence is indeed observed for some
-real return data, e.g. the ``Energy'', ``Consumer Staples'' and
-``Information Technology'' sectors of the S\&P 500 index; see 
-Figure~\ref{fig:sectors_parameters}.
-\par
-A particularly interesting case occurs  when 
-$F_X$ is symmetric, i.e., when $\alpha = \beta$, $K =K'$ and $p=0.5$.
-Then \eqref{eq:pareto} turns into 
-\begin{eqnarray}
-  \wt u(F_X, \phi) &=& {\alpha \over 2} K^\alpha (1 + b)
-  \int_{0}^\infty {
-    u(C(x)) \left[
-      1 - {b \over 1 + b}\1_{\{x \ge q\}}
-      \right]
-    + u(C(-x))
-    \over
-    (K + x)^{\alpha + 1}
-  } dx
-  - b u(\delta v) F_X(q)\,.\nonumber\\
-  \label{eq:ipf}
-\end{eqnarray} 
-This situation is not covered by Lemma~\ref{thrm:I}:
-for the proof of the latter result we used Lemma \ref{lemma:I} whose
-assumptions are not satisfied in the present situation.
-Indeed, the integrand 
-\[
-U_{\text{all}}(x) =
-u(C(x)) \left[
-  1 - {b \over 1 + b}\1(x \ge q)
-  \right]
-+ u(C(-x))
-\]
-is not monotone.\footnote{To see this we may plug \eqref{eq:hjyr} 
-in $U_{\text{all}}$ and re-write it as
-\[U_{\text{all}}(x) =
-  -{\phi^{-\xi} \over \xi} \Big\{
-  \underbrace{
-    \Big(1 - {b \over 1 + b}\1(x \ge q)\Big)
-    \big[a + e^x\big]^{-\xi}
-    +
-    \Big[a + e^{-x}\Big]^{-\xi}
-  }_{U(x)} \Big\}\,,
-\]
-where $a = (1 - \phi) e^r/\phi$.
-Direct computation gives
-\begin{eqnarray*}
-  U'(x)
-  &=&
-  -{(a + e^x)^{-\xi - 1} \xi e^x
-    \over
-    1 + b \1(x \ge  q)
-  } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}\,,
-  \qquad x \neq q
-\end{eqnarray*} 
-The \fct\ $U(x)$, hence $U_{\text{all}}$, 
-is not monotone because $U'(x)>0$ for all large $x$ while $U(x)$ decreases
-in a small neighborhood of $q$.}
-\par
-We resort to numerical methods to gain some understanding of
-how $\tilde u(F_X, \phi)$ changes with $\alpha$.
-The value of $\hat\phi$ can be calculated
-by numerical integration and optimization with respect to $\phi$ for
-given values of $K, K', \alpha, \beta$.  This is shown in 
-Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\
-\eqref{eq:hjyr} both for fixed $K', \beta$ and for $K=K'$, $\alpha=\beta$.  
-The corresponding values $\tilde u_{\rm max}(\alpha, K)$
-are shown in Figure \ref{fig:preference_pareto}.
-If $K'$ and $\beta$ are fixed, both $\wt u_{\rm max}(\alpha, K)$ 
-and $\hat\phi$ increase  with $\alpha$ and decrease with $K$. 
-This is in agreement with Lemma~\ref{thrm:I}. 
-In contrast, when $K=K'$ are equal to $\alpha=\beta$
-$\wt u_{\rm max}(\alpha, K)$ decreases with $\alpha$ 
-but is rather insensitive with respect to $K$. On the
-other hand, $\hat\phi$ is not monotone with respect to $\alpha$ or
-$K$. For each fixed $K$, it peaks at an $\alpha$-value somewhere below
-1. For realistic values $\alpha\in (2, 4)$, $\hat\phi$ is a
-small value below 5\%.
-Since $\wt u_{\rm max}(\alpha, K)$ decreases with
-$\alpha$, investors who seek to maximize $\wt u_{\rm max}(\alpha, K)$
-will prefer the smallest $\alpha$ in the market, resulting in similar
-values of $\alpha$ for different equities.
-\begin{figure}[htb!]
-  \begin{minipage}{0.25\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1_A.pdf}    
-  \end{minipage}\hfill
-  \begin{minipage}{0.25\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_pareto4_A.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.25\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
-  \end{minipage}\hfill
-  \begin{minipage}{0.25\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
-  \end{minipage}
- \caption{
-   {\em The 1st and 2nd} graphs show $\hat\phi$, the optimal equity
-   allocation as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
-   \eqref{eq:pareto} for fixed $K'=0.012$, $\beta = 1.4$.
-   {\em The 3rd and 4th} graphs show $\hat\phi$ as a \fct\ of $\alpha$
-   and $K$ with $\beta = \alpha$ and $K' = K$.
-   We choose the utility \fct\ $u$ from \eqref{eq:hjyr} for $\xi = 1/2$
-   and $\xi = 4$, $b = 0.01$ in all cases.
-  }
-  \label{fig:phi_hat_pareto}
-\end{figure}
-
-\begin{minipage}{0.5\linewidth}
-  \includegraphics[width=\textwidth]{power_utilities.png}
-\end{minipage}\hfill
-\begin{minipage}{0.42\textwidth}
-  The economitrical differences between the two utility functions
-  are illustrated in the graph to the left.
-  We see that
-  $u(x)$ grows slower and saturates later for small $\xi$ than for
-  large $\xi$. The latter case represents an agent who is more
-  tolerant about low consumption and seeks wealth more
-  aggressively. In other words, he is less risk-averse than one with a
-  larger $\xi$. Such an agent will therefore invest more heavily in
-  equity, as shown in Figure~\ref{fig:phi_hat_pareto}. Note that,
-  while $\hat \phi$ changes nearly in the same in both graphs, the
-  scales in the graphcs are very different.
-\end{minipage}
-
-\begin{figure}[htb!]
-  \begin{minipage}{0.25\linewidth}
-    \includegraphics[width=\textwidth]{preference_pareto5e-1_A.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.25\linewidth}
-    \includegraphics[width=\textwidth]{preference_pareto4_A.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.25\linewidth}
-    \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.25\linewidth}
-    \includegraphics[width=\textwidth]{preference_pareto4.pdf}
-  \end{minipage}
-  \caption{
-   {\em The 1st and 2nd} graphs show
-   $\tilde u_{\rm max}(\alpha, K)$, as a \fct\ of $\alpha$ and $K$
-   in the two-sided Pareto model \eqref{eq:pareto} with $K'=0.012$,
-   $\beta = 1.4$.
-   Clearly, $\tilde u_{\rm max}(\alpha, K)$ increases with $\alpha$
-   and decreases with $K$ when $K'$ and $\beta$ are fixed.
-   {\em The 3rd and 4th} graphs show
-   $\tilde u_{\rm max}(\alpha, K)$ as a \fct\ of $\alpha$
-   and $K$ with $\beta = \alpha$ and $K' = K$.
-   We choose the utility \fct\ $u$ from \eqref{eq:hjyr} for $\xi = 1/2$
-   and $\xi = 4$. $b = 0.01$ in all cases.
-  }
-  \label{fig:preference_pareto}
-\end{figure}
-
-If the parameter $b$ is very small the GDA preference is closely
-approximated by the mean-utility preference, corresponding to $b = 0$. 
-In this case, we show in the proof of Lemma \ref{lemma:III} that the function
-$U_{\text{all}}(x)$ may increase or decrease depending on particular conditions
-on the values of $\xi$ and $(1 - \phi) e^r / \phi$:
-\begin{enumerate}
-\item If $\max\{a, 1\} < \xi$, $U_{\text{all}}(\cdot)$ is
-  monotone decreasing.
-\item If $a < \xi < 1$ and $(a + y_-)/(a y_- + 1) <
-  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
-  decreasing.
-\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotone
-  increasing.
-\item If $1 < \xi < a$ and $(a + y_+)/(a y_+ + 1) >
-  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
-  increasing.
-\item In other case, $U_{\text{all}}(\cdot)$ is not monotone.
-\end{enumerate}
-where
-\begin{eqnarray*}
-a &=& {
-  (1 - \phi) e^r
-  \over
-  \phi
-} \\
-y_\pm &=& {
-  a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
-  \over
-  a (\xi - 1)
-}
-\end{eqnarray*}
-Moreover, it is also easily checked that, when
-$x \in (0, K(e^{1/\alpha} - 1)]$,
-the density function in the integral of \eqref{eq:ipf} increases
-with $\alpha$; when $x \in (K(e^{1/\alpha} - 1), \infty)$ it 
-decreases with $\alpha$. Following the arguments for Lemma~\ref{thrm:I},
-and applying Lemma \ref{lemma:I}, it can be seen that
-$\wt u_{\rm max}(\alpha, K)$ increases/decreases with $\alpha$ when
-$U_{\text{all}}(x)$  decreases/increases.
-
-\section{Conclusion}
-\label{sec:4}
-{\red Casper, we believe you are the most knowledgeable to conclude.}
-\appendix
-\section{A monotonicity lemma}
-\setcounter{equation}{0}
-\begin{lemma} \label{lemma:I}
-  Assume distribution function $F(x, \theta)$ parameterized by 
-  $\theta \in \Theta \subseteq \mathbb R$ has support
-  $(a, b) \subseteq \mathbb R$,
-  and in addition $F(x, \theta)$ has density
-  function $f(x, \theta)$ that is differentiable with respect to
-  $\theta$ for all $\theta \in \Theta$.
-  Let $X \sim F$ and assume function $h(\cdot)$ is defined on $(a, b)$
-  and is monotone throughout this interval.
-  Moreover, we assume $h(x)$ and $f(x, \theta)$ satisfy
-  \begin{equation}\label{eq:frt}
-%    \E |h(X)| < \infty, \quad
-    \int_a^b \left| {\pd f(x, \theta) \over \pd \theta} \right| dx
-    < \infty \text { and }
-    \int_a^b \left| h(x) {\pd f(x, \theta) \over \pd \theta} \right| dx
-    < \infty
-  \end{equation}
-  Then the following holds true:
-  \begin{enumerate}
-  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
-    $\frac{\pd f}{\pd \theta}(x, \theta) > 0$ for $x \in (a, x_0)$ while
-    $\frac{\pd f}{\pd \theta}(x, \theta) < 0$ for $x \in (x_0, b)$, then
-    \[
-    \frac{\pd \E h(X)}{\pd \theta} > 0
-    \]
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    $\frac{\pd f}{\pd \theta}(x, \theta) < 0$ for $x \in (a, x_0)$  while
-    $\frac{\pd f}{\pd \theta}(x, \theta) > 0$ for $x \in (x_0, b)$, then
-    \[
-    \frac{\pd \E h(X)}{\pd \theta} > 0
-    \]
-  \end{enumerate}
-\end{lemma}
-\begin{remark}
-  \label{remark:I}
-  Two other cases follow trivially from lemma \ref{lemma:I}:
-  \begin{enumerate}
-  \item If $h(\cdot)$ is increasing and $\frac{\pd f}{\pd \theta}$
-    satisfies the
-    same conditions of the 1st case of lemma \ref{lemma:I},
-    $\frac{\pd \E h(X)}{\pd \theta} < 0$. This immediately follows
-    from applying 1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
-  \item By the same argument, if $h(\cdot)$ is decreasing and
-    ${\pd f \over \pd \theta}$ satisfies the same conditions of the
-    2nd case of lemma \ref{lemma:I}, ${\pd \E h(X) \over \pd \theta} < 0$.
-  \end{enumerate}
-\end{remark}
-
-\begin{proof}
-  Firstly, by dominated convergence theorem, conditions \eqref{eq:frt}
-  imply, for all $S \subseteq (a, b)$,
-  \begin{eqnarray*}
-    {\pd \over \pd \theta}\int_S f(x, \theta) dx
-    &=&
-    \int_S {\pd \over \pd \theta} f(x, \theta) dx \\
-    {\pd \over \pd \theta}\int_S h(x) f(x, \theta) dx
-    &=&
-    \int_S h(x) {\pd \over \pd \theta} f(x, \theta) dx \\
-  \end{eqnarray*}
-  Thus we have
-  \begin{eqnarray*}
-    {\pd \E h(X) \over \pd \theta}
-    &=&
-    {\pd \over \pd \theta}
-    \int_a^b h(x)
-    f(x, \theta) dx \\
-    &=& \int_a^b h(x)
-    {\pd f \over \pd \theta}(x, \theta) dx \\
-    &=& \underbrace{\int_a^{x_0}
-      h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_1}
-    + \underbrace{\int_{x_0}^b
-      h(x) {\pd f \over \pd \theta}(x,
-      \theta) dx}_{I_2}
-  \end{eqnarray*}
-  $x_0$ being located in the interior of $(a, b)$ and
-  $h(\cdot)$ being monotone imply $h(x_0) < \infty$.
-  \begin{enumerate}
-  \item When $h(x)$ is decreasing on $(a, b)$ and ${\pd f \over \pd
-    \theta}(x, \theta) > 0$ on $(a, x_0)$
-    \[
-    I_1 > h(x_0) \int_a^{x_0}
-    {\pd f \over \pd \theta}(x, \theta) dx
-    \]
-    Similarly, because ${\pd f \over \pd \theta}(x, \theta) < 0$ for
-    $x \in (x_0, b)$ and $h(x)$ is decreasing, we have
-    \begin{eqnarray*}
-      I_2 &=& \int_{x_0}^b -h(x_0)
-      \left|{\pd f \over \pd \theta}(x, \theta) \right| dx
-      > -h(x_0)
-      \int_{x_0}^b \left| 
-        {\pd f \over \pd \theta}(x, \theta)
-      \right| dx
-    \end{eqnarray*}
-    Finally we have
-    \begin{eqnarray*}
-      {\pd \E h(X) \over \pd \theta}
-      > h(x_0) \int_a^b
-      {\pd f \over \pd \theta}(x, \theta) dx
-      = h(x_0) {\partial \over \partial \theta}
-      \int_a^b f(x, \theta) dx
-      = 0
-    \end{eqnarray*}
-  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
-    ${\pd f \over \pd \theta}(x_0, \theta) < 0$ on $(a, x_0)$  while
-    ${\pd f \over \pd \theta}(x_0, \theta) > 0$ on $(x_0, b)$, 
-    by similar arguments, one can show
-    \[
-    {\pd \E h(X) \over \pd \theta}> 0
-    \]
-\end{enumerate}
-\end{proof}
-
-\section{When equity returns follow Student's t-distribution}
-\setcounter{equation}{0}
-It is a common practice to use Student's t-distribution to model the
-stationary distribution of equity returns. So it is of interest to
-find out what implications this distribution has when it is combined
-with the PDA preference. Formally we assume
-\[
-f(x; \alpha) = c(\alpha) \left(
-  1 + {x^2 \over \alpha}
-\right)^{-(\alpha + 1)/2}
-\]
-where $\alpha > 1$ and
-\[
-c(\alpha) = {
-  \Gamma({\alpha + 1 \over 2})
-  \over
-  \Gamma(\alpha/2) \sqrt{\alpha \pi}
-}
-\]
-In the same way as for \eqref{eq:ipf}, we can write $\wt u(F, \phi)$
-as
-\begin{small}
-  \begin{eqnarray}
-    \wt u(F, \phi)
-    &=&
-    (1 + b)
-    \int_{0}^{\infty}
-    \underbrace{
-      \left\{
-      u(C(x)) \left[
-        1 - {b \over 1 + b}\1_{\{x \ge q\}}
-        \right]
-      + u(C(-x))
-      \right\}
-    }_{U_{\text{all}}}
-    f(x, \alpha) dx - b u(\delta v) F_X(q)
-    \label{eq:t1}
-  \end{eqnarray}
-\end{small}
-where $C(\cdot)$ is defined in \eqref{eq:xxie1}. As shown in lemma
-\ref{lemma:III}, when $b = 0$ and $u(\cdot)$
-takes the power-form of \eqref{eq:hjyr}, $U_{\text{all}}$ is monotone
-depending on the values of $\xi$ and $(1 - \phi) e^r / \phi$. As given
-in lemma \ref{lemma:II}, there is a point $x_0 > 0$ such that
-${\pd f \over \pd \alpha}(x_0, \alpha) = 0$ and
-$\forall x \in (0, x_0),
-{\pd f \over  \pd \alpha}(x, \alpha) > 0$ and
-$\forall x \in (x_0, \infty)$,
-${\pd f \over \pd \alpha}(x, \alpha) < 0$. Thus it remains to verify
-$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$ and
-$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
-if we are to apply lemma \ref{lemma:I}.
-
-As computed in the proof of lemma \ref{lemma:II}, ${\pd f \over \pd
-  \alpha}(x, \alpha)$ is given by \eqref{eq:xxie4.1}. It is also shown
-there ${d \over d \alpha} c(\alpha) > 0$. Thus for
-$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
-it suffices to show
-\[
-\int_ 0^\infty {
-x^2
-\over
-(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
-} dx < \infty
-\]
-We may write
-\begin{eqnarray*}
-&& \int_ 0^\infty {
-x^2
-\over
-(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
-} dx \\
-&=& \left(\int_0^1 + \int_1^\infty \right) {
-x^2
-\over
-(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
-} dx \\
-&=& I_1 + I_2
-\end{eqnarray*}
-Clearly
-\[
-I_1 < \int_0^1 {1 \over \alpha} dx < \infty
-\]
-while
-\begin{eqnarray*}
-  I_2 &=& \int_1^\infty {1 \over 1 + \alpha/x^2}
-  {1 \over (1/x^2 + 1/\alpha)^{(\alpha+1) / 2}} {dx \over x^{\alpha +
-      1}} \\
-  &<& \int_1^\infty \alpha^{(\alpha+1) / 2} {dx \over x^{\alpha + 1}}
-  < \infty
-\end{eqnarray*}
-So we conclude
-$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
-To see
-$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$, we note
-\begin{eqnarray*}
-  \xi |U_{\text{all}}(x)|
-  &<& [(1 - \phi) e^r + e^x]^{-\xi} + [(1 - \phi) e^r + e^{-x}]^{-\xi} \\
-  &<& 1 + (1 - \phi)^{-\xi} e^{-r \xi}
-\end{eqnarray*}
-Since $\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$,
-it follows from the above inequality $\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
-Thus by lemma \ref{lemma:I}, $\wt u_{\rm max}$ is monotone
-increasing/decreasing with $\alpha$ when $U_{\text{all}}(\cdot)$ is
-monotone decreasing/increasing. Accordingly, an investor guided by the
-utility function will seek the largest/smallest $\alpha$ observed in
-the market.
-
-If however $b > 0$, Lemma \ref{lemma:I} is not applicable
-anymore. Nonetheless, numerical analysis lends some insight.
-As shown in figure \ref{fig:htfg}, $\hat\phi$ is monotone increasing
-for all 4 values of $b$, while $\tilde u_{\rm max}(\alpha)$ is
-increasing with $\alpha$ when $b$ is relatively large, but
-decreasing with $\alpha$ when $b$ is small. We note that a sizable
-value of $b$ indicates a conservative, risk-averse investor.
-
-\begin{figure}[htb!]
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
-  \end{minipage}
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{phi_hat_b_t_power4.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\linewidth}
-    \includegraphics[width=\textwidth]{U_b_t_power4.pdf}
-  \end{minipage}
-  \caption{$\hat\phi$ (left) and ${\pd \tilde u_{\rm max} \over \pd
-      \alpha}$ (right). {\em top:} $\xi = 1/2$. {\em bottom:} $\xi =
-    4$.
-  }
-  \label{fig:htfg}
-\end{figure}
-\begin{lemma}
-  \label{lemma:II}
-  Let $f$ denotes the density function of the Student's
-  t-distribution, i.e.
-  \[
-  f(x; \alpha) = c(\alpha) \left(
-    1 + {x^2 \over \alpha}
-  \right)^{-(\alpha + 1)/2}
-  \]
-  where $\alpha > 1$ and
-  \[
-  c(\alpha) = {
-    \Gamma({\alpha + 1 \over 2})
-    \over
-    \Gamma(\alpha/2) \sqrt{\alpha \pi}
-  }
-  \]
-  Then there exists $x_0 > 0$ such that ${\pd f \over
-    \pd \alpha}(x_0, \alpha) = 0$ and $\forall x \in (0, x_0), {\pd f \over
-    \pd \alpha}(x, \alpha) > 0$ and $\forall x \in (x_0, \infty), {\pd f
-    \over \pd \alpha}(x, \alpha) < 0$.
-\end{lemma}
-\begin{proof}
-  Straightforward computation gives
-  \begin{eqnarray}
-    {\pd f(x, \alpha) \over \pd \alpha} &=& {
-      c(\alpha) x^2 (\alpha + 1) + (2 \alpha x^2 + 2 \alpha^2) c'(\alpha)
-      -
-      \alpha c(\alpha) (x^2 + \alpha) \log(1 + x^2/\alpha)
-      \over
-      2 \alpha (x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
-    } \nonumber \\
-    &:=& {
-      P(x^2, \alpha)
-      \over
-      2 \alpha (x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
-    }
-    \label{eq:xxie4.1}
-  \end{eqnarray}
-
-  While the denominator of the right side of ${\pd f(x, \alpha) \over \pd \alpha}$ is
-  always positive, its numerator $P(x^2, \alpha)$ has a single root:
-  \begin{equation}
-    \label{eq:xxie5}
-    x_0^2 = \alpha\exp\left\{
-      W\left[
-        -\left(1 + {1 \over \alpha}\right)
-        e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
-      \right]
-      + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)}
-    \right\} - \alpha
-  \end{equation}
-  where $W(\cdot)$ is the principle branch of the Lambert $W$
-  function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
-  the right side of \eqref{eq:xxie5} for positivity, we first note
-  $c'(\alpha) > 0$:
-  \[
-  c'(\alpha) = {
-    \pi \Gamma(\alpha/2 + 1/2) \left\{
-      \alpha \left[ \Psi(\alpha/2 + 1/2) - \Psi(\alpha/2)\right] - 1
-    \right\}
-    \over
-    2 \Gamma(\alpha/2) (\pi \alpha)^{3/2}
-  }
-  \]
-  where $\Psi(\cdot)$ is the digamma function:
-  \[
-  \Psi(x) = {d \log[\Gamma(x)] \over d x}
-  \]
-  \begin{minipage}{0.48\textwidth}
-    $\Psi(x)$ \linebreak
-    \includegraphics[width=\textwidth]{digamma.png}
-  \end{minipage}\hfill
-  \begin{minipage}{0.5\textwidth}
-    As shown in the figure to the left, $\Psi(x)$ is increasing
-    for $x > 0$. This immediately follows from the series
-    representation
-    \[
-    \Psi(x + 1) = -\gamma + \sum_{n=1}^\infty {x \over n(n + x)}
-    \quad x \neq -1, -2, -3,\dots
-    \]
-    which gives
-    \[
-    {\pd \Psi(x + 1) \over \pd x}
-    = \sum_{n=1}^\infty {1 \over (n + x)^2} > 0
-    \]
-    See Abramowitz and Stegun \cite{abramowitz1972handbook}, p.259,
-    formula 6.3.16. Therefore $\Psi(\alpha/2 + 1/2) - \Psi(\alpha/2) > 0$.
-    So we have
-    \begin{eqnarray*}
-      && \alpha \left[
-        \Psi(\alpha/2 + 1/2) - \Psi(\alpha/2)
-      \right] - 1 \\
-      &\geq& 1 \times \left[
-        \Psi(1/2 + 1/2) - \Psi(1/2)
-      \right] - 1 \\
-      &=& \log(4) - \log(e) \\
-      &>& 0
-    \end{eqnarray*}
-    Thus $c'(\alpha) > 0$. Furthermore, we recall
-    $W(\cdot)$ is increasing on its principle branch. So
-  \end{minipage}
-  \begin{eqnarray*}
-    &&
-    W\left[
-      -\left( 1 + {1 \over \alpha} \right)
-      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
-    \right]
-    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \\
-    &>& 
-    W\left[
-      -\left( 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \right)
-      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
-    \right]
-    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \\
-    &=& W(-y e^{-y}) + y
-  \end{eqnarray*}
-  where
-  \[
-  y = 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} > 1
-  \]
-  Now notice
-  \[
-  \log(y e^{-y}) = \log(y) - y
-  \]
-  is a decreasing function for $y > 1$. Thus $-y e^{-y}$ is an
-  increasing function. Hence we have
-  \begin{eqnarray*}
-    W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
-  \end{eqnarray*}
-  Now it is clear
-  \[
-  \alpha\exp\left\{
-    W\left[
-      -\left(1 + {1 \over \alpha}\right)
-      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
-    \right]
-    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)}
-  \right\} - \alpha > 0
-  \]
-  Now that we have established that ${\pd f \over \pd \alpha}(x,
-  \alpha) = 0$ has a single positive root, it remains to determine the
-  sign of ${\pd f \over \pd \alpha}(x, \alpha)$ on the two sides of
-  the root. For this purpose we observe
-  \begin{equation}
-    \label{eq:xxie5.1}
-    P(0, \alpha) = 2 \alpha^2 c'(\alpha) > 0
-  \end{equation}
-  So we want to investigate ${\pd P \over \pd x}(x, \alpha)$:
-  \begin{eqnarray}
-    {\pd P \over \pd x}(x, \alpha) &=&
-    2 \alpha c'(\alpha) + c(\alpha) - \alpha c(\alpha) \log\left(
-      1 + {x \over \alpha}
-    \right)
-    \label{eq:xxie6}
-  \end{eqnarray}
-  Clearly, $\frac{\pd P}{\pd x}(0, \alpha) > 0$. Hence from
-  \eqref{eq:xxie5.1} and
-  \eqref{eq:xxie6} it is clear
-  \begin{equation*}
-    \text{sign}\left[
-      {\pd f \over \pd\alpha}(x, \alpha)
-    \right]
-    = \left\{
-    \begin{array}{rl}
-      1 & 0 < x < x_0 \\
-      -1 & x > x_0
-    \end{array}
-    \right.
-  \end{equation*}
-  where $x_0$ is the positive root of \eqref{eq:xxie5}.
-\end{proof}
-
-\section{Proof of Lemma~\ref{thrm:I}}
-\setcounter{equation}{0}
-\label{sec:thrmI_proof}
-\begin{proof}
-  Let
-  \begin{equation}
-    \label{eq:xxie2}
-    \hat \phi := \underset{0 < \phi \leq 1}{\text{argmax}}\,
-    \wt u(F_X, \phi)
-  \end{equation}
-  We have
-  \[
-  \tilde u_{\rm max}(F_X) = \wt u(F_X, \hat\phi)
-  \]
-  It follows
-  \begin{eqnarray}
-    {d \tilde u_{\rm max}(F_X) \over d \alpha}
-    &=&
-    \left.{\pd \wt u(\alpha, \phi) \over \pd \alpha}\right|_{\phi = \hat\phi}
-    + \left.{\pd \wt u(\alpha, \phi) \over \pd \phi}\right|_{\phi = \hat\phi}
-    {\pd \hat\phi \over \pd \alpha}
-    \label{eq:xxie3}
-  \end{eqnarray}
-  The definition \eqref{eq:xxie2} implies for all $\alpha$
-  \begin{equation}
-    \label{eq:xxie4}
-    \left.{\pd \wt u(\alpha, \phi) \over \pd \phi}\right|_{\phi = \hat\phi} = 0
-  \end{equation}
-  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
-  the first term is positive. From \eqref{eq:xxie1.0}, it follows
-  \begin{eqnarray*}
-    {\pd \wt u(\alpha, \phi) \over \pd \alpha}
-    = {\partial \over \partial \alpha}\E[u((1 - \phi) e^r + \phi e^x) \1_{\{X < 0\}}]
-  \end{eqnarray*}
-  The function $u((1 - \phi) e^r + \phi e^x)$ is obviously
-  increasing with $x$. It follows
-  \begin{eqnarray*}
-    {\pd f_X(x; \alpha, K) \over \pd \alpha}
-    = {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} 
-    = - {K^\alpha \over (K - x)^{\alpha + 1}}
-    \left[
-      \alpha
-      \log\left(
-        1 - {x \over K}
-      \right) - 1
-    \right]
-  \end{eqnarray*}
-  It is easily checked
-  \[
-  {\partial \over \partial \alpha}
-  {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
-  \left\{
-    \begin{array}{ll}
-      > 0 & \text{ when } x < K(1 - e^{1/\alpha}) < 0 \\
-      < 0 & \text{ when } K(1 - e^{1/\alpha}) < x < 0
-    \end{array}
-  \right.
-  \]
-% that when $x < K(1 - e^{1/\alpha}) < 0$,
-%   \[
-%   {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
-%   \]
-%   and when $K(1 - e^{1/\alpha}) < x < 0$,
-%   \[
-%   {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
-%   \]
-  This is the second case of lemma \ref{lemma:I}. So we have
-  ${\pd \wt u(\alpha, \phi) \over \pd \alpha} > 0$.
-  As for ${\pd \tilde u_{\rm max}(F_X) \over \pd K}$, by the same
-  argument, it suffices to show ${\pd \wt u(K, \phi) \over \pd K} < 0$.
-  We have
-  \[
-  {\pd \wt u(K, \phi) \over \pd K}
-  = {\partial \over \partial K} \E \left[
-    u((1 - \phi) e^r + \phi e^x) \1_{\{X < 0\}}
-  \right] 
-  \]
-  and
-  \begin{eqnarray*}
-    {\pd f_X(x; \alpha, K) \over \pd K}
-    = {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
-   = -\alpha K^{\alpha - 1} {
-      \alpha x + K
-      \over
-      (K - x)^{\alpha + 2}
-    }
-  \end{eqnarray*}
-  Clearly,
-  \[
-  {\partial \over \partial K}
-  {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
-  \left\{
-    \begin{array}{ll}
-      > 0 & \text{ when } x < -K/\alpha < 0\\
-      < 0 & \text{ when } -K/\alpha < x < 0
-    \end{array}
-  \right.
-  \]
-  % Clearly, when $x < -K/\alpha$
-  % \[
-  % {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
-  % \]
-  % and when $-K/\alpha < x < 0$
-  % \[
-  % {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
-  % \]
-  So by the 1st case of Remark \ref{remark:I} we conclude $  {\pd \wt u(K, \phi) \over \pd K} < 0$.
-\end{proof}
-
-\section{Lemma~\ref{lemma:III}}
-\setcounter{equation}{0}
-\begin{lemma}\label{lemma:III}
-Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:hjyr} and
-\eqref{eq:xxie1} respectively. Define
-\begin{eqnarray*}
-U_{\text{all}} &=& u(C(x))
-+ u(C(-x))
-\quad
-x \geq 0 \\
-a &=& {
-  (1 - \phi) e^r
-  \over
-  \phi
-} \\
-y_\pm &=& {
-  a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
-  \over
-  a (\xi - 1)
-}
-\end{eqnarray*}
-The following holds true:
-\begin{enumerate}
-\item If $\max\{a, 1\} < \xi$, $U_{\text{all}}(\cdot)$ is
-  monotone decreasing.
-\item If $a < \xi < 1$ and $(a + y_-)/(a y_- + 1) <
-  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
-  decreasing.
-\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotone
-  increasing.
-\item If $1 < \xi < a$ and $(a + y_+)/(a y_+ + 1) >
-  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
-  increasing.
-\item In other case, $U_{\text{all}}(\cdot)$ is not monotone.
-\end{enumerate}
-\end{lemma}
-\begin{proof}
-  It is convenient to re-write $U_{\text{all}}$ as
-  \[
-  U_{\text{all}} =
-  -{\phi^{-\xi} \over \xi} \left\{
-    \underbrace{
-      \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
-      +
-      \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
-    }_{U(x)}
-  \right\}
-  \]
-  Thus the monotonicity of $U_{\text{all}}(x)$ is the opposite of $U(x)$. For
-  convenience of writing, let $a = (1 - \phi)e^r/\phi$.
-  First of all, we find the conditions on which $U(x)$ is monotone.
-  Direct differentiation yields
-  \begin{eqnarray*}
-    {\pd U(x) \over \pd x}
-    &=&
-    -{
-      (a + e^x)^{-\xi - 1} \xi e^x
-    } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}
-  \end{eqnarray*}
-  ${\pd U(x) \over \pd x} = 0$ is equivalent to
-  \begin{eqnarray*}
-    {a + y \over a y + 1} &=& y^{1 -\xi \over 1+\xi} \\
-    \underbrace{\log(a + y) - \log(a y + 1)}_{f(y)} &=&
-    {1 -\xi \over 1+\xi} \log(y)
-  \end{eqnarray*}
-  where we have defined $y = e^x$, and $f(y)$, $g(y)$ as above. Observe
-  \begin{equation}
-    \label{eq:f-g-der}
-    {\pd (f -g)(y) \over \pd y} = {
-      a (\xi - 1) y^2 + 2 (\xi - a^2) y + a(\xi - 1)
-      \over
-      (1 + \xi) (a + y) (a y + 1) y
-    }
-  \end{equation}
-  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots when $\xi \ne 1$
-  \begin{eqnarray*}
-    y_\pm &=& {
-      a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
-      \over
-      a (\xi - 1)
-    }
-  \end{eqnarray*}
-  \begin{enumerate}
-  \item If $\min\{1, \xi\} < a < \max\{1, \xi\}$, $y_\pm$ are not
-    real, ${\pd (f -g)(y) \over \pd y} = 0$ has no solution on
-    $(1, \infty)$; $(f -g)(y)$ is monotone on $(1, \infty)$.
-    \begin{enumerate}
-    \item If in addition $\xi > 1$, i.e. $1 < a < \xi$, $(f - g)(y)$
-      is monotone increasing on $(1, \infty)$, because the
-      coefficient of the $y^2$ term of the numerator of
-      \eqref{eq:f-g-der}, i.e. $a (\xi - 1)$ is positive. We note
-      $f(1) = g(1) = 0$; thus on $(1, \infty)$, there is no solution
-      to $f(y) = g(y)$. It can be concluded that $U(\cdot)$ is
-      monotone on $(0, \infty)$. Furthermore
-        \begin{eqnarray*}
-          \left.{\pd U(x) \over \pd x}\right|_{x=0} &=& 0 \\
-        \end{eqnarray*}
-        For a small $\epsilon > 0$, the sign of ${\pd U(x) \over \pd
-          x}$ on $(0, \epsilon)$ is thus the same as
-        $\left.{\pd^2 U(x) \over \pd x^2}\right|_{x=0}$:
-        \[
-        \left.{\pd^2 U(x) \over \pd x^2}\right|_{x=0} =
-        2 (a + 1)^{-\xi - 2} (\xi - a) > 0
-        \]
-        Thus ${\pd U(x) \over \pd x} > 0$ for $x \in (0, \infty)$;
-        $U_{\text{all}}$ is monotone decreasing.
-    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar
-      argument as in the previous case, $U_{\text{all}}$ is monotone
-      increasing on $(0, \infty)$.
-    \end{enumerate}
-  \item If $a < \min\{1, \xi\}$ and $\xi > 1$, i.e. $a < 1 < \xi$, it
-    is clear
-    \[
-    y_- = {
-      a^2 - \xi - \sqrt{(a^2 - 1) (a^2 - \xi^2)}
-      \over
-      a (\xi - 1)
-    } < 0
-    \]
-    It remains to compare $y_+$ with $1$ to determine whether
-    ${\pd (f -g)(y) \over \pd y} = 0$ has a solution on $(1, \infty)$.
-    Assume $y_+ > 1$. Then
-    \begin{eqnarray}
-      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1)
-      \nonumber \\
-      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0 \label{eq:mi6}
-    \end{eqnarray}
-    This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
-    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotone increasing
-    on $(1, \infty)$. Following the same analysis as in the case
-    (1.a), one can see $U_{\text{all}}$ is monotone decreasing.
-
-  \item If $a < \min\{1, \xi\}$ and $\xi < 1$, i.e. $a < \xi < 1$, it
-    is clear $y_- > 0$ and $y_+ < y_-$. Moreover, $y_- > 1$ is equivalent
-    to
-    \[
-    \sqrt{(a^2 - 1) (a^2 - \xi^2)} > a^2 - \xi a + a - \xi
-    = (a + 1) (a - \xi)
-    \]
-    The last inequality is obviously true in this case. So $y_- >
-    1$. ${\pd (f-g)(y) \over \pd y}$ has a maximum at
-    $y_- \in (1, \infty)$. We know $(f - g)(y_-)$ is a maximum because
-    \eqref{eq:f-g-der} shows that, for $y > y_-$,
-    ${\pd (f-g)(y) \over \pd y} < 0$.
-
-    If $(f-g)(y_-) < 0$, $(f-g)(y) = 0$ has no solution on $(1,
-    \infty)$. Hence $U(\cdot)$ is monotone on both $(0, \infty)$.
-    The same analysis as in the case (1.a) shows $U_{\text{all}}$
-    is monotone decreasing on $(0, \infty)$.
-
-    If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
-    \infty)$. So $U(\cdot)$ is not monotone.
-
-    \item $a > \max\{1, \xi\}$ and $\xi > 1$, i.e. $1 < \xi < a$. By
-      the same argument that leads to \eqref{eq:mi6}, we see
-      $y_+ > 1$. From \eqref{eq:f-g-der} it is clear $(f-g)(y_+)$ is a
-      minimum. If $(f-g)(y_+) > 0$, there is no solution to
-      $(f-g)(y) = 0$ on $(1, \infty)$; $U(x)$ is monotone. By the
-      same analysis as in the case (1.a), we know $U(x)$ is
-      monotone decreasing and $U_{\text{all}}$ is monotone increasing.
-
-      if $(f-g)(y_+) < 0$, there must be a solution to $(f-g)(y) = 0$
-      on $(y_+, \infty)$. $U_{\text{all}}$ is not monotone.
-  \end{enumerate}
-\end{proof}
+\input{PaperBody}
 

### 2017-07-21 phd/Chapterextremes.tex
Added my own thesis.

@@
 
-\begin{abstract}{We  provide some  asymptotic theory for the largest eigenvalues of a sample covariance matrix
-of a $p$-dimensional time series where the dimension $p=p_n$ converges to infinity when the sample size $n$ increases.
-We give a short overview of the literature on the topic both in the light- and heavy-tailed cases when the data have
-finite (infinite) fourth moment, respectively.
-Our main focus is on the heavy-tailed case. In this case, one has a theory for the point process of the normalized eigenvalues
-of the sample covariance matrix in the iid case but also when rows and columns of the data are linearly dependent.
-We provide limit results for the weak convergence of these point processes to Poisson or cluster Poisson processes. Based o
-this convergence we can also derive the limit laws of various functionals of the ordered eigenvalues such as the
-joint convergence of a finite number of the largest order statistics, the joint limit law of the largest eigenvalue and the trace,
-limit laws for successive ratios of ordered eigenvalues,
-etc. We also develop some
-limit theory for the singular values of the sample autocovariance matrices and their sums of squares. The theory is illustrated
-for simulated data and for the components of the S\&P 500 stock index.
-\medskip
-
-\noindent {\bf Keywords:} Regular variation, sample covariance matrix, dependent entries,
-largest  eigenvalues, trace, point process
-  convergence, cluster Poisson limit,
-infinite variance stable limit, Fr\'echet distribution.}
-\end{abstract}
-
-\newpage
-\section[Estimation of the largest eigenvalues]{Estimation of the largest eigenvalues: an overview in the iid case}\label{sec:motivationch3}
-\subsection{The light-tailed case}
-%In recent years, \asy\ theory for large random matrices has a
-%attracted a
-%great deal of attention; see for example the monographs
-%Bai and Silverstein \cite{bai:silverstein:2010} and Anderson et al.
-%\cite{anderson:guionnet:zeitouni:2008}. The literature about {\em
-%  heavy-tailed random matrices} is rather sparse. Here and in what
-%follows, we call a random matrix {\em heavy-tailed} if its entries have
-%\ds s with \regvary\ tails, typically with tail index below 4.
-%The eigenvalues of heavy-tailed matrices
-%with iid entries and dimension $p\times n$, where $p=p_n\to\infty$ and $p/n\to
-%\gamma\in (0,\infty)$, were studied by Soshnikov \cite{soshnikov:2004,
-%  soshnikov:2006} and  Auffinger et
-%al. \cite{auffinger:arous:peche:2009}. In the latter reference and in
-%Belinschi et al. \cite{belinschi:dembo:guionnet:2009},
-%the sample covariance matrices of iid heavy-tailed \seq s were also studied.
-%Bose et al. \cite{bose:hazra:saha:2009,bose:hazra:saha:2010}
-%investigated the spectral norm of circulant-type heavy-tailed matrices
-%and Ben Arous and Guionnet \cite{arous:guionnet:2008}
-%studied the limits of heavy-tailed random Wigner matrices with
-%infinite variance.
-One of the exciting new areas of statistics is concerned with analyses of large data sets.
-For such data one often studies the dependence structure via covariances and correlations.
-In this paper we focus on one aspect: the estimation of the eigenvalues of the covariance matrix of a multivariate  \ts\
-when the dimension $p$ of the series increases with the sample size $n$.
-In particular, we are interested in limit theory for the largest eigenvalues of the sample covariance matrix.
-This theory is closely related to topics from classical \evt\ such as
-maximum domains of attraction with the corresponding normalizing and centering constants for maxima; cf.
-Embrechts et al.~\cite{embrechts:kluppelberg:mikosch:1997}, Resnick~\cite{resnick:2007,resnick:1987}.
-Moreover, \pp\ \con\ with limiting Poisson and cluster Poisson processes enters in a natural way when one describes
-the joint \con\ of the largest eigenvalues of the sample covariance matrix. Large deviation techniques find applications,
-linking \evt\ with random walk theory and \pp\ \con . The objective of this paper is to illustrate some of the main
-developments in random matrix theory for the particular case of the sample covariance matrix of multivariate \ts\ with
-independent or dependent entries. We give special emphasis to the heavy-tailed case when \evt\ enters in a rather
-straightforward way.
-%We do not cover other parts of random matrix theory such as limit theory for the eigen-spectrum of random matrices with
-%iid entries or other special types of matrices. Also in this case one can find many parallels with \evt.
-\par
-Classical multivariate \tsa\ deals with
-observations which assume values in a $p$-dimensional space
-where $p$ is ``relatively small'' compared to the sample size $n$.  With
-the availability of large data sets $p$ can be ``large''
-relative to $n$. One of the possible con\seq s is that standard asymptotics (such as the \clt )
-break down and may even cause misleading results.
-\par
-The dependence
-structure in multivariate data is often summarized by the covariance matrix which is typically
-estimated by its sample analog.  For example, principal component analysis (PCA)
-extracts principal component vectors corresponding to the largest
-eigenvalues of the sample covariance matrix. The magnitudes of these eigenvalues provide an empirical \ms\ of the importance
-of these components.
-\par
-If $p,n$ are fixed, a column of the $p\times n$ data matrix
-\beao
-\X =\X_n = \big(X_{it}\big)_{i=1,\ldots,p;t=1,\ldots,n}\,
-\eeao
-represents an observation of a $p$-dimensional \ts\ model with unknown parameters.
-In this section we assume that the real-valued entries $X_{it}$ are iid, unless mentioned otherwise, and we write $X$ for a generic element.
-One challenge is to infer information about the parameters
-from the eigenvalues $\lambda_1,\ldots,\lambda_p$
-of the {\em sample covariance matrix} $\X\X'$. In the notation we suppress the dependence of $(\la_i)$ on $n$ and $p$.
-If $p$ and $n$ are finite and the columns of $\X$ are iid and multivariate normal,
-Muirhead \cite{muirhead} derived a (rather complicated) formula for the joint distribution of the eigenvalues $(\lambda_i)$.
-%Therefore a theoretical solution is already available in the finite $n$ case.
-%The expression, however, contains an integral over the orthogonal group $\mathbb{O}(p)$.
-%\begin{figure}[htb!]
-%  \centering
- % \includegraphics[scale=0.5]{eigen_normal_qqplot1.pdf}
- % \caption{Largest eigenvalues of $500$ simulated sample covariance matrices against standard normal quantiles.}
- % \label{fig:QQ}
-%\end{figure}
-
-For $p$ fixed and $\nto$, assuming
-$\X$ has centered normal entries and a diagonal covariance matrix $\Sigma$,
-Anderson~\cite{anderson:1963}  derived the joint \asy\ density of $(\lambda_1, \ldots, \lambda_p)$.
-%If the largest eigenvalue of the underlying covariance matrix appears with multiplicity one, he
-%showed that the largest eigenvalue of the sample covariave matrix
-%is asymptotically normal.
-We quote from Johnstone~\cite{johnstone:2001}:
-``The classic paper by Anderson~\cite{anderson:1963} gives the limiting joint distribution of the roots, but the
-marginal distribution of the largest eigenvalue is hard to extract even in the null case'' (i.e., when the covariance matrix $\Sigma$ is
-proportional to the identity matrix).
-%If the largest entry of $\Sigma$ exceeds its second largest, one obtains a central limit type theorem for the largest eigenvalue of the sample covariance matrix.
-%In Figure~\ref{fig:QQ} we chose $n=10000, p=50$ and $\Sigma =\diag(1024,1,\ldots,1)$, and plot the properly normalized and
-%standardized largest eigenvalue against standard normal quantiles.
-
-It turns out that limit theory for the largest eigenvalues becomes ``easier''
-when the dimension $p$ increases with $n$.
-Over the last 15 years there has been increasing interest in the case when $p=p_n\to\infty$ as $\nto$. In most of
-the literature (exceptions are El Karoui \cite{elkaroui:2003}, Davis et al. \cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} and Heiny and Mikosch
-\cite{heiny:mikosch:2015:iid})
-one assumes that $p$ and $n$ grow at the same rate:
-\beam\label{eq:gammach3}
-\dfrac p n\to \gamma\qquad \mbox{for some $\gamma \in (0,\infty)$.}
-\eeam
-%One explanation for the commonly used condition \eqref{eq:gamma} is that much of the theory was developed for $n \times n$ {\em Wigner matrices}, which is a Hermitean matrix $H$ such that $\{ H_{ij}:i<j \}$ and $\{ H_{ii}:0\le i\le n \}$ are two sets of iid random variables.
-If $p =\gamma n$, it turns out that the limit theory for the eigenvalues of Wigner matrices and sample covariance matrices is closely connected by the convergence of the empirical spectral distributions to the semicircle and the Mar\v cenko--Pastur law, respectively; see Bai and Silverstein \cite{bai:silverstein:2010}, Chapters 1-3.
-
-In random matrix theory,  the convergence of the {\em empirical spectral distributions} $(F_{n^{-1}\X \X'})$ of a sequence $(n^{-1}\X \X')$ of non-negative definite matrices is the principle object of study.
-The empirical spectral distribution $F_{n^{-1}\X \X'}$ is constructed from the eigenvalues via
-\begin{equation*}
-F_{n^{-1}\X \X'}(x)= \frac{1}{p}\; \# \{ 1\le j\le p : n^{-1} \lambda_j \le x \}, \quad x\in \R,\quad n\ge1.
-\end{equation*}
-In the literature convergence results for the sequence of empirical spectral distributions are established under the assumption that $p$ and $n$ grow at the same rate.
-Suppose that the iid entries $Z_{it}$ have mean $0$ and variance $1$.
-If \eqref{eq:gammach3} holds, then, with probability one, $(F_{n^{-1}\X \X'})$ converges weakly to the celebrated Mar\v cenko--Pastur law $F_\gamma$.
-If $\gamma \in (0,1]$,  $F_\gamma$  has density,
-\begin{eqnarray}\label{eq:MPch3}
-f_\gamma(x) =
-\left\{\begin{array}{ll}
-\frac{1}{2\pi x\gamma} \sqrt{(b-x)(x-a)} \,, & \mbox{if } a\le x \le b, \\
- 0 \,, & \mbox{otherwise,}
-\end{array}\right.
-\end{eqnarray}\noindent
-where $a=(1-\sqrt{\gamma})^2$ and $b=(1+\sqrt{\gamma})^2$.
-If $\gamma>1$, the \MP law is a mixture of a point mass at $0$ and the density function $f_{1/\gamma}$ with weights $1-1/\gamma$ and $1/\gamma$, respectively.
-The point mass at $0$ is intuitively explained by the fact that, with probability $1$, $\min(p,n)$ eigenvalues $\lambda_i$ are non-zero.
-When $n=(1/\gamma) \; p$ and $\gamma >1$ one sees that the proportion of non-zero eigenvalues of the sample covariance matrix is $1/\gamma$ while the proportion of zero eigenvalues is $1-1/\gamma$.
-
-While the finite second moment is the central assumption to obtain the Mar\v cenko--Pastur law as the limiting spectral distribution, the finite fourth moment plays a crucial role when studying the largest eigenvalues
-\beam\label{eq:orderch3}
-\la_{(1)}\ge \cdots \ge \la_{(p)}
-\eeam
-of $\X\X'$, where we suppress the dependence on $n$ in the notation.
-
-\begin{figure}[htb!]
-  \centering
-  \subfigure[Standard normal entries]{
-    \includegraphics[scale=0.34]{normal-TW.pdf}
-  }
-  \subfigure[Entry distribution: $\P(X=\sqrt{3}) = \P(X=-\sqrt{3} ) =
-  1/6$, $\P(X=0)=2/3$. Note $\E X = 0$, { $\E[ X^2] = 1$, $\E [X^3] = 0$ and
-  $\E[ X^4] = 3$, i.e., the first 4 moments of $X$ match those of the standard normal \ds .}]{
-    \includegraphics[scale=0.34]{3point-TW.pdf}
-  }
-  \caption{Sample density function of the largest eigenvalue compared
-    with the Tracy--Widom density function. The data matrix $ \X$ has
-    dimension $200 \times 1000$. An ensemble of 2000 matrices is
-    simulated.}
-  \label{fig:normal-3point-TW}
-\end{figure}
-
-Assuming \eqref{eq:gammach3} and iid entries $X_{it}$ with zero mean, unit variance and finite fourth moment,  Geman~\cite{geman}
-showed that
-\beam\label{eq:gemanch3}
-\dfrac {\la_{(1)}}{n} \stas \big(1+\sqrt{\gamma}\big)^2\,,\qquad \nto\,.
-\eeam
-Johnstone \cite{johnstone:2001} complemented this \slln\ by the corresponding \clt\ in the special case of iid standard normal
-entries:
-\beam\label{eq:tcch3}
- n^{2/3}\,\dfrac{(\sqrt{\gamma})^{1/3}}{\big(1+\sqrt{\gamma}\big)^{4/3}}\Big(\dfrac {\la_{(1)}}{n} -
-\big(1+\sqrt{\tfrac pn }\big)^2\Big)
-\std {\rm TW}\,,
-\eeam
-where the limiting \rv\ has a {\em Tracy--Widom \ds} of order 1. Notice that the centering
-$\big(1+\sqrt{\tfrac pn }\big)^2$ can in general not be replaced by $(1+\sqrt{\gamma})^2$.
-This \ds\ is ubiquitous in random matrix theory.
-%It is defined via some ordinary differential equation; we refer to \cite{tracy:widom:2012} for a definition and properties.
-Its distribution function $F_1$ is given by
-\begin{equation*}
-F_1(s) = \exp\Big\{
-  -\frac{1}{2} \int_{s}^\infty [
-    q(x) + (x - s) q^2(x)
- ] \dint x
-\Big\}\,,
-\end{equation*}
-where $q(x)$ is the unique solution to the Painlev\'e II differential
-equation
-\begin{equation*}
-  q''(x) = xq(x) + 2 q^3(x)\,,
-\end{equation*}
-where $ q(x)\sim {\rm Ai}(x)$ as $x \to \infty$ and Ai$(\cdot)$ is the Airy kernel; see Tracy and Widom~\cite{tracy:widom:2012} for details.
-We notice that the rate $n^{2/3}$ compares favorably to the $\sqrt{n}$-rate in the classical \clt\ for sums
-of iid finite variance \rv s.
-The calculation of the spectrum is facilitated by the fact that the distribution of
-the classical Gaussian matrix ensembles is invariant under orthogonal transformations. The corresponding
-computation for non-invariant matrices with non-Gaussian entries is more complicated and was a major challenge for several years; a first step was made by Johansson \cite{johansson}.
-Johnstone's result was extended to matrices $\X$ with iid non-Gaussian entries
-by Tao and Vu \cite[Theorem~1.16]{tao09b}. Assuming that the first four moments of the entry \ds\ match  those of
-the standard normal \ds , they showed \eqref{eq:tcch3} by
-employing {\em Lindeberg's replacement method}, i.e., the iid non-Gaussian entries are replaced
-step-by-step by iid Gaussian ones.
-This approach is well-known from summation theory for \seq s of iid \rv s.
-Tao and Vu's result is a consequence of the so-called {\em Four Moment Theorem}, which describes the insensitivity of the eigenvalues with respect to changes in the distribution of the entries.
-To some extent (modulo the strong moment matching conditions) it shows the universality of Johnstone's limit
-result \eqref{eq:tcch3}. Later we will deal with entries with infinite fourth moment. In this case, the weak limit
-for the normalized largest eigenvalue $\la_{(1)}$ is distinct from the Tracy--Widom \ds : the classical Fr\'echet extreme value
-\ds\ appears.
-In Figure~\ref{fig:normal-3point-TW} we illustrate how the Tracy--Widom approximation works for Gaussian and non-Gaussian entries of $\X$
-and in Figure~\ref{fig:MyDist-Frechet} we also illustrate that this approach fails when $\E[X^4]=\infty$.
-
-Figure \ref{fig:normal-3point-TW} compares the sample density function
-of the properly normalized largest eigenvalue estimated from 2000 simulated sample covariance matrices $\X\X'$ ($n=1000, p=200$) with
-the Tracy--Widom density. If  $X$ has infinite fourth moment and further regularity conditions on the tail hold then the
-Tracy--Widom limiting law needs to be replaced by the \Frechet
-distribution; see Section~\ref{subsec:1.2} for details. Figure \ref{fig:MyDist-Frechet} illustrates this fact with a
-simulated ensemble whose entries are distributed according to
-the heavy-tailed distribution from \eqref{eq:distrsimch3} below with $\alpha = 1.6$.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.5]{MyDist-Frechet.pdf}
-  \caption{Sample distribution function of the largest eigenvalue $\la_{(1)}$
-    compared to the \Frechet distribution (solid line) with $\alpha=1.6$. The data matrices have dimension
-    $200 \times 1000$ and iid entries with infinite fourth moment. The results are based on 2000 replicates.}
-  \label{fig:MyDist-Frechet}
-\end{figure}
-
-\subsection{The heavy-tailed case}\label{subsec:1.2}
-So far we focused on ``light-tailed'' $\X$
-in the sense that its entries have finite fourth moment.  However, there is statistical evidence that the assumption
-of finite fourth moment may be violated when dealing with data from
-insurance, finance or telecommunications. We illustrate this fact
-in Figure~\ref{fig:SP500_tail_indices} where we show the pairs $(\alpha_L,\alpha_U)$ of
-lower and upper tail indices
-of $p=478$  log-return series composing
-the S\&P 500 index estimated from $n=1,345$ daily observations from 01/04/2010 to 02/28/2015.
-This means we assume for every row  $(X_{it})_{t=1,\ldots,n}$ of $\X$ that the tails behave like
-\beao
-\P(X_{it}>x)\sim c_U\,x^{-\alpha_U}\qquad\mbox{and}\qquad \P(X_{it}<-x)\sim c_L\,x^{-\alpha_L}\,,\qquad \xto\,,
-\eeao
-for non-negative constants $c_L,c_U$. We apply the Hill estimator (see Embrechts et al.~\cite{embrechts:kluppelberg:mikosch:1997}, p.~330,
-de Haan and Ferreira \cite{dehaan:ferreira:2006}, p.~69)
-to the \ts\ of the gains and losses in a naive way,
-neglecting the dependence and non-stationarity in the data; we also omit confidence bands.
-From the figure it is evident that the majority of the return series have tail indices below
-four, corresponding to an infinite fourth moment.
-\begin{figure}[htb!]
-\centering
-\includegraphics[scale=0.5]{SP500_tail_indices.pdf}
-\caption{Tail indices of log-returns of 478 \ts\ from the S\&P 500 index. The values  $(\hat \alpha_L,\hat \alpha_U)$ of the
-lower and upper tail indices are provided by
-Hill's estimator.  We also draw the line $\hat\alpha_U=\hat\alpha_L$.}
-\label{fig:SP500_tail_indices}
-\end{figure}
-The behavior of the largest eigenvalue $\la_{(1)}$
-changes dramatically when $\X$ has infinite fourth moment.
-Bai and Silverstein~\cite{baisilv} proved for an $n\times n$ matrix $\X$ with iid centered entries
-that
-\beam\label{eq:wdfrch3}
-\limsup_{\nto} \dfrac{\la_{(1)}}{n}=\infty \qquad {\rm a.s.}
-\eeam
-This is in stark contrast to Geman's result \eqref{eq:gemanch3}.
-%It is also unlikely that the
-%Tracy--Widom \ds\ yields a good approximation to the \ds\ of the largest eigenvalue of the sample covariance matrix;
-%see Figure~\ref{fig:MyDist-Frechet} for some simulation evidence.
-\par
-In the heavy-tailed case it is common to assume a {\em \regvar\ condition:}
-\beam\label{eq:regvarch3}
-\P(X>x)\sim p_+\,\dfrac{L(x)}{x^{\alpha}}\qquad\mbox{and}\qquad \P(X<-x) \sim p_-\,\dfrac{L(x)}{x^\alpha}\,,\qquad \xto\,,
-\eeam
-where $ p_\pm$ are non-negative constants \sth\ $p_++p_-=1$ and $L$ is a \slvary\ \fct . In particular, if $\alpha<4$ we have $\E [X^4]= \infty$. The \regvar\ condition
-on $X$ (we will also refer to $X$ as a \regvary\ \rv ) is needed for proving \asy\ theory for
-the eigenvalues of $\X\X'$. This is similar to proving limit theory for sums
-of iid \rv s with infinite variance stable limits; see for example Feller~\cite{feller}.
-\par
-
-In \eqref{eq:MPch3} we have seen that the sequence $(F_{n^{-1}\X \X'})$ of empirical spectral distributions converges to the Mar\v cenko--Pastur law if the centered iid entries possess a finite second moment.
-Now we will discuss the situation when the entries are still iid and centered, but have an infinite variance.
-Here we assume the entries to be regularly varying with index $\alpha \in (0,2)$.
-Assuming \eqref{eq:gammach3} with $\gamma\in (0,1]$ in this infinite variance case, Belinschi et al.~\cite[Theorem~1.10]{belinschi:dembo:guionnet:2009} showed that the sequence $(F_{a_{n+p}^{-2}\X \X'})$ converges with probability one to a non-random probability measure with density $\rho_{\alpha}^\gamma$ satisfying
-\begin{equation*}
-\rho_{\alpha}^\gamma(x) x^{1+\alpha/2} \to \frac{\alpha \gamma}{2(1+\gamma)}, \quad x \to \infty,
-\end{equation*}
-see also Ben Arous and Guionnet \cite[Theorem~1.6]{arous:guionnet:2008}.
-The normalization $(a_k)$ is chosen \sth\ $\P(|X|>a_k)\sim k^{-1}$ as $k\to\infty$.
-An application of the Potter bounds (see Bingham et al.
-\cite[p.~25]{bingham:goldie:teugels:1987}) shows that $a_{n+p}^2/n \to \infty$.
-
-It is interesting to note that there is a phase change in the extreme eigenvalues in going from finite to infinite fourth moment, while the phase change occurs for the empirical spectral distribution going from finite to infinite variance.
-%In the case of finite fourth moment, \eqref{eq:geman} and \eqref{eq:MP} imply that $\lambda_{(k)}/n$ has to converge to the right endpoint of the Mar\v cenko--Pastur law for any finite $k$.
-However, if we assume an infinite fourth moment but finite second moment the empirical spectral distributions  $(F_{{n}^{-1}\X \X'})$ still converge to the Mar\v cenko--Pastur law with the same support, while $\lambda_{(1)}/n$ will tend to infintity due to \eqref{eq:wdfr}.
-This means that the proportion of the normalized eigenvalues taking values larger than $(1+\sqrt{\gamma})^2$ has to be negligible; otherwise we would have a contradiction to the convergence of the empirical spectral distributions to the Mar\v cenko--Pastur law.
-
-%Finally, if the second moment is infinity, i.e.~regularly varying entries with index $\alpha \in (0,2)$, then the much larger values of the eigenvalues are visible in the limiting spectral distribution even when using a much stronger normalization \cite[Theorem~1.10]{belinschi:dembo:guionnet:2009}.
-\par
-
-The theory for the largest eigenvalues of sample covariance matrices with heavy tails is less developed than in the light-tailed case.
-Pioneering work for $\la_{(1)}$ in the case of iid \regvary\ entries $X_{it}$ with index $\alpha\in (0,2)$
-is due to Soshnikov~\cite{soshnikov:2004,soshnikov:2006}. He showed the \pp\ \con\
-\beam\label{eq:nn}
-N_n=\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i} \std N=\sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
-\eeam
-under the growth condition \eqref{eq:gammach3} on $(p_n)$.
-%The normalization $(a_k)$ is chosen \st\ $\P(|X|>a_k)\sim k^{-1}$ as $k\to\infty$.
-Here
-\beam\label{eq:Gamma}
-\Gamma_i=E_1+\cdots + E_i\,,\qquad i\ge 1\,,
-\eeam
-and $(E_i)$ is an iid standard exponential \seq . In other words, $N$ is a Poisson point process on $(0,\infty)$ with mean \ms\
-$\mu(x,\infty)= x^{-\alpha/2}$, $x>0$. Convergence in \ds\ of \pp es is understood in the sense of weak \con\
-in the space of point \ms s equipped with the vague topology; see Resnick \cite{resnick:2007,resnick:1987}.
-We can easily derive the limiting distribution of $a_{np}^{-2} \lambda_{(k)}$ for fixed $k\ge 1$ from \eqref{eq:nn}:
-\begin{equation*}
-\begin{split}
-\lim_{\nto}\P(a_{np}^{-2} \lambda_{(k)}\le x)&= \lim_{\nto}\P(N_n(x,\infty)<k)
-=  \P(N(x,\infty)<k) =\P(\Gamma_k^{-2/\alpha}\le x)\\ &= \sum_{s=0}^{k-1} \frac{\big(\mu(x,\infty)\big)^s}{s!} \e^{-\mu(x,\infty)}, \quad  x>0.
-\end{split}
-\end{equation*}
-In particular,
-\beao
-\dfrac{\la_{(1)}}{a_{np}^2}\std \Gamma_1^{-\alpha/2}\,,\qquad \nto\,,
-\eeao
-where the limit has {\em Fr\'echet \ds } with parameter $\alpha/2$ and distribution function
-\beao
-\Phi_{\alpha/2}(x) =\ex^{-x^{-\alpha/2}}\,,\qquad x>0\,.
-\eeao
-
-\noindent We mention that the tail balance condition \eqref{eq:regvarch3} may be replaced in this case by the weaker assumption
-$\P(|X|>x)= L(x) x^{-\alpha}$ for a \slvary\ \fct\ $L$. Indeed, it follows from the proof %(see also the comments below)
-that only the squares $X_{it}^2$ contribute to the \pp\ limits of the eigenvalues $(\la_i)$. A con\seq\ of the
-\cmt\ and \eqref{eq:nn} is the joint \con\ of the upper order statistics: for any $k\ge 1$,
-\beao
-a_{np}^{-2} \big(\la_{(1)},\ldots,\la_{(k)}\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,,\qquad \nto\,.
-\eeao
-
-It follows from standard theory
-for \pp es with iid points (e.g. Resnick \cite{resnick:2007,resnick:1987})
-that \eqref{eq:nn} remains valid if
-we replace  $N_n$  by the \pp\ $\sum_{i=1}^p\sum_{t=1}^n\vep_{X_{it}^2/a_{np}^2}$. Then we also have for any $k\ge 1$,
-\beam\label{eq:nna}
-a_{np}^{-2} \big(X_{(1),np}^2,\ldots,X_{(k),np}^2\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,,\qquad \nto\,,
-\eeam
-where
-\beao
-X_{(1),np}^2\ge \cdots \ge X_{(np),np}^2
-\eeao
-denote the order statistics of $(X_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$.
-\par
-Auffinger et al.~\cite{auffinger:arous:peche:2009} showed that \eqref{eq:nn}
-remains valid under the \regvar\ condition \eqref{eq:regvarch3} for  $\alpha\in (2,4)$, the growth condition
-\eqref{eq:gammach3} on $(p_n)$ and the
-additional assumption $\E [X]=0$. Of course, \eqref{eq:nna} remains valid as well.
-Davis et al.~\cite{davis:pfaffel:stelzer:2014} extended these results to the case when the rows of $\X$
-are iid linear processes with iid \regvary\ noise. The Poisson \pp\ \con\ result of \eqref{eq:nn} remains valid in this case.
-Different limit processes can only be expected if there is dependence across rows and columns.
-\par
-In what follows, we refer to the  {\em heavy-tailed case} when we assume the \regvar\ condition \eqref{eq:regvarch3}
-for some $\alpha\in (0,4)$.%{\blue This hypothesis is confirmed in}
-
-\subsection{Overview}
-
-The primary objective of this overview  is to make a connection between extreme value theory and the behavior of the largest eigenvalues of sample covariance matrices from heavy-tailed multivariate time series.
-For time series that are linearly dependent through time and across rows, it turns out that the extreme eigenvalues are essentially determined by the extreme order statistics from an array of iid random variables.
-The asymptotic behavior of the extreme eigenvalues is then derived routinely from classical extreme value theory.
-As such, explicit joint distributions of the extreme order statistics can be given which yield a plethora of ancillary results.
-%These include the limit behavior of ratios of consecutive extreme eigenvalues, ratios of extreme order statistics to the sample variance, etc..
-%%The underlying multivariate \ts\ will assumed to be heavy-tailed and linearly dependent through time and across the rows.We consider asymptotic theory for the largest eigenvalues of a sample covariance matrix,
-%and we will also address the largest singular values of the sample autocovariance matrix.
-%To our knowledge, the present paper is the first to consider bonafide dependence among the components in the time series which renders a multivariate analysis, such as PCA, meaningful.
-Convergence of the point process of extreme eigenvalues, properly normalized, plays a central role in establishing the main results.
-\par
-In Section~\ref{sec:2ch3} we continue the study of the case when the data matrix $\X$ consists of iid heavy-tailed entries.
-We will consider power-law growth rates on the dimension $(p_n)$ that is more general than prescribed by \eqref{eq:gammach3}.
-In Section~\ref{sec:model} we introduce a model for $X_{it}$ which allows for linear dependence across the rows and through time.
-The point process convergence of normalized eigenvalues is presented in Section~\ref{sec:mainresult}.
-This result lays the foundation for new insight into the spectral behavior of the sample covariance matrix, which is the content of Section~\ref{sec:samplecov}.
-
-Sections~\ref{sec:samplecov} and \ref{sec:possemidef} are devoted to {\em sample autocovariance matrices}.
-Motivated by \cite{lam:yao}, we study the eigenvalues of sums of transformed matrices and illustrate the results in two examples.
-These results are applied to the time series of S\&P 500 in Section \ref{sec:sp500}.
-%The technical proofs are collected in Section~\ref{sec:proof}. 
-Appendix~\ref{appendix:Ach3} contains useful facts about regular variation and point processes.
-
-\section[General growth rates for $p_n$]{General growth rates for $p_n$ in the iid heavy-tailed case}\label{sec:2ch3}
-This section is based on ideas in Heiny and Mikosch \cite{heiny:mikosch:2015:iid} where one can also find detailed proofs.
-\subsection*{Growth conditions on $(p_n)$}\label{subsec:pn}
-In many applications it is not realistic to assume
-that the dimension $p$ of the data and the sample size $n$ grow at the same rate.
-The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009} already indicate that the value $\gamma$ in the growth
-rate~\eqref{eq:gammach3} does not appear in the \ds al limits.
-This obervation is in contrast to the light-tailed case; see \eqref{eq:gemanch3}
-and \eqref{eq:tcch3}.  Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}
-and Heiny and Mikosch ~\cite{heiny:mikosch:2015:iid} allowed for more general rates for
-$p_n\to\infty$ than linear growth in $n$.
-Recall that $p=p_n\to\infty$ is the number of rows in the
-matrix $\X_n$. We need to specify the growth rate of $(p_n)$ to ensure
-a non-degenerate limit distribution of the normalized singular values of the sample autocovariance
-matrices. To be precise, we assume
-\begin{equation}\label{eq:pch3}
-p=p_n=n^\beta \ell(n), \quad n\ge1,\tag{$C_p(\beta)$}
-\end{equation}
-where $\ell$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume $\ell(n) \to \infty$.
-Condition~\ref{eq:pch3} is more general than the growth conditions in the literature; see
-\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
-\begin{theorem}\label{thm:intro}
-Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvarch3} for some
-$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Let $(p_n)$ be an integer sequence
-satisfying \ref{eq:pch3} with $\beta\ge 0$. In addition, we require
-\begin{equation}\label{Cbeta}
-\min(\beta,\beta^{-1})\in (\alpha/2-1,1] \qquad \mbox{ for } \alpha\in [2,4), \tag{$\widetilde{C}_\beta(\alpha)$}
-\end{equation}
-Then
-\beam\label{eq:2}
-\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
-\eeam
-where the \con\ holds in the space of point \ms s with state space $(0,\infty)$ equipped with the vague toplogy; see Resnick \cite{resnick:2007}.
-\end{theorem}
-
-\subsubsection*{A discussion of the case $\beta\in [0,1]$}
-We mentioned earlier that in the heavy-tailed case, limit theory for the
-largest eigenvalues of the sample covariance matrix is rather insensitive to
-the growth rate of $(p_n)$ and that the limits are essentially
-determined by the diagonal of this matrix. This is confirmed by the following result.
-
-\begin{proposition}\label{prop:offdiagonalch3}
-Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvarch3} for some
-$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Then for any \seq\ $(p_n)$
-satisfying \ref{eq:pch3} with $\beta\in [0,1]$ we have
-\beao
-a_{np}^{-2} \twonorm{\X \X' - \diag(\X \X')}\stp 0\,,\qquad\nto\,,
-\eeao
-where $\| \cdot\|_2$ denotes the spectral norm; see \eqref{specnorm} for its definition.
-\end{proposition}
-
-\par
-Proposition~\ref{prop:offdiagonalch3} is not unexpected for two reasons:
-\begin{itemize}
-\item
-It is well-known from classical theory (see Embrechts and Veraverbeke \cite{embrechts:veraverbeke:1982}) that for any iid \regvary\ non-negative \rv s
-$Y,Y'$ with index  $\alpha'>0$, $Y\,Y'$ is \regvary\ with index $\alpha'$ while $Y^2$ is \regvary\ with index $\alpha'/2$.
-Therefore $X^2$ and $X_{11}X_{12}$  are \regvary\ with indices $\alpha/2$ and $\alpha$, respectively.
-\item
-The aforementioned tail behavior is inherited by the entries of $\X\X'$ in the following sense.
-By virtue of Nagaev-type \ld\ results for an iid \regvary\ \seq\ $(Y_i)$ with index $\alpha'\in (0,2)$ where we also assume that
-$\E [Y_0]=0$ if $\E[|Y_0|]<\infty$
-(see Theorem~\ref{thm:nagaevch3}) %S.V. Nagaev \cite{nagaev:1979} and
-%Cline and Hsing \cite{cline:hsing:1998})
-we have that $\P(Y_1+\cdots +Y_n >b_n)/(n \,\P(|Y_0|>b_n))$ converges to a non-negative constant
-provided $b_n/a_n'\to\infty$, where $\P(|Y_0|>a_n')\sim n^{-1}$ as $\nto$. As a con\seq\ of the tail behaviors of $X_{it}^2$ and $X_{it}X_{jt}$ for $i\ne j$
-and Nagaev's results we have for $(b_n)$ \sth\ $b_n/a_n^2\to\infty$,
-\beam \label{eq:dfsdfjl}
-\dfrac{\P\big((\X\X')_{ij}> b_n\big)}{\P\big((\X\X')_{ii}- c_n> b_n\big)}\sim \dfrac{n\,\P( X_{11}X_{12}>b_n)}{n\,\P(X^2>b_n)}\to 0\,,\qquad \nto\,,
-\eeam
-where $c_n=0$ or $n\,\E[X^2]$ according as $\alpha\in (0,2)$ or $\alpha\in (2,4)$.
-This means that the diagonal and off-diagonal entries of
-$\X\X'$ inherit the tails of $X_{it}^2$ and $X_{it}X_{jt}$, $i\ne j$, respectively, above the high threshold $b_n$.
-\end{itemize}
-\par
-Proposition~\ref{prop:offdiagonalch3} has some immediate con\seq s for the approximation of the eigenvalues
-of $\X\X'$ by those of ${\rm diag}(\X\X')$. Indeed, let $C$ be a symmetric $p\times p$ matrix with
-eigenvalues $\la_1(C),\ldots,\la_p(C)$ and
-ordered eigenvalues
-\beam \label{eq:weyl}
-\la_{(1)}(C)\ge \cdots \ge \la_{(p)}(C)\,.
-\eeam
-Then for any symmetric $p\times p$ matrices $A,B$, by {\em Weyl's inequality} (see Bhatia \cite{bhatia:1997}),
-\beao
-\max_{i=1,\ldots,p}\big|\la_{(i)}(A+B)-\la_{(i)}(A)\big|\le \|B\|_2\,.
-\eeao
-If we now choose
-$A+B=\X\X'$ and $A= \diag (\X\X')$ we obtain the following result.
-\begin{corollary}\label{cor:687ch3}
-Under the conditions of Proposition~\ref{prop:offdiagonalch3},
-\beao
-a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\diag(\X\X'))\big|\stp 0\,,\quad\nto \,.
-\eeao
-\end{corollary}
-Thus the problem of deriving limit theory for the order statistics of $\X\X'$ has been reduced to limit theory for the order
-statistics of the iid row-sums
-\beao
-D_i^{\rightarrow}= (\X\X')_{ii}=\sum_{t=1}^n X_{it}^2\,,\qquad i=1,\ldots,p\,,
-\eeao
-which are the eigenvalues of $\diag(\X\X')$. This theory is completely described by the \pp es constructed from the points $D_i^\rightarrow/a_{np}^2$ $i=1,\ldots,p$. Necessary
-and sufficient conditions for the weak \con\ of these \pp es are provided by  Lemma~\ref{lem:pprch3}
-which in combination with the Nagaev-type \ld\ results of  Theorem~\ref{thm:nagaevch3}
-yield the following result; see also Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
-\begin{lemma}\label{lem:ppch3}
-Assume the conditions of Proposition~\ref{prop:offdiagonalch3} hold. Then
-\beao
-\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i^\rightarrow-c_n)}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
-\eeao
-where $(\Gamma_i)$ is defined in \eqref{eq:Gamma} and $c_n=0$ if $\E[D^\rightarrow]=\infty$ and $c_n=\E[D^\rightarrow]$ otherwise.
-\end{lemma}
-In this result, centering is only needed for $\alpha\in [2,4)$ when $n/a_{np}^2\not\to 0$. Under the additional condition \ref{Cbeta},
-%\begin{equation}\label{Cbetaor}
-%\beta\in (\alpha/2-1,1], \quad \mbox{ for } \alpha\in [2,4),
-%\end{equation}
-$n/a_{np}^2\to 0$ in view of the Potter bounds; see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}.
-Combining Lemma~\ref{lem:ppch3} and Corollary~\ref{cor:687ch3}, we conclude that Theorem~\ref{thm:intro} holds for $\beta\in [0,1]$.
-
-%In  Davis et al. \cite{davis:mikosch:pfaffel;2015} relation \eqref{eq:2} was derived for $\alpha\in (0,1)$ even for \seq s
-%$p_n\to\infty$ \st\ $p_n= O(\ex^{s_n/n})$, where $s_n=o(n)$ but with stronger limitations on the growth of $(p_n)$ for $[1,4]$.
-%In the case $\alpha\in (2,4]$ they proved \eqref{eq:center} with $(\la_i)$ replaced by $(\la_i(\X\X'-n\E[ \X\X']))$. It is surprising that
-%\eqref{eq:center} also holds since, in general, $\la_i(\X\X'-\E [\X\X'])\not = \la_i-n\,\E X$.
-
-\subsubsection*{Extension to general $\beta$}%\label{subsec:generalb}
-Next we explain that it suffices to consider only the case $\beta\in [0,1]$ and how to proceed when $\beta>1$.
-The main reason is that
-the $p \times p$ sample covariance matrix
-$\X\X'$ and the $n \times n$ matrix $\X'\X$  have the same rank and their non-zero eigenvalues coincide; see Bhatia \cite[p.~64]{bhatia:1997}. When proving limit
-theory for the eigenvalues  of the sample covariance matrix one may switch to $\X'\X$ and vice versa,
-hereby interchanging the roles of $p$ and $n$.
-By switching to $\X'\X$, one basically replaces $\beta$ by $\beta^{-1}$.
-Since $\min(\beta,\beta^{-1})\in [0,1]$ for any $ \beta \ge 0$, one can assume without loss of generality that $\beta\in [0,1]$.
-This trick allows one to extend results for
-$(p_n)$ satisfying \ref{eq:pch3} with $\beta\in [0,1]$  to $\beta>1$.
-We illustrate this approach by providing the direct analogs of Proposition~\ref{prop:offdiagonalch3} and Corollary~\ref{cor:687ch3}.
-
-\begin{proposition}\label{prop:offdiagonal1}
-Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvarch3} for some
-$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Then for any \seq\ $(p_n)$
-satisfying \ref{eq:pch3} with $\beta>1$ we have
-\beao
-a_{np}^{-2} \twonorm{\X' \X - \diag(\X' \X)}\stp 0\,,\qquad\nto\,,
-\eeao
-where $\| \cdot\|_2$ denotes the spectral norm.
-\end{proposition}
-Note that for $\beta>1$ we have $\lim_{\nto} p/n= \infty$. This means that
-$\X'\X$ has  asymptotically a much smaller dimension than $\X\X'$ and therefore it is more convenient to work with $\X' \X$
-when bounding the spectral norm.
-\begin{corollary}\label{cor:6871}
-Under the conditions of Proposition~\ref{prop:offdiagonal1},
-\beao
-a_{np}^{-2}\,\max_{i=1,\ldots,n}\big|\la_{(i)}-\la_{(i)}(\diag(\X'\X))\big|\stp 0\,,\quad\nto \,.
-\eeao
-\end{corollary}
-Now, Theorem~\ref{thm:intro} for $\beta > 1$ is a consequence of Corollary~\ref{cor:6871}.
-\par
-
-%-----------------------------------------------------------------------
-\section{Introducing dependence between the rows and columns}\label{sec:model}
-For details on the results of this section, we refer to Davis et al.~\cite{davis:mikosch:pfaffel:2015}, Heiny and Mikosch \cite{heiny:mikosch:2015:iid} and Heiny et al.~\cite{heiny:mikosch:2016:noniid}.
-
-\subsection{The model}
-When dealing with covariance matrices of a multivariate \ts\ $(\X_n)$ it is rather natural to assume dependence
-between the entries $X_{it}$.
-In this section we introduce a model which allows for {\em linear dependence}
-between the rows and columns of $\X$:
-\begin{equation}\label{eq:1}
-X_{it}=\sum_{l\in \Z}\sum_{k\in \Z} h_{kl} Z_{i-k,t-l}\,,\qquad i,t\in\Z\,,
-\end{equation}
-where $(Z_{it})_{i,t\in \Z}$ is a field of iid \rv s and $(h_{kl})_{k,l\in\Z}$ is an array of real numbers.
-Of course, linear dependence is restrictive in some sense. However, the particular dependence structure allows one to
-determine those ingredients in the sample covariance matrix which contribute to its largest eigenvalues.
-If the series in \eqref{eq:1} converges a.s. $(X_{it})$ constitutes a strictly stationary random field.
-We denote generic elements of the $Z$- and $X$-fields by $Z$ and $X$, respectively. We assume that $Z$ is \regvary\ in the sense that
-\begin{equation}\label{eq:27}
-\P(Z>x)\sim p_+ \dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad  \P(Z\le -x)\sim p_-
-\dfrac{L(x)}{x^{\alpha}}\,,\qquad \xto\,,
-\end{equation}
-for some tail index $\alpha>0$, constants $p_+,p_-\ge 0$ with $p_++p_-=1$ and a \slvary\ $L$. We will assume $\E[ Z]=0$ whenever $\E [Z^2]<\infty$.
-Moreover, we require the summability condition
-\begin{equation}\label{eq:2a}
-\sum_{l \in \Z} \sum_{k\in \Z} |h_{kl}|^{\delta} <\infty
-\end{equation}
-for some $\delta\in (0,\min({\alpha/2},1))$ which ensures
-the a.s.~absolute convergence of the series in \eqref{eq:1}. Under the conditions \eqref{eq:27} and \eqref{eq:2a}, the marginal and
-\fidi s of the field $(X_{it})$ are \regvary\ with index $\alpha$; see
-Embrechts et al. \cite{embrechts:kluppelberg:mikosch:1997}, Appendix A3.3. Therefore we also refer to $(X_{it})$ and $(Z_{it})$
-as \regvary\ fields.
-\par
-The model \eqref{eq:1} was introduced by Davis et al. \cite{davis:pfaffel:stelzer:2014}, assuming the rows iid, and in the
-present form by Davis et al. \cite{davis:mikosch:pfaffel:2015}.
-\subsection{Sample covariance and autocovariance matrices}
-From the field $(X_{it})$ we construct the $p\times n$ matrices
-\begin{equation}\label{eq:26}
-\X_n(s)= (X_{i,t+s})_{i=1,\ldots,p;t=1,\ldots,n}\,,\qquad s=0,1,2,\ldots\,,
-\end{equation}
-As before, we will write $\X=\X_n(0)$.
-Now we can introduce the (non-normalized)
-{\em sample autocovariance matrices}
-\beam\label{eq:sample}
-\X_n(0)\X_n(s)'\,,\qquad s=0,1,2,\ldots\,.
-\eeam
-We will refer to $s$ as the {\em lag}. For $s=0$, we obtain the {\em sample covariance matrix.}
-In what follows, we will be interested in the asymptotic behavior (of \fct s) of the eigen- and singular values of the
-sample covariance and autocovariance matrices in the heavy-tailed case. Recall that the {\em singular values} of a matrix $A$ are the square roots of the
-eigenvalues of the non-negative definite matrix $AA'$ and its {\em spectral norm} $\twonorm{A}$ is its largest singular value. We notice that $\X_n(0)\X_n(s)'$ is not symmetric
-and therefore its eigenvalues can be complex. To avoid this situation, we use
-%\beao
-%\X_n(0)\X_n(s)' +\X_n(s)\X_n(0)'
-%\eeao
-%(this matrix is symmetric and therefore it has real eigenvalues) or use
-the squares
-\beam\label{eq:squares}
-\X_n(0)\X_n(s)' \X_n(s)\X_n(0)'
-\eeam
-%(this matrix is  non-negative definite  and therefore it has non-negative eigenvalues).
-whose eigenvalues are the squares of the singular values of $\X_n(0)\X_n(s)'$.
-The idea of using the sample autocovariance matrices and \fct s of their squares \eqref{eq:squares}
-originates from a paper by Lam and Yao~\cite{lam:yao} who used a model different from \eqref{eq:1}.
-This idea is quite natural in the context of \tsa .
-\par
-In Theorem~\ref{thm:mains} below, we provide a general approximation result for the ordered singular values of the sample
-autocovariance matrices in the heavy-tailed case. This result is rather technical. To formulate it we introduce further notation.
-As before, $p=p_n$ is any integer \seq\ converging to infinity.
-\subsection{More notation}
-Important roles are played by the quantities $(Z_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$ and their order statistics
- \begin{equation}\label{eq:zorderch3}
-Z_{(1),np}^2 \ge Z_{(2),np}^2 \ge  \ldots \ge Z_{(np),np}^2, \qquad n,p\ge 1\,.
-\end{equation}
-As important are the row-sums
-\begin{equation}\label{eq:llch3}
-D_i^\rightarrow=D_i^{(n),\rightarrow}=\sum_{t=1}^n Z_{it}^2\,, \qquad
-i=1,\ldots,p;\quad  n=1,2,\ldots\,,
-\end{equation} with generic element $D^\rightarrow$ and their ordered values
-\beam\label{eq:help6ch3}
-D_{(1)}^\rightarrow=D_{L_1}^\rightarrow\ge \cdots \ge D_{(p)}^\rightarrow=D_{L_p}^\rightarrow\,,
-\eeam
-where we assume without loss of generality that $(L_1,\ldots,L_p)$ is
-a permutation of $(1,\ldots,p)$ for fixed~$n$.
-\par
-Finally, we introduce the column-sums
-\beao
-D_t^\downarrow=D_t^{(n),\downarrow}= \sum_{i=1}^p Z_{it}^2\,,\qquad t=1,\ldots,n;\; \quad p=1,2,\ldots\,,
-\eeao
-with generic element $D^\downarrow$ and we also adapt the notation from \eqref{eq:help6ch3} to these quantities.
-\subsubsection*{Matrix norms} For any $p\times n $ matrix $\A=(a_{ij})$, we will use the following norms:
-\begin {itemize}
-\item
-{\em Spectral norm:}
-\beam\label{specnorm}
-\|\A\|_2=\sqrt{\la_{(1)}(\A\A')}\,,
-\eeam
-\item
-{\em Frobenius norm:}
-\beao
-\frobnorm{\A}= \Big( \sum_{i=1}^p \sum_{j=1}^n |a_{ij}|^2\Big)^{1/2}\,.
-\eeao
-%\item
-%{\em Max-row sum norm:}
-%\beao
-%\inftynorm{A}= \max_{i=1,\ldots,p} \sum_{j=1}^n |a_{ij}|\,.
-%\eeao
-%\item
-%{\em Max-column sum norm:}
-%\beao
-%\|{A}\|_1 = \max_{j=1,\ldots,n} \sum_{i=1}^p |a_{ij}|\,.
-%\eeao
-\end{itemize}
-We will frequently make use of the bound $\|\A\|_2\le \|\A\|_ F$. Standard references for matrix norms are \cite{belitski,bhatia:1997, horn, shores}.
-\subsubsection*{Singular values of the sample autocovariance matrices} Fix integers $n\ge 1$ and $s\ge 0$. We recycle the $\la$-notation
-for the singular values
-$\la_1(s), \ldots, \la_p(s)$ of the sample autocovariance matrix $\X_n(0)\X_n(s)'$, suppressing the dependence on $n$.
-Correspondingly, the  order statistics are denoted by
-\begin{equation}\label{eq:sigma}
-\la_{(1)}(s) \ge \cdots \ge \la_{(p)}(s)\,.
-\end{equation}
-When $s=0$ we typically write $\la_i$ instead of $\la_i(0)$.
-\subsubsection*{The matrix $\M(s)$} We introduce some auxiliary matrices
-derived from the coefficient matrix $\H=(h_{kl})_{k,l\in \Z}$:
-\beao
-\H(s)=(h_{k,l+s})_{k,l\in \Z}, \qquad \M(s)= \H(0)\H(s)'\qquad s\ge 0\,.
-\eeao
-Notice that
-\begin{equation}\label{eq:m}
-(\M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z .
-\end{equation}
-We denote the ordered singular values of $\M(s)$ by
-\begin{equation}\label{eq:v1}
-v_1(s) \ge v_2(s) \ge\cdots \,.
-\end{equation}
-Let $r(s)$ be the rank of $\M(s)$ so that $v_{r(s)}(s)>0$ while $v_{r(s)+1}(s)=0$ if $r(s)$ is
-finite, otherwise $v_i(s)>0$ for all $i$. We also write $r=r(0)$.
-\par
-Under the summability condition \eqref{eq:2a} on $(h_{kl})$ for fixed $s\ge 0$,
-\beam\label{eq:tracea}
-\sum_{i=1}^\infty (v_i(s))^2 &=&  \frobnorm{\M(s)}^2= \sum_{i,j\in \Z} \sum_{l_1,l_2 \in \Z} h_{i,l_1} h_{j,l_1+s}h_{i,l_2}
-h_{j,l_2+s}\nonumber\\
-&\le& c \,\Big(\sum_{l_1,l_2 \in \Z} \sum_{i\in \Z} |h_{i,l_1} h_{i,l_2}|\Big)^2
-\le c \,\sum_{l_1\in \Z} \sum_{i \in \Z} |h_{i,l_1}| <\infty\,.
-\eeam
-Therefore all singular values $v_i(s)$ are finite and the ordering
-  \eqref{eq:v1}
-is justified.
-\par
-{\em Here and in what follows, we write $c$ for any constant whose value is not of interest.}
-
-\subsubsection*{Normalizing sequence}
-We define $(a_k)$ by
-\beao
-\P(|Z|>a_k)\sim k^{-1}\,,\qquad k\to\infty\,,
-\eeao
-and choose the normalizing \seq\ for the singular values as $(a_{np}^2)$ for suitable \seq s $p=p_n\to\infty$.
-\subsubsection*{Approximations to singular values}\label{subsec:defdelta}
-We will give approximations to the singular values $\la_i(s)$  in
-terms of the  $p$ largest ordered values for $s\ge 0$,
-\beao
-&&\delta_{(1)}(s)\ge \cdots \ge \delta_{(p)}(s)\,,\\
-&&\gamma_{(1)}^\rightarrow(s)\ge \cdots \ge \gamma_{(p)}^\rightarrow(s)\,,\\
-&&\gamma_{(1)}^\downarrow(s)\ge \cdots \ge \gamma_{(n)}^\downarrow(s)\,,
-\eeao
-from the sets
-\beao
-&&\big\{Z_{(i),np}^2 v_j(s)\,, i=1,\ldots,p\,;j=1,2,\ldots\big\}\,,\notag\\
-&&\big\{D_{i}^\rightarrow v_j(s), i=1,\ldots,p\,;j=1,2,\ldots\big\}\,,\notag\\
-&&\big\{D_{t}^\downarrow v_j(s), t=1,\ldots,n\,;j=1,2,\ldots\big\}\,,
-\eeao
-respectively.
-%---------------------------------------------------------------------------
-\subsection{Approximation of the singular values}\label{sec:mainresult}
-%Furthermore, let $\tau_{(1)}(s)  \ge \ldots \ge \tau_{(p)}(s)$ be  the singular values of $\tfrac{1}{2}(\X_n(0)\X_n(s)'+\X_n(s)\X_n(0)' )$ and let $\tilde{\tau}_{(1)}(s) \ge  \ldots \ge\tilde{\tau}_{(p)}(s)$ be the singular values of
-%\begin{equation*}
-%\tfrac{1}{2}(\X_n(0)\X_n(s)'-\E[\X_n(0)\X_n(s)']+\X_n(s)\X_n(0)' -\E[\X_n(s)\X_n(0)'])
-%\end{equation*}
-In the following result we povide some useful approximations to the singular values of the sample autocovariance matrices of
-the linear model \eqref{eq:1}.
-\begin{theorem}\label{thm:mains}
-Consider the linear process \eqref{eq:1} under
-\begin{itemize}
-\item
-the \regvar\ condition \eqref{eq:27}
-for some $\alpha\in (0,4)$,
-\item the centering condition
-$\E[Z]=0$ if $\E[|Z|]<\infty$,
-\item
-the summability condition
-\eqref{eq:2a} on the coefficient matrix  $(h_{kl})$,
-\item
-the growth condition \ref{eq:pch3} on $(p_n)$ for some $\beta\ge 0$.
-\end{itemize}
-Then the following statements hold for $s\ge 0$:
-\begin{enumerate}
-\item We consider two disjoint cases:
-$\alpha \in (0,2)$ and $\beta\in (0,\infty)$, or
-$\alpha\in [2,4)$ and $\beta$ satisfying \ref{Cbeta}. Then
-\begin{equation}\label{eq:mains1}
-a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\delta_{(i)}(s)| \cip 0, \quad \nto.
-\end{equation}
-\item
-Assume $\beta\in [0,1]$.
-If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta\in (\alpha/2-1,1]$ then
-\beao
-a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\gamma_{(i)}^\rightarrow(s)| \cip 0, \quad \nto.
-\eeao
-Assume $\beta>1$. If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta^{-1}\in(\alpha/2-1,1]$. Then
-\beao
-a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\gamma_{(i)}^\downarrow(s)| \cip 0, \quad \nto.
-\eeao
-\end{enumerate}
-\end{theorem}
-\begin{remark}\em
-The proof of Theorem~\ref{thm:mains} is given in Heiny et al.~\cite{heiny:mikosch:2016:noniid}. Part (2) of this  result
-with more restrictive conditions on the growth rate of $(p_n)$ is contained in Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
-These proofs are very technical and lengthy.
-\end{remark}
-\begin{remark}\em
-If we consider a random array $(h_{kl})$ independent of  $(X_{it})$ and
-assume that the summability
-condition \eqref{eq:2a} holds a.s.
-then Theorem~\ref{thm:mains} remains valid conditionally on
-$(h_{kl})$, hence  unconditionally in $\P$-probability; see also \cite{davis:mikosch:pfaffel:2015}.\end{remark}
-\subsection{Point process \con }\label{sec:lkasfj}
-Theorem~\ref{thm:mains}  and arguments similar to the proofs in Davis et al. \cite{davis:mikosch:pfaffel:2015}
-enable one to derive the weak \con\ of
-the point processes of the normalized singular values. Recall
-the representation of the points $(\Gamma_i)$ of a unit rate homogeneous Poisson process on $(0,\infty)$
-given in \eqref{eq:Gamma}. For $s\ge 0$, we define the point processes of the normalized singular values:
-\begin{equation}\label{eq:ppdef}
-N_n^{\lambda,s}=\sum_{i=1}^p \vep_{a_{np}^{-2}(\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s))} \,.
-\end{equation}
-
-\begin{theorem}\label{cor:1ch3}
-Assume the conditions of Theorem~\ref{thm:mains}.
-Then $(N_n^{\lambda,s})$ converge weakly in the space of point measures
-with state space $(0,\infty)^{s+1}$ equipped with the vague topology.
-If either $\alpha \in (0,2]$, $\E[Z^2]=\infty$ and $\beta \ge 0$,
-or $\alpha \in [2,4)$, $\E[Z^2]<\infty$ and \ref{Cbeta} hold then
-\begin{equation}\label{eq:pp}
-N_n^{\lambda,s} \cid N= \sum_{i=1}^\infty
-\sum_{j=1}^{\infty} \vep_{\Gamma_i^{-2/\alpha} (v_j(0),\ldots,v_j(s))}, \qquad \nto.
-\end{equation}
-\end{theorem}
-
-\begin{proof}
-Regular variation of $Z^2$ is equivalent to
-\begin{equation}\label{eq:wwa}
-n\,p\, \P(a_{np}^{-2} Z^2  \in \cdot ) \civ \mu(\cdot),
-\end{equation}
-where $\civ$ denotes vague convergence of Radon measures on  $(0,\infty)$ and the \ms\ $\mu$ is given by $\mu(x,\infty)=x^{-\alpha/2}$, $x>0$.
-In view of Resnick \cite{resnick:1987}, Proposition~3.21, \eqref{eq:wwa} is equivalent
-to the weak \con\ of the following \pp es:
-\begin{equation*}
-\sum_{i=1}^p\sum_{t=1}^n \vep_{a_{np}^{-2} Z_{it}^2}=\sum_{i=1}^{np} \vep_{a_{np}^{-2} Z^2_{(i),np}}\cid
-\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha} }=\widetilde{N}\,,\quad
-\nto\,,
-\end{equation*}
-where the limit $\widetilde{N}$ is a Poisson random measure (PRM) with state space $(0,\infty)$ and
-mean measure~$\mu$.
-\par
-Since $a_{np}^{-2} Z^2_{(p),np}\cip 0$ as $\nto$,
-the point processes $\sum_{i=1}^{p} \vep_{a_{np}^{-2} Z^2_{(i),np}}$ converge weakly to the same PRM:
-\begin{equation}\label{eq:ppc}
-\sum_{i=1}^{p} \vep_{a_{np}^{-2} Z^2_{(i),np}}\cid
-\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha} }\,,\quad
-\nto\,.
-\end{equation}
-A continuous mapping argument together with the fact that $\sum_{i=1}^\infty (v_i(s))^2<\infty$ (see  \eqref{eq:tracea})
-shows that
-\begin{equation*}
-\sum_{j=1}^{ \infty}\sum_{i=1}^p\vep_{a_{np}^{-2} Z^2_{(i),np}(v_j(0),\ldots,v_j(s))}\cid
-\sum_{j=1}^{ \infty}\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha}  (v_j(0),\ldots,v_j(s))}\,.
-\end{equation*}
-If the assumptions of part (1) of Theorem~\ref{thm:mains} are satisfied
-an application of \eqref{eq:mains1} (also recalling the definition of $(\delta_{(i)}(s))$) shows that \eqref{eq:ppc} remains
-valid with the points  $(a_{np}^{-2} Z^2_{(i),np}(v_j(0),\ldots,v_j(s)))$ replaced by
-$(a_{np}^{-2} (\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s))$.
-\par
-The only cases which are not covered by Theorem~\ref{thm:mains}(1) are $\alpha\in (0,2)$, $\beta=0$ and $\alpha=2$, $\E[Z^2]=\infty$, $\beta\ge0$.
-In these cases we get from Theorem~\ref{thm:nagaevch3} that
-\begin{equation*}
- p \, \P(a_{np}^{-2} D^\rightarrow>x ) \sim p\,n\, \P( Z^2>a_{np}^{2} x )\to \mu(x,\infty)\,,\quad x>0\,,
-\end{equation*}
-i.e., $
- p \, \P(a_{np}^{-2} D^\rightarrow\in \cdot ) \stackrel{v}{\rightarrow} \mu(\cdot)$.
-It follows from Lemma~\ref{lem:pprch3} that
-$\sum_{i=1}^p\vep_{a_{np}^{-2} D_i^\rightarrow}\std \widetilde{N}$. As before,
-a continuous mapping argument in combination with
-the approximation obtained in Theorem~\ref{thm:mains}(2)
-justifies the replacement of the points $(a_{np}^{-2} D_{(i)}^\rightarrow(v_j(0),\ldots,v_j(s)))$ by
-$(a_{np}^{-2} (\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s)))$ in the case $\beta\in [0,1]$. If $\beta>1$ one has to work with the
-quantities $(D_i^\downarrow)_{i=1,\ldots,n}$ instead of  $(D_i^\rightarrow)_{i=1,\ldots,p}$ and one may follow the same argument as above.
-This finishes the proof.
-\end{proof}
-
-\section{Some applications}
-\subsection{Sample covariance matrices}\label{sec:samplecov}%\setcounter{equation}{0}
-The sample covariance matrix $\X_n(0)\X_n(0)'=\X\X'$ is a non-negative definite matrix. Therefore
-its eigenvalues and singular values coincide.  Moreover, $v_j=v_j(0)$, $j\ge 1$, are the eigenvalues of $\M=\M(0)$.
-\par
-Theorem~\ref{thm:mains}(1) yields an approximation of the ordered eigenvalues $(\la_{(i)})$ of $\X\X'$ by
-the quantities $(\delta_{(i)})$ which are derived from the order statistics of $(Z_{it}^2)$.
-Part (2) of this result provides an approximation of $(\la_{(i)})$ by the quantities
-$(\gamma_{(i)}^{\rightarrow/\downarrow})$ which are derived from the order statistics
-of the partial sums $(D_i^{\rightarrow/\downarrow})$.
-\par
-In the following example we illustrate the
-quality of the two approximations.
-\begin{example}\label{ex:xiao}\em
-We choose a Pareto-type \ds\ for $Z$ with density
-\beam\label{eq:distrsimch3}
-f_Z(x) =
-\left\{\begin{array}{cc}
- \frac{\alpha}{(4|x|)^{\alpha + 1}}\,, & \mbox{if } |x| > 1/4 \\
-1\,, & \mbox{otherwise.}
-\end{array}\right.
-\eeam
-We simulated $20,000$ matrices $\X_n$ for $n=1,000$ and $p=200$ whose
-iid entries have this density. We assume $\beta=1$.
-Note that $\M=\M(0)$ has rank one and $v_1=1$.
-The estimated densities of the deviations $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ and
-$a_{np}^{-2}(\lambda_{(1)}-Z^2_{(1),np})$  based on the simulations are shown in
-Figure~\ref{fig:lambda_comparisonch3}. The approximation error is very small indeed.
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.40]{lambda_comparison.pdf}
-  \caption{Density of the approximation errors for the eigenvalues of $a_{np}^{-2}\X\X'$.
-The entries of $\X$ are iid with density \eqref{eq:distrsimch3} and $\alpha=1.6$.}
-  \label{fig:lambda_comparisonch3}
-\end{figure}
-According to the theory,
-\beao
-a_{np}^{-2}\sup_i|D_{(i)}^\rightarrow-\lambda_{(i)}| +a_{np}^{-2}\sup_i |Z^2_{(i),np}-\lambda_{(i)}|\stp 0\,,
-\eeao
-but  for finite $n$ the $(D_{(i)}^\rightarrow)$ sequence yields a better approximation to $(\la_{(i)})$. By construction,
-the considered differences
-have a tendency to be positive but Figure~\ref{fig:lambda_comparisonch3} also shows that the median of the
-approximation error for $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ is almost zero.
-\end{example}
-\par
-Theorem~\ref{cor:1ch3} and the continuous mapping theorem immediately yield results about
-the joint convergence of the largest eigenvalues of the matrices $a_{np}^{-2}\X_n\X_n'$ for $\alpha\in (0,2)$ and
-$\alpha\in (2,4)$ when $\beta$ satisfies \ref{Cbeta}.
-For fixed $k\ge 1$ one gets
-\begin{equation*}
-a_{np}^{-2} \big(\la_{(1)},\ldots,\la_{(k)}\big)\cid
-\big(d_{(1)},\ldots, d_{(k)}\big)\,,
-\end{equation*}
-where $d_{(1)}\ge \cdots\ge d_{(k)}$ are the $k$ largest ordered
-values of the set $\{\Gamma_i^{-2/\alpha} v_j, i=1,2,\ldots,j=1,\ldots,r\}$.
-The continuous mapping theorem yields for $k\ge 1$,
-\begin{equation}\label{eq:u}
-\dfrac{\la_{(1)}}{\la_{(1)}+\cdots+ \la_{(k)}}\cid
-\dfrac{d_{(1)}}{d_{(1)}+\cdots+ d_{(k)}}\,,\quad \nto\,.
-\end{equation}
-
-An application of the continuous mapping theorem to  the distributional convergence of the point processes in Theorem~\ref{cor:1ch3} in the spirit of Resnick
-\cite{resnick:2007}, Theorem 7.1, also yields the following result; see Davis et al. \cite{davis:mikosch:pfaffel:2015} for a proof and a similar result in the case $\alpha \in (2,4)$.
-\begin{corollary}\label{cor:1q} Assume the conditions of Theorem~\ref{thm:mains}.
-If $\alpha \in (0,2]$ and $\E[Z^2]=\infty$, then
-\begin{equation*}
-a_{np}^{-2}\Big(\la_{(1)},\sum_{i=1}^p \la_i\Big) \cid
-\Big( v_1\,\Gamma_1^{-2/\alpha}\,,\sum_{j=1}^r v_j\,\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}\Big)\,,
-\end{equation*}
-where $\Gamma_1^{-2/\alpha}$ is Fr\'echet $\Phi_{\alpha/2}$-distributed.
-and $\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}$ has the distribution
-of a positive $\alpha/2$-stable random variable.
-In particular,
-\begin{equation}\label{eq:limit}
-\dfrac{\la_{(1)}}{\la_{1}+\cdots+ \la_{p}}\cid
-\dfrac{v_1}{\sum_{j=1}^r v_j}\;
-\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty\Gamma_i^{-2/\alpha}}\,,\quad
-\nto\,.
-\end{equation}
-\end{corollary}
-
-\begin{remark}\label{rem:4.5}\em
-The ratio
-\begin{equation*}
-\dfrac{\la_{(1)}+\cdots +\la_{(k)}}{\la_1+\cdots+\la_p}, \qquad k\ge 1\,,
-\end{equation*}
-plays an important role in PCA. It reflects the proportion of the total variance in the data that we can explain by the first $k$ principal components.
-It follows from Corollary~\ref{cor:1q} that for fixed $k\ge 1$,
-\begin{equation*}
-\dfrac{\la_{(1)}+\cdots \la_{(k)}}{\la_1+\cdots+\la_p}\cid
-\dfrac{d_{(1)}+\cdots +d_{(k)}} {d_{(1)}+d_{(2)}+\cdots}\,.
-\end{equation*}
-Unfortunately, the limiting variable does in general not have a clean
-form.
-An exception is the case when $r=1$; see Example~\ref{exam:separable}.
-Also notice that the trace of $\X\X'$ coincides with $\la_1+\cdots+\la_p$.
-\end{remark}
-
-To illustrate the theory we consider a simple moving average example
-taken from Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
-\begin{example} \rm
-\begin{figure}[htb!]
-  \centering
-  \subfigure[iid data]{
-    \includegraphics[scale=0.34]{lambda_1_2_Gaps_iid.pdf}
-  }
-  \subfigure[data from model \eqref{ex:4.6}]{
-    \includegraphics[scale=0.34]{lambda_1_2_Gaps.pdf}
-  }
-  \caption{Distribution function of $(\lambda_{(1)} - \lambda_{(2)})/\lambda_{(1)}$ for iid data (left) and
-    data generated from the model \eqref{ex:4.6} (right). In each graph we compare the empirical \ds\ \fct\
-(dotted line, based on
-1000 simulations of $200 \times 1000$ matrices with $Z$-\ds\ \eqref{eq:distrsimch3}) with the theoretical curve (solid line).}
-  \label{fig:ProbMass}
-\end{figure}
-Assume that $\alpha\in (0,2)$ and
-\begin{equation}\label{ex:4.6}
-X_{it}= Z_{it}+ Z_{i,t-1}-2 (Z_{i-1,t}- Z_{i-1,t-1})\,,\quad i,t\in\Z\,.
-\end{equation}
-In this case, the non-zero entries of $\H$ are
-\begin{equation*}
-h_{00}=1, h_{01}=1,h_{10}=-2 \quad \mbox{ and }\quad h_{11}=2.
-\end{equation*}
-Hence $\M=\H\H'$ has the positive eigenvalues
-$v_1=8$ and $v_2=2$. The limit point process in \eqref{eq:pp} is
-\begin{equation*}
-N=\sum_{i=1}^\infty \vep_{8\Gamma_i^{-2/\alpha}}+ \sum_{i=1}^\infty \vep_{2\Gamma_i^{-2/\alpha}}\,,
-\end{equation*}
-so that
-\begin{equation*}
-a_{np}^{-2}\,\big(\la_{(1)},\la_{(2)}\big) \cid
-\big(8\Gamma_{1}^{-2/\alpha}, 2\Gamma_1^{-2/\alpha}\vee 8
-\Gamma_2^{-2/\alpha}\big)\,.
-\end{equation*}
-Using the fact that $U=\Gamma_1/\Gamma_2$ has a uniform \ds\ on $(0,1)$ we calculate
-\begin{equation*}
-\P(2\Gamma_1^{-2/\alpha}>8\Gamma_2^{-2/\alpha})= \P(\Gamma_1/\Gamma_2<2^{-\alpha})= 2^{-\alpha} \in (1/4,1).
-\end{equation*}
-In particular, we have for the normalized spectral gap
-\begin{equation*}
-a_{np}^{-2} \big(\la_{(1)}-\la_{(2)}\big)\cid
-6 \,\Gamma_1^{-2/\alpha} \1_{\{\Gamma_1 4^{\alpha/2 }<\Gamma_2\}}+
-8\,
-\big(\Gamma_1^{-2/\alpha}-\Gamma_2^{-2/\alpha}\big)\1_{\{\Gamma_14^{\alpha/2
-  }> \Gamma_2\}}
-\end{equation*}
-and for the self-normalized spectral gap (see also Example~\ref{exam:spectralgap} for a detailed analysis)
-\begin{equation*}
-\begin{split}
-\dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}}&\cid
-\dfrac{6}{8} \, \1_{\{\Gamma_1 2^{\alpha }<\Gamma_2\}}+
-\big(1-(\Gamma_1/\Gamma_2)^{2/\alpha}\big)\1_{\{\Gamma_12^{\alpha
-  }> \Gamma_2\}}\\
-& =\dfrac{3}{4}\,  \1_{\{U 2^{\alpha }<1\}}+
-\big(1-U^{2/\alpha}\big)\1_{\{U2^{\alpha
-  }> 1\}}=Y\,.
-\end{split}
-\end{equation*}
-The limit  distribution of the spectral gap
-has an atom at $3/4$ with probability $2^{-\alpha}$, i.e.,~$\P(Y=3/4)=2^{-\alpha}$, and
-\begin{equation*}
-\P(Y\le x)= 1- (1-x)^{\alpha/2},\qquad x\in (0,3/4).
-\end{equation*}
-In the iid case the limit distribution of the self-normalized spectral gap has distribution function $F(x) = 1- (1-x)^{\alpha/2}$ for $x\in [0,1]$.
-This means that the atom disappears if the entries are iid.
-Figure~\ref{fig:ProbMass} compares the distribution function of $Y$
-with $F$ for $\alpha=0.6$; the atom at $3/4$ is clearly visible.
-
-Along the same lines, we also have
-\begin{equation*}
-(a_{np}^{-2}\lambda_{(1)},\lambda_{(2)}/\lambda_{(1)}) \cid(
-8\,\Gamma_1^{-2/\alpha},\frac{1}{4}\,\1_{\{U<2^{-\alpha}\}}+U^{2/\alpha}\,\1_{\{U\ge
-  2^{-\alpha}\}})\,
-\end{equation*}
-and hence the limit distribution of $\lambda_{(2)}/\lambda_{(1)}$ is supported on $[1/4,1)$ with  mass of $2^{-\alpha}$ at $1/4$.
-The histogram of the ratio $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ based on 1000 replications from the model \eqref{ex:4.6} with noise given by a $t$-distribution with $\alpha=1.5$ degrees of freedom, $n=1000$ and $p=200$ is displayed in Figure \ref{fig:1}.
-\begin{figure} [h]
-\begin{center}
-\includegraphics[scale=.35]{histogram.pdf}
-\end{center}
-\caption{Histogram based on 1000 replications of $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ from model \eqref{ex:4.6}.}
-\label{fig:1}
-\end{figure} Observing that $2^{-\alpha}=0.3536\ldots$,
-the histogram is remarkably close to what one would expect from a sample from the truncated
-uniform \ds , $2^{-\alpha}\, \1_{\{U<2^{-\alpha}\}}+U\, \1_{\{U\ge 2^{-\alpha} \}}$.
-The mass of the limiting discrete component of the ratio can be much
-larger if one conditions on $a_{np}^{-2}\lambda_{(1)}$ being large.
-Specifically, for any $\epsilon\in (0,1/4)$ and $x>0$,
-\begin{equation*}
-\lim_{n\to\infty} \P(\epsilon<\lambda_{(2)}/\lambda_{(1)}\le 1/4|\lambda_{(1)}>a_{np}^2x)
-=\P(\Gamma_1/\Gamma_2\le 2^{-\alpha}|\Gamma_1<(x/8)^{-\alpha/2})=G(x)\,.
-\end{equation*}
-The function $G$ approaches $1$ as $x\to\infty$ indicating the speed at which the
-two largest eigenvalues get linearly related; see Figure \ref{fig:2} for a graph of $G$ in the case $\alpha=1.5$.
-\begin{figure} [h]
-\begin{center}
-\includegraphics[scale=.35]{Hfunction.pdf}
-\end{center}
-\caption{Graph of $G(x)=\P(\Gamma_1/\Gamma_2\le2^{-\alpha}|\Gamma_1<(x/8)^{-\alpha/2})$ when $\alpha=1.5$.}
-\label{fig:2}
-\end{figure}
-In addition, from Remark~\ref{rem:4.5}, we also have
-\begin{equation*}
-\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \cid \dfrac{4}{5}\,
-\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}}\,.
-\end{equation*}
-Clearly, the limit \rv\ is stochastically smaller than what one would get in the iid case; see \eqref{eq:limit}.
-\end{example}
-\begin{example}\label{exam:spectralgap}\rm
-The previous example also illustrates the behavior of the two largest eigenvalues in the general case
-when the rank $r$ of the matrix $\M$ is larger than one.
-We have in general
-\begin{equation*}
-\dfrac{\lambda_{(2)}}{\lambda_{(1)}}\cid
-\frac{v_2}{v_1}\,\1_{\{U<(v_2/v_1)^{\alpha/2}\}}+U^{2/\alpha}\,\1_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
-\end{equation*}
-In particular, the limiting
-{\em self-normalized spectral gap} has \rep
-\beao
-\dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}} \cid \dfrac{v_1-v_2}{v_1}\,\1_{\{U<(v_2/v_1)^{\alpha/2}\}}+(1-U^{2/\alpha})\,\1_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
-\eeao
-The limiting variable assumes values in $(0,1-v_2/v_1]$  and has an atom at the right end-point.
-This is in contrast to the iid case
-and to the case when $r=1$ (hence $v_2=0$) including the case of iid rows and the separable case; see Example~\ref{exam:separable}.
-\end{example}
-\begin{example}\label{exam:separable}
-{\em We consider the separable case when  $h_{kl}=\theta_kc_l$, $k,l \in \Z$, where $(c_l)$, $(\theta_k)$ are
-  real sequences such that the conditions on $(h_{kl})$ in
-  Theorem~\ref{thm:mains} hold. In this case,
-\begin{equation*}
-\M= \sum_{l\in \Z} c_l^2\;\; (\theta_i\theta_j)_{i,j \in \Z}\,.
-\end{equation*}
-Note that $r=1$ with the only non-negative eigenvalue
-\beao
-v_1=\sum_{l\in \Z} c_l^2\;\;\sum_{k\in \Z}\theta_k^2\,.
-\eeao
-In this
-case, the limiting point process in Theorem~\ref{cor:1ch3} is a PRM on
-$(0,\infty)$ with mean measure of $(y,\infty)$ given by $(v_1/
-y)^{\alpha/2}$, $y>0$. The normalized eigenvalues have similar asymptotic behavior as
-in the case of iid entries. For example, the log-spacings have the same limit as in the iid case for fixed $k$,
-\beao
-\big(\log \la_{(1)}-\log \la_{(2)},\ldots,\log \la_{(k+1)}-\log \la_{(k)}\big)\std
--\dfrac{2}{\alpha}\,\big(\log(\Gamma_1/\Gamma_2),\ldots,\log (\Gamma_{k}/\Gamma_{k+1})\big)\,.
-\eeao
-The same observation applies to the ratio of the largest eigenvalue and the trace in the case $\alpha\in (0,2)$:
-\beao
-\dfrac{\la_{(1)}}{{\rm tr} (\X \X')}=\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \std
-\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}}\,.
-\eeao
-We also mentioned in Example~\ref{exam:spectralgap} that the \ds al limit of the self-normalized spectral gap has no atom
-as in the iid case.
-}
-\end{example}
-
-\subsection{S\&P 500 data}\label{sec:sp500}
-We conduct a short analysis of the largest eigenvalues of the univariate log-return time series which compose
-the S\&P 500 stock index; see Section~\ref{subsec:1.2} for a description of the data.
-Although there is strong empirical evidence that these univariate series have power-law tails
-(see Figure~\ref{fig:SP500_tail_indices}) we  do not expect that they
-have the same tail index. One way to proceed would be to ignore this fact because the tail indices are
-in a close range and the differences are due to large sampling errors for estimating such quantities.  One could also collect \ts\ with similar
-tail indices in the same group.
-In this case, the dimension $p$ decreases. This grouping would be a rather arbitrary classification method.
-We have chosen a third way: to use rank transforms.
-This approach has its merits because it aims at standardizing the tails but it also has a major disadvantage: one destroys
-the covariance structure underlying the data.
-\par
-Given a $p\times n$ matrix $(R_{it})_{i = 1, \cdots, p;t=1,\cdots, n}$,
-we construct a matrix $\X$ via the rank transforms
-\begin{equation*}
-  X_{it} = -\Big[
-    \log \Big(\frac{1}{n+1} \sum_{\tau=1}^n \1_{\{R_{i\tau} \leq R_{it}\}} \Big)
-  \Big]^{-1} \,,\qquad i=1,\ldots,p;t=1,\ldots,n\,.
-\end{equation*}
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.550]{EigenRatioSP500_ranks_log_50_shown.pdf}
-  \caption{The logarithms of the ratios $\lambda_{(i+1)} / \lambda_{(i)}$ for the S\&P 500 series after rank transform.
-We also show the 1, 50 and 99\% quantiles (bottom, middle, top lines, respectively) of the variables
-$\log((\Gamma_i/ \Gamma_{i+1})^{2})$. }
-  \label{fig:EigenRatio}
-\end{figure}
-\begin{figure}[htb!]
-  \centering
-    \includegraphics[scale=0.550]{EigenRatioSP500_log_50_shown.pdf}
-  \caption{The logarithms of the ratios $\lambda_{(i+1)} / \lambda_{(i)}$ for the original (non-rank transformed) S\&P 500 log-return data.
-We also show the 1, 50 and 99\% quantiles (bottom, middle, top lines, respectively) of the variables
-$\log((\Gamma_i/ \Gamma_{i+1})^{2/2.3})$; see also Figure~\ref{fig:EigenRatio} for comparison.}\label{fig:22}
-\end{figure}
-
-%\begin{figure}[htb!]
-%  \centering
-%    \includegraphics[scale=0.550]{EigenRatioSP500_lin2_250_shown.pdf}
-%  \caption{{\green Same as Figure \ref{fig:EigenRatio} but on original scale, no logarithm. Choose one.}}
-%\end{figure}
-
-If the rows $R_{i1}, \ldots, R_{in}$ were iid (or, more generally, stationary ergodic)
-with a continuous \ds\  then
-the averages under the logarithm would be  asymptotically uniform
-on $(0,1)$ as $n \to \infty$. Hence $X_{it}$ would  be \asy ally standard
-Fr\'echet $\Phi_1$-distributed.
-In what follows, we assume that the aforementioned univariate \ts\ of the S\&P 500 index have undergone the rank transform and that
-their marginal \ds s are close to $\Phi_1$; we
-always use the symbol $\X$ for the resulting multivariate series.
-\par
-In Figure~\ref{fig:EigenRatio} we show the ratios of the consecutive ordered eigenvalues $(\la_{(i+1)}/\la_{(i)})$
-of the matrix $\X\X'$. This graph shows the rather surprising fact that the ratios are close to one even for small values $i$.
-We also show the 1, 50 and 99 \% quantiles of the variables $((\Gamma_{i}/\Gamma_{i+1})^{2/\alpha})$ calculated from the formula
-\beam\label{eq:iid}
-\P\big((\Gamma_{i} /\Gamma_{i+1})^{2/\alpha}\le x \big) = x^{i \cdot \alpha/2}, \quad x\in (0,1)\,.
-\eeam
-For increasing $i$, the \ds\ is concentrated closely to 1, in agreement with the \slln\ which yields
-$\Gamma_{i} /\Gamma_{i+1}\stas 1$ as $i\to\infty$.
-The \asy\ \ds s \eqref{eq:iid} correspond to the case when the matrix
-$\M$ has rank $r=1$. It includes the iid and separable cases;
-see Example~\ref{exam:separable}. The shown \asy\ quantiles are in agreement with the rank $r=1$ hypothesis.
-\par
-For comparison, in Figure~\ref{fig:22} we also show the ratios  $(\la_{(i+1)}/\la_{(i)})$ for the 
-non-rank transformed S\&P 500 data and the 1, 50 and 99\% quantiles of the variables
-$\log((\Gamma_i/ \Gamma_{i+1})^{2/\alpha})$, where we choose $\alpha=2.3$ motivated by the estimated tail indices in 
-Figure~\ref{fig:SP500_tail_indices}.
-The two graphs in Figure~\ref{fig:EigenRatio} and Figure~\ref{fig:22} are quite similar but the smallest ratios for the original data are slightly larger than
-for the rank-transformed data.
-
-%--------------------------------------------------------------------------------------
-\subsection{Sums of squares of sample autocovariance matrices}\label{sec:possemidef}
-In this section we consider some additive \fct s
-of the squares of $\A_n(s)=\X_n(0)\X_n(s)'$ given by $\A_n(s)\A_n(s)'$ for $s=0,1,\ldots$. By definition of the singular values of a matrix
-(see  \eqref{eq:sigma}), the non-negative definite
-matrix $\A_n(s)\A_n(s)'$ has eigenvalues $(\la_i^2(s))_{i=1,\ldots,p}$.
-\par
-The following result is a corollary of Theorem~\ref{thm:mains}.
-\begin{proposition}\label{thm:mainstr} Consider the linear process \eqref{eq:1} under
-the conditions of Theorem~\ref{thm:mains}. Then the following statements hold for $s\ge 0$:
-\begin{enumerate}
-\item[$(1)$]
-We consider two disjoint cases:
-$\alpha \in (0,2)$ and $\beta\in (0,\infty)$, or
-$\alpha\in [2,4)$ and $\beta$ satisfying \ref{Cbeta}. Then
-\beao
-a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-\delta_{(i)}^2(s)| \cip 0, \quad \nto.
-\eeao
-\item[$(2)$]
-Assume $\beta\in [0,1]$.
-If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta \in (\alpha/2-1,1]$, then
-\beao
-a_{np}^{-4} \max_{i=1,\ldots,p} |\la_{(i)}^2(s)-(\gamma_{(i)}^\rightarrow(s))^2| \cip 0, \quad \nto.
-\eeao
-Assume $\beta>1$. If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta^{-1} \in (\alpha/2-1,1]$. Then
-\beao
-a_{np}^{-4} \max_{i=1,\ldots,p} |\la_{(i)}^2(s)-(\gamma_{(i)}^\downarrow(s))^2| \cip 0, \quad \nto.
-\eeao
-\end{enumerate}
-\end{proposition}
-To the best of our knowledge, sums of squares of sample autocovariance matrices were used first in the paper by Lam and Yao
-\cite{lam:yao}; their \ts\ model is quite different from ours.
-\begin{proof}
-Part (1).
-The proof follows from Theorem~\ref{thm:mains} if we can show that
-\beao
-a_{np}^{-2}\max_{i=1,\ldots,p} \big(\la_{(i)}(s)+\delta_{(i)}(s) \big)=O_\P(1)\,\quad \nto\,.
-\eeao
-We have by Theorem~\ref{cor:1ch3},
-\beam \label{eq:gdfg}
-a_{np}^{-2}\max_{i=1\ldots,p} \la_{(i)}(s)=a_{np}^{-2}\la_{(1)}(s) \cid c\, \xi_{\alpha/2}\,,
-\eeam
-where $\xi_{\alpha/2}$ has a $\Phi_{\alpha/2}$ \ds . In view of Theorem~\ref{thm:mains}(1) we also have
-\beao
-a_{np}^{-2}\max_{i=1\ldots,p} \delta_{(i)}(s)\cid c\, \xi_{\alpha/2}\,.
-\eeao
-Therefore, again using Theorem~\ref{thm:mains}(1), we have
-\beao
-\lefteqn{a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-\delta_{(i)}^2(s)|}\\
-&\le &\big[a_{np}^{-2} \max_{i=1,\ldots,p} |\lambda_{(i)}(s)-\delta_{(i)}(s)|\big]\,
-\big[a_{np}^{-2} \max_{i=1,\ldots,p}\big ( |\lambda_{(i)}(s)|+|\delta_{(i)}(s)|\big)\big]
-\cip 0, \quad \nto.
-\eeao
-This proves part (1).\\[1mm]
-Part (2).
-Now assume $\beta\in [0,1]$ and $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta\in (\alpha/2-1,1]$.
-Then \eqref{eq:gdfg} is still true and we have by Theorem~\ref{thm:mains}(2) and Theorem~\ref{cor:1ch3}
-\beao
-a_{np}^{-2}\max_{i=1\ldots,p} \gamma_{(i)}^{\rightarrow}(s)\cid c\, \xi_{\alpha/2}\,.
-\eeao
-We then have
-\beao
-\lefteqn{a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-(\gamma_{(i)}^\rightarrow(s))^2|}\\
-&\le &\big[a_{np}^{-2} \max_{i=1,\ldots,p} |\lambda_{(i)}(s) -\gamma_{(i)}^\rightarrow(s)|\big]\,
-\big[a_{np}^{-2} \max_{i=1,\ldots,p}\big ( \lambda_{(i)}(s)+\gamma_{(i)}^\rightarrow(s)
-\big)\big]
-\cip 0\,, \, \nto.
-\eeao
-The proof of the remaining part is similar and therefore omitted.
-\end{proof}
-Now, using Proposition~\ref{thm:mainstr} and a continuous mapping argument, we can show
-limit theory for the eigenvalues
-\beao
-w_{(1)}(s_0,s_1)\ge \cdots \ge w_{(p)}(s_0,s_1)\,,\qquad 0\le s_0\le s_1\,,
-\eeao
-of the non-negative definite random matrices
-\begin{equation}\label{eq:sumA}
-\sum_{s=s_0}^{s_1} \A_n(s) \A_n(s)'\,.
-\end{equation}
-
-\begin{proposition}\label{prop:sumsmal}
-Assume $0\le s_0\le s_1$ and the conditions of Theorem~\ref{thm:mains} hold.
-If $\alpha \in (0,4)$ and $\beta\in (0,1] \cap  (\alpha/2-1,1]$ then
-\begin{equation*}
-a_{np}^{-4} \max_{i=1,\ldots,p} |w_{(i)}(s_0,s_1)-\omega_{(i)}(s_0,s_1)| \cip 0, \quad \nto,
-\end{equation*}
-where $\omega_{(i)}(s_0,s_1)$ are the ordered values of the set $\{Z_{(i),np}^4 v_j(s_0,s_1), i=1,\ldots,p;j=1,2,\ldots\}$
-and $(v_j(s_0,s_1))$ are the ordered eigenvalues of $ \sum_{s=s_0}^{s_1} \M(s)\M(s)'$.
-\end{proposition}
-
-\begin{example}\label{exam:additive}{\em
-\begin{figure}[htb!]
-  \centering
-  \subfigure[]
-{
-    \includegraphics[scale=0.34]{eigen_sum.pdf}
-    \label{fig:LamYao:a}
-  }
-  \subfigure[] {
-    \includegraphics[scale=0.31]{eigen_sum_ratio.pdf}
-    \label{fig:LamYao:b}
-  }
-  \caption{The largest eigenvalues of the sums of the squared autocovariance
-    matrices compared with the sums of the largest eigenvalues of these matrices
-    for the S\&P 500 data for different values $s_1$. The two values are surprisingly close to each other; mind the scale
-of the $y$-axis. We also show their ratios.}
-  \label{fig:LamYao}
-\end{figure}
-Recall the separable case from Example~\ref{exam:separable}, i.e.,
-$h_{kl}=\theta_kc_l$, $k,l \ge 0$, where $(c_l)$, $(\theta_k)$ are
-real sequences such that the conditions on $(h_{kl})$ in
-Theorem~\ref{thm:mains} hold.
-%Let $\alpha \in (0,2)$ and set
-%\begin{equation*}
-%\theta = (\theta_0, \theta_1,\theta_2,\ldots)' \quad \mbox{and} \quad
-%c= (c_0,c_1,c_2, \ldots)'.
-%\end{equation*}
-Write $\Theta_{ij}=\theta_i \theta_j$. It is symmetric and has rank one; the only non-zero
-eigenvalue is $\gamma_\theta(0)=\sum_{k=0}^\infty \theta_k^2$. Hence $\Theta$  is non-negative definite.
-We get from \eqref{eq:m} that
-\begin{equation*}
-\M(s)=\gamma_c(s)\, \Theta, \quad s\ge 0\,,
-\end{equation*}
-where
-\beao
-\gamma_c(s)=\sum_{l=0}^\infty c_lc_{l+s}\,,\qquad s\ge 0\,.
-\eeao
-The matrix $\M(s)$ has the only non-zero eigenvalue $\gamma_c(s)\gamma_\theta(0)$.
-The factors $(\gamma_c(s))$ can be positive or negative; they constitute the autocovariance \fct\ of a stationary linear process
-with coefficients $(c_l)$.
-Accordingly, $\M(s)$ is either non-negative or non-positive definite. %This explains why a symmetrization of $\M(s)$ is irrelevant.
-Now we consider the non-negative definite matrix
-\beao
-\sum_{s=s_0}^{s_1} \M(s)\,\M(s)'= \sum_{s=s_0}^{s_1}\gamma_c^2(s)\,\Theta\Theta'\,.
-\eeao
-This matrix has rank $1$ and its largest eigenvalue is given by
-\begin{equation*}
-C_{c,\theta}(s_0,s_1)=\sum_{s=s_0}^{s_1}\gamma_c^2(s)\,\gamma_\theta^2(0)\,.
-\end{equation*}
-An application of
-Proposition~\ref{prop:sumsmal} yields that the ordered eigenvalues of the matrix $a_{np}^{-4}\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'$
-are uniformly approximated by the quantities
-\begin{equation}\label{eq:drtgdfg}
-a_{np}^{-4} Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\,,\qquad  i=1,\ldots,p\,.
-\end{equation}
-Since
-\beao
-C_{c,\theta}(s_0,s_1)= \sum_{i=s_0}^{s_1} C_{c,\theta}(i,i)
-\eeao
-one gets the remarkable property that
-\beao
-&&a_{np}^{-4}\max_{i=1,\ldots,p} \Big| \la_{(i)}\Big(\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'\Big)- Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\Big|\\
-&=&a_{np}^{-4}\max_{i=1,\ldots,p} \Big|\sum_{s=s_0}^{s_1}\la_{(i)}(\A_n(s)\A_n(s)')- Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\Big|+o_P(1)\,.
-\eeao
-In particular, for $s_1\ge s_0$ we get the weak \con\ of the \pp es towards a PRM:
-\begin{equation*}
-\begin{split}
-\sum_{i=1}^p &\varepsilon_{a_{np}^{-4}\Big(\la_{i}\Big(\sum_{s=s_0}^{s_0}\A_n(s)\A_n(s)'\big)\,,\ldots,\la_{i}\big(\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'\big)\Big)}\\
-&\std \sum_{i=1}^\infty \varepsilon_{\Gamma_i^{-4/\alpha} \Big(C_{c,\theta}(s_0,s_0),\ldots,C_{c,\theta}(s_0,s_1)\Big)}\,,\quad \nto\,.
-\end{split}
-\end{equation*}
-
-}
-\end{example}
-\begin{example}\em
-In Figure~\ref{fig:LamYao} we calculate the largest eigenvalues\\
-$\la_{(1)}\big(\sum_{s=0}^{s_1}\A_n(s)\A_n(s)'\big)$ for $s_1=0,\ldots,5$ as well as the sums of the largest eigenvalues
-$\sum_{s=0}^{s_1}\la_{(1)}\big(\A_n(s)\A_n(s)'\big)$
-the log-return series from the S\&P 500 index described in Section~\ref{subsec:1.2}. The data are not rank-transformed.
-We notice that
-the two values are surprisingly close across the values $s_0=0,\ldots,5$. This phenomenon
-could be explained by the structure of the eigenvalues in Example~\ref{exam:additive}.
-Also note that the largest eigenvalue $\A_n(0)\A_n(0)'$ makes a major contribution to the values in Figure~\ref{fig:LamYao};
-the contribution of the squares $\A_n(s)\A_n(s)'$, $s=1,\ldots,5,$ to the largest eigenvalue of the sum of squares is less substantial.
-%Figure~\ref{fig:LamYao}(b) shows the same calculations for the matrices studied in Example~\ref{exam:4.6cont}.
-Here cancellation is possible (see \eqref{eq:Xrgsdf}) which is indicated by the fact that the red line does not follow the blue line as well as in (a).
-\end{example}
-%\section{\red Some conclusions?}\setcounter{equation}{0}
-%{\em Say something about EVT and largest eigenvalues}
-
-\section{Auxiliary results}\label{appendix:Ach3}
-
-Let $(Z_i)$ be iid copies of $Z$ whose distribution satisfies
-\begin{equation*}
-\P(Z>x)\sim p_+ \dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad  \P(Z\le -x)\sim p_-
-\dfrac{L(x)}{x^{\alpha}} \,,\quad \xto\,,
-\end{equation*}
- for some tail index $\alpha>0$,
-where $p_+,p_-\ge 0$ with $p_++p_-=1$ and $L$ is a slowly varying function.
-If $\E[|Z|]<\infty$ also assume $\E[Z]=0$.
-The product $Z_1Z_2$ is regular varying with the same index $\alpha$ and $\P(|Z_1Z_2|>x)= x^{-\alpha} L_1(x)$, where $L_1$ is slowly varying function different from $L$;
-see Embrechts and Goldie \cite{embrechts:goldie:1980}.
-Write
-\begin{equation*}
-S_n=Z_1+\cdots +Z_n\,,\quad n\ge 1,
-\end{equation*} and consider a sequence $(a_n)$ such that $\P(|Z|>a_n)\sim n^{-1}$.
-
-\subsection{Large deviation results}
-The following theorem can be found in
-Nagaev \cite{nagaev:1979} and Cline and Hsing
-\cite{cline:hsing:1998} for $\alpha>2$ and $\alpha\le 2$,
-respectively; see also  Denisov et al.~\cite{denisov:dieker:shneer:2008}.
-\begin{theorem}\label{thm:nagaevch3}
-Under the assumptions on the iid sequence $(Z_t)$
-given above the following relation holds
-\begin{equation*}
-\sup_{x\ge c_n}\left|\dfrac{\P(S_n>x )}{n\P(|Z|>x)} -p_+ \right|\to 0\,,
-\end{equation*}
-where $(c_n)$ is any sequence satisfying $c_n/a_n\to  \infty$ for
-$\alpha\le 2$ and $c_n\ge \sqrt{(\alpha-2)n\log n}$ for $\alpha>2$.
-\end{theorem}
-
-\subsection{A point process convergence result}
-Assume that the conditions at the beginning of Appendix \ref{appendix:Ach3} hold.
-Consider a sequence of iid copies $(S_{n}^{(t)})_{t=1,2,\ldots}$
-of $S_n$ and the sequence of point processes
-\begin{equation*}
-N_n= \sum_{t=1}^{p} \vep_{a_{np}^{-1} S_{n}^{(t)}}, \quad n=1,2,\ldots\,,
-\end{equation*}
-for an integer sequence $p=p_n\to\infty$. We assume that the state space of the
-point processes $N_n$ is $\overline{\R}_0=[\R\cup\{\pm \infty\}]\backslash \{0\}$.
-\begin{lemma}\label{lem:pprch3}
-Assume $\alpha \in (0,2)$ and the
-conditions of  Appendix~\ref{appendix:Ach3}
-on the iid sequence $(Z_t)$ and the normalizing sequence $(a_n)$. Then the limit relation
-$N_n\cid N$ holds in the space of point measures on $\overline{\R}_0$
-equipped with the vague topology (see
-\cite{resnick:1987,resnick:2007})
-for a  Poisson random measure $N$ with state space $\overline{\R}_0$ and intensity measure $\mu_\alpha(dx)=\alpha |x|^{-\alpha-1} (p_+ \1_{\{x>0\}}+ p_- \1_{\{x<0\}}) dx$.
-\end{lemma}
-\begin{proof}
-According to Resnick \cite{resnick:1987}, Proposition 3.21, we need to
-show that
-$p\, \P(a_{np}^{-1}S_n\in \cdot)\civ \mu_\alpha
-$,
-where $\civ$ denotes vague convergence of Radon measures on  $\overline{\R}_0$.
-Observe that we have $a_{np}/a_n\to\infty$ as $\nto$. This fact and
-$\alpha\in (0,2)$ allow one to apply
-Theorem~\ref{thm:nagaevch3}:
-\begin{equation*}
-\dfrac{\P( S_n >x a_{np})}{n\,\P(|Z|>  a_{np})}\to p_+ x^{-\alpha}
-\quad \mbox{and}\quad \dfrac{\P( S_n \le -x a_{np})}{n\,\P(|Z|>
-  a_{np})}\to p_-\, x^{-\alpha}\,,\quad x>0\,.
-\end{equation*}
-On the other hand, $n\,\P(|Z|>  a_{np})\sim p^{-1}$ as $\nto$.
-This proves the lemma.
-\end{proof}
-
-%\section*{Acknowledgments}
-
-%We thank Olivier Wintenberger for reading the manuscript and fruitful discussions.
+\input{../papers/DavisMikoschHeinyXie/PaperBody}

### 2017-07-21 phd/Chapterintroduction.tex
Added my own thesis.

@@
 
-In this thesis we study the largest and smallest eigenvalues of high-dimensional sample covariance and correlation matrices of heavy-tailed time series.
+\section{Regular Variation}
+The concept of regular variation is defined by the following scaling
+property: If a function $f(\cdot)$ satisfies
+\[
+\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
+\]
+then we say $f(x)$ is regularly varying with index $\alpha$.
+Distribution functions, say $F(\cdot)$, whose survival function
+$\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
+observed in a variety of phenomena, for example, phase transition of
+matter, black body radiation, neuronal avalanches and stock returns.
+When expanded to multiple dimensions, the scaling property is better
+described in terms of weak convergence to a spectral measure: If a
+random vector $V$ satisfies
+\[
+\lim_{x \to \infty}
+{
+  \P(V/|V| \in \cdot, |V| > c x)
+  \over
+  \P(|V| > x)
+}
+\overset{w}{\to} c^{-\alpha} \mu_\alpha(\cdot)
+\]
+then we say the components of $V$ is jointly regularly varying with
+index $\alpha$. Here $\mu_\alpha(\cdot)$ is a probability measure on
+the unit sphere \cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
+is jointly regularly varying with index $\alpha$, then each component
+of it and each linear combination of its components are regularly
+varying with the same index $\alpha$. This follows from Feller
+\cite{feller}, p. 278. Cf. also Jessen and Mikosch
+\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et al
+\cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.
 
-\section{Random Matrix Theory}\label{sec:rmt}
+Thus estimating the tail index $\alpha$ is particularly important for
+understanding the behaviour of a heavy-tailed series. The earliest
+method proposed for this purpose is due to Hill \cite{hill1975simple}:
+\[
+\hat \alpha_H = \left[
+  {1 \over k} \sum_{i=1}^k \log \left(
+  X_{(i)} \over X_{(k+1)}
+  \right)
+  \right]^{-1}
+\]
+where $X_1, X_2, \dots$ is the series whose tail index is the subject
+of estimation, and $X_{(i)}$ is the $i$-th upper order statistics of
+this series. Several authors have contributed to showing the
+weak consistency of the estimator $\hat \alpha_H$,
+i.e. if $k \to \infty, k/n \to 0$ as $n \to \infty$, then
+$\hat \alpha_H \overset{P}{\to} \alpha$.
 
-The field of Random Matrix Theory (RMT) is concerned with the spectral properties of high-dimensional random matrices. Its development was motivated by applications.
-In quantum mechanics, for example, the energy levels of particles in a large system can be characterized by the eigenvalues of a random infinite-dimensional Hermitean operator $\bfW$ on a Hilbert space.
-It is common to work with discretizations in a finite-dimensional space.
-In this case $\bfW$ turns into a high-dimensional Hermitean random matrix.
-Such matrices are called Wigner matrices named after Eugene Paul Wigner.
-He proved in the $1950$s that
-if $\bfW$ has independent standard normal entries on and above the diagonal, then
-the expected empirical distribution of the eigenvalues of $\bfW$ tends to the so-called semi-circle law defined in \eqref{eq:semi-circle}, as the dimension of $\bfW$ goes to infinity; see for example \cite{wigner:1955,wigner:1957}.
-Since then a great variety of asymptotic results has been proved for various classes of random matrices under different assumptions on the distribution of the entries and their dependence structure.
-The discovery of many results was triggered by the enormous improvement of computation power which led to numerous conjectures.
-In the second half of the $20$th century, the research on asymptotic spectral properties of large-dimensional random matrices attracted considerable interest among physicists, computer scientists and mathematicians.
-A breakthrough in the theory of spectral distributions of sample covariance matrices was achieved by Mar\v cenko and Pastur in $1967$. 
-%In $2010$, the most famous conjecture about the circular law was solved in \cite{tao:vu:krishnapur:2010}.
-
-For many years the main focus of research in RMT has been on limiting spectral distributions.
-More recently, the focus turned to linear spectral statistics, eigenvectors, limiting distributions of extreme eigenvalues and their spacings.
-This thesis addresses these four topics in the special setting of random matrices with  heavy-tailed entries.
-
-RMT is a versatile and useful tool in many fields of modern sciences that are faced with high-dimensional data sets.
-It  employs techniques from probability theory, multivariate statistics, number theory and combinatorics.
-Moreover, RMT finds applications in quantum physics, signal processing, wireless communications and finance; see Bai et al.~\cite{bai:fang:liang:2014} for more detailed examples.
-%``Big data'' is a popular topic nowadays. To meet the demand for data analysis and dimension reduction techniques RMT brings together
-
-\subsection{Limiting spectral distributions}
-
-For any random $p\times p$ matrix $\bfA$ with real eigenvalues $\lambda_1(\bfA),\ldots,\lambda_p(\bfA)$ the {\em empirical spectral distribution} is defined by 
-\begin{equation*}
-F_{\bfA}(x)= \frac{1}{p}\; \sum_{i=1}^p \1_{\{ \lambda_i(\bfA)\le x \}}, \qquad x\in  \R\,.
-\end{equation*}
-
-A major problem in random matrix theory is to find the weak limit of $(F_{\bfA_n})$, the so-called {\em limiting spectral distribution}, for suitable sequences of Hermitean $p\times p$ matrices $(\bfA_n)$.
-By weak convergence of a sequence of probability distributions $(F_{\bfA_n})$
-to a \pro y \ds\ $F$, we mean $\lim_{\nto} F_{\bfA_n}(x)=F(x) \,\as$ for all continuity points of $F$.
-Although the eigenvalues of $\bfA_n$ are continuous functions of the entries of $\bfA_n$ there are no closed-form expressions if the dimension is larger than $4$.
-Therefore methods to identify and characterize the limiting spectral distribution are needed.
-We will briefly discuss the two most common ones: the method of moments and Stieltjes transforms.
-\par
-
-By Lemma B.3 in \cite{bai:silverstein:2010}, a distribution function $F$ is uniquely characterized by its sequence of moments 
-\begin{equation*}
-\beta_k=\int_0^{\infty} x^k \dint F (x)\,, \quad k=1,2,\ldots
-\end{equation*}
-if Carleman's condition $\sum_{k=1}^{\infty} \beta_{2k}^{-1/(2k)}=\infty$ is satisfied. In this case, weak convergence of $(F_{\bfA_n})$ to $F$ is equivalent to the convergence of moments, that is
-\begin{equation}\label{eq:mmom}
-\beta_k(\bfA_n)= \int_0^{\infty} x^k \dint F_{\bfA_n} (x)=\frac{1}{p} \tr(\bfA_n^k) \to \beta_k\,,\quad \nto, \,k=1,2,\ldots
-\end{equation}
-In many cases the calculation of $\tr(\bfA_n^k)$ is very demanding. Often its mean and variance are estimated by combinatorial techniques.
-On the positive side, if $F$ has finite support, Carleman's condition holds automatically.
-\par
-
-Another useful tool is the {\em Stieltjes transform}
-of the empirical spectral distribution $F_{\bfA}$:
-\begin{equation*}
-s_{F_{\bfA}}(z)= \int_{\R} \frac{1}{x-z} \dint F_{\bfA}(x) = \frac{1}{p} \tr\big((\bfA -z \bfI)^{-1}\big)\,, \quad z\in \C^+\,,
-\end{equation*}
-where $\C^+$ denotes the complex numbers with positive imaginary part. 
-Weak convergence of $(F_{\bfA_n})$ to a distribution function $F$ is equivalent to $s_{F_{\bfA_n}}(z) \to s_F(z)$ a.s. for all $z\in \C^+$.
-Notice that the Stieltjes transform $s_F$ determines a distribution function $F$ at all continuity points $a,b$ of $F$ via
-\begin{equation*}
-F(b)-F(a)= \lim_{v\to 0^+} \frac{1}{\pi} \int_a^b \Im s_F(x+\mathrm{i} v) \dint x\,.
-\end{equation*}
-
-\section{Sample covariance matrices}
-For a sample of $n$ column vectors $\x_1,\ldots,\x_n$ of a $p$-dimensional time series the (non-normalized) {\em sample covariance matrix} is usually defined as
-\begin{equation*}
-\bfS=\sum_{i=1}^n (\x_i-\overline{\x})(\x_i-\overline{\x})'= \sum_{i=1}^n \x_i \x_i' - n\overline{\x}\,\overline{\x}'= \X\X' - n\overline{\x}\,\overline{\x}'\,,
-\end{equation*}  
-where $\overline{\x}=n^{-1}\sum_i \x_i$ is the sample mean and 
-\begin{equation*}
-\X=(\x_1, \ldots, \x_n) =(X_{it})_{i=1,\ldots,p;\,t=1,\ldots,n}
-\end{equation*}
-is the {\em data matrix}.
-The sample covariance matrix is of crucial importance in multivariate statistics, for instance in principal component analysis, canonical correlation analysis, multivariate regression, factor analysis, hypothesis testing and discriminant analysis.
-The case of multivariate normal observations has played a particular role in the development of statistical theory.
-Already in $1928$, Wishart \cite{wishart:1928} studied sample covariance matrices with normal entries.
-Through the $20$th century non-asymptotic procedures for Gaussian observations such as Fisher's test, Student's test, and the analysis of variance were developed.
-In practice, however, observations are often not normally distributed and asymptotic methods based on limit theorems for certain model parameters are employed instead of exact results which are difficult to obtain.
-
-Most of the classical limit theorems are derived under the assumption that the dimension $p$ is fixed and the sample size $n$ goes to infinity. 
-If the assumptions of the law of large numbers are satisfied, then $n^{-1} \bfS$ converges \as~to the covariance matrix $\Sigma$ of $\x_1$.
-
-If $p$ is moderately large, it is known that $n^{-1} \bfS$ ceases to be a good estimate for $\Sigma$.
-This means that classical methods based on fixed dimension and large sample limits may lead to wrong conclusions when applied to high-dimensional data.
-One would need appropriate adjustments.
-RMT provides limit theory in the case of large $p$.
-New statistical methods can be built on these results.
-
-In RMT one assumes that $p=p_n$ grows with $n$. The most common condition in the literature is
-\begin{equation}\label{Cp}
-\lim_{\nto} \frac{p_n}{n}\to \gamma\in (0,\infty)\,. %\tag{$G_{\gamma}$}
-\end{equation}
-The asymptotic spectral behaviors of two large matrices are the same if their difference is of finite rank.
-Therefore we will refer to $\X\X'$ as the {\em sample covariance matrix} from now on.
-Indeed, by the rank inequality (see Bai and Silverstein \cite[Theorem A.44]{bai:silverstein:2010}) we have for the supremum norm
-\begin{equation*}
-\| F_{n^{-1}\bfS}-F_{n^{-1}\X\X'} \| \le p^{-1} \mathrm{rank}((\overline{\x},\ldots,\overline{\x}))\,,
-\end{equation*}
-which means that the respective limiting spectral distributions coincide under \eqref{Cp}.
-For the same reason we can assume without loss of generality that the entries $(X_{it})$ are centered whenever the expectations exist.
-\par
-
-The limiting spectral distribution of normalized sample covariance matrices was found by Mar\v cenko and Pastur.
-
-\begin{theorem}[Debashis and Aue \cite{debashis:aue:2014}]\label{thm:mpch1}
-Suppose that $\X$ has iid entries with common mean and variance $1$.
-If \eqref{Cp} holds, then, with probability one, $(F_{n^{-1}\X \X'})$ converges weakly to a non-random distribution, the so-called  Mar\v cenko--Pastur law $F_\gamma$.
-If $\gamma \in (0,1]$,  $F_\gamma$  has density,
-\begin{eqnarray}\label{eq:MPch1}
-f_\gamma(x) =
-\left\{\begin{array}{ll}
-\frac{1}{2\pi x\gamma} \sqrt{(b-x)(x-a)} \,, & \mbox{if } a\le x \le b, \\
- 0 \,, & \mbox{otherwise,}
-\end{array}\right.
-\end{eqnarray}\noindent
-where $a=(1-\sqrt{\gamma})^2$ and $b=(1+\sqrt{\gamma})^2$.
-If $\gamma>1$, the \MP law is a mixture of a point mass at $0$ and the density function $f_{1/\gamma}$ with weights $1-1/\gamma$ and $1/\gamma$, respectively.
-\end{theorem}
-
-The \MP law describes the global behavior of the eigenvalues of $\X \X'$.
-Theorem \ref{thm:mpch1} quantifies the spread of the eigenvalues around their mean $1$.
-Note that the range of the deviation increases when $\gamma$ increases from $0$ to $\infty$.
-
-If $p/n\to 0$, the limiting spectral distribution in Theorem \ref{thm:mpch1} is the Dirac measure at $1$.
-After an appropriate transformation of the sample covariance matrix one can obtain the semi-circle law defined in \eqref{eq:semi-circle} as a non-degenerate limiting spectral distribution in this case.
-
-The crucial assumptions in Theorem \ref{thm:mpch1} are the finiteness of the variance and that $p$ and $n$ tend to infinity at the same rate.
-By Theorem 2.8 in Bai \cite{Bai:1999}, the conclusion still holds if the entries are independent, have common mean and satisfy the Lindeberg-type condition
-\begin{equation*}
-\lim_{\nto} \frac{1}{\delta^2 n p}\sum_{i,t} \E[X_{it}^2 \1_{\{|X_{it}|>\delta \sqrt{n}  \}}] =0\,,\quad \delta>0\,.
-\end{equation*}
-\par
-
-Many important test statistics in multivariate analysis are functions of the eigenvalues
-$\la_{(1)}\ge \cdots \ge \la_{(p)}$ of the sample covariance matrix and can be expressed by means of the empirical spectral distribution of $n^{-1}\X\X'$.
-
-\begin{example}{\em 
-Let $(X_{it})$ be iid standard normal and consider
-\begin{equation*}
-V_n:=\log(\det (n^{-1}\X\X'))=
-\log \prod_{i=1}^p \frac{\lambda_{(i)}}{n} =  p \int_{0}^{\infty} \log x \;\dint F_{n^{-1}\X\X'}(x)\,.
-\end{equation*}
-On the one hand, if $p$ is fixed we know from Example 1.1 in Yao et a.~\cite{yao:zheng:bai:2015} that 
-\begin{equation}\label{eq:srgdrgd}
-\sqrt{\frac{n}{p}}\, V_n \cid Y\sim N(0,2)\,,\quad \nto\,.
-\end{equation}
-On the other hand, if $p/n \to \gamma\in (0,1)$, one can use Theorem \ref{thm:mpch1} to obtain asymptotic values for $V_n$. On gets \as
-\begin{equation*}
-\frac{V_n}{p}\to \int_{0}^{\infty} \log x \;\dint F_{\gamma}(x)= \frac{\gamma-1}{\gamma} \log(1-\gamma)-1<0\,,
-\end{equation*}
-which implies $\sqrt{n/p}\, V_n \to -\infty$ \as~In view of \eqref{eq:srgdrgd} the asymptotic behavior of the test statistic $V_n$ crucially depends on the dimension $p$.}
-\end{example}
-
-A characterization of the limiting spectral distribution of sample covariance matrices with  general population covariance has been derived for many settings.
-As an example we state Theorem 1.1 in Bai and Zhou \cite{bai:zhou:2008}.
-%a result that we will use in Chapter \ref{ch:corr} to find the limiting spectral distribution of sample correlation matrices.
-\begin{theorem}\label{them:baizhou}
-Assume \eqref{Cp} and the following conditions.
-\begin{itemize}
-\item For all $k$, $\E[X_{jk}X_{lk}]=T_{lj}$, and for any non-random $p\times p$ matrix $\bfB$ with bounded norm, $\E[(\x_k'\bfB \x_k -\tr(\bfB \bfT))^2]=o(n^{-2})$, where $\bfT=\bfT_n=(T_{jl})$ and $\x_k$ are the columns of $\X$.
-\item The norm of $\bfT_n$ is uniformly bounded and $F_{\bfT_n}$ tends to a non-random probability distribution $H$.
-\end{itemize}
-Then, with probability $1$, $F_{n^{-1}\X\X'}$ tends to a probability distribution, whose Stieltjes transform $m(z)$ satisfies
-\begin{equation}\label{eq:mpequation}
-m(z)=\int_{\R} \frac{1}{t(1-\gamma-\gamma z m(z))-z} \dint H(t)\,.
-\end{equation}
-If $\underline{m}(z)=-(1-\gamma)/z+\gamma m(z)$, then \eqref{eq:mpequation} becomes
-\begin{equation}\label{eq:silversteineq}
-z=-\frac{1}{\underline{m}(z)}+\gamma \int_{\R} \frac{1}{1+\underline{m}(z) \,t} \dint H(t)\,.
-\end{equation}
-\end{theorem}
-For historical reasons, \eqref{eq:mpequation} is often called {\em \MP equation}.
-In practice, the inversion of such integral equations can be very difficult.
-Therefore, a characterization of the limiting spectral distribution via \eqref{eq:mpequation} is of limited use.
-For numerical procedures, the {\em Silverstein equation} \eqref{eq:silversteineq} is sometimes preferred.
-Roughly speaking, the only known explicit examples of non-degenerate limiting spectral distributions are the \MP law, the circular law, the semi-circle law and the multivariate $F$-matrix.
-In contrast, the number of existence results is huge.
-We refer to the discussion in Yao et al.~\cite{yao:zheng:bai:2015} for further details.
-\medskip
-
-In Theorem \ref{thm:mpch1} we have seen that the sequence $(F_{n^{-1}\X \X'})$ converges to the Mar\v cenko--Pastur law if the iid entries possess a finite second moment.
-Now we will discuss the situation when the entries are still iid, but have an infinite variance.
-Here we assume the entries to be regularly varying with index $\alpha \in (0,2)$; see \eqref{eq:regvarch1} on page \pageref{eq:regvarch1} for the definition of regular variation.
-Assuming \eqref{Cp} with $\gamma\in (0,1]$ in this infinite variance case, Belinschi et al.~\cite[Theorem~1.10]{belinschi:dembo:guionnet:2009} showed that the sequence $(F_{a_{n+p}^{-2}\X \X'})$ converges with probability one to a non-random probability measure with density $\rho_{\alpha}^\gamma$ satisfying
-\begin{equation}\label{eq:esd}
-\rho_{\alpha}^\gamma(x) x^{1+\alpha/2} \to \frac{\alpha \gamma}{2(1+\gamma)}, \quad x \to \infty\,;
-\end{equation}
-see also Ben Arous and Guionnet \cite[Theorem~1.6]{arous:guionnet:2008}.
-Here the normalization $(a_k)$ is defined \sth\ 
-\begin{equation}\label{eq:defanp}
-\P(|X|>a_k)\sim k^{-1}\,,\quad k\to\infty\,.
-\end{equation}
-An application of the Potter bounds (see Bingham et al.
-\cite[p.~25]{bingham:goldie:teugels:1987}) shows that $a_{n+p}^2/n \to \infty$.
-To the best of our knowledge, explicit expressions or computational methods for the limiting spectral distribution in the infinite variance case are not available at this moment.
-
-\section{Limits of extreme eigenvalues under finite fourth moment}
-
-After the limiting spectral distribution for sample covariance matrices had been found, the focus shifted to the asymptotic behavior of the largest and smallest eigenvalues $\la_{(1)}$ and $\la_{(p)}$, respectively, of $\X\X'$.
-In this section, we assume that the entries of the data matrix $\X$ are iid with generic element $X$.
-Furthermore, suppose $p\le n$; otherwise $\la_{(p)}=0$ since $\X\X'$ has at most $\min(n,p)$ non-zero eigenvalues.
-We will discuss the setting $\E[X^4]<\infty$, while the case $\E[X^4]=\infty$ is treated in Section \ref{sec:contr}.
-
-Under condition \eqref{Cp} with $\gamma\le 1$, one can infer from Theorem \ref{thm:mpch1} that
-\begin{equation}\label{eq:conlcu}
-\liminf_{n\to \infty}\frac{\la_{(1)}}{n} \ge (1+\sqrt{\gamma})^2\quad \mbox{and}\quad
-\limsup_{n\to \infty}\frac{\la_{(p)}}{n} \le (1-\sqrt{\gamma})^2\quad \as 
-\end{equation}
-We follow Bai and Silverstein \cite{bai:silverstein:2010} and derive necessary conditions for the \as~convergence of $n^{-1}\la_{(1)}$.
-Since the largest diagonal entry of a matrix is bounded by its largest eigenvalue, we have
-\begin{equation}\label{eq:simplebo}
-\frac{\la_{(1)}}{n}\ge \max_{i=1,\ldots,p} \frac{1}{n} \sum_{t=1}^n X_{it}^2\,.
-\end{equation}
-If $\E[X^4]=\infty$, then by Lemma B.25 in \cite{bai:silverstein:2010}
-\begin{equation}\label{eq:sghdfgdfs}
-\limsup_{n\to \infty}\max_{i=1,\ldots,p} \frac{1}{n} \sum_{t=1}^n X_{it}^2 =\infty \quad \as
-\end{equation}
-%Next note that Theorem \ref{thm:mpch1} holds if $\E[X]=c\neq 0$. 
-If $\E[X^4]<\infty$ and $\E[X]=c\neq 0$, then
-\begin{equation*}
-\frac{1}{\sqrt{n}}\twonorm{\X}\ge \frac{1}{\sqrt{n}}\twonorm{\E[\X]}  -\frac{1}{\sqrt{n}}\twonorm{\X-\E[\X]}
-\ge \frac{|c| p}{\sqrt{n}}  - \frac{1}{\sqrt{n}}\twonorm{\X-\E[\X]} \to\infty \quad \as,
-\end{equation*}
-where for any matrix $\bfA$, $\twonorm{\bfA}$ denotes its spectral norm, i.e., its largest singular value.
-This and \eqref{eq:sghdfgdfs} show that $\E[X]=0$ and $\E[X^4]<\infty$ are necessary conditions for the \as~convergence of $n^{-1}\la_{(1)}$.
-
-{\em In what follows, we assume $\E[X]=0$ and $\E[X^2]=1$, whenever the respective moments exist.}
-
-\subsection{Sample covariance matrices}
-Extending work by Geman \cite{geman},
-Bai et al.~\cite{bai:yin:krishnaiah:1988} 
-showed under \eqref{Cp} that
-\beam\label{eq:gemanch1}
-\dfrac {\la_{(1)}}{n} \stas \big(1+\sqrt{\gamma}\big)^2\,,\qquad \nto\,,
-\eeam
-which is the optimal result in view of \eqref{eq:sghdfgdfs}. Later Bai and Yin \cite{bai:yin:1993} proved the following result under the additional assumption $\gamma\in (0,1)$:
-\beam\label{eq:jo45}
-\limsup_{n\to \infty} \twonorm{n^{-1} \X\X'-(1+\gamma) \bfI} \le 2 \sqrt{\gamma}\,\quad \as
-\eeam
-Because of 
-\begin{equation*}
-\twonorm{n^{-1} \X\X'-(1+\gamma) \bfI} = \max \{ \la_{(1)}/n- (1+\gamma), -\la_{(p)}/n+(1+\gamma)\}\,,
-\end{equation*}
-equations \eqref{eq:jo45} and \eqref{eq:conlcu} imply 
-\begin{equation}\label{eq:sdfgsdfdf}
-\lim_{n\to \infty}\frac{\la_{(1)}}{n} = (1+\sqrt{\gamma})^2\quad \mbox{and}\quad
-\lim_{n\to \infty}\frac{\la_{(p)}}{n} = (1-\sqrt{\gamma})^2\quad \as \,
-\end{equation}
-The approach based on \eqref{eq:jo45} provides a lower bound on the smallest eigenvalue.
-Unfortunately, one cannot gain any information about the minimal conditions for the existence of a limit of $\la_{(p)}/n$ since the method treats $\la_{(1)}$ and $\la_{(p)}$ simultaneously.
-Therefore it can (at best) only be applied under $\E[X^4]<\infty$.
-Consequently, one loses sharpness of the conditions for the limit of $\la_{(p)}/n$.
-It was finally discovered by Tikhomirov \cite{tikhomirov:2015} that the \as~limit of $ \la_{(p)}/n$ is given by \eqref{eq:sdfgsdfdf} if $\E[X^2]=1$, whereas higher moments can be infinite.
-\par
-
-Under suitable moment assumptions, $\la_{(1)}$ and $\la_{(p)}$ possess {\em Tracy--Widom} fluctuations around their almost sure limits. For instance, 
-Johnstone \cite{johnstone:2001} complemented \eqref{eq:sdfgsdfdf} by the corresponding \clt\ in the special case of iid standard normal 
-entries:
-\begin{equation*}%\label{eq:john}
- n^{2/3}\,\dfrac{(\sqrt{\gamma})^{1/3}}{\big(1+\sqrt{\gamma}\big)^{4/3}}\Big(\dfrac {\la_{(1)}}{n} -
-\big(1+\sqrt{\tfrac pn }\big)^2\Big)
-\std \xi\,,
-\end{equation*}
-where the limiting \rv\ has a {\em Tracy--Widom \ds} of order~$1$.
-Its distribution function $G_1$ is given by
-\begin{equation*}
-G_1(s) = \exp\Big\{
-  -\frac{1}{2} \int_{s}^\infty [
-    q(x) + (x - s) q^2(x)
- ] \dint x
-\Big\}\,,
-\end{equation*}
-where $q(x)$ is the unique solution to the Painlev\'e II differential
-equation
-\begin{equation*}
-  q''(x) = xq(x) + 2 q^3(x)\,,
-\end{equation*}
-where $ q(x)\sim {\rm Ai}(x)$ as $x \to \infty$ and Ai$(\cdot)$ is the Airy kernel; see Tracy and Widom~\cite{tracy:widom:2012} for details.
-
-\subsection{Sample correlation matrices}
-
-In comparison with the eigenvalues of $\X\X'$, much less is known about the ordered eigenvalues
-\begin{equation*}
-\mu_{(1)} \ge \cdots \ge\mu_{(p)}\,
-\end{equation*} 
-of the {\em sample correlation matrix} $\bfR=\Y\Y'$ with entries
-\begin{equation}\label{eq:corrRch1}
-R_{ij}=\sum_{t=1}^n \frac{X_{it}X_{jt}}{\sqrt{D_i} \sqrt{D_j}} = \sum_{t=1}^n Y_{it}Y_{jt}\,, \quad i,j=1,\ldots,p\,.
-\end{equation}
-In this thesis we will often make use of the notation $\Y=(Y_{it})=(X_{it}/\sqrt{D_i})$ and
-\begin{equation}\label{eq:Dch1}
-D_i=D_i^{(n)}=\sum_{t=1}^n X_{it}^2\,, \qquad
-i=1,\ldots,p;\;  n\ge 1\,.
-\end{equation}
-
-With $\bfF= \diag (1/D_1,\ldots,1/D_p)$, we have $\bfR={\bfF}^{1/2} {\X}{\X}'{\bfF}^{1/2}$ which has the same eigenvalues as ${\X}{\X}'{\bfF}$. Weyl's inequality (see \cite{bhatia:1997}) yields
-\begin{equation}\label{eq:lamuch1}
-\begin{split}
-\max_{i=1,\ldots,p} |\mu_{(i)}- n^{-1} \la_{(i)}| &\le \twonorm{{\X}{\X}'{\bfF}-n^{-1}{\X}{\X}' }\\
-&\le n^{-1} \twonorm{{\X}{\X}'}  \twonorm{n \bfF- \bfI}\\
-&= n^{-1} \la_{(1)} \max_{i=1,\ldots,p} \Big| \frac{n}{D_i} -1 \Big|\,,
-\end{split}
-\end{equation}
-which converges \as~to $0$ if $\E[X^4]<\infty$; see Chapter~\ref{ch:corr} for details.
-Under \eqref{Cp} with $\gamma\in(0,1]$,  Jiang \cite{jiang:2004}, and Xiao and Zhou \cite{xiao:zhou:2010} used this approach to derive 
-\begin{equation*}
-\mu_{(1)} \to (1+\sqrt{\gamma})^2\, \quad \text{ and } \quad  \mu_{(p)} \to (1-\sqrt{\gamma})^2\, \quad \as
-\end{equation*}
-
-\section{Eigenvectors}
-
-Eigenvectors of large random matrices and graphs play an essential role in statistical analysis, physics and computer science.
-Many properties of a matrix or a graph can be deduced from its eigenvectors.
-Popular algorithms for data analysis such as spectral clustering, principal component analysis, PageRank and community detection are based on the eigenvector-eigenvalue decomposition of a matrix.
-
-If the data matrix $\X$ has iid standard normal entries, then $\X\X'$ is a {\em Wishart matrix}, whose eigenvectors are well studied; see for example Bai and Silverstein \cite[Ch.~10]{bai:silverstein:2010}.
-Due to the invariance of $\X\X'$ under orthogonal transformations, its matrix of properly normalized eigenvectors is Haar distributed, i.e., the distribution is uniform on the space of orthogonal $p\times p$ matrices.
-This result has been extended to different classes of matrices $\X$ by direct comparison with Wishart matrices.
-Such statements are called universality results and they often require that the new entry distribution is in some sense similar to the standard normal distribution.
-This is often achieved by moment conditions.
-Silverstin \cite{silverstein:1984} showed that the matrix of eigenvectors is asymptotically Haar distributed as $p/n\to \gamma\in(0,\infty)$ if the first four moments of the iid entries coincide with those of the standard normal distribution.
-On the one hand, this means that eigenvectors of $\X\X'$ are completely unstructured for light-tailed entry distributions.
-On the other hand, the extreme eigenvalues converge to constants \as
-\par
-
-Although eigenvectors play a minor role in this thesis, we summarize some results from the literature on the light-tailed case.
-Consequently, our approximations of eigenvectors in the heavy-tailed case are put into context.
-The majority of studies on eigenvectors of large random matrices is conducted on Wigner matrices $\bfW=(W_{ij})$.
-They are symmetric, real-valued, $n\times n$ matrices with entries $W_{ij}, 1\le i \le j \le n$, that are iid, mean zero and unit variance random variables.
-
-Roughly speaking, the spectral properties of the square of a Wigner matrix and a sample covariance matrix with $p=n$ are quite similar if they share the same sufficiently light-tailed entry distribution.
-For the purpose of exposition, we look at $\bfW^2=\bfW \bfW'$ instead of $\X\X'$.
-Indeed, apart from the additional symmetry restriction for $\bfW$ they have the same structure.
-
-By Theorem 2.1 in \cite{bai:fang:liang:2014}, the limiting spectral distribution of $(\bfW /\sqrt{n})$ is the {\em semi-circle law} $G$ with density
-\begin{equation}\label{eq:semi-circle}
-g(x)=\frac{1}{2\pi} \sqrt{4-x^2}\, \1_{\{ |x|\le 2 \}}\,.
-\end{equation}
-The semi-circle law and the \MP law $F_{\gamma}$ with $\gamma=1$ are linked in the following way: if $Y\sim G$ then $Y^2\sim F_1$.
-
-Since $\bfW^2$ and $\bfW$ have the same eigenvectors it will be sufficient to study the latter.
-In the remainder of this section we list some properties of $\bfv_1, \ldots, \bfv_n$, where $\bfv_i$ is an eigenvector associated to the $i$th largest eigenvalue of $\bfW$.
-In addition, we assume $\bfv_i$ are unit vectors, i.e.
-$\ltwonorm{\bfv_i}=1$, and that the first non-zero component of $\bfv_i$ is positive.
-
-Often the entries are assumed to be {\em sub-exponential}. We call a random variable $W$ {\em sub-exponential} with exponent $\alpha>0$ if there exists a constant $\beta>0$ such that 
-\begin{equation*}
-\P(|W|>x)\le \beta \exp(-x^\alpha /\beta)\,,\quad x>0\,.
-\end{equation*}
-
-Theorem 1.4 in O'Rourke et al.~\cite{orourke:vu:wang:2016} focuses on Wigner matrices with sub-exponential entries with exponent $\alpha$.
-There exists a constant $C_{\alpha}>0$ such that the probability that the spectrum of $\bfW$ is simple and that every coordinate of every $\bfv_i$ is non-zero is at least $1-C_{\alpha} n^{-\alpha}$.
-By our sign convention, the eigenvectors are unique with high probability.
-In particular, if $W_{11}$ is standard normal, $\bfv_i$ is uniformly distributed on
-\begin{equation*}
-\mathcal{S}_+^{n-1}:=\{ \x=(x_1,\ldots,x_n)': \ltwonorm{\x}=1 \text{ and } x_1>0 \}\,.
-\end{equation*}
-An eigenvector of a Wigner matrix with light-tailed entry distribution (for instance sub-exponential) behaves like a random vector uniformly distributed on $\mathcal{S}_+^{n-1}$.
-More precise quantitative statements are difficult to obtain.
-For details we refer to \cite{orourke:vu:wang:2016}.
-
-If $\bfv=(v_1,\ldots,v_n)'$ is a random vector uniformly distributed on $\mathcal{S}_+^{n-1}$, probabilistic bounds on its coordinates are available.
-By Theorem 2.1 in \cite{orourke:vu:wang:2016}, we have
-\begin{equation}\label{eq:coordinates}
-v_{\text{max}} := \max_{i=1,\ldots,n} |v_i| \le C\sqrt{\frac{\log n}{n}}\quad \text{ and } \quad v_{\text{min}} := \min_{i=1,\ldots,n} |v_i|\ge \frac{c}{n^{3/2}}
-\end{equation}
-with probability $1-o(1)$ for any $C>1$ and $c\in [0,1)$.
-
-In RMT it is common to study the so-called {\em bulk} and {\em edge} spectra separately.
-For $\vep\in (0,1)$ one distinguishes between the {\em bulk} eigenvectors $\bfv_i, i\in \{1\le t\le n: \vep n \le t \le (1-\vep) n\} :=B_\vep$ and the {\em edge} eigenvectors $\bfv_i, i \in \{1,\ldots,n\} \backslash B_\vep$.
-The associated eigenvalues are usually referred to as bulk and edge spectrum, respectively.
-Roughly speaking, the limiting spectral distribution of a sequence of random matrices depends on the bulk spectrum, while the edge spectrum influences the behavior of functionals of the eigenvalues.
-\par
-
-The behavior of the largest coordinates of $\bfv_1,\ldots, \bfv_n$ was studied in the case of certain sub-exponential entries with exponent $\alpha=2$ whose first four moments match those of the standard normal distribution.
-By Theorems 4.1 and 4.3 in \cite{orourke:vu:wang:2016}, one has for any bulk eigenvector $\bfv_i$ and appropriate constants $c_1,c_2>0$,
-\begin{equation*}
-c_1 \sqrt{\frac{\log n}{n}} \le \bfv_{i,\text{max}}\le c_2 \sqrt{\frac{\log n}{n}}
-\end{equation*}
-with probability $1-o(1)$.
-This result is astonishingly precise.
-Note that up to logarithmic corrections $\bfv_{i,\text{max}}$ is of the smallest possible magnitude $n^{-1/2}$.
-This property is called {\em complete delocalization}; see \cite{rudelson:vershynin:2015}.
-For edge eigenvectors, however, it is proved that for a constant $c_3>0$,
-\begin{equation*}
-\bfv_{i,\text{max}}\le c_3 \frac{\log n}{\sqrt{n}}\,
-\end{equation*}
-with probability $1-o(1)$.
-The optimal bound in the edge case remains an open problem.
-Additionally, by Corollary 5.4 in \cite{orourke:vu:wang:2016} the $\ell_p$-norms, $1\le p\le2$, of the $(\bfv_i)$ are of the same order of magnitude:
-\begin{equation}\label{eq:lpefgs}
-c_0 n^{1/p-1/2}\le \min_{i=1,\ldots,n} \| \bfv_i\|_{\ell_p} \le \max_{i=1,\ldots,n} \| \bfv_i\|_{\ell_p}  \le C_0 n^{1/p-1/2}
-\end{equation}
-with probability $1-o(1)$ for positive constants $c_0,C_0$.
-\par
-
-If the entry distribution of $\bfW$ has heavy tails, the behavior of its eigenvectors is completely different.
-In \cite{benaych:peche}, Benaych-Georges and P\'{e}ch\'{e} assumed $W_{11}$ to be regularly varying with index $\alpha\in (0,4)$; see \eqref{eq:regvarch1} on page \pageref{eq:regvarch1}, and provided an approximation of $\bfv_k$ for any fixed $k$.
-From their asymptotic result one can deduce that $\bfv_{k,\text{max}}$ converges to $1/\sqrt{2}$ in probability.
-Moreover, asymptotically there are only two coordinates of $\bfv_k$ with non-zero mass.
-Both are of magnitude $1/\sqrt{2}$.
-This is the opposite of complete delocalization: {\em complete localization}.
-The number of non-zero coordinates is bounded.
-
-Note that in the presence of heavy tails the eigenvectors of $\bfW$ and $\X\X'$ are very different; see \eqref{eq:eigenvec} in Section \ref{sec:contr}.
+Mason \cite{mason:1982} first proved that the estimator
+was consistent when $X_1, X_2, \dots$ were iid; later Rootz\'en,
+Leadbetter and de Haan \cite{rootzen:leadbetter:dehaan1992} and Hsing
+\cite{hsing:1991} proved its consistency when $X_1, X_2, \dots$ were
+weakly dependent; and Resnick and St{\u a}ric{\u a}
+\cite{resnick:starica:1995, resnick:starica:1997} proved its
+consistency when $X_t$ was a linear process.
 
@@
 
-In this section we summarize our contribution to the spectral theory of high-dimensional sample covariance and correlation matrices.
-We focus on the case where the entries of the $(p\times n)$-dimensional data matrix $\X$ have an infinite fourth moment.
+In this section we summarize our results from the research papers.
 
-\subsection{Sample covariance matrices: the iid case}
+\subsection{Tail parameters of equity return series}
+We have established that, in the case of an equity return series with
+two-sided, functionally independent Pareto tails, investor
+preference functionals are monotone increasing/decreasing with the
+tail index/scale parameters. Thus in a market dominated by such
+equities, the investors would pursue the largest tail index in the
+market, leading to a shared common tail index for all equities.
 
-While the finite second moment is the central assumption to obtain the Mar\v cenko--Pastur law as the limiting spectral distribution, the finite fourth moment plays a crucial role when studying the eigenvalues
-\beam\label{eq:orderch1}
-\la_{(1)}\ge \cdots \ge \la_{(p)}
-\eeam
-of the sample covariance matrix $\X\X'$. 
-Unless stated otherwise, the entries $(X_{it})$ are iid regularly varying random variables with index $\alpha\in (0,4)$ (see \eqref{eq:regvarch1}) and $X$ is a generic random variable with the same distribution.
-This implies $\E[X^4]=\infty$.
-Here and in what follows, we will refer to this setting as the heavy-tailed case, in contrast to the light-tailed case in which $\E[X^4]$ is finite.
-Moreover, we assume $\E[X]=0$ and $\E[X^2]=1$, whenever $\E[X^2]<\infty$.
+The empirical results presented in section \ref{sec:1} suggest this
+may well be the case for the ``Consumer Staples'' sector of S\&P 500,
+given the Hill estimates of tail indices shown in figure \ref{fig:1}
+and the largely positive results of tests for equal tail indices shown
+in figure \ref{fig:PairTest}.
 
-We normalize the eigenvalues $(\la_{(i)})$ by $(a_{np}^2)$ where the \seq\
-$(a_k)$ is chosen \sth 
-\beao
-\P(|X|>a_k)\sim k^{-1}\,,\quad k\to\infty.
-\eeao
-Standard theory for \regvary\ \fct s (e.g. Bingham et al. \cite{bingham:goldie:teugels:1987}, Feller \cite{feller}) yields that $a_n=n^{1/\alpha} \ell(n)$ where $\ell$ is a slowly varying function.
-Assuming the usual growth condition \eqref{Cp} for $p$,
-the Potter bounds (see \cite[p.~25]{bingham:goldie:teugels:1987}) yield for $\alpha\in (0,4)$ that
-\begin{equation}\label{eq:toinfch1}
- \frac{a_{np}^2}{n} \sim \frac{n^{4/\alpha} \gamma^{2/\alpha}\,\ell^2(n^2 \gamma)}{n} \to \infty, \qquad \nto\,,
-\end{equation}
-i.e., the normalization $a_{np}^2$ is stronger than $n$.
+On the other hand, we have also seen that, when the left and the right
+tails have the same indices, investor preference over the equity has
+more sophisticated variations in the parameters' space including the
+tail parameters of the equity, the interest rate, the investor's risk
+apetite as captured by his utility function, and his threshold of
+disappointment.
 
-By \eqref{eq:simplebo}, we have $\la_{(1)}\ge X_{(1),np}^2$, where $X_{(1),np}^2\ge \cdots \ge X_{(np),np}^2$
-denote the order statistics of $(X_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$. Classical extreme value theory yields that $a_{np}^{-2} X_{(1),np}^2$ converges weakly to a  
-{\em \Frechet distribution} with parameter~$\alpha/2$:
-\beam\label{eq:frechetch1}
-\Phi_{\alpha/2}(x) =\ex^{-x^{-\alpha/2}}\,,\qquad x>0\,.
-\eeam
-\par
-
-The theory for the largest eigenvalues of sample covariance matrices with heavy tails is less developed than in the light-tailed case.
-Pioneering work for $\la_{(1)}$ under the growth condition \eqref{Cp} and $\alpha\in (0,2)$
-is due to Soshnikov~\cite{soshnikov:2004,soshnikov:2006}. For $k\ge 1$ fixed, he showed that
-\begin{equation}\label{eq:sosh}
-\frac{\la_{(m)}}{X_{(m),np}^2}\cip 1\,,\quad \nto\,, \,1\le m\le k\,,
-\end{equation}
-which reveals that the limiting distribution of $a_{np}^{-2} X_{(1),np}^2$ is the \Frechet distribution \eqref{eq:frechetch1}. Furthermore he proved the \pp\ \con\
-\beam\label{eq:nnch1}
-N_n=\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i} \std N_{\Gamma}=\sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,.
-\eeam
-Here $\vep_y$ is the Dirac measure at $y$,
-\beam\label{eq:Gammach1}
-\Gamma_i=E_1+\cdots + E_i\,,\qquad i\ge 1\,,
-\eeam
-and $(E_i)$ is a sequence of iid standard exponential random variables. In other words, $N_{\Gamma}$ is a Poisson point process on $(0,\infty)$ with mean \ms\
-$\mu(x,\infty)= x^{-\alpha/2}$, \mbox{$x>0$}. Convergence in \ds\ of \pp es is understood in the sense of weak \con\
-in the space of point \ms s equipped with the vague topology; see Resnick \cite{resnick:2007,resnick:1987}.
-
-Later Auffinger et al.~\cite{auffinger:arous:peche:2009} established \eqref{eq:nnch1} also for $\alpha \in [2,4)$. In their proofs they used truncation techniques and a combinatorial approach.
-
-\subsubsection*{General growth rates for $p_n$}
-
-In many applications it is not realistic to assume
-that the dimension $p$ of the data and the sample size $n$ grow at the same rate.
-In the light-tailed case little is known when $p$ and $n$ grow at different rates, i.e., $\lim p/n \in\{ 0,\infty\}$. 
-Notable exceptions are El Karoui \cite{elkaroui:2003} who proved that Johnstone's result in \cite{johnstone:2001} (assuming iid standard normal entries)
-remains valid when $p/n\to 0$ or $n/p\to\infty$, and P{\'e}ch{\'e} \cite{peche:2009} who showed universality results for the largest eigenvalues of some sample covariance matrices with non-Gaussian entries.
-
-The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009} already indicate that the value $\gamma$ in the usual growth
-rate~\eqref{Cp} does not appear in the \ds al limits.
-In the heavy-tailed case, more general growth of $p$ than prescribed by \eqref{Cp} has been used in Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
-In Chapters \ref{ch:iid} and \ref{ch:extremes} we will consider power-law growth rates on the dimension $(p_n)$. To be precise, we assume an integer sequence
-\begin{equation}\label{eq:pch1}
-p=p_n=n^\beta l(n), \quad n\ge1\,,%\tag{$C_p(\beta)$}
-\end{equation}
-where $l$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume \mbox{$l(n) \to \infty$}.
-Our condition~\eqref{eq:pch1} is more general than the growth conditions in the literature; see
-\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
-
-Note that the matrices $\X\X'$ and $\X'\X$ have the same non-zero eigenvalues.
-Therefore it is sufficient to consider $\beta\in [0,1]$.
-For details we refer to Chapters \ref{ch:iid} and \ref{ch:extremes}.
-
-\subsubsection*{Our contribution}
-
-%It is interesting to note that there is a phase change of the behavior of the extreme eigenvalues when going from finite to infinite fourth moment, while a similar phase change occurs for the empirical spectral distribution when going from finite to infinite variance.
-
-In the heavy-tailed case and under the growth condition \eqref{eq:pch1} with $\beta \in [0,1]$ we prove with considerable technical effort that  
-\begin{equation}\label{eq:offdiagonalch1}
-a_{np}^{-2} \twonorm{\X \X' - \diag(\X \X')}\stp 0\,,\qquad\nto\,,
-\end{equation}
-where $\diag(\X \X')$ denotes the diagonal of $\X\X'$; see Theorem \ref{prop:offdiagonal}.
-
-The employed techniques originate from extreme value analysis and large deviation theory. The $(i,j)$ entry of $\X\X'$ is 
-\begin{equation*}
-(\X\X')_{ij}=\sum_{t=1}^n X_{it}X_{jt}\,,\quad i,j=1\ldots,p\,.
-\end{equation*}
-From Embrechts and Veraverbeke \cite{embrechts:veraverbeke:1982} we know that  $X^2$ and $X_{11}X_{12}$  are \regvary\ with indices $\alpha/2$ and $\alpha$, respectively.
-By large deviation theory (see \eqref{eq:dfsdfjl}), the diagonal and off-diagonal entries of
-$\X\X'$ inherit the tails of $X_{it}^2$ and $X_{it}X_{jt}$, $i\ne j$, respectively, above some high threshold.
-Therefore the random variables in the diagonal of $\X\X'$ have the heaviest tail.
-They dominate the spectral behavior of $\X\X'$ and thus \eqref{eq:offdiagonalch1} is not unexpected.
-\par
-
-Equation \eqref{eq:offdiagonalch1} has some immediate con\seq s for the approximation of the eigenvalues
-of $\X\X'$ by those of ${\rm diag}(\X\X')$. Indeed, let $\bfC$ be a symmetric $p\times p$ matrix with
-eigenvalues 
-\beam \label{eq:weylch1}
-\la_{(1)}(\bfC)\ge \cdots \ge \la_{(p)}(\bfC)\,.
-\eeam
-Then for any symmetric $p\times p$ matrices $\bfA,\bfB$, by {\em Weyl's inequality} (see Bhatia \cite{bhatia:1997}),
-\beao
-\max_{i=1,\ldots,p}\big|\la_{(i)}(\bfA+\bfB)-\la_{(i)}(\bfA)\big|\le \|\bfB\|_2\,.
-\eeao
-If we now choose
-$\bfA+\bfB=\X\X'$ and $\bfA= \diag (\X\X')$ we obtain
-\begin{equation}\label{eq:appch1}
-a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\diag(\X\X'))\big|\stp 0\,,\quad\nto \,.
-\end{equation}
-
-Thus the problem of deriving limit theory for the order statistics of $\X\X'$ has been reduced to limit theory for the order
-statistics of the iid row-sums
-\beao
-D_i= (\X\X')_{ii}=\sum_{t=1}^n X_{it}^2\,,\qquad i=1,\ldots,p\,,
-\eeao
-which are the eigenvalues of $\diag(\X\X')$. 
-This theory is completely described by the \pp es constructed from the points $D_i/a_{np}^2$, $i=1,\ldots,p$. Necessary
-and sufficient conditions for the weak \con\ of these \pp es are provided by  Lemma~\ref{lem:pprch3}.
-In combination with the Nagaev-type \ld\ results of  Theorem~\ref{thm:nagaevch3} they yield the following result under \eqref{eq:pch1}:
-\begin{equation}\label{eq:ppdiagch1}
-\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i-c_n)}\std N_{\Gamma}\,,\qquad \nto\,,
-\end{equation}
-where $N_{\Gamma}$ was defined in \eqref{eq:nnch1} and  $c_n=0$ if $\E[D]=\infty$ and $c_n=\E[D]=n \E[Z^2]$ otherwise.
-Note that the centering $c_n$ in the finite variance case can be avoided if $n/a_{np}^2\to 0$. The latter condition is satisfied if
-\begin{equation}\label{eq:b}
-\beta>\alpha/2-1\,.
-\end{equation}
-
-Combining \eqref{eq:offdiagonalch1}, \eqref{eq:appch1} and \eqref{eq:ppdiagch1}, we conclude that \eqref{eq:nnch1} holds under the general growth rate \eqref{eq:pch1} with $\beta\in [0,1]$, where for $\alpha\in [2,4)$ one needs to require \eqref{eq:b}.
-
-The limiting point process \eqref{eq:nnch1} yields a plethora of ancillary results. For example, one can easily derive the limiting distribution of $a_{np}^{-2} \lambda_{(k)}$ for fixed $k\ge 1$:
-\begin{equation*}
-\begin{split}
-\lim_{\nto}\P(a_{np}^{-2} \lambda_{(k)}\le x)&= \lim_{\nto}\P(N_n(x,\infty)<k)
-=  \P(N(x,\infty)<k)\\ &=\P(\Gamma_k^{-2/\alpha}\le x)= \sum_{s=0}^{k-1} \frac{\big(x^{-\alpha/2}\big)^s}{s!} \e^{-x^{-\alpha/2}}, \quad  x>0\,.
-\end{split}
-\end{equation*}
-Another immediate con\seq\ of \eqref{eq:nnch1} is
-\beam \label{eq:helpgamma1ch1}
-a_{np}^{-2}\big(\la_{(1)},\ldots,\la_{(k)}\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)
-\eeam
-for any fixed $k\ge 1$ and $\alpha\in (0,2]$. In Chapter \ref{ch:extremes} we show for $\alpha \in (2,4)$
-\beam \label{eq:helpgamma2ch1}
-a_{np}^{-2}\big(\la_{(1)}-n \,\E[Z^2],\ldots,\la_{(k)}-n\, \E[Z^2]\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,.
-\eeam
-Equations \eqref{eq:helpgamma1ch1} and \eqref{eq:helpgamma2ch1} yield that for $\alpha \in (0,4)$ and any fixed $k\ge 1$,
-\beam \label{eq:helpgamma3ch1}
-a_{np}^{-2}\big(\la_{(1)}\!-\!
-\la_{(2)},\ldots, \la_{(k)}\!- \!\la_{(k+1)}\big) \!\std \!\big(\Gamma_1^{-2/\alpha} \!- \!
-\Gamma_{2}^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha} \!
-- \!
-\Gamma_{k+1}^{-2/\alpha} \big).
-\eeam
-Related results were also derived for linear spectral statistics such as the trace $a_{np}^{-2}(\la_1+\cdots+\la_p)$. 
-We refer to Chapter \ref{ch:extremes} and Davis et al.~\cite{davis:mikosch:pfaffel:2015} for details on the proofs and more examples.
-\medskip
-
-In the case of  fixed $p$, Janssen et al.~\cite{janssen:mikosch:rezapour:xie:2016} related the limiting distribution of the eigenvalues $(\la_{(i)})$ to stable distributions.
-They also used \eqref{eq:offdiagonalch1} and \eqref{eq:appch1}.
-In this case it is clear that for example
-\begin{equation*}
-\frac{\la_{(2)}}{X_{(2),np}^2}\cip 1\,,\quad \nto\,, 
-\end{equation*}
-cannot hold.
-If $X_{(1),np}^2$ and $X_{(2),np}^2$ lie in the same row of $\X$, then they appear on the same spot on the diagonal of $\X\X'$.
-Then $X_{(2),np}^2$ cannot be used for the approxiamtion of $\la_{(2)}$ in view of \eqref{eq:appch1}.
-Indeed, the probability that this happens is approximately $1/p$ which does not tend to $0$ if $p$ is fixed.
-This is in contrast to \eqref{eq:sosh}.
-
-Recall that if $p$ is fixed, one has to use the approximation of the eigenvalues provided by \eqref{eq:appch1}, while if $p/n \to \gamma$ one can use either \eqref{eq:appch1} or \eqref{eq:sosh}.
-There is a phase change of the behavior of $(\la_{(i)})$ when going from finite $p$ to $p$ proportional to $n$.
-In our condition \eqref{eq:pch1} the growth of $p$ is essentially described by the parameter $\beta$.
-
-Under \eqref{eq:pch1} our Theorem \ref{thm:iidmain} asserts that 
-\beam\label{eq:eigch1}
-a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-X_{(i),np}^2\big|\stp 0\,,
-\eeam
-provided \eqref{eq:b} holds.
-Due to its simplicity, \eqref{eq:eigch1} is an elegant result.
-It reveals, for example, that the largest eigenvalue of a high-dimensional heavy-tailed matrix behaves like the maximum of its iid entries.
-
-In view of Lemma \ref{thm:keyresult}, condition \eqref{eq:b} describes the precise $\beta$-region, up to the slowly varying function in the tail of $X$, where $\max_i a_{np}^{-2}|D_{L_i}-X_{(i),np}^2|$ is sufficiently small.
-Here $L_i$ encodes the location of the $i$th largest diagonal element of $\X\X'$; see \eqref{eq:help6} for the formal definition of $L_i$.
-Therefore the critical value of $\beta$ at which the aforementioned phase change occurs is  $\max(0,\alpha/2-1)$.
-\medskip
-
-The study of eigenvectors of heavy-tailed sample covariance matrices is a fresh topic which has not been explored in the literature listed here.
-Our approximation of $\X\X'$ in \eqref{eq:offdiagonalch1} and the limiting distributions of the spacings \eqref{eq:helpgamma3ch1} can be applied to approximate the unit eigenvectors $(\bfv_j)$ of $\X\X'$, where $\bfv_{{j}}$ is associated to $\la_{(j)}$.
-As for the eigenvectors of Wigner matrices we use the convention that their first non-zero coordinate is positive.
-From \eqref{eq:offdiagonalch1} we know that $\X\X'$ is approximated in spectral norm by $\diag(\X\X')$.
-The unit eigenvectors of $\diag(\X\X')$ are the canonical basis vectors $\bfe_j\in\R^p$, $j=1,\dots,p$.
-
-By Theorem \ref{thm:eigenvec}, $(\bfe_j)$ approximate the eigenvectors $(\bfv_j)$.
-For $\beta\in [0,1]$ and any fixed $k\ge 1$, we have
-\begin{equation}\label{eq:eigenvec}
-\ltwonorm{\bfv_k - \bfe_{L_k}} \cip 0\,, \quad \nto \,.
-\end{equation}
-
-\subsection{Sample covariance matrices: the non-iid case}
-
-Davis et al.~\cite{davis:pfaffel:stelzer:2014} extended the results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009}
-to the case where the rows of $\X$
-are iid linear processes with iid \regvary\ noise.
-After a multiplication of the mean measure $\mu$ by a constant the Poisson \pp\ \con\ result of \eqref{eq:nnch1} remains valid.
-Pfaffel and Schlemm \cite{pfaffel:schlemm:2011} described the Stieltjes transform of the limiting spectral distribution in this model.
-Different limit processes can only be expected if there is dependence in both directions:
-in Chapter~\ref{ch:extremes} we use a model for $(X_{it})$ which allows for linear dependence across the rows and through time (see also \cite{davis:mikosch:pfaffel:2015}): 
-\begin{equation}\label{eq:1ch1}
-X_{it}=\sum_{l\in \Z}\sum_{k\in \Z} h_{kl} Z_{i-k,t-l}\,,\qquad i,t\in\Z\,,
-\end{equation}
-where $(Z_{it})_{i,t\in \Z}$ is a field of iid regularly varying \rv s with index $\alpha\in (0,4)$ and $(h_{kl})_{k,l\in\Z}$ is an array of real numbers.
-Moreover, we require the summability condition
-\begin{equation}\label{eq:2ach1}
-\sum_{l \in \Z} \sum_{k\in \Z} |h_{kl}|^{\delta} <\infty
-\end{equation}
-for some $\delta\in (0,\min({\alpha/2},1))$ which ensures
-the a.s.~absolute convergence of the series in \eqref{eq:1ch1}. Under the condition \eqref{eq:2ach1}, the marginal and
-\fidi s of the field $(X_{it})$ are \regvary\ with index $\alpha$; see
-Embrechts et al. \cite{embrechts:kluppelberg:mikosch:1997}, Appendix A3.3.
-
-From the field $(X_{it})$ we construct the $p\times n$ matrices
-\begin{equation*}
-\X_n(s)= (X_{i,t+s})_{i=1,\ldots,p;t=1,\ldots,n}\,,\qquad s=0,1,2,\ldots\,,
-\end{equation*}
-As before, we will write $\X=\X_n(0)$.
-Now we can introduce the (non-normalized)
-{\em sample autocovariance matrices}
-\beao%\label{eq:samplech1}
-\X_n(0)\X_n(s)'\,,\qquad s=0,1,2,\ldots\,.
-\eeao
-We will refer to $s$ as the {\em lag}. For $s=0$, we obtain the {\em sample covariance matrix.}
-In \cite{heiny:mikosch:2016:noniid1} (see page \pageref{eq:regvarch1}) and Chapter \ref{ch:extremes}, we study the asymptotic behavior of the eigen- and singular values of the
-sample covariance and autocovariance matrices under the growth condition \eqref{eq:pch1}.
-
-Theorem~\ref{thm:mains} provides a general approximation result for the ordered singular values of $\X_n(0)\X_n(s)'$.
-Their behavior is determined by the sums $(\sum_{t=1}^n Z_{it}^2)$ and the singular values of the matrix $\M(s)$ given by
-\begin{equation*}
-(\M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z \,.
-\end{equation*}
-In Section \ref{sec:lkasfj} the limiting point process of the singular values of $\X_n(0)\X_n(s)'$ is derived. 
-Finally, we mention that our paper \cite{heiny:mikosch:2016:noniid1} deals with the eigenvectors of $\X_n(0)\X_n(s)'$; see also \eqref{eq:eigenvec}.
-One obtains more interesting structures than in the latter result.
-In fact, by choosing $(h_{kl})$ accordingly one can obtain arbitrary eigenvectors.
-This constitutes a valuable property in principal component analysis.
-
-\subsection{Sample correlation matrices}
-
-In Chapter \ref{ch:corr} we study the spectrum of the sample correlation matrix $\bfR$ defined in \eqref{eq:corrRch1}.
-We assume that the underlying data matrix $\X$ has iid centered entries and the usual growth condition \eqref{Cp} holds.
-Recall the notation $Y_{it}=X_{it}/\sqrt{D_i}$ from p.~\pageref{eq:corrRch1}.
-We will sometimes write $(Y_1, \ldots, Y_n)= (Y_{11}, \ldots, Y_{1n})$ and $Y=Y_1$.
-\par
-
-Consider the following domain of attraction type-condition for the Gaussian law:
-\beam\label{eq:condXch1}
-\E \big[ Y_{1}Y_{2} \big] = o(n^{-2}) \quad  \text{ and } \quad  \E \big[ Y^4 \big] = o(n^{-1})\,,\qquad \nto\,.
-\eeam
-By Gin{\'e} et al.~\cite{gine:goetze:mason:1997}, condition \eqref{eq:condXch1} holds if the \ds\ of 
-$X$ is in the domain of attraction of the normal law.
-We use Theorem \ref{them:baizhou} to  show that under \eqref{eq:condXch1} the sequence $(F_{\bfR})$ converges weakly to the \MP law $F_{\gamma}$ defined in \eqref{eq:MPch1}; see Theorem \ref{thm:mpcorrelation}.
-We prove that the condition \eqref{eq:condXch1} is necessary.
-When \eqref{eq:condXch1} is not valid, the limiting spectral distribution of $(F_{\bfR})$ (if it exists) must have mean $1$, by virtue of the method of moments (see \eqref{eq:mmom}).
-This follows from the fact that the diagonal elements of $\bfR$ are $1$.
-This together with our approximation of $\E [\beta_k(\bfR)]$ provides some information about this distribution; compare also with \eqref{eq:esd} for the sample covariance case.
-\par
-
-Our analysis of the almost sure convergence of the extreme eigenvalues $\mu_{(1)}$ and $\mu_{(p)}$ of $\bfR$ is carried out for symmetric $X$. Then condition \eqref{eq:condXch1} turns into 
-\beam\label{eq:condXb}
-n \, \E \big[ Y^4 \big] \to 0\,,\qquad \nto\,.
-\eeam
-Theorem \ref{thm:mu1convergence} asserts
-\begin{equation}\label{eq:drtgdrghdr1ch1}
-\mu_{(1)} \to (1+\sqrt{\gamma})^2\, \quad \text{ and } \quad  \mu_{(p)} \to (1-\sqrt{\gamma})^2\, \quad \as
-\end{equation}
-under some condition \eqref{eq:assumptionq} on p.~\pageref{eq:assumptionq} which is slightly more restrictive than \eqref{eq:condXb}.
-Condition \eqref{eq:assumptionq} essentially means that the convergence rate of $n \, \E \big[ Y^4 \big]$ is at least logarithmic.
-A detailed discussion is given in Section \ref{sec:2ch4}.
-\par
-
-Our proof requires an adequate bound on $\E[\mu_{(1)}^{k_n}]$, where $k_n\to \infty$. To this end, we use the inequality
-\begin{equation*}
-\E[\mu_{(1)}^{k_n}]\le \E[\tr \bfR^{k_n}] %=p \E\big[\int x^{k_n} F_{\bfR}(dx)\big]
-= \sum_{i_1,\ldots,i_{k_n}=1}^p \sum_{t_1,\ldots,t_{k_n}=1}^n \E[ Y_{i_1t_{k_n}} Y_{i_1t_1}  \cdots Y_{i_{k_n}t_{{k_n}-1}} Y_{i_{k_n}t_{k_n}}  ]
-\end{equation*}
-and determine those summands on the \rhs~which are largest when weighted by their multiplicities.
-Employing our {\em Path-Shortening Algorithm}, which is a novel technique that efficiently uses the inherent structure of sample correlation matrices, their contribution is calculated explicitly.
-The other summands can --with considerable technical effort-- be controlled by \eqref{eq:assumptionq}.
-Note that because of the identity $\E[\tr \bfR^{k_n}] =p\, \E\big[\int x^{k_n} F_{\bfR}(dx)\big]$ the behavior of the moments of the empirical spectral distribution is closely linked to the above upper bound.
-
-Equation \eqref{eq:drtgdrghdr1ch1} indicates that the a.s. convergence of the extreme eigenvalues of $\bfR$
-does not depend on the finiteness of the fourth or even second moments. 
-This is in stark contrast to the a.s. behavior of $n^{-1}\lambda_{(1)}$, the largest eigenvalue of the sample covariance matrix 
-$n^{-1}\bfX\bfX'$. Note that there is a phase transition of the a.s. \asy\ behavior of the  
-extreme eigenvalues at the border between finite and infinite fourth moment of $X$, 
-while such a transition occurs for the empirical spectral distribution at the border between finite and infinite variance.
-\par
-
-The eigenvalues of sample correlation matrices exhibit a  ``more robust'' behavior than their sample covariance analogs. 
-This is perhaps not surprising in view of the {\em self-normalizing property} of sample correlations.
-Self-normalization also has the advantage that one does not have to worry about the correct normalization.
-This is a crucial problem in the study of sample covariance matrices in the case of an infinite fourth moment 
-where one needs a normalization stronger than the classical one; see \eqref{eq:toinfch1}.
-A simulation study in Section~\ref{sec:3} shows that the \asy\ results for $\mu_{(1)}, \mu_{(p)}, \la_{(1)}$ and $\la_{(p)}$ work nicely.
-They can be used to design new statistical tests; see for example~\eqref{eq:forch1}.
-
-\section{Outlook}
-
-The main objective of our work was to find explicit limiting distributions of the eigenvalues of large random matrices and functionals thereof.
-Thus our theory can be applied in a straightforward way.
-Sections \ref{sec:rmt}-\ref{sec:contr} listed fields where our results can be used.
-Another example is the analysis of high-frequency data which receives significant interest; see for example Podolskij and Heinrich \cite{heinrich:podolskij:2014}, and  Xia and Zheng \cite{xia:zheng:2014,xia:zheng:2016}.
-
-For practical purposes it is important to work with arbitrary population covariance matrices.
-Numerous generalizations and estimation techniques have been developed.
-For many models the limiting spectral distribution can only be characterized in terms of an integral equation (=\MP equation) for its Stieltjes transform.
-Explicit solutions are more involved; see the discussion after Theorem \ref{them:baizhou}.
-In Dobriban \cite{dobriban:2015}, an algorithm for calculating the spectral distribution based on certain approximate integral equations for its Stieltjes transform was presented.
-Contributions like this one breathe life into abstract theoretical results.
-Research in this direction will attract major interest from the industry.
-
-From a more theoretical point of view, it is interesting to study models with heavy tails in which the asymptotic behavior of the sample covariance matrix $\X\X'$ is not dominated by the squares of the entries of $\X$.
-Moreover, different tail indices of the rows of $\X$ could make the model more appealing to practitioners.
-In Chapter \ref{ch:extremes}, we will see that a simple transformation, such as the rank transform, does not entirely overcome this issue.
-
-As regards the sample correlation matrix our methods can be applied to more general models such as spiked covariance/correlation structures.
-Our technical results in Section~\ref{sec:5.3} are of independent interest.
-They provide a {\em Path-Shortening Algorithm} for the calculation of bounds for the very high moments of
-$\mu_{(1)}$. This technique is novel and will be of further use for proving results
-in Random Matrix Theory.
-We also conjecture that \eqref{eq:drtgdrghdr1ch1} may be proved under \eqref{eq:condXch1} only.
-
-Finally, in our working papers \cite{heiny:mikosch:2016:noniid1} and \cite{heiny:mikosch:2016:stochvol1} (see page \pageref{eq:regvarch1}), we analyze  sample autocovariance matrices and provide limit theory for high-dimensional stochastic
-volatility matrices.
-Again, we utilize the large deviations approach propagated in Chapters \ref{ch:iid} and \ref{ch:extremes}, now for dependent heavy-tailed time series.
-The corresponding large deviations theory is available in Mikosch and Wintenberger \cite{mikosch:wintenberger:2016}.
+We also acknowledge that our model of the market and the investor is a
+simple one, not accoounting for the dependence between equities, nor
+the categorization of investors and their interactions. These are
+potential topics of future work.

### 2017-07-21 phd/MyThesis.tex
Added my own thesis.

@@
 \input{Preamble}
+\usepackage[numbers,sort&compress]{natbib}
+
+%% \usepackage[toc,page]{appendix}
+\usepackage{appendix}
+\graphicspath{
+        {../papers/FX/}{../papers/Jeffrey1/}
+        {../papers/DavisMikoschHeinyXie/}
+}
 
 \begin{document}
-
-%\includepdf{delete}
 
@@
 
-This thesis is concerned with asymptotic properties of the eigenvalues of high-dimensional sample covariance and correlation matrices under an infinite fourth moment of the entries.
-
-In the first part, we study the joint distributional convergence of the largest eigenvalues of the sample covariance matrix of a $p$-dimensional heavy-tailed time series when $p$ converges to infinity together with the sample size $n$.
-We generalize the growth rates of $p$ existing in the literature.
-Assuming a regular variation condition with tail index $\alpha<4$, we employ a large deviations approach to show that the extreme eigenvalues are essentially determined by the extreme order statistics from an array of iid random variables.
-The asymptotic behavior of the extreme eigenvalues is then derived routinely from classical extreme value theory.
-The resulting approximations are strikingly simple considering the high dimension of the problem at hand.
-
-We develop a theory for the point process of the normalized eigenvalues
-of the sample covariance matrix in the case where rows and columns of the data are linearly dependent.
-Based on the weak convergence of this point process we derive the limit laws of various functionals of the eigenvalues.
-
-In the second part, we show that the largest and smallest eigenvalues 
-of a high-dimensional sample correlation matrix possess almost sure non-random limits if the truncated variance of the entry distribution is ``almost slowly varying'', a condition we describe via moment properties of self-normalized sums.
-We compare the behavior of the eigenvalues of the sample covariance and sample correlation matrices and argue that the latter seems more robust, in particular in the case of infinite fourth moment.
+This thesis is about vanity of life.
 
 \section*{Resum{\'e}}
-%\addcontentsline{toc}{section}{Resum{\'e}}
-
-%Fokus for afhandlingen ligger i / Afhandlingen fokuserer på 
-Denne afhandling beskæftiger sig med de asymptotiske egenskaber af egenværdierne for højdimensionale empiriske korrelations- og kovariansmatricer, under antagelse af at matrixindgangene har uendeligt fjerde moment.
-
-I første del undersøger vi konvergens i fordeling af de 
-%principiale? 
-største egenværdier for den observerede kovarians af en $p$-dimensional tunghalet tidsrække, når
-%/hvor 
-$p$ sammen med stikprøvestørrelsen går mod uendelig.  
-Vi generaliserer de eksisterende vækstrater 
-af $p$ 
-fra litteraturen.  Under antagelse af regulær variation med haleindeks $\alpha < 4$, bruger vi en 
-%"store afvigelser?"
-large deviations-tilgang for at vise at de ekstremale egenværdier er essentielt bestemt ud fra de største værdier i et array af iid. stokastiske variable.
-Herefter udleder vi rutinemæssigt de ekstremale egenværdiers asymptotiske egenskaber ved brug af klassisk 
-%ekstremal
-ekstremværditeori.
-
-Vi fremfører en teori
-%? 
-for punktprocesser af de normalisererde egenværdier af den empiriske kovariansmatrix i tilfældet, hvor dens rækker og søjler er afhængige.
-Med udgangspunkt i, at denne punktproces konverger svagt, udleder vi grænsemomenter af
-%diverse?  
-forskellige funktionaler af egenværdierne.
-
-I den anden del af afhandlingen viser vi at de største og mindste egenværdier fra en højdimensional korrelationsmatrix har næsten sikre grænser, hvis den trunkerede varians af indgangsfordelingen er ,,næsten langsom varierende``, en betingelse som kan indentificeres ud fra momentegenskaber af selv-normaliserende summer.
-Vi sammenligner, hvordan egenværdierne for
-%hhv.
-henholdsvis den empiriske korrelations- og den empiriske kovariansmatrix opfører sig, og argumenterer for at sidstnævnte tilfælde er mere robust, hvilket særligt gælder i tilfældet med uendeligt fjerde moment.
 
@@
 
-First and foremost, I would like to thank my supervisor Thomas Mikosch for many fascinating discussions, his guidance and experience.
-I could not have wished for a better person to work with and learn from.
-Furthermore I am grateful for the support I received from my co-supervisor Olivier Wintenberger.
-
-I want to thank Richard Davis and the Statistics Department of Columbia University in the City of New York, where I spent a research visit in $2016$, for their great hospitality.
-The group meetings on Wednesdays led to a vivid exchange of ideas.
-I would also like to thank Claudia Kl\"uppelberg. Her advice is highly appreciated.
-
-Next, I want to thank Gennady Samorodnitsky and Mark Podolskij for inspiring discussions. I am also grateful to Jeffrey Collamore, Jesper Lund Pedersen, Mogens Steffensen and Jostein Paulsen.
-
-It is my special pleasure to thank our lunch group at the University of Copenhagen which was an integral part of daily university life.
-I also want to thank my fellow PhD students at the University of Copenhagen and Columbia University for many fun hours.
-Special thanks go to all my office mates and in particular Xie Xiaolei.
-
-I acknowledge the support by Danish Research Council Grant DFF-4002-00435 ``Large random matrices with heavy tails and dependence''.
-
-Last but not least, I want to thank my family for the unconditional love and the invaluable support I have received over the years.
-
+Thank you all for various things.
 \bigskip
 
-{\em Johannes Heiny \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}
+{\em Xie Xiaolei\qquad \qquad \qquad \qquad \qquad Copenhagen, September $2017$.}
 
@@
 \addcontentsline{toc}{section}{Summary}
-
-This PhD thesis provides asymptotic theory for the eigenvalues of the sample covariance matrix of a high-dimensional time series with infinite fourth moment.
-Its main part consists of the first three of the following research papers.
-They were written from March $2014$ until December $2016$.
+This PhD thesis provides results from analysis of heavy-tailed time
+series. Its main part consists of the following research papers
+written between October $2014$ and Augest $2017$.
 
 \begin{enumerate}
-\bibitem[P1]{heiny:mikosch:2015:iid1}
-{\sc Heiny, J., and Mikosch, T.}
-\newblock Eigenvalues and eigenvectors of heavy-tailed sample covariance
-  matrices with general growth rates: the iid case.
-\newblock {\em Stochastic Process. Appl.\/} (2016), 29. \href{http://www.sciencedirect.com/science/article/pii/S0304414916301934}{\blue{[pdf]}}
-
 \bibitem[P2]{davis:mikosch:heiny:xie:20151}
@@
 \newblock {\em Extremes 19}, 3 (2016), 517--547. \href{http://link.springer.com/article/10.1007/s10687-016-0251-7}{\blue{[pdf]}}
-
-\bibitem[P3]{heiny:mikosch:2016:corr11}
-{\sc Heiny, J., and Mikosch, T.}
-\newblock Almost sure convergence of the largest and smallest eigenvalues of high-dimensional sample correlation matrices under infinite fourth moment.
-\newblock {\em Submitted for publication}.
-
-\bibitem[P4]{heiny:mikosch:2016:noniid1}
-{\sc Heiny, J., and Mikosch, T.}
-\newblock Limit theory for the singular values of the sample autocovariance
-  matrix function of multivariate time series.
-\newblock {\em In preparation}.
-
-\bibitem[P5]{heiny:mikosch:2016:stochvol1}
-{\sc Heiny, J., and Mikosch, T.}
-\newblock Asymptotic theory for high-dimensional stochastic volatility matrices.
-\newblock {\em In preparation}.
 \end{enumerate}
 
-The aforementioned infinite fourth moment is ensured by a {\em regular variation condition}. We say that a random variable $X$ and its distribution are {\em regularly varying with index} $\alpha>0$ if
-\begin{equation}\label{eq:regvarch1}
-\P(X>x)\sim p_+\,\dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad \P(X<-x) \sim p_-\,\dfrac{L(x)}{x^\alpha}\,,\qquad \xto\,,\tag{$RV_{\alpha}$}
-\end{equation}
-where $ p_\pm$ are non-negative constants \sth\ $p_++p_-=1$ and $L$ is a \slvary\ \fct .
-In particular, if $\alpha<4$ we have $\E [X^4]= \infty$.
-The \regvar\ condition is needed for proving \asy\ theory for the eigenvalues of the sample covariance matrix.
-Moreover, we will often use the concept of {\em heavy tails}.
-A distribution is called {\em heavy-tailed} if certain moments are infinite.
-By construction, any regularly varying distribution is heavy-tailed.
-\par
+Now we explain the structure of this thesis.
 
-Now we explain the structure of this thesis.
-In Chapter \ref{ch:intr} we provide an introduction to Random Matrix Theory and present the classical results in the light-tailed case.
-We give examples of high-dimensional statistical inference problems and indicate how the asymptotic theory applies.
-The empirical distribution of the eigenvalues of the widely used sample covariance matrix is studied.
-Furthermore the \as~limits of its largest and smallest eigenvalues are identified under finite fourth moment.
-Then we explain the typical behavior of its eigenvectors.
-{\em Section \ref{sec:contr} constitutes the main part of Chapter~\ref{ch:intr}.} There we present the contribution of this thesis and the novelties of \cite{heiny:mikosch:2015:iid1, davis:mikosch:heiny:xie:20151, heiny:mikosch:2016:corr11}.
-
-Chapters \ref{ch:iid}-\ref{ch:corr} consist of these $3$ papers, respectively.
-Each chapter is self-contained with its own introduction and references.
-Chapter~\ref{ch:intr} can be used as a joint introduction to Chapters \ref{ch:iid}-\ref{ch:corr}.
-\par
-In Chapter~\ref{ch:iid} we study the joint \ds al \con\ of the largest eigenvalues of the sample covariance matrix
-of a $p$-dimensional \ts\ with iid entries when $p$ converges to infinity together with the sample size $n$.
-We generalize the growth rates of $p$ in the literature.
-Assuming the regular variation condition with $\alpha<4$, we employ a large deviations approach to show
-that only the diagonal of the sample covariance matrix is relevant for the \asy\ behavior of the largest eigenvalues and the corresponding eigenvectors.
-The resulting approximations are strikingly simple considering the high dimension of the problem at hand.
-
-In Chapter~\ref{ch:extremes} we generalize the results from the iid case that were presented in the previous chapter. 
-We develop a theory for the \pp\ of the normalized eigenvalues
-of the sample covariance matrix in the case when rows and columns of the data are linearly dependent.
-We provide limit results for the weak \con\ of these \pp es to Poisson or cluster Poisson processes. Based on
-this \con\ we can also derive the limit laws of various \fct als of the ordered eigenvalues such as the
-joint \con\ of a finite number of the largest order statistics, the joint limit law of the largest eigenvalue and the trace,
-limit laws for successive ratios of ordered eigenvalues,
-etc. We also develop some
-limit theory for the singular values of the sample autocovariance matrices and their sums of squares. The theory is illustrated
-for simulated data and for the components of the S\&P 500 stock index.
-Further generalizations of the results of this chapter are made in \cite{heiny:mikosch:2016:noniid1} and \cite{heiny:mikosch:2016:stochvol1}, but are not part of this thesis.
-
-In Chapter~\ref{ch:corr}, we show that the largest and smallest eigenvalues 
-of a sample correlation matrix stemming from $n$ independent observations of a $p$-dimensional 
-time series with iid components converge almost surely to $(1+\sqrt{\gamma})^2$ and $(1-\sqrt{\gamma})^2$, 
-respectively, as $\nto$, if $p/n\to \gamma \in (0,1]$ and the truncated variance of the entry 
-distribution is ``almost slowly varying'', a condition we describe via moment 
-properties of self-normalized sums. We compare the behavior of the eigenvalues of the 
-sample covariance and sample correlation matrices and argue that the latter seems more robust, 
-in particular in the case of infinite fourth moment. We briefly address some practical issues 
-for the estimation of extreme eigenvalues in a simulation study.
-
-Chapter~\ref{ch:corr} is the most technical one of this thesis.
-%In addition to the submitted version we added Section~\ref{sec:anexample} which contains an attempt to adapt the widely employed truncation techniques to self-normalized sums with the goal of verifying our crucial distributional assumption.
-In our proofs we use the method of moments combined with a Path-Shortening Algorithm, 
-which efficiently uses the structure of sample correlation matrices, 
-to calculate precise bounds for matrix norms. We believe that this new approach could be of further use in Random Matrix Theory. 
-\bigskip
-
-{\em Johannes Heiny \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}
+{\em Xie Xiaolei \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}
 
@@
 
-\include{Chapterintroduction}
+\def\phdthesis{1}
+\input{Chapterintroduction}
 %------------------------------------------------------------------------------------
-\include{Chapteriidpaper}
-\include{Chapterextremes}
-\include{Chaptercorrelation}
+\input{ChapterTailParameters}
+%% \input{TailParameters}
+\input{Chapterextremes}
 
-\bibliography{libraryjohannes}
+% \bibliography{libraryjohannes}
+\bibliography{../thesis/econophysics}
 

### 2017-07-21 papers/DavisMikoschHeinyXie/PaperBody.tex
Add the Extremes paper.

\begin{abstract}
We  provide some  asymptotic theory for the largest eigenvalues of a sample covariance matrix
of a $p$-dimensional \ts\ where the dimension $p=p_n$ converges to infinity when the sample size $n$ increases.
We give a short overview of the literature on the topic both in the light- and heavy-tailed cases when the data have
finite (infinite) fourth moment, respectively.
Our main focus is on the heavy-tailed case. In this case, one has a theory for the \pp\ of the normalized eigenvalues
of the sample covariance matrix in the iid case but also when rows and columns of the data are linearly dependent.
We provide limit results for the weak \con\ of these \pp es to Poisson or cluster Poisson processes. Based on
this \con\ we can also derive the limit laws of various \fct als of the ordered eigenvalues such as the
joint \con\ of a finite number of the largest order statistics, the joint limit law of the largest eigenvalue and the trace,
limit laws for successive ratios of ordered eigenvalues,
etc. We also develop some
limit theory for the singular values of the sample autocovariance matrices and their sums of squares. The theory is illustrated
for simulated data and for the components of the S\&P 500 stock index.
\end{abstract}
{\bf keywords: }{Regular variation, sample covariance matrix, dependent
  entries, largest  eigenvalues, trace, point process convergence,
  cluster Poisson limit, infinite variance stable limit, Fr\'echet
  distribution.}
{\bf subject class: }{Primary 60B20; Secondary 60F05 60F10 60G10 60G55 60G70}

%\tableofcontents

\section{Estimation of the largest eigenvalues: an overview in the iid case}\label{sec:motivation}\setcounter{equation}{0}
\subsection{The light-tailed case}
%In recent years, \asy\ theory for large random matrices has a
%attracted a
%great deal of attention; see for example the monographs
%Bai and Silverstein \cite{bai:silverstein:2010} and Anderson et al.
%\cite{anderson:guionnet:zeitouni:2008}. The literature about {\em
%  heavy-tailed random matrices} is rather sparse. Here and in what
%follows, we call a random matrix {\em heavy-tailed} if its entries have
%\ds s with \regvary\ tails, typically with tail index below 4.
%The eigenvalues of heavy-tailed matrices
%with iid entries and dimension $p\times n$, where $p=p_n\to\infty$ and $p/n\to
%\gamma\in (0,\infty)$, were studied by Soshnikov \cite{soshnikov:2004,
%  soshnikov:2006} and  Auffinger et
%al. \cite{auffinger:arous:peche:2009}. In the latter reference and in
%Belinschi et al. \cite{belinschi:dembo:guionnet:2009},
%the sample covariance matrices of iid heavy-tailed \seq s were also studied.
%Bose et al. \cite{bose:hazra:saha:2009,bose:hazra:saha:2010}
%investigated the spectral norm of circulant-type heavy-tailed matrices
%and Ben Arous and Guionnet \cite{arous:guionnet:2008}
%studied the limits of heavy-tailed random Wigner matrices with
%infinite variance.
One of the exciting new areas of statistics is concerned with analyses of large data sets.
For such data one often studies the dependence structure via covariances and correlations.
In this paper we focus on one aspect: the estimation of the eigenvalues of the covariance matrix of a multivariate  \ts\
when the dimension $p$ of the series increases with the sample size $n$. In particular, we are interested in limit theory for the largest eigenvalues of the sample covariance matrix. This theory is closely related to topics from classical \evt\ such as
maximum domains of attraction with the corresponding normalizing and centering constants for maxima; cf. Embrechts et al.~\cite{embrechts:klueppelberg:mikosch:1997}, Resnick~\cite{resnick:2007,resnick:1987}.
Moreover, \pp\ \con\ with limiting Poisson and cluster Poisson processes enters in a natural way when one describes
the joint \con\ of the largest eigenvalues of the sample covariance matrix. Large deviation techniques find applications,
linking \evt\ with random walk theory and \pp\ \con . The objective of this paper is to illustrate some of the main
developments in random matrix theory for the particular case of the sample covariance matrix of multivariate \ts\ with
independent or dependent entries. We give special emphasis to the heavy-tailed case when \evt\ enters in a rather
straightforward way.
%We do not cover other parts of random matrix theory such as limit theory for the eigen-spectrum of random matrices with
%iid entries or other special types of matrices. Also in this case one can find many parallels with \evt.
\par
Classical multivariate \tsa\ deals with
observations which assume values in a $p$-dimensional space
where $p$ is ``relatively small'' compared to the sample size $n$.  With
the availability of large data sets $p$ can be ``large''
relative to $n$. One of the possible con\seq s is that standard asymptotics (such as the \clt )
break down and may even cause misleading results.
\par
The dependence
structure in multivariate data is often summarized by the covariance matrix which is typically
estimated by its sample analog.  For example, principal component analysis (PCA)
extracts principal component vectors corresponding to the largest
eigenvalues of the sample covariance matrix. The magnitudes of these eigenvalues provide an empirical \ms\ of the importance
of these components.
\par
If $p,n$ are fixed, a column of the $p\times n$ data matrix
\beao
\X =\X_n = \big(X_{it}\big)_{i=1,\ldots,p;t=1,\ldots,n}\,
\eeao
represents an observation of a $p$-dimensional \ts\ model with unknown parameters.
In this section we assume that the real-valued entries $X_{it}$ are iid, unless mentioned otherwise, and we write $X$ for a generic element.
One challenge is to infer information about the parameters
from the eigenvalues $\lambda_1,\ldots,\lambda_p$
of the {\em sample covariance matrix} $\X\X'$. In the notation we suppress the dependence of $(\la_i)$ on $n$ and $p$.
If $p$ and $n$ are finite and the columns of $\X$ are iid and multivariate normal,
Muirhead \cite{muirhead} derived a (rather complicated) formula for the joint distribution of the eigenvalues $(\lambda_i)$.
%Therefore a theoretical solution is already available in the finite $n$ case.
%The expression, however, contains an integral over the orthogonal group $\mathbb{O}(p)$.
%\begin{figure}[htb!]
%  \centering
 % \includegraphics[scale=0.5]{eigen_normal_qqplot1.pdf}
 % \caption{Largest eigenvalues of $500$ simulated sample covariance matrices against standard normal quantiles.}
 % \label{fig:QQ}
%\end{figure}

For $p$ fixed and $\nto$, assuming
$\X$ has centered normal entries and a diagonal covariance matrix $\Sigma$,
Anderson~\cite{anderson:1963}  derived the joint \asy\ density of $(\lambda_1, \ldots, \lambda_p)$.
%If the largest eigenvalue of the underlying covariance matrix appears with multiplicity one, he
%showed that the largest eigenvalue of the sample covariave matrix
%is asymptotically normal.
We quote from Johnstone~\cite{johnstone:2001}:
``The classic paper by Anderson~\cite{anderson:1963} gives the limiting joint distribution of the roots, but the
marginal distribution of the largest eigenvalue is hard to extract even in the null case'' (i.e., when the covariance matrix $\Sigma$ is
proportional to the identity matrix).
%If the largest entry of $\Sigma$ exceeds its second largest, one obtains a central limit type theorem for the largest eigenvalue of the sample covariance matrix.
%In Figure~\ref{fig:QQ} we chose $n=10000, p=50$ and $\Sigma =\diag(1024,1,\ldots,1)$, and plot the properly normalized and
%standardized largest eigenvalue against standard normal quantiles.

It turns out that limit theory for the largest eigenvalues becomes ``easier''
when the dimension $p$ increases with $n$.
Over the last 15 years there has been increasing interest in the case when $p=p_n\to\infty$ as $\nto$. In most of
the literature (exceptions are El Karoui \cite{elkaroui:2003}, Davis et al. \cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} and Heiny and Mikosch
\cite{heiny:mikosch:2015:iid})
one assumes that $p$ and $n$ grow at the same rate:
\beam\label{eq:gamma}
\dfrac p n\to \gamma\qquad \mbox{for some $\gamma \in (0,\infty)$.}
\eeam
%One explanation for the commonly used condition \eqref{eq:gamma} is that much of the theory was developed for $n \times n$ {\em Wigner matrices}, which is a Hermitean matrix $H$ such that $\{ H_{ij}:i<j \}$ and $\{ H_{ii}:0\le i\le n \}$ are two sets of iid random variables. If $p =\gamma n$, it turns out that the limit theory for the eigenvalues of Wigner matrices and sample covariance matrices is closely connected by the convergence of the empirical spectral distributions to the semicircle and the Mar\v cenko--Pastur law, respectively; see Bai and Silverstein \cite{bai:silverstein:2010}, Chapters 1-3.

In random matrix theory,  the convergence of the {\em empirical spectral distributions} $(F_{n^{-1}\X \X'})$ of a sequence $(n^{-1}\X \X')$ of non-negative definite matrices is the principle object of study. The empirical spectral distribution $F_{n^{-1}\X \X'}$ is constructed from the eigenvalues via
\begin{equation*}
F_{n^{-1}\X \X'}(x)= \frac{1}{p}\; \# \{ 1\le j\le p : n^{-1} \lambda_j \le x \}, \quad x\in \R,\quad n\ge1.
\end{equation*}
In the literature convergence results for the sequence of empirical spectral distributions are established under the assumption that $p$ and $n$ grow at the same rate.
Suppose that the iid entries $Z_{it}$ have mean $0$ and variance $1$. If \eqref{eq:gamma} holds, then, with probability one, $(F_{n^{-1}\X \X'})$ converges to the celebrated Mar\v cenko--Pastur law with absolutely  continuous part given by the density,
\begin{eqnarray}\label{eq:MP}
f_\gamma(x) =
\left\{\begin{array}{cc}
\frac{1}{2\pi x\gamma} \sqrt{(b-x)(x-a)} \,, & \mbox{if } a\le x \le b, \\
 0 \,, & \mbox{otherwise,}
\end{array}\right.
\end{eqnarray}\noindent
where $a=(1-\sqrt{\gamma})^2$ and $b=(1+\sqrt{\gamma})^2$. The Mar\v cenko--Pastur law has a point mass $1-1/\gamma$ at the origin if $\gamma>1$, cf. Bai and Silverstein~\cite[Chapter~3]{bai:silverstein:2010}. The point mass at zero is intuitively explained by the fact that, with probability $1$, $\min(p,n)$ eigenvalues $\lambda_i$ are non-zero. When $n=(1/\gamma) \; p$ and $\gamma >1$ one sees that the proportion of non-zero eigenvalues of the sample covariance matrix is $1/\gamma$ while the proportion of zero eigenvalues is $1-1/\gamma$.

While the finite second moment is the central assumption to obtain the Mar\v cenko--Pastur law as the limiting spectral distribution, the finite fourth moment plays a crucial role when studying the largest eigenvalues
\beam\label{eq:order}
\la_{(1)}\ge \cdots \ge \la_{(p)}
\eeam
of $\X\X'$, where we suppress the dependence on $n$ in the notation.

\begin{figure}[htb!]
  \centering
  \subfigure[Standard normal entries]{
    \includegraphics[scale=0.4]{normal-TW.pdf}
  }
  \subfigure[Entry distribution: $\P(X=\sqrt{3}) = \P(X=-\sqrt{3} ) =
  1/6$, $\P(X=0)=2/3$. Note $\E X = 0$, { $\E[ X^2] = 1$, $\E [X^3] = 0$ and
  $\E[ X^4] = 3$, i.e., the first 4 moments of $X$ match those of the standard normal \ds .}]{
    \includegraphics[scale=0.4]{3point-TW.pdf}
  }
  \caption{Sample density function of the largest eigenvalue compared
    with the Tracy--Widom density function. The data matrix $ \X$ has
    dimension $200 \times 1000$. An ensemble of 2000 matrices is
    simulated.}
  \label{fig:normal-3point-TW}
\end{figure}

Assuming \eqref{eq:gamma} and iid entries $X_{it}$ with zero mean, unit variance and finite fourth moment,  Geman~\cite{geman}
showed that
\beam\label{eq:geman}
\dfrac {\la_{(1)}}{n} \stas \big(1+\sqrt{\gamma}\big)^2\,,\qquad \nto\,.
\eeam
Johnstone \cite{johnstone:2001} complemented this \slln\ by the corresponding \clt\ in the special case of iid standard normal
entries:
\beam\label{eq:tc}
 n^{2/3}\,\dfrac{(\sqrt{\gamma})^{1/3}}{\big(1+\sqrt{\gamma}\big)^{4/3}}\Big(\dfrac {\la_{(1)}}{n} -
\big(1+\sqrt{\tfrac pn }\big)^2\Big)
\std {\rm TW}\,,
\eeam
where the limiting \rv\ has a {\em Tracy--Widom \ds} of order 1. Notice that the centering
$\big(1+\sqrt{\tfrac pn }\big)^2$ can in general not be replaced by $(1+\sqrt{\gamma})^2$.
This \ds\ is ubiquitous in random matrix theory.
%It is defined via some ordinary differential equation; we refer to \cite{tracy:widom:2012} for a definition and properties.
Its distribution function $F_1$ is given by
\begin{equation*}
F_1(s) = \exp\Big\{
  -\frac{1}{2} \int_{s}^\infty [
    q(x) + (x - s) q^2(x)
 ] \dint x
\Big\}\,,
\end{equation*}
where $q(x)$ is the unique solution to the Painlev\'e II differential
equation
\begin{equation*}
  q''(x) = xq(x) + 2 q^3(x)\,,
\end{equation*}
where $ q(x)\sim {\rm Ai}(x)$ as $x \to \infty$ and Ai$(\cdot)$ is the Airy kernel; see Tracy and Widom~\cite{tracy:widom:2012} for details.
We notice that the rate $n^{2/3}$ compares favorably to the $\sqrt{n}$-rate in the classical \clt\ for sums
of iid finite variance \rv s.
The calculation of the spectrum is facilitated by the fact that the distribution of
the classical Gaussian matrix ensembles is invariant under orthogonal transformations. The corresponding
computation for non-invariant matrices with non-Gaussian entries is more complicated and was a major challenge for several years; a first step was made by Johansson \cite{johansson}.
Johnstone's result was extended to matrices $\X$ with iid non-Gaussian entries
by Tao and Vu \cite[Theorem~1.16]{tao09b}. Assuming that the first four moments of the entry \ds\ match  those of
the standard normal \ds , they showed \eqref{eq:tc} by
employing {\em Lindeberg's replacement method}, i.e., the iid non-Gaussian entries are replaced
step-by-step by iid Gaussian ones.
This approach is well-known from summation theory for \seq s of iid \rv s. Tao and Vu's result is a consequence of the so-called {\em Four Moment Theorem}, which describes the insensitivity of the eigenvalues with respect to changes in the distribution of the entries. To some extent (modulo the strong moment matching conditions) it shows the universality of Johnstone's limit
result \eqref{eq:tc}. Later we will deal with entries with infinite fourth moment. In this case, the weak limit
for the normalized largest eigenvalue $\la_{(1)}$ is distinct from the Tracy--Widom \ds : the classical Fr\'echet extreme value
\ds\ appears.
In Figure~\ref{fig:normal-3point-TW} we illustrate how the Tracy--Widom approximation works for Gaussian and non-Gaussian entries of $\X$
and in Figure~\ref{fig:MyDist-Frechet} we also illustrate that this approach fails when $\E[X^4]=\infty$.

Figure \ref{fig:normal-3point-TW} compares the sample density function
of the properly normalized largest eigenvalue estimated from 2000 simulated sample covariance matrices $\X\X'$ ($n=1000, p=200$) with
the Tracy--Widom density. If  $X$ has infinite fourth moment and further regularity conditions on the tail hold then the
Tracy--Widom limiting law needs to be replaced by the \Frechet
distribution; see Section~\ref{subsec:1.2} for details. Figure \ref{fig:MyDist-Frechet} illustrates this fact with a
simulated ensemble whose entries are distributed according to
the heavy-tailed distribution from \eqref{eq:distrsim} below with $\alpha = 1.6$.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{MyDist-Frechet.pdf}
  \caption{Sample distribution function of the largest eigenvalue $\la_{(1)}$
    compared to the \Frechet distribution (solid line) with $\alpha=1.6$. The data matrices have dimension
    $200 \times 1000$ and iid entries with infinite fourth moment. The results are based on 2000 replicates.}
  \label{fig:MyDist-Frechet}
\end{figure}

\subsection{The heavy-tailed case}\label{subsec:1.2}
So far we focused on ``light-tailed'' $\X$
in the sense that its entries have finite fourth moment.  However, there is statistical evidence that the assumption
of finite fourth moment may be violated when dealing with data from
insurance, finance or telecommunications. We illustrate this fact
in Figure~\ref{fig:SP500_tail_indices} where we show the pairs $(\alpha_L,\alpha_U)$ of
lower and upper tail indices
of $p=478$  log-return series composing
the S\&P 500 index estimated from $n=1,345$ daily observations from 01/04/2010 to 02/28/2015.
This means we assume for every row  $(X_{it})_{t=1,\ldots,n}$ of $\X$ that the tails behave like
\beao
\P(X_{it}>x)\sim c_U\,x^{-\alpha_U}\qquad\mbox{and}\qquad \P(X_{it}<-x)\sim c_L\,x^{-\alpha_L}\,,\qquad \xto\,,
\eeao
for non-negative constants $c_L,c_U$. We apply the Hill estimator (see Embrechts et al.~\cite{embrechts:klueppelberg:mikosch:1997}, p.~330,
de Haan and Ferreira \cite{dehaan:ferreira:2006}, p.~69)
to the \ts\ of the gains and losses in a naive way,
neglecting the dependence and non-stationarity in the data; we also omit confidence bands.
From the figure it is evident that the majority of the return series have tail indices below
four, corresponding to an infinite fourth moment.
\begin{figure}[htb!]
\centering
\includegraphics[scale=0.5]{SP500_tail_indices.pdf}
\caption{Tail indices of log-returns of 478 \ts\ from the S\&P 500 index. The values  $(\hat \alpha_L,\hat \alpha_U)$ of the
lower and upper tail indices are provided by
Hill's estimator.  We also draw the line $\hat\alpha_U=\hat\alpha_L$.}
\label{fig:SP500_tail_indices}
\end{figure}
The behavior of the largest eigenvalue $\la_{(1)}$
changes dramatically when $\X$ has infinite fourth moment.
Bai and Silverstein~\cite{baisilv} proved for an $n\times n$ matrix $\X$ with iid centered entries
that
\beam\label{eq:wdfr}
\limsup_{\nto} \dfrac{\la_{(1)}}{n}=\infty \qquad {\rm a.s.}
\eeam
This is in stark contrast to Geman's result \eqref{eq:geman}.
%It is also unlikely that the
%Tracy--Widom \ds\ yields a good approximation to the \ds\ of the largest eigenvalue of the sample covariance matrix;
%see Figure~\ref{fig:MyDist-Frechet} for some simulation evidence.
\par
In the heavy-tailed case it is common to assume a {\em \regvar\ condition:}
\beam\label{eq:regvar}
\P(X>x)\sim p_+\,\dfrac{L(x)}{x^{\alpha}}\qquad\mbox{and}\qquad \P(X<-x) \sim p_-\,\dfrac{L(x)}{x^\alpha}\,,\qquad \xto\,,
\eeam
where $ p_\pm$ are non-negative constants \st\ $p_++p_-=1$ and $L$ is a \slvary\ \fct . In particular, if $\alpha<4$ we have $\E [X^4]= \infty$. The \regvar\ condition
on $X$ (we will also refer to $X$ as a \regvary\ \rv ) is needed for proving \asy\ theory for
the eigenvalues of $\X\X'$. This is similar to proving limit theory for sums
of iid \rv s with infinite variance stable limits; see for example Feller~\cite{feller}.
\par

In \eqref{eq:MP} we have seen that the sequence $(F_{n^{-1}\X \X'})$ of empirical spectral distributions converges to the Mar\v cenko--Pastur law if the centered iid entries possess a finite second moment. Now we will discuss the situation when the entries are still iid and centered, but have an infinite variance. Here we assume the entries to be regularly varying with index $\alpha \in (0,2)$.
Assuming \eqref{eq:gamma} with $\gamma\in (0,1]$ in this infinite variance case, Belinschi et al.~\cite[Theorem~1.10]{belinschi:dembo:guionnet:2009} showed that the sequence $(F_{a_{n+p}^{-2}\X \X'})$ converges with probability one to a non-random probability measure with density $\rho_{\alpha}^\gamma$ satisfying
\begin{equation*}
\rho_{\alpha}^\gamma(x) x^{1+\alpha/2} \to \frac{\alpha \gamma}{2(1+\gamma)}, \quad x \to \infty,
\end{equation*}
see also Ben Arous and Guionnet \cite[Theorem~1.6]{arous:guionnet:2008}.
The normalization $(a_k)$ is chosen \st\ $\P(|X|>a_k)\sim k^{-1}$ as $k\to\infty$. An application of the Potter bounds (see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}) shows that $a_{n+p}^2/n \to \infty$.

It is interesting to note that there is a phase change in the extreme eigenvalues in going from finite to infinite fourth moment, while the phase change occurs for the empirical spectral distribution going from finite to infinite variance.
%In the case of finite fourth moment, \eqref{eq:geman} and \eqref{eq:MP} imply that $\lambda_{(k)}/n$ has to converge to the right endpoint of the Mar\v cenko--Pastur law for any finite $k$. However, if we assume an infinite fourth moment but finite second moment the empirical spectral distributions  $(F_{{n}^{-1}\X \X'})$ still converge to the Mar\v cenko--Pastur law with the same support, while $\lambda_{(1)}/n$ will tend to infintity due to \eqref{eq:wdfr}. This means that the proportion of the normalized eigenvalues taking values larger than $(1+\sqrt{\gamma})^2$ has to be negligible; otherwise we would have a contradiction to the convergence of the empirical spectral distributions to the Mar\v cenko--Pastur law.

%Finally, if the second moment is infinity, i.e.~regularly varying entries with index $\alpha \in (0,2)$, then the much larger values of the eigenvalues are visible in the limiting spectral distribution even when using a much stronger normalization \cite[Theorem~1.10]{belinschi:dembo:guionnet:2009}.
\par

The theory for the largest eigenvalues of sample covariance matrices with heavy tails is less developed than in the light-tailed case.
Pioneering work for $\la_{(1)}$ in the case of iid \regvary\ entries $X_{it}$ with index $\alpha\in (0,2)$
is due to Soshnikov~\cite{soshnikov:2004,soshnikov:2006}. He showed the \pp\ \con\
\beam\label{eq:nn}
N_n=\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i} \std N=\sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeam
under the growth condition \eqref{eq:gamma} on $(p_n)$.
%The normalization $(a_k)$ is chosen \st\ $\P(|X|>a_k)\sim k^{-1}$ as $k\to\infty$.
Here
\beam\label{eq:Gamma}
\Gamma_i=E_1+\cdots + E_i\,,\qquad i\ge 1\,,
\eeam
and $(E_i)$ is an iid standard exponential \seq . In other words, $N$ is a Poisson point process on $(0,\infty)$ with mean \ms\
$\mu(x,\infty)= x^{-\alpha/2}$, $x>0$. Convergence in \ds\ of \pp es is understood in the sense of weak \con\
in the space of point \ms s equipped with the vague topology; see Resnick \cite{resnick:2007,resnick:1987}.
We can easily derive the limiting distribution of $a_{np}^{-2} \lambda_{(k)}$ for fixed $k\ge 1$ from \eqref{eq:nn}:
\begin{equation*}
\begin{split}
\lim_{\nto}\P(a_{np}^{-2} \lambda_{(k)}\le x)&= \lim_{\nto}\P(N_n(x,\infty)<k)
=  \P(N(x,\infty)<k) =\P(\Gamma_k^{-2/\alpha}\le x)\\ &= \sum_{s=0}^{k-1} \frac{\big(\mu(x,\infty)\big)^s}{s!} \e^{-\mu(x,\infty)}, \quad  x>0.
\end{split}
\end{equation*}
In particular,
\beao
\dfrac{\la_{(1)}}{a_{np}^2}\std \Gamma_1^{-\alpha/2}\,,\qquad \nto\,,
\eeao
where the limit has {\em Fr\'echet \ds } with parameter $\alpha/2$ and distribution function
\beao
\Phi_{\alpha/2}(x) =\ex^{-x^{-\alpha/2}}\,,\qquad x>0\,.
\eeao

\noindent We mention that the tail balance condition \eqref{eq:regvar} may be replaced in this case by the weaker assumption
$\P(|X|>x)= L(x) x^{-\alpha}$ for a \slvary\ \fct\ $L$. Indeed, it follows from the proof %(see also the comments below)
that only the squares $X_{it}^2$ contribute to the \pp\ limits of the eigenvalues $(\la_i)$. A con\seq\ of the
\cmt\ and \eqref{eq:nn} is the joint \con\ of the upper order statistics: for any $k\ge 1$,
\beao
a_{np}^{-2} \big(\la_{(1)},\ldots,\la_{(k)}\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,,\qquad \nto\,.
\eeao

It follows from standard theory
for \pp es with iid points (e.g. Resnick \cite{resnick:2007,resnick:1987})
that \eqref{eq:nn} remains valid if
we replace  $N_n$  by the \pp\ $\sum_{i=1}^p\sum_{t=1}^n\vep_{X_{it}^2/a_{np}^2}$. Then we also have for any $k\ge 1$,
\beam\label{eq:nna}
a_{np}^{-2} \big(X_{(1),np}^2,\ldots,X_{(k),np}^2\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,,\qquad \nto\,,
\eeam
where
\beao
X_{(1),np}^2\ge \cdots \ge X_{(np),np}^2
\eeao
denote the order statistics of $(X_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$.
\par
Auffinger et al.~\cite{auffinger:arous:peche:2009} showed that \eqref{eq:nn}
remains valid under the \regvar\ condition \eqref{eq:regvar} for  $\alpha\in (2,4)$, the growth condition
\eqref{eq:gamma} on $(p_n)$ and the
additional assumption $\E [X]=0$. Of course, \eqref{eq:nna} remains valid as well.
Davis et al.~\cite{davis:pfaffel:stelzer:2014} extended these results to the case when the rows of $\X$
are iid linear processes with iid \regvary\ noise. The Poisson \pp\ \con\ result of \eqref{eq:nn} remains valid in this case.
Different limit processes can only be expected if there is dependence across rows and columns.
\par
In what follows, we refer to the  {\em heavy-tailed case} when we assume the \regvar\ condition \eqref{eq:regvar}
for some $\alpha\in (0,4)$.%{\blue This hypothesis is confirmed in}

\subsection{Overview}

The primary objective of this overview  is to make a connection between extreme value theory and the behavior of the largest eigenvalues of sample covariance matrices from heavy-tailed multivariate time series.  For time series that are linearly dependent through time and across rows, it turns out that the extreme eigenvalues are essentially determined by the extreme order statistics from an array of iid random variables.  The asymptotic behavior of the extreme eigenvalues is then derived routinely from classical extreme value theory.  As such, explicit joint distributions of the extreme order statistics can be given which yield a plethora of ancillary results.  
%These include the limit behavior of ratios of consecutive extreme eigenvalues, ratios of extreme order statistics to the sample variance, etc..
%%The underlying multivariate \ts\ will assumed to be heavy-tailed and linearly dependent through time and across the rows.We consider asymptotic theory for the largest eigenvalues of a sample covariance matrix,
%and we will also address the largest singular values of the sample autocovariance matrix.
%To our knowledge, the present paper is the first to consider bonafide dependence among the components in the time series which renders a multivariate analysis, such as PCA, meaningful.
Convergence of the point process of extreme eigenvalues, properly normalized, plays a central role in establishing the main results.
\par
In Section~\ref{sec:2} we continue the study of the case when the data matrix $\X$ consists of iid heavy-tailed entries.  We will consider power-law growth rates on the dimension $(p_n)$ that is more general than prescribed by \eqref{eq:gamma}.
In Section~\ref{sec:model} we introduce a model for $X_{it}$ which allows for linear dependence across the rows and through time.
The point process convergence of normalized eigenvalues is presented in Section~\ref{sec:mainresult}.
This result lays the foundation for new insight into the spectral behavior of the sample covariance matrix, which is the content of Section~\ref{sec:samplecov}.

Sections~\ref{sec:samplecov} and \ref{sec:possemidef} are devoted to {\em sample autocovariance matrices}.  Motivated by \cite{lam:yao}, we study the eigenvalues of sums of transformed matrices and illustrate the results in two examples. These results are applied to the time series of S\&P 500 in Section \ref{sec:sp500}.
%The technical proofs are collected in Section~\ref{sec:proof}. 
Appendix~\ref{appendix:A} contains useful facts about regular variation and point processes.

\section{General growth rates for $p_n$ in the iid heavy-tailed case}\label{sec:2}\setcounter{equation}{0}
This section is based on ideas in Heiny and Mikosch \cite{heiny:mikosch:2015:iid} where one can also find detailed proofs.
\subsection*{Growth conditions on $(p_n)$}\label{subsec:pn}
In many applications it is not realistic to assume
that the dimension $p$ of the data and the sample size $n$ grow at the same rate.
The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009} already indicate that the value $\gamma$ in the growth
rate~\eqref{eq:gamma} does not appear in the \ds al limits.
This obervation is in contrast to the light-tailed case; see \eqref{eq:geman}
and \eqref{eq:tc}.  Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}
and Heiny and Mikosch ~\cite{heiny:mikosch:2015:iid} allowed for more general rates for
$p_n\to\infty$ than linear growth in $n$.
Recall that $p=p_n\to\infty$ is the number of rows in the
matrix $\X_n$. We need to specify the growth rate of $(p_n)$ to ensure
a non-degenerate limit distribution of the normalized singular values of the sample autocovariance
matrices. To be precise, we assume
\begin{equation}\label{eq:p}
p=p_n=n^\beta \ell(n), \quad n\ge1,\tag{$C_p(\beta)$}
\end{equation}
where $\ell$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume $\ell(n) \to \infty$.
Condition~\ref{eq:p} is more general than the growth conditions in the literature; see
\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
\begin{theorem}\label{thm:intro}
Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvar} for some
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Let $(p_n)$ be an integer sequence
satisfying \ref{eq:p} with $\beta\ge 0$. In addition, we require
\begin{equation}\label{Cbeta}
\min(\beta,\beta^{-1})\in (\alpha/2-1,1] \qquad \mbox{ for } \alpha\in [2,4), \tag{$\widetilde{C}_\beta(\alpha)$}
\end{equation}
Then
\beam\label{eq:2}
\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeam
where the \con\ holds in the space of point \ms s with state space $(0,\infty)$ equipped with the vague toplogy; see Resnick \cite{resnick:2007}.
\end{theorem}

\subsubsection*{A discussion of the case $\beta\in [0,1]$}
We mentioned earlier that in the heavy-tailed case, limit theory for the
largest eigenvalues of the sample covariance matrix is rather insensitive to
the growth rate of $(p_n)$ and that the limits are essentially
determined by the diagonal of this matrix. This is confirmed by the following result.

\begin{proposition}\label{prop:offdiagonal}
Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvar} for some
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Then for any \seq\ $(p_n)$
satisfying \ref{eq:p} with $\beta\in [0,1]$ we have
\beao
a_{np}^{-2} \twonorm{\X \X' - \diag(\X \X')}\stp 0\,,\qquad\nto\,,
\eeao
where $\| \cdot\|_2$ denotes the spectral norm; see \eqref{specnorm} for its definition.
\end{proposition}

\par
Proposition~\ref{prop:offdiagonal} is not unexpected for two reasons:
\begin{itemize}
\item
It is well-known from classical theory (see Embrechts and Veraverbeke \cite{embrechts:veraverbeke:1982}) that for any iid \regvary\ non-negative \rv s
$Y,Y'$ with index  $\alpha'>0$, $Y\,Y'$ is \regvary\ with index $\alpha'$ while $Y^2$ is \regvary\ with index $\alpha'/2$.
Therefore $X^2$ and $X_{11}X_{12}$  are \regvary\ with indices $\alpha/2$ and $\alpha$, respectively.
\item
The aforementioned tail behavior is inherited by the entries of $\X\X'$ in the following sense.
By virtue of Nagaev-type \ld\ results for an iid \regvary\ \seq\ $(Y_i)$ with index $\alpha'\in (0,2)$ where we also assume that
$\E [Y_0]=0$ if $\E[|Y_0|]<\infty$
(see Theorem~\ref{thm:nagaev}) %S.V. Nagaev \cite{nagaev:1979} and
%Cline and Hsing \cite{cline:hsing:1998})
we have that $\P(Y_1+\cdots +Y_n >b_n)/(n \,\P(|Y_0|>b_n))$ converges to a non-negative constant
provided $b_n/a_n'\to\infty$, where $\P(|Y_0|>a_n')\sim n^{-1}$ as $\nto$. As a con\seq\ of the tail behaviors of $X_{it}^2$ and $X_{it}X_{jt}$ for $i\ne j$
and Nagaev's results we have for $(b_n)$ \st\ $b_n/a_n^2\to\infty$,
\beao
\dfrac{\P\big((\X\X')_{ij}> b_n\big)}{\P\big((\X\X')_{ii}- c_n> b_n\big)}\sim \dfrac{n\,\P( X_{11}X_{12}>b_n)}{n\,\P(X^2>b_n)}\to 0\,,\qquad \nto\,,
\eeao
where $c_n=0$ or $n\,\E[X^2]$ according as $\alpha\in (0,2)$ or $\alpha\in (2,4)$.
This means that the diagonal and off-diagonal entries of
$\X\X'$ inherit the tails of $X_{it}^2$ and $X_{it}X_{jt}$, $i\ne j$, respectively, above the high threshold $b_n$.
\end{itemize}
\par
Proposition~\ref{prop:offdiagonal} has some immediate con\seq s for the approximation of the eigenvalues
of $\X\X'$ by those of ${\rm diag}(\X\X')$. Indeed, let $C$ be a symmetric $p\times p$ matrix with
eigenvalues $\la_1(C),\ldots,\la_p(C)$ and
ordered eigenvalues
\beam \label{eq:weyl}
\la_{(1)}(C)\ge \cdots \ge \la_{(p)}(C)\,.
\eeam
Then for any symmetric $p\times p$ matrices $A,B$, by {\em Weyl's inequality} (see Bhatia \cite{bhatia:1997}),
\beao
\max_{i=1,\ldots,p}\big|\la_{(i)}(A+B)-\la_{(i)}(A)\big|\le \|B\|_2\,.
\eeao
If we now choose
$A+B=\X\X'$ and $A= \diag (\X\X')$ we obtain the following result.
\begin{corollary}\label{cor:687}
Under the conditions of Proposition~\ref{prop:offdiagonal},
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\diag(\X\X'))\big|\stp 0\,,\quad\nto \,.
\eeao
\end{corollary}
Thus the problem of deriving limit theory for the order statistics of $\X\X'$ has been reduced to limit theory for the order
statistics of the iid row-sums
\beao
D_i^{\rightarrow}= (\X\X')_{ii}=\sum_{t=1}^n X_{it}^2\,,\qquad i=1,\ldots,p\,,
\eeao
which are the eigenvalues of $\diag(\X\X')$. This theory is completely described by the \pp es constructed from the points $D_i^\rightarrow/a_{np}^2$ $i=1,\ldots,p$. Necessary
and sufficient conditions for the weak \con\ of these \pp es are provided by  Lemma~\ref{lem:ppr}
which in combination with the Nagaev-type \ld\ results of  Theorem~\ref{thm:nagaev}
yield the following result; see also Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
\begin{lemma}\label{lem:pp}
Assume the conditions of Proposition~\ref{prop:offdiagonal} hold. Then
\beao
\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i^\rightarrow-c_n)}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeao
where $(\Gamma_i)$ is defined in \eqref{eq:Gamma} and $c_n=0$ if $\E[D^\rightarrow]=\infty$ and $c_n=\E[D^\rightarrow]$ otherwise.
\end{lemma}
In this result, centering is only needed for $\alpha\in [2,4)$ when $n/a_{np}^2\not\to 0$. Under the additional condition \ref{Cbeta},
%\begin{equation}\label{Cbetaor}
%\beta\in (\alpha/2-1,1], \quad \mbox{ for } \alpha\in [2,4),
%\end{equation}
$n/a_{np}^2\to 0$ in view of the Potter bounds; see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}.
Combining Lemma~\ref{lem:pp} and Corollary~\ref{cor:687}, we conclude that Theorem~\ref{thm:intro} holds for $\beta\in [0,1]$.

%In  Davis et al. \cite{davis:mikosch:pfaffel;2015} relation \eqref{eq:2} was derived for $\alpha\in (0,1)$ even for \seq s
%$p_n\to\infty$ \st\ $p_n= O(\ex^{s_n/n})$, where $s_n=o(n)$ but with stronger limitations on the growth of $(p_n)$ for $[1,4]$.
%In the case $\alpha\in (2,4]$ they proved \eqref{eq:center} with $(\la_i)$ replaced by $(\la_i(\X\X'-n\E[ \X\X']))$. It is surprising that
%\eqref{eq:center} also holds since, in general, $\la_i(\X\X'-\E [\X\X'])\not = \la_i-n\,\E X$.

\subsubsection*{Extension to general $\beta$}%\label{subsec:generalb}
Next we explain that it suffices to consider only the case $\beta\in [0,1]$ and how to proceed when $\beta>1$.
The main reason is that
the $p \times p$ sample covariance matrix
$\X\X'$ and the $n \times n$ matrix $\X'\X$  have the same rank and their non-zero eigenvalues coincide; see Bhatia \cite[p.~64]{bhatia:1997}. When proving limit
theory for the eigenvalues  of the sample covariance matrix one may switch to $\X'\X$ and vice versa,
hereby interchanging the roles of $p$ and $n$. By switching to $\X'\X$, one basically replaces $\beta$ by $\beta^{-1}$. Since $\min(\beta,\beta^{-1})\in [0,1]$ for any $ \beta \ge 0$, one can assume without loss of generality that $\beta\in [0,1]$.  This trick allows one to extend results for
$(p_n)$ satisfying \ref{eq:p} with $\beta\in [0,1]$  to $\beta>1$. We illustrate this approach by providing the direct analogs of Proposition~\ref{prop:offdiagonal} and Corollary~\ref{cor:687}.

\begin{proposition}\label{prop:offdiagonal1}
Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvar} for some
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Then for any \seq\ $(p_n)$
satisfying \ref{eq:p} with $\beta>1$ we have
\beao
a_{np}^{-2} \twonorm{\X' \X - \diag(\X' \X)}\stp 0\,,\qquad\nto\,,
\eeao
where $\| \cdot\|_2$ denotes the spectral norm.
\end{proposition}
Note that for $\beta>1$ we have $\lim_{\nto} p/n= \infty$. This means that
$\X'\X$ has  asymptotically a much smaller dimension than $\X\X'$ and therefore it is more convenient to work with $\X' \X$
when bounding the spectral norm.
\begin{corollary}\label{cor:6871}
Under the conditions of Proposition~\ref{prop:offdiagonal1},
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,n}\big|\la_{(i)}-\la_{(i)}(\diag(\X'\X))\big|\stp 0\,,\quad\nto \,.
\eeao
\end{corollary}
Now, Theorem~\ref{thm:intro} for $\beta > 1$ is a consequence of Corollary~\ref{cor:6871}.
\par

%-----------------------------------------------------------------------
\section{Introducing dependence between the rows and columns}\label{sec:model}\setcounter{equation}{0}
For details on the results of this section, we refer to Davis et al.~\cite{davis:mikosch:pfaffel:2015}, Heiny and Mikosch \cite{heiny:mikosch:2015:iid} and Heiny et al.~\cite{heiny:mikosch:2016:noniid}.

\subsection{The model}
When dealing with covariance matrices of a multivariate \ts\ $(\X_n)$ it is rather natural to assume dependence
between the entries $X_{it}$.
In this section we introduce a model which allows for {\em linear dependence}
between the rows and columns of $\X$:
\begin{equation}\label{eq:1}
X_{it}=\sum_{l\in \Z}\sum_{k\in \Z} h_{kl} Z_{i-k,t-l}\,,\qquad i,t\in\Z\,,
\end{equation}
where $(Z_{it})_{i,t\in \Z}$ is a field of iid \rv s and $(h_{kl})_{k,l\in\Z}$ is an array of real numbers.
Of course, linear dependence is restrictive in some sense. However, the particular dependence structure allows one to
determine those ingredients in the sample covariance matrix which contribute to its largest eigenvalues.
If the series in \eqref{eq:1} converges a.s. $(X_{it})$ constitutes a strictly stationary random field.
We denote generic elements of the $Z$- and $X$-fields by $Z$ and $X$, respectively. We assume that $Z$ is \regvary\ in the sense that
\begin{equation}\label{eq:27}
\P(Z>x)\sim p_+ \dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad  \P(Z\le -x)\sim p_-
\dfrac{L(x)}{x^{\alpha}}\,,\qquad \xto\,,
\end{equation}
for some tail index $\alpha>0$, constants $p_+,p_-\ge 0$ with $p_++p_-=1$ and a \slvary\ $L$. We will assume $\E[ Z]=0$ whenever $\E [Z^2]<\infty$.
Moreover, we require the summability condition
\begin{equation}\label{eq:2a}
\sum_{l \in \Z} \sum_{k\in \Z} |h_{kl}|^{\delta} <\infty
\end{equation}
for some $\delta\in (0,\min({\alpha/2},1))$ which ensures
the a.s.~absolute convergence of the series in \eqref{eq:1}. Under the conditions \eqref{eq:27} and \eqref{eq:2a}, the marginal and
\fidi s of the field $(X_{it})$ are \regvary\ with index $\alpha$; see
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, Appendix A3.3. Therefore we also refer to $(X_{it})$ and $(Z_{it})$
as \regvary\ fields.
\par
The model \eqref{eq:1} was introduced by Davis et al. \cite{davis:pfaffel:stelzer:2014}, assuming the rows iid, and in the
present form by Davis et al. \cite{davis:mikosch:pfaffel:2015}.
\subsection{Sample covariance and autocovariance matrices}
From the field $(X_{it})$ we construct the $p\times n$ matrices
\begin{equation}\label{eq:26}
\X_n(s)= (X_{i,t+s})_{i=1,\ldots,p;t=1,\ldots,n}\,,\qquad s=0,1,2,\ldots\,,
\end{equation}
As before, we will write $\X=\X_n(0)$.
Now we can introduce the (non-normalized)
{\em sample autocovariance matrices}
\beam\label{eq:sample}
\X_n(0)\X_n(s)'\,,\qquad s=0,1,2,\ldots\,.
\eeam
We will refer to $s$ as the {\em lag}. For $s=0$, we obtain the {\em sample covariance matrix.}
In what follows, we will be interested in the asymptotic behavior (of \fct s) of the eigen- and singular values of the
sample covariance and autocovariance matrices in the heavy-tailed case. Recall that the {\em singular values} of a matrix $A$ are the square roots of the
eigenvalues of the non-negative definite matrix $AA'$ and its {\em spectral norm} $\twonorm{A}$ is its largest singular value. We notice that $\X_n(0)\X_n(s)'$ is not symmetric
and therefore its eigenvalues can be complex. To avoid this situation, we use
%\beao
%\X_n(0)\X_n(s)' +\X_n(s)\X_n(0)'
%\eeao
%(this matrix is symmetric and therefore it has real eigenvalues) or use
the squares
\beam\label{eq:squares}
\X_n(0)\X_n(s)' \X_n(s)\X_n(0)'
\eeam
%(this matrix is  non-negative definite  and therefore it has non-negative eigenvalues).
whose eigenvalues are the squares of the singular values of $\X_n(0)\X_n(s)'$.
The idea of using the sample autocovariance matrices and \fct s of their squares \eqref{eq:squares}
originates from a paper by Lam and Yao~\cite{lam:yao} who used a model different from \eqref{eq:1}.
This idea is quite natural in the context of \tsa .
\par
In Theorem~\ref{thm:mains} below, we provide a general approximation result for the ordered singular values of the sample
autocovariance matrices in the heavy-tailed case. This result is rather technical. To formulate it we introduce further notation.
As before, $p=p_n$ is any integer \seq\ converging to infinity.
\subsection{More notation}
Important roles are played by the quantities $(Z_{it}^2)_{i=1,\ldots,p;t=1,\ldots,n}$ and their order statistics
 \begin{equation}\label{eq:zorder}
Z_{(1),np}^2 \ge Z_{(2),np}^2 \ge  \ldots \ge Z_{(np),np}^2, \qquad n,p\ge 1\,.
\end{equation}
As important are the row-sums
\begin{equation}\label{eq:ll}
D_i^\rightarrow=D_i^{(n),\rightarrow}=\sum_{t=1}^n Z_{it}^2\,, \qquad
i=1,\ldots,p;\quad  n=1,2,\ldots\,,
\end{equation} with generic element $D^\rightarrow$ and their ordered values
\beam\label{eq:help6}
D_{(1)}^\rightarrow=D_{L_1}^\rightarrow\ge \cdots \ge D_{(p)}^\rightarrow=D_{L_p}^\rightarrow\,,
\eeam
where we assume without loss of generality that $(L_1,\ldots,L_p)$ is
a permutation of $(1,\ldots,p)$ for fixed~$n$.
\par
Finally, we introduce the column-sums
\beao
D_t^\downarrow=D_t^{(n),\downarrow}= \sum_{i=1}^p Z_{it}^2\,,\qquad t=1,\ldots,n;\; \quad p=1,2,\ldots\,,
\eeao
with generic element $D^\downarrow$ and we also adapt the notation from \eqref{eq:help6} to these quantities.
\subsubsection*{Matrix norms} For any $p\times n $ matrix $\A=(a_{ij})$, we will use the following norms:
\begin {itemize}
\item
{\em Spectral norm:}
\beam\label{specnorm}
\|\A\|_2=\sqrt{\la_{(1)}(\A\A')}\,,
\eeam
\item
{\em Frobenius norm:}
\beao
\frobnorm{\A}= \Big( \sum_{i=1}^p \sum_{j=1}^n |a_{ij}|^2\Big)^{1/2}\,.
\eeao
%\item
%{\em Max-row sum norm:}
%\beao
%\inftynorm{A}= \max_{i=1,\ldots,p} \sum_{j=1}^n |a_{ij}|\,.
%\eeao
%\item
%{\em Max-column sum norm:}
%\beao
%\|{A}\|_1 = \max_{j=1,\ldots,n} \sum_{i=1}^p |a_{ij}|\,.
%\eeao
\end{itemize}
We will frequently make use of the bound $\|\A\|_2\le \|\A\|_ F$. Standard references for matrix norms are \cite{belitski,bhatia:1997, horn, shores}.
\subsubsection*{Singular values of the sample autocovariance matrices} Fix integers $n\ge 1$ and $s\ge 0$. We recycle the $\la$-notation
for the singular values
$\la_1(s), \ldots, \la_p(s)$ of the sample autocovariance matrix $\X_n(0)\X_n(s)'$, suppressing the dependence on $n$.
Correspondingly, the  order statistics are denoted by
\begin{equation}\label{eq:sigma}
\la_{(1)}(s) \ge \cdots \ge \la_{(p)}(s)\,.
\end{equation}
When $s=0$ we typically write $\la_i$ instead of $\la_i(0)$.
\subsubsection*{The matrix $\M(s)$} We introduce some auxiliary matrices
derived from the coefficient matrix $\H=(h_{kl})_{k,l\in \Z}$:
\beao
\H(s)=(h_{k,l+s})_{k,l\in \Z}, \qquad \M(s)= \H(0)\H(s)'\qquad s\ge 0\,.
\eeao
Notice that
\begin{equation}\label{eq:m}
(\M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z .
\end{equation}
We denote the ordered singular values of $\M(s)$ by
\begin{equation}\label{eq:v1}
v_1(s) \ge v_2(s) \ge\cdots \,.
\end{equation}
Let $r(s)$ be the rank of $\M(s)$ so that $v_{r(s)}(s)>0$ while $v_{r(s)+1}(s)=0$ if $r(s)$ is
finite, otherwise $v_i(s)>0$ for all $i$. We also write $r=r(0)$.
\par
Under the summability condition \eqref{eq:2a} on $(h_{kl})$ for fixed $s\ge 0$,
\beam\label{eq:tracea}
\sum_{i=1}^\infty (v_i(s))^2 &=&  \frobnorm{\M(s)}^2= \sum_{i,j\in \Z} \sum_{l_1,l_2 \in \Z} h_{i,l_1} h_{j,l_1+s}h_{i,l_2}
h_{j,l_2+s}\nonumber\\
&\le& c \,\Big(\sum_{l_1,l_2 \in \Z} \sum_{i\in \Z} |h_{i,l_1} h_{i,l_2}|\Big)^2
\le c \,\sum_{l_1\in \Z} \sum_{i \in \Z} |h_{i,l_1}| <\infty\,.
\eeam
Therefore all singular values $v_i(s)$ are finite and the ordering
  \eqref{eq:v1}
is justified.
\par
{\em Here and in what follows, we write $c$ for any constant whose value is not of interest.}

\subsubsection*{Normalizing sequence}
We define $(a_k)$ by
\beao
\P(|Z|>a_k)\sim k^{-1}\,,\qquad k\to\infty\,,
\eeao
and choose the normalizing \seq\ for the singular values as $(a_{np}^2)$ for suitable \seq s $p=p_n\to\infty$.
\subsubsection*{Approximations to singular values}\label{subsec:defdelta}
We will give approximations to the singular values $\la_i(s)$  in
terms of the  $p$ largest ordered values for $s\ge 0$,
\beao
&&\delta_{(1)}(s)\ge \cdots \ge \delta_{(p)}(s)\,,\\
&&\gamma_{(1)}^\rightarrow(s)\ge \cdots \ge \gamma_{(p)}^\rightarrow(s)\,,\\
&&\gamma_{(1)}^\downarrow(s)\ge \cdots \ge \gamma_{(n)}^\downarrow(s)\,,
\eeao
from the sets
\beao
&&\big\{Z_{(i),np}^2 v_j(s)\,, i=1,\ldots,p\,;j=1,2,\ldots\big\}\,,\notag\\
&&\big\{D_{i}^\rightarrow v_j(s), i=1,\ldots,p\,;j=1,2,\ldots\big\}\,,\notag\\
&&\big\{D_{t}^\downarrow v_j(s), t=1,\ldots,n\,;j=1,2,\ldots\big\}\,,
\eeao
respectively.
%---------------------------------------------------------------------------
\subsection{Approximation of the singular values}\label{sec:mainresult}
%Furthermore, let $\tau_{(1)}(s)  \ge \ldots \ge \tau_{(p)}(s)$ be  the singular values of $\tfrac{1}{2}(\X_n(0)\X_n(s)'+\X_n(s)\X_n(0)' )$ and let $\tilde{\tau}_{(1)}(s) \ge  \ldots \ge\tilde{\tau}_{(p)}(s)$ be the singular values of
%\begin{equation*}
%\tfrac{1}{2}(\X_n(0)\X_n(s)'-\E[\X_n(0)\X_n(s)']+\X_n(s)\X_n(0)' -\E[\X_n(s)\X_n(0)'])
%\end{equation*}
In the following result we povide some useful approximations to the singular values of the sample autocovariance matrices of
the linear model \eqref{eq:1}.
\begin{theorem}\label{thm:mains}
Consider the linear process \eqref{eq:1} under
\begin{itemize}
\item
the \regvar\ condition \eqref{eq:27}
for some $\alpha\in (0,4)$,
\item the centering condition
$\E[Z]=0$ if $\E[|Z|]<\infty$,
\item
the summability condition
\eqref{eq:2a} on the coefficient matrix  $(h_{kl})$,
\item
the growth condition \ref{eq:p} on $(p_n)$ for some $\beta\ge 0$.
\end{itemize}
Then the following statements hold for $s\ge 0$:
\begin{enumerate}
\item We consider two disjoint cases:
$\alpha \in (0,2)$ and $\beta\in (0,\infty)$, or
$\alpha\in [2,4)$ and $\beta$ satisfying \ref{Cbeta}. Then
\begin{equation}\label{eq:mains1}
a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\delta_{(i)}(s)| \cip 0, \quad \nto.
\end{equation}
\item
Assume $\beta\in [0,1]$.
If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta\in (\alpha/2-1,1]$ then
\beao
a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\gamma_{(i)}^\rightarrow(s)| \cip 0, \quad \nto.
\eeao
Assume $\beta>1$. If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta^{-1}\in(\alpha/2-1,1]$. Then
\beao
a_{np}^{-2} \max_{i=1,\ldots,p} |\la_{(i)}(s)-\gamma_{(i)}^\downarrow(s)| \cip 0, \quad \nto.
\eeao
\end{enumerate}
\end{theorem}
\begin{remark}\em
The proof of Theorem~\ref{thm:mains} is given in Heiny et al.~\cite{heiny:mikosch:2016:noniid}. Part (2) of this  result
with more restrictive conditions on the growth rate of $(p_n)$ is contained in Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
These proofs are very technical and lengthy.
\end{remark}
\begin{remark}\em
If we consider a random array $(h_{kl})$ independent of  $(X_{it})$ and
assume that the summability
condition \eqref{eq:2a} holds a.s.
then Theorem~\ref{thm:mains} remains valid conditionally on
$(h_{kl})$, hence  unconditionally in $\P$-probability; see also \cite{davis:mikosch:pfaffel:2015}.\end{remark}
\subsection{Point process \con }
Theorem~\ref{thm:mains}  and arguments similar to the proofs in Davis et al. \cite{davis:mikosch:pfaffel:2015}
enable one to derive the weak \con\ of
the point processes of the normalized singular values. Recall
the representation of the points $(\Gamma_i)$ of a unit rate homogeneous Poisson process on $(0,\infty)$
given in \eqref{eq:Gamma}. For $s\ge 0$, we define the point processes of the normalized singular values:
\begin{equation}\label{eq:ppdef}
N_n^{\lambda,s}=\sum_{i=1}^p \vep_{a_{np}^{-2}(\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s))} \,.
\end{equation}

\begin{theorem}\label{cor:1}
Assume the conditions of Theorem~\ref{thm:mains}.
Then $(N_n^{\lambda,s})$ converge weakly in the space of point measures
with state space $(0,\infty)^{s+1}$ equipped with the vague topology.
If either $\alpha \in (0,2]$, $\E[Z^2]=\infty$ and $\beta \ge 0$,
or $\alpha \in [2,4)$, $\E[Z^2]<\infty$ and \ref{Cbeta} hold then
\begin{equation}\label{eq:pp}
N_n^{\lambda,s} \cid N= \sum_{i=1}^\infty
\sum_{j=1}^{\infty} \vep_{\Gamma_i^{-2/\alpha} (v_j(0),\ldots,v_j(s))}, \qquad \nto.
\end{equation}
\end{theorem}

\begin{proof}
Regular variation of $Z^2$ is equivalent to
\begin{equation}\label{eq:wwa}
n\,p\, \P(a_{np}^{-2} Z^2  \in \cdot ) \civ \mu(\cdot),
\end{equation}
where $\civ$ denotes vague convergence of Radon measures on  $(0,\infty)$ and the \ms\ $\mu$ is given by $\mu(x,\infty)=x^{-\alpha/2}$, $x>0$.
In view of Resnick \cite{resnick:1987}, Proposition~3.21, \eqref{eq:wwa} is equivalent
to the weak \con\ of the following \pp es:
\begin{equation*}
\sum_{i=1}^p\sum_{t=1}^n \vep_{a_{np}^{-2} Z_{it}^2}=\sum_{i=1}^{np} \vep_{a_{np}^{-2} Z^2_{(i),np}}\cid
\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha} }=\widetilde{N}\,,\quad
\nto\,,
\end{equation*}
where the limit $\widetilde{N}$ is a Poisson random measure (PRM) with state space $(0,\infty)$ and
mean measure~$\mu$.
\par
Since $a_{np}^{-2} Z^2_{(p),np}\cip 0$ as $\nto$,
the point processes $\sum_{i=1}^{p} \vep_{a_{np}^{-2} Z^2_{(i),np}}$ converge weakly to the same PRM:
\begin{equation}\label{eq:ppc}
\sum_{i=1}^{p} \vep_{a_{np}^{-2} Z^2_{(i),np}}\cid
\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha} }\,,\quad
\nto\,.
\end{equation}
A continuous mapping argument together with the fact that $\sum_{i=1}^\infty (v_i(s))^2<\infty$ (see  \eqref{eq:tracea})
shows that
\begin{equation*}
\sum_{j=1}^{ \infty}\sum_{i=1}^p\vep_{a_{np}^{-2} Z^2_{(i),np}(v_j(0),\ldots,v_j(s))}\cid
\sum_{j=1}^{ \infty}\sum_{i=1}^\infty\vep_{\Gamma_i^{-2/\alpha}  (v_j(0),\ldots,v_j(s))}\,.
\end{equation*}
If the assumptions of part (1) of Theorem~\ref{thm:mains} are satisfied
an application of \eqref{eq:mains1} (also recalling the definition of $(\delta_{(i)}(s))$) shows that \eqref{eq:ppc} remains
valid with the points  $(a_{np}^{-2} Z^2_{(i),np}(v_j(0),\ldots,v_j(s)))$ replaced by
$(a_{np}^{-2} (\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s))$.
\par
The only cases which are not covered by Theorem~\ref{thm:mains}(1) are $\alpha\in (0,2)$, $\beta=0$ and $\alpha=2$, $\E[Z^2]=\infty$, $\beta\ge0$.
In these cases we get from Theorem~\ref{thm:nagaev} that
\begin{equation*}
 p \, \P(a_{np}^{-2} D^\rightarrow>x ) \sim p\,n\, \P( Z^2>a_{np}^{2} x )\to \mu(x,\infty)\,,\quad x>0\,,
\end{equation*}
i.e., $
 p \, \P(a_{np}^{-2} D^\rightarrow\in \cdot ) \stackrel{v}{\rightarrow} \mu(\cdot)$.
It follows from Lemma~\ref{lem:ppr} that
$\sum_{i=1}^p\vep_{a_{np}^{-2} D_i^\rightarrow}\std \widetilde{N}$. As before,
a continuous mapping argument in combination with
the approximation obtained in Theorem~\ref{thm:mains}(2)
justifies the replacement of the points $(a_{np}^{-2} D_{(i)}^\rightarrow(v_j(0),\ldots,v_j(s)))$ by
$(a_{np}^{-2} (\lambda_{(i)}(0),\ldots,\lambda_{(i)}(s)))$ in the case $\beta\in [0,1]$. If $\beta>1$ one has to work with the
quantities $(D_i^\downarrow)_{i=1,\ldots,n}$ instead of  $(D_i^\rightarrow)_{i=1,\ldots,p}$ and one may follow the same argument as above.
This finishes the proof.
\end{proof}

\section{Some applications}\setcounter{equation}{0}
\subsection{Sample covariance matrices}\label{sec:samplecov}%\setcounter{equation}{0}
The sample covariance matrix $\X_n(0)\X_n(0)'=\X\X'$ is a non-negative definite matrix. Therefore
its eigenvalues and singular values coincide.  Moreover, $v_j=v_j(0)$, $j\ge 1$, are the eigenvalues of $\M=\M(0)$.
\par
Theorem~\ref{thm:mains}(1) yields an approximation of the ordered eigenvalues $(\la_{(i)})$ of $\X\X'$ by
the quantities $(\delta_{(i)})$ which are derived from the order statistics of $(Z_{it}^2)$.
Part (2) of this result provides an approximation of $(\la_{(i)})$ by the quantities
$(\gamma_{(i)}^{\rightarrow/\downarrow})$ which are derived from the order statistics
of the partial sums $(D_i^{\rightarrow/\downarrow})$.
\par
In the following example we illustrate the
quality of the two approximations.
\begin{example}\label{ex:xiao}\em
We choose a Pareto-type \ds\ for $Z$ with density
\beam\label{eq:distrsim}
f_Z(x) =
\left\{\begin{array}{cc}
 \frac{\alpha}{(4|x|)^{\alpha + 1}}\,, & \mbox{if } |x| > 1/4 \\
1\,, & \mbox{otherwise.}
\end{array}\right.
\eeam
We simulated $20,000$ matrices $\X_n$ for $n=1,000$ and $p=200$ whose
iid entries have this density. We assume $\beta=1$.
Note that $\M=\M(0)$ has rank one and $v_1=1$.
The estimated densities of the deviations $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ and
$a_{np}^{-2}(\lambda_{(1)}-Z^2_{(1),np})$  based on the simulations are shown in
Figure~\ref{fig:lambda_comparison}. The approximation error is very small indeed.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.40]{lambda_comparison.pdf}
  \caption{Density of the approximation errors for the eigenvalues of $a_{np}^{-2}\X\X'$.
The entries of $\X$ are iid with density \eqref{eq:distrsim} and $\alpha=1.6$.}
  \label{fig:lambda_comparison}
\end{figure}
According to the theory,
\beao
a_{np}^{-2}\sup_i|D_{(i)}^\rightarrow-\lambda_{(i)}| +a_{np}^{-2}\sup_i |Z^2_{(i),np}-\lambda_{(i)}|\stp 0\,,
\eeao
but  for finite $n$ the $(D_{(i)}^\rightarrow)$ sequence yields a better approximation to $(\la_{(i)})$. By construction,
the considered differences
have a tendency to be positive but Figure~\ref{fig:lambda_comparison} also shows that the median of the
approximation error for $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ is almost zero.
\end{example}
\par
Theorem~\ref{cor:1} and the continuous mapping theorem immediately yield results about
the joint convergence of the largest eigenvalues of the matrices $a_{np}^{-2}\X_n\X_n'$ for $\alpha\in (0,2)$ and
$\alpha\in (2,4)$ when $\beta$ satisfies \ref{Cbeta}.
For fixed $k\ge 1$ one gets
\begin{equation*}
a_{np}^{-2} \big(\la_{(1)},\ldots,\la_{(k)}\big)\cid
\big(d_{(1)},\ldots, d_{(k)}\big)\,,
\end{equation*}
where $d_{(1)}\ge \cdots\ge d_{(k)}$ are the $k$ largest ordered
values of the set $\{\Gamma_i^{-2/\alpha} v_j, i=1,2,\ldots,j=1,\ldots,r\}$.
The continuous mapping theorem yields for $k\ge 1$,
\begin{equation}\label{eq:u}
\dfrac{\la_{(1)}}{\la_{(1)}+\cdots+ \la_{(k)}}\cid
\dfrac{d_{(1)}}{d_{(1)}+\cdots+ d_{(k)}}\,,\quad \nto\,.
\end{equation}

An application of the continuous mapping theorem to  the distributional convergence of the point processes in Theorem~\ref{cor:1} in the spirit of Resnick
\cite{resnick:2007}, Theorem 7.1, also yields the following result; see Davis et al. \cite{davis:mikosch:pfaffel:2015} for a proof and a similar result in the case $\alpha \in (2,4)$.
\begin{corollary}\label{cor:1q} Assume the conditions of Theorem~\ref{thm:mains}.
If $\alpha \in (0,2]$ and $\E[Z^2]=\infty$, then
\begin{equation*}
a_{np}^{-2}\Big(\la_{(1)},\sum_{i=1}^p \la_i\Big) \cid
\Big( v_1\,\Gamma_1^{-2/\alpha}\,,\sum_{j=1}^r v_j\,\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}\Big)\,,
\end{equation*}
where $\Gamma_1^{-2/\alpha}$ is Fr\'echet $\Phi_{\alpha/2}$-distributed.
and $\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}$ has the distribution
of a positive $\alpha/2$-stable random variable.
In particular,
\begin{equation}\label{eq:limit}
\dfrac{\la_{(1)}}{\la_{1}+\cdots+ \la_{p}}\cid
\dfrac{v_1}{\sum_{j=1}^r v_j}\;
\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty\Gamma_i^{-2/\alpha}}\,,\quad
\nto\,.
\end{equation}
\end{corollary}

\begin{remark}\label{rem:4.5}\em
The ratio
\begin{equation*}
\dfrac{\la_{(1)}+\cdots +\la_{(k)}}{\la_1+\cdots+\la_p}, \qquad k\ge 1\,,
\end{equation*}
plays an important role in PCA. It reflects the proportion of the total variance in the data that we can explain by the first $k$ principal components.
It follows from Corollary~\ref{cor:1q} that for fixed $k\ge 1$,
\begin{equation*}
\dfrac{\la_{(1)}+\cdots \la_{(k)}}{\la_1+\cdots+\la_p}\cid
\dfrac{d_{(1)}+\cdots +d_{(k)}} {d_{(1)}+d_{(2)}+\cdots}\,.
\end{equation*}
Unfortunately, the limiting variable does in general not have a clean
form.
An exception is the case when $r=1$; see Example~\ref{exam:separable}.
Also notice that the trace of $\X\X'$ coincides with $\la_1+\cdots+\la_p$.
\end{remark}

To illustrate the theory we consider a simple moving average example
taken from Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
\begin{example} \rm
\begin{figure}[htb!]
  \centering
  \subfigure[iid data]{
    \includegraphics[scale=0.4]{lambda_1_2_Gaps_iid.pdf}
  }
  \subfigure[data from model \eqref{ex:4.6}]{
    \includegraphics[scale=0.4]{lambda_1_2_Gaps.pdf}
  }
  \caption{Distribution function of $(\lambda_{(1)} - \lambda_{(2)})/\lambda_{(1)}$ for iid data (left) and
    data generated from the model \eqref{ex:4.6} (right). In each graph we compare the empirical \ds\ \fct\
(dotted line, based on
1000 simulations of $200 \times 1000$ matrices with $Z$-\ds\ \eqref{eq:distrsim}) with the theoretical curve (solid line).}
  \label{fig:ProbMass}
\end{figure}
Assume that $\alpha\in (0,2)$ and
\begin{equation}\label{ex:4.6}
X_{it}= Z_{it}+ Z_{i,t-1}-2 (Z_{i-1,t}- Z_{i-1,t-1})\,,\quad i,t\in\Z\,.
\end{equation}
In this case, the non-zero entries of $\H$ are
\begin{equation*}
h_{00}=1, h_{01}=1,h_{10}=-2 \quad \mbox{ and }\quad h_{11}=2.
\end{equation*}
Hence $\M=\H\H'$ has the positive eigenvalues
$v_1=8$ and $v_2=2$. The limit point process in \eqref{eq:pp} is
\begin{equation*}
N=\sum_{i=1}^\infty \vep_{8\Gamma_i^{-2/\alpha}}+ \sum_{i=1}^\infty \vep_{2\Gamma_i^{-2/\alpha}}\,,
\end{equation*}
so that
\begin{equation*}
a_{np}^{-2}\,\big(\la_{(1)},\la_{(2)}\big) \cid
\big(8\Gamma_{1}^{-2/\alpha}, 2\Gamma_1^{-2/\alpha}\vee 8
\Gamma_2^{-2/\alpha}\big)\,.
\end{equation*}
Using the fact that $U=\Gamma_1/\Gamma_2$ has a uniform \ds\ on $(0,1)$ we calculate
\begin{equation*}
\P(2\Gamma_1^{-2/\alpha}>8\Gamma_2^{-2/\alpha})= \P(\Gamma_1/\Gamma_2<2^{-\alpha})= 2^{-\alpha} \in (1/4,1).
\end{equation*}
In particular, we have for the normalized spectral gap
\begin{equation*}
a_{np}^{-2} \big(\la_{(1)}-\la_{(2)}\big)\cid
6 \,\Gamma_1^{-2/\alpha} \1_{\{\Gamma_1 4^{\alpha/2 }<\Gamma_2\}}+
8\,
\big(\Gamma_1^{-2/\alpha}-\Gamma_2^{-2/\alpha}\big)\1_{\{\Gamma_14^{\alpha/2
  }> \Gamma_2\}}
\end{equation*}
and for the self-normalized spectral gap (see also Example~\ref{exam:spectralgap} for a detailed analysis)
\begin{equation*}
\begin{split}
\dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}}&\cid
\dfrac{6}{8} \, \1_{\{\Gamma_1 2^{\alpha }<\Gamma_2\}}+
\big(1-(\Gamma_1/\Gamma_2)^{2/\alpha}\big)\1_{\{\Gamma_12^{\alpha
  }> \Gamma_2\}}\\
& =\dfrac{3}{4}\,  \1_{\{U 2^{\alpha }<1\}}+
\big(1-U^{2/\alpha}\big)\1_{\{U2^{\alpha
  }> 1\}}=Y\,.
\end{split}
\end{equation*}
The limit  distribution of the spectral gap
has an atom at $3/4$ with probability $2^{-\alpha}$, i.e.,~$\P(Y=3/4)=2^{-\alpha}$, and
\begin{equation*}
\P(Y\le x)= 1- (1-x)^{\alpha/2},\qquad x\in (0,3/4).
\end{equation*}
In the iid case the limit distribution of the self-normalized spectral gap has distribution function $F(x) = 1- (1-x)^{\alpha/2}$ for $x\in [0,1]$. This means that the atom disappears if the entries are iid. Figure~\ref{fig:ProbMass} compares the distribution function of $Y$
with $F$ for $\alpha=0.6$; the atom at $3/4$ is clearly visible.

Along the same lines, we also have
\begin{equation*}
(a_{np}^{-2}\lambda_{(1)},\lambda_{(2)}/\lambda_{(1)}) \cid(
8\,\Gamma_1^{-2/\alpha},\frac{1}{4}\,\1_{\{U<2^{-\alpha}\}}+U^{2/\alpha}\,\1_{\{U\ge
  2^{-\alpha}\}})\,
\end{equation*}
and hence the limit distribution of $\lambda_{(2)}/\lambda_{(1)}$ is supported on $[1/4,1)$ with  mass of $2^{-\alpha}$ at $1/4$.  The histogram of the ratio $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ based on 1000 replications from the model \eqref{ex:4.6} with noise given by a $t$-distribution with $\alpha=1.5$ degrees of freedom, $n=1000$ and $p=200$ is displayed in Figure \ref{fig:1}.
\begin{figure} [h]
\begin{center}
\includegraphics[scale=.35]{histogram.pdf}
\end{center}
\caption{Histogram based on 1000 replications of $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ from model \eqref{ex:4.6}.}
\label{fig:1}
\end{figure} Observing that $2^{-\alpha}=0.3536\ldots$,
the histogram is remarkably close to what one would expect from a sample from the truncated
uniform \ds , $2^{-\alpha}\, \1_{\{U<2^{-\alpha}\}}+U\, \1_{\{U\ge 2^{-\alpha} \}}$.
The mass of the limiting discrete component of the ratio can be much
larger if one conditions on $a_{np}^{-2}\lambda_{(1)}$ being large.
Specifically, for any $\epsilon\in (0,1/4)$ and $x>0$,
\begin{equation*}
\lim_{n\to\infty} \P(\epsilon<\lambda_{(2)}/\lambda_{(1)}\le 1/4|\lambda_{(1)}>a_{np}^2x)
=\P(\Gamma_1/\Gamma_2\le 2^{-\alpha}|\Gamma_1<(x/8)^{-\alpha/2})=G(x)\,.
\end{equation*}
The function $G$ approaches $1$ as $x\to\infty$ indicating the speed at which the
two largest eigenvalues get linearly related; see Figure \ref{fig:2} for a graph of $G$ in the case $\alpha=1.5$.
\begin{figure} [h]
\begin{center}
\includegraphics[scale=.35]{Hfunction.pdf}
\end{center}
\caption{Graph of $G(x)=\P(\Gamma_1/\Gamma_2\le2^{-\alpha}|\Gamma_1<(x/8)^{-\alpha/2})$ when $\alpha=1.5$.}
\label{fig:2}
\end{figure}
In addition, from Remark~\ref{rem:4.5}, we also have
\begin{equation*}
\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \cid \dfrac{4}{5}\,
\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}}\,.
\end{equation*}
Clearly, the limit \rv\ is stochastically smaller than what one would get in the iid case; see \eqref{eq:limit}.
\end{example}
\begin{example}\label{exam:spectralgap}\rm
The previous example also illustrates the behavior of the two largest eigenvalues in the general case
when the rank $r$ of the matrix $\M$ is larger than one.
We have in general
\begin{equation*}
\dfrac{\lambda_{(2)}}{\lambda_{(1)}}\cid
\frac{v_2}{v_1}\,\1_{\{U<(v_2/v_1)^{\alpha/2}\}}+U^{2/\alpha}\,\1_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
\end{equation*}
In particular, the limiting
{\em self-normalized spectral gap} has \rep
\beao
\dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}} \cid \dfrac{v_1-v_2}{v_1}\,\1_{\{U<(v_2/v_1)^{\alpha/2}\}}+(1-U^{2/\alpha})\,\1_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
\eeao
The limiting variable assumes values in $(0,1-v_2/v_1]$  and has an atom at the right end-point.
This is in contrast to the iid case
and to the case when $r=1$ (hence $v_2=0$) including the case of iid rows and the separable case; see Example~\ref{exam:separable}.
\end{example}
\begin{example}\label{exam:separable}
{\em We consider the separable case when  $h_{kl}=\theta_kc_l$, $k,l \in \Z$, where $(c_l)$, $(\theta_k)$ are
  real sequences such that the conditions on $(h_{kl})$ in
  Theorem~\ref{thm:mains} hold. In this case,
\begin{equation*}
\M= \sum_{l\in \Z} c_l^2\;\; (\theta_i\theta_j)_{i,j \in \Z}\,.
\end{equation*}
Note that $r=1$ with the only non-negative eigenvalue
\beao
v_1=\sum_{l\in \Z} c_l^2\;\;\sum_{k\in \Z}\theta_k^2\,.
\eeao
In this
case, the limiting point process in Theorem~\ref{cor:1} is a PRM on
$(0,\infty)$ with mean measure of $(y,\infty)$ given by $(v_1/
y)^{\alpha/2}$, $y>0$. The normalized eigenvalues have similar asymptotic behavior as
in the case of iid entries. For example, the log-spacings have the same limit as in the iid case for fixed $k$,
\beao
\big(\log \la_{(1)}-\log \la_{(2)},\ldots,\log \la_{(k+1)}-\log \la_{(k)}\big)\std
-\dfrac{2}{\alpha}\,\big(\log(\Gamma_1/\Gamma_2),\ldots,\log (\Gamma_{k}/\Gamma_{k+1})\big)\,.
\eeao
The same observation applies to the ratio of the largest eigenvalue and the trace in the case $\alpha\in (0,2)$:
\beao
\dfrac{\la_{(1)}}{{\rm tr} (\X \X')}=\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \std
\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}}\,.
\eeao
We also mentioned in Example~\ref{exam:spectralgap} that the \ds al limit of the self-normalized spectral gap has no atom
as in the iid case.
}
\end{example}

\subsection{S\&P 500 data}\label{sec:sp500}
We conduct a short analysis of the largest eigenvalues of the univariate log-return time series which compose
the S\&P 500 stock index; see Section~\ref{subsec:1.2} for a description of the data.
Although there is strong empirical evidence that these univariate series have power-law tails
(see Figure~\ref{fig:SP500_tail_indices}) we  do not expect that they
have the same tail index. One way to proceed would be to ignore this fact because the tail indices are
in a close range and the differences are due to large sampling errors for estimating such quantities.  One could also collect \ts\ with similar
tail indices in the same group.
In this case, the dimension $p$ decreases. This grouping would be a rather arbitrary classification method.
We have chosen a third way: to use rank transforms.
This approach has its merits because it aims at standardizing the tails but it also has a major disadvantage: one destroys
the covariance structure underlying the data.
\par
Given a $p\times n$ matrix $(R_{it})_{i = 1, \cdots, p;t=1,\cdots, n}$,
we construct a matrix $\X$ via the rank transforms
\begin{equation*}
  X_{it} = -\Big[
    \log \Big(\frac{1}{n+1} \sum_{\tau=1}^n \1_{\{R_{i\tau} \leq R_{it}\}} \Big)
  \Big]^{-1} \,,\qquad i=1,\ldots,p;t=1,\ldots,n\,.
\end{equation*}
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.550]{EigenRatioSP500_ranks_log_50_shown.pdf}
  \caption{The logarithms of the ratios $\lambda_{(i+1)} / \lambda_{(i)}$ for the S\&P 500 series after rank transform.
We also show the 1, 50 and 99\% quantiles (bottom, middle, top lines, respectively) of the variables
$\log((\Gamma_i/ \Gamma_{i+1})^{2})$. }
  \label{fig:EigenRatio}
\end{figure}
\begin{figure}[htb!]
  \centering
    \includegraphics[scale=0.550]{EigenRatioSP500_log_50_shown.pdf}
  \caption{The logarithms of the ratios $\lambda_{(i+1)} / \lambda_{(i)}$ for the original (non-rank transformed) S\&P 500 log-return data.
We also show the 1, 50 and 99\% quantiles (bottom, middle, top lines, respectively) of the variables
$\log((\Gamma_i/ \Gamma_{i+1})^{2/2.3})$; see also Figure~\ref{fig:EigenRatio} for comparison.}\label{fig:22}
\end{figure}

%\begin{figure}[htb!]
%  \centering
%    \includegraphics[scale=0.550]{EigenRatioSP500_lin2_250_shown.pdf}
%  \caption{{\green Same as Figure \ref{fig:EigenRatio} but on original scale, no logarithm. Choose one.}}
%\end{figure}

If the rows $R_{i1}, \ldots, R_{in}$ were iid (or, more generally, stationary ergodic)
with a continuous \ds\  then
the averages under the logarithm would be  asymptotically uniform
on $(0,1)$ as $n \to \infty$. Hence $X_{it}$ would  be \asy ally standard
Fr\'echet $\Phi_1$-distributed.
In what follows, we assume that the aforementioned univariate \ts\ of the S\&P 500 index have undergone the rank transform and that
their marginal \ds s are close to $\Phi_1$; we
always use the symbol $\X$ for the resulting multivariate series.
\par
In Figure~\ref{fig:EigenRatio} we show the ratios of the consecutive ordered eigenvalues $(\la_{(i+1)}/\la_{(i)})$
of the matrix $\X\X'$. This graph shows the rather surprising fact that the ratios are close to one even for small values $i$.
We also show the 1, 50 and 99 \% quantiles of the variables $((\Gamma_{i}/\Gamma_{i+1})^{2/\alpha})$ calculated from the formula
\beam\label{eq:iid}
\P\big((\Gamma_{i} /\Gamma_{i+1})^{2/\alpha}\le x \big) = x^{i \cdot \alpha/2}, \quad x\in (0,1)\,.
\eeam
For increasing $i$, the \ds\ is concentrated closely to 1, in agreement with the \slln\ which yields
$\Gamma_{i} /\Gamma_{i+1}\stas 1$ as $i\to\infty$.
The \asy\ \ds s \eqref{eq:iid} correspond to the case when the matrix
$\M$ has rank $r=1$. It includes the iid and separable cases;
see Example~\ref{exam:separable}. The shown \asy\ quantiles are in agreement with the rank $r=1$ hypothesis.
\par
For comparison, in Figure~\ref{fig:22} we also show the ratios  $(\la_{(i+1)}/\la_{(i)})$ for the 
non-rank transformed S\&P 500 data and the 1, 50 and 99\% quantiles of the variables
$\log((\Gamma_i/ \Gamma_{i+1})^{2/\alpha})$, where we choose $\alpha=2.3$ motivated by the estimated tail indices in 
Figure~\ref{fig:SP500_tail_indices}.
The two graphs in Figure~\ref{fig:EigenRatio} and Figure~\ref{fig:22} are quite similar but the smallest ratios for the original data are slightly larger than
for the rank-transformed data.

%--------------------------------------------------------------------------------------
\subsection{Sums of squares of sample autocovariance matrices}\label{sec:possemidef}
In this section we consider some additive \fct s
of the squares of $\A_n(s)=\X_n(0)\X_n(s)'$ given by $\A_n(s)\A_n(s)'$ for $s=0,1,\ldots$. By definition of the singular values of a matrix
(see  \eqref{eq:sigma}), the non-negative definite
matrix $\A_n(s)\A_n(s)'$ has eigenvalues $(\la_i^2(s))_{i=1,\ldots,p}$.
\par
The following result is a corollary of Theorem~\ref{thm:mains}.
\begin{proposition}\label{thm:mainstr} Consider the linear process \eqref{eq:1} under
the conditions of Theorem~\ref{thm:mains}. Then the following statements hold for $s\ge 0$:
\begin{enumerate}
\item[$(1)$]
We consider two disjoint cases:
$\alpha \in (0,2)$ and $\beta\in (0,\infty)$, or
$\alpha\in [2,4)$ and $\beta$ satisfying \ref{Cbeta}. Then
\beao
a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-\delta_{(i)}^2(s)| \cip 0, \quad \nto.
\eeao
\item[$(2)$]
Assume $\beta\in [0,1]$.
If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta \in (\alpha/2-1,1]$, then
\beao
a_{np}^{-4} \max_{i=1,\ldots,p} |\la_{(i)}^2(s)-(\gamma_{(i)}^\rightarrow(s))^2| \cip 0, \quad \nto.
\eeao
Assume $\beta>1$. If $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta^{-1} \in (\alpha/2-1,1]$. Then
\beao
a_{np}^{-4} \max_{i=1,\ldots,p} |\la_{(i)}^2(s)-(\gamma_{(i)}^\downarrow(s))^2| \cip 0, \quad \nto.
\eeao
\end{enumerate}
\end{proposition}
To the best of our knowledge, sums of squares of sample autocovariance matrices were used first in the paper by Lam and Yao
\cite{lam:yao}; their \ts\ model is quite different from ours.
\begin{proof}
Part (1).
The proof follows from Theorem~\ref{thm:mains} if we can show that
\beao
a_{np}^{-2}\max_{i=1,\ldots,p} \big(\la_{(i)}(s)+\delta_{(i)}(s) \big)=O_\P(1)\,\quad \nto\,.
\eeao
We have by Theorem~\ref{cor:1},
\beam \label{eq:gdfg}
a_{np}^{-2}\max_{i=1\ldots,p} \la_{(i)}(s)=a_{np}^{-2}\la_{(1)}(s) \cid c\, \xi_{\alpha/2}\,,
\eeam
where $\xi_{\alpha/2}$ has a $\Phi_{\alpha/2}$ \ds . In view of Theorem~\ref{thm:mains}(1) we also have
\beao
a_{np}^{-2}\max_{i=1\ldots,p} \delta_{(i)}(s)\cid c\, \xi_{\alpha/2}\,.
\eeao
Therefore, again using Theorem~\ref{thm:mains}(1), we have
\beao
\lefteqn{a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-\delta_{(i)}^2(s)|}\\
&\le &\big[a_{np}^{-2} \max_{i=1,\ldots,p} |\lambda_{(i)}(s)-\delta_{(i)}(s)|\big]\,
\big[a_{np}^{-2} \max_{i=1,\ldots,p}\big ( |\lambda_{(i)}(s)|+|\delta_{(i)}(s)|\big)\big]
\cip 0, \quad \nto.
\eeao
This proves part (1).\\[1mm]
Part (2). Now assume $\beta\in [0,1]$ and $\alpha \in (0,2]$, $\E[Z^2]=\infty$ or $\alpha\in [2,4)$, $\E [Z^2]<\infty$ and $\beta\in (\alpha/2-1,1]$. Then \eqref{eq:gdfg} is still true and we have by Theorem~\ref{thm:mains}(2) and Theorem~\ref{cor:1}
\beao
a_{np}^{-2}\max_{i=1\ldots,p} \gamma_{(i)}^{\rightarrow}(s)\cid c\, \xi_{\alpha/2}\,.
\eeao
We then have
\beao
\lefteqn{a_{np}^{-4} \max_{i=1,\ldots,p} |\lambda_{(i)}^2(s)-(\gamma_{(i)}^\rightarrow(s))^2|}\\
&\le &\big[a_{np}^{-2} \max_{i=1,\ldots,p} |\lambda_{(i)}(s) -\gamma_{(i)}^\rightarrow(s)|\big]\,
\big[a_{np}^{-2} \max_{i=1,\ldots,p}\big ( \lambda_{(i)}(s)+\gamma_{(i)}^\rightarrow(s)
\big)\big]
\cip 0\,, \qquad \nto.
\eeao
The proof of the remaining part is similar and therefore omitted.
\end{proof}
Now, using Proposition~\ref{thm:mainstr} and a continuous mapping argument, we can show
limit theory for the eigenvalues
\beao
w_{(1)}(s_0,s_1)\ge \cdots \ge w_{(p)}(s_0,s_1)\,,\qquad 0\le s_0\le s_1\,,
\eeao
of the non-negative definite random matrices
\begin{equation}\label{eq:sumA}
\sum_{s=s_0}^{s_1} \A_n(s) \A_n(s)'\,.
\end{equation}

\begin{proposition}\label{prop:sumsmal}
Assume $0\le s_0\le s_1$ and the conditions of Theorem~\ref{thm:mains} hold.
If $\alpha \in (0,4)$ and $\beta\in (0,1] \cap  (\alpha/2-1,1]$ then
\begin{equation*}
a_{np}^{-4} \max_{i=1,\ldots,p} |w_{(i)}(s_0,s_1)-\omega_{(i)}(s_0,s_1)| \cip 0, \quad \nto,
\end{equation*}
where $\omega_{(i)}(s_0,s_1)$ are the ordered values of the set $\{Z_{(i),np}^4 v_j(s_0,s_1), i=1,\ldots,p;j=1,2,\ldots\}$
and $(v_j(s_0,s_1))$ are the ordered eigenvalues of $ \sum_{s=s_0}^{s_1} \M(s)\M(s)'$.
\end{proposition}

\begin{example}\label{exam:additive}{\em
\begin{figure}[htb!]
  \centering
  \subfigure[]
{
    \includegraphics[scale=0.4]{eigen_sum.pdf}
    \label{fig:LamYao:a}
  }
  \subfigure[] {
    \includegraphics[scale=0.375]{eigen_sum_ratio.pdf}
    \label{fig:LamYao:b}
  }
  \caption{The largest eigenvalues of the sums of the squared autocovariance
    matrices compared with the sums of the largest eigenvalues of these matrices
    for the S\&P 500 data for different values $s_1$. The two values are surprisingly close to each other; mind the scale
of the $y$-axis. We also show their ratios.}
  \label{fig:LamYao}
\end{figure}
Recall the separable case from Example~\ref{exam:separable}, i.e.,
$h_{kl}=\theta_kc_l$, $k,l \ge 0$, where $(c_l)$, $(\theta_k)$ are
real sequences such that the conditions on $(h_{kl})$ in
Theorem~\ref{thm:mains} hold.
%Let $\alpha \in (0,2)$ and set
%\begin{equation*}
%\theta = (\theta_0, \theta_1,\theta_2,\ldots)' \quad \mbox{and} \quad
%c= (c_0,c_1,c_2, \ldots)'.
%\end{equation*}
Write $\Theta_{ij}=\theta_i \theta_j$. It is symmetric and has rank one; the only non-zero
eigenvalue is $\gamma_\theta(0)=\sum_{k=0}^\infty \theta_k^2$. Hence $\Theta$  is non-negative definite.
We get from \eqref{eq:m} that
\begin{equation*}
\M(s)=\gamma_c(s)\, \Theta, \quad s\ge 0\,,
\end{equation*}
where
\beao
\gamma_c(s)=\sum_{l=0}^\infty c_lc_{l+s}\,,\qquad s\ge 0\,.
\eeao
The matrix $\M(s)$ has the only non-zero eigenvalue $\gamma_c(s)\gamma_\theta(0)$.
The factors $(\gamma_c(s))$ can be positive or negative; they constitute the autocovariance \fct\ of a stationary linear process
with coefficients $(c_l)$.
Accordingly, $\M(s)$ is either non-negative or non-positive definite. %This explains why a symmetrization of $\M(s)$ is irrelevant.
Now we consider the non-negative definite matrix
\beao
\sum_{s=s_0}^{s_1} \M(s)\,\M(s)'= \sum_{s=s_0}^{s_1}\gamma_c^2(s)\,\Theta\Theta'\,.
\eeao
This matrix has rank $1$ and its largest eigenvalue is given by
\begin{equation*}
C_{c,\theta}(s_0,s_1)=\sum_{s=s_0}^{s_1}\gamma_c^2(s)\,\gamma_\theta^2(0)\,.
\end{equation*}
An application of
Proposition~\ref{prop:sumsmal} yields that the ordered eigenvalues of $a_{np}^{-4}\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'$
are uniformly approximated by the quantities
\begin{equation}\label{eq:drtgdfg}
a_{np}^{-4} Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\,,\qquad  i=1,\ldots,p\,.
\end{equation}
Since
\beao
C_{c,\theta}(s_0,s_1)= \sum_{i=s_0}^{s_1} C_{c,\theta}(i,i)
\eeao
one gets the remarkable property that
\beao
&&a_{np}^{-4}\max_{i=1,\ldots,p} \Big| \la_{(i)}\Big(\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'\Big)- Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\Big|\\
&=&a_{np}^{-4}\max_{i=1,\ldots,p} \Big|\sum_{s=s_0}^{s_1}\la_{(i)}(\A_n(s)\A_n(s)')- Z_{(i),np}^4 C_{c,\theta}(s_0,s_1)\Big|+o_P(1)\,.
\eeao
In particular, for $s_1\ge s_0$ we get the weak \con\ of the \pp es towards a PRM:
\beao
\sum_{i=1}^p \varepsilon_{a_{np}^{-4}\Big(\la_{i}\Big(\sum_{s=s_0}^{s_0}\A_n(s)\A_n(s)'\big)\,,\ldots,\la_{i}\big(\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'\big)\Big)}
\std \sum_{i=1}^\infty \varepsilon_{\Gamma_i^{-4/\alpha} \Big(C_{c,\theta}(s_0,s_0),\ldots,C_{c,\theta}(s_0,s_1)\Big)}\,.
\eeao

}
\end{example}
\begin{example}\em
In Figure~\ref{fig:LamYao} we calculate the largest eigenvalues
$\la_{(1)}\big(\sum_{s=0}^{s_1}\A_n(s)\A_n(s)'\big)$ for $s_1=0,\ldots,5$ as well as the sums of the largest eigenvalues
$\sum_{s=0}^{s_1}\la_{(1)}\big(\A_n(s)\A_n(s)'\big)$
the log-return series from the S\&P 500 index described in Section~\ref{subsec:1.2}. The data are not rank-transformed.
We notice that
the two values are surprisingly close across the values $s_0=0,\ldots,5$. This phenomenon
could be explained by the structure of the eigenvalues in Example~\ref{exam:additive}.
Also note that the largest eigenvalue $\A_n(0)\A_n(0)'$ makes a major contribution to the values in Figure~\ref{fig:LamYao};
the contribution of the squares $\A_n(s)\A_n(s)'$, $s=1,\ldots,5,$ to the largest eigenvalue of the sum of squares is less substantial.
%Figure~\ref{fig:LamYao}(b) shows the same calculations for the matrices studied in Example~\ref{exam:4.6cont}. Here cancellation is possible (see \eqref{eq:Xrgsdf}) which is indicated by the fact that the red line does not follow the blue line as well as in (a).
\end{example}
%\section{\red Some conclusions?}\setcounter{equation}{0}
%{\em Say something about EVT and largest eigenvalues}

\ifx\phdthesis\undefined
\appendix
\else
\begin{subappendices}
\fi

\section{Auxiliary results}\label{appendix:A}\setcounter{equation}{0}

Let $(Z_i)$ be iid copies of $Z$ whose distribution satisfies
\begin{equation*}
\P(Z>x)\sim p_+ \dfrac{L(x)}{x^{\alpha}}\quad\mbox{and}\quad  \P(Z\le -x)\sim p_-
\dfrac{L(x)}{x^{\alpha}} \,,\quad \xto\,,
\end{equation*}
 for some tail index $\alpha>0$,
where $p_+,p_-\ge 0$ with $p_++p_-=1$ and $L$ is a slowly varying function. If $\E[|Z|]<\infty$ also assume $\E[Z]=0$. The product $Z_1Z_2$ is regular varying with the same index $\alpha$ and $\P(|Z_1Z_2|>x)= x^{-\alpha} L_1(x)$, where $L_1$ is slowly varying function different from $L$;
see Embrechts and Goldie \cite{embrechts:goldie:1980}.
Write
\begin{equation*}
S_n=Z_1+\cdots +Z_n\,,\quad n\ge 1,
\end{equation*} and consider a sequence $(a_n)$ such that $\P(|Z|>a_n)\sim n^{-1}$.

\subsection{Large deviation results}
The following theorem can be found in
Nagaev \cite{nagaev:1979} and Cline and Hsing
\cite{cline:hsing:1998} for $\alpha>2$ and $\alpha\le 2$,
respectively; see also  Denisov et al.~\cite{denisov:dieker:shneer:2008}.
\begin{theorem}\label{thm:nagaev}
Under the assumptions on the iid sequence $(Z_t)$
given above the following relation holds
\begin{equation*}
\sup_{x\ge c_n}\left|\dfrac{\P(S_n>x )}{n\P(|Z|>x)} -p_+ \right|\to 0\,,
\end{equation*}
where $(c_n)$ is any sequence satisfying $c_n/a_n\to  \infty$ for
$\alpha\le 2$ and $c_n\ge \sqrt{(\alpha-2)n\log n}$ for $\alpha>2$.
\end{theorem}

\subsection{A point process convergence result}
Assume that the conditions at the beginning of Appendix \ref{appendix:A} hold.
Consider a sequence of iid copies $(S_{n}^{(t)})_{t=1,2,\ldots}$
of $S_n$ and the sequence of point processes
\begin{equation*}
N_n= \sum_{t=1}^{p} \vep_{a_{np}^{-1} S_{n}^{(t)}}, \quad n=1,2,\ldots\,,
\end{equation*}
for an integer sequence $p=p_n\to\infty$. We assume that the state space of the
point processes $N_n$ is $\overline{\R}_0=[\R\cup\{\pm \infty\}]\backslash \{0\}$.
\begin{lemma}\label{lem:ppr}
Assume $\alpha \in (0,2)$ and the
conditions of  Appendix~\ref{appendix:A}
on the iid sequence $(Z_t)$ and the normalizing sequence $(a_n)$. Then the limit relation
$N_n\cid N$ holds in the space of point measures on $\overline{\R}_0$
equipped with the vague topology (see
\cite{resnick:1987,resnick:2007})
for a  Poisson random measure $N$ with state space $\overline{\R}_0$ and intensity measure $\mu_\alpha(dx)=\alpha |x|^{-\alpha-1} (p_+ \1_{\{x>0\}}+ p_- \1_{\{x<0\}}) dx$.
\end{lemma}
\begin{proof}
According to Resnick \cite{resnick:1987}, Proposition 3.21, we need to
show that
$p\, \P(a_{np}^{-1}S_n\in \cdot)\civ \mu_\alpha
$,
where $\civ$ denotes vague convergence of Radon measures on  $\overline{\R}_0$.
Observe that we have $a_{np}/a_n\to\infty$ as $\nto$. This fact and
$\alpha\in (0,2)$ allow one to apply
Theorem~\ref{thm:nagaev}:
\begin{equation*}
\dfrac{\P( S_n >x a_{np})}{n\,\P(|Z|>  a_{np})}\to p_+ x^{-\alpha}
\quad \mbox{and}\quad \dfrac{\P( S_n \le -x a_{np})}{n\,\P(|Z|>
  a_{np})}\to p_-\, x^{-\alpha}\,,\quad x>0\,.
\end{equation*}
On the other hand, $n\,\P(|Z|>  a_{np})\sim p^{-1}$ as $\nto$.
This proves the lemma.
\end{proof}

\ifx\phdthesis\undefined
\else
\end{subappendices}
\fi

### 2017-07-21 papers/DavisMikoschHeinyXie/s21.tex
Add the Extremes paper.

%last edited 10.11.2015
\documentclass[11pt,reqno]{amsart}
\usepackage{amsmath,epsfig,graphicx,color}
%\usepackage[english,ngerman]{babel}
%\usepackage[latin1]{inputenc}
\usepackage{ifthen}
%\usepackage{dsfont}
%\usepackage{shadethm}
%\usepackage{framed}
%\usepackage{pstricks}
%\usepackage{graphics}
%\usepackage{tocbibind}
%\usepackage{showkeys}
%\usepackage{units}
\usepackage{amssymb,latexsym}
\usepackage{subfigure}
%\usepackage{amsfonts}
%\usepackage{makeidx,showidx}
%\usepackage[sc]{mathpazo}
%\linespread{1.05}
%\usepackage[backref=page]{hyperref}
\usepackage{hyperref}
\hypersetup{
urlcolor=black,
  menucolor=black,
  citecolor=black,
  anchorcolor=black,
  filecolor=black,
  linkcolor=black,
  colorlinks=true,
}
\newcommand{\rep}{representation}
\newcommand{\ld}{large deviation}
\renewcommand{\H}{{\mathbf H}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\cmt}{continuous mapping theorem}
\newcommand{\fct}{function}
\newcommand{\slvary}{slowly varying}
\newcommand{\regvar}{regular variation}
\newcommand{\regvary}{regularly varying}
\newcommand{\st}{such that}
\newcommand{\stas}{\stackrel{\rm a.s.}{\rightarrow}}
\newcommand{\std}{\stackrel{\rm d}{\rightarrow}}
\newcommand{\stp}{\stackrel{\P}{\rightarrow}}
\newcommand{\la}{\lambda}
\newcommand{\ds}{distribution}
\newcommand{\beao}{\begin{eqnarray*}}
\newcommand{\eeao}{\end{eqnarray*}}
\newcommand{\beam}{\begin{eqnarray}}
\newcommand{\eeam}{\end{eqnarray}}
\newcommand{\red}{\color{darkred}}
\newcommand{\blue}{\color{darkblue}}
\newcommand{\green}{\color{darkgreen}}
\newcommand{\black}{\color{black}}
\definecolor{darkblue}{rgb}{.1, 0.1,.8}
\definecolor{darkgreen}{rgb}{0,0.8,0.2}
\definecolor{darkred}{rgb}{.8, .1,.1}
\textwidth 6.50in
\topmargin -0.50in
\oddsidemargin 0in
\evensidemargin 0in
\textheight 9.00in
%\pagestyle{plain}

\newcommand{\wt}{\widetilde}
\newcommand{\bco}{\begin{corrolary}}
\newcommand{\eco}{\end{corrolary}}
\newcommand{\wrt}{with respect to}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\1}{\mathbf{1}}
%\newcommand{\1}{\mathds{1}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\Fo}{\bar{F}}
\renewcommand{\b}[1]{\boldsymbol{#1}}
\newcommand{\Rq}{\mkern 1.5mu\overline{\mkern-1.5mu\R\mkern-3.0mu}\mkern 1.5mu}
\newcommand{\Frechet}{Fr\'{e}chet }
\renewcommand{\Finv}{F^{\gets}}
\DeclareMathOperator{\e}{e}
\newcommand{\inv}[1]{#1^{\gets}}
\newcommand{\x}{{\mathbf x}}
\newcommand{\y}{{\mathbf y}}
\newcommand{\X}{{\mathbf X}}
\newcommand{\Y}{{\mathbf Y}}
\newcommand{\M}{{\mathbf M}}
\newcommand{\bfZ}{{\mathbf Z}}
\newcommand{\0}{\boldsymbol{0}}
\newcommand{\dint}{\,\mathrm{d}}
\newcommand{\mB}{\mathcal{B}}
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\twonorm}[1]{\|#1\|_2}
\newcommand{\inftynorm}[1]{\|#1\|_\infty}
\newcommand{\frobnorm}[1]{\|#1\|_F}
\newcommand{\vep}{\varepsilon}
\newcommand{\nto}{n \to \infty}
\newcommand{\xto}{x \to \infty}
\newcommand{\lhs}{left-hand side}
\newcommand{\rhs}{right-hand side}
\newcommand{\ts}{time series}
\newcommand{\tsa}{\ts\ analysis}
\newcommand{\fidi}{finite-dimensional distribution}
\newcommand{\rv}{random variable}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\Zbar}{\overline{Z}}
\newcommand{\diag}{\operatorname{diag}}
%\renewcommand*{\backref}[1]{}
%\renewcommand*{\backrefalt}[4]{%
%    \ifcase #1 (Not cited.)%
%    \or        (Cited on page~#2.)%
%    \else      (Cited on pages~#2.)%
%    \fi}

\newcommand{\4}{\mathchoice{\mskip1.5mu}{\mskip1.5mu}{}{}}
\newcommand{\5}{\mathchoice{\mskip-1.5mu}{\mskip-1.5mu}{}{}}
\newcommand{\2}{\penalty250\mskip\thickmuskip\mskip-\thinmuskip} % after comma

\newcommand{\levy}{L\'evy}
\newcommand{\slln}{strong law of large numbers}
\newcommand{\clt}{central limit theorem}
\newcommand{\sde}{stochastic differential equation}
\newcommand{\It}{It\^o}
\newcommand{\sta}{St\u aric\u a}
\newcommand{\ex}{{\rm e}\,}

\def\theequation{\thesection.\arabic{equation}}
\def\tag{\refstepcounter{equation}\leqno }
\def\neqno{\refstepcounter{equation}\leqno(\thesection.\arabic{equation})}

\newtheorem{lemma}{Lemma}[section]
\newtheorem{figur}[lemma]{Figure}
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{proposition}[lemma]{Proposition}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{example}[lemma]{Example}
\newtheorem{exercise}[lemma]{Exercise}
\newtheorem{remark}[lemma]{Remark}
\newtheorem{fig}[lemma]{Figure}
\newtheorem{tab}[lemma]{Table}
\newtheorem{conjecture}[lemma]{Conjecture}

\newcommand{\cid}{\stackrel{d}{\rightarrow}}
\newcommand{\cip}{\stackrel{\P}{\rightarrow}}
\newcommand{\civ}{\stackrel{v}{\rightarrow}}
\newcommand{\cas}{\stackrel{\rm a.s.}{\rightarrow}}
\newcommand{\simsl}{\stackrel{sl.}{\sim}}

\newcommand{\A}{{\mathbf A}}
\newcommand{\var}{{\rm var}}
\newcommand{\med}{{\rm med}}
\newcommand{\cov}{{\rm cov}}
\newcommand{\corr}{{\rm corr}}
\newcommand{\as}{{\rm a.s.}}
\newcommand{\io}{{\rm i.o.}}
\newcommand{\Holder}{H\"older}
\newcommand{\evt}{extreme value theory}
\newcommand{\cadlag}{c\`adl\`ag}
\newcommand{\pp}{point process}
\newcommand{\con}{convergence}
\newcommand{\seq}{sequence}
\newcommand{\ms}{measure}
\newcommand{\asy}{asymptotic}
%--------------------------------------------------------------------------------------------
\begin{document}
\today
%\bibliographystyle{acm}
\title[Extreme value analysis for the sample autocovariance matrices of time series]
{Extreme value analysis for the sample autocovariance matrices of heavy-tailed multivariate time series}
\thanks{Richard Davis
was supported by ARO MURI grant W911NF-12-1-0385. Thomas Mikosch's and Johannes Heiny's research is partly supported by the Danish Research Council Grant DFF-4002-00435 ``Large random matrices with heavy tails and dependence''.}
\author[Richard A. Davis]{Richard A. Davis}
\author[Johannes Heiny]{Johannes Heiny}
\author[Thomas Mikosch]{Thomas Mikosch}
\author[Xiaolei Xie]{Xiaolei Xie}
\address{Department of Statistics,
Columbia University,
1255 Amsterdam Ave.
New York, NY 10027, U.S.A.}
\email{rdavis@stat.columbia.edu\,, www.stat.columbia.edu/$\sim$rdavis}
\address{Department  of Mathematics,
University of Copenhagen,
Universitetsparken 5,
DK-2100 Copenhagen,
Denmark}
\email{johannes.heiny@math.ku.dk}
\email{mikosch@math.ku.dk\,, www.math.ku.dk/$\sim$mikosch}
\email{xie@math.ku.dk}

\maketitle

\input{PaperBody}

\section*{Acknowledgments}\setcounter{equation}{0}

We thank Olivier Wintenberger for reading the manuscript and fruitful discussions.

%\bibliography{libraryjohannes}
\begin{thebibliography}{10}

\bibitem{anderson:1963}
{\sc Anderson, T.~W.}
\newblock Asymptotic theory for principal component analysis.
\newblock {\em Ann. Math. Statist. 34\/} (1963), 122--148.

\bibitem{auffinger:arous:peche:2009}
{\sc Auffinger, A., Ben~Arous, G., and P{\'e}ch{\'e}, S.}
\newblock Poisson convergence for the largest eigenvalues of heavy tailed
  random matrices.
\newblock {\em Ann. Inst. Henri Poincar\'e Probab. Stat. 45}, 3 (2009),
  589--610.

\bibitem{bai:silverstein:2010}
{\sc Bai, Z., and Silverstein, J.~W.}
\newblock {\em Spectral Analysis of Large Dimensional Random Matrices},
  second~ed.
\newblock Springer Series in Statistics. Springer, New York, 2010.

\bibitem{baisilv}
{\sc Bai, Z.~D., Silverstein, J.~W., and Yin, Y.~Q.}
\newblock A note on the largest eigenvalue of a large-dimensional sample
  covariance matrix.
\newblock {\em J. Multivariate Anal. 26}, 2 (1988), 166--168.

\bibitem{belinschi:dembo:guionnet:2009}
{\sc Belinschi, S., Dembo, A., and Guionnet, A.}
\newblock Spectral measure of heavy tailed band and covariance random matrices.
\newblock {\em Comm. Math. Phys. 289}, 3 (2009), 1023--1055.

\bibitem{belitski}
{\sc Belitski{\u\i}, G.~R., and Lyubich, Y.~I.}
\newblock {\em Matrix Norms and their Applications}, vol.~36 of {\em Operator
  Theory: Advances and Applications}.
\newblock Birkh\"auser Verlag, Basel, 1988.

\bibitem{arous:guionnet:2008}
{\sc Ben~Arous, G., and Guionnet, A.}
\newblock The spectrum of heavy tailed random matrices.
\newblock {\em Comm. Math. Phys. 278}, 3 (2008), 715--751.

\bibitem{bhatia:1997}
{\sc Bhatia, R.}
\newblock {\em Matrix Analysis}, vol.~169 of {\em Graduate Texts in
  Mathematics}.
\newblock Springer-Verlag, New York, 1997.

\bibitem{bingham:goldie:teugels:1987}
{\sc Bingham, N.~H., Goldie, C.~M., and Teugels, J.~L.}
\newblock {\em Regular Variation}, vol.~27 of {\em Encyclopedia of Mathematics
  and its Applications}.
\newblock Cambridge University Press, Cambridge, 1987.

\bibitem{cline:hsing:1998}
{\sc Cline, D. B.~H., and Hsing, T.}
\newblock Large deviation probabilities for sums of random variables with heavy
  or subexponential tails.
\newblock {\em Technical report. Statistics Dept., Texas A\&M University.\/}
  (1998).

\bibitem{davis:mikosch:pfaffel:2015}
{\sc Davis, R.~A., Mikosch, T., and Pfaffel, O.}
\newblock Asymptotic theory for the sample covariance matrix of a heavy-tailed
  multivariate time series.
\newblock {\em Stochastic Process. Appl.\/} (2015).

\bibitem{davis:pfaffel:stelzer:2014}
{\sc Davis, R.~A., Pfaffel, O., and Stelzer, R.}
\newblock Limit theory for the largest eigenvalues of sample covariance
  matrices with heavy-tails.
\newblock {\em Stochastic Process. Appl. 124}, 1 (2014), 18--50.

\bibitem{dehaan:ferreira:2006}
{\sc de~Haan, L., and Ferreira, A.}
\newblock {\em Extreme Value Theory: An Introduction}.
\newblock Springer Series in Operations Research and Financial Engineering.
  Springer, New York, 2006.

\bibitem{denisov:dieker:shneer:2008}
{\sc Denisov, D., Dieker, A.~B., and Shneer, V.}
\newblock Large deviations for random walks under subexponentiality: the
  big-jump domain.
\newblock {\em Ann. Probab. 36}, 5 (2008), 1946--1991.

\bibitem{elkaroui:2003}
{\sc El~Karoui, N.}
\newblock On the largest eigenvalue of wishart matrices with identity
  covariance when n,p and p/n tend to infinity.
\newblock {\em Available at {\tt http://arxiv.org/abs/math/0309355}\/} (2003).

\bibitem{embrechts:goldie:1980}
{\sc Embrechts, P., and Goldie, C.~M.}
\newblock On closure and factorization properties of subexponential and related
  distributions.
\newblock {\em J. Austral. Math. Soc. Ser. A 29}, 2 (1980), 243--256.

\bibitem{embrechts:klueppelberg:mikosch:1997}
{\sc Embrechts, P., Kl{\"u}ppelberg, C., and Mikosch, T.}
\newblock {\em
  \href{http://www.springer.com/mathematics/quantitative+finance/book/978-3-540-60931-5}{Modelling
  Extremal Events for Insurance and Finance}}, vol.~33 of {\em Applications of
  Mathematics (New York)}.
\newblock Springer, Berlin, 1997.

\bibitem{embrechts:veraverbeke:1982}
{\sc Embrechts, P., and Veraverbeke, N.}
\newblock Estimates for the probability of ruin with special emphasis on the
  possibility of large claims.
\newblock {\em Insurance Math. Econom. 1}, 1 (1982), 55--72.

\bibitem{feller}
{\sc Feller, W.}
\newblock {\em An Introduction to Probability Theory and its Applications.
  {V}ol. {II}}.
\newblock John Wiley \& Sons, Inc., New York-London-Sydney, 1966.

\bibitem{geman}
{\sc Geman, S.}
\newblock A limit theorem for the norm of random matrices.
\newblock {\em Ann. Probab. 8}, 2 (1980), 252--261.

\bibitem{heiny:mikosch:2015:iid}
{\sc Heiny, J., and Mikosch, T.}
\newblock Eigenvalues and eigenvectors of heavy-tailed sample covariance
  matrices with general growth rates: the iid case.
\newblock {\em Submitted\/} (2015).

\bibitem{heiny:mikosch:2016:noniid}
{\sc Heiny, J., Mikosch, T., and Davis, R.~A.}
\newblock Limit theory for the singular values of the sample autocovariance
  matrix function of multivariate time series.
\newblock {\em Work in progress\/} (2015).

\bibitem{horn}
{\sc Horn, R.~A., and Johnson, C.~R.}
\newblock {\em Matrix Analysis}, second~ed.
\newblock Cambridge University Press, Cambridge, 2013.

\bibitem{johansson}
{\sc Johansson, K.}
\newblock Universality of the local spacing distribution in certain ensembles
  of {H}ermitian {W}igner matrices.
\newblock {\em Comm. Math. Phys. 215}, 3 (2001), 683--705.

\bibitem{johnstone:2001}
{\sc Johnstone, I.~M.}
\newblock On the distribution of the largest eigenvalue in principal components
  analysis.
\newblock {\em Ann. Statist. 29}, 2 (2001), 295--327.

\bibitem{lam:yao}
{\sc Lam, C., and Yao, Q.}
\newblock Factor modeling for high-dimensional time series: inference for the
  number of factors.
\newblock {\em Ann. Statist. 40}, 2 (2012), 694--726.

\bibitem{muirhead}
{\sc Muirhead, R.~J.}
\newblock {\em Aspects of Multivariate Statistical Theory}.
\newblock John Wiley \& Sons, Inc., New York, 1982.
\newblock Wiley Series in Probability and Mathematical Statistics.

\bibitem{nagaev:1979}
{\sc Nagaev, S.~V.}
\newblock Large deviations of sums of independent random variables.
\newblock {\em Ann. Probab. 7}, 5 (1979), 745--789.

\bibitem{resnick:2007}
{\sc Resnick, S.~I.}
\newblock {\em Heavy-Tail Phenomena: Probabilistic and Statistical Modeling}.
\newblock Springer Series in Operations Research and Financial Engineering.
  Springer, New York, 2007.

\bibitem{resnick:1987}
{\sc Resnick, S.~I.}
\newblock {\em Extreme Values, Regular Variation and Point Processes}.
\newblock Springer Series in Operations Research and Financial Engineering.
  Springer, New York, 2008.
\newblock Reprint of the 1987 original.

\bibitem{shores}
{\sc Shores, T.~S.}
\newblock {\em Applied Linear Algebra and Matrix Analysis}.
\newblock Undergraduate Texts in Mathematics. Springer, New York, 2007.

\bibitem{soshnikov:2004}
{\sc Soshnikov, A.}
\newblock Poisson statistics for the largest eigenvalues of {W}igner random
  matrices with heavy tails.
\newblock {\em Electron. Comm. Probab. 9\/} (2004), 82--91 (electronic).

\bibitem{soshnikov:2006}
{\sc Soshnikov, A.}
\newblock Poisson statistics for the largest eigenvalues in random matrix
  ensembles.
\newblock In {\em Mathematical physics of quantum mechanics}, vol.~690 of {\em
  Lecture Notes in Phys.} Springer, Berlin, 2006, pp.~351--364.

\bibitem{tao09b}
{\sc Tao, T., and Vu, V.}
\newblock Random matrices: universality of local eigenvalue statistics up to
  the edge.
\newblock {\em Comm. Math. Phys. 298}, 2 (2010), 549--572.

\bibitem{tracy:widom:2012}
{\sc Tracy, C.~A., and Widom, H.}
\newblock Distribution functions for largest eigenvalues and their
  applications.
\newblock In {\em Proceedings of the {I}nternational {C}ongress of
  {M}athematicians, {V}ol. {I} ({B}eijing, 2002)\/} (2002), Higher Ed. Press,
  Beijing, pp.~587--596.

\end{thebibliography}
\end{document}
%Use this \end{document} to get the version without the proof after the bibliography

\section{A result for the largest singular value in the iid case}\setcounter{equation}{0}

Next we briefly state a rather general version of a result 
in Heiny and Mikosch \cite{heiny:mikosch:2015:iid} which will be relevant when we consider dependent entries of $\X$.

The singular values of a matrix $A$ are the positive roots of the eigenvalues of $AA'$. The spectral norm $\twonorm{\cdot}$ of $A$ is defined as its largest singular value.

 We consider a double array $(X_{it})_{i,t \in \Z}$ 
of iid random variables satisfying the \regvar\ condition \eqref{eq:regvar} for some  
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E X=0$. In our proofs we will need all $p\times n$ blocks that appear in the $X$-array. For $u,s\in \Z$, write 
\beao%\label{eq:Ygsrg}
\X(u,s)=\X_n(u,s)&=&(X_{i-u,t-s})_{i=1,\ldots,p;t=1,\ldots,n}\,,\\
(\X(0,0)\X(u,s)')_{ij}& =& \sum_{t=1}^n X_{i,t} \,X_{j-u,t-s},\qquad i,j=1,\ldots,p\,.
\eeao

\begin{theorem}\label{thm:iidjohannes}
Assume $u,s\in \Z$. Let $\xi_{\alpha/2}$ be a $\Phi_{\alpha/2}$-distributed \rv .
\begin{enumerate}
\item Assume one of the following conditions:
\begin{itemize}
\item $\alpha \in (0,2)$ and \ref{eq:p} for $\beta\ge 0$, 
\item $\alpha \in [2,4)$ and \ref{eq:p} for $\beta$ such that $\min(\beta,\beta^{-1})\in (\alpha/2-1,1]$.
\end{itemize}
Then 
\begin{equation*}
a_{np}^{-2} \twonorm{\X(0,0)\X(u,s)'} \cid 0 \,\1(s\ne 0)+ \xi_{\alpha/2}\,\1(s=0)\,.
\end{equation*}
For $\gamma<\alpha/2$,
the \seq\ $(a_{np}^{-2\gamma} \twonorm{\X(0,0)\X(u,s)'}^\gamma)_{n\ge 1}$ is uniformly integrable.
\item 
Assume $\alpha \in (2,4)$ and \ref{eq:p} for $\beta\in [0,1]$. 
Then 
\begin{equation*}
a_{np}^{-2} \twonorm{\X(0,0)\X(u,s)'-\E[\X(0,0)\X(u,s)']} \cid 0 \,\1(s\ne 0)+ \xi_{\alpha/2}\,\1(s=0)\,.
\end{equation*}
For $\gamma<\alpha/2$,
the \seq\ $(a_{np}^{-2\gamma} \twonorm{\X(0,0)\X(u,s)'- \E[\X(0,0)\X(u,s)']}^\gamma)_{n\ge 1}$ is uniformly integrable.
\end{enumerate}
\end{theorem}

\section{Further Karamata theory}\setcounter{equation}{0}

\begin{proposition}
Let $(c_n)$ be the threshold sequence in Theorem~\ref{thm:nagaev} for a
given $\alpha>0$,  and let $(d_n)$ be such that
$d_n/c_n\to\infty$ for $\alpha>2$ and $d_n=c_n$ for $\alpha\le 2$. Assume $\gamma>\alpha$. 
Then we have  for $\delta \in (0,1)$ and a sequence $x_n\ge d_n$
\begin{equation}\label{eq:equiv}
\E[|x_n^{-1} S_n|^\gamma \1_{\{\delta x_n \le |S_n|\le x_n\}}] \sim \frac{\alpha}{\gamma -\alpha} (1- \delta^{\gamma-\alpha}) n \P(|Z|>x_n), \quad \nto.
\end{equation}
Moreover, 
\begin{equation}\label{eq:iec}
\E[|x_n^{-1}S_n|^\gamma \1_{\{d_n \le  |S_n|\le x_n
  \}}]\le   \frac{\alpha}{\gamma-\alpha}\,n \,\P(|Z|>x_n)\,,
\end{equation}
and 
\begin{equation}\label{eq:iec1}
\E[|x_n^{-1}S_n|^\gamma \1_{\{|S_n|\le x_n
  \}}] \le 	\frac{\alpha}{\gamma-\alpha}\,n \,\P(|Z|>x_n) + \Big(\frac{d_n}{x}\Big)^\gamma \,.
\end{equation}
\end{proposition}

\begin{proof}
We use the notation $Y_n:=|x_n^{-1} S_n|$. Since $Y_n^\gamma \1_{\{Y_n\le 1\}}$ is a positive random variable one can write 
\begin{equation*}
\E[Y_n^\gamma \1_{\{\delta \le Y_n \le 1\}}]=\int_0^\infty \P(Y_n^\gamma \1_{\{\delta \le Y_n \le 1\}}>y) \dint y.
\end{equation*}
The probability inside the integral is 
\begin{equation*}
\begin{split}
\P(Y_n^\gamma \1_{\{\delta \le Y_n \le 1\}}>y) &= \P(Y_n > \max \{y^{1/\gamma},\delta \}, y\le 1)\\
&= \begin{cases}
0 & \text{if } y > 1,\\ 
\P(Y_n>y^{1/\gamma}) - \P(Y_n>1) & \text{if } 1 \ge y \ge \delta^\gamma,\\
\P(Y_n>\delta) - \P(Y_n>1) & \text{if } y \le \delta^\gamma.
\end{cases}
\end{split}
\end{equation*}
Therefore, using the uniform convergence result in Theorem~\ref{thm:nagaev}, we conclude that 
\begin{equation*}
\begin{split}
\int_0^\infty \P(Y_n^\gamma \1_{\{\delta \le Y_n \le 1\}}>y) \dint y&=
\delta^\gamma (\P(Y_n>\delta)- \P(Y_n>1))\\
&\quad + \int_{\delta^\gamma}^1 \P(Y_n>y^{1/\gamma}) \, \dint y- (1-\delta^\gamma)  \P(Y_n>1)\\
&=\delta^\gamma \P(|S_n/x_n|>\delta)- \P(|S_n/x_n|>1) \\
&\quad + \int_{\delta^\gamma}^1 \frac{\P(|S_n/x_n|>y^{1/\gamma})}{n \P(|Z/x_n|> y^{1/\gamma})}\,
\, [n \P(|Z/x_n|> y^{1/\gamma})]\, \dint y\\
&\sim \delta^{\gamma-\alpha} n \P(|Z|>x_n) -n \P(|Z|>x_n) + \int_{\delta^\gamma}^1 y^{-\frac{\alpha}{\gamma}} \dint y \, n \P(|Z|>x_n)\\
&= \frac{\alpha}{\gamma-\alpha} (1- \delta^{\gamma-\alpha}) n \P(|Z|>x_n)           \,,\quad \nto\,.
\end{split}
\end{equation*}
The proof of \eqref{eq:iec} is analogous. For \eqref{eq:iec1} we find that
\begin{equation*}
\begin{split}
\E[|x_n^{-1}S_n|^\gamma \1_{\{|S_n|\le x
  \}}] &=  \E[|x^{-1}S_n|^\gamma \1_{\{d_n \le  |S_n|\le x_n
  \}}] + \E[|x_n^{-1}S_n|^\gamma \1_{\{  |S_n| < d_n\}}]\\
&\le 	\frac{\alpha}{\gamma-\alpha}\,n \,\P(|Z|>x_n) + \Big(\frac{d_n}{x_n}\Big)^\gamma.
\end{split}
\end{equation*}
\end{proof}

%\newpage
\section{Proof of Theorem~\ref{thm:mains}}\label{sec:proof}\setcounter{equation}{0}

To increase the readability of our arguments we will proof the results for $h_{kl}=0$ if $\min(k,l)<0$. The extension to general two-sided filters is straightforward.

Let $s\in \N_0$. The $(i,j)$th entry of $\X_n(0)\X_n(s)'$ is
\begin{equation*}
\Big(\X_n(0)
\X_n(s)'\Big)_{ij}= \sum_{l_1,l_2=0}^\infty
\sum_{k_1, k_2=0}^{\infty} h_{k_1,l_1}  h_{k_2,l_2}\sum_{t=1}^n Z_{i-k_1,t-l_1} Z_{j-k_2,t+s-l_2}\,, \quad i,j=1, \ldots,p.
\end{equation*}
We approximate this matrix in three steps.

%-----------------------------------------------------------------------
\subsection{Step 1}

We show that it suffices to deal with
the finite moving averages
\begin{equation*}
X_{it}^{(m)}(s)=\sum_{l=0}^m\sum_{k=0}^m h_{kl}
Z_{i-k,t+s-l}\,,\quad m\ge 1\,.
\end{equation*}
and
the corresponding matrices
$\X_n^{(m)}(s)=(X_{i,t}^{(m)}(s))_{i=1,\ldots,p,t=1,\ldots,n}$.
The $(i,j)$th element of\\ $\X_n^{(m)}(0) \X_n^{(m)}(s)'$ can be decomposed as follows:
\begin{equation*}
\begin{split}
\sum_{t=1}^n X_{it}^{(m)}(0)X_{jt}^{(m)}(s)
&= \sum_{l_1,l_2=0}^m
\sum_{k_1,k_2=0}^m h_{k_1,l_1}h_{k_2,l_2}\sum_{t=1}^n
Z_{i-k_1,t-l_1} Z_{j-k_2,t+s-l_2}\\
&= \sum_{l=0}^m
\sum_{k=0}^m h_{k,l}h_{k+j-i,l+s}\sum_{t=1}^n
Z_{i-k,t-l}^2\\
& + \Big( \sum_{l_1,l_2=0;l_2\neq l_1+s}^m
\sum_{k_1,k_2=0}^m h_{k_1,l_1}h_{k_2,l_2}\sum_{t=1}^n
Z_{i-k_1,t-l_1} Z_{j-k_2,t+s-l_2}\\
& + \sum_{l=0}^m
\sum_{k_1,k_2=0;i-k_1\neq j-k_2}^m h_{k_1,l}h_{k_2,l+s}\sum_{t=1}^n
Z_{i-k_1,t-l} Z_{j-k_2,t-l} \Big)\\
&= (\wt{X}^{(m)}_{n}(s))_{ij}+I_{ij}\,.
\end{split}
\end{equation*}

\begin{lemma} \label{lem:0}
Assume the conditions of Theorem~\ref{thm:mains} and let $\alpha\in (0,4)$. 
\begin{itemize}
\item[($a$)] If $\beta \in [0,1] \cap (\alpha/2-1,1]$,
then
\begin{equation*}
\lim_{m\to\infty} \limsup_{\nto} \P(a_{np}^{-2}\|\X_n(0)
\X_n(s)'-\wt{X}^{(m)}_{n}(s) \|_2>\vep)=0\,,\quad \vep>0\,.
\end{equation*}
\item[($b$)] If $\alpha\in (2,4)$ and $\beta \in [0,1]$,
then for all $\vep>0$ 
\begin{equation*}
\begin{split}
\lim_{m\to\infty} \limsup_{\nto} \P(a_{np}^{-2}\|\X_n(0)
\X_n(s)'- \E[\X_n(0)\X_n(s)'] -\wt{X}^{(m)}_{n}(s)+\E[\wt{X}^{(m)}_{n}(s)] \|_2>\vep)=0\,.
\end{split}
\end{equation*}
\end{itemize}
\end{lemma}

\begin{proof}[Proof of Lemma~\ref{lem:0}]
Observe that
\begin{equation*}
\Big(\X_n(0)\X_n(s)'-\wt{X}^{(m)}_{n}(s)\Big)_{ij}= I_{ij}+
\sum_{l_1\vee l_2 \vee k_1\vee k_2>m} h_{k_1,l_1}  h_{k_2,l_2}\sum_{t=1}^n Z_{i-k_1,t-l_1} Z_{j-k_2,t+s-l_2}.
\end{equation*}
We start with part ($a$). Let 
\begin{equation*}
Y(p,n,k,l)=(Z_{i-k,t-l})_{i=1,\ldots,p,t=1,\ldots,n}, \quad k,l \in \Z.
\end{equation*}
 We have
\begin{equation*}
\begin{split}
\P( & \twonorm{\X_n(0)\X_n(s)'-\wt{X}^{(m)}_{n}(s)}>ca_{np}^{2}) \le \P(\twonorm{I}>ca_{np}^{2})\\
&+ \P\Big( \sum_{l_1\vee l_2 \vee k_1\vee k_2>m}  |h_{k_1,l_1}h_{k_2,l_2}| \twonorm{Y(p,n,k_1,l_1)Y(p,n,k_2,l_2-s)'}>c a_{np}^2 \Big) = P_1+P_2
\end{split}
\end{equation*}

If $\alpha \in (2,4)$, set $\gamma=1$. If $\alpha \in (0,2]$, choose $\gamma<\alpha/2$. Let $\beta \in [0,1] \cap (\alpha/2-1,1]$. By virtue of Theorem~\ref{thm:iidjohannes}
\begin{equation}\label{eq:kjd}
\lim_{\nto}\E[a_{np}^{-2\gamma}  \twonorm{Y(p,n,0,0)Y(p,n,k,l)'}^{\gamma}]\le \E[\Psi^\gamma]<\infty, \quad k,l \in \Z,
\end{equation}
for a \rv\ $\Psi$ which is \Frechet distributed with parameter $\alpha/2$.
An application of Markov's inequality yields
\begin{equation*}
\begin{split}
P_2 
&\le c a_{np}^{-2\gamma}\E\Big[ \Big( \sum_{l_1\vee l_2 \vee k_1\vee k_2>m}  |h_{k_1,l_1}h_{k_2,l_2}| \twonorm{Y(p,n,k_1,l_1)Y(p,n,k_2,l_2-s)'}  \Big)^\gamma \Big]\\
&\le c \sum_{l_1\vee l_2 \vee k_1\vee k_2>m}  |h_{k_1,l_1}h_{k_2,l_2}|^{\gamma} \E[a_{np}^{-2\gamma} \twonorm{Y(p,n,k_1,l_1)Y(p,n,k_2,l_2-s)'}^{ \gamma}]\to 0, \quad n,m\to \infty,
\end{split}
\end{equation*}
%We also used that fact that for any $X,Y$ with the same distribution it holds $\E[XY]\le \E[X^2]$. 
because of \eqref{eq:2a}. For $P_1$ the approach is similar. It is important to use Theorem~\ref{thm:iidjohannes}, namely that for $k,l\in \Z$, $k\neq 0$
\begin{equation*}
a_{np}^{-2}\twonorm{Y(p,n,0,0)Y(p,n,k,l)'} \cip 0 \quad \mbox{and} \quad
\lim_{\nto}\E[a_{np}^{-2\gamma}  \twonorm{Y(p,n,0,0)Y(p,n,k,l)'}^{\gamma}]=0
\end{equation*}

For part ($b$) we have to additionally substract the expectation. Let $\alpha\in (2,4)$ and $\beta \in [0,1]$. Choose $\gamma$ as above. Theorem~\ref{thm:iidjohannes} tells us that 
\begin{equation*}
\lim_{\nto}\E[a_{np}^{-2\gamma}  \twonorm{Y(p,n,0,0)Y(p,n,k,l)'-\E[Y(p,n,0,0)Y(p,n,k,l)']}^{\gamma}]\le \E[\Psi^\gamma]<\infty, \quad k,l \in \Z.
\end{equation*}
The rest of the proof is completely analogous to part ($a$).
\end{proof}

%-----------------------------------------------------------------------
\subsection{Step 2}

Let $M^{(m)}(s)$ be given as
\begin{equation*}
(M^{(m)}(s))_{ij} = \sum_{l=0}^m h_{il} h_{j,l+s}, \quad i,j=0, \ldots ,m
\end{equation*}\noindent
and $0$ otherwise. Denote the ordered singular values of ${M}^{(m)}(s)$ by
\begin{equation*}
v_1^{(m)}(s)\ge\cdots \ge v_{m+1}^{(m)}(s)\,,
\end{equation*}\noindent
and let $r_m(s)$ be the rank of $M^{(m)}(s)$.
\par
Next define the $p\times p$ matrices $M_a^{(m)}$, $a\in\Z$, via
\begin{eqnarray}\label{eq:matrixm}
(M_{a}^{(m)}(s))_{ij}= \left\{\begin{array}{ll}
(M^{(m)}(s))_{i-a,j-a}\,,&i,j=a,\ldots,m+a\,,\\
0\,,&\mbox{otherwise}
\end{array}\right.\,, \quad i,j=1,\ldots,p\,.
\end{eqnarray}\noindent
For $a\ge 1$, $M_a^{(m)}(s)$ has rank $r_m(s)$ and has the same
singular values $v_1^{(m)}(s),\ldots,v_{m+1}^{(m)}(s)$ as $M^{(m)}(s)$.

\begin{lemma}\label{lem:1}
Assume the conditions of Theorem~\ref{thm:mains} and let $\alpha\in (0,2)\cup (2,4)$. 
\begin{itemize}
\item[($a$)] If $\beta \in [0,1] \cap (\alpha/2-1,1]$,
then
\begin{equation*}
a_{np}^{-2}\Big\| \wt{X}^{(m)}_{n}(s) -\sum_{a=1}^p
  D_a M_a^{(m)}(s)\Big\|_2 \cip 0, \quad \nto.
\end{equation*}
\item[($b$)] If $\alpha\in (2,4)$ and $\beta \in [0,1]$,
then
\begin{equation*}
a_{np}^{-2}\Big\| \wt{X}^{(m)}_{n}(s)- \E[\wt{X}^{(m)}_{n}(s)] -\sum_{a=1}^p
  (D_a-\E[D]) M_a^{(m)}(s)\Big\|_2 \cip 0, \quad \nto.
\end{equation*}
\end{itemize}
\end{lemma}

\begin{proof}
We start with $\alpha\in(0,2)$ and $\beta \in[0,1]$. For this aim we study
\begin{equation}\label{eq:help1}
\begin{split}
a_{np}^{-2} \big\| \wt{X}^{(m)}_{n}(s) -\sum_{a=1}^p
  D_a M_a^{(m)}(s) \big\|_2 &\le a_{np}^{-2}\big\|\wt{X}^{(m)}_{n}(s) -\sum_{a=-m}^p
  D_a M_a^{(m)}(s)\big\|_2
 +a_{np}^{-2} \big\|\sum_{a=-m}^0 D_a M_a^{(m)}(s)\big\|_2.
\end{split}
\end{equation}
We have
\begin{eqnarray}\label{eq:help2}
a_{np}^{-2} \big\|\sum_{a=-m}^0 D_a M_a^{(m)}(s)\big\|_2\le a_{np}^{-2}\max_{i=-m,\ldots,0}D_i\;
\sum_{a=-m}^0\|M_a^{(m)}(s)\|_2\cip 0\,.
\end{eqnarray}\noindent
This follows because $\|M_a^{(m)}(s)\|_2< \infty$ for each $a$ and since
$a_{n}^{-2} D_i\cid \xi_{\alpha/2}$ for an $\alpha/2$-stable \rv\
$\xi_{\alpha/2}$ as $\nto$, hence  $a_{np}^{-2} D_i\cip 0$ by virtue
of $p_n\to\infty$. Observe
\begin{equation}\label{eq:ghjgh}
 \Big( \wt{X}^{(m)}_{n}(s) -\sum_{a=-m}^p
  D_a M_a^{(m)}(s)\Big)_{ij}
    = \sum_{t=1}^n \sum_{k=0}^m \sum_{l=0}^m h_{kl}h_{j-i+k,l}
    Z_{i-k, t-l}^2 - \sum_{k=i-p}^{i+m} D_{i-k} (M_{i-k}^{(m)}(s))_{ij}
\end{equation}
  and note that $(M_{i-k}^{(m)}(s))_{ij}$ is non-zero only if $i-k \leq i \leq
  i-k+m$, that is, if $0 \leq k \leq m$. This and \eqref{eq:matrixm} can be used to write the right-hand side of \eqref{eq:ghjgh} as
\begin{equation}\label{eq:lgfh}
\sum_{k=0}^m \sum_{l=0}^m h_{kl} h_{j-i+k, l+s}\left(
      \sum_{t=1}^l Z_{i-k, t-l}^2 - \sum_{t=n-l+1}^n Z_{i-k, t}^2
    \right) = I_{ij}^{(1)}-I_{ij}^{(2)}.
\end{equation}
For $a_{np}^{-2}\big\|\wt{X}^{(m)}_{n}(s) -\sum_{a=-m}^p
  D_a M_a^{(m)}(s)\big\|_2 \cip 0$ it suffices to show that
\begin{equation*}
a_{np}^{-2} \|I^{(i)}\|_2\cip 0\,,\quad i=1,2.
\end{equation*}\noindent
We will show the limit relation for $i=1$; the case $i=2$ is
analogous. In the sequel we interpret $h_{kl}$ as $0$ for $\max(k,l)>m$ or $\min(k,l)<0$. For the non-symmetric $I^{(1)}$, we have $\twonorm{I^{(1)}}\le \max(\inftynorm{I^{(1)}}, \|{I^{(1)}}\|_1)$, where
\begin{equation*}
\inftynorm{I^{(1)}}= \max_{i=1,\ldots,p} \sum_{j=1}^p |I^{(1)}_{ij}| \quad \mbox{ and } \|{I^{(1)}}\|_1 = \max_{j=1,\ldots,p} \sum_{i=1}^p |I^{(1)}_{ij}|
\end{equation*}
For $a_{np}^{-2} \twonorm{I^{(1)}}\cip 0$ it thus suffices to show that $a_{np}^{-2} \inftynorm{I^{(1)}},a_{np}^{-2}\|{I^{(1)}}\|_1$  both converge to zero in probability. We have
\begin{equation*}
\begin{split}
\inftynorm{I^{(1)}}&= \max_{i=1,\ldots,p} \sum_{j=1}^p \Big|\sum_{k,l=0}^m h_{kl} h_{j-i+k, l+s} \sum_{t=1}^l Z_{i-k, t-l}^2\Big|\\
&\le c \max_{i=1,\ldots,p}\sum_{k,l=0}^m  \sum_{t=1}^m Z_{i-k, t-l}^2
\end{split}
\end{equation*}
Thus, for a sequence $(Z_t)$ of iid copies of $Z$, we get
\begin{equation*}
\begin{split}
\P(\inftynorm{I^{(1)}}>c a_{np}^2) &\le p\P\Big(\sum_{t=1}^{(m+1)^3} Z_t^2>c a_{np}^2 \Big)\\
&\sim p (m+1)^3 P(Z^2 >c a_{np}^2) \to 0, \quad \nto.
\end{split}
\end{equation*}
Similarly one obtains
\begin{equation*}
\begin{split}
\P(\|{I^{(1)}}\|_1>c a_{np}^2) &\le \P\Big(\sum_{i=1}^p\sum_{t=1}^{(m+1)^3} Z_{ti}^2>c a_{np}^2 \Big)\\
&\sim p (m+1)^3 P(Z^2 >c a_{np}^2) \to 0, \quad \nto.
\end{split}
\end{equation*}
This finishes the proof for $\alpha\in(0,2)$ and $\beta \in[0,1]$.
\par

If $\alpha\in(2,4)$ and $\beta \in[0,1]$, one has to rewrite \eqref{eq:lgfh}:
\begin{equation}\label{eq:vovof}
\sum_{k=0}^m \sum_{l=0}^m h_{kl} h_{j-i+k, l+s}\left(
      \sum_{t=1}^l (Z_{i-k, t-l}^2-\E[Z^2]) - \sum_{t=n-l+1}^n (Z_{i-k, t}^2-\E[Z^2])
    \right) = \wt{I}_{ij}^{(1)}-\wt{I}_{ij}^{(2)}.
\end{equation}
and then proceed as before, the only difference being that $Z^2$ is replaced by $|Z^2-\E[Z^2]|$.
\par

If $\alpha\in(2,4)$ and $\beta \in (\alpha/2-1,1]$, we have to additionally  justify why \eqref{eq:help2} still holds. In general, for $\alpha\in(2,4)$ we only know that $a_{np}^{-2} |D_1-\E[D]| \cip 0$. However,
\begin{equation*}
\frac{\E[D]}{a_{np}^2}  =\E[Z^2] \frac{n}{a_{np}^2} \le \E[Z^2]  n^{1-\frac{2(1+\beta)}{\alpha}+\delta}
\end{equation*}
converges to $0$ if $\beta>\alpha/2-1$, where $\delta >0$ can be chosen arbitrarily small. Therefore the centering is negligible and $a_{np}^{-2} D_1 \cip 0$.  The proof is complete.
\end{proof}

%-----------------------------------------------------------------------
\subsection{Step 3}

\noindent\textbf{The case $\alpha\in (0,2)$.} From \eqref{eq:help6} recall  the definition of the order statistics
\begin{equation*}
D_{(p)}=D_{L_p}<\cdots  <  D_{(1)}=D_{L_1}\quad \as
\end{equation*}\noindent
of the iid sequence $D_1,\ldots,D_p$. % defined in \eqref{eq:ll}.
Here we assume without loss of generality that there are no ties in the
sample.
Otherwise, if two or more of the $D_i$'s are equal, randomize the
corresponding $L_i$'s over the respective indices.

We choose an integer sequence $k=k_p\to\infty$ such that $k_p^2=o(p)$ as $\nto$
and define the event
\begin{equation}\label{eq:an}
A_n=\{|L_i-L_j|>m+1\,, i,j=1,\ldots,k\,, i\ne j\}\,.
\end{equation}\noindent
Since the $D_i$'s are iid, $L_1,\ldots,L_k$ have a uniform distribution
on the set of
distinct $k$-tuples from $(1,\ldots,p)$ and
\begin{equation}\label{eq:31}
\P(A_n^c)\le k(k-1) \dfrac{pm (p-2)\ldots (p-k+1)}{p(p-1)\ldots (p-k+1)}\le
\dfrac{k^2 m}{p-1}\to 0\,,\quad \nto\,.
\end{equation}\noindent
On the event $A_n$, the matrix $\sum_{i=1}^k D_{L_i}  M_{L_i}^{(m)}(s)$ is
block diagonal and has singular values in the set 
$\{D_{(i)}v_j^{(m)}(s):i=1,\ldots,k,j=1,\ldots,m+1\} \cup\{0\}$.
In this step of the proof we approximate $\sum_{i=1}^p D_i M_i^{(m)}(s)$
by the matrix $\sum_{i=1}^k D_{L_i}  M_{L_i}(s)^{(m)}(s)$ which is block
diagonal with high probability.\smallskip\\

\noindent \textbf{The case $\alpha\in (2,4)$.} Recall that the order
statistics of $\widetilde{D}_i=|D_i-\E[D]|$, $i=1,\ldots,p,$ are denoted by
\begin{equation}\label{eq:ww}
\widetilde{D}_{(p)}=\widetilde{D}_{\ell_p}\le \cdots \le \widetilde{D}_{(1)}=\widetilde{D}_{\ell_1}\quad \as ,
\end{equation}
where we again assume without loss of generality that there are no
ties in the sample.
We choose an integer sequence $k=k_p\to\infty$ such that $k^2_p=o(p)$ as
$\nto$ and define the event
\begin{equation}\label{eq:antilde}
\widetilde{A}_n=\{|\ell_i-\ell_j|>m+1\,,i,j=1,\ldots,k\,,i\ne j\}\,.
\end{equation}
As for $A_n$, we have
\begin{equation}\label{eq:antildec}
P(\widetilde{A}_n^c)\le \dfrac{k^2 m}{p-1}\to 0\,,\quad \nto\,.
\end{equation}
On the event $\widetilde{A}_n$, the matrix
$\sum_{i=1}^k (D_{\ell_i}-\E[D]) M_{\ell_i}^{(m)}(s)$ is block diagonal and its
singular values are the $p$ largest values in the set $\{\widetilde{D}_{\ell_i} v_j^{(m)}(s): i=1,\ldots,k, j=1,\ldots,m+1\}\cup\{0\}$.

\begin{lemma}\label{lem:2}
Assume the conditions of Theorem~\ref{thm:mains} and let $\alpha\in (0,2)\cup (2,4)$. Consider an integer sequence  $(k_p)$ such that
$k_p\to\infty$ and $k_p^2=o(p)$ as $\nto$.
\begin{itemize}
\item[($a$)] If $\beta \in [0,1] \cap (\alpha/2-1,1]$,
then
\begin{equation*}
a_{np}^{-2}\Big\|\sum_{i=1}^p
  D_i M_i^{(m)}(s) -\sum_{i=1}^k
  D_{L_i} M_{L_i}^{(m)}(s)\Big\|_2\cip 0\,,
\quad \nto \,.
\end{equation*}
\item[($b$)] If $\alpha\in (2,4)$ and $\beta \in [0,1]$,
then
\begin{equation*}
a_{np}^{-2}\Big\| \sum_{i=1}^p (D_i-\E[D]) M_i^{(m)}(s)- \sum_{i=1}^k (D_{\ell_i}-\E[D]) M_{\ell_i}^{(m)}(s)\Big\|_2 \cip 0, \quad \nto.
\end{equation*}
\end{itemize}
\end{lemma}

\begin{proof}
\textbf{The case $\alpha\in (0,2)$.} Let us start with $\alpha\in (0,2)$ and $\beta \in [0,1]$. We have
\begin{equation*}
a_{np}^{-2}\Big(\sum_{i=1}^p
  D_i M_i^{(m)}(s) -\sum_{i=1}^k
  D_{L_i} M_{L_i}^{(m)}(s)\Big)
=a_{np}^{-2}\sum_{i=k+1}^p
  D_{L_i} M_{L_i}^{(m)}(s)\,,
\end{equation*}\noindent
and therefore it suffices to show that the \rhs\ converges to zero in
probability. Then for $\delta>0$,
\begin{equation}\label{eq:98}
\P\Big(a_{np}^{-2}\Big\|\sum_{i=k+1}^p D_{L_i}
M_{L_i}^{(m)}(s)\Big\|_2>\delta\Big)
\le \P\Big(c\, a_{np}^{-2}\sum_{i=k+1}^p D_{(i)}>\delta\Big)\,.
\end{equation}\noindent
We will show that the \rhs\ converges to zero as $\nto$.
From Lemma~\ref{lem:ppr} we conclude that
\begin{equation*}
\sum_{i=1}^p \vep_{a_{np}^{-2} D_i}\cid \sum_{i=1}^
\infty\vep_{\Gamma_i^{-2/\alpha}}\,,\quad \nto\,,
\end{equation*}\noindent
where $(\Gamma_i)$ is an increasing enumeration of the points of a homogeneous Poisson
process on $(0,\infty)$. A continuous mapping
argument
(Resnick
\cite{resnick:2007}, Theorem 7.1)
shows that for every $\gamma>0$,
\begin{equation*}
a_{np}^{-2}
\Big(\sum_{i=1}^p D_i  \1_{\{a_{np}^{-2}D_i>\gamma\}}\,,
\sum_{i=1}^p D_i\Big)
\cid
\Big(\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}
\1_{\{\Gamma_i^{-2/\alpha}>\gamma\}}\,,
\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}\Big)
\end{equation*}\noindent
provided
\begin{equation}\label{eq:35}
\lim_{\xi \downarrow 0}\limsup_{\nto } p\, a_{np}^{-2}\,\E
[D \1_{\{a_{np}^{-2}D\le \xi\}}] =0,\end{equation}\noindent
and hence
\begin{equation*}
a_{np}^{-2}
\sum_{i=1}^p D_i  \1_{\{a_{np}^{-2}D_i\le \gamma\}}
\cid \sum_{i=1}^\infty \Gamma_i^{-2/\alpha}
\1_{\{\Gamma_i^{-2/\alpha}\le \gamma\}}\,.
\end{equation*}\noindent
But we also have
\begin{equation*}
a_{np}^{-2}\sum_{i=k+1}^p D_{(i)}=a_{np}^{-2}\sum_{i=1}^p D_i
\1_{\{a_{np}^{-2}D_i< a_{np}^{-2} D_{(k)}\}}\,,
\end{equation*}\noindent
and $ a_{np}^{-2} D_{(k)}\cip0$ as $\nto$.
For $\gamma>0$, we write
$B_n=\{a_{np}^{-2} D_{(k)}\le \gamma\}$. Then $\P(B_n^c)\to 0$ as
$\nto$
and
\begin{equation*}
\begin{split}
\P\Big( c a_{np}^{-2}\sum_{i=k+1}^p D_{(i)}>\delta\Big)
&\le
\P\Big(B_n\cap \{c\, a_{np}^{-2}\sum_{i=1}^p D_i
\1_{\{a_{np}^{-2}D_i\le  \gamma\}}>\delta\}\Big)+o(1)\\
&\le \P\Big(c\,a_{np}^{-2}\sum_{i=1}^p D_i
\1_{\{a_{np}^{-2}D_i\le  \gamma\}}>\delta\Big)+o(1)\\
&\to \P\Big(c \,\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}
\1_{\{\Gamma_i^{-2/\alpha}\le \gamma\}}>\delta\Big)\,,\quad\nto\,,\\
&\to 0\,,\quad \gamma \downarrow 0\,.
\end{split}
\end{equation*}\noindent
Therefore the \rhs\ in \eqref{eq:98} converges to zero if
\eqref{eq:35} holds.
\par
Thus it remains to show \eqref{eq:35}. By Karamata's theorem, as $\nto$,
\begin{equation*}
p\, a_{np}^{-2}\, \E[D\1_{\{ D\le \xi a_{np}^2
  \}}]\le np\,a_{np}^{-2}\, \E[Z^2 \1_{\{Z^2\le a_{np}^2\xi \}}]
\sim \xi\, np\, \P(Z^2>a_{np}^2 \xi)\sim \xi^{1-\alpha/2}\,,
\end{equation*}\noindent
and the \rhs\ converges to zero as $\xi\downarrow 0$.
Then \eqref{eq:35} follows and the proof for $\alpha \in (0,2)$ is finished.\medskip\\

\textbf{The case $\alpha\in (2,4)$.} Next, we will consider $\alpha\in (2,4)$ and $\beta \in [0,1]$.
As a  first step in the proof we show the following relation for
every $\delta>0$,
\begin{equation}\label{eq:200a}
\lim_{\gamma\downarrow 0}\limsup_{\nto}\P\Big(
a_{np}^{-2}\Big\|\sum_{i=1}^p (D_i-\E[D]) \1_{\{a_{np}^{-2}|D_i-\E[D]|\le \gamma \}} M_{i}^{(m)}(s)\Big\|_2>\delta\Big)=0\,.%\nonumber\\
\end{equation}
We observe that we can divide the index set
  $\{1,\ldots,p \}$ of the involved sums into disjoint subsets
$I_1=\{1,m+2,2m+3,\ldots\}\cap\{1,\ldots,p\}$,
$I_2=\{2,m+3,2m+4,\ldots\}\cap\{1,\ldots,p\}$, etc. For $\gamma>0$,
due to the
construction of the matrices $M_i^{(m)}(s)$ (see \eqref{eq:matrixm}), the sums
\begin{equation*}
T_j=a_{np}^{-2}\sum_{i\in I_j} (D_i-\E[D])\1_{\{a_{np}^{-2}|D_i-\E[D]|\le
  \gamma\}}
M_i^{(m)}(s)\,, \quad j=1,\ldots,m+1\,,
\end{equation*}
constitute block diagonal matrices with norm
\begin{equation*}
\|T_j\|_2=a_{np}^{-2}\max_{i \in I_j}  |D_i-\E[D]|\1_{\{a_{np}^{-2}|D_i-\E[D]|\le
  \gamma\}} \;v_1^{(m)}(s)\,.
\end{equation*}
An application of Markov's inequality yields
\begin{eqnarray}\label{eq:105}\lefteqn{
\P\Big(
a_{np}^{-2}\Big\|\sum_{i=1}^p
(D_{i}-\E[D]) \1_{\{a_{np}^{-2}|D_i-\E[D]|\le \gamma \}}
M_{i}^{(m)}(s)\Big\|_2>\delta\Big)}\nonumber\\
&\le &
\sum_{j=1}^{m+1} \P\Big(\|T_j\|_2>\delta/(m+1)\Big)\nonumber\\
&\le & c\, a_{np}^{-4}\sum_{j=1}^{m+1} \E [\max_{i \in I_j}  |D_i-\E[D]|^2\1_{\{a_{np}^{-2}|D_i-\E[D]|\le
  \gamma\}}]\nonumber\\
&\le & c\, a_{np}^{-4}\sum_{i=1}^p \E [|D_i-\E[D]|^2\1_{\{a_{np}^{-2}|D_i-\E[D]|\le
  \gamma\}}]\nonumber\\
&=& c\,p  a_{np}^{-4}\E [|D-\E[D]|^2 \1_{\{a_{np}^{-2}|D-\E[D]|\le \gamma\}}]\,.
\end{eqnarray}
An application of \eqref{eq:iec} yields for $d_n=a_n^2 s_n$,
  any sequence $(s_n)$ such that $s_n\to\infty$,
\begin{equation*}
p\,a_{np}^{-4}\E [|D-\E[D]|^2 \1_{\{d_n \le
  |D-\E[D]|\le a_{np}^2 \gamma\}}] \le  c np\, \gamma^2 \P(Z^2>a_{np}^2
\gamma) \sim c\gamma^{(4-\alpha)/2}\,,\quad \nto\,,
\end{equation*}
and the \rhs\ converges to zero as $\gamma\to 0$. On the other hand,
for $\epsilon>0$ arbitrarily small and $s_n\to\infty$ sufficiently slowly,
\begin{equation*}
p\,a_{np}^{-4}\E [|D-\E[D]|^2 \1_{\{
  |D-\E[D]|\le d_n \}}] \le   p s_n^2 (a_n/a_{np})^4\le c\,s_n^2 p^{1-
  4/\alpha+\epsilon}\to 0\,,\quad \nto\,.
\end{equation*}
Thus we proved that \eqref{eq:105} converges to zero by first letting
$\nto$ and then $\gamma\to 0$.
This proves \eqref{eq:200a}.
In view of \eqref{eq:200a}, the lemma is proven if we can show that
for every $\delta>0$ and  $\gamma\in (0,\gamma_0(\delta))$ for a
sufficiently small $\gamma_0(\delta)$,
\begin{equation}\label{eq:201}
\lim_{\nto}\P\Big(
a_{np}^{-2}\Big\|\sum_{i=1}^p (D_i-\E[D]) M_i^{(m)}(s) \1_{\{|D_i-\E[D]|>\gamma a_{np}^2\}} -\sum_{i=1}^k
(D_{\ell_i}-\E[D])  M_{\ell_i}^{(m)}(s)\Big\|_2>\delta\Big)=0\,.%\nonumber\\
\end{equation}
If
\begin{equation*}
\begin{split}
0&\ne
\sum_{i=1}^p (D_i-\E[D])
 M_i^{(m)}(s) \1_{\{|D_i-\E[D]|>\gamma a_{np}^2\}}
-\sum_{i=1}^k
(D_{\ell_i}-\E[D]) \1_{\{|D_{\ell_i}-\E[D]|>\gamma a_{np}^2\}} M_{\ell_i}^{(m)}(s)\\
&= \sum_{i=k+1}^p (D_{\ell_i}-\E[D]) M_{\ell_i}^{(m)}(s) \1_{\{|D_{\ell_i}-\E[D]|>\gamma
  a_{np}^2\}}  \,,
\end{split}
\end{equation*}
then $\widetilde{D}_{(k+1)}=|D_{\ell_{k+1}}-\E[D]|>\gamma a_{np}^2$ and therefore
\begin{equation*}
N_n(\gamma)=\#\{ i\le p : |D_i-\E[D]|>\gamma a_{np}^2\}>k\,.
\end{equation*}
However, for fixed $\gamma>0$, in view of Theorem~\ref{thm:nagaev}, as
$\nto$,
\begin{equation*}
\P(N_n(\gamma)>k)\le k^{-1} \E[N_n(\gamma)]=k^{-1} p\, \P(|D-\E[D]|>\gamma
a_{np}^2)
\sim k^{-1}\gamma^{-\alpha/2} np \, \P(Z^2>a_{np}^2)\to 0\,.
\end{equation*}
Thus we have proven that
\begin{equation*}
\begin{split}
\lim_{\nto} &\P\Big(
a_{np}^{-2}\Big\|\sum_{i=1}^p (D_i-\E[D]) M_i^{(m)}(s) \1_{\{|D_i-\E[D]|>\gamma a_{np}^2\}}\\
& -\sum_{i=1}^k (D_{\ell_i}-\E[D])  M_{\ell_i}^{(m)}(s) \1_{\{|D_{\ell_i}-\E[D]|>\gamma a_{np}^2\}} \Big\|_2>\delta\Big)=0\,.
\end{split}
\end{equation*}
Relation \eqref{eq:201}
is proven if we can show that for $\delta>0$,
 \begin{equation}\label{eq:k2}
\lim_{\nto}\P\Big(a_{np}^{-2}\Big\|\sum_{i=1}^k
(D_{\ell_i}-\E[D]) \1_{\{|D_{\ell_i}-\E[D]|\le \gamma a_{np}^2\}} M_{\ell_i}^{(m)}(s)\Big\|_2>\delta
\Big)\to 0\,,
\end{equation}
as $\gamma \downarrow 0$.
On the event $\widetilde{A}_n$ defined in \eqref{eq:antilde},  $a_{np}^{-2}\sum_{i=1}^k
(D_{\ell_i}-\E[D]) \1_{\{|D_{\ell_i}-\E[D]|\le \gamma a_{np}^2\}} M_{\ell_i}^{(m)}(s)$ is
block diagonal and therefore its spectral norm is
\begin{equation*}
a_{np}^{-2} \, v_1^{(m)}(s)\,
  \max_{i=1,\ldots,k}|D_{\ell_i}-\E[D]| \1_{\{|D_{\ell_i}-\E[D]|\le
  \gamma a_{np}^2\}}\le  c \,\gamma\,.
\end{equation*}
We also observe that $\P(\widetilde{A}_n^c)\to 0$; see \eqref{eq:antildec}.
Then \eqref{eq:k2} is immediate and the lemma is proven. \medskip\\

\textbf{The case $\alpha\in (2,4)$ and $\beta\in (\alpha/2,1]$.} In this situation we need to prove that
\begin{equation*}
a_{np}^{-2}\Big\|\sum_{i=1}^p
  D_i M_i^{(m)}(s) -\sum_{i=1}^k
  D_{L_i} M_{L_i}^{(m)}(s)\Big\|_2\cip 0\,,
\quad \nto \,.
\end{equation*}
We have 
\begin{equation*}
\begin{split}
\Big\|\sum_{i=1}^p
  D_i M_i^{(m)}(s) -\sum_{i=1}^k
  D_{L_i} M_{L_i}^{(m)}(s)\Big\|_2 &\le \Big\|\sum_{i=1}^p
  (D_i-\E[D]) M_i^{(m)}(s) -\sum_{i=1}^k
  (D_{L_i}-\E[D]) M_{L_i}^{(m)}(s)\Big\|_2\\
	&\quad + \E[D]\Big\| \sum_{i=k+1}^p M_{L_i}^{(m)}(s)  \Big\|_2.
\end{split}
\end{equation*}
Since the $M_{i}^{(m)}(s), i=1,\ldots,p$, consist of block matrices of size $m$ shifted by $i$, at most $2m$ of them can overlap. By Cauchy's interlacing theorem, see \cite[Lemma~22]{tao:vu:2012},
\begin{equation*}
\frac{\E[D]}{a_{np}^2} \Big\| \sum_{i=k+1}^p M_{L_i}^{(m)}(s)  \Big\|_2 \le c m \frac{\E[D]}{a_{np}^2} \to 0, \quad \nto.
\end{equation*}
Now we can follow the lines of the proof in the case $\alpha \in (2,4)$ to show the desired result.
\end{proof}
\begin{remark}
The case $\alpha =2$ in Lemma~\ref{lem:2}. Here one has to distinguish whether $E[Z^2]$ is infinite or finite. We omit the technical details.
%If $E[Z^2]=\infty$, then the function
%\begin{equation*}
%f(x)=\E[Z^2 \1_{\{ |Z| \le x \}}]
%\end{equation*}
%is slowly varying. After follow the lines of the proof in the case $\alpha \in (2,4)$.
\end{remark}

%-----------------------------------------------------------------------
\subsection{Final argument}

\begin{proof}[Proof of Theorem~\ref{thm:mains}]
\noindent {\bf The case $\alpha\in (0,2)$.} On $A_n$
defined in \eqref{eq:an},
the matrix $\sum_{i=1}^k
  D_{L_i} M_{L_i}^{(m)}(s)$ is block diagonal
and its
singular values are the $p$ largest values in the set
\begin{equation}\label{eq:drghrf}
\{D_{L_i} v_j^{(m)}(s)=D_{(i)}v_j^{(m)}(s):  i=1,\ldots,k, j=1,\ldots,m+1\}\cup\{0\}
\end{equation}
Because of $a_{np}^{-2} D_{(k)}\cip 0$ we can write $i=1, \ldots, p$ in \eqref{eq:drghrf}.
The corresponding largest $p$ ordered values of them are denoted by
$\gamma_{(1)}^{(m)}(s)\ge \cdots\ge \gamma_{(p)}^{(m)}(s)$.
Combining
Lemmas~\ref{lem:0}--\ref{lem:2} with Weyl's inequality for
singular values (see Bhatia \cite{bhatia:1997})
and recalling that $\P(A_n^c)\to 0$ as $\nto$, we have
\begin{equation}
\lim_{m\to\infty}\limsup_{\nto}\P\Big(a_{np}^{-2}
\max_{i\le p}\Big|
\lambda_{(i)}(s)-\gamma_{(i)}^{(m)}(s)\Big|>\vep\Big)=0\,,\quad \vep>0\,.
\end{equation}
Finally, we observe that
\begin{equation*}
a_{np}^{-2}\max_{i\le p} \big|\gamma_{(i)}^{(m)}(s)-\gamma_{(i)}(s)\big|\le
a_{np}^{-2}\max_{i\le p} D_i\; \max_{i\le p}\big|v_i^{(m)}(s)-v_i(s)\big|=o_\P(1) \,,
\end{equation*}
since $a_{np}^{-2}\max_{i\le p} D_i\cid \Gamma_1^{-\alpha/2}$
and $v_i^{(m)}\to v_i$ uniformly in $i$ because both sequences are monotone.

\noindent {\bf The case $\alpha\in (0,4)$ and $\beta \in (0,1] \cup (\alpha/2-1,1]$.}
The additional step is to replace $D_{(i)}$ by $Z_{(i),np}^2$. We see that
\begin{equation*}
\max_{j\in \N} v_j\max_{i\le p} a_{np}^{-2}\big|D_{(i)}-Z_{(i),np}^2 \big|= c a_{np}^{-2}\big|D_{(i)}-Z_{(i),np}^2 \big| =o_\P(1),
\end{equation*}
as shown in \cite{heiny:mikosch:2015:iid}.

\noindent {\bf The case $\alpha\in (2,4)$.} On $\widetilde{A}_n$ defined in  \eqref{eq:antilde},
the matrix $\sum_{i=1}^k(D_{\ell_i}-\E[D])M_{\ell_i}^{(m)}(s)$ is block
diagonal and 
its
singular values are the $p$ largest values in the set $\{|D_{\ell_i}-\E[D]| v_j^{(m)}(s): i=1,\ldots,k, j=1,\ldots,m+1\}\cup\{0\}$.
The corresponding ordered values are denoted by
$\widetilde{\gamma}_{(1)}^{(m)}(s)\ge
\cdots \ge \widetilde{\gamma}_{(p)}^{(m)}(s)$.
Combining Lemmas~\ref{lem:0}--\ref{lem:2} with Weyl's inequality
and recalling that $\P(\widetilde{A}_n^c)\to 0$, we have
\begin{equation*}
\lim_{m\to\infty}\limsup_{\nto}\P\Big(
a_{np}^{-2}
\max_{i\le p}\Big| \widetilde{\lambda}_{(i)}(s)-\widetilde{\gamma}_{(i)}^{(m)}(s)\Big|
>\vep\Big)=0\,,\quad\vep>0\,.
\end{equation*}
As before, one observes that
\begin{equation*}
a_{np}^{-2}\max_{i\le p} \big|\widetilde{\gamma}_{(i)}^{(m)}(s)-\widetilde{\gamma}_{(i)}(s)\big|\le
a_{np}^{-2}\max_{i\le p} |D_i-\E[D]|\; \max_{i\le p}\big|v_i^{(m)}-v_i\big|=o_\P(1) \,,
\end{equation*}

\noindent {\bf When $\alpha=2$} the second moment of $Z$ can be infinite or finite. If $\E[Z^2]= \infty$ one proceeds as in the case $\alpha\in (0,2)$, otherwise as in the case $\alpha\in (2,4)$.
This finishes the proof of Theorem~\ref{thm:mains}.
\end{proof}

%\end{document}
% Use this end document to get the complete proof after the bibliography.

### 2017-07-21 papers/FX/PaperBody.tex
Add the Extremes paper.

\begin{abstract}
%We consider an investor with preferences that accord with Generalized
%Disappointment Aversion. Such an investor cares about downside risk and
%we assume he recognizes the heavy tail feature of asset return
%distributions. We argue that when a market is dominated by rational
%investors of this kind, the return distributions of equities that are
%actively traded in this market must have very similar tail-indices. We
%also show, in contrast, the scale parameters of the return
%distributions may differ hugely from one another.
%On the other hand, whether or not all equities in a multivariate model
%have the same tail-index is a dividing issue for multivariate GARCH
%models proposed in the literature. Therefore, it is important to analyze
%data of real equity returns and see how close to each other the
%tail-indices actually are.
%In this work empirical results are also presented and they appear
%to support the conclusion that the tail-indices are very similar,
%with respect to the confidence bands of estimation.
\end{abstract}

\section{Introduction}\setcounter{equation}{0}
It is one of the stylized facts of financial econometrics that 
returns of speculative prices are {\em heavy-tailed.} There is no agreement
in the literature about how heavy these tails really are. For example, Barndorff-Nielsen and Shephard
\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails
which are comparable with those of a gamma \ds . On the other
hand, tails of returns have been studied in great detail 
in the extreme value community. Among extreme value specialists
there is general agreement that returns $X_t$ have tails of power-law-type, i.e.,
\beam\label{eq:1}
\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad \P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
\eeam  
where $c_{\pm}$, $\alpha_{\rm up}$ and $\alpha_{\rm low}$ are positive constants.\footnote{Here and in what
  follows, $f(x)\sim g(x)$ for positive \fct s $f$ and $g$ means 
that $f(x)/g(x)\to 1$ as $\xto$.} See for example, Embrechts et
al. \cite{embrechts:klueppelberg:mikosch:1997}, Jansen and de Vries
\cite{jansen1991frequency}, Mikosch \cite{mikosch:2003}, Resnick
\cite{resnick:2007}.  
In the extreme value literature it is common to replace the constants $c_\pm$ by suitable
{\em slowly varying} \fct s; cf. Embrechts et
al. \cite{embrechts:klueppelberg:mikosch:1997}, Chapter 3. In this
paper, for the  
sake of argument, we stick to the condition \eqref{eq:1}.
\par
There are some good theoretical reasons for the appearance of power-law tails in situations where certain moments 
of data are believed to be infinite. Tails of type \eqref{eq:1}
describe the maximum domain of attraction of the  
Fr\'echet \ds\ $\Phi_{\alpha_{\rm up}}(x)=\exp(-x^{-\alpha_{\rm up}})$ for $x>0$, i.e., scaled maxima of an iid \seq\ $(X_t)$ with
upper tail described in \eqref{eq:1} converge in \ds\ to $\Phi_{\alpha_{\rm up}}$.
Equivalently, power-law tails are prescribed by the 
generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e., for a suitable positive scaling \fct\ $a(u)$,
\beao
\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha_{\rm up})^{-\alpha_{\rm up}}\,,\qquad u\to\infty\,.
\eeao
The aforementioned results 
are considered very natural for iid and weakly dependent strictly stationary \seq s of random variables $(X_t)$; 
in the world of extremes they are the analogs
of the \clt\ from the world of sums.
\par
In the literature on extremes for return data one finds the 
statement that {\em estimated} values $\hat \alpha_{\rm up}$ and $\hat \alpha_{\rm low}$ 
of the tail-indices $\alpha_{\rm up}$ and $\alpha_{\rm low}$,
respectively, typically have the tendency that $\hat \alpha_{\rm
  up}>\hat \alpha_{\rm low}$. 
This observation is often explained by the fact that investors are
more prone to negative than to positive news in the market. 
Moreover, in the literature the {\em estimated} tail-indices $\hat \alpha$ (both in the left and right tails) 
are typically found in the range $(2,4)$. For an illustration, see
Figure~\ref{fig:1} where estimates $\hat \alpha_{\rm low}$ 
in three sectors of the Standard \& Poors 500 index  are shown. The
estimates are based on 1304 observations of daily return 
data from 4 January 2010 to 31 December 2014.
\par
When looking at Figure~\ref{fig:1} one might ask the following questions:
\begin{itemize}
\item
In view of the wide \asy\ confidence bands for the estimators of tail-indices, 
are the tail-indices from different series really distinct?
\item
Are there some {\em theoretical} reasons supporting the fact that the tail-indices from different series are {\em not} 
distinct?
\end{itemize}
In this paper, we try to find some answers to these questions. 
\par
The estimator of the tail-index $\alpha>0$  in the model
\beao
\P(X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeao
favored in the literature  is the {\em Hill estimator}; the graphs in 
Figure~\ref{fig:1} are based on this estimator. We introduce this estimator in Section~\ref{sec:1} and discuss 
some of its virtues and vices. In addition to tail-index estimation we also discuss the related problem of
estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). 
%{\red check} It turns out 
%that the simultaneous estimation of the tail-index and the scale parameter are strongly related.
\par
In Section~\ref{sec:2} we discuss the theoretical problem of appearance of power-law tails in models 
for daily or, more generally, low-frequency
return data. In particular, in Section~\ref{subsec:garch} we address the power-law tails of univariate and multivariate GARCH models as potential models
for a set of return data from distinct assets. As a matter of fact, under mild conditions, the aforementioned models
have power-law tails due to their relation with so-called {\em stochastic recurrence equations}. Moreover, some of the {\em standard} multivariate 
GARCH models as the CCC ensure that the component-wise marginal \ds s have power-law tails with the same index.
\par
In Section~\ref{sec:3} we discuss an economic argument for the fact that return data of similar assets
(such as return series in a given sector of the S\&P 500 index) may have tail-indices which are close to each other.
We argue based on a  utility \fct\ approach. We explicitly recognize the behavioral
concern for downside risk in an investor's evaluation of a portfolio
using the framework of Generalized Disappointment Aversion (GDA)
introduced by Routledge and Zin \cite{routledge2010generalized}. GDA
is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991} who derived DA from first principles (axiomatic).

In Section~\ref{sec:4} we summarize the discussion of the previous
sections.

\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
In this section, we assume the model \eqref{eq:1} for the tails of the marginal
\ds\ of a univariate return series $(X_t)$. For the sake of argument, we assume
that this series constitutes a strictly stationary \seq . In what follows, we focus
on the left tail of the \ds , i.e., on the losses. 
However, it is common to present the tail-index estimators
for positive data. Therefore we will multiply the losses $X_t$ by
minus one, swapping the negative with the positive values.
For simplicity, we also suppress subscripts in the notation:
\beam\label{eq:1a}
\P(-X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeam
where we assume that the two parameters -- the tail-index  $\alpha$ and the scale parameter
$c$ -- are positive. They play crucial roles for the understanding of the risk hidden in the data, hence 
for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
useful description of the risk, for example in terms of high quantiles such as
Value-at-Risk. Alternatively, these parameters can be
used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
%With these motives in mind, we present a survey of the
%tail-indices and scale parameters of 3 sectors, namely ``Energy'',
%``Consumer Staples'' and ``Information Technology'', of the S\&P 500
%index.

\subsection{Hill estimates of lower tail-indices}\label{sec:Hill}
Various estimators of the tail-index $\alpha$  in the model
\eqref{eq:1a} have been proposed in the literature;
see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, de
Haan and Ferreira \cite{haan:ferreira:2006}, Resnick
\cite{resnick:2007}. The most popular among them was introduced  by
Hill \cite{hill1975simple}.
Given a sample $-X_1,\ldots,-X_n$ whose marginal \ds\ satisfies  \eqref{eq:1a}, calculate
the order statistics $X_{(1)}\le \cdots\le X_{(n)}$  and construct the
{\em Hill estimator}:
\beao%\label{eq:Hill_index}
  \hat \alpha_k = \Big(
    {1 \over k} \sum_{i=1}^k \log {X_{(n-i+1)} \over X_{(n-k)}}
    \Big)^{-1}\,.
\eeao
Here $k$  is the number of upper order statistics in the sample used
for the estimation. The estimator $\hat \alpha_k$ is 
a maximum-likelihood estimator of $\alpha$ based on the $k$ upper
order statistics in the pure Pareto model (recall that 
we multiplied the data by minus one)
\beam\label{eq:3}
\P(-X_t>x)= \dfrac{K^\alpha}{x^\alpha}\,,\qquad x>K\,,
\eeam
under the hypothesis that we do not know the (high) threshold value
$K$. The estimator has ``good'' theoretical properties such as 
\asy\ consistency and \asy\ normality. These properties hold under
strict stationarity assumptions on the data; Drees and Rootz\'en 
\cite{drees:rootzen:2010} give perhaps most general conditions for
dependent \seq s and de Haan and Ferreira \cite{haan:ferreira:2006} provide
a complete \asy\ theory in the iid case.
\par
A major problem for Hill estimation is the choice of the number $k$ of
upper order statistics. As  a matter of fact,
if $k$ is too large the order statistics are too close to the center
of the \ds\ of the $-X_t$, leading to a bias
of the estimator. On the other hand, by construction, $\hat \alpha_k$ is an average of $k$ log-differences of the 
data. Therefore, the variance of the estimator is the larger the
smaller $k$. For these reasons, \asy\ theory requires 
to choose $k=k_n$ \st\ $k_n\to\infty$ and $k_n/n\to 0$ as $\nto$. This
fact does not make the estimation of $\alpha$ an  
easy matter: one has to choose a ``small'' value $k$ which is not
``too large''. For practical purposes,
a so-called {\em Hill plot} is recommended where $\hat \alpha_k$ is
plotted for a variety of $k$-values, corresponding to
some high  quantile $X_{(n-k)}$ of the data. Then $k$ is chosen from a
region in the plot where $\hat \alpha_k$ is relatively
stable. For example, in Figure~\ref{fig:1} we have chosen $k=50$ from
a sample of size $n=1304$, corresponding to the 96\%-quantile of the
data. In general, the estimation of the tail-index is an art and
requires some expertise; for some guidance
see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997},
Resnick \cite{resnick:2007} and Drees et al. 
\cite{drees:resnick:2000}. 
\par
In Figure~\ref{fig:1},
we exhibit 95\% \asy\ confidence bands derived from  the \clt
\beam\label{eq:2}
\sqrt k\, \big(\hat \alpha_k - \alpha\big) \std N(0, \alpha^2)\,,
\eeam
i.e., $\hat \alpha_k$ is \asy ally unbiased and has variance
$\alpha^2/k$. Since $k/n\to 0$ this means that the 
confidence bands are significantly larger than the classical
$1/\sqrt{n}$-rates. This fact is one explanation for the fact that it is
difficult to say something meaningful about the true value of
$\alpha$. There exist various other reasons why one should
not have 100\% trust in the confidence bands shown in
Figure~\ref{fig:1}. Indeed, \eqref{eq:2}
holds under rather subtle {\em second order conditions} on the tail
$\P(X_t>x)$ which cannot be verified on data. However, 
given a theoretical model such as the GARCH, these conditions can be
verified based on the theoretical properties of the model. If they are not satisfied the 
Hill estimator may exhibit significant bias; see Embrechts et
al. \cite{embrechts:klueppelberg:mikosch:1997} and Resnick \cite{resnick:1987} for illustrations 
of this fact leading to so-called ``Hill horror plots''. Moreover, the
Hill estimator is rather sensitive to non-stationarity 
of the data and to dependence. For example, results in Drees \cite{drees:2008}, and Drees and
Rootz\'en \cite{drees:rootzen:2010} show that the \asy\ variance of the
Hill estimator can be significantly larger than in the iid case. 
Since return data are dependent, the \asy\ confidence bands should be
even wider than exhibited in Figure~\ref{fig:1}. 
Again, only under he assumption of a concrete model like GARCH these
confidence bands can be evaluated and therefore the bands shown in Figure~\ref{fig:1}
just show some benchmark which holds in the iid case and under additional conditions on the tail \asy s.
\begin{figure}[htb!]
  \begin{minipage}{1.0\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
    {Energy_lower.pdf}
  \end{minipage}
  \begin{minipage}{1.0\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
    {Consumer_Staples_lower.pdf}
  \end{minipage}
  \begin{minipage}{1.0\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
    {Information_Technology_lower.pdf}
  \end{minipage}
  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail-indices $\alpha$ of
    daily return series in sectors of the S\&P 500
    index. The data span
    from 1 January 2010 to 31 December 2014 and comprise $n=1304$ observations.
The graphs from top to bottom correspond to the ``Energy'',
    ``Consumer Staples'' and ``Information Technology'' sectors.
    Each circle corresponds to a Hill estimate $\hat\alpha_{50}$; the gray
    triangles above and below it mark the 97.5\% and 2.5\% quantiles
    of its approximate normal distribution; see \eqref{eq:2} and the discussion following it for an 
interpretation.
    The lower and upper red lines mark the medians of the 2.5\% 
    and 97.5\% quantiles, respectively, evaluated from all stocks in the sector.
    The data are taken from {\it Yahoo Finance}; the labels on
    the horizontal axes are Yahoo symbols of the stocks. 
  }\label{fig:1}
\end{figure}

In Figure~\ref{fig:1} we see significant overlap of the confidence intervals of the Hill
estimates of the losses in the ``Energy'' and ``Consumer Staples''
sectors of the S\&P 500 index, as well as those of a 
large portion of losses in the ``Information Technology'' sector.
This fact indicates that the returns in each sector may 
have comparable tail-indices.
\par
Hoga's \cite{hoga:2016} test about the change of extreme quantiles
in a sample
may provide some further insight about how similar these tail-indices are.
Different tail-indices are likely to result in different
extreme quantiles. Nevertheless, changes in the extreme quantiles may also
result from changing scale parameters in the tail. Therefore  we first investigate the scale
parameters of daily stock returns in the same sectors of S\&P 500 before we apply the test.

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha_k} X_{(n-k)}\,.
\eeao
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, Theorem 2.2.1), one can show
\beao
  \sqrt{k}\, (\hat K_k - K) &\std& N\big(
    0, (K /\alpha)^2\big)\quad\mbox{and}\quad
  \sqrt {k} \,(\hat K_k^\alpha - K^\alpha) \overset{d}{\to} N\big(
    0, K^{2\alpha}  \big)\,,
\end{eqnarray*}
where the tail-index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of $\hat K_k$
and $\hat K_k^{\hat\alpha}$ can be constructed.   
Estimates $\hat K_k$ in the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {CS_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {IT_K.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {CS_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {IT_scale.pdf}
  \end{minipage}
  \caption{\small Estimates of $\hat K_k$ (top) and $\hat K_k^{\hat
      \alpha}$ (bottom) on $\log_{10}$-scale of stocks in sectors
    of the S\&P 500 index. The estimates are ordered according to the 
    corresponding estimated $\alpha$-values.
    The points are the estimated values, the bars 
    the \asy\ 95\%-confidence intervals; the confidence bands of the
    corresponding Hill estimates $\hat \alpha_k$ of these sectors are
    shown in Figure~\ref{fig:1}. 
  }
  \label{fig:sectors_parameters}
\end{figure}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd row of
Figure~\ref{fig:sectors_parameters}.
\par
Secondly, it appears that there is positive dependence between the
values of $\alpha$ and $K$.
As argued in Section~\ref{sec:pareto_tail}, this is consistent with
the assumption that the return series have Pareto tails on both sides
with the tail parameters on each side independent of those on the
other.

\par
Thirdly, Figure~\ref{fig:sectors_parameters} shows that, on
average, the values of $K$
in the ``Energy'' and ``Information Technology'' sectors are  larger
than those in the ``Consumer Staples'' sector.  For a given loss \pro y, 
a larger value of $K$ implies that large losses are more
probable. Thus one can conclude that these two sectors are
considerably riskier than the ``Consumer Staples'' sector. {\red Casper:
we leave this sentence to your discretion.} This is of course a
confirmation of one's economic instinct.

Yet another indication from Figure \ref{fig:sectors_parameters} is
that, while the ``Energy'' and the ``Information Technology'' sectors
are similar in riskiness, the dependence  between $\alpha$ and $K$ is
stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, when moving along
a curve of equal preference in the direction of increasing $\alpha$, the parameter $K$ also
increases. So the strong positive dependence seen in the ``Energy''
sector suggests that these stocks might have very similar investor
preferences. This in turn may be attributed to stronger business
relations between the energy enterprises. While two IT companies
may provide a variety of products and services and do not depend on each
other, two energy companies are more likely to depend on each other
via relations of supplier and customer or otherwise to compete with each
other if they are on the same link of the chain of energy production
and distribution.

\subsection{A test for equal tail-indices based on Hill estimation}
Suppose we have two independent strictly stationary positive series $X_1, \dots, X_n $ and
$Y_1, \dots, Y_n$ with corresponding distribution functions $F_X$ and
$F_Y$ that have power-law tails  with indices $\alpha_X$ and
$\alpha_Y$, respectively. From \eqref{eq:2} one can deduce
\begin{equation}
  \label{eq:x3}
  \sqrt k
  \begin{pmatrix}
    {\hat \alpha_X - \alpha_X} \\
    {\hat \alpha_Y - \alpha_Y} \\
  \end{pmatrix} \overset{d}{\to}
  \begin{pmatrix}
    Z_X \\
    Z_Y
  \end{pmatrix}
  \sim
  N\left(
    0, \text{diag}(\alpha_X^2, \alpha_Y^2)
  \right)
\end{equation}
where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimators of
$\alpha_X$ and $\alpha_Y$; we suppress their dependence on $k$. Then it follows from the continuous mapping
theorem
\begin{equation}
  \label{eq:x1}
  \sqrt k [(\hat \alpha_X -\alpha_X) - (\hat \alpha_Y - \alpha_Y)]
  \overset{d}{\to}
  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)\,.
\end{equation}
This relation allows one to construct an \asy\ test under 
the null hypothesis $\alpha_X = \alpha_Y$ and with test statistic $\hat \alpha_X - \hat \alpha_Y$.
We apply this test to the equities in the ``Energy'',
``Consumer Staples'' and ``Information Technology''
sectors of the S\&P 500 index. The results are shown in the top row of
Figure~\ref{fig:PairTest}.
\begin{figure}[htb!]
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{HillTest_Energy.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{HillTest_CS.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{HillTest_IT.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_Energy_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_CS_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_IT_pair.pdf}
  \end{minipage}
  \caption{{\em Top row}: Test for pairwise equality of tail-indices
    of losses in the  ``Energy'', ``Consumer Staples'' and
    ``Information Technology'' 
    sectors of S\&P 500. The test statistic in $\hat\alpha_X-\hat \alpha_Y$ is based
    on Hill estimates 
    of $\alpha_X$ and $\alpha_Y$. 
    The green, blue and red points correspond to pairs of stock in a sector
    when the test statistic is outside the intervals $[q_{0.075},q_{0.925}]$,
    $[q_{0.05},q_{0.95}]$,  $[q_{0.025},q_{0.975}]$, respectively, where
    $q_p$ is the $p$-quantile of the limiting 
    $N(0,\alpha_X^2+\alpha_Y^2)$-\ds\ of the test statistic in
    \eqref{eq:x1}. Grey points stand for pairs for which the test
    statistic is inside $[q_{0.075},q_{0.925}]$. 
    {\em Bottom row:} Test for changing tail-index or scale parameter
    of losses
    using Hoga's test based on concatenated series of pairs of
    stocks. The green, blue and red points
    correspond to pairs of stock in a sector 
    when the test statistic $T_n$ exceeds the $85\%$-, $90\%$-,
    $95\%$-quantile of the limit \ds .
  Grey points stand for pairs for which the test statistic is below
  the \asy\ $85\%$-quantile. Black points represent 
    pairs for which the computation of $T_n$ fails for given precision  
    requirements and time limits.
    The same number (50) of upper order statistics is used for both tests.}
  \label{fig:PairTest} 
\end{figure}
They indicate that tail-indices of equities in the
``Energy'' or ``Information Technology'' sectors are more variable 
than in the ``Consumer Staples'' sector, as the null
hypothesis is rejected more often for members of these two former sectors.
Moreover, these figures suggest that the test based on the Hill estimator
is quite powerful in distinguishing between tail-indices. In contrast to the test presented in Section~\ref{sec:Hoga}
the present test results in  more rejections for the ``Energy'' and the
``Information Technology'' sectors.
\par
As a caution, one should bear in mind that 
\eqref{eq:x3} is valid on condition that the $X$- and $Y$-
series are independent of each other (or weakly dependent on each other), which is generally untrue for
two return series in the same market.

\subsection{A test for a change in the extreme tail}\label{sec:Hoga}
Here we apply a test from
a recent paper by Hoga \cite{hoga:2016}. This test has been developed
for a different kind of problem. Given a strictly stationary
\ts\ $X_1,\ldots,X_n$ with a marginal \ds\ $F$ with right power-law
tail, the goal is to test whether there is a structural break of the
{\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to
zero. If the tail-index {\em or} the scale parameter in a \ds\ of type
\eqref{eq:3} change inside a sample, then it is likely that the
extreme quantiles change as well. We will test for a change of tail-index or scale parameter in this indirect way.
\par
The null hypothesis of the test in \cite{hoga:2016} is that there is no change of the extreme quantiles $F^{-1}(1-p)$ 
for $p=p_n\to 0$ in any subsample with indices
$t\in (n\,t_0,n(1-t_0))$ where $t_0$ is a fixed number in $(0,0.5)$. Writing $\hat x_p(a,b)$ for an estimator of the extreme $(1-p)$-quantile 
based on the subsample with indices $t\in (na,nb)$, the test statistic is given by
\beam\label{eq:4}
T_n = \sup_{s \in [t_0, 1 - t_0]}
  \dfrac{  \big[s (1 - s) \log \big(\hat x_p(0, s)/\hat x_p(s, 1)\big)
    \big]^2}{
    \int_{t_0}^s\big[r \log \big( \hat x_p(0, r)/\hat x_p(0, s)
      \big)
    \big]^2 dr
    +
    \int_{s}^{1 - t_0}
    \big[
      (1 - r) \log \big(
      \hat x_p(r, 1)/
      \hat x_p(s, 1)
      \big)
    \big]^2 dr}\nonumber\\
\eeam
Under the null hypothesis, $(T_n)$ converges to a complicated \fct al
of Brownian motion\ on $[0,1]$; the \asy\ quantiles need to be
evaluated by simulation.
\par
When applied to our problem we would like to test 
whether there is a change of the tail-index {\em or} scale parameter in \eqref{eq:3} in each of the S\&P 500 
series in the distinct sectors. We also  want to get some indication about a possible change of tail-index or
scale parameter from one series to another within a given sector. For this reason, we choose any pair of series
within a sector and concatenate each of the paired series. Then we run the test on the concatenated series.
Of course, despite the fact that we test changes of tail-index or scale parameter 
{\em in a very indirect way} -- there may be many other reasons for the change of extreme quantiles in a sample -- 
we also concatenate two rather distinct series. Even if we assume that the two series come from related models
(such as GARCH), the parameters of these models will in general not be the same. Moreover, the concatenation
of two strictly stationary \ts\ is in general not strictly stationary. Therefore we have to be careful
with interpretations of the results of the tests.
\par 
In Figure~\ref{fig:Hoga_Single} we show the values of the test statistic $T_n$ (horizontal bars)
for $t_0=0.1$ and daily return series of stock in the ``Energy'' and ``Consumer Staples'' sectors of the S\&P
500 index. The ``null hypothesis'' is that the tail-index and scale parameter remain the
same throughout the selected period of time. For most stocks, the hypothesis cannot be rejected even at the 85\% level.
This fact may be an indication that the \ds\ inside a series is rather homogeneous.
Alternatively, it may show that the power of the test is very low. A possible reason for this suspicion is that the \con\ rate
of $(T_n)$ to its limit is very slow, i.e., the \asy\ \ds\ is not representative for the \ds\ of $T_n$ for the chosen $n$; for some
simulation evidence, see below.
\par
To check whether any pair of stocks shares the same tail-index and scale parameter we concatenate
any two series and apply the aforementioned test on the concatenated series. For the ``Energy'' and the
``Consumer Staples'' sector we summarize the results in the bottom row of Figure~\ref{fig:PairTest}.
These graphs show that the ``null
hypothesis'' of an equal tail-index and scale parameter is rejected for more pairs in the
``Energy'' sector than it is for those in the ``Consumer Staples''
sector. This suggests that lower tail-indices of stocks in the
``Energy'' sector are more spread out than those of the ``Consumer
Staples'' sector. Also observe that while 3 stocks, say A, B and C, test in favor of
the relations $\alpha_A = \alpha_B, \alpha_A \neq \alpha_C$, it often
happens that another test on B and C is supportive of $\alpha_B = \alpha_C$. Again,
this is due to the limited power of the test. Based
on such results, one may guess that $\alpha_B$ lies between $\alpha_A$ and
$\alpha_C$. The test is unable to recognize the smaller differences
between $\alpha_A, \alpha_B$ on one hand and between $\alpha_B, \alpha_C$ on the other hand.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_Energy_Single.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]
    {Hoga_CS_Single.pdf}
  \end{minipage}
  \caption{Test statistic $T_n$ from \eqref{eq:4} for the  stocks in the ``Energy'' and
    ``Consumer Staples'' sectors of S\&P 500. The green, blue and red
    lines correspond to the 85\%, 90\% and 95\% quantiles of the limit \ds\ of $T_n$.  They are derived
by simulations from the limit \ds .}
  \label{fig:Hoga_Single}
\end{figure}

To get an idea about the power of the test we run it on a
sample concatenated from two independent iid samples of the same size
$n=1304$ as the S\&P 500 series. Both pieces are $t$-distributed with
distinct degrees of freedom. The results are shown in
Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
larger the minimum tail-index in the concatenated pair.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{t_sim_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
  \end{minipage}
  \caption{
    {\em Left}: Test  of concatenated $t$-samples with different
    degrees of freedom $\alpha$. Numbers on the axes are the degrees of freedom
    in the subsamples. For an interpretation of the colored bullets,
    see the caption for the bottom row of Figure~\ref{fig:PairTest}. The graph shows the limited power
    of the test. In particular, if both degrees of freedom are
    relatively large it loses the capability of distinguishing between the \ds s.
    {\em Right}: Comparison of the \asy\ distribution of the test statistic
    $T_n$ in \eqref{eq:4} under the null hypothesis and the \ds\ of
    $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees
    of freedom.
  }
  \label{fig:t_sim_pair}
\end{figure}

A major problem of this test is the \asy\ distribution of the test statistic under the
null hypothesis.  The rate at which the finite-sample distribution tends to its limit is not known.
To find out about this problem we compared the \ds s of $T_{1304}$ 
for $t$-distributed $X_t$ with  $\alpha=3$ and $\alpha=4$ degrees of freedom with
the limit \ds\ of $T_n$.
The estimated density functions are shown on the right of Figure~\ref{fig:t_sim_pair}.
As seen in the graph, the asymptotic distribution assigns
significantly more mass to the tail than the \ds s of $T_n$ do. For
comparison, we list a few quantiles of these \ds s in
Table~\ref{tab:HogaAsymptotic}, showing major differences between the
\asy\ and finite-sample \ds s.
\begin{table}[htb!]
  \centering
  \begin{tabular}{l|c|c|c|r}
    & \multicolumn{4}{c}{Quantiles} \\[2mm]
    \hline
    Distribution of $X_t$& 80\% & 85\% & 90\% & 95\% \\
    \hline
    Asymptotic & 47.48 & 59.61 & 78.90 & 113.12 \\
    t(3)  & 23.30 & 28.27 & 35.04 & 46.75 \\
    t(4)  & 25.76 & 31.13 & 38.81 & 56.32\\[2mm]
  \end{tabular}
  \caption{Quantiles of the test statistic $T_n$ for $n=1304$
    $t$-distributed samples with $\alpha=3$ and $\alpha=4$ 
degrees of freedom
    as well as the corresponding quantiles for the limiting \ds\ of
    $T_n$. In particular, there are huge differences between the three
    \ds s  for the higher quantiles. 
    }
  \label{tab:HogaAsymptotic}
\end{table}

Figure~\ref{fig:PairTest:permuted} points at another shortcoming
of the test: we show $T_n$ for an 
arbitrarily chosen random permutation of the concatenated data from two 
different stocks. In this case the null hypothesis that two series
have the same tail-index is rejected much more often, as a comparison
with the bottom graphs of
Figure~\ref{fig:PairTest} shows. If the data in the concatenated 
series were iid a random permutation would not change the 
\ds\ of $T_n$. Thus the value of the test statistic $T_n$  
strongly depends on the dependence structure of the underlying data
and therefore a test based on $T_n$ may be misleading.

\begin{figure}[htb!]
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_Energy_pair_permuted.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_CS_pair_permuted.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_IT_pair_permuted.pdf}
  \end{minipage}
  \caption{
    Test for changing tail-index or scale parameter
    of losses
    using Hoga's test based on concatenated series of pairs of
    stocks.
    A random permutation is applied to the observations of each
    series.
    The green, blue and red points
    correspond to pairs of stock in a sector 
    when the test statistic $T_n$ exceeds the $85\%$-, $90\%$-,
    $95\%$-quantile of the limit \ds .
    Grey points stand for pairs for which the test statistic is below
    the \asy\ $85\%$-quantile. Black points represent 
    pairs for which the computation of $T_n$ fails for given precision  
    requirements and time limits.
  }
  \label{fig:PairTest:permuted} 
\end{figure}

\newpage
\section{Some theoretical arguments for equality of tail-indices}\setcounter{equation}{0}\label{sec:2}
\subsection{Multivariate GARCH models whose components have  equal tail-indices}\label{subsec:garch}
Among the models for returns the generalized autoregressive conditionally heteroscedastic (GARCH) model
is certainly most popular because it is parsimonious, captures various of the stylized facts of real return data
and can also be modified in various directions to capture specific behavior of \ts\ such a asymmetry, skewness, long memory; 
see for example Andersen et al. \cite{andersen:davis:kreiss:mikosch:2009}, Part 1, for a collection of results
on GARCH-type models.
The original  {\em univariate} GARCH model of Bollerslev \cite{bollerslev:1986} is a stochastic volatility model of the type $X_t=\sigma_t\,Z_t$, where 
$(Z_t)$ is an iid mean-zero unit-variance \seq . In the simple case of a GARCH the {\em squared volatility} satisfies the {\em stochastic recurrence equation} 
\beam\label{eq:5}
\sigma_t^2= \alpha_0+\alpha_1 X_{t-1}^2+\beta_1\sigma_{t-1}^2=\alpha_0+(\alpha_1Z_{t-1}^2+\beta_1)\sigma_{t-1}^2\,,\quad t\in\bbz\,.
\eeam
Here $\alpha_0>0$, $\alpha_1,\beta_1$ are non-negative constants. For suitable choices of $\alpha_1,\beta_1$ the equation \eqref{eq:5}
can be solved and the solution $(\sigma_t^2)$ constitutes a strictly stationary \seq , implying that $(X_t)$ 
is strictly stationary itself. A remarkable property of the process $(\sigma_t)$ is that it has a power-law tail 
of the form
\beam\label{eq:6}
\P(\sigma_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeam
for some positive $c>0$ and a positive tail-index $\alpha$ which is the unique solution of the equation 
$\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$
provided that the solution exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
follows by an application of the Kesten-Goldie theorem; see Kesten \cite{kesten:1973}, Goldie \cite{goldie:1991}, cf. 
Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} for a 
recent textbook treatment. The latter result ensures power-law tails  for
the strictly stationary  solution $(Y_t)$ to the stochastic recurrence equation\ 
\beam\label{eq:7}
Y_t= A_t\,Y_{t-1}+B_t\,,\qquad t\in\bbz\,,
\eeam
for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^{\alpha/2}]=1$.  
In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2+\beta_1$ and $B_t=\alpha_0$ to achieve
\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016}, p.~275)
it follows that
\beao
\P(\pm X_t>x)\sim \E[(Z_t)_\pm^\alpha]\,\P(\sigma_t>x)\,,\qquad \xto\,, 
\eeao
implying power-laws for the right and left tails of $X_t$ caused by the power-law tail of $\sigma_t$.

A GARCH process of the order $(p,q)$ can be embedded in a multivariate
equation of the type \eqref{eq:7}, where $(\bfA_t)$ are iid random
matrices and $\bfB=\bfB_t$ is a constant vector. Again, the Kesten
theory \cite{kesten:1973} applies, implying that the marginal and
\fidi s of the GARCH process are \regvary\ with a positive index
$\alpha$. We refrain from explaining the notion of multivariate
\regvar\ which is needed in this context. For further details, see
Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} where the
Kesten theorem and \regvar\ of GARCH processes are explained in detail.
\par
There exist various extensions of the univariate GARCH
model to the multivariate case. For the sake of argument, we stick here to the 
{\em constant conditional correlation} (CCC) model of Bollerslev
\cite{bollerslev:1990} and Jeantheau \cite{jeantheau:1998}, and we only consider a special bivariate case.
It is the model 
\beao\bfX_t=
\left(\barr{l}X_{1,t}\\
X_{2,t}\earr\right)= \left(\barr{ll}\sigma_{1,t}& 0\\
0&\sigma_{2,t}\earr\right)\,\left(\barr{l}Z_{1,t}\\Z_{2,t}\earr\right)=\Sigma_t\,\bfZ_t\,,\qquad t\in\bbz\,.
\eeao
Thus both return components $X_{i,t}$ have the form of a univariate
stochastic volatility model $X_{i,t}=\sigma_{i,t}Z_{i,t}$ 
with non-negative volatility $\sigma_{i,t}$ and an iid bivariate noise \seq\ $(\bfZ_t)$ with zero mean and unit variance components.
We also have the specification
\beam\label{eq:8}
\bfY_t=\left(\barr{l}\sigma^2_{1,t}  \\  
\sigma^2_{2,t}\earr
\right)
&=& \left(
\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)
+\left(\barr{cc}\alpha_{11} & \alpha_{12}  \\
      \alpha_{21} & \alpha_{22}\earr \right)\, 
\left(\barr{l}X_{1,t-1}^2  \\X_{2,t-1}^2   \earr\right)
 + \left(\barr{cc}\beta_{11} & \beta_{12}  \\\beta_{21} & \beta_{22} \earr
 \right)\,\left(\barr{c}\sigma^2_{1,t-1}  \\\sigma^2_{2,t-1}\earr
  \right)\nonumber\\
&=& \left(
\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)+\left(\barr{cc}\alpha_{11}Z_{1,t-1}^2+\beta_{11}&\alpha_{12}Z_{2,t-1}^2+
\beta_{12}\\
\alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
\earr\right)\,\left(\barr{l}\sigma_{1,t-1}^2\\\sigma_{2,t-1}^2\earr
\right)\,,
\eeam
for positive $\alpha_{0i}$ and suitable non-negative 
$\alpha_{ij},\beta_{ij}$, $i,j=1,2$.
Writing
\beao
\bfB_t= \left(
\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)\quad\mbox{and}\quad
\bfA_t=\left(\barr{cc}\alpha_{11}Z_{1,t-1}^2+\beta_{11}&\alpha_{12}Z_{2,t-1}^2+
\beta_{12}\\
\alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
\earr\right)\,,
\eeao
we see that we are again in the framework of a stochastic recurrence equation\ 
but this time for vector-valued $\bfB_t$ and matrix-valued $\bfA_t$:
\beam\label{eq:jan6b}
\bfY_t=\bfA_t\,\bfY_{t-1}+\bfB_t\,,\qquad t\in\bbz\,.
\eeam
Kesten \cite{kesten:1973} also provided the corresponding theory  
for stationarity and tails in this case. \sta\ \cite{starica:1999}
dealt with the corresponding problems for CCC-GARCH processes,
making use of the theory in Kesten \cite{kesten:1973},
Bougerol and Picard \cite{bougerol:picard:1992}
and its
specification to the tails of GARCH models 
in Basrak et al.~\cite{basrak:davis:mikosch:2002}. \sta\ \cite{starica:1999} assumed the 
Kesten conditions for the matrices $\bfA_t$. These conditions ensure that the product matrices $\bfA_1\cdots\bfA_n$ 
have positive entries for sufficiently large $n$. Then Kesten's theory implies that
all components of the vector $\bfX_t$ have power-law tails with the same index $\alpha$ and also
that the \fidi s of the process $(\bfX_t)$ are \regvary\ with index $\alpha$. 
\par
Various GARCH modifications are derived by considering linear combinations of CCC-GARCH models.
The property of multivariate \regvar\ of multivariate GARCH ensures that, after linear transformations,  
the new process in all components has again power-law tails with the same index as the original GARCH process; see 
Basrak et al.~\cite{basrak:davis:mikosch:2002}. 
Models which are constructed in this way are
the Orthogonal GARCH model of
Alexander and Chibumba \cite{alexander:chibumba:1996}, its
generalization GO-GARCH by van der Weide \cite{Weide2002},  the Full Factor GARCH model of Vrontos et al.
\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
model of Lanne and Saikkonen  \cite{lanne2007modelling}. These
models are characterized by their treatment of each series as a linear
combination of factors, and each of the factors is modeled as a GARCH
process; see Silvennoinen and Ter\"asvirt\"a \cite{silventeras:2009}.
\par
Not all choices of $\alpha$- and $\beta$-parameters in the model \eqref{eq:8} allow for an
application of the Kesten theory. For example, assume that only the
diagonal elements $\alpha_{ii}$ and $\beta_{ii}$ are positive.
Then $\bfA_t$ is diagonal and, hence, the condition that
$\bfA_1\cdots\bfA_n$ have positive entries for sufficiently large $n$ 
cannot be satisfied. In the latter situation, both $(X_{1,t})$ and
$(X_{2,t})$ are univariate GARCH processes. Assuming the 
conditions of the univariate Kesten-Goldie theorem for each component
process, $(X_{1,t})$ and $(X_{2,t})$ have power-law tails 
with indices $\kappa_1$ and $\kappa_2$, respectively,  given by the solutions to the equations 
$\E [(\alpha_{ii} Z_{i,t}^2+\beta_{ii})^{\kappa_i/2}]=1$, $i=1,2$. In
this model, one can introduce dependence between the two component
series $(X_{1,t})$ and $(X_{2,t})$ by assuming dependence between the
noise variables $Z_{1,t}$ and $Z_{2,t}$. Another situation when the
Kesten theory fails 
appears when $\bfA_t$ is an upper or lower triangle matrix: then the
products  $\bfA_1\cdots\bfA_n$ are always of the same triangular
type. 
Similar remarks apply when one considers a CCC model in general
dimension. Of course, one may argue that the latter models 
are not natural: they are degenerate since they do not allow 
for a linear relationship
between all squared volatilities on a given day.

\subsection{A utility based argument for equal tail-indices}\label{sec:3}
In this section we give an argument based on economic theory that
suggests equality of  tail-indices for equity return series.
We follow an approach by Routledge and Zin
\cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky
payoff $C$ of an investor and assume that it has a continuous \ds\ on
$(0,\infty)$ with distribution function $F_C$. Let $u$ be a utility \fct\ assumed to be
increasing and concave on $(0,\infty)$.  Following Routledge and Zin
\cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\,,
  %&=&
  %\E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$.
  Here $v$ can be thought of as the {\em certainty payoff} equivalent to the
 risky payoff $C$; $\delta$ tunes the {\em threshold of 
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappointment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
\par
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:hjyr}
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
\eeam

For the sake of argument, we assume that an investor initially has one
unit of wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
bond with interest rate $r>0$ and $\phi$ units in a risky asset with
return $X$ over one time unit, i.e.,
\beam\label{eq:xxie1}
  C(X) = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
\eeam
Then we have 
\beao
\wt u(F_X, \phi) &=& \E [u(C)] + b\, \E \big[u(C)\1(C \le  \delta v)\big] - b \,u(\delta\, v) F_X(q)\,,
\eeao
where $F_X$ is the distribution function of $X$  and
\beao
  q = %q(r, \phi) &:=& \log \left( {
      %\delta v - (1 - \phi) e^r
      %\over
      %\phi
    % \right) 
\log\Big(
    \ex^r + {\delta \,v - \ex^r \over \phi}
  \Big)\,.
\eeao
Note that $C \le \delta v$ \fif\ $X \le  q$. 
\par
Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher (on average) return from the risky asset than he is guaranteed from the
riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
$q > r $. For given $b,\delta,v$, the \fct al $\wt u$  depends only on $\phi$ and $F_X$,
$\wt u=\wt u (F_X, \phi)$.
We assume that
\beao
\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
\eeao
exists and that the maximum is achieved at a unique $\hat\phi \in (0, 1)$.

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
we assume the following toy model. 
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where
$\alpha, \beta > 0$, $K,K' > 0$, $0 < p < 1$. 
We also write $f_X$ for the density function of $F_X$.
\par
We have
\beam\label{eq:xxie1.0}
 \wt u(F_X, \phi)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x <  q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
\par
We observe the following property 
whose proof is given in Appendix~\ref{sec:thrmI_proof}. 
% which also applies to the special case 
%\eqref{eq:xxie1.0}. 
\ble\label{thrm:I}
Assume the two-sided Pareto model \eqref{eq:pareto},
that there is no \fct al relationship between $\alpha,K$ and $\beta,K'$
and the utility \fct\ $u$ is increasing and differentiable. Then
$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and
$\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.
\ele
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with
$K$. Therefore there is a curve of  equal preference on the $(\alpha,
K)$-plane. Moving along this curve in the direction of increasing
$\alpha$, one expects the values of $K$ to increase too, i.e., the
estimated values of $\alpha$ and $K$ 
should appear positively dependent. Figure~\ref{fig:preference_pareto}
illustrates this scenario for $\xi = 1/2, 4$ for the power-utility
\fct\ \eqref{eq:hjyr}.
In fact, this positive dependence is indeed observed for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure~\ref{fig:sectors_parameters}.
\par
A particularly interesting case occurs  when 
$F_X$ is symmetric, i.e., when $\alpha = \beta$, $K =K'$ and $p=0.5$.
Then \eqref{eq:pareto} turns into 
\begin{eqnarray}
  \wt u(F_X, \phi) &=& {\alpha \over 2} K^\alpha (1 + b)
  \int_{0}^\infty {
    u(C(x)) \left[
      1 - {b \over 1 + b}\1_{\{x \ge q\}}
      \right]
    + u(C(-x))
    \over
    (K + x)^{\alpha + 1}
  } dx
  - b u(\delta v) F_X(q)\,.\nonumber\\
  \label{eq:ipf}
\end{eqnarray} 
This situation is not covered by Lemma~\ref{thrm:I}:
for the proof of the latter result we used Lemma \ref{lemma:I} whose
assumptions are not satisfied in the present situation.
Indeed, the integrand 
\[
U_{\text{all}}(x) =
u(C(x)) \left[
  1 - {b \over 1 + b}\1(x \ge q)
  \right]
+ u(C(-x))
\]
is not monotone.\footnote{To see this we may plug \eqref{eq:hjyr} 
in $U_{\text{all}}$ and re-write it as
\[U_{\text{all}}(x) =
  -{\phi^{-\xi} \over \xi} \Big\{
  \underbrace{
    \Big(1 - {b \over 1 + b}\1(x \ge q)\Big)
    \big[a + e^x\big]^{-\xi}
    +
    \Big[a + e^{-x}\Big]^{-\xi}
  }_{U(x)} \Big\}\,,
\]
where $a = (1 - \phi) e^r/\phi$.
Direct computation gives
\begin{eqnarray*}
  U'(x)
  &=&
  -{(a + e^x)^{-\xi - 1} \xi e^x
    \over
    1 + b \1(x \ge  q)
  } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}\,,
  \qquad x \neq q
\end{eqnarray*} 
The \fct\ $U(x)$, hence $U_{\text{all}}$, 
is not monotone because $U'(x)>0$ for all large $x$ while $U(x)$ decreases
in a small neighborhood of $q$.}
\par
We resort to numerical methods to gain some understanding of
how $\tilde u(F_X, \phi)$ changes with $\alpha$.
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for
given values of $K, K', \alpha, \beta$.  This is shown in 
Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\
\eqref{eq:hjyr} both for fixed $K', \beta$ and for $K=K'$, $\alpha=\beta$.  
The corresponding values $\tilde u_{\rm max}(\alpha, K)$
are shown in Figure \ref{fig:preference_pareto}.
If $K'$ and $\beta$ are fixed, both $\wt u_{\rm max}(\alpha, K)$ 
and $\hat\phi$ increase  with $\alpha$ and decrease with $K$. 
This is in agreement with Lemma~\ref{thrm:I}. 
In contrast, when $K=K'$ and $\alpha=\beta$
$\wt u_{\rm max}(\alpha, K)$ decreases with $\alpha$ 
but is rather insensitive with respect to $K$. On the
other hand, $\hat\phi$ is not monotone with respect to $\alpha$ or
$K$. For each fixed $K$, it peaks at an $\alpha$-value somewhere below
1. For realistic values $\alpha\in (2, 4)$, $\hat\phi$ is a
small value below 5\%.
Since $\wt u_{\rm max}(\alpha, K)$ decreases with
$\alpha$, investors who seek to maximize $\wt u_{\rm max}(\alpha, K)$
will prefer the smallest $\alpha$ in the market, resulting in similar
values of $\alpha$ for different equities.
\begin{figure}[htb!]
  \begin{minipage}{0.25\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1_A.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.25\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4_A.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.25\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.25\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
 \caption{
   {\em The 1st and 2nd} graphs show $\hat\phi$, the optimal equity
   allocation as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
   \eqref{eq:pareto} for fixed $K'=0.012$, $\beta = 1.4$.
   {\em The 3rd and 4th} graphs show $\hat\phi$ as a \fct\ of $\alpha$
   and $K$ with $\beta = \alpha$ and $K' = K$.
   We choose the utility \fct\ $u$ from \eqref{eq:hjyr} for $\xi = 1/2$
   and $\xi = 4$, $b = 0.01$ in all cases.
  }
  \label{fig:phi_hat_pareto}
\end{figure}

\begin{minipage}{0.5\linewidth}
  \includegraphics[width=\textwidth]{power_utilities.png}
\end{minipage}\hfill
\begin{minipage}{0.42\textwidth}
  The economitrical differences between the two utility functions
  are illustrated in the graph to the left.
  We see that
  $u(x)$ grows slower and saturates later for small $\xi$ than for
  large $\xi$. The latter case represents an agent who is more
  tolerant about low consumption and seeks wealth more
  aggressively. In other words, he is less risk-averse than one with a
  larger $\xi$. Such an agent will therefore invest more heavily in
  equity, as shown in Figure~\ref{fig:phi_hat_pareto}. Note that,
  while $\hat \phi$ changes nearly in the same in both graphs, the
  scales in the graphcs are very different.
\end{minipage}

\begin{figure}[htb!]
  \begin{minipage}{0.25\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto5e-1_A.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.25\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto4_A.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.25\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.25\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto4.pdf}
  \end{minipage}
  \caption{
   {\em The 1st and 2nd} graphs show
   $\tilde u_{\rm max}(\alpha, K)$, as a \fct\ of $\alpha$ and $K$
   in the two-sided Pareto model \eqref{eq:pareto} with $K'=0.012$,
   $\beta = 1.4$.
   Clearly, $\tilde u_{\rm max}(\alpha, K)$ increases with $\alpha$
   and decreases with $K$ when $K'$ and $\beta$ are fixed.
   {\em The 3rd and 4th} graphs show
   $\tilde u_{\rm max}(\alpha, K)$ as a \fct\ of $\alpha$
   and $K$ with $\beta = \alpha$ and $K' = K$.
   We choose the utility \fct\ $u$ from \eqref{eq:hjyr} for $\xi = 1/2$
   and $\xi = 4$. $b = 0.01$ in all cases.
  }
  \label{fig:preference_pareto}
\end{figure}

If the parameter $b$ is very small the GDA preference is closely
approximated by the mean-utility preference, corresponding to $b = 0$. 
In this case, we show in the proof of Lemma \ref{lemma:III} that the function
$U_{\text{all}}(x)$ may increase or decrease depending on particular conditions
on the values of $\xi$ and $(1 - \phi) e^r / \phi$:
\begin{enumerate}
\item If $\max\{a, 1\} < \xi$, $U_{\text{all}}(\cdot)$ is
  monotone decreasing.
\item If $a < \xi < 1$ and $(a + y_-)/(a y_- + 1) <
  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
  decreasing.
\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotone
  increasing.
\item If $1 < \xi < a$ and $(a + y_+)/(a y_+ + 1) >
  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
  increasing.
\item In other case, $U_{\text{all}}(\cdot)$ is not monotone.
\end{enumerate}
where
\begin{eqnarray*}
a &=& {
  (1 - \phi) e^r
  \over
  \phi
} \\
y_\pm &=& {
  a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
  \over
  a (\xi - 1)
}
\end{eqnarray*}
Moreover, it is also easily checked that, when
$x \in (0, K(e^{1/\alpha} - 1)]$,
the density function in the integral of \eqref{eq:ipf} increases
with $\alpha$; when $x \in (K(e^{1/\alpha} - 1), \infty)$ it 
decreases with $\alpha$. Following the arguments for Lemma~\ref{thrm:I},
and applying Lemma \ref{lemma:I}, it can be seen that
$\wt u_{\rm max}(\alpha, K)$ increases/decreases with $\alpha$ when
$U_{\text{all}}(x)$  decreases/increases.

\section{Conclusion}
\label{sec:4}
We have established that, in the case of an equity return series with
two-sided, functionally independent Pareto tails, investor
preference functionals are monotone increasing/decreasing with the
tail index/scale parameters. Thus in a market dominated by such
equities, the investors would pursue the largest tail index in the
market, leading to a shared common tail index for all equities.

The empirical results presented in section \ref{sec:1} suggest this
may well be the case for the ``Consumer Staples'' sector of S\&P 500,
given the Hill estimates of tail indices shown in figure \ref{fig:1}
and the largely positive results of tests for equal tail indices shown
in figure \ref{fig:PairTest}.

On the other hand, we have also seen that, when the left and the right
tails have the same indices, investor preference over the equity has
more sophisticated variations in the parameters' space including the
tail parameters of the equity, the interest rate, the investor's risk
apetite as captured by his utility function, and his threshold of
disappointment.

We also acknowledge that our model of the market and the investor is a
simple one, not accoounting for the dependence between equities, nor
the categorization of investors and their interactions. These are
potential topics of future work.

\ifx\phdthesis\undefined
\appendix
\else
\begin{subappendices}
\fi

\section{A monotonicity lemma}
\setcounter{equation}{0}
\begin{lemma} \label{lemma:I}
  Assume distribution function $F(x, \theta)$ parameterized by 
  $\theta \in \Theta \subseteq \mathbb R$ has support
  $(a, b) \subseteq \mathbb R$,
  and in addition $F(x, \theta)$ has density
  function $f(x, \theta)$ that is differentiable with respect to
  $\theta$ for all $\theta \in \Theta$.
  Let $X \sim F$ and assume function $h(\cdot)$ is defined on $(a, b)$
  and is monotone throughout this interval.
  Moreover, we assume $h(x)$ and $f(x, \theta)$ satisfy
  \begin{equation}\label{eq:frt}
%    \E |h(X)| < \infty, \quad
    \int_a^b \left| {\pd f(x, \theta) \over \pd \theta} \right| dx
    < \infty \text { and }
    \int_a^b \left| h(x) {\pd f(x, \theta) \over \pd \theta} \right| dx
    < \infty
  \end{equation}
  Then the following holds true:
  \begin{enumerate}
  \item If $h(\cdot)$ is decreasing and $\exists x_0 \in (a, b)$ such that
    $\frac{\pd f}{\pd \theta}(x, \theta) > 0$ for $x \in (a, x_0)$ while
    $\frac{\pd f}{\pd \theta}(x, \theta) < 0$ for $x \in (x_0, b)$, then
    \[
    \frac{\pd \E h(X)}{\pd \theta} > 0
    \]
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    $\frac{\pd f}{\pd \theta}(x, \theta) < 0$ for $x \in (a, x_0)$  while
    $\frac{\pd f}{\pd \theta}(x, \theta) > 0$ for $x \in (x_0, b)$, then
    \[
    \frac{\pd \E h(X)}{\pd \theta} > 0
    \]
  \end{enumerate}
\end{lemma}
\begin{remark}
  \label{remark:I}
  Two other cases follow trivially from lemma \ref{lemma:I}:
  \begin{enumerate}
  \item If $h(\cdot)$ is increasing and $\frac{\pd f}{\pd \theta}$
    satisfies the
    same conditions of the 1st case of lemma \ref{lemma:I},
    $\frac{\pd \E h(X)}{\pd \theta} < 0$. This immediately follows
    from applying 1st case of lemma \ref{lemma:I} to $-h(\cdot)$.
  \item By the same argument, if $h(\cdot)$ is decreasing and
    ${\pd f \over \pd \theta}$ satisfies the same conditions of the
    2nd case of lemma \ref{lemma:I}, ${\pd \E h(X) \over \pd \theta} < 0$.
  \end{enumerate}
\end{remark}

\begin{proof}
  Firstly, by dominated convergence theorem, conditions \eqref{eq:frt}
  imply, for all $S \subseteq (a, b)$,
  \begin{eqnarray*}
    {\pd \over \pd \theta}\int_S f(x, \theta) dx
    &=&
    \int_S {\pd \over \pd \theta} f(x, \theta) dx \\
    {\pd \over \pd \theta}\int_S h(x) f(x, \theta) dx
    &=&
    \int_S h(x) {\pd \over \pd \theta} f(x, \theta) dx \\
  \end{eqnarray*}
  Thus we have
  \begin{eqnarray*}
    {\pd \E h(X) \over \pd \theta}
    &=&
    {\pd \over \pd \theta}
    \int_a^b h(x)
    f(x, \theta) dx \\
    &=& \int_a^b h(x)
    {\pd f \over \pd \theta}(x, \theta) dx \\
    &=& \underbrace{\int_a^{x_0}
      h(x) {\pd f \over \pd \theta}(x, \theta) dx}_{I_1}
    + \underbrace{\int_{x_0}^b
      h(x) {\pd f \over \pd \theta}(x,
      \theta) dx}_{I_2}
  \end{eqnarray*}
  $x_0$ being located in the interior of $(a, b)$ and
  $h(\cdot)$ being monotone imply $h(x_0) < \infty$.
  \begin{enumerate}
  \item When $h(x)$ is decreasing on $(a, b)$ and ${\pd f \over \pd
    \theta}(x, \theta) > 0$ on $(a, x_0)$
    \[
    I_1 > h(x_0) \int_a^{x_0}
    {\pd f \over \pd \theta}(x, \theta) dx
    \]
    Similarly, because ${\pd f \over \pd \theta}(x, \theta) < 0$ for
    $x \in (x_0, b)$ and $h(x)$ is decreasing, we have
    \begin{eqnarray*}
      I_2 &=& \int_{x_0}^b -h(x_0)
      \left|{\pd f \over \pd \theta}(x, \theta) \right| dx
      > -h(x_0)
      \int_{x_0}^b \left| 
        {\pd f \over \pd \theta}(x, \theta)
      \right| dx
    \end{eqnarray*}
    Finally we have
    \begin{eqnarray*}
      {\pd \E h(X) \over \pd \theta}
      > h(x_0) \int_a^b
      {\pd f \over \pd \theta}(x, \theta) dx
      = h(x_0) {\partial \over \partial \theta}
      \int_a^b f(x, \theta) dx
      = 0
    \end{eqnarray*}
  \item If $h(\cdot)$ is increasing and $\exists x_0 \in (a, b)$ such that 
    ${\pd f \over \pd \theta}(x_0, \theta) < 0$ on $(a, x_0)$  while
    ${\pd f \over \pd \theta}(x_0, \theta) > 0$ on $(x_0, b)$, 
    by similar arguments, one can show
    \[
    {\pd \E h(X) \over \pd \theta}> 0
    \]
\end{enumerate}
\end{proof}

\section{When equity returns follow Student's t-distribution}
\setcounter{equation}{0}
It is a common practice to use Student's t-distribution to model the
stationary distribution of equity returns. So it is of interest to
find out what implications this distribution has when it is combined
with the PDA preference. Formally we assume
\[
f(x; \alpha) = c(\alpha) \left(
  1 + {x^2 \over \alpha}
\right)^{-(\alpha + 1)/2}
\]
where $\alpha > 1$ and
\[
c(\alpha) = {
  \Gamma({\alpha + 1 \over 2})
  \over
  \Gamma(\alpha/2) \sqrt{\alpha \pi}
}
\]
In the same way as for \eqref{eq:ipf}, we can write $\wt u(F, \phi)$
as
\begin{small}
  \begin{eqnarray}
    \wt u(F, \phi)
    &=&
    (1 + b)
    \int_{0}^{\infty}
    \underbrace{
      \left\{
      u(C(x)) \left[
        1 - {b \over 1 + b}\1_{\{x \ge q\}}
        \right]
      + u(C(-x))
      \right\}
    }_{U_{\text{all}}}
    f(x, \alpha) dx - b u(\delta v) F_X(q)
    \label{eq:t1}
  \end{eqnarray}
\end{small}
where $C(\cdot)$ is defined in \eqref{eq:xxie1}. As shown in lemma
\ref{lemma:III}, when $b = 0$ and $u(\cdot)$
takes the power-form of \eqref{eq:hjyr}, $U_{\text{all}}$ is monotone
depending on the values of $\xi$ and $(1 - \phi) e^r / \phi$. As given
in lemma \ref{lemma:II}, there is a point $x_0 > 0$ such that
${\pd f \over \pd \alpha}(x_0, \alpha) = 0$ and
$\forall x \in (0, x_0),
{\pd f \over  \pd \alpha}(x, \alpha) > 0$ and
$\forall x \in (x_0, \infty)$,
${\pd f \over \pd \alpha}(x, \alpha) < 0$. Thus it remains to verify
$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$ and
$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
if we are to apply lemma \ref{lemma:I}.

As computed in the proof of lemma \ref{lemma:II}, ${\pd f \over \pd
  \alpha}(x, \alpha)$ is given by \eqref{eq:xxie4.1}. It is also shown
there ${d \over d \alpha} c(\alpha) > 0$. Thus for
$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$
it suffices to show
\[
\int_ 0^\infty {
x^2
\over
(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
} dx < \infty
\]
We may write
\begin{eqnarray*}
&& \int_ 0^\infty {
x^2
\over
(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
} dx \\
&=& \left(\int_0^1 + \int_1^\infty \right) {
x^2
\over
(x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
} dx \\
&=& I_1 + I_2
\end{eqnarray*}
Clearly
\[
I_1 < \int_0^1 {1 \over \alpha} dx < \infty
\]
while
\begin{eqnarray*}
  I_2 &=& \int_1^\infty {1 \over 1 + \alpha/x^2}
  {1 \over (1/x^2 + 1/\alpha)^{(\alpha+1) / 2}} {dx \over x^{\alpha +
      1}} \\
  &<& \int_1^\infty \alpha^{(\alpha+1) / 2} {dx \over x^{\alpha + 1}}
  < \infty
\end{eqnarray*}
So we conclude
$\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
To see
$\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$, we note
\begin{eqnarray*}
  \xi |U_{\text{all}}(x)|
  &<& [(1 - \phi) e^r + e^x]^{-\xi} + [(1 - \phi) e^r + e^{-x}]^{-\xi} \\
  &<& 1 + (1 - \phi)^{-\xi} e^{-r \xi}
\end{eqnarray*}
Since $\int_0^\infty |{\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$,
it follows from the above inequality $\int_0^\infty |U_{\text{all}}(x){\pd f \over \pd \alpha}(x, \alpha)| dx < \infty$.
Thus by lemma \ref{lemma:I}, $\wt u_{\rm max}$ is monotone
increasing/decreasing with $\alpha$ when $U_{\text{all}}(\cdot)$ is
monotone decreasing/increasing. Accordingly, an investor guided by the
utility function will seek the largest/smallest $\alpha$ observed in
the market.

If however $b > 0$, Lemma \ref{lemma:I} is not applicable
anymore. Nonetheless, numerical analysis lends some insight.
As shown in figure \ref{fig:htfg}, $\hat\phi$ is monotone increasing
for all 4 values of $b$, while $\tilde u_{\rm max}(\alpha)$ is
increasing with $\alpha$ when $b$ is relatively large, but
decreasing with $\alpha$ when $b$ is small. We note that a sizable
value of $b$ indicates a conservative, risk-averse investor.

\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_b_t_power4.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{U_b_t_power4.pdf}
  \end{minipage}
  \caption{$\hat\phi$ (left) and ${\pd \tilde u_{\rm max} \over \pd
      \alpha}$ (right). {\em top:} $\xi = 1/2$. {\em bottom:} $\xi =
    4$.
  }
  \label{fig:htfg}
\end{figure}
\begin{lemma}
  \label{lemma:II}
  Let $f$ denotes the density function of the Student's
  t-distribution, i.e.
  \[
  f(x; \alpha) = c(\alpha) \left(
    1 + {x^2 \over \alpha}
  \right)^{-(\alpha + 1)/2}
  \]
  where $\alpha > 1$ and
  \[
  c(\alpha) = {
    \Gamma({\alpha + 1 \over 2})
    \over
    \Gamma(\alpha/2) \sqrt{\alpha \pi}
  }
  \]
  Then there exists $x_0 > 0$ such that ${\pd f \over
    \pd \alpha}(x_0, \alpha) = 0$ and $\forall x \in (0, x_0), {\pd f \over
    \pd \alpha}(x, \alpha) > 0$ and $\forall x \in (x_0, \infty), {\pd f
    \over \pd \alpha}(x, \alpha) < 0$.
\end{lemma}
\begin{proof}
  Straightforward computation gives
  \begin{eqnarray}
    {\pd f(x, \alpha) \over \pd \alpha} &=& {
      c(\alpha) x^2 (\alpha + 1) + (2 \alpha x^2 + 2 \alpha^2) c'(\alpha)
      -
      \alpha c(\alpha) (x^2 + \alpha) \log(1 + x^2/\alpha)
      \over
      2 \alpha (x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
    } \nonumber \\
    &:=& {
      P(x^2, \alpha)
      \over
      2 \alpha (x^2 + \alpha) (1 + x^2 / \alpha)^{\alpha/2 + 1/2}
    }
    \label{eq:xxie4.1}
  \end{eqnarray}

  While the denominator of the right side of ${\pd f(x, \alpha) \over \pd \alpha}$ is
  always positive, its numerator $P(x^2, \alpha)$ has a single root:
  \begin{equation}
    \label{eq:xxie5}
    x_0^2 = \alpha\exp\left\{
      W\left[
        -\left(1 + {1 \over \alpha}\right)
        e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
      \right]
      + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)}
    \right\} - \alpha
  \end{equation}
  where $W(\cdot)$ is the principle branch of the Lambert $W$
  function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
  the right side of \eqref{eq:xxie5} for positivity, we first note
  $c'(\alpha) > 0$:
  \[
  c'(\alpha) = {
    \pi \Gamma(\alpha/2 + 1/2) \left\{
      \alpha \left[ \Psi(\alpha/2 + 1/2) - \Psi(\alpha/2)\right] - 1
    \right\}
    \over
    2 \Gamma(\alpha/2) (\pi \alpha)^{3/2}
  }
  \]
  where $\Psi(\cdot)$ is the digamma function:
  \[
  \Psi(x) = {d \log[\Gamma(x)] \over d x}
  \]
  \begin{minipage}{0.48\textwidth}
    $\Psi(x)$ \linebreak
    \includegraphics[width=\textwidth]{digamma.png}
  \end{minipage}\hfill
  \begin{minipage}{0.5\textwidth}
    As shown in the figure to the left, $\Psi(x)$ is increasing
    for $x > 0$. This immediately follows from the series
    representation
    \[
    \Psi(x + 1) = -\gamma + \sum_{n=1}^\infty {x \over n(n + x)}
    \quad x \neq -1, -2, -3,\dots
    \]
    which gives
    \[
    {\pd \Psi(x + 1) \over \pd x}
    = \sum_{n=1}^\infty {1 \over (n + x)^2} > 0
    \]
    See Abramowitz and Stegun \cite{abramowitz1972handbook}, p.259,
    formula 6.3.16. Therefore $\Psi(\alpha/2 + 1/2) - \Psi(\alpha/2) > 0$.
    So we have
    \begin{eqnarray*}
      && \alpha \left[
        \Psi(\alpha/2 + 1/2) - \Psi(\alpha/2)
      \right] - 1 \\
      &\geq& 1 \times \left[
        \Psi(1/2 + 1/2) - \Psi(1/2)
      \right] - 1 \\
      &=& \log(4) - \log(e) \\
      &>& 0
    \end{eqnarray*}
    Thus $c'(\alpha) > 0$. Furthermore, we recall
    $W(\cdot)$ is increasing on its principle branch. So
  \end{minipage}
  \begin{eqnarray*}
    &&
    W\left[
      -\left( 1 + {1 \over \alpha} \right)
      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
    \right]
    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \\
    &>& 
    W\left[
      -\left( 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \right)
      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
    \right]
    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} \\
    &=& W(-y e^{-y}) + y
  \end{eqnarray*}
  where
  \[
  y = 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)} > 1
  \]
  Now notice
  \[
  \log(y e^{-y}) = \log(y) - y
  \]
  is a decreasing function for $y > 1$. Thus $-y e^{-y}$ is an
  increasing function. Hence we have
  \begin{eqnarray*}
    W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
  \end{eqnarray*}
  Now it is clear
  \[
  \alpha\exp\left\{
    W\left[
      -\left(1 + {1 \over \alpha}\right)
      e^{-1 - 2 c'(\alpha)/c(\alpha) - 1/\alpha}
    \right]
    + 1 + {1 \over \alpha} + {2 c'(\alpha) \over c(\alpha)}
  \right\} - \alpha > 0
  \]
  Now that we have established that ${\pd f \over \pd \alpha}(x,
  \alpha) = 0$ has a single positive root, it remains to determine the
  sign of ${\pd f \over \pd \alpha}(x, \alpha)$ on the two sides of
  the root. For this purpose we observe
  \begin{equation}
    \label{eq:xxie5.1}
    P(0, \alpha) = 2 \alpha^2 c'(\alpha) > 0
  \end{equation}
  So we want to investigate ${\pd P \over \pd x}(x, \alpha)$:
  \begin{eqnarray}
    {\pd P \over \pd x}(x, \alpha) &=&
    2 \alpha c'(\alpha) + c(\alpha) - \alpha c(\alpha) \log\left(
      1 + {x \over \alpha}
    \right)
    \label{eq:xxie6}
  \end{eqnarray}
  Clearly, $\frac{\pd P}{\pd x}(0, \alpha) > 0$. Hence from
  \eqref{eq:xxie5.1} and
  \eqref{eq:xxie6} it is clear
  \begin{equation*}
    \text{sign}\left[
      {\pd f \over \pd\alpha}(x, \alpha)
    \right]
    = \left\{
    \begin{array}{rl}
      1 & 0 < x < x_0 \\
      -1 & x > x_0
    \end{array}
    \right.
  \end{equation*}
  where $x_0$ is the positive root of \eqref{eq:xxie5}.
\end{proof}

\section{Proof of Lemma~\ref{thrm:I}}
\setcounter{equation}{0}
\label{sec:thrmI_proof}
\begin{proof}
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \underset{0 < \phi \leq 1}{\text{argmax}}\,
    \wt u(F_X, \phi)
  \end{equation}
  We have
  \[
  \tilde u_{\rm max}(F_X) = \wt u(F_X, \hat\phi)
  \]
  It follows
  \begin{eqnarray}
    {d \tilde u_{\rm max}(F_X) \over d \alpha}
    &=&
    \left.{\pd \wt u(\alpha, \phi) \over \pd \alpha}\right|_{\phi = \hat\phi}
    + \left.{\pd \wt u(\alpha, \phi) \over \pd \phi}\right|_{\phi = \hat\phi}
    {\pd \hat\phi \over \pd \alpha}
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies for all $\alpha$
  \begin{equation}
    \label{eq:xxie4}
    \left.{\pd \wt u(\alpha, \phi) \over \pd \phi}\right|_{\phi = \hat\phi} = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  the first term is positive. From \eqref{eq:xxie1.0}, it follows
  \begin{eqnarray*}
    {\pd \wt u(\alpha, \phi) \over \pd \alpha}
    = {\partial \over \partial \alpha}\E[u((1 - \phi) e^r + \phi e^x) \1_{\{X < 0\}}]
  \end{eqnarray*}
  The function $u((1 - \phi) e^r + \phi e^x)$ is obviously
  increasing with $x$. It follows
  \begin{eqnarray*}
    {\pd f_X(x; \alpha, K) \over \pd \alpha}
    = {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} 
    = - {K^\alpha \over (K - x)^{\alpha + 1}}
    \left[
      \alpha
      \log\left(
        1 - {x \over K}
      \right) - 1
    \right]
  \end{eqnarray*}
  It is easily checked
  \[
  {\partial \over \partial \alpha}
  {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
  \left\{
    \begin{array}{ll}
      > 0 & \text{ when } x < K(1 - e^{1/\alpha}) < 0 \\
      < 0 & \text{ when } K(1 - e^{1/\alpha}) < x < 0
    \end{array}
  \right.
  \]
% that when $x < K(1 - e^{1/\alpha}) < 0$,
%   \[
%   {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
%   \]
%   and when $K(1 - e^{1/\alpha}) < x < 0$,
%   \[
%   {\partial \over \partial \alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
%   \]
  This is the second case of lemma \ref{lemma:I}. So we have
  ${\pd \wt u(\alpha, \phi) \over \pd \alpha} > 0$.
  As for ${\pd \tilde u_{\rm max}(F_X) \over \pd K}$, by the same
  argument, it suffices to show ${\pd \wt u(K, \phi) \over \pd K} < 0$.
  We have
  \[
  {\pd \wt u(K, \phi) \over \pd K}
  = {\partial \over \partial K} \E \left[
    u((1 - \phi) e^r + \phi e^x) \1_{\{X < 0\}}
  \right] 
  \]
  and
  \begin{eqnarray*}
    {\pd f_X(x; \alpha, K) \over \pd K}
    = {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
   = -\alpha K^{\alpha - 1} {
      \alpha x + K
      \over
      (K - x)^{\alpha + 2}
    }
  \end{eqnarray*}
  Clearly,
  \[
  {\partial \over \partial K}
  {\alpha K^\alpha \over (K - x)^{\alpha + 1}}
  \left\{
    \begin{array}{ll}
      > 0 & \text{ when } x < -K/\alpha < 0\\
      < 0 & \text{ when } -K/\alpha < x < 0
    \end{array}
  \right.
  \]
  % Clearly, when $x < -K/\alpha$
  % \[
  % {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  % \]
  % and when $-K/\alpha < x < 0$
  % \[
  % {\partial \over \partial K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  % \]
  So by the 1st case of Remark \ref{remark:I} we conclude $  {\pd \wt u(K, \phi) \over \pd K} < 0$.
\end{proof}

\section{Lemma~\ref{lemma:III}}
\setcounter{equation}{0}
\begin{lemma}\label{lemma:III}
Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:hjyr} and
\eqref{eq:xxie1} respectively. Define
\begin{eqnarray*}
U_{\text{all}} &=& u(C(x))
+ u(C(-x))
\quad
x \geq 0 \\
a &=& {
  (1 - \phi) e^r
  \over
  \phi
} \\
y_\pm &=& {
  a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
  \over
  a (\xi - 1)
}
\end{eqnarray*}
The following holds true:
\begin{enumerate}
\item If $\max\{a, 1\} < \xi$, $U_{\text{all}}(\cdot)$ is
  monotone decreasing.
\item If $a < \xi < 1$ and $(a + y_-)/(a y_- + 1) <
  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
  decreasing.
\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotone
  increasing.
\item If $1 < \xi < a$ and $(a + y_+)/(a y_+ + 1) >
  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotone
  increasing.
\item In other case, $U_{\text{all}}(\cdot)$ is not monotone.
\end{enumerate}
\end{lemma}
\begin{proof}
  It is convenient to re-write $U_{\text{all}}$ as
  \[
  U_{\text{all}} =
  -{\phi^{-\xi} \over \xi} \left\{
    \underbrace{
      \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
      +
      \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
    }_{U(x)}
  \right\}
  \]
  Thus the monotonicity of $U_{\text{all}}(x)$ is the opposite of $U(x)$. For
  convenience of writing, let $a = (1 - \phi)e^r/\phi$.
  First of all, we find the conditions on which $U(x)$ is monotone.
  Direct differentiation yields
  \begin{eqnarray*}
    {\pd U(x) \over \pd x}
    &=&
    -{
      (a + e^x)^{-\xi - 1} \xi e^x
    } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}
  \end{eqnarray*}
  ${\pd U(x) \over \pd x} = 0$ is equivalent to
  \begin{eqnarray*}
    {a + y \over a y + 1} &=& y^{1 -\xi \over 1+\xi} \\
    \underbrace{\log(a + y) - \log(a y + 1)}_{f(y)} &=&
    {1 -\xi \over 1+\xi} \log(y)
  \end{eqnarray*}
  where we have defined $y = e^x$, and $f(y)$, $g(y)$ as above. Observe
  \begin{equation}
    \label{eq:f-g-der}
    {\pd (f -g)(y) \over \pd y} = {
      a (\xi - 1) y^2 + 2 (\xi - a^2) y + a(\xi - 1)
      \over
      (1 + \xi) (a + y) (a y + 1) y
    }
  \end{equation}
  ${\pd (f -g)(y) \over \pd y} = 0$ has two roots when $\xi \ne 1$
  \begin{eqnarray*}
    y_\pm &=& {
      a^2 - \xi \pm \sqrt{(a^2 - 1) (a^2 - \xi^2)}
      \over
      a (\xi - 1)
    }
  \end{eqnarray*}
  \begin{enumerate}
  \item If $\min\{1, \xi\} < a < \max\{1, \xi\}$, $y_\pm$ are not
    real, ${\pd (f -g)(y) \over \pd y} = 0$ has no solution on
    $(1, \infty)$; $(f -g)(y)$ is monotone on $(1, \infty)$.
    \begin{enumerate}
    \item If in addition $\xi > 1$, i.e. $1 < a < \xi$, $(f - g)(y)$
      is monotone increasing on $(1, \infty)$, because the
      coefficient of the $y^2$ term of the numerator of
      \eqref{eq:f-g-der}, i.e. $a (\xi - 1)$ is positive. We note
      $f(1) = g(1) = 0$; thus on $(1, \infty)$, there is no solution
      to $f(y) = g(y)$. It can be concluded that $U(\cdot)$ is
      monotone on $(0, \infty)$. Furthermore
        \begin{eqnarray*}
          \left.{\pd U(x) \over \pd x}\right|_{x=0} &=& 0 \\
        \end{eqnarray*}
        For a small $\epsilon > 0$, the sign of ${\pd U(x) \over \pd
          x}$ on $(0, \epsilon)$ is thus the same as
        $\left.{\pd^2 U(x) \over \pd x^2}\right|_{x=0}$:
        \[
        \left.{\pd^2 U(x) \over \pd x^2}\right|_{x=0} =
        2 (a + 1)^{-\xi - 2} (\xi - a) > 0
        \]
        Thus ${\pd U(x) \over \pd x} > 0$ for $x \in (0, \infty)$;
        $U_{\text{all}}$ is monotone decreasing.
    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar
      argument as in the previous case, $U_{\text{all}}$ is monotone
      increasing on $(0, \infty)$.
    \end{enumerate}
  \item If $a < \min\{1, \xi\}$ and $\xi > 1$, i.e. $a < 1 < \xi$, it
    is clear
    \[
    y_- = {
      a^2 - \xi - \sqrt{(a^2 - 1) (a^2 - \xi^2)}
      \over
      a (\xi - 1)
    } < 0
    \]
    It remains to compare $y_+$ with $1$ to determine whether
    ${\pd (f -g)(y) \over \pd y} = 0$ has a solution on $(1, \infty)$.
    Assume $y_+ > 1$. Then
    \begin{eqnarray}
      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1)
      \nonumber \\
      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0 \label{eq:mi6}
    \end{eqnarray}
    This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotone increasing
    on $(1, \infty)$. Following the same analysis as in the case
    (1.a), one can see $U_{\text{all}}$ is monotone decreasing.

  \item If $a < \min\{1, \xi\}$ and $\xi < 1$, i.e. $a < \xi < 1$, it
    is clear $y_- > 0$ and $y_+ < y_-$. Moreover, $y_- > 1$ is equivalent
    to
    \[
    \sqrt{(a^2 - 1) (a^2 - \xi^2)} > a^2 - \xi a + a - \xi
    = (a + 1) (a - \xi)
    \]
    The last inequality is obviously true in this case. So $y_- >
    1$. ${\pd (f-g)(y) \over \pd y}$ has a maximum at
    $y_- \in (1, \infty)$. We know $(f - g)(y_-)$ is a maximum because
    \eqref{eq:f-g-der} shows that, for $y > y_-$,
    ${\pd (f-g)(y) \over \pd y} < 0$.

    If $(f-g)(y_-) < 0$, $(f-g)(y) = 0$ has no solution on $(1,
    \infty)$. Hence $U(\cdot)$ is monotone on both $(0, \infty)$.
    The same analysis as in the case (1.a) shows $U_{\text{all}}$
    is monotone decreasing on $(0, \infty)$.

    If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
    \infty)$. So $U(\cdot)$ is not monotone.

    \item $a > \max\{1, \xi\}$ and $\xi > 1$, i.e. $1 < \xi < a$. By
      the same argument that leads to \eqref{eq:mi6}, we see
      $y_+ > 1$. From \eqref{eq:f-g-der} it is clear $(f-g)(y_+)$ is a
      minimum. If $(f-g)(y_+) > 0$, there is no solution to
      $(f-g)(y) = 0$ on $(1, \infty)$; $U(x)$ is monotone. By the
      same analysis as in the case (1.a), we know $U(x)$ is
      monotone decreasing and $U_{\text{all}}$ is monotone increasing.

      if $(f-g)(y_+) < 0$, there must be a solution to $(f-g)(y) = 0$
      on $(y_+, \infty)$. $U_{\text{all}}$ is not monotone.
  \end{enumerate}
\end{proof}

\ifx\phdthesis\undefined
\else
\end{subappendices}
\fi

### 2017-07-21 papers/DavisMikoschHeinyXie/PaperBody.tex
Added some paragraph in the introduction.

@@
 \par
-In Section~\ref{sec:2} we continue the study of the case when the data matrix $\X$ consists of iid heavy-tailed entries.
+In Section~\ref{sec:Davis2} we continue the study of the case when the data matrix $\X$ consists of iid heavy-tailed entries.
 We will consider power-law growth rates on the dimension $(p_n)$ that is more general than prescribed by \eqref{eq:gamma}.
@@
 
-\section{General growth rates for $p_n$ in the iid heavy-tailed case}\label{sec:2}\setcounter{equation}{0}
+\section{General growth rates for $p_n$ in the iid heavy-tailed case}
+\label{sec:Davis2}\setcounter{equation}{0}
 This section is based on ideas in Heiny and Mikosch \cite{heiny:mikosch:2015:iid} where one can also find detailed proofs.
@@
 Then
-\beam\label{eq:2}
+\beam\label{eq:Davis2}
 \sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
@@
 
-%In  Davis et al. \cite{davis:mikosch:pfaffel;2015} relation \eqref{eq:2} was derived for $\alpha\in (0,1)$ even for \seq s
+%In  Davis et al. \cite{davis:mikosch:pfaffel;2015} relation \eqref{eq:Davis2} was derived for $\alpha\in (0,1)$ even for \seq s
 %$p_n\to\infty$ \st\ $p_n= O(\ex^{s_n/n})$, where $s_n=o(n)$ but with stronger limitations on the growth of $(p_n)$ for $[1,4]$.
@@
 between the rows and columns of $\X$:
-\begin{equation}\label{eq:1}
+\begin{equation}\label{eq:Davis1}
 X_{it}=\sum_{l\in \Z}\sum_{k\in \Z} h_{kl} Z_{i-k,t-l}\,,\qquad i,t\in\Z\,,
@@
 determine those ingredients in the sample covariance matrix which contribute to its largest eigenvalues.
-If the series in \eqref{eq:1} converges a.s. $(X_{it})$ constitutes a strictly stationary random field.
+If the series in \eqref{eq:Davis1} converges a.s. $(X_{it})$ constitutes a strictly stationary random field.
 We denote generic elements of the $Z$- and $X$-fields by $Z$ and $X$, respectively. We assume that $Z$ is \regvary\ in the sense that
@@
 for some $\delta\in (0,\min({\alpha/2},1))$ which ensures
-the a.s.~absolute convergence of the series in \eqref{eq:1}. Under the conditions \eqref{eq:27} and \eqref{eq:2a}, the marginal and
+the a.s.~absolute convergence of the series in \eqref{eq:Davis1}. Under the conditions \eqref{eq:27} and \eqref{eq:2a}, the marginal and
 \fidi s of the field $(X_{it})$ are \regvary\ with index $\alpha$; see
@@
 \par
-The model \eqref{eq:1} was introduced by Davis et al. \cite{davis:pfaffel:stelzer:2014}, assuming the rows iid, and in the
+The model \eqref{eq:Davis1} was introduced by Davis et al. \cite{davis:pfaffel:stelzer:2014}, assuming the rows iid, and in the
 present form by Davis et al. \cite{davis:mikosch:pfaffel:2015}.
@@
 The idea of using the sample autocovariance matrices and \fct s of their squares \eqref{eq:squares}
-originates from a paper by Lam and Yao~\cite{lam:yao} who used a model different from \eqref{eq:1}.
+originates from a paper by Lam and Yao~\cite{lam:yao} who used a model different from \eqref{eq:Davis1}.
 This idea is quite natural in the context of \tsa .
@@
 In the following result we povide some useful approximations to the singular values of the sample autocovariance matrices of
-the linear model \eqref{eq:1}.
+the linear model \eqref{eq:Davis1}.
 \begin{theorem}\label{thm:mains}
-Consider the linear process \eqref{eq:1} under
+Consider the linear process \eqref{eq:Davis1} under
 \begin{itemize}
@@
 or $\alpha \in [2,4)$, $\E[Z^2]<\infty$ and \ref{Cbeta} hold then
-\begin{equation}\label{eq:pp}
+\begin{equation}\label{eq:Davispp}
 N_n^{\lambda,s} \cid N= \sum_{i=1}^\infty
@@
 Hence $\M=\H\H'$ has the positive eigenvalues
-$v_1=8$ and $v_2=2$. The limit point process in \eqref{eq:pp} is
+$v_1=8$ and $v_2=2$. The limit point process in \eqref{eq:Davispp} is
 \begin{equation*}
@@
 and hence the limit distribution of $\lambda_{(2)}/\lambda_{(1)}$ is supported on $[1/4,1)$ with  mass of $2^{-\alpha}$ at $1/4$.
-The histogram of the ratio $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ based on 1000 replications from the model \eqref{ex:4.6} with noise given by a $t$-distribution with $\alpha=1.5$ degrees of freedom, $n=1000$ and $p=200$ is displayed in Figure \ref{fig:1}.
+The histogram of the ratio $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ based on 1000 replications from the model \eqref{ex:4.6} with noise given by a $t$-distribution with $\alpha=1.5$ degrees of freedom, $n=1000$ and $p=200$ is displayed in Figure \ref{fig:Davis1}.
 \begin{figure} [h]
@@
 \caption{Histogram based on 1000 replications of $\left(\lambda_{(2)}/\lambda_{(1)}\right)^{2/\alpha}$ from model \eqref{ex:4.6}.}
-\label{fig:1}
+\label{fig:Davis1}
 \end{figure} Observing that $2^{-\alpha}=0.3536\ldots$,
@@
 The following result is a corollary of Theorem~\ref{thm:mains}.
-\begin{proposition}\label{thm:mainstr} Consider the linear process \eqref{eq:1} under
+\begin{proposition}\label{thm:mainstr} Consider the linear process \eqref{eq:Davis1} under
 the conditions of Theorem~\ref{thm:mains}. Then the following statements hold for $s\ge 0$:

### 2017-07-21 phd/Chapterintroduction.tex
Added some paragraph in the introduction.

@@
 observed in a variety of phenomena, for example, phase transition of
-matter, black body radiation, neuronal avalanches and stock returns.
-When expanded to multiple dimensions, the scaling property is better
-described in terms of weak convergence to a spectral measure: If a
-random vector $V$ satisfies
+matter and black body radiation as studied in physics, neuronal
+avalanches in biology, claim sizes of insurance mathematics and stock
+returns in finance. The last application is indeed the focus of this
+thesis.
+
+When expanded to multiple dimensions, the scaling property of regular
+variation is better described in terms of weak convergence to a
+spectral measure: If a random vector $V$ satisfies
 \[
@@
 
+When considered as the multiplicative inverse of the parameter of the
+{\em Generalized Extreme Value} distribution, there are other methods
+in the literature for estimating the tail index, e.g. Pickand's 
+estimator \cite{pickands1975statistical} $\hat \alpha_P$ and the
+Deckers-Einmahl-de Haan estimator $\hat \alpha_{\text{DEH}}$
+\begin{eqnarray*}
+{1 \over \hat \alpha_P}
+&=&
+{1 \over \log 2}
+\log {
+  X_{(k)} - X_{(2k)}
+  \over
+X_{(2k)} - X_{(4k)}  
+} \\
+{1 \over \hat \alpha_{\text{DEH}}}
+&=&
+1 + {1 \over \alpha_H} + {1 \over 2} \left(
+  {H \over \alpha_H^2} - 1
+\right)^{-1}
+\end{eqnarray*}
+where it is similarly assumed $k \to \infty$ and $k/n \to 0$ as $n \to
+\infty$. The Deckers-Einmahl-de Haan estimator makes use of Hill's
+estimator $\hat \alpha_H$ and computes
+\[
+H = \left[
+  {1 \over k} \sum_{i=1}^k \log \left(
+    { X_{(i)} \over X_{(k+1)} }
+  \right)^2
+\right]^{-1}
+\]
+Apparently $1/H$ can be interpreted as the 2nd empirical moment of
+$\log(X_{(i)}/X_{(k+1)})$ for $i \leq k$.
+A major drawback of Pickand's and Deckers-Einmahl-de Haan's estimators
+is that, when applied to estimating the tail index, they discard the
+information that the tail index is always positive, hence resulting in
+a larger confidence band compared with that obtained for Hill's
+estimator. Therefore we stick to Hill's estimator in the empirical
+work included in this thesis.
+
+One of the most important dynamical mechanisms that lead to regularly
+varying r.v. is stochastic recursion of the following form:
+\[
+X_t = A_t X_{t-1} + B_t
+\]
+where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
+random matrix and $B_t$ is a $d$-dimensional vector, random or
+deterministic.
+
+\section{Stochastic Recurrence Equation}
+
 \section{Contribution of this thesis}\label{sec:contr}

### 2017-07-24 phd/Chapterintroduction.tex
Revised the introduction section of the thesis. Started programming for the main importance sampling algorithm.

@@
 
+\section{Stochastic Recurrence Equation}
 One of the most important dynamical mechanisms that lead to regularly
@@
 random matrix and $B_t$ is a $d$-dimensional vector, random or
-deterministic.
+deterministic. The sequence $\{A_i\}_{i=1,2,\dots}$ and
+$\{B_i\}_{i=1,2,\dots}$  are iid and independent of each other.
 
-\section{Stochastic Recurrence Equation}
+Kesten \cite{kesten:1973} showed that, when $A_t$ and $B_t$ were almost
+surely non-negative, had no row or column of only zeros, and there was
+a positive probability that $B_t$ was strictly positive, the strictly
+stationary solution to the equation
+$V \overset{d}{=} A V + B$ had power-law tails
+for its marginal distributions, assuming the following conditions (M)
+and (A):
+\begin{itemize}
+\item Condition (M)
+  \begin{enumerate}
+  \item The top Lyapunov exponent
+    \[
+    \gamma = \inf_{n \geq 1} {1 \over n}\E \log \|A_n \cdots A_1\|
+    \]
+    is negative.
+  \item There exists $\xi > 0$ such that
+    \[
+    1 = \lambda(\xi) = \lim_{n \to \infty} {1 \over n} \log \E \|A_n \cdots A_1\|^\xi
+    \]
+  \item $\E (\|A_1\|^\xi \log^+\|A_1\|) < \infty$
+  \item $\E |B_1|^\xi < \infty$
+  \end{enumerate}
+\item Condition (A) : The group generated by
+  \[
+  \{\log\rho(s): s = A_n \cdots A_1 \text{ for some } n \geq 1\}
+  \]
+  is dense in $\reals$, where $\rho(s)$ denotes the spectral
+  radius of matrix $s$.
+\end{itemize}
+Upon these conditions, Kesten's theorem gives
+\begin{equation*}
+  u^\xi \P(u^{-1} V \in \cdot) \overset{v}{\to} \mu_\xi(\cdot)
+\end{equation*}
+where $\mu_\xi$ is a non-null Radon measure on
+$\reals^d_+ \setminus \{0\}$ with the property
+$\mu_\xi(a A) = a^{-\xi} \mu_\xi(A)$.
+
+Recently, Collamore and Mentemeier \cite{collamore:mentemeier:2016}
+extended Kesten's result and gave an explicit expression for $\mu_\xi$:
+\begin{equation}
+  \label{eq:CollamoreMentemeierIntro}
+  \lim_{u \to \infty} u^{\xi} \E \left[
+    f(u^{-1} V)
+    \right]
+  =
+  {C \over \lambda'(\xi)}  
+  \int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
+\end{equation}
+where $C$ is a constant (cf. eq.(2.15) of Collamore and Mentemeier
+\cite{collamore:mentemeier:2016}), $f(\cdot)$ is any bounded
+continuous function on $\reals^{d}_+ \setminus \{0\}$  and
+$\ell_\xi$ is a probability measure on $\sphere_+^{d-1}$. Its definition
+is also found in \eqref{eq:eigenmeasure}.
+
+From \eqref{eq:CollamoreMentemeierIntro} a representation for $\mu_\xi
+(\cdot)$ immediately follows  
+\[
+\mu_\xi (\cdot) = {C \over \lambda'(\xi)} \mathcal L_\xi(\cdot)
+\]
+Here $\mathcal L_\xi$ is a non-null Radon measure  on
+$\reals^d_+ \setminus \{0\}$ that satisfies, for all
+bounded continuous function $f(\cdot)$ on
+$\reals_+^{d} \setminus \{0\}$:
+\[
+\int_{\reals_+^d\setminus \{0\}} f(x) \mathcal L_\xi(dx)
+=
+\int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
+\]
 

### 2017-07-24 phd/MyThesis.tex
Revised the introduction section of the thesis. Started programming for the main importance sampling algorithm.

@@
 \usepackage[numbers,sort&compress]{natbib}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+
+\newtheorem{case}{case}
 
@@
 & \\
-Assessment committee: &  Prof.~Catalin Starica\\
+Assessment committee: &  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
 & University of Neuchatel, CH\\
@@
 \input{ChapterTailParameters}
-%% \input{TailParameters}
+\input{ChapterImportanceSampling}
 \input{Chapterextremes}

### 2017-07-24 cpp/ImportanceSampling/garch21.hpp

@@ -1,58 +1,24 @@
 #ifndef XXIE_GARCH21_H
 #define XXIE_GARCH21_H
-#include <array>
-#include <queue>
-#include <random>
 
-#define URNG random_device
+#include "GarchPQ.hpp"
 
-using namespace std;
-
-class garch21
+class Garch21 : public GarchPQ
 {
-protected:
-    double tail_index;
-//    random_device dev;
 public:
-    class F_Set
-    {
-    public:
-	garch21 *markov;
-	double b;
-	double rho;
-	array<double, 2> x_interval;
-	array<double, 2> eta_interval;
-	F_Set(garch21 *markov);
-	bool includes(const array<double, 2> &arg) const;
-    };
+    Garch21(const vector<double> &alpha, const vector<double> &beta,
+	    double tail_index_sup = 5.0);
+    const double tail_index;
+    double r(double) const;
+    double draw_z(bool shift) const;
 
-    class nu_dist
-    {
-    public:
-	/* The nu measure of the F set */
-	double delta;
-	double c1, c2, c3;
-	garch21 *markov;
-	array<double, 2> draw(URNG& dev);
-	double density(const array<double, 2> &arg) const;
+protected:
+    vector<double> pool;
+    typedef array<double, 2> funval;
+    vector<funval> eigenfunction;
+    void right_eigenfunction(unsigned n);
 
-	double eta_draw(URNG& dev);
-	double eta_marginal_density(double eta);
-	double eta_proposal_draw(URNG& dev);
-	double eta_proposal_density(double eta);
-	nu_dist(garch21 *markov);
-    };
-
-    double a0, a1, a2, b1;
-    array<double, 2> C;
-    F_Set F;
-    nu_dist nu;
-    garch21(array<double, 4> &params);
-    inline bool C_includes(array<double, 2> arg) ;
-    double kernel_density(const array<double, 2> &arg, double x0) const;
-    array<double, 2> forward(URNG& dev, double x0, bool orig = true);
-    array<double, 2> simulate_path(void);
-    double compute_tail_index(size_t beg_line, size_t end_line);
+    double q_shifted
 };
 
 #endif

### 2017-07-24 cpp/ImportanceSampling/right_eigenfunction.cpp

@@ -81,6 +81,7 @@ double right_eigenfunction(const struct garch21_param &param,
     return *p;
 }
 
+#ifndef MAIN_ALGO
 int main(int argc, char **argv)
 {
     // struct garch21_param param = {
@@ -125,3 +126,4 @@ int main(int argc, char **argv)
     
     return 0;
 }
+#endif

### 2017-07-31 papers/Anja/PaperBody.tex
Added the paper with Anja to the thesis.

\begin{abstract}
We consider a multivariate heavy-tailed stochastic volatility model
and analyze the large-sample behavior of its sample covariance
matrix. We study the limiting behavior of its entries in the
infinite-variance case and derive results for the ordered eigenvalues
and corresponding eigenvectors. Essentially, we consider two different
cases where the tail behavior either stems from the iid\ innovations
of the process or from its volatility sequence. In both cases, we make
use of a large deviations technique for regularly varying time series
to derive multivariate $\alpha$-stable limit distributions of the
sample covariance matrix. While we show that in the case of
heavy-tailed innovations the limiting behavior resembles that of
completely independent observations, we also derive that in the case
of a heavy-tailed volatility sequence the possible limiting behavior
is more diverse, i.e.\ allowing for dependencies in the limiting
distributions which are determined by the structure of the underlying
volatility sequence.
\end{abstract}
% \keywords{Regular variation, sample covariance matrix, dependent entries,
% largest eigenvalues, eigenvectors, stochastic volatility}

% \subjclass{Primary 60B20; Secondary 60F05 60G10 60G70 62M10}

% \maketitle

\section{Introduction}\label{sec:intro}
\subsection{Background and Motivation}\label{subsec:motivation}

The study of sample covariance matrices is fundamental for the
analysis of dependence in multivariate time series. Besides from
providing estimators for variances and covariances of the observations
(in case of their existence), the sample covariance matrices are a
starting point for dimension reduction methods like principal
component analysis. Accordingly, the special structure of sample
covariance matrices and their largest eigenvalues has been intensively
studied in random matrix theory, starting with iid\ Gaussian
observations and more recently extending results to arbitrary
distributions which satisfy some moment assumptions like in the four
moment theorem of Tao and Vu \cite{tao:vu:2012}.

However, with respect to the analysis of financial time series, such a
moment assumption is often not suitable. Instead, in this work, we
will analyze the large sample behavior of sample covariance matrices
under the assumption that the marginal distributions of our
observations are regularly varying with index $\alpha<4$ which implies
that fourth moments do not exist. In this case, we would expect the
largest eigenvalues of the sample covariance matrix to inherit
heavy-tailed behavior as well; see for example Ben Arous and Guionnet
\cite{benarous:guionnet:2007}, 
Auffinger et al. \cite{auffinger:arous:peche:2009}, Soshnikov
\cite{soshnikov:2004,soshnikov:2006}, Davis et
al. \cite{davis:mikosch:heiny:xie:2015}, Heiny and Mikosch
\cite{heiny:mikosch:2016} for the case of iid entries. Furthermore, in
the context of financial time series we have to allow for dependencies
both over time and between different components and indeed it is the
very aim of the analysis to discover and test for these dependencies
from the resulting sample covariance matrix as has for example been done in  
Plerou et al. \cite{plerou:et:al:2002} and Davis et
al. \cite{davis:pfaffel:stelzer:2014,davis:mikosch:pfaffel:2016}. 
The detection of dependencies among assets also plays a crucial role
in portfolio optimization based on multi-factor prizing models, where
principal component analysis is one way to derive the main driving
factors of a portfolio; cf.\ Campbell et
al. \cite{campbell:lo:macKinlay:1997} and recent work by Lam and Yao
\cite{lam:yao:2012}.
\par
The literature on the asymptotic behavior of sample covariance
matrices derived from dependent heavy-tailed data is, however,
relatively sparse up till now. Starting with the analysis of the
sample autocorrelation of univariate linear heavy-tailed time series
in Davis and Resnick \cite{davis:resnick:1985, davis:resnick:1986},  
the theory has recently been extended to multivariate heavy-tailed time series with linear structure in 
Davis et al. \cite{davis:pfaffel:stelzer:2014,davis:mikosch:pfaffel:2016}, cf.\ also the recent 
survey article by Davis et
al. \cite{davis:mikosch:heiny:xie:2015}. But most of the standard
models for financial time series 
such as GARCH and stochastic volatility models
are non-linear. In this paper we will therefore focus on a class of
multivariate stochastic volatility models of the form
\beam\label{Eq:intro:sv}
X_{it}=\sigma_{it}\,Z_{it}\,,\qquad t\in\bbz\,, \quad 1\leq i \leq p,
\eeam
where $(Z_{it})$ is an iid random field independent of a strictly
stationary ergodic field $(\sigma_{it})$ of non-negative \rv s;
see Section \ref{sec:model:anja} for further details.  Stochastic
volatility models have been studied in detail in the financial \ts\
literature; see for example Andersen et
al. \cite{andersen:davis:kreiss:mikosch:2009}, Part II. They are among
the simplest models allowing for conditional heteroscedasticity of a
time series. In view of independence between the $Z$- and
$\sigma$-fields dependence conditions on $(X_{it})$ are imposed only
via the stochastic volatility $(\sigma_{it})$. Often it is assumed
that $(\log \sigma_{it})$ has a linear structure, most often Gaussian.
\par
In this paper we are interested in the case when the marginal and
\fidi s of $(X_{it})$ have power-law tails.
Due to independence between  $(\sigma_{it})$ and $(Z_{it})$ heavy
tails of $(X_{it})$ can be due to the $Z$- or the $\sigma$-field. 
Here we will consider two cases: (1) the tails of $Z$ dominate the
right tail of $\sigma$ and (2) the right tail of $\sigma$ dominates
the tail of $Z$. The third case when both $\sigma$ and $Z$ have heavy
tails and are tail-equivalent will not be considered in this
paper. Case (1) is typically more simple to handle; see Davis and Mikosch
\cite{davis:mikosch:1999,davis:mikosch:2001,davis:mikosch:2009} for
\evt , \pp\ \con\ and central limit theory 
with infinite variance stable limits. Case (2) is more subtle as
regards the tails of the \fidi s. The literature on stochastic
volatility models with a heavy-tailed volatility sequence is so far
sparse but the interest in these models has been growing recently; see
Mikosch and Rezapour \cite{mikosch:rezapour:2013}, Kulik and Soulier
\cite{kulik:soulier:2015} and Jan\ss en and Drees
\cite{janssen:drees:2016}. In particular, it has been shown that these
models offer a lot of flexibility with regard to the extremal
dependence structure of the time series, ranging from asymptotic
dependence of consecutive observations (cf.\
\cite{mikosch:rezapour:2013}) to asymptotic independence of varying
degrees (cf.\ \cite{kulik:soulier:2015} and
\cite{janssen:drees:2016}).

\subsection{Aims, main results and structure} 
After introducing the general model in Section \ref{sec:model:anja} we
first deal with the case of heavy-tailed innovations and a
light-tailed volatility sequence in Section \ref{sec:case1}. The first
step in our analysis is to describe the extremal structure of the
corresponding process by deriving its so-called tail process; see 
Section~\ref{subsec:rvts} and Proposition~\ref{prop:regvarsv}. This
allows one to apply an infinite variance stable \clt\  from Mikosch and Wintenberger 
\cite{mikosch:wintenberger:2016} (see Appendix \ref{App:A}) to derive
the joint limiting behavior of the entries of the sample covariance
matrix of this model. This leads to the main results  
in the first case: Theorems~\ref{the:1} and \ref{lem:1}. They say,
roughly speaking, that all values on the off-diagonals of the sample
covariance matrix are negligible compared to the values on the
diagonals. Furthermore, the values on the diagonal converge, under
suitable normalization, to independent $\alpha$-stable random
variables, so the limiting behavior of this class of stochastic
volatility  models is quite similar to the case of iid heavy-tailed
random variables. This fairly tractable structure allows us also to
derive explicit results about the asymptotic behavior of the ordered
eigenvalues and corresponding eigenvectors which can be
found in Sections \ref{subsec:eigen} and \ref{subsec:appl}. In
particular, we will see that in this model  
the eigenvectors are basically the unit canonical basis vectors which
describe a very weak form of extremal dependence. With a view towards
portfolio analysis, our assumptions imply that large movements of the
market are mainly driven by one single asset, where each asset is
equally likely to be this extreme driving force.

In the second case of a heavy-tailed volatility sequence combined with
light-tailed innovations, which we analyze in Section \ref{Sec:case2},
we see that the range of possible limiting behaviors of the entries of
the sample covariance matrix is more diverse and depends on the
specific structure of the underlying volatility process. We make the
common assumption that our volatility process is log-linear, where we
distinguish between two different cases for the corresponding
innovation distribution of this process. Again, for both cases, we
first derive the specific form of the corresponding tail process (see
Proposition~\ref{prop:rv:2})
which then allows us to derive the limiting behavior of the sample
covariance matrix entries, leading to the main results in the second
case: Theorems~\ref{thm:1} and \ref{thm:2}. We show that the sample
covariance matrix can feature non-negligible off-diagonal components,
therefore clearly distinguishing from the iid\ case, if we assume that
the innovations of the log-linear volatility process are convolution
equivalent. We discuss concrete examples for both model specifications
and the corresponding implications for the asymptotic behavior of
ordered eigenvalues and corresponding eigenvectors at the end of
Section \ref{Sec:case2}.

Section \ref{sec:simulation} contains a small simulation study which
illustrates our results for both cases and also includes a real-life
data example for comparison. From the foreign exchange rate data that
we use, it is notable that the corresponding sample covariance matrix
features a relatively large gap between the largest and the second
largest eigenvalue and that the eigenvector corresponding to the
largest eigenvalue is fairly spread out, i.e.,\ all its components are
of a similar order of magnitude. This implies that the model discussed
in Section~\ref{sec:case1} may not be that suitable to catch the
extremal dependence of this data, and that there is not one single
component that is most affected by extreme movements but instead all
assets are affected in a similar way. We perform simulations for three
different specifications of models from Sections~\ref{sec:case1} and
\ref{Sec:case2}. They illustrate that the models analyzed 
in Section~\ref{Sec:case2} are capable of exhibiting more diverse
asymptotic behaviors of the sample covariance matrix and in particular
non-localized dominant eigenvectors.

Some useful results for the (joint) tail and extremal behavior of
random products are gathered in Appendix~\ref{App:B}.
These results may be of independent interest when studying the
extremes of multivariate stochastic volatility models with possibly
distinct tail indices. We mention in passing that there is great
interest in non-linear models for log-returns of speculative prices
when the number of assets $p$ increases with the sample size $n$. We
understand our analysis as a first step in this direction.

\section{The model}\label{sec:model:anja}\setcounter{equation}{0}
We consider a stochastic volatility model
\beam\label{eq:1a:anja}
X_{it}=\sigma_{it}\,Z_{it}\,,\qquad i,t\in\bbz\,,
\eeam
where $(Z_{it})$ is an iid field independent of a strictly stationary
ergodic field $(\sigma_{it})$ of non-negative \rv s.
We write $Z$, $\sigma$, $X$ for generic elements of the $Z$-,
$\sigma$- and $X$-fields \st\ $\sigma$ and $Z$ are independent.
A special case appears when $\sigma>0$ is a constant: then $(X_{it})$ constitutes an iid field.

For the stochastic volatility model as in \eqref{Eq:intro:sv} we construct the 
multivariate \ts\ 
\beam\label{eq:multvts}
 \bfX_t= (X_{1t},\ldots,X_{pt})', \;\;\; t \in \mathbb{Z},
\eeam 
for a given dimension $p\ge 1$. For $n\geq 1$ we write $\bfX^n={\rm
  vec}\big((\bfX_t)_{t=1, \ldots, n}\big) \in \mathbb{R}^{p \times n}$
and consider the non-normalized sample covariance matrix
\beam\label{eq:esses}
\bfX^n(\bfX^n)'= (S_{ij})_{i,j=1,\ldots,p}\,,\qquad S_{ij}=
\sum_{t=1}^n X_{it}X_{jt}\,,\qquad S_i=S_{ii}\,.
\eeam

\subsection{Case (1): $Z$ dominates the tail}
We assume that $Z$ is \regvary\ with index $\alpha>0$, i.e.,
\beam\label{eq:10b}
\P(Z>x)\sim p_+\,\dfrac{L(x)}{x^\alpha}\qquad \mbox{and}\qquad
\P(Z<-x)\sim p_-\,\dfrac{L(x)}{x^\alpha}\,,\qquad \xto\,,
\eeam
where $p_+$ and $p_-$ are non-negative numbers with $p_++p_-=1$ and $L$ is a \slvary\ \fct . 
If we assume $\E[\sigma^{\alpha+\delta}]<\infty$ for some $\delta>0$ then, in view of a result 
by Breiman \cite{breiman:1965} (see also Lemma~\ref{lem:product}), it follows that
\beam\label{eq:br1}
\P(X>x)\sim \E[\sigma^\alpha]\,\P(Z>x)\quad \mbox{and}\quad
\P(X<-x)\sim \E[\sigma^\alpha]\,\P(Z<-x)\,,\qquad \xto\,, 
\eeam
i.e., $X$ is \regvary\ with index $\alpha$. Moreover, we know from a
result by Embrechts and Goldie \cite{embrechts:goldie:1980} 
that for independent copies $Z_1$ and $Z_2$ of $Z$, $Z_1Z_2$ is again
\regvary\ with index $\alpha$; cf. Lemma~\ref{lem:product}. Therefore,
using again 
Breiman's result under the 
condition that $\E[(\sigma_{i0}\sigma_{j0})^{\alpha+\delta}\I(i\ne
j)+\sigma_{i0}^{\alpha+\delta}]<\infty$ for some $\delta>0$, we have
\beam\label{eq:br2}
\P(\pm X_{it}\, X_{jt}>x) \sim
\left\{\begin{array}{ll}\E[(\sigma_{it}\,\sigma_{jt})^{\alpha}]\,\P(\pm
    Z_i\,Z_j>x)& i\ne j\,,\\[2mm]
 \E[\sigma^{\alpha}]\,\P(Z^2>x)& i=j\,,
\end{array}\qquad \xto\,.
\right.
\eeam

\subsection{Case (2): $\sigma$ dominates the tail}\label{subsec:Case:2}
We assume that $\sigma\ge 0$ is \regvary\ with some index $\alpha>0$: for some \slvary\ \fct\ $\ell$,
\beao
\P(\sigma>x)= x^{-\alpha}\,\ell(x)\,,
\eeao 
and $\E [|Z|^{\alpha+\delta}]<\infty$ for some $\delta>0$. Now the Breiman result yields
\beao
\P(X>x) \sim \E [Z_+^\alpha]\,\P(\sigma>x)\qquad \mbox{and} \qquad
\P(X<-x) \sim \E[Z_-^\alpha]\,\P(\sigma>x)\,,\qquad \xto\,.
\eeao
Since we are also interested in the tail behavior of the products $X_{it}X_{jt}$ we need to
be more precise about the joint \ds\ of the \seq s $(\sigma_{it})$. We assume
\beam\label{eq:sv2model}
\sigma_{it}= \exp\Big(\sum_{k,l=-\infty}^\infty \psi_{kl} \,\eta_{i-k,t-l}\Big)\,,\qquad i,t\in\bbz\,,
\eeam
where $(\psi_{kl})$ is a field of non-negative numbers (at least one
of them being positive) \st\ (without loss of generality)
$\max_{kl}\psi_{kl}=1$
and 
$(\eta_{it})$ is an iid random field  \st\ a generic element $\eta$ satisfies
\beam\label{Eq:tail:eta}
\P\big(\ex^{\eta}>x)= x^{-\alpha}\,L(x)\,, 
\eeam
for some $\alpha>0$ and a \slvary\ \fct\ $L$. 
We also assume  
$\sum_{k,l}\psi_{kl}<\infty$ to ensure absolute summability of 
$\log \sigma_{it}$. 
A distribution of $\eta$ that fits into this scheme is for example the
exponential distribution; cf.\ also Rootz\'{e}n \cite{rootzen:1986}
for further examples and extreme value theory for linear processes of
the form $\sum_{l=-\infty}^\infty \psi_{l} \,\eta_{t-l}$.

\subsection{Regularly varying \seq s}\label{subsec:rvts}
In Sections~\ref{subsec:regvar} and \ref{subsec:regvar2} 
we will elaborate on the joint tail behavior of the \seq s
$(\sigma_{it})$, $(X_{it})$, $(\sigma_{it}\sigma_{jt})$, 
and $(X_{it}X_{jt})$. We will show that, under suitable conditions,
these \seq s are \regvary\ with positive indices.

%\par
The notion of a {\em univariate \regvary\ \seq } was introduced by
Davis and Hsing \cite{davis:hsing:1995}. 
Its extension to the multivariate case does not represent
difficulties; see Davis and Mikosch \cite{davis:mikosch:2009b}.
An $\bbr^d$-valued strictly stationary \seq\ $(\bfY_t)$ is {\em
  \regvary\ with index $\gamma>0$} if each of the vectors 
$(\bfY_t)_{t=0,\ldots,h}$, $h\ge 0$, is \regvary\ with index $\gamma$,
i.e., there exist non-null Radon s $\mu_h$ on 
$[-\infty,\infty]^{d(h+1)}\backslash\{\bf0\}$ which are homogeneous of
order $-\gamma$ such that

\beam \label{eq:multvtsvague}
\dfrac{\P(x^{-1} (\bfY_t)_{t=0,\ldots,h}\in \cdot)}{\P(\|\bfY_0\|>x)} \stv \mu_h(\cdot)\,.
\eeam
Here $\stv$ denotes vague \con\ on the Borel $\sigma$-field of
$[-\infty,\infty]^{d(h+1)}\backslash\{\bf0\}$ and $\|\cdot\|$ denotes
any given norm; see Resnick's books \cite{resnick:1987,resnick:2007}
as general references to multivariate \regvar.
%\par

Following Basrak and Segers \cite{basrak:segers:2009}, 
an $\bbr^d$-valued strictly stationary \seq\ $(\bfY_t)$ is \regvary\
with index $\gamma>0$ \fif\ there exists a \seq\ of $\bbr^d$-valued
random vectors $(\bfTh_h)$ independent of a Pareto($\gamma$) \rv\ 
$Y$, i.e., $\P(Y>x)=x^{-\gamma}$, $x>1$,  \st\ for any $k\ge 0$, 
\beam\label{eq:basseg}
\P( x^{-1} (\bfY_0,\ldots,\bfY_k)\in \cdot \mid \|\bfY_0\|>x)\stw
\P\big(Y\,(\bfTh_0,\ldots,\bfTh_k)\in\cdot\big)\,,\qquad
\xto\,.
\eeam

We call $(\bfTh_h)$ the {\em spectral tail process} of $(\bfY_t)$ and
$(Y \bfTh_h)$ the {\em tail process}. We will use both defining
properties (i.e.,\ \eqref{eq:multvtsvague} and \eqref{eq:basseg}) of a
\regvary\ \seq.

\section{Case (1): $Z$ dominates the tail}\label{sec:case1}\setcounter{equation}{0}
\subsection{Regular variation of the stochastic volatility model and
  its product processes}\label{subsec:regvar}
\begin{proposition}\label{prop:regvarsv}
We assume the stochastic volatility model \eqref{eq:1a:anja} and that $Z$
is \regvary\ with index $\alpha>0$ in the sense of \eqref{eq:10b}.
\begin{enumerate}
\item
If $\E[\sigma^{\alpha+\vep}]<\infty$ for some $\vep>0$
the \seq\ $(X_{it})_{t \in\bbz}$ is \regvary\ with index $\alpha$ and
the corresponding spectral tail process $(\Theta^i_h)_{h\ge 1}$
vanishes.
\item
For any $i\ne j$, if
$\E[(\sigma_{i0}\sigma_{j0})^{\alpha+\vep}]<\infty$ for some $\vep>0$
then the \seq\ $(X_{it}X_{jt})$ is \regvary\ with index $\alpha$ and
the corresponding spectral tail process $(\Theta_h^{ij})_{h\ge 1}$ vanishes.
\end{enumerate}
\end{proposition}
\bre
If  $\E[(\sigma_{ik}\sigma_{jl})^{\alpha+\vep_{ik,jl}}]<\infty$ for
some $\vep_{ik,jl}>0$ and any $(i,k)\ne (j,l)$ it is also possible to
show the joint \regvar\ of the processes $(X_{it}X_{jt})$, $i\ne j$,
with index $\alpha$. The description of the corresponding spectral
tail process is slightly tedious. It is not needed for the purposes of
this paper and therefore omitted.
\ere
\begin{proof} Regular variation of the marginal \ds s of  $(X_{it})$
  and $(X_{it}X_{jt})$ follows from Breiman's result; see
  \eqref{eq:br1} and \eqref{eq:br2}. As regards the \regvar\ of the
  \fidi s of $(X_{it})$, we have for $h\ge 1$,
\beao
\P(|X_{ih}|>x\mid |X_{i0}|>x)&=&\dfrac{\P(\min(|X_{i0}|,|X_{ih}|)>x)}{\P(|X_{i0}|>x)}\\
&\le &\dfrac{\P(\max(\sigma_{i0},\sigma_{ih}) \min(|Z_{i0}|,|Z_{ih}|)>x)}{\P(|X_{i0}|>x)}\to 0\,,\qquad \xto\,.
\eeao
In the last step we used Markov's inequality together with the moment
condition $\E[\sigma^{\alpha+\vep}]<\infty$ and the fact that 
$\min(|Z_{i0}|,|Z_{ih}|)$ is \regvary\ with index $2\alpha$.
This means that $\Theta^i_h=0$ for $h\ge 1$. 
\par
Similarly, for $i\ne j$, $h\ge 1$,
\beao
\P(|X_{ih}X_{jh}|>x\mid |X_{i0}X_{j0}|>x)&\le
&\dfrac{\P(\max(\sigma_{i0}\sigma_{j0},\sigma_{ih}\sigma_{jh})
  \min(|Z_{i0}Z_{j0}|, |Z_{ih}Z_{jh}|)>x)}{\P(|X_{i0}X_{j0}|>x)}\to 0\,.
\eeao
In the last step we again used Markov's inequality, the fact that $Z_{i0}Z_{j0}$ is \regvary\ with index $\alpha$
(see  Embrechts and Goldie \cite{embrechts:goldie:1980}; cf.\
Lemma~\ref{lem:product}(1) below), hence
$\min(|Z_{i0}Z_{j0}|,|Z_{ih}Z_{jh}|)$ is \regvary\ with
index~$2\alpha$, and the moment condition $\E
[(\sigma_{i0}\sigma_{j0})^{\alpha+\vep}]<\infty$.
Hence $\Theta_h^{ij}=0$ for $i\ne j$, $h\ge 1$.
\end{proof}
\subsection{Infinite variance stable limit theory for the stochastic
  volatility model and its product processes}\label{subsec:limit}
\bth\label{the:1}
Consider the stochastic volatility model \eqref{eq:1a:anja} and assume the following conditions:
\begin{enumerate}
\item
$Z$ is \regvary\ with index $\alpha\in (0,4) \setminus \{2\}$.
\item
$\big((\sigma_{it})_{t=1,2,\ldots}\big)_{i=1,\ldots,p}$  is strongly
mixing with rate \fct\ $(\alpha_h)$ \st\ for some $\delta>0$,
\beam\label{eq:mixi}
\sum_{h=0}^\infty \alpha_h^{\delta/(2+\delta)}<\infty\,.
\eeam
\item
The moment condition
\beam\label{eq:moment}
\E[\sigma^{2\max(2+\delta,\alpha+\epsilon)} ]<\infty
\eeam
holds for the same $\delta>0$ as in \eqref{eq:mixi} and some $\epsilon>0$. 
\end{enumerate}
Then 
\beam\label{eq:feller}
a_n^{-2}\big(S_{1}-c_n,\ldots,S_{p}-c_n\big) \std (\xi_{1,\alpha/2},\ldots,\xi_{p,\alpha/2})\,,
\eeam
where $(\xi_{i,\alpha/2})$ are iid $\alpha/2$-stable \rv s which are totally skewed to the right,
\beam\label{eq:6:anja}
c_n=\left\{\begin{array}{ll}
0& \alpha \in (0,2)\,,\\
n \,\E[X^2] &\alpha\in (2,4)\,,
\end{array}
\right.
\eeam
and $(a_n)$ satisfies $n\,\P(|X|>a_n)\to 1$ as $\nto$.
\ethe
\bre\label{rem:feller}
From classical limit theory (see Feller \cite{feller}, Petrov \cite{petrov:1995}) we know 
that \eqref{eq:feller} holds for an iid random field $(X_{it})$ with
\regvary\ $X$ with index $\alpha\in (0,4)$. In the case $\alpha=2$ one
needs the special centering $c_n=n\,\E [X^2 \I(|X|\le a_n)]$ which
often leads to some additional technical difficulties. For this reason
we typically exclude this case in the sequel.
\ere
\bre\label{rem:mix}
It follows from standard theory that $\alpha$-mixing of
$(\sigma_{it})$ with rate \fct\ $(\alpha_h)$ implies $\alpha$-mixing
of $(X_{it})$ with rate \fct\ $(4\alpha_h)$; see Davis and Mikosch
\cite{davis:mikosch:2009}.
\ere
\begin{proof} 
Recall the definition of $(\bfX_t)$ from \eqref{eq:multvts}. We will
verify the conditions of Theorem~\ref{thm:mikwin} 
for $\bfX_t^2=(X_{it}^2)_{i=1,\ldots,p}$, $t=0,1,2,\ldots$.\\[2mm]
(1) We start by verifying the \regvar\ condition for $(\bfX_t)$;  see  \eqref{eq:basseg}.
We will determine the \seq\ $(\bfTh_h)$ corresponding to $(\bfX_t)$.
We have for $t\ge 1$, with the max-norm $\|\cdot\|$,
\beao
\P\big(\|\bfX_t\|> x \mid \|\bfX_0\|>x\big)&\le &
\dfrac{\P\big( \|\bfX_t\|> x\,, \cup_{i=1}^p\{|X_{i0}|>x\}\big)}{\P(\|\bfX_0\|>x)}\\
&\le &\sum_{i=1}^p \dfrac{\P\big(  \|\bfX_t\|> x\,, |X_{i0}|>x\big)}{\P(\|\bfX_0\|>x)}\\
&\le &\sum_{i=1}^p \sum_{j=1}^p\dfrac{\P\big( |X_{jt}|> x\,, |X_{i0}|>x\big)}{\P(|X|>x)}\\
&\le &\sum_{i=1}^p \sum_{j=1}^p\dfrac{\P\big(
  \max(\sigma_{jt},\sigma_{i0})\min(|Z_{jt}|,|Z_{i0}|)>
  x\big)}{\P(\sigma|Z|>x)}\,.
\eeao
We observe that by Breiman's result and in view of the moment
condition \eqref{eq:moment}, for $t\ge 1$ and some positive constant
$c$,
\beao
\dfrac{\P\big( \max(\sigma_{jt},\sigma_{i0})\min(|Z_{jt}|,|Z_{i0}|)>
  x\big)}{\P(\sigma|Z|>x)} \sim c\, \dfrac{\P(\min(|Z_{jt}|,|Z_{i0}|)>
  x)}{\P(|Z|>x)}\,,
\eeao
and the \rhs\ converges to zero as $\xto$.
We conclude that $\bfTh_h=\bf0$ for $h\ge 1$. We also have for $i\ne j$,
\beao
\dfrac{\P(|X_{i0}|>x\,,|X_{j0}|>x)}{\P(|X|>x)} \leq \dfrac{\P\big(
  \max(\sigma_{i0},\sigma_{j0})\min(|Z_{i0}|,|Z_{j0}|)>
  x\big)}{\P(\sigma|Z|>x)} \to 0\,,\qquad \xto \,.
\eeao
Then, in a similar way, one can show
\beam\label{eq:Theta}
\P(\bfX_0/\|\bfX_{0}\|\in \cdot \mid \|\bfX_0\|>x)
&\stw & \P(\bfTh_0 \in \cdot) =\dfrac  1p \sum_{i=1}^p \big(p_+\vep_{\bfe_i}(\cdot)+p_-\vep_{-\bfe_i}(\cdot)\big)\,.
\eeam
where $\bfe_i$ are the canonical basis vectors in $\bbr^p$,
$\vep_{\bfx}$ is Dirac measure at $\bfx$ and $p_\pm$ are the tail 
balance factors in \eqref{eq:10b}. 
\par
We conclude that the spectral tail process $(\bfTh_h^{(2)})$ of $(\bfX_t^2)$ is given by 
$\bfTh_h^{(2)}=\bf0$ for $h\ge 1$ and from \eqref{eq:Theta} we also have
\beam\label{Eq:Theta_0:measure}
\P(\bfTh_0^{(2)}\in\cdot) =\dfrac  1p \sum_{i=1}^p \vep_{\bfe_i}(\cdot)\,.
\eeam
In particular, the condition $\sum_{i=1}^\infty  \E [\|\bfTh_i^{(2)}\|]<\infty$  in Theorem~\ref{thm:mikwin}(4)
is trivially satisfied.\\[2mm]
(2) Next we want to prove the mixing condition~\eqref{eq:chfa} for the \seq\ $(\bfX_t^2)$.  
We start by observing that there are integer \seq s  $(l_n)$ and
$(m_n)$ \st\ $k_n\,\alpha_{l_n}\to 0$, $l_n=o(m_n)$ and
$m_n=o(n)$. Then we  also have for any $\gamma>0$,
\beam\label{eq:neg}
k_n\,\P\big( \sum_{t=1}^{l_n} \bfX_t^2 \I(\|\bfX_t\|>\vep a_n)>\gamma
a_n^2 \big)\le k_n\,l_n\,\P(\|\bfX_t\|>\vep a_n)\le c\,
l_n/m_n=o(1)\,.\eeam
Relation \eqref{eq:chfa} turns into
\beao
\E \ex^{i \bfs 'a_n^{-2} \sum_{t=1}^n \bfX_t^2 \I(\|\bfX_t\|>\vep
  a_n)}- \big(\E \ex^{i\bfs' a_n^{-2}\sum_{t=1}^{m_n} \bfX_t^2
  \I(\|\bfX_t\|>\vep a_n)}\big)^{k_n}\to 0\,,\qquad \bfs\in\bbr^p\,.
\eeao
In view of \eqref{eq:neg} it is not difficult to see that we can
replace the sum in the former characteristic function by the sum over the index set
$J_n=\{1,\ldots,m_n-l_n,m_n+1,\ldots,2m_n-l_n,\ldots,\}\subset
\{1,\ldots,n\}$ and in the latter characteristic function by the sum
over the index set $\{1,\ldots,m_n-l_n\}$. 
Without loss of generality we may assume that $n/m_n$ is an
integer. Thus it remains to show that the following difference
converges to zero for every $\bfs\in\bbr^p$:
\beao\lefteqn{\Big|
\E \big[\ex^{i \bfs 'a_n^{-2} \sum_{t\in J_n} \bfX_t^2 \I(\|\bfX_t\|>\vep a_n)}\big]- 
\Big(\E\big[ \ex^{i\bfs' a_n^{-2}\sum_{t=1}^{m_n-l_n} \bfX_t^2
  \I(\|\bfX_t\|>\vep a_n)}\big]\Big)^{k_n}\Big|}\\&=&\Big|
\sum_{v=1}^{k_n} \E\Big[
\prod_{j=1}^{v-1} \ex^{i \bfs 'a_n^{-2} \sum_{t=(j-1)m_n+1}^{jm_n-l_n} \bfX_t^2 
\I(\|\bfX_t\|>\vep a_n)}\\&&\times  \big(\ex^{i \bfs 'a_n^{-2} \sum_{t=(v-1)m_n+1}^{vm_n-l_n} \bfX_t^2 
\I(\|\bfX_t\|>\vep a_n)} -\E \big[\ex^{i \bfs 'a_n^{-2} \sum_{t=(v-1)m_n+1}^{vm_n-l_n} \bfX_t^2 
\I(\|\bfX_t\|>\vep a_n)}\big] \big)\Big]\\&&\times
\prod_{j=v+1}^{k_n} \E \big[\ex^{i \bfs 'a_n^{-2}
  \sum_{t=(j-1)m_n+1}^{jm_n-l_n} \bfX_t^2
\I(\|\bfX_t\|>\vep a_n)}\big]\Big|\,.  
\eeao 
In view of a standard inequality for covariances of strongly mixing
\seq s of bounded \rv s (see Doukhan \cite{doukhan:1994}, p.~3) the
\rhs\ is bounded by $c\,k_n\alpha_{l_n}$ which converges to zero by
construction. Here and in what follows, $c$ stands for any
positive constant whose value is not of interest. Its value may change from line to line.
This finishes the proof of the mixing condition.\\[2mm]
(3)~Next we check the anti-clustering condition \eqref{eq:acl} for
$(\bfX_t)$ with normalization $(a_n)$, implying the corresponding 
condition for $(\bfX_t^2)$ with normalization $(a_n^2)$.
By similar methods as for part (1) of the proof, assuming that $\|\cdot\|$ is the max-norm, we have 
\beao\lefteqn{
\P\big(\max_{t=l,\ldots,m_n} \|\bfX_t\|>\gamma a_n\mid \|\bfX_0\|>\gamma a_n\big)}\\&\le &
\sum_{t=l}^{m_n} \P\big( \|\bfX_t\|>\gamma a_n\mid \|\bfX_0\|>\gamma a_n\big)\\
&\le &c\,\sum_{t=l}^{m_n} \sum_{i=1}^p\sum_{j=1}^p\dfrac{\P\big(
  |X_{it}|>\gamma a_n\,,|X_{j0}|>\gamma a_n\big)}{\P(|Z|>\gamma
  a_n)}\\
&\le &c\,\sum_{t=l}^{m_n} \sum_{i=1}^p\sum_{j=1}^p
\dfrac{\P\big(\max (\sigma_{it},\sigma_{j0})\min (|Z_{it}|,|Z_{j0}|)>\gamma a_n\big)}{\P(|Z|>\gamma a_n)}\\
&\le &c\,\sum_{t=l}^{m_n} \sum_{i=1}^p\sum_{j=1}^p
\dfrac{\P\big(\sigma_{it}\min (|Z_{it}|,|Z_{j0}|)>\gamma a_n\big)}{\P(|Z|>\gamma a_n)}\,.
%+\dfrac{\P\big(\sigma_{0j}\min (|Z_{it}|,|Z_{0j}|)>\delta a_n\big)}{\P(|Z|>\delta a_n)}\Big]
\eeao
By stationarity the \pro ies on the \rhs\ do not depend on $t\ge
l$. Therefore and by Breiman's result, the \rhs\ is bounded by
\beao
c\, m_n \dfrac{\P\big(\min (|Z_{it}|,|Z_{j0}|)>\gamma
  a_n\big)}{\P(|Z|>\gamma a_n)}=O((m_n/n) [n\,\,\P(|Z|>a_n)])=o(1)\,.
\eeao
This proves \eqref{eq:acl} for $(\bfX_t)$.\\[2mm]
(4)
Next we check the vanishing small values condition \eqref{eq:vansm} for the partial sums of 
$(\bfX_t^2)$ and $\alpha\in (2,4)$. 
It is not difficult to see that it suffices to prove the corresponding result for the component
processes:
\beam\label{eq:van2}
&&\lim_{\vep\downarrow 0}\limsup_{\nto}\P\Big(\Big|\sum_{t=1}^n
\big(X_{it}^2\I(|X_{it}|\le \vep a_n)-\E [X_{it}^2\I(|X_{it}|\le \vep
a_n)]\big)\Big|>\gamma a_n^2\Big)=0\,,\\&& \quad
\gamma>0\,,\;i=1,\ldots,p\,.\nonumber
\eeam
We have 
\beao\lefteqn{
a_n^{-2}\sum_{t=1} ^n\sigma_{it}^2
\E\big[Z_{it}^2\I(|X_{it}|\le \vep a_n)\mid \sigma_{it}] -a_n^{-2}\,n\,\E[X_{it}^2\I(|X_{it}|\le \vep a_n)]}\\
&=&  a_n^{-2}\sum_{t=1} ^n(\sigma_{it}^2 -\E[\sigma_{it}^2])\, \E[Z^2] -
a_n^{-2} \sum_{t=1} ^n \big(\sigma_{it}^2 \E[Z_{it}^2 \I(|X_{it}|>\vep
a_n)\mid \sigma_{it}] - \E[X_{it}^2 \I(|X_{it}|>\vep
a_n)]\big)\\&=&I_1+I_2\,.
\eeao
The \seq\ $(\sigma_{it}^2)$ satisfies the \clt\ with normalization
$\sqrt{n}$. This follows from Ibragimov's \clt\ for strongly mixing
\seq\ 
whose rate \fct\ $(\alpha_h)$ satisfies \eqref{eq:mixi} and has moment
$\E [\sigma^{2(2+\delta))}]<\infty$ (see \eqref{eq:moment}); 
cf. Doukhan \cite{doukhan:1994}, p.~45. We know that
$\sqrt{n}/a_n^2\to 0$ for $\alpha\in (2,4)$. Therefore $I_1\stp 0$.
We also have
\beao
\E[I_2^2]&\le& \,\dfrac n {a_n^{4}} \E \big[\sigma^4 (\E
[Z^2\I(|X|>\vep a_n)\mid \sigma])^2\big]\\&& + 2\, \dfrac {n}{a_n^{4}}
\sum_{h=1}^n |\cov(\sigma_{i0}^2 \E\big[Z_{i0}^2\I(|X_{i0}^2|> \vep
a_n)\mid \sigma_{i0}], \sigma_{ih}^2 \E\big[Z_{ih}^2\I(|X_{ih}^2|>
\vep a_n)\mid \sigma_{ih}])| \\
&=&I_3+I_4\,.
\eeao
%By Karamata's theorem 
In view of the moment conditions on $\sigma$ and since  $\E[Z^2]<\infty$,
$I_3\le c (n/a_n^4)\to 0$. In view of Doukhan \cite{doukhan:1994}, Theorem 3 on p.~9,  we have
\beao
I_4&\le& c\,\dfrac n {a_n^{4}}\sum_{h=1}^n
\alpha_h^{\delta/(2+\delta)}
(\E|\sigma|^{2(2+\delta)})^{2/(2+\delta)}\to 0\,.
\eeao
Thus it suffices for \eqref{eq:van2} to prove
\beao
\lim_{\vep\downarrow 0}\limsup_{\nto} \P\Big(\Big|\sum_{t=1}^n \big(\sigma_{it}^2\E[Z_{it}^2\I(|X_{it}|
\le \vep a_n)\mid\sigma_{it}] - X_{it}^2\I(|X_{it}|\le \vep a_n)\big)\Big|>\gamma\,a_n^2\Big)=0\,,\qquad \gamma>0\,.
\eeao
The summands are independent and centered, conditional on the
$\sigma$-field generated by $(\sigma_{it})_{t=1,\ldots,n}$. An
application of \v Cebyshev's inequality 
conditional on this $\sigma$-field and 
%independent centered random variable (see Petrov \cite{petrov:1995}, 2.6.20 on p.~82) 
Karamata's theorem yield, as $\nto$,
\beao&&
\E\Big[\P\Big(\Big|\sum_{t=1}^n \big(\sigma_{it}^2\E[Z_{it}^2\I(|X_{it}|
\le \vep a_n)\mid\sigma_{it}] - X_{it}^2\I(|X_{it}|\le \vep
a_n)\big)\Big|>\gamma\,a_n^2\big| (\sigma_{is})\Big)\Big]\\
&\le &
c\, a_n^{-4}\E\Big[\sum_{t=1}^n \var(X_{it}^2\I(|X_{it}|\le \vep a_n)\mid \sigma_{it})\mid (\sigma_{is})\Big]\\
&\le &c\,n\,\vep^{4}\,
%\Big[\E\big[\sigma_{it}^{2\nu} \big[(\E[Z_{it}^2\I(|X_{it}|\le \vep a_n)\mid \sigma_{it}])^\nu+
\E[|X/(\vep a_n)|^{4}\I(|X|\le \vep a_n)]%\mid \sigma_{it}\big]\big]\Big]\\
\to  c\,\vep^{4-\alpha}\,.
\eeao
The \rhs\ converges to zero as $\vep\downarrow 0$.

This proves that all assumptions of Theorem \ref{thm:mikwin} are satisfied. 
Therefore the random variables on the \lhs\ of \eqref{eq:feller}
converge to an $\alpha$-stable random vector with log characteristic
function
\begin{eqnarray*}
&&\int_0^\infty \E\big[\ex^{i\,y\,\bft'\sum_{j=0}^\infty
  \bfTh_j^{(2)}}- \ex^{i\,y\,\bft'\sum_{j=1}^\infty
  \bfTh_j^{(2)}}-i\,y\,\bft'\I_{(1,2)}(\alpha/2)\big]\,
d(-y^{\alpha/2})\\
&=&\sum_{j=1}^p \frac{1}{p}\int_0^\infty \E\big[\ex^{i\,y\,t_j}-i\,y\,t_j\I_{(1,2)}(\alpha/2)\big]\,
d(-y^{\alpha/2})\,,\qquad \bft=(t_1, \ldots, t_p)'\in\bbr^p,
\end{eqnarray*}
where we used \eqref{Eq:Theta_0:measure} and that $\bfTh_h^{(2)}=\bf0$
for $h\ge 1$. One easily checks that all summands in this expression
are homogeneous functions in $t_j$ of degree $\alpha/2$.  Therefore,
the limiting random vector in \eqref{eq:feller} has the same
distribution as the sum $\sum_{j=1}^p \bfe_j \xi_{j,\alpha/2}$ for iid
$\xi_{j,\alpha/2}$ which are $\alpha/2$-stable and totally skewed to
the right (because all the summands in $S_j$ are non-negative).
\end{proof}

\subsection{Eigenvalues of the sample covariance matrix}\label{subsec:eigen}
We have the following approximations:
\bth\label{lem:1}
Assume that one of the following conditions holds:
\begin{enumerate}
\item
$(X_{it})$ is an iid field of \regvary\ \rv s with index 
$\alpha\in (0,4)$. If \ $\E[|X|]<\infty$ we also assume $\E[X]=0$.
\item
$(X_{it})$ is a stochastic volatility model \eqref{eq:1a:anja} satisfying
the \regvar , mixing and moment conditions of Theorem~\ref{the:1}.
If $\E [|Z|]<\infty$ we also assume $\E[Z]=0$. 
\end{enumerate}
Then, with $\mathbf{X}^n$ as in \eqref{eq:esses},
\beao
&&a_n^{-2} \twonorm{\bfX^n(\bfX^n)'-\diag(\bfX^n(\bfX^n)')}\stp 0\,,
\eeao
where $\twonorm{\cdot}$ is the spectral norm and 
$(a_n)$ is a \seq\ \st\ $n\,\P(|X|>a_n)\to 1$.
\ethe
\begin{proof} Part (1). Recall that for a 
$p\times p$ matrix $\bfA$ we have $\twonorm{\bfA}\le \frobnorm{\bfA}$,
where $\frobnorm{\cdot}$ denotes the Frobenius norm.
Hence
\beam\label{eq:1:anja}
a_n^{-4} \twonorm{\bfX^n(\bfX^n)'-\diag(\bfX^n(\bfX^n)')}^2&\le &
a_n^{-4} \frobnorm{\bfX^n(\bfX^n)'-\diag(\bfX^n(\bfX^n)')}^2\nonumber
\\
&=& \sum_{1\le i\ne j\le p} \big(a_n^{-2}S_{ij}\big)^2\,.
\eeam
In view of the assumptions, $(X_{it}\,X_{jt})_{t=1,2,\ldots}$, $i\ne
j$, is an iid \seq\ of \regvary\ \rv s with index $\alpha$ which is 
also centered if $\E[|X|]<\infty$. We consider two different cases.\\
{\em The case $\alpha\in (0,2)$.} According to classical limit theory
(see Feller \cite{feller}, Petrov \cite{petrov:1995}) we have for
$i\ne j$,
$b_n^{-1}S_{ij}\std \xi_{\alpha}$, (see \eqref{eq:esses} for the definition of $S_{ij}$)
where $\xi_{\alpha}$ is an $\alpha$-stable \rv\ and $(b_n)$ is 
chosen \st\ $n\,\P(|X_1X_2|>b_n)\to 1$ for independent copies $X_1,X_2$ of $X$.
Since $(b_n)$ and $(a_n^2)$ are \regvary\ with indices 
$1/\alpha$ and $2/\alpha$, respectively, the \rhs\ in \eqref{eq:1:anja} converges to zero in \pro y.\\[1mm]
{\em The case $\alpha\in [2,4)$.} In this case the \ds\ of $X_1X_2$ is
in the domain of attraction of the normal law.
Since $X_1X_2$ has mean zero we can apply classical limit theory (see
Feller \cite{feller}, Petrov \cite{petrov:1995}) to conclude that 
$
b_n^{-1} S_{ij} \std N\,,
$
where $(b_n)$ is \regvary\ with index $1/2$ and $N$ is centered
Gaussian. Since $b_n/a_n^2\to 0$ we again conclude that the \rhs\ of
\eqref{eq:1:anja}
converges to zero in \pro y.\\[2mm]
Part (2). We again appeal to \eqref{eq:1:anja}.
Let $\gamma < \min(2,\alpha)$. Then we have for $i \neq j$, using the
independence of $(X_{it}X_{jt})$ conditional on
$((\sigma_{it},\sigma_{jt}))$ and that the distribution of $Z$ is
centered if its first absolute moments exists, that
\beao
a_n^{-2\gamma} \E\Big[\big|S_{ij}\big|^\gamma\mid ((\sigma_{it},\sigma_{jt}))\Big]
&\le &c\,\dfrac{n}{a_n^{2\gamma}} \dfrac 1 n \sum_{t=1}^n (\sigma_{it}\sigma_{jt})^{\gamma}(\E|Z|^\gamma)^2\, ,
\eeao
cf.\ von Bahr and Ess\'een \cite{bahr:esseen:1965} and Petrov~\cite{petrov:1995}, 2.6.20 on p.~82.
In view of the moment condition \eqref{eq:moment} we have 
$ \E[(\sigma_{i}\sigma_{j})^{\gamma}]<\infty$ and $n/a_n^{2\gamma}\to
0$ if we choose $\gamma$ sufficiently close to $\min(2,\alpha)$. Then
the \rhs\ 
converges to zero in view of the ergodic theorem. An application of the conditional Markov inequality 
of order $\gamma$ yields 
$
a_n^{-2} S_{ij}\stp 0\,.
$
This proves the theorem.
\end{proof}
\begin{corollary}\label{cor:sv1}
Assume that 
$(X_{it})$ is either
\begin{enumerate}
\item
an iid field of \regvary\ \rv s with index $\alpha\in (0,4)$ and $\E [X]=0$ if $\E [|X|]< \infty$, or
\item 
a stochastic volatility model of \regvary\ \rv s with index $\alpha\in
(0,4)\setminus \{2\}$ satisfying the conditions of
Theorem~\ref{lem:1}(2).
\end{enumerate}
Then
\beao
a_n^{-2}\max_{i=1,\ldots,p} \big|\la_{(i)}- S_{(i)}\big|\stp 0\,,
\eeao
where $(\la_i)$ are the eigenvalues of $\bfX^n(\bfX^n)'$,
$\la_{(1)}\ge \cdots\ge \la_{(p)}$ are their ordered values  
and $S_{(1)}\ge \cdots \ge S_{(p)}$ are the ordered values of $S_1,\ldots,S_p$ defined in \eqref{eq:esses}. 
In particular, we have 
\beam\label{eq:2a:anja}
a_n^{-2} \big(\la_{(1)}-c_n,\ldots,\la_{(p)}-c_n\big) \std \big(\xi_{(1),\alpha/2},\ldots,\xi_{(p),\alpha/2}\big)\,,
\eeam
where $(c_n)$ is defined in \eqref{eq:6:anja} for $\alpha\ne 2$ and in Remark~\ref{rem:feller} for $\alpha=2$,
$(\xi_{i,\alpha/2})$ are iid $\alpha/2$-stable \rv s given in
Theorem~\ref{the:1} for the stochastic volatility model and in
Remark~\ref{rem:feller} for the iid field,
and $\xi_{(1),\alpha/2}\ge \cdots\ge \xi_{(p),\alpha/2}$ are their
ordered values. 
\end{corollary}
\begin{proof}
We have by Weyl's inequality (see Bhatia \cite{bhatia:1997}) and  Theorem~\ref{lem:1},
\beam\label{eq:weyl1}
a_n^{-2}\max_{i=1,\ldots,p} \big|\la_{(i)}- S_{(i)}\big|&\le &
a_n^{-2} \twonorm{\bfX^n(\bfX^n)'-\diag(\bfX^n(\bfX^n)')}\stp 0\,.
\eeam
If $(X_{it})$ is an iid random field (see Remark~\ref{rem:feller}) or
a stochastic volatility model satisfying the conditions of
Theorem~\ref{lem:1}(2)
we have \eqref{eq:feller}. Then \eqref{eq:weyl1} implies \eqref{eq:2a:anja}.
\end{proof}
\bre\label{rm:many}\rm 
If $\alpha\in (2,4)$ we have $\E[X^2]<\infty$. Therefore
\eqref{eq:2a:anja} reads as 
\beam\label{eq:rem:many}
\dfrac{n}{a_n^2}\,\big(\dfrac{\la_{(i)}}{n}- \E[X^2]\big)_{i=1,\ldots,p}\std (\xi_{(i),\alpha/2})_{i=1,\ldots,p}\,.
\eeam
We notice that $n/a_n^2\to \infty$ for $\alpha\in (2,4)$ since
$(n/a_n^2)$ is \regvary\ with index $1-2/\alpha$. In particular, if
${\rm tr}(\bfX^n(\bfX^n)')$ denotes the trace of $\bfX^n(\bfX^n)'$ we
have for $i\le p$,
\beam\label{eq:ratio}
\dfrac{\la_{(i)}}{{\rm tr}(\bfX^n(\bfX^n)')}&=& \dfrac{\la_{(i)}/n}{(\la_1+\cdots +\la_p)/n}
\stp \dfrac  1p\,.
%\dfrac{\la_{(1)}+\cdots + \la_{(i)}}{{\rm tr}(\bfX\bfX')}&\stp& \dfrac  ip\,.
\eeam
\par
The joint \asy\ distribution of the ordered eigenvalues $(\la_{(i)})$
is easily calculated from the distribution of a totally skewed
$\alpha/2$-stable \rv\ $\xi_{1,\alpha/2}$; in particular, the limit of
$(a_n^{-2}(\la_{(1)}-c_n))$ has the \ds\ of
$\max(\xi_{1,\alpha/2},\ldots,\xi_{p,\alpha/2})$.
\par
For applications, it is more natural to replace the \rv s $X_{it}$ by their mean-centered versions $X_{it}-\ov X_i$,
where $\ov X_{i}= (1/n) \sum_{t=1}^n X_{it}$, instead of assuming that
they have mean zero. The previous results remain valid for the
sample-mean centered \rv s $X_{it}$, also in the case when $X$ has
infinite first moment.
\ere

\subsection{Some applications: Limit results for ordered eigenvalues
  and eigenvectors of the sample covariance matrix}\label{subsec:appl} 
In what follows, we assume the conditions of Corollary~\ref{cor:sv1}.
\subsubsection{Spacings}\label{ssection:spacing}
Using the joint \con\ of the normalized ordered eigenvalues
$(\la_{(i)})$ we can calculate the limit of the spectral gaps:
\beam\label{eq:order:anja}
\big(\dfrac{\la_{(i)}-\la_{(i+1)}}{a_n^2}\big)_{i=1,\ldots,p-1} \std
\big(\xi_{(i),\alpha/2}-\xi_{(i+1),\alpha/2}\big)_{i=1,\ldots,p-1}\,.
\eeam
\par
We notice that the ordered values $\xi_{(i),\alpha/2}$ and linear \fct
als thereof (such as $\xi_{(i),\alpha/2}-\xi_{(i+1),\alpha/2}$) are
again jointly \regvary\ with index $\alpha/2$. 
This is due to the continuous mapping theorem for \regvary\ vectors; see Hult and 
Lindskog \cite{hult:lindskog:2005,hult:lindskog:2006}, cf. 
Jessen and Mikosch \cite{jessen:mikosch:2006}.
\subsubsection{Trace}
For the trace of $\bfX^n(\bfX^n)'$ we have
\beao
a_n^{-2} \big({\rm tr}(\bfX^n(\bfX^n)')-p\,c_n\big)&=& a_{n}^{-2}\sum_{i=1}^p (S_i-c_n)\\
&=& a_{n}^{-2}\sum_{i=1}^p (\la_i-c_n)\std
\xi_{1,\alpha/2}+\cdots+\xi_{p,\alpha/2}\overset{d}{=}
p^{2/\alpha}\xi_{1,\alpha/2}\,.
\eeao
Moreover, we have the joint \con\ of the normalized and centered
$(\la_{(i)})$ and ${\rm tr}(\bfX^n(\bfX^n)')=\la_1+\cdots +\la_p$. In
particular, we have the self-normalized limit relations
\beao
\big(\dfrac{\la_{(i)}-c_n}{{\rm tr}(\bfX^n(\bfX^n)')-p\,c_n}\big)_{i=1,\ldots,p}&\std& 
\big(\dfrac{\xi_{(i),\alpha/2}}{\xi_{1,\alpha/2}+\cdots+\xi_{p,\alpha/2}}\big)_{i=1,\ldots,p}\,,
\eeao
and for $\alpha\in (2,4)$, by the \slln ,
\beao
\dfrac{np}{a_n^2}\,\Big(\dfrac{\la_{(i)}-c_n}{{\rm
    tr}(\bfX^n(\bfX^n)')}\Big)_{i=1,\ldots,p}&\std&
\dfrac{\xi_{(i),\alpha/2}}{\E [X^2]}\,.
\eeao
\subsubsection{Determinant}
Since $\la_i-c_n$ are the eigenvalues of $\bfX^n(\bfX^n)'-c_n \bfI_p$,
where $\bfI_p$ is the $p\times p$ identity matrix,
we obtain for the determinant
\beao
{\rm det} \big(a_n^{-2}(\bfX^n(\bfX^n)'-c_n \,\bfI_p)\big)&=& \prod_{i=1}^p a_n^{-2} (\la_{(i)}-c_n)\\
&\std& \xi_{(1),\alpha/2}\cdots \xi_{(p),\alpha/2}=\xi_{1,\alpha/2}\cdots \xi_{p,\alpha/2} \,. 
\eeao
For $\alpha\in (2,4)$, we also have 
\beao
\dfrac 1{a_n^2 c_n^{p-1}} \big({\rm det} (\bfX^n(\bfX^n)')- c_n^p\big)&=&
\sum_{i=1}^p  a_n^{-2}\big(\la_{(i)}-c_n\big) \prod_{j=1}^{i-1} \dfrac{\la_{(j)}}{c_n}\\
&\std &\sum_{i=1}^p \xi_{(i),\alpha/2}= \sum_{i=1}^p\xi_{i,\alpha/2} \overset{d}{=} p^{2/\alpha}\,\xi_{1,\alpha/2}\,,
\eeao
where we used \eqref{eq:rem:many}. 
\subsubsection{Eigenvectors}\label{Subsub:eigenvectors}
It is also possible to localize the eigenvectors of the matrix
$a_n^{-2}\bfX^n(\bfX^n)'$. Since this matrix is approximated
by its diagonal in spectral norm, one may expect that the unit
eigenvectors of the original matrix are close to the canonical basis 
vectors. We can write
\beao
a_n^{-2}\bfX^n(\bfX^n)'\bfe_{L_j}= a_n^{-2}S_{(j)}\,\bfe_{L_j}+ \vep_n\,\bfW\,,
\eeao
where $\bfW$ is a unit vector orthogonal to $\bfe_{L_j}$, $L_j$ is the index of $S_{(j)}=S_{L_j}$ and 
\beao
\vep_n= a_n^{-2}\|\big(\bfX^n(\bfX^n)'- S_{(j)}\big) \bfe_{L_j}\|_{\ell_2}\stp 0\,,
\eeao
from Theorem \ref{lem:1} and by equivalence of all matrix norms. 
According to Proposition A.1 in Benaych-Georges and Pech\'e
\cite{benaych:peche:2014}, there is an eigenvalue $a_n^{-2}\la_{(j)}$
of  $a_n^{-2}\bfX^n(\bfX^n)'$ in some $\vep_n$-neighborhood of  $a_n^{-2}S_{(j)}$. Define
\beao\Omega_n=\{a_n^{-2}|\la_{(j)}-\la_{(l)}|>d_n\,,l\ne j \}\,,
\eeao
for $d_n=k\vep_n$ for any fixed $k>1$. Then
$\lim_{\nto}\P(\Omega_n)=1$ because of \eqref{eq:order:anja} and $d_n\stp 0$.
Hence, for large $n$, $a_n^{-2}\la_{(j)}$ and $a_n^{-2}\la_{(l)}$ have
distance at least $d_n$ with high \pro y. Another application of 
Proposition A.1 in \cite{benaych:peche:2014} yields that the unit
eigenvector $\bfV$ associated with $a_n^{-2}\la_{(j)}$ 
satisfies the relation
\beao
\limsup_{\nto}\P\big(\|\bfV- V_{L_j}\bfe_{L_j}\|_{\ell_2}>\delta\big)&\le &
\limsup_{\nto}\P\big(\{\|\bfV- V_{L_j}\bfe_{L_j}\|_{\ell_2}>\delta\}\cap \Omega_n\big)+
\limsup_{\nto}\P(\Omega_n^c)\\
&\le &\limsup_{\nto}\P\big( \{2\,\vep_n /(d_n-\vep_n)>\delta\}\cap \Omega_n\big)\\
%& \le & \limsup_{\nto}\P\big( 2\,\vep /(d-\vep)>\delta\big)\\
&= & \I_{\{2/(k-1)>\delta\}}\,.
\eeao
For any fixed $\delta>0$, the \rhs\ is zero for sufficiently large
$k$. Since both $\bfV$ and $\bfe_{L_j}$ are unit eigenvectors
this means that 
$\|\bfV- \bfe_{L_j}\|_{\ell_2}\stp 0$.
\subsubsection{Sample correlation matrix}
In Remark~\ref{rm:many} we mentioned that we can replace the variables
$X_{it}$ by their sample-mean centered versions $X_{it}-\ov X_i$
without changing the \asy\ theory. Similarly, one may be interested in transforming the $X_{it}$ as follows:
\beao
\wt X_{it}= \dfrac{X_{it}- \ov X_i}{\wh \sigma_i}\,,\qquad 
\wh \sigma_i^2= \sum_{t=1}^n (X_{it}-\ov X_i)^2\,. 
\eeao
Then the matrix
\beao
\wt\bfX^n(\wt \bfX^n)'= \big(\sum_{t=1}^n\wt X_{it}\wt X_{jt}\big)_{i,j=1,\ldots,p}\,,
\eeao
is  the sample correlation matrix. We write $\wt \la_i$, $i=1,\ldots,p$, for the eigenvalues
of $\wt \bfX^n(\wt \bfX^n)'$ and $\wt \la_{(1)}\ge \cdots \ge \wt \la_{(p)}$ for their ordered values. 
\par
We notice that the entries of this matrix are all bounded in modulus by one.
In particular, the diagonal consists of ones.
We do not have a complete  limit theory for the eigenvalues  $\wt \la_i$.
We restrict ourselves to iid $(X_{it})$ to explain the differences.
\begin{lemma}
Assume that 
$(X_{it})$ is an iid field of \rv s. 
%If $\E[|X|]<\infty$ we also assume $\E[X]=0$.
\begin{enumerate}
\item If $\E[X^2]<\infty$ then 
\beao
\sqrt{n} \max_{i=1,\ldots,p} |\wt \la_{i}-1|= O_\P(1)\,.
\eeao
\item If $X$ is \regvary\ with index $\alpha\in (0,2)$ then
\beao
\dfrac{a_n^2}{b_n} \max_{i=1,\ldots,p} |\wt \la_{i}-1|=O_\P(1)\,,
\eeao
where $(a_n)$ and $(b_n)$  are chosen \st\ $\P(|X|>a_n)\sim
\P(|X_1X_2|>b_n) \sim n^{-1}$ for iid copies $X_1,X_2$ of $X$.
\end{enumerate}
\end{lemma}
\bre
Notice that the lemma implies $\wt \la_i\stp 1$ for $i=1,\ldots,p$, and the analog of
relation \eqref{eq:ratio} remains valid. 
\ere
\begin{proof} Part(1) We assume without loss of generality that $1=\E [X^2]$.
Then by classical limit theory,
\beao
\sqrt{n}\big(\wt \bfX^n(\wt \bfX^n)'-{\rm diag} (\wt \bfX^n(\wt
\bfX^n)')\big)&=& \sqrt{n}\big(\wt \bfX^n(\wt \bfX^n)'-\bfI_p\big)\\
&=& \Big(\I(i\ne j)\dfrac{n^{-1/2}\sum_{t=1}^n (X_{it}-\ov X_{i})(X_{jt}-\ov X_j)}{(\wh \sigma_i/\sqrt{n}) 
(\wh\sigma_j/\sqrt{n})}\Big)\\
&\std & \big(N_{ij}\I(i\ne j)\big)\,,
\eeao
where $N_{ij}$, $1\le i<j\le n$, are iid $N(0,1)$ and $N_{ij}=N_{ji}$. By Weyl's inequality,
\beao
\sqrt{n} \max_{i=1,\ldots,p}\Big|\wt \la_{(i)}-1\Big|\le\sqrt{n} \|\wt \bfX^n(\wt \bfX^n)'-\bfI_p\|_2=O_\P(1)\,.
\eeao
Part(2) If $X$ is \regvary\ with index $\alpha\in (0,2)$, we have that
$(a_n^{-2} \wh \sigma_i^2)$ converges to a vector of iid positive
$\alpha/2$-stable
\rv s $(\xi_i)$, while for every $i\ne j $, $b_n^{-1}\sum_{t=1}^n 
(X_{it}-\ov X_{i})\,(X_{jt}-\ov X_{j})\std \xi_{ij}$ and the limit $\xi_{ij}$ is $\alpha$-stable.
Then by Weyl's inequality
\beao
\dfrac{a_n^2}{b_n} \max_{i=1,\ldots,p}\Big|\wt \la_{(i)}-1\Big|\le
\dfrac{a_n^2}{b_n}\|\wt \bfX^n(\wt \bfX^n)'-\bfI_p\|_2=O_\P(1)\,.
\eeao
\end{proof}

\section{Case (2): $\sigma$ dominates the tail}\label{Sec:case2}\setcounter{equation}{0}
In this section we assume the conditions of Case (2); see Section \ref{subsec:Case:2}. Our goal is to derive results analogous to Case (1): \regvar\ 
of $(X_{it})$, infinite variance limits for $S_{ij}$ and limit theory for the eigenvalues of the corresponding sample 
covariance matrices. It turns out that this case offers a wider spectrum of possible limit behaviors and that we have to further distinguish our assumptions about the distribution of $\eta$. So, in addition to \eqref{Eq:tail:eta} we assume that either
\begin{equation}\label{Eq:inf:mom:eta} \E[\ex^{\eta \alpha}]=\infty\end{equation}
or 
\begin{equation}\label{Eq:conv:eq:eta} \lim_{x \to \infty}\frac{\P(\eta_1+\eta_2>x)}{\P(\eta_1>x)}=c \in (0,\infty) \;\; \Leftrightarrow\;\; \lim_{x \to \infty}\frac{\P(\ex^{\eta_1}\cdot \ex^{\eta_2}>x)}{\P(\ex^{\eta_1}>x)}=c \in (0,\infty)\end{equation}
hold, where $\eta_1$ and $\eta_2$ are independent copies of $\eta$.
\bre Following  Cline \cite{cline:1986}, we call 
the distribution of a \rv\ $\eta$ {\it convolution equivalent} if  $\ex^\eta$ is regularly varying and
relation \eqref{Eq:conv:eq:eta} holds. The assumptions \eqref{Eq:inf:mom:eta} and \eqref{Eq:conv:eq:eta} are mutually exclusive, since the only possible finite limit $c$ in \eqref{Eq:conv:eq:eta} is given by $c=2\E[\ex^{\eta \alpha}]$; see Davis and Resnick \cite{davis:resnick:1986}. There are, however, 
regularly varying distributions of $\ex^\eta$ which satisfy $\E[\ex^{\eta \alpha}]<\infty$ but not
\eqref{Eq:conv:eq:eta}. An example is given in Cline \cite{cline:1986}, p.\ 538;  see also Lemma~\ref{lem:product}(3) 
for a necessary and sufficient condition ensuring \eqref{Eq:conv:eq:eta}.
\ere
As we will see later, relations \eqref{Eq:inf:mom:eta} and \eqref{Eq:conv:eq:eta} cause
rather distinct limit behavior of the sample covariance matrix. In particular, \eqref{Eq:conv:eq:eta}
allows for non-vanishing off-diagonal elements of the normalized sample  covariance matrices, in contrast to Case (1). 
\par
For notational simplicity, define
\beao
\psi=\max_{k,l} \psi_{kl}\qquad \mbox{and}\qquad \Lambda=\{(k,l): \psi_{kl}=\psi\}\,.
\eeao
Recall that for convenience we assume that $\psi=1$; if the latter condition does not hold we can replace  
(without loss of generality) the \rv s $\eta_{kl}$ by $\psi \eta_{kl}$ and the coefficients $\psi_{kl}$ by $\psi_{kl}/\psi$.
For given $(i,j)$, we define 
\begin{equation} \label{eq:psi:ij:def}
\psi^{ij}= \max_{k,l}\, (\psi_{kl}+\psi_{k+i-j,l})\,.
\end{equation}
Notice that $1\le \psi^{ij}\le 2$. For $d\ge 1$, we write $\mathbf{i}=(i_1, \ldots, i_d), \mathbf{j}=(j_1, \ldots, j_d)$ for elements of $\mathbb{Z}^d$. For given  $\mathbf{i}$ and  $\mathbf{j}$ we also define
\beao
\psi^{\mathbf{i},\mathbf{j}}= \max_{1 \leq l \leq d} \psi^{i_l,j_l}.
\eeao

\subsection{Regular variation}\label{subsec:regvar2}
We start by showing that the volatility sequences are regularly varying.
\begin{proposition}\label{prop:rv:1} Under the aforementioned conditions and conventions (including that 
either \eqref{Eq:inf:mom:eta} or \eqref{Eq:conv:eq:eta} hold), 
\begin{enumerate}
\item
each of the \seq s 
$(\sigma_{it})_{t \in \mathbb{Z}}$, $i=1,2,\ldots$, is \regvary\ with index $\alpha$,
\item
each of the \seq s $(\sigma_{it}\sigma_{jt})_{t \in \mathbb{Z}}$, $i,j=1,2,\ldots$, is \regvary\ with corresponding index  $\alpha/\psi^{ij}$,
\item
For $d\geq 1$ and $\mathbf{i},\mathbf{j}\in\bbz$, the $d$-variate \seq\ 
$((\sigma_{i_k,t}\sigma_{j_k,t})_{1\leq k \leq d})_{t \in \mathbb{Z}}$ is regularly 
varying with index $\alpha/\psi^{\mathbf{i},\mathbf{j}}$.
\end{enumerate}
\end{proposition}
\bre Part (3) of the proposition possibly includes degenerate cases in the sense that for some choices of $(i_k,j_k)$, 
$(\sigma_{i_k,t}\sigma_{j_k,t})$ is \regvary\ with index $\alpha/\psi^{i_k,j_k}> \alpha/\psi^{\mathbf{i},\mathbf{j}}$. 
\par
Part (3) implies (2) in the case $d=1$. Part (2) implies (1) by setting $i=j$ and observing that, 
by non-negativity of $\sigma$,  \regvar\ of $(\sigma_{it}^2)$
with index $\alpha/2$ is equivalent to \regvar\ of $(\sigma_{it})$ with index $\alpha$. 
\ere
% \bre We notice that $(\sigma_{it}^2)_{t=1,2,\ldots}$ is \regvary\ with index $\alpha/(2\psi)$. If $m(i,j)>2$ for some $i\ne j$ then
% $(\sigma_{it}\sigma_{jt})_{t=1,2,\ldots}$ is \regvary\ with a smaller index than $(\sigma_{it}^2)_{t=1,2,\ldots}$. If $\Lambda$ consists of exactly one element
% then $m(i,j)=1$ for $i\ne j$.
% \ere
\begin{proof} To give some intuition we start with the proof of the marginal regular variation of $\sigma$, although it is just a special case of (1).
We have
\beam\label{eq:oo}
\sigma_{it}= \ex^{ \sum_{(k,l)\in\Lambda}\eta_{i-k,t-l}}\,
\ex^{\sum _{(k,l)\not \in\Lambda} \psi_{kl} \eta_{i-k,t-l}}=:
\sigma_{it,\Lambda}\sigma_{it,\Lambda^c}.
\eeam
%{\blue where we note that $\sigma_{it,\Lambda}$ and $\sigma_{it,\Lambda^c}$ are independent for each $i,t$.}
We first verify that $\sigma=\sigma_\Lambda\sigma_{\Lambda^c}$ is \regvary\ with index 
$\alpha$. Since $|\Lambda|<\infty$ by our assumptions, and in view of Embrechts and Goldie \cite{embrechts:goldie:1980}, Corollary on p.~245, cf. also Lemma~\ref{lem:product}(1) below,
the product
$\sigma_\Lambda$ is \regvary\ with index $\alpha$. % hence $\sigma_\Lambda$ is
%\regvary\ with index $\alpha/\psi$.
The \rv\ $\sigma_{\Lambda^c}$ is independent of $\sigma_\Lambda$. Similarly to
Mikosch and Rezapour \cite{mikosch:rezapour:2013} (see also the end of this proof for a similar argumentation) one can show that 
$\sigma_{\Lambda^c}$ has moment of order $\alpha+\vep$ 
for sufficiently small positive $\vep$. Therefore, by Breiman's lemma \cite{breiman:1965},
\beao
\P(\sigma >x)\sim \E [\sigma_{\Lambda^c}^{\alpha}]\,\P(\sigma_\Lambda>x)\,,\qquad \xto\,.
\eeao
This proves \regvar\ with index $\alpha$ of the marginal \ds s of $(\sigma_{it})$.
\par
In the remainder of the proof we focus on (3).
%Statement (2) is then a special case of (3) with $d=1$ and statement (1) follows by the non-negativity of $\sigma$ from (2) by setting $i=j$. 
%In the following let $d\geq 1$ and set $\mathbf{i}=\{i_1, \ldots, i_d\}, \mathbf{j}=\{j_1, \ldots, j_d\},$ $\mathbf{t}=\{t_1, \ldots, t_d\}$ all be subsets of $\mathbb{Z}$. Define
% $$ \psi^{\mathbf{i},\mathbf{j}}= \max_{1 \leq l \leq d} \psi^{i_l,j_l}.$$
For a given choice of $\mathbf{i, j, t}\in \bbz^d$, we write
% $$ \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^{(l)}=\{m,n: \psi_{m+i_l,n+t_l}+\psi_{m+j_l,n+t_l}= \psi^{\mathbf{i},\mathbf{j}}\}, \;\;\; 1 \leq l \leq d, $$
%and 
\begin{equation}\label{Eq:Lambda:set:def} \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}=\{(m,n ): \psi_{i_l-m,t_l-n}+\psi_{j_l-m,t_l-n}= \psi^{\mathbf{i},\mathbf{j}}\; \mbox{for some } 1\leq l \leq d\}.\end{equation}
We will show that the random vector $(\sigma_{i_1,t_1}\sigma_{j_1,t_1}, \ldots, \sigma_{i_d,t_d}\sigma_{j_d,t_d})=:\boldsymbol{\sigma}'$ is regularly varying with index $\alpha/\psi^{\mathbf{i},\mathbf{j}}$ which proves (3). Note that
\begin{eqnarray*}
\sigma_{i,t}\sigma_{j,t}&=&\prod_{(k,l)}\exp(\psi_{kl}\eta_{i-k,t-l})\prod_{(k',l' )}\exp(\psi_{k'l'}\eta_{j-k',t-l'}) \\
&=& \prod_{(m,n )}\exp((\psi_{i-m,t-n}+\psi_{j-m,t-n})\eta_{m,n})
\end{eqnarray*}
and write
\begin{align} \nonumber \boldsymbol{\sigma}&= \mbox{diag}\left(\left(
  \begin{matrix}
  \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c}\ex^{\eta_{m,n}(\psi_{i_1-m,t_1-n}+\psi_{j_1-m,t_1-n})} \\ \vdots \\ \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c}\ex^{\eta_{m,n}(\psi_{i_d-m,t_d-n}+\psi_{j_d-m,t_d-n})}
  \end{matrix}
\right)'\right)\cdot \left(\begin{matrix}
                   \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}}\ex^{\eta_{m,n}(\psi_{i_1-m,t_1-n}+\psi_{j_1-m,t_1-n})} \\ \vdots \\ \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}}\ex^{\eta_{m,n}(\psi_{i_d-m,t_d-n}+\psi_{j_d-m,t_d-n})}
                   \end{matrix}
 \right)\\
\label{Eq:factors:sigma} &=: \mathbf{A}\, \mathbf{Z},
\end{align}
where ${\rm diag}((a_1,\ldots,a_k))$ is any diagonal matrix with diagonal elements $a_1,\ldots,a_k$. 
We notice that $\bfA$ and $\bfZ$ are independent.
\par
Consider iid copies $(Y_j)$ of $\ex^\eta$. There exist suitable numbers $(a_{ij})_{1\le i\le d,1\le j\le p}$ with $p= |\Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}|$
\st\ the components of $\mathbf{Z}$ have \rep\ in \ds\ 
$\prod_{j=1}^{p}Y_j^{a_{ij}}$, $1 \leq i \leq d $. By assumption, $Y_j$ is regularly varying with index $\alpha$ and
satisfies either assumption \eqref{Eq:conv:equivalent} or $\E[Y_j^\alpha]=\infty$. 
Furthermore, for each $j$ there exists one $1\leq i \leq d$ 
such that $a_{ij}=a_{\max}=\psi^{\mathbf{i},\mathbf{j}}$ by the definition of $\Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}$. 
An application of Proposition \ref{Pr:genRVforproducts} shows that $\bfZ$ is \regvary\ 
with index $\alpha/\psi^{\mathbf{i},\mathbf{j}}$ and limit measure $\mu_{\mathbf{Z}}$ which is given as $\mu$ in Proposition \ref{Pr:genRVforproducts} (ii) (if \eqref{Eq:inf:mom:eta} holds) or Proposition \ref{Pr:genRVforproducts} (i) (if \eqref{Eq:conv:eq:eta} holds). Now, choose $\epsilon, \delta>0$ such that
$$\frac{\psi_{i_l-m,t_l-n}+\psi_{j_l-m,t_l-n}}{\psi^{\mathbf{i},\mathbf{j}}}(1+\delta)<1-\epsilon, \;\;\;  (m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c,\;\;1\leq l \leq d,$$
which is possible by the definition of $\Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}$ and the summability constraint on the coefficients. Then we have 
\begin{eqnarray*}&&\E\left[\|\mathbf{A}\|_{\mbox{\scriptsize op}}^{\alpha(1+\delta)/\psi^{\mathbf{i},\mathbf{j}}}\right]\\
&\leq & \sum_{l=1}^d \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c} \E\left[\ex^{\eta_{m,n}\alpha(1+\delta)(\psi_{i_l-m,t_l-n}+\psi_{j_l-m,t_l-n})/\psi^{\mathbf{i},\mathbf{j}}}\right]\\
&\leq& \sum_{l=1}^d \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c} \E\left[\ex^{\eta_{m,n}\alpha(1-\epsilon)}\right]^{(1+\delta)(\psi_{i_l-m,t_l-n}+\psi_{j_l-m,t_l-n})/((1-\epsilon)\psi^{\mathbf{i},\mathbf{j}})} <\infty,
\end{eqnarray*}
where we used Jensen's inequality for the penultimate step and the 
summability condition of the coefficients for the final one. Thus we have verified all conditions of the 
multivariate Breiman lemma in Basrak et al.~\cite{basrak:davis:mikosch:2002a}, implying that
$\boldsymbol{\sigma}$ inherits \regvar\  from $\bfZ$ with corresponding
index $\alpha/\psi^{\mathbf{i},\mathbf{j}}$ and limit measure
$\mu_{\boldsymbol{\sigma}}(\cdot)=\E[\mu_{\mathbf{Z}}(\mathbf{A}^{-1}\cdot)]$.
%with $\mathbf{A}$ as in \eqref{Eq:factors:sigma}.
\end{proof}
\begin{proposition}\label{prop:rv:2} Assume that the aforementioned conditions (including either \eqref{Eq:inf:mom:eta} or \eqref{Eq:conv:eq:eta}) hold and that in addition $\E[|Z|^{\alpha+\delta}]<\infty$ for some $\delta>0$. Then the following statements hold: 
\begin{enumerate}
\item
Each of the \seq s 
$(X_{it})_{t \in \mathbb{Z}}$, $i \in \mathbb{Z}$, is \regvary\ with index $\alpha$.\\[1mm]
If  \eqref{Eq:inf:mom:eta} holds then the corresponding spectral 
tail process satisfies $\Theta_t^{i}=0$ a.s., $t\ge 1$, and
$ \P(\Theta^{i}_0=\pm 1)=\E[Z_{\pm}^{\alpha}]/\E[|Z|^{\alpha}]$.\\[1mm]
 %\; \P(\theta_0^{(i)}=-1)=\frac{\E[(Z^-)^{\alpha}]}{\E[|Z|^{\alpha}]}
%\qquad \bfTh_h^{i}=0 \;\as,  h \geq 1\,.
If \eqref{Eq:conv:eq:eta} holds, then for any Borel set $B=B_0\times\cdots \times B_n\subset \bbr^{n+1}$,
\begin{align}\label{Eq:theta:law:1}
  \P((\Theta^{i}_t)_{t=0,\ldots,n}
  \in B)=\sum_{(u,v) \in \Lambda_i^{(0)}}\frac{1}{|\Lambda_i^{(0)}
    |}\frac{\E\left[\I\left(\left(\I((u,v) \in
            \Lambda_i^{(t)})\frac{X_{it}}{|X_{i0}|}\right)_{t=0, \ldots,
          n} \in B
      \right)|X_{i0}|^{\alpha}\right]}{\E[|X_{i0}|^{\alpha}]},
\end{align}
where $\Lambda_i^{(t)}=\{(u,v):\psi_{i-u,t-v}=1 \}, t=0, \ldots, n$. 
\item
Each of the \seq s $(X_{it}X_{jt})_{t \in \mathbb{Z}}$, $i,j\in \mathbb{Z}$, is \regvary\ with index  $\alpha/\psi^{ij}$.\\[1mm] 
If \eqref{Eq:inf:mom:eta} holds then the corresponding spectral tail process satisfies $\Theta_t^{ij}=0$ a.s., $t\ge 1$, and
$\P(\Theta^{ij}_0=\pm 1)=\E[(Z_iZ_j)_{\pm}^{\alpha/\psi^{ij}}]/\E[|Z_iZ_j|^{\alpha/\psi^{ij}}]$.\\[1mm]
%, \; \P(\theta_0^{(ij)}=-1)=\frac{\E[((Z_iZ_j)^-)^{\alpha/\psi^{ij}}]}{\E[|Z_iZ_j|^{\alpha/\psi^{ij}}]}, $$
If \eqref{Eq:conv:eq:eta} holds, then for any Borel set $B=B_0\times\cdots \times B_n\subset \bbr^{n+1}$, 
\begin{align}  
&\P(( \Theta^{ij}_t)_{t=0,\ldots,n} \in B) \nonumber \\
 \label{Eq:theta:law:2}=&\sum_{(u,v) \in
   \Lambda_{i,j}^{(0)}}\frac{1}{|\Lambda_{i,j}^{(0)}
   |}\frac{\E\left[\I\left(\left(\I((u,v) \in
         \Lambda_{i,j}^{(t)})\frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0,
         \ldots, n} \in B
     \right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]}{\E[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}]},
\end{align}
where $\Lambda_{i,j}^{(t)}=\{(u,v):\psi_{i-u,t-v}+\psi_{j-u,t-v}=\psi^{ij}\}, t=0, \ldots, n$. 
%and the random variables $\ex^{\eta_{u,v}^{(u,v)}}$,$(\ex^{\eta_{l,m}^{(u,v)}})_{(l,m)\neq(u,v)}$, $(\tilde{Z}_{i,t},\tilde{Z}_{j,t})_{t=0, \ldots, n}$ have joint $\nu_{\alpha} \otimes (\otimes_{(l,m)\neq(u,v)} P^{\ex^{\eta}}) \otimes (\otimes_{t=0, \ldots, n}P^{(Z_{i},Z_{j})})$-density 
%\begin{eqnarray*}&& f^{(u,v)}(y_{u,v},(y_{l,m})_{(l,m) \neq (u,v)}, (z_{i,t},z_{j,t})_{t=0, \ldots, n})\\
%&=&\frac{\I(\prod_{(l,m)}y_{l,m}^{\psi_{i-l,-m}+\psi_{j-l,-m}}|z_{i,0}z_{j,0}|>1)}{\E[(\prod_{(l,m)\neq(u,v)}\ex^{\eta_{l,m}^{(u,v)}(\psi_{i-l,-m}+\psi_{j-l,-m})}|\tilde{Z}_{i,0}\tilde{Z}_{j,0}|)^{\alpha/\psi^{ij}}]},
%\end{eqnarray*}
\item
For $d\geq 1$ and $\mathbf{i},\mathbf{j}\in \bbz^d$, the $d$-variate \seq\ $((X_{i_kt}X_{j_kt})_{1\leq k \leq d})_{t \in \mathbb{Z}}$ 
is jointly regularly varying with index $\alpha/\psi^{\mathbf{i},\mathbf{j}}$.
\end{enumerate}
\end{proposition}
\bre\label{rem:77} \begin{enumerate}
      \item Equation \eqref{Eq:theta:law:1} shows that in this case the distribution 
of $(\Theta^{i}_t)_{t \geq 0}$ is a mixture of $|\Lambda_i^{(0)}|$ distributions, where 
each distribution gets the weight $1/|\Lambda_i^{(0)}|$. 
Heuristically speaking, a distribution in this mixture that corresponds 
to a specific $(u,v) \in \Lambda_i^{(0)}$ has interpretation as the distribution of $(X_{it}/|X_{i0}|)_{t \geq 0}$, given that we have seen an extreme observation of $|X_{i0}|$ caused by an extreme realization of $\ex^{\eta_{u,v}}$.
The variables $\ex^{\eta_{u,v}}$, $(u,v) \in \Lambda_i^{(0)}$, are those which have a maximum exponent (equal to 1) 
in the product $\prod_{(u,v)}\exp(\psi_{i-u,-v}\eta_{u,v})=\sigma_{i0}$. They are therefore the factors which are most likely 
to make $\sigma_{i0}$, hence $X_{i0}$, extreme. 
\par      
An analogous interpretation can be derived from \eqref{Eq:theta:law:2} for the distribution of $(\Theta^{ij}_t)_{t \geq 0}$. 
      \item Note that for fixed $i, j$, the inner indicator functions in \eqref{Eq:theta:law:1} and \eqref{Eq:theta:law:2} 
are positive only for finitely many $t$. Hence there are only finitely many $t\geq 1$ such that 
$\P(\Theta_t^{i}\neq 0)>0$ and $\P(\Theta_t^{(ij)}\neq 0)>0$.
      \item Using similar techniques as in the proof of cases (1) and (2) below, one can
also give an explicit expression for the resulting $d$-dimensional spectral tail process 
of $((X_{i_kt}X_{j_kt})_{1\leq k \leq d})_{t \in \mathbb{Z}}$ in (3). 
However, due to its complexity, we refrain from stating it here.
     \end{enumerate}
\ere
\begin{proof} 
We start by showing that all mentioned sequences 
are regularly varying. Exemplarily, we show this for case (2). 
Very similar arguments can be used for the two other cases.
For $n \geq 0$ write 
$$ \left(X_{it}X_{jt}\right)'_{t=0, \ldots, n}=\mbox{diag}\left((Z_{it}Z_{jt})_{t=0, \ldots, n}'\right)\cdot
\left(\sigma_{it}\sigma_{jt}\right)'_{t=0, \ldots, n}\,.$$
Since $\psi^{ij} \geq 1$ our moment assumption on $Z$ implies that 
$\E[|Z|^{\alpha/\psi^{ij}+\delta}]<\infty$ for some $\delta>0$. Then  Proposition \ref{prop:rv:1} 
allows us to apply the aforementioned multivariate Breiman lemma, yielding  the  
regular variation of the vector $(X_{it}X_{jt})_{t=0, \ldots, n}$ with index $\alpha/\psi^{ij}$. From the first definition given in Section \ref{subsec:rvts}, this implies the regular variation of the sequence.
\par
As for the derivation of the explicit form of the spectral tail process in (1) and (2), we restrict ourselves to derive the distribution of the spectral tail process $(\Theta_t^{ij})_{t \geq 0}$ in part (2); part (1) is similar.
\par
If $\mu_n^{{\boldsymbol{\sigma}}^{ij}}$ denotes the vague limit measure of 
$(\sigma_{i,0}\sigma_{j,0}, \ldots, \sigma_{i,n}\sigma_{j,n})'$ the multivariate Breiman lemma yields 
the vague limit measure $\mu_n^{\bfX^{ij}}$ of $(X_{i,0}X_{j,0}, \ldots, X_{i,n}X_{j,n})'$ given by
\begin{eqnarray}\nonumber \mu_n^{\bfX^{ij}}(B)&=& c\,\E\left[\mu_n^{\boldsymbol{\sigma}^{ij}}(\times_{t=0}^n (B_t/(Z_{it}Z_{jt})))\right] \\
&=& \label{Eq:mu:X} c\,\E\left[\tilde{\mu}_n^{\boldsymbol{\sigma}^{ij}}\left(\times_{t=0}^n \left(B_t \bigg/ \left(Z_{it}Z_{jt}\prod_{(u,v) \in \Lambda_{i,j,n}^c}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}\right)\right)\right)\right] 
\end{eqnarray}
for any $\mu_n^{\bfX^{ij}}$-continuity  Borel set $B=\times_{t=0}^n B_t \in [-\infty,\infty]^{n+1} \setminus \{\mathbf{0}\}$ bounded away from $\mathbf{0}$, $\Lambda_{i,j,n}$ is equal to $\Lambda_{\mathbf{i}, \mathbf{j}, \mathbf{t}}$ 
as defined in \eqref{Eq:Lambda:set:def} with $\mathbf{i}=(i, \ldots, i), \mathbf{j}=(j, \ldots, j), \mathbf{t}=(0, \ldots ,n)$, and $\tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n$ 
is the limit measure of the \regvary\ vector
\beam \label{Eq:prod:sigma:tilde}
\Big(\prod_{(u,v) \in \Lambda_{i,j,n}}\ex^{\eta_{u,v}\,(\psi_{i-u,t-v}+\psi_{j-u,t-v} )}\Big)_{t=0, \ldots, n}\,,
\eeam
 see the proof of 
Proposition~\ref{prop:rv:1}. The distribution of the tail process of $(X_{it}X_{jt})$ (cf.\ Section \ref{subsec:rvts}) is then determined by
\begin{eqnarray}\label{Eq:theta:law} \P((Y \Theta^{ij}_t)_{t=0, \ldots, n} \in B)&=&\lim_{x \to \infty} \frac{\P((X_{it}X_{jt}/x)_{t=0, \ldots, n} \in B, |X_{i0}X_{j0}|/x>1)}{\P(|X_{i0}X_{j0}|/x>1)}\\
\nonumber &=&\frac{\mu_n^{\bfX^{ij}}(B \cap \big([-\infty,\infty]\backslash [-1,1]\times 
[-\infty,\infty]^n)\big)}{\mu_n^{\bfX^{ij}}\big([-\infty,\infty]\backslash [-1,1]\times [-\infty,\infty]^n\big)}\,.
\end{eqnarray}
The concrete forms of $\tilde{\mu}_n^{\boldsymbol{\sigma}^{ij}}$, hence of $\mu_n^{\mathbf{X}^{ij}}$, 
now depend on whether \eqref{Eq:inf:mom:eta} or \eqref{Eq:conv:eq:eta} holds.
\par
We first assume \eqref{Eq:inf:mom:eta}. Note that
$\Lambda_{i,j,n}=\cup_{t=0}^n \Lambda_{i,j}^{(t)},$ where $\Lambda_{i,j}^{(t)}=\{(u,v): \psi_{i-u,t-v}+\psi_{j-u,t-v}=\psi^{ij}\}$. Indeed, 
we easily see that $\Lambda_{i,j}^{(t)}=\Lambda_{i,j}^{(0)}+(0,t)$, $t=1, \ldots, n$. 
We apply Proposition~\ref{Pr:genRVforproducts}(ii) 
to derive the specific form of the limit measure $\tilde{\mu}_n^{\boldsymbol{\sigma}^{ij}}$ of \eqref{Eq:prod:sigma:tilde}. 
Each component of this vector contains $|\Lambda_{i,j}^{(0)}|$ factors 
with maximal exponent $\psi^{ij}$. For the $t$-th component, those are the factors 
$\exp(\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v}))$, $(u,v) \in \Lambda_{ij}^{(t)}$. Hence $p_{\mbox{\scriptsize eff}}=|\Lambda_{i,j}^{(0)}|$ and 
$P_{\mbox{\scriptsize eff}}=\{\Lambda_{i,j}^{(0)}+(0,t), t=0, \ldots, n\}$. By \eqref{Eq:RV:prod:2}, 
the measure $\tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n$, up to a constant multiple, is given by
\begin{eqnarray*}
 \tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n(B)&=& c \sum_{s=0}^n \int_0^\infty \P\bigg(\bigg(\I(\psi_{i-u,t-v}+\psi_{j-u,t-v}=\psi^{ij} \, \forall\, (u,v) \in \Lambda_{i,j}^{(s)})z^{\psi^{ij}}\\
 && \hspace{1.7cm} \prod_{(u,v) \in \Lambda_{i,j,n} \setminus \Lambda_{i,j}^{(s)}}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}\bigg)_{0 \leq t \leq n} \in B\bigg)\nu_{\alpha}(dz)\\
 &=& c \sum_{s=0}^n \int_0^\infty \P\bigg(\bigg(\I(t=s)z^{\psi^{ij}}\prod_{(u,v) \in \Lambda_{i,j,n} \setminus \Lambda_{i,j}^{(s)}}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}\bigg)_{0 \leq t \leq n} \in B\bigg)\nu_{\alpha}(dz),
\end{eqnarray*}
where $\nu_\alpha(dx)=\alpha x^{-\alpha-1}dx$. 
The $s$-th measure in the sum above is concentrated on the $s$-th axis. Therefore the limit measure $\tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n$ is
concentrated on the axes.
By \eqref{Eq:mu:X}, this implies that $\mu_n^{\bfX^{ij}}$ is concentrated on the axes as well. Therefore
$\mu_n^{\bfX^{ij}}(B \cap ([-\infty,\infty]\backslash [-1,1])\times [-\infty,\infty]^n)=0$ as soon as one $B_i, 1 \leq i \leq n,$ in $B=\times_{i=0}^n B_i$ is bounded away from 0. 
With \eqref{Eq:theta:law} this gives $Y \Theta_t^{ij}=0$ a.s. for $t \geq 1$ and therefore $\Theta_t^{ij}=0$ a.s. for $t \geq 1$. 
The law of $\Theta_0^{ij}$ follows from the univariate Breiman lemma. 
\par
Next assume \eqref{Eq:conv:eq:eta}. By Proposition~\ref{Pr:genRVforproducts}(i),  
the vague limit measure $\tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n$ is up to a constant given by 
\begin{eqnarray*} && \tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n(B) \\
&=&\hspace{-0.3cm} \sum_{(u,v) \in \Lambda_{i,j,n}}\int_0^\infty \P\bigg(\bigg(\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}\prod_{\substack{(\tilde{u},\tilde{v}) \in \Lambda_{i,j,n} \\ (\tilde{u},\tilde{v}) \neq (u,v)}}\ex^{(\psi_{i-\tilde{u},t-\tilde{v}}+\psi_{j-\tilde{u},t-\tilde{v}})\eta_{\tilde{u},\tilde{v}}}\bigg)_{t=0, \ldots ,n} \in B \bigg)\nu_{\alpha}(dz).\end{eqnarray*}                                                                                                
For sets $B$ such that $B \cap (\{0\} \times [-\infty,\infty]^n)=\emptyset$ it suffices thereby to sum only 
over  $(u,v) \in \Lambda_{i,j}^{(0)}$ instead over all $(u,v) \in \Lambda_{i,j,n}=\cup_{t=0}^n\Lambda_{i,j}^{(t)}$.
For these sets we have by Breiman's lemma (cf.\ \eqref{Eq:mu:X}),
\begin{align*} & \mu^{\bfX^{ij}}_n(B)/c\\
= &\sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\int\limits_0^\infty \P\bigg((\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}\prod\limits_{(\tilde{u},\tilde{v}) \neq (u,v)}\ex^{(\psi_{i-\tilde{u},t-\tilde{v}}+\psi_{j-\tilde{u},t-\tilde{v}})\eta_{\tilde{u},\tilde{v}}} Z_{it}Z_{jt})_{t=0, \ldots ,n} \in B \bigg)\nu_{\alpha}(dz)\\
= &\sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\int\limits_0^\infty \P\bigg((\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}X_{it}X_{jt}\ex^{-\psi^{ij}\eta_{u,v}})_{t=0, \ldots ,n} \in B \bigg)\nu_{\alpha}(dz),\end{align*}
where we used that if $(u,v) \in \Lambda_{i,j}^{(t)},$ then 
$$ \prod_{(\tilde{u},\tilde{v}) \neq (u,v)}\ex^{(\psi_{i-\tilde{u},t-\tilde{v}}+\psi_{j-\tilde{u},t-\tilde{v}})\eta_{\tilde{u},\tilde{v}}}= \frac{\sigma_{it}\sigma_{jt}}{\ex^{(\psi_{i-u,t-v}+\psi_{j-u,t-v})\eta_{u,v}}}=\frac{\sigma_{it}\sigma_{jt}}{\ex^{\psi^{ij}\eta_{u,v}}}. $$  
Fubini's Theorem and a substitution finally simplify this expression to 
\begin{eqnarray*} &&  \sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_0^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}X_{it}X_{jt}\ex^{-\psi^{ij}\eta_{u,v}}\right)_{t=0, \ldots, n} \in B\right)\nu_\alpha(dz)\right]  \\
&=& \sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_0^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\nu_{\frac{\alpha}{\psi^{ij}}}(dy)\right].
\end{eqnarray*}
Note that the range of the inner integral in the last expression can be changed 
from $(0, \infty)$ to $(1,\infty)$, if $B \cap [-1,1] \times [-\infty,\infty]^n = \emptyset$. 
Therefore, by writing
\beao
\tilde{B}_0=B_0 \setminus [-1,1] \,, \quad \tilde{B}_t=B_t, \quad t \geq 1,\quad \tilde B=
\times_{t=0}^n\tilde B_t\,,
\eeao 
we get from \eqref{Eq:theta:law} that
\begin{eqnarray*}
\lefteqn{P\big((Y \Theta^{ij}_t)_{t=0,\ldots,n} \in B\big)}\\&=&\frac{\mu^{\bfX^{ij}}_n(\tilde{B})}{\mu^{\bfX^{ij}}_n(
([-\infty,\infty]\backslash [-1,1])\times [-\infty,\infty]^n)}\\
&=& \frac{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_1^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\nu_{\frac{\alpha}{\psi^{ij}}}(dy)\right]}{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\right]} \\
&=& \sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\frac{1}{| \Lambda_{i,j}^{(0)}|}\frac{\E\left[\I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})Y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]}{\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]},
\end{eqnarray*}
where $Y$ is a Pareto($\alpha/\psi^{ij}$) random variable, independent of all other random variables in the expression. For the last equation, we expanded both numerator and denominator by multiplying with $\E(e^{\alpha \eta_{u,v}})$, noting that for $(u,v) \in \Lambda_{i,j}^{(0)}$ the random variable $e^{\alpha \eta_{u,v}}$ is independent both of the indicator function and of $|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}e^{-\alpha \eta_{u,v}}$. From the law of the tail process $(Y \Theta_t^{ij})$ we can now see that the law of the spectral tail process $(\Theta_t^{ij})$ satisfies \eqref{Eq:theta:law:2}.
\end{proof}

\subsection{Infinite variance stable limit theory for the stochastic volatility model and its product processes}\label{subsec:limit2}
In the following result we provide central limit theory with infinite variance stable limits for the sums $S_{ij}$; see \eqref{eq:esses}.
\bth\label{thm:1}
We consider the stochastic volatility model \eqref{eq:1a:anja} and assume the special form of $(\sigma_{it})$ given in \eqref{eq:sv2model} with $\psi=1$. 
For given $(i,j)$, define a \seq\ $(b_n)$ \st\ 
$n\,\P(|X_{i0}X_{j0}|>b_n)\to 1$ as $\nto$.
Assume the following conditions:
\begin{enumerate}
\item 
The conditions of Proposition~\ref{prop:rv:2} hold, ensuring that $\E[|Z|^{\alpha/\psi^{ij}+\vep}]<\infty$ for some $\vep>0$ and
$(X_{it}X_{jt})$ is \regvary\ with index $\alpha/\psi^{ij}$ and spectral tail process $(\Theta_h^{ij})$.
%\item 
%$\E[|Z|^{\alpha/\psi^{ij}+\vep}]<\infty$ for some $\vep>0$ {\red(can be deleted since $\E[|Z|^{\alpha+\vep}]<\infty$ is already assumed in Proposition~\ref{prop:rv:2})},
\item
$(\sigma_{it}\sigma_{jt})$ is $\alpha$-mixing with rate \fct\ $(\alpha_h)$ and there exists $\delta>0$ such that $\alpha_n=o(n^{-\delta})$.
\item Either 
\item[(i)]$\alpha/\psi^{ij}< 1$, or 
\item[(ii)] 
$i\ne j$, $\alpha/\psi^{ij}\in [1,2)$ and $Z$ is symmetric, or 
\item[(iii)] $i=j$,
$\alpha/\psi^{ii}=\alpha/2 \in (1,2)$ and the mixing rate in (2)
satisfies 
$\sup _n n\,\sum_{h=r_n}^\infty \alpha_h<\infty$ for some integer \seq\ $(r_n)$ \st\ $nr_n/b_n^2\to 0$ as $\nto$.
\end{enumerate}
Then 
\beam\label{eq:rra}
b_n^{-1}(S_{ij}-c_n) \std \xi_{ij,\alpha/\psi^{ij}}\,,
\eeam
where $\xi_{ij,\alpha/\psi^{ij}}$ is a totally skewed to the right $\alpha/\psi^{ij}$-stable \rv\ and 
\beao
c_n=\left\{\barr{ll}n\, \E [X^2]& \mbox{$i=j$ and $\alpha\in (2,4)$}\,,\\
0& i\ne j \mbox{ or } \alpha/\psi^{ij}<1 \,,\earr\right.
\eeao
\ethe
\bre\label{reM:c1}
\begin{enumerate}
%Using Breiman's result and the moment condition $\E [|Z|^{\alpha/\psi^{ij}+\vep}]<\infty$ for some $\vep>0$,
%\regvar\ of $(\sigma_{it}\sigma_{jt})$ with index $\alpha/\psi^{ij}$ implies \regvar\ of $(X_{it}X_{jt})$ with the same index. 
%Sufficient conditions for \regvar\ of $(X_{it}X_{jt})$ with index $\alpha/\psi^{ij}$ are given in Proposition~\ref{prop:rv:2}.
\item
If $(\alpha_h)$ decays  at an exponential rate one can choose $r_n= C\log n$ for a sufficiently large constant $C$. Then
  $\sup _n n\,\sum_{h=r_n}^\infty \alpha_h<\infty$ and $nr_n/b_n^2\to 0$ hold. These conditions are also satisfied if
$\alpha_h\le c n^{-(1+\gamma)}$ for some $\gamma>0$, $r_n=C n^\xi$  for some
$\xi>0$ and $1/\gamma\le \xi< 2\psi^{ij}/\alpha-1$.
\item
The \seq\ $(X_{it}X_{jt})$ inherits $\alpha$-mixing from $(\sigma_{it}\sigma_{jt})$; see Remark~\ref{rem:mix}.
\item It is possible to prove joint \con\ for $1\le i,j\le p$ in \eqref{eq:rra}. Due to different tail behavior for
distinct $(i,j)$ the normalizing \seq s $(b_n)=(b_n^{ij})$ typically increase to infinity at different rates. Then it is only of interest
to consider the joint \con\ of those $S_{ij}$ whose summands $X_{it}X_{jt}$ have the same tail index $\alpha/\psi^{ ij}$. More
precisely, it suffices to consider those $S_{ij}$ with the property that $X_{it}X_{jt}$ is tail-equivalent to $X_{it}^2$. 
The joint \con\ follows in a similar way as in the proof below, by observing
that Theorem~\ref{thm:mikwin} is a multivariate limit result. 
The joint limit of $S_{ij}$ in \eqref{eq:rra} with equivalent tails of index $\tilde{\alpha}$ (say) is
jointly $\tilde{\alpha}$-stable with possible dependencies in the limit vector. 
\item
The strongest normalization is needed for $S_i=S_{ii}$. Recall that the summands $X_{it}^2$ of $S_i$ are \regvary\ with index 
$\alpha/2$, i.e., $\psi^{ii}=2$. Let $(a_n)$ be \st\ $n\,\P(|X|>a_n)\to 1$.
Under the conditions of Theorem~\ref{thm:1}, we have that $a_n^{-2}(S_i-c_n)\std \xi_{i,\alpha/2}$, $i=1,\ldots,p$ 
for a jointly $\alpha/2$-stable limit.  If
$\alpha/2<\alpha/\psi^{ij}$ for some $i\ne j$, then $b_n/a_n^2\to 0$, hence $a_n^{-2} S_{ij}\stp 0$. It is possible that $X_{it}X_{jt}$
is regularly varying with index $\alpha/2$ but nevertheless $b_n/a_n^2\to 0$; see Example~\ref{exam:1} 
which deals with the case $\E [\ex^{\alpha \eta}]=\infty$.
\end{enumerate}
\ere
\begin{proof}
We apply Theorem~\ref{thm:mikwin} to the \seq\ $(X_{it}X_{jt})$, cf.\ also Remark~\ref{rem:case:alpha:1}.\\[2mm] 
(1) The \regvar\ condition on $(X_{it}X_{jt})$ with index $\alpha/\psi^{ij}$ is satisfied by assumption. Moreover, $\Theta_h=0$ for 
sufficiently large $h$; see Remark~\ref{rem:77}.\\[2mm]
(2) The assumption about the mixing coefficients in condition (2) implies that for a 
sufficiently small $\varepsilon \in (0,1)$ and $m_n=n^{1-\varepsilon}$ there exists an integer sequence $l_n=o(m_n)$ such that $k_n \alpha_{l_n} \to 0$. For this choice of $m_n$ and $l_n$, the proof of the mixing condition for the sums of the truncated variables
\beao
\un S_{ij} =\sum_{t=1}^n X_{it}X_{jt} \I( |X_{it}X_{jt}|>\vep b_n)
\eeao
is now analogous to the proof of the corresponding property in Theorem~\ref{the:1}. \\[2mm]
(3) We want to show that
\beam\label{eq:ac}
\lim_{l\to\infty}\limsup_{\nto} n\,\sum_{t=l}^{m_n}\,\P\big(|X_{it}X_{jt}|>b_n\,,|X_{i0}X_{j0}|>b_n\big)=0\,
\eeam
for $m_n=n^{1-\varepsilon}$ as above.
Write
$$ 
\sigma_{it}\sigma_{jt}= \prod_{(m,n)}\exp((\psi_{i-m,t-n}+\psi_{j-m,t-n})\eta_{m,n})
$$
and set $\Lambda_{\varepsilon,t}=\{(m,n): \psi_{i-m,t-n}+\psi_{j-m,t-n}\geq 8^{-1}\psi^{ij}\varepsilon\}$, $t \in \mathbb{Z}$. 
Without loss of generality we assume that $l$ is so large that $\Lambda_{\varepsilon,t} \cap \Lambda_{\varepsilon,0}$ is empty
for all $t \geq l$. Then write for $t \geq l$, 
\beao
\sigma_{it}\sigma_{jt}=
\sigma_{it,jt,\Lambda_{\varepsilon,t}}\cdot \sigma_{it,jt,\Lambda_{\varepsilon,0}} \cdot \sigma_{it,jt,\Lambda_{\varepsilon,0,t}^c},\;\;\; \sigma_{i0}\sigma_{j0}=\sigma_{i0,j0,\Lambda_{\varepsilon,0}}\cdot \sigma_{i0,j0,\Lambda_{\varepsilon,t}} \cdot \sigma_{i0,j0,\Lambda_{\varepsilon,0,t}^c},\eeao
where
$$\sigma_{it_1,jt_1,\Lambda_{\varepsilon,t_2}}= \prod_{(m,n) \in \Lambda_{\varepsilon,t_2}}\exp((\psi_{i-m,t_1-n}+\psi_{j-m,t_1-n})\eta_{m,n}).$$
We conclude that $(\sigma_{it,jt,\Lambda_{\varepsilon,t}}$,$\sigma_{it,jt,\Lambda_{\varepsilon,0}}$, $\sigma_{i0,j0,\Lambda_{\varepsilon,0}}$, $\sigma_{i0,j0,\Lambda_{\varepsilon,t}})$ and $(\sigma_{it,jt,\Lambda_{\varepsilon,0,t}^c}$,$\sigma_{i0,j0,\Lambda_{\varepsilon,0,t}^c})$
are independent. We have
\beao\lefteqn{
\P\big(|X_{it}X_{jt}|>b_n\,,|X_{i0}X_{j0}|>b_n\big)}\\
&\le & \P\big(\max(|Z_{i0}Z_{j0}|,|Z_{it}Z_{jt}|)\,\max(\sigma_{it,jt,\Lambda_{\varepsilon,0,t}^c},\sigma_{i0,j0,\Lambda_{\varepsilon,0,t}^c})\, \\&&
\min(\sigma_{i0,j0,\Lambda_{\varepsilon,0}} \sigma_{i0,j0,\Lambda_{\varepsilon,t}},\sigma_{it,jt,\Lambda_{\varepsilon,t}}\sigma_{it,jt,\Lambda_{\varepsilon,0}})>b_n\big)\,.
%&\le &
%\P\big(M^2\max(\sigma_{i0,\Lambda_2^c}\sigma_{j0,\Lambda_2^c},\sigma_{it,\Lambda_2^c}\sigma_{jt,\Lambda_2^c})\, \min(\sigma_{i0,\Lambda}\sigma_{j0,\Lambda},\sigma_{it,\Lambda}\sigma_{jt,\Lambda})>b_n\big)\\
%&&+\P\big(\{\max(\sigma_{i0,\Lambda^c}\sigma_{j0,\Lambda^c},\sigma_{it,\Lambda^c}\sigma_{jt,\Lambda^c})\, \min(\sigma_{i0,\Lambda}\sigma_{j0,\Lambda},\sigma_{it,\Lambda}\sigma_{jt,\Lambda})>b_n\}\\&&\cap 
%\big(\{\sigma_{i0,\Lambda_1^c}\sigma_{j0,\Lambda_1^c}>M\}\cup \{\sigma_{it,\Lambda_1^c}\sigma_{jt,\Lambda_1^c}>M\}\big)=:J_{1t}+
%J_{2t}\,.
\eeao
The \ds\ of $\max(\sigma_{it,jt,\Lambda_{\varepsilon,0,t}^c},\sigma_{i0,j0,\Lambda_{\varepsilon,0,t}^c})$ 
is stochastically dominated uniformly for $t\geq l$ by a \ds\ which has moment of order $8 \alpha/(\psi^{ij}\varepsilon)>2 \alpha/\psi^{ij}$. Furthermore,
\begin{eqnarray*}
 && \min(\sigma_{i0,j0,\Lambda_{\varepsilon,0}} \sigma_{i0,j0,\Lambda_{\varepsilon,t}},\sigma_{it,jt,\Lambda_{\varepsilon,t}}\sigma_{it,jt,\Lambda_{\varepsilon,0}})\\
 &\le & \min\Big(\prod_{(m,n) \in \Lambda_{\varepsilon,0} \cup \Lambda_{\varepsilon,t}}\exp((\psi_{i-m,-n}+\psi_{j-m,-n})(\eta_{m,n})_+),\\
 && \hspace{1cm}\prod_{(m,n) \in \Lambda_{\varepsilon,0} \cup \Lambda_{\varepsilon,t}}\exp((\psi_{i-m,t-n}+\psi_{j-m,t-n})(\eta_{m,n})_+)\Big)\,\\
 &\le& \min\Big(\prod_{(m,n) \in \Lambda_{\varepsilon,0}}\exp(\psi^{ij}(\eta_{m,n})_+)\prod_{(m',n') \in \Lambda_{\varepsilon,t}}\exp(8^{-1}\psi^{ij}\varepsilon (\eta_{m',n'})_+),\\
 && \hspace{1cm}\prod_{(m',n') \in \Lambda_{\varepsilon,t}}\exp(\psi^{ij}(\eta_{m',n'})_+)\prod_{(m,n) \in \Lambda_{\varepsilon,0}}\exp(8^{-1}\psi^{ij}\varepsilon (\eta_{m,n})_+)\Big)\, \\
 &\leq& \min\Big(\prod_{(m,n) \in \Lambda_{\varepsilon,0}}\exp((\psi^{ij}+8^{-1}\psi^{ij}\varepsilon)(\eta_{m,n})_+),
 \prod_{(m,n) \in \Lambda_{\varepsilon,t}}\exp((\psi^{ij}+8^{-1}\psi^{ij}\varepsilon)(\eta_{m,n})_+)\Big).
 \end{eqnarray*}
The \rhs\ is \regvary\ with index $2\alpha/(\psi^{ij}(1+8^{-1}{\vep}))$.
A stochastic domination argument and an application of Breiman's lemma show that uniformly for $l\le t\le m_n$,
$$ m_n\, n\, \P\big(|X_{it}X_{jt}|>b_n\,,|X_{i0}X_{j0}|>b_n\big)=n^{2-\varepsilon} o\left(b_n^{-2\alpha/(\psi^{ij}(1+4^{-1}\epsilon))}\right)=n^{2-\varepsilon}o(n^{-2/(1+2^{-1}\varepsilon)})=o(1)$$
which yields \eqref{eq:ac}.\\[2mm]
(4) We check the vanishing small values condition. For any fixed $\delta$, we write
\beao
\ov {X_{it}X_{jt}}&=& X_{it}X_{jt} \I(|X_{it}X_{jt}|\le \delta b_n)\,,\qquad i\ne j\,,\\
\ov {X_{it}^2}&=& X_{it}^2 \I(X_{it}^2\le \delta b_n) - \E[X_{it}^2 \I(X_{it}^2\le \delta b_n)] \,,\\
\ov S_{ij}&=& \sum_{t=1}^n \ov {X_{it}X_{jt}}\,,\qquad \ov S_i=\ov S_{ii}\,.
\eeao
%Assume $\alpha/\psi^{ij}\in(0,1)$.
%We have by Karamata's theorem for any $\gamma>0$ as $\nto$.
%\beao
%\P(|\ov S_{ij}|>\gamma b_n)&\le &
%(\gamma b_n)^{-1} \E [\ov S_{ij}]\\&\le& n\,(\gamma b_n)^{-1}\,\E[|\ov {X_{it}X_{jt}}|]\\
%&\sim & \gamma^{-1} \delta^{1-\alpha/\psi^{ij}}\,,
%\eeao
%and the \rhs\ converges to zero for $\delta\downarrow 0$. {\red Can be deleted since index is less than 1 and vanishing small values condition does not need to hold.}
Assume $\alpha/\psi^{ij}\in [1,2)$, $i\ne j$. Then, by symmetry of the \rv s $Z_{it}$ and Karamata's theorem for any $\gamma>0$ as $\nto$,
$\E[\ov S_{ij}]=0$ and
\beao
\P(|\ov S_{ij}|>\gamma b_n)&\le &
(\gamma b_n)^{-2} \var(\ov S_{ij})\\
&=& n\,(\gamma b_n)^{-2} \E[(\ov {X_{it}X_{jt}})^2]\\
&\sim &\gamma^{-2}\,\delta^{2-\alpha}\,,
\eeao
and the \rhs\ converges to zero as $\delta\downarrow 0$.
\par
For $i=j$ and $\alpha/\psi^{ii}>1$ we need a different argument. 
We have by \v Cebyshev's inequality,
\beao
\P(|\ov S_{i} |>\gamma \,b_n)&\le & \gamma^{-2} b_n^{-2} \var\big(\ov S_{i}\big)\\
&=&  \gamma^{-2}\,(n/b_n^2)\,\sum_{|h|<n} (1-h/n)\,\cov(\ov {X_{i0}^2}, \ov {X_{ih}^2})\,.
\eeao
For $|h|\le h_0$ for any fixed $h_0$, $(n/b_n^2)  |\cov(\ov {X_{i0}^2}, \ov {X_{ih}^2})|$ vanishes by letting first $\nto$ and then $\delta\downarrow 0$.
This follows by Karamata's theorem. Standard bounds for the covariance \fct\ of an $\alpha$-mixing \seq\ 
(see Doukhan \cite{doukhan:1994}, p.~3)
yield
\beao
(n/b_n^2)\sum_{r_n\le |h|<n}|\cov(\ov {X_{i0}^2}, \ov {X_{ih}^2})| 
&\le & c\,\delta^2 n\,\sum_{r_n\le |h|<n}\alpha_h\,,
\eeao
where $r_n\to\infty$ is chosen \st\ $\sup_n\,n\,\sum_{r_n\le |h|<\infty} \alpha_h<\infty$ and $nr_n/b_n^2 \to 0$. The \rhs\ converges to zero by first letting $\nto$ and then
$\delta\downarrow 0$. It remains to show that
\beao
I_n=(n/b_n^2)\sum_{h_0< |h|\le r_n}(1-h/n)\,\cov(\ov {X_{i0}^2}, \ov {X_{ih}^2})
\eeao
is \asy ally negligible. We have
\beao
|I_n|&\le& (n/b_n^2)\sum_{h_0< |h|\le r_n} \E [X_{i0}^2\,X_{ih}^2\I(X_{i0}^2\le \delta b_n,X_{ih}^2\le \delta b_n)] + 
c\, n\,r_n/b_n^2\\
&\le & (n/b_n^2)\sum_{h_0< |h|\le r_n} \E [X_{i0}^2\,X_{ih}^2] +o(1)\,,
\eeao 
where we used that $n\,r_n/b_n^2\to 0$.
We will show that the summands on the \rhs\ are uniformly bounded by a constant if $h_0$ is sufficiently large. Then $\lim_{\nto} I_n=0$. 
\par
We observe that by H\"older's inequality,
\beao
\E [X_{i0}^2\,X_{ih}^2]&=& c\, \E [\sigma_{i0}^2\,\sigma_{ih}^2]\\
&=&c\,\E \big[
\ex^{2\sum_{(k,l)\in \Gamma_{\xi}}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}  \ex^{2\sum_{(k,l) \not \in \Gamma_\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]\\
&\le &c\,\big(\E \big[
\ex^{2r\sum_{(k,l)\in \Gamma_\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]\big)^{1/r} \big(\E\big[ \ex^{2s\sum_{(k,l) \not \in \Gamma_\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]\big)^{1/s}\,,
\eeao
where $\Gamma_\xi=\{(k,l): \psi_{ik}>\xi\}$ for some positive $\xi,s,t$ \st\ $1/r+1/s=1$. Since $\sigma_{i0}^2$ has moments up to order $\alpha/\psi^{ii}\in (1,2)$ and 
$(\eta_{i-k,-l})_{(k,l)\in\Gamma_\xi}$ and $(\eta_{i-k,h-l})_{(k,l)\in \Gamma_\xi}$ are independent for sufficiently large $h$ we can choose $r>1$ close to
one \st\ $\E \big[
\ex^{2r\sum_{(k,l)\in \Gamma\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]$ is finite. This implies that we choose $s$ sufficiently large. On the other hand,
for fixed $s$ we can make $\xi$ so small that $\E\big[ \ex^{2s\sum_{(k,l) \not \in \Gamma\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]$ is finite and uniformly bounded for
sufficiently large $h$. Fine tuning $\xi$ and $s$, we may conclude that $\lim_{\nto}I_n=0$ as desired.

By Theorem~\ref{thm:mikwin} and Remark~\ref{rem:case:alpha:1} the result now follows; see  also 
the end of the proof of Theorem~\ref{the:1} for the form of the resulting limit law.
\end{proof}

\begin{example}\label{exam:1}
We assume that $\E [\ex^{\alpha \eta}]=\infty$, hence $\ex^{2\eta}$ does not have a finite $\alpha/2$-th moment. 
Using Lemma~\ref{lem:product}(5), calculation shows that for $i\ne j$ with $\psi^{ij}=2$,
\beam\label{eq:null}
\lim_{\xto}\dfrac{\P(|X_{i0}\,X_{j0}|>x)}{\P(X^2>x)}=0
\eeam
Define $(a_n)$ \st\ $n\,\P(|X|>a_n)\to 1$. We may conclude from \eqref{eq:null} and Theorem~\ref{thm:1} that for $i\ne j$ we have 
$a_n^{-2} S_{ij}\stp 0$ although 
both $X_{i0}\,X_{j0}$ and $X^2$ are \regvary\ with index $\alpha/2$.
\par
By Theorem~\ref{thm:1} and Remark~\ref{reM:c1} we conclude that
\beam\label{eq:aa}
a_n^{-2} (S_i-c_n)_{i=1,\ldots,p} \std (\xi_{i,\alpha/2})_{i=1,\ldots,p}\,,
\eeam 
where the limit vector consists of $\alpha/2$-stable components. The spectral tail process $(\bfTh_h)_{h\ge 1}$ 
of the \seq\ $\bfX_t=(X_{1t},\ldots,X_{pt})'$, $t=1,2,\ldots$, vanishes. This follows by an argument similar to the proofs of 
Propositions~\ref{prop:rv:2} and \ref{Pr:genRVforproducts} under condition \eqref{Eq:inf:mom:eta}. 
A similar argument also yields that
\beao
\lim_{\xto}\dfrac{\P(|X_{i0}|>x\,,|X_{j0}|>x)}{\P(|X|>x)}=0\,,\qquad i\ne j\,.
\eeao
Therefore the the \ds\ of $\bfTh_0$ is concentrated on the axes and has the same form as $\boldsymbol{\Theta}_0^{(2)}$ in \eqref{Eq:Theta_0:measure}. As in the proof of Theorem~\ref{the:1} this implies that the limit random vector in \eqref{eq:aa} has iid components.
\par
We conclude that the limit theory for $S_{ij}$, $1\le i,j\le p$, are very essentially the same 
in Case (1) and in Case (2) when the additional
condition $\E [\ex^{\alpha \eta}]=\infty$ holds.
\end{example}
\begin{example}\label{exam:2}
Assume that \eqref{Eq:conv:eq:eta} holds. We may conclude from Theorem~\ref{thm:1}
that $a_n^{-2} S_{ij}\stp 0$ for $i\ne j$ if $\psi^{ij}<2$. The crucial difference to the previous case appears
when $\psi^{ij}=2$ for some $i\ne j$. In this case, not only the $(a_n^{-2}(S_i-c_n))$, $i=1,2,\ldots,$ have 
totally skewed to the right $\alpha/2$-stable limits 
but we also have $a_n^{-2} S_{ij}\std \xi_{ij,\alpha/2}$ for non-degenerate $\alpha/2$-stable $\xi_{ij,\alpha/2}$. From \eqref{eq:psi:ij:def}
we conclude that if $\psi^{ij}=2$ appears then $\psi^{i'j'}=2$ for all $|i'-j'|=|i-j|$. This means that non-degenerate limits may appear not only
on the diagonal of the matrix $a_n^{-2}(S_{ij}-c_n)$ but also along full sub-diagonals.
\par
In this case, the distribution of $\boldsymbol{\Theta}_0$ from the spectral tail process of the 
\seq\ $\bfX_t=(X_{1t},\ldots,X_{pt})'$ does not have to be concentrated on the axes---in contrast to Example~\ref{exam:1}. 
This implies that the limiting $\alpha/2$-stable random variables $\xi_{i,\alpha/2}, i=1, \ldots, p,$ are in general not independent. However, similar to the arguments at the end of the proof of Theorem~\ref{the:1}, one can show that the distribution of the limiting random vector $(\xi_{i,\alpha/2})_{i=1,\ldots,p}$ is the convolution of distributions of $\alpha/2$-stable random vectors which concentrate on 
hyperplanes of $\mathbb{R}^p$ of dimension less or equal than $|\{(m,n): \psi_{mn}=1\}|$. 
\end{example}

\subsection{The eigenvalues of the sample covariance matrix of a multivariate stochastic volatility model}\label{subsec:eigen2}
In this section we provide some results for the eigenvalues of the  sample covariance matrix $\bfX^n(\bfX^n) '$
under the conditions of  Theorem~\ref{thm:1}.
We introduce the sets 
\beao
\Gamma_p= \{(i,j):1\le i,j\le p\;\text{such that}\; \psi^{ij}=2\}\,,\qquad \Gamma_p^c= \{(i,j):1\le i,j\le p\}\backslash \Gamma_p\\
\eeao 
and let $(a_n)$ be \st\ $n\,\P(|X|>a_n)\to 1$. 
\bth\label{thm:2}
Assume that the conditions of Theorem~\ref{thm:1} hold for $(X_{it},X_{jt})$, $1\le i,j\le p$, and $\alpha\in (0,4)$.
Then
\beao
a_n^{-2} \big\|\bfX^n(\bfX^n)'- \wt \bfX^n\big\|_2\stp 0\,,\qquad \nto\,,
\eeao
where $\wt \bfX^n$ is a $p\times p$ matrix with entries
\beao
\wt X_{ij}=\sum_{t=1}^n X_{it}X_{jt} \I( (i,j)\in \Gamma_p)\,,\qquad 1\le i,j\le p\,.
\eeao
Moreover, if $\E [\ex^{\alpha\eta}]=\infty$ we also have
\beao
a_n^{-2} \big\|\bfX^n(\bfX^n)'- {\rm diag}(\bfX^n(\bfX^n)')\big\|_2\stp 0\,,\qquad \nto\,.
\eeao
\ethe
\begin{proof} We have
\beao
a_n^{-4} \big\|\bfX^n(\bfX^n)'- \wt \bfX^n\big\|_2^2 \le 
\sum_{(i,j)\in \Gamma_p^c} \big(a_n^{-2}S_{ij}\big)^2 \,.
\eeao
For $(i,j) \in \Gamma_p^c$ we have $i\ne j$ and the \seq\ $(X_{it}X_{jt})$ is \regvary\ with index $\alpha/\psi^{ij}>\alpha/2$. 
In view of Theorem~\ref{thm:1} the \rhs\
converges to zero in \pro y. 
\par
In the case when $\E[\ex^{\alpha\eta}]=\infty$ we learned in Example~\ref{exam:1} that $a_n^{-2}S_{ij}\stp 0$ whenever $i\ne j$. This concludes
the proof.
\end{proof}
For any $p\times p$ non-negative definite matrix $\bfA$ write $\la_i(\bfA)$, $i=1,\ldots,p$, for its eigenvalues and 
$\la_{(1)}(\bfA)\ge \cdots \ge \la_{(p)}(\bfA)$ for their ordered values. For the eigenvalues of $\bfX^n(\bfX^n)'$ we keep the previous
notation $(\la_i)$,
\begin{corollary}
Assume the conditions of Theorem~\ref{thm:2} and $\alpha\in (0,4)\backslash \{2\}$. Then 
\beam\label{eq:xy}
a_n^{-2}\max_{i=1,\ldots,p} \big|\la_{(i)}-\la_{(i)}(\wt \bfX^{n})\big|\stp 0\,.
\eeam
and
\beam\label{eq:90}
a_n^{-2}\Big(\la_{(i)} -n\,\E[X^2] 1(\alpha\in (2,4))\Big)_{i=1,\ldots,p}
\std \Big(\la_{(i)}\big((\xi_{kl,\alpha/2} \I((k,l)\in \Gamma_p))_{1\le k,l\le p}\big)\Big)_{i=1,\ldots,p} \,,\nonumber\\
\eeam
where $(\xi_{ij,\alpha/2})_{(i,j)\in \Gamma_p}$ are jointly $\alpha/2$-stable (possibly degenerate for $i\ne j$) \rv s.
Moreover, in the case when $\E [\ex^{\alpha\eta}]=\infty$ we have
\beam\label{eq:90a}
a_n^{-2}\Big(\la_{(i)} -n\,\E[X^2] 1(\alpha\in (2,4))\Big)_{i=1,\ldots,p}
\std \big(\xi_{(i),\alpha/2}\big)_{i=1,\ldots,p}\,,
\eeam
where $(\xi_{i,\alpha/2})_{i=1,\ldots,p}$ are iid totally skewed to the right $\alpha/2$-stable \rv s with order statistics $\xi_{(1),\alpha/2}\ge \cdots\ge \xi_{(p),\alpha/2}$.
\end{corollary}
\begin{proof} Relation \eqref{eq:xy} is an immediate con\seq\ of Theorem~\ref{thm:2} and Weyl's inequality; 
see Bhatia \cite{bhatia:1997}. We conclude  from Theorem~\ref{thm:1} and Remark~\ref{reM:c1}(3) that
\beam\label{eq:nn:anja}
a_n^{-2}\big(S_{ij}  - n\,\E[X^2]\,\I (\alpha\in (2,4))\big)_{(i,j)\in \Gamma_p}\std \big(\xi_{ij,\alpha/2}\big)_{(i,j)\in \Gamma_p} \,.
\eeam
Then \eqref{eq:90} follows. Relation \eqref{eq:90a} is a special case of \eqref{eq:90}. If $\E [\ex^{\alpha\eta}]=\infty$ then, 
in view of Example~\ref{exam:1}, only the diagonal elements in \eqref{eq:nn:anja} have non-degenerate iid $\alpha/2$-stable limits.
\end{proof}
\subsubsection*{Some conclusions}
By virtue of this corollary and in view of Section~\ref{subsec:eigen} the results for the eigenvalues 
in Case (1) and in Case (2) when $\E [\ex^{\alpha\eta}]=\infty$ are very much the same. Moreover, the results in 
Section~\ref{subsec:appl} remain valid in the latter case.
\par
If \eqref{Eq:conv:eq:eta} holds, Case (2) is quite different from Case (1); see Example~\ref{exam:2}.
In this case not only the diagonal of the matrix $\bfX^n(\bfX^n)'$ determines the
\asy\ behavior of its eigenvalues and eigenvectors. Indeed, if $\psi^{ij}=2$ for some $i\ne j$, then
at least two sub-diagonals of $\bfX^n(\bfX^n)'$ have non-degenerate $\alpha/2$-limits and these sub-diagonals together
with the diagonal determine the \asy\ behavior of the eigenspectrum. The limiting diagonal elements are dependent
in contrast to Case (1). This fact and the presence of sub-diagonals are challenges if one wants to calculate
the limit \ds s of the eigenvalues and eigenvectors.
\section{Simulations and data example}\label{sec:simulation}\setcounter{equation}{0}
In this section we illustrate the behavior of sample covariance matrices for moderate sample sizes
for the models discussed in Sections~\ref{sec:case1} and \ref{Sec:case2} 
and we compare them with a real-life data example. These data consist of
1567 daily log-returns of foreign exchange (FX) rates from
18 currencies against the Swedish Kroner (SEK) from January 4th 2010 to April 1st 2016, as made available by the Swedish National Bank. To start with, the Hill estimators of the tail indices $\alpha_{ij}, 1 \leq i,j, \leq 18,$ of the cross 
products $X_{it}X_{jt}, 1 \leq i,j, \leq 18,$ are visualized in Figure
\ref{Fig:Hill}.

\begin{figure}[h]
  \begin{minipage}{0.55\linewidth}
    \includegraphics[width=1.0\linewidth]
    {FX_HillEstimates.pdf}
  \end{minipage}
  \begin{minipage}[0.42\linewidth]{0.42\linewidth}
    In particular, 
    the Hill estimators on the diagonal (corresponding to the series
    $X_{it}^2, 1 \leq i \leq 18$) of the values $\alpha_i/2$, where
    $\alpha_i$ is the tail index of the $i$th currency, are of similar
    size although not identical. Even if all series had the same tail
    index the Hill estimator exhibits high statistical uncertainty which
    even increases for serially dependent data, cf.\ Drees
    \cite{drees:2003}.
  \end{minipage}
  \caption{Estimated tail indices of cross products for the FX rates
    of 18 currencies against SEK. The indices are derived by Hill
    estimators with threshold equal to the 97\%-quantile of $n=1567$
    observations.}
  \label{Fig:Hill}
\end{figure}

A way to make the data more homogeneous in their tails is to
rank-transform their marginals to the same \ds . We do, however,
refrain from such a  transformation to keep the correlation structure
of the original data unchanged.

It is clearly visible that some off-diagonal components of the matrix
have an estimated tail index  which is comparable to the on-diagonal
elements.  This implies that the tails of the corresponding
off-diagonal entries $S_{ij}, i\ne j$, of the sample covariance matrix
may  be of a similar magnitude as the on-diagonal entries $S_i$. This
is in stark contrast to the  asymptotic behavior of the models
analyzed in Section \ref{sec:case1}.

\begin{figure}[ht]
  \begin{minipage}{0.48\linewidth}
    \includegraphics[width=\linewidth]{FX_eigenvalues.pdf}
  \end{minipage}
  \begin{minipage}{0.48\linewidth}
    \includegraphics[width=\linewidth]{FX_eigenvector1.pdf}
  \end{minipage}
% \subcaptionbox{Based on FX rate data of 18 foreign currencies against SEK.\label{eigen:data}}
% {\includegraphics[width=6.5cm]{FX_eigenvalues.pdf}
% \includegraphics[width=6.5cm]{FX_eigenvector1.pdf}}
% \caption{Normalized and ordered eigenvalues (left) and eigenvector
% corresponding to largest eigenvalue (right) of real and simulated
% data, with $n=1567, p=18$.}\label{fig:1}
  \caption{Normalized and ordered eigenvalues (left) and eigenvector
    corresponding to largest eigenvalue (right) of real and
    simulated data, with $n=1567, p=18$. Based on FX rate data of 18
    foreign currencies against SEK.}
  \label{eigen:data}
\end{figure}

% \ContinuedFloat
% \subcaptionbox{Based on a stochastic volatility model with heavy-tailed innovation sequence.\label{eigen:model1}}
% {\includegraphics[width=6.5cm]{HeavyTailed_eigenvalues.pdf}
% \includegraphics[width=6.5cm]{HeavyTailed_eigenvector1.pdf}}
% \subcaptionbox{Based on a stochastic volatility model with heavy-tailed volatility sequence that satisfies assumptions of Example \ref{exam:1}. \label{eigen:model2}}
% {\includegraphics[width=6.5cm]{LightTailedCase1_eigenvalues.pdf}
% \includegraphics[width=6.5cm]{LightTailedCase1_eigenvector1.pdf}}
% \subcaptionbox{Based on a stochastic volatility model with heavy-tailed volatility sequence that satisfies assumptions of Example \ref{exam:2}.  \label{eigen:model3}}
% {\includegraphics[width=6.5cm]{LightTailedCase2_eigenvalues.pdf}
% \includegraphics[width=6.5cm]{LightTailedCase2_eigenvector1.pdf}}
\begin{figure}[htb]
  \centering
  \begin{minipage}{0.48\linewidth}
    \includegraphics[width=1.0\linewidth]{HeavyTailed_eigenvalues.pdf}
  \end{minipage}
  \begin{minipage}{0.48\linewidth}
    \includegraphics[width=1.0\linewidth]{HeavyTailed_eigenvector1.pdf}      
  \end{minipage}
  \caption{Based on a stochastic volatility model with heavy-tailed
    innovation sequence.}
  \label{eigen:model1}
\end{figure}

\begin{figure}
  \centering
  \begin{minipage}{0.48\linewidth}
    \includegraphics[width=6.5cm]{LightTailedCase1_eigenvalues.pdf}
  \end{minipage}
  \begin{minipage}{0.48\linewidth}
    \includegraphics[width=6.5cm]{LightTailedCase1_eigenvector1.pdf}    
  \end{minipage}
  \caption{Based on a stochastic volatility model with heavy-tailed
    volatility sequence that satisfies assumptions of Example
    \ref{exam:1}}
  \label{eigen:model2}
\end{figure}

\begin{figure}[htb]
  \centering
  \begin{minipage}{0.48\linewidth}
    \includegraphics[width=6.5cm]{LightTailedCase2_eigenvalues.pdf}
  \end{minipage}
  \begin{minipage}{0.48\linewidth}
    \includegraphics[width=6.5cm]{LightTailedCase2_eigenvector1.pdf}
  \end{minipage}
  \caption{Based on a stochastic volatility model with heavy-tailed
    volatility sequence that satisfies assumptions of Example
    \ref{exam:2}}
  \label{eigen:model3}
\end{figure}

Figure~\ref{eigen:data} shows the ordered eigenvalues of the sample
covariance matrix (normalized by its trace) 
and the eigenvector of the FX rate data corresponding to the largest eigenvalue.
There exists a notable spectral gap 
between the largest and second largest eigenvalues and the unit
eigenvector corresponding to the largest eigenvector 
has all positive and non-vanishing components.
For comparison and to illustrate the variety of the models discussed
above we also plot corresponding realizations of three model
specifications from Sections~\ref{sec:case1} and \ref{Sec:case2}. 
In all cases we choose $p=18$ and $n=1567$ in accordance with the data example. 
We assume throughout a moving average structure in the log-volatility
process $\log \sigma_{it}$ in \eqref{eq:1a:anja}. More specifically,
\begin{equation}\label{Eq:MA18} \sigma_{it}=\exp(\sum_{k=1}^{18}\eta_{i-k,t}), \;\;\; 1 \leq i \leq 18, \; t \in \mathbb{Z}. \end{equation}
In accordance with the model properties discussed in Section \ref{sec:case1}, we first assume iid\ standard Gaussian $\eta_{i,t}$ and iid\ $Z_{it}$ with a Student-$t$ distribution with $t=3$ degrees of freedom. 
Figure \ref{eigen:model1} shows the normalized eigenvalues and the first unit eigenvector from a realization of this model. We notice a relatively large gap between the first and second eigenvalue and, in accordance with Section \ref{Subsub:eigenvectors}, we see that the first unit eigenvector is relatively close to a unit basis vector. 
Figure \ref{eigen:model2} shows the corresponding realizations for the model \eqref{Eq:MA18} with a specification according to Example \ref{exam:1}, i.e.,\ Exponential(3)-distributed iid\ $\eta_{i,t}$ (meaning that $\P(\eta_{i,t}>x)=\exp(-3x), x \geq 0$, 
which implies $\alpha=3$ and $\E[\ex^{3 \eta}]=\infty$) and iid\ standard Gaussian $Z_{it}$. Compared to the first simulated model, we see a slower decay in the magnitude of the ordered eigenvalues and a more spread out first unit eigenvector. This 
observation illustrates that although the limit behavior of this model and the one analyzed 
before should be very similar (cf. Example \ref{exam:1}), convergence to the prescribed limit appears slower for the heavy-tailed volatility sequence than for the heavy-tailed innovations. 
Finally, Figure \ref{eigen:model3} shows a simulation drawn from \eqref{Eq:MA18} where the $\eta_{i,t}$ are iid\ such that $\P(\eta_{i,t}>x) \sim x^{-2}\exp(-3x), x \to \infty,$ and the $Z_{it}$ are iid\ standard Gaussian. Again, $\alpha=3$, but direct calculations show that the distribution of $\eta_{i,t}$ is convolution equivalent, i.e.,\ it satisfies \eqref{Eq:conv:eq:eta} instead of \eqref{Eq:inf:mom:eta}. The graphs are in line with the analysis in Example \ref{exam:2} and illustrate a very spread out dominant eigenvector. We note that while none of the three very simple models analyzed in the simulations above is able to fully describe
the behavior of the analyzed data, the two models with heavy-tailed volatility and light-tailed innovations are able 
to explain a non-concentrated first unit eigenvector of the sample covariance matrix and 
therefore non-negligible dependence between components as seen in the data.

\section*{Acknowledgements}
Thomas Mikosch's and Xiaolei Xie's research is partly supported by the Danish Research Council Grant DFF-4002-00435 ``Large random matrices with heavy tails and dependence''. Parts of the paper were written when Mohsen Rezapour visited the Department of Mathematics at the University of Copenhagen December 2015--January 2016. He would like to thank the Department of Mathematics for hospitality.

\section{Some $\alpha$-stable limit theory}\setcounter{equation}{0}\label{App:A}
In this paper, we make frequently use of
Theorem 4.3 in Mikosch and Wintenberger \cite{mikosch:wintenberger:2016} which we quote for convenience:
\begin{theorem}\label{thm:mikwin}
Let $(\bfY_t)$ be an $\bbr^p$-valued strictly stationary \seq , $\bfS_n=\bfY_1+\cdots +\bfY_n$ and $(a_n)$ be \st\ $n\,\P(\|\bfY\|>a_n)\to 1$.
Also write for $\vep>0$, $\ov \bfY_t = \bfY_t \I(\|{\mathbf Y}_t\|\le \vep a_n)$, $\underline \bfY_t= \bfY_t-\ov  \bfY_t $ and
\beao
\ov \bfS_{l,n}= \sum_{t=1}^l \ov \bfY_t \qquad  \un \bfS_{l,n}= \sum_{t=1}^l \un \bfY_t\,.
\eeao
Assume the following conditions:
\begin{enumerate}
\item 
$(\bfY_t)$ is \regvary\ with index $\alpha\in (0,2)  \setminus\{1\}$ and spectral tail process $(\bfTh_j)$.
\item
A mixing condition holds: there exists an integer \seq\ $m_n\to\infty$ \st\ $k_n= [n/m_n]\to \infty$
and 
\beam\label{eq:chfa}
\E \ex^{i\bft'\underline \bfS_n/a_n} - \Big(\E \ex^{i\bft'\underline \bfS_{m_n,n}/a_n}\Big)^{k_n} \to 0\,,\qquad \nto\,,\qquad \bft\in\bbr^p\,.
\eeam
\item An anti-clustering condition holds:
\beam\label{eq:acl}
\lim_{l\to\infty} \limsup_{\nto} \P\big(\max_{t=l,\ldots,m_n} \|\bfY_t\|>\delta a_n\mid \|\bfY_0\|>\delta a_n\big)=0\,,\qquad \delta>0\,
\eeam
for the same sequence $(m_n)$ as in (2).
\item If $\alpha\in (1,2)$, in addition $\E[\mathbf{Y}]=\mathbf{0}$ and the vanishing small values condition holds: 
\beam\label{eq:vansm}
\lim_{\vep\downarrow 0}\limsup_{\nto} \P\big(a_n^{-1} \|\ov \bfS_n-\E [\ov \bfS_n]\|>\delta \big)=0\,,\qquad \delta>0\,
\eeam
and $\sum_{i=1}^\infty \E[\|\bfTh_i\|]<\infty$.
\end{enumerate}
Then $a_n^{-1} \bfS_n\std \xi_\alpha$ for an $\alpha$-stable $\bbr^p$-valued vector $\xi_\alpha$ with log characteristic function
\beam\label{eq:chfid}
\int_0^\infty \E\big[\ex^{i\,y\,\bft'\sum_{j=0}^\infty \bfTh_j}- \ex^{i\,y\,\bft'\sum_{j=1}^\infty \bfTh_j}-i\,y\,\bft'\I_{(1,2)}(\alpha)\big]\,
d(-y^{\alpha})\,,\qquad \bft\in\bbr^p\,.
\eeam
\end{theorem}\noindent 
\begin{remark}\label{rem:case:alpha:1}
If we additionally assume that $\mathbf{Y}$ is symmetric, which implies $\E[\ov \bfY]=\mathbf{0}$, then the statement 
of the theorem also holds for $\alpha=1$.
\end{remark}

\section{(Joint) Tail behavior for products of \regvary\ \\ \rv s}\label{App:B}
In this paper, we make frequently use of the tail behavior of products of non-negative independent \rv s $X$ and $Y$.
In particular, we are interested in conditions for the existence of the limit
\beam\label{eq:rr}
\lim_{\xto}\dfrac{\P(XY>x)}{\P(X>x)}=q\,.
\eeam 
for some $q\in [0,\infty]$.
We quote some of these results
for convenience.
\ble\label{lem:product} Let $X$ and $Y$ be independent \rv s.
\begin{enumerate}
\item 
If $X$ and $Y$ are \regvary\ with index $\alpha>0$ then $XY$ is \regvary\ with the same index.
\item
If $X$ is \regvary\ with index $\alpha>0$ and $\E[Y^{\alpha+\vep}]<\infty$ for some $\vep>0$ then \eqref{eq:rr} holds with 
$q=\E[Y^\alpha]$.
\item
If  $X$ and $Y$ are iid \regvary\ with index $\alpha>0$ and $\E[Y^\alpha]<\infty$, then \eqref{eq:rr} holds with
$q=2 \E[|Y|^\alpha]$ iff
\beam\label{eq:hh}
\lim_{M\to\infty}\limsup_{\xto}\dfrac{\P(XY>x,M<Y\le x/M )}{\P(X>x)}=0\,.
\eeam
\item
If $X$ and $Y$ are \regvary\ with index $\alpha>0$, $\E[Y^\alpha+ X^\alpha]<\infty$, $\lim_{\xto} \P(Y>x)/\P(X>x)=0$ and 
\eqref{eq:hh} holds, then \eqref{eq:rr} holds with $q=\E[|Y|^\alpha]$.
\item
Assume that $\E[|Y|^\alpha]=\infty$. Then  \eqref{eq:rr} holds with $q=\infty$.
\end{enumerate}
\ele
\begin{proof} (1) This is proved in Embrechts and Goldie \cite{embrechts:goldie:1980}.\\
(2) This is Breiman's \cite{breiman:1965} result.\\
(3) This is Proposition~3.1 in Davis and Resnick \cite{davis:resnick:1985}.\\
(4) This part is proved similarly to (3); we borrow the ideas from \cite{davis:resnick:1985}.
For $M>0$ we have the following decomposition
\beao
\dfrac{\P(XY>x)}{\P(X>x)}&=& \dfrac{\P(XY>x,Y\le M)}{\P(X>x)}+\dfrac{\P(XY>x,M<Y\le x/M)}{\P(X>x)}
+\dfrac{\P(XY>x, Y>x/M)}{\P(X>x)}\\
&\sim & \E[Y^\alpha \I(Y\le M)]+ \dfrac{\P(XY>x,M<Y\le x/M)}{\P(X>x)} + \E [(X\wedge M)^\alpha] \dfrac{\P(Y>x)}{\P(X>x)}\\
&=&  \E[Y^\alpha \I(Y\le M)] + \dfrac{\P(XY>x,M<Y\le x/M)}{\P(X>x)} +o(1)\,.
\eeao
Here we applied Breiman's result twice. The second term vanishes by virtue of \eqref{eq:hh}. Thus $q=  \E[Y^\alpha]$.\\
(5) The same argument as for (4) yields as $\xto$,
\beao
\dfrac{\P(XY>x)}{\P(X>x)}&\ge & \dfrac{\P(XY>x,Y\le M)}{\P(X>x)}\sim \E[Y^\alpha \I(Y\le M)]\,.
\eeao  
Then \eqref{eq:rr} with $q=\infty$ is immediate.
\end{proof}
\ble\label{Lem:splitup} Let $Y_1, \ldots, Y_p \geq 0$ be iid regularly varying \rv s with index $\alpha>0$. Assume that
\begin{equation}\label{Eq:conv:equivalent} \lim_{t \to \infty} \frac{\P(Y_1\cdot Y_2>t)}{\P(Y_1>t)}=c \in (0,\infty)\,.
\end{equation}
Then for any $a_1, \ldots, a_p \geq 0$ such that $a_{\max}:=\max_{j=1, \ldots, p}a_j>0$ and any $v>0$ we have
\begin{eqnarray}\label{Eq:lemma:prod:1} \lim_{t \to \infty} \frac{\P(\prod_{i=1}^p Y_i^{a_i}>vt)}{\P(Y_1^{a_{\max}}>t)}= \sum_{j:a_j=a_{\max}} \lim_{s \to 0} \lim_{t \to \infty} \frac{\P(\prod_{i=1}^p Y_i^{a_i}>vt, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)} \end{eqnarray}
and
\begin{eqnarray}\label{Eq:lemma:prod:2}  \lim_{s \to 0} \limsup_{t \to \infty}\frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, \max_{j=1, \ldots, p}Y_j^{a_{\max}}\leq st) }{\P(Y_1^{a_{\max}}>t)}=0. \end{eqnarray}
\begin{proof} 
In view of Davis and Resnick \cite{davis:resnick:1986} the only possible value for $c$ in \eqref{Eq:conv:equivalent} is $2 \E[Y_1^\alpha]$ (which implies that $\E[Y_1^\alpha]<\infty$). Furthermore, we note that the product $\prod_{j:a_j=a_{\max}}Y_j^{a_j}$ is regularly varying with index $-\alpha/a_{\max}$; see Embrechts and Goldie \cite{embrechts:goldie:1980}, Corollary on p.~245. By Breiman's lemma this implies that
\begin{eqnarray*}
&& \lim_{t \to \infty} \frac{\P(\prod_{i=1}^p Y_i^{a_i}>vt)}{\P(Y_1^{a_{\max}}>t)}\\
&=& \lim_{t \to \infty} \frac{\P(Y_1^{a_{\max}}>vt)}{\P(Y_1^{a_{\max}}>t)}\frac{\P(\prod_{i=1}^p Y_i^{a_i}>vt)}{\P(Y_1^{a_{\max}}>vt)}\\
&=& v^{-\alpha/a_{\max}}  \Big( \prod_{j:a_j \neq a_{\max}} \E[Y_j^{\alpha a_j/a_{\max}}] \Big)\lim_{t \to \infty} \frac{\P(\prod_{j:a_j=a_{\max}} Y_j^{a_{\max}}>vt)}{\P(Y_1^{a_{\max}}>vt)}.
\end{eqnarray*}
By Lemma 2.5 in Embrechts and Goldie \cite{embrechts:goldie:1982} 
(cf.\ also  Chover, Ney and Wainger \cite{chover:ney:wainger:1973}) this equals 
$$ v^{-\alpha/a_{max}} \Big(\prod_{j:a_j\neq a_{\max}} \E[Y_j^{\alpha a_j/a_{\max}}] \Big) |\{j:a_j = a_{\max} \}|\,\E[Y_1^\alpha]^{ |\{j:a_j = a_{\max} \}|-1}. $$
On the other hand, we have
\begin{eqnarray*}
&& \sum_{j:a_j=a_{\max}} \lim_{s \to 0} \lim_{t \to \infty} \frac{\P(\prod_{i=1}^p Y_i^{a_i}>vt, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)} \\
&=& \sum_{j:a_j=a_{\max}} \lim_{s \to 0} \lim_{t \to \infty} \frac{\P(Y_j^{a_{\max}}\min(s^{-1},v^{-1}\prod_{k\neq j}Y_k^{a_k})>t)}{\P(Y_j^{a_{\max}}>t)} \\
&=&  \sum_{j:a_j=a_{\max}} \lim_{s \to 0} \E[(\min(s^{-1},v^{-1}\prod_{k \neq j}Y_k^{a_k}))^{\alpha/a_{\max}}] \\
&=&  v^{-\alpha/a_{\max}} \sum_{j:a_j=a_{\max}} \prod_{k \neq j}\E[Y_k^{\alpha a_k/a_{\max}}] \\
&=& v^{-\alpha/a_{\max}}  \Big(\prod_{j:a_j\neq a_{\max}} \E[Y_j^{\alpha a_j/a_{\max}}] \Big)  |\{j: a_j=a_{\max}\}|\, \E[Y_1^\alpha]^{|\{j:a_j = a_{\max} \}|-1},
\end{eqnarray*}
where we applied Breiman's lemma in the second step to the bounded random variable $\min(s^{-1},$ $v^{-1}\prod_{k\neq j}Y_k^{a_k})$, and the monotone convergence theorem in the penultimate step. This proves \eqref{Eq:lemma:prod:1}. To prove \eqref{Eq:lemma:prod:2} note that 
for $s>0$,
\begin{eqnarray*}  \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt) }{\P(Y_1^{a_{\max}}>t)}  &\geq& \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, \max_{j=1, \ldots, p}Y_j^{a_{\max}}\leq st) }{\P(Y_1^{a_{\max}}>t)} \\
&& + \sum_{j:a_j =a_{\max}} \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_j^{a_{\max}}>st) }{\P(Y_1^{a_{\max}}>t)}\\
&& -\frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_{j_1}^{a_{\max}}>st, Y_{j_2}^{a_{\max}}>st \mbox{ for some } j_1\neq j_2) }{\P(Y_1^{a_{\max}}>t)}, \;\; s>0.
\end{eqnarray*} 
The last summand on the \rhs\  
converges to 0 as $t \to \infty$ by independence of the $Y_j's$. Moreover,  the left-hand term and the second term 
on the \rhs\ become equal by first $t \to \infty$ and then $s \to 0$, in view of \eqref{Eq:lemma:prod:1}. Therefore 
the first right-hand term vanishes by first $t \to \infty$ and then $s \to 0$. This proves the statement.
\end{proof}
\ele
\begin{proposition}\label{Pr:genRVforproducts} Let $Y_1, \ldots, Y_p \geq 0$ be iid\ regularly varying with index $\alpha$ and 
$(a_{ij}) \in [0,\infty)^{n\times p}, n,p \geq 1,$ be such that $\max_{1\leq i \leq n} a_{ik}=a_{\max}:=\max_{i,j}a_{ij}>0$ for any $1 \leq k \leq p$. 
\begin{itemize}
\item[(i)] Assume that \eqref{Eq:conv:equivalent} holds. Then the random vector
\begin{equation}\label{Eq:def:vector:Y} \mathbf{Y}:=\big(\prod_{j=1}^pY_j^{a_{ij}}\big)_{1 \leq i \leq n}
\end{equation}
is regularly varying with index $\alpha/a_{\max}$. Furthermore, up to a constant the limit measure $\mu$ 
of $\bfY$ is given by  $\sum_{j=1}^p \mu_j,$ where for any Borel set $B \in [0,\infty]^n$ bounded away from $\mathbf{0}$ and $\nu_\alpha(dz)= \alpha z^{-\alpha-1}dz$,
\begin{eqnarray}
\label{Eq:RV:prod:1}\mu_j(B)&=&\int_0^\infty \P\bigg(\bigg(\I(a_{ij}=a_{\max})z^{a_{\max}}\prod_{k \neq j}Y_k^{a_{ik}}\bigg)_{1 \leq i \leq n} \in B\bigg)\nu_{\alpha}(dz). 
\end{eqnarray}
\item[(ii)] Assume that $\E[Y_1^\alpha]=\infty$. Set 
\beao
p_{\mbox{\scriptsize eff}}&:=&\max_i |\{1\leq j \leq p:a_{ij}=a_{\max}\}|\,,\\
P_{\mbox{\scriptsize eff}}&:=&\{A \subset \{1, \ldots, p\}: |A|=p_{\mbox{\scriptsize eff}} \wedge \, \exists \, i:\, \forall\, j \in A: a_{ij}=a_{\max}\}\,.
\eeao Then the random vector $\mathbf{Y}$ in \eqref{Eq:def:vector:Y}  is regularly varying with index $\alpha/a_{\max}$. 
Furthermore,  up to a constant the limit measure $\mu$ of $\bfY$ is equal to $\sum_{A \in P_{\mbox{\scriptsize eff}}} \mu_A,$ where 
for any Borel set $B \in [0,\infty]^n$ bounded away from $\mathbf{0}$,
\begin{eqnarray}
\label{Eq:RV:prod:2} \mu_A(B)&=&\int_0^\infty \P\bigg(\bigg(\I(a_{ij}=a_{\max} \, \forall\, j \in A)z^{a_{\max}}\prod_{k \notin A}Y_k^{a_{ik}}\bigg)_{1 \leq i \leq n} \in B\bigg)\nu_{\alpha}(dz)\,.
\end{eqnarray}

\end{itemize}
\end{proposition}
\begin{proof}% \leavevmode
%\begin{itemize}
(i)~Let $B \in [0,\infty]^n$ be a Borel set bounded away from $\mathbf{0}$. For $s>0$ we have
               \begin{eqnarray}
                \label{Eq:three:summands}  \frac{\P(\mathbf{Y} \in tB)}{\P(Y_1^{a_{\max}}>t)} &=&  \frac{\P(\mathbf{Y} \in tB, \max_{j=1, \ldots, p}Y_j^{a_{\max}}\leq st)}{\P(Y_1^{a_{\max}}>t)}
                 + \sum_{j=1}^p  \frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)} \nonumber\\
                && - \frac{\P(\mathbf{Y} \in tB, Y_{j_1}^{a_{\max}}>st, Y_{j_2}^{a_{\max}}>st, \mbox{ for some } j_1 \neq j_2)}{\P(Y_1^{a_{\max}}>t)}.
               \end{eqnarray}
Since $B$ is bounded away from $\mathbf{0}$, there exists $v>0$ and $1 \leq i \leq n$ such that $B \subset \{(x_1, \ldots, x_n) \in [0,\infty]^n: x_i>v\}$. From Lemma \ref{Lem:splitup}, \eqref{Eq:lemma:prod:2} the first summand in \eqref{Eq:three:summands} therefore tends to 0 by first $t \to \infty$ and then $s \to 0$. Furthermore, the third summand converges to zero as $t \to \infty$ by independence of the $Y_j's$. We are thus left to show
$$ \lim_{s \searrow 0}\lim_{t \to \infty} \frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}=\mu_j(B), \;\;\; 1 \leq j \leq p, $$
with $\mu_j$ as in \eqref{Eq:RV:prod:1}. For $s>0$ write
\begin{eqnarray*} && \lim_{t \to \infty}\frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}\\
&=& s^{-\alpha/{a_{\max}}} \lim_{t\to \infty} \P(\mathbf{Y} \in tB \mid Y_j^{a_{\max}}>st) \\
&=& s^{-\alpha/{a_{\max}}} \lim_{t \to \infty} \P\left(\left(\left(\frac{Y_j^{a_{\max}}}{st}\right)^{\frac{a_{ij}}{a_{\max}}}s^{\frac{a_{ij}}{a_{\max}}}t^{\frac{a_{ij}}{a_{\max}}-1}\prod_{k\neq j} Y_k^{a_{ik}}\right)_{1\leq i \leq n} \in B \; \bigg| \; Y_j^{a_{\max}}>st \right)\\
&=& s^{-\alpha/{a_{\max}}} \int_1^\infty \P\left(\left(\I(a_{ij}=a_{\max})sy\prod_{k \neq j}Y_k^{a_{ik}}\right)_{1 \leq i \leq n} \in B\right)\nu_{\alpha/a_{\max}}(dy).
\end{eqnarray*}
Substituting $sy$ by $z$ in the integral finally gives
$$ \lim_{s \searrow 0}\lim_{t \to \infty} \frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}=\int_0^\infty \P\left(\left(\I(a_{ij}=a_{\max})z^{a_{\max}}\prod_{k \neq j}Y_k^{a_{ik}}\right)_{1 \leq i \leq n} \in B\right)\nu_{\alpha}(dz). $$
(ii) Note first that under our assumptions for any $1\leq n_1 < n_2 \leq p$,
\begin{eqnarray}\nonumber \lim_{t \to \infty}\frac{\P(\prod_{j=1}^{n_2} Y_j>t)}{\P(\prod_{j=1}^{n_1} Y_j>t)}
\label{Eq:non:conv:equiv} &=& \lim_{t \to \infty} \int_0^\infty \frac{\P(\prod_{j=1}^{n_1} Y_j>t/y)}{\P(\prod_{j=1}^{n_1} Y_j>t)}P^{\prod_{j=n_1+1}^{n_2}Y_j}(dy)\\
&\geq& \E\left[\prod_{j=n_1+1}^{n_2}Y_j^\alpha\right]=\infty
\end{eqnarray}
by Fatou's lemma and the regular variation of $\prod_{j=1}^{n_1}Y_j$. Write now
\begin{equation}\label{Eq:split:sum}\mathbf{Y}=\sum_{\substack{1 \leq i \leq n \\ |\{j:a_{ij}=a_{\max}\}|=p_{\mbox{\scriptsize eff}}}} \prod_{j=1}^p Y_j^{a_{ij}} \mathbf{e}_i + \sum_{\substack{1 \leq i \leq n \\ |\{j:a_{ij}=a_{\max}\}|<p_{\mbox{\scriptsize eff}}}} \prod_{j=1}^p Y_j^{a_{ij}} \mathbf{e}_i,
\end{equation}
where $\mathbf{e}_i$ stands for the $i$-th unit vector. The first sum can also be written as
\begin{equation}\label{Eq:split:YA}\sum_{A \in P_{\mbox{\scriptsize eff}}}\mbox{diag}((\I(a_{ij}=a_{\max} \, \forall\, j \in A)\prod_{k \notin A}Y_k^{a_{ik}})_{1 \leq i \leq n})\prod_{j \in A}Y_j^{a_{\max}}=:\sum_{A \in P_{\mbox{\scriptsize eff}}}\mathbf{Y}^A, 
\end{equation}
where for each summand the random matrix and the random factor are independent and for the non-zero entries of the matrix we have $a_{ik}<a_{\max}$ since $k \notin A$. Thus, by the multivariate version of Breiman's lemma each $\mathbf{Y}^A$ is a multivariate regularly varying vector with limit measure $\mu_A$ (up to a constant multiplier) as in \eqref{Eq:RV:prod:2} and normalizing function $P(\prod_{i=1}^{p_{\mbox{\scriptsize eff}}}Y_i^{a_{\max}}>x)$. Furthermore, for $A, A' \in P_{\mbox{\scriptsize eff}}$ with $A \neq A'$ and $i,i'$ such that $a_{ij}=a_{\max} \, \forall\, j \in A$ and $a_{i'j}=a_{\max} \, \forall\, j \in A'$ we have
\begin{eqnarray}\label{Eq:asymp:ind:YA} && \frac{\P(\mathbf{Y}^A_i>x,\mathbf{Y}^{A'}_{i'}>x)}{\P(\prod_{i=1}^{p_{\mbox{\scriptsize eff}}}Y_i^{a_{\max}}>x)}\\
\nonumber &=& \frac{\P((\prod_{j \in A \cap A'}Y_j)^{a_{\max}}\prod_{j \in (A \cap A')^c}Y_j^{a_{ij}}>x,(\prod_{j \in A \cap A'}Y_j)^{a_{\max}}\prod_{j \in (A \cap A')^c}Y_j^{a_{i'j}}>x)}{\P(\prod_{i=1}^{p_{\mbox{\scriptsize eff}}}Y_i^{a_{\max}}>x)}\,. 
\end{eqnarray}
By Jan\ss en and Drees \cite{janssen:drees:2016}, Theorem 4.2 (in connection with Remark 4.3 (ii) and the minor 
change that our random variables are regularly varying with index $\alpha$ instead of $1$), the numerator behaves asymptotically like $\P((\prod_{j \in A \cap A'}Y_j)^{a_{\max}}>x)$, since $\kappa_0=a_{\max}^{-1}, \kappa_{j}=0, j \in (A \cap A')^c$ is the unique non-negative optimal solution to
$$ \kappa_0+\sum_{j \in (A \cap A')^c} \kappa_j \to \min ! $$
under
$$\kappa_0 a_{\max}+\sum_{j \in (A \cap A')^c}\kappa_j a_{ij}\geq 1, \;\;\;  \kappa_0 a_{\max}+\sum_{j \in (A \cap A')^c}\kappa_j a_{i'j}\geq 1.$$
This is because $\min(a_{ij},a_{i'j})<a_{\max}$ and $\max(a_{ij},a_{i'j})\leq a_{\max}$ for all $j \in (A \cap A')^c$. Since $A \neq A'$, we have $|A \cap A'|<p_{\mbox{\scriptsize eff}}$ and thus, by \eqref{Eq:non:conv:equiv}, the expression \eqref{Eq:asymp:ind:YA} converges to 0 as $x \to \infty$. Therefore, each component of $\mathbf{Y}^A$ is asymptotically independent of each component of $\mathbf{Y}^{A'}$ and thus the sum in \eqref{Eq:split:YA} is multivariate regularly varying with limit measure $\sum_{A \in P_{\mbox{\scriptsize eff}}}\mu_A$ and normalizing function $\P(\prod_{i=1}^{p_{\mbox{\scriptsize eff}}}Y_i^{a_{\max}}>x)$. Since the second sum in \eqref{Eq:split:sum} consists by \eqref{Eq:non:conv:equiv} only of random vectors for which $\P(\|\prod_{j=1}^p Y_j^{a_{ij}} \mathbf{e}_i\|>x)=\P(\prod_{j=1}^p Y_j^{a_{ij}}>x) =o(\P(\prod_{i=1}^{p_{\mbox{\scriptsize eff}}}Y_i^{a_{\max}}>x))$, we have that $\mathbf{Y}$ is regularly varying with index $\alpha/a_{\max}$ and limit measure $\sum_{A \in P_{\mbox{\scriptsize eff}}}\mu_A$ by Lemma 3.12 in Jessen and Mikosch \cite{jessen:mikosch:2006}. 
\end{proof}

### 2017-07-31 cpp/ImportanceSampling/garch21.cpp

@@ -38,7 +38,7 @@ double interpolate_fun
 Garch21::Garch21(const vector<double> &alpha, const vector<double> &beta,
 		 double tail_index_sup)
     :GarchPQ(alpha, beta),
-     tail_index(find_tail_index(vector<double>({0, tail_index_sup}).data()))
+     tail_index(find_tail_index(vector<double>({1, tail_index_sup}).data()))
 {
     random_device randev;
     chi_squared_distribution<double> chi2;
@@ -119,11 +119,10 @@ void Garch21::simulate_sample_path
 {
 }
 
-void Garch21::right_eigenfunction
+double Garch21::right_eigenfunction
 (double index, vector<funval> &eigenfunction) const
 {
     double ang_max = atan(1/alpha[1]);
-    double flag;
     size_t n = eigenfunction.size();
     vec angles(n);
     mat P(n, n);
@@ -132,58 +131,45 @@ void Garch21::right_eigenfunction
     double c1 = pow(pow(alpha[2], 2) + pow(beta[0], 2), 0.5);
     double ang1 = atan(beta[0] / alpha[2]);
     
-    do {
-	vec eigenfun(n);
-	vec tangents(n);
-	vec cosines(n);
+    vec eigenfun(n);
+    vec tangents(n);
+    vec cosines(n);
 
-	double d = ang_max/(n + 1);
-	P.set_size(n, n);
-	angles.set_size(n);
-	for (unsigned i = 0; i < n; i++) {
-	    angles[i] = d * (i + 1);
-	    tangents[i] = tan(angles[i]);
-	    cosines[i] = cos(angles[i]);
+    double d = ang_max/(n + 1);
+    P.set_size(n, n);
+    angles.set_size(n);
+    for (unsigned i = 0; i < n; i++) {
+	angles[i] = d * (i + 1);
+	tangents[i] = tan(angles[i]);
+	cosines[i] = cos(angles[i]);
+    }
+
+    for (unsigned i = 0; i < n; i++) {
+	for (unsigned j = 0; j < n; j++) {
+	    double t1 = sin(angles[i] + ang1);
+	    double t2 = sqrt(tangents[i] * alpha[2] + beta[0]);
+	    double t3 = (alpha[2] * tangents[i] + beta[0]);
+	    t3 *= tangents[j] / (1 - tangents[j] * alpha[1]) / 2;
+	    t3 = exp(-t3);
+
[…]

### 2017-08-01 cpp/ImportanceSampling/garch21.hpp
Implemented some functions for GARCH21.

@@ -2,7 +2,6 @@
 #define XXIE_GARCH21_H
 
 #include "GarchPQ.hpp"
-
 typedef array<double, 2> funval;
 // typedef vector<funval> emp_fun;
 
@@ -14,7 +13,10 @@ class Garch21 : public GarchPQ
 public:
     Garch21(const vector<double> &alpha, const vector<double> &beta,
 	    double tail_index_sup = 5.0);
+
+    ~Garch21(void);
     const double tail_index;
+    
     double quantile(double u, double angle) const;
     double draw_z2(void) const;
     double draw_z2(double) const;
@@ -26,13 +28,17 @@ public:
     double right_eigenfunction
     (double index, vector<funval> &eigenfunction) const;
 
-    double M_fun(double theta);
+    // double M_fun(double theta);
     double b_fun(double theta);
     
 protected:
     vector<double> pool;
-    // vector<funval> eigenfunction;
-
+    struct eigenfunction {
+	double kappa;
+	double lambda_kappa;
+	vector<funval> *r_kappa;
+    };
+    vector<struct eigenfunction> eigenfunctions;
 };
 
 

### 2017-08-01 phd/MyThesis.tex
Revised the importance sampling paper.
Added the paper with Anja to the thesis.
Debugged the program for finding M.

@@
 \input{Preamble}
+\input{../physics_common}
+
 \usepackage[numbers,sort&compress]{natbib}
@@
 \usepackage{algpseudocode}
-
-\newtheorem{case}{case}
-
-%% \usepackage[toc,page]{appendix}
 \usepackage{appendix}
@@
         {../papers/FX/}{../papers/Jeffrey1/}
+        {../papers/Anja/}
         {../papers/DavisMikoschHeinyXie/}
@@
 
-This thesis is about vanity of life.
+This thesis is about heavy-tailed time series.
 
@@
 
-Thank you all for various things.
+I owe my gratitude to many people and especially to my supervisors Prof. Mikosch and Prof. Collamore.
 \bigskip
@@
 \input{Chapterextremes}
+\input{ChapterBernoulli}
 
@@
 \phantom{bla}
+
 \end{document}

### 2017-08-01 cpp/ImportanceSampling/garch21.hpp

@@ -10,12 +10,26 @@ double interpolate_fun
 
 class Garch21 : public GarchPQ
 {
+protected:
+    const size_t nbr_eigenfunctions;
+    const size_t nbr_eigenfunction_points;
+    vector<double> pool;
+    struct eigenfunction {
+	double kappa;
+	double lambda_kappa;
+	vector<funval> *r_kappa;
+    };
+    vector<eigenfunction> eigenfunctions;
+    void compute_eigenfunctions(void);
+    double compute_M(void);
+
 public:
     Garch21(const vector<double> &alpha, const vector<double> &beta,
 	    double tail_index_sup = 5.0);
-
     ~Garch21(void);
+
     const double tail_index;
+    const double M;
     
     double quantile(double u, double angle) const;
     double draw_z2(void) const;
@@ -28,17 +42,7 @@ public:
     double right_eigenfunction
     (double index, vector<funval> &eigenfunction) const;
 
-    // double M_fun(double theta);
-    double b_fun(double theta);
-    
-protected:
-    vector<double> pool;
-    struct eigenfunction {
-	double kappa;
-	double lambda_kappa;
-	vector<funval> *r_kappa;
-    };
-    vector<struct eigenfunction> eigenfunctions;
+
 };
 
 

### 2017-08-02 cpp/ImportanceSampling/garch21.hpp
On the importance sampling implementation.

@@ -19,6 +19,10 @@ protected:
 	double lambda_kappa;
 	vector<funval> *r_kappa;
     };
+    struct chain_state {
+	double angle; // X = (cos(angle), sin(angle))
+	double addive; // V = X*exp(l)
+    };
     vector<eigenfunction> eigenfunctions;
     void compute_eigenfunctions(void);
     double compute_M(void);
@@ -34,8 +38,8 @@ public:
     double quantile(double u, double angle) const;
     double draw_z2(void) const;
     double draw_z2(double) const;
-    void simulate_sample_path(double threshold, unsigned int n,
-			      vector<vec> &path) const;
+    void simulate_path(vector<vec> &path) const;
+    double estimate_prob(double u);
 
     double Lyapunov(size_t n, size_t iterations);
 

### 2017-08-02 papers/FX/PaperBody.tex
Revised the introduction section of the thesis. Started programming for the main importance sampling algorithm.
Some minor revision.

@@
 in the ``Energy'' and ``Information Technology'' sectors are  larger
-than those in the ``Consumer Staples'' sector.  For a given loss \pro y, 
-a larger value of $K$ implies that large losses are more
+than those in the ``Consumer Staples'' sector.  For a given loss
+probability, a larger value of $K$ implies that large losses are more
 probable. Thus one can conclude that these two sectors are
-considerably riskier than the ``Consumer Staples'' sector. {\red Casper:
-we leave this sentence to your discretion.} This is of course a
-confirmation of one's economic instinct.
+considerably riskier than the ``Consumer Staples'' sector. This is of
+course a confirmation of one's economic instinct.
 
@@
   equity, as shown in Figure~\ref{fig:phi_hat_pareto}. Note that,
-  while $\hat \phi$ changes nearly in the same in both graphs, the
+  while $\hat \phi$ changes nearly in the same way in both graphs, the
   scales in the graphcs are very different.

### 2017-08-02 papers/FX/PaperBody.tex
Some textual revision.

@@
 in the ``Energy'' and ``Information Technology'' sectors are  larger
-than those in the ``Consumer Staples'' sector.  For a given loss
-probability, a larger value of $K$ implies that large losses are more
+than those in the ``Consumer Staples'' sector.  For a given loss probability, 
+a larger value of $K$ implies that large losses are more
 probable. Thus one can conclude that these two sectors are
@@
   equity, as shown in Figure~\ref{fig:phi_hat_pareto}. Note that,
-  while $\hat \phi$ changes nearly in the same way in both graphs, the
+  while $\hat \phi$ changes nearly in the same in both graphs, the
   scales in the graphcs are very different.
@@
   increasing.
-\item In other case, $U_{\text{all}}(\cdot)$ is not monotone.
+\item In other cases, $U_{\text{all}}(\cdot)$ is not monotone.
 \end{enumerate}

### 2017-08-03 cpp/ImportanceSampling/GarchPQ.cpp
Coded the importance sampling algo.

@@ -150,8 +150,8 @@ double GarchPQ::find_tail_index
 	lb = gsl_root_fsolver_x_lower(solver);
 	ub = gsl_root_fsolver_x_upper(solver);
 	status = gsl_root_test_interval(lb, ub, 0.0, 1.0e-2);
-	if (status == GSL_SUCCESS)
-	    cout << "Tail index found: xi = " << xi << endl;
+	// if (status == GSL_SUCCESS)
+	//     cout << "Tail index found: xi = " << xi << endl;
     } while (status == GSL_CONTINUE && iter < max_iter);
     gsl_root_fsolver_free(solver);
     if (status != GSL_SUCCESS) {

### 2017-08-03 cpp/ImportanceSampling/garch21.cpp
Implemented the GARCH(2,1) importance sampling algorithm.

@@ -35,12 +35,6 @@ double interpolate_fun
     }
 }
 
-// Garch21::chain_state::chain_state(const vec &v)
-//     :angle(atan(v(1)/v(0))), additive(log(norm(v)))
-// {
-//     copy(v.begin(), v.end(), V.begin());
-// }
-
 void Garch21::compute_eigenfunctions(void)
 {
     eigenfunctions.resize(nbr_eigenfunctions);
@@ -111,32 +105,29 @@ Garch21::Garch21(const vector<double> &alpha,
     chi_squared_distribution<double> chi2;
     // Prepare the ppol
     pool.resize(10000u);
+    A_matrices.resize(pool.size());
+    
 #pragma omp parallel for
-    for (auto i = pool.begin(); i != pool.end(); ++i) {
-	*i = chi2(randev);
+    for (size_t i = 0; i < pool.size(); ++i) {
+	pool[i] = chi2(randev);
+	A_matrices[i].resize(2, 2);
+	gen_rand_matrix(pool[i], A_matrices[i]);
     }
     sort(pool.begin(), pool.end());
-    r_xi.resize(1000);
+
+    r_xi.resize(100);
     right_eigenfunction(tail_index, r_xi);
 }
 
 double Garch21::quantile(double u, double angle) const
 {
     assert(u > 0 && u < 1);
-    vector<funval> equantile(pool.size());
+    vector<double> ecdf(pool.size());
     vec x = {cos(angle), sin(angle)};
-    static vector<mat> matrices;
     
-    if (matrices.empty()) {
-#pragma omp parallel for
-	for (unsigned int i = 0; i < pool.size(); ++i) {
-	    gen_rand_matrix(pool[i], matrices[i]);
-	}
-    }
-
 #pragma omp parallel for
     for (unsigned int i = 0; i < pool.size(); ++i) {
-	vec y = matrices[i] * x;
+	vec y = A_matrices[i] * x;
 	double len = norm(y);
 	double theta = acos(y[0]/len);
 	double inc =
@@ -144,26 +135,27 @@ double Garch21::quantile(double u, double angle) const
 	    interpolate_fun(theta, r_xi) /
 	    interpolate_fun(angle, r_xi) /
 	    pool.size();
-	equantile[i][1] = pool[i];
-	equantile[i][0] = inc;
+	ecdf[i] = inc;
     }
[…]

### 2017-08-04 phd/Chapterintroduction.tex
Some minor arrangements.

@@
 
-Kesten \cite{kesten:1973} showed that, when $A_t$ and $B_t$ were almost
-surely non-negative, had no row or column of only zeros, and there was
-a positive probability that $B_t$ was strictly positive, the strictly
-stationary solution to the equation
-$V \overset{d}{=} A V + B$ had power-law tails
+Kesten \cite{kesten:1973} showed that, when $A_t$ are almost
+surely non-negative, has no row or column of only zeros, and
+$B_t$ is almost surely non-negative and is not equal to the null
+vector with probability 1, then the strictly stationary solution to
+the equation $V \overset{d}{=} A V + B$ has power-law tails
 for its marginal distributions, assuming the following conditions (M)
@@
 \end{equation}
-where $C$ is a constant (cf. eq.(2.15) of Collamore and Mentemeier
+where $C$ is a constant (cf. (2.15) of Collamore and Mentemeier
 \cite{collamore:mentemeier:2016}), $f(\cdot)$ is any bounded

### 2017-08-04 phd/Chapterintroduction.tex
Revised the PhD thesis.

@@
 \chapter{Introduction}\label{ch:intr}
+The phenomena of heavy-tailedness is widely observed in many
+disciplins of science, for example, phase transition of matter and
+black body radiation as studied in physics, neuronal avalanches in
+biology, claim sizes of insurance mathematics and stock returns in
+finance. The last application is indeed the focus of this thesis. To
+discuss the phenomena in precise terms, we introduce the concept of
+regular variation.
 
@@
 The concept of regular variation is defined by the following scaling
-property: If a function $f(\cdot)$ satisfies
+property: if a function $f(\cdot)$ satisfies
 \[
 \lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
+\quad
+\forall c > 0
 \]
 then we say $f(x)$ is regularly varying with index $\alpha$.
-Distribution functions, say $F(\cdot)$, whose survival function
-$\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
-observed in a variety of phenomena, for example, phase transition of
-matter and black body radiation as studied in physics, neuronal
-avalanches in biology, claim sizes of insurance mathematics and stock
-returns in finance. The last application is indeed the focus of this
-thesis.
-
+% Distribution functions, say $F(\cdot)$, whose survival function
+% $\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
+% observed in a variety of phenomena, 
 When expanded to multiple dimensions, the scaling property of regular
-variation is better described in terms of weak convergence to a
-spectral measure: If a random vector $V$ satisfies
+variation is better described in terms of vague convergence to a
+spectral measure $\mu_\alpha$: if a random vector $V$ satisfies
 \[
-\lim_{x \to \infty}
 {
@@
   \P(|V| > x)
-}
-\overset{w}{\to} c^{-\alpha} \mu_\alpha(\cdot)
+},
+\overset{v}{\to} c^{-\alpha} \mu_\alpha(\cdot)
+\quad
+x \to \infty, \forall c > 0
 \]
-then we say the components of $V$ is jointly regularly varying with
-index $\alpha$. Here $\mu_\alpha(\cdot)$ is a probability measure on
-the unit sphere \cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
-is jointly regularly varying with index $\alpha$, then each component
+then we say the $V$ is regularly varying with index $\alpha$. Here
+$\mu_\alpha(\cdot)$ is a probability measure on the unit sphere
+\cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
+is regularly varying with index $\alpha$, then each component
 of it and each linear combination of its components are regularly
@@
 \cite{feller}, p. 278. Cf. also Jessen and Mikosch
-\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et al
+\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et. al.
 \cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.
 
-Thus estimating the tail index $\alpha$ is particularly important for
-understanding the behaviour of a heavy-tailed series. The earliest
+Clearly, estimating the tail index $\alpha$ is particularly important
+for understanding the behaviour of a heavy-tailed series. A standard
 method proposed for this purpose is due to Hill \cite{hill1975simple}:
@@
 
-When considered as the multiplicative inverse of the parameter of the
-{\em Generalized Extreme Value} distribution, there are other methods
-in the literature for estimating the tail index, e.g. Pickand's 
-estimator \cite{pickands1975statistical} $\hat \alpha_P$ and the
-Deckers-Einmahl-de Haan estimator $\hat \alpha_{\text{DEH}}$
-\begin{eqnarray*}
-{1 \over \hat \alpha_P}
-&=&
-{1 \over \log 2}
-\log {
-  X_{(k)} - X_{(2k)}
-  \over
-X_{(2k)} - X_{(4k)}  
-} \\
-{1 \over \hat \alpha_{\text{DEH}}}
-&=&
-1 + {1 \over \alpha_H} + {1 \over 2} \left(
-  {H \over \alpha_H^2} - 1
-\right)^{-1}
-\end{eqnarray*}
-where it is similarly assumed $k \to \infty$ and $k/n \to 0$ as $n \to
-\infty$. The Deckers-Einmahl-de Haan estimator makes use of Hill's
-estimator $\hat \alpha_H$ and computes
-\[
-H = \left[
-  {1 \over k} \sum_{i=1}^k \log \left(
-    { X_{(i)} \over X_{(k+1)} }
-  \right)^2
-\right]^{-1}
-\]
-Apparently $1/H$ can be interpreted as the 2nd empirical moment of
-$\log(X_{(i)}/X_{(k+1)})$ for $i \leq k$.
-A major drawback of Pickand's and Deckers-Einmahl-de Haan's estimators
-is that, when applied to estimating the tail index, they discard the
-information that the tail index is always positive, hence resulting in
-a larger confidence band compared with that obtained for Hill's
-estimator. Therefore we stick to Hill's estimator in the empirical
-work included in this thesis.
+% When considered as the multiplicative inverse of the parameter of the
+% {\em Generalized Extreme Value} distribution, there are other methods
+% in the literature for estimating the tail index, e.g. Pickand's 
+% estimator \cite{pickands1975statistical} $\hat \alpha_P$ and the
+% Deckers-Einmahl-de Haan estimator $\hat \alpha_{\text{DEH}}$
+% \begin{eqnarray*}
+% {1 \over \hat \alpha_P}
+% &=&
+% {1 \over \log 2}
+% \log {
+%   X_{(k)} - X_{(2k)}
+%   \over
+% X_{(2k)} - X_{(4k)}  
+% } \\
+% {1 \over \hat \alpha_{\text{DEH}}}
+% &=&
+% 1 + {1 \over \alpha_H} + {1 \over 2} \left(
+%   {H \over \alpha_H^2} - 1
+% \right)^{-1}
+% \end{eqnarray*}
+% where it is similarly assumed $k \to \infty$ and $k/n \to 0$ as $n \to
+% \infty$. The Deckers-Einmahl-de Haan estimator makes use of Hill's
+% estimator $\hat \alpha_H$ and computes
+% \[
+% H = \left[
+%   {1 \over k} \sum_{i=1}^k \log \left(
+%     { X_{(i)} \over X_{(k+1)} }
+%   \right)^2
+% \right]^{-1}
+% \]
+% Apparently $1/H$ can be interpreted as the 2nd empirical moment of
+% $\log(X_{(i)}/X_{(k+1)})$ for $i \leq k$.
+% A major drawback of Pickand's and Deckers-Einmahl-de Haan's estimators
+% is that, when applied to estimating the tail index, they discard the
+% information that the tail index is always positive, hence resulting in
+% a larger confidence band compared with that obtained for Hill's
+% estimator. Therefore we stick to Hill's estimator in the empirical
+% work included in this thesis.
 

### 2017-08-04 phd/MyThesis.tex
Some minor arrangements.
Revised the PhD thesis.

@@
 & \\
-Assessment committee: &  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
-& University of Neuchatel, CH\\
+
+Assessment committee:  & Lecturer.~Jesper Lund Pedersen (chairman)\\
+& University of Copenhagen\\
+& \\
+
+&  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
+& University of Neuchatel, CH \\
 & \\
@@
 & Prof.~Olivier Wintenberger\\
-& University of Marie \& Pierre Curie, Paris\\
-& \\
-
-& Lecturer.~Jesper Lund Pedersen (chairman)\\
-& University of Copenhagen\\
+& University of Marie \& Pierre Curie, Paris
 \end{tabular}
@@
 \input{ChapterImportanceSampling}
+\input{ChapterBernoulli}
 \input{Chapterextremes}
-\input{ChapterBernoulli}
 

### 2017-08-07 phd/Chapterintroduction.tex
progress on the importance sampling paper. Added to the introduction section of the thesis.

@@
 The phenomena of heavy-tailedness is widely observed in many
-disciplins of science, for example, phase transition of matter and
+disciplines of science, for example, phase transition of matter and
 black body radiation as studied in physics, neuronal avalanches in
@@
 
+Figure \ref{fig:thjyuj} shows the Hill estimates of the tail indices
+of daily stock return series from 3 sectors of the
+{\em Standard \& Poor 500} index \footnote{An American stock index
+  comprising 500 or so companies.}. The 2.5\% and 97.5\% quantiles of
+the asymptotic normal distribution of the estimates are also
+given. One can see the confidence bands are fairly large compared with
+the estimated values. This certainly raises the question of how
+similar they really are and if/how their variations can be explained by
+economical arguments. We explore this topic in chapter
+\ref{ch:TailParameters}.
+
+\begin{figure}[htb!]
+  \begin{minipage}{1.0\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+    {Energy_lower.pdf}
+  \end{minipage}
+  \begin{minipage}{1.0\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+    {Consumer_Staples_lower.pdf}
+  \end{minipage}
+  \begin{minipage}{1.0\linewidth}
+    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
+    {Information_Technology_lower.pdf}
+  \end{minipage}
+  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower
+    tail-indices $\alpha$ of daily return series in sectors of the S\&P 500
+    index. The data span from 1 January 2010 to 31 December 2014 and
+    comprise $n=1304$ observations.
+    The graphs from top to bottom correspond to the ``Energy'',
+    ``Consumer Staples'' and ``Information Technology'' sectors.
+    Each circle corresponds to a Hill estimate $\hat\alpha_{50}$; the gray
+    triangles above and below it mark the 97.5\% and 2.5\% quantiles
+    of its approximate normal distribution; see \eqref{eq:2} and the
+    discussion following it for an interpretation.
+    The lower and upper red lines mark the medians of the 2.5\% 
+    and 97.5\% quantiles, respectively, evaluated from all stocks in
+    the sector.
+    The data are taken from {\it Yahoo Finance}; the labels on
+    the horizontal axes are Yahoo symbols of the stocks. 
+  }\label{fig:thjyuj}
+\end{figure}
+
+Random variables with regularly varying tails have some very nice
+features: if $X_1$ and $X_2$ are both regularly varying with indices
+$\alpha_1$ and $\alpha_2$, respectively, then $a_1 X_1 + a_2 X_2$ is
+regularly varying with index $\min\{\alpha_1, \alpha_2\}$. Moreover
+if $X_1, X_2$ are iid,
+$\P(a_1 X_1 + a_2 X_2 > u) \sim \P(a_1 X_1 > u) + \P(a_2 X_2 > u)$.
+
+Now consider $N$ return series $X_{i,t}$, $i=1,2,\dots, p$. Suppose
+each of these series is a linear combination of $K$ factors $Y_{i,t}$,
+$i=1,2,\dots,K$, the $i$-th of which is regularly varying with index
+$\alpha_i$. Then by the summation property, each and every $X_i$ is
+regularly varying with index $\min_{1 \leq i \leq K} \alpha_i$.
+Since in practice a factor $Y_{i,t}$ is found as a linear combination
+of $X_{i,t}$, the observed time series, following an eigenvector of
+the sample covariance matrix of $X_{i,t}$, it is important to
+understand the eigensystem of this matrix. This topic indeed
+constitutes chapters \ref{ch:bernoulli} and \ref{ch:extremes} of this
+thesis.
+
+When a product of random variables, say $X_1 X_2$, involves a random
+variable with regularly varying tails, a useful result is that of
+Breiman: assume $X_1$ is regularly varying with index $\alpha$ and
+there exists $\epsilon > 0$ such that
+$\E |X_2|^{\alpha + \epsilon} < \infty$.
+Then $\P(X_1 X_2 > x) \sim \E|X_2|^{\alpha} \P(X_1 > x)$.
+More generally, if $X_1, X_2$ are regularly varying with the same tail
+index $\alpha$ or if $\P(X_2 > x) = o(\P(X_1 > x))$, then $X_1 X_2$ is
+regularly varying with index $\alpha$.
+
+For an extensive summary of the regular variation properties of
+functions of regularly varying random variables, see Mikosch and
+Jessen \cite{JessenMikosch2006}.
+
 % When considered as the multiplicative inverse of the parameter of the
@@
 varying r.v. is stochastic recursion of the following form:
-\[
+\begin{equation}
+  \label{eq:rhjyu}
   X_t = A_t X_{t-1} + B_t
-\]
+\end{equation}
 where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
@@
 
-Recently, Collamore and Mentemeier \cite{collamore:mentemeier:2016}
-extended Kesten's result and gave an explicit expression for $\mu_\xi$:
-\begin{equation}
-  \label{eq:CollamoreMentemeierIntro}
-  \lim_{u \to \infty} u^{\xi} \E \left[
-    f(u^{-1} V)
-    \right]
-  =
-  {C \over \lambda'(\xi)}  
-  \int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
-\end{equation}
-where $C$ is a constant (cf. (2.15) of Collamore and Mentemeier
-\cite{collamore:mentemeier:2016}), $f(\cdot)$ is any bounded
-continuous function on $\reals^{d}_+ \setminus \{0\}$  and
-$\ell_\xi$ is a probability measure on $\sphere_+^{d-1}$. Its definition
-is also found in \eqref{eq:eigenmeasure}.
+%% Recently, Collamore and Mentemeier \cite{collamore:mentemeier:2016}
+%% extended Kesten's result and gave an explicit expression for $\mu_\xi$:
+%% \begin{equation}
+%%   \label{eq:CollamoreMentemeierIntro}
+%%   \lim_{u \to \infty} u^{\xi} \E \left[
+%%     f(u^{-1} V)
+%%     \right]
+%%   =
+%%   {C \over \lambda'(\xi)}  
+%%   \int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
+%% \end{equation}
+%% where $C$ is a constant (cf. (2.15) of Collamore and Mentemeier
+%% \cite{collamore:mentemeier:2016}), $f(\cdot)$ is any bounded
+%% continuous function on $\reals^{d}_+ \setminus \{0\}$  and
+%% $\ell_\xi$ is a probability measure on $\sphere_+^{d-1}$. Its definition
+%% is also found in \eqref{eq:eigenmeasure}.
 
-From \eqref{eq:CollamoreMentemeierIntro} a representation for $\mu_\xi
-(\cdot)$ immediately follows  
+%% From \eqref{eq:CollamoreMentemeierIntro} a representation for $\mu_\xi
+%% (\cdot)$ immediately follows  
+%% \[
+%% \mu_\xi (\cdot) = {C \over \lambda'(\xi)} \mathcal L_\xi(\cdot)
+%% \]
+%% Here $\mathcal L_\xi$ is a non-null Radon measure  on
+%% $\reals^d_+ \setminus \{0\}$ that satisfies, for all
+%% bounded continuous function $f(\cdot)$ on
+%% $\reals_+^{d} \setminus \{0\}$:
+%% \[
+%% \int_{\reals_+^d\setminus \{0\}} f(x) \mathcal L_\xi(dx)
+%% =
+%% \int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
+%% \]
+
+\section{GARCH models}
+Introduced by Bollerslev in 1986, GARCH models have been hugely
+popular in modelling volatility of financial time series and have
+inspired numerous variants. Its original form is the following the
+recurrence equation:
 \[
-\mu_\xi (\cdot) = {C \over \lambda'(\xi)} \mathcal L_\xi(\cdot)
+\sigma_t^2 = \omega + \sum_{i=1}^p \alpha_i R_{t-i}^2 +
+\sum_{j=1}^q \beta_j \sigma_{t-j}^2
 \]
-Here $\mathcal L_\xi$ is a non-null Radon measure  on
-$\reals^d_+ \setminus \{0\}$ that satisfies, for all
-bounded continuous function $f(\cdot)$ on
-$\reals_+^{d} \setminus \{0\}$:
+where $\R_t$ is a return series, e.g. stock returns, foreign exchange
+rates, interest rates, etc; $\sigma_t^2$ is the variance of the
+distribution of $R_t$ conditional on $\{(R_i,
+\sigma_i^2)\}_{i=0}^{t-1}$. $\omega, \{\alpha_i\}_{i=1}^p,
+\{\beta_i\}_{i=1}^q$ are constant parameters of the model. Clearly,
+the GARCH($p,q$) recurrence equation is of the form of
+\eqref{eq:rhjyu}. So, with appropriate conditions, including
+$\sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1$, $\sigma_t^2$ is
+shown to be a positive Harris Markov chain (cf. Bollerslev
+\cite{bollerslev:1986} and Buraczewski et al
+\cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
+has regularly varying tails. The tail index, call it $\xi$, is given by
 \[
-\int_{\reals_+^d\setminus \{0\}} f(x) \mathcal L_\xi(dx)
-=
-\int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
+\lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\xi = 0
+\]
+where $A_1, A_2, \dots, A_n$ are iid matrices whose entries are
+functions of $\{\alpha_i\}_{i=1}^p$ and $\{\beta_i\}_{i=1}^q$:
+\[
+A_i =
+\begin{pmatrix}
+  \alpha_1 Z_{t-1}^2 + \beta_1 & \beta_2 & \cdots &
+  \beta_{q-1} & \beta_q & \alpha_2 & \alpha_3 &
+  \cdots & \alpha_{p-1} & \alpha_p\\
+  1 & 0 & \cdots & 
+  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
+  \vdots & \vdots & \ddots & 
+  \vdots & \vdots & \vdots & \vdots &
+  \ddots & \vdots & \vdots \\
+  0 & 0 & \cdots &
+  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
+  0 & 0 & \cdots &
+  1 & 0 & 0 & 0 & \cdots & 0 & 0 \\
+  Z_{t-1}^2 & 0 & \cdots &
+  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
+  0 & 0 & \cdots &
+  0 & 0 & 1 & 0 & \cdots & 0 & 0 \\
+  \vdots & \vdots & \ddots &
+  \vdots & \vdots & \vdots & \vdots &
+  \ddots & \vdots & \vdots \\
+  0 & 0 & \cdots &
+  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\    
+  0 & 0 & \cdots &
+  0 & 0 & 0 & 0 & \cdots & 1 & 0 \\    
+\end{pmatrix}
+\begin{pmatrix}
+  \sigma_{t-1}^2 \\
+  \sigma_{t-2}^2 \\
+  \vdots \\
+  \sigma_{t-q+1}^2 \\
+  \sigma_{t-q}^2 \\
+  R_{t-2}^2 \\
+  R_{t-3}^2 \\
+  \vdots \\
+  R_{t-p+1}^2 \\
+  R_{t-p}^2
+\end{pmatrix}
 \]

### 2017-08-07 cpp/ImportanceSampling/garch21.hpp

@@ -41,7 +41,7 @@ public:
     double draw_z2(double) const;
     void simulate_path(vector<vec> &path) const;
     pair<double, size_t> sample_estimator(const vec &V0, double u);
-    double estimate_prob(double u);
+    array<double,2> estimate_prob(double u, size_t);
 
     double Lyapunov(size_t n, size_t iterations);
 

### 2017-08-07 cpp/ImportanceSampling/main.cpp

@@ -172,10 +172,12 @@ int main(int argc, char* argv[])
      * Simulate GARCH(1,1) processes.
      */
     double coef[] = {
-	1.0e-7, 0.11, 0.88
+	// 1.0e-7, 0.11, 0.88
+	5.641683e-06, 1.611969e-01, 7.287080e-01 
     };
     Garch1D<double> garch11(coef[0], coef[1], coef[2], stol(argv[1]));
-    printf("Lambda(%.4f) = %.4f\n", garch11.xi, log(garch11.moment_func(garch11.xi, 0, 0)));
+    printf("Lambda(%.4f) = %.4f\n",
+	   garch11.xi, log(garch11.moment_func(garch11.xi, 0, 0))); 
     for (double x = stod(argv[2]); x <= stod(argv[3]); x += 0.1) {
 	printf("%.4f    %.4f\n", x, log(garch11.moment_func(x)));
     }

### 2017-08-07 phd/Chapterintroduction.tex
Finished the description of the algorithm.

@@
 random matrix and $B_t$ is a $d$-dimensional vector, random or
-deterministic. The sequence $\{A_i\}_{i=1,2,\dots}$ and
-$\{B_i\}_{i=1,2,\dots}$  are iid and independent of each other.
+deterministic.
 
@@
 the GARCH($p,q$) recurrence equation is of the form of
-\eqref{eq:rhjyu}. So, with appropriate conditions, including
+\eqref{eq:rhjyu}. So, with appropriate conditions including
 $\sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1$, $\sigma_t^2$ is
@@
 \end{pmatrix}
-\begin{pmatrix}
-  \sigma_{t-1}^2 \\
-  \sigma_{t-2}^2 \\
-  \vdots \\
-  \sigma_{t-q+1}^2 \\
-  \sigma_{t-q}^2 \\
-  R_{t-2}^2 \\
-  R_{t-3}^2 \\
-  \vdots \\
-  R_{t-p+1}^2 \\
-  R_{t-p}^2
-\end{pmatrix}
 \]

### 2017-08-08 phd/Chapterintroduction.tex
Corrected the introduction chapter of the thesis.

@@
 The concept of regular variation is defined by the following scaling
-property: if a function $f(\cdot)$ satisfies
+property: if a function $f$ on $(0, \infty)$ satisfies
 \[
@@
 \]
-then we say $f(x)$ is regularly varying with index $\alpha$.
+then we say $f$ is regularly varying with index $\alpha$.
+$f$ can be written in the form $f(x) = {L(x) \over x^\alpha}$, where
+$L(x)$ is a slowly varying function, i.e.
+$\lim_{x \to \infty} {L(c x) \over L(x)} = 1$, $\forall c > 0$.
+We call a random variable $X$ regularly varying with index
+$\alpha \geq 0$ if it satisfies the tail balance condition
+\[
+\P(X > x) \sim p_+ {L(x) \over x^\alpha},\quad
+\P(X < -x) \sim p_- {L(x) \over x^\alpha}, \quad
+\text{for some } p_+, p_- > 0,\; p_+ + p_- = 1
+\]
+
 % Distribution functions, say $F(\cdot)$, whose survival function
@@
   \P(|V| > x)
-},
+}
 \overset{v}{\to} c^{-\alpha} \mu_\alpha(\cdot)
 \quad
-x \to \infty, \forall c > 0
+x \to \infty, \forall c > 0,
 \]
-then we say the $V$ is regularly varying with index $\alpha$. Here
-$\mu_\alpha(\cdot)$ is a probability measure on the unit sphere
+then we say $V$ is regularly varying with index $\alpha$. Here
+$\mu_\alpha$ is a probability measure on the unit sphere
 \cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
 is regularly varying with index $\alpha$, then each component
-of it and each linear combination of its components are regularly
+and each linear combination of its components are regularly
 varying with the same index $\alpha$. This follows from Feller
@@
 
-Clearly, estimating the tail index $\alpha$ is particularly important
-for understanding the behaviour of a heavy-tailed series. A standard
-method proposed for this purpose is due to Hill \cite{hill1975simple}:
+Clearly, estimating the tail index $\alpha$ of a sequence
+$X_1, X_2, \dots$ of regularly varying variables is particularly
+important for understanding the behaviour of a heavy-tailed series. A
+standard method proposed for this purpose is due to Hill
+\cite{hill1975simple}:
 \[
@@
 \]
-where $X_1, X_2, \dots$ is the series whose tail index is the subject
-of estimation, and $X_{(i)}$ is the $i$-th upper order statistics of
-this series. Several authors have contributed to showing the
-weak consistency of the estimator $\hat \alpha_H$,
-i.e. if $k \to \infty, k/n \to 0$ as $n \to \infty$, then
-$\hat \alpha_H \overset{P}{\to} \alpha$.
+where $X_1, X_2, \dots, X_n$ is a sample whose tail index is the
+subject of estimation, and $X_{(i)}$ is the $i$-th upper order
+statistic of the sample. Several authors have contributed to showing
+the weak consistency and asymptotic normality of the estimator
+$\hat \alpha_H$, under the assumptions $k \to \infty, k/n \to 0$
+as $n \to \infty$.
 
-Mason \cite{mason:1982} first proved that the estimator
-was consistent when $X_1, X_2, \dots$ were iid; later Rootz\'en,
-Leadbetter and de Haan \cite{rootzen:leadbetter:dehaan1992} and Hsing
-\cite{hsing:1991} proved its consistency when $X_1, X_2, \dots$ were
-weakly dependent; and Resnick and St{\u a}ric{\u a}
-\cite{resnick:starica:1995, resnick:starica:1997} proved its
-consistency when $X_t$ was a linear process.
+%% Mason \cite{mason:1982} first proved that the estimator
+%% was consistent when $X_1, X_2, \dots$ were iid; later Rootz\'en,
+%% Leadbetter and de Haan \cite{rootzen:leadbetter:dehaan1992} and Hsing
+%% \cite{hsing:1991} proved its consistency when $X_1, X_2, \dots$ were
+%% weakly dependent; and Resnick and St{\u a}ric{\u a}
+%% \cite{resnick:starica:1995, resnick:starica:1997} proved its
+%% consistency when $X_t$ was a linear process.
 
@@
 
-Now consider $N$ return series $X_{i,t}$, $i=1,2,\dots, p$. Suppose
-each of these series is a linear combination of $K$ factors $Y_{i,t}$,
-$i=1,2,\dots,K$, the $i$-th of which is regularly varying with index
+Now consider $p$ return series $X_{i,t}$, $i=1,2,\dots, p$. Suppose
+each of these series is a linear combination of $K$ factors $Y_{j,t}$,
+$j=1,2,\dots,K$, the $j$-th of which is regularly varying with index
 $\alpha_i$. Then by the summation property, each and every $X_i$ is

### 2017-08-08 phd/Chapterintroduction.tex
Wrote some texts in the introduction section of the thesis.

@@
 \cite{hill1975simple}:
-\[
+\begin{equation}
+  \label{eq:fhhty}
   \hat \alpha_H = \left[
@@
   \right]^{-1}
-\]
+\end{equation}
 where $X_1, X_2, \dots, X_n$ is a sample whose tail index is the
@@
 $\alpha_1$ and $\alpha_2$, respectively, then $a_1 X_1 + a_2 X_2$ is
-regularly varying with index $\min\{\alpha_1, \alpha_2\}$. Moreover
-if $X_1, X_2$ are iid,
+regularly varying with index $\min\{\alpha_1,
+\alpha_2\}$ (cf. Mikosch and Jessen
+\cite{JessenMikosch2006}). Moreover, if $X_1, X_2$ are iid,
 $\P(a_1 X_1 + a_2 X_2 > u) \sim \P(a_1 X_1 > u) + \P(a_2 X_2 > u)$.
 
-Now consider $p$ return series $X_{i,t}$, $i=1,2,\dots, p$. Suppose
-each of these series is a linear combination of $K$ factors $Y_{j,t}$,
-$j=1,2,\dots,K$, the $j$-th of which is regularly varying with index
-$\alpha_i$. Then by the summation property, each and every $X_i$ is
-regularly varying with index $\min_{1 \leq i \leq K} \alpha_i$.
-Since in practice a factor $Y_{i,t}$ is found as a linear combination
-of $X_{i,t}$, the observed time series, following an eigenvector of
-the sample covariance matrix of $X_{i,t}$, it is important to
-understand the eigensystem of this matrix. This topic indeed
-constitutes chapters \ref{ch:bernoulli} and \ref{ch:extremes} of this
-thesis.
+Now consider $p$ return series
+$X_{i,t}$, $i=1,2,\dots, p; t=1,\dots,n$.
+Suppose each of these series is a linear combination of $K$ factors
+$Y_{j,t}$, $j=1,2,\dots,K$, the $j$-th of which is regularly varying
+with index $\alpha_j$. Then by the summation property, each and every
+$X_i$ is regularly varying with index $\min_{1 \leq j \leq K} \alpha_j$.
+In practice a factor $Y_{i,t}$ is estimated as
+$\hat Y_{j,t} = \sum_{i=1}^p a_{j, i} X_{i,t}$, where
+$(a_{j, 1}, a_{j, 2}, \dots, a_{j, p})^\top$ is the $j$-th eigenvector
+of the sample covariance matrix of
+$\{X_{i,t}\}, i=1,\dots,p; t=1,\dots,n$, i.e. the eigenvector
+associated with the $j$-th largest eigenvalue. 
+For this reason, it is important to understand the eigensystem of the
+sample covariance matrix of $\{X_{i,t}\}$. This topic is addressed in
+chapters \ref{ch:bernoulli} and \ref{ch:extremes} of this thesis.
 
-When a product of random variables, say $X_1 X_2$, involves a random
-variable with regularly varying tails, a useful result is that of
-Breiman: assume $X_1$ is regularly varying with index $\alpha$ and
-there exists $\epsilon > 0$ such that
-$\E |X_2|^{\alpha + \epsilon} < \infty$.
-Then $\P(X_1 X_2 > x) \sim \E|X_2|^{\alpha} \P(X_1 > x)$.
+When a product of positive random variables, say $X_1 X_2$, involves
+one with regularly varying tails, a useful result is that of
+Breiman \cite{breiman:1965}: assume $X_1$ is regularly varying with
+index $\alpha$ and there exists $\epsilon > 0$ such that
+$\E X_2^{\alpha + \epsilon} < \infty$.
+Then $\P(X_1 X_2 > x) \sim \E X_2^{\alpha} \P(X_1 > x)$.
 More generally, if $X_1, X_2$ are regularly varying with the same tail
@@
   \label{eq:rhjyu}
-  X_t = A_t X_{t-1} + B_t
+  X_t = A_t X_{t-1} + B_t, \quad t \in \mathbb Z
 \end{equation}
@@
 random matrix and $B_t$ is a $d$-dimensional vector, random or
-deterministic.
+deterministic. The sequence $\{(A_t, B_t)\}_{t \in \mathbb Z}$ are
+iid. The stationary solution to \eqref{eq:rhjyu} satisfies the fixed-point
+equation $X \overset{d}{=} A X + B$, where $X$ and $(A, B)$ are
+generic elements of the $\{X_t\}_{t \in \mathbb Z}$ and
+$(A_t, B_t)_{t \in \mathbb Z}$ sequences.
 
@@
 $B_t$ is almost surely non-negative and is not equal to the null
-vector with probability 1, then the strictly stationary solution to
-the equation $V \overset{d}{=} A V + B$ has power-law tails
-for its marginal distributions, assuming the following conditions (M)
-and (A):
+vector with probability 1, then the solution $X$ to the equation
+$X \overset{d}{=} A X + B$ is regularly varying with a positive index
+$\alpha$, assuming the following conditions (M) and (A):
 \begin{itemize}
@@
     is negative.
-  \item There exists $\xi > 0$ such that
+  \item There exists $\alpha > 0$ such that
     \[
-    1 = \lambda(\xi) = \lim_{n \to \infty} {1 \over n} \log \E \|A_n \cdots A_1\|^\xi
+    1 = \lambda(\alpha) = \lim_{n \to \infty} {1 \over n} \log \E \|A_n \cdots A_1\|^\alpha
     \]
-  \item $\E (\|A_1\|^\xi \log^+\|A_1\|) < \infty$
-  \item $\E |B_1|^\xi < \infty$
+  \item $\E (\|A_1\|^\alpha \log^+\|A_1\|) < \infty$
+  \item $\E |B_1|^\alpha < \infty$
   \end{enumerate}
@@
 \begin{equation*}
-  u^\xi \P(u^{-1} V \in \cdot) \overset{v}{\to} \mu_\xi(\cdot)
+  u^\alpha \P(u^{-1} V \in \cdot) \overset{v}{\to} \nu_\alpha(\cdot)
 \end{equation*}
-where $\mu_\xi$ is a non-null Radon measure on
+where $\nu_\alpha$ is a non-null Radon measure on
 $\reals^d_+ \setminus \{0\}$ with the property
-$\mu_\xi(a A) = a^{-\xi} \mu_\xi(A)$.
-
-%% Recently, Collamore and Mentemeier \cite{collamore:mentemeier:2016}
-%% extended Kesten's result and gave an explicit expression for $\mu_\xi$:
-%% \begin{equation}
-%%   \label{eq:CollamoreMentemeierIntro}
-%%   \lim_{u \to \infty} u^{\xi} \E \left[
-%%     f(u^{-1} V)
-%%     \right]
-%%   =
-%%   {C \over \lambda'(\xi)}  
-%%   \int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
-%% \end{equation}
-%% where $C$ is a constant (cf. (2.15) of Collamore and Mentemeier
-%% \cite{collamore:mentemeier:2016}), $f(\cdot)$ is any bounded
-%% continuous function on $\reals^{d}_+ \setminus \{0\}$  and
-%% $\ell_\xi$ is a probability measure on $\sphere_+^{d-1}$. Its definition
-%% is also found in \eqref{eq:eigenmeasure}.
-
-%% From \eqref{eq:CollamoreMentemeierIntro} a representation for $\mu_\xi
-%% (\cdot)$ immediately follows  
-%% \[
-%% \mu_\xi (\cdot) = {C \over \lambda'(\xi)} \mathcal L_\xi(\cdot)
-%% \]
-%% Here $\mathcal L_\xi$ is a non-null Radon measure  on
-%% $\reals^d_+ \setminus \{0\}$ that satisfies, for all
-%% bounded continuous function $f(\cdot)$ on
-%% $\reals_+^{d} \setminus \{0\}$:
-%% \[
-%% \int_{\reals_+^d\setminus \{0\}} f(x) \mathcal L_\xi(dx)
-%% =
-%% \int_{\sphere^{d-1}_+ \times R} e^{-\xi s} f(e^s x) \ell_\xi(dx) ds
-%% \]
+$\nu_\alpha(a A) = a^{-\alpha} \nu_\alpha(A)$.
 
 \section{GARCH models}
-Introduced by Bollerslev in 1986, GARCH models have been hugely
-popular in modelling volatility of financial time series and have
-inspired numerous variants. Its original form is the following the
-recurrence equation:
-\[
-\sigma_t^2 = \omega + \sum_{i=1}^p \alpha_i R_{t-i}^2 +
+Introduced by Bollerslev \cite{bollerslev:1986} in 1986,
+{\em Generalized Autoregressive Conditional Heteroscedasticity}
+(GARCH) models have been hugely popular for modelling volatility of
+financial time series and have inspired numerous variants.
+A GARCH($p, q$) model is a stationary process
+$\{X_t\}_{t \in \mathbb Z}$satisfying
+\begin{eqnarray*}
+  X_t &=& \sigma_t Z_t \\
+  \sigma_t^2 &=& \omega + \sum_{i=1}^p \alpha_i X_{t-i}^2 +
   \sum_{j=1}^q \beta_j \sigma_{t-j}^2
-\]
-where $\R_t$ is a return series, e.g. stock returns, foreign exchange
-rates, interest rates, etc; $\sigma_t^2$ is the variance of the
-distribution of $R_t$ conditional on $\{(R_i,
-\sigma_i^2)\}_{i=0}^{t-1}$. $\omega, \{\alpha_i\}_{i=1}^p,
-\{\beta_i\}_{i=1}^q$ are constant parameters of the model. Clearly,
-the GARCH($p,q$) recurrence equation is of the form of
-\eqref{eq:rhjyu}. So, with appropriate conditions including
-$\sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1$, $\sigma_t^2$ is
-shown to be a positive Harris Markov chain (cf. Bollerslev
-\cite{bollerslev:1986} and Buraczewski et al
+\end{eqnarray*}
+where $\{X_t\}$ is a return series, e.g. stock returns, foreign exchange
+rates, interest rates, etc; $\{Z_t\}$ is a iid, mean 0, unit-variance
+sequence; and $\sigma_t^2$ is the variance of the
+distribution of $X_t$ conditional on $\{(X_i,
+\sigma_i^2)\}_{i=0}^{t-1}$; $\omega, \{\alpha_i\}_{i=1}^p,
+\{\beta_i\}_{i=1}^q$ are non-negative parameters of the
+model. The GARCH($p,q$) recurrence equation is of the form of
+\eqref{eq:rhjyu}. With appropriate conditions, $\sigma_t^2$ is
+shown to be a positive Harris recurrent Markov chain (cf. Bollerslev
+\cite{bollerslev:1986} and Buraczewski et al.
 \cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
-has regularly varying tails. The tail index, call it $\xi$, is given by
-\[
-\lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\xi = 0
-\]
-where $A_1, A_2, \dots, A_n$ are iid matrices whose entries are
-functions of $\{\alpha_i\}_{i=1}^p$ and $\{\beta_i\}_{i=1}^q$:
+has regularly varying tails. The tail index $\alpha$, is given by
+\begin{equation}
+  \label{eq:grgbg}
+  \lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\alpha = 0,
+\end{equation}
+where $\{A_i\}_{i \in \mathbb Z}$ are iid matrices whose entries are
+functions of $\{\alpha_i\}_{i=1}^p$, $\{\beta_i\}_{i=1}^q$ and
+$\{Z_t^2\}$:
 \[
@@
 \]
+While GARCH models have been very successful for modelling financial
+time series, it does have its drawbacks. For example, the tail
+index is very sensitive to the model parameters $\{\alpha_i\}_{i=1}^p$ and
+$\{\beta_i\}_{i=1}^q$. In applications, these parameters need to be
+estimated from a sample and are all ways uncertain to an extent, there
+can be a significant discrepancy between the tail index implied by
+\eqref{eq:grgbg} and the Hill estimate \eqref{eq:fhhty}.
+
+There exist various extensions of the univariate GARCH
+model to the multivariate case. The most notable one is perhaps the
+{\em constant conditional correlation} (CCC) model of Bollerslev
+\cite{bollerslev:1990} and Jeantheau \cite{jeantheau:1998}. In the
+bivariate case, CCC is the model 
+\beao\bfX_t=
+\left(\barr{l}X_{1,t}\\
+X_{2,t}\earr\right)= \left(\barr{ll}\sigma_{1,t}& 0\\
+0&\sigma_{2,t}\earr\right)\,\left(\barr{l}Z_{1,t}\\Z_{2,t}\earr\right)=\Sigma_t\,\bfZ_t\,,\qquad
+t\in\bbz\,.
+\eeao
+Thus both return components $X_{i,t}$ have the form of a univariate
+stochastic volatility model $X_{i,t}=\sigma_{i,t}Z_{i,t}$ 
+with non-negative volatility $\sigma_{i,t}$ and an iid bivariate noise
+\seq\ $(\bfZ_t)$ with zero mean and unit variance components.
+We also have the specification
+\begin{small}
+\beam\label{eq:8:mikosch}
+\bfY_t=\left(\barr{l}\sigma^2_{1,t}  \\  
+\sigma^2_{2,t}\earr
+\right)
+&=& \left(
+\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)
++\left(\barr{cc}\alpha_{11} & \alpha_{12}  \\
+      \alpha_{21} & \alpha_{22}\earr \right)\, 
+\left(\barr{l}X_{1,t-1}^2  \\X_{2,t-1}^2   \earr\right)
+ + \left(\barr{cc}\beta_{11} & \beta_{12}  \\\beta_{21} & \beta_{22} \earr
+ \right)\,\left(\barr{c}\sigma^2_{1,t-1}  \\\sigma^2_{2,t-1}\earr
+  \right)\nonumber\\
+&=& \left(
+\barr{l}\alpha_{01}  \\\alpha_{02}
+\earr\right)+\left(\barr{cc}\alpha_{11}Z_{1,t-1}^2+\beta_{11}&\alpha_{12}Z_{2,t-1}^2+
+\beta_{12}\\
+\alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
+\earr\right)\,\left(\barr{l}\sigma_{1,t-1}^2\\\sigma_{2,t-1}^2\earr
+\right)\,,
+\eeam
+for positive $\alpha_{0i}$ and suitable non-negative 
+$\alpha_{ij},\beta_{ij}$, $i,j=1,2$.
+Writing
+\beao
+\bfB_t= \left(
+\barr{l}\alpha_{01}  \\\alpha_{02}   \earr\right)\quad\mbox{and}\quad
+\bfA_t=\left(\barr{cc}\alpha_{11}Z_{1,t-1}^2+\beta_{11}&\alpha_{12}Z_{2,t-1}^2+
+\beta_{12}\\
+\alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
+\earr\right)\,,
+\eeao
+\end{small}
+we see that we are again in the framework of a stochastic recurrence equation\ 
+but this time for vector-valued $\bfB_t$ and matrix-valued $\bfA_t$:
+\beam\label{eq:jan6b:mikosch}
+\bfY_t=\bfA_t\,\bfY_{t-1}+\bfB_t\,,\qquad t\in\bbz\,.
+\eeam
+Kesten \cite{kesten:1973} also provided the corresponding theory  
+for stationarity and tails in this case. \sta\ \cite{starica:1999}
+dealt with the corresponding problems for CCC-GARCH processes,
+making use of the theory in Kesten \cite{kesten:1973},
+Bougerol and Picard \cite{bougerol:picard:1992}
+and its
+specification to the tails of GARCH models 
+in Basrak et al.~\cite{basrak:davis:mikosch:2002}. \sta\ \cite{starica:1999} assumed the 
+Kesten conditions for the matrices $\bfA_t$. These conditions ensure
+that the product matrices $\bfA_1\cdots\bfA_n$ have positive entries
+for sufficiently large $n$. Then Kesten's theory implies that all
+components of the vector $\bfX_t$ have power-law tails with the same
+index $\alpha$ and also that the \fidi s of the process $(\bfX_t)$ are
+\regvary\ with index $\alpha$.
+\par
+Various GARCH modifications are derived by considering linear
+combinations of CCC-GARCH models. The property of multivariate
+\regvar\ of multivariate GARCH ensures that, after linear
+transformations,  the new process in all components has again
+power-law tails with the same index as the original GARCH process; see
+Basrak et al.~\cite{basrak:davis:mikosch:2002}.
+Models which are constructed in this way are
+the Orthogonal GARCH model of
+Alexander and Chibumba \cite{alexander:chibumba:1996}, its
+generalization GO-GARCH by van der Weide \cite{Weide2002},  the Full
+Factor GARCH model of Vrontos et al. \cite{vrontos2003full} and the
+Generalized Orthogonal Factor GARCH model of Lanne and Saikkonen
+\cite{lanne2007modelling}. These models are characterized by their
+treatment of each series as a linear combination of factors, and each
+of the factors is modeled as a GARCH process; see Silvennoinen and
+Ter\"asvirt\"a \cite{silventeras:2009}.
+\par
+Not all choices of $\alpha$- and $\beta$-parameters in the model
+\eqref{eq:8:mikosch} allow for an application of the Kesten theory. For
+example, assume that only the diagonal elements $\alpha_{ii}$ and
+$\beta_{ii}$ are positive. Then $\bfA_t$ is diagonal and, hence, the
+condition that $\bfA_1\cdots\bfA_n$ have positive entries for
+sufficiently large $n$  cannot be satisfied. In the latter situation,
+both $(X_{1,t})$ and $(X_{2,t})$ are univariate GARCH
+processes. Assuming the  conditions of the univariate Kesten-Goldie
+theorem for each component process, $(X_{1,t})$ and $(X_{2,t})$ have
+power-law tails  with indices $\alpha_1$ and $\alpha_2$, respectively,
+given by the solutions to the equations  $\E [(\alpha_{ii}
+Z_{i,t}^2+\beta_{ii})^{\alpha_i/2}]=1$, $i=1,2$. In this model, one
+can introduce dependence between the two component series $(X_{1,t})$
+and $(X_{2,t})$ by assuming dependence between the noise variables
+$Z_{1,t}$ and $Z_{2,t}$. Another situation when the Kesten theory fails 
+appears when $\bfA_t$ is an upper or lower triangle matrix: then the
+products  $\bfA_1\cdots\bfA_n$ are always of the same triangular
+type. 
+Similar remarks apply when one considers a CCC model in general
+dimension. Of course, one may argue that the latter models 
+are not natural: they are degenerate since they do not allow 
+for a linear relationship between all squared volatilities on a given
+day.
+
+\section{Stochastic Volatility models}
 

### 2017-08-09 phd/Chapterintroduction.tex
Added texts to the introduction section of the phd thesis.

@@
 Upon these conditions, Kesten's theorem gives
-\begin{equation*}
+\begin{equation}
+  \label{eq:jorg}
   u^\alpha \P(u^{-1} V \in \cdot) \overset{v}{\to} \nu_\alpha(\cdot)
-\end{equation*}
+\end{equation}
 where $\nu_\alpha$ is a non-null Radon measure on
@@
 $\nu_\alpha(a A) = a^{-\alpha} \nu_\alpha(A)$.
+
+In addition to non-negative matrices, two other classes of random
+matrices have been shown to lead to power-law tails via the recurrence
+relation \eqref{eq:garchpq_sre}. Alsmeyer and Mentemeier
+\cite{alsmeyer:mentemeier:2012} considered invertible matrices whose
+distribution has a density. Let $M(d, \reals)$ denote the space of
+$d \times d$ matrices with real entries that are invertible with
+probability 1. They replaced Kesten's condition
+$\E (\|A\|^\xi \log^+\|A\|) < \infty$ with a stronger
+counterpart
+$\E [\|A\|^\xi (\log^+\|A\| + \log \|A^{-1} \|)] < \infty$,
+and lifted the condition (A). In addition, they assumed
+\begin{enumerate}
+  \item The Markov chain $X_n$ on $\sphere^{d-1}$, namely
+    $X_n = A_n X_{n-1} / |A_n X_{n-1}|$, is irreducible, i.e. for any open
+    set $U \subset \sphere^{d-1}$ and any $u \in \sphere^{d-1}$, 
+    $\exists n \geq 1$ such that $\P(X_n u \in U) > 0$.
+  \item There exist $N \geq 1$, $c, \epsilon > 0$ and an invertible
+    matrix $\bar A \in M(d, \reals)$ such that for any set
+    $C \subset M(d, \reals)$, it holds true
+    $\P(A_N \cdots A_1 \in C) \geq c |B_\epsilon(\bar A) \cap C|$,
+    where $|\cdot|$ denotes the Lebesgue measure.
+\end{enumerate}
+These assumptions are termed conditions (id). Furthermore, they assumed
+that there was no point in $\reals^d$ such that the recurrence
+equation \eqref{eq:garchpq_sre} was stuck at this point with probability 1: 
+$\P(A X + B = X) < 1$ for all $X \in \reals^d$ and all $A \in M(d, \reals)$.
+With these assumptions, they showed
+\[
+\lim_{u \to \infty} u^\xi \P(\inn{x, V} > u) = e_\xi(x)
+\]
+where $x \in \sphere^{d - 1}$ and $e_\xi(\cdot)$ is a continuous
+function $\sphere^{d-1} \to \reals_+$.
+
+The second of the (id) conditions, which is satisfied when the
+distribution of $A$ has a Lebesgue density, can actually be lifted if
+stronger moment conditions are imposed on $A$ and $B$, and in
+addition, a proximity condition is satisfied by the support of
+$A$. This is the result of Guivarc'h and Le Page, et al.
+\cite{guivarc:page:2016}. Let $G_A$ denote the semi-group generated
+by $\{\Pi_n: \Pi_n = A_n \cdots A_1, A_i \in M(d, \reals)\}$. The
+authors assumed
+\begin{enumerate}
+  \item There is no finite union $W$ of proper subspaces of $\reals^d$
+    that satisfies $\forall a \in G_A, a W = W$.
+  \item $G_A$ contains a proximal element, i.e. an element $a$ whose
+    largest singular value is an algebraically simple eigenvalue of $a$.
+\end{enumerate}
+These two assumptions are termed (ip) conditions. Replacing the (id)
+conditions of Alsmeyer and Mentemeier with (ip) and the moment
+conditions of the former with
+\[
+\E \|A\|^{\xi + \delta} < \infty, \quad
+\E (\|A\|^{\xi} \|A^{-1}\|^{\delta}) < \infty, \quad
+\E (|B|^{\xi + \delta} < \infty),\quad
+\text{ for some }\delta > 0
+\]
+Guivarc'h and Le Page et al. proved the same vague convergence result
+of \eqref{eq:jorg}.
+
+In the special case of GARCH($p, q$) processes, which are discussed in
+more details in the next section, Bollerslev \cite{bollerslev:1986}
+showed that the fixed-point equation $X \overset{d}{=} A X + B$ has a
+unique, strictly stationary solution $X$ with finite variance if and
+only if
+\begin{equation}
+  \sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1
+  \label{eq:bollerslev:intro}
+\end{equation}
 
@@
 $\{\beta_i\}_{i=1}^q$. In applications, these parameters need to be
-estimated from a sample and are all ways uncertain to an extent, there
-can be a significant discrepancy between the tail index implied by
-\eqref{eq:grgbg} and the Hill estimate \eqref{eq:fhhty}.
+estimated from a sample and are all ways uncertain to some extent. For
+this reason, there can be a significant discrepancy between the tail
+index obtained via \eqref{eq:grgbg} with the model parameters
+substituted for their sample estimates and the Hill estimate
+\eqref{eq:fhhty}.
 
@@
 \end{small}
-we see that we are again in the framework of a stochastic recurrence equation\ 
-but this time for vector-valued $\bfB_t$ and matrix-valued $\bfA_t$:
+we see that we are again in the framework of a stochastic recurrence
+equation but this time for vector-valued $\bfB_t$ and matrix-valued
+$\bfA_t$:
 \beam\label{eq:jan6b:mikosch}
@@
 \section{Stochastic Volatility models}
+With the availability of high-frequency data, a different approach
+than that of GARCH has been popularized for modelling volatility of
+financial time series and has led to greatly improved accuracy of
+prediction. This is the approach of stochastic volatility models.
+
+In the pioneering work of Clark \cite{clark:1973}, the author modelled
+the logarithmic price $X_t$ as a subordinated stochastic process:
+$X_t = V_{\tau_t}$, where $V_i$ is a normally distributed random
+variable with mean 0, variance $\sigma^2 i$ and independent
+increments. $\tau_t$ is a real-valued, non-negative, non-decreasing
+sequence with $\tau_0 = 0$. It models a time change. As pointed out by
+Shepard and Andersen \cite{Shephard:Andersen:2009}, the log-price
+process $X_t$ is serially uncorrelated although potentially dependent,
+provided that $V_t$ and $\tau_t$ are independent.
+
+Later authors, e.g. Back \cite{back:1991} chose to model the log-price
+process as a semimartingale, with increments of the Martingale
+component modelled as a product process:
+\begin{eqnarray*}
+  Y_t &=& Y_0 + A_t + M_t \\
+  M_t - M_{t-1} &=& X_t = \sigma_t Z_t
+\end{eqnarray*}
+where $M_t$ is a semimartingale, $A_t$ is a finite variation process
+and $M_t$ is a Martingale. $\sigma_t$ is non-negative and $Z_t$ is an
+iid process with 0 mean and unit variance. A convenient choice of
+$\sigma_t$ is
+\[
+\log\sigma_t = \sum_{l \in \mathbb Z} \psi_l \eta_{t-l}
+\]
+where $\psi_l$ is a sequence of non-negative numbers with at least one
+positive element and $0 \leq \psi_l \leq 1$.
+
+Clearly, if either
 

### 2017-08-09 phd/Chapterintroduction.tex
Revised the introduction part of the phd thesis.

@@
 We call a random variable $X$ regularly varying with index
-$\alpha \geq 0$ if it satisfies the tail balance condition
+$\alpha \geq 0$ if it satisfies the tail balance condition in the
+limit $x \to \infty$.
 \[
@@
 {
-  \P(V/|V| \in \cdot, |V| > c x)
+  \P(X/|X| \in \cdot, |X| > c x)
   \over
-  \P(|V| > x)
+  \P(|X| > x)
 }
-\overset{v}{\to} c^{-\alpha} \mu_\alpha(\cdot)
+\overset{v}{\to} c^{-\alpha} \mu_\alpha(\cdot),
 \quad
@@
 \]
-then we say $V$ is regularly varying with index $\alpha$. Here
+then we say $X$ is regularly varying with index $\alpha$. Here
 $\mu_\alpha$ is a probability measure on the unit sphere
-\cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
-is regularly varying with index $\alpha$, then each component
+\cite{buraczewski:damek:mikosch:2016}. It is called the spectral 
+measure of $X$ and $\alpha$ is again the tail index.
+If $X$ is regularly varying with index $\alpha$, then each component
 and each linear combination of its components are regularly
@@
     \right)
-  \right]^{-1}
+  \right]^{-1},
 \end{equation}
@@
 of daily stock return series from 3 sectors of the
-{\em Standard \& Poor 500} index \footnote{An American stock index
-  comprising 500 or so companies.}. The 2.5\% and 97.5\% quantiles of
+{\em Standard \& Poor's 500} index \footnote{An American stock index
+  comprising around 500 companies.}. The 2.5\% and 97.5\% quantiles of
 the asymptotic normal distribution of the estimates are also
@@
 similar they really are and if/how their variations can be explained by
-economical arguments. We explore this topic in chapter
+economic arguments. We explore this topic in chapter
 \ref{ch:TailParameters}.
@@
 with index $\alpha_j$. Then by the summation property, each and every
-$X_i$ is regularly varying with index $\min_{1 \leq j \leq K} \alpha_j$.
+$\{X_i\}$ is regularly varying with index $\min_{1 \leq j \leq K} \alpha_j$.
 In practice a factor $Y_{i,t}$ is estimated as
@@
 
-When a product of positive random variables, say $X_1 X_2$, involves
-one with regularly varying tails, a useful result is that of
-Breiman \cite{breiman:1965}: assume $X_1$ is regularly varying with
-index $\alpha$ and there exists $\epsilon > 0$ such that
+When a product of independent positive random variables, say $X_1
+X_2$, involves one with regularly varying tails, a useful result is
+that of Breiman \cite{breiman:1965}: assume $X_1$ is regularly varying
+with index $\alpha$ and there exists $\epsilon > 0$ such that
 $\E X_2^{\alpha + \epsilon} < \infty$.
@@
 probability 1. They replaced Kesten's condition
-$\E (\|A\|^\xi \log^+\|A\|) < \infty$ with a stronger
+$\E (\|A\|^\alpha \log^+\|A\|) < \infty$ with a stronger
 counterpart
-$\E [\|A\|^\xi (\log^+\|A\| + \log \|A^{-1} \|)] < \infty$,
+$\E [\|A\|^\alpha (\log^+\|A\| + \log \|A^{-1} \|)] < \infty$,
 and lifted the condition (A). In addition, they assumed
@@
 equation \eqref{eq:garchpq_sre} was stuck at this point with probability 1: 
-$\P(A X + B = X) < 1$ for all $X \in \reals^d$ and all $A \in M(d, \reals)$.
-With these assumptions, they showed
+$\P(A X + B = X) < 1$ for all $X \in \reals^d$. With these
+assumptions, they showed
 \[
-\lim_{u \to \infty} u^\xi \P(\inn{x, V} > u) = e_\xi(x)
+\lim_{u \to \infty} u^\alpha \P(\inn{x, V} > u) = e_\alpha(x)
 \]
-where $x \in \sphere^{d - 1}$ and $e_\xi(\cdot)$ is a continuous
-function $\sphere^{d-1} \to \reals_+$.
+where $x \in \sphere^{d - 1}$ and $e_\alpha(\cdot)$ is a continuous
+function on $\sphere^{d-1}$.
 
@@
 addition, a proximity condition is satisfied by the support of
-$A$. This is the result of Guivarc'h and Le Page, et al.
+$A$. This is the result of Guivarc'h and Le Page.
 \cite{guivarc:page:2016}. Let $G_A$ denote the semi-group generated
@@
 \[
-\E \|A\|^{\xi + \delta} < \infty, \quad
-\E (\|A\|^{\xi} \|A^{-1}\|^{\delta}) < \infty, \quad
-\E (|B|^{\xi + \delta} < \infty),\quad
+\E \|A\|^{\alpha + \delta} < \infty, \quad
+\E (\|A\|^{\alpha} \|A^{-1}\|^{\delta}) < \infty, \quad
+\E (|B|^{\alpha + \delta} < \infty),\quad
 \text{ for some }\delta > 0
 \]
-Guivarc'h and Le Page et al. proved the same vague convergence result
-of \eqref{eq:jorg}.
+Guivarc'h and Le Page proved the vague convergence result of \eqref{eq:jorg}.
 
-In the special case of GARCH($p, q$) processes, which are discussed in
-more details in the next section, Bollerslev \cite{bollerslev:1986}
-showed that the fixed-point equation $X \overset{d}{=} A X + B$ has a
-unique, strictly stationary solution $X$ with finite variance if and
-only if
+Bollerslev \cite{bollerslev:1986} showed that the fixed-point equation
+$Y \overset{d}{=} A Y + B$ has a unique, strictly stationary solution
+$Y$ with finite variance if and only if
 \begin{equation}
@@
 where $\{X_t\}$ is a return series, e.g. stock returns, foreign exchange
-rates, interest rates, etc; $\{Z_t\}$ is a iid, mean 0, unit-variance
-sequence; and $\sigma_t^2$ is the variance of the
-distribution of $X_t$ conditional on $\{(X_i,
-\sigma_i^2)\}_{i=0}^{t-1}$; $\omega, \{\alpha_i\}_{i=1}^p,
-\{\beta_i\}_{i=1}^q$ are non-negative parameters of the
-model. The GARCH($p,q$) recurrence equation is of the form of
-\eqref{eq:rhjyu}. With appropriate conditions, $\sigma_t^2$ is
-shown to be a positive Harris recurrent Markov chain (cf. Bollerslev
-\cite{bollerslev:1986} and Buraczewski et al.
+rates, interest rates, etc; $\{Z_t\}$ is an iid, mean 0, unit-variance
+sequence, $\sigma_t^2$ is the variance of the distribution of $X_t$
+conditional on
+$\{(X_i, \sigma_i^2)\}_{i=0}^{t-1}$; $\omega, \{\alpha_i\}_{i=1}^p,
+\{\beta_i\}_{i=1}^q$
+are non-negative parameters of the model. The GARCH($p,q$) recurrence
+equation is of the form of \eqref{eq:rhjyu}. With appropriate
+conditions, $\sigma_t^2$ is shown to be a positive Harris recurrent
+Markov chain (cf. Bollerslev \cite{bollerslev:1986} and Buraczewski et al.
 \cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
@@
 $\{\beta_i\}_{i=1}^q$. In applications, these parameters need to be
-estimated from a sample and are all ways uncertain to some extent. For
+estimated from a sample and are always uncertain to some extent. For
 this reason, there can be a significant discrepancy between the tail
@@
 In the pioneering work of Clark \cite{clark:1973}, the author modelled
-the logarithmic price $X_t$ as a subordinated stochastic process:
-$X_t = V_{\tau_t}$, where $V_i$ is a normally distributed random
+the logarithmic price $Y_t$ as a subordinated stochastic process:
+$Y_t = V_{\tau_t}$, where $V_i$ is a normally distributed random
 variable with mean 0, variance $\sigma^2 i$ and independent
@@
 Shepard and Andersen \cite{Shephard:Andersen:2009}, the log-price
-process $X_t$ is serially uncorrelated although potentially dependent,
+process $Y_t$ is serially uncorrelated although potentially dependent,
 provided that $V_t$ and $\tau_t$ are independent.
@@
 \end{eqnarray*}
-where $M_t$ is a semimartingale, $A_t$ is a finite variation process
-and $M_t$ is a Martingale. $\sigma_t$ is non-negative and $Z_t$ is an
-iid process with 0 mean and unit variance. A convenient choice of
-$\sigma_t$ is
+where $A_t$ is a finite variation process and $M_t$ is a Martingale
+and hence $Y_t$ is a semimartingale . $\sigma_t$ is non-negative and
+$Z_t$ is an iid process with 0 mean and unit variance. A convenient
+choice of $\sigma_t$ is
 \[
@@
 positive element and $0 \leq \psi_l \leq 1$.
-
-Clearly, if either
 

### 2017-08-10 phd/Chapterintroduction.tex
Revised the introduction section.

@@
 Later authors, e.g. Back \cite{back:1991} chose to model the log-price
-process as a semimartingale, with increments of the Martingale
+process as a semimartingale, with increments of the martingale
 component modelled as a product process:
@@
 \end{eqnarray*}
-where $A_t$ is a finite variation process and $M_t$ is a Martingale
-and hence $Y_t$ is a semimartingale . $\sigma_t$ is non-negative and
+where $A_t$ is a finite variation process and $M_t$ is a martingale
+and hence $Y_t$ is a semimartingale. $\sigma_t$ is non-negative and
 $Z_t$ is an iid process with 0 mean and unit variance. A convenient
@@
 \]
-where $\psi_l$ is a sequence of non-negative numbers with at least one
-positive element and $0 \leq \psi_l \leq 1$.
+where $-1 \leq \{\psi_l\}_{l \in \mathbb Z} \leq 1$
+is a sequence of real numbers with at least one non-zero element.
 

### 2017-08-10 phd/Chapterintroduction.tex
Revised the introduction chapter. reimplemented the sampling algorithm using resampling.

@@
 
-\section{Regular Variation}
+\section{Regular Variation}\label{sec:intro_regvar}
 The concept of regular variation is defined by the following scaling
@@
 
+In addition to Breiman's result, the following is also well-known,
+assuming $X_1, X_2$ are positive independent random variables and
+$X_1$ is regularly varying with index $\alpha$:
+\begin{enumerate}
+\item if $X_2$ is regularly varying with index $\alpha$ or
+  $\lim_{x \to \infty} {P(X_2 > x) \over \P(X_1 > x)} = 0$, then $X_1
+  X_2$ is regularly varying with index $\alpha$.
+\item if $X_1, X_2$ are iid and $\E X_1^\alpha = \infty$, then
+  ${\P(X_1 X_2 > x) \over \P(X_1 > x)} \to \infty$
+\item if $X_1, X_2$ are iid and $\E X_1^\alpha < \infty$, then the
+  only possible limit of ${\P(X_1 X_2 > x) \over \P(X_1 > x)}$ is
+  $2 \E X_1^\alpha$.
+\end{enumerate}
 For an extensive summary of the regular variation properties of
@@
 choice of $\sigma_t$ is
-\[
+\begin{equation}
+  \label{eq:rfht}
   \log\sigma_t = \sum_{l \in \mathbb Z} \psi_l \eta_{t-l}  
-\]
+\end{equation}
 where $-1 \leq \{\psi_l\}_{l \in \mathbb Z} \leq 1$
-is a sequence of real numbers with at least one non-zero element.
+is a sequence of real numbers with at least one non-zero element,
+$\{\eta_t\}$ is a iid sequence of random variables. By Kolmogorov's
+3-series theorem, the infinite series above converges if and only if
+$\sum_{l \in \mathbb Z} \psi_l^2 < \infty$.
+In particular, if $\{\eta_i\}$ is normally distributed with mean 0,
+$\{\sigma_t\}$ is stationary and its marginal distributions are
+log-normal.
+
+If, however, the sequence $\{Z_t\}$ or $\{\sigma_t\}$ is regularly
+varying with index $\alpha$ and some additional conditions are
+satisfied, $X_t$ is also regularly varying with the same
+index. Specifically, if $Z_t$ is regularly varying and $\sigma_t$ has
+a lighter tail, the conclusion follows from Breiman's lemma. See
+\S\ref{sec:case1} of Jan\ss en et al. i.e. chapter \ref{ch:bernoulli}
+of this thesis for more details.
+
+If, instead, $\sigma_t$ has a heavier tail than $Z_t$, $X_{i,t}$ is
+regularly varying assuming the following conditions:
+\begin{itemize}
+\item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
+  is a generic element of the iid sequence $\{\eta_t\}$, and
+  \begin{equation}
+    \label{eq:dfhty}
+    \E e^{\alpha \eta} = \infty, \quad\text{ or }
+    {\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)} \to c \in (0, \infty)
+  \end{equation}
+    where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
+    iid sequence $\{\eta_t\}$.
+\end{itemize}
+Note the only finite limit for
+${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ in $(0, \infty)$ is
+$\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
+are mutually exclusive. In chapter \ref{ch:bernoulli} we will see that
+these two conditions lead to rather different behavior for
+eigenvalues of the sample covariance matrix of stochastic volatility
+processes of the form \eqref{eq:rfht}.
+
+There are a few advantages in using stochastic volatility models. For
+one thing, they are among the simplest models allowing for conditional
+heteroscedasticity; for another thing,
 

### 2017-08-11 papers/FX/PaperBody.tex
Revised my phd thesis.

--- 2017-08-02 papers/FX/PaperBody.tex
@@
 \begin{abstract}
-%We consider an investor with preferences that accord with Generalized
-%Disappointment Aversion. Such an investor cares about downside risk and
-%we assume he recognizes the heavy tail feature of asset return
-%distributions. We argue that when a market is dominated by rational
-%investors of this kind, the return distributions of equities that are
-%actively traded in this market must have very similar tail-indices. We
-%also show, in contrast, the scale parameters of the return
+We consider an investor with preferences that accord with Generalized
+Disappointment Aversion (GDA). Such an investor cares about downside
+risk and we assume he recognizes the heavy tail feature of asset return
+distributions. We argue that when a market is dominated by rational
+investors of this kind, the return distributions of equities that are
+actively traded in this market may have nearly equal tail-indices due
+to monotonicity of the GDA preference with respect to the tail index.
+
+We estimate tail indices and scale parameters of S\&P 500 stocks and
+test the hypothesis that two given stock return series have the same
+tail index.
+% We also show, in contrast, the scale parameters of the return
 %distributions may differ hugely from one another.

### 2017-08-11 phd/Chapterintroduction.tex
Revised my phd thesis.

@@
 \cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
-has regularly varying tails. The tail index $\alpha$, is given by
+has regularly varying tails. The tail index $\alpha$ is given by
 \begin{equation}
@@
 component modelled as a product process:
-\begin{eqnarray*}
-  Y_t &=& Y_0 + A_t + M_t \\
+\begin{eqnarray}
+  Y_t &=& Y_0 + A_t + M_t \nonumber \\
   M_t - M_{t-1} &=& X_t = \sigma_t Z_t
-\end{eqnarray*}
+  \label{eq:rtght}
+\end{eqnarray}
 where $A_t$ is a finite variation process and $M_t$ is a martingale
@@
 is a sequence of real numbers with at least one non-zero element,
-$\{\eta_t\}$ is a iid sequence of random variables. By Kolmogorov's
-3-series theorem, the infinite series above converges if and only if
-$\sum_{l \in \mathbb Z} \psi_l^2 < \infty$.
-In particular, if $\{\eta_i\}$ is normally distributed with mean 0,
-$\{\sigma_t\}$ is stationary and its marginal distributions are
-log-normal.
+$\{\eta_t\}_{t \in \mathbb Z}$ is a iid sequence of random
+variables. By Kolmogorov's 3-series theorem, the infinite series above
+converges if and only if $\sum_{l \in \mathbb Z} \psi_l^2 < \infty$.
+In particular, if $\{\eta_i\}_{i \in \mathbb Z}$ is normally
+distributed with 0 mean, $\{\sigma_t\}_{t \in \mathbb Z}$ is
+stationary and its marginal distributions are log-normal.
 
-If, however, the sequence $\{Z_t\}$ or $\{\sigma_t\}$ is regularly
-varying with index $\alpha$ and some additional conditions are
-satisfied, $X_t$ is also regularly varying with the same
-index. Specifically, if $Z_t$ is regularly varying and $\sigma_t$ has
-a lighter tail, the conclusion follows from Breiman's lemma. See
-\S\ref{sec:case1} of Jan\ss en et al. i.e. chapter \ref{ch:bernoulli}
-of this thesis for more details.
+If, however, the sequence $\{Z_t\}_{t \in \mathbb Z}$ or
+$\{\sigma_t\}_{t \in \mathbb Z}$ is regularly varying with index
+$\alpha$ and some additional conditions are satisfied, $X_t$ is also
+regularly varying with the same index. Specifically, if $Z_t$ is
+regularly varying and $\sigma_t$ has a lighter tail, the conclusion
+follows from Breiman's lemma. See \S\ref{sec:case1} of Jan\ss en et
+al. i.e. chapter \ref{ch:bernoulli} of this thesis for more details.
 
-If, instead, $\sigma_t$ has a heavier tail than $Z_t$, $X_{i,t}$ is
-regularly varying assuming the following conditions:
+If, instead, $\sigma_t$ has a heavier tail than $Z_t$, then $X_{i,t}$
+is regularly varying assuming the following conditions:
 \begin{itemize}
 \item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
-  is a generic element of the iid sequence $\{\eta_t\}$, and
+  is a generic element of the iid sequence
+  $\{\eta_t\}_{t \in \mathbb Z}$, and
   \begin{equation}
@@
     where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
-    iid sequence $\{\eta_t\}$.
+    iid sequence $\{\eta_t\}_{t \in \mathbb Z}$.
 \end{itemize}
 Note the only finite limit for
-${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ in $(0, \infty)$ is
+${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ is
 $\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
@@
 one thing, they are among the simplest models allowing for conditional
-heteroscedasticity; for another thing,
+heteroscedasticity (cf. Andersen et
+al. \cite{andersen:davis:kreiss:mikosch:2009}); for another thing,
+they greatly improve the accuracy of predicting future volatilities
+by taking advantage of high frequency data. Specifically, for the
+semimartingale $Y_t$ we have
+\begin{equation}
+  \label{eq:jieg}
+  [Y_{\delta}]_t
+  =
+  \sum_{j=1}^{\floor{t/\delta}}
+  (Y_{j \delta} - Y_{(j-1)\delta})^2
+  \overset{P}{\to} [Y]_t,
+  \quad
+  \delta \to 0.
+\end{equation}
+The object $[Y_\delta]_t$ is called the realized quadratic variation
+in time series literature.
+The convergence in probability follows directly from the definition of
+quadratic variation.
+%% When $\E Z_t = 0$, $[Y]_t = [M]_t$ 
+Furthermore, Jacod \cite{jacod:1994} and
+Barndorff-Nielsen and Shephard \cite{barndorff:shephard:2002} proved a
+central limit theorem:
+\[
+  {[Y_\delta]_t - [Y]_t \over \sqrt{2 \delta \int_0^t \sigma_s^2 ds}}
+  \overset{d}{\to} N(0, 1),
+  \quad \delta \to 0
+\]
+Moreover, the structure $X_t = M_{t} - M_{t-1} = \sigma_t Z_t$ gives
+$[Y_\delta]_t \overset{P}{\to} \int_0^t \sigma_s^2 ds$. Meanwhile,
+It\^o's formula gives, for semimartingale $Y_t$,
+\begin{eqnarray}
+  Y_t^2 &=& [Y]_t + 2 \int_0^t Y_s d Y_s \nonumber \\
+  &=& [Y]_t + 2 \int_0^t Y_s dA_s + 2 \int_0^t Y_s dM_s \nonumber \\
+  \E Y_t^2 &=& \E [Y]_t + 2 \E \left( \int_0^t Y_s dA_s \right) \nonumber\\
+  &\approx& \E [Y]_t, \quad \text{when $t$ is small} \label{eq:hytre}
+\end{eqnarray}
+\eqref{eq:hytre} and \eqref{eq:jieg} shows $\E Y_t^2$, or in other
+words, the forecast of future squared return can be obtained as
+$\E [Y_\delta]_t$, i.e. the forecast of future realized quadratic
+variation. Since past realized quadratic variations are observable,
+forcast of future $[Y_\delta]_t$ can be obtained using e.g. an ARFIMA
+model.
 

### 2017-08-11 papers/FX/PaperBody.tex
Added an abstract to the FX paper.

@@
 to monotonicity of the GDA preference with respect to the tail index.
-
-We estimate tail indices and scale parameters of S\&P 500 stocks and
+We give conditions upon which the GDA preference is monotone and hence
+suggests an equal tail index for all actively traded stocks.
+
+We also estimate tail indices and scale parameters of S\&P 500 stocks and
 test the hypothesis that two given stock return series have the same
-tail index.
+tail index. The results vary across different sectors of the index.
+
 % We also show, in contrast, the scale parameters of the return
@@
 returns of speculative prices are {\em heavy-tailed.} There is no agreement
-in the literature about how heavy these tails really are. For example, Barndorff-Nielsen and Shephard
-\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails
+in the literature about how heavy these tails really are. For example,
+Barndorff-Nielsen and Shephard \cite{barndorff:shephard:2001} and
+Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails
 which are comparable with those of a gamma \ds . On the other
@@
 \beam\label{eq:1}
-\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad \P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
+\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad
+\P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
 \eeam  
-where $c_{\pm}$, $\alpha_{\rm up}$ and $\alpha_{\rm low}$ are positive constants.\footnote{Here and in what
-  follows, $f(x)\sim g(x)$ for positive \fct s $f$ and $g$ means 
-that $f(x)/g(x)\to 1$ as $\xto$.} See for example, Embrechts et
+where $c_{\pm}$, $\alpha_{\rm up}$ and $\alpha_{\rm low}$ are positive
+constants.\footnote{Here and in what follows, $f(x)\sim g(x)$ for
+  positive \fct s $f$ and $g$ means that $f(x)/g(x)\to 1$ as $\xto$.}
+See for example, Embrechts et
 al. \cite{embrechts:klueppelberg:mikosch:1997}, Jansen and de Vries
 \cite{jansen1991frequency}, Mikosch \cite{mikosch:2003}, Resnick 
-\cite{resnick:2007}.  
-In the extreme value literature it is common to replace the constants $c_\pm$ by suitable
-{\em slowly varying} \fct s; cf. Embrechts et
-al. \cite{embrechts:klueppelberg:mikosch:1997}, Chapter 3. In this
-paper, for the  
-sake of argument, we stick to the condition \eqref{eq:1}.
-\par
-There are some good theoretical reasons for the appearance of power-law tails in situations where certain moments 
+\cite{resnick:2007}. In the extreme value literature it is common to
+replace the constants $c_\pm$ by suitable {\em slowly varying} \fct s;
+cf. Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997},
+Chapter 3. In this paper, for the sake of argument, we stick to the condition \eqref{eq:1}.
+\par
+There are some good theoretical reasons for the appearance of
+power-law tails in situations where certain moments 
 of data are believed to be infinite. Tails of type \eqref{eq:1}
 describe the maximum domain of attraction of the  
-Fr\'echet \ds\ $\Phi_{\alpha_{\rm up}}(x)=\exp(-x^{-\alpha_{\rm up}})$ for $x>0$, i.e., scaled maxima of an iid \seq\ $(X_t)$ with
+Fr\'echet \ds\ $\Phi_{\alpha_{\rm up}}(x)=\exp(-x^{-\alpha_{\rm up}})$
+for $x>0$, i.e., scaled maxima of an iid \seq\ $(X_t)$ with
 upper tail described in \eqref{eq:1} converge in \ds\ to $\Phi_{\alpha_{\rm up}}$.
 Equivalently, power-law tails are prescribed by the 
-generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e., for a suitable positive scaling \fct\ $a(u)$,
+generalized Pareto \ds\ which is the limit \ds\ of the excesses of
+$X_t$ above high thresholds, i.e., for a suitable positive scaling
+\fct\ $a(u)$,
 \beao
-\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha_{\rm up})^{-\alpha_{\rm up}}\,,\qquad u\to\infty\,.
+\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha_{\rm up})^{-\alpha_{\rm
+    up}}\,,\qquad u\to\infty\,.
 \eeao

### 2017-08-11 phd/Chapterintroduction.tex
Added an abstract to the FX paper.

@@
 potential topics of future work.
+
+\subsection{Importance sampling estimator of GARCH(p,q) rare event probability}
+  We propose an efficient importance sampling estimator for the rare
+  event probability $\P(|V| > u)$ where $V$ is a vector following the
+  stationary distribution of a GARCH($p, q$) process. Recall
+  $V_t = A_t V_{t-1} + B$ is the matrix recurrence equation of a
+  GARCH($p, q$) process. We emanate the process $\{V_t\}$ from a set
+  $\mathcal C = \{V: |V| \leq M\}$ for some predefined positive
+  constant $M$, and then we introduce a dual change of measure for the
+  originally iid matrices $\{A_t\}$: for
+  $t < T_u = \min\{i > 0: |V_i| > u\}$, we exponentially tilt the
+  distribution of $A_t | X_{t-1}$, where $X_0 = V_0$ and
+  $X_t = \prod_{i=1}^t X_0/|\prod_{i=1}^t X_0|$, so that
+  $|A_t X_{t-1}|^\xi$ is more likely to take on large values and hence
+  $|V_t|$ is more likely to exceed $u$. Once the exeedance has
+  happened, we change the
+  distribution of $\{A_t\}$ back to the original and continue the
+  process until $V_t$ returns to the set $\mathcal C$. Along each
+  simulated path emenating from $\mathcal C$ and ending in
+  $\mathcal C$ we compute $N_u = \sum_{i=1}^K \1{|V_i| > u}$, where
+  $K = \min\{i > 0: |V_i| \leq M\}$.
+
+  The pursuit estimate of $P(|V|>u)$ is then a weighted average of
+  $N_u$ computed along each path. The weight depends on the path.
+
+\subsection{Eigenvalues of the sample covariance matrix of a   
+stochastic volatility model}
+We consider a multivariate heavy-tailed stochastic volatility model
+and analyze the large-sample behavior of its sample covariance
+matrix. We study the limiting behavior of its entries in the
+infinite-variance case and derive results for the ordered eigenvalues
+and corresponding eigenvectors. Essentially, we consider two different
+cases where the tail behavior either stems from the iid\ innovations
+of the process or from its volatility sequence. In both cases, we make
+use of a large deviations technique for regularly varying time series
+to derive multivariate $\alpha$-stable limit distributions of the
+sample covariance matrix. While we show that in the case of
+heavy-tailed innovations the limiting behavior resembles that of
+completely independent observations, we also derive that in the case
+of a heavy-tailed volatility sequence the possible limiting behavior
+is more diverse, i.e.\ allowing for dependencies in the limiting
+distributions which are determined by the structure of the underlying
+volatility sequence.
+
+\subsection{Extreme value analysis for the sample autocovariance
+  matrices of time series}
+We  provide some  asymptotic theory for the largest eigenvalues of a
+sample covariance matrix of a $p$-dimensional \ts\ where the dimension
+$p=p_n$ converges to infinity when the sample size $n$ increases. 
+We give a short overview of the literature on the topic both in the
+light- and heavy-tailed cases when the data have finite (infinite)
+fourth moment, respectively. Our main focus is on the heavy-tailed
+case. In this case, one has a theory for the \pp\ of the normalized
+eigenvalues of the sample covariance matrix in the iid case but also
+when rows and columns of the data are linearly dependent. We provide
+limit results for the weak \con\ of these \pp es to Poisson or cluster
+Poisson processes. Based on this \con\ we can also derive the limit
+laws of various \fct als of the ordered eigenvalues such as the 
+joint \con\ of a finite number of the largest order statistics, the
+joint limit law of the largest eigenvalue and the trace, limit laws
+for successive ratios of ordered eigenvalues, etc. We also develop
+some limit theory for the singular values of the sample autocovariance
+matrices and their sums of squares. The theory is illustrated for
+simulated data and for the components of the S\&P 500 stock index.

### 2017-08-11 phd/Chapterintroduction.tex
Thomas is not happy about the contribution part.

@@
 with index $\alpha_j$. Then by the summation property, each and every
-$\{X_i\}$ is regularly varying with index $\min_{1 \leq j \leq K} \alpha_j$.
+$\{X_{i,t}\}$ is regularly varying with index $\min_{1 \leq j \leq K} \alpha_j$.
 In practice a factor $Y_{i,t}$ is estimated as
@@
 One of the most important dynamical mechanisms that lead to regularly
-varying r.v. is stochastic recursion of the following form:
+varying random variables is a stochastic recursion of the following form:
 \begin{equation}
@@
 random matrix and $B_t$ is a $d$-dimensional vector, random or
-deterministic. The sequence $\{(A_t, B_t)\}_{t \in \mathbb Z}$ are
+deterministic. The sequence $\{(A_t, B_t)\}_{t \in \mathbb Z}$ is
 iid. The stationary solution to \eqref{eq:rhjyu} satisfies the fixed-point
@@
   \label{eq:jorg}
-  u^\alpha \P(u^{-1} V \in \cdot) \overset{v}{\to} \nu_\alpha(\cdot)
+  u^\alpha \P(u^{-1} X \in \cdot) \overset{v}{\to} \nu_\alpha(\cdot)
 \end{equation}
@@
 $\reals^d_+ \setminus \{0\}$ with the property
-$\nu_\alpha(a A) = a^{-\alpha} \nu_\alpha(A)$.
+$\forall a > 0, \nu_\alpha(a A) = a^{-\alpha} \nu_\alpha(A)$.
 
@@
 \begin{enumerate}
-  \item The Markov chain $X_n$ on $\sphere^{d-1}$, namely
-    $X_n = A_n X_{n-1} / |A_n X_{n-1}|$, is irreducible, i.e. for any open
-    set $U \subset \sphere^{d-1}$ and any $u \in \sphere^{d-1}$, 
-    $\exists n \geq 1$ such that $\P(X_n u \in U) > 0$.
+  \item For any open set $U \subset \sphere^{d-1}$ and any
+    $u \in \sphere^{d-1}$,  $\exists n \geq 1$ such that
+    \[
+    \P\left({\prod_{i=1} A_i u
+        \over
+        \left|\prod_{i=1} A_i u\right|}\in U
+    \right) > 0.
+    \]
   \item There exist $N \geq 1$, $c, \epsilon > 0$ and an invertible
@@
 \[
-\lim_{u \to \infty} u^\alpha \P(\inn{x, V} > u) = e_\alpha(x)
+\lim_{u \to \infty} u^\alpha \P(\inn{x, X} > u) = e_\alpha(x)
 \]
@@
 
-Bollerslev \cite{bollerslev:1986} showed that the fixed-point equation
-$Y \overset{d}{=} A Y + B$ has a unique, strictly stationary solution
-$Y$ with finite variance if and only if
-\begin{equation}
-  \sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1
-  \label{eq:bollerslev:intro}
-\end{equation}
+% Bollerslev \cite{bollerslev:1986} showed that the fixed-point equation
+% $Y \overset{d}{=} A Y + B$ has a unique, strictly stationary solution
+% $Y$ with finite variance if and only if
+% \begin{equation}
+%   \sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1
+%   \label{eq:bollerslev:intro}
+% \end{equation}
 
@@
 $\{(X_i, \sigma_i^2)\}_{i=0}^{t-1}$; $\omega, \{\alpha_i\}_{i=1}^p,
-\{\beta_i\}_{i=1}^q$
-are non-negative parameters of the model. The GARCH($p,q$) recurrence
-equation is of the form of \eqref{eq:rhjyu}. With appropriate
-conditions, $\sigma_t^2$ is shown to be a positive Harris recurrent
-Markov chain (cf. Bollerslev \cite{bollerslev:1986} and Buraczewski et al.
+\{\beta_i\}_{i=1}^q$ are non-negative parameters of the model.
+Written in matrix form, as shown in \eqref{eq:jopger}, the
+GARCH($p,q$) recurrence  equation is of the form of
+\eqref{eq:rhjyu}. With appropriate conditions,
+$(\sigma_t^2, \dots, \sigma_{t-q+1}^2, X_{t-1}^2, \dots, X_{t - p
+  +1}^2)^\top$ is shown to be a positive Harris recurrent Markov chain
+(cf. Bollerslev \cite{bollerslev:1986} and Buraczewski et al.
 \cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
@@
 the logarithmic price $Y_t$ as a subordinated stochastic process:
-$Y_t = V_{\tau_t}$, where $V_i$ is a normally distributed random
-variable with mean 0, variance $\sigma^2 i$ and independent
-increments. $\tau_t$ is a real-valued, non-negative, non-decreasing
+$Y_t = V_{\tau_t}$, where $V_i/\sigma^2$ is a standard Brownian
+motion. $\tau_t$ is a real-valued, non-negative, non-decreasing
 sequence with $\tau_0 = 0$. It models a time change. As pointed out by
@@
 \end{equation}
-where $-1 \leq \{\psi_l\}_{l \in \mathbb Z} \leq 1$
-is a sequence of real numbers with at least one non-zero element,
-$\{\eta_t\}_{t \in \mathbb Z}$ is a iid sequence of random
+where $\{\psi_l\}_{l \in \mathbb Z}$ is a sequence of real numbers
+with at least one non-zero element,
+$\{\eta_t\}_{t \in \mathbb Z}$ is an iid sequence of random
 variables. By Kolmogorov's 3-series theorem, the infinite series above
-converges if and only if $\sum_{l \in \mathbb Z} \psi_l^2 < \infty$.
+converges if and only if $\sum_{l \in \mathbb Z} \psi_l^2 < \infty$,
+and in this case $\{\sigma_{t}\}$ is stationary.
 In particular, if $\{\eta_i\}_{i \in \mathbb Z}$ is normally
-distributed with 0 mean, $\{\sigma_t\}_{t \in \mathbb Z}$ is
-stationary and its marginal distributions are log-normal.
+distributed with zero mean, $\{\sigma_t\}_{t \in \mathbb Z}$ has
+log-normal marginal distributions.
 
-If, however, the sequence $\{Z_t\}_{t \in \mathbb Z}$ or
-$\{\sigma_t\}_{t \in \mathbb Z}$ is regularly varying with index
+If, however, the sequences $\{Z_t\}_{t \in \mathbb Z}$ or
+$\{\sigma_t\}_{t \in \mathbb Z}$ are regularly varying with index
 $\alpha$ and some additional conditions are satisfied, $X_t$ is also
@@
 
-If, instead, $\sigma_t$ has a heavier tail than $Z_t$, then $X_{i,t}$
-is regularly varying assuming the following conditions:
-\begin{itemize}
-\item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
-  is a generic element of the iid sequence
-  $\{\eta_t\}_{t \in \mathbb Z}$, and
-  \begin{equation}
-    \label{eq:dfhty}
-    \E e^{\alpha \eta} = \infty, \quad\text{ or }
-    {\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)} \to c \in (0, \infty)
-  \end{equation}
-    where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
-    iid sequence $\{\eta_t\}_{t \in \mathbb Z}$.
-\end{itemize}
-Note the only finite limit for
-${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ is
-$\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
-are mutually exclusive. In chapter \ref{ch:bernoulli} we will see that
-these two conditions lead to rather different behavior for
-eigenvalues of the sample covariance matrix of stochastic volatility
-processes of the form \eqref{eq:rfht}.
+% If, instead, $\sigma_t$ has a heavier tail than $Z_t$, then $X_{i,t}$
+% is regularly varying assuming the following conditions:
+% \begin{itemize}
+% \item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
+%   is a generic element of the iid sequence
+%   $\{\eta_t\}_{t \in \mathbb Z}$, and
+%   \begin{equation}
+%     \label{eq:dfhty}
+%     \E e^{\alpha \eta} = \infty, \quad\text{ or }
+%     {\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)} \to c \in (0, \infty)
+%   \end{equation}
+%     where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
+%     iid sequence $\{\eta_t\}_{t \in \mathbb Z}$.
+% \end{itemize}
+% Note the only finite limit for
+% ${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ is
+% $\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
+% are mutually exclusive. In chapter \ref{ch:bernoulli} we will see that
+% these two conditions lead to rather different behavior for
+% eigenvalues of the sample covariance matrix of stochastic volatility
+% processes of the form \eqref{eq:rfht}.
 
@@
 \end{eqnarray}
-\eqref{eq:hytre} and \eqref{eq:jieg} shows $\E Y_t^2$, or in other
+\eqref{eq:hytre} and \eqref{eq:jieg} show that $\E Y_t^2$, or in other
 words, the forecast of future squared return can be obtained as
 $\E [Y_\delta]_t$, i.e. the forecast of future realized quadratic
-variation. Since past realized quadratic variations are observable,
-forcast of future $[Y_\delta]_t$ can be obtained using e.g. an ARFIMA
-model.
+variation.
+
+So far we have been discussing univariate stochastic volatility
+models. In a multivariate context, we may model a $p$-variate process
+as
+\begin{eqnarray*}
+  X_{i,t} &=& \sigma_{i,t} Z_{i,t}, \quad i = 1,2,\dots,p\\
+  \log \sigma_{i, t}
+  &=&
+  \sum_{k,l \in \mathbb Z} \psi_{k,l} \eta_{i-k, t-l}
+\end{eqnarray*}
+where $Z_{i,t}$ is an iid field of random variables, and the
+coefficients $\{\psi_{k,l}\}$ are real and satisfy
+$\sum_{k, l \in \mathbb Z} |\psi_{k,l}| < \infty$.
+% Since past realized quadratic variations are observable,
+% forcast of future $[Y_\delta]_t$ can be obtained using e.g. an ARFIMA
+% model.
 
@@
 tail parameters of the equity, the interest rate, the investor's risk
-apetite as captured by his utility function, and his threshold of
+appetite as captured by his utility function, and his threshold of
 disappointment.

### 2017-08-21 papers/DavisMikoschHeinyXie/PaperBody.tex
Revised the phd thesis.

@@
 a_{np}^{-2} \big(\la_{(1)}-\la_{(2)}\big)\cid
-6 \,\Gamma_1^{-2/\alpha} \1_{\{\Gamma_1 4^{\alpha/2 }<\Gamma_2\}}+
+6 \,\Gamma_1^{-2/\alpha} \I_{\{\Gamma_1 4^{\alpha/2 }<\Gamma_2\}}+
 8\,
-\big(\Gamma_1^{-2/\alpha}-\Gamma_2^{-2/\alpha}\big)\1_{\{\Gamma_14^{\alpha/2
+\big(\Gamma_1^{-2/\alpha}-\Gamma_2^{-2/\alpha}\big)\I_{\{\Gamma_14^{\alpha/2
   }> \Gamma_2\}}
@@
 \dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}}&\cid
-\dfrac{6}{8} \, \1_{\{\Gamma_1 2^{\alpha }<\Gamma_2\}}+
-\big(1-(\Gamma_1/\Gamma_2)^{2/\alpha}\big)\1_{\{\Gamma_12^{\alpha
+\dfrac{6}{8} \, \I_{\{\Gamma_1 2^{\alpha }<\Gamma_2\}}+
+\big(1-(\Gamma_1/\Gamma_2)^{2/\alpha}\big)\I_{\{\Gamma_12^{\alpha
   }> \Gamma_2\}}\\
-& =\dfrac{3}{4}\,  \1_{\{U 2^{\alpha }<1\}}+
-\big(1-U^{2/\alpha}\big)\1_{\{U2^{\alpha
+& =\dfrac{3}{4}\,  \I_{\{U 2^{\alpha }<1\}}+
+\big(1-U^{2/\alpha}\big)\I_{\{U2^{\alpha
   }> 1\}}=Y\,.
@@
 (a_{np}^{-2}\lambda_{(1)},\lambda_{(2)}/\lambda_{(1)}) \cid(
-8\,\Gamma_1^{-2/\alpha},\frac{1}{4}\,\1_{\{U<2^{-\alpha}\}}+U^{2/\alpha}\,\1_{\{U\ge
+8\,\Gamma_1^{-2/\alpha},\frac{1}{4}\,\I_{\{U<2^{-\alpha}\}}+U^{2/\alpha}\,\I_{\{U\ge
   2^{-\alpha}\}})\,
@@
 the histogram is remarkably close to what one would expect from a sample from the truncated
-uniform \ds , $2^{-\alpha}\, \1_{\{U<2^{-\alpha}\}}+U\, \1_{\{U\ge 2^{-\alpha} \}}$.
+uniform \ds , $2^{-\alpha}\, \I_{\{U<2^{-\alpha}\}}+U\, \I_{\{U\ge 2^{-\alpha} \}}$.
 The mass of the limiting discrete component of the ratio can be much
@@
 \dfrac{\lambda_{(2)}}{\lambda_{(1)}}\cid
-\frac{v_2}{v_1}\,\1_{\{U<(v_2/v_1)^{\alpha/2}\}}+U^{2/\alpha}\,\1_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
+\frac{v_2}{v_1}\,\I_{\{U<(v_2/v_1)^{\alpha/2}\}}+U^{2/\alpha}\,\I_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
 \end{equation*}
@@
 \beao
-\dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}} \cid \dfrac{v_1-v_2}{v_1}\,\1_{\{U<(v_2/v_1)^{\alpha/2}\}}+(1-U^{2/\alpha})\,\1_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
+\dfrac{\la_{(1)}-\la_{(2)}}{\la_{(1)}} \cid \dfrac{v_1-v_2}{v_1}\,\I_{\{U<(v_2/v_1)^{\alpha/2}\}}+(1-U^{2/\alpha})\,\I_{\{U\ge (v_2/v_1)^{\alpha/2}\}}\,.
 \eeao
@@
   X_{it} = -\Big[
-    \log \Big(\frac{1}{n+1} \sum_{\tau=1}^n \1_{\{R_{i\tau} \leq R_{it}\}} \Big)
+    \log \Big(\frac{1}{n+1} \sum_{\tau=1}^n \I_{\{R_{i\tau} \leq R_{it}\}} \Big)
   \Big]^{-1} \,,\qquad i=1,\ldots,p;t=1,\ldots,n\,.
@@
 \cite{resnick:1987,resnick:2007})
-for a  Poisson random measure $N$ with state space $\overline{\R}_0$ and intensity measure $\mu_\alpha(dx)=\alpha |x|^{-\alpha-1} (p_+ \1_{\{x>0\}}+ p_- \1_{\{x<0\}}) dx$.
+for a  Poisson random measure $N$ with state space $\overline{\R}_0$ and intensity measure $\mu_\alpha(dx)=\alpha |x|^{-\alpha-1} (p_+ \I_{\{x>0\}}+ p_- \I_{\{x<0\}}) dx$.
 \end{lemma}

### 2017-08-21 phd/Chapterintroduction.tex
Revised the phd thesis.

--- 2017-08-11 phd/Chapterintroduction.tex
@@
 \quad
-\forall c > 0
+\forall c > 0,
 \]
@@
 variation is better described in terms of vague convergence to a
-spectral measure $\mu_\alpha$: if a random vector $V$ satisfies
+spectral measure $\mu_\alpha$: if a random vector $X$ satisfies
 \[
@@
 $X_1, X_2, \dots$ of regularly varying variables is particularly
-important for understanding the behaviour of a heavy-tailed series. A
+important for understanding the behavior of a heavy-tailed series. A
 standard method proposed for this purpose is due to Hill
@@
 $\hat \alpha_H$, under the assumptions $k \to \infty, k/n \to 0$
-as $n \to \infty$.
+as $n \to \infty$. See theorem 6.4.6 of Embrechts et. al.
+\cite{embrechts:klueppelberg:mikosch:1997} for details.
 
@@
 of daily stock return series from 3 sectors of the
-{\em Standard \& Poor's 500} index \footnote{An American stock index
+{\em Standard \& Poor's 500} index\footnote{An American stock index
   comprising around 500 companies.}. The 2.5\% and 97.5\% quantiles of
@@
 similar they really are and if/how their variations can be explained by
-economic arguments. We explore this topic in chapter
-\ref{ch:TailParameters}.
+economic arguments.
+
+%% We explore this topic in chapter \ref{ch:TailParameters}.
 
@@
 features: if $X_1$ and $X_2$ are both regularly varying with indices
-$\alpha_1$ and $\alpha_2$, respectively, then $a_1 X_1 + a_2 X_2$ is
-regularly varying with index $\min\{\alpha_1,
-\alpha_2\}$ (cf. Mikosch and Jessen
+$\alpha_1$ and $\alpha_2$, respectively, then $a_1 X_1 + a_2 X_2$,
+for $a_1, a_2 > 0$, is regularly varying with index
+$\min\{\alpha_1, \alpha_2\}$ (cf. Mikosch and Jessen
 \cite{JessenMikosch2006}). Moreover, if $X_1, X_2$ are iid,
@@
 Jessen \cite{JessenMikosch2006}.
+
+The notion of a regularly varying strictly stationary process is also of
+interest. Originally introduced by Davis and Hsing
+\cite{davis:hsing:1995} for univariate processes, the concept was
+extended to multivariate processes by Davis and Mikosch
+\cite{davis:mikosch:2009b}: An $\reals^d$-valued strictly stationary
+process $X_t$ is said to be regularly varying with index $\alpha$, if
+for each $h \geq 0$, 
+\[
+  {
+    \P\left[x^{-1}(X_0, X_1, \dots, X_h) \in \cdot \right]
+    \over
+    \P(|X_0| > x)
+  } \overset{v}{\to} \nu_\alpha(\cdot),
+  \quad x \to \infty
+\]
+where $\nu_\alpha$ a non-null Radon measure on
+$\reals^{d(h+1)} \setminus \{0\}$ that
+is homogeneous of order $-\alpha$, i.e.
+$\forall c > 0, \forall S \subset \reals^{d(h+1)}\setminus\{0\}, \nu_\alpha(c S) = c^{-\alpha} \nu_\alpha(S)$. $|\cdot|$ could be any given vector norm.
+
+Basrak and Segers \cite{basrak:segers:2009} proved an equivalence
+condition for regular variation of
+a strictly stationary process: an $\reals^d$-valued strictly stationary
+process $\{X_t\}$ is regularly varying with index $\alpha$ if 
+there exists an $\reals^d$-valued sequence $\{\Theta_i\}_{i = 0,1,\ldots}$
+and a Pareto($\alpha$) random variable $Z$, i.e.
+$\P(Z > x) = x^{-\alpha}, \forall x \geq 1$, independent of
+$\{\Theta_i\}$
+\[
+\P\left[\left.
+  x^{-1}(X_0, X_1, \ldots) \in \cdot
+  \right| |X_0| > x
+  \right]
+  \overset{w}{\to}
+  \P(Z(\Theta_1, \Theta_2, \ldots) \in \cdot),
+  \quad
+  x \to \infty.
+\]
+The sequences $\{\Theta_i\}_{i = 0,1,\ldots}$ and
+$\{Z \Theta_i\}_{i = 0,1,\ldots}$ are called the spectral tail process
+and the tail process of $\{X_t\}$, respectively.
+
+A recent remarkable result from Mikosch and Wintenberger
+\cite{mikosch:wintenberger:2016} regarding
+strictly stationary regularly varying processes is quoted below:
+\begin{theorem}\label{thm:mikwin:intro}
+Let $(Y_t)$ be an $\bbr^d$-valued strictly stationary sequence , $S_n=Y_1+\cdots +Y_n$ and $(a_n)$ be such that $n\,\P(|Y|>a_n)\to 1$.
+Also write for $\vep>0$, $\ov Y_t = Y_t \I(|Y_t|\le \vep a_n)$, $\underline Y_t= Y_t-\ov  Y_t $ and
+\beao
+\ov S_{l,n}= \sum_{t=1}^l \ov Y_t \qquad  \un S_{l,n}= \sum_{t=1}^l \un Y_t\,.
+\eeao
+Assume the following conditions:
+\begin{enumerate}
+\item 
+$(Y_t)$ is regularly varying with index $\alpha\in (0,2)  \setminus\{1\}$ and spectral tail process $(\Theta_j)$.
+\item
+A mixing condition holds: there exists an integer \seq\ $m_n\to\infty$ \st\ $k_n= [n/m_n]\to \infty$
+and 
+\beam\label{eq:chfa:intro}
+\E \ex^{it'\underline S_n/a_n} - \Big(\E \ex^{it'\underline S_{m_n,n}/a_n}\Big)^{k_n} \to 0\,,\qquad \nto\,,\qquad t\in\bbr^d\,.
+\eeam
+\item An anti-clustering condition holds:
+\beam\label{eq:acl:intro}
+\lim_{l\to\infty} \limsup_{\nto} \P\big(\max_{t=l,\ldots,m_n} |Y_t|>\delta a_n\mid |Y_0|>\delta a_n\big)=0\,,\qquad \delta>0\,
+\eeam
+for the same sequence $(m_n)$ as in (2).
+\item If $\alpha\in (1,2)$, in addition $\E[Y]=0$ and the vanishing small values condition holds: 
+\beam\label{eq:vansm:intro}
+\lim_{\vep\downarrow 0}\limsup_{\nto} \P\big(a_n^{-1} |\ov S_n-\E [\ov S_n]|>\delta \big)=0\,,\qquad \delta>0\,
+\eeam
+and $\sum_{i=1}^\infty \E[|\Theta_i|]<\infty$.
+\end{enumerate}
+Then $a_n^{-1} S_n\std \xi_\alpha$ for an $\alpha$-stable $\bbr^d$-valued
+vector $\xi_\alpha$ with log characteristic function
+\begin{small}
+  \beam\label{eq:chfid:intro}
+  \int_0^\infty \E\left[
+    \exp\left(i\,y\,t'\sum_{j=0}^\infty \Theta_j\right)-
+    \exp\left(i\,y\,t'\sum_{j=1}^\infty \Theta_j\right)
+    -i\,y\,t'\I_{(1,2)}(\alpha)
+    \right]\,
+  d(-y^{\alpha})\,,\qquad t\in\bbr^d\,.
+  \eeam
+\end{small}
+\end{theorem}\noindent 
+\begin{remark}\label{rem:case:alpha:1:intro}
+If we additionally assume that $Y$ is symmetric, which implies $\E[\ov
+  Y]=\mathbf{0}$, then the statement of the theorem also holds for
+$\alpha=1$.
+\end{remark}
+
+Using theorem \ref{thm:mikwin:intro} we prove in chapter
+\ref{ch:bernoulli} joint convergence in distribution of eigenvalues of
+a sample covariance matrix of stochastic volatility processes (see
+\S\ref{sec:SV:intro}) to $\alpha/2$-stable random variables.
 
@@
 \section{Stochastic Recurrence Equation}
+\label{sec:tyhyjyt}
 One of the most important dynamical mechanisms that lead to regularly
-varying r.v. is stochastic recursion of the following form:
+varying random vectors is a stochastic recursion of the following form:
 \begin{equation}
   \label{eq:rhjyu}
-  X_t = A_t X_{t-1} + B_t, \quad t \in \mathbb Z
+  X_t = A_t X_{t-1} + B_t, \quad t \in \integers
 \end{equation}
@@
 random matrix and $B_t$ is a $d$-dimensional vector, random or
-deterministic. The sequence $\{(A_t, B_t)\}_{t \in \mathbb Z}$ are
+deterministic. The sequence $\{(A_t, B_t)\}_{t \in \integers}$ is
 iid. The stationary solution to \eqref{eq:rhjyu} satisfies the fixed-point
 equation $X \overset{d}{=} A X + B$, where $X$ and $(A, B)$ are
-generic elements of the $\{X_t\}_{t \in \mathbb Z}$ and
-$(A_t, B_t)_{t \in \mathbb Z}$ sequences.
+generic elements of the $\{X_t\}_{t \in \integers}$ and
+$\{(A_t, B_t)\}_{t \in \integers}$ sequences.
 
-Kesten \cite{kesten:1973} showed that, when $A_t$ are almost
+Kesten \cite{kesten:1973} showed that, when $A_t$ is almost
 surely non-negative, has no row or column of only zeros, and
@@
     \[
-    1 = \lambda(\alpha) = \lim_{n \to \infty} {1 \over n} \log \E \|A_n \cdots A_1\|^\alpha
+    1 = \lambda(\alpha) = \lim_{n \to \infty} {1 \over n} \log \E
+    \|A_n \cdots A_1\|^\alpha
     \]
@@
 equation \eqref{eq:garchpq_sre} was stuck at this point with probability 1: 
-$\P(A X + B = X) < 1$ for all $X \in \reals^d$. With these
+$\P(A x + B = x) < 1$ for all $x \in \reals^d$. With these
 assumptions, they showed
 \[
-\lim_{u \to \infty} u^\alpha \P(\inn{x, V} > u) = e_\alpha(x)
+\lim_{u \to \infty} u^\alpha \P(\inn{x, X} > u) = e_\alpha(x)
 \]
@@
 addition, a proximity condition is satisfied by the support of
-$A$. This is the result of Guivarc'h and Le Page.
+$A$. This is the result of Guivarc'h and Le Page
 \cite{guivarc:page:2016}. Let $G_A$ denote the semi-group generated
@@
 \[
-\E \|A\|^{\alpha + \delta} < \infty, \quad
+\E (\|A\|^{\alpha + \delta}) < \infty, \quad
 \E (\|A\|^{\alpha} \|A^{-1}\|^{\delta}) < \infty, \quad
-\E (|B|^{\alpha + \delta} < \infty),\quad
-\text{ for some }\delta > 0
+\E (|B|^{\alpha + \delta}) < \infty \quad
+\text{ for some }\delta > 0,
 \]
-Guivarc'h and Le Page proved the vague convergence result of \eqref{eq:jorg}.
+Guivarc'h and Le Page proved the vague convergence result of
+\eqref{eq:jorg}.
 
-Bollerslev \cite{bollerslev:1986} showed that the fixed-point equation
-$Y \overset{d}{=} A Y + B$ has a unique, strictly stationary solution
-$Y$ with finite variance if and only if
-\begin{equation}
-  \sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1
-  \label{eq:bollerslev:intro}
-\end{equation}
+% Bollerslev \cite{bollerslev:1986} showed that the fixed-point equation
+% $Y \overset{d}{=} A Y + B$ has a unique, strictly stationary solution
+% $Y$ with finite variance if and only if
+% \begin{equation}
+%   \sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1
+%   \label{eq:bollerslev:intro}
+% \end{equation}
 
 \section{GARCH models}
+\label{sec:garch:intro}
 Introduced by Bollerslev \cite{bollerslev:1986} in 1986,
@@
 A GARCH($p, q$) model is a stationary process
-$\{X_t\}_{t \in \mathbb Z}$satisfying
-\begin{eqnarray*}
+$\{X_t\}_{t \in \integers}$ satisfying
+\begin{eqnarray}
+  \label{eq:garch:intro}
   X_t &=& \sigma_t Z_t \\
@@
   \sum_{j=1}^q \beta_j \sigma_{t-j}^2
-\end{eqnarray*}
-where $\{X_t\}$ is a return series, e.g. stock returns, foreign exchange
-rates, interest rates, etc; $\{Z_t\}$ is an iid, mean 0, unit-variance
-sequence, $\sigma_t^2$ is the variance of the distribution of $X_t$
-conditional on
-$\{(X_i, \sigma_i^2)\}_{i=0}^{t-1}$; $\omega, \{\alpha_i\}_{i=1}^p,
-\{\beta_i\}_{i=1}^q$
-are non-negative parameters of the model. The GARCH($p,q$) recurrence
-equation is of the form of \eqref{eq:rhjyu}. With appropriate
-conditions, $\sigma_t^2$ is shown to be a positive Harris recurrent
-Markov chain (cf. Bollerslev \cite{bollerslev:1986} and Buraczewski et al.
+\end{eqnarray}
+where $\{X_t\}$ is a model for return series of stock prices,
+foreign exchange rates, interest rates, etc; $\{Z_t\}$ is an iid,
+mean 0, unit-variance sequence, $\sigma_t^2$ is the variance of the
+distribution of $X_t$ conditional on
+$\{(X_i, \sigma_i^2)\}_{i < t}$; $\omega, \{\alpha_i\}_{i=1}^p,
+\{\beta_i\}_{i=1}^q$ are non-negative parameters of the model.
+Written in matrix form, as shown in \eqref{eq:jopger}, the
+GARCH($p,q$) recurrence  equation is of the form of
+\eqref{eq:rhjyu}. With appropriate conditions,
+$(\sigma_t^2, \dots, \sigma_{t-q+1}^2, X_{t-1}^2, \dots, X_{t - p +1}^2)$
+is shown to be a positive Harris recurrent Markov chain
+(cf. Bollerslev \cite{bollerslev:1986} and Buraczewski et al.
 \cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
@@
   \label{eq:grgbg}
-  \lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\alpha = 0,
+  \Lambda(\alpha) = \lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\alpha = 0,
 \end{equation}
-where $\{A_i\}_{i \in \mathbb Z}$ are iid matrices whose entries are
+where $\{A_i\}_{i \in \integers}$ are iid matrices whose entries are
 functions of $\{\alpha_i\}_{i=1}^p$, $\{\beta_i\}_{i=1}^q$ and
 $\{Z_t^2\}$:
-\[
-A_i =
+\begin{equation}
+  \label{eq:A_matrix:intro}
+  A_t =
   \begin{pmatrix}
@@
   \end{pmatrix}
-\]
+\end{equation}
 While GARCH models have been very successful for modelling financial
-time series, it does have its drawbacks. For example, the tail
+time series, they do have their drawbacks. For example, the tail
 index is very sensitive to the model parameters $\{\alpha_i\}_{i=1}^p$ and
@@
 \alpha_{21}Z_{1,t-1}^2+\beta_{21}& \alpha_{22}Z_{2,t-1}^2+\beta_{22}
-\earr\right)\,,
+\earr\right),
 \eeao
@@
 
-\section{Stochastic Volatility models}
+\section{Stochastic volatility models}
+\label{sec:SV:intro}
 With the availability of high-frequency data, a different approach
@@
 the logarithmic price $Y_t$ as a subordinated stochastic process:
-$Y_t = V_{\tau_t}$, where $V_i$ is a normally distributed random
-variable with mean 0, variance $\sigma^2 i$ and independent
-increments. $\tau_t$ is a real-valued, non-negative, non-decreasing
+$Y_t = V_{\tau_t}$, where $V_i$ is a Brownian motion.
+$\tau_t$ is a real-valued, non-negative, non-decreasing
 sequence with $\tau_0 = 0$. It models a time change. As pointed out by
-Shepard and Andersen \cite{Shephard:Andersen:2009}, the log-price
-process $Y_t$ is serially uncorrelated although potentially dependent,
+Shephard and Andersen \cite{Shephard:Andersen:2009}, the log-price
+process $\{Y_t\}$ is serially uncorrelated although potentially dependent,
 provided that $V_t$ and $\tau_t$ are independent.
@@
 \end{eqnarray}
-where $A_t$ is a finite variation process and $M_t$ is a martingale
-and hence $Y_t$ is a semimartingale. $\sigma_t$ is non-negative and
-$Z_t$ is an iid process with 0 mean and unit variance. A convenient
-choice of $\sigma_t$ is
+where $\{A_t\}$ is a finite-variation process and $\{M_t\}$ is a
+martingale and hence $\{Y_t\}$ is a semimartingale.
+$\{\sigma_t\}$ is non-negative and $Z_t$ is an iid process with zero mean
+and unit variance. A convenient choice of $\sigma_t$ is
 \begin{equation}
   \label{eq:rfht}
-  \log\sigma_t = \sum_{l \in \mathbb Z} \psi_l \eta_{t-l}  
+  \log\sigma_t = \sum_{l \in \integers} \psi_l \eta_{t-l}
+  \quad
+  t \in \integers,
 \end{equation}
-where $-1 \leq \{\psi_l\}_{l \in \mathbb Z} \leq 1$
-is a sequence of real numbers with at least one non-zero element,
-$\{\eta_t\}_{t \in \mathbb Z}$ is a iid sequence of random
+where $\{\psi_l\}_{l \in \integers}$ is a sequence of real numbers
+with at least one non-zero element,
+$\{\eta_t\}_{t \in \integers}$ is an iid sequence of random
 variables. By Kolmogorov's 3-series theorem, the infinite series above
-converges if and only if $\sum_{l \in \mathbb Z} \psi_l^2 < \infty$.
-In particular, if $\{\eta_i\}_{i \in \mathbb Z}$ is normally
-distributed with 0 mean, $\{\sigma_t\}_{t \in \mathbb Z}$ is
-stationary and its marginal distributions are log-normal.
+converges if and only if $\sum_{l \in \integers} \psi_l^2 < \infty$,
+and in this case $\{\sigma_{t}\}$ is stationary.
+In particular, if $\{\eta_i\}_{i \in \integers}$ is normally
+distributed with zero mean, $\{\sigma_t\}_{t \in \integers}$ has
+log-normal marginal distributions.
 
-If, however, the sequence $\{Z_t\}_{t \in \mathbb Z}$ or
-$\{\sigma_t\}_{t \in \mathbb Z}$ is regularly varying with index
-$\alpha$ and some additional conditions are satisfied, $X_t$ is also
-regularly varying with the same index. Specifically, if $Z_t$ is
-regularly varying and $\sigma_t$ has a lighter tail, the conclusion
-follows from Breiman's lemma. See \S\ref{sec:case1} of Jan\ss en et
-al. i.e. chapter \ref{ch:bernoulli} of this thesis for more details.
+If, however, the sequences $\{Z_t\}_{t \in \integers}$ or
+$\{\sigma_t\}_{t \in \integers}$ are regularly varying with index
+$\alpha$ and some additional conditions are satisfied, $\{X_t\}$
+is also regularly varying with the same index. Specifically,
+if $\{Z_t\}$ is regularly varying and $\{\sigma_t\}$ has a lighter
+tail, the conclusion follows from Breiman's lemma \cite{breiman:1965}.
+See \S\ref{sec:case1} of Jan\ss en et al. i.e. chapter
+\ref{ch:bernoulli} of this thesis for more details.
 
-If, instead, $\sigma_t$ has a heavier tail than $Z_t$, then $X_{i,t}$
-is regularly varying assuming the following conditions:
-\begin{itemize}
-\item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
-  is a generic element of the iid sequence
-  $\{\eta_t\}_{t \in \mathbb Z}$, and
-  \begin{equation}
-    \label{eq:dfhty}
-    \E e^{\alpha \eta} = \infty, \quad\text{ or }
-    {\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)} \to c \in (0, \infty)
-  \end{equation}
-    where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
-    iid sequence $\{\eta_t\}_{t \in \mathbb Z}$.
-\end{itemize}
-Note the only finite limit for
-${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ is
-$\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
-are mutually exclusive. In chapter \ref{ch:bernoulli} we will see that
-these two conditions lead to rather different behavior for
-eigenvalues of the sample covariance matrix of stochastic volatility
-processes of the form \eqref{eq:rfht}.
+% If, instead, $\sigma_t$ has a heavier tail than $Z_t$, then $X_{i,t}$
+% is regularly varying assuming the following conditions:
+% \begin{itemize}
+% \item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
+%   is a generic element of the iid sequence
+%   $\{\eta_t\}_{t \in \integers}$, and
+%   \begin{equation}
+%     \label{eq:dfhty}
+%     \E e^{\alpha \eta} = \infty, \quad\text{ or }
+%     {\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)} \to c \in (0, \infty)
+%   \end{equation}
+%     where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
+%     iid sequence $\{\eta_t\}_{t \in \integers}$.
+% \end{itemize}
+% Note the only finite limit for
+% ${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ is
+% $\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
+% are mutually exclusive. In chapter \ref{ch:bernoulli} we will see that
+% these two conditions lead to rather different behavior for
+% eigenvalues of the sample covariance matrix of stochastic volatility
+% processes of the form \eqref{eq:rfht}.
 
-There are a few advantages in using stochastic volatility models. For
-one thing, they are among the simplest models allowing for conditional
-heteroscedasticity (cf. Andersen et
-al. \cite{andersen:davis:kreiss:mikosch:2009}); for another thing,
+There are a few advantages in using stochastic volatility models.
+They are among the simplest models allowing for conditional
+heteroscedasticity (cf. Andersen et al.
+\cite{andersen:davis:kreiss:mikosch:2009}); nevertheless,
 they greatly improve the accuracy of predicting future volatilities
@@
   \overset{d}{\to} N(0, 1),
-  \quad \delta \to 0
+  \quad \delta \to 0.
 \]
@@
   Y_t^2 &=& [Y]_t + 2 \int_0^t Y_s d Y_s \nonumber \\
-  &=& [Y]_t + 2 \int_0^t Y_s dA_s + 2 \int_0^t Y_s dM_s \nonumber \\
+  &=& [Y]_t + 2 \int_0^t Y_s dA_s + 2 \int_0^t Y_s dM_s, \nonumber \\
   \E Y_t^2 &=& \E [Y]_t + 2 \E \left( \int_0^t Y_s dA_s \right) \nonumber\\
-  &\approx& \E [Y]_t, \quad \text{when $t$ is small} \label{eq:hytre}
+  &\approx& \E [Y]_t, \quad \text{when $t$ is small}. \label{eq:hytre}
 \end{eqnarray}
-\eqref{eq:hytre} and \eqref{eq:jieg} shows $\E Y_t^2$, or in other
-words, the forecast of future squared return can be obtained as
+\eqref{eq:hytre} and \eqref{eq:jieg} show that $\E Y_t^2$, or in other
+words, the forecast of future squared return, can be obtained as
 $\E [Y_\delta]_t$, i.e. the forecast of future realized quadratic
-variation. Since past realized quadratic variations are observable,
-forcast of future $[Y_\delta]_t$ can be obtained using e.g. an ARFIMA
-model.
+variation.
+% Since past realized quadratic variations are observable,
+% forcast of future $[Y_\delta]_t$ can be obtained using e.g. an ARFIMA
+% model.
 
-\section{Contribution of this thesis}\label{sec:contr}
+\section{Contributions of this thesis}\label{sec:contr}
 
@@
 \subsection{Tail parameters of equity return series}
+In chapter \ref{ch:TailParameters} we consider a
+minimal market where a riskless bond and an equity are the only assets
+available to investors. We model the investor's preference of the
+equity with {\em Generalized Disappoint Aversion (GDA)}, an idea envisaged
+by Gul \cite{gul:1991} and generalized by Routledge and Zin
+\cite{routledge2010generalized}. Specifically, in the GDA theory a
+rational investor's behavior is characterized by his attempt to
+maximize the GDA functional $U(F)$:
+\[
+U(F) = \E_F[u(C)] - b \E_F[u(\delta) - u(C)\1{C < \delta}]
+\]
+where $C$ is the investor's wealth evaluated at a fixed date, $\delta$
+is the level of disappointment -- if $C$ falls below $\delta$, the
+investor becomes disappointed. $b$ parameterizes the growth of his
+disappointment. $F$ is the distribution function of the return of the
+investor's portfolio. In the aforementioned minimal market,
+$F$ is the distribution function of the equity's return. The subscript
+$F$ of $\E_F$ reminds us that the expectation is taken with respect to
+the distribution funciton $F$.
+
 We have established that, in the case of an equity return series with
-two-sided, functionally independent Pareto tails, investor
-preference functionals are monotone increasing/decreasing with the
+two-sided, functionally independent Pareto tails, GDA preference
+functionals are monotone increasing/decreasing with the
 tail index/scale parameters. Thus in a market dominated by such
@@
 may well be the case for the ``Consumer Staples'' sector of S\&P 500,
-given the Hill estimates of tail indices shown in figure \ref{fig:1}
-and the largely positive results of tests for equal tail indices shown
-in figure \ref{fig:PairTest}.
+given the Hill estimates of tail indices shown in figure
+\ref{fig:thjyuj} and the largely positive results of tests for equal
+tail indices shown in figure \ref{fig:PairTest}.
 
@@
 tails have the same indices, investor preference over the equity has
-more sophisticated variations in the parameters' space including the
+more sophisticated variations in the parameter space including the
 tail parameters of the equity, the interest rate, the investor's risk
-apetite as captured by his utility function, and his threshold of
+appetite as captured by his utility function, and his threshold of
 disappointment.
@@
 We also acknowledge that our model of the market and the investor is a
-simple one, not accoounting for the dependence between equities, nor
+simple one, not accounting for the dependence between equities, nor
 the categorization of investors and their interactions. These are
 potential topics of future work.
+
+\subsection{Importance sampling estimator of GARCH(p,q) rare event
+  probability}
+In \S\ref{sec:tyhyjyt} we have seen how power-law tails can
+arise e.g. by Kesten's theorem in the stationary distribution of a
+Markov chain $\{V_t\}_{t \geq 0}$ described by a stochastic recurrence
+equation. In \S\ref{sec:garch:intro} we have introduced GARCH($p,q$)
+processes as examples of such Markov chains. By Kesten's theorem, the
+stationary distribution of a GARCH($p,q$) process has power-law tails
+asymptotically, i.e.
+$\P(|V| > x) = c x^{-\alpha} + o(x^{-\alpha}), c > 0, \alpha > 0$ as
+$x \to \infty$. While a nice result, this theorem does not allow us to
+compute, in precision, the probability $\P(|V| > x)$. A numerical
+procedure is needed for this purpose. A straightforward approach is of
+course direct Monte Carlo: we simulate the first $n$ steps of
+$\{V_t\}_{t \geq 0}$ and approximate $\P(|V| > x)$ as
+\begin{equation}
+  \label{eq:trhythyth}
+  \P(|V| > x) \approx {1 \over n - K} \sum_{t={n-K+1}}^n\1{|V_t| > x}
+\end{equation}
+where we discard the first $K$ steps of the simulated sample path so
+that the empirical distribution of $\{V_t\}_{t \geq n-K+1}$ is closer
+to the stationary distribution $\pi$ of $\{V_t\}$.
+
+The difficulty of this naive approach is that, when $x$ is large,
+the event $\{|V| > x\}$ happens very rarely, making the variance of
+the estimate too big to be of any use. A method to overcome this
+difficulty is importance sampling: we introduce a Markov additive
+process $\{(Y_t, S_t)\}_{t \geq 0}, |Y_t| = 1$ on the unit sphere:
+\begin{eqnarray*}
+  B &=& (\omega, 0, \cdots, 0) \\
+  V_t &=& (\sigma_t^2, \sigma_{t-1}^2, \cdots, \sigma_{t-q+1}^2,
+  X_{t-1}^2, \cdots, X_{t-p+1}^2) = A_t V_{t-1} + B\\
+  Y_t &=& {A_t \cdots A_1 V_0 \over |A_t \cdots A_1 V_0|}
+  = {A_t Y_{t-1} \over |A_t Y_{t-1}|} = A_t \cdot Y_{t-1}\\
+  l_t &=& \log |A_t Y_{t-1}| \\
+  S_t &=& \log |A_t \cdots A_1 V_0| - \log |V_0|
+  = \sum_{i=1}^t l_i
+\end{eqnarray*}
+where $X_t = \sigma_t Z_t$ (see \eqref{eq:garch:intro}). The matrices
+$\{A_t\}_{t \geq 1}$ is defined by \eqref{eq:A_matrix:intro}.
+For more details, see \S\ref{sec:IS_Results}.
+$V_t$, our main object of interest, is a function of the path
+$\{(Y_i, l_i)\}_{1 \leq i \leq t}$. To increase the chance of
+observing $\{|V_t| > x\}$, we adopt a dual change of the transition
+kernle of $\{(Y_t, l_t)\}_{t \geq 0}$. Before the first occurence of
+$\{|V_t| > x\}$ we simulate $(Y_t, l_t)$ according to a shifted
+transition kernel, whose induced probability measure is denoted
+$\P^\alpha$:
+\begin{eqnarray}
+  &&
+  \P^\alpha \left[
+    (Y_t, l_t) \in dy \times dl | Y_{t-1} = w
+    \right] \nonumber \\
+  &=&
+  e^{\alpha l} {r_\alpha(y) \over r_\alpha(w)}
+  \P\left[
+    (Y_t, l_t) \in dy \times dl | Y_{t-1} = w
+    \right] \label{eq:pgrtgpd}
+\end{eqnarray}
+where $\P$ denotes the probability with respect to the probability
+measure $\pi$, the stationary probability measure of $\{V_t\}$.
+$r_\alpha$ is a right eigenfunction of the operator
+$\mathscr P^\alpha$, which is defined by its action on
+a test function $g: \sphere^{d-1} \to \reals_+$:
+\[
+\mathscr P^\alpha g(x) = \int_{\text{dom}(Z)} |A(z^2) x|^\alpha
+g(A(z^2) \cdot x) f_{Z}(z) dz.
+\]
+$r_\alpha$ is the right eigenfunction of $\mathscr P^\alpha$
+associated with the eigenvalue
+$\lambda(\alpha) = e^{\Lambda(\alpha)} = 1$, i.e.
+$\mathscr P^\alpha r_\alpha(x) = r_\alpha(x)$.
+The function $\Lambda(\alpha)$ is defined by \eqref{eq:grgbg},
+$f_{Z}$ is the density function of $Z$ and $A(z^2)$ is the matrix
+\eqref{eq:A_matrix:intro} with $Z_{t-1}^2$ substituted for $z^2$.
+
+Conditional on $(Y_{t-1}, l_{t-1})$, the only source of randomness to
+$Y_t$ comes from $Z_{t-1}$. Hence the shift of conditional probability
+distribution shown in \eqref{eq:pgrtgpd} is equivalent to shifting the
+distribution of $Z_{t-1}$. We have
+\[
+\P^\alpha (Z_{t-1} \in dz | Y_{t-1} = w) = |A(z^2) w|^\alpha {
+  r_\alpha(A(z^2) \cdot w) \over r_\alpha(w)
+} \P(Z_{t-1} \in dz).
+\]
+Note that $\{Z_t\}$ is an iid sequence in the original measure.
+An expected value with respect to $\P^\alpha$ is related to its
+counterpart with respect to $\P$ via
+\begin{equation}
+  \label{eq:pkjrkpo}
+  \E [g(Y_t, l_t)] = \E^\alpha \left[
+    g(Y_t, l_t) e^{-\alpha l_t} {r_\alpha(Y_{t-1}) \over r_\alpha(Y_t)}
+    \right]
+\end{equation}
+Let's define $T_x = \min\{t \geq 1: |V_t| > x\}$. 
+Once the first excursion of $|V_t|$ above $x$ has occurred, i.e.
+$t > T_x$, we change the transition kernel back to the original and
+continue the simulation until the process returns to a designated set
+$\mathcal C = \{v: |v| \leq M\}$, where
+the positive number $M$ is chosen in accordance with the function
+$\Lambda$. We denote the successive times of $\{V_t\}$ returning to the
+set $\mathcal C$ as $0 = K_0  < K_1 < K_2 < \ldots$. It can be shown that
+$\{(V_{K_{m+1}}, \sum_{i=K_m}^{K_{m+1} - 1} \1{|V_i| > x})\}_{m \geq 0}$
+is a positive Harris recurrent Markov chain for all $x \geq 0$.
+Let $N_x = \sum_{i=0}^{K_1 - 1} \1{|V_i| > x}$. We show by the law of large
+numbers for Markov chains
+\[
+\P(|V| > x) = \pi(\mathcal C) \E_\eta(N_x)
+\]
+where $\eta$ is $\pi$ restricted to the set $\mathcal C$, i.e.
+$\forall S \subseteq \mathcal C, \eta(S) = \pi(S)/\pi(\mathcal C)$.
+$\E_\eta$ means the expectation is taken only on condition
+$V_0 \sim \eta$. Finally by \eqref{eq:pkjrkpo},
+\[
+\P(|V| > x) = \pi(\mathcal C) \E_\eta^{\mathcal D} \left[
+  N_x |A_{T_x} \cdots A_1 V_0|^{-\alpha}
+  {r_\alpha(Y_0) \over r_\alpha(Y_{T_x})}
+  \1{T_x < K_1}
+  \right]
+  =
+  \E_\eta^{\alpha} \mathcal E_x
+\]
+where $\E_\eta^{\mathcal D}$, $\mathcal D$ for ``dual'', is to remind us
+that the expectation is taken with respect to the shifted transition kernel
+as given by \eqref{eq:pkjrkpo}, and then with the original transition kernel.
+Here
+\[
+\mathcal E_x
+=
+\pi(\mathcal C)
+N_x |A_{T_x} \cdots A_1 V_0|^{-\alpha}
+{r_\alpha(Y_0) \over r_\alpha(Y_{T_x})}
+\1{T_x < K_1}
+\]
+is our importance sampling estimator. In \S\ref{sec:efficiency} we show
+\[
+\limsup_{x \to \infty} {
+  \E_\eta^{\mathcal D} \mathcal E_x^2
+  \over
+  [\P(|V| > x)]^2
+} < \infty
+\]
+In plain words, the estimator $\mathcal E_x$ has bounded relative error.
+
+%% We propose an efficient importance sampling estimator for the rare
+%%   event probability $\P(|V| > u)$ where $V$ is a vector following the
+%%   stationary distribution of a GARCH($p, q$) process. Recall
+%%   $V_t = A_t V_{t-1} + B$ is the matrix recurrence equation of a
+%%   GARCH($p, q$) process. We emanate the process $\{V_t\}$ from a set
+%%   $\mathcal C = \{V: |V| \leq M\}$ for some predefined positive
+%%   constant $M$, and then we introduce a dual change of measure for the
+%%   originally iid matrices $\{A_t\}$: for
+%%   $t < T_u = \min\{i > 0: |V_i| > u\}$, we exponentially tilt the
+%%   distribution of $A_t | X_{t-1}$, where $X_0 = V_0$ and
+%%   $X_t = \prod_{i=1}^t X_0/|\prod_{i=1}^t X_0|$, so that
+%%   $|A_t X_{t-1}|^\xi$ is more likely to take on large values and hence
+%%   $|V_t|$ is more likely to exceed $u$. Once the exeedance has
+%%   happened, we change the distribution of $\{A_t\}$ back to the
+%%   original and continue the process until $V_t$ returns to the set
+%%   $\mathcal C$. Along each simulated path emenating from $\mathcal C$
+%%   and ending in $\mathcal C$ we compute $N_u = \sum_{i=1}^K \1{|V_i| >
+%%     u}$, where $K = \min\{i > 0: |V_i| \leq M\}$.
+
+%%   The pursuit estimate of $P(|V|>u)$ is then a weighted average of
+%%   $N_u$ computed along each path. The weight depends on the path.
+
+\subsection{Eigenvalues of the sample covariance matrix of a   
+  stochastic volatility model}
+In \S\ref{sec:SV:intro} we have introduced stochastic volatility
+models as well as their connection to realized quadratic
+variation and their improved predictive power derived thereby. But
+there we have only discussed univariate models. In fact, the
+generalization to multivariate models is rather straightforward. In
+chapter \ref{ch:bernoulli} we adopt the following multivariate model:
+\begin{eqnarray*}
+  X_{i,t} &=& \sigma_{i,t} Z_{i,t} \quad
+  1 \leq i \leq p,\;
+  t \in \integers, \\
+  \log \sigma_{i, t}
+  &=&
+  \sum_{k,l \in \integers} \psi_{k,l} \eta_{i-k, t-l}\quad
+  1 \leq i \leq p,\;
+  t \in \integers,
+\end{eqnarray*}
+where $\{Z_{i,t}\}$ and $\{\eta_{i,t}\}$ are iid fields of random
+variables. The distribution of $\eta$, a generic element of
+$\{\eta_{i,t}\}$, satisfies $\P(e^\eta > x) \sim x^{-\alpha} L(x)$,
+where $\alpha > 0$ and $L(x)$ is a slowly varying function. The
+coefficients $\{\psi_{k,l}\}$ are real and satisfy
+$\sum_{k, l \in \integers} |\psi_{k,l}| < \infty$.
+
+Depending on the tails of $\{\sigma_{i,t}\}$ and $\{Z_{i,t}\}$,
+two situations can arise. When $\{Z_{i,t}\}_{t \in \integers}$
+is a regularly varying sequence with index $\alpha \in (0, 4)$
+and dominates the tail, we show that each of the sequence
+$\{X_{i,t}\}_{t \in \integers}, 1 \leq i \leq p$ and each of
+$\{X_{i,t} X_{j,t}\}_{t \in \integers}, 1\leq i,j \leq p$
+are regularly varying with index $\alpha$, assuming suitable
+conditions on $\{\sigma_{i,t}\}$.
+
+Define the matrix $\mtx X$
+\[
+\mtx X = \{X_{i,t}\}_{1 \leq i \leq p,\;1 \leq t \leq n}
+\]
+and let $\mtx X'$ denote the transpose of $\mtx X$. Then
+\[
+\mtx{XX'} =
+\left\{\sum_{t=1}^n X_{i,t} X_{j,t}\right\}_{1 \leq i,j \leq p}.
+\]
+Using theorem \ref{thm:mikwin:intro} of Mikosch and Wintenberger
+\cite{mikosch:wintenberger:2016}
+(see theorem \ref{thm:mikwin:intro} of this thesis), we prove 
+\[
+a_n^{-2} \left(
+\sum_{t=1}^n X_{i,t}^2 - n \I_{(2,4)}(\alpha) \E X^2
+\right)_{1 \leq i \leq p}
+\overset{d}{\to}
+(\xi_{i, \alpha/2})_{1 \leq i \leq p}
+\]
+where $\{a_n\}_{n\geq 1}$ is such that $n\P(|X| > a_n) \to 1$
+as $n \to \infty$,
+and $\{\xi_{i, \alpha/2}\}_{1 \leq i \leq p}$ is an iid sequence of
+$\alpha/2$-stable random variable.
+See chapter \ref{ch:bernoulli} for details.
+Built on this result, we show that $\mtx{XX'}$
+is approximated by its diagonal:
+\begin{equation}
+  \label{eq:tyi7k}
+  a_n^{-2} \|\mtx{X X'} - \diag(\mtx{X X'}) \|
+  \overset{P}{\to} 0.
+\end{equation}
+where $\|\cdot\|$ denotes the spectral norm. Following
+\eqref{eq:tyi7k}, we have
+\[
+a_n^{-2} (\lambda_{(1)}, \ldots, \lambda_{(p)})
+\overset{d}{\to}
+(\xi_{(1), \alpha/2}, \dots, \xi_{(p), \alpha/2})
+\]
+where $\lambda_{(i)}$ is the $i$-th upper order statistic of the
+eigenvalues of the matrix $\mtx{XX'}$ and $\xi_{(i), \alpha/2}$ is
+the $i$-th upper order statistic of the iid sequence
+$\{\xi_{i, \alpha/2}\}_{1 \leq i \leq p}$.
+
+When $\sigma_{i,t}$ dominates the tail of $X_{i,t} = \sigma_{i,t} Z_{i,t}$
+and satisfies a few more
+technical conditions, we show that each of the sequences
+$\{\sigma_{i,t}\}_{t \in \integers,\; 1 \leq i \leq p}$
+is regularly varying with index $\alpha$. Moreover, in contrast to the
+previous case, we show that the sequence
+$\{\sigma_{i,t} \sigma_{j,t}\}_{t \in \integers, 1 \leq i,j \leq p}$
+is regularly varying with index $\alpha/\psi^{i,j}$, where
+$\psi^{i,j} = \max_{k,l}(\psi_{k,l} + \psi_{k+i-j, l})$.
+For $d \geq 1$, the $d$-variate sequence
+$\{(\sigma_{i_k, t} \sigma_{j_k, t})_{1 \leq k \leq d}\}_{t \in \integers}$
+is regularly varying with index
+$\alpha /(\max_{1 \leq k \leq d} \psi^{i_k, j_k})$.
+
+This result then allows us to approximate the matrix of $\mtx{XX'}$
+by
+$\{\tilde X_{i,j}\}_{1 \leq i,j \leq p}$, where
+\[
+\tilde X_{i,j} = \sum_{t=1}^n X_{i,t} X_{j,t}
+\I_{\{1 \leq i,j \leq p, \psi^{i,j} = 2\}}
+\]
+A notable difference from the previous case is that the matrix
+$a_{n}^{-2}\{\tilde X_{i,j}\}_{1 \leq i,j \leq p}$ can have non-vanishing
+values on its off-diagonal entries in the limit
+$n \to \infty$, implying its eigenvalues in this limit may not be
+solely determined by its diagonal entries.
+
+\subsection{Extreme value analysis for the sample autocovariance matrices of time series}
+Jan\ss en et al. \cite{janssen:mikosch:rezapour:xie:2016} investigated
+the sample covariance matrix
+$\mathbf X = \{\sum_{t=1}^n X_{i,t} X_{j,t}\}_{1 \leq i,j \leq p}$ of
+multivariate {\em stochastic volatility models} assuming the dimension
+of the matrix $p$ is fixed. It is also of interest to look into the
+sample covariance and the sample auto-covariance matrix when the dimension
+$p$ tends to infinity at some rate as the number of observations $n$ tends
+to infinity. This is the subject of Davis et al.
+\cite{davis:mikosch:heiny:xie:2015} which we summarize in this section.
+
+We are interested in the model
+\begin{eqnarray*}
+  X_{i,t} &=& \sum_{k,l \in \integers} h_{k,l} Z_{i-k, t-l},
+  \quad i,t \in \integers \label{eq:tyhjuj}
+\end{eqnarray*}
+where $\{Z_{i,t}\}_{i,t \in \integers}$ is a field of iid random
+variables and $h_{k,l}$ is an array of real coefficients. We
+assume $\{Z_{i,t}\}$ is regularly varying with index $\alpha$
+and
+\[
+\sum_{k,l \in \integers} |h_{k,l}|^\delta < \infty
+\]
+for some $\delta \in (0, \min\{\alpha/2, 1\})$. This conditions ensures
+that the infinite series in \eqref{eq:tyhjuj} is almost surely
+absolutely convergent. Since each $\{X_{i,t}\}_{t \in \integers}$
+is a linear combination of the sequences $\{Z_{i,t}\}_{t \in \integers}$
+that are regularly varying with index $\alpha$, each of
+$\{X_{i,t}\}_{t \in \integers}$ is regularly varying with index $\alpha$.
+
+Define matrices
+\[
+\mathbf X(s) = \left\{X_{i, t+s}\right\}_{1 \leq i \leq p;\; 1 \leq t \leq n}
+\quad
+s \geq 0;\;
+{p \over n^\beta \ell(n)} \to \infty, n \to \infty
+\]
+where $\beta \geq 0$ and $\ell$ is a slowly varying function.
+If $\beta = 0$, we assume in addition $\ell(n) \to \infty, n\to\infty$.
+We intend to understand the behavior of the singular values of the
+matrix $\mathbf X(0) \mathbf X(s)$, i.e. the eigenvalues of
+$\mathbf X(0) \mathbf X(s) \mathbf X'(s) \mathbf X'(0)$.
+
+Corresponding to the time-lagged matrices $\mathbf X(s)$, we also introduce
+the time-lagged coefficient matrices. Define
+\[
+\mathbf H(s) = \{h_{k, l+s}\}_{k, l \in \integers}
+\]
+and denote by $v_1(s) \geq v_2(s) \geq \cdots$ the singular values of
+the matrix
+$\mathbf H(0) \mathbf H(s) = \{\sum_{l \in \integers} h_{i,l} h_{j,l+s}\}_{i,j \in \integers}$.
+We have established the following point process convergence result
+upon appropriate conditions (see theorem \ref{cor:1} of chapter
+\ref{ch:extremes}):
+\[
+\sum_{i=1}^p \varepsilon_{a_{np}^{-2}(\lambda_{(i)}(0), \ldots, \lambda_{(i)}(s))}
+\overset{v}{\to}
+\sum_{i=1}^\infty \sum_{j=1}^\infty
+\varepsilon_{\Gamma_i^{-2/\alpha}(v_j(0), \ldots, v_j(s))}.
+\]
+Here $\varepsilon_{x}$ is the Dirac measure with unit mass at $x$ and
+$\Gamma_i = \sum_{k=1}^i E_k$, where $\{E_k\}_{1 \leq k \leq i}$ is
+an iid sequence of Exp($1$) random variables. In particular, we see
+$a_{np}^{-2}\lambda_{(1)}(0) \overset{d}{\to} E_1^{-2/\alpha} v_1(0)$, i.e.
+asymptotically $a_{np}^{-2}\lambda_{(1)}(0)$ has scaled
+Fréchet($\alpha/2$) distribution.
+
+%% We  provide some  asymptotic theory for the largest eigenvalues of a
+%% sample covariance matrix of a $p$-dimensional \ts\ where the dimension
+%% $p=p_n$ converges to infinity when the sample size $n$ increases. 
+%% We give a short overview of the literature on the topic both in the
+%% light- and heavy-tailed cases when the data have finite (infinite)
+%% fourth moment, respectively. Our main focus is on the heavy-tailed
+%% case. In this case, one has a theory for the \pp\ of the normalized
+%% eigenvalues of the sample covariance matrix in the iid case but also
+%% when rows and columns of the data are linearly dependent. We provide
+%% limit results for the weak \con\ of these \pp es to Poisson or cluster
+%% Poisson processes. Based on this \con\ we can also derive the limit
+%% laws of various \fct als of the ordered eigenvalues such as the 
+%% joint \con\ of a finite number of the largest order statistics, the
+%% joint limit law of the largest eigenvalue and the trace, limit laws
+%% for successive ratios of ordered eigenvalues, etc. We also develop
+%% some limit theory for the singular values of the sample autocovariance
+%% matrices and their sums of squares. The theory is illustrated for
+%% simulated data and for the components of the S\&P 500 stock index.

### 2017-08-21 phd/MyThesis.tex
Revised my phd thesis.
Revised the phd thesis.

@@
 \usepackage{appendix}
+\usepackage{multirow}
 \graphicspath{
@@
 \hspace*{1.5cm} Universitetsparken 5\\
-\hspace*{1.5cm} 2100 København Ø\\
+\hspace*{1.5cm} 2100 Copenhagen\\
 \hspace*{1.5cm} Denmark\\
@@
 & \\
-
-Assessment committee:  & Lecturer.~Jesper Lund Pedersen (chairman)\\
+Co-supervisor: & Prof.~Rolf Poulsen\\
+& University of Copenhagen\\
+& \\
+Assessment committee:  & Associate Prof.~Jesper Lund Pedersen (chairman)\\
 & University of Copenhagen\\
@@
 &  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
-& University of Neuchatel, CH \\
+& University of Neuchatel\\
 & \\
@@
 \noindent \hspace*{0.5cm} Copyright:\\
-\hspace*{0.5cm} Chapters 1-4: \copyright Johannes Heiny (according to Danish legislation)\\
-\hspace*{0.5cm} Chapter 2: \copyright Elsevier\\
+\hspace*{0.5cm} Chapters 1-2: \copyright Xiaolei Xie (according to Danish legislation)\\
+\hspace*{0.5cm} Chapter 3: \copyright Bernoulli\\
+\hspace*{0.5cm} Chapter 4: \copyright Elsevier\\
 
-\noindent \hspace*{0.5cm} ISBN: $978$-$87$-$7078$-$921$-$9$
+% \noindent \hspace*{0.5cm} ISBN: $978$-$87$-$7078$-$921$-$9$
 
@@
 
-This thesis is about heavy-tailed time series.
+This thesis is about analysis of heavy-tailed time series. We discuss tail
+properties of real-world equity return series and investigate the
+possibility that a single tail index is shared by all return series of
+actively traded equities in a market. Conditions for this hypothesis to
+be true are identified.
+
+We study the eigenvalues and eigenvectors of sample covariance and
+sample auto-covariance matrices of multivariate heavy-tailed time series,
+and particularly for time series with very high dimensions. Asymptotic
+approximations of the eigenvalues and eigenvectors of such matrices
+are found and expressed in terms of the parameters of the dependence
+structure, among others.
+
+Furthermore, we study an importance sampling method for estimating
+rare-event probabilities of multivariate heavy-tailed time series
+generated by matrix recursion. We show that the proposed algorithm is
+efficient in the sense that its relative error remains bounded as the
+probability of interest tends to 0. We make use of exponential
+twisting of the transition kernel of an {\em Markov additive process},
+and take advantage of asymptotic theories on products of positive
+random matrices.
 
 \section*{Resum{\'e}}
+Denne afhandling handler om analyse af tunge tailed tidsserier. Vi
+diskuterer halen Egenskaber af real-world egenkapital retur serie og
+undersøge mulighed for at et enkelt haleindeks deles af alle
+returserier af aktivt handlede aktier i et marked. Betingelser for
+denne hypotese til Være sande er identificeret.
 
+Vi studerer egenværdier og egenvektorer af observerede kovarians og
+observerede auto-kovarians matricer af multivariate heavy-tailed
+tidsserier, og især til tidsserier med meget høje dimensioner.
+Asymptotisk tilnærmelser af egenværdierne og egenvektorer af
+sådanne matricer findes og udtrykkes i afhængighedens parametre
+struktur, blandt andre.
+
+Vi studerer også en importance samplings metod til estimering af
+sjældne begivenheder sandsynligheder for multivariate heavy-tailed
+tidsserier genereret af matrix recursion. Vi viser, at den foreslåede
+algoritme er effektiv i den forstand, at dens relative fejl forbliver
+afgrænset som sandsynligheden af interesse har tendens til 0. Vi gør
+brug af eksponentiel vridning af overgangskernen til en {\em Markov
+additive process} og tage fordel af asymptotiske teorier om produkter
+af positiv tilfældig matricer.
 \newpage
@@
 \addcontentsline{toc}{section}{Acknowledgments}
+I owe my gratitude to many people and especially to my supervisors
+Prof. Thomas Mikosch and Prof. Jeffrey Collamore. Their insight and
+guidance are indispensible throughout the scientific endeavor that
+culminates in this thesis. Meanwhile, I cannot be more grateful to
+Prof. Mogens Steffensen -- without his financial
+support from the HIPERFIT program, my PhD project would not have
+been feasible.
 
-I owe my gratitude to many people and especially to my supervisors Prof. Mikosch and Prof. Collamore.
+I also feel deeply indebted to my co-authors Prof. Richard A. Davis,
+Prof. Casper de Vries, Prof. Anja Jan\ss en and Dr. Johannes Heiny.
+Collaboration with them is not only fruitful but enjoyable. Those moments
+of relaxed discussions are to be remembered forever. Dr. Johannes Heiny
+was also my office-mate. I truly enjoyed those casual conversations
+with him about maths and life in general.
+
+To the professors and PhD students in the Mathematics department of
+Copenhagen University and especially to Prof. Jostein Paulsen, I owe a
+big ``thank you'' too. I really enjoyed working with him as a teaching
+assistant and at the same time learned a lot of fantastic maths from him
+and his book manuscript.
+
+This is also the opportunity for me to thank Dr. Dan Cheng and
+Dr. Xuan Leng, whom I first met at EPFL Lausanne and later at TU Delft.
+Discussions with them have been greatly helpful and inspiring to me.
+
+Last but not least, I would like to thank the PhD committee for spending
+their precious time to review and assess this thesis.
 \bigskip
 
-{\em Xie Xiaolei\qquad \qquad \qquad \qquad \qquad Copenhagen, September $2017$.}
+{\em Xiaolei Xie\qquad \qquad \qquad \qquad \qquad Copenhagen, September $2017$.}
 
@@
 \begin{enumerate}
-\bibitem[P2]{davis:mikosch:heiny:xie:20151}
+%% \bibitem{mikosch:deVries:xie:2016}
+%% {\sc de Vries, C. and Mikosch, T. and Xie, X.}
+%% \newblock Do return series have power-law tails with the same
+%%   index?
+%% \newblock working paper.
+
+%% \bibitem{collamore:vidyashankar:xie:2017}
+%% {\sc Collamore, Jeffrey and Vidyashankar, Anand and Xie, Xiaolei}
+%% \newblock Rare event simulation for GARCH(p,q) processes.
+%% \newblock working paper.
+
+%% \bibitem{janssen:mikosch:rezapour:xie:2016}
+%% {\sc Janssen, A. and Mikosch, T. and Mohsen, R. and X., Xie}
+%% \newblock The eigenvalues of the sample covariance matrix of a multivariate heavy-tailed stochastic volatility model.
+%% \newblock {\em Bernoulli Journal}, To appear
+
+%% \bibitem[P4]{davis:mikosch:heiny:xie:2015}
+%% {\sc Davis, R.~A., Heiny, J., Mikosch, T., and Xie, X.}
+%% \newblock Extreme value analysis for the sample autocovariance matrices of heavy-tailed multivariate time series.
+%% \newblock {\em Extremes 19}, 3 (2016), 517--547. \href{http://link.springer.com/article/10.1007/s10687-016-0251-7}
+\bibitem[P1]{mikosch:deVries:xie:2016}
+{\sc Mikosch, T., de Vries, C. and Xie, X.}
+\newblock Do return series have power-law tails with the same index?
+\newblock {\em Technical Report}
+
+\bibitem[P2]{collamore:vidyashankar:xie:2017}
+{\sc Collamore, J., Vidyashankar, A. and Xie, X.}
+\newblock Rare event simulation for GARCH(p,q) processes.
+\newblock {\em Technical Report}.
+
+\bibitem[P3]{janssen:mikosch:rezapour:xie:2016:intro}
+{\sc Janssen, A., Mikosch, T., Mohsen, R. and Xie, X.}
+\newblock Limit theory for the singular values of the sample autocovariance
+  matrix function of multivariate time series.
+\newblock {\em Bernoulli, to appear}.
+
+\bibitem[P4]{davis:mikosch:heiny:xie:20151}
 {\sc Davis, R.~A., Heiny, J., Mikosch, T., and Xie, X.}
@@
 
-Now we explain the structure of this thesis.
-
-{\em Xie Xiaolei \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}
+%% {\em Xiaolei Xie \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}
 

### 2017-08-21 papers/Anja/PaperBody.tex
Updated the thesis manuscript.

@@
 Similarly, for $i\ne j$, $h\ge 1$,
+\begin{small}
 \beao
@@
 \eeao
+\end{small}
 In the last step we again used Markov's inequality, the fact that $Z_{i0}Z_{j0}$ is \regvary\ with index $\alpha$
@@
 processes:
+\begin{small}
 \beam\label{eq:van2}
@@
 \eeam
+\end{small}
 We have 
+\begin{small}
 \beao\lefteqn{
@@
 \eeao
+\end{small}
 The \seq\ $(\sigma_{it}^2)$ satisfies the \clt\ with normalization
@@
 Thus it suffices for \eqref{eq:van2} to prove
+\begin{small}
 \beao
@@
 \eeao
+\end{small}
 The summands are independent and centered, conditional on the
@@
 \beao
-\limsup_{\nto}\P\big(\|\bfV- V_{L_j}\bfe_{L_j}\|_{\ell_2}>\delta\big)&\le &
+&&\limsup_{\nto}\P\big(\|\bfV- V_{L_j}\bfe_{L_j}\|_{\ell_2}>\delta\big) \\
+&\le &
 \limsup_{\nto}\P\big(\{\|\bfV- V_{L_j}\bfe_{L_j}\|_{\ell_2}>\delta\}\cap \Omega_n\big)+
@@
 and write
-\begin{align} \nonumber \boldsymbol{\sigma}&= \mbox{diag}\left(\left(
+\begin{equation}
+  \label{Eq:factors:sigma}
+  \boldsymbol\sigma = \mathbf A\, \mathbf Z
+\end{equation}
+where
+\begin{eqnarray*}
+  \mathbf A
+  &=&
+  \mbox{diag}\left(\left(
       \begin{matrix}
-  \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c}\ex^{\eta_{m,n}(\psi_{i_1-m,t_1-n}+\psi_{j_1-m,t_1-n})} \\ \vdots \\ \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c}\ex^{\eta_{m,n}(\psi_{i_d-m,t_d-n}+\psi_{j_d-m,t_d-n})}
+        \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c}
+        \ex^{\eta_{m,n}(\psi_{i_1-m,t_1-n}+\psi_{j_1-m,t_1-n})} \\
+        \vdots \\
+        \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}^c}
+        \ex^{\eta_{m,n}(\psi_{i_d-m,t_d-n}+\psi_{j_d-m,t_d-n})}
   \end{matrix}
-\right)'\right)\cdot \left(\begin{matrix}
-                   \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}}\ex^{\eta_{m,n}(\psi_{i_1-m,t_1-n}+\psi_{j_1-m,t_1-n})} \\ \vdots \\ \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}}\ex^{\eta_{m,n}(\psi_{i_d-m,t_d-n}+\psi_{j_d-m,t_d-n})}
+\right)'\right), \\
+\mathbf Z
+&=&
+\left(
+  \begin{matrix}
+    \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}}
+    \ex^{\eta_{m,n}(\psi_{i_1-m,t_1-n}+\psi_{j_1-m,t_1-n})} \\
+    \vdots \\
+    \prod\limits_{(m,n) \in \Lambda_{\mathbf{i},\mathbf{j},\mathbf{t}}}
+    \ex^{\eta_{m,n}(\psi_{i_d-m,t_d-n}+\psi_{j_d-m,t_d-n})}
   \end{matrix}
- \right)\\
-\label{Eq:factors:sigma} &=: \mathbf{A}\, \mathbf{Z},
-\end{align}
+\right).
+\end{eqnarray*}
 where ${\rm diag}((a_1,\ldots,a_k))$ is any diagonal matrix with diagonal elements $a_1,\ldots,a_k$. 
@@
 If \eqref{Eq:conv:eq:eta} holds, then for any Borel set $B=B_0\times\cdots \times B_n\subset \bbr^{n+1}$,
-\begin{align}\label{Eq:theta:law:1}
-  \P((\Theta^{i}_t)_{t=0,\ldots,n}
-  \in B)=\sum_{(u,v) \in \Lambda_i^{(0)}}\frac{1}{|\Lambda_i^{(0)}
+\begin{eqnarray}
+  && \P((\Theta^{i}_t)_{t=0,\ldots,n} \in B) \label{Eq:theta:law:1} \\
+  &=&
+  \sum_{(u,v) \in \Lambda_i^{(0)}}\frac{1}{|\Lambda_i^{(0)}
     |}\frac{\E\left[\I\left(\left(\I((u,v) \in
@@
       \right)|X_{i0}|^{\alpha}\right]}{\E[|X_{i0}|^{\alpha}]},
-\end{align}
+  \nonumber
+\end{eqnarray}
 where $\Lambda_i^{(t)}=\{(u,v):\psi_{i-u,t-v}=1 \}, t=0, \ldots, n$. 
@@
 If \eqref{Eq:conv:eq:eta} holds, then for any Borel set $B=B_0\times\cdots \times B_n\subset \bbr^{n+1}$, 
-\begin{align}  
-&\P(( \Theta^{ij}_t)_{t=0,\ldots,n} \in B) \nonumber \\
- \label{Eq:theta:law:2}=&\sum_{(u,v) \in
+\begin{eqnarray}
+  && \P(( \Theta^{ij}_t)_{t=0,\ldots,n} \in B) \label{Eq:theta:law:2} \\
+  &=&
+  \sum_{(u,v) \in
    \Lambda_{i,j}^{(0)}}\frac{1}{|\Lambda_{i,j}^{(0)}
@@
      \right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]}{\E[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}]},
-\end{align}
+ \nonumber
+\end{eqnarray}
 where $\Lambda_{i,j}^{(t)}=\{(u,v):\psi_{i-u,t-v}+\psi_{j-u,t-v}=\psi^{ij}\}, t=0, \ldots, n$. 
-%and the random variables $\ex^{\eta_{u,v}^{(u,v)}}$,$(\ex^{\eta_{l,m}^{(u,v)}})_{(l,m)\neq(u,v)}$, $(\tilde{Z}_{i,t},\tilde{Z}_{j,t})_{t=0, \ldots, n}$ have joint $\nu_{\alpha} \otimes (\otimes_{(l,m)\neq(u,v)} P^{\ex^{\eta}}) \otimes (\otimes_{t=0, \ldots, n}P^{(Z_{i},Z_{j})})$-density
-%\begin{eqnarray*}&& f^{(u,v)}(y_{u,v},(y_{l,m})_{(l,m) \neq (u,v)}, (z_{i,t},z_{j,t})_{t=0, \ldots, n})\\
-%&=&\frac{\I(\prod_{(l,m)}y_{l,m}^{\psi_{i-l,-m}+\psi_{j-l,-m}}|z_{i,0}z_{j,0}|>1)}{\E[(\prod_{(l,m)\neq(u,v)}\ex^{\eta_{l,m}^{(u,v)}(\psi_{i-l,-m}+\psi_{j-l,-m})}|\tilde{Z}_{i,0}\tilde{Z}_{j,0}|)^{\alpha/\psi^{ij}}]},
-%\end{eqnarray*}
 \item
@@
 \begin{eqnarray}\nonumber \mu_n^{\bfX^{ij}}(B)&=& c\,\E\left[\mu_n^{\boldsymbol{\sigma}^{ij}}(\times_{t=0}^n (B_t/(Z_{it}Z_{jt})))\right] \\
-&=& \label{Eq:mu:X} c\,\E\left[\tilde{\mu}_n^{\boldsymbol{\sigma}^{ij}}\left(\times_{t=0}^n \left(B_t \bigg/ \left(Z_{it}Z_{jt}\prod_{(u,v) \in \Lambda_{i,j,n}^c}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}\right)\right)\right)\right]
+&=& \label{Eq:mu:X} c\,\E\left[\tilde{\mu}_n^{\boldsymbol{\sigma}^{ij}}\left(\times_{t=0}^n
+    \left( {
+      B_t Z_{it}^{-1} Z_{jt}^{-1}
+      \over
+      \prod_{(u,v) \in \Lambda_{i,j,n}^c}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}
+  }\right)
+\right)\right] 
 \end{eqnarray}
@@
 \begin{eqnarray*}
- \tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n(B)&=& c \sum_{s=0}^n \int_0^\infty \P\bigg(\bigg(\I(\psi_{i-u,t-v}+\psi_{j-u,t-v}=\psi^{ij} \, \forall\, (u,v) \in \Lambda_{i,j}^{(s)})z^{\psi^{ij}}\\
- && \hspace{1.7cm} \prod_{(u,v) \in \Lambda_{i,j,n} \setminus \Lambda_{i,j}^{(s)}}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}\bigg)_{0 \leq t \leq n} \in B\bigg)\nu_{\alpha}(dz)\\
- &=& c \sum_{s=0}^n \int_0^\infty \P\bigg(\bigg(\I(t=s)z^{\psi^{ij}}\prod_{(u,v) \in \Lambda_{i,j,n} \setminus \Lambda_{i,j}^{(s)}}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}\bigg)_{0 \leq t \leq n} \in B\bigg)\nu_{\alpha}(dz),
+ && \tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n(B) \\
+ &=& c \sum_{s=0}^n \int_0^\infty \P\bigg[\bigg(\I(\psi_{i-u,t-v}+\psi_{j-u,t-v}=\psi^{ij} \, \forall\, (u,v) \in \Lambda_{i,j}^{(s)})z^{\psi^{ij}}\\
+ && \prod_{(u,v) \in \Lambda_{i,j,n} \setminus \Lambda_{i,j}^{(s)}}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}\bigg)_{0 \leq t \leq n} \in B\bigg]\nu_{\alpha}(dz)\\
+ &=& c \sum_{s=0}^n \int_0^\infty \P\bigg[\bigg(\I(t=s)z^{\psi^{ij}} \\
+ && \prod_{(u,v) \in \Lambda_{i,j,n} \setminus \Lambda_{i,j}^{(s)}}\ex^{\eta_{u,v}(\psi_{i-u,t-v}+\psi_{j-u,t-v})}\bigg)_{0 \leq t \leq n} \in B\bigg]\nu_{\alpha}(dz),
 \end{eqnarray*}
@@
 the vague limit measure $\tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n$ is up to a constant given by 
-\begin{eqnarray*} && \tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n(B) \\
-&=&\hspace{-0.3cm} \sum_{(u,v) \in \Lambda_{i,j,n}}\int_0^\infty \P\bigg(\bigg(\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}\prod_{\substack{(\tilde{u},\tilde{v}) \in \Lambda_{i,j,n} \\ (\tilde{u},\tilde{v}) \neq (u,v)}}\ex^{(\psi_{i-\tilde{u},t-\tilde{v}}+\psi_{j-\tilde{u},t-\tilde{v}})\eta_{\tilde{u},\tilde{v}}}\bigg)_{t=0, \ldots ,n} \in B \bigg)\nu_{\alpha}(dz).\end{eqnarray*}
+\begin{eqnarray*}
+  && \tilde{\mu}^{\boldsymbol{\sigma}^{ij}}_n(B) \\
+  &=& \sum_{(u,v) \in \Lambda_{i,j,n}}\int_0^\infty \P\bigg[\bigg(\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}} \\
+  && \prod_{\substack{(\tilde{u},\tilde{v}) \in \Lambda_{i,j,n} \\ (\tilde{u},\tilde{v}) \neq (u,v)}}\ex^{(\psi_{i-\tilde{u},t-\tilde{v}}+\psi_{j-\tilde{u},t-\tilde{v}})\eta_{\tilde{u},\tilde{v}}}\bigg)_{t=0, \ldots ,n} \in B \bigg]\nu_{\alpha}(dz).
+\end{eqnarray*}                                                                                                
 For sets $B$ such that $B \cap (\{0\} \times [-\infty,\infty]^n)=\emptyset$ it suffices thereby to sum only 
@@
 For these sets we have by Breiman's lemma (cf.\ \eqref{Eq:mu:X}),
-\begin{align*} & \mu^{\bfX^{ij}}_n(B)/c\\
-= &\sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\int\limits_0^\infty \P\bigg((\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}\prod\limits_{(\tilde{u},\tilde{v}) \neq (u,v)}\ex^{(\psi_{i-\tilde{u},t-\tilde{v}}+\psi_{j-\tilde{u},t-\tilde{v}})\eta_{\tilde{u},\tilde{v}}} Z_{it}Z_{jt})_{t=0, \ldots ,n} \in B \bigg)\nu_{\alpha}(dz)\\
-= &\sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\int\limits_0^\infty \P\bigg((\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}X_{it}X_{jt}\ex^{-\psi^{ij}\eta_{u,v}})_{t=0, \ldots ,n} \in B \bigg)\nu_{\alpha}(dz),\end{align*}
+\begin{eqnarray*}
+  && \mu^{\bfX^{ij}}_n(B)/c\\
+  &=&
+  \sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\int\limits_0^\infty \P\bigg[(\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}} \\
+  &&
+  \prod\limits_{(\tilde{u},\tilde{v}) \neq (u,v)}\ex^{(\psi_{i-\tilde{u},t-\tilde{v}}+\psi_{j-\tilde{u},t-\tilde{v}})\eta_{\tilde{u},\tilde{v}}} Z_{it}Z_{jt})_{t=0, \ldots ,n} \in B \bigg]\nu_{\alpha}(dz)\\
+  &=&
+  \sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\int\limits_0^\infty \P\bigg((\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}X_{it}X_{jt}\ex^{-\psi^{ij}\eta_{u,v}})_{t=0, \ldots ,n} \in B \bigg)\nu_{\alpha}(dz),
+\end{eqnarray*}
 where we used that if $(u,v) \in \Lambda_{i,j}^{(t)},$ then 
@@
 Fubini's Theorem and a substitution finally simplify this expression to 
-\begin{eqnarray*} &&  \sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_0^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}X_{it}X_{jt}\ex^{-\psi^{ij}\eta_{u,v}}\right)_{t=0, \ldots, n} \in B\right)\nu_\alpha(dz)\right]  \\
-&=& \sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_0^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\nu_{\frac{\alpha}{\psi^{ij}}}(dy)\right].
+\begin{eqnarray*}
+&&
+\sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_0^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})z^{\psi^{ij}}X_{it}X_{jt}\ex^{-\psi^{ij}\eta_{u,v}}\right)_{t=0, \ldots, n} \in B\right)\nu_\alpha(dz)\right]  \\
+&=&
+\sum_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\bigg[\int_0^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right) \\
+&&
+|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\nu_{\frac{\alpha}{\psi^{ij}}}(dy)\bigg].
 \end{eqnarray*}
@@
 A stochastic domination argument and an application of Breiman's lemma show that uniformly for $l\le t\le m_n$,
+\begin{small}
 $$ m_n\, n\, \P\big(|X_{it}X_{jt}|>b_n\,,|X_{i0}X_{j0}|>b_n\big)=n^{2-\varepsilon} o\left(b_n^{-2\alpha/(\psi^{ij}(1+4^{-1}\epsilon))}\right)=n^{2-\varepsilon}o(n^{-2/(1+2^{-1}\varepsilon)})=o(1)$$  
+\end{small}
 which yields \eqref{eq:ac}.\\[2mm]
@@
 &\le &c\,\big(\E \big[
-\ex^{2r\sum_{(k,l)\in \Gamma_\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]\big)^{1/r} \big(\E\big[ \ex^{2s\sum_{(k,l) \not \in \Gamma_\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]\big)^{1/s}\,,
+\ex^{2r\sum_{(k,l)\in \Gamma_\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]\big)^{1/r} \\
+&&
+\big(\E\big[ \ex^{2s\sum_{(k,l) \not \in \Gamma_\xi}\psi_{kl}(\eta_{i-k,-l}+\eta_{i-k,h-l})}\big]\big)^{1/s}\,,
 \eeao
@@
 \beam\label{eq:90}
-a_n^{-2}\Big(\la_{(i)} -n\,\E[X^2] 1(\alpha\in (2,4))\Big)_{i=1,\ldots,p}
-\std \Big(\la_{(i)}\big((\xi_{kl,\alpha/2} \I((k,l)\in \Gamma_p))_{1\le k,l\le p}\big)\Big)_{i=1,\ldots,p} \,,\nonumber\\
+&& a_n^{-2}\Big(\la_{(i)} -n\,\E[X^2] 1(\alpha\in (2,4))\Big)_{i=1,\ldots,p} \\
+&\std&
+\Big(\la_{(i)}\big((\xi_{kl,\alpha/2} \I((k,l)\in \Gamma_p))_{1\le k,l\le p}\big)\Big)_{i=1,\ldots,p} \,,\nonumber
 \eeam
@@
 For $M>0$ we have the following decomposition
+\begin{small}
 \beao
@@
 \eeao
+\end{small}
 Here we applied Breiman's result twice. The second term vanishes by virtue of \eqref{eq:hh}. Thus $q=  \E[Y^\alpha]$.\\
@@
 for any Borel set $B \in [0,\infty]^n$ bounded away from $\mathbf{0}$,
+\begin{footnotesize}
 \begin{eqnarray}
@@
 \end{eqnarray}
+\end{footnotesize}
 
@@
 (i)~Let $B \in [0,\infty]^n$ be a Borel set bounded away from $\mathbf{0}$. For $s>0$ we have
+\begin{small}
 \begin{eqnarray}
@@
 \end{eqnarray}
+\end{small}
 Since $B$ is bounded away from $\mathbf{0}$, there exists $v>0$ and $1 \leq i \leq n$ such that $B \subset \{(x_1, \ldots, x_n) \in [0,\infty]^n: x_i>v\}$.
@@
 Substituting $sy$ by $z$ in the integral finally gives
+\begin{small}
 $$ \lim_{s \searrow 0}\lim_{t \to \infty} \frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}=\int_0^\infty \P\left(\left(\I(a_{ij}=a_{\max})z^{a_{\max}}\prod_{k \neq j}Y_k^{a_{ik}}\right)_{1 \leq i \leq n} \in B\right)\nu_{\alpha}(dz).
 $$
+\end{small}
 (ii) Note first that under our assumptions for any $1\leq n_1 < n_2 \leq p$,

### 2017-08-21 papers/DavisMikoschHeinyXie/PaperBody.tex
Updated the thesis manuscript.

@@
 \end{equation*}
-An application of
-Proposition~\ref{prop:sumsmal} yields that the ordered eigenvalues of $a_{np}^{-4}\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'$
+An application of Proposition~\ref{prop:sumsmal} yields that the ordered eigenvalues of
+\[
+a_{np}^{-4}\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'
+\]
+
 are uniformly approximated by the quantities
@@
 \beao
-\sum_{i=1}^p \varepsilon_{a_{np}^{-4}\Big(\la_{i}\Big(\sum_{s=s_0}^{s_0}\A_n(s)\A_n(s)'\big)\,,\ldots,\la_{i}\big(\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'\big)\Big)}
-\std \sum_{i=1}^\infty \varepsilon_{\Gamma_i^{-4/\alpha} \Big(C_{c,\theta}(s_0,s_0),\ldots,C_{c,\theta}(s_0,s_1)\Big)}\,.
+&& \sum_{i=1}^p \varepsilon_{a_{np}^{-4}\Big(\la_{i}\Big(\sum_{s=s_0}^{s_0}\A_n(s)\A_n(s)'\big)\,,\ldots,\la_{i}\big(\sum_{s=s_0}^{s_1}\A_n(s)\A_n(s)'\big)\Big)} \\
+&\std&
+\sum_{i=1}^\infty \varepsilon_{\Gamma_i^{-4/\alpha} \Big(C_{c,\theta}(s_0,s_0),\ldots,C_{c,\theta}(s_0,s_1)\Big)}\,.
 \eeao

### 2017-08-21 phd/Chapterintroduction.tex
Updated the thesis manuscript.

@@
 \chapter{Introduction}\label{ch:intr}
-The phenomena of heavy-tailedness is widely observed in many
+The phenomenon of heavy-tailedness is widely observed in many
 disciplines of science, for example, phase transition of matter and
@@
 \P(X < -x) \sim p_- {L(x) \over x^\alpha}, \quad
-\text{for some } p_+, p_- > 0,\; p_+ + p_- = 1
+\text{for some } p_+, p_- \geq 0,\; p_+ + p_- = 1
 \]
@@
 $\hat \alpha_H$, under the assumptions $k \to \infty, k/n \to 0$
-as $n \to \infty$. See theorem 6.4.6 of Embrechts et. al.
+as $n \to \infty$. See Theorem 6.4.6 of Embrechts et. al.
 \cite{embrechts:klueppelberg:mikosch:1997} for details.
@@
 Random variables with regularly varying tails have some very nice
-features: if $X_1$ and $X_2$ are both regularly varying with indices
-$\alpha_1$ and $\alpha_2$, respectively, then $a_1 X_1 + a_2 X_2$,
+features: if $X_1$ and $X_2$ are both positive and regularly varying
+with indices $\alpha_1$ and $\alpha_2$, respectively, then $a_1 X_1 + a_2 X_2$,
 for $a_1, a_2 > 0$, is regularly varying with index
@@
 $\hat Y_{j,t} = \sum_{i=1}^p a_{j, i} X_{i,t}$, where
-$(a_{j, 1}, a_{j, 2}, \dots, a_{j, p})^\top$ is the $j$-th eigenvector
+$(a_{j, 1}, a_{j, 2}, \dots, a_{j, p})'$ is the $j$-th eigenvector
 of the sample covariance matrix of
@@
 associated with the $j$-th largest eigenvalue. 
-For this reason, it is important to understand the eigensystem of the
+For this reason, it is important to understand the eigen system of the
 sample covariance matrix of $\{X_{i,t}\}$. This topic is addressed in
@@
 extended to multivariate processes by Davis and Mikosch
-\cite{davis:mikosch:2009b}: An $\reals^d$-valued strictly stationary
+\cite{davis:mikosch:2009b}: an $\reals^d$-valued strictly stationary
 process $X_t$ is said to be regularly varying with index $\alpha$, if
@@
     \P(|X_0| > x)
-  } \overset{v}{\to} \nu_\alpha(\cdot),
+  } \overset{v}{\to} \nu^{(h)}_\alpha(\cdot),
   \quad x \to \infty
 \]
-where $\nu_\alpha$ a non-null Radon measure on
+where $\nu_\alpha^{(h)}$ a non-null Radon measure on
 $\reals^{d(h+1)} \setminus \{0\}$ that
 is homogeneous of order $-\alpha$, i.e.
-$\forall c > 0, \forall S \subset \reals^{d(h+1)}\setminus\{0\}, \nu_\alpha(c S) = c^{-\alpha} \nu_\alpha(S)$. $|\cdot|$ could be any given vector norm.
+$\forall c > 0, \forall S \subset \reals^{d(h+1)}\setminus\{0\}, \nu_\alpha^{(h)}(c S) = c^{-\alpha} \nu_\alpha^{(h)}(S)$. $|\cdot|$ could be any given vector norm.
 
@@
 a strictly stationary process: an $\reals^d$-valued strictly stationary
-process $\{X_t\}$ is regularly varying with index $\alpha$ if 
+process $\{X_t\}$ is regularly varying with index $\alpha$ if and only if
 there exists an $\reals^d$-valued sequence $\{\Theta_i\}_{i = 0,1,\ldots}$
@@
 \P\left[\left.
-  x^{-1}(X_0, X_1, \ldots) \in \cdot
+  x^{-1}(X_0, X_1, \ldots, X_h) \in \cdot
   \right| |X_0| > x
@@
   \overset{w}{\to}
-  \P(Z(\Theta_1, \Theta_2, \ldots) \in \cdot),
+  \P(Z(\Theta_1, \ldots, \Theta_h) \in \cdot),
   \quad
-  x \to \infty.
+  x \to \infty, \forall h \geq 0
 \]
@@
 \eeam
-for the same sequence $(m_n)$ as in (2).
+for the same sequence $(m_n)$ as in 2.
 \item If $\alpha\in (1,2)$, in addition $\E[Y]=0$ and the vanishing small values condition holds: 
@@
   \end{enumerate}
-\item Condition (A) : The group generated by
+\item Condition (A) : the group generated by
   \[
@@
   \label{eq:jorg}
-  u^\alpha \P(u^{-1} X \in \cdot) \overset{v}{\to} \nu_\alpha(\cdot)
+  u^\alpha \P(u^{-1} X \in \cdot) \overset{v}{\to} \nu_\alpha^{(h)}(\cdot)
 \end{equation}
-where $\nu_\alpha$ is a non-null Radon measure on
+where $\nu_\alpha^{(h)}$ is a non-null Radon measure on
 $\reals^d_+ \setminus \{0\}$ with the property
-$\forall a > 0, \nu_\alpha(a A) = a^{-\alpha} \nu_\alpha(A)$.
+$\forall a > 0, \nu_\alpha^{(h)}(a A) = a^{-\alpha} \nu_\alpha^{(h)}(A)$.
 
@@
 \cite{alsmeyer:mentemeier:2012} considered invertible matrices whose
-distribution has a density. Let $M(d, \reals)$ denote the space of
+distribution has a density. Let $M(d, \reals)$ denote a metric space of
 $d \times d$ matrices with real entries that are invertible with
@@
     $\P(A_N \cdots A_1 \in C) \geq c |B_\epsilon(\bar A) \cap C|$,
-    where $|\cdot|$ denotes the Lebesgue measure.
+    where $|\cdot|$ denotes the Lebesgue measure and
+    $B_\epsilon(\bar A)$ is the ball with radius $\epsilon$ centered
+    at $\bar A$.
 \end{enumerate}
@@
 \begin{enumerate}
-  \item There is no finite union $W$ of proper subspaces of $\reals^d$
+  \item There is no finite union $W$ of proper sub-spaces of $\reals^d$
     that satisfies $\forall a \in G_A, a W = W$.
@@
 {\em Generalized Autoregressive Conditional Heteroscedasticity}
-(GARCH) models have been hugely popular for modelling volatility of
+(GARCH) models have been hugely popular for modeling volatility of
 financial time series and have inspired numerous variants.
@@
   \label{eq:garch:intro}
-  X_t &=& \sigma_t Z_t \\
+  X_t &=& \sigma_t Z_t, \\
   \sigma_t^2 &=& \omega + \sum_{i=1}^p \alpha_i X_{t-i}^2 +
-  \sum_{j=1}^q \beta_j \sigma_{t-j}^2
+  \sum_{j=1}^q \beta_j \sigma_{t-j}^2,
 \end{eqnarray}
@@
 \eqref{eq:rhjyu}. With appropriate conditions,
-$(\sigma_t^2, \dots, \sigma_{t-q+1}^2, X_{t-1}^2, \dots, X_{t - p +1}^2)$
+\[
+(\sigma_t^2, \dots, \sigma_{t-q+1}^2, X_{t-1}^2, \dots, X_{t - p +1}^2)
+\]
 is shown to be a positive Harris recurrent Markov chain
@@
 \end{equation}
-While GARCH models have been very successful for modelling financial
+While GARCH models have been very successful for modeling financial
 time series, they do have their drawbacks. For example, the tail
@@
 With the availability of high-frequency data, a different approach
-than that of GARCH has been popularized for modelling volatility of
+than that of GARCH has been popularized for modeling volatility of
 financial time series and has led to greatly improved accuracy of
@@
 
-In the pioneering work of Clark \cite{clark:1973}, the author modelled
+In the pioneering work of Clark \cite{clark:1973}, the author modeled
 the logarithmic price $Y_t$ as a subordinated stochastic process:
-$Y_t = V_{\tau_t}$, where $V_i$ is a Brownian motion.
+$Y_t = V_{\tau_t}, t\geq 0$, where $V_i$ is a Brownian motion.
 $\tau_t$ is a real-valued, non-negative, non-decreasing
@@
 Later authors, e.g. Back \cite{back:1991} chose to model the log-price
-process as a semimartingale, with increments of the martingale
-component modelled as a product process:
-\begin{eqnarray}
-  Y_t &=& Y_0 + A_t + M_t \nonumber \\
-  M_t - M_{t-1} &=& X_t = \sigma_t Z_t
-  \label{eq:rtght}
-\end{eqnarray}
+process as a semi-martingale, with increments of the martingale
+component modeled as a product process:
+\begin{eqnarray*}
+  Y_t &=& Y_0 + A_t + M_t,
+\end{eqnarray*}
 where $\{A_t\}$ is a finite-variation process and $\{M_t\}$ is a
-martingale and hence $\{Y_t\}$ is a semimartingale.
-$\{\sigma_t\}$ is non-negative and $Z_t$ is an iid process with zero mean
-and unit variance. A convenient choice of $\sigma_t$ is
+martingale and hence $\{Y_t\}$ is a semi-martingale.
+$\{\sigma_t\}$ is non-negative and $\{Z_t\}$ is an iid process with
+zero mean and unit variance. $\{\sigma_t\}$ and $\{Z_t\}$ are
+independent of each other.
+
+For discrete time models, Taylor \cite{taylor:1982} was the first to
+propose a product process for modeling the martingale part $M_t$ of
+$Y_t$:
+\[
+M_t - M_{t-1}= \sigma_t Z_t, \quad t \in \integers
+\]
+A convenient choice of $\sigma_t$ is
 \begin{equation}
@@
 $\{\eta_t\}_{t \in \integers}$ is an iid sequence of random
-variables. By Kolmogorov's 3-series theorem, the infinite series above
+variables with zero-mean and finite variance.
+By Kolmogorov's 3-series theorem, the infinite series above
 converges if and only if $\sum_{l \in \integers} \psi_l^2 < \infty$,
@@
 tail, the conclusion follows from Breiman's lemma \cite{breiman:1965}.
-See \S\ref{sec:case1} of Jan\ss en et al. i.e. chapter
+See \S\ref{sec:case1} of Jan\ss en et
+al. \cite{janssen:mikosch:rezapour:xie:2016}, i.e. chapter
 \ref{ch:bernoulli} of this thesis for more details.
@@
 by taking advantage of high frequency data. Specifically, for the
-semimartingale $Y_t$ we have
+semi-martingale $\{Y_t\}$ we have
 \begin{equation}
@@
 The convergence in probability follows directly from the definition of
-quadratic variation.
+quadratic variation $[Y]_t$:
+\[
+[Y]_t = \lim_{\Delta t \to 0} \sum_{i=1}^{n} (Y_{t_i} - Y_{t_{i-1}})^2
+\]
+where $0 = t_0 < t_1 < t_2 < \cdots < t_n = t$ and
+$\Delta t = \sup_{1\leq i \leq n} (t_i - t_{i-1})$.
 %% When $\E Z_t = 0$, $[Y]_t = [M]_t$ 
@@
 \]
-Moreover, the structure $X_t = M_{t} - M_{t-1} = \sigma_t Z_t$ gives
+If the process of expected return has continuous sample path and $M_t$
+is a stochastic volatility process, it has been shown $[Y]_t = [M]_t$ and
 $[Y_\delta]_t \overset{P}{\to} \int_0^t \sigma_s^2 ds$. Meanwhile,
-It\^o's formula gives, for semimartingale $Y_t$,
+It\^o's formula gives, for semi-martingale $Y_t$,
 \begin{eqnarray}
@@
 available to investors. We model the investor's preference of the
-equity with {\em Generalized Disappoint Aversion (GDA)}, an idea envisaged
-by Gul \cite{gul:1991} and generalized by Routledge and Zin
+equity with {\em Generalized Disappointment Aversion (GDA)}, an idea
+envisaged by Gul \cite{gul:1991} and generalized by Routledge and Zin
 \cite{routledge2010generalized}. Specifically, in the GDA theory a
@@
 is the level of disappointment -- if $C$ falls below $\delta$, the
-investor becomes disappointed. $b$ parameterizes the growth of his
+investor becomes disappointed. $b$ parametrizes the growth of his
 disappointment. $F$ is the distribution function of the return of the
@@
 $F$ of $\E_F$ reminds us that the expectation is taken with respect to
-the distribution funciton $F$.
+the distribution function $F$.
 
@@
 \begin{eqnarray*}
-  B &=& (\omega, 0, \cdots, 0) \\
+  B &=& (\omega, 0, \cdots, 0)', \\
   V_t &=& (\sigma_t^2, \sigma_{t-1}^2, \cdots, \sigma_{t-q+1}^2,
-  X_{t-1}^2, \cdots, X_{t-p+1}^2) = A_t V_{t-1} + B\\
+  X_{t-1}^2, \cdots, X_{t-p+1}^2)' = A_t V_{t-1} + B,\\
   Y_t &=& {A_t \cdots A_1 V_0 \over |A_t \cdots A_1 V_0|}
-  = {A_t Y_{t-1} \over |A_t Y_{t-1}|} = A_t \cdot Y_{t-1}\\
-  l_t &=& \log |A_t Y_{t-1}| \\
-  S_t &=& \log |A_t \cdots A_1 V_0| - \log |V_0|
-  = \sum_{i=1}^t l_i
+  = {A_t Y_{t-1} \over |A_t Y_{t-1}|},\\
+  l_t &=& \log |A_t Y_{t-1}|, \\
+  S_t &=& \log |A_t \cdots A_1 V_0|
+  = \sum_{i=1}^t l_i + \log |V_0|,
 \end{eqnarray*}
-where $X_t = \sigma_t Z_t$ (see \eqref{eq:garch:intro}). The matrices
-$\{A_t\}_{t \geq 1}$ is defined by \eqref{eq:A_matrix:intro}.
-For more details, see \S\ref{sec:IS_Results}.
-$V_t$, our main object of interest, is a function of the path
+where $X_t = \sigma_t Z_t$ (see \eqref{eq:garch:intro}). For
+convergence we also use the notation 
+\[
+A_t \cdot Y_{t-1} := {A_t Y_{t-1} \over |A_t Y_{t-1}|}.
+\]
+The matrices $\{A_t\}_{t \geq 1}$ are defined by
+\eqref{eq:A_matrix:intro}. For more details, see
+\S\ref{sec:IS_Results}.
+$V_t$, our object of main interest, is a function of the path
 $\{(Y_i, l_i)\}_{1 \leq i \leq t}$. To increase the chance of
 observing $\{|V_t| > x\}$, we adopt a dual change of the transition
-kernle of $\{(Y_t, l_t)\}_{t \geq 0}$. Before the first occurence of
+kernel of $\{(Y_t, l_t)\}_{t \geq 0}$. Before the first occurrence of
 $\{|V_t| > x\}$ we simulate $(Y_t, l_t)$ according to a shifted
@@
     (Y_t, l_t) \in dy \times dl | Y_{t-1} = w
-    \right] \label{eq:pgrtgpd}
+    \right], \label{eq:pgrtgpd}
 \end{eqnarray}
@@
     g(Y_t, l_t) e^{-\alpha l_t} {r_\alpha(Y_{t-1}) \over r_\alpha(Y_t)}
-    \right]
+    \right].
 \end{equation}
-Let's define $T_x = \min\{t \geq 1: |V_t| > x\}$. 
+Let us define $T_x = \min\{t \geq 1: |V_t| > x\}$. 
 Once the first excursion of $|V_t|$ above $x$ has occurred, i.e.
@@
 $\Lambda$. We denote the successive times of $\{V_t\}$ returning to the
-set $\mathcal C$ as $0 = K_0  < K_1 < K_2 < \ldots$. It can be shown that
+set $\mathcal C$ as $0 = K_0  < K_1 < K_2 < \cdots$. It can be shown that
 $\{(V_{K_{m+1}}, \sum_{i=K_m}^{K_{m+1} - 1} \1{|V_i| > x})\}_{m \geq 0}$
@@
 \[
-\P(|V| > x) = \pi(\mathcal C) \E_\eta^{\mathcal D} \left[
-  N_x |A_{T_x} \cdots A_1 V_0|^{-\alpha}
-  {r_\alpha(Y_0) \over r_\alpha(Y_{T_x})}
-  \1{T_x < K_1}
-  \right]
-  =
-  \E_\eta^{\alpha} \mathcal E_x
-\]
-where $\E_\eta^{\mathcal D}$, $\mathcal D$ for ``dual'', is to remind us
-that the expectation is taken with respect to the shifted transition kernel
-as given by \eqref{eq:pkjrkpo}, and then with the original transition kernel.
-Here
+\P(|V| > x) = \pi(\mathcal C) \E_\eta N_x = \E_\eta^{\alpha} \mathcal E_x
+\]
+where
 \[
@@
 \]
-is our importance sampling estimator. In \S\ref{sec:efficiency} we show
+is our importance sampling estimator. $\E_\eta^{\mathcal D}$,
+$\mathcal D$ for ``dual'', is to remind us that the expectation is
+taken with respect to the shifted transition kernel as given by
+\eqref{eq:pkjrkpo}, and then with respect to the original transition
+kernel. In \S\ref{sec:efficiency} we show
 \[
@@
 \]
-In plain words, the estimator $\mathcal E_x$ has bounded relative error.
-
+In plain words, the estimator $\mathcal E_x$ has bounded relative
+error. See Asmussen and Glynn \cite{opac-b1123521}, \S 1 for
+definition of {\em bounded relative error}.
+The method presented in this paper is a multivariate generalization of
+Collamore et al. \cite{collamore2014}. The reader is referred to it
+for the treatment of the one-dimensional problem.
 %% We propose an efficient importance sampling estimator for the rare
@@
 where $\{Z_{i,t}\}$ and $\{\eta_{i,t}\}$ are iid fields of random
-variables. The distribution of $\eta$, a generic element of
-$\{\eta_{i,t}\}$, satisfies $\P(e^\eta > x) \sim x^{-\alpha} L(x)$,
+variables. They are independent of each other. The distribution of
+$\eta$, a generic element of $\{\eta_{i,t}\}$, satisfies
+$\P(e^\eta > x) \sim x^{-\alpha} L(x)$,
 where $\alpha > 0$ and $L(x)$ is a slowly varying function. The
@@
 is a regularly varying sequence with index $\alpha \in (0, 4)$
-and dominates the tail, we show that each of the sequence
+and dominates the tail, we show that each of the sequences
 $\{X_{i,t}\}_{t \in \integers}, 1 \leq i \leq p$ and each of
@@
 (\xi_{i, \alpha/2})_{1 \leq i \leq p}
+\]
+and
+\[
+a_n^{-2} \sum_{t=1}^n X_{i,t} X_{j,t} \overset{P}{\to} 0
+\quad
+\text{for } i \neq j.
 \]
@@
 
-\subsection{Extreme value analysis for the sample autocovariance matrices of time series}
+\subsection{Extreme value analysis for the sample auto-covariance
+  matrices of time series}
 Jan\ss en et al. \cite{janssen:mikosch:rezapour:xie:2016} investigated
-the sample covariance matrix
-$\mathbf X = \{\sum_{t=1}^n X_{i,t} X_{j,t}\}_{1 \leq i,j \leq p}$ of
-multivariate {\em stochastic volatility models} assuming the dimension
+the sample covariance matrix of the time series
+$\{X_{i,t}\}_{t \in \integers}$, $1 \leq i \leq p$ for
+{\em stochastic volatility models} assuming the dimension
 of the matrix $p$ is fixed. It is also of interest to look into the
@@
 We are interested in the model
-\begin{eqnarray*}
+\begin{eqnarray}
   X_{i,t} &=& \sum_{k,l \in \integers} h_{k,l} Z_{i-k, t-l},
   \quad i,t \in \integers \label{eq:tyhjuj}
-\end{eqnarray*}
+\end{eqnarray}
 where $\{Z_{i,t}\}_{i,t \in \integers}$ is a field of iid random
 variables and $h_{k,l}$ is an array of real coefficients. We
-assume $\{Z_{i,t}\}$ is regularly varying with index $\alpha$
-and
+assume $\{Z_{i,t}\}$ is regularly varying with index
+$\alpha \in (0, 4)$ and
 \[
@@
 \quad
-s \geq 0;\;
-{p \over n^\beta \ell(n)} \to \infty, n \to \infty
+s \geq 0,
+\]
+and assume that
+\[
+{p \over n^\beta \ell(n)} \to \infty, n \to \infty,
 \]
@@
 We intend to understand the behavior of the singular values of the
-matrix $\mathbf X(0) \mathbf X(s)$, i.e. the eigenvalues of
+matrix $\mathbf X(0) \mathbf X(s)$, i.e. the square roots of the
+eigenvalues of
 $\mathbf X(0) \mathbf X(s) \mathbf X'(s) \mathbf X'(0)$.
@@
 \]
-Here $\varepsilon_{x}$ is the Dirac measure with unit mass at $x$ and
+where ${a_{np}}_{n,p \in \integers}$ is a sequence of positive real
+numbers such that $\P(Z > a_{np}) = \ell(np) (np)^{-\alpha}$.
+$\varepsilon_{x}$ is the Dirac measure with unit mass at $x$ and
 $\Gamma_i = \sum_{k=1}^i E_k$, where $\{E_k\}_{1 \leq k \leq i}$ is
@@
 $a_{np}^{-2}\lambda_{(1)}(0) \overset{d}{\to} E_1^{-2/\alpha} v_1(0)$, i.e.
-asymptotically $a_{np}^{-2}\lambda_{(1)}(0)$ has scaled
+asymptotically $a_{np}^{-2}\lambda_{(1)}(0)$ has a scaled
 Fréchet($\alpha/2$) distribution.
+
+These results extend and generalize previous work by Soshnikov
+\cite{soshnikov:2004} and Auffinger et
+al. \cite{auffinger:arous:peche:2009} who deal with the case of iid
+$X_{i,t}$, assuming regular variation with $\alpha \in (0, 4)$. This
+work also generalizes Davis, Pfaffel and Stelzer
+\cite{davis:pfaffel:stelzer:2014} as well as Davis, Mikosch and
+Pfaffel \cite{davis:mikosch:oliver:2016} who deal with the linear
+model \eqref{eq:tyhjuj} under suboptimal conditions on the growth rate
+of $p \to \infty$.
+
+In this paper, large deviation results for sums of regularly varying
+random variables are consequently used; see Nagaev
+\cite{nagaev1979}. This is in contrast to the papers by Soshnikov
+\cite{soshnikov:2004}  and Auffinger et
+al. \cite{auffinger:arous:peche:2009} who only deal with the case when
+$p/n \to \gamma \in (0, \infty)$. The large deviation approach allows
+one to determine the dominating parts of the sample covariance matrix
+$\mathbf {XX'}$ and the sample auto-covariance matrices. Typically
+these parts are given by functionals of $\{Z_{i,t}^2\}$.
+
+In the case of a finite 4th moment of $\{X_{i,t}\}$, the theory
+changes completely. A typical result in this case is proved by
+Johnstone \cite{johnstone:2001}: if $p/n \to \gamma \in (0, \infty)$,
+then
+\[
+{\lambda_{(1)} - \mu_{np} \over \sigma_{np}}
+\overset{d}{\to}
+W_1 \sim F_1,
+\]
+where the constants $\mu_{np}$ and $\sigma_{np}$ are given by
+\begin{eqnarray*}
+  \mu_{np} &=& (\sqrt{n-1} + \sqrt p)^2, \\
+  \sigma_{np} &=& (\sqrt{n-1} + \sqrt p)
+  \left(
+    {1 \over \sqrt{n - 1}} + {1 \over \sqrt p}
+  \right)^{1/3}.
+\end{eqnarray*}
+$W_1$ is a Tracy-Widom random variable with distribution
+function $F_1$ defined by
+\[
+F_1(s) = \exp \left\{
+  -{1 \over 2} \int_s^\infty
+  [q(x) + (x - s) q^2(x)] dx
+\right\}.
+\]
+$q$ is defined as the solution to the Painlevé II differential
+equation
+\begin{eqnarray*}
+  q''(x) &=& x q(x) + 2 q^3(x), \\
+q(x) &\sim& \text{Ai}(x)
+\quad \text{as } x \to \infty.
+\end{eqnarray*}
+where $\text{Ai}(x)$ is the Airy function.
+In the heavy-tailed case considered in this thesis, asymptotic results
+about the eigenvalues are easier to derive than in the light-tailed
+case, i.e. when 4th moments are finite.
 

### 2017-08-21 phd/MyThesis.tex
Updated the thesis manuscript.

@@
 \usepackage{appendix}
+\usepackage{pstricks, pst-barcode}
 \usepackage{multirow}
@@
 & Prof.~Olivier Wintenberger\\
-& University of Marie \& Pierre Curie, Paris
+& University Marie \& Pierre Curie, Paris
 \end{tabular}
@@
 efficient in the sense that its relative error remains bounded as the
-probability of interest tends to 0. We make use of exponential
+probability of interest tends to zero. We make use of exponential
 twisting of the transition kernel of an {\em Markov additive process},
@@
 \section*{Resum{\'e}}
-Denne afhandling handler om analyse af tunge tailed tidsserier. Vi
-diskuterer halen Egenskaber af real-world egenkapital retur serie og
-undersøge mulighed for at et enkelt haleindeks deles af alle
-returserier af aktivt handlede aktier i et marked. Betingelser for
-denne hypotese til Være sande er identificeret.
+Denne afhandling handler om analyse af tidsrækker med tunge haler. Vi 
+diskuterer haleegenskaber af real-world equity return serie og
+undersøge muligheden for at et enkelt haleindeks deles af alle
+return serier af aktivt handlede aktier i et marked. Betingelser for
+denne hypotese til være sande er identificeret.
 
-Vi studerer egenværdier og egenvektorer af observerede kovarians og
-observerede auto-kovarians matricer af multivariate heavy-tailed
-tidsserier, og især til tidsserier med meget høje dimensioner.
-Asymptotisk tilnærmelser af egenværdierne og egenvektorer af
-sådanne matricer findes og udtrykkes i afhængighedens parametre
+Vi studerer egenværdier og egenvektorer af observerede kovarians- og
+auto-kovarians- matricer af multivariate tung-halede
+tidsrækker, og især til tidsrækker med meget høje dimensioner.
+Asymptotiske tilnærmelser af egenværdierne og egenvektorer af
+sådanne matricer findes og udtrykkes i afhængighedens parameter
 struktur, blandt andre.
 
-Vi studerer også en importance samplings metod til estimering af
-sjældne begivenheder sandsynligheder for multivariate heavy-tailed
-tidsserier genereret af matrix recursion. Vi viser, at den foreslåede
+Vi studerer også en importance sampling metode til estimering
+sandsynligheder af sjældne begivenheder for multivariate tung-halede
+tidsrækker genereret af matrice rekursioner. Vi viser, at den foreslåede
 algoritme er effektiv i den forstand, at dens relative fejl forbliver
-afgrænset som sandsynligheden af interesse har tendens til 0. Vi gør
+afgrænset som sandsynligheden af interesse har tendens til nul. Vi gør
 brug af eksponentiel vridning af overgangskernen til en {\em Markov
 additive process} og tage fordel af asymptotiske teorier om produkter
-af positiv tilfældig matricer.
+af positive tilfældige matricer.
 \newpage
@@
 Prof. Thomas Mikosch and Prof. Jeffrey Collamore. Their insight and
-guidance are indispensible throughout the scientific endeavor that
+guidance are indispensable throughout the scientific endeavor that
 culminates in this thesis. Meanwhile, I cannot be more grateful to
@@
 
-To the professors and PhD students in the Mathematics department of
+To the professors and PhD students in the Mathematics Department of
 Copenhagen University and especially to Prof. Jostein Paulsen, I owe a
@@
 
-Last but not least, I would like to thank the PhD committee for spending
+Last but not least, I would like to thank the PhD Committee for spending
 their precious time to review and assess this thesis.
@@
 series. Its main part consists of the following research papers
-written between October $2014$ and Augest $2017$.
+written between October $2014$ and August $2017$.
 
@@
 {\sc Janssen, A., Mikosch, T., Mohsen, R. and Xie, X.}
-\newblock Limit theory for the singular values of the sample autocovariance
+\newblock Limit theory for the singular values of the sample auto-covariance
   matrix function of multivariate time series.
@@
 {\sc Davis, R.~A., Heiny, J., Mikosch, T., and Xie, X.}
-\newblock Extreme value analysis for the sample autocovariance matrices of
+\newblock Extreme value analysis for the sample auto-covariance matrices of
   heavy-tailed multivariate time series.

### 2017-08-25 papers/Anja/PaperBody.tex
Probably the final revision of the thesis.

@@
 heavy-tailed behavior as well; see for example Ben Arous and Guionnet
-\cite{benarous:guionnet:2007}, 
+\cite{arous:guionnet:2008}, 
 Auffinger et al. \cite{auffinger:arous:peche:2009}, Soshnikov
@@
 
-\section{The model}\label{sec:model:anja}\setcounter{equation}{0}
+\section{The model}\label{sec:model:anja}
+% \setcounter{equation}{0}
 We consider a stochastic volatility model
@@
 
-\section{Case (1): $Z$ dominates the tail}\label{sec:case1}\setcounter{equation}{0}
+\section{Case (1): $Z$ dominates the tail}\label{sec:case1}
+% \setcounter{equation}{0}
 \subsection{Regular variation of the stochastic volatility model and
@@
 
-\section{Case (2): $\sigma$ dominates the tail}\label{Sec:case2}\setcounter{equation}{0}
+\section{Case (2): $\sigma$ dominates the tail}\label{Sec:case2}
+% \setcounter{equation}{0}
 In this section we assume the conditions of Case (2); see Section \ref{subsec:Case:2}. Our goal is to derive results analogous to Case (1): \regvar\ 
@@
 \begin{eqnarray*}
-\lefteqn{P\big((Y \Theta^{ij}_t)_{t=0,\ldots,n} \in B\big)}\\&=&\frac{\mu^{\bfX^{ij}}_n(\tilde{B})}{\mu^{\bfX^{ij}}_n(
+\lefteqn{\P\big((Y \Theta^{ij}_t)_{t=0,\ldots,n} \in B\big)}\\
+&=&
+\frac{\mu^{\bfX^{ij}}_n(\tilde{B})}{\mu^{\bfX^{ij}}_n(
 ([-\infty,\infty]\backslash [-1,1])\times [-\infty,\infty]^n)}\\
-&=& \frac{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_1^\infty \I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\nu_{\frac{\alpha}{\psi^{ij}}}(dy)\right]}{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\right]} \\
-&=& \sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\frac{1}{| \Lambda_{i,j}^{(0)}|}\frac{\E\left[\I\left(\left(\I((u,v) \in \Lambda_{i,j}^{(t)})Y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\right)_{t=0, \ldots, n} \in B\right)|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]}{\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]},
+&=&
+\frac{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[\int_1^\infty
+\I_B\left(y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\I_{\Lambda_{i,j}^{(t)}}(u,v)\right)_{t=0, \ldots, n}|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\nu_{\frac{\alpha}{\psi^{ij}}}(dy)\right]}{\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\ex^{-\alpha \eta_{u,v}}\right]} \\
+&=&
+\sum\limits_{(u,v) \in \Lambda_{i,j}^{(0)}}\frac{1}{| \Lambda_{i,j}^{(0)}|}\frac{\E\left[
+\I_B\left(y \frac{X_{it}X_{jt}}{|X_{i0}X_{j0}|}\I_{\Lambda_{i,j}^{(t)}}(u,v)\right)_{t=0, \ldots, n}|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]}{\E\left[|X_{i0}X_{j0}|^{\alpha/\psi^{ij}}\right]},
 \end{eqnarray*}
@@
 the limit \ds s of the eigenvalues and eigenvectors.
-\section{Simulations and data example}\label{sec:simulation}\setcounter{equation}{0}
+\section{Simulations and data example}\label{sec:simulation}
+% \setcounter{equation}{0}
 In this section we illustrate the behavior of sample covariance matrices for moderate sample sizes
@@
 
-\section{Some $\alpha$-stable limit theory}\setcounter{equation}{0}\label{App:A}
+\section{Some $\alpha$-stable limit theory}
+\label{App:A}
+% \setcounter{equation}{0}
 In this paper, we make frequently use of
@@
 for $s>0$,
-\begin{eqnarray*}  \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt) }{\P(Y_1^{a_{\max}}>t)}  &\geq& \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, \max_{j=1, \ldots, p}Y_j^{a_{\max}}\leq st) }{\P(Y_1^{a_{\max}}>t)} \\
-&& + \sum_{j:a_j =a_{\max}} \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_j^{a_{\max}}>st) }{\P(Y_1^{a_{\max}}>t)}\\
-&& -\frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_{j_1}^{a_{\max}}>st, Y_{j_2}^{a_{\max}}>st \mbox{ for some } j_1\neq j_2) }{\P(Y_1^{a_{\max}}>t)}, \;\; s>0.
+\begin{eqnarray*}
+  && \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt) }{\P(Y_1^{a_{\max}}>t)}  \\
+  &\geq&
+  \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, \max_{j=1, \ldots, p}Y_j^{a_{\max}}\leq st) }{\P(Y_1^{a_{\max}}>t)} \\
+&&
++ \sum_{j:a_j =a_{\max}} \frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_j^{a_{\max}}>st) }{\P(Y_1^{a_{\max}}>t)}\\
+&&
+-\frac{\P(\prod_{i=1}^pY_i^{a_i}>vt, Y_{j_1}^{a_{\max}}>st, Y_{j_2}^{a_{\max}}>st \mbox{ for some } j_1\neq j_2) }{\P(Y_1^{a_{\max}}>t)}, \;\; s>0.
 \end{eqnarray*} 
@@
 with $\mu_j$ as in \eqref{Eq:RV:prod:1}. For $s>0$ write
-\begin{eqnarray*} && \lim_{t \to \infty}\frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}\\
+\begin{small}
+  \begin{eqnarray*}
+    &&
+    \lim_{t \to \infty}\frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}\\
     &=& s^{-\alpha/{a_{\max}}} \lim_{t\to \infty} \P(\mathbf{Y} \in tB \mid Y_j^{a_{\max}}>st) \\
@@
   \end{eqnarray*}
+\end{small}
 Substituting $sy$ by $z$ in the integral finally gives
-\begin{small}
-$$ \lim_{s \searrow 0}\lim_{t \to \infty} \frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)}=\int_0^\infty \P\left(\left(\I(a_{ij}=a_{\max})z^{a_{\max}}\prod_{k \neq j}Y_k^{a_{ik}}\right)_{1 \leq i \leq n} \in B\right)\nu_{\alpha}(dz).
-$$
-\end{small}
+\begin{eqnarray*}
+&&
+\lim_{s \searrow 0}\lim_{t \to \infty} \frac{\P(\mathbf{Y} \in tB, Y_j^{a_{\max}}>st)}{\P(Y_1^{a_{\max}}>t)} \\
+&=&
+\int_0^\infty \P\left(\left(\I(a_{ij}=a_{\max})z^{a_{\max}}\prod_{k \neq j}Y_k^{a_{ik}}\right)_{1 \leq i \leq n} \in B\right)
+\nu_{\alpha}(dz).   
+\end{eqnarray*}
 (ii) Note first that under our assumptions for any $1\leq n_1 < n_2 \leq p$,

### 2017-08-25 papers/DavisMikoschHeinyXie/PaperBody.tex
Probably the final revision of the thesis.

@@
 
-\section{Estimation of the largest eigenvalues: an overview in the iid case}\label{sec:motivation}\setcounter{equation}{0}
+%% \section{Estimation of the largest eigenvalues: an overview in the iid case}\label{sec:motivation}
+\section{Estimation of the largest eigenvalues}
+\label{sec:motivation}
+% \setcounter{equation}{0}
 \subsection{The light-tailed case}
@@
 Over the last 15 years there has been increasing interest in the case when $p=p_n\to\infty$ as $\nto$. In most of
-the literature (exceptions are El Karoui \cite{elkaroui:2003}, Davis et al. \cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} and Heiny and Mikosch
-\cite{heiny:mikosch:2015:iid})
+the literature (exceptions are El Karoui \cite{elkaroui:2003}, Davis et al. \cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} and Heiny and Mikosch
+\cite{heiny:mikosch:2016})
 one assumes that $p$ and $n$ grow at the same rate:
@@
 Sections~\ref{sec:samplecov} and \ref{sec:possemidef} are devoted to {\em sample autocovariance matrices}.
-Motivated by \cite{lam:yao}, we study the eigenvalues of sums of transformed matrices and illustrate the results in two examples.
+Motivated by \cite{lam:yao:2012}, we study the eigenvalues of sums of transformed matrices and illustrate the results in two examples.
 These results are applied to the time series of S\&P 500 in Section \ref{sec:sp500}.
@@
 \section{General growth rates for $p_n$ in the iid heavy-tailed case}
-\label{sec:Davis2}\setcounter{equation}{0}
-This section is based on ideas in Heiny and Mikosch \cite{heiny:mikosch:2015:iid} where one can also find detailed proofs.
+\label{sec:Davis2}
+% \setcounter{equation}{0}
+This section is based on ideas in Heiny and Mikosch \cite{heiny:mikosch:2016} where one can also find detailed proofs.
 \subsection*{Growth conditions on $(p_n)$}\label{subsec:pn}
@@
 This obervation is in contrast to the light-tailed case; see \eqref{eq:geman}
-and \eqref{eq:tc}.  Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}
-and Heiny and Mikosch ~\cite{heiny:mikosch:2015:iid} allowed for more general rates for
+and \eqref{eq:tc}.  Davis et al.~\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014}
+and Heiny and Mikosch ~\cite{heiny:mikosch:2016} allowed for more general rates for
 $p_n\to\infty$ than linear growth in $n$.
@@
 Condition~\ref{eq:p} is more general than the growth conditions in the literature; see
-\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
+\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014}.
 \begin{theorem}\label{thm:intro}
@@
 which in combination with the Nagaev-type \ld\ results of  Theorem~\ref{thm:nagaev}
-yield the following result; see also Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
+yield the following result; see also Davis et al.~\cite{davis:mikosch:pfaffel:2016}.
 \begin{lemma}\label{lem:pp}
@@
 %-----------------------------------------------------------------------
-\section{Introducing dependence between the rows and columns}\label{sec:model}\setcounter{equation}{0}
-For details on the results of this section, we refer to Davis et al.~\cite{davis:mikosch:pfaffel:2015}, Heiny and Mikosch \cite{heiny:mikosch:2015:iid} and Heiny et al.~\cite{heiny:mikosch:2016:noniid}.
+\section{Introducing dependence between the rows and columns}\label{sec:model}
+% \setcounter{equation}{0}
+For details on the results of this section, we refer to Davis et al.~\cite{davis:mikosch:pfaffel:2016}, Heiny and Mikosch \cite{heiny:mikosch:2016} and Heiny et al.~\cite{heiny:mikosch:2016:noniid}.
 
@@
 The model \eqref{eq:Davis1} was introduced by Davis et al. \cite{davis:pfaffel:stelzer:2014}, assuming the rows iid, and in the
-present form by Davis et al. \cite{davis:mikosch:pfaffel:2015}.
+present form by Davis et al. \cite{davis:mikosch:pfaffel:2016}.
 \subsection{Sample covariance and autocovariance matrices}
@@
 The idea of using the sample autocovariance matrices and \fct s of their squares \eqref{eq:squares}
-originates from a paper by Lam and Yao~\cite{lam:yao} who used a model different from \eqref{eq:Davis1}.
+originates from a paper by Lam and Yao~\cite{lam:yao:2012} who used a model different from \eqref{eq:Davis1}.
 This idea is quite natural in the context of \tsa .
@@
 When $s=0$ we typically write $\la_i$ instead of $\la_i(0)$.
-\subsubsection*{The matrix $\M(s)$} We introduce some auxiliary matrices
+\subsubsection*{The matrix $\mathbf M(s)$} We introduce some auxiliary matrices
 derived from the coefficient matrix $\H=(h_{kl})_{k,l\in \Z}$:
 \beao
-\H(s)=(h_{k,l+s})_{k,l\in \Z}, \qquad \M(s)= \H(0)\H(s)'\qquad s\ge 0\,.
+\H(s)=(h_{k,l+s})_{k,l\in \Z}, \qquad \mathbf M(s)= \H(0)\H(s)'\qquad s\ge 0\,.
 \eeao
@@
 \begin{equation}\label{eq:m}
-(\M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z .
-\end{equation}
-We denote the ordered singular values of $\M(s)$ by
+(\mathbf M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z .
+\end{equation}
+We denote the ordered singular values of $\mathbf M(s)$ by
 \begin{equation}\label{eq:v1}
@@
 \end{equation}
-Let $r(s)$ be the rank of $\M(s)$ so that $v_{r(s)}(s)>0$ while $v_{r(s)+1}(s)=0$ if $r(s)$ is
+Let $r(s)$ be the rank of $\mathbf M(s)$ so that $v_{r(s)}(s)>0$ while $v_{r(s)+1}(s)=0$ if $r(s)$ is
 finite, otherwise $v_i(s)>0$ for all $i$. We also write $r=r(0)$.
@@
 \beam\label{eq:tracea}
-\sum_{i=1}^\infty (v_i(s))^2 &=&  \frobnorm{\M(s)}^2= \sum_{i,j\in \Z} \sum_{l_1,l_2 \in \Z} h_{i,l_1} h_{j,l_1+s}h_{i,l_2}
+\sum_{i=1}^\infty (v_i(s))^2 &=&  \frobnorm{\mathbf M(s)}^2= \sum_{i,j\in \Z} \sum_{l_1,l_2 \in \Z} h_{i,l_1} h_{j,l_1+s}h_{i,l_2}
 h_{j,l_2+s}\nonumber\\
@@
 The proof of Theorem~\ref{thm:mains} is given in Heiny et al.~\cite{heiny:mikosch:2016:noniid}. Part (2) of this  result
-with more restrictive conditions on the growth rate of $(p_n)$ is contained in Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
+with more restrictive conditions on the growth rate of $(p_n)$ is contained in Davis et al.~\cite{davis:mikosch:pfaffel:2016}.
 These proofs are very technical and lengthy.
@@
 then Theorem~\ref{thm:mains} remains valid conditionally on
-$(h_{kl})$, hence  unconditionally in $\P$-probability; see also \cite{davis:mikosch:pfaffel:2015}.\end{remark}
+$(h_{kl})$, hence  unconditionally in $\P$-probability; see also \cite{davis:mikosch:pfaffel:2016}.\end{remark}
 \subsection{Point process \con }
-Theorem~\ref{thm:mains}  and arguments similar to the proofs in Davis et al. \cite{davis:mikosch:pfaffel:2015}
+Theorem~\ref{thm:mains}  and arguments similar to the proofs in Davis et al. \cite{davis:mikosch:pfaffel:2016}
 enable one to derive the weak \con\ of
@@
 
-\section{Some applications}\setcounter{equation}{0}
-\subsection{Sample covariance matrices}\label{sec:samplecov}%\setcounter{equation}{0}
+\section{Some applications}
+% \setcounter{equation}{0}
+\subsection{Sample covariance matrices}\label{sec:samplecov}
 The sample covariance matrix $\X_n(0)\X_n(0)'=\X\X'$ is a non-negative definite matrix. Therefore
-its eigenvalues and singular values coincide.  Moreover, $v_j=v_j(0)$, $j\ge 1$, are the eigenvalues of $\M=\M(0)$.
+its eigenvalues and singular values coincide.  Moreover, $v_j=v_j(0)$, $j\ge 1$, are the eigenvalues of $\mathbf M=\mathbf M(0)$.
 \par
@@
 iid entries have this density. We assume $\beta=1$.
-Note that $\M=\M(0)$ has rank one and $v_1=1$.
+Note that $\mathbf M=\mathbf M(0)$ has rank one and $v_1=1$.
 The estimated densities of the deviations $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ and
@@
 An application of the continuous mapping theorem to  the distributional convergence of the point processes in Theorem~\ref{cor:1} in the spirit of Resnick
-\cite{resnick:2007}, Theorem 7.1, also yields the following result; see Davis et al. \cite{davis:mikosch:pfaffel:2015} for a proof and a similar result in the case $\alpha \in (2,4)$.
+\cite{resnick:2007}, Theorem 7.1, also yields the following result; see Davis et al. \cite{davis:mikosch:pfaffel:2016} for a proof and a similar result in the case $\alpha \in (2,4)$.
 \begin{corollary}\label{cor:1q} Assume the conditions of Theorem~\ref{thm:mains}.
@@
 To illustrate the theory we consider a simple moving average example
-taken from Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
+taken from Davis et al.~\cite{davis:mikosch:pfaffel:2016}.
 \begin{example} \rm
@@
 \end{equation*}
-Hence $\M=\H\H'$ has the positive eigenvalues
+Hence $\mathbf M=\H\H'$ has the positive eigenvalues
 $v_1=8$ and $v_2=2$. The limit point process in \eqref{eq:Davispp} is
@@
 The previous example also illustrates the behavior of the two largest eigenvalues in the general case
-when the rank $r$ of the matrix $\M$ is larger than one.
+when the rank $r$ of the matrix $\mathbf M$ is larger than one.
 We have in general
@@
 \begin{equation*}
-\M= \sum_{l\in \Z} c_l^2\;\; (\theta_i\theta_j)_{i,j \in \Z}\,.
+\mathbf M= \sum_{l\in \Z} c_l^2\;\; (\theta_i\theta_j)_{i,j \in \Z}\,.
 \end{equation*}
@@
 The \asy\ \ds s \eqref{eq:iid} correspond to the case when the matrix
-$\M$ has rank $r=1$. It includes the iid and separable cases;
+$\mathbf M$ has rank $r=1$. It includes the iid and separable cases;
 see Example~\ref{exam:separable}. The shown \asy\ quantiles are in agreement with the rank $r=1$ hypothesis.
@@
 To the best of our knowledge, sums of squares of sample autocovariance matrices were used first in the paper by Lam and Yao
-\cite{lam:yao}; their \ts\ model is quite different from ours.
+\cite{lam:yao:2012}; their \ts\ model is quite different from ours.
 \begin{proof}
@@
 \end{equation*}
-where $\omega_{(i)}(s_0,s_1)$ are the ordered values of the set $\{Z_{(i),np}^4 v_j(s_0,s_1), i=1,\ldots,p;j=1,2,\ldots\}$
-and $(v_j(s_0,s_1))$ are the ordered eigenvalues of $ \sum_{s=s_0}^{s_1} \M(s)\M(s)'$.
+where $\omega_{(i)}(s_0,s_1)$ are the ordered values of the set
+$\{Z_{(i),np}^4 v_j(s_0,s_1), i=1,\ldots,p;j=1,2,\ldots\}$
+and $(v_j(s_0,s_1))$ are the ordered eigenvalues of $ \sum_{s=s_0}^{s_1} \mathbf M(s)\mathbf M(s)'$.
 \end{proposition}
@@
 \begin{equation*}
-\M(s)=\gamma_c(s)\, \Theta, \quad s\ge 0\,,
+\mathbf M(s)=\gamma_c(s)\, \Theta, \quad s\ge 0\,,
 \end{equation*}
@@
 \eeao
-The matrix $\M(s)$ has the only non-zero eigenvalue $\gamma_c(s)\gamma_\theta(0)$.
+The matrix $\mathbf M(s)$ has the only non-zero eigenvalue $\gamma_c(s)\gamma_\theta(0)$.
 The factors $(\gamma_c(s))$ can be positive or negative; they constitute the autocovariance \fct\ of a stationary linear process
 with coefficients $(c_l)$.
-Accordingly, $\M(s)$ is either non-negative or non-positive definite. %This explains why a symmetrization of $\M(s)$ is irrelevant.
+Accordingly, $\mathbf M(s)$ is either non-negative or non-positive definite. %This explains why a symmetrization of $\mathbf M(s)$ is irrelevant.
 Now we consider the non-negative definite matrix
 \beao
-\sum_{s=s_0}^{s_1} \M(s)\,\M(s)'= \sum_{s=s_0}^{s_1}\gamma_c^2(s)\,\Theta\Theta'\,.
+\sum_{s=s_0}^{s_1} \mathbf M(s)\,\mathbf M(s)'= \sum_{s=s_0}^{s_1}\gamma_c^2(s)\,\Theta\Theta'\,.
 \eeao
@@
 the contribution of the squares $\A_n(s)\A_n(s)'$, $s=1,\ldots,5,$ to the largest eigenvalue of the sum of squares is less substantial.
-%Figure~\ref{fig:LamYao}(b) shows the same calculations for the matrices studied in Example~\ref{exam:4.6cont}.
-Here cancellation is possible (see \eqref{eq:Xrgsdf}) which is indicated by the fact that the red line does not follow the blue line as well as in (a).
 \end{example}
-%\section{\red Some conclusions?}\setcounter{equation}{0}
-%{\em Say something about EVT and largest eigenvalues}
 
@@
 
-\section{Auxiliary results}\label{appendix:A}\setcounter{equation}{0}
+\section{Auxiliary results}\label{appendix:A}
+% \setcounter{equation}{0}
 

### 2017-08-25 papers/DavisMikoschHeinyXie/s21.tex
Probably the final revision of the thesis.

@@
 
-\bibitem{davis:mikosch:pfaffel:2015}
+\bibitem{davis:mikosch:pfaffel:2016}
 {\sc Davis, R.~A., Mikosch, T., and Pfaffel, O.}
@@
 
-\bibitem{heiny:mikosch:2015:iid}
+\bibitem{heiny:mikosch:2016}
 {\sc Heiny, J., and Mikosch, T.}
@@
 
-\bibitem{lam:yao}
+\bibitem{lam:yao:2012}
 {\sc Lam, C., and Yao, Q.}
@@
 Next we briefly state a rather general version of a result 
-in Heiny and Mikosch \cite{heiny:mikosch:2015:iid} which will be relevant when we consider dependent entries of $\X$.
+in Heiny and Mikosch \cite{heiny:mikosch:2016} which will be relevant when we consider dependent entries of $\X$.
 
@@
 \end{equation*}
-as shown in \cite{heiny:mikosch:2015:iid}.
+as shown in \cite{heiny:mikosch:2016}.
 

### 2017-08-25 papers/FX/PaperBody.tex
Revised the phd thesis.
Updated the thesis manuscript.
Probably the final revision of the thesis.

@@
 
-\section{Introduction}\setcounter{equation}{0}
+\section{Introduction}
+% \setcounter{equation}{0}
 It is one of the stylized facts of financial econometrics that 
@@
 
-\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
+\section{Power-law tails of return series: some empirical results}\label{sec:1}
+% \setcounter{equation}{0}
 In this section, we assume the model \eqref{eq:1} for the tails of the marginal
@@
 see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, de
-Haan and Ferreira \cite{haan:ferreira:2006}, Resnick
+Haan and Ferreira \cite{dehaan:ferreira:2006}, Resnick
 \cite{resnick:2007}. The most popular among them was introduced  by
@@
 \cite{drees:rootzen:2010} give perhaps most general conditions for
-dependent \seq s and de Haan and Ferreira \cite{haan:ferreira:2006} provide
+dependent \seq s and de Haan and Ferreira \cite{dehaan:ferreira:2006} provide
 a complete \asy\ theory in the iid case.
@@
 Using the asymptotic normality property of upper order statistics
-(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, Theorem 2.2.1), one can show
+(cf. de Haan and Ferreira~\cite{dehaan:ferreira:2006}, Theorem 2.2.1), one can show
 \beao
@@
 $t\in (n\,t_0,n(1-t_0))$ where $t_0$ is a fixed number in $(0,0.5)$. Writing $\hat x_p(a,b)$ for an estimator of the extreme $(1-p)$-quantile 
-based on the subsample with indices $t\in (na,nb)$, the test statistic is given by
+based on the subsample with indices $t\in (na,nb)$, the test statistic
+is given by
+\begin{small}
   \beam\label{eq:4}
@@
   \eeam
+\end{small}
 Under the null hypothesis, $(T_n)$ converges to a complicated \fct al
@@
 \newpage
-\section{Some theoretical arguments for equality of tail-indices}\setcounter{equation}{0}\label{sec:2}
+\section{Some theoretical arguments for equality of tail-indices}
+\label{sec:2}
+% \setcounter{equation}{0}
 \subsection{Multivariate GARCH models whose components have  equal tail-indices}\label{subsec:garch}
@@
 We also have the specification
+\begin{small}
   \beam\label{eq:8}
@@
   \eeam
+\end{small}
 for positive $\alpha_{0i}$ and suitable non-negative 
@@
   {1 + b \1(x <  q) \over (K' + x)^{\beta + 1}} \,dx
-  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
+  \nonumber \\
+  && -b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
   \eeam
@@
     u(C(x)) \left[
-      1 - {b \over 1 + b}\1_{\{x \ge q\}}
+      1 - {b \over 1 + b}\1{x \ge q}
       \right]
@@
     (K + x)^{\alpha + 1}
-  } dx
-  - b u(\delta v) F_X(q)\,.\nonumber\\
+  } dx \nonumber \\
+  && - b u(\delta v) F_X(q)\,.
   \label{eq:ipf}
@@
 \section{A monotonicity lemma}
-\setcounter{equation}{0}
+% \setcounter{equation}{0}
 \begin{lemma} \label{lemma:I}
@@
 \section{When equity returns follow Student's t-distribution}
-\setcounter{equation}{0}
+% \setcounter{equation}{0}
 It is a common practice to use Student's t-distribution to model the
@@
       u(C(x)) \left[
-        1 - {b \over 1 + b}\1_{\{x \ge q\}}
+        1 - {b \over 1 + b}\1{x \ge q}
         \right]
@@
     }_{U_{\text{all}}}
-    f(x, \alpha) dx - b u(\delta v) F_X(q)
-    \label{eq:t1}
+    f(x, \alpha) dx \nonumber \\
+    && - b u(\delta v) F_X(q) \label{eq:t1}
   \end{eqnarray}
@@
   \]
-  \begin{minipage}{0.48\textwidth}
+  \begin{minipage}{0.44\textwidth}
     $\Psi(x)$ \linebreak
@@
   \end{minipage}\hfill
-  \begin{minipage}{0.5\textwidth}
+  \begin{minipage}{0.55\textwidth}
     As shown in the figure to the left, $\Psi(x)$ is increasing
@@
     representation
+    \begin{small}
     \[
     \Psi(x + 1) = -\gamma + \sum_{n=1}^\infty {x \over n(n + x)}
-    \quad x \neq -1, -2, -3,\dots
-    \]
+    \quad x \neq -1, -2, -3,\ldots
+    \]
+    \end{small}
     which gives
@@
 \section{Proof of Lemma~\ref{thrm:I}}
-\setcounter{equation}{0}
+% \setcounter{equation}{0}
 \label{sec:thrmI_proof}
@@
     {\pd \wt u(\alpha, \phi) \over \pd \alpha}
-    = {\partial \over \partial \alpha}\E[u((1 - \phi) e^r + \phi e^x) \1_{\{X < 0\}}]
+    = {\partial \over \partial \alpha}\E[u((1 - \phi) e^r + \phi e^x) \1{X < 0}]
   \end{eqnarray*}
@@
     \begin{array}{ll}
-      > 0 & \text{ when } x < K(1 - e^{1/\alpha}) < 0 \\
-      < 0 & \text{ when } K(1 - e^{1/\alpha}) < x < 0
+      < 0 & \text{ when } x < K(1 - e^{1/\alpha}) < 0 \\
+      > 0 & \text{ when } K(1 - e^{1/\alpha}) < x < 0
     \end{array}
@@
   = {\partial \over \partial K} \E \left[
-    u((1 - \phi) e^r + \phi e^x) \1_{\{X < 0\}}
+    u((1 - \phi) e^r + \phi e^x) \1{X < 0}
   \right] 
@@
 \section{Lemma~\ref{lemma:III}}
-\setcounter{equation}{0}
+% \setcounter{equation}{0}
 \begin{lemma}\label{lemma:III}

### 2017-08-25 phd/Chaptercorrelation.tex
Probably the final revision of the thesis.

@@
 
-Asymptotic theory for the eigenvalues of $\X\X'$ in the case of an entry distribution with infinite fourth moment was studied in \cite{soshnikov:2004,soshnikov:2006,auffinger:arous:peche:2009} in the cases when $p/n\to \gamma \in (0,\infty)$, while the authors of \cite{davis:mikosch:heiny:xie:2015,heiny:mikosch:2015:iid} allowed nearly arbitrary growth of the dimension $p$.
+Asymptotic theory for the eigenvalues of $\X\X'$ in the case of an entry distribution with infinite fourth moment was studied in \cite{soshnikov:2004,soshnikov:2006,auffinger:arous:peche:2009} in the cases when $p/n\to \gamma \in (0,\infty)$, while the authors of \cite{davis:mikosch:heiny:xie:2015,heiny:mikosch:2016} allowed nearly arbitrary growth of the dimension $p$.
 In their model, the entries of $\X$ are regularly varying with index $\alpha>0$, implying that
@@
 
-In view of \cite{davis:pfaffel:stelzer:2014,davis:mikosch:pfaffel:2015,davis:mikosch:heiny:xie:2015} the behavior of the top eigenvalues is reasonably well understood in the case of linear dependence among the $X_{it}$ and $\E[X^4]=\infty$.
+In view of \cite{davis:pfaffel:stelzer:2014,davis:mikosch:pfaffel:2016,davis:mikosch:heiny:xie:2015} the behavior of the top eigenvalues is reasonably well understood in the case of linear dependence among the $X_{it}$ and $\E[X^4]=\infty$.
 If $\E[X^4]<\infty$, similar arguments to \eqref{eq:lamu} can be developed to show that methods for sample covariance matrices can be applied to sample correlation matrices; see for example \cite{elkaroui:2009}.
@@
 For example, $\lambda_{(1)}/n=35.3196$ while $(1+\sqrt{\gamma})^2$ is only $1.7325$.
-By \cite{auffinger:arous:peche:2009,heiny:mikosch:2015:iid,davis:mikosch:heiny:xie:2015}, the properly normalized $\lambda_{(1)}$ converges to a \Frechet distributed random variable.
+By \cite{auffinger:arous:peche:2009,heiny:mikosch:2016,davis:mikosch:heiny:xie:2015}, the properly normalized $\lambda_{(1)}$ converges to a \Frechet distributed random variable.
 The correct normalization is roughly $n^{4/2.1}$ and hence it is expected that $\lambda_{(1)}/n$ is separated from the bulk, whose behavior ultimately determines the limiting spectral distribution, which is the \MP law with parameter $\gamma=0.1$.

### 2017-08-25 phd/Chapteriidpaper.tex
Probably the final revision of the thesis.

@@
 from which one can easily infer the joint limiting distribution of the $k$ largest eigenvalues.  
-Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} 
+Davis et al.~\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} 
 extended these results allowing for more general growth of $p$ 
@@
 are heavy-tailed and allow for dependence through the rows and across the columns; see 
-Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} for some recent attempts 
+Davis et al.~\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} for some recent attempts 
 when the entries satisfy some linear dependence conditions. In the iid case, these results are covered by the present paper
-and we also show that they remain valid under much more general growth conditions than in \cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
+and we also show that they remain valid under much more general growth conditions than in \cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014}.
 In particular, we make clear that centering of the sample covariance matrix (as assumed in
-\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} when $Z$ has a finite second moment) is not needed.
+\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} when $Z$ has a finite second moment) is not needed.
 Thus, our techniques are applicable under rather general dependence structures.
@@
 {\em sample autocovariance matrices}  which are a generalization of the traditional autocovariance function for time series to high-dimensional matrices.
-In applications, the analysis of sample autocovariance matrices for different lags might help to detect dependencies in the data; see Lam and Yao \cite{lam:yao} for related work.
+In applications, the analysis of sample autocovariance matrices for different lags might help to detect dependencies in the data; see Lam and Yao \cite{lam:yao:2012} for related work.
  We conclude with Appendix~\ref{appendix:A} which contains useful facts about regular variation and point processes.
@@
 This observation is in contrast to the light-tailed case; see \eqref{eq:MP} 
-and \eqref{eq:geman}.  Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} allowed for more general rates for
+and \eqref{eq:geman}.  Davis et al.~\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} allowed for more general rates for
 $p_n\to\infty$ than linear growth in $n$. {\em However, they could not completely solve the technical difficulties arising with general growth rates of~$p$.}
@@
 Condition~\ref{eq:p} is more general than the growth conditions in the literature; see 
-\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
+\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014}.
 
@@
 result whose proof will also be given in Section~\ref{sec:proofs};
-a version of this theorem was proved in Davis et al.~\cite{davis:mikosch:pfaffel:2015} 
+a version of this theorem was proved in Davis et al.~\cite{davis:mikosch:pfaffel:2016} 
 under more restrictive conditions on the growth rate of $(p_n)$.
@@
 \begin{remark} \rm
-Similar results were used in the proofs of Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2015}.
+Similar results were used in the proofs of Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2016}.
 We also mention that the centering $c_n$ in the finite variance case can be avoided if $n/a_{np}^2\to 0$. 
@@
 
-Theorem~\ref{thm:iidmain}  and arguments similar to the proofs in Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2015}
+Theorem~\ref{thm:iidmain}  and arguments similar to the proofs in Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2016}
 enable one to derive the weak \con\ of
@@
 \eeao 
-We refer to Davis et al.~\cite{davis:mikosch:pfaffel:2015} for details on the proofs and more examples.
+We refer to Davis et al.~\cite{davis:mikosch:pfaffel:2016} for details on the proofs and more examples.
 

### 2017-08-25 phd/Chapterintroduction.tex
Fix a minor error.
Probably the final revision of the thesis.

@@
 
-\section{Regular Variation}\label{sec:intro_regvar}
+\section{Regular variation}\label{sec:intro_regvar}
 The concept of regular variation is defined by the following scaling
@@
 \cite{feller}, p. 278. Cf. also Jessen and Mikosch
-\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et. al.
+\cite{jessen:mikosch:2006}, lemma 3.1, and Embrechts et. al.
 \cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.
@@
 $\min\{\alpha_1, \alpha_2\}$ (cf. Mikosch and Jessen
-\cite{JessenMikosch2006}). Moreover, if $X_1, X_2$ are iid,
+\cite{jessen:mikosch:2006}). Moreover, if $X_1, X_2$ are iid,
 $\P(a_1 X_1 + a_2 X_2 > u) \sim \P(a_1 X_1 > u) + \P(a_2 X_2 > u)$.
@@
 functions of regularly varying random variables, see Mikosch and
-Jessen \cite{JessenMikosch2006}.
+Jessen \cite{jessen:mikosch:2006}.
 
@@
 
-\section{Stochastic Recurrence Equation}
+\section{Stochastic recurrence equation}
 \label{sec:tyhyjyt}
@@
   \label{eq:rfht}
-  \log\sigma_t = \sum_{l \in \integers} \psi_l \eta_{t-l}
+  \log\sigma_t = \sum_{l \in \integers} \psi_l \eta_{t-l},
   \quad
-  t \in \integers,
+  t \in \integers
 \end{equation}
@@
 If the process of expected return has continuous sample path and $M_t$
-is a stochastic volatility process, it has been shown $[Y]_t = [M]_t$ and
-$[Y_\delta]_t \overset{P}{\to} \int_0^t \sigma_s^2 ds$. Meanwhile,
-It\^o's formula gives, for semi-martingale $Y_t$,
+is a stochastic volatility process, it has been shown that
+$[Y]_t = [M]_t$ and
+$[Y_\delta]_t \overset{P}{\to} \int_0^t \sigma_s^2 ds$.
+Meanwhile, It\^o's formula gives, for a semi-martingale $\{Y_t\}$,
 \begin{eqnarray}
@@
 \[
-\P(|V| > x) = \pi(\mathcal C) \E_\eta N_x = \E_\eta^{\alpha} \mathcal E_x
+\P(|V| > x) = \pi(\mathcal C) \E_\eta N_x
+= \E_\eta^{\mathcal D} \mathcal E_x,
 \]
@@
 In \S\ref{sec:SV:intro} we have introduced stochastic volatility
-models as well as their connection to realized quadratic
+models and discussed their connection to realized quadratic
 variation and their improved predictive power derived thereby. But
@@
 $\{X_{i,t}\}_{t \in \integers}, 1 \leq i \leq p$ and each of
-$\{X_{i,t} X_{j,t}\}_{t \in \integers}, 1\leq i,j \leq p$
+$\{X_{i,t} X_{j,t}\}_{t \in \integers}, 1\leq i < j \leq p$
 are regularly varying with index $\alpha$, assuming suitable
@@
 \]
-for some $\delta \in (0, \min\{\alpha/2, 1\})$. This conditions ensures
+for some $\delta \in (0, \min\{\alpha/2, 1\})$. This condition ensures
 that the infinite series in \eqref{eq:tyhjuj} is almost surely
@@
 \sum_{i=1}^p \varepsilon_{a_{np}^{-2}(\lambda_{(i)}(0), \ldots, \lambda_{(i)}(s))}
-\overset{v}{\to}
+\overset{d}{\to}
 \sum_{i=1}^\infty \sum_{j=1}^\infty
@@
 \]
-where ${a_{np}}_{n,p \in \integers}$ is a sequence of positive real
-numbers such that $\P(Z > a_{np}) = \ell(np) (np)^{-\alpha}$.
+where $\{a_{np}\}_{n,p \in \integers}$ is a sequence of positive real
+numbers such that $n p \P(Z > a_{np}) \to 1, n \to \infty$.
 $\varepsilon_{x}$ is the Dirac measure with unit mass at $x$ and
@@
 $X_{i,t}$, assuming regular variation with $\alpha \in (0, 4)$. This
-work also generalizes Davis, Pfaffel and Stelzer
-\cite{davis:pfaffel:stelzer:2014} as well as Davis, Mikosch and
-Pfaffel \cite{davis:mikosch:oliver:2016} who deal with the linear
-model \eqref{eq:tyhjuj} under suboptimal conditions on the growth rate
-of $p \to \infty$.
+work also generalizes Davis et al. \cite{davis:pfaffel:stelzer:2014}
+as well as Davis et al. \cite{davis:mikosch:pfaffel:2016} who deal with
+the linear model \eqref{eq:tyhjuj} under suboptimal conditions on the
+growth rate of $p \to \infty$.
 
@@
 random variables are consequently used; see Nagaev
-\cite{nagaev1979}. This is in contrast to the papers by Soshnikov
+\cite{nagaev:1979}. This is in contrast to the papers by Soshnikov
 \cite{soshnikov:2004}  and Auffinger et

### 2017-08-25 phd/MyThesis.tex
Probably the final revision of the thesis.

@@
 \hspace*{0.5cm} Chapter 3: \copyright Bernoulli\\
-\hspace*{0.5cm} Chapter 4: \copyright Elsevier\\
+\hspace*{0.5cm} Chapter 4: \copyright Elsevier
 
-% \noindent \hspace*{0.5cm} ISBN: $978$-$87$-$7078$-$921$-$9$
+% \begin{comment}
+% {	\pagestyle{empty}
+% 	\setlength{\topmargin}{0pt}
+% 	\setlength{\headheight}{0pt}
+% 	\setlength{\headsep}{0pt}
+% 	\setlength{\footskip}{0pt}
 
-\begin{comment}
-{	\pagestyle{empty}
-	\setlength{\topmargin}{0pt}
-	\setlength{\headheight}{0pt}
-	\setlength{\headsep}{0pt}
-	\setlength{\footskip}{0pt}
+% 	\setlength{\parskip}{12pt}
 
-	\setlength{\parskip}{12pt}
+% 	%\begin{center}
+% 		%\vspace*{2.5cm}
+%  \large \textbf{\textsc{Johannes Heiny}}
 
-	%\begin{center}
-		%\vspace*{2.5cm}
- \large \textbf{\textsc{Johannes Heiny}}
+% 		\vspace*{1cm}
 
-		\vspace*{1cm}
+% 	%\end{center}
 
-	%\end{center}
+% 	\vspace*{\fill}}
+% \end{comment}
 
-	\vspace*{\fill}}
-\end{comment}
-
-\chapter*{}
-\vspace*{-2cm}
+% \newpage
+% \chapter*{}
+% \vspace*{-2cm}
 \section*{Abstract}
@@
 \section*{Resum{\'e}}
-Denne afhandling handler om analyse af tidsrækker med tunge haler. Vi 
-diskuterer haleegenskaber af real-world equity return serie og
-undersøge muligheden for at et enkelt haleindeks deles af alle
-return serier af aktivt handlede aktier i et marked. Betingelser for
+Denne afhandling handler om en analyse af tidsrækker med tunge
+haler. Vi diskuterer haleegenskaber af real-world equity return series
+og undersøger muligheden for at et enkelt haleindeks deles af alle
+return seriers af aktivt handlede aktier i et marked. Betingelser for
 denne hypotese til være sande er identificeret.
@@
 Asymptotiske tilnærmelser af egenværdierne og egenvektorer af
-sådanne matricer findes og udtrykkes i afhængighedens parameter
-struktur, blandt andre.
+sådanne matricer findes og udtrykkes i afhængighedens
+parameterstruktur.
 
@@
 algoritme er effektiv i den forstand, at dens relative fejl forbliver
-afgrænset som sandsynligheden af interesse har tendens til nul. Vi gør
+begrænset, när sandsynligheden af interesse konvergerer til nul. Vi gør
 brug af eksponentiel vridning af overgangskernen til en {\em Markov
-additive process} og tage fordel af asymptotiske teorier om produkter
+additive process} og tager fordel af asymptotiske teorier om produkter
 af positive tilfældige matricer.
@@
 \newpage
+\vspace*{\fill}
+\newpage
+
 \tableofcontents
